### 2017-09-13 astronum_2017/astronum-amrex.tex
astronum 2017 paper

\documentclass[letterpaper]{jpconf}

\usepackage{graphicx}

\usepackage{hyperref}

\usepackage{cite}

\newcommand{\maestro}{{\sffamily Maestro}}
\newcommand{\castro}{{\sffamily Castro}}
\newcommand{\starkiller}{{\sffamily StarKiller}}
\newcommand{\nyx}{{\sffamily Nyx}}
\newcommand{\amrex}{{\sffamily AMReX}}

\begin{document}

\title{Meeting the Challenges of Modeling Astrophysical Thermonuclear Explosions:
\castro, \maestro, and the \amrex\ Astrophysics Suite}

\author{M. Zingale$^1$,
        A.~S. Almgren$^2$,
        M.~G. Barrios Sazo$^1$,
        V.~E. Beckner$^2$,
        J.~B. Bell$^2$,
        B. Friesen$^2$,
        A.~M. Jacobs$^3$,
        M.~P. Katz$^4$,
        C.~M. Malone$^5$,
        A.~J. Nonaka$^2$,
        D. Willcox$^1$, and
        W. Zhang$^2$}

\address{$^1$Department of Physics and Astronomy, Stony Brook
  University, Stony Brook, NY 11794-3800 USA}

\address{$^2$Center for Computational Sciences and Engineering,
  Lawrence Berkeley National Lab, Berkeley, CA 94720 USA}

\address{$^3$Department of Physics and Astronomy, Michigan State
  University, East Lansing, Michigan 48824 USA}

\address{$^4$Nvidia Corporation}

\address{$^5$Los Alamos National Laboratory, Los Alamos, NM, 87545 USA}

\begin{abstract}
We describe the \amrex\ suite of astrophysics codes and their
application to modeling problems in stellar astrophysics.
\maestro\ is tuned to efficiently model subsonic convective flows
while \castro\ models the highly compressible flows associated with
stellar explosions.  Both are built on the block-structured adaptive
mesh refinement library \amrex.  Together, these codes enable a
thorough investigation of stellar phenomena, including Type Ia
supernovae and X-ray bursts.  We describe these science applications
and the approach we are taking to make these codes performant on
current and future many-core and GPU-based architectures.
\end{abstract}

\section{Introduction}

Astrophysical explosions come in many flavors: gravitational and
thermonuclear supernovae, unstable burning on the surface of compact
objects, and explosive ignition of burning stages in stellar
evolution.  Accurate modeling of these events requires the coupling of
hydrodynamics, gravity, thermonuclear reactions, and in some cases,
radiation and magnetic fields.  Further, these environments are
characterized by a wide range of lengthscales, from the size of the
star or binary system down to the burning zone width and dissipation
scales.  Temporal scales are equally impressive---stellar evolution
occurs over 10s of millions to billions of years, the simmering phase
leading up to explosions lasts millenia to days or hours, and the
explosion can be over in seconds to hours.  The radiation leakage,
which leads to the observables we see lasts from hours to months.

No single algorithm meets all of the demands imposed by these events.
Instead, we advance our understanding of these events by piecing
together simulations of different phases of the evolution from
different codes.  Here we discuss our simulation codes, \maestro\ and
\castro\ designed to perform three-dimensional models of the early
subsonic evolution leading to runaway and the subsequent expolosion.
Together this suite of codes allows us to address many problems in
stellar and nuclear astrophysics.  We describe some of the design
details, the current architecture of the code, and some applications
below.

\section{Science drivers and challenges}

Our interests are thermonuclear explosions, including Type Ia
supernovae, X-ray bursts, and novae.  The basic ingredients for these
events are thermonuclear energy release and a degenerate equation of
state that allows a runaway to build without a pressure response.
Most of the current models for these events are characterized by a
long timescale ``simmmering'' phase where reactions heat the star or
layer and drive convection.  Eventually, reactions become vigorous
enough that a runaway takes place, perhaps with an accompanying
burning front that spreads through the star.

\subsection{Type Ia supernovae}

The largest open theoretical question for SNe Ia is---what is the
nature of the progenitor?  About 20 years ago, the community had
mostly converged up the single-degenerate scenario---a
Chandrasekhar-mass C/O white dwarf that accretes from its companion,
eventually leading to runaway at the center that burns through the
star (see~\cite{hillebrandtniemeyer2000} for the state of the field
at that time).  Since then, the wealth of observations has show that
there is a lot of diversity in SNe Ia, and searches for progenitor
systems have not shown that Chandra-mass white dwarfs can explain all
SNe Ia.  Today, merging white dwarfs (the double degenerate scenario)
have become the most popular model.  Other progenitors, like He
burning on the surface of a sub-Chandra white dwarf have also seen
interest in explaining some of the observed diversity.

There are open questions in all of these scenarios which can be
addressed through simulation.  For the Chandra and sub-Chandra models,
what is the distribution (spatial and temporal) of the hotspots setup
by turbulent convection that give rise to burning fronts.  Can you
create a detonation in the He layer in the double detonation model?
For double degenerates, does the burning take place promptly?  Is it
possible to accurate model a detonation in this case, given our
resolution constraints?  These are some of the questions we seek to
answer.

\subsection{XRBs}

X-ray bursts, the burning of accreted H/He on the surface of a neutron
star, can be important probes of neutron star structure.  Interpreting
observations requests that we understand what we are seeing, which can
be influenced by the products of the burning and how the burning
spreads across the star.  Many efforts have focused on different
aspects of these events.  One-dimensional models capture the
energetics well and inform us about the nucleosynthesis
\cite{woosley-xrb}.  Global models show the importance of rotation in
confining the burning \cite{SPIT_ETAL02}, why models inspired from
atmospheric science can explore the vertical structure
\cite{cavecchi:2012}.  However, the resolution differences from the
scale of the burning to a reasonable fraction of the neutron star
surface has prevented detailed explorations of the burning in resolved
calculations and how it feeds back on the flame structure and
propagation.

\subsection{Requirements}

conservation of gravity / angular momentum
                strong coupling of hydrodynamics and burning
                long timescale evolution

                3-d, turbulence, instabilities, rotation, ...

\section{AMReX Astrophysics Suite}

Our suite of application codes is built on the
\amrex\ block-structured adaptive mesh refinement (AMR) library.  The
basic programming model is that \amrex\ manages the grid
data-structures and parallelism, and calls the computational kernels
on a patch-by-patch basis.  AMReX supports subcycling in time, which
we use in \castro.

\amrex\ uses a hybrid approach to parallelism.  Tiling is used to
logically subdivide AMR patches to increase the parallelism for
OpenMP, without incurring the memory overheads that using many more
smaller grids would bring~\cite{tiling}.  This strategy is especially
important for many core architectures like the Intel Phi.  Ongoing development
is being done to support GPU offloading, using managed memory provided
by the latest generations of GPUs.

There are many application codes built on \amrex, including those in
combustion, ...  In astrophysics, these include \maestro\ and
\castro\ for stellar and nuclear astrophysics applications and
\nyx~\cite{nyx} for cosmological applications.  Here we focus on the
former two.

\subsection{\maestro}

\maestro~\cite{MAESTRO:Multilevel} is a low Mach number stellar
hydrodynamics code.  \maestro\ has been applied to convection in the
Chandrasekhar-mass model for Type Ia supernovae (SNe Ia)
\cite{ZABNW:IV,wdconvect,wdturb}, the sub-Chandra model for SNe Ia
\cite{subchandra,subchandra2} X-ray bursts~\cite{xrb,xrb2,xrb3}, and
convection in massive stars \cite{ms_cc}.

\maestro\ decomposes the state into a one-dimensional hydrostatic base
state and a three-dimensional Cartesian state that evolves the
departures from HSE.  A constraint equation is derived by requiring
that the pressure everywhere be close to the background hydrostatic
state.  The constraint acts to enforce instantaneous acoustic
equilibriation, effectively filtering soundwaves from the system,
while retaining the compressibility effects due to the background
stratification of the star and local heat release, as well as the
hydrostatic adjustment of the star.  In this fashion, it is more
general than traditional anelastic methods.  Low speed methods have
seen a lot of recent development~\cite{kleinpauluis,vasil:2013}, which
has been incorporated into \maestro.  The next development effort for
\maestro\ will be to allow for moderate rotation of the star.

\subsection{\castro}

\castro~\cite{castro,castroII,castroIII} is a fully-compressible
radiation hydrodynamics code that supports arbitrary equations of
state, nuclear reaction networks, and Poisson gravity.  \castro\ has
been applied to core-collapse supernovae~\cite{castro-ccsne},
population III pair-instability
supernovae~\cite{castro-pairinstability}, and white dwarf mergers as a
model for SNe Ia~\cite{wdmergerI}.  Recent work has focused on
improving energy conservation with gravitational and rotational
sources and new integration methods that more accurately couple hydro
and reactions.

For \maestro\ simulations that evolve from the subsonic regime to the
sonic regime, we have demonstrated the ability to restart the
calculations in \castro\ to continue the evolution into the sonic
regime~\cite{scidac-petascale,malone:2014} (in this case, for Chandra
model SNe Ia).

\subsection{\starkiller\ Microphysics}

\maestro\ and \castro\ share the same microphysics, available as the
\starkiller\ astrophysics github
project\footnote{\url{https://github.com/StarKiller-astro/microphysics/}}.
This includes equations of state and nuclear reaction networks.  The
reaction networks are written such that that rates and integration
strategy are decoupled, allowing us to change the integration strategy
for a set of rates.

\subsection{Open source and reproducibility}

All of our simulation codes are open source and follow a fully open
development model---the development git repos are hosted on
github\footnote{\url{https://github.com/AMReX-Astro/}}, available for
anyone to see and contribute to using issues and pull-requests.
{\em All source files, model files, input parameters, etc.\ for any
published science results are also available in the code repos.}  When
feasible, the git hashes for the published results are included in paper
acknowledgements.

\section{Parallel Performance and GPUs}

GPUs

     . show Don's plot

\begin{figure}[t]
\centering
\includegraphics[width=0.5\linewidth]{titan_xrb_scaling}
\caption{\label{fig:maestro_scaling} blah}
\end{figure}

Figure~\ref{fig:maestro_scaling} shows strong scaling for \maestro\ on the 3-d
XRB problem.

show new scaling

Our latest focus has been on GPUs

\section{Science results}

\begin{figure}[t]
\centering
\includegraphics[height=1.8in]{subch_h}
\includegraphics[height=1.8in]{xrb_compact}
\includegraphics[height=1.8in]{wdmerger_08030_new}
\caption{\label{fig:current-runs} (left) Convective plumes in a
  \maestro\ sub-Ch calculation. (center) Vertical velocity showing the
  convective structure in a \maestro\ XRB calculation. (right)
  Snapshot of a \castro\ simulation of the merger of two white dwarfs,
  with 0.90 and 0.81 solar masses. The contours represent density
  levels.}
\end{figure}

Figure~\ref{fig:current-runs} shows some of our most recent science simulations.

\ack The work at Stony Brook was supported by DOE/Office of Nuclear
Physics grant DE-FG02-87ER40317 and NSF award AST-1211563.  An award
of computer time was provided by the Innovative and Novel
Computational Impact on Theory and Experiment (INCITE) program.  This
research used resources of the Oak Ridge Leadership Computing Facility
at the Oak Ridge National Laboratory, which is supported by the Office
of Science of the U.S. Department of Energy under Contract
No.\ DE-AC05-00OR22725.  This research used resources of the National
Energy Research Scientific Computing Center, which is supported by the
Office of Science of the U.S. Department of Energy under Contract
No.\ DE-AC02-05CH11231.  Results in this paper were obtained using the
high-performance LIred computing system at the Institute for Advanced
Computational Science at Stony Brook University, which was obtained
through the Empire State Development grant NYS \#28451.  The GPU work
benefited from a grant of an Titan X Pascal board from Nvidia Corp
through the GPU Grant Program.

\section*{References}

\bibliographystyle{iopart-num}
\bibliography{ws}

\end{document}

### 2017-09-14 astronum_2017/astronum-amrex.tex
a few updates

@@
 
-Figure~\ref{fig:current-runs} shows some of our most recent science simulations.
+Figure~\ref{fig:current-runs} shows some of our most recent science
+simulations.  The left panel is an image of convection in the helium
+layer on a sub-Chandra white dwarf.  This is part of a study of the
+early stages of the double detonation SNe Ia model.  Using \maestro,
+we are able to model the convection in the He layer for many turnover
+times and saw a range of outcomes depending on the mass of the white
+dwarf and He layer, including a global He-nova-like behavior where the
+entire layer runsway together and localized
+burning~\cite{subchandra2}.
+
+The middle figure shows convection in a H/He layer on a neutron star,
+as a model of the early burning in an XRB.  This \maestro\ model was
+the first 3D model of convection for this problem~\cite{xrb3}.
+
+The rightmost image is the coalesced remains of the merger of a
+0.9~$M_\odot$ and 0.6~$M_\odot$ WD.
 

### 2017-10-02 astronum_2017/astronum-amrex.tex
more work on this proceedings

@@
 \newcommand{\amrex}{{\sffamily AMReX}}
+
+\usepackage{color}
+\setlength{\marginparwidth}{0.75in}
+\newcommand{\MarginPar}[1]{\marginpar{\vskip-\baselineskip\raggedright\tiny\sffamily\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}
 
@@
 eventually leading to runaway at the center that burns through the
-star (see~\cite{hillebrandtniemeyer2000} for the state of the field
-at that time).  Since then, the wealth of observations has show that
+star (see~\cite{hillebrandtniemeyer2000} for the state of the field at
+that time).  Since then, the wealth of observations has show that
 there is a lot of diversity in SNe Ia, and searches for progenitor
-systems have not shown that Chandra-mass white dwarfs can explain all
+systems have not shown that Chandra-mass white dwarfs can explain most
 SNe Ia.  Today, merging white dwarfs (the double degenerate scenario)
-have become the most popular model.  Other progenitors, like He
-burning on the surface of a sub-Chandra white dwarf have also seen
+have perhaps become the most popular model.  Other progenitors, like
+He burning on the surface of a sub-Chandra white dwarf have also seen
 interest in explaining some of the observed diversity.
+See~\cite{araa-maoz} for a review.
 
@@
 what is the distribution (spatial and temporal) of the hotspots setup
-by turbulent convection that give rise to burning fronts.  Can you
-create a detonation in the He layer in the double detonation model?
-For double degenerates, does the burning take place promptly?  Is it
-possible to accurate model a detonation in this case, given our
-resolution constraints?  These are some of the questions we seek to
-answer.
+by turbulent convection that give rise to burning fronts?  A
+longstanding question with the Chandra model is, can a deflagration
+transition to a detonation during the burning front propagation
+through the star.  For the sub-Chandra double detonation model, is it
+possible to create a detonation in the thin surface He layer?  For
+double degenerates, does the burning take place promptly? and can we
+avoid an accretion-induced collapse to a neutron star.  For many of
+these, we have the question of whether it is possible to accurately
+model the ignition of a detonation with the spatial resolution we can
+attain?  These are some of the questions we seek to answer.
 
@@
 star, can be important probes of neutron star structure.  Interpreting
-observations requests that we understand what we are seeing, which can
+observations requires that we understand what we are seeing, which can
 be influenced by the products of the burning and how the burning
@@
 \cite{woosley-xrb}.  Global models show the importance of rotation in
-confining the burning \cite{SPIT_ETAL02}, why models inspired from
+confining the burning \cite{SPIT_ETAL02}, while models inspired from
 atmospheric science can explore the vertical structure
@@
 calculations and how it feeds back on the flame structure and
-propagation.
+propagation.  Algorithms and computing architectures are starting to
+reach the point where we can span the gaps in spatial scales between
+these calculations to provide a better understanding of the ignition
+and propagation of the burning front, and the nucleosynthesis produced.
 
@@
 
-conservation of gravity / angular momentum
-                strong coupling of hydrodynamics and burning
-                long timescale evolution
-
-                3-d, turbulence, instabilities, rotation, ...
+These problems share common algorithmic requirements: strong coupling
+between hydrodynamics and burning, support for a general equation of state,
+self-gravity, including isolated boundary conditions, and long timescale
+evolution for the convective phases.  All of these problems are inherently
+three-dimensional, as turbulence, fluid instabilities, and rotation affect
+the dynamics.  Conservation is important as well, suggesting approaches
+which implement gravitational and rotation sources conservatively, and
+methods for improving angular momentum conservation.
 
@@
 \amrex\ block-structured adaptive mesh refinement (AMR) library.  The
-basic programming model is that \amrex\ manages the grid
-data-structures and parallelism, and calls the computational kernels
-on a patch-by-patch basis.  AMReX supports subcycling in time, which
-we use in \castro.
+basic programming model has \amrex\ managing the grid data-structures
+and parallel communications, and it calls the computational kernels on
+a patch-by-patch basis.  The core library is written in C++ with the
+computational kernels written in Fortran\footnote{Currently
+  \maestro\ is using an older pure-Fortran interface, but this will be
+  ported to the C++ AMReX this coming year.}---this allows us to take
+advantage of the strengths of both languages.AMReX supports subcycling
+in time, which we use in \castro.
 
-\amrex\ uses a hybrid approach to parallelism.  Tiling is used to
-logically subdivide AMR patches to increase the parallelism for
+\amrex\ uses a hybrid MPI + OpenMP approach to parallelism.
+Distribution of grid patches to nodes using MPI provides a natural
+coarse-grained approach to distributing the computational work, while
+threading of loops over zones in a grid using OpenMP provides
+effective fine-grained parallelization.  For \castro\, tiling is used
+to logically subdivide AMR patches to increase the parallelism for
 OpenMP, without incurring the memory overheads that using many more
 smaller grids would bring~\cite{tiling}.  This strategy is especially
-important for many core architectures like the Intel Phi.  Ongoing development
-is being done to support GPU offloading, using managed memory provided
-by the latest generations of GPUs.
+important for many core architectures like the Intel Phi.  Ongoing
+development is being done to support GPU offloading, using managed
+memory provided by the latest generations of GPUs.
 
 There are many application codes built on \amrex, including those in
-combustion, ...  In astrophysics, these include \maestro\ and
+combustion, ... \MarginPar{domains?} In astrophysics, these include \maestro\ and
 \castro\ for stellar and nuclear astrophysics applications and
@@
 \maestro~\cite{MAESTRO:Multilevel} is a low Mach number stellar
-hydrodynamics code.  \maestro\ has been applied to convection in the
-Chandrasekhar-mass model for Type Ia supernovae (SNe Ia)
+hydrodynamics code designed for efficiently modeling convection in
+stars.  \maestro\ decomposes the state into a one-dimensional
+hydrostatic base state and a three-dimensional Cartesian state that
+evolves the departures from HSE.  A constraint equation is derived by
+requiring that the pressure everywhere be close to the background
+hydrostatic state.  The constraint acts to enforce instantaneous
+acoustic equilibriation, effectively filtering soundwaves from the
+system, while retaining the compressibility effects due to the
+background stratification of the star and local heat release, as well
+as the hydrostatic adjustment of the star.  In this fashion, it is
+more general than traditional anelastic methods.  The timestep
+constraint for these equations depends only on the fluid velocity, not
+the sound speed, enabling much larger timesteps than compressible
+codes for highly subsonic flows.
+
+The state is advanced using a second-order accurate projection method.
+Fluid quantities are advected using an unsplit Godunov method, with
+reactions incorporated via operator splitting.  The provisional
+velocities are then projected onto the space that satisfies the
+divergence constraint.  The projections involve elliptic solves,
+solving a variable-coefficient Poisson problem, computed using a
+multigrid algorithm.  Low speed methods have seen a lot of recent
+development~\cite{kleinpauluis,vasil:2013}, which has been
+incorporated into \maestro.  The next development effort for
+\maestro\ will be to allow for moderate rotation of the star.
+
+\maestro\ has been applied to convection in the Chandrasekhar-mass
+model for Type Ia supernovae (SNe Ia)
 \cite{ZABNW:IV,wdconvect,wdturb}, the sub-Chandra model for SNe Ia
@@
 convection in massive stars \cite{ms_cc}.
-
-\maestro\ decomposes the state into a one-dimensional hydrostatic base
-state and a three-dimensional Cartesian state that evolves the
-departures from HSE.  A constraint equation is derived by requiring
-that the pressure everywhere be close to the background hydrostatic
-state.  The constraint acts to enforce instantaneous acoustic
-equilibriation, effectively filtering soundwaves from the system,
-while retaining the compressibility effects due to the background
-stratification of the star and local heat release, as well as the
-hydrostatic adjustment of the star.  In this fashion, it is more
-general than traditional anelastic methods.  Low speed methods have
-seen a lot of recent development~\cite{kleinpauluis,vasil:2013}, which
-has been incorporated into \maestro.  The next development effort for
-\maestro\ will be to allow for moderate rotation of the star.
 
@@
 radiation hydrodynamics code that supports arbitrary equations of
-state, nuclear reaction networks, and Poisson gravity.  \castro\ has
-been applied to core-collapse supernovae~\cite{castro-ccsne},
-population III pair-instability
+state, nuclear reaction networks, and Poisson gravity using geometric
+multigrid.  The main hydrodynamics scheme in \castro\ is an unsplit
+piecewise parabolic method.  Recent developments have improved
+conservation with gravity sources and rotation~\cite{wdmergerI}.  The
+radiation solver in \castro\ uses the flux-limited diffusion
+approximation for gray or multigroup radiation.  The integration
+algorithm on the grid hierarchy is a recursive procedure in which
+coarse grids are advanced in time, fine grids are advanced multiple
+steps to reach the same time as the coarse grids and the data at
+different levels are then synchronized. The synchronization for
+self-gravity is similar to the algorithm introduced by
+\cite{miniati-colella}.
+
+\castro\ has been applied to core-collapse
+supernovae~\cite{castro-ccsne}, population III pair-instability
 supernovae~\cite{castro-pairinstability}, and white dwarf mergers as a
@@
 sources and new integration methods that more accurately couple hydro
-and reactions.
-
-For \maestro\ simulations that evolve from the subsonic regime to the
-sonic regime, we have demonstrated the ability to restart the
-calculations in \castro\ to continue the evolution into the sonic
-regime~\cite{scidac-petascale,malone:2014} (in this case, for Chandra
-model SNe Ia).
+and reactions.  For \maestro\ simulations that evolve from the
+subsonic regime to the sonic regime, we have demonstrated the ability
+to restart the calculations in \castro\ to continue the evolution into
+the sonic regime~\cite{scidac-petascale,malone:2014} (in this case,
+for Chandra model SNe Ia).
 
@@
 acknowledgements.
+
+regression testing
 

### 2017-10-05 astronum_2017/scaling/summitdev/gpu_scaling.py
some scaling #s

@@ -1,0 +1,21 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+matplotlib.rcParams.update({'font.size': 14})
+
+data = np.loadtxt("starlord-2017-10-05.txt")
+
+nodes = data[:,0]
+GPUs = data[:,1]
+probsize = data[:,2]
+FOM = data[:,3]
+
+num_gpus = nodes*GPUs
+
+plt.scatter(num_gpus, FOM)
+
+plt.xlabel("# of GPUs", fontsize="medium")
+plt.ylabel(r"zones/$\mu$s updated", fontsize="medium")
+
+plt.tight_layout()
+plt.savefig("summitdev_scaling.pdf")

### 2017-10-05 astronum_2017/astronum-amrex.tex
more reproducibile
update fig
more plots

@@
 \newcommand{\starkiller}{{\sffamily StarKiller}}
+\newcommand{\starlord}{{\sffamily StarLord}}
 \newcommand{\nyx}{{\sffamily Nyx}}
@@
 anyone to see and contribute to using issues and pull-requests.
+New changes are put into the {\tt development} branch in each repo.
+Nightly regression testing is used to ensure that no new bugs were
+introduced.  Once a month, {\tt development} is merged into {\tt
+  master}.
 {\em All source files, model files, input parameters, etc.\ for any
@@
 acknowledgements.
-
-regression testing
 
@@
 \centering
-\includegraphics[width=0.5\linewidth]{titan_xrb_scaling}
-\caption{\label{fig:maestro_scaling} blah}
+\includegraphics[width=0.48\linewidth]{titan_edison_maestro_scaling}
+\includegraphics[width=0.48\linewidth]{summitdev_scaling}
+\caption{\label{fig:scaling} (left) \maestro\ strong scaling on NERSC edison
+and OCLF titan for a 3-d XRB problem.  Two different problem sizes are show.
+(right) \starlord\ scaling across nodes on the OLCF summitdev platform, using
+up to 4 GPUs per node.}
 \end{figure}

### 2017-10-05 astronum_2017/astronum-amrex.tex
more work

@@
 
-GPUs
-
-     . show Don's plot
+A key design goal of our application codes is performance portability.
+We want the same kernels to run on clusters, manycore machines (like
+Intel KNL), and GPU-based machines.  Our development has balanced this
+need with architecture-specific optimizations to maximize code reuse.
 
@@
 
-Figure~\ref{fig:maestro_scaling} shows strong scaling for \maestro\ on the 3-d
-XRB problem.
+Figure~\ref{fig:scaling} shows strong scaling for \maestro\ on the 3-d
+XRB problem.  We see that the code scales well to $\mathcal{O}(10^4)$
+processors.  The main limitation to the scaling at the moment are the
+multigrid solves used to enforce the projection (in particular the
+nodal solver).  \maestro\ also does not yet take advantage of the
+tiling capabilities offered in \amrex.  These will be addressed in the
+future as we port \maestro\ to the C++ \amrex\ codebase to take advantage
+of ongoing development.
 
-show new scaling
+Castro scaling...
 
-Our latest focus has been on GPUs
+Our latest focus has been on GPU ports of our application codes.  A
+small proxy app, \starlord, was created from \castro\ with just the
+hydrodynamics and stellar equation of state.  It uses a simple
+method-of-lines formulation of the hydrodynamics and advects 13
+nuclear species in addition to the hydrodynamics.  To offload work on
+GPUs, GPU support was added directly into \amrex.  In \amrex, grids
+consist of a number of boxes that enable domain decomposition for
+parallelism.  An iterator loops over all of the boxes at the same
+level of refinement and passes a data pointer into a Fortran kernel
+function where the work is done. For the simulation state data that
+resides in each box, we have modified the memory allocator so that it
+can use a CUDA allocator (either device or managed memory).  The
+domain iterator is configured to handle data motion to and from the
+device, so that the compute kernels can operate on data which is
+presumed to already be there, and the computation is decoupled from
+the memory management. Computation on the data can then be performed
+with OpenACC, CUDA Fortran, or (more recently) OpenMP 4.5. We have
+also built CUDA compute support into \amrex\ so that a compute kernel
+can be transparently operated on using CUDA Fortran without
+substantially modifying the kernels (and we anticipate using a similar
+strategy to use OpenACC and/or OpenMP in the future). This helps
+ensure that we can continue to maintain performance portability in our
+simulation codes, by decoupling the physics algorithms from the
+backend support used to implement them on various compute
+architectures.\MarginPar{add CPU time for reference}
+
+Figure~\ref{fig:scaling} shows the performance of \starlord\ on the
+summitdev platform at OLCF\footnote{summitdev consists of 2 IBM Power8
+  processors and 4 Nvidia Pascal GPUs per node.}.  The highest GPU
+count (128) corresponds to 32 nodes on summitdev.  We see a nearly
+linear speedup with the number of GPUs, indicating good weak scaling
+on this machine.  Efforts are underway to complete the port of the GPU
+developments into \castro.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.48\linewidth]{CUDA-Fortran-VODE-Integrator}
+\begin{minipage}[b]{0.48\linewidth}
+\caption{\label{fig:scaling} (left) \maestro\ strong scaling on NERSC edison
+and OCLF titan for a 3-d XRB problem.  Two different problem sizes are show.
+(right) \starlord\ scaling across nodes on the OLCF summitdev platform, using
+up to 4 GPUs per node.\vspace{2em}}
+\end{minipage}
+\end{figure}
+
+A parallel effort is porting our microphysics, in particular reaction
+networks to GPUs.  Our strategy is to do the entire ODE integration on
+the GPU, i.e., the data for a patch of zones is passed to the GPU, all
+righthand size and Jacobian evaluations and the timestepping itself is
+done on the GPU and once all zones are burned, we access the data as
+needed on the CPU.  To enable this, we ported our workhorse ODE
+integrator (VODE~\cite{vode}) to CUDA Fortran. This required extensive
+rewrites of the internals of VODE, which was originally written using
+Fortran 77 syntax unsupported by CUDA Fortran. Accelerating VODE with
+CUDA Fortran has proven successful, and we now see significant
+performance gains with the CUDA version of our reaction networks, even
+for moderate sized networks (a 13-isotope standard network).
+Figure~\ref{fig:cudaode} shows the speed-ups on a GPU vs.\ single CPU
+core.
 
@@
 \begin{figure}[t]
-\centering
-\includegraphics[height=1.8in]{subch_h}
+\centering \includegraphics[height=1.8in]{subch_h}
 \includegraphics[height=1.8in]{xrb_compact}
@@
 of Science of the U.S. Department of Energy under Contract
-No.\ DE-AC05-00OR22725.  This research used resources of the National
-Energy Research Scientific Computing Center, which is supported by the
-Office of Science of the U.S. Department of Energy under Contract
-No.\ DE-AC02-05CH11231.  Results in this paper were obtained using the
-high-performance LIred computing system at the Institute for Advanced
-Computational Science at Stony Brook University, which was obtained
-through the Empire State Development grant NYS \#28451.  The GPU work
-benefited from a grant of an Titan X Pascal board from Nvidia Corp
-through the GPU Grant Program.
+No.\ DE-AC05-00OR22725.  We appreciate the efforts of the OLCF (and in
+particular Fernanda Foertter) in organizing the GPU hackathons.  This
+research used resources of the National Energy Research Scientific
+Computing Center, which is supported by the Office of Science of the
+U.S. Department of Energy under Contract No.\ DE-AC02-05CH11231.
+Results in this paper were obtained using the high-performance LIred
+computing system at the Institute for Advanced Computational Science
+at Stony Brook University, which was obtained through the Empire State
+Development grant NYS \#28451.  The GPU work benefited from a grant of
+an Titan X Pascal board from Nvidia Corp through the GPU Grant
+Program.
 

### 2017-10-05 astronum_2017/scaling/summitdev/gpu_scaling.py

@@ -12,8 +12,9 @@ FOM = data[:,3]
 
 num_gpus = nodes*GPUs
 
-plt.scatter(num_gpus, FOM)
 
+plt.scatter(num_gpus, FOM)
+plt.plot(num_gpus, num_gpus/num_gpus[0]*FOM[0], ls=":")
 plt.xlabel("# of GPUs", fontsize="medium")
 plt.ylabel(r"zones/$\mu$s updated", fontsize="medium")
 

### 2017-10-05 astronum_2017/astronum-amrex.tex
acks

@@
 project\footnote{\url{https://github.com/StarKiller-astro/microphysics/}}.
-This includes equations of state and nuclear reaction networks.  The
-reaction networks are written such that that rates and integration
-strategy are decoupled, allowing us to change the integration strategy
-for a set of rates.
+This includes equations of state and nuclear reaction
+networks\footnote{Several of the reaction network righthand sides and
+  the EOS were provided from Frank Timmes' {\em cococubed} software
+  instruments page
+  \url{http://cococubed.asu.edu/code_pages/codes.shtml}.  We thank him
+  for making them available.}.  The reaction networks are written such
+that that rates and integration strategy are decoupled, allowing us to
+change the integration strategy for a set of rates.  They are also
+written to be threadsafe and with GPUs in mind (more on that
+below).  The goal of \starkiller\ is to evolve the code so that it is
+able to be used in a variety of nuclear astrophysics codes, not just
+those discussed here.
 
@@
 an Titan X Pascal board from Nvidia Corp through the GPU Grant
-Program.
+Program.  This research has made use of NASA's Astrophysics Data
+System Bibliographic Services.
 

### 2017-10-05 astronum_2017/astronum-amrex.tex
new fig

@@
 \includegraphics[width=0.48\linewidth]{titan_edison_maestro_scaling}
-\includegraphics[width=0.48\linewidth]{summitdev_scaling}
+\includegraphics[width=0.48\linewidth]{summitdev_scaling} \\
+\includegraphics[width=0.48\linewidth]{castro_scaling}
 \caption{\label{fig:scaling} (left) \maestro\ strong scaling on NERSC edison
@@
 
-Castro scaling...
+Castro scaling is a placeholder...
 
@@
 \begin{minipage}[b]{0.48\linewidth}
-\caption{\label{fig:scaling} (left) \maestro\ strong scaling on NERSC edison
-and OCLF titan for a 3-d XRB problem.  Two different problem sizes are show.
-(right) \starlord\ scaling across nodes on the OLCF summitdev platform, using
-up to 4 GPUs per node.\vspace{2em}}
+\caption{\label{fig:cudaode} GPU speed up over single CPU core for two
+  reaction networks.  The 13-isotope alpha-chain is a
+  potential network for the sub-Chandra simulations we propose here.
+  The 9-isotope Urca network is used for our Urca simulations.  A
+  variety of grid sizes were used, $16^3$, $32^3$, $64^3$, and
+  $128^3$.  In all cases, we see significant speed up on the GPU
+  vs.\ a single core.  These speed-up numbers are from our testbed
+  workstation with a Titan X Pascal board and Intel Xeon X5650 processors.}
 \end{minipage}
@@
 
-\section{Science results}
+\section{Some science results}
 
@@
 
-Figure~\ref{fig:current-runs} shows some of our most recent science
+Figure~\ref{fig:current-runs} shows some of our recent science
 simulations.  The left panel is an image of convection in the helium
@@
 as a model of the early burning in an XRB.  This \maestro\ model was
-the first 3D model of convection for this problem~\cite{xrb3}.
+the first 3D model of convection for this problem~\cite{xrb3}.  This
+study showed that the convective field became fully turbulent,
+achieving a Kolmogorov spectrum, and the overall dynamics was very
+different than our earlier 2-d simulations.
 
 The rightmost image is the coalesced remains of the merger of a
-0.9~$M_\odot$ and 0.6~$M_\odot$ WD.
+0.9~$M_\odot$ and 0.6~$M_\odot$ WD performed with \castro.  This used
+the developments from \cite{wdmergerI}.  Our primary focus with this
+suite of simulations is understanding the numerical sensitivity of
+mergers and collisions on the burning that takes place.
+
+\section{Summary and future development}
+
+We described our suite of astrophysics codes built on the
+\amrex\ framework.  These codes were developed to model problems in
+stellar astrophysics spanning from low speed convection to explosive
+burning.  A major theme of the codes is the open development model,
+with the code development done on github and all problem files needed
+to recreate any science results freely available.  Future development
+efforts for \maestro\ include higher-order hydrodynamics and
+time-integration and rotation.  For \castro, we are investigating
+strong coupling between hydrodynamics and reactions, new solvers
+(including MHD), and finishing the GPU port.
 

### 2017-10-06 astronum_2017/scaling/cori/cori_scaling.py
add Cori stuff

@@ -1,0 +1,51 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+matplotlib.rcParams.update({'font.size': 14})
+
+data = np.loadtxt("single_level.txt")
+
+nodes = data[:,0]
+MPI = data[:,1]
+OMP = data[:,2]
+time = data[:,3]
+
+num_threads = MPI*OMP
+
+plt.scatter(nodes, time, label="coarse level only")
+plt.plot(nodes, nodes[0]/nodes*time[0], ls=":")
+
+
+data = np.loadtxt("one_amr.txt")
+
+nodes = data[:,0]
+MPI = data[:,1]
+OMP = data[:,2]
+time = data[:,3]
+
+num_threads = MPI*OMP
+
+plt.scatter(nodes, time, label="coarse + 1 AMR level")
+plt.plot(nodes, nodes[0]/nodes*time[0], ls=":")
+
+plt.xlim(10, 256)
+
+plt.xlabel("# of nodes", fontsize="medium")
+plt.ylabel(r"avg time to update step", fontsize="medium")
+
+plt.legend(frameon=False)
+
+ax = plt.gca()
+ax.set_xscale("log")
+ax.set_yscale("log")
+
+threads_per_node = 68*4
+ax2 = ax.twiny()
+ax2.set_xlim(10*threads_per_node, 256*threads_per_node)
+ax2.set_xlabel("# of threads")
+
+f = plt.gcf()
+f.set_size_inches(8, 6)
+
+plt.tight_layout()
+plt.savefig("cori_scaling.pdf")

### 2017-10-09 astronum_2017/scaling/castro/sedov_scaling.py
add scaling

@@ -1,0 +1,46 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+class ScalingRun(object):
+    def __init__(self, MPI=1, OMP=1, max_grid=1, nzones=1, max_level=1, time=0.0, std=0.0):
+        self.MPI = MPI
+        self.OMP = OMP
+        self.cores = MPI * OMP
+        self.max_grid = max_grid
+        self.nzones = nzones
+        self.max_level = max_level
+        self.time = time
+        self.std = std
+
+
+def trend_line(c, t):
+    trend = t[0]*c[0]/c[:]
+    return trend
+
+runs = []
+
+data = np.loadtxt("castro-sedov-scaling.txt")
+
+for row in data:
+    runs.append(ScalingRun(MPI=row[0], OMP=row[1], max_grid=row[4], 
+                           nzones=row[5], max_level=row[6], time=row[7], std=row[8]))
+
+
+sizes = set([q.nzones for q in runs])
+
+for nz in sizes:
+    nz_runs = [q for q in runs if q.nzones == nz]
+    c = [q.cores for q in nz_runs]
+    t = [q.time for q in nz_runs]
+    err = [q.std for q in nz_runs]
+
+    color="C{:1d}".format(int(nz % len(sizes)))
+    plt.errorbar(c, t, yerr=err, fmt="o", color=color)
+    plt.plot(c, trend_line(c, t), ls=":", color=color)
+
+plt.xscale("log")
+plt.yscale("log")
+
+plt.xlabel("cores")
+plt.ylabel("avg. time / step")
+plt.savefig("sedov_scaling.png")

### 2017-10-09 astronum_2017/scaling/castro/wdmerger_scaling.py

@@ -1,0 +1,58 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+class ScalingRun(object):
+    def __init__(self, MPI=1, OMP=1, max_grid=1, nzones=1, max_level=1, time=0.0, std=0.0):
+        self.MPI = MPI
+        self.OMP = OMP
+        self.cores = MPI * OMP
+        self.max_grid = max_grid
+        self.nzones = nzones
+        self.max_level = max_level
+        self.time = time
+        self.std = std
+
+
+def trend_line(c, t):
+    cnew = np.array(list(set(c)))
+    print(cnew)
+    trend = t[0]*c[0]/cnew[:]
+    return cnew, trend
+
+runs = []
+
+data = np.loadtxt("castro-wdmerger-scaling.txt")
+
+for row in data:
+    runs.append(ScalingRun(MPI=row[0], OMP=row[1], max_grid=row[4], 
+                           nzones=row[5], max_level=row[6], time=row[7], std=row[8]))
+
+
+sizes = set([q.nzones for q in runs])
+#sizes = [512]
+levels = [0, 1]#set([int(q.max_level) for q in runs])
+
+markers = ["o", "^", "s"]
+
+for nl in levels:
+    for i, nz in enumerate(sizes):
+        nz_runs = [q for q in runs if q.nzones == nz and q.max_level == nl]
+        if len(nz_runs) == 0:
+            continue
+
+        c = [q.cores for q in nz_runs]
+        t = [q.time for q in nz_runs]
+        err = [q.std for q in nz_runs]
+
+        color="C{:1d}".format(int(i % len(levels)))
+        print(color, nz)
+        plt.errorbar(c, t, yerr=err, fmt=markers[nl], color=color)
+        ctrend, trend = trend_line(c, t)
+        plt.plot(ctrend, trend, ls=":", color=color)
+
+plt.xscale("log")
+plt.yscale("log")
+
+plt.xlabel("cores")
+plt.ylabel("avg. time / step")
+plt.savefig("wdmerger_scaling.png", dpi=150)

### 2017-10-09 astronum_2017/scaling/castro/sedov_scaling.py
add scaling

@@ -43,4 +43,4 @@ plt.yscale("log")
 
 plt.xlabel("cores")
 plt.ylabel("avg. time / step")
-plt.savefig("sedov_scaling.png")
+plt.savefig("sedov_scaling.pdf", dpi=150, bbox_inches="tight")

### 2017-10-09 astronum_2017/scaling/castro/wdmerger_scaling.py

@@ -55,4 +55,4 @@ plt.yscale("log")
 
 plt.xlabel("cores")
 plt.ylabel("avg. time / step")
-plt.savefig("wdmerger_scaling.png", dpi=150)
+plt.savefig("wdmerger_scaling.pdf", dpi=150, bbox_inches="tight")

### 2017-10-09 astronum_2017/scaling/maestro/scaling.py
maestro plot

@@ -1,0 +1,177 @@
+import math
+import numpy as np
+import matplotlib.pyplot as plt
+
+def scaling():
+
+    plt.rcParams.update({'xtick.labelsize': 14,
+                         'ytick.labelsize': 14,
+                         'font.size': 14})
+
+    plt.rc("axes", linewidth=1.5)
+    plt.rc("lines", markeredgewidth=1.5)
+
+
+    # titan data
+
+    data = np.loadtxt("scaling-titan.txt")
+
+    N_mpi = data[:,0]
+    N_omp = data[:,1]
+
+    t_adv = data[:,2]
+    t_MAC = data[:,3]
+    t_nodal = data[:,4]
+    t_react = data[:,5]
+    t_misc = data[:,6]
+    t_total = data[:,7]
+    std_total = data[:,8]
+    width = data[:,9]
+
+    cores = N_mpi * N_omp
+
+
+    # we ran 2 tests, one that was 384 zones wide, and one that was
+    # 768 zones wide we'll plot these in different colors, with hollow
+    # symbols for pure MPI
+
+    idx_384 = width[:] == 384
+    idx_MPI_384 = np.logical_and(width[:] == 384, N_omp[:] == 1)
+    idx_OMP_384 = np.logical_and(width[:] == 384, N_omp[:] != 1)
+
+    plt.errorbar(cores[idx_MPI_384], t_total[idx_MPI_384], yerr=std_total[idx_MPI_384],
+                 marker="o", markeredgecolor="C0", markerfacecolor="none",
+                 markersize=7, markeredgewidth=1.25,
+                 color="C0", ls="none")
+
+    plt.errorbar(cores[idx_OMP_384], t_total[idx_OMP_384], yerr=std_total[idx_OMP_384],
+                 marker="o", markeredgecolor="C0", markersize=7, markeredgewidth=1.25,
+                 color="C0", ls="none")
+
+    plt.plot(cores[idx_384], t_total[idx_384], color="C0", ls="-", lw=1.5)
+
+
+    idx_768 = width[:] == 768
+    idx_MPI_768 = np.logical_and(width[:] == 768, N_omp[:] == 1)
+    idx_OMP_768 = np.logical_and(width[:] == 768, N_omp[:] != 1)
+
+    plt.errorbar(cores[idx_MPI_768], t_total[idx_MPI_768], yerr=std_total[idx_MPI_768],
[…]

### 2017-10-09 astronum_2017/scaling/castro/sedov_scaling.py
style

@@ -1,6 +1,13 @@
 import numpy as np
 import matplotlib.pyplot as plt
 
+plt.rcParams.update({'xtick.labelsize': 14,
+                     'ytick.labelsize': 14,
+                     'font.size': 14})
+
+plt.rc("axes", linewidth=1.5)
+plt.rc("lines", markeredgewidth=1.5)
+
 class ScalingRun(object):
     def __init__(self, MPI=1, OMP=1, max_grid=1, nzones=1, max_level=1, time=0.0, std=0.0):
         self.MPI = MPI
@@ -26,21 +33,43 @@ for row in data:
                            nzones=row[5], max_level=row[6], time=row[7], std=row[8]))
 
 
-sizes = set([q.nzones for q in runs])
+sizes = sorted(list(set([int(q.nzones) for q in runs])))
 
-for nz in sizes:
+for i, nz in enumerate(sizes):
     nz_runs = [q for q in runs if q.nzones == nz]
     c = [q.cores for q in nz_runs]
     t = [q.time for q in nz_runs]
     err = [q.std for q in nz_runs]
 
-    color="C{:1d}".format(int(nz % len(sizes)))
+    color="C{:1d}".format(int(i % len(sizes)))
     plt.errorbar(c, t, yerr=err, fmt="o", color=color)
     plt.plot(c, trend_line(c, t), ls=":", color=color)
 
 plt.xscale("log")
 plt.yscale("log")
 
-plt.xlabel("cores")
+plt.ylim(1, 50)
+plt.xlabel("number of cores")
 plt.ylabel("avg. time / step")
+
+f
+or i, nz in enumerate(sizes):
+    color="C{:1d}".format(int(i % len(sizes)))
+    legs.append(plt.Line2D((0,1),(0,0), color=color))
+    legnames.append(r"${}^3$".format(nz))
+
+plt.legend(legs, legnames, frameon=False,
+           fontsize="small", numpoints=1, loc=3)
+
+
+ax = plt.gca()
+plt.text(0.95, 0.95, "Castro 3-d Sedov (pure hydro)", 
+         fontsize="small", horizontalalignment="right", transform = ax.transAxes)
+
+
+plt.tight_layout()
+
+f = plt.gcf()
+f.set_size_inches(8, 6)
+
 plt.savefig("sedov_scaling.pdf", dpi=150, bbox_inches="tight")

### 2017-10-09 astronum_2017/scaling/castro/wdmerger_scaling.py

@@ -1,6 +1,13 @@
 import numpy as np
 import matplotlib.pyplot as plt
 
+plt.rcParams.update({'xtick.labelsize': 14,
+                     'ytick.labelsize': 14,
+                     'font.size': 14})
+
+plt.rc("axes", linewidth=1.5)
+plt.rc("lines", markeredgewidth=1.5)
+
 class ScalingRun(object):
     def __init__(self, MPI=1, OMP=1, max_grid=1, nzones=1, max_level=1, time=0.0, std=0.0):
         self.MPI = MPI
@@ -14,7 +21,7 @@ class ScalingRun(object):
 
 
 def trend_line(c, t):
-    cnew = np.array(list(set(c)))
+    cnew = np.array(sorted(list(set(c))))
     print(cnew)
     trend = t[0]*c[0]/cnew[:]
     return cnew, trend
@@ -28,7 +35,7 @@ for row in data:
                            nzones=row[5], max_level=row[6], time=row[7], std=row[8]))
 
 
-sizes = set([q.nzones for q in runs])
+sizes = set([int(q.nzones) for q in runs])
 #sizes = [512]
 levels = [0, 1]#set([int(q.max_level) for q in runs])
 
@@ -53,6 +60,30 @@ for nl in levels:
 plt.xscale("log")
 plt.yscale("log")
 
-plt.xlabel("cores")
+plt.ylim(1, 200.)
+
+plt.xlabel("number of cores")
 plt.ylabel("avg. time / step")
+
+# custom legend
+legs = []
+legnames = []
+
+for i, nz in enumerate(sizes):
+    color="C{:1d}".format(int(i % len(sizes)))
+    legs.append(plt.Line2D((0,1),(0,0), color=color))
+    legnames.append(r"${}^3$".format(nz))
+
+plt.legend(legs, legnames, frameon=False,
+           fontsize="small", numpoints=1, loc=3)
+
+ax = plt.gca()
+plt.text(0.95, 0.95, "Castro 3-d wdmerger (hydro + Poisson gravity)", 
+         fontsize="small", horizontalalignment="right", transform = ax.transAxes)
+
+plt.tight_layout()
+
+f = plt.gcf()
+f.set_size_inches(8, 6)
+
 plt.savefig("wdmerger_scaling.pdf", dpi=150, bbox_inches="tight")

### 2017-10-09 astronum_2017/scaling/maestro/scaling.py
style

@@ -163,13 +163,13 @@ def scaling():
     plt.ylabel("avg. time / step")
 
     ax = plt.gca()
-    plt.text(0.9, 0.9, "Maestro 3-d XRB", fontsize="small", horizontalalignment="right", transform = ax.transAxes)
+    plt.text(0.95, 0.95, "Maestro 3-d XRB", fontsize="small", horizontalalignment="right", transform = ax.transAxes)
 
     plt.tight_layout()
 
     f = plt.gcf()
     f.set_size_inches(8, 6)
-    plt.savefig("titan_edison_maestro_scaling.png")
+
     plt.savefig("titan_edison_maestro_scaling.pdf", dpi=150, bbox_inches="tight")
 
 

### 2017-10-09 astronum_2017/scaling/castro/sedov_scaling.py
fix

@@ -52,8 +52,11 @@ plt.ylim(1, 50)
 plt.xlabel("number of cores")
 plt.ylabel("avg. time / step")
 
-f
-or i, nz in enumerate(sizes):
+# custom legend
+legs = []
+legnames = []
+
+for i, nz in enumerate(sizes):
     color="C{:1d}".format(int(i % len(sizes)))
     legs.append(plt.Line2D((0,1),(0,0), color=color))
     legnames.append(r"${}^3$".format(nz))

### 2017-10-09 astronum_2017/astronum-amrex.tex
add plot
updated scaling
more progress

@@
 \includegraphics[width=0.48\linewidth]{titan_edison_maestro_scaling}
-\includegraphics[width=0.48\linewidth]{summitdev_scaling} \\
-\includegraphics[width=0.48\linewidth]{castro_scaling}
-\caption{\label{fig:scaling} (left) \maestro\ strong scaling on NERSC edison
-and OCLF titan for a 3-d XRB problem.  Two different problem sizes are show.
-(right) \starlord\ scaling across nodes on the OLCF summitdev platform, using
-up to 4 GPUs per node.}
+\begin{minipage}[b]{0.48\linewidth}
+\caption{\label{fig:maestro_scaling} \maestro\ strong scaling on NERSC
+  edison and OLCF titan for a 3-d XRB problem.  Two different problem
+  sizes are shown.\vspace{2em}}
+\end{minipage}
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.48\linewidth]{sedov_scaling}
+\includegraphics[width=0.48\linewidth]{wdmerger_scaling}
+\caption{\label{fig:castro-scaling} (left) \castro\ strong scaling on OLCF titan
+for a pure hydro (Sedov) problem; (right) \castro\ strong scaling
+on OLCF titan for a hydro + Poisson gravity (wdmerger) problem.}
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.48\linewidth]{summitdev_scaling}
+\includegraphics[width=0.48\linewidth]{cori_scaling}
+\caption{\label{fig:knl-gpu-scaling} \castro\ scaling on new architectures.}
 \end{figure}
@@
 
-Castro scaling is a placeholder...
+Figure~\ref{fig:castro-scaling}
 
@@
 
-Figure~\ref{fig:scaling} shows the performance of \starlord\ on the
+Figure~\ref{fig:knl-gpu-scaling} shows the performance of \starlord\ on the
 summitdev platform at OLCF\footnote{summitdev consists of 2 IBM Power8

### 2017-10-09 astronum_2017/scaling/cori/cori_scaling.py
update some style

@@ -1,7 +1,12 @@
 import numpy as np
 import matplotlib.pyplot as plt
-import matplotlib
-matplotlib.rcParams.update({'font.size': 14})
+
+plt.rcParams.update({'xtick.labelsize': 14,
+                     'ytick.labelsize': 14,
+                     'font.size': 14})
+
+plt.rc("axes", linewidth=1.5)
+plt.rc("lines", markeredgewidth=1.5)
 
 data = np.loadtxt("single_level.txt")
 
@@ -33,7 +38,7 @@ plt.xlim(10, 256)
 plt.xlabel("# of nodes", fontsize="medium")
 plt.ylabel(r"avg time to update step", fontsize="medium")
 
-plt.legend(frameon=False)
+plt.legend(frameon=False, loc=3)
 
 ax = plt.gca()
 ax.set_xscale("log")
@@ -44,8 +49,13 @@ ax2 = ax.twiny()
 ax2.set_xlim(10*threads_per_node, 256*threads_per_node)
 ax2.set_xlabel("# of threads")
 
+ax = plt.gca()
+plt.text(0.95, 0.95, "Castro KNL scaling (wdmerger)", 
+         fontsize="small", horizontalalignment="right", transform = ax.transAxes)
+
+
 f = plt.gcf()
 f.set_size_inches(8, 6)
 
 plt.tight_layout()
-plt.savefig("cori_scaling.pdf")
+plt.savefig("cori_scaling.pdf", dpi=150, bbox_inches="tight")

### 2017-10-09 astronum_2017/scaling/summitdev/gpu_scaling.py

@@ -1,7 +1,13 @@
 import numpy as np
 import matplotlib.pyplot as plt
-import matplotlib
-matplotlib.rcParams.update({'font.size': 14})
+
+plt.rcParams.update({'xtick.labelsize': 14,
+                     'ytick.labelsize': 14,
+                     'font.size': 14})
+
+plt.rc("axes", linewidth=1.5)
+plt.rc("lines", markeredgewidth=1.5)
+
 
 data = np.loadtxt("starlord-2017-10-05.txt")
 
@@ -15,8 +21,17 @@ num_gpus = nodes*GPUs
 
 plt.scatter(num_gpus, FOM)
 plt.plot(num_gpus, num_gpus/num_gpus[0]*FOM[0], ls=":")
-plt.xlabel("# of GPUs", fontsize="medium")
+plt.xlabel("number of GPUs", fontsize="medium")
 plt.ylabel(r"zones/$\mu$s updated", fontsize="medium")
 
+ax = plt.gca()
+plt.text(0.95, 0.95, "StarLord GPU performance", 
+         fontsize="small", horizontalalignment="right", transform = ax.transAxes)
+
+
 plt.tight_layout()
-plt.savefig("summitdev_scaling.pdf")
+
+f = plt.gcf()
+f.set_size_inches(8, 6)
+
+plt.savefig("summitdev_scaling.pdf", dpi=150, bbox_inches="tight")

### 2017-10-11 astronum_2017/scaling/castro/sedov_scaling.py
some updates

@@ -33,17 +33,24 @@ for row in data:
                            nzones=row[5], max_level=row[6], time=row[7], std=row[8]))
 
 
-sizes = sorted(list(set([int(q.nzones) for q in runs])))
+sizes = set([int(q.nzones) for q in runs])
+levels = [0, 1]#set([int(q.max_level) for q in runs])
 
-for i, nz in enumerate(sizes):
-    nz_runs = [q for q in runs if q.nzones == nz]
-    c = [q.cores for q in nz_runs]
-    t = [q.time for q in nz_runs]
-    err = [q.std for q in nz_runs]
+markers = ["o", "^", "s"]
 
-    color="C{:1d}".format(int(i % len(sizes)))
-    plt.errorbar(c, t, yerr=err, fmt="o", color=color)
-    plt.plot(c, trend_line(c, t), ls=":", color=color)
+for nl in levels:
+    for i, nz in enumerate(sizes):
+        nz_runs = [q for q in runs if q.nzones == nz and q.max_level == nl]
+        if len(nz_runs) == 0:
+            continue
+
+        c = [q.cores for q in nz_runs]
+        t = [q.time for q in nz_runs]
+        err = [q.std for q in nz_runs]
+
+        color="C{:1d}".format(int(i % len(sizes)))
+        plt.errorbar(c, t, yerr=err, fmt=markers[nl], color=color)
+        plt.plot(c, trend_line(c, t), ls=":", color=color)
 
 plt.xscale("log")
 plt.yscale("log")

### 2017-10-11 astronum_2017/scaling/castro/sedov_scaling.py
updated figs

@@ -55,7 +55,7 @@ for nl in levels:
 plt.xscale("log")
 plt.yscale("log")
 
-plt.ylim(1, 50)
+plt.ylim(1, 100)
 plt.xlabel("number of cores")
 plt.ylabel("avg. time / step")
 

### 2017-10-11 astronum_2017/astronum-amrex.tex
add comment about domain volume

@@
   edison and OLCF titan for a 3-d XRB problem.  Two different problem
-  sizes are shown.\vspace{2em}}
+  sizes are shown.  We see excellent strong scaling to high core counts
+  for this problem.\vspace{2em}}
 \end{minipage}
@@
 \includegraphics[width=0.48\linewidth]{wdmerger_scaling}
-\caption{\label{fig:castro-scaling} (left) \castro\ strong scaling on OLCF titan
-for a pure hydro (Sedov) problem; (right) \castro\ strong scaling
-on OLCF titan for a hydro + Poisson gravity (wdmerger) problem.}
+\caption{\label{fig:castro-scaling} (left) \castro\ strong scaling on
+  OLCF titan for a pure hydro (Sedov w/ real EOS) problem.  The
+  different colors represent different base resolutions.  We see
+  excellent strong and weak scaling for the single level runs.  For
+  the $1024^3$ run, we also ran with one $2\times$ jump level of
+  refinment (triangles) and see good strong scaling.  The variability
+  (shown by the error bars) at high core counts shows we are becoming
+  work-starved. (right) \castro\ strong scaling on OLCF titan for a
+  hydro + Poisson gravity (wdmerger) problem.  This problem uses a
+  multipole solver to determine Dirichlet boundary conditions
+  representing an isolated mass distribution, and then geometric
+  multigrid to solve for the potential in the interior.  Two coarse
+  grid sizes are shown, and demonstrate great strong scaling.  We also
+  look at a single level of refinement (a $4\times$ jump) on top of
+  this coarse grid.}
 \end{figure}
@@
 \includegraphics[width=0.48\linewidth]{cori_scaling}
-\caption{\label{fig:knl-gpu-scaling} \castro\ scaling on new architectures.}
+\caption{\label{fig:knl-gpu-scaling} \castro\ scaling on new
+  architectures.  The left figure shows the \castro\ GPU proxy app
+  (\starlord) on OLCF summitdev.  We see good scaling behavior to 64
+  GPUs in this test of a pure hydro problem (Sedov w/ real EOS).  On the
+right, we explore performance on NERSC Cori (and Intel KNL platform).}
 \end{figure}
@@
 Figure~\ref{fig:scaling} shows strong scaling for \maestro\ on the 3-d
-XRB problem.  We see that the code scales well to $\mathcal{O}(10^4)$
-processors.  The main limitation to the scaling at the moment are the
-multigrid solves used to enforce the projection (in particular the
-nodal solver).  \maestro\ also does not yet take advantage of the
-tiling capabilities offered in \amrex.  These will be addressed in the
-future as we port \maestro\ to the C++ \amrex\ codebase to take advantage
-of ongoing development.
+XRB problem.  We ran on both OLCF titan and NERSC edison.  We see that
+the code scales well to $\mathcal{O}(10^4)$ processors.  The main
+limitation to the scaling at the moment are the multigrid solves used
+to enforce the projection (in particular the nodal solver).
+\maestro\ also does not yet take advantage of the tiling capabilities
+offered in \amrex.  These will be addressed in the future as we port
+\maestro\ to the C++ \amrex\ codebase to take advantage of ongoing
+development.
 
-Figure~\ref{fig:castro-scaling}
+Figure~\ref{fig:castro-scaling} shows scaling behavior for a pure
+hydro + real EOS problem and a hydro / real EOS, + self-gravity
+problem.  All tests were run on OLCF titan.  The AMD processors in
+titan can be used in a 8 or 16 cores / node configuration.  For all
+these runs, we ran in the 16 cores / node mode.  All runs also used
+the PGI 17.7 compilers.  We see excellent scaling behavior for the
+pure hydro problem.  For the self-gravity problem, we scale well with
+a uniform grid, but with a $4\times$ jump for the AMR level on top of
+this, we do worse at higher core counts.  This is a challenging
+problem, as only 3\% of the domain volume is refined.  For a given
+core count, there are a number of different combinations of MPI ranks
+and OpenMP threads we can use.  In general, with multilevel problems,
+we found the best performance with fewer MPI ranks and more OpenMP
+threads.
 
@@
 
-Figure~\ref{fig:knl-gpu-scaling} shows the performance of \starlord\ on the
-summitdev platform at OLCF\footnote{summitdev consists of 2 IBM Power8
-  processors and 4 Nvidia Pascal GPUs per node.}.  The highest GPU
-count (128) corresponds to 32 nodes on summitdev.  We see a nearly
-linear speedup with the number of GPUs, indicating good weak scaling
-on this machine.  Efforts are underway to complete the port of the GPU
-developments into \castro.
+Figure~\ref{fig:knl-gpu-scaling} shows the performance of
+\starlord\ on the summitdev platform at OLCF\footnote{summitdev
+  consists of 2 IBM Power8 processors and 4 Nvidia Pascal GPUs per
+  node.}.  The highest GPU count (128) corresponds to 32 nodes on
+summitdev.  We see a nearly linear speedup with the number of GPUs,
+indicating good weak scaling on this machine.  Efforts are underway to
+complete the port of the GPU developments into \castro.  This figure
+also shows our performance on the Intel KNL manycore processors (using
+NERSC Cori).  This is for the same problem as the self-gravity test on
+titan. \MarginPar{need more info: base grid?}  The KNL chips in Cori
+have 68 cores than can be run with 1, 2, or 4 hardware threads each.
+For all these runs, we ran with 4 threads / core (effectively 272
+threads / chip).
 

### 2017-10-12 astronum_2017/scaling/castro/sedov_scaling.py
more scaling

@@ -34,11 +34,12 @@ for row in data:
 
 
 sizes = set([int(q.nzones) for q in runs])
-levels = [0, 1]#set([int(q.max_level) for q in runs])
+levels = set([int(q.max_level) for q in runs])
 
 markers = ["o", "^", "s"]
 
 for nl in levels:
+    
     for i, nz in enumerate(sizes):
         nz_runs = [q for q in runs if q.nzones == nz and q.max_level == nl]
         if len(nz_runs) == 0:
@@ -49,13 +50,13 @@ for nl in levels:
         err = [q.std for q in nz_runs]
 
         color="C{:1d}".format(int(i % len(sizes)))
-        plt.errorbar(c, t, yerr=err, fmt=markers[nl], color=color)
+        plt.errorbar(c, t, yerr=err, fmt=markers[min(nl, len(markers)-1)], color=color)
         plt.plot(c, trend_line(c, t), ls=":", color=color)
 
 plt.xscale("log")
 plt.yscale("log")
 
-plt.ylim(1, 100)
+plt.ylim(1, 500)
 plt.xlabel("number of cores")
 plt.ylabel("avg. time / step")
 

### 2017-10-12 astronum_2017/scaling/cori/wdmerger_scaling.py
numbers updated for runs with MPI multiples of 16 in cori, using gcc. Still need to add the plot for intel compilers

@@ -1,0 +1,91 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+plt.rcParams.update({'xtick.labelsize': 14,
+                     'ytick.labelsize': 14,
+                     'font.size': 14})
+
+plt.rc("axes", linewidth=1.5)
+plt.rc("lines", markeredgewidth=1.5)
+
+class ScalingRun(object):
+    def __init__(self, MPI=1, OMP=1, max_grid=1, nzones=1, max_level=1, time=0.0, std=0.0):
+        self.MPI = MPI
+        self.OMP = OMP
+        self.cores = MPI * OMP
+        self.max_grid = max_grid
+        self.nzones = nzones
+        self.max_level = max_level
+        self.time = time
+       # self.std = std
+
+
+def trend_line(c, t):
+    cnew = np.array(sorted(list(set(c))))
+    print(cnew)
+    trend = t[0]*c[0]/cnew[:]
+    return cnew, trend
+
+runs = []
+
+data = np.loadtxt("castro-wdmerger-scaling.txt")
+
+
+for row in data:
+    runs.append(ScalingRun(MPI=row[0], OMP=row[1], max_grid=row[4], 
+                           nzones=row[5], max_level=row[6], time=row[7]))
+
+
+sizes = set([int(q.nzones) for q in runs])
+#sizes = [512]
+levels = [0, 1]#set([int(q.max_level) for q in runs])
+
+markers = ["o", "^", "s"]
+
+for nl in levels:
+    for i, nz in enumerate(sizes):
+        nz_runs = [q for q in runs if q.nzones == nz and q.max_level == nl]
+        if len(nz_runs) == 0:
+            continue
+
+        c = [q.cores for q in nz_runs]
+        t = [q.time for q in nz_runs]
+        #err = [q.std for q in nz_runs]
+
+        color="C{:1d}".format(int(i % len(levels)))
+        print(color, nz)
+        plt.scatter(c,t,c=color,marker=markers[nl])
+        #plt.errorbar(c, t, yerr=err, fmt=markers[nl], color=color)
+        ctrend, trend = trend_line(c, t)
+        plt.plot(ctrend, trend, ls=":", color=color)
+
+plt.xscale("log")
+plt.yscale("log")
+
+plt.ylim(1, 2000.)
+
+plt.xlabel("number of cores")
+plt.ylabel("avg. time / step")
+
+# custom legend
[…]

### 2017-10-13 astronum_2017/scaling/cori/cori_wdmerger_scaling.py
rename

@@ -62,7 +62,7 @@ for nl in levels:
 plt.xscale("log")
 plt.yscale("log")
 
-plt.ylim(1, 2000.)
+plt.ylim(10, 2000.)
 
 plt.xlabel("number of cores")
 plt.ylabel("avg. time / step")
@@ -88,4 +88,4 @@ plt.tight_layout()
 f = plt.gcf()
 f.set_size_inches(8, 6)
 
-plt.savefig("wdmerger_scaling.pdf", dpi=150, bbox_inches="tight")
+plt.savefig("cori_scaling.pdf", dpi=150, bbox_inches="tight")

### 2017-10-13 astronum_2017/scaling/cori/cori_wdmerger_scaling.py
clean up the script

@@ -9,9 +9,10 @@ plt.rc("axes", linewidth=1.5)
 plt.rc("lines", markeredgewidth=1.5)
 
 class ScalingRun(object):
-    def __init__(self, MPI=1, OMP=1, max_grid=1, nzones=1, max_level=1, time=0.0, std=0.0):
+    def __init__(self, MPI=1, OMP=1, max_grid=1, nzones=1, nodes=1, max_level=1, time=0.0, std=0.0):
         self.MPI = MPI
         self.OMP = OMP
+        self.nodes = nodes
         self.cores = MPI * OMP
         self.max_grid = max_grid
         self.nzones = nzones
@@ -32,7 +33,7 @@ data = np.loadtxt("castro-wdmerger-scaling.txt")
 
 
 for row in data:
-    runs.append(ScalingRun(MPI=row[0], OMP=row[1], max_grid=row[4], 
+    runs.append(ScalingRun(MPI=row[0], OMP=row[1], nodes=row[3], max_grid=row[4], 
                            nzones=row[5], max_level=row[6], time=row[7]))
 
 
@@ -48,24 +49,33 @@ for nl in levels:
         if len(nz_runs) == 0:
             continue
 
-        c = [q.cores for q in nz_runs]
+        n = [q.nodes for q in nz_runs]
         t = [q.time for q in nz_runs]
         #err = [q.std for q in nz_runs]
 
         color="C{:1d}".format(int(i % len(levels)))
         print(color, nz)
-        plt.scatter(c,t,c=color,marker=markers[nl])
-        #plt.errorbar(c, t, yerr=err, fmt=markers[nl], color=color)
-        ctrend, trend = trend_line(c, t)
-        plt.plot(ctrend, trend, ls=":", color=color)
+        plt.scatter(n, t, c=color, marker=markers[nl])
+        #plt.errorbar(n, t, yerr=err, fmt=markers[nl], color=color)
+        ntrend, trend = trend_line(n, t)
+        plt.plot(ntrend, trend, ls=":", color=color)
 
 plt.xscale("log")
 plt.yscale("log")
 
 plt.ylim(10, 2000.)
 
-plt.xlabel("number of cores")
+plt.xlabel("number of nodes")
 plt.ylabel("avg. time / step")
+
+threads_per_node = 68*4
+ax = plt.gca()
+ax.set_xlim(4, 256)
+
+ax2 = ax.twiny()
[…]

### 2017-10-13 astronum_2017/astronum-amrex.tex

@@
-\documentclass[letterpaper]{jpconf}
+\documentclass[a4paper]{jpconf}
 
@@
 \newcommand{\amrex}{{\sffamily AMReX}}
+
+\newcommand{\cpp}{C\nolinebreak\hspace{-.05em}\raisebox{.4ex}{\tiny\bf +}\nolinebreak\hspace{-.10em}\raisebox{.4ex}{\tiny\bf +}}
 
@@
 different codes.  Here we discuss our simulation codes, \maestro\ and
-\castro\ designed to perform three-dimensional models of the early
+\castro, designed to perform three-dimensional models of the early
 subsonic evolution leading to runaway and the subsequent expolosion.
@@
 Our interests are thermonuclear explosions, including Type Ia
-supernovae, X-ray bursts, and novae.  The basic ingredients for these
-events are thermonuclear energy release and a degenerate equation of
-state that allows a runaway to build without a pressure response.
-Most of the current models for these events are characterized by a
-long timescale ``simmmering'' phase where reactions heat the star or
-layer and drive convection.  Eventually, reactions become vigorous
-enough that a runaway takes place, perhaps with an accompanying
-burning front that spreads through the star.
+supernovae (SNe Ia), X-ray bursts (XRBs), and novae.  The basic
+ingredients for these events are thermonuclear energy release and a
+degenerate equation of state that allows a runaway to build without a
+pressure response.  Most of the current models for these events are
+characterized by a long timescale ``simmmering'' phase where reactions
+heat the star or layer and drive convection.  Eventually, reactions
+become vigorous enough that a runaway takes place, perhaps with an
+accompanying burning front that spreads through the star.
 
@@
 avoid an accretion-induced collapse to a neutron star.  For many of
-these, we have the question of whether it is possible to accurately
-model the ignition of a detonation with the spatial resolution we can
-attain?  These are some of the questions we seek to answer.
+these, we also need to address numerical issues, such as whether it is
+possible to accurately model the ignition of a detonation with the
+spatial resolution we can attain?  These are some of the questions we
+seek to answer.
 
@@
 and parallel communications, and it calls the computational kernels on
-a patch-by-patch basis.  The core library is written in C++ with the
+a patch-by-patch basis.  The core library is written in \cpp\ with the
 computational kernels written in Fortran\footnote{Currently
   \maestro\ is using an older pure-Fortran interface, but this will be
-  ported to the C++ AMReX this coming year.}---this allows us to take
+  ported to the \cpp\ AMReX this coming year.}---this allows us to take
 advantage of the strengths of both languages.AMReX supports subcycling
@@
 development~\cite{kleinpauluis,vasil:2013}, which has been
-incorporated into \maestro.  The next development effort for
-\maestro\ will be to allow for moderate rotation of the star.
+incorporated into \maestro.
 
@@
 multigrid.  The main hydrodynamics scheme in \castro\ is an unsplit
-piecewise parabolic method.  Recent developments have improved
-conservation with gravity sources and rotation~\cite{wdmergerI}.  The
-radiation solver in \castro\ uses the flux-limited diffusion
-approximation for gray or multigroup radiation.  The integration
-algorithm on the grid hierarchy is a recursive procedure in which
-coarse grids are advanced in time, fine grids are advanced multiple
-steps to reach the same time as the coarse grids and the data at
-different levels are then synchronized. The synchronization for
-self-gravity is similar to the algorithm introduced by
-\cite{miniati-colella}.
+piecewise parabolic method.  The radiation solver in \castro\ uses the
+flux-limited diffusion approximation for gray or multigroup radiation.
+The integration algorithm on the grid hierarchy is a recursive
+procedure in which coarse grids are advanced in time, fine grids are
+advanced multiple steps to reach the same time as the coarse grids and
+the data at different levels are then synchronized. The
+synchronization for self-gravity is similar to the algorithm
+introduced by \cite{miniati-colella}.  Recent developments in
+\castro\ include a deferred-correction method of coupling
+hydrodynamics and reactions, a conservative gravity and rotation
+source formulation~\cite{wdmergerI}, and retry mechanism to redo a
+step based on criteria evaluated during the integration.\MarginPar{more new stuff?}
 
@@
 change the integration strategy for a set of rates.  They are also
-written to be threadsafe and with GPUs in mind (more on that
-below).  The goal of \starkiller\ is to evolve the code so that it is
-able to be used in a variety of nuclear astrophysics codes, not just
-those discussed here.
+written to be threadsafe and with GPUs in mind (more on that below).
+The goal of \starkiller\ is to create a set of community microphysics
+solvers that can be used in a variety of nuclear astrophysics codes,
+not just those discussed here.
 
@@
 anyone to see and contribute to using issues and pull-requests.
-New changes are put into the {\tt development} branch in each repo.
-Nightly regression testing is used to ensure that no new bugs were
-introduced.  Once a month, {\tt development} is merged into {\tt
-  master}.
-{\em All source files, model files, input parameters, etc.\ for any
-published science results are also available in the code repos.}  When
-feasible, the git hashes for the published results are included in paper
-acknowledgements.
+Additionally, we have mailing lists for discussions and asking for
+help.  Several branches are used in our workflow.  New changes are put
+into the {\tt development} branch in each repo.  Nightly regression
+testing is used to ensure that no new bugs were introduced.  Once a
+month, {\tt development} is merged into {\tt master}.  Finally, {\em all source
+  files, model files, input parameters, etc.\ for any published
+  science results are also available in the code repos.}  When
+feasible, the git hashes for the published results are included in
+paper acknowledgements.
 
@@
 
-Figure~\ref{fig:scaling} shows strong scaling for \maestro\ on the 3-d
-XRB problem.  We ran on both OLCF titan and NERSC edison.  We see that
-the code scales well to $\mathcal{O}(10^4)$ processors.  The main
-limitation to the scaling at the moment are the multigrid solves used
-to enforce the projection (in particular the nodal solver).
-\maestro\ also does not yet take advantage of the tiling capabilities
-offered in \amrex.  These will be addressed in the future as we port
-\maestro\ to the C++ \amrex\ codebase to take advantage of ongoing
-development.
+Figure~\ref{fig:maestro_scaling} shows strong scaling for \maestro\ on the 3-d
+XRB problem.  This is a typical \maestro\ application~\cite{xrb3}, where burning
+can be a significant part of the overall evolution.  We ran on both
+OLCF titan and NERSC edison.  We see that the code scales well to
+$\mathcal{O}(10^4)$ processors.  The main limitation to the scaling at
+the moment are the multigrid solves used to enforce the projection (in
+particular the nodal solver).  \maestro\ also does not yet take
+advantage of the tiling capabilities offered in \amrex.  These will be
+addressed in the future as we port \maestro\ to the
+\cpp\ \amrex\ codebase to take advantage of ongoing development.
 
@@
 nuclear species in addition to the hydrodynamics.  To offload work on
-GPUs, GPU support was added directly into \amrex.  In \amrex, grids
-consist of a number of boxes that enable domain decomposition for
-parallelism.  An iterator loops over all of the boxes at the same
-level of refinement and passes a data pointer into a Fortran kernel
-function where the work is done. For the simulation state data that
-resides in each box, we have modified the memory allocator so that it
-can use a CUDA allocator (either device or managed memory).  The
-domain iterator is configured to handle data motion to and from the
-device, so that the compute kernels can operate on data which is
-presumed to already be there, and the computation is decoupled from
-the memory management. Computation on the data can then be performed
-with OpenACC, CUDA Fortran, or (more recently) OpenMP 4.5. We have
-also built CUDA compute support into \amrex\ so that a compute kernel
-can be transparently operated on using CUDA Fortran without
-substantially modifying the kernels (and we anticipate using a similar
-strategy to use OpenACC and/or OpenMP in the future). This helps
-ensure that we can continue to maintain performance portability in our
-simulation codes, by decoupling the physics algorithms from the
-backend support used to implement them on various compute
-architectures.\MarginPar{add CPU time for reference}
+GPUs, GPU support was added directly into \amrex\footnote{This support
+  is currently on a feature branch in the git repo, awaiting merge
+  into {\tt development}.}.  In \amrex, grids consist of a number of
+boxes that enable domain decomposition for parallelism.  An iterator
+loops over all of the boxes at the same level of refinement and passes
+a data pointer into a Fortran kernel function where the work is
+done. For the simulation state data that resides in each box, we have
+modified the memory allocator so that it can use a CUDA allocator
+(either device or managed memory).  The domain iterator is configured
+to handle data motion to and from the device, so that the compute
+kernels can operate on data which is presumed to already be there, and
+the computation is decoupled from the memory management. Computation
+on the data can then be performed with OpenACC, CUDA Fortran, or (more
+recently) OpenMP 4.5. We have also built CUDA compute support into
+\amrex\ so that a compute kernel can be transparently operated on
+using CUDA Fortran without substantially modifying the kernels (and we
+anticipate using a similar strategy to use OpenACC and/or OpenMP in
+the future). This helps ensure that we can continue to maintain
+performance portability in our simulation codes, by decoupling the
+physics algorithms from the backend support used to implement them on
+various compute architectures.\MarginPar{add CPU time for reference}
 
@@
 entire layer runsway together and localized
-burning~\cite{subchandra2}.
+burning~\cite{subchandra2}.  Work continues to characterize the ignition.
 
@@
 achieving a Kolmogorov spectrum, and the overall dynamics was very
-different than our earlier 2-d simulations.
+different than our earlier 2-d simulations.  This calculation acts as 
+a bridge to our next set of studies that will look at larger scales.
 
@@
 suite of simulations is understanding the numerical sensitivity of
-mergers and collisions on the burning that takes place.
+mergers and collisions on the burning that takes place.  This work
+is ongoing.
 
@@
 We described our suite of astrophysics codes built on the
-\amrex\ framework.  These codes were developed to model problems in
-stellar astrophysics spanning from low speed convection to explosive
-burning.  A major theme of the codes is the open development model,
-with the code development done on github and all problem files needed
-to recreate any science results freely available.  Future development
-efforts for \maestro\ include higher-order hydrodynamics and
-time-integration and rotation.  For \castro, we are investigating
-strong coupling between hydrodynamics and reactions, new solvers
-(including MHD), and finishing the GPU port.
+\amrex\ block-structured adaptive mesh refinement framework.  These
+codes were developed to model problems in stellar astrophysics
+spanning from low speed convection to explosive burning.  A major
+theme of the codes is the open development model, with the code
+development done on github and all problem files needed to recreate
+any science results freely available.  Future development efforts for
+\maestro\ include higher-order hydrodynamics and time-integration and
+rotation.  For \castro, we are investigating stronger coupling between
+hydrodynamics and reactions, new solvers (including MHD), and
+finishing the GPU port.
 

### 2017-10-14 astronum_2017/scaling/cori/cori_scaling.py
add intel + legend

@@ -1,61 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-
-plt.rcParams.update({'xtick.labelsize': 14,
-                     'ytick.labelsize': 14,
-                     'font.size': 14})
-
-plt.rc("axes", linewidth=1.5)
-plt.rc("lines", markeredgewidth=1.5)
-
-data = np.loadtxt("single_level.txt")
-
-nodes = data[:,0]
-MPI = data[:,1]
-OMP = data[:,2]
-time = data[:,3]
-
-num_threads = MPI*OMP
-
-plt.scatter(nodes, time, label="coarse level only")
-plt.plot(nodes, nodes[0]/nodes*time[0], ls=":")
-
-
-data = np.loadtxt("one_amr.txt")
-
-nodes = data[:,0]
-MPI = data[:,1]
-OMP = data[:,2]
-time = data[:,3]
-
-num_threads = MPI*OMP
-
-plt.scatter(nodes, time, label="coarse + 1 AMR level")
-plt.plot(nodes, nodes[0]/nodes*time[0], ls=":")
-
-plt.xlim(10, 256)
-
-plt.xlabel("# of nodes", fontsize="medium")
-plt.ylabel(r"avg time to update step", fontsize="medium")
-
-plt.legend(frameon=False, loc=3)
-
-ax = plt.gca()
-ax.set_xscale("log")
-ax.set_yscale("log")
-
-threads_per_node = 68*4
-ax2 = ax.twiny()
-ax2.set_xlim(10*threads_per_node, 256*threads_per_node)
-ax2.set_xlabel("# of threads")
-
-ax = plt.gca()
-plt.text(0.95, 0.95, "Castro KNL scaling (wdmerger)", 
-         fontsize="small", horizontalalignment="right", transform = ax.transAxes)
-
-
-f = plt.gcf()
-f.set_size_inches(8, 6)
-
-plt.tight_layout()
-plt.savefig("cori_scaling.pdf", dpi=150, bbox_inches="tight")

### 2017-10-14 astronum_2017/scaling/cori/cori_wdmerger_scaling.py

@@ -9,7 +9,7 @@ plt.rc("axes", linewidth=1.5)
 plt.rc("lines", markeredgewidth=1.5)
 
 class ScalingRun(object):
-    def __init__(self, MPI=1, OMP=1, max_grid=1, nzones=1, nodes=1, max_level=1, time=0.0, std=0.0):
+    def __init__(self, MPI=1, OMP=1, max_grid=1, nzones=1, nodes=1, max_level=1, time=0.0, std=0.0, compiler=""):
         self.MPI = MPI
         self.OMP = OMP
         self.nodes = nodes
@@ -18,6 +18,7 @@ class ScalingRun(object):
         self.nzones = nzones
         self.max_level = max_level
         self.time = time
+        self.compiler = compiler
        # self.std = std
 
 
@@ -34,7 +35,7 @@ data = np.loadtxt("castro-wdmerger-scaling.txt")
 
 for row in data:
     runs.append(ScalingRun(MPI=row[0], OMP=row[1], nodes=row[3], max_grid=row[4], 
-                           nzones=row[5], max_level=row[6], time=row[7]))
+                           nzones=row[5], max_level=row[6], time=row[7], compiler=row[8]))
 
 
 sizes = set([int(q.nzones) for q in runs])
@@ -42,23 +43,31 @@ sizes = set([int(q.nzones) for q in runs])
 levels = [0, 1]#set([int(q.max_level) for q in runs])
 
 markers = ["o", "^", "s"]
+gnu_colors = ["C0", "C1"]
+intel_colors = ["C9", "C3"]
 
-for nl in levels:
-    for i, nz in enumerate(sizes):
-        nz_runs = [q for q in runs if q.nzones == nz and q.max_level == nl]
-        if len(nz_runs) == 0:
-            continue
+compiler = {0: "gnu", 1: "intel"}
 
-        n = [q.nodes for q in nz_runs]
-        t = [q.time for q in nz_runs]
-        #err = [q.std for q in nz_runs]
+for comp in compiler:
+    for nl in levels:
+        for i, nz in enumerate(sizes):
+            nz_runs = [q for q in runs if q.nzones == nz and q.max_level == nl and q.compiler == comp]
+            if len(nz_runs) == 0:
+                continue
 
-        color="C{:1d}".format(int(i % len(levels)))
-        print(color, nz)
-        plt.scatter(n, t, c=color, marker=markers[nl])
[…]

### 2017-10-14 astronum_2017/scaling/maestro/scaling.py
update style

@@ -36,35 +36,19 @@ def scaling():
     # symbols for pure MPI
 
     idx_384 = width[:] == 384
-    idx_MPI_384 = np.logical_and(width[:] == 384, N_omp[:] == 1)
-    idx_OMP_384 = np.logical_and(width[:] == 384, N_omp[:] != 1)
 
-    plt.errorbar(cores[idx_MPI_384], t_total[idx_MPI_384], yerr=std_total[idx_MPI_384],
-                 marker="o", markeredgecolor="C0", markerfacecolor="none",
-                 markersize=7, markeredgewidth=1.25,
-                 color="C0", ls="none")
+    plt.errorbar(cores[idx_384], t_total[idx_384], yerr=std_total[idx_384],
+                 marker="o", markersize=7, color="C0", ls="none")
 
-    plt.errorbar(cores[idx_OMP_384], t_total[idx_OMP_384], yerr=std_total[idx_OMP_384],
-                 marker="o", markeredgecolor="C0", markersize=7, markeredgewidth=1.25,
-                 color="C0", ls="none")
-
-    plt.plot(cores[idx_384], t_total[idx_384], color="C0", ls="-", lw=1.5)
+    #plt.plot(cores[idx_384], t_total[idx_384], color="C0", ls="-", lw=1.5)
 
 
     idx_768 = width[:] == 768
-    idx_MPI_768 = np.logical_and(width[:] == 768, N_omp[:] == 1)
-    idx_OMP_768 = np.logical_and(width[:] == 768, N_omp[:] != 1)
 
-    plt.errorbar(cores[idx_MPI_768], t_total[idx_MPI_768], yerr=std_total[idx_MPI_768],
-                 marker="^", markeredgecolor="C0", markerfacecolor="none",
-                 markersize=7, markeredgewidth=1.25,
-                 color="C0", ls="none")
+    plt.errorbar(cores[idx_768], t_total[idx_768], yerr=std_total[idx_768],
+                 marker="^", markersize=7, color="C0", ls="none")
 
-    plt.errorbar(cores[idx_OMP_768], t_total[idx_OMP_768], yerr=std_total[idx_OMP_768],
-                 marker="^", markeredgecolor="C0", markersize=7, markeredgewidth=1.25,
-                 color="C0", ls="none")
-
-    plt.plot(cores[idx_768], t_total[idx_768], color="C0", ls="-", lw=1.5)
[…]

### 2017-10-14 astronum_2017/scaling/castro/sedov_scaling.py
more legends

@@ -36,7 +36,7 @@ for row in data:
 sizes = set([int(q.nzones) for q in runs])
 levels = set([int(q.max_level) for q in runs])
 
-markers = ["o", "^", "s"]
+markers = ["o", "s", "^"]
 
 for nl in levels:
     
@@ -70,8 +70,20 @@ for i, nz in enumerate(sizes):
     legs.append(plt.Line2D((0,1),(0,0), color=color))
     legnames.append(r"${}^3$".format(nz))
 
+legs.append(plt.Line2D((0,1),(0,0), color="k",
+                       marker="o", markeredgecolor="k", markerfacecolor="k", linestyle="none"))
+legnames.append("no AMR")
+
+legs.append(plt.Line2D((0,1),(0,0), color="k",
+                       marker="s", markeredgecolor="k", markerfacecolor="k",  linestyle="none"))
+legnames.append("base + one 2x level")
+
+legs.append(plt.Line2D((0,1),(0,0), color="k",
+                       marker="^", markeredgecolor="k", markerfacecolor="k",  linestyle="none"))
+legnames.append("base + one 4x level")
+
 plt.legend(legs, legnames, frameon=False,
-           fontsize="small", numpoints=1, loc=3)
+           fontsize="11", numpoints=1, loc=3, ncol=2)
 
 
 ax = plt.gca()

### 2017-10-14 astronum_2017/scaling/castro/wdmerger_scaling.py

@@ -74,8 +74,16 @@ for i, nz in enumerate(sizes):
     legs.append(plt.Line2D((0,1),(0,0), color=color))
     legnames.append(r"${}^3$".format(nz))
 
+legs.append(plt.Line2D((0,1),(0,0), color="k",
+                       marker="o", markeredgecolor="k", markerfacecolor="k", linestyle="none"))
+legnames.append("no AMR")
+
+legs.append(plt.Line2D((0,1),(0,0), color="k",
+                       marker="^", markeredgecolor="k", markerfacecolor="k",  linestyle="none"))
+legnames.append("base + one 4x level")
+
 plt.legend(legs, legnames, frameon=False,
-           fontsize="small", numpoints=1, loc=3)
+           fontsize="11", numpoints=1, loc=3, ncol=2)
 
 ax = plt.gca()
 plt.text(0.95, 0.95, "Castro 3-d wdmerger (hydro + Poisson gravity)", 

### 2017-10-14 astronum_2017/astronum-amrex.tex
some more scaling discussion

@@
 
-Figure~\ref{fig:maestro_scaling} shows strong scaling for \maestro\ on the 3-d
-XRB problem.  This is a typical \maestro\ application~\cite{xrb3}, where burning
-can be a significant part of the overall evolution.  We ran on both
-OLCF titan and NERSC edison.  We see that the code scales well to
-$\mathcal{O}(10^4)$ processors.  The main limitation to the scaling at
-the moment are the multigrid solves used to enforce the projection (in
-particular the nodal solver).  \maestro\ also does not yet take
-advantage of the tiling capabilities offered in \amrex.  These will be
-addressed in the future as we port \maestro\ to the
-\cpp\ \amrex\ codebase to take advantage of ongoing development.
+Figure~\ref{fig:maestro_scaling} shows strong scaling for \maestro\ on
+the 3-d XRB problem.  This is a typical
+\maestro\ application~\cite{xrb3}, where burning can be a significant
+part of the overall evolution.  We ran on both OLCF titan and NERSC
+edison.  We see that the code scales well to $\mathcal{O}(10^4)$
+processors.  The upturn at the end of the scaling curves on titan
+reflect the change from 1 MPI task / 8 OpenMP threads per NUMA node to
+1 MPI task / 16 OpenMP threads per compute node (2 NUMA nodes).  The
+main limitation to the scaling at the moment are the multigrid solves
+used to enforce the projection (in particular the nodal solver).
+\maestro\ also does not yet take advantage of the tiling capabilities
+offered in \amrex.  These will be addressed in the future as we port
+\maestro\ to the \cpp\ \amrex\ codebase to take advantage of ongoing
+development.
 
@@
 problem.  All tests were run on OLCF titan.  The AMD processors in
-titan can be used in a 8 or 16 cores / node configuration.  For all
+titan can be used in an 8 or 16 cores / node configuration.  For all
 these runs, we ran in the 16 cores / node mode.  All runs also used

### 2017-10-14 astronum_2017/astronum-amrex.tex
typo

@@
   \maestro\ is using an older pure-Fortran interface, but this will be
-  ported to the \cpp\ AMReX this coming year.}---this allows us to take
-advantage of the strengths of both languages.AMReX supports subcycling
+  ported to the \cpp\ \amrex\ this coming year.}---this allows us to take
+advantage of the strengths of both languages. \amrex\ supports subcycling
 in time, which we use in \castro.

### 2017-10-16 astronum_2017/scaling/cori/cori_wdmerger_scaling.py
update legend

@@ -108,7 +108,7 @@ legnames.append("no AMR")
 
 legs.append(plt.Line2D((0,1),(0,0), color="k",
                        marker="^", markeredgecolor="k", markerfacecolor="k",  linestyle="none"))
-legnames.append("base + one 2x level")
+legnames.append("base + one 4x level")
 
 plt.legend(legs, legnames, frameon=False,
            fontsize="11", numpoints=1, loc=3, ncol=3)

### 2017-10-19 astronum_2017/astronum-amrex.tex
add NERSC affil

@@
         J.~B. Bell$^2$,
-        B. Friesen$^2$,
-        A.~M. Jacobs$^3$,
-        M.~P. Katz$^4$,
-        C.~M. Malone$^5$,
+        B. Friesen$^3,2$,
+        A.~M. Jacobs$^4$,
+        M.~P. Katz$^5$,
+        C.~M. Malone$^6$,
         A.~J. Nonaka$^2$,
@@
 
-\address{$^3$Department of Physics and Astronomy, Michigan State
+\address{$^3$National Energy Research Scientific Computing Center,
+  Lawrence Berkeley National Lab, Berkeley, CA 94720 USA}
+
+\address{$^4$Department of Physics and Astronomy, Michigan State
   University, East Lansing, Michigan 48824 USA}
 
-\address{$^4$Nvidia Corporation}
+\address{$^5$Nvidia Corporation}
 
-\address{$^5$Los Alamos National Laboratory, Los Alamos, NM, 87545 USA}
+\address{$^6$Los Alamos National Laboratory, Los Alamos, NM, 87545 USA}
 

### 2017-10-20 astronum_2017/astronum-amrex.tex
John's edits

--- 2017-10-14 astronum_2017/astronum-amrex.tex
@@
 there is a lot of diversity in SNe Ia, and searches for progenitor
-systems have not shown that Chandra-mass white dwarfs can explain most
-SNe Ia.  Today, merging white dwarfs (the double degenerate scenario)
-have perhaps become the most popular model.  Other progenitors, like
-He burning on the surface of a sub-Chandra white dwarf have also seen
-interest in explaining some of the observed diversity.
-See~\cite{araa-maoz} for a review.
+systems have strongly suggested that Chandra-mass white dwarfs cannot
+explain most SNe Ia.  Today, merging white dwarfs (the double
+degenerate scenario) have perhaps become the most popular model.
+Other progenitors, like He burning on the surface of a sub-Chandra
+white dwarf have also seen interest in explaining some of the observed
+diversity.  See~\cite{araa-maoz} for a review.
 
-There are open questions in all of these scenarios which can be
+There are open questions in all of these scenarios that can be
 addressed through simulation.  For the Chandra and sub-Chandra models,
@@
 and parallel communications, and it calls the computational kernels on
-a patch-by-patch basis.  The core library is written in \cpp\ with the
+a patch-by-patch basis.  The core library is written in \cpp\ with 
 computational kernels written in Fortran\footnote{Currently
@@
 coarse-grained approach to distributing the computational work, while
-threading of loops over zones in a grid using OpenMP provides
-effective fine-grained parallelization.  For \castro\, tiling is used
-to logically subdivide AMR patches to increase the parallelism for
-OpenMP, without incurring the memory overheads that using many more
-smaller grids would bring~\cite{tiling}.  This strategy is especially
-important for many core architectures like the Intel Phi.  Ongoing
-development is being done to support GPU offloading, using managed
-memory provided by the latest generations of GPUs.
+subdividing patches into tiles and using threads to parallelize over
+tiles using OpenMP provides effective fine-scale parallelization and
+amortizes thread overhead over large units of work.  For \castro\,
+this tiling logically subdivides AMR patches to increase the
+parallelism for OpenMP, without incurring a large increase in metadata
+that would be associated with using smaller patches in a flat MPI
+mode~\cite{tiling}.  This strategy is especially important for many
+core architectures like the Intel Phi.  Ongoing development is being
+done to support GPU offloading, using managed memory provided by the
+latest generations of GPUs.
 
 There are many application codes built on \amrex, including those in
-combustion, ... \MarginPar{domains?} In astrophysics, these include \maestro\ and
-\castro\ for stellar and nuclear astrophysics applications and
-\nyx~\cite{nyx} for cosmological applications.  Here we focus on the
-former two.
+combustion, multiphase flow, accelerator design, and microfluidics. In
+astrophysics, these include \maestro\ and \castro\ for stellar and
+nuclear astrophysics applications and \nyx~\cite{nyx} for cosmological
+applications.  Here we focus on the former two.
 
@@
 velocities are then projected onto the space that satisfies the
-divergence constraint.  The projections involve elliptic solves,
-solving a variable-coefficient Poisson problem, computed using a
-multigrid algorithm.  Low speed methods have seen a lot of recent
-development~\cite{kleinpauluis,vasil:2013}, which has been
+divergence constraint.  The projections require solving a variable
+coefficient elliptic equations, which is done numerically using a
+multigrid algorithm.  A number of recent advances in low Mach
+number modeling~\cite{kleinpauluis,vasil:2013} have been
 incorporated into \maestro.
@@
 introduced by \cite{miniati-colella}.  Recent developments in
-\castro\ include a deferred-correction method of coupling
+\castro\ include a spectral-deferred correction method of coupling
 hydrodynamics and reactions, a conservative gravity and rotation
@@
   for making them available.}.  The reaction networks are written such
-that that rates and integration strategy are decoupled, allowing us to
-change the integration strategy for a set of rates.  They are also
-written to be threadsafe and with GPUs in mind (more on that below).
-The goal of \starkiller\ is to create a set of community microphysics
-solvers that can be used in a variety of nuclear astrophysics codes,
-not just those discussed here.
+that the rates and integration strategy are decoupled, allowing us to
+change the integration strategy for a given set of rates.  They are
+also written to be threadsafe and with GPUs in mind (more on that
+below).  The goal of \starkiller\ is to create a set of community
+microphysics solvers that can be used in a variety of nuclear
+astrophysics codes, not just those discussed here.
 
@@
   excellent strong and weak scaling for the single level runs.  For
-  the $1024^3$ run, we also ran with one $2\times$ jump level of
-  refinment (triangles) and see good strong scaling.  The variability
-  (shown by the error bars) at high core counts shows we are becoming
-  work-starved. (right) \castro\ strong scaling on OLCF titan for a
-  hydro + Poisson gravity (wdmerger) problem.  This problem uses a
-  multipole solver to determine Dirichlet boundary conditions
-  representing an isolated mass distribution, and then geometric
-  multigrid to solve for the potential in the interior.  Two coarse
-  grid sizes are shown, and demonstrate great strong scaling.  We also
-  look at a single level of refinement (a $4\times$ jump) on top of
-  this coarse grid.}
+  the $1024^3$ run, we also ran with one level of refinment by a
+  factor of 2 (triangles) or a factor of 4 (squares), and see good
+  strong scaling.  The variability (shown by the error bars) at high
+  core counts shows we are becoming work-starved. (right)
+  \castro\ strong scaling on OLCF titan for a hydro + Poisson gravity
+  (wdmerger) problem.  This problem uses a multipole solver to
+  determine Dirichlet boundary conditions representing an isolated
+  mass distribution, and then geometric multigrid to solve for the
+  potential in the interior.  Two coarse grid sizes are shown, and
+  demonstrate great strong scaling.  We also look at a single level of
+  refinement (by a factor of 4) on top of this coarse grid.}
 \end{figure}
@@
 1 MPI task / 16 OpenMP threads per compute node (2 NUMA nodes).  The
-main limitation to the scaling at the moment are the multigrid solves
+main limitation to the scaling at the moment is the multigrid solves
 used to enforce the projection (in particular the nodal solver).
@@
 pure hydro problem.  For the self-gravity problem, we scale well with
-a uniform grid, but with a $4\times$ jump for the AMR level on top of
-this, we do worse at higher core counts.  This is a challenging
-problem, as only 3\% of the domain volume is refined.  For a given
-core count, there are a number of different combinations of MPI ranks
-and OpenMP threads we can use.  In general, with multilevel problems,
-we found the best performance with fewer MPI ranks and more OpenMP
-threads.
+a uniform grid, but with a factor of 4 refinement for the AMR level on
+top of this, we observe degraded performance at higher core counts.
+This is a challenging problem, as only 3\% of the domain volume is
+refined.  For a given core count, there are a number of different
+combinations of MPI ranks and OpenMP threads we can use.  In general,
+with multilevel problems, we found the best performance with fewer MPI
+ranks and more OpenMP threads.
 
@@
 to handle data motion to and from the device, so that the compute
-kernels can operate on data which is presumed to already be there, and
+kernels can operate on data that is presumed to already be there, and
 the computation is decoupled from the memory management. Computation
@@
 righthand size and Jacobian evaluations and the timestepping itself is
-done on the GPU and once all zones are burned, we access the data as
+done on the GPU and once the burning in all zones is completed, we access the data as
 needed on the CPU.  To enable this, we ported our workhorse ODE
@@
 
-We described our suite of astrophysics codes built on the
+We have described our suite of astrophysics codes built on the
 \amrex\ block-structured adaptive mesh refinement framework.  These
@@
 \ack The work at Stony Brook was supported by DOE/Office of Nuclear
-Physics grant DE-FG02-87ER40317 and NSF award AST-1211563.  An award
-of computer time was provided by the Innovative and Novel
-Computational Impact on Theory and Experiment (INCITE) program.  This
-research used resources of the Oak Ridge Leadership Computing Facility
-at the Oak Ridge National Laboratory, which is supported by the Office
-of Science of the U.S. Department of Energy under Contract
+Physics grant DE-FG02-87ER40317 and NSF award AST-1211563.  The work
+at LBNL was supported by the DOE Office of Advanced Scientific
+Computing Research under Contract No, DE-AC02-05CH11231. An award of
+computer time was provided by the Innovative and Novel Computational
+Impact on Theory and Experiment (INCITE) program.  This research used
+resources of the Oak Ridge Leadership Computing Facility at the Oak
+Ridge National Laboratory, which is supported by the Office of Science
+of the U.S. Department of Energy under Contract
 No.\ DE-AC05-00OR22725.  We appreciate the efforts of the OLCF (and in

### 2017-10-20 astronum_2017/astronum-amrex.tex
caption fix
some small edits

@@
 \castro, designed to perform three-dimensional models of the early
-subsonic evolution leading to runaway and the subsequent expolosion.
+subsonic evolution leading to runaway and the subsequent explosion.
 Together this suite of codes allows us to address many problems in
@@
 pressure response.  Most of the current models for these events are
-characterized by a long timescale ``simmmering'' phase where reactions
+characterized by a long timescale ``simmering'' phase where reactions
 heat the star or layer and drive convection.  Eventually, reactions
@@
 possible to create a detonation in the thin surface He layer?  For
-double degenerates, does the burning take place promptly? and can we
-avoid an accretion-induced collapse to a neutron star.  For many of
-these, we also need to address numerical issues, such as whether it is
+double degenerates, does the burning take place promptly? Can we
+avoid an accretion-induced collapse to a neutron star?  For many of
+these investigations, we need to address numerical issues such as 
+whether it is
 possible to accurately model the ignition of a detonation with the
-spatial resolution we can attain?  These are some of the questions we
+spatial resolution we can attain.  These are some of the questions we
 seek to answer.
@@
 computational kernels written in Fortran\footnote{Currently
-  \maestro\ is using an older pure-Fortran interface, but this will be
-  ported to the \cpp\ \amrex\ this coming year.}---this allows us to take
+  \maestro\ is written in pure-Fortran, but will be
+  ported to the updated \cpp\ \amrex\ framework this coming year.}---this allows us to take
 advantage of the strengths of both languages. \amrex\ supports subcycling
@@
 tiles using OpenMP provides effective fine-scale parallelization and
-amortizes thread overhead over large units of work.  For \castro\,
-this tiling logically subdivides AMR patches to increase the
-parallelism for OpenMP, without incurring a large increase in metadata
-that would be associated with using smaller patches in a flat MPI
-mode~\cite{tiling}.  This strategy is especially important for many
-core architectures like the Intel Phi.  Ongoing development is being
+amortizes thread overhead over large units of work.  Additionally,
+tiling does not incur the large increase in metadata
+associated with using smaller patches in a flat MPI
+mode~\cite{tiling} since we can use fewer, larger patches.
+This strategy is especially important for
+many-core architectures like the Intel Phi.  Ongoing development is being
 done to support GPU offloading, using managed memory provided by the
@@
 applications.  Here we focus on the former two.
+\MarginPar{ajn - I want to put a grid picture in this doc somewhere}
 
@@
 hydrodynamics code designed for efficiently modeling convection in
-stars.  \maestro\ decomposes the state into a one-dimensional
+stars.  \maestro\ decomposes the state variables into a one-dimensional
 hydrostatic base state and a three-dimensional Cartesian state that
-evolves the departures from HSE.  A constraint equation is derived by
+models the deviation from hydrostatic equilibrium.
+A constraint equation is derived by
 requiring that the pressure everywhere be close to the background
-hydrostatic state.  The constraint acts to enforce instantaneous
-acoustic equilibriation, effectively filtering soundwaves from the
+hydrostatic pressure.  The constraint acts to enforce instantaneous
+acoustic equilibration, effectively filtering soundwaves from the
 system, while retaining the compressibility effects due to the
@@
 feasible, the git hashes for the published results are included in
-paper acknowledgements.
+paper acknowledgments.
 
@@
 A key design goal of our application codes is performance portability.
-We want the same kernels to run on clusters, manycore machines (like
+We want the same kernels to run on clusters, manycore machines (e.g.
 Intel KNL), and GPU-based machines.  Our development has balanced this
@@
   excellent strong and weak scaling for the single level runs.  For
-  the $1024^3$ run, we also ran with one level of refinment by a
+  the $1024^3$ run, we also ran with one level of refinement by a
   factor of 2 (triangles) or a factor of 4 (squares), and see good
@@
   GPUs in this test of a pure hydro problem (Sedov w/ real EOS).  On the
-right, we explore performance on NERSC Cori (and Intel KNL platform).}
+right, we explore performance on NERSC cori (and Intel KNL platform).}
 \end{figure}
@@
 used to enforce the projection (in particular the nodal solver).
-\maestro\ also does not yet take advantage of the tiling capabilities
-offered in \amrex.  These will be addressed in the future as we port
-\maestro\ to the \cpp\ \amrex\ codebase to take advantage of ongoing
-development.
+Also, \maestro\ currently uses a simpler fine-grained approach to parallelism
+where planes in the $z$-direction are divided among OpenMP threads.
+As we port \maestro\ to the \cpp\ \amrex\ code base we will take
+advantage of ongoing development for increased tile size control.
 
-Figure~\ref{fig:castro-scaling} shows scaling behavior for a pure
+Figure~\ref{fig:castro-scaling} shows \castro\ scaling behavior for a pure
 hydro + real EOS problem and a hydro / real EOS, + self-gravity
@@
   into {\tt development}.}.  In \amrex, grids consist of a number of
-boxes that enable domain decomposition for parallelism.  An iterator
-loops over all of the boxes at the same level of refinement and passes
+boxes that enable domain decomposition for parallelism.\MarginPar{unclear what the difference between a grid and box is here}
+An iterator loops over all of the boxes at the same level of refinement and passes
 a data pointer into a Fortran kernel function where the work is
@@
 also shows our performance on the Intel KNL manycore processors (using
-NERSC Cori).  This is for the same problem as the self-gravity test on
-titan. \MarginPar{need more info: base grid?}  The KNL chips in Cori
+NERSC cori).  This is for the same problem as the self-gravity test on
+titan. \MarginPar{need more info: base grid?}  The KNL chips in cori
 have 68 cores than can be run with 1, 2, or 4 hardware threads each.
@@
   reaction networks.  The 13-isotope alpha-chain is a
-  potential network for the sub-Chandra simulations we propose here.
+  potential network for the sub-Chandra simulations.
   The 9-isotope Urca network is used for our Urca simulations.  A
@@
 dwarf and He layer, including a global He-nova-like behavior where the
-entire layer runsway together and localized
-burning~\cite{subchandra2}.  Work continues to characterize the ignition.
+entire layer runs way together and localized
+burning~\cite{subchandra2}.\MarginPar{incomplete sentence}
+We are performing further studies to characterize the ignition.
 

### 2017-10-20 astronum_2017/astronum-amrex.tex
added AMR grid structure

@@
 \maestro\ is written in pure-Fortran, but will be
-  ported to the updated \cpp\ \amrex\ framework this coming year.}---this allows us to take
+ ported to the updated \cpp\ \amrex\ framework this coming year.}
+---this allows us to take
 advantage of the strengths of both languages. \amrex\ supports subcycling
 in time, which we use in \castro.
+\begin{figure}[htb]
+\centering
+\includegraphics[width=0.48\linewidth]{ppsn-amr.png}
+\begin{minipage}[b]{0.48\linewidth}
+\caption{\label{fig:amr_grid} Colormap of density and the associated AMR 
+grid structure in a pulsational pair-instability supernova simulation.  
+Image courtesy of Ken Chen \cite{chen2014two}.}
+\end{minipage}
+\end{figure}
 
@@
 applications.  Here we focus on the former two.
-\MarginPar{ajn - I want to put a grid picture in this doc somewhere}
 
@@
 
-\begin{figure}[t]
+\begin{figure}[htb]
 \centering
@@
 
-\begin{figure}[t]
+\begin{figure}[htb]
 \centering \includegraphics[height=1.8in]{subch_h}

### 2017-10-20 astronum_2017/astronum-amrex.tex
reset the figures to the top
add plot

@@
 in time, which we use in \castro.
-\begin{figure}[htb]
+\begin{figure}[t]
 \centering
-\includegraphics[width=0.48\linewidth]{ppsn-amr.png}
-\begin{minipage}[b]{0.48\linewidth}
-\caption{\label{fig:amr_grid} Colormap of density and the associated AMR 
-grid structure in a pulsational pair-instability supernova simulation.  
-Image courtesy of Ken Chen \cite{chen2014two}.}
+\includegraphics[width=0.58\linewidth]{slice_grid}
+\begin{minipage}[b]{0.38\linewidth}
+\caption{\label{fig:amr_grid} A slice through a
+  \castro\ three-dimensional simulation of merging white dwarfs
+  showing the 2-level grid structure.  The boxes making up the base
+  grid are drawn in white and the boxes making up the finer grid are drawn in black.  Note
+  that the finer boxes can overlap multiple coarse boxes.\vspace{2em}}
 \end{minipage}
@@
 
-\begin{figure}[htb]
+\begin{figure}[t]
 \centering
@@
 
-\begin{figure}[htb]
+\begin{figure}[t]
 \centering \includegraphics[height=1.8in]{subch_h}

### 2017-10-21 astronum_2017/astronum-amrex.tex
update with Brian's comments

@@
         J.~B. Bell$^2$,
-        B. Friesen$^3,2$,
+        B. Friesen$^{3,2}$,
         A.~M. Jacobs$^4$,
@@
 radiation and magnetic fields.  Further, these environments are
-characterized by a wide range of lengthscales, from the size of the
+characterized by a wide range of length scales, from the size of the
 star or binary system down to the burning zone width and dissipation
@@
 occurs over 10s of millions to billions of years, the simmering phase
-leading up to explosions lasts millenia to days or hours, and the
+leading up to explosions lasts hours or days to millenia, and the
 explosion can be over in seconds to hours.  The radiation leakage,
@@
 
-The largest open theoretical question for SNe Ia is---what is the
-nature of the progenitor?  About 20 years ago, the community had
-mostly converged up the single-degenerate scenario---a
+Among the most significant open questions for SNe Ia is the identity of the
+progenitor.  About 20 years ago, the community had
+mostly converged upon the single-degenerate scenario---a
 Chandrasekhar-mass C/O white dwarf that accretes from its companion,
@@
 star (see~\cite{hillebrandtniemeyer2000} for the state of the field at
-that time).  Since then, the wealth of observations has show that
+that time).  Since then, the wealth of observations has indicated that
 there is a lot of diversity in SNe Ia, and searches for progenitor
@@
 Other progenitors, like He burning on the surface of a sub-Chandra
-white dwarf have also seen interest in explaining some of the observed
+white dwarf, have also seen interest in explaining some of the observed
 diversity.  See~\cite{araa-maoz} for a review.
@@
 by turbulent convection that give rise to burning fronts?  A
-longstanding question with the Chandra model is, can a deflagration
-transition to a detonation during the burning front propagation
+longstanding question with the Chandra model is whether a deflagration
+can transition to a detonation during the burning front propagation
 through the star.  For the sub-Chandra double detonation model, is it
-possible to create a detonation in the thin surface He layer?  For
-double degenerates, does the burning take place promptly? Can we
-avoid an accretion-induced collapse to a neutron star?  For many of
-these investigations, we need to address numerical issues such as 
-whether it is
-possible to accurately model the ignition of a detonation with the
-spatial resolution we can attain.  These are some of the questions we
-seek to answer.
+not clear whether it is possible to create a detonation in the thin
+surface He layer.  For double degenerates, it is still unresolved
+whether the burning takes place promptly or after some delay.  However
+it proceeds, we need to avoid an accretion-induced collapse to a
+neutron star.  For many of these investigations, we need to address
+numerical issues such as whether it is possible to accurately model
+the ignition of a detonation with the spatial resolution we can
+attain.  These are some of the questions we seek to answer.
 
@@
   grid are drawn in white and the boxes making up the finer grid are drawn in black.  Note
-  that the finer boxes can overlap multiple coarse boxes.\vspace{2em}}
+  that the finer boxes can overlap multiple coarse boxes, and not all
+  boxes are the same size.\vspace{2em}}
 \end{minipage}
@@
 This strategy is especially important for
-many-core architectures like the Intel Phi.  Ongoing development is being
+many-core architectures like the Intel Xeon Phi.  Ongoing development is being
 done to support GPU offloading, using managed memory provided by the
-latest generations of GPUs.
+latest generations of GPUs.  Figure~\ref{fig:amr_grid} shows an
+example of a 2-level grid.
 
@@
 We want the same kernels to run on clusters, manycore machines (e.g.
-Intel KNL), and GPU-based machines.  Our development has balanced this
+Intel Xeon Phi), and GPU-based machines.  Our development has balanced this
 need with architecture-specific optimizations to maximize code reuse.
@@
 \caption{\label{fig:maestro_scaling} \maestro\ strong scaling on NERSC
-  edison and OLCF titan for a 3-d XRB problem.  Two different problem
+  Edison and OLCF Titan for a 3-d XRB problem.  Two different problem
   sizes are shown.  We see excellent strong scaling to high core counts
@@
 \caption{\label{fig:castro-scaling} (left) \castro\ strong scaling on
-  OLCF titan for a pure hydro (Sedov w/ real EOS) problem.  The
+  OLCF Titan for a pure hydro (Sedov w/ real EOS) problem.  The
   different colors represent different base resolutions.  We see
@@
   core counts shows we are becoming work-starved. (right)
-  \castro\ strong scaling on OLCF titan for a hydro + Poisson gravity
+  \castro\ strong scaling on OLCF Titan for a hydro + Poisson gravity
   (wdmerger) problem.  This problem uses a multipole solver to
@@
   GPUs in this test of a pure hydro problem (Sedov w/ real EOS).  On the
-right, we explore performance on NERSC cori (and Intel KNL platform).}
+right, we explore performance on NERSC Cori (and Intel Xeon Phi platform).}
 \end{figure}
@@
 \maestro\ application~\cite{xrb3}, where burning can be a significant
-part of the overall evolution.  We ran on both OLCF titan and NERSC
-edison.  We see that the code scales well to $\mathcal{O}(10^4)$
-processors.  The upturn at the end of the scaling curves on titan
+part of the overall evolution.  We ran on both OLCF Titan and NERSC
+Edison.  We see that the code scales well to $\mathcal{O}(10^4)$
+processors.  The upturn at the end of the scaling curves on Titan
 reflect the change from 1 MPI task / 8 OpenMP threads per NUMA node to
@@
 hydro + real EOS problem and a hydro / real EOS, + self-gravity
-problem.  All tests were run on OLCF titan.  The AMD processors in
-titan can be used in an 8 or 16 cores / node configuration.  For all
+problem.  All tests were run on OLCF Titan.  The AMD processors in
+Titan can be used in an 8 or 16 cores / node configuration.  For all
 these runs, we ran in the 16 cores / node mode.  All runs also used
@@
 complete the port of the GPU developments into \castro.  This figure
-also shows our performance on the Intel KNL manycore processors (using
-NERSC cori).  This is for the same problem as the self-gravity test on
-titan. \MarginPar{need more info: base grid?}  The KNL chips in cori
+also shows our performance on the Intel Xeon Phi manycore processors (using
+NERSC Cori).  This is for the same problem as the self-gravity test on
+Titan.  The Intel Xeon Phi chips in Cori
 have 68 cores than can be run with 1, 2, or 4 hardware threads each.
-For all these runs, we ran with 4 threads / core (effectively 272
+For all these runs, we ran with 4 threads / core (272
 threads / chip).
@@
 \end{figure}
+
+\MarginPar{we should update this to a more modern processor---Don can you run
+this on summitdev?}
 

### 2017-10-24 astronum_2017/astronum-amrex.tex
wordsmithing

@@
 \castro, designed to perform three-dimensional models of the early
-subsonic evolution leading to runaway and the subsequent explosion.
-Together this suite of codes allows us to address many problems in
-stellar and nuclear astrophysics.  We describe some of the design
-details, the current architecture of the code, and some applications
-below.
+subsonic evolution leading to runaway and the subsequent explosion,
+respectively.  Together this suite of codes allows us to address many
+problems in stellar and nuclear astrophysics.  We describe some of the
+design details, the current architecture of the code, and some
+applications below.
 
@@
 addressed through simulation.  For the Chandra and sub-Chandra models,
-what is the distribution (spatial and temporal) of the hotspots setup
+what is the distribution (spatial and temporal) of the hotspots set up
 by turbulent convection that give rise to burning fronts?  A
@@
 
-\subsection{XRBs}
+\subsection{X-ray Bursts}
 
-X-ray bursts, the burning of accreted H/He on the surface of a neutron
-star, can be important probes of neutron star structure.  Interpreting
-observations requires that we understand what we are seeing, which can
-be influenced by the products of the burning and how the burning
-spreads across the star.  Many efforts have focused on different
-aspects of these events.  One-dimensional models capture the
+X-ray bursts --- the burning of accreted H/He on the surface of a
+neutron star --- can be important probes of neutron star structure.
+Interpreting observations requires that we understand what we are
+seeing, which can be influenced by the products of the burning and how
+the burning spreads across the star.  Many efforts have focused on
+different aspects of these events.  One-dimensional models capture the
 energetics well and inform us about the nucleosynthesis
@@
 scale of the burning to a reasonable fraction of the neutron star
-surface has prevented detailed explorations of the burning in resolved
-calculations and how it feeds back on the flame structure and
-propagation.  Algorithms and computing architectures are starting to
+surface has prevented detailed explorations of the burning and how it
+feeds back on the flame structure and propagation in resolved
+calculations.  Algorithms and computing architectures are starting to
 reach the point where we can span the gaps in spatial scales between
 these calculations to provide a better understanding of the ignition
-and propagation of the burning front, and the nucleosynthesis produced.
+and propagation of the burning front, and the nucleosynthesis
+produced.
 
@@
 the dynamics.  Conservation is important as well, suggesting approaches
-which implement gravitational and rotation sources conservatively, and
+that implement gravitational and rotation sources conservatively, and
 methods for improving angular momentum conservation.
@@
 divergence constraint.  The projections require solving a variable
-coefficient elliptic equations, which is done numerically using a
-multigrid algorithm.  A number of recent advances in low Mach
-number modeling~\cite{kleinpauluis,vasil:2013} have been
-incorporated into \maestro.
+coefficient elliptic equation, which is done numerically using a
+multigrid algorithm.  A number of recent advances in low Mach number
+modeling~\cite{kleinpauluis,vasil:2013} have been incorporated into
+\maestro.
 
 \maestro\ has been applied to convection in the Chandrasekhar-mass
-model for Type Ia supernovae (SNe Ia)
-\cite{ZABNW:IV,wdconvect,wdturb}, the sub-Chandra model for SNe Ia
-\cite{subchandra,subchandra2} X-ray bursts~\cite{xrb,xrb2,xrb3}, and
-convection in massive stars \cite{ms_cc}.
+model for SNe Ia \cite{ZABNW:IV,wdconvect,wdturb}, the sub-Chandra
+model for SNe Ia \cite{subchandra,subchandra2},
+XRBs~\cite{xrb,xrb2,xrb3}, and convection in massive stars
+\cite{ms_cc}.
 
@@
 source formulation~\cite{wdmergerI}, and retry mechanism to redo a
-step based on criteria evaluated during the integration.\MarginPar{more new stuff?}
+step based on criteria evaluated during the
+integration.\MarginPar{more new stuff?}
 
@@
 improving energy conservation with gravitational and rotational
-sources and new integration methods that more accurately couple hydro
+sources, and new integration methods that more accurately couple hydro
 and reactions.  For \maestro\ simulations that evolve from the
@@
 \maestro\ and \castro\ share the same microphysics, available as the
-\starkiller\ astrophysics github
+\starkiller\ astrophysics GitHub
 project\footnote{\url{https://github.com/StarKiller-astro/microphysics/}}.
@@
 development model---the development git repos are hosted on
-github\footnote{\url{https://github.com/AMReX-Astro/}}, available for
+GitHub\footnote{\url{https://github.com/AMReX-Astro/}}, available for
 anyone to see and contribute to using issues and pull-requests.
@@
 
-Figure~\ref{fig:castro-scaling} shows \castro\ scaling behavior for a pure
-hydro + real EOS problem and a hydro / real EOS, + self-gravity
-problem.  All tests were run on OLCF Titan.  The AMD processors in
-Titan can be used in an 8 or 16 cores / node configuration.  For all
-these runs, we ran in the 16 cores / node mode.  All runs also used
-the PGI 17.7 compilers.  We see excellent scaling behavior for the
-pure hydro problem.  For the self-gravity problem, we scale well with
-a uniform grid, but with a factor of 4 refinement for the AMR level on
-top of this, we observe degraded performance at higher core counts.
-This is a challenging problem, as only 3\% of the domain volume is
-refined.  For a given core count, there are a number of different
-combinations of MPI ranks and OpenMP threads we can use.  In general,
-with multilevel problems, we found the best performance with fewer MPI
-ranks and more OpenMP threads.
+Figure~\ref{fig:castro-scaling} shows \castro\ scaling behavior for a
+pure hydro problem with a real EOS and a hydro + self-gravity problem
+with a real EOS.  All tests were run on OLCF Titan.  The AMD
+processors in Titan can be used in an 8 or 16 cores / node
+configuration.  For all these runs, we ran in the 16 cores / node
+mode.  All runs also used the PGI 17.7 compilers.  We see excellent
+scaling behavior for the pure hydro problem.  For the self-gravity
+problem, we scale well with a uniform grid, but with a factor of 4
+refinement for the AMR level on top of this, we observe degraded
+performance at higher core counts.  This is a challenging problem, as
+only 3\% of the domain volume is refined.  For a given core count,
+there are a number of different combinations of MPI ranks and OpenMP
+threads we can use.  In general, with multilevel problems, we found
+the best performance with fewer MPI ranks and more OpenMP threads.
 
@@
 
-A parallel effort is porting our microphysics, in particular reaction
-networks to GPUs.  Our strategy is to do the entire ODE integration on
-the GPU, i.e., the data for a patch of zones is passed to the GPU, all
-righthand size and Jacobian evaluations and the timestepping itself is
-done on the GPU and once the burning in all zones is completed, we access the data as
-needed on the CPU.  To enable this, we ported our workhorse ODE
-integrator (VODE~\cite{vode}) to CUDA Fortran. This required extensive
-rewrites of the internals of VODE, which was originally written using
-Fortran 77 syntax unsupported by CUDA Fortran. Accelerating VODE with
-CUDA Fortran has proven successful, and we now see significant
-performance gains with the CUDA version of our reaction networks, even
-for moderate sized networks (a 13-isotope standard network).
-Figure~\ref{fig:cudaode} shows the speed-ups on a GPU vs.\ single CPU
-core.
+A parallel effort is porting our microphysics --- in particular
+reaction networks --- to GPUs.  Our strategy is to do the entire ODE
+integration on the GPU, i.e., the data for a patch of zones is passed
+to the GPU, all righthand size and Jacobian evaluations and the
+timestepping itself is done on the GPU, and once the burning in all
+zones is completed, we access the data as needed on the CPU.  To
+enable this, we ported our workhorse ODE integrator (VODE~\cite{vode})
+to CUDA Fortran. This required extensive rewrites of the internals of
+VODE, which was originally written using Fortran 77 syntax unsupported
+by CUDA Fortran. Accelerating VODE with CUDA Fortran has proven
+successful, and we now see significant performance gains with the CUDA
+version of our reaction networks, even for moderate-sized networks (a
+13-isotope standard network).  Figure~\ref{fig:cudaode} shows the
+speed-ups on a GPU vs.\ single CPU core.
 
@@
 theme of the codes is the open development model, with the code
-development done on github and all problem files needed to recreate
+development done on GitHub and all problem files needed to recreate
 any science results freely available.  Future development efforts for

### 2017-10-24 astronum_2017/astronum-amrex.tex
marginpar about other castro works, if wanted

@@
 supernovae~\cite{castro-pairinstability}, and white dwarf mergers as a
-model for SNe Ia~\cite{wdmergerI}.  Recent work has focused on
+model for SNe Ia~\cite{wdmergerI}. \MarginPar{There are others: Ma's
+  Chandra model of SNe Ia; Lovegrove's shock breakout; Moll's WD
+  merger; Moll's double det in subCh.} Recent work has focused on
 improving energy conservation with gravitational and rotational

### 2017-10-24 astronum_2017/astronum-amrex.tex
removed mention of weak scaling as we never show any of these results

@@
   different colors represent different base resolutions.  We see
-  excellent strong and weak scaling for the single level runs.  For
-  the $1024^3$ run, we also ran with one level of refinement by a
-  factor of 2 (triangles) or a factor of 4 (squares), and see good
-  strong scaling.  The variability (shown by the error bars) at high
-  core counts shows we are becoming work-starved. (right)
-  \castro\ strong scaling on OLCF Titan for a hydro + Poisson gravity
-  (wdmerger) problem.  This problem uses a multipole solver to
-  determine Dirichlet boundary conditions representing an isolated
-  mass distribution, and then geometric multigrid to solve for the
+  excellent strong scaling for the single level runs.  For the
+  $1024^3$ run, we also ran with one level of refinement by a factor
+  of 2 (triangles) or a factor of 4 (squares), and see good strong
+  scaling.  The variability (shown by the error bars) at high core
+  counts shows we are becoming work-starved. (right) \castro\ strong
+  scaling on OLCF Titan for a hydro + Poisson gravity (wdmerger)
+  problem.  This problem uses a multipole solver to determine
+  Dirichlet boundary conditions representing an isolated mass
+  distribution, and then geometric multigrid to solve for the
   potential in the interior.  Two coarse grid sizes are shown, and

### 2017-10-24 astronum_2017/astronum-amrex.tex
address Chris's marginpar

--- 2017-10-24 astronum_2017/astronum-amrex.tex
@@
 \castro\ has been applied to core-collapse
-supernovae~\cite{castro-ccsne}, population III pair-instability
-supernovae~\cite{castro-pairinstability}, and white dwarf mergers as a
-model for SNe Ia~\cite{wdmergerI}.  Recent work has focused on
-improving energy conservation with gravitational and rotational
-sources, and new integration methods that more accurately couple hydro
-and reactions.  For \maestro\ simulations that evolve from the
-subsonic regime to the sonic regime, we have demonstrated the ability
-to restart the calculations in \castro\ to continue the evolution into
-the sonic regime~\cite{scidac-petascale,malone:2014} (in this case,
-for Chandra model SNe Ia).
+supernovae~\cite{castro-ccsne}, radiative shock breakout in
+supernovae~\cite{lovegrove:2017}, population III pair-instability
+supernovae~\cite{castro-pairinstability}, the Chandra model for
+SNe~Ia~\cite{ma:2013}, the sub-Chandra SNe Ia model~\cite{moll:2013}, and white dwarf mergers as a model for SNe
+Ia~\cite{moll:2014,wdmergerI}. Recent work has focused on improving energy
+conservation with gravitational and rotational sources, and new
+integration methods that more accurately couple hydro and reactions.
+For \maestro\ simulations that evolve from the subsonic regime to the
+sonic regime, we have demonstrated the ability to restart the
+calculations in \castro\ to continue the evolution into the sonic
+regime~\cite{scidac-petascale,malone:2014} (in this case, for Chandra
+model SNe Ia).
 

### 2017-10-24 astronum_2017/astronum-amrex.tex
more updates

@@
   is currently on a feature branch in the git repo, awaiting merge
-  into {\tt development}.}.  In \amrex, grids consist of a number of
-boxes that enable domain decomposition for parallelism.\MarginPar{unclear what the difference between a grid and box is here}
-An iterator loops over all of the boxes at the same level of refinement and passes
-a data pointer into a Fortran kernel function where the work is
-done. For the simulation state data that resides in each box, we have
-modified the memory allocator so that it can use a CUDA allocator
-(either device or managed memory).  The domain iterator is configured
-to handle data motion to and from the device, so that the compute
-kernels can operate on data that is presumed to already be there, and
-the computation is decoupled from the memory management. Computation
-on the data can then be performed with OpenACC, CUDA Fortran, or (more
-recently) OpenMP 4.5. We have also built CUDA compute support into
-\amrex\ so that a compute kernel can be transparently operated on
-using CUDA Fortran without substantially modifying the kernels (and we
-anticipate using a similar strategy to use OpenACC and/or OpenMP in
-the future). This helps ensure that we can continue to maintain
-performance portability in our simulation codes, by decoupling the
-physics algorithms from the backend support used to implement them on
-various compute architectures.\MarginPar{add CPU time for reference}
+  into {\tt development}.}.  In \amrex, an iterator loops over all of
+the boxes at the same level of refinement and passes a data pointer
+into a Fortran kernel function where the work is done. For the
+simulation state data that resides in each box, we have modified the
+memory allocator so that it can use a CUDA allocator (either device or
+managed memory).  The domain iterator is configured to handle data
+motion to and from the device, so that the compute kernels can operate
+on data that is presumed to already be there, and the computation is
+decoupled from the memory management. Computation on the data can then
+be performed with OpenACC, CUDA Fortran, or (more recently) OpenMP
+4.5. We have also built CUDA compute support into \amrex\ so that a
+compute kernel can be transparently operated on using CUDA Fortran
+without substantially modifying the kernels (and we anticipate using a
+similar strategy to use OpenACC and/or OpenMP in the future). This
+helps ensure that we can continue to maintain performance portability
+in our simulation codes, by decoupling the physics algorithms from the
+backend support used to implement them on various compute
+architectures.\MarginPar{add CPU time for reference}
 
@@
 times and saw a range of outcomes depending on the mass of the white
-dwarf and He layer, including a global He-nova-like behavior where the
+dwarf and He layer, including a both nova-like behavior where the
 entire layer runs way together and localized
-burning~\cite{subchandra2}.\MarginPar{incomplete sentence}
+burning ignited in a small region~\cite{subchandra2}.
 We are performing further studies to characterize the ignition.
@@
 an Titan X Pascal board from Nvidia Corp through the GPU Grant
-Program.  This research has made use of NASA's Astrophysics Data
-System Bibliographic Services.
+Program.  Visualization were done using \cite{yt}.  This research has
+made use of NASA's Astrophysics Data System Bibliographic Services.
 

### 2017-10-25 astronum_2017/astronum-amrex.tex
update some paper refs

@@
   architectures.  The left figure shows the \castro\ GPU proxy app
-  (\starlord) on OLCF summitdev.  We see good scaling behavior to 64
+  (\starlord) on OLCF Summitdev.  We see good scaling behavior to 64
   GPUs in this test of a pure hydro problem (Sedov w/ real EOS).  On the
@@
 Figure~\ref{fig:knl-gpu-scaling} shows the performance of
-\starlord\ on the summitdev platform at OLCF\footnote{summitdev
+\starlord\ on the Summitdev platform at OLCF\footnote{Summitdev
   consists of 2 IBM Power8 processors and 4 Nvidia Pascal GPUs per
   node.}.  The highest GPU count (128) corresponds to 32 nodes on
-summitdev.  We see a nearly linear speedup with the number of GPUs,
+Summitdev.  We see a nearly linear speedup with the number of GPUs,
 indicating good weak scaling on this machine.  Efforts are underway to

### 2017-10-27 astronum_2017/astronum-amrex.tex
Take a pass through

@@
 
-\address{$^5$Nvidia Corporation}
+\address{$^5$NVIDIA Corporation, 2788 San Tomas Expressway,
+  Santa Clara, CA, 95050 USA}
 
@@
 star (see~\cite{hillebrandtniemeyer2000} for the state of the field at
-that time).  Since then, the wealth of observations has indicated that
+that time).  Since then, a wealth of observations has indicated that
 there is a lot of diversity in SNe Ia, and searches for progenitor
@@
 hydrodynamics and reactions, a conservative gravity and rotation
-source formulation~\cite{wdmergerI}, and retry mechanism to redo a
+source formulation~\cite{wdmergerI}, and a retry mechanism to redo a
 step based on criteria evaluated during the
-integration.\MarginPar{more new stuff?}
+integration.
 
@@
 simulation state data that resides in each box, we have modified the
-memory allocator so that it can use a CUDA allocator (either device or
+memory allocator so that it can use a CUDA allocator (mainly relying on
 managed memory).  The domain iterator is configured to handle data
@@
 backend support used to implement them on various compute
-architectures.\MarginPar{add CPU time for reference}
+architectures.
 
@@
 \starlord\ on the Summitdev platform at OLCF\footnote{Summitdev
-  consists of 2 IBM Power8 processors and 4 Nvidia Pascal GPUs per
+  consists of 2 IBM Power8 processors and 4 NVIDIA Pascal GPUs per
   node.}.  The highest GPU count (128) corresponds to 32 nodes on
 Summitdev.  We see a nearly linear speedup with the number of GPUs,
-indicating good weak scaling on this machine.  Efforts are underway to
+indicating good weak scaling on this machine. A single P100 GPU achieves
+a performance approximately 2.5 times that of the 20 Power8 cores. Efforts are underway to
 complete the port of the GPU developments into \castro.  This figure
@@
 \end{figure}
-
-\MarginPar{we should update this to a more modern processor---Don can you run
-this on summitdev?}
 
@@
 U.S. Department of Energy under Contract No.\ DE-AC02-05CH11231.
-Results in this paper were obtained using the high-performance LIred
-computing system at the Institute for Advanced Computational Science
-at Stony Brook University, which was obtained through the Empire State
-Development grant NYS \#28451.  The GPU work benefited from a grant of
-an Titan X Pascal board from Nvidia Corp through the GPU Grant
-Program.  Visualization were done using \cite{yt}.  This research has
+The authors would like to thank Stony Brook Research Computing
+and Cyberinfrastructure, and the Institute for Advanced Computational
+Science at Stony Brook University for access to the high-performance
+LIred and SeaWulf computing systems, the latter of which was made
+possible by a \$1.4M National Science Foundation grant (\#1531492).
+The GPU work benefited from a grant of an Titan X Pascal board
+from NVIDIA Corporation through the GPU Grant Program.
+Visualizations were done using \cite{yt}.  This research has
 made use of NASA's Astrophysics Data System Bibliographic Services.

### 2017-10-28 astronum_2017/scaling/castro/sedov_scaling.py
start of new scaling numbers with the new MLMG solver (on Castro mlmg branch). THese use the Cray compilers

@@ -8,6 +8,8 @@ plt.rcParams.update({'xtick.labelsize': 14,
 plt.rc("axes", linewidth=1.5)
 plt.rc("lines", markeredgewidth=1.5)
 
+plt.style.use("seaborn-darkgrid")
+
 class ScalingRun(object):
     def __init__(self, MPI=1, OMP=1, max_grid=1, nzones=1, max_level=1, time=0.0, std=0.0):
         self.MPI = MPI
@@ -97,3 +99,4 @@ f = plt.gcf()
 f.set_size_inches(8, 6)
 
 plt.savefig("sedov_scaling.pdf", dpi=150, bbox_inches="tight")
+plt.savefig("sedov_scaling.png", dpi=150, bbox_inches="tight")

### 2017-10-28 astronum_2017/astronum-amrex.tex
update plots

@@
   scaling.  The variability (shown by the error bars) at high core
-  counts shows we are becoming work-starved. (right) \castro\ strong
+  counts shows we are becoming work-starved.  These runs used the PGI 17.7
+  compilers.  (right) \castro\ strong
   scaling on OLCF Titan for a hydro + Poisson gravity (wdmerger)
@@
   demonstrate great strong scaling.  We also look at a single level of
-  refinement (by a factor of 4) on top of this coarse grid.}
+  refinement (by a factor of 4) on top of this coarse grid.  These runs 
+  used the Cray 8.5.7 compilers.}
 \end{figure}
@@
 configuration.  For all these runs, we ran in the 16 cores / node
-mode.  All runs also used the PGI 17.7 compilers.  We see excellent
+mode.  We see excellent
 scaling behavior for the pure hydro problem.  For the self-gravity
-problem, we scale well with a uniform grid, but with a factor of 4
+problem, we scale very well with a uniform grid, but with a factor of 4
 refinement for the AMR level on top of this, we observe degraded
@@
 from NVIDIA Corporation through the GPU Grant Program.
-Visualizations were done using \cite{yt}.  This research has
+Visualizations were done using yt~\cite{yt}.  This research has
 made use of NASA's Astrophysics Data System Bibliographic Services.

### 2017-10-28 astronum_2017/scaling/castro/wdmerger_scaling.py

@@ -28,7 +28,7 @@ def trend_line(c, t):
 
 runs = []
 
-data = np.loadtxt("castro-wdmerger-scaling.txt")
+data = np.loadtxt("castro-wdmerger-newmg.txt")
 
 for row in data:
     runs.append(ScalingRun(MPI=row[0], OMP=row[1], max_grid=row[4], 

### 2017-10-28 astronum_2017/astronum-amrex.tex
fix formatting

@@
 
-X-ray bursts --- the burning of accreted H/He on the surface of a
-neutron star --- can be important probes of neutron star structure.
+X-ray bursts---the burning of accreted H/He on the surface of a
+neutron star---can be important probes of neutron star structure.
 Interpreting observations requires that we understand what we are
@@
 
-A parallel effort is porting our microphysics --- in particular
-reaction networks --- to GPUs.  Our strategy is to do the entire ODE
+A parallel effort is porting our microphysics---in particular
+reaction networks---to GPUs.  Our strategy is to do the entire ODE
 integration on the GPU, i.e., the data for a patch of zones is passed

### 2017-10-30 astronum_2017/astronum-amrex.tex
Small changes after one pass.

@@
 explain most SNe Ia.  Today, merging white dwarfs (the double
-degenerate scenario) have perhaps become the most popular model.
+degenerate scenario) has perhaps become the most popular model.
 Other progenitors, like He burning on the surface of a sub-Chandra
@@
 can transition to a detonation during the burning front propagation
-through the star.  For the sub-Chandra double detonation model, is it
+through the star.  For the sub-Chandra double detonation model, it is
 not clear whether it is possible to create a detonation in the thin
@@
 A constraint equation is derived by
-requiring that the pressure everywhere be close to the background
+requiring that the pressure everywhere is close to the background
 hydrostatic pressure.  The constraint acts to enforce instantaneous

### 2017-11-01 astronum_2017/astronum-amrex.tex
final edits

@@
 supernovae~\cite{castro-pairinstability}, the Chandra model for
-SNe~Ia~\cite{ma:2013}, the sub-Chandra SNe Ia model~\cite{moll:2013}, and white dwarf mergers as a model for SNe
-Ia~\cite{moll:2014,wdmergerI}. Recent work has focused on improving energy
-conservation with gravitational and rotational sources, and new
-integration methods that more accurately couple hydro and reactions.
-For \maestro\ simulations that evolve from the subsonic regime to the
-sonic regime, we have demonstrated the ability to restart the
-calculations in \castro\ to continue the evolution into the sonic
-regime~\cite{scidac-petascale,malone:2014} (in this case, for Chandra
-model SNe Ia).
+SNe~Ia~\cite{ma:2013}, the sub-Chandra SNe Ia model~\cite{moll:2013},
+and white dwarf mergers as a model for SNe
+Ia~\cite{moll:2014,wdmergerI}.  For \maestro\ simulations that evolve
+from the subsonic regime to the sonic regime, we have demonstrated the
+ability to restart the calculations in \castro\ to continue the
+evolution into the sonic regime~\cite{scidac-petascale,malone:2014}
+(in this case, for Chandra model SNe Ia).
 
@@
 \maestro\ and \castro\ share the same microphysics, available as the
-\starkiller\ astrophysics GitHub
+\starkiller\ Microphysics GitHub
 project\footnote{\url{https://github.com/StarKiller-astro/microphysics/}}.
@@
 configuration.  For all these runs, we ran in the 16 cores / node
-mode.  We see excellent
-scaling behavior for the pure hydro problem.  For the self-gravity
-problem, we scale very well with a uniform grid, but with a factor of 4
-refinement for the AMR level on top of this, we observe degraded
-performance at higher core counts.  This is a challenging problem, as
-only 3\% of the domain volume is refined.  For a given core count,
-there are a number of different combinations of MPI ranks and OpenMP
-threads we can use.  In general, with multilevel problems, we found
-the best performance with fewer MPI ranks and more OpenMP threads.
+mode.  We see excellent scaling behavior for the pure hydro problem to
+$\mathcal{O}(10^5)$ cores.  For the self-gravity problem, we scale
+very well with a uniform grid, but with a factor of 4 refinement for
+the AMR level on top of this, we observe degraded performance at
+higher core counts.  This is a challenging problem, as only 3\% of the
+domain volume is refined.  For a given core count, there are a number
+of different combinations of MPI ranks and OpenMP threads we can use.
+In general, with multilevel problems, we found the best performance
+with fewer MPI ranks and more OpenMP threads.
 
@@
 For all these runs, we ran with 4 threads / core (272
-threads / chip).
+threads / chip).  We have not yet focused on optimizing \castro\ for the
+Intel Xeon Phi architecture.
 
@@
 \centering
-\includegraphics[width=0.48\linewidth]{CUDA-Fortran-VODE-Integrator}
+\includegraphics[width=0.48\linewidth]{cuda-vode-speedup}
 \begin{minipage}[b]{0.48\linewidth}
-\caption{\label{fig:cudaode} GPU speed up over single CPU core for two
+\caption{\label{fig:cudaode} GPU speed up over single CPU core for three
+  different reaction networks.
   reaction networks.  The 13-isotope alpha-chain is a
-  potential network for the sub-Chandra simulations.
-  The 9-isotope Urca network is used for our Urca simulations.  A
+  potential network for the sub-Chandra simulations;
+  the 9-isotope Urca network is used for our Urca simulations;
+  and the 3 isotope carbon network was used for our original \maestro\
+  white dwarf convection calculations.  A
   variety of grid sizes were used, $16^3$, $32^3$, $64^3$, and
-  $128^3$.  In all cases, we see significant speed up on the GPU
-  vs.\ a single core.  These speed-up numbers are from our testbed
-  workstation with a Titan X Pascal board and Intel Xeon X5650 processors.}
+  $128^3$.  In all cases, we see a good  speed up on the GPU
+  vs.\ a single core.  These speed-up numbers are from the
+  Summitdev machine, comparing a single Tesla P100 GPU to a single Power8 CPU core.}
 \end{minipage}
@@
 13-isotope standard network).  Figure~\ref{fig:cudaode} shows the
-speed-ups on a GPU vs.\ single CPU core.
+speed-ups on a GPU vs.\ single CPU core.  The main issue with scaling
+is running out of local stack memory per thread with larger networks.
+Reducing the memory footprint is a near-term goal for this work.  This
+test problem also shows a lot of thread divergence due to the widely
+differing thermodynamic conditions in the zones that are burning.
 

### 2017-11-01 astronum_2017/astronum-amrex.tex
typo
Added my middle initial.
righthand size -> righthand side.

@@
         A.~J. Nonaka$^2$,
-        D. Willcox$^1$, and
+        D.~E. Willcox$^1$, and
         W. Zhang$^2$}
@@
   GPUs in this test of a pure hydro problem (Sedov w/ real EOS).  On the
-right, we explore performance on NERSC Cori (and Intel Xeon Phi platform).}
+right, we explore performance on NERSC Cori (an Intel Xeon Phi platform).}
 \end{figure}
@@
 integration on the GPU, i.e., the data for a patch of zones is passed
-to the GPU, all righthand size and Jacobian evaluations and the
+to the GPU, all righthand side and Jacobian evaluations and the
 timestepping itself is done on the GPU, and once the burning in all

### 2017-12-04 astronum_2017/astronum-amrex.tex
fix some typos

@@
 systems have strongly suggested that Chandra-mass white dwarfs cannot
-explain most SNe Ia.  Today, merging white dwarfs (the double
+explain most SNe Ia.  Today, a pair of merging white dwarfs (the double
 degenerate scenario) has perhaps become the most popular model.
@@
 Titan.  The Intel Xeon Phi chips in Cori
-have 68 cores than can be run with 1, 2, or 4 hardware threads each.
+have 68 cores that can be run with 1, 2, or 4 hardware threads each.
 For all these runs, we ran with 4 threads / core (272

### 2018-09-11 MAESTROeX_I/paper.tex
add some more sections + Alice

%% This manuscript uses the AASTeX v6.2 LaTeX 2e macros
%%
%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls 
%% (Copyright 2000-2015).  See the classfile for details.

%% AASTeX requires revtex4-1.cls (http://publish.aps.org/revtex4/) and
%% other external packages (latexsym, graphicx, amssymb, longtable, and epsf).
%% All of these external packages should already be present in the modern TeX 
%% distributions.  If not they can also be obtained at www.ctan.org.

%% The first piece of markup in an AASTeX v6.x document is the \documentclass
%% command. LaTeX will ignore any data that comes before this command. The 
%% documentclass can take an optional argument to modify the output style.
%% The command below calls the preprint style  which will produce a tightly 
%% typeset, one-column, single-spaced document.  It is the default and thus
%% does not need to be explicitly stated.
%%
%%
%% using aastex version 6.2
\documentclass{aastex62}

%% The default is a single spaced, 10 point font, single spaced article.
%% There are 5 other style options available via an optional argument. They
%% can be envoked like this:
%%
%% \documentclass[argument]{aastex62}
%% 
%% where the layout options are:
%%
%%  twocolumn   : two text columns, 10 point font, single spaced article.
%%                This is the most compact and represent the final published
%%                derived PDF copy of the accepted manuscript from the publisher
%%  manuscript  : one text column, 12 point font, double spaced article.
%%  preprint    : one text column, 12 point font, single spaced article.  
%%  preprint2   : two text columns, 12 point font, single spaced article.
%%  modern      : a stylish, single text column, 12 point font, article with
%% 		  wider left and right margins. This uses the Daniel
%% 		  Foreman-Mackey and David Hogg design.
%%  RNAAS       : Preferred style for Research Notes which are by design 
%%                lacking an abstract and brief. DO NOT use \begin{abstract}
%%                and \end{abstract} with this style.
%%
%% Note that you can submit to the AAS Journals in any of these 6 styles.
%%
%% There are other optional arguments one can envoke to allow other stylistic
%% actions. The available options are:
%%
%%  astrosymb    : Loads Astrosymb font and define \astrocommands. 
%%  tighten      : Makes baselineskip slightly smaller, only works with 
%%                 the twocolumn substyle.
%%  times        : uses times font instead of the default
%%  linenumbers  : turn on lineno package.
%%  trackchanges : required to see the revision mark up and print its output
%%  longauthor   : Do not use the more compressed footnote style (default) for 
%%                 the author/collaboration/affiliations. Instead print all
%%                 affiliation information after each name. Creates a much
%%                 long author list but may be desirable for short author papers
%%
%% these can be used in any combination, e.g.
%%
%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex62}
%%
%% AASTeX v6.* now includes \hyperref support. While we have built in specific
%% defaults into the classfile you can manually override them with the
%% \hypersetup command. For example,
%%
%%\hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
%%
%% will change the color of the internal links to red, the links to the
%% bibliography to green, the file links to cyan, and the external links to
%% magenta. Additional information on \hyperref options can be found here:
%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
%%
%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%
\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}

% for non-stacked fractions
\newcommand{\sfrac}[2]{\mathchoice
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptscriptfont0 #1}\kern-.2em/
   \kern-.15em\lower.25ex\hbox{\the\scriptscriptfont0 #2}}
  {#1\!/#2}}

\newcommand{\myhalf}{\sfrac{1}{2}}
\newcommand{\thalf}{\sfrac{3}{2}}

\newcommand{\eb}{{\bf{e}}}
\newcommand{\Ub}{{\bf{U}}}
\newcommand{\Ubt}{\widetilde{\Ub}}
\newcommand{\Vb}{{\bf{V}}}
\newcommand{\xb}{{\bf{x}}}

\newcommand{\dr}{\Delta r}
\newcommand{\dt}{\Delta t}

\newcommand{\etarho}{\eta_\rho}
\newcommand{\gammaonebar}{\overline{\Gamma_1}}
\newcommand{\Hext}{H_{\rm ext}}
\newcommand{\Hnuc}{H_{\rm nuc}}
\newcommand{\omegadot}{\dot\omega}
\newcommand{\pred}{{\rm pred}}
\newcommand{\Sbar}{\overline{S}}

\newcommand{\inp}{\mathrm{in}}
\newcommand{\outp}{\mathrm{out}}
\newcommand{\nph}{{n+\myhalf}}
\newcommand{\nmh}{{n-\myhalf}}
\newcommand{\ow}{\overline{w_0}}
\newcommand{\dw}{\delta w_0}
\newcommand{\uadvone}{\Ub^{\mathrm{ADV},\star}}
\newcommand{\uadvonedag}{\Ub^{\mathrm{ADV},\dagger,\star}}
\newcommand{\uadvtwo}{\Ub^{\mathrm{ADV}}}
\newcommand{\uadvtwodag}{\Ub^{\mathrm{ADV},\dagger}}
\newcommand{\gcc}{\mathrm{g~cm^{-3} }}

% for the red MarginPars
\usepackage{color}
% make the MarginPars look pretty
\setlength{\marginparwidth}{0.5in}
\newcommand{\MarginPar}[1]{\marginpar{\vskip-\baselineskip\raggedright\tiny\sffamily
\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
\received{XXX X, XXXX}
\revised{XXX X, XXXX}
\accepted{XXX X, XXXX}
%% Command to document which AAS Journal the manuscript was submitted to.
%% Adds "Submitted to " the arguement.
\submitjournal{ApJ}

%% Mark up commands to limit the number of authors on the front page.
%% Note that in AASTeX v6.2 a \collaboration call (see below) counts as
%% an author in this case.
%
%\AuthorCollaborationLimit=3
%
%% Will only show Schwarz, Muench and "the AAS Journals Data Scientist 
%% collaboration" on the front page of this example manuscript.
%%
%% Note that all of the author will be shown in the published article.
%% This feature is meant to be used prior to acceptance to make the
%% front end of a long author article more manageable. Please do not use
%% this functionality for manuscripts with less than 20 authors. Conversely,
%% please do use this when the number of authors exceeds 40.
%%
%% Use \allauthors at the manuscript end to show the full author list.
%% This command should only be used with \AuthorCollaborationLimit is used.

%% The following command can be used to set the latex table counters.  It
%% is needed in this document because it uses a mix of latex tabular and
%% AASTeX deluxetables.  In general it should not be needed.
%\setcounter{table}{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
\shorttitle{MAESTROeX Low Mach Number Astrophysics}
\shortauthors{Fan et al.}
%%
%% You can add a light gray and diagonal water-mark to the first page 
%% with this command:
% \watermark{text}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
%% long you can control the water-mark size with:
%  \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% This is the end of the preamble.  Indicate the beginning of the
%% manuscript itself with \begin{document}.

\begin{document}

\title{Title}

%% LaTeX will automatically break titles if they run longer than
%% one line. However, you may use \\ to force a line break if
%% you desire. In v6.2 you can include a footnote in the title.

%% A significant change from earlier AASTEX versions is in the structure for 
%% calling author and affilations. The change was necessary to implement 
%% autoindexing of affilations which prior was a manual process that could 
%% easily be tedious in large author manuscripts.
%%
%% The \author command is the same as before except it now takes an optional
%% arguement which is the 16 digit ORCID. The syntax is:
%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
%%
%% This will hyperlink the author name to the author's ORCID page. Note that
%% during compilation, LaTeX will do some limited checking of the format of
%% the ID to make sure it is valid.
%%
%% Use \affiliation for affiliation information. The old \affil is now aliased
%% to \affiliation. AASTeX v6.2 will automatically index these in the header.
%% When a duplicate is found its index will be the same as its previous entry.
%%
%% Note that \altaffilmark and \altaffiltext have been removed and thus 
%% can not be used to document secondary affiliations. If they are used latex
%% will issue a specific error message and quit. Please use multiple 
%% \affiliation calls for to document more than one affiliation.
%%
%% The new \altaffiliation can be used to indicate some secondary information
%% such as fellowships. This command produces a non-numeric footnote that is
%% set away from the numeric \affiliation footnotes.  NOTE that if an
%% \altaffiliation command is used it must come BEFORE the \affiliation call,
%% right after the \author command, in order to place the footnotes in
%% the proper location.
%%
%% Use \email to set provide email addresses. Each \email will appear on its
%% own line so you can put multiple email address in one \email call. A new
%% \correspondingauthor command is available in V6.2 to identify the
%% corresponding author of the manuscript. It is the author's responsibility
%% to make sure this name is also in the author list.
%%
%% While authors can be grouped inside the same \author and \affiliation
%% commands it is better to have a single author for each. This allows for
%% one to exploit all the new benefits and should make book-keeping easier.
%%
%% If done correctly the peer review system will be able to
%% automatically put the author and affiliation information from the manuscript
%% and save the corresponding author the trouble of entering it by hand.

\correspondingauthor{Doreen Fan}
\email{DFan@lbl.gov}

\author{Doreen Fan}
\affil{Lawrence Berkelely National Laboratory \\
Center for Computational Sciences and Engineering \\
One Cyclotron Road, MS 50A-3111 \\
Berkeley, CA 94720, USA}

\author{Andrew Nonaka}
\affil{Lawrence Berkelely National Laboratory \\
Center for Computational Sciences and Engineering \\
One Cyclotron Road, MS 50A-3111 \\
Berkeley, CA 94720, USA}

\author{Ann Almgren}
\affil{Lawrence Berkelely National Laboratory \\
Center for Computational Sciences and Engineering \\
One Cyclotron Road, MS 50A-3111 \\
Berkeley, CA 94720, USA}

\author[0000-0002-1530-781X]{Alice Harpole}
\affil{Stony Brook University \\
Department of Physics and Astronomy \\
Stony Brook, NY 11794-3800, USA}

\author[0000-0001-8401-030X]{Michael Zingale}
\affil{Stony Brook University \\
Department of Physics and Astronomy \\
Stony Brook, NY 11794-3800, USA}

%% Note that the \and command from previous versions of AASTeX is now
%% depreciated in this version as it is no longer necessary. AASTeX 
%% automatically takes care of all commas and "and"s between authors names.

%% AASTeX 6.2 has the new \collaboration and \nocollaboration commands to
%% provide the collaboration status of a group of authors. These commands 
%% can be used either before or after the list of corresponding authors. The
%% argument for \collaboration is the collaboration identifier. Authors are
%% encouraged to surround collaboration identifiers with ()s. The 
%% \nocollaboration command takes no argument and exists to indicate that
%% the nearby authors are not part of surrounding collaborations.

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
Insert abstract here.

\end{abstract}

%% Keywords should appear after the \end{abstract} command. 
%% See the online documentation for the full list of available subject
%% keywords and the rules for their use.
\keywords{insert keywords here}

%% From the front matter, we move on to the body of the paper.
%% Sections are demarcated by \section and \subsection, respectively.
%% Observe the use of the LaTeX \label
%% command after the \subsection to give a symbolic KEY to the
%% subsection for cross-referencing in a \ref command.
%% You can use LaTeX's \ref and \label commands to keep track of
%% cross-references to sections, equations, tables, and figures.
%% That way, if you change the order of any elements, LaTeX will
%% automatically renumber them.
%%
%% We recommend that authors also use the natbib \citep
%% and \citet commands to identify citations.  The citations are
%% tied to the reference list via symbolic KEYs. The KEY corresponds
%% to the KEY in the \bibitem in the reference list below.

\section{Introduction} \label{sec:intro}

Sample citation \cite{MAESTRO_I,MAESTRO_II,MAESTRO_III,MAESTRO_IV,MAESTRO_V}

%% This command is needed to show the entire author+affilation list when
%% the collaboration and author truncation commands are used.  It has to
%% go at the end of the manuscript.
%\allauthors

%% Include this line if you are using the \added, \replaced, \deleted
%% commands to see a summary list of all changes at the end of the article.
%\listofchanges

\section{Using Exact Spherical Base States on Irregular Grid} \label{sec:exactbasestate}

\subsection{Potential Problems}
\begin{enumerate}
\item How do we define \verb|r_edge_loc| for irregularly-spaced base states whose cell-centered values are defined at $r_m=\Delta x\sqrt{\frac{3}{4} + 2m}$?
\item We would stil need to approximate the values at edges when advecting the base states (i.e. $\rho_0^{\nph,\pred}$). This may affect the nice properties we would get with exact cell-centered base state values.
\item Excessive memory storage may be required (\verb|nr_irreg >> nr_fine|). 
\end{enumerate}

\subsection{A Possible Solution}
We could choose to no longer advect the base states for both predictor and corrector steps. This would solve the problems with computing state values at cell edges. Instead, we propose to only update the base states $\rho_0$ and $p_0$ by computing the average of the full state values after advecting the full states. This would result in the following changes in the algorithm from Paper V:
%
\begin{enumerate}
\item Steps 4 and 8 will be greatly simplified, leading to a more efficient algorithm.
\item There is no longer a need for $w_0$, so we can use the full velocity $\Ub$ instead of $\Ubt$. As a consequence, $\eta$ and $\psi$ also do not need to be computed. This has the added advantage of saving some memory storage by removing many of the terms needed when we decided to split terms into its base state and perturbation state in our previous implementation.
\item By placing the base states exactly at cell-centers, we eliminate interpolation errors and simplify the existing code.
\end{enumerate}
%

\subsection{Main Algorithm Description}\label{Sec:Main Algorithm Description}
We now describe the proposed algorithm, making frequent use of the
shorthand notations mentioned in Paper V, Appendix A3.  In summary, in the predictor step ({\bf
  Steps 2-5}) we use an estimate of the expansion term, $S$, to
compute a preliminary solution at the new time level, denoted with an
``$n+1,\star$'' superscript.  In the corrector step ({\bf Steps 6-9}),
we use the results from the predictor step to compute a more accurate
expansion term, and compute the final solution at the new time level,
denoted with an ``$n+1$'' superscript.  We use Strang-splitting to
achieve second-order accuracy in time.

\begin{description}

%--------------------------------------------------------------------------
% STEP 0
%--------------------------------------------------------------------------

\item[Step 0.] {\em Initialization}
The initialization step only occurs at the beginning of the simulation. 
The initialization step computes initial values for $\pi^{-\myhalf}$, $S^0$, and
$S^1$ for use in time stepping routine.  We pass through {\bf Step 1} -- {\bf Step 12},
but only keep these values. 
The initial values for $\Ub^0, \rho^0, (\rho h)^0, X_k^0, T^0, 
\rho_0^0, p_0^0$, and $\overline{\Gamma}_1^0$ are specified from the problem-dependent 
initial conditions.  The initial time step, $\dt^0$, is computed as in 
Paper III.

%--------------------------------------------------------------------------
% STEP 1
%--------------------------------------------------------------------------
\item[Step 1.] {\em React the full state through the first time interval of $\dt / 2.$}

Call {\bf React State}$[\rho^n, (\rho h)^n, X_k^n, T^n, (\rho\Hext)^n, p_0^n] \rightarrow [\rho^{(1)},(\rho h)^{(1)},X_k^{(1)},T^{(1)},(\rho \omegadot_k)^{(1)},(\rho \Hnuc)^{(1)}]$.

%--------------------------------------------------------------------------
% STEP 2
%--------------------------------------------------------------------------

\item[Step 2.] {\em Compute the provisional time-centered expansion, 
    $S^{\nph,\star}$.}

We compute an estimate for the time-centered expansion term in the velocity 
divergence constraint.  Following \citet{Bell:2004}, we extrapolate
to the half-time using $S$ at the previous and current
time steps,
\begin{equation}
S^{\nph,\star} = S^n + \frac{\dt^n}{2} \frac{S^n - S^{n-1}}{\dt^{n-1}}.
\end{equation}
Note that in the first time step we average $S^0$ and $S^1$ from the
initialization step.

%--------------------------------------------------------------------------
% STEP 3
%--------------------------------------------------------------------------
\item[Step 3.] {\em Construct the provisional time-centered advective velocity on 
edges, $\uadvone$.}

Using the equation below,
% 
\begin{equation}
\frac{\partial\Ub}{\partial t} = -\Ub\cdot\nabla\Ub - \frac{1}{\rho}\nabla\pi - \frac{\rho-\rho_0}{\rho}g\eb_r, \label{eq:u_update}
\end{equation}
%
we compute time-centered edge velocities, $\uadvonedag$, using 
$\Ub^n$.  The $\dagger$ superscript refers to the 
fact that the predicted velocity field does not satisfy the divergence 
constraint.  We then construct $\uadvone$ from $\uadvonedag$
using a MAC projection.  
We note that $\uadvone$ satisfies the divergence constraint 
\begin{equation}
\nabla \cdot \left(\beta_0^n \uadvone\right) = \beta_0^n \left(S^{\nph,\star} - \frac{1}{\overline{\Gamma}_1^np_0^n}\frac{p_0^n-p_0^{n-1}}{\Delta t^{n-1}} + \chi^n\right),
\end{equation}
\begin{equation}
 \beta_0^n = \beta_0 \left(\rho_0^n, p_0^n, \overline{\Gamma}_1^n\right),
\end{equation}
\begin{equation}
\chi^n = \frac{f}{\overline{\Gamma}_1^n p_0^n}\frac{p_{\rm eos}^n(\rho,h,X_k) - p_0^n}{\Delta t^n}
\end{equation}
where $\beta_0$ is computed as described in Appendix C of Paper III.
\MarginPar{add volume discrepancy stuff}

%--------------------------------------------------------------------------
% STEP 4
%--------------------------------------------------------------------------
\item[Step 4.] {\em Advect the full state through a time interval of $\dt.$}

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \Alph{enumi}}}

\item Update $(\rho X_k)$ using a discretized version of
%
\begin{equation}
\frac{\partial(\rho X_k)}{\partial t} = -\nabla\cdot(\rho X_k\Ub) + \rho\omegadot_k,
\end{equation}
%
omitting the reaction terms, which were already 
accounted for in {\bf React State}.  The update consists of two steps:

\begin{enumerate}
\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}

\item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\pred}$,
  for the conservative update of $(\rho X_k)^{(1)}$ using 
  $\Ub = \uadvone$ and $\rho_0^n$. 
\MarginPar{We may not need $\rho_0$ if we predict $\rho X$ or $\rho$ and $X$.}

\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2),\star}$ using
\begin{equation}
(\rho X_k)^{(2),\star} = (\rho X_k)^{(1)} 
  - \dt \left\{ \nabla \cdot \left[ \uadvone (\rho X_k)^{\nph,\pred} \right] \right\},
\end{equation}
\begin{equation}
\rho^{(2),\star} = \sum_k (\rho X_k)^{(2),\star},
\qquad
X_k^{(2),\star} = (\rho X_k)^{(2),\star} / \rho^{(2),\star}.
\end{equation}

\end{enumerate}

\item Update $\rho_0$ by setting $\rho_0^{n+1,\star} =$ {\bf Avg}$(\rho^{(2),\star})$.

\item Update $p_0$ by calling 
{\bf Enforce HSE}$[p_0^n,\rho_0^{n+1,\star}] \rightarrow [p_0^{n+1,\star}]$.

\item Update the enthalpy using a discretized version of equation
%
\begin{equation} 
\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc + \rho\Hext,
\end{equation}
%
again omitting the reaction and heating terms
since we already accounted for
them in {\bf React State}.  This equation takes the form:
\begin{equation}
\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h) + \frac{\partial p_0}{\partial t} + (\Ub \cdot \eb_r) \frac{\partial p_0}{\partial r}.
\end{equation}
For spherical geometry, we solve the
analytically equivalent form,
\begin{equation}
\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h) + \frac{\partial p_0}{\partial t} + \nabla \cdot (\Ub p_0) - p_0 \nabla \cdot \Ub.
\end{equation}
The update consists of two steps:

\begin{enumerate}
\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}

\item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\pred},$
  for the conservative update of $(\rho h)^{(1)}$
  using $\Ub = \uadvone$ and $(\rho h)^n$,

\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$.
\begin{description}
\item For planar geometry,
\begin{eqnarray}
(\rho h)^{(2),\star} 
&=& (\rho h)^{(1)} \nonumber \\
&&- \dt^n \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\ 
&& + \dt^n \left(\uadvone \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^{n} \nonumber \\ 
&& + \dt^n \left( \frac{p_0^n - p_0^{n-1}}{\dt^{n-1}} \right),
\end{eqnarray}

\item For spherical geometry,
\begin{eqnarray}
(\rho h)^{(2),\star} 
&=& (\rho h)^{(1)} \nonumber \\
&&- \dt^n \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\ 
&& + \dt^n \left \{ \nabla \cdot \left (\uadvone p_0^{n} \right ) - p_0^{n} \nabla \cdot \uadvone \right \} \nonumber \\
&& + \dt^n \left( \frac{p_0^n - p_0^{n-1}}{\dt^{n-1}} \right),
\end{eqnarray}
\end{description}

\end{enumerate}

Note that during the first time step we use $\dt^n$ instead of $\dt^{n-1}$.
Then, for each Cartesian cell where $\rho^{(2),\star} < \rho_\mathrm{cutoff}$,
we recompute enthalpy using
\begin{equation}
(\rho h)^{(2),\star} = \rho^{(2),\star}h\left(\rho^{(2),\star},p_0^{n+1,\star},X_k^{(2),\star}\right).
\end{equation}

\item Update the temperature using the equation of state:
$T^{(2),\star} = 
  T(\rho^{(2),\star}, h^{(2),\star}, X_k^{(2),\star})$ (planar geometry) or
$T^{(2),\star} = 
  T(\rho^{(2),\star}, p_0^{n+1,\star}, X_k^{(2),\star})$ (spherical geometry).
\end{enumerate}

%--------------------------------------------------------------------------
% STEP 5
%--------------------------------------------------------------------------
\item[Step 5.] {\em React the full state through a second time interval of $\dt / 2.$}

Call {\bf React State}$[ \rho^{(2),\star},(\rho h)^{(2),\star}, X_k^{(2),\star}, T^{(2),\star}, 
(\rho\Hext)^{(2),\star}, p_0^{n+1,\star}]$\\
\phantom{ }\hfill $\rightarrow [ \rho^{n+1,\star},(\rho h)^{n+1,\star}, X_k^{n+1,\star}, T^{n+1,\star}, (\rho \omegadot_k)^{(2),\star}, (\rho \Hnuc)^{(2),\star} ].$

%--------------------------------------------------------------------------
% STEP 6
%--------------------------------------------------------------------------
\item[Step 6.] {\em Compute the time-centered expansion, $S^{\nph,\star}$.}

\item First, compute $S^{n+1,\star}$ with
\begin{equation}
S^{n+1,\star} =  -\sigma  \sum_k  \xi_k  (\omegadot_k)^{(2),\star}  + \frac{1}{\rho^{n+1,\star} p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2),\star} + \sigma \Hnuc^{(2),\star} + \sigma \Hext^{(2),\star},
\end{equation}
  where $(\omegadot_k)^{(2),\star} = (\rho \omegadot_k)^{(2),\star} /
  \rho^{(2),\star}$ and the thermodynamic quantities are defined using
  $\rho^{n+1,\star}, X_k^{n+1,\star},$ and $T^{n+1,\star}$ as inputs to
  the equation of state.  Then, define
\begin{equation}
 S^{\nph} = \frac{S^n + S^{n+1,\star}}{2},
\end{equation}

%--------------------------------------------------------------------------
% STEP 7
%--------------------------------------------------------------------------
\item[Step 7.] {\em Construct the time-centered advective velocity on edges, $\uadvtwo$.}

The procedure to construct $\uadvtwodag$ is identical to the procedure
for computing $\uadvonedag$ in {\bf Step 3}, but uses 
the updated value $S^{\nph}$ rather than $S^{\nph,\star}$.  
We note that $\uadvtwo$ satisfies the divergence constraint 
\begin{equation}
\nabla \cdot \left(\beta_0^{\nph} \uadvtwo\right) =
\beta_0^{\nph} \left(S^{\nph} - \frac{1}{\overline{\Gamma}_1^{n+\myhalf}p_0^n}\frac{p_0^{n+1,\star}-p_0^n}{\Delta t^n} + \chi^{n+\myhalf}\right),
\end{equation}
\begin{equation}
\beta_0^{\nph} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2};
\qquad
 \beta_0^{n+1,\star} = \beta_0 \left(\rho_0^{n+1,\star}, p_0^{n+1,\star}, \overline{\Gamma}_1^{n+1,\star}\right),
\end{equation}
\begin{equation}
\overline{\Gamma}_1^{\nph} = \frac{ \overline{\Gamma}_1^n +  \overline{\Gamma}_1^{n+1,\star} }{2}.
\end{equation}
\begin{equation}
\chi^{n+\myhalf} = \chi^n + \frac{f}{\overline{\Gamma}_1^{n+1,\star} p_0^{n+1,\star}}\frac{p_{\rm eos}^{n+1,\star}(\rho,h,X_k) - p_0^{n+1,\star}}{\Delta t^n}
\end{equation}
\MarginPar{add volume discrepancy stuff}

%--------------------------------------------------------------------------
% STEP 8
%--------------------------------------------------------------------------
\item[Step 8.] {\em Advect the base state and full state through a time interval of $\dt.$}

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \Alph{enumi}}}

\item Update $(\rho X_k)$.  This step is identical to {\bf Step 4A} except we use
  the updated values $\uadvtwo$ and $\rho_0^{(2a)}$ rather than 
  $\uadvone$ and $\rho_0^{(2a),\star}$.  In particular:

\begin{enumerate}
\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}

\item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\pred}$,
  for the conservative update of $(\rho X_k)^{(1)}$ using $\Ub = \uadvtwo$ and 
  $\rho_0^n$.

\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2)}$ using
\begin{equation}
(\rho X_k)^{(2)} = (\rho X_k)^{(1)} 
- \dt \left\{ \nabla \cdot \left[\uadvtwo (\rho X_k)^{\nph,\pred} \right] \right\},
\end{equation}
\begin{equation}
\rho^{(2)} = \sum_k (\rho X_k)^{(2)},
\qquad
X_k^{(2)} = (\rho X_k)^{(2)} / \rho^{(2)}.
\end{equation}

\end{enumerate}

\item Update $\rho_0$ by setting $\rho_0^{n+1} =$ {\bf Avg}$(\rho^{(2)})$.

\item Update $p_0$ by calling
{\bf Enforce HSE}$[p_0^n,\rho_0^{n+1}] \rightarrow [p_0^{n+1}]$.

\item Update the enthalpy.  This step is identical to {\bf Step 4D} except we use
  the updated values $\uadvtwo, \rho_0^{n+1}, (\rho h)_0^{n+1}$, and $p_0^{n+\myhalf}$ 
  rather than 
  $\uadvone, \rho_0^{n+1,\star}, (\rho h)_0^{n+1,\star}$, and $p_0^n$.  
  In particular:
\MarginPar{Again, we may not need $\rho_0$ and $(\rho h)_0$.}

\begin{enumerate}
\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}

\item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\pred},$
  for the conservative update of $(\rho h)^{(1)}$.

\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2)}$.
\begin{description}
\item For planar geometry,
\begin{eqnarray}
(\rho h)^{(2)} 
&=& (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvtwo (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
&& + \dt \left(\uadvtwo \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^\nph \nonumber \\
&& + \dt \left( \frac{p_0^{n+1} - p_0^n}{\dt^n} \right),
\end{eqnarray}

\item For spherical geometry,
\begin{eqnarray}
(\rho h)^{(2)} 
&=& (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvtwo (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
&& + \dt \left[ \nabla \cdot \left (\uadvtwo p_0^{\nph} \right ) - p_0^{\nph} \nabla \cdot \uadvtwo \right] \nonumber \\
&& + \dt \left( \frac{p_0^{n+1} - p_0^n}{\dt^n} \right),
\end{eqnarray}
\end{description}
where $p_0^\nph$ is defined as $p_0^\nph = (p_0^n+p_0^{n+1})/2$.

\end{enumerate}

Then, for each Cartesian cell where $\rho^{(2)} < \rho_\mathrm{cutoff}$, we recompute enthalpy using
\begin{equation}
(\rho h)^{(2)} = \rho^{(2)}h\left(\rho^{(2)},p_0^{n+1},X_k^{(2)}\right).
\end{equation}

\item Update the temperature using the equation of state:
$T^{(2)} = 
   T(\rho^{(2)}, h^{(2)}, X_k^{(2)})$ (planar geometry) or
$T^{(2)} = 
   T(\rho^{(2)}, p_0^{n+1}, X_k^{(2)})$ (spherical geometry).
\end{enumerate}

%--------------------------------------------------------------------------
% STEP 9
%--------------------------------------------------------------------------
\item[Step 9.] {\em React the full state through a second time interval of $\dt / 2.$}

Call {\bf React State}$[\rho^{(2)},(\rho h)^{(2)}, X_k^{(2)},T^{(2)}, (\rho\Hext)^{(2)}, p_0^{n+1}] \rightarrow [\rho^{n+1}, (\rho h)^{n+1}, X_k^{n+1}, T^{n+1}, (\rho \omegadot_k)^{(2)}, (\rho \Hnuc)^{(2)} ].$

%--------------------------------------------------------------------------
% STEP 10
%--------------------------------------------------------------------------
\item[Step 10.] {\em Define the new time expansion, $S^{n+1}$, and $\overline{\Gamma}_1^{n+1}$.}

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \Alph{enumi}}}
\item Define
\begin{equation}
  S^{n+1} =  -\sigma  \sum_k  \xi_k (\omegadot_k)^{(2)}  + \sigma \Hnuc^{(2)} +
  \frac{1}{\rho^{n+1} p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2)}  
   + \sigma \Hext^{(2)},
\end{equation}
where $(\omegadot_k)^{(2)} = (\rho \omegadot_k)^{(2)} / \rho^{(2)}$
and the thermodynamic quantities are defined using $\rho^{n+1}$,
$X_k^{n+1}$, and $T^{n+1}$ as inputs to the equation of state.

\item Define
\begin{equation}
\overline{\Gamma}_1^{n+1} = {\rm{\bf Avg}}\left[\Gamma_1\left(\rho^{n+1}, p_0^{n+1}, 
X_k^{n+1}\right) \right].
\end{equation}

\end{enumerate}

%--------------------------------------------------------------------------
% STEP 11
%--------------------------------------------------------------------------
\item[Step 11.] {\em Update the velocity}.

First, we compute the time-centered edge velocities, $\Ub^{\nph,\pred}$.
Then, we define
\begin{equation}
\rho^\nph = \frac{\rho^n + \rho^{n+1}}{2}, \qquad \rho_0^\nph = \frac{\rho_0^n + \rho_0^{n+1}}{2}.
\end{equation}
We update the velocity field $\Ub^n$ to $\Ub^{n+1,\dagger}$ by discretizing 
equation (\ref{eq:u_update}) as
\begin{eqnarray}
\Ub^{n+1,\dagger} 
&=& \Ub^n - \dt \left[\uadvtwo \cdot \nabla \Ub^{\nph,\pred} \right] \nonumber \\
&& - \dt \left[ - \frac{1}{\rho^\nph} \mathbf{G} \pi^\nmh + \frac{\left(\rho^\nph-\rho_0^\nph\right)}{\rho^\nph} g^{\nph} \eb_r \right],\nonumber \\
\end{eqnarray}
where $\mathbf{G}$ approximates a cell-centered gradient from nodal
data.  Again, the $\dagger$ superscript refers 
to the fact that the updated velocity does not satisfy the divergence 
constraint.

Finally, we use an approximate nodal projection to define $\Ub^{n+1}$
from $\Ub^{n+1,\dagger},$  such that $\Ub^{n+1}$ approximately
satisfies 
\begin{equation}
\nabla \cdot \left(\beta_0^{n+1} \Ub^{n+1} \right) 
= \beta_0^{n+1} S^{n+1},
\end{equation}
\MarginPar{I think we need the 1/gamma1bar part too}
where 
\begin{equation}
\beta_0^{n+1} = \beta \left(\rho_0^{n+1}, p_0^{n+1}, \overline{\Gamma}_1^{n+1}, g^{n+1}\right).
\end{equation}
As part of the projection we also define the new-time perturbational pressure,
$\pi^\nph.$  This projection necessarily differs from the MAC projection used in 
{\bf Step 3} and {\bf Step 7} because the velocities in those steps are defined
on edges and $\Ub^{n+1}$ is defined at cell centers, requiring different divergence
and gradient operators.  Details of the approximate projection are given in Paper III.

%--------------------------------------------------------------------------
% STEP 12
%--------------------------------------------------------------------------
\item[Step 12.] {\em Compute a new $\dt.$}

Compute $\dt$ for the next time step with a similar procedure described in
\S 3.4 of Paper III. However, we are using $\Ub^{n+1}$
as computed in {\bf Step 11} instead of $w_0$ and $\Ubt^{n+1}$ as in Paper V.
The following constraints need to be satisfied.

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \Alph{enumi}}}
\item Use standard CFL condition for explicit methods using CFL factor $\sigma^{\mathrm{CFL}}$. 
  For a calculation in $n_{\mathrm{dim}}$ dimensions ($n_{\mathrm{dim}}=2$ or 3), 
  the first constraint is 
\begin{equation}
\dt_{\Ub} = \sigma^\mathrm{CFL}  \min_{i=1\ldots n_\mathrm{dim}} \left \{ \dt_i \right \}\enskip,
\end{equation}
  where
\begin{equation}
\dt_i = \min_{\bf x}  \left \{ \frac{\Delta x_i}{\left | \Ub \right |}  \right \}\enskip.
\end{equation}

\item Use the forcing terms rather than the velocities. 
  This constraint is necessary when a calculation is started from rest, 
  since in that case the velocity-based time step would be infinite. Define
\begin{equation}
\dt_{\rm F} = \min_{i=1\ldots n_\mathrm{dim}} \left \{ \dt_{\rm F_i} \right \}\enskip,
\end{equation}
where
\begin{equation}
\dt_{\rm F_i} = \sqrt{2 \Delta x_i / {F_i}_\mathrm{max}}\enskip,
\end{equation}
where ${F_i}_\mathrm{max}$ is the maximum buoyancy force in the
$i^\mathrm{th}$ coordinate direction.

\item To prevent local expansion from numerically emptying a cell, 
  we require that the density be reduced by no more than 40\% in a single time step.
  This constraint is expressed as $\dt \leq \dt_S$ where 
\begin{equation}
 \Delta t_S \; (\nabla \cdot \Ub)  \leq 0.40 \enskip ,
\end{equation}
  in every cell.

\end{enumerate}

\end{description}

\noindent This completes one step of the algorithm.

\section{Test Problems}

\section{Performance and Scaling}

show OpenMP performance, maybe some comparisons with old Maestro?

\section{Conclusions and Future Work}

future stuff: rotation, SDC, high-order, GPUs

\acknowledgements

\bibliographystyle{aasjournal.bst}
\bibliography{references}

\end{document}

### 2018-09-16 astronum_2018/astronum-xrb.tex
add astronum 2018

\documentclass[a4paper]{jpconf}

\usepackage{graphicx}

\usepackage{hyperref}

\usepackage{cite}

\usepackage{amsmath}

\usepackage{bm}

\newcommand{\maestro}{{\sffamily Maestro}}
\newcommand{\castro}{{\sffamily Castro}}
\newcommand{\starkiller}{{\sffamily StarKiller}}
\newcommand{\starlord}{{\sffamily StarLord}}
\newcommand{\nyx}{{\sffamily Nyx}}
\newcommand{\amrex}{{\sffamily AMReX}}

\newcommand{\Uc}{{\,\bm{\mathcal{U}}}}
\newcommand{\Advs}[1]{\boldsymbol{\mathcal{A}} \left(#1\right)}

\newcommand{\cpp}{C\nolinebreak\hspace{-.05em}\raisebox{.4ex}{\tiny\bf +}\nolinebreak\hspace{-.10em}\raisebox{.4ex}{\tiny\bf +}}

\usepackage{color}
\setlength{\marginparwidth}{0.75in}
\newcommand{\MarginPar}[1]{\marginpar{\vskip-\baselineskip\raggedright\tiny\sffamily\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}

\newcommand{\apj}{Astrophysical Journal}
\newcommand{\aap}{Astronomy and Astrophysics}
\newcommand{\mnras}{Monthly Notices of the Royal Astronomical Society}
\newcommand{\prd}{Physical Review D}
\newcommand{\apjs}{Astrophysical Journal Supplement}

\begin{document}

\title{Toward Resolved Simulations of Burning Fronts in Thermonuclear
       X-ray Bursts}

\author{M. Zingale$^1$,
        K.~Eiden$^1$,
        Y.~Cavecchi$^2$,
        A.~Harpole$^1$,
        M.~P. Katz$^5$,
        J.~B. Bell$^2$,
        M.~Chang$^1$,
        C.~M. Malone$^6$, and
        W. Zhang$^2$}

\address{$^1$Department of Physics and Astronomy, Stony Brook
  University, Stony Brook, NY 11794-3800 USA}

\address{$^2$Center for Computational Sciences and Engineering,
  Lawrence Berkeley National Lab, Berkeley, CA 94720 USA}

\address{$^3$National Energy Research Scientific Computing Center,
  Lawrence Berkeley National Lab, Berkeley, CA 94720 USA}

\address{$^4$Department of Physics and Astronomy, Michigan State
  University, East Lansing, Michigan 48824 USA}

\address{$^5$NVIDIA Corporation, 2788 San Tomas Expressway,
  Santa Clara, CA, 95050 USA}

\address{$^6$Los Alamos National Laboratory, Los Alamos, NM, 87545 USA}

\begin{abstract}
We discuss the challenges of modeling X-ray bursts, review the different
calculations done to date, and discuss our new set of simulations.  We also
describe algorithmic improvements that may help in the future to offset some
of the expense of these simulations.
\end{abstract}

\section{Introduction}

X-ray bursts (XRBs) are fascinating astrophysical explosions.  A
neutron star accretes fuel from its binary companion, building up just
a thin layer ($\sim 5$--$10$~m) before the immense gravitational
acceleration at the neutron star surface compresses and heats the fuel
to the point of thermonuclear runaway.  A brief flash of X-rays
follows, the accreted layer diffuses the heat produced from reactions,
and the process repeats.  We can observe multiple bursts from a single
source, and the information their lightcurves encodes tells us about
the underlying neutron star, and ultimately the nuclear equation of
state.  A challenging aspect of interpreting the lightcurves is
understanding the radiation transport through the hot ash layers.  In
particular, the composition can affect the effective temperature,
which in turn affects the radius we measure for the neutron
star~\cite{suleimanov:2011}.  Convection in the burning layer can even
bring ashes up to the photosphere which alters what we
see~\cite{kajava:2017}.  Accurate measurements of neutron star radii
from XRBs could greatly constrain the nuclear equation of
state~\cite{steiner:2010,ozel:2010}.  Finally, brightness oscillations
and other features seen in lightcurves are interpreted as arising from
the spreading of a hotspot across the neutron star---demonstrating
that the burst begins in a localized
region~\cite{bhattacharyya:2006,bhattacharyya:2007}.
See~\cite{galloway:2017} for a recent review of XRBs.

Simulations of XRBs can help us to understand the observations, but
many technical challenges remain.  The primary difficulties are simply
the range of length and timescales involved.  At the largest scale,
the neutron star is 10~km and the Rossby lengthscale, where rotation
and laterial pressure gradients balance is $\sim
1$~km~\cite{SPIT_ETAL02}.  At the smallest scales, a conductive He
flame has a thickness of 10s of centimeters~\cite{Timmes00} and the
pressure scale height of the fuel layer is meters.  To capture the
rise of the XRB, we would need to model seconds, but the flame itself
is subsonic, so long timescale evolution is difficult.

Present and past simulations of XRBs employ a variety of
approximations to get past these difficulties.  The variety of
approximations used enables a complementary picture of XRBs to be built up, 
and has greatly advanced our understand of these systems.  
The breadth of simulations to date has been impressive.  We summarize
the different approaches below.

One-dimensional simulations~\cite{woosley-xrb,fisker:2008} assume
spherical symmetry and model the burst through the vertical column
depth of fuel.  These simulations can capture the luminosities and
burst recurrence times well.  The can utilize very large nuclear
reaction networks and have been used to understand the nucleosynthesis
and rp-process, as well as to understand how rate uncertainties can
affect the outcomes of the bursts.  Many successive bursts can be
modeled and these simulations can show us the evolution of the ash
layer with time.  The approximation of 1D however means that we cannot
learn about lateral variations across the neutron star, including
flame spreading.

Global shallow water simulations~\cite{SPIT_ETAL02} capture the
largescale spreading of the burning front by using a very simple
vertical structure.  These simulations were instrumental in showing
that the Coriolis force plays an important role in confining a
spreading region and the flame spreading can explain brightness
oscillations during the rise of the burst lightcurve.

Our group and others have modeled the convection preceeding the
runaway using low Mach number hydrodynamic methods, for pure helium
bursts~\cite{Lin:2006,xrb} and mixed hydrogen/helium
bursts~\cite{xrb2} in 2d, and full 3d simulations of mixed
bursts~\cite{xrb3}, and followed the development of the turbulence.
These simulations can give an understand of the pre-burning front
regime, including the strength and character of the turbulence, but
cannot model largescale lateral variations because of the implicit
assumptions built into the low Mach number model~\cite{ABRZ:I}.

While detonations are computationally easier to model than flames (we
largely get the speed through the jump conditions in the Riemann
problem without resolving the structure), detonations require extreme
conditions not found in XRBs~\cite{ZINGALE_ETAL01,harpole:2018}.
This means that calculations that want to capture the nucleosynthesis
yields from spreading burning fronts in XRBs need to model a deflagration.

Flame spreading was first modeled in a series of calculations
employing an algorithm from atmospheric science, where vertical
hydrostatic equilibrium is enforced and wide-aspect ratio zones give a
horizontal CFL number that is large enough to get long timescale
evolution~\cite{cavecchi:2012}.  These calculations... \MarginPar{Yuri can probably best summarize}

A longstanding goal is to perform simulations where we resolve the
flame structure, to allow us to accurately capture the
nucleosynthesis, and watch it move across the neutron star surface.
We show some in-progress calculations that are a step toward this (but
still have approximations) and also discuss what new algorithmic
techniques might be needed to realize this goal with the advent of
exascale computing.

\section{Resolved Flame Studies}

We have begun a set of flame spreading calculations using the
compressible hydrodynamics code Castro~\cite{castro}, part of the
AMReX astrophysics suite~\cite{astronum:2017}.  \castro\ uses adaptive
mesh refinement, supports a general equation of state and reaction
networks, has a conservative gravity formulation~\cite{wdmergerI}, and
is optimized to run on current supercomputers.  All of our code base
is open source and available on
github\footnote{\url{http://github.com/amrex-astro/}}, and all of the
solvers, problem setup, initial models, and analysis scripts for each
of our science problems are distributed with the codes.

Short description of problem setup and approximations
rotation, network, flame boosting

Describe resolution needs and flame structure, Figure~\ref{fig:xrb}

\begin{figure}
\centering
\includegraphics[width=0.75\linewidth]{xrb}
\caption{\label{fig:xrb} Helium flame spreading across the surface of
  a neutron star.  Shown are temperature (top), the velocity out of the
  simulation plane (middle), and nuclear energy generation rate
  (bottom).  }
\end{figure}

\section{Estimates for Full Star Simulations}

A fully resolved calculation of burning over the entire neutron star
is out of reach, and will be for some time.

time estimate...

\MarginPar{Alice and Mike will work on this}

\section{Algorithmic Developments}

Algorithmic improvements are needed to push the scale of simulations
to the point where we can model XRBs at the full star scale with
resolved nuclear physics.  Here we discuss a few possibilities.

Machine architectures are evolving, with heterogeneous architectures
becoming more common.  Our recent development in \castro\ has focused
on moving the hydrodynamics to GPUs~\cite{astronum:2017}.  The XRB
problem is an ideal case for this, since hydrodynamics with constant
gravity (for plane-parallel domains) ports in a straightforward manner
to GPUs.  We can expect GPUs to perform 10--20$\times$ faster than
CPUs on modern supercomputer nodes.  This performance increase will
directly benefit XRB simulations.  GPU offloading of the reaction networks
will also allow us to explore larger networks and more detailed nuclear
physics.

If we continue to resolve the flame structure, then we need algroithms
that can do so less expensively.  We can be smarter about our grid
refinement strategy, putting the finest grids only at the very base of
the burning layer, where the flame is the thinnest.  Using
well-balanced schemes can help us retain HSE in the lower refined
regions, reducing the number of zones needed to maintain fidelity.  We
can also explore initialize newly created fine grids ahread of the burning
with an understanding of hydrostatic equilibrium to allow us to only refine
the peak burning as the flame moves across the star.

Another way we can meet the resolution requirements is to use a more
accurate hydrodynamic method.  In particular, we are developing fully
fourth-order (in space and time) coupling of hydrodynamics and
reactions using the method of spectral deferred corrections in
\castro, instead of the more commonly employed Strang splitting.  This
provides us with two benefits.  First, the improved coupling actually
reduces the stiffness of the reactions, requiring fewer righthand side
calls and therefore reducing the expense of the
reactions. \MarginPar{need some refs here} Second, by moving to fourth
order, we may also be able to reduce our resolution requirements
needed for converged flames, and since computational work scales like
$(\Delta x)^4$ in 3D, this can save a lot of computer time.

As an example of the improvements in coupling between hydro and
reactions, Figure~\ref{fig:sdc} shows the mass fraction of helium
behind a detonation over the course of 3 timesteps.  The points
represent individual calls to the righthand side function of our
reaction network (a 13-isotope $\alpha$-chain), and the density of
points indicates how hard the integrator is working.  We use the same
tolerances for both the Strang-split method and the SDC method.  For
the SDC method, we are using a second-order accurate method.  We
predict a time-centered advection term, $\Advs{\Uc}^{n+1/2}$ using
standard unsplit Godunov methods, but explicitly include a reactive
source term in the tracing of the interface states.  We can then solve
the reactive system, using this advective term as a
piecewise-constant-in-time source
\begin{equation}
\frac{d{\Uc}}{dt} = -\Advs{\Uc}^{n+1/2} + {\bf R}(\Uc)
\end{equation}
This system can be integrated with standard ODE methods.  To achieve
second-order accuracy, we need to iterate, using the updated $\Uc$ to
create a reactive source included in the predictor for the advective
update, and then reintegrate the system.  The result of this iteration
is that the advection sees the effects of reactions over the timestep
and the reactions see the effects of advection.  For the zone tracked
in Figure~\ref{fig:sdc}, each SDC iteration called the righthand side
function of the network 5 times less than the Strang case.  
We are continuing to develop this new integration strategy, with full
fourth-order in space and time reacting hydrodynamics methods to
follow shortly.

\begin{figure}
\centering
\includegraphics[width=0.75\linewidth]{sdc_plot}
\caption{\label{fig:sdc} Helium mass fraction as a function of time over three timesteps.}
\end{figure}

Other ways to make the simulation less expensive can include a subgrid model,
even something as simple as subzone burning.\MarginPar{need the paper ref fo rthis}

Finally, we can try hybrid model approaches where the fully
compressible simulation can act as a subgrid model.  \MarginPar{this ties into Ann's stuff, and also the stuff Alice was doing}

\ack The work at Stony Brook was supported by DOE/Office of Nuclear
Physics grant DE-FG02-87ER40317 and contract 7418390 with Lawrence
Berkeley National Laboratory as part of the Exascale Compute Project
ExaStar collaboration..  The work at LBNL was supported by the DOE
Office of Advanced Scientific Computing Research under Contract No,
DE-AC02-05CH11231. An award of computer time was provided by the
Innovative and Novel Computational Impact on Theory and Experiment
(INCITE) program.  This research used resources of the Oak Ridge
Leadership Computing Facility at the Oak Ridge National Laboratory,
which is supported by the Office of Science of the U.S. Department of
Energy under Contract No.\ DE-AC05-00OR22725.  This research used
resources of the National Energy Research Scientific Computing Center,
which is supported by the Office of Science of the U.S. Department of
Energy under Contract No.\ DE-AC02-05CH11231.  Visualizations were
done using yt~\cite{yt}.  This research has made use of NASA's
Astrophysics Data System Bibliographic Services.

\section*{References}

\bibliographystyle{iopart-num}
\bibliography{ws}

\end{document}

### 2018-09-16 astronum_2018/plots/sdc_plot.py

@@ -1,0 +1,59 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+# read the Strang data into a numpy array
+strang = np.loadtxt("strang_diag.out")
+
+# now manually read it looking for the times that indicate when the Strang halves began
+strang_starts = []
+strang_second = []
+with open("strang_diag.out") as f:
+    for line in f:
+        if line.startswith(" # start of step"):
+            _, time = line.split("=")
+            strang_starts.append(float(time))
+        elif line.startswith(" # second half of strang"):
+            _, time = line.split("=")
+            strang_second.append(float(time))
+
+strang_starts = sorted(set(strang_starts))
+strang_second = sorted(set(strang_second))
+
+print(strang_starts)
+
+# now get the SDC data
+sdc = np.loadtxt("sdc_diag.out")
+
+# we have 2 SDC iterations here, so be able to separate them
+idx_iter0 = sdc[:,1] == 0
+idx_iter1 = sdc[:,1] == 1
+
+
+# make the plots
+plt.scatter(strang[:,0], strang[:,2], marker="x", s=15, label="Strang splitting")
+plt.scatter(sdc[idx_iter1,0], sdc[idx_iter1,3], marker="o", s=15, label="SDC")
+
+# draw timestep boundaries
+for time in strang_starts:
+    plt.plot([time, time], [0,1], color="k", linewidth=2)
+
+for time in strang_second:
+    plt.plot([time, time], [0,1], color="0.5", linestyle=":", linewidth=2)
+
+plt.xlim(1.8685965733085516e-05,1.8999075234060896e-05)
+plt.ylim(0.5872, 0.5884)
+
+plt.legend(frameon=False)
+
+plt.tight_layout()
+
+f = plt.gcf()
+f.set_size_inches(9.0, 7.0)
+
+plt.xlabel("time (s)", fontsize="large")
+plt.ylabel(r"$X({}^4\mathrm{He})$", fontsize="large")
+
+plt.savefig("sdc_plot.png")
+
+print("sdc points = ", len(sdc[idx_iter1,0]))
+print("strang points = ", len(strang[:,0]))

### 2018-09-16 astronum_2018/astronum-xrb.tex
some updates

@@
 \newcommand{\Uc}{{\,\bm{\mathcal{U}}}}
+\newcommand{\Adv}[1]{{\left [\boldsymbol{\mathcal{A}} \left(#1\right)\right]}}
+\newcommand{\Advt}[1]{{\left [\mathcal{\tilde{A}} \left(#1\right)\right]}}
 \newcommand{\Advs}[1]{\boldsymbol{\mathcal{A}} \left(#1\right)}
@@
         A.~Harpole$^1$,
-        M.~P. Katz$^5$,
         J.~B. Bell$^2$,
         M.~Chang$^1$,
-        C.~M. Malone$^6$, and
+        M.~P. Katz$^5$,
+        C.~M. Malone$^6$, 
+        D.~E. Willcox$^2$, and
         W. Zhang$^2$}
@@
 
-Describe resolution needs and flame structure, Figure~\ref{fig:xrb}
+Describe resolution needs and flame structure, Figure~\ref{fig:xrb} \MarginPar{Mike will update this}
 
@@
 
-As an example of the improvements in coupling between hydro and
+As an example of the improvements in coupling between hydro and \MarginPar{Don can add some comments about startup costs?}
 reactions, Figure~\ref{fig:sdc} shows the mass fraction of helium
@@
 the SDC method, we are using a second-order accurate method.  We
-predict a time-centered advection term, $\Advs{\Uc}^{n+1/2}$ using
+predict a time-centered advection term, $\Adv{\Uc}^{n+1/2}_i$ using
 standard unsplit Godunov methods, but explicitly include a reactive
@@
 \begin{equation}
-\frac{d{\Uc}}{dt} = -\Advs{\Uc}^{n+1/2} + {\bf R}(\Uc)
+\frac{d{\Uc_i}}{dt} = -\Adv{\Uc}^{n+1/2}_i + {\bf R}(\Uc_i)
 \end{equation}
@@
 
+\section{Summary}
+
 \ack The work at Stony Brook was supported by DOE/Office of Nuclear

### 2018-09-23 astronum_2018/astronum-xrb.tex
add Andy
merge Yuri's changes

@@
         K.~Eiden$^1$,
-        Y.~Cavecchi$^2$,
+        Y.~Cavecchi$^{2,3}$,
         A.~Harpole$^1$,
-        J.~B. Bell$^2$,
+        J.~B. Bell$^4$,
         M.~Chang$^1$,
-        M.~P. Katz$^5$,
-        C.~M. Malone$^6$, 
-        D.~E. Willcox$^2$, and
-        W. Zhang$^2$}
+        M.~P. Katz$^7$,
+        C.~M. Malone$^8$, 
+        A.~J. Nonaka$^4$,
+        D.~E. Willcox$^4$, and
+        W. Zhang$^4$}
 
@@
 
-\address{$^2$Center for Computational Sciences and Engineering,
+\address{$^2$Department of Astrophysical Sciences, Princeton University,
+  Peyton Hall, Princeton, NJ 08544, USA}
+
+\address{$^3$Mathematical Sciences and STAG Research Centre,
+  University of Southampton, SO17 1BJ, UK}
+
+\address{$^4$Center for Computational Sciences and Engineering,
   Lawrence Berkeley National Lab, Berkeley, CA 94720 USA}
 
-\address{$^3$National Energy Research Scientific Computing Center,
+\address{$^5$National Energy Research Scientific Computing Center,
   Lawrence Berkeley National Lab, Berkeley, CA 94720 USA}
 
-\address{$^4$Department of Physics and Astronomy, Michigan State
+\address{$^6$Department of Physics and Astronomy, Michigan State
   University, East Lansing, Michigan 48824 USA}
 
-\address{$^5$NVIDIA Corporation, 2788 San Tomas Expressway,
+\address{$^7$NVIDIA Corporation, 2788 San Tomas Expressway,
   Santa Clara, CA, 95050 USA}
 
-\address{$^6$Los Alamos National Laboratory, Los Alamos, NM, 87545 USA}
+\address{$^8$Los Alamos National Laboratory, Los Alamos, NM, 87545 USA}
 
@@
 state~\cite{steiner:2010,ozel:2010}.  Finally, brightness oscillations
+during the rising phase of the bursts
 and other features seen in lightcurves are interpreted as arising from
@@
 the neutron star is 10~km and the Rossby lengthscale, where rotation
-and laterial pressure gradients balance is $\sim
+and lateral pressure gradients balance is $\sim
 1$~km~\cite{SPIT_ETAL02}.  At the smallest scales, a conductive He
@@
 approximations used enables a complementary picture of XRBs to be built up, 
-and has greatly advanced our understand of these systems.  
+and has greatly advanced our understanding of these systems.  
 The breadth of simulations to date has been impressive.  We summarize
@@
 
-One-dimensional simulations~\cite{woosley-xrb,fisker:2008} assume
+One-dimensional simulations, e.g. \cite{woosley-xrb,fisker:2008}, assume
 spherical symmetry and model the burst through the vertical column
@@
 Global shallow water simulations~\cite{SPIT_ETAL02} capture the
-largescale spreading of the burning front by using a very simple
+large scale spreading of the burning front by using a very simple
 vertical structure.  These simulations were instrumental in showing
@@
 regime, including the strength and character of the turbulence, but
-cannot model largescale lateral variations because of the implicit
+cannot model large scale lateral variations because of the implicit
 assumptions built into the low Mach number model~\cite{ABRZ:I}.
@@
 hydrostatic equilibrium is enforced and wide-aspect ratio zones give a
-horizontal CFL number that is large enough to get long timescale
-evolution~\cite{cavecchi:2012}.  These calculations... \MarginPar{Yuri can probably best summarize}
+horizontal CFL number that is large enough to obtain long timescale
+evolution~\cite{cavecchi:2012}.
+
+These calculations display key features of the propagation
+mechanism, namely a balance between hydrodynamics and the pure
+flame physics. The flame proceed mainly via conduction, thus being a
+deflagration. The flame front, however, is not vertical but inclined,
+the angle with the horizontal being very small: $\theta \sim
+H/R_{\rm{Ro}} \sim 10^{-3}$. Here $H$ is the scale height and
+$R_{\rm{Ro}}$ is the Rossby radius. It is the hydrodynamical balance
+between the Coriolis force, gravity and pressure gradients that
+determines the Rossby radius and therefore the inclination angle of the
+flame front. The extended surface of the flame front is what leads to
+propagations speed of order $10^5 $ km/s in good agreement with
+observations. Due to its dependence on the Coriolis force, the flame
+speed changes across the surface of the NS (being faster at the
+equator), leading to potentially detectable effects if the rise of the
+burst is observed with enough resolution
+\cite{art-2015-cavecchi-etal}. The NS systems exhibiting XRBs are
+expected to have magnetic fields in the range $10^7 - 10^{10}$
+G. Inclusion of a vertical magnetic field proved to have a strong
+influence on the nature and the speed of flame propagation
+\cite{art-2016-cavecchi-etal}, due to the interaction between the
+Coriolis force and the magnetic tension, which further acts towards
+determining the inclination angle of the flame front.
 
@@
 nucleosynthesis, and watch it move across the neutron star surface.
+The results of these simulations will enable strict comparison to
+observations made with the new generation of X-ray telescopes such as
+\textit{eXTP} and \textit{STROBE--X}
+\cite{art-2017-wilhod-etal,art-2016-zhang-etal} which will offer
+high collecting area and time resolution.
 We show some in-progress calculations that are a step toward this (but
@@
   simulation plane (middle), and nuclear energy generation rate
-  (bottom).  }
+  (bottom).
+  The out of plane velocity contours indicate where the Coriolis force
+  has created a hurricane-like wind structure which confines the flame,
+  leading to an inclined front. It is at such interface that the
+  temperature and burning rate are highest.
+  }
 \end{figure}
@@
 the burning layer, where the flame is the thinnest.  Using
-well-balanced schemes can help us retain HSE in the lower refined
-regions, reducing the number of zones needed to maintain fidelity.  We
-can also explore initialize newly created fine grids ahread of the burning
-with an understanding of hydrostatic equilibrium to allow us to only refine
-the peak burning as the flame moves across the star.
+well-balanced schemes can help us retain hydrostatic equilibrium (HSE)
+in the lower refined regions, reducing the number of zones needed to
+maintain fidelity.  We can also explore initializing newly created fine
+grids ahead of the burning with an understanding of hydrostatic
+equilibrium to allow us to only refine the peak burning as the flame
+moves across the star.
 
@@
 fourth-order (in space and time) coupling of hydrodynamics and
-reactions using the method of spectral deferred corrections in
+reactions using the method of spectral deferred corrections (SDC) in
 \castro, instead of the more commonly employed Strang splitting.  This
@@
 standard unsplit Godunov methods, but explicitly include a reactive
-source term in the tracing of the interface states.  We can then solve
+source term in the tracing of the interface states, ${\bf R}(\Uc_i)$.  We can then solve
 the reactive system, using this advective term as a
@@
 \begin{equation}
+\label{eq:syst}
 \frac{d{\Uc_i}}{dt} = -\Adv{\Uc}^{n+1/2}_i + {\bf R}(\Uc_i)
@@
 \includegraphics[width=0.75\linewidth]{sdc_plot}
-\caption{\label{fig:sdc} Helium mass fraction as a function of time over three timesteps.}
+\caption{\label{fig:sdc} Helium mass fraction as a function of time over three timesteps.
+  The points represent individual calls to the
+  righthand side function of our reaction network (see Eq. \ref{eq:syst}) and the density of
+  points indicates how hard the integrator is working.}
 \end{figure}
@@
 Berkeley National Laboratory as part of the Exascale Compute Project
-ExaStar collaboration..  The work at LBNL was supported by the DOE
+ExaStar collaboration.  The work at LBNL was supported by the DOE
 Office of Advanced Scientific Computing Research under Contract No,
@@
 Innovative and Novel Computational Impact on Theory and Experiment
-(INCITE) program.  This research used resources of the Oak Ridge
-Leadership Computing Facility at the Oak Ridge National Laboratory,
-which is supported by the Office of Science of the U.S. Department of
-Energy under Contract No.\ DE-AC05-00OR22725.  This research used
-resources of the National Energy Research Scientific Computing Center,
-which is supported by the Office of Science of the U.S. Department of
-Energy under Contract No.\ DE-AC02-05CH11231.  Visualizations were
-done using yt~\cite{yt}.  This research has made use of NASA's
-Astrophysics Data System Bibliographic Services.
+(INCITE) program. YC is supported by the European Union Horizon 2020
+research and innovation programme under the Marie Sklodowska-Curie
+Global Fellowship grant agreement No 703916. This research used
+resources of the Oak Ridge Leadership Computing Facility at the Oak
+Ridge National Laboratory, which is supported by the Office of Science
+of the U.S. Department of Energy under Contract
+No.\ DE-AC05-00OR22725.  This research used resources of the National
+Energy Research Scientific Computing Center, which is supported by the
+Office of Science of the U.S. Department of Energy under Contract
+No.\ DE-AC02-05CH11231.  Visualizations were done using yt~\cite{yt}.
+This research has made use of NASA's Astrophysics Data System
+Bibliographic Services.
 

### 2018-09-24 astronum_2018/astronum-xrb.tex
Minor typo corrections and comments

@@
 the range of length and timescales involved.  At the largest scale,
-the neutron star is 10~km and the Rossby lengthscale, where rotation
-and lateral pressure gradients balance is $\sim
+the neutron star radius is 10~km and the Rossby lengthscale, where rotation
+and lateral pressure gradients balance, is $\sim
 1$~km~\cite{SPIT_ETAL02}.  At the smallest scales, a conductive He
@@
 pressure scale height of the fuel layer is meters.  To capture the
-rise of the XRB, we would need to model seconds, but the flame itself
+rise of the XRB\MarginPar{Define rise?}, we would need to model seconds, but the flame itself
 is subsonic, so long timescale evolution is difficult.
@@
 depth of fuel.  These simulations can capture the luminosities and
-burst recurrence times well.  The can utilize very large nuclear
+burst recurrence times well.  They can utilize very large nuclear
 reaction networks and have been used to understand the nucleosynthesis
@@
 bursts~\cite{xrb3}, and followed the development of the turbulence.
-These simulations can give an understand of the pre-burning front
+These simulations can give an understanding of the pre-burning front
 regime, including the strength and character of the turbulence, but
@@
 mechanism, namely a balance between hydrodynamics and the pure
-flame physics. The flame proceed mainly via conduction, thus being a
+flame physics. The flame proceeds mainly via conduction, thus being a
 deflagration. The flame front, however, is not vertical but inclined,
@@
 flame front. The extended surface of the flame front is what leads to
-propagations speed of order $10^5 $ km/s in good agreement with
+propagation speeds of order $10^5 $~km/s, in good agreement with
 observations. Due to its dependence on the Coriolis force, the flame
@@
 \cite{art-2015-cavecchi-etal}. The NS systems exhibiting XRBs are
-expected to have magnetic fields in the range $10^7 - 10^{10}$
-G. Inclusion of a vertical magnetic field proved to have a strong
+expected to have magnetic fields in the range $10^7 - 10^{10}$~G.
+Inclusion of a vertical magnetic field proved to have a strong
 influence on the nature and the speed of flame propagation

### 2018-09-26 astronum_2018/astronum-xrb.tex
Add IH affiliation.
Minor rephrasing or restructuring.

@@
 that the Coriolis force plays an important role in confining a
-spreading region and the flame spreading can explain brightness
+spreading region and that the flame spreading can explain brightness
 oscillations during the rise of the burst lightcurve.
@@
 evolution~\cite{cavecchi:2012}.
-
+%
 These calculations display key features of the propagation
@@
 burst is observed with enough resolution
-\cite{art-2015-cavecchi-etal}. The NS systems exhibiting XRBs are
+\cite{art-2015-cavecchi-etal}.
+
+The NS systems exhibiting XRBs are
 expected to have magnetic fields in the range $10^7 - 10^{10}$~G.
@@
 high collecting area and time resolution.
-We show some in-progress calculations that are a step toward this (but
-still have approximations) and also discuss what new algorithmic
+We show some in-progress calculations that are a step toward this, discuss
+their remaining approximations, and also discuss what new algorithmic
 techniques might be needed to realize this goal with the advent of
@@
   has created a hurricane-like wind structure which confines the flame,
-  leading to an inclined front. It is at such interface that the
+  leading to an inclined front. It is at such an interface that the
   temperature and burning rate are highest.

### 2018-09-26 astronum_2018/astronum-xrb.tex
Being slightly more explicit about scales.
I think this makes defining "rise" unnecessary.

--- 2018-09-24 astronum_2018/astronum-xrb.tex
@@
 many technical challenges remain.  The primary difficulties are simply
-the range of length and timescales involved.  At the largest scale,
+the range of length and timescales involved.  At the largest lengthscale,
 the neutron star radius is 10~km and the Rossby lengthscale, where rotation
@@
 pressure scale height of the fuel layer is meters.  To capture the
-rise of the XRB\MarginPar{Define rise?}, we would need to model seconds, but the flame itself
+rise timescale of the XRB, we would need to model seconds, but the flame itself
 is subsonic, so long timescale evolution is difficult.

### 2018-09-26 astronum_2018/astronum-xrb.tex

--- 2018-09-26 astronum_2018/astronum-xrb.tex
@@
         M.~Chang$^1$,
+        I.Hawke$^3$,
         M.~P. Katz$^7$,

### 2018-09-26 astronum_2018/astronum-xrb.tex
update

@@
 
+Within the realm of compressible hydrodynamics, there are a number of other techniques that 
+should be explored. 
+smarter refinement with AMR + hydrostatic refinement
+well-balanced method to allow ultralow resolution to slide along with the burning front
+higher-order
+This are all
+
 Other ways to make the simulation less expensive can include a subgrid model,
 even something as simple as subzone burning.\MarginPar{need the paper ref fo rthis}
+This is harder...
 
@@
 compressible simulation can act as a subgrid model.  \MarginPar{this ties into Ann's stuff, and also the stuff Alice was doing}
+Multiscale methods can combine different hydro models together to allow 
+for a full star simulation with resolved burning.  Examples include...
 

### 2018-09-26 astronum_2018/astronum-xrb.tex
more simulation work

@@
         M.~Chang$^1$,
-        I.Hawke$^3$,
+        I.~Hawke$^3$,
         M.~P. Katz$^7$,
@@
 \begin{abstract}
-We discuss the challenges of modeling X-ray bursts, review the different
-calculations done to date, and discuss our new set of simulations.  We also
-describe algorithmic improvements that may help in the future to offset some
-of the expense of these simulations.
+We discuss the challenges of modeling X-ray bursts in
+multi-dimensions, review the different calculations done to date, and
+discuss our new set of ongoing simulations.  We also describe
+algorithmic improvements that may help in the future to offset some of
+the expense of these simulations, and describe what may be possible
+with exascale computing.
 \end{abstract}
@@
 reaction networks and have been used to understand the nucleosynthesis
-and rp-process, as well as to understand how rate uncertainties can
+and rp-process~\cite{shatz_rp}, as well as to understand how rate uncertainties can
 affect the outcomes of the bursts.  Many successive bursts can be
@@
 evolution~\cite{cavecchi:2012}.
-%
 These calculations display key features of the propagation
@@
 The NS systems exhibiting XRBs are
-expected to have magnetic fields in the range $10^7 - 10^{10}$~G.
+expected to have magnetic fields in the range $10^7$--$10^{10}$~G.
 Inclusion of a vertical magnetic field proved to have a strong
@@
 
-Short description of problem setup and approximations
-rotation, network, flame boosting
+We setup a problem by creating a pair of 1D hydrostatic XRB models
+representing the post- and pre-flame states and apply these laterally
+across the domain to setup a hot region at the left of our domain in
+equilibrium with the cool model.  Thermal diffusion and nuclear
+reactions quickly setup a laterally propagating flame that begins to
+move through our domain.  We resolve the diffusion scale, using a
+resolution of 10~cm, and use adaptive mesh refinement (two jumps of
+$4\times$) to keep a buffer of low density material at low resolution
+above the surface to allow for expansion.  To make the lengthscales
+tractable, we rotate the neutron star at 2000~Hz---much faster than
+expected.  This helps confine the spreading region to a more
+reasonable horizontal region.  Nevertheless, our simulation uses 12288
+zones laterally, on the finest grid.  We use the stellar
+conducitivities from \cite{Timmes00} and a 13-isotopic alpha-chain
+reaction network.  To make the timescales easier, we boost the flame
+speed by a factor of 10, by boosting both the conductivity and energy
+generation rate.
 
-Describe resolution needs and flame structure, Figure~\ref{fig:xrb} \MarginPar{Mike will update this}
+Figure~\ref{fig:xrb} \MarginPar{Mike will update this}
 

### 2018-09-26 astronum_2018/astronum-xrb.tex
updated fig

@@
 
-Figure~\ref{fig:xrb} \MarginPar{Mike will update this}
+Figure~\ref{fig:xrb} shows the flame structure.  The top panel shows
+the temperature.  The initial perturbation reached to $2\times
+10^4~\mathrm{cm}$, so at this point, the extend of the burnt region
+has nearly doubled.  The flame structure is quite wide
 
@@
 \centering
-\includegraphics[width=0.75\linewidth]{xrb}
+\includegraphics[width=0.95\linewidth]{xrb}
 \caption{\label{fig:xrb} Helium flame spreading across the surface of
@@
 
-Within the realm of compressible hydrodynamics, there are a number of other techniques that 
-should be explored. 
-smarter refinement with AMR + hydrostatic refinement
-well-balanced method to allow ultralow resolution to slide along with the burning front
-higher-order
-This are all
+Within the realm of compressible hydrodynamics, there are a number of
+other techniques that should be explored.  smarter refinement with AMR
++ hydrostatic refinement well-balanced method to allow ultralow
+resolution to slide along with the burning front higher-order This are
+all
 
-Other ways to make the simulation less expensive can include a subgrid model,
-even something as simple as subzone burning.\MarginPar{need the paper ref fo rthis}
-This is harder...
+Other ways to make the simulation less expensive can include a subgrid
+model, Flame models are commonly used for Type Ia supernovae.  There
+however, the flame is thin compared to the scale height, so a simple
+model can represent the unresolved flame on the grid.  For the XRB,
+the flame thickness is a non-trivial fraction of the pressure scale
+height and further, the laminar flame properties will change strongly
+over the height of the burning layer (see~\cite{Timmes00}).  This
+makes traditional flame models a poor fit.  Methods that use even
+something as simple as subzone burning.  This is harder...
 

### 2018-09-26 astronum_2018/astronum-xrb.tex
more plot work

@@
 representing the post- and pre-flame states and apply these laterally
-across the domain to setup a hot region at the left of our domain in
-equilibrium with the cool model.  Thermal diffusion and nuclear
+across the domain to setup a hot region at the left end of our domain
+in equilibrium with the cool model.  Thermal diffusion and nuclear
 reactions quickly setup a laterally propagating flame that begins to
@@
 speed by a factor of 10, by boosting both the conductivity and energy
-generation rate.
+generation rate.  This is an approximation we hope to relax in the
+near future.
 
-Figure~\ref{fig:xrb} shows the flame structure.  The top panel shows
-the temperature.  The initial perturbation reached to $2\times
-10^4~\mathrm{cm}$, so at this point, the extend of the burnt region
-has nearly doubled.  The flame structure is quite wide
+Figure~\ref{fig:xrb} shows the flame structure for an initial 2D run.
+The top panel shows the temperature.  The initial perturbation reached
+to $2\times 10^4~\mathrm{cm}$, so at this point, the extent of the
+burnt region has nearly doubled.  The flame structure is quite wide,
+since there are multiple burning stages beyond just helium to carbon, and
+this shows up in the composition.
 

### 2018-09-26 astronum_2018/astronum-xrb.tex
more text

@@
 \newcommand{\Advs}[1]{\boldsymbol{\mathcal{A}} \left(#1\right)}
+
+\newcommand{\isot}[2]{$^{#2}\mathrm{#1}$}
+
+\newcommand{\gcc}{\mathrm{g~cm^{-3} }}
+\newcommand{\cms}{\mathrm{cm~s^{-1} }}
 
@@
 
-Figure~\ref{fig:xrb} shows the flame structure for an initial 2D run.
-The top panel shows the temperature.  The initial perturbation reached
-to $2\times 10^4~\mathrm{cm}$, so at this point, the extent of the
-burnt region has nearly doubled.  The flame structure is quite wide,
-since there are multiple burning stages beyond just helium to carbon, and
-this shows up in the composition.
+Figure~\ref{fig:xrb} shows the flame structure for an initial 2D run
+at 0.007~s.  The burning layer is only about 10~m high.  Underneath is
+\isot{Ni}{56}, representative of the underlying neutron star, and
+above it is very a low density ($10^{-4}~\gcc$) buffer.  Only a
+portion of the vertical extent is show.  The top panel shows the
+temperature.  The initial perturbation reached to $2\times
+10^4~\mathrm{cm}$, so at this point, the extent of the burnt region
+has nearly doubled.  The flame structure is quite wide, since there
+are multiple burning stages beyond just helium to carbon, and this
+shows up in the composition.  This is seen more clearly in the lower
+panel which shows the mean molecular weight of the nuclei ($\bar{A}$).
+We see the further behind the head of the flame we are, the heavier
+the ash nuclei.  The second panel shows the energy generation rate.
+It is strongest just above the neutron star surface, behind the flame.
+We also see that the head of the flame is ``detached'' from the base
+of the atmosphere, with the strongest burning at the head of the flame
+slightly higher up in the atmosphere.  This can also be more clearly
+seen in the last panel that shows the ash structure.  This appears
+reminiscent of the flame structure seen in the earlier simulations of
+\cite{cavecchi:2012} \MarginPar{Yuri, would you agree?}  The third
+panel shows the velocity through the plane of the simulation---this is
+induced by the Coriolis force as the burning front spreads laterally.
+A tight hurricane structure has setup as a result of the flame
+spreading.  This is the geostrophic balance discussed in
+\cite{SPIT_ETAL02}.  Finally, as already mentioned, the last panel
+shows the ash composition.  These calculations are ongoing and will be
+the subject of a detailed study in the near future.
 

### 2018-09-26 astronum_2018/astronum-xrb.tex
more discussion of cost

@@
 
-\section{Estimates for Full Star Simulations}
+\section{Large Scale Simulations and Algorithmic Developments}
 
-A fully resolved calculation of burning over the entire neutron star
-is out of reach, and will be for some time.
-
-time estimate...
-
-\MarginPar{Alice and Mike will work on this}
-
-\section{Algorithmic Developments}
+The 2D calculations shown in the previous section require about 10k
+node-hours of computational time to get to about 10~ms (those
+particular simulations were run at NERSC).  The GPU version of the
+Castro hydrodynamics solver, under active development, is about
+$10\times$ faster on a comparable node, so a GPU calculation could
+bring the 2D cost down to 1k node-hours.  Most of this time is spent
+in the reactions\MarginPar{verify fraction}.  A 3D version of this
+calculation is estimated to require 5M node-hours on CPUs or 0.5M
+node-hours using GPUs---this is about the size of an
+annual \MarginPar{double check \#s} allocation.  Additionally, this is
+still with the boosted flame, but we expect that once the flame is
+developed, the boosting will not be as critical, due to hydrodynamic
+effects.  We plan to test this soon.  In any case, this is less than
+1\% of the neutron star surface, so a fully resolved calculation of
+burning over the entire neutron star is out of reach, and will be for
+some time.
 

### 2018-09-26 astronum_2018/astronum-xrb.tex
some updates

--- 2018-09-26 astronum_2018/astronum-xrb.tex
@@
         K.~Eiden$^1$,
-        Y.~Cavecchi$^{2,3}$,
+        Y.~Cavecchi$^2$,
         A.~Harpole$^1$,
-        J.~B. Bell$^4$,
+        J.~B. Bell$^2$,
         M.~Chang$^1$,
-        I.~Hawke$^3$,
-        M.~P. Katz$^7$,
-        C.~M. Malone$^8$,
-        A.~J. Nonaka$^4$,
-        D.~E. Willcox$^4$, and
-        W. Zhang$^4$}
+        M.~P. Katz$^5$,
+        C.~M. Malone$^6$, 
+        A.~J. Nonaka$^2$,
+        D.~E. Willcox$^2$, and
+        W. Zhang$^2$}
 
@@
 
-\address{$^2$Department of Astrophysical Sciences, Princeton University,
-  Peyton Hall, Princeton, NJ 08544, USA}
-
-\address{$^3$Mathematical Sciences and STAG Research Centre,
-  University of Southampton, SO17 1BJ, UK}
-
-\address{$^4$Center for Computational Sciences and Engineering,
+\address{$^2$Center for Computational Sciences and Engineering,
   Lawrence Berkeley National Lab, Berkeley, CA 94720 USA}
 
-\address{$^5$National Energy Research Scientific Computing Center,
+\address{$^3$National Energy Research Scientific Computing Center,
   Lawrence Berkeley National Lab, Berkeley, CA 94720 USA}
 
-\address{$^6$Department of Physics and Astronomy, Michigan State
+\address{$^4$Department of Physics and Astronomy, Michigan State
   University, East Lansing, Michigan 48824 USA}
 
-\address{$^7$NVIDIA Corporation, 2788 San Tomas Expressway,
+\address{$^5$NVIDIA Corporation, 2788 San Tomas Expressway,
   Santa Clara, CA, 95050 USA}
 
-\address{$^8$Los Alamos National Laboratory, Los Alamos, NM, 87545 USA}
+\address{$^6$Los Alamos National Laboratory, Los Alamos, NM, 87545 USA}
 
 \begin{abstract}
-We discuss the challenges of modeling X-ray bursts in
-multi-dimensions, review the different calculations done to date, and
-discuss our new set of ongoing simulations.  We also describe
-algorithmic improvements that may help in the future to offset some of
-the expense of these simulations, and describe what may be possible
-with exascale computing.
+We discuss the challenges of modeling X-ray bursts, review the different
+calculations done to date, and discuss our new set of simulations.  We also
+describe algorithmic improvements that may help in the future to offset some
+of the expense of these simulations.
 \end{abstract}
@@
 state~\cite{steiner:2010,ozel:2010}.  Finally, brightness oscillations
-during the rising phase of the bursts
 and other features seen in lightcurves are interpreted as arising from
@@
 many technical challenges remain.  The primary difficulties are simply
-the range of length and timescales involved.  At the largest lengthscale,
-the neutron star radius is 10~km and the Rossby lengthscale, where rotation
-and lateral pressure gradients balance, is $\sim
+the range of length and timescales involved.  At the largest scale,
+the neutron star is 10~km and the Rossby lengthscale, where rotation
+and laterial pressure gradients balance is $\sim
 1$~km~\cite{SPIT_ETAL02}.  At the smallest scales, a conductive He
@@
 pressure scale height of the fuel layer is meters.  To capture the
-rise timescale of the XRB, we would need to model seconds, but the flame itself
+rise of the XRB, we would need to model seconds, but the flame itself
 is subsonic, so long timescale evolution is difficult.
@@
 approximations to get past these difficulties.  The variety of
-approximations used enables a complementary picture of XRBs to be built up,
-and has greatly advanced our understanding of these systems.
+approximations used enables a complementary picture of XRBs to be
+built up, and has greatly advanced our understand of these systems.
 The breadth of simulations to date has been impressive.  We summarize
@@
 
-One-dimensional simulations, e.g. \cite{woosley-xrb,fisker:2008}, assume
+One-dimensional simulations~\cite{woosley-xrb,fisker:2008} assume
 spherical symmetry and model the burst through the vertical column
 depth of fuel.  These simulations can capture the luminosities and
-burst recurrence times well.  They can utilize very large nuclear
+burst recurrence times well.  The can utilize very large nuclear
 reaction networks and have been used to understand the nucleosynthesis
-and rp-process~\cite{shatz_rp}, as well as to understand how rate uncertainties can
+and rp-process, as well as to understand how rate uncertainties can
 affect the outcomes of the bursts.  Many successive bursts can be
@@
 Global shallow water simulations~\cite{SPIT_ETAL02} capture the
-large scale spreading of the burning front by using a very simple
+largescale spreading of the burning front by using a very simple
 vertical structure.  These simulations were instrumental in showing
 that the Coriolis force plays an important role in confining a
-spreading region and that the flame spreading can explain brightness
+spreading region and the flame spreading can explain brightness
 oscillations during the rise of the burst lightcurve.
@@
 bursts~\cite{xrb3}, and followed the development of the turbulence.
-These simulations can give an understanding of the pre-burning front
+These simulations can give an understand of the pre-burning front
 regime, including the strength and character of the turbulence, but
-cannot model large scale lateral variations because of the implicit
+cannot model largescale lateral variations because of the implicit
 assumptions built into the low Mach number model~\cite{ABRZ:I}.
@@
 hydrostatic equilibrium is enforced and wide-aspect ratio zones give a
-horizontal CFL number that is large enough to obtain long timescale
-evolution~\cite{cavecchi:2012}.
-These calculations display key features of the propagation
-mechanism, namely a balance between hydrodynamics and the pure
-flame physics. The flame proceeds mainly via conduction, thus being a
-deflagration. The flame front, however, is not vertical but inclined,
-the angle with the horizontal being very small: $\theta \sim
-H/R_{\rm{Ro}} \sim 10^{-3}$. Here $H$ is the scale height and
-$R_{\rm{Ro}}$ is the Rossby radius. It is the hydrodynamical balance
-between the Coriolis force, gravity and pressure gradients that
-determines the Rossby radius and therefore the inclination angle of the
-flame front. The extended surface of the flame front is what leads to
-propagation speeds of order $10^5 $~km/s, in good agreement with
-observations. Due to its dependence on the Coriolis force, the flame
-speed changes across the surface of the NS (being faster at the
-equator), leading to potentially detectable effects if the rise of the
-burst is observed with enough resolution
-\cite{art-2015-cavecchi-etal}.
-
-The NS systems exhibiting XRBs are
-expected to have magnetic fields in the range $10^7$--$10^{10}$~G.
-Inclusion of a vertical magnetic field proved to have a strong
-influence on the nature and the speed of flame propagation
-\cite{art-2016-cavecchi-etal}, due to the interaction between the
-Coriolis force and the magnetic tension, which further acts towards
-determining the inclination angle of the flame front.
+horizontal CFL number that is large enough to get long timescale
+evolution~\cite{cavecchi:2012}.  These calculations... \MarginPar{Yuri
+  can probably best summarize}
 
@@
 nucleosynthesis, and watch it move across the neutron star surface.
-The results of these simulations will enable strict comparison to
-observations made with the new generation of X-ray telescopes such as
-\textit{eXTP} and \textit{STROBE--X}
-\cite{art-2017-wilhod-etal,art-2016-zhang-etal} which will offer
-high collecting area and time resolution.
-We show some in-progress calculations that are a step toward this, discuss
-their remaining approximations, and also discuss what new algorithmic
+We show some in-progress calculations that are a step toward this (but
+still have approximations) and also discuss what new algorithmic
 techniques might be needed to realize this goal with the advent of
@@
 
-We setup a problem by creating a pair of 1D hydrostatic XRB models
-representing the post- and pre-flame states and apply these laterally
-across the domain to setup a hot region at the left end of our domain
-in equilibrium with the cool model.  Thermal diffusion and nuclear
-reactions quickly setup a laterally propagating flame that begins to
-move through our domain.  We resolve the diffusion scale, using a
-resolution of 10~cm, and use adaptive mesh refinement (two jumps of
-$4\times$) to keep a buffer of low density material at low resolution
-above the surface to allow for expansion.  To make the lengthscales
-tractable, we rotate the neutron star at 2000~Hz---much faster than
-expected.  This helps confine the spreading region to a more
-reasonable horizontal region.  Nevertheless, our simulation uses 12288
-zones laterally, on the finest grid.  We use the stellar
-conducitivities from \cite{Timmes00} and a 13-isotopic alpha-chain
-reaction network.  To make the timescales easier, we boost the flame
-speed by a factor of 10, by boosting both the conductivity and energy
-generation rate.  This is an approximation we hope to relax in the
-near future.
-
-Figure~\ref{fig:xrb} shows the flame structure for an initial 2D run.
-The top panel shows the temperature.  The initial perturbation reached
-to $2\times 10^4~\mathrm{cm}$, so at this point, the extent of the
-burnt region has nearly doubled.  The flame structure is quite wide,
-since there are multiple burning stages beyond just helium to carbon, and
-this shows up in the composition.
+As described above, current simulations of XRBs must approximate the
+system somehow.  Our goal is resolve the flame structure and
+understand the details of the nucleosynthesis.  This sets the needed
+spatial resolution.  We use a 13-isotope alpha-chain, which should
+capture the energetics for a helium flame.  To make the range of
+lengthscales tractable, we compress the horizontal scale by
+considering a neutron star that rotates faster than normal.  This
+reduces the Rossby length.  We also artificially boost the flame speed
+by a factor of 10 to reduce the length of time needed.  We expect that
+once a steady-state burning front develops that the flame speed should
+be faster than the laminar speed.  The flame is initiated by placing a
+hot region in equilibrium on the left part of the domain.  Reactions
+then heat the fuel and thermal diffusion propagates the flame
+laterally across the neutron star surface.
 
@@
 \centering
-\includegraphics[width=0.95\linewidth]{xrb}
+\includegraphics[width=0.75\linewidth]{xrb}
 \caption{\label{fig:xrb} Helium flame spreading across the surface of
@@
   simulation plane (middle), and nuclear energy generation rate
-  (bottom).
-  The out of plane velocity contours indicate where the Coriolis force
-  has created a hurricane-like wind structure which confines the flame,
-  leading to an inclined front. It is at such an interface that the
-  temperature and burning rate are highest.
-  }
+  (bottom).  }
 \end{figure}
+
+Figure~\ref{fig:xrb} shows a snapshot in time
 
@@
 the burning layer, where the flame is the thinnest.  Using
-well-balanced schemes can help us retain hydrostatic equilibrium (HSE)
-in the lower refined regions, reducing the number of zones needed to
-maintain fidelity.  We can also explore initializing newly created fine
-grids ahead of the burning with an understanding of hydrostatic
-equilibrium to allow us to only refine the peak burning as the flame
-moves across the star.
+well-balanced schemes can help us retain HSE in the lower refined
+regions, reducing the number of zones needed to maintain fidelity.  We
+can also explore initialize newly created fine grids ahread of the burning
+with an understanding of hydrostatic equilibrium to allow us to only refine
+the peak burning as the flame moves across the star.
 
@@
 fourth-order (in space and time) coupling of hydrodynamics and
-reactions using the method of spectral deferred corrections (SDC) in
+reactions using the method of spectral deferred corrections in
 \castro, instead of the more commonly employed Strang splitting.  This
@@
 standard unsplit Godunov methods, but explicitly include a reactive
-source term in the tracing of the interface states, ${\bf R}(\Uc_i)$.  We can then solve
+source term in the tracing of the interface states.  We can then solve
 the reactive system, using this advective term as a
@@
 \begin{equation}
-\label{eq:syst}
 \frac{d{\Uc_i}}{dt} = -\Adv{\Uc}^{n+1/2}_i + {\bf R}(\Uc_i)
@@
 \includegraphics[width=0.75\linewidth]{sdc_plot}
-\caption{\label{fig:sdc} Helium mass fraction as a function of time over three timesteps.
-  The points represent individual calls to the
-  righthand side function of our reaction network (see Eq. \ref{eq:syst}) and the density of
-  points indicates how hard the integrator is working.}
+\caption{\label{fig:sdc} Helium mass fraction as a function of time over three timesteps.}
 \end{figure}
 
-Within the realm of compressible hydrodynamics, there are a number of
-other techniques that should be explored.  smarter refinement with AMR
-+ hydrostatic refinement well-balanced method to allow ultralow
-resolution to slide along with the burning front higher-order This are
-all
-
-Other ways to make the simulation less expensive can include a subgrid
-model, Flame models are commonly used for Type Ia supernovae.  There
-however, the flame is thin compared to the scale height, so a simple
-model can represent the unresolved flame on the grid.  For the XRB,
-the flame thickness is a non-trivial fraction of the pressure scale
-height and further, the laminar flame properties will change strongly
-over the height of the burning layer (see~\cite{Timmes00}).  This
-makes traditional flame models a poor fit.  Methods that use even
-something as simple as subzone burning.  This is harder...
+Other ways to make the simulation less expensive can include a subgrid model,
+even something as simple as subzone burning.\MarginPar{need the paper ref fo rthis}
 
@@
 compressible simulation can act as a subgrid model.  \MarginPar{this ties into Ann's stuff, and also the stuff Alice was doing}
-Multiscale methods can combine different hydro models together to allow 
-for a full star simulation with resolved burning.  Examples include...
 
@@
 Berkeley National Laboratory as part of the Exascale Compute Project
-ExaStar collaboration.  The work at LBNL was supported by the DOE
+ExaStar collaboration..  The work at LBNL was supported by the DOE
 Office of Advanced Scientific Computing Research under Contract No,
@@
 Innovative and Novel Computational Impact on Theory and Experiment
-(INCITE) program. YC is supported by the European Union Horizon 2020
-research and innovation programme under the Marie Sklodowska-Curie
-Global Fellowship grant agreement No 703916. This research used
-resources of the Oak Ridge Leadership Computing Facility at the Oak
-Ridge National Laboratory, which is supported by the Office of Science
-of the U.S. Department of Energy under Contract
-No.\ DE-AC05-00OR22725.  This research used resources of the National
-Energy Research Scientific Computing Center, which is supported by the
-Office of Science of the U.S. Department of Energy under Contract
-No.\ DE-AC02-05CH11231.  Visualizations were done using yt~\cite{yt}.
-This research has made use of NASA's Astrophysics Data System
-Bibliographic Services.
+(INCITE) program.  This research used resources of the Oak Ridge
+Leadership Computing Facility at the Oak Ridge National Laboratory,
+which is supported by the Office of Science of the U.S. Department of
+Energy under Contract No.\ DE-AC05-00OR22725.  This research used
+resources of the National Energy Research Scientific Computing Center,
+which is supported by the Office of Science of the U.S. Department of
+Energy under Contract No.\ DE-AC02-05CH11231.  Visualizations were
+done using yt~\cite{yt}.  This research has made use of NASA's
+Astrophysics Data System Bibliographic Services.
 

### 2018-09-26 astronum_2018/astronum-xrb.tex
better caption

--- 2018-09-26 astronum_2018/astronum-xrb.tex
@@
 \caption{\label{fig:xrb} Helium flame spreading across the surface of
-  a neutron star.  Shown are temperature (top), the velocity out of the
-  simulation plane (middle), and nuclear energy generation rate
-  (bottom).
-  The out of plane velocity contours indicate where the Coriolis force
-  has created a hurricane-like wind structure which confines the flame,
+  a neutron star.  Shown are temperature (top), the nuclear energy
+  generation rate (second), the velocity out of the simulation plane
+  (third) and the mean molecular weight of the ash (bottom).  The out
+  of plane velocity contours indicate where the Coriolis force has
+  created a hurricane-like wind structure which confines the flame,
   leading to an inclined front. It is at such an interface that the
-  temperature and burning rate are highest.
-  }
+  temperature and burning rate are highest.  }
 \end{figure}
+
+Figure~\ref{fig:xrb} shows a snapshot in time
 

### 2018-09-27 astronum_2018/astronum-xrb.tex
minor edit
fix acks
clean up some descriptions

--- 2018-09-26 astronum_2018/astronum-xrb.tex
@@
 reaction networks and have been used to understand the nucleosynthesis
-and rp-process~\cite{shatz_rp}, as well as to understand how rate uncertainties can
+and rp-process~\cite{schatz:rp1999,rpprocess,schatz_rp}, as well as to understand how rate uncertainties can
 affect the outcomes of the bursts.  Many successive bursts can be
@@
 observations made with the new generation of X-ray telescopes such as
-\textit{eXTP} and \textit{STROBE--X}
+\textit{eXTP}\/ and \textit{STROBE-X}\/
 \cite{art-2017-wilhod-etal,art-2016-zhang-etal} which will offer
@@
 
-We setup a problem by creating a pair of 1D hydrostatic XRB models
+We setup a problem by creating a pair of 1D hydrostatic XRB models (hot and cool)
 representing the post- and pre-flame states and apply these laterally
@@
 portion of the vertical extent is show.  The top panel shows the
-temperature.  The initial perturbation reached to $2\times
+temperature structure.  The initial perturbation reached to $2\times
 10^4~\mathrm{cm}$, so at this point, the extent of the burnt region
@@
 are multiple burning stages beyond just helium to carbon, and this
-shows up in the composition.  This is seen more clearly in the lower
+shows up in the composition.  This is seen more clearly in the next 
 panel which shows the mean molecular weight of the nuclei ($\bar{A}$).
 We see the further behind the head of the flame we are, the heavier
-the ash nuclei.  The second panel shows the energy generation rate.
+the ash nuclei.  We also see where that the head of the flame
+appears lifted off of the base of the burning layer---unburned
+material underlies the spreading flame.  This appears
+reminiscent of the flame structure seen in the earlier simulations of
+\cite{cavecchi:2012}. \MarginPar{Yuri, would you agree?} The third panel shows the energy generation rate.
 It is strongest just above the neutron star surface, behind the flame.
-We also see that the head of the flame is ``detached'' from the base
+We again see that the head of the flame is detached from the base
 of the atmosphere, with the strongest burning at the head of the flame
-slightly higher up in the atmosphere.  This can also be more clearly
-seen in the last panel that shows the ash structure.  This appears
-reminiscent of the flame structure seen in the earlier simulations of
-\cite{cavecchi:2012} \MarginPar{Yuri, would you agree?}  The third
+slightly higher up in the atmosphere.   Finally, the bottom
 panel shows the velocity through the plane of the simulation---this is
@@
 spreading.  This is the geostrophic balance discussed in
-\cite{SPIT_ETAL02}.  Finally, as already mentioned, the last panel
-shows the ash composition.  These calculations are ongoing and will be
+\cite{SPIT_ETAL02}.    These calculations are ongoing and will be
 the subject of a detailed study in the near future.
@@
 \caption{\label{fig:xrb} Helium flame spreading across the surface of
-  a neutron star.  Shown are temperature (top), the velocity out of the
-  simulation plane (middle), and nuclear energy generation rate
-  (bottom).
-  The out of plane velocity contours indicate where the Coriolis force
-  has created a hurricane-like wind structure which confines the flame,
+  a neutron star.  Shown are temperature (top), the mean molecular
+  weight of the ash (second), the nuclear energy generation rate
+  (third), and the velocity out of the simulation plane (bottom).  The
+  out of plane velocity contours indicate where the Coriolis force has
+  created a hurricane-like wind structure which confines the flame,
   leading to an inclined front. It is at such an interface that the
-  temperature and burning rate are highest.
-  }
+  temperature and burning rate are highest.  }
 \end{figure}
@@
 Office of Advanced Scientific Computing Research under Contract No,
-DE-AC02-05CH11231. An award of computer time was provided by the
-Innovative and Novel Computational Impact on Theory and Experiment
-(INCITE) program. YC is supported by the European Union Horizon 2020
+DE-AC02-05CH11231. YC is supported by the European Union Horizon 2020
 research and innovation programme under the Marie Sklodowska-Curie
-Global Fellowship grant agreement No 703916. This research used
-resources of the Oak Ridge Leadership Computing Facility at the Oak
-Ridge National Laboratory, which is supported by the Office of Science
-of the U.S. Department of Energy under Contract
-No.\ DE-AC05-00OR22725.  This research used resources of the National
-Energy Research Scientific Computing Center, which is supported by the
-Office of Science of the U.S. Department of Energy under Contract
+Global Fellowship grant agreement No 703916.  An award of computer
+time was provided by the Innovative and Novel Computational Impact on
+Theory and Experiment (INCITE) program. This research used resources
+of the Oak Ridge Leadership Computing Facility at the Oak Ridge
+National Laboratory, which is supported by the Office of Science of
+the U.S. Department of Energy under Contract No.\ DE-AC05-00OR22725.
+This research used resources of the National Energy Research
+Scientific Computing Center, which is supported by the Office of
+Science of the U.S. Department of Energy under Contract
 No.\ DE-AC02-05CH11231.  Visualizations were done using yt~\cite{yt}.

### 2018-09-27 astronum_2018/astronum-xrb.tex
more cites

@@
 \textit{eXTP}\/ and \textit{STROBE-X}\/
-\cite{art-2017-wilhod-etal,art-2016-zhang-etal} which will offer
+\cite{art-2017-wilhod-etal,art-2016-zhang-etal,intZand2018} which will offer
 high collecting area and time resolution.
@@
 the burning layer, where the flame is the thinnest.  Using
-well-balanced schemes can help us retain hydrostatic equilibrium (HSE)
+well-balanced schemes~\cite{kappeli:2016} can help us retain hydrostatic equilibrium (HSE)
 in the lower refined regions, reducing the number of zones needed to
@@
 makes traditional flame models a poor fit.  Methods that use even
-something as simple as subzone burning.  This is harder...
+something as simple as subzone burning (e.g.~\cite{Wang2012190}).  This is harder...
 

### 2018-09-27 astronum_2018/astronum-xrb.tex
some algorithm work

@@
 node-hours of computational time to get to about 10~ms (those
-particular simulations were run at NERSC).  The GPU version of the
-Castro hydrodynamics solver, under active development, is about
-$10\times$ faster on a comparable node, so a GPU calculation could
-bring the 2D cost down to 1k node-hours.  Most of this time is spent
-in the reactions\MarginPar{verify fraction}.  A 3D version of this
-calculation is estimated to require 5M node-hours on CPUs or 0.5M
-node-hours using GPUs---this is about the size of an
-annual \MarginPar{double check \#s} allocation.  Additionally, this is
-still with the boosted flame, but we expect that once the flame is
-developed, the boosting will not be as critical, due to hydrodynamic
-effects.  We plan to test this soon.  In any case, this is less than
+particular simulations were run at NERSC).  Most of this time is spent
+in the reactions\MarginPar{verify fraction}.  We really need to do 3D.
+There the shear from the hurricane structure acting along the flame
+front may cause instabilities that can affect the flame propagation.
+This is something we can test with our current simulation methodology.
+A 3D version of this calculation is estimated to require 5M node-hours
+on CPUs---this is about the size of an annual \MarginPar{double check
+  \#s} allocation.  Additionally, this is still with the boosted
+flame, but we expect that once the flame is developed, the boosting
+will not be as critical, due to hydrodynamic effects.  We plan to test
+this soon.  We can bootstrap the 3D calculation by mapping the 2D
+axisymmetric calculation, once the flame is firmly established, into
+3D, rotating about the symmetry axis.  This will cut down the
+computational time by about one-half.  In any case, this is less than
 1\% of the neutron star surface, so a fully resolved calculation of
@@
 gravity (for plane-parallel domains) ports in a straightforward manner
-to GPUs.  We can expect GPUs to perform 10--20$\times$ faster than
-CPUs on modern supercomputer nodes.  This performance increase will
-directly benefit XRB simulations.  GPU offloading of the reaction networks
-will also allow us to explore larger networks and more detailed nuclear
-physics.
+to GPUs.  The current GPU version of the \castro hydrodynamics solver
+is about $10\times$ faster using GPUs on a node than the entire node
+of CPU cores.  We see similar speed ups with reaction networks, with
+these GPU ports under active development.  This means a GPU
+calculation could lower the above costs by a factor of 10.  GPU
+offloading of the reaction networks will also allow us to explore
+larger networks and more detailed nuclear physics.
 
-If we continue to resolve the flame structure, then we need algroithms
+If we continue to resolve the flame structure, then we need algorithms
 that can do so less expensively.  We can be smarter about our grid
@@
 the burning layer, where the flame is the thinnest.  Using
-well-balanced schemes~\cite{kappeli:2016} can help us retain hydrostatic equilibrium (HSE)
-in the lower refined regions, reducing the number of zones needed to
-maintain fidelity.  We can also explore initializing newly created fine
-grids ahead of the burning with an understanding of hydrostatic
-equilibrium to allow us to only refine the peak burning as the flame
-moves across the star.
+well-balanced schemes~\cite{kappeli:2016} can help us retain
+hydrostatic equilibrium (HSE) in the lower refined regions, reducing
+the number of zones needed to maintain fidelity.  We can also explore
+initializing newly created fine grids ahead of the burning with an
+understanding of hydrostatic equilibrium to allow us to only refine
+the peak burning as the flame moves across the star.
 

### 2018-09-27 astronum_2018/astronum-xrb.tex
more work on the algorithm section

@@
 We have begun a set of flame spreading calculations using the
-compressible hydrodynamics code Castro~\cite{castro}, part of the
+compressible hydrodynamics code \castro~\cite{castro}, part of the
 AMReX astrophysics suite~\cite{astronum:2017}.  \castro\ uses adaptive
@@
 gravity (for plane-parallel domains) ports in a straightforward manner
-to GPUs.  The current GPU version of the \castro hydrodynamics solver
+to GPUs.  The current GPU version of the \castro\ hydrodynamics solver
 is about $10\times$ faster using GPUs on a node than the entire node
@@
 
-If we continue to resolve the flame structure, then we need algorithms
-that can do so less expensively.  We can be smarter about our grid
-refinement strategy, putting the finest grids only at the very base of
-the burning layer, where the flame is the thinnest.  Using
-well-balanced schemes~\cite{kappeli:2016} can help us retain
-hydrostatic equilibrium (HSE) in the lower refined regions, reducing
-the number of zones needed to maintain fidelity.  We can also explore
-initializing newly created fine grids ahead of the burning with an
-understanding of hydrostatic equilibrium to allow us to only refine
-the peak burning as the flame moves across the star.
-
 Another way we can meet the resolution requirements is to use a more
-accurate hydrodynamic method.  In particular, we are developing fully
+accurate hydrodynamic method.  We are developing fully
 fourth-order (in space and time) coupling of hydrodynamics and
@@
 
-As an example of the improvements in coupling between hydro and \MarginPar{Don can add some comments about startup costs?}
+As an example of the improvements in coupling between hydro
+and \MarginPar{Don can add some comments about startup costs?}
 reactions, Figure~\ref{fig:sdc} shows the mass fraction of helium
@@
 tolerances for both the Strang-split method and the SDC method.  For
-the SDC method, we are using a second-order accurate method.  We
-predict a time-centered advection term, $\Adv{\Uc}^{n+1/2}_i$ using
-standard unsplit Godunov methods, but explicitly include a reactive
-source term in the tracing of the interface states, ${\bf R}(\Uc_i)$.  We can then solve
-the reactive system, using this advective term as a
-piecewise-constant-in-time source
+the SDC method, we are using a second-order accurate method based on
+\cite{SDC-old}.  We predict a time-centered advection term,
+$\Adv{\Uc}^{n+1/2}_i$ using standard unsplit Godunov methods, but
+explicitly include a reactive source term in the tracing of the
+interface states, ${\bf R}(\Uc_i)$.  We can then solve the reactive
+system, using this advective term as a piecewise-constant-in-time
+source
 \begin{equation}
@@
 fourth-order in space and time reacting hydrodynamics methods to
-follow shortly.
+follow shortly.  The XRB simulations are our target application.
 
@@
 \includegraphics[width=0.75\linewidth]{sdc_plot}
-\caption{\label{fig:sdc} Helium mass fraction as a function of time over three timesteps.
-  The points represent individual calls to the
-  righthand side function of our reaction network (see Eq. \ref{eq:syst}) and the density of
-  points indicates how hard the integrator is working.}
+\caption{\label{fig:sdc} Helium mass fraction as a function of time
+  over three timesteps.  The points represent individual calls to the
+  righthand side function of our reaction network (see
+  Eq. \ref{eq:syst}) and the density of points indicates how hard the
+  integrator is working.}
 \end{figure}
-
-Within the realm of compressible hydrodynamics, there are a number of
-other techniques that should be explored.  smarter refinement with AMR
-+ hydrostatic refinement well-balanced method to allow ultralow
-resolution to slide along with the burning front higher-order This are
-all
 
@@
 makes traditional flame models a poor fit.  Methods that use even
-something as simple as subzone burning (e.g.~\cite{Wang2012190}).  This is harder...
+something as simple as subzone burning (e.g.~\cite{Wang2012190}) may
+help capture the energetics with lower resolution.  This can be
+implemented using the AMR machinery.  This may help a little, but will
+still be insufficient for full star models.
 
-Finally, we can try hybrid model approaches where the fully
-compressible simulation can act as a subgrid model.  \MarginPar{this ties into Ann's stuff, and also the stuff Alice was doing}
-Multiscale methods can combine different hydro models together to allow 
-for a full star simulation with resolved burning.  Examples include...
+If we continue to resolve the flame structure, then we need to be
+smarter (and more aggressive) about our grid refinement strategy.
+Currently, we put the entire fuel layer on the finest grid, but for
+the material ahead of the flame, we could keep it at lower resolution
+until the flame reaches it.  The main obstacle is that we need to keep
+the fuel ahead of the flame in hydrostatic equilibrium, and keeping it
+refined accomplishes this well.  Using well-balanced
+schemes~\cite{kappeli:2016} can help us retain hydrostatic equilibrium
+(HSE) in the lower refined regions, reducing the number of zones
+needed to maintain fidelity, but we would still need a way to regrid
+the layer ahead of the flame to higher resolution to accurately
+capture the burning dynamics.  To accomplish this, we envision
+building a refinment strategy that does interpolation of the data onto
+the newly refined grids enforcing hydrostatic equilibrium.  This can
+allow us to keep only a narrow fully refined region that ``slides''
+along the surface of the neutron star keeping pace with the advancing
+flame.  Such a scheme would greatly reduce the computational costs.
+This is a longer term development.
+
+All of the above developments still involve fully compressible
+hydrodynamics, but multiscale methods where we couple different hydrodynamics
+solvers together in a single simulation may ultimately be needed to model the full star while resolving the burning front. \MarginPar{Alice and Ian contribute here}
+
+%% Finally, we can try hybrid model approaches where the fully
+%% compressible simulation can act as a subgrid model.  
+%% Multiscale methods can combine different hydro models together to allow 
+%% for a full star simulation with resolved burning.  Examples include...
 

### 2018-09-27 astronum_2018/astronum-xrb.tex
First draft of stuff on multiscale

@@
 All of the above developments still involve fully compressible
-hydrodynamics, but multiscale methods where we couple different hydrodynamics
+hydrodynamics, but multiscale methods~\cite{weinan2011principles} where we couple different hydrodynamics
 solvers together in a single simulation may ultimately be needed to model the full star while resolving the burning front. \MarginPar{Alice and Ian contribute here}
+Such techniques rely on it being possible to decompose the dynamics of the system
+into two scales, e.g.~a short and a long lengthscale, and that the processes at each
+scale do not interact with each other. In the context of X-ray bursts, one possibility for
+multiscale modelling would be to model the large scale dynamics using a shallow water model
+(like that of~\cite{SPIT_ETAL02}) to capture the effects of the Coriolis force, and use
+a fully compressible model in the vicinity of the burning front to capture
+the turbulent burning processes. Given that the computational costs of the 2d shallow water model
+would be negligible compared to those of the compressible model, the overall cost of 
+this multiscale model would be determined by the size of the compressible region only. This
+approach would therefore allow us to capture dynamics across the full star
+without sacrificing resolution of the flame.
 

### 2018-09-27 astronum_2018/astronum-xrb.tex

@@
 hydrodynamics, but multiscale methods~\cite{weinan2011principles} where we couple different hydrodynamics
-solvers together in a single simulation may ultimately be needed to model the full star while resolving the burning front. \MarginPar{Alice and Ian contribute here}
+solvers together in a single simulation may ultimately be needed to model the full star while resolving the burning front.
+%\MarginPar{Alice and Ian contribute here}
 Such techniques rely on it being possible to decompose the dynamics of the system
 into two scales, e.g.~a short and a long lengthscale, and that the processes at each
-scale do not interact with each other. In the context of X-ray bursts, one possibility for
+scale are at most weakly coupled with each other. In the context of X-ray bursts, one possibility for
 multiscale modelling would be to model the large scale dynamics using a shallow water model
 (like that of~\cite{SPIT_ETAL02}) to capture the effects of the Coriolis force, and use
+a Low Mach model or
 a fully compressible model in the vicinity of the burning front to capture
@@
 approach would therefore allow us to capture dynamics across the full star
-without sacrificing resolution of the flame.
+without sacrificing resolution around the flame.
+An example of coupling compressible and Low Mach models within the mesh
+refinement framework used by our codes, but in a terrestrial context,
+is~\cite{Motheau2018}. An initial implemenation, \cite{Harpole2018}, in an
+astrophysical context, coupling shallow water to compressible models in both
+Newtonian and relativistic gravity, indicates that considerable computational
+efficiencies are possible, but that close attention to the coupling between the
+models is needed.
 

### 2018-09-27 astronum_2018/astronum-xrb.tex
Minor typos

@@
 zones laterally, on the finest grid.  We use the stellar
-conducitivities from \cite{Timmes00} and a 13-isotopic alpha-chain
+conductivities from \cite{Timmes00} and a 13-isotopic alpha-chain
 reaction network.  To make the timescales easier, we boost the flame
@@
 above it is very a low density ($10^{-4}~\gcc$) buffer.  Only a
-portion of the vertical extent is show.  The top panel shows the
+portion of the vertical extent is shown.  The top panel shows the
 temperature structure.  The initial perturbation reached to $2\times
@@
 panel which shows the mean molecular weight of the nuclei ($\bar{A}$).
-We see the further behind the head of the flame we are, the heavier
-the ash nuclei.  We also see where that the head of the flame
+We see that the further behind the head of the flame we are, the heavier
+the ash nuclei.  We also see where the head of the flame
 appears lifted off of the base of the burning layer---unburned
@@
 Other ways to make the simulation less expensive can include a subgrid
-model, Flame models are commonly used for Type Ia supernovae.  There
+model. Flame models are commonly used for Type Ia supernovae.  There
 however, the flame is thin compared to the scale height, so a simple
@@
 capture the burning dynamics.  To accomplish this, we envision
-building a refinment strategy that does interpolation of the data onto
+building a refinement strategy that does interpolation of the data onto
 the newly refined grids enforcing hydrostatic equilibrium.  This can
@@
 into two scales, e.g.~a short and a long lengthscale, and that the processes at each
-scale are at most weakly coupled with each other. In the context of X-ray bursts, one possibility for
+scale are at most weakly coupled with each other. In the context of XRBs, one possibility for
 multiscale modelling would be to model the large scale dynamics using a shallow water model
 (like that of~\cite{SPIT_ETAL02}) to capture the effects of the Coriolis force, and use
-a Low Mach model or
+a low Mach number or
 a fully compressible model in the vicinity of the burning front to capture
-the turbulent burning processes. Given that the computational costs of the 2d shallow water model
-would be negligible compared to those of the compressible model, the overall cost of
-this multiscale model would be determined by the size of the compressible region only. This
+the turbulent burning processes. Given that the computational costs of the 2D shallow water model
+would be negligible compared to those of the low Mach/compressible model, the overall cost of
+this multiscale model would be determined by the size of the low Mach/compressible region only. This
 approach would therefore allow us to capture dynamics across the full star
 without sacrificing resolution around the flame.
-An example of coupling compressible and Low Mach models within the mesh
-refinement framework used by our codes, but in a terrestrial context,
+An example of coupling compressible and low Mach number models implemented using the \amrex~framework, but in a terrestrial context,
 is~\cite{Motheau2018}. An initial implemenation, \cite{Harpole2018}, in an

### 2018-09-28 astronum_2018/astronum-xrb.tex
Suggested reworking of section 3.
Try to link each potential improvement to its impact on number of
zones, or timesteps, or computational cost per zone.

Add some back-of-the-envelope estimates of the benefits where
possible.

@@
 node-hours of computational time to get to about 10~ms (those
-particular simulations were run at NERSC).  Most of this time is spent
-in the reactions\MarginPar{verify fraction}.  We really need to do 3D.
-There the shear from the hurricane structure acting along the flame
-front may cause instabilities that can affect the flame propagation.
-This is something we can test with our current simulation methodology.
-A 3D version of this calculation is estimated to require 5M node-hours
+particular simulations were run at NERSC).  We expect qualitatively different
+results in 3D calculations: for example, the shear from the hurricane structure
+acting along the flame front may cause instabilities that can affect the flame
+propagation.
+With our current simulation methodology,
+a 3D version of this calculation is estimated to require 5M node-hours
 on CPUs---this is about the size of an annual \MarginPar{double check
-  \#s} allocation.  Additionally, this is still with the boosted
-flame, but we expect that once the flame is developed, the boosting
-will not be as critical, due to hydrodynamic effects.  We plan to test
-this soon.  We can bootstrap the 3D calculation by mapping the 2D
-axisymmetric calculation, once the flame is firmly established, into
-3D, rotating about the symmetry axis.  This will cut down the
-computational time by about one-half.  In any case, this is less than
+  \#s} allocation.
+%Additionally, this is still with the boosted flame, but we expect that once the flame is developed, the boosting will not be as critical, due to hydrodynamic effects.
+We plan to test this soon.
+This simulation covers less than
 1\% of the neutron star surface, so a fully resolved calculation of
-burning over the entire neutron star is out of reach, and will be for
-some time.
+burning over the entire neutron star is out of reach: a naive scaling
+calculation suggest that 15 years of Moore's Law improvements are required.
+%
+% The calculation above is
+% log_2 [ (4 pi 10^2) / (1.23^2) ] * 1.5
+% for a NS of 10km radius vs the computational domain of 1.23km^2
+%
 
-Algorithmic improvements are needed to push the scale of simulations
+Improvements are needed to push the scale of simulations
 to the point where we can model XRBs at the full star scale with
-resolved nuclear physics.  Here we discuss a few possibilities.
+resolved nuclear physics.
+As the computational cost is roughly a product of the number of zones, the
+number of zone updates, and the cost per zone update, we need to look for
+improvements in each of these. The number of zones depends on the physical
+model employed, the meshing strategy, and the accuracy required. The number of
+zone updates depends on the smallest zone size and the largest characteristic
+speed in the model. The computational cost per zone update is dominated by the
+time spent in the reactions\MarginPar{verify fraction}. We discuss a few
+possible improvements that can come both from implementing new algorithms and
+from exploiting new hardware.
+
+%For example, the number of zones can be reduced
+%by symmetry considerations. We can bootstrap the 3D calculation by mapping the
+%2D axisymmetric calculation, once the flame is firmly established, into
+%3D, rotating about the symmetry axis.  This will cut down the computational
+%time by about one-half.
 
@@
 
-Another way we can meet the resolution requirements is to use a more
+Another way we can reduce the number of zones for given accuracy is to use a more
 accurate hydrodynamic method.  We are developing fully
@@
 order, we may also be able to reduce our resolution requirements
-needed for converged flames, and since computational work scales like
-$(\Delta x)^4$ in 3D, this can save a lot of computer time.
+needed for converged flames. For smooth wave solutions we may expect to need
+around half the number of zones per dimension, and since computational work
+scales like $(\Delta x)^4$ in 3D, this could lower costs by a factor of around
+10.
+%
+% This calculation is based on the number of points per wavelength required,
+% to resolve a periodic wave, which (I believe) goes from 12 (2nd order) to
+% about 7 (4th order). However, this is very idealized, and I'm not sure that
+% it carries over to the reactions, so this may well be optimistic.
+%
 
@@
 along the surface of the neutron star keeping pace with the advancing
-flame.  Such a scheme would greatly reduce the computational costs.
-This is a longer term development.
+flame.  When using mesh refinement the overall costs are dominated by the
+number of zones at the finest resolution, so such a scheme would greatly reduce
+the computational costs. This is a longer term development.
 
@@
 without sacrificing resolution around the flame.
+As with the more aggressive gridding strategy above, the hybrid approach
+reduces the number of zones with the largest computational cost through mesh
+refinement, and also reduces the computational costs where possible by
+employing cheaper physical models in regions where this is possible.
 An example of coupling compressible and low Mach number models implemented using the \amrex~framework, but in a terrestrial context,

### 2018-09-28 astronum_2018/astronum-xrb.tex
Minor rewording of some lines, fixed typo

--- 2018-09-27 astronum_2018/astronum-xrb.tex
@@
 flame has a thickness of 10s of centimeters~\cite{Timmes00} and the
-pressure scale height of the fuel layer is meters.  To capture the
+pressure scale height of the fuel layer is on the order of meters.  To capture the
 rise timescale of the XRB, we would need to model seconds, but the flame itself
@@
 
-Our group and others have modeled the convection preceeding the
-runaway using low Mach number hydrodynamic methods, for pure helium
+Our group and others have modeled the convection preceding the
+runaway using low Mach number hydrodynamic methods for pure helium
 bursts~\cite{Lin:2006,xrb} and mixed hydrogen/helium
-bursts~\cite{xrb2} in 2d, and full 3d simulations of mixed
-bursts~\cite{xrb3}, and followed the development of the turbulence.
+bursts~\cite{xrb2} in 2D, as well as full 3D simulations of mixed
+bursts~\cite{xrb3}, and followed the development of turbulence.
 These simulations can give an understanding of the pre-burning front
@@
 is open source and available on
-github\footnote{\url{http://github.com/amrex-astro/}}, and all of the
+GitHub\footnote{\url{http://github.com/amrex-astro/}}, and all of the
 solvers, problem setup, initial models, and analysis scripts for each
@@
 across the domain to setup a hot region at the left end of our domain
-in equilibrium with the cool model.  Thermal diffusion and nuclear
-reactions quickly setup a laterally propagating flame that begins to
+in equilibrium with the cool model on the right.  Thermal diffusion and nuclear
+reactions quickly produce a laterally propagating flame that begins to
 move through our domain.  We resolve the diffusion scale, using a
@@
 expected.  This helps confine the spreading region to a more
-reasonable horizontal region.  Nevertheless, our simulation uses 12288
+reasonable horizontal domain.  Nevertheless, our simulation uses 12288
 zones laterally, on the finest grid.  We use the stellar
@@
 reaction network.  To make the timescales easier, we boost the flame
-speed by a factor of 10, by boosting both the conductivity and energy
-generation rate.  This is an approximation we hope to relax in the
-near future.
+speed by a factor of 10 by scaling both the conductivity and energy
+generation rate by the same factor.  This is an approximation we hope to
+relax in the near future.
 

### 2018-09-29 astronum_2018/astronum-xrb.tex
fix cites

--- 2018-09-27 astronum_2018/astronum-xrb.tex
@@
 
-\usepackage{cite}
+\usepackage[sort&compress,numbers]{natbib}
+
+%\usepackage{cite}
 
@@
 reaction networks and have been used to understand the nucleosynthesis
-and rp-process~\cite{schatz:rp1999,rpprocess,schatz_rp}, as well as to understand how rate uncertainties can
+and rp-process~\cite{shatz_rp}, as well as to understand how rate uncertainties can
 affect the outcomes of the bursts.  Many successive bursts can be
@@
 observations made with the new generation of X-ray telescopes such as
-\textit{eXTP}\/ and \textit{STROBE-X}\/
-\cite{art-2017-wilhod-etal,art-2016-zhang-etal,intZand2018} which will offer
+\textit{eXTP} and \textit{STROBE--X}
+\cite{art-2017-wilhod-etal,art-2016-zhang-etal} which will offer
 high collecting area and time resolution.
@@
 We have begun a set of flame spreading calculations using the
-compressible hydrodynamics code \castro~\cite{castro}, part of the
+compressible hydrodynamics code Castro~\cite{castro}, part of the
 AMReX astrophysics suite~\cite{astronum:2017}.  \castro\ uses adaptive
@@
 
-We setup a problem by creating a pair of 1D hydrostatic XRB models (hot and cool)
+We setup a problem by creating a pair of 1D hydrostatic XRB models
 representing the post- and pre-flame states and apply these laterally
@@
 zones laterally, on the finest grid.  We use the stellar
-conductivities from \cite{Timmes00} and a 13-isotopic alpha-chain
+conducitivities from \cite{Timmes00} and a 13-isotopic alpha-chain
 reaction network.  To make the timescales easier, we boost the flame
@@
 above it is very a low density ($10^{-4}~\gcc$) buffer.  Only a
-portion of the vertical extent is shown.  The top panel shows the
-temperature structure.  The initial perturbation reached to $2\times
+portion of the vertical extent is show.  The top panel shows the
+temperature.  The initial perturbation reached to $2\times
 10^4~\mathrm{cm}$, so at this point, the extent of the burnt region
@@
 are multiple burning stages beyond just helium to carbon, and this
-shows up in the composition.  This is seen more clearly in the next
+shows up in the composition.  This is seen more clearly in the lower
 panel which shows the mean molecular weight of the nuclei ($\bar{A}$).
-We see that the further behind the head of the flame we are, the heavier
-the ash nuclei.  We also see where the head of the flame
-appears lifted off of the base of the burning layer---unburned
-material underlies the spreading flame.  This appears
+We see the further behind the head of the flame we are, the heavier
+the ash nuclei.  The second panel shows the energy generation rate.
+It is strongest just above the neutron star surface, behind the flame.
+We also see that the head of the flame is ``detached'' from the base
+of the atmosphere, with the strongest burning at the head of the flame
+slightly higher up in the atmosphere.  This can also be more clearly
+seen in the last panel that shows the ash structure.  This appears
 reminiscent of the flame structure seen in the earlier simulations of
-\cite{cavecchi:2012}. \MarginPar{Yuri, would you agree?} The third panel shows the energy generation rate.
-It is strongest just above the neutron star surface, behind the flame.
-We again see that the head of the flame is detached from the base
-of the atmosphere, with the strongest burning at the head of the flame
-slightly higher up in the atmosphere.   Finally, the bottom
+\cite{cavecchi:2012} \MarginPar{Yuri, would you agree?}  The third
 panel shows the velocity through the plane of the simulation---this is
@@
 spreading.  This is the geostrophic balance discussed in
-\cite{SPIT_ETAL02}.    These calculations are ongoing and will be
+\cite{SPIT_ETAL02}.  Finally, as already mentioned, the last panel
+shows the ash composition.  These calculations are ongoing and will be
 the subject of a detailed study in the near future.
@@
 \caption{\label{fig:xrb} Helium flame spreading across the surface of
-  a neutron star.  Shown are temperature (top), the mean molecular
-  weight of the ash (second), the nuclear energy generation rate
-  (third), and the velocity out of the simulation plane (bottom).  The
-  out of plane velocity contours indicate where the Coriolis force has
-  created a hurricane-like wind structure which confines the flame,
+  a neutron star.  Shown are temperature (top), the velocity out of the
+  simulation plane (middle), and nuclear energy generation rate
+  (bottom).
+  The out of plane velocity contours indicate where the Coriolis force
+  has created a hurricane-like wind structure which confines the flame,
   leading to an inclined front. It is at such an interface that the
-  temperature and burning rate are highest.  }
+  temperature and burning rate are highest.
+  }
 \end{figure}
@@
 node-hours of computational time to get to about 10~ms (those
-particular simulations were run at NERSC).  Most of this time is spent
-in the reactions\MarginPar{verify fraction}.  We really need to do 3D.
-There the shear from the hurricane structure acting along the flame
-front may cause instabilities that can affect the flame propagation.
-This is something we can test with our current simulation methodology.
-A 3D version of this calculation is estimated to require 5M node-hours
-on CPUs---this is about the size of an annual \MarginPar{double check
-  \#s} allocation.  Additionally, this is still with the boosted
-flame, but we expect that once the flame is developed, the boosting
-will not be as critical, due to hydrodynamic effects.  We plan to test
-this soon.  We can bootstrap the 3D calculation by mapping the 2D
-axisymmetric calculation, once the flame is firmly established, into
-3D, rotating about the symmetry axis.  This will cut down the
-computational time by about one-half.  In any case, this is less than
+particular simulations were run at NERSC).  The GPU version of the
+Castro hydrodynamics solver, under active development, is about
+$10\times$ faster on a comparable node, so a GPU calculation could
+bring the 2D cost down to 1k node-hours.  Most of this time is spent
+in the reactions\MarginPar{verify fraction}.  A 3D version of this
+calculation is estimated to require 5M node-hours on CPUs or 0.5M
+node-hours using GPUs---this is about the size of an
+annual \MarginPar{double check \#s} allocation.  Additionally, this is
+still with the boosted flame, but we expect that once the flame is
+developed, the boosting will not be as critical, due to hydrodynamic
+effects.  We plan to test this soon.  In any case, this is less than
 1\% of the neutron star surface, so a fully resolved calculation of
@@
 gravity (for plane-parallel domains) ports in a straightforward manner
-to GPUs.  The current GPU version of the \castro\ hydrodynamics solver
-is about $10\times$ faster using GPUs on a node than the entire node
-of CPU cores.  We see similar speed ups with reaction networks, with
-these GPU ports under active development.  This means a GPU
-calculation could lower the above costs by a factor of 10.  GPU
-offloading of the reaction networks will also allow us to explore
-larger networks and more detailed nuclear physics.
+to GPUs.  We can expect GPUs to perform 10--20$\times$ faster than
+CPUs on modern supercomputer nodes.  This performance increase will
+directly benefit XRB simulations.  GPU offloading of the reaction networks
+will also allow us to explore larger networks and more detailed nuclear
+physics.
+
+If we continue to resolve the flame structure, then we need algroithms
+that can do so less expensively.  We can be smarter about our grid
+refinement strategy, putting the finest grids only at the very base of
+the burning layer, where the flame is the thinnest.  Using
+well-balanced schemes can help us retain hydrostatic equilibrium (HSE)
+in the lower refined regions, reducing the number of zones needed to
+maintain fidelity.  We can also explore initializing newly created fine
+grids ahead of the burning with an understanding of hydrostatic
+equilibrium to allow us to only refine the peak burning as the flame
+moves across the star.
 
 Another way we can meet the resolution requirements is to use a more
-accurate hydrodynamic method.  We are developing fully
+accurate hydrodynamic method.  In particular, we are developing fully
 fourth-order (in space and time) coupling of hydrodynamics and
@@
 
-As an example of the improvements in coupling between hydro
-and \MarginPar{Don can add some comments about startup costs?}
+As an example of the improvements in coupling between hydro and \MarginPar{Don can add some comments about startup costs?}
 reactions, Figure~\ref{fig:sdc} shows the mass fraction of helium
@@
 tolerances for both the Strang-split method and the SDC method.  For
-the SDC method, we are using a second-order accurate method based on
-\cite{SDC-old}.  We predict a time-centered advection term,
-$\Adv{\Uc}^{n+1/2}_i$ using standard unsplit Godunov methods, but
-explicitly include a reactive source term in the tracing of the
-interface states, ${\bf R}(\Uc_i)$.  We can then solve the reactive
-system, using this advective term as a piecewise-constant-in-time
-source
+the SDC method, we are using a second-order accurate method.  We
+predict a time-centered advection term, $\Adv{\Uc}^{n+1/2}_i$ using
+standard unsplit Godunov methods, but explicitly include a reactive
+source term in the tracing of the interface states, ${\bf R}(\Uc_i)$.  We can then solve
+the reactive system, using this advective term as a
+piecewise-constant-in-time source
 \begin{equation}
@@
 fourth-order in space and time reacting hydrodynamics methods to
-follow shortly.  The XRB simulations are our target application.
+follow shortly.
 
@@
 \includegraphics[width=0.75\linewidth]{sdc_plot}
-\caption{\label{fig:sdc} Helium mass fraction as a function of time
-  over three timesteps.  The points represent individual calls to the
-  righthand side function of our reaction network (see
-  Eq. \ref{eq:syst}) and the density of points indicates how hard the
-  integrator is working.}
+\caption{\label{fig:sdc} Helium mass fraction as a function of time over three timesteps.
+  The points represent individual calls to the
+  righthand side function of our reaction network (see Eq. \ref{eq:syst}) and the density of
+  points indicates how hard the integrator is working.}
 \end{figure}
 
+Within the realm of compressible hydrodynamics, there are a number of
+other techniques that should be explored.  smarter refinement with AMR
++ hydrostatic refinement well-balanced method to allow ultralow
+resolution to slide along with the burning front higher-order This are
+all
+
 Other ways to make the simulation less expensive can include a subgrid
-model. Flame models are commonly used for Type Ia supernovae.  There
+model, Flame models are commonly used for Type Ia supernovae.  There
 however, the flame is thin compared to the scale height, so a simple
@@
 makes traditional flame models a poor fit.  Methods that use even
-something as simple as subzone burning (e.g.~\cite{Wang2012190}) may
-help capture the energetics with lower resolution.  This can be
-implemented using the AMR machinery.  This may help a little, but will
-still be insufficient for full star models.
+something as simple as subzone burning.  This is harder...
 
-If we continue to resolve the flame structure, then we need to be
-smarter (and more aggressive) about our grid refinement strategy.
-Currently, we put the entire fuel layer on the finest grid, but for
-the material ahead of the flame, we could keep it at lower resolution
-until the flame reaches it.  The main obstacle is that we need to keep
-the fuel ahead of the flame in hydrostatic equilibrium, and keeping it
-refined accomplishes this well.  Using well-balanced
-schemes~\cite{kappeli:2016} can help us retain hydrostatic equilibrium
-(HSE) in the lower refined regions, reducing the number of zones
-needed to maintain fidelity, but we would still need a way to regrid
-the layer ahead of the flame to higher resolution to accurately
-capture the burning dynamics.  To accomplish this, we envision
-building a refinement strategy that does interpolation of the data onto
-the newly refined grids enforcing hydrostatic equilibrium.  This can
-allow us to keep only a narrow fully refined region that ``slides''
-along the surface of the neutron star keeping pace with the advancing
-flame.  Such a scheme would greatly reduce the computational costs.
-This is a longer term development.
-
-All of the above developments still involve fully compressible
-hydrodynamics, but multiscale methods~\cite{weinan2011principles} where we couple different hydrodynamics
-solvers together in a single simulation may ultimately be needed to model the full star while resolving the burning front.
-%\MarginPar{Alice and Ian contribute here}
-Such techniques rely on it being possible to decompose the dynamics of the system
-into two scales, e.g.~a short and a long lengthscale, and that the processes at each
-scale are at most weakly coupled with each other. In the context of XRBs, one possibility for
-multiscale modelling would be to model the large scale dynamics using a shallow water model
-(like that of~\cite{SPIT_ETAL02}) to capture the effects of the Coriolis force, and use
-a low Mach number or
-a fully compressible model in the vicinity of the burning front to capture
-the turbulent burning processes. Given that the computational costs of the 2D shallow water model
-would be negligible compared to those of the low Mach/compressible model, the overall cost of
-this multiscale model would be determined by the size of the low Mach/compressible region only. This
-approach would therefore allow us to capture dynamics across the full star
-without sacrificing resolution around the flame.
-An example of coupling compressible and low Mach number models implemented using the \amrex~framework, but in a terrestrial context,
-is~\cite{Motheau2018}. An initial implemenation, \cite{Harpole2018}, in an
-astrophysical context, coupling shallow water to compressible models in both
-Newtonian and relativistic gravity, indicates that considerable computational
-efficiencies are possible, but that close attention to the coupling between the
-models is needed.
-
-%% Finally, we can try hybrid model approaches where the fully
-%% compressible simulation can act as a subgrid model.
-%% Multiscale methods can combine different hydro models together to allow
-%% for a full star simulation with resolved burning.  Examples include...
+Finally, we can try hybrid model approaches where the fully
+compressible simulation can act as a subgrid model.  \MarginPar{this ties into Ann's stuff, and also the stuff Alice was doing}
+Multiscale methods can combine different hydro models together to allow 
+for a full star simulation with resolved burning.  Examples include...
 
@@
 Office of Advanced Scientific Computing Research under Contract No,
-DE-AC02-05CH11231. YC is supported by the European Union Horizon 2020
+DE-AC02-05CH11231. An award of computer time was provided by the
+Innovative and Novel Computational Impact on Theory and Experiment
+(INCITE) program. YC is supported by the European Union Horizon 2020
 research and innovation programme under the Marie Sklodowska-Curie
-Global Fellowship grant agreement No 703916.  An award of computer
-time was provided by the Innovative and Novel Computational Impact on
-Theory and Experiment (INCITE) program. This research used resources
-of the Oak Ridge Leadership Computing Facility at the Oak Ridge
-National Laboratory, which is supported by the Office of Science of
-the U.S. Department of Energy under Contract No.\ DE-AC05-00OR22725.
-This research used resources of the National Energy Research
-Scientific Computing Center, which is supported by the Office of
-Science of the U.S. Department of Energy under Contract
+Global Fellowship grant agreement No 703916. This research used
+resources of the Oak Ridge Leadership Computing Facility at the Oak
+Ridge National Laboratory, which is supported by the Office of Science
+of the U.S. Department of Energy under Contract
+No.\ DE-AC05-00OR22725.  This research used resources of the National
+Energy Research Scientific Computing Center, which is supported by the
+Office of Science of the U.S. Department of Energy under Contract
 No.\ DE-AC02-05CH11231.  Visualizations were done using yt~\cite{yt}.

### 2018-09-29 astronum_2018/astronum-xrb.tex
typo
Merge Yuri's edits

--- 2018-09-28 astronum_2018/astronum-xrb.tex
@@
 \isot{Ni}{56}, representative of the underlying neutron star, and
-above it is very a low density ($10^{-4}~\gcc$) buffer.  Only a
+above it is a very low density ($10^{-4}~\gcc$) buffer.  Only a
 portion of the vertical extent is shown.  The top panel shows the
@@
 We see that the further behind the head of the flame we are, the heavier
-the ash nuclei.  We also see where the head of the flame
+the ash nuclei.  We also see the head of the flame
 appears lifted off of the base of the burning layer---unburned
-material underlies the spreading flame.  This appears
+material underlies the spreading flame.  This is
 reminiscent of the flame structure seen in the earlier simulations of
-\cite{cavecchi:2012}. \MarginPar{Yuri, would you agree?} The third panel shows the energy generation rate.
+\cite{cavecchi:2012}.  The third panel shows the energy generation rate.
 It is strongest just above the neutron star surface, behind the flame.
@@
 panel shows the velocity through the plane of the simulation---this is
-induced by the Coriolis force as the burning front spreads laterally.
+induced by the Coriolis force as the burning fluid expands and begins to spreads laterally.
 A tight hurricane structure has setup as a result of the flame

### 2018-09-29 astronum_2018/astronum-xrb.tex
some minor edits

@@
 The 2D calculations shown in the previous section require about 10k
-node-hours of computational time to get to about 10~ms (those
-particular simulations were run at NERSC).  We expect qualitatively different
+node-hours of computational time to get to about 10~ms (running
+on the Edison machine at NERSC).  We expect qualitatively different
 results in 3D calculations: for example, the shear from the hurricane structure
@@
 With our current simulation methodology,
-a 3D version of this calculation is estimated to require 5M node-hours
-on CPUs---this is about the size of an annual \MarginPar{double check
-  \#s} allocation.
+a 3D version of this calculation is estimated to require 10M node-hours
+or more---this is about the size of an annualallocation.
 %Additionally, this is still with the boosted flame, but we expect that once the flame is developed, the boosting will not be as critical, due to hydrodynamic effects.
-We plan to test this soon.
 This simulation covers less than
@@
 speed in the model. The computational cost per zone update is dominated by the
-time spent in the reactions\MarginPar{verify fraction}. We discuss a few
+time spent in the reactions and equation of state. We discuss a few
 possible improvements that can come both from implementing new algorithms and

### 2018-09-30 astronum_2018/plots/aprox19/sdc_plot.py
another plot

@@ -1,0 +1,66 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+
+#mpl.rcParams['xtick.labelsize'] = "small"
+
+# read the Strang data into a numpy array
+strang = np.loadtxt("strang_diag.out")
+
+# now manually read it looking for the times that indicate when the Strang halves began
+strang_starts = []
+strang_second = []
+with open("strang_diag.out") as f:
+    for line in f:
+        if line.startswith(" # start of step"):
+            _, time = line.split("=")
+            strang_starts.append(float(time))
+        elif line.startswith(" # second half of strang"):
+            _, time = line.split("=")
+            strang_second.append(float(time))
+
+strang_starts = sorted(set(strang_starts))
+strang_second = sorted(set(strang_second))
+
+print(strang_starts)
+
+# now get the SDC data
+sdc = np.loadtxt("sdc_diag.out")
+
+# we have 2 SDC iterations here, so be able to separate them
+idx_iter0 = sdc[:,1] == 0
+idx_iter1 = sdc[:,1] == 1
+
+
+# make the plots
+plt.scatter(strang[:,0], strang[:,2], marker="x", s=15, label="Strang splitting")
+plt.scatter(sdc[idx_iter1,0], sdc[idx_iter1,3], marker="o", s=15, label="SDC")
+
+# draw timestep boundaries
+for time in strang_starts:
+    plt.plot([time, time], [0,1], color="k", linewidth=2)
+
+for time in strang_second:
+    plt.plot([time, time], [0,1], color="0.5", linestyle=":", linewidth=2)
+
+plt.xlim(1.7960120964633856E-005, 1.8157726447146557E-005)
+plt.ylim(0.57, 0.595)
+
+plt.legend(frameon=False)
+
+plt.tight_layout()
+
+f = plt.gcf()
+f.set_size_inches(9.0, 7.0)
+
+plt.xlabel("time (s)", fontsize="large")
+plt.ylabel(r"$X({}^4\mathrm{He})$", fontsize="large")
+
+plt.ticklabel_format(style='sci', axis='x', scilimits=(0,0))
+
+plt.tight_layout()
+
+plt.savefig("sdc_plot.png")
+
+print("sdc points = ", len(sdc[idx_iter1,0]))
+print("strang points = ", len(strang[:,0]))

### 2018-09-30 astronum_2018/astronum-xrb.tex
tweak size
a little more on the SDC

@@
 \centering
-\includegraphics[width=0.95\linewidth]{xrb}
+\includegraphics[width=\linewidth]{xrb}
 \caption{\label{fig:xrb} Helium flame spreading across the surface of
@@
 reactions, Figure~\ref{fig:sdc} shows the mass fraction of helium
-behind a detonation over the course of 3 timesteps.  The points
+behind a detonation over the course of 2 timesteps.  For this test,
+a 19 isotope reaction network was used.  The points
 represent individual calls to the righthand side function of our
-reaction network (a 13-isotope $\alpha$-chain), and the density of
+reaction network, and the density of
 points indicates how hard the integrator is working.  We use the same
@@
 in Figure~\ref{fig:sdc}, each SDC iteration called the righthand side
-function of the network 5 times less than the Strang case.
-We are continuing to develop this new integration strategy, with full
-fourth-order in space and time reacting hydrodynamics methods to
-follow shortly.  The XRB simulations are our target application.
+function of the network 2 times less than the Strang case.  This
+savings is dependent on the network and thermodynamic state, but our
+tests thus far show that the SDC methods need fewer network calls to
+integrate a timestep.  We also clearly see how the Strang system
+evolves in a discontinuous fashion.  We are continuing to develop this
+new integration strategy, with full fourth-order in space and time
+reacting hydrodynamics methods to follow shortly.  The XRB simulations
+are our target application.
 
@@
 \caption{\label{fig:sdc} Helium mass fraction as a function of time
-  over three timesteps.  The points represent individual calls to the
+  over two timesteps for a detonation test problem with a 19-isotope
+  reaction network.  The points represent individual calls to the
   righthand side function of our reaction network (see
   Eq. \ref{eq:syst}) and the density of points indicates how hard the
-  integrator is working.}
+  integrator is working.  For the Strang case, we see that the first
+  half of the reactions brings us off of the smooth solution
+  (represented by the SDC curve), then the advection overcorrects,
+  resulting in a discontinuous jump, and finally the second half of
+  the reactions works to bring the state back to the solution curve.}
 \end{figure}

### 2018-09-30 astronum_2018/astronum-xrb.tex
Remove explicit References section.
As the bibliography command puts one in on my system, so it duplicates.

Also a couple of tiny typos.

@@
 a 3D version of this calculation is estimated to require 10M node-hours
-or more---this is about the size of an annualallocation.
+or more---this is about the size of an annual allocation.
 %Additionally, this is still with the boosted flame, but we expect that once the flame is developed, the boosting will not be as critical, due to hydrodynamic effects.
@@
 burning over the entire neutron star is out of reach: a naive scaling
-calculation suggest that 15 years of Moore's Law improvements are required.
+calculation suggests that 15 years of Moore's Law improvements are required.
 %
@@
 
-\section*{References}
+%\section*{References}
 

### 2018-09-30 astronum_2018/astronum-xrb.tex
another pass

@@
 reactions using the method of spectral deferred corrections (SDC) in
-\castro, instead of the more commonly employed Strang splitting.  This
+\castro, instead of the more commonly employed Strang splitting~\cite{strang:1968}.  This
 provides us with two benefits.  First, the improved coupling actually
@@
 calls and therefore reducing the expense of the
-reactions. \MarginPar{need some refs here} Second, by moving to fourth
+reactions.  Second, by moving to fourth
 order, we may also be able to reduce our resolution requirements
@@
   righthand side function of our reaction network (see
-  Eq. \ref{eq:syst}) and the density of points indicates how hard the
+  Eq.~\ref{eq:syst}) and the density of points indicates how hard the
   integrator is working.  For the Strang case, we see that the first
@@
 scale are at most weakly coupled with each other. In the context of XRBs, one possibility for
-multiscale modelling would be to model the large scale dynamics using a shallow water model
+multiscale modeling would be to model the large scale dynamics using a shallow water model
 (like that of~\cite{SPIT_ETAL02}) to capture the effects of the Coriolis force, and use
@@
 a fully compressible model in the vicinity of the burning front to capture
-the turbulent burning processes. Given that the computational costs of the 2D shallow water model
+the turbulent burning processes. Given that the computational costs of the shallow water model
 would be negligible compared to those of the low Mach/compressible model, the overall cost of
@@
 An example of coupling compressible and low Mach number models implemented using the \amrex~framework, but in a terrestrial context,
-is~\cite{Motheau2018}. An initial implemenation, \cite{Harpole2018}, in an
+is~\cite{Motheau2018}. An initial implementation, \cite{Harpole2018}, in an
 astrophysical context, coupling shallow water to compressible models in both
@@
 \section{Summary}
+
+XRBs are multiscale, multiphysics problems that are challenging to
+model.  Nevertheless, significant progress has been made in
+multidimensional models of these events, through a set of
+complementary algorithmic approaches.  Future advances in computer
+architectures and algorithms will allow for even more realistic models
+of these events, and help us connect to observations, ultimately telling
+us about the neutron star itself.
 

### 2018-10-01 astronum_2018/astronum-xrb.tex
typo fix

@@
 panel shows the velocity through the plane of the simulation---this is
-induced by the Coriolis force as the burning fluid expands and begins to spreads laterally.
+induced by the Coriolis force as the burning fluid expands and begins to spread laterally.
 A tight hurricane structure has setup as a result of the flame

### 2018-10-01 astronum_2018/astronum-xrb.tex
another typo

@@
   (third), and the velocity out of the simulation plane (bottom).  The
-  out of plane velocity contours indicate where the Coriolis force has
+  out-of-plane velocity contours indicate where the Coriolis force has
   created a hurricane-like wind structure which confines the flame,

### 2018-10-02 astronum_2018/astronum-xrb.tex
Add paragraph about VODE timestepping and the restart experiment.

@@
 \newcommand{\amrex}{{\sffamily AMReX}}
+\newcommand{\vode}{{\sffamily VODE}}
 
@@
 As an example of the improvements in coupling between hydro
-and \MarginPar{Don can add some comments about startup costs?}
-reactions, Figure~\ref{fig:sdc} shows the mass fraction of helium
+and reactions, Figure~\ref{fig:sdc} shows the mass fraction of helium
 behind a detonation over the course of 2 timesteps.  For this test,
@@
 are our target application.
+
+For the ODE integrations shown we use the variable-order,
+variable-step integrator \vode\ \cite{vode} with its implementation of
+the backward-differentiation formulas. The integration order
+corresponds to the order of the interpolating polynomials used to
+predict the solution in each integrator step. We note that for both
+Strang splitting and SDC, \vode\ must start the integration at first
+order and correspondingly small time steps since no higher-order
+information about the solution is available. That is, our reaction
+network right-hand sides provide the first time derivative of the
+solution but no higher derivatives. Based on testing with a 13-isotope
+approximate alpha-chain network, \vode\ takes many small timesteps
+(e.g. $\approx 200$) before it can stably increase to a high order
+representation of the solution, but after it does so, it can often
+complete the integration in only a handful of steps (e.g.
+$\approx 5-6$). Observing that during a \castro\ timestep, the first
+Strang step is a direct continuation of the integration in the second
+Strang step of the previous \castro\ timestep, we explored saving the
+\vode\ state following the second Strang step and restarting \vode\
+during the first Strang step. This approach allowed us to start the
+first Strang step with a high order representation of the solution and
+complete it in very few \vode\ timesteps (e.g. $<10$ steps), speeding
+up the \castro\ timestep advance by 50\% for our detonation test
+problem using the 13-isotope network. However, this came at the memory
+cost of storing $7 N_{eq} + 90$ scalar variables on our grid, where
+for our networks $N_{eq} = N_{spec} + 2$ (in addition to the number of
+species $N_{spec}$ we also integrate temperature and specific
+energy). For a large-scale science problem with a large network, this
+could quickly become prohibitive, so we are exploring ways to mitigate
+either the memory requirements for restarting the integration or the
+small timestep start up costs for the integration method.
 

### 2018-10-02 astronum_2018/astronum-xrb.tex
update refs`

--- 2018-10-01 astronum_2018/astronum-xrb.tex
@@
         I.~Hawke$^3$,
-        M.~P. Katz$^7$,
-        C.~M. Malone$^8$,
+        M.~P. Katz$^5$,
+        C.~M. Malone$^6$,
         A.~J. Nonaka$^4$,
@@
 
-\address{$^5$National Energy Research Scientific Computing Center,
-  Lawrence Berkeley National Lab, Berkeley, CA 94720 USA}
-
-\address{$^6$Department of Physics and Astronomy, Michigan State
-  University, East Lansing, Michigan 48824 USA}
-
-\address{$^7$NVIDIA Corporation, 2788 San Tomas Expressway,
+\address{$^5$NVIDIA Corporation, 2788 San Tomas Expressway,
   Santa Clara, CA, 95050 USA}
 
-\address{$^8$Los Alamos National Laboratory, Los Alamos, NM, 87545 USA}
+\address{$^6$Los Alamos National Laboratory, Los Alamos, NM, 87545 USA}
 
@@
 panel shows the velocity through the plane of the simulation---this is
-induced by the Coriolis force as the burning fluid expands and begins to spread laterally.
+induced by the Coriolis force as the burning fluid expands and begins to spreads laterally.
 A tight hurricane structure has setup as a result of the flame

### 2018-10-03 astronum_2018/astronum-xrb.tex
a little smoothing of the ODE paragraph

@@
 \usepackage{amsmath}
-
+\usepackage{amssymb}
 \usepackage{bm}
@@
 variable-step integrator \vode\ \cite{vode} with its implementation of
-the backward-differentiation formulas. The integration order
-corresponds to the order of the interpolating polynomials used to
-predict the solution in each integrator step. We note that for both
+the backward-differentiation formulas.
+% The integration order
+%corresponds to the order of the interpolating polynomials used to
+%predict the solution in each integrator step. 
+We note that for both
 Strang splitting and SDC, \vode\ must start the integration at first
-order and correspondingly small time steps since no higher-order
-information about the solution is available. That is, our reaction
-network right-hand sides provide the first time derivative of the
-solution but no higher derivatives. Based on testing with a 13-isotope
+order and correspondingly small timesteps since no higher-order
+information about the solution is available (our reaction
+network right-hand sides provide only the first time derivative of the
+solution). Based on testing with a 13-isotope
 approximate alpha-chain network, \vode\ takes many small timesteps
-(e.g. $\approx 200$) before it can stably increase to a high order
+(e.g.\ $\sim 200$) before it can stably increase to a high order
 representation of the solution, but after it does so, it can often
-complete the integration in only a handful of steps (e.g.
-$\approx 5-6$). Observing that during a \castro\ timestep, the first
-Strang step is a direct continuation of the integration in the second
-Strang step of the previous \castro\ timestep, we explored saving the
-\vode\ state following the second Strang step and restarting \vode\
-during the first Strang step. This approach allowed us to start the
-first Strang step with a high order representation of the solution and
-complete it in very few \vode\ timesteps (e.g. $<10$ steps), speeding
-up the \castro\ timestep advance by 50\% for our detonation test
-problem using the 13-isotope network. However, this came at the memory
-cost of storing $7 N_{eq} + 90$ scalar variables on our grid, where
-for our networks $N_{eq} = N_{spec} + 2$ (in addition to the number of
-species $N_{spec}$ we also integrate temperature and specific
-energy). For a large-scale science problem with a large network, this
-could quickly become prohibitive, so we are exploring ways to mitigate
-either the memory requirements for restarting the integration or the
-small timestep start up costs for the integration method.
+complete the integration in only a handful of steps (e.g.\ $\sim
+5$--$6$). Observing that during a \castro\ timestep, the first Strang
+step is a direct continuation of the integration in the second Strang
+step of the previous \castro\ timestep, we explored saving the
+\vode\ state following the second Strang step and restarting
+\vode\ during the first Strang step. This approach allowed us to start
+the first Strang step with a high order representation of the solution
+and complete it in very few \vode\ timesteps ($\lesssim 10$ steps),
+speeding up the \castro\ timestep advance by 50\% for our detonation
+test problem using the 13-isotope network. However, this came at the
+memory cost of storing $7 N_{eq} + 90$ scalar variables on our grid,
+where for our networks $N_{eq} = N_{spec} + 2$ (in addition to the
+number of species $N_{spec}$ we also integrate temperature and
+specific energy). For a large-scale science problem with a large
+network, this could quickly become prohibitive, so we are exploring
+ways to mitigate either the memory requirements for restarting the
+integration or the small timestep start up costs for the integration
+method.
 

### 2018-11-04 gpus/paper.tex
start of the GPU paper

%\documentclass[12pt, preprint]{aastex}
\documentclass[preprint2,times]{aastex6}
%\documentclass[preprint,times]{aastex6}

% these lines seem necessary for pdflatex to get the paper size right
\pdfpagewidth 8.5in
\pdfpageheight 11.0in

\usepackage{epsf,color,amsmath}

\usepackage{cancel}

\newcommand{\sfrac}[2]{\mathchoice%
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
    \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
    \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptscriptfont0 #1}\kern-.2em/
    \kern-.15em\lower.25ex\hbox{\the\scriptscriptfont0 #2}} {#1\!/#2}}

\newcommand{\castro}{{\sf Castro}}

\newcommand{\nablab}{{\mathbf{\nabla}}}
\newcommand{\Ub}{\mathbf{U}}
\newcommand{\gb}{\mathbf{g}}
\newcommand{\omegadot}{\dot{\omega}}
\newcommand{\Sdot}{\dot{S}}
\newcommand{\ddx}[1]{{\frac{{\partial#1}}{\partial x}}}
\newcommand{\ddt}[1]{{\frac{{\partial#1}}{\partial t}}}
\newcommand{\odt}[1]{{\frac{{d#1}}{dt}}}
\newcommand{\divg}[1]{{\nablab \cdot \left (#1\right)}}

\newcommand{\Ic}{\mathcal{I}}
\newcommand{\smax}{{s_\mathrm{max}}}

\usepackage{bm}

\newcommand{\Uc}{{\bm{\mathcal{U}}}}
\newcommand{\Fb}{\mathbf{F}}
\newcommand{\Sc}{\mathbf{S}}

\newcommand{\xv}{{(x)}}
\newcommand{\yv}{{(y)}}
\newcommand{\zv}{{(z)}}

\newcommand{\ex}{{\bf e}_x}
\newcommand{\ey}{{\bf e}_y}
\newcommand{\ez}{{\bf e}_z}

\newcommand{\Ab}{{\bf A}}
\newcommand{\Sq}{{\bf S}_\qb}
\newcommand{\Sqhydro}{{\Sq^{\mathrm{hydro}}}}
\newcommand{\qb}{{\bf q}}

\newcommand{\Shydro}{{{\bf S}^{\mathrm{hydro}}}}
\newcommand{\Rb}{{\bf R}}
\newcommand{\Rq}{{\bf R}}
\newcommand{\Adv}[1]{{\left [\mathcal{A} \left(#1\right)\right]}}
\newcommand{\Advt}[1]{{\left [\mathcal{\tilde{A}} \left(#1\right)\right]}}
\newcommand{\Advs}[1]{{\mathcal{A} \left(#1\right)}}

\setlength{\marginparwidth}{0.75in}
\newcommand{\MarginPar}[1]{\marginpar{\vskip-\baselineskip\raggedright\tiny\sffamily\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}

\begin{document}
%======================================================================
% Title
%======================================================================
\title{CASTRO: A New Compressible Astrophysical Solver. IV. Performance Portability}

\shorttitle{}
\shortauthors{}

\author{Castro developers / M.~P.~Katz et al.}
\affil{}
\email{}

%======================================================================
% Abstract and Keywords
%======================================================================
\begin{abstract}
blah
\end{abstract}

\keywords{hydrodynamics---methods: numerical}

%======================================================================
% Introduction
%======================================================================
\section{Introduction}\label{Sec:Introduction}

\citep{castro}

Existing GPU-enabled astrophysical hydro codes include
\citet{gamer,cholla,fargo3d,pekkila:2017}

GPU-enabled integration has been explored by \citep{brock:2015}.

\section{Performance portable compute kernels}

Describe mfiter structure

Describe MPI, MPI+OpenMP w/ tiling, MPI + CUDA

\begin{figure}
\centering
\includegraphics[height=0.25\textheight]{gpu_1} \\
\includegraphics[height=0.25\textheight]{gpu_2} \\
\includegraphics[height=0.25\textheight]{gpu_3}
\caption{\label{fig:loops} caption}
\end{figure}

\section{Scaling}

\acknowledgements \castro\ is freely available at
\url{http://github.com/AMReX-Astro/Castro}.  The work at Stony Brook
was supported by DOE/Office of Nuclear Physics grant DE-FG02-87ER40317
and NSF award AST-1211563.  An award of computer time was provided by
the Innovative and Novel Computational Impact on Theory and Experiment
(INCITE) program.  This research used resources of the Oak Ridge
Leadership Computing Facility at the Oak Ridge National Laboratory,
which is supported by the Office of Science of the U.S. Department of
Energy under Contract No. DE-AC05-00OR22725.  We thank NVIDIA Corporation
for the donation of a Titan X Pascal used in this research.

%======================================================================
% References
%======================================================================

\bibliographystyle{apj}
\bibliography{ws}

\end{document}

### 2018-11-04 gpus/domain.py

@@ -1,0 +1,161 @@
+import matplotlib.pyplot as plt
+import random
+import matplotlib as mpl
+from matplotlib import cm
+
+class Box(object):
+
+    def __init__(self, ll, uu, nx, ny, grid_color="0.5", ng=0):
+
+        self.ll = ll
+        self.uu = uu
+        self.nx = nx
+        self.ny = ny
+        self.grid_color = grid_color
+        self.ng = ng
+
+        self.dx = (self.uu[0] - self.ll[0])/self.nx
+        self.dy = (self.uu[1] - self.ll[1])/self.ny
+
+    def draw(self):
+
+        # draw the frame
+        plt.plot([self.ll[0], self.ll[0], self.uu[0], self.uu[0], self.ll[0]],
+                 [self.ll[1], self.uu[1], self.uu[1], self.ll[1], self.ll[1]],
+                 color=self.grid_color, lw=2)
+
+        # draw the x grid lines
+        for n in range(1, self.nx):
+            plt.plot([self.ll[0] + n*self.dx, self.ll[0] + n*self.dx],
+                     [self.ll[1], self.uu[1]],
+                     color=self.grid_color, ls=":", lw=1)
+
+        # draw the y grid lines
+        for n in range(1, self.ny):
+            plt.plot([self.ll[0], self.uu[0]],
+                     [self.ll[1] + n*self.dy, self.ll[1] + n*self.dy],
+                     color=self.grid_color, ls=":", lw=1)
+
+        # ghostcells?
+        if self.ng > 0:
+            xmin = self.ll[0] - self.ng*self.dx
+            xmax = self.uu[0] + self.ng*self.dx
+            ymin = self.ll[1] - self.ng*self.dy
+            ymax = self.uu[1] + self.ng*self.dy
+
+            plt.plot([xmin, xmin, xmax, xmax, xmin],
+                     [ymin, ymax, ymax, ymin, ymin],
+                     ls="--", color="r")
+
+
+    def tile(self, num_x):
+
+        tile_nx = int(self.nx / num_x)
+        if self.nx % num_x != 0:
+            tile_nx += 1
+
+        for n in range(num_x):
[…]

### 2018-11-04 gpus/paper.tex
more outline
software

@@
 
+\subsection{Summary of changes required}
+
+The main hydrodynamics solver in \castro\ is an unsplit corner
+transport upwind \citep{ppmunsplit} piecewise parabolic method
+\citep{ppm} scheme.  This uses full corner coupling, requiring a 
+number of transverse Riemann problem solutions.
+
+\section{Reaction networks}
+
 \section{Scaling}
+
+Our current focus is on offloading the hydrodynamics and reactions to
+GPUs, so we will achieve the best GPU performance for problems without
+self-gravity.
+
+\software{MPICH, GCC, Castro, python, matplotlib}
 

### 2018-11-04 gpus/paper.tex
some abstract

@@
 \newcommand{\castro}{{\sf Castro}}
+\newcommand{\maestro}{{\sf Maestro}}
 
@@
 \begin{abstract}
-blah
+We describe a performance portable version of the \castro\ hydrodynamics
+code that is able to achieve excellent scaling on both CPU and GPU-based
+supercomputers.
 \end{abstract}
@@
 
+\subsection{Pure hydrodynamics}
+
+\subsection{Reaction test}
+
+\subsection{X-ray burst}
+
+This problem uses constant gravity, so all of the physics can be done on the GPUs.
+
+\section{Summary}
+
+Future work: gravity on GPUs (multigrid or multipole), port of \maestro\ following
+the same ideas.
+
 \software{MPICH, GCC, Castro, python, matplotlib}

### 2018-11-05 gpus/paper.tex
more notes
more work

@@
 %\documentclass[12pt, preprint]{aastex}
-\documentclass[preprint2,times]{aastex6}
+\documentclass[twocolumn,times]{aastex62}
 %\documentclass[preprint,times]{aastex6}
@@
 
-\citep{castro}
+is a compressible (radiation-)hydrodynamics simulation code built on
+the AMReX adaptive mesh refinement framework \citep{castro,castro2,castro3}.  It 
+includes full self-gravity using multigrid methods with isolated boundaries
+\cite{katz:2016}, works with general equations of state \citep{zingalekatz}
+and arbitrary nuclear reaction networks.
 
@@
 GPU-enabled integration has been explored by \citep{brock:2015}.
+
+Discuss frameworks vs.\ doing it outselves
+
+Discuss OpenMP, OpenACC, and CUDA.
 

### 2018-11-20 MAESTROeX_I/paper.tex
possible sign error

@@
 \end{eqnarray}
+\MarginPar{I don't think there should be a minus sign in front of the $-\frac{1}{\rho}\mathbf{G} \pi$ term as this has been moved outside the brackets.}
 where $\mathbf{G}$ approximates a cell-centered gradient from nodal

### 2018-11-20 MAESTROeX_I/paper.tex
add governing equations

@@
 %\listofchanges
+
+\section{Governing Equations}
+\begin{eqnarray}
+\frac{\partial(\rho X_k)}{\partial t} &=& -\nabla\cdot(\rho X_k\Ub) + \rho\omegadot_k,\label{eq:species}\\
+\frac{\partial\Ub}{\partial t} &=& -\Ub\cdot\nabla\Ub  - \frac{1}{\rho}\nabla\pi - \frac{\rho-\rho_0}{\rho} g\eb_r,\label{eq:momentum}\\
+\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc + \rho\Hext,\label{eq:enthalpy}
+\end{eqnarray}
+
+\begin{equation}
+\nabla\cdot(\beta_0\Ub) = \beta_0\left(S - \frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial t}\right),\label{eq:U divergence}
+\end{equation}
+\MarginPar{This is different from the Jacobs 2016 paper where we switched to alt\_energy\_fix.}
+where $\beta_0$ is a density-like variable that carries background stratification, defined as
+\begin{equation}
+\beta_0(r,t) = \rho_0(0,t)\exp\left(\int_0^r\frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial r'}dr'\right).
+\end{equation}
+The expansion term, 
+$S$, incorporates local 
+compressibility effects due to heat release from reactions, compositional changes, and 
+external sources,
+\begin{equation}
+S = -\sigma\sum_k\xi_k\omegadot_k + \frac{1}{\rho p_\rho}\sum_k p_{X_k}\omegadot_k + \sigma\Hnuc + \sigma\Hext.\label{eq:S}
+\end{equation}
 

### 2018-11-20 MAESTROeX_I/paper.tex
marginpars

@@
 Call {\bf React State}$[\rho^n, (\rho h)^n, X_k^n, T^n, (\rho\Hext)^n, p_0^n] \rightarrow [\rho^{(1)},(\rho h)^{(1)},X_k^{(1)},T^{(1)},(\rho \omegadot_k)^{(1)},(\rho \Hnuc)^{(1)}]$.
+\MarginPar{AJN: I think we can simplify the description here... this is somewhat confusing now.}
 
@@
   $\Ub = \uadvone$ and $\rho_0^n$.
-\MarginPar{We may not need $\rho_0$ if we predict $\rho X$ or $\rho$ and $X$.}
+\MarginPar{We may not need $\rho_0$ if we predict $\rho X$ or $\rho$ and $X$.
+AJN: If we end up keeping prediction of $\rho'$ and $X$, and $(\rho h)'$ we need to summarize this, particularly since $\rho_0$ and $(\rho h)_0$ treatment differs from previous papers.}
 

### 2018-11-20 MAESTROeX_I/paper.tex
fix references

@@
 
-Sample citation \cite{MAESTRO_I,MAESTRO_II,MAESTRO_III,MAESTRO_IV,MAESTRO_V}
+The key numerical developments of the MAESTRO algorithm are presented in a series of papers which we refer to as papers I-V (\cite{MAESTRO_I,MAESTRO_II,MAESTRO_III,MAESTRO_IV,MAESTRO_V}, respectively).
 
@@
 \section{Governing Equations}
+The derivation of the low Mach number hydrodynamics equations is given in Papers I-III.
+The resulting equations are
 \begin{eqnarray}
@@
 \end{eqnarray}
+where $\rho$, $\Ub$, and $h$ are the mass density,
+velocity and specific enthalpy, respectively, and
+$X_k$ are the mass fractions of species $k$ with associated
+production rate $\omegadot_k$.  The species are constrained
+such that $\sum_k X_k = 1$ giving $\rho = \sum_k (\rho X_k)$ and
+\begin{equation}
+\frac{\partial\rho}{\partial t} = -\nabla\cdot(\rho\Ub).
+\end{equation}
+The source terms $\Hext$ and $\Hnuc$ are the external heating rate and nuclear energy 
+generation rate per unit mass.  The pressure is decomposed into a hydrostatic base state
+ pressure, $p_0 = p_0(r,t)$, and a dynamic pressure, $\pi = \pi(\xb,t)$, such that 
+$|\pi|/p_0 = \mathcal{O}(M^2)$ (we use $\xb$ to represent the Cartesian coordinate 
+directions of the full state and $r$ to represent the radial coordinate direction for 
+the base state).  We also define a base state density, $\rho_0 = \rho_0(r,t)$, 
+which is in hydrostatic equilibrium with $p_0$, i.e., 
+$\nabla p_0 = -\rho_0 g\eb_r$, where $g=g(r,t)$ is
+the magnitude of the gravitational acceleration and $\eb_r$ is the unit vector in the
+outward radial direction.
 
+Mathematically, this system must still be closed by the equation of state which we
+express as a divergence constraint on the velocity field (see Paper III),
 \begin{equation}
@@
 \begin{equation}
-\beta_0(r,t) = \rho_0(0,t)\exp\left(\int_0^r\frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial r'}dr'\right).
+\beta_0(r,t) = \rho_0(0,t)\exp\left(\int_0^r\frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial r'}dr'\right),
 \end{equation}
-The expansion term, 
-$S$, incorporates local 
-compressibility effects due to heat release from reactions, compositional changes, and 
-external sources,
+where $\gammaonebar$ is the lateral average of $\Gamma_1 = d(\log p)/d(\log\rho)$ at constant entropy.
+The expansion term, $S$, incorporates local compressibility effects due to heat release from reactions, compositional changes, and external sources,
 \begin{equation}
-S = -\sigma\sum_k\xi_k\omegadot_k + \frac{1}{\rho p_\rho}\sum_k p_{X_k}\omegadot_k + \sigma\Hnuc + \sigma\Hext.\label{eq:S}
+S = -\sigma\sum_k\xi_k\omegadot_k + \frac{1}{\rho p_\rho}\sum_k p_{X_k}\omegadot_k + \sigma\Hnuc + \sigma\Hext,\label{eq:S}
 \end{equation}
+where $p_{X_k} \equiv \left. \partial p / \partial X_k
+\right|_{\rho,T,X_{j,j\ne k}}$, $\xi_k \equiv \left. \partial h /
+\partial X_k \right |_{p,T,X_{j,j\ne k}},
+p_\rho \equiv \left.
+\partial p/\partial \rho \right |_{T, X_k}$, and $\sigma \equiv
+p_T/(\rho c_p p_\rho)$, with $p_T \equiv \left. \partial p / \partial
+T \right|_{\rho, X_k}$ and $c_p \equiv \left.  \partial h / \partial T
+\right|_{p,X_k}$ is the specific heat at constant pressure.
+
+Previously we adopted an approach where we split the velocity into a base state component, $w_0$, and a local velocity $\Ubt$, so that
+\begin{equation}
+\Ub = w_0(r,t)\eb_r + \Ubt(\xb,t).
+\end{equation}
+We used $w_0$ to provide an estimate of the base state density evolution over a time step.
+In our new simplified approach, we instead use a predictor-corrector approach for evolving the base state.
 

### 2018-12-19 MAESTROeX_I/paper.tex
Include computation of etarho and psi in numerical scheme

@@
 
+\subsection{Motivation}
+We would like to revamp our previous MAESTRO code by moving to the AMReX framework
+so that we can take advantage of all of the latest features of the AMReX libraries. 
+In addition, we want to modify and simplify the existing numerical scheme 
+to both encourage new users and to develop a method that will lend itself more 
+easily to incorporating SDC and other high-order methods. 
+Irregular base state spacing is considered in order to improve spatial resolution 
+that have been seen in the previous MAESTRO code in some test cases.
+
 \section{Governing Equations}
@@
 The initialization step only occurs at the beginning of the simulation.
-The initialization step computes initial values for $\pi^{-\myhalf}$, $S^0$, and
-$S^1$ for use in time stepping routine.  We pass through {\bf Step 1} -- {\bf Step 12},
-but only keep these values.
+The initialization step computes initial values for $\pi^{-\myhalf}$, , $\eta_{\rho}^{-\myhalf}$, 
+$\psi^{-\myhalf}$, $S^0$, and $S^1$ for use in time stepping routine.  
+We pass through {\bf Step 1} -- {\bf Step 12}, but only keep these values.
 The initial values for $\Ub^0, \rho^0, (\rho h)^0, X_k^0, T^0,
@@
 \begin{equation}
-\nabla \cdot \left(\beta_0^n \uadvone\right) = \beta_0^n \left(S^{\nph,\star} - \frac{1}{\overline{\Gamma}_1^np_0^n}\frac{p_0^n-p_0^{n-1}}{\Delta t^{n-1}} + \chi^n\right),
+\nabla \cdot \left(\beta_0^n \uadvone\right) = \beta_0^n \left(S^{\nph,\star} - \frac{1}{\overline{\Gamma}_1^np_0^n}\psi^{n-\myhalf} + \chi^n \right),
 \end{equation}
@@
 where $\beta_0$ is computed as described in Appendix C of Paper III.
-\MarginPar{add volume discrepancy stuff}
+
+It should also be noted that we have found it better to approximate 
+the temporal changes in base pressure as 
+\begin{equation}
+\frac{\partial p_0}{\partial t} \approx \psi = \eta_{\rho} g
+\end{equation}
+instead of using the lagged pressure changes from previous time $t^{n-1}$ 
+in the predictor step.
 
@@
 && + \dt^n \left(\uadvone \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^{n} \nonumber \\
-&& + \dt^n \left( \frac{p_0^n - p_0^{n-1}}{\dt^{n-1}} \right),
+&& + \dt^n \psi^{n-\myhalf},
 \end{eqnarray}
@@
 && + \dt^n \left \{ \nabla \cdot \left (\uadvone p_0^{n} \right ) - p_0^{n} \nabla \cdot \uadvone \right \} \nonumber \\
-&& + \dt^n \left( \frac{p_0^n - p_0^{n-1}}{\dt^{n-1}} \right),
+&& + \dt^n \psi^{n-\myhalf},
 \end{eqnarray}
@@
 
-Note that during the first time step we use $\dt^n$ instead of $\dt^{n-1}$.
 Then, for each Cartesian cell where $\rho^{(2),\star} < \rho_\mathrm{cutoff}$,
@@
 the updated value $S^{\nph}$ rather than $S^{\nph,\star}$.
+Also, the temporal changes of $p_0$ are now computed directly instead of 
+approximated using $\psi$.
 We note that $\uadvtwo$ satisfies the divergence constraint
@@
 \nabla \cdot \left(\beta_0^{\nph} \uadvtwo\right) =
-\beta_0^{\nph} \left(S^{\nph} - \frac{1}{\overline{\Gamma}_1^{n+\myhalf}p_0^n}\frac{p_0^{n+1,\star}-p_0^n}{\Delta t^n} + \chi^{n+\myhalf}\right),
+\beta_0^{\nph} \left(S^{\nph} - \frac{1}{\overline{\Gamma}_1^{n+\myhalf}p_0^{n+1,\star}}\frac{p_0^{n+1,\star}-p_0^n}{\Delta t^n} + \chi^{n+\myhalf}\right),
 \end{equation}
@@
 \end{equation}
-\MarginPar{add volume discrepancy stuff}
 
@@
   $\uadvone, \rho_0^{n+1,\star}, (\rho h)_0^{n+1,\star}$, and $p_0^n$.
+  Also, the temporal changes of $p_0$ are now computed directly instead of 
+  approximated using $\psi$.
   In particular:
@@
    T(\rho^{(2)}, p_0^{n+1}, X_k^{(2)})$ (spherical geometry).
+
+\item Define a radial cell-centered $\etarho^{\nph}$.
+
+\begin{description}
+\item For planar geometry, $\etarho = \overline{\rho'(\Ub\cdot\eb_r)}$,
+\begin{equation}
+ \etarho^{\nph} =  {\rm {\bf Avg}} \sum_k \left[ \left(\uadvtwo \cdot \eb_r \right) (\rho X_k)^{\nph,\pred} \right]
+\end{equation}
+\item For spherical geometry, first construct 
+$\etarho^{{\rm cart},\nph} = [\rho'(\Ub\cdot\eb_r)]^{\nph}$ on Cartesian cell centers using:
+\begin{equation}
+\etarho^{{\rm cart},\nph} = \left[\left(\frac{\rho^{(1)}+\rho^{(2)}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{n+1}}{2}\right)\right] \cdot \left( \uadvtwo \cdot \eb_r \right).
+\end{equation}
+Then,
+\begin{equation}
+\etarho^{\nph,\star} = {\rm {\bf Avg}}\left(\etarho^{{\rm cart},\nph}\right).
+\end{equation}
+\end{description}
+
+\item Compute $\psi^{\nph}$: 
+\begin{equation}
+\psi_j^{\nph} = \eta_{\rho,j}^{\nph} g.
+\end{equation}
+
 \end{enumerate}
@@
 &=& \Ub^n - \dt \left[\uadvtwo \cdot \nabla \Ub^{\nph,\pred} \right] \nonumber \\
-&& - \dt \left[ - \frac{1}{\rho^\nph} \mathbf{G} \pi^\nmh + \frac{\left(\rho^\nph-\rho_0^\nph\right)}{\rho^\nph} g^{\nph} \eb_r \right],\nonumber \\
+&& - \dt \left[ \frac{1}{\rho^\nph} \mathbf{G} \pi^\nmh + \frac{\left(\rho^\nph-\rho_0^\nph\right)}{\rho^\nph} g^{\nph} \eb_r \right],\nonumber \\
 \end{eqnarray}
-\MarginPar{I don't think there should be a minus sign in front of the $-\frac{1}{\rho}\mathbf{G} \pi$ term as this has been moved outside the brackets.}
 where $\mathbf{G}$ approximates a cell-centered gradient from nodal
@@
 \end{equation}
-\MarginPar{I think we need the 1/gamma1bar part too}
 where

### 2018-12-19 MAESTROeX_I/paper.tex
add refs, some intro

@@
 
-\title{Title}
+\title{MAESTROeX}
 
@@
 
-\correspondingauthor{Doreen Fan}
+\correspondingauthor{MAESTROeX}
 \email{DFan@lbl.gov}
@@
 \begin{abstract}
-Insert abstract here.
+We present MAESTROeX, a next-generation solver for low Mach number astrophysical flows.
+Here we summarize implementation details as well new spatiotemporal integration schemes
+developed for increased accuracy and a simpler base state treatment.  The code is written
+in the AMReX software framework for massively parallel block-structured adaptive mesh
+refinement applications.
 
@@
 
+Many astrophysical flows are highly subsonic; sound waves carry sufficiently little energy that they do not significantly affect the convective dynamics of the system.
+In many of these flows, long-time convective dynamics are of interest.
+Thus, using a traditional compressible astrophysical solver is highly inefficient.
+One approach to this problem is to use low Mach number hydrodynamic models.
+
 The key numerical developments of the MAESTRO algorithm are presented in a series of papers which we refer to as papers I-V (\cite{MAESTRO_I,MAESTRO_II,MAESTRO_III,MAESTRO_IV,MAESTRO_V}, respectively).
-
-%% This command is needed to show the entire author+affilation list when
-%% the collaboration and author truncation commands are used.  It has to
-%% go at the end of the manuscript.
-%\allauthors
-
-%% Include this line if you are using the \added, \replaced, \deleted
-%% commands to see a summary list of all changes at the end of the article.
-%\listofchanges
+Applications include the convective phase preceding Chandrasekhar mass models for type Ia supernovae \cite{MAESTRO_convection,MAESTRO_AMR,MAESTRO_CASTRO}, 
+convection in massivve stars \cite{Gilet:2013},
+sub-Chandrasekhar white dwarfs \cite{subChandra_I,subChandra_II}, and
+Type I X-ray bursts \cite{XRB_I,XRB_II,XRB_III}.
 

### 2018-12-19 MAESTROeX_I/paper.tex
minor

@@
 
+\section{Implementation and Performance}
+show strong and/or weak scaling performance (with OpenMP threads/tiling), comparisons with old MAESTRO?
+
 \section{Test Problems}
-
-\section{Performance and Scaling}
-
-show OpenMP performance, maybe some comparisons with old Maestro?
+white dwarf convection runs at varying resolution
 
@@
 
-future stuff: rotation, SDC, high-order, GPUs
+future stuff
+science: rotation, solar physics, MHD
+algorithm: SDC, high-order
+implementation: GPUs
 
 \acknowledgements
+
+This work was supported by
 

### 2018-12-30 astronum_2018/plots/aprox19_new/sdc_plot.py
new run

@@ -1,0 +1,66 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+
+#mpl.rcParams['xtick.labelsize'] = "small"
+
+# read the Strang data into a numpy array
+strang = np.loadtxt("strang_diag.out")
+
+# now manually read it looking for the times that indicate when the Strang halves began
+strang_starts = []
+strang_second = []
+with open("strang_diag.out") as f:
+    for line in f:
+        if line.startswith(" # start of step"):
+            _, time = line.split("=")
+            strang_starts.append(float(time))
+        elif line.startswith(" # second half of strang"):
+            _, time = line.split("=")
+            strang_second.append(float(time))
+
+strang_starts = sorted(set(strang_starts))
+strang_second = sorted(set(strang_second))
+
+print(strang_starts)
+
+# now get the SDC data
+sdc = np.loadtxt("sdc_diag.out")
+
+# we have 2 SDC iterations here, so be able to separate them
+idx_iter0 = sdc[:,1] == 0
+idx_iter1 = sdc[:,1] == 1
+
+
+# make the plots
+plt.scatter(strang[:,0], strang[:,2], marker="x", s=15, label="Strang splitting")
+plt.scatter(sdc[idx_iter1,0], sdc[idx_iter1,3], marker="o", s=15, label="SDC")
+
+# draw timestep boundaries
+for time in strang_starts:
+    plt.plot([time, time], [0,1], color="k", linewidth=2)
+
+for time in strang_second:
+    plt.plot([time, time], [0,1], color="0.5", linestyle=":", linewidth=2)
+
+plt.xlim(2.0178527608932999E-005, 2.0378470571915846E-005)
+plt.ylim(0.535, 0.5365)
+
+plt.legend(frameon=False)
+
+plt.tight_layout()
+
+f = plt.gcf()
+f.set_size_inches(9.0, 7.0)
+
+plt.xlabel("time (s)", fontsize="large")
+plt.ylabel(r"$X({}^4\mathrm{He})$", fontsize="large")
+
+plt.ticklabel_format(style='sci', axis='x', scilimits=(0,0))
+
+plt.tight_layout()
+
+plt.savefig("sdc_plot.png")
+
+print("sdc points = ", len(sdc[idx_iter1,0]))
+print("strang points = ", len(strang[:,0]))

### 2018-12-31 astronum_2018/astronum-xrb.tex
add e-mail
some figs
updated version
add note about GPUs

@@
 \address{$^6$Los Alamos National Laboratory, Los Alamos, NM, 87545 USA}
+
+\ead{michael.zingale@stonybrook.edu}
 
@@
 
-\begin{figure}
+\begin{figure}[t]
 \centering
@@
   temperature and burning rate are highest.  }
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.9\linewidth]{xrb_profiles}
+\caption{\label{fig:xrb_profiles} Vertical profiles at several laterial distances
+  from the vertical axis through the plots shown in Figure~\ref{fig:xrb}.}
 \end{figure}
@@
 is about $10\times$ faster using GPUs on a node than the entire node
-of CPU cores.  We see similar speed ups with reaction networks, with
-these GPU ports under active development.  This means a GPU
-calculation could lower the above costs by a factor of 10.  GPU
-offloading of the reaction networks will also allow us to explore
-larger networks and more detailed nuclear physics.
+of CPU cores (comparing 4 NVIDIA Volta GPUs to 20 IBM Power9 cores).
+We see similar speed ups with reaction networks, with these GPU ports
+under active development.  This means a GPU calculation could lower
+the above costs by a factor of 10.  GPU offloading of the reaction
+networks will also allow us to explore larger networks and more
+detailed nuclear physics.
 
@@
 reacting hydrodynamics methods to follow shortly.  The XRB simulations
-are our target application.
+are our target application.  We note that there has been some
+development in the community on implicit time-integration methods for
+compressible flow, but for the XRB problem, the wide range of Mach
+number seen in the domain (exceeding unity near the top of the
+atmosphere, ahead of the flame) removes much of the advantages of
+implicit timestepping, so we focus on explicit methods.
 
@@
 All of the above developments still involve fully compressible
-hydrodynamics, but multiscale methods~\cite{weinan2011principles} where we couple different hydrodynamics
-solvers together in a single simulation may ultimately be needed to model the full star while resolving the burning front.
+hydrodynamics, but multiscale methods~\cite{weinan2011principles}
+where we couple different hydrodynamics solvers together in a single
+simulation may ultimately be needed to model the full star while
+resolving the burning front.
 %\MarginPar{Alice and Ian contribute here}
-Such techniques rely on it being possible to decompose the dynamics of the system
-into two scales, e.g.~a short and a long lengthscale, and that the processes at each
-scale are at most weakly coupled with each other. In the context of XRBs, one possibility for
-multiscale modeling would be to model the large scale dynamics using a shallow water model
-(like that of~\cite{SPIT_ETAL02}) to capture the effects of the Coriolis force, and use
-a low Mach number or
-a fully compressible model in the vicinity of the burning front to capture
-the turbulent burning processes. Given that the computational costs of the shallow water model
-would be negligible compared to those of the low Mach/compressible model, the overall cost of
-this multiscale model would be determined by the size of the low Mach/compressible region only. This
-approach would therefore allow us to capture dynamics across the full star
-without sacrificing resolution around the flame.
-As with the more aggressive gridding strategy above, the hybrid approach
-reduces the number of zones with the largest computational cost through mesh
-refinement, and also reduces the computational costs where possible by
-employing cheaper physical models in regions where this is possible.
-An example of coupling compressible and low Mach number models implemented using the \amrex~framework, but in a terrestrial context,
-is~\cite{Motheau2018}. An initial implementation, \cite{Harpole2018}, in an
-astrophysical context, coupling shallow water to compressible models in both
-Newtonian and relativistic gravity, indicates that considerable computational
-efficiencies are possible, but that close attention to the coupling between the
-models is needed.
+Such techniques rely on it being possible to decompose the dynamics of
+the system into two scales, e.g.~a short and a long lengthscale, and
+that the processes at each scale are at most weakly coupled with each
+other. In the context of XRBs, one possibility for multiscale modeling
+would be to model the large scale dynamics using a shallow water model
+(like that of~\cite{SPIT_ETAL02}) to capture the effects of the
+Coriolis force, and use a low Mach number or a fully compressible
+model in the vicinity of the burning front to capture the turbulent
+burning processes. Given that the computational costs of the shallow
+water model would be negligible compared to those of the low
+Mach/compressible model, the overall cost of this multiscale model
+would be determined by the size of the low Mach/compressible region
+only. This approach would therefore allow us to capture dynamics
+across the full star without sacrificing resolution around the flame.
+As with the more aggressive gridding strategy above, the hybrid
+approach reduces the number of zones with the largest computational
+cost through mesh refinement, and also reduces the computational costs
+where possible by employing cheaper physical models in regions where
+this is possible.  An example of coupling compressible and low Mach
+number models implemented using the \amrex~framework, but in a
+terrestrial context, is~\cite{Motheau2018}. An initial implementation,
+\cite{Harpole2018}, in an astrophysical context, coupling shallow
+water to compressible models in both Newtonian and relativistic
+gravity, indicates that considerable computational efficiencies are
+possible, but that close attention to the coupling between the models
+is needed.
 

### 2018-12-31 astronum_2018/astronum-xrb.tex
more referee addressing

--- 2018-10-03 astronum_2018/astronum-xrb.tex
@@
 horizontal CFL number that is large enough to obtain long timescale
-evolution~\cite{cavecchi:2012}.
-These calculations display key features of the propagation
-mechanism, namely a balance between hydrodynamics and the pure
-flame physics. The flame proceeds mainly via conduction, thus being a
-deflagration. The flame front, however, is not vertical but inclined,
-the angle with the horizontal being very small: $\theta \sim
-H/R_{\rm{Ro}} \sim 10^{-3}$. Here $H$ is the scale height and
-$R_{\rm{Ro}}$ is the Rossby radius. It is the hydrodynamical balance
-between the Coriolis force, gravity and pressure gradients that
-determines the Rossby radius and therefore the inclination angle of the
-flame front. The extended surface of the flame front is what leads to
-propagation speeds of order $10^5 $~km/s, in good agreement with
-observations. Due to its dependence on the Coriolis force, the flame
-speed changes across the surface of the NS (being faster at the
-equator), leading to potentially detectable effects if the rise of the
-burst is observed with enough resolution
-\cite{art-2015-cavecchi-etal}.
+evolution~\cite{cavecchi:2012}.  These calculations display key
+features of the propagation mechanism, namely a balance between
+hydrodynamics and the pure flame physics. The flame proceeds mainly
+via conduction, thus being a deflagration. The flame front, however,
+is not vertical but inclined, the angle with the horizontal being very
+small: $\theta \sim H/R_{\rm{Ro}} \sim 10^{-3}$. Here $H$ is the scale
+height and $R_{\rm{Ro}}$ is the Rossby radius. It is the
+hydrodynamical balance between the Coriolis force, gravity and
+pressure gradients that determines the Rossby radius and therefore the
+inclination angle of the flame front. The extended surface of the
+flame front is what leads to propagation speeds of order $10^5 $~km/s,
+in good agreement with observations. Due to its dependence on the
+Coriolis force, the flame speed changes across the surface of the NS
+(being faster at the equator), leading to potentially detectable
+effects if the rise of the burst is observed with enough resolution
+\cite{art-2015-cavecchi-etal}.  We note that no calculations have
+considered the effects of the flame interacting with preexisting
+turbulence ahead of the flame from simmering that may be occurring
+there, as we characterized in \cite{xrb3}.  Flame-turbulence
+interactions have long been seen as a means to accelerate a flame in
+Type Ia supernovae, so future studies should look at turbulence in the
+layer.  Capturing turbulence, of course, requires high resolution, so
+the ideas described here will be needed.
 
@@
 Figure~\ref{fig:xrb} shows the flame structure for an initial 2D run
-at 0.007~s.  The burning layer is only about 10~m high.  Underneath is
-\isot{Ni}{56}, representative of the underlying neutron star, and
-above it is a very low density ($10^{-4}~\gcc$) buffer.  Only a
-portion of the vertical extent is shown.  The top panel shows the
-temperature structure.  The initial perturbation reached to $2\times
-10^4~\mathrm{cm}$, so at this point, the extent of the burnt region
-has nearly doubled.  The flame structure is quite wide, since there
-are multiple burning stages beyond just helium to carbon, and this
-shows up in the composition.  This is seen more clearly in the next
-panel which shows the mean molecular weight of the nuclei ($\bar{A}$).
-We see that the further behind the head of the flame we are, the heavier
-the ash nuclei.  We also see the head of the flame
-appears lifted off of the base of the burning layer---unburned
-material underlies the spreading flame.  This is
-reminiscent of the flame structure seen in the earlier simulations of
-\cite{cavecchi:2012}.  The third panel shows the energy generation rate.
-It is strongest just above the neutron star surface, behind the flame.
-We again see that the head of the flame is detached from the base
-of the atmosphere, with the strongest burning at the head of the flame
-slightly higher up in the atmosphere.   Finally, the bottom
-panel shows the velocity through the plane of the simulation---this is
-induced by the Coriolis force as the burning fluid expands and begins to spread laterally.
-A tight hurricane structure has setup as a result of the flame
-spreading.  This is the geostrophic balance discussed in
-\cite{SPIT_ETAL02}.    These calculations are ongoing and will be
-the subject of a detailed study in the near future.
+at 0.007~s.  Vertical profiles at a few radial distances from the
+origin are shown for the same quantities in
+Figure~\ref{fig:xrb-profiles}.  The burning layer is only about 10~m
+high.  Underneath is \isot{Ni}{56}, representative of the underlying
+neutron star, and above it is a very low density ($10^{-4}~\gcc$)
+buffer.  Only a portion of the vertical extent is shown.  The top
+panel shows the temperature structure.  The initial perturbation
+reached to $2\times 10^4~\mathrm{cm}$, so at this point, the extent of
+the burnt region has nearly doubled.  The flame structure is quite
+wide, since there are multiple burning stages beyond just helium to
+carbon, and this shows up in the composition.  This is seen more
+clearly in the next panel which shows the mean molecular weight of the
+nuclei ($\bar{A}$).  We see that the further behind the head of the
+flame we are, the heavier the ash nuclei.  We also see the head of the
+flame appears lifted off of the base of the burning layer---unburned
+material underlies the spreading flame.  This is reminiscent of the
+flame structure seen in the earlier simulations of
+\cite{cavecchi:2012}.  The third panel shows the energy generation
+rate.  It is strongest just above the neutron star surface, behind the
+flame.  We again see that the head of the flame is detached from the
+base of the atmosphere, with the strongest burning at the head of the
+flame slightly higher up in the atmosphere.  Finally, the bottom panel
+shows the velocity through the plane of the simulation---this is
+induced by the Coriolis force as the burning fluid expands and begins
+to spread laterally.  A tight hurricane structure has setup as a
+result of the flame spreading.  This is the geostrophic balance
+discussed in \cite{SPIT_ETAL02}.  These calculations are ongoing and
+will be the subject of a detailed study in the near future.
 
-\begin{figure}
+\begin{figure}[t]
 \centering
@@
   temperature and burning rate are highest.  }
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.8\linewidth]{xrb_profiles}
+\caption{\label{fig:xrb_profiles} Vertical profiles at several laterial distances
+  from the vertical axis through the plots shown in Figure~\ref{fig:xrb}.  These
+  profiles sample the region behind and ahead of the flame.}
 \end{figure}

### 2018-12-31 astronum_2018/astronum-xrb.tex
fix typo
corrected typo in figure caption

@@
 \includegraphics[width=0.8\linewidth]{xrb_profiles}
-\caption{\label{fig:xrb_profiles} Vertical profiles at several laterial distances
+\caption{\label{fig:xrb_profiles} Vertical profiles at several radial distances
   from the vertical axis through the plots shown in Figure~\ref{fig:xrb}.  These

### 2018-12-31 astronum_2018/astronum-xrb.tex

--- 2018-12-31 astronum_2018/astronum-xrb.tex
@@
 Figure~\ref{fig:xrb} shows the flame structure for an initial 2D run
-at 0.007~s.  The burning layer is only about 10~m high.  Underneath is
-\isot{Ni}{56}, representative of the underlying neutron star, and
-above it is a very low density ($10^{-4}~\gcc$) buffer.  Only a
-portion of the vertical extent is shown.  The top panel shows the
-temperature structure.  The initial perturbation reached to $2\times
-10^4~\mathrm{cm}$, so at this point, the extent of the burnt region
-has nearly doubled.  The flame structure is quite wide, since there
-are multiple burning stages beyond just helium to carbon, and this
-shows up in the composition.  This is seen more clearly in the next
-panel which shows the mean molecular weight of the nuclei ($\bar{A}$).
-We see that the further behind the head of the flame we are, the heavier
-the ash nuclei.  We also see the head of the flame
-appears lifted off of the base of the burning layer---unburned
-material underlies the spreading flame.  This is
-reminiscent of the flame structure seen in the earlier simulations of
-\cite{cavecchi:2012}.  The third panel shows the energy generation rate.
-It is strongest just above the neutron star surface, behind the flame.
-We again see that the head of the flame is detached from the base
-of the atmosphere, with the strongest burning at the head of the flame
-slightly higher up in the atmosphere.   Finally, the bottom
-panel shows the velocity through the plane of the simulation---this is
-induced by the Coriolis force as the burning fluid expands and begins to spread laterally.
-A tight hurricane structure has setup as a result of the flame
-spreading.  This is the geostrophic balance discussed in
-\cite{SPIT_ETAL02}.    These calculations are ongoing and will be
-the subject of a detailed study in the near future.
+at 0.007~s.  Vertical profiles at a few radial distances from the
+origin are shown for the same quantities in
+Figure~\ref{fig:xrb-profiles}.  The burning layer is only about 10~m
+high.  Underneath is \isot{Ni}{56}, representative of the underlying
+neutron star, and above it is a very low density ($10^{-4}~\gcc$)
+buffer.  Only a portion of the vertical extent is shown.  The top
+panel shows the temperature structure.  The initial perturbation
+reached to $2\times 10^4~\mathrm{cm}$, so at this point, the extent of
+the burnt region has nearly doubled.  The flame structure is quite
+wide, since there are multiple burning stages beyond just helium to
+carbon, and this shows up in the composition.  This is seen more
+clearly in the next panel which shows the mean molecular weight of the
+nuclei ($\bar{A}$).  We see that the further behind the head of the
+flame we are, the heavier the ash nuclei.  We also see the head of the
+flame appears lifted off of the base of the burning layer---unburned
+material underlies the spreading flame.  This is reminiscent of the
+flame structure seen in the earlier simulations of
+\cite{cavecchi:2012}.  The third panel shows the energy generation
+rate.  It is strongest just above the neutron star surface, behind the
+flame.  We again see that the head of the flame is detached from the
+base of the atmosphere, with the strongest burning at the head of the
+flame slightly higher up in the atmosphere.  Finally, the bottom panel
+shows the velocity through the plane of the simulation---this is
+induced by the Coriolis force as the burning fluid expands and begins
+to spread laterally.  A tight hurricane structure has setup as a
+result of the flame spreading.  This is the geostrophic balance
+discussed in \cite{SPIT_ETAL02}.  These calculations are ongoing and
+will be the subject of a detailed study in the near future.
 
@@
 \caption{\label{fig:xrb_profiles} Vertical profiles at several laterial distances
-  from the vertical axis through the plots shown in Figure~\ref{fig:xrb}.}
+  from the vertical axis through the plots shown in Figure~\ref{fig:xrb}.  These
+  profiles sample the region behind and ahead of the flame.}
 \end{figure}

### 2018-12-31 astronum_2018/astronum-xrb.tex
fix ref
refix the caption

@@
 origin are shown for the same quantities in
-Figure~\ref{fig:xrb-profiles}.  The burning layer is only about 10~m
+Figure~\ref{fig:xrb_profiles}.  The burning layer is only about 10~m
 high.  Underneath is \isot{Ni}{56}, representative of the underlying
@@
 \includegraphics[width=0.9\linewidth]{xrb_profiles}
-\caption{\label{fig:xrb_profiles} Vertical profiles at several laterial distances
+\caption{\label{fig:xrb_profiles} Vertical profiles at several radial distances
   from the vertical axis through the plots shown in Figure~\ref{fig:xrb}.  These

### 2019-02-11 MAESTROeX_I/paper.tex
small updates

--- 2018-12-19 MAESTROeX_I/paper.tex
@@
 
-\title{MAESTROeX}
+\title{MAESTROeX: A Massively Parallel Low Mach Number Astrophysical Solver}
 
@@
 
-\author{Doreen Fan}
+\author[0000-0002-3246-4315]{Duoming Fan}
 \affil{Lawrence Berkelely National Laboratory \\
@@
 
-\author{Andrew Nonaka}
+\author[0000-0003-1791-0265]{Andrew Nonaka}
 \affil{Lawrence Berkelely National Laboratory \\
@@
 
-\author{Ann Almgren}
+\author[0000-0003-2103-312X]{Ann S. Almgren}
 \affil{Lawrence Berkelely National Laboratory \\
@@
 \begin{abstract}
-We present MAESTROeX, a next-generation solver for low Mach number astrophysical flows.
-Here we summarize implementation details as well new spatiotemporal integration schemes
-developed for increased accuracy and a simpler base state treatment.  The code is written
-in the AMReX software framework for massively parallel block-structured adaptive mesh
-refinement applications.
-
+We present MAESTROeX, a massively parallel solver for low Mach number astrophysical flows.
+This new solver leverages the algorithmic capabilities of the previous MAESTRO code
+while taking advantage of the next-generation AMReX software framework block-structured
+adaptive mesh refinement applications.
+Using our previous studies on white dwarf convection as a benchmark, we characterize
+the performance of our new code.
+We have also developed new temporal integration schemes that have a much simpler treatment
+of the base state, while allowing for more straightforward extensions to state-of-the
+art, high-order multiphysics coupling schemes.
+Also, we present an improved spatial coupling scheme between the hydrostatic base
+state and the Cartesian grid full-state.
 \end{abstract}
@@
 
-Many astrophysical flows are highly subsonic; sound waves carry sufficiently little energy that they do not significantly affect the convective dynamics of the system.
+Many astrophysical flows are highly subsonic; 
+sound waves carry sufficiently little energy that they do not significantly affect the convective dynamics of the system.
 In many of these flows, long-time convective dynamics are of interest.
@@
 One approach to this problem is to use low Mach number hydrodynamic models.
+In a low Mach number approach, sound waves are eliminated from the governing equations by requiring that the
+thermodynamic variables evolve subject to a specified ambient pressure that varies only as a function of altitude and time.
+In the low Mach equations, the equation of state is recast as a divergence constraint on velocity, requiring an
+elliptic solve for pressure in the spirit of projection methods for incompressible flow.
 
@@
 
-\section{Performance and Scaling}
+\subsection{Scaling}
+Scaling with OpenMP
 
-show OpenMP performance, maybe some comparisons with old Maestro?
+\subsection{Performance Compared to Original MAESTRO}
+Compare to original MAESTRO
+
+\section{Test Problems}
+white dwarf convection runs with 3 algorithms (original, new temporal, new temporal + irregular base state)
 
@@
 
-future stuff: rotation, SDC, high-order, GPUs
+science: rotation, solar physics, MHD
+algorithm: SDC, high-order
+implementation: GPUs
 
 \acknowledgements
+
+This work was supported by SciDAC... (contract number)
+
+This research used resources of the National Energy Research Scientific Computing Center (NERSC), a U.S.
+Department of Energy Office of Science User Facility operated under Contract No.
+DE-AC02-05CH11231.
 

### 2019-02-11 MAESTROeX_I/paper.tex
minor

@@
 The derivation of the low Mach number hydrodynamics equations is given in Papers I-III.
+We take the standard equations of reacting, compressible flow, and recast the equation
+of state as a divergence constraint on the velocity field.  This is done by taking
+the Langrangian derivative of the equation of state for pressure as a function of the
+density, composition, and enthalpy, enforcing the constraint that the pressure is
+a prescribed function of altitutde and time based on the hydrostatic equilibrium condition,
+and substituting the equations for species and enthalpy in the resulting Lagrangian derivative.
 The resulting equations are
 \begin{eqnarray}
+\frac{\partial\Ub}{\partial t} &=& -\Ub\cdot\nabla\Ub  - \frac{1}{\rho}\nabla\pi - \frac{\rho-\rho_0}{\rho} g\eb_r,\label{eq:momentum}\\
 \frac{\partial(\rho X_k)}{\partial t} &=& -\nabla\cdot(\rho X_k\Ub) + \rho\omegadot_k,\label{eq:species}\\
-\frac{\partial\Ub}{\partial t} &=& -\Ub\cdot\nabla\Ub  - \frac{1}{\rho}\nabla\pi - \frac{\rho-\rho_0}{\rho} g\eb_r,\label{eq:momentum}\\
 \frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc + \rho\Hext,\label{eq:enthalpy}
 \end{eqnarray}
-where $\rho$, $\Ub$, and $h$ are the mass density,
+\begin{equation}
+\nabla\cdot(\beta_0\Ub) = \beta_0\left(S - \frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial t}\right).\label{eq:U divergence}
+\end{equation}
+Here $\rho$, $\Ub$, and $h$ are the mass density,
 velocity and specific enthalpy, respectively, and
@@
 
-Mathematically, this system must still be closed by the equation of state which we
-express as a divergence constraint on the velocity field (see Paper III),
-\begin{equation}
-\nabla\cdot(\beta_0\Ub) = \beta_0\left(S - \frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial t}\right),\label{eq:U divergence}
-\end{equation}
+Mathematically, equations (\ref{eq:momentum})-(\ref{eq:enthalpy}) must still be closed by the equation of state which we
+express as a divergence constraint on the velocity field (\ref{eq:U divergence}),
 \MarginPar{This is different from the Jacobs 2016 paper where we switched to alt\_energy\_fix.}
@@
 
-This work was supported by SciDAC... (contract number)
-
+The work at LBNL was supported by the U.S. Department of Energy's Scientific Discovery Through Advanced Computing (SciDAC) program under contract No. DE-AC02-05CH11231.
 This research used resources of the National Energy Research Scientific Computing Center (NERSC), a U.S.

### 2019-02-11 MAESTROeX_I/paper.tex
remove rho Hext from paper... not using external heating.  attempting to simplify text since algorithm is already so complex

@@
 \newcommand{\gammaonebar}{\overline{\Gamma_1}}
-\newcommand{\Hext}{H_{\rm ext}}
 \newcommand{\Hnuc}{H_{\rm nuc}}
@@
 \frac{\partial(\rho X_k)}{\partial t} &=& -\nabla\cdot(\rho X_k\Ub) + \rho\omegadot_k,\label{eq:species}\\
-\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc + \rho\Hext,\label{eq:enthalpy}
+\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc,\label{eq:enthalpy}
 \end{eqnarray}
@@
 \end{equation}
-The source terms $\Hext$ and $\Hnuc$ are the external heating rate and nuclear energy 
-generation rate per unit mass.  The pressure is decomposed into a hydrostatic base state
+Here $\Hnuc$ is the nuclear energy generation rate per unit mass.
+The pressure is decomposed into a hydrostatic base state
  pressure, $p_0 = p_0(r,t)$, and a dynamic pressure, $\pi = \pi(\xb,t)$, such that 
@@
 \begin{equation}
-S = -\sigma\sum_k\xi_k\omegadot_k + \frac{1}{\rho p_\rho}\sum_k p_{X_k}\omegadot_k + \sigma\Hnuc + \sigma\Hext,\label{eq:S}
+S = -\sigma\sum_k\xi_k\omegadot_k + \frac{1}{\rho p_\rho}\sum_k p_{X_k}\omegadot_k + \sigma\Hnuc,\label{eq:S}
 \end{equation}
@@
 
-\item[Step 0.] {\em Initialization}
+\item[Step 0] {\em Initialization}
 The initialization step only occurs at the beginning of the simulation.
@@
 %--------------------------------------------------------------------------
-\item[Step 1.] {\em React the full state through the first time interval of $\dt / 2.$}
+\item[Step 1] {\em React the full state through the first time interval of $\dt / 2.$}
 
-Call {\bf React State}$[\rho^n, (\rho h)^n, X_k^n, T^n, (\rho\Hext)^n, p_0^n] \rightarrow [\rho^{(1)},(\rho h)^{(1)},X_k^{(1)},T^{(1)},(\rho \omegadot_k)^{(1)},(\rho \Hnuc)^{(1)}]$.
+Call {\bf React State}$[\rho^n, (\rho h)^n, X_k^n, T^n, p_0^n] \rightarrow [\rho^{(1)},(\rho h)^{(1)},X_k^{(1)},T^{(1)},(\rho \omegadot_k)^{(1)},(\rho \Hnuc)^{(1)}]$.
 \MarginPar{AJN: I think we can simplify the description here... this is somewhat confusing now.}
@@
 
-\item[Step 2.] {\em Compute the provisional time-centered expansion,
+\item[Step 2] {\em Compute the provisional time-centered expansion,
     $S^{\nph,\star}$.}
@@
 %--------------------------------------------------------------------------
-\item[Step 3.] {\em Construct the provisional time-centered advective velocity on
+\item[Step 3] {\em Construct the provisional time-centered advective velocity on
 edges, $\uadvone$.}
@@
 %--------------------------------------------------------------------------
-\item[Step 4.] {\em Advect the full state through a time interval of $\dt.$}
+\item[Step 4] {\em Advect the full state through a time interval of $\dt.$}
 
@@
 \begin{equation}
-\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc + \rho\Hext,
+\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc,
 \end{equation}
@@
 %--------------------------------------------------------------------------
-\item[Step 5.] {\em React the full state through a second time interval of $\dt / 2.$}
+\item[Step 5] {\em React the full state through a second time interval of $\dt / 2.$}
 
-Call {\bf React State}$[ \rho^{(2),\star},(\rho h)^{(2),\star}, X_k^{(2),\star}, T^{(2),\star},
-(\rho\Hext)^{(2),\star}, p_0^{n+1,\star}]$\\
+Call {\bf React State}$[ \rho^{(2),\star},(\rho h)^{(2),\star}, X_k^{(2),\star}, T^{(2),\star}, p_0^{n+1,\star}]$\\
 \phantom{ }\hfill $\rightarrow [ \rho^{n+1,\star},(\rho h)^{n+1,\star}, X_k^{n+1,\star}, T^{n+1,\star}, (\rho \omegadot_k)^{(2),\star}, (\rho \Hnuc)^{(2),\star} ].$
@@
 %--------------------------------------------------------------------------
-\item[Step 6.] {\em Compute the time-centered expansion, $S^{\nph,\star}$.}
+\item[Step 6] {\em Compute the time-centered expansion, $S^{\nph,\star}$.}
 
@@
 \begin{equation}
-S^{n+1,\star} =  -\sigma  \sum_k  \xi_k  (\omegadot_k)^{(2),\star}  + \frac{1}{\rho^{n+1,\star} p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2),\star} + \sigma \Hnuc^{(2),\star} + \sigma \Hext^{(2),\star},
+S^{n+1,\star} =  -\sigma  \sum_k  \xi_k  (\omegadot_k)^{(2),\star}  + \frac{1}{\rho^{n+1,\star} p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2),\star} + \sigma \Hnuc^{(2),\star},
 \end{equation}
@@
 %--------------------------------------------------------------------------
-\item[Step 7.] {\em Construct the time-centered advective velocity on edges, $\uadvtwo$.}
+\item[Step 7] {\em Construct the time-centered advective velocity on edges, $\uadvtwo$.}
 
@@
 %--------------------------------------------------------------------------
-\item[Step 8.] {\em Advect the base state and full state through a time interval of $\dt.$}
+\item[Step 8] {\em Advect the base state and full state through a time interval of $\dt.$}
 
@@
 %--------------------------------------------------------------------------
-\item[Step 9.] {\em React the full state through a second time interval of $\dt / 2.$}
+\item[Step 9] {\em React the full state through a second time interval of $\dt / 2.$}
 
-Call {\bf React State}$[\rho^{(2)},(\rho h)^{(2)}, X_k^{(2)},T^{(2)}, (\rho\Hext)^{(2)}, p_0^{n+1}] \rightarrow [\rho^{n+1}, (\rho h)^{n+1}, X_k^{n+1}, T^{n+1}, (\rho \omegadot_k)^{(2)}, (\rho \Hnuc)^{(2)} ].$
+Call {\bf React State}$[\rho^{(2)},(\rho h)^{(2)}, X_k^{(2)},T^{(2)}, p_0^{n+1}] \rightarrow [\rho^{n+1}, (\rho h)^{n+1}, X_k^{n+1}, T^{n+1}, (\rho \omegadot_k)^{(2)}, (\rho \Hnuc)^{(2)} ].$
 
@@
 %--------------------------------------------------------------------------
-\item[Step 10.] {\em Define the new time expansion, $S^{n+1}$, and $\overline{\Gamma}_1^{n+1}$.}
+\item[Step 10] {\em Define the new time expansion, $S^{n+1}$, and $\overline{\Gamma}_1^{n+1}$.}
 
@@
   S^{n+1} =  -\sigma  \sum_k  \xi_k (\omegadot_k)^{(2)}  + \sigma \Hnuc^{(2)} +
-  \frac{1}{\rho^{n+1} p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2)}
-   + \sigma \Hext^{(2)},
+  \frac{1}{\rho^{n+1} p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2)},
 \end{equation}
@@
 %--------------------------------------------------------------------------
-\item[Step 11.] {\em Update the velocity}.
+\item[Step 11] {\em Update the velocity}.
 
@@
 %--------------------------------------------------------------------------
-\item[Step 12.] {\em Compute a new $\dt.$}
+\item[Step 12] {\em Compute a new $\dt.$}
 

### 2019-02-11 MAESTROeX_I/paper.tex
more simplifications.  need to discuss etarho and psi

@@
 
+\MarginPar{AJN: eliminated references to volume discrepancy for now; we'd only use it for planar at the moment, don't want to open up that can of worms here}
+
 \begin{description}
@@
 
-Call {\bf React State}$[\rho^n, (\rho h)^n, X_k^n, T^n, p_0^n] \rightarrow [\rho^{(1)},(\rho h)^{(1)},X_k^{(1)},T^{(1)},(\rho \omegadot_k)^{(1)},(\rho \Hnuc)^{(1)}]$.
+Call {\bf React State}$[(\rho X_k)^n, (\rho h)^n, p_0^n] \rightarrow [(\rho X_k)^{(1)}, (\rho h)^{(1)}, (\rho \omegadot_k)^{(1)}, (\rho \Hnuc)^{(1)}]$.
 \MarginPar{AJN: I think we can simplify the description here... this is somewhat confusing now.}
@@
 
-Using the equation below,
-%
-\begin{equation}
-\frac{\partial\Ub}{\partial t} = -\Ub\cdot\nabla\Ub - \frac{1}{\rho}\nabla\pi - \frac{\rho-\rho_0}{\rho}g\eb_r, \label{eq:u_update}
-\end{equation}
-%
+Using equation (\ref{eq:momentum}), 
 we compute time-centered edge velocities, $\uadvonedag$, using
@@
 \begin{equation}
-\nabla \cdot \left(\beta_0^n \uadvone\right) = \beta_0^n \left(S^{\nph,\star} - \frac{1}{\overline{\Gamma}_1^np_0^n}\psi^{n-\myhalf} + \chi^n \right),
+\nabla \cdot \left(\beta_0^n \uadvone\right) = \beta_0^n \left(S^{\nph,\star} - \frac{1}{\overline{\Gamma}_1^np_0^n}\psi^{n-\myhalf} \right),
 \end{equation}
@@
  \beta_0^n = \beta_0 \left(\rho_0^n, p_0^n, \overline{\Gamma}_1^n\right),
-\end{equation}
-\begin{equation}
-\chi^n = \frac{f}{\overline{\Gamma}_1^n p_0^n}\frac{p_{\rm eos}^n(\rho,h,X_k) - p_0^n}{\Delta t^n}
 \end{equation}
@@
 
-Call {\bf React State}$[ \rho^{(2),\star},(\rho h)^{(2),\star}, X_k^{(2),\star}, T^{(2),\star}, p_0^{n+1,\star}]$\\
-\phantom{ }\hfill $\rightarrow [ \rho^{n+1,\star},(\rho h)^{n+1,\star}, X_k^{n+1,\star}, T^{n+1,\star}, (\rho \omegadot_k)^{(2),\star}, (\rho \Hnuc)^{(2),\star} ].$
+Call {\bf React State}$[ (\rho X_k)^{(2),\star}, (\rho h)^{(2),\star}, p_0^{n+1,\star}]$\\
+\phantom{ }\hfill $\rightarrow [ (\rho X_k)^{n+1,\star}, (\rho h)^{n+1,\star}, (\rho \omegadot_k)^{(2),\star}, (\rho \Hnuc)^{(2),\star} ].$
 
@@
 \nabla \cdot \left(\beta_0^{\nph} \uadvtwo\right) =
-\beta_0^{\nph} \left(S^{\nph} - \frac{1}{\overline{\Gamma}_1^{n+\myhalf}p_0^{n+1,\star}}\frac{p_0^{n+1,\star}-p_0^n}{\Delta t^n} + \chi^{n+\myhalf}\right),
+\beta_0^{\nph} \left(S^{\nph} - \frac{1}{\overline{\Gamma}_1^{n+\myhalf}p_0^{n+1,\star}}\frac{p_0^{n+1,\star}-p_0^n}{\Delta t^n}\right),
 \end{equation}
@@
 \overline{\Gamma}_1^{\nph} = \frac{ \overline{\Gamma}_1^n +  \overline{\Gamma}_1^{n+1,\star} }{2}.
-\end{equation}
-\begin{equation}
-\chi^{n+\myhalf} = \chi^n + \frac{f}{\overline{\Gamma}_1^{n+1,\star} p_0^{n+1,\star}}\frac{p_{\rm eos}^{n+1,\star}(\rho,h,X_k) - p_0^{n+1,\star}}{\Delta t^n}
 \end{equation}
@@
 
-Call {\bf React State}$[\rho^{(2)},(\rho h)^{(2)}, X_k^{(2)},T^{(2)}, p_0^{n+1}] \rightarrow [\rho^{n+1}, (\rho h)^{n+1}, X_k^{n+1}, T^{n+1}, (\rho \omegadot_k)^{(2)}, (\rho \Hnuc)^{(2)} ].$
+Call {\bf React State}$[(\rho X_k)^{(2)},(\rho h)^{(2)}, p_0^{n+1}] \rightarrow [(\rho X_k)^{n+1}, (\rho h)^{n+1}, (\rho \omegadot_k)^{(2)}, (\rho \Hnuc)^{(2)}].$
 
@@
 We update the velocity field $\Ub^n$ to $\Ub^{n+1,\dagger}$ by discretizing
-equation (\ref{eq:u_update}) as
+equation (\ref{eq:momentum}) as
 \begin{eqnarray}

### 2019-02-12 MAESTROeX_I/paper.tex
stuff

@@
 %% keywords and the rules for their use.
-\keywords{insert keywords here}
+\keywords{convection, hydrodynamics, methods: numerical, nuclear reactions, nucleosynthesis, abundances, supernovae: general}
 
@@
 
-Many astrophysical flows are highly subsonic; 
-sound waves carry sufficiently little energy that they do not significantly affect the convective dynamics of the system.
-In many of these flows, long-time convective dynamics are of interest.
-Thus, using a traditional compressible astrophysical solver is highly inefficient.
-One approach to this problem is to use low Mach number hydrodynamic models.
-In a low Mach number approach, sound waves are eliminated from the governing equations by requiring that the
-thermodynamic variables evolve subject to a specified ambient pressure that varies only as a function of altitude and time.
+Many astrophysical flows are highly subsonic; sound waves carry sufficiently little energy that they do not significantly affect the convective dynamics of the system.
+In many of these flows, modeling long-time convective dynamics are of interest, and numerical approaches based on compressible hydrodynamics are intractable, even on modern supercomputers.
+One approach to this problem is to use low Mach number models.
+In a low Mach number approach, sound waves are eliminated from the governing equations, and the resulting model can be numerically integrated with much larger time steps than a compressible model.
+
+Previously, we have developed the low Mach number astrophysical solver, MAESTRO.
+MAESTRO is unique in two ways.
+First, the low Mach number model is specifically derived for astrophysical settings.
+Second, MAESTRO incorporates the use of a stratified background (or base) state density and pressure held in hydrostatic equilibrium, and also vary as a function of altitude and time.
+MAESTRO is suitable for for spherical stars, as well as planar simulations of local dynamics.
+
+by requiring that the thermodynamic variables evolve subject to a specified ambient pressure that varies only as a function of altitude and time.
 In the low Mach equations, the equation of state is recast as a divergence constraint on velocity, requiring an
@@
 
-The key numerical developments of the MAESTRO algorithm are presented in a series of papers which we refer to as papers I-V (\cite{MAESTRO_I,MAESTRO_II,MAESTRO_III,MAESTRO_IV,MAESTRO_V}, respectively).
-Applications include the convective phase preceding Chandrasekhar mass models for type Ia supernovae \cite{MAESTRO_convection,MAESTRO_AMR,MAESTRO_CASTRO}, 
-convection in massivve stars \cite{Gilet:2013},
-sub-Chandrasekhar white dwarfs \cite{subChandra_I,subChandra_II}, and
-Type I X-ray bursts \cite{XRB_I,XRB_II,XRB_III}.
+(overview of MAESTRO numerics) finite volume, AMR. include definition of planar and spherical.
 
-\subsection{Motivation}
+The key numerical developments of the original MAESTRO algorithm are presented in a series of papers which we refer to as Papers I-V:
+\begin{itemize}
+\item In Paper I \citep{MAESTRO_I}, we derive the low Mach number equation set from the fully compressible equations.
+\item In Paper II \citep{MAESTRO_II}, we incorporate the effects atmospheric expansion through the use of a time-dependent background state.
+\item In Paper III \citep{MAESTRO_III}, we incorporate reactions and the associated coupling to the hydrodynamics.
+\item In Paper IV \citep{MAESTRO_IV}, we describe our treatment of spherical stars in a three-dimensional Cartesian geometry.
+\item In Paper V \citep{MAESTRO_V}, we describe the use of block-structured adaptive mesh refinement to focus spatial resolution in regions of interest.
+\end{itemize}
+
+Since then, there have been many scientific investigations using MAESTRO, which include additional algorithmic enhancements.  Topics include:
+\begin{itemize}
+\item The convective phase preceding Chandrasekhar mass models for type Ia supernovae \citep{MAESTRO_convection,MAESTRO_AMR,MAESTRO_CASTRO}.
+\item Convection in massive stars \citep{Gilet:2013}.
+\item Sub-Chandrasekhar white dwarfs \citep{subChandra_I,subChandra_II}.
+In the latter paper we introduced an optional modification to the momentum equation of velocity constraint that conserves total energy, can give results closer to compressible codes of convective dynamics, including low density regions at the surface of a star.
+\item Type I X-ray bursts \citep{XRB_I,XRB_II,XRB_III}.
+In \cite{XRB_I} we discuss the modifications required for implicit thermal conduction, as well as an additional ``volume discrepancy'' modification to the velocity constraint to force the species and enthalpy to evolve in a manner consistent with the background pressure.
+\end{itemize}
+
 We would like to revamp our previous MAESTRO code by moving to the AMReX framework
@@
 \subsection{Main Algorithm Description}\label{Sec:Main Algorithm Description}
+\MarginPar{talk about cutoff densities}
 We now describe the proposed algorithm, making frequent use of the

### 2019-02-12 MAESTROeX_I/paper.tex
references

@@
 science: rotation, solar physics, MHD
-algorithm: SDC, high-order
-implementation: GPUs
+
+algorithm: SDC \cite{dutt2000spectral}, high-order, multi-implicit \cite{bourlioux2003high}.
+Consider SDC ideas used for terrestrial combustion \cite{pazner2016high,nonaka2018conservative}
+
+implementation: GPUs.  Forward cite next CASTRO paper with GPU strategy.
 

### 2019-02-12 MAESTROeX_I/paper.tex
add official AMReX citation

@@
 
-We would like to revamp our previous MAESTRO code by moving to the AMReX framework
+We would like to revamp our previous MAESTRO code by moving to the AMReX \cite{AMReX} framework
 so that we can take advantage of all of the latest features of the AMReX libraries. 

### 2019-02-12 MAESTROeX_I/paper.tex
updates

--- 2019-02-12 MAESTROeX_I/paper.tex
@@
 We present MAESTROeX, a massively parallel solver for low Mach number astrophysical flows.
-This new solver leverages the algorithmic capabilities of the previous MAESTRO code
-while taking advantage of the next-generation AMReX software framework block-structured
-adaptive mesh refinement applications.
-Using our previous studies on white dwarf convection as a benchmark, we characterize
-the performance of our new code.
-We have also developed new temporal integration schemes that have a much simpler treatment
-of the base state, while allowing for more straightforward extensions to state-of-the
-art, high-order multiphysics coupling schemes.
-Also, we present an improved spatial coupling scheme between the hydrostatic base
-state and the Cartesian grid full-state.
+Our model equations allow for long-time integration for highly subsonic flows compared to compressible approaches.
+The code leverages the new AMReX framework for block-structured adaptive mesh refinement calculations, and also provides additional algorithmic features from the original MAESTRO code.
+We present a new temporal integration option that is much simpler than our previous approach while retaining the same order order of accuracy.
+We also provide a more accurate spatial discretization for mapping the base state onto the full state Cartesian grid.
+Using our previous studies on the convective phase of single-degenerate progenitor models of Type Ia supernovae, we characterize the performance of the code and validate the new algorithmic features.
 \end{abstract}
@@
 One approach to this problem is to use low Mach number models.
-In a low Mach number approach, sound waves are eliminated from the governing equations, and the resulting model can be numerically integrated with much larger time steps than a compressible model.
+In a low Mach number approach, sound waves are eliminated from the governing equations while retaining compressibilitiy effects due to, e.g., nuclear energy release, compositional changes, and thermal diffusion.
+The resulting model can be numerically integrated with much larger time steps than a compressible model.
 
 Previously, we have developed the low Mach number astrophysical solver, MAESTRO.
-MAESTRO is unique in two ways.
-First, the low Mach number model is specifically derived for astrophysical settings.
-Second, MAESTRO incorporates the use of a stratified background (or base) state density and pressure held in hydrostatic equilibrium, and also vary as a function of altitude and time.
+Low Mach number models have been developed for a variety of contexts including combustion \citep{day2000numerical}, atmospherics \citep{duarte2015low}, and elastic solids \citep{abbate2017all}.
+The low Mach number model in MAESTRO is unique in that it is specifically designed for astrophysical settings with significant atmospheric stratification.
+Central to the algorithm is a stratified background (or base) state density and pressure held in hydrostatic equilibrium, and also vary as a function of altitude and time.
 MAESTRO is suitable for for spherical stars, as well as planar simulations of local dynamics.
-
-by requiring that the thermodynamic variables evolve subject to a specified ambient pressure that varies only as a function of altitude and time.
-In the low Mach equations, the equation of state is recast as a divergence constraint on velocity, requiring an
-elliptic solve for pressure in the spirit of projection methods for incompressible flow.
-
-(overview of MAESTRO numerics) finite volume, AMR. include definition of planar and spherical.
 
@@
 
-We would like to revamp our previous MAESTRO code by moving to the AMReX framework
-so that we can take advantage of all of the latest features of the AMReX libraries. 
-In addition, we want to modify and simplify the existing numerical scheme 
-to both encourage new users and to develop a method that will lend itself more 
-easily to incorporating SDC and other high-order methods. 
-Irregular base state spacing is considered in order to improve spatial resolution 
-that have been seen in the previous MAESTRO code in some test cases.
+In this paper, we characterize the performance of the MAESTROeX algorithm compared to the previous implementation.
+We also develop simpler, alternative temporal integrators, with a longer term goal of developing high-order coupling schemes.
+We also present a new treatment of the base state mapping to Cartesian grids for spherical problsm.
 
@@
 
+\subsection{Convergence Test}
+3D reacting bubble rise; to demonstrate 2nd-order convergence of the new temporal algorithm.
+
 \subsection{Scaling}
@@
 
-\section{Test Problems}
+\subsection{White Dwarf Convection}
 white dwarf convection runs with 3 algorithms (original, new temporal, new temporal + irregular base state)

### 2019-02-13 MAESTROeX_I/paper.tex
simplifying notation... trying to make algorithm more human-readable

@@
 \newcommand{\etarho}{\eta_\rho}
-\newcommand{\gammaonebar}{\overline{\Gamma_1}}
+\newcommand{\gammaonebar}{\overline{\Gamma}_1}
 \newcommand{\Hnuc}{H_{\rm nuc}}
@@
 
-Call {\bf React State}$[ (\rho X_k)^{(2),\star}, (\rho h)^{(2),\star}, p_0^{n+1,\star}]$\\
-\phantom{ }\hfill $\rightarrow [ (\rho X_k)^{n+1,\star}, (\rho h)^{n+1,\star}, (\rho \omegadot_k)^{(2),\star}, (\rho \Hnuc)^{(2),\star} ].$
+Call {\bf React State}$[ (\rho X_k)^{(2),\star}, (\rho h)^{(2),\star}, p_0^{n+1,\star}] 
+\rightarrow 
+[ (\rho X_k)^{n+1,\star}, (\rho h)^{n+1,\star}, (\rho \omegadot_k)^{n+1,\star}, (\rho \Hnuc)^{n+1,\star} ].$
 
@@
 
-\item First, compute $S^{n+1,\star}$ with
+First, compute $S^{n+1,\star}$ with
 \begin{equation}
-S^{n+1,\star} =  -\sigma  \sum_k  \xi_k  (\omegadot_k)^{(2),\star}  + \frac{1}{\rho^{n+1,\star} p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2),\star} + \sigma \Hnuc^{(2),\star},
+S^{n+1,\star} =  \left(-\sigma  \sum_k  \xi_k  \omegadot_k  + \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k + \sigma \Hnuc\right)^{n+1,\star}.
 \end{equation}
-  where $(\omegadot_k)^{(2),\star} = (\rho \omegadot_k)^{(2),\star} /
-  \rho^{(2),\star}$ and the thermodynamic quantities are defined using
-  $\rho^{n+1,\star}, X_k^{n+1,\star},$ and $T^{n+1,\star}$ as inputs to
-  the equation of state.  Then, define
+  Then, define
 \begin{equation}
@@
 
-Call {\bf React State}$[(\rho X_k)^{(2)},(\rho h)^{(2)}, p_0^{n+1}] \rightarrow [(\rho X_k)^{n+1}, (\rho h)^{n+1}, (\rho \omegadot_k)^{(2)}, (\rho \Hnuc)^{(2)}].$
+Call {\bf React State}$[(\rho X_k)^{(2)},(\rho h)^{(2)}, p_0^{n+1}] \rightarrow [(\rho X_k)^{n+1}, (\rho h)^{n+1}, (\rho \omegadot_k)^{n+1}, (\rho \Hnuc)^{n+1}].$
 
@@
 \begin{equation}
-  S^{n+1} =  -\sigma  \sum_k  \xi_k (\omegadot_k)^{(2)}  + \sigma \Hnuc^{(2)} +
-  \frac{1}{\rho^{n+1} p_\rho} \sum_k p_{X_k}  ({\omegadot}_k)^{(2)},
+  S^{n+1} =  \left(-\sigma  \sum_k  \xi_k \omegadot_k  + \sigma \Hnuc +
+  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  \omegadot_k\right)^{n+1}.
 \end{equation}
-where $(\omegadot_k)^{(2)} = (\rho \omegadot_k)^{(2)} / \rho^{(2)}$
-and the thermodynamic quantities are defined using $\rho^{n+1}$,
-$X_k^{n+1}$, and $T^{n+1}$ as inputs to the equation of state.
 

### 2019-02-13 MAESTROeX_I/paper.tex
major simplifications to algorithm description... trying to make this human-readable again

@@
 \subsection{Main Algorithm Description}\label{Sec:Main Algorithm Description}
+New algorithm...
+overframework is similar but now more splitting $w_0$ dynamics...
+eliminates the need to explicitly evolve the base state and compute the average expansion...
+instead we can just use a predictor-corrector treatment of the base state while retaining the same formal order of accuracy...
+
+We now describe the new temporal integration scheme, which has been greatly simplified
+from Paper V without formal loss of order of accuracy.
+Note that this temporal integration scheme is valid for the original base state mapping (with constant base state grid spacing), or the new irregularly spaced base state mapping.
+
+At the beginning of each time step we have the state,
+$(\Ub,\rho X_k,\rho h,\rho_0,p_0)^n$, and $\pi^{n-\myhalf}$.
+At any time, the associated density, composition, and enthalpy can be trivially computed using, e.g.,
+\begin{equation}
+\rho^n = \sum_k(\rho X_k)^n, \quad
+X_k^n = (\rho X_k)^n / \rho^n, \quad
+h^n = (\rho h)^n / \rho^n.
+\end{equation}
+Temperature is computed using the equation of state, e.g.,
+\footnote{As described in Paper V, for planar problems we compute temperature using $h$ instead of $p_0$, since we have successfully developed planar volume discrepancy schemes to effectively couple the enthalpy to the rest of the solution.
+We are still exploring this option for spherical stars.}
+\begin{equation}
+T^n = T(\rho,p_0,X_k)^n.
+\end{equation}
+Similarly, $\gammaonebar$ and $\beta_0$ are computed using $\rho,p_0$, and $X_k$ (see Appendix A of Paper I and Appendix C of Paper III for details).
+
+The overal flow of the algorithm is a second-order Strang Splitting approach for the splitting of the reaction and advection steps.
+Furthermore, we use a predictor-corrector approach within this splitting scheme to achieve second-order accurate treatment of the constraing equation.
+To summarize:
+\begin{itemize}
+\item In {\bf Step 1} we react the thermodynamic variables over $\Delta t/2$.
+\item In {\bf Steps 2-4} we advect the thermodynamic variables over $\Delta t$.
+\item In {\bf Step 5} we react the thermodynamic variables over $\Delta t/2$.
+\item In {\bf Steps 6-8} we redo the advection in {\bf Steps 2-4} but are able to use the trapezoidal rule to time-center certain quantities such as $S$, $\rho_0$, etc.
+\item In {\bf Step 9} we redo the reactions from {\bf Step 5} using the improved results for the corrector advection step.
+\item In {\bf Steps 10-11} we update the velocity.
+\end{itemize}
+
+Each time step satisfies the standard advective CFL condition,
+\begin{equation}
+\Delta t = \sigma^{\rm CFL} \min_i(\Delta x / U_i),
+\end{equation}
+where for our simulations we typically use $\sigma^{\rm CFL}\sim 0.5$.
+There are additional constraints on the time involving the buoyancy force (commonly used when the velocity is approximately zero at the start of some simulations) and the divergence constraint (to prevent too much mass evacuation from a cell in a time step); see Section 3.4 in Paper III for details.
+
 \MarginPar{talk about cutoff densities}
+
+\sout{
 We now describe the proposed algorithm, making frequent use of the
@@
 denoted with an ``$n+1$'' superscript.  We use Strang-splitting to
-achieve second-order accuracy in time.
+achieve second-order accuracy in time.}
 
@@
 \item[Step 0] {\em Initialization}
+\sout{
 The initialization step only occurs at the beginning of the simulation.
@@
 $\psi^{-\myhalf}$, $S^0$, and $S^1$ for use in time stepping routine.  
-We pass through {\bf Step 1} -- {\bf Step 12}, but only keep these values.
+We pass through {\bf Step 1} -- {\bf Step 11}, but only keep these values.
 The initial values for $\Ub^0, \rho^0, (\rho h)^0, X_k^0, T^0,
-\rho_0^0, p_0^0$, and $\overline{\Gamma}_1^0$ are specified from the problem-dependent
+\rho_0^0, p_0^0$, and $\gammaonebar^0$ are specified from the problem-dependent
 initial conditions.  The initial time step, $\dt^0$, is computed as in
 Paper III.
+}
 
@@
 \begin{equation}
-\nabla \cdot \left(\beta_0^n \uadvone\right) = \beta_0^n \left(S^{\nph,\star} - \frac{1}{\overline{\Gamma}_1^np_0^n}\psi^{n-\myhalf} \right),
+\nabla \cdot \left(\beta_0^n \uadvone\right) = \beta_0^n \left(S^{\nph,\star} - \frac{1}{\gammaonebar^np_0^n}\psi^{n-\myhalf} \right).
 \end{equation}
-\begin{equation}
- \beta_0^n = \beta_0 \left(\rho_0^n, p_0^n, \overline{\Gamma}_1^n\right),
-\end{equation}
-where $\beta_0$ is computed as described in Appendix C of Paper III.
 
@@
   - \dt \left\{ \nabla \cdot \left[ \uadvone (\rho X_k)^{\nph,\pred} \right] \right\},
-\end{equation}
-\begin{equation}
-\rho^{(2),\star} = \sum_k (\rho X_k)^{(2),\star},
-\qquad
-X_k^{(2),\star} = (\rho X_k)^{(2),\star} / \rho^{(2),\star}.
 \end{equation}
@@
 
-\item Update the temperature using the equation of state:
-$T^{(2),\star} =
-  T(\rho^{(2),\star}, h^{(2),\star}, X_k^{(2),\star})$ (planar geometry) or
-$T^{(2),\star} =
-  T(\rho^{(2),\star}, p_0^{n+1,\star}, X_k^{(2),\star})$ (spherical geometry).
 \end{enumerate}
@@
 \nabla \cdot \left(\beta_0^{\nph} \uadvtwo\right) =
-\beta_0^{\nph} \left(S^{\nph} - \frac{1}{\overline{\Gamma}_1^{n+\myhalf}p_0^{n+1,\star}}\frac{p_0^{n+1,\star}-p_0^n}{\Delta t^n}\right),
+\beta_0^{\nph} \left(S^{\nph} - \frac{1}{\gammaonebar^{n+\myhalf}p_0^{n+1,\star}}\frac{p_0^{n+1,\star}-p_0^n}{\Delta t^n}\right),
 \end{equation}
+where
 \begin{equation}
-\beta_0^{\nph} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2};
+\beta_0^{\nph} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2},
 \qquad
- \beta_0^{n+1,\star} = \beta_0 \left(\rho_0^{n+1,\star}, p_0^{n+1,\star}, \overline{\Gamma}_1^{n+1,\star}\right),
-\end{equation}
-\begin{equation}
-\overline{\Gamma}_1^{\nph} = \frac{ \overline{\Gamma}_1^n +  \overline{\Gamma}_1^{n+1,\star} }{2}.
+\gammaonebar^{\nph} = \frac{ \gammaonebar^n +  \gammaonebar^{n+1,\star} }{2}.
 \end{equation}
@@
 \item Update $(\rho X_k)$.  This step is identical to {\bf Step 4A} except we use
-  the updated values $\uadvtwo$ and $\rho_0^{(2a)}$ rather than
-  $\uadvone$ and $\rho_0^{(2a),\star}$.  In particular:
+  the updated values $\uadvtwo$ and $\rho_0^{n+1,\star}$ rather than
+  $\uadvone$ and $\rho_0^{n+1,\star}$.  In particular:
 
@@
 
-\item Update the temperature using the equation of state:
-$T^{(2)} =
-   T(\rho^{(2)}, h^{(2)}, X_k^{(2)})$ (planar geometry) or
-$T^{(2)} =
-   T(\rho^{(2)}, p_0^{n+1}, X_k^{(2)})$ (spherical geometry).
-
 \item Define a radial cell-centered $\etarho^{\nph}$.
@@
 %--------------------------------------------------------------------------
-\item[Step 10] {\em Define the new time expansion, $S^{n+1}$, and $\overline{\Gamma}_1^{n+1}$.}
+\item[Step 10] {\em Define the new time expansion, $S^{n+1}$.}
 
@@
   \frac{1}{\rho p_\rho} \sum_k p_{X_k}  \omegadot_k\right)^{n+1}.
-\end{equation}
-
-\item Define
-\begin{equation}
-\overline{\Gamma}_1^{n+1} = {\rm{\bf Avg}}\left[\Gamma_1\left(\rho^{n+1}, p_0^{n+1},
-X_k^{n+1}\right) \right].
 \end{equation}
@@
 \nabla \cdot \left(\beta_0^{n+1} \Ub^{n+1} \right)
-= \beta_0^{n+1} S^{n+1},
-\end{equation}
-where
-\begin{equation}
-\beta_0^{n+1} = \beta \left(\rho_0^{n+1}, p_0^{n+1}, \overline{\Gamma}_1^{n+1}, g^{n+1}\right).
+= \beta_0^{n+1} S^{n+1}.
 \end{equation}
@@
 and gradient operators.  Details of the approximate projection are given in Paper III.
-
-%--------------------------------------------------------------------------
-% STEP 12
-%--------------------------------------------------------------------------
-\item[Step 12] {\em Compute a new $\dt.$}
-
-Compute $\dt$ for the next time step with a similar procedure described in
-\S 3.4 of Paper III. However, we are using $\Ub^{n+1}$
-as computed in {\bf Step 11} instead of $w_0$ and $\Ubt^{n+1}$ as in Paper V.
-The following constraints need to be satisfied.
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \Alph{enumi}}}
-\item Use standard CFL condition for explicit methods using CFL factor $\sigma^{\mathrm{CFL}}$.
-  For a calculation in $n_{\mathrm{dim}}$ dimensions ($n_{\mathrm{dim}}=2$ or 3),
-  the first constraint is
-\begin{equation}
-\dt_{\Ub} = \sigma^\mathrm{CFL}  \min_{i=1\ldots n_\mathrm{dim}} \left \{ \dt_i \right \}\enskip,
-\end{equation}
-  where
-\begin{equation}
-\dt_i = \min_{\bf x}  \left \{ \frac{\Delta x_i}{\left | \Ub \right |}  \right \}\enskip.
-\end{equation}
-
-\item Use the forcing terms rather than the velocities.
-  This constraint is necessary when a calculation is started from rest,
-  since in that case the velocity-based time step would be infinite. Define
-\begin{equation}
-\dt_{\rm F} = \min_{i=1\ldots n_\mathrm{dim}} \left \{ \dt_{\rm F_i} \right \}\enskip,
-\end{equation}
-where
-\begin{equation}
-\dt_{\rm F_i} = \sqrt{2 \Delta x_i / {F_i}_\mathrm{max}}\enskip,
-\end{equation}
-where ${F_i}_\mathrm{max}$ is the maximum buoyancy force in the
-$i^\mathrm{th}$ coordinate direction.
-
-\item To prevent local expansion from numerically emptying a cell,
-  we require that the density be reduced by no more than 40\% in a single time step.
-  This constraint is expressed as $\dt \leq \dt_S$ where
-\begin{equation}
- \Delta t_S \; (\nabla \cdot \Ub)  \leq 0.40 \enskip ,
-\end{equation}
-  in every cell.
-
-\end{enumerate}
 

### 2019-02-14 MAESTROeX_I/paper.tex
minor

@@
 \subsection{Main Algorithm Description}\label{Sec:Main Algorithm Description}
-New algorithm...
-overframework is similar but now more splitting $w_0$ dynamics...
-eliminates the need to explicitly evolve the base state and compute the average expansion...
-instead we can just use a predictor-corrector treatment of the base state while retaining the same formal order of accuracy...
+Our new temporal integration scheme is much simpler than the scheme from Paper V.
+The primary reason is that we no longer split the velocity field into a perturbational and base state components.  
+Previously, we used the base state velocity to predict the expansion of the base state.  
+In this new scheme we use a simpler, predictor-corrector approach to the base state density and pressure that no longer requires the complex algorithm from Paper V, yet still retains the same overall second-order accuracy.
 
@@
 \item In {\bf Steps 2-4} we advect the thermodynamic variables over $\Delta t$.
-\item In {\bf Step 5} we react the thermodynamic variables over $\Delta t/2$.
+Specifically, we compute an estimate for the expansion term, $S$, project the velocities so they satisfy the divergence constraint, and then advect the thermodynamic variables.
+\item In {\bf Step 5} we react the thermodynamic variables over $\Delta t/2$.\footnote{Note that we could skip to the velocity advance in {\bf Steps 10-11}, however the overall scheme would be only first-order in time, so {\bf Steps 6-9} can be thought of as a trapezoidal corrector step.}
 \item In {\bf Steps 6-8} we redo the advection in {\bf Steps 2-4} but are able to use the trapezoidal rule to time-center certain quantities such as $S$, $\rho_0$, etc.
@@
 
-\MarginPar{talk about cutoff densities}
+At the beginning of each simulation, we define $(\Ub,\rho X_k,\rho h)$...
 
-\sout{
-We now describe the proposed algorithm, making frequent use of the
-shorthand notations mentioned in Paper V, Appendix A3.  In summary, in the predictor step ({\bf
-  Steps 2-5}) we use an estimate of the expansion term, $S$, to
-compute a preliminary solution at the new time level, denoted with an
-``$n+1,\star$'' superscript.  In the corrector step ({\bf Steps 6-9}),
-we use the results from the predictor step to compute a more accurate
-expansion term, and compute the final solution at the new time level,
-denoted with an ``$n+1$'' superscript.  We use Strang-splitting to
-achieve second-order accuracy in time.}
+\MarginPar{mention cutoff densities}
 
@@
 \begin{description}
-
-%--------------------------------------------------------------------------
-% STEP 0
-%--------------------------------------------------------------------------
-
-\item[Step 0] {\em Initialization}
-\sout{
-The initialization step only occurs at the beginning of the simulation.
-The initialization step computes initial values for $\pi^{-\myhalf}$, , $\eta_{\rho}^{-\myhalf}$, 
-$\psi^{-\myhalf}$, $S^0$, and $S^1$ for use in time stepping routine.  
-We pass through {\bf Step 1} -- {\bf Step 11}, but only keep these values.
-The initial values for $\Ub^0, \rho^0, (\rho h)^0, X_k^0, T^0,
-\rho_0^0, p_0^0$, and $\gammaonebar^0$ are specified from the problem-dependent
-initial conditions.  The initial time step, $\dt^0$, is computed as in
-Paper III.
-}
 
@@
 
-\noindent This completes one step of the algorithm.
+This completes one step of the algorithm.
+
+To initialize the simulation
+
+%--------------------------------------------------------------------------
+% STEP 0
+%--------------------------------------------------------------------------
+{\em Initialization}
+\sout{
+The initialization step only occurs at the beginning of the simulation.
+The initialization step computes initial values for $\pi^{-\myhalf}$, , $\eta_{\rho}^{-\myhalf}$, 
+$\psi^{-\myhalf}$, $S^0$, and $S^1$ for use in time stepping routine.  
+We pass through {\bf Step 1} -- {\bf Step 11}, but only keep these values.
+The initial values for $\Ub^0, \rho^0, (\rho h)^0, X_k^0, T^0,
+\rho_0^0, p_0^0$, and $\gammaonebar^0$ are specified from the problem-dependent
+initial conditions.  The initial time step, $\dt^0$, is computed as in
+Paper III.
+}
 

### 2019-02-14 gpus/paper.tex
add ref
add paper
add a code
fix figure
add an author list

@@
 
-\author{Castro developers / M.~P.~Katz et al.}
-\affil{}
-\email{}
+\correspondingauthor{Michael Zingale}
+
+\collaboration{the Castro Development Team}
+
+\author[0000-0003-2103-312X]{Ann S. Almgren}
+\affil{Center for Computational Sciences and Engineering, Lawrence Berkeley National Laboratory}
+
+\author[0000-0002-3185-9809]{Maria Barrios Sazo}
+\affil{Department of Physics and Astronomy, Stony Brook University}
+
+\author{Kevin Gott}
+\affil{National Energy Research Scientific Computing Center}
+
+\author[0000-0002-1530-781X]{Alice Harpole}
+\affil{Department of Physics and Astronomy, Stony Brook University}
+
+\author[0000-0003-0439-4556]{Max P.~Katz}
+\affil{NVIDIA Corporation}
+
+\author[0000-0003-2300-5165]{Donald Willcox}
+\affil{Center for Computational Sciences and Engineering, Lawrence Berkeley National Laboratory}
+
+\author[0000-0001-8092-1974]{Weiqun Zhang}
+\affil{Center for Computational Sciences and Engineering, Lawrence Berkeley National Laboratory}
+
+\author[0000-0001-8401-030X]{Michael Zingale}
+\affil{Department of Physics and Astronomy, Stony Brook University}
+\email{michael.zingale@stonybrook.edu}
 
@@
 Existing GPU-enabled astrophysical hydro codes include
-\citet{gamer,cholla,fargo3d,pekkila:2017}
+\citet{gamer,cholla,fargo3d,pekkila:2017,caplan:2018,racz:2018,liska:2018,goz:2018}
 
@@
 
+\section{The Castro hydrodynamics code}
+
+Disucss changes since last paper
+
+Discuss open development model
+
 \section{Performance portable compute kernels}
@@
 
-\begin{figure}
+\begin{figure*}
 \centering
-\includegraphics[height=0.25\textheight]{gpu_1} \\
-\includegraphics[height=0.25\textheight]{gpu_2} \\
-\includegraphics[height=0.25\textheight]{gpu_3}
+\includegraphics[width=0.3\textwidth]{gpu_1} 
+\includegraphics[width=0.3\textwidth]{gpu_2} 
+\includegraphics[width=0.3\textwidth]{gpu_3}
 \caption{\label{fig:loops} caption}
-\end{figure}
+\end{figure*}
 

### 2019-02-14 gpus/paper.tex

@@
 
-is a compressible (radiation-)hydrodynamics simulation code built on
+\castro\ is a compressible (radiation-)hydrodynamics simulation code built on
 the AMReX adaptive mesh refinement framework \citep{castro,castro2,castro3}.  It 
 includes full self-gravity using multigrid methods with isolated boundaries
-\cite{katz:2016}, works with general equations of state \citep{zingalekatz}
-and arbitrary nuclear reaction networks.
+\citep{katz:2016}, works with general equations of state \citep{zingalekatz}
+and arbitrary nuclear reaction networks.  Castro has been used for science investigations of
+...
 
-Existing GPU-enabled astrophysical hydro codes include
+The original computing architectures targeted were
+
+With GPU accelerators becoming increasingly dominant, algorithms need to be ported.
+There are a number of descriptions of GPU-enabled astrophysical hydro codes, including
 \citet{gamer,cholla,fargo3d,pekkila:2017,caplan:2018,racz:2018,liska:2018,goz:2018}
@@
 
+In this paper, we describe our approach to performance portability.
+Our goal is to have the same computational kernels run efficiently on
+CPUs (including manycore processors) and GPUs.  We describe the work
+we needed to do to enable this portability and show performance
+metrics for a reacting flow problem in astrophysics.
+
 \section{The Castro hydrodynamics code}
@@
 
-Discuss open development model
+\castro\ is developed using a fully open-development model, with the
+code managed in git and hosted on
+github\footnote{\url{https://github.com/amrex-astro/Castro/}}.
+Contributions from the community are welcomed via github issues and
+pull requests.  For the science projects that the core development
+team are working on, all problem setup, input files, auxillary data,
+and anything else needed for the science is stored in the main github
+repo, allowing for the community to contribute to the science and
+reproduce results as they are published.
 

### 2019-02-15 gpus/paper.tex
minor

@@
 
-Discuss frameworks vs.\ doing it outselves
-
-Discuss OpenMP, OpenACC, and CUDA.
-
 In this paper, we describe our approach to performance portability.
@@
 \section{Performance portable compute kernels}
+
+\subsection{Offloading to GPUs}
+
+CUDA Fortran is not as well supported as CUDA C/C++.  In particular,
+CUDA Fortran does not automatically create both the host and device
+kernel, necessitating one to do this themselves.  Usually this means
+having two copies of a source routine, one marked up with device
+attributes.  Keeping them in sync essentially requires that this
+action be scripted, but in our experience, this scripting is fragile
+because of the need to deal with Fortran syntax, preprocessor
+directives, ...  Thus the pure CUDA route for the compute kernels makes the most sense
+with C/C++ kernels, and not Fortran kernels.
+
+OpenACC
+
+OpenMP
+
+Discuss frameworks vs.\ doing it outselves
+
+\subsection{Example of advancing a grid}
 
@@
 
-\software{MPICH, GCC, Castro, python, matplotlib}
+\software{MPICH, GCC, Castro, AMReX, python, matplotlib}
+
+\facilities{OLCF, NERSC}
 
@@
 Energy under Contract No. DE-AC05-00OR22725.  We thank NVIDIA Corporation
-for the donation of a Titan X Pascal used in this research.
+for the donation of a Titan X Pascal and Titan V used in this research.
 

### 2019-03-09 MAESTROeX_I/paper.tex
updates/simplifications.  need to talk about projection splitting still

@@
 
-\correspondingauthor{MAESTROeX}
-\email{DFan@lbl.gov}
+\correspondingauthor{Doreen Fan; DFan@lbl.gov}
 
@@
 
-Previously we adopted an approach where we split the velocity into a base state component, $w_0$, and a local velocity $\Ubt$, so that
+Previously we adopted an approach where we split the velocity into a base state component, $w_0(r,t)$, 
+and a local velocity $\Ubt(\xb,t)$, so that
 \begin{equation}
-\Ub = w_0(r,t)\eb_r + \Ubt(\xb,t).
+\Ub = \Ubt(\xb,t) + w_0(r,t)\eb_r,
 \end{equation}
@@
 \item Steps 4 and 8 will be greatly simplified, leading to a more efficient algorithm.
-\item There is no longer a need for $w_0$, so we can use the full velocity $\Ub$ instead of $\Ubt$.
-As a consequence, $\eta$ and $\psi$ also do not need to be computed.
-This has the added advantage of saving some memory storage by removing many of the terms needed when we decided to split terms into its base state and perturbation state in our previous implementation.
+\item \sout{There is no longer a need for $w_0$, so we can use the full velocity $\Ub$ instead of $\Ubt$.
+As a consequence, $\eta$ does not need to be computed.
+This has the added advantage of saving some memory storage by removing many of the terms needed when we decided to split terms into its base state and perturbation state in our previous implementation.} We split the velocity for the projection to handle the cutoff.
 \item By placing the base states exactly at cell-centers, we eliminate interpolation errors and simplify the existing code.
@@
 
-At the beginning of each time step we have the state,
-$(\Ub,\rho X_k,\rho h,\rho_0,p_0)^n$, and $\pi^{n-\myhalf}$.
+At the beginning of each time step we have the cell-centered state,
+$(\Ub,\rho X_k,\rho h,\rho_0,p_0)^n$, and nodal state, $\pi^{n-\myhalf}$.
 At any time, the associated density, composition, and enthalpy can be trivially computed using, e.g.,
@@
 \begin{equation}
-T^n = T(\rho,p_0,X_k)^n.
+T = T(\rho,p_0,X_k),
 \end{equation}
-Similarly, $\gammaonebar$ and $\beta_0$ are computed using $\rho,p_0$, and $X_k$ (see Appendix A of Paper I and Appendix C of Paper III for details).
+and ($\gammaonebar,\beta_0)$ are computed from $(\rho,p_0,X_k)$ (see Appendix A of Paper I and Appendix C of Paper III for details).
 
@@
 \begin{equation}
-\nabla \cdot \left(\beta_0^n \uadvone\right) = \beta_0^n \left(S^{\nph,\star} - \frac{1}{\gammaonebar^np_0^n}\psi^{n-\myhalf} \right).
+\nabla \cdot \left(\beta_0^n \uadvone\right) = \beta_0^n \left(S^{\nph,\star} - \frac{1}{\gammaonebar^np_0^n}\frac{\partial p_0}{\partial t} \right).
 \end{equation}
-
-It should also be noted that we have found it better to approximate 
-the temporal changes in base pressure as 
-\begin{equation}
-\frac{\partial p_0}{\partial t} \approx \psi = \eta_{\rho} g
-\end{equation}
-instead of using the lagged pressure changes from previous time $t^{n-1}$ 
-in the predictor step.
+In practice we solve for.
 
@@
 &&- \dt^n \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \dt^n \left(\uadvone \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^{n} \nonumber \\
-&& + \dt^n \psi^{n-\myhalf},
+&& + \dt^n \left(\frac{p_0^n - p_0^{n-1}}{\dt^{n-1}}\right) \nonumber \\
+&& + \dt^n \left(\uadvone \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^{n}
 \end{eqnarray}
@@
 &&- \dt^n \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \dt^n \left \{ \nabla \cdot \left (\uadvone p_0^{n} \right ) - p_0^{n} \nabla \cdot \uadvone \right \} \nonumber \\
-&& + \dt^n \psi^{n-\myhalf},
+&& + \dt^n \left(\frac{p_0^n - p_0^{n-1}}{\dt^{n-1}}\right) \nonumber \\
+&& + \dt^n \left \{ \nabla \cdot \left (\uadvone p_0^{n} \right ) - p_0^{n} \nabla \cdot \uadvone \right \}
 \end{eqnarray}
@@
 the updated value $S^{\nph}$ rather than $S^{\nph,\star}$.
-Also, the temporal changes of $p_0$ are now computed directly instead of 
-approximated using $\psi$.
 We note that $\uadvtwo$ satisfies the divergence constraint
@@
 \nabla \cdot \left(\beta_0^{\nph} \uadvtwo\right) =
-\beta_0^{\nph} \left(S^{\nph} - \frac{1}{\gammaonebar^{n+\myhalf}p_0^{n+1,\star}}\frac{p_0^{n+1,\star}-p_0^n}{\Delta t^n}\right),
+\beta_0^{\nph} \left(S^{\nph} - \frac{1}{\gammaonebar^{n+\myhalf}p_0^{n+\myhalf}}\frac{\partial p_0}{\partial t}\right),
 \end{equation}
@@
 \gammaonebar^{\nph} = \frac{ \gammaonebar^n +  \gammaonebar^{n+1,\star} }{2}.
+\qquad
+p_0^{\nph} = \frac{ p_0^n + p_0^{n+1,\star} }{2},
 \end{equation}
@@
 \end{equation}
-\begin{equation}
-\rho^{(2)} = \sum_k (\rho X_k)^{(2)},
-\qquad
-X_k^{(2)} = (\rho X_k)^{(2)} / \rho^{(2)}.
-\end{equation}
 
@@
   $\uadvone, \rho_0^{n+1,\star}, (\rho h)_0^{n+1,\star}$, and $p_0^n$.
-  Also, the temporal changes of $p_0$ are now computed directly instead of 
-  approximated using $\psi$.
   In particular:
@@
 &=& (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvtwo (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \dt \left(\uadvtwo \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^\nph \nonumber \\
-&& + \dt \left( \frac{p_0^{n+1} - p_0^n}{\dt^n} \right),
+&& + \dt \left( \frac{p_0^{n+1} - p_0^n}{\dt} \right)\nonumber \\
+&& + \dt \left(\uadvtwo \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^\nph,
 \end{eqnarray}
@@
 &=& (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvtwo (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \dt \left[ \nabla \cdot \left (\uadvtwo p_0^{\nph} \right ) - p_0^{\nph} \nabla \cdot \uadvtwo \right] \nonumber \\
-&& + \dt \left( \frac{p_0^{n+1} - p_0^n}{\dt^n} \right),
+&& + \dt \left( \frac{p_0^{n+1} - p_0^n}{\dt} \right)\nonumber \\
+&& + \dt \left[ \nabla \cdot \left (\uadvtwo p_0^{\nph} \right ) - p_0^{\nph} \nabla \cdot \uadvtwo \right],
 \end{eqnarray}
@@
 \end{description}
-
-\item Compute $\psi^{\nph}$: 
-\begin{equation}
-\psi_j^{\nph} = \eta_{\rho,j}^{\nph} g.
-\end{equation}
 
@@
 equation (\ref{eq:momentum}) as
-\begin{eqnarray}
+\begin{equation}
 \Ub^{n+1,\dagger}
-&=& \Ub^n - \dt \left[\uadvtwo \cdot \nabla \Ub^{\nph,\pred} \right] \nonumber \\
-&& - \dt \left[ \frac{1}{\rho^\nph} \mathbf{G} \pi^\nmh + \frac{\left(\rho^\nph-\rho_0^\nph\right)}{\rho^\nph} g^{\nph} \eb_r \right],\nonumber \\
-\end{eqnarray}
-where $\mathbf{G}$ approximates a cell-centered gradient from nodal
-data.  Again, the $\dagger$ superscript refers
+= \Ub^n - \dt \left[\uadvtwo \cdot \nabla \Ub^{\nph,\pred} \right]
+ - \dt \left[ \frac{1}{\rho^\nph} \nabla \pi^\nmh + \frac{\left(\rho^\nph-\rho_0^\nph\right)}{\rho^\nph} g^{\nph} \eb_r \right],
+\end{equation}
+Again, the $\dagger$ superscript refers
 to the fact that the updated velocity does not satisfy the divergence
@@
 \begin{equation}
-\nabla \cdot \left(\beta_0^{n+1} \Ub^{n+1} \right)
-= \beta_0^{n+1} S^{n+1}.
+\nabla \cdot \left(\beta_0^{n+1} \Ub^{n+1} \right) = \beta_0^{n+1} \left( S^{n+1} - \frac{1}{\gammaonebar^{n+1}p_0^{n+1}}\frac{\partial p_0}{\partial t} \right).
 \end{equation}
@@
 The initialization step only occurs at the beginning of the simulation.
-The initialization step computes initial values for $\pi^{-\myhalf}$, , $\eta_{\rho}^{-\myhalf}$, 
-$\psi^{-\myhalf}$, $S^0$, and $S^1$ for use in time stepping routine.  
+The initialization step computes initial values for 
+$\pi^{-\myhalf}, \eta_{\rho}^{-\myhalf}, S^0$, and $S^1$ for use in time stepping routine.  
 We pass through {\bf Step 1} -- {\bf Step 11}, but only keep these values.

### 2019-03-09 MAESTROeX_I/paper.tex
small updates

@@
 
-The overal flow of the algorithm is a second-order Strang Splitting approach for the splitting of the reaction and advection steps.
-Furthermore, we use a predictor-corrector approach within this splitting scheme to achieve second-order accurate treatment of the constraing equation.
+The overall flow of the algorithm is a second-order Strang splitting approach for the coupling of the reactions and advection of the thermodynamic variables.  
+We use a predictor-corrector approach within the Strang splitting scheme to achieve second-order accuracy in time.
+We integrate the velocity using a standard second-order projection method to enforce the divergence constraint.
 To summarize:
@@
 \begin{equation}
-S^{\nph,\star} = S^n + \frac{\dt^n}{2} \frac{S^n - S^{n-1}}{\dt^{n-1}}.
+S^{\nph,\star} = S^n + \frac{\dt^n}{2} \left(\frac{S^n - S^{n-1}}{\dt^{n-1}}\right).
 \end{equation}

### 2019-03-11 MAESTROeX_I/paper.tex
new spherical make_w0 idea

@@
 
+\subsection{Computing $w_0$ For Spherical Problems}
+Recall that we want to solve
+\begin{equation}
+\frac{1}{r^2}\frac{\partial}{\partial r}(r^2 w_0) = \Sbar - \frac{1}{\gammaonebar p_0}\left(\frac{\partial p_0}{\partial t} + w_0\frac{\partial p_0}{\partial r}\right)
+\end{equation}
+First, note that $\partial p_0/\partial r = \rho_0 g$.
+The key observation here is that we can compute $\partial p_0/\partial t$ directly instead of making complicated substitutions involving $\eta_\rho$ to eliminate it.
+Let's discretize this as follows:
+\begin{equation}
+\frac{1}{r_j^2}\left[\left(r^2 w_0\right)_{j+\myhalf} - \left(r^2 w_0\right)_{j-\myhalf}\right] = 
+\Sbar_j - \frac{1}{\left(\gammaonebar p_0\right)_j}\left[\left(\frac{\partial p_0}{\partial t}\right)_j + \left(\frac{w_{0,j-\myhalf}+w_{0,j+\myhalf}}{2}\right)(\rho_0 g)_j\right]
+\end{equation}
+Rearranging this to solve for $w_{0,j+\myhalf}$ gives us an explicit update:
+\begin{equation}
+\left[\frac{r_{j+\myhalf}^2}{r_j^2} - \frac{(\rho_0 g)_j}{2\left(\gammaonebar p_0\right)_j}\right] w_{0,j+\myhalf} =
+\left[\frac{r_{j-\myhalf}^2}{r_j^2} + \frac{(\rho_0 g)_j}{2\left(\gammaonebar p_0\right)_j}\right] w_{0,j-\myhalf}
++ \Sbar_j - \frac{1}{\left(\gammaonebar p_0\right)_j}\left(\frac{\partial p_0}{\partial t}\right)_j
+\end{equation}
+The interface to the routine in the algorithm description is
+{\bf Make $\mathbf{w_0}$}$[\rho_0,p_0,\partial p_0/\partial t,\gammaonebar,\Sbar] \rightarrow [w_0]$
+
 \section{Test Problems}

### 2019-03-11 MAESTROeX_I/paper.tex
make w0 spherical

@@
 \end{equation}
+As before, once $\rho_0$ falls below $\rho_{\rm cutoff}$, we hold $r^2 w_0$ constant.\\ \\
 The interface to the routine in the algorithm description is

### 2019-03-12 MAESTROeX_I/paper.tex
sign error
another sign error
add a TODO

@@
 \section{Introduction} \label{sec:intro}
+
+TODO list:
+\begin{itemize}
+\item Weak scaling tests of original algorithm, comparing MAESTRO to MAESTROeX.  Use cori knl, 4 MPI per node, 16 threads per MPI
+\item Run at effective $512^3$ resolution (maybe to ignition?...
+need to see how max T trends go) (1) classic MAESTRO, (2) MAESTROeX with original algorithm, (3) MAESTROeX with new temporal integrator, (4) MAESTROeX with new temporal integrator and irregular dr, (5) MAESTROeX with original algorithm and 3-levels of AMR, (6) MAESTROeX with new temporal integrator and 3 levels of AMR.
+\item (perhaps) a demonstration of what goes wrong when you don't split the velocity dynamics in the projection
+\end{itemize}
 
@@
 \end{equation}
-First, note that $\partial p_0/\partial r = \rho_0 g$.
+First, note that $\partial p_0/\partial r = -\rho_0 g$.
 The key observation here is that we can compute $\partial p_0/\partial t$ directly instead of making complicated substitutions involving $\eta_\rho$ to eliminate it.
@@
 \frac{1}{r_j^2}\left[\left(r^2 w_0\right)_{j+\myhalf} - \left(r^2 w_0\right)_{j-\myhalf}\right] = 
-\Sbar_j - \frac{1}{\left(\gammaonebar p_0\right)_j}\left[\left(\frac{\partial p_0}{\partial t}\right)_j + \left(\frac{w_{0,j-\myhalf}+w_{0,j+\myhalf}}{2}\right)(\rho_0 g)_j\right]
+\Sbar_j - \frac{1}{\left(\gammaonebar p_0\right)_j}\left[\left(\frac{\partial p_0}{\partial t}\right)_j - \left(\frac{w_{0,j-\myhalf}+w_{0,j+\myhalf}}{2}\right)(\rho_0 g)_j\right]
 \end{equation}

### 2019-03-12 MAESTROeX_I/paper.tex
added some advection references; moved stuff to appendix B (notes that won't make the final paper) also testing email notification to maestro-development

@@
 
-Using equation (\ref{eq:momentum}), 
+The construction of face-centered time-centered states used to discretize the
+advection terms for velocity, species, and enthalpy, are performed using
+standard multidimensional corner transport upwind 
+\cite{colella1990multidimensional,saltzman1994unsplit} with piecewise-parabolic
+one-dimensional tracing \cite{colella1984piecewise}.  In this step,
+using equation (\ref{eq:momentum}), 
 we compute time-centered edge velocities, $\uadvonedag$, using
@@
 
-\item Define a radial cell-centered $\etarho^{\nph}$.
-
-\begin{description}
-\item For planar geometry, $\etarho = \overline{\rho'(\Ub\cdot\eb_r)}$,
-\begin{equation}
- \etarho^{\nph} =  {\rm {\bf Avg}} \sum_k \left[ \left(\uadvtwo \cdot \eb_r \right) (\rho X_k)^{\nph,\pred} \right]
-\end{equation}
-\item For spherical geometry, first construct 
-$\etarho^{{\rm cart},\nph} = [\rho'(\Ub\cdot\eb_r)]^{\nph}$ on Cartesian cell centers using:
-\begin{equation}
-\etarho^{{\rm cart},\nph} = \left[\left(\frac{\rho^{(1)}+\rho^{(2)}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{n+1}}{2}\right)\right] \cdot \left( \uadvtwo \cdot \eb_r \right).
-\end{equation}
-Then,
-\begin{equation}
-\etarho^{\nph,\star} = {\rm {\bf Avg}}\left(\etarho^{{\rm cart},\nph}\right).
-\end{equation}
-\end{description}
-
 \end{enumerate}
@@
 
-\subsection{Computing $w_0$ For Spherical Problems}
-Recall that we want to solve
-\begin{equation}
-\frac{1}{r^2}\frac{\partial}{\partial r}(r^2 w_0) = \Sbar - \frac{1}{\gammaonebar p_0}\left(\frac{\partial p_0}{\partial t} + w_0\frac{\partial p_0}{\partial r}\right)
-\end{equation}
-First, note that $\partial p_0/\partial r = -\rho_0 g$.
-The key observation here is that we can compute $\partial p_0/\partial t$ directly instead of making complicated substitutions involving $\eta_\rho$ to eliminate it.
-Let's discretize this as follows:
-\begin{equation}
-\frac{1}{r_j^2}\left[\left(r^2 w_0\right)_{j+\myhalf} - \left(r^2 w_0\right)_{j-\myhalf}\right] = 
-\Sbar_j - \frac{1}{\left(\gammaonebar p_0\right)_j}\left[\left(\frac{\partial p_0}{\partial t}\right)_j - \left(\frac{w_{0,j-\myhalf}+w_{0,j+\myhalf}}{2}\right)(\rho_0 g)_j\right]
-\end{equation}
-Rearranging this to solve for $w_{0,j+\myhalf}$ gives us an explicit update:
-\begin{equation}
-\left[\frac{r_{j+\myhalf}^2}{r_j^2} - \frac{(\rho_0 g)_j}{2\left(\gammaonebar p_0\right)_j}\right] w_{0,j+\myhalf} =
-\left[\frac{r_{j-\myhalf}^2}{r_j^2} + \frac{(\rho_0 g)_j}{2\left(\gammaonebar p_0\right)_j}\right] w_{0,j-\myhalf}
-+ \Sbar_j - \frac{1}{\left(\gammaonebar p_0\right)_j}\left(\frac{\partial p_0}{\partial t}\right)_j
-\end{equation}
-As before, once $\rho_0$ falls below $\rho_{\rm cutoff}$, we hold $r^2 w_0$ constant.\\ \\
-The interface to the routine in the algorithm description is
-{\bf Make $\mathbf{w_0}$}$[\rho_0,p_0,\partial p_0/\partial t,\gammaonebar,\Sbar] \rightarrow [w_0]$
-
 \section{Test Problems}
@@
 
+\appendix
+\section{Projection Details}
+
+After {\bf Step 8D}, we define a radial cell-centered $\etarho^{\nph}$.
+
+\begin{description}
+\item For planar geometry, $\etarho = \overline{\rho'(\Ub\cdot\eb_r)}$,
+\begin{equation}
+ \etarho^{\nph} =  {\rm {\bf Avg}} \sum_k \left[ \left(\uadvtwo \cdot \eb_r \right) (\rho X_k)^{\nph,\pred} \right]
+\end{equation}
+\item For spherical geometry, first construct 
+$\etarho^{{\rm cart},\nph} = [\rho'(\Ub\cdot\eb_r)]^{\nph}$ on Cartesian cell centers using:
+\begin{equation}
+\etarho^{{\rm cart},\nph} = \left[\left(\frac{\rho^{(1)}+\rho^{(2)}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{n+1}}{2}\right)\right] \cdot \left( \uadvtwo \cdot \eb_r \right).
+\end{equation}
+Then,
+\begin{equation}
+\etarho^{\nph,\star} = {\rm {\bf Avg}}\left(\etarho^{{\rm cart},\nph}\right).
+\end{equation}
+\end{description}
+
+\section{Old Notes}
+
+\section{Computing $w_0$ For Spherical Problems}
+Recall that we want to solve
+\begin{equation}
+\frac{1}{r^2}\frac{\partial}{\partial r}(r^2 w_0) = \Sbar - \frac{1}{\gammaonebar p_0}\left(\frac{\partial p_0}{\partial t} + w_0\frac{\partial p_0}{\partial r}\right)
+\end{equation}
+First, note that $\partial p_0/\partial r = -\rho_0 g$.
+The key observation here is that we can compute $\partial p_0/\partial t$ directly instead of making complicated substitutions involving $\eta_\rho$ to eliminate it.
+Let's discretize this as follows:
+\begin{equation}
+\frac{1}{r_j^2}\left[\left(r^2 w_0\right)_{j+\myhalf} - \left(r^2 w_0\right)_{j-\myhalf}\right] = 
+\Sbar_j - \frac{1}{\left(\gammaonebar p_0\right)_j}\left[\left(\frac{\partial p_0}{\partial t}\right)_j - \left(\frac{w_{0,j-\myhalf}+w_{0,j+\myhalf}}{2}\right)(\rho_0 g)_j\right]
+\end{equation}
+Rearranging this to solve for $w_{0,j+\myhalf}$ gives us an explicit update:
+\begin{equation}
+\left[\frac{r_{j+\myhalf}^2}{r_j^2} - \frac{(\rho_0 g)_j}{2\left(\gammaonebar p_0\right)_j}\right] w_{0,j+\myhalf} =
+\left[\frac{r_{j-\myhalf}^2}{r_j^2} + \frac{(\rho_0 g)_j}{2\left(\gammaonebar p_0\right)_j}\right] w_{0,j-\myhalf}
++ \Sbar_j - \frac{1}{\left(\gammaonebar p_0\right)_j}\left(\frac{\partial p_0}{\partial t}\right)_j
+\end{equation}
+As before, once $\rho_0$ falls below $\rho_{\rm cutoff}$, we hold $r^2 w_0$ constant.\\ \\
+The interface to the routine in the algorithm description is
+{\bf Make $\mathbf{w_0}$}$[\rho_0,p_0,\partial p_0/\partial t,\gammaonebar,\Sbar] \rightarrow [w_0]$
+
 \acknowledgements

### 2019-03-12 MAESTROeX_I/paper.tex
point to XRB_III for Godunov details

@@
 advection terms for velocity, species, and enthalpy, are performed using
-standard multidimensional corner transport upwind 
-\cite{colella1990multidimensional,saltzman1994unsplit} with piecewise-parabolic
-one-dimensional tracing \cite{colella1984piecewise}.  In this step,
-using equation (\ref{eq:momentum}), 
+a standard multidimensional corner transport upwind approach
+\citep{colella1990multidimensional,saltzman1994unsplit} with piecewise-parabolic
+one-dimensional tracing \citep{colella1984piecewise}.  The full details of this
+Godunov advection approach for all steps in this algorithm are described 
+in Appendix A of \cite{XRB_III}.
+
+In this step, using equation (\ref{eq:momentum}), 
 we compute time-centered edge velocities, $\uadvonedag$, using
@@
 \item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\pred}$,
-  for the conservative update of $(\rho X_k)^{(1)}$ using
-  $\Ub = \uadvone$ and $\rho_0^n$.
+  for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
 \MarginPar{We may not need $\rho_0$ if we predict $\rho X$ or $\rho$ and $X$.
@@
   for the conservative update of $(\rho h)^{(1)}$
-  using $\Ub = \uadvone$ and $(\rho h)^n$,
+  using using a Godunov approach \citep{XRB_III}.
 
@@
 
-The procedure to construct $\uadvtwodag$ is identical to the procedure
+The procedure to construct $\uadvtwodag$ is identical to the Godunov procedure
 for computing $\uadvonedag$ in {\bf Step 3}, but uses
@@
 \item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\pred}$,
-  for the conservative update of $(\rho X_k)^{(1)}$ using $\Ub = \uadvtwo$ and
-  $\rho_0^n$.
+  for the conservative update of $(\rho X_k)^{(1)}$ usingusing a Godunov approach \citep{XRB_III}.
 
@@
 \item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\pred},$
-  for the conservative update of $(\rho h)^{(1)}$.
+  for the conservative update of $(\rho h)^{(1)}$ using a Godunov approach \citep{XRB_III}..
 
@@
 
-First, we compute the time-centered edge velocities, $\Ub^{\nph,\pred}$.
-Then, we define
+First, we compute the time-centered edge velocities, $\Ub^{\nph,\pred}$
+using a Godunov approach \citep{XRB_III}. Then, we define
 \begin{equation}

### 2019-03-13 MAESTROeX_I/paper.tex
simplified description of Enforce HSE, updated input arguments to only take in rho_0 (since p_0 was only used to get the pressure outside the star, may as well just simply state that)

@@
 
+There are a few key numerical modules we use in each time step.
+\begin{itemize}
+\item {\bf Enforce HSE} computes the base state pressure, $p_0$, from a base state density, $\rho_0$ by integrating the hydrostatic equilibrium condition in 1D (see equation (A10) in \cite{MAESTRO_V}).
+The base state pressure remains equal to a constant value from the location of a prescribed cutoff density outward for the entire simulation.
+\item {\bf React State}
+\end{itemize}
+
 Each time step satisfies the standard advective CFL condition,
@@
 \item Update $p_0$ by calling
-{\bf Enforce HSE}$[p_0^n,\rho_0^{n+1,\star}] \rightarrow [p_0^{n+1,\star}]$.
+{\bf Enforce HSE}$[\rho_0^{n+1,\star}] \rightarrow [p_0^{n+1,\star}]$.
 
@@
 \item Update $p_0$ by calling
-{\bf Enforce HSE}$[p_0^n,\rho_0^{n+1}] \rightarrow [p_0^{n+1}]$.
+{\bf Enforce HSE}$[\rho_0^{n+1}] \rightarrow [p_0^{n+1}]$.
 

### 2019-03-13 MAESTROeX_I/paper.tex
description of React State

@@
 The base state pressure remains equal to a constant value from the location of a prescribed cutoff density outward for the entire simulation.
-\item {\bf React State}
+\item {\bf React State} advances the species and enthalpy due to reactions over $\Delta t/2$ by integrating
+\begin{equation}
+\frac{dX_k}{dt} = \dot\omega_k(\rho,X_k,T); \qquad
+\frac{dT}{dt} = \frac{1}{c_p}\left(-\sum_k\xi_k\dot\omega_k + H_{\rm nuc}\right)
+\end{equation}
+The inputs are the species, enthalpy, and base state pressure, and the outpus are the species, enthalpy, reaction rates, and nuclear energy generation rate.
+See \cite{MAESTRO_III,MAESTRO_V} for details.
 \end{itemize}
@@
 Call {\bf React State}$[(\rho X_k)^n, (\rho h)^n, p_0^n] \rightarrow [(\rho X_k)^{(1)}, (\rho h)^{(1)}, (\rho \omegadot_k)^{(1)}, (\rho \Hnuc)^{(1)}]$.
-\MarginPar{AJN: I think we can simplify the description here... this is somewhat confusing now.}
 
@@
   for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
-\MarginPar{We may not need $\rho_0$ if we predict $\rho X$ or $\rho$ and $X$.
-AJN: If we end up keeping prediction of $\rho'$ and $X$, and $(\rho h)'$ we need to summarize this, particularly since $\rho_0$ and $(\rho h)_0$ treatment differs from previous papers.}
+\MarginPar{mention $\rho'$ and $X$ advection; $(\rho h)'$ advection}
 
@@
 \item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\pred}$,
-  for the conservative update of $(\rho X_k)^{(1)}$ usingusing a Godunov approach \citep{XRB_III}.
+  for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
 
@@
   In particular:
-\MarginPar{Again, we may not need $\rho_0$ and $(\rho h)_0$.}
 

### 2019-03-13 MAESTROeX_I/paper.tex
first cut at etarho and split projection discussion

@@
 \begin{equation}
-\nabla \cdot \left(\beta_0^n \uadvone\right) = \beta_0^n \left(S^{\nph,\star} - \frac{1}{\gammaonebar^np_0^n}\frac{\partial p_0}{\partial t} \right).
+\nabla \cdot \left(\beta_0^n \uadvone\right) = \beta_0^n \left[S^{\nph,\star} - \frac{1}{\gammaonebar^np_0^n}\left(\frac{\partial p_0}{\partial t}\right)^{n-\myhalf} \right].\label{eq:div1}
 \end{equation}
@@
 \nabla \cdot \left(\beta_0^{\nph} \uadvtwo\right) =
-\beta_0^{\nph} \left(S^{\nph} - \frac{1}{\gammaonebar^{n+\myhalf}p_0^{n+\myhalf}}\frac{\partial p_0}{\partial t}\right),
+\beta_0^{\nph} \left[S^{\nph} - \frac{1}{\gammaonebar^{n+\myhalf}p_0^{n+\myhalf}}\left(\frac{\partial p_0}{\partial t}\right)^{n+\myhalf}\right],\label{eq:div2}
 \end{equation}
@@
 from $\Ub^{n+1,\dagger},$  such that $\Ub^{n+1}$ approximately
-satisfies
+satisfies the divergence constraint,
 \begin{equation}
-\nabla \cdot \left(\beta_0^{n+1} \Ub^{n+1} \right) = \beta_0^{n+1} \left( S^{n+1} - \frac{1}{\gammaonebar^{n+1}p_0^{n+1}}\frac{\partial p_0}{\partial t} \right).
+\nabla \cdot \left(\beta_0^{n+1} \Ub^{n+1} \right) = \beta_0^{n+1} \left[ S^{n+1} - \frac{1}{\gammaonebar^{n+1}p_0^{n+1}}\left(\frac{\partial p_0}{\partial t}\right)^{n+\myhalf}\right].\label{eq:div3}
 \end{equation}
@@
 \section{Projection Details}
+To enforce the divergence constraint in {\bf Step 3, 7}, and {\bf 11}, we use a projection method analgous to the methods originally developed for incompressible flow.
 
-After {\bf Step 8D}, we define a radial cell-centered $\etarho^{\nph}$.
+Based on our past experience in the MAESTRO project, we have found it useful to split the velocity dynamics into a perturbational and base state velocity,
+\begin{equation}
+\Ub = \Ubt(\xb,t) + w_0(r,t)\eb_r,
+\end{equation}
+solve for each term separately, and immediately combine them to find a full velocity that satisfies the constraint.
+We take that approach here, primarily because it allows us to enforce a boundary condition on $w_0$ at the edge of the star (i.e., the cutoff density location where we hold density constant).
+Namely, to enforce that $r^2 w_0$ remain constant is difficult to do when solving for the full velocity.
+So in practice, we solve the constraint over the lateral average,
+\begin{equation}
+\nabla\cdot\left(\beta_0\Ubt\right) = \beta_0(S-\overline{S})
+\end{equation}
+and sepratately solve for $w_0$ using 
+\begin{equation}
+\nabla\cdot(\beta_0w_0\eb_r) = \beta_0\left(\overline{S} - \frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial t}\right).
+\end{equation}
+To find $w_0$ we integrate in 1D using the procedure in Appendix B of \cite{MAESTRO_V}.
 
-\begin{description}
-\item For planar geometry, $\etarho = \overline{\rho'(\Ub\cdot\eb_r)}$,
-\begin{equation}
- \etarho^{\nph} =  {\rm {\bf Avg}} \sum_k \left[ \left(\uadvtwo \cdot \eb_r \right) (\rho X_k)^{\nph,\pred} \right]
-\end{equation}
-\item For spherical geometry, first construct 
-$\etarho^{{\rm cart},\nph} = [\rho'(\Ub\cdot\eb_r)]^{\nph}$ on Cartesian cell centers using:
-\begin{equation}
-\etarho^{{\rm cart},\nph} = \left[\left(\frac{\rho^{(1)}+\rho^{(2)}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{n+1}}{2}\right)\right] \cdot \left( \uadvtwo \cdot \eb_r \right).
-\end{equation}
-Then,
-\begin{equation}
-\etarho^{\nph,\star} = {\rm {\bf Avg}}\left(\etarho^{{\rm cart},\nph}\right).
-\end{equation}
-\end{description}
+Note that in this approach, we estimated the time-derivative of the pressure in part by examining how laterally averaged $\rho'=\rho-\rho_0$ changed over time (quantified by $\eta_\rho = \overline{\rho'\Ub\cdot\eb_r}$.
+After evolving the species, we compute this term in the same way as reported in \cite{MAESTRO_V} for use in computing $w_0$.
 
@@
 
+After {\bf Step 8D}, we define a radial cell-centered $\etarho^{\nph}$.
+
+\begin{description}
+\item For planar geometry, $\etarho = \overline{\rho'(\Ub\cdot\eb_r)}$,
+\begin{equation}
+ \etarho^{\nph} =  {\rm {\bf Avg}} \sum_k \left[ \left(\uadvtwo \cdot \eb_r \right) (\rho X_k)^{\nph,\pred} \right]
+\end{equation}
+\item For spherical geometry, first construct 
+$\etarho^{{\rm cart},\nph} = [\rho'(\Ub\cdot\eb_r)]^{\nph}$ on Cartesian cell centers using:
+\begin{equation}
+\etarho^{{\rm cart},\nph} = \left[\left(\frac{\rho^{(1)}+\rho^{(2)}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{n+1}}{2}\right)\right] \cdot \left( \uadvtwo \cdot \eb_r \right).
+\end{equation}
+Then,
+\begin{equation}
+\etarho^{\nph,\star} = {\rm {\bf Avg}}\left(\etarho^{{\rm cart},\nph}\right).
+\end{equation}
+\end{description}
+
 \acknowledgements

### 2019-03-13 MAESTROeX_I/paper.tex
clean up alg description more

@@
 &=& (\rho h)^{(1)} \nonumber \\
-&&- \dt^n \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \dt^n \left(\frac{p_0^n - p_0^{n-1}}{\dt^{n-1}}\right) \nonumber \\
-&& + \dt^n \left(\uadvone \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^{n}
+&&- \dt \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
+&& + \left(p_0^{n+1,*} - p_0^n\right) \nonumber \\
+&& + \dt \left(\uadvone \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^{n}
 \end{eqnarray}
@@
 &=& (\rho h)^{(1)} \nonumber \\
-&&- \dt^n \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \dt^n \left(\frac{p_0^n - p_0^{n-1}}{\dt^{n-1}}\right) \nonumber \\
-&& + \dt^n \left \{ \nabla \cdot \left (\uadvone p_0^{n} \right ) - p_0^{n} \nabla \cdot \uadvone \right \}
+&&- \dt \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
+&& + \left(p_0^{n+1,*} - p_0^n\right) \nonumber \\
+&& + \dt \left \{ \nabla \cdot \left (\uadvone p_0^{n+\myhalf} \right ) - p_0^{n+\myhalf} \nabla \cdot \uadvone \right \}
 \end{eqnarray}
 \end{description}
+where we define $p_0^\nph$ as $p_0^\nph = (p_0^n+p_0^{n+1,*})/2$.
 
@@
 \gammaonebar^{\nph} = \frac{ \gammaonebar^n +  \gammaonebar^{n+1,\star} }{2}.
-\qquad
-p_0^{\nph} = \frac{ p_0^n + p_0^{n+1,\star} }{2},
 \end{equation}
@@
 &=& (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvtwo (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \dt \left( \frac{p_0^{n+1} - p_0^n}{\dt} \right)\nonumber \\
+&& + \left( p_0^{n+1} - p_0^n \right)\nonumber \\
 && + \dt \left(\uadvtwo \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^\nph,
@@
 &=& (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvtwo (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \dt \left( \frac{p_0^{n+1} - p_0^n}{\dt} \right)\nonumber \\
+&& + \left(p_0^{n+1} - p_0^n\right)\nonumber \\
 && + \dt \left[ \nabla \cdot \left (\uadvtwo p_0^{\nph} \right ) - p_0^{\nph} \nabla \cdot \uadvtwo \right],
@@
 \end{description}
-where $p_0^\nph$ is defined as $p_0^\nph = (p_0^n+p_0^{n+1})/2$.
+where we now define $p_0^\nph$ as $p_0^\nph = (p_0^n+p_0^{n+1})/2$.
 
@@
 
-\section{Old Notes}
+\section{Old Notes (not for inclusion in paper)}
 
-\section{Computing $w_0$ For Spherical Problems}
+\subsection{$\eta_\rho$ Notes}
+After {\bf Step 8D}, we define a radial cell-centered $\etarho^{\nph}$.
+
+\begin{description}
+\item For planar geometry, $\etarho = \overline{\rho'(\Ub\cdot\eb_r)}$,
+\begin{equation}
+ \etarho^{\nph} =  {\rm {\bf Avg}} \sum_k \left[ \left(\uadvtwo \cdot \eb_r \right) (\rho X_k)^{\nph,\pred} \right]
+\end{equation}
+\item For spherical geometry, first construct 
+$\etarho^{{\rm cart},\nph} = [\rho'(\Ub\cdot\eb_r)]^{\nph}$ on Cartesian cell centers using:
+\begin{equation}
+\etarho^{{\rm cart},\nph} = \left[\left(\frac{\rho^{(1)}+\rho^{(2)}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{n+1}}{2}\right)\right] \cdot \left( \uadvtwo \cdot \eb_r \right).
+\end{equation}
+Then,
+\begin{equation}
+\etarho^{\nph,\star} = {\rm {\bf Avg}}\left(\etarho^{{\rm cart},\nph}\right).
+\end{equation}
+\end{description}
+
+\subsection{Computing $w_0$ For Spherical Problems}
 Recall that we want to solve
@@
 
-After {\bf Step 8D}, we define a radial cell-centered $\etarho^{\nph}$.
-
-\begin{description}
-\item For planar geometry, $\etarho = \overline{\rho'(\Ub\cdot\eb_r)}$,
-\begin{equation}
- \etarho^{\nph} =  {\rm {\bf Avg}} \sum_k \left[ \left(\uadvtwo \cdot \eb_r \right) (\rho X_k)^{\nph,\pred} \right]
-\end{equation}
-\item For spherical geometry, first construct 
-$\etarho^{{\rm cart},\nph} = [\rho'(\Ub\cdot\eb_r)]^{\nph}$ on Cartesian cell centers using:
-\begin{equation}
-\etarho^{{\rm cart},\nph} = \left[\left(\frac{\rho^{(1)}+\rho^{(2)}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{n+1}}{2}\right)\right] \cdot \left( \uadvtwo \cdot \eb_r \right).
-\end{equation}
-Then,
-\begin{equation}
-\etarho^{\nph,\star} = {\rm {\bf Avg}}\left(\etarho^{{\rm cart},\nph}\right).
-\end{equation}
-\end{description}
-
 \acknowledgements

### 2019-03-13 MAESTROeX_I/paper.tex
more re-org

@@
 We take the standard equations of reacting, compressible flow, and recast the equation
-of state as a divergence constraint on the velocity field.  This is done by taking
+of state (EOS) as a divergence constraint on the velocity field.  This is done by taking
 the Langrangian derivative of the equation of state for pressure as a function of the
@@
 
-\section{Using Exact Spherical Base States on Irregular Grid} \label{sec:exactbasestate}
+\section{Numerical Algorithm}
 
-\subsection{Potential Problems}
-\begin{enumerate}
-\item How do we define \verb|r_edge_loc| for irregularly-spaced base states whose cell-centered values are defined at $r_m=\Delta x\sqrt{\frac{3}{4} + 2m}$?
-\item We would stil need to approximate the values at edges when advecting the base states (i.e.
-$\rho_0^{\nph,\pred}$).
-This may affect the nice properties we would get with exact cell-centered base state values.
-\item Excessive memory storage may be required (\verb|nr_irreg >> nr_fine|).
-\end{enumerate}
+\subsection{Spatial Discretization}
 
-\subsection{A Possible Solution}
-We could choose to no longer advect the base states for both predictor and corrector steps.
-This would solve the problems with computing state values at cell edges.
-Instead, we propose to only update the base states $\rho_0$ and $p_0$ by computing the average of the full state values after advecting the full states.
-This would result in the following changes in the algorithm from Paper V:
-%
-\begin{enumerate}
-\item Steps 4 and 8 will be greatly simplified, leading to a more efficient algorithm.
-\item \sout{There is no longer a need for $w_0$, so we can use the full velocity $\Ub$ instead of $\Ubt$.
-As a consequence, $\eta$ does not need to be computed.
-This has the added advantage of saving some memory storage by removing many of the terms needed when we decided to split terms into its base state and perturbation state in our previous implementation.} We split the velocity for the projection to handle the cutoff.
-\item By placing the base states exactly at cell-centers, we eliminate interpolation errors and simplify the existing code.
-\end{enumerate}
-%
+(description of base state mapping in planar, and origianl spherical)
 
-\subsection{Main Algorithm Description}\label{Sec:Main Algorithm Description}
+\subsection{Exact Base State Mapping} \label{sec:exactbasestate}
+
+\subsection{Temporal Integration Scheme}\label{Sec:Temporal Integration Scheme}
 Our new temporal integration scheme is much simpler than the scheme from Paper V.
@@
 \begin{itemize}
-\item {\bf Enforce HSE} computes the base state pressure, $p_0$, from a base state density, $\rho_0$ by integrating the hydrostatic equilibrium condition in 1D (see equation (A10) in \cite{MAESTRO_V}).
+\item {\bf Average} computes...
+\item {\bf Enforce HSE} computes the base state pressure, $p_0$, from a base state density, $\rho_0$ by integrating the hydrostatic equilibrium condition in 1D (see equation (A10) in Paper V).
 The base state pressure remains equal to a constant value from the location of a prescribed cutoff density outward for the entire simulation.
@@
 The inputs are the species, enthalpy, and base state pressure, and the outpus are the species, enthalpy, reaction rates, and nuclear energy generation rate.
-See \cite{MAESTRO_III,MAESTRO_V} for details.
+See Papers III and V for details.
 \end{itemize}
@@
 
-\item Update $\rho_0$ by setting $\rho_0^{n+1,\star} =$ {\bf Avg}$(\rho^{(2),\star})$.
+\item Update $\rho_0$ by setting $\rho_0^{n+1,\star} =$ {\bf Average}$(\rho^{(2),\star})$.
 
@@
 \end{enumerate}
-
-Then, for each Cartesian cell where $\rho^{(2),\star} < \rho_\mathrm{cutoff}$,
-we recompute enthalpy using
-\begin{equation}
-(\rho h)^{(2),\star} = \rho^{(2),\star}h\left(\rho^{(2),\star},p_0^{n+1,\star},X_k^{(2),\star}\right).
-\end{equation}
+Note that outside the cutoff density region, we compute enthalpy using the EOS.
 
@@
 
-\item Update $\rho_0$ by setting $\rho_0^{n+1} =$ {\bf Avg}$(\rho^{(2)})$.
+\item Update $\rho_0$ by setting $\rho_0^{n+1} =$ {\bf Average}$(\rho^{(2)})$.
 
@@
 \item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\pred},$
-  for the conservative update of $(\rho h)^{(1)}$ using a Godunov approach \citep{XRB_III}..
+  for the conservative update of $(\rho h)^{(1)}$ using a Godunov approach \citep{XRB_III}.
 
@@
 \end{enumerate}
-
-Then, for each Cartesian cell where $\rho^{(2)} < \rho_\mathrm{cutoff}$, we recompute enthalpy using
-\begin{equation}
-(\rho h)^{(2)} = \rho^{(2)}h\left(\rho^{(2)},p_0^{n+1},X_k^{(2)}\right).
-\end{equation}
+Note that outside the cutoff density region, we compute enthalpy using the EOS.
 
@@
 \end{equation}
-To find $w_0$ we integrate in 1D using the procedure in Appendix B of \cite{MAESTRO_V}.
+To find $w_0$ we integrate in 1D using the procedure in Appendix B of Paper V.
 
 Note that in this approach, we estimated the time-derivative of the pressure in part by examining how laterally averaged $\rho'=\rho-\rho_0$ changed over time (quantified by $\eta_\rho = \overline{\rho'\Ub\cdot\eb_r}$.
-After evolving the species, we compute this term in the same way as reported in \cite{MAESTRO_V} for use in computing $w_0$.
+After evolving the species, we compute this term in the same way as reported in Paper V for use in computing $w_0$.
 
 \section{Old Notes (not for inclusion in paper)}
+
+\subsection{Exact Base State Notes}
+
+\subsubsection{Potential Problems}
+\begin{enumerate}
+\item How do we define \verb|r_edge_loc| for irregularly-spaced base states whose cell-centered values are defined at $r_m=\Delta x\sqrt{\frac{3}{4} + 2m}$?
+\item We would stil need to approximate the values at edges when advecting the base states (i.e.
+$\rho_0^{\nph,\pred}$).
+This may affect the nice properties we would get with exact cell-centered base state values.
+\item Excessive memory storage may be required (\verb|nr_irreg >> nr_fine|).
+\end{enumerate}
+
+\subsubsection{A Possible Solution}
+We could choose to no longer advect the base states for both predictor and corrector steps.
+This would solve the problems with computing state values at cell edges.
+Instead, we propose to only update the base states $\rho_0$ and $p_0$ by computing the average of the full state values after advecting the full states.
+This would result in the following changes in the algorithm from Paper V:
+%
+\begin{enumerate}
+\item Steps 4 and 8 will be greatly simplified, leading to a more efficient algorithm.
+\item \sout{There is no longer a need for $w_0$, so we can use the full velocity $\Ub$ instead of $\Ubt$.
+As a consequence, $\eta$ does not need to be computed.
+This has the added advantage of saving some memory storage by removing many of the terms needed when we decided to split terms into its base state and perturbation state in our previous implementation.} We split the velocity for the projection to handle the cutoff.
+\item By placing the base states exactly at cell-centers, we eliminate interpolation errors and simplify the existing code.
+\end{enumerate}
+%
 
@@
 \begin{equation}
- \etarho^{\nph} =  {\rm {\bf Avg}} \sum_k \left[ \left(\uadvtwo \cdot \eb_r \right) (\rho X_k)^{\nph,\pred} \right]
+ \etarho^{\nph} =  {\rm {\bf Average}} \sum_k \left[ \left(\uadvtwo \cdot \eb_r \right) (\rho X_k)^{\nph,\pred} \right]
 \end{equation}
@@
 \begin{equation}
-\etarho^{\nph,\star} = {\rm {\bf Avg}}\left(\etarho^{{\rm cart},\nph}\right).
+\etarho^{\nph,\star} = {\rm {\bf Average}}\left(\etarho^{{\rm cart},\nph}\right).
 \end{equation}

### 2019-03-13 MAESTROeX_I/paper.tex
more paper re-org

@@
 
-Previously we adopted an approach where we split the velocity into a base state component, $w_0(r,t)$, 
-and a local velocity $\Ubt(\xb,t)$, so that
-\begin{equation}
-\Ub = \Ubt(\xb,t) + w_0(r,t)\eb_r,
-\end{equation}
-We used $w_0$ to provide an estimate of the base state density evolution over a time step.
-In our new simplified approach, we instead use a predictor-corrector approach for evolving the base state.
-
 \section{Numerical Algorithm}
@@
 \subsection{Temporal Integration Scheme}\label{Sec:Temporal Integration Scheme}
-Our new temporal integration scheme is much simpler than the scheme from Paper V.
-The primary reason is that we no longer split the velocity field into a perturbational and base state components.  
-Previously, we used the base state velocity to predict the expansion of the base state.  
+Previously we adopted an approach where we split the velocity into a base state component, $w_0(r,t)$, 
+and a local velocity $\Ubt(\xb,t)$, so that
+\begin{equation}
+\Ub = \Ubt(\xb,t) + w_0(r,t)\eb_r,
+\end{equation}
+We used $w_0$ to provide an estimate of the base state density evolution over a time step.
+
+Our new temporal integration scheme does not use this splitting, and is much simpler than the scheme from Paper V.
 In this new scheme we use a simpler, predictor-corrector approach to the base state density and pressure that no longer requires the complex algorithm from Paper V, yet still retains the same overall second-order accuracy.
+Also, we now evolve the full-state velocity directly rather than the perturbational velocity, eliminating the need to compute additional forcing terms that arise due to the splitting.
 
@@
 where for our simulations we typically use $\sigma^{\rm CFL}\sim 0.5$.
-There are additional constraints on the time involving the buoyancy force (commonly used when the velocity is approximately zero at the start of some simulations) and the divergence constraint (to prevent too much mass evacuation from a cell in a time step); see Section 3.4 in Paper III for details.
+There are additional constraints on the timestep that are typically much less restrictive than the advective CFL including
+the acceleration due to the buoyancy force (commonly used when the velocity is approximately zero at the start of some simulations) 
+and the local magnitude of the divergence constraint (to prevent too much mass evacuation from a cell in a time step); see Section 3.4 in Paper III for details.
 

### 2019-03-13 MAESTROeX_I/paper.tex
minor

@@
 
+\subsection{Adaptive Mesh Refinement}
+
 \section{Test Problems}
 
-\subsection{Convergence Test}
-3D reacting bubble rise; to demonstrate 2nd-order convergence of the new temporal algorithm.
+\subsection{AMR Performance}
+Show 3D reacting bubble rise performance.
+Show 3-level spherical performance (stats won't be as good but explain why: refining a larger percentage of grid)
 
@@
 
-\subsection{Performance Compared to Original MAESTRO}
-Compare to original MAESTRO
-
 \subsection{White Dwarf Convection}
 white dwarf convection runs with 3 algorithms (original, new temporal, new temporal + irregular base state)
+
+Show 3-level wdconvect
 
@@
 \begin{equation}
-\frac{1}{r_j^2}\left[\left(r^2 w_0\right)_{j+\myhalf} - \left(r^2 w_0\right)_{j-\myhalf}\right] = 
+\frac{1}{r_j^2}\left[\frac{\left(r^2 w_0\right)_{j+\myhalf} - \left(r^2 w_0\right)_{j-\myhalf}}{\Delta r}\right] = 
 \Sbar_j - \frac{1}{\left(\gammaonebar p_0\right)_j}\left[\left(\frac{\partial p_0}{\partial t}\right)_j - \left(\frac{w_{0,j-\myhalf}+w_{0,j+\myhalf}}{2}\right)(\rho_0 g)_j\right]
@@
 \begin{equation}
-\left[\frac{r_{j+\myhalf}^2}{r_j^2} - \frac{(\rho_0 g)_j}{2\left(\gammaonebar p_0\right)_j}\right] w_{0,j+\myhalf} =
-\left[\frac{r_{j-\myhalf}^2}{r_j^2} + \frac{(\rho_0 g)_j}{2\left(\gammaonebar p_0\right)_j}\right] w_{0,j-\myhalf}
+\left[\frac{r_{j+\myhalf}^2}{r_j^2\Delta r} - \frac{(\rho_0 g)_j}{2\left(\gammaonebar p_0\right)_j}\right] w_{0,j+\myhalf} =
+\left[\frac{r_{j-\myhalf}^2}{r_j^2\Delta r} + \frac{(\rho_0 g)_j}{2\left(\gammaonebar p_0\right)_j}\right] w_{0,j-\myhalf}
 + \Sbar_j - \frac{1}{\left(\gammaonebar p_0\right)_j}\left(\frac{\partial p_0}{\partial t}\right)_j

### 2019-03-13 MAESTROeX_I/algorithm.tex
split paper into multiple tex files for multi-user edits

--- 2019-03-13 MAESTROeX_I/paper.tex
@@
-%% This manuscript uses the AASTeX v6.2 LaTeX 2e macros
-%%
-%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls
-%% (Copyright 2000-2015).  See the classfile for details.
-
-%% AASTeX requires revtex4-1.cls (http://publish.aps.org/revtex4/) and
-%% other external packages (latexsym, graphicx, amssymb, longtable, and epsf).
-%% All of these external packages should already be present in the modern TeX
-%% distributions.  If not they can also be obtained at www.ctan.org.
-
-%% The first piece of markup in an AASTeX v6.x document is the \documentclass
-%% command. LaTeX will ignore any data that comes before this command. The
-%% documentclass can take an optional argument to modify the output style.
-%% The command below calls the preprint style  which will produce a tightly
-%% typeset, one-column, single-spaced document.  It is the default and thus
-%% does not need to be explicitly stated.
-%%
-%%
-%% using aastex version 6.2
-\documentclass{aastex62}
-
-%% The default is a single spaced, 10 point font, single spaced article.
-%% There are 5 other style options available via an optional argument. They
-%% can be envoked like this:
-%%
-%% \documentclass[argument]{aastex62}
-%%
-%% where the layout options are:
-%%
-%%  twocolumn   : two text columns, 10 point font, single spaced article.
-%%                This is the most compact and represent the final published
-%%                derived PDF copy of the accepted manuscript from the publisher
-%%  manuscript  : one text column, 12 point font, double spaced article.
-%%  preprint    : one text column, 12 point font, single spaced article.
-%%  preprint2   : two text columns, 12 point font, single spaced article.
-%%  modern      : a stylish, single text column, 12 point font, article with
-%% 		  wider left and right margins. This uses the Daniel
-%% 		  Foreman-Mackey and David Hogg design.
-%%  RNAAS       : Preferred style for Research Notes which are by design
-%%                lacking an abstract and brief. DO NOT use \begin{abstract}
-%%                and \end{abstract} with this style.
-%%
-%% Note that you can submit to the AAS Journals in any of these 6 styles.
-%%
-%% There are other optional arguments one can envoke to allow other stylistic
-%% actions. The available options are:
-%%
-%%  astrosymb    : Loads Astrosymb font and define \astrocommands.
-%%  tighten      : Makes baselineskip slightly smaller, only works with
-%%                 the twocolumn substyle.
-%%  times        : uses times font instead of the default
-%%  linenumbers  : turn on lineno package.
-%%  trackchanges : required to see the revision mark up and print its output
-%%  longauthor   : Do not use the more compressed footnote style (default) for
-%%                 the author/collaboration/affiliations. Instead print all
-%%                 affiliation information after each name. Creates a much
-%%                 long author list but may be desirable for short author papers
-%%
-%% these can be used in any combination, e.g.
-%%
-%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex62}
-%%
-%% AASTeX v6.* now includes \hyperref support. While we have built in specific
-%% defaults into the classfile you can manually override them with the
-%% \hypersetup command. For example,
-%%
-%%\hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
-%%
-%% will change the color of the internal links to red, the links to the
-%% bibliography to green, the file links to cyan, and the external links to
-%% magenta. Additional information on \hyperref options can be found here:
-%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
-%%
-%% If you want to create your own macros, you can do so
-%% using \newcommand. Your macros should appear before
-%% the \begin{document} command.
-%%
-\newcommand{\vdag}{(v)^\dagger}
-\newcommand\aastex{AAS\TeX}
-\newcommand\latex{La\TeX}
-
-% for non-stacked fractions
-\newcommand{\sfrac}[2]{\mathchoice
-  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
-  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
-  {\kern0em\raise.5ex\hbox{\the\scriptscriptfont0 #1}\kern-.2em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptscriptfont0 #2}}
-  {#1\!/#2}}
-
-\newcommand{\myhalf}{\sfrac{1}{2}}
-\newcommand{\thalf}{\sfrac{3}{2}}
-
-\newcommand{\eb}{{\bf{e}}}
-\newcommand{\Ub}{{\bf{U}}}
-\newcommand{\Ubt}{\widetilde{\Ub}}
-\newcommand{\Vb}{{\bf{V}}}
-\newcommand{\xb}{{\bf{x}}}
-
-\newcommand{\dr}{\Delta r}
-\newcommand{\dt}{\Delta t}
-
-\newcommand{\etarho}{\eta_\rho}
-\newcommand{\gammaonebar}{\overline{\Gamma}_1}
-\newcommand{\Hnuc}{H_{\rm nuc}}
-\newcommand{\omegadot}{\dot\omega}
-\newcommand{\pred}{{\rm pred}}
-\newcommand{\Sbar}{\overline{S}}
-
-\newcommand{\inp}{\mathrm{in}}
-\newcommand{\outp}{\mathrm{out}}
-\newcommand{\nph}{{n+\myhalf}}
-\newcommand{\nmh}{{n-\myhalf}}
-\newcommand{\ow}{\overline{w_0}}
-\newcommand{\dw}{\delta w_0}
-\newcommand{\uadvone}{\Ub^{\mathrm{ADV},\star}}
-\newcommand{\uadvonedag}{\Ub^{\mathrm{ADV},\dagger,\star}}
-\newcommand{\uadvtwo}{\Ub^{\mathrm{ADV}}}
-\newcommand{\uadvtwodag}{\Ub^{\mathrm{ADV},\dagger}}
-\newcommand{\gcc}{\mathrm{g~cm^{-3} }}
-
-% for the red MarginPars
-\usepackage{color}
-% make the MarginPars look pretty
-\setlength{\marginparwidth}{0.5in}
-\newcommand{\MarginPar}[1]{\marginpar{\vskip-\baselineskip\raggedright\tiny\sffamily
-\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}
-
-%% Reintroduced the \received and \accepted commands from AASTeX v5.2
-\received{XXX X, XXXX}
-\revised{XXX X, XXXX}
-\accepted{XXX X, XXXX}
-%% Command to document which AAS Journal the manuscript was submitted to.
-%% Adds "Submitted to " the arguement.
-\submitjournal{ApJ}
-
-%% Mark up commands to limit the number of authors on the front page.
-%% Note that in AASTeX v6.2 a \collaboration call (see below) counts as
-%% an author in this case.
-%
-%\AuthorCollaborationLimit=3
-%
-%% Will only show Schwarz, Muench and "the AAS Journals Data Scientist
-%% collaboration" on the front page of this example manuscript.
-%%
-%% Note that all of the author will be shown in the published article.
-%% This feature is meant to be used prior to acceptance to make the
-%% front end of a long author article more manageable. Please do not use
-%% this functionality for manuscripts with less than 20 authors. Conversely,
-%% please do use this when the number of authors exceeds 40.
-%%
-%% Use \allauthors at the manuscript end to show the full author list.
-%% This command should only be used with \AuthorCollaborationLimit is used.
-
-%% The following command can be used to set the latex table counters.  It
-%% is needed in this document because it uses a mix of latex tabular and
-%% AASTeX deluxetables.  In general it should not be needed.
-%\setcounter{table}{1}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% The following section outlines numerous optional output that
-%% can be displayed in the front matter or as running meta-data.
-%%
-%% If you wish, you may supply running head information, although
-%% this information may be modified by the editorial offices.
-\shorttitle{MAESTROeX Low Mach Number Astrophysics}
-\shortauthors{Fan et al.}
-%%
-%% You can add a light gray and diagonal water-mark to the first page
-%% with this command:
-% \watermark{text}
-%% where "text", e.g. DRAFT, is the text to appear.  If the text is
-%% long you can control the water-mark size with:
-%  \setwatermarkfontsize{dimension}
-%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% This is the end of the preamble.  Indicate the beginning of the
-%% manuscript itself with \begin{document}.
-
-\begin{document}
-
-\title{MAESTROeX: A Massively Parallel Low Mach Number Astrophysical Solver}
-
-%% LaTeX will automatically break titles if they run longer than
-%% one line. However, you may use \\ to force a line break if
-%% you desire. In v6.2 you can include a footnote in the title.
-
-%% A significant change from earlier AASTEX versions is in the structure for
-%% calling author and affilations. The change was necessary to implement
-%% autoindexing of affilations which prior was a manual process that could
-%% easily be tedious in large author manuscripts.
-%%
-%% The \author command is the same as before except it now takes an optional
-%% arguement which is the 16 digit ORCID. The syntax is:
-%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
-%%
-%% This will hyperlink the author name to the author's ORCID page. Note that
-%% during compilation, LaTeX will do some limited checking of the format of
-%% the ID to make sure it is valid.
-%%
-%% Use \affiliation for affiliation information. The old \affil is now aliased
-%% to \affiliation. AASTeX v6.2 will automatically index these in the header.
-%% When a duplicate is found its index will be the same as its previous entry.
-%%
-%% Note that \altaffilmark and \altaffiltext have been removed and thus
-%% can not be used to document secondary affiliations. If they are used latex
-%% will issue a specific error message and quit. Please use multiple
-%% \affiliation calls for to document more than one affiliation.
-%%
-%% The new \altaffiliation can be used to indicate some secondary information
-%% such as fellowships. This command produces a non-numeric footnote that is
-%% set away from the numeric \affiliation footnotes.  NOTE that if an
-%% \altaffiliation command is used it must come BEFORE the \affiliation call,
-%% right after the \author command, in order to place the footnotes in
-%% the proper location.
-%%
-%% Use \email to set provide email addresses. Each \email will appear on its
-%% own line so you can put multiple email address in one \email call. A new
-%% \correspondingauthor command is available in V6.2 to identify the
-%% corresponding author of the manuscript. It is the author's responsibility
-%% to make sure this name is also in the author list.
-%%
-%% While authors can be grouped inside the same \author and \affiliation
-%% commands it is better to have a single author for each. This allows for
-%% one to exploit all the new benefits and should make book-keeping easier.
-%%
-%% If done correctly the peer review system will be able to
-%% automatically put the author and affiliation information from the manuscript
-%% and save the corresponding author the trouble of entering it by hand.
-
-\correspondingauthor{Doreen Fan; DFan@lbl.gov}
-
-\author[0000-0002-3246-4315]{Duoming Fan}
-\affil{Lawrence Berkelely National Laboratory \\
-Center for Computational Sciences and Engineering \\
-One Cyclotron Road, MS 50A-3111 \\
-Berkeley, CA 94720, USA}
-
-\author[0000-0003-1791-0265]{Andrew Nonaka}
-\affil{Lawrence Berkelely National Laboratory \\
-Center for Computational Sciences and Engineering \\
-One Cyclotron Road, MS 50A-3111 \\
-Berkeley, CA 94720, USA}
-
-\author[0000-0003-2103-312X]{Ann S. Almgren}
-\affil{Lawrence Berkelely National Laboratory \\
-Center for Computational Sciences and Engineering \\
-One Cyclotron Road, MS 50A-3111 \\
-Berkeley, CA 94720, USA}
-
-\author[0000-0002-1530-781X]{Alice Harpole}
-\affil{Stony Brook University \\
-Department of Physics and Astronomy \\
-Stony Brook, NY 11794-3800, USA}
-
-\author[0000-0001-8401-030X]{Michael Zingale}
-\affil{Stony Brook University \\
-Department of Physics and Astronomy \\
-Stony Brook, NY 11794-3800, USA}
-
-%% Note that the \and command from previous versions of AASTeX is now
-%% depreciated in this version as it is no longer necessary. AASTeX
-%% automatically takes care of all commas and "and"s between authors names.
-
-%% AASTeX 6.2 has the new \collaboration and \nocollaboration commands to
-%% provide the collaboration status of a group of authors. These commands
-%% can be used either before or after the list of corresponding authors. The
-%% argument for \collaboration is the collaboration identifier. Authors are
-%% encouraged to surround collaboration identifiers with ()s. The
-%% \nocollaboration command takes no argument and exists to indicate that
-%% the nearby authors are not part of surrounding collaborations.
-
-%% Mark off the abstract in the ``abstract'' environment.
-\begin{abstract}
-We present MAESTROeX, a massively parallel solver for low Mach number astrophysical flows.
-Our model equations allow for long-time integration for highly subsonic flows compared to compressible approaches.
-The code leverages the new AMReX framework for block-structured adaptive mesh refinement calculations, and also provides additional algorithmic features from the original MAESTRO code.
-We present a new temporal integration option that is much simpler than our previous approach while retaining the same order order of accuracy.
-We also provide a more accurate spatial discretization for mapping the base state onto the full state Cartesian grid.
-Using our previous studies on the convective phase of single-degenerate progenitor models of Type Ia supernovae, we characterize the performance of the code and validate the new algorithmic features.
-\end{abstract}
-
-%% Keywords should appear after the \end{abstract} command.
-%% See the online documentation for the full list of available subject
-%% keywords and the rules for their use.
-\keywords{convection, hydrodynamics, methods: numerical, nuclear reactions, nucleosynthesis, abundances, supernovae: general}
-
-%% From the front matter, we move on to the body of the paper.
-%% Sections are demarcated by \section and \subsection, respectively.
-%% Observe the use of the LaTeX \label
-%% command after the \subsection to give a symbolic KEY to the
-%% subsection for cross-referencing in a \ref command.
-%% You can use LaTeX's \ref and \label commands to keep track of
-%% cross-references to sections, equations, tables, and figures.
-%% That way, if you change the order of any elements, LaTeX will
-%% automatically renumber them.
-%%
-%% We recommend that authors also use the natbib \citep
-%% and \citet commands to identify citations.  The citations are
-%% tied to the reference list via symbolic KEYs. The KEY corresponds
-%% to the KEY in the \bibitem in the reference list below.
-
-\section{Introduction} \label{sec:intro}
-
-TODO list:
-\begin{itemize}
-\item Weak scaling tests of original algorithm, comparing MAESTRO to MAESTROeX.  Use cori knl, 4 MPI per node, 16 threads per MPI
-\item Run at effective $512^3$ resolution (maybe to ignition?...
-need to see how max T trends go) (1) classic MAESTRO, (2) MAESTROeX with original algorithm, (3) MAESTROeX with new temporal integrator, (4) MAESTROeX with new temporal integrator and irregular dr, (5) MAESTROeX with original algorithm and 3-levels of AMR, (6) MAESTROeX with new temporal integrator and 3 levels of AMR.
-\item (perhaps) a demonstration of what goes wrong when you don't split the velocity dynamics in the projection
-\end{itemize}
-
-Many astrophysical flows are highly subsonic; sound waves carry sufficiently little energy that they do not significantly affect the convective dynamics of the system.
-In many of these flows, modeling long-time convective dynamics are of interest, and numerical approaches based on compressible hydrodynamics are intractable, even on modern supercomputers.
-One approach to this problem is to use low Mach number models.
-In a low Mach number approach, sound waves are eliminated from the governing equations while retaining compressibilitiy effects due to, e.g., nuclear energy release, compositional changes, and thermal diffusion.
-The resulting model can be numerically integrated with much larger time steps than a compressible model.
-
-Previously, we have developed the low Mach number astrophysical solver, MAESTRO.
-Low Mach number models have been developed for a variety of contexts including combustion \citep{day2000numerical}, atmospherics \citep{duarte2015low}, and elastic solids \citep{abbate2017all}.
-The low Mach number model in MAESTRO is unique in that it is specifically designed for astrophysical settings with significant atmospheric stratification.
-Central to the algorithm is a stratified background (or base) state density and pressure held in hydrostatic equilibrium, and also vary as a function of altitude and time.
-MAESTRO is suitable for for spherical stars, as well as planar simulations of local dynamics.
-
-The key numerical developments of the original MAESTRO algorithm are presented in a series of papers which we refer to as Papers I-V:
-\begin{itemize}
-\item In Paper I \citep{MAESTRO_I}, we derive the low Mach number equation set from the fully compressible equations.
-\item In Paper II \citep{MAESTRO_II}, we incorporate the effects atmospheric expansion through the use of a time-dependent background state.
-\item In Paper III \citep{MAESTRO_III}, we incorporate reactions and the associated coupling to the hydrodynamics.
-\item In Paper IV \citep{MAESTRO_IV}, we describe our treatment of spherical stars in a three-dimensional Cartesian geometry.
-\item In Paper V \citep{MAESTRO_V}, we describe the use of block-structured adaptive mesh refinement to focus spatial resolution in regions of interest.
-\end{itemize}
-
-Since then, there have been many scientific investigations using MAESTRO, which include additional algorithmic enhancements.  Topics include:
-\begin{itemize}
-\item The convective phase preceding Chandrasekhar mass models for type Ia supernovae \citep{MAESTRO_convection,MAESTRO_AMR,MAESTRO_CASTRO}.
-\item Convection in massive stars \citep{Gilet:2013}.
-\item Sub-Chandrasekhar white dwarfs \citep{subChandra_I,subChandra_II}.
-In the latter paper we introduced an optional modification to the momentum equation of velocity constraint that conserves total energy, can give results closer to compressible codes of convective dynamics, including low density regions at the surface of a star.
-\item Type I X-ray bursts \citep{XRB_I,XRB_II,XRB_III}.
-In \cite{XRB_I} we discuss the modifications required for implicit thermal conduction, as well as an additional ``volume discrepancy'' modification to the velocity constraint to force the species and enthalpy to evolve in a manner consistent with the background pressure.
-\end{itemize}
-
-In this paper, we characterize the performance of the MAESTROeX algorithm compared to the previous implementation.
-We also develop simpler, alternative temporal integrators, with a longer term goal of developing high-order coupling schemes.
-We also present a new treatment of the base state mapping to Cartesian grids for spherical problsm.
-
-\section{Governing Equations}
-The derivation of the low Mach number hydrodynamics equations is given in Papers I-III.
-We take the standard equations of reacting, compressible flow, and recast the equation
-of state (EOS) as a divergence constraint on the velocity field.  This is done by taking
-the Langrangian derivative of the equation of state for pressure as a function of the
-density, composition, and enthalpy, enforcing the constraint that the pressure is
-a prescribed function of altitutde and time based on the hydrostatic equilibrium condition,
-and substituting the equations for species and enthalpy in the resulting Lagrangian derivative.
-The resulting equations are
-\begin{eqnarray}
-\frac{\partial\Ub}{\partial t} &=& -\Ub\cdot\nabla\Ub  - \frac{1}{\rho}\nabla\pi - \frac{\rho-\rho_0}{\rho} g\eb_r,\label{eq:momentum}\\
-\frac{\partial(\rho X_k)}{\partial t} &=& -\nabla\cdot(\rho X_k\Ub) + \rho\omegadot_k,\label{eq:species}\\
-\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc,\label{eq:enthalpy}
-\end{eqnarray}
-\begin{equation}
-\nabla\cdot(\beta_0\Ub) = \beta_0\left(S - \frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial t}\right).\label{eq:U divergence}
-\end{equation}
-Here $\rho$, $\Ub$, and $h$ are the mass density,
-velocity and specific enthalpy, respectively, and
-$X_k$ are the mass fractions of species $k$ with associated
-production rate $\omegadot_k$.  The species are constrained
-such that $\sum_k X_k = 1$ giving $\rho = \sum_k (\rho X_k)$ and
-\begin{equation}
-\frac{\partial\rho}{\partial t} = -\nabla\cdot(\rho\Ub).
-\end{equation}
-Here $\Hnuc$ is the nuclear energy generation rate per unit mass.
-The pressure is decomposed into a hydrostatic base state
- pressure, $p_0 = p_0(r,t)$, and a dynamic pressure, $\pi = \pi(\xb,t)$, such that 
-$|\pi|/p_0 = \mathcal{O}(M^2)$ (we use $\xb$ to represent the Cartesian coordinate 
-directions of the full state and $r$ to represent the radial coordinate direction for 
-the base state).  We also define a base state density, $\rho_0 = \rho_0(r,t)$, 
-which is in hydrostatic equilibrium with $p_0$, i.e., 
-$\nabla p_0 = -\rho_0 g\eb_r$, where $g=g(r,t)$ is
-the magnitude of the gravitational acceleration and $\eb_r$ is the unit vector in the
-outward radial direction.
-
-Mathematically, equations (\ref{eq:momentum})-(\ref{eq:enthalpy}) must still be closed by the equation of state which we
-express as a divergence constraint on the velocity field (\ref{eq:U divergence}),
-\MarginPar{This is different from the Jacobs 2016 paper where we switched to alt\_energy\_fix.}
-where $\beta_0$ is a density-like variable that carries background stratification, defined as
-\begin{equation}
-\beta_0(r,t) = \rho_0(0,t)\exp\left(\int_0^r\frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial r'}dr'\right),
-\end{equation}
-where $\gammaonebar$ is the lateral average of $\Gamma_1 = d(\log p)/d(\log\rho)$ at constant entropy.
-The expansion term, $S$, incorporates local compressibility effects due to heat release from reactions, compositional changes, and external sources,
-\begin{equation}
-S = -\sigma\sum_k\xi_k\omegadot_k + \frac{1}{\rho p_\rho}\sum_k p_{X_k}\omegadot_k + \sigma\Hnuc,\label{eq:S}
-\end{equation}
-where $p_{X_k} \equiv \left. \partial p / \partial X_k
-\right|_{\rho,T,X_{j,j\ne k}}$, $\xi_k \equiv \left. \partial h /
-\partial X_k \right |_{p,T,X_{j,j\ne k}},
-p_\rho \equiv \left.
-\partial p/\partial \rho \right |_{T, X_k}$, and $\sigma \equiv
-p_T/(\rho c_p p_\rho)$, with $p_T \equiv \left. \partial p / \partial
-T \right|_{\rho, X_k}$ and $c_p \equiv \left.  \partial h / \partial T
-\right|_{p,X_k}$ is the specific heat at constant pressure.
-
 \section{Numerical Algorithm}
@@
 \subsection{Adaptive Mesh Refinement}
-
-\section{Test Problems}
-
-\subsection{AMR Performance}
-Show 3D reacting bubble rise performance.
-Show 3-level spherical performance (stats won't be as good but explain why: refining a larger percentage of grid)
-
-\subsection{Scaling}
-Scaling with OpenMP
-
-\subsection{White Dwarf Convection}
-white dwarf convection runs with 3 algorithms (original, new temporal, new temporal + irregular base state)
-
-Show 3-level wdconvect
-
-\section{Conclusions and Future Work}
-
-science: rotation, solar physics, MHD
-
-algorithm: SDC \cite{dutt2000spectral}, high-order, multi-implicit \cite{bourlioux2003high}.
-Consider SDC ideas used for terrestrial combustion \cite{pazner2016high,nonaka2018conservative}
-
-implementation: GPUs.  Forward cite next CASTRO paper with GPU strategy.
-
-\appendix
-\section{Projection Details}
-To enforce the divergence constraint in {\bf Step 3, 7}, and {\bf 11}, we use a projection method analgous to the methods originally developed for incompressible flow.
-
-Based on our past experience in the MAESTRO project, we have found it useful to split the velocity dynamics into a perturbational and base state velocity,
-\begin{equation}
-\Ub = \Ubt(\xb,t) + w_0(r,t)\eb_r,
-\end{equation}
-solve for each term separately, and immediately combine them to find a full velocity that satisfies the constraint.
-We take that approach here, primarily because it allows us to enforce a boundary condition on $w_0$ at the edge of the star (i.e., the cutoff density location where we hold density constant).
-Namely, to enforce that $r^2 w_0$ remain constant is difficult to do when solving for the full velocity.
-So in practice, we solve the constraint over the lateral average,
-\begin{equation}
-\nabla\cdot\left(\beta_0\Ubt\right) = \beta_0(S-\overline{S})
-\end{equation}
-and sepratately solve for $w_0$ using 
-\begin{equation}
-\nabla\cdot(\beta_0w_0\eb_r) = \beta_0\left(\overline{S} - \frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial t}\right).
-\end{equation}
-To find $w_0$ we integrate in 1D using the procedure in Appendix B of Paper V.
-
-Note that in this approach, we estimated the time-derivative of the pressure in part by examining how laterally averaged $\rho'=\rho-\rho_0$ changed over time (quantified by $\eta_\rho = \overline{\rho'\Ub\cdot\eb_r}$.
-After evolving the species, we compute this term in the same way as reported in Paper V for use in computing $w_0$.
-
-\section{Old Notes (not for inclusion in paper)}
-
-\subsection{Exact Base State Notes}
-
-\subsubsection{Potential Problems}
-\begin{enumerate}
-\item How do we define \verb|r_edge_loc| for irregularly-spaced base states whose cell-centered values are defined at $r_m=\Delta x\sqrt{\frac{3}{4} + 2m}$?
-\item We would stil need to approximate the values at edges when advecting the base states (i.e.
-$\rho_0^{\nph,\pred}$).
-This may affect the nice properties we would get with exact cell-centered base state values.
-\item Excessive memory storage may be required (\verb|nr_irreg >> nr_fine|).
-\end{enumerate}
-
-\subsubsection{A Possible Solution}
-We could choose to no longer advect the base states for both predictor and corrector steps.
-This would solve the problems with computing state values at cell edges.
-Instead, we propose to only update the base states $\rho_0$ and $p_0$ by computing the average of the full state values after advecting the full states.
-This would result in the following changes in the algorithm from Paper V:
-%
-\begin{enumerate}
-\item Steps 4 and 8 will be greatly simplified, leading to a more efficient algorithm.
-\item \sout{There is no longer a need for $w_0$, so we can use the full velocity $\Ub$ instead of $\Ubt$.
-As a consequence, $\eta$ does not need to be computed.
-This has the added advantage of saving some memory storage by removing many of the terms needed when we decided to split terms into its base state and perturbation state in our previous implementation.} We split the velocity for the projection to handle the cutoff.
-\item By placing the base states exactly at cell-centers, we eliminate interpolation errors and simplify the existing code.
-\end{enumerate}
-%
-
-\subsection{$\eta_\rho$ Notes}
-After {\bf Step 8D}, we define a radial cell-centered $\etarho^{\nph}$.
-
-\begin{description}
-\item For planar geometry, $\etarho = \overline{\rho'(\Ub\cdot\eb_r)}$,
-\begin{equation}
- \etarho^{\nph} =  {\rm {\bf Average}} \sum_k \left[ \left(\uadvtwo \cdot \eb_r \right) (\rho X_k)^{\nph,\pred} \right]
-\end{equation}
-\item For spherical geometry, first construct 
-$\etarho^{{\rm cart},\nph} = [\rho'(\Ub\cdot\eb_r)]^{\nph}$ on Cartesian cell centers using:
-\begin{equation}
-\etarho^{{\rm cart},\nph} = \left[\left(\frac{\rho^{(1)}+\rho^{(2)}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{n+1}}{2}\right)\right] \cdot \left( \uadvtwo \cdot \eb_r \right).
-\end{equation}
-Then,
-\begin{equation}
-\etarho^{\nph,\star} = {\rm {\bf Average}}\left(\etarho^{{\rm cart},\nph}\right).
-\end{equation}
-\end{description}
-
-\subsection{Computing $w_0$ For Spherical Problems}
-Recall that we want to solve
-\begin{equation}
-\frac{1}{r^2}\frac{\partial}{\partial r}(r^2 w_0) = \Sbar - \frac{1}{\gammaonebar p_0}\left(\frac{\partial p_0}{\partial t} + w_0\frac{\partial p_0}{\partial r}\right)
-\end{equation}
-First, note that $\partial p_0/\partial r = -\rho_0 g$.
-The key observation here is that we can compute $\partial p_0/\partial t$ directly instead of making complicated substitutions involving $\eta_\rho$ to eliminate it.
-Let's discretize this as follows:
-\begin{equation}
-\frac{1}{r_j^2}\left[\frac{\left(r^2 w_0\right)_{j+\myhalf} - \left(r^2 w_0\right)_{j-\myhalf}}{\Delta r}\right] = 
-\Sbar_j - \frac{1}{\left(\gammaonebar p_0\right)_j}\left[\left(\frac{\partial p_0}{\partial t}\right)_j - \left(\frac{w_{0,j-\myhalf}+w_{0,j+\myhalf}}{2}\right)(\rho_0 g)_j\right]
-\end{equation}
-Rearranging this to solve for $w_{0,j+\myhalf}$ gives us an explicit update:
-\begin{equation}
-\left[\frac{r_{j+\myhalf}^2}{r_j^2\Delta r} - \frac{(\rho_0 g)_j}{2\left(\gammaonebar p_0\right)_j}\right] w_{0,j+\myhalf} =
-\left[\frac{r_{j-\myhalf}^2}{r_j^2\Delta r} + \frac{(\rho_0 g)_j}{2\left(\gammaonebar p_0\right)_j}\right] w_{0,j-\myhalf}
-+ \Sbar_j - \frac{1}{\left(\gammaonebar p_0\right)_j}\left(\frac{\partial p_0}{\partial t}\right)_j
-\end{equation}
-As before, once $\rho_0$ falls below $\rho_{\rm cutoff}$, we hold $r^2 w_0$ constant.\\ \\
-The interface to the routine in the algorithm description is
-{\bf Make $\mathbf{w_0}$}$[\rho_0,p_0,\partial p_0/\partial t,\gammaonebar,\Sbar] \rightarrow [w_0]$
-
-\acknowledgements
-
-The work at LBNL was supported by the U.S. Department of Energy's Scientific Discovery Through Advanced Computing (SciDAC) program under contract No. DE-AC02-05CH11231.
-This research used resources of the National Energy Research Scientific Computing Center (NERSC), a U.S.
-Department of Energy Office of Science User Facility operated under Contract No.
-DE-AC02-05CH11231.
-
-\bibliographystyle{aasjournal.bst}
-\bibliography{references}
-
-\end{document}

### 2019-03-13 MAESTROeX_I/appendix.tex
split paper into multiple tex files for multi-user edits

--- 2019-03-13 MAESTROeX_I/paper.tex
@@
-%% This manuscript uses the AASTeX v6.2 LaTeX 2e macros
-%%
-%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls
-%% (Copyright 2000-2015).  See the classfile for details.
-
-%% AASTeX requires revtex4-1.cls (http://publish.aps.org/revtex4/) and
-%% other external packages (latexsym, graphicx, amssymb, longtable, and epsf).
-%% All of these external packages should already be present in the modern TeX
-%% distributions.  If not they can also be obtained at www.ctan.org.
-
-%% The first piece of markup in an AASTeX v6.x document is the \documentclass
-%% command. LaTeX will ignore any data that comes before this command. The
-%% documentclass can take an optional argument to modify the output style.
-%% The command below calls the preprint style  which will produce a tightly
-%% typeset, one-column, single-spaced document.  It is the default and thus
-%% does not need to be explicitly stated.
-%%
-%%
-%% using aastex version 6.2
-\documentclass{aastex62}
-
-%% The default is a single spaced, 10 point font, single spaced article.
-%% There are 5 other style options available via an optional argument. They
-%% can be envoked like this:
-%%
-%% \documentclass[argument]{aastex62}
-%%
-%% where the layout options are:
-%%
-%%  twocolumn   : two text columns, 10 point font, single spaced article.
-%%                This is the most compact and represent the final published
-%%                derived PDF copy of the accepted manuscript from the publisher
-%%  manuscript  : one text column, 12 point font, double spaced article.
-%%  preprint    : one text column, 12 point font, single spaced article.
-%%  preprint2   : two text columns, 12 point font, single spaced article.
-%%  modern      : a stylish, single text column, 12 point font, article with
-%% 		  wider left and right margins. This uses the Daniel
-%% 		  Foreman-Mackey and David Hogg design.
-%%  RNAAS       : Preferred style for Research Notes which are by design
-%%                lacking an abstract and brief. DO NOT use \begin{abstract}
-%%                and \end{abstract} with this style.
-%%
-%% Note that you can submit to the AAS Journals in any of these 6 styles.
-%%
-%% There are other optional arguments one can envoke to allow other stylistic
-%% actions. The available options are:
-%%
-%%  astrosymb    : Loads Astrosymb font and define \astrocommands.
-%%  tighten      : Makes baselineskip slightly smaller, only works with
-%%                 the twocolumn substyle.
-%%  times        : uses times font instead of the default
-%%  linenumbers  : turn on lineno package.
-%%  trackchanges : required to see the revision mark up and print its output
-%%  longauthor   : Do not use the more compressed footnote style (default) for
-%%                 the author/collaboration/affiliations. Instead print all
-%%                 affiliation information after each name. Creates a much
-%%                 long author list but may be desirable for short author papers
-%%
-%% these can be used in any combination, e.g.
-%%
-%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex62}
-%%
-%% AASTeX v6.* now includes \hyperref support. While we have built in specific
-%% defaults into the classfile you can manually override them with the
-%% \hypersetup command. For example,
-%%
-%%\hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
-%%
-%% will change the color of the internal links to red, the links to the
-%% bibliography to green, the file links to cyan, and the external links to
-%% magenta. Additional information on \hyperref options can be found here:
-%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
-%%
-%% If you want to create your own macros, you can do so
-%% using \newcommand. Your macros should appear before
-%% the \begin{document} command.
-%%
-\newcommand{\vdag}{(v)^\dagger}
-\newcommand\aastex{AAS\TeX}
-\newcommand\latex{La\TeX}
-
-% for non-stacked fractions
-\newcommand{\sfrac}[2]{\mathchoice
-  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
-  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
-  {\kern0em\raise.5ex\hbox{\the\scriptscriptfont0 #1}\kern-.2em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptscriptfont0 #2}}
-  {#1\!/#2}}
-
-\newcommand{\myhalf}{\sfrac{1}{2}}
-\newcommand{\thalf}{\sfrac{3}{2}}
-
-\newcommand{\eb}{{\bf{e}}}
-\newcommand{\Ub}{{\bf{U}}}
-\newcommand{\Ubt}{\widetilde{\Ub}}
-\newcommand{\Vb}{{\bf{V}}}
-\newcommand{\xb}{{\bf{x}}}
-
-\newcommand{\dr}{\Delta r}
-\newcommand{\dt}{\Delta t}
-
-\newcommand{\etarho}{\eta_\rho}
-\newcommand{\gammaonebar}{\overline{\Gamma}_1}
-\newcommand{\Hnuc}{H_{\rm nuc}}
-\newcommand{\omegadot}{\dot\omega}
-\newcommand{\pred}{{\rm pred}}
-\newcommand{\Sbar}{\overline{S}}
-
-\newcommand{\inp}{\mathrm{in}}
-\newcommand{\outp}{\mathrm{out}}
-\newcommand{\nph}{{n+\myhalf}}
-\newcommand{\nmh}{{n-\myhalf}}
-\newcommand{\ow}{\overline{w_0}}
-\newcommand{\dw}{\delta w_0}
-\newcommand{\uadvone}{\Ub^{\mathrm{ADV},\star}}
-\newcommand{\uadvonedag}{\Ub^{\mathrm{ADV},\dagger,\star}}
-\newcommand{\uadvtwo}{\Ub^{\mathrm{ADV}}}
-\newcommand{\uadvtwodag}{\Ub^{\mathrm{ADV},\dagger}}
-\newcommand{\gcc}{\mathrm{g~cm^{-3} }}
-
-% for the red MarginPars
-\usepackage{color}
-% make the MarginPars look pretty
-\setlength{\marginparwidth}{0.5in}
-\newcommand{\MarginPar}[1]{\marginpar{\vskip-\baselineskip\raggedright\tiny\sffamily
-\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}
-
-%% Reintroduced the \received and \accepted commands from AASTeX v5.2
-\received{XXX X, XXXX}
-\revised{XXX X, XXXX}
-\accepted{XXX X, XXXX}
-%% Command to document which AAS Journal the manuscript was submitted to.
-%% Adds "Submitted to " the arguement.
-\submitjournal{ApJ}
-
-%% Mark up commands to limit the number of authors on the front page.
-%% Note that in AASTeX v6.2 a \collaboration call (see below) counts as
-%% an author in this case.
-%
-%\AuthorCollaborationLimit=3
-%
-%% Will only show Schwarz, Muench and "the AAS Journals Data Scientist
-%% collaboration" on the front page of this example manuscript.
-%%
-%% Note that all of the author will be shown in the published article.
-%% This feature is meant to be used prior to acceptance to make the
-%% front end of a long author article more manageable. Please do not use
-%% this functionality for manuscripts with less than 20 authors. Conversely,
-%% please do use this when the number of authors exceeds 40.
-%%
-%% Use \allauthors at the manuscript end to show the full author list.
-%% This command should only be used with \AuthorCollaborationLimit is used.
-
-%% The following command can be used to set the latex table counters.  It
-%% is needed in this document because it uses a mix of latex tabular and
-%% AASTeX deluxetables.  In general it should not be needed.
-%\setcounter{table}{1}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% The following section outlines numerous optional output that
-%% can be displayed in the front matter or as running meta-data.
-%%
-%% If you wish, you may supply running head information, although
-%% this information may be modified by the editorial offices.
-\shorttitle{MAESTROeX Low Mach Number Astrophysics}
-\shortauthors{Fan et al.}
-%%
-%% You can add a light gray and diagonal water-mark to the first page
-%% with this command:
-% \watermark{text}
-%% where "text", e.g. DRAFT, is the text to appear.  If the text is
-%% long you can control the water-mark size with:
-%  \setwatermarkfontsize{dimension}
-%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% This is the end of the preamble.  Indicate the beginning of the
-%% manuscript itself with \begin{document}.
-
-\begin{document}
-
-\title{MAESTROeX: A Massively Parallel Low Mach Number Astrophysical Solver}
-
-%% LaTeX will automatically break titles if they run longer than
-%% one line. However, you may use \\ to force a line break if
-%% you desire. In v6.2 you can include a footnote in the title.
-
-%% A significant change from earlier AASTEX versions is in the structure for
-%% calling author and affilations. The change was necessary to implement
-%% autoindexing of affilations which prior was a manual process that could
-%% easily be tedious in large author manuscripts.
-%%
-%% The \author command is the same as before except it now takes an optional
-%% arguement which is the 16 digit ORCID. The syntax is:
-%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
-%%
-%% This will hyperlink the author name to the author's ORCID page. Note that
-%% during compilation, LaTeX will do some limited checking of the format of
-%% the ID to make sure it is valid.
-%%
-%% Use \affiliation for affiliation information. The old \affil is now aliased
-%% to \affiliation. AASTeX v6.2 will automatically index these in the header.
-%% When a duplicate is found its index will be the same as its previous entry.
-%%
-%% Note that \altaffilmark and \altaffiltext have been removed and thus
-%% can not be used to document secondary affiliations. If they are used latex
-%% will issue a specific error message and quit. Please use multiple
-%% \affiliation calls for to document more than one affiliation.
-%%
-%% The new \altaffiliation can be used to indicate some secondary information
-%% such as fellowships. This command produces a non-numeric footnote that is
-%% set away from the numeric \affiliation footnotes.  NOTE that if an
-%% \altaffiliation command is used it must come BEFORE the \affiliation call,
-%% right after the \author command, in order to place the footnotes in
-%% the proper location.
-%%
-%% Use \email to set provide email addresses. Each \email will appear on its
-%% own line so you can put multiple email address in one \email call. A new
-%% \correspondingauthor command is available in V6.2 to identify the
-%% corresponding author of the manuscript. It is the author's responsibility
-%% to make sure this name is also in the author list.
-%%
-%% While authors can be grouped inside the same \author and \affiliation
-%% commands it is better to have a single author for each. This allows for
-%% one to exploit all the new benefits and should make book-keeping easier.
-%%
-%% If done correctly the peer review system will be able to
-%% automatically put the author and affiliation information from the manuscript
-%% and save the corresponding author the trouble of entering it by hand.
-
-\correspondingauthor{Doreen Fan; DFan@lbl.gov}
-
-\author[0000-0002-3246-4315]{Duoming Fan}
-\affil{Lawrence Berkelely National Laboratory \\
-Center for Computational Sciences and Engineering \\
-One Cyclotron Road, MS 50A-3111 \\
-Berkeley, CA 94720, USA}
-
-\author[0000-0003-1791-0265]{Andrew Nonaka}
-\affil{Lawrence Berkelely National Laboratory \\
-Center for Computational Sciences and Engineering \\
-One Cyclotron Road, MS 50A-3111 \\
-Berkeley, CA 94720, USA}
-
-\author[0000-0003-2103-312X]{Ann S. Almgren}
-\affil{Lawrence Berkelely National Laboratory \\
-Center for Computational Sciences and Engineering \\
-One Cyclotron Road, MS 50A-3111 \\
-Berkeley, CA 94720, USA}
-
-\author[0000-0002-1530-781X]{Alice Harpole}
-\affil{Stony Brook University \\
-Department of Physics and Astronomy \\
-Stony Brook, NY 11794-3800, USA}
-
-\author[0000-0001-8401-030X]{Michael Zingale}
-\affil{Stony Brook University \\
-Department of Physics and Astronomy \\
-Stony Brook, NY 11794-3800, USA}
-
-%% Note that the \and command from previous versions of AASTeX is now
-%% depreciated in this version as it is no longer necessary. AASTeX
-%% automatically takes care of all commas and "and"s between authors names.
-
-%% AASTeX 6.2 has the new \collaboration and \nocollaboration commands to
-%% provide the collaboration status of a group of authors. These commands
-%% can be used either before or after the list of corresponding authors. The
-%% argument for \collaboration is the collaboration identifier. Authors are
-%% encouraged to surround collaboration identifiers with ()s. The
-%% \nocollaboration command takes no argument and exists to indicate that
-%% the nearby authors are not part of surrounding collaborations.
-
-%% Mark off the abstract in the ``abstract'' environment.
-\begin{abstract}
-We present MAESTROeX, a massively parallel solver for low Mach number astrophysical flows.
-Our model equations allow for long-time integration for highly subsonic flows compared to compressible approaches.
-The code leverages the new AMReX framework for block-structured adaptive mesh refinement calculations, and also provides additional algorithmic features from the original MAESTRO code.
-We present a new temporal integration option that is much simpler than our previous approach while retaining the same order order of accuracy.
-We also provide a more accurate spatial discretization for mapping the base state onto the full state Cartesian grid.
-Using our previous studies on the convective phase of single-degenerate progenitor models of Type Ia supernovae, we characterize the performance of the code and validate the new algorithmic features.
-\end{abstract}
-
-%% Keywords should appear after the \end{abstract} command.
-%% See the online documentation for the full list of available subject
-%% keywords and the rules for their use.
-\keywords{convection, hydrodynamics, methods: numerical, nuclear reactions, nucleosynthesis, abundances, supernovae: general}
-
-%% From the front matter, we move on to the body of the paper.
-%% Sections are demarcated by \section and \subsection, respectively.
-%% Observe the use of the LaTeX \label
-%% command after the \subsection to give a symbolic KEY to the
-%% subsection for cross-referencing in a \ref command.
-%% You can use LaTeX's \ref and \label commands to keep track of
-%% cross-references to sections, equations, tables, and figures.
-%% That way, if you change the order of any elements, LaTeX will
-%% automatically renumber them.
-%%
-%% We recommend that authors also use the natbib \citep
-%% and \citet commands to identify citations.  The citations are
-%% tied to the reference list via symbolic KEYs. The KEY corresponds
-%% to the KEY in the \bibitem in the reference list below.
-
-\section{Introduction} \label{sec:intro}
-
-TODO list:
-\begin{itemize}
-\item Weak scaling tests of original algorithm, comparing MAESTRO to MAESTROeX.  Use cori knl, 4 MPI per node, 16 threads per MPI
-\item Run at effective $512^3$ resolution (maybe to ignition?...
-need to see how max T trends go) (1) classic MAESTRO, (2) MAESTROeX with original algorithm, (3) MAESTROeX with new temporal integrator, (4) MAESTROeX with new temporal integrator and irregular dr, (5) MAESTROeX with original algorithm and 3-levels of AMR, (6) MAESTROeX with new temporal integrator and 3 levels of AMR.
-\item (perhaps) a demonstration of what goes wrong when you don't split the velocity dynamics in the projection
-\end{itemize}
-
-Many astrophysical flows are highly subsonic; sound waves carry sufficiently little energy that they do not significantly affect the convective dynamics of the system.
-In many of these flows, modeling long-time convective dynamics are of interest, and numerical approaches based on compressible hydrodynamics are intractable, even on modern supercomputers.
-One approach to this problem is to use low Mach number models.
-In a low Mach number approach, sound waves are eliminated from the governing equations while retaining compressibilitiy effects due to, e.g., nuclear energy release, compositional changes, and thermal diffusion.
-The resulting model can be numerically integrated with much larger time steps than a compressible model.
-
-Previously, we have developed the low Mach number astrophysical solver, MAESTRO.
-Low Mach number models have been developed for a variety of contexts including combustion \citep{day2000numerical}, atmospherics \citep{duarte2015low}, and elastic solids \citep{abbate2017all}.
-The low Mach number model in MAESTRO is unique in that it is specifically designed for astrophysical settings with significant atmospheric stratification.
-Central to the algorithm is a stratified background (or base) state density and pressure held in hydrostatic equilibrium, and also vary as a function of altitude and time.
-MAESTRO is suitable for for spherical stars, as well as planar simulations of local dynamics.
-
-The key numerical developments of the original MAESTRO algorithm are presented in a series of papers which we refer to as Papers I-V:
-\begin{itemize}
-\item In Paper I \citep{MAESTRO_I}, we derive the low Mach number equation set from the fully compressible equations.
-\item In Paper II \citep{MAESTRO_II}, we incorporate the effects atmospheric expansion through the use of a time-dependent background state.
-\item In Paper III \citep{MAESTRO_III}, we incorporate reactions and the associated coupling to the hydrodynamics.
-\item In Paper IV \citep{MAESTRO_IV}, we describe our treatment of spherical stars in a three-dimensional Cartesian geometry.
-\item In Paper V \citep{MAESTRO_V}, we describe the use of block-structured adaptive mesh refinement to focus spatial resolution in regions of interest.
-\end{itemize}
-
-Since then, there have been many scientific investigations using MAESTRO, which include additional algorithmic enhancements.  Topics include:
-\begin{itemize}
-\item The convective phase preceding Chandrasekhar mass models for type Ia supernovae \citep{MAESTRO_convection,MAESTRO_AMR,MAESTRO_CASTRO}.
-\item Convection in massive stars \citep{Gilet:2013}.
-\item Sub-Chandrasekhar white dwarfs \citep{subChandra_I,subChandra_II}.
-In the latter paper we introduced an optional modification to the momentum equation of velocity constraint that conserves total energy, can give results closer to compressible codes of convective dynamics, including low density regions at the surface of a star.
-\item Type I X-ray bursts \citep{XRB_I,XRB_II,XRB_III}.
-In \cite{XRB_I} we discuss the modifications required for implicit thermal conduction, as well as an additional ``volume discrepancy'' modification to the velocity constraint to force the species and enthalpy to evolve in a manner consistent with the background pressure.
-\end{itemize}
-
-In this paper, we characterize the performance of the MAESTROeX algorithm compared to the previous implementation.
-We also develop simpler, alternative temporal integrators, with a longer term goal of developing high-order coupling schemes.
-We also present a new treatment of the base state mapping to Cartesian grids for spherical problsm.
-
-\section{Governing Equations}
-The derivation of the low Mach number hydrodynamics equations is given in Papers I-III.
-We take the standard equations of reacting, compressible flow, and recast the equation
-of state (EOS) as a divergence constraint on the velocity field.  This is done by taking
-the Langrangian derivative of the equation of state for pressure as a function of the
-density, composition, and enthalpy, enforcing the constraint that the pressure is
-a prescribed function of altitutde and time based on the hydrostatic equilibrium condition,
-and substituting the equations for species and enthalpy in the resulting Lagrangian derivative.
-The resulting equations are
-\begin{eqnarray}
-\frac{\partial\Ub}{\partial t} &=& -\Ub\cdot\nabla\Ub  - \frac{1}{\rho}\nabla\pi - \frac{\rho-\rho_0}{\rho} g\eb_r,\label{eq:momentum}\\
-\frac{\partial(\rho X_k)}{\partial t} &=& -\nabla\cdot(\rho X_k\Ub) + \rho\omegadot_k,\label{eq:species}\\
-\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc,\label{eq:enthalpy}
-\end{eqnarray}
-\begin{equation}
-\nabla\cdot(\beta_0\Ub) = \beta_0\left(S - \frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial t}\right).\label{eq:U divergence}
-\end{equation}
-Here $\rho$, $\Ub$, and $h$ are the mass density,
-velocity and specific enthalpy, respectively, and
-$X_k$ are the mass fractions of species $k$ with associated
-production rate $\omegadot_k$.  The species are constrained
-such that $\sum_k X_k = 1$ giving $\rho = \sum_k (\rho X_k)$ and
-\begin{equation}
-\frac{\partial\rho}{\partial t} = -\nabla\cdot(\rho\Ub).
-\end{equation}
-Here $\Hnuc$ is the nuclear energy generation rate per unit mass.
-The pressure is decomposed into a hydrostatic base state
- pressure, $p_0 = p_0(r,t)$, and a dynamic pressure, $\pi = \pi(\xb,t)$, such that 
-$|\pi|/p_0 = \mathcal{O}(M^2)$ (we use $\xb$ to represent the Cartesian coordinate 
-directions of the full state and $r$ to represent the radial coordinate direction for 
-the base state).  We also define a base state density, $\rho_0 = \rho_0(r,t)$, 
-which is in hydrostatic equilibrium with $p_0$, i.e., 
-$\nabla p_0 = -\rho_0 g\eb_r$, where $g=g(r,t)$ is
-the magnitude of the gravitational acceleration and $\eb_r$ is the unit vector in the
-outward radial direction.
-
-Mathematically, equations (\ref{eq:momentum})-(\ref{eq:enthalpy}) must still be closed by the equation of state which we
-express as a divergence constraint on the velocity field (\ref{eq:U divergence}),
-\MarginPar{This is different from the Jacobs 2016 paper where we switched to alt\_energy\_fix.}
-where $\beta_0$ is a density-like variable that carries background stratification, defined as
-\begin{equation}
-\beta_0(r,t) = \rho_0(0,t)\exp\left(\int_0^r\frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial r'}dr'\right),
-\end{equation}
-where $\gammaonebar$ is the lateral average of $\Gamma_1 = d(\log p)/d(\log\rho)$ at constant entropy.
-The expansion term, $S$, incorporates local compressibility effects due to heat release from reactions, compositional changes, and external sources,
-\begin{equation}
-S = -\sigma\sum_k\xi_k\omegadot_k + \frac{1}{\rho p_\rho}\sum_k p_{X_k}\omegadot_k + \sigma\Hnuc,\label{eq:S}
-\end{equation}
-where $p_{X_k} \equiv \left. \partial p / \partial X_k
-\right|_{\rho,T,X_{j,j\ne k}}$, $\xi_k \equiv \left. \partial h /
-\partial X_k \right |_{p,T,X_{j,j\ne k}},
-p_\rho \equiv \left.
-\partial p/\partial \rho \right |_{T, X_k}$, and $\sigma \equiv
-p_T/(\rho c_p p_\rho)$, with $p_T \equiv \left. \partial p / \partial
-T \right|_{\rho, X_k}$ and $c_p \equiv \left.  \partial h / \partial T
-\right|_{p,X_k}$ is the specific heat at constant pressure.
-
-\section{Numerical Algorithm}
-
-\subsection{Spatial Discretization}
-
-(description of base state mapping in planar, and origianl spherical)
-
-\subsection{Exact Base State Mapping} \label{sec:exactbasestate}
-
-\subsection{Temporal Integration Scheme}\label{Sec:Temporal Integration Scheme}
-Previously we adopted an approach where we split the velocity into a base state component, $w_0(r,t)$, 
-and a local velocity $\Ubt(\xb,t)$, so that
-\begin{equation}
-\Ub = \Ubt(\xb,t) + w_0(r,t)\eb_r,
-\end{equation}
-We used $w_0$ to provide an estimate of the base state density evolution over a time step.
-
-Our new temporal integration scheme does not use this splitting, and is much simpler than the scheme from Paper V.
-In this new scheme we use a simpler, predictor-corrector approach to the base state density and pressure that no longer requires the complex algorithm from Paper V, yet still retains the same overall second-order accuracy.
-Also, we now evolve the full-state velocity directly rather than the perturbational velocity, eliminating the need to compute additional forcing terms that arise due to the splitting.
-
-We now describe the new temporal integration scheme, which has been greatly simplified
-from Paper V without formal loss of order of accuracy.
-Note that this temporal integration scheme is valid for the original base state mapping (with constant base state grid spacing), or the new irregularly spaced base state mapping.
-
-At the beginning of each time step we have the cell-centered state,
-$(\Ub,\rho X_k,\rho h,\rho_0,p_0)^n$, and nodal state, $\pi^{n-\myhalf}$.
-At any time, the associated density, composition, and enthalpy can be trivially computed using, e.g.,
-\begin{equation}
-\rho^n = \sum_k(\rho X_k)^n, \quad
-X_k^n = (\rho X_k)^n / \rho^n, \quad
-h^n = (\rho h)^n / \rho^n.
-\end{equation}
-Temperature is computed using the equation of state, e.g.,
-\footnote{As described in Paper V, for planar problems we compute temperature using $h$ instead of $p_0$, since we have successfully developed planar volume discrepancy schemes to effectively couple the enthalpy to the rest of the solution.
-We are still exploring this option for spherical stars.}
-\begin{equation}
-T = T(\rho,p_0,X_k),
-\end{equation}
-and ($\gammaonebar,\beta_0)$ are computed from $(\rho,p_0,X_k)$ (see Appendix A of Paper I and Appendix C of Paper III for details).
-
-The overall flow of the algorithm is a second-order Strang splitting approach for the coupling of the reactions and advection of the thermodynamic variables.  
-We use a predictor-corrector approach within the Strang splitting scheme to achieve second-order accuracy in time.
-We integrate the velocity using a standard second-order projection method to enforce the divergence constraint.
-To summarize:
-\begin{itemize}
-\item In {\bf Step 1} we react the thermodynamic variables over $\Delta t/2$.
-\item In {\bf Steps 2-4} we advect the thermodynamic variables over $\Delta t$.
-Specifically, we compute an estimate for the expansion term, $S$, project the velocities so they satisfy the divergence constraint, and then advect the thermodynamic variables.
-\item In {\bf Step 5} we react the thermodynamic variables over $\Delta t/2$.\footnote{Note that we could skip to the velocity advance in {\bf Steps 10-11}, however the overall scheme would be only first-order in time, so {\bf Steps 6-9} can be thought of as a trapezoidal corrector step.}
-\item In {\bf Steps 6-8} we redo the advection in {\bf Steps 2-4} but are able to use the trapezoidal rule to time-center certain quantities such as $S$, $\rho_0$, etc.
-\item In {\bf Step 9} we redo the reactions from {\bf Step 5} using the improved results for the corrector advection step.
-\item In {\bf Steps 10-11} we update the velocity.
-\end{itemize}
-
-There are a few key numerical modules we use in each time step.
-\begin{itemize}
-\item {\bf Average} computes...
-\item {\bf Enforce HSE} computes the base state pressure, $p_0$, from a base state density, $\rho_0$ by integrating the hydrostatic equilibrium condition in 1D (see equation (A10) in Paper V).
-The base state pressure remains equal to a constant value from the location of a prescribed cutoff density outward for the entire simulation.
-\item {\bf React State} advances the species and enthalpy due to reactions over $\Delta t/2$ by integrating
-\begin{equation}
-\frac{dX_k}{dt} = \dot\omega_k(\rho,X_k,T); \qquad
-\frac{dT}{dt} = \frac{1}{c_p}\left(-\sum_k\xi_k\dot\omega_k + H_{\rm nuc}\right)
-\end{equation}
-The inputs are the species, enthalpy, and base state pressure, and the outpus are the species, enthalpy, reaction rates, and nuclear energy generation rate.
-See Papers III and V for details.
-\end{itemize}
-
-Each time step satisfies the standard advective CFL condition,
-\begin{equation}
-\Delta t = \sigma^{\rm CFL} \min_i(\Delta x / U_i),
-\end{equation}
-where for our simulations we typically use $\sigma^{\rm CFL}\sim 0.5$.
-There are additional constraints on the timestep that are typically much less restrictive than the advective CFL including
-the acceleration due to the buoyancy force (commonly used when the velocity is approximately zero at the start of some simulations) 
-and the local magnitude of the divergence constraint (to prevent too much mass evacuation from a cell in a time step); see Section 3.4 in Paper III for details.
-
-At the beginning of each simulation, we define $(\Ub,\rho X_k,\rho h)$...
-
-\MarginPar{mention cutoff densities}
-
-\MarginPar{AJN: eliminated references to volume discrepancy for now; we'd only use it for planar at the moment, don't want to open up that can of worms here}
-
-\begin{description}
-
-%--------------------------------------------------------------------------
-% STEP 1
-%--------------------------------------------------------------------------
-\item[Step 1] {\em React the full state through the first time interval of $\dt / 2.$}
-
-Call {\bf React State}$[(\rho X_k)^n, (\rho h)^n, p_0^n] \rightarrow [(\rho X_k)^{(1)}, (\rho h)^{(1)}, (\rho \omegadot_k)^{(1)}, (\rho \Hnuc)^{(1)}]$.
-
-%--------------------------------------------------------------------------
-% STEP 2
-%--------------------------------------------------------------------------
-
-\item[Step 2] {\em Compute the provisional time-centered expansion,
-    $S^{\nph,\star}$.}
-
-We compute an estimate for the time-centered expansion term in the velocity
-divergence constraint.  Following \citet{Bell:2004}, we extrapolate
-to the half-time using $S$ at the previous and current
-time steps,
-\begin{equation}
-S^{\nph,\star} = S^n + \frac{\dt^n}{2} \left(\frac{S^n - S^{n-1}}{\dt^{n-1}}\right).
-\end{equation}
-Note that in the first time step we average $S^0$ and $S^1$ from the
-initialization step.
-
-%--------------------------------------------------------------------------
-% STEP 3
-%--------------------------------------------------------------------------
-\item[Step 3] {\em Construct the provisional time-centered advective velocity on
-edges, $\uadvone$.}
-
-The construction of face-centered time-centered states used to discretize the
-advection terms for velocity, species, and enthalpy, are performed using
-a standard multidimensional corner transport upwind approach
-\citep{colella1990multidimensional,saltzman1994unsplit} with piecewise-parabolic
-one-dimensional tracing \citep{colella1984piecewise}.  The full details of this
-Godunov advection approach for all steps in this algorithm are described 
-in Appendix A of \cite{XRB_III}.
-
-In this step, using equation (\ref{eq:momentum}), 
-we compute time-centered edge velocities, $\uadvonedag$, using
-$\Ub^n$.  The $\dagger$ superscript refers to the
-fact that the predicted velocity field does not satisfy the divergence
-constraint.  We then construct $\uadvone$ from $\uadvonedag$
-using a MAC projection.
-We note that $\uadvone$ satisfies the divergence constraint
-\begin{equation}
-\nabla \cdot \left(\beta_0^n \uadvone\right) = \beta_0^n \left[S^{\nph,\star} - \frac{1}{\gammaonebar^np_0^n}\left(\frac{\partial p_0}{\partial t}\right)^{n-\myhalf} \right].\label{eq:div1}
-\end{equation}
-In practice we solve for.
-
-%--------------------------------------------------------------------------
-% STEP 4
-%--------------------------------------------------------------------------
-\item[Step 4] {\em Advect the full state through a time interval of $\dt.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \Alph{enumi}}}
-
-\item Update $(\rho X_k)$ using a discretized version of
-%
-\begin{equation}
-\frac{\partial(\rho X_k)}{\partial t} = -\nabla\cdot(\rho X_k\Ub) + \rho\omegadot_k,
-\end{equation}
-%
-omitting the reaction terms, which were already
-accounted for in {\bf React State}.  The update consists of two steps:
-
-\begin{enumerate}
-\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-\item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\pred}$,
-  for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
-\MarginPar{mention $\rho'$ and $X$ advection; $(\rho h)'$ advection}
-
-\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2),\star}$ using
-\begin{equation}
-(\rho X_k)^{(2),\star} = (\rho X_k)^{(1)}
-  - \dt \left\{ \nabla \cdot \left[ \uadvone (\rho X_k)^{\nph,\pred} \right] \right\},
-\end{equation}
-
-\end{enumerate}
-
-\item Update $\rho_0$ by setting $\rho_0^{n+1,\star} =$ {\bf Average}$(\rho^{(2),\star})$.
-
-\item Update $p_0$ by calling
-{\bf Enforce HSE}$[\rho_0^{n+1,\star}] \rightarrow [p_0^{n+1,\star}]$.
-
-\item Update the enthalpy using a discretized version of equation
-%
-\begin{equation}
-\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc,
-\end{equation}
-%
-again omitting the reaction and heating terms
-since we already accounted for
-them in {\bf React State}.  This equation takes the form:
-\begin{equation}
-\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h) + \frac{\partial p_0}{\partial t} + (\Ub \cdot \eb_r) \frac{\partial p_0}{\partial r}.
-\end{equation}
-For spherical geometry, we solve the
-analytically equivalent form,
-\begin{equation}
-\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h) + \frac{\partial p_0}{\partial t} + \nabla \cdot (\Ub p_0) - p_0 \nabla \cdot \Ub.
-\end{equation}
-The update consists of two steps:
-
-\begin{enumerate}
-\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-\item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\pred},$
-  for the conservative update of $(\rho h)^{(1)}$
-  using using a Godunov approach \citep{XRB_III}.
-
-\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$.
-\begin{description}
-\item For planar geometry,
-\begin{eqnarray}
-(\rho h)^{(2),\star}
-&=& (\rho h)^{(1)} \nonumber \\
-&&- \dt \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \left(p_0^{n+1,*} - p_0^n\right) \nonumber \\
-&& + \dt \left(\uadvone \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^{n}
-\end{eqnarray}
-
-\item For spherical geometry,
-\begin{eqnarray}
-(\rho h)^{(2),\star}
-&=& (\rho h)^{(1)} \nonumber \\
-&&- \dt \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \left(p_0^{n+1,*} - p_0^n\right) \nonumber \\
-&& + \dt \left \{ \nabla \cdot \left (\uadvone p_0^{n+\myhalf} \right ) - p_0^{n+\myhalf} \nabla \cdot \uadvone \right \}
-\end{eqnarray}
-\end{description}
-where we define $p_0^\nph$ as $p_0^\nph = (p_0^n+p_0^{n+1,*})/2$.
-
-\end{enumerate}
-Note that outside the cutoff density region, we compute enthalpy using the EOS.
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 5
-%--------------------------------------------------------------------------
-\item[Step 5] {\em React the full state through a second time interval of $\dt / 2.$}
-
-Call {\bf React State}$[ (\rho X_k)^{(2),\star}, (\rho h)^{(2),\star}, p_0^{n+1,\star}] 
-\rightarrow 
-[ (\rho X_k)^{n+1,\star}, (\rho h)^{n+1,\star}, (\rho \omegadot_k)^{n+1,\star}, (\rho \Hnuc)^{n+1,\star} ].$
-
-%--------------------------------------------------------------------------
-% STEP 6
-%--------------------------------------------------------------------------
-\item[Step 6] {\em Compute the time-centered expansion, $S^{\nph,\star}$.}
-
-First, compute $S^{n+1,\star}$ with
-\begin{equation}
-S^{n+1,\star} =  \left(-\sigma  \sum_k  \xi_k  \omegadot_k  + \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k + \sigma \Hnuc\right)^{n+1,\star}.
-\end{equation}
-  Then, define
-\begin{equation}
- S^{\nph} = \frac{S^n + S^{n+1,\star}}{2},
-\end{equation}
-
-%--------------------------------------------------------------------------
-% STEP 7
-%--------------------------------------------------------------------------
-\item[Step 7] {\em Construct the time-centered advective velocity on edges, $\uadvtwo$.}
-
-The procedure to construct $\uadvtwodag$ is identical to the Godunov procedure
-for computing $\uadvonedag$ in {\bf Step 3}, but uses
-the updated value $S^{\nph}$ rather than $S^{\nph,\star}$.
-We note that $\uadvtwo$ satisfies the divergence constraint
-\begin{equation}
-\nabla \cdot \left(\beta_0^{\nph} \uadvtwo\right) =
-\beta_0^{\nph} \left[S^{\nph} - \frac{1}{\gammaonebar^{n+\myhalf}p_0^{n+\myhalf}}\left(\frac{\partial p_0}{\partial t}\right)^{n+\myhalf}\right],\label{eq:div2}
-\end{equation}
-where
-\begin{equation}
-\beta_0^{\nph} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2},
-\qquad
-\gammaonebar^{\nph} = \frac{ \gammaonebar^n +  \gammaonebar^{n+1,\star} }{2}.
-\end{equation}
-
-%--------------------------------------------------------------------------
-% STEP 8
-%--------------------------------------------------------------------------
-\item[Step 8] {\em Advect the base state and full state through a time interval of $\dt.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \Alph{enumi}}}
-
-\item Update $(\rho X_k)$.  This step is identical to {\bf Step 4A} except we use
-  the updated values $\uadvtwo$ and $\rho_0^{n+1,\star}$ rather than
-  $\uadvone$ and $\rho_0^{n+1,\star}$.  In particular:
-
-\begin{enumerate}
-\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-\item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\pred}$,
-  for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
-
-\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2)}$ using
-\begin{equation}
-(\rho X_k)^{(2)} = (\rho X_k)^{(1)}
-- \dt \left\{ \nabla \cdot \left[\uadvtwo (\rho X_k)^{\nph,\pred} \right] \right\},
-\end{equation}
-
-\end{enumerate}
-
-\item Update $\rho_0$ by setting $\rho_0^{n+1} =$ {\bf Average}$(\rho^{(2)})$.
-
-\item Update $p_0$ by calling
-{\bf Enforce HSE}$[\rho_0^{n+1}] \rightarrow [p_0^{n+1}]$.
-
-\item Update the enthalpy.  This step is identical to {\bf Step 4D} except we use
-  the updated values $\uadvtwo, \rho_0^{n+1}, (\rho h)_0^{n+1}$, and $p_0^{n+\myhalf}$
-  rather than
-  $\uadvone, \rho_0^{n+1,\star}, (\rho h)_0^{n+1,\star}$, and $p_0^n$.
-  In particular:
-
-\begin{enumerate}
-\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-\item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\pred},$
-  for the conservative update of $(\rho h)^{(1)}$ using a Godunov approach \citep{XRB_III}.
-
-\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2)}$.
-\begin{description}
-\item For planar geometry,
-\begin{eqnarray}
-(\rho h)^{(2)}
-&=& (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvtwo (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \left( p_0^{n+1} - p_0^n \right)\nonumber \\
-&& + \dt \left(\uadvtwo \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^\nph,
-\end{eqnarray}
-
-\item For spherical geometry,
-\begin{eqnarray}
-(\rho h)^{(2)}
-&=& (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvtwo (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \left(p_0^{n+1} - p_0^n\right)\nonumber \\
-&& + \dt \left[ \nabla \cdot \left (\uadvtwo p_0^{\nph} \right ) - p_0^{\nph} \nabla \cdot \uadvtwo \right],
-\end{eqnarray}
-\end{description}
-where we now define $p_0^\nph$ as $p_0^\nph = (p_0^n+p_0^{n+1})/2$.
-
-\end{enumerate}
-Note that outside the cutoff density region, we compute enthalpy using the EOS.
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 9
-%--------------------------------------------------------------------------
-\item[Step 9] {\em React the full state through a second time interval of $\dt / 2.$}
-
-Call {\bf React State}$[(\rho X_k)^{(2)},(\rho h)^{(2)}, p_0^{n+1}] \rightarrow [(\rho X_k)^{n+1}, (\rho h)^{n+1}, (\rho \omegadot_k)^{n+1}, (\rho \Hnuc)^{n+1}].$
-
-%--------------------------------------------------------------------------
-% STEP 10
-%--------------------------------------------------------------------------
-\item[Step 10] {\em Define the new time expansion, $S^{n+1}$.}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \Alph{enumi}}}
-\item Define
-\begin{equation}
-  S^{n+1} =  \left(-\sigma  \sum_k  \xi_k \omegadot_k  + \sigma \Hnuc +
-  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  \omegadot_k\right)^{n+1}.
-\end{equation}
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 11
-%--------------------------------------------------------------------------
-\item[Step 11] {\em Update the velocity}.
-
-First, we compute the time-centered edge velocities, $\Ub^{\nph,\pred}$
-using a Godunov approach \citep{XRB_III}. Then, we define
-\begin{equation}
-\rho^\nph = \frac{\rho^n + \rho^{n+1}}{2}, \qquad \rho_0^\nph = \frac{\rho_0^n + \rho_0^{n+1}}{2}.
-\end{equation}
-We update the velocity field $\Ub^n$ to $\Ub^{n+1,\dagger}$ by discretizing
-equation (\ref{eq:momentum}) as
-\begin{equation}
-\Ub^{n+1,\dagger}
-= \Ub^n - \dt \left[\uadvtwo \cdot \nabla \Ub^{\nph,\pred} \right]
- - \dt \left[ \frac{1}{\rho^\nph} \nabla \pi^\nmh + \frac{\left(\rho^\nph-\rho_0^\nph\right)}{\rho^\nph} g^{\nph} \eb_r \right],
-\end{equation}
-Again, the $\dagger$ superscript refers
-to the fact that the updated velocity does not satisfy the divergence
-constraint.
-
-Finally, we use an approximate nodal projection to define $\Ub^{n+1}$
-from $\Ub^{n+1,\dagger},$  such that $\Ub^{n+1}$ approximately
-satisfies the divergence constraint,
-\begin{equation}
-\nabla \cdot \left(\beta_0^{n+1} \Ub^{n+1} \right) = \beta_0^{n+1} \left[ S^{n+1} - \frac{1}{\gammaonebar^{n+1}p_0^{n+1}}\left(\frac{\partial p_0}{\partial t}\right)^{n+\myhalf}\right].\label{eq:div3}
-\end{equation}
-As part of the projection we also define the new-time perturbational pressure,
-$\pi^\nph.$  This projection necessarily differs from the MAC projection used in
-{\bf Step 3} and {\bf Step 7} because the velocities in those steps are defined
-on edges and $\Ub^{n+1}$ is defined at cell centers, requiring different divergence
-and gradient operators.  Details of the approximate projection are given in Paper III.
-
-\end{description}
-
-This completes one step of the algorithm.
-
-To initialize the simulation
-
-%--------------------------------------------------------------------------
-% STEP 0
-%--------------------------------------------------------------------------
-{\em Initialization}
-\sout{
-The initialization step only occurs at the beginning of the simulation.
-The initialization step computes initial values for 
-$\pi^{-\myhalf}, \eta_{\rho}^{-\myhalf}, S^0$, and $S^1$ for use in time stepping routine.  
-We pass through {\bf Step 1} -- {\bf Step 11}, but only keep these values.
-The initial values for $\Ub^0, \rho^0, (\rho h)^0, X_k^0, T^0,
-\rho_0^0, p_0^0$, and $\gammaonebar^0$ are specified from the problem-dependent
-initial conditions.  The initial time step, $\dt^0$, is computed as in
-Paper III.
-}
-
-\subsection{Adaptive Mesh Refinement}
-
-\section{Test Problems}
-
-\subsection{AMR Performance}
-Show 3D reacting bubble rise performance.
-Show 3-level spherical performance (stats won't be as good but explain why: refining a larger percentage of grid)
-
-\subsection{Scaling}
-Scaling with OpenMP
-
-\subsection{White Dwarf Convection}
-white dwarf convection runs with 3 algorithms (original, new temporal, new temporal + irregular base state)
-
-Show 3-level wdconvect
-
-\section{Conclusions and Future Work}
-
-science: rotation, solar physics, MHD
-
-algorithm: SDC \cite{dutt2000spectral}, high-order, multi-implicit \cite{bourlioux2003high}.
-Consider SDC ideas used for terrestrial combustion \cite{pazner2016high,nonaka2018conservative}
-
-implementation: GPUs.  Forward cite next CASTRO paper with GPU strategy.
-
 \appendix
@@
 After evolving the species, we compute this term in the same way as reported in Paper V for use in computing $w_0$.
+
+\MarginPar{mention alt energy fix for projection}
 
@@
 {\bf Make $\mathbf{w_0}$}$[\rho_0,p_0,\partial p_0/\partial t,\gammaonebar,\Sbar] \rightarrow [w_0]$
-
-\acknowledgements
-
-The work at LBNL was supported by the U.S. Department of Energy's Scientific Discovery Through Advanced Computing (SciDAC) program under contract No. DE-AC02-05CH11231.
-This research used resources of the National Energy Research Scientific Computing Center (NERSC), a U.S.
-Department of Energy Office of Science User Facility operated under Contract No.
-DE-AC02-05CH11231.
-
-\bibliographystyle{aasjournal.bst}
-\bibliography{references}
-
-\end{document}

### 2019-03-13 MAESTROeX_I/conclusions.tex
split paper into multiple tex files for multi-user edits

--- 2019-03-13 MAESTROeX_I/paper.tex
@@
-%% This manuscript uses the AASTeX v6.2 LaTeX 2e macros
-%%
-%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls
-%% (Copyright 2000-2015).  See the classfile for details.
-
-%% AASTeX requires revtex4-1.cls (http://publish.aps.org/revtex4/) and
-%% other external packages (latexsym, graphicx, amssymb, longtable, and epsf).
-%% All of these external packages should already be present in the modern TeX
-%% distributions.  If not they can also be obtained at www.ctan.org.
-
-%% The first piece of markup in an AASTeX v6.x document is the \documentclass
-%% command. LaTeX will ignore any data that comes before this command. The
-%% documentclass can take an optional argument to modify the output style.
-%% The command below calls the preprint style  which will produce a tightly
-%% typeset, one-column, single-spaced document.  It is the default and thus
-%% does not need to be explicitly stated.
-%%
-%%
-%% using aastex version 6.2
-\documentclass{aastex62}
-
-%% The default is a single spaced, 10 point font, single spaced article.
-%% There are 5 other style options available via an optional argument. They
-%% can be envoked like this:
-%%
-%% \documentclass[argument]{aastex62}
-%%
-%% where the layout options are:
-%%
-%%  twocolumn   : two text columns, 10 point font, single spaced article.
-%%                This is the most compact and represent the final published
-%%                derived PDF copy of the accepted manuscript from the publisher
-%%  manuscript  : one text column, 12 point font, double spaced article.
-%%  preprint    : one text column, 12 point font, single spaced article.
-%%  preprint2   : two text columns, 12 point font, single spaced article.
-%%  modern      : a stylish, single text column, 12 point font, article with
-%% 		  wider left and right margins. This uses the Daniel
-%% 		  Foreman-Mackey and David Hogg design.
-%%  RNAAS       : Preferred style for Research Notes which are by design
-%%                lacking an abstract and brief. DO NOT use \begin{abstract}
-%%                and \end{abstract} with this style.
-%%
-%% Note that you can submit to the AAS Journals in any of these 6 styles.
-%%
-%% There are other optional arguments one can envoke to allow other stylistic
-%% actions. The available options are:
-%%
-%%  astrosymb    : Loads Astrosymb font and define \astrocommands.
-%%  tighten      : Makes baselineskip slightly smaller, only works with
-%%                 the twocolumn substyle.
-%%  times        : uses times font instead of the default
-%%  linenumbers  : turn on lineno package.
-%%  trackchanges : required to see the revision mark up and print its output
-%%  longauthor   : Do not use the more compressed footnote style (default) for
-%%                 the author/collaboration/affiliations. Instead print all
-%%                 affiliation information after each name. Creates a much
-%%                 long author list but may be desirable for short author papers
-%%
-%% these can be used in any combination, e.g.
-%%
-%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex62}
-%%
-%% AASTeX v6.* now includes \hyperref support. While we have built in specific
-%% defaults into the classfile you can manually override them with the
-%% \hypersetup command. For example,
-%%
-%%\hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
-%%
-%% will change the color of the internal links to red, the links to the
-%% bibliography to green, the file links to cyan, and the external links to
-%% magenta. Additional information on \hyperref options can be found here:
-%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
-%%
-%% If you want to create your own macros, you can do so
-%% using \newcommand. Your macros should appear before
-%% the \begin{document} command.
-%%
-\newcommand{\vdag}{(v)^\dagger}
-\newcommand\aastex{AAS\TeX}
-\newcommand\latex{La\TeX}
-
-% for non-stacked fractions
-\newcommand{\sfrac}[2]{\mathchoice
-  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
-  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
-  {\kern0em\raise.5ex\hbox{\the\scriptscriptfont0 #1}\kern-.2em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptscriptfont0 #2}}
-  {#1\!/#2}}
-
-\newcommand{\myhalf}{\sfrac{1}{2}}
-\newcommand{\thalf}{\sfrac{3}{2}}
-
-\newcommand{\eb}{{\bf{e}}}
-\newcommand{\Ub}{{\bf{U}}}
-\newcommand{\Ubt}{\widetilde{\Ub}}
-\newcommand{\Vb}{{\bf{V}}}
-\newcommand{\xb}{{\bf{x}}}
-
-\newcommand{\dr}{\Delta r}
-\newcommand{\dt}{\Delta t}
-
-\newcommand{\etarho}{\eta_\rho}
-\newcommand{\gammaonebar}{\overline{\Gamma}_1}
-\newcommand{\Hnuc}{H_{\rm nuc}}
-\newcommand{\omegadot}{\dot\omega}
-\newcommand{\pred}{{\rm pred}}
-\newcommand{\Sbar}{\overline{S}}
-
-\newcommand{\inp}{\mathrm{in}}
-\newcommand{\outp}{\mathrm{out}}
-\newcommand{\nph}{{n+\myhalf}}
-\newcommand{\nmh}{{n-\myhalf}}
-\newcommand{\ow}{\overline{w_0}}
-\newcommand{\dw}{\delta w_0}
-\newcommand{\uadvone}{\Ub^{\mathrm{ADV},\star}}
-\newcommand{\uadvonedag}{\Ub^{\mathrm{ADV},\dagger,\star}}
-\newcommand{\uadvtwo}{\Ub^{\mathrm{ADV}}}
-\newcommand{\uadvtwodag}{\Ub^{\mathrm{ADV},\dagger}}
-\newcommand{\gcc}{\mathrm{g~cm^{-3} }}
-
-% for the red MarginPars
-\usepackage{color}
-% make the MarginPars look pretty
-\setlength{\marginparwidth}{0.5in}
-\newcommand{\MarginPar}[1]{\marginpar{\vskip-\baselineskip\raggedright\tiny\sffamily
-\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}
-
-%% Reintroduced the \received and \accepted commands from AASTeX v5.2
-\received{XXX X, XXXX}
-\revised{XXX X, XXXX}
-\accepted{XXX X, XXXX}
-%% Command to document which AAS Journal the manuscript was submitted to.
-%% Adds "Submitted to " the arguement.
-\submitjournal{ApJ}
-
-%% Mark up commands to limit the number of authors on the front page.
-%% Note that in AASTeX v6.2 a \collaboration call (see below) counts as
-%% an author in this case.
-%
-%\AuthorCollaborationLimit=3
-%
-%% Will only show Schwarz, Muench and "the AAS Journals Data Scientist
-%% collaboration" on the front page of this example manuscript.
-%%
-%% Note that all of the author will be shown in the published article.
-%% This feature is meant to be used prior to acceptance to make the
-%% front end of a long author article more manageable. Please do not use
-%% this functionality for manuscripts with less than 20 authors. Conversely,
-%% please do use this when the number of authors exceeds 40.
-%%
-%% Use \allauthors at the manuscript end to show the full author list.
-%% This command should only be used with \AuthorCollaborationLimit is used.
-
-%% The following command can be used to set the latex table counters.  It
-%% is needed in this document because it uses a mix of latex tabular and
-%% AASTeX deluxetables.  In general it should not be needed.
-%\setcounter{table}{1}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% The following section outlines numerous optional output that
-%% can be displayed in the front matter or as running meta-data.
-%%
-%% If you wish, you may supply running head information, although
-%% this information may be modified by the editorial offices.
-\shorttitle{MAESTROeX Low Mach Number Astrophysics}
-\shortauthors{Fan et al.}
-%%
-%% You can add a light gray and diagonal water-mark to the first page
-%% with this command:
-% \watermark{text}
-%% where "text", e.g. DRAFT, is the text to appear.  If the text is
-%% long you can control the water-mark size with:
-%  \setwatermarkfontsize{dimension}
-%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% This is the end of the preamble.  Indicate the beginning of the
-%% manuscript itself with \begin{document}.
-
-\begin{document}
-
-\title{MAESTROeX: A Massively Parallel Low Mach Number Astrophysical Solver}
-
-%% LaTeX will automatically break titles if they run longer than
-%% one line. However, you may use \\ to force a line break if
-%% you desire. In v6.2 you can include a footnote in the title.
-
-%% A significant change from earlier AASTEX versions is in the structure for
-%% calling author and affilations. The change was necessary to implement
-%% autoindexing of affilations which prior was a manual process that could
-%% easily be tedious in large author manuscripts.
-%%
-%% The \author command is the same as before except it now takes an optional
-%% arguement which is the 16 digit ORCID. The syntax is:
-%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
-%%
-%% This will hyperlink the author name to the author's ORCID page. Note that
-%% during compilation, LaTeX will do some limited checking of the format of
-%% the ID to make sure it is valid.
-%%
-%% Use \affiliation for affiliation information. The old \affil is now aliased
-%% to \affiliation. AASTeX v6.2 will automatically index these in the header.
-%% When a duplicate is found its index will be the same as its previous entry.
-%%
-%% Note that \altaffilmark and \altaffiltext have been removed and thus
-%% can not be used to document secondary affiliations. If they are used latex
-%% will issue a specific error message and quit. Please use multiple
-%% \affiliation calls for to document more than one affiliation.
-%%
-%% The new \altaffiliation can be used to indicate some secondary information
-%% such as fellowships. This command produces a non-numeric footnote that is
-%% set away from the numeric \affiliation footnotes.  NOTE that if an
-%% \altaffiliation command is used it must come BEFORE the \affiliation call,
-%% right after the \author command, in order to place the footnotes in
-%% the proper location.
-%%
-%% Use \email to set provide email addresses. Each \email will appear on its
-%% own line so you can put multiple email address in one \email call. A new
-%% \correspondingauthor command is available in V6.2 to identify the
-%% corresponding author of the manuscript. It is the author's responsibility
-%% to make sure this name is also in the author list.
-%%
-%% While authors can be grouped inside the same \author and \affiliation
-%% commands it is better to have a single author for each. This allows for
-%% one to exploit all the new benefits and should make book-keeping easier.
-%%
-%% If done correctly the peer review system will be able to
-%% automatically put the author and affiliation information from the manuscript
-%% and save the corresponding author the trouble of entering it by hand.
-
-\correspondingauthor{Doreen Fan; DFan@lbl.gov}
-
-\author[0000-0002-3246-4315]{Duoming Fan}
-\affil{Lawrence Berkelely National Laboratory \\
-Center for Computational Sciences and Engineering \\
-One Cyclotron Road, MS 50A-3111 \\
-Berkeley, CA 94720, USA}
-
-\author[0000-0003-1791-0265]{Andrew Nonaka}
-\affil{Lawrence Berkelely National Laboratory \\
-Center for Computational Sciences and Engineering \\
-One Cyclotron Road, MS 50A-3111 \\
-Berkeley, CA 94720, USA}
-
-\author[0000-0003-2103-312X]{Ann S. Almgren}
-\affil{Lawrence Berkelely National Laboratory \\
-Center for Computational Sciences and Engineering \\
-One Cyclotron Road, MS 50A-3111 \\
-Berkeley, CA 94720, USA}
-
-\author[0000-0002-1530-781X]{Alice Harpole}
-\affil{Stony Brook University \\
-Department of Physics and Astronomy \\
-Stony Brook, NY 11794-3800, USA}
-
-\author[0000-0001-8401-030X]{Michael Zingale}
-\affil{Stony Brook University \\
-Department of Physics and Astronomy \\
-Stony Brook, NY 11794-3800, USA}
-
-%% Note that the \and command from previous versions of AASTeX is now
-%% depreciated in this version as it is no longer necessary. AASTeX
-%% automatically takes care of all commas and "and"s between authors names.
-
-%% AASTeX 6.2 has the new \collaboration and \nocollaboration commands to
-%% provide the collaboration status of a group of authors. These commands
-%% can be used either before or after the list of corresponding authors. The
-%% argument for \collaboration is the collaboration identifier. Authors are
-%% encouraged to surround collaboration identifiers with ()s. The
-%% \nocollaboration command takes no argument and exists to indicate that
-%% the nearby authors are not part of surrounding collaborations.
-
-%% Mark off the abstract in the ``abstract'' environment.
-\begin{abstract}
-We present MAESTROeX, a massively parallel solver for low Mach number astrophysical flows.
-Our model equations allow for long-time integration for highly subsonic flows compared to compressible approaches.
-The code leverages the new AMReX framework for block-structured adaptive mesh refinement calculations, and also provides additional algorithmic features from the original MAESTRO code.
-We present a new temporal integration option that is much simpler than our previous approach while retaining the same order order of accuracy.
-We also provide a more accurate spatial discretization for mapping the base state onto the full state Cartesian grid.
-Using our previous studies on the convective phase of single-degenerate progenitor models of Type Ia supernovae, we characterize the performance of the code and validate the new algorithmic features.
-\end{abstract}
-
-%% Keywords should appear after the \end{abstract} command.
-%% See the online documentation for the full list of available subject
-%% keywords and the rules for their use.
-\keywords{convection, hydrodynamics, methods: numerical, nuclear reactions, nucleosynthesis, abundances, supernovae: general}
-
-%% From the front matter, we move on to the body of the paper.
-%% Sections are demarcated by \section and \subsection, respectively.
-%% Observe the use of the LaTeX \label
-%% command after the \subsection to give a symbolic KEY to the
-%% subsection for cross-referencing in a \ref command.
-%% You can use LaTeX's \ref and \label commands to keep track of
-%% cross-references to sections, equations, tables, and figures.
-%% That way, if you change the order of any elements, LaTeX will
-%% automatically renumber them.
-%%
-%% We recommend that authors also use the natbib \citep
-%% and \citet commands to identify citations.  The citations are
-%% tied to the reference list via symbolic KEYs. The KEY corresponds
-%% to the KEY in the \bibitem in the reference list below.
-
-\section{Introduction} \label{sec:intro}
-
-TODO list:
-\begin{itemize}
-\item Weak scaling tests of original algorithm, comparing MAESTRO to MAESTROeX.  Use cori knl, 4 MPI per node, 16 threads per MPI
-\item Run at effective $512^3$ resolution (maybe to ignition?...
-need to see how max T trends go) (1) classic MAESTRO, (2) MAESTROeX with original algorithm, (3) MAESTROeX with new temporal integrator, (4) MAESTROeX with new temporal integrator and irregular dr, (5) MAESTROeX with original algorithm and 3-levels of AMR, (6) MAESTROeX with new temporal integrator and 3 levels of AMR.
-\item (perhaps) a demonstration of what goes wrong when you don't split the velocity dynamics in the projection
-\end{itemize}
-
-Many astrophysical flows are highly subsonic; sound waves carry sufficiently little energy that they do not significantly affect the convective dynamics of the system.
-In many of these flows, modeling long-time convective dynamics are of interest, and numerical approaches based on compressible hydrodynamics are intractable, even on modern supercomputers.
-One approach to this problem is to use low Mach number models.
-In a low Mach number approach, sound waves are eliminated from the governing equations while retaining compressibilitiy effects due to, e.g., nuclear energy release, compositional changes, and thermal diffusion.
-The resulting model can be numerically integrated with much larger time steps than a compressible model.
-
-Previously, we have developed the low Mach number astrophysical solver, MAESTRO.
-Low Mach number models have been developed for a variety of contexts including combustion \citep{day2000numerical}, atmospherics \citep{duarte2015low}, and elastic solids \citep{abbate2017all}.
-The low Mach number model in MAESTRO is unique in that it is specifically designed for astrophysical settings with significant atmospheric stratification.
-Central to the algorithm is a stratified background (or base) state density and pressure held in hydrostatic equilibrium, and also vary as a function of altitude and time.
-MAESTRO is suitable for for spherical stars, as well as planar simulations of local dynamics.
-
-The key numerical developments of the original MAESTRO algorithm are presented in a series of papers which we refer to as Papers I-V:
-\begin{itemize}
-\item In Paper I \citep{MAESTRO_I}, we derive the low Mach number equation set from the fully compressible equations.
-\item In Paper II \citep{MAESTRO_II}, we incorporate the effects atmospheric expansion through the use of a time-dependent background state.
-\item In Paper III \citep{MAESTRO_III}, we incorporate reactions and the associated coupling to the hydrodynamics.
-\item In Paper IV \citep{MAESTRO_IV}, we describe our treatment of spherical stars in a three-dimensional Cartesian geometry.
-\item In Paper V \citep{MAESTRO_V}, we describe the use of block-structured adaptive mesh refinement to focus spatial resolution in regions of interest.
-\end{itemize}
-
-Since then, there have been many scientific investigations using MAESTRO, which include additional algorithmic enhancements.  Topics include:
-\begin{itemize}
-\item The convective phase preceding Chandrasekhar mass models for type Ia supernovae \citep{MAESTRO_convection,MAESTRO_AMR,MAESTRO_CASTRO}.
-\item Convection in massive stars \citep{Gilet:2013}.
-\item Sub-Chandrasekhar white dwarfs \citep{subChandra_I,subChandra_II}.
-In the latter paper we introduced an optional modification to the momentum equation of velocity constraint that conserves total energy, can give results closer to compressible codes of convective dynamics, including low density regions at the surface of a star.
-\item Type I X-ray bursts \citep{XRB_I,XRB_II,XRB_III}.
-In \cite{XRB_I} we discuss the modifications required for implicit thermal conduction, as well as an additional ``volume discrepancy'' modification to the velocity constraint to force the species and enthalpy to evolve in a manner consistent with the background pressure.
-\end{itemize}
-
-In this paper, we characterize the performance of the MAESTROeX algorithm compared to the previous implementation.
-We also develop simpler, alternative temporal integrators, with a longer term goal of developing high-order coupling schemes.
-We also present a new treatment of the base state mapping to Cartesian grids for spherical problsm.
-
-\section{Governing Equations}
-The derivation of the low Mach number hydrodynamics equations is given in Papers I-III.
-We take the standard equations of reacting, compressible flow, and recast the equation
-of state (EOS) as a divergence constraint on the velocity field.  This is done by taking
-the Langrangian derivative of the equation of state for pressure as a function of the
-density, composition, and enthalpy, enforcing the constraint that the pressure is
-a prescribed function of altitutde and time based on the hydrostatic equilibrium condition,
-and substituting the equations for species and enthalpy in the resulting Lagrangian derivative.
-The resulting equations are
-\begin{eqnarray}
-\frac{\partial\Ub}{\partial t} &=& -\Ub\cdot\nabla\Ub  - \frac{1}{\rho}\nabla\pi - \frac{\rho-\rho_0}{\rho} g\eb_r,\label{eq:momentum}\\
-\frac{\partial(\rho X_k)}{\partial t} &=& -\nabla\cdot(\rho X_k\Ub) + \rho\omegadot_k,\label{eq:species}\\
-\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc,\label{eq:enthalpy}
-\end{eqnarray}
-\begin{equation}
-\nabla\cdot(\beta_0\Ub) = \beta_0\left(S - \frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial t}\right).\label{eq:U divergence}
-\end{equation}
-Here $\rho$, $\Ub$, and $h$ are the mass density,
-velocity and specific enthalpy, respectively, and
-$X_k$ are the mass fractions of species $k$ with associated
-production rate $\omegadot_k$.  The species are constrained
-such that $\sum_k X_k = 1$ giving $\rho = \sum_k (\rho X_k)$ and
-\begin{equation}
-\frac{\partial\rho}{\partial t} = -\nabla\cdot(\rho\Ub).
-\end{equation}
-Here $\Hnuc$ is the nuclear energy generation rate per unit mass.
-The pressure is decomposed into a hydrostatic base state
- pressure, $p_0 = p_0(r,t)$, and a dynamic pressure, $\pi = \pi(\xb,t)$, such that 
-$|\pi|/p_0 = \mathcal{O}(M^2)$ (we use $\xb$ to represent the Cartesian coordinate 
-directions of the full state and $r$ to represent the radial coordinate direction for 
-the base state).  We also define a base state density, $\rho_0 = \rho_0(r,t)$, 
-which is in hydrostatic equilibrium with $p_0$, i.e., 
-$\nabla p_0 = -\rho_0 g\eb_r$, where $g=g(r,t)$ is
-the magnitude of the gravitational acceleration and $\eb_r$ is the unit vector in the
-outward radial direction.
-
-Mathematically, equations (\ref{eq:momentum})-(\ref{eq:enthalpy}) must still be closed by the equation of state which we
-express as a divergence constraint on the velocity field (\ref{eq:U divergence}),
-\MarginPar{This is different from the Jacobs 2016 paper where we switched to alt\_energy\_fix.}
-where $\beta_0$ is a density-like variable that carries background stratification, defined as
-\begin{equation}
-\beta_0(r,t) = \rho_0(0,t)\exp\left(\int_0^r\frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial r'}dr'\right),
-\end{equation}
-where $\gammaonebar$ is the lateral average of $\Gamma_1 = d(\log p)/d(\log\rho)$ at constant entropy.
-The expansion term, $S$, incorporates local compressibility effects due to heat release from reactions, compositional changes, and external sources,
-\begin{equation}
-S = -\sigma\sum_k\xi_k\omegadot_k + \frac{1}{\rho p_\rho}\sum_k p_{X_k}\omegadot_k + \sigma\Hnuc,\label{eq:S}
-\end{equation}
-where $p_{X_k} \equiv \left. \partial p / \partial X_k
-\right|_{\rho,T,X_{j,j\ne k}}$, $\xi_k \equiv \left. \partial h /
-\partial X_k \right |_{p,T,X_{j,j\ne k}},
-p_\rho \equiv \left.
-\partial p/\partial \rho \right |_{T, X_k}$, and $\sigma \equiv
-p_T/(\rho c_p p_\rho)$, with $p_T \equiv \left. \partial p / \partial
-T \right|_{\rho, X_k}$ and $c_p \equiv \left.  \partial h / \partial T
-\right|_{p,X_k}$ is the specific heat at constant pressure.
-
-Previously we adopted an approach where we split the velocity into a base state component, $w_0(r,t)$, 
-and a local velocity $\Ubt(\xb,t)$, so that
-\begin{equation}
-\Ub = \Ubt(\xb,t) + w_0(r,t)\eb_r,
-\end{equation}
-We used $w_0$ to provide an estimate of the base state density evolution over a time step.
-In our new simplified approach, we instead use a predictor-corrector approach for evolving the base state.
-
-\section{Numerical Algorithm}
-
-\subsection{Spatial Discretization}
-
-(description of base state mapping in planar, and origianl spherical)
-
-\subsection{Exact Base State Mapping} \label{sec:exactbasestate}
-
-\subsection{Temporal Integration Scheme}\label{Sec:Temporal Integration Scheme}
-Our new temporal integration scheme is much simpler than the scheme from Paper V.
-The primary reason is that we no longer split the velocity field into a perturbational and base state components.  
-Previously, we used the base state velocity to predict the expansion of the base state.  
-In this new scheme we use a simpler, predictor-corrector approach to the base state density and pressure that no longer requires the complex algorithm from Paper V, yet still retains the same overall second-order accuracy.
-
-We now describe the new temporal integration scheme, which has been greatly simplified
-from Paper V without formal loss of order of accuracy.
-Note that this temporal integration scheme is valid for the original base state mapping (with constant base state grid spacing), or the new irregularly spaced base state mapping.
-
-At the beginning of each time step we have the cell-centered state,
-$(\Ub,\rho X_k,\rho h,\rho_0,p_0)^n$, and nodal state, $\pi^{n-\myhalf}$.
-At any time, the associated density, composition, and enthalpy can be trivially computed using, e.g.,
-\begin{equation}
-\rho^n = \sum_k(\rho X_k)^n, \quad
-X_k^n = (\rho X_k)^n / \rho^n, \quad
-h^n = (\rho h)^n / \rho^n.
-\end{equation}
-Temperature is computed using the equation of state, e.g.,
-\footnote{As described in Paper V, for planar problems we compute temperature using $h$ instead of $p_0$, since we have successfully developed planar volume discrepancy schemes to effectively couple the enthalpy to the rest of the solution.
-We are still exploring this option for spherical stars.}
-\begin{equation}
-T = T(\rho,p_0,X_k),
-\end{equation}
-and ($\gammaonebar,\beta_0)$ are computed from $(\rho,p_0,X_k)$ (see Appendix A of Paper I and Appendix C of Paper III for details).
-
-The overall flow of the algorithm is a second-order Strang splitting approach for the coupling of the reactions and advection of the thermodynamic variables.  
-We use a predictor-corrector approach within the Strang splitting scheme to achieve second-order accuracy in time.
-We integrate the velocity using a standard second-order projection method to enforce the divergence constraint.
-To summarize:
-\begin{itemize}
-\item In {\bf Step 1} we react the thermodynamic variables over $\Delta t/2$.
-\item In {\bf Steps 2-4} we advect the thermodynamic variables over $\Delta t$.
-Specifically, we compute an estimate for the expansion term, $S$, project the velocities so they satisfy the divergence constraint, and then advect the thermodynamic variables.
-\item In {\bf Step 5} we react the thermodynamic variables over $\Delta t/2$.\footnote{Note that we could skip to the velocity advance in {\bf Steps 10-11}, however the overall scheme would be only first-order in time, so {\bf Steps 6-9} can be thought of as a trapezoidal corrector step.}
-\item In {\bf Steps 6-8} we redo the advection in {\bf Steps 2-4} but are able to use the trapezoidal rule to time-center certain quantities such as $S$, $\rho_0$, etc.
-\item In {\bf Step 9} we redo the reactions from {\bf Step 5} using the improved results for the corrector advection step.
-\item In {\bf Steps 10-11} we update the velocity.
-\end{itemize}
-
-There are a few key numerical modules we use in each time step.
-\begin{itemize}
-\item {\bf Average} computes...
-\item {\bf Enforce HSE} computes the base state pressure, $p_0$, from a base state density, $\rho_0$ by integrating the hydrostatic equilibrium condition in 1D (see equation (A10) in Paper V).
-The base state pressure remains equal to a constant value from the location of a prescribed cutoff density outward for the entire simulation.
-\item {\bf React State} advances the species and enthalpy due to reactions over $\Delta t/2$ by integrating
-\begin{equation}
-\frac{dX_k}{dt} = \dot\omega_k(\rho,X_k,T); \qquad
-\frac{dT}{dt} = \frac{1}{c_p}\left(-\sum_k\xi_k\dot\omega_k + H_{\rm nuc}\right)
-\end{equation}
-The inputs are the species, enthalpy, and base state pressure, and the outpus are the species, enthalpy, reaction rates, and nuclear energy generation rate.
-See Papers III and V for details.
-\end{itemize}
-
-Each time step satisfies the standard advective CFL condition,
-\begin{equation}
-\Delta t = \sigma^{\rm CFL} \min_i(\Delta x / U_i),
-\end{equation}
-where for our simulations we typically use $\sigma^{\rm CFL}\sim 0.5$.
-There are additional constraints on the time involving the buoyancy force (commonly used when the velocity is approximately zero at the start of some simulations) and the divergence constraint (to prevent too much mass evacuation from a cell in a time step); see Section 3.4 in Paper III for details.
-
-At the beginning of each simulation, we define $(\Ub,\rho X_k,\rho h)$...
-
-\MarginPar{mention cutoff densities}
-
-\MarginPar{AJN: eliminated references to volume discrepancy for now; we'd only use it for planar at the moment, don't want to open up that can of worms here}
-
-\begin{description}
-
-%--------------------------------------------------------------------------
-% STEP 1
-%--------------------------------------------------------------------------
-\item[Step 1] {\em React the full state through the first time interval of $\dt / 2.$}
-
-Call {\bf React State}$[(\rho X_k)^n, (\rho h)^n, p_0^n] \rightarrow [(\rho X_k)^{(1)}, (\rho h)^{(1)}, (\rho \omegadot_k)^{(1)}, (\rho \Hnuc)^{(1)}]$.
-
-%--------------------------------------------------------------------------
-% STEP 2
-%--------------------------------------------------------------------------
-
-\item[Step 2] {\em Compute the provisional time-centered expansion,
-    $S^{\nph,\star}$.}
-
-We compute an estimate for the time-centered expansion term in the velocity
-divergence constraint.  Following \citet{Bell:2004}, we extrapolate
-to the half-time using $S$ at the previous and current
-time steps,
-\begin{equation}
-S^{\nph,\star} = S^n + \frac{\dt^n}{2} \left(\frac{S^n - S^{n-1}}{\dt^{n-1}}\right).
-\end{equation}
-Note that in the first time step we average $S^0$ and $S^1$ from the
-initialization step.
-
-%--------------------------------------------------------------------------
-% STEP 3
-%--------------------------------------------------------------------------
-\item[Step 3] {\em Construct the provisional time-centered advective velocity on
-edges, $\uadvone$.}
-
-The construction of face-centered time-centered states used to discretize the
-advection terms for velocity, species, and enthalpy, are performed using
-a standard multidimensional corner transport upwind approach
-\citep{colella1990multidimensional,saltzman1994unsplit} with piecewise-parabolic
-one-dimensional tracing \citep{colella1984piecewise}.  The full details of this
-Godunov advection approach for all steps in this algorithm are described 
-in Appendix A of \cite{XRB_III}.
-
-In this step, using equation (\ref{eq:momentum}), 
-we compute time-centered edge velocities, $\uadvonedag$, using
-$\Ub^n$.  The $\dagger$ superscript refers to the
-fact that the predicted velocity field does not satisfy the divergence
-constraint.  We then construct $\uadvone$ from $\uadvonedag$
-using a MAC projection.
-We note that $\uadvone$ satisfies the divergence constraint
-\begin{equation}
-\nabla \cdot \left(\beta_0^n \uadvone\right) = \beta_0^n \left[S^{\nph,\star} - \frac{1}{\gammaonebar^np_0^n}\left(\frac{\partial p_0}{\partial t}\right)^{n-\myhalf} \right].\label{eq:div1}
-\end{equation}
-In practice we solve for.
-
-%--------------------------------------------------------------------------
-% STEP 4
-%--------------------------------------------------------------------------
-\item[Step 4] {\em Advect the full state through a time interval of $\dt.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \Alph{enumi}}}
-
-\item Update $(\rho X_k)$ using a discretized version of
-%
-\begin{equation}
-\frac{\partial(\rho X_k)}{\partial t} = -\nabla\cdot(\rho X_k\Ub) + \rho\omegadot_k,
-\end{equation}
-%
-omitting the reaction terms, which were already
-accounted for in {\bf React State}.  The update consists of two steps:
-
-\begin{enumerate}
-\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-\item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\pred}$,
-  for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
-\MarginPar{mention $\rho'$ and $X$ advection; $(\rho h)'$ advection}
-
-\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2),\star}$ using
-\begin{equation}
-(\rho X_k)^{(2),\star} = (\rho X_k)^{(1)}
-  - \dt \left\{ \nabla \cdot \left[ \uadvone (\rho X_k)^{\nph,\pred} \right] \right\},
-\end{equation}
-
-\end{enumerate}
-
-\item Update $\rho_0$ by setting $\rho_0^{n+1,\star} =$ {\bf Average}$(\rho^{(2),\star})$.
-
-\item Update $p_0$ by calling
-{\bf Enforce HSE}$[\rho_0^{n+1,\star}] \rightarrow [p_0^{n+1,\star}]$.
-
-\item Update the enthalpy using a discretized version of equation
-%
-\begin{equation}
-\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc,
-\end{equation}
-%
-again omitting the reaction and heating terms
-since we already accounted for
-them in {\bf React State}.  This equation takes the form:
-\begin{equation}
-\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h) + \frac{\partial p_0}{\partial t} + (\Ub \cdot \eb_r) \frac{\partial p_0}{\partial r}.
-\end{equation}
-For spherical geometry, we solve the
-analytically equivalent form,
-\begin{equation}
-\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h) + \frac{\partial p_0}{\partial t} + \nabla \cdot (\Ub p_0) - p_0 \nabla \cdot \Ub.
-\end{equation}
-The update consists of two steps:
-
-\begin{enumerate}
-\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-\item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\pred},$
-  for the conservative update of $(\rho h)^{(1)}$
-  using using a Godunov approach \citep{XRB_III}.
-
-\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$.
-\begin{description}
-\item For planar geometry,
-\begin{eqnarray}
-(\rho h)^{(2),\star}
-&=& (\rho h)^{(1)} \nonumber \\
-&&- \dt \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \left(p_0^{n+1,*} - p_0^n\right) \nonumber \\
-&& + \dt \left(\uadvone \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^{n}
-\end{eqnarray}
-
-\item For spherical geometry,
-\begin{eqnarray}
-(\rho h)^{(2),\star}
-&=& (\rho h)^{(1)} \nonumber \\
-&&- \dt \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \left(p_0^{n+1,*} - p_0^n\right) \nonumber \\
-&& + \dt \left \{ \nabla \cdot \left (\uadvone p_0^{n+\myhalf} \right ) - p_0^{n+\myhalf} \nabla \cdot \uadvone \right \}
-\end{eqnarray}
-\end{description}
-where we define $p_0^\nph$ as $p_0^\nph = (p_0^n+p_0^{n+1,*})/2$.
-
-\end{enumerate}
-Note that outside the cutoff density region, we compute enthalpy using the EOS.
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 5
-%--------------------------------------------------------------------------
-\item[Step 5] {\em React the full state through a second time interval of $\dt / 2.$}
-
-Call {\bf React State}$[ (\rho X_k)^{(2),\star}, (\rho h)^{(2),\star}, p_0^{n+1,\star}] 
-\rightarrow 
-[ (\rho X_k)^{n+1,\star}, (\rho h)^{n+1,\star}, (\rho \omegadot_k)^{n+1,\star}, (\rho \Hnuc)^{n+1,\star} ].$
-
-%--------------------------------------------------------------------------
-% STEP 6
-%--------------------------------------------------------------------------
-\item[Step 6] {\em Compute the time-centered expansion, $S^{\nph,\star}$.}
-
-First, compute $S^{n+1,\star}$ with
-\begin{equation}
-S^{n+1,\star} =  \left(-\sigma  \sum_k  \xi_k  \omegadot_k  + \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k + \sigma \Hnuc\right)^{n+1,\star}.
-\end{equation}
-  Then, define
-\begin{equation}
- S^{\nph} = \frac{S^n + S^{n+1,\star}}{2},
-\end{equation}
-
-%--------------------------------------------------------------------------
-% STEP 7
-%--------------------------------------------------------------------------
-\item[Step 7] {\em Construct the time-centered advective velocity on edges, $\uadvtwo$.}
-
-The procedure to construct $\uadvtwodag$ is identical to the Godunov procedure
-for computing $\uadvonedag$ in {\bf Step 3}, but uses
-the updated value $S^{\nph}$ rather than $S^{\nph,\star}$.
-We note that $\uadvtwo$ satisfies the divergence constraint
-\begin{equation}
-\nabla \cdot \left(\beta_0^{\nph} \uadvtwo\right) =
-\beta_0^{\nph} \left[S^{\nph} - \frac{1}{\gammaonebar^{n+\myhalf}p_0^{n+\myhalf}}\left(\frac{\partial p_0}{\partial t}\right)^{n+\myhalf}\right],\label{eq:div2}
-\end{equation}
-where
-\begin{equation}
-\beta_0^{\nph} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2},
-\qquad
-\gammaonebar^{\nph} = \frac{ \gammaonebar^n +  \gammaonebar^{n+1,\star} }{2}.
-\end{equation}
-
-%--------------------------------------------------------------------------
-% STEP 8
-%--------------------------------------------------------------------------
-\item[Step 8] {\em Advect the base state and full state through a time interval of $\dt.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \Alph{enumi}}}
-
-\item Update $(\rho X_k)$.  This step is identical to {\bf Step 4A} except we use
-  the updated values $\uadvtwo$ and $\rho_0^{n+1,\star}$ rather than
-  $\uadvone$ and $\rho_0^{n+1,\star}$.  In particular:
-
-\begin{enumerate}
-\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-\item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\pred}$,
-  for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
-
-\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2)}$ using
-\begin{equation}
-(\rho X_k)^{(2)} = (\rho X_k)^{(1)}
-- \dt \left\{ \nabla \cdot \left[\uadvtwo (\rho X_k)^{\nph,\pred} \right] \right\},
-\end{equation}
-
-\end{enumerate}
-
-\item Update $\rho_0$ by setting $\rho_0^{n+1} =$ {\bf Average}$(\rho^{(2)})$.
-
-\item Update $p_0$ by calling
-{\bf Enforce HSE}$[\rho_0^{n+1}] \rightarrow [p_0^{n+1}]$.
-
-\item Update the enthalpy.  This step is identical to {\bf Step 4D} except we use
-  the updated values $\uadvtwo, \rho_0^{n+1}, (\rho h)_0^{n+1}$, and $p_0^{n+\myhalf}$
-  rather than
-  $\uadvone, \rho_0^{n+1,\star}, (\rho h)_0^{n+1,\star}$, and $p_0^n$.
-  In particular:
-
-\begin{enumerate}
-\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-\item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\pred},$
-  for the conservative update of $(\rho h)^{(1)}$ using a Godunov approach \citep{XRB_III}.
-
-\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2)}$.
-\begin{description}
-\item For planar geometry,
-\begin{eqnarray}
-(\rho h)^{(2)}
-&=& (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvtwo (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \left( p_0^{n+1} - p_0^n \right)\nonumber \\
-&& + \dt \left(\uadvtwo \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^\nph,
-\end{eqnarray}
-
-\item For spherical geometry,
-\begin{eqnarray}
-(\rho h)^{(2)}
-&=& (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvtwo (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \left(p_0^{n+1} - p_0^n\right)\nonumber \\
-&& + \dt \left[ \nabla \cdot \left (\uadvtwo p_0^{\nph} \right ) - p_0^{\nph} \nabla \cdot \uadvtwo \right],
-\end{eqnarray}
-\end{description}
-where we now define $p_0^\nph$ as $p_0^\nph = (p_0^n+p_0^{n+1})/2$.
-
-\end{enumerate}
-Note that outside the cutoff density region, we compute enthalpy using the EOS.
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 9
-%--------------------------------------------------------------------------
-\item[Step 9] {\em React the full state through a second time interval of $\dt / 2.$}
-
-Call {\bf React State}$[(\rho X_k)^{(2)},(\rho h)^{(2)}, p_0^{n+1}] \rightarrow [(\rho X_k)^{n+1}, (\rho h)^{n+1}, (\rho \omegadot_k)^{n+1}, (\rho \Hnuc)^{n+1}].$
-
-%--------------------------------------------------------------------------
-% STEP 10
-%--------------------------------------------------------------------------
-\item[Step 10] {\em Define the new time expansion, $S^{n+1}$.}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \Alph{enumi}}}
-\item Define
-\begin{equation}
-  S^{n+1} =  \left(-\sigma  \sum_k  \xi_k \omegadot_k  + \sigma \Hnuc +
-  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  \omegadot_k\right)^{n+1}.
-\end{equation}
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 11
-%--------------------------------------------------------------------------
-\item[Step 11] {\em Update the velocity}.
-
-First, we compute the time-centered edge velocities, $\Ub^{\nph,\pred}$
-using a Godunov approach \citep{XRB_III}. Then, we define
-\begin{equation}
-\rho^\nph = \frac{\rho^n + \rho^{n+1}}{2}, \qquad \rho_0^\nph = \frac{\rho_0^n + \rho_0^{n+1}}{2}.
-\end{equation}
-We update the velocity field $\Ub^n$ to $\Ub^{n+1,\dagger}$ by discretizing
-equation (\ref{eq:momentum}) as
-\begin{equation}
-\Ub^{n+1,\dagger}
-= \Ub^n - \dt \left[\uadvtwo \cdot \nabla \Ub^{\nph,\pred} \right]
- - \dt \left[ \frac{1}{\rho^\nph} \nabla \pi^\nmh + \frac{\left(\rho^\nph-\rho_0^\nph\right)}{\rho^\nph} g^{\nph} \eb_r \right],
-\end{equation}
-Again, the $\dagger$ superscript refers
-to the fact that the updated velocity does not satisfy the divergence
-constraint.
-
-Finally, we use an approximate nodal projection to define $\Ub^{n+1}$
-from $\Ub^{n+1,\dagger},$  such that $\Ub^{n+1}$ approximately
-satisfies the divergence constraint,
-\begin{equation}
-\nabla \cdot \left(\beta_0^{n+1} \Ub^{n+1} \right) = \beta_0^{n+1} \left[ S^{n+1} - \frac{1}{\gammaonebar^{n+1}p_0^{n+1}}\left(\frac{\partial p_0}{\partial t}\right)^{n+\myhalf}\right].\label{eq:div3}
-\end{equation}
-As part of the projection we also define the new-time perturbational pressure,
-$\pi^\nph.$  This projection necessarily differs from the MAC projection used in
-{\bf Step 3} and {\bf Step 7} because the velocities in those steps are defined
-on edges and $\Ub^{n+1}$ is defined at cell centers, requiring different divergence
-and gradient operators.  Details of the approximate projection are given in Paper III.
-
-\end{description}
-
-This completes one step of the algorithm.
-
-To initialize the simulation
-
-%--------------------------------------------------------------------------
-% STEP 0
-%--------------------------------------------------------------------------
-{\em Initialization}
-\sout{
-The initialization step only occurs at the beginning of the simulation.
-The initialization step computes initial values for 
-$\pi^{-\myhalf}, \eta_{\rho}^{-\myhalf}, S^0$, and $S^1$ for use in time stepping routine.  
-We pass through {\bf Step 1} -- {\bf Step 11}, but only keep these values.
-The initial values for $\Ub^0, \rho^0, (\rho h)^0, X_k^0, T^0,
-\rho_0^0, p_0^0$, and $\gammaonebar^0$ are specified from the problem-dependent
-initial conditions.  The initial time step, $\dt^0$, is computed as in
-Paper III.
-}
-
-\section{Test Problems}
-
-\subsection{Convergence Test}
-3D reacting bubble rise; to demonstrate 2nd-order convergence of the new temporal algorithm.
-
-\subsection{Scaling}
-Scaling with OpenMP
-
-\subsection{Performance Compared to Original MAESTRO}
-Compare to original MAESTRO
-
-\subsection{White Dwarf Convection}
-white dwarf convection runs with 3 algorithms (original, new temporal, new temporal + irregular base state)
-
 \section{Conclusions and Future Work}
@@
 
-\appendix
-\section{Projection Details}
-To enforce the divergence constraint in {\bf Step 3, 7}, and {\bf 11}, we use a projection method analgous to the methods originally developed for incompressible flow.
-
-Based on our past experience in the MAESTRO project, we have found it useful to split the velocity dynamics into a perturbational and base state velocity,
-\begin{equation}
-\Ub = \Ubt(\xb,t) + w_0(r,t)\eb_r,
-\end{equation}
-solve for each term separately, and immediately combine them to find a full velocity that satisfies the constraint.
-We take that approach here, primarily because it allows us to enforce a boundary condition on $w_0$ at the edge of the star (i.e., the cutoff density location where we hold density constant).
-Namely, to enforce that $r^2 w_0$ remain constant is difficult to do when solving for the full velocity.
-So in practice, we solve the constraint over the lateral average,
-\begin{equation}
-\nabla\cdot\left(\beta_0\Ubt\right) = \beta_0(S-\overline{S})
-\end{equation}
-and sepratately solve for $w_0$ using 
-\begin{equation}
-\nabla\cdot(\beta_0w_0\eb_r) = \beta_0\left(\overline{S} - \frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial t}\right).
-\end{equation}
-To find $w_0$ we integrate in 1D using the procedure in Appendix B of Paper V.
-
-Note that in this approach, we estimated the time-derivative of the pressure in part by examining how laterally averaged $\rho'=\rho-\rho_0$ changed over time (quantified by $\eta_\rho = \overline{\rho'\Ub\cdot\eb_r}$.
-After evolving the species, we compute this term in the same way as reported in Paper V for use in computing $w_0$.
-
-\section{Old Notes (not for inclusion in paper)}
-
-\subsection{Exact Base State Notes}
-
-\subsubsection{Potential Problems}
-\begin{enumerate}
-\item How do we define \verb|r_edge_loc| for irregularly-spaced base states whose cell-centered values are defined at $r_m=\Delta x\sqrt{\frac{3}{4} + 2m}$?
-\item We would stil need to approximate the values at edges when advecting the base states (i.e.
-$\rho_0^{\nph,\pred}$).
-This may affect the nice properties we would get with exact cell-centered base state values.
-\item Excessive memory storage may be required (\verb|nr_irreg >> nr_fine|).
-\end{enumerate}
-
-\subsubsection{A Possible Solution}
-We could choose to no longer advect the base states for both predictor and corrector steps.
-This would solve the problems with computing state values at cell edges.
-Instead, we propose to only update the base states $\rho_0$ and $p_0$ by computing the average of the full state values after advecting the full states.
-This would result in the following changes in the algorithm from Paper V:
-%
-\begin{enumerate}
-\item Steps 4 and 8 will be greatly simplified, leading to a more efficient algorithm.
-\item \sout{There is no longer a need for $w_0$, so we can use the full velocity $\Ub$ instead of $\Ubt$.
-As a consequence, $\eta$ does not need to be computed.
-This has the added advantage of saving some memory storage by removing many of the terms needed when we decided to split terms into its base state and perturbation state in our previous implementation.} We split the velocity for the projection to handle the cutoff.
-\item By placing the base states exactly at cell-centers, we eliminate interpolation errors and simplify the existing code.
-\end{enumerate}
-%
-
-\subsection{$\eta_\rho$ Notes}
-After {\bf Step 8D}, we define a radial cell-centered $\etarho^{\nph}$.
-
-\begin{description}
-\item For planar geometry, $\etarho = \overline{\rho'(\Ub\cdot\eb_r)}$,
-\begin{equation}
- \etarho^{\nph} =  {\rm {\bf Average}} \sum_k \left[ \left(\uadvtwo \cdot \eb_r \right) (\rho X_k)^{\nph,\pred} \right]
-\end{equation}
-\item For spherical geometry, first construct 
-$\etarho^{{\rm cart},\nph} = [\rho'(\Ub\cdot\eb_r)]^{\nph}$ on Cartesian cell centers using:
-\begin{equation}
-\etarho^{{\rm cart},\nph} = \left[\left(\frac{\rho^{(1)}+\rho^{(2)}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{n+1}}{2}\right)\right] \cdot \left( \uadvtwo \cdot \eb_r \right).
-\end{equation}
-Then,
-\begin{equation}
-\etarho^{\nph,\star} = {\rm {\bf Average}}\left(\etarho^{{\rm cart},\nph}\right).
-\end{equation}
-\end{description}
-
-\subsection{Computing $w_0$ For Spherical Problems}
-Recall that we want to solve
-\begin{equation}
-\frac{1}{r^2}\frac{\partial}{\partial r}(r^2 w_0) = \Sbar - \frac{1}{\gammaonebar p_0}\left(\frac{\partial p_0}{\partial t} + w_0\frac{\partial p_0}{\partial r}\right)
-\end{equation}
-First, note that $\partial p_0/\partial r = -\rho_0 g$.
-The key observation here is that we can compute $\partial p_0/\partial t$ directly instead of making complicated substitutions involving $\eta_\rho$ to eliminate it.
-Let's discretize this as follows:
-\begin{equation}
-\frac{1}{r_j^2}\left[\left(r^2 w_0\right)_{j+\myhalf} - \left(r^2 w_0\right)_{j-\myhalf}\right] = 
-\Sbar_j - \frac{1}{\left(\gammaonebar p_0\right)_j}\left[\left(\frac{\partial p_0}{\partial t}\right)_j - \left(\frac{w_{0,j-\myhalf}+w_{0,j+\myhalf}}{2}\right)(\rho_0 g)_j\right]
-\end{equation}
-Rearranging this to solve for $w_{0,j+\myhalf}$ gives us an explicit update:
-\begin{equation}
-\left[\frac{r_{j+\myhalf}^2}{r_j^2} - \frac{(\rho_0 g)_j}{2\left(\gammaonebar p_0\right)_j}\right] w_{0,j+\myhalf} =
-\left[\frac{r_{j-\myhalf}^2}{r_j^2} + \frac{(\rho_0 g)_j}{2\left(\gammaonebar p_0\right)_j}\right] w_{0,j-\myhalf}
-+ \Sbar_j - \frac{1}{\left(\gammaonebar p_0\right)_j}\left(\frac{\partial p_0}{\partial t}\right)_j
-\end{equation}
-As before, once $\rho_0$ falls below $\rho_{\rm cutoff}$, we hold $r^2 w_0$ constant.\\ \\
-The interface to the routine in the algorithm description is
-{\bf Make $\mathbf{w_0}$}$[\rho_0,p_0,\partial p_0/\partial t,\gammaonebar,\Sbar] \rightarrow [w_0]$
-
 \acknowledgements
@@
 DE-AC02-05CH11231.
-
-\bibliographystyle{aasjournal.bst}
-\bibliography{references}
-
-\end{document}

### 2019-03-13 MAESTROeX_I/equations.tex
split paper into multiple tex files for multi-user edits

--- 2019-03-13 MAESTROeX_I/paper.tex
@@
-%% This manuscript uses the AASTeX v6.2 LaTeX 2e macros
-%%
-%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls
-%% (Copyright 2000-2015).  See the classfile for details.
-
-%% AASTeX requires revtex4-1.cls (http://publish.aps.org/revtex4/) and
-%% other external packages (latexsym, graphicx, amssymb, longtable, and epsf).
-%% All of these external packages should already be present in the modern TeX
-%% distributions.  If not they can also be obtained at www.ctan.org.
-
-%% The first piece of markup in an AASTeX v6.x document is the \documentclass
-%% command. LaTeX will ignore any data that comes before this command. The
-%% documentclass can take an optional argument to modify the output style.
-%% The command below calls the preprint style  which will produce a tightly
-%% typeset, one-column, single-spaced document.  It is the default and thus
-%% does not need to be explicitly stated.
-%%
-%%
-%% using aastex version 6.2
-\documentclass{aastex62}
-
-%% The default is a single spaced, 10 point font, single spaced article.
-%% There are 5 other style options available via an optional argument. They
-%% can be envoked like this:
-%%
-%% \documentclass[argument]{aastex62}
-%%
-%% where the layout options are:
-%%
-%%  twocolumn   : two text columns, 10 point font, single spaced article.
-%%                This is the most compact and represent the final published
-%%                derived PDF copy of the accepted manuscript from the publisher
-%%  manuscript  : one text column, 12 point font, double spaced article.
-%%  preprint    : one text column, 12 point font, single spaced article.
-%%  preprint2   : two text columns, 12 point font, single spaced article.
-%%  modern      : a stylish, single text column, 12 point font, article with
-%% 		  wider left and right margins. This uses the Daniel
-%% 		  Foreman-Mackey and David Hogg design.
-%%  RNAAS       : Preferred style for Research Notes which are by design
-%%                lacking an abstract and brief. DO NOT use \begin{abstract}
-%%                and \end{abstract} with this style.
-%%
-%% Note that you can submit to the AAS Journals in any of these 6 styles.
-%%
-%% There are other optional arguments one can envoke to allow other stylistic
-%% actions. The available options are:
-%%
-%%  astrosymb    : Loads Astrosymb font and define \astrocommands.
-%%  tighten      : Makes baselineskip slightly smaller, only works with
-%%                 the twocolumn substyle.
-%%  times        : uses times font instead of the default
-%%  linenumbers  : turn on lineno package.
-%%  trackchanges : required to see the revision mark up and print its output
-%%  longauthor   : Do not use the more compressed footnote style (default) for
-%%                 the author/collaboration/affiliations. Instead print all
-%%                 affiliation information after each name. Creates a much
-%%                 long author list but may be desirable for short author papers
-%%
-%% these can be used in any combination, e.g.
-%%
-%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex62}
-%%
-%% AASTeX v6.* now includes \hyperref support. While we have built in specific
-%% defaults into the classfile you can manually override them with the
-%% \hypersetup command. For example,
-%%
-%%\hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
-%%
-%% will change the color of the internal links to red, the links to the
-%% bibliography to green, the file links to cyan, and the external links to
-%% magenta. Additional information on \hyperref options can be found here:
-%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
-%%
-%% If you want to create your own macros, you can do so
-%% using \newcommand. Your macros should appear before
-%% the \begin{document} command.
-%%
-\newcommand{\vdag}{(v)^\dagger}
-\newcommand\aastex{AAS\TeX}
-\newcommand\latex{La\TeX}
-
-% for non-stacked fractions
-\newcommand{\sfrac}[2]{\mathchoice
-  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
-  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
-  {\kern0em\raise.5ex\hbox{\the\scriptscriptfont0 #1}\kern-.2em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptscriptfont0 #2}}
-  {#1\!/#2}}
-
-\newcommand{\myhalf}{\sfrac{1}{2}}
-\newcommand{\thalf}{\sfrac{3}{2}}
-
-\newcommand{\eb}{{\bf{e}}}
-\newcommand{\Ub}{{\bf{U}}}
-\newcommand{\Ubt}{\widetilde{\Ub}}
-\newcommand{\Vb}{{\bf{V}}}
-\newcommand{\xb}{{\bf{x}}}
-
-\newcommand{\dr}{\Delta r}
-\newcommand{\dt}{\Delta t}
-
-\newcommand{\etarho}{\eta_\rho}
-\newcommand{\gammaonebar}{\overline{\Gamma}_1}
-\newcommand{\Hnuc}{H_{\rm nuc}}
-\newcommand{\omegadot}{\dot\omega}
-\newcommand{\pred}{{\rm pred}}
-\newcommand{\Sbar}{\overline{S}}
-
-\newcommand{\inp}{\mathrm{in}}
-\newcommand{\outp}{\mathrm{out}}
-\newcommand{\nph}{{n+\myhalf}}
-\newcommand{\nmh}{{n-\myhalf}}
-\newcommand{\ow}{\overline{w_0}}
-\newcommand{\dw}{\delta w_0}
-\newcommand{\uadvone}{\Ub^{\mathrm{ADV},\star}}
-\newcommand{\uadvonedag}{\Ub^{\mathrm{ADV},\dagger,\star}}
-\newcommand{\uadvtwo}{\Ub^{\mathrm{ADV}}}
-\newcommand{\uadvtwodag}{\Ub^{\mathrm{ADV},\dagger}}
-\newcommand{\gcc}{\mathrm{g~cm^{-3} }}
-
-% for the red MarginPars
-\usepackage{color}
-% make the MarginPars look pretty
-\setlength{\marginparwidth}{0.5in}
-\newcommand{\MarginPar}[1]{\marginpar{\vskip-\baselineskip\raggedright\tiny\sffamily
-\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}
-
-%% Reintroduced the \received and \accepted commands from AASTeX v5.2
-\received{XXX X, XXXX}
-\revised{XXX X, XXXX}
-\accepted{XXX X, XXXX}
-%% Command to document which AAS Journal the manuscript was submitted to.
-%% Adds "Submitted to " the arguement.
-\submitjournal{ApJ}
-
-%% Mark up commands to limit the number of authors on the front page.
-%% Note that in AASTeX v6.2 a \collaboration call (see below) counts as
-%% an author in this case.
-%
-%\AuthorCollaborationLimit=3
-%
-%% Will only show Schwarz, Muench and "the AAS Journals Data Scientist
-%% collaboration" on the front page of this example manuscript.
-%%
-%% Note that all of the author will be shown in the published article.
-%% This feature is meant to be used prior to acceptance to make the
-%% front end of a long author article more manageable. Please do not use
-%% this functionality for manuscripts with less than 20 authors. Conversely,
-%% please do use this when the number of authors exceeds 40.
-%%
-%% Use \allauthors at the manuscript end to show the full author list.
-%% This command should only be used with \AuthorCollaborationLimit is used.
-
-%% The following command can be used to set the latex table counters.  It
-%% is needed in this document because it uses a mix of latex tabular and
-%% AASTeX deluxetables.  In general it should not be needed.
-%\setcounter{table}{1}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% The following section outlines numerous optional output that
-%% can be displayed in the front matter or as running meta-data.
-%%
-%% If you wish, you may supply running head information, although
-%% this information may be modified by the editorial offices.
-\shorttitle{MAESTROeX Low Mach Number Astrophysics}
-\shortauthors{Fan et al.}
-%%
-%% You can add a light gray and diagonal water-mark to the first page
-%% with this command:
-% \watermark{text}
-%% where "text", e.g. DRAFT, is the text to appear.  If the text is
-%% long you can control the water-mark size with:
-%  \setwatermarkfontsize{dimension}
-%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% This is the end of the preamble.  Indicate the beginning of the
-%% manuscript itself with \begin{document}.
-
-\begin{document}
-
-\title{MAESTROeX: A Massively Parallel Low Mach Number Astrophysical Solver}
-
-%% LaTeX will automatically break titles if they run longer than
-%% one line. However, you may use \\ to force a line break if
-%% you desire. In v6.2 you can include a footnote in the title.
-
-%% A significant change from earlier AASTEX versions is in the structure for
-%% calling author and affilations. The change was necessary to implement
-%% autoindexing of affilations which prior was a manual process that could
-%% easily be tedious in large author manuscripts.
-%%
-%% The \author command is the same as before except it now takes an optional
-%% arguement which is the 16 digit ORCID. The syntax is:
-%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
-%%
-%% This will hyperlink the author name to the author's ORCID page. Note that
-%% during compilation, LaTeX will do some limited checking of the format of
-%% the ID to make sure it is valid.
-%%
-%% Use \affiliation for affiliation information. The old \affil is now aliased
-%% to \affiliation. AASTeX v6.2 will automatically index these in the header.
-%% When a duplicate is found its index will be the same as its previous entry.
-%%
-%% Note that \altaffilmark and \altaffiltext have been removed and thus
-%% can not be used to document secondary affiliations. If they are used latex
-%% will issue a specific error message and quit. Please use multiple
-%% \affiliation calls for to document more than one affiliation.
-%%
-%% The new \altaffiliation can be used to indicate some secondary information
-%% such as fellowships. This command produces a non-numeric footnote that is
-%% set away from the numeric \affiliation footnotes.  NOTE that if an
-%% \altaffiliation command is used it must come BEFORE the \affiliation call,
-%% right after the \author command, in order to place the footnotes in
-%% the proper location.
-%%
-%% Use \email to set provide email addresses. Each \email will appear on its
-%% own line so you can put multiple email address in one \email call. A new
-%% \correspondingauthor command is available in V6.2 to identify the
-%% corresponding author of the manuscript. It is the author's responsibility
-%% to make sure this name is also in the author list.
-%%
-%% While authors can be grouped inside the same \author and \affiliation
-%% commands it is better to have a single author for each. This allows for
-%% one to exploit all the new benefits and should make book-keeping easier.
-%%
-%% If done correctly the peer review system will be able to
-%% automatically put the author and affiliation information from the manuscript
-%% and save the corresponding author the trouble of entering it by hand.
-
-\correspondingauthor{Doreen Fan; DFan@lbl.gov}
-
-\author[0000-0002-3246-4315]{Duoming Fan}
-\affil{Lawrence Berkelely National Laboratory \\
-Center for Computational Sciences and Engineering \\
-One Cyclotron Road, MS 50A-3111 \\
-Berkeley, CA 94720, USA}
-
-\author[0000-0003-1791-0265]{Andrew Nonaka}
-\affil{Lawrence Berkelely National Laboratory \\
-Center for Computational Sciences and Engineering \\
-One Cyclotron Road, MS 50A-3111 \\
-Berkeley, CA 94720, USA}
-
-\author[0000-0003-2103-312X]{Ann S. Almgren}
-\affil{Lawrence Berkelely National Laboratory \\
-Center for Computational Sciences and Engineering \\
-One Cyclotron Road, MS 50A-3111 \\
-Berkeley, CA 94720, USA}
-
-\author[0000-0002-1530-781X]{Alice Harpole}
-\affil{Stony Brook University \\
-Department of Physics and Astronomy \\
-Stony Brook, NY 11794-3800, USA}
-
-\author[0000-0001-8401-030X]{Michael Zingale}
-\affil{Stony Brook University \\
-Department of Physics and Astronomy \\
-Stony Brook, NY 11794-3800, USA}
-
-%% Note that the \and command from previous versions of AASTeX is now
-%% depreciated in this version as it is no longer necessary. AASTeX
-%% automatically takes care of all commas and "and"s between authors names.
-
-%% AASTeX 6.2 has the new \collaboration and \nocollaboration commands to
-%% provide the collaboration status of a group of authors. These commands
-%% can be used either before or after the list of corresponding authors. The
-%% argument for \collaboration is the collaboration identifier. Authors are
-%% encouraged to surround collaboration identifiers with ()s. The
-%% \nocollaboration command takes no argument and exists to indicate that
-%% the nearby authors are not part of surrounding collaborations.
-
-%% Mark off the abstract in the ``abstract'' environment.
-\begin{abstract}
-We present MAESTROeX, a massively parallel solver for low Mach number astrophysical flows.
-Our model equations allow for long-time integration for highly subsonic flows compared to compressible approaches.
-The code leverages the new AMReX framework for block-structured adaptive mesh refinement calculations, and also provides additional algorithmic features from the original MAESTRO code.
-We present a new temporal integration option that is much simpler than our previous approach while retaining the same order order of accuracy.
-We also provide a more accurate spatial discretization for mapping the base state onto the full state Cartesian grid.
-Using our previous studies on the convective phase of single-degenerate progenitor models of Type Ia supernovae, we characterize the performance of the code and validate the new algorithmic features.
-\end{abstract}
-
-%% Keywords should appear after the \end{abstract} command.
-%% See the online documentation for the full list of available subject
-%% keywords and the rules for their use.
-\keywords{convection, hydrodynamics, methods: numerical, nuclear reactions, nucleosynthesis, abundances, supernovae: general}
-
-%% From the front matter, we move on to the body of the paper.
-%% Sections are demarcated by \section and \subsection, respectively.
-%% Observe the use of the LaTeX \label
-%% command after the \subsection to give a symbolic KEY to the
-%% subsection for cross-referencing in a \ref command.
-%% You can use LaTeX's \ref and \label commands to keep track of
-%% cross-references to sections, equations, tables, and figures.
-%% That way, if you change the order of any elements, LaTeX will
-%% automatically renumber them.
-%%
-%% We recommend that authors also use the natbib \citep
-%% and \citet commands to identify citations.  The citations are
-%% tied to the reference list via symbolic KEYs. The KEY corresponds
-%% to the KEY in the \bibitem in the reference list below.
-
-\section{Introduction} \label{sec:intro}
-
-TODO list:
-\begin{itemize}
-\item Weak scaling tests of original algorithm, comparing MAESTRO to MAESTROeX.  Use cori knl, 4 MPI per node, 16 threads per MPI
-\item Run at effective $512^3$ resolution (maybe to ignition?...
-need to see how max T trends go) (1) classic MAESTRO, (2) MAESTROeX with original algorithm, (3) MAESTROeX with new temporal integrator, (4) MAESTROeX with new temporal integrator and irregular dr, (5) MAESTROeX with original algorithm and 3-levels of AMR, (6) MAESTROeX with new temporal integrator and 3 levels of AMR.
-\item (perhaps) a demonstration of what goes wrong when you don't split the velocity dynamics in the projection
-\end{itemize}
-
-Many astrophysical flows are highly subsonic; sound waves carry sufficiently little energy that they do not significantly affect the convective dynamics of the system.
-In many of these flows, modeling long-time convective dynamics are of interest, and numerical approaches based on compressible hydrodynamics are intractable, even on modern supercomputers.
-One approach to this problem is to use low Mach number models.
-In a low Mach number approach, sound waves are eliminated from the governing equations while retaining compressibilitiy effects due to, e.g., nuclear energy release, compositional changes, and thermal diffusion.
-The resulting model can be numerically integrated with much larger time steps than a compressible model.
-
-Previously, we have developed the low Mach number astrophysical solver, MAESTRO.
-Low Mach number models have been developed for a variety of contexts including combustion \citep{day2000numerical}, atmospherics \citep{duarte2015low}, and elastic solids \citep{abbate2017all}.
-The low Mach number model in MAESTRO is unique in that it is specifically designed for astrophysical settings with significant atmospheric stratification.
-Central to the algorithm is a stratified background (or base) state density and pressure held in hydrostatic equilibrium, and also vary as a function of altitude and time.
-MAESTRO is suitable for for spherical stars, as well as planar simulations of local dynamics.
-
-The key numerical developments of the original MAESTRO algorithm are presented in a series of papers which we refer to as Papers I-V:
-\begin{itemize}
-\item In Paper I \citep{MAESTRO_I}, we derive the low Mach number equation set from the fully compressible equations.
-\item In Paper II \citep{MAESTRO_II}, we incorporate the effects atmospheric expansion through the use of a time-dependent background state.
-\item In Paper III \citep{MAESTRO_III}, we incorporate reactions and the associated coupling to the hydrodynamics.
-\item In Paper IV \citep{MAESTRO_IV}, we describe our treatment of spherical stars in a three-dimensional Cartesian geometry.
-\item In Paper V \citep{MAESTRO_V}, we describe the use of block-structured adaptive mesh refinement to focus spatial resolution in regions of interest.
-\end{itemize}
-
-Since then, there have been many scientific investigations using MAESTRO, which include additional algorithmic enhancements.  Topics include:
-\begin{itemize}
-\item The convective phase preceding Chandrasekhar mass models for type Ia supernovae \citep{MAESTRO_convection,MAESTRO_AMR,MAESTRO_CASTRO}.
-\item Convection in massive stars \citep{Gilet:2013}.
-\item Sub-Chandrasekhar white dwarfs \citep{subChandra_I,subChandra_II}.
-In the latter paper we introduced an optional modification to the momentum equation of velocity constraint that conserves total energy, can give results closer to compressible codes of convective dynamics, including low density regions at the surface of a star.
-\item Type I X-ray bursts \citep{XRB_I,XRB_II,XRB_III}.
-In \cite{XRB_I} we discuss the modifications required for implicit thermal conduction, as well as an additional ``volume discrepancy'' modification to the velocity constraint to force the species and enthalpy to evolve in a manner consistent with the background pressure.
-\end{itemize}
-
-In this paper, we characterize the performance of the MAESTROeX algorithm compared to the previous implementation.
-We also develop simpler, alternative temporal integrators, with a longer term goal of developing high-order coupling schemes.
-We also present a new treatment of the base state mapping to Cartesian grids for spherical problsm.
-
 \section{Governing Equations}
@@
 \right|_{p,X_k}$ is the specific heat at constant pressure.
-
-Previously we adopted an approach where we split the velocity into a base state component, $w_0(r,t)$, 
-and a local velocity $\Ubt(\xb,t)$, so that
-\begin{equation}
-\Ub = \Ubt(\xb,t) + w_0(r,t)\eb_r,
-\end{equation}
-We used $w_0$ to provide an estimate of the base state density evolution over a time step.
-In our new simplified approach, we instead use a predictor-corrector approach for evolving the base state.
-
-\section{Numerical Algorithm}
-
-\subsection{Spatial Discretization}
-
-(description of base state mapping in planar, and origianl spherical)
-
-\subsection{Exact Base State Mapping} \label{sec:exactbasestate}
-
-\subsection{Temporal Integration Scheme}\label{Sec:Temporal Integration Scheme}
-Our new temporal integration scheme is much simpler than the scheme from Paper V.
-The primary reason is that we no longer split the velocity field into a perturbational and base state components.  
-Previously, we used the base state velocity to predict the expansion of the base state.  
-In this new scheme we use a simpler, predictor-corrector approach to the base state density and pressure that no longer requires the complex algorithm from Paper V, yet still retains the same overall second-order accuracy.
-
-We now describe the new temporal integration scheme, which has been greatly simplified
-from Paper V without formal loss of order of accuracy.
-Note that this temporal integration scheme is valid for the original base state mapping (with constant base state grid spacing), or the new irregularly spaced base state mapping.
-
-At the beginning of each time step we have the cell-centered state,
-$(\Ub,\rho X_k,\rho h,\rho_0,p_0)^n$, and nodal state, $\pi^{n-\myhalf}$.
-At any time, the associated density, composition, and enthalpy can be trivially computed using, e.g.,
-\begin{equation}
-\rho^n = \sum_k(\rho X_k)^n, \quad
-X_k^n = (\rho X_k)^n / \rho^n, \quad
-h^n = (\rho h)^n / \rho^n.
-\end{equation}
-Temperature is computed using the equation of state, e.g.,
-\footnote{As described in Paper V, for planar problems we compute temperature using $h$ instead of $p_0$, since we have successfully developed planar volume discrepancy schemes to effectively couple the enthalpy to the rest of the solution.
-We are still exploring this option for spherical stars.}
-\begin{equation}
-T = T(\rho,p_0,X_k),
-\end{equation}
-and ($\gammaonebar,\beta_0)$ are computed from $(\rho,p_0,X_k)$ (see Appendix A of Paper I and Appendix C of Paper III for details).
-
-The overall flow of the algorithm is a second-order Strang splitting approach for the coupling of the reactions and advection of the thermodynamic variables.  
-We use a predictor-corrector approach within the Strang splitting scheme to achieve second-order accuracy in time.
-We integrate the velocity using a standard second-order projection method to enforce the divergence constraint.
-To summarize:
-\begin{itemize}
-\item In {\bf Step 1} we react the thermodynamic variables over $\Delta t/2$.
-\item In {\bf Steps 2-4} we advect the thermodynamic variables over $\Delta t$.
-Specifically, we compute an estimate for the expansion term, $S$, project the velocities so they satisfy the divergence constraint, and then advect the thermodynamic variables.
-\item In {\bf Step 5} we react the thermodynamic variables over $\Delta t/2$.\footnote{Note that we could skip to the velocity advance in {\bf Steps 10-11}, however the overall scheme would be only first-order in time, so {\bf Steps 6-9} can be thought of as a trapezoidal corrector step.}
-\item In {\bf Steps 6-8} we redo the advection in {\bf Steps 2-4} but are able to use the trapezoidal rule to time-center certain quantities such as $S$, $\rho_0$, etc.
-\item In {\bf Step 9} we redo the reactions from {\bf Step 5} using the improved results for the corrector advection step.
-\item In {\bf Steps 10-11} we update the velocity.
-\end{itemize}
-
-There are a few key numerical modules we use in each time step.
-\begin{itemize}
-\item {\bf Average} computes...
-\item {\bf Enforce HSE} computes the base state pressure, $p_0$, from a base state density, $\rho_0$ by integrating the hydrostatic equilibrium condition in 1D (see equation (A10) in Paper V).
-The base state pressure remains equal to a constant value from the location of a prescribed cutoff density outward for the entire simulation.
-\item {\bf React State} advances the species and enthalpy due to reactions over $\Delta t/2$ by integrating
-\begin{equation}
-\frac{dX_k}{dt} = \dot\omega_k(\rho,X_k,T); \qquad
-\frac{dT}{dt} = \frac{1}{c_p}\left(-\sum_k\xi_k\dot\omega_k + H_{\rm nuc}\right)
-\end{equation}
-The inputs are the species, enthalpy, and base state pressure, and the outpus are the species, enthalpy, reaction rates, and nuclear energy generation rate.
-See Papers III and V for details.
-\end{itemize}
-
-Each time step satisfies the standard advective CFL condition,
-\begin{equation}
-\Delta t = \sigma^{\rm CFL} \min_i(\Delta x / U_i),
-\end{equation}
-where for our simulations we typically use $\sigma^{\rm CFL}\sim 0.5$.
-There are additional constraints on the time involving the buoyancy force (commonly used when the velocity is approximately zero at the start of some simulations) and the divergence constraint (to prevent too much mass evacuation from a cell in a time step); see Section 3.4 in Paper III for details.
-
-At the beginning of each simulation, we define $(\Ub,\rho X_k,\rho h)$...
-
-\MarginPar{mention cutoff densities}
-
-\MarginPar{AJN: eliminated references to volume discrepancy for now; we'd only use it for planar at the moment, don't want to open up that can of worms here}
-
-\begin{description}
-
-%--------------------------------------------------------------------------
-% STEP 1
-%--------------------------------------------------------------------------
-\item[Step 1] {\em React the full state through the first time interval of $\dt / 2.$}
-
-Call {\bf React State}$[(\rho X_k)^n, (\rho h)^n, p_0^n] \rightarrow [(\rho X_k)^{(1)}, (\rho h)^{(1)}, (\rho \omegadot_k)^{(1)}, (\rho \Hnuc)^{(1)}]$.
-
-%--------------------------------------------------------------------------
-% STEP 2
-%--------------------------------------------------------------------------
-
-\item[Step 2] {\em Compute the provisional time-centered expansion,
-    $S^{\nph,\star}$.}
-
-We compute an estimate for the time-centered expansion term in the velocity
-divergence constraint.  Following \citet{Bell:2004}, we extrapolate
-to the half-time using $S$ at the previous and current
-time steps,
-\begin{equation}
-S^{\nph,\star} = S^n + \frac{\dt^n}{2} \left(\frac{S^n - S^{n-1}}{\dt^{n-1}}\right).
-\end{equation}
-Note that in the first time step we average $S^0$ and $S^1$ from the
-initialization step.
-
-%--------------------------------------------------------------------------
-% STEP 3
-%--------------------------------------------------------------------------
-\item[Step 3] {\em Construct the provisional time-centered advective velocity on
-edges, $\uadvone$.}
-
-The construction of face-centered time-centered states used to discretize the
-advection terms for velocity, species, and enthalpy, are performed using
-a standard multidimensional corner transport upwind approach
-\citep{colella1990multidimensional,saltzman1994unsplit} with piecewise-parabolic
-one-dimensional tracing \citep{colella1984piecewise}.  The full details of this
-Godunov advection approach for all steps in this algorithm are described 
-in Appendix A of \cite{XRB_III}.
-
-In this step, using equation (\ref{eq:momentum}), 
-we compute time-centered edge velocities, $\uadvonedag$, using
-$\Ub^n$.  The $\dagger$ superscript refers to the
-fact that the predicted velocity field does not satisfy the divergence
-constraint.  We then construct $\uadvone$ from $\uadvonedag$
-using a MAC projection.
-We note that $\uadvone$ satisfies the divergence constraint
-\begin{equation}
-\nabla \cdot \left(\beta_0^n \uadvone\right) = \beta_0^n \left[S^{\nph,\star} - \frac{1}{\gammaonebar^np_0^n}\left(\frac{\partial p_0}{\partial t}\right)^{n-\myhalf} \right].\label{eq:div1}
-\end{equation}
-In practice we solve for.
-
-%--------------------------------------------------------------------------
-% STEP 4
-%--------------------------------------------------------------------------
-\item[Step 4] {\em Advect the full state through a time interval of $\dt.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \Alph{enumi}}}
-
-\item Update $(\rho X_k)$ using a discretized version of
-%
-\begin{equation}
-\frac{\partial(\rho X_k)}{\partial t} = -\nabla\cdot(\rho X_k\Ub) + \rho\omegadot_k,
-\end{equation}
-%
-omitting the reaction terms, which were already
-accounted for in {\bf React State}.  The update consists of two steps:
-
-\begin{enumerate}
-\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-\item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\pred}$,
-  for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
-\MarginPar{mention $\rho'$ and $X$ advection; $(\rho h)'$ advection}
-
-\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2),\star}$ using
-\begin{equation}
-(\rho X_k)^{(2),\star} = (\rho X_k)^{(1)}
-  - \dt \left\{ \nabla \cdot \left[ \uadvone (\rho X_k)^{\nph,\pred} \right] \right\},
-\end{equation}
-
-\end{enumerate}
-
-\item Update $\rho_0$ by setting $\rho_0^{n+1,\star} =$ {\bf Average}$(\rho^{(2),\star})$.
-
-\item Update $p_0$ by calling
-{\bf Enforce HSE}$[\rho_0^{n+1,\star}] \rightarrow [p_0^{n+1,\star}]$.
-
-\item Update the enthalpy using a discretized version of equation
-%
-\begin{equation}
-\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc,
-\end{equation}
-%
-again omitting the reaction and heating terms
-since we already accounted for
-them in {\bf React State}.  This equation takes the form:
-\begin{equation}
-\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h) + \frac{\partial p_0}{\partial t} + (\Ub \cdot \eb_r) \frac{\partial p_0}{\partial r}.
-\end{equation}
-For spherical geometry, we solve the
-analytically equivalent form,
-\begin{equation}
-\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h) + \frac{\partial p_0}{\partial t} + \nabla \cdot (\Ub p_0) - p_0 \nabla \cdot \Ub.
-\end{equation}
-The update consists of two steps:
-
-\begin{enumerate}
-\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-\item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\pred},$
-  for the conservative update of $(\rho h)^{(1)}$
-  using using a Godunov approach \citep{XRB_III}.
-
-\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$.
-\begin{description}
-\item For planar geometry,
-\begin{eqnarray}
-(\rho h)^{(2),\star}
-&=& (\rho h)^{(1)} \nonumber \\
-&&- \dt \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \left(p_0^{n+1,*} - p_0^n\right) \nonumber \\
-&& + \dt \left(\uadvone \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^{n}
-\end{eqnarray}
-
-\item For spherical geometry,
-\begin{eqnarray}
-(\rho h)^{(2),\star}
-&=& (\rho h)^{(1)} \nonumber \\
-&&- \dt \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \left(p_0^{n+1,*} - p_0^n\right) \nonumber \\
-&& + \dt \left \{ \nabla \cdot \left (\uadvone p_0^{n+\myhalf} \right ) - p_0^{n+\myhalf} \nabla \cdot \uadvone \right \}
-\end{eqnarray}
-\end{description}
-where we define $p_0^\nph$ as $p_0^\nph = (p_0^n+p_0^{n+1,*})/2$.
-
-\end{enumerate}
-Note that outside the cutoff density region, we compute enthalpy using the EOS.
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 5
-%--------------------------------------------------------------------------
-\item[Step 5] {\em React the full state through a second time interval of $\dt / 2.$}
-
-Call {\bf React State}$[ (\rho X_k)^{(2),\star}, (\rho h)^{(2),\star}, p_0^{n+1,\star}] 
-\rightarrow 
-[ (\rho X_k)^{n+1,\star}, (\rho h)^{n+1,\star}, (\rho \omegadot_k)^{n+1,\star}, (\rho \Hnuc)^{n+1,\star} ].$
-
-%--------------------------------------------------------------------------
-% STEP 6
-%--------------------------------------------------------------------------
-\item[Step 6] {\em Compute the time-centered expansion, $S^{\nph,\star}$.}
-
-First, compute $S^{n+1,\star}$ with
-\begin{equation}
-S^{n+1,\star} =  \left(-\sigma  \sum_k  \xi_k  \omegadot_k  + \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k + \sigma \Hnuc\right)^{n+1,\star}.
-\end{equation}
-  Then, define
-\begin{equation}
- S^{\nph} = \frac{S^n + S^{n+1,\star}}{2},
-\end{equation}
-
-%--------------------------------------------------------------------------
-% STEP 7
-%--------------------------------------------------------------------------
-\item[Step 7] {\em Construct the time-centered advective velocity on edges, $\uadvtwo$.}
-
-The procedure to construct $\uadvtwodag$ is identical to the Godunov procedure
-for computing $\uadvonedag$ in {\bf Step 3}, but uses
-the updated value $S^{\nph}$ rather than $S^{\nph,\star}$.
-We note that $\uadvtwo$ satisfies the divergence constraint
-\begin{equation}
-\nabla \cdot \left(\beta_0^{\nph} \uadvtwo\right) =
-\beta_0^{\nph} \left[S^{\nph} - \frac{1}{\gammaonebar^{n+\myhalf}p_0^{n+\myhalf}}\left(\frac{\partial p_0}{\partial t}\right)^{n+\myhalf}\right],\label{eq:div2}
-\end{equation}
-where
-\begin{equation}
-\beta_0^{\nph} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2},
-\qquad
-\gammaonebar^{\nph} = \frac{ \gammaonebar^n +  \gammaonebar^{n+1,\star} }{2}.
-\end{equation}
-
-%--------------------------------------------------------------------------
-% STEP 8
-%--------------------------------------------------------------------------
-\item[Step 8] {\em Advect the base state and full state through a time interval of $\dt.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \Alph{enumi}}}
-
-\item Update $(\rho X_k)$.  This step is identical to {\bf Step 4A} except we use
-  the updated values $\uadvtwo$ and $\rho_0^{n+1,\star}$ rather than
-  $\uadvone$ and $\rho_0^{n+1,\star}$.  In particular:
-
-\begin{enumerate}
-\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-\item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\pred}$,
-  for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
-
-\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2)}$ using
-\begin{equation}
-(\rho X_k)^{(2)} = (\rho X_k)^{(1)}
-- \dt \left\{ \nabla \cdot \left[\uadvtwo (\rho X_k)^{\nph,\pred} \right] \right\},
-\end{equation}
-
-\end{enumerate}
-
-\item Update $\rho_0$ by setting $\rho_0^{n+1} =$ {\bf Average}$(\rho^{(2)})$.
-
-\item Update $p_0$ by calling
-{\bf Enforce HSE}$[\rho_0^{n+1}] \rightarrow [p_0^{n+1}]$.
-
-\item Update the enthalpy.  This step is identical to {\bf Step 4D} except we use
-  the updated values $\uadvtwo, \rho_0^{n+1}, (\rho h)_0^{n+1}$, and $p_0^{n+\myhalf}$
-  rather than
-  $\uadvone, \rho_0^{n+1,\star}, (\rho h)_0^{n+1,\star}$, and $p_0^n$.
-  In particular:
-
-\begin{enumerate}
-\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-\item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\pred},$
-  for the conservative update of $(\rho h)^{(1)}$ using a Godunov approach \citep{XRB_III}.
-
-\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2)}$.
-\begin{description}
-\item For planar geometry,
-\begin{eqnarray}
-(\rho h)^{(2)}
-&=& (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvtwo (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \left( p_0^{n+1} - p_0^n \right)\nonumber \\
-&& + \dt \left(\uadvtwo \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^\nph,
-\end{eqnarray}
-
-\item For spherical geometry,
-\begin{eqnarray}
-(\rho h)^{(2)}
-&=& (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvtwo (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \left(p_0^{n+1} - p_0^n\right)\nonumber \\
-&& + \dt \left[ \nabla \cdot \left (\uadvtwo p_0^{\nph} \right ) - p_0^{\nph} \nabla \cdot \uadvtwo \right],
-\end{eqnarray}
-\end{description}
-where we now define $p_0^\nph$ as $p_0^\nph = (p_0^n+p_0^{n+1})/2$.
-
-\end{enumerate}
-Note that outside the cutoff density region, we compute enthalpy using the EOS.
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 9
-%--------------------------------------------------------------------------
-\item[Step 9] {\em React the full state through a second time interval of $\dt / 2.$}
-
-Call {\bf React State}$[(\rho X_k)^{(2)},(\rho h)^{(2)}, p_0^{n+1}] \rightarrow [(\rho X_k)^{n+1}, (\rho h)^{n+1}, (\rho \omegadot_k)^{n+1}, (\rho \Hnuc)^{n+1}].$
-
-%--------------------------------------------------------------------------
-% STEP 10
-%--------------------------------------------------------------------------
-\item[Step 10] {\em Define the new time expansion, $S^{n+1}$.}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \Alph{enumi}}}
-\item Define
-\begin{equation}
-  S^{n+1} =  \left(-\sigma  \sum_k  \xi_k \omegadot_k  + \sigma \Hnuc +
-  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  \omegadot_k\right)^{n+1}.
-\end{equation}
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 11
-%--------------------------------------------------------------------------
-\item[Step 11] {\em Update the velocity}.
-
-First, we compute the time-centered edge velocities, $\Ub^{\nph,\pred}$
-using a Godunov approach \citep{XRB_III}. Then, we define
-\begin{equation}
-\rho^\nph = \frac{\rho^n + \rho^{n+1}}{2}, \qquad \rho_0^\nph = \frac{\rho_0^n + \rho_0^{n+1}}{2}.
-\end{equation}
-We update the velocity field $\Ub^n$ to $\Ub^{n+1,\dagger}$ by discretizing
-equation (\ref{eq:momentum}) as
-\begin{equation}
-\Ub^{n+1,\dagger}
-= \Ub^n - \dt \left[\uadvtwo \cdot \nabla \Ub^{\nph,\pred} \right]
- - \dt \left[ \frac{1}{\rho^\nph} \nabla \pi^\nmh + \frac{\left(\rho^\nph-\rho_0^\nph\right)}{\rho^\nph} g^{\nph} \eb_r \right],
-\end{equation}
-Again, the $\dagger$ superscript refers
-to the fact that the updated velocity does not satisfy the divergence
-constraint.
-
-Finally, we use an approximate nodal projection to define $\Ub^{n+1}$
-from $\Ub^{n+1,\dagger},$  such that $\Ub^{n+1}$ approximately
-satisfies the divergence constraint,
-\begin{equation}
-\nabla \cdot \left(\beta_0^{n+1} \Ub^{n+1} \right) = \beta_0^{n+1} \left[ S^{n+1} - \frac{1}{\gammaonebar^{n+1}p_0^{n+1}}\left(\frac{\partial p_0}{\partial t}\right)^{n+\myhalf}\right].\label{eq:div3}
-\end{equation}
-As part of the projection we also define the new-time perturbational pressure,
-$\pi^\nph.$  This projection necessarily differs from the MAC projection used in
-{\bf Step 3} and {\bf Step 7} because the velocities in those steps are defined
-on edges and $\Ub^{n+1}$ is defined at cell centers, requiring different divergence
-and gradient operators.  Details of the approximate projection are given in Paper III.
-
-\end{description}
-
-This completes one step of the algorithm.
-
-To initialize the simulation
-
-%--------------------------------------------------------------------------
-% STEP 0
-%--------------------------------------------------------------------------
-{\em Initialization}
-\sout{
-The initialization step only occurs at the beginning of the simulation.
-The initialization step computes initial values for 
-$\pi^{-\myhalf}, \eta_{\rho}^{-\myhalf}, S^0$, and $S^1$ for use in time stepping routine.  
-We pass through {\bf Step 1} -- {\bf Step 11}, but only keep these values.
-The initial values for $\Ub^0, \rho^0, (\rho h)^0, X_k^0, T^0,
-\rho_0^0, p_0^0$, and $\gammaonebar^0$ are specified from the problem-dependent
-initial conditions.  The initial time step, $\dt^0$, is computed as in
-Paper III.
-}
-
-\section{Test Problems}
-
-\subsection{Convergence Test}
-3D reacting bubble rise; to demonstrate 2nd-order convergence of the new temporal algorithm.
-
-\subsection{Scaling}
-Scaling with OpenMP
-
-\subsection{Performance Compared to Original MAESTRO}
-Compare to original MAESTRO
-
-\subsection{White Dwarf Convection}
-white dwarf convection runs with 3 algorithms (original, new temporal, new temporal + irregular base state)
-
-\section{Conclusions and Future Work}
-
-science: rotation, solar physics, MHD
-
-algorithm: SDC \cite{dutt2000spectral}, high-order, multi-implicit \cite{bourlioux2003high}.
-Consider SDC ideas used for terrestrial combustion \cite{pazner2016high,nonaka2018conservative}
-
-implementation: GPUs.  Forward cite next CASTRO paper with GPU strategy.
-
-\appendix
-\section{Projection Details}
-To enforce the divergence constraint in {\bf Step 3, 7}, and {\bf 11}, we use a projection method analgous to the methods originally developed for incompressible flow.
-
-Based on our past experience in the MAESTRO project, we have found it useful to split the velocity dynamics into a perturbational and base state velocity,
-\begin{equation}
-\Ub = \Ubt(\xb,t) + w_0(r,t)\eb_r,
-\end{equation}
-solve for each term separately, and immediately combine them to find a full velocity that satisfies the constraint.
-We take that approach here, primarily because it allows us to enforce a boundary condition on $w_0$ at the edge of the star (i.e., the cutoff density location where we hold density constant).
-Namely, to enforce that $r^2 w_0$ remain constant is difficult to do when solving for the full velocity.
-So in practice, we solve the constraint over the lateral average,
-\begin{equation}
-\nabla\cdot\left(\beta_0\Ubt\right) = \beta_0(S-\overline{S})
-\end{equation}
-and sepratately solve for $w_0$ using 
-\begin{equation}
-\nabla\cdot(\beta_0w_0\eb_r) = \beta_0\left(\overline{S} - \frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial t}\right).
-\end{equation}
-To find $w_0$ we integrate in 1D using the procedure in Appendix B of Paper V.
-
-Note that in this approach, we estimated the time-derivative of the pressure in part by examining how laterally averaged $\rho'=\rho-\rho_0$ changed over time (quantified by $\eta_\rho = \overline{\rho'\Ub\cdot\eb_r}$.
-After evolving the species, we compute this term in the same way as reported in Paper V for use in computing $w_0$.
-
-\section{Old Notes (not for inclusion in paper)}
-
-\subsection{Exact Base State Notes}
-
-\subsubsection{Potential Problems}
-\begin{enumerate}
-\item How do we define \verb|r_edge_loc| for irregularly-spaced base states whose cell-centered values are defined at $r_m=\Delta x\sqrt{\frac{3}{4} + 2m}$?
-\item We would stil need to approximate the values at edges when advecting the base states (i.e.
-$\rho_0^{\nph,\pred}$).
-This may affect the nice properties we would get with exact cell-centered base state values.
-\item Excessive memory storage may be required (\verb|nr_irreg >> nr_fine|).
-\end{enumerate}
-
-\subsubsection{A Possible Solution}
-We could choose to no longer advect the base states for both predictor and corrector steps.
-This would solve the problems with computing state values at cell edges.
-Instead, we propose to only update the base states $\rho_0$ and $p_0$ by computing the average of the full state values after advecting the full states.
-This would result in the following changes in the algorithm from Paper V:
-%
-\begin{enumerate}
-\item Steps 4 and 8 will be greatly simplified, leading to a more efficient algorithm.
-\item \sout{There is no longer a need for $w_0$, so we can use the full velocity $\Ub$ instead of $\Ubt$.
-As a consequence, $\eta$ does not need to be computed.
-This has the added advantage of saving some memory storage by removing many of the terms needed when we decided to split terms into its base state and perturbation state in our previous implementation.} We split the velocity for the projection to handle the cutoff.
-\item By placing the base states exactly at cell-centers, we eliminate interpolation errors and simplify the existing code.
-\end{enumerate}
-%
-
-\subsection{$\eta_\rho$ Notes}
-After {\bf Step 8D}, we define a radial cell-centered $\etarho^{\nph}$.
-
-\begin{description}
-\item For planar geometry, $\etarho = \overline{\rho'(\Ub\cdot\eb_r)}$,
-\begin{equation}
- \etarho^{\nph} =  {\rm {\bf Average}} \sum_k \left[ \left(\uadvtwo \cdot \eb_r \right) (\rho X_k)^{\nph,\pred} \right]
-\end{equation}
-\item For spherical geometry, first construct 
-$\etarho^{{\rm cart},\nph} = [\rho'(\Ub\cdot\eb_r)]^{\nph}$ on Cartesian cell centers using:
-\begin{equation}
-\etarho^{{\rm cart},\nph} = \left[\left(\frac{\rho^{(1)}+\rho^{(2)}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{n+1}}{2}\right)\right] \cdot \left( \uadvtwo \cdot \eb_r \right).
-\end{equation}
-Then,
-\begin{equation}
-\etarho^{\nph,\star} = {\rm {\bf Average}}\left(\etarho^{{\rm cart},\nph}\right).
-\end{equation}
-\end{description}
-
-\subsection{Computing $w_0$ For Spherical Problems}
-Recall that we want to solve
-\begin{equation}
-\frac{1}{r^2}\frac{\partial}{\partial r}(r^2 w_0) = \Sbar - \frac{1}{\gammaonebar p_0}\left(\frac{\partial p_0}{\partial t} + w_0\frac{\partial p_0}{\partial r}\right)
-\end{equation}
-First, note that $\partial p_0/\partial r = -\rho_0 g$.
-The key observation here is that we can compute $\partial p_0/\partial t$ directly instead of making complicated substitutions involving $\eta_\rho$ to eliminate it.
-Let's discretize this as follows:
-\begin{equation}
-\frac{1}{r_j^2}\left[\left(r^2 w_0\right)_{j+\myhalf} - \left(r^2 w_0\right)_{j-\myhalf}\right] = 
-\Sbar_j - \frac{1}{\left(\gammaonebar p_0\right)_j}\left[\left(\frac{\partial p_0}{\partial t}\right)_j - \left(\frac{w_{0,j-\myhalf}+w_{0,j+\myhalf}}{2}\right)(\rho_0 g)_j\right]
-\end{equation}
-Rearranging this to solve for $w_{0,j+\myhalf}$ gives us an explicit update:
-\begin{equation}
-\left[\frac{r_{j+\myhalf}^2}{r_j^2} - \frac{(\rho_0 g)_j}{2\left(\gammaonebar p_0\right)_j}\right] w_{0,j+\myhalf} =
-\left[\frac{r_{j-\myhalf}^2}{r_j^2} + \frac{(\rho_0 g)_j}{2\left(\gammaonebar p_0\right)_j}\right] w_{0,j-\myhalf}
-+ \Sbar_j - \frac{1}{\left(\gammaonebar p_0\right)_j}\left(\frac{\partial p_0}{\partial t}\right)_j
-\end{equation}
-As before, once $\rho_0$ falls below $\rho_{\rm cutoff}$, we hold $r^2 w_0$ constant.\\ \\
-The interface to the routine in the algorithm description is
-{\bf Make $\mathbf{w_0}$}$[\rho_0,p_0,\partial p_0/\partial t,\gammaonebar,\Sbar] \rightarrow [w_0]$
-
-\acknowledgements
-
-The work at LBNL was supported by the U.S. Department of Energy's Scientific Discovery Through Advanced Computing (SciDAC) program under contract No. DE-AC02-05CH11231.
-This research used resources of the National Energy Research Scientific Computing Center (NERSC), a U.S.
-Department of Energy Office of Science User Facility operated under Contract No.
-DE-AC02-05CH11231.
-
-\bibliographystyle{aasjournal.bst}
-\bibliography{references}
-
-\end{document}

### 2019-03-13 MAESTROeX_I/intro.tex
split paper into multiple tex files for multi-user edits

--- 2019-03-13 MAESTROeX_I/paper.tex
@@
-%% This manuscript uses the AASTeX v6.2 LaTeX 2e macros
-%%
-%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls
-%% (Copyright 2000-2015).  See the classfile for details.
-
-%% AASTeX requires revtex4-1.cls (http://publish.aps.org/revtex4/) and
-%% other external packages (latexsym, graphicx, amssymb, longtable, and epsf).
-%% All of these external packages should already be present in the modern TeX
-%% distributions.  If not they can also be obtained at www.ctan.org.
-
-%% The first piece of markup in an AASTeX v6.x document is the \documentclass
-%% command. LaTeX will ignore any data that comes before this command. The
-%% documentclass can take an optional argument to modify the output style.
-%% The command below calls the preprint style  which will produce a tightly
-%% typeset, one-column, single-spaced document.  It is the default and thus
-%% does not need to be explicitly stated.
-%%
-%%
-%% using aastex version 6.2
-\documentclass{aastex62}
-
-%% The default is a single spaced, 10 point font, single spaced article.
-%% There are 5 other style options available via an optional argument. They
-%% can be envoked like this:
-%%
-%% \documentclass[argument]{aastex62}
-%%
-%% where the layout options are:
-%%
-%%  twocolumn   : two text columns, 10 point font, single spaced article.
-%%                This is the most compact and represent the final published
-%%                derived PDF copy of the accepted manuscript from the publisher
-%%  manuscript  : one text column, 12 point font, double spaced article.
-%%  preprint    : one text column, 12 point font, single spaced article.
-%%  preprint2   : two text columns, 12 point font, single spaced article.
-%%  modern      : a stylish, single text column, 12 point font, article with
-%% 		  wider left and right margins. This uses the Daniel
-%% 		  Foreman-Mackey and David Hogg design.
-%%  RNAAS       : Preferred style for Research Notes which are by design
-%%                lacking an abstract and brief. DO NOT use \begin{abstract}
-%%                and \end{abstract} with this style.
-%%
-%% Note that you can submit to the AAS Journals in any of these 6 styles.
-%%
-%% There are other optional arguments one can envoke to allow other stylistic
-%% actions. The available options are:
-%%
-%%  astrosymb    : Loads Astrosymb font and define \astrocommands.
-%%  tighten      : Makes baselineskip slightly smaller, only works with
-%%                 the twocolumn substyle.
-%%  times        : uses times font instead of the default
-%%  linenumbers  : turn on lineno package.
-%%  trackchanges : required to see the revision mark up and print its output
-%%  longauthor   : Do not use the more compressed footnote style (default) for
-%%                 the author/collaboration/affiliations. Instead print all
-%%                 affiliation information after each name. Creates a much
-%%                 long author list but may be desirable for short author papers
-%%
-%% these can be used in any combination, e.g.
-%%
-%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex62}
-%%
-%% AASTeX v6.* now includes \hyperref support. While we have built in specific
-%% defaults into the classfile you can manually override them with the
-%% \hypersetup command. For example,
-%%
-%%\hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
-%%
-%% will change the color of the internal links to red, the links to the
-%% bibliography to green, the file links to cyan, and the external links to
-%% magenta. Additional information on \hyperref options can be found here:
-%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
-%%
-%% If you want to create your own macros, you can do so
-%% using \newcommand. Your macros should appear before
-%% the \begin{document} command.
-%%
-\newcommand{\vdag}{(v)^\dagger}
-\newcommand\aastex{AAS\TeX}
-\newcommand\latex{La\TeX}
-
-% for non-stacked fractions
-\newcommand{\sfrac}[2]{\mathchoice
-  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
-  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
-  {\kern0em\raise.5ex\hbox{\the\scriptscriptfont0 #1}\kern-.2em/
-   \kern-.15em\lower.25ex\hbox{\the\scriptscriptfont0 #2}}
-  {#1\!/#2}}
-
-\newcommand{\myhalf}{\sfrac{1}{2}}
-\newcommand{\thalf}{\sfrac{3}{2}}
-
-\newcommand{\eb}{{\bf{e}}}
-\newcommand{\Ub}{{\bf{U}}}
-\newcommand{\Ubt}{\widetilde{\Ub}}
-\newcommand{\Vb}{{\bf{V}}}
-\newcommand{\xb}{{\bf{x}}}
-
-\newcommand{\dr}{\Delta r}
-\newcommand{\dt}{\Delta t}
-
-\newcommand{\etarho}{\eta_\rho}
-\newcommand{\gammaonebar}{\overline{\Gamma}_1}
-\newcommand{\Hnuc}{H_{\rm nuc}}
-\newcommand{\omegadot}{\dot\omega}
-\newcommand{\pred}{{\rm pred}}
-\newcommand{\Sbar}{\overline{S}}
-
-\newcommand{\inp}{\mathrm{in}}
-\newcommand{\outp}{\mathrm{out}}
-\newcommand{\nph}{{n+\myhalf}}
-\newcommand{\nmh}{{n-\myhalf}}
-\newcommand{\ow}{\overline{w_0}}
-\newcommand{\dw}{\delta w_0}
-\newcommand{\uadvone}{\Ub^{\mathrm{ADV},\star}}
-\newcommand{\uadvonedag}{\Ub^{\mathrm{ADV},\dagger,\star}}
-\newcommand{\uadvtwo}{\Ub^{\mathrm{ADV}}}
-\newcommand{\uadvtwodag}{\Ub^{\mathrm{ADV},\dagger}}
-\newcommand{\gcc}{\mathrm{g~cm^{-3} }}
-
-% for the red MarginPars
-\usepackage{color}
-% make the MarginPars look pretty
-\setlength{\marginparwidth}{0.5in}
-\newcommand{\MarginPar}[1]{\marginpar{\vskip-\baselineskip\raggedright\tiny\sffamily
-\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}
-
-%% Reintroduced the \received and \accepted commands from AASTeX v5.2
-\received{XXX X, XXXX}
-\revised{XXX X, XXXX}
-\accepted{XXX X, XXXX}
-%% Command to document which AAS Journal the manuscript was submitted to.
-%% Adds "Submitted to " the arguement.
-\submitjournal{ApJ}
-
-%% Mark up commands to limit the number of authors on the front page.
-%% Note that in AASTeX v6.2 a \collaboration call (see below) counts as
-%% an author in this case.
-%
-%\AuthorCollaborationLimit=3
-%
-%% Will only show Schwarz, Muench and "the AAS Journals Data Scientist
-%% collaboration" on the front page of this example manuscript.
-%%
-%% Note that all of the author will be shown in the published article.
-%% This feature is meant to be used prior to acceptance to make the
-%% front end of a long author article more manageable. Please do not use
-%% this functionality for manuscripts with less than 20 authors. Conversely,
-%% please do use this when the number of authors exceeds 40.
-%%
-%% Use \allauthors at the manuscript end to show the full author list.
-%% This command should only be used with \AuthorCollaborationLimit is used.
-
-%% The following command can be used to set the latex table counters.  It
-%% is needed in this document because it uses a mix of latex tabular and
-%% AASTeX deluxetables.  In general it should not be needed.
-%\setcounter{table}{1}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% The following section outlines numerous optional output that
-%% can be displayed in the front matter or as running meta-data.
-%%
-%% If you wish, you may supply running head information, although
-%% this information may be modified by the editorial offices.
-\shorttitle{MAESTROeX Low Mach Number Astrophysics}
-\shortauthors{Fan et al.}
-%%
-%% You can add a light gray and diagonal water-mark to the first page
-%% with this command:
-% \watermark{text}
-%% where "text", e.g. DRAFT, is the text to appear.  If the text is
-%% long you can control the water-mark size with:
-%  \setwatermarkfontsize{dimension}
-%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% This is the end of the preamble.  Indicate the beginning of the
-%% manuscript itself with \begin{document}.
-
-\begin{document}
-
-\title{MAESTROeX: A Massively Parallel Low Mach Number Astrophysical Solver}
-
-%% LaTeX will automatically break titles if they run longer than
-%% one line. However, you may use \\ to force a line break if
-%% you desire. In v6.2 you can include a footnote in the title.
-
-%% A significant change from earlier AASTEX versions is in the structure for
-%% calling author and affilations. The change was necessary to implement
-%% autoindexing of affilations which prior was a manual process that could
-%% easily be tedious in large author manuscripts.
-%%
-%% The \author command is the same as before except it now takes an optional
-%% arguement which is the 16 digit ORCID. The syntax is:
-%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
-%%
-%% This will hyperlink the author name to the author's ORCID page. Note that
-%% during compilation, LaTeX will do some limited checking of the format of
-%% the ID to make sure it is valid.
-%%
-%% Use \affiliation for affiliation information. The old \affil is now aliased
-%% to \affiliation. AASTeX v6.2 will automatically index these in the header.
-%% When a duplicate is found its index will be the same as its previous entry.
-%%
-%% Note that \altaffilmark and \altaffiltext have been removed and thus
-%% can not be used to document secondary affiliations. If they are used latex
-%% will issue a specific error message and quit. Please use multiple
-%% \affiliation calls for to document more than one affiliation.
-%%
-%% The new \altaffiliation can be used to indicate some secondary information
-%% such as fellowships. This command produces a non-numeric footnote that is
-%% set away from the numeric \affiliation footnotes.  NOTE that if an
-%% \altaffiliation command is used it must come BEFORE the \affiliation call,
-%% right after the \author command, in order to place the footnotes in
-%% the proper location.
-%%
-%% Use \email to set provide email addresses. Each \email will appear on its
-%% own line so you can put multiple email address in one \email call. A new
-%% \correspondingauthor command is available in V6.2 to identify the
-%% corresponding author of the manuscript. It is the author's responsibility
-%% to make sure this name is also in the author list.
-%%
-%% While authors can be grouped inside the same \author and \affiliation
-%% commands it is better to have a single author for each. This allows for
-%% one to exploit all the new benefits and should make book-keeping easier.
-%%
-%% If done correctly the peer review system will be able to
-%% automatically put the author and affiliation information from the manuscript
-%% and save the corresponding author the trouble of entering it by hand.
-
-\correspondingauthor{Doreen Fan; DFan@lbl.gov}
-
-\author[0000-0002-3246-4315]{Duoming Fan}
-\affil{Lawrence Berkelely National Laboratory \\
-Center for Computational Sciences and Engineering \\
-One Cyclotron Road, MS 50A-3111 \\
-Berkeley, CA 94720, USA}
-
-\author[0000-0003-1791-0265]{Andrew Nonaka}
-\affil{Lawrence Berkelely National Laboratory \\
-Center for Computational Sciences and Engineering \\
-One Cyclotron Road, MS 50A-3111 \\
-Berkeley, CA 94720, USA}
-
-\author[0000-0003-2103-312X]{Ann S. Almgren}
-\affil{Lawrence Berkelely National Laboratory \\
-Center for Computational Sciences and Engineering \\
-One Cyclotron Road, MS 50A-3111 \\
-Berkeley, CA 94720, USA}
-
-\author[0000-0002-1530-781X]{Alice Harpole}
-\affil{Stony Brook University \\
-Department of Physics and Astronomy \\
-Stony Brook, NY 11794-3800, USA}
-
-\author[0000-0001-8401-030X]{Michael Zingale}
-\affil{Stony Brook University \\
-Department of Physics and Astronomy \\
-Stony Brook, NY 11794-3800, USA}
-
-%% Note that the \and command from previous versions of AASTeX is now
-%% depreciated in this version as it is no longer necessary. AASTeX
-%% automatically takes care of all commas and "and"s between authors names.
-
-%% AASTeX 6.2 has the new \collaboration and \nocollaboration commands to
-%% provide the collaboration status of a group of authors. These commands
-%% can be used either before or after the list of corresponding authors. The
-%% argument for \collaboration is the collaboration identifier. Authors are
-%% encouraged to surround collaboration identifiers with ()s. The
-%% \nocollaboration command takes no argument and exists to indicate that
-%% the nearby authors are not part of surrounding collaborations.
-
-%% Mark off the abstract in the ``abstract'' environment.
-\begin{abstract}
-We present MAESTROeX, a massively parallel solver for low Mach number astrophysical flows.
-Our model equations allow for long-time integration for highly subsonic flows compared to compressible approaches.
-The code leverages the new AMReX framework for block-structured adaptive mesh refinement calculations, and also provides additional algorithmic features from the original MAESTRO code.
-We present a new temporal integration option that is much simpler than our previous approach while retaining the same order order of accuracy.
-We also provide a more accurate spatial discretization for mapping the base state onto the full state Cartesian grid.
-Using our previous studies on the convective phase of single-degenerate progenitor models of Type Ia supernovae, we characterize the performance of the code and validate the new algorithmic features.
-\end{abstract}
-
-%% Keywords should appear after the \end{abstract} command.
-%% See the online documentation for the full list of available subject
-%% keywords and the rules for their use.
-\keywords{convection, hydrodynamics, methods: numerical, nuclear reactions, nucleosynthesis, abundances, supernovae: general}
-
-%% From the front matter, we move on to the body of the paper.
-%% Sections are demarcated by \section and \subsection, respectively.
-%% Observe the use of the LaTeX \label
-%% command after the \subsection to give a symbolic KEY to the
-%% subsection for cross-referencing in a \ref command.
-%% You can use LaTeX's \ref and \label commands to keep track of
-%% cross-references to sections, equations, tables, and figures.
-%% That way, if you change the order of any elements, LaTeX will
-%% automatically renumber them.
-%%
-%% We recommend that authors also use the natbib \citep
-%% and \citet commands to identify citations.  The citations are
-%% tied to the reference list via symbolic KEYs. The KEY corresponds
-%% to the KEY in the \bibitem in the reference list below.
-
 \section{Introduction} \label{sec:intro}
@@
 We also present a new treatment of the base state mapping to Cartesian grids for spherical problsm.
-
-\section{Governing Equations}
-The derivation of the low Mach number hydrodynamics equations is given in Papers I-III.
-We take the standard equations of reacting, compressible flow, and recast the equation
-of state (EOS) as a divergence constraint on the velocity field.  This is done by taking
-the Langrangian derivative of the equation of state for pressure as a function of the
-density, composition, and enthalpy, enforcing the constraint that the pressure is
-a prescribed function of altitutde and time based on the hydrostatic equilibrium condition,
-and substituting the equations for species and enthalpy in the resulting Lagrangian derivative.
-The resulting equations are
-\begin{eqnarray}
-\frac{\partial\Ub}{\partial t} &=& -\Ub\cdot\nabla\Ub  - \frac{1}{\rho}\nabla\pi - \frac{\rho-\rho_0}{\rho} g\eb_r,\label{eq:momentum}\\
-\frac{\partial(\rho X_k)}{\partial t} &=& -\nabla\cdot(\rho X_k\Ub) + \rho\omegadot_k,\label{eq:species}\\
-\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc,\label{eq:enthalpy}
-\end{eqnarray}
-\begin{equation}
-\nabla\cdot(\beta_0\Ub) = \beta_0\left(S - \frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial t}\right).\label{eq:U divergence}
-\end{equation}
-Here $\rho$, $\Ub$, and $h$ are the mass density,
-velocity and specific enthalpy, respectively, and
-$X_k$ are the mass fractions of species $k$ with associated
-production rate $\omegadot_k$.  The species are constrained
-such that $\sum_k X_k = 1$ giving $\rho = \sum_k (\rho X_k)$ and
-\begin{equation}
-\frac{\partial\rho}{\partial t} = -\nabla\cdot(\rho\Ub).
-\end{equation}
-Here $\Hnuc$ is the nuclear energy generation rate per unit mass.
-The pressure is decomposed into a hydrostatic base state
- pressure, $p_0 = p_0(r,t)$, and a dynamic pressure, $\pi = \pi(\xb,t)$, such that 
-$|\pi|/p_0 = \mathcal{O}(M^2)$ (we use $\xb$ to represent the Cartesian coordinate 
-directions of the full state and $r$ to represent the radial coordinate direction for 
-the base state).  We also define a base state density, $\rho_0 = \rho_0(r,t)$, 
-which is in hydrostatic equilibrium with $p_0$, i.e., 
-$\nabla p_0 = -\rho_0 g\eb_r$, where $g=g(r,t)$ is
-the magnitude of the gravitational acceleration and $\eb_r$ is the unit vector in the
-outward radial direction.
-
-Mathematically, equations (\ref{eq:momentum})-(\ref{eq:enthalpy}) must still be closed by the equation of state which we
-express as a divergence constraint on the velocity field (\ref{eq:U divergence}),
-\MarginPar{This is different from the Jacobs 2016 paper where we switched to alt\_energy\_fix.}
-where $\beta_0$ is a density-like variable that carries background stratification, defined as
-\begin{equation}
-\beta_0(r,t) = \rho_0(0,t)\exp\left(\int_0^r\frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial r'}dr'\right),
-\end{equation}
-where $\gammaonebar$ is the lateral average of $\Gamma_1 = d(\log p)/d(\log\rho)$ at constant entropy.
-The expansion term, $S$, incorporates local compressibility effects due to heat release from reactions, compositional changes, and external sources,
-\begin{equation}
-S = -\sigma\sum_k\xi_k\omegadot_k + \frac{1}{\rho p_\rho}\sum_k p_{X_k}\omegadot_k + \sigma\Hnuc,\label{eq:S}
-\end{equation}
-where $p_{X_k} \equiv \left. \partial p / \partial X_k
-\right|_{\rho,T,X_{j,j\ne k}}$, $\xi_k \equiv \left. \partial h /
-\partial X_k \right |_{p,T,X_{j,j\ne k}},
-p_\rho \equiv \left.
-\partial p/\partial \rho \right |_{T, X_k}$, and $\sigma \equiv
-p_T/(\rho c_p p_\rho)$, with $p_T \equiv \left. \partial p / \partial
-T \right|_{\rho, X_k}$ and $c_p \equiv \left.  \partial h / \partial T
-\right|_{p,X_k}$ is the specific heat at constant pressure.
-
-Previously we adopted an approach where we split the velocity into a base state component, $w_0(r,t)$, 
-and a local velocity $\Ubt(\xb,t)$, so that
-\begin{equation}
-\Ub = \Ubt(\xb,t) + w_0(r,t)\eb_r,
-\end{equation}
-We used $w_0$ to provide an estimate of the base state density evolution over a time step.
-In our new simplified approach, we instead use a predictor-corrector approach for evolving the base state.
-
-\section{Numerical Algorithm}
-
-\subsection{Spatial Discretization}
-
-(description of base state mapping in planar, and origianl spherical)
-
-\subsection{Exact Base State Mapping} \label{sec:exactbasestate}
-
-\subsection{Temporal Integration Scheme}\label{Sec:Temporal Integration Scheme}
-Our new temporal integration scheme is much simpler than the scheme from Paper V.
-The primary reason is that we no longer split the velocity field into a perturbational and base state components.  
-Previously, we used the base state velocity to predict the expansion of the base state.  
-In this new scheme we use a simpler, predictor-corrector approach to the base state density and pressure that no longer requires the complex algorithm from Paper V, yet still retains the same overall second-order accuracy.
-
-We now describe the new temporal integration scheme, which has been greatly simplified
-from Paper V without formal loss of order of accuracy.
-Note that this temporal integration scheme is valid for the original base state mapping (with constant base state grid spacing), or the new irregularly spaced base state mapping.
-
-At the beginning of each time step we have the cell-centered state,
-$(\Ub,\rho X_k,\rho h,\rho_0,p_0)^n$, and nodal state, $\pi^{n-\myhalf}$.
-At any time, the associated density, composition, and enthalpy can be trivially computed using, e.g.,
-\begin{equation}
-\rho^n = \sum_k(\rho X_k)^n, \quad
-X_k^n = (\rho X_k)^n / \rho^n, \quad
-h^n = (\rho h)^n / \rho^n.
-\end{equation}
-Temperature is computed using the equation of state, e.g.,
-\footnote{As described in Paper V, for planar problems we compute temperature using $h$ instead of $p_0$, since we have successfully developed planar volume discrepancy schemes to effectively couple the enthalpy to the rest of the solution.
-We are still exploring this option for spherical stars.}
-\begin{equation}
-T = T(\rho,p_0,X_k),
-\end{equation}
-and ($\gammaonebar,\beta_0)$ are computed from $(\rho,p_0,X_k)$ (see Appendix A of Paper I and Appendix C of Paper III for details).
-
-The overall flow of the algorithm is a second-order Strang splitting approach for the coupling of the reactions and advection of the thermodynamic variables.  
-We use a predictor-corrector approach within the Strang splitting scheme to achieve second-order accuracy in time.
-We integrate the velocity using a standard second-order projection method to enforce the divergence constraint.
-To summarize:
-\begin{itemize}
-\item In {\bf Step 1} we react the thermodynamic variables over $\Delta t/2$.
-\item In {\bf Steps 2-4} we advect the thermodynamic variables over $\Delta t$.
-Specifically, we compute an estimate for the expansion term, $S$, project the velocities so they satisfy the divergence constraint, and then advect the thermodynamic variables.
-\item In {\bf Step 5} we react the thermodynamic variables over $\Delta t/2$.\footnote{Note that we could skip to the velocity advance in {\bf Steps 10-11}, however the overall scheme would be only first-order in time, so {\bf Steps 6-9} can be thought of as a trapezoidal corrector step.}
-\item In {\bf Steps 6-8} we redo the advection in {\bf Steps 2-4} but are able to use the trapezoidal rule to time-center certain quantities such as $S$, $\rho_0$, etc.
-\item In {\bf Step 9} we redo the reactions from {\bf Step 5} using the improved results for the corrector advection step.
-\item In {\bf Steps 10-11} we update the velocity.
-\end{itemize}
-
-There are a few key numerical modules we use in each time step.
-\begin{itemize}
-\item {\bf Average} computes...
-\item {\bf Enforce HSE} computes the base state pressure, $p_0$, from a base state density, $\rho_0$ by integrating the hydrostatic equilibrium condition in 1D (see equation (A10) in Paper V).
-The base state pressure remains equal to a constant value from the location of a prescribed cutoff density outward for the entire simulation.
-\item {\bf React State} advances the species and enthalpy due to reactions over $\Delta t/2$ by integrating
-\begin{equation}
-\frac{dX_k}{dt} = \dot\omega_k(\rho,X_k,T); \qquad
-\frac{dT}{dt} = \frac{1}{c_p}\left(-\sum_k\xi_k\dot\omega_k + H_{\rm nuc}\right)
-\end{equation}
-The inputs are the species, enthalpy, and base state pressure, and the outpus are the species, enthalpy, reaction rates, and nuclear energy generation rate.
-See Papers III and V for details.
-\end{itemize}
-
-Each time step satisfies the standard advective CFL condition,
-\begin{equation}
-\Delta t = \sigma^{\rm CFL} \min_i(\Delta x / U_i),
-\end{equation}
-where for our simulations we typically use $\sigma^{\rm CFL}\sim 0.5$.
-There are additional constraints on the time involving the buoyancy force (commonly used when the velocity is approximately zero at the start of some simulations) and the divergence constraint (to prevent too much mass evacuation from a cell in a time step); see Section 3.4 in Paper III for details.
-
-At the beginning of each simulation, we define $(\Ub,\rho X_k,\rho h)$...
-
-\MarginPar{mention cutoff densities}
-
-\MarginPar{AJN: eliminated references to volume discrepancy for now; we'd only use it for planar at the moment, don't want to open up that can of worms here}
-
-\begin{description}
-
-%--------------------------------------------------------------------------
-% STEP 1
-%--------------------------------------------------------------------------
-\item[Step 1] {\em React the full state through the first time interval of $\dt / 2.$}
-
-Call {\bf React State}$[(\rho X_k)^n, (\rho h)^n, p_0^n] \rightarrow [(\rho X_k)^{(1)}, (\rho h)^{(1)}, (\rho \omegadot_k)^{(1)}, (\rho \Hnuc)^{(1)}]$.
-
-%--------------------------------------------------------------------------
-% STEP 2
-%--------------------------------------------------------------------------
-
-\item[Step 2] {\em Compute the provisional time-centered expansion,
-    $S^{\nph,\star}$.}
-
-We compute an estimate for the time-centered expansion term in the velocity
-divergence constraint.  Following \citet{Bell:2004}, we extrapolate
-to the half-time using $S$ at the previous and current
-time steps,
-\begin{equation}
-S^{\nph,\star} = S^n + \frac{\dt^n}{2} \left(\frac{S^n - S^{n-1}}{\dt^{n-1}}\right).
-\end{equation}
-Note that in the first time step we average $S^0$ and $S^1$ from the
-initialization step.
-
-%--------------------------------------------------------------------------
-% STEP 3
-%--------------------------------------------------------------------------
-\item[Step 3] {\em Construct the provisional time-centered advective velocity on
-edges, $\uadvone$.}
-
-The construction of face-centered time-centered states used to discretize the
-advection terms for velocity, species, and enthalpy, are performed using
-a standard multidimensional corner transport upwind approach
-\citep{colella1990multidimensional,saltzman1994unsplit} with piecewise-parabolic
-one-dimensional tracing \citep{colella1984piecewise}.  The full details of this
-Godunov advection approach for all steps in this algorithm are described 
-in Appendix A of \cite{XRB_III}.
-
-In this step, using equation (\ref{eq:momentum}), 
-we compute time-centered edge velocities, $\uadvonedag$, using
-$\Ub^n$.  The $\dagger$ superscript refers to the
-fact that the predicted velocity field does not satisfy the divergence
-constraint.  We then construct $\uadvone$ from $\uadvonedag$
-using a MAC projection.
-We note that $\uadvone$ satisfies the divergence constraint
-\begin{equation}
-\nabla \cdot \left(\beta_0^n \uadvone\right) = \beta_0^n \left[S^{\nph,\star} - \frac{1}{\gammaonebar^np_0^n}\left(\frac{\partial p_0}{\partial t}\right)^{n-\myhalf} \right].\label{eq:div1}
-\end{equation}
-In practice we solve for.
-
-%--------------------------------------------------------------------------
-% STEP 4
-%--------------------------------------------------------------------------
-\item[Step 4] {\em Advect the full state through a time interval of $\dt.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \Alph{enumi}}}
-
-\item Update $(\rho X_k)$ using a discretized version of
-%
-\begin{equation}
-\frac{\partial(\rho X_k)}{\partial t} = -\nabla\cdot(\rho X_k\Ub) + \rho\omegadot_k,
-\end{equation}
-%
-omitting the reaction terms, which were already
-accounted for in {\bf React State}.  The update consists of two steps:
-
-\begin{enumerate}
-\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-\item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\pred}$,
-  for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
-\MarginPar{mention $\rho'$ and $X$ advection; $(\rho h)'$ advection}
-
-\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2),\star}$ using
-\begin{equation}
-(\rho X_k)^{(2),\star} = (\rho X_k)^{(1)}
-  - \dt \left\{ \nabla \cdot \left[ \uadvone (\rho X_k)^{\nph,\pred} \right] \right\},
-\end{equation}
-
-\end{enumerate}
-
-\item Update $\rho_0$ by setting $\rho_0^{n+1,\star} =$ {\bf Average}$(\rho^{(2),\star})$.
-
-\item Update $p_0$ by calling
-{\bf Enforce HSE}$[\rho_0^{n+1,\star}] \rightarrow [p_0^{n+1,\star}]$.
-
-\item Update the enthalpy using a discretized version of equation
-%
-\begin{equation}
-\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc,
-\end{equation}
-%
-again omitting the reaction and heating terms
-since we already accounted for
-them in {\bf React State}.  This equation takes the form:
-\begin{equation}
-\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h) + \frac{\partial p_0}{\partial t} + (\Ub \cdot \eb_r) \frac{\partial p_0}{\partial r}.
-\end{equation}
-For spherical geometry, we solve the
-analytically equivalent form,
-\begin{equation}
-\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h) + \frac{\partial p_0}{\partial t} + \nabla \cdot (\Ub p_0) - p_0 \nabla \cdot \Ub.
-\end{equation}
-The update consists of two steps:
-
-\begin{enumerate}
-\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-\item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\pred},$
-  for the conservative update of $(\rho h)^{(1)}$
-  using using a Godunov approach \citep{XRB_III}.
-
-\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$.
-\begin{description}
-\item For planar geometry,
-\begin{eqnarray}
-(\rho h)^{(2),\star}
-&=& (\rho h)^{(1)} \nonumber \\
-&&- \dt \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \left(p_0^{n+1,*} - p_0^n\right) \nonumber \\
-&& + \dt \left(\uadvone \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^{n}
-\end{eqnarray}
-
-\item For spherical geometry,
-\begin{eqnarray}
-(\rho h)^{(2),\star}
-&=& (\rho h)^{(1)} \nonumber \\
-&&- \dt \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \left(p_0^{n+1,*} - p_0^n\right) \nonumber \\
-&& + \dt \left \{ \nabla \cdot \left (\uadvone p_0^{n+\myhalf} \right ) - p_0^{n+\myhalf} \nabla \cdot \uadvone \right \}
-\end{eqnarray}
-\end{description}
-where we define $p_0^\nph$ as $p_0^\nph = (p_0^n+p_0^{n+1,*})/2$.
-
-\end{enumerate}
-Note that outside the cutoff density region, we compute enthalpy using the EOS.
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 5
-%--------------------------------------------------------------------------
-\item[Step 5] {\em React the full state through a second time interval of $\dt / 2.$}
-
-Call {\bf React State}$[ (\rho X_k)^{(2),\star}, (\rho h)^{(2),\star}, p_0^{n+1,\star}] 
-\rightarrow 
-[ (\rho X_k)^{n+1,\star}, (\rho h)^{n+1,\star}, (\rho \omegadot_k)^{n+1,\star}, (\rho \Hnuc)^{n+1,\star} ].$
-
-%--------------------------------------------------------------------------
-% STEP 6
-%--------------------------------------------------------------------------
-\item[Step 6] {\em Compute the time-centered expansion, $S^{\nph,\star}$.}
-
-First, compute $S^{n+1,\star}$ with
-\begin{equation}
-S^{n+1,\star} =  \left(-\sigma  \sum_k  \xi_k  \omegadot_k  + \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k + \sigma \Hnuc\right)^{n+1,\star}.
-\end{equation}
-  Then, define
-\begin{equation}
- S^{\nph} = \frac{S^n + S^{n+1,\star}}{2},
-\end{equation}
-
-%--------------------------------------------------------------------------
-% STEP 7
-%--------------------------------------------------------------------------
-\item[Step 7] {\em Construct the time-centered advective velocity on edges, $\uadvtwo$.}
-
-The procedure to construct $\uadvtwodag$ is identical to the Godunov procedure
-for computing $\uadvonedag$ in {\bf Step 3}, but uses
-the updated value $S^{\nph}$ rather than $S^{\nph,\star}$.
-We note that $\uadvtwo$ satisfies the divergence constraint
-\begin{equation}
-\nabla \cdot \left(\beta_0^{\nph} \uadvtwo\right) =
-\beta_0^{\nph} \left[S^{\nph} - \frac{1}{\gammaonebar^{n+\myhalf}p_0^{n+\myhalf}}\left(\frac{\partial p_0}{\partial t}\right)^{n+\myhalf}\right],\label{eq:div2}
-\end{equation}
-where
-\begin{equation}
-\beta_0^{\nph} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2},
-\qquad
-\gammaonebar^{\nph} = \frac{ \gammaonebar^n +  \gammaonebar^{n+1,\star} }{2}.
-\end{equation}
-
-%--------------------------------------------------------------------------
-% STEP 8
-%--------------------------------------------------------------------------
-\item[Step 8] {\em Advect the base state and full state through a time interval of $\dt.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \Alph{enumi}}}
-
-\item Update $(\rho X_k)$.  This step is identical to {\bf Step 4A} except we use
-  the updated values $\uadvtwo$ and $\rho_0^{n+1,\star}$ rather than
-  $\uadvone$ and $\rho_0^{n+1,\star}$.  In particular:
-
-\begin{enumerate}
-\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-\item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\pred}$,
-  for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
-
-\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2)}$ using
-\begin{equation}
-(\rho X_k)^{(2)} = (\rho X_k)^{(1)}
-- \dt \left\{ \nabla \cdot \left[\uadvtwo (\rho X_k)^{\nph,\pred} \right] \right\},
-\end{equation}
-
-\end{enumerate}
-
-\item Update $\rho_0$ by setting $\rho_0^{n+1} =$ {\bf Average}$(\rho^{(2)})$.
-
-\item Update $p_0$ by calling
-{\bf Enforce HSE}$[\rho_0^{n+1}] \rightarrow [p_0^{n+1}]$.
-
-\item Update the enthalpy.  This step is identical to {\bf Step 4D} except we use
-  the updated values $\uadvtwo, \rho_0^{n+1}, (\rho h)_0^{n+1}$, and $p_0^{n+\myhalf}$
-  rather than
-  $\uadvone, \rho_0^{n+1,\star}, (\rho h)_0^{n+1,\star}$, and $p_0^n$.
-  In particular:
-
-\begin{enumerate}
-\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-\item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\pred},$
-  for the conservative update of $(\rho h)^{(1)}$ using a Godunov approach \citep{XRB_III}.
-
-\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2)}$.
-\begin{description}
-\item For planar geometry,
-\begin{eqnarray}
-(\rho h)^{(2)}
-&=& (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvtwo (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \left( p_0^{n+1} - p_0^n \right)\nonumber \\
-&& + \dt \left(\uadvtwo \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^\nph,
-\end{eqnarray}
-
-\item For spherical geometry,
-\begin{eqnarray}
-(\rho h)^{(2)}
-&=& (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvtwo (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \left(p_0^{n+1} - p_0^n\right)\nonumber \\
-&& + \dt \left[ \nabla \cdot \left (\uadvtwo p_0^{\nph} \right ) - p_0^{\nph} \nabla \cdot \uadvtwo \right],
-\end{eqnarray}
-\end{description}
-where we now define $p_0^\nph$ as $p_0^\nph = (p_0^n+p_0^{n+1})/2$.
-
-\end{enumerate}
-Note that outside the cutoff density region, we compute enthalpy using the EOS.
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 9
-%--------------------------------------------------------------------------
-\item[Step 9] {\em React the full state through a second time interval of $\dt / 2.$}
-
-Call {\bf React State}$[(\rho X_k)^{(2)},(\rho h)^{(2)}, p_0^{n+1}] \rightarrow [(\rho X_k)^{n+1}, (\rho h)^{n+1}, (\rho \omegadot_k)^{n+1}, (\rho \Hnuc)^{n+1}].$
-
-%--------------------------------------------------------------------------
-% STEP 10
-%--------------------------------------------------------------------------
-\item[Step 10] {\em Define the new time expansion, $S^{n+1}$.}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \Alph{enumi}}}
-\item Define
-\begin{equation}
-  S^{n+1} =  \left(-\sigma  \sum_k  \xi_k \omegadot_k  + \sigma \Hnuc +
-  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  \omegadot_k\right)^{n+1}.
-\end{equation}
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 11
-%--------------------------------------------------------------------------
-\item[Step 11] {\em Update the velocity}.
-
-First, we compute the time-centered edge velocities, $\Ub^{\nph,\pred}$
-using a Godunov approach \citep{XRB_III}. Then, we define
-\begin{equation}
-\rho^\nph = \frac{\rho^n + \rho^{n+1}}{2}, \qquad \rho_0^\nph = \frac{\rho_0^n + \rho_0^{n+1}}{2}.
-\end{equation}
-We update the velocity field $\Ub^n$ to $\Ub^{n+1,\dagger}$ by discretizing
-equation (\ref{eq:momentum}) as
-\begin{equation}
-\Ub^{n+1,\dagger}
-= \Ub^n - \dt \left[\uadvtwo \cdot \nabla \Ub^{\nph,\pred} \right]
- - \dt \left[ \frac{1}{\rho^\nph} \nabla \pi^\nmh + \frac{\left(\rho^\nph-\rho_0^\nph\right)}{\rho^\nph} g^{\nph} \eb_r \right],
-\end{equation}
-Again, the $\dagger$ superscript refers
-to the fact that the updated velocity does not satisfy the divergence
-constraint.
-
-Finally, we use an approximate nodal projection to define $\Ub^{n+1}$
-from $\Ub^{n+1,\dagger},$  such that $\Ub^{n+1}$ approximately
-satisfies the divergence constraint,
-\begin{equation}
-\nabla \cdot \left(\beta_0^{n+1} \Ub^{n+1} \right) = \beta_0^{n+1} \left[ S^{n+1} - \frac{1}{\gammaonebar^{n+1}p_0^{n+1}}\left(\frac{\partial p_0}{\partial t}\right)^{n+\myhalf}\right].\label{eq:div3}
-\end{equation}
-As part of the projection we also define the new-time perturbational pressure,
-$\pi^\nph.$  This projection necessarily differs from the MAC projection used in
-{\bf Step 3} and {\bf Step 7} because the velocities in those steps are defined
-on edges and $\Ub^{n+1}$ is defined at cell centers, requiring different divergence
-and gradient operators.  Details of the approximate projection are given in Paper III.
-
-\end{description}
-
-This completes one step of the algorithm.
-
-To initialize the simulation
-
-%--------------------------------------------------------------------------
-% STEP 0
-%--------------------------------------------------------------------------
-{\em Initialization}
-\sout{
-The initialization step only occurs at the beginning of the simulation.
-The initialization step computes initial values for 
-$\pi^{-\myhalf}, \eta_{\rho}^{-\myhalf}, S^0$, and $S^1$ for use in time stepping routine.  
-We pass through {\bf Step 1} -- {\bf Step 11}, but only keep these values.
-The initial values for $\Ub^0, \rho^0, (\rho h)^0, X_k^0, T^0,
-\rho_0^0, p_0^0$, and $\gammaonebar^0$ are specified from the problem-dependent
-initial conditions.  The initial time step, $\dt^0$, is computed as in
-Paper III.
-}
-
-\section{Test Problems}
-
-\subsection{Convergence Test}
-3D reacting bubble rise; to demonstrate 2nd-order convergence of the new temporal algorithm.
-
-\subsection{Scaling}
-Scaling with OpenMP
-
-\subsection{Performance Compared to Original MAESTRO}
-Compare to original MAESTRO
-
-\subsection{White Dwarf Convection}
-white dwarf convection runs with 3 algorithms (original, new temporal, new temporal + irregular base state)
-
-\section{Conclusions and Future Work}
-
-science: rotation, solar physics, MHD
-
-algorithm: SDC \cite{dutt2000spectral}, high-order, multi-implicit \cite{bourlioux2003high}.
-Consider SDC ideas used for terrestrial combustion \cite{pazner2016high,nonaka2018conservative}
-
-implementation: GPUs.  Forward cite next CASTRO paper with GPU strategy.
-
-\appendix
-\section{Projection Details}
-To enforce the divergence constraint in {\bf Step 3, 7}, and {\bf 11}, we use a projection method analgous to the methods originally developed for incompressible flow.
-
-Based on our past experience in the MAESTRO project, we have found it useful to split the velocity dynamics into a perturbational and base state velocity,
-\begin{equation}
-\Ub = \Ubt(\xb,t) + w_0(r,t)\eb_r,
-\end{equation}
-solve for each term separately, and immediately combine them to find a full velocity that satisfies the constraint.
-We take that approach here, primarily because it allows us to enforce a boundary condition on $w_0$ at the edge of the star (i.e., the cutoff density location where we hold density constant).
-Namely, to enforce that $r^2 w_0$ remain constant is difficult to do when solving for the full velocity.
-So in practice, we solve the constraint over the lateral average,
-\begin{equation}
-\nabla\cdot\left(\beta_0\Ubt\right) = \beta_0(S-\overline{S})
-\end{equation}
-and sepratately solve for $w_0$ using 
-\begin{equation}
-\nabla\cdot(\beta_0w_0\eb_r) = \beta_0\left(\overline{S} - \frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial t}\right).
-\end{equation}
-To find $w_0$ we integrate in 1D using the procedure in Appendix B of Paper V.
-
-Note that in this approach, we estimated the time-derivative of the pressure in part by examining how laterally averaged $\rho'=\rho-\rho_0$ changed over time (quantified by $\eta_\rho = \overline{\rho'\Ub\cdot\eb_r}$.
-After evolving the species, we compute this term in the same way as reported in Paper V for use in computing $w_0$.
-
-\section{Old Notes (not for inclusion in paper)}
-
-\subsection{Exact Base State Notes}
-
-\subsubsection{Potential Problems}
-\begin{enumerate}
-\item How do we define \verb|r_edge_loc| for irregularly-spaced base states whose cell-centered values are defined at $r_m=\Delta x\sqrt{\frac{3}{4} + 2m}$?
-\item We would stil need to approximate the values at edges when advecting the base states (i.e.
-$\rho_0^{\nph,\pred}$).
-This may affect the nice properties we would get with exact cell-centered base state values.
-\item Excessive memory storage may be required (\verb|nr_irreg >> nr_fine|).
-\end{enumerate}
-
-\subsubsection{A Possible Solution}
-We could choose to no longer advect the base states for both predictor and corrector steps.
-This would solve the problems with computing state values at cell edges.
-Instead, we propose to only update the base states $\rho_0$ and $p_0$ by computing the average of the full state values after advecting the full states.
-This would result in the following changes in the algorithm from Paper V:
-%
-\begin{enumerate}
-\item Steps 4 and 8 will be greatly simplified, leading to a more efficient algorithm.
-\item \sout{There is no longer a need for $w_0$, so we can use the full velocity $\Ub$ instead of $\Ubt$.
-As a consequence, $\eta$ does not need to be computed.
-This has the added advantage of saving some memory storage by removing many of the terms needed when we decided to split terms into its base state and perturbation state in our previous implementation.} We split the velocity for the projection to handle the cutoff.
-\item By placing the base states exactly at cell-centers, we eliminate interpolation errors and simplify the existing code.
-\end{enumerate}
-%
-
-\subsection{$\eta_\rho$ Notes}
-After {\bf Step 8D}, we define a radial cell-centered $\etarho^{\nph}$.
-
-\begin{description}
-\item For planar geometry, $\etarho = \overline{\rho'(\Ub\cdot\eb_r)}$,
-\begin{equation}
- \etarho^{\nph} =  {\rm {\bf Average}} \sum_k \left[ \left(\uadvtwo \cdot \eb_r \right) (\rho X_k)^{\nph,\pred} \right]
-\end{equation}
-\item For spherical geometry, first construct 
-$\etarho^{{\rm cart},\nph} = [\rho'(\Ub\cdot\eb_r)]^{\nph}$ on Cartesian cell centers using:
-\begin{equation}
-\etarho^{{\rm cart},\nph} = \left[\left(\frac{\rho^{(1)}+\rho^{(2)}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{n+1}}{2}\right)\right] \cdot \left( \uadvtwo \cdot \eb_r \right).
-\end{equation}
-Then,
-\begin{equation}
-\etarho^{\nph,\star} = {\rm {\bf Average}}\left(\etarho^{{\rm cart},\nph}\right).
-\end{equation}
-\end{description}
-
-\subsection{Computing $w_0$ For Spherical Problems}
-Recall that we want to solve
-\begin{equation}
-\frac{1}{r^2}\frac{\partial}{\partial r}(r^2 w_0) = \Sbar - \frac{1}{\gammaonebar p_0}\left(\frac{\partial p_0}{\partial t} + w_0\frac{\partial p_0}{\partial r}\right)
-\end{equation}
-First, note that $\partial p_0/\partial r = -\rho_0 g$.
-The key observation here is that we can compute $\partial p_0/\partial t$ directly instead of making complicated substitutions involving $\eta_\rho$ to eliminate it.
-Let's discretize this as follows:
-\begin{equation}
-\frac{1}{r_j^2}\left[\left(r^2 w_0\right)_{j+\myhalf} - \left(r^2 w_0\right)_{j-\myhalf}\right] = 
-\Sbar_j - \frac{1}{\left(\gammaonebar p_0\right)_j}\left[\left(\frac{\partial p_0}{\partial t}\right)_j - \left(\frac{w_{0,j-\myhalf}+w_{0,j+\myhalf}}{2}\right)(\rho_0 g)_j\right]
-\end{equation}
-Rearranging this to solve for $w_{0,j+\myhalf}$ gives us an explicit update:
-\begin{equation}
-\left[\frac{r_{j+\myhalf}^2}{r_j^2} - \frac{(\rho_0 g)_j}{2\left(\gammaonebar p_0\right)_j}\right] w_{0,j+\myhalf} =
-\left[\frac{r_{j-\myhalf}^2}{r_j^2} + \frac{(\rho_0 g)_j}{2\left(\gammaonebar p_0\right)_j}\right] w_{0,j-\myhalf}
-+ \Sbar_j - \frac{1}{\left(\gammaonebar p_0\right)_j}\left(\frac{\partial p_0}{\partial t}\right)_j
-\end{equation}
-As before, once $\rho_0$ falls below $\rho_{\rm cutoff}$, we hold $r^2 w_0$ constant.\\ \\
-The interface to the routine in the algorithm description is
-{\bf Make $\mathbf{w_0}$}$[\rho_0,p_0,\partial p_0/\partial t,\gammaonebar,\Sbar] \rightarrow [w_0]$
-
-\acknowledgements
-
-The work at LBNL was supported by the U.S. Department of Energy's Scientific Discovery Through Advanced Computing (SciDAC) program under contract No. DE-AC02-05CH11231.
-This research used resources of the National Energy Research Scientific Computing Center (NERSC), a U.S.
-Department of Energy Office of Science User Facility operated under Contract No.
-DE-AC02-05CH11231.
-
-\bibliographystyle{aasjournal.bst}
-\bibliography{references}
-
-\end{document}

### 2019-03-13 MAESTROeX_I/paper.tex
split paper into multiple tex files for multi-user edits

--- 2019-03-13 MAESTROeX_I/paper.tex
@@
 
-\section{Introduction} \label{sec:intro}
+\input{intro.tex}
 
-TODO list:
-\begin{itemize}
-\item Weak scaling tests of original algorithm, comparing MAESTRO to MAESTROeX.  Use cori knl, 4 MPI per node, 16 threads per MPI
-\item Run at effective $512^3$ resolution (maybe to ignition?...
-need to see how max T trends go) (1) classic MAESTRO, (2) MAESTROeX with original algorithm, (3) MAESTROeX with new temporal integrator, (4) MAESTROeX with new temporal integrator and irregular dr, (5) MAESTROeX with original algorithm and 3-levels of AMR, (6) MAESTROeX with new temporal integrator and 3 levels of AMR.
-\item (perhaps) a demonstration of what goes wrong when you don't split the velocity dynamics in the projection
-\end{itemize}
+\input{equations.tex}
 
-Many astrophysical flows are highly subsonic; sound waves carry sufficiently little energy that they do not significantly affect the convective dynamics of the system.
-In many of these flows, modeling long-time convective dynamics are of interest, and numerical approaches based on compressible hydrodynamics are intractable, even on modern supercomputers.
-One approach to this problem is to use low Mach number models.
-In a low Mach number approach, sound waves are eliminated from the governing equations while retaining compressibilitiy effects due to, e.g., nuclear energy release, compositional changes, and thermal diffusion.
-The resulting model can be numerically integrated with much larger time steps than a compressible model.
+\input{algorithm.tex}
 
-Previously, we have developed the low Mach number astrophysical solver, MAESTRO.
-Low Mach number models have been developed for a variety of contexts including combustion \citep{day2000numerical}, atmospherics \citep{duarte2015low}, and elastic solids \citep{abbate2017all}.
-The low Mach number model in MAESTRO is unique in that it is specifically designed for astrophysical settings with significant atmospheric stratification.
-Central to the algorithm is a stratified background (or base) state density and pressure held in hydrostatic equilibrium, and also vary as a function of altitude and time.
-MAESTRO is suitable for for spherical stars, as well as planar simulations of local dynamics.
+\input{results.tex}
 
-The key numerical developments of the original MAESTRO algorithm are presented in a series of papers which we refer to as Papers I-V:
-\begin{itemize}
-\item In Paper I \citep{MAESTRO_I}, we derive the low Mach number equation set from the fully compressible equations.
-\item In Paper II \citep{MAESTRO_II}, we incorporate the effects atmospheric expansion through the use of a time-dependent background state.
-\item In Paper III \citep{MAESTRO_III}, we incorporate reactions and the associated coupling to the hydrodynamics.
-\item In Paper IV \citep{MAESTRO_IV}, we describe our treatment of spherical stars in a three-dimensional Cartesian geometry.
-\item In Paper V \citep{MAESTRO_V}, we describe the use of block-structured adaptive mesh refinement to focus spatial resolution in regions of interest.
-\end{itemize}
+\input{conclusions.tex}
 
-Since then, there have been many scientific investigations using MAESTRO, which include additional algorithmic enhancements.  Topics include:
-\begin{itemize}
-\item The convective phase preceding Chandrasekhar mass models for type Ia supernovae \citep{MAESTRO_convection,MAESTRO_AMR,MAESTRO_CASTRO}.
-\item Convection in massive stars \citep{Gilet:2013}.
-\item Sub-Chandrasekhar white dwarfs \citep{subChandra_I,subChandra_II}.
-In the latter paper we introduced an optional modification to the momentum equation of velocity constraint that conserves total energy, can give results closer to compressible codes of convective dynamics, including low density regions at the surface of a star.
-\item Type I X-ray bursts \citep{XRB_I,XRB_II,XRB_III}.
-In \cite{XRB_I} we discuss the modifications required for implicit thermal conduction, as well as an additional ``volume discrepancy'' modification to the velocity constraint to force the species and enthalpy to evolve in a manner consistent with the background pressure.
-\end{itemize}
-
-In this paper, we characterize the performance of the MAESTROeX algorithm compared to the previous implementation.
-We also develop simpler, alternative temporal integrators, with a longer term goal of developing high-order coupling schemes.
-We also present a new treatment of the base state mapping to Cartesian grids for spherical problsm.
-
-\section{Governing Equations}
-The derivation of the low Mach number hydrodynamics equations is given in Papers I-III.
-We take the standard equations of reacting, compressible flow, and recast the equation
-of state (EOS) as a divergence constraint on the velocity field.  This is done by taking
-the Langrangian derivative of the equation of state for pressure as a function of the
-density, composition, and enthalpy, enforcing the constraint that the pressure is
-a prescribed function of altitutde and time based on the hydrostatic equilibrium condition,
-and substituting the equations for species and enthalpy in the resulting Lagrangian derivative.
-The resulting equations are
-\begin{eqnarray}
-\frac{\partial\Ub}{\partial t} &=& -\Ub\cdot\nabla\Ub  - \frac{1}{\rho}\nabla\pi - \frac{\rho-\rho_0}{\rho} g\eb_r,\label{eq:momentum}\\
-\frac{\partial(\rho X_k)}{\partial t} &=& -\nabla\cdot(\rho X_k\Ub) + \rho\omegadot_k,\label{eq:species}\\
-\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc,\label{eq:enthalpy}
-\end{eqnarray}
-\begin{equation}
-\nabla\cdot(\beta_0\Ub) = \beta_0\left(S - \frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial t}\right).\label{eq:U divergence}
-\end{equation}
-Here $\rho$, $\Ub$, and $h$ are the mass density,
-velocity and specific enthalpy, respectively, and
-$X_k$ are the mass fractions of species $k$ with associated
-production rate $\omegadot_k$.  The species are constrained
-such that $\sum_k X_k = 1$ giving $\rho = \sum_k (\rho X_k)$ and
-\begin{equation}
-\frac{\partial\rho}{\partial t} = -\nabla\cdot(\rho\Ub).
-\end{equation}
-Here $\Hnuc$ is the nuclear energy generation rate per unit mass.
-The pressure is decomposed into a hydrostatic base state
- pressure, $p_0 = p_0(r,t)$, and a dynamic pressure, $\pi = \pi(\xb,t)$, such that 
-$|\pi|/p_0 = \mathcal{O}(M^2)$ (we use $\xb$ to represent the Cartesian coordinate 
-directions of the full state and $r$ to represent the radial coordinate direction for 
-the base state).  We also define a base state density, $\rho_0 = \rho_0(r,t)$, 
-which is in hydrostatic equilibrium with $p_0$, i.e., 
-$\nabla p_0 = -\rho_0 g\eb_r$, where $g=g(r,t)$ is
-the magnitude of the gravitational acceleration and $\eb_r$ is the unit vector in the
-outward radial direction.
-
-Mathematically, equations (\ref{eq:momentum})-(\ref{eq:enthalpy}) must still be closed by the equation of state which we
-express as a divergence constraint on the velocity field (\ref{eq:U divergence}),
-\MarginPar{This is different from the Jacobs 2016 paper where we switched to alt\_energy\_fix.}
-where $\beta_0$ is a density-like variable that carries background stratification, defined as
-\begin{equation}
-\beta_0(r,t) = \rho_0(0,t)\exp\left(\int_0^r\frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial r'}dr'\right),
-\end{equation}
-where $\gammaonebar$ is the lateral average of $\Gamma_1 = d(\log p)/d(\log\rho)$ at constant entropy.
-The expansion term, $S$, incorporates local compressibility effects due to heat release from reactions, compositional changes, and external sources,
-\begin{equation}
-S = -\sigma\sum_k\xi_k\omegadot_k + \frac{1}{\rho p_\rho}\sum_k p_{X_k}\omegadot_k + \sigma\Hnuc,\label{eq:S}
-\end{equation}
-where $p_{X_k} \equiv \left. \partial p / \partial X_k
-\right|_{\rho,T,X_{j,j\ne k}}$, $\xi_k \equiv \left. \partial h /
-\partial X_k \right |_{p,T,X_{j,j\ne k}},
-p_\rho \equiv \left.
-\partial p/\partial \rho \right |_{T, X_k}$, and $\sigma \equiv
-p_T/(\rho c_p p_\rho)$, with $p_T \equiv \left. \partial p / \partial
-T \right|_{\rho, X_k}$ and $c_p \equiv \left.  \partial h / \partial T
-\right|_{p,X_k}$ is the specific heat at constant pressure.
-
-Previously we adopted an approach where we split the velocity into a base state component, $w_0(r,t)$, 
-and a local velocity $\Ubt(\xb,t)$, so that
-\begin{equation}
-\Ub = \Ubt(\xb,t) + w_0(r,t)\eb_r,
-\end{equation}
-We used $w_0$ to provide an estimate of the base state density evolution over a time step.
-In our new simplified approach, we instead use a predictor-corrector approach for evolving the base state.
-
-\section{Numerical Algorithm}
-
-\subsection{Spatial Discretization}
-
-(description of base state mapping in planar, and origianl spherical)
-
-\subsection{Exact Base State Mapping} \label{sec:exactbasestate}
-
-\subsection{Temporal Integration Scheme}\label{Sec:Temporal Integration Scheme}
-Our new temporal integration scheme is much simpler than the scheme from Paper V.
-The primary reason is that we no longer split the velocity field into a perturbational and base state components.  
-Previously, we used the base state velocity to predict the expansion of the base state.  
-In this new scheme we use a simpler, predictor-corrector approach to the base state density and pressure that no longer requires the complex algorithm from Paper V, yet still retains the same overall second-order accuracy.
-
-We now describe the new temporal integration scheme, which has been greatly simplified
-from Paper V without formal loss of order of accuracy.
-Note that this temporal integration scheme is valid for the original base state mapping (with constant base state grid spacing), or the new irregularly spaced base state mapping.
-
-At the beginning of each time step we have the cell-centered state,
-$(\Ub,\rho X_k,\rho h,\rho_0,p_0)^n$, and nodal state, $\pi^{n-\myhalf}$.
-At any time, the associated density, composition, and enthalpy can be trivially computed using, e.g.,
-\begin{equation}
-\rho^n = \sum_k(\rho X_k)^n, \quad
-X_k^n = (\rho X_k)^n / \rho^n, \quad
-h^n = (\rho h)^n / \rho^n.
-\end{equation}
-Temperature is computed using the equation of state, e.g.,
-\footnote{As described in Paper V, for planar problems we compute temperature using $h$ instead of $p_0$, since we have successfully developed planar volume discrepancy schemes to effectively couple the enthalpy to the rest of the solution.
-We are still exploring this option for spherical stars.}
-\begin{equation}
-T = T(\rho,p_0,X_k),
-\end{equation}
-and ($\gammaonebar,\beta_0)$ are computed from $(\rho,p_0,X_k)$ (see Appendix A of Paper I and Appendix C of Paper III for details).
-
-The overall flow of the algorithm is a second-order Strang splitting approach for the coupling of the reactions and advection of the thermodynamic variables.  
-We use a predictor-corrector approach within the Strang splitting scheme to achieve second-order accuracy in time.
-We integrate the velocity using a standard second-order projection method to enforce the divergence constraint.
-To summarize:
-\begin{itemize}
-\item In {\bf Step 1} we react the thermodynamic variables over $\Delta t/2$.
-\item In {\bf Steps 2-4} we advect the thermodynamic variables over $\Delta t$.
-Specifically, we compute an estimate for the expansion term, $S$, project the velocities so they satisfy the divergence constraint, and then advect the thermodynamic variables.
-\item In {\bf Step 5} we react the thermodynamic variables over $\Delta t/2$.\footnote{Note that we could skip to the velocity advance in {\bf Steps 10-11}, however the overall scheme would be only first-order in time, so {\bf Steps 6-9} can be thought of as a trapezoidal corrector step.}
-\item In {\bf Steps 6-8} we redo the advection in {\bf Steps 2-4} but are able to use the trapezoidal rule to time-center certain quantities such as $S$, $\rho_0$, etc.
-\item In {\bf Step 9} we redo the reactions from {\bf Step 5} using the improved results for the corrector advection step.
-\item In {\bf Steps 10-11} we update the velocity.
-\end{itemize}
-
-There are a few key numerical modules we use in each time step.
-\begin{itemize}
-\item {\bf Average} computes...
-\item {\bf Enforce HSE} computes the base state pressure, $p_0$, from a base state density, $\rho_0$ by integrating the hydrostatic equilibrium condition in 1D (see equation (A10) in Paper V).
-The base state pressure remains equal to a constant value from the location of a prescribed cutoff density outward for the entire simulation.
-\item {\bf React State} advances the species and enthalpy due to reactions over $\Delta t/2$ by integrating
-\begin{equation}
-\frac{dX_k}{dt} = \dot\omega_k(\rho,X_k,T); \qquad
-\frac{dT}{dt} = \frac{1}{c_p}\left(-\sum_k\xi_k\dot\omega_k + H_{\rm nuc}\right)
-\end{equation}
-The inputs are the species, enthalpy, and base state pressure, and the outpus are the species, enthalpy, reaction rates, and nuclear energy generation rate.
-See Papers III and V for details.
-\end{itemize}
-
-Each time step satisfies the standard advective CFL condition,
-\begin{equation}
-\Delta t = \sigma^{\rm CFL} \min_i(\Delta x / U_i),
-\end{equation}
-where for our simulations we typically use $\sigma^{\rm CFL}\sim 0.5$.
-There are additional constraints on the time involving the buoyancy force (commonly used when the velocity is approximately zero at the start of some simulations) and the divergence constraint (to prevent too much mass evacuation from a cell in a time step); see Section 3.4 in Paper III for details.
-
-At the beginning of each simulation, we define $(\Ub,\rho X_k,\rho h)$...
-
-\MarginPar{mention cutoff densities}
-
-\MarginPar{AJN: eliminated references to volume discrepancy for now; we'd only use it for planar at the moment, don't want to open up that can of worms here}
-
-\begin{description}
-
-%--------------------------------------------------------------------------
-% STEP 1
-%--------------------------------------------------------------------------
-\item[Step 1] {\em React the full state through the first time interval of $\dt / 2.$}
-
-Call {\bf React State}$[(\rho X_k)^n, (\rho h)^n, p_0^n] \rightarrow [(\rho X_k)^{(1)}, (\rho h)^{(1)}, (\rho \omegadot_k)^{(1)}, (\rho \Hnuc)^{(1)}]$.
-
-%--------------------------------------------------------------------------
-% STEP 2
-%--------------------------------------------------------------------------
-
-\item[Step 2] {\em Compute the provisional time-centered expansion,
-    $S^{\nph,\star}$.}
-
-We compute an estimate for the time-centered expansion term in the velocity
-divergence constraint.  Following \citet{Bell:2004}, we extrapolate
-to the half-time using $S$ at the previous and current
-time steps,
-\begin{equation}
-S^{\nph,\star} = S^n + \frac{\dt^n}{2} \left(\frac{S^n - S^{n-1}}{\dt^{n-1}}\right).
-\end{equation}
-Note that in the first time step we average $S^0$ and $S^1$ from the
-initialization step.
-
-%--------------------------------------------------------------------------
-% STEP 3
-%--------------------------------------------------------------------------
-\item[Step 3] {\em Construct the provisional time-centered advective velocity on
-edges, $\uadvone$.}
-
-The construction of face-centered time-centered states used to discretize the
-advection terms for velocity, species, and enthalpy, are performed using
-a standard multidimensional corner transport upwind approach
-\citep{colella1990multidimensional,saltzman1994unsplit} with piecewise-parabolic
-one-dimensional tracing \citep{colella1984piecewise}.  The full details of this
-Godunov advection approach for all steps in this algorithm are described 
-in Appendix A of \cite{XRB_III}.
-
-In this step, using equation (\ref{eq:momentum}), 
-we compute time-centered edge velocities, $\uadvonedag$, using
-$\Ub^n$.  The $\dagger$ superscript refers to the
-fact that the predicted velocity field does not satisfy the divergence
-constraint.  We then construct $\uadvone$ from $\uadvonedag$
-using a MAC projection.
-We note that $\uadvone$ satisfies the divergence constraint
-\begin{equation}
-\nabla \cdot \left(\beta_0^n \uadvone\right) = \beta_0^n \left[S^{\nph,\star} - \frac{1}{\gammaonebar^np_0^n}\left(\frac{\partial p_0}{\partial t}\right)^{n-\myhalf} \right].\label{eq:div1}
-\end{equation}
-In practice we solve for.
-
-%--------------------------------------------------------------------------
-% STEP 4
-%--------------------------------------------------------------------------
-\item[Step 4] {\em Advect the full state through a time interval of $\dt.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \Alph{enumi}}}
-
-\item Update $(\rho X_k)$ using a discretized version of
-%
-\begin{equation}
-\frac{\partial(\rho X_k)}{\partial t} = -\nabla\cdot(\rho X_k\Ub) + \rho\omegadot_k,
-\end{equation}
-%
-omitting the reaction terms, which were already
-accounted for in {\bf React State}.  The update consists of two steps:
-
-\begin{enumerate}
-\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-\item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\pred}$,
-  for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
-\MarginPar{mention $\rho'$ and $X$ advection; $(\rho h)'$ advection}
-
-\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2),\star}$ using
-\begin{equation}
-(\rho X_k)^{(2),\star} = (\rho X_k)^{(1)}
-  - \dt \left\{ \nabla \cdot \left[ \uadvone (\rho X_k)^{\nph,\pred} \right] \right\},
-\end{equation}
-
-\end{enumerate}
-
-\item Update $\rho_0$ by setting $\rho_0^{n+1,\star} =$ {\bf Average}$(\rho^{(2),\star})$.
-
-\item Update $p_0$ by calling
-{\bf Enforce HSE}$[\rho_0^{n+1,\star}] \rightarrow [p_0^{n+1,\star}]$.
-
-\item Update the enthalpy using a discretized version of equation
-%
-\begin{equation}
-\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc,
-\end{equation}
-%
-again omitting the reaction and heating terms
-since we already accounted for
-them in {\bf React State}.  This equation takes the form:
-\begin{equation}
-\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h) + \frac{\partial p_0}{\partial t} + (\Ub \cdot \eb_r) \frac{\partial p_0}{\partial r}.
-\end{equation}
-For spherical geometry, we solve the
-analytically equivalent form,
-\begin{equation}
-\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h) + \frac{\partial p_0}{\partial t} + \nabla \cdot (\Ub p_0) - p_0 \nabla \cdot \Ub.
-\end{equation}
-The update consists of two steps:
-
-\begin{enumerate}
-\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-\item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\pred},$
-  for the conservative update of $(\rho h)^{(1)}$
-  using using a Godunov approach \citep{XRB_III}.
-
-\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$.
-\begin{description}
-\item For planar geometry,
-\begin{eqnarray}
-(\rho h)^{(2),\star}
-&=& (\rho h)^{(1)} \nonumber \\
-&&- \dt \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \left(p_0^{n+1,*} - p_0^n\right) \nonumber \\
-&& + \dt \left(\uadvone \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^{n}
-\end{eqnarray}
-
-\item For spherical geometry,
-\begin{eqnarray}
-(\rho h)^{(2),\star}
-&=& (\rho h)^{(1)} \nonumber \\
-&&- \dt \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \left(p_0^{n+1,*} - p_0^n\right) \nonumber \\
-&& + \dt \left \{ \nabla \cdot \left (\uadvone p_0^{n+\myhalf} \right ) - p_0^{n+\myhalf} \nabla \cdot \uadvone \right \}
-\end{eqnarray}
-\end{description}
-where we define $p_0^\nph$ as $p_0^\nph = (p_0^n+p_0^{n+1,*})/2$.
-
-\end{enumerate}
-Note that outside the cutoff density region, we compute enthalpy using the EOS.
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 5
-%--------------------------------------------------------------------------
-\item[Step 5] {\em React the full state through a second time interval of $\dt / 2.$}
-
-Call {\bf React State}$[ (\rho X_k)^{(2),\star}, (\rho h)^{(2),\star}, p_0^{n+1,\star}] 
-\rightarrow 
-[ (\rho X_k)^{n+1,\star}, (\rho h)^{n+1,\star}, (\rho \omegadot_k)^{n+1,\star}, (\rho \Hnuc)^{n+1,\star} ].$
-
-%--------------------------------------------------------------------------
-% STEP 6
-%--------------------------------------------------------------------------
-\item[Step 6] {\em Compute the time-centered expansion, $S^{\nph,\star}$.}
-
-First, compute $S^{n+1,\star}$ with
-\begin{equation}
-S^{n+1,\star} =  \left(-\sigma  \sum_k  \xi_k  \omegadot_k  + \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k + \sigma \Hnuc\right)^{n+1,\star}.
-\end{equation}
-  Then, define
-\begin{equation}
- S^{\nph} = \frac{S^n + S^{n+1,\star}}{2},
-\end{equation}
-
-%--------------------------------------------------------------------------
-% STEP 7
-%--------------------------------------------------------------------------
-\item[Step 7] {\em Construct the time-centered advective velocity on edges, $\uadvtwo$.}
-
-The procedure to construct $\uadvtwodag$ is identical to the Godunov procedure
-for computing $\uadvonedag$ in {\bf Step 3}, but uses
-the updated value $S^{\nph}$ rather than $S^{\nph,\star}$.
-We note that $\uadvtwo$ satisfies the divergence constraint
-\begin{equation}
-\nabla \cdot \left(\beta_0^{\nph} \uadvtwo\right) =
-\beta_0^{\nph} \left[S^{\nph} - \frac{1}{\gammaonebar^{n+\myhalf}p_0^{n+\myhalf}}\left(\frac{\partial p_0}{\partial t}\right)^{n+\myhalf}\right],\label{eq:div2}
-\end{equation}
-where
-\begin{equation}
-\beta_0^{\nph} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2},
-\qquad
-\gammaonebar^{\nph} = \frac{ \gammaonebar^n +  \gammaonebar^{n+1,\star} }{2}.
-\end{equation}
-
-%--------------------------------------------------------------------------
-% STEP 8
-%--------------------------------------------------------------------------
-\item[Step 8] {\em Advect the base state and full state through a time interval of $\dt.$}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \Alph{enumi}}}
-
-\item Update $(\rho X_k)$.  This step is identical to {\bf Step 4A} except we use
-  the updated values $\uadvtwo$ and $\rho_0^{n+1,\star}$ rather than
-  $\uadvone$ and $\rho_0^{n+1,\star}$.  In particular:
-
-\begin{enumerate}
-\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-\item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\pred}$,
-  for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
-
-\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2)}$ using
-\begin{equation}
-(\rho X_k)^{(2)} = (\rho X_k)^{(1)}
-- \dt \left\{ \nabla \cdot \left[\uadvtwo (\rho X_k)^{\nph,\pred} \right] \right\},
-\end{equation}
-
-\end{enumerate}
-
-\item Update $\rho_0$ by setting $\rho_0^{n+1} =$ {\bf Average}$(\rho^{(2)})$.
-
-\item Update $p_0$ by calling
-{\bf Enforce HSE}$[\rho_0^{n+1}] \rightarrow [p_0^{n+1}]$.
-
-\item Update the enthalpy.  This step is identical to {\bf Step 4D} except we use
-  the updated values $\uadvtwo, \rho_0^{n+1}, (\rho h)_0^{n+1}$, and $p_0^{n+\myhalf}$
-  rather than
-  $\uadvone, \rho_0^{n+1,\star}, (\rho h)_0^{n+1,\star}$, and $p_0^n$.
-  In particular:
-
-\begin{enumerate}
-\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}
-
-\item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\pred},$
-  for the conservative update of $(\rho h)^{(1)}$ using a Godunov approach \citep{XRB_III}.
-
-\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2)}$.
-\begin{description}
-\item For planar geometry,
-\begin{eqnarray}
-(\rho h)^{(2)}
-&=& (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvtwo (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \left( p_0^{n+1} - p_0^n \right)\nonumber \\
-&& + \dt \left(\uadvtwo \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^\nph,
-\end{eqnarray}
-
-\item For spherical geometry,
-\begin{eqnarray}
-(\rho h)^{(2)}
-&=& (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvtwo (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \left(p_0^{n+1} - p_0^n\right)\nonumber \\
-&& + \dt \left[ \nabla \cdot \left (\uadvtwo p_0^{\nph} \right ) - p_0^{\nph} \nabla \cdot \uadvtwo \right],
-\end{eqnarray}
-\end{description}
-where we now define $p_0^\nph$ as $p_0^\nph = (p_0^n+p_0^{n+1})/2$.
-
-\end{enumerate}
-Note that outside the cutoff density region, we compute enthalpy using the EOS.
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 9
-%--------------------------------------------------------------------------
-\item[Step 9] {\em React the full state through a second time interval of $\dt / 2.$}
-
-Call {\bf React State}$[(\rho X_k)^{(2)},(\rho h)^{(2)}, p_0^{n+1}] \rightarrow [(\rho X_k)^{n+1}, (\rho h)^{n+1}, (\rho \omegadot_k)^{n+1}, (\rho \Hnuc)^{n+1}].$
-
-%--------------------------------------------------------------------------
-% STEP 10
-%--------------------------------------------------------------------------
-\item[Step 10] {\em Define the new time expansion, $S^{n+1}$.}
-
-\begin{enumerate}
-\renewcommand{\theenumi}{{\bf \Alph{enumi}}}
-\item Define
-\begin{equation}
-  S^{n+1} =  \left(-\sigma  \sum_k  \xi_k \omegadot_k  + \sigma \Hnuc +
-  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  \omegadot_k\right)^{n+1}.
-\end{equation}
-
-\end{enumerate}
-
-%--------------------------------------------------------------------------
-% STEP 11
-%--------------------------------------------------------------------------
-\item[Step 11] {\em Update the velocity}.
-
-First, we compute the time-centered edge velocities, $\Ub^{\nph,\pred}$
-using a Godunov approach \citep{XRB_III}. Then, we define
-\begin{equation}
-\rho^\nph = \frac{\rho^n + \rho^{n+1}}{2}, \qquad \rho_0^\nph = \frac{\rho_0^n + \rho_0^{n+1}}{2}.
-\end{equation}
-We update the velocity field $\Ub^n$ to $\Ub^{n+1,\dagger}$ by discretizing
-equation (\ref{eq:momentum}) as
-\begin{equation}
-\Ub^{n+1,\dagger}
-= \Ub^n - \dt \left[\uadvtwo \cdot \nabla \Ub^{\nph,\pred} \right]
- - \dt \left[ \frac{1}{\rho^\nph} \nabla \pi^\nmh + \frac{\left(\rho^\nph-\rho_0^\nph\right)}{\rho^\nph} g^{\nph} \eb_r \right],
-\end{equation}
-Again, the $\dagger$ superscript refers
-to the fact that the updated velocity does not satisfy the divergence
-constraint.
-
-Finally, we use an approximate nodal projection to define $\Ub^{n+1}$
-from $\Ub^{n+1,\dagger},$  such that $\Ub^{n+1}$ approximately
-satisfies the divergence constraint,
-\begin{equation}
-\nabla \cdot \left(\beta_0^{n+1} \Ub^{n+1} \right) = \beta_0^{n+1} \left[ S^{n+1} - \frac{1}{\gammaonebar^{n+1}p_0^{n+1}}\left(\frac{\partial p_0}{\partial t}\right)^{n+\myhalf}\right].\label{eq:div3}
-\end{equation}
-As part of the projection we also define the new-time perturbational pressure,
-$\pi^\nph.$  This projection necessarily differs from the MAC projection used in
-{\bf Step 3} and {\bf Step 7} because the velocities in those steps are defined
-on edges and $\Ub^{n+1}$ is defined at cell centers, requiring different divergence
-and gradient operators.  Details of the approximate projection are given in Paper III.
-
-\end{description}
-
-This completes one step of the algorithm.
-
-To initialize the simulation
-
-%--------------------------------------------------------------------------
-% STEP 0
-%--------------------------------------------------------------------------
-{\em Initialization}
-\sout{
-The initialization step only occurs at the beginning of the simulation.
-The initialization step computes initial values for 
-$\pi^{-\myhalf}, \eta_{\rho}^{-\myhalf}, S^0$, and $S^1$ for use in time stepping routine.  
-We pass through {\bf Step 1} -- {\bf Step 11}, but only keep these values.
-The initial values for $\Ub^0, \rho^0, (\rho h)^0, X_k^0, T^0,
-\rho_0^0, p_0^0$, and $\gammaonebar^0$ are specified from the problem-dependent
-initial conditions.  The initial time step, $\dt^0$, is computed as in
-Paper III.
-}
-
-\section{Test Problems}
-
-\subsection{Convergence Test}
-3D reacting bubble rise; to demonstrate 2nd-order convergence of the new temporal algorithm.
-
-\subsection{Scaling}
-Scaling with OpenMP
-
-\subsection{Performance Compared to Original MAESTRO}
-Compare to original MAESTRO
-
-\subsection{White Dwarf Convection}
-white dwarf convection runs with 3 algorithms (original, new temporal, new temporal + irregular base state)
-
-\section{Conclusions and Future Work}
-
-science: rotation, solar physics, MHD
-
-algorithm: SDC \cite{dutt2000spectral}, high-order, multi-implicit \cite{bourlioux2003high}.
-Consider SDC ideas used for terrestrial combustion \cite{pazner2016high,nonaka2018conservative}
-
-implementation: GPUs.  Forward cite next CASTRO paper with GPU strategy.
-
-\appendix
-\section{Projection Details}
-To enforce the divergence constraint in {\bf Step 3, 7}, and {\bf 11}, we use a projection method analgous to the methods originally developed for incompressible flow.
-
-Based on our past experience in the MAESTRO project, we have found it useful to split the velocity dynamics into a perturbational and base state velocity,
-\begin{equation}
-\Ub = \Ubt(\xb,t) + w_0(r,t)\eb_r,
-\end{equation}
-solve for each term separately, and immediately combine them to find a full velocity that satisfies the constraint.
-We take that approach here, primarily because it allows us to enforce a boundary condition on $w_0$ at the edge of the star (i.e., the cutoff density location where we hold density constant).
-Namely, to enforce that $r^2 w_0$ remain constant is difficult to do when solving for the full velocity.
-So in practice, we solve the constraint over the lateral average,
-\begin{equation}
-\nabla\cdot\left(\beta_0\Ubt\right) = \beta_0(S-\overline{S})
-\end{equation}
-and sepratately solve for $w_0$ using 
-\begin{equation}
-\nabla\cdot(\beta_0w_0\eb_r) = \beta_0\left(\overline{S} - \frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial t}\right).
-\end{equation}
-To find $w_0$ we integrate in 1D using the procedure in Appendix B of Paper V.
-
-Note that in this approach, we estimated the time-derivative of the pressure in part by examining how laterally averaged $\rho'=\rho-\rho_0$ changed over time (quantified by $\eta_\rho = \overline{\rho'\Ub\cdot\eb_r}$.
-After evolving the species, we compute this term in the same way as reported in Paper V for use in computing $w_0$.
-
-\section{Old Notes (not for inclusion in paper)}
-
-\subsection{Exact Base State Notes}
-
-\subsubsection{Potential Problems}
-\begin{enumerate}
-\item How do we define \verb|r_edge_loc| for irregularly-spaced base states whose cell-centered values are defined at $r_m=\Delta x\sqrt{\frac{3}{4} + 2m}$?
-\item We would stil need to approximate the values at edges when advecting the base states (i.e.
-$\rho_0^{\nph,\pred}$).
-This may affect the nice properties we would get with exact cell-centered base state values.
-\item Excessive memory storage may be required (\verb|nr_irreg >> nr_fine|).
-\end{enumerate}
-
-\subsubsection{A Possible Solution}
-We could choose to no longer advect the base states for both predictor and corrector steps.
-This would solve the problems with computing state values at cell edges.
-Instead, we propose to only update the base states $\rho_0$ and $p_0$ by computing the average of the full state values after advecting the full states.
-This would result in the following changes in the algorithm from Paper V:
-%
-\begin{enumerate}
-\item Steps 4 and 8 will be greatly simplified, leading to a more efficient algorithm.
-\item \sout{There is no longer a need for $w_0$, so we can use the full velocity $\Ub$ instead of $\Ubt$.
-As a consequence, $\eta$ does not need to be computed.
-This has the added advantage of saving some memory storage by removing many of the terms needed when we decided to split terms into its base state and perturbation state in our previous implementation.} We split the velocity for the projection to handle the cutoff.
-\item By placing the base states exactly at cell-centers, we eliminate interpolation errors and simplify the existing code.
-\end{enumerate}
-%
-
-\subsection{$\eta_\rho$ Notes}
-After {\bf Step 8D}, we define a radial cell-centered $\etarho^{\nph}$.
-
-\begin{description}
-\item For planar geometry, $\etarho = \overline{\rho'(\Ub\cdot\eb_r)}$,
-\begin{equation}
- \etarho^{\nph} =  {\rm {\bf Average}} \sum_k \left[ \left(\uadvtwo \cdot \eb_r \right) (\rho X_k)^{\nph,\pred} \right]
-\end{equation}
-\item For spherical geometry, first construct 
-$\etarho^{{\rm cart},\nph} = [\rho'(\Ub\cdot\eb_r)]^{\nph}$ on Cartesian cell centers using:
-\begin{equation}
-\etarho^{{\rm cart},\nph} = \left[\left(\frac{\rho^{(1)}+\rho^{(2)}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{n+1}}{2}\right)\right] \cdot \left( \uadvtwo \cdot \eb_r \right).
-\end{equation}
-Then,
-\begin{equation}
-\etarho^{\nph,\star} = {\rm {\bf Average}}\left(\etarho^{{\rm cart},\nph}\right).
-\end{equation}
-\end{description}
-
-\subsection{Computing $w_0$ For Spherical Problems}
-Recall that we want to solve
-\begin{equation}
-\frac{1}{r^2}\frac{\partial}{\partial r}(r^2 w_0) = \Sbar - \frac{1}{\gammaonebar p_0}\left(\frac{\partial p_0}{\partial t} + w_0\frac{\partial p_0}{\partial r}\right)
-\end{equation}
-First, note that $\partial p_0/\partial r = -\rho_0 g$.
-The key observation here is that we can compute $\partial p_0/\partial t$ directly instead of making complicated substitutions involving $\eta_\rho$ to eliminate it.
-Let's discretize this as follows:
-\begin{equation}
-\frac{1}{r_j^2}\left[\left(r^2 w_0\right)_{j+\myhalf} - \left(r^2 w_0\right)_{j-\myhalf}\right] = 
-\Sbar_j - \frac{1}{\left(\gammaonebar p_0\right)_j}\left[\left(\frac{\partial p_0}{\partial t}\right)_j - \left(\frac{w_{0,j-\myhalf}+w_{0,j+\myhalf}}{2}\right)(\rho_0 g)_j\right]
-\end{equation}
-Rearranging this to solve for $w_{0,j+\myhalf}$ gives us an explicit update:
-\begin{equation}
-\left[\frac{r_{j+\myhalf}^2}{r_j^2} - \frac{(\rho_0 g)_j}{2\left(\gammaonebar p_0\right)_j}\right] w_{0,j+\myhalf} =
-\left[\frac{r_{j-\myhalf}^2}{r_j^2} + \frac{(\rho_0 g)_j}{2\left(\gammaonebar p_0\right)_j}\right] w_{0,j-\myhalf}
-+ \Sbar_j - \frac{1}{\left(\gammaonebar p_0\right)_j}\left(\frac{\partial p_0}{\partial t}\right)_j
-\end{equation}
-As before, once $\rho_0$ falls below $\rho_{\rm cutoff}$, we hold $r^2 w_0$ constant.\\ \\
-The interface to the routine in the algorithm description is
-{\bf Make $\mathbf{w_0}$}$[\rho_0,p_0,\partial p_0/\partial t,\gammaonebar,\Sbar] \rightarrow [w_0]$
-
-\acknowledgements
-
-The work at LBNL was supported by the U.S. Department of Energy's Scientific Discovery Through Advanced Computing (SciDAC) program under contract No. DE-AC02-05CH11231.
-This research used resources of the National Energy Research Scientific Computing Center (NERSC), a U.S.
-Department of Energy Office of Science User Facility operated under Contract No.
-DE-AC02-05CH11231.
+\input{appendix.tex}
 

### 2019-03-13 MAESTROeX_I/results.tex
split paper into multiple tex files for multi-user edits

\section{Performance and Validation}

\subsection{AMR Performance}
Show 3D reacting bubble rise performance.
Show 3-level spherical performance (stats won't be as good but explain why: refining a larger percentage of grid)

\subsection{Scaling}
Scaling with OpenMP

\subsection{White Dwarf Convection}
white dwarf convection runs with 3 algorithms (original, new temporal, new temporal + irregular base state)

Show 3-level wdconvect

### 2019-03-13 MAESTROeX_I/equations.tex
more updates

@@
 \section{Governing Equations}
-The derivation of the low Mach number hydrodynamics equations is given in Papers I-III.
-We take the standard equations of reacting, compressible flow, and recast the equation
-of state (EOS) as a divergence constraint on the velocity field.  This is done by taking
-the Langrangian derivative of the equation of state for pressure as a function of the
-density, composition, and enthalpy, enforcing the constraint that the pressure is
-a prescribed function of altitutde and time based on the hydrostatic equilibrium condition,
-and substituting the equations for species and enthalpy in the resulting Lagrangian derivative.
-The resulting equations are
+Low Mach number models for reacting flow were originally derived using asymptotic analysis
+\citep{rehm1978equations,majda1985derivation} and used in terrestrial combustion applications
+\citep{knio1999semi,day2000numerical}.  These models have been extended to nuclear flames
+in astrophysical environments using adaptive algorithms in space and time \citep{Bell:2004}.
+In Papers I-III, we derived a model and algorithm suitable for stratified astrophysical flow.
+In our model, we take the standard equations of reacting, compressible flow, and recast the equation
+of state (EOS) as a divergence constraint on the velocity field.
+The resulting model is a series of evolution equations for mass, momentum, and energy, subject
+to this constraint:
 \begin{eqnarray}
@@
 
-Mathematically, equations (\ref{eq:momentum})-(\ref{eq:enthalpy}) must still be closed by the equation of state which we
-express as a divergence constraint on the velocity field (\ref{eq:U divergence}),
-\MarginPar{This is different from the Jacobs 2016 paper where we switched to alt\_energy\_fix.}
+Mathematically, equations (\ref{eq:momentum})-(\ref{eq:enthalpy}) must still be closed by an equation of state.  
+This is done by taking the Langrangian derivative of the equation of state for pressure as a function of the density, composition, and enthalpy, requiring that the pressure is a prescribed function of altitutde and time based on the hydrostatic equilibrium condition.
+Details are given in Papers I and II.
+After substuting the equations for mass and energy into the Lagrangian derivative, the result
+is a divergence constraint on the velocity field (\ref{eq:U divergence}),
 where $\beta_0$ is a density-like variable that carries background stratification, defined as

### 2019-03-13 MAESTROeX_I/intro.tex
more updates

@@
 The resulting model can be numerically integrated with much larger time steps than a compressible model.
+Low Mach number models have been developed for a variety of contexts including combustion \citep{day2000numerical}, terrestrial atmospheric modeling \citep{duarte2015low}, 
+and elastic solids \citep{abbate2017all}.
 
-Previously, we have developed the low Mach number astrophysical solver, MAESTRO.
-Low Mach number models have been developed for a variety of contexts including combustion \citep{day2000numerical}, atmospherics \citep{duarte2015low}, and elastic solids \citep{abbate2017all}.
+Previously, we developed the low Mach number astrophysical solver, MAESTRO.
 The low Mach number model in MAESTRO is unique in that it is specifically designed for astrophysical settings with significant atmospheric stratification.
 Central to the algorithm is a stratified background (or base) state density and pressure held in hydrostatic equilibrium, and also vary as a function of altitude and time.
-MAESTRO is suitable for for spherical stars, as well as planar simulations of local dynamics.
+MAESTRO is a structured-grid, finite-volume code that utlizes adaptive mesh refinement (AMR) to refine grids locally in space.
+MAESTRO is suitable for for full spherical stars, as well as planar simulations of dynamics within localized regions of a star.
 
@@
 \item Sub-Chandrasekhar white dwarfs \citep{subChandra_I,subChandra_II}.
-In the latter paper we introduced an optional modification to the momentum equation of velocity constraint that conserves total energy, can give results closer to compressible codes of convective dynamics, including low density regions at the surface of a star.
+%  In the latter paper we introduced an optional modification to the momentum equation of velocity constraint that conserves total energy, can give results closer to compressible codes of convective dynamics, including low density regions at the surface of a star.
 \item Type I X-ray bursts \citep{XRB_I,XRB_II,XRB_III}.
-In \cite{XRB_I} we discuss the modifications required for implicit thermal conduction, as well as an additional ``volume discrepancy'' modification to the velocity constraint to force the species and enthalpy to evolve in a manner consistent with the background pressure.
+%  In \cite{XRB_I} we discuss the modifications required for implicit thermal conduction, as well as an additional ``volume discrepancy'' modification to the velocity constraint to force the species and enthalpy to evolve in a manner consistent with the background pressure.
 \end{itemize}
@@
 We also develop simpler, alternative temporal integrators, with a longer term goal of developing high-order coupling schemes.
-We also present a new treatment of the base state mapping to Cartesian grids for spherical problsm.
+We also present a new treatment of the base state mapping to Cartesian grids for spherical problems.
+
+The resulting code is publicly available on GitHub at {\url https://github.com/AMReX-Astro/MAESTROeX} and uses the microphysics 
+libraries at {\url https://github.com/starkiller-astro/Microphysics} and the AMReX software library for block-structured adaptive mesh refinement
+{\url https://github.com/AMReX-Codes/amrex}.

### 2019-03-13 MAESTROeX_I/intro.tex
more intro

@@
 
-In this paper, we characterize the performance of the MAESTROeX algorithm compared to the previous implementation.
-We also develop simpler, alternative temporal integrators, with a longer term goal of developing high-order coupling schemes.
-We also present a new treatment of the base state mapping to Cartesian grids for spherical problems.
+We also present new algorithmic methodology that improves upon Paper V in a number of ways.
+First, the overall temporal algorithm has been greatly simplified without compromising accuracy.
+The key design decision was to replace the predicted evolution of the base state with 
+a predictor-corrector approach.  Not only does this greatly simplify the dynamics of the base
+state, but this treatment is more amenable to higher-order multiphysics coupling strategies
+based on method-of-lines integration.  Second, we now provide additional spatial mapping
+procedures that reduces, and in some cases eliminates, mapping error from the one-dimensional
+base state to the Cartesian grid state for spherical problems.
+Finally, the code has been completely ported to a new software framework suitable for modern manycore supercomputers.
+The resulting MAESTROeX is implemented in the C++/F90 AMReX software library.  MAESTROeX uses MPI+OpenMP parallelism
+and scales to XXX MPI processes, each capable of spawning tens of threads.
+\MarginPar{quantify XXX.  I think it's 5,000-10,000 MPIs with potentially hundreds of threads}
 

### 2019-03-14 MAESTROeX_I/algorithm.tex
more updates
updates, re-org

@@
 
-\subsection{Spatial Discretization}
+\subsection{Spatial Discretization}\label{Sec:Spatial}
+MAESTRO contains variables defined on Cartesian grids 
+one-dimensional base state stored on radial arrays.
 
-(description of base state mapping in planar, and origianl spherical)
+ associated with square (in 2D) or cubic (in 3D)
 
-\subsection{Exact Base State Mapping} \label{sec:exactbasestate}
+(description of base state mapping in planar, and origianl spherical, and new spherical)
+
+Mapping from the base state to the full state
+
+Mapping a Cartesian variable to a one-dimensional radial array uses the lateral averating operator.
+
+One of the key numerical modules is the ``lateral average'', which computes the average over a 
+layer of a Cartesian grid variable onto a one-dimensional radial array.
+In planar geometries, this is a straightforward arithmetic average of cells at
+a particular height since the radial cell centers are in alignment
+with the Cartesian grid cell centers.
+However for spherical problems, the procedure is much more complicated.
+In Section 4 of Paper V, we describe how there is a finite, easily computable set of radii that any Cartesian cell-center can map to.  
+Specifically, for every Cartesian cell, there exists an integer $m$ such that the radius to the center of the star is given by $\hat{r}_m=\Delta x\sqrt{0.75+2m}$.
+We average by binning all the cells associated with each radii and taking the average.
+Then we interpolate this data onto a one-dimensional radial array.  Previously, MAESTRO only allowed for radial arrays with a constant $\Delta r$ (typically equal to $\Delta x/5$.
+Here we present a new option to retain an irregularly-spaced radial array to eliminate mapping errors back onto the Cartesian grid state.
+Consider a spherical star in hydrostatic equlibrium at rest.  In the absense of reactions, the star should remain at rest.
+The buoyancy forcing term in the momentum equation contains $\rho-\rho_0$.
+With the original scheme, interpolation errors in computing $\rho_0$ by averaging would cause artifical acceleration in the velocity field due to the interpolation error from the Cartesian grid to and from the radial base state.
+By retaining the radial base state as an irregularly spaced array, the effects due to interpolation error are completely eliminated.
 
@@
 Temperature is computed using the equation of state, e.g.,
-\footnote{As described in Paper V, for planar problems we compute temperature using $h$ instead of $p_0$, since we have successfully developed planar volume discrepancy schemes to effectively couple the enthalpy to the rest of the solution.
+\footnote{As described in Paper V, for planar problems we compute temperature using $h$ instead of $p_0$, since we have successfully developed planar volume discrepancy schemes to effectively couple the enthalpy to the rest of the solution; see \cite{XRB_I}.
 We are still exploring this option for spherical stars.}
@@
 \begin{itemize}
-\item {\bf Average} computes...
-\item {\bf Enforce HSE} computes the base state pressure, $p_0$, from a base state density, $\rho_0$ by integrating the hydrostatic equilibrium condition in 1D (see equation (A10) in Paper V).
+\item {\bf Average}$[\phi]\rightarrow[\overline\phi]$ computes the lateral average of a quantity over a layer at constant radius $r$, as described above in Section \ref{Sec:Spatial}.
+\item {\bf Enforce HSE}$[\rho_0]\rightarrow[p_0]$ computes the base state pressure, $p_0$, from a base state density, $\rho_0$ by integrating the hydrostatic equilibrium condition in 1D (see equation (A10) in Paper V).
 The base state pressure remains equal to a constant value from the location of a prescribed cutoff density outward for the entire simulation.
-\item {\bf React State} advances the species and enthalpy due to reactions over $\Delta t/2$ by integrating
+\item {\bf React State}$[(\rho X_k)^{\rm in},(\rho h)^{\rm in},p_0]\rightarrow[(\rho X_k)^{\rm out},(\rho h)^{\rm out},(\rho\dot\omega),(\rho\Hnuc]$ integrates the species and enthalpy due to reactions over $\Delta t/2$ by solving
 \begin{equation}
 \frac{dX_k}{dt} = \dot\omega_k(\rho,X_k,T); \qquad
-\frac{dT}{dt} = \frac{1}{c_p}\left(-\sum_k\xi_k\dot\omega_k + H_{\rm nuc}\right)
+\frac{dT}{dt} = \frac{1}{c_p}\left(-\sum_k\xi_k\dot\omega_k + H_{\rm nuc})\right)
 \end{equation}
@@
 There are additional constraints on the timestep that are typically much less restrictive than the advective CFL including
-the acceleration due to the buoyancy force (commonly used when the velocity is approximately zero at the start of some simulations) 
+the acceleration due to the buoyancy force (sometimes in effect when the velocity is approximately zero at the start of some simulations) 
 and the local magnitude of the divergence constraint (to prevent too much mass evacuation from a cell in a time step); see Section 3.4 in Paper III for details.
@@
 \MarginPar{mention cutoff densities}
-
-\MarginPar{AJN: eliminated references to volume discrepancy for now; we'd only use it for planar at the moment, don't want to open up that can of worms here}
 
@@
 
-\item Update $\rho_0$ by setting $\rho_0^{n+1,\star} =$ {\bf Average}$(\rho^{(2),\star})$.
+\item Update $\rho_0$ by calling {\bf Average}$[\rho^{(2),\star}]\rightarrow[\rho_0^{n+1,\star}]$.
 
-\item Update $p_0$ by calling
-{\bf Enforce HSE}$[\rho_0^{n+1,\star}] \rightarrow [p_0^{n+1,\star}]$.
+\item Update $p_0$ by calling {\bf Enforce HSE}$[\rho_0^{n+1,\star}] \rightarrow [p_0^{n+1,\star}]$.
 
@@
 
-\item Update $\rho_0$ by setting $\rho_0^{n+1} =$ {\bf Average}$(\rho^{(2)})$.
+\item Update $\rho_0$ by calling {\bf Average}$[\rho^{(2)}]\rightarrow[\rho_0^{n+1}]$.
 
-\item Update $p_0$ by calling
-{\bf Enforce HSE}$[\rho_0^{n+1}] \rightarrow [p_0^{n+1}]$.
+\item Update $p_0$ by calling {\bf Enforce HSE}$[\rho_0^{n+1}] \rightarrow [p_0^{n+1}]$.
 
@@
 \subsection{Adaptive Mesh Refinement}
+Talk about AMR algorithm and AMReX computational framework.

### 2019-03-14 MAESTROeX_I/appendix.tex
updates, re-org

@@
 {\bf Make $\mathbf{w_0}$}$[\rho_0,p_0,\partial p_0/\partial t,\gammaonebar,\Sbar] \rightarrow [w_0]$
+
+\subsection{TODO List}
+
+TODO list:
+\begin{itemize}
+\item Weak scaling tests of original algorithm, comparing MAESTRO to MAESTROeX.  Use cori knl, 4 MPI per node, 16 threads per MPI
+\item Run at effective $512^3$ resolution (maybe to ignition?...
+need to see how max T trends go) (1) classic MAESTRO, (2) MAESTROeX with original algorithm, (3) MAESTROeX with new temporal integrator, (4) MAESTROeX with new temporal integrator and irregular dr, (5) MAESTROeX with original algorithm and 3-levels of AMR, (6) MAESTROeX with new temporal integrator and 3 levels of AMR.
+\item (perhaps) a demonstration of what goes wrong when you don't split the velocity dynamics in the projection
+\end{itemize}

### 2019-03-14 MAESTROeX_I/intro.tex
updates, re-org

@@
 \section{Introduction} \label{sec:intro}
-
-TODO list:
-\begin{itemize}
-\item Weak scaling tests of original algorithm, comparing MAESTRO to MAESTROeX.  Use cori knl, 4 MPI per node, 16 threads per MPI
-\item Run at effective $512^3$ resolution (maybe to ignition?...
-need to see how max T trends go) (1) classic MAESTRO, (2) MAESTROeX with original algorithm, (3) MAESTROeX with new temporal integrator, (4) MAESTROeX with new temporal integrator and irregular dr, (5) MAESTROeX with original algorithm and 3-levels of AMR, (6) MAESTROeX with new temporal integrator and 3 levels of AMR.
-\item (perhaps) a demonstration of what goes wrong when you don't split the velocity dynamics in the projection
-\end{itemize}
-
 Many astrophysical flows are highly subsonic; sound waves carry sufficiently little energy that they do not significantly affect the convective dynamics of the system.

### 2019-03-14 MAESTROeX_I/algorithm.tex
initialization

@@
 \item In {\bf Steps 2-4} we advect the thermodynamic variables over $\Delta t$.
-Specifically, we compute an estimate for the expansion term, $S$, project the velocities so they satisfy the divergence constraint, and then advect the thermodynamic variables.
-\item In {\bf Step 5} we react the thermodynamic variables over $\Delta t/2$.\footnote{Note that we could skip to the velocity advance in {\bf Steps 10-11}, however the overall scheme would be only first-order in time, so {\bf Steps 6-9} can be thought of as a trapezoidal corrector step.}
+Specifically, we compute an estimate for the expansion term, $S$, project the face-centered, time-centered velocities so they satisfy the divergence constraint, and then advect the thermodynamic variables.
+\item In {\bf Step 5} we react the thermodynamic variables over $\Delta t/2$.\footnote{After this step we could skip to the velocity advance in {\bf Steps 10-11}, however the overall scheme would be only first-order in time, so {\bf Steps 6-9} can be thought of as a trapezoidal corrector step.}
 \item In {\bf Steps 6-8} we redo the advection in {\bf Steps 2-4} but are able to use the trapezoidal rule to time-center certain quantities such as $S$, $\rho_0$, etc.
 \item In {\bf Step 9} we redo the reactions from {\bf Step 5} using the improved results for the corrector advection step.
-\item In {\bf Steps 10-11} we update the velocity.
+\item In {\bf Steps 10-11} we and update the velocity, compute the expansion term, $S$, and project the cell-centered velocities.
 \end{itemize}
@@
 The base state pressure remains equal to a constant value from the location of a prescribed cutoff density outward for the entire simulation.
-\item {\bf React State}$[(\rho X_k)^{\rm in},(\rho h)^{\rm in},p_0]\rightarrow[(\rho X_k)^{\rm out},(\rho h)^{\rm out},(\rho\dot\omega),(\rho\Hnuc]$ integrates the species and enthalpy due to reactions over $\Delta t/2$ by solving
+\item {\bf React State}$[(\rho X_k)^{\rm in},(\rho h)^{\rm in},p_0]\rightarrow[(\rho X_k)^{\rm out},(\rho h)^{\rm out},(\rho\dot\omega),(\rho\Hnuc)]$ integrates the species and enthalpy due to reactions over $\Delta t/2$ by solving
 \begin{equation}
@@
 
-This completes one step of the algorithm.
-
-To initialize the simulation
-
-%--------------------------------------------------------------------------
-% STEP 0
-%--------------------------------------------------------------------------
+This completes one step of the algorithm.\\ \\
 {\em Initialization}
-\sout{
-The initialization step only occurs at the beginning of the simulation.
-The initialization step computes initial values for 
-$\pi^{-\myhalf}, \eta_{\rho}^{-\myhalf}, S^0$, and $S^1$ for use in time stepping routine.  
-We pass through {\bf Step 1} -- {\bf Step 11}, but only keep these values.
-The initial values for $\Ub^0, \rho^0, (\rho h)^0, X_k^0, T^0,
-\rho_0^0, p_0^0$, and $\gammaonebar^0$ are specified from the problem-dependent
-initial conditions.  The initial time step, $\dt^0$, is computed as in
-Paper III.
-}
+To initialize the simulation we use the same procedure described in Paper III.
+We set initial values for $\Ub, \rho X_k$, and $\rho h$ and perform a sequence of projections 
+(to ensure the velocity field satisfies the divergence constraint) 
+followed by a small number of steps of the temporal advancement scheme to iteratively 
+find initial values for $\pi^{n-\myhalf}$ and $S^0$ and $S^1$ for use in the first time step.
 

### 2019-03-15 MAESTROeX_I/algorithm.tex
more refs

@@
 The base state pressure remains equal to a constant value from the location of a prescribed cutoff density outward for the entire simulation.
+\MarginPar{add note about changes for irregular}
 \item {\bf React State}$[(\rho X_k)^{\rm in},(\rho h)^{\rm in},p_0]\rightarrow[(\rho X_k)^{\rm out},(\rho h)^{\rm out},(\rho\dot\omega),(\rho\Hnuc)]$ integrates the species and enthalpy due to reactions over $\Delta t/2$ by solving
@@
 \end{equation}
-In practice we solve for.
+In practice, for all the projections in the algorithm, we solve a split dynamics
+system for a base state velocity, and a perturbational velocity, so that we can properly
+enforce boundary conditions at the edge of the star.
+Refer to Appendix \ref{Sec:Projection} for details.
 
@@
   for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
-\MarginPar{mention $\rho'$ and $X$ advection; $(\rho h)'$ advection}
+  As described in Paper V, we predict $\rho'^n=\rho^n-\rho_0^n$ and $X_k^n$ to edges
+  and here we spatially interpolate $\rho_0^n$ to edges to assemble the fluxes.
 
@@
 \end{equation}
-For spherical geometry, we solve the
-analytically equivalent form,
+For spherical geometry, we solve the analytically equivalent form,
 \begin{equation}
@@
 \item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\pred},$
-  for the conservative update of $(\rho h)^{(1)}$
-  using using a Godunov approach \citep{XRB_III}.
+  for the conservative update of $(\rho h)^{(1)}$ using using a Godunov approach \citep{XRB_III}.
+  As described in Paper V, we predict $(\rho h)'^n=(\rho h)^n-(\rho h)_0^n$ to edges,
+  where $(\rho h)_0^n$ is obtained by calling {\bf Average}$[(\rho h)^n]\rightarrow[(\rho h)_0]$,
+  and here we spatially interpolate $(\rho h)_0^n$ to edges to assemble the fluxes.
 
@@
   for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
+  Again, we predict $\rho'^n=\rho^n-\rho_0^n$ and $X_k^n$ to edges
+  and here we spatially interpolate $(\rho_0^n+\rho_0^{n+1,*})/2$ to edges to assemble the fluxes.
 
@@
   for the conservative update of $(\rho h)^{(1)}$ using a Godunov approach \citep{XRB_III}.
+  Again, we predict $(\rho h)'^n=(\rho h)^n-(\rho h)_0^n$ to edges
+  and here we spatially interpolate $[(\rho h)_0^n)+(\rho h)_0^{n+1,*}]/2$ to edges to assemble the fluxes.
 

### 2019-03-15 MAESTROeX_I/appendix.tex
more refs

@@
 \appendix
-\section{Projection Details}
-To enforce the divergence constraint in {\bf Step 3, 7}, and {\bf 11}, we use a projection method analgous to the methods originally developed for incompressible flow.
+\section{Projection Details}\label{Sec:Projection}
+To enforce the divergence constraint in {\bf Step 3, 7}, and {\bf 11}, we use a projection method analgous to the methods originally developed for incompressible flow \citep{almgren1998conservative,bell1989second}.
 
@@
 
-\MarginPar{mention alt energy fix for projection}
+\MarginPar{mention alt energy fix for projection?  also mention changes in $w_0$ computation for irregular base state}
 

### 2019-03-18 MAESTROeX_I/algorithm.tex
updates

@@
 \end{itemize}
-
-Each time step satisfies the standard advective CFL condition,
+Each time step is constrained by the standard advective CFL condition,
 \begin{equation}
@@
 \end{equation}
-where for our simulations we typically use $\sigma^{\rm CFL}\sim 0.5$.
+where for our simulations we typically use $\sigma^{\rm CFL}\sim 0.7$ and the minimum is taken over all spatial directions over all cells.
 There are additional constraints on the timestep that are typically much less restrictive than the advective CFL including
@@
 
-At the beginning of each simulation, we define $(\Ub,\rho X_k,\rho h)$...
-
+To initialize the simulation we use the same procedure described in Paper III.
+At the beginning of each simulation, we define $(\Ub,\rho X_k,\rho h)$.
+We set initial values for $\Ub, \rho X_k$, and $\rho h$ and perform a sequence of projections 
+(to ensure the velocity field satisfies the divergence constraint) 
+followed by a small number of steps of the temporal advancement scheme to iteratively 
+find initial values for $\pi^{n-\myhalf}$ and $S^0$ and $S^1$ for use in the first time step.
+The new temporal integration strategy in MAESTROeX contains the following steps:
 \MarginPar{mention cutoff densities}
@@
 \end{description}
+This completes one step of the algorithm.
 
-This completes one step of the algorithm.\\ \\
-{\em Initialization}
-To initialize the simulation we use the same procedure described in Paper III.
-We set initial values for $\Ub, \rho X_k$, and $\rho h$ and perform a sequence of projections 
-(to ensure the velocity field satisfies the divergence constraint) 
-followed by a small number of steps of the temporal advancement scheme to iteratively 
-find initial values for $\pi^{n-\myhalf}$ and $S^0$ and $S^1$ for use in the first time step.
-
-\subsection{Adaptive Mesh Refinement}
-Talk about AMR algorithm and AMReX computational framework.
+Our approach to adaptive mesh refinement is algorithmically the same as the treatment described in Section 5 of Paper V; we refer the reader there for details.
+We note that for spherical problems, AMR is only available for the case of a regularly-spaced base state.

### 2019-03-18 MAESTROeX_I/intro.tex
updates

@@
 The resulting MAESTROeX is implemented in the C++/F90 AMReX software library.  MAESTROeX uses MPI+OpenMP parallelism
-and scales to XXX MPI processes, each capable of spawning tens of threads.
-\MarginPar{quantify XXX.  I think it's 5,000-10,000 MPIs with potentially hundreds of threads}
+and scales well to over 10,000 MPI processes, with each MPI process supporting tens of threads
 

### 2019-03-18 MAESTROeX_I/conclusions.tex
more refs
more ferences

@@
 \section{Conclusions and Future Work}
+Our new implementation of the MAESTRO code in the AMReX software framework
 
-science: rotation, solar physics, MHD
+Our future scientific investigations include convection in massive rotating stars, \citep{heger2000presupernova}, the convective Urca process in white dwarfs \citep{willcox2016type}, solar physics \citep{wood2018self}, and magnetohydrodynamics \citep{wood2015three,wood2011sun}.
 
-algorithm: SDC \cite{dutt2000spectral}, high-order, multi-implicit \cite{bourlioux2003high}.
-Consider SDC ideas used for terrestrial combustion \cite{pazner2016high,nonaka2018conservative}
+Our future algorithmic enhancements include more accurate and higher-order multiphysics coupling strategies byased on spectral deferred corrections \cite{dutt2000spectral,bourlioux2003high}.
+This framework has been successfully used in terrestrial combustion \cite{pazner2016high,nonaka2018conservative}
 
-implementation: GPUs.  Forward cite next CASTRO paper with GPU strategy.
+Our software enhancements will include GPU implementation.
+\MarginPar{Forward cite CASTRO paper; add a few sentences on GPU strategy}
 

### 2019-03-18 MAESTROeX_I/intro.tex
more ferences

@@
 Finally, the code has been completely ported to a new software framework suitable for modern manycore supercomputers.
-The resulting MAESTROeX is implemented in the C++/F90 AMReX software library.  MAESTROeX uses MPI+OpenMP parallelism
-and scales well to over 10,000 MPI processes, with each MPI process supporting tens of threads
-
+The resulting MAESTROeX is implemented in the C++/F90 AMReX public software library \cite{AMReX}.  
+MAESTROeX uses MPI+OpenMP parallelism and scales well to over 10,000 MPI processes, with each MPI process supporting tens of threads
 The resulting code is publicly available on GitHub at {\url https://github.com/AMReX-Astro/MAESTROeX} and uses the microphysics 

### 2019-03-19 MAESTROeX_I/results.tex
note about scaling

@@
 \subsection{Scaling}
-Scaling with OpenMP
+In Figure XXX we show weak scaling results for a spherical, full-star sub-chandrasekhar mass white dwarf calculation.
+The x-axis represents total core count (in this case, the total number of OpenMP threads), and the y-axis is the wallclock time per time step.
+These simulations were performed using the NERSC cori system on the Intel Xeon Phi (KNL) partition.
+We split each KNL node to contain 4 MPI processes with 16 threads each.
+The computational domain for each simulation is divided into $64^3$ grid cells, and we assign 1 MPI process to each grid.
+We used simulations ranging from $256^3$ grid cells (64 MPI processes, corresponding to 1024 total threads) up to $1536^3$ grid cells (13,824 MPI processes, corresponding to 221,184 total threads).
+Note that the largest simulation used roughly 36\% of the entire computational system.
+We see that the total increas in wallclock time from the smallest to largest simulation is roughly 41\%, which is quite remarkable given that there are 3 linear solves per time step.
+\MarginPar{Making new plots splitting out projections vs. everything else.}
 

### 2019-03-22 gpus/paper.tex
update with BCs
lots of paper refs

@@
 There are a number of descriptions of GPU-enabled astrophysical hydro codes, including
-\citet{gamer,cholla,fargo3d,pekkila:2017,caplan:2018,racz:2018,liska:2018,goz:2018}
+\citet{gamer,cholla,fargo3d,pekkila:2017,caplan:2018,racz:2018,liska:2018,goz:2018,wang:2010} \MarginPar{todo: group by AMR/no-AMR, GPU exclusive, Kokkos, ...}
 
@@
 
-Disucss changes since last paper
+\castro\ development began with \cite{castro}, with the goal of
+creating an astrophysical hydrodynamics code using a modern adaptive
+mesh refinement library (BoxLib at the time) and unsplit hydrodynamics
+techniques.  The initial application area was nuclear astrophysics,
+including core-collapse supernovae, so it was designed to work with a
+general equation of state~\cite{zingalekatz} and reaction network, and
+with radiation transport (through flux limited diffusion,
+\citealt{castro2,castro3}).  \castro\ also acted as the compressible
+counterpart to our low Mach number hydrodynamics code \maestro,
+sharing much of the same infrastructure, and allowing for simulations
+to be restarted from the low Mach regime into the compressible
+regime~\cite{malone:2014}.  \castro\ has been applied to various
+progenitor models of Type Ia
+supernova~\cite{malone:2014,katz:2016,polin:2019}, exoplanet
+dynamics~\cite{ryu:2018}, core-collapse and population III
+supernovae~\cite{chen:2014,dolence:2015,chen:2017}, and X-ray bursts~\cite{astronum:2018}..
 
@@
 
+Boundary conditions require special attention---threadblock must include
+all zones that participate in the boundary.
+
 \section{Reaction networks}

### 2019-03-22 gpus/paper.tex
more intro work

@@
 
+Supercomputer architectures have evolved a lot over the past
+decades~\citep{bell:2015}.  Shared memory vector machines with custom
+processors dominated early, displaced in the early 1990s by
+distributed memory, massively parallel architectures built upon
+commodity CPUs where the message passing interface (MPI) became the
+dominant library for achieving parallelism.  As core counts increased,
+threading because increasingly important, and OpenMP became the
+standard for threading in high-performance computing, with hybrid
+MPI+OpenMP approaches to parallelism.  In the race to increase
+performance while keeping electrical power consumption reasonable,
+accelerators were increasingly turned to, include GPUs and many-core
+chips like the Intel Phi series.  The simulation codes we use in
+astrophysics need to keep pace with these developments, or they will
+not be able to run on the next generation of supercomputers.  In
+particular, GPUs are dominant now, so existing algorithms need to be
+ported to run on GPUs, or new codes need to be written from scratch
+targeting the new architecture.
+
+There are a number of descriptions of GPU-enabled astrophysical hydro
+codes, including
+\citet{gamer,cholla,fargo3d,pekkila:2017,caplan:2018,racz:2018,liska:2018,goz:2018,wang:2010} \MarginPar{todo:
+  group by AMR/no-AMR, GPU exclusive, Kokkos, ...}
+
+GPU-enabled integration has been explored by \citep{brock:2015}.
+
+In this paper, we describe our approach to performance portability
+with the \castro\ multiphysics simulation code..  Our goal is to have
+the same computational kernels run efficiently on CPUs (including
+manycore processors) and GPUs.  We describe the work we needed to do
+to enable this portability and show performance metrics for a reacting
+flow problem in astrophysics.
+
+\subsection{The Castro hydrodynamics code}
+
 \castro\ is a compressible (radiation-)hydrodynamics simulation code built on
@@
 The original computing architectures targeted were
-
-With GPU accelerators becoming increasingly dominant, algorithms need to be ported.
-There are a number of descriptions of GPU-enabled astrophysical hydro codes, including
-\citet{gamer,cholla,fargo3d,pekkila:2017,caplan:2018,racz:2018,liska:2018,goz:2018,wang:2010} \MarginPar{todo: group by AMR/no-AMR, GPU exclusive, Kokkos, ...}
-
-GPU-enabled integration has been explored by \citep{brock:2015}.
-
-In this paper, we describe our approach to performance portability.
-Our goal is to have the same computational kernels run efficiently on
-CPUs (including manycore processors) and GPUs.  We describe the work
-we needed to do to enable this portability and show performance
-metrics for a reacting flow problem in astrophysics.
-
-\section{The Castro hydrodynamics code}
 

### 2019-03-22 gpus/paper.tex
added a few more references to astro papers that talk about GPUs

--- 2019-02-15 gpus/paper.tex
@@
 There are a number of descriptions of GPU-enabled astrophysical hydro codes, including
-\citet{gamer,cholla,fargo3d,pekkila:2017,caplan:2018,racz:2018,liska:2018,goz:2018}
+\citet{gamer,cholla,fargo3d,pekkila:2017,caplan:2018,racz:2018,liska:2018,goz:2018,Padioleau2019,Bryan2014,Kulikov2014}
 

### 2019-03-22 gpus/paper.tex

--- 2019-03-22 gpus/paper.tex
@@
 codes, including
-\citet{gamer,cholla,fargo3d,pekkila:2017,caplan:2018,racz:2018,liska:2018,goz:2018,wang:2010} \MarginPar{todo:
+\citet{gamer,cholla,fargo3d,pekkila:2017,caplan:2018,racz:2018,liska:2018,goz:2018,wang:2010,Padioleau2019,Bryan2014,Kulikov2014} \MarginPar{todo:
   group by AMR/no-AMR, GPU exclusive, Kokkos, ...}
@@
 dynamics~\cite{ryu:2018}, core-collapse and population III
-supernovae~\cite{chen:2014,dolence:2015,chen:2017}, and X-ray bursts~\cite{astronum:2018}..
+supernovae~\cite{chen:2014,dolence:2015,chen:2017}, and X-ray bursts~\cite{astronum:2018}.
 

### 2019-03-22 gpus/paper.tex
more writing

--- 2019-03-22 gpus/paper.tex
@@
 \newcommand{\maestro}{{\sf Maestro}}
+\newcommand{\maestroex}{{\sf MAESTROeX}}
 
@@
 
+\newcommand{\fab}{{\sf Fab}}
+\newcommand{\multifab}{{\sf MultiFab}}
+
 \begin{document}
@@
 %======================================================================
-\title{CASTRO: A New Compressible Astrophysical Solver. IV. Performance Portability}
+\title{CASTRO: A New Compressible Astrophysical Solver. IV. Performance Portability and GPUs}
 
-\shorttitle{}
+\shorttitle{CASTRO Performance Portability and GPUs}
 \shortauthors{}
@@
 
-GPU-enabled integration has been explored by \citep{brock:2015}.
+GPU-enabled integration has been explored by \citep{brock:2015}.\MarginPar{XNet paper?}
 
 In this paper, we describe our approach to performance portability
-with the \castro\ multiphysics simulation code..  Our goal is to have
+with the \castro\ multiphysics simulation code.  Our goal is to have
 the same computational kernels run efficiently on CPUs (including
@@
 
-\castro\ is a compressible (radiation-)hydrodynamics simulation code built on
-the AMReX adaptive mesh refinement framework \citep{castro,castro2,castro3}.  It 
-includes full self-gravity using multigrid methods with isolated boundaries
-\citep{katz:2016}, works with general equations of state \citep{zingalekatz}
-and arbitrary nuclear reaction networks.  Castro has been used for science investigations of
-...
-
-The original computing architectures targeted were
-
 \castro\ development began with \cite{castro}, with the goal of
 creating an astrophysical hydrodynamics code using a modern adaptive
-mesh refinement library (BoxLib at the time) and unsplit hydrodynamics
-techniques.  The initial application area was nuclear astrophysics,
-including core-collapse supernovae, so it was designed to work with a
-general equation of state~\cite{zingalekatz} and reaction network, and
-with radiation transport (through flux limited diffusion,
-\citealt{castro2,castro3}).  \castro\ also acted as the compressible
-counterpart to our low Mach number hydrodynamics code \maestro,
-sharing much of the same infrastructure, and allowing for simulations
-to be restarted from the low Mach regime into the compressible
-regime~\cite{malone:2014}.  \castro\ has been applied to various
+mesh refinement library with space- and time-refinement (BoxLib at the
+time) and unsplit hydrodynamics techniques.  \castro\ now uses the
+AMReX adaptive mesh refinement library to handle the computational
+grid, memory management, and parallelism, and the \castro\ and AMReX
+developers work closely together to The initial application area was
+nuclear astrophysics, including core-collapse supernovae, so it was
+designed to work with a general equation of state~\cite{zingalekatz}
+and arbitrary nuclear reaction networks, and with radiation transport
+(through flux limited diffusion, \citealt{castro2,castro3}).  It
+includes full self-gravity using multigrid methods with isolated
+boundaries \citep{katz:2016}.  Finally, \castro\ also acted as the
+compressible counterpart to our low Mach number hydrodynamics code
+\maestro, sharing much of the same infrastructure, and allowing for
+simulations to be restarted from the low Mach regime into the
+compressible regime~\cite{malone:2014}.
+
+\castro\ has been used for science investigations of various different
 progenitor models of Type Ia
@@
 dynamics~\cite{ryu:2018}, core-collapse and population III
-supernovae~\cite{chen:2014,dolence:2015,chen:2017}, and X-ray bursts~\cite{astronum:2018}..
+supernovae~\cite{chen:2014,dolence:2015,chen:2017}, and X-ray
+bursts~\cite{astronum:2018}..
 
@@
 
+In \castro\ the domain is decomposed into a number of grids at various
+levels of refinement.  A single grid and its associated data is called
+a \fab, and its location in space is specified uniquely by the index
+of its lower left cell, typically called {\tt lo}, and its upper right
+cell, {\tt hi}, in a global index space for that level of refinement.
+The collection of all the \fab s at a given level of refinement is
+called a \multifab.  Time advancement proceeds by advancing all of the
+grids at the coarsest level through a timestep $\Delta t$, and then
+the grids at the next refinement level through {\tt rr} timesteps of
+$\Delta t/\mathtt{rr}$, where {\tt rr} is the refinement ratio between
+the levels, typically 2 or 4.  At that point, the two levels are at
+the same time and a synchronization is done, as discussed in \cite{castro}.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=0.33\textwidth]{gpu_1} 
+\includegraphics[width=0.33\textwidth]{gpu_2} 
+\includegraphics[width=0.33\textwidth]{gpu_3}
+\caption{\label{fig:loops} caption}
+\end{figure*}
+
+The main drivers in Castro, including the memory management and
+parallelism are written in C++.  The vast majority of the compute
+kernels are written in modern Fortran, although with time we are
+converting these to C++ as well.
+
 \subsection{Offloading to GPUs}
+
+GPUs achieve their performance by having a large number of threads
+(1000s) that can execute the same instructions on multiple pieces of
+data (a SIMD approach). \MarginPar{Max will not like this desciption.}
+In modern supercomputers, like Summit at OLCF, more than 95\% of the
+theoretical floating point performance is from the
+GPUs \MarginPar{cite?}.  The main bottleneck with these hybrid systems
+is copying the data back and forth between the main memory accessible
+to the CPU and the memory on the GPU.  Therefore, to get the best
+perfomance, we want to copy the data once onto the GPUs, operate as
+much as possible on it while it is on the GPUs, and then copy back to
+CPUs only when needed.  To simplify this data transfer, we use CUDA
+managed memory.\MarginPar{more here}.  This approach will require us
+to write all of our code in a manner that can be run on the GPUs.  We
+discuss some of the issues here in this section.
+
+There are several approaches to offloading a compute kernel to GPUs,
+including the directive-based OpenACC and OpenMP, and the
+NVIDIA-specific CUDA library.  For the present work, we use CUDA,
+although our approach can mix the use of the directive-based
+offloading instead of using pure CUDA.  This is something we will
+explore in the future.
 
@@
 Describe MPI, MPI+OpenMP w/ tiling, MPI + CUDA
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.3\textwidth]{gpu_1} 
-\includegraphics[width=0.3\textwidth]{gpu_2} 
-\includegraphics[width=0.3\textwidth]{gpu_3}
-\caption{\label{fig:loops} caption}
-\end{figure*}
 
@@
 
-Future work: gravity on GPUs (multigrid or multipole), port of \maestro\ following
-the same ideas.
+Self-gravity is the next important physics module that needs to be
+developed for GPUs.  \castro\ uses either geometric multigrid,
+implemented via AMReX, or a multipole expansion to solve for the
+gravitational potential.  With Strang splitting of the reactions, it
+is possible to asynchronously solve for the gravitational potential on
+CPUs while solving the second-half of the reaction system on GPUs,
+since the operator split approach does not change density via
+reactions.  This is the approach we will explore in the near future
+for self-gravitating reacting flows.
+
+We will also begin using the same ideas described here to port the
+hydrodynamics and reactions in \maestroex\ to GPUs.  The projection of
+the velocity field will remain on CPUs in the near future.
 
@@
 
-\acknowledgements \castro\ is freely available at
+\acknowledgements \castro\ is available at
 \url{http://github.com/AMReX-Astro/Castro}.  The work at Stony Brook
@@
 which is supported by the Office of Science of the U.S. Department of
-Energy under Contract No. DE-AC05-00OR22725.  We thank NVIDIA Corporation
-for the donation of a Titan X Pascal and Titan V used in this research.
+Energy under Contract No. DE-AC05-00OR22725.  In particular, we thank
+OLCF for early access to the Summitdev platform, where we could
+develop and debug our GPU approach.  We also thank NVIDIA Corporation
+for the donation of a Titan X Pascal and Titan V used in this
+research.  The GPU development of \castro\ benefited greatly from
+numerous GPU hackathons arranged by OLCF.
 

### 2019-03-22 gpus/paper.tex

@@
 
+\begin{figure*}
+\centering
+\includegraphics[width=0.3\textwidth]{gpu_1}
+\includegraphics[width=0.3\textwidth]{gpu_2}
+\includegraphics[width=0.3\textwidth]{gpu_3}
+\caption{\label{fig:loops} caption}
+\end{figure*}
+
 \subsection{Summary of changes required}

### 2019-03-22 gpus/paper.tex
fix double fig
added more references

--- 2019-03-22 gpus/paper.tex
@@
 codes, including
-\citet{gamer,cholla,fargo3d,pekkila:2017,caplan:2018,racz:2018,liska:2018,goz:2018,wang:2010,Padioleau2019,Bryan2014,Kulikov2014} \MarginPar{todo:
-  group by AMR/no-AMR, GPU exclusive, Kokkos, ...}
+\citet{gamer,cholla,fargo3d,pekkila:2017,caplan:2018,racz:2018,liska:2018,goz:2018,wang:2010,Padioleau2019,Bryan2014,Kulikov2014,Bedorf2012,Schive2018,Potter2016} \MarginPar{todo:
+  group by AMR/no-AMR, GPU exclusive, Kokkos \citep{CarterEdwards2014}, ...}
 

### 2019-03-22 gpus/paper.tex

--- 2019-03-22 gpus/paper.tex
@@
 codes, including
-\citet{gamer,cholla,fargo3d,pekkila:2017,caplan:2018,racz:2018,liska:2018,goz:2018,wang:2010} \MarginPar{todo:
-  group by AMR/no-AMR, GPU exclusive, Kokkos, ...}
+\citet{gamer,cholla,fargo3d,pekkila:2017,caplan:2018,racz:2018,liska:2018,goz:2018,wang:2010,Padioleau2019,Bryan2014,Kulikov2014,Bedorf2012,Schive2018,Potter2016} \MarginPar{todo:
+  group by AMR/no-AMR, GPU exclusive, Kokkos \citep{CarterEdwards2014}, ...}
 

### 2019-03-22 gpus/paper.tex
more text

--- 2019-03-22 gpus/paper.tex
@@
 accelerators were increasingly turned to, include GPUs and many-core
-chips like the Intel Phi series.  The simulation codes we use in
-astrophysics need to keep pace with these developments, or they will
+chips like the Intel Phi series.  Throughout this evolution, the simulation codes we use in
+astrophysics had to continue adapt to keep pace with the new computing paradigms, or they will
 not be able to run on the next generation of supercomputers.  In
@@
 In this paper, we describe our approach to performance portability
-with the \castro\ multiphysics simulation code.  Our goal is to have
-the same computational kernels run efficiently on CPUs (including
-manycore processors) and GPUs.  We describe the work we needed to do
-to enable this portability and show performance metrics for a reacting
+with the \castro\ multiphysics simulation code.  We choose not to use
+a framework like Kokkos but instead we implement the mechanism for
+offloading compute kernels to GPUs inside of the AMReX adaptive mesh
+refinement library itself.  Our goal is to have the same computational
+kernels run efficiently on CPUs (including manycore processors) and
+GPUs.  Early GPU results with a version of \castro\ were desribed in
+\citet{astronum:2017}.  Here, we describe the work we needed to do to
+enable this portability and show performance metrics for a reacting
 flow problem in astrophysics.
@@
 progenitor models of Type Ia
-supernova~\cite{malone:2014,katz:2016,polin:2019}, exoplanet
-dynamics~\cite{ryu:2018}, core-collapse and population III
-supernovae~\cite{chen:2014,dolence:2015,chen:2017}, and X-ray
-bursts~\cite{astronum:2018}..
+supernova~\citep{malone:2014,katz:2016,polin:2019}, exoplanet
+dynamics~\citep{ryu:2018}, core-collapse and population III
+supernovae~\citep{chen:2014,dolence:2015,chen:2017}, and X-ray
+bursts~\citep{astronum:2018}.
 
@@
 \includegraphics[width=0.33\textwidth]{gpu_3}
-\caption{\label{fig:loops} caption}
+\caption{\label{fig:loops} Examples of distributing the work inside a
+  grid patch.  On the left is the pure MPI approach.  An entire patch
+  is assigned to a MPI task and the work loop runs over all of the
+  cells in the grid, $\mathtt{lo:hi}$.  At the center is the
+  tile-based MPI+OpenMP approach.  Again the grid is owned by an MPI
+  task, but now it is logically subdivided into files (two shown
+  here), and each tile is assigned to an OpenMP thread.  Each thread
+  loops only over its tile, where the bounds are now given as
+  $\mathtt{lo_{tile}:hi_{tile}}$.  Finally on the right is a GPU
+  approach.  As with the others, the grid is owned by an MPI task but
+  now each zone is assigned to a different GPU thread on the
+  accelerator.  This is achieved by calling the same compute kernels,
+  with $\mathtt{lo_{gpu}:hi_{gpu}}$ now corresponding to a single cell.}
 \end{figure*}
@@
 because of the need to deal with Fortran syntax, preprocessor
-directives, ...  Thus the pure CUDA route for the compute kernels makes the most sense
-with C/C++ kernels, and not Fortran kernels.
-
-OpenACC
-
-OpenMP
-
-Discuss frameworks vs.\ doing it outselves
+directives, ...  Thus the pure CUDA route for the compute kernels
+makes the most sense with C/C++ kernels, and not Fortran kernels.
 
@@
 Describe MPI, MPI+OpenMP w/ tiling, MPI + CUDA
-
-\begin{figure*}
-\centering
-\includegraphics[width=0.3\textwidth]{gpu_1}
-\includegraphics[width=0.3\textwidth]{gpu_2}
-\includegraphics[width=0.3\textwidth]{gpu_3}
-\caption{\label{fig:loops} caption}
-\end{figure*}
 
@@
 
-\section{Scaling}
+The reaction update requires solving a system of ordinary differential
+equations (ODEs).  Astrophysical reaction networks are notoriously
+stiff, so implicit integrators are used (see,
+e.g.\ \citealt{timmes:1999}).  \castro\ uses the VODE integrator
+\citep{vode} which implements a variable order backward difference
+implicit method to advance the system.  This requires the network to
+provide a Jacobian as well as a routine to evaluate the righthand side
+of the reaction ODE system.  It is these evaluations that dominate the
+floating point work.  For integrating the system on GPUs, our approach
+is to put the entire ODE integration on GPUs: the evaluation of the
+righthand side and Jacobian as well as the driver which subcycles in
+time to integrate the system to the desired tolerance.  VODE is
+written in old Fortran, with numerous deprecated features that are not
+supported in CUDA Fortran (like {\tt go to} statements and {\tt
+  common} blocks).  We updated the code to eliminate these unsupported
+statements, using more modern Fortran constructs.  The resulting
+codebase is called VODE90 and serves as our GPU integrator.  \MarginPar{more here from Don}
+
+\section{Performance}
+
+On summit, we create a resource set with 1 MPI, 1 GPU, and 7 threads,
+and there are 6 resource sets per node.
+
+Do we also want to run on Cori?
 
@@
 
-This problem uses constant gravity, so all of the physics can be done on the GPUs.
+The {\tt flame\_wave} problem in \castro\ was described in
+\cite{astronum:2018}.  This problem explores the lateral propagation
+of a helium deflagration across a neutron star in a plane-parallel
+atmosphere.  The Castro PPM CTU hydrodynamics solver is used, with a
+constant gravitational acceleration, and the problem is solved in the
+corotating frame, where the Coriolis forces acts to balance the
+lateral spreading in the atmosphere via geostrophic balance.  For the
+microphysics, we use the general stellar equation of state of
+\cite{timmes_swesty:2000} and the stellar conductivity described in
+\cite{Timmes00}.  The science goal of the problem is to understand the
+dynamics of a resolved burning front in X-ray bursts.  For this work,
+this problem serves as an ideal reacting flow problem to benchmark
+GPUs with.  All of the necessary physics: hydrodynamics, equation of
+state, thermal conductivities, reactions, hydrodynamic source terms,
+and thermal diffusion are run on the GPU.  We look at two different
+reaction networks, a 7-isotope alpha network and a 13-isotope alpha
+network~\cite{iso7}.
 

### 2019-03-22 gpus/paper.tex
update bib style
updates

@@
-%\documentclass[12pt, preprint]{aastex}
 \documentclass[twocolumn,times]{aastex62}
-%\documentclass[preprint,times]{aastex6}
 
@@
 
+\usepackage[T1]{fontenc}
 \usepackage{epsf,color,amsmath}
@@
 nuclear astrophysics, including core-collapse supernovae, so it was
-designed to work with a general equation of state~\cite{zingalekatz}
+designed to work with a general equation of state~\citep{zingalekatz}
 and arbitrary nuclear reaction networks, and with radiation transport
@@
 simulations to be restarted from the low Mach regime into the
-compressible regime~\cite{malone:2014}.
+compressible regime~\citep{malone:2014}.
 
@@
 microphysics, we use the general stellar equation of state of
-\cite{timmes_swesty:2000} and the stellar conductivity described in
-\cite{Timmes00}.  The science goal of the problem is to understand the
+\citep{timmes_swesty:2000} and the stellar conductivity described in
+\citep{Timmes00}.  The science goal of the problem is to understand the
 dynamics of a resolved burning front in X-ray bursts.  For this work,
@@
 
-\bibliographystyle{apj}
+\bibliographystyle{aasjournal}
 \bibliography{ws}

### 2019-03-22 gpus/paper.tex
some fixes

@@
 
+To understand the GPU performance of the reactions, we run a simple pure reaction test.  \MarginPar{is this {\tt test\_react}?}
+
 \subsection{X-ray burst}
 
-The {\tt flame\_wave} problem in \castro\ was described in
+The {\tt flame\_wave} problem\footnote{Found in {\tt Castro/Exec/science/flame\_wave/}} in \castro\ was described in
 \cite{astronum:2018}.  This problem explores the lateral propagation
@@
 reaction networks, a 7-isotope alpha network and a 13-isotope alpha
-network~\cite{iso7}.
+network~\citep{iso7}.
 
@@
 which is supported by the Office of Science of the U.S. Department of
-Energy under Contract No. DE-AC05-00OR22725.  In particular, we thank
+Energy under Contract No.\ DE-AC05-00OR22725.  In particular, we thank
 OLCF for early access to the Summitdev platform, where we could

### 2019-03-22 gpus/paper.tex
some updates to summary

@@
 kernels run efficiently on CPUs (including manycore processors) and
-GPUs.  Early GPU results with a version of \castro\ were desribed in
+GPUs.  Early GPU results with a version of \castro\ were described in
 \citet{astronum:2017}.  Here, we describe the work we needed to do to
@@
 
+The Nyx cosmological code branched from \castro\ some time ago, but
+they share a lot of the same compute kernels, so it should porting the
+GPU-based hydro solver to Nyx is straightforward (and already in
+progress).
+
 We will also begin using the same ideas described here to port the

### 2019-03-25 MAESTROeX_I/results.tex
scaling figures

@@
 \subsection{Scaling}
-In Figure XXX we show weak scaling results for a spherical, full-star sub-chandrasekhar mass white dwarf calculation.
-The x-axis represents total core count (in this case, the total number of OpenMP threads), and the y-axis is the wallclock time per time step.
+In Figure XXX we show weak scaling results for convection preceding ignition in a spherical, full-star sub-Chandrasekhar mass white dwarf.
+These results are representative of simulations used in actual scientific studies; see \cite{MAESTRO_convection,MAESTRO_AMR}.
+For this scaling study we use a spatially uniform grid (no AMR).
+
 These simulations were performed using the NERSC cori system on the Intel Xeon Phi (KNL) partition.
+
+We compare the original BoxLib-based MAESTRO implementation to the AMReX MAESTROeX implementation with the classic algorithm (original temporal integrator).
+
+We also include a plot of MAESTROeX without base state evolution.
+
+The y-axis is the wallclock time per time step, and the x-axis represents total core count (in this case, the total number of OpenMP threads).
+
 We split each KNL node to contain 4 MPI processes with 16 threads each.
+We note that KNL can support up to 272 hardware threads per node (i.e., 68 hardware threads per MPI process since here we assign 4 MPI processes per node), but testing reveals that using $64^3$ grids, additional threads do not decrease the wallclock time.
+Thus, the more accurate measure of weak scaling is to consider the number of MPI processes, since the scaling plot would look virtually identical for larger thread counts.
+
+In principle, we could use larger grids and use more OpenMP threads per MPI process, but the overall wallclock time would increase.
+
 The computational domain for each simulation is divided into $64^3$ grid cells, and we assign 1 MPI process to each grid.
@@
 We see that the total increas in wallclock time from the smallest to largest simulation is roughly 41\%, which is quite remarkable given that there are 3 linear solves per time step.
+
+Include a plot showing the linear solver scaling, and a discussion about how the bottleneck for typical MAESTRO simulations is the averaging operator.
+
 \MarginPar{Making new plots splitting out projections vs. everything else.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=2.75in]{./figs/MAESTRO_scaling1}
+\includegraphics[width=2.75in]{./figs/MAESTRO_scaling2}
+\caption{\label{fig:scaling} (Left) Weak scaling results for a spherical, full-star sub-Chandrasekhar mass white dwarf calculation using the original MAESTRO code, MAESTROeX, and MAESTROeX with base state evolution disabled.
+Shown is the average wallclock time per time step.
+(Right) Weak scaling results showing the average wallclock time per time step spent in the cell-centered and nodal linear solvers within a full time step of the aforementioned simulations.}
+\end{center}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2019-03-25 MAESTROeX_I/results.tex
saving progress

@@
 
-\subsection{AMR Performance}
-Show 3D reacting bubble rise performance.
-Show 3-level spherical performance (stats won't be as good but explain why: refining a larger percentage of grid)
-
-\subsection{Scaling}
-In Figure XXX we show weak scaling results for convection preceding ignition in a spherical, full-star sub-Chandrasekhar mass white dwarf.
-These results are representative of simulations used in actual scientific studies; see \cite{MAESTRO_convection,MAESTRO_AMR}.
+\subsection{Performance and Scaling}
+In Figure \ref{fig:scaling} we show weak scaling results for simulations of convection preceding ignition in a spherical, full-star sub-Chandrasekhar mass white dwarf.
+We emphasize that these results are obtained using simulations used in actual scientific studies; see \cite{MAESTRO_convection,MAESTRO_AMR}.
 For this scaling study we use a spatially uniform grid (no AMR).
@@
 
-We compare the original BoxLib-based MAESTRO implementation to the AMReX MAESTROeX implementation with the classic algorithm (original temporal integrator).
+We compare the original BoxLib-based MAESTRO implementation to the AMReX MAESTReX implementation with the same algorithm.
+We also include a plot of MAESTROeX without base state evolution.
 
-We also include a plot of MAESTROeX without base state evolution.
+For the new temporal algorithm, and the option for irregular base state mapping, <COMMENT HERE ABOUT RUNTIME>
 
@@
 
-\MarginPar{Making new plots splitting out projections vs. everything else.}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\subsection{AMR Performance}
+Show 3D reacting bubble rise performance.
+Show 3-level spherical performance (stats won't be as good but explain why: refining a larger percentage of grid)
+Images of grid configuration.
+
 \subsection{White Dwarf Convection}

### 2019-03-25 MAESTROeX_I/results.tex
updates

@@
 \subsection{Performance and Scaling}
-In Figure \ref{fig:scaling} we show weak scaling results for simulations of convection preceding ignition in a spherical, full-star sub-Chandrasekhar mass white dwarf.
-We emphasize that these results are obtained using simulations used in actual scientific studies; see \cite{MAESTRO_convection,MAESTRO_AMR}.
-For this scaling study we use a spatially uniform grid (no AMR).
+We perform weak scaling tests for simulations of convection preceding ignition in a spherical, full-star sub-Chandrasekhar mass white dwarf.
+The simulation setup remains the same as reported in Section 3 of \cite{MAESTRO_AMR} and originally used in \cite{MAESTRO_convection}, and thus we emphasize that these scaling tests are performed using meaningful, scientific calculations.
+Here, we perform simulations using $256^3, 512^3, 768^3, 1024^3, 1280^3$, and $1536^3$ grid cells on a spatially uniform grid (no AMR).
+We divide each simulation into $64^3$ grids, so this simulations contain between 64 grids ($256^3$) and 13,824 grids ($1536^3$).
+These simulations were performed using the NERSC cori system on the Intel Xeon Phi (KNL) partition.
+Each node contains 68 cores, each capable of supporting up to 4 hardware threads (i.e., a maximum of 272 hardware threads per node).
+For these tests, we assign 4 MPI tasks to each node, and 16 OpenMP threads per MPI process.
+For $64^3$ grids we discovered that using more than 16 OpenMP threads did not decrease the wallclock time due to a lack of work available per grid; in principle one could use larger grids, fewer MPI processes, and more threads per MPI process, however the overall wallclock time would increase except for extremely large numbers of MPI processes (beyond the range we tested here).
 
-These simulations were performed using the NERSC cori system on the Intel Xeon Phi (KNL) partition.
-
-We compare the original BoxLib-based MAESTRO implementation to the AMReX MAESTReX implementation with the same algorithm.
+In Figure \ref{fig:scaling} we compare the wallclock time per time step for the original BoxLib-based MAESTRO implementation to the AMReX MAESTReX implementation with the same algorithm.
 We also include a plot of MAESTROeX without base state evolution.
-
-For the new temporal algorithm, and the option for irregular base state mapping, <COMMENT HERE ABOUT RUNTIME>
 
@@
 Include a plot showing the linear solver scaling, and a discussion about how the bottleneck for typical MAESTRO simulations is the averaging operator.
+
+\MarginPar{Include commentary about new temporal algorithm, and new temporal algorithm with irregular base state}
+For the new temporal algorithm, and the option for irregular base state mapping, <COMMENT HERE ABOUT RUNTIME>
 

### 2019-03-25 MAESTROeX_I/results.tex
expanded scaling discussion

@@
 For these tests, we assign 4 MPI tasks to each node, and 16 OpenMP threads per MPI process.
-For $64^3$ grids we discovered that using more than 16 OpenMP threads did not decrease the wallclock time due to a lack of work available per grid; in principle one could use larger grids, fewer MPI processes, and more threads per MPI process, however the overall wallclock time would increase except for extremely large numbers of MPI processes (beyond the range we tested here).
-
-In Figure \ref{fig:scaling} we compare the wallclock time per time step for the original BoxLib-based MAESTRO implementation to the AMReX MAESTReX implementation with the same algorithm.
-We also include a plot of MAESTROeX without base state evolution.
-
-The y-axis is the wallclock time per time step, and the x-axis represents total core count (in this case, the total number of OpenMP threads).
-
-We split each KNL node to contain 4 MPI processes with 16 threads each.
-We note that KNL can support up to 272 hardware threads per node (i.e., 68 hardware threads per MPI process since here we assign 4 MPI processes per node), but testing reveals that using $64^3$ grids, additional threads do not decrease the wallclock time.
+Each MPI task is assigned to a single grid, so our tests use between 64 and 13,824 MPI processes (i.e., between 1,024 and 221,184 total OpenMP threads).
+For $64^3$ grids we discovered that using more than 16 OpenMP threads did not decrease the wallclock time due to a lack of work available per grid; in principle one could use larger grids, fewer MPI processes, and more threads per MPI process to obtain a flatter weak scaling curve, however the overall wallclock time would increase except for extremely large numbers of MPI processes (beyond the range we tested here).
 Thus, the more accurate measure of weak scaling is to consider the number of MPI processes, since the scaling plot would look virtually identical for larger thread counts.
-
-In principle, we could use larger grids and use more OpenMP threads per MPI process, but the overall wallclock time would increase.
-
-The computational domain for each simulation is divided into $64^3$ grid cells, and we assign 1 MPI process to each grid.
-We used simulations ranging from $256^3$ grid cells (64 MPI processes, corresponding to 1024 total threads) up to $1536^3$ grid cells (13,824 MPI processes, corresponding to 221,184 total threads).
 Note that the largest simulation used roughly 36\% of the entire computational system.
-We see that the total increas in wallclock time from the smallest to largest simulation is roughly 41\%, which is quite remarkable given that there are 3 linear solves per time step.
-
-Include a plot showing the linear solver scaling, and a discussion about how the bottleneck for typical MAESTRO simulations is the averaging operator.
-
-\MarginPar{Include commentary about new temporal algorithm, and new temporal algorithm with irregular base state}
-For the new temporal algorithm, and the option for irregular base state mapping, <COMMENT HERE ABOUT RUNTIME>
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+In the left panel of Figure \ref{fig:scaling} we compare the wallclock time per time step as a function of total core count (in this case, the total number of OpenMP threads) for the original FBoxLib-based MAESTRO implementation to the AMReX MAESTROeX implementation.
+These tests were performed using the original temporal integration strategy in \cite{MAESTRO_V}, noting that the new temporal integration scheme gives very similar performance results.\MarginPar{quantify this?...
+and also irregular dr case?}
+We also include a plot of MAESTROeX without base state evolution.
+Comparing the original and new implementations, we see similar scaling results except for the largest simulation, where MAESTROeX performs better.
+We see that the increase in wallclock time from the smallest to largest simulation is roughly 42\%.
+We also note that without base state evolution, the code runs 14\% faster for small problems, and scale mush better
+However, for the case without base state evolution, the overall runtime is faster and the increase in wallclock time from the smallest to largest simulation is roughly 13\%.
+This is quite remarkable since there are 3 linear solver per time step (2 cell-centered Poisson solves used in the MAC projection, and a nodal Poisson solve used to compute the updated cell-centered velocities).
+Contrary to our prior assumptions, the linear solves are not the primary scaling bottleneck in this code.
+In the right panel of Figure \ref{fig:scaling}, we isolate the wallclock time required for these linear solves and see that (i) the linear solves only use 20-23\% of the total computational time, and (ii) the increase in the solver wallclock time from the smallest to largest simulation is only 28\%.
+Further profiling reveals that the primary scaling bottleneck is the average operator.
+The averaging operator requires binning the sum of Cartesian data onto one-dimensional arrays holding every possible mapping radius.
+This amounts to at least 24,384 points (for the $256^3$ simulation) up to 883,584 point (for the $1536^3$ simulation).
+The averaging operator requires a global sum reduction over all processors, and the communication of this data is the primary scaling bottleneck.
+For the simulation with base state evolution, this averaging operator is only called once per time step (as opposed to 14 times per time step when base state evolution is included).
+The difference in total wallclock times with and without base state evolution is almost entirely due to the averaging.
+Note that as expected, advection, reactions, and calls to the equation of state scale almost perfectly, since there is only a single parallel communication call to fill ghost cells.
+
 \subsection{AMR Performance}

### 2019-03-25 MAESTROeX_I/algorithm.tex
simplify alg description

@@
 
+In stratified low Mach number models, due to the extreme variation in density, the velocity can become very large in low density regions at the edge of the star.
+Throughout Papers II-V, we have employed two techniques to help control these dynamics without significantly affecting the dynamics in the convective region.
+First, we use a cutoff density technique, where we hold the density constant outside a specified radius (typically near where the density is $\sim$4 orders of magnitude smaller than the largest densities in the simulation.
+Second, we employ a sponge technique where we artificially damp the velocities near and beyond the cutoff region.
+For more details, refer to Paper V and the previous references cited within.
+
 To initialize the simulation we use the same procedure described in Paper III.
@@
 The new temporal integration strategy in MAESTROeX contains the following steps:
-\MarginPar{mention cutoff densities}
-
 \begin{description}
@@
 %
-again omitting the reaction and heating terms
-since we already accounted for
+again omitting the reaction terms since we already accounted for
 them in {\bf React State}.  This equation takes the form:
@@
 
-\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$.
-\begin{description}
-\item For planar geometry,
-\begin{eqnarray}
+\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$ using
+\begin{equation}
 (\rho h)^{(2),\star}
-&=& (\rho h)^{(1)} \nonumber \\
-&&- \dt \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \left(p_0^{n+1,*} - p_0^n\right) \nonumber \\
-&& + \dt \left(\uadvone \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^{n}
-\end{eqnarray}
-
-\item For spherical geometry,
-\begin{eqnarray}
-(\rho h)^{(2),\star}
-&=& (\rho h)^{(1)} \nonumber \\
-&&- \dt \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \left(p_0^{n+1,*} - p_0^n\right) \nonumber \\
-&& + \dt \left \{ \nabla \cdot \left (\uadvone p_0^{n+\myhalf} \right ) - p_0^{n+\myhalf} \nabla \cdot \uadvone \right \}
-\end{eqnarray}
-\end{description}
-where we define $p_0^\nph$ as $p_0^\nph = (p_0^n+p_0^{n+1,*})/2$.
-
+= (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} + \Delta t\frac{Dp_0}{Dt}
+\end{equation}
+where here
+\begin{equation}
+\frac{Dp_0}{Dt} =
+\begin{cases}
+\frac{p_0^{n+1,*} - p_0^n}{\Delta t} + \left(\uadvone \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^{n}& {\rm (planar)} \\
+\frac{p_0^{n+1,*} - p_0^n}{\Delta t} + \left[ \nabla \cdot \left (\uadvone p_0^{\nph} \right ) - p_0^{\nph} \nabla \cdot \uadvone \right]& {\rm (spherical)}
+\end{cases}
+,
+\end{equation}
+and $p_0^\nph = (p_0^n+p_0^{n+1,*})/2$.
 \end{enumerate}
-Note that outside the cutoff density region, we compute enthalpy using the EOS.
-
 \end{enumerate}
@@
 \nabla \cdot \left(\beta_0^{\nph} \uadvtwo\right) =
-\beta_0^{\nph} \left[S^{\nph} - \frac{1}{\gammaonebar^{n+\myhalf}p_0^{n+\myhalf}}\left(\frac{\partial p_0}{\partial t}\right)^{n+\myhalf}\right],\label{eq:div2}
+\beta_0^{\nph} \left[S^{\nph} - \frac{1}{\gammaonebar^{\nph}p_0^{\nph}}\left(\frac{\partial p_0}{\partial t}\right)^{\nph}\right],\label{eq:div2}
 \end{equation}
@@
 \item Update the enthalpy.  This step is identical to {\bf Step 4D} except we use
-  the updated values $\uadvtwo, \rho_0^{n+1}, (\rho h)_0^{n+1}$, and $p_0^{n+\myhalf}$
+  the updated values $\uadvtwo, \rho_0^{n+1}, (\rho h)_0^{n+1}$, and $p_0^{\nph}$
   rather than
@@
 \item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2)}$.
-\begin{description}
-\item For planar geometry,
-\begin{eqnarray}
+\begin{equation}
 (\rho h)^{(2)}
-&=& (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvtwo (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \left( p_0^{n+1} - p_0^n \right)\nonumber \\
-&& + \dt \left(\uadvtwo \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^\nph,
-\end{eqnarray}
-
-\item For spherical geometry,
-\begin{eqnarray}
-(\rho h)^{(2)}
-&=& (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvtwo (\rho h)^{\nph,\pred} \right] \right\} \nonumber \\
-&& + \left(p_0^{n+1} - p_0^n\right)\nonumber \\
-&& + \dt \left[ \nabla \cdot \left (\uadvtwo p_0^{\nph} \right ) - p_0^{\nph} \nabla \cdot \uadvtwo \right],
-\end{eqnarray}
-\end{description}
-where we now define $p_0^\nph$ as $p_0^\nph = (p_0^n+p_0^{n+1})/2$.
-
+= (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvtwo (\rho h)^{\nph,\pred} \right] \right\} + \Delta t\frac{Dp_0}{Dt}
+\end{equation}
+where here
+\begin{equation}
+\frac{Dp_0}{Dt} =
+\begin{cases}
+\frac{p_0^{n+1} - p_0^n}{\Delta t} + \left(\uadvtwo \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^{n}& {\rm (planar)} \\
+\frac{p_0^{n+1} - p_0^n}{\Delta t} + \left[ \nabla \cdot \left (\uadvtwo p_0^{\nph} \right ) - p_0^{\nph} \nabla \cdot \uadvtwo \right]& {\rm (spherical)}
+\end{cases}
+,
+\end{equation}
+and $p_0^\nph = (p_0^n+p_0^{n+1})/2$.
 \end{enumerate}
-Note that outside the cutoff density region, we compute enthalpy using the EOS.
-
 \end{enumerate}
@@
 \begin{equation}
-\nabla \cdot \left(\beta_0^{n+1} \Ub^{n+1} \right) = \beta_0^{n+1} \left[ S^{n+1} - \frac{1}{\gammaonebar^{n+1}p_0^{n+1}}\left(\frac{\partial p_0}{\partial t}\right)^{n+\myhalf}\right].\label{eq:div3}
+\nabla \cdot \left(\beta_0^{n+1} \Ub^{n+1} \right) = \beta_0^{n+1} \left[ S^{n+1} - \frac{1}{\gammaonebar^{n+1}p_0^{n+1}}\left(\frac{\partial p_0}{\partial t}\right)^{\nph}\right].\label{eq:div3}
 \end{equation}

### 2019-03-28 gpus/paper.tex
translate slack discussion

@@
 
+\section{Runtime Considerations}
+
+Volta can have as many as 160k threads in flight at any one time,
+whereas a $32^3$ box of cells is 30k.  This means that we haven't
+exposed enough parallelism to keep the device busy.  Generally we
+launch each box in a separate stream so it's often OK to do this, but
+other times this is a bottleneck, and we need bigger boxes for best
+performance.
+
 \section{Performance}

### 2019-04-02 MAESTROeX_I/intro.tex
updates
added paragraph on recent low Mach papers

@@
 
+In recent years, a number of alternative approaches to modelling low Mach number flows have been developed. These include semi-implicit all-Mach number solvers, where the Euler equations
+are split into an acoustic part and an advective part \citep{Degond2009,Cordier2012,Haack2012,Chalons2016,Kwatra2009,Happenhofer2013,Padioleau2019}. The fast acoustic waves are then
+solved using implicit time integration, while the slow material waves are solved explicitly. Another approach is to use preconditioned all-Mach number solvers
+\citep{Miczek2014,Barsukow2016}, where the numerical flux is multiplied by a preconditioning matrix. This reduces the stiffness of the system at low Mach numbers, while retaining the
+correct scaling behavior. In the reduced speed of sound technique (RSST) and related methods, the speed of sound is artificially reduced by including a suitable scaling factor in the
+continuity equation, reducing the restriction on the size of the timestep \citep{Rempel2005,Hotta2012,Wang2015,Takeyama2017,Iijima2018}.
+The MUSIC code uses fully implicit time integration for the compressible Euler equations, which therefore allows for arbitrarily large timesteps \citep{Viallet2011,Viallet2015,Goffrey2016}.
+
 Previously, we developed the low Mach number astrophysical solver, MAESTRO.
@@
 MAESTRO is a structured-grid, finite-volume code that utlizes adaptive mesh refinement (AMR) to refine grids locally in space.
+This code was developed in the pure-Fortran90 FBoxLib framework (see the FBoxLib repository in \cite{AMReX}) for block-structured AMR applications.
 MAESTRO is suitable for for full spherical stars, as well as planar simulations of dynamics within localized regions of a star.

### 2019-04-02 MAESTROeX_I/intro.tex
reordered references

@@
 In recent years, a number of alternative approaches to modelling low Mach number flows have been developed. These include semi-implicit all-Mach number solvers, where the Euler equations
-are split into an acoustic part and an advective part \citep{Degond2009,Cordier2012,Haack2012,Chalons2016,Kwatra2009,Happenhofer2013,Padioleau2019}. The fast acoustic waves are then
+are split into an acoustic part and an advective part \citep{Kwatra2009,Degond2009,Cordier2012,Haack2012,Happenhofer2013,Chalons2016,Padioleau2019}. The fast acoustic waves are then
 solved using implicit time integration, while the slow material waves are solved explicitly. Another approach is to use preconditioned all-Mach number solvers

### 2019-04-02 MAESTROeX_I/algorithm.tex

@@
 \subsection{Spatial Discretization}\label{Sec:Spatial}
+\MarginPar{needs work}
 MAESTRO contains variables defined on Cartesian grids 
@@
 We used $w_0$ to provide an estimate of the base state density evolution over a time step.
+Our new temporal integration scheme does not use this splitting, and is much simpler than the scheme from Paper V since we evolve the velocity directly rather than the perturbational and base state velocity separately.
+This eliminates
+Additionally, the new scheme uses a simpler predictor-corrector approach to the base state density and pressure that no longer requires the complex algorithm from Paper V, yet still retains the same overall second-order accuracy.
 
-Our new temporal integration scheme does not use this splitting, and is much simpler than the scheme from Paper V.
-In this new scheme we use a simpler, predictor-corrector approach to the base state density and pressure that no longer requires the complex algorithm from Paper V, yet still retains the same overall second-order accuracy.
-Also, we now evolve the full-state velocity directly rather than the perturbational velocity, eliminating the need to compute additional forcing terms that arise due to the splitting.
-
-We now describe the new temporal integration scheme, which has been greatly simplified
-from Paper V without formal loss of order of accuracy.
+We now describe the new temporal integration scheme, which has been greatly simplified from Paper V without formal loss of order of accuracy.
 Note that this temporal integration scheme is valid for the original base state mapping (with constant base state grid spacing), or the new irregularly spaced base state mapping.

### 2019-04-02 MAESTROeX_I/results.tex

@@
 In the left panel of Figure \ref{fig:scaling} we compare the wallclock time per time step as a function of total core count (in this case, the total number of OpenMP threads) for the original FBoxLib-based MAESTRO implementation to the AMReX MAESTROeX implementation.
-These tests were performed using the original temporal integration strategy in \cite{MAESTRO_V}, noting that the new temporal integration scheme gives very similar performance results.\MarginPar{quantify this?...
-and also irregular dr case?}
+These tests were performed using the original temporal integration strategy in \cite{MAESTRO_V}, noting that the new temporal integration with and without the irregular base state gives essentially the same results.
 We also include a plot of MAESTROeX without base state evolution.
@@
 The averaging operator requires binning the sum of Cartesian data onto one-dimensional arrays holding every possible mapping radius.
-This amounts to at least 24,384 points (for the $256^3$ simulation) up to 883,584 point (for the $1536^3$ simulation).
+This amounts to at least 24,384 double precision values (for the $256^3$ simulation) up to 883,584 values (for the $1536^3$ simulation).
 The averaging operator requires a global sum reduction over all processors, and the communication of this data is the primary scaling bottleneck.

### 2019-04-04 MAESTROeX_I/algorithm.tex
Some spelling corrections

@@
 
-(description of base state mapping in planar, and origianl spherical, and new spherical)
+(description of base state mapping in planar, and original spherical, and new spherical)
 
@@
 
-Mapping a Cartesian variable to a one-dimensional radial array uses the lateral averating operator.
+Mapping a Cartesian variable to a one-dimensional radial array uses the lateral averaging operator.
 

### 2019-04-04 MAESTROeX_I/intro.tex
Some spelling corrections

@@
 Central to the algorithm is a stratified background (or base) state density and pressure held in hydrostatic equilibrium, and also vary as a function of altitude and time.
-MAESTRO is a structured-grid, finite-volume code that utlizes adaptive mesh refinement (AMR) to refine grids locally in space.
+MAESTRO is a structured-grid, finite-volume code that utilizes adaptive mesh refinement (AMR) to refine grids locally in space.
 This code was developed in the pure-Fortran90 FBoxLib framework (see the FBoxLib repository in \cite{AMReX}) for block-structured AMR applications.
-MAESTRO is suitable for for full spherical stars, as well as planar simulations of dynamics within localized regions of a star.
+MAESTRO is suitable for full spherical stars, as well as planar simulations of dynamics within localized regions of a star.
 
@@
 \item In Paper I \citep{MAESTRO_I}, we derive the low Mach number equation set from the fully compressible equations.
-\item In Paper II \citep{MAESTRO_II}, we incorporate the effects atmospheric expansion through the use of a time-dependent background state.
+\item In Paper II \citep{MAESTRO_II}, we incorporate the effects of atmospheric expansion through the use of a time-dependent background state.
 \item In Paper III \citep{MAESTRO_III}, we incorporate reactions and the associated coupling to the hydrodynamics.
@@
 The resulting MAESTROeX is implemented in the C++/F90 AMReX public software library \cite{AMReX}.
-MAESTROeX uses MPI+OpenMP parallelism and scales well to over 10,000 MPI processes, with each MPI process supporting tens of threads
+MAESTROeX uses MPI+OpenMP parallelism and scales well to over 10,000 MPI processes, with each MPI process supporting tens of threads.
 The resulting code is publicly available on GitHub at {\url https://github.com/AMReX-Astro/MAESTROeX} and uses the microphysics

### 2019-04-08 xrb/paper.tex
move this over from astro_papers

%\documentclass[12pt, preprint]{aastex}
\documentclass[twocolumn,times,tighten]{aastex62}

\usepackage{epsf,color,amsmath}

\usepackage{cancel}

\newcommand{\sfrac}[2]{\mathchoice%
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
    \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
    \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptscriptfont0 #1}\kern-.2em/
    \kern-.15em\lower.25ex\hbox{\the\scriptscriptfont0 #2}} {#1\!/#2}}

\newcommand{\myhalf}{\sfrac{1}{2}}
\newcommand{\nph}{{n+\myhalf}}
\newcommand{\nmh}{{n-\myhalf}}

\newcommand{\inp}{\mathrm{in}}
\newcommand{\outp}{\mathrm{out}}

% boldsymbol means bold italic
\newcommand{\eb}{{\bf{e}}}
\newcommand{\Ub}{{\bf{U}}}
\newcommand{\xb}{{\bf{x}}}
\newcommand{\kb}{{\bf{k}}}
\newcommand{\Vb}{{\bf{V}}_n}
\newcommand{\Vbhat}{{\bf{\widehat{V}}}_n}
\newcommand{\Omegab}{{\bf{\Omega}}}
\newcommand{\gb}{{\bf{g}}}
\newcommand{\rb}{{\bf{r}}}

\newcommand{\pb}{p_\mathrm{base}}
\newcommand{\epsdot}{\dot{\epsilon}}
\newcommand{\qburn}{q_\mathrm{burn}}
\newcommand{\rt}{\tilde{r}_0}

\newcommand{\nablab}{\mathbf{\nabla}}
\newcommand{\dt}{\Delta\ t}

\newcommand{\omegadot}{\dot{\omega}}

\newcommand{\Hext}{H_{\rm ext}}
\newcommand{\Hnuc}{H_{\rm nuc}}
\newcommand{\kth}{k_{\rm th}}

\newcommand{\Gammaonebar}{\overline{\Gamma}_1}
\newcommand{\Sbar}{\overline{S}}

\newcommand{\etarho}{\eta_\rho}
\newcommand{\etarhoh}{\eta_{\rho~h}}

\newcommand{\Ubt}{\widetilde{\Ub}}
\newcommand{\wt}{\widetilde{w}}

\newcommand{\He}{$^4$He}
\newcommand{\C}{$^{12}$C}
\newcommand{\Fe}{$^{56}$Fe}

\newcommand{\isot}[2]{$^{#2}\mathrm{#1}$}
\newcommand{\isotm}[2]{{}^{#2}\mathrm{#1}}

\newcommand{\maestro}{{\sf MAESTRO}}
\newcommand{\castro}{{\sf Castro}}
\newcommand{\pynucastro}{{\sf pynucastro}}

\newcommand{\avg}[1]{\overline{#1}}
\newcommand{\avgtwod}[1]{\langle~#1 \rangle}
\newcommand{\rms}[2]{\left(\delta#1\right)_{r_{#2}}}
\newcommand{\mymax}[1]{\left(#1\right)_{\rm max}}

\newcommand{\Tb}{\ensuremath{T_\mathrm{base}}}
\newcommand{\gcc}{\mathrm{g~cm^{-3} }}
\newcommand{\cms}{\mathrm{cm~s^{-1} }}

\newcommand{\half}{\frac{1}{2}}

\setlength{\marginparwidth}{0.5in}
\newcommand{\MarginPar}[1]{\marginpar{\vskip-\baselineskip\raggedright\tiny\sffamily\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}

\begin{document}
%======================================================================
% Title
%======================================================================
\title{Dynamics of Laterally Propagating Flames in X-ray Bursts}

\shorttitle{Lateral Flame Dynamics}
\shortauthors{Eiden et al.}

\author{Kiran Eiden}
\affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
             Stony Brook, NY 11794-3800}

\author[0000-0001-8401-030X]{Michael Zingale}
\affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
             Stony Brook, NY 11794-3800}

\author{Yuri Cavecchi}
\affiliation{Princeton University}

\author{Max P.\ Katz}
\affiliation{Nvidia Corp}

%
\author{Chris M.\ Malone}
\affiliation{Los Alamos National Laborary, Los Alamos, NM}

\author{Weiqun Zhang}
\affiliation{Lawrence Berkeley National Laboratory, Berkeley, CA}

\correspondingauthor{Michael Zingale}
\email{Michael.Zingale@stonybrook.edu}

%======================================================================
% Abstract and Keywords
%======================================================================
\begin{abstract}
We investigate the structure of laterally-propagating flames through
the highly-stratified burning layer in an X-ray burst.  Simple
two-dimensional simulations of flame propagation are performed through
a rotating plane-parallel atmosphere, exploring the structure of the
flame.  We discuss the approximations needed to capture the length and
time scales at play in an X-ray burst.  Our studies complement the
other multidimensional studies of burning in X-ray bursts.
\end{abstract}

\keywords{convection---hydrodynamics---methods: numerical---stars: neutron---X-rays: bursts}

%======================================================================
% Introduction
%======================================================================
\section{Introduction}\label{Sec:Introduction}

XRB intro

Evidence for localized burning \MarginPar{refs}

One dimensional studies of XRBs are very successful in predicting the
lightcurves and recurrance times (see, e.g., \citealt{woosley-xrb}).
These assume spherical symmetry and thus cannot capture the effects of
localized burning spreading across the neutron star.  These studies
have also been used to explore the sensitivity of the burst
observables to reaction rates\MarginPar{ref}.

Multidimensional simulations of burning on a neutron star are more
challenging, and thus a number of approximations have been made in the
past.

Laterally propagating detonations were modeled by
\citet{fryxellwoosley82,hedet}, but these would only be applicable at
very high densities, since it is difficult to detonation helium at the
densities found in normal XRBs, and even harder to detonation hydrogen
because of the waiting times for weak reactions.  This means that
detonations may only really be applicable to superbursts (where carbon
is the reactant).

Global multidimensional studies were performed by \citet{spitkovsky2002}, where
it was demonstrated that the Coriolis force plays an important role in 
confining the burning as it spreads across the neutron star surface.  These
calculations used the shallow-water approximation, so the vertical details
of the atmosphere structure we not modeled.

Small-domain studies of convective burning in XRBs preceding flame
development have been done in two-dimensions \citep{lin:2006,xrb,xrb2}
and three-dimensions \citep{xrb3d}.  These calculations used low Mach
number methods that approximate the hydrodynamics equations to filter
soundswaves, enabling large timesteps and efficient modeling of
subsonic convection.  While these calculations could not support
the lateral differences needed for flame spreading, they can help
understand the role that convection plays in distributing the initial
burning products vertically throughout the neutron star atmosphere.

\cite{cavecchi:2013} \MarginPar{and others}

Challenges: timescale for flame propagation, lengthscales (flame
width, Rossby scale), reaction network size.\MarginPar{from talk}

Detonations are in some sense easier to model, since their structure
follows from the jump conditions which are solved with shock-capturing
schemes.  For deflagrations, we either need to resolve the structure
or use a flame model.  Flame models usually assume that the flame
structure is thin compared to the size of the system (see, e.g., XX\MarginPar{refs} for applications to Type Ia supernovae).  For the XRB however, the flame thickness is comparable to the scale
height of the atmosphere, so we cannot use the approximations.

The goal of this study is to understand what numerical and physical
approximations are needed to perform a full hydrodynamical,
multidimensional simulation of flame propagation through the
atmosphere on the neutron star.  For simplicity in this first set of
calculations, we will use pure helium composition.  These studies
complement the other multidimensional work described above in helping
us build a picture of the dynamics of X-ray bursts.

%======================================================================
% Numerical Approach
%======================================================================
\section{Numerical Approach}\label{Sec:numerics}

All simulations are performed with the \castro\ hydrodynamics
code~(\citealt{castro}; see also \citealt{astronum:2017} for a recent
description).  We evolve the system of fully compressible Euler
equations for reacting flow:
\begin{eqnarray}
\frac{\partial( \rho X_k)}{\partial t} &=& - \nabla\cdot (\rho \Ub X_k) + \rho \omegadot_k \\
%
\frac{\partial (\rho \Ub)}{\partial t} &=& - \nabla\cdot (\rho \Ub \Ub) - \nabla p +
    \rho \gb \nonumber \\
  &&-2 \rho \Omegab\times \Ub - \Omegab \times (\Omegab \times \rb) \\
%
\frac{\partial (\rho E)}{\partial t} &=& - \nabla \cdot (\rho \Ub E + p\Ub) +
    \nabla \cdot \kth \nabla T + \rho \epsdot +\nonumber\\
  && \rho \Ub \cdot \gb - \rho (\Omegab \cdot \rb)(\Omegab \cdot \Ub) + \rho |\Omegab|^2 (\Ub \cdot \rb)
\end{eqnarray}
Here, $\rho$ is the mass density, $\Ub$ is the velocity, and $E$ is
the specific total energy, which is related to the specific internal
energy as $e = E - |\Ub|^2/2$.  Species are described by mass
fractions, $X_k$ (such that $\sum_k X_k = 1$), and creation rates,
$\omegadot$, and are related to the total specific energy generation
rate, $\epsdot$.  The total mass conservation,
\begin{equation}
\frac{\partial \rho}{\partial t} = - \nabla\cdot (\rho \Ub)
\end{equation}
implies $\sum_k \omegadot_k = 0$.  \castro\ uses an unsplit piecewise
parabolic method (PPM) with characteristic tracing for solving the
hydrodynamics.  Reactions are incorporated via Strang splitting,
giving a method that is overall second-order accurate in space and
time.

 An equation of state of the form $p = p(\rho, e,
\{X_k\})$ completes the thermodynamic description of the
system.  Finally, $\kth$ is the thermal conductivity and $T$ is the
temperature.

We work in 2D axisymmetric coordinates, but we advective a third
component of velocity, coming out of the simulation plane, that
participates in the Coriolis force (sometimes described as a 2.5D
simulation).  We will take the \castro\ $x$-coordinate to be the
cylindrical $r$ coordinate with corresponding velocity $u$, the
\castro\ $y$-coordinate to be the cylindrical $z$ coordinate with
corresponding velocity $v$, and the \castro\ $z$-coordinate to be the
cylindrical $\theta$ coordinate, with corresponding velocity $w$.  A
righthanded coordinate system has positive $w$ pointing out of the
page.  We will take $\Omegab = \Omega_0 \hat{\bf y}$ for the angular
rotation rate, and $\gb = -g \hat{\bf y}$ for the gravitational
acceleration, with $g$ constant.  With these choices, the Coriolis
force is:
\begin{equation}
-2\rho \Omegab \times \Ub =
   -2\rho \left ( \Omegab w \hat{x} - \Omegab u \hat{z} \right )
\end{equation}
We will neglect the centrifugal force---with our plane-parallel
geometry, this will act only in the lateral direction, and is not
expected to greatly affect the dynamics.  Carrying the Coriolis force
allows us to capture the geostrophic balance\MarginPar{ref} that sets up via lateral
hydrostatic equilibrium.

Writing the momentum equation in terms of the $u$, $v$, and $w$
components, and neglecting the centrifugal force, we have:
\begin{align}
\frac{\partial (\rho u)}{\partial t} + \nabla \cdot (\rho u \Ub) +
     \frac{\partial p}{\partial x} &= -2\rho \Omega w  \\
%
\frac{\partial (\rho v)}{\partial t} + \nabla \cdot (\rho v \Ub) +
     \frac{\partial p}{\partial y} &= -\rho g \\
%
\frac{\partial (\rho w)}{\partial t} + \nabla \cdot (\rho w \Ub) +
     \cancelto{0}{\frac{\partial p}{\partial z}} &=
    2\rho \Omega u 
\end{align}
where we cancel $\partial p/\partial z$ because there are no
variations in the azimuthal direction.  This allows us to recast the
$w$-velocity equation as a simple advection equation:
\begin{equation}
\frac{\partial w}{\partial t} + \Ub \cdot \nabla w = 2\Omega u
\end{equation}
In our geometry, the flame will propagate from left to right, so $u$
will be positive and the Coriolis force results in $w > 0$ (out of the
simulation plane).  The algorithmic implementation of rotation in
\castro\ is described in \cite{wdmergerI}.

We use a general stellar equation of state with nuclei (treated as an
ideal gas), photons, and degenerate/relativistic electrons, as
described in \cite{timmes_swesty:2000}.  We use a simple reaction
network consisting of 4 nuclei, \isot{He}{4}, \isot{C}{12},
\isot{O}{16}, \isot{Fe}{56}.  Only the first 3 are evolved, with the
triple-$\alpha$ rate and $\isotm{C}{12}(\alpha, \gamma)\isotm{O}{16}$.
Screening of the reactions is also included.  The network is provided
in our open-source microphysics
routines\footnote{\url{https://github.com/starkiller-astro/Microphysics}}.

We note that we do not explicitly model viscosity.  The reactions will
provide the smallscale cutoff to the instabilities and turbulence at
the flame front.  We also do not include species
diffusion---astrophysical flames tend to have large Lewis numbers, so
this is not expected to be important~\citep{timmeswoosley:1992}.
Finally, we use the thermal conductivities described in
\citet{Timmes00}.

All simulations use adaptive mesh refinement to refine on the
atmosphere (leaving the space between the top of the atmosphere and
upper boundary at low resolution).  In addition to the base grid, we
use up to 3 refinement levels, the first two a factor of 4 finer than
the previous and the last a factor of 2 finer than the previous.
\castro\ subcycles in time, so the finer grids are evolved with a
finer timestep than the coarse grids.  Occasionally the timestep
chosen at the start of a cycle will violate the CFL condition during
the advance of the finer grids.  In this case, we restart the finer
grid evolution with a smaller timestep, subcycling within the larger
timestep hierarchy.  The finest grid in most of our simulations (base
+ 2 refined levels) corresponds to $4096\times 1024$ zones.

Thermal diffusion is modeled explicitly, using a predictor-corrector
scheme to achieve second-order accuracy.  A verification test of the
diffusion scheme is shown in Appendix~\ref{app:diffusion}.  The
explicit thermal diffusion requires a timestep limiter of the form:
\begin{equation}
\Delta t_\mathrm{diff} \le \frac{1}{2} \frac{\Delta x^2}{\mathcal{D}}
\end{equation}
where $\mathcal{D} = \kth/(\rho c_v)$ is the thermal diffusivity.  The
diffusivity increases rapidly at the top of the atmosphere, making
these low density regions determine the overall timestep for the
simulations.  Therefore, we disable thermal conduction at low
densities where it is not expected to be important.

We use hydrostatic boundary conditions on the lower boundary, using
a discretized hydrostatic equilibrium equation of the form:
\begin{equation}
\label{eq:hse}
p_i = p_{i-1} + \frac{1}{2} \Delta y (\rho_i + \rho_{i-1}) g
\end{equation}
holding the temperature constant in the ghost cells.  This is solved
together with the equation of state.  The velocity is reflected here.
This procedure follows the form described in \citet{ppm-hse}.  The
left boundary is reflecting and the right boundary is a zero-gradient
outflow.  The top boundary sets the state to simply the conditions
in our outer buffer region of the initial model (see below) with the
normal velocity set to the larger of zero or the velocity at the top
of the domain (this prevents incoming velocities at the top) and the
transverse velocities set to zero.

sponge\MarginPar{write}

In the discussions below, we will use the Castro coordinate 
names, $(x, y)$ in our notation.

%======================================================================
% Flame
%======================================================================
\section{Flame Properties}\label{Sec:Flame}

\input{flame_table_1d.tex}

To understand the time and length scales involved in flame
propagation, we do a 1D simulation of a laminar flame using our
microphysics.  Figure~\ref{fig:flame} shows the flame properties for a
He flame using the standard conductivities and our simple reaction
network.  This flame had a density of $2\times 10^6~\gcc$ and
temperature of $5\times 10^7$~K.  We note that this flame speed is
about $2\times 10^4~\cms$, the flame width is about 20~cm, and it
takes about 0.1~s to setting into a sustained flame.  Note this speed
is quite small compared to the speed estimated in
\citet{spitkovsky2002}.  We expected a multidimensional flame to
accelerate due to hydrodynamics interactions (wrinkling, turbulence
interactions, directed flows feeding fuel into the flame, ...).

We measure the flame width as:
\begin{equation}
\lambda_f \equiv \frac{\Delta T}{\max\{|\nabla T|\}}
\end{equation}

\begin{figure*}[t]
\plottwo{flame}{speed}
\caption{\label{fig:flame} Our ``standard'' flame.}
\end{figure*}

To make the simulation feasible, we artifically speed up the flame.
Dimensional analysis shows that the flame speed scales as:\MarginPar{ref}
\begin{equation}
v_f \approx \sqrt{k_\mathrm{th} \dot{\epsilon}}
\end{equation}
and the flame width scales as:
\begin{equation}
\lambda_f = \sqrt{\frac{k_\mathrm{th}}{\dot{\epsilon}}}
\end{equation}
To boost the flame, while keeping the thickness the same, we can
multiply both the burning rate and the conductivity by the same
factor.  We choose 10 for each, to give a flame 10$\times$ faster.

Experience with modeling resolved flames suggests that we need a
spatial resolution, $\Delta x$ of $\lambda_f/\Delta x \sim 5$.  These
conditions represent the bottom of the He layer.  As the density
decreases with altitude, the flame thickness increases and the flame
speed decreases, so we will easily resolve the flame structure throughout
the rest of the atmosphere.

%======================================================================
% Initial Model
%======================================================================
\section{Initial Model}\label{Sec:inital_model}

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{model}
\caption{\label{fig:sketch} A schematic of our model setup.}
\end{figure}

We wish to create an initial atmosphere consisting of a hot
``post-flame'' region and a cooler atmosphere that the flame will
laterally propagate into.  We put the hot region at the very left of
the domain (the origin of the axisymmetric coordinates).  To create
these initial conditions, w produce two different hydrostatic models,
a ``hot'' model that will represent the perturbation that drives the
flame and a ``cool'' model that will represent the state ahead of the
flame.  These will have different scale heights.  To create these
models, we break the vertical structure of the atmosphere into 4
layers: the underlying star, a ramp-up to the base of the
accreted atmosphere, a fuel layer representing the bulk of the
atmosphere where the flame will propagate, and an outer, low density,
isothermal buffer above the atmosphere that allows us
to capture expansion and explosive dynamics.  This is sketched in
Figure~\ref{fig:sketch}.

The temperature profile in the star and ramp region is given as:
\begin{equation}
T = T_\star + \frac{1}{2} (T_\mathrm{hi} - T_\star) \left [ 1 + \tanh\left( \frac{\tilde{y}}{2 \delta_\mathrm{atm}} \right ) \right ]
\end{equation}
with
\begin{equation}
\tilde{y} = y - H_\star - \frac{3}{2} \delta_\mathrm{atm}
\end{equation}
Here, $\delta_\mathrm{atm}$ is a characteristic width of the transition
ramp, $T_\mathrm{hi}$ is the highest temperature in the HSE model---it
will represent the base of the fuel layer.

The mass fractions use this same profile, switching from a set
describing the underlying star, ${X_k}_\star$, and the set for the
accreted material, ${X_k}_\mathrm{atm}$, which is used in the
isentropic and outer regions.  Note, since the profile above is
linear in $X$, if the initial mass fractions sum to one, then the blended
mass fractions in the ramp region also sum to one.

We specify the density, $\rho_\mathrm{int} = \rho(y = H_\star)$ as the
starting point for the integration of hydrostatic equilibrium.  This
is just below the ramp-up region---this ensures that regardless of
what the peak temperature ($T_\mathrm{hi}$) is, the state beneath the
ramp-up region remains unchanged.  Therefore, we will still be in
lateral equilibrium in the star region.  We will denote the density
where $T = T_\mathrm{hi}$ as $\rho_\mathrm{fuel}$.

Creating the model involves specifying $T_\star$, $T_\mathrm{hi}$,
$T_\mathrm{lo}$, $\rho_\mathrm{int}$, $H_\star$,
$\delta_\mathrm{atm}$, ${X_k}_\star$, and ${X_k}_\mathrm{atm}$.  We
then integrate outwards from the base of the ramp region ($y =
H_\star$), enforcing the discrete form of hydrostatic equilibrium,
Eq.~\ref{eq:hse}.  Integrating upwards, we would find $p_i$ and
$\rho_i$ using a Newton solve together either with the temperature
specified, $T_i = T(p_i, \rho_i, \{X_k\}_i)$ (in the isothermal, ramp,
and buffer layers) or constant entropy, $s_i = s(p_i, \rho_i,
\{X_k\}_i)$, in the fuel layer.  In each case, using the equation of
state.  This follows the procedures described in \citet{ppm-hse}.  We
use the constant temperature for all $y < H_\star +
3\delta_\mathrm{atm}$.  Above this, we switch to isentropic until the
temperature drops to a floor value, $T_\mathrm{lo}$, at which point we
again keep the temperature constant.  The integration of the
atmosphere continues until the density falls to a low density cutoff,
$\rho_\mathrm{cutoff}$.  The material above this height is taken to
have constant density and temperature.

\begin{deluxetable}{lcc}
\tablecaption{\label{table:params} Initial model parameters.}
\tablehead{\colhead{parameter} & \colhead{cool} & \colhead{hot}}
\startdata
$T_\star$       & \multicolumn{2}{c}{$5\times 10^7$~K} \\
$T_\mathrm{hi}$ & $2\times 10^8$~K & $1.2\times 10^9$~K \\
$T_\mathrm{lo}$ & \multicolumn{2}{c}{$10^7$~K} \\
$\rho_\mathrm{int}$ & \multicolumn{2}{c}{$3.43\times 10^6~\gcc$} \\
$\rho_\mathrm{fuel}$\tablenotemark{a} & $2.21\times 10^6~\gcc$ & $1.27\times 10^6~\gcc$ \\
$\rho_\mathrm{cutoff}$ & \multicolumn{2}{c}{$10^{-4}~\gcc$} \\
$g$             & \multicolumn{2}{c}{$-1.5\times 10^{14}~\mathrm{cm~s^{-2}}$} \\
$H_\star$       & \multicolumn{2}{c}{5000~cm} \\
$\delta_\mathrm{atm}$ & \multicolumn{2}{c}{50~cm} \\
$X_\star(\isotm{Fe}{56})\tablenotemark{b}$ & \multicolumn{2}{c}{1.0} \\
$X_\mathrm{atm}(\isotm{He}{4})\tablenotemark{b}$ & \multicolumn{2}{c}{1.0} \\
\enddata
%
\tablenotetext{a}{This is not an input parameter, but instead is
  computed during integration.  We list it here for reference.}
%
\tablenotetext{b}{All other species are taken as 0.}
\end{deluxetable}

The choice of factors in front of $\delta_\mathrm{atm}$ were designed
to make sure the peak $T$ is attained at the desired density 
of the burning layer.

We create two models, a ``cool'' model representing the atmosphere
ahead of the flame and a ``hot'' (or perturbed) model that will
initiate the flame by driving the burning.  The parameters we use for
the model generation are listed in Table~\ref{table:params}.

We blend the hot and cold models laterally to produce the perturbation
needed to initiate a localized flame, with the hot model at the
origin of the axisymmetric geometry.  The blending is done as:
\begin{align}
p(x,y) &= f(x) p_\mathrm{hot}(y) + [1-f(x)] p_\mathrm{cool}(y) \\
\rho(x,y) &= f(x) \rho_\mathrm{hot}(y) + [1-f(x)] \rho_\mathrm{cool}(y) \\
X_k(x,y) &= f(x) {X_k}_\mathrm{hot}(y) + [1-f(x)] {X_k}_\mathrm{cool}(y)
\end{align}
with
\begin{equation}
f(x) = \begin{cases}
     1 & x < x_\mathrm{pert} \\
   1 - \frac{x - x_\mathrm{pert}}{\delta_\mathrm{blend}} & x_\mathrm{pert} \le x \le x_\mathrm{pert} + \delta_\mathrm{blend} \\
     0 & x > x_\mathrm{pert} + \delta_\mathrm{blend}
\end{cases}
\end{equation}
Since the equation of hydrostatic equilibrium is linear and our
blending is a linear combination of two models in hydrostatic
equilibrum, the blended model is also in equilibrium initially.  We choose
$x_\mathrm{pert} = 1.024\times 10^4$~cm and $\delta_\mathrm{blend} = 2048$~cm.  Once
the blended model is constructed, we compute $T(y)$ and $(\rho e)(y)$
from the equation of state.  \MarginPar{show}

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{initial_model_paper}
\caption{\label{fig:initial_models} Our ``cool'' (solid) and ``hot''
  (dashed) initial models, showing both the density and temperature.}
\end{figure}

%======================================================================
% Results
%======================================================================
\section{Simulations and Results}\label{Sec:results}

Things to try to make the model feasible:
\begin{itemize}
\item Start the initial perturbation location near the Rossby radius
\item Hotter ``hot'' model
\item Aprox13
\item Higher diffusion cutoff lessens the timestep restriction
\end{itemize}

All simulations were run with in a domain $8.192\times
10^4~\mathrm{cm} \times 2.048\times 10^4~\mathrm{cm}$.  With our
default refinement, this gives a resolution of $20~\mathrm{cm}$ on the
finest level.  As a comparison, the resolution in our detonation
models was $98~\mathrm{cm}$ and in our multidimensional convection
studies was $6~\mathrm{cm}$.

We specify the rotation period, $P_\mathrm{rot}$, which is related to
the angular rotation rate as $\Omega = 2\pi / P_\mathrm{rot}$.  Our
default rotation period is $0.0005~\mathrm{s}$ (2000 times per
second).  This is quite fast, but picked to make the Rossby length
small enough to fit in the spatial scales we can model.  With this
rotation rate, the Rossby length is
\begin{equation}
L_R = \frac{\sqrt{g H_0}}{\Omega} \sim 3\times 10^4~\mathrm{cm}
\end{equation}
\MarginPar{we should make the width of the hot region closer to the
  Rossby length} using a scale height $H_0 = 10^3~\mathrm{cm}$.  This
is less than half our domain width.

We note that the entire simulation framework for these calculations is 
freely available in the \castro\ github repository.

\subsection{Resolution}

Our default resolution puts $\sim 50$ zones vertically in the ``cool''
model atmosphere height.  To understand the effects of resolution, we
also perform a run with one additional level, with a $2\times$ jump
refining just the region where the density is $10^4~\gcc < \rho \le
4\times 10^4~\gcc$.  On this level, the grid resolution is
$10~\mathrm{cm}$.

\subsection{Rotation Rate}

We also try a rotation period of $P_\mathrm{rot} \sim 10^{-4}~\mathrm{s}$.

\subsection{Reaction Network}

We use aprox13 instead of the small network to understand the
differences.  Energy comes out faster.

\subsection{Boosting}

To meet reasonable timescales, we boosted the microphysical inputs to
the flame to result in a faster flame.  Here we look at the effect of
this choice.

\subsection{Sponge}

\section{Discussion}

Full hydrodynamics models of flame spreading in XRB are possible with certain approximations.

CPU cost

Accurate simulations require
\begin{enumerate}
\item Resolving the scale height of the atmosphere

\item Thin transition between underlying neutron star and atmosphere, to ensure that the peak temperature
hits right at the proper base density
\end{enumerate}

The main approximations we made were
\begin{enumerate}
\item A higher than normal rotation rate
\item A boosted flame
\item Simplified reaction network
\item A 2.5D model for the flow
\end{enumerate}
The first two of these were needed to reduce the spatial and temporal
scales needed to model, making the simulations feasible.  Using the
simulation framework developed here, these both can be relaxed in the
future, at the cost of more computer time.  The same goes for 2.5D
vs.\ full 3D---the only difference is computer time, and our future
calculations will explore the 3D evolution and compare to these.  In particular,
in 3D we will be able to explore shear instabilities at the flame front.
Larger networks are a straightforward change, and already supported in
\castro\ using the \pynucastro\ framework~\citep{pynucastro} and JINA
ReacLib rate database~\citep{reaclib}.

These studies complement the

Other future work includes mixed H/He bursts.  This will require a
different reaction network, and perhaps different (more relaxed)
resolution requirements.

On the algorithmic front, we are exploring a change of our
hydrodynamics and reactive coupling to eliminate splitting error from
the traditional Strang splitting used in \castro\ and other
hydrodynamics codes.  One advantage of this is that the new algorithm
would be more amenable to offloading onto GPUs, a development task
that is underway.  Additionally, we will move to fourth-order accurate
methods and incorporate well-balanced reconstruction to better capture
the burning.

MHD, radiation transport

\acknowledgements \castro\ is open-source and  freely available at
\url{http://github.com/AMReX-Astro/Castro}.  The problem setups used
here are available in the git repo as {\tt flame} and {\tt
  flame\_wave}.  The work at Stony Brook was supported by DOE/Office
of Nuclear Physics grant DE-FG02-87ER40317 and NSF award AST-1211563.
Work at LANL was done under the auspices of the National Nuclear
Security Administration of the U.S. Department of Energy at Los Alamos
National Laboratory under Contract No.\ DE-AC52-06NA25396.
This research used resources of the National Energy Research
Scientific Computing Center, a DOE Office of Science User Facility
supported by the Office of Science of the U.~S.\ Department of Energy
under Contract No.\ DE-AC02-05CH11231.

\facilities{NERSC}

\software{ADS (http://adsabs.harvard.edu/),
          AMReX (\url{https://github.com/AMReX-Codes/amrex}),
          Castro \citep{castro},
          GCC (\url{https://gcc.gnu.org/}),
          linux (\url{https://www.kernel.org/}),
          matplotlib (\citealt{Hunter:2007}, \url{http://matplotlib.org/}),
          NumPy \citep{numpy,numpy2},
          python (\url{https://www.python.org/}),
          valgrind \citep{valgrind},
          yt \citep{yt}}

\appendix

\section{Diffusion Tests}
\label{app:diffusion}

Show diffusion test problem and convergence.

%======================================================================
% References
%======================================================================

\bibliographystyle{aasjournal}
\bibliography{ws}

\end{document}

### 2019-04-08 xrb/model.py

@@ -1,0 +1,63 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+z_max = 10
+z = np.linspace(0, z_max, 200)
+
+H_star = 4
+delta_atm = 1
+z_isentropic = 3
+
+T_star = 1.0
+T_hi = 2.0
+T_lo = 1.2
+
+T = T_star*np.ones_like(z)
+
+idx = z > H_star
+T[idx] = (T_hi - T_star)/delta_atm*(z[idx] - H_star)  + T_star
+
+plt.plot([H_star, H_star], [0.9, 2.1], color="C0", ls=":")
+plt.text(H_star/2, 1.8, "star layer", color="C0", horizontalalignment="center", fontsize="small")
+
+plt.text(-0.1, T_star, r"$T_\star$", horizontalalignment="right")
+
+idx = z > H_star + delta_atm
+T[idx] = (T_lo - T_hi)/z_isentropic*(z[idx] - (H_star + delta_atm)) + T_hi
+
+plt.plot([H_star + delta_atm, H_star + delta_atm], [0.9, 2.1], color="C0", ls=":")
+plt.text(H_star+delta_atm/2, 1.8, "ramp\nlayer", color="C0", horizontalalignment="center", fontsize="small")
+plt.plot([0, H_star + delta_atm], [T_hi, T_hi], color="0.5", ls=":")
+plt.text(-0.1, T_hi, r"$T_\mathrm{hi}$", horizontalalignment="right")
+
+T[z > H_star + delta_atm + z_isentropic] = T_lo
+
+plt.plot([H_star + delta_atm + z_isentropic, H_star + delta_atm + z_isentropic], [0.9, 2.1], color="C0", ls=":")
+plt.text(H_star+delta_atm + z_isentropic/2, 1.8, "fuel\nlayer", color="C0", horizontalalignment="center", fontsize="small")
+
+plt.plot([0, H_star + delta_atm + z_isentropic], [T_lo, T_lo], color="0.5", ls=":")
+plt.text(-0.1, T_lo, r"$T_\mathrm{lo}$", horizontalalignment="right")
+
+
+plt.text(H_star+delta_atm + z_isentropic + 0.5*(z_max - (H_star+delta_atm+z_isentropic)),
+         1.8, "outer\nbuffer", color="C0", horizontalalignment="center", fontsize="small")
+plt.plot(z, T, color="C1")
+plt.axis("off")
+
[…]

### 2019-04-08 xrb/plots/initial_model/initial_model.py

@@ -1,0 +1,79 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+def initial_model():
+
+    cool_model = "flame_wave_new_cool.out"
+    hot_model = "flame_wave_new_hot.out"
+
+    cool_data = np.loadtxt(cool_model)
+    hot_data = np.loadtxt(hot_model)
+
+
+    # make the plots
+    r_min = 0
+    r_max = 10000.0
+
+    r_min = 4500
+    r_max = 7000
+
+    #-------------------------------------------------------------------------
+    # density and temperature plots
+    #-------------------------------------------------------------------------
+    dens_min = 1.e2
+    dens_max = 5.e8
+
+    sp = plt.subplot(111)
+
+    sp.set_yscale('log')
+
+
+    # density
+    plt.plot(cool_data[:,0], cool_data[:,1], color="C0")
+    plt.plot(hot_data[:,0], hot_data[:,1], color="C0", ls="--")
+
+    plt.xlabel(r"y (cm)")
+    plt.ylabel(r"density (g cm$^{-3}$)", color="C0")
+
+    plt.xlim(r_min, r_max)
+    plt.ylim(dens_min, dens_max)
+
+    ax = plt.gca()
+
+    # the offset text is the 1.e8 that appears under the axis labels when
+    # doing scientific notation
+    #ax.tick_params(labeltop='off')
+    #ax.tick_params(labelbottom='off')
+    ax.xaxis.offsetText.set_visible(False)
+
+    # temperature
+    sp2 = plt.twinx()
+    sp2.set_yscale('log')
+
+    # original
+    plt.plot(cool_data[:,0], cool_data[:,2], color="C1")
+    plt.plot(hot_data[:,0], hot_data[:,2], color="C1", ls="--")
+
+    sp2.yaxis.tick_right()
+    plt.axis(labelcolor="C1")
+    plt.ylabel(r"temperature (K)", color="C1")
+
+    plt.xlabel(r"y (cm)")
+
+    plt.xlim(r_min, r_max)
+    plt.ylim(8.e6, 2.e9)
+
+
+    #-------------------------------------------------------------------------
+    #plt.subplots_adjust(hspace=0.1)
+
+    f = plt.gcf()
+    f.set_size_inches(6.0,6.0)
+
+    f.tight_layout()
+
+    plt.savefig("initial_model_paper.pdf", bbox_inches='tight')
+
+if __name__== "__main__":
+    initial_model()
+

### 2019-04-08 xrb/plots/max_density/dens_temp.py

@@ -1,0 +1,27 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import glob
+
+for filename in glob.glob("*.out"):
+    data = np.loadtxt(filename)
+
+    plt.plot(data[:,0], data[:,2]/data[0,2], label="{}".format(filename.split(".")[0]))
+
+
+plt.ylim(1.0, 1.4)
+plt.legend(fontsize="small", frameon=False)
+
+plt.savefig("max_dens.png")
+
+plt.clf()
+
+for filename in glob.glob("*.out"):
+    data = np.loadtxt(filename)
+    if len(data[0,:]) == 5:
+        plt.plot(data[:,0], data[:,4]/data[0,4], label="{}".format(filename.split(".")[0]))
+
+
+plt.ylim(0.6, 1.4)
+plt.legend(fontsize="small", frameon=False)
+
+plt.savefig("max_temp.png")

### 2019-04-10 MAESTROeX_I/algorithm.tex
Add figures in results section

@@
 We average by binning all the cells associated with each radii and taking the average.
-Then we interpolate this data onto a one-dimensional radial array.  Previously, MAESTRO only allowed for radial arrays with a constant $\Delta r$ (typically equal to $\Delta x/5$.
+Then we interpolate this data onto a one-dimensional radial array.  Previously, MAESTRO only allowed for radial arrays with a constant $\Delta r$ (typically equal to $\Delta x/5$).
 Here we present a new option to retain an irregularly-spaced radial array to eliminate mapping errors back onto the Cartesian grid state.
@@
 The buoyancy forcing term in the momentum equation contains $\rho-\rho_0$.
-With the original scheme, interpolation errors in computing $\rho_0$ by averaging would cause artifical acceleration in the velocity field due to the interpolation error from the Cartesian grid to and from the radial base state.
+With the original scheme, interpolation errors in computing $\rho_0$ by averaging would cause artificial acceleration in the velocity field due to the interpolation error from the Cartesian grid to and from the radial base state.
 By retaining the radial base state as an irregularly spaced array, the effects due to interpolation error are completely eliminated.

### 2019-04-10 MAESTROeX_I/results.tex
Add figures in results section

@@
 \subsection{AMR Performance}
-Show 3D reacting bubble rise performance.
-Show 3-level spherical performance (stats won't be as good but explain why: refining a larger percentage of grid)
-Images of grid configuration.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=2.75in]{./figs/reacting_bubble_amr} 
+$\qquad$
+\includegraphics[width=2.75in]{./figs/wdconvect_amr_3grid}
+\caption{\label{fig:amr_grids} Initial grid structures with two levels of refinement. 
+         The red, green, and black lines represent grids of increasing refinement. 
+         (Left) Profile of $T - \bar{T}$ for a hot bubble in a white dwarf environment.
+         (Right) Region of star where $\rho\ge 10^5 \text{ g cm}^{-3}$ in a full white dwarf star simulation. }
+\end{center}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=2.75in]{./figs/reacting_bubble_result} 
+$\qquad$
+\includegraphics[width=2.75in]{./figs/reacting_bubble_amr_result}
+\caption{\label{fig:bubble_results} Time-lapse cross-section of a hot bubble in a white dwarf environment at 
+         $t = 0$, 1.25, and 2.5 s for (left) single-level simulation, and 
+         (right) adaptive simulation at the same effective resolution.}
+\end{center}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2019-04-13 MAESTROeX_I/results.tex
AMR wdconvect plot

@@
 
+Reacting bubble simulation at effective resolution of $128^2 \times 1024$ run on Cori Haswell with 48 processors took 33.54 s per time step (averaged over 10 time steps) to run on single-level grid and only 3.70 s on adaptive grid with two levels of refinement.
+This 89\% decrease in runtime is mostly due to the fact that only 6.25\% of the cells (1,048,576 out of $128^2 \times 1024$ cells) are refined at the the finest level for the AMR run.
+
 \subsection{White Dwarf Convection}
@@
 
-Show 3-level wdconvect
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=3.5in]{./figs/wdconvect_amr_Tmax} 
+\caption{\label{fig:wdconvect_amr_Tmax} Peak temperature, $T_{\text{peak}}$, in a white dwarf from $t=0$ to $350$ s  
+for grids with effective resolution of $256^3$. We can see that the adaptive grids with two levels of refinement 
+give very similar solution trends compared to the single-level grid.}
+\end{center}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Full white dwarf star simulation at effective resolution of $512^3$ on Cori KNL with 512 processors and 16 threads took 26.01 s per time step (averaged over 10 time steps) using the original temporal scheme, and 22.66 s using the new scheme.
+On an adaptive grid with two levels of refinement, the simulation took 34.69 s because the refinement region was set all the way to the edge of the star.

### 2019-04-13 xrb/paper.tex
update authors
add ack

@@
 \author[0000-0001-8401-030X]{Michael Zingale}
+\affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
+             Stony Brook, NY 11794-3800}
+
+\author[0000-0002-1530-781X]{Alice Harpole}
 \affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
@@
   flame\_wave}.  The work at Stony Brook was supported by DOE/Office
-of Nuclear Physics grant DE-FG02-87ER40317 and NSF award AST-1211563.
-Work at LANL was done under the auspices of the National Nuclear
-Security Administration of the U.S. Department of Energy at Los Alamos
-National Laboratory under Contract No.\ DE-AC52-06NA25396.
-This research used resources of the National Energy Research
-Scientific Computing Center, a DOE Office of Science User Facility
-supported by the Office of Science of the U.~S.\ Department of Energy
-under Contract No.\ DE-AC02-05CH11231.
+of Nuclear Physics grant DE-FG02-87ER40317.  This research used
+resources of the National Energy Research Scientific Computing Center,
+a DOE Office of Science User Facility supported by the Office of
+Science of the U.~S.\ Department of Energy under Contract
+No.\ DE-AC02-05CH11231.  This research used resources of the Oak Ridge
+Leadership Computing Facility at the Oak Ridge National Laboratory,
+which is supported by the Office of Science of the U.S. Department of
+Energy under Contract No. DE-AC05-00OR22725, awarded through the DOE
+INCITE program.
 
-\facilities{NERSC}
+\facilities{NERSC, OLCF}
 

### 2019-04-19 MAESTROeX_I/results.tex
add results for long wdconvect runs until ignition

@@
 
-Reacting bubble simulation at effective resolution of $128^2 \times 1024$ run on Cori Haswell with 48 processors took 33.54 s per time step (averaged over 10 time steps) to run on single-level grid and only 3.70 s on adaptive grid with two levels of refinement.
+Reacting bubble simulation at effective resolution of $128^2 \times 1024$ run on Cori Haswell with 48 processors took approximately 33.5 s per time step (averaged over 10 time steps) to run on single-level grid using either the original or new temporal algorithm, and only 3.7 s on adaptive grid with two levels of refinement.
 This 89\% decrease in runtime is mostly due to the fact that only 6.25\% of the cells (1,048,576 out of $128^2 \times 1024$ cells) are refined at the the finest level for the AMR run.
@@
 \subsection{White Dwarf Convection}
+
 white dwarf convection runs with 3 algorithms (original, new temporal, new temporal + irregular base state)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=2.75in]{./figs/wdconvect_128_maxT} 
+\includegraphics[width=2.75in]{./figs/wdconvect_128_maxMach}
+\caption{\label{fig:wdconvect_128_maxval} (left) Peak temperature, $T_{\text{peak}}$, and (right) peak Mach number 
+         in a white dwarf until time of ignition at resolution of $128^3$ for three different MAESTROeX algorithms.}
+\end{center}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \caption{\label{fig:wdconvect_amr_Tmax} Peak temperature, $T_{\text{peak}}$, in a white dwarf from $t=0$ to $350$ s 
-for grids with effective resolution of $256^3$. We can see that the adaptive grids with two levels of refinement 
-give very similar solution trends compared to the single-level grid.}
+         for grids with effective resolution of $512^3$. We can see that the adaptive grids with two levels of 
+         refinement give very similar solution trends compared to the single-level grid.}
 \end{center}

### 2019-04-21 MAESTROeX_I/equations.tex

@@
 The resulting model is a series of evolution equations for mass, momentum, and energy, subject
-to this constraint:
+to these constraints:
 \begin{eqnarray}
@@
 Mathematically, equations (\ref{eq:momentum})-(\ref{eq:enthalpy}) must still be closed by an equation of state.  
-This is done by taking the Langrangian derivative of the equation of state for pressure as a function of the density, composition, and enthalpy, requiring that the pressure is a prescribed function of altitutde and time based on the hydrostatic equilibrium condition.
+This is done by taking the Langrangian derivative of the equation of state for pressure as a function of the density, composition, and enthalpy, requiring that the pressure is a prescribed function of altitude and time based on the hydrostatic equilibrium condition.
 Details are given in Papers I and II.
-After substuting the equations for mass and energy into the Lagrangian derivative, the result
+After substituting the equations for mass and energy into the Lagrangian derivative, the result
 is a divergence constraint on the velocity field (\ref{eq:U divergence}),

### 2019-04-30 MAESTROeX_I/equations.tex
paper updates

--- 2019-03-13 MAESTROeX_I/equations.tex
@@
-\section{Governing Equations}
+\section{Governing Equations}\label{sec:equations}
 Low Mach number models for reacting flow were originally derived using asymptotic analysis
@@
 The resulting model is a series of evolution equations for mass, momentum, and energy, subject
-to this constraint:
+to to an additional constraint on velocity discussed below:
 \begin{eqnarray}
@@
 \frac{\partial(\rho X_k)}{\partial t} &=& -\nabla\cdot(\rho X_k\Ub) + \rho\omegadot_k,\label{eq:species}\\
-\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc,\label{eq:enthalpy}
+\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc.\label{eq:enthalpy}
 \end{eqnarray}
-\begin{equation}
-\nabla\cdot(\beta_0\Ub) = \beta_0\left(S - \frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial t}\right).\label{eq:U divergence}
-\end{equation}
 Here $\rho$, $\Ub$, and $h$ are the mass density,
@@
 directions of the full state and $r$ to represent the radial coordinate direction for 
-the base state).  We also define a base state density, $\rho_0 = \rho_0(r,t)$, 
-which is in hydrostatic equilibrium with $p_0$, i.e., 
+the base state).
+We also define a base state density, $\rho_0 = \rho_0(r,t)$, that represents the lateral average (see Section \ref{Sec:Spatial}) of $\rho$ and is in hydrostatic equilibrium with $p_0$, i.e.,
 $\nabla p_0 = -\rho_0 g\eb_r$, where $g=g(r,t)$ is
@@
 
-Mathematically, equations (\ref{eq:momentum})-(\ref{eq:enthalpy}) must still be closed by an equation of state.  
-This is done by taking the Langrangian derivative of the equation of state for pressure as a function of the density, composition, and enthalpy, requiring that the pressure is a prescribed function of altitutde and time based on the hydrostatic equilibrium condition.
-Details are given in Papers I and II.
-After substuting the equations for mass and energy into the Lagrangian derivative, the result
+Mathematically, equations (\ref{eq:momentum})-(\ref{eq:enthalpy}) must still be closed by an EOS.  
+This is done by taking the Langrangian derivative of the EOS for pressure as a function of the density, composition, and enthalpy, requiring that the pressure is a prescribed function of altitude and time based on the hydrostatic equilibrium condition.
+Details are given in Papers I and II, but to summarize, after substituting the equations for mass and energy into the Lagrangian derivative, the result
 is a divergence constraint on the velocity field (\ref{eq:U divergence}),
-where $\beta_0$ is a density-like variable that carries background stratification, defined as
+\begin{equation}
+\nabla\cdot(\beta_0\Ub) = \beta_0\left(S - \frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial t}\right).\label{eq:U divergence}
+\end{equation}
+Here $\beta_0$ is a density-like variable that carries background stratification, defined as
 \begin{equation}
@@
 \right|_{p,X_k}$ is the specific heat at constant pressure.
+To summarize, we model equations (\ref{eq:momentum})-(\ref{eq:enthalpy})
+subject to (\ref{eq:U divergence}) and the hydrostatic equilibrium condition.

### 2019-04-30 MAESTROeX_I/intro.tex
paper updates

@@
 The resulting model can be numerically integrated with much larger time steps than a compressible model.
+
 Low Mach number models have been developed for a variety of contexts including combustion \citep{day2000numerical}, terrestrial atmospheric modeling \citep{duarte2015low},
 and elastic solids \citep{abbate2017all}.
-
-In recent years, a number of alternative approaches to modelling low Mach number flows have been developed. These include semi-implicit all-Mach number solvers, where the Euler equations
-are split into an acoustic part and an advective part \citep{Kwatra2009,Degond2009,Cordier2012,Haack2012,Happenhofer2013,Chalons2016,Padioleau2019}. The fast acoustic waves are then
-solved using implicit time integration, while the slow material waves are solved explicitly. Another approach is to use preconditioned all-Mach number solvers
-\citep{Miczek2014,Barsukow2016}, where the numerical flux is multiplied by a preconditioning matrix. This reduces the stiffness of the system at low Mach numbers, while retaining the
-correct scaling behavior. In the reduced speed of sound technique (RSST) and related methods, the speed of sound is artificially reduced by including a suitable scaling factor in the
-continuity equation, reducing the restriction on the size of the timestep \citep{Rempel2005,Hotta2012,Wang2015,Takeyama2017,Iijima2018}.
+In recent years, a number of alternative approaches to modeling low Mach number astrophysical flows have been developed.
+These include semi-implicit all-Mach number solvers, where the Euler equations are split into an acoustic part and an advective part \citep{Kwatra2009,Degond2009,Cordier2012,Haack2012,Happenhofer2013,Chalons2016,Padioleau2019}.
+The fast acoustic waves are then solved using implicit time integration, while the slow material waves are solved explicitly.
+Another approach is to use preconditioned all-Mach number solvers \citep{Miczek2014,Barsukow2016}, where the numerical flux is multiplied by a preconditioning matrix.
+This reduces the stiffness of the system at low Mach numbers, while retaining the correct scaling behavior.
+In the reduced speed of sound technique (RSST) and related methods, the speed of sound is artificially reduced by including a suitable scaling factor in the continuity equation, reducing the restriction on the size of the timestep \citep{Rempel2005,Hotta2012,Wang2015,Takeyama2017,Iijima2018}.
 The MUSIC code uses fully implicit time integration for the compressible Euler equations, which therefore allows for arbitrarily large timesteps \citep{Viallet2011,Viallet2015,Goffrey2016}.
@@
 Previously, we developed the low Mach number astrophysical solver, MAESTRO.
-The low Mach number model in MAESTRO is unique in that it is specifically designed for astrophysical settings with significant atmospheric stratification.
+Unlike several of the references above, MAESTRO is not an all-Mach solver, but is suitable for flows where the Mach number is small ($\sim 0.1$ or smaller).  
+Furthermore, the low Mach number model in MAESTRO is specifically designed for astrophysical settings with significant atmospheric stratification.
+This includes full spherical stars, as well as planar simulations of dynamics within localized regions of a star.
+MAESTRO uses an explicit Godunov approach for advection, the stiff ODE VODE package \citep{vode} for reactions, and multigrid-based linear solvers for the pressure-projection steps, and is thus time-step limited by an advective CFL constraint based on the fluid velocity, not the sound speed.
 Central to the algorithm is a stratified background (or base) state density and pressure held in hydrostatic equilibrium, and also vary as a function of altitude and time.
@@
 This code was developed in the pure-Fortran90 FBoxLib framework (see the FBoxLib repository in \cite{AMReX}) for block-structured AMR applications.
-MAESTRO is suitable for full spherical stars, as well as planar simulations of dynamics within localized regions of a star.
 
@@
 state, but this treatment is more amenable to higher-order multiphysics coupling strategies
-based on method-of-lines integration.  Second, we now provide additional spatial mapping
+based on method-of-lines integration.
+In particular, schemes based on spectral deferred corrections \citep{dutt2000spectral} have been used to generate high-order temporal integrators for problems of reactive flow \citep{dutt2000spectral} and low Mach number combustion \citep{pazner2016high,nonaka2018conservative}
+Second, we now provide additional spatial mapping
 procedures that reduces, and in some cases eliminates, mapping error from the one-dimensional
@@
 MAESTROeX uses MPI+OpenMP parallelism and scales well to over 10,000 MPI processes, with each MPI process supporting tens of threads.
-The resulting code is publicly available on GitHub at {\url https://github.com/AMReX-Astro/MAESTROeX} and uses the microphysics
-libraries at {\url https://github.com/starkiller-astro/Microphysics} and the AMReX software library for block-structured adaptive mesh refinement
-{\url https://github.com/AMReX-Codes/amrex}.
+The resulting code is publicly available on GitHub ({\tt https://github.com/AMReX-Astro/MAESTROeX}) and uses the Starkiller-Astro microphysics libraries ({\tt https://github.com/starkiller-astro/Microphysics}) and the AMReX software library for block-structured adaptive mesh refinement ({\tt https://github.com/AMReX-Codes/amrex}).
+
+The rest of this paper is organized as follows.
+In Section \ref{sec:equations} we review our model for stratified low Mach number astrophysical flow.
+In Section \ref{eq:algorithm} we present our numerical algorithm in detail.
+In Section \ref{sec:results} we validate our new approach and examine the performance of our algorithm on full spherical star problems used in previous scientific investigations.
+Finally, we conclude in Section \ref{sec:conclusions}.

### 2019-04-30 MAESTROeX_I/paper.tex
more updates
paper updates

@@
 \hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}
+
+\usepackage{url}
+
+\usepackage{mathtools}
 
@@
 Our model equations allow for long-time integration for highly subsonic flows compared to compressible approaches.
-The code leverages the new AMReX framework for block-structured adaptive mesh refinement calculations, and also provides additional algorithmic features from the original MAESTRO code.
-We present a new temporal integration option that is much simpler than our previous approach while retaining the same order order of accuracy.
-We also provide a more accurate spatial discretization for mapping the base state onto the full state Cartesian grid.
+The code leverages the new AMReX framework for block-structured adaptive mesh refinement calculations allowing for improved scalability of the predecessor code.
+We present a new, simplified temporal integration scheme that retains the same order of accuracy as our previous code.
+We also explore the use of alternative spatial mapping of the one-dimensional base state onto the full Cartesian grid.
 Using our previous studies on the convective phase of single-degenerate progenitor models of Type Ia supernovae, we characterize the performance of the code and validate the new algorithmic features.

### 2019-04-30 MAESTROeX_I/algorithm.tex
paper updates
more updates

@@
-\section{Numerical Algorithm}
+\section{Numerical Algorithm}\label{eq:algorithm}
 
@@
 constraint.  We then construct $\uadvone$ from $\uadvonedag$
-using a MAC projection.
+using a MAC projection (see Appendix \ref{Sec:Projection} for details).
 We note that $\uadvone$ satisfies the divergence constraint
@@
 the updated value $S^{\nph}$ rather than $S^{\nph,\star}$.
+The $\dagger$ superscript refers to the
+fact that the predicted velocity field does not satisfy the divergence
+constraint.  We then construct $\uadvtwo$ from $\uadvtwodag$
+using a MAC projection (see Appendix \ref{Sec:Projection} for details).
 We note that $\uadvtwo$ satisfies the divergence constraint
@@
 
-Finally, we use an approximate nodal projection to define $\Ub^{n+1}$
+Finally, we use an approximate projection to define $\Ub^{n+1}$
 from $\Ub^{n+1,\dagger},$  such that $\Ub^{n+1}$ approximately
@@
 \end{equation}
-As part of the projection we also define the new-time perturbational pressure,
-$\pi^\nph.$  This projection necessarily differs from the MAC projection used in
+As part of the projection we also define the new-time perturbational pressure, $\pi^\nph$.
+This pressure $\pi$ is a nodal quantity, and thus this projection step is often referred to as a nodal projection.
+This projection necessarily differs from the MAC projection used in
 {\bf Step 3} and {\bf Step 7} because the velocities in those steps are defined
 on edges and $\Ub^{n+1}$ is defined at cell centers, requiring different divergence
-and gradient operators.  Details of the approximate projection are given in Paper III.
+and gradient operators.
+
+Refer to Appendix \ref{Sec:Projection} for more details about the projection steps.
 

### 2019-04-30 MAESTROeX_I/appendix.tex
more ferences
more updates

@@
 To enforce the divergence constraint in {\bf Step 3, 7}, and {\bf 11}, we use a projection method analgous to the methods originally developed for incompressible flow \citep{almgren1998conservative,bell1989second}.
+The basic idea is to decompose the velocity field into a divergence-free component and a curl-free (gradient of a scalar field) component by solving a variable-coefficient Poisson equation for the scalar field.
+The details for the MAC projection in {\bf Step 3} and {\bf Step 7} are given in Appendix B of Paper III.  The details of the nodal projection in {\bf Step 11} are given in Section 3.2 of Paper III.
+We note that in the nodal projection, the gradient of the scalar field is used to update the perturbational pressure, $\pi$.
 
@@
 \end{equation}
+To solve for $\Ubt$ we use a projection method, which involves the solution of a variable-coefficient Poisson solver to extract the curl-free component of the unprojected velocity.
+Note that MAESTRO contains alternate low Mach number formulations that conserve total energy in stratified systems, with minimal changes to the code (see Appendix A of \cite{subChandra_II} for details).
 To find $w_0$ we integrate in 1D using the procedure in Appendix B of Paper V.
-
-Note that in this approach, we estimated the time-derivative of the pressure in part by examining how laterally averaged $\rho'=\rho-\rho_0$ changed over time (quantified by $\eta_\rho = \overline{\rho'\Ub\cdot\eb_r}$.
+Note that in this approach, we estimated the time-derivative of the pressure in part by examining how laterally averaged $\rho'=\rho-\rho_0$ changed over time (quantified by $\eta_\rho = \overline{\rho'\Ub\cdot\eb_r}$).
 After evolving the species, we compute this term in the same way as reported in Paper V for use in computing $w_0$.
-
-\MarginPar{mention alt energy fix for projection?  also mention changes in $w_0$ computation for irregular base state}
+\MarginPar{mention changes in $w_0$ computation for irregular base state. Add graphic of temperature falling due to incorrect splitting?}
 

### 2019-04-30 MAESTROeX_I/algorithm.tex
more updates

@@
 \begin{equation}
-\Ub = \Ubt(\xb,t) + w_0(r,t)\eb_r,
+\Ub = \Ubt(\xb,t) + w_0(r,t)\eb_r.
 \end{equation}
 We used $w_0$ to provide an estimate of the base state density evolution over a time step.
-Our new temporal integration scheme does not use this splitting, and is much simpler than the scheme from Paper V since we evolve the velocity directly rather than the perturbational and base state velocity separately.
-This eliminates
-Additionally, the new scheme uses a simpler predictor-corrector approach to the base state density and pressure that no longer requires the complex algorithm from Paper V, yet still retains the same overall second-order accuracy.
+This resulted in some unnecessary complications to the temporal integration scheme.
+Our new temporal integration scheme does not use this splitting, and is much simpler than the scheme from Paper V since we evolve the velocity directly rather than a more complex evolution equation for the perturbational velocity.
+Additionally, the new scheme uses a simpler predictor-corrector approach to the base state density and pressure that no longer requires evolution equations and the associated discretizations for the base state, greatly simplifying the algorithm while retaining the same overall second-order accuracy in time.
 
-We now describe the new temporal integration scheme, which has been greatly simplified from Paper V without formal loss of order of accuracy.
-Note that this temporal integration scheme is valid for the original base state mapping (with constant base state grid spacing), or the new irregularly spaced base state mapping.
-
+We now describe the new temporal integration scheme, which we note can be used for the original base state mapping (with constant base state grid spacing) as well as the new irregularly spaced base state mapping.
 At the beginning of each time step we have the cell-centered state,
@@
 \item {\bf Average}$[\phi]\rightarrow[\overline\phi]$ computes the lateral average of a quantity over a layer at constant radius $r$, as described above in Section \ref{Sec:Spatial}.
-\item {\bf Enforce HSE}$[\rho_0]\rightarrow[p_0]$ computes the base state pressure, $p_0$, from a base state density, $\rho_0$ by integrating the hydrostatic equilibrium condition in 1D (see equation (A10) in Paper V).
+\item {\bf Enforce HSE}$[\rho_0]\rightarrow[p_0]$ computes the base state pressure, $p_0$, from a base state density, $\rho_0$ by integrating the hydrostatic equilibrium condition in one dimension; see equation (A10) in Paper V.
 The base state pressure remains equal to a constant value from the location of a prescribed cutoff density outward for the entire simulation.
@@
 
-To initialize the simulation we use the same procedure described in Paper III.
-At the beginning of each simulation, we define $(\Ub,\rho X_k,\rho h)$.
-We set initial values for $\Ub, \rho X_k$, and $\rho h$ and perform a sequence of projections 
-(to ensure the velocity field satisfies the divergence constraint) 
-followed by a small number of steps of the temporal advancement scheme to iteratively 
-find initial values for $\pi^{n-\myhalf}$ and $S^0$ and $S^1$ for use in the first time step.
-The new temporal integration strategy in MAESTROeX contains the following steps:
+The temporal integration scheme contains the following steps:
 \begin{description}
@@
 
+To initialize the simulation we use the same procedure described in Paper III.
+At the beginning of each simulation, we define $(\Ub,\rho X_k,\rho h)$.
+We set initial values for $\Ub, \rho X_k$, and $\rho h$ and perform a sequence of projections 
+(to ensure the velocity field satisfies the divergence constraint) 
+followed by a small number of steps of the temporal advancement scheme to iteratively 
+find initial values for $\pi^{n-\myhalf}$ and $S^0$ and $S^1$ for use in the first time step.
+
 Our approach to adaptive mesh refinement is algorithmically the same as the treatment described in Section 5 of Paper V; we refer the reader there for details.

### 2019-04-30 MAESTROeX_I/equations.tex
more updates

@@
 We also define a base state density, $\rho_0 = \rho_0(r,t)$, that represents the lateral average (see Section \ref{Sec:Spatial}) of $\rho$ and is in hydrostatic equilibrium with $p_0$, i.e.,
-$\nabla p_0 = -\rho_0 g\eb_r$, where $g=g(r,t)$ is
-the magnitude of the gravitational acceleration and $\eb_r$ is the unit vector in the
-outward radial direction.
+\begin{equation}
+\nabla p_0 = -\rho_0 g\eb_r, \label{eq:HSE}
+\end{equation}
+where $g=g(r,t)$ is the magnitude of the gravitational acceleration and $\eb_r$ is the unit vector in the outward radial direction.
 
@@
 \right|_{p,X_k}$ is the specific heat at constant pressure.
-To summarize, we model equations (\ref{eq:momentum})-(\ref{eq:enthalpy})
-subject to (\ref{eq:U divergence}) and the hydrostatic equilibrium condition.
+
+To summarize, we model evolution equations for momentum, mass, and energy, (\ref{eq:momentum})-(\ref{eq:enthalpy}) subject to a divergence constraint on the velocity (\ref{eq:U divergence}) and the hydrostatic equilibrium condition (\ref{eq:HSE}).

### 2019-04-30 MAESTROeX_I/intro.tex
more updates

@@
 Unlike several of the references above, MAESTRO is not an all-Mach solver, but is suitable for flows where the Mach number is small ($\sim 0.1$ or smaller).  
-Furthermore, the low Mach number model in MAESTRO is specifically designed for astrophysical settings with significant atmospheric stratification.
+Furthermore, the low Mach number model in MAESTRO is specifically designed for, but not limited to, astrophysical settings with significant atmospheric stratification.
 This includes full spherical stars, as well as planar simulations of dynamics within localized regions of a star.
@@
 In Section \ref{sec:equations} we review our model for stratified low Mach number astrophysical flow.
-In Section \ref{eq:algorithm} we present our numerical algorithm in detail.
+In Section \ref{eq:algorithm} we present our numerical algorithm in detail, highlighting the new temporal integration scheme as well as spatial base state mapping options.
 In Section \ref{sec:results} we validate our new approach and examine the performance of our algorithm on full spherical star problems used in previous scientific investigations.

### 2019-04-30 MAESTROeX_I/paper.tex
more updates

@@
 We present MAESTROeX, a massively parallel solver for low Mach number astrophysical flows.
-Our model equations allow for long-time integration for highly subsonic flows compared to compressible approaches.
-The code leverages the new AMReX framework for block-structured adaptive mesh refinement calculations allowing for improved scalability of the predecessor code.
-We present a new, simplified temporal integration scheme that retains the same order of accuracy as our previous code.
+The underlying low Mach number equation set allows for efficient, long-time integration for highly subsonic flows compared to compressible approaches.
+MAESTROeX is suitable for modeling full spherical stars as well as well as planar simulations of dynamics within localized regions of a star,
+and can robustly handle several orders of magnitude of density and pressure stratification.
+We present a new, greatly simplified temporal integration scheme that retains the same order of accuracy as our previous approaches.
 We also explore the use of alternative spatial mapping of the one-dimensional base state onto the full Cartesian grid.
-Using our previous studies on the convective phase of single-degenerate progenitor models of Type Ia supernovae, we characterize the performance of the code and validate the new algorithmic features.
+The code leverages the new AMReX software framework for block-structured adaptive mesh refinement applications, allowing for scalability to large fractions of leadership-class machines.
+Using our previous studies on the convective phase of single-degenerate progenitor models of Type Ia supernovae as a guide, we characterize the performance of the code and validate the new algorithmic features.
 \end{abstract}

### 2019-04-30 MAESTROeX_I/intro.tex
reworked intro

@@
 In a low Mach number approach, sound waves are eliminated from the governing equations while retaining compressibilitiy effects due to, e.g., nuclear energy release, compositional changes, and thermal diffusion.
-The resulting model can be numerically integrated with much larger time steps than a compressible model.
+When the Mach number (the ratio of the characteristic fluid velocity over the characteristic sound speed; Ma $= U/c$) is small,
+the resulting system can be numerically integrated with much larger time steps than a compressible model.
+Specifically, the time step increase is at least a factor of $\sim 1/{\rm Ma}$ larger.
+Each time step is more computationally expensive due to the presence of additional linear solves,
+but for many problems of interest the overall gains in efficiency can easily be an order of magnitude or more.
 
@@
 and elastic solids \citep{abbate2017all}.
-In recent years, a number of alternative approaches to modeling low Mach number astrophysical flows have been developed.
+For astrophysical applications, a number of approaches to modeling low Mach number flows have been developed in recent years.
 These include semi-implicit all-Mach number solvers, where the Euler equations are split into an acoustic part and an advective part \citep{Kwatra2009,Degond2009,Cordier2012,Haack2012,Happenhofer2013,Chalons2016,Padioleau2019}.
@@
 Previously, we developed the low Mach number astrophysical solver, MAESTRO.
+MAESTRO is a block-structured, Cartesian grid finite-volume, adaptive mesh refinement (AMR) code that has been successfully used for many years for a number of applications, detailed below.
 Unlike several of the references above, MAESTRO is not an all-Mach solver, but is suitable for flows where the Mach number is small ($\sim 0.1$ or smaller).  
@@
 This includes full spherical stars, as well as planar simulations of dynamics within localized regions of a star.
-MAESTRO uses an explicit Godunov approach for advection, the stiff ODE VODE package \citep{vode} for reactions, and multigrid-based linear solvers for the pressure-projection steps, and is thus time-step limited by an advective CFL constraint based on the fluid velocity, not the sound speed.
-Central to the algorithm is a stratified background (or base) state density and pressure held in hydrostatic equilibrium, and also vary as a function of altitude and time.
-MAESTRO is a structured-grid, finite-volume code that utilizes adaptive mesh refinement (AMR) to refine grids locally in space.
-This code was developed in the pure-Fortran90 FBoxLib framework (see the FBoxLib repository in \cite{AMReX}) for block-structured AMR applications.
+MAESTRO uses an explicit Godunov approach for advection, a stiff ODE package (VODE) \citep{vode} for reactions, and multigrid-based linear solvers for the pressure-projection steps, and is thus time-step limited by an advective CFL constraint based on the fluid velocity, not the sound speed.
+The time-advancement strategy uses Strang splitting to couple the thermodynamic variables, and a second-order projection method to integrate the velocity equations subject to a divergence constraint.
+Central to the algorithm is a time-varying, one-dimensional stratified background (or base) state density and pressure held in hydrostatic equilibrium that couple to the full Cartesian grid solution.
+The original MAESTRO code was developed in the pure-Fortran90 FBoxLib software framework (see the FBoxLib repository in \cite{AMReX}) for block-structured AMR applications.
 
@@
 \begin{itemize}
-\item In Paper I \citep{MAESTRO_I}, we derive the low Mach number equation set from the fully compressible equations.
+\item In Paper I \citep{MAESTRO_I}, we derive the low Mach number equation set for stratified environments from the fully compressible equations.
 \item In Paper II \citep{MAESTRO_II}, we incorporate the effects of atmospheric expansion through the use of a time-dependent background state.
@@
 
-Since then, there have been many scientific investigations using MAESTRO, which include additional algorithmic enhancements.  Topics include:
+Since then, there have been many scientific investigations using MAESTRO, which have included additional algorithmic enhancements.  Topics include:
 \begin{itemize}
@@
 
-We also present new algorithmic methodology that improves upon Paper V in a number of ways.
+In this paper, we present new algorithmic methodology that improves upon Paper V in a number of ways.
 First, the overall temporal algorithm has been greatly simplified without compromising accuracy.
-The key design decision was to replace the predicted evolution of the base state with
-a predictor-corrector approach.  Not only does this greatly simplify the dynamics of the base
+The key design decisions were to eliminate the splitting of the velocity into average and perturbational components, 
+and also to replace the hydrodynamic evolution of the base state with a predictor-corrector approach.
+Not only does this greatly simplify the dynamics of the base
 state, but this treatment is more amenable to higher-order multiphysics coupling strategies
 based on method-of-lines integration.  
-In particular, schemes based on spectral deferred corrections \citep{dutt2000spectral} have been used to generate high-order temporal integrators for problems of reactive flow \citep{dutt2000spectral} and low Mach number combustion \citep{pazner2016high,nonaka2018conservative}
-Second, we now provide additional spatial mapping
-procedures that reduces, and in some cases eliminates, mapping error from the one-dimensional
-base state to the Cartesian grid state for spherical problems.
+In particular, schemes based on spectral deferred corrections \citep{dutt2000spectral} have been used to generate high-order temporal integrators for problems of reactive flow \citep{dutt2000spectral} and low Mach number combustion \citep{pazner2016high,nonaka2018conservative}.
+Second, we explore the effects of alternative spatial mapping routines for coupling the base state and the Cartesian grid state in spherical problems.
 Finally, the code has been completely ported to a new software framework suitable for modern manycore supercomputers.
-The resulting MAESTROeX is implemented in the C++/F90 AMReX public software library \cite{AMReX}.
+We examine the performance of our new MAESTROeX implementation in the new C++/F90 AMReX public software library \cite{AMReX}.
 MAESTROeX uses MPI+OpenMP parallelism and scales well to over 10,000 MPI processes, with each MPI process supporting tens of threads.
-The resulting code is publicly available on GitHub ({\tt https://github.com/AMReX-Astro/MAESTROeX}) and uses the Starkiller-Astro microphysics libraries ({\tt https://github.com/starkiller-astro/Microphysics}) and the AMReX software library for block-structured adaptive mesh refinement ({\tt https://github.com/AMReX-Codes/amrex}).
+The resulting code is publicly available on GitHub ({\tt https://github.com/AMReX-Astro/MAESTROeX}),
+uses the Starkiller-Astro microphysics libraries ({\tt https://github.com/starkiller-astro/Microphysics}),
+and the AMReX software library for block-structured adaptive mesh refinement ({\tt https://github.com/AMReX-Codes/amrex}).
 
@@
 In Section \ref{sec:results} we validate our new approach and examine the performance of our algorithm on full spherical star problems used in previous scientific investigations.
-Finally, we conclude in Section \ref{sec:conclusions}.
+We conclude in Section \ref{sec:conclusions}.

### 2019-04-30 MAESTROeX_I/equations.tex
pass at equations

@@
 in astrophysical environments using adaptive algorithms in space and time \citep{Bell:2004}.
-In Papers I-III, we derived a model and algorithm suitable for stratified astrophysical flow.
+In Papers I-III, we extended this work by deriving a model and algorithm suitable for stratified astrophysical flow.
 In our model, we take the standard equations of reacting, compressible flow, and recast the equation
@@
 The resulting model is a series of evolution equations for mass, momentum, and energy, subject
-to to an additional constraint on velocity discussed below:
+to to an additional constraint on velocity.  The evolution equations are
 \begin{eqnarray}
@@
 Here $\Hnuc$ is the nuclear energy generation rate per unit mass.
-The pressure is decomposed into a hydrostatic base state
+The total pressure is decomposed into a one-dimensional hydrostatic base state
  pressure, $p_0 = p_0(r,t)$, and a dynamic pressure, $\pi = \pi(\xb,t)$, such that 
-$|\pi|/p_0 = \mathcal{O}(M^2)$ (we use $\xb$ to represent the Cartesian coordinate 
+$p = p_0 + \pi$ and $|\pi|/p_0 = \mathcal{O}({\rm Ma}^2)$ (we use $\xb$ to represent the Cartesian coordinate 
 directions of the full state and $r$ to represent the radial coordinate direction for 
 the base state).
-We also define a base state density, $\rho_0 = \rho_0(r,t)$, that represents the lateral average (see Section \ref{Sec:Spatial}) of $\rho$ and is in hydrostatic equilibrium with $p_0$, i.e.,
+We also define a one-dimensional base state density, $\rho_0 = \rho_0(r,t)$, that represents the lateral average (see Section \ref{Sec:Spatial}) of $\rho$ and is in hydrostatic equilibrium with $p_0$, i.e.,
 \begin{equation}
@@
 where $g=g(r,t)$ is the magnitude of the gravitational acceleration and $\eb_r$ is the unit vector in the outward radial direction. 
+Thermal diffusion is not discussed in this paper, but we have previously described the modifications to the original algorithm required 
+for implicit thermal diffusion in \cite{XRB_I}; inclusion of these effects in the new algorithm presented here is straightforward.
 
-Mathematically, equations (\ref{eq:momentum})-(\ref{eq:enthalpy}) must still be closed by an EOS.  
-This is done by taking the Langrangian derivative of the EOS for pressure as a function of the density, composition, and enthalpy, requiring that the pressure is a prescribed function of altitude and time based on the hydrostatic equilibrium condition.
-Details are given in Papers I and II, but to summarize, after substituting the equations for mass and energy into the Lagrangian derivative, the result
-is a divergence constraint on the velocity field (\ref{eq:U divergence}),
+Mathematically, equations (\ref{eq:momentum})-(\ref{eq:enthalpy}) must still be closed by the EOS.  
+This is done by taking the Langrangian derivative of the EOS for pressure as a function of the thermodynamic variables, 
+substituting in the equations of motion for mass and energy,
+and requiring that the pressure is a prescribed function of altitude and time based on the hydrostatic equilibrium condition.
+See Papers I and II for details of this derivation.
+The resulting equation is a divergence constraint on the velocity field,
 \begin{equation}
@@
 where $\gammaonebar$ is the lateral average of $\Gamma_1 = d(\log p)/d(\log\rho)$ at constant entropy.
-The expansion term, $S$, incorporates local compressibility effects due to heat release from reactions, compositional changes, and external sources,
+The expansion term, $S$, incorporates local compressibility effects due to compositional changes and heat release from reactions,
 \begin{equation}

### 2019-05-01 MAESTROeX_I/algorithm.tex
started updates to section 3

@@
 \MarginPar{needs work}
-MAESTRO contains variables defined on Cartesian grids 
-one-dimensional base state stored on radial arrays.
+The spatial discretization and adaptive mesh refinement methodology remains unchanged from Paper V.
+We now summarize some of the key points here before describing the new temporal integrator in the next section.
+We recommend the reader review Section 3 of Paper V for further details.
 
- associated with square (in 2D) or cubic (in 3D)
-
-(description of base state mapping in planar, and original spherical, and new spherical)
-
-Mapping from the base state to the full state
-
-Mapping a Cartesian variable to a one-dimensional radial array uses the lateral averaging operator.
-
+We shall refer to local atmospheric flows in two and three dimensions as problems in ``planar'' geometry, and full-star flows
+in three dimensions as problems in ``spherical'' geometry. The solution in both cases consists of the Cartesian grid solution
+and the one-dimensional base state solution.
+Figure \ref{Fig:BaseGrid} illustrates the relationship between the base state and the Cartesian grid state for both planar and spherical geometries in the presence of spatially adaptive grids.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[tb]
+\centering
+\includegraphics[height=2.0in]{./figs/base_grid} \hspace{0.5in}
+\includegraphics[height=2.0in]{./figs/base_spherical}
+\caption{\label{Fig:BaseGrid}  
+(Left) For multi-level problems in planar geometry, we force a direct alignment
+between the radial array cell centers and the Cartesian grid cell centers by 
+allowing the radial base state spacing to change with space and time.
+(Right) For multi-level problems in spherical geometry, since there is no direct alignment
+between the radial array cell centers and the Cartesian grid cell centers, we choose to fix
+the radial base state spacing across levels.}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 One of the key numerical modules is the ``lateral average'', which computes the average over a 
@@
 We average by binning all the cells associated with each radii and taking the average.
-Then we interpolate this data onto a one-dimensional radial array.  Previously, MAESTRO only allowed for radial arrays with a constant $\Delta r$ (typically equal to $\Delta x/5$).
+Then we interpolate this data onto a one-dimensional radial array.  
+Previously, MAESTRO only allowed for radial arrays with a constant $\Delta r$ (typically equal to $\Delta x/5$).
 Here we present a new option to retain an irregularly-spaced radial array to eliminate mapping errors back onto the Cartesian grid state.
@@
 By retaining the radial base state as an irregularly spaced array, the effects due to interpolation error are completely eliminated.
+\MarginPar{Doreen you had a great figure of this mapping in one of your talks}
 
 \subsection{Temporal Integration Scheme}\label{Sec:Temporal Integration Scheme}
+We now describe the new temporal integration scheme, which we note can be used for the original base state mapping (with constant base state grid spacing) as well as the new irregularly spaced base state mapping.
 Previously we adopted an approach where we split the velocity into a base state component, $w_0(r,t)$, 
@@
 
-We now describe the new temporal integration scheme, which we note can be used for the original base state mapping (with constant base state grid spacing) as well as the new irregularly spaced base state mapping.
 At the beginning of each time step we have the cell-centered state,
@@
 The inputs are the species, enthalpy, and base state pressure, and the outpus are the species, enthalpy, reaction rates, and nuclear energy generation rate.
-See Papers III and V for details.
+See Paper III for details.
 \end{itemize}

### 2019-05-01 MAESTROeX_I/algorithm.tex
projection, conclusions

@@
 in Appendix A of \cite{XRB_III}.
-
-In this step, using equation (\ref{eq:momentum}), 
-we compute time-centered edge velocities, $\uadvonedag$, using
-$\Ub^n$.  The $\dagger$ superscript refers to the
-fact that the predicted velocity field does not satisfy the divergence
-constraint.  We then construct $\uadvone$ from $\uadvonedag$
-using a MAC projection (see Appendix \ref{Sec:Projection} for details).
-We note that $\uadvone$ satisfies the divergence constraint
+Here we use equation (\ref{eq:momentum}) to compute time-centered edge velocities, $\uadvonedag$.
+The $\dagger$ superscript refers to the fact that the predicted velocity field does not satisfy the divergence constraint,
 \begin{equation}
@@
 \end{equation}
-In practice, for all the projections in the algorithm, we solve a split dynamics
-system for a base state velocity, and a perturbational velocity, so that we can properly
-enforce boundary conditions at the edge of the star.
-Refer to Appendix \ref{Sec:Projection} for details.
+ We project $\uadvonedag$ onto the space of velocities that satisfy the constraint to obtain $\uadvone$.
+Each projection step in the algorithm involves the solution of a variable-coefficient Poisson solve using multigrid.
+The details of this ``MAC'' projection are provided in Appendix \ref{Sec:Projection}.
 
@@
 the updated value $S^{\nph}$ rather than $S^{\nph,\star}$.
-The $\dagger$ superscript refers to the
-fact that the predicted velocity field does not satisfy the divergence
-constraint.  We then construct $\uadvtwo$ from $\uadvtwodag$
-using a MAC projection (see Appendix \ref{Sec:Projection} for details).
-We note that $\uadvtwo$ satisfies the divergence constraint
+The $\dagger$ superscript refers to the fact that the predicted velocity field does not satisfy the divergence constraint,
 \begin{equation}
@@
 \end{equation}
-where
+with
 \begin{equation}
 \beta_0^{\nph} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2},
+\gammaonebar^{\nph} = \frac{ \gammaonebar^n +  \gammaonebar^{n+1,\star} }{2}.
 \qquad
-\gammaonebar^{\nph} = \frac{ \gammaonebar^n +  \gammaonebar^{n+1,\star} }{2}.
 \end{equation}
+we project $\uadvtwodag$ onto the space of velocities that satisfy the constraint to obtain $\uadvtwo$ using a MAC projection (see Appendix \ref{Sec:Projection}).
 
@@
 Again, the $\dagger$ superscript refers
-to the fact that the updated velocity does not satisfy the divergence
-constraint.
-
-Finally, we use an approximate projection to define $\Ub^{n+1}$
-from $\Ub^{n+1,\dagger},$  such that $\Ub^{n+1}$ approximately
-satisfies the divergence constraint,
+to the fact that the updated velocity does not satisfy the divergence constraint,
 \begin{equation}
@@
 \end{equation}
-As part of the projection we also define the new-time perturbational pressure, $\pi^\nph$.
-This pressure $\pi$ is a nodal quantity, and thus this projection step is often referred to as a nodal projection.
+We use an approximate projection to project $\Ub^{n+1,\dagger}$ onto the space of velocities that satisfy the constraint to obtain $\Ub^{n+1}$ using a ``nodal'' projection.
 This projection necessarily differs from the MAC projection used in
@@
 and gradient operators.
-
-Refer to Appendix \ref{Sec:Projection} for more details about the projection steps.
+Furthermore, as part of the nodal projection, we also define a nodal new-time perturbational pressure, $\pi^\nph$.
+Refer to Appendix \ref{Sec:Projection} for more details.
 

### 2019-05-01 MAESTROeX_I/conclusions.tex
paper updates
projection, conclusions

@@
-\section{Conclusions and Future Work}
-Our new implementation of the MAESTRO code in the AMReX software framework
-
+\section{Conclusions and Future Work}\label{sec:conclusions}
+We have developed a new temporal integrator and spatial mapping options into our low Mach number solver, MAESTROeX.
+The new AMReX-enabled code scales well on large fractions of supercomputers with multicore architectures.
+Future software enhancements will include GPU implementation.
+In particular, the AMReX-based companion code, the compressible CASTRO code \citep{CASTRO}, has recently ported hydrodynamics and reactions to GPUs \citep{CASTRO_GPU}.
+We plan to leverage the newly implemented mechanisms for offloading compute kernels to GPUs inside of the AMReX software library itself.ment library itself.
 Our future scientific investigations include convection in massive rotating stars, \citep{heger2000presupernova}, the convective Urca process in white dwarfs \citep{willcox2016type}, solar physics \citep{wood2018self}, and magnetohydrodynamics \citep{wood2015three,wood2011sun}.
-
-Our future algorithmic enhancements include more accurate and higher-order multiphysics coupling strategies byased on spectral deferred corrections \cite{dutt2000spectral,bourlioux2003high}.
-This framework has been successfully used in terrestrial combustion \cite{pazner2016high,nonaka2018conservative}
-
-Our software enhancements will include GPU implementation.
-\MarginPar{Forward cite CASTRO paper; add a few sentences on GPU strategy}
+Our future algorithmic enhancements include more accurate and higher-order multiphysics coupling strategies byased on spectral deferred corrections \citep{dutt2000spectral,bourlioux2003high}.
+This framework has been successfully used in terrestrial combustion \citep{pazner2016high,nonaka2018conservative}
 

### 2019-05-03 MAESTROeX_I/results.tex
paper updates
Updated wdconvect AMR results when refined tighter around the convective region

@@
-\section{Performance and Validation}
+\section{Performance and Validation}\label{sec:results}
 
@@
 Reacting bubble simulation at effective resolution of $128^2 \times 1024$ run on Cori Haswell with 48 processors took approximately 33.5 s per time step (averaged over 10 time steps) to run on single-level grid using either the original or new temporal algorithm, and only 3.7 s on adaptive grid with two levels of refinement.
-This 89\% decrease in runtime is mostly due to the fact that only 6.25\% of the cells (1,048,576 out of $128^2 \times 1024$ cells) are refined at the the finest level for the AMR run.
+This 89\% decrease in runtime is mostly due to the fact that only 6.25\% of the cells (1,048,576 out of $128^2 \times 1024$ cells) are refined at the finest level for the AMR run.
 
@@
 \begin{center}
-\includegraphics[width=3.5in]{./figs/wdconvect_amr_Tmax} 
-\caption{\label{fig:wdconvect_amr_Tmax} Peak temperature, $T_{\text{peak}}$, in a white dwarf from $t=0$ to $350$ s 
+\includegraphics[width=4.0in]{./figs/wdconvect_amr_Tmax} 
+\caption{\label{fig:wdconvect_amr_Tmax} Peak temperature, $T_{\text{peak}}$, in a white dwarf from $t=0$ to $400$ s 
          for grids with effective resolution of $512^3$. We can see that the adaptive grids with two levels of 
@@
 Full white dwarf star simulation at effective resolution of $512^3$ on Cori KNL with 512 processors and 16 threads took 26.01 s per time step (averaged over 10 time steps) using the original temporal scheme, and 22.66 s using the new scheme.
-On an adaptive grid with two levels of refinement, the simulation took 34.69 s because the refinement region was set all the way to the edge of the star.
+On an adaptive grid with two levels of refinement, 8.9\% of the cells (1,493,504 out of $256^3$ cells) are refined at the first level and 5.3\% (7,176,192 out of $512^3$ cells) at the second.
+The simulation took 11.55 s and 12.48 s using the orginial and new temporal schemes, respectively.

### 2019-05-13 xrb/paper.tex
some tweaks

@@
 %\documentclass[12pt, preprint]{aastex}
-\documentclass[twocolumn,times,tighten]{aastex62}
+\documentclass[preprint,times,tighten]{aastex62}
 
@@
 \setlength{\marginparwidth}{0.5in}
-\newcommand{\MarginPar}[1]{\marginpar{\vskip-\baselineskip\raggedright\tiny\sffamily\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}
+
+\newcommand{\MarginPar}[1]{
+    \marginpar{\vskip-\baselineskip%
+               \raggedright%
+               \tiny\sffamily%
+               {\color{red}\hrule%
+               \smallskip%
+               #1\par%
+               \smallskip%
+               \hrule}}%
+}
+
+\newcommand{\AssignTo}[1]{
+    \marginpar{\vskip-\baselineskip%
+               \raggedright%
+               \tiny\sffamily%
+               {\color{blue}\hrule%
+               \smallskip%
+               #1\par%
+               \smallskip%
+               \hrule}}%
+}
 
@@
 
+\AssignTo{Mike, Alice}
+
 XRB intro
@@
 detonations may only really be applicable to superbursts (where carbon
-is the reactant).
+is the reactant). \MarginPar{ref}
 
@@
 Challenges: timescale for flame propagation, lengthscales (flame
-width, Rossby scale), reaction network size.\MarginPar{from talk}
+width, Rossby scale), reaction network size.\MarginPar{summarize AstroNum 2018 paper}
 
@@
 parabolic method (PPM) with characteristic tracing for solving the
-hydrodynamics.  Reactions are incorporated via Strang splitting,
-giving a method that is overall second-order accurate in space and
-time.
+hydrodynamics~\cite{ppm,millercolella:2002}, generalized to an
+arbitrary equation of state~\cite{zingalekatz}.  Reactions are
+incorporated via Strang splitting~\cite{strang:1968}, giving a method that is overall
+second-order accurate in space and time.
 
@@
 system.  Finally, $\kth$ is the thermal conductivity and $T$ is the
-temperature.
+temperature.  Diffusion is treated explicitly in time.
 

### 2019-05-13 xrb/paper.tex
more work

@@
 
-We work in 2D axisymmetric coordinates, but we advective a third
-component of velocity, coming out of the simulation plane, that
-participates in the Coriolis force (sometimes described as a 2.5D
-simulation).  We will take the \castro\ $x$-coordinate to be the
-cylindrical $r$ coordinate with corresponding velocity $u$, the
-\castro\ $y$-coordinate to be the cylindrical $z$ coordinate with
-corresponding velocity $v$, and the \castro\ $z$-coordinate to be the
-cylindrical $\theta$ coordinate, with corresponding velocity $w$.  A
-righthanded coordinate system has positive $w$ pointing out of the
-page.  We will take $\Omegab = \Omega_0 \hat{\bf y}$ for the angular
-rotation rate, and $\gb = -g \hat{\bf y}$ for the gravitational
-acceleration, with $g$ constant.  With these choices, the Coriolis
-force is:
+Since the neutron star rotates, we work in a corotating frame, with
+angular velocity $\Omegab$.  Further, for the two-dimensional
+simulations presented here, we work in axisymmetric coordinates,
+but we advective a third component of velocity, coming out of the
+simulation plane, that participates in the Coriolis force (sometimes
+described as a 2.5D simulation).  We will take the
+\castro\ $x$-coordinate to be the cylindrical radial coordinate with
+corresponding velocity $u$, the \castro\ $y$-coordinate to be the
+cylindrical vertical coordinate with corresponding velocity $v$, and the
+\castro\ $z$-coordinate to be the cylindrical azimuthal coordinate,
+with corresponding velocity $w$.  A righthanded coordinate system has
+positive $w$ pointing out of the page.  We will take $\Omegab =
+\Omega_0 \hat{\bf y}$ for the angular rotation rate, and $\gb = -g
+\hat{\bf y}$ for the gravitational acceleration, with $g$ constant.
+With these choices, the Coriolis force is:
 \begin{equation}
@@
 expected to greatly affect the dynamics.  Carrying the Coriolis force
-allows us to capture the geostrophic balance\MarginPar{ref} that sets up via lateral
-hydrostatic equilibrium.
+allows us to capture the geostrophic balance\MarginPar{ref} that sets
+up via lateral hydrostatic equilibrium.
 
@@
 ideal gas), photons, and degenerate/relativistic electrons, as
-described in \cite{timmes_swesty:2000}.  We use a simple reaction
-network consisting of 4 nuclei, \isot{He}{4}, \isot{C}{12},
-\isot{O}{16}, \isot{Fe}{56}.  Only the first 3 are evolved, with the
-triple-$\alpha$ rate and $\isotm{C}{12}(\alpha, \gamma)\isotm{O}{16}$.
-Screening of the reactions is also included.  The network is provided
-in our open-source microphysics
-routines\footnote{\url{https://github.com/starkiller-astro/Microphysics}}.
+described in \cite{timmes_swesty:2000}.  For reactions, we use a 13
+isotope alpha chain, derived from the {\tt aprox13}
+network~\citep{timmes_aprox13}.  For one of our runs, we use the
+smaller 7-isotope network described in \citet{iso7}.  We integrate the
+network using the VODE integration package~\citep{vode}, and out
+implementation is provided in the StarKiller Microphysics
+source~\citep{starkiller}.
 

### 2019-05-13 xrb/paper.tex
more updates

@@
 upper boundary at low resolution).  In addition to the base grid, we
-use up to 3 refinement levels, the first two a factor of 4 finer than
-the previous and the last a factor of 2 finer than the previous.
-\castro\ subcycles in time, so the finer grids are evolved with a
-finer timestep than the coarse grids.  Occasionally the timestep
-chosen at the start of a cycle will violate the CFL condition during
-the advance of the finer grids.  In this case, we restart the finer
-grid evolution with a smaller timestep, subcycling within the larger
-timestep hierarchy.  The finest grid in most of our simulations (base
-+ 2 refined levels) corresponds to $4096\times 1024$ zones.
+use up to 3 refinement levels, the first one a factor of 4 finer than
+the previous and the remaining each a factor of 2 finer than the
+previous.  \castro\ subcycles in time, so the finer grids are evolved
+with a finer timestep than the coarse grids.  Occasionally the
+timestep chosen at the start of a cycle will violate the CFL condition
+during the advance of the finer grids.  In this case, we restart the
+finer grid evolution with a smaller timestep, subcycling within the
+larger timestep hierarchy.  We use a CFL number of 0.8 for our
+simulations.  \MarginPar{true for all?} The base grid for our standard
+simulations is $768\times 192$ zones and the corresponding finest grid
+when 3 refinement levels are added would be $12288\times 3072$ zones.
+Our standard domain is $1.2288\times 10^5~\mathrm{cm} \times
+3.072\times 10^4~\mathrm{cm}$, so we have $10~\mathrm{cm}$ resolution
+on the finest grid.  We only refine the fuel layer in the left half of
+the domain at the highest resolution (and only down to densities of
+$2.5\times 10^4~\gcc$), since this is where we expect the flame to
+propagate.  At the start of the simulation, $3.4\%$ of the domain is
+at the finest resolution.  This increases to $7.4\%$ by the end of the
+simulation, because of the increase in the scale height of the atmosphere
+behind the flame.
 
@@
 \plottwo{flame}{speed}
-\caption{\label{fig:flame} Our ``standard'' flame.}
+\caption{\label{fig:flame} Our ``standard'' flame. {\color{red} this should be redone for the boosted flame}}
 \end{figure*}

### 2019-05-13 xrb/paper.tex
some more updates

@@
 
-\cite{cavecchi:2013} \MarginPar{and others}
+\cite{cavecchi:2013,art-2015-cavecchi-etal,art-2016-cavecchi-etal} \MarginPar{need a summary}
 
@@
 
-sponge\MarginPar{write}
+sponge\AssignTo{Mike will write}
 
@@
 
-\input{flame_table_1d.tex}
-
-To understand the time and length scales involved in flame
-propagation, we do a 1D simulation of a laminar flame using our
-microphysics.  Figure~\ref{fig:flame} shows the flame properties for a
-He flame using the standard conductivities and our simple reaction
-network.  This flame had a density of $2\times 10^6~\gcc$ and
-temperature of $5\times 10^7$~K.  We note that this flame speed is
-about $2\times 10^4~\cms$, the flame width is about 20~cm, and it
-takes about 0.1~s to setting into a sustained flame.  Note this speed
-is quite small compared to the speed estimated in
-\citet{spitkovsky2002}.  We expected a multidimensional flame to
-accelerate due to hydrodynamics interactions (wrinkling, turbulence
-interactions, directed flows feeding fuel into the flame, ...).
-
-We measure the flame width as:
-\begin{equation}
-\lambda_f \equiv \frac{\Delta T}{\max\{|\nabla T|\}}
-\end{equation}
-
-\begin{figure*}[t]
-\plottwo{flame}{speed}
-\caption{\label{fig:flame} Our ``standard'' flame. {\color{red} this should be redone for the boosted flame}}
-\end{figure*}
-
-To make the simulation feasible, we artifically speed up the flame.
-Dimensional analysis shows that the flame speed scales as:\MarginPar{ref}
+The speed and thickness of a laminar helium flame is determined by the
+energy generation rate and conductivity and scale roughly as \MarginPar{need a ref}
 \begin{equation}
@@
 \end{equation}
-To boost the flame, while keeping the thickness the same, we can
-multiply both the burning rate and the conductivity by the same
-factor.  We choose 10 for each, to give a flame 10$\times$ faster.
+At the densities we consider in this simulation, the pure He flame
+speed is quite slow and would require a prohibitively long integration
+time to see significant evolution of the burning.  We therefore
+consider boosted flames in this first paper, to accelerate the burning
+and allow us to understand the qualitative effects of laterally
+propagating flames.  To boost the flame, while keeping the thickness
+the same, we can multiply both the burning rate and the conductivity
+by the same factor.  For our standard calculations, we choose 10 for
+each, to give a flame 10$\times$ faster.  We call this the ``10/10''
+flame.  We will also do a simulation with the reactions and
+conductivity both boosted by 5, the ``5/5'' flame.
 
+To understand the time and length scales involved in flame
+propagation, we do a 1D simulation of a laminar flame using our
+microphysics.  Figure~\ref{fig:flame} shows the flame properties for a \AssignTo{Kiran and Mike will remake with the 10/10 boosted flame}
+He flame using the standard conductivities and our simple reaction
+network.  This flame had a density of $2\times 10^6~\gcc$ and \MarginPar{double check}
+temperature of $5\times 10^7$~K.  We note that this flame speed is
+about $2\times 10^4~\cms$, the flame width is about 20~cm, and it
+takes about 0.1~s to settle into a sustained flame.  Note this speed
+is quite small compared to the speed estimated in
+\citet{spitkovsky2002}.  We expected a multidimensional flame to
+accelerate due to hydrodynamics interactions (wrinkling, turbulence
+interactions, directed flows feeding fuel into the flame, ...).
+
+\begin{figure*}[t]
+\plottwo{flame}{speed}
+\caption{\label{fig:flame} Our ``standard'' flame. {\color{red} this should be redone for the boosted flame}}
+\end{figure*}
+
+We measure the flame width as:
+\begin{equation}
+\lambda_f \equiv \frac{\Delta T}{\max\{|\nabla T|\}}
+\end{equation}
 Experience with modeling resolved flames suggests that we need a
-spatial resolution, $\Delta x$ of $\lambda_f/\Delta x \sim 5$.  These
+spatial resolution, $\Delta x$ of $\lambda_f/\Delta x \sim 5$~\citep{SNld}.  These
 conditions represent the bottom of the He layer.  As the density
@@
 the rest of the atmosphere.
+
+Flame properties are show in Table XX \MarginPar{we'll redo the table with 10/10 and 5/5/ and unboosted flame}
+
+\input{flame_table_1d.tex}
 
@@
 \centering
-\includegraphics[width=\linewidth]{model}
+\plotone{model}
 \caption{\label{fig:sketch} A schematic of our model setup.}
@@
 blending is a linear combination of two models in hydrostatic
-equilibrum, the blended model is also in equilibrium initially.  We choose
+equilibrum, the blended model is also in vertical equilibrium initially.  We choose
 $x_\mathrm{pert} = 1.024\times 10^4$~cm and $\delta_\mathrm{blend} = 2048$~cm.  Once
@@
 \centering
-\includegraphics[width=\linewidth]{initial_model_paper}
+\plotone{initial_model_paper}
 \caption{\label{fig:initial_models} Our ``cool'' (solid) and ``hot''
@@
 Energy under Contract No. DE-AC05-00OR22725, awarded through the DOE
-INCITE program.
+INCITE program.  This research has made use of NASA's Astrophysics
+Data System Bibliographic Services.
 
@@
 
-\software{ADS (http://adsabs.harvard.edu/),
-          AMReX (\url{https://github.com/AMReX-Codes/amrex}),
+\software{AMReX (\url{https://github.com/AMReX-Codes/amrex}),
           Castro \citep{castro},
@@
           valgrind \citep{valgrind},
+          VODE \citep{vode},
           yt \citep{yt}}

### 2019-05-14 xrb/paper.tex
fix some equations
add some section assignments

@@
 \newcommand{\castro}{{\sf Castro}}
+\newcommand{\amrex}{{\sf AMReX}}
 \newcommand{\pynucastro}{{\sf pynucastro}}
@@
 arbitrary equation of state~\cite{zingalekatz}.  Reactions are
-incorporated via Strang splitting~\cite{strang:1968}, giving a method that is overall
-second-order accurate in space and time.
+incorporated via Strang splitting~\cite{strang:1968}, giving a method
+that is overall second-order accurate in space and time.
+\castro\ uses the \amrex\ adaptive mesh refinement
+library~\cite{amrex_joss} to manage a hierarchy of grids at different
+resolutions.
 
@@
 -2\rho \Omegab \times \Ub =
-   -2\rho \left ( \Omegab w \hat{x} - \Omegab u \hat{z} \right )
+   -2\rho \left ( \Omega_0 w \hat{x} - \Omega_0 u \hat{z} \right )
 \end{equation}
@@
 \frac{\partial (\rho u)}{\partial t} + \nabla \cdot (\rho u \Ub) +
-     \frac{\partial p}{\partial x} &= -2\rho \Omega w  \\
+     \frac{\partial p}{\partial x} &= -2\rho \Omega_0 w  \\
 %
@@
      \cancelto{0}{\frac{\partial p}{\partial z}} &=
-    2\rho \Omega u 
+    2\rho \Omega_0 u 
 \end{align}
@@
 \begin{equation}
-\frac{\partial w}{\partial t} + \Ub \cdot \nabla w = 2\Omega u
+\frac{\partial w}{\partial t} + \Ub \cdot \nabla w = 2\Omega_0 u
 \end{equation}
@@
 
+\subsection{General Features}
+
+\AssignTo{Mike, Kiran, Alice}
+
+Here we'll show the large scale dynamics, colormap plots, etc. and discuss the trends.
+
+\subsection{Flame Acceleration}
+
+\AssignTo{Kiran}
+
+Describe the diagnostics used to measure the flame speed and compare to laminar.
+
+\subsection{Boosting}
+
+\AssignTo{Kiran, Mike}
+
+To meet reasonable timescales, we boosted the microphysical inputs to
+the flame to result in a faster flame.  Here we look at the effect of
+this choice.
+
+\subsection{Entrainment, Flow Features}
+
+\AssignTo{Alice, Don}
+
+Show the phase plots and whatever it was that Don cooked up.
+
+\subsection{Rotation Rate}
+
+\AssignTo{Mike, Kiran}
+
+We also try a rotation period of $P_\mathrm{rot} \sim 10^{-4}~\mathrm{s}$.
+
+\subsection{Reaction Network}
+
+\AssignTo{Mike, Kiran}
+
+We look at iso7 instead of aprox13
+
 \subsection{Resolution}
+
+\MarginPar{do we want to do a low res run?}
 
@@
 
-\subsection{Rotation Rate}
+\section{Discussion}
 
-We also try a rotation period of $P_\mathrm{rot} \sim 10^{-4}~\mathrm{s}$.
-
-\subsection{Reaction Network}
-
-We use aprox13 instead of the small network to understand the
-differences.  Energy comes out faster.
-
-\subsection{Boosting}
-
-To meet reasonable timescales, we boosted the microphysical inputs to
-the flame to result in a faster flame.  Here we look at the effect of
-this choice.
-
-\subsection{Sponge}
-
-\section{Discussion}
+\AssignTo{everyone}
 
@@
 
-MHD, radiation transport
+MHD, radiation transport, GPUs
 
@@
 
-\software{AMReX (\url{https://github.com/AMReX-Codes/amrex}),
+\software{AMReX \citep{amrex_joss},
           Castro \citep{castro},

### 2019-05-14 xrb/paper.tex
some intro refs

@@
 
+\author{Donald Willcox}
+\affiliation{Lawrence Berkeley National Laboratory, Berkeley, CA}
+
 \author{Yuri Cavecchi}
@@
 %
-\author{Chris M.\ Malone}
-\affiliation{Los Alamos National Laborary, Los Alamos, NM}
 
@@
 
-XRB intro
-
-Evidence for localized burning \MarginPar{refs}
+X-ray bursts (XRBs) are thermonuclear explosions of an accreted H or
+He layer on the surface of a neutron star (see~\citealt{galloway:2017}
+for a review).  Observations of bursts can help constrain the
+properties of the underlying neutron star, helping to illuminate the
+nuclear equation of state~\citep{steiner:2010,ozel:2010}.  Extensive
+observations of XRBs have provided evidence that the burning begins in
+a localized region and spreads over the surface of the neutron
+star~\citep{bhattacharyya:2006,bhattacharyya:2007}.
 
@@
 
-Global multidimensional studies were performed by \citet{spitkovsky2002}, where
-it was demonstrated that the Coriolis force plays an important role in 
-confining the burning as it spreads across the neutron star surface.  These
-calculations used the shallow-water approximation, so the vertical details
-of the atmosphere structure we not modeled.
+Global multidimensional studies were performed by
+\citet{spitkovsky2002}, where it was demonstrated that the Coriolis
+force plays an important role in confining the burning as it spreads
+across the neutron star surface.  These calculations used the
+shallow-water approximation, so the vertical details of the atmosphere
+structure we not modeled.
 

### 2019-05-15 xrb/paper.tex
fix some cites

@@
 parabolic method (PPM) with characteristic tracing for solving the
-hydrodynamics~\cite{ppm,millercolella:2002}, generalized to an
-arbitrary equation of state~\cite{zingalekatz}.  Reactions are
-incorporated via Strang splitting~\cite{strang:1968}, giving a method
+hydrodynamics~\citep{ppm,millercolella:2002}, generalized to an
+arbitrary equation of state~\citep{zingalekatz}.  Reactions are
+incorporated via Strang splitting~\citep{strang:1968}, giving a method
 that is overall second-order accurate in space and time.
 \castro\ uses the \amrex\ adaptive mesh refinement
-library~\cite{amrex_joss} to manage a hierarchy of grids at different
+library~\citep{amrex_joss} to manage a hierarchy of grids at different
 resolutions.

### 2019-05-15 gpus/paper.tex
some tweaks
add note
Frame not-Kokkos positively, not negatively

@@
 \newcommand{\multifab}{{\sf MultiFab}}
+
+\received{XXX X, XXXX}
+\revised{XXX X, XXXX}
+\accepted{XXX X, XXXX}
+%% Command to document which AAS Journal the manuscript was submitted to.
+%% Adds "Submitted to " the arguement.
+\submitjournal{ApJ}
 
@@
 In this paper, we describe our approach to performance portability
-with the \castro\ multiphysics simulation code.  We choose not to use
-a framework like Kokkos but instead we implement the mechanism for
-offloading compute kernels to GPUs inside of the AMReX adaptive mesh
-refinement library itself.  Our goal is to have the same computational
+with the \castro\ multiphysics simulation code.  Similar to the approaches
+of Kokkos and RAJA, we have implemented a mechanism for offloading compute
+kernels to GPUs inside of the AMReX adaptive mesh refinement library.
+Our goal is to have the same computational
 kernels run efficiently on CPUs (including manycore processors) and
@@
 converting these to C++ as well.
+
+needed to change to variable sizes known at compile time
 

### 2019-05-23 MAESTROeX_I/intro.tex
add cite to JOSS AMReX
some comments

@@
 \section{Introduction} \label{sec:intro}
-Many astrophysical flows are highly subsonic; sound waves carry sufficiently little energy that they do not significantly affect the convective dynamics of the system.
-In many of these flows, modeling long-time convective dynamics are of interest, and numerical approaches based on compressible hydrodynamics are intractable, even on modern supercomputers.
-One approach to this problem is to use low Mach number models.
-In a low Mach number approach, sound waves are eliminated from the governing equations while retaining compressibilitiy effects due to, e.g., nuclear energy release, compositional changes, and thermal diffusion.
-When the Mach number (the ratio of the characteristic fluid velocity over the characteristic sound speed; Ma $= U/c$) is small,
-the resulting system can be numerically integrated with much larger time steps than a compressible model.
-Specifically, the time step increase is at least a factor of $\sim 1/{\rm Ma}$ larger.
-Each time step is more computationally expensive due to the presence of additional linear solves,
-but for many problems of interest the overall gains in efficiency can easily be an order of magnitude or more.
+Many astrophysical flows are highly subsonic.  In this regime,
+sound waves carry sufficiently little energy that they do not
+significantly affect the convective dynamics of the system.  In many
+of these flows, modeling long-time convective dynamics are of
+interest, and numerical approaches based on compressible hydrodynamics
+are intractable, even on modern supercomputers.  One approach to this
+problem is to use low Mach number models.  In a low Mach number
+approach, sound waves are eliminated from the governing equations
+while retaining compressibilitiy effects due to, e.g., nuclear energy
+release, stratification, compositional changes, and thermal diffusion.  When the Mach
+number (the ratio of the characteristic fluid velocity over the
+characteristic sound speed; Ma $= U/c$) is small, the resulting system
+can be numerically integrated with much larger time steps than a
+compressible model.  Specifically, the time step increase is at least
+a factor of $\sim 1/{\rm Ma}$ larger.  Each time step is more
+computationally expensive due to the presence of additional linear
+solves, but for many problems of interest the overall gains in
+efficiency can easily be an order of magnitude or more.
 
@@
 based on method-of-lines integration.  
-In particular, schemes based on spectral deferred corrections \citep{dutt2000spectral} have been used to generate high-order temporal integrators for problems of reactive flow \citep{dutt2000spectral} and low Mach number combustion \citep{pazner2016high,nonaka2018conservative}.
+In particular, schemes based on spectral deferred corrections have been used to generate high-order temporal integrators for problems of reactive flow \citep{dutt2000spectral} and low Mach number combustion \citep{pazner2016high,nonaka2018conservative}.
 Second, we explore the effects of alternative spatial mapping routines for coupling the base state and the Cartesian grid state in spherical problems.
-Finally, the code has been completely ported to a new software framework suitable for modern manycore supercomputers.
-We examine the performance of our new MAESTROeX implementation in the new C++/F90 AMReX public software library \cite{AMReX}.
+Finally, the code has been completely ported to a new software framework suitable for modern manycore and GPU-based supercomputers.
+We examine the performance of our new MAESTROeX implementation in the new C++/F90 AMReX public software library \cite{AMReX,AMReX_JOSS}.
 MAESTROeX uses MPI+OpenMP parallelism and scales well to over 10,000 MPI processes, with each MPI process supporting tens of threads.

### 2019-05-23 MAESTROeX_I/equations.tex
add our grants

@@
 \end{eqnarray}
-Here $\rho$, $\Ub$, and $h$ are the mass density,
+Here $\rho$, $\Ub$, and $h$ are the mass density,  \MarginPar{that's not the momentum equation we solve.  It doesn't have the energy fix.}
 velocity and specific enthalpy, respectively, and

### 2019-05-23 MAESTROeX_I/algorithm.tex
some tweaks

@@
 Figure \ref{Fig:BaseGrid} illustrates the relationship between the base state and the Cartesian grid state for both planar and spherical geometries in the presence of spatially adaptive grids.
+\MarginPar{did this fig come from one of the earlier papers?
+if so, we will need to acknowledge that paper and make sure ApJ knows.}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{equation}
-Temperature is computed using the equation of state, e.g.,
-\footnote{As described in Paper V, for planar problems we compute temperature using $h$ instead of $p_0$, since we have successfully developed planar volume discrepancy schemes to effectively couple the enthalpy to the rest of the solution; see \cite{XRB_I}.
-We are still exploring this option for spherical stars.}
+Temperature is computed using the equation of state\footnote{As described in Paper V, for planar problems we compute temperature using $h$ instead of $p_0$, since we have successfully developed planar volume discrepancy schemes to effectively couple the enthalpy to the rest of the solution; see \cite{XRB_I}.
+We are still exploring this option for spherical stars.}, e.g.,
 \begin{equation}
@@
 \item In {\bf Step 1} we react the thermodynamic variables over $\Delta t/2$.
-\item In {\bf Steps 2-4} we advect the thermodynamic variables over $\Delta t$.
+\item In {\bf Steps 2--4} we advect the thermodynamic variables over $\Delta t$.
 Specifically, we compute an estimate for the expansion term, $S$, project the face-centered, time-centered velocities so they satisfy the divergence constraint, and then advect the thermodynamic variables.
-\item In {\bf Step 5} we react the thermodynamic variables over $\Delta t/2$.\footnote{After this step we could skip to the velocity advance in {\bf Steps 10-11}, however the overall scheme would be only first-order in time, so {\bf Steps 6-9} can be thought of as a trapezoidal corrector step.}
-\item In {\bf Steps 6-8} we redo the advection in {\bf Steps 2-4} but are able to use the trapezoidal rule to time-center certain quantities such as $S$, $\rho_0$, etc.
+\item In {\bf Step 5} we react the thermodynamic variables over $\Delta t/2$\footnote{After this step we could skip to the velocity advance in {\bf Steps 10--11}, however the overall scheme would be only first-order in time, so {\bf Steps 6-9} can be thought of as a trapezoidal corrector step.}.
+\item In {\bf Steps 6--8} we redo the advection in {\bf Steps 2--4} but are able to use the trapezoidal rule to time-center certain quantities such as $S$, $\rho_0$, etc.
 \item In {\bf Step 9} we redo the reactions from {\bf Step 5} using the improved results for the corrector advection step.
-\item In {\bf Steps 10-11} we and update the velocity, compute the expansion term, $S$, and project the cell-centered velocities.
+\item In {\bf Steps 10--11} we and update the velocity, compute the expansion term, $S$, and project the cell-centered velocities.
 \end{itemize}
@@
 Throughout Papers II-V, we have employed two techniques to help control these dynamics without significantly affecting the dynamics in the convective region.
-First, we use a cutoff density technique, where we hold the density constant outside a specified radius (typically near where the density is $\sim$4 orders of magnitude smaller than the largest densities in the simulation.
+First, we use a cutoff density technique, where we hold the density constant outside a specified radius (typically near where the density is $\sim$4 orders of magnitude smaller than the largest densities in the simulation).
 Second, we employ a sponge technique where we artificially damp the velocities near and beyond the cutoff region.

### 2019-05-29 xrb/paper.tex
Added some references and fixed some typos

@@
 have also been used to explore the sensitivity of the burst
-observables to reaction rates\MarginPar{ref}.
+observables to accretion and reaction rates~\citep{Jose2010a,Lampe2016}\MarginPar{ref}.
 
@@
 \citet{fryxellwoosley82,hedet}, but these would only be applicable at
-very high densities, since it is difficult to detonation helium at the
-densities found in normal XRBs, and even harder to detonation hydrogen
+very high densities, since it is difficult to detonate helium at the
+densities found in normal XRBs, and even harder to detonate hydrogen
 because of the waiting times for weak reactions.  This means that
 detonations may only really be applicable to superbursts (where carbon
-is the reactant). \MarginPar{ref}
+is the reactant)~\citep{Weinberg2006b,Weinberg2007}. \MarginPar{ref}
 
@@
 across the neutron star surface.  These calculations used the
-shallow-water approximation, so the vertical details of the atmosphere
+shallow-water approximation, so the vertical details of the atmosphere's
 structure we not modeled.
@@
 number methods that approximate the hydrodynamics equations to filter
-soundswaves, enabling large timesteps and efficient modeling of
+soundwaves, enabling large timesteps and efficient modeling of
 subsonic convection.  While these calculations could not support
@@
 or use a flame model.  Flame models usually assume that the flame
-structure is thin compared to the size of the system (see, e.g., XX\MarginPar{refs} for applications to Type Ia supernovae).  For the XRB however, the flame thickness is comparable to the scale
-height of the atmosphere, so we cannot use the approximations.
+structure is thin compared to the size of the system (see, e.g., \citet{Ropke2007}\MarginPar{refs} for applications to Type Ia supernovae).
+For XRBs however, the flame thickness is comparable to the scale
+height of the atmosphere, so we cannot use these approximations.
 
@@
 multidimensional simulation of flame propagation through the
-atmosphere on the neutron star.  For simplicity in this first set of
+atmosphere on a neutron star.  For simplicity in this first set of
 calculations, we will use pure helium composition.  These studies
 complement the other multidimensional work described above in helping
-us build a picture of the dynamics of X-ray bursts.
+us to build a picture of the dynamics of X-ray bursts.
 
@@
 simulations presented here, we work in axisymmetric coordinates,
-but we advective a third component of velocity, coming out of the
+but we advect a third component of velocity, coming out of the
 simulation plane, that participates in the Coriolis force (sometimes
@@
 smaller 7-isotope network described in \citet{iso7}.  We integrate the
-network using the VODE integration package~\citep{vode}, and out
+network using the VODE integration package~\citep{vode}, and our
 implementation is provided in the StarKiller Microphysics

### 2019-06-03 MAESTROeX_I/algorithm.tex
Latest plots and updates

@@
 Then we interpolate this data onto a one-dimensional radial array.
+
 Previously, MAESTRO only allowed for radial arrays with a constant $\Delta r$ (typically equal to $\Delta x/5$).
@@
 By retaining the radial base state as an irregularly spaced array, the effects due to interpolation error are completely eliminated.
-\MarginPar{Doreen you had a great figure of this mapping in one of your talks}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[tb]
+\centering
+\includegraphics[height=2.0in]{./figs/base_spherical_new}
+\caption{\label{Fig:NewBaseGrid}  
+A direct mapping between the radial array cell centers (red squares) and the Cartesian grid cell centers (blue crosses) 
+is enforced by computing the average of the grid cell centers that share the same radial distance from 
+the center of the domain.}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 = \Ub^n - \dt \left[\uadvtwo \cdot \nabla \Ub^{\nph,\pred} \right]
- - \dt \left[ \frac{1}{\rho^\nph} \nabla \pi^\nmh + \frac{\left(\rho^\nph-\rho_0^\nph\right)}{\rho^\nph} g^{\nph} \eb_r \right],
+ - \dt \left[ \frac{\beta_0^\nph}{\rho^\nph} \nabla \left( \frac{\pi^\nmh}{\beta_0^\nmh} \right) + \frac{\left(\rho^\nph-\rho_0^\nph\right)}{\rho^\nph} g^{\nph} \eb_r \right],
 \end{equation}

### 2019-06-03 MAESTROeX_I/paper.tex
add cite to JOSS AMReX
add our grants

@@
 and can robustly handle several orders of magnitude of density and pressure stratification.
-We present a new, greatly simplified temporal integration scheme that retains the same order of accuracy as our previous approaches.
+Previously, we have described the development of the predecessor of MAESTROeX, called MAESTRO, in a series of papers. 
+Here, we present a new, greatly simplified temporal integration scheme that retains the same order of accuracy as our previous approaches. 
 We also explore the use of alternative spatial mapping of the one-dimensional base state onto the full Cartesian grid.

### 2019-06-03 MAESTROeX_I/results.tex

@@
 
-white dwarf convection runs with 3 algorithms (original, new temporal, new temporal + irregular base state)
+white dwarf convection runs with 3 algorithms (original, new temporal, new temporal + variable-spaced base state)
 
@@
 \begin{center}
-\includegraphics[width=2.75in]{./figs/wdconvect_128_maxT} 
+\includegraphics[width=2.75in]{./figs/wdconvect_128_maxT}  \hspace{0.5in}
 \includegraphics[width=2.75in]{./figs/wdconvect_128_maxMach}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[hbt]
+\begin{center}
+\includegraphics[width=2.7in]{./figs/wdconvect_256_maxT}  \hspace{0.5in}
+\includegraphics[width=2.75in]{./figs/wdconvect_512_maxT}
+\caption{\label{fig:wdconvect_Tmax} Peak temperature, $T_{\text{peak}}$  in a white dwarf at 
+         (left) resolution of $256^3$ until time of ignition and (right) resolution of $512^3$ until $t=1000$ 
+         for three different MAESTROeX algorithms. Note that the variable-spaced (d$r$) solution agrees better 
+         with the evenly-spaced solution as the resolution increases.}
+\end{center}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{AMR}
 

### 2019-06-03 MAESTROeX_I/algorithm.tex
some comments

--- 2019-05-23 MAESTROeX_I/algorithm.tex
@@
 %
-omitting the reaction terms, which were already
+omitting the reaction terms, which were already \MarginPar{since we are omitting the reaction terms, we should just leave them out of the Eq.}
 accounted for in {\bf React State}.  The update consists of two steps:
@@
 \begin{equation}
-\beta_0^{\nph} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2},
+\beta_0^{\nph} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2}, \quad
 \gammaonebar^{\nph} = \frac{ \gammaonebar^n +  \gammaonebar^{n+1,\star} }{2}.
@@
 \item Update the enthalpy.  This step is identical to {\bf Step 4D} except we use
-  the updated values $\uadvtwo, \rho_0^{n+1}, (\rho h)_0^{n+1}$, and $p_0^{\nph}$
+  the updated values $\uadvtwo$, $\rho_0^{n+1}$, $(\rho h)_0^{n+1}$, and $p_0^{\nph}$
   rather than
-  $\uadvone, \rho_0^{n+1,\star}, (\rho h)_0^{n+1,\star}$, and $p_0^n$.
+  $\uadvone, \rho_0^{n+1,\star}$, $(\rho h)_0^{n+1,\star}$, and $p_0^n$.
   In particular:

### 2019-06-03 MAESTROeX_I/equations.tex
some tweaks

--- 2019-04-30 MAESTROeX_I/equations.tex
@@
 \begin{eqnarray}
-\frac{\partial\Ub}{\partial t} &=& -\Ub\cdot\nabla\Ub  - \frac{1}{\rho}\nabla\pi - \frac{\rho-\rho_0}{\rho} g\eb_r,\label{eq:momentum}\\
 \frac{\partial(\rho X_k)}{\partial t} &=& -\nabla\cdot(\rho X_k\Ub) + \rho\omegadot_k,\label{eq:species}\\
+\frac{\partial\Ub}{\partial t} &=& -\Ub\cdot\nabla\Ub  - \frac{\beta_0}{\rho}\nabla\left(\frac{\pi}{\beta_0}\right) - \frac{\rho-\rho_0}{\rho} g\eb_r,\label{eq:momentum}\\
 \frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc.\label{eq:enthalpy}
@@
 $X_k$ are the mass fractions of species $k$ with associated
-production rate $\omegadot_k$.  The species are constrained
-such that $\sum_k X_k = 1$ giving $\rho = \sum_k (\rho X_k)$ and
+production rate $\omegadot_k$.  
+The species are constrained such that $\sum_k X_k = 1$ giving $\rho = \sum_k (\rho X_k)$ and
 \begin{equation}
@@
 Here $\Hnuc$ is the nuclear energy generation rate per unit mass. 
+Note that the momentum equation above enforces conservation of total energy in the 
+low Mach number system in the absence of external heating or viscous terms \citep{Vasil2013}, 
+and has been previously demonstrated in modeling sub-Chandrasekhar white dwarfs using MAESTRO \citep{subChandra_II}. 
 The total pressure is decomposed into a one-dimensional hydrostatic base state
@@
 \end{equation}
-where $\gammaonebar$ is the lateral average of $\Gamma_1 = d(\log p)/d(\log\rho)$ at constant entropy.
+where $\gammaonebar$ is the lateral average of $\Gamma_1 = d(\log p)/d(\log\rho) |_s$ (evaluated with entropy, $s$, constant).
 The expansion term, $S$, incorporates local compressibility effects due to compositional changes and heat release from reactions,
@@
 \end{equation}
-where $p_{X_k} \equiv \left. \partial p / \partial X_k
-\right|_{\rho,T,X_{j,j\ne k}}$, $\xi_k \equiv \left. \partial h /
-\partial X_k \right |_{p,T,X_{j,j\ne k}},
-p_\rho \equiv \left.
-\partial p/\partial \rho \right |_{T, X_k}$, and $\sigma \equiv
-p_T/(\rho c_p p_\rho)$, with $p_T \equiv \left. \partial p / \partial
+where 
+$p_{X_k} \equiv \left. \partial p / \partial X_k \right|_{\rho,T,X_{j,j\ne k}}$, 
+$\xi_k \equiv \left. \partial h /\partial X_k \right |_{p,T,X_{j,j\ne k}}$,
+$p_\rho \equiv \left.\partial p/\partial \rho \right |_{T, X_k}$, and 
+$\sigma \equiv p_T/(\rho c_p p_\rho)$, with $p_T \equiv \left. \partial p / \partial
 T \right|_{\rho, X_k}$ and $c_p \equiv \left.  \partial h / \partial T

### 2019-06-03 xrb/paper.tex

@@
 the domain (the origin of the axisymmetric coordinates).  To create
-these initial conditions, w produce two different hydrostatic models,
+these initial conditions, we produce two different hydrostatic models,
 a ``hot'' model that will represent the perturbation that drives the
@@
 
+Baroclinicity is calculated as
+\begin{equation}
+    \boldsymbol{\psi} = \frac{1}{\rho^2} \nablab p \times \nablab \rho,
+\end{equation}
+and shows the misalignment of the local density and pressure gradients
+\citep{Malone2014a}. As we are considering a 2d system, the component of the
+baroclinicity we consider here (out of the plane) reflects the misalignment
+of the fields in the plane of the simulation. Figure~\ref{fig:baroclinicity}
+shows that the baroclinicity peaks along the flame front. This baroclinicity
+generates vorticity, which in turn entrains material along the surface of the
+flame.
+
+\begin{figure}[t]
+\centering
+\plotone{plots/results/baroclinicity}
+\caption{\label{fig:baroclinicity} Baroclinicity. This plot shows
+$\ln \left(\mathbf{\psi}\right)$ at time $t = 0.02~\mathrm{s}$.}
+\end{figure}
+
+We show some phase plots in Figure~\ref{fig:phase_plots}. The phase plot of the
+energy generation rate against the as a function of
+the $x$- and $y$-velocities shows that energy is preferentially generated in
+regions with negative $x$-velocity. This most likely corresponds to material
+being entrained along the underside of the flame front. There are loops of
+points of similar $\dot{e}_\mathrm{nuc}$ in the outer edges of the plot.
+These are likely to correspond to the vortices that appear within the flame,
+and show the fluid moving in a circular motion in $u-v$ phase space.
+
+The phase plot of the energy generation rate as a function of the $x$-velocity
+and the density demonstrates that the energy generation rate peaks at
+$\rho \sim 3 \times 10^5~\mathrm{g/cm}^3$, at the base of the flame.
+\MarginPar{I'm guessing? It's at least towards the base}
+
+\begin{figure}[t]
+\centering
+\plottwo{plots/results/vx_vy_129139}{plots/results/u_rho_129139}
+\caption{\label{fig:phase_plots} Phase plots at time $t = 0.02~\mathrm{s}$.
+\emph{Left}: Phase plot showing the energy generation rate as a function of
+the $x$- and $y$-velocities. \emph{Right}: phase plot showing the energy
+generation rate as a function of the $x$-velocity and the density.}
+\end{figure}
+
+Figure~\ref{fig:abar_temp} shows the energy generation rate as a function of the
+temperature and $\bar{A}$. \MarginPar{I'm assuming will be defined somewhere above}
+The energy generation rate peaks at low $\bar{A}$, where there is a high fraction
+of unburnt material. The burning raises the temperature of this material, such
+that the peak temperature coincides with the peak energy generation rate. The
+material then cools again as the burning converts the fuel into ashes,
+increasing $\bar{A}$ and reducing the energy generation rate as there is
+less available fuel. Along the base of the plot we see cool unburnt fluid and
+ashes. In the center of the plot there are thin `trails' in phase space, which
+could correspond to less common reaction pathways in the reaction network.
+
+\begin{figure}[t]
+\centering
+\plotone{plots/results/abar_temp_129139}
+\caption{\label{fig:abar_temp} Phase plot of energy generation rate as a function of $\bar{A}$ and temperature at time $t = 0.02~\mathrm{s}$.}
+\end{figure}
+
 \subsection{Rotation Rate}

### 2019-06-04 MAESTROeX_I/appendix.tex
Added figures to appendix A

@@
 \MarginPar{mention changes in $w_0$ computation for irregular base state. Add graphic of temperature falling due to incorrect splitting?}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[htb]
+\begin{center}
+\begin{tabular}{l r}
+\multirow{4}{3.5in}[34.5mm]{ \includegraphics[width=3.0in]{./figs/wdconvect_256_splitU} } & \multicolumn{1}{c}{\includegraphics[width=2.35in]{./figs/magvel_full_XY}} \\
+& \multicolumn{1}{c}{\begin{footnotesize} (a) $|\mathbf{U}|$, solved using $\mathbf{U}$ \end{footnotesize}} \\[1.em]
+& \multicolumn{1}{c}{\includegraphics[width=2.35in]{./figs/magvel_split_XY}} \\ 
+& \multicolumn{1}{c}{\begin{footnotesize} (b) $|\mathbf{U}|$, solved using $w_0 + \tilde{\mathbf{U}}$ \end{footnotesize}} \\
+\end{tabular}
+\caption{\label{fig:wdconvect_splitU} (Left) Peak temperature, $T_{\text{peak}}$ in a white dwarf at resolution of $256^3$ 
+         during initial transition. (Right) Velocity magnitude solved using the full velocity $\mathbf{U}$ in projection step 
+         results in large values outside the density cutoff region that is not seen with using the split velocities.}
+\end{center}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2019-06-04 MAESTROeX_I/intro.tex
add some more refs

@@
 
-Low Mach number models have been developed for a variety of contexts including combustion \citep{day2000numerical}, terrestrial atmospheric modeling \citep{duarte2015low},
-and elastic solids \citep{abbate2017all}.
-For astrophysical applications, a number of approaches to modeling low Mach number flows have been developed in recent years.
-These include semi-implicit all-Mach number solvers, where the Euler equations are split into an acoustic part and an advective part \citep{Kwatra2009,Degond2009,Cordier2012,Haack2012,Happenhofer2013,Chalons2016,Padioleau2019}.
-The fast acoustic waves are then solved using implicit time integration, while the slow material waves are solved explicitly.
-Another approach is to use preconditioned all-Mach number solvers \citep{Miczek2014,Barsukow2016}, where the numerical flux is multiplied by a preconditioning matrix.
-This reduces the stiffness of the system at low Mach numbers, while retaining the correct scaling behavior.
-In the reduced speed of sound technique (RSST) and related methods, the speed of sound is artificially reduced by including a suitable scaling factor in the continuity equation, reducing the restriction on the size of the timestep \citep{Rempel2005,Hotta2012,Wang2015,Takeyama2017,Iijima2018}.
-The MUSIC code uses fully implicit time integration for the compressible Euler equations, which therefore allows for arbitrarily large timesteps \citep{Viallet2011,Viallet2015,Goffrey2016}.
+Low Mach number models have been developed for a variety of contexts
+including combustion \citep{day2000numerical}, terrestrial atmospheric
+modeling \citep{durran:1989,oneill:2014,duarte2015low}, and elastic
+solids \citep{abbate2017all}.  For astrophysical applications, a
+number of approaches to modeling low Mach number flows have been
+developed in recent years.  These include methods atmospheric model similar to
+what is presented here~\cite{Lin:2006}.  There are also  semi-implicit all-Mach
+number solvers, where the Euler equations are split into an acoustic
+part and an advective part
+\citep{Kwatra2009,Degond2009,Cordier2012,Haack2012,Happenhofer2013,Chalons2016,Padioleau2019}.
+The fast acoustic waves are then solved using implicit time
+integration, while the slow material waves are solved explicitly.
+Another approach is to use preconditioned all-Mach number solvers
+\citep{Miczek2014,Barsukow2016}, where the numerical flux is
+multiplied by a preconditioning matrix.  This reduces the stiffness of
+the system at low Mach numbers, while retaining the correct scaling
+behavior. In the reduced speed of sound technique (RSST) and related
+methods, the speed of sound is artificially reduced by including a
+suitable scaling factor in the continuity equation, reducing the
+restriction on the size of the timestep
+\citep{Rempel2005,Hotta2012,Wang2015,Takeyama2017,Iijima2018}.
+Finally, there are fully implicit time integration codes for the
+compressible Euler equations
+\citep{Viallet2011,kifonidis:2012,Viallet2015,Goffrey2016}.  The MUSIC code
+uses fully implicit time integration for the compressible Euler
+equations, which therefore allows for arbitrarily large timesteps.
 
@@
 Central to the algorithm is a time-varying, one-dimensional stratified background (or base) state density and pressure held in hydrostatic equilibrium that couple to the full Cartesian grid solution.
-The original MAESTRO code was developed in the pure-Fortran90 FBoxLib software framework (see the FBoxLib repository in \cite{AMReX}) for block-structured AMR applications.
+The original MAESTRO code was developed in the pure-Fortran 90 FBoxLib software framework (available in the same github organization as AMReX).
 
@@
 \item The convective phase preceding Chandrasekhar mass models for type Ia supernovae \citep{MAESTRO_convection,MAESTRO_AMR,MAESTRO_CASTRO}.
-\item Convection in massive stars \citep{Gilet:2013}.
+\item Convection in massive stars \citep{Gilet:2013,gilkis:2016}.
 \item Sub-Chandrasekhar white dwarfs \citep{subChandra_I,subChandra_II}.

### 2019-06-04 MAESTROeX_I/equations.tex
made it through section 2

@@
 in astrophysical environments using adaptive algorithms in space and time \citep{Bell:2004}.
-In Papers I-III, we extended this work by deriving a model and algorithm suitable for stratified astrophysical flow.
+In Papers I-III, we extended this work and the atmospheric model by \citet{durran:1989} by deriving a model and algorithm suitable for stratified astrophysical flow.
 In our model, we take the standard equations of reacting, compressible flow, and recast the equation
@@
 $X_k$ are the mass fractions of species $k$ with associated
-production rate $\omegadot_k$.  
+production rate $\omegadot_k$ and energy release per time per unit mass $\Hnuc$.
 The species are constrained such that $\sum_k X_k = 1$ giving $\rho = \sum_k (\rho X_k)$ and
@@
 \end{equation}
-Here $\Hnuc$ is the nuclear energy generation rate per unit mass. 
 Note that the momentum equation above enforces conservation of total energy in the 
-low Mach number system in the absence of external heating or viscous terms \citep{Vasil2013}, 
+low Mach number system in the absence of external heating or viscous terms \citep{kleinpauluis,Vasil2013}, 
 and has been previously demonstrated in modeling sub-Chandrasekhar white dwarfs using MAESTRO \citep{subChandra_II}. 
@@
 
-Mathematically, equations (\ref{eq:momentum})-(\ref{eq:enthalpy}) must still be closed by the EOS.  
+Mathematically, equations \ref{eq:momentum}--\ref{eq:enthalpy} must still be closed by the EOS.  
 This is done by taking the Langrangian derivative of the EOS for pressure as a function of the thermodynamic variables, 
@@
 
-To summarize, we model evolution equations for momentum, mass, and energy, (\ref{eq:momentum})-(\ref{eq:enthalpy}) subject to a divergence constraint on the velocity (\ref{eq:U divergence}) and the hydrostatic equilibrium condition (\ref{eq:HSE}).
+To summarize, we model evolution equations for momentum, mass, and energy, equations \ref{eq:momentum}--\ref{eq:enthalpy} subject to a divergence constraint on the velocity, equation \ref{eq:U divergence}, and the hydrostatic equilibrium condition, equation \ref{eq:HSE}.

### 2019-06-04 MAESTROeX_I/intro.tex

@@
 Finally, the code has been completely ported to a new software framework suitable for modern manycore and GPU-based supercomputers.
-We examine the performance of our new MAESTROeX implementation in the new C++/F90 AMReX public software library \cite{AMReX,AMReX_JOSS}.
+We examine the performance of our new MAESTROeX implementation in the new C++/F90 AMReX public software library \citep{AMReX,AMReX_JOSS}.
 MAESTROeX uses MPI+OpenMP parallelism and scales well to over 10,000 MPI processes, with each MPI process supporting tens of threads.
 The resulting code is publicly available on GitHub ({\tt https://github.com/AMReX-Astro/MAESTROeX}),
-uses the Starkiller-Astro microphysics libraries ({\tt https://github.com/starkiller-astro/Microphysics}),
+uses the Starkiller-Astro microphysics libraries \citep{starkiller},
 and the AMReX software library for block-structured adaptive mesh refinement ({\tt https://github.com/AMReX-Codes/amrex}).

### 2019-06-04 MAESTROeX_I/algorithm.tex
updated appendix

--- 2019-06-03 MAESTROeX_I/algorithm.tex
@@
 Figure \ref{Fig:BaseGrid} illustrates the relationship between the base state and the Cartesian grid state for both planar and spherical geometries in the presence of spatially adaptive grids. 
-\MarginPar{did this fig come from one of the earlier papers?
-if so, we will need to acknowledge that paper and make sure ApJ knows.}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 between the radial array cell centers and the Cartesian grid cell centers, we choose to fix
-the radial base state spacing across levels.}
+the radial base state spacing across levels. Reprinted from Paper V \citep{MAESTRO_V}. }
 \end{figure}

### 2019-06-04 MAESTROeX_I/results.tex
updated appendix

@@
 
-white dwarf convection runs with 3 algorithms (original, new temporal, new temporal + variable-spaced base state)
+white dwarf convection runs with 3 algorithms (original, new temporal, new temporal + irregularly-spaced base state)
 
@@
          (left) resolution of $256^3$ until time of ignition and (right) resolution of $512^3$ until $t=1000$ 
-         for three different MAESTROeX algorithms. Note that the variable-spaced (d$r$) solution agrees better 
-         with the evenly-spaced solution as the resolution increases.}
+         for three different MAESTROeX algorithms. Note that the irregularly-spaced (d$r$) solution agrees better 
+         with the uniformly-spaced solution as the resolution increases.}
 \end{center}

### 2019-06-05 MAESTROeX_I/appendix.tex

--- 2019-04-30 MAESTROeX_I/appendix.tex
@@
 Namely, to enforce that $r^2 w_0$ remain constant is difficult to do when solving for the full velocity.
+This is demonstrated in Figure \ref{fig:wdconvect_splitU} (right) where the velocity magnitude is observed to incorrectly increase outside the cutoff density radius when we use the full velocity in the nodal projection.
+The resulting peak temperature also dips significantly as seen in Figure \ref{fig:wdconvect_splitU} (left).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[htb]
+\begin{center}
+\begin{tabular}{l r}
+\multirow{4}{3.5in}[34.5mm]{ \includegraphics[width=3.0in]{./figs/wdconvect_256_splitU} } & \multicolumn{1}{c}{\includegraphics[width=2.35in]{./figs/magvel_full_XY}} \\
+& \multicolumn{1}{c}{\begin{footnotesize} (a) $|\mathbf{U}|$, solved using $\mathbf{U}$ \end{footnotesize}} \\[1.em]
+& \multicolumn{1}{c}{\includegraphics[width=2.35in]{./figs/magvel_split_XY}} \\ 
+& \multicolumn{1}{c}{\begin{footnotesize} (b) $|\mathbf{U}|$, solved using $\tilde{\mathbf{U}}+w_0$ \end{footnotesize}} \\
+\end{tabular}
+\caption{\label{fig:wdconvect_splitU} (Left) Peak temperature, $T_{\text{peak}}$ in a white dwarf at resolution of $256^3$ 
+         during initial transition. (Right) Velocity magnitude solved using full velocity $\mathbf{U}$ in the projection step 
+         results in large values outside the density cutoff region that is not seen when we split the velocity dynamics.}
+\end{center}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 So in practice, we solve the constraint over the lateral average,
@@
 \end{equation}
-and sepratately solve for $w_0$ using 
+and separately solve for $w_0$ using 
 \begin{equation}
@@
 Note that MAESTRO contains alternate low Mach number formulations that conserve total energy in stratified systems, with minimal changes to the code (see Appendix A of \cite{subChandra_II} for details).
-To find $w_0$ we integrate in 1D using the procedure in Appendix B of Paper V.
+To find $w_0$ we integrate in 1D using the procedure in Appendix B of Paper V, keeping in mind that the base state spacing ($\Delta r$) should be computed using the appropriate cell-edge and cell-center locations when using irregularly-spaced base state.
 Note that in this approach, we estimated the time-derivative of the pressure in part by examining how laterally averaged $\rho'=\rho-\rho_0$ changed over time (quantified by $\eta_\rho = \overline{\rho'\Ub\cdot\eb_r}$).
-After evolving the species, we compute this term in the same way as reported in Paper V for use in computing $w_0$.
-\MarginPar{mention changes in $w_0$ computation for irregular base state. Add graphic of temperature falling due to incorrect splitting?}
-
-\section{Old Notes (not for inclusion in paper)}
-
-\subsection{Exact Base State Notes}
-
-\subsubsection{Potential Problems}
-\begin{enumerate}
-\item How do we define \verb|r_edge_loc| for irregularly-spaced base states whose cell-centered values are defined at $r_m=\Delta x\sqrt{\frac{3}{4} + 2m}$?
-\item We would stil need to approximate the values at edges when advecting the base states (i.e.
-$\rho_0^{\nph,\pred}$).
-This may affect the nice properties we would get with exact cell-centered base state values.
-\item Excessive memory storage may be required (\verb|nr_irreg >> nr_fine|).
-\end{enumerate}
-
-\subsubsection{A Possible Solution}
-We could choose to no longer advect the base states for both predictor and corrector steps.
-This would solve the problems with computing state values at cell edges.
-Instead, we propose to only update the base states $\rho_0$ and $p_0$ by computing the average of the full state values after advecting the full states.
-This would result in the following changes in the algorithm from Paper V:
-%
-\begin{enumerate}
-\item Steps 4 and 8 will be greatly simplified, leading to a more efficient algorithm.
-\item \sout{There is no longer a need for $w_0$, so we can use the full velocity $\Ub$ instead of $\Ubt$.
-As a consequence, $\eta$ does not need to be computed.
-This has the added advantage of saving some memory storage by removing many of the terms needed when we decided to split terms into its base state and perturbation state in our previous implementation.} We split the velocity for the projection to handle the cutoff.
-\item By placing the base states exactly at cell-centers, we eliminate interpolation errors and simplify the existing code.
-\end{enumerate}
-%
-
-\subsection{$\eta_\rho$ Notes}
-After {\bf Step 8D}, we define a radial cell-centered $\etarho^{\nph}$.
+After evolving the species ({\bf Steps 4A/8A}), we compute this term in a similar fashion as reported in Paper V with the slight difference in using the full velocity instead of the split velocities.
+For example, after {\bf Step 8A}, we define a radial cell-centered $\etarho^{\nph}$,
 
@@
 \end{description}
-
-\subsection{Computing $w_0$ For Spherical Problems}
-Recall that we want to solve
-\begin{equation}
-\frac{1}{r^2}\frac{\partial}{\partial r}(r^2 w_0) = \Sbar - \frac{1}{\gammaonebar p_0}\left(\frac{\partial p_0}{\partial t} + w_0\frac{\partial p_0}{\partial r}\right)
-\end{equation}
-First, note that $\partial p_0/\partial r = -\rho_0 g$.
-The key observation here is that we can compute $\partial p_0/\partial t$ directly instead of making complicated substitutions involving $\eta_\rho$ to eliminate it.
-Let's discretize this as follows:
-\begin{equation}
-\frac{1}{r_j^2}\left[\frac{\left(r^2 w_0\right)_{j+\myhalf} - \left(r^2 w_0\right)_{j-\myhalf}}{\Delta r}\right] = 
-\Sbar_j - \frac{1}{\left(\gammaonebar p_0\right)_j}\left[\left(\frac{\partial p_0}{\partial t}\right)_j - \left(\frac{w_{0,j-\myhalf}+w_{0,j+\myhalf}}{2}\right)(\rho_0 g)_j\right]
-\end{equation}
-Rearranging this to solve for $w_{0,j+\myhalf}$ gives us an explicit update:
-\begin{equation}
-\left[\frac{r_{j+\myhalf}^2}{r_j^2\Delta r} - \frac{(\rho_0 g)_j}{2\left(\gammaonebar p_0\right)_j}\right] w_{0,j+\myhalf} =
-\left[\frac{r_{j-\myhalf}^2}{r_j^2\Delta r} + \frac{(\rho_0 g)_j}{2\left(\gammaonebar p_0\right)_j}\right] w_{0,j-\myhalf}
-+ \Sbar_j - \frac{1}{\left(\gammaonebar p_0\right)_j}\left(\frac{\partial p_0}{\partial t}\right)_j
-\end{equation}
-As before, once $\rho_0$ falls below $\rho_{\rm cutoff}$, we hold $r^2 w_0$ constant.\\ \\
-The interface to the routine in the algorithm description is
-{\bf Make $\mathbf{w_0}$}$[\rho_0,p_0,\partial p_0/\partial t,\gammaonebar,\Sbar] \rightarrow [w_0]$
-
-\subsection{TODO List}
-
-TODO list:
-\begin{itemize}
-\item Weak scaling tests of original algorithm, comparing MAESTRO to MAESTROeX.  Use cori knl, 4 MPI per node, 16 threads per MPI
-\item Run at effective $512^3$ resolution (maybe to ignition?...
-need to see how max T trends go) (1) classic MAESTRO, (2) MAESTROeX with original algorithm, (3) MAESTROeX with new temporal integrator, (4) MAESTROeX with new temporal integrator and irregular dr, (5) MAESTROeX with original algorithm and 3-levels of AMR, (6) MAESTROeX with new temporal integrator and 3 levels of AMR.
-\item (perhaps) a demonstration of what goes wrong when you don't split the velocity dynamics in the projection
-\end{itemize}

### 2019-06-05 MAESTROeX_I/algorithm.tex
Updated section 3, numerical algorithm

--- 2019-06-04 MAESTROeX_I/algorithm.tex
@@
 \subsection{Spatial Discretization}\label{Sec:Spatial}
-\MarginPar{needs work}
 The spatial discretization and adaptive mesh refinement methodology remains unchanged from Paper V.
@@
 In Section 4 of Paper V, we describe how there is a finite, easily computable set of radii that any Cartesian cell-center can map to.  
-Specifically, for every Cartesian cell, there exists an integer $m$ such that the radius to the center of the star is given by $\hat{r}_m=\Delta x\sqrt{0.75+2m}$.
+Specifically, for every Cartesian cell, there exists an integer $m$ such that the radius to the center of the star is given by 
+\begin{equation}\label{eqn:radii}
+\hat{r}_m=\Delta x\sqrt{0.75+2m}.
+\end{equation}
 We average by binning all the cells associated with each radii and taking the average.
@@
 Here we present a new option to retain an irregularly-spaced radial array to eliminate mapping errors back onto the Cartesian grid state.
-Consider a spherical star in hydrostatic equlibrium at rest.  In the absense of reactions, the star should remain at rest.
+The idea is to follow the same method performed on planar geometries by simply computing an arithmetic average of cells that share the same radial distance from the center of the star. 
+Figure \ref{Fig:NewBaseGrid} illustrates the relationship between the Cartesian grid state and the one-dimensional array base state with this new mapping technique.
+Again, these radial distances, and hence the cell-center locations in the new base state array, can be predetermined using equation (\ref{eqn:radii}).
+Also note that $\Delta r$ decreases as the base state moves further from the center of the star, which results in far more total cells in the irregularly-spaced array than the previous uniformly-spaced array.
+A potential benefit to eliminating the mapping error is to consider a spherical star in hydrostatic equlibrium at rest.  In the absense of reactions, the star should remain at rest.
 The buoyancy forcing term in the momentum equation contains $\rho-\rho_0$.
@@
 \centering
-\includegraphics[height=2.0in]{./figs/base_spherical_new}
+\includegraphics[height=2.5in]{./figs/base_spherical_new}
 \caption{\label{Fig:NewBaseGrid}  
@@
 is enforced by computing the average of the grid cell centers that share the same radial distance from 
-the center of the domain.}
+the center of the star.}
 \end{figure}
@@
 \item {\bf Average}$[\phi]\rightarrow[\overline\phi]$ computes the lateral average of a quantity over a layer at constant radius $r$, as described above in Section \ref{Sec:Spatial}.
-\item {\bf Enforce HSE}$[\rho_0]\rightarrow[p_0]$ computes the base state pressure, $p_0$, from a base state density, $\rho_0$ by integrating the hydrostatic equilibrium condition in one dimension; see equation (A10) in Paper V.
+\item {\bf Enforce HSE}$[\rho_0]\rightarrow[p_0]$ computes the base state pressure, $p_0$, from a base state density, $\rho_0$ by integrating the hydrostatic equilibrium condition in one dimension.
+This follows equation (A10) in Paper V with the exception of computing a non-constant $\Delta r_{j+1/2} = r_{j+1} - r_j$ for cell-edge $j+1/2$ if using irregularly-spaced base states.
 The base state pressure remains equal to a constant value from the location of a prescribed cutoff density outward for the entire simulation.
-\MarginPar{add note about changes for irregular}
 \item {\bf React State}$[(\rho X_k)^{\rm in},(\rho h)^{\rm in},p_0]\rightarrow[(\rho X_k)^{\rm out},(\rho h)^{\rm out},(\rho\dot\omega),(\rho\Hnuc)]$ integrates the species and enthalpy due to reactions over $\Delta t/2$ by solving

### 2019-06-05 MAESTROeX_I/results.tex
Updated figure in appendix and moved figures in results.

@@
 \begin{center}
-\includegraphics[width=2.75in]{./figs/MAESTRO_scaling1}
-\includegraphics[width=2.75in]{./figs/MAESTRO_scaling2}
+\includegraphics[width=3.0in]{./figs/MAESTRO_scaling1} \hspace{0.5em}
+\includegraphics[width=3.0in]{./figs/MAESTRO_scaling2}
 \caption{\label{fig:scaling} (Left) Weak scaling results for a spherical, full-star sub-Chandrasekhar mass white dwarf calculation using the original MAESTRO code, MAESTROeX, and MAESTROeX with base state evolution disabled.
@@
 \begin{center}
-\includegraphics[width=2.75in]{./figs/reacting_bubble_amr} 
-$\qquad$
-\includegraphics[width=2.75in]{./figs/wdconvect_amr_3grid}
+\includegraphics[width=2.5in]{./figs/reacting_bubble_amr} \hspace{2.5em}
+\includegraphics[width=2.5in]{./figs/wdconvect_amr_3grid}
 \caption{\label{fig:amr_grids} Initial grid structures with two levels of refinement. 
@@
 \begin{center}
-\includegraphics[width=2.75in]{./figs/reacting_bubble_result} 
-$\qquad$
-\includegraphics[width=2.75in]{./figs/reacting_bubble_amr_result}
+\includegraphics[width=2.5in]{./figs/reacting_bubble_result} \hspace{2.5em}
+\includegraphics[width=2.5in]{./figs/reacting_bubble_amr_result}
 \caption{\label{fig:bubble_results} Time-lapse cross-section of a hot bubble in a white dwarf environment at 
@@
 This 89\% decrease in runtime is mostly due to the fact that only 6.25\% of the cells (1,048,576 out of $128^2 \times 1024$ cells) are refined at the finest level for the AMR run.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=4.0in]{./figs/wdconvect_amr_Tmax} 
+\caption{\label{fig:wdconvect_amr_Tmax} Peak temperature, $T_{\text{peak}}$, in a white dwarf from $t=0$ to $400$ s 
+         for grids with effective resolution of $512^3$. We can see that the adaptive grids with two levels of 
+         refinement give very similar solution trends compared to the single-level grid.}
+\end{center}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Full white dwarf star simulation at effective resolution of $512^3$ on Cori KNL with 512 processors and 16 threads took 26.01 s per time step (averaged over 10 time steps) using the original temporal scheme, and 22.66 s using the new scheme.
+On an adaptive grid with two levels of refinement, 8.9\% of the cells (1,493,504 out of $256^3$ cells) are refined at the first level and 5.3\% (7,176,192 out of $512^3$ cells) at the second.
+The simulation took 11.55 s and 12.48 s using the orginial and new temporal schemes, respectively.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{AMR}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[htb]
-\begin{center}
-\includegraphics[width=4.0in]{./figs/wdconvect_amr_Tmax} 
-\caption{\label{fig:wdconvect_amr_Tmax} Peak temperature, $T_{\text{peak}}$, in a white dwarf from $t=0$ to $400$ s 
-         for grids with effective resolution of $512^3$. We can see that the adaptive grids with two levels of 
-         refinement give very similar solution trends compared to the single-level grid.}
-\end{center}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Full white dwarf star simulation at effective resolution of $512^3$ on Cori KNL with 512 processors and 16 threads took 26.01 s per time step (averaged over 10 time steps) using the original temporal scheme, and 22.66 s using the new scheme.
-On an adaptive grid with two levels of refinement, 8.9\% of the cells (1,493,504 out of $256^3$ cells) are refined at the first level and 5.3\% (7,176,192 out of $512^3$ cells) at the second.
-The simulation took 11.55 s and 12.48 s using the orginial and new temporal schemes, respectively.

### 2019-06-06 MAESTROeX_I/results.tex
Updated results section; still need to finish a subsection

@@
 
-\subsection{Performance and Scaling}
+\subsection{Performance and Scaling}\label{sec:scaling}
 We perform weak scaling tests for simulations of convection preceding ignition in a spherical, full-star sub-Chandrasekhar mass white dwarf.
@@
 
+\subsection{White Dwarf Convection}
+
+white dwarf convection runs with 3 algorithms (original, new temporal, new temporal + irregularly-spaced base state)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=2.75in]{./figs/wdconvect_256_maxT}  \hspace{0.5in}
+\includegraphics[width=2.75in]{./figs/wdconvect_256_maxMach}
+\caption{\label{fig:wdconvect_128_maxval} (left) Peak temperature, $T_{\text{peak}}$, and (right) peak Mach number 
+         in a white dwarf until time of ignition at resolution of $256^3$ for three different MAESTROeX algorithms.}
+\end{center}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[hbt]
+\begin{center}
+\includegraphics[width=3.25in]{./figs/wdconvect_compare_Tmax}
+\caption{\label{fig:wdconvect_Tmax} Peak temperature, $T_{\text{peak}}$  in a white dwarf at grid resolutions of 
+         $256^3$ (dotted line) and $512^3$ (solid line) until $t=1000$ for uniform (d$x$) and irregular (d$r$) base state spacing. 
+         Note that the irregularly-spaced solution agrees better 
+         with the uniformly-spaced solution as the resolution increases.}
+\end{center}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{AMR Performance}
@@
 
+To test the performance of MAESTROeX for adaptive, three-dimensional simulations, we make use of the AMReX framework to track a localized feature over time.
+Figure \ref{fig:amr_grids} illustrates the initial grid structures with two levels of refinement for both planar and spherical geometries where the grid is refined according to the temperature and density profiles, respectively.
+For each of the problems we tested, the single-level simulation was run using the original temporal scheme and the adaptive simulations using the original and new temporal algorithms.
+We want to show that the adaptive simulation can give the similar results as the single-level simulation and in a more computationally efficient manner.
+
+In the planar case, we use the same problem setup for a hot bubble rising in a white-dwarf environment as described in Section 6 of Paper V.
+Here we use a domain size of $3.6\times 10^7$ cm by $3.6\times 10^7$ cm by $2.88\times 10^8$ cm, and allow the grid structure to change with time.
+The single-level simulation at a resolution of $128^2 \times 1024$ was run on Cori Haswell with 48 processors and took approximately 33.5 s per time step (averaged over 10 time steps) using either the original or new temporal algorithm.
+The adaptive simulation has a resolution of $32^2 \times 256$ at the coarsest level, resulting in the same effective resolution at the finest level as the single-level simulation.
+We tag cells that satisfy $T-\bar{T} > 3\times 10^7$ K as well as all cell at that height.
+The adaptive run took only 3.7 s per time step, and this 89\% decrease in runtime is mostly due to the fact that initially only 6.25\% of the cells (1,048,576 out of $128^2 \times 1024$ cells) are refined at the finest level.
+Figure \ref{fig:bubble_results} shows a series of planar slices of the temperature profile at time intervals of 1.25 s, and verifies that the adaptive simulation is able to capture the same dynamics as the single-level simulation at much lower computational cost.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Reacting bubble simulation at effective resolution of $128^2 \times 1024$ run on Cori Haswell with 48 processors took approximately 33.5 s per time step (averaged over 10 time steps) to run on single-level grid using either the original or new temporal algorithm, and only 3.7 s on adaptive grid with two levels of refinement.
-This 89\% decrease in runtime is mostly due to the fact that only 6.25\% of the cells (1,048,576 out of $128^2 \times 1024$ cells) are refined at the finest level for the AMR run.
+We continue to use the full-star white dwarf problem described in Section \ref{sec:scaling} to test adaptive simulations on spherical geometry.
+The simulation on a single-level grid of $512^3$ resolution took 26.01 s per time step (again averaged over 10 time steps) using the original temporal scheme, and 22.66 s using the new scheme.
+On the adaptive grid with two levels of refinement, 8.9\% of the cells (1,493,504 out of $256^3$ cells) are refined at the first level and 5.3\% (7,176,192 out of $512^3$ cells) at the second.
+The adaptive simulation took 11.55 s and 12.48 s using the orginal and new temporal schemes, respectively.
+Both simulations are computed to $t=400$ s and we choose to use the peak temperature as the diagnostics to compare the results.
+Figure \ref{fig:wdconvect_amr_Tmax} shows the evolution of the peak temperature for all three runs and shows that the adaptive simulation gives the same qualitative result as the single-level simulation.
+We do not expect the curves to match up exactly because the governing equations are highly nonlinear, and slight differences in the solution caused by solver tolerance and discretization error can change the details of the results.
+Each simulation was run on Cori KNL with 512 processors and 16 threads.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Full white dwarf star simulation at effective resolution of $512^3$ on Cori KNL with 512 processors and 16 threads took 26.01 s per time step (averaged over 10 time steps) using the original temporal scheme, and 22.66 s using the new scheme.
-On an adaptive grid with two levels of refinement, 8.9\% of the cells (1,493,504 out of $256^3$ cells) are refined at the first level and 5.3\% (7,176,192 out of $512^3$ cells) at the second.
-The simulation took 11.55 s and 12.48 s using the orginial and new temporal schemes, respectively.
-
-\subsection{White Dwarf Convection}
-
-white dwarf convection runs with 3 algorithms (original, new temporal, new temporal + irregularly-spaced base state)
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[htb]
-\begin{center}
-\includegraphics[width=2.75in]{./figs/wdconvect_128_maxT}  \hspace{0.5in}
-\includegraphics[width=2.75in]{./figs/wdconvect_128_maxMach}
-\caption{\label{fig:wdconvect_128_maxval} (left) Peak temperature, $T_{\text{peak}}$, and (right) peak Mach number 
-         in a white dwarf until time of ignition at resolution of $128^3$ for three different MAESTROeX algorithms.}
-\end{center}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[hbt]
-\begin{center}
-\includegraphics[width=2.7in]{./figs/wdconvect_256_maxT}  \hspace{0.5in}
-\includegraphics[width=2.75in]{./figs/wdconvect_512_maxT}
-\caption{\label{fig:wdconvect_Tmax} Peak temperature, $T_{\text{peak}}$  in a white dwarf at 
-         (left) resolution of $256^3$ until time of ignition and (right) resolution of $512^3$ until $t=1000$ 
-         for three different MAESTROeX algorithms. Note that the irregularly-spaced (d$r$) solution agrees better 
-         with the uniformly-spaced solution as the resolution increases.}
-\end{center}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-06-06 MAESTROeX_I/results.tex
Finished results section

@@
 
-\subsection{White Dwarf Convection}
+\subsection{White Dwarf Convection}\label{sec:whitedwarf}
 
-white dwarf convection runs with 3 algorithms (original, new temporal, new temporal + irregularly-spaced base state)
+To explore the accuracy and efficiency of the new temporal algorithm, we now analyze in detail three-dimensional, full-star calculations of convection preceding ignition in a sub-Chandrasekhar mass white dwarf.
+Again, we refer the reader to \cite{MAESTRO_AMR} and \cite{MAESTRO_convection} for setup details.
+We implement both uniformly- and irregularly-spaced base state with the new temporal algorithm, while only uniform base state spacing is used in the original algorithm.
+As in Section 3 of \cite{MAESTRO_AMR}, we choose the peak temperature and peak Mach number as the two diagnostics to compare the simulations.
+Figure \ref{fig:wdconvect_256_maxvar} shows the evolution of both peak temperature and peak Mach number until time of ignition on a single-level grid with resolution of $256^3$.
+The simulation using the new temporal scheme with uniformly-spaced base state gives the same qualitative results as the original scheme, and predicts very similar time of ignition ($t=7810$ s compared to $t=7850$ s for original algorithm).
+The simulation using the new temporal scheme with irregularly-spaced base state displays a slightly different peak temperature behavior during the initial transition period $t<150$ s, which results in the difference between the curves post transition.
+We strongly suspect that this is a result of using different initial model inputs for the star in which the resolution near the center of the star is much coarser with the irregular spacing than the uniform spacing.
+Fortunately, the simulation with irregular base state spacing still follows the same trend as with uniform spacing, and the star is shown to ignite at an earlier time $t=6840$ s.
+
+Figure \ref{fig:wdconvect_Tmax} shows the peak temperature evolution over the first 1000 s on two grids of differing resolutions, $256^3$ and $512^3$.
+As previously suspected, the simulation with irregularly-spaced base state agrees much closer to the results from using uniform spacing as the resolution is increased.
+This is most likely due to the increased resolution of the initial star models for the irregular-spaced base state that is now able to properly capture the same properties near the center of the star as its uniformly-spaced counterpart.
+This is especially important when computing the base state pressure from base state density since we enforce a constant base state pressure outside of a prescribed cutoff density radius that could potentially be quite different with the two types of base state spacing.
 
@@
 \includegraphics[width=2.75in]{./figs/wdconvect_256_maxMach}
-\caption{\label{fig:wdconvect_128_maxval} (left) Peak temperature, $T_{\text{peak}}$, and (right) peak Mach number 
+\caption{\label{fig:wdconvect_256_maxvar} (left) Peak temperature, $T_{\text{peak}}$, and (right) peak Mach number 
          in a white dwarf until time of ignition at resolution of $256^3$ for three different MAESTROeX algorithms.}
@@
 
+In terms of efficiency, all three simulations on $256^3$ single-level grid were run on Cori haswell with 64 processors and 8 threads per core.
+The simulation using the new temporal algorithm took 6.75 s per time step with uniformly-spaced base state, which is 13\% faster than the 7.77 s per time step when using the original scheme, and 9.72 s with irregularly-spaced base state.
+The 25\% increase in run time of the irregularly-spaced algorithm is mainly caused by the much larger base state array and could be significantly reduced if we consider truncating the base state beyond the cutoff density radius.
+
 \subsection{AMR Performance}
@@
 
-To test the performance of MAESTROeX for adaptive, three-dimensional simulations, we make use of the AMReX framework to track a localized feature over time.
+We now test the performance of MAESTROeX for adaptive, three-dimensional simulations to track a localized feature over time.
 Figure \ref{fig:amr_grids} illustrates the initial grid structures with two levels of refinement for both planar and spherical geometries where the grid is refined according to the temperature and density profiles, respectively.
@@
 
-We continue to use the full-star white dwarf problem described in Section \ref{sec:scaling} to test adaptive simulations on spherical geometry.
-The simulation on a single-level grid of $512^3$ resolution took 26.01 s per time step (again averaged over 10 time steps) using the original temporal scheme, and 22.66 s using the new scheme.
-On the adaptive grid with two levels of refinement, 8.9\% of the cells (1,493,504 out of $256^3$ cells) are refined at the first level and 5.3\% (7,176,192 out of $512^3$ cells) at the second.
-The adaptive simulation took 11.55 s and 12.48 s using the orginal and new temporal schemes, respectively.
+We continue to use the full-star white dwarf problem described in Section \ref{sec:whitedwarf} to test adaptive simulations on spherical geometry.
+The adaptive grid is refined twice by tagging the density at $\rho < 5\times 10^7$ g cm$^{-3}$ on the first level and $\rho < 10^8$ g cm$^{-3}$ on the second level.
+The simulation on a single-level grid of $512^3$ resolution took 26.01 s per time step (again averaged over 10 time steps).
+The adaptive grid has a resolution of $128^3$ at the coarsest level and an effective resolution of $512^3$ at the finest level.
+On this grid, 8.9\% of the cells (1,493,504 out of $256^3$ cells) are refined at the first level and 5.3\% (7,176,192 out of $512^3$ cells) at the second.
+The adaptive simulation took 11.55 s, resulting in more than a factor of 2 in speedup.
 Both simulations are computed to $t=400$ s and we choose to use the peak temperature as the diagnostics to compare the results.
@@
 We do not expect the curves to match up exactly because the governing equations are highly nonlinear, and slight differences in the solution caused by solver tolerance and discretization error can change the details of the results.
-Each simulation was run on Cori KNL with 512 processors and 16 threads.
+Each simulation was run on Cori KNL with 512 processors and 16 threads per core.
 

### 2019-06-06 MAESTROeX_I/algorithm.tex
Updated results section; still need to finish a subsection

@@
 Our approach to adaptive mesh refinement is algorithmically the same as the treatment described in Section 5 of Paper V; we refer the reader there for details.
-We note that for spherical problems, AMR is only available for the case of a regularly-spaced base state.
+We note that for spherical problems, AMR is only available for the case of a uniformly-spaced base state.

### 2019-06-06 MAESTROeX_I/appendix.tex

@@
 \begin{center}
-\begin{tabular}{l r}
-\multirow{4}{3.5in}[34.5mm]{ \includegraphics[width=3.0in]{./figs/wdconvect_256_splitU} } & \multicolumn{1}{c}{\includegraphics[width=2.35in]{./figs/magvel_full_XY}} \\
-& \multicolumn{1}{c}{\begin{footnotesize} (a) $|\mathbf{U}|$, solved using $\mathbf{U}$ \end{footnotesize}} \\[1.em]
-& \multicolumn{1}{c}{\includegraphics[width=2.35in]{./figs/magvel_split_XY}} \\ 
-& \multicolumn{1}{c}{\begin{footnotesize} (b) $|\mathbf{U}|$, solved using $\tilde{\mathbf{U}}+w_0$ \end{footnotesize}} \\
+\begin{tabular}{l c}
+\multirow{4}{3.25in}[30mm]{ \includegraphics[width=3.0in]{./figs/wdconvect_256_splitU} } & \multicolumn{1}{c}{\includegraphics[width=2.35in]{./figs/magvel_full_XY}} \\
+& \multicolumn{1}{c}{\begin{footnotesize} (a) Velocity magnitude, solved using full $\mathbf{U}$ \end{footnotesize}} \\[1.em]
+& \multicolumn{1}{c}{\includegraphics[width=2.75in]{./figs/magvel_lineout_X}} \\ 
+& \multicolumn{1}{c}{\begin{footnotesize} (b) Comparison along white line \end{footnotesize}} \\
 \end{tabular}
@@
          during initial transition. (Right) Velocity magnitude solved using full velocity $\mathbf{U}$ in the projection step 
-         results in large values outside the density cutoff region that is not seen when we split the velocity dynamics.}
+         results in much larger values outside the density cutoff region, $\rho < 10^5 \text{ g cm}^{-3}$, that is not seen when 
+         we split the velocity dynamics.}
 \end{center}

### 2019-06-06 MAESTROeX_I/conclusions.tex
add our grants

@@
 In particular, the AMReX-based companion code, the compressible CASTRO code \citep{CASTRO}, has recently ported hydrodynamics and reactions to GPUs \citep{CASTRO_GPU}.
-We plan to leverage the newly implemented mechanisms for offloading compute kernels to GPUs inside of the AMReX software library itself.ment library itself.
+We plan to leverage the newly implemented mechanisms for offloading compute kernels to GPUs inside of the AMReX software library itself.
 Our future scientific investigations include convection in massive rotating stars, \citep{heger2000presupernova}, the convective Urca process in white dwarfs \citep{willcox2016type}, solar physics \citep{wood2018self}, and magnetohydrodynamics \citep{wood2015three,wood2011sun}.
@@
 
-The work at LBNL was supported by the U.S. Department of Energy's Scientific Discovery Through Advanced Computing (SciDAC) program under contract No. DE-AC02-05CH11231.
-This research used resources of the National Energy Research Scientific Computing Center (NERSC), a U.S.
-Department of Energy Office of Science User Facility operated under Contract No.
-DE-AC02-05CH11231.
+The work at LBNL was supported by the U.S. Department of Energy's
+Scientific Discovery Through Advanced Computing (SciDAC) program under
+contract No. DE-AC02-05CH11231.  The work at Stony Brook was supported
+by DOE/Office of Nuclear Physics grant DE-FG02-87ER40317 and through
+the SciDAC program DOE grant DE-SC0017955.  This research used
+resources of the National Energy Research Scientific Computing Center
+(NERSC), a U.S. Department of Energy Office of Science User Facility
+operated under Contract No. DE-AC02-05CH11231.
+
+\software{AMReX \citep{AMReX, AMReX_JOSS}, StarKiller Microphysics \citep{starkiller}}
+\facilities{NERSC}

### 2019-06-06 MAESTROeX_I/results.tex
fixed typo

@@
 We continue to use the full-star white dwarf problem described in Section \ref{sec:whitedwarf} to test adaptive simulations on spherical geometry.
-The adaptive grid is refined twice by tagging the density at $\rho < 5\times 10^7$ g cm$^{-3}$ on the first level and $\rho < 10^8$ g cm$^{-3}$ on the second level.
+The adaptive grid is refined twice by tagging the density at $\rho > 5\times 10^7$ g cm$^{-3}$ on the first level and $\rho > 10^8$ g cm$^{-3}$ on the second level.
 The simulation on a single-level grid of $512^3$ resolution took 26.01 s per time step (again averaged over 10 time steps).

### 2019-06-06 MAESTROeX_I/results.tex
change line colors in figure 5 to match figure 4

@@
 Figure \ref{fig:wdconvect_256_maxvar} shows the evolution of both peak temperature and peak Mach number until time of ignition on a single-level grid with resolution of $256^3$.
-The simulation using the new temporal scheme with uniformly-spaced base state gives the same qualitative results as the original scheme, and predicts very similar time of ignition ($t=7810$ s compared to $t=7850$ s for original algorithm).
+The simulation using the new temporal scheme with uniformly-spaced base state gives the same qualitative results as the original scheme, and predicts a similar time of ignition ($t=7810$ s compared to $t=7850$ s for original algorithm).
 The simulation using the new temporal scheme with irregularly-spaced base state displays a slightly different peak temperature behavior during the initial transition period $t<150$ s, which results in the difference between the curves post transition.
@@
 Figure \ref{fig:wdconvect_Tmax} shows the peak temperature evolution over the first 1000 s on two grids of differing resolutions, $256^3$ and $512^3$.
-As previously suspected, the simulation with irregularly-spaced base state agrees much closer to the results from using uniform spacing as the resolution is increased.
-This is most likely due to the increased resolution of the initial star models for the irregular-spaced base state that is now able to properly capture the same properties near the center of the star as its uniformly-spaced counterpart.
+As previously suspected, the simulation using irregularly-spaced base state agrees much closer to the results from using uniform spacing as the resolution is increased.
+This is most likely due to the increased resolution of the initial star model for the irregular-spaced base state that is now able to properly capture the same properties near the center of the star as its uniformly-spaced counterpart.
 This is especially important when computing the base state pressure from base state density since we enforce a constant base state pressure outside of a prescribed cutoff density radius that could potentially be quite different with the two types of base state spacing.
@@
 The adaptive simulation has a resolution of $32^2 \times 256$ at the coarsest level, resulting in the same effective resolution at the finest level as the single-level simulation.
-We tag cells that satisfy $T-\bar{T} > 3\times 10^7$ K as well as all cell at that height.
+We tag cells that satisfy $T-\bar{T} > 3\times 10^7$ K as well as all cells at that height.
 The adaptive run took only 3.7 s per time step, and this 89\% decrease in runtime is mostly due to the fact that initially only 6.25\% of the cells (1,048,576 out of $128^2 \times 1024$ cells) are refined at the finest level.
@@
 The adaptive simulation took 11.55 s, resulting in more than a factor of 2 in speedup.
-Both simulations are computed to $t=400$ s and we choose to use the peak temperature as the diagnostics to compare the results.
+Both simulations are computed to $t=400$ s and we choose to use the peak temperature as the diagnostic to compare the results.
 Figure \ref{fig:wdconvect_amr_Tmax} shows the evolution of the peak temperature for all three runs and shows that the adaptive simulation gives the same qualitative result as the single-level simulation.

### 2019-06-07 MAESTROeX_I/intro.tex
sec 1 updates

@@
 number of approaches to modeling low Mach number flows have been
-developed in recent years.  These include methods atmospheric model similar to
-what is presented here~\cite{Lin:2006}.  There are also  semi-implicit all-Mach
+developed in recent years.  One of the more similar approaches to ours is
+\cite{Lin:2006}, however this approach is only first-order accurate and does
+not account for atmospheric expansion.  There are also semi-implicit all-Mach
 number solvers, where the Euler equations are split into an acoustic
@@
 Previously, we developed the low Mach number astrophysical solver, MAESTRO.
-MAESTRO is a block-structured, Cartesian grid finite-volume, adaptive mesh refinement (AMR) code that has been successfully used for many years for a number of applications, detailed below.
-Unlike several of the references above, MAESTRO is not an all-Mach solver, but is suitable for flows where the Mach number is small ($\sim 0.1$ or smaller).  
-Furthermore, the low Mach number model in MAESTRO is specifically designed for, but not limited to, astrophysical settings with significant atmospheric stratification.
-This includes full spherical stars, as well as planar simulations of dynamics within localized regions of a star.
-MAESTRO uses an explicit Godunov approach for advection, a stiff ODE package (VODE) \citep{vode} for reactions, and multigrid-based linear solvers for the pressure-projection steps, and is thus time-step limited by an advective CFL constraint based on the fluid velocity, not the sound speed.
-The time-advancement strategy uses Strang splitting to couple the thermodynamic variables, and a second-order projection method to integrate the velocity equations subject to a divergence constraint.
-Central to the algorithm is a time-varying, one-dimensional stratified background (or base) state density and pressure held in hydrostatic equilibrium that couple to the full Cartesian grid solution.
-The original MAESTRO code was developed in the pure-Fortran 90 FBoxLib software framework (available in the same github organization as AMReX).
+MAESTRO is a block-structured, Cartesian grid finite-volume, adaptive mesh refinement (AMR) 
+code that has been successfully used for many years for a number of applications, detailed below.
+Unlike several of the references above, MAESTRO is not an all-Mach solver, but is suitable for 
+flows where the Mach number is small ($\sim 0.1$ or smaller).  
+Furthermore, the low Mach number model in MAESTRO is specifically designed for, but not limited 
+to, astrophysical settings with significant atmospheric stratification.
+This includes full spherical stars, as well as planar simulations of dynamics within localized 
+regions of a star.
+The numerical methodology relies on an explicit Godunov approach for advection, a stiff ODE 
+solver for reactions (VODE) \citep{vode}, and multigrid-based linear solvers for the 
+pressure-projection steps.  This, the time step is limited by an advective CFL constraint based on the 
+fluid velocity, not the sound speed.
+Central to the algorithm are time-varying, one-dimensional stratified background (or base) state 
+density and pressure fields that are held in hydrostatic equilibrium.
+The base state density couples to the full state solution through buyoancy terms in the momentum equation,
+and the base state pressure couples to the full state solution by constraining the evolution of the
+thermodynamic variables to match this pressure.
+The time-advancement strategy uses a predictor-corrector strategy for evolving the base state,
+Strang splitting to integrate the thermodynamic variables, and a 
+second-order projection method to integrate the velocity equations subject to a divergence constraint.
+The original MAESTRO code was developed in the pure-Fortran 90 FBoxLib software framework, whereas
+MAESTROeX is developed in the C++/F90 AMReX framework \citep{AMReX,AMReX_JOSS}.
 
-The key numerical developments of the original MAESTRO algorithm are presented in a series of papers which we refer to as Papers I-V:
+The key numerical developments of the original MAESTRO algorithm are presented in a series of 
+papers which we refer to as Papers I-V:
 \begin{itemize}
-\item In Paper I \citep{MAESTRO_I}, we derive the low Mach number equation set for stratified environments from the fully compressible equations.
-\item In Paper II \citep{MAESTRO_II}, we incorporate the effects of atmospheric expansion through the use of a time-dependent background state.
-\item In Paper III \citep{MAESTRO_III}, we incorporate reactions and the associated coupling to the hydrodynamics.
-\item In Paper IV \citep{MAESTRO_IV}, we describe our treatment of spherical stars in a three-dimensional Cartesian geometry.
-\item In Paper V \citep{MAESTRO_V}, we describe the use of block-structured adaptive mesh refinement to focus spatial resolution in regions of interest.
+\item In Paper I \citep{MAESTRO_I}, we derive the low Mach number equation set for stratified 
+environments from the fully compressible equations.
+\item In Paper II \citep{MAESTRO_II}, we incorporate the effects of atmospheric expansion 
+through the use of a time-dependent background state.
+\item In Paper III \citep{MAESTRO_III}, we incorporate reactions and the associated coupling 
+to the hydrodynamics.
+\item In Paper IV \citep{MAESTRO_IV}, we describe our treatment of spherical stars in a 
+three-dimensional Cartesian geometry.
+\item In Paper V \citep{MAESTRO_V}, we describe the use of block-structured adaptive mesh 
+refinement to focus spatial resolution in regions of interest.
 \end{itemize}
@@
 In this paper, we present new algorithmic methodology that improves upon Paper V in a number of ways.
-First, the overall temporal algorithm has been greatly simplified without compromising accuracy.
+First, the overall temporal algorithm has been greatly simplified without compromising second-order accuracy.
 The key design decisions were to eliminate the splitting of the velocity into average and perturbational components, 
@@
 based on method-of-lines integration.  
-In particular, schemes based on spectral deferred corrections have been used to generate high-order temporal integrators for problems of reactive flow \citep{dutt2000spectral} and low Mach number combustion \citep{pazner2016high,nonaka2018conservative}.
-Second, we explore the effects of alternative spatial mapping routines for coupling the base state and the Cartesian grid state in spherical problems.
-Finally, the code has been completely ported to a new software framework suitable for modern manycore and GPU-based supercomputers.
-We examine the performance of our new MAESTROeX implementation in the new C++/F90 AMReX public software library \citep{AMReX,AMReX_JOSS}.
+In particular, schemes based on deferred corrections have been used to generate high-order temporal integrators for problems of reactive flow \citep{dutt2000spectral} and low Mach number combustion \citep{pazner2016high,nonaka2018conservative}.
+Second, we explore the effects of alternative spatial mapping routines for coupling the base state and the Cartesian grid state for spherical problems.
+Finally, we examine the performance of our new MAESTROeX implementation in the new C++/F90 AMReX public software library \citep{AMReX,AMReX_JOSS}.
 MAESTROeX uses MPI+OpenMP parallelism and scales well to over 10,000 MPI processes, with each MPI process supporting tens of threads.
 The resulting code is publicly available on GitHub ({\tt https://github.com/AMReX-Astro/MAESTROeX}),
-uses the Starkiller-Astro microphysics libraries \citep{starkiller},
-and the AMReX software library for block-structured adaptive mesh refinement ({\tt https://github.com/AMReX-Codes/amrex}).
+uses the Starkiller-Astro microphysics libraries \citep{starkiller} ({\tt https://github.com/starkiller-astro}) ,
+as well as AMReX ({\tt https://github.com/AMReX-Codes/amrex}).
 

### 2019-06-07 MAESTROeX_I/paper.tex
Added figures to appendix A
sec 1 updates

@@
 \usepackage{mathtools}
+
+\usepackage{multirow}
 
@@
 We also explore the use of alternative spatial mapping of the one-dimensional base state onto the full Cartesian grid.
-The code leverages the new AMReX software framework for block-structured adaptive mesh refinement applications, allowing for scalability to large fractions of leadership-class machines.
+The code leverages the new AMReX software framework for block-structured adaptive mesh refinement (AMR) applications, allowing for scalability to large fractions of leadership-class machines.
 Using our previous studies on the convective phase of single-degenerate progenitor models of Type Ia supernovae as a guide, we characterize the performance of the code and validate the new algorithmic features.
+Like MAESTRO, MAESTROeX is fully open source.
 \end{abstract}

### 2019-06-07 MAESTROeX_I/equations.tex
sec 2 readthrough

@@
 In Papers I-III, we extended this work and the atmospheric model by \citet{durran:1989} by deriving a model and algorithm suitable for stratified astrophysical flow.
-In our model, we take the standard equations of reacting, compressible flow, and recast the equation
-of state (EOS) as a divergence constraint on the velocity field.
+We take the standard equations of reacting, compressible flow, and recast the equation
+of state (EOS) as a divergence constraint on the velocity field, as explained below.
 The resulting model is a series of evolution equations for mass, momentum, and energy, subject
@@
 \end{equation}
-Note that the momentum equation above enforces conservation of total energy in the 
-low Mach number system in the absence of external heating or viscous terms \citep{kleinpauluis,Vasil2013}, 
-and has been previously demonstrated in modeling sub-Chandrasekhar white dwarfs using MAESTRO \citep{subChandra_II}. 
 The total pressure is decomposed into a one-dimensional hydrostatic base state
@@
 the base state).  
+By comparing the momentum equation (\ref{eq:momentum}) to the momentum equation used in equation (2) in Paper V, we
+note that we are using a formulation that enforces conservation of total energy in the 
+low Mach number system in the absence of external heating or viscous terms \citep{kleinpauluis,Vasil2013}.
+We have previously validated this approach in modeling sub-Chandrasekhar white dwarfs using MAESTRO \citep{subChandra_II}. 
 We also define a one-dimensional base state density, $\rho_0 = \rho_0(r,t)$, that represents the lateral average (see Section \ref{Sec:Spatial}) of $\rho$ and is in hydrostatic equilibrium with $p_0$, i.e.,
@@
 where $g=g(r,t)$ is the magnitude of the gravitational acceleration and $\eb_r$ is the unit vector in the outward radial direction. 
+Here $\beta_0$ is a density-like variable that carries background stratification, defined as
+\begin{equation}
+\beta_0(r,t) = \rho_0(0,t)\exp\left(\int_0^r\frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial r'}dr'\right),
+\end{equation}
+where $\gammaonebar$ is the lateral average of $\Gamma_1 = d(\log p)/d(\log\rho) |_s$ (evaluated with entropy, $s$, constant).
 Thermal diffusion is not discussed in this paper, but we have previously described the modifications to the original algorithm required 
@@
 
-Mathematically, equations \ref{eq:momentum}--\ref{eq:enthalpy} must still be closed by the EOS.  
+Mathematically, equations (\ref{eq:momentum})-(\ref{eq:enthalpy}) must still be closed by the EOS.  
 This is done by taking the Langrangian derivative of the EOS for pressure as a function of the thermodynamic variables, 
@@
 \end{equation}
-Here $\beta_0$ is a density-like variable that carries background stratification, defined as
-\begin{equation}
-\beta_0(r,t) = \rho_0(0,t)\exp\left(\int_0^r\frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial r'}dr'\right),
-\end{equation}
-where $\gammaonebar$ is the lateral average of $\Gamma_1 = d(\log p)/d(\log\rho) |_s$ (evaluated with entropy, $s$, constant).
 The expansion term, $S$, incorporates local compressibility effects due to compositional changes and heat release from reactions,
@@
 
-To summarize, we model evolution equations for momentum, mass, and energy, equations \ref{eq:momentum}--\ref{eq:enthalpy} subject to a divergence constraint on the velocity, equation \ref{eq:U divergence}, and the hydrostatic equilibrium condition, equation \ref{eq:HSE}.
+To summarize, we model evolution equations for momentum, mass, and energy, equations (\ref{eq:momentum})-(\ref{eq:enthalpy}) subject to a divergence constraint on the velocity, equation (\ref{eq:U divergence}), and the hydrostatic equilibrium condition, equation (\ref{eq:HSE}).

### 2019-06-07 MAESTROeX_I/algorithm.tex
sec 3.1 fixes

--- 2019-06-05 MAESTROeX_I/algorithm.tex
@@
 We shall refer to local atmospheric flows in two and three dimensions as problems in ``planar'' geometry, and full-star flows
-in three dimensions as problems in ``spherical'' geometry. The solution in both cases consists of the Cartesian grid solution
+in three dimensions as problems in ``spherical'' geometry.
+The solution in both cases consists of the Cartesian grid solution
 and the one-dimensional base state solution.
@@
 (Left) For multi-level problems in planar geometry, we force a direct alignment
-between the radial array cell centers and the Cartesian grid cell centers by 
+between the base state cell centers and the Cartesian grid cell centers by 
 allowing the radial base state spacing to change with space and time.
 (Right) For multi-level problems in spherical geometry, since there is no direct alignment
-between the radial array cell centers and the Cartesian grid cell centers, we choose to fix
+between the base state cell centers and the Cartesian grid cell centers, we choose to fix
 the radial base state spacing across levels. Reprinted from Paper V \citep{MAESTRO_V}. }
@@
 One of the key numerical modules is the ``lateral average'', which computes the average over a 
-layer of a Cartesian grid variable onto a one-dimensional radial array.
-In planar geometries, this is a straightforward arithmetic average of cells at
-a particular height since the radial cell centers are in alignment
+layer of constant altitude of a Cartesian grid variable onto a one-dimensional base state.
+In planar geometries, this is a straightforward arithmetic average of values in cells at
+a particular height since the base state cell centers are in alignment
 with the Cartesian grid cell centers.
 However for spherical problems, the procedure is much more complicated.
-In Section 4 of Paper V, we describe how there is a finite, easily computable set of radii that any Cartesian cell-center can map to.  
-Specifically, for every Cartesian cell, there exists an integer $m$ such that the radius to the center of the star is given by 
-\begin{equation}\label{eqn:radii}
-\hat{r}_m=\Delta x\sqrt{0.75+2m}.
+In Section 4 of Paper V, we describe how there is a finite, easily computable set of radii that any 
+there-dimensional Cartesian cell-center can map to.  
+Specifically, for every three-dimensional Cartesian cell, there exists an integer $m$ such that the radius to the center of the star is given by 
+\begin{equation}
+\hat{r}_m=\Delta x\sqrt{0.75+2m}.\label{eqn:radii}
 \end{equation}
-We average by binning all the cells associated with each radii and taking the average.
-Then we interpolate this data onto a one-dimensional radial array.
+Figure \ref{Fig:NewBaseGrid} is a two-dimensional illustration 
+(two dimensions is chosen in the figure for ease of exposition; this mapping is only used for three-dimensional spherical stars) 
+of the relationship between the Cartesian grid state and the one-dimensional base state array.
+We compute the lateral average by first summing the values in all the cells associated with each radius,
+dividing by the number of contributing cells to obtain the arithmetic average, and then use quadratic interpolation to map this data onto a one-dimensional base state.
+Previously, for spherical problems MAESTRO only allowed for a base state with constant $\Delta r$ (typically equal to $\Delta x/5$).
 
-Previously, MAESTRO only allowed for radial arrays with a constant $\Delta r$ (typically equal to $\Delta x/5$).
-Here we present a new option to retain an irregularly-spaced radial array to eliminate mapping errors back onto the Cartesian grid state.
-The idea is to follow the same method performed on planar geometries by simply computing an arithmetic average of cells that share the same radial distance from the center of the star. 
-Figure \ref{Fig:NewBaseGrid} illustrates the relationship between the Cartesian grid state and the one-dimensional array base state with this new mapping technique.
-Again, these radial distances, and hence the cell-center locations in the new base state array, can be predetermined using equation (\ref{eqn:radii}).
-Also note that $\Delta r$ decreases as the base state moves further from the center of the star, which results in far more total cells in the irregularly-spaced array than the previous uniformly-spaced array.
-A potential benefit to eliminating the mapping error is to consider a spherical star in hydrostatic equlibrium at rest.  In the absense of reactions, the star should remain at rest.
-The buoyancy forcing term in the momentum equation contains $\rho-\rho_0$.
-With the original scheme, interpolation errors in computing $\rho_0$ by averaging would cause artificial acceleration in the velocity field due to the interpolation error from the Cartesian grid to and from the radial base state.
-By retaining the radial base state as an irregularly spaced array, the effects due to interpolation error are completely eliminated.
+The companion ``fill'' module maps a base state array onto the full Cartesian grid state.
+For planar problems, direct injection can be used due to the perfect alignment of the base state and Cartesian grid state.
+For spherical problems, quadratic interpolation of the base state is used to assign values to each Cartesian cell center.
+
+In this paper we explore a new option to retain an irregularly-spaced base state to eliminate mapping errors from the ``fill'' module.
+For the lateral average, as before we first sum the values in all the cells associated with each radius and divide
+by the number of contributing cells to obtain the arithmetic average.  However we do not map this onto a regularly spaced
+base state and retain the use of this irregularly-spaced base state.
+The advantage with this approach is that the fill module does not require any interpolation.
+A potential benefit to eliminating the mapping error is to consider a spherical star in hydrostatic equlibrium at rest.
+In the absense of reactions, the star should remain at rest.
+The buoyancy forcing term in the momentum equation contains $\rho-\rho_0$.  With the original scheme, interpolation errors 
+in computing $\rho_0$ by averaging would cause artificial acceleration in the velocity field due to the interpolation error 
+from the Cartesian grid to and from the radial base state.  By retaining the radial base state as an irregularly spaced 
+array, the effects due to interpolation error are nearly eliminated, but not completely eliminated since there are still 
+machine precision effects resulting from averaging a large number of numerical values.
+We note that $\Delta r$ decreases as the base state moves further from the center of the star, 
+which results in far more total cells in the irregularly-spaced array than the previous uniformly-spaced array.
 
@@
 \caption{\label{Fig:NewBaseGrid}  
-A direct mapping between the radial array cell centers (red squares) and the Cartesian grid cell centers (blue crosses) 
+A direct mapping between the base state cell centers (red squares) and the Cartesian grid cell centers (blue crosses) 
 is enforced by computing the average of the grid cell centers that share the same radial distance from 

### 2019-06-07 MAESTROeX_I/algorithm.tex
changes up to the 12 step description

@@
 \section{Numerical Algorithm}\label{eq:algorithm}
-
 \subsection{Spatial Discretization}\label{Sec:Spatial}
@@
 One of the key numerical modules is the ``lateral average'', which computes the average over a 
-layer of constant altitude of a Cartesian grid variable onto a one-dimensional base state.
+layer of constant radius of a Cartesian grid variable onto a one-dimensional base state.
 In planar geometries, this is a straightforward arithmetic average of values in cells at
@@
 \end{equation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[tb]
+\centering
+\includegraphics[height=2.5in]{./figs/base_spherical_new}
+\caption{\label{Fig:NewBaseGrid}  
+A direct mapping between the base state cell centers (red squares) and the Cartesian grid cell centers (blue crosses) 
+is enforced by computing the average of the grid cell centers that share the same radial distance from 
+the center of the star.}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 Figure \ref{Fig:NewBaseGrid} is a two-dimensional illustration 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}[tb]
-\centering
-\includegraphics[height=2.5in]{./figs/base_spherical_new}
-\caption{\label{Fig:NewBaseGrid}  
-A direct mapping between the base state cell centers (red squares) and the Cartesian grid cell centers (blue crosses) 
-is enforced by computing the average of the grid cell centers that share the same radial distance from 
-the center of the star.}
-\end{figure}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \subsection{Temporal Integration Scheme}\label{Sec:Temporal Integration Scheme}
@@
 \begin{equation}
-\Ub = \Ubt(\xb,t) + w_0(r,t)\eb_r.
+\Ub = \Ubt(\xb,t) + w_0(r,t)\eb_r,
 \end{equation}
+where $\eb_r$ is the normal vector in the outward radial direction.
 We used $w_0$ to provide an estimate of the base state density evolution over a time step.
@@
 Our new temporal integration scheme does not use this splitting, and is much simpler than the scheme from Paper V since we evolve the velocity directly rather than a more complex evolution equation for the perturbational velocity.
-Additionally, the new scheme uses a simpler predictor-corrector approach to the base state density and pressure that no longer requires evolution equations and the associated discretizations for the base state, greatly simplifying the algorithm while retaining the same overall second-order accuracy in time.
+Additionally, the new scheme uses a simpler predictor-corrector approach to the base state density and pressure that no 
+longer requires evolution equations and the associated numerical discretizations to update the base state, greatly 
+simplifying the algorithm while retaining the same overall second-order accuracy in time.
 
-At the beginning of each time step we have the cell-centered state,
-$(\Ub,\rho X_k,\rho h,\rho_0,p_0)^n$, and nodal state, $\pi^{n-\myhalf}$.
+At the beginning of each time step we have the cell-centered Cartesian grid state,
+$(\Ub,\rho X_k,\rho h)^n$, and nodal Cartesian grid state, $\pi^{n-\myhalf}$, and base state $(\rho_0,p_0)^n$.
 At any time, the associated density, composition, and enthalpy can be trivially computed using, e.g.,
@@
 \end{equation}
-Temperature is computed using the equation of state\footnote{As described in Paper V, for planar problems we compute temperature using $h$ instead of $p_0$, since we have successfully developed planar volume discrepancy schemes to effectively couple the enthalpy to the rest of the solution; see \cite{XRB_I}.
+Temperature is computed using the equation of state\footnote{As described in Paper V, for planar problems we compute temperature using $h$ instead of $p_0$, since we have successfully developed volume discrepancy schemes to effectively couple the enthalpy to the rest of the solution; see \cite{XRB_I}.
 We are still exploring this option for spherical stars.}, e.g.,
-\begin{equation}
-T = T(\rho,p_0,X_k),
-\end{equation}
-and ($\gammaonebar,\beta_0)$ are computed from $(\rho,p_0,X_k)$ (see Appendix A of Paper I and Appendix C of Paper III for details).
+$T = T(\rho,p_0,X_k)$, where $p_0$ has been mapped to the Cartesian grid using the fill module,
+and ($\gammaonebar,\beta_0)$ are similarly computed from $(\rho,p_0,X_k)$;
+see Appendix A of Paper I and Appendix C of Paper III for details on how these terms are computed.
 
-The overall flow of the algorithm is a second-order Strang splitting approach for the coupling of the reactions and advection of the thermodynamic variables.  
-We use a predictor-corrector approach within the Strang splitting scheme to achieve second-order accuracy in time.
-We integrate the velocity using a standard second-order projection method to enforce the divergence constraint.
+The overall flow of the algorithm begins with a second-order Strang splitting approach to integrate the advection-reaction system for 
+the thermodynamic variables $(\rho X_k, \rho h)$, followed by a second-order projection methodology to integrate the velocities subject to a divergence constraint.
+Within the thermodynamic variable update we use a predictor-corrector approach to achieve second-order accuracy in time.
 To summarize:
@@
 \item In {\bf Steps 2--4} we advect the thermodynamic variables over $\Delta t$.
-Specifically, we compute an estimate for the expansion term, $S$, project the face-centered, time-centered velocities so they satisfy the divergence constraint, and then advect the thermodynamic variables.
+Specifically, we compute an estimate for the expansion term, $S$, compute face-centered, time-centered velocities that satisfy the divergence constraint, and then advect the thermodynamic variables.
 \item In {\bf Step 5} we react the thermodynamic variables over $\Delta t/2$\footnote{After this step we could skip to the velocity advance in {\bf Steps 10--11}, however the overall scheme would be only first-order in time, so {\bf Steps 6-9} can be thought of as a trapezoidal corrector step.}.
 \item In {\bf Steps 6--8} we redo the advection in {\bf Steps 2--4} but are able to use the trapezoidal rule to time-center certain quantities such as $S$, $\rho_0$, etc.
-\item In {\bf Step 9} we redo the reactions from {\bf Step 5} using the improved results for the corrector advection step.
-\item In {\bf Steps 10--11} we and update the velocity, compute the expansion term, $S$, and project the cell-centered velocities.
+\item In {\bf Step 9} we redo the reactions from {\bf Step 5} beginning with the improved results from {\bf Steps 6--8}.
+\item In {\bf Steps 10--11} we advect the velocity, and then project these velocities so they satisfy the divergence constraint.
 \end{itemize}
@@
 \item {\bf Enforce HSE}$[\rho_0]\rightarrow[p_0]$ computes the base state pressure, $p_0$, from a base state density, $\rho_0$ by integrating the hydrostatic equilibrium condition in one dimension. 
-This follows equation (A10) in Paper V with the exception of computing a non-constant $\Delta r_{j+1/2} = r_{j+1} - r_j$ for cell-edge $j+1/2$ if using irregularly-spaced base states.
-The base state pressure remains equal to a constant value from the location of a prescribed cutoff density outward for the entire simulation.
+This follows equation (A10) in Paper V, noting that for the irregularly spaced base state case, $\Delta r$ is not constant, where $\Delta r_{j+1/2} = r_{j+1} - r_j$ for cell-edge $j+1/2$.  
+The base state pressure remains equal to a constant value at the location of a prescribed cutoff density outward for the entire simulation.
 \item {\bf React State}$[(\rho X_k)^{\rm in},(\rho h)^{\rm in},p_0]\rightarrow[(\rho X_k)^{\rm out},(\rho h)^{\rm out},(\rho\dot\omega),(\rho\Hnuc)]$ integrates the species and enthalpy due to reactions over $\Delta t/2$ by solving

### 2019-06-07 MAESTROeX_I/algorithm.tex
readthrough of sec 3

@@
 \begin{itemize}
-\item In {\bf Step 1} we react the thermodynamic variables over $\Delta t/2$.
+\item In {\bf Step 1} we react the thermodynamic variables over the first $\Delta t/2$ interval.
 \item In {\bf Steps 2--4} we advect the thermodynamic variables over $\Delta t$.
 Specifically, we compute an estimate for the expansion term, $S$, compute face-centered, time-centered velocities that satisfy the divergence constraint, and then advect the thermodynamic variables.
-\item In {\bf Step 5} we react the thermodynamic variables over $\Delta t/2$\footnote{After this step we could skip to the velocity advance in {\bf Steps 10--11}, however the overall scheme would be only first-order in time, so {\bf Steps 6-9} can be thought of as a trapezoidal corrector step.}.
+\item In {\bf Step 5} we react the thermodynamic variables over the second $\Delta t/2$ interval\footnote{After this step we could skip to the velocity advance in {\bf Steps 10--11}, however the overall scheme would be only first-order in time, so {\bf Steps 6-9} can be thought of as a trapezoidal corrector step.}.
 \item In {\bf Steps 6--8} we redo the advection in {\bf Steps 2--4} but are able to use the trapezoidal rule to time-center certain quantities such as $S$, $\rho_0$, etc.
@@
 
-The temporal integration scheme contains the following steps:
+Beginning with $(\Ub,\rho X_k,\rho h)^n$, $\pi^{n-\myhalf}$, and $(\rho_0,p_0)^n$,
+the temporal integration scheme contains the following steps:
 \begin{description}
@@
 %--------------------------------------------------------------------------
-\item[Step 1] {\em React the full state through the first time interval of $\dt / 2.$}
+\item[Step 1] {\em React the thermodynamic variables over the first $\Delta t / 2$ interval.}
 
@@
 
-\item[Step 2] {\em Compute the provisional time-centered expansion,
-    $S^{\nph,\star}$.}
+\item[Step 2] {\em Compute the time-centered expansion term, $S^{\nph,\star}$.}
 
@@
 %--------------------------------------------------------------------------
-\item[Step 3] {\em Construct the provisional time-centered advective velocity on
-edges, $\uadvone$.}
+\item[Step 3] {\em Construct a time-centered advective velocity on edges, $\uadvone$.}
 
@@
 a standard multidimensional corner transport upwind approach
-\citep{colella1990multidimensional,saltzman1994unsplit} with piecewise-parabolic
+\citep{colella1990multidimensional,saltzman1994unsplit} with the piecewise-parabolic method (PPM)
 one-dimensional tracing \citep{colella1984piecewise}.  The full details of this
@@
 %--------------------------------------------------------------------------
-\item[Step 4] {\em Advect the full state through a time interval of $\dt.$}
+\item[Step 4] {\em Advect the thermodynamic variables over a time interval of $\dt.$}
 
@@
   for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
-  As described in Paper V, we predict $\rho'^n=\rho^n-\rho_0^n$ and $X_k^n$ to edges
+  As described in Paper V, for robust numerical slope limiting we predict 
+  $\rho'^n=\rho^n-\rho_0^n$ and $X_k^n$ to edges
   and here we spatially interpolate $\rho_0^n$ to edges to assemble the fluxes.
@@
 \begin{equation}
-\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h) + \frac{\partial p_0}{\partial t} + (\Ub \cdot \eb_r) \frac{\partial p_0}{\partial r}.
+\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\rho h\Ub) + \frac{\partial p_0}{\partial t} + (\Ub \cdot \eb_r) \frac{\partial p_0}{\partial r}.
 \end{equation}
@@
 \begin{equation}
-\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\Ub \rho h) + \frac{\partial p_0}{\partial t} + \nabla \cdot (\Ub p_0) - p_0 \nabla \cdot \Ub.
+\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\rho h\Ub) + \frac{\partial p_0}{\partial t} + \nabla \cdot (\Ub p_0) - p_0 \nabla \cdot \Ub.
 \end{equation}
@@
   for the conservative update of $(\rho h)^{(1)}$ using using a Godunov approach \citep{XRB_III}.
-  As described in Paper V, we predict $(\rho h)'^n=(\rho h)^n-(\rho h)_0^n$ to edges,
+  As described in Paper V, for robust numerical slope limiting 
+  we predict $(\rho h)'^n=(\rho h)^n-(\rho h)_0^n$ to edges,
   where $(\rho h)_0^n$ is obtained by calling {\bf Average}$[(\rho h)^n]\rightarrow[(\rho h)_0]$,
@@
 %--------------------------------------------------------------------------
-\item[Step 5] {\em React the full state through a second time interval of $\dt / 2.$}
+\item[Step 5] {\em React the thermodynamic variables over the second $\Delta t / 2$ interval.}
 
@@
 %--------------------------------------------------------------------------
-\item[Step 6] {\em Compute the time-centered expansion, $S^{\nph,\star}$.}
+\item[Step 6] {\em Compute the time-centered expansion term, $S^{\nph,\star}$.}
 
@@
 %--------------------------------------------------------------------------
-\item[Step 7] {\em Construct the time-centered advective velocity on edges, $\uadvtwo$.}
+\item[Step 7] {\em Construct a time-centered advective velocity on edges, $\uadvtwo$.}
 
@@
 %--------------------------------------------------------------------------
-\item[Step 8] {\em Advect the base state and full state through a time interval of $\dt.$}
+\item[Step 8] {\em Advect the thermodynamic variables over a time interval of $\dt.$}
 
@@
 %--------------------------------------------------------------------------
-\item[Step 9] {\em React the full state through a second time interval of $\dt / 2.$}
+\item[Step 9] {\em React the thermodynamic variables over the second $\Delta t / 2$ interval.}
 
@@
 %--------------------------------------------------------------------------
-\item[Step 10] {\em Define the new time expansion, $S^{n+1}$.}
+\item[Step 10] {\em Define the new-time expansion term, $S^{n+1}$.}
 

### 2019-06-07 MAESTROeX_I/results.tex
sec 4 readthrough

@@
 We see that the increase in wallclock time from the smallest to largest simulation is roughly 42\%.
-We also note that without base state evolution, the code runs 14\% faster for small problems, and scale mush better
-However, for the case without base state evolution, the overall runtime is faster and the increase in wallclock time from the smallest to largest simulation is roughly 13\%.
+We also note that without base state evolution, the code runs 14\% faster for small problems, and scales much better with wallclock time from the smallest to largest simulation increasing by only 13\%.
 This is quite remarkable since there are 3 linear solver per time step (2 cell-centered Poisson solves used in the MAC projection, and a nodal Poisson solve used to compute the updated cell-centered velocities).
@@
 Further profiling reveals that the primary scaling bottleneck is the average operator.
-The averaging operator requires binning the sum of Cartesian data onto one-dimensional arrays holding every possible mapping radius.
+The averaging operator requires collecting the sum of Cartesian data onto one-dimensional arrays holding every possible mapping radius.
 This amounts to at least 24,384 double precision values (for the $256^3$ simulation) up to 883,584 values (for the $1536^3$ simulation).
@@
 
-To explore the accuracy and efficiency of the new temporal algorithm, we now analyze in detail three-dimensional, full-star calculations of convection preceding ignition in a sub-Chandrasekhar mass white dwarf.
+To explore the accuracy of the new temporal algorithm, we now analyze in detail three-dimensional, full-star calculations of convection preceding ignition in a sub-Chandrasekhar mass white dwarf.
 Again, we refer the reader to \cite{MAESTRO_AMR} and \cite{MAESTRO_convection} for setup details.
@@
 Figure \ref{fig:wdconvect_Tmax} shows the peak temperature evolution over the first 1000 s on two grids of differing resolutions, $256^3$ and $512^3$.
+Limited allocations prevented us from running this simulation further.
 As previously suspected, the simulation using irregularly-spaced base state agrees much closer to the results from using uniform spacing as the resolution is increased.
 This is most likely due to the increased resolution of the initial star model for the irregular-spaced base state that is now able to properly capture the same properties near the center of the star as its uniformly-spaced counterpart.
-This is especially important when computing the base state pressure from base state density since we enforce a constant base state pressure outside of a prescribed cutoff density radius that could potentially be quite different with the two types of base state spacing.
+This is especially important when computing the base state pressure from base state density since we enforce a constant base state pressure outside of a prescribed cutoff density radius that could potentially be quite different with the two types of base state spacing at lower resolutions.
 
@@
 The 25\% increase in run time of the irregularly-spaced algorithm is mainly caused by the much larger base state array and could be significantly reduced if we consider truncating the base state beyond the cutoff density radius.
+\MarginPar{why is the new temporal with constant dr faster?}
 
@@
 
-We now test the performance of MAESTROeX for adaptive, three-dimensional simulations to track a localized feature over time.
+We now test the performance of MAESTROeX for adaptive, three-dimensional simulations to track localized regions of interest over time.
 Figure \ref{fig:amr_grids} illustrates the initial grid structures with two levels of refinement for both planar and spherical geometries where the grid is refined according to the temperature and density profiles, respectively.
@@
          $t = 0$, 1.25, and 2.5 s for (left) single-level simulation, and 
-         (right) adaptive simulation at the same effective resolution.}
+         (right) adaptive simulation at the same effective resolution.  The red, green, and black boxes indicate grids of increasing resolution}
 \end{center}

### 2019-06-07 MAESTROeX_I/algorithm.tex
sec 4 readthrough
typos, spell check

@@
 The advantage with this approach is that the fill module does not require any interpolation.
-A potential benefit to eliminating the mapping error is to consider a spherical star in hydrostatic equlibrium at rest.
-In the absense of reactions, the star should remain at rest.
+A potential benefit to eliminating the mapping error is to consider a spherical star in hydrostatic equilibrium at rest.
+In the absence of reactions, the star should remain at rest.
 The buoyancy forcing term in the momentum equation contains $\rho-\rho_0$.  With the original scheme, interpolation errors 
@@
 \end{equation}
-The inputs are the species, enthalpy, and base state pressure, and the outpus are the species, enthalpy, reaction rates, and nuclear energy generation rate.
+The inputs are the species, enthalpy, and base state pressure, and the outputs are the species, enthalpy, reaction rates, and nuclear energy generation rate.
 See Paper III for details.
@@
 where for our simulations we typically use $\sigma^{\rm CFL}\sim 0.7$ and the minimum is taken over all spatial directions over all cells.
-There are additional constraints on the timestep that are typically much less restrictive than the advective CFL including
+There are additional constraints on the time step that are typically much less restrictive than the advective CFL including
 the acceleration due to the buoyancy force (sometimes in effect when the velocity is approximately zero at the start of some simulations) 
@@
 Our approach to adaptive mesh refinement is algorithmically the same as the treatment described in Section 5 of Paper V; we refer the reader there for details.
+MAESTROeX supports refinement ratios of 2 between levels.
 We note that for spherical problems, AMR is only available for the case of a uniformly-spaced base state.

### 2019-06-07 MAESTROeX_I/appendix.tex
typos, spell check

@@
 \section{Projection Details}\label{Sec:Projection}
-To enforce the divergence constraint in {\bf Step 3, 7}, and {\bf 11}, we use a projection method analgous to the methods originally developed for incompressible flow \citep{almgren1998conservative,bell1989second}.
+To enforce the divergence constraint in {\bf Step 3, 7}, and {\bf 11}, we use a projection method analogous to the methods originally developed for incompressible flow \citep{almgren1998conservative,bell1989second}.
 The basic idea is to decompose the velocity field into a divergence-free component and a curl-free (gradient of a scalar field) component by solving a variable-coefficient Poisson equation for the scalar field.

### 2019-06-07 MAESTROeX_I/conclusions.tex
typos, spell check

@@
 Our future scientific investigations include convection in massive rotating stars, \citep{heger2000presupernova}, the convective Urca process in white dwarfs \citep{willcox2016type}, solar physics \citep{wood2018self}, and magnetohydrodynamics \citep{wood2015three,wood2011sun}.
-Our future algorithmic enhancements include more accurate and higher-order multiphysics coupling strategies byased on spectral deferred corrections \citep{dutt2000spectral,bourlioux2003high}.
+Our future algorithmic enhancements include more accurate and higher-order multiphysics coupling strategies based on spectral deferred corrections \citep{dutt2000spectral,bourlioux2003high}.
 This framework has been successfully used in terrestrial combustion \citep{pazner2016high,nonaka2018conservative}

### 2019-06-07 MAESTROeX_I/equations.tex
typos, spell check

@@
 Mathematically, equations (\ref{eq:momentum})-(\ref{eq:enthalpy}) must still be closed by the EOS.  
-This is done by taking the Langrangian derivative of the EOS for pressure as a function of the thermodynamic variables, 
+This is done by taking the Lagrangian derivative of the EOS for pressure as a function of the thermodynamic variables, 
 substituting in the equations of motion for mass and energy,

### 2019-06-07 MAESTROeX_I/intro.tex
typos, spell check

@@
 suitable scaling factor in the continuity equation, reducing the
-restriction on the size of the timestep
+restriction on the size of the time step
 \citep{Rempel2005,Hotta2012,Wang2015,Takeyama2017,Iijima2018}.
@@
 uses fully implicit time integration for the compressible Euler
-equations, which therefore allows for arbitrarily large timesteps.
+equations, which therefore allows for arbitrarily large time steps.
 
@@
 density and pressure fields that are held in hydrostatic equilibrium.
-The base state density couples to the full state solution through buyoancy terms in the momentum equation,
+The base state density couples to the full state solution through buoyancy terms in the momentum equation,
 and the base state pressure couples to the full state solution by constraining the evolution of the

### 2019-06-08 MAESTROeX_I/appendix.tex
appendix

@@
 To enforce the divergence constraint in {\bf Step 3, 7}, and {\bf 11}, we use a projection method analogous to the methods originally developed for incompressible flow \citep{almgren1998conservative,bell1989second}.
-The basic idea is to decompose the velocity field into a divergence-free component and a curl-free (gradient of a scalar field) component by solving a variable-coefficient Poisson equation for the scalar field.
+The basic idea is to decompose the velocity field into a part that satisfies the divergence-satisfying component and a curl-free (gradient of a scalar field) component by solving a variable-coefficient Poisson equation for the scalar field.
 The details for the MAC projection in {\bf Step 3} and {\bf Step 7} are given in Appendix B of Paper III.  The details of the nodal projection in {\bf Step 11} are given in Section 3.2 of Paper III.
@@
 Namely, to enforce that $r^2 w_0$ remain constant is difficult to do when solving for the full velocity.
-This is demonstrated in Figure \ref{fig:wdconvect_splitU} (right) where the velocity magnitude is observed to incorrectly increase outside the cutoff density radius when we use the full velocity in the nodal projection.
-The resulting peak temperature also dips significantly as seen in Figure \ref{fig:wdconvect_splitU} (left).
+This is demonstrated in Figure \ref{fig:wdconvect_splitU} (right) where the velocity magnitude is observed to incorrectly increase outside the cutoff density radius when we solve for the full velocity in the nodal projection.
+The resulting peak temperature also dips significantly as seen in Figure \ref{fig:wdconvect_splitU} (left), presumably because the dynamics of the overall expansion of the star are not being captured correctly.
 
@@
 
-So in practice, we solve the constraint over the lateral average,
+In practice, we solve the constraint over the lateral average,
 \begin{equation}
@@
 \end{equation}
-To solve for $\Ubt$ we use a projection method, which involves the solution of a variable-coefficient Poisson solver to extract the curl-free component of the unprojected velocity.
+To solve for $\Ubt$ we use a projection method, which involves the solution of a variable-coefficient Poisson solver to extract the curl-free component of the unprojected velocity, leaving a velocity field that satisfies the divergence constraint.
 Note that MAESTRO contains alternate low Mach number formulations that conserve total energy in stratified systems, with minimal changes to the code (see Appendix A of \cite{subChandra_II} for details).
-To find $w_0$ we integrate in 1D using the procedure in Appendix B of Paper V, keeping in mind that the base state spacing ($\Delta r$) should be computed using the appropriate cell-edge and cell-center locations when using irregularly-spaced base state.
+To find $w_0$ we integrate in one dimension using the procedure in Appendix B of Paper V, keeping in mind that the base state spacing ($\Delta r$) should be computed using the appropriate cell-edge and cell-center locations when using irregularly-spaced base state.
 Note that in this approach, we estimated the time-derivative of the pressure in part by examining how laterally averaged $\rho'=\rho-\rho_0$ changed over time (quantified by $\eta_\rho = \overline{\rho'\Ub\cdot\eb_r}$).
-After evolving the species ({\bf Steps 4A/8A}), we compute this term in a similar fashion as reported in Paper V with the slight difference in using the full velocity instead of the split velocities.
+After evolving the species ({\bf Steps 4A/8A}), we compute this term as reported in Paper V.
 For example, after {\bf Step 8A}, we define a radial cell-centered $\etarho^{\nph}$,

### 2019-06-11 MAESTROeX_I/results.tex
fixed a couple of typos

@@
 Here, we perform simulations using $256^3, 512^3, 768^3, 1024^3, 1280^3$, and $1536^3$ grid cells on a spatially uniform grid (no AMR).
-We divide each simulation into $64^3$ grids, so this simulations contain between 64 grids ($256^3$) and 13,824 grids ($1536^3$).
+We divide each simulation into $64^3$ grids, so these simulations contain between 64 grids ($256^3$) and 13,824 grids ($1536^3$).
 These simulations were performed using the NERSC cori system on the Intel Xeon Phi (KNL) partition.
@@
 We also note that without base state evolution, the code runs 14\% faster for small problems, and scales much better with wallclock time from the smallest to largest simulation increasing by only 13\%.
-This is quite remarkable since there are 3 linear solver per time step (2 cell-centered Poisson solves used in the MAC projection, and a nodal Poisson solve used to compute the updated cell-centered velocities).
+This is quite remarkable since there are 3 linear solves per time step (2 cell-centered Poisson solves used in the MAC projection, and a nodal Poisson solve used to compute the updated cell-centered velocities).
 Contrary to our prior assumptions, the linear solves are not the primary scaling bottleneck in this code.
@@
 For each of the problems we tested, the single-level simulation was run using the original temporal scheme and the adaptive simulations using the original and new temporal algorithms.
-We want to show that the adaptive simulation can give the similar results as the single-level simulation and in a more computationally efficient manner.
+We want to show that the adaptive simulation can give similar results to the single-level simulation and in a more computationally efficient manner.
 

### 2019-06-23 MAESTROeX_I/paper.tex

@@
 
-\correspondingauthor{Doreen Fan; DFan@lbl.gov}
+\correspondingauthor{Duoming Fan; DFan@lbl.gov}
 
 \author[0000-0002-3246-4315]{Duoming Fan}
-\affil{Lawrence Berkelely National Laboratory \\
+\affil{Lawrence Berkeley National Laboratory \\
 Center for Computational Sciences and Engineering \\
@@
 \author[0000-0003-1791-0265]{Andrew Nonaka}
-\affil{Lawrence Berkelely National Laboratory \\
+\affil{Lawrence Berkeley National Laboratory \\
 Center for Computational Sciences and Engineering \\
@@
 \author[0000-0003-2103-312X]{Ann S. Almgren}
-\affil{Lawrence Berkelely National Laboratory \\
+\affil{Lawrence Berkeley National Laboratory \\
 Center for Computational Sciences and Engineering \\

### 2019-07-03 MAESTROeX_I/results.tex
add note about problem

@@
 We perform weak scaling tests for simulations of convection preceding ignition in a spherical, full-star sub-Chandrasekhar mass white dwarf.
+\MarginPar{is this really the sub-Chandra stuff?
+or was it wdconvect, which is Chandra?}
 The simulation setup remains the same as reported in Section 3 of \cite{MAESTRO_AMR} and originally used in \cite{MAESTRO_convection}, and thus we emphasize that these scaling tests are performed using meaningful, scientific calculations.

### 2019-07-08 astronum_2019_MAESTROeX/astronum-maestroex.tex
Added astronum 2019 maestroex proceedings files

--- 2018-12-31 astronum_2018/astronum-xrb.tex
@@
 \newcommand{\maestro}{{\sffamily Maestro}}
+\newcommand{\maestroex}{{\sffamily MAESTROeX}}
 \newcommand{\castro}{{\sffamily Castro}}
-\newcommand{\starkiller}{{\sffamily StarKiller}}
-\newcommand{\starlord}{{\sffamily StarLord}}
-\newcommand{\nyx}{{\sffamily Nyx}}
 \newcommand{\amrex}{{\sffamily AMReX}}
-\newcommand{\vode}{{\sffamily VODE}}
 
@@
 
-\title{Toward Resolved Simulations of Burning Fronts in Thermonuclear
-       X-ray Bursts}
+\title{Modelling low Mach number stellar hydrodynamics with MAESTROeX}
 
-\author{M. Zingale$^1$,
-        K.~Eiden$^1$,
-        Y.~Cavecchi$^{2,3}$,
-        A.~Harpole$^1$,
-        J.~B. Bell$^4$,
-        M.~Chang$^1$,
-        I.~Hawke$^3$,
-        M.~P. Katz$^5$,
-        C.~M. Malone$^6$,
-        A.~J. Nonaka$^4$,
-        D.~E. Willcox$^4$, and
-        W. Zhang$^4$}
+\author{A.~Harpole$^1$,
+        D. Fan$^2$,
+        M.~P. Katz$^3$,
+        A.~J. Nonaka$^2$,
+        D.~E. Willcox$^2$, and
+        M. Zingale$^1$,}
 
@@
 
-\address{$^2$Department of Astrophysical Sciences, Princeton University,
-  Peyton Hall, Princeton, NJ 08544, USA}
-
-\address{$^3$Mathematical Sciences and STAG Research Centre,
-  University of Southampton, SO17 1BJ, UK}
-
-\address{$^4$Center for Computational Sciences and Engineering,
+\address{$^2$Center for Computational Sciences and Engineering,
   Lawrence Berkeley National Lab, Berkeley, CA 94720 USA}
 
-\address{$^5$NVIDIA Corporation, 2788 San Tomas Expressway,
+\address{$^3$NVIDIA Corporation, 2788 San Tomas Expressway,
   Santa Clara, CA, 95050 USA}
 
-\address{$^6$Los Alamos National Laboratory, Los Alamos, NM, 87545 USA}
-
-\ead{michael.zingale@stonybrook.edu}
+\ead{alice.harpole@stonybrook.edu}
 
 \begin{abstract}
-We discuss the challenges of modeling X-ray bursts in
-multi-dimensions, review the different calculations done to date, and
-discuss our new set of ongoing simulations.  We also describe
-algorithmic improvements that may help in the future to offset some of
-the expense of these simulations, and describe what may be possible
-with exascale computing.
+abstract abstract abstract
 \end{abstract}
@@
 
-X-ray bursts (XRBs) are fascinating astrophysical explosions.  A
-neutron star accretes fuel from its binary companion, building up just
-a thin layer ($\sim 5$--$10$~m) before the immense gravitational
-acceleration at the neutron star surface compresses and heats the fuel
-to the point of thermonuclear runaway.  A brief flash of X-rays
-follows, the accreted layer diffuses the heat produced from reactions,
-and the process repeats.  We can observe multiple bursts from a single
-source, and the information their lightcurves encodes tells us about
-the underlying neutron star, and ultimately the nuclear equation of
-state.  A challenging aspect of interpreting the lightcurves is
-understanding the radiation transport through the hot ash layers.  In
-particular, the composition can affect the effective temperature,
-which in turn affects the radius we measure for the neutron
-star~\cite{suleimanov:2011}.  Convection in the burning layer can even
-bring ashes up to the photosphere which alters what we
-see~\cite{kajava:2017}.  Accurate measurements of neutron star radii
-from XRBs could greatly constrain the nuclear equation of
-state~\cite{steiner:2010,ozel:2010}.  Finally, brightness oscillations
-during the rising phase of the bursts
-and other features seen in lightcurves are interpreted as arising from
-the spreading of a hotspot across the neutron star---demonstrating
-that the burst begins in a localized
-region~\cite{bhattacharyya:2006,bhattacharyya:2007}.
-See~\cite{galloway:2017} for a recent review of XRBs.
+intro intro intro
 
-Simulations of XRBs can help us to understand the observations, but
-many technical challenges remain.  The primary difficulties are simply
-the range of length and timescales involved.  At the largest lengthscale,
-the neutron star radius is 10~km and the Rossby lengthscale, where rotation
-and lateral pressure gradients balance, is $\sim
-1$~km~\cite{SPIT_ETAL02}.  At the smallest scales, a conductive He
-flame has a thickness of 10s of centimeters~\cite{Timmes00} and the
-pressure scale height of the fuel layer is on the order of meters.  To capture the
-rise timescale of the XRB, we would need to model seconds, but the flame itself
-is subsonic, so long timescale evolution is difficult.
+\section{Low Mach Number hydrodynamics}
+slow fluids in space are hard to model
 
-Present and past simulations of XRBs employ a variety of
-approximations to get past these difficulties.  The variety of
-approximations used enables a complementary picture of XRBs to be built up,
-and has greatly advanced our understanding of these systems.
-The breadth of simulations to date has been impressive.  We summarize
-the different approaches below.
+\section{MAESTROeX}
+\maestroex\ is our awesome new code for modelling slow fluids in space
 
-One-dimensional simulations, e.g. \cite{woosley-xrb,fisker:2008}, assume
-spherical symmetry and model the burst through the vertical column
-depth of fuel.  These simulations can capture the luminosities and
-burst recurrence times well.  They can utilize very large nuclear
-reaction networks and have been used to understand the nucleosynthesis
-and rp-process~\cite{schatz:rp1999,rpprocess,schatz_rp}, as well as to understand how rate uncertainties can
-affect the outcomes of the bursts.  Many successive bursts can be
-modeled and these simulations can show us the evolution of the ash
-layer with time.  The approximation of 1D however means that we cannot
-learn about lateral variations across the neutron star, including
-flame spreading.
-
-Global shallow water simulations~\cite{SPIT_ETAL02} capture the
-large scale spreading of the burning front by using a very simple
-vertical structure.  These simulations were instrumental in showing
-that the Coriolis force plays an important role in confining a
-spreading region and that the flame spreading can explain brightness
-oscillations during the rise of the burst lightcurve.
-
-Our group and others have modeled the convection preceding the
-runaway using low Mach number hydrodynamic methods for pure helium
-bursts~\cite{Lin:2006,xrb} and mixed hydrogen/helium
-bursts~\cite{xrb2} in 2D, as well as full 3D simulations of mixed
-bursts~\cite{xrb3}, and followed the development of turbulence.
-These simulations can give an understanding of the pre-burning front
-regime, including the strength and character of the turbulence, but
-cannot model large scale lateral variations because of the implicit
-assumptions built into the low Mach number model~\cite{ABRZ:I}.
-
-While detonations are computationally easier to model than flames (we
-largely get the speed through the jump conditions in the Riemann
-problem without resolving the structure), detonations require extreme
-conditions not found in XRBs~\cite{ZINGALE_ETAL01,harpole:2018}.
-This means that calculations that want to capture the nucleosynthesis
-yields from spreading burning fronts in XRBs need to model a deflagration.
-
-Flame spreading was first modeled in a series of calculations
-employing an algorithm from atmospheric science, where vertical
-hydrostatic equilibrium is enforced and wide-aspect ratio zones give a
-horizontal CFL number that is large enough to obtain long timescale
-evolution~\cite{cavecchi:2012}.  These calculations display key
-features of the propagation mechanism, namely a balance between
-hydrodynamics and the pure flame physics. The flame proceeds mainly
-via conduction, thus being a deflagration. The flame front, however,
-is not vertical but inclined, the angle with the horizontal being very
-small: $\theta \sim H/R_{\rm{Ro}} \sim 10^{-3}$. Here $H$ is the scale
-height and $R_{\rm{Ro}}$ is the Rossby radius. It is the
-hydrodynamical balance between the Coriolis force, gravity and
-pressure gradients that determines the Rossby radius and therefore the
-inclination angle of the flame front. The extended surface of the
-flame front is what leads to propagation speeds of order $10^5 $~km/s,
-in good agreement with observations. Due to its dependence on the
-Coriolis force, the flame speed changes across the surface of the NS
-(being faster at the equator), leading to potentially detectable
-effects if the rise of the burst is observed with enough resolution
-\cite{art-2015-cavecchi-etal}.  We note that no calculations have
-considered the effects of the flame interacting with preexisting
-turbulence ahead of the flame from simmering that may be occurring
-there, as we characterized in \cite{xrb3}.  Flame-turbulence
-interactions have long been seen as a means to accelerate a flame in
-Type Ia supernovae, so future studies should look at turbulence in the
-layer.  Capturing turbulence, of course, requires high resolution, so
-the ideas described here will be needed.
-
-The NS systems exhibiting XRBs are
-expected to have magnetic fields in the range $10^7$--$10^{10}$~G.
-Inclusion of a vertical magnetic field proved to have a strong
-influence on the nature and the speed of flame propagation
-\cite{art-2016-cavecchi-etal}, due to the interaction between the
-Coriolis force and the magnetic tension, which further acts towards
-determining the inclination angle of the flame front.
-
-A longstanding goal is to perform simulations where we resolve the
-flame structure, to allow us to accurately capture the
-nucleosynthesis, and watch it move across the neutron star surface.
-The results of these simulations will enable strict comparison to
-observations made with the new generation of X-ray telescopes such as
-\textit{eXTP}\/ and \textit{STROBE-X}\/
-\cite{art-2017-wilhod-etal,art-2016-zhang-etal,intZand2018} which will offer
-high collecting area and time resolution.
-We show some in-progress calculations that are a step toward this, discuss
-their remaining approximations, and also discuss what new algorithmic
-techniques might be needed to realize this goal with the advent of
-exascale computing.
-
-\section{Resolved Flame Studies}
-
-We have begun a set of flame spreading calculations using the
-compressible hydrodynamics code \castro~\cite{castro}, part of the
-AMReX astrophysics suite~\cite{astronum:2017}.  \castro\ uses adaptive
-mesh refinement, supports a general equation of state and reaction
-networks, has a conservative gravity formulation~\cite{wdmergerI}, and
-is optimized to run on current supercomputers.  All of our code base
-is open source and available on
-GitHub\footnote{\url{http://github.com/amrex-astro/}}, and all of the
-solvers, problem setup, initial models, and analysis scripts for each
-of our science problems are distributed with the codes.
-
-We setup a problem by creating a pair of 1D hydrostatic XRB models (hot and cool)
-representing the post- and pre-flame states and apply these laterally
-across the domain to setup a hot region at the left end of our domain
-in equilibrium with the cool model on the right.  Thermal diffusion and nuclear
-reactions quickly produce a laterally propagating flame that begins to
-move through our domain.  We resolve the diffusion scale, using a
-resolution of 10~cm, and use adaptive mesh refinement (two jumps of
-$4\times$) to keep a buffer of low density material at low resolution
-above the surface to allow for expansion.  To make the lengthscales
-tractable, we rotate the neutron star at 2000~Hz---much faster than
-expected.  This helps confine the spreading region to a more
-reasonable horizontal domain.  Nevertheless, our simulation uses 12288
-zones laterally, on the finest grid.  We use the stellar
-conductivities from \cite{Timmes00} and a 13-isotopic alpha-chain
-reaction network.  To make the timescales easier, we boost the flame
-speed by a factor of 10 by scaling both the conductivity and energy
-generation rate by the same factor.  This is an approximation we hope to
-relax in the near future.
-
-Figure~\ref{fig:xrb} shows the flame structure for an initial 2D run
-at 0.007~s.  Vertical profiles at a few radial distances from the
-origin are shown for the same quantities in
-Figure~\ref{fig:xrb-profiles}.  The burning layer is only about 10~m
-high.  Underneath is \isot{Ni}{56}, representative of the underlying
-neutron star, and above it is a very low density ($10^{-4}~\gcc$)
-buffer.  Only a portion of the vertical extent is shown.  The top
-panel shows the temperature structure.  The initial perturbation
-reached to $2\times 10^4~\mathrm{cm}$, so at this point, the extent of
-the burnt region has nearly doubled.  The flame structure is quite
-wide, since there are multiple burning stages beyond just helium to
-carbon, and this shows up in the composition.  This is seen more
-clearly in the next panel which shows the mean molecular weight of the
-nuclei ($\bar{A}$).  We see that the further behind the head of the
-flame we are, the heavier the ash nuclei.  We also see the head of the
-flame appears lifted off of the base of the burning layer---unburned
-material underlies the spreading flame.  This is reminiscent of the
-flame structure seen in the earlier simulations of
-\cite{cavecchi:2012}.  The third panel shows the energy generation
-rate.  It is strongest just above the neutron star surface, behind the
-flame.  We again see that the head of the flame is detached from the
-base of the atmosphere, with the strongest burning at the head of the
-flame slightly higher up in the atmosphere.  Finally, the bottom panel
-shows the velocity through the plane of the simulation---this is
-induced by the Coriolis force as the burning fluid expands and begins
-to spread laterally.  A tight hurricane structure has setup as a
-result of the flame spreading.  This is the geostrophic balance
-discussed in \cite{SPIT_ETAL02}.  These calculations are ongoing and
-will be the subject of a detailed study in the near future.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=\linewidth]{xrb}
-\caption{\label{fig:xrb} Helium flame spreading across the surface of
-  a neutron star.  Shown are temperature (top), the mean molecular
-  weight of the ash (second), the nuclear energy generation rate
-  (third), and the velocity out of the simulation plane (bottom).  The
-  out-of-plane velocity contours indicate where the Coriolis force has
-  created a hurricane-like wind structure which confines the flame,
-  leading to an inclined front. It is at such an interface that the
-  temperature and burning rate are highest.  }
-\end{figure}
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=0.8\linewidth]{xrb_profiles}
-\caption{\label{fig:xrb_profiles} Vertical profiles at several radial distances
-  from the vertical axis through the plots shown in Figure~\ref{fig:xrb}.  These
-  profiles sample the region behind and ahead of the flame.}
-\end{figure}
-
-\section{Large Scale Simulations and Algorithmic Developments}
-
-The 2D calculations shown in the previous section require about 10k
-node-hours of computational time to get to about 10~ms (running
-on the Edison machine at NERSC).  We expect qualitatively different
-results in 3D calculations: for example, the shear from the hurricane structure
-acting along the flame front may cause instabilities that can affect the flame
-propagation.
-With our current simulation methodology,
-a 3D version of this calculation is estimated to require 10M node-hours
-or more---this is about the size of an annual allocation.
-%Additionally, this is still with the boosted flame, but we expect that once the flame is developed, the boosting will not be as critical, due to hydrodynamic effects.
-This simulation covers less than
-1\% of the neutron star surface, so a fully resolved calculation of
-burning over the entire neutron star is out of reach: a naive scaling
-calculation suggests that 15 years of Moore's Law improvements are required.
-%
-% The calculation above is
-% log_2 [ (4 pi 10^2) / (1.23^2) ] * 1.5
-% for a NS of 10km radius vs the computational domain of 1.23km^2
-%
-
-Improvements are needed to push the scale of simulations
-to the point where we can model XRBs at the full star scale with
-resolved nuclear physics.
-As the computational cost is roughly a product of the number of zones, the
-number of zone updates, and the cost per zone update, we need to look for
-improvements in each of these. The number of zones depends on the physical
-model employed, the meshing strategy, and the accuracy required. The number of
-zone updates depends on the smallest zone size and the largest characteristic
-speed in the model. The computational cost per zone update is dominated by the
-time spent in the reactions and equation of state. We discuss a few
-possible improvements that can come both from implementing new algorithms and
-from exploiting new hardware.
-
-%For example, the number of zones can be reduced
-%by symmetry considerations. We can bootstrap the 3D calculation by mapping the
-%2D axisymmetric calculation, once the flame is firmly established, into
-%3D, rotating about the symmetry axis.  This will cut down the computational
-%time by about one-half.
-
-Machine architectures are evolving, with heterogeneous architectures
-becoming more common.  Our recent development in \castro\ has focused
-on moving the hydrodynamics to GPUs~\cite{astronum:2017}.  The XRB
-problem is an ideal case for this, since hydrodynamics with constant
-gravity (for plane-parallel domains) ports in a straightforward manner
-to GPUs.  The current GPU version of the \castro\ hydrodynamics solver
-is about $10\times$ faster using GPUs on a node than the entire node
-of CPU cores.  We see similar speed ups with reaction networks, with
-these GPU ports under active development.  This means a GPU
-calculation could lower the above costs by a factor of 10.  GPU
-offloading of the reaction networks will also allow us to explore
-larger networks and more detailed nuclear physics.
-
-Another way we can reduce the number of zones for given accuracy is to use a more
-accurate hydrodynamic method.  We are developing fully
-fourth-order (in space and time) coupling of hydrodynamics and
-reactions using the method of spectral deferred corrections (SDC) in
-\castro, instead of the more commonly employed Strang splitting~\cite{strang:1968}.  This
-provides us with two benefits.  First, the improved coupling actually
-reduces the stiffness of the reactions, requiring fewer righthand side
-calls and therefore reducing the expense of the
-reactions.  Second, by moving to fourth
-order, we may also be able to reduce our resolution requirements
-needed for converged flames. For smooth wave solutions we may expect to need
-around half the number of zones per dimension, and since computational work
-scales like $(\Delta x)^4$ in 3D, this could lower costs by a factor of around
-10.
-%
-% This calculation is based on the number of points per wavelength required,
-% to resolve a periodic wave, which (I believe) goes from 12 (2nd order) to
-% about 7 (4th order). However, this is very idealized, and I'm not sure that
-% it carries over to the reactions, so this may well be optimistic.
-%
-
-As an example of the improvements in coupling between hydro
-and reactions, Figure~\ref{fig:sdc} shows the mass fraction of helium
-behind a detonation over the course of 2 timesteps.  For this test,
-a 19 isotope reaction network was used.  The points
-represent individual calls to the righthand side function of our
-reaction network, and the density of
-points indicates how hard the integrator is working.  We use the same
-tolerances for both the Strang-split method and the SDC method.  For
-the SDC method, we are using a second-order accurate method based on
-\cite{SDC-old}.  We predict a time-centered advection term,
-$\Adv{\Uc}^{n+1/2}_i$ using standard unsplit Godunov methods, but
-explicitly include a reactive source term in the tracing of the
-interface states, ${\bf R}(\Uc_i)$.  We can then solve the reactive
-system, using this advective term as a piecewise-constant-in-time
-source
-\begin{equation}
-\label{eq:syst}
-\frac{d{\Uc_i}}{dt} = -\Adv{\Uc}^{n+1/2}_i + {\bf R}(\Uc_i)
-\end{equation}
-This system can be integrated with standard ODE methods.  To achieve
-second-order accuracy, we need to iterate, using the updated $\Uc$ to
-create a reactive source included in the predictor for the advective
-update, and then reintegrate the system.  The result of this iteration
-is that the advection sees the effects of reactions over the timestep
-and the reactions see the effects of advection.  For the zone tracked
-in Figure~\ref{fig:sdc}, each SDC iteration called the righthand side
-function of the network 2 times less than the Strang case.  This
-savings is dependent on the network and thermodynamic state, but our
-tests thus far show that the SDC methods need fewer network calls to
-integrate a timestep.  We also clearly see how the Strang system
-evolves in a discontinuous fashion.  We are continuing to develop this
-new integration strategy, with full fourth-order in space and time
-reacting hydrodynamics methods to follow shortly.  The XRB simulations
-are our target application.
-
-For the ODE integrations shown we use the variable-order,
-variable-step integrator \vode\ \cite{vode} with its implementation of
-the backward-differentiation formulas.
-% The integration order
-%corresponds to the order of the interpolating polynomials used to
-%predict the solution in each integrator step. 
-We note that for both
-Strang splitting and SDC, \vode\ must start the integration at first
-order and correspondingly small timesteps since no higher-order
-information about the solution is available (our reaction
-network right-hand sides provide only the first time derivative of the
-solution). Based on testing with a 13-isotope
-approximate alpha-chain network, \vode\ takes many small timesteps
-(e.g.\ $\sim 200$) before it can stably increase to a high order
-representation of the solution, but after it does so, it can often
-complete the integration in only a handful of steps (e.g.\ $\sim
-5$--$6$). Observing that during a \castro\ timestep, the first Strang
-step is a direct continuation of the integration in the second Strang
-step of the previous \castro\ timestep, we explored saving the
-\vode\ state following the second Strang step and restarting
-\vode\ during the first Strang step. This approach allowed us to start
-the first Strang step with a high order representation of the solution
-and complete it in very few \vode\ timesteps ($\lesssim 10$ steps),
-speeding up the \castro\ timestep advance by 50\% for our detonation
-test problem using the 13-isotope network. However, this came at the
-memory cost of storing $7 N_{eq} + 90$ scalar variables on our grid,
-where for our networks $N_{eq} = N_{spec} + 2$ (in addition to the
-number of species $N_{spec}$ we also integrate temperature and
-specific energy). For a large-scale science problem with a large
-network, this could quickly become prohibitive, so we are exploring
-ways to mitigate either the memory requirements for restarting the
-integration or the small timestep start up costs for the integration
-method.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.75\linewidth]{sdc_plot}
-\caption{\label{fig:sdc} Helium mass fraction as a function of time
-  over two timesteps for a detonation test problem with a 19-isotope
-  reaction network.  The points represent individual calls to the
-  righthand side function of our reaction network (see
-  Eq.~\ref{eq:syst}) and the density of points indicates how hard the
-  integrator is working.  For the Strang case, we see that the first
-  half of the reactions brings us off of the smooth solution
-  (represented by the SDC curve), then the advection overcorrects,
-  resulting in a discontinuous jump, and finally the second half of
-  the reactions works to bring the state back to the solution curve.}
-\end{figure}
-
-Other ways to make the simulation less expensive can include a subgrid
-model. Flame models are commonly used for Type Ia supernovae.  There
-however, the flame is thin compared to the scale height, so a simple
-model can represent the unresolved flame on the grid.  For the XRB,
-the flame thickness is a non-trivial fraction of the pressure scale
-height and further, the laminar flame properties will change strongly
-over the height of the burning layer (see~\cite{Timmes00}).  This
-makes traditional flame models a poor fit.  Methods that use even
-something as simple as subzone burning (e.g.~\cite{Wang2012190}) may
-help capture the energetics with lower resolution.  This can be
-implemented using the AMR machinery.  This may help a little, but will
-still be insufficient for full star models.
-
-If we continue to resolve the flame structure, then we need to be
-smarter (and more aggressive) about our grid refinement strategy.
-Currently, we put the entire fuel layer on the finest grid, but for
-the material ahead of the flame, we could keep it at lower resolution
-until the flame reaches it.  The main obstacle is that we need to keep
-the fuel ahead of the flame in hydrostatic equilibrium, and keeping it
-refined accomplishes this well.  Using well-balanced
-schemes~\cite{kappeli:2016} can help us retain hydrostatic equilibrium
-(HSE) in the lower refined regions, reducing the number of zones
-needed to maintain fidelity, but we would still need a way to regrid
-the layer ahead of the flame to higher resolution to accurately
-capture the burning dynamics.  To accomplish this, we envision
-building a refinement strategy that does interpolation of the data onto
-the newly refined grids enforcing hydrostatic equilibrium.  This can
-allow us to keep only a narrow fully refined region that ``slides''
-along the surface of the neutron star keeping pace with the advancing
-flame.  When using mesh refinement the overall costs are dominated by the
-number of zones at the finest resolution, so such a scheme would greatly reduce
-the computational costs. This is a longer term development.
-
-All of the above developments still involve fully compressible
-hydrodynamics, but multiscale methods~\cite{weinan2011principles} where we couple different hydrodynamics
-solvers together in a single simulation may ultimately be needed to model the full star while resolving the burning front.
-%\MarginPar{Alice and Ian contribute here}
-Such techniques rely on it being possible to decompose the dynamics of the system
-into two scales, e.g.~a short and a long lengthscale, and that the processes at each
-scale are at most weakly coupled with each other. In the context of XRBs, one possibility for
-multiscale modeling would be to model the large scale dynamics using a shallow water model
-(like that of~\cite{SPIT_ETAL02}) to capture the effects of the Coriolis force, and use
-a low Mach number or
-a fully compressible model in the vicinity of the burning front to capture
-the turbulent burning processes. Given that the computational costs of the shallow water model
-would be negligible compared to those of the low Mach/compressible model, the overall cost of
-this multiscale model would be determined by the size of the low Mach/compressible region only. This
-approach would therefore allow us to capture dynamics across the full star
-without sacrificing resolution around the flame.
-As with the more aggressive gridding strategy above, the hybrid approach
-reduces the number of zones with the largest computational cost through mesh
-refinement, and also reduces the computational costs where possible by
-employing cheaper physical models in regions where this is possible.
-An example of coupling compressible and low Mach number models implemented using the \amrex~framework, but in a terrestrial context,
-is~\cite{Motheau2018}. An initial implementation, \cite{Harpole2018}, in an
-astrophysical context, coupling shallow water to compressible models in both
-Newtonian and relativistic gravity, indicates that considerable computational
-efficiencies are possible, but that close attention to the coupling between the
-models is needed.
-
-%% Finally, we can try hybrid model approaches where the fully
-%% compressible simulation can act as a subgrid model.
-%% Multiscale methods can combine different hydro models together to allow
-%% for a full star simulation with resolved burning.  Examples include...
+\section{GPUs}
+GPUs can make things faster
 
@@
 
-XRBs are multiscale, multiphysics problems that are challenging to
-model.  Nevertheless, significant progress has been made in
-multidimensional models of these events, through a set of
-complementary algorithmic approaches.  Future advances in computer
-architectures and algorithms will allow for even more realistic models
-of these events, and help us connect to observations, ultimately telling
-us about the neutron star itself.
+Summary
 
-\ack The work at Stony Brook was supported by DOE/Office of Nuclear
-Physics grant DE-FG02-87ER40317 and contract 7418390 with Lawrence
-Berkeley National Laboratory as part of the Exascale Compute Project
-ExaStar collaboration.  The work at LBNL was supported by the DOE
-Office of Advanced Scientific Computing Research under Contract No,
-DE-AC02-05CH11231. YC is supported by the European Union Horizon 2020
-research and innovation programme under the Marie Sklodowska-Curie
-Global Fellowship grant agreement No 703916.  An award of computer
-time was provided by the Innovative and Novel Computational Impact on
-Theory and Experiment (INCITE) program. This research used resources
-of the Oak Ridge Leadership Computing Facility at the Oak Ridge
-National Laboratory, which is supported by the Office of Science of
-the U.S. Department of Energy under Contract No.\ DE-AC05-00OR22725.
-This research used resources of the National Energy Research
-Scientific Computing Center, which is supported by the Office of
-Science of the U.S. Department of Energy under Contract
-No.\ DE-AC02-05CH11231.  Visualizations were done using yt~\cite{yt}.
-This research has made use of NASA's Astrophysics Data System
-Bibliographic Services.
+\ack acknowledgements here ~\cite{yt}
+
+% The work at Stony Brook was supported by DOE/Office of Nuclear
+% Physics grant DE-FG02-87ER40317 and contract 7418390 with Lawrence
+% Berkeley National Laboratory as part of the Exascale Compute Project
+% ExaStar collaboration.  The work at LBNL was supported by the DOE
+% Office of Advanced Scientific Computing Research under Contract No,
+% DE-AC02-05CH11231. An award of computer
+% time was provided by the Innovative and Novel Computational Impact on
+% Theory and Experiment (INCITE) program. This research used resources
+% of the Oak Ridge Leadership Computing Facility at the Oak Ridge
+% National Laboratory, which is supported by the Office of Science of
+% the U.S. Department of Energy under Contract No.\ DE-AC05-00OR22725.
+% This research used resources of the National Energy Research
+% Scientific Computing Center, which is supported by the Office of
+% Science of the U.S. Department of Energy under Contract
+% No.\ DE-AC02-05CH11231.  Visualizations were done using yt~\cite{yt}.
+% This research has made use of NASA's Astrophysics Data System
+% Bibliographic Services.
 

### 2019-07-09 MAESTROeX_I/algorithm.tex
add comment for velocity splitting

@@
 Our new temporal integration scheme does not use this splitting, and is much simpler than the scheme from Paper V since we evolve the velocity directly rather than a more complex evolution equation for the perturbational velocity.
+\MarginPar{Should emphasize that velocity splitting is still done for projection steps, but full velocity is used to update advection and reactions. Also mention role of $\eta_{\rho}$ here(?)}
 Additionally, the new scheme uses a simpler predictor-corrector approach to the base state density and pressure that no 

### 2019-07-11 MAESTROeX_I/results.tex
updated amr results and added comments

@@
 Here we use a domain size of $3.6\times 10^7$ cm by $3.6\times 10^7$ cm by $2.88\times 10^8$ cm, and allow the grid structure to change with time.
-The single-level simulation at a resolution of $128^2 \times 1024$ was run on Cori Haswell with 48 processors and took approximately 33.5 s per time step (averaged over 10 time steps) using either the original or new temporal algorithm.
+The single-level simulation at a resolution of $128^2 \times 1024$ was run on Cori haswell with 48 processors and took approximately 33.5 s per time step (averaged over 10 time steps) using either the original or new temporal algorithm.
 The adaptive simulation has a resolution of $32^2 \times 256$ at the coarsest level, resulting in the same effective resolution at the finest level as the single-level simulation.
@@
 We continue to use the full-star white dwarf problem described in Section \ref{sec:whitedwarf} to test adaptive simulations on spherical geometry.
-The adaptive grid is refined twice by tagging the density at $\rho > 5\times 10^7$ g cm$^{-3}$ on the first level and $\rho > 10^8$ g cm$^{-3}$ on the second level.
-The simulation on a single-level grid of $512^3$ resolution took 26.01 s per time step (again averaged over 10 time steps).
+The adaptive grid is refined twice by tagging the density at $\rho > 10^5$ g cm$^{-3}$ on the first level and $\rho > 10^8$ g cm$^{-3}$ on the second level.
+\MarginPar{possibly mention that we found the code encounters numerical difficulties when the tagging values are too close to each other in subsequent levels of refinement}
+The simulation on a single-level grid of $512^3$ resolution took 12.7 s per time step (again averaged over 10 time steps).
 The adaptive grid has a resolution of $128^3$ at the coarsest level and an effective resolution of $512^3$ at the finest level.
-On this grid, 8.9\% of the cells (1,493,504 out of $256^3$ cells) are refined at the first level and 5.3\% (7,176,192 out of $512^3$ cells) at the second.
-The adaptive simulation took 11.55 s, resulting in more than a factor of 2 in speedup.
-Both simulations are computed to $t=400$ s and we choose to use the peak temperature as the diagnostic to compare the results.
+On this grid, 27.8\% of the cells (4,666,880 out of $256^3$ cells) are refined at the first level and 5.3\% (7,176,192 out of $512^3$ cells) at the second.
+The adaptive simulation took 5.61 s, resulting in more than a factor of 2 in speedup.
+Both simulations are computed to $t=2000$ s and we choose to use the peak temperature as the diagnostic to compare the results.
 Figure \ref{fig:wdconvect_amr_Tmax} shows the evolution of the peak temperature for all three runs and shows that the adaptive simulation gives the same qualitative result as the single-level simulation.
 We do not expect the curves to match up exactly because the governing equations are highly nonlinear, and slight differences in the solution caused by solver tolerance and discretization error can change the details of the results.
-Each simulation was run on Cori KNL with 512 processors and 16 threads per core.
+Each simulation was run on Cori haswell with 512 processors and 4 threads per core.
 
@@
 \includegraphics[width=4.0in]{./figs/wdconvect_amr_Tmax}
-\caption{\label{fig:wdconvect_amr_Tmax} Peak temperature, $T_{\text{peak}}$, in a white dwarf from $t=0$ to $400$ s
+\caption{\label{fig:wdconvect_amr_Tmax} Peak temperature, $T_{\text{peak}}$, in a white dwarf from $t=0$ to $2000$ s
          for grids with effective resolution of $512^3$. We can see that the adaptive grids with two levels of

### 2019-07-26 MAESTROeX_I/results.tex
addressed all comments in results section

@@
 \subsection{Performance and Scaling}\label{sec:scaling}
-We perform weak scaling tests for simulations of convection preceding ignition in a spherical, full-star sub-Chandrasekhar mass white dwarf.
-\MarginPar{is this really the sub-Chandra stuff?
-or was it wdconvect, which is Chandra?}
+We perform weak scaling tests for simulations of convection preceding ignition in a spherical, full-star Chandrasekhar mass white dwarf. 
 The simulation setup remains the same as reported in Section 3 of \cite{MAESTRO_AMR} and originally used in \cite{MAESTRO_convection}, and thus we emphasize that these scaling tests are performed using meaningful, scientific calculations.
@@
 \includegraphics[width=3.0in]{./figs/MAESTRO_scaling2}
-\caption{\label{fig:scaling} (Left) Weak scaling results for a spherical, full-star sub-Chandrasekhar mass white dwarf calculation using the original MAESTRO code, MAESTROeX, and MAESTROeX with base state evolution disabled.
+\caption{\label{fig:scaling} (Left) Weak scaling results for a spherical, full-star white dwarf calculation using the original MAESTRO code, MAESTROeX, and MAESTROeX with base state evolution disabled.
 Shown is the average wallclock time per time step.
@@
 
-To explore the accuracy of the new temporal algorithm, we now analyze in detail three-dimensional, full-star calculations of convection preceding ignition in a sub-Chandrasekhar mass white dwarf.
+To explore the accuracy of the new temporal algorithm, we now analyze in detail three-dimensional, full-star calculations of convection preceding ignition in a white dwarf.
 Again, we refer the reader to \cite{MAESTRO_AMR} and \cite{MAESTRO_convection} for setup details.
@@
 
-In terms of efficiency, all three simulations on $256^3$ single-level grid were run on Cori haswell with 64 processors and 8 threads per core.
-The simulation using the new temporal algorithm took 6.75 s per time step with uniformly-spaced base state, which is 13\% faster than the 7.77 s per time step when using the original scheme, and 9.72 s with irregularly-spaced base state.
-The 25\% increase in run time of the irregularly-spaced algorithm is mainly caused by the much larger base state array and could be significantly reduced if we consider truncating the base state beyond the cutoff density radius.
-\MarginPar{why is the new temporal with constant dr faster?}
+In terms of efficiency, all three simulations on $256^3$ single-level grid were run on Cori haswell with 64 processors and 8 threads per core and their run times were compared.
+As a result of simplifying the algorithm by eliminating the evolution equations for the base state density and pressure, the simulation using the new temporal algorithm took only 6.75 s per time step with uniformly-spaced base state, which is 13\% faster than the 7.77 s per time step when using the original scheme.
+However, we do observe a 25\% increase in a run time of 9.72 s when using irregularly-spaced base state.
+This can be explained by the irregularly-spaced base state array being much larger in size than its regularly-spaced counterpart, and thus require additional computation time.
+One possible strategy to significantly reduce this run time is to consider truncating the base state beyond the cutoff density radius.
 
@@
 The adaptive grid is refined twice by tagging the density at $\rho > 10^5$ g cm$^{-3}$ on the first level and $\rho > 10^8$ g cm$^{-3}$ on the second level.
-\MarginPar{possibly mention that we found the code encounters numerical difficulties when the tagging values are too close to each other in subsequent levels of refinement}
+These tagging values are determined by referring to Paper V, but we have found that the code may encounter numerical difficulties when the tagging values are too close to each other in subsequent levels of refinement.
 The simulation on a single-level grid of $512^3$ resolution took 12.7 s per time step (again averaged over 10 time steps).

### 2019-07-28 xrb/paper.tex
Additions to flame acceleration section

@@
 
-\author{Kiran Eiden}
+\author[0000-0001-6191-4285]{Kiran Eiden}
 \affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
@@
 
-Describe the diagnostics used to measure the flame speed and compare to laminar.
+To measure the propagation rate of the burning front, we track the location beyond the hottest part of the flame where $\dot{e}_\mathrm{nuc}$ first drops to less than $0.1 \%$ of maximum.
+Using $0.1 \%$ of the peak value rather than the peak value itself helps to mitigate fluctuations caused by localized fluid motions, as the reduced value lies in a less turbulent part of the burning region.
+We also average $\dot{e}_\mathrm{nuc}$ over the vertical coordinate to further dampen these fluctuations.
+After an initial transient period typically spanning $\sim 3$ ms, the flame settles into a state of steady propagation.
+The position data here are well fitted by a linear function of time (as seen in Figure \ref{fig:flame_speed}), and the resulting slope gives the velocity of the flame front.
+
+Table \ref{table:flame_speeds} gives the flame speed measured in each multidimensional simulation, as well as laminar speeds obtained from any corresponding 1D runs.
+The 2D flames propagate at speeds about an order of magnitude faster than their 1D counterparts.
+The increase in flame speed is likely a product of hydrodynamical effects such as turbulence, wrinkling, and convective cycles, which bring cooler fuel from ahead of the front into the hottest part of the burning region.
+
+\begin{deluxetable}{lcc}
+	\tablecaption{\label{table:flame_speeds}Turbulent and laminar flame speeds.}
+	\tablehead{\colhead{run} & \colhead{$s_T$ (km s$^{-1}$)} & \colhead{$s_L$ (km s$^{-1}$)}}
+	\startdata
+	10/10 boost & $9.46 \pm 0.04$  & $1.06 \pm 0.01$ \\
+	5/5 boost & $3.97 \pm 0.01$ & $0.56 \pm 0.01$ \\
+	iso7 network & $7.66 \pm 0.04$ & $-$ \\
+	1000 Hz & $26.1 \pm 0.20$ & $-$ \\
+	\enddata
+	%
+\end{deluxetable}
+
+\begin{figure}[t]
+\centering
+\plotone{plots/results/speed_10_10.pdf}
+\caption{\label{fig:flame_speed}The position of the burning front for the 10/10 flame as a function of time. The dashed line shows a linear least squares fit for $t \gtrsim 10$ ms.}
+\end{figure}
 
@@
 
-We also try a rotation period of $P_\mathrm{rot} \sim 10^{-4}~\mathrm{s}$.
+We also try a rotation period of $P_\mathrm{rot} \sim 10^{-3}~\mathrm{s}$.
 

### 2019-07-29 MAESTROeX_I/algorithm.tex
updated amr results and added comments
address comments in algorithm section

--- 2019-06-23 MAESTROeX_I/algorithm.tex
@@
 For the lateral average, as before we first sum the values in all the cells associated with each radius and divide
-by the number of contributing cells to obtain the arithmetic average.  However we do not map this onto a regularly spaced
+by the number of contributing cells to obtain the arithmetic average.  However we do not map this onto a uniformly spaced
 base state and retain the use of this irregularly-spaced base state.
@@
 \subsection{Temporal Integration Scheme}\label{Sec:Temporal Integration Scheme}
-We now describe the new temporal integration scheme, which we note can be used for the original base state mapping (with constant base state grid spacing) as well as the new irregularly spaced base state mapping.
+We now describe the new temporal integration scheme, which we note can be used for the original base state mapping (with uniform base state grid spacing) as well as the new irregularly spaced base state mapping.
 Previously we adopted an approach where we split the velocity into a base state component, $w_0(r,t)$, 
@@
 \begin{equation}
-\Ub = \Ubt(\xb,t) + w_0(r,t)\eb_r,
+\Ub = \Ubt(\xb,t) + w_0(r,t)\eb_r, \label{eq:velsplit}
 \end{equation}
@@
 This resulted in some unnecessary complications to the temporal integration scheme.
-Our new temporal integration scheme does not use this splitting, and is much simpler than the scheme from Paper V since we evolve the velocity directly rather than a more complex evolution equation for the perturbational velocity.
+Our new temporal integration scheme uses full velocities to evolve advection and reactions instead, and only uses the above splitting to satisfy the velocity divergence constraints due to boundary considerations.
+This produces a cleaner scheme than the one from Paper V since we use the velocity directly rather than more complex terms involving the perturbational velocity.
 Additionally, the new scheme uses a simpler predictor-corrector approach to the base state density and pressure that no 
@@
 Each projection step in the algorithm involves the solution of a variable-coefficient Poisson solve using multigrid.
+Note that we still employ velocity-splitting as described by equation (\ref{eq:velsplit}) for this step 
+in order to enforce the appropriate behavior of the system near the edge of the star as determined by the cutoff density.
 The details of this ``MAC'' projection are provided in Appendix \ref{Sec:Projection}.
@@
 \begin{equation}
-\frac{\partial(\rho X_k)}{\partial t} = -\nabla\cdot(\rho X_k\Ub) + \rho\omegadot_k,
+\frac{\partial(\rho X_k)}{\partial t} = -\nabla\cdot(\rho X_k\Ub),
 \end{equation}
 %
-omitting the reaction terms, which were already \MarginPar{since we are omitting the reaction terms, we should just leave them out of the Eq.}
+where the reaction terms have been omitted since they were already 
 accounted for in {\bf React State}.  The update consists of two steps:

### 2019-07-29 MAESTROeX_I/results.tex
address comments in algorithm section

@@
 As a result of simplifying the algorithm by eliminating the evolution equations for the base state density and pressure, the simulation using the new temporal algorithm took only 6.75 s per time step with uniformly-spaced base state, which is 13\% faster than the 7.77 s per time step when using the original scheme.
-However, we do observe a 25\% increase in a run time of 9.72 s when using irregularly-spaced base state.
-This can be explained by the irregularly-spaced base state array being much larger in size than its regularly-spaced counterpart, and thus require additional computation time.
-One possible strategy to significantly reduce this run time is to consider truncating the base state beyond the cutoff density radius.
+However, we do observe a 25\% increase in run time of 9.72 s when using irregularly-spaced base state with the new algorithm.
+This can be explained by the irregularly-spaced base state array being much larger in size than its uniformly-spaced counterpart, and thus require additional computation time.
+One possible strategy to significantly reduce the run time is to consider truncating the base state beyond the cutoff density radius.
 
@@
 The adaptive grid is refined twice by tagging the density at $\rho > 10^5$ g cm$^{-3}$ on the first level and $\rho > 10^8$ g cm$^{-3}$ on the second level.
-These tagging values are determined by referring to Paper V, but we have found that the code may encounter numerical difficulties when the tagging values are too close to each other in subsequent levels of refinement.
+These tagging values have been shown to work well previously in Paper V, but we have found that the code may encounter numerical difficulties when the tagging values are too close to each other in subsequent levels of refinement.
 The simulation on a single-level grid of $512^3$ resolution took 12.7 s per time step (again averaged over 10 time steps).

### 2019-08-01 MAESTROeX_I/algorithm.tex
minor changes and comments through the alg description

@@
 One of the key numerical modules is the ``lateral average'', which computes the average over a 
-layer of constant radius of a Cartesian grid variable onto a one-dimensional base state.
+layer of constant radius of a Cartesian grid variable and stores the result in a one-dimensional base state array.
 In planar geometries, this is a straightforward arithmetic average of values in cells at
@@
 there-dimensional Cartesian cell-center can map to.  
-Specifically, for every three-dimensional Cartesian cell, there exists an integer $m$ such that the radius to the center of the star is given by 
+Specifically, for every three-dimensional Cartesian cell, there exists an integer $m$ such that the distance
+from the cell center to the center of the star is given by 
 \begin{equation}
@@
 For the lateral average, as before we first sum the values in all the cells associated with each radius and divide
-by the number of contributing cells to obtain the arithmetic average.  However we do not map this onto a uniformly spaced
+by the number of contributing cells to obtain the arithmetic average.  However we do not interpolate this onto a uniformly spaced
 base state and retain the use of this irregularly-spaced base state.
@@
 \subsection{Temporal Integration Scheme}\label{Sec:Temporal Integration Scheme}
-We now describe the new temporal integration scheme, which we note can be used for the original base state mapping (with uniform base state grid spacing) as well as the new irregularly spaced base state mapping.
+We now describe the new temporal integration scheme, noting that it can be used for the original base state mapping 
+(with uniform base state grid spacing) as well as the new irregularly spaced base state mapping.
 Previously we adopted an approach where we split the velocity into a base state component, $w_0(r,t)$, 
@@
 We used $w_0$ to provide an estimate of the base state density evolution over a time step.
-This resulted in some unnecessary complications to the temporal integration scheme.
-Our new temporal integration scheme uses full velocities to evolve advection and reactions instead, and only uses the above splitting to satisfy the velocity divergence constraints due to boundary considerations.
-This produces a cleaner scheme than the one from Paper V since we use the velocity directly rather than more complex terms involving the perturbational velocity.
+This resulted in some unnecessary complications to the temporal integration scheme including
+base state advection modules for density, enthalpy, and velocity, as well as more 
+cumbersome split velocity dynamics evolution equations.
+Our new temporal integration scheme uses full velocities for scalar and velocity advection,
+and only uses the above splitting to satisfy the velocity divergence constraint due to boundary considerations
+at the edge of the star.
+This results in a much simpler numerical scheme than the one from Paper V 
+since we use the velocity directly rather than more complex terms involving the perturbational velocity.
 Additionally, the new scheme uses a simpler predictor-corrector approach to the base state density and pressure that no 
-longer requires evolution equations and the associated numerical discretizations to update the base state, greatly 
+longer requires evolution equations and numerical discretizations to update the base state, greatly 
 simplifying the algorithm while retaining the same overall second-order accuracy in time.
@@
 and ($\gammaonebar,\beta_0)$ are similarly computed from $(\rho,p_0,X_k)$;
-see Appendix A of Paper I and Appendix C of Paper III for details on how these terms are computed.
+see Appendix A of Paper I and Appendix C of Paper III for details on how $\beta_0$ is computed.
 
@@
 \item In {\bf Step 9} we redo the reactions from {\bf Step 5} beginning with the improved results from {\bf Steps 6--8}.
-\item In {\bf Steps 10--11} we advect the velocity, and then project these velocities so they satisfy the divergence constraint.
+\item In {\bf Steps 10--11} we advect the velocity, and then project these velocities so they satisfy the divergence constraint while updating $\pi$.
 \end{itemize}
@@
 The base state pressure remains equal to a constant value at the location of a prescribed cutoff density outward for the entire simulation.
-\item {\bf React State}$[(\rho X_k)^{\rm in},(\rho h)^{\rm in},p_0]\rightarrow[(\rho X_k)^{\rm out},(\rho h)^{\rm out},(\rho\dot\omega),(\rho\Hnuc)]$ integrates the species and enthalpy due to reactions over $\Delta t/2$ by solving
+\item {\bf React State}$[(\rho X_k)^{\rm in},(\rho h)^{\rm in},p_0]\rightarrow[(\rho X_k)^{\rm out},(\rho h)^{\rm out},(\rho\dot\omega),(\rho\Hnuc)]$ 
+uses VODE \citep{vode} to integrate the species and enthalpy due to reactions over $\Delta t/2$ by solving
 \begin{equation}
@@
 In stratified low Mach number models, due to the extreme variation in density, the velocity can become very large in low density regions at the edge of the star.
-Throughout Papers II-V, we have employed two techniques to help control these dynamics without significantly affecting the dynamics in the convective region.
+These large velocities can severely affect the time step, so
+throughout Papers II-V, we have employed two techniques to help control these dynamics without significantly 
+affecting the dynamics in the convective region.
 First, we use a cutoff density technique, where we hold the density constant outside a specified radius (typically near where the density is $\sim$4 orders of magnitude smaller than the largest densities in the simulation).

### 2019-08-01 MAESTROeX_I/equations.tex
minor changes and comments through the alg description

@@
 We take the standard equations of reacting, compressible flow, and recast the equation
-of state (EOS) as a divergence constraint on the velocity field, as explained below.
+of state (EOS) as a divergence constraint on the velocity field.
 The resulting model is a series of evolution equations for mass, momentum, and energy, subject
-to to an additional constraint on velocity.  The evolution equations are
+to an additional constraint on velocity.  The evolution equations are
 \begin{eqnarray}
@@
 $p = p_0 + \pi$ and $|\pi|/p_0 = \mathcal{O}({\rm Ma}^2)$ (we use $\xb$ to represent the Cartesian coordinate
-directions of the full state and $r$ to represent the radial coordinate direction for 
-the base state).  
+directions of the full state and $r$ to represent the radial coordinate direction for the base state).
+One way to mathematically think of the difference between $p_0$ and $\pi$ is that $\pi$ controls the velocity evolution 
+in a way that forces the thermodynamic variables $(\rho,h,X_k)$ to evolve in a manner that is consistent with the EOS and $p_0$.\MarginPar{AJN - added this sentence; does it make sense?}
+
 By comparing the momentum equation (\ref{eq:momentum}) to the momentum equation used in equation (2) in Paper V, we

### 2019-08-01 MAESTROeX_I/intro.tex
fixed a couple of typos
minor changes and comments through the alg description

@@
 solver for reactions (VODE) \citep{vode}, and multigrid-based linear solvers for the
-pressure-projection steps.  This, the time step is limited by an advective CFL constraint based on the 
+pressure-projection steps.  Thus, the time step is limited by an advective CFL constraint based on the 
 fluid velocity, not the sound speed.
@@
 thermodynamic variables to match this pressure.
-The time-advancement strategy uses a predictor-corrector strategy for evolving the base state,
-Strang splitting to integrate the thermodynamic variables, and a 
-second-order projection method to integrate the velocity equations subject to a divergence constraint.
+The time-advancement strategy uses Strang splitting to integrate the thermodynamic variables, a
+second-order projection method to integrate the velocity subject to a divergence constraint,
+and a velocity splitting scheme that uses a radially-averaged velocity to hydrodynamically evolve the base state.\MarginPar{AJN changed this sentence.}
 The original MAESTRO code was developed in the pure-Fortran 90 FBoxLib software framework, whereas
@@
 based on method-of-lines integration.
-In particular, schemes based on deferred corrections have been used to generate high-order temporal integrators for problems of reactive flow \citep{dutt2000spectral} and low Mach number combustion \citep{pazner2016high,nonaka2018conservative}.
+In particular, schemes based on deferred corrections \citep{dutt2000spectral} have been used to generate 
+high-order temporal integrators for problems of reactive flow and low Mach number combustion \citep{pazner2016high,nonaka2018conservative}.
 Second, we explore the effects of alternative spatial mapping routines for coupling the base state and the Cartesian grid state for spherical problems.

### 2019-08-01 MAESTROeX_I/algorithm.tex
in algorithm description, replace "edge" with "face" where appropriate

@@
 \item {\bf Enforce HSE}$[\rho_0]\rightarrow[p_0]$ computes the base state pressure, $p_0$, from a base state density, $\rho_0$ by integrating the hydrostatic equilibrium condition in one dimension. 
-This follows equation (A10) in Paper V, noting that for the irregularly spaced base state case, $\Delta r$ is not constant, where $\Delta r_{j+1/2} = r_{j+1} - r_j$ for cell-edge $j+1/2$.  
+This follows equation (A10) in Paper V, noting that for the irregularly spaced base state case, $\Delta r$ is not constant, where $\Delta r_{j+1/2} = r_{j+1} - r_j$ for cell face with index $j+1/2$.  
 The base state pressure remains equal to a constant value at the location of a prescribed cutoff density outward for the entire simulation.
@@
 %--------------------------------------------------------------------------
-\item[Step 3] {\em Construct a time-centered advective velocity on edges, $\uadvone$.}
+\item[Step 3] {\em Construct a face-centered, time-centered advective velocity, $\uadvone$.}
 
@@
 in Appendix A of \cite{XRB_III}.
-Here we use equation (\ref{eq:momentum}) to compute time-centered edge velocities, $\uadvonedag$.
+Here we use equation (\ref{eq:momentum}) to compute face-centered, time-centered velocities, $\uadvonedag$.
 The $\dagger$ superscript refers to the fact that the predicted velocity field does not satisfy the divergence constraint,
@@
 
-\item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\pred}$,
+\item Compute the face-centered, time-centered species, $(\rho X_k)^{\nph,\pred}$,
   for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
   As described in Paper V, for robust numerical slope limiting we predict 
-  $\rho'^n=\rho^n-\rho_0^n$ and $X_k^n$ to edges
-  and here we spatially interpolate $\rho_0^n$ to edges to assemble the fluxes.
+  $\rho'^n=\rho^n-\rho_0^n$ and $X_k^n$ to faces
+  and here we spatially interpolate $\rho_0^n$ to faces to assemble the fluxes.
 
@@
 
-\item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\pred},$
+\item Compute the face-centered, time-centered enthalpy, $(\rho h)^{\nph,\pred},$
   for the conservative update of $(\rho h)^{(1)}$ using using a Godunov approach \citep{XRB_III}.
   As described in Paper V, for robust numerical slope limiting 
-  we predict $(\rho h)'^n=(\rho h)^n-(\rho h)_0^n$ to edges,
+  we predict $(\rho h)'^n=(\rho h)^n-(\rho h)_0^n$ to faces,
   where $(\rho h)_0^n$ is obtained by calling {\bf Average}$[(\rho h)^n]\rightarrow[(\rho h)_0]$,
-  and here we spatially interpolate $(\rho h)_0^n$ to edges to assemble the fluxes.
+  and here we spatially interpolate $(\rho h)_0^n$ to faces to assemble the fluxes.
 
@@
 %--------------------------------------------------------------------------
-\item[Step 7] {\em Construct a time-centered advective velocity on edges, $\uadvtwo$.}
+\item[Step 7] {\em Construct a face-centered, time-centered advective velocity, $\uadvtwo$.}
 
@@
 
-\item Compute the time-centered species edge states, $(\rho X_k)^{\nph,\pred}$,
+\item Compute the face-centered, time-centered species, $(\rho X_k)^{\nph,\pred}$,
   for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
-  Again, we predict $\rho'^n=\rho^n-\rho_0^n$ and $X_k^n$ to edges
-  and here we spatially interpolate $(\rho_0^n+\rho_0^{n+1,*})/2$ to edges to assemble the fluxes.
+  Again, we predict $\rho'^n=\rho^n-\rho_0^n$ and $X_k^n$ to faces
+  and here we spatially interpolate $(\rho_0^n+\rho_0^{n+1,*})/2$ to faces to assemble the fluxes.
 
@@
 
-\item Compute the time-centered enthalpy edge state, $(\rho h)^{\nph,\pred},$
+\item Compute the face-centered, time-centered enthalpy, $(\rho h)^{\nph,\pred},$
   for the conservative update of $(\rho h)^{(1)}$ using a Godunov approach \citep{XRB_III}.
-  Again, we predict $(\rho h)'^n=(\rho h)^n-(\rho h)_0^n$ to edges
-  and here we spatially interpolate $[(\rho h)_0^n)+(\rho h)_0^{n+1,*}]/2$ to edges to assemble the fluxes.
+  Again, we predict $(\rho h)'^n=(\rho h)^n-(\rho h)_0^n$ to faces
+  and here we spatially interpolate $[(\rho h)_0^n)+(\rho h)_0^{n+1,*}]/2$ to faces to assemble the fluxes.
 
@@
 
-First, we compute the time-centered edge velocities, $\Ub^{\nph,\pred}$
+First, we compute the face-centered, time-centered velocities, $\Ub^{\nph,\pred}$
 using a Godunov approach \citep{XRB_III}. Then, we define
@@
 {\bf Step 3} and {\bf Step 7} because the velocities in those steps are defined
-on edges and $\Ub^{n+1}$ is defined at cell centers, requiring different divergence
+on faces and $\Ub^{n+1}$ is defined at cell centers, requiring different divergence
 and gradient operators.

### 2019-08-01 MAESTROeX_I/algorithm.tex
minor changes to sec 3 algorithm description

@@
 
-\item Compute the face-centered, time-centered species, $(\rho X_k)^{\nph,\pred}$,
+\item Compute the face-centered, time-centered species, $(\rho X_k)^{\nph,\pred,\star}$,
   for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
@@
 (\rho X_k)^{(2),\star} = (\rho X_k)^{(1)}
-  - \dt \left\{ \nabla \cdot \left[ \uadvone (\rho X_k)^{\nph,\pred} \right] \right\},
+  - \dt \left\{ \nabla \cdot \left[ \uadvone (\rho X_k)^{\nph,\pred,\star} \right] \right\},
 \end{equation}
@@
 
-\item Compute the face-centered, time-centered enthalpy, $(\rho h)^{\nph,\pred},$
+\item Compute the face-centered, time-centered enthalpy, $(\rho h)^{\nph,\pred,\star},$
   for the conservative update of $(\rho h)^{(1)}$ using using a Godunov approach \citep{XRB_III}.
@@
   we predict $(\rho h)'^n=(\rho h)^n-(\rho h)_0^n$ to faces,
-  where $(\rho h)_0^n$ is obtained by calling {\bf Average}$[(\rho h)^n]\rightarrow[(\rho h)_0]$,
+  where $(\rho h)_0^n$ is obtained by calling {\bf Average}$[(\rho h)^n]\rightarrow[(\rho h)_0^n]$,
   and here we spatially interpolate $(\rho h)_0^n$ to faces to assemble the fluxes.
@@
 (\rho h)^{(2),\star}
-= (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred} \right] \right\} + \Delta t\frac{Dp_0}{Dt}
+= (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred,\star} \right] \right\} + \Delta t\frac{Dp_0}{Dt}
 \end{equation}
@@
   the updated values $\uadvtwo$ and $\rho_0^{n+1,\star}$ rather than
-  $\uadvone$ and $\rho_0^{n+1,\star}$.  In particular:
+  $\uadvone$ and $\rho_0^n$.  In particular:
 
@@
   Again, we predict $\rho'^n=\rho^n-\rho_0^n$ and $X_k^n$ to faces
-  and here we spatially interpolate $(\rho_0^n+\rho_0^{n+1,*})/2$ to faces to assemble the fluxes.
+  but here we spatially interpolate $(\rho_0^n+\rho_0^{n+1,*})/2$ to faces to assemble the fluxes.
 
@@
 \item Update the enthalpy.  This step is identical to {\bf Step 4D} except we use
-  the updated values $\uadvtwo$, $\rho_0^{n+1}$, $(\rho h)_0^{n+1}$, and $p_0^{\nph}$
+  the updated values $\uadvtwo$, $\rho_0^{n+1}$, $(\rho h)_0^{n+1}$, and $p_0^{n+1}$
   rather than
-  $\uadvone, \rho_0^{n+1,\star}$, $(\rho h)_0^{n+1,\star}$, and $p_0^n$.
+  $\uadvone, \rho_0^n$, $(\rho h)_0^n$, and $p_0^n$.
   In particular:
@@
   Again, we predict $(\rho h)'^n=(\rho h)^n-(\rho h)_0^n$ to faces
-  and here we spatially interpolate $[(\rho h)_0^n)+(\rho h)_0^{n+1,*}]/2$ to faces to assemble the fluxes.
+  but here we spatially interpolate $[(\rho h)_0^n)+(\rho h)_0^{n+1,*}]/2$ to faces to assemble the fluxes.
 
@@
 First, we compute the face-centered, time-centered velocities, $\Ub^{\nph,\pred}$
-using a Godunov approach \citep{XRB_III}. Then, we define
-\begin{equation}
-\rho^\nph = \frac{\rho^n + \rho^{n+1}}{2}, \qquad \rho_0^\nph = \frac{\rho_0^n + \rho_0^{n+1}}{2}.
-\end{equation}
-We update the velocity field $\Ub^n$ to $\Ub^{n+1,\dagger}$ by discretizing
+using a Godunov approach \citep{XRB_III}. Then, we update
+the velocity field $\Ub^n$ to $\Ub^{n+1,\dagger}$ by discretizing
 equation (\ref{eq:momentum}) as
@@
  - \dt \left[ \frac{\beta_0^\nph}{\rho^\nph} \nabla \left( \frac{\pi^\nmh}{\beta_0^\nmh} \right) + \frac{\left(\rho^\nph-\rho_0^\nph\right)}{\rho^\nph} g^{\nph} \eb_r \right],
+\end{equation}
+where
+\begin{equation}
+\rho^\nph = \frac{\rho^n + \rho^{n+1}}{2}, \qquad \rho_0^\nph = \frac{\rho_0^n + \rho_0^{n+1}}{2}.
 \end{equation}
@@
 
-Our approach to adaptive mesh refinement is algorithmically the same as the treatment described in Section 5 of Paper V; we refer the reader there for details.
+Our approach to adaptive mesh refinement is algorithmically the same as the treatment described 
+in Section 5 of Paper V; we refer the reader there for details.
 MAESTROeX supports refinement ratios of 2 between levels.

### 2019-08-02 MAESTROeX_I/appendix.tex
Some touchups
finished pass

@@
 \end{tabular}
-\caption{\label{fig:wdconvect_splitU} (Left) Peak temperature, $T_{\text{peak}}$ in a white dwarf at resolution of $256^3$ 
-         during initial transition. (Right) Velocity magnitude solved using full velocity $\mathbf{U}$ in the projection step 
-         results in much larger values outside the density cutoff region, $\rho < 10^5 \text{ g cm}^{-3}$, that is not seen when 
-         we split the velocity dynamics.}
+\caption{\label{fig:wdconvect_splitU} When solving for the full velocity $\mathbf{U}$ in the nodal projection step instead of the split velocity 
+  formulation, we illustrate
+  (left) the peak temperature, $T_{\text{peak}}$ in a white dwarf at resolution of $256^3$ during the early evolution of the star and
+  (right) the magnitude of velocity at early time. We observe much larger velocities outside the density cutoff region, $\rho < 10^5 \text{ g cm}^{-3}$, 
+  that are not seen when we split the velocity dynamics.}
 \end{center}
@@
 \begin{equation}
-\etarho^{\nph,\star} = {\rm {\bf Average}}\left(\etarho^{{\rm cart},\nph}\right).
+\etarho^{\nph} = {\rm {\bf Average}}\left(\etarho^{{\rm cart},\nph}\right).
 \end{equation}

### 2019-08-02 MAESTROeX_I/results.tex
finished pass

@@
 The simulation using the new temporal scheme with irregularly-spaced base state displays a slightly different peak temperature behavior during the initial transition period $t<150$ s, which results in the difference between the curves post transition.
-We strongly suspect that this is a result of using different initial model inputs for the star in which the resolution near the center of the star is much coarser with the irregular spacing than the uniform spacing.
+We strongly suspect that this is a result of using a different initial model file (the resolution near the center of the star is much coarser with the irregular spacing than the uniform spacing).
 Fortunately, the simulation with irregular base state spacing still follows the same trend as with uniform spacing, and the star is shown to ignite at an earlier time $t=6840$ s.
@@
 Limited allocations prevented us from running this simulation further.
-As previously suspected, the simulation using irregularly-spaced base state agrees much closer to the results from using uniform spacing as the resolution is increased.
-This is most likely due to the increased resolution of the initial star model for the irregular-spaced base state that is now able to properly capture the same properties near the center of the star as its uniformly-spaced counterpart.
-This is especially important when computing the base state pressure from base state density since we enforce a constant base state pressure outside of a prescribed cutoff density radius that could potentially be quite different with the two types of base state spacing at lower resolutions.
+As previously suspected, the simulation using irregularly-spaced base state agrees much closer to the results from using uniform spacing as the resolution increases.
+This is most likely due to the increased resolution of the initial model, which more closely matches the uniformly-spaced counterpart.
+This is especially important when computing the base state pressure from base state density, which is particularly sensitive to coarse resolution near the center of the star.
+\MarginPar{AJN: Simplified this paragraph - OK?}
 
@@
 However, we do observe a 25\% increase in run time of 9.72 s when using irregularly-spaced base state with the new algorithm.
-This can be explained by the irregularly-spaced base state array being much larger in size than its uniformly-spaced counterpart, and thus require additional computation time.
+This can be explained by the irregularly-spaced base state array being much larger in size than its uniformly-spaced counterpart, and thus require additional communication and computation time.
 One possible strategy to significantly reduce the run time is to consider truncating the base state beyond the cutoff density radius.

### 2019-08-02 MAESTROeX_I/equations.tex
MZ read through

@@
 \end{equation}
-where $\gammaonebar$ is the lateral average of $\Gamma_1 = d(\log p)/d(\log\rho) |_s$ (evaluated with entropy, $s$, constant).
+where $\gammaonebar$ is the lateral average of $\Gamma_1 = d(\log p)/d(\log\rho) |_s$ (evaluated with entropy, $s$, constant).  We explored the effect of
+replacing $\Gamma_1$ with $\gammaonebar$ as well as a correction term in paper III. \MarginPar{MZ added this sentence--good to remember this}
 Thermal diffusion is not discussed in this paper, but we have previously described the modifications to the original algorithm required

### 2019-08-02 MAESTROeX_I/intro.tex
MZ read through

@@
 The numerical methodology relies on an explicit Godunov approach for advection, a stiff ODE
-solver for reactions (VODE) \citep{vode}, and multigrid-based linear solvers for the
+solver for reactions (VODE, \citealt{vode}), and multigrid-based linear solvers for the
 pressure-projection steps.  Thus, the time step is limited by an advective CFL constraint based on the 
@@
 MAESTROeX uses MPI+OpenMP parallelism and scales well to over 10,000 MPI processes, with each MPI process supporting tens of threads.
-The resulting code is publicly available on GitHub ({\tt https://github.com/AMReX-Astro/MAESTROeX}),
-uses the Starkiller-Astro microphysics libraries \citep{starkiller} ({\tt https://github.com/starkiller-astro}) ,
-as well as AMReX ({\tt https://github.com/AMReX-Codes/amrex}).
+The resulting code is publicly available on GitHub (\url{https://github.com/AMReX-Astro/MAESTROeX}),
+uses the Starkiller-Astro microphysics libraries (\citealt{starkiller}, \url{https://github.com/starkiller-astro}) ,
+as well as AMReX (\url{https://github.com/AMReX-Codes/amrex}).
 

### 2019-08-02 MAESTROeX_I/equations.tex
removed comments

--- 2019-08-01 MAESTROeX_I/equations.tex
@@
 One way to mathematically think of the difference between $p_0$ and $\pi$ is that $\pi$ controls the velocity evolution 
-in a way that forces the thermodynamic variables $(\rho,h,X_k)$ to evolve in a manner that is consistent with the EOS and $p_0$.\MarginPar{AJN - added this sentence; does it make sense?}
+in a way that forces the thermodynamic variables $(\rho,h,X_k)$ to evolve in a manner that is consistent with the EOS and $p_0$.
 
@@
 \end{equation}
-where $\gammaonebar$ is the lateral average of $\Gamma_1 = d(\log p)/d(\log\rho) |_s$ (evaluated with entropy, $s$, constant).
+where $\gammaonebar$ is the lateral average of $\Gamma_1 = d(\log p)/d(\log\rho) |_s$ (evaluated with entropy, $s$, constant).  We explored the effect of
+replacing $\Gamma_1$ with $\gammaonebar$ as well as a correction term in paper III. 
 Thermal diffusion is not discussed in this paper, but we have previously described the modifications to the original algorithm required

### 2019-08-02 MAESTROeX_I/intro.tex
removed comments

@@
 second-order projection method to integrate the velocity subject to a divergence constraint,
-and a velocity splitting scheme that uses a radially-averaged velocity to hydrodynamically evolve the base state.\MarginPar{AJN changed this sentence.}
+and a velocity splitting scheme that uses a radially-averaged velocity to hydrodynamically evolve the base state.
 The original MAESTRO code was developed in the pure-Fortran 90 FBoxLib software framework, whereas

### 2019-08-02 MAESTROeX_I/results.tex
removed comments

@@
 Limited allocations prevented us from running this simulation further.
-As previously suspected, the simulation using irregularly-spaced base state agrees much closer to the results from using uniform spacing as the resolution increases.
+As previously suspected, the simulation using irregularly-spaced base state agrees much closer with the results from using uniform spacing as the resolution increases.
 This is most likely due to the increased resolution of the initial model, which more closely matches the uniformly-spaced counterpart.
 This is especially important when computing the base state pressure from base state density, which is particularly sensitive to coarse resolution near the center of the star.
-\MarginPar{AJN: Simplified this paragraph - OK?}
 

### 2019-08-10 MAESTROeX_I/submit/paper.tex
some submission files for arxiv

%% This manuscript uses the AASTeX v6.2 LaTeX 2e macros
%%
%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls
%% (Copyright 2000-2015).  See the classfile for details.

%% AASTeX requires revtex4-1.cls (http://publish.aps.org/revtex4/) and
%% other external packages (latexsym, graphicx, amssymb, longtable, and epsf).
%% All of these external packages should already be present in the modern TeX
%% distributions.  If not they can also be obtained at www.ctan.org.

%% The first piece of markup in an AASTeX v6.x document is the \documentclass
%% command. LaTeX will ignore any data that comes before this command. The
%% documentclass can take an optional argument to modify the output style.
%% The command below calls the preprint style  which will produce a tightly
%% typeset, one-column, single-spaced document.  It is the default and thus
%% does not need to be explicitly stated.
%%
%%
%% using aastex version 6.2
\documentclass{aastex62}

%% The default is a single spaced, 10 point font, single spaced article.
%% There are 5 other style options available via an optional argument. They
%% can be envoked like this:
%%
%% \documentclass[argument]{aastex62}
%%
%% where the layout options are:
%%
%%  twocolumn   : two text columns, 10 point font, single spaced article.
%%                This is the most compact and represent the final published
%%                derived PDF copy of the accepted manuscript from the publisher
%%  manuscript  : one text column, 12 point font, double spaced article.
%%  preprint    : one text column, 12 point font, single spaced article.
%%  preprint2   : two text columns, 12 point font, single spaced article.
%%  modern      : a stylish, single text column, 12 point font, article with
%% 		  wider left and right margins. This uses the Daniel
%% 		  Foreman-Mackey and David Hogg design.
%%  RNAAS       : Preferred style for Research Notes which are by design
%%                lacking an abstract and brief. DO NOT use \begin{abstract}
%%                and \end{abstract} with this style.
%%
%% Note that you can submit to the AAS Journals in any of these 6 styles.
%%
%% There are other optional arguments one can envoke to allow other stylistic
%% actions. The available options are:
%%
%%  astrosymb    : Loads Astrosymb font and define \astrocommands.
%%  tighten      : Makes baselineskip slightly smaller, only works with
%%                 the twocolumn substyle.
%%  times        : uses times font instead of the default
%%  linenumbers  : turn on lineno package.
%%  trackchanges : required to see the revision mark up and print its output
%%  longauthor   : Do not use the more compressed footnote style (default) for
%%                 the author/collaboration/affiliations. Instead print all
%%                 affiliation information after each name. Creates a much
%%                 long author list but may be desirable for short author papers
%%
%% these can be used in any combination, e.g.
%%
%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex62}
%%
%% AASTeX v6.* now includes \hyperref support. While we have built in specific
%% defaults into the classfile you can manually override them with the
%% \hypersetup command. For example,
%%
%%\hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
%%
%% will change the color of the internal links to red, the links to the
%% bibliography to green, the file links to cyan, and the external links to
%% magenta. Additional information on \hyperref options can be found here:
%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
%%
%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%
\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}

% for non-stacked fractions
\newcommand{\sfrac}[2]{\mathchoice
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
   \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptscriptfont0 #1}\kern-.2em/
   \kern-.15em\lower.25ex\hbox{\the\scriptscriptfont0 #2}}
  {#1\!/#2}}

\newcommand{\myhalf}{\sfrac{1}{2}}
\newcommand{\thalf}{\sfrac{3}{2}}

\newcommand{\eb}{{\bf{e}}}
\newcommand{\Ub}{{\bf{U}}}
\newcommand{\Ubt}{\widetilde{\Ub}}
\newcommand{\Vb}{{\bf{V}}}
\newcommand{\xb}{{\bf{x}}}

\newcommand{\dr}{\Delta r}
\newcommand{\dt}{\Delta t}

\newcommand{\etarho}{\eta_\rho}
\newcommand{\gammaonebar}{\overline{\Gamma}_1}
\newcommand{\Hnuc}{H_{\rm nuc}}
\newcommand{\omegadot}{\dot\omega}
\newcommand{\pred}{{\rm pred}}
\newcommand{\Sbar}{\overline{S}}

\newcommand{\inp}{\mathrm{in}}
\newcommand{\outp}{\mathrm{out}}
\newcommand{\nph}{{n+\myhalf}}
\newcommand{\nmh}{{n-\myhalf}}
\newcommand{\ow}{\overline{w_0}}
\newcommand{\dw}{\delta w_0}
\newcommand{\uadvone}{\Ub^{\mathrm{ADV},\star}}
\newcommand{\uadvonedag}{\Ub^{\mathrm{ADV},\dagger,\star}}
\newcommand{\uadvtwo}{\Ub^{\mathrm{ADV}}}
\newcommand{\uadvtwodag}{\Ub^{\mathrm{ADV},\dagger}}
\newcommand{\gcc}{\mathrm{g~cm^{-3} }}

% for the red MarginPars
\usepackage{color}
% make the MarginPars look pretty
\setlength{\marginparwidth}{0.5in}
\newcommand{\MarginPar}[1]{\marginpar{\vskip-\baselineskip\raggedright\tiny\sffamily
\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}

\usepackage{url}

\usepackage{mathtools}

\usepackage{multirow}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
\received{Aug 5, 2019}
\revised{XXX X, XXXX}
\accepted{XXX X, XXXX}
%% Command to document which AAS Journal the manuscript was submitted to.
%% Adds "Submitted to " the arguement.
\submitjournal{ApJ}

%% Mark up commands to limit the number of authors on the front page.
%% Note that in AASTeX v6.2 a \collaboration call (see below) counts as
%% an author in this case.
%
%\AuthorCollaborationLimit=3
%
%% Will only show Schwarz, Muench and "the AAS Journals Data Scientist
%% collaboration" on the front page of this example manuscript.
%%
%% Note that all of the author will be shown in the published article.
%% This feature is meant to be used prior to acceptance to make the
%% front end of a long author article more manageable. Please do not use
%% this functionality for manuscripts with less than 20 authors. Conversely,
%% please do use this when the number of authors exceeds 40.
%%
%% Use \allauthors at the manuscript end to show the full author list.
%% This command should only be used with \AuthorCollaborationLimit is used.

%% The following command can be used to set the latex table counters.  It
%% is needed in this document because it uses a mix of latex tabular and
%% AASTeX deluxetables.  In general it should not be needed.
%\setcounter{table}{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% The following section outlines numerous optional output that
%% can be displayed in the front matter or as running meta-data.
%%
%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
\shorttitle{MAESTROeX Low Mach Number Astrophysics}
\shortauthors{Fan et al.}
%%
%% You can add a light gray and diagonal water-mark to the first page
%% with this command:
% \watermark{text}
%% where "text", e.g. DRAFT, is the text to appear.  If the text is
%% long you can control the water-mark size with:
%  \setwatermarkfontsize{dimension}
%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% This is the end of the preamble.  Indicate the beginning of the
%% manuscript itself with \begin{document}.

\begin{document}

\title{MAESTROeX: A Massively Parallel Low Mach Number Astrophysical Solver}

%% LaTeX will automatically break titles if they run longer than
%% one line. However, you may use \\ to force a line break if
%% you desire. In v6.2 you can include a footnote in the title.

%% A significant change from earlier AASTEX versions is in the structure for
%% calling author and affilations. The change was necessary to implement
%% autoindexing of affilations which prior was a manual process that could
%% easily be tedious in large author manuscripts.
%%
%% The \author command is the same as before except it now takes an optional
%% arguement which is the 16 digit ORCID. The syntax is:
%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
%%
%% This will hyperlink the author name to the author's ORCID page. Note that
%% during compilation, LaTeX will do some limited checking of the format of
%% the ID to make sure it is valid.
%%
%% Use \affiliation for affiliation information. The old \affil is now aliased
%% to \affiliation. AASTeX v6.2 will automatically index these in the header.
%% When a duplicate is found its index will be the same as its previous entry.
%%
%% Note that \altaffilmark and \altaffiltext have been removed and thus
%% can not be used to document secondary affiliations. If they are used latex
%% will issue a specific error message and quit. Please use multiple
%% \affiliation calls for to document more than one affiliation.
%%
%% The new \altaffiliation can be used to indicate some secondary information
%% such as fellowships. This command produces a non-numeric footnote that is
%% set away from the numeric \affiliation footnotes.  NOTE that if an
%% \altaffiliation command is used it must come BEFORE the \affiliation call,
%% right after the \author command, in order to place the footnotes in
%% the proper location.
%%
%% Use \email to set provide email addresses. Each \email will appear on its
%% own line so you can put multiple email address in one \email call. A new
%% \correspondingauthor command is available in V6.2 to identify the
%% corresponding author of the manuscript. It is the author's responsibility
%% to make sure this name is also in the author list.
%%
%% While authors can be grouped inside the same \author and \affiliation
%% commands it is better to have a single author for each. This allows for
%% one to exploit all the new benefits and should make book-keeping easier.
%%
%% If done correctly the peer review system will be able to
%% automatically put the author and affiliation information from the manuscript
%% and save the corresponding author the trouble of entering it by hand.

\correspondingauthor{Duoming Fan; DFan@lbl.gov}

\author[0000-0002-3246-4315]{Duoming Fan}
\affil{Lawrence Berkeley National Laboratory \\
Center for Computational Sciences and Engineering \\
One Cyclotron Road, MS 50A-3111 \\
Berkeley, CA 94720, USA}

\author[0000-0003-1791-0265]{Andrew Nonaka}
\affil{Lawrence Berkeley National Laboratory \\
Center for Computational Sciences and Engineering \\
One Cyclotron Road, MS 50A-3111 \\
Berkeley, CA 94720, USA}

\author[0000-0003-2103-312X]{Ann S. Almgren}
\affil{Lawrence Berkeley National Laboratory \\
Center for Computational Sciences and Engineering \\
One Cyclotron Road, MS 50A-3111 \\
Berkeley, CA 94720, USA}

\author[0000-0002-1530-781X]{Alice Harpole}
\affil{Stony Brook University \\
Department of Physics and Astronomy \\
Stony Brook, NY 11794-3800, USA}

\author[0000-0001-8401-030X]{Michael Zingale}
\affil{Stony Brook University \\
Department of Physics and Astronomy \\
Stony Brook, NY 11794-3800, USA}

%% Note that the \and command from previous versions of AASTeX is now
%% depreciated in this version as it is no longer necessary. AASTeX
%% automatically takes care of all commas and "and"s between authors names.

%% AASTeX 6.2 has the new \collaboration and \nocollaboration commands to
%% provide the collaboration status of a group of authors. These commands
%% can be used either before or after the list of corresponding authors. The
%% argument for \collaboration is the collaboration identifier. Authors are
%% encouraged to surround collaboration identifiers with ()s. The
%% \nocollaboration command takes no argument and exists to indicate that
%% the nearby authors are not part of surrounding collaborations.

%% Mark off the abstract in the ``abstract'' environment.
\begin{abstract}
We present MAESTROeX, a massively parallel solver for low Mach number astrophysical flows.
The underlying low Mach number equation set allows for efficient, long-time integration for highly subsonic flows compared to compressible approaches.
MAESTROeX is suitable for modeling full spherical stars as well as well as planar simulations of dynamics within localized regions of a star,
and can robustly handle several orders of magnitude of density and pressure stratification.
Previously, we have described the development of the predecessor of MAESTROeX, called MAESTRO, in a series of papers. 
Here, we present a new, greatly simplified temporal integration scheme that retains the same order of accuracy as our previous approaches. 
We also explore the use of alternative spatial mapping of the one-dimensional base state onto the full Cartesian grid.
The code leverages the new AMReX software framework for block-structured adaptive mesh refinement (AMR) applications, allowing for scalability to large fractions of leadership-class machines.
Using our previous studies on the convective phase of single-degenerate progenitor models of Type Ia supernovae as a guide, we characterize the performance of the code and validate the new algorithmic features.  Like MAESTRO, MAESTROeX is fully open source.
\end{abstract}

%% Keywords should appear after the \end{abstract} command.
%% See the online documentation for the full list of available subject
%% keywords and the rules for their use.
\keywords{convection, hydrodynamics, methods: numerical, nuclear reactions, nucleosynthesis, abundances, supernovae: general}

%% From the front matter, we move on to the body of the paper.
%% Sections are demarcated by \section and \subsection, respectively.
%% Observe the use of the LaTeX \label
%% command after the \subsection to give a symbolic KEY to the
%% subsection for cross-referencing in a \ref command.
%% You can use LaTeX's \ref and \label commands to keep track of
%% cross-references to sections, equations, tables, and figures.
%% That way, if you change the order of any elements, LaTeX will
%% automatically renumber them.
%%
%% We recommend that authors also use the natbib \citep
%% and \citet commands to identify citations.  The citations are
%% tied to the reference list via symbolic KEYs. The KEY corresponds
%% to the KEY in the \bibitem in the reference list below.

\section{Introduction} \label{sec:intro}
Many astrophysical flows are highly subsonic.  In this regime,
sound waves carry sufficiently little energy that they do not
significantly affect the convective dynamics of the system.  In many
of these flows, modeling long-time convective dynamics are of
interest, and numerical approaches based on compressible hydrodynamics
are intractable, even on modern supercomputers.  One approach to this
problem is to use low Mach number models.  In a low Mach number
approach, sound waves are eliminated from the governing equations
while retaining compressibilitiy effects due to, e.g., nuclear energy
release, stratification, compositional changes, and thermal diffusion.  When the Mach
number (the ratio of the characteristic fluid velocity over the
characteristic sound speed; Ma $= U/c$) is small, the resulting system
can be numerically integrated with much larger time steps than a
compressible model.  Specifically, the time step increase is at least
a factor of $\sim 1/{\rm Ma}$ larger.  Each time step is more
computationally expensive due to the presence of additional linear
solves, but for many problems of interest the overall gains in
efficiency can easily be an order of magnitude or more.

Low Mach number models have been developed for a variety of contexts
including combustion \citep{day2000numerical}, terrestrial atmospheric
modeling \citep{durran:1989,oneill:2014,duarte2015low}, and elastic
solids \citep{abbate2017all}.  For astrophysical applications, a
number of approaches to modeling low Mach number flows have been
developed in recent years.  One of the more similar approaches to ours is
\cite{Lin:2006}, however this approach is only first-order accurate and does
not account for atmospheric expansion.  There are also semi-implicit all-Mach
number solvers, where the Euler equations are split into an acoustic
part and an advective part
\citep{Kwatra2009,Degond2009,Cordier2012,Haack2012,Happenhofer2013,Chalons2016,Padioleau2019}.
The fast acoustic waves are then solved using implicit time
integration, while the slow material waves are solved explicitly.
Another approach is to use preconditioned all-Mach number solvers
\citep{Miczek2014,Barsukow2016}, where the numerical flux is
multiplied by a preconditioning matrix.  This reduces the stiffness of
the system at low Mach numbers, while retaining the correct scaling
behavior. In the reduced speed of sound technique (RSST) and related
methods, the speed of sound is artificially reduced by including a
suitable scaling factor in the continuity equation, reducing the
restriction on the size of the time step
\citep{Rempel2005,Hotta2012,Wang2015,Takeyama2017,Iijima2018}.
Finally, there are fully implicit time integration codes for the
compressible Euler equations
\citep{Viallet2011,kifonidis:2012,Viallet2015,Goffrey2016}.  The MUSIC code
uses fully implicit time integration for the compressible Euler
equations, which therefore allows for arbitrarily large time steps.

Previously, we developed the low Mach number astrophysical solver, MAESTRO.
MAESTRO is a block-structured, Cartesian grid finite-volume, adaptive mesh refinement (AMR)
code that has been successfully used for many years for a number of applications, detailed below.
Unlike several of the references above, MAESTRO is not an all-Mach solver, but is suitable for
flows where the Mach number is small ($\sim 0.1$ or smaller).
Furthermore, the low Mach number model in MAESTRO is specifically designed for, but not limited
to, astrophysical settings with significant atmospheric stratification.
This includes full spherical stars, as well as planar simulations of dynamics within localized
regions of a star.
The numerical methodology relies on an explicit Godunov approach for advection, a stiff ODE
solver for reactions (VODE, \citealt{vode}), and multigrid-based linear solvers for the
pressure-projection steps.  Thus, the time step is limited by an advective CFL constraint based on the 
fluid velocity, not the sound speed.
Central to the algorithm are time-varying, one-dimensional stratified background (or base) state
density and pressure fields that are held in hydrostatic equilibrium.
The base state density couples to the full state solution through buoyancy terms in the momentum equation,
and the base state pressure couples to the full state solution by constraining the evolution of the
thermodynamic variables to match this pressure.
The time-advancement strategy uses Strang splitting to integrate the thermodynamic variables, a
second-order projection method to integrate the velocity subject to a divergence constraint,
and a velocity splitting scheme that uses a radially-averaged velocity to hydrodynamically evolve the base state.
The original MAESTRO code was developed in the pure-Fortran 90 FBoxLib software framework, whereas
MAESTROeX is developed in the C++/F90 AMReX framework \citep{AMReX,AMReX_JOSS}.

The key numerical developments of the original MAESTRO algorithm are presented in a series of
papers which we refer to as Papers I-V:
\begin{itemize}
\item In Paper I \citep{MAESTRO_I}, we derive the low Mach number equation set for stratified
environments from the fully compressible equations.
\item In Paper II \citep{MAESTRO_II}, we incorporate the effects of atmospheric expansion
through the use of a time-dependent background state.
\item In Paper III \citep{MAESTRO_III}, we incorporate reactions and the associated coupling
to the hydrodynamics.
\item In Paper IV \citep{MAESTRO_IV}, we describe our treatment of spherical stars in a
three-dimensional Cartesian geometry.
\item In Paper V \citep{MAESTRO_V}, we describe the use of block-structured adaptive mesh
refinement to focus spatial resolution in regions of interest.
\end{itemize}

Since then, there have been many scientific investigations using MAESTRO, which have included additional algorithmic enhancements.  Topics include:
\begin{itemize}
\item The convective phase preceding Chandrasekhar mass models for type Ia supernovae \citep{MAESTRO_convection,MAESTRO_AMR,MAESTRO_CASTRO}.
\item Convection in massive stars \citep{Gilet:2013,gilkis:2016}.
\item Sub-Chandrasekhar white dwarfs \citep{subChandra_I,subChandra_II}.
%  In the latter paper we introduced an optional modification to the momentum equation of velocity constraint that conserves total energy, can give results closer to compressible codes of convective dynamics, including low density regions at the surface of a star.
\item Type I X-ray bursts \citep{XRB_I,XRB_II,XRB_III}.
%  In \cite{XRB_I} we discuss the modifications required for implicit thermal conduction, as well as an additional ``volume discrepancy'' modification to the velocity constraint to force the species and enthalpy to evolve in a manner consistent with the background pressure.
\end{itemize}

In this paper, we present new algorithmic methodology that improves upon Paper V in a number of ways.
First, the overall temporal algorithm has been greatly simplified without compromising second-order accuracy.
The key design decisions were to eliminate the splitting of the velocity into average and perturbational components,
and also to replace the hydrodynamic evolution of the base state with a predictor-corrector approach.
Not only does this greatly simplify the dynamics of the base
state, but this treatment is more amenable to higher-order multiphysics coupling strategies
based on method-of-lines integration.
In particular, schemes based on deferred corrections \citep{dutt2000spectral} have been used to generate 
high-order temporal integrators for problems of reactive flow and low Mach number combustion \citep{pazner2016high,nonaka2018conservative}.
Second, we explore the effects of alternative spatial mapping routines for coupling the base state and the Cartesian grid state for spherical problems.
Finally, we examine the performance of our new MAESTROeX implementation in the new C++/F90 AMReX public software library \citep{AMReX,AMReX_JOSS}.
MAESTROeX uses MPI+OpenMP parallelism and scales well to over 10,000 MPI processes, with each MPI process supporting tens of threads.
The resulting code is publicly available on GitHub (\url{https://github.com/AMReX-Astro/MAESTROeX}),
uses the Starkiller-Astro microphysics libraries (\citealt{starkiller}, \url{https://github.com/starkiller-astro}) ,
as well as AMReX (\url{https://github.com/AMReX-Codes/amrex}).

The rest of this paper is organized as follows.
In Section \ref{sec:equations} we review our model for stratified low Mach number astrophysical flow.
In Section \ref{eq:algorithm} we present our numerical algorithm in detail, highlighting the new temporal integration scheme as well as spatial base state mapping options.
In Section \ref{sec:results} we validate our new approach and examine the performance of our algorithm on full spherical star problems used in previous scientific investigations.
We conclude in Section \ref{sec:conclusions}.

\section{Governing Equations}\label{sec:equations}
Low Mach number models for reacting flow were originally derived using asymptotic analysis
\citep{rehm1978equations,majda1985derivation} and used in terrestrial combustion applications
\citep{knio1999semi,day2000numerical}.  These models have been extended to nuclear flames
in astrophysical environments using adaptive algorithms in space and time \citep{Bell:2004}.
In Papers I-III, we extended this work and the atmospheric model by \citet{durran:1989} by deriving a model and algorithm suitable for stratified astrophysical flow.
We take the standard equations of reacting, compressible flow, and recast the equation
of state (EOS) as a divergence constraint on the velocity field.
The resulting model is a series of evolution equations for mass, momentum, and energy, subject
to an additional constraint on velocity.  The evolution equations are
\begin{eqnarray}
\frac{\partial(\rho X_k)}{\partial t} &=& -\nabla\cdot(\rho X_k\Ub) + \rho\omegadot_k,\label{eq:species}\\
\frac{\partial\Ub}{\partial t} &=& -\Ub\cdot\nabla\Ub  - \frac{\beta_0}{\rho}\nabla\left(\frac{\pi}{\beta_0}\right) - \frac{\rho-\rho_0}{\rho} g\eb_r,\label{eq:momentum}\\
\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc.\label{eq:enthalpy}
\end{eqnarray}
Here $\rho$, $\Ub$, and $h$ are the mass density,
velocity and specific enthalpy, respectively, and
$X_k$ are the mass fractions of species $k$ with associated
production rate $\omegadot_k$ and energy release per time per unit mass $\Hnuc$.
The species are constrained such that $\sum_k X_k = 1$ giving $\rho = \sum_k (\rho X_k)$ and
\begin{equation}
\frac{\partial\rho}{\partial t} = -\nabla\cdot(\rho\Ub).
\end{equation}
The total pressure is decomposed into a one-dimensional hydrostatic base state
 pressure, $p_0 = p_0(r,t)$, and a dynamic pressure, $\pi = \pi(\xb,t)$, such that
$p = p_0 + \pi$ and $|\pi|/p_0 = \mathcal{O}({\rm Ma}^2)$ (we use $\xb$ to represent the Cartesian coordinate
directions of the full state and $r$ to represent the radial coordinate direction for the base state).
One way to mathematically think of the difference between $p_0$ and $\pi$ is that $\pi$ controls the velocity evolution 
in a way that forces the thermodynamic variables $(\rho,h,X_k)$ to evolve in a manner that is consistent with the EOS and $p_0$.

By comparing the momentum equation (\ref{eq:momentum}) to the momentum equation used in equation (2) in Paper V, we
note that we are using a formulation that enforces conservation of total energy in the
low Mach number system in the absence of external heating or viscous terms \citep{kleinpauluis,Vasil2013}.
We have previously validated this approach in modeling sub-Chandrasekhar white dwarfs using MAESTRO \citep{subChandra_II}.
We also define a one-dimensional base state density, $\rho_0 = \rho_0(r,t)$, that represents the lateral average (see Section \ref{Sec:Spatial}) of $\rho$ and is in hydrostatic equilibrium with $p_0$, i.e.,
\begin{equation}
\nabla p_0 = -\rho_0 g\eb_r, \label{eq:HSE}
\end{equation}
where $g=g(r,t)$ is the magnitude of the gravitational acceleration and $\eb_r$ is the unit vector in the outward radial direction.
Here $\beta_0$ is a density-like variable that carries background stratification, defined as
\begin{equation}
\beta_0(r,t) = \rho_0(0,t)\exp\left(\int_0^r\frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial r'}dr'\right),
\end{equation}
where $\gammaonebar$ is the lateral average of $\Gamma_1 = d(\log p)/d(\log\rho) |_s$ (evaluated with entropy, $s$, constant).  We explored the effect of
replacing $\Gamma_1$ with $\gammaonebar$ as well as a correction term in paper III. 
Thermal diffusion is not discussed in this paper, but we have previously described the modifications to the original algorithm required
for implicit thermal diffusion in \cite{XRB_I}; inclusion of these effects in the new algorithm presented here is straightforward.

Mathematically, equations (\ref{eq:momentum})-(\ref{eq:enthalpy}) must still be closed by the EOS.
This is done by taking the Lagrangian derivative of the EOS for pressure as a function of the thermodynamic variables,
substituting in the equations of motion for mass and energy,
and requiring that the pressure is a prescribed function of altitude and time based on the hydrostatic equilibrium condition.
See Papers I and II for details of this derivation.
The resulting equation is a divergence constraint on the velocity field,
\begin{equation}
\nabla\cdot(\beta_0\Ub) = \beta_0\left(S - \frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial t}\right).\label{eq:U divergence}
\end{equation}
The expansion term, $S$, incorporates local compressibility effects due to compositional changes and heat release from reactions,
\begin{equation}
S = -\sigma\sum_k\xi_k\omegadot_k + \frac{1}{\rho p_\rho}\sum_k p_{X_k}\omegadot_k + \sigma\Hnuc,\label{eq:S}
\end{equation}
where
$p_{X_k} \equiv \left. \partial p / \partial X_k \right|_{\rho,T,X_{j,j\ne k}}$,
$\xi_k \equiv \left. \partial h /\partial X_k \right |_{p,T,X_{j,j\ne k}}$,
$p_\rho \equiv \left.\partial p/\partial \rho \right |_{T, X_k}$, and
$\sigma \equiv p_T/(\rho c_p p_\rho)$, with $p_T \equiv \left. \partial p / \partial
T \right|_{\rho, X_k}$ and $c_p \equiv \left.  \partial h / \partial T
\right|_{p,X_k}$ is the specific heat at constant pressure.

To summarize, we model evolution equations for momentum, mass, and energy, equations (\ref{eq:momentum})-(\ref{eq:enthalpy}) subject to a divergence constraint on the velocity, equation (\ref{eq:U divergence}), and the hydrostatic equilibrium condition, equation (\ref{eq:HSE}).

\section{Numerical Algorithm}\label{eq:algorithm}
\subsection{Spatial Discretization}\label{Sec:Spatial}
The spatial discretization and adaptive mesh refinement methodology remains unchanged from Paper V.
We now summarize some of the key points here before describing the new temporal integrator in the next section.
We recommend the reader review Section 3 of Paper V for further details.

We shall refer to local atmospheric flows in two and three dimensions as problems in ``planar'' geometry, and full-star flows
in three dimensions as problems in ``spherical'' geometry.
The solution in both cases consists of the Cartesian grid solution
and the one-dimensional base state solution.
Figure \ref{Fig:BaseGrid} illustrates the relationship between the base state and the Cartesian grid state for both planar and spherical geometries in the presence of spatially adaptive grids. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[tb]
\centering
\includegraphics[height=2.0in]{base_grid.eps} \hspace{0.5in}
\includegraphics[height=2.0in]{base_spherical.eps}
\caption{\label{Fig:BaseGrid}  
(Left) For multi-level problems in planar geometry, we force a direct alignment
between the base state cell centers and the Cartesian grid cell centers by 
allowing the radial base state spacing to change with space and time.
(Right) For multi-level problems in spherical geometry, since there is no direct alignment
between the base state cell centers and the Cartesian grid cell centers, we choose to fix
the radial base state spacing across levels. Reprinted from Paper V \citep{MAESTRO_V}. }
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
One of the key numerical modules is the ``lateral average'', which computes the average over a 
layer of constant radius of a Cartesian grid variable and stores the result in a one-dimensional base state array.
In planar geometries, this is a straightforward arithmetic average of values in cells at
a particular height since the base state cell centers are in alignment
with the Cartesian grid cell centers.
However for spherical problems, the procedure is much more complicated.
In Section 4 of Paper V, we describe how there is a finite, easily computable set of radii that any 
there-dimensional Cartesian cell-center can map to.  
Specifically, for every three-dimensional Cartesian cell, there exists an integer $m$ such that the distance
from the cell center to the center of the star is given by 
\begin{equation}
\hat{r}_m=\Delta x\sqrt{0.75+2m}.\label{eqn:radii}
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[tb]
\centering
\includegraphics[height=2.5in]{base_spherical_new.eps}
\caption{\label{Fig:NewBaseGrid}  
A direct mapping between the base state cell centers (red squares) and the Cartesian grid cell centers (blue crosses) 
is enforced by computing the average of the grid cell centers that share the same radial distance from 
the center of the star.}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Figure \ref{Fig:NewBaseGrid} is a two-dimensional illustration 
(two dimensions is chosen in the figure for ease of exposition; this mapping is only used for three-dimensional spherical stars) 
of the relationship between the Cartesian grid state and the one-dimensional base state array.
We compute the lateral average by first summing the values in all the cells associated with each radius,
dividing by the number of contributing cells to obtain the arithmetic average, and then use quadratic interpolation to map this data onto a one-dimensional base state.
Previously, for spherical problems MAESTRO only allowed for a base state with constant $\Delta r$ (typically equal to $\Delta x/5$).

The companion ``fill'' module maps a base state array onto the full Cartesian grid state.
For planar problems, direct injection can be used due to the perfect alignment of the base state and Cartesian grid state.
For spherical problems, quadratic interpolation of the base state is used to assign values to each Cartesian cell center.

In this paper we explore a new option to retain an irregularly-spaced base state to eliminate mapping errors from the ``fill'' module.
For the lateral average, as before we first sum the values in all the cells associated with each radius and divide
by the number of contributing cells to obtain the arithmetic average.  However we do not interpolate this onto a uniformly spaced
base state and retain the use of this irregularly-spaced base state.
The advantage with this approach is that the fill module does not require any interpolation.
A potential benefit to eliminating the mapping error is to consider a spherical star in hydrostatic equilibrium at rest.
In the absence of reactions, the star should remain at rest.
The buoyancy forcing term in the momentum equation contains $\rho-\rho_0$.  With the original scheme, interpolation errors 
in computing $\rho_0$ by averaging would cause artificial acceleration in the velocity field due to the interpolation error 
from the Cartesian grid to and from the radial base state.  By retaining the radial base state as an irregularly spaced 
array, the effects due to interpolation error are nearly eliminated, but not completely eliminated since there are still 
machine precision effects resulting from averaging a large number of numerical values.
We note that $\Delta r$ decreases as the base state moves further from the center of the star, 
which results in far more total cells in the irregularly-spaced array than the previous uniformly-spaced array.

\subsection{Temporal Integration Scheme}\label{Sec:Temporal Integration Scheme}
We now describe the new temporal integration scheme, noting that it can be used for the original base state mapping 
(with uniform base state grid spacing) as well as the new irregularly spaced base state mapping.
Previously we adopted an approach where we split the velocity into a base state component, $w_0(r,t)$, 
and a local velocity $\Ubt(\xb,t)$, so that
\begin{equation}
\Ub = \Ubt(\xb,t) + w_0(r,t)\eb_r, \label{eq:velsplit}
\end{equation}
where $\eb_r$ is the normal vector in the outward radial direction.
We used $w_0$ to provide an estimate of the base state density evolution over a time step.
This resulted in some unnecessary complications to the temporal integration scheme including
base state advection modules for density, enthalpy, and velocity, as well as more 
cumbersome split velocity dynamics evolution equations.
Our new temporal integration scheme uses full velocities for scalar and velocity advection,
and only uses the above splitting to satisfy the velocity divergence constraint due to boundary considerations
at the edge of the star.
This results in a much simpler numerical scheme than the one from Paper V 
since we use the velocity directly rather than more complex terms involving the perturbational velocity.
Additionally, the new scheme uses a simpler predictor-corrector approach to the base state density and pressure that no 
longer requires evolution equations and numerical discretizations to update the base state, greatly 
simplifying the algorithm while retaining the same overall second-order accuracy in time.

At the beginning of each time step we have the cell-centered Cartesian grid state,
$(\Ub,\rho X_k,\rho h)^n$, and nodal Cartesian grid state, $\pi^{n-\myhalf}$, and base state $(\rho_0,p_0)^n$.
At any time, the associated density, composition, and enthalpy can be trivially computed using, e.g.,
\begin{equation}
\rho^n = \sum_k(\rho X_k)^n, \quad
X_k^n = (\rho X_k)^n / \rho^n, \quad
h^n = (\rho h)^n / \rho^n.
\end{equation}
Temperature is computed using the equation of state\footnote{As described in Paper V, for planar problems we compute temperature using $h$ instead of $p_0$, since we have successfully developed volume discrepancy schemes to effectively couple the enthalpy to the rest of the solution; see \cite{XRB_I}.  We are still exploring this option for spherical stars.}, e.g.,
$T = T(\rho,p_0,X_k)$, where $p_0$ has been mapped to the Cartesian grid using the fill module,
and ($\gammaonebar,\beta_0)$ are similarly computed from $(\rho,p_0,X_k)$;
see Appendix A of Paper I and Appendix C of Paper III for details on how $\beta_0$ is computed.

The overall flow of the algorithm begins with a second-order Strang splitting approach to integrate the advection-reaction system for 
the thermodynamic variables $(\rho X_k, \rho h)$, followed by a second-order projection methodology to integrate the velocities subject to a divergence constraint.  Within the thermodynamic variable update we use a predictor-corrector approach to achieve second-order accuracy in time.
To summarize:
\begin{itemize}
\item In {\bf Step 1} we react the thermodynamic variables over the first $\Delta t/2$ interval.
\item In {\bf Steps 2--4} we advect the thermodynamic variables over $\Delta t$.  Specifically, we compute an estimate for the expansion term, $S$, compute face-centered, time-centered velocities that satisfy the divergence constraint, and then advect the thermodynamic variables.
\item In {\bf Step 5} we react the thermodynamic variables over the second $\Delta t/2$ interval\footnote{After this step we could skip to the velocity advance in {\bf Steps 10--11}, however the overall scheme would be only first-order in time, so {\bf Steps 6-9} can be thought of as a trapezoidal corrector step.}.
\item In {\bf Steps 6--8} we redo the advection in {\bf Steps 2--4} but are able to use the trapezoidal rule to time-center certain quantities such as $S$, $\rho_0$, etc.
\item In {\bf Step 9} we redo the reactions from {\bf Step 5} beginning with the improved results from {\bf Steps 6--8}.
\item In {\bf Steps 10--11} we advect the velocity, and then project these velocities so they satisfy the divergence constraint while updating $\pi$.
\end{itemize}

There are a few key numerical modules we use in each time step.
\begin{itemize}
\item {\bf Average}$[\phi]\rightarrow[\overline\phi]$ computes the lateral average of a quantity over a layer at constant radius $r$, as described above in Section \ref{Sec:Spatial}.
\item {\bf Enforce HSE}$[\rho_0]\rightarrow[p_0]$ computes the base state pressure, $p_0$, from a base state density, $\rho_0$ by integrating the hydrostatic equilibrium condition in one dimension. 
This follows equation (A10) in Paper V, noting that for the irregularly spaced base state case, $\Delta r$ is not constant, where $\Delta r_{j+1/2} = r_{j+1} - r_j$ for cell face with index $j+1/2$.  
The base state pressure remains equal to a constant value at the location of a prescribed cutoff density outward for the entire simulation.
\item {\bf React State}$[(\rho X_k)^{\rm in},(\rho h)^{\rm in},p_0]\rightarrow[(\rho X_k)^{\rm out},(\rho h)^{\rm out},(\rho\dot\omega),(\rho\Hnuc)]$ 
uses VODE \citep{vode} to integrate the species and enthalpy due to reactions over $\Delta t/2$ by solving
\begin{equation}
\frac{dX_k}{dt} = \dot\omega_k(\rho,X_k,T); \qquad
\frac{dT}{dt} = \frac{1}{c_p}\left(-\sum_k\xi_k\dot\omega_k + H_{\rm nuc})\right)
\end{equation}
The inputs are the species, enthalpy, and base state pressure, and the outputs are the species, enthalpy, reaction rates, and nuclear energy generation rate.
See Paper III for details.
\end{itemize}
Each time step is constrained by the standard advective CFL condition,
\begin{equation}
\Delta t = \sigma^{\rm CFL} \min_i(\Delta x / U_i),
\end{equation}
where for our simulations we typically use $\sigma^{\rm CFL}\sim 0.7$ and the minimum is taken over all spatial directions over all cells.
There are additional constraints on the time step that are typically much less restrictive than the advective CFL including
the acceleration due to the buoyancy force (sometimes in effect when the velocity is approximately zero at the start of some simulations) 
and the local magnitude of the divergence constraint (to prevent too much mass evacuation from a cell in a time step); see Section 3.4 in Paper III for details.

In stratified low Mach number models, due to the extreme variation in density, the velocity can become very large in low density regions at the edge of the star.
These large velocities can severely affect the time step, so
throughout Papers II-V, we have employed two techniques to help control these dynamics without significantly 
affecting the dynamics in the convective region.
First, we use a cutoff density technique, where we hold the density constant outside a specified radius (typically near where the density is $\sim$4 orders of magnitude smaller than the largest densities in the simulation).
Second, we employ a sponge technique where we artificially damp the velocities near and beyond the cutoff region.
For more details, refer to Paper V and the previous references cited within.

Beginning with $(\Ub,\rho X_k,\rho h)^n$, $\pi^{n-\myhalf}$, and $(\rho_0,p_0)^n$,
the temporal integration scheme contains the following steps:
\begin{description}

%--------------------------------------------------------------------------
% STEP 1
%--------------------------------------------------------------------------
\item[Step 1] {\em React the thermodynamic variables over the first $\Delta t / 2$ interval.}

Call {\bf React State}$[(\rho X_k)^n, (\rho h)^n, p_0^n] \rightarrow [(\rho X_k)^{(1)}, (\rho h)^{(1)}, (\rho \omegadot_k)^{(1)}, (\rho \Hnuc)^{(1)}]$.

%--------------------------------------------------------------------------
% STEP 2
%--------------------------------------------------------------------------

\item[Step 2] {\em Compute the time-centered expansion term, $S^{\nph,\star}$.}

We compute an estimate for the time-centered expansion term in the velocity
divergence constraint.  Following \citet{Bell:2004}, we extrapolate
to the half-time using $S$ at the previous and current
time steps,
\begin{equation}
S^{\nph,\star} = S^n + \frac{\dt^n}{2} \left(\frac{S^n - S^{n-1}}{\dt^{n-1}}\right).
\end{equation}
Note that in the first time step we average $S^0$ and $S^1$ from the
initialization step.

%--------------------------------------------------------------------------
% STEP 3
%--------------------------------------------------------------------------
\item[Step 3] {\em Construct a face-centered, time-centered advective velocity, $\uadvone$.}

The construction of face-centered time-centered states used to discretize the
advection terms for velocity, species, and enthalpy, are performed using
a standard multidimensional corner transport upwind approach
\citep{colella1990multidimensional,saltzman1994unsplit} with the piecewise-parabolic method (PPM)
one-dimensional tracing \citep{colella1984piecewise}.  The full details of this
Godunov advection approach for all steps in this algorithm are described 
in Appendix A of \cite{XRB_III}.
Here we use equation (\ref{eq:momentum}) to compute face-centered, time-centered velocities, $\uadvonedag$.
The $\dagger$ superscript refers to the fact that the predicted velocity field does not satisfy the divergence constraint,
\begin{equation}
\nabla \cdot \left(\beta_0^n \uadvone\right) = \beta_0^n \left[S^{\nph,\star} - \frac{1}{\gammaonebar^np_0^n}\left(\frac{\partial p_0}{\partial t}\right)^{n-\myhalf} \right].\label{eq:div1}
\end{equation}
 We project $\uadvonedag$ onto the space of velocities that satisfy the constraint to obtain $\uadvone$.
Each projection step in the algorithm involves the solution of a variable-coefficient Poisson solve using multigrid.
Note that we still employ velocity-splitting as described by equation (\ref{eq:velsplit}) for this step 
in order to enforce the appropriate behavior of the system near the edge of the star as determined by the cutoff density.
The details of this ``MAC'' projection are provided in Appendix \ref{Sec:Projection}.

%--------------------------------------------------------------------------
% STEP 4
%--------------------------------------------------------------------------
\item[Step 4] {\em Advect the thermodynamic variables over a time interval of $\dt.$}

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \Alph{enumi}}}

\item Update $(\rho X_k)$ using a discretized version of
%
\begin{equation}
\frac{\partial(\rho X_k)}{\partial t} = -\nabla\cdot(\rho X_k\Ub),
\end{equation}
%
where the reaction terms have been omitted since they were already 
accounted for in {\bf React State}.  The update consists of two steps:

\begin{enumerate}
\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}

\item Compute the face-centered, time-centered species, $(\rho X_k)^{\nph,\pred,\star}$,
  for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
  As described in Paper V, for robust numerical slope limiting we predict 
  $\rho'^n=\rho^n-\rho_0^n$ and $X_k^n$ to faces
  and here we spatially interpolate $\rho_0^n$ to faces to assemble the fluxes.

\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2),\star}$ using
\begin{equation}
(\rho X_k)^{(2),\star} = (\rho X_k)^{(1)}
  - \dt \left\{ \nabla \cdot \left[ \uadvone (\rho X_k)^{\nph,\pred,\star} \right] \right\},
\end{equation}

\end{enumerate}

\item Update $\rho_0$ by calling {\bf Average}$[\rho^{(2),\star}]\rightarrow[\rho_0^{n+1,\star}]$.

\item Update $p_0$ by calling {\bf Enforce HSE}$[\rho_0^{n+1,\star}] \rightarrow [p_0^{n+1,\star}]$.

\item Update the enthalpy using a discretized version of equation
%
\begin{equation}
\frac{\partial(\rho h)}{\partial t} = -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc,
\end{equation}
%
again omitting the reaction terms since we already accounted for
them in {\bf React State}.  This equation takes the form:
\begin{equation}
\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\rho h\Ub) + \frac{\partial p_0}{\partial t} + (\Ub \cdot \eb_r) \frac{\partial p_0}{\partial r}.
\end{equation}
For spherical geometry, we solve the analytically equivalent form,
\begin{equation}
\frac{\partial (\rho h)}{\partial t}  = - \nabla \cdot (\rho h\Ub) + \frac{\partial p_0}{\partial t} + \nabla \cdot (\Ub p_0) - p_0 \nabla \cdot \Ub.
\end{equation}
The update consists of two steps:

\begin{enumerate}
\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}

\item Compute the face-centered, time-centered enthalpy, $(\rho h)^{\nph,\pred,\star},$
  for the conservative update of $(\rho h)^{(1)}$ using using a Godunov approach \citep{XRB_III}.
  As described in Paper V, for robust numerical slope limiting 
  we predict $(\rho h)'^n=(\rho h)^n-(\rho h)_0^n$ to faces,
  where $(\rho h)_0^n$ is obtained by calling {\bf Average}$[(\rho h)^n]\rightarrow[(\rho h)_0^n]$,
  and here we spatially interpolate $(\rho h)_0^n$ to faces to assemble the fluxes.

\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2),\star}$ using
\begin{equation}
(\rho h)^{(2),\star}
= (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvone (\rho h)^{\nph,\pred,\star} \right] \right\} + \Delta t\frac{Dp_0}{Dt}
\end{equation}
where here
\begin{equation}
\frac{Dp_0}{Dt} =
\begin{cases}
\frac{p_0^{n+1,*} - p_0^n}{\Delta t} + \left(\uadvone \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^{n}& {\rm (planar)} \\
\frac{p_0^{n+1,*} - p_0^n}{\Delta t} + \left[ \nabla \cdot \left (\uadvone p_0^{\nph} \right ) - p_0^{\nph} \nabla \cdot \uadvone \right]& {\rm (spherical)}
\end{cases}
,
\end{equation}
and $p_0^\nph = (p_0^n+p_0^{n+1,*})/2$.
\end{enumerate}
\end{enumerate}

%--------------------------------------------------------------------------
% STEP 5
%--------------------------------------------------------------------------
\item[Step 5] {\em React the thermodynamic variables over the second $\Delta t / 2$ interval.}

Call {\bf React State}$[ (\rho X_k)^{(2),\star}, (\rho h)^{(2),\star}, p_0^{n+1,\star}] 
\rightarrow 
[ (\rho X_k)^{n+1,\star}, (\rho h)^{n+1,\star}, (\rho \omegadot_k)^{n+1,\star}, (\rho \Hnuc)^{n+1,\star} ].$

%--------------------------------------------------------------------------
% STEP 6
%--------------------------------------------------------------------------
\item[Step 6] {\em Compute the time-centered expansion term, $S^{\nph,\star}$.}

First, compute $S^{n+1,\star}$ with
\begin{equation}
S^{n+1,\star} =  \left(-\sigma  \sum_k  \xi_k  \omegadot_k  + \frac{1}{\rho p_\rho} \sum_k p_{X_k}  {\omegadot}_k + \sigma \Hnuc\right)^{n+1,\star}.
\end{equation}
  Then, define
\begin{equation}
 S^{\nph} = \frac{S^n + S^{n+1,\star}}{2},
\end{equation}

%--------------------------------------------------------------------------
% STEP 7
%--------------------------------------------------------------------------
\item[Step 7] {\em Construct a face-centered, time-centered advective velocity, $\uadvtwo$.}

The procedure to construct $\uadvtwodag$ is identical to the Godunov procedure
for computing $\uadvonedag$ in {\bf Step 3}, but uses
the updated value $S^{\nph}$ rather than $S^{\nph,\star}$.
The $\dagger$ superscript refers to the fact that the predicted velocity field does not satisfy the divergence constraint,
\begin{equation}
\nabla \cdot \left(\beta_0^{\nph} \uadvtwo\right) =
\beta_0^{\nph} \left[S^{\nph} - \frac{1}{\gammaonebar^{\nph}p_0^{\nph}}\left(\frac{\partial p_0}{\partial t}\right)^{\nph}\right],\label{eq:div2}
\end{equation}
with
\begin{equation}
\beta_0^{\nph} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2}, \quad
\gammaonebar^{\nph} = \frac{ \gammaonebar^n +  \gammaonebar^{n+1,\star} }{2}.
\qquad
\end{equation}
we project $\uadvtwodag$ onto the space of velocities that satisfy the constraint to obtain $\uadvtwo$ using a MAC projection (see Appendix \ref{Sec:Projection}).

%--------------------------------------------------------------------------
% STEP 8
%--------------------------------------------------------------------------
\item[Step 8] {\em Advect the thermodynamic variables over a time interval of $\dt.$}

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \Alph{enumi}}}

\item Update $(\rho X_k)$.  This step is identical to {\bf Step 4A} except we use
  the updated values $\uadvtwo$ and $\rho_0^{n+1,\star}$ rather than
  $\uadvone$ and $\rho_0^n$.  In particular:

\begin{enumerate}
\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}

\item Compute the face-centered, time-centered species, $(\rho X_k)^{\nph,\pred}$,
  for the conservative update of $(\rho X_k)^{(1)}$ using a Godunov approach \citep{XRB_III}.
  Again, we predict $\rho'^n=\rho^n-\rho_0^n$ and $X_k^n$ to faces
  but here we spatially interpolate $(\rho_0^n+\rho_0^{n+1,*})/2$ to faces to assemble the fluxes.

\item Evolve $(\rho X_k)^{(1)} \rightarrow (\rho X_k)^{(2)}$ using
\begin{equation}
(\rho X_k)^{(2)} = (\rho X_k)^{(1)}
- \dt \left\{ \nabla \cdot \left[\uadvtwo (\rho X_k)^{\nph,\pred} \right] \right\},
\end{equation}

\end{enumerate}

\item Update $\rho_0$ by calling {\bf Average}$[\rho^{(2)}]\rightarrow[\rho_0^{n+1}]$.

\item Update $p_0$ by calling {\bf Enforce HSE}$[\rho_0^{n+1}] \rightarrow [p_0^{n+1}]$.

\item Update the enthalpy.  This step is identical to {\bf Step 4D} except we use
  the updated values $\uadvtwo$, $\rho_0^{n+1}$, $(\rho h)_0^{n+1}$, and $p_0^{n+1}$
  rather than
  $\uadvone, \rho_0^n$, $(\rho h)_0^n$, and $p_0^n$.
  In particular:

\begin{enumerate}
\renewcommand{\labelenumii}{{\bf \roman{enumii}}.}

\item Compute the face-centered, time-centered enthalpy, $(\rho h)^{\nph,\pred},$
  for the conservative update of $(\rho h)^{(1)}$ using a Godunov approach \citep{XRB_III}.
  Again, we predict $(\rho h)'^n=(\rho h)^n-(\rho h)_0^n$ to faces
  but here we spatially interpolate $[(\rho h)_0^n)+(\rho h)_0^{n+1,*}]/2$ to faces to assemble the fluxes.

\item Evolve $(\rho h)^{(1)} \rightarrow (\rho h)^{(2)}$.
\begin{equation}
(\rho h)^{(2)}
= (\rho h)^{(1)} - \dt \left\{ \nabla \cdot \left[ \uadvtwo (\rho h)^{\nph,\pred} \right] \right\} + \Delta t\frac{Dp_0}{Dt}
\end{equation}
where here
\begin{equation}
\frac{Dp_0}{Dt} =
\begin{cases}
\frac{p_0^{n+1} - p_0^n}{\Delta t} + \left(\uadvtwo \cdot \eb_r\right) \left(\frac{\partial p_0}{\partial r} \right)^{n}& {\rm (planar)} \\
\frac{p_0^{n+1} - p_0^n}{\Delta t} + \left[ \nabla \cdot \left (\uadvtwo p_0^{\nph} \right ) - p_0^{\nph} \nabla \cdot \uadvtwo \right]& {\rm (spherical)}
\end{cases}
,
\end{equation}
and $p_0^\nph = (p_0^n+p_0^{n+1})/2$.
\end{enumerate}
\end{enumerate}

%--------------------------------------------------------------------------
% STEP 9
%--------------------------------------------------------------------------
\item[Step 9] {\em React the thermodynamic variables over the second $\Delta t / 2$ interval.}

Call {\bf React State}$[(\rho X_k)^{(2)},(\rho h)^{(2)}, p_0^{n+1}] \rightarrow [(\rho X_k)^{n+1}, (\rho h)^{n+1}, (\rho \omegadot_k)^{n+1}, (\rho \Hnuc)^{n+1}].$

%--------------------------------------------------------------------------
% STEP 10
%--------------------------------------------------------------------------
\item[Step 10] {\em Define the new-time expansion term, $S^{n+1}$.}

\begin{enumerate}
\renewcommand{\theenumi}{{\bf \Alph{enumi}}}
\item Define
\begin{equation}
  S^{n+1} =  \left(-\sigma  \sum_k  \xi_k \omegadot_k  + \sigma \Hnuc +
  \frac{1}{\rho p_\rho} \sum_k p_{X_k}  \omegadot_k\right)^{n+1}.
\end{equation}

\end{enumerate}

%--------------------------------------------------------------------------
% STEP 11
%--------------------------------------------------------------------------
\item[Step 11] {\em Update the velocity}.

First, we compute the face-centered, time-centered velocities, $\Ub^{\nph,\pred}$
using a Godunov approach \citep{XRB_III}. Then, we update
the velocity field $\Ub^n$ to $\Ub^{n+1,\dagger}$ by discretizing
equation (\ref{eq:momentum}) as
\begin{equation}
\Ub^{n+1,\dagger}
= \Ub^n - \dt \left[\uadvtwo \cdot \nabla \Ub^{\nph,\pred} \right]
 - \dt \left[ \frac{\beta_0^\nph}{\rho^\nph} \nabla \left( \frac{\pi^\nmh}{\beta_0^\nmh} \right) + \frac{\left(\rho^\nph-\rho_0^\nph\right)}{\rho^\nph} g^{\nph} \eb_r \right],
\end{equation}
where
\begin{equation}
\rho^\nph = \frac{\rho^n + \rho^{n+1}}{2}, \qquad \rho_0^\nph = \frac{\rho_0^n + \rho_0^{n+1}}{2}.
\end{equation}
Again, the $\dagger$ superscript refers
to the fact that the updated velocity does not satisfy the divergence constraint,
\begin{equation}
\nabla \cdot \left(\beta_0^{n+1} \Ub^{n+1} \right) = \beta_0^{n+1} \left[ S^{n+1} - \frac{1}{\gammaonebar^{n+1}p_0^{n+1}}\left(\frac{\partial p_0}{\partial t}\right)^{\nph}\right].\label{eq:div3}
\end{equation}
We use an approximate projection to project $\Ub^{n+1,\dagger}$ onto the space of velocities that satisfy the constraint to obtain $\Ub^{n+1}$ using a ``nodal'' projection.
This projection necessarily differs from the MAC projection used in
{\bf Step 3} and {\bf Step 7} because the velocities in those steps are defined
on faces and $\Ub^{n+1}$ is defined at cell centers, requiring different divergence
and gradient operators.
Furthermore, as part of the nodal projection, we also define a nodal new-time perturbational pressure, $\pi^\nph$.
Refer to Appendix \ref{Sec:Projection} for more details.

\end{description}
This completes one step of the algorithm.

To initialize the simulation we use the same procedure described in Paper III.
At the beginning of each simulation, we define $(\Ub,\rho X_k,\rho h)$.
We set initial values for $\Ub, \rho X_k$, and $\rho h$ and perform a sequence of projections 
(to ensure the velocity field satisfies the divergence constraint) 
followed by a small number of steps of the temporal advancement scheme to iteratively 
find initial values for $\pi^{n-\myhalf}$ and $S^0$ and $S^1$ for use in the first time step.

Our approach to adaptive mesh refinement is algorithmically the same as the treatment described 
in Section 5 of Paper V; we refer the reader there for details.
MAESTROeX supports refinement ratios of 2 between levels.
We note that for spherical problems, AMR is only available for the case of a uniformly-spaced base state.

\section{Performance and Validation}\label{sec:results}

\subsection{Performance and Scaling}\label{sec:scaling}
We perform weak scaling tests for simulations of convection preceding ignition in a spherical, full-star Chandrasekhar mass white dwarf. 
The simulation setup remains the same as reported in Section 3 of \cite{MAESTRO_AMR} and originally used in \cite{MAESTRO_convection}, and thus we emphasize that these scaling tests are performed using meaningful, scientific calculations.
Here, we perform simulations using $256^3, 512^3, 768^3, 1024^3, 1280^3$, and $1536^3$ grid cells on a spatially uniform grid (no AMR).
We divide each simulation into $64^3$ grids, so these simulations contain between 64 grids ($256^3$) and 13,824 grids ($1536^3$).
These simulations were performed using the NERSC cori system on the Intel Xeon Phi (KNL) partition.
Each node contains 68 cores, each capable of supporting up to 4 hardware threads (i.e., a maximum of 272 hardware threads per node).
For these tests, we assign 4 MPI tasks to each node, and 16 OpenMP threads per MPI process.
Each MPI task is assigned to a single grid, so our tests use between 64 and 13,824 MPI processes (i.e., between 1,024 and 221,184 total OpenMP threads).
For $64^3$ grids we discovered that using more than 16 OpenMP threads did not decrease the wallclock time due to a lack of work available per grid; in principle one could use larger grids, fewer MPI processes, and more threads per MPI process to obtain a flatter weak scaling curve, however the overall wallclock time would increase except for extremely large numbers of MPI processes (beyond the range we tested here).
Thus, the more accurate measure of weak scaling is to consider the number of MPI processes, since the scaling plot would look virtually identical for larger thread counts.
Note that the largest simulation used roughly 36\% of the entire computational system.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htb]
\begin{center}
\includegraphics[width=3.0in]{MAESTRO_scaling1.eps} \hspace{0.5em}
\includegraphics[width=3.0in]{MAESTRO_scaling2.eps}
\caption{\label{fig:scaling} (Left) Weak scaling results for a spherical, full-star white dwarf calculation using the original MAESTRO code, MAESTROeX, and MAESTROeX with base state evolution disabled.  Shown is the average wallclock time per time step.
(Right) Weak scaling results showing the average wallclock time per time step spent in the cell-centered and nodal linear solvers within a full time step of the aforementioned simulations.}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In the left panel of Figure \ref{fig:scaling} we compare the wallclock time per time step as a function of total core count (in this case, the total number of OpenMP threads) for the original FBoxLib-based MAESTRO implementation to the AMReX MAESTROeX implementation.
These tests were performed using the original temporal integration strategy in \cite{MAESTRO_V}, noting that the new temporal integration with and without the irregular base state gives essentially the same results.
We also include a plot of MAESTROeX without base state evolution.
Comparing the original and new implementations, we see similar scaling results except for the largest simulation, where MAESTROeX performs better.
We see that the increase in wallclock time from the smallest to largest simulation is roughly 42\%.
We also note that without base state evolution, the code runs 14\% faster for small problems, and scales much better with wallclock time from the smallest to largest simulation increasing by only 13\%.
This is quite remarkable since there are 3 linear solves per time step (2 cell-centered Poisson solves used in the MAC projection, and a nodal Poisson solve used to compute the updated cell-centered velocities).
Contrary to our prior assumptions, the linear solves are not the primary scaling bottleneck in this code.
In the right panel of Figure \ref{fig:scaling}, we isolate the wallclock time required for these linear solves and see that (i) the linear solves only use 20-23\% of the total computational time, and (ii) the increase in the solver wallclock time from the smallest to largest simulation is only 28\%.
Further profiling reveals that the primary scaling bottleneck is the average operator.
The averaging operator requires collecting the sum of Cartesian data onto one-dimensional arrays holding every possible mapping radius.
This amounts to at least 24,384 double precision values (for the $256^3$ simulation) up to 883,584 values (for the $1536^3$ simulation).
The averaging operator requires a global sum reduction over all processors, and the communication of this data is the primary scaling bottleneck.
For the simulation with base state evolution, this averaging operator is only called once per time step (as opposed to 14 times per time step when base state evolution is included).
The difference in total wallclock times with and without base state evolution is almost entirely due to the averaging.
Note that as expected, advection, reactions, and calls to the equation of state scale almost perfectly, since there is only a single parallel communication call to fill ghost cells.

\subsection{White Dwarf Convection}\label{sec:whitedwarf}

To explore the accuracy of the new temporal algorithm, we now analyze in detail three-dimensional, full-star calculations of convection preceding ignition in a white dwarf. Again, we refer the reader to \cite{MAESTRO_AMR} and \cite{MAESTRO_convection} for setup details. We implement both uniformly- and irregularly-spaced base state with the new temporal algorithm, while only uniform base state spacing is used in the original algorithm. As in Section 3 of \cite{MAESTRO_AMR}, we choose the peak temperature and peak Mach number as the two diagnostics to compare the simulations. Figure \ref{fig:wdconvect_256_maxvar} shows the evolution of both peak temperature and peak Mach number until time of ignition on a single-level grid with resolution of $256^3$. The simulation using the new temporal scheme with uniformly-spaced base state gives the same qualitative results as the original scheme, and predicts a similar time of ignition ($t=7810$ s compared to $t=7850$ s for original algorithm). The simulation using the new temporal scheme with irregularly-spaced base state displays a slightly different peak temperature behavior during the initial transition period $t<150$ s, which results in the difference between the curves post transition. We strongly suspect that this is a result of using a different initial model file (the resolution near the center of the star is much coarser with the irregular spacing than the uniform spacing). Fortunately, the simulation with irregular base state spacing still follows the same trend as with uniform spacing, and the star is shown to ignite at an earlier time $t=6840$ s.

Figure \ref{fig:wdconvect_Tmax} shows the peak temperature evolution over the first 1000 s on two grids of differing resolutions, $256^3$ and $512^3$.
Limited allocations prevented us from running this simulation further.
As previously suspected, the simulation using irregularly-spaced base state agrees much closer with the results from using uniform spacing as the resolution increases.
This is most likely due to the increased resolution of the initial model, which more closely matches the uniformly-spaced counterpart.
This is especially important when computing the base state pressure from base state density, which is particularly sensitive to coarse resolution near the center of the star.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htb]
\begin{center}
\includegraphics[width=2.75in]{wdconvect_256_maxT.eps}  \hspace{0.5in}
\includegraphics[width=2.75in]{wdconvect_256_maxMach.eps}
\caption{\label{fig:wdconvect_256_maxvar} (left) Peak temperature, $T_{\text{peak}}$, and (right) peak Mach number
         in a white dwarf until time of ignition at resolution of $256^3$ for three different MAESTROeX algorithms.}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[hbt]
\begin{center}
\includegraphics[width=3.25in]{wdconvect_compare_Tmax.eps}
\caption{\label{fig:wdconvect_Tmax} Peak temperature, $T_{\text{peak}}$  in a white dwarf at grid resolutions of
         $256^3$ (dotted line) and $512^3$ (solid line) until $t=1000$ for uniform (d$x$) and irregular (d$r$) base state spacing.
         Note that the irregularly-spaced solution agrees better
         with the uniformly-spaced solution as the resolution increases.}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In terms of efficiency, all three simulations on $256^3$ single-level grid were run on Cori haswell with 64 processors and 8 threads per core and their run times were compared. As a result of simplifying the algorithm by eliminating the evolution equations for the base state density and pressure, the simulation using the new temporal algorithm took only 6.75 s per time step with uniformly-spaced base state, which is 13\% faster than the 7.77 s per time step when using the original scheme. However, we do observe a 25\% increase in run time of 9.72 s when using irregularly-spaced base state with the new algorithm. This can be explained by the irregularly-spaced base state array being much larger in size than its uniformly-spaced counterpart, and thus require additional communication and computation time. One possible strategy to significantly reduce the run time is to consider truncating the base state beyond the cutoff density radius.

\subsection{AMR Performance}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htb]
\begin{center}
\includegraphics[width=2.5in]{reacting_bubble_amr.eps} \hspace{2.5em}
\includegraphics[width=2.5in]{wdconvect_amr_3grid.eps}
\caption{\label{fig:amr_grids} Initial grid structures with two levels of refinement.
         The red, green, and black lines represent grids of increasing refinement.
         (Left) Profile of $T - \bar{T}$ for a hot bubble in a white dwarf environment.
         (Right) Region of star where $\rho\ge 10^5 \text{ g cm}^{-3}$ in a full white dwarf star simulation. }
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We now test the performance of MAESTROeX for adaptive, three-dimensional simulations to track localized regions of interest over time. Figure \ref{fig:amr_grids} illustrates the initial grid structures with two levels of refinement for both planar and spherical geometries where the grid is refined according to the temperature and density profiles, respectively. For each of the problems we tested, the single-level simulation was run using the original temporal scheme and the adaptive simulations using the original and new temporal algorithms. We want to show that the adaptive simulation can give similar results to the single-level simulation and in a more computationally efficient manner.

In the planar case, we use the same problem setup for a hot bubble rising in a white-dwarf environment as described in Section 6 of Paper V. Here we use a domain size of $3.6\times 10^7$ cm by $3.6\times 10^7$ cm by $2.88\times 10^8$ cm, and allow the grid structure to change with time. The single-level simulation at a resolution of $128^2 \times 1024$ was run on Cori haswell with 48 processors and took approximately 33.5 s per time step (averaged over 10 time steps) using either the original or new temporal algorithm. The adaptive simulation has a resolution of $32^2 \times 256$ at the coarsest level, resulting in the same effective resolution at the finest level as the single-level simulation. We tag cells that satisfy $T-\bar{T} > 3\times 10^7$ K as well as all cells at that height. The adaptive run took only 3.7 s per time step, and this 89\% decrease in runtime is mostly due to the fact that initially only 6.25\% of the cells (1,048,576 out of $128^2 \times 1024$ cells) are refined at the finest level. Figure \ref{fig:bubble_results} shows a series of planar slices of the temperature profile at time intervals of 1.25 s, and verifies that the adaptive simulation is able to capture the same dynamics as the single-level simulation at much lower computational cost.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htb]
\begin{center}
\includegraphics[width=2.5in]{reacting_bubble_result.eps} \hspace{2.5em}
\includegraphics[width=2.5in]{reacting_bubble_amr_result.eps}
\caption{\label{fig:bubble_results} Time-lapse cross-section of a hot bubble in a white dwarf environment at
         $t = 0$, 1.25, and 2.5 s for (left) single-level simulation, and
         (right) adaptive simulation at the same effective resolution.  The red, green, and black boxes indicate grids of increasing resolution}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We continue to use the full-star white dwarf problem described in Section \ref{sec:whitedwarf} to test adaptive simulations on spherical geometry. The adaptive grid is refined twice by tagging the density at $\rho > 10^5$ g cm$^{-3}$ on the first level and $\rho > 10^8$ g cm$^{-3}$ on the second level. These tagging values have been shown to work well previously in Paper V, but we have found that the code may encounter numerical difficulties when the tagging values are too close to each other in subsequent levels of refinement.
The simulation on a single-level grid of $512^3$ resolution took 12.7 s per time step (again averaged over 10 time steps).  The adaptive grid has a resolution of $128^3$ at the coarsest level and an effective resolution of $512^3$ at the finest level. On this grid, 27.8\% of the cells (4,666,880 out of $256^3$ cells) are refined at the first level and 5.3\% (7,176,192 out of $512^3$ cells) at the second. The adaptive simulation took 5.61 s, resulting in more than a factor of 2 in speedup. Both simulations are computed to $t=2000$ s and we choose to use the peak temperature as the diagnostic to compare the results. Figure \ref{fig:wdconvect_amr_Tmax} shows the evolution of the peak temperature for all three runs and shows that the adaptive simulation gives the same qualitative result as the single-level simulation. We do not expect the curves to match up exactly because the governing equations are highly nonlinear, and slight differences in the solution caused by solver tolerance and discretization error can change the details of the results. Each simulation was run on Cori haswell with 512 processors and 4 threads per core.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htb]
\begin{center}
\includegraphics[width=4.0in]{wdconvect_amr_Tmax.eps}
\caption{\label{fig:wdconvect_amr_Tmax} Peak temperature, $T_{\text{peak}}$, in a white dwarf from $t=0$ to $2000$ s
         for grids with effective resolution of $512^3$. We can see that the adaptive grids with two levels of
         refinement give very similar solution trends compared to the single-level grid.}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions and Future Work}\label{sec:conclusions}
We have developed a new temporal integrator and spatial mapping options into our low Mach number solver, MAESTROeX.
The new AMReX-enabled code scales well on large fractions of supercomputers with multicore architectures.
Future software enhancements will include GPU implementation.
In particular, the AMReX-based companion code, the compressible CASTRO code \citep{CASTRO}, has recently ported hydrodynamics and reactions to GPUs \citep{CASTRO_GPU}.
We plan to leverage the newly implemented mechanisms for offloading compute kernels to GPUs inside of the AMReX software library itself.
Our future scientific investigations include convection in massive rotating stars, \citep{heger2000presupernova}, the convective Urca process in white dwarfs \citep{willcox2016type}, solar physics \citep{wood2018self}, and magnetohydrodynamics \citep{wood2015three,wood2011sun}.
Our future algorithmic enhancements include more accurate and higher-order multiphysics coupling strategies based on spectral deferred corrections \citep{dutt2000spectral,bourlioux2003high}.
This framework has been successfully used in terrestrial combustion \citep{pazner2016high,nonaka2018conservative}

\acknowledgements

The work at LBNL was supported by the U.S. Department of Energy's
Scientific Discovery Through Advanced Computing (SciDAC) program under
contract No. DE-AC02-05CH11231.  The work at Stony Brook was supported
by DOE/Office of Nuclear Physics grant DE-FG02-87ER40317 and through
the SciDAC program DOE grant DE-SC0017955.  This research used
resources of the National Energy Research Scientific Computing Center
(NERSC), a U.S. Department of Energy Office of Science User Facility
operated under Contract No. DE-AC02-05CH11231.

\software{AMReX \citep{AMReX, AMReX_JOSS}, StarKiller Microphysics \citep{starkiller}}
\facilities{NERSC}

\appendix
\section{Projection Details}\label{Sec:Projection}
To enforce the divergence constraint in {\bf Step 3, 7}, and {\bf 11}, we use a projection method analogous to the methods originally developed for incompressible flow \citep{almgren1998conservative,bell1989second}.
The basic idea is to decompose the velocity field into a part that satisfies the divergence-satisfying component and a curl-free (gradient of a scalar field) component by solving a variable-coefficient Poisson equation for the scalar field.
The details for the MAC projection in {\bf Step 3} and {\bf Step 7} are given in Appendix B of Paper III.  The details of the nodal projection in {\bf Step 11} are given in Section 3.2 of Paper III.
We note that in the nodal projection, the gradient of the scalar field is used to update the perturbational pressure, $\pi$.

Based on our past experience in the MAESTRO project, we have found it useful to split the velocity dynamics into a perturbational and base state velocity,
\begin{equation}
\Ub = \Ubt(\xb,t) + w_0(r,t)\eb_r,
\end{equation}
solve for each term separately, and immediately combine them to find a full velocity that satisfies the constraint.  We take that approach here, primarily because it allows us to enforce a boundary condition on $w_0$ at the edge of the star (i.e., the cutoff density location where we hold density constant).  Namely, to enforce that $r^2 w_0$ remain constant is difficult to do when solving for the full velocity. 
This is demonstrated in Figure \ref{fig:wdconvect_splitU} (right) where the velocity magnitude is observed to incorrectly increase outside the cutoff density radius when we solve for the full velocity in the nodal projection.
The resulting peak temperature also dips significantly as seen in Figure \ref{fig:wdconvect_splitU} (left), presumably because the dynamics of the overall expansion of the star are not being captured correctly.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htb]
\begin{center}
\begin{tabular}{l c}
\multirow{4}{3.25in}[30mm]{ \includegraphics[width=3.0in]{wdconvect_256_splitU.eps} } & \multicolumn{1}{c}{\includegraphics[width=2.35in]{magvel_full_XY.eps}} \\
& \multicolumn{1}{c}{\begin{footnotesize} (a) Velocity magnitude, solved using full $\mathbf{U}$ \end{footnotesize}} \\[1.em]
& \multicolumn{1}{c}{\includegraphics[width=2.75in]{magvel_lineout_X.eps}} \\ 
& \multicolumn{1}{c}{\begin{footnotesize} (b) Comparison along white line \end{footnotesize}} \\
\end{tabular}
\caption{\label{fig:wdconvect_splitU} When solving for the full velocity $\mathbf{U}$ in the nodal projection step instead of the split velocity 
  formulation, we illustrate
  (left) the peak temperature, $T_{\text{peak}}$ in a white dwarf at resolution of $256^3$ during the early evolution of the star and
  (right) the magnitude of velocity at early time. We observe much larger velocities outside the density cutoff region, $\rho < 10^5 \text{ g cm}^{-3}$, 
  that are not seen when we split the velocity dynamics.}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In practice, we solve the constraint over the lateral average,
\begin{equation}
\nabla\cdot\left(\beta_0\Ubt\right) = \beta_0(S-\overline{S})
\end{equation}
and separately solve for $w_0$ using 
\begin{equation}
\nabla\cdot(\beta_0w_0\eb_r) = \beta_0\left(\overline{S} - \frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial t}\right).
\end{equation}
To solve for $\Ubt$ we use a projection method, which involves the solution of a variable-coefficient Poisson solver to extract the curl-free component of the unprojected velocity, leaving a velocity field that satisfies the divergence constraint.
Note that MAESTRO contains alternate low Mach number formulations that conserve total energy in stratified systems, with minimal changes to the code (see Appendix A of \cite{subChandra_II} for details).
To find $w_0$ we integrate in one dimension using the procedure in Appendix B of Paper V, keeping in mind that the base state spacing ($\Delta r$) should be computed using the appropriate cell-edge and cell-center locations when using irregularly-spaced base state.
Note that in this approach, we estimated the time-derivative of the pressure in part by examining how laterally averaged $\rho'=\rho-\rho_0$ changed over time (quantified by $\eta_\rho = \overline{\rho'\Ub\cdot\eb_r}$).  After evolving the species ({\bf Steps 4A/8A}), we compute this term as reported in Paper V. For example, after {\bf Step 8A}, we define a radial cell-centered $\etarho^{\nph}$,

\begin{description}
\item For planar geometry, $\etarho = \overline{\rho'(\Ub\cdot\eb_r)}$,
\begin{equation}
 \etarho^{\nph} =  {\rm {\bf Average}} \sum_k \left[ \left(\uadvtwo \cdot \eb_r \right) (\rho X_k)^{\nph,\pred} \right]
\end{equation}
\item For spherical geometry, first construct 
$\etarho^{{\rm cart},\nph} = [\rho'(\Ub\cdot\eb_r)]^{\nph}$ on Cartesian cell centers using:
\begin{equation}
\etarho^{{\rm cart},\nph} = \left[\left(\frac{\rho^{(1)}+\rho^{(2)}}{2}\right)-\left(\frac{\rho_0^n+\rho_0^{n+1}}{2}\right)\right] \cdot \left( \uadvtwo \cdot \eb_r \right).
\end{equation}
Then,
\begin{equation}
\etarho^{\nph} = {\rm {\bf Average}}\left(\etarho^{{\rm cart},\nph}\right).
\end{equation}
\end{description}

\bibliographystyle{aasjournal.bst}
\bibliography{references}

\end{document}

### 2019-08-13 xrb/paper.tex
switch to aastex 6.3
update ORCiDs
update the end discussion

@@
 %\documentclass[12pt, preprint]{aastex}
-\documentclass[preprint,times,tighten]{aastex62}
+\documentclass[preprint,times,tighten]{aastex63}
 
@@
 
-\author{Donald Willcox}
+\author[0000-0003-2300-5165]{Donald Willcox}
 \affiliation{Lawrence Berkeley National Laboratory, Berkeley, CA}
@@
 
-\author{Max P.\ Katz}
-\affiliation{Nvidia Corp}
+\author[0000-0003-0439-4556]{Max P.\ Katz}
+\affiliation{NVIDIA Corp}
 
@@
 
-\author{Weiqun Zhang}
+\author[0000-0001-8092-1974]{Weiqun Zhang}
 \affiliation{Lawrence Berkeley National Laboratory, Berkeley, CA}
@@
 
-On the algorithmic front, we are exploring a change of our
-hydrodynamics and reactive coupling to eliminate splitting error from
-the traditional Strang splitting used in \castro\ and other
-hydrodynamics codes.  One advantage of this is that the new algorithm
-would be more amenable to offloading onto GPUs, a development task
-that is underway.  Additionally, we will move to fourth-order accurate
-methods and incorporate well-balanced reconstruction to better capture
-the burning.
+There are several changes we will pursue to the algorithm used to
+model these XRBs.  First, as shown in \citet{castro_sdc}, we have
+developed a fourth-order (in space and time) method for coupling
+hydrodynamics and reactions that should greatly improve the accuracy
+of the simulations.  We expect that by using this new high-order
+algorithm we can drop a level of refinement from the simulations while
+still accurately modeling the evolution.  We have also ported
+\castro\ to GPUs, giving an order of magnitude performance boost on
+nodes with both CPUs and GPUs.  The GPU-enabled algorithm will
+increase the performance, allowing us to run bigger domains and/or
+eliminate the boosting used here.
 
-MHD, radiation transport, GPUs
+MHD, radiation transport
 

### 2019-08-16 xrb/paper.tex
add the time series plot
add a table of simulations:

@@
   Rossby length} using a scale height $H_0 = 10^3~\mathrm{cm}$.  This
-is less than half our domain width.
+is less than half our domain width.  Table~\ref{table:sim_names} lists
+the simulations run and their parameters.
 
@@
 freely available in the \castro\ github repository.
+
+\begin{deluxetable}{lccccc}
+\tablecaption{\label{table:sim_names} Simulaiton parameters.}
+\tablehead{\colhead{name} & \colhead{reaction boost} & \colhead{conductivity boost} & \colhead{fine grid resolution} & \colhead{rotation rate}  & \colhead{network}}
+\startdata
+10/10 & 10 & 10 & 10~cm & 2000~Hz & {\tt aprox13} \\
+10/10 slow & 10 & 10 & 10~cm & 1000~Hz & {\tt aprox13} \\
+10/10 lores & 10 & 10 & 20~cm & 2000~Hz & {\tt aprox13} \\
+5/5 & 5 & 5 & 10~cm & 2000~Hz & {\tt aprox13} \\
+10/10 iso7 & 10 & 10 & 10~cm & 2000~Hz & {\tt iso7} \\
+\enddata
+\end{deluxetable}
 
@@
 
+Figure~\ref{fig:time_series} shows the time
+
 Here we'll show the large scale dynamics, colormap plots, etc. and discuss the trends.
+
+\begin{figure}[h]
+\centering
+\epsscale{0.95}
+\plotone{time_series}
+\caption{\label{fig:time_series} Time series of the mean molecular weight of the 
+flame for our standard 10/10 flame.}
+\end{figure}
 

### 2019-08-17 xrb/paper.tex
more work on the paper

--- 2019-08-13 xrb/paper.tex
@@
 freely available in the \castro\ github repository.
+
+\begin{deluxetable}{lccccc}
+\tablecaption{\label{table:sim_names} Simulaiton parameters.}
+\tablehead{\colhead{name} & \colhead{reaction boost} & \colhead{conductivity boost} & \colhead{fine grid resolution} & \colhead{rotation rate}  & \colhead{network}}
+\startdata
+10/10 & 10 & 10 & 10~cm & 2000~Hz & {\tt aprox13} \\
+10/10-slow & 10 & 10 & 10~cm & 1000~Hz & {\tt aprox13} \\
+10/10-lores & 10 & 10 & 20~cm & 2000~Hz & {\tt aprox13} \\
+5/5 & 5 & 5 & 10~cm & 2000~Hz & {\tt aprox13} \\
+10/10-iso7 & 10 & 10 & 10~cm & 2000~Hz & {\tt iso7} \\
+\enddata
+\end{deluxetable}
 
@@
 
+Figure~\ref{fig:time_series} shows the time
+
 Here we'll show the large scale dynamics, colormap plots, etc. and discuss the trends.
+
+% made with time_series.py in flame_wave_boost_10_10_paper with --skip 100
+\begin{figure}[h]
+\centering
+\plotone{time_series}
+\caption{\label{fig:time_series} Time series of the mean molecular weight of the 
+flame for our standard 10/10 simulation.}
+\end{figure}
+
+% made with overview.py
+\begin{figure}[h]
+\centering
+\plotone{flame_boost_10_10_overview}
+\caption{\label{fig:10_10_overview} Temperature, energy generation rate, and out-of-plane velocity for the 10/10 simulation.}
+\end{figure}
 
@@
 this choice.
+
+\begin{figure}[h]
+\centering
+\plotone{flame_wave_boost_5_5_slice}
+\caption{\label{fig:5_5_overview} Temperature, mean molecular weight, energy generation rate, and out-of-plane velocity for the 5/5 simulation.}
+\end{figure}
 

### 2019-08-19 xrb/plots/initial_model/initial_model.py
update initial model stuff

@@ -14,8 +14,8 @@ def initial_model():
     r_min = 0
     r_max = 10000.0
 
-    r_min = 4500
-    r_max = 7000
+    r_min = 1500
+    r_max = 4500
 
     #-------------------------------------------------------------------------
     # density and temperature plots

### 2019-08-29 astronum_2019_MAESTROeX/astronum-maestroex.tex
first draft of first few paragraphs of maestroex astronum paper

@@
 \newcommand{\castro}{{\sffamily Castro}}
+\newcommand{\nyx}{{\sffamily Nyx}}
 \newcommand{\amrex}{{\sffamily AMReX}}
+\newcommand{\amrexastro}{{\sffamily AMReX-Astro}}
+\newcommand{\fboxlib}{{\sffamily FBoxLib}}
 
+\newcommand{\Ub}{{\,\bm{U}}}
 \newcommand{\Uc}{{\,\bm{\mathcal{U}}}}
@@
 \newcommand{\Advs}[1]{\boldsymbol{\mathcal{A}} \left(#1\right)}
+\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
+\newcommand{\md}[2]{\frac{\mathrm{D} #1}{\mathrm{D} #2}}
 
@@
         D.~E. Willcox$^2$, and
-        M. Zingale$^1$,}
+        M. Zingale$^1$}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{abstract}
-abstract abstract abstract
+Convective flows in the interiors of stars are challenging to model using conventional compressible hydrodynamics codes due to their low Mach number, which severely limits the timestep that can be taken in simulations.
+\maestroex\ is an open source low Mach number stellar hydrodynamics code that allows much larger timesteps to be taken, therefore enabling systems to be modelled for much longer periods of time.
+This is particularly important for the problem of convection in the cores of rotating massive stars prior to core collapse.
+To fully capture the dynamics, it is necessary to model these systems in 3d at high resolution over many rotational periods, but this has been too computationally expensive for existing, conventional schemes.
+
+We shall present an overview of \maestroex's current capabilities, describe ongoing work to incorporate the effects of rotation and discuss how we are optimizing the code to run on GPUs. 
 \end{abstract}
 
-\section{Introduction}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-intro intro intro
+\section{Introduction} \label{sec:intro}
 
-\section{Low Mach Number hydrodynamics}
-slow fluids in space are hard to model
+For many flows in astrophysical systems, the magnitude of the fluid velocity is much less than the soundspeed.
+Consequently, the Mach number, $\textrm{Ma} = |\Ub| / c_s \ll 1$.
+Such low Mach number flows are challenging to model using standard compressible schemes, where the maximum size of the timestep is determined by the CFL condition: for a grid with cell spacing $\Delta x$, the timestep $\Delta t <  \Delta x / \max(|\Ub|+c_s)$.
+In low Mach number flows, this condition is dominated by the contributions of the sound speed, with the result that many fine timesteps are required to achieve high spatial resolution.
 
-\section{MAESTROeX}
-\maestroex\ is our awesome new code for modelling slow fluids in space
+This restriction can be lifted by using sound-proof methods.
+These methods involve modifying the fluid equations, modifying the computational algorithm and/or modifying the flow parameters in order to allow much larger timesteps to be used.
+In \maestroex, we use the \emph{low Mach number approximation}.
+An overview of this shall be given in \sref{sec:low_mach_hydro}, with further details of the \maestroex~code and our development workflow given in \sref{sec:maestroex} and \sref{sec:workflow}.
 
-\section{GPUs}
+An example of an astrophysical low Mach number flow is convection within the interiors of massive stars.
+Prior to core collapse, the interiors of massive stars consist of a series of convective burning shells, separated by inert, non-convective shells.
+Moving from the outermost layers inwards, the shells consist of heavier and heavier elements, with the core burning elements up to Fe.
+An accurate picture of the structure of such stars in the minutes before core collapse is important for supernova modelling.
+The composition and structure of the star prior to collapsing provides the initial data for these models, so this needs to be accurate if the supernova models are to be trusted.
+Modelling this convection is challenging using conventional compressible schemes because the domain size is very large (i.e.
+a significant fraction of the interior of the entire star), high resolution is required in order to properly capture the turbulent mixing that occurs at shell boundaries, and long time periods are required (multiple convective turnover times).
+In \sref{sec:rotation}, we shall describe ongoing work to model this problem in \maestroex, in particular describing how we are implementing rotation.
+
+The latest generation of supercomputers coming on line are relying more and more on GPU architectures in order to increase the performance whilst minimizing the power requirements.
+For HPC codes to best exploit the top machines, it is therefore necessary to port these codes so that they can run on GPUs.
+In \sref{sec:gpus}, we describe our work to port \maestroex~to run on GPUs and demonstrate its performance.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Low Mach number hydrodynamics} \label{sec:low_mach_hydro}
+
+Sound-proof methods for modelling low Mach number flows can take a variety of forms.
+One technique is to use preconditioners in order to reduce the stiffness of the equations, allowing larger timesteps to be used CITE.
+Another technique is to modify the flow parameters, artificially boosting the speed of the flow without changing its behavior so that the system evolves faster and fewer timesteps are required CITE.
+
+The technique that we use in \maestroex~is to modify the fluid equations themselves so as to filter out the soundwaves.
+This is a similar approach to the incompressible CITE and anelastic CITE approximations, however the approximation that we use (variously called the low Mach number approximation  CITE and the generalized pseudo-incompressible approximation CITE) allows background stratification and large pressure perturbations due to heating and changes in composition.
+This is achieved by decomposing the pressure into a 1d hydrostatic base state, $p_0 = p_0(r, t)$, and a dynamic pressure perturbation, $\pi = \pi(r, \bm{x}, t)$, such that $p(r, \bm{x}, t) = p_0 + \pi$.
+In order for the approximation to hold, $|\pi|/p_0 = O(\mathrm{Ma}^2)$.
+The low Mach number fluid equations are given by
+\begin{align}
+    \pd{\left(\rho X_k\right)}{t} &= - \nabla\cdot\left(\rho X_k \Ub \right) + \rho\dot{\omega}_k, \\
+    \pd{\Ub}{t} &= - \Ub\cdot\nabla\Ub - \frac{\beta_0}{\rho}\nabla\left(\frac{\pi}{\beta_0}\right) - \frac{\rho - \rho_0}{\rho}\bm{e}_r,\\
+    \pd{\left(\rho h\right)}{t} &= -\nabla\cdot\left(\rho h \Ub\right) + \md{p_0}{t} + \rho H_{\mathrm{nuc}},
+\end{align}
+where $\rho$, $\Ub$ and $h$ are the mass density, fluid velocity and specific enthalpy, $X_k$ and $\dot{\omega}_k$ are the species mass fraction and production rate of species $k$, and $H_{\mathrm{nuc}}$ is the energy release per time per unit mass.
+From the base state pressure, we can define a 1d base state density, $\rho_0 = \rho_0(r, t)$, which represents the lateral average and is in hydrostatic equilibrium with $p_0$:
+\begin{equation}
+    \nabla p_0 = -\rho_0 g \bm{e}_r,
+\end{equation}
+where $g = g(r,t)$ is the magnitude of the gravitational acceleration, and $\bm{e}_r$ is the radial unit vector.
+The background stratification is captured by introducing a buoyancy-like term, $\beta_0$.
+It is defined as
+\begin{equation}
+    \beta_0(r,t) = \rho_0(0,t) \exp\left(\int_0^r dr'\, \frac{1}{\bar{\Gamma}_1p_0}\pd{p_0}{r'} \right),
+\end{equation}
+where $\bar{\Gamma}_1$ is the lateral average of the first adiabatic exponent, $\Gamma_1 = d(\ln p)/d(\ln \rho)|_s$, and $s$ is the entropy.
+The equations are closed by casting the equation of state as a velocity divergence constraint.
+This is done by taking the Lagrangian derivative of the EoS for pressure as a function of the thermodynamic variables, substituting in the equations of motion for mass and energy, and requiring that the pressure is described by a function of $r$ and $t$ based on the condition of hydrostatic equilibrium.
+Details of this derivation can be found in CITE.
+The constraint is given by
+\begin{equation}
+    \nabla\cdot\left(\beta_0\Ub\right) = \beta_0 \left(S - \frac{1}{\bar{\Gamma}_1 p_0}\pd{p_0}{t} \right).
+\end{equation}
+Here, $S$ is an expansion term which describes local compressibility effects due to changes in composition and heating from reactions:
+\begin{equation}
+    S = -\sigma \sum_k \xi_k\dot{\omega}_k + \frac{1}{\rho p_\rho}\sum_k p_{X_k}\dot{\omega}_k + \sigma H_{\mathrm{nuc}},
+\end{equation}
+where we define 
+\begin{align*}
+    p_{X_k} \equiv \left.\pd{p}{X_k}\right|_{\rho,T,X_{j,j\neq k}},\qquad
+     \xi_k&\equiv \left.\pd{h}{X_k}\right|_{p, T,X_{j,j\neq k}},\qquad
+     p_\rho\equiv \left.\pd{p}{\rho}\right|_{T, X_k},\\
+     \sigma \equiv \frac{p_T}{\rho c_p p_\rho}, \qquad
+     p_T&\equiv \left.\pd{p}{T}\right|_{\rho, X_k} \quad\mathrm{and}\quad 
+     c_p\equiv \left.\pd{h}{T}\right|_{p, X_k}.
+\end{align*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{MAESTROeX} \label{sec:maestroex}
+
+% \maestroex\ is our awesome new code for modelling slow fluids in space
+
+Prior to \maestroex, we developed the low Mach number code \maestro.
+Like \maestroex, \maestro~is a block-structured adaptive mesh refinement (AMR) code for modelling low Mach number astrophysical codes.
+It was developed over a series of papers CITE.
+The system of low Mach number equations are solved using an explicit Godunov approach for the advection, a stiff ODE solver for the reactions (VODE, CITE), and multigrid linear solvers for the pressure projection steps.
+Strang spitting CITE is used to integrate the thermodynamic variables, a second order projection method to integrate the velocity subject to the divergence constraint, and a velocity splitting scheme to hydrodynamically evolve the base state.
+The original \maestro~code was developed using the pure Fortran 90 \fboxlib~software framework CITE; \maestroex~instead uses the C++/Fortran 90 \amrex~framework CITE.
+
+\maestro~has been used to model a number of astrophysical systems, including convection in white dwarfs prior to type Ia supernovae CITE, convection in massive stars CITE and type I X-ray bursts CITE.
+
+As well as being built using the new \amrex~framework, the numerical algorithm implemented in \maestroex~improves upon the original \maestro~algorithm in a number of ways.
+The temporal integration method has been greatly simplified without compromising the second order accuracy, and a new spherical base state mapping has been implemented in order to reduce mapping errors between spherical and Cartesian grids.
+\amrex~has also allowed us to use MPI+OpenMP parallelism, and has been shown to scale well to over 10,000 MPI processes.
+
+Further details of the algorithm implemented in \maestroex~and its performance can be found in \cite{Fan2019}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{AMReX-Astro development workflow} \label{sec:workflow}
+
+% this section sounds kind of preachy
+
+\maestroex~is part of the \amrexastro~suite of open source adaptive mesh refinement hydrodynamics codes for astrophysical flows.
+Other codes in this family include \castro, an astrophysical radiation hydrodynamics simulation code, and \nyx, an N-body hydrodynamics cosmological simulation code.
+All three codes are developed using the \amrex~software framework, and share much in common in terms of their development, structure and numerical methods.
+Developers of the different codes work closely together (in fact many of the developers work on more than one of the codes), with their shared structure and framework meaning that new features developed in one code can easily be replicated in the others.
+Examples of this include the SDC solvers CITE and GPU port that have recently been implemented in \castro, and are currently in the process of being implemented in \maestroex.
+
+All codes in the \amrexastro~suite are open source, with all development carried out in public repositories hosted on GitHub.
+We believe that having both the codes and the development process completely open promotes good scientific practices, as it means that results from our simulations are reproducible and the codes used to produce them can be investigated.\MarginPar{I feel like there is a better word here than investigated} Others can use the codes for their own scientific investigations, and it is possible to adapt all or parts of the codes to suite new problems.
+Using version control allows us to keep a record of the codes' development process, helping us to track down bugs and meaning that new and existing developers can learn from previous mistakes.
+The development branches of the codes are tested nightly using a test suite of problems, checking that any new additions to the code have significantly changed any of the solutions or significantly slowed down the performance.
+
+New versions of the codes are released on the first of each month, and are archived on Zenodo (which also provides us with DOIs, further enhancing the reproducibility of our results).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Rotation} \label{sec:rotation}
+rotation mixes things up
+
+\begin{itemize}
+\item Ignore centrifugal force - for slow rotation can neglect centrifugal, only include Coriolis force
+\item Introduce new pressure to balance centrifugal potential $$\rho \nabla \Phi_{\text{eff}} = \nabla p_0 + \nabla p_1 = \rho \nabla \phi + \frac{1}{2}\rho \nabla\Omega^2 r^2$$
+\item Rewrite base state in terms of equipotentials $$p_0(r) \rightarrow p_0(\Phi_{\text{eff}}) $$
+\item 2d base state $$p_0(r)  \rightarrow p_0(r, \theta)$$
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{GPUs} \label{sec:gpus}
 GPUs can make things faster
 
-\section{Summary}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Summary} \label{sec:summary}
 
@@
 
-\ack acknowledgements here ~\cite{yt}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\ack acknowledgements here % ~\cite{yt}
 
@@
 
-%\section*{References}
-
 \bibliographystyle{iopart-num}

### 2019-08-29 astronum_2019_MAESTROeX/astronum-maestroex.tex
(very) rough draft of the rest of the maestroex astronum paper

@@
 \section{Rotation} \label{sec:rotation}
-rotation mixes things up
+Currently, \maestroex~is able to model spherically symmetric systems, however it has no support for rotation.
+We wish to model convection in the interiors of massive stars, and it is known that these stars often rotate and that trotation can have have significant effect on mixing at convective shell boundaries.
+We are therefore currently exploring several possible ways of implementing rotation in \maestroex.
 
-\begin{itemize}
-\item Ignore centrifugal force - for slow rotation can neglect centrifugal, only include Coriolis force
-\item Introduce new pressure to balance centrifugal potential $$\rho \nabla \Phi_{\text{eff}} = \nabla p_0 + \nabla p_1 = \rho \nabla \phi + \frac{1}{2}\rho \nabla\Omega^2 r^2$$
-\item Rewrite base state in terms of equipotentials $$p_0(r) \rightarrow p_0(\Phi_{\text{eff}}) $$
-\item 2d base state $$p_0(r)  \rightarrow p_0(r, \theta)$$
-\end{itemize}
+The simplest way to include rotation is to ignore the effects of the centrifugal force, and only include the Coriolis force.
+This was done in \maestro, and should be a reasonable approximation for slowly rotating systems.
+
+Another way that rotation could be incorporated would be to introduce a new `rotational pressure term', $p_1$, in order to balance the centrifugal potential:
+\begin{equation}
+    \nabla p_0 + \nabla p_1 = \rho \nabla \phi + \frac{1}{2}\rho \nabla\Omega^2 r^2 = \rho \nabla \Phi_{\text{eff}},
+\end{equation}
+where $\phi = \phi(r)$ is the gravitational potential (so $\nabla \phi(r) = -g \bm{e}_r$), $\Omega$ is the rotational frequency, and $\Phi_{\text{eff}}$ is the new effective gravitational potential.
+
+In \maestroex, we currently write the 1d base state as a function of the radial coordinate, $r$.
+For a spherical system, this therefore assumes the base state to be spherically symmetric.
+However, a star that is rotating at a significant rate will no longer be spherical: it will instead become oblique, bulging around the equator.
+In order to capture the star's deformation without sacrificing the 1d base state, we can rewrite the base state pressure as a function of the effective potential: $p_0(r)\rightarrow p_0(\Phi_{\text{eff}})$.
+
+The massive stars that we are interested in modelling are believed to be relatively slow rotators, so it is likely that one of the simpler approaches outlined above should be sufficient.
+However, for modelling e.g.~type I X-ray bursts, it may be that a more complex approach is needed.
+Type I X-ray bursts are produced by thermonuclear deflagrations in the liquid surface layers of fast rotating neutron stars.
+These stars typically rotate at frequencies of $\sim$300\textendash600 Hz, a significant fraction of the stars' breakup velocity.
+This produces non-negligible deformation of the star about the equator, and may cause surface flows with a significant dependence on the meriodional angle.
+To properly capture the effects of rotation in such a system, it may therefore be necessary to use a 2d base state, $p_0(r)\rightarrow p_0(r, \theta)$.
+
+% \begin{itemize}
+% \item Ignore centrifugal force \textemdash~for slow rotation can neglect centrifugal, only include Coriolis force
+% \item Introduce new pressure to balance centrifugal potential $$\rho \nabla \Phi_{\text{eff}} = \nabla p_0 + \nabla p_1 = \rho \nabla \phi + \frac{1}{2}\rho \nabla\Omega^2 r^2$$
+% \item Rewrite base state in terms of equipotentials $$p_0(r) \rightarrow p_0(\Phi_{\text{eff}}) $$
+% \item 2d base state $$p_0(r)  \rightarrow p_0(r, \theta)$$
+% \end{itemize}
 
@@
 \section{GPUs} \label{sec:gpus}
-GPUs can make things faster
+
+\begin{itemize}
+    \item Why GPUs? Big computers use them.
+    \item GPUs in \amrex: it does everything for you.
+Also important to note performance portability: GPU-ported AMReX codes still run on other architectures, optimizations for GPU often improve performance on MPI+OpenMP as well.
+    \item GPUs in \maestroex: ported most of the source terms, working on the hydro. 1d base state causing headaches, especially for big problems where we suspect it may not fit in GPU memory.
+    \item Performance plots? Do these at the hackathon?
+\end{itemize}
 
@@
 
-Summary
+\maestroex~is an open source code for modelling low Mach number astrophysical codes.
+We use the low Mach number approximation, which allows us to use a much larger timestep than conventional compressible schemes by filtering out the soundwaves.
+Unlike other sound-proof methods, we still retain background stratification and large pressure perturbations from local compositional changes and heating, both of which are particularly important for modelling atmospheres and burning.
+
+\maestroex~is a new and improved version of our previous code \maestro.
+It is based on \amrex~(rather than \fboxlib), which means we can exploit its powerful new solvers and ongoing improvements and don't have to write so much Fortran.
+It is part of the \amrexastro~suite of open source adaptive mesh refinement astrophysical hydrodynamics codes.
+New features developed in other codes in the family can therefore be easily implemented in \maestroex.
+As development is open, results from our simulations can be reproduced and monthly archiving to Zenodo ensures software sustainability.
+
+We're currently working on implementing rotation in \maestroex.
+For modelling the interiors of massive stars (which rotate relatively slowly), we may be able to get away with neglecting the centrifugal force and simply add the Coriolis force as a source term.
+However, this is no fun so we're also looking at doing things like modelling the centrifugal force as an effective pressure term, rewriting the base state as a function of the effective gravitational potential and using a 2d base state which is also a function of the meriodional angle.
+
+Copying the work done for \castro, we're currently in the process of porting \maestroex~to run on GPUs so that we can exploit the latest supercomputer architectures.
+So far, we've ported source terms and are in the process of porting the hydro.
 
@@
 
-\ack acknowledgements here % ~\cite{yt}
+\ack Visualizations were probably done using~\cite{yt}.
 

### 2019-08-29 astronum_2019_MAESTROeX/astronum-maestroex.tex
add in lots of citations

@@
 Sound-proof methods for modelling low Mach number flows can take a variety of forms.
-One technique is to use preconditioners in order to reduce the stiffness of the equations, allowing larger timesteps to be used CITE.
-Another technique is to modify the flow parameters, artificially boosting the speed of the flow without changing its behavior so that the system evolves faster and fewer timesteps are required CITE.
+One technique is to use preconditioners in order to reduce the stiffness of the equations, allowing larger timesteps to be used \cite{Miczek2014,Barsukow2016}.
+Another technique is to modify the flow parameters, artificially boosting the speed of the flow without changing its behavior so that the system evolves faster and fewer timesteps are required \citep{Rempel2005,Hotta2012}.
+Fully implicit time integration codes, such as those used by \cite{Viallet2011,Viallet2015,Goffrey2016}, can use arbitratily timesteps, so are no longer restricted by the CFL condition.
 
 The technique that we use in \maestroex~is to modify the fluid equations themselves so as to filter out the soundwaves.
-This is a similar approach to the incompressible CITE and anelastic CITE approximations, however the approximation that we use (variously called the low Mach number approximation  CITE and the generalized pseudo-incompressible approximation CITE) allows background stratification and large pressure perturbations due to heating and changes in composition.
+This is a similar approach to the incompressible \cite{Boussinesq1901} and anelastic \cite{Ogura1962a,Gough1968,Durran1989} approximations, however the approximation that we use (variously called the low Mach number approximation \cite{Day2000,Almgren2006a,Nonaka2010} and the generalized pseudo-incompressible approximation \cite{Vasil2013}) allows background stratification and large pressure perturbations due to heating and changes in composition.
 This is achieved by decomposing the pressure into a 1d hydrostatic base state, $p_0 = p_0(r, t)$, and a dynamic pressure perturbation, $\pi = \pi(r, \bm{x}, t)$, such that $p(r, \bm{x}, t) = p_0 + \pi$.
@@
 \begin{equation}
-    \beta_0(r,t) = \rho_0(0,t) \exp\left(\int_0^r dr'\, \frac{1}{\bar{\Gamma}_1p_0}\pd{p_0}{r'} \right),
+    \beta_0(r,t) = \rho_0(0,t) \exp\left(\int_0^r dr'\, \frac{1}{\overline{\Gamma}_1p_0}\pd{p_0}{r'} \right),
 \end{equation}
-where $\bar{\Gamma}_1$ is the lateral average of the first adiabatic exponent, $\Gamma_1 = d(\ln p)/d(\ln \rho)|_s$, and $s$ is the entropy.
+where $\overline{\Gamma}_1$ is the lateral average of the first adiabatic exponent, $\Gamma_1 \equiv d(\ln p)/d(\ln \rho)|_s$, and $s$ is the entropy.
 The equations are closed by casting the equation of state as a velocity divergence constraint.
 This is done by taking the Lagrangian derivative of the EoS for pressure as a function of the thermodynamic variables, substituting in the equations of motion for mass and energy, and requiring that the pressure is described by a function of $r$ and $t$ based on the condition of hydrostatic equilibrium.
-Details of this derivation can be found in CITE.
+Details of this derivation can be found in \cite{Almgren2006a,Almgren2006b}.
 The constraint is given by
 \begin{equation}
-    \nabla\cdot\left(\beta_0\Ub\right) = \beta_0 \left(S - \frac{1}{\bar{\Gamma}_1 p_0}\pd{p_0}{t} \right).
+    \nabla\cdot\left(\beta_0\Ub\right) = \beta_0 \left(S - \frac{1}{\overline{\Gamma}_1 p_0}\pd{p_0}{t} \right).
 \end{equation}
@@
 Like \maestroex, \maestro~is a block-structured adaptive mesh refinement (AMR) code for modelling low Mach number astrophysical codes.
-It was developed over a series of papers CITE.
+It was developed over a series of papers \cite{Almgren2006a,Almgren2006b,Almgren2008a,Zingale2009,Nonaka2010}.
 The system of low Mach number equations are solved using an explicit Godunov approach for the advection, a stiff ODE solver for the reactions (VODE, CITE), and multigrid linear solvers for the pressure projection steps.
 Strang spitting CITE is used to integrate the thermodynamic variables, a second order projection method to integrate the velocity subject to the divergence constraint, and a velocity splitting scheme to hydrodynamically evolve the base state.
-The original \maestro~code was developed using the pure Fortran 90 \fboxlib~software framework CITE; \maestroex~instead uses the C++/Fortran 90 \amrex~framework CITE.
+The original \maestro~code was developed using the pure Fortran 90 \fboxlib~software framework \cite{Zhang2016}; \maestroex~instead uses the C++/Fortran 90 \amrex~framework \cite{Zhang2019}.
 
-\maestro~has been used to model a number of astrophysical systems, including convection in white dwarfs prior to type Ia supernovae CITE, convection in massive stars CITE and type I X-ray bursts CITE.
+\maestro~has been used to model a number of astrophysical systems, including convection in white dwarfs prior to type Ia supernovae \cite{Zingale2011,Nonaka2011,Malone2014a,Zingale2013,Jacobs2016}, convection in massive stars \cite{Gilet2013} and type I X-ray bursts \cite{Malone2011,Malone2014,Zingale2015}.
 
@@
 \maestroex~is part of the \amrexastro~suite of open source adaptive mesh refinement hydrodynamics codes for astrophysical flows.
-Other codes in this family include \castro, an astrophysical radiation hydrodynamics simulation code, and \nyx, an N-body hydrodynamics cosmological simulation code.
+Other codes in this family include \castro, an astrophysical radiation hydrodynamics simulation code \cite{Almgren2010}, and \nyx, an N-body hydrodynamics cosmological simulation code \cite{Almgren2013}.
 All three codes are developed using the \amrex~software framework, and share much in common in terms of their development, structure and numerical methods.
 Developers of the different codes work closely together (in fact many of the developers work on more than one of the codes), with their shared structure and framework meaning that new features developed in one code can easily be replicated in the others.
-Examples of this include the SDC solvers CITE and GPU port that have recently been implemented in \castro, and are currently in the process of being implemented in \maestroex.
+Examples of this include the SDC solvers \cite{Zingale2019} and GPU port that have recently been implemented in \castro, and are currently in the process of being implemented in \maestroex.
 

### 2019-09-02 astronum_2019_MAESTROeX/astronum-maestroex.tex
fixed bibliography entries

@@
 
-All codes in the \amrexastro~suite are open source, with all development carried out in public repositories hosted on GitHub.
+All codes in the \amrexastro~suite are open source, with all development carried out in public repositories hosted on GitHub\footnote{\url{https://github.com/AMReX-Astro}}.
 We believe that having both the codes and the development process completely open promotes good scientific practices, as it means that results from our simulations are reproducible and the codes used to produce them can be investigated.\MarginPar{I feel like there is a better word here than investigated} Others can use the codes for their own scientific investigations, and it is possible to adapt all or parts of the codes to suite new problems.
@@
 
-New versions of the codes are released on the first of each month, and are archived on Zenodo (which also provides us with DOIs, further enhancing the reproducibility of our results).
+New versions of the codes are released on the first of each month, and are archived on Zenodo\footnote{\url{https://zenodo.org/}} (which also provides us with DOIs, further enhancing the reproducibility of our results).
 

### 2019-09-02 astronum_2019_MAESTROeX/astronum-maestroex.tex
added GPU section text

--- 2019-08-29 astronum_2019_MAESTROeX/astronum-maestroex.tex
@@
 The composition and structure of the star prior to collapsing provides the initial data for these models, so this needs to be accurate if the supernova models are to be trusted.
-Modelling this convection is challenging using conventional compressible schemes because the domain size is very large (i.e.
-a significant fraction of the interior of the entire star), high resolution is required in order to properly capture the turbulent mixing that occurs at shell boundaries, and long time periods are required (multiple convective turnover times).
+Modelling this convection is challenging using conventional compressible schemes because the domain size is very large (i.e.~a significant fraction of the interior of the entire star), high resolution is required in order to properly capture the turbulent mixing that occurs at shell boundaries, and long time periods are required (multiple convective turnover times).
 In \sref{sec:rotation}, we shall describe ongoing work to model this problem in \maestroex, in particular describing how we are implementing rotation.
@@
 \section{Rotation} \label{sec:rotation}
+
 Currently, \maestroex~is able to model spherically symmetric systems, however it has no support for rotation.
@@
 
-\begin{itemize}
-    \item Why GPUs? Big computers use them.
-    \item GPUs in \amrex: it does everything for you.
+% \begin{itemize}
+%     \item Why GPUs? Big computers use them.
+%     \item GPUs in \amrex: it does everything for you.
 Also important to note performance portability: GPU-ported AMReX codes still run on other architectures, optimizations for GPU often improve performance on MPI+OpenMP as well.
-    \item GPUs in \maestroex: ported most of the source terms, working on the hydro. 1d base state causing headaches, especially for big problems where we suspect it may not fit in GPU memory.
-    \item Performance plots? Do these at the hackathon?
-\end{itemize}
+%     \item GPUs in \maestroex: ported most of the source terms, working on the hydro. 1d base state causing headaches, especially for big problems where we suspect it may not fit in GPU memory.
+%     \item Performance plots? Do these at the hackathon?
+% \end{itemize}
+
+\MarginPar{I feel like someone who knows what they're talking about should write/seriously edit this section}
+The latest generation of supercomputers to come online (e.g.~Summit, soon Perlmutter and Frontier) rely heavily on GPUs in order to achieve high performance while keeping energy consumption to a minimum (GPUs consume considerably less power per flop than CPUs).
+This trend is set to continue: supercomputers are moving to new architectures in order to increase their performance, rather than simply adding more and more CPUs.
+In order for HPC codes to be able to exploit these latest machines, it is therefore becoming increasingly necessary to port codes to run on GPUs.
+
+We have begun porting \maestroex~to GPUs, leveraging the GPU capabilities built into \amrex~and following the lessons learnt porting \castro~to GPUs.
+In the \amrex~framework, routines can be ported to GPUs by simply inserting a few macros into the code.
+The boilerplate code to required to compile this for GPU is then generated by a custom preprocessor prior to compilation.
+CUDA managed memory is used to take care of data transfers between the CPU and GPU, so no explicit copies of data to/from the CPU and GPU are required.
+One of the key emphases of the \amrex~approach to GPUs is to maintain performance portability: porting code to GPUs should not come at the expense of the code's performance on CPU-only machines.
+Because GPU-specific code is created by the preprocessor prior to compilation, it is not necessary to maintain separate versions of the code for different architectures.
+By changing the compiler flags, the same bit of code can be compiled for serial, MPI-only, OpenMP-only, MPI+OpenMP, GPU-only or MPI+GPU.
+In order to achieve the best performance on the GPU, some additional optimization is required.
+However, it has been found that these optimizations often improve the performance when the code is run on the CPU only as well.
+More details to appear in a forthcoming \amrexastro~GPU paper.
+
+So far we have ported most of the source terms and hydrodynamics in \maestroex~to run on GPUs.
+We have found that one of the challenges moving forwards is dealing with the 1d base state.
+Currently, all AMR subgrids have a copy of the entire 1d base state.
+This is not a problem for small problems where the 1d base state does not contain many cells compared to the number of cells in a single subgrid, however as the problem size and/or resolution gets larger, the size of the 1d base state grows to the point it is now significantly larger than the size of a single subgrid.
+Compared to CPUs, GPUs typically have less memory.
+For larger problems, it is therefore not possible to store the entire base state on the GPU.
+This significantly reduces the performance of the GPU parts of the code, as it means data will be copied back and forth between the CPU and GPU (a slow process).
+To alleviate this problem, we are looking at ways to reduce the amount of 1d base state data that needs to be stored on the GPU at any one time.
+For example, in some routines we map the 1d base state to the 2d grid \textendash for these, we can move this operation to the CPU so that the GPU does not require the 1d base state.
+It may be necessary for us to develop a way to split up the 1d base state, so that subgrids only receive the portions of the base state necessary for their calculations.
+
+Insert some performance plots here (do these at the hackathon?)
 
@@
 So far, we've ported source terms and are in the process of porting the hydro.
+
+More future work: SDC, more science, etc.
 

### 2019-09-11 xrb/paper.tex
update table numbers
Updated flame speeds and flame speed plots

@@
 \plottwo{flame}{speed}
-\caption{\label{fig:flame} Our ``standard'' flame. {\color{red} this should be redone for the boosted flame}}
+\caption{\label{fig:flame} Our ``standard'' flame.}
 \end{figure*}
@@
 
+\iffalse
 \input{flame_table_1d.tex}
+\fi
 
@@
 \startdata
-$T_\star$       & \multicolumn{2}{c}{$5\times 10^7$~K} \\
-$T_\mathrm{hi}$ & $2\times 10^8$~K & $1.2\times 10^9$~K \\
-$T_\mathrm{lo}$ & \multicolumn{2}{c}{$10^7$~K} \\
+$T_\star$       & \multicolumn{2}{c}{$10^8$~K} \\
+$T_\mathrm{hi}$ & $2\times 10^8$~K & $1.4\times 10^9$~K \\
+$T_\mathrm{lo}$ & \multicolumn{2}{c}{$8\times 10^6$~K} \\
 $\rho_\mathrm{int}$ & \multicolumn{2}{c}{$3.43\times 10^6~\gcc$} \\
-$\rho_\mathrm{fuel}$\tablenotemark{a} & $2.21\times 10^6~\gcc$ & $1.27\times 10^6~\gcc$ \\
+$\rho_\mathrm{fuel}$\tablenotemark{a} & $2.36\times 10^6~\gcc$ & $1.20\times 10^6~\gcc$ \\
 $\rho_\mathrm{cutoff}$ & \multicolumn{2}{c}{$10^{-4}~\gcc$} \\
 $g$             & \multicolumn{2}{c}{$-1.5\times 10^{14}~\mathrm{cm~s^{-2}}$} \\
-$H_\star$       & \multicolumn{2}{c}{5000~cm} \\
+$H_\star$       & \multicolumn{2}{c}{2000~cm} \\
 $\delta_\mathrm{atm}$ & \multicolumn{2}{c}{50~cm} \\
-$X_\star(\isotm{Fe}{56})\tablenotemark{b}$ & \multicolumn{2}{c}{1.0} \\
+$X_\star(\isotm{Ni}{56})\tablenotemark{b}$ & \multicolumn{2}{c}{1.0} \\
 $X_\mathrm{atm}(\isotm{He}{4})\tablenotemark{b}$ & \multicolumn{2}{c}{1.0} \\
@@
 	\startdata
-	10/10 boost & $9.46 \pm 0.04$  & $1.06 \pm 0.01$ \\
-	5/5 boost & $3.97 \pm 0.01$ & $0.56 \pm 0.01$ \\
-	iso7 network & $7.66 \pm 0.04$ & $-$ \\
-	1000 Hz & $26.1 \pm 0.20$ & $-$ \\
+	10/10 boost & $9.18 \pm 0.03$  & $1.06 \pm 0.01$ \\
+	5/5 boost & $4.00 \pm 0.01$ & $0.56 \pm 0.01$ \\
+	iso7 network & $7.56 \pm 0.02$ & $-$ \\
+	lowres & $9.33 \pm 0.04$ & $-$ \\
+	1000 Hz & $18.6 \pm 0.16$ & $-$ \\
 	\enddata
@@
 \centering
-\plotone{plots/results/speed_10_10.pdf}
-\caption{\label{fig:flame_speed}The position of the burning front for the 10/10 flame as a function of time. The dashed line shows a linear least squares fit for $t \gtrsim 10$ ms.}
+\plotone{plots/results/speedplot_all.pdf}
+\caption{\label{fig:flame_speed}The position of the burning front for each simulation run as a function of time. The dashed lines show linear least squares fits for $t \gtrsim 6$ ms.}
 \end{figure}

### 2019-09-16 MAESTROeX_I/paper.tex
updated keywords to UAT concepts as required by AAS
update to aastex63 style

@@
-%% This manuscript uses the AASTeX v6.2 LaTeX 2e macros
+%% Beginning of file 'sample63.tex'
+%%
+%% Modified 2019 June
+%%
+%% This is a sample manuscript marked up using the
+%% AASTeX v6.3 LaTeX 2e macros.
 %%
@@
 %%
-%% using aastex version 6.2
-\documentclass{aastex62}
+%% using aastex version 6.3
+\documentclass{aastex63}
 
@@
 %% There are 5 other style options available via an optional argument. They
-%% can be envoked like this:
+%% can be invoked like this:
 %%
-%% \documentclass[argument]{aastex62}
+%% \documentclass[arguments]{aastex63}
 %% 
@@
 %%
-%% There are other optional arguments one can envoke to allow other stylistic
+%% There are other optional arguments one can invoke to allow other stylistic
 %% actions. The available options are:
@@
 %%                  affiliation information after each name. Creates a much 
-%%                 long author list but may be desirable for short author papers
+%%                  longer author list but may be desirable for short 
+%%                  author papers.
+%% twocolappendix : make 2 column appendix.
+%%   anonymous    : Do not show the authors, affiliations and acknowledgments 
+%%                  for dual anonymous review.
 %%
@@
 %%
-%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex62}
+%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex63}
 %%
@@
 %%
-%%\hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
+%% \hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
 %%
@@
 %%
+%% Note that in v6.3 "bookmarks" has been changed to "true" in hyperref
+%% to improve the accessibility of the compiled pdf file.
+%%
 %% If you want to create your own macros, you can do so
@@
 %%
+
 \newcommand{\vdag}{(v)^\dagger}
@@
 %% Command to document which AAS Journal the manuscript was submitted to.
-%% Adds "Submitted to " the arguement.
+%% Adds "Submitted to " the argument.
 \submitjournal{ApJ}
 
-%% Mark up commands to limit the number of authors on the front page.
-%% Note that in AASTeX v6.2 a \collaboration call (see below) counts as
-%% an author in this case.
+%% For manuscript that include authors in collaborations, AASTeX v6.3
+%% builds on the \collaboration command to allow greater freedom to 
+%% keep the traditional author+affiliation information but only show
+%% subsets. The \collaboration command now must appear AFTER the group
+%% of authors in the collaboration and it takes TWO arguments. The last
+%% is still the collaboration identifier. The text given in this
+%% argument is what will be shown in the manuscript. The first argument
+%% is the number of author above the \collaboration command to show with
+%% the collaboration text. If there are authors that are not part of any
+%% collaboration the \nocollaboration command is used. This command takes
+%% one argument which is also the number of authors above to show. A
+%% dashed line is shown to indicate no collaboration. This example manuscript
+%% shows how these commands work to display specific set of authors 
+%% on the front page.
+%%
+%% For manuscript without any need to use \collaboration the 
+%% \AuthorCollaborationLimit command from v6.2 can still be used to 
+%% show a subset of authors.
 %
-%\AuthorCollaborationLimit=3
+%\AuthorCollaborationLimit=2
 %
-%% Will only show Schwarz, Muench and "the AAS Journals Data Scientist
-%% collaboration" on the front page of this example manuscript.
+%% will only show Schwarz & Muench on the front page of the manuscript
+%% (assuming the \collaboration and \nocollaboration commands are
+%% commented out).
 %%
@@
 %% with this command:
-% \watermark{text}
+%% \watermark{text}
 %% where "text", e.g. DRAFT, is the text to appear.  If the text is 
 %% long you can control the water-mark size with:
-%  \setwatermarkfontsize{dimension}
+%% \setwatermarkfontsize{dimension}
 %% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
@@
 %% one line. However, you may use \\ to force a line break if
-%% you desire. In v6.2 you can include a footnote in the title.
+%% you desire. In v6.3 you can include a footnote in the title.
 
 %% A significant change from earlier AASTEX versions is in the structure for 
-%% calling author and affilations. The change was necessary to implement
-%% autoindexing of affilations which prior was a manual process that could
+%% calling author and affiliations. The change was necessary to implement 
+%% auto-indexing of affiliations which prior was a manual process that could 
 %% easily be tedious in large author manuscripts.
@@
 %% The \author command is the same as before except it now takes an optional
-%% arguement which is the 16 digit ORCID. The syntax is:
+%% argument which is the 16 digit ORCID. The syntax is:
 %% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
@@
 %% during compilation, LaTeX will do some limited checking of the format of
-%% the ID to make sure it is valid.
+%% the ID to make sure it is valid. If the "orcid-ID.png" image file is 
+%% present or in the LaTeX pathway, the OrcID icon will appear next to
+%% the authors name.
 %%
 %% Use \affiliation for affiliation information. The old \affil is now aliased
-%% to \affiliation. AASTeX v6.2 will automatically index these in the header.
+%% to \affiliation. AASTeX v6.3 will automatically index these in the header.
 %% When a duplicate is found its index will be the same as its previous entry.
@@
 %% own line so you can put multiple email address in one \email call. A new
-%% \correspondingauthor command is available in V6.2 to identify the
+%% \correspondingauthor command is available in V6.3 to identify the
 %% corresponding author of the manuscript. It is the author's responsibility
@@
 
-\correspondingauthor{Duoming Fan; DFan@lbl.gov}
+\correspondingauthor{Duoming Fan}
+\email{DFan@lbl.gov}
 
 \author[0000-0002-3246-4315]{Duoming Fan}
-\affil{Lawrence Berkeley National Laboratory \\
+\affiliation{Lawrence Berkeley National Laboratory \\
 Center for Computational Sciences and Engineering \\
@@
 \author[0000-0003-1791-0265]{Andrew Nonaka}
-\affil{Lawrence Berkeley National Laboratory \\
+\affiliation{Lawrence Berkeley National Laboratory \\
 Center for Computational Sciences and Engineering \\
@@
 \author[0000-0003-2103-312X]{Ann S. Almgren}
-\affil{Lawrence Berkeley National Laboratory \\
+\affiliation{Lawrence Berkeley National Laboratory \\
 Center for Computational Sciences and Engineering \\
@@
 \author[0000-0002-1530-781X]{Alice Harpole}
-\affil{Stony Brook University \\
+\affiliation{Stony Brook University \\
 Department of Physics and Astronomy \\
@@
 \author[0000-0001-8401-030X]{Michael Zingale}
-\affil{Stony Brook University \\
+\affiliation{Stony Brook University \\
 Department of Physics and Astronomy \\
@@
 
-%% AASTeX 6.2 has the new \collaboration and \nocollaboration commands to
+%% AASTeX 6.3 has the new \collaboration and \nocollaboration commands to
 %% provide the collaboration status of a group of authors. These commands 
@@
 %% Mark off the abstract in the ``abstract'' environment.
+
 \begin{abstract}
@@
 %% keywords and the rules for their use.
-\keywords{convection, hydrodynamics, methods: numerical, nuclear reactions, nucleosynthesis, abundances, supernovae: general}
+\keywords{Stellar convective zones, Hydrodynamics, Computational methods, 
+Nuclear astrophysics, Nucleosynthesis, Nuclear abundances, Type Ia supernovae}
 
@@
 
-\bibliographystyle{aasjournal.bst}
-\bibliography{references}
+%% For this sample we use BibTeX plus aasjournals.bst to generate the
+%% the bibliography. The sample63.bib file was populated from ADS. To
+%% get the citations to show in the compiled file do the following:
+%%
+%% pdflatex sample63.tex
+%% bibtext sample63
+%% pdflatex sample63.tex
+%% pdflatex sample63.tex
+
+\bibliography{references}{}
+\bibliographystyle{aasjournal}
+
+%% This command is needed to show the entire author+affiliation list when
+%% the collaboration and author truncation commands are used.  It has to
+%% go at the end of the manuscript.
+%\allauthors
+
+%% Include this line if you are using the \added, \replaced, \deleted
+%% commands to see a summary list of all changes at the end of the article.
+%\listofchanges
 
 \end{document}
+
+% End of file `sample63.tex'.

### 2019-09-17 MAESTROeX_I/intro.tex
starting on response

@@
 interest, and numerical approaches based on compressible hydrodynamics
-are intractable, even on modern supercomputers.  One approach to this
+are intractable, even on modern supercomputers.  \replaced{One approach to this
 problem is to use low Mach number models.  In a low Mach number
-approach, sound waves are eliminated from the governing equations
+approach,}{Our approach to this problem is to use a low Mach number model where} 
+sound waves are eliminated from the governing equations
 while retaining compressibilitiy effects due to, e.g., nuclear energy
-release, stratification, compositional changes, and thermal diffusion.  When the Mach
-number (the ratio of the characteristic fluid velocity over the
+release, stratification, compositional changes, and thermal diffusion.
+When the Mach number (the ratio of the characteristic fluid velocity over the
 characteristic sound speed; Ma $= U/c$) is small, the resulting system
@@
 
-Low Mach number models have been developed for a variety of contexts
+Low Mach number models \added{that do not contain acoustic waves}
+have been developed for a variety of contexts
 including combustion \citep{day2000numerical}, terrestrial atmospheric
 modeling \citep{durran:1989,oneill:2014,duarte2015low}, and elastic
-solids \citep{abbate2017all}.  For astrophysical applications, a
+solids \citep{abbate2017all}.  
+For astrophysical applications, a
 number of approaches to modeling low Mach number flows have been
@@
 \cite{Lin:2006}, however this approach is only first-order accurate and does
-not account for atmospheric expansion.  There are also semi-implicit all-Mach
+not account for atmospheric expansion.
+\added{Other approaches model acoustic wave propagation with various strategies
+to efficiently and robustly handle the low Mach number limit.}
+There are \deleted{also} semi-implicit all-Mach
 number solvers, where the Euler equations are split into an acoustic
@@
 the system at low Mach numbers, while retaining the correct scaling
-behavior. In the reduced speed of sound technique (RSST) and related
+behavior \added{in an acoustic limit of the Euler equations that 
+retains $\mathcal{O}$(Ma) pressure fluctuations}.
+In the reduced speed of sound technique (RSST) and related
 methods, the speed of sound is artificially reduced by including a
@@
 compressible Euler equations
-\citep{Viallet2011,kifonidis:2012,Viallet2015,Goffrey2016}.  The MUSIC code
+\citep{kifonidis:2012,Viallet2015,Goffrey2016}.
+\explain{removed outdated Viallet2011 citation and corrected the typo in the Viallet2015 citation}
+The MUSIC code
 uses fully implicit time integration for the compressible Euler

### 2019-09-17 MAESTROeX_I/algorithm.tex
more responses

@@
 in computing $\rho_0$ by averaging would cause artificial acceleration in the velocity field due to the interpolation error 
-from the Cartesian grid to and from the radial base state.  By retaining the radial base state as an irregularly spaced 
-array, the effects due to interpolation error are nearly eliminated, but not completely eliminated since there are still 
-machine precision effects resulting from averaging a large number of numerical values.
+from the Cartesian grid to and from the radial base state.  By retaining the radial base state as an irregularly spaced array, 
+\replaced{the effects due to interpolation error are nearly eliminated, but not completely eliminated since there are still machine precision 
+effects resulting from averaging a large number of numerical values.}
+{the truncation error in the mapping scheme is completely eliminated.  The only source of error is due to machine precision effects resulting 
+from averaging a large number of values, which is not significant over the course of any simulation.}
 We note that $\Delta r$ decreases as the base state moves further from the center of the star, 
@@
 \item {\bf React State}$[(\rho X_k)^{\rm in},(\rho h)^{\rm in},p_0]\rightarrow[(\rho X_k)^{\rm out},(\rho h)^{\rm out},(\rho\dot\omega),(\rho\Hnuc)]$ 
-uses VODE \citep{vode} to integrate the species and enthalpy due to reactions over $\Delta t/2$ by solving
+uses 
+\added{the multistep variable-coefficient Adams-Moulton and Backward Differentiation Formula (BDF) methods in the}
+VODE \citep{vode} 
+\added{package}
+to integrate the species and enthalpy due to reactions over $\Delta t/2$ by solving
 \begin{equation}

### 2019-09-17 astronum_2019_MAESTROeX/astronum-maestroex.tex
small updates

@@
 For many flows in astrophysical systems, the magnitude of the fluid velocity is much less than the soundspeed.
-Consequently, the Mach number, $\textrm{Ma} = |\Ub| / c_s \ll 1$.
-Such low Mach number flows are challenging to model using standard compressible schemes, where the maximum size of the timestep is determined by the CFL condition: for a grid with cell spacing $\Delta x$, the timestep $\Delta t <  \Delta x / \max(|\Ub|+c_s)$.
-In low Mach number flows, this condition is dominated by the contributions of the sound speed, with the result that many fine timesteps are required to achieve high spatial resolution.
+Consequently, the Mach number, the ratio of the fluid velocity to the sound speed, is much less than 1: $\textrm{Ma} = |\Ub| / c_s \ll 1$.
+Such low Mach number flows are challenging to model with standard compressible schemes using explicit timestepping methods, where the maximum size of the timestep is determined by the Courant-Friedrich-Lewy (CFL) condition.
+This states that for a grid with cell spacing $\Delta x$, the timestep $\Delta t <  \Delta x / \max(|\Ub|+c_s)$.
+For low Mach number flows, this condition is dominated by the contributions of the sound speed, with the result that many fine timesteps are required to achieve high spatial resolution.
+Long timescale, high resolution simulations therefore become extremely computationally expensive.
 
 This restriction can be lifted by using sound-proof methods.
-These methods involve modifying the fluid equations, modifying the computational algorithm and/or modifying the flow parameters in order to allow much larger timesteps to be used.
-In \maestroex, we use the \emph{low Mach number approximation}.
+These methods can involve modifying the fluid equations, modifying the computational algorithm and/or modifying the flow parameters in order to allow much larger timesteps to be used.
+In \maestroex, we use the \emph{low Mach number approximation}, a limit of the compressible Euler equations which effectively filters out sound waves.
 An overview of this shall be given in \sref{sec:low_mach_hydro}, with further details of the \maestroex~code and our development workflow given in \sref{sec:maestroex} and \sref{sec:workflow}.
@@
 An example of an astrophysical low Mach number flow is convection within the interiors of massive stars.
-Prior to core collapse, the interiors of massive stars consist of a series of convective burning shells, separated by inert, non-convective shells.
-Moving from the outermost layers inwards, the shells consist of heavier and heavier elements, with the core burning elements up to Fe.
-An accurate picture of the structure of such stars in the minutes before core collapse is important for supernova modelling.
-The composition and structure of the star prior to collapsing provides the initial data for these models, so this needs to be accurate if the supernova models are to be trusted.
+Prior to core collapse, the interiors of massive stars consist of a sequence of convective burning shells, separated by inert, non-convective shells.
+Moving from the outermost layers inwards, these shells consist of heavier and heavier elements, with the core burning elements up to ${}^{56}$Fe.
+An accurate description of the structure of such stars in the minutes before core collapse is important for supernova modelling.
+The composition and structure of the star prior to collapse provides the initial data for these models, so this needs to be accurate if the resulting supernova models are to be trusted.
 Modelling this convection is challenging using conventional compressible schemes because the domain size is very large (i.e.~a significant fraction of the interior of the entire star), high resolution is required in order to properly capture the turbulent mixing that occurs at shell boundaries, and long time periods are required (multiple convective turnover times).
-In \sref{sec:rotation}, we shall describe ongoing work to model this problem in \maestroex, in particular describing how we are implementing rotation.
+In \sref{sec:rotation}, we shall describe ongoing work to model this problem in \maestroex, in particular describing how we are adding rotation to our existing scheme.
 
-The latest generation of supercomputers coming on line are relying more and more on GPU architectures in order to increase the performance whilst minimizing the power requirements.
-For HPC codes to best exploit the top machines, it is therefore necessary to port these codes so that they can run on GPUs.
+The latest generation of supercomputers coming online are relying more and more on GPU architectures in order to increase performance whilst minimizing power requirements.
+For HPC codes to best exploit the most powerful machines, it is therefore becoming increasingly necessary to port these codes so that they can run on GPUs.
 In \sref{sec:gpus}, we describe our work to port \maestroex~to run on GPUs and demonstrate its performance.
@@
 Another technique is to modify the flow parameters, artificially boosting the speed of the flow without changing its behavior so that the system evolves faster and fewer timesteps are required \citep{Rempel2005,Hotta2012}.
-Fully implicit time integration codes, such as those used by \cite{Viallet2011,Viallet2015,Goffrey2016}, can use arbitratily timesteps, so are no longer restricted by the CFL condition.
+Fully implicit time integration codes, such as those used by \cite{Viallet2011,Viallet2015,Goffrey2016}, are no longer restricted by the CFL condition and so can use arbitratily large timesteps.
 
 The technique that we use in \maestroex~is to modify the fluid equations themselves so as to filter out the soundwaves.
-This is a similar approach to the incompressible \cite{Boussinesq1901} and anelastic \cite{Ogura1962a,Gough1968,Durran1989} approximations, however the approximation that we use (variously called the low Mach number approximation \cite{Day2000,Almgren2006a,Nonaka2010} and the generalized pseudo-incompressible approximation \cite{Vasil2013}) allows background stratification and large pressure perturbations due to heating and changes in composition.
-This is achieved by decomposing the pressure into a 1d hydrostatic base state, $p_0 = p_0(r, t)$, and a dynamic pressure perturbation, $\pi = \pi(r, \bm{x}, t)$, such that $p(r, \bm{x}, t) = p_0 + \pi$.
-In order for the approximation to hold, $|\pi|/p_0 = O(\mathrm{Ma}^2)$.
+This is a similar approach to the incompressible \cite{Boussinesq1901} and anelastic \cite{Ogura1962a,Gough1968,Durran1989} approximations, however the approximation that we use (variously called the low Mach number approximation \cite{Day2000,Almgren2006a,Nonaka2010} and the generalized pseudo-incompressible approximation \cite{Vasil2013}) allows for background stratification and large pressure perturbations due to heating and changes in composition.
+This is achieved by decomposing the pressure into a 1d hydrostatic base state, $p_0 = p_0(r, t)$, and a dynamic pressure perturbation, $\pi = \pi(r, \bm{x}, t)$, such that the full pressure is given by $p(r, \bm{x}, t) = p_0(r, t) + \pi(r, \bm{x}, t)$.
+In order for the approximation to hold, it is necessary that $|\pi|/p_0 = O(\mathrm{Ma}^2)$.
+
 The low Mach number fluid equations are given by 
@@
 where $\overline{\Gamma}_1$ is the lateral average of the first adiabatic exponent, $\Gamma_1 \equiv d(\ln p)/d(\ln \rho)|_s$, and $s$ is the entropy.
-The equations are closed by casting the equation of state as a velocity divergence constraint.
+The equations are closed by casting the equation of state (EoS) as a velocity divergence constraint.
 This is done by taking the Lagrangian derivative of the EoS for pressure as a function of the thermodynamic variables, substituting in the equations of motion for mass and energy, and requiring that the pressure is described by a function of $r$ and $t$ based on the condition of hydrostatic equilibrium.
 Details of this derivation can be found in \cite{Almgren2006a,Almgren2006b}.
-The constraint is given by
+This constraint is given by
 \begin{equation}
@@
 \end{equation}
-where we define 
+where we define the following thermodynamic quantities as
 \begin{align*}
@@
 Like \maestroex, \maestro~is a block-structured adaptive mesh refinement (AMR) code for modelling low Mach number astrophysical codes.
-It was developed over a series of papers \cite{Almgren2006a,Almgren2006b,Almgren2008a,Zingale2009,Nonaka2010}.
-The system of low Mach number equations are solved using an explicit Godunov approach for the advection, a stiff ODE solver for the reactions (VODE, CITE), and multigrid linear solvers for the pressure projection steps.
-Strang spitting CITE is used to integrate the thermodynamic variables, a second order projection method to integrate the velocity subject to the divergence constraint, and a velocity splitting scheme to hydrodynamically evolve the base state.
+Its development is described in a series of papers: \cite{Almgren2006a,Almgren2006b,Almgren2008a,Zingale2009,Nonaka2010}.
+In \maestro, the system of low Mach number equations is solved using an explicit Godunov approach for the advection, a stiff ODE solver for the reactions (VODE, CITE), and multigrid linear solvers for the pressure projection steps.
+Strang splitting CITE is used to integrate the thermodynamic variables, a second order projection method to integrate the velocity subject to the divergence constraint, and a velocity splitting scheme to hydrodynamically evolve the base state.
 The original \maestro~code was developed using the pure Fortran 90 \fboxlib~software framework \cite{Zhang2016}; \maestroex~instead uses the C++/Fortran 90 \amrex~framework \cite{Zhang2019}.
@@
 
-As well as being built using the new \amrex~framework, the numerical algorithm implemented in \maestroex~improves upon the original \maestro~algorithm in a number of ways.
+The numerical algorithm implemented in \maestroex~improves upon the original \maestro~algorithm in a number of ways.
 The temporal integration method has been greatly simplified without compromising the second order accuracy, and a new spherical base state mapping has been implemented in order to reduce mapping errors between spherical and Cartesian grids.
-\amrex~has also allowed us to use MPI+OpenMP parallelism, and has been shown to scale well to over 10,000 MPI processes.
+Utilising the \amrex~framework has allowed us to implement MPI+OpenMP parallelism, which has been shown to scale well to over 10,000 MPI processes.
 
@@
 We believe that having both the codes and the development process completely open promotes good scientific practices, as it means that results from our simulations are reproducible and the codes used to produce them can be investigated.\MarginPar{I feel like there is a better word here than investigated} Others can use the codes for their own scientific investigations, and it is possible to adapt all or parts of the codes to suite new problems.
-Using version control allows us to keep a record of the codes' development process, helping us to track down bugs and meaning that new and existing developers can learn from previous mistakes.
-The development branches of the codes are tested nightly using a test suite of problems, checking that any new additions to the code have significantly changed any of the solutions or significantly slowed down the performance.
+Using version control allows us to keep a record of the codes' development process, helping us to track down bugs and means that new and existing developers can learn from previous mistakes.
+The development branches of the codes are tested nightly using a test suite of problems, checking that any new additions to the code have not significantly changed any of the solutions or slowed down the performance.
 
-New versions of the codes are released on the first of each month, and are archived on Zenodo\footnote{\url{https://zenodo.org/}} (which also provides us with DOIs, further enhancing the reproducibility of our results).
+New versions of the codes are released on the first of each month, and are archived on Zenodo\footnote{\url{https://zenodo.org/}}.
+This also provides us with DOIs (digital object identifiers), further enhancing the reproducibility of our results and ensuring the sustainability of the codes used to produce them.
 
@@
 
-Currently, \maestroex~is able to model spherically symmetric systems, however it has no support for rotation.
-We wish to model convection in the interiors of massive stars, and it is known that these stars often rotate and that trotation can have have significant effect on mixing at convective shell boundaries.
+\maestroex~is currently able to model spherically symmetric systems, however it has no support for rotation.
+We wish to model convection in the interiors of massive stars, and it is known that these stars often have non-negligible rotational frequencies and that this rotation can have have significant effect on mixing at convective shell boundaries.
 We are therefore currently exploring several possible ways of implementing rotation in \maestroex.
@@
 In order to capture the star's deformation without sacrificing the 1d base state, we can rewrite the base state pressure as a function of the effective potential: $p_0(r)\rightarrow p_0(\Phi_{\text{eff}})$.
+We can then rewrite the other base state quantities and the system of low Mach number equations as functions of the new coordinates $(\Phi_\text{eff}, \bm{x}, t)$.
 
-The massive stars that we are interested in modelling are believed to be relatively slow rotators, so it is likely that one of the simpler approaches outlined above should be sufficient.
-However, for modelling e.g.~type I X-ray bursts, it may be that a more complex approach is needed.
+The massive stars that we are interested in modelling are believed to be relatively slow rotators\MarginPar{insert some citation/quantify this}, so it is likely that one of the simpler approaches outlined above should be sufficient.
+However, for modelling e.g.~type I X-ray bursts, it may be that a more sophisticated approach is needed.
 Type I X-ray bursts are produced by thermonuclear deflagrations in the liquid surface layers of fast rotating neutron stars.
@@
 This produces non-negligible deformation of the star about the equator, and may cause surface flows with a significant dependence on the meriodional angle.
-To properly capture the effects of rotation in such a system, it may therefore be necessary to use a 2d base state, $p_0(r)\rightarrow p_0(r, \theta)$.
+To properly capture the effects of rotation in such a system, it may therefore be necessary to use a 2d base state, with the base state pressure a function of both the radius and the meriodional angle $p_0(r)\rightarrow p_0(r, \theta)$.
 
@@
 The boilerplate code to required to compile this for GPU is then generated by a custom preprocessor prior to compilation.
-CUDA managed memory is used to take care of data transfers between the CPU and GPU, so no explicit copies of data to/from the CPU and GPU are required.
+CUDA managed memory is used to take care of data transfers between the CPU and GPU, so no explicit copies of data to/from the CPU and GPU are required, significantly reducing the extra code required to port the code.
 One of the key emphases of the \amrex~approach to GPUs is to maintain performance portability: porting code to GPUs should not come at the expense of the code's performance on CPU-only machines.
@@
 Currently, all AMR subgrids have a copy of the entire 1d base state.
-This is not a problem for small problems where the 1d base state does not contain many cells compared to the number of cells in a single subgrid, however as the problem size and/or resolution gets larger, the size of the 1d base state grows to the point it is now significantly larger than the size of a single subgrid.
+This is not a problem for small problems where the 1d base state does not contain many cells compared to the number of cells in a single subgrid, however as the problem size and/or resolution gets larger, the size of the 1d base state grows to the point that it is now significantly larger than the size of a single subgrid.
 Compared to CPUs, GPUs typically have less memory.
 For larger problems, it is therefore not possible to store the entire base state on the GPU.
-This significantly reduces the performance of the GPU parts of the code, as it means data will be copied back and forth between the CPU and GPU (a slow process).
+This significantly reduces the performance of the GPU parts of the code, as it will increase the amount of data will be copied back and forth between the CPU and GPU (a slow process).
 To alleviate this problem, we are looking at ways to reduce the amount of 1d base state data that needs to be stored on the GPU at any one time.
-For example, in some routines we map the 1d base state to the 2d grid \textendash for these, we can move this operation to the CPU so that the GPU does not require the 1d base state.
-It may be necessary for us to develop a way to split up the 1d base state, so that subgrids only receive the portions of the base state necessary for their calculations.
+For example, in some routines we map the 1d base state to the 3d grid \textendash~for these, we can move this operation to the CPU so that the GPU does not require the 1d base state.
+It may be necessary for us to develop a way to split up the 1d base state, so that subgrids only receive the sections of the base state necessary for their calculations (rather than the base state for the entire problem domain).
 
@@
 
-\maestroex~is an open source code for modelling low Mach number astrophysical codes.
-We use the low Mach number approximation, which allows us to use a much larger timestep than conventional compressible schemes by filtering out the soundwaves.
-Unlike other sound-proof methods, we still retain background stratification and large pressure perturbations from local compositional changes and heating, both of which are particularly important for modelling atmospheres and burning.
+\maestroex~is an open source code for modelling low Mach number astrophysical flows.
+It uses the low Mach number approximation, which allows it to use a much larger timestep than conventional compressible schemes by filtering out the soundwaves.
+Unlike other sound-proof methods, we still retain background stratification and large pressure perturbations due to local compositional changes and heating, both of which are particularly important for modelling atmospheres and burning.
 
 \maestroex~is a new and improved version of our previous code \maestro.
-It is based on \amrex~(rather than \fboxlib), which means we can exploit its powerful new solvers and ongoing improvements and don't have to write so much Fortran.
-It is part of the \amrexastro~suite of open source adaptive mesh refinement astrophysical hydrodynamics codes.
+It is based on \amrex~(rather than \fboxlib), which allows us to exploit its powerful new solvers and make use of ongoing improvements.
+\maestroex~is part of the \amrexastro~suite of open source adaptive mesh refinement astrophysical hydrodynamics codes (which also includes \castro~and \nyx).
 New features developed in other codes in the family can therefore be easily implemented in \maestroex.
-As development is open, results from our simulations can be reproduced and monthly archiving to Zenodo ensures software sustainability.
+As development is open, results from our simulations can be reproduced, and monthly archiving to Zenodo ensures software sustainability.
 
@@
 For modelling the interiors of massive stars (which rotate relatively slowly), we may be able to get away with neglecting the centrifugal force and simply add the Coriolis force as a source term.
-However, this is no fun so we're also looking at doing things like modelling the centrifugal force as an effective pressure term, rewriting the base state as a function of the effective gravitational potential and using a 2d base state which is also a function of the meriodional angle.
+However, this will not be sufficient for faster rotating systems (e.g.~for modelling Type I X-ray bursts on the surfaces of millisecond pulsars), so we are also exploring other methods including modelling the centrifugal force as an effective pressure term, rewriting the base state as a function of the effective gravitational potential and using a 2d base state which is also a function of the meriodional angle.
 
-Copying the work done for \castro, we're currently in the process of porting \maestroex~to run on GPUs so that we can exploit the latest supercomputer architectures.
+We're currently in the process of porting \maestroex~to run on GPUs, using our experience doing the same for ~castro.
+This will enable us exploit the latest supercomputer architectures.
 So far, we've ported source terms and are in the process of porting the hydro.

### 2019-09-17 MAESTROeX_I/intro.tex

@@
 
+\explain{re-ordered discussion of other codes and approaches;
+kept text intact unless indicated otherwise}
 Low Mach number models \added{that do not contain acoustic waves}
@@
 not account for atmospheric expansion.
-\added{Other approaches model acoustic wave propagation with various strategies
-to efficiently and robustly handle the low Mach number limit.}
+\added{There are several other approaches that retain the effects of acoustic wave 
+propagation with various strategies to efficiently and robustly handle 
+the low Mach number limit.}
+In the reduced speed of sound technique (RSST) and related
+methods, the speed of sound is artificially reduced by including a
+suitable scaling factor in the continuity equation, reducing the
+restriction on the size of the time step
+\citep{Rempel2005,Hotta2012,Wang2015,Takeyama2017,Iijima2018}.
 There are \deleted{also} semi-implicit all-Mach
@@
 integration, while the slow material waves are solved explicitly.
-Another approach is to use preconditioned all-Mach number solvers
-\citep{Miczek2014,Barsukow2016}, where the numerical flux is
-multiplied by a preconditioning matrix.  This reduces the stiffness of
-the system at low Mach numbers, while retaining the correct scaling
-behavior \added{in an acoustic limit of the Euler equations that 
-retains $\mathcal{O}$(Ma) pressure fluctuations}.
-In the reduced speed of sound technique (RSST) and related
-methods, the speed of sound is artificially reduced by including a
-suitable scaling factor in the continuity equation, reducing the
-restriction on the size of the time step
-\citep{Rempel2005,Hotta2012,Wang2015,Takeyama2017,Iijima2018}.
-Finally, there are fully implicit time integration codes for the
+There are also fully implicit time integration codes for the
 compressible Euler equations
@@
 equations, which therefore allows for arbitrarily large time steps.
+\added{To resolve potential pressure scaling issues in very low Mach number flow
+in explicit or implicit schemes, }
+another approach is to use preconditioned all-Mach number solvers
+\citep{Miczek2014,Barsukow2016}
+\added{to properly capture pressure fluctuations in the low Mach number regime
+(see \citealt{guillard2017behaviour} for discussion).}
+\replaced{, where}{The numerical flux is
+multiplied by a preconditioning matrix}\replaced{.  This}{ that} reduces the stiffness of
+the system at low Mach numbers, while retaining the correct scaling
+behavior \added{in an acoustic limit of the Euler equations that 
+retains $\mathcal{O}$(Ma$^2$) pressure fluctuations}.
 

### 2019-09-17 MAESTROeX_I/intro.tex
fix acoustic limit statement

@@
 behavior \added{in an acoustic limit of the Euler equations that 
-retains $\mathcal{O}$(Ma$^2$) pressure fluctuations}.
+retains $\mathcal{O}$(Ma) pressure fluctuations}.
 

### 2019-09-17 MAESTROeX_I/algorithm.tex
referee comment about alternate time step restrictions

@@
 where for our simulations we typically use $\sigma^{\rm CFL}\sim 0.7$ and the minimum is taken over all spatial directions over all cells.
-There are additional constraints on the time step that are typically much less restrictive than the advective CFL including
+\replaced{There are additional constraints on the time step that are typically much less restrictive than the advective CFL including
 the acceleration due to the buoyancy force (sometimes in effect when the velocity is approximately zero at the start of some simulations) 
-and the local magnitude of the divergence constraint (to prevent too much mass evacuation from a cell in a time step); see Section 3.4 in Paper III for details.
+and the local magnitude of the divergence constraint (to prevent too much mass evacuation from a cell in a time step);}
+{For simulations that are initialized with zero velocity, we use the buoyancy force in the momentum equation to give a proper time scale
+to start the simulation;} see Section 3.4 in Paper III for details.
 

### 2019-09-17 MAESTROeX_I/appendix.tex
updated figs 4, 5, 8, 9 to have larger fonts

@@
 & \multicolumn{1}{c}{\begin{footnotesize} (a) Velocity magnitude, solved using full $\mathbf{U}$ \end{footnotesize}} \\[1.em]
-& \multicolumn{1}{c}{\includegraphics[width=2.75in]{./figs/magvel_lineout_X}} \\ 
+& \multicolumn{1}{c}{\includegraphics[width=2.5in]{./figs/magvel_lineout_X}} \\ 
 & \multicolumn{1}{c}{\begin{footnotesize} (b) Comparison along white line \end{footnotesize}} \\

### 2019-09-17 MAESTROeX_I/results.tex
enlarged font in fig 3.  and also enlarged fig 3 itself
figure size
other figure sizes

@@
 \begin{center}
-\includegraphics[width=3.0in]{./figs/MAESTRO_scaling1} \hspace{0.5em}
-\includegraphics[width=3.0in]{./figs/MAESTRO_scaling2}
+\includegraphics[width=3.25in]{./figs/MAESTRO_scaling1} \hspace{0.5em}
+\includegraphics[width=3.25in]{./figs/MAESTRO_scaling2}
 \caption{\label{fig:scaling} (Left) Weak scaling results for a spherical, full-star white dwarf calculation using the original MAESTRO code, MAESTROeX, and MAESTROeX with base state evolution disabled.
@@
 \begin{center}
-\includegraphics[width=2.75in]{./figs/wdconvect_256_maxT}  \hspace{0.5in}
-\includegraphics[width=2.75in]{./figs/wdconvect_256_maxMach}
-\caption{\label{fig:wdconvect_256_maxvar} (left) Peak temperature, $T_{\text{peak}}$, and (right) peak Mach number
+\includegraphics[width=3.25in]{./figs/wdconvect_256_maxT}  \hspace{0.5in}
+\includegraphics[width=3.25in]{./figs/wdconvect_256_maxMach}
+\caption{\label{fig:wdconvect_256_maxvar} (Left) Peak temperature, $T_{\text{peak}}$, and (Right) peak Mach number
          in a white dwarf until time of ignition at resolution of $256^3$ for three different MAESTROeX algorithms.}
@@
          (Left) Profile of $T - \bar{T}$ for a hot bubble in a white dwarf environment.
-         (Right) Region of star where $\rho\ge 10^5 \text{ g cm}^{-3}$ in a full white dwarf star simulation. }
+         (right) Region of star where $\rho\ge 10^5 \text{ g cm}^{-3}$ in a full white dwarf star simulation. }
 \end{center}
@@
 \caption{\label{fig:bubble_results} Time-lapse cross-section of a hot bubble in a white dwarf environment at
-         $t = 0$, 1.25, and 2.5 s for (left) single-level simulation, and
+         $t = 0$, 1.25, and 2.5 s for (Left) single-level simulation, and
          (right) adaptive simulation at the same effective resolution.  The red, green, and black boxes indicate grids of increasing resolution}

### 2019-09-29 astronum_2019_Castro/astronum-castro-2019.tex
start of the AstroNum 2019 paper using last year's as the template

--- 2018-12-31 astronum_2018/astronum-xrb.tex
@@
 effects if the rise of the burst is observed with enough resolution
-\cite{art-2015-cavecchi-etal}.  We note that no calculations have
-considered the effects of the flame interacting with preexisting
-turbulence ahead of the flame from simmering that may be occurring
-there, as we characterized in \cite{xrb3}.  Flame-turbulence
-interactions have long been seen as a means to accelerate a flame in
-Type Ia supernovae, so future studies should look at turbulence in the
-layer.  Capturing turbulence, of course, requires high resolution, so
-the ideas described here will be needed.
+\cite{art-2015-cavecchi-etal}.  We note that there may be preexisting
+turbulence ahead of the flame from simmering occurring
+there, and no calculations have considered the effects of the flame
+interacting with this turbulence, as we characterized in \cite{xrb3}.
+Flame-turbulence interactions have long been seen as a means to
+accelerate a flame in Type Ia supernovae, so future studies should
+look at turbulence in the layer.  Capturing turbulence, of course,
+requires high resolution, so the ideas described here will be needed.
 
@@
 is about $10\times$ faster using GPUs on a node than the entire node
-of CPU cores (comparing 4 NVIDIA Volta GPUs to 20 IBM Power9 cores).
+of CPU cores (comparing 4 NVIDIA Volta GPUs to a dual-socket IBM Power9 CPU using 40 cores).
 We see similar speed ups with reaction networks, with these GPU ports

### 2019-09-29 astronum_2019_Castro/astronum-castro-2019.tex
start afresh

--- 2018-12-31 astronum_2018/astronum-xrb.tex
@@
 
-\title{Toward Resolved Simulations of Burning Fronts in Thermonuclear
-       X-ray Bursts}
+\title{The Castro AMR Simulation Code: Current and Future Developments}
 
 \author{M. Zingale$^1$,
+        J.~B. Bell$^4$,
         K.~Eiden$^1$,
-        Y.~Cavecchi$^{2,3}$,
         A.~Harpole$^1$,
-        J.~B. Bell$^4$,
-        M.~Chang$^1$,
-        I.~Hawke$^3$,
         M.~P. Katz$^5$,
-        C.~M. Malone$^6$,
         A.~J. Nonaka$^4$,
@@
 
-\address{$^6$Los Alamos National Laboratory, Los Alamos, NM, 87545 USA}
-
 \ead{michael.zingale@stonybrook.edu}
@@
 \begin{abstract}
-We discuss the challenges of modeling X-ray bursts in
-multi-dimensions, review the different calculations done to date, and
-discuss our new set of ongoing simulations.  We also describe
-algorithmic improvements that may help in the future to offset some of
-the expense of these simulations, and describe what may be possible
-with exascale computing.
+We discuss 
 \end{abstract}
@@
 
-X-ray bursts (XRBs) are fascinating astrophysical explosions.  A
-neutron star accretes fuel from its binary companion, building up just
-a thin layer ($\sim 5$--$10$~m) before the immense gravitational
-acceleration at the neutron star surface compresses and heats the fuel
-to the point of thermonuclear runaway.  A brief flash of X-rays
-follows, the accreted layer diffuses the heat produced from reactions,
-and the process repeats.  We can observe multiple bursts from a single
-source, and the information their lightcurves encodes tells us about
-the underlying neutron star, and ultimately the nuclear equation of
-state.  A challenging aspect of interpreting the lightcurves is
-understanding the radiation transport through the hot ash layers.  In
-particular, the composition can affect the effective temperature,
-which in turn affects the radius we measure for the neutron
-star~\cite{suleimanov:2011}.  Convection in the burning layer can even
-bring ashes up to the photosphere which alters what we
-see~\cite{kajava:2017}.  Accurate measurements of neutron star radii
-from XRBs could greatly constrain the nuclear equation of
-state~\cite{steiner:2010,ozel:2010}.  Finally, brightness oscillations
-during the rising phase of the bursts
-and other features seen in lightcurves are interpreted as arising from
-the spreading of a hotspot across the neutron star---demonstrating
-that the burst begins in a localized
-region~\cite{bhattacharyya:2006,bhattacharyya:2007}.
-See~\cite{galloway:2017} for a recent review of XRBs.
-
-Simulations of XRBs can help us to understand the observations, but
-many technical challenges remain.  The primary difficulties are simply
-the range of length and timescales involved.  At the largest lengthscale,
-the neutron star radius is 10~km and the Rossby lengthscale, where rotation
-and lateral pressure gradients balance, is $\sim
-1$~km~\cite{SPIT_ETAL02}.  At the smallest scales, a conductive He
-flame has a thickness of 10s of centimeters~\cite{Timmes00} and the
-pressure scale height of the fuel layer is on the order of meters.  To capture the
-rise timescale of the XRB, we would need to model seconds, but the flame itself
-is subsonic, so long timescale evolution is difficult.
-
-Present and past simulations of XRBs employ a variety of
-approximations to get past these difficulties.  The variety of
-approximations used enables a complementary picture of XRBs to be built up,
-and has greatly advanced our understanding of these systems.
-The breadth of simulations to date has been impressive.  We summarize
-the different approaches below.
-
-One-dimensional simulations, e.g. \cite{woosley-xrb,fisker:2008}, assume
-spherical symmetry and model the burst through the vertical column
-depth of fuel.  These simulations can capture the luminosities and
-burst recurrence times well.  They can utilize very large nuclear
-reaction networks and have been used to understand the nucleosynthesis
-and rp-process~\cite{schatz:rp1999,rpprocess,schatz_rp}, as well as to understand how rate uncertainties can
-affect the outcomes of the bursts.  Many successive bursts can be
-modeled and these simulations can show us the evolution of the ash
-layer with time.  The approximation of 1D however means that we cannot
-learn about lateral variations across the neutron star, including
-flame spreading.
-
-Global shallow water simulations~\cite{SPIT_ETAL02} capture the
-large scale spreading of the burning front by using a very simple
-vertical structure.  These simulations were instrumental in showing
-that the Coriolis force plays an important role in confining a
-spreading region and that the flame spreading can explain brightness
-oscillations during the rise of the burst lightcurve.
-
-Our group and others have modeled the convection preceding the
-runaway using low Mach number hydrodynamic methods for pure helium
-bursts~\cite{Lin:2006,xrb} and mixed hydrogen/helium
-bursts~\cite{xrb2} in 2D, as well as full 3D simulations of mixed
-bursts~\cite{xrb3}, and followed the development of turbulence.
-These simulations can give an understanding of the pre-burning front
-regime, including the strength and character of the turbulence, but
-cannot model large scale lateral variations because of the implicit
-assumptions built into the low Mach number model~\cite{ABRZ:I}.
-
-While detonations are computationally easier to model than flames (we
-largely get the speed through the jump conditions in the Riemann
-problem without resolving the structure), detonations require extreme
-conditions not found in XRBs~\cite{ZINGALE_ETAL01,harpole:2018}.
-This means that calculations that want to capture the nucleosynthesis
-yields from spreading burning fronts in XRBs need to model a deflagration.
-
-Flame spreading was first modeled in a series of calculations
-employing an algorithm from atmospheric science, where vertical
-hydrostatic equilibrium is enforced and wide-aspect ratio zones give a
-horizontal CFL number that is large enough to obtain long timescale
-evolution~\cite{cavecchi:2012}.  These calculations display key
-features of the propagation mechanism, namely a balance between
-hydrodynamics and the pure flame physics. The flame proceeds mainly
-via conduction, thus being a deflagration. The flame front, however,
-is not vertical but inclined, the angle with the horizontal being very
-small: $\theta \sim H/R_{\rm{Ro}} \sim 10^{-3}$. Here $H$ is the scale
-height and $R_{\rm{Ro}}$ is the Rossby radius. It is the
-hydrodynamical balance between the Coriolis force, gravity and
-pressure gradients that determines the Rossby radius and therefore the
-inclination angle of the flame front. The extended surface of the
-flame front is what leads to propagation speeds of order $10^5 $~km/s,
-in good agreement with observations. Due to its dependence on the
-Coriolis force, the flame speed changes across the surface of the NS
-(being faster at the equator), leading to potentially detectable
-effects if the rise of the burst is observed with enough resolution
-\cite{art-2015-cavecchi-etal}.  We note that no calculations have
-considered the effects of the flame interacting with preexisting
-turbulence ahead of the flame from simmering that may be occurring
-there, as we characterized in \cite{xrb3}.  Flame-turbulence
-interactions have long been seen as a means to accelerate a flame in
-Type Ia supernovae, so future studies should look at turbulence in the
-layer.  Capturing turbulence, of course, requires high resolution, so
-the ideas described here will be needed.
-
-The NS systems exhibiting XRBs are
-expected to have magnetic fields in the range $10^7$--$10^{10}$~G.
-Inclusion of a vertical magnetic field proved to have a strong
-influence on the nature and the speed of flame propagation
-\cite{art-2016-cavecchi-etal}, due to the interaction between the
-Coriolis force and the magnetic tension, which further acts towards
-determining the inclination angle of the flame front.
-
-A longstanding goal is to perform simulations where we resolve the
-flame structure, to allow us to accurately capture the
-nucleosynthesis, and watch it move across the neutron star surface.
-The results of these simulations will enable strict comparison to
-observations made with the new generation of X-ray telescopes such as
-\textit{eXTP}\/ and \textit{STROBE-X}\/
-\cite{art-2017-wilhod-etal,art-2016-zhang-etal,intZand2018} which will offer
-high collecting area and time resolution.
-We show some in-progress calculations that are a step toward this, discuss
-their remaining approximations, and also discuss what new algorithmic
-techniques might be needed to realize this goal with the advent of
-exascale computing.
-
-\section{Resolved Flame Studies}
-
-We have begun a set of flame spreading calculations using the
-compressible hydrodynamics code \castro~\cite{castro}, part of the
-AMReX astrophysics suite~\cite{astronum:2017}.  \castro\ uses adaptive
-mesh refinement, supports a general equation of state and reaction
-networks, has a conservative gravity formulation~\cite{wdmergerI}, and
-is optimized to run on current supercomputers.  All of our code base
-is open source and available on
-GitHub\footnote{\url{http://github.com/amrex-astro/}}, and all of the
-solvers, problem setup, initial models, and analysis scripts for each
-of our science problems are distributed with the codes.
-
-We setup a problem by creating a pair of 1D hydrostatic XRB models (hot and cool)
-representing the post- and pre-flame states and apply these laterally
-across the domain to setup a hot region at the left end of our domain
-in equilibrium with the cool model on the right.  Thermal diffusion and nuclear
-reactions quickly produce a laterally propagating flame that begins to
-move through our domain.  We resolve the diffusion scale, using a
-resolution of 10~cm, and use adaptive mesh refinement (two jumps of
-$4\times$) to keep a buffer of low density material at low resolution
-above the surface to allow for expansion.  To make the lengthscales
-tractable, we rotate the neutron star at 2000~Hz---much faster than
-expected.  This helps confine the spreading region to a more
-reasonable horizontal domain.  Nevertheless, our simulation uses 12288
-zones laterally, on the finest grid.  We use the stellar
-conductivities from \cite{Timmes00} and a 13-isotopic alpha-chain
-reaction network.  To make the timescales easier, we boost the flame
-speed by a factor of 10 by scaling both the conductivity and energy
-generation rate by the same factor.  This is an approximation we hope to
-relax in the near future.
-
-Figure~\ref{fig:xrb} shows the flame structure for an initial 2D run
-at 0.007~s.  Vertical profiles at a few radial distances from the
-origin are shown for the same quantities in
-Figure~\ref{fig:xrb_profiles}.  The burning layer is only about 10~m
-high.  Underneath is \isot{Ni}{56}, representative of the underlying
-neutron star, and above it is a very low density ($10^{-4}~\gcc$)
-buffer.  Only a portion of the vertical extent is shown.  The top
-panel shows the temperature structure.  The initial perturbation
-reached to $2\times 10^4~\mathrm{cm}$, so at this point, the extent of
-the burnt region has nearly doubled.  The flame structure is quite
-wide, since there are multiple burning stages beyond just helium to
-carbon, and this shows up in the composition.  This is seen more
-clearly in the next panel which shows the mean molecular weight of the
-nuclei ($\bar{A}$).  We see that the further behind the head of the
-flame we are, the heavier the ash nuclei.  We also see the head of the
-flame appears lifted off of the base of the burning layer---unburned
-material underlies the spreading flame.  This is reminiscent of the
-flame structure seen in the earlier simulations of
-\cite{cavecchi:2012}.  The third panel shows the energy generation
-rate.  It is strongest just above the neutron star surface, behind the
-flame.  We again see that the head of the flame is detached from the
-base of the atmosphere, with the strongest burning at the head of the
-flame slightly higher up in the atmosphere.  Finally, the bottom panel
-shows the velocity through the plane of the simulation---this is
-induced by the Coriolis force as the burning fluid expands and begins
-to spread laterally.  A tight hurricane structure has setup as a
-result of the flame spreading.  This is the geostrophic balance
-discussed in \cite{SPIT_ETAL02}.  These calculations are ongoing and
-will be the subject of a detailed study in the near future.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=\linewidth]{xrb}
-\caption{\label{fig:xrb} Helium flame spreading across the surface of
-  a neutron star.  Shown are temperature (top), the mean molecular
-  weight of the ash (second), the nuclear energy generation rate
-  (third), and the velocity out of the simulation plane (bottom).  The
-  out-of-plane velocity contours indicate where the Coriolis force has
-  created a hurricane-like wind structure which confines the flame,
-  leading to an inclined front. It is at such an interface that the
-  temperature and burning rate are highest.  }
-\end{figure}
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=0.9\linewidth]{xrb_profiles}
-\caption{\label{fig:xrb_profiles} Vertical profiles at several radial distances
-  from the vertical axis through the plots shown in Figure~\ref{fig:xrb}.  These
-  profiles sample the region behind and ahead of the flame.}
-\end{figure}
-
-\section{Large Scale Simulations and Algorithmic Developments}
-
-The 2D calculations shown in the previous section require about 10k
-node-hours of computational time to get to about 10~ms (running
-on the Edison machine at NERSC).  We expect qualitatively different
-results in 3D calculations: for example, the shear from the hurricane structure
-acting along the flame front may cause instabilities that can affect the flame
-propagation.
-With our current simulation methodology,
-a 3D version of this calculation is estimated to require 10M node-hours
-or more---this is about the size of an annual allocation.
-%Additionally, this is still with the boosted flame, but we expect that once the flame is developed, the boosting will not be as critical, due to hydrodynamic effects.
-This simulation covers less than
-1\% of the neutron star surface, so a fully resolved calculation of
-burning over the entire neutron star is out of reach: a naive scaling
-calculation suggests that 15 years of Moore's Law improvements are required.
-%
-% The calculation above is
-% log_2 [ (4 pi 10^2) / (1.23^2) ] * 1.5
-% for a NS of 10km radius vs the computational domain of 1.23km^2
-%
-
-Improvements are needed to push the scale of simulations
-to the point where we can model XRBs at the full star scale with
-resolved nuclear physics.
-As the computational cost is roughly a product of the number of zones, the
-number of zone updates, and the cost per zone update, we need to look for
-improvements in each of these. The number of zones depends on the physical
-model employed, the meshing strategy, and the accuracy required. The number of
-zone updates depends on the smallest zone size and the largest characteristic
-speed in the model. The computational cost per zone update is dominated by the
-time spent in the reactions and equation of state. We discuss a few
-possible improvements that can come both from implementing new algorithms and
-from exploiting new hardware.
-
-%For example, the number of zones can be reduced
-%by symmetry considerations. We can bootstrap the 3D calculation by mapping the
-%2D axisymmetric calculation, once the flame is firmly established, into
-%3D, rotating about the symmetry axis.  This will cut down the computational
-%time by about one-half.
-
-Machine architectures are evolving, with heterogeneous architectures
-becoming more common.  Our recent development in \castro\ has focused
-on moving the hydrodynamics to GPUs~\cite{astronum:2017}.  The XRB
-problem is an ideal case for this, since hydrodynamics with constant
-gravity (for plane-parallel domains) ports in a straightforward manner
-to GPUs.  The current GPU version of the \castro\ hydrodynamics solver
-is about $10\times$ faster using GPUs on a node than the entire node
-of CPU cores (comparing 4 NVIDIA Volta GPUs to 20 IBM Power9 cores).
-We see similar speed ups with reaction networks, with these GPU ports
-under active development.  This means a GPU calculation could lower
-the above costs by a factor of 10.  GPU offloading of the reaction
-networks will also allow us to explore larger networks and more
-detailed nuclear physics.
-
-Another way we can reduce the number of zones for given accuracy is to use a more
-accurate hydrodynamic method.  We are developing fully
-fourth-order (in space and time) coupling of hydrodynamics and
-reactions using the method of spectral deferred corrections (SDC) in
-\castro, instead of the more commonly employed Strang splitting~\cite{strang:1968}.  This
-provides us with two benefits.  First, the improved coupling actually
-reduces the stiffness of the reactions, requiring fewer righthand side
-calls and therefore reducing the expense of the
-reactions.  Second, by moving to fourth
-order, we may also be able to reduce our resolution requirements
-needed for converged flames. For smooth wave solutions we may expect to need
-around half the number of zones per dimension, and since computational work
-scales like $(\Delta x)^4$ in 3D, this could lower costs by a factor of around
-10.
-%
-% This calculation is based on the number of points per wavelength required,
-% to resolve a periodic wave, which (I believe) goes from 12 (2nd order) to
-% about 7 (4th order). However, this is very idealized, and I'm not sure that
-% it carries over to the reactions, so this may well be optimistic.
-%
-
-As an example of the improvements in coupling between hydro
-and reactions, Figure~\ref{fig:sdc} shows the mass fraction of helium
-behind a detonation over the course of 2 timesteps.  For this test,
-a 19 isotope reaction network was used.  The points
-represent individual calls to the righthand side function of our
-reaction network, and the density of
-points indicates how hard the integrator is working.  We use the same
-tolerances for both the Strang-split method and the SDC method.  For
-the SDC method, we are using a second-order accurate method based on
-\cite{SDC-old}.  We predict a time-centered advection term,
-$\Adv{\Uc}^{n+1/2}_i$ using standard unsplit Godunov methods, but
-explicitly include a reactive source term in the tracing of the
-interface states, ${\bf R}(\Uc_i)$.  We can then solve the reactive
-system, using this advective term as a piecewise-constant-in-time
-source
-\begin{equation}
-\label{eq:syst}
-\frac{d{\Uc_i}}{dt} = -\Adv{\Uc}^{n+1/2}_i + {\bf R}(\Uc_i)
-\end{equation}
-This system can be integrated with standard ODE methods.  To achieve
-second-order accuracy, we need to iterate, using the updated $\Uc$ to
-create a reactive source included in the predictor for the advective
-update, and then reintegrate the system.  The result of this iteration
-is that the advection sees the effects of reactions over the timestep
-and the reactions see the effects of advection.  For the zone tracked
-in Figure~\ref{fig:sdc}, each SDC iteration called the righthand side
-function of the network 2 times less than the Strang case.  This
-savings is dependent on the network and thermodynamic state, but our
-tests thus far show that the SDC methods need fewer network calls to
-integrate a timestep.  We also clearly see how the Strang system
-evolves in a discontinuous fashion.  We are continuing to develop this
-new integration strategy, with full fourth-order in space and time
-reacting hydrodynamics methods to follow shortly.  The XRB simulations
-are our target application.  We note that there has been some
-development in the community on implicit time-integration methods for
-compressible flow, but for the XRB problem, the wide range of Mach
-number seen in the domain (exceeding unity near the top of the
-atmosphere, ahead of the flame) removes much of the advantages of
-implicit timestepping, so we focus on explicit methods.
-
-For the ODE integrations shown we use the variable-order,
-variable-step integrator \vode\ \cite{vode} with its implementation of
-the backward-differentiation formulas.
-% The integration order
-%corresponds to the order of the interpolating polynomials used to
-%predict the solution in each integrator step. 
-We note that for both
-Strang splitting and SDC, \vode\ must start the integration at first
-order and correspondingly small timesteps since no higher-order
-information about the solution is available (our reaction
-network right-hand sides provide only the first time derivative of the
-solution). Based on testing with a 13-isotope
-approximate alpha-chain network, \vode\ takes many small timesteps
-(e.g.\ $\sim 200$) before it can stably increase to a high order
-representation of the solution, but after it does so, it can often
-complete the integration in only a handful of steps (e.g.\ $\sim
-5$--$6$). Observing that during a \castro\ timestep, the first Strang
-step is a direct continuation of the integration in the second Strang
-step of the previous \castro\ timestep, we explored saving the
-\vode\ state following the second Strang step and restarting
-\vode\ during the first Strang step. This approach allowed us to start
-the first Strang step with a high order representation of the solution
-and complete it in very few \vode\ timesteps ($\lesssim 10$ steps),
-speeding up the \castro\ timestep advance by 50\% for our detonation
-test problem using the 13-isotope network. However, this came at the
-memory cost of storing $7 N_{eq} + 90$ scalar variables on our grid,
-where for our networks $N_{eq} = N_{spec} + 2$ (in addition to the
-number of species $N_{spec}$ we also integrate temperature and
-specific energy). For a large-scale science problem with a large
-network, this could quickly become prohibitive, so we are exploring
-ways to mitigate either the memory requirements for restarting the
-integration or the small timestep start up costs for the integration
-method.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.75\linewidth]{sdc_plot}
-\caption{\label{fig:sdc} Helium mass fraction as a function of time
-  over two timesteps for a detonation test problem with a 19-isotope
-  reaction network.  The points represent individual calls to the
-  righthand side function of our reaction network (see
-  Eq.~\ref{eq:syst}) and the density of points indicates how hard the
-  integrator is working.  For the Strang case, we see that the first
-  half of the reactions brings us off of the smooth solution
-  (represented by the SDC curve), then the advection overcorrects,
-  resulting in a discontinuous jump, and finally the second half of
-  the reactions works to bring the state back to the solution curve.}
-\end{figure}
-
-Other ways to make the simulation less expensive can include a subgrid
-model. Flame models are commonly used for Type Ia supernovae.  There
-however, the flame is thin compared to the scale height, so a simple
-model can represent the unresolved flame on the grid.  For the XRB,
-the flame thickness is a non-trivial fraction of the pressure scale
-height and further, the laminar flame properties will change strongly
-over the height of the burning layer (see~\cite{Timmes00}).  This
-makes traditional flame models a poor fit.  Methods that use even
-something as simple as subzone burning (e.g.~\cite{Wang2012190}) may
-help capture the energetics with lower resolution.  This can be
-implemented using the AMR machinery.  This may help a little, but will
-still be insufficient for full star models.
-
-If we continue to resolve the flame structure, then we need to be
-smarter (and more aggressive) about our grid refinement strategy.
-Currently, we put the entire fuel layer on the finest grid, but for
-the material ahead of the flame, we could keep it at lower resolution
-until the flame reaches it.  The main obstacle is that we need to keep
-the fuel ahead of the flame in hydrostatic equilibrium, and keeping it
-refined accomplishes this well.  Using well-balanced
-schemes~\cite{kappeli:2016} can help us retain hydrostatic equilibrium
-(HSE) in the lower refined regions, reducing the number of zones
-needed to maintain fidelity, but we would still need a way to regrid
-the layer ahead of the flame to higher resolution to accurately
-capture the burning dynamics.  To accomplish this, we envision
-building a refinement strategy that does interpolation of the data onto
-the newly refined grids enforcing hydrostatic equilibrium.  This can
-allow us to keep only a narrow fully refined region that ``slides''
-along the surface of the neutron star keeping pace with the advancing
-flame.  When using mesh refinement the overall costs are dominated by the
-number of zones at the finest resolution, so such a scheme would greatly reduce
-the computational costs. This is a longer term development.
-
-All of the above developments still involve fully compressible
-hydrodynamics, but multiscale methods~\cite{weinan2011principles}
-where we couple different hydrodynamics solvers together in a single
-simulation may ultimately be needed to model the full star while
-resolving the burning front.
-%\MarginPar{Alice and Ian contribute here}
-Such techniques rely on it being possible to decompose the dynamics of
-the system into two scales, e.g.~a short and a long lengthscale, and
-that the processes at each scale are at most weakly coupled with each
-other. In the context of XRBs, one possibility for multiscale modeling
-would be to model the large scale dynamics using a shallow water model
-(like that of~\cite{SPIT_ETAL02}) to capture the effects of the
-Coriolis force, and use a low Mach number or a fully compressible
-model in the vicinity of the burning front to capture the turbulent
-burning processes. Given that the computational costs of the shallow
-water model would be negligible compared to those of the low
-Mach/compressible model, the overall cost of this multiscale model
-would be determined by the size of the low Mach/compressible region
-only. This approach would therefore allow us to capture dynamics
-across the full star without sacrificing resolution around the flame.
-As with the more aggressive gridding strategy above, the hybrid
-approach reduces the number of zones with the largest computational
-cost through mesh refinement, and also reduces the computational costs
-where possible by employing cheaper physical models in regions where
-this is possible.  An example of coupling compressible and low Mach
-number models implemented using the \amrex~framework, but in a
-terrestrial context, is~\cite{Motheau2018}. An initial implementation,
-\cite{Harpole2018}, in an astrophysical context, coupling shallow
-water to compressible models in both Newtonian and relativistic
-gravity, indicates that considerable computational efficiencies are
-possible, but that close attention to the coupling between the models
-is needed.
-
-%% Finally, we can try hybrid model approaches where the fully
-%% compressible simulation can act as a subgrid model.
-%% Multiscale methods can combine different hydro models together to allow
-%% for a full star simulation with resolved burning.  Examples include...
+Castro~\cite{castro}
 
@@
 
-XRBs are multiscale, multiphysics problems that are challenging to
-model.  Nevertheless, significant progress has been made in
-multidimensional models of these events, through a set of
-complementary algorithmic approaches.  Future advances in computer
-architectures and algorithms will allow for even more realistic models
-of these events, and help us connect to observations, ultimately telling
-us about the neutron star itself.
-
-\ack The work at Stony Brook was supported by DOE/Office of Nuclear
-Physics grant DE-FG02-87ER40317 and contract 7418390 with Lawrence
-Berkeley National Laboratory as part of the Exascale Compute Project
-ExaStar collaboration.  The work at LBNL was supported by the DOE
-Office of Advanced Scientific Computing Research under Contract No,
-DE-AC02-05CH11231. YC is supported by the European Union Horizon 2020
-research and innovation programme under the Marie Sklodowska-Curie
-Global Fellowship grant agreement No 703916.  An award of computer
-time was provided by the Innovative and Novel Computational Impact on
-Theory and Experiment (INCITE) program. This research used resources
-of the Oak Ridge Leadership Computing Facility at the Oak Ridge
-National Laboratory, which is supported by the Office of Science of
-the U.S. Department of Energy under Contract No.\ DE-AC05-00OR22725.
-This research used resources of the National Energy Research
-Scientific Computing Center, which is supported by the Office of
-Science of the U.S. Department of Energy under Contract
-No.\ DE-AC02-05CH11231.  Visualizations were done using yt~\cite{yt}.
-This research has made use of NASA's Astrophysics Data System
-Bibliographic Services.
+We did stuff
 

### 2019-09-29 astronum_2019_Castro/astronum-castro-2019.tex
more work on the paper

@@
 \author{M. Zingale$^1$,
+        A.~S.~Almgren$^2$,
+        M.~Barrios Sazo$^1$,
         J.~B. Bell$^4$,
@@
         A.~Harpole$^1$,
-        M.~P. Katz$^5$,
-        A.~J. Nonaka$^4$,
-        D.~E. Willcox$^4$, and
-        W. Zhang$^4$}
+        M.~P. Katz$^3$,
+        A.~J. Nonaka$^2$,
+        D.~E. Willcox$^2$, and
+        W. Zhang$^2$}
 
@@
 
-\address{$^2$Department of Astrophysical Sciences, Princeton University,
-  Peyton Hall, Princeton, NJ 08544, USA}
-
-\address{$^3$Mathematical Sciences and STAG Research Centre,
-  University of Southampton, SO17 1BJ, UK}
-
-\address{$^4$Center for Computational Sciences and Engineering,
+\address{$^2$Center for Computational Sciences and Engineering,
   Lawrence Berkeley National Lab, Berkeley, CA 94720 USA}
 
-\address{$^5$NVIDIA Corporation, 2788 San Tomas Expressway,
+\address{$^3$NVIDIA Corporation, 2788 San Tomas Expressway,
   Santa Clara, CA, 95050 USA}
@@
 
-Castro~\cite{castro}
+The \castro\ astrophysical simulation code~\cite{castro} is designed
+for modeling problems in nuclear astrophysics, with the ability to
+accurately capture the interplay between hydrodynamics, reactions,
+gravity, and radiation in stars with complex equations of state.
+Since \castro\ was first development, there have been a number of
+enhancements to the code base, expanding its applicability to a new
+range of scientific problems.
 
-\section{Summary}
+\castro\ has been applied to models of Type Ia supernovae,
+core-collapse supernovae, pair-instability supernova, and most
+recently X-ray bursts.  A common challenge in modeling these events is
+the range of length and timescales involved.
 
-We did stuff
+To capture length scales, \castro\ uses adaptive mesh refinement,
+through the \amrex\ library~\cite{amrex_joss}.
+
+Approximations are usually made for multi-dimensional simulations.
+
+\section{Modeling Reactive Flow}
+
+The recent focus in \castro\ has been on high-fidelity simulations of
+reactive flow.  In \cite{castro:sdc} we introduced a new time
+integration strategy, spectral deferred corrections (SDC), that
+eliminates the coupling error introduced by commonly used operator
+splitting techniques (see the discussion in \cite{astronum:2018} for a
+graphical illustration of splitting error).  The SDC algorithm used in
+\castro\ follows the ideas of \cite{dutt:2000,minion:2003}, and uses
+low order explicit advection and implicit reaction updates in a
+correction equation that when applied iteratively achieves high-order
+time-accuracy.  \castro\ implements both a second-order and
+fourth-order method (in space and time) coupling reactions and
+hydrodynamics.  At the moment, the method is limited to single levels.
+
+\section{Performance Portability}
+
+The original approach to parallelism in \castro\ was MPI + OpenMP,
+with scaling on manycore architectures achieve using a tiling approach
+to OpenMP~\cite{tiling}.
+
+\section{Example Application}
+
+The SDC approach was demonstrated in \cite{castro:sdc} through a large
+suite of convergence tests and a thermonuclear flame problem.
+
+\section{Future Developments}
+
+MHD
+
+Other SDC quadrature schemes / detonations
+
+SDC multilevel + XRBs
 

### 2019-09-29 astronum_2019_Castro/astronum-castro-2019.tex
add acks

@@
 
+\ack The work at Stony Brook was supported by DOE/Office of Nuclear
+Physics grant DE-FG02-87ER40317 and contract 7418390 with Lawrence
+Berkeley National Laboratory as part of the Exascale Compute Project
+ExaStar collaboration.  This research was supported by the Exascale
+Computing Project (17-SC-20-SC), a collaborative effort of the
+U.S. Department of Energy Office of Science and the National Nuclear
+Security Administration.  The work at LBNL was supported by the DOE
+Office of Advanced Scientific Computing Research under Contract No,
+DE-AC02-05CH11231.  An award of computer time was provided by the
+Innovative and Novel Computational Impact on Theory and Experiment
+(INCITE) program. This research used resources of the Oak Ridge
+Leadership Computing Facility at the Oak Ridge National Laboratory,
+which is supported by the Office of Science of the U.S. Department of
+Energy under Contract No.\ DE-AC05-00OR22725.  This research used
+resources of the National Energy Research Scientific Computing Center,
+which is supported by the Office of Science of the U.S. Department of
+Energy under Contract No.\ DE-AC02-05CH11231.  Visualizations were
+done using yt~\cite{yt}.  This research has made use of NASA's
+Astrophysics Data System Bibliographic Services.
+
 %\section*{References}

### 2019-09-29 astronum_2019_Castro/astronum-castro-2019.tex
updates

@@
 \newcommand{\isot}[2]{$^{#2}\mathrm{#1}$}
+\newcommand{\avg}[1]{{\left \langle #1 \right \rangle}}
+\newcommand{\Rbs}[1]{{\bf R} \left ( #1 \right )}
+\newcommand{\Ic}{\bm{\mathcal{I}}}
 
@@
 correction equation that when applied iteratively achieves high-order
-time-accuracy.  \castro\ implements both a second-order and
-fourth-order method (in space and time) coupling reactions and
-hydrodynamics.  At the moment, the method is limited to single levels.
+time-accuracy.
+
+In an operator split implementation, reaction and burning proceed
+without knowing about the other process.  In contrast, the update in
+the SDC formulation explicitly couples the processes together.  In \castro\ we consider
+reactions and hydrodynamics, and write our conservative system as:
+\begin{equation}
+\Uc_t = \Advs{\Uc} + \Rbs{\Uc}
+\end{equation}
+where $\Uc$ is the vector of conserved quantities, $\Advs{\Uc}$ is the
+advective term (the divergence of the hydrodynamic fluxes along with
+hydrodynamic sources), and $\Rbs{\Uc}$ are the reactive source terms.
+The update is done at specific temporal nodes, the number of location
+of which are picked to give the desired temporal accuracy.  The update
+from one time node $m$ to $m+1$ appears as:
+    \begin{align}
+      \label{eq:sdc:general}
+      \avg{\Uc}^{{m+1},(k+1)} = \avg{\Uc}^{m,(k+1)}
+            &+ \delta t_m \left [ \avg{\Advs{\Uc}}^{m,(k+1)} - \avg{\Advs{\Uc}}^{m,(k)} \right ] \nonumber \\
+            &+ \delta t_m \left [ \avg{\Rbs{\Uc}}^{{m+1},(k+1)} - \avg{\Rbs{\Uc}}^{{m+1},(k)} \right ] \nonumber \\
+            &+ \int_{t^m}^{t^{m+1}} dt  \left (\avg{\Advs{\Uc}}^{(k)} + \avg{\Rbs{\Uc}}^{(k)}\right)
+   \end{align}
+We use a finite-volume formalism, so $\avg{\Uc}$ is the average of the
+conserved state in a zone, $\avg{\Advs{\Uc}}^m$ is the average
+advective term in a zone at time $t^m$, and $\avg{\Rbs{\Uc}}^{k+1}$ is
+the average of the reactive source at time $t^{m+1}$.  The second
+superscript on each term, $(k)$ or $(k+1)$, is the iteration.  This
+update is an implicit equation for the new state,
+$\avg{\Uc}^{{m+1},(k+1)}$.  The last term in the update is an integral
+over the sources constructed from the previous iteration's values at
+each time node.  Each iteration of the SDC method increases the
+temporal order of accuracy by one, up to the order of accuracy with
+which the integral is constructed.  \castro\ implements both a
+second-order method (using a trapezoid rule for the integral) and
+fourth-order method in space and time (using a Simpsons rule for the
+integral and the spatial reconstruction of
+\cite{mccorquodalecolella}).  At the moment, the method is limited to
+single levels.
 
@@
 with scaling on manycore architectures achieve using a tiling approach
-to OpenMP~\cite{tiling}.
+to OpenMP~\cite{tiling}.  More recently we've ported \castro\ to GPUs
+using CUDA, using the same computational kernels as the MPI + OpenMP
+version.  This reuse keeps the code base managable---we don't need
+separate kernels for each architecture---while allowing us to take
+advantage of current and next generation supercomputers.  All of the
+solvers needed to run our core science problems run on GPUs: the main
+unsplit PPM hydrodynamics scheme~\cite{ppm,millercolella:2002}, self-gravity via multigrid with
+isolated boundary conditions, thermal diffusion, and nuclear
+reactions.  Our approach is to put the data on GPUs and then run all
+of the kernels on the GPUs, minimizing data movement.  This has given
+us enormous speed-ups.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.8\linewidth]{wdmerger_gpu}
+\caption{\label{fig:gpu} blah}
+\end{figure}
 

### 2019-09-29 astronum_2019_Castro/astronum-castro-2019.tex
more work

@@
 \begin{abstract}
-We discuss 
+We describe the \castro\ astrophysics simulation code, focusing on the
+recent developments in coupling hydrodynamics and reactions and describing
+upcoming developments.
 \end{abstract}
@@
 
+We've run XRB on GPUs and with SDC
+
 \section{Future Developments}
 
-MHD
+The development of SDC in \castro\ continues, with the current focus
+on modeling detonations.  Reactions in astrophysical detonations are
+especially stiff and challenging to model when the ash state reaches
+nuclear statistical equilibrium.  We are exploring different
+quadrature schemes for the integral in Equation~\ref{eq:sdc} to model
+detonations in the SDC framework.  We are also working on extending
+the SDC integration methodology to adaptive mesh refinement with
+subcycling in time.
 
-Other SDC quadrature schemes / detonations
+To complement the existing suite of hydrodynamics solver in \castro,
+an MHD solver is under development and expected to be merged into the
+main branch soon.  We will use the experiences learned with the
+\castro\ hydrodynamics solver to port this solver to GPUs and fit into
+the SDC framework.
 
-SDC multilevel + XRBs
+All of these developments fit into our goal of modeling XRBs.
 

### 2019-10-01 MAESTROeX_I/intro.tex
fix ref
minor

@@
 of these flows, modeling long-time convective dynamics are of
-interest, and numerical approaches based on compressible hydrodynamics
+interest, and numerical approaches based on \added{explicit} compressible hydrodynamics
 are intractable, even on modern supercomputers.  \replaced{One approach to this
@@
 another approach is to use preconditioned all-Mach number solvers
-\citep{Miczek2014,Barsukow2016}
+\citep{Miczek2015,Barsukow2016}
 \added{to properly capture pressure fluctuations in the low Mach number regime

### 2019-10-01 astronum_2019_Castro/astronum-castro-2019.tex
more progress

@@
 gravity, and radiation in stars with complex equations of state.
-Since \castro\ was first development, there have been a number of
+Since \castro\ was first developed, there have been a number of
 enhancements to the code base, expanding its applicability to a new
-range of scientific problems.
+range of scientific problems.  The most active development presently
+focuses on new methods of time integration, with better coupling of
+physical processes, and GPU performance.  Here we discuss some of
+these features.
 
 \castro\ has been applied to models of Type Ia supernovae,
-core-collapse supernovae, pair-instability supernova, and most
-recently X-ray bursts.  A common challenge in modeling these events is
-the range of length and timescales involved.
+core-collapse supernovae, pair-instability supernovae, exoplanet
+atmospheres, and most recently X-ray bursts (see, e.g.,
+\cite{castro-ccsne,castro-pairinstability,polin:2019,wdmergerI} for
+some example science applications).  A common challenge in modeling
+these events is the range of length and timescales involved.
 

### 2019-10-01 astronum_2019_Castro/astronum-castro-2019.tex

@@
 
-\newcommand{\maestro}{{\sffamily Maestro}}
+\newcommand{\maestroex}{{\sffamily MAESTROeX}}
 \newcommand{\castro}{{\sffamily Castro}}
@@
 some example science applications).  A common challenge in modeling
-these events is the range of length and timescales involved.
+these events is the range of length and timescales involved.  To
+capture length scales, \castro\ uses adaptive mesh refinement, through
+the \amrex\ library~\cite{amrex_joss}.  This allows us to focus the
+computational effort of regions where burning or the flow is
+important.  To address the range of timescales involved in
+astrophysical explosions, we have developed a low Mach number
+hydrodynamics code, \maestroex~\cite{maestroex}, built on the same
+framework as \castro, that can model the subsonic convection that
+often precedes explosive events.  Both codes are open source and
+freely available on
+github\footnote{\url{https://github.com/AMReX-Astro/}}.
 
-To capture length scales, \castro\ uses adaptive mesh refinement,
-through the \amrex\ library~\cite{amrex_joss}.
-
-Approximations are usually made for multi-dimensional simulations.
+In the next sections, we discuss \castro's new approach to coupling
+hydrodynamics and reactions and our efforts on performance portavbility.
+We illustrate these new developments with our X-ray burst problem.
 
@@
 
+In modeling stellar explosions, it is often the case that the hydrodynamics method
+and the nuclear reaction network require different timesteps in order to produce stable
+and accurate models.  For the stiff nuclear reactions we encounter in stars, implicit ODE
+method are often used together with operator splitting to evolve the nuclear reactions separately
+from the hydrodynamics.  In the limit of small timesteps, the two processes are well-coupled together, but 
 The recent focus in \castro\ has been on high-fidelity simulations of

### 2019-10-01 astronum_2019_Castro/astronum-castro-2019.tex
more work

@@
 
-In modeling stellar explosions, it is often the case that the hydrodynamics method
-and the nuclear reaction network require different timesteps in order to produce stable
-and accurate models.  For the stiff nuclear reactions we encounter in stars, implicit ODE
-method are often used together with operator splitting to evolve the nuclear reactions separately
-from the hydrodynamics.  In the limit of small timesteps, the two processes are well-coupled together, but 
-The recent focus in \castro\ has been on high-fidelity simulations of
-reactive flow.  In \cite{castro:sdc} we introduced a new time
-integration strategy, spectral deferred corrections (SDC), that
-eliminates the coupling error introduced by commonly used operator
-splitting techniques (see the discussion in \cite{astronum:2018} for a
-graphical illustration of splitting error).  The SDC algorithm used in
-\castro\ follows the ideas of \cite{dutt:2000,minion:2003}, and uses
-low order explicit advection and implicit reaction updates in a
-correction equation that when applied iteratively achieves high-order
-time-accuracy.
+In modeling stellar explosions, it is often the case that the
+hydrodynamics method and the nuclear reaction network require
+different timesteps in order to produce stable and accurate models.
+For the stiff nuclear reactions we encounter in stars, implicit ODE
+method are often used together with operator splitting to evolve the
+nuclear reactions separately from the hydrodynamics.  In the limit of
+small timesteps, the two processes are well-coupled together, but
+these conditions are not always met in simulations.  The recent focus
+in \castro\ has been on high-fidelity simulations of reactive flow.
+In \cite{castro:sdc} we introduced a new time integration strategy,
+spectral deferred corrections (SDC), that eliminates the coupling
+error introduced by commonly used operator splitting techniques (see
+the discussion in \cite{astronum:2018} for a graphical illustration of
+splitting error).
 
-In an operator split implementation, reaction and burning proceed
-without knowing about the other process.  In contrast, the update in
-the SDC formulation explicitly couples the processes together.  In \castro\ we consider
-reactions and hydrodynamics, and write our conservative system as:
+The SDC algorithm used in \castro\ follows the ideas of
+\cite{dutt:2000,minion:2003}, and uses low order explicit advection
+and implicit reaction updates in a correction equation that when
+applied iteratively achieves high-order time-accuracy.  In an operator
+split implementation, reaction and burning proceed without knowing
+about the other process.  In contrast, the update in the SDC
+formulation explicitly couples the processes together.  In \castro\ we
+consider reactions and hydrodynamics, and write our conservative
+system as:
 \begin{equation}
@@
 temporal order of accuracy by one, up to the order of accuracy with
-which the integral is constructed.  \castro\ implements both a
-second-order method (using a trapezoid rule for the integral) and
-fourth-order method in space and time (using a Simpsons rule for the
-integral and the spatial reconstruction of
-\cite{mccorquodalecolella}).  At the moment, the method is limited to
-single levels.
+which the integral is constructed.
+
+\castro\ implements both a second-order method (using a trapezoid rule
+for the integral) and fourth-order method in space and time (using a
+Simpsons rule for the integral and the spatial reconstruction of
+\cite{mccorquodalecolella}).  We demonstrated that \castro\ achieves
+the expected convergence on a wide variety of problems in
+\cite{castro:sdc}.  At the moment, the method is limited to single
+levels, but work is underway to extend this methodology to AMR.  For
+problems where burning is imprtant and can dominate the computational
+expense, we expect these new SDC method will become the preferred
+integration technique in the future.
 

### 2019-10-01 astronum_2019_Castro/astronum-castro-2019.tex
some more work

@@
 with scaling on manycore architectures achieve using a tiling approach
-to OpenMP~\cite{tiling}.  More recently we've ported \castro\ to GPUs
-using CUDA, using the same computational kernels as the MPI + OpenMP
-version.  This reuse keeps the code base managable---we don't need
-separate kernels for each architecture---while allowing us to take
-advantage of current and next generation supercomputers.  All of the
-solvers needed to run our core science problems run on GPUs: the main
-unsplit PPM hydrodynamics scheme~\cite{ppm,millercolella:2002}, self-gravity via multigrid with
+to OpenMP~\cite{tiling,astronum:2017}.  The general approach is to
+have MPI distribute the AMR grids across nodes and have OpenMP threads
+work on regions of each grid (tiles) by passing the box describing the
+tile into the computational kernel.  More recently we've ported
+\castro\ to GPUs using CUDA, using the same computational kernels as
+the MPI + OpenMP version.  When run on GPUs, each CUDA thread handles
+the update of a single zone, simply by passing that zone index into
+the computational kernel.  This reuse keeps the code base
+managable---we don't need separate kernels for each
+architecture---while allowing us to take advantage of current and next
+generation supercomputers.  All of the solvers needed to run our core
+science problems run on GPUs: the main unsplit PPM hydrodynamics
+scheme~\cite{ppm,millercolella:2002}, self-gravity via multigrid with
 isolated boundary conditions, thermal diffusion, and nuclear
@@
 of the kernels on the GPUs, minimizing data movement.  This has given
-us enormous speed-ups.
+us enormous speed-ups.  Figure~ref{fig:gpu} shows weak scaling on the
+OLCF Summit machine for a WD merger problem~\cite{wdmergerI}.  When
+using CPUs, we use 42 IBM Power 9 cores per node, while when using
+GPUs we use 6 Nvidia Volta GPUs per node.  We see that the code
+performance is about $10\times$ higher per node when using GPUs, while
+both show excellent weak scaling.
 
@@
 \includegraphics[width=0.8\linewidth]{wdmerger_gpu}
-\caption{\label{fig:gpu} blah}
+\caption{\label{fig:gpu} \castro\ weak scaling on the OLCF Summit
+  machine for the white dwarf merger problem.  On each node we use
+  either 42 IBM Power 9 CPU cores or 6 Nvidia Volta GPUs.  Ideal
+  scaling is shown for both cases as the solid line.}
 \end{figure}

### 2019-10-01 astronum_2019_MAESTROeX/astronum-maestroex.tex
add gpu performance plot

@@
 
-Insert some performance plots here (do these at the hackathon?)
+\begin{figure}\label{fig:gpu_speedup}
+    \centering
+
+    \includegraphics[width=\textwidth]{plots/gpu_speedup}
+
+    \caption{Comparison of the performance of functions on the CPU vs the GPU.
+The plot shows the average execution time for several functions, recorded for a run of the 3d \texttt{reacting\_bubble} problem on Summit.
+Both tests were performed using a single node, with the CPU test using 42 MPI processes, each with 4 OpenMP threads, and the GPU test using 6 MPI processes, each process associated with a single GPU.
+Note that the $x$-axis scale is logarithmic.}
+
+\end{figure}
 
@@
 
-We're currently in the process of porting \maestroex~to run on GPUs, using our experience doing the same for ~castro.
+We're currently in the process of porting \maestroex~to run on GPUs, using our experience doing the same for \castro.
 This will enable us exploit the latest supercomputer architectures.

### 2019-10-01 astronum_2019_MAESTROeX/astronum-maestroex.tex
minor comments/changes

@@
 This states that for a grid with cell spacing $\Delta x$, the timestep $\Delta t <  \Delta x / \max(|\Ub|+c_s)$.
-For low Mach number flows, this condition is dominated by the contributions of the sound speed, with the result that many fine timesteps are required to achieve high spatial resolution.
+For low Mach number flows, this condition is dominated by the contributions of the sound speed, with the result that many fine timesteps are required when using high spatial resolution.
 Long timescale, high resolution simulations therefore become extremely computationally expensive.
@@
 The technique that we use in \maestroex~is to modify the fluid equations themselves so as to filter out the soundwaves.
-This is a similar approach to the incompressible \cite{Boussinesq1901} and anelastic \cite{Ogura1962a,Gough1968,Durran1989} approximations, however the approximation that we use (variously called the low Mach number approximation \cite{Day2000,Almgren2006a,Nonaka2010} and the generalized pseudo-incompressible approximation \cite{Vasil2013}) allows for background stratification and large pressure perturbations due to heating and changes in composition.
-This is achieved by decomposing the pressure into a 1d hydrostatic base state, $p_0 = p_0(r, t)$, and a dynamic pressure perturbation, $\pi = \pi(r, \bm{x}, t)$, such that the full pressure is given by $p(r, \bm{x}, t) = p_0(r, t) + \pi(r, \bm{x}, t)$.
-In order for the approximation to hold, it is necessary that $|\pi|/p_0 = O(\mathrm{Ma}^2)$.
+This is a similar approach to the incompressible \cite{Boussinesq1901} and anelastic \cite{Ogura1962a,Gough1968,Durran1989} approximations, however the approximation that we use (called the low Mach number approximation \cite{Day2000,Almgren2006a,Nonaka2010} and the generalized pseudo-incompressible approximation \cite{Vasil2013}) allows for background stratification and large density and temperature perturbations due to heating and changes in composition.
+This is achieved by decomposing the pressure into a 1d hydrostatic base state, $p_0 = p_0(r, t)$, and a dynamic pressure perturbation, $\pi = \pi(r, \bm{x}, t)$, such that the full pressure is given by $p(r, \bm{x}, t) = p_0(r, t) + \pi(\bm{x}, t)$.
+In the low Mach number regime, asymptotic analysis shows that $|\pi|/p_0 = O(\mathrm{Ma}^2)$.
 
@@
 The temporal integration method has been greatly simplified without compromising the second order accuracy, and a new spherical base state mapping has been implemented in order to reduce mapping errors between spherical and Cartesian grids.
-Utilising the \amrex~framework has allowed us to implement MPI+OpenMP parallelism, which has been shown to scale well to over 10,000 MPI processes.
+Utilising the \amrex~framework has allowed us to implement MPI+OpenMP (with tiling \cite{Zhang2016}) parallelism, which has been shown to scale well to over 10,000 MPI processes.
 
@@
 Developers of the different codes work closely together (in fact many of the developers work on more than one of the codes), with their shared structure and framework meaning that new features developed in one code can easily be replicated in the others.
-Examples of this include the SDC solvers \cite{Zingale2019} and GPU port that have recently been implemented in \castro, and are currently in the process of being implemented in \maestroex.
+A prime example of this is the GPU port capability that has recently been implemented in \castro, and are currently in the process of being implemented in \maestroex.
+\MarginPar{mention starkiller astrophysics}
 
@@
 This was done in \maestro, and should be a reasonable approximation for slowly rotating systems.
+\MarginPar{In Zingale 2011, equation (2) has the Coriolis and centrifugal terms.  Include that equation here?}
 
@@
 
-We have begun porting \maestroex~to GPUs, leveraging the GPU capabilities built into \amrex~and following the lessons learnt porting \castro~to GPUs.
+We have begun porting \maestroex~to GPUs, leveraging the GPU capabilities built into \amrex~and following the lessons learned porting \castro~to GPUs.
 In the \amrex~framework, routines can be ported to GPUs by simply inserting a few macros into the code.

### 2019-10-02 astronum_2019_Castro/astronum-castro-2019.tex
mor eedits

@@
 
-\section{Example Application}
+\section{Example Application: X-ray bursts}
 
-The SDC approach was demonstrated in \cite{castro:sdc} through a large
-suite of convergence tests and a thermonuclear flame problem.
+Both the new SDC method and GPU offloading are important to our
+science goals, in particular the problem of modeling a flame spreading
+across the neutron star during an X-ray burst (XRB).  We discussed the
+computational challenges of modeling these events in
+\cite{astronum:2018}.  In short: the largest scale we need to model is
+the size of the neutron star, or atleast several times the scale at
+which rotation and the lateral flame spreading balance (the Rossby
+length).  The smallest scale we need to model is dictated by
+accurately capturing the burning---we are interested in both the
+energy release and nucleosynthesis.  Our first set of two-dimensional simulations
+(Eiden et al.\ in preparation) used two physical approximations to
+make the problem more tractable.  First, we used a higher rotation
+rate than expected in order to reduce the Rossby length, allowing us
+to model a smaller region of that star.  Second, we artificially
+boosted the speed of the flame to reduce the duration we need to
+model.  Our next set of calculations will relax these approximations.
 
-We've run XRB on GPUs and with SDC
+With the new fourth-order accurate reactive hydro solver, we believe
+we can capture the dynamics of the spreading flame accurately with one
+less refinement level.  This will allow us to expand to larger domains
+while keeping the memory demands reasonable.  Figure~XX shows a
+simulation of the flame spreading done with the second-order SDC
+solver (the 4th order solver does not yet work with axisymmetry),
+demonstrating that the SDC framework can evolve this system.
+
+The large increase in performance of the code when run on GPUs enables
+us to do away with the flame boosting.  Preliminary estimates show
 

### 2019-10-02 astronum_2019_MAESTROeX/astronum-maestroex.tex
updated GPU plot to include speedups

--- 2019-10-01 astronum_2019_MAESTROeX/astronum-maestroex.tex
@@
 
-\begin{figure}\label{fig:gpu_speedup}
+\begin{figure}
     \centering
@@
 Both tests were performed using a single node, with the CPU test using 42 MPI processes, each with 4 OpenMP threads, and the GPU test using 6 MPI processes, each process associated with a single GPU.
-Note that the $x$-axis scale is logarithmic.}
+Note that the $x$-axis scale is logarithmic.}\label{fig:gpu_speedup}
 
@@
 
-\ack Visualizations were probably done using~\cite{yt}.
+\ack Figure~\ref{fig:gpu_speedup} was generated using \texttt{Jupyter} \citep{Kluyver2016} and \texttt{matplotlib} \citep{Hunter2007}.
 
@@
 % Science of the U.S. Department of Energy under Contract
-% No.\ DE-AC02-05CH11231.  Visualizations were done using yt~\cite{yt}.
-% This research has made use of NASA's Astrophysics Data System
-% Bibliographic Services.
+% No.\ DE-AC02-05CH11231.
 

### 2019-10-02 astronum_2019_MAESTROeX/astronum-maestroex.tex
updated abstract; some other minor edits

--- 2019-10-01 astronum_2019_MAESTROeX/astronum-maestroex.tex
@@
 \begin{abstract}
-Convective flows in the interiors of stars are challenging to model using conventional compressible hydrodynamics codes due to their low Mach number, which severely limits the timestep that can be taken in simulations.
+Modelling long-time convective flows in the interiors of stars is extremeley challenging using conventional compressible hydrodynamics codes due to the acoustic timestep limitation.
+Many of these flows are in the low Mach number regime, which allows us to exploit the relationship between acoustic and advective time scales to develop a more computationally efficient approach.
 \maestroex\ is an open source low Mach number stellar hydrodynamics code that allows much larger timesteps to be taken, therefore enabling systems to be modelled for much longer periods of time.
 This is particularly important for the problem of convection in the cores of rotating massive stars prior to core collapse.
-To fully capture the dynamics, it is necessary to model these systems in 3d at high resolution over many rotational periods, but this has been too computationally expensive for existing, conventional schemes.
-
-We shall present an overview of \maestroex's current capabilities, describe ongoing work to incorporate the effects of rotation and discuss how we are optimizing the code to run on GPUs. 
+To fully capture the dynamics, it is necessary to model these systems in three dimensions at high resolution over many rotational periods.
+We present an overview of \maestroex's current capabilities, describe ongoing work to incorporate the effects of rotation and discuss how we are optimizing the code to run on GPUs.
 \end{abstract}
@@
 This produces non-negligible deformation of the star about the equator, and may cause surface flows with a significant dependence on the meriodional angle.
-To properly capture the effects of rotation in such a system, it may therefore be necessary to use a 2d base state, with the base state pressure a function of both the radius and the meriodional angle $p_0(r)\rightarrow p_0(r, \theta)$.
+To properly capture the effects of rotation in such a system, it may therefore be necessary to use a two-dimensional base state, with the base state pressure a function of both the radius and the meriodional angle $p_0(r)\rightarrow p_0(r, \theta)$.
 
@@
 % \item Rewrite base state in terms of equipotentials $$p_0(r) \rightarrow p_0(\Phi_{\text{eff}}) $$
-% \item 2d base state $$p_0(r)  \rightarrow p_0(r, \theta)$$
+% \item two-dimensional base state $$p_0(r)  \rightarrow p_0(r, \theta)$$
 % \end{itemize}
@@
 The latest generation of supercomputers to come online (e.g.~Summit, soon Perlmutter and Frontier) rely heavily on GPUs in order to achieve high performance while keeping energy consumption to a minimum (GPUs consume considerably less power per flop than CPUs).
-This trend is set to continue: supercomputers are moving to new architectures in order to increase their performance, rather than simply adding more and more CPUs.
+This trend is set to continue: supercomputers are moving to new architectures in order to increase performance rather than simply adding more and more CPUs.
 In order for HPC codes to be able to exploit these latest machines, it is therefore becoming increasingly necessary to port codes to run on GPUs.
@@
 By changing the compiler flags, the same bit of code can be compiled for serial, MPI-only, OpenMP-only, MPI+OpenMP, GPU-only or MPI+GPU.
-In order to achieve the best performance on the GPU, some additional optimization is required.
-However, it has been found that these optimizations often improve the performance when the code is run on the CPU only as well.
+In order to achieve the best performance on the GPU, we require some additional optimization of the computational kernals.
+However, we find that these optimizations often improve the performance when the code is run on the CPU only as well.
 More details to appear in a forthcoming \amrexastro~GPU paper.
@@
 So far we have ported most of the source terms and hydrodynamics in \maestroex~to run on GPUs.
-We have found that one of the challenges moving forwards is dealing with the 1d base state.
-Currently, all AMR subgrids have a copy of the entire 1d base state.
-This is not a problem for small problems where the 1d base state does not contain many cells compared to the number of cells in a single subgrid, however as the problem size and/or resolution gets larger, the size of the 1d base state grows to the point that it is now significantly larger than the size of a single subgrid.
+GPU support for the linear solvers is becoming increasingly available from the \amrex~team.
+We have found that one of the challenges moving forwards is dealing with the one-dimensional base state.
+Currently, each MPI process holds a copy of the entire base state.
+This is not a problem for small problems where the base state does not contain many cells compared to the number of cells in a single subgrid, however as the problem size and/or resolution gets larger, the size of the base state grows to the point that it is now significantly larger than the size of a single subgrid.
 Compared to CPUs, GPUs typically have less memory.
@@
 This significantly reduces the performance of the GPU parts of the code, as it will increase the amount of data will be copied back and forth between the CPU and GPU (a slow process).
-To alleviate this problem, we are looking at ways to reduce the amount of 1d base state data that needs to be stored on the GPU at any one time.
-For example, in some routines we map the 1d base state to the 3d grid \textendash~for these, we can move this operation to the CPU so that the GPU does not require the 1d base state.
-It may be necessary for us to develop a way to split up the 1d base state, so that subgrids only receive the sections of the base state necessary for their calculations (rather than the base state for the entire problem domain).
+To alleviate this problem, we are looking at ways to reduce the amount of base state data that needs to be stored on the GPU at any one time.
+For example, in some routines we map the base state to the three-dimensional grid \textendash~for these, we can move this operation to the CPU so that the GPU does not require the base state.
+It may be necessary for us to develop a way to split up the base state, so that subgrids only receive the sections of the base state necessary for their calculations (rather than the base state for the entire problem domain).
 
@@
     \caption{Comparison of the performance of functions on the CPU vs the GPU.
-The plot shows the average execution time for several functions, recorded for a run of the 3d \texttt{reacting\_bubble} problem on Summit.
+The plot shows the average execution time for several functions, recorded for a run of the three-dimensional \texttt{reacting\_bubble} problem on Summit.
 Both tests were performed using a single node, with the CPU test using 42 MPI processes, each with 4 OpenMP threads, and the GPU test using 6 MPI processes, each process associated with a single GPU.
@@
 For modelling the interiors of massive stars (which rotate relatively slowly), we may be able to get away with neglecting the centrifugal force and simply add the Coriolis force as a source term.
-However, this will not be sufficient for faster rotating systems (e.g.~for modelling Type I X-ray bursts on the surfaces of millisecond pulsars), so we are also exploring other methods including modelling the centrifugal force as an effective pressure term, rewriting the base state as a function of the effective gravitational potential and using a 2d base state which is also a function of the meriodional angle.
+However, this will not be sufficient for faster rotating systems (e.g.~for modelling Type I X-ray bursts on the surfaces of millisecond pulsars), so we are also exploring other methods including modelling the centrifugal force as an effective pressure term, rewriting the base state as a function of the effective gravitational potential and using a two-dimensional base state which is also a function of the meriodional angle.
 

### 2019-10-02 astronum_2019_MAESTROeX/astronum-maestroex.tex

@@
 
-\begin{figure}\label{fig:gpu_speedup}
+\begin{figure}
     \centering
-
     \includegraphics[width=\textwidth]{plots/gpu_speedup}
-
     \caption{Comparison of the performance of functions on the CPU vs the GPU.
-The plot shows the average execution time for several functions, recorded for a run of the three-dimensional \texttt{reacting\_bubble} problem on Summit.
+The plot shows the average execution time for several functions, recorded for a run of the 3d \texttt{reacting\_bubble} problem on Summit.
 Both tests were performed using a single node, with the CPU test using 42 MPI processes, each with 4 OpenMP threads, and the GPU test using 6 MPI processes, each process associated with a single GPU.
-Note that the $x$-axis scale is logarithmic.}
-
+Note that the $x$-axis scale is logarithmic.}\label{fig:gpu_speedup}
 \end{figure}

### 2019-10-02 astronum_2019_Castro/astronum-castro-2019.tex
finish a first draft

@@
 \begin{abstract}
-We describe the \castro\ astrophysics simulation code, focusing on the
-recent developments in coupling hydrodynamics and reactions and describing
-upcoming developments.
+We describe recent developments to the \castro\ astrophysics
+simulation code, focusing on new features that enable our simulations
+of X-ray bursts.  Two highlights of \castro's ongoing development are
+the new integration technique to couple hydrodynamics and reactions to
+high order and GPU offloading.  We discuss how these features will
+help offset some of the computational expense in X-ray burst models.
 \end{abstract}
@@
 enhancements to the code base, expanding its applicability to a new
-range of scientific problems.  The most active development presently
-focuses on new methods of time integration, with better coupling of
-physical processes, and GPU performance.  Here we discuss some of
-these features.
+range of scientific problems.  Throughout this development,
+\castro\ has strived to perform well on modern supercomputers,
+adapting to trends in hardware and programming models, while
+maintaining portability across architectures.
 
@@
 these events is the range of length and timescales involved.  To
-capture length scales, \castro\ uses adaptive mesh refinement, through
+capture length scales, \castro\ uses adaptive mesh refinement (AMR), through
 the \amrex\ library~\cite{amrex_joss}.  This allows us to focus the
@@
 
-In the next sections, we discuss \castro's new approach to coupling
-hydrodynamics and reactions and our efforts on performance portavbility.
-We illustrate these new developments with our X-ray burst problem.
+The most active development presently focuses on new methods of time
+integration, with better coupling of physical processes, and GPU
+performance.  In these next sections we discuss these features and
+their impact on our ongoing X-ray burst simulations.
 
@@
 For the stiff nuclear reactions we encounter in stars, implicit ODE
-method are often used together with operator splitting to evolve the
+methods are often used together with operator splitting to evolve the
 nuclear reactions separately from the hydrodynamics.  In the limit of
@@
    \end{align}
-We use a finite-volume formalism, so $\avg{\Uc}$ is the average of the
+We use a finite-volume formalism, so $\avg{\Uc}$ is the spatial average of the
 conserved state in a zone, $\avg{\Advs{\Uc}}^m$ is the average
@@
 the average of the reactive source at time $t^{m+1}$.  The second
-superscript on each term, $(k)$ or $(k+1)$, is the iteration.  This
+superscript on each term, $(k)$ or $(k+1)$, is the iteration counter.  This
 update is an implicit equation for the new state,
@@
 levels, but work is underway to extend this methodology to AMR.  For
-problems where burning is imprtant and can dominate the computational
+problems where burning is important and can dominate the computational
 expense, we expect these new SDC method will become the preferred
@@
 of the kernels on the GPUs, minimizing data movement.  This has given
-us enormous speed-ups.  Figure~ref{fig:gpu} shows weak scaling on the
+us enormous speed-ups.  Figure~\ref{fig:gpu} shows weak scaling on the
 OLCF Summit machine for a WD merger problem~\cite{wdmergerI}.  When
@@
 \cite{astronum:2018}.  In short: the largest scale we need to model is
-the size of the neutron star, or atleast several times the scale at
-which rotation and the lateral flame spreading balance (the Rossby
+the size of the neutron star, or at least several times the scale at
+which rotation and the lateral pressure gradient balance (the Rossby
 length).  The smallest scale we need to model is dictated by
@@
 energy release and nucleosynthesis.  Our first set of two-dimensional simulations
-(Eiden et al.\ in preparation) used two physical approximations to
+(\cite{astronum:2018}, Eiden et al.\ in preparation) used two physical approximations to
 make the problem more tractable.  First, we used a higher rotation
@@
 less refinement level.  This will allow us to expand to larger domains
-while keeping the memory demands reasonable.  Figure~XX shows a
-simulation of the flame spreading done with the second-order SDC
-solver (the 4th order solver does not yet work with axisymmetry),
-demonstrating that the SDC framework can evolve this system.
+while keeping the memory demands reasonable.  Full science simulations
+will begin once we port the SDC framework to AMR and generalize the
+fourth order solver to axisymmetric geometries.  Proof-of-concept single
+level runs are running now, but the lack of AMR makes the very
+expensive.
 
 The large increase in performance of the code when run on GPUs enables
-us to do away with the flame boosting.  Preliminary estimates show
+us to do away with the flame boosting.  All of the physics solvers
+needed for the XRB simulations run on the GPUs: the hydrodynamics,
+explicit diffusion, realistic equation of state and conductivities,
+and the 13-isotope He burning network we are using.  Preliminary
+estimates show that we can expect a speed-up of atleast $5\times$ when
+comparing the NERSC Cori Haswell nodes (using CPUs) to the OLCF Summit
+nodes (using GPUs).  Work underway with further optimize the GPU code,
+especially the reaction networks.
 
@@
 
-The development of SDC in \castro\ continues, with the current focus
-on modeling detonations.  Reactions in astrophysical detonations are
-especially stiff and challenging to model when the ash state reaches
-nuclear statistical equilibrium.  We are exploring different
-quadrature schemes for the integral in Equation~\ref{eq:sdc} to model
-detonations in the SDC framework.  We are also working on extending
+Since its inception \castro\ has undergone steady development to
+enable new science investigations and take advantage of new
+supercomputer architectures.  The development of SDC, summarized here,
+continues, with the current focus on enabling our XRB studies.  We are
+also investigating astrophysical detonations with SDC to understand
+how well the method works with nuclear statistical equilibrium.  For
+the detonation work, we are exploring different quadrature schemes for
+the integral in Equation~\ref{eq:sdc:general} that are more amenable
+to highly-stiff reaction processes.  We are also working on extending
 the SDC integration methodology to adaptive mesh refinement with
@@
 
-All of these developments fit into our goal of modeling XRBs.
+Altogether, these developments will allow us to begin our first
+three-dimensional realistic XRB calculations by the end of the year,
+with parameter studies to follow.  These calculations will help us
+understand the nucleosynthesis during XRBs and its impact on
+observables and the interpretation of lightcurves and neutron star
+structure.
 

### 2019-10-02 astronum_2019_Castro/astronum-castro-2019.tex
fixed some minor typos

@@
 the \amrex\ library~\cite{amrex_joss}.  This allows us to focus the
-computational effort of regions where burning or the flow is
+computational effort on regions where burning or the flow is
 important.  To address the range of timescales involved in
@@
 hydrodynamic sources), and $\Rbs{\Uc}$ are the reactive source terms.
-The update is done at specific temporal nodes, the number of location
+The update is done at specific temporal nodes, the number of locations
 of which are picked to give the desired temporal accuracy.  The update
@@
 conserved state in a zone, $\avg{\Advs{\Uc}}^m$ is the average
-advective term in a zone at time $t^m$, and $\avg{\Rbs{\Uc}}^{k+1}$ is
+advective term in a zone at time $t^m$, and $\avg{\Rbs{\Uc}}^{k+1}$ \MarginPar{I think this should be $\avg{\Rbs{\Uc}}^{m+1}$?}is
 the average of the reactive source at time $t^{m+1}$.  The second
@@
 problems where burning is important and can dominate the computational
-expense, we expect these new SDC method will become the preferred
+expense, we expect this new SDC method will become the preferred
 integration technique in the future.
@@
 The original approach to parallelism in \castro\ was MPI + OpenMP,
-with scaling on manycore architectures achieve using a tiling approach
+with scaling on manycore architectures achieved using a tiling approach
 to OpenMP~\cite{tiling,astronum:2017}.  The general approach is to
@@
 fourth order solver to axisymmetric geometries.  Proof-of-concept single
-level runs are running now, but the lack of AMR makes the very
+level runs are running now, but the lack of AMR makes them very
 expensive.
@@
 and the 13-isotope He burning network we are using.  Preliminary
-estimates show that we can expect a speed-up of atleast $5\times$ when
+estimates show that we can expect a speed-up of at least $5\times$ when
 comparing the NERSC Cori Haswell nodes (using CPUs) to the OLCF Summit
-nodes (using GPUs).  Work underway with further optimize the GPU code,
+nodes (using GPUs).  Work is underway to further optimize the GPU code,
 especially the reaction networks.

### 2019-10-02 astronum_2019_MAESTROeX/astronum-maestroex.tex
add grants

@@
 
-\ack Figure~\ref{fig:gpu_speedup} was generated using \texttt{Jupyter} \citep{Kluyver2016} and \texttt{matplotlib} \citep{Hunter2007}.
+\ack Figure~\ref{fig:gpu_speedup} was generated using \texttt{Jupyter}
+\citep{Kluyver2016} and \texttt{matplotlib} \citep{Hunter2007}.  The
+work at Stony Brook was supported by the SciDAC program DOE grant
+DE-SC0017955 and DOE/Office of Nuclear Physics grant
+DE-FG02-87ER40317.  The work at LBNL was supported by the DOE Office
+of Advanced Scientific Computing Research under Contract No,
+DE-AC02-05CH11231.  An award of computer time was provided by the
+Innovative and Novel Computational Impact on Theory and Experiment
+(INCITE) program. This research used resources of the Oak Ridge
+Leadership Computing Facility at the Oak Ridge National Laboratory,
+which is supported by the Office of Science of the U.S. Department of
+Energy under Contract No.\ DE-AC05-00OR22725.  This research used
+resources of the National Energy Research Scientific Computing Center,
+which is supported by the Office of Science of the U.S. Department of
+Energy under Contract No.\ DE-AC02-05CH11231. This research has made
+use of NASA's Astrophysics Data System Bibliographic Services.
 

### 2019-10-02 astronum_2019_MAESTROeX/astronum-maestroex.tex
add VODE cite

@@
 Its development is described in a series of papers: \cite{Almgren2006a,Almgren2006b,Almgren2008a,Zingale2009,Nonaka2010}.
-In \maestro, the system of low Mach number equations is solved using an explicit Godunov approach for the advection, a stiff ODE solver for the reactions (VODE, CITE), and multigrid linear solvers for the pressure projection steps.
+In \maestro, the system of low Mach number equations is solved using an explicit Godunov approach for the advection, a stiff ODE solver for the reactions (VODE~\cite{vode}), and multigrid linear solvers for the pressure projection steps.
 Strang splitting CITE is used to integrate the thermodynamic variables, a second order projection method to integrate the velocity subject to the divergence constraint, and a velocity splitting scheme to hydrodynamically evolve the base state.

### 2019-10-02 astronum_2019_MAESTROeX/astronum-maestroex.tex
add cite

@@
 In \maestro, the system of low Mach number equations is solved using an explicit Godunov approach for the advection, a stiff ODE solver for the reactions (VODE~\cite{vode}), and multigrid linear solvers for the pressure projection steps.
-Strang splitting CITE is used to integrate the thermodynamic variables, a second order projection method to integrate the velocity subject to the divergence constraint, and a velocity splitting scheme to hydrodynamically evolve the base state.
+Strang splitting~\cite{strang:1968} is used to integrate the thermodynamic variables, a second order projection method to integrate the velocity subject to the divergence constraint, and a velocity splitting scheme to hydrodynamically evolve the base state.
 The original \maestro~code was developed using the pure Fortran 90 \fboxlib~software framework \cite{Zhang2016}; \maestroex~instead uses the C++/Fortran 90 \amrex~framework \cite{Zhang2019}.

### 2019-10-02 astronum_2019_MAESTROeX/astronum-maestroex.tex
some updates

@@
 \maestroex~is currently able to model spherically symmetric systems, however it has no support for rotation.
-We wish to model convection in the interiors of massive stars, and it is known that these stars often have non-negligible rotational frequencies and that this rotation can have have significant effect on mixing at convective shell boundaries.
+\MarginPar{I would say it has no support for rotation that breaks the spherical symmetry} We wish to model convection in the interiors of massive stars, and it is known that these stars often have non-negligible rotational frequencies and that this rotation can have have significant effect on mixing at convective shell boundaries.
 We are therefore currently exploring several possible ways of implementing rotation in \maestroex.
@@
 \MarginPar{I feel like someone who knows what they're talking about should write/seriously edit this section}
-The latest generation of supercomputers to come online (e.g.~Summit, soon Perlmutter and Frontier) rely heavily on GPUs in order to achieve high performance while keeping energy consumption to a minimum (GPUs consume considerably less power per flop than CPUs).
+The latest generation of supercomputers to come online (e.g.~ OLCF Summit, soon NERSC Perlmutter and OLCF Frontier) rely heavily on GPUs in order to achieve high performance while keeping energy consumption to a minimum (GPUs consume considerably less power per flop than CPUs).
 This trend is set to continue: supercomputers are moving to new architectures in order to increase performance rather than simply adding more and more CPUs.
@@
 By changing the compiler flags, the same bit of code can be compiled for serial, MPI-only, OpenMP-only, MPI+OpenMP, GPU-only or MPI+GPU.
-In order to achieve the best performance on the GPU, we require some additional optimization of the computational kernals.
+In order to achieve the best performance on the GPU, we require some additional optimization of the computational kernels.
 However, we find that these optimizations often improve the performance when the code is run on the CPU only as well.

### 2019-10-02 astronum_2019_MAESTROeX/astronum-maestroex.tex
added citations, wrote paragraph about the speedup figure

--- 2019-10-02 astronum_2019_MAESTROeX/astronum-maestroex.tex
@@
 \newcommand{\amrex}{{\sffamily AMReX}}
+\newcommand{\starkiller}{{\sffamily Starkiller Microphysics}}
 \newcommand{\amrexastro}{{\sffamily AMReX-Astro}}
@@
 The technique that we use in \maestroex~is to modify the fluid equations themselves so as to filter out the soundwaves.
-This is a similar approach to the incompressible \cite{Boussinesq1901} and anelastic \cite{Ogura1962a,Gough1968,Durran1989} approximations, however the approximation that we use (called the low Mach number approximation \cite{Day2000,Almgren2006a,Nonaka2010} and the generalized pseudo-incompressible approximation \cite{Vasil2013}) allows for background stratification and large density and temperature perturbations due to heating and changes in composition.
+This is a similar approach to the incompressible \cite{Boussinesq1901} and anelastic \cite{Ogura1962a,Gough1968,Durran1989} approximations, however the approximation that we use (called the low Mach number approximation \cite{Day2000,Almgren2006a,Nonaka2010} and the generalised pseudo-incompressible approximation \cite{Vasil2013}) allows for background stratification and large density and temperature perturbations due to heating and changes in composition.
 This is achieved by decomposing the pressure into a 1d hydrostatic base state, $p_0 = p_0(r, t)$, and a dynamic pressure perturbation, $\pi = \pi(r, \bm{x}, t)$, such that the full pressure is given by $p(r, \bm{x}, t) = p_0(r, t) + \pi(\bm{x}, t)$.
@@
 Its development is described in a series of papers: \cite{Almgren2006a,Almgren2006b,Almgren2008a,Zingale2009,Nonaka2010}.
-In \maestro, the system of low Mach number equations is solved using an explicit Godunov approach for the advection, a stiff ODE solver for the reactions (VODE~\cite{vode}), and multigrid linear solvers for the pressure projection steps.
-Strang splitting CITE is used to integrate the thermodynamic variables, a second order projection method to integrate the velocity subject to the divergence constraint, and a velocity splitting scheme to hydrodynamically evolve the base state.
+In \maestro, the system of low Mach number equations is solved using an explicit Godunov approach for the advection, a stiff ODE solver for the reactions (VODE, \cite{vode}), and multigrid linear solvers for the pressure projection steps.
+Strang splitting \cite{strang1968} is used to integrate the thermodynamic variables, a second order projection method to integrate the velocity subject to the divergence constraint, and a velocity splitting scheme to hydrodynamically evolve the base state.
 The original \maestro~code was developed using the pure Fortran 90 \fboxlib~software framework \cite{Zhang2016}; \maestroex~instead uses the C++/Fortran 90 \amrex~framework \cite{Zhang2019}.
@@
 All three codes are developed using the \amrex~software framework, and share much in common in terms of their development, structure and numerical methods.
+\castro~and \maestroex~share a common set of microphysics solvers provided by the \starkiller~library\footnote{\url{https://github.com/starkiller-astro/Microphysics}}.
 Developers of the different codes work closely together (in fact many of the developers work on more than one of the codes), with their shared structure and framework meaning that new features developed in one code can easily be replicated in the others.
 A prime example of this is the GPU port capability that has recently been implemented in \castro, and are currently in the process of being implemented in \maestroex.
-\MarginPar{mention starkiller astrophysics}
+% \MarginPar{mention starkiller astrophysics}
 
@@
 By changing the compiler flags, the same bit of code can be compiled for serial, MPI-only, OpenMP-only, MPI+OpenMP, GPU-only or MPI+GPU.
-In order to achieve the best performance on the GPU, we require some additional optimization of the computational kernals.
+In order to achieve the best performance on the GPU, we require some additional optimization of the computational kernels.
 However, we find that these optimizations often improve the performance when the code is run on the CPU only as well.
-More details to appear in a forthcoming \amrexastro~GPU paper.
+More details of the approach used to port our codes to GPU are to appear in a forthcoming \amrexastro~GPU paper.
 
@@
 
+As mentioned, our work porting \maestroex~to run on GPUs is still ongoing, however in \ref{fig:gpu_speedup}, we show the speedup we have achieved so far for a number of individual functions that have been offloaded to GPU.
+The plot shows the average execution time of each function, recorded for the 3d \texttt{reacting\_bubble} problem on the OLCF Summit machine.
+Both tests were run on a single node, consisting of two IBM POWER9 processors (together providing 42 physical CPU cores) and six NVIDIA Volta GPUs.
+The CPU test used 42 MPI processes, each with 4 OpenMP threads, and the GPU test used 6 MPI processes, each process associated with a single GPU.
+Note that the plot's $x$-axis scale is logarithmic.
+As can be seen, large speedups in excess of 30-40$\times$ were achieved for several of the functions when run on the GPU.
+Once we have finished porting the remaining routines to GPU and further optimised our code, we hope to achieve this sort of speedup for the entire timestep.
+
 \begin{figure}
@@
 Both tests were performed using a single node, with the CPU test using 42 MPI processes, each with 4 OpenMP threads, and the GPU test using 6 MPI processes, each process associated with a single GPU.
-Note that the $x$-axis scale is logarithmic.}\label{fig:gpu_speedup}
+Note that the $x$-axis scale is logarithmic.
+On the right, we have listed the speedup for each function when run on GPUs.}\label{fig:gpu_speedup}
 \end{figure}
@@
 
-\ack Figure~\ref{fig:gpu_speedup} was generated using \texttt{Jupyter}
-\citep{Kluyver2016} and \texttt{matplotlib} \citep{Hunter2007}.  The
-work at Stony Brook was supported by the SciDAC program DOE grant
-DE-SC0017955 and DOE/Office of Nuclear Physics grant
-DE-FG02-87ER40317.  The work at LBNL was supported by the DOE Office
-of Advanced Scientific Computing Research under Contract No,
-DE-AC02-05CH11231.  An award of computer time was provided by the
-Innovative and Novel Computational Impact on Theory and Experiment
-(INCITE) program. This research used resources of the Oak Ridge
-Leadership Computing Facility at the Oak Ridge National Laboratory,
-which is supported by the Office of Science of the U.S. Department of
-Energy under Contract No.\ DE-AC05-00OR22725.  This research used
-resources of the National Energy Research Scientific Computing Center,
-which is supported by the Office of Science of the U.S. Department of
-Energy under Contract No.\ DE-AC02-05CH11231. This research has made
-use of NASA's Astrophysics Data System Bibliographic Services.
+\ack Figure~\ref{fig:gpu_speedup} was generated using \texttt{Jupyter} \citep{Kluyver2016} and \texttt{matplotlib} \citep{Hunter2007}.
 

### 2019-10-02 astronum_2019_Castro/astronum-castro-2019.tex

@@
 conserved state in a zone, $\avg{\Advs{\Uc}}^m$ is the average
-advective term in a zone at time $t^m$, and $\avg{\Rbs{\Uc}}^{k+1}$ \MarginPar{I think this should be $\avg{\Rbs{\Uc}}^{m+1}$?}is
+advective term in a zone at time $t^m$, and $\avg{\Rbs{\Uc}}^{m+1}$
 the average of the reactive source at time $t^{m+1}$.  The second

### 2019-10-02 astronum_2019_MAESTROeX/astronum-maestroex.tex

--- 2019-10-02 astronum_2019_MAESTROeX/astronum-maestroex.tex
@@
 However, we find that these optimizations often improve the performance when the code is run on the CPU only as well.
-More details to appear in a forthcoming \amrexastro~GPU paper.
+More details of the approach used to port our codes to GPU are to appear in a forthcoming \amrexastro~GPU paper.
 

### 2019-10-02 astronum_2019_MAESTROeX/astronum-maestroex.tex
fix spacing

@@
 \MarginPar{I feel like someone who knows what they're talking about should write/seriously edit this section}
-The latest generation of supercomputers to come online (e.g.~ OLCF Summit, soon NERSC Perlmutter and OLCF Frontier) rely heavily on GPUs in order to achieve high performance while keeping energy consumption to a minimum (GPUs consume considerably less power per flop than CPUs).
+The latest generation of supercomputers to come online (e.g.\ OLCF Summit, soon NERSC Perlmutter and OLCF Frontier) rely heavily on GPUs in order to achieve high performance while keeping energy consumption to a minimum (GPUs consume considerably less power per flop than CPUs).
 This trend is set to continue: supercomputers are moving to new architectures in order to increase performance rather than simply adding more and more CPUs.

### 2019-10-02 astronum_2019_MAESTROeX/astronum-maestroex.tex
finished writing summary

@@
 
-\maestroex~is currently able to model spherically symmetric systems, however it has no support for rotation.
-\MarginPar{I would say it has no support for rotation that breaks the spherical symmetry} We wish to model convection in the interiors of massive stars, and it is known that these stars often have non-negligible rotational frequencies and that this rotation can have have significant effect on mixing at convective shell boundaries.
+\maestroex~is currently able to model spherically symmetric systems, however it has no support for rotation which breaks the spherical symmetry.
+We wish to model convection in the interiors of massive stars, and it is known that these stars often have non-negligible rotational frequencies and that this rotation can have have significant effect on mixing at convective shell boundaries.
 We are therefore currently exploring several possible ways of implementing rotation in \maestroex.
+
+As described in \cite{Zingale2011}, to incorporate rotation in our equation set, we add the Coriolis and centrifugal terms to the velocity evolution equation:
+\begin{equation}
+    \pd{\Ub}{t} = -\Ub\cdot\nabla\Ub - \frac{\beta_0}{\rho}\nabla\left(\frac{\pi}{\beta_0}\right) - \frac{\rho - \rho_0}{\rho}\bm{e}_r - 2\boldsymbol{\Omega} \times \Ub - \boldsymbol{\Omega} \times (\boldsymbol{\Omega} \times \bm{r}),
+\end{equation}
+where $\boldsymbol{\Omega}$ is the angular velocity, $\bm{F}_\text{Coriolis}=- 2\boldsymbol{\Omega} \times \Ub$ is the Coriolis force and $\bm{F}_\text{centrifugal}= - \boldsymbol{\Omega} \times (\boldsymbol{\Omega} \times \bm{r})$ the centrifugal force.
 
@@
 This was done in \maestro, and should be a reasonable approximation for slowly rotating systems.
-\MarginPar{In Zingale 2011, equation (2) has the Coriolis and centrifugal terms.  Include that equation here?}
+% \MarginPar{In Zingale 2011, equation (2) has the Coriolis and centrifugal terms.  Include that equation here?}
 
@@
 
-\MarginPar{I feel like someone who knows what they're talking about should write/seriously edit this section}
-The latest generation of supercomputers to come online (e.g.\ OLCF Summit, soon NERSC Perlmutter and OLCF Frontier) rely heavily on GPUs in order to achieve high performance while keeping energy consumption to a minimum (GPUs consume considerably less power per flop than CPUs).
+% \MarginPar{I feel like someone who knows what they're talking about should write/seriously edit this section}
+The latest generation of supercomputers to come online (e.g.~ OLCF Summit, soon NERSC Perlmutter and OLCF Frontier) rely heavily on GPUs in order to achieve high performance while keeping energy consumption to a minimum (GPUs consume considerably less power per flop than CPUs).
 This trend is set to continue: supercomputers are moving to new architectures in order to increase performance rather than simply adding more and more CPUs.
@@
 
-As mentioned, our work porting \maestroex~to run on GPUs is still ongoing, however in \ref{fig:gpu_speedup}, we show the speedup we have achieved so far for a number of individual functions that have been offloaded to GPU.
+As mentioned, our work porting \maestroex~to run on GPUs is still ongoing, however in Figure~\ref{fig:gpu_speedup} we show the speedup we have achieved so far for a number of individual functions that have been offloaded to GPU.
 The plot shows the average execution time of each function, recorded for the 3d \texttt{reacting\_bubble} problem on the OLCF Summit machine.
@@
 
-More future work: SDC, more science, etc.
+Work has recently begun on implementing the new time integration strategy, spectral deferred corrections (SDC) that has recently been implemented in \castro~\cite{castro:sdc}.
+This method eliminates the coupling error between source terms and hydrodynamics incurred in operator splitting techniques.
+It is hoped that this shall be particularly useful for our simulations of rotating massive stars, where we have found energetic reactions in the star's core to be challenging to model using our current methods.
 

### 2019-10-02 astronum_2019_MAESTROeX/astronum-maestroex.tex

@@
 % \MarginPar{I feel like someone who knows what they're talking about should write/seriously edit this section}
-The latest generation of supercomputers to come online (e.g.~ OLCF Summit, soon NERSC Perlmutter and OLCF Frontier) rely heavily on GPUs in order to achieve high performance while keeping energy consumption to a minimum (GPUs consume considerably less power per flop than CPUs).
+The latest generation of supercomputers to come online (e.g.\ OLCF Summit, soon NERSC Perlmutter and OLCF Frontier) rely heavily on GPUs in order to achieve high performance while keeping energy consumption to a minimum (GPUs consume considerably less power per flop than CPUs).
 This trend is set to continue: supercomputers are moving to new architectures in order to increase performance rather than simply adding more and more CPUs.

### 2019-10-02 astronum_2019_MAESTROeX/astronum-maestroex.tex
comment

@@
 
+\MarginPar{I didn't think we did Zenodo yet}
 New versions of the codes are released on the first of each month, and are archived on Zenodo\footnote{\url{https://zenodo.org/}}.

### 2019-10-03 astronum_2019_MAESTROeX/astronum-maestroex.tex

@@
 This is a similar approach to the incompressible \cite{Boussinesq1901} and anelastic \cite{Ogura1962a,Gough1968,Durran1989} approximations, however the approximation that we use (called the low Mach number approximation \cite{Day2000,Almgren2006a,Nonaka2010} and the generalised pseudo-incompressible approximation \cite{Vasil2013}) allows for background stratification and large density and temperature perturbations due to heating and changes in composition.
-This is achieved by decomposing the pressure into a 1d hydrostatic base state, $p_0 = p_0(r, t)$, and a dynamic pressure perturbation, $\pi = \pi(r, \bm{x}, t)$, such that the full pressure is given by $p(r, \bm{x}, t) = p_0(r, t) + \pi(\bm{x}, t)$.
+This is achieved by decomposing the pressure into a one-dimensional hydrostatic base state, $p_0 = p_0(r, t)$, and a dynamic pressure perturbation, $\pi = \pi(\bm{x}, t)$, such that the full pressure is given by $p(\bm{x}, t) = p_0(r, t) + \pi(\bm{x}, t)$.
 In the low Mach number regime, asymptotic analysis shows that $|\pi|/p_0 = O(\mathrm{Ma}^2)$.
@@
 where $\rho$, $\Ub$ and $h$ are the mass density, fluid velocity and specific enthalpy, $X_k$ and $\dot{\omega}_k$ are the species mass fraction and production rate of species $k$, and $H_{\mathrm{nuc}}$ is the energy release per time per unit mass.
-From the base state pressure, we can define a 1d base state density, $\rho_0 = \rho_0(r, t)$, which represents the lateral average and is in hydrostatic equilibrium with $p_0$:
+From the base state pressure, we can define a one-dimensional base state density, $\rho_0 = \rho_0(r, t)$, which represents the lateral average and is in hydrostatic equilibrium with $p_0$:
 \begin{equation}
@@
 All codes in the \amrexastro~suite are open source, with all development carried out in public repositories hosted on GitHub\footnote{\url{https://github.com/AMReX-Astro}}.
-We believe that having both the codes and the development process completely open promotes good scientific practices, as it means that results from our simulations are reproducible and the codes used to produce them can be investigated.\MarginPar{I feel like there is a better word here than investigated} Others can use the codes for their own scientific investigations, and it is possible to adapt all or parts of the codes to suite new problems.
+We believe that having both the codes and the development process completely open promotes good scientific practices, as it means that results from our simulations are reproducible and the codes used to produce them can be examined and improved by the community.
+Others can use the codes for their own scientific investigations, and it is possible to adapt all or parts of the codes to suite new problems.
 Using version control allows us to keep a record of the codes' development process, helping us to track down bugs and means that new and existing developers can learn from previous mistakes.
@@
 
-As described in \cite{Zingale2011}, to incorporate rotation in our equation set, we add the Coriolis and centrifugal terms to the velocity evolution equation:
+As described in \cite{Zingale2011}, to incorporate rotation in our equation set, we add the Coriolis and/or centrifugal terms to the velocity evolution equation:
 \begin{equation}
@@
 where $\boldsymbol{\Omega}$ is the angular velocity, $\bm{F}_\text{Coriolis}=- 2\boldsymbol{\Omega} \times \Ub$ is the Coriolis force and $\bm{F}_\text{centrifugal}= - \boldsymbol{\Omega} \times (\boldsymbol{\Omega} \times \bm{r})$ the centrifugal force.
+Note that for slowly rotating systems it is a reasonable approximation to ignore the centrifugal force.
 
-The simplest way to include rotation is to ignore the effects of the centrifugal force, and only include the Coriolis force.
+%The simplest way to include rotation is to ignore the effects of the centrifugal force, and only include the Coriolis force.
 This was done in \maestro, and should be a reasonable approximation for slowly rotating systems.
@@
 
-In \maestroex, we currently write the 1d base state as a function of the radial coordinate, $r$.
+In \maestroex, we currently write the one-dimensional base state as a function of the radial coordinate, $r$.
 For a spherical system, this therefore assumes the base state to be spherically symmetric.
 However, a star that is rotating at a significant rate will no longer be spherical: it will instead become oblique, bulging around the equator.
-In order to capture the star's deformation without sacrificing the 1d base state, we can rewrite the base state pressure as a function of the effective potential: $p_0(r)\rightarrow p_0(\Phi_{\text{eff}})$.
+In order to capture the star's deformation without sacrificing the base state, we can rewrite the base state pressure as a function of the effective potential: $p_0(r)\rightarrow p_0(\Phi_{\text{eff}})$.
 We can then rewrite the other base state quantities and the system of low Mach number equations as functions of the new coordinates $(\Phi_\text{eff}, \bm{x}, t)$.
@@
 To properly capture the effects of rotation in such a system, it may therefore be necessary to use a two-dimensional base state, with the base state pressure a function of both the radius and the meriodional angle $p_0(r)\rightarrow p_0(r, \theta)$.
+\MarginPar{should meriodional be meridional (throughout text)?}
 
@@
 Also important to note performance portability: GPU-ported AMReX codes still run on other architectures, optimizations for GPU often improve performance on MPI+OpenMP as well.
-%     \item GPUs in \maestroex: ported most of the source terms, working on the hydro. 1d base state causing headaches, especially for big problems where we suspect it may not fit in GPU memory.
+%     \item GPUs in \maestroex: ported most of the source terms, working on the hydro.
+one-dimensional base state causing headaches, especially for big problems where we suspect it may not fit in GPU memory.
 %     \item Performance plots? Do these at the hackathon?
@@
 As mentioned, our work porting \maestroex~to run on GPUs is still ongoing, however in Figure~\ref{fig:gpu_speedup} we show the speedup we have achieved so far for a number of individual functions that have been offloaded to GPU.
-The plot shows the average execution time of each function, recorded for the 3d \texttt{reacting\_bubble} problem on the OLCF Summit machine.
+The plot shows the average execution time of each function, recorded for the three-dimensional \texttt{reacting\_bubble} problem on the OLCF Summit machine.
 Both tests were run on a single node, consisting of two IBM POWER9 processors (together providing 42 physical CPU cores) and six NVIDIA Volta GPUs.
@@
 It uses the low Mach number approximation, which allows it to use a much larger timestep than conventional compressible schemes by filtering out the soundwaves.
-Unlike other sound-proof methods, we still retain background stratification and large pressure perturbations due to local compositional changes and heating, both of which are particularly important for modelling atmospheres and burning.
+Unlike other sound-proof methods, we still retain background stratification and large density and temperature perturbations due to local compositional changes and heating, both of which are particularly important for modelling atmospheres and burning.
 
@@
 New features developed in other codes in the family can therefore be easily implemented in \maestroex.
-As development is open, results from our simulations can be reproduced, and monthly archiving to Zenodo ensures software sustainability.
+As development is open, results from our simulations can be reproduced, and monthly archiving to Zenodo ensures software sustainability.\MarginPar{Zenodo?}
 

### 2019-10-03 astronum_2019_MAESTROeX/astronum-maestroex.tex
minor changes

@@
 
-\MarginPar{I didn't think we did Zenodo yet}
-New versions of the codes are released on the first of each month, and are archived on Zenodo\footnote{\url{https://zenodo.org/}}.
+% \MarginPar{I didn't think we did Zenodo yet}
+New versions of the codes are released on the first of each month.
+In the case of \castro, these versions are then archived on Zenodo\footnote{\url{https://zenodo.org/}}.
 This also provides us with DOIs (digital object identifiers), further enhancing the reproducibility of our results and ensuring the sustainability of the codes used to produce them.
+We intend to extend this archiving procedure to the other \amrexastro~codes in the near future.
 
@@
 These stars typically rotate at frequencies of $\sim$300\textendash600 Hz, a significant fraction of the stars' breakup velocity.
-This produces non-negligible deformation of the star about the equator, and may cause surface flows with a significant dependence on the meriodional angle.
-To properly capture the effects of rotation in such a system, it may therefore be necessary to use a two-dimensional base state, with the base state pressure a function of both the radius and the meriodional angle $p_0(r)\rightarrow p_0(r, \theta)$.
-\MarginPar{should meriodional be meridional (throughout text)?}
+This produces non-negligible deformation of the star about the equator, and may cause surface flows with a significant dependence on the meridional angle.
+To properly capture the effects of rotation in such a system, it may therefore be necessary to use a two-dimensional base state, with the base state pressure a function of both the radius and the meridional angle $p_0(r)\rightarrow p_0(r, \theta)$.
 
@@
 New features developed in other codes in the family can therefore be easily implemented in \maestroex.
-As development is open, results from our simulations can be reproduced, and monthly archiving to Zenodo ensures software sustainability.\MarginPar{Zenodo?}
+As development is open, results from our simulations can be reproduced, and monthly versioning ensures software sustainability.
 
@@
 For modelling the interiors of massive stars (which rotate relatively slowly), we may be able to get away with neglecting the centrifugal force and simply add the Coriolis force as a source term.
-However, this will not be sufficient for faster rotating systems (e.g.~for modelling Type I X-ray bursts on the surfaces of millisecond pulsars), so we are also exploring other methods including modelling the centrifugal force as an effective pressure term, rewriting the base state as a function of the effective gravitational potential and using a two-dimensional base state which is also a function of the meriodional angle.
+However, this will not be sufficient for faster rotating systems (e.g.~for modelling Type I X-ray bursts on the surfaces of millisecond pulsars), so we are also exploring other methods including modelling the centrifugal force as an effective pressure term, rewriting the base state as a function of the effective gravitational potential and using a two-dimensional base state which is also a function of the meridional angle.
 

### 2019-10-03 astronum_2019_MAESTROeX/astronum-maestroex.tex
added citation for massive star rotation rates

@@
 
-The massive stars that we are interested in modelling are believed to be relatively slow rotators\MarginPar{insert some citation/quantify this}, so it is likely that one of the simpler approaches outlined above should be sufficient.
+The massive stars that we are interested in modelling are believed to be relatively slow rotators, with the rotation rate in the core prior to collapse no more than a few percent of the Kepler frequency \cite{heger2004presupernova}, so it is likely that one of the simpler approaches outlined above should be sufficient.
 However, for modelling e.g.~type I X-ray bursts, it may be that a more sophisticated approach is needed.

### 2019-10-03 astronum_2019_MAESTROeX/astronum-maestroex.tex
fix spelling

@@
 \begin{abstract}
-Modelling long-time convective flows in the interiors of stars is extremeley challenging using conventional compressible hydrodynamics codes due to the acoustic timestep limitation.
+Modelling long-time convective flows in the interiors of stars is extremely challenging using conventional compressible hydrodynamics codes due to the acoustic timestep limitation.
 Many of these flows are in the low Mach number regime, which allows us to exploit the relationship between acoustic and advective time scales to develop a more computationally efficient approach.

### 2019-10-03 astronum_2019_MAESTROeX/astronum-maestroex.tex
Add missing factor of g in eqs 2 and 8.
remove extra "to"
arbitratily->arbitrarily

@@
 Another technique is to modify the flow parameters, artificially boosting the speed of the flow without changing its behavior so that the system evolves faster and fewer timesteps are required \citep{Rempel2005,Hotta2012}.
-Fully implicit time integration codes, such as those used by \cite{Viallet2011,Viallet2015,Goffrey2016}, are no longer restricted by the CFL condition and so can use arbitratily large timesteps.
+Fully implicit time integration codes, such as those used by \cite{Viallet2011,Viallet2015,Goffrey2016}, are no longer restricted by the CFL condition and so can use arbitrarily large timesteps.
 
@@
     \pd{\left(\rho X_k\right)}{t} &= - \nabla\cdot\left(\rho X_k \Ub \right) + \rho\dot{\omega}_k, \\
-    \pd{\Ub}{t} &= - \Ub\cdot\nabla\Ub - \frac{\beta_0}{\rho}\nabla\left(\frac{\pi}{\beta_0}\right) - \frac{\rho - \rho_0}{\rho}\bm{e}_r,\\
+    \pd{\Ub}{t} &= - \Ub\cdot\nabla\Ub - \frac{\beta_0}{\rho}\nabla\left(\frac{\pi}{\beta_0}\right) - \frac{\rho - \rho_0}{\rho} g \bm{e}_r,\\
     \pd{\left(\rho h\right)}{t} &= -\nabla\cdot\left(\rho h \Ub\right) + \md{p_0}{t} + \rho H_{\mathrm{nuc}},
@@
 \begin{equation}
-    \pd{\Ub}{t} = -\Ub\cdot\nabla\Ub - \frac{\beta_0}{\rho}\nabla\left(\frac{\pi}{\beta_0}\right) - \frac{\rho - \rho_0}{\rho}\bm{e}_r - 2\boldsymbol{\Omega} \times \Ub - \boldsymbol{\Omega} \times (\boldsymbol{\Omega} \times \bm{r}),
+    \pd{\Ub}{t} = -\Ub\cdot\nabla\Ub - \frac{\beta_0}{\rho}\nabla\left(\frac{\pi}{\beta_0}\right) - \frac{\rho - \rho_0}{\rho} g \bm{e}_r - 2\boldsymbol{\Omega} \times \Ub - \boldsymbol{\Omega} \times (\boldsymbol{\Omega} \times \bm{r}),
 \end{equation}
@@
 In the \amrex~framework, routines can be ported to GPUs by simply inserting a few macros into the code.
-The boilerplate code to required to compile this for GPU is then generated by a custom preprocessor prior to compilation.
+The boilerplate code required to compile this for GPU is then generated by a custom preprocessor prior to compilation.
 CUDA managed memory is used to take care of data transfers between the CPU and GPU, so no explicit copies of data to/from the CPU and GPU are required, significantly reducing the extra code required to port the code.

### 2019-10-03 astronum_2019_MAESTROeX/astronum-maestroex.tex
suite new problems -> suit new problems
Add a sentence about the ignition simple network integration for the reacting bubble test problem to give context for the burn timing.

@@
 We believe that having both the codes and the development process completely open promotes good scientific practices, as it means that results from our simulations are reproducible and the codes used to produce them can be examined and improved by the community.
-Others can use the codes for their own scientific investigations, and it is possible to adapt all or parts of the codes to suite new problems.
+Others can use the codes for their own scientific investigations, and it is possible to adapt all or parts of the codes to suit new problems.
 Using version control allows us to keep a record of the codes' development process, helping us to track down bugs and means that new and existing developers can learn from previous mistakes.
@@
 The plot shows the average execution time of each function, recorded for the three-dimensional \texttt{reacting\_bubble} problem on the OLCF Summit machine.
+This problem uses a simple reaction network modelling $^{12}\mathrm{C} + {}^{12}\mathrm{C} \rightarrow {}^{24}\mathrm{Mg}$ integrated on the GPUs using the CUDA Fortran port of VODE \citep{vode} described in \cite{Zingale2018}.
 Both tests were run on a single node, consisting of two IBM POWER9 processors (together providing 42 physical CPU cores) and six NVIDIA Volta GPUs.

### 2019-10-03 astronum_2019_MAESTROeX/astronum-maestroex.tex

@@
 where $\rho$, $\Ub$ and $h$ are the mass density, fluid velocity and specific enthalpy, $X_k$ and $\dot{\omega}_k$ are the species mass fraction and production rate of species $k$, and $H_{\mathrm{nuc}}$ is the energy release per time per unit mass.
-From the base state pressure, we can define a one-dimensional base state density, $\rho_0 = \rho_0(r, t)$, which represents the lateral average and is in hydrostatic equilibrium with $p_0$:
+\maestroex~defines the base state pressure $p_0$ to be consistent with the one-dimensional base state density, $\rho_0 = \rho_0(r, t)$, which represents the lateral average and is in hydrostatic equilibrium with $p_0$:
 \begin{equation}

### 2019-10-04 MAESTROeX_I/appendix.tex

@@
 \begin{tabular}{l c}
-\multirow{4}{3.25in}[30mm]{ \includegraphics[width=3.0in]{./figs/wdconvect_256_splitU} } & \multicolumn{1}{c}{\includegraphics[width=2.35in]{./figs/magvel_full_XY}} \\
+\multirow{4}{3.25in}[30mm]{ \includegraphics[width=3.25in]{./figs/wdconvect_256_splitU} } & \multicolumn{1}{c}{\includegraphics[width=2.75in]{./figs/magvel_full_XY}} \\
 & \multicolumn{1}{c}{\begin{footnotesize} (a) Velocity magnitude, solved using full $\mathbf{U}$ \end{footnotesize}} \\[1.em]
-& \multicolumn{1}{c}{\includegraphics[width=2.5in]{./figs/magvel_lineout_X}} \\ 
+& \multicolumn{1}{c}{\includegraphics[width=3.25in]{./figs/magvel_lineout_X}} \\ 
 & \multicolumn{1}{c}{\begin{footnotesize} (b) Comparison along white line \end{footnotesize}} \\
@@
   formulation, we illustrate
-  (left) the peak temperature, $T_{\text{peak}}$ in a white dwarf at resolution of $256^3$ during the early evolution of the star and
-  (right) the magnitude of velocity at early time. We observe much larger velocities outside the density cutoff region, $\rho < 10^5 \text{ g cm}^{-3}$, 
+  (Left) the peak temperature, $T_{\text{peak}}$ in a white dwarf at resolution of $256^3$ during the early evolution of the star and
+  (Right) the magnitude of velocity at early time. We observe much larger velocities outside the density cutoff region, $\rho < 10^5 \text{ g cm}^{-3}$, 
   that are not seen when we split the velocity dynamics.}

### 2019-10-04 MAESTROeX_II/paper.tex

--- 2019-09-16 MAESTROeX_I/paper.tex
@@
 %% using aastex version 6.3
+%\documentclass[trackchanges]{aastex63}
 \documentclass{aastex63}
@@
 \newcommand{\gcc}{\mathrm{g~cm^{-3} }}
+\newcommand{\kth}{{k_{\mathrm{th}}}}
 
@@
 %% this information may be modified by the editorial offices.
-\shorttitle{MAESTROeX Low Mach Number Astrophysics}
+\shorttitle{SDC Scheme for Low Mach Number Flows}
 \shortauthors{Fan et al.}
@@
 
-\title{MAESTROeX: A Massively Parallel Low Mach Number Astrophysical Solver}
+\title{Spectral Deferred Corrections Scheme for Low Mach Number Flows}
 
@@
 \begin{abstract}
-We present MAESTROeX, a massively parallel solver for low Mach number astrophysical flows.
-The underlying low Mach number equation set allows for efficient, long-time integration for highly subsonic flows compared to compressible approaches.
-MAESTROeX is suitable for modeling full spherical stars as well as well as planar simulations of dynamics within localized regions of a star,
-and can robustly handle several orders of magnitude of density and pressure stratification.
-Previously, we have described the development of the predecessor of MAESTROeX, called MAESTRO, in a series of papers. 
-Here, we present a new, greatly simplified temporal integration scheme that retains the same order of accuracy as our previous approaches. 
-We also explore the use of alternative spatial mapping of the one-dimensional base state onto the full Cartesian grid.
-The code leverages the new AMReX software framework for block-structured adaptive mesh refinement (AMR) applications, allowing for scalability to large fractions of leadership-class machines.
-Using our previous studies on the convective phase of single-degenerate progenitor models of Type Ia supernovae as a guide, we characterize the performance of the code and validate the new algorithmic features.
-Like MAESTRO, MAESTROeX is fully open source.
+  We describe a spectral deferred corrections scheme for MAESTROeX, an adaptive
+  low Mach number hydrodynamics code that allows for efficient, long-time integration. 
 \end{abstract}
@@
 
-\input{intro.tex}
+\section{Introduction} \label{sec:intro}
 
-\input{equations.tex}
+\section{Numerical Considerations and Algorithm} \label{sec:equations}
 
-\input{algorithm.tex}
+MAESTROeX integrates the following system of equations, 
+%%
+\begin{eqnarray}
+\frac{\partial\Ub}{\partial t} &=& -\Ub\cdot\nabla\Ub  - \frac{\beta_0}{\rho}\nabla\left(\frac{\pi}{\beta_0}\right) - \frac{\rho-\rho_0}{\rho} g\eb_r,\label{eq:momentum}\\
+\frac{\partial(\rho X_k)}{\partial t} &=& -\nabla\cdot(\rho X_k\Ub) + \rho\omegadot_k,\label{eq:species}\\
+\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc + \nabla\cdot\kth\nabla T,\label{eq:enthalpy}
+\end{eqnarray}
+%%
+along with a velocity constraint equation,
+%
+\begin{equation}
+\nabla\cdot(\beta_0\Ub) = \beta_0\left(S - \frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial t}\right), \label{eq:U divergence}
+\end{equation}
+%
+as described in previous papers \citep{MAESTRO_V,MAESTRO_VI}. Note that we have included thermal diffusion
+in the governing equations with the term 
+\begin{equation}
+\nabla\cdot\kth\nabla T = \nabla\cdot\frac{\kth}{c_p}\nabla h - \sum_k\nabla\cdot\frac{\xi_k k_{\rm th}}{c_p}\nabla X_k - \nabla\cdot\frac{h_p k_{\rm th}}{c_p}\nabla p_0 .
+\end{equation}
+The species are also constrained such that $\sum_k X_k = 1$ giving $\rho = \sum_k (\rho X_k)$ and
+%
+\begin{equation}
+\frac{\partial\rho}{\partial t} = -\nabla\cdot(\rho\Ub).
+\end{equation}
+%
 
-\input{results.tex}
+In past papers, MAESTROeX advance the thermodynamic variables by coupling advection, diffusion,
+and reactions using Strang splitting.  Specifically, we integrate the reaction terms 
+over half a time step ignoring contributions from advection and diffusion,
+then from this intermediate state we integrate the advection and diffusion terms over
+a full time step (while ignoring reactions), and finally integrate the reactions over
+a half time step (ignoring advection and diffusion). For problems where
+the reactions and/or diffusion greatly alter the energy balance or composition 
+as compared to advection, this operator splitting approach can lead to large 
+splitting errors and highly inaccurate solutions.
+This issue is can be particularly exasperating for low Mach number methods that can
+theoretically take large advection-based time steps.
 
-\input{conclusions.tex}
+An alternate approach to advancing the thermodynamic variables is to use
+spectral deferred corrections (SDC). SDC is an iterative scheme used to couple the various processes
+together, with each process seeing an  approximation of the other processes as a source term.
+The SDC algorithm converges to an integral representation of the solution in
+time that couples all of the processes together in a self-consistent fashion.
+It has been shown to be more accurate and efficient than Strang splitting in a
+terrestrial, non-stratified, low Mach number reacting flow solver \cite{nonaka2012sdc},
+so we believe an SDC version of MAESTROeX would greatly improve the efficiency of the code.
+\\
 
-\input{appendix.tex}
+\emph{As a first attempt, we will work on coupling advection and reactions only
+via SDC, with a base state that is fixed in time, but not space.}
+\\
+
+The time-advancement is divided into three major steps.
+The first step is the predictor, where we integrate the thermodynamic variables, $(\rho,\rho X_k,\rho h)$, over the full time step.
+The second step is corrector, where we use the results from the predictor to perform a more accurate temporal integration of the thermodynamic variables.
+The third step is the velocity and dynamic pressure update.\\
+
+{\bf Step 1:} ({\it Compute advection velocities})\\ \\
+Use $\Ub^n$ and a second-order Godunov method to compute time-centered edge velocities, $\uadvonedag$, with time-lagged dynamic pressure and explicit buoyancy as forcing terms.
+The $\dagger$ superscript indicates that this field does not satisfy the divergence constraint.
+Compute $S^{n+\myhalf,\star}$ by extrapolating in time,
+\begin{equation}
+S^{n+\myhalf,\star} = S^n + \frac{\Delta t^n}{2}\frac{S^n - S^{n-1}}{\Delta t^{n-1}},
+\end{equation}
+and project $\uadvonedag$ to obtain $\uadvone$, which satisfies
+\begin{equation}
+\nabla\cdot\left(\beta_0^n\uadvone\right) = S^{n+\myhalf,\star}.
+\end{equation}
+
+{\bf Step 2:} ({\it Predictor})\\ \\
+ In this step, we integrate $(\rho, \rho X_k, \rho h)$ over the full time step.
+The quantities $(S, \beta_0, k_{\rm th}, c_p, \xi_k, h_p)^n$ are computed from the the thermodynamic variables at $t^n$.
+This step is divided into several sub-steps:\\
+
+{\bf Step 2A:} ({\it Compute advective flux divergences})\\ \\
+Use $\uadvone$ and a second-order Godunov integrator to compute time-centered edge states, $(\rho X_k, \rho h)^{n+\myhalf,(0)}$, with time-lagged reactions ($I^{\rm lagged} = I^{(j_{\rm max})}$ from the previous time step), explicit diffusion, and time-centered thermodynamic pressure as source terms.
+Define the advective flux divergences as
+\begin{eqnarray}
+A_{\rho X_k}^{(0)} &=& -\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf,{(0)}}\uadvone\right],\\
+A_{\rho h}^{(0)} &=& -\nabla\cdot\left[\left(\rho h\right)^{n+\myhalf,(0)}\uadvone\right] + \uadvone\cdot\nabla p_0.
+\end{eqnarray}
+Next, use these fluxes to compute the time-advanced density,
+\begin{equation}
+\frac{\rho^{n+1} - \rho^n}{\Delta t} = \sum_k A_{\rho X_k}^{(0)}.
+\end{equation}
+Then, compute preliminary, time-advanced species using
+\begin{equation}
+\frac{\rho^{n+1}\widehat{X}_k^{n+1,(0)} - (\rho X_k)^n}{\Delta t} = A_{\rho X_k}^{(0)} + I_{\rho X_k}^{\rm lagged}.\label{eq:sdc species 2}
+\end{equation}
+
+{\bf Step 2B:} ({\it Compute diffusive flux divergence})\\ \\
+Solve a Crank-Nicolson-type diffusion equation for $\widehat{h}^{n+1,(0)}$, using transport coefficients evaluated at $t^n$ everywhere,
+\begin{eqnarray}
+\frac{\rho^{n+1}\widehat{h}^{n+1,(0)} - (\rho h)^n}{\Delta t} &=& A_{\rho h}^{(0)} + I_{\rho h}^{\rm lagged}\nonumber\\
+&& + \frac{1}{2}\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n + \nabla\cdot\frac{\kth^n}{c_p^n}\nabla \widehat{h}^{n+1,(0)}\right)\nonumber\\
+&& - \frac{1}{2}\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla\widehat{X}_k^{n+1,(0)}\right)\nonumber\\
+&& - \frac{1}{2}\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0\right),\label{eq:sdc enthalpy 2}
+\end{eqnarray}
+
+{\bf Step 2C:} ({\it Advance thermodynamic variables})\\ \\
+Define $Q_{\rho X_k}^{(0)}$ as the right hand side of (\ref{eq:sdc species 2}) without the $I_{\rho X_k}^{\rm lagged}$ term, and define $Q_{\rho h}^{(0)}$ as the right hand side of (\ref{eq:sdc enthalpy 2}) without the $I_{\rho h}^{\rm lagged}$ term.
+Use {\tt VODE} to integrate (\ref{eq:species}) and (\ref{eq:enthalpy}) over $\Delta t$ to advance $(\rho X_k, \rho h)^n$ to $(\rho X_k, \rho h)^{n+1,(0)}$ using the piecewise-constant advection and diffusion source terms:
+\begin{eqnarray}
+\frac{\partial(\rho X_k)}{\partial t} &=& Q_{\rho X_k}^{(0)} + \rho\dot\omega_k\\
+\frac{\partial(\rho h)}{\partial t} &=& Q_{\rho h}^{(0)} + \rho\Hnuc.
+\end{eqnarray}
+At this point we can define $I_{\rho X_k}^{(0)}$ and $I_{\rho h}^{(0)}$, or whatever term we need depending on our species and enthalpy edge state prediction types, for use in the corrector step.
+In our first implementation, we are predicting $\rho X_k$ and $\rho h$, in which case we define:
+\begin{eqnarray}
+I_{\rho X_k}^{(0)} &=& \frac{(\rho X_k)^{n+1,(0)} - (\rho X_k)^n}{\Delta t} - Q_{\rho X_k}^{(0)}\\
+I_{\rho h}^{(0)} &=& \frac{(\rho h)^{n+1,(0)} - (\rho h)^n}{\Delta t} - Q_{\rho h}^{(0)}.
+\end{eqnarray}
+
+{\bf Step 3:} ({\it Update advection velocities})\\ \\
+First, compute $S^{n+\myhalf}$ and $\beta_0^{n+\myhalf}$ using
+\begin{equation}
+S^{n+\myhalf} = \frac{S^n + S^{n+1,(0)}}{2}, \qquad \beta_0^{n+\myhalf} = \frac{\beta_0^n + \beta_0^{n+1,(0)}}{2}.
+\end{equation}
+Then, project $\uadvtwodag$ to obtain $\uadvtwo$, which satisfies
+\begin{equation}
+\nabla\cdot\left(\beta_0^{n+\myhalf}\uadvtwo\right) = S^{n+\myhalf}.
+\end{equation}
+
+{\bf Step 4:} ({\it Corrector Loop})\\ \\
+We loop over this step from $j=1,j_{\rm max}$ times.
+In the corrector, we use the time-advanced state from the predictor to perform a more accurate integration of the thermodynamic variables.
+The quantities $(S, \beta_0, k_{\rm th}, c_p, \xi_k, h_p)^{n+1,(j-1)}$ are computed from $(\rho,\rho X_k,\rho h)^{n+1,(j-1)}$.
+This step is divided into several sub-steps:\\
+
+{\bf Step 4A:} ({\it Compute advective flux divergences})\\ \\
+Use $\uadvtwo$ and a second-order Godunov integrator to compute time-centered edge states, $(\rho X_k, \rho h)^{n+\myhalf}$, with iteratively-lagged reactions ($I^{(j-1)}$), explicit diffusion, and time-centered thermodynamic pressure as source terms.
+Define the advective flux divergences as
+\begin{eqnarray}
+A_{\rho X_k}^{(j)} &=& -\nabla\cdot\left[\left(\rho X_k\right)^{n+\myhalf,(j)}\uadvtwo\right],\\
+A_{\rho h}^{(j)} &=& -\nabla\cdot\left[\left(\rho h\right)^{n+\myhalf,(j)}\uadvtwo\right] + \uadvtwo\cdot\nabla p_0.
+\end{eqnarray}
+Then, compute preliminary, time-advanced species using
+\begin{equation}
+\frac{\rho^{n+1}\widehat{X}_k^{n+1,(j)} - (\rho X_k)^n}{\Delta t} = A_{\rho X_k}^{(j)} + I_{\rho X_k}^{(j-1)}.\label{eq:sdc species 3}
+\end{equation}
+
+{\bf Step 4B:} ({\it Compute diffusive flux divergence})\\ \\
+Solve a backward-Euler-type correction equation for $\widehat{h}^{n+1,(j)}$,
+\begin{eqnarray}
+\frac{\rho^{n+1}\widehat{h}^{n+1,(j)} - (\rho h)^n}{\Delta t} &=& A_{\rho h}^{(j)} + I_{\rho h}^{(j-1)}\nonumber\\
+&& + \nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla\widehat{h}^{n+1,(j)} + \frac{1}{2}\left(\nabla\cdot\frac{\kth^n}{c_p^n}\nabla h^n - \nabla\cdot\frac{\kth^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla h^{n+1,(j-1)}\right)\nonumber\\
+&& - \frac{1}{2}\left(\sum_k\nabla\cdot\frac{\xi_k^n k_{\rm th}^n}{c_p^n}\nabla X_k^n + \sum_k\nabla\cdot\frac{\xi_k^{n+1,(j-1)} k_{\rm th}^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla\widehat{X}_k^{n+1,(j)}\right)\nonumber\\
+&& - \frac{1}{2}\left(\nabla\cdot\frac{h_p^n k_{\rm th}^n}{c_p^n}\nabla p_0 + \nabla\cdot\frac{h_p^{n+1,(j-1)}k_{\rm th}^{n+1,(j-1)}}{c_p^{n+1,(j-1)}}\nabla p_0\right),\label{eq:sdc enthalpy 3}
+\end{eqnarray}
+
+{\bf Step 4C:} ({\it Advance thermodynamic variables})\\ \\
+Define $Q_{\rho X_k}^{(j)}$ as the right hand side of (\ref{eq:sdc species 3}) without the $I_{\rho X_k}^{(j-1)}$ term, and define $Q_{\rho h}^{(j)}$ as the right hand side of (\ref{eq:sdc enthalpy 3}) without the $I_{\rho h}^{(j-1)}$ term.
+Use {\tt VODE} to integrate (\ref{eq:species}) and (\ref{eq:enthalpy}) over $\Delta t$ to advance $(\rho X_k, \rho h)^n$ to $(\rho X_k, \rho h)^{n+1,(j)}$ using the piecewise-constant advection and diffusion source terms:
+\begin{eqnarray}
+\frac{\partial(\rho X_k)}{\partial t} &=& Q_{\rho X_k}^{(j)} + \rho\dot\omega_k\\
+\frac{\partial(\rho h)}{\partial t} &=& Q_{\rho h}^{(j)} + \rho\Hnuc.
+\end{eqnarray}
+At this point we can define $I_{\rho X_k}^{(j)}$, $I_{\rho h}^{(j)}$, and any other $I$ terms we need depending on 
+our species and enthalpy edge state prediction types, for use in the predictor in the next time step.  In our first implementation, we are predicting $\rho X_k$ and $\rho h$, in which case we define:
+\begin{eqnarray}
+I_{\rho X_k}^{(j)} &=& \frac{(\rho X_k)^{n+1,(j)} - (\rho X_k)^n}{\Delta t} - Q_{\rho X_k}^{(j)}\\
+I_{\rho h}^{(j)} &=& \frac{(\rho h)^{n+1,(j)} - (\rho h)^n}{\Delta t} - Q_{\rho h}^{(j)}.
+\end{eqnarray}
+
+{\bf Step 5:} ({\it Advance velocity and dynamic pressure})\\ \\
+Similar to the original MAESTROeX algorithm, we first compute the face-centered, time-centered velocities, $\Ub^{\nph,\pred}$
+using a Godunov approach \citep{XRB_III}. Then, we update
+the velocity field $\Ub^n$ to $\Ub^{n+1,\dagger}$ by discretizing
+equation (\ref{eq:momentum}) as
+\begin{equation}
+\Ub^{n+1,\dagger}
+= \Ub^n - \dt \left[\uadvtwo \cdot \nabla \Ub^{\nph,\pred} \right]
+ - \dt \left[ \frac{\beta_0^\nph}{\rho^\nph} \nabla \left( \frac{\pi^\nmh}{\beta_0^\nmh} \right) + \frac{\left(\rho^\nph-\rho_0^\nph\right)}{\rho^\nph} g^{\nph} \eb_r \right],
+\end{equation}
+where
+\begin{equation}
+\rho^\nph = \frac{\rho^n + \rho^{n+1}}{2}, \qquad \rho_0^\nph = \frac{\rho_0^n + \rho_0^{n+1}}{2}.
+\end{equation}
+Again, the $\dagger$ superscript refers
+to the fact that the updated velocity does not satisfy the divergence constraint,
+\begin{equation}
+\nabla \cdot \left(\beta_0^{n+1} \Ub^{n+1} \right) = \beta_0^{n+1} \left[ S^{n+1} - \frac{1}{\gammaonebar^{n+1}p_0^{n+1}}\left(\frac{\partial p_0}{\partial t}\right)^{\nph}\right].\label{eq:div3}
+\end{equation}
+We use an approximate projection to project $\Ub^{n+1,\dagger}$ onto the space of velocities that satisfy the constraint to obtain $\Ub^{n+1}$ using a ``nodal'' projection.
+This projection necessarily differs from the MAC projection used in
+{\bf Step 1} and {\bf Step 3} because the velocities in those steps are defined
+on faces and $\Ub^{n+1}$ is defined at cell centers, requiring different divergence
+and gradient operators..
+
+\section{Numerical Algorithm} \label{sec:algorithm}
+
+\section{Results} \label{sec:results}
 
@@
 %\listofchanges
+\listofchanges
 

### 2019-10-04 astronum_2019_MAESTROeX/astronum-maestroex.tex

@@
 To fully capture the dynamics, it is necessary to model these systems in three dimensions at high resolution over many rotational periods.
-We present an overview of \maestroex's current capabilities, describe ongoing work to incorporate the effects of rotation and discuss how we are optimizing the code to run on GPUs.
+We present an overview of \maestroex's current capabilities, describe ongoing work to incorporate the effects of rotation and discuss how we are optimising the code to run on GPUs.
 \end{abstract}
@@
 One technique is to use preconditioners in order to reduce the stiffness of the equations, allowing larger timesteps to be used \cite{Miczek2014,Barsukow2016}.
-Another technique is to modify the flow parameters, artificially boosting the speed of the flow without changing its behavior so that the system evolves faster and fewer timesteps are required \citep{Rempel2005,Hotta2012}.
+Another technique is to modify the flow parameters, artificially boosting the speed of the flow without changing its behaviour so that the system evolves faster and fewer timesteps are required \citep{Rempel2005,Hotta2012}.
 Fully implicit time integration codes, such as those used by \cite{Viallet2011,Viallet2015,Goffrey2016}, are no longer restricted by the CFL condition and so can use arbitrarily large timesteps.
@@
 By changing the compiler flags, the same bit of code can be compiled for serial, MPI-only, OpenMP-only, MPI+OpenMP, GPU-only or MPI+GPU.
-In order to achieve the best performance on the GPU, we require some additional optimization of the computational kernels.
-However, we find that these optimizations often improve the performance when the code is run on the CPU only as well.
+In order to achieve the best performance on the GPU, we require some additional optimisation of the computational kernels.
+However, we find that these optimisations often improve the performance when the code is run on the CPU only as well.
 More details of the approach used to port our codes to GPU are to appear in a forthcoming \amrexastro~GPU paper.
@@
 The plot shows the average execution time of each function, recorded for the three-dimensional \texttt{reacting\_bubble} problem on the OLCF Summit machine.
-This problem uses a simple reaction network modelling $^{12}\mathrm{C} + {}^{12}\mathrm{C} \rightarrow {}^{24}\mathrm{Mg}$ integrated on the GPUs using the CUDA Fortran port of VODE \citep{vode} described in \cite{Zingale2018}.
+This problem uses a simple reaction network modelling $^{12}\mathrm{C} + {}^{12}\mathrm{C} \rightarrow {}^{24}\mathrm{Mg}$ that can be integrated on the GPUs using the CUDA Fortran port of VODE \citep{vode} described in \cite{Zingale2018}.
 Both tests were run on a single node, consisting of two IBM POWER9 processors (together providing 42 physical CPU cores) and six NVIDIA Volta GPUs.

### 2019-10-04 xrb/paper.tex
added more marginpar's

@@
 \author{Yuri Cavecchi}
-\affiliation{Princeton University}
+\affiliation{Department of Astrophysical Sciences, Princeton University, Peyton Hall, Princeton, NJ 08544}
+\affiliation{Mathematical Sciences and STAG Research Centre, University of Southampton, SO17 1BJ}
 
@@
 expected to greatly affect the dynamics.  Carrying the Coriolis force
-allows us to capture the geostrophic balance\MarginPar{ref} that sets
+allows us to capture the geostrophic balance\MarginPar{ref, Spitkovsky 2002?} that sets
 up via lateral hydrostatic equilibrium.
@@
 larger timestep hierarchy.  We use a CFL number of 0.8 for our
-simulations.  \MarginPar{true for all?} The base grid for our standard
+simulations.  
+% \MarginPar{true for all?} 
+The base grid for our standard
 simulations is $768\times 192$ zones and the corresponding finest grid
@@
 
-The speed and thickness of a laminar helium flame is determined by the
-energy generation rate and conductivity and scale roughly as \MarginPar{need a ref}
+The speed and thickness of a laminar helium flame are determined by the
+energy generation rate and conductivity, and scale roughly as \MarginPar{need a ref}
 \begin{equation}
@@
 propagation, we do a 1D simulation of a laminar flame using our
-microphysics.  Figure~\ref{fig:flame} shows the flame properties for a \AssignTo{Kiran and Mike will remake with the 10/10 boosted flame}
+microphysics.  Figure~\ref{fig:flame} shows the flame properties for a 
+% \AssignTo{Kiran and Mike will remake with the 10/10 boosted flame}
 He flame using the standard conductivities and our simple reaction
-network.  This flame had a density of $2\times 10^6~\gcc$ and \MarginPar{double check}
+network.  This flame had a density of $2\times 10^6~\gcc$ and 
+% \MarginPar{double check}
 temperature of $5\times 10^7$~K.  We note that this flame speed is
@@
 
-\begin{figure}[t]
-\centering
-\plotone{model}
-\caption{\label{fig:sketch} A schematic of our model setup.}
-\end{figure}
+% \begin{figure}[t]
+% \centering
+% \plotone{model}
+% \caption{\label{fig:sketch} A schematic of our model setup.}
+% \end{figure}
 
@@
 isothermal buffer above the atmosphere that allows us
-to capture expansion and explosive dynamics.  This is sketched in
-Figure~\ref{fig:sketch}.
+to capture expansion and explosive dynamics.  
+% This is sketched in
+% Figure~\ref{fig:sketch}.
 
@@
 We note that the entire simulation framework for these calculations is
-freely available in the \castro\ github repository.
+freely available in the \castro\ github repository\footnote{\url{https://github.com/AMReX-Astro/Castro}}.
 
 \begin{deluxetable}{lccccc}
-\tablecaption{\label{table:sim_names} Simulaiton parameters.}
+\tablecaption{\label{table:sim_names} Simulation parameters.}
 \tablehead{\colhead{name} & \colhead{reaction boost} & \colhead{conductivity boost} & \colhead{fine grid resolution} & \colhead{rotation rate}  & \colhead{network}}
@@
 
-Figure~\ref{fig:time_series} shows the time
+Figure~\ref{fig:time_series} shows the time \MarginPar{this plot needs a description}
 
@@
 
-\MarginPar{do we want to do a low res run?}
+\MarginPar{talk about the low res run?}
 
@@
 
-These studies complement the
+These studies complement the\MarginPar{???? Finish this sentence}
 
@@
 There are several changes we will pursue to the algorithm used to
-model these XRBs.  First, as shown in \citet{castro_sdc}, we have
+model these XRBs.  First, as shown in \citet{castro-sdc}, we have
 developed a fourth-order (in space and time) method for coupling

### 2019-10-05 astronum_2019_MAESTROeX/astronum-maestroex.tex
changes to GPU section

@@
 \newcommand{\fboxlib}{{\sffamily FBoxLib}}
+\newcommand{\boxlib}{{\sffamily BoxLib}}
 
@@
 Strang splitting~\cite{strang:1968} is used to integrate the thermodynamic variables, a second order projection method to integrate the velocity subject to the divergence constraint, and a velocity splitting scheme to hydrodynamically evolve the base state.
-The original \maestro~code was developed using the pure Fortran 90 \fboxlib~software framework \cite{Zhang2016}; \maestroex~instead uses the C++/Fortran 90 \amrex~framework \cite{Zhang2019}.
+The original \maestro~code was developed using the Fortran 90 interface of the \boxlib~software framework \cite{Zhang2016}; \maestroex~instead uses the C++/Fortran 90 \amrex~framework \cite{Zhang2019}.
 
@@
 Developers of the different codes work closely together (in fact many of the developers work on more than one of the codes), with their shared structure and framework meaning that new features developed in one code can easily be replicated in the others.
-A prime example of this is the GPU port capability that has recently been implemented in \castro, and are currently in the process of being implemented in \maestroex.
+A prime example of this is the GPU porting capability that has recently been implemented in \castro, and is currently in the process of being implemented in \maestroex.
 % \MarginPar{mention starkiller astrophysics}
@@
 Currently, each MPI process holds a copy of the entire base state.
-This is not a problem for small problems where the base state does not contain many cells compared to the number of cells in a single subgrid, however as the problem size and/or resolution gets larger, the size of the base state grows to the point that it is now significantly larger than the size of a single subgrid.
-Compared to CPUs, GPUs typically have less memory.
-For larger problems, it is therefore not possible to store the entire base state on the GPU.
-This significantly reduces the performance of the GPU parts of the code, as it will increase the amount of data will be copied back and forth between the CPU and GPU (a slow process).
-To alleviate this problem, we are looking at ways to reduce the amount of base state data that needs to be stored on the GPU at any one time.
-For example, in some routines we map the base state to the three-dimensional grid \textendash~for these, we can move this operation to the CPU so that the GPU does not require the base state.
+As the problem size and/or resolution gets larger, the size of the base state grows.
+Data transfers between the CPU and GPU are typically a slow process, so as the base state grows in size, the cost of these transfers becomes more significant and reduces the performance overall.
+As we are still in the process of porting our code to GPUs, there still remain some functions operating on the base state which run on the CPU.
+Every time one of these is called, if the necessary base state data is currently on the GPU, it is copied back to the CPU (and then back again next time it is required for a GPU-based calculation).
+
+To alleviate this problem, we are looking at ways to reduce both the number and size of these data transfers.
+For example, in some routines we map the base state to the three-dimensional grid \textendash~for these, we can move this operation to the CPU so that the GPU does not require the entire 1d base state to be copied over.
 It may be necessary for us to develop a way to split up the base state, so that subgrids only receive the sections of the base state necessary for their calculations (rather than the base state for the entire problem domain).
@@
 This problem uses a simple reaction network modelling $^{12}\mathrm{C} + {}^{12}\mathrm{C} \rightarrow {}^{24}\mathrm{Mg}$ that can be integrated on the GPUs using the CUDA Fortran port of VODE \citep{vode} described in \cite{Zingale2018}.
-Both tests were run on a single node, consisting of two IBM POWER9 processors (together providing 42 physical CPU cores) and six NVIDIA Volta GPUs.
+Both tests were run on a single node, consisting of two IBM POWER9 processors (together providing 42 physical CPU cores) and six NVIDIA Volta GPUs, and compiled using the PGI compiler version 19.4.
 The CPU test used 42 MPI processes, each with 4 OpenMP threads, and the GPU test used 6 MPI processes, each process associated with a single GPU.

### 2019-10-10 astronum_2019_Castro/astronum-castro-2019.tex
fix typo

@@
 the computational kernel.  This reuse keeps the code base
-managable---we don't need separate kernels for each
+manageable---we don't need separate kernels for each
 architecture---while allowing us to take advantage of current and next

### 2019-10-10 astronum_2019_Castro/astronum-castro-2019.tex
Reword Summit vs Cori Haswell comparison to make it clear which direction we are quoting a speedup.

@@
 and the 13-isotope He burning network we are using.  Preliminary
-estimates show that we can expect a speed-up of at least $5\times$ when
-comparing the NERSC Cori Haswell nodes (using CPUs) to the OLCF Summit
-nodes (using GPUs).  Work is underway to further optimize the GPU code,
+estimates show that we can expect a speed-up of at least $5\times$ for the OLCF Summit
+nodes (using GPUs) over the NERSC Cori Haswell nodes (using CPUs).
+Work is underway to further optimize the GPU code,
 especially the reaction networks.

### 2019-10-10 astronum_2019_Castro/astronum-castro-2019.tex
revise awkward train of and :-)

@@
 understand the nucleosynthesis during XRBs and its impact on
-observables and the interpretation of lightcurves and neutron star
+observables along with the interpretation of lightcurves and neutron star
 structure.

### 2019-10-10 astronum_2019_Castro/astronum-castro-2019.tex
Some day people will remember to capitalize NVIDIA

@@
 \address{$^3$NVIDIA Corporation, 2788 San Tomas Expressway,
-  Santa Clara, CA, 95050 USA}
+  Santa Clara, CA, 95051 USA}
 
@@
 using CPUs, we use 42 IBM Power 9 cores per node, while when using
-GPUs we use 6 Nvidia Volta GPUs per node.  We see that the code
+GPUs we use 6 NVIDIA Volta GPUs per node.  We see that the code
 performance is about $10\times$ higher per node when using GPUs, while
@@
   machine for the white dwarf merger problem.  On each node we use
-  either 42 IBM Power 9 CPU cores or 6 Nvidia Volta GPUs.  Ideal
+  either 42 IBM Power 9 CPU cores or 6 NVIDIA Volta GPUs.  Ideal
   scaling is shown for both cases as the solid line.}

### 2019-10-10 astronum_2019_Castro/astronum-castro-2019.tex
soften

@@
 problems where burning is important and can dominate the computational
-expense, we expect this new SDC method will become the preferred
+expense, we hope this new SDC method will become the preferred
 integration technique in the future.

### 2019-10-11 xrb/paper.tex
some intro work

@@
 Multidimensional simulations of burning on a neutron star are more
-challenging, and thus a number of approximations have been made in the
-past.
+difficult, with both the temporal and spatial scales presenting
+challenges~\citep{astronum_2018}.  For the spatial scales, we need to
+resolve the reaction zone, $\mathcal{O}(10~\mbox{cm})$ or smaller, the
+scale height of the atmosphere, $\mathcal{O}(500~\mbox{cm})$, and the
+Rossby scale where the Coriolis force balances the lateral pressure
+gradient, $\mathcal{O}(10^5~\mbox{cm})$~\citep{spitkovsky2002}.  In
+the temporal scales, capturing the rise, $\mathcal{O}(1~\mbox{s})$,
+and the decay of the burst, $\mathcal{O}(10~\mbox{s})$, and the
+accretion period between bursts, $\mathcal{O}(10^4~\mbox{s})$ is
+beyond the ability of multidimensional hydro codes currently.
+Nevertheless, significant progress has been made in understanding the
+multidimensional nature of XRBs, through various approximations.x
 
@@
 detonations may only really be applicable to superbursts (where carbon
-is the reactant)~\citep{Weinberg2006b,Weinberg2007}. \MarginPar{ref}
+is the reactant)~\citep{Weinberg2006b,Weinberg2007}.
 
@@
 across the neutron star surface.  These calculations used the
-shallow-water approximation, so the vertical details of the atmosphere's
-structure we not modeled.
+shallow-water approximation, so the vertical details of the
+atmosphere's structure we not modeled.  The model showed that the
+horizontal pressure gradient between the ash and fuel can be important
+in accelerating the burning front.
 
@@
 \cite{cavecchi:2013,art-2015-cavecchi-etal,art-2016-cavecchi-etal} \MarginPar{need a summary}
-
-Challenges: timescale for flame propagation, lengthscales (flame
-width, Rossby scale), reaction network size.\MarginPar{summarize AstroNum 2018 paper}
 

### 2019-10-11 xrb/paper.tex
more updates

@@
 \author{Yuri Cavecchi}
-\affiliation{Department of Astrophysical Sciences, Princeton University, Peyton Hall, Princeton, NJ 08544}
 \affiliation{Mathematical Sciences and STAG Research Centre, University of Southampton, SO17 1BJ}
@@
 \affiliation{NVIDIA Corp}
-
-%
 
@@
 $x_\mathrm{pert} = 1.024\times 10^4$~cm and $\delta_\mathrm{blend} = 2048$~cm.  Once
-the blended model is constructed, we compute $T(y)$ and $(\rho e)(y)$
-from the equation of state.  \MarginPar{show}
+the blended model is constructed, we compute $T(x,y)$ and $(\rho e)(x,y)$
+from the equation of state,
+\begin{align}
+  T(x,y) &= T(\rho(x,y), p(x,y), X_k(x,y)) \\
+  (\rho e)(x,y) &= \rho(x, y) \cdot e(\rho(x,y), p(x,y), X_k(x,y)) 
+\end{align}
 

### 2019-10-13 xrb/paper.tex
do a pass through the intro

@@
 nuclear equation of state~\citep{steiner:2010,ozel:2010}.  Extensive
-observations of XRBs have provided evidence that the burning begins in
-a localized region and spreads over the surface of the neutron
-star~\citep{bhattacharyya:2006,bhattacharyya:2007}.
+observations of brightness oscillations during the rise have provided
+evidence that the burning begins in a localized region and spreads
+over the surface of the neutron
+star~\citep{bhattacharyya:2006,bhattacharyya:2007,chakraborty:2014}.
 
@@
 have also been used to explore the sensitivity of the burst
-observables to accretion and reaction rates~\citep{Jose2010a,Lampe2016}\MarginPar{ref}.
+observables to accretion and reaction
+rates~\citep{cyburt:2010,Jose2010a,Lampe2016}, and model individual
+bursts~\citep{johnston:2019}.
 
@@
 difficult, with both the temporal and spatial scales presenting
-challenges~\citep{astronum_2018}.  For the spatial scales, we need to
+challenges (see \citealt{astronum_2018} for an overview).  For the spatial scales, we need to
 resolve the reaction zone, $\mathcal{O}(10~\mbox{cm})$ or smaller, the
@@
 Rossby scale where the Coriolis force balances the lateral pressure
-gradient, $\mathcal{O}(10^5~\mbox{cm})$~\citep{spitkovsky2002}.  In
+gradient, $\mathcal{O}(10^5~\mbox{cm})$~\citep{spitkovsky2002}.  For
 the temporal scales, capturing the rise, $\mathcal{O}(1~\mbox{s})$,
@@
 understand the role that convection plays in distributing the initial
-burning products vertically throughout the neutron star atmosphere.
+burning products vertically throughout the neutron star atmosphere as well
+as the nature of any turbulence the burning front might encounter.
 
@@
 
-Detonations are in some sense easier to model, since their structure
-follows from the jump conditions which are solved with shock-capturing
-schemes.  For deflagrations, we either need to resolve the structure
-or use a flame model.  Flame models usually assume that the flame
-structure is thin compared to the size of the system (see, e.g., \citet{Ropke2007}\MarginPar{refs} for applications to Type Ia supernovae).
-For XRBs however, the flame thickness is comparable to the scale
-height of the atmosphere, so we cannot use these approximations.
+For deflagrations, we either need to resolve the structure of the
+reaction zone or use a flame model.  Flame models usually assume that
+the flame structure is thin compared to the size of the system (see,
+e.g., \citet{Ropke2007} for applications to Type Ia supernovae).  For
+XRBs however, the flame thickness is comparable to the scale height of
+the atmosphere, so we cannot use these approximations.  Accurate
+models of flames therefore require that we resolve the thermal width,
+which is $\mathcal{O}(10~cm)$ for helium flames~\citep{Timmes00}.
 

### 2019-10-14 xrb/paper.tex
add ref

@@
 The speed and thickness of a laminar helium flame are determined by the
-energy generation rate and conductivity, and scale roughly as \MarginPar{need a ref}
+energy generation rate and conductivity, and scale roughly as 
 \begin{equation}
@@
 \end{equation}
-and the flame width scales as:
+and
 \begin{equation}
@@
 \end{equation}
+\citep{orourke:1979}.
 At the densities we consider in this simulation, the pure He flame

### 2019-10-14 xrb/paper.tex
write section on the sponge

@@
 
-\AssignTo{Mike, Alice}
-
 X-ray bursts (XRBs) are thermonuclear explosions of an accreted H or
@@
 models of flames therefore require that we resolve the thermal width,
-which is $\mathcal{O}(10~cm)$ for helium flames~\citep{Timmes00}.
+which is $\mathcal{O}(10~\mbox{cm})$ for helium flames~\citep{Timmes00}.
 
@@
 
- An equation of state of the form $p = p(\rho, e,
-\{X_k\})$ completes the thermodynamic description of the
-system.  Finally, $\kth$ is the thermal conductivity and $T$ is the
-temperature.  Diffusion is treated explicitly in time.
+An equation of state of the form $p = p(\rho, e, X_k)$ completes the
+thermodynamic description of the system.  Finally, $\kth$ is the
+thermal conductivity and $T$ is the temperature.  Diffusion is treated
+explicitly in time.
 
@@
 allows us to capture the geostrophic balance\MarginPar{ref, Spitkovsky 2002?} that sets
-up via lateral hydrostatic equilibrium.
+up via lateral hydrostatic equilibrium.  In the discussions below, we will use the Castro coordinate
+names, $(x, y)$ in our notation.
 
@@
 
-sponge\AssignTo{Mike will write}
-
-In the discussions below, we will use the Castro coordinate
-names, $(x, y)$ in our notation.
+When we begin the simulation, there is a transient phase as the flame
+gets established.  Material that is forced upward will encounter the
+steep density gradient at the top of the atmosphere and accelerate as
+it is blown out of the atmosphere.  Eventually this material will fall
+back to the top of the atmosphere.  In this paper, we are mostly
+concerned with the behavior of the flame and not any material that is
+violently blown out of the top of the atmosphere, so we apply a sponge
+to this region.  This is similar to what we've done in \cite{xrb3d}.
+The sponge drives the velocity of the material in the low
+density \MarginPar{add mathematical form of sponge} regions at the top
+of the atmosphere to zero.  This sponging helps increase our timestep
+as well.
 

### 2019-10-14 xrb/paper.tex
add refs
update table

@@
 \begin{equation}
-v_f \approx \sqrt{k_\mathrm{th} \dot{\epsilon}}
+v_f \approx \sqrt{k_\mathrm{th} \dot{\epsilon}} \qquad
+\lambda_f \approx \sqrt{\frac{k_\mathrm{th}}{\dot{\epsilon}}}
 \end{equation}
-and
-\begin{equation}
-\lambda_f = \sqrt{\frac{k_\mathrm{th}}{\dot{\epsilon}}}
-\end{equation}
-\citep{orourke:1979}.
+\citep{orourke:1979,khokhlov:1993}.
 At the densities we consider in this simulation, the pure He flame
@@
 
-Things to try to make the model feasible:
-\begin{itemize}
-\item Start the initial perturbation location near the Rossby radius
-\item Hotter ``hot'' model
-\item Aprox13
-\item Higher diffusion cutoff lessens the timestep restriction
-\end{itemize}
-
 All simulations were run with in a domain $8.192\times
@@
 
-\begin{deluxetable}{lccccc}
+\begin{deluxetable}{lcccccc}
 \tablecaption{\label{table:sim_names} Simulation parameters.}
-\tablehead{\colhead{name} & \colhead{reaction boost} & \colhead{conductivity boost} & \colhead{fine grid resolution} & \colhead{rotation rate}  & \colhead{network}}
+\tablehead{\colhead{name} &
+           \colhead{reaction} &
+           \colhead{conductivity} &
+           \colhead{fine grid} &
+           \colhead{rotation} &
+           \colhead{domain size} &
+           \colhead{network} \\
+%
+           \colhead{} &
+           \colhead{boost} &
+           \colhead{boost} &
+           \colhead{resolution} &
+           \colhead{rate} &
+           \colhead{} &
+           \colhead{}
+}
 \startdata
-10/10 & 10 & 10 & 10~cm & 2000~Hz & {\tt aprox13} \\
-10/10-slow & 10 & 10 & 10~cm & 1000~Hz & {\tt aprox13} \\
-10/10-lores & 10 & 10 & 20~cm & 2000~Hz & {\tt aprox13} \\
-5/5 & 5 & 5 & 10~cm & 2000~Hz & {\tt aprox13} \\
-10/10-iso7 & 10 & 10 & 10~cm & 2000~Hz & {\tt iso7} \\
+10/10 & 10 & 10 & 10~cm & 2000~Hz & $1.2288\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$ & {\tt aprox13} \\
+10/10-slow & 10 & 10 & 10~cm & 1000~Hz & $1.8432\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$ &{\tt aprox13} \\
+10/10-lores & 10 & 10 & 20~cm & 2000~Hz & $1.2288\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$ &{\tt aprox13} \\
+5/5 & 5 & 5 & 10~cm & 2000~Hz & $1.2288\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$  &{\tt aprox13} \\
+10/10-iso7 & 10 & 10 & 10~cm & 2000~Hz & $1.2288\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$ &{\tt iso7} \\
 \enddata
@@
 \centering
-\plotone{flame_boost_10_10_overview}
+\plotone{flame_wave_boost_10_10_overview}
 \caption{\label{fig:10_10_overview} Temperature, energy generation rate, and out-of-plane velocity for the 10/10 simulation.}
+\end{figure}
+
+% made with slice_multi_crop.py
+\begin{figure}[h]
+\centering
+\plotone{flame_wave_boost_10_10_lowres_slice}
+\caption{\label{fig:10_10_lowres} Temperature, mean molecular weight, energy generation rate, and out-of-plane velocity for the 10/10 low resolution simulation.}
 \end{figure}

### 2019-10-14 xrb/paper.tex
some results work

@@
 
-All simulations were run with in a domain $8.192\times
-10^4~\mathrm{cm} \times 2.048\times 10^4~\mathrm{cm}$.  With our
-default refinement, this gives a resolution of $20~\mathrm{cm}$ on the
-finest level.  As a comparison, the resolution in our detonation
-models was $98~\mathrm{cm}$ and in our multidimensional convection
-studies was $6~\mathrm{cm}$.
-
-We specify the rotation period, $P_\mathrm{rot}$, which is related to
-the angular rotation rate as $\Omega = 2\pi / P_\mathrm{rot}$.  Our
-default rotation period is $0.0005~\mathrm{s}$ (2000 times per
-second).  This is quite fast, but picked to make the Rossby length
-small enough to fit in the spatial scales we can model.  With this
-rotation rate, the Rossby length is
+Table~\ref{table:sim_names} summarizes the simulations that we ran.
+The majority of them used a reaction rate boosting of $10$ and a
+conductivity boosting of $10$, which should increase the flame speed
+by a factor of $10$.  Most simulations used a resolution of
+$10~\mbox{cm}$ and a domain width of a little more than one kilometer.  
+We use an artifically high rotation rate of
+$2000~\mbox{Hz}$, which gives a Rossby length of
 \begin{equation}
@@
 \end{equation}
-\MarginPar{we should make the width of the hot region closer to the
-  Rossby length} using a scale height $H_0 = 10^3~\mathrm{cm}$.  This
-is less than half our domain width.  Table~\ref{table:sim_names} lists
-the simulations run and their parameters.
-
-We note that the entire simulation framework for these calculations is
-freely available in the \castro\ github repository\footnote{\url{https://github.com/AMReX-Astro/Castro}}.
+using a scale height $H_0 = 10^3~\mathrm{cm}$.  This is about one
+quarter of domain width.  The run at $20~\mbox{cm}$ resolution used
+one fewer level of refinement.  The slower rotating case using a
+slightly wider domain to accommodate the expected larger Rossby
+length.  We note that the entire simulation framework for these
+calculations is freely available in the \castro\ github
+repository\footnote{\url{https://github.com/AMReX-Astro/Castro}}.
 

### 2019-10-14 xrb/paper.tex
updated caption of baroclinicity plot
more results

@@
 calculations is freely available in the \castro\ github
-repository\footnote{\url{https://github.com/AMReX-Astro/Castro}}.
+repository\footnote{\url{https://github.com/AMReX-Astro/Castro}, using
+  the {\tt flame\_wave} setup.}.  In the discussions below, we'll use
+the simulation names defined in the table to refer to specific runs
+and we'll use the 10/10 run as the reference calculation.
 
@@
 
-\AssignTo{Mike, Kiran, Alice}
-
-Figure~\ref{fig:time_series} shows the time \MarginPar{this plot needs a description}
-
-Here we'll show the large scale dynamics, colormap plots, etc. and discuss the trends.
+Figure~\ref{fig:time_series} shows the time evolution of the 10/10
+simulation, focusing on the mean-molecular weight,
+\begin{equation}
+\bar{A} = \left ( \sum_k \frac{X_k}{A_k} \right )^{-1}
+\end{equation}
+In each frame the buffer of \isot{Ni}{56} that serves as the
+underlying neutron star is seen spanning the bottom of the domain.
+Above that, the composition begins as pure \isot{He}{4}, but as the
+simulation progresses, the flame processes this to heavier nuclei,
+increasing $\bar{A}$.  By about $10~\mbox{ms}$, we see the flame front
+is reasonable well-defined.  We see that the bottom of the burning
+front is listed off of the base of the atmosphere, greatly increasing
+the surface area of the burning compared with a perfectly vertical
+flame front.  By $20~\mbox{ms}$ the flame has moved out substantially
+and we are beginning to see a gradient in the composition of the ash,
+with the heavier nuclei furthest behind the flame.  The boosting of
+the burning likely artificially increases this effect.
 
@@
 
+Figure~\ref{fig:10_10_overview} shows the temperature, energy
+generation rate, and $w$ component of the velocity (the out-of-plane
+velocity induced by the Coriolis force).  This latter field
+illustrates the hurricane effect setup by the laterally spreading
+burning front.  In the energy generation rate plot, we see that the
+burning is mostly concentrated toward the bottom of the layer, as
+expected since the density is greatest there.  We see that the peak of
+the burning has moved off of the symmetry axis, demonstrating that the
+burning front is propagating to the right.  In the temperature plot,
+we see the effect of our refinement criteria focusing only on the
+atmosphere where we are most dense, with an artificial drop in the
+temperature at the refinement boundary likely due to the thermal
+diffusion.  As we'll see when the lower resolution case, we do not
+believe that this affects the results.
+
 % made with overview.py
@@
 \plotone{flame_wave_boost_10_10_overview}
-\caption{\label{fig:10_10_overview} Temperature, energy generation rate, and out-of-plane velocity for the 10/10 simulation.}
+\caption{\label{fig:10_10_overview} Temperature, energy generation rate, and out-of-plane velocity for the 10/10 simulation at 20~ms.}
 \end{figure}
+
+Our default resolution puts $\sim 100$ zones vertically in the
+``cool'' \MarginPar{double check} model atmosphere height.  To
+understand the effects of resolution, we also perform a run with one
+fewer level, giving at $20~\mbox{cm}$ resolution overall.  Figure~\ref{fig:10_10_lowres}
+shows the fields at 20~ms.
 
@@
 \plotone{flame_wave_boost_10_10_lowres_slice}
-\caption{\label{fig:10_10_lowres} Temperature, mean molecular weight, energy generation rate, and out-of-plane velocity for the 10/10 low resolution simulation.}
+\caption{\label{fig:10_10_lowres} Temperature, mean molecular weight,
+  energy generation rate, and out-of-plane velocity for the 10/10 low
+  resolution simulation at 20~ms.}
 \end{figure}
@@
 
-\subsection{Resolution}
-
-\MarginPar{talk about the low res run?}
-
-Our default resolution puts $\sim 50$ zones vertically in the ``cool''
-model atmosphere height.  To understand the effects of resolution, we
-also perform a run with one additional level, with a $2\times$ jump
-refining just the region where the density is $10^4~\gcc < \rho \le
-4\times 10^4~\gcc$.  On this level, the grid resolution is
-$10~\mathrm{cm}$.
-
 \section{Discussion}

### 2019-10-14 xrb/paper.tex
more results

@@
 \centering
+\epsscale{0.75}
 \plotone{initial_model_paper}
@@
   (dashed) initial models, showing both the density and temperature.}
+\epsscale{1.0}
 \end{figure}
@@
 
+A final feature worth noting is the ash that seems to move along the
+surface at a higher velocity than the flame, via surface gravity
+waves.  The sponging that we perform is likely damping this some, and
+the method by which we initialize the flame may induce a larger
+transient than in nature, nevertheless, this surface ash is intriguing
+because it affects the composition of the photosphere ahead of the
+burning front, potentially changing our interpretation of
+observations.  This is something that will be explored more fully in
+the future.
+
 % made with overview.py
@@
 \caption{\label{fig:baroclinicity} Baroclinicity. This plot shows
-$\ln \left(\mathbf{\psi}\right)$ at time $t = 0.02~\mathrm{s}$.}
+$\ln \left(\mathbf{\psi}\right)$ at time $t = 0.02~\mathrm{s}$ for the 10/10 simulation.}
 \end{figure}

### 2019-10-14 xrb/paper.tex
more work on results

@@
 
-Our default resolution puts $\sim 100$ zones vertically in the
+Our default resolution puts $\sim 80$ zones vertically in the
 ``cool'' \MarginPar{double check} model atmosphere height.  To
 understand the effects of resolution, we also perform a run with one
-fewer level, giving at $20~\mbox{cm}$ resolution overall.  Figure~\ref{fig:10_10_lowres}
-shows the fields at 20~ms.
+fewer level, giving at $20~\mbox{cm}$ resolution overall (this is our
+10/10-lores run).  Figure~\ref{fig:10_10_lowres} shows the fields at
+20~ms.  The structure is largely the same as the 10/10 run, with
+largely the same flame shape and position, and the same structure in
+the energy generation rate.  Since we do not have a jump in refinement
+right below the atmosphere, we don't see the strong diffusive effect
+there, but we do see some cooling in the \isot{Ni}{56} region.  This
+is likely a resolution effect, again due to the strong gradient in
+temperature at the base of the atmosphere.
 
@@
 
+\subsection{Effects of our Approximations}
+
+The results above all used a boosting of 10/10.  To see how the
+results are sensitivity to our boosting we ran a simulation with a
+boosting of 5/5.  This is shown in Figure~\ref{fig:5_5_overview}.  The
+results look qualitatively the same---a laterally propagating flame
+develops that is lifted off of the bottom of the fuel layer.  The
+flame has not moved out as far as in the 10/10 simulation, simply
+because there is less energy release, but we expect that if we were to
+run this out twice as long, the flame would have advanced to the
+position seen in our 10/10 runs.  The good agreement in the structure
+of the flame seen with the lower boosting gives us confidence that the
+overall aspects of the flame structure and acceleration we are seeing
+are robust to the approximations we make.
+
+\begin{figure}[h]
+\centering
+\plotone{flame_wave_boost_5_5_slice}
+\caption{\label{fig:5_5_overview} Temperature, mean molecular weight, energy generation rate, and out-of-plane velocity for the 5/5 simulation at 20~ms.}
+\end{figure}
+
+We can also look at the
+
 \subsection{Flame Acceleration}
 
-\AssignTo{Kiran}
+To measure the propagation rate of the burning front, we track the
+location beyond the hottest part of the flame where
+$\dot{e}_\mathrm{nuc}$ first drops to less than $0.1 \%$ of
+maximum. Using $0.1 \%$ of the peak value rather than the peak value
+itself helps to mitigate fluctuations caused by localized fluid
+motions, as the reduced value lies in a less turbulent part of the
+burning region. We also average $\dot{e}_\mathrm{nuc}$ over the
+vertical coordinate to further dampen these fluctuations. After an
+initial transient period typically spanning $\sim 3$ ms, the flame
+settles into a state of steady propagation. The position data here are
+well fitted by a linear function of time (as seen in Figure
+\ref{fig:flame_speed}), and the resulting slope gives the velocity of
+the flame front.
 
-To measure the propagation rate of the burning front, we track the location beyond the hottest part of the flame where $\dot{e}_\mathrm{nuc}$ first drops to less than $0.1 \%$ of maximum.
-Using $0.1 \%$ of the peak value rather than the peak value itself helps to mitigate fluctuations caused by localized fluid motions, as the reduced value lies in a less turbulent part of the burning region.
-We also average $\dot{e}_\mathrm{nuc}$ over the vertical coordinate to further dampen these fluctuations.
-After an initial transient period typically spanning $\sim 3$ ms, the flame settles into a state of steady propagation.
-The position data here are well fitted by a linear function of time (as seen in Figure \ref{fig:flame_speed}), and the resulting slope gives the velocity of the flame front.
-
-Table \ref{table:flame_speeds} gives the flame speed measured in each multidimensional simulation, as well as laminar speeds obtained from any corresponding 1D runs.
-The 2D flames propagate at speeds about an order of magnitude faster than their 1D counterparts.
-The increase in flame speed is likely a product of hydrodynamical effects such as turbulence, wrinkling, and convective cycles, which bring cooler fuel from ahead of the front into the hottest part of the burning region.
+Table \ref{table:flame_speeds} gives the flame speed measured in each
+multidimensional simulation, as well as laminar speeds obtained from
+any corresponding 1D runs. The 2D flames propagate at speeds about an
+order of magnitude faster than their 1D counterparts. The increase in
+flame speed is likely a product of hydrodynamical effects such as
+turbulence, wrinkling, and convective cycles, which bring cooler fuel
+from ahead of the front into the hottest part of the burning region.
 
@@
 
-\subsection{Boosting}
-
-\AssignTo{Kiran, Mike}
-
-To meet reasonable timescales, we boosted the microphysical inputs to
-the flame to result in a faster flame.  Here we look at the effect of
-this choice.
-
-\begin{figure}[h]
-\centering
-\plotone{flame_wave_boost_5_5_slice}
-\caption{\label{fig:5_5_overview} Temperature, mean molecular weight, energy generation rate, and out-of-plane velocity for the 5/5 simulation.}
-\end{figure}
+{\color{red} Should say qualitatively what the plot shows}
 
@@
 \end{figure}
-
-\subsection{Rotation Rate}
-
-\AssignTo{Mike, Kiran}
-
-We also try a rotation period of $P_\mathrm{rot} \sim 10^{-3}~\mathrm{s}$.
-
-\subsection{Reaction Network}
-
-\AssignTo{Mike, Kiran}
-
-We look at iso7 instead of aprox13
 

### 2019-10-14 xrb/paper.tex
added reference

--- 2019-10-14 xrb/paper.tex
@@
 properties of the underlying neutron star, helping to illuminate the
-nuclear equation of state~\citep{steiner:2010,ozel:2010}.  Extensive
+nuclear equation of state~\citep{steiner:2010,ozel2016masses}.  Extensive
 observations of brightness oscillations during the rise have provided
@@
 calculations is freely available in the \castro\ github
-repository\footnote{\url{https://github.com/AMReX-Astro/Castro}, using
-  the {\tt flame\_wave} setup.}.  In the discussions below, we'll use
-the simulation names defined in the table to refer to specific runs
-and we'll use the 10/10 run as the reference calculation.
+repository\footnote{\url{https://github.com/AMReX-Astro/Castro}}.
 
@@
 
-Figure~\ref{fig:time_series} shows the time evolution of the 10/10
-simulation, focusing on the mean-molecular weight,
-\begin{equation}
-\bar{A} = \left ( \sum_k \frac{X_k}{A_k} \right )^{-1}
-\end{equation}
-In each frame the buffer of \isot{Ni}{56} that serves as the
-underlying neutron star is seen spanning the bottom of the domain.
-Above that, the composition begins as pure \isot{He}{4}, but as the
-simulation progresses, the flame processes this to heavier nuclei,
-increasing $\bar{A}$.  By about $10~\mbox{ms}$, we see the flame front
-is reasonable well-defined.  We see that the bottom of the burning
-front is listed off of the base of the atmosphere, greatly increasing
-the surface area of the burning compared with a perfectly vertical
-flame front.  By $20~\mbox{ms}$ the flame has moved out substantially
-and we are beginning to see a gradient in the composition of the ash,
-with the heavier nuclei furthest behind the flame.  The boosting of
-the burning likely artificially increases this effect.
+\AssignTo{Mike, Kiran, Alice}
+
+Figure~\ref{fig:time_series} shows the time \MarginPar{this plot needs a description}
+
+Here we'll show the large scale dynamics, colormap plots, etc. and discuss the trends.
 
@@
 
-Figure~\ref{fig:10_10_overview} shows the temperature, energy
-generation rate, and $w$ component of the velocity (the out-of-plane
-velocity induced by the Coriolis force).  This latter field
-illustrates the hurricane effect setup by the laterally spreading
-burning front.  In the energy generation rate plot, we see that the
-burning is mostly concentrated toward the bottom of the layer, as
-expected since the density is greatest there.  We see that the peak of
-the burning has moved off of the symmetry axis, demonstrating that the
-burning front is propagating to the right.  In the temperature plot,
-we see the effect of our refinement criteria focusing only on the
-atmosphere where we are most dense, with an artificial drop in the
-temperature at the refinement boundary likely due to the thermal
-diffusion.  As we'll see when the lower resolution case, we do not
-believe that this affects the results.
-
 % made with overview.py
@@
 \plotone{flame_wave_boost_10_10_overview}
-\caption{\label{fig:10_10_overview} Temperature, energy generation rate, and out-of-plane velocity for the 10/10 simulation at 20~ms.}
+\caption{\label{fig:10_10_overview} Temperature, energy generation rate, and out-of-plane velocity for the 10/10 simulation.}
 \end{figure}
-
-Our default resolution puts $\sim 100$ zones vertically in the
-``cool'' \MarginPar{double check} model atmosphere height.  To
-understand the effects of resolution, we also perform a run with one
-fewer level, giving at $20~\mbox{cm}$ resolution overall.  Figure~\ref{fig:10_10_lowres}
-shows the fields at 20~ms.
 
@@
 \plotone{flame_wave_boost_10_10_lowres_slice}
-\caption{\label{fig:10_10_lowres} Temperature, mean molecular weight,
-  energy generation rate, and out-of-plane velocity for the 10/10 low
-  resolution simulation at 20~ms.}
+\caption{\label{fig:10_10_lowres} Temperature, mean molecular weight, energy generation rate, and out-of-plane velocity for the 10/10 low resolution simulation.}
 \end{figure}
@@
 We look at iso7 instead of aprox13
+
+\subsection{Resolution}
+
+\MarginPar{talk about the low res run?}
+
+Our default resolution puts $\sim 50$ zones vertically in the ``cool''
+model atmosphere height.  To understand the effects of resolution, we
+also perform a run with one additional level, with a $2\times$ jump
+refining just the region where the density is $10^4~\gcc < \rho \le
+4\times 10^4~\gcc$.  On this level, the grid resolution is
+$10~\mathrm{cm}$.
 

### 2019-10-14 xrb/paper.tex
add another plot

@@
 
-We can also look at the
+We can also look at the effect of the network size.
+Figure~\ref{fig:network} shows a comparison of the 10/10 boosting run
+with the standard 13-isotope ``aprox13'' network and the reduced
+7-isotope ``iso7'' network.  We see that the flame in the aprox13 case
+is slightly more advanced and has a higher $\bar{A}$ than the iso7
+case.  We see in the next section that these two networks give largely
+the same flame speed.  The reduced network size saves a lot of memory
+which will be important when we transition to 3-d simulations.
+
+\begin{figure}[h]
+\centering
+\plotone{network_compare}
+\caption{\label{fig:network} Mean molecular weight at 15~ms comparing a run with
+  aprox13 vs a run with the iso7 network.}
+\end{figure}
+
+The final approximation to explore is our choice of rotation rate.
 

### 2019-10-14 xrb/paper.tex
minor changes

--- 2019-10-14 xrb/paper.tex
@@
 is reasonable well-defined.  We see that the bottom of the burning
-front is listed off of the base of the atmosphere, greatly increasing
+front is lifted off of the base of the atmosphere, greatly increasing
 the surface area of the burning compared with a perfectly vertical
@@
 with the heavier nuclei furthest behind the flame.  The boosting of
-the burning likely artificially increases this effect.
+the burning likely artificially increases this effect. \MarginPar{Doesn't the increased flame speed of the 10/10 run vs. the 5/5 run show this?}
 
@@
 velocity induced by the Coriolis force).  This latter field
-illustrates the hurricane effect setup by the laterally spreading
+illustrates the hurricane effect set up by the laterally spreading
 burning front.  In the energy generation rate plot, we see that the
@@
 temperature at the refinement boundary likely due to the thermal
-diffusion.  As we'll see when the lower resolution case, we do not
+diffusion. \MarginPar{Remind the reader what this was/where they should be looking in the plot to see this temperature drop} As we'll see when the lower resolution case, we do not
 believe that this affects the results.
@@
 surface at a higher velocity than the flame, via surface gravity
-waves.  The sponging that we perform is likely damping this some, and
+waves.  The sponging that we perform is likely damping this to some extent, and
 the method by which we initialize the flame may induce a larger
@@
 ``cool'' \MarginPar{double check} model atmosphere height.  To
-understand the effects of resolution, we also perform a run with one
-fewer level, giving at $20~\mbox{cm}$ resolution overall (this is our
+understand the effects of resolution, we also performed a run with one
+fewer level, giving a $20~\mbox{cm}$ resolution overall (this is our
 10/10-lores run).  Figure~\ref{fig:10_10_lowres} shows the fields at
@@
 We show some phase plots in Figure~\ref{fig:phase_plots}. The phase plot of the
-energy generation rate against the as a function of
-the $x$- and $y$-velocities shows that energy is preferentially generated in
-regions with negative $x$-velocity. This most likely corresponds to material
-being entrained along the underside of the flame front. There are loops of
+energy generation rate as a function of
+the $r$- and $z$-velocities shows that energy is preferentially generated in
+regions with negative $r$-velocity.
+This most likely corresponds to the burning along the underside of the flame front, where fuel has been entrained along the surface of the flame.
+This is further supported by the phase plot of the energy generation rate as a function of the $r$-velocity
+and the density, which shows that the energy generation rate peaks at
+$\rho \sim 3 \times 10^5~\mathrm{g/cm}^3$, at the base of the flame.
+
+In the first plot, there are ``loops" of
 points of similar $\dot{e}_\mathrm{nuc}$ in the outer edges of the plot.
@@
 and show the fluid moving in a circular motion in $u-v$ phase space.
-
-The phase plot of the energy generation rate as a function of the $x$-velocity
-and the density demonstrates that the energy generation rate peaks at
-$\rho \sim 3 \times 10^5~\mathrm{g/cm}^3$, at the base of the flame.
-\MarginPar{I'm guessing? It's at least towards the base}
 
@@
 \emph{Left}: Phase plot showing the energy generation rate as a function of
-the $x$- and $y$-velocities. \emph{Right}: phase plot showing the energy
-generation rate as a function of the $x$-velocity and the density.}
+the $r$- and $z$-velocities. The black cross shows the location of $u = v = 0$. \emph{Right}: phase plot showing the energy
+generation rate as a function of the $r$-velocity and the density.}
 \end{figure}

### 2019-10-14 xrb/paper.tex
Updates to 1D plots and speedplot
Also made some edits to the flame properties section.

@@
 % \MarginPar{double check}
-temperature of $5\times 10^7$~K.  We note that this flame speed is
-about $2\times 10^4~\cms$, the flame width is about 20~cm, and it
-takes about 0.1~s to settle into a sustained flame.  Note this speed
+temperature of $5\times 10^7$~K.  We observe that this flame speed is
+about $10^5~\cms$, the flame width is about 40~cm, and it
+takes about 3~ms to settle into a sustained flame.  Note that this speed
 is quite small compared to the speed estimated in
-\citet{spitkovsky2002}.  We expected a multidimensional flame to
+\citet{spitkovsky2002}.  We expect a multidimensional flame to
 accelerate due to hydrodynamics interactions (wrinkling, turbulence
@@
 \begin{figure*}[t]
-\plottwo{flame}{speed}
-\caption{\label{fig:flame} Our ``standard'' flame.}
+\plottwo{flame.pdf}{speed.pdf}
+\caption{\label{fig:flame} Time-evolution of the 10$\times$ boosted 1D laminar flame.
+The left plot shows temperature and nuclear energy generation profiles at 11 different time points, while the right plot shows flame propagation speed and flame thickness as functions of time.}
 \end{figure*}
@@
 
-Flame properties are show in Table XX \MarginPar{we'll redo the table with 10/10 and 5/5/ and unboosted flame}
-
-\iffalse
-\input{flame_table_1d.tex}
-\fi
+Flame propagation rates from both the 1D laminar flames and full 2D simulations are given in Table \ref{table:flame_speeds}.
+
+\begin{deluxetable}{lcc}
+	\tablecaption{\label{table:flame_speeds}Turbulent and laminar flame speeds.}
+	\tablehead{\colhead{run} & \colhead{$s_T$ (km s$^{-1}$)} & \colhead{$s_L$ (km s$^{-1}$)}}
+	\startdata
+	10/10 boost & $9.18 \pm 0.03$  & $1.06 \pm 0.01$ \\
+	5/5 boost & $4.00 \pm 0.01$ & $0.56 \pm 0.01$ \\
+	iso7 network & $7.56 \pm 0.02$ & $-$ \\
+	lowres & $9.33 \pm 0.04$ & $-$ \\
+	1000 Hz & $18.6 \pm 0.16$ & $-$ \\
+	\enddata
+	%
+\end{deluxetable}
 
@@
 
-\begin{deluxetable}{lcc}
-	\tablecaption{\label{table:flame_speeds}Turbulent and laminar flame speeds.}
-	\tablehead{\colhead{run} & \colhead{$s_T$ (km s$^{-1}$)} & \colhead{$s_L$ (km s$^{-1}$)}}
-	\startdata
-	10/10 boost & $9.18 \pm 0.03$  & $1.06 \pm 0.01$ \\
-	5/5 boost & $4.00 \pm 0.01$ & $0.56 \pm 0.01$ \\
-	iso7 network & $7.56 \pm 0.02$ & $-$ \\
-	lowres & $9.33 \pm 0.04$ & $-$ \\
-	1000 Hz & $18.6 \pm 0.16$ & $-$ \\
-	\enddata
-	%
-\end{deluxetable}
-
 \begin{figure}[t]

### 2019-10-14 xrb/paper.tex
add diffusion figure

--- 2019-10-14 xrb/paper.tex
@@
 is reasonable well-defined.  We see that the bottom of the burning
-front is lifted off of the base of the atmosphere, greatly increasing
+front is listed off of the base of the atmosphere, greatly increasing
 the surface area of the burning compared with a perfectly vertical
@@
 with the heavier nuclei furthest behind the flame.  The boosting of
-the burning likely artificially increases this effect. \MarginPar{Doesn't the increased flame speed of the 10/10 run vs. the 5/5 run show this?}
+the burning likely artificially increases this effect.
 
@@
 velocity induced by the Coriolis force).  This latter field
-illustrates the hurricane effect set up by the laterally spreading
+illustrates the hurricane effect setup by the laterally spreading
 burning front.  In the energy generation rate plot, we see that the
@@
 temperature at the refinement boundary likely due to the thermal
-diffusion. \MarginPar{Remind the reader what this was/where they should be looking in the plot to see this temperature drop} As we'll see when the lower resolution case, we do not
+diffusion.  As we'll see when the lower resolution case, we do not
 believe that this affects the results.
@@
 surface at a higher velocity than the flame, via surface gravity
-waves.  The sponging that we perform is likely damping this to some extent, and
+waves.  The sponging that we perform is likely damping this some, and
 the method by which we initialize the flame may induce a larger
@@
 ``cool'' \MarginPar{double check} model atmosphere height.  To
-understand the effects of resolution, we also performed a run with one
-fewer level, giving a $20~\mbox{cm}$ resolution overall (this is our
+understand the effects of resolution, we also perform a run with one
+fewer level, giving at $20~\mbox{cm}$ resolution overall (this is our
 10/10-lores run).  Figure~\ref{fig:10_10_lowres} shows the fields at
@@
 The final approximation to explore is our choice of rotation rate.
+{\color{red} need to add discussion + plot}
 
@@
 We show some phase plots in Figure~\ref{fig:phase_plots}. The phase plot of the
-energy generation rate as a function of
-the $r$- and $z$-velocities shows that energy is preferentially generated in
-regions with negative $r$-velocity.
-This most likely corresponds to the burning along the underside of the flame front, where fuel has been entrained along the surface of the flame.
-This is further supported by the phase plot of the energy generation rate as a function of the $r$-velocity
-and the density, which shows that the energy generation rate peaks at
-$\rho \sim 3 \times 10^5~\mathrm{g/cm}^3$, at the base of the flame.
-
-In the first plot, there are ``loops" of
+energy generation rate against the as a function of
+the $x$- and $y$-velocities shows that energy is preferentially generated in
+regions with negative $x$-velocity. This most likely corresponds to material
+being entrained along the underside of the flame front. There are loops of
 points of similar $\dot{e}_\mathrm{nuc}$ in the outer edges of the plot.
@@
 and show the fluid moving in a circular motion in $u-v$ phase space.
+
+The phase plot of the energy generation rate as a function of the $x$-velocity
+and the density demonstrates that the energy generation rate peaks at
+$\rho \sim 3 \times 10^5~\mathrm{g/cm}^3$, at the base of the flame.
+\MarginPar{I'm guessing? It's at least towards the base}
 
@@
 \emph{Left}: Phase plot showing the energy generation rate as a function of
-the $r$- and $z$-velocities. The black cross shows the location of $u = v = 0$. \emph{Right}: phase plot showing the energy
-generation rate as a function of the $r$-velocity and the density.}
+the $x$- and $y$-velocities. \emph{Right}: phase plot showing the energy
+generation rate as a function of the $x$-velocity and the density.}
 \end{figure}
@@
 
+\begin{figure}[t]
+\centering
+\plotone{diffusion_temp_amr}
+\caption{\label{fig:diffusion} blah}
+\end{figure}
+
 %======================================================================

### 2019-10-14 xrb/paper.tex
added more stuff to discussion

--- 2019-10-14 xrb/paper.tex
@@
 
-Full hydrodynamics models of flame spreading in XRB are possible with certain approximations.
-
-CPU cost
-
-Accurate simulations require
-\begin{enumerate}
-\item Resolving the scale height of the atmosphere
-
-\item Thin transition between underlying neutron star and atmosphere, to ensure that the peak temperature
-hits right at the proper base density
-\end{enumerate}
-
-The main approximations we made were
-\begin{enumerate}
-\item A higher than normal rotation rate
-\item A boosted flame
-\item Simplified reaction network
-\item A 2.5D model for the flow
-\end{enumerate}
-The first two of these were needed to reduce the spatial and temporal
-scales needed to model, making the simulations feasible.  Using the
-simulation framework developed here, these both can be relaxed in the
-future, at the cost of more computer time.  The same goes for 2.5D
+We have shown the results of our fully hydrodynamical, multidimensional simulations of flame propagation through the atmosphere of a neutron star.
+By using the fully compressible hydrodynamics equations, we are able to capture the vertical dynamics of the system.
+To accurately model the flame propagation, it is necessary to sufficiently resolve the scale height of the atmosphere.
+It is also important that there is a thin transition between the underlying neutron star and the atmosphere, to ensure that the peak temperature occurs at the correct base density.
+
+In order to satisfy these requirements whilst allowing the simulations to be computationally feasible, we used several approximations: a higher than normal rotation rate, a boosted flame, a simplified reaction network, and a 2.5D model for the flow.
+The first two of these were needed to reduce the spatial and temporal scales needed to model.
+Using the simulation framework developed here, these both can be relaxed in the future, at the cost of more computer time.
+The same goes for 2.5D
 vs.\ full 3D---the only difference is computer time, and our future
-calculations will explore the 3D evolution and compare to these.  In particular,
+calculations will explore the 3D evolution and compare to the 2.5D simulations presented here. In particular,
 in 3D we will be able to explore shear instabilities at the flame front.
@@
 
-These studies complement the\MarginPar{???? Finish this sentence}
-
-Other future work includes mixed H/He bursts.  This will require a
+\MarginPar{Relate our work to previous work}
+
+In addition to extending our models to 3D and relaxing the approximations used in this study, in the future we plan to perform a number of additional further studies.
+These include investigating the effects of ignition latitude and different initial models.
+We also plan to model mixed H/He bursts.
+This will require a
 different reaction network, and perhaps different (more relaxed)
 resolution requirements.
-
-There are several changes we will pursue to the algorithm used to
+A long term goal is to include magnetohydrodynamics in our models.
+Although the magnetic fields of neutron stars exhibiting X-ray bursts are relatively weak, with $B \leq 10^8 - 10^9$G \citep{mukherjee2015magnetic} (at least compared to e.g.
+magnetars, which have $B\sim 10^{15}$G), it has been found by \citet{art-2016-cavecchi-etal} that even weak magnetic fields could have a non-negligible effect on the flame propagation, reducing confinement due to the Coriolis force and leading to increased flame speeds.
+Ultimately, we wish to link our simulations to observed light curves.
+To do this, we will need to explore radiation transport in order to model how the burst energy propagates through the outer layers of the neutron star atmosphere.
+
+% Full hydrodynamics models of flame spreading in XRB are possible with certain approximations.
+
+% CPU cost
+
+% Accurate simulations require
+% \begin{enumerate}
+% \item Resolving the scale height of the atmosphere
+
+% \item Thin transition between underlying neutron star and atmosphere, to ensure that the peak temperature
+% hits right at the proper base density
+% \end{enumerate}
+
+% The main approximations we made were
+% \begin{enumerate}
+% \item A higher than normal rotation rate
+% \item A boosted flame
+% \item Simplified reaction network
+% \item A 2.5D model for the flow
+% \end{enumerate}
+% The first two of these were needed to reduce the spatial and temporal
+% scales needed to model, making the simulations feasible.  Using the
+% simulation framework developed here, these both can be relaxed in the
+% future, at the cost of more computer time.  The same goes for 2.5D
+% vs.\ full 3D---the only difference is computer time, and our future
+% calculations will explore the 3D evolution and compare to these.  In particular,
+% in 3D we will be able to explore shear instabilities at the flame front.
+% Larger networks are a straightforward change, and already supported in
+% \castro\ using the \pynucastro\ framework~\citep{pynucastro} and JINA
+% ReacLib rate database~\citep{reaclib}.
+
+% These studies complement the\MarginPar{???? Finish this sentence}
+
+% Other future work includes mixed H/He bursts.  This will require a
+% different reaction network, and perhaps different (more relaxed)
+% resolution requirements.
+
+In addition to exploring other aspects of the XRB physics, there are several changes we will pursue to the algorithm used to
 model these XRBs.  First, as shown in \citet{castro-sdc}, we have
@@
 
-MHD, radiation transport
+% MHD, radiation transport
 

### 2019-10-14 xrb/paper.tex

@@
 from ahead of the front into the hottest part of the burning region.
-
-\begin{deluxetable}{lcc}
-	\tablecaption{\label{table:flame_speeds}Turbulent and laminar flame speeds.}
-	\tablehead{\colhead{run} & \colhead{$s_T$ (km s$^{-1}$)} & \colhead{$s_L$ (km s$^{-1}$)}}
-	\startdata
-	10/10 boost & $9.18 \pm 0.03$  & $1.06 \pm 0.01$ \\
-	5/5 boost & $4.00 \pm 0.01$ & $0.56 \pm 0.01$ \\
-	iso7 network & $7.56 \pm 0.02$ & $-$ \\
-	lowres & $9.33 \pm 0.04$ & $-$ \\
-	1000 Hz & $18.6 \pm 0.16$ & $-$ \\
-	\enddata
-	%
-\end{deluxetable}
 

### 2019-10-14 xrb/paper.tex
start of a table
add some more diffusion stuff

@@
 \end{figure}
+
+The initial model is created on a uniform grid at the resolution
+corresponding to the finest level of refinement.  At regions of the
+atmosphere that are not refined to the finest level, we interpolate
+density, pressure, and composition on the grid and then obtain the
+temperature from the EOS.  Because of the strong degeneracy at the
+high densities at the bottom of the domain, the numerical error in the
+interpolation can create a temperature jump there.\MarginPar{discuss better}
 
@@
 atmosphere where we are most dense, with an artificial drop in the
-temperature at the refinement boundary likely due to the thermal
-diffusion. \MarginPar{Remind the reader what this was/where they should be looking in the plot to see this temperature drop} As we'll see when the lower resolution case, we do not
-believe that this affects the results.
+temperature at the refinement boundary due to the interpolation
+procedure of our initial model which focused on pressure and density
+rather than temperature. \MarginPar{Remind the reader what this
+  was/where they should be looking in the plot to see this temperature
+  drop} As we'll see when the lower resolution case, we do not believe
+that this affects the results.
 
@@
 the energy generation rate.  Since we do not have a jump in refinement
-right below the atmosphere, we don't see the strong diffusive effect
-there, but we do see some cooling in the \isot{Ni}{56} region.  This
-is likely a resolution effect, again due to the strong gradient in
-temperature at the base of the atmosphere.
+right below the atmosphere, we don't see the temperature feature from
+the initial model mapping there, but we do see some cooling in the
+\isot{Ni}{56} region.  This is likely a resolution effect, again due
+to the strong gradient in temperature at the base of the atmosphere.
 
@@
 
-Show diffusion test problem and convergence.
+\begin{deluxetable}{rL}
+	\tablecaption{\label{table:diffusion} Convergence of diffusion test problem.}
+	\tablehead{\colhead{base resolution} & \colhead{L-inf error}}
+	\startdata
+\cutinhead{no refinement}
+	32  & 5.995\times 10^{-3} \\
+        64  & 1.477\times 10^{-3} \\
+       128  & 3.664\times 10^{-4} \\
+       256  & 9.139\times 10^{-5} \\
+\cutinhead{one level of refinement}
+        32  & 1.480\times 10^{-3} \\
+        64  & 3.685\times 10^{-4} \\
+       128  & 9.406\times 10^{-5} \\
+       256  & 2.446\times 10^{-5} \\
+	\enddata
+	%
+\end{deluxetable}
+
+Table~\ref{table:diffusion} shows the L-inf error against the analytic
+solution for the diffusion test problem.  The tests were run in two
+fashions: no AMR, and AMR with one level of refinement (a jump of
+$2\times$).  In both cases we see second-order convergence of the
+error.  Figure~\ref{fig:diffusion} shows the temperature at the end of
+the simulation of the $64\times 128$ base grid plus one level of
+refinement simulation, with the grids shown to illustrate where the
+refinement jump is.
 
@@
 \plotone{diffusion_temp_amr}
-\caption{\label{fig:diffusion} blah}
+\caption{\label{fig:diffusion} Temperature for the diffusion test problem at $10^{-3}~\mbox{s}$
+for a diffusion test with a $64\times 128$ zone base grid and one level of refinement.}
 \end{figure}

### 2019-10-14 xrb/paper.tex
do the appendix

@@
 
+To test our implementation of diffusion we created a unit test in
+\castro\ that just uses the diffusion solver.  This test uses the
+standard Gaussian initial conditions---the diffusion of a Gaussian
+remains Gaussian, with a lower amplitude and wider width.  This is run
+with a constant diffusion coefficent.  The point of this test is to
+demonstrate that the implementation, done in a predictor-corrector
+form is second-order accurate.
+
+We use a 2-d axisymmetric geometry, with a reflection boundary on the symmetry axis (left)
+and Neumann boundaries on all other sides.  In this geometry, the solution behaves like
+a spherical problem, and has the form:
+\begin{equation}
+  T(r) = T_1 + (T_2 - T_1) \left (\frac{t_0}{t + t_0} \right )^{3/2} e^{-r^2/(4 D (t + t_0))}
+\end{equation}
+where $t_0$ is a small time used to set the initial width of the
+Gaussian, $r$ is the distance from the origin, and $D$ is the
+diffusion coefficient.  $T_1$ is the ambient temperature and $T_2$ is
+the temperature of the peak of the Gaussian.  We use $D = 1$, $T_1 = 1$, $T_2 = 2$,
+and $t_0 = 0.001$.
+
 \begin{deluxetable}{rL}

### 2019-10-17 xrb/paper.tex
some updates

@@
 We investigate the structure of laterally-propagating flames through
-the highly-stratified burning layer in an X-ray burst.  Simple
+the highly-stratified burning layer in an X-ray burst.  Simple \MarginPar{is simple the right word here?}
 two-dimensional simulations of flame propagation are performed through
@@
 source~\citep{starkiller}.
-
 We note that we do not explicitly model viscosity.  The reactions will
@@
 simulation, because of the increase in the scale height of the atmosphere
-behind the flame.
+behind the flame.  \MarginPar{add a pic of the grid to show refinement}
 

### 2019-10-19 xrb/paper.tex
some comments

@@
 % \AssignTo{Kiran and Mike will remake with the 10/10 boosted flame}
-He flame using the standard conductivities and our simple reaction
+He flame using the standard conductivities and our simple reaction \MarginPar{standard?  this is a boosted flame, right?}
 network.  This flame had a density of $2\times 10^6~\gcc$ and 
@@
 Flame propagation rates from both the 1D laminar flames and full 2D simulations are given in Table \ref{table:flame_speeds}.
+\MarginPar{I think we should only report the laminar speeds here and defer the turbulent speeds until the section where we talk about how they are computed}
 
@@
 models, we break the vertical structure of the atmosphere into 4
-layers: the underlying star, a ramp-up to the base of the
+layers: the underlying neutron star, a ramp-up to the base of the
 accreted atmosphere, a fuel layer representing the bulk of the
@@
 \end{equation}
-Here, $\delta_\mathrm{atm}$ is a characteristic width of the transition
-ramp, $T_\mathrm{hi}$ is the highest temperature in the HSE model---it
-will represent the base of the fuel layer.
+Here, $\delta_\mathrm{atm}$ is a characteristic width of the
+transition ramp, $T_\star$ is the temperature of the neutron star, and
+$T_\mathrm{hi}$ is the highest temperature in the HSE model---it will
+represent the base of the fuel layer.
 
@@
 initiate the flame by driving the burning.  The parameters we use for
-the model generation are listed in Table~\ref{table:params}.
+the model generation are listed in Table~\ref{table:params}. \MarginPar{reference and discuss the figure of initial model here}
 
@@
 equilibrum, the blended model is also in vertical equilibrium initially.  We choose
-$x_\mathrm{pert} = 1.024\times 10^4$~cm and $\delta_\mathrm{blend} = 2048$~cm.  Once
+$x_\mathrm{pert} = 1.024\times 10^4$~cm and $\delta_\mathrm{blend} = 2048$~cm.  Once \MarginPar{double check these numbers}
 the blended model is constructed, we compute $T(x,y)$ and $(\rho e)(x,y)$
@@
 density, pressure, and composition on the grid and then obtain the
-temperature from the EOS.  Because of the strong degeneracy at the
-high densities at the bottom of the domain, the numerical error in the
-interpolation can create a temperature jump there.\MarginPar{discuss better}
+temperature from the EOS.
 
@@
 % made with time_series.py in flame_wave_boost_10_10_paper with --skip 100
-\begin{figure}[h]
+\begin{figure}[t]
 \centering
@@
 we see the effect of our refinement criteria focusing only on the
-atmosphere where we are most dense, with an artificial drop in the
-temperature at the refinement boundary due to the interpolation
-procedure of our initial model which focused on pressure and density
-rather than temperature. \MarginPar{Remind the reader what this
+atmosphere where we are most dense, with an artificial change in the
+temperature at the refinement boundary due in part to the construction
+of the initial model using the fine grid resolution for HSE. \MarginPar{Remind the reader what this
   was/where they should be looking in the plot to see this temperature
@@
 % made with overview.py
-\begin{figure}[h]
+\begin{figure}[t]
 \centering
@@
 to the strong gradient in temperature at the base of the atmosphere.
+The strong agreement in the low resolution run to the 10/10 run gives
+us confidence that we are capturing the flame physics properly.
 
 % made with slice_multi_crop.py
-\begin{figure}[h]
+\begin{figure}[t]
 \centering
@@
 
-\begin{figure}[h]
+\begin{figure}[t]
 \centering
@@
 To measure the propagation rate of the burning front, we track the
-location beyond the hottest part of the flame where
+location beyond the hottest part of the flame \MarginPar{what does this mean?} where
 $\dot{e}_\mathrm{nuc}$ first drops to less than $0.1 \%$ of
@@
 motions, as the reduced value lies in a less turbulent part of the
-burning region. We also average $\dot{e}_\mathrm{nuc}$ over the
+burning region. We also average $\dot{e}_\mathrm{nuc}$ over the  \MarginPar{I think we should discuss the averaging first, before we discuss the peak value}
 vertical coordinate to further dampen these fluctuations. After an
@@
 Figure~\ref{fig:abar_temp} shows the energy generation rate as a function of the
-temperature and $\bar{A}$. \MarginPar{I'm assuming will be defined somewhere above}
+temperature and $\bar{A}$. 
 The energy generation rate peaks at low $\bar{A}$, where there is a high fraction
@@
 
-In addition to extending our models to 3D and relaxing the approximations used in this study, in the future we plan to perform a number of additional further studies.
-These include investigating the effects of ignition latitude and different initial models.
-We also plan to model mixed H/He bursts.
-This will require a
-different reaction network, and perhaps different (more relaxed)
-resolution requirements.
-A long term goal is to include magnetohydrodynamics in our models.
-Although the magnetic fields of neutron stars exhibiting X-ray bursts are relatively weak, with $B \leq 10^8 - 10^9$G \citep{mukherjee2015magnetic} (at least compared to e.g.
-magnetars, which have $B\sim 10^{15}$G), it has been found by \citet{art-2016-cavecchi-etal} that even weak magnetic fields could have a non-negligible effect on the flame propagation, reducing confinement due to the Coriolis force and leading to increased flame speeds.
-Ultimately, we wish to link our simulations to observed light curves.
-To do this, we will need to explore radiation transport in order to model how the burst energy propagates through the outer layers of the neutron star atmosphere.
+In addition to extending our models to 3D and relaxing the
+approximations used in this study, in the future we plan to perform a
+number of additional further studies. These include investigating the
+effects of ignition latitude and different initial models. We also
+plan to model mixed H/He bursts. This will require a different
+reaction network, and perhaps different (more relaxed) resolution
+requirements. A long term goal is to include magnetohydrodynamics in
+our models. Although the magnetic fields of neutron stars exhibiting
+X-ray bursts are relatively weak, with $B \leq 10^8 - 10^9$G
+\citep{mukherjee2015magnetic} (at least compared to e.g. magnetars,
+which have $B\sim 10^{15}$G), it has been found by
+\citet{art-2016-cavecchi-etal} that even weak magnetic fields could
+have a non-negligible effect on the flame propagation, reducing
+confinement due to the Coriolis force and leading to increased flame
+speeds. The effect of magnetic fields on the conductivities and
+anisotropic conduction might also be important.  Ultimately, we wish
+to link our simulations to observed light curves. To do this, we will
+need to explore radiation transport in order to model how the burst
+energy propagates through the outer layers of the neutron star
+atmosphere.
 
@@
 \begin{equation}
-  T(r) = T_1 + (T_2 - T_1) \left (\frac{t_0}{t + t_0} \right )^{3/2} e^{-r^2/(4 D (t + t_0))}
+  T(r,t) = T_1 + (T_2 - T_1) \left (\frac{t_0}{t + t_0} \right )^{3/2} e^{-r^2/(4 D (t + t_0))}
 \end{equation}
@@
 \centering
+\epsscale{0.5}
 \plotone{diffusion_temp_amr}
@@
 for a diffusion test with a $64\times 128$ zone base grid and one level of refinement.}
+\epsscale{1.0}
 \end{figure}

### 2019-10-26 xrb/paper.tex
Added some comments and rephrased a few sentences

@@
 We investigate the structure of laterally-propagating flames through
-the highly-stratified burning layer in an X-ray burst.  Simple \MarginPar{is simple the right word here?}
+the highly-stratified burning layer in an X-ray burst.
+Simple \MarginPar{is simple the right word here?} \MarginPar{It might be better to say simulations are performed through a simplified atmosphere?}
 two-dimensional simulations of flame propagation are performed through
@@
 flame.  We discuss the approximations needed to capture the length and
-time scales at play in an X-ray burst.  Our studies complement the
+time scales at play in an X-ray burst.  Our studies complement 
 other multidimensional studies of burning in X-ray bursts.
@@
 
-X-ray bursts (XRBs) are thermonuclear explosions of an accreted H or
+X-ray bursts (XRBs) are thermonuclear explosions in an accreted H or
 He layer on the surface of a neutron star (see~\citealt{galloway:2017}
-for a review).  Observations of bursts can help constrain the
+for a review).  Observations of bursts can assist in constraining the
 properties of the underlying neutron star, helping to illuminate the
 nuclear equation of state~\citep{steiner:2010,ozel2016masses}.  Extensive
-observations of brightness oscillations during the rise have provided
+observations of brightness oscillations during the rise (the initial phase of the burst where the observed flux rapidly increases) have provided
 evidence that the burning begins in a localized region and spreads
@@
 
-One dimensional studies of XRBs are very successful in predicting the
-lightcurves and recurrance times (see, e.g., \citealt{woosley-xrb}).
+One dimensional studies of XRBs have been very successful in predicting the
+lightcurves and recurrence times (see, e.g., \citealt{woosley-xrb}).
 These assume spherical symmetry and thus cannot capture the effects of
@@
 observables to accretion and reaction
-rates~\citep{cyburt:2010,Jose2010a,Lampe2016}, and model individual
+rates~\citep{cyburt:2010,Jose2010a,Lampe2016}, and to model individual
 bursts~\citep{johnston:2019}.
@@
 the temporal scales, capturing the rise, $\mathcal{O}(1~\mbox{s})$,
-and the decay of the burst, $\mathcal{O}(10~\mbox{s})$, and the
-accretion period between bursts, $\mathcal{O}(10^4~\mbox{s})$ is
-beyond the ability of multidimensional hydro codes currently.
+and the decay of the burst, $\mathcal{O}(10~\mbox{s})$, as well as the
+accretion period between bursts, $\mathcal{O}(10^4~\mbox{s})$, is
+currently beyond the ability of multidimensional hydro codes.
 Nevertheless, significant progress has been made in understanding the
-multidimensional nature of XRBs, through various approximations.x
+multidimensional nature of XRBs, through various approximations.
 
 Laterally propagating detonations were modeled by
-\citet{fryxellwoosley82,hedet}, but these would only be applicable at
-very high densities, since it is difficult to detonate helium at the
+\citet{fryxellwoosley82,hedet}. However, since it is difficult to detonate helium at the
 densities found in normal XRBs, and even harder to detonate hydrogen
-because of the waiting times for weak reactions.  This means that
+because of the waiting times for weak reactions, these would only occur at
+very high densities.  This means that
 detonations may only really be applicable to superbursts (where carbon
@@
 shallow-water approximation, so the vertical details of the
-atmosphere's structure we not modeled.  The model showed that the
+atmosphere's structure we not captured.  Their model showed that the
 horizontal pressure gradient between the ash and fuel can be important
@@
 and three-dimensions \citep{xrb3d}.  These calculations used low Mach
-number methods that approximate the hydrodynamics equations to filter
+number methods, which approximate the hydrodynamics equations to filter
 soundwaves, enabling large timesteps and efficient modeling of
@@
 burning products vertically throughout the neutron star atmosphere as well
-as the nature of any turbulence the burning front might encounter.
+as the nature of any turbulence the burning front might encounter as it propagates through the atmosphere.
 
@@
 the atmosphere, so we cannot use these approximations.  Accurate
-models of flames therefore require that we resolve the thermal width,
+models of flames in XRBs therefore require that we resolve the thermal width,
 which is $\mathcal{O}(10~\mbox{cm})$ for helium flames~\citep{Timmes00}.
@@
 The goal of this study is to understand what numerical and physical
-approximations are needed to perform a full hydrodynamical,
+approximations are required to perform a full hydrodynamical,
 multidimensional simulation of flame propagation through the
-atmosphere on a neutron star.  For simplicity in this first set of
-calculations, we will use pure helium composition.  These studies
-complement the other multidimensional work described above in helping
+atmosphere of a neutron star.  For simplicity in this first set of
+calculations, we will use a pure helium composition.  These studies
+complement the prior multidimensional studies described above in helping
 us to build a picture of the dynamics of X-ray bursts.
@@
 \end{eqnarray}
-Here, $\rho$ is the mass density, $\Ub$ is the velocity, and $E$ is
+Here, $\rho$ is the mass density, $\Ub$ is the velocity, $p$ is the pressure and $E$ is
 the specific total energy, which is related to the specific internal
@@
 thermal conductivity and $T$ is the temperature.  Diffusion is treated
-explicitly in time.
+explicitly in time. \MarginPar{Have not defined what r is in eqs 1-3, and Omega and g are only defined with respect to eq. 5}
 
@@
 ideal gas), photons, and degenerate/relativistic electrons, as
-described in \cite{timmes_swesty:2000}.  For reactions, we use a 13
-isotope alpha chain, derived from the {\tt aprox13}
+described in \cite{timmes_swesty:2000}.  To model reactions, we use a 
+13-isotope alpha chain, derived from the {\tt aprox13}
 network~\citep{timmes_aprox13}.  For one of our runs, we use the
@@
 previous.  \castro\ subcycles in time, so the finer grids are evolved
-with a finer timestep than the coarse grids.  Occasionally the
+with a finer timestep than the coarse grids.  Occasionally, the
 timestep chosen at the start of a cycle will violate the CFL condition
-during the advance of the finer grids.  In this case, we restart the
+during the advancement of the finer grids.  In this case, we restart the
 finer grid evolution with a smaller timestep, subcycling within the
@@
 The base grid for our standard
-simulations is $768\times 192$ zones and the corresponding finest grid
+simulations is $768\times 192$ zones, and the equivalent finest grid
 when 3 refinement levels are added would be $12288\times 3072$ zones.
 Our standard domain is $1.2288\times 10^5~\mathrm{cm} \times
-3.072\times 10^4~\mathrm{cm}$, so we have $10~\mathrm{cm}$ resolution
+3.072\times 10^4~\mathrm{cm}$, corresponding to $10~\mathrm{cm}$ resolution
 on the finest grid.  We only refine the fuel layer in the left half of
@@
 where $\mathcal{D} = \kth/(\rho c_v)$ is the thermal diffusivity.  The
-diffusivity increases rapidly at the top of the atmosphere, making
-these low density regions determine the overall timestep for the
+diffusivity increases rapidly at the top of the atmosphere, causing
+these low density regions to determine the overall timestep for the
 simulations.  Therefore, we disable thermal conduction at low
@@
 \end{equation}
-holding the temperature constant in the ghost cells.  This is solved
-together with the equation of state.  The velocity is reflected here.
+and holding the temperature constant in the ghost cells.  This is solved
+together with the equation of state.  The velocity is reflected at this boundary.
 This procedure follows the form described in \citet{ppm-hse}.  The
@@
 outflow.  The top boundary sets the state to simply the conditions
-in our outer buffer region of the initial model (see below) with the
+in our outer buffer region of the initial model (see below), with the
 normal velocity set to the larger of zero or the velocity at the top
@@
 violently blown out of the top of the atmosphere, so we apply a sponge
-to this region.  This is similar to what we've done in \cite{xrb3d}.
+to this region.  This is similar to the method we previously used in \cite{xrb3d}.
 The sponge drives the velocity of the material in the low
@@
 and allow us to understand the qualitative effects of laterally
-propagating flames.  To boost the flame, while keeping the thickness
+propagating flames.  To boost the flame while keeping the thickness
 the same, we can multiply both the burning rate and the conductivity
 by the same factor.  For our standard calculations, we choose 10 for
-each, to give a flame 10$\times$ faster.  We call this the ``10/10''
+each, to give a 10$\times$ faster flame speed.  We call this the ``10/10''
 flame.  We will also do a simulation with the reactions and
@@
 
-The mass fractions use this same profile, switching from a set
+The species mass fractions use this same profile, switching from a set
 describing the underlying star, ${X_k}_\star$, and the set for the
@@
 H_\star$), enforcing the discrete form of hydrostatic equilibrium,
-Eq.~\ref{eq:hse}.  Integrating upwards, we would find $p_i$ and
-$\rho_i$ using a Newton solve together either with the temperature
+Eq.~\ref{eq:hse}.  Integrating upwards, we find $p_i$ and
+$\rho_i$ using a Newton solve together with the equation of state, with either the temperature
 specified, $T_i = T(p_i, \rho_i, \{X_k\}_i)$ (in the isothermal, ramp,
 and buffer layers) or constant entropy, $s_i = s(p_i, \rho_i,
-\{X_k\}_i)$, in the fuel layer.  In each case, using the equation of
-state.  This follows the procedures described in \citet{ppm-hse}.  We
-use the constant temperature for all $y < H_\star +
-3\delta_\mathrm{atm}$.  Above this, we switch to isentropic until the
+\{X_k\}_i)$, in the fuel layer.  This follows the procedures described in \citet{ppm-hse}.  We
+use a constant temperature for all $y < H_\star +
+3\delta_\mathrm{atm}$.  Above this, we switch to an isentropic atmosphere until the
 temperature drops to a floor value, $T_\mathrm{lo}$, at which point we
@@
 ahead of the flame and a ``hot'' (or perturbed) model that will
-initiate the flame by driving the burning.  The parameters we use for
+initiate the flame by driving the burning.  \MarginPar{This has already been said at the start of this section} The parameters we use for
 the model generation are listed in Table~\ref{table:params}. \MarginPar{reference and discuss the figure of initial model here}
@@
 
-Table~\ref{table:sim_names} summarizes the simulations that we ran.
+Table~\ref{table:sim_names} summarizes the simulations that we ran. \MarginPar{This table has different model names to the previous table}
 The majority of them used a reaction rate boosting of $10$ and a
@@
 using a scale height $H_0 = 10^3~\mathrm{cm}$.  This is about one
-quarter of domain width.  The run at $20~\mbox{cm}$ resolution used
+quarter of the domain width.  The run at $20~\mbox{cm}$ resolution used
 one fewer level of refinement.  The slower rotating case using a
@@
 \begin{equation}
-\bar{A} = \left ( \sum_k \frac{X_k}{A_k} \right )^{-1}
+\bar{A} = \left ( \sum_k \frac{X_k}{A_k} \right )^{-1}.
 \end{equation}
@@
 the method by which we initialize the flame may induce a larger
-transient than in nature, nevertheless, this surface ash is intriguing
+transient than in nature. Nevertheless, this surface ash is intriguing
 because it affects the composition of the photosphere ahead of the
@@
 understand the effects of resolution, we also performed a run with one
-fewer level, giving a $20~\mbox{cm}$ resolution overall (this is our
-10/10-lores run).  Figure~\ref{fig:10_10_lowres} shows the fields at
+fewer refinement level, giving a $20~\mbox{cm}$ resolution overall (this is our
+10/10-lores run).  Figure~\ref{fig:10_10_lowres} shows the fields for this run at
 20~ms.  The structure is largely the same as the 10/10 run, with
@@
 The results above all used a boosting of 10/10.  To see how the
-results are sensitivity to our boosting we ran a simulation with a
+results are sensitive to this boosting, we ran a simulation with a reduced
 boosting of 5/5.  This is shown in Figure~\ref{fig:5_5_overview}.  The
@@
 
-We can also look at the effect of the network size.
+We also considered the effect of the network size.
 Figure~\ref{fig:network} shows a comparison of the 10/10 boosting run
@@
 case.  We see in the next section that these two networks give largely
-the same flame speed.  The reduced network size saves a lot of memory
-which will be important when we transition to 3-d simulations.
+the same flame speed.  The reduced network size saves a lot of memory,
+which will be useful when we transition to 3-d simulations.
 
@@
 location beyond the hottest part of the flame \MarginPar{what does this mean?} where
-$\dot{e}_\mathrm{nuc}$ first drops to less than $0.1 \%$ of
-maximum. Using $0.1 \%$ of the peak value rather than the peak value
+$\dot{e}_\mathrm{nuc}$ first drops to less than $0.1 \%$ of its
+maximum value. Using $0.1 \%$ of the peak value rather than the peak value
 itself helps to mitigate fluctuations caused by localized fluid
@@
 multidimensional simulation, as well as laminar speeds obtained from
-any corresponding 1D runs. The 2D flames propagate at speeds about an
+corresponding 1D runs. The 2D flames propagate at speeds about an
 order of magnitude faster than their 1D counterparts. The increase in
-flame speed is likely a product of hydrodynamical effects such as
+flame speed is likely a product of the larger flame surface area and hydrodynamical effects such as
 turbulence, wrinkling, and convective cycles, which bring cooler fuel
@@
 and shows the misalignment of the local density and pressure gradients
-\citep{Malone2014a}. As we are considering a 2d system, the component of the
+\citep{Malone2014a}.
+As we are considering a 2d \MarginPar{We've used at least three different ways of writing 2-d/2D/2d so far.
+These should be changed so we used one style throughout.}system, the component of the
 baroclinicity we consider here (out of the plane) reflects the misalignment
@@
 
-We show some phase plots in Figure~\ref{fig:phase_plots}. The phase plot of the
+To further demonstrate the relationships between different properties of the flame, in Figure~\ref{fig:phase_plots} we present some phase plots of the energy generation rate. The phase plot of the
 energy generation rate as a function of
-the $r$- and $z$-velocities shows that energy is preferentially generated in
-regions with negative $r$-velocity.
-This most likely corresponds to the burning along the underside of the flame front, where fuel has been entrained along the surface of the flame.
-This is further supported by the phase plot of the energy generation rate as a function of the $r$-velocity
+the $x$- and $y$-velocities shows that energy is preferentially generated in
+regions with negative $x$-velocity.
+This most likely corresponds to the burning along the underside of the flame front, where fuel has been entrained along the surface of the flame and so is moving in the opposite direction to the direction the flame is propagating in.
+This is further supported by the phase plot of the energy generation rate as a function of the $x$-velocity
 and the density, which shows that the energy generation rate peaks at
@@
 These are likely to correspond to the vortices that appear within the flame,
-and show the fluid moving in a circular motion in $u-v$ phase space.
+where the fluid moves in a circular motion in $u-v$ phase space.
 
@@
 \emph{Left}: Phase plot showing the energy generation rate as a function of
-the $r$- and $z$-velocities. The black cross shows the location of $u = v = 0$. \emph{Right}: phase plot showing the energy
-generation rate as a function of the $r$-velocity and the density.}
+the $x$- and $y$-velocities. The black cross shows the location of $u = v = 0$. \emph{Right}: phase plot showing the energy
+generation rate as a function of the $x$-velocity and the density.}
 \end{figure}
@@
 In order to satisfy these requirements whilst allowing the simulations to be computationally feasible, we used several approximations: a higher than normal rotation rate, a boosted flame, a simplified reaction network, and a 2.5D model for the flow.
-The first two of these were needed to reduce the spatial and temporal scales needed to model.
+The first two of these were used to reduce the model's spatial and temporal scales.
 Using the simulation framework developed here, these both can be relaxed in the future, at the cost of more computer time.
@@
 
-In addition to extending our models to 3D and relaxing the
-approximations used in this study, in the future we plan to perform a
-number of additional further studies. These include investigating the
-effects of ignition latitude and different initial models. We also
-plan to model mixed H/He bursts. This will require a different
-reaction network, and perhaps different (more relaxed) resolution
-requirements. A long term goal is to include magnetohydrodynamics in
-our models. Although the magnetic fields of neutron stars exhibiting
-X-ray bursts are relatively weak, with $B \leq 10^8 - 10^9$G
-\citep{mukherjee2015magnetic} (at least compared to e.g. magnetars,
-which have $B\sim 10^{15}$G), it has been found by
-\citet{art-2016-cavecchi-etal} that even weak magnetic fields could
-have a non-negligible effect on the flame propagation, reducing
-confinement due to the Coriolis force and leading to increased flame
-speeds. The effect of magnetic fields on the conductivities and
-anisotropic conduction might also be important.  Ultimately, we wish
-to link our simulations to observed light curves. To do this, we will
-need to explore radiation transport in order to model how the burst
-energy propagates through the outer layers of the neutron star
-atmosphere.
+In addition to extending our models to 3D and relaxing the approximations used in 
+this study, in the future we plan to perform a number of additional further studies. 
+These include investigating the effects of ignition latitude and different initial 
+models. We also plan to model mixed H/He bursts. This will require a
+different reaction network, and perhaps different (more relaxed)
+resolution requirements. A long term goal is to include magnetohydrodynamics in our 
+models. Although the magnetic fields of neutron stars exhibiting X-ray bursts are 
+relatively weak, with $B \lesssim 10^8 - 10^9$G \citep{mukherjee2015magnetic} (at
+ least compared to e.g. magnetars, which have $B\lesssim 10^{15}$G), it has been 
+ found by \citet{art-2016-cavecchi-etal} that even weak magnetic fields could have 
+ a non-negligible effect on the flame propagation, reducing confinement due to the 
+ Coriolis force and leading to increased flame speeds. Ultimately, we wish to link 
+ our simulations to observed light curves. To do this, we will need to explore 
+ radiation transport in order to model how the burst energy propagates through the 
+ outer layers of the neutron star atmosphere.
 
@@
 
-In addition to exploring other aspects of the XRB physics, there are several changes we will pursue to the algorithm used to
-model these XRBs.  First, as shown in \citet{castro-sdc}, we have
+In addition to exploring other aspects of the XRB physics, there are several changes to the algorithm used to
+model these XRBs we will pursue.  First, as shown in \citet{castro-sdc}, we have
 developed a fourth-order (in space and time) method for coupling
@@
 
-To test our implementation of diffusion we created a unit test in
+To test our implementation of diffusion, we created a unit test in
 \castro\ that just uses the diffusion solver.  This test uses the
@@
 with a constant diffusion coefficent.  The point of this test is to
-demonstrate that the implementation, done in a predictor-corrector
+demonstrate that the implementation done in a predictor-corrector
 form is second-order accurate.
@@
 $2\times$).  In both cases we see second-order convergence of the
-error.  Figure~\ref{fig:diffusion} shows the temperature at the end of
+error.\MarginPar{This would be clearer if there was a third column showing the ratios of the errors?}  Figure~\ref{fig:diffusion} shows the temperature at the end of
 the simulation of the $64\times 128$ base grid plus one level of

### 2019-11-04 xrb/paper.tex
Fix merge conflicts
Added grid and streamline plots
The streamline plot needs an explanation in the paper.

@@
 
-\cite{cavecchi:2013,art-2015-cavecchi-etal,art-2016-cavecchi-etal} \MarginPar{need a summary}
+\cite{cavecchi:2013,art-2015-cavecchi-etal,art-2016-cavecchi-etal,Cavecchi2017,Cavecchi2019} \MarginPar{need a summary}
 
@@
 atmosphere (leaving the space between the top of the atmosphere and
-upper boundary at low resolution).  In addition to the base grid, we
-use up to 3 refinement levels, the first one a factor of 4 finer than
-the previous and the remaining each a factor of 2 finer than the
-previous.  \castro\ subcycles in time, so the finer grids are evolved
-with a finer timestep than the coarse grids.  Occasionally, the
-timestep chosen at the start of a cycle will violate the CFL condition
-during the advancement of the finer grids.  In this case, we restart the
-finer grid evolution with a smaller timestep, subcycling within the
-larger timestep hierarchy.  We use a CFL number of 0.8 for our
-simulations.  
+upper boundary at low resolution). As seen in Figure \ref{fig:grids}, 
+we use up to 3 refinement levels in addition to the base grid, the first
+one a factor of 4 finer than the previous and the remaining each a
+factor of 2 finer than the previous. \castro\ subcycles in time, so the 
+finer grids are evolved with a finer timestep than the coarse grids. 
+Occasionally, the timestep chosen at the start of a cycle will violate
+the CFL condition during the advancement of the finer grids. In this
+case, we restart the finer grid evolution with a smaller timestep,
+subcycling within the larger timestep hierarchy.  We use a CFL number
+of 0.8 for our simulations.  
 % \MarginPar{true for all?} 
@@
 simulation, because of the increase in the scale height of the atmosphere
-behind the flame.  \MarginPar{add a pic of the grid to show refinement}
+behind the flame.
+
+\begin{figure}[t]
+	\plotone{grids.pdf}
+	\caption{\label{fig:grids} Slice through a simulation showing the four-level
+		grid structure. Note that the boxes shown are not the simulation zones, which
+		span $10~\mathrm{cm}$ at the finest level, but subdomains containing
+		approximately equal numbers of zones that are each managed by a separate
+		MPI process. A coarse base grid extending to the upper boundary is drawn
+		in white, with the green, blue, and black grids showing the jumps in
+		refinement needed to fully resolve the fuel layer and underlying neutron star.}
+\end{figure}
 
@@
 \plottwo{flame.pdf}{speed.pdf}
-\caption{\label{fig:flame} Time-evolution of the 10$\times$ boosted 1D laminar flame.
-The left plot shows temperature and nuclear energy generation profiles at 11 different time points, while the right plot shows flame propagation speed and flame thickness as functions of time.}
+\caption{\label{fig:flame} Time-evolution of the 10$\times$ boosted 1D
+	laminar flame. The left plot shows temperature and nuclear energy
+	generation profiles at 11 different time points, while the right plot
+	shows flame propagation speed and flame thickness as functions of time.}
 \end{figure*}
@@
 To measure the propagation rate of the burning front, we track the
-location beyond the hottest part of the flame \MarginPar{what does this mean?} where
+location beyond the hottest part of the flame\MarginPar{what does this mean?} where
 $\dot{e}_\mathrm{nuc}$ first drops to less than $0.1 \%$ of its
@@
 
-Show the phase plots and whatever it was that Don cooked up.
+{\color{red} Show whatever it was that Don cooked up. Also need to discuss the streamline plot.}
 
@@
 $\ln \left(\mathbf{\psi}\right)$ at time $t = 0.02~\mathrm{s}$ for the 10/10 simulation.}
+\end{figure}
+
+\begin{figure}[t]
+	\centering
+	\plotone{streamlines.pdf}
+	\caption{\label{fig:streamlines} Streamlines showing velocity field at time $t = 0.01~\mathrm{s}$ for the 10/10 simulation.}
 \end{figure}

### 2019-11-16 xrb2/paper.tex
paper II stub

--- 2019-10-19 xrb/paper.tex
@@
 %======================================================================
-\title{Dynamics of Laterally Propagating Flames in X-ray Bursts}
+\title{Dynamics of Laterally Propagating Flames in X-ray Bursts. II. Realistic Burning}
 
-\shorttitle{Lateral Flame Dynamics}
-\shortauthors{Eiden et al.}
+\shorttitle{Lateral Flame Dynamics II}
+\shortauthors{Harpole et al.}
+
+\author[0000-0002-1530-781X]{Alice Harpole}
+\affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
+             Stony Brook, NY 11794-3800}
 
@@
 \author[0000-0001-8401-030X]{Michael Zingale}
-\affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
-             Stony Brook, NY 11794-3800}
-
-\author[0000-0002-1530-781X]{Alice Harpole}
 \affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
@@
 
-\correspondingauthor{Michael Zingale}
-\email{Michael.Zingale@stonybrook.edu}
+\correspondingauthor{Alice Harpole}
+\email{alice.harpolee@stonybrook.edu}
 
@@
 \begin{abstract}
-We investigate the structure of laterally-propagating flames through
-the highly-stratified burning layer in an X-ray burst.  Simple \MarginPar{is simple the right word here?}
-two-dimensional simulations of flame propagation are performed through
-a rotating plane-parallel atmosphere, exploring the structure of the
-flame.  We discuss the approximations needed to capture the length and
-time scales at play in an X-ray burst.  Our studies complement the
-other multidimensional studies of burning in X-ray bursts.
+We continue to investigate laterally propagating flames in XRBs...
 \end{abstract}
@@
 
-X-ray bursts (XRBs) are thermonuclear explosions of an accreted H or
-He layer on the surface of a neutron star (see~\citealt{galloway:2017}
-for a review).  Observations of bursts can help constrain the
-properties of the underlying neutron star, helping to illuminate the
-nuclear equation of state~\citep{steiner:2010,ozel2016masses}.  Extensive
-observations of brightness oscillations during the rise have provided
-evidence that the burning begins in a localized region and spreads
-over the surface of the neutron
-star~\citep{bhattacharyya:2006,bhattacharyya:2007,chakraborty:2014}.
-
-One dimensional studies of XRBs are very successful in predicting the
-lightcurves and recurrance times (see, e.g., \citealt{woosley-xrb}).
-These assume spherical symmetry and thus cannot capture the effects of
-localized burning spreading across the neutron star.  These studies
-have also been used to explore the sensitivity of the burst
-observables to accretion and reaction
-rates~\citep{cyburt:2010,Jose2010a,Lampe2016}, and model individual
-bursts~\citep{johnston:2019}.
-
-Multidimensional simulations of burning on a neutron star are more
-difficult, with both the temporal and spatial scales presenting
-challenges (see \citealt{astronum_2018} for an overview).  For the spatial scales, we need to
-resolve the reaction zone, $\mathcal{O}(10~\mbox{cm})$ or smaller, the
-scale height of the atmosphere, $\mathcal{O}(500~\mbox{cm})$, and the
-Rossby scale where the Coriolis force balances the lateral pressure
-gradient, $\mathcal{O}(10^5~\mbox{cm})$~\citep{spitkovsky2002}.  For
-the temporal scales, capturing the rise, $\mathcal{O}(1~\mbox{s})$,
-and the decay of the burst, $\mathcal{O}(10~\mbox{s})$, and the
-accretion period between bursts, $\mathcal{O}(10^4~\mbox{s})$ is
-beyond the ability of multidimensional hydro codes currently.
-Nevertheless, significant progress has been made in understanding the
-multidimensional nature of XRBs, through various approximations.x
-
-Laterally propagating detonations were modeled by
-\citet{fryxellwoosley82,hedet}, but these would only be applicable at
-very high densities, since it is difficult to detonate helium at the
-densities found in normal XRBs, and even harder to detonate hydrogen
-because of the waiting times for weak reactions.  This means that
-detonations may only really be applicable to superbursts (where carbon
-is the reactant)~\citep{Weinberg2006b,Weinberg2007}.
-
-Global multidimensional studies were performed by
-\citet{spitkovsky2002}, where it was demonstrated that the Coriolis
-force plays an important role in confining the burning as it spreads
-across the neutron star surface.  These calculations used the
-shallow-water approximation, so the vertical details of the
-atmosphere's structure we not modeled.  The model showed that the
-horizontal pressure gradient between the ash and fuel can be important
-in accelerating the burning front.
-
-Small-domain studies of convective burning in XRBs preceding flame
-development have been done in two-dimensions \citep{lin:2006,xrb,xrb2}
-and three-dimensions \citep{xrb3d}.  These calculations used low Mach
-number methods that approximate the hydrodynamics equations to filter
-soundwaves, enabling large timesteps and efficient modeling of
-subsonic convection.  While these calculations could not support
-the lateral differences needed for flame spreading, they can help
-understand the role that convection plays in distributing the initial
-burning products vertically throughout the neutron star atmosphere as well
-as the nature of any turbulence the burning front might encounter.
-
-\cite{cavecchi:2013,art-2015-cavecchi-etal,art-2016-cavecchi-etal} \MarginPar{need a summary}
-
-For deflagrations, we either need to resolve the structure of the
-reaction zone or use a flame model.  Flame models usually assume that
-the flame structure is thin compared to the size of the system (see,
-e.g., \citet{Ropke2007} for applications to Type Ia supernovae).  For
-XRBs however, the flame thickness is comparable to the scale height of
-the atmosphere, so we cannot use these approximations.  Accurate
-models of flames therefore require that we resolve the thermal width,
-which is $\mathcal{O}(10~\mbox{cm})$ for helium flames~\citep{Timmes00}.
-
-The goal of this study is to understand what numerical and physical
-approximations are needed to perform a full hydrodynamical,
-multidimensional simulation of flame propagation through the
-atmosphere on a neutron star.  For simplicity in this first set of
-calculations, we will use pure helium composition.  These studies
-complement the other multidimensional work described above in helping
-us to build a picture of the dynamics of X-ray bursts.
-
 %======================================================================
@@
 
-All simulations are performed with the \castro\ hydrodynamics
-code~(\citealt{castro}; see also \citealt{astronum:2017} for a recent
-description).  We evolve the system of fully compressible Euler
-equations for reacting flow:
-\begin{eqnarray}
-\frac{\partial( \rho X_k)}{\partial t} &=& - \nabla\cdot (\rho \Ub X_k) + \rho \omegadot_k \\
-%
-\frac{\partial (\rho \Ub)}{\partial t} &=& - \nabla\cdot (\rho \Ub \Ub) - \nabla p +
-    \rho \gb \nonumber \\
-  &&-2 \rho \Omegab\times \Ub - \Omegab \times (\Omegab \times \rb) \\
-%
-\frac{\partial (\rho E)}{\partial t} &=& - \nabla \cdot (\rho \Ub E + p\Ub) +
-    \nabla \cdot \kth \nabla T + \rho \epsdot +\nonumber\\
-  && \rho \Ub \cdot \gb - \rho (\Omegab \cdot \rb)(\Omegab \cdot \Ub) + \rho |\Omegab|^2 (\Ub \cdot \rb)
-\end{eqnarray}
-Here, $\rho$ is the mass density, $\Ub$ is the velocity, and $E$ is
-the specific total energy, which is related to the specific internal
-energy as $e = E - |\Ub|^2/2$.  Species are described by mass
-fractions, $X_k$ (such that $\sum_k X_k = 1$), and creation rates,
-$\omegadot$, and are related to the total specific energy generation
-rate, $\epsdot$.  The total mass conservation,
-\begin{equation}
-\frac{\partial \rho}{\partial t} = - \nabla\cdot (\rho \Ub)
-\end{equation}
-implies $\sum_k \omegadot_k = 0$.  \castro\ uses an unsplit piecewise
-parabolic method (PPM) with characteristic tracing for solving the
-hydrodynamics~\citep{ppm,millercolella:2002}, generalized to an
-arbitrary equation of state~\citep{zingalekatz}.  Reactions are
-incorporated via Strang splitting~\citep{strang:1968}, giving a method
-that is overall second-order accurate in space and time.
-\castro\ uses the \amrex\ adaptive mesh refinement
-library~\citep{amrex_joss} to manage a hierarchy of grids at different
-resolutions.
-
-An equation of state of the form $p = p(\rho, e, X_k)$ completes the
-thermodynamic description of the system.  Finally, $\kth$ is the
-thermal conductivity and $T$ is the temperature.  Diffusion is treated
-explicitly in time.
-
-Since the neutron star rotates, we work in a corotating frame, with
-angular velocity $\Omegab$.  Further, for the two-dimensional
-simulations presented here, we work in axisymmetric coordinates,
-but we advect a third component of velocity, coming out of the
-simulation plane, that participates in the Coriolis force (sometimes
-described as a 2.5D simulation).  We will take the
-\castro\ $x$-coordinate to be the cylindrical radial coordinate with
-corresponding velocity $u$, the \castro\ $y$-coordinate to be the
-cylindrical vertical coordinate with corresponding velocity $v$, and the
-\castro\ $z$-coordinate to be the cylindrical azimuthal coordinate,
-with corresponding velocity $w$.  A righthanded coordinate system has
-positive $w$ pointing out of the page.  We will take $\Omegab =
-\Omega_0 \hat{\bf y}$ for the angular rotation rate, and $\gb = -g
-\hat{\bf y}$ for the gravitational acceleration, with $g$ constant.
-With these choices, the Coriolis force is:
-\begin{equation}
--2\rho \Omegab \times \Ub =
-   -2\rho \left ( \Omega_0 w \hat{x} - \Omega_0 u \hat{z} \right )
-\end{equation}
-We will neglect the centrifugal force---with our plane-parallel
-geometry, this will act only in the lateral direction, and is not
-expected to greatly affect the dynamics.  Carrying the Coriolis force
-allows us to capture the geostrophic balance\MarginPar{ref, Spitkovsky 2002?} that sets
-up via lateral hydrostatic equilibrium.  In the discussions below, we will use the Castro coordinate
-names, $(x, y)$ in our notation.
-
-Writing the momentum equation in terms of the $u$, $v$, and $w$
-components, and neglecting the centrifugal force, we have:
-\begin{align}
-\frac{\partial (\rho u)}{\partial t} + \nabla \cdot (\rho u \Ub) +
-     \frac{\partial p}{\partial x} &= -2\rho \Omega_0 w  \\
-%
-\frac{\partial (\rho v)}{\partial t} + \nabla \cdot (\rho v \Ub) +
-     \frac{\partial p}{\partial y} &= -\rho g \\
-%
-\frac{\partial (\rho w)}{\partial t} + \nabla \cdot (\rho w \Ub) +
-     \cancelto{0}{\frac{\partial p}{\partial z}} &=
-    2\rho \Omega_0 u
-\end{align}
-where we cancel $\partial p/\partial z$ because there are no
-variations in the azimuthal direction.  This allows us to recast the
-$w$-velocity equation as a simple advection equation:
-\begin{equation}
-\frac{\partial w}{\partial t} + \Ub \cdot \nabla w = 2\Omega_0 u
-\end{equation}
-In our geometry, the flame will propagate from left to right, so $u$
-will be positive and the Coriolis force results in $w > 0$ (out of the
-simulation plane).  The algorithmic implementation of rotation in
-\castro\ is described in \cite{wdmergerI}.
-
-We use a general stellar equation of state with nuclei (treated as an
-ideal gas), photons, and degenerate/relativistic electrons, as
-described in \cite{timmes_swesty:2000}.  For reactions, we use a 13
-isotope alpha chain, derived from the {\tt aprox13}
-network~\citep{timmes_aprox13}.  For one of our runs, we use the
-smaller 7-isotope network described in \citet{iso7}.  We integrate the
-network using the VODE integration package~\citep{vode}, and our
-implementation is provided in the StarKiller Microphysics
-source~\citep{starkiller}.
-We note that we do not explicitly model viscosity.  The reactions will
-provide the smallscale cutoff to the instabilities and turbulence at
-the flame front.  We also do not include species
-diffusion---astrophysical flames tend to have large Lewis numbers, so
-this is not expected to be important~\citep{timmeswoosley:1992}.
-Finally, we use the thermal conductivities described in
-\citet{Timmes00}.
-
-All simulations use adaptive mesh refinement to refine on the
-atmosphere (leaving the space between the top of the atmosphere and
-upper boundary at low resolution).  In addition to the base grid, we
-use up to 3 refinement levels, the first one a factor of 4 finer than
-the previous and the remaining each a factor of 2 finer than the
-previous.  \castro\ subcycles in time, so the finer grids are evolved
-with a finer timestep than the coarse grids.  Occasionally the
-timestep chosen at the start of a cycle will violate the CFL condition
-during the advance of the finer grids.  In this case, we restart the
-finer grid evolution with a smaller timestep, subcycling within the
-larger timestep hierarchy.  We use a CFL number of 0.8 for our
-simulations.  
-% \MarginPar{true for all?} 
-The base grid for our standard
-simulations is $768\times 192$ zones and the corresponding finest grid
-when 3 refinement levels are added would be $12288\times 3072$ zones.
-Our standard domain is $1.2288\times 10^5~\mathrm{cm} \times
-3.072\times 10^4~\mathrm{cm}$, so we have $10~\mathrm{cm}$ resolution
-on the finest grid.  We only refine the fuel layer in the left half of
-the domain at the highest resolution (and only down to densities of
-$2.5\times 10^4~\gcc$), since this is where we expect the flame to
-propagate.  At the start of the simulation, $3.4\%$ of the domain is
-at the finest resolution.  This increases to $7.4\%$ by the end of the
-simulation, because of the increase in the scale height of the atmosphere
-behind the flame.  \MarginPar{add a pic of the grid to show refinement}
-
-Thermal diffusion is modeled explicitly, using a predictor-corrector
-scheme to achieve second-order accuracy.  A verification test of the
-diffusion scheme is shown in Appendix~\ref{app:diffusion}.  The
-explicit thermal diffusion requires a timestep limiter of the form:
-\begin{equation}
-\Delta t_\mathrm{diff} \le \frac{1}{2} \frac{\Delta x^2}{\mathcal{D}}
-\end{equation}
-where $\mathcal{D} = \kth/(\rho c_v)$ is the thermal diffusivity.  The
-diffusivity increases rapidly at the top of the atmosphere, making
-these low density regions determine the overall timestep for the
-simulations.  Therefore, we disable thermal conduction at low
-densities where it is not expected to be important.
-
-We use hydrostatic boundary conditions on the lower boundary, using
-a discretized hydrostatic equilibrium equation of the form:
-\begin{equation}
-\label{eq:hse}
-p_i = p_{i-1} + \frac{1}{2} \Delta y (\rho_i + \rho_{i-1}) g
-\end{equation}
-holding the temperature constant in the ghost cells.  This is solved
-together with the equation of state.  The velocity is reflected here.
-This procedure follows the form described in \citet{ppm-hse}.  The
-left boundary is reflecting and the right boundary is a zero-gradient
-outflow.  The top boundary sets the state to simply the conditions
-in our outer buffer region of the initial model (see below) with the
-normal velocity set to the larger of zero or the velocity at the top
-of the domain (this prevents incoming velocities at the top) and the
-transverse velocities set to zero.
-
-When we begin the simulation, there is a transient phase as the flame
-gets established.  Material that is forced upward will encounter the
-steep density gradient at the top of the atmosphere and accelerate as
-it is blown out of the atmosphere.  Eventually this material will fall
-back to the top of the atmosphere.  In this paper, we are mostly
-concerned with the behavior of the flame and not any material that is
-violently blown out of the top of the atmosphere, so we apply a sponge
-to this region.  This is similar to what we've done in \cite{xrb3d}.
-The sponge drives the velocity of the material in the low
-density \MarginPar{add mathematical form of sponge} regions at the top
-of the atmosphere to zero.  This sponging helps increase our timestep
-as well.
-
-%======================================================================
-% Flame
-%======================================================================
-\section{Flame Properties}\label{Sec:Flame}
-
-The speed and thickness of a laminar helium flame are determined by the
-energy generation rate and conductivity, and scale roughly as 
-\begin{equation}
-v_f \approx \sqrt{k_\mathrm{th} \dot{\epsilon}} \qquad
-\lambda_f \approx \sqrt{\frac{k_\mathrm{th}}{\dot{\epsilon}}}
-\end{equation}
-\citep{orourke:1979,khokhlov:1993}.
-At the densities we consider in this simulation, the pure He flame
-speed is quite slow and would require a prohibitively long integration
-time to see significant evolution of the burning.  We therefore
-consider boosted flames in this first paper, to accelerate the burning
-and allow us to understand the qualitative effects of laterally
-propagating flames.  To boost the flame, while keeping the thickness
-the same, we can multiply both the burning rate and the conductivity
-by the same factor.  For our standard calculations, we choose 10 for
-each, to give a flame 10$\times$ faster.  We call this the ``10/10''
-flame.  We will also do a simulation with the reactions and
-conductivity both boosted by 5, the ``5/5'' flame.
-
-To understand the time and length scales involved in flame
-propagation, we do a 1D simulation of a laminar flame using our
-microphysics.  Figure~\ref{fig:flame} shows the flame properties for a 
-% \AssignTo{Kiran and Mike will remake with the 10/10 boosted flame}
-He flame using the standard conductivities and our simple reaction \MarginPar{standard?  this is a boosted flame, right?}
-network.  This flame had a density of $2\times 10^6~\gcc$ and 
-% \MarginPar{double check}
-temperature of $5\times 10^7$~K.  We observe that this flame speed is
-about $10^5~\cms$, the flame width is about 40~cm, and it
-takes about 3~ms to settle into a sustained flame.  Note that this speed
-is quite small compared to the speed estimated in
-\citet{spitkovsky2002}.  We expect a multidimensional flame to
-accelerate due to hydrodynamics interactions (wrinkling, turbulence
-interactions, directed flows feeding fuel into the flame, ...).
-
-\begin{figure*}[t]
-\plottwo{flame.pdf}{speed.pdf}
-\caption{\label{fig:flame} Time-evolution of the 10$\times$ boosted 1D laminar flame.
-The left plot shows temperature and nuclear energy generation profiles at 11 different time points, while the right plot shows flame propagation speed and flame thickness as functions of time.}
-\end{figure*}
-
-We measure the flame width as:
-\begin{equation}
-\lambda_f \equiv \frac{\Delta T}{\max\{|\nabla T|\}}
-\end{equation}
-Experience with modeling resolved flames suggests that we need a
-spatial resolution, $\Delta x$ of $\lambda_f/\Delta x \sim 5$~\citep{SNld}.  These
-conditions represent the bottom of the He layer.  As the density
-decreases with altitude, the flame thickness increases and the flame
-speed decreases, so we will easily resolve the flame structure throughout
-the rest of the atmosphere.
-
-Flame propagation rates from both the 1D laminar flames and full 2D simulations are given in Table \ref{table:flame_speeds}.
-\MarginPar{I think we should only report the laminar speeds here and defer the turbulent speeds until the section where we talk about how they are computed}
-
-\begin{deluxetable}{lcc}
-	\tablecaption{\label{table:flame_speeds}Turbulent and laminar flame speeds.}
-	\tablehead{\colhead{run} & \colhead{$s_T$ (km s$^{-1}$)} & \colhead{$s_L$ (km s$^{-1}$)}}
-	\startdata
-	10/10 boost & $9.18 \pm 0.03$  & $1.06 \pm 0.01$ \\
-	5/5 boost & $4.00 \pm 0.01$ & $0.56 \pm 0.01$ \\
-	iso7 network & $7.56 \pm 0.02$ & $-$ \\
-	lowres & $9.33 \pm 0.04$ & $-$ \\
-	1000 Hz & $18.6 \pm 0.16$ & $-$ \\
-	\enddata
-	%
-\end{deluxetable}
-
-%======================================================================
-% Initial Model
-%======================================================================
-\section{Initial Model}\label{Sec:inital_model}
-
-% \begin{figure}[t]
-% \centering
-% \plotone{model}
-% \caption{\label{fig:sketch} A schematic of our model setup.}
-% \end{figure}
-
-We wish to create an initial atmosphere consisting of a hot
-``post-flame'' region and a cooler atmosphere that the flame will
-laterally propagate into.  We put the hot region at the very left of
-the domain (the origin of the axisymmetric coordinates).  To create
-these initial conditions, we produce two different hydrostatic models,
-a ``hot'' model that will represent the perturbation that drives the
-flame and a ``cool'' model that will represent the state ahead of the
-flame.  These will have different scale heights.  To create these
-models, we break the vertical structure of the atmosphere into 4
-layers: the underlying neutron star, a ramp-up to the base of the
-accreted atmosphere, a fuel layer representing the bulk of the
-atmosphere where the flame will propagate, and an outer, low density,
-isothermal buffer above the atmosphere that allows us
-to capture expansion and explosive dynamics.  
-% This is sketched in
-% Figure~\ref{fig:sketch}.
-
-The temperature profile in the star and ramp region is given as:
-\begin{equation}
-T = T_\star + \frac{1}{2} (T_\mathrm{hi} - T_\star) \left [ 1 + \tanh\left( \frac{\tilde{y}}{2 \delta_\mathrm{atm}} \right ) \right ]
-\end{equation}
-with
-\begin{equation}
-\tilde{y} = y - H_\star - \frac{3}{2} \delta_\mathrm{atm}
-\end{equation}
-Here, $\delta_\mathrm{atm}$ is a characteristic width of the
-transition ramp, $T_\star$ is the temperature of the neutron star, and
-$T_\mathrm{hi}$ is the highest temperature in the HSE model---it will
-represent the base of the fuel layer.
-
-The mass fractions use this same profile, switching from a set
-describing the underlying star, ${X_k}_\star$, and the set for the
-accreted material, ${X_k}_\mathrm{atm}$, which is used in the
-isentropic and outer regions.  Note, since the profile above is
-linear in $X$, if the initial mass fractions sum to one, then the blended
-mass fractions in the ramp region also sum to one.
-
-We specify the density, $\rho_\mathrm{int} = \rho(y = H_\star)$ as the
-starting point for the integration of hydrostatic equilibrium.  This
-is just below the ramp-up region---this ensures that regardless of
-what the peak temperature ($T_\mathrm{hi}$) is, the state beneath the
-ramp-up region remains unchanged.  Therefore, we will still be in
-lateral equilibrium in the star region.  We will denote the density
-where $T = T_\mathrm{hi}$ as $\rho_\mathrm{fuel}$.
-
-Creating the model involves specifying $T_\star$, $T_\mathrm{hi}$,
-$T_\mathrm{lo}$, $\rho_\mathrm{int}$, $H_\star$,
-$\delta_\mathrm{atm}$, ${X_k}_\star$, and ${X_k}_\mathrm{atm}$.  We
-then integrate outwards from the base of the ramp region ($y =
-H_\star$), enforcing the discrete form of hydrostatic equilibrium,
-Eq.~\ref{eq:hse}.  Integrating upwards, we would find $p_i$ and
-$\rho_i$ using a Newton solve together either with the temperature
-specified, $T_i = T(p_i, \rho_i, \{X_k\}_i)$ (in the isothermal, ramp,
-and buffer layers) or constant entropy, $s_i = s(p_i, \rho_i,
-\{X_k\}_i)$, in the fuel layer.  In each case, using the equation of
-state.  This follows the procedures described in \citet{ppm-hse}.  We
-use the constant temperature for all $y < H_\star +
-3\delta_\mathrm{atm}$.  Above this, we switch to isentropic until the
-temperature drops to a floor value, $T_\mathrm{lo}$, at which point we
-again keep the temperature constant.  The integration of the
-atmosphere continues until the density falls to a low density cutoff,
-$\rho_\mathrm{cutoff}$.  The material above this height is taken to
-have constant density and temperature.
-
-\begin{deluxetable}{lcc}
-\tablecaption{\label{table:params} Initial model parameters.}
-\tablehead{\colhead{parameter} & \colhead{cool} & \colhead{hot}}
-\startdata
-$T_\star$       & \multicolumn{2}{c}{$10^8$~K} \\
-$T_\mathrm{hi}$ & $2\times 10^8$~K & $1.4\times 10^9$~K \\
-$T_\mathrm{lo}$ & \multicolumn{2}{c}{$8\times 10^6$~K} \\
-$\rho_\mathrm{int}$ & \multicolumn{2}{c}{$3.43\times 10^6~\gcc$} \\
-$\rho_\mathrm{fuel}$\tablenotemark{a} & $2.36\times 10^6~\gcc$ & $1.20\times 10^6~\gcc$ \\
-$\rho_\mathrm{cutoff}$ & \multicolumn{2}{c}{$10^{-4}~\gcc$} \\
-$g$             & \multicolumn{2}{c}{$-1.5\times 10^{14}~\mathrm{cm~s^{-2}}$} \\
-$H_\star$       & \multicolumn{2}{c}{2000~cm} \\
-$\delta_\mathrm{atm}$ & \multicolumn{2}{c}{50~cm} \\
-$X_\star(\isotm{Ni}{56})\tablenotemark{b}$ & \multicolumn{2}{c}{1.0} \\
-$X_\mathrm{atm}(\isotm{He}{4})\tablenotemark{b}$ & \multicolumn{2}{c}{1.0} \\
-\enddata
-%
-\tablenotetext{a}{This is not an input parameter, but instead is
-  computed during integration.  We list it here for reference.}
-%
-\tablenotetext{b}{All other species are taken as 0.}
-\end{deluxetable}
-
-The choice of factors in front of $\delta_\mathrm{atm}$ were designed
-to make sure the peak $T$ is attained at the desired density
-of the burning layer.
-
-We create two models, a ``cool'' model representing the atmosphere
-ahead of the flame and a ``hot'' (or perturbed) model that will
-initiate the flame by driving the burning.  The parameters we use for
-the model generation are listed in Table~\ref{table:params}. \MarginPar{reference and discuss the figure of initial model here}
-
-We blend the hot and cold models laterally to produce the perturbation
-needed to initiate a localized flame, with the hot model at the
-origin of the axisymmetric geometry.  The blending is done as:
-\begin{align}
-p(x,y) &= f(x) p_\mathrm{hot}(y) + [1-f(x)] p_\mathrm{cool}(y) \\
-\rho(x,y) &= f(x) \rho_\mathrm{hot}(y) + [1-f(x)] \rho_\mathrm{cool}(y) \\
-X_k(x,y) &= f(x) {X_k}_\mathrm{hot}(y) + [1-f(x)] {X_k}_\mathrm{cool}(y)
-\end{align}
-with
-\begin{equation}
-f(x) = \begin{cases}
-     1 & x < x_\mathrm{pert} \\
-   1 - \frac{x - x_\mathrm{pert}}{\delta_\mathrm{blend}} & x_\mathrm{pert} \le x \le x_\mathrm{pert} + \delta_\mathrm{blend} \\
-     0 & x > x_\mathrm{pert} + \delta_\mathrm{blend}
-\end{cases}
-\end{equation}
-Since the equation of hydrostatic equilibrium is linear and our
-blending is a linear combination of two models in hydrostatic
-equilibrum, the blended model is also in vertical equilibrium initially.  We choose
-$x_\mathrm{pert} = 1.024\times 10^4$~cm and $\delta_\mathrm{blend} = 2048$~cm.  Once \MarginPar{double check these numbers}
-the blended model is constructed, we compute $T(x,y)$ and $(\rho e)(x,y)$
-from the equation of state,
-\begin{align}
-  T(x,y) &= T(\rho(x,y), p(x,y), X_k(x,y)) \\
-  (\rho e)(x,y) &= \rho(x, y) \cdot e(\rho(x,y), p(x,y), X_k(x,y)) 
-\end{align}
-
-\begin{figure}[t]
-\centering
-\epsscale{0.75}
-\plotone{initial_model_paper}
-\caption{\label{fig:initial_models} Our ``cool'' (solid) and ``hot''
-  (dashed) initial models, showing both the density and temperature.}
-\epsscale{1.0}
-\end{figure}
-
-The initial model is created on a uniform grid at the resolution
-corresponding to the finest level of refinement.  At regions of the
-atmosphere that are not refined to the finest level, we interpolate
-density, pressure, and composition on the grid and then obtain the
-temperature from the EOS.
-
 %======================================================================
@@
 \section{Simulations and Results}\label{Sec:results}
-
-Table~\ref{table:sim_names} summarizes the simulations that we ran.
-The majority of them used a reaction rate boosting of $10$ and a
-conductivity boosting of $10$, which should increase the flame speed
-by a factor of $10$.  Most simulations used a resolution of
-$10~\mbox{cm}$ and a domain width of a little more than one kilometer.  
-We use an artifically high rotation rate of
-$2000~\mbox{Hz}$, which gives a Rossby length of
-\begin{equation}
-L_R = \frac{\sqrt{g H_0}}{\Omega} \sim 3\times 10^4~\mathrm{cm}
-\end{equation}
-using a scale height $H_0 = 10^3~\mathrm{cm}$.  This is about one
-quarter of domain width.  The run at $20~\mbox{cm}$ resolution used
-one fewer level of refinement.  The slower rotating case using a
-slightly wider domain to accommodate the expected larger Rossby
-length.  We note that the entire simulation framework for these
-calculations is freely available in the \castro\ github
-repository\footnote{\url{https://github.com/AMReX-Astro/Castro}, using
-  the {\tt flame\_wave} setup.}.  In the discussions below, we'll use
-the simulation names defined in the table to refer to specific runs
-and we'll use the 10/10 run as the reference calculation.
-
-\begin{deluxetable}{lcccccc}
-\tablecaption{\label{table:sim_names} Simulation parameters.}
-\tablehead{\colhead{name} &
-           \colhead{reaction} &
-           \colhead{conductivity} &
-           \colhead{fine grid} &
-           \colhead{rotation} &
-           \colhead{domain size} &
-           \colhead{network} \\
-%
-           \colhead{} &
-           \colhead{boost} &
-           \colhead{boost} &
-           \colhead{resolution} &
-           \colhead{rate} &
-           \colhead{} &
-           \colhead{}
-}
-\startdata
-10/10 & 10 & 10 & 10~cm & 2000~Hz & $1.2288\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$ & {\tt aprox13} \\
-10/10-slow & 10 & 10 & 10~cm & 1000~Hz & $1.8432\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$ &{\tt aprox13} \\
-10/10-lores & 10 & 10 & 20~cm & 2000~Hz & $1.2288\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$ &{\tt aprox13} \\
-5/5 & 5 & 5 & 10~cm & 2000~Hz & $1.2288\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$  &{\tt aprox13} \\
-10/10-iso7 & 10 & 10 & 10~cm & 2000~Hz & $1.2288\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$ &{\tt iso7} \\
-\enddata
-\end{deluxetable}
-
-\subsection{General Features}
-
-Figure~\ref{fig:time_series} shows the time evolution of the 10/10
-simulation, focusing on the mean-molecular weight,
-\begin{equation}
-\bar{A} = \left ( \sum_k \frac{X_k}{A_k} \right )^{-1}
-\end{equation}
-In each frame the buffer of \isot{Ni}{56} that serves as the
-underlying neutron star is seen spanning the bottom of the domain.
-Above that, the composition begins as pure \isot{He}{4}, but as the
-simulation progresses, the flame processes this to heavier nuclei,
-increasing $\bar{A}$.  By about $10~\mbox{ms}$, we see the flame front
-is reasonable well-defined.  We see that the bottom of the burning
-front is lifted off of the base of the atmosphere, greatly increasing
-the surface area of the burning compared with a perfectly vertical
-flame front.  By $20~\mbox{ms}$ the flame has moved out substantially
-and we are beginning to see a gradient in the composition of the ash,
-with the heavier nuclei furthest behind the flame.  The boosting of
-the burning likely artificially increases this effect. \MarginPar{Doesn't the increased flame speed of the 10/10 run vs. the 5/5 run show this?}
-
-% made with time_series.py in flame_wave_boost_10_10_paper with --skip 100
-\begin{figure}[t]
-\centering
-\plotone{time_series}
-\caption{\label{fig:time_series} Time series of the mean molecular weight of the 
-flame for our standard 10/10 simulation.}
-\end{figure}
-
-Figure~\ref{fig:10_10_overview} shows the temperature, energy
-generation rate, and $w$ component of the velocity (the out-of-plane
-velocity induced by the Coriolis force).  This latter field
-illustrates the hurricane effect set up by the laterally spreading
-burning front.  In the energy generation rate plot, we see that the
-burning is mostly concentrated toward the bottom of the layer, as
-expected since the density is greatest there.  We see that the peak of
-the burning has moved off of the symmetry axis, demonstrating that the
-burning front is propagating to the right.  In the temperature plot,
-we see the effect of our refinement criteria focusing only on the
-atmosphere where we are most dense, with an artificial change in the
-temperature at the refinement boundary due in part to the construction
-of the initial model using the fine grid resolution for HSE. \MarginPar{Remind the reader what this
-  was/where they should be looking in the plot to see this temperature
-  drop} As we'll see when the lower resolution case, we do not believe
-that this affects the results.
-
-A final feature worth noting is the ash that seems to move along the
-surface at a higher velocity than the flame, via surface gravity
-waves.  The sponging that we perform is likely damping this to some extent, and
-the method by which we initialize the flame may induce a larger
-transient than in nature, nevertheless, this surface ash is intriguing
-because it affects the composition of the photosphere ahead of the
-burning front, potentially changing our interpretation of
-observations.  This is something that will be explored more fully in
-the future.
-
-% made with overview.py
-\begin{figure}[t]
-\centering
-\plotone{flame_wave_boost_10_10_overview}
-\caption{\label{fig:10_10_overview} Temperature, energy generation rate, and out-of-plane velocity for the 10/10 simulation at 20~ms.}
-\end{figure}
-
-Our default resolution puts $\sim 80$ zones vertically in the
-``cool'' \MarginPar{double check} model atmosphere height.  To
-understand the effects of resolution, we also performed a run with one
-fewer level, giving a $20~\mbox{cm}$ resolution overall (this is our
-10/10-lores run).  Figure~\ref{fig:10_10_lowres} shows the fields at
-20~ms.  The structure is largely the same as the 10/10 run, with
-largely the same flame shape and position, and the same structure in
-the energy generation rate.  Since we do not have a jump in refinement
-right below the atmosphere, we don't see the temperature feature from
-the initial model mapping there, but we do see some cooling in the
-\isot{Ni}{56} region.  This is likely a resolution effect, again due
-to the strong gradient in temperature at the base of the atmosphere.
-The strong agreement in the low resolution run to the 10/10 run gives
-us confidence that we are capturing the flame physics properly.
-
-% made with slice_multi_crop.py
-\begin{figure}[t]
-\centering
-\plotone{flame_wave_boost_10_10_lowres_slice}
-\caption{\label{fig:10_10_lowres} Temperature, mean molecular weight,
-  energy generation rate, and out-of-plane velocity for the 10/10 low
-  resolution simulation at 20~ms.}
-\end{figure}
-
-\subsection{Effects of our Approximations}
-
-The results above all used a boosting of 10/10.  To see how the
-results are sensitivity to our boosting we ran a simulation with a
-boosting of 5/5.  This is shown in Figure~\ref{fig:5_5_overview}.  The
-results look qualitatively the same---a laterally propagating flame
-develops that is lifted off of the bottom of the fuel layer.  The
-flame has not moved out as far as in the 10/10 simulation, simply
-because there is less energy release, but we expect that if we were to
-run this out twice as long, the flame would have advanced to the
-position seen in our 10/10 runs.  The good agreement in the structure
-of the flame seen with the lower boosting gives us confidence that the
-overall aspects of the flame structure and acceleration we are seeing
-are robust to the approximations we make.
-
-\begin{figure}[t]
-\centering
-\plotone{flame_wave_boost_5_5_slice}
-\caption{\label{fig:5_5_overview} Temperature, mean molecular weight, energy generation rate, and out-of-plane velocity for the 5/5 simulation at 20~ms.}
-\end{figure}
-
-We can also look at the effect of the network size.
-Figure~\ref{fig:network} shows a comparison of the 10/10 boosting run
-with the standard 13-isotope ``aprox13'' network and the reduced
-7-isotope ``iso7'' network.  We see that the flame in the aprox13 case
-is slightly more advanced and has a higher $\bar{A}$ than the iso7
-case.  We see in the next section that these two networks give largely
-the same flame speed.  The reduced network size saves a lot of memory
-which will be important when we transition to 3-d simulations.
-
-\begin{figure}[t]
-\centering
-\plotone{network_compare}
-\caption{\label{fig:network} Mean molecular weight at 15~ms comparing a run with
-  aprox13 vs a run with the iso7 network.}
-\end{figure}
-
-The final approximation to explore is our choice of rotation rate.
-{\color{red} need to add discussion + plot}
-
-\subsection{Flame Acceleration}
-
-To measure the propagation rate of the burning front, we track the
-location beyond the hottest part of the flame \MarginPar{what does this mean?} where
-$\dot{e}_\mathrm{nuc}$ first drops to less than $0.1 \%$ of
-maximum. Using $0.1 \%$ of the peak value rather than the peak value
-itself helps to mitigate fluctuations caused by localized fluid
-motions, as the reduced value lies in a less turbulent part of the
-burning region. We also average $\dot{e}_\mathrm{nuc}$ over the  \MarginPar{I think we should discuss the averaging first, before we discuss the peak value}
-vertical coordinate to further dampen these fluctuations. After an
-initial transient period typically spanning $\sim 3$ ms, the flame
-settles into a state of steady propagation. The position data here are
-well fitted by a linear function of time (as seen in Figure
-\ref{fig:flame_speed}), and the resulting slope gives the velocity of
-the flame front.
-
-Table \ref{table:flame_speeds} gives the flame speed measured in each
-multidimensional simulation, as well as laminar speeds obtained from
-any corresponding 1D runs. The 2D flames propagate at speeds about an
-order of magnitude faster than their 1D counterparts. The increase in
-flame speed is likely a product of hydrodynamical effects such as
-turbulence, wrinkling, and convective cycles, which bring cooler fuel
-from ahead of the front into the hottest part of the burning region.
-
-\begin{figure}[t]
-\centering
-\plotone{plots/results/speedplot_all.pdf}
-\caption{\label{fig:flame_speed}The position of the burning front for each simulation run as a function of time. The dashed lines show linear least squares fits for $t \gtrsim 6$ ms.}
-\end{figure}
-
-{\color{red} Should say qualitatively what the plot shows}
-
-\subsection{Entrainment, Flow Features}
-
-\AssignTo{Alice, Don}
-
-Show the phase plots and whatever it was that Don cooked up.
-
-Baroclinicity is calculated as
-\begin{equation}
-    \boldsymbol{\psi} = \frac{1}{\rho^2} \nablab p \times \nablab \rho,
-\end{equation}
-and shows the misalignment of the local density and pressure gradients
-\citep{Malone2014a}. As we are considering a 2d system, the component of the
-baroclinicity we consider here (out of the plane) reflects the misalignment
-of the fields in the plane of the simulation. Figure~\ref{fig:baroclinicity}
-shows that the baroclinicity peaks along the flame front. This baroclinicity
-generates vorticity, which in turn entrains material along the surface of the
-flame.
-
-\begin{figure}[t]
-\centering
-\plotone{plots/results/baroclinicity}
-\caption{\label{fig:baroclinicity} Baroclinicity. This plot shows
-$\ln \left(\mathbf{\psi}\right)$ at time $t = 0.02~\mathrm{s}$ for the 10/10 simulation.}
-\end{figure}
-
-We show some phase plots in Figure~\ref{fig:phase_plots}. The phase plot of the
-energy generation rate as a function of
-the $r$- and $z$-velocities shows that energy is preferentially generated in
-regions with negative $r$-velocity.
-This most likely corresponds to the burning along the underside of the flame front, where fuel has been entrained along the surface of the flame.
-This is further supported by the phase plot of the energy generation rate as a function of the $r$-velocity
-and the density, which shows that the energy generation rate peaks at
-$\rho \sim 3 \times 10^5~\mathrm{g/cm}^3$, at the base of the flame.
-
-In the first plot, there are ``loops" of
-points of similar $\dot{e}_\mathrm{nuc}$ in the outer edges of the plot.
-These are likely to correspond to the vortices that appear within the flame,
-and show the fluid moving in a circular motion in $u-v$ phase space.
-
-\begin{figure}[t]
-\centering
-\plottwo{plots/results/vx_vy_129139}{plots/results/u_rho_129139}
-\caption{\label{fig:phase_plots} Phase plots at time $t = 0.02~\mathrm{s}$.
-\emph{Left}: Phase plot showing the energy generation rate as a function of
-the $r$- and $z$-velocities. The black cross shows the location of $u = v = 0$. \emph{Right}: phase plot showing the energy
-generation rate as a function of the $r$-velocity and the density.}
-\end{figure}
-
-Figure~\ref{fig:abar_temp} shows the energy generation rate as a function of the
-temperature and $\bar{A}$. 
-The energy generation rate peaks at low $\bar{A}$, where there is a high fraction
-of unburnt material. The burning raises the temperature of this material, such
-that the peak temperature coincides with the peak energy generation rate. The
-material then cools again as the burning converts the fuel into ashes,
-increasing $\bar{A}$ and reducing the energy generation rate as there is
-less available fuel. Along the base of the plot we see cool unburnt fluid and
-ashes. In the center of the plot there are thin `trails' in phase space, which
-could correspond to less common reaction pathways in the reaction network.
-
-\begin{figure}[t]
-\centering
-\plotone{plots/results/abar_temp_129139}
-\caption{\label{fig:abar_temp} Phase plot of energy generation rate as a function of $\bar{A}$ and temperature at time $t = 0.02~\mathrm{s}$.}
-\end{figure}
-
-\section{Discussion}
-
-\AssignTo{everyone}
-
-We have shown the results of our fully hydrodynamical, multidimensional simulations of flame propagation through the atmosphere of a neutron star.
-By using the fully compressible hydrodynamics equations, we are able to capture the vertical dynamics of the system.
-To accurately model the flame propagation, it is necessary to sufficiently resolve the scale height of the atmosphere.
-It is also important that there is a thin transition between the underlying neutron star and the atmosphere, to ensure that the peak temperature occurs at the correct base density.
-
-In order to satisfy these requirements whilst allowing the simulations to be computationally feasible, we used several approximations: a higher than normal rotation rate, a boosted flame, a simplified reaction network, and a 2.5D model for the flow.
-The first two of these were needed to reduce the spatial and temporal scales needed to model.
-Using the simulation framework developed here, these both can be relaxed in the future, at the cost of more computer time.
-The same goes for 2.5D
-vs.\ full 3D---the only difference is computer time, and our future
-calculations will explore the 3D evolution and compare to the 2.5D simulations presented here. In particular,
-in 3D we will be able to explore shear instabilities at the flame front.
-Larger networks are a straightforward change, and already supported in
-\castro\ using the \pynucastro\ framework~\citep{pynucastro} and JINA
-ReacLib rate database~\citep{reaclib}.
-
-\MarginPar{Relate our work to previous work}
-
-In addition to extending our models to 3D and relaxing the
-approximations used in this study, in the future we plan to perform a
-number of additional further studies. These include investigating the
-effects of ignition latitude and different initial models. We also
-plan to model mixed H/He bursts. This will require a different
-reaction network, and perhaps different (more relaxed) resolution
-requirements. A long term goal is to include magnetohydrodynamics in
-our models. Although the magnetic fields of neutron stars exhibiting
-X-ray bursts are relatively weak, with $B \leq 10^8 - 10^9$G
-\citep{mukherjee2015magnetic} (at least compared to e.g. magnetars,
-which have $B\sim 10^{15}$G), it has been found by
-\citet{art-2016-cavecchi-etal} that even weak magnetic fields could
-have a non-negligible effect on the flame propagation, reducing
-confinement due to the Coriolis force and leading to increased flame
-speeds. The effect of magnetic fields on the conductivities and
-anisotropic conduction might also be important.  Ultimately, we wish
-to link our simulations to observed light curves. To do this, we will
-need to explore radiation transport in order to model how the burst
-energy propagates through the outer layers of the neutron star
-atmosphere.
-
-% Full hydrodynamics models of flame spreading in XRB are possible with certain approximations.
-
-% CPU cost
-
-% Accurate simulations require
-% \begin{enumerate}
-% \item Resolving the scale height of the atmosphere
-
-% \item Thin transition between underlying neutron star and atmosphere, to ensure that the peak temperature
-% hits right at the proper base density
-% \end{enumerate}
-
-% The main approximations we made were
-% \begin{enumerate}
-% \item A higher than normal rotation rate
-% \item A boosted flame
-% \item Simplified reaction network
-% \item A 2.5D model for the flow
-% \end{enumerate}
-% The first two of these were needed to reduce the spatial and temporal
-% scales needed to model, making the simulations feasible.  Using the
-% simulation framework developed here, these both can be relaxed in the
-% future, at the cost of more computer time.  The same goes for 2.5D
-% vs.\ full 3D---the only difference is computer time, and our future
-% calculations will explore the 3D evolution and compare to these.  In particular,
-% in 3D we will be able to explore shear instabilities at the flame front.
-% Larger networks are a straightforward change, and already supported in
-% \castro\ using the \pynucastro\ framework~\citep{pynucastro} and JINA
-% ReacLib rate database~\citep{reaclib}.
-
-% These studies complement the\MarginPar{???? Finish this sentence}
-
-% Other future work includes mixed H/He bursts.  This will require a
-% different reaction network, and perhaps different (more relaxed)
-% resolution requirements.
-
-In addition to exploring other aspects of the XRB physics, there are several changes we will pursue to the algorithm used to
-model these XRBs.  First, as shown in \citet{castro-sdc}, we have
-developed a fourth-order (in space and time) method for coupling
-hydrodynamics and reactions that should greatly improve the accuracy
-of the simulations.  We expect that by using this new high-order
-algorithm we can drop a level of refinement from the simulations while
-still accurately modeling the evolution.  We have also ported
-\castro\ to GPUs, giving an order of magnitude performance boost on
-nodes with both CPUs and GPUs.  The GPU-enabled algorithm will
-increase the performance, allowing us to run bigger domains and/or
-eliminate the boosting used here.
-
-% MHD, radiation transport
 
@@
 
-\appendix
-
-\section{Diffusion Tests}
-\label{app:diffusion}
-
-To test our implementation of diffusion we created a unit test in
-\castro\ that just uses the diffusion solver.  This test uses the
-standard Gaussian initial conditions---the diffusion of a Gaussian
-remains Gaussian, with a lower amplitude and wider width.  This is run
-with a constant diffusion coefficent.  The point of this test is to
-demonstrate that the implementation, done in a predictor-corrector
-form is second-order accurate.
-
-We use a 2-d axisymmetric geometry, with a reflection boundary on the symmetry axis (left)
-and Neumann boundaries on all other sides.  In this geometry, the solution behaves like
-a spherical problem, and has the form:
-\begin{equation}
-  T(r,t) = T_1 + (T_2 - T_1) \left (\frac{t_0}{t + t_0} \right )^{3/2} e^{-r^2/(4 D (t + t_0))}
-\end{equation}
-where $t_0$ is a small time used to set the initial width of the
-Gaussian, $r$ is the distance from the origin, and $D$ is the
-diffusion coefficient.  $T_1$ is the ambient temperature and $T_2$ is
-the temperature of the peak of the Gaussian.  We use $D = 1$, $T_1 = 1$, $T_2 = 2$,
-and $t_0 = 0.001$.
-
-\begin{deluxetable}{rL}
-	\tablecaption{\label{table:diffusion} Convergence of diffusion test problem.}
-	\tablehead{\colhead{base resolution} & \colhead{L-inf error}}
-	\startdata
-\cutinhead{no refinement}
-	32  & 5.995\times 10^{-3} \\
-        64  & 1.477\times 10^{-3} \\
-       128  & 3.664\times 10^{-4} \\
-       256  & 9.139\times 10^{-5} \\
-\cutinhead{one level of refinement}
-        32  & 1.480\times 10^{-3} \\
-        64  & 3.685\times 10^{-4} \\
-       128  & 9.406\times 10^{-5} \\
-       256  & 2.446\times 10^{-5} \\
-	\enddata
-	%
-\end{deluxetable}
-
-Table~\ref{table:diffusion} shows the L-inf error against the analytic
-solution for the diffusion test problem.  The tests were run in two
-fashions: no AMR, and AMR with one level of refinement (a jump of
-$2\times$).  In both cases we see second-order convergence of the
-error.  Figure~\ref{fig:diffusion} shows the temperature at the end of
-the simulation of the $64\times 128$ base grid plus one level of
-refinement simulation, with the grids shown to illustrate where the
-refinement jump is.
-
-\begin{figure}[t]
-\centering
-\epsscale{0.5}
-\plotone{diffusion_temp_amr}
-\caption{\label{fig:diffusion} Temperature for the diffusion test problem at $10^{-3}~\mbox{s}$
-for a diffusion test with a $64\times 128$ zone base grid and one level of refinement.}
-\epsscale{1.0}
-\end{figure}
-
 %======================================================================

### 2019-11-24 xrb/paper.tex
Edited caption for plot showing AMR grid

@@
 
-\author[0000-0001-8092-1974]{Weiqun Zhang}
-\affiliation{Lawrence Berkeley National Laboratory, Berkeley, CA}
-
-\correspondingauthor{Michael Zingale}
-\email{Michael.Zingale@stonybrook.edu}
+\correspondingauthor{Kiran Eiden}
+\email{kiran.eiden@stonybrook.edu}
 
@@
 
-\cite{cavecchi:2013,art-2015-cavecchi-etal,art-2016-cavecchi-etal,Cavecchi2017,Cavecchi2019} \MarginPar{need a summary}
+The first vertically resolved simulations of lateral deflagrations
+were obtained by \citet{cavecchi:2013} who showed how the Coriolis
+force creates a geometrical configurations that increases the flame
+speed set by conduction by a factor $\sim ~ L_r / H$, where $L_r$ is the
+Rossby radius and $H$ is the scale height of the burning layer. The
+effect on flame propagation of changing Coriolis confinement across
+the surface was explored in \citet{art-2015-cavecchi-etal}, while \citet{art-2016-cavecchi-etal}
+showed how the magnetic field tension, opposing the Coriolis
+force, can either speed up or slow down the flame by changing the
+flame front orizontal extent. Finally, \citet{Cavecchi2019}
+studied the effects in 3D of the baroclinic instability at the flame
+front, measuring flames up to 10 times faster than in the 2D case.
 
@@
     \rho \gb \nonumber \\
-  &&-2 \rho \Omegab\times \Ub - \Omegab \times (\Omegab \times \rb) \\
+  &&-2 \rho \Omegab\times \Ub - \rho \Omegab \times (\Omegab \times \rb) \\
 %
@@
 	\plotone{grids.pdf}
-	\caption{\label{fig:grids} Slice through a simulation showing the four-level
-		grid structure. Note that the boxes shown are not the simulation zones, which
-		span $10~\mathrm{cm}$ at the finest level, but subdomains containing
-		approximately equal numbers of zones that are each managed by a separate
-		MPI process. A coarse base grid extending to the upper boundary is drawn
-		in white, with the green, blue, and black grids showing the jumps in
-		refinement needed to fully resolve the fuel layer and underlying neutron star.}
+	\caption{\label{fig:grids}A slice through part of a 2D simulation showing the four-level
+		grid structure. Note that the boxes shown are not the simulation zones, which span
+		$10~\mathrm{cm}$ at the finest level, but subdomains containing approximately equal
+		numbers of zones that are each managed by a separate MPI process. A coarse base grid
+		extending to the upper boundary is drawn in white, with the green, blue, and black
+		grids showing the jumps in refinement needed to fully resolve the fuel layer and
+		underlying neutron star.}
 \end{figure}
@@
 \begin{equation}
-v_f \approx \sqrt{k_\mathrm{th} \dot{\epsilon}} \qquad
-\lambda_f \approx \sqrt{\frac{k_\mathrm{th}}{\dot{\epsilon}}}
+\label{eq:flame_scaling}
+v_f \approx \sqrt{\kth \dot{\epsilon}} \qquad
+\lambda_f \approx \sqrt{\frac{\kth}{\dot{\epsilon}}}
 \end{equation}
@@
 	laminar flame. The left plot shows temperature and nuclear energy
-	generation profiles at 11 different time points, while the right plot
+	generation profiles at 11 different times, while the right plot
 	shows flame propagation speed and flame thickness as functions of time.}
@@
 
-\begin{deluxetable}{lcc}
-	\tablecaption{\label{table:flame_speeds}Turbulent and laminar flame speeds.}
-	\tablehead{\colhead{run} & \colhead{$s_T$ (km s$^{-1}$)} & \colhead{$s_L$ (km s$^{-1}$)}}
+\begin{deluxetable}{lc}
+	\tablecaption{\label{table:flame_speeds} Laminar flame speeds.}
+	\tablehead{\colhead{run} & \colhead{$s_L$ (km s$^{-1}$)}}
 	\startdata
-	10/10 boost & $9.18 \pm 0.03$  & $1.06 \pm 0.01$ \\
-	5/5 boost & $4.00 \pm 0.01$ & $0.56 \pm 0.01$ \\
-	iso7 network & $7.56 \pm 0.02$ & $-$ \\
-	lowres & $9.33 \pm 0.04$ & $-$ \\
-	1000 Hz & $18.6 \pm 0.16$ & $-$ \\
+	10/10 boost & $1.06 \pm 0.01$ \\
+	5/5 boost & $0.56 \pm 0.01$ \\
 	\enddata
@@
 Eq.~\ref{eq:hse}.  Integrating upwards, we find $p_i$ and
-$\rho_i$ using a Newton solve together with the equation of state, with either the temperature
+$\rho_i$ using a Newton-Raphson solver together with the equation of state, with either the temperature
 specified, $T_i = T(p_i, \rho_i, \{X_k\}_i)$ (in the isothermal, ramp,
@@
 
-Table \ref{table:flame_speeds} gives the flame speed measured in each
+\begin{deluxetable}{lc}
+	\tablecaption{\label{table:flame_speeds_multid} Flame speeds measured from 2D calculations.}
+	\tablehead{\colhead{run} & \colhead{$s_\mathrm{front}$ (km s$^{-1}$)}} 
+	\startdata
+	10/10 boost & $9.18 \pm 0.03$ \\
+	5/5 boost & $4.00 \pm 0.01$ \\
+	iso7 network & $7.56 \pm 0.02$ \\
+	lowres & $9.33 \pm 0.04$ \\
+	1000 Hz & $18.6 \pm 0.16$ \\
+	\enddata
+	%
+\end{deluxetable}
+
+Table \ref{table:flame_speeds_multid} gives the flame speed measured in each
 multidimensional simulation, as well as laminar speeds obtained from
@@
 
-{\color{red} Should say qualitatively what the plot shows}
+As we see in Figure~\ref{fig:flame_speed}, all of the various
+approximations behave as we would expect.  The 1000~Hz run goes 2
+times faster than the 2000~Hz run, as expected by the reduced rossby
+radius \citep{cavecchi:2013}).  The 5/5 boost run is half the speed of
+the 10/10 boost run, also as expected, consistent with the 1D laminar
+flames and behaving as predicted by Eq.~\ref{eq:flame_scaling}.  This
+confirms that the role of the Coriolis force is really to boost the
+laminar flame by the geometrical/hydrodynamical effect expected ($v_f
+= v_l * L_r / H$), similarly to what the 1000 Hz simulation
+said.
+
+%Seeing the proofs both changing the Coriolis force AND the
+%laminar flame properties I think is very cool.
 
@@
 and shows the misalignment of the local density and pressure gradients
-\citep{Malone2014a}.
-As we are considering a 2d \MarginPar{We've used at least three different ways of writing 2-d/2D/2d so far.
-These should be changed so we used one style throughout.}system, the component of the
-baroclinicity we consider here (out of the plane) reflects the misalignment
-of the fields in the plane of the simulation. Figure~\ref{fig:baroclinicity}
-shows that the baroclinicity peaks along the flame front. This baroclinicity
-generates vorticity, which in turn entrains material along the surface of the
-flame.
+\citep{Malone2014a}. As we are considering a 2d \MarginPar{We've used
+  at least three different ways of writing 2-d/2D/2d so far. These
+  should be changed so we used one style throughout.}system, the
+component of the baroclinicity we consider here (out of the plane)
+reflects the misalignment of the fields in the plane of the
+simulation. Figure~\ref{fig:baroclinicity} shows that the
+baroclinicity peaks along the flame front. This baroclinicity
+generates vorticity, which in turn entrains material along the surface
+of the flame (see also \citet{cavecchi:2013}). This same vorticity,
+in 3D, should perturb the flame front and further increase flame speed
+\citep{Cavecchi2019}.
 
@@
 
-To further demonstrate the relationships between different properties of the flame, in Figure~\ref{fig:phase_plots} we present some phase plots of the energy generation rate. The phase plot of the
-energy generation rate as a function of
-the $x$- and $y$-velocities shows that energy is preferentially generated in
-regions with negative $x$-velocity.
-This most likely corresponds to the burning along the underside of the flame front, where fuel has been entrained along the surface of the flame and so is moving in the opposite direction to the direction the flame is propagating in.
-This is further supported by the phase plot of the energy generation rate as a function of the $x$-velocity
-and the density, which shows that the energy generation rate peaks at
-$\rho \sim 3 \times 10^5~\mathrm{g/cm}^3$, at the base of the flame.
+To further demonstrate the relationships between different properties
+of the flame, in Figure~\ref{fig:phase_plots} we present some phase
+plots of the energy generation rate. The phase plot of the energy
+generation rate as a function of the $x$- and $y$-velocities shows
+that energy is preferentially generated in regions with negative
+$x$-velocity. This most likely corresponds to the burning along the
+underside of the flame front, where fuel has been entrained along the
+surface of the flame and so is moving in the opposite direction to the
+direction the flame is propagating in. This is further supported by
+the phase plot of the energy generation rate as a function of the
+$x$-velocity and the density, which shows that the energy generation
+rate peaks at $\rho \sim 3 \times 10^5~\mathrm{g/cm}^3$, at the base
+of the flame.  The energy generation rate also depends strongly on density.
 
@@
 calculations will explore the 3D evolution and compare to the 2.5D simulations presented here. In particular,
-in 3D we will be able to explore shear instabilities at the flame front.
+in 3D we will be able to explore shear instabilities at the flame front.  We can also capture the 
+baroclinic instability \citep{Cavecchi2019} and the competition between it and shear.
 Larger networks are a straightforward change, and already supported in
@@
   flame\_wave}.  The work at Stony Brook was supported by DOE/Office
-of Nuclear Physics grant DE-FG02-87ER40317.  This research used
-resources of the National Energy Research Scientific Computing Center,
-a DOE Office of Science User Facility supported by the Office of
-Science of the U.~S.\ Department of Energy under Contract
-No.\ DE-AC02-05CH11231.  This research used resources of the Oak Ridge
-Leadership Computing Facility at the Oak Ridge National Laboratory,
-which is supported by the Office of Science of the U.S. Department of
-Energy under Contract No. DE-AC05-00OR22725, awarded through the DOE
-INCITE program.  This research has made use of NASA's Astrophysics
-Data System Bibliographic Services.
+of Nuclear Physics grant DE-FG02-87ER40317.  YC was supported by the
+European Union Horizon 2020 research and innovation program under
+the Marie Sklodowska-Curie Global Fellowship grant agreement No.\
+703916.  This research used resources of the National Energy Research
+Scientific Computing Center, a DOE Office of Science User Facility
+supported by the Office of Science of the U.~S.\ Department of Energy
+under Contract No.\ DE-AC02-05CH11231.  This research used resources
+of the Oak Ridge Leadership Computing Facility at the Oak Ridge
+National Laboratory, which is supported by the Office of Science of
+the U.S. Department of Energy under Contract No. DE-AC05-00OR22725,
+awarded through the DOE INCITE program.  This research has made use of
+NASA's Astrophysics Data System Bibliographic Services.
 

### 2019-11-25 xrb/paper.tex
Some additional comments, edits for XRB paper

@@
 burning products vertically throughout the neutron star atmosphere as well
-as the nature of any turbulence the burning front might encounter as it propagates through the atmosphere.
+as the nature of any turbulence the burning front might encounter as it propagates
+through the atmosphere.
 
 The first vertically resolved simulations of lateral deflagrations
-were obtained by \citet{cavecchi:2013} who showed how the Coriolis
-force creates a geometrical configurations that increases the flame
-speed set by conduction by a factor $\sim ~ L_r / H$, where $L_r$ is the
+were obtained by \citet{cavecchi:2013}, who showed how the Coriolis
+force creates a geometrical configuration that increases the flame
+speed set by conduction by a factor $\sim ~ L_R / H$, where $L_R$ is the
 Rossby radius and $H$ is the scale height of the burning layer. The
-effect on flame propagation of changing Coriolis confinement across
-the surface was explored in \citet{art-2015-cavecchi-etal}, while \citet{art-2016-cavecchi-etal}
-showed how the magnetic field tension, opposing the Coriolis
-force, can either speed up or slow down the flame by changing the
-flame front orizontal extent. Finally, \citet{Cavecchi2019}
+effect of changing Coriolis confinement across the surface on the flame 
+propagation was explored in \citet{art-2015-cavecchi-etal},
+while \citet{art-2016-cavecchi-etal} showed how magnetic field tension,
+opposing the Coriolis force, can either speed up or slow down the flame by
+changing the horizontal extent of the flame front. Finally, \citet{Cavecchi2019}
 studied the effects in 3D of the baroclinic instability at the flame
@@
 $\omegadot$, and are related to the total specific energy generation
-rate, $\epsdot$.  The total mass conservation,
+rate, $\epsdot$\MarginPar{Is notation consistent between E, e, epsilondot, and enucdot?}.  The total mass conservation,
 \begin{equation}
@@
 variations in the azimuthal direction.  This allows us to recast the
-$w$-velocity equation as a simple advection equation:
+$w$-velocity equation as a simple advection equation:\MarginPar{In conjunction with total mass conservation?}
 \begin{equation}
@@
 described in \cite{timmes_swesty:2000}.  To model reactions, we use a 
-13-isotope alpha chain, derived from the {\tt aprox13}
+13-isotope alpha chain network derived from the {\tt aprox13}
 network~\citep{timmes_aprox13}.  For one of our runs, we use the
@@
 
+{\color{red} We should decide on a notation for 2D and laminar flame speeds.}
+
 The speed and thickness of a laminar helium flame are determined by the
@@
 speed decreases, so we will easily resolve the flame structure throughout
-the rest of the atmosphere.
-
-Flame propagation rates from both the 1D laminar flames and full 2D simulations are given in Table \ref{table:flame_speeds}.
-\MarginPar{I think we should only report the laminar speeds here and defer the turbulent speeds until the section where we talk about how they are computed}
+the rest of the atmosphere. Propagation rates of the 1D laminar flames are given
+in Table \ref{table:flame_speeds_1d}.
 
 \begin{deluxetable}{lc}
-	\tablecaption{\label{table:flame_speeds} Laminar flame speeds.}
+	\tablecaption{\label{table:flame_speeds_1d} Laminar flame speeds.}
 	\tablehead{\colhead{run} & \colhead{$s_L$ (km s$^{-1}$)}}
@@
 
-Table~\ref{table:sim_names} summarizes the simulations that we ran. \MarginPar{This table has different model names to the previous table}
+Table~\ref{table:sim_names} summarizes the simulations that we ran. \MarginPar{This table has different model names from the flame speed table}
 The majority of them used a reaction rate boosting of $10$ and a
@@
 burning front is propagating to the right.  In the temperature plot,
-we see the effect of our refinement criteria focusing only on the
-atmosphere where we are most dense, with an artificial change in the
+we see the effect of our refinement criteria focusing only on the part of
+the atmosphere where we are most dense, with an artificial change in the
 temperature at the refinement boundary due in part to the construction
@@
   was/where they should be looking in the plot to see this temperature
-  drop} As we'll see when the lower resolution case, we do not believe
+  drop} As we will see in the lower resolution case, we do not believe
 that this affects the results.
@@
 Figure~\ref{fig:network} shows a comparison of the 10/10 boosting run
-with the standard 13-isotope ``aprox13'' network and the reduced
+with the standard 13-isotope ``aprox13'' network and the reduced\MarginPar{We use {\tt tt} instead of quotes for the networks in the numerical approach section.}
 7-isotope ``iso7'' network.  We see that the flame in the aprox13 case
@@
 approximations behave as we would expect.  The 1000~Hz run goes 2
-times faster than the 2000~Hz run, as expected by the reduced rossby
+times faster than the 2000~Hz run, as expected by the reduced Rossby
 radius \citep{cavecchi:2013}).  The 5/5 boost run is half the speed of
@@
 laminar flame by the geometrical/hydrodynamical effect expected ($v_f
-= v_l * L_r / H$), similarly to what the 1000 Hz simulation
+= v_l * L_R / H$), similarly to what the 1000 Hz simulation
 said.
@@
 generates vorticity, which in turn entrains material along the surface
-of the flame (see also \citet{cavecchi:2013}). This same vorticity,
-in 3D, should perturb the flame front and further increase flame speed
+of the flame (see also \citet{cavecchi:2013}). In 3D, his same vorticity
+should perturb the flame front and further increase the flame speed.
 \citep{Cavecchi2019}.
@@
 remains Gaussian, with a lower amplitude and wider width.  This is run
-with a constant diffusion coefficent.  The point of this test is to
+with a constant diffusion coefficient. The point of this test is to
 demonstrate that the implementation done in a predictor-corrector

### 2019-11-25 xrb/paper.tex
Integrate Yuri's comments

@@
 
-\author{Yuri Cavecchi}
+\author[0000-0002-6447-3603]{Yuri Cavecchi}
 \affiliation{Mathematical Sciences and STAG Research Centre, University of Southampton, SO17 1BJ}
@@
 shallow-water approximation, so the vertical details of the
-atmosphere's structure we not captured.  Their model showed that the
+atmosphere's structure were not captured.  Their model showed that the
 horizontal pressure gradient between the ash and fuel can be important
@@
 force creates a geometrical configuration that increases the flame
-speed set by conduction by a factor $\sim ~ L_R / H$, where $L_R$ is the
+speed set by conduction by a factor $\sim L_R / H$, where $L_R$ is the
 Rossby radius and $H$ is the scale height of the burning layer. The
@@
 takes about 3~ms to settle into a sustained flame.  Note that this speed
-is quite small compared to the speed estimated in
+is quite small compared to the speeds of $\sim 10^6~\cms$ estimated in
 \citet{spitkovsky2002}.  We expect a multidimensional flame to
@@
 \begin{deluxetable}{lc}
-	\tablecaption{\label{table:flame_speeds_multid} Flame speeds measured from 2D calculations.}
+	\tablecaption{\label{table:flame_speeds_multid} Flame speeds measured in 2D calculations.}
 	\tablehead{\colhead{run} & \colhead{$s_\mathrm{front}$ (km s$^{-1}$)}} 
@@
 multidimensional simulation, as well as laminar speeds obtained from
-corresponding 1D runs. The 2D flames propagate at speeds about an
-order of magnitude faster than their 1D counterparts. The increase in
-flame speed is likely a product of the larger flame surface area and hydrodynamical effects such as
-turbulence, wrinkling, and convective cycles, which bring cooler fuel
-from ahead of the front into the hottest part of the burning region.
+corresponding 1D runs. {\color{red} Should we include the laminar speeds again for reference?}
+The 2D flames propagate at speeds about an order of magnitude faster than their 1D counterparts. 
+The increase in flame speed is likely a product of the larger flame surface area and
+hydrodynamical effects such as turbulence, wrinkling, and convective cycles, which
+bring cooler fuel from ahead of the front into the hottest part of the burning region.
 

### 2019-11-26 xrb/paper.tex
some comments

@@
 %======================================================================
-\title{Dynamics of Laterally Propagating Flames in X-ray Bursts}
+\title{Dynamics of Laterally Propagating Flames in X-ray Bursts. I. Burning Front Structure}
 
@@
 the highly-stratified burning layer in an X-ray burst.
-Simple \MarginPar{is simple the right word here?} \MarginPar{It might be better to say simulations are performed through a simplified atmosphere?}
-two-dimensional simulations of flame propagation are performed through
-a rotating plane-parallel atmosphere, exploring the structure of the
-flame.  We discuss the approximations needed to capture the length and
-time scales at play in an X-ray burst.  Our studies complement 
+Two-dimensional hydrodynamics simulations of flame propagation are
+performed through a rotating plane-parallel atmosphere, exploring the
+structure of the flame.  We discuss the approximations needed to
+capture the length and time scales at play in an X-ray burst and
+describe the flame acceleration observed.  Our studies complement
 other multidimensional studies of burning in X-ray bursts.
@@
 Laterally propagating detonations were modeled by
-\citet{fryxellwoosley82,hedet}. However, since it is difficult to detonate helium at the
+\citet{fryxellwoosley82} and \citet{hedet}. However, since it is difficult to detonate helium at the
 densities found in normal XRBs, and even harder to detonate hydrogen
@@
 \end{eqnarray}
-Here, $\rho$ is the mass density, $\Ub$ is the velocity, $p$ is the pressure and $E$ is
-the specific total energy, which is related to the specific internal
-energy as $e = E - |\Ub|^2/2$.  Species are described by mass
-fractions, $X_k$ (such that $\sum_k X_k = 1$), and creation rates,
-$\omegadot$, and are related to the total specific energy generation
-rate, $\epsdot$\MarginPar{Is notation consistent between E, e, epsilondot, and enucdot?}.  The total mass conservation,
+Here, $\rho$ is the mass density, $\Ub$ is the velocity, $p$ is the
+pressure and $E$ is the specific total energy, which is related to the
+specific internal energy as $e = E - |\Ub|^2/2$.  The forcing in the
+momentum equation includes gravity, described by gravitation
+acceleration $\gb$ and rotational forces described by angular velocity
+$\Omegab$, with $\rb$ the position vector from the origin.
+Species are described by mass fractions, $X_k$ (such that $\sum_k X_k
+= 1$), and creation rates, $\omegadot$, and are related to the total
+specific energy generation rate, $\epsdot$\MarginPar{Is notation
+  consistent between E, e, epsilondot, and enucdot?}.  The total mass
+conservation,
 \begin{equation}
@@
 \end{equation}
-implies $\sum_k \omegadot_k = 0$.  \castro\ uses an unsplit piecewise
+implies $\sum_k \omegadot_k = 0$.  An equation of state of the form $p = p(\rho, e, X_k)$ completes the
+thermodynamic description of the system.  Thermal diffusion is 
+described by a thermal conductivity $\kth$ and temperature $T$.
+
+\castro\ uses an unsplit piecewise
 parabolic method (PPM) with characteristic tracing for solving the
@@
 
-An equation of state of the form $p = p(\rho, e, X_k)$ completes the
-thermodynamic description of the system.  Finally, $\kth$ is the
-thermal conductivity and $T$ is the temperature.  Diffusion is treated
-explicitly in time. \MarginPar{Have not defined what r is in eqs 1-3, and Omega and g are only defined with respect to eq. 5}
-
-Since the neutron star rotates, we work in a corotating frame, with
-angular velocity $\Omegab$.  Further, for the two-dimensional
+Since the neutron star rotates, we work in a corotating frame, taking
+the angular velocity $\Omegab$ to be constant.  Further, for the two-dimensional
 simulations presented here, we work in axisymmetric coordinates,
@@
 expected to greatly affect the dynamics.  Carrying the Coriolis force
-allows us to capture the geostrophic balance\MarginPar{ref, Spitkovsky 2002?} that sets
-up via lateral hydrostatic equilibrium.  In the discussions below, we will use the Castro coordinate
-names, $(x, y)$ in our notation.
+allows us to capture the geostrophic balance that sets up via lateral
+hydrostatic equilibrium \citep{spitkovsky2002}.  In the discussions
+below, we will use the Castro coordinate names, $(x, y)$ in our
+notation.
 

### 2019-11-26 xrb/paper.tex
add sponge

@@
 violently blown out of the top of the atmosphere, so we apply a sponge
-to this region.  This is similar to the method we previously used in \cite{xrb3d}.
+to this region.  This is similar to the method we previously used in \cite{xrb2}, and takes
+the form of a source term to the momentum and energy equations of the form:
+\begin{align}
+{\bf S}_{\rho \Ub} &= \rho \Ub \frac{f}{\Delta t} \\
+S_{\rho E} &= \rho \Ub \cdot {\bf S}_{\rho \Ub}
+\end{align}
+with the sponge forcing $f$ dependent on the density.  We define the sponge
+shape as:
+\begin{equation}
+s = \left \{
+     \begin{array}{cc}
+            0   & \rho > \rho_\mathrm{upper} \\
+            \frac{1}{2}
+                \left [ 1 - \cos \left ( \frac{\pi (\rho - \rho_\mathrm{upper})}{\Delta \rho} \right ) \right ]  & \rho_\mathrm{upper} \ge \rho > \rho_\mathrm{lower} \\
+                     1 & \rho < \rho_\mathrm{lower} 
+             \end{array} \right .
+\end{equation}
+Here $\rho_\mathrm{upper}$ and $\rho_\mathrm{lower}$ are the densities where the sponge transitions to being fully applied.
+We take $\rho_\mathrm{upper} = 10^2~\gcc$ and $\rho_\mathrm{lower} = 1~\gcc$, with $\Delta \rho = \rho_\mathrm{upper} - \rho_\mathrm{lower}$.
+The sponge update is done implicitly to get the effective forcing, $f$:
+\begin{equation}
+f = -\left [ 1 - \frac{1}{1 + \alpha s} \right ]
+\end{equation}
+with $\alpha = \Delta t/\tau_\mathrm{sponge}$.  Here
+$\tau_\mathrm{sponge}$ is the timescale over which the sponge acts.
+We take $\tau_\mathrm{sponge} = 10^{-7}~\mathrm{s}$.
 The sponge drives the velocity of the material in the low
-density \MarginPar{add mathematical form of sponge} regions at the top
+density regions at the top
 of the atmosphere to zero.  This sponging helps increase our timestep

### 2019-11-26 xrb/paper.tex
more edits

@@
 propagation, we do a 1D simulation of a laminar flame using our
-microphysics.  Figure~\ref{fig:flame} shows the flame properties for a 
+microphysics.  Figure~\ref{fig:flame} shows the flame thermodynamic
+profile and properties for the 10/10 flame using
 % \AssignTo{Kiran and Mike will remake with the 10/10 boosted flame}
-He flame using the standard conductivities and our simple reaction \MarginPar{standard?  this is a boosted flame, right?}
-network.  This flame had a density of $2\times 10^6~\gcc$ and 
-% \MarginPar{double check}
+our conductivities and the {\tt aprox13} reaction network.  This flame
+had a density of $2\times 10^6~\gcc$ and
 temperature of $5\times 10^7$~K.  We observe that this flame speed is
@@
 is quite small compared to the speeds of $\sim 10^6~\cms$ estimated in
-\citet{spitkovsky2002}.  We expect a multidimensional flame to
+\citet{spitkovsky2002}.  Table~\ref{table:flame_speeds_1d} gives the properties
+of the 10/10 and 5/5 laminar flames.  We expect a multidimensional flame to
 accelerate due to hydrodynamics interactions (wrinkling, turbulence
@@
 speed decreases, so we will easily resolve the flame structure throughout
-the rest of the atmosphere. Propagation rates of the 1D laminar flames are given
-in Table \ref{table:flame_speeds_1d}.
+the rest of the atmosphere.
 
@@
 have constant density and temperature.
+The choice of factors in front of $\delta_\mathrm{atm}$ were designed
+to make sure the peak $T$ is attained at the desired density
+of the burning layer.
 
@@
 
-The choice of factors in front of $\delta_\mathrm{atm}$ were designed
-to make sure the peak $T$ is attained at the desired density
-of the burning layer.
-
 We create two models, a ``cool'' model representing the atmosphere
@@
 the model generation are listed in Table~\ref{table:params}. \MarginPar{reference and discuss the figure of initial model here}
-
 We blend the hot and cold models laterally to produce the perturbation
@@
 equilibrum, the blended model is also in vertical equilibrium initially.  We choose
-$x_\mathrm{pert} = 1.024\times 10^4$~cm and $\delta_\mathrm{blend} = 2048$~cm.  Once \MarginPar{double check these numbers}
+$x_\mathrm{pert} = 1.024\times 10^4$~cm and $\delta_\mathrm{blend} = 2048$~cm.  Once
 the blended model is constructed, we compute $T(x,y)$ and $(\rho e)(x,y)$
@@
 
-Table~\ref{table:sim_names} summarizes the simulations that we ran. \MarginPar{This table has different model names from the flame speed table}
+To assess the sensitivity of the flame propagation to the various approximations
+we made, we ran a suite of simulations. 
+Table~\ref{table:sim_names} summarizes these simulations. \MarginPar{This table has different model names from the flame speed table}
 The majority of them used a reaction rate boosting of $10$ and a

### 2019-12-01 xrb/paper.tex
More edits, discussed streamline plot
Still need to talk about rotation, and show Don's stuff. The initial
model and general features sections need a little bit more work.

@@
 nuclear equation of state~\citep{steiner:2010,ozel2016masses}.  Extensive
-observations of brightness oscillations during the rise (the initial phase of the burst where the observed flux rapidly increases) have provided
+observations of brightness oscillations during the rise (the initial phase
+of the burst where the observed flux rapidly increases) have provided
 evidence that the burning begins in a localized region and spreads
@@
 = 1$), and creation rates, $\omegadot$, and are related to the total
-specific energy generation rate, $\epsdot$\MarginPar{Is notation
-  consistent between E, e, epsilondot, and enucdot?}.  The total mass
+specific energy generation rate, $\epsdot$.  The total mass
 conservation,
@@
 variations in the azimuthal direction.  This allows us to recast the
-$w$-velocity equation as a simple advection equation:\MarginPar{In conjunction with total mass conservation?}
+$w$-velocity equation as a simple advection equation:
 \begin{equation}
@@
 	\plotone{grids.pdf}
-	\caption{\label{fig:grids}A slice through part of a 2D simulation showing the four-level
+	\caption{\label{fig:grids}A section of a 2D simulation showing the four-level
 		grid structure. Note that the boxes shown are not the simulation zones, which span
@@
 
-{\color{red} We should decide on a notation for 2D and laminar flame speeds.}
-
 The speed and thickness of a laminar helium flame are determined by the
@@
 \label{eq:flame_scaling}
-v_f \approx \sqrt{\kth \dot{\epsilon}} \qquad
-\lambda_f \approx \sqrt{\frac{\kth}{\dot{\epsilon}}}
+s_L \approx \sqrt{\kth \epsdot} \qquad
+\lambda_L \approx \sqrt{\frac{\kth}{\epsdot}}
 \end{equation}
@@
 
-We measure the flame width as:
-\begin{equation}
-\lambda_f \equiv \frac{\Delta T}{\max\{|\nabla T|\}}
+We measure the laminar flame width as:
+\begin{equation}
+\lambda_L \equiv \frac{\Delta T}{\max\{|\nabla T|\}}
 \end{equation}
 Experience with modeling resolved flames suggests that we need a
-spatial resolution, $\Delta x$ of $\lambda_f/\Delta x \sim 5$~\citep{SNld}.  These
+spatial resolution, $\Delta x$ of $\lambda_L/\Delta x \sim 5$~\citep{SNld}.  These
 conditions represent the bottom of the He layer.  As the density
@@
 we made, we ran a suite of simulations. 
-Table~\ref{table:sim_names} summarizes these simulations. \MarginPar{This table has different model names from the flame speed table}
+Table~\ref{table:sim_names} summarizes these simulations.
 The majority of them used a reaction rate boosting of $10$ and a
@@
 10/10 & 10 & 10 & 10~cm & 2000~Hz & $1.2288\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$ & {\tt aprox13} \\
-10/10-slow & 10 & 10 & 10~cm & 1000~Hz & $1.8432\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$ &{\tt aprox13} \\
-10/10-lores & 10 & 10 & 20~cm & 2000~Hz & $1.2288\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$ &{\tt aprox13} \\
 5/5 & 5 & 5 & 10~cm & 2000~Hz & $1.2288\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$  &{\tt aprox13} \\
 10/10-iso7 & 10 & 10 & 10~cm & 2000~Hz & $1.2288\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$ &{\tt iso7} \\
+10/10-lowres & 10 & 10 & 20~cm & 2000~Hz & $1.2288\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$ &{\tt aprox13} \\
+10/10-1000 Hz & 10 & 10 & 10~cm & 1000~Hz & $1.8432\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$ &{\tt aprox13} \\
 \enddata
@@
 fewer refinement level, giving a $20~\mbox{cm}$ resolution overall (this is our
-10/10-lores run).  Figure~\ref{fig:10_10_lowres} shows the fields for this run at
+10/10-lowres run).  Figure~\ref{fig:10_10_lowres} shows the fields for this run at
 20~ms.  The structure is largely the same as the 10/10 run, with
@@
 Figure~\ref{fig:network} shows a comparison of the 10/10 boosting run
-with the standard 13-isotope ``aprox13'' network and the reduced\MarginPar{We use {\tt tt} instead of quotes for the networks in the numerical approach section.}
-7-isotope ``iso7'' network.  We see that the flame in the aprox13 case
-is slightly more advanced and has a higher $\bar{A}$ than the iso7
+with the standard 13-isotope {\tt aprox13} network and the reduced
+7-isotope {\tt iso7} network.  We see that the flame in the {\tt aprox13}
+case is slightly more advanced and has a higher $\bar{A}$ than the {\tt iso7}
 case.  We see in the next section that these two networks give largely
 the same flame speed.  The reduced network size saves a lot of memory,
-which will be useful when we transition to 3-d simulations.
+which will be useful when we transition to 3D simulations.
 
@@
 \caption{\label{fig:network} Mean molecular weight at 15~ms comparing a run with
-  aprox13 vs a run with the iso7 network.}
+  {\tt aprox13} to a run with the {\tt iso7} network.}
 \end{figure}
@@
 
-To measure the propagation rate of the burning front, we track the
-location beyond the hottest part of the flame\MarginPar{what does this mean?} where
-$\dot{e}_\mathrm{nuc}$ first drops to less than $0.1 \%$ of its
-maximum value. Using $0.1 \%$ of the peak value rather than the peak value
-itself helps to mitigate fluctuations caused by localized fluid
-motions, as the reduced value lies in a less turbulent part of the
-burning region. We also average $\dot{e}_\mathrm{nuc}$ over the  \MarginPar{I think we should discuss the averaging first, before we discuss the peak value}
-vertical coordinate to further dampen these fluctuations. After an
-initial transient period typically spanning $\sim 3$ ms, the flame
-settles into a state of steady propagation. The position data here are
-well fitted by a linear function of time (as seen in Figure
-\ref{fig:flame_speed}), and the resulting slope gives the velocity of
-the flame front.
+To measure the propagation rate of the burning front, we first collapse our nuclear
+energy generation rate ($\dot{e}_\mathrm{nuc}$) data at each time into a 1D radial
+profile by averaging over the vertical coordinate. We then take the peak
+$\dot{e}_\mathrm{nuc}$ value across all profiles to provide a fixed reference point.
+We define the position of the flame front to be the location ahead of hottest part of
+the flame where $\dot{e}_\mathrm{nuc}$ first drops to $< 0.1 \%$ of the global maximum. 
+This corresponds roughly to the leading edge of the burning region. Averaging
+over the vertical coordinate helps to reduce sensitivity to localized fluid motions, as does
+tracking the $0.1 \%$ contour rather than a local maximum. As we see in Figure
+\ref{fig:flame_speed}, the flame settles into a state of steady propagation after an initial
+transient period spanning $\sim 3$ ms. The position data here are well fitted by a linear
+function of time, and the resulting slope gives the velocity of the flame front.
 
@@
 	\startdata
-	10/10 boost & $9.18 \pm 0.03$ \\
-	5/5 boost & $4.00 \pm 0.01$ \\
-	iso7 network & $7.56 \pm 0.02$ \\
-	lowres & $9.33 \pm 0.04$ \\
-	1000 Hz & $18.6 \pm 0.16$ \\
+	10/10 & $9.18 \pm 0.03$ \\
+	5/5 & $4.00 \pm 0.01$ \\
+	10/10-iso7 & $7.56 \pm 0.02$ \\
+	10/10-lowres & $9.33 \pm 0.04$ \\
+	10/10-1000 Hz & $18.6 \pm 0.16$ \\
 	\enddata
@@
 
-Table \ref{table:flame_speeds_multid} gives the flame speed measured in each
-multidimensional simulation, as well as laminar speeds obtained from
-corresponding 1D runs. {\color{red} Should we include the laminar speeds again for reference?}
-The 2D flames propagate at speeds about an order of magnitude faster than their 1D counterparts. 
-The increase in flame speed is likely a product of the larger flame surface area and
-hydrodynamical effects such as turbulence, wrinkling, and convective cycles, which
-bring cooler fuel from ahead of the front into the hottest part of the burning region.
+Table \ref{table:flame_speeds_multid} gives the flame speed measured in each multidimensional simulation.
+The 2D flames propagate at speeds about an order of magnitude faster than their 1D counterparts (Table \ref{table:flame_speeds_1d}).
+The increase in flame speed is likely a product of the larger flame surface area and hydrodynamical effects such as turbulence,
+wrinkling, and convective cycles, which bring cooler fuel from ahead of the front into the hottest part of the burning region.
 
@@
 
-As we see in Figure~\ref{fig:flame_speed}, all of the various
-approximations behave as we would expect.  The 1000~Hz run goes 2
-times faster than the 2000~Hz run, as expected by the reduced Rossby
-radius \citep{cavecchi:2013}).  The 5/5 boost run is half the speed of
-the 10/10 boost run, also as expected, consistent with the 1D laminar
-flames and behaving as predicted by Eq.~\ref{eq:flame_scaling}.  This
-confirms that the role of the Coriolis force is really to boost the
-laminar flame by the geometrical/hydrodynamical effect expected ($v_f
-= v_l * L_R / H$), similarly to what the 1000 Hz simulation
-said.
+All of the various approximations have the expected effects. The flame in the 5/5
+boost run propagates at about half the speed of the 10/10 flame. This is consistent
+with the 1D laminar flames, and is the behavior predicted by Eq.~\ref{eq:flame_scaling}. The
+1000~Hz run goes 2 times faster than the 10/10 2000~Hz run, as expected from the reduced
+Rossby radius \citep{cavecchi:2013}. This confirms that the role of the Coriolis
+force is to limit the rate of flame spreading by the anticipated geometrical/hydrodynamical
+effect ($s_\mathrm{front} = s_L * L_R / H$). Reducing the resolution had minimal impact on
+the flame speed -- the low resolution line is right on top of the fit for the standard run
+in Figure \ref{fig:flame_speed}, with their slopes differing by only a few percent. Using a
+smaller network also produces similar behavior to the standard run, although there is a small
+reduction in speed owing to less energetic burning.
 
@@
 
-\AssignTo{Alice, Don}
-
-{\color{red} Show whatever it was that Don cooked up. Also need to discuss the streamline plot.}
+{\color{red} Show whatever it was that Don cooked up.}
 
@@
 and shows the misalignment of the local density and pressure gradients
-\citep{Malone2014a}. As we are considering a 2d \MarginPar{We've used
-  at least three different ways of writing 2-d/2D/2d so far. These
-  should be changed so we used one style throughout.}system, the
+\citep{Malone2014a}. As we are considering a 2D system, the
 component of the baroclinicity we consider here (out of the plane)
@@
 generates vorticity, which in turn entrains material along the surface
-of the flame (see also \citet{cavecchi:2013}). In 3D, his same vorticity
-should perturb the flame front and further increase the flame speed.
+of the flame (see also \citealt{cavecchi:2013}). In 3D, this same vorticity
+should perturb the flame front and further increase the flame speed
 \citep{Cavecchi2019}.
@@
 \end{figure}
+
+Vortical motions are also evident in a direct visualization of the velocity field, as seen in
+Figure \ref{fig:streamlines}. We observe turbulence in the wake of the flame and in the
+unburnt fuel ahead of the front, while a convective cycle sets up near the hottest part of the
+burning region. The cycle draws the cooler fluid near the interface into the center of the flame
+and drives hot ash out towards the instability at its surface, helping to facilitate the flame
+spreading. Convective mixing should still be important in 3D, but we would expect much more
+complicated flow patterns, with a greater contribution from small-scale features \citep{xrb3d}.
 
@@
 \section{Discussion}
-
-\AssignTo{everyone}
 
@@
 
-We use a 2-d axisymmetric geometry, with a reflection boundary on the symmetry axis (left)
+We use a 2D axisymmetric geometry, with a reflection boundary on the symmetry axis (left)
 and Neumann boundaries on all other sides.  In this geometry, the solution behaves like

### 2019-12-02 xrb/paper.tex
Reformatting
add new plot

@@
   was/where they should be looking in the plot to see this temperature
-  drop} As we will see in the lower resolution case, we do not believe
-that this affects the results.
+  drop} As we will see in the lower resolution case, this does not
+affect the results.
 
@@
 
-The final approximation to explore is our choice of rotation rate.
-{\color{red} need to add discussion + plot}
+The final approximation to explore is our choice of rotation rate.  We
+ran a simulation at 1000 Hz for 12.6~ms (shorter than the 20~ms we
+used for the other runs).  We also used a larger domain, $1.8432\times
+10^5~\mathrm{cm} \times 3.072\times 10^4~\mathrm{cm}$ to account for
+the larger Rossby radius.  Figure~\ref{fig:10_10_slow} shows the flame
+structure.  Overall it looks much like the faster rotator.  In the
+next section, we explore the effect of rotation on the flame
+acceleration.
+
+\begin{figure}[t]
+\centering
+\plotone{flame_wave_slow}
+\caption{\label{fig:10_10_slow} Temperature, mean molecular weight, energy generation rate, and out-of-plane velocity for the 10/10-1000 Hz simulation at 12.6~ms.}
+\end{figure}
 
@@
 
-Table \ref{table:flame_speeds_multid} gives the flame speed measured in each multidimensional simulation.
-The 2D flames propagate at speeds about an order of magnitude faster than their 1D counterparts (Table \ref{table:flame_speeds_1d}).
-The increase in flame speed is likely a product of the larger flame surface area and hydrodynamical effects such as turbulence,
-wrinkling, and convective cycles, which bring cooler fuel from ahead of the front into the hottest part of the burning region.
+Table \ref{table:flame_speeds_multid} gives the flame speed measured in each multidimensional
+simulation. The 2D flames propagate at speeds about an order of magnitude faster than their 1D
+counterparts (Table \ref{table:flame_speeds_1d}). The increase in flame speed is likely a product
+of the larger flame surface area and hydrodynamical effects such as turbulence, wrinkling, and
+convective cycles, which bring cooler fuel from ahead of the front into the hottest part of the
+burning region.
 

### 2019-12-02 xrb/paper.tex
finish my pass

@@
 
-We have shown the results of our fully hydrodynamical, multidimensional simulations of flame propagation through the atmosphere of a neutron star.
-By using the fully compressible hydrodynamics equations, we are able to capture the vertical dynamics of the system.
-To accurately model the flame propagation, it is necessary to sufficiently resolve the scale height of the atmosphere.
-It is also important that there is a thin transition between the underlying neutron star and the atmosphere, to ensure that the peak temperature occurs at the correct base density.
-
-In order to satisfy these requirements whilst allowing the simulations to be computationally feasible, we used several approximations: a higher than normal rotation rate, a boosted flame, a simplified reaction network, and a 2.5D model for the flow.
-The first two of these were used to reduce the model's spatial and temporal scales.
-Using the simulation framework developed here, these both can be relaxed in the future, at the cost of more computer time.
-The same goes for 2.5D
-vs.\ full 3D---the only difference is computer time, and our future
-calculations will explore the 3D evolution and compare to the 2.5D simulations presented here. In particular,
-in 3D we will be able to explore shear instabilities at the flame front.  We can also capture the 
-baroclinic instability \citep{Cavecchi2019} and the competition between it and shear.
-Larger networks are a straightforward change, and already supported in
-\castro\ using the \pynucastro\ framework~\citep{pynucastro} and JINA
-ReacLib rate database~\citep{reaclib}.
+We have shown the results of our fully hydrodynamical,
+multidimensional simulations of flame propagation through the
+atmosphere of a neutron star. By using the fully compressible
+hydrodynamics equations, we are able to capture the vertical dynamics
+of the system. To accurately model the flame propagation, it is
+necessary to sufficiently resolve the scale height of the
+atmosphere. It is also important that there is a thin transition
+between the underlying neutron star and the atmosphere, to ensure that
+the peak temperature occurs at the correct base density.
+
+In order to satisfy these requirements whilst allowing the simulations
+to be computationally feasible, we used several approximations: a
+higher than normal rotation rate, a boosted flame, a simplified
+reaction network, and a 2D axisymmetric model for the flow. The first two of
+these were used to reduce the model's spatial and temporal
+scales. Using the simulation framework developed here, these both can
+be relaxed in the future, at the cost of more computer time. The same
+goes for 2D axisymmetry vs.\ full 3D---the only difference is computer time, and
+our future calculations will explore the 3D evolution and compare to
+the 2D simulations presented here. In particular, in 3D we will be
+able to explore shear instabilities at the flame front.  We can also
+capture the baroclinic instability \citep{Cavecchi2019} and the
+competition between it and shear.  Larger networks are a
+straightforward change, and already supported in \castro\ using the
+\pynucastro\ framework~\citep{pynucastro} and JINA ReacLib rate
+database~\citep{reaclib}.
 
@@
 
-In addition to extending our models to 3D and relaxing the approximations used in 
-this study, in the future we plan to perform a number of additional further studies. 
-These include investigating the effects of ignition latitude and different initial 
-models. We also plan to model mixed H/He bursts. This will require a
-different reaction network, and perhaps different (more relaxed)
-resolution requirements. A long term goal is to include magnetohydrodynamics in our 
-models. Although the magnetic fields of neutron stars exhibiting X-ray bursts are 
-relatively weak, with $B \lesssim 10^8 - 10^9$G \citep{mukherjee2015magnetic} (at
- least compared to e.g. magnetars, which have $B\lesssim 10^{15}$G), it has been 
- found by \citet{art-2016-cavecchi-etal} that even weak magnetic fields could have 
- a non-negligible effect on the flame propagation, reducing confinement due to the 
- Coriolis force and leading to increased flame speeds. Ultimately, we wish to link 
- our simulations to observed light curves. To do this, we will need to explore 
- radiation transport in order to model how the burst energy propagates through the 
- outer layers of the neutron star atmosphere.
+In addition to extending our models to 3D and relaxing the
+approximations used in this study, in the future we plan to perform a
+number of additional further studies.  These include investigating the
+effects of ignition latitude and different initial models. We also
+plan to model mixed H/He bursts. This will require a different
+reaction network, and perhaps different (more relaxed) resolution
+requirements. A long term goal is to include magnetohydrodynamics in
+our models. Although the magnetic fields of neutron stars exhibiting
+X-ray bursts are relatively weak, with $B \lesssim 10^8 - 10^9$G
+\citep{mukherjee2015magnetic} (at least compared to e.g. magnetars,
+which have $B\lesssim 10^{15}$G), it has been found by
+\citet{art-2016-cavecchi-etal} that even weak magnetic fields could
+have a non-negligible effect on the flame propagation, reducing
+confinement due to the Coriolis force and leading to increased flame
+speeds. Ultimately, we wish to link our simulations to observed light
+curves. To do this, we will need to explore radiation transport in
+order to model how the burst energy propagates through the outer
+layers of the neutron star atmosphere.
 
@@
 \castro\ to GPUs, giving an order of magnitude performance boost on
-nodes with both CPUs and GPUs.  The GPU-enabled algorithm will
-increase the performance, allowing us to run bigger domains and/or
-eliminate the boosting used here.
+nodes with both CPUs and GPUs.  Flames without any boosting are currently running
+using the new GPU-enabled \castro.
 

### 2019-12-02 xrb/paper.tex
add convergence to the diffusion table

@@
 \usepackage{epsf,color,amsmath}
-
+\usepackage{multirow}
 \usepackage{cancel}
@@
 
-\begin{deluxetable}{rL}
+\begin{deluxetable}{rLc}
 	\tablecaption{\label{table:diffusion} Convergence of diffusion test problem.}
-	\tablehead{\colhead{base resolution} & \colhead{L-inf error}}
+	\tablehead{\colhead{base resolution} & \colhead{L-inf error} & \colhead{convergence rate\tablenotemark{a}}}
 	\startdata
 \cutinhead{no refinement}
-	32  & 5.995\times 10^{-3} \\
-        64  & 1.477\times 10^{-3} \\
-       128  & 3.664\times 10^{-4} \\
-       256  & 9.139\times 10^{-5} \\
+	32  & 5.995\times 10^{-3} & \multirow{2}{*}{2.04} \\
+        64  & 1.477\times 10^{-3} & \multirow{2}{*}{2.01} \\
+       128  & 3.664\times 10^{-4} & \multirow{2}{*}{2.00} \\
+       256  & 9.139\times 10^{-5} & \\
 \cutinhead{one level of refinement}
-        32  & 1.480\times 10^{-3} \\
-        64  & 3.685\times 10^{-4} \\
-       128  & 9.406\times 10^{-5} \\
-       256  & 2.446\times 10^{-5} \\
+        32  & 1.480\times 10^{-3} & \multirow{2}{*}{2.01} \\
+        64  & 3.685\times 10^{-4} & \multirow{2}{*}{1.97} \\
+       128  & 9.406\times 10^{-5} & \multirow{2}{*}{1.94} \\
+       256  & 2.446\times 10^{-5} & \\
 	\enddata
 	%
+\tablenotetext{a}{Convergence, $n$, is defined between 2 successive resolutions as $n = \log_2(e_{2h}/e_h)$,
+ where $e_{2h}$ is the error for the coarser resolution run and $e_h$ is the error for the finer resolution run.}
 \end{deluxetable}

### 2019-12-03 xrb/paper.tex
add Simons affiliation

@@
              Stony Brook, NY 11794-3800}
+\affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010}
 
@@
   flame\_wave}.  The work at Stony Brook was supported by DOE/Office
-of Nuclear Physics grant DE-FG02-87ER40317.  YC was supported by the
+of Nuclear Physics grant DE-FG02-87ER40317 and the SciDAC program DOE grant DE-SC0017955.  MZ acknowledges support
+from the Simons Foundation.  YC was supported by the
 European Union Horizon 2020 research and innovation program under

### 2019-12-03 xrb/paper.tex
addresses marginpar

@@
 $2\times$).  In both cases we see second-order convergence of the
-error.\MarginPar{This would be clearer if there was a third column showing the ratios of the errors?}  Figure~\ref{fig:diffusion} shows the temperature at the end of
+error.  Figure~\ref{fig:diffusion} shows the temperature at the end of
 the simulation of the $64\times 128$ base grid plus one level of

### 2019-12-06 xrb/paper.tex
final readthrough

@@
 At the densities we consider in this simulation, the pure He flame
-speed is quite slow and would require a prohibitively long integration
-time to see significant evolution of the burning.  We therefore
+speed is quite slow and would require long integration
+times to see significant evolution of the burning.  We therefore
 consider boosted flames in this first paper, to accelerate the burning
@@
 flame.  These will have different scale heights.  To create these
-models, we break the vertical structure of the atmosphere into 4
-layers: the underlying neutron star, a ramp-up to the base of the
-accreted atmosphere, a fuel layer representing the bulk of the
-atmosphere where the flame will propagate, and an outer, low density,
+models, we break the vertical structure of the atmosphere into four
+layers: (1) the underlying neutron star, (2) a ramp-up to the base of the
+accreted atmosphere, (3) a fuel layer representing the bulk of the
+atmosphere where the flame will propagate, and (4) an outer, low density,
 isothermal buffer above the atmosphere that allows us
@@
 \begin{equation}
-T = T_\star + \frac{1}{2} (T_\mathrm{hi} - T_\star) \left [ 1 + \tanh\left( \frac{\tilde{y}}{2 \delta_\mathrm{atm}} \right ) \right ]
+T(y) = T_\star + \frac{1}{2} (T_\mathrm{hi} - T_\star) \left [ 1 + \tanh\left( \frac{\tilde{y}}{2 \delta_\mathrm{atm}} \right ) \right ]
 \end{equation}
@@
 $T_\mathrm{lo}$, $\rho_\mathrm{int}$, $H_\star$,
-$\delta_\mathrm{atm}$, ${X_k}_\star$, and ${X_k}_\mathrm{atm}$.  We
+$\delta_\mathrm{atm}$, ${X_k}_\star$, ${X_k}_\mathrm{atm}$, and $g$.  We
 then integrate outwards from the base of the ramp region ($y =
@@
 of the burning layer.
+The parameters we use for
+the model generation are listed in Table~\ref{table:params} and the 
+initial model profiles are showing in Figure~\ref{fig:initial_models}.
 
@@
 $X_\mathrm{atm}(\isotm{He}{4})\tablenotemark{b}$ & \multicolumn{2}{c}{1.0} \\
+$g$             & \multicolumn{2}{c}{$-1.5\times 10^{14}~\mathrm{cm~s^{-2}}$} \\
 \enddata
@@
 
-We create two models, a ``cool'' model representing the atmosphere
-ahead of the flame and a ``hot'' (or perturbed) model that will
-initiate the flame by driving the burning.  \MarginPar{This has already been said at the start of this section} The parameters we use for
-the model generation are listed in Table~\ref{table:params}. \MarginPar{reference and discuss the figure of initial model here}
 We blend the hot and cold models laterally to produce the perturbation
@@
 quarter of the domain width.  The run at $20~\mbox{cm}$ resolution used
-one fewer level of refinement.  The slower rotating case using a
+one fewer level of refinement.  The slower rotating case (1000~Hz)  uses a
 slightly wider domain to accommodate the expected larger Rossby
@@
 with the heavier nuclei furthest behind the flame.  The boosting of
-the burning likely artificially increases this effect. \MarginPar{Doesn't the increased flame speed of the 10/10 run vs. the 5/5 run show this?}
+the burning likely artificially increases this effect, as we'll see in the 5/5 case below.
 
@@
 temperature at the refinement boundary due in part to the construction
-of the initial model using the fine grid resolution for HSE. \MarginPar{Remind the reader what this
-  was/where they should be looking in the plot to see this temperature
-  drop} As we will see in the lower resolution case, this does not
+of the initial model using the fine grid resolution for HSE.  As we will see in the lower resolution case, this does not
 affect the results.
@@
 Our default resolution puts $\sim 80$ zones vertically in the
-``cool'' \MarginPar{double check} model atmosphere height.  To
+``cool''  model atmosphere height.  To
 understand the effects of resolution, we also performed a run with one
@@
 run this out twice as long, the flame would have advanced to the
-position seen in our 10/10 runs.  The good agreement in the structure
+position seen in our 10/10 runs.  The ash is also not as evolved to as high of an $\bar{A}$.  The good agreement in the structure
 of the flame seen with the lower boosting gives us confidence that the
@@
 
-The final approximation to explore is our choice of rotation rate.  We
-ran a simulation at 1000 Hz for 12.6~ms (shorter than the 20~ms we
+The final approximation to explore is our choice of rotation rate.  We ran the 10/10-1000~Hz
+simulation for 12.6~ms (shorter than the 20~ms we
 used for the other runs).  We also used a larger domain, $1.8432\times
@@
 
-{\color{red} Show whatever it was that Don cooked up.}
-
-Baroclinicity is calculated as
+We explore the baroclinic instability by looking at the magnitude of the 
+baroclinicity, calculated as
 \begin{equation}
@@
 and shows the misalignment of the local density and pressure gradients
-\citep{Malone2014a}. As we are considering a 2D system, the
+(we also explored this in \citep{Malone2014a}). As we are considering a 2D system, the
 component of the baroclinicity we consider here (out of the plane)
@@
 
-\MarginPar{Relate our work to previous work}
-
 In addition to extending our models to 3D and relaxing the
@@
 plan to model mixed H/He bursts. This will require a different
-reaction network, and perhaps different (more relaxed) resolution
+reaction network, and perhaps different resolution
 requirements. A long term goal is to include magnetohydrodynamics in
@@
 nodes with both CPUs and GPUs.  Flames without any boosting are currently running
-using the new GPU-enabled \castro.
+using the new GPU-enabled \castro and will be the focus of the next study.
 
@@
 Gaussian, $r$ is the distance from the origin, and $D$ is the
-diffusion coefficient.  $T_1$ is the ambient temperature and $T_2$ is
+diffusion coefficient (see, e.g., \citet{SWE_MYRA09}).  $T_1$ is the ambient temperature and $T_2$ is
 the temperature of the peak of the Gaussian.  We use $D = 1$, $T_1 = 1$, $T_2 = 2$,

### 2019-12-06 xrb/paper.tex
Fixed wording so we no longer imply that we are using 1 grid per MPI rank.

@@
 		$10~\mathrm{cm}$ at the finest level, but subdomains containing approximately equal
-		numbers of zones that are each managed by a separate MPI process. A coarse base grid
+        numbers of zones that are distributed across MPI processes. A coarse base grid
 		extending to the upper boundary is drawn in white, with the green, blue, and black

### 2019-12-06 xrb/paper.tex
Make signs for g consistent in HSE equation and parameter table.
Fix typo.
Add MarginPar asking to rename the Flame Acceleration section.

@@
 \label{eq:hse}
-p_i = p_{i-1} + \frac{1}{2} \Delta y (\rho_i + \rho_{i-1}) g
+    p_i = p_{i-1} + \frac{1}{2} \Delta y (\rho_i + \rho_{i-1}) \gb \cdot \hat{\bf y}
 \end{equation}
@@
 $\rho_\mathrm{cutoff}$ & \multicolumn{2}{c}{$10^{-4}~\gcc$} \\
-$g$             & \multicolumn{2}{c}{$-1.5\times 10^{14}~\mathrm{cm~s^{-2}}$} \\
 $H_\star$       & \multicolumn{2}{c}{2000~cm} \\
@@
 $X_\mathrm{atm}(\isotm{He}{4})\tablenotemark{b}$ & \multicolumn{2}{c}{1.0} \\
-$g$             & \multicolumn{2}{c}{$-1.5\times 10^{14}~\mathrm{cm~s^{-2}}$} \\
+$\gb$             & \multicolumn{2}{c}{$-1.5\times 10^{14}~\mathrm{cm~s^{-2}} \hat{\bf y}$} \\
 \enddata
@@
 increasing $\bar{A}$.  By about $10~\mbox{ms}$, we see the flame front
-is reasonable well-defined.  We see that the bottom of the burning
+is reasonably well-defined.  We see that the bottom of the burning
 front is lifted off of the base of the atmosphere, greatly increasing
@@
 
-To measure the propagation rate of the burning front, we first collapse our nuclear
+To measure the propagation \MarginPar{Can we rename this section "Flame Propagation"? The burning front in our simulations seems to move at a constant velocity, and we don't discuss acceleration.}
+rate of the burning front, we first collapse our nuclear
 energy generation rate ($\dot{e}_\mathrm{nuc}$) data at each time into a 1D radial

### 2019-12-06 xrb/paper.tex
Fix wording where we talk about the relative flame speed for the 1000 Hz and 2000 Hz cases. We are saying 1000 Hz run goes 2x faster and in the next sentence mention the reduced Rossby radius. This 1000 Hz run has an increased Rossby radius relative to 2000 Hz, and a larger burning surface area consistent with the linear scaling between flame propagation speed and the ratio of the burning front area to vertical area in the Cavecchi approximate model for flame speed.

@@
 with the 1D laminar flames, and is the behavior predicted by Eq.~\ref{eq:flame_scaling}. The
-1000~Hz run goes 2 times faster than the 10/10 2000~Hz run, as expected from the reduced
-Rossby radius \citep{cavecchi:2013}. This confirms that the role of the Coriolis
+1000~Hz run goes 2 times faster than the 10/10 2000~Hz run, as expected from the
+inverse relation between rotation rate and the ratio of the burning front area to vertical area \citep{cavecchi:2013}.
+This confirms that the role of the Coriolis
 force is to limit the rate of flame spreading by the anticipated geometrical/hydrodynamical

### 2019-12-06 xrb/paper.tex
Add MarginPar to ask about magenta grid colors for Figure 1.

@@
 upper boundary at low resolution). As seen in Figure \ref{fig:grids}, 
+\MarginPar{Can we change the black grids to magenta? Black is hard to see against the dark red and black background.}
 we use up to 3 refinement levels in addition to the base grid, the first

### 2019-12-08 xrb3/paper.tex
stub for paper III
add NVIDIA ack

--- 2019-11-16 xrb2/paper.tex
@@
 %======================================================================
-\title{Dynamics of Laterally Propagating Flames in X-ray Bursts. II. Realistic Burning}
+\title{Dynamics of Laterally Propagating Flames in X-ray Bursts. III. Mixed H/He Flames}
 
-\shorttitle{Lateral Flame Dynamics II}
-\shortauthors{Harpole et al.}
+\shorttitle{Lateral Flame Dynamics III}
+%% \shortauthors{Harpole et al.}
 
-\author[0000-0002-1530-781X]{Alice Harpole}
-\affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
-             Stony Brook, NY 11794-3800}
+%% \author[0000-0002-1530-781X]{Alice Harpole}
+%% \affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
+%%              Stony Brook, NY 11794-3800}
 
-\author[0000-0001-6191-4285]{Kiran Eiden}
-\affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
-             Stony Brook, NY 11794-3800}
+%% \author[0000-0001-6191-4285]{Kiran Eiden}
+%% \affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
+%%              Stony Brook, NY 11794-3800}
 
-\author[0000-0001-8401-030X]{Michael Zingale}
-\affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
-             Stony Brook, NY 11794-3800}
+%% \author[0000-0001-8401-030X]{Michael Zingale}
+%% \affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
+%%              Stony Brook, NY 11794-3800}
 
-\author[0000-0003-2300-5165]{Donald Willcox}
-\affiliation{Lawrence Berkeley National Laboratory, Berkeley, CA}
+%% \author[0000-0003-2300-5165]{Donald Willcox}
+%% \affiliation{Lawrence Berkeley National Laboratory, Berkeley, CA}
 
-\author{Yuri Cavecchi}
-\affiliation{Mathematical Sciences and STAG Research Centre, University of Southampton, SO17 1BJ}
+%% \author{Yuri Cavecchi}
+%% \affiliation{Mathematical Sciences and STAG Research Centre, University of Southampton, SO17 1BJ}
 
-\author[0000-0003-0439-4556]{Max P.\ Katz}
-\affiliation{NVIDIA Corp}
+%% \author[0000-0003-0439-4556]{Max P.\ Katz}
+%% \affiliation{NVIDIA Corp}
 
-\author[0000-0001-8092-1974]{Weiqun Zhang}
-\affiliation{Lawrence Berkeley National Laboratory, Berkeley, CA}
+%% \author[0000-0001-8092-1974]{Weiqun Zhang}
+%% \affiliation{Lawrence Berkeley National Laboratory, Berkeley, CA}
 
-\correspondingauthor{Alice Harpole}
-\email{alice.harpolee@stonybrook.edu}
+%% \correspondingauthor{Alice Harpole}
+%% \email{alice.harpolee@stonybrook.edu}
 
@@
 Energy under Contract No. DE-AC05-00OR22725, awarded through the DOE
-INCITE program.  This research has made use of NASA's Astrophysics
-Data System Bibliographic Services.
+INCITE program.  We thank NVIDIA Corporation for the donation of a
+Titan X and Titan V GPU through their academic grant program.  This
+research has made use of NASA's Astrophysics Data System Bibliographic
+Services.
 

### 2019-12-09 xrb/paper.tex
Grammar

@@
 specific internal energy as $e = E - |\Ub|^2/2$.  The forcing in the
-momentum equation includes gravity, described by gravitation
-acceleration $\gb$ and rotational forces described by angular velocity
+momentum equation includes gravity, described by gravitational
+acceleration $\gb$, and rotational forces, described by angular velocity
 $\Omegab$, with $\rb$ the position vector from the origin.
@@
 accelerate due to hydrodynamics interactions (wrinkling, turbulence
-interactions, directed flows feeding fuel into the flame, ...).
+interactions, directed flows feeding fuel into the flame, etc.).
 
@@
 Figure~\ref{fig:time_series} shows the time evolution of the 10/10
-simulation, focusing on the mean-molecular weight,
+simulation, focusing on the mean molecular weight,
 \begin{equation}

### 2019-12-09 xrb/paper.tex
Updated grid vis

--- 2019-12-06 xrb/paper.tex
@@
         numbers of zones that are distributed across MPI processes. A coarse base grid
-		extending to the upper boundary is drawn in white, with the green, blue, and black
-		grids showing the jumps in refinement needed to fully resolve the fuel layer and
-		underlying neutron star.}
+		extending to the upper boundary is drawn in white, with the magenta, green, and
+		blue grids showing the jumps in refinement needed to fully resolve the fuel layer
+		and underlying neutron star.}
 \end{figure}
@@
 
-\subsection{Flame Acceleration}
-
-To measure the propagation \MarginPar{Can we rename this section "Flame Propagation"? The burning front in our simulations seems to move at a constant velocity, and we don't discuss acceleration.}
-rate of the burning front, we first collapse our nuclear
+\subsection{Flame Propagation}
+
+To measure the propagation rate of the burning front, we first collapse our nuclear
 energy generation rate ($\dot{e}_\mathrm{nuc}$) data at each time into a 1D radial
@@
 1000~Hz run goes 2 times faster than the 10/10 2000~Hz run, as expected from the
-inverse relation between rotation rate and the ratio of the burning front area to vertical area \citep{cavecchi:2013}.
+inverse relation between rotation rate and the ratio of Rossby length to scale height \citep{cavecchi:2013}.
 This confirms that the role of the Coriolis

### 2019-12-09 xrb/paper.tex
Modify wording in Flame Propagation section

@@
 1000~Hz run goes 2 times faster than the 10/10 2000~Hz run, as expected from the
-inverse relation between rotation rate and the ratio of Rossby length to scale height \citep{cavecchi:2013}.
+inverse relation between rotation rate and the ratio of burning front area to scale height \citep{cavecchi:2013}.
 This confirms that the role of the Coriolis

### 2019-12-09 xrb/paper.tex
keywords

--- 2019-12-09 xrb/paper.tex
@@
 
-\keywords{convection---hydrodynamics---methods: numerical---stars: neutron---X-rays: bursts}
+\keywords{X-ray bursts (1814), Nucleosynthesis (1131), Hydrodynamical simulations (767), Hydrodynamics (1963), Neutron stars (1108), Open source software (1866), Computational methods (1965)}
 
@@
         numbers of zones that are distributed across MPI processes. A coarse base grid
-		extending to the upper boundary is drawn in white, with the magenta, green, and
-		blue grids showing the jumps in refinement needed to fully resolve the fuel layer
-		and underlying neutron star.}
+		extending to the upper boundary is drawn in white, with the green, blue, and black
+		grids showing the jumps in refinement needed to fully resolve the fuel layer and
+		underlying neutron star.}
 \end{figure}
@@
 
-\subsection{Flame Propagation}
-
-To measure the propagation rate of the burning front, we first collapse our nuclear
+\subsection{Flame Acceleration}
+
+To measure the propagation \MarginPar{Can we rename this section "Flame Propagation"? The burning front in our simulations seems to move at a constant velocity, and we don't discuss acceleration.}
+rate of the burning front, we first collapse our nuclear
 energy generation rate ($\dot{e}_\mathrm{nuc}$) data at each time into a 1D radial
@@
 1000~Hz run goes 2 times faster than the 10/10 2000~Hz run, as expected from the
-inverse relation between rotation rate and the ratio of Rossby length to scale height \citep{cavecchi:2013}.
+inverse relation between rotation rate and the ratio of the burning front area to vertical area \citep{cavecchi:2013}.
 This confirms that the role of the Coriolis

### 2019-12-09 xrb/paper.tex
close marginpar

--- 2019-12-09 xrb/paper.tex
@@
 upper boundary at low resolution). As seen in Figure \ref{fig:grids}, 
-\MarginPar{Can we change the black grids to magenta? Black is hard to see against the dark red and black background.}
 we use up to 3 refinement levels in addition to the base grid, the first

### 2019-12-10 xrb2/paper.tex
add some todo
add NVIDIA ack
some refs

@@
 
+We use the simulation framework introduced in \citet{flame_wave1}.
+The main difference is that we no longer boost the flames by adjusting
+the reaction rates and conductivities.  By porting \castro\ to GPUs,
+we are able to run the unboosted flame simulations with fewer
+resources than the boosted simulations used in our earlier paper.
+
 %======================================================================
@@
 \section{Simulations and Results}\label{Sec:results}
+
+Things to show:
+
+vertical profiles of the atmosphere behind the flame---how does it compare to isentropic?
+
+what is the composition on the photosphere across the flame
+
+different latitudes
 

### 2019-12-14 MAESTROeX_II/paper.tex
add Don -- he's been doing all the SDC interface work
add ref

@@
 
+\author[0000-0003-2300-5165]{Donald Willcox}
+\affiliation{Lawrence Berkeley National Laboratory \\
+Center for Computational Sciences and Engineering \\
+One Cyclotron Road, MS 50A-3111 \\
+Berkeley, CA 94720, USA}
+
 \author[0000-0001-8401-030X]{Michael Zingale}
@@
 It has been shown to be more accurate and efficient than Strang splitting in a
-terrestrial, non-stratified, low Mach number reacting flow solver \cite{nonaka2012sdc},
-so we believe an SDC version of MAESTROeX would greatly improve the efficiency of the code.
+terrestrial, non-stratified, low Mach number reacting flow solver \cite{nonaka2012sdc}.
+SDC methods have also been demonstrated for compressible astrophysical flows \cite{castro_sdc}.
+Based on this past work, we believe an SDC version of MAESTROeX would greatly improve the efficiency of the code.
 \\

### 2019-12-15 xrb2/paper.tex
update ORCiD

@@
 
+different initial models?
+
 \acknowledgements \castro\ is open-source and freely available at
@@
   flame\_wave}.  The work at Stony Brook was supported by DOE/Office
-of Nuclear Physics grant DE-FG02-87ER40317.  This research used
-resources of the National Energy Research Scientific Computing Center,
-a DOE Office of Science User Facility supported by the Office of
-Science of the U.~S.\ Department of Energy under Contract
-No.\ DE-AC02-05CH11231.  This research used resources of the Oak Ridge
-Leadership Computing Facility at the Oak Ridge National Laboratory,
-which is supported by the Office of Science of the U.S. Department of
-Energy under Contract No. DE-AC05-00OR22725, awarded through the DOE
-INCITE program.  We thank NVIDIA Corporation for the donation of a
-Titan X and Titan V GPU through their academic grant program.  This
-research has made use of NASA's Astrophysics Data System Bibliographic
-Services.
+of Nuclear Physics grant DE-FG02-87ER40317 and the SciDAC program DOE
+grant DE-SC0017955.  MZ acknowledges support from the Simons
+Foundation.  This research used resources of the National Energy
+Research Scientific Computing Center, a DOE Office of Science User
+Facility supported by the Office of Science of the U.~S.\ Department
+of Energy under Contract No.\ DE-AC02-05CH11231.  This research used
+resources of the Oak Ridge Leadership Computing Facility at the Oak
+Ridge National Laboratory, which is supported by the Office of Science
+of the U.S. Department of Energy under Contract No. DE-AC05-00OR22725,
+awarded through the DOE INCITE program.  We thank NVIDIA Corporation
+for the donation of a Titan X and Titan V GPU through their academic
+grant program.  This research has made use of NASA's Astrophysics Data
+System Bibliographic Services.
 

### 2020-01-01 MAESTROeX_II/paper.tex
add ref

@@
 Define $Q_{\rho X_k}^{(0)}$ as the right hand side of (\ref{eq:sdc species 2}) without the $I_{\rho X_k}^{\rm lagged}$ term, and define $Q_{\rho h}^{(0)}$ as the right hand side of (\ref{eq:sdc enthalpy 2}) without the $I_{\rho h}^{\rm lagged}$ term.
-Use {\tt VODE} to integrate (\ref{eq:species}) and (\ref{eq:enthalpy}) over $\Delta t$ to advance $(\rho X_k, \rho h)^n$ to $(\rho X_k, \rho h)^{n+1,(0)}$ using the piecewise-constant advection and diffusion source terms:
+Use {\tt VODE} \citep{vode} to integrate (\ref{eq:species}) and (\ref{eq:enthalpy}) over $\Delta t$ to advance $(\rho X_k, \rho h)^n$ to $(\rho X_k, \rho h)^{n+1,(0)}$ using the piecewise-constant advection and diffusion source terms:
 \begin{eqnarray}

### 2020-01-02 xrb2/paper.tex
move over keywords

@@
 
-\author[0000-0001-8092-1974]{Weiqun Zhang}
-\affiliation{Lawrence Berkeley National Laboratory, Berkeley, CA}
-
 \correspondingauthor{Alice Harpole}
@@
 
-We use the simulation framework introduced in \citet{flame_wave1}.
+We use the Castro hydrodynamics code \citep{castro} and the simulation framework introduced in \citet{flame_wave1}.
 The main difference is that we no longer boost the flames by adjusting

### 2020-01-05 grand_challenge/astro_grand_challenge.tex
nsf white paper

\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}

\usepackage{mathpazo}

\usepackage{fancyhdr}
\pagestyle{fancy}
\renewcommand{\headrulewidth}{0pt}

\rfoot{\footnotesize \em Future Directions in Extreme Scale\\ Computing for Scientific Grand Challenges}

\begin{document}

\begin{center}
{\Large Challenges in Modeling Astrophysical Thermonuclear Explosions} \\[0.25em]
Michael Zingale (Stony Brook)
\end{center}

The end states of stellar evolution result in exotic compact objects:
white dwarfs, neutron stars, and black holes.  When these remnants are
in binary systems, interactions with the companion star can lead to a
variety of different thermonuclear explosions: classical novae, X-ray
bursts (XRBs), and thermonuclear (Type Ia) supernovae (SN Ia).  These
explosions are important sites of nucleosynthesis in the galaxy.  XRBs
allow us to probe the physics of dense nuclear matter, and SN Ia are
important distant indicators for measuring the expansion history of
the Universe.  Simulations on these events has been ongoing for
decades and allow us to understand the physics of the explosion,
interpret the observations, and test different progenitor models for
viability.

Modeling these events require multiscale and multiphysics simulations.
Spatial scales range from the size of the white dwarf or neutron star
down to the burning or dissipation scale.  Temporary scales range from
the millions of years (SN Ia) to hours (XRBs) of evolution that
brought the star to the brink of explosions to the seconds-long
explosion.  Models must incorporate hydrodynamics, nuclear reactions,
gravity, magnetic fields, and radiation.

algorithms: hydrodynamics, finite-volume, AMR

new developments: low speed / implicit, integration techniques, coupling

performance: GPUs \& CPUs

Example: XRBs

\end{document}

### 2020-01-05 grand_challenge/astro_grand_challenge.tex
update after a pass

@@
 
+\usepackage{hyperref}
 \usepackage{mathpazo}
@@
 variety of different thermonuclear explosions: classical novae, X-ray
-bursts (XRBs), and thermonuclear (Type Ia) supernovae (SN Ia).  These
-explosions are important sites of nucleosynthesis in the galaxy.  XRBs
-allow us to probe the physics of dense nuclear matter, and SN Ia are
+bursts (XRBs), and thermonuclear (Type Ia) supernovae (SN Ia).  SN Ia
+and novae are important sites of nucleosynthesis in the galaxy; XRBs
+allow us to probe the physics of dense nuclear matter; and SN Ia are
 important distant indicators for measuring the expansion history of
-the Universe.  Simulations on these events has been ongoing for
-decades and allow us to understand the physics of the explosion,
+the Universe.  Simulations on these events have been ongoing for
+decades and allow us to understand the physics of the explosions,
 interpret the observations, and test different progenitor models for
-viability.
+viability.  Next generation supercomputers will allow us to push these
+simulations to unprecedented realism.
 
-Modeling these events require multiscale and multiphysics simulations.
-Spatial scales range from the size of the white dwarf or neutron star
+Modeling these events requires multiscale and multiphysics simulations.
+Spatial scales range from the size of the compact object
 down to the burning or dissipation scale.  Temporary scales range from
@@
 explosion.  Models must incorporate hydrodynamics, nuclear reactions,
-gravity, magnetic fields, and radiation.
+gravity, thermal conduction, magnetic fields, and radiation.
 
-algorithms: hydrodynamics, finite-volume, AMR
+While a variety of approximations are used in astrophysics to model
+hydrodynamic flows, for stellar explosions, conservative finite-volume
+methods are the most popular.  Adaptive mesh refinement (AMR) allows
+us to increase the range of lengthscales captures.  Our group has been
+developing the AMReX Astrophysics Suite of open-source AMR
+multiphysics hydrodynamics
+codes\footnote{\url{https://github.com/AMReX-Astro/}} for over a
+decade.  Two codes in particular address stellar explosions: MAESTROeX
+is a low Mach number hydrodynamics code for modeling the highly
+subsonic stellar convection leading up to explosion and Castro is a
+fully compressible radiation hydrodynamics code that efficiently
+models the explosive phase.  These codes share a common framework
+(adaptive mesh library, microphysics routines).
 
-new developments: low speed / implicit, integration techniques, coupling
+Advances in our understanding of these events have been made both
+through the development of new algorithms (e.g.\ low-speed solvers) as
+well as new computer architectures.  Today's supercomputers are
+dominated by GPU accelerators, as will the next generation of
+DOE supercomputers.  Over the past year,
+we've ported the AMReX Astrophysics Suite to GPUs, enabling a large
+performance gain and opening new areas for scientific exploration.  An
+important part of that process is performance portability.  Our path
+allows us to use the exact same compute kernels on CPUs and GPUs,
+eliminating the burden of maintaining redundant code paths for
+different architectures.
 
-performance: GPUs \& CPUs
+To highlight the way that new computer architectures enable new
+science, we focus on our current work on XRBs.  Our interest is in
+modeling the nucleosynthesis as a burning front propagates laterally
+through the accreted fuel layer on a neutron star.  Due to the range
+of length and timescales, various approximations have been used in the
+literature.  Our group started performing two-dimensional
+hydrodynamical simulations of the flame propagating using the NERSC
+Cori and OLCF Titan machines over a year ago.  These calculations were
+the first to perform full hydrodynamics in both the lateral and
+vertical directions, allowing us to understand the complex dynamics of
+the flame front with realistic microphysics.  The simulations were
+challenging, and we needed to make approximations to both the length
+scales (using artificially high neutron star rotation rates to compress
+the lateral scales) and time scales (boosting the flame speed).
+Nevertheless, we were able to complete the first resolved calculations
+of hydrodynamical flames in X-ray bursts.  Taking advantage of GPUs,
+we are now able to eliminate these approximations in our current runs on the
+OLCF Summit machine.  The same code is used,
+but running entirely on the GPUs, it is 5--10$\times$ faster than
+CPUs.  We are on target to perform the first three-dimensional
+simulations this year as well as to explore the more interesting case
+of mixed H/He bursts, which involve more complicated nuclear
+reactions.
 
-Example: XRBs
+The present generation of supercomputers allow us to do simulations
+that were simply not possible with the previous generation.  Future
+computers will give the same boost.  With the next generation
+machines, we will be able to routinely do 3D simulations, cover a
+larger portion of the neutron star surface, increase the size of our
+reaction networks, and couple radiation.These will be the first global
+simulations of burning fronts on neutron stars that capture the
+realistic physics and allow us to connect directly to observations.
+
+New architectures bring new programming challenges, and to date we've
+managed to keep pace with these developments.  A key point is that
+funding needs to be provided to aid in preparing codes, not just for
+the science, especially for novel architectures.  NSF should recognize
+this need.  To ensure the largest benefit to the community,
+application codes for these machines should be Open Source to the
+fullest extend allowed.
 

### 2020-01-05 grand_challenge/astro_grand_challenge.tex
some feedback

@@
 models the explosive phase.  These codes share a common framework
-(adaptive mesh library, microphysics routines).
+(adaptive mesh library, microphysics routines), are written in C++ and Fortran,
+and use MPI+OpenMP or MPI+CUDA.
 
@@
 hydrodynamical simulations of the flame propagating using the NERSC
-Cori and OLCF Titan machines over a year ago.  These calculations were
+Cori and OLCF Titan (on CPUs)  machines over a year ago.  These calculations were
 the first to perform full hydrodynamics in both the lateral and
@@
 the science, especially for novel architectures.  NSF should recognize
-this need.  To ensure the largest benefit to the community,
+this need.  The software stack on the next generation machine is as
+important as the hardware.  It should support the popular
+community-supported programming models today and have wide support for
+C/C++ and Fortran.  To ensure the largest benefit to the community,
 application codes for these machines should be Open Source to the

### 2020-01-12 grand_challenge/astro_grand_challenge.tex
change size
final version

@@
 \usepackage{hyperref}
-\usepackage{mathpazo}
+\usepackage{times}
+
+\usepackage[para]{footmisc}
 
@@
 \begin{center}
-{\Large Challenges in Modeling Astrophysical Thermonuclear Explosions} \\[0.25em]
+{\large Challenges in Modeling Astrophysical Thermonuclear Explosions} \\[0.25em]
 Michael Zingale (Stony Brook)
@@
 allow us to probe the physics of dense nuclear matter; and SN Ia are
-important distant indicators for measuring the expansion history of
+important distance indicators for measuring the expansion history of
 the Universe.  Simulations on these events have been ongoing for
@@
 Spatial scales range from the size of the compact object
-down to the burning or dissipation scale.  Temporary scales range from
+down to the burning or dissipation scale.  Temporal scales range from
 the millions of years (SN Ia) to hours (XRBs) of evolution that
-brought the star to the brink of explosions to the seconds-long
+brought the star to the brink of explosion to the seconds-long
 explosion.  Models must incorporate hydrodynamics, nuclear reactions,
@@
 methods are the most popular.  Adaptive mesh refinement (AMR) allows
-us to increase the range of lengthscales captures.  Our group has been
-developing the AMReX Astrophysics Suite of open-source AMR
+us to increase the range of length scales captured.  Our group develops
+the AMReX Astrophysics Suite of open-source AMR
 multiphysics hydrodynamics
-codes\footnote{\url{https://github.com/AMReX-Astro/}} for over a
-decade.  Two codes in particular address stellar explosions: MAESTROeX
+codes\footnote{\url{https://github.com/AMReX-Astro/}}.  Two codes in particular address stellar explosions: MAESTROeX
 is a low Mach number hydrodynamics code for modeling the highly
@@
 models the explosive phase.  These codes share a common framework
-(adaptive mesh library, microphysics routines), are written in C++ and Fortran,
+(AMR library, microphysics routines), are written in C++ and Fortran,
 and use MPI+OpenMP or MPI+CUDA.
@@
 through the accreted fuel layer on a neutron star.  Due to the range
-of length and timescales, various approximations have been used in the
-literature.  Our group started performing two-dimensional
+of length and time scales, various approximations have been used in the
+literature.  Our group started performing 2D
 hydrodynamical simulations of the flame propagating using the NERSC
@@
 the lateral scales) and time scales (boosting the flame speed).
-Nevertheless, we were able to complete the first resolved calculations
-of hydrodynamical flames in X-ray bursts.  Taking advantage of GPUs,
+  Taking advantage of GPUs,
 we are now able to eliminate these approximations in our current runs on the

### 2020-01-28 astronum_2019_Castro/astronum-castro-2019.tex
some work on addressing referees

@@
 
+\newcommand{\Ub}{\mathbf{U}}
+\newcommand{\Fb}{\mathbf{F}}
+
+\newcommand{\ddx}[1]{{\frac{{\partial#1}}{\partial x}}}
+\newcommand{\ddy}[1]{{\frac{{\partial#1}}{\partial y}}}
+\newcommand{\ddz}[1]{{\frac{{\partial#1}}{\partial z}}}
+\newcommand{\omegadot}{\dot{\omega}}
 \newcommand{\isot}[2]{$^{#2}\mathrm{#1}$}
@@
 hydrodynamic sources), and $\Rbs{\Uc}$ are the reactive source terms.
-The update is done at specific temporal nodes, the number of locations
-of which are picked to give the desired temporal accuracy.  The update
-from one time node $m$ to $m+1$ appears as:
+The conserved state has the form:
+\begin{equation}
+\Uc = (\rho, \rho X_k, \rho u, \rho v, \rho w, \rho E)^\intercal
+\end{equation}
+where $\rho$ is the mass density, $X_k$ are the mass fractions of the
+nuclear species, $\Ub = (u, v, w)$ is the Cartesian velocity vector,
+and $E$ is the total specific energy, related to the specific internal
+energy, $e$, as $E = e + |\Ub|^2/2$.  The advective term takes the form:
+\begin{equation}
+\Advs{\Uc} = -\nabla \cdot \Fb (\Ub) + {\bf H}
+\end{equation}
+where $\Fb(\Ub)$ are the hydrodynamical fluxes:
+\begin{equation}
+\Fb \cdot \hat{\bf x} = \left (\begin{array}{c} \rho u \\ \rho X_k u \\ \rho u^2 + p \\ \rho v u \\ \rho w u \\ (\rho E + p) u  \end{array} \right )
+\quad
+\Fb \cdot \hat{\bf y} = \left (\begin{array}{c} \rho v \\ \rho X_k v \\ \rho u v \\ \rho v^2 + p \\ \rho w v \\ (\rho E + p) v  \end{array} \right )
+\quad
+\Fb \cdot \hat{\bf z} = \left (\begin{array}{c} \rho w \\ \rho X_k w \\ \rho u w \\ \rho v w \\ \rho w^2 + p \\ (\rho E + p) w  \end{array} \right )
+\end{equation}
+Here, $p$ is the pressure, which can be evaluated through our equation
+of state, $p = p(\rho, e, X_k)$.  Note that we do not include
+transport terms in this system, but it is straightforward to add in a
+thermal diffusion flux.  ${\bf H}$ captures any hydrodynamical source
+terms (e.g., gravity).  Finally the reactive sources take the form:
+\begin{equation}
+\Rbs{\Uc} = (0, \rho \omegadot_k, 0, 0, 0, \rho \dot{\epsilon})^\intercal
+\end{equation}
+where $\omegadot_k$ is the creation rate of species $k$ and $\dot{\epsilon}$
+is the specific energy generation rate from reactions.
+
+In the SDC formalism~\cite{dutt:2000,minion:2003}, we start by writing our system in integral form:
+\begin{equation}
+\Uc^{n+1} = \Uc^n + \int_{t^n}^{t^{n+1}} \left [\Advs{\Uc} + \Rbs{\Uc} \right ] dt
+\end{equation}
+We next introduce the idea of iteratively approximating the solution.
+We'll use the superscript $k$ to denote the current iteration and
+$k+1$ for the new iteration.  We rewrite the integral as:
+\begin{align}
+\Uc^{n+1,(k+1)} = \Uc^{n,(k+1)} +
+   &\int_{t^n}^{t^{n+1}} \left \{ \left [ \Advs{\Uc}^{(k+1)} - \Advs{\Uc}^{(k)} \right ]
+   + \left [ \Rbs{\Uc}^{(k+1)} - \Rbs{\Uc}^{(k)} \right ] \right \} dt + \nonumber \\
+  & \int_{t^n}^{t^{n+1}} \left [\Advs{\Uc}^{(k)} + \Rbs{\Uc}^{(k)} \right ] dt
+\end{align}
+For the first integral, we approximate these terms using a first order
+update.  For advection, we simply use the value at the starting time
+(analogous to a forward Euler approximation) and for the reactions, we
+treat them implicitly by using the value at the ending time (analogous
+to a backward Euler approximation).  To gain temporal accuracy, we
+further subdivide the time into $M+1$ nodes, and perform this update
+over all time nodes for each iteration.  Denoting the time nodes with
+the index $m$, the updated from one time node $m$ to $m+1$ appears as:
     \begin{align}
@@
    \end{align}
-We use a finite-volume formalism, so $\avg{\Uc}$ is the spatial average of the
-conserved state in a zone, $\avg{\Advs{\Uc}}^m$ is the average
-advective term in a zone at time $t^m$, and $\avg{\Rbs{\Uc}}^{m+1}$
-the average of the reactive source at time $t^{m+1}$.  The second
-superscript on each term, $(k)$ or $(k+1)$, is the iteration counter.  This
-update is an implicit equation for the new state,
-$\avg{\Uc}^{{m+1},(k+1)}$.  The last term in the update is an integral
-over the sources constructed from the previous iteration's values at
-each time node.  Each iteration of the SDC method increases the
-temporal order of accuracy by one, up to the order of accuracy with
-which the integral is constructed.
+where $\delta t_m = t^{m+1} - t^m$ is the time difference between nodes.
+We use a finite-volume formalism, so $\avg{\Uc}$ is the spatial
+average of the conserved state in a zone, $\avg{\Advs{\Uc}}^m$ is the
+average advective term in a zone at time $t^m$, and
+$\avg{\Rbs{\Uc}}^{m+1}$ the average of the reactive source at time
+$t^{m+1}$.  The second superscript on each term, $(k)$ or $(k+1)$, is
+the iteration counter.  This update is an implicit equation for the
+new state, $\avg{\Uc}^{{m+1},(k+1)}$.  The last term in the update is
+an integral over the sources constructed from the previous iteration's
+values at each time node.  Each iteration of the SDC method increases
+the temporal order of accuracy by one, up to the order of accuracy
+with which the integral is constructed.
 
@@
 energy release and nucleosynthesis.  Our first set of two-dimensional simulations
-(\cite{astronum:2018}, Eiden et al.\ in preparation) used two physical approximations to
+(\cite{astronum:2018,eiden:2019} used two physical approximations to
 make the problem more tractable.  First, we used a higher rotation

### 2020-01-28 astronum_2019_Castro/astronum-castro-2019.tex
more cleaning for referee

@@
 further subdivide the time into $M+1$ nodes, and perform this update
-over all time nodes for each iteration.  Denoting the time nodes with
+over all time nodes for each iteration.  Finally, in \castro\ we use a finite-volume
+discretion, so we'll write the state as $\avg{\Uc}$ to indicate that it is an average over the zone.
+Denoting the time nodes with
 the index $m$, the updated from one time node $m$ to $m+1$ appears as:
@@
 where $\delta t_m = t^{m+1} - t^m$ is the time difference between nodes.
-We use a finite-volume formalism, so $\avg{\Uc}$ is the spatial
-average of the conserved state in a zone, $\avg{\Advs{\Uc}}^m$ is the
+Here, $\avg{\Advs{\Uc}}^m$ is the
 average advective term in a zone at time $t^m$, and

### 2020-01-31 astronum_2019_Castro/astronum-castro-2019.tex
update convergence

@@
 \newcommand{\isot}[2]{$^{#2}\mathrm{#1}$}
+\newcommand{\isotm}[2]{{}^{#2}\mathrm{#1}}
+
 \newcommand{\avg}[1]{{\left \langle #1 \right \rangle}}
@@
 the expected convergence on a wide variety of problems in
-\cite{castro:sdc}.  At the moment, the method is limited to single
+\cite{castro:sdc}.  Table~\ref{table:convergence} summaries the
+convergence of the standard unsplit CTU PPM hydrodynamics algorithm
+with Strang splitting and the second- and fourth-order SDC methods on
+a reacting acoustic pulse test problem.
+
+At the moment, the method is limited to single
 levels, but work is underway to extend this methodology to AMR.  For
@@
 integration technique in the future.
+
+\begin{table}[t]
+\renewcommand{\arraystretch}{1.25}
+\centering
+\begin{tabular}{llllll}
+{\bf field} & {$\epsilon_{64 \rightarrow 128}$} & 
+{\bf rate} & {$\epsilon_{128\rightarrow 256}$} & 
+{\bf rate} & {$\epsilon_{256\rightarrow 512}$} \\
+\hline
+\multicolumn{6}{c}{\bf Strang + CTU} \\
+\hline
+ $\rho X(\isotm{C}{12})$     & $1.717 \times 10^{17}$  & 1.945  & $4.458 \times 10^{16}$  & 2.194  & $9.745 \times 10^{15}$  \\
+ $\rho X(\isotm{O}{16})$     & $1.717 \times 10^{14}$  & 1.648  & $5.479 \times 10^{13}$  & 1.898  & $1.471 \times 10^{13}$  \\
+\hline
+\multicolumn{6}{c}{\bf 2nd order SDC} \\
+\hline
+ $\rho X(\isotm{C}{12})$     & $1.042 \times 10^{17}$  & 2.032  & $2.546 \times 10^{16}$  & 2.019  & $6.281 \times 10^{15}$  \\
+ $\rho X(\isotm{O}{16})$     & $1.564 \times 10^{14}$  & 1.935  & $4.090 \times 10^{13}$  & 2.003  & $1.020 \times 10^{13}$  \\
+\hline
+\multicolumn{6}{c}{\bf 4th order SDC} \\
+\hline
+ $\rho X(\isotm{C}{12})$     & $8.789 \times 10^{15}$  & 3.798  & $6.319 \times 10^{14}$  & 3.911  & $4.201 \times 10^{13}$  \\
+ $\rho X(\isotm{O}{16})$     & $1.294 \times 10^{13}$  & 3.765  & $9.518 \times 10^{11}$  & 3.872  & $6.501 \times 10^{10}$  \\
+\hline
+\end{tabular}
+\renewcommand{\arraystretch}{1.0}
+\caption{\label{table:convergence} Convergence of carbon mass density
+  ($\rho X(\isotm{C}{12})$) and oxygen mass density ($\rho
+  X(\isotm{O}{16})$) for the Strang-CTU algorithm and 2nd and 4th
+  order SDC.  Data taken from \cite{castro:sdc}.}
+\end{table}
 

### 2020-02-03 astronum_2019_Castro/astronum-castro-2019.tex
update ref response

@@
 with Strang splitting and the second- and fourth-order SDC methods on
-a reacting acoustic pulse test problem.
+a reacting acoustic pulse test problem modeling the 3-$\alpha$ and
+$\isotm{C}{12}(\alpha,\gamma)\isotm{O}{16}$ reactions.  We see that
+the Strang CTU method has trouble achieving second order accuracy,
+while the second-order SDC method converges second order.  The
+fourth-order SDC method converges almost fourth order overall.  Full
+details of this problem and more extensive convergence results are
+given in \cite{castro:sdc}.
 
@@
 \section{Example Application: X-ray bursts}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=\linewidth]{flame_wave_smallplt131070_slice}
+\caption{\label{fig:xrb} A snapshot of a 20 cm XRB flame calculation showing
+the mean molecular weight.  The flame is moving from left to right, and we see
+a gradient in the ash behind the flame, as the burning makes heavier nuclei.  
+Calculation from \cite{eiden:2019}.}
+\end{figure}
 

### 2020-02-04 xrb/submit/paper.tex
suybmitted version
A few basic changes to address some comments by the referee

--- 2019-12-09 xrb/paper.tex
@@
 the lateral differences needed for flame spreading, they can help
-understand the role that convection plays in distributing the initial
+us to understand the role that convection plays in distributing the initial
 burning products vertically throughout the neutron star atmosphere as well
@@
 $\dot{e}_\mathrm{nuc}$ value across all profiles to provide a fixed reference point.
-We define the position of the flame front to be the location ahead of hottest part of
+We define the position of the flame front to be the location to the right of hottest part of
 the flame where $\dot{e}_\mathrm{nuc}$ first drops to $< 0.1 \%$ of the global maximum. 
@@
 \centering
-\plotone{plots/results/speedplot_all.pdf}
+\plotone{speedplot_all.pdf}
 \caption{\label{fig:flame_speed}The position of the burning front for each simulation run as a function of time. The dashed lines show linear least squares fits for $t \gtrsim 6$ ms.}
@@
 and shows the misalignment of the local density and pressure gradients
-(we also explored this in \citep{Malone2014a}). As we are considering a 2D system, the
+(we also explored this in \citealt{Malone2014a}). As we are considering a 2D system, the
 component of the baroclinicity we consider here (out of the plane)
@@
 \centering
-\plotone{plots/results/baroclinicity}
+\plotone{baroclinicity}
 \caption{\label{fig:baroclinicity} Baroclinicity. This plot shows
@@
 \centering
-\plottwo{plots/results/vx_vy_129139}{plots/results/u_rho_129139}
+\plottwo{vx_vy_129139}{u_rho_129139}
 \caption{\label{fig:phase_plots} Phase plots at time $t = 0.02~\mathrm{s}$.
@@
 \centering
-\plotone{plots/results/abar_temp_129139}
+\plotone{abar_temp_129139}
 \caption{\label{fig:abar_temp} Phase plot of energy generation rate as a function of $\bar{A}$ and temperature at time $t = 0.02~\mathrm{s}$.}

### 2020-02-04 xrb/submit/paper.tex
Added new data for the 1000 Hz run
The flame speed dropped a bit - by about 10%. The true uncertainties on
the speeds should probably be a bit larger, but I am not certain how to
quantify them. Regardless, it is still nearly twice as fast as the
standard run.

@@
 	10/10-lowres & $9.33 \pm 0.04$ \\
-	10/10-1000 Hz & $18.6 \pm 0.16$ \\
+	10/10-1000 Hz & $16.5 \pm 0.21$ \\
 	\enddata
@@
 with the 1D laminar flames, and is the behavior predicted by Eq.~\ref{eq:flame_scaling}. The
-1000~Hz run goes 2 times faster than the 10/10 2000~Hz run, as expected from the
+1000~Hz run goes nearly 2 times faster than the 10/10 2000~Hz run, as expected from the
 inverse relation between rotation rate and the ratio of burning front area to scale height \citep{cavecchi:2013}.

### 2020-02-04 xrb/submit/paper.tex
Address comment about T_{lo} not being defined

@@
 $T_\mathrm{hi}$ is the highest temperature in the HSE model---it will
-represent the base of the fuel layer.
+represent the base of the fuel layer. We also define a floor temperature,
+$T_\mathrm{lo}$, which the atmosphere reaches and is held constant at in
+the outer, isothermal layer.
 
@@
 3\delta_\mathrm{atm}$.  Above this, we switch to an isentropic atmosphere until the
-temperature drops to a floor value, $T_\mathrm{lo}$, at which point we
+temperature drops to $T_\mathrm{lo}$, at which point we
 again keep the temperature constant.  The integration of the

### 2020-02-04 astronum_2019_Castro/astronum-castro-2019.tex
update
Fixed typo in Castro astronum proceedings.

@@
 over all time nodes for each iteration.  Finally, in \castro\ we use a finite-volume
-discretion, so we'll write the state as $\avg{\Uc}$ to indicate that it is an average over the zone.
+discretization, so we'll write the state as $\avg{\Uc}$ to indicate that it is an average over the zone.
 Denoting the time nodes with
@@
 boosted the speed of the flame to reduce the duration we need to
-model.  Our next set of calculations will relax these approximations.
+model.  Figure~\ref{fig:xrb} shows the structure of the flame.  Our next set of calculations will relax these approximations.
 

### 2020-02-07 astronum_2019_Castro/astronum-castro-2019.tex
Fix typo: updated -> update
fix typo
Fix typo: summaries -> summarizes

@@
 Denoting the time nodes with
-the index $m$, the updated from one time node $m$ to $m+1$ appears as:
+the index $m$, the update from one time node $m$ to $m+1$ appears as:
     \begin{align}
@@
 the expected convergence on a wide variety of problems in
-\cite{castro:sdc}.  Table~\ref{table:convergence} summaries the
+\cite{castro:sdc}.  Table~\ref{table:convergence} summarizes the
 convergence of the standard unsplit CTU PPM hydrodynamics algorithm

### 2020-02-07 astronum_2019_Castro/astronum-castro-2019.tex
Fix typo: Exascale Compute Project -> Exascale Computing Project

@@
 Physics grant DE-FG02-87ER40317 and contract 7418390 with Lawrence
-Berkeley National Laboratory as part of the Exascale Compute Project
+Berkeley National Laboratory as part of the Exascale Computing Project
 ExaStar collaboration.  This research was supported by the Exascale

### 2020-02-08 astronum_2019_Castro/astronum-castro-2019.tex
update table
update table caption

@@
 \begin{equation}
-\Advs{\Uc} = -\nabla \cdot \Fb (\Ub) + {\bf H}
+\Advs{\Uc} = -\nabla \cdot \Fb (\Uc) + {\bf H}
 \end{equation}
-where $\Fb(\Ub)$ are the hydrodynamical fluxes:
+where $\Fb(\Uc)$ are the hydrodynamical fluxes:
 \begin{equation}
@@
 \centering
-\begin{tabular}{llllll}
-{\bf field} & {$\epsilon_{64 \rightarrow 128}$} & 
-{\bf rate} & {$\epsilon_{128\rightarrow 256}$} & 
-{\bf rate} & {$\epsilon_{256\rightarrow 512}$} \\
+\begin{tabular}{lcc}
+{\bf field} & {$\log_2 \frac{\epsilon_{64 \rightarrow 128}}{\epsilon_{128\rightarrow 256}}$} &
+{$\log_2 \frac{\epsilon_{128\rightarrow 256}}{\epsilon_{256\rightarrow 512}}$} \\[0.5em]
 \hline
-\multicolumn{6}{c}{\bf Strang + CTU} \\
+\multicolumn{3}{c}{\bf Strang + CTU} \\
 \hline
- $\rho X(\isotm{C}{12})$     & $1.717 \times 10^{17}$  & 1.945  & $4.458 \times 10^{16}$  & 2.194  & $9.745 \times 10^{15}$  \\
- $\rho X(\isotm{O}{16})$     & $1.717 \times 10^{14}$  & 1.648  & $5.479 \times 10^{13}$  & 1.898  & $1.471 \times 10^{13}$  \\
+ $\rho X(\isotm{C}{12})$     & 1.945  & 2.194  \\
+ $\rho X(\isotm{O}{16})$     & 1.648  & 1.898  \\
 \hline
-\multicolumn{6}{c}{\bf 2nd order SDC} \\
+\multicolumn{3}{c}{\bf 2nd order SDC} \\
 \hline
- $\rho X(\isotm{C}{12})$     & $1.042 \times 10^{17}$  & 2.032  & $2.546 \times 10^{16}$  & 2.019  & $6.281 \times 10^{15}$  \\
- $\rho X(\isotm{O}{16})$     & $1.564 \times 10^{14}$  & 1.935  & $4.090 \times 10^{13}$  & 2.003  & $1.020 \times 10^{13}$  \\
+ $\rho X(\isotm{C}{12})$     & 2.032  & 2.019  \\
+ $\rho X(\isotm{O}{16})$     & 1.935  & 2.003  \\
 \hline
-\multicolumn{6}{c}{\bf 4th order SDC} \\
+\multicolumn{3}{c}{\bf 4th order SDC} \\
 \hline
- $\rho X(\isotm{C}{12})$     & $8.789 \times 10^{15}$  & 3.798  & $6.319 \times 10^{14}$  & 3.911  & $4.201 \times 10^{13}$  \\
- $\rho X(\isotm{O}{16})$     & $1.294 \times 10^{13}$  & 3.765  & $9.518 \times 10^{11}$  & 3.872  & $6.501 \times 10^{10}$  \\
+ $\rho X(\isotm{C}{12})$     & 3.798  & 3.911  \\
+ $\rho X(\isotm{O}{16})$     & 3.765  & 3.872  \\
 \hline
@@
   X(\isotm{O}{16})$) for the Strang-CTU algorithm and 2nd and 4th
-  order SDC.  Data taken from \cite{castro:sdc}.}
+  order SDC for a simple reacting acoustic pulse.  Here $\epsilon_{64\rightarrow 128}$ is the norm of the 
+  difference between a $64^2$ and $128^2$ simulation, and similar for the other $\epsilon$s.
+The convergence rate is the $\log_2$ of the ratio of errors between different resolutions.
+Data taken from \cite{castro:sdc}.}
 \end{table}

### 2020-02-11 xrb/submit/paper.tex
Added additional paragraph on Yuri's papers to intro

@@
 front, measuring flames up to 10 times faster than in the 2D case.
+
+However, the computational expense of the problem necessitated several approximations
+to make this set of simulations \citep{cavecchi:2013, art-2015-cavecchi-etal, art-2016-cavecchi-etal, Cavecchi2019}
+tractable. The first was the assumption of vertical hydrostatic equilibrium, which allowed them
+to filter out vertically propagating sound waves that would otherwise have limited the
+timestep. Although this enabled them to reach the simulation times (on the order of seconds) needed
+to capture the rise of the burst, they could not model vortical fluid motions driven by effects such
+as turbulence and convection. These can facilitate mixing and help to accelerate the propagation of
+the flame front. They also include only the triple-$\alpha$ process in their implementation of
+nuclear burning, neglecting the other reaction chains at play in a pure He burst. Our simulations
+relax the assumption of vertical hydrostatic equilibrium and implement a more complete set
+of nuclear reactions.
 

### 2020-02-18 xrb/submit/paper.tex
Update paragraph comparing with Yuri's papers

--- 2020-02-04 xrb/submit/paper.tex
@@
 front, measuring flames up to 10 times faster than in the 2D case.
+
+The computational expense of the problem necessitated several approximations
+to make this set of simulations \citep{cavecchi:2013, art-2015-cavecchi-etal, art-2016-cavecchi-etal, Cavecchi2019}
+tractable. The first was the assumption of vertical hydrostatic equilibrium, which
+allowed them to filter out vertically propagating sound waves that would otherwise
+have limited the timestep. This enabled them to reach the simulation times (on the
+order of seconds) needed to capture the rise of the burst, but they could not model
+vortical fluid motions driven by effects such as turbulence and convection. These can
+facilitate mixing and help to accelerate the propagation of the flame front. They also
+included only the triple-$\alpha$ process in their implementation of nuclear burning,
+and use a fixed, constant opacity across the entire domain for each simulation run (although
+they explore the dependence of the flame propagation speed on this opacity value).
+Our simulations relax the assumption of vertical hydrostatic equilibrium, and implement
+a more complete set of nuclear reactions and opacities. However, this comes at the expense
+of a different set of approximations.
 
@@
 generates vorticity, which in turn entrains material along the surface
-of the flame (see also \citealt{cavecchi:2013}). In 3D, this same vorticity
+of the flame (as seen in \citealt{cavecchi:2013}). In 3D, this same vorticity
 should perturb the flame front and further increase the flame speed

### 2020-02-18 xrb/submit/paper.tex
Added paragraph with flame speed comparison

@@
 Our simulations relax the assumption of vertical hydrostatic equilibrium, and implement
-a more complete set of nuclear reactions and opacities. However, this comes at the expense
-of a different set of approximations.
+a more complete set of nuclear reactions and opacities. However, as discussed later, this
+comes at the expense of a different set of approximations.
 
@@
 
+\citet{cavecchi:2013} ran simulations at rotation rates of $450$ Hz exploring a range of opacities.
+Accounting for the boosting, we would expect to see our 10/10 flame propagate at $\sim 2.25$ times
+the speed of these $450$ Hz runs, but instead the speeds range from $4.5$ to $7$ times those reported in
+\citet{cavecchi:2013}. The additional boost may come from cyclic fluid motions that would have been
+disallowed in the $450$ Hz runs by the enforcement of vertical hydrostatic equilibrium (we already saw
+that hydrodynamic effects can accelerate the burning front in our comparison with the 1D flames). The
+inclusion of reaction chains beyond may also increase energy generation and consequently the
+propagation speed.
+
 %Seeing the proofs both changing the Coriolis force AND the

### 2020-02-18 xrb/submit/paper.tex

@@
 %\documentclass[12pt, preprint]{aastex}
-\documentclass[preprint,times,tighten]{aastex63}
+\documentclass[trackchanges,preprint,times,tighten]{aastex63}
 
@@
 the lateral differences needed for flame spreading, they can help
-us to understand the role that convection plays in distributing the initial
+\added{us to} understand the role that convection plays in distributing the initial
 burning products vertically throughout the neutron star atmosphere as well
@@
 
-The computational expense of the problem necessitated several approximations
+\added{The computational expense of the problem necessitated several approximations
 to make this set of simulations \citep{cavecchi:2013, art-2015-cavecchi-etal, art-2016-cavecchi-etal, Cavecchi2019}
@@
 a more complete set of nuclear reactions and opacities. However, as discussed later, this
-comes at the expense of a different set of approximations.
+comes at the expense of a different set of approximations.}
 
@@
 $T_\mathrm{hi}$ is the highest temperature in the HSE model---it will
-represent the base of the fuel layer. We also define a floor temperature,
+represent the base of the fuel layer. \added{We also define a floor temperature,
 $T_\mathrm{lo}$, which the atmosphere reaches and is held constant at in
-the outer, isothermal layer.
+the outer, isothermal layer.}
 
@@
 3\delta_\mathrm{atm}$.  Above this, we switch to an isentropic atmosphere until the
-temperature drops to $T_\mathrm{lo}$, at which point we
+temperature drops to \deleted{a floor value, }$T_\mathrm{lo}$, at which point we
 again keep the temperature constant.  The integration of the
@@
 $\dot{e}_\mathrm{nuc}$ value across all profiles to provide a fixed reference point.
-We define the position of the flame front to be the location to the right of hottest part of
+We define the position of the flame front to be the location \replaced{ahead}{to the right} of hottest part of
 the flame where $\dot{e}_\mathrm{nuc}$ first drops to $< 0.1 \%$ of the global maximum. 
@@
 \begin{deluxetable}{lc}
-	\tablecaption{\label{table:flame_speeds_multid} Flame speeds measured in 2D calculations.}
+	\tablecaption{\label{table:flame_speeds_multid} Flame speeds measured in 2D calculations.\explain{Updated last entry with some additional data.}}
 	\tablehead{\colhead{run} & \colhead{$s_\mathrm{front}$ (km s$^{-1}$)}} 
@@
 \plotone{speedplot_all.pdf}
-\caption{\label{fig:flame_speed}The position of the burning front for each simulation run as a function of time. The dashed lines show linear least squares fits for $t \gtrsim 6$ ms.}
+\caption{\label{fig:flame_speed}The position of the burning front for each simulation run as a function of time.
+The dashed lines show linear least squares fits for $t \gtrsim 6$ ms.
+\explain{We've updated this with an additional ms of data from the 1000 Hz run.}}
 \end{figure}
@@
 with the 1D laminar flames, and is the behavior predicted by Eq.~\ref{eq:flame_scaling}. The
-1000~Hz run goes nearly 2 times faster than the 10/10 2000~Hz run, as expected from the
-inverse relation between rotation rate and the ratio of burning front area to scale height \citep{cavecchi:2013}.
+1000~Hz run goes \added{nearly} 2 times faster than the 10/10 2000~Hz run, as expected from the
+inverse relation between rotation rate and the ratio of burning front area to scale height
+\replaced{\citep{cavecchi:2013}}{(and as observed in \citealt{cavecchi:2013})}.
 This confirms that the role of the Coriolis
 force is to limit the rate of flame spreading by the anticipated geometrical/hydrodynamical
-effect ($s_\mathrm{front} = s_L * L_R / H$). Reducing the resolution had minimal impact on
-the flame speed -- the low resolution line is right on top of the fit for the standard run
-in Figure \ref{fig:flame_speed}, with their slopes differing by only a few percent. Using a
+effect ($s_\mathrm{front} = s_L * L_R / H$). Reducing the resolution
+had minimal impact on the flame speed -- the low resolution line is right on top of the fit for the
+standard run in Figure \ref{fig:flame_speed}, with their slopes differing by only a few percent. Using a
 smaller network also produces similar behavior to the standard run, although there is a small
@@
 
+\added{
 \citet{cavecchi:2013} ran simulations at rotation rates of $450$ Hz exploring a range of opacities.
@@
 inclusion of reaction chains beyond may also increase energy generation and consequently the
-propagation speed.
+propagation speed.}
 

### 2020-02-18 xrb/submit/paper.tex
update initial model

@@
 \begin{equation}
-T(y) = T_\star + \frac{1}{2} (T_\mathrm{hi} - T_\star) \left [ 1 + \tanh\left( \frac{\tilde{y}}{2 \delta_\mathrm{atm}} \right ) \right ]
+T(z) = T_\star + \frac{1}{2} (T_\mathrm{hi} - T_\star) \left [ 1 + \tanh\left( \frac{\tilde{z}}{2 \delta_\mathrm{atm}} \right ) \right ]
 \end{equation}
@@
 \begin{equation}
-\tilde{y} = y - H_\star - \frac{3}{2} \delta_\mathrm{atm}
+\tilde{z} = z - H_\star - \frac{3}{2} \delta_\mathrm{atm}
 \end{equation}
@@
 
-We specify the density, $\rho_\mathrm{int} = \rho(y = H_\star)$ as the
+We specify the density, $\rho_\mathrm{int} = \rho(z = H_\star)$ as the
 starting point for the integration of hydrostatic equilibrium.  This
@@
 $T_\mathrm{lo}$, $\rho_\mathrm{int}$, $H_\star$,
-$\delta_\mathrm{atm}$, ${X_k}_\star$, ${X_k}_\mathrm{atm}$, and $g$.  We
-then integrate outwards from the base of the ramp region ($y =
+$\delta_\mathrm{atm}$, ${X_k}_\star$, ${X_k}_\mathrm{atm}$, and $g$.
+We then integrate outwards from the base of the ramp region ($z =
 H_\star$), enforcing the discrete form of hydrostatic equilibrium,
-Eq.~\ref{eq:hse}.  Integrating upwards, we find $p_i$ and
-$\rho_i$ using a Newton-Raphson solver together with the equation of state, with either the temperature
-specified, $T_i = T(p_i, \rho_i, \{X_k\}_i)$ (in the isothermal, ramp,
-and buffer layers) or constant entropy, $s_i = s(p_i, \rho_i,
-\{X_k\}_i)$, in the fuel layer.  This follows the procedures described in \citet{ppm-hse}.  We
-use a constant temperature for all $y < H_\star +
-3\delta_\mathrm{atm}$.  Above this, we switch to an isentropic atmosphere until the
-temperature drops to \deleted{a floor value, }$T_\mathrm{lo}$, at which point we
-again keep the temperature constant.  The integration of the
-atmosphere continues until the density falls to a low density cutoff,
-$\rho_\mathrm{cutoff}$.  The material above this height is taken to
-have constant density and temperature.
-The choice of factors in front of $\delta_\mathrm{atm}$ were designed
-to make sure the peak $T$ is attained at the desired density
-of the burning layer.
-The parameters we use for
-the model generation are listed in Table~\ref{table:params} and the 
-initial model profiles are showing in Figure~\ref{fig:initial_models}.
+Eq.~\ref{eq:hse}.  Integrating upwards, we find $p_i$ and $\rho_i$
+using a Newton-Raphson solver together with the equation of state,
+with either the temperature specified, $T_i = T(p_i, \rho_i,
+\{X_k\}_i)$ (in the isothermal, ramp, and buffer layers) or constant
+entropy, $s_i = s(p_i, \rho_i, \{X_k\}_i)$, in the fuel layer.  This
+follows the procedures described in \citet{ppm-hse}.  We use a
+constant temperature for all $z < H_\star + 3\delta_\mathrm{atm}$.
+Above this, we switch to an isentropic atmosphere until the
+temperature drops to \deleted{a floor value, }$T_\mathrm{lo}$, at
+which point we again keep the temperature constant.  The integration
+of the atmosphere continues until the density falls to a low density
+cutoff, $\rho_\mathrm{cutoff}$.  The material above this height is
+taken to have constant density and temperature.  The choice of factors
+in front of $\delta_\mathrm{atm}$ were designed to make sure the peak
+$T$ is attained at the desired density of the burning layer.  The
+parameters we use for the model generation are listed in
+Table~\ref{table:params} and the initial model profiles are showing in
+Figure~\ref{fig:initial_models}.
 
@@
 $X_\mathrm{atm}(\isotm{He}{4})\tablenotemark{b}$ & \multicolumn{2}{c}{1.0} \\
-$\gb$             & \multicolumn{2}{c}{$-1.5\times 10^{14}~\mathrm{cm~s^{-2}} \hat{\bf y}$} \\
+$\gb$             & \multicolumn{2}{c}{$-1.5\times 10^{14}~\mathrm{cm~s^{-2}} \hat{\bf z}$} \\
 \enddata
@@
 \begin{align}
-p(x,y) &= f(x) p_\mathrm{hot}(y) + [1-f(x)] p_\mathrm{cool}(y) \\
-\rho(x,y) &= f(x) \rho_\mathrm{hot}(y) + [1-f(x)] \rho_\mathrm{cool}(y) \\
-X_k(x,y) &= f(x) {X_k}_\mathrm{hot}(y) + [1-f(x)] {X_k}_\mathrm{cool}(y)
+p(r,z) &= f(r) p_\mathrm{hot}(z) + [1-f(r)] p_\mathrm{cool}(z) \\
+\rho(r,z) &= f(r) \rho_\mathrm{hot}(z) + [1-f(r)] \rho_\mathrm{cool}(z) \\
+X_k(r,z) &= f(r) {X_k}_\mathrm{hot}(z) + [1-f(r)] {X_k}_\mathrm{cool}(z)
 \end{align}
@@
 \begin{equation}
-f(x) = \begin{cases}
-     1 & x < x_\mathrm{pert} \\
-   1 - \frac{x - x_\mathrm{pert}}{\delta_\mathrm{blend}} & x_\mathrm{pert} \le x \le x_\mathrm{pert} + \delta_\mathrm{blend} \\
-     0 & x > x_\mathrm{pert} + \delta_\mathrm{blend}
+f(r) = \begin{cases}
+     1 & r < r_\mathrm{pert} \\
+   1 - \frac{r - r_\mathrm{pert}}{\delta_\mathrm{blend}} & r_\mathrm{pert} \le r \le r_\mathrm{pert} + \delta_\mathrm{blend} \\
+     0 & r > r_\mathrm{pert} + \delta_\mathrm{blend}
 \end{cases}
@@
 equilibrum, the blended model is also in vertical equilibrium initially.  We choose
-$x_\mathrm{pert} = 1.024\times 10^4$~cm and $\delta_\mathrm{blend} = 2048$~cm.  Once
-the blended model is constructed, we compute $T(x,y)$ and $(\rho e)(x,y)$
+$r_\mathrm{pert} = 1.024\times 10^4$~cm and $\delta_\mathrm{blend} = 2048$~cm.  Once
+the blended model is constructed, we compute $T(r,z)$ and $(\rho e)(r,z)$
 from the equation of state,
 \begin{align}
-  T(x,y) &= T(\rho(x,y), p(x,y), X_k(x,y)) \\
-  (\rho e)(x,y) &= \rho(x, y) \cdot e(\rho(x,y), p(x,y), X_k(x,y)) 
+  T(r,z) &= T(\rho(r,z), p(r,z), X_k(r,z)) \\
+  (\rho e)(r,z) &= \rho(r, z) \cdot e(\rho(r,z), p(r,z), X_k(r,z)) 
 \end{align}

### 2020-02-25 xrb/submit/paper.tex
Rewrite equations in cylindrical coords

@@
 simulation plane, that participates in the Coriolis force (sometimes
-described as a 2.5D simulation).  We will take the
-\castro\ $x$-coordinate to be the cylindrical radial coordinate with
-corresponding velocity $u$, the \castro\ $y$-coordinate to be the
+described as a 2.5D simulation).  We will take $r$ to be the cylindrical
+radial coordinate with corresponding velocity $u$, $z$ to be the
 cylindrical vertical coordinate with corresponding velocity $v$, and the
-\castro\ $z$-coordinate to be the cylindrical azimuthal coordinate,
-with corresponding velocity $w$.  A righthanded coordinate system has
-positive $w$ pointing out of the page.  We will take $\Omegab =
-\Omega_0 \hat{\bf y}$ for the angular rotation rate, and $\gb = -g
-\hat{\bf y}$ for the gravitational acceleration, with $g$ constant.
-With these choices, the Coriolis force is:
+$\theta$ to be the cylindrical azimuthal coordinate with corresponding velocity
+$w$. As $r$ and $z$ define the simulation plane, we adopt the nonstandard ordering
+($r$, $z$, $\theta$), for which a righthanded coordinate system has positive $w$
+pointing out of the page.  We will take $\Omegab = \Omega_0 \hat{\bf z}$ for the
+angular rotation rate, and $\gb = -g \hat{\bf z}$ for the gravitational acceleration,
+with $g$ constant. With these choices, the Coriolis force is:
 \begin{equation}
 -2\rho \Omegab \times \Ub =
-   -2\rho \left ( \Omega_0 w \hat{x} - \Omega_0 u \hat{z} \right )
+   -2\rho \left ( \Omega_0 w \hat{\bf r} - \Omega_0 u \hat{\boldsymbol{\theta}} \right )
 \end{equation}
@@
 allows us to capture the geostrophic balance that sets up via lateral
-hydrostatic equilibrium \citep{spitkovsky2002}.  In the discussions
-below, we will use the Castro coordinate names, $(x, y)$ in our
-notation.
+hydrostatic equilibrium \citep{spitkovsky2002}.
 
@@
 \frac{\partial (\rho u)}{\partial t} + \nabla \cdot (\rho u \Ub) +
-     \frac{\partial p}{\partial x} &= -2\rho \Omega_0 w  \\
+     \frac{\partial p}{\partial r} &= -2\rho \Omega_0 w  \\
 %
 \frac{\partial (\rho v)}{\partial t} + \nabla \cdot (\rho v \Ub) +
-     \frac{\partial p}{\partial y} &= -\rho g \\
+     \frac{\partial p}{\partial z} &= -\rho g \\
 %
 \frac{\partial (\rho w)}{\partial t} + \nabla \cdot (\rho w \Ub) +
-     \cancelto{0}{\frac{\partial p}{\partial z}} &=
+     \frac{1}{r}\cancelto{0}{\frac{\partial p}{\partial \theta}} &=
     2\rho \Omega_0 u
 \end{align}
-where we cancel $\partial p/\partial z$ because there are no
+where we cancel $\partial p/\partial \theta$ because there are no
 variations in the azimuthal direction.  This allows us to recast the

### 2020-02-25 xrb/submit/paper.tex
Make opening sentence of paragraph comparing to Yuri's papers clearer

@@
 \added{The computational expense of the problem necessitated several approximations
-to make this set of simulations \citep{cavecchi:2013, art-2015-cavecchi-etal, art-2016-cavecchi-etal, Cavecchi2019}
-tractable. The first was the assumption of vertical hydrostatic equilibrium, which
-allowed them to filter out vertically propagating sound waves that would otherwise
+to make the simulations of \citet{cavecchi:2013, art-2015-cavecchi-etal, art-2016-cavecchi-etal}
+and \citet{Cavecchi2019} tractable. The first was the assumption of vertical hydrostatic equilibrium,
+which allowed them to filter out vertically propagating sound waves that would otherwise
 have limited the timestep. This enabled them to reach the simulation times (on the

### 2020-02-25 xrb/submit/paper.tex
Change coordinate in HSE equation for lower boundary
latest changes
Fixed typo, added a few comments to diff

@@
 \section{Numerical Approach}\label{Sec:numerics}
+
+\explain{We've updated the equations in this section to use cylindrical
+coordinates.}
 
@@
 $\theta$ to be the cylindrical azimuthal coordinate with corresponding velocity
-$w$. As $r$ and $z$ define the simulation plane, we adopt the nonstandard ordering
-($r$, $z$, $\theta$), for which a righthanded coordinate system has positive $w$
+$w$. \added{As $r$ and $z$ define the simulation plane, we adopt the nonstandard ordering
+($r$, $z$, $\theta$)}, for which a righthanded coordinate system has positive $w$
 pointing out of the page.  We will take $\Omegab = \Omega_0 \hat{\bf z}$ for the
@@
 \begin{equation}
-\Delta t_\mathrm{diff} \le \frac{1}{2} \frac{\Delta x^2}{\mathcal{D}}
-\end{equation}
-where $\mathcal{D} = \kth/(\rho c_v)$ is the thermal diffusivity.  The
+\Delta t_\mathrm{diff} \le \frac{1}{2} \frac{h^2}{\mathcal{D}}
+\end{equation}
+where $\mathcal{D} = \kth/(\rho c_v)$ is the thermal diffusivity and $h$ is the width of a cell (either $\Delta r$ or $\Delta z$).  The
 diffusivity increases rapidly at the top of the atmosphere, causing
@@
 \label{eq:hse}
-    p_i = p_{i-1} + \frac{1}{2} \Delta y (\rho_i + \rho_{i-1}) \gb \cdot \hat{\bf y}
+    p_i = p_{i-1} + \frac{1}{2} \Delta z (\rho_i + \rho_{i-1}) \gb \cdot \hat{\bf z}
 \end{equation}
@@
 \section{Initial Model}\label{Sec:inital_model}
+\explain{We've updated the equations in this section to use cylindrical
+coordinates.}
 
@@
 that hydrodynamic effects can accelerate the burning front in our comparison with the 1D flames). The
-inclusion of reaction chains beyond may also increase energy generation and consequently the
+inclusion of reaction chains beyond triple-$\alpha$ may also increase energy generation and consequently the
 propagation speed.}

### 2020-02-27 xrb/paper.tex
Clarify wording where we specify where in the initial model we set the temperature vs the entropy.

@@
 $\rho_i$ using a Newton-Raphson solver together with the equation of state, with either the temperature
-specified, $T_i = T(p_i, \rho_i, \{X_k\}_i)$ (in the isothermal, ramp,
-and buffer layers) or constant entropy, $s_i = s(p_i, \rho_i,
+specified, $T_i = T(p_i, \rho_i, \{X_k\}_i)$ (in the neutron star, ramp,
+and isothermal buffer layers) or constant entropy, $s_i = s(p_i, \rho_i,
 \{X_k\}_i)$, in the fuel layer.  This follows the procedures described in \citet{ppm-hse}.  We

### 2020-02-27 xrb/paper.tex
Added sentence discussing running the 10/10 iso7 simulation to 15 ms instead of 20 ms with aprox13.
Add space after the typeset code name "Castro" in section 6.

@@
 case.  We see in the next section that these two networks give largely
-the same flame speed.  The reduced network size saves a lot of memory,
+the same flame speed.
+We only ran the 10/10 simulation with the {\tt iso7} reaction network to 15~ms,
+and we observe this was sufficient to establish its steady flame speed for comparison to our 10/10 simulation with the {\tt aprox13} network.
+The reduced network size saves a lot of memory,
 which will be useful when we transition to 3D simulations.
@@
 nodes with both CPUs and GPUs.  Flames without any boosting are currently running
-using the new GPU-enabled \castro and will be the focus of the next study.
+using the new GPU-enabled \castro\ and will be the focus of the next study.
 

### 2020-02-27 xrb/paper.tex
Add space between magnetic field values and the G symbol for Gauss in section 6.

@@
 our models. Although the magnetic fields of neutron stars exhibiting
-X-ray bursts are relatively weak, with $B \lesssim 10^8 - 10^9$G
+X-ray bursts are relatively weak, with $B \lesssim 10^8 - 10^9$~G
 \citep{mukherjee2015magnetic} (at least compared to e.g. magnetars,
-which have $B\lesssim 10^{15}$G), it has been found by
+which have $B\lesssim 10^{15}$~G), it has been found by
 \citet{art-2016-cavecchi-etal} that even weak magnetic fields could

### 2020-03-03 xrb/submit/paper.tex
Incorporate Yuri's comments

@@
 to make the simulations of \citet{cavecchi:2013, art-2015-cavecchi-etal, art-2016-cavecchi-etal}
-and \citet{Cavecchi2019} tractable. The first was the assumption of vertical hydrostatic equilibrium,
-which allowed them to filter out vertically propagating sound waves that would otherwise
-have limited the timestep. This enabled them to reach the simulation times (on the
-order of seconds) needed to capture the rise of the burst, but they could not model
-vortical fluid motions driven by effects such as turbulence and convection. These can
-facilitate mixing and help to accelerate the propagation of the flame front. They also
-included only the triple-$\alpha$ process in their implementation of nuclear burning,
-and use a fixed, constant opacity across the entire domain for each simulation run (although
-they explore the dependence of the flame propagation speed on this opacity value).
+and \citet{Cavecchi2019} tractable. The first of these approximations was the assumption of
+vertical hydrostatic equilibrium, which allowed them to use a less refined grid and filter out
+vertically propagating sound waves that would otherwise have limited the timestep. This
+enabled them to reach the simulation times (on the order of seconds) needed to capture the rise
+of the burst, but they could not fully resolve vortical fluid motions driven by effects such as
+turbulence and convection. These can facilitate mixing and help to accelerate the propagation of
+the flame front. They also included only the triple-$\alpha$ process in their implementation of
+nuclear burning, and used a fixed, constant opacity across the entire domain for each simulation
+run (although they explored the dependence of the flame propagation speed on this opacity value).
 Our simulations relax the assumption of vertical hydrostatic equilibrium, and implement
@@
 \citet{cavecchi:2013}. The additional boost may come from cyclic fluid motions that would have been
-disallowed in the $450$ Hz runs by the enforcement of vertical hydrostatic equilibrium (we already saw
+underresolved in the $450$ Hz runs due to the enforcement of vertical hydrostatic equilibrium (we already saw
 that hydrodynamic effects can accelerate the burning front in our comparison with the 1D flames). The

### 2020-03-03 xrb/submit/paper.tex
Migrate Don's changes over from parent dir
Sorry for the confusion

@@
 with either the temperature specified, $T_i = T(p_i, \rho_i,
-\{X_k\}_i)$ (in the isothermal, ramp, and buffer layers) or constant
+\{X_k\}_i)$ (in the neutron star, ramp, and isothermal buffer layers) or constant
 entropy, $s_i = s(p_i, \rho_i, \{X_k\}_i)$, in the fuel layer.  This
@@
 case.  We see in the next section that these two networks give largely
-the same flame speed.  The reduced network size saves a lot of memory,
-which will be useful when we transition to 3D simulations.
+the same flame speed.  We only ran the simulation with the {\tt iso7}
+reaction network to 15~ms, and we observe this was sufficient to establish
+a steady flame speed for comparison to our 10/10 simulation with the {\tt aprox13}
+network. The reduced network size saves a lot of memory, which will be useful when we
+transition to 3D simulations.
 
@@
 with the 1D laminar flames, and is the behavior predicted by Eq.~\ref{eq:flame_scaling}. The
-1000~Hz run goes \added{nearly} 2 times faster than the 10/10 2000~Hz run, as expected from the
+1000~Hz run goes nearly 2 times faster than the 10/10 2000~Hz run, as expected from the
 inverse relation between rotation rate and the ratio of burning front area to scale height
@@
 our models. Although the magnetic fields of neutron stars exhibiting
-X-ray bursts are relatively weak, with $B \lesssim 10^8 - 10^9$G
+X-ray bursts are relatively weak, with $B \lesssim 10^8 - 10^9~$G
 \citep{mukherjee2015magnetic} (at least compared to e.g. magnetars,
-which have $B\lesssim 10^{15}$G), it has been found by
+which have $B\lesssim 10^{15}~$G), it has been found by
 \citet{art-2016-cavecchi-etal} that even weak magnetic fields could

### 2020-03-03 astronum_2019_MAESTROeX/astronum-maestroex.tex
responded to referee's report

@@
 \usepackage[sort&compress,numbers]{natbib}
-
-%\usepackage{cite}
 
@@
 
-For many flows in astrophysical systems, the magnitude of the fluid velocity is much less than the soundspeed.
+For many flows in astrophysical systems, the magnitude of the fluid velocity is much less than the soundspeed \cite{Bell2004,Hujeirat2009,Kupka2017}.
 Consequently, the Mach number, the ratio of the fluid velocity to the sound speed, is much less than 1: $\textrm{Ma} = |\Ub| / c_s \ll 1$.
@@
 
-This restriction can be lifted by using sound-proof methods.
+This restriction can be lifted by using sound-proof methods (see \cite{Kupka2017},~Section~4 for a review).
 These methods can involve modifying the fluid equations, modifying the computational algorithm and/or modifying the flow parameters in order to allow much larger timesteps to be used.
-In \maestroex, we use the \emph{low Mach number approximation}, a limit of the compressible Euler equations which effectively filters out sound waves.
+In \maestroex\ \cite{Fan2019,Fan2019JOSS}, we use the \emph{low Mach number approximation}, a limit of the compressible Euler equations which effectively filters out sound waves.
 An overview of this shall be given in \sref{sec:low_mach_hydro}, with further details of the \maestroex~code and our development workflow given in \sref{sec:maestroex} and \sref{sec:workflow}.
 
-An example of an astrophysical low Mach number flow is convection within the interiors of massive stars.
+An example of an astrophysical low Mach number flow is convection within the interiors of massive stars \cite{Kupka2017}.
 Prior to core collapse, the interiors of massive stars consist of a sequence of convective burning shells, separated by inert, non-convective shells.
@@
 where $\rho$, $\Ub$ and $h$ are the mass density, fluid velocity and specific enthalpy, $X_k$ and $\dot{\omega}_k$ are the species mass fraction and production rate of species $k$, and $H_{\mathrm{nuc}}$ is the energy release per time per unit mass.
+The Lagrangian derivative is defined as $D/Dt = \partial/\partial t + \nabla\cdot$.
 \maestroex~defines the base state pressure $p_0$ to be consistent with the one-dimensional base state density, $\rho_0 = \rho_0(r, t)$, which represents the lateral average and is in hydrostatic equilibrium with $p_0$:
@@
 
-% \maestroex\ is our awesome new code for modelling slow fluids in space
-
 Prior to \maestroex, we developed the low Mach number code \maestro.
@@
 
-% this section sounds kind of preachy
-
 \maestroex~is part of the \amrexastro~suite of open source adaptive mesh refinement hydrodynamics codes for astrophysical flows.
@@
 A prime example of this is the GPU porting capability that has recently been implemented in \castro, and is currently in the process of being implemented in \maestroex.
-% \MarginPar{mention starkiller astrophysics}
 
@@
 
-% \MarginPar{I didn't think we did Zenodo yet}
 New versions of the codes are released on the first of each month.
@@
 
-%The simplest way to include rotation is to ignore the effects of the centrifugal force, and only include the Coriolis force.
-This was done in \maestro, and should be a reasonable approximation for slowly rotating systems.
-% \MarginPar{In Zingale 2011, equation (2) has the Coriolis and centrifugal terms.  Include that equation here?}
-
 Another way that rotation could be incorporated would be to introduce a new `rotational pressure term', $p_1$, in order to balance the centrifugal potential:
@@
 
-% \begin{itemize}
-% \item Ignore centrifugal force \textemdash~for slow rotation can neglect centrifugal, only include Coriolis force
-% \item Introduce new pressure to balance centrifugal potential $$\rho \nabla \Phi_{\text{eff}} = \nabla p_0 + \nabla p_1 = \rho \nabla \phi + \frac{1}{2}\rho \nabla\Omega^2 r^2$$
-% \item Rewrite base state in terms of equipotentials $$p_0(r) \rightarrow p_0(\Phi_{\text{eff}}) $$
-% \item two-dimensional base state $$p_0(r)  \rightarrow p_0(r, \theta)$$
-% \end{itemize}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-% \begin{itemize}
-%     \item Why GPUs? Big computers use them.
-%     \item GPUs in \amrex: it does everything for you.
-Also important to note performance portability: GPU-ported AMReX codes still run on other architectures, optimizations for GPU often improve performance on MPI+OpenMP as well.
-%     \item GPUs in \maestroex: ported most of the source terms, working on the hydro.
-one-dimensional base state causing headaches, especially for big problems where we suspect it may not fit in GPU memory.
-%     \item Performance plots? Do these at the hackathon?
-% \end{itemize}
-
-% \MarginPar{I feel like someone who knows what they're talking about should write/seriously edit this section}
 The latest generation of supercomputers to come online (e.g.\ OLCF Summit, soon NERSC Perlmutter and OLCF Frontier) rely heavily on GPUs in order to achieve high performance while keeping energy consumption to a minimum (GPUs consume considerably less power per flop than CPUs).
@@
     \centering
-    \includegraphics[width=\textwidth]{plots/gpu_speedup}
+    \includegraphics[width=\textwidth]{gpu_speedup}
     \caption{Comparison of the performance of functions on the CPU vs the GPU.
@@
 
-% The work at Stony Brook was supported by DOE/Office of Nuclear
-% Physics grant DE-FG02-87ER40317 and contract 7418390 with Lawrence
-% Berkeley National Laboratory as part of the Exascale Compute Project
-% ExaStar collaboration.  The work at LBNL was supported by the DOE
-% Office of Advanced Scientific Computing Research under Contract No,
-% DE-AC02-05CH11231. An award of computer
-% time was provided by the Innovative and Novel Computational Impact on
-% Theory and Experiment (INCITE) program. This research used resources
-% of the Oak Ridge Leadership Computing Facility at the Oak Ridge
-% National Laboratory, which is supported by the Office of Science of
-% the U.S. Department of Energy under Contract No.\ DE-AC05-00OR22725.
-% This research used resources of the National Energy Research
-% Scientific Computing Center, which is supported by the Office of
-% Science of the U.S. Department of Energy under Contract
-% No.\ DE-AC02-05CH11231.
-
 \bibliographystyle{iopart-num}

### 2020-03-03 xrb/submit/paper.tex

@@
 case.  We see in the next section that these two networks give largely
-the same flame speed.  We only ran the simulation with the {\tt iso7}
-reaction network to 15~ms, and we observe this was sufficient to establish
+the same flame speed. \added{We only ran the simulation with the {\tt iso7}
+reaction network out to 15~ms, and we observe this was sufficient to establish
 a steady flame speed for comparison to our 10/10 simulation with the {\tt aprox13}
-network. The reduced network size saves a lot of memory, which will be useful when we
+network.} The reduced network size saves a lot of memory, which will be useful when we
 transition to 3D simulations.
@@
 Table \ref{table:flame_speeds_multid} gives the flame speed measured in each multidimensional
-simulation. The 2D flames propagate at speeds about an order of magnitude faster than their 1D
-counterparts (Table \ref{table:flame_speeds_1d}). The increase in flame speed is likely a product
-of the larger flame surface area and hydrodynamical effects such as turbulence, wrinkling, and
-convective cycles, which bring cooler fuel from ahead of the front into the hottest part of the
-burning region.
+simulation. \added{Uncertainties in the velocities are the standard errors of the slopes returned by the
+least-squares fits, assuming uncertainties in front position on the order of the coarse grid size.}
+The 2D flames propagate at speeds about an order of magnitude faster than their 1D counterparts
+(Table \ref{table:flame_speeds_1d}). The increase in flame speed is likely a product of the larger
+flame surface area and hydrodynamical effects such as turbulence, wrinkling, and convective cycles,
+which bring cooler fuel from ahead of the front into the hottest part of the burning region.
 

### 2020-03-20 sdc_applications/paper.tex
add the Castro SDC applications paper moved from astro_papers

\documentclass[times,modern]{aastex62}

% these lines seem necessary for pdflatex to get the paper size right
\pdfpagewidth 8.5in
\pdfpageheight 11.0in

\usepackage[T1]{fontenc}
\usepackage{epsf,color,amsmath}

\usepackage{cancel}

\newcommand{\sfrac}[2]{\mathchoice%
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
    \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
    \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptscriptfont0 #1}\kern-.2em/
    \kern-.15em\lower.25ex\hbox{\the\scriptscriptfont0 #2}} {#1\!/#2}}

\newcommand{\castro}{{\sf Castro}}
\newcommand{\maestro}{{\sf Maestro}}
\newcommand{\flash}{{\sf Flash}}
\newcommand{\amrex}{{\sf AMReX}}

\newcommand{\isot}[2]{$^{#2}\mathrm{#1}$}
\newcommand{\isotm}[2]{{}^{#2}\mathrm{#1}}

\newcommand{\gcc}{\mathrm{g~cm^{-3} }}
\newcommand{\cms}{\mathrm{cm~s^{-1} }}

\newcommand{\nablab}{{\mathbf{\nabla}}}
\newcommand{\Ub}{\mathbf{U}}
\newcommand{\gb}{\mathbf{g}}
\newcommand{\omegadot}{\dot{\omega}}
\newcommand{\Sdot}{\dot{S}}
\newcommand{\ddx}[1]{{\frac{{\partial#1}}{\partial x}}}
\newcommand{\ddt}[1]{{\frac{{\partial#1}}{\partial t}}}
\newcommand{\odt}[1]{{\frac{{d#1}}{dt}}}
\newcommand{\divg}[1]{{\nablab \cdot \left (#1\right)}}
\newcommand{\inp}{{\mathrm{input}}}
\newcommand{\dedr}{\left . {\partial{}e}/{\partial\rho}\right |_{T, X_k}}
\newcommand{\dedrd}{\left . \frac{\partial{}e}{\partial\rho}\right |_{T, X_k}}
\newcommand{\dedX}{\left . {\partial{}e}/{\partial{}X_k} \right |_{\rho, T}}
\newcommand{\dedXd}{\left . \frac{\partial{}e}{\partial{}X_k} \right |_{\rho, T, X_{j,j\ne k}}}
\newcommand{\dedT}{\left . {\partial{}e}/{\partial{}T} \right |_{\rho,X_k}}
\newcommand{\dedTd}{\left . \frac{\partial{}e}{\partial{}T} \right |_{\rho,X_k}}

\newcommand{\Ic}{{\boldsymbol{\mathcal{I}}}}
\newcommand{\smax}{{s_\mathrm{max}}}
\newcommand{\kth}{k_\mathrm{th}}
\usepackage{bm}

\newcommand{\Uc}{{\,\bm{\mathcal{U}}}}
\newcommand{\Fb}{\mathbf{F}}
\newcommand{\Sc}{\mathbf{S}}

\newcommand{\xv}{{(x)}}
\newcommand{\yv}{{(y)}}
\newcommand{\zv}{{(z)}}

\newcommand{\ex}{{\bf e}_x}
\newcommand{\ey}{{\bf e}_y}
\newcommand{\ez}{{\bf e}_z}

\newcommand{\Ab}{{\bf A}}
\newcommand{\Sq}{{\bf S}_\qb}
\newcommand{\Sqhydro}{{\Sq^{\mathrm{hydro}}}}
\newcommand{\qb}{{\bf q}}

\newcommand{\Shydro}{{{\bf H}}}
\newcommand{\Hb}{{\bf H}}
\newcommand{\Rb}{{\bf R}}
\newcommand{\Rq}{{\bf R}}
\newcommand{\Adv}[1]{{\left [\boldsymbol{\mathcal{A}} \left(#1\right)\right]}}
\newcommand{\Advt}[1]{{\left [\boldsymbol{\mathcal{\tilde{A}}} \left(#1\right)\right]}}
\newcommand{\Advs}[1]{\boldsymbol{\mathcal{A}} \left(#1\right)}

\newcommand{\avg}[1]{{\left \langle #1 \right \rangle}}

\setlength{\marginparwidth}{0.75in}
\newcommand{\MarginPar}[1]{\marginpar{\vskip-\baselineskip\raggedright\tiny\sffamily\hrule\smallskip{\color{red}#1}\par\smallskip\hrule}}

\begin{document}
%======================================================================
% Title
%======================================================================
%\title{A Simple Spectral Deferred Correction Method to Couple Nuclear Burning and Hydrodynamics for Detonations}

\title{So you've decided to detonate helium}

\shorttitle{Detonations and NSE}
\shortauthors{Zingale et al.}

%% \author[0000-0001-8401-030X]{M.~Zingale}
%% \affiliation{Dept.\ of Physics and Astronomy, Stony Brook University, 
%%        Stony Brook, NY 11794-3800}

%% \author[0000-0003-0439-4556]{M.~P.~Katz}
%% \affiliation{Nvidia Corp}

%% \author[0000-0002-5749-334X]{J.~B.~Bell}
%% \affiliation{Center for Computational Sciences and Engineering, Lawrence Berkeley National Laboratory, Berkeley, CA  94720}

%% \author[0000-0003-1791-0265]{A.~J.~Nonaka}
%% \affiliation{Center for Computational Sciences and Engineering, Lawrence Berkeley National Laboratory, Berkeley, CA  94720}

%% \author[0000-0001-8092-1974]{W.~Zhang}
%% \affiliation{Center for Computational Sciences and Engineering, Lawrence Berkeley National Laboratory, Berkeley, CA  94720}

%% \correspondingauthor{Michael Zingale}
%% \email{michael.zingale@stonybrook.edu}

%======================================================================
% Abstract and Keywords
%======================================================================
\begin{abstract}
Helium detonations can be challenging to model because of the
difficulty in accurately coupling hydrodynamics and reactions.  This
is particularly acute in the ash of the detonation where nuclear
statistical equilibrium is established.  We look a several different
methods of coupling explicit hydrodynamics and stiff reaction sources,
applied to detonations which reach NSE conditions.  This is all done
in the framework of the \castro\ hydrodynamics code, and all algorithm
implementations are freely available.
\end{abstract}

\keywords{hydrodynamics---methods: numerical}

%======================================================================
% Introduction
%======================================================================
\section{Introduction}\label{Sec:Introduction}

Reactive flow can be challenging to model when the timescale for
changes in the nuclear abundances due to reactions is comparable to
the hydrodynamical timescales.  Traditional methods of coupling
hydrodynamics and reactions used in astrophysics use operator
splitting---each physical process acts on the output of the previous
process.  This makes it easy to add physics in a modular way to a
simulation code, but competition between physical processes can cause
the coupling to breakdown.  In general, the error in the splitting
controlled by the timestep.

A particular difficult phase of evolution to model is the nuclear
statistic equilibrium that sets in with helium burning for
temperatures in excess of $4\times 10^9~\mathrm{K}$.  Physically, the
forward and reverse rates should balance leading to an equilibrium.
With operator splitting, an NSE region will have a large positive flow
through the network in a zone in one step followed by a large negative
flow over the next timestep, as the code struggles to produce an
equilibrium.  This large change in abundances (and large alternately
positive and negative energy generation rates) can be a challenge for
a code.  The easiest way to improve the coupling is to cut the
timestep, but this makes simulations prohibitively expensive.
Sometimes the burning is simply halted on a zone-by-zone basis when
NSE conditions are reached (e.g., as in \citealt{hedet}).
Alternately, at high temperatures, a reaction network can be replaced
with a table of NSE abundances and the zone's composition set through
table look-ups\MarginPar{example? GCD?}.

In this paper, we look at some different approaches to coupling
hydrodynamics and reactions, using a helium detonation that attains
NSE as our testbed.  We use the \castro\ hydrodynamics code for all of
our numerical experiments, and all of the code to reproduce the results
in this paper are in the \castro\ github repository.

\section{Numerical Methodology}

We solve the Euler equations for compressible, reacting flow.  We'll focus on
one-dimensional flow, and the system appears as:
\begin{equation}
\frac{\partial \Uc_t}{\partial t}  + \frac{\partial \Fb^\xv (\Uc)}{\partial x} = \Sc(\Uc)
\end{equation}
where $\Uc = (\rho, (\rho X_k), (\rho u), (\rho E))^\intercal$ are
the conserved fluid quantities: mass density, $\rho$, velocity, $u$,
specific total energy, $E$, and nuclear
species mass fractions, $X_k$.  The specific total energy relates to
the specific internal energy as $E = e + u^2/2$.  The
corresponding fluxes are
\begin{equation}
\Fb^\xv (\Uc) = \left ( \begin{array}{c}
   \rho u \\ \rho X_k u \\ \rho u u + p \\  \rho u E + u p
   \end{array} \right )
\end{equation}
Here the pressure, $p$, enters, and is found via the equation of state,
\begin{equation}
p = p(\rho, X_k, e)
\end{equation}

Finally, $\Sc(\Uc)$ are the source terms, which we decompose into
hydrodynamical sources, $\Hb$ (like gravity), and reactive sources,
$\Rb$,
\begin{equation}
  \Sc(\Uc) = \Hb(\Uc) + \Rb(\Uc)
\end{equation}
In this paper, we will not consider any hydrodynamical sources, but we
will carry the term $\Hb$ around for completeness.  The reactive
sources take the form:
\begin{equation}
  \Rb(\Uc) = \left ( \begin{array}{c}
     0 \\ \rho \omegadot_k \\ 0 \\ \rho \dot{S}
  \end{array} \right )
\end{equation}
where $\omegadot_k$ is the creation rate for species $k$ and $\dot{S}$
is the energy generation rate per unit mass.

Sometimes it is preferable to work with the primitive variables,  $\qb
= (\rho, X_k, u, p, (\rho e))^\intercal$.  Here, the system appears
as:
\begin{equation}
\qb_t + \Ab^\xv(\qb) \qb_x  = \Sq
\end{equation}
with the matrix $\Ab^\xv$ giving the coefficients of the spatial derivatives
of the primitive variables:
\begin{equation}
\Ab^\xv(\qb) = \left ( \begin{array}{ccccc}
    u & 0 & \rho & 0 & 0 \\
    0 & u & 0    & 0 & 0 \\
    0 & 0 & u    & 1/\rho & 0 \\
    0 & 0 & \Gamma_1 p & u & 0 \\
    0 & 0 & \rho h & 0 & u
  \end{array} \right )
\end{equation}
where $\Gamma_1$ is an adiabatic index, $\Gamma_1 = d\log p/d\log\rho
|_s$.  Note, the primitive state has two thermodynamic quantities, $p$
and $(\rho e)$, to more efficiently handle the general equation of
state in the Riemann solver, as described in \citet{castro}, but
alternate formulations are possible \citep{colellaglaz:1985}.
The source term vector, $\Sq$, can again be decomposed into hydrodynamic
sources (now in terms of the primitive variables) and reaction terms,
\begin{equation}
  \Sq = \Sqhydro + \Rb(q)
\end{equation}
with
\begin{equation}
\Rb(q) = \left ( \begin{array}{c}
     0 \\
     \omegadot_k \\
     0 \\
     \Gamma_1 p \sigma \Sdot \\
     \rho \Sdot
   \end{array} \right )
\end{equation}
where
\begin{equation}
\sigma \equiv \frac{\partial p/\partial T |_\rho}{\rho c_p \partial p/\partial \rho |_T}
\end{equation}
and $c_p$ is the specific heat at constant pressure, $c_p = \partial
h/\partial T |_p$.  A derivation of this source for the pressure
equation can be found in \cite{ABNZ:III}.  We note that this source is
algebraically identical to that shown in Eq.~25 of \cite{castro}.

We use the \castro\ hydrodynamics code~\cite{castro}, together with
the corner transport upwind piecewise parabolic method
hydrodynamics~\citep{millercolella:2002}.  This is a finite-volume
method that uses characteristic tracing to predict a time-centered
flux through the interfaces of the grid zones.  For pure
hydrodynamics, these time-centered fluxes result in second-order
accurate time integration.  With reactions, we want to couple
hydrodynamics and the reaction sources to second order.  Nuclear
reaction sources are stiff, and need to be integrated using implicit
methods for stabilty.  Operator splitting is traditionally employed
here, and our goal in this paper is to look at several different
approaches for splitting.

We describe several time-integration techniques for coupling
hydrodynamics and reactions in the next sections.

\subsection{Strang Splitting}

\subsection{Timestep Limiters and Retry Mechanism}

In addition to the standard CFL timestep limiter for explicit
hydrodynamics, \castro\ can also enforcing timestep limiters based on
the energy generation or abundance changes over a timestep.

\castro\ has the ability to reject a timestep if it detects a failure
and retry with smaller timesteps (subcycling to make up the original
required timestep).  Among the conditions that can trigger this are
density failing below zero during advection, the VODE integration
failing to converge in the implicit solve, or violation of one of the
timestep limiters during the step.  At the moment, this retry
mechanism only works with the CTU hydrodynamics + Strang splitting.
We enable it here, since, as will be shown shortly, this combination
of methods has the most difficulty in integrating the detonation.

\subsection{Balanced Splitting}

A variation on Strang splitting called (re-)balanced splitting was
developed in \citet{speth:2013}.

\subsection{SDC Method}

A second-order accurate time integration method using a simple
deferred correction strategy was shown in \citep{nonaka:2016}.  A
version of this was also implemented in the \maestro\ low-Mach number
hydrodynamics code some time ago.  A compressible version, for use in
\castro\ is slightly different due to the need to do some operations
on the conserved variable state and some on the primitive variable
state.  We describe the application of this to the equations of
compressible hydrodynamics below:

\begin{itemize}

\item {\em Initialization}

  \begin{itemize}
  \item Set the first iterate of the primitive-variable reactive source terms from the
    last step:
    \begin{equation}
      \Ic^{n+1/2,(0)}_{\qb} = \Ic^{n-1/2,(\smax)}_{\qb}
    \end{equation}
  %% \item Solve the Poisson problem for the initial gravitational potential:
  %%   \begin{equation}
  %%     \nabla^2 \Phi^n = 4\pi G \rho^n
  %%   \end{equation}

  %% \item Set the first guess at the new time potential as
  %%   $\Phi^{n+1,(0)} = \Phi^n$.

  \end{itemize}

\item {\em Iterate}

  Iterate from $k = 1, \ldots, \smax$.  For second-order accuracy,
  $\smax = 2$ is sufficient.

  \begin{itemize}
  \item {\em Create the Advective Update Term}

    \begin{itemize}
    \item convert $\Uc \rightarrow \qb$.  This is an algebraic transformation,
      but will require the EOS.

    \item predict $\qb$ to the interfaces at $n+1/2$ using the CTU PPM
      method.  The source terms, $\Sq$, used in the prediction are:
      \begin{equation}
        \Sq = \Sqhydro^{n+1/2} + \Ic_\qb^{n+1/2,(k-1)}
      \end{equation}
      Here we use the iteratively lagged integrals of the primitive variable
      terms accounting only for reactions, $\Ic_\qb^{n+1/2,(k-1)}$, as the
      reactive source.  Any hydrodynamic source terms are time-centered
      using the previous iteration:
      \begin{equation}
        \Sqhydro^{n+1/2} = \frac{1}{2} \left ( \Sqhydro^n + \Sqhydro^{n+1,(k-1)} \right )
      \end{equation}

    \item solve the Riemann problem at each interface to get a unique
      conserved state on each interface, $\Uc^{n+1/2}_{i+1/2}$

    \item construct the advective update terms, first ignoring the hydrodynamics sources:
      $\Advt{\Uc}^{n+1/2}_{i,j,k}$, e.g., as:
      \begin{align}
        \Advt{\Uc}^{n+1/2,(k)}_{i,j,k} =
          &- \frac{\Fb^\xv(\Uc^{n+1/2}_{i+1/2}) - \Fb^\xv(\Uc^{n+1/2}_{i-1/2})}{\Delta x} 
      \end{align}
    Now the conservative hydrodynamics source terms are computed by first updating to the
    new state with advection only\footnote{For a source like gravity, this update can be done first for $\rho$ and then define the new momentum source using $\rho^{\star\star}$ and likewise for energy}, as:
    \begin{equation}
      \Uc^{\star\star} = \Uc^n + \Delta t \Advt{\Uc}^{n+1/2,(k)}
    \end{equation}
    %% first updating the density to the new state with advection only as:
    %% \begin{equation}
    %%   \rho^{\star\star} = \rho^n + \Delta t \Advt{\rho}^{n+1/2,(k)}
    %% \end{equation}
    %% then constructing the momentum source term:
    %% \begin{equation}
    %%   {\bf S}_{\rho\Ub}^{n+1/2} = \frac{1}{2} (\rho^n \gb^n + \rho^{\star\star} \gb^{n+1,(k-1)})
    %% \end{equation}
    %% Then updating the momentum with only advection as:
    %% \begin{equation}
    %%   (\rho \Ub)^{\star\star} = (\rho \Ub)^n + \Delta t \Advt{\rho\Ub}^{n+1/2,(k)} + \Delta t {\bf S}_{\rho\Ub}^{n+1/2}
    %% \end{equation}
    %% and finally constructing the energy source:
    %% \begin{equation}
    %%   S_{\rho E}^{n+1/2} = \frac{1}{2} \left [ (\rho\Ub)^n \cdot \gb^n + (\rho\Ub)^{\star\star} \cdot \gb^{n+1,(k-1)}
    %%     \right ]
    %% \end{equation}

    The final advective update term is then:
    \begin{equation}
      \Adv{\Uc}^{n+1/2,(k)}_{i,j,k} = \Advt{\Uc}^{n+1/2,(k)}_{i,j,k} + \Shydro^{n+1/2}
    \end{equation}
    with
    \begin{equation}
      \Shydro^{n+1/2} = \left ( \begin{array}{c}
                     0 \\ 0 \\
                    {\bf S}_{\rho\Ub}^{n+1/2} \cdot \ex \\
                    S_{\rho E}^{n+1/2} \end{array} \right )
    \end{equation}

    \MarginPar{if $\Shydro$ depended on the outcome of the burn, we might need to do things differently here}

    \end{itemize}

  \item {\em Update the System Using a Method of Lines Integration}

    \begin{itemize}
    \item Define the hydro source terms---these will be time-centered,
      but piecewise-constant in time throughout the integration.

    \item With the spatial discretization done, our system appears as a
      system of ODEs: \MarginPar{what is the right form of acceleration
        here? should we do piecewise linear in time?}
      \begin{equation}
        \left . \odt{\Uc} \right |_{i,j,k} = \Adv{\Uc}^{n+1/2,(k)}_{i,j,k} + \Rb(\Uc)
      \end{equation}

      As we are integrating this system we need to get the
      temperature, $T$, for the rate evaluations.  We construct this
      by subtracting the kinetic energy from the total energy to get
      the specific internal energy, $e$, and then calling the equation
      of state.

      Note that our advection terms are piecewise constant \MarginPar{need to think about this more}
      approximations, so when we integrate our system, their
      contributions will be linear in time.  Since $(\rho X_k)$ sees a
      linear advective term, $\rho$ does as well.  $(\rho \Ub)$ and
      $(\rho E)$ will also have linear contributions in time from
      advection.  We can see that the internal energy will also have a
      linear contribution by recognizing that the kinetic energy term,
      $K = (\rho |\Ub|^2)/\rho$ is linear in time.
      Since the reactions don't affect $\rho$ and $(\rho \Ub)$, we can
      algebraically update these using this piecewise linear behavior:
      \begin{align}
        \rho(t) &= \rho^n + \Adv{\rho}^{n+1/2,(k)} \, t \\
        (\rho \Ub)(t) &= (\rho \Ub)^n + \Adv{(\rho \Ub)}^{n+1/2,(k)} \, t
      \end{align}
      At the end of the integration, our new state is $\Uc^{n+1,(k)}$.

    \end{itemize}

  \item {\em Compute the Reactive Source Terms.}

    We now seek the $\Ic$'s that capture the effect of just the
    reaction sources on the state variables for the next iteration.
    For the conserved quantities, these would simply be: \MarginPar{fix signs on A}
    \begin{equation}
      \Ic^{(k)}_{\Uc} = \frac{\Uc^{n+1,(k)} - \Uc^n}{\Delta t} - \Adv{\Uc}^{n+1/2,(k)}
    \end{equation}
    However, for our primitive variables, which are used in the
    prediction, we need to construct the required source terms more
    carefully.  We want:
    \begin{equation}
      \label{eq:Iq}
      \Ic^{(k)}_{\qb} = \frac{\qb^{n+1,(k)} - \qb^n}{\Delta t} - \Adv{\qb}^{n+1/2,(k)}
    \end{equation}
    but we need the advective update for $\qb$, which we have not
    constructed.  Additionally, we cannot simply use the equation of
    state on $\Ic^{(k)}_{\Uc}$ since this is a time-derivative and
    does not represent a well-defined state in itself.  Instead, we
    derive $\Ic^{(k)}_{\qb}$ via a multi-step process.  We first find
    the conservative state as if it were updated only with advection:
    \begin{equation}
      \Uc^\star = \Uc^n + \Delta t \Adv{\Uc}^{n+1/2,(k)}
    \end{equation}
    and then construct the corresponding primitive variable state via an algebraic transform,
    $\Uc^\star \rightarrow \qb^\star$.
    This allows us to define the advective update for $\qb$ as:
    \begin{equation}
      \Adv{\qb}^{n+1/2,(k)} = \frac{\qb^\star - \qb^n}{\Delta t}
    \end{equation}
    Defining the primitive state corresponding to the fully-updated
    conserved state via an algebraic transform, $\Uc^{n+1,(k)}
    \rightarrow \qb^{n+1,(k)}$, we can construct $\Ic^{(k)}_{\qb}$ as given
    in Eq.~\ref{eq:Iq}.
    Putting all of this together, we see:
    \begin{equation}
      \Ic^{(k)}_{\qb} = \frac{\qb^{n+1,(k)} - \qb^\star}{\Delta t}
    \end{equation}

  %% \item {\em Solve for the New Gravitational Potential.}

  %%   We solve
  %%   \begin{equation}
  %%     \nabla^2 \Phi^{n+1,(k)} = 4\pi G \rho^{n+1,(k)}
  %%   \end{equation}

  \end{itemize}

\end{itemize}

\subsection{Old Primitive Stuff}

and our primitive system as:
\begin{equation}
\qb_t = \Advs{\qb} + \Rb(\qb)
\end{equation}
with
\begin{equation}
\Advs{\qb} = -\Ab^\xv(\qb) \qb_x - \Ab^\yv(\qb) \qb_y - \Ab^\zv(\qb) \qb_z + \Sqhydro
\end{equation}

\section{Detonations}

Our focus here is modeling helium detonations.  Previously, we
showed~\cite{katzzingale:2019} that accurate modeling the ignition of
a detonations requires a resolution such that $\tau_s < \tau_e$, where
$\tau_s = \Delta x/c_s$ is the sound crossing time of a zone and
$\tau_e = e/\dot{e}$ is the nuclear energy injection timescale.  For
the detonations explored here, we will work in that regime.

Helium detonations can be important in different models of Type Ia
supernova.  In the double detonation model for Type Ia supernovae, a
helium detonation is used to trigger a carbon detonation in the
underlying white dwarf.  White dwarf collisions or mergers can have
helium detonations if there is a thin layer of helium on the surface
of the white dwarfs.  We will model two densities, $2\times
10^6~\gcc$, e.g., as used in \citet{glasner:2018} for double
detonation models, and $10^7~\gcc$ to see the effects at higher
densities.

Effect of number of SDC iterations

Effect of timestep

Effect of resolution

Difference densities?

\begin{figure}[t]
\centering
\plotone{strang_timestep}
\caption{\label{strang_timestep} Effect of timestep size on the detonation using Strang splitting.}
\end{figure}

\begin{figure}[t]
\centering
\plotone{sdc_iters}
\caption{\label{sdc_iters} Effect of number of SDC iterations detonation.}
\end{figure}

\begin{figure}[t]
\centering
\plotone{sdc_timestep}
\caption{\label{sdc_timestep} Effect of timestep size on the detonation using the SDC method.}
\end{figure}

\begin{figure}[t]
\centering
\plotone{strang_sdc_compare}
\caption{\label{strang_vs_sdc} Comparison of Strang and SDC on the detonation problem.}
\end{figure}

Something about expense?  We can just compare CPU time.

\section{Summary}

We presented a simple spectral deferred corrections scheme for coupling 
hydrodynamics and reactions.

\acknowledgements \castro\ is freely available at
\url{http://github.com/AMReX-Astro/Castro}.  All of the code and problem
setups used here are available in the git repo.  The work at Stony
Brook was supported by DOE/Office of Nuclear Physics grant
DE-FG02-87ER40317.

\software{\amrex\ (\url{https://github.com/AMReX-Codes/amrex}),
          \castro\ \citep{castro},
          GNU Compiler Collection (\url{https://gcc.gnu.org/}),
          Linux (\url{https://www.kernel.org}),
          matplotlib (\citealt{Hunter:2007},  \url{http://matplotlib.org/})
          NumPy \citep{numpy,numpy2},
          python (\url{https://www.python.org/})
         }

\appendix

\section{Jacobian}

To solve the reaction system implicitly, the ODE solver needs the Jacobian,
$\partial \Rb/\partial \Uc$.  We follow the method of \cite{castro_sdc}
and factor this into two pieces,
\begin{equation}
{\bf J} = \partial \Rb/\partial {\bf w } \partial {\bf w}/\partial \Uc
\end{equation}.

Writing this out for two species, $X_\alpha$ and $X_\beta$, we have
\begin{equation}
\Uc = \left ( \begin{array}{c} \rho X_\alpha \\ \rho X_\beta \\ \rho E \\ \rho e \end{array} \right )
\end{equation}
We take the intermediate state to be ${\bf w} = (\rho, X_\alpha, X_\beta,
T)^\intercal$.

The Jacobian transformation $\partial \Uc/\partial {\bf w}$ is:
\begin{equation}
\frac{\partial \Uc}{\partial {\bf w}} = \left (
   \begin{array}{cccc}
       X_\alpha & \rho & 0 & 0\\
       X_\beta & 0 & \rho & 0 \\
       \rho e_\rho  + e + K &
                 \rho  e_{X_\alpha} & \rho e_{X_\beta} &
                 \rho e_T \\
       \rho e_\rho  + e  &
                 \rho  e_{X_\alpha} & \rho e_{X_\beta} &
                 \rho e_T
     \end{array}\right)
\end{equation}
where we use the following notation for compactness:
\begin{equation}
e_\rho = \dedrd \qquad 
e_T = \dedTd \qquad
e_{X_k} = \dedXd 
\end{equation}
and write the kinetic energy as $K = |\Ub|^2/2$.  We get the inverse
(computed via SymPy, see the included Jupyter notebook) is:

\begin{equation}
\renewcommand{\arraystretch}{1.5}
\frac{\partial {\bf w}}{\partial \Uc} = \left (
  \begin{array}{cccc}
   0  & 0  &  \frac{1}{K} & -\frac{1}{K} \\
   \frac{1}{\rho} & 0 & - \frac{X_\alpha}{\rho K} & \frac{X_\alpha}{\rho K} \\
   0 & \frac{1}{\rho} &  - \frac{X_\beta}{\rho K} & \frac{X_\beta}{\rho K} \\
   -\frac{e_{X_\alpha}}{\rho e_T} & -\frac{e_{X_\beta}}{\rho e_T} & 
   \frac{\sum_k X_{k} e_{X_k} - \rho e_\rho - e }{\rho e_T K}  &
   -\frac{\sum_k X_{k} e_{X_k} - \rho e_\rho - e - K}{ \rho e_T K}  \\
   \end{array}\right)
\renewcommand{\arraystretch}{1}
\end{equation}

The reaction vector is
\begin{equation}
\Rb = \left (  \begin{array}{c} \rho \omegadot_\alpha \\ \rho \omegadot_\beta \\ \rho \Sdot \\ \rho \Sdot \end{array} \right )
\end{equation}
and the Jacobian is computed as $\partial \Rb/\partial {\bf w}$:
\begin{equation}
\frac{\partial \Rb}{\partial {\bf w}} = \left (
  \begin{array}{cccc}
     \omegadot_\alpha + \rho \frac{\partial \omegadot_\alpha}{\partial \rho} &
     \rho \frac{\partial \omegadot_\alpha}{\partial X_\alpha} &
     \rho \frac{\partial \omegadot_\alpha}{\partial X_\beta} & \rho \frac{\partial \omegadot_\alpha}{\partial T} \\
      %
     \omegadot_\beta + \rho \frac{\partial \omegadot_\beta}{\partial \rho} &
     \rho \frac{\partial \omegadot_\beta}{\partial X_\alpha} &
     \rho \frac{\partial \omegadot_\beta}{\partial X_\beta} & \rho \frac{\partial \omegadot_\beta}{\partial T} \\
     %
     \Sdot + \rho \frac{\partial \Sdot}{\partial \rho} & \rho \frac{\partial \Sdot}{\partial X_\alpha} & \rho \frac{\partial \Sdot}{\partial X_\beta} & \rho \frac{\partial \Sdot}{\partial T} \\
     %
     \Sdot + \rho \frac{\partial \Sdot}{\partial \rho} & \rho \frac{\partial \Sdot}{\partial X_\alpha} & \rho \frac{\partial \Sdot}{\partial X_\beta} & \rho \frac{\partial \Sdot}{\partial T} \\
  \end{array}
  \right )
\end{equation}

%======================================================================
% References
%======================================================================

\bibliographystyle{aasjournal}
\bibliography{ws}

\end{document}

### 2020-03-21 sdc_applications/paper.tex
some minor edits

@@
 \shorttitle{Detonations and NSE}
-\shortauthors{Zingale et al.}
+\shortauthors{}
 
@@
 required timestep).  Among the conditions that can trigger this are
-density failing below zero during advection, the VODE integration
+density failing below zero during advection, the ODE integration
 failing to converge in the implicit solve, or violation of one of the
@@
 
-\subsection{SDC Method}
+\subsection{Spectral Deferred Corrections}
+
+The basic idea of spectral deferred corrections is break the timestep
+into a number of temporal nodes and use low order difference methods
+in a correction equation that, upon iteration, achieves a high-order
+solution coupling the different physics together.
+
+\subsection{Simplified-SDC Method}
 
 A second-order accurate time integration method using a simple
-deferred correction strategy was shown in \citep{nonaka:2016}.  A
+deferred correction strategy was shown in \citep{SDC-old}.  A
 version of this was also implemented in the \maestro\ low-Mach number
@@
 
-\section{Detonations}
+\section{Simulations}
+
+\subsection{Flames}
+
+ECSNe flames
+
+\subsection{Detonations}
 
@@
 
+\subsection{Hydro and Gravity}
+
+Merging stars
+
 \section{Summary}
@@
 
-\software{\amrex\ (\url{https://github.com/AMReX-Codes/amrex}),
+\software{\amrex\ \citep{amrex_joss},
           \castro\ \citep{castro},
@@
 \begin{equation}
-{\bf J} = \partial \Rb/\partial {\bf w } \partial {\bf w}/\partial \Uc
+{\bf J} = \frac{\partial \Rb}{\partial {\bf w }} \frac{\partial {\bf w}}{\partial \Uc}
 \end{equation}.

### 2020-04-30 sdc_applications/paper.tex
updates

@@
 
+Even though it has no reactive sources, we include $\rho$ in our conservative state
+for the purposes of computing the Jacobian (we denote the conserved state used for the
+Jacobian $\Uc^\prime$.  
 Writing this out for two species, $X_\alpha$ and $X_\beta$, we have
 \begin{equation}
-\Uc = \left ( \begin{array}{c} \rho X_\alpha \\ \rho X_\beta \\ \rho E \\ \rho e \end{array} \right )
+\Uc^\prime = \left ( \begin{array}{c} \rho \\ \rho X_\alpha \\ \rho X_\beta \\ \rho E \\ \rho e \end{array} \right )
 \end{equation}
 We take the intermediate state to be ${\bf w} = (\rho, X_\alpha, X_\beta,
-T)^\intercal$.
+K, T)^\intercal$, where $K$ is the kinetic energy:
+\begin{equation}
+K = \frac{1}{2} |\Ub|^2
+\end{equation}
 
-The Jacobian transformation $\partial \Uc/\partial {\bf w}$ is:
+The Jacobian transformation $\partial \Uc^\prime/\partial {\bf w}$ is:
 \begin{equation}
 \frac{\partial \Uc}{\partial {\bf w}} = \left (
-   \begin{array}{cccc}
-       X_\alpha & \rho & 0 & 0\\
-       X_\beta & 0 & \rho & 0 \\
+   \begin{array}{ccccc}
+       1 & 0 & 0 & 0 & 0 \\
+       X_\alpha & \rho & 0 & 0 & 0 \\
+       X_\beta & 0 & \rho & 0 & 0  \\
        \rho e_\rho  + e + K &
-                 \rho  e_{X_\alpha} & \rho e_{X_\beta} &
+                 \rho  e_{X_\alpha} & \rho e_{X_\beta} & \rho &
                  \rho e_T \\
        \rho e_\rho  + e  &
-                 \rho  e_{X_\alpha} & \rho e_{X_\beta} &
+                 \rho  e_{X_\alpha} & \rho e_{X_\beta} & 0 &
                  \rho e_T
@@
 \renewcommand{\arraystretch}{1.5}
-\frac{\partial {\bf w}}{\partial \Uc} = \left (
-  \begin{array}{cccc}
-   0  & 0  &  \frac{1}{K} & -\frac{1}{K} \\
-   \frac{1}{\rho} & 0 & - \frac{X_\alpha}{\rho K} & \frac{X_\alpha}{\rho K} \\
-   0 & \frac{1}{\rho} &  - \frac{X_\beta}{\rho K} & \frac{X_\beta}{\rho K} \\
-   -\frac{e_{X_\alpha}}{\rho e_T} & -\frac{e_{X_\beta}}{\rho e_T} & 
-   \frac{\sum_k X_{k} e_{X_k} - \rho e_\rho - e }{\rho e_T K}  &
-   -\frac{\sum_k X_{k} e_{X_k} - \rho e_\rho - e - K}{ \rho e_T K}  \\
+\frac{\partial {\bf w}}{\partial \Uc^\prime} = \left (
+  \begin{array}{ccccc}
+   1  & 0 & 0 & 0 & 0 \\
+   - \frac{X_\alpha}{\rho} & \frac{1}{\rho} & 0 & 0 & 0 \\
+   - \frac{X_\beta}{\rho} & 0 & \frac{1}{\rho} & 0 & 0 \\
+   - \frac{K}{\rho} & 0 & 0 & \frac{1}{\rho} & -\frac{1}{\rho} \\
+   \frac{\sum_k X_{k} e_{X_k} - \rho e_\rho - e}{ \rho e_T} &
+    -\frac{e_{X_\alpha}}{\rho e_T} & -\frac{e_{X_\beta}}{\rho e_T} & 0 & \frac{1}{\rho e_T} \\
    \end{array}\right)
@@
 \begin{equation}
-\Rb = \left (  \begin{array}{c} \rho \omegadot_\alpha \\ \rho \omegadot_\beta \\ \rho \Sdot \\ \rho \Sdot \end{array} \right )
+\Rb(\Uc^\prime) = \left (  \begin{array}{c} 0 \\ \rho \omegadot_\alpha \\ \rho \omegadot_\beta \\ \rho \Sdot \\ \rho \Sdot \end{array} \right )
 \end{equation}
@@
 \frac{\partial \Rb}{\partial {\bf w}} = \left (
-  \begin{array}{cccc}
+  \begin{array}{ccccc}
+     0 & 0 & 0 & 0 & 0 \\
      \omegadot_\alpha + \rho \frac{\partial \omegadot_\alpha}{\partial \rho} &
      \rho \frac{\partial \omegadot_\alpha}{\partial X_\alpha} &
-     \rho \frac{\partial \omegadot_\alpha}{\partial X_\beta} & \rho \frac{\partial \omegadot_\alpha}{\partial T} \\
+     \rho \frac{\partial \omegadot_\alpha}{\partial X_\beta} & 0 & \rho \frac{\partial \omegadot_\alpha}{\partial T} \\
       %
@@
      \rho \frac{\partial \omegadot_\beta}{\partial X_\alpha} &
-     \rho \frac{\partial \omegadot_\beta}{\partial X_\beta} & \rho \frac{\partial \omegadot_\beta}{\partial T} \\
+     \rho \frac{\partial \omegadot_\beta}{\partial X_\beta} & 0 &  \rho \frac{\partial \omegadot_\beta}{\partial T} \\
      %
-     \Sdot + \rho \frac{\partial \Sdot}{\partial \rho} & \rho \frac{\partial \Sdot}{\partial X_\alpha} & \rho \frac{\partial \Sdot}{\partial X_\beta} & \rho \frac{\partial \Sdot}{\partial T} \\
+     \Sdot + \rho \frac{\partial \Sdot}{\partial \rho} & \rho \frac{\partial \Sdot}{\partial X_\alpha} & \rho \frac{\partial \Sdot}{\partial X_\beta} & 0 & \rho \frac{\partial \Sdot}{\partial T} \\
      %
-     \Sdot + \rho \frac{\partial \Sdot}{\partial \rho} & \rho \frac{\partial \Sdot}{\partial X_\alpha} & \rho \frac{\partial \Sdot}{\partial X_\beta} & \rho \frac{\partial \Sdot}{\partial T} \\
+     \Sdot + \rho \frac{\partial \Sdot}{\partial \rho} & \rho \frac{\partial \Sdot}{\partial X_\alpha} & \rho \frac{\partial \Sdot}{\partial X_\beta} & 0 & \rho \frac{\partial \Sdot}{\partial T} \\
   \end{array}

### 2020-05-07 sdc_applications/paper.tex
some notes

@@
 
+\subsection{Reaction Convergence}
+
 \subsection{Flames}
@@
 ECSNe flames
+
+aprox13 flames
+
+simplified-SDC takes 3x more timesteps than Strang
+
+look at number of right hand side calls
 
@@
 
+\subsection{Massive Star}
+
+we need to look at the number of RHS calls spatially
+
+also need to look at tolerances in the burner
+
+analytic Jacobian run is much faster
+
+what does Strang run do
+
+\subsection{X-ray bursts}
+
+crashed with too many subcycles.  Need to rerun with new Jacobian
+
 \subsection{Hydro and Gravity}
@@
 Merging stars
+
+it doesn't run more than a few steps with simplified SDC
+
+has a small density error and too many subcycles after ~ 128 steps
 

### 2020-06-26 xrb2/paper.tex
some updates
sketched intro for xrb2 paper

@@
 %======================================================================
-\title{Dynamics of Laterally Propagating Flames in X-ray Bursts. II. Realistic Burning}
+\title{Dynamics of Laterally Propagating Flames in X-ray Bursts. II. Realistic Burning \& Rotation}
 
@@
 We continue to investigate laterally propagating flames in XRBs...
+
 \end{abstract}
@@
 \section{Introduction}\label{Sec:Introduction}
+
+Unlike our previous paper, all runs are now conducted without artificially boosting the speed of the flame in order to reduce the computational cost.
+This is possible because of the work done to offload the code to GPUs, where it runs significantly faster.
+We can therefore run the tests without any boosting.
+
+We investigate the effect of rotation rate on the flame.
+With the exception of (citation to the weird 11 Hz flame), most observations of XRBs come from sources with rotation rates of 200-600Hz \MarginPar{check whether this is true for XRBs or just burst oscillations}.
+There are a number of factors that could explain this lack of observations below 200 Hz.
+It could be that there is some physical process which inhibits the flame ignition and/or spread at lower rotation rates.
+It could be that burst oscillations at lower rotation rates are smaller in amplitude and therefore more difficult to detect.
+It could be that it's not got anything to do with the flame at all, but that neutron stars in accreting low mass X-ray binaries rarely have rotation rates below 200 Hz \MarginPar{look into this?}.
+
+Insert some stuff here about previous work on rotation
+
+In the following sections, we conduct a series of simulations at various rotation rates to investigate its effect on the flame.
+We find that at lower rotation rates, the flame itself becomes harder to ignite.
+We discuss the implications that this may have for burst physics and observations.
 
@@
 
+We add the geometric source terms from \citet{bernard-champmartin}
+
 %======================================================================
@@
 
+lateral profiles of X vs distance behind the flame
+
 vertical profiles of the atmosphere behind the flame---how does it compare to isentropic?
@@
 
-different latitudes
+different initial models?
 
-different initial models?
+different rotation rates
+
+look at flame speed, shape of flame, amount of mixing (e.g. it looks like slower rotation rates have less mixed flames as the hurricane is less strong.)
 

### 2020-08-03 xrb2/paper.tex
add some figurest p
added rotation references

@@
 We investigate the effect of rotation rate on the flame.
-With the exception of (citation to the weird 11 Hz flame), most observations of XRBs come from sources with rotation rates of 200-600Hz \MarginPar{check whether this is true for XRBs or just burst oscillations}.
+With the exception of (citation to the weird 11 Hz flame), most observations of XRBs come from sources with rotation rates of 200-600Hz \citep{bilous:2019,galloway:2020} \MarginPar{check whether this is true for XRBs or just burst oscillations}.
 There are a number of factors that could explain this lack of observations below 200 Hz.

### 2020-08-03 xrb2/paper.tex
added phase plots and outlined discussion

@@
 
+\begin{figure}[t]
+\centering
+\plotone{time_series_500}
+\caption{Time series of the 500 Hz run showing $\bar{A}$.}
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\plotone{time_series_1000}
+\caption{Time series of the 1000 Hz run showing $\bar{A}$.}
+\end{figure}
+
 lateral profiles of X vs distance behind the flame
@@
 look at flame speed, shape of flame, amount of mixing (e.g. it looks like slower rotation rates have less mixed flames as the hurricane is less strong.)
+
+Looking at the phase plots Figures~\ref{fig:urho}-\ref{fig:abar}, we see:
+\begin{itemize}
+    \item From the $u$-$\rho$ and $u$-$v$ plots, we see that the burning occurs in a region of high density where the fluid is moving in a circular motion.
+From the previous plots (see: streamlines?), this most likely corresponds to the leading edge of the flame.
+This feature is not (yet?) developed in the 250~Hz flame.
+(will this ever develop, or will it just fizzle out completely?)
+    \item From the $\bar{A}$-$\rho$ plots in Figure \ref{fig:abar}, we can see that as the rotation rate increases, so does the peak temperature.
+We believe that the tracks in the plot correspond to reaction pathways in the reaction network.
+In the 1000~Hz plot, these tracks are somewhat disrupted, possibly due to the more vigorous mixing of the vortex at the flame front.
+(if this continues or if the rotation was cranked even higher, would the reactions get so disrupted that the burning is snuffed out, similar to what happens in combustion when the turbulent vortices become so small they penetrate the reaction zone and completely disrupt it).
+\end{itemize}
+
+\begin{figure}[t]
+    \centering
+    \plotone{urho_plot_0_1}
+    \caption{\label{fig:urho}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and 1000~Hz runs.}
+\end{figure}
+
+\begin{figure}[t]
+    \centering
+    \plotone{uv_plot_0_1}
+    \caption{\label{fig:uv}Phase plot of the horizontal $x$-velocity $u$, vertical $y$-velocity and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and 1000~Hz runs.}
+\end{figure}
+
+\begin{figure}[t]
+    \centering
+    \plotone{abar_plot_0_1}
+    \caption{\label{fig:abar}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and 1000~Hz runs.}
+\end{figure}
 

### 2020-08-03 xrb2/paper.tex
typos

@@
 \correspondingauthor{Alice Harpole}
-\email{alice.harpolee@stonybrook.edu}
+\email{alice.harpole@stonybrook.edu}
 
@@
 
-Looking at the phase plots Figures~\ref{fig:urho}-\ref{fig:abar}, we see:
+Looking at the phase plots in Figures~\ref{fig:urho}-\ref{fig:abar}, we see:
 \begin{itemize}
@@
 (will this ever develop, or will it just fizzle out completely?)
-    \item From the $\bar{A}$-$\rho$ plots in Figure \ref{fig:abar}, we can see that as the rotation rate increases, so does the peak temperature.
+    \item From the $\bar{A}$-$T$ plots in Figure \ref{fig:abar}, we can see that as the rotation rate increases, so does the peak temperature.
 We believe that the tracks in the plot correspond to reaction pathways in the reaction network.
 In the 1000~Hz plot, these tracks are somewhat disrupted, possibly due to the more vigorous mixing of the vortex at the flame front.
-(if this continues or if the rotation was cranked even higher, would the reactions get so disrupted that the burning is snuffed out, similar to what happens in combustion when the turbulent vortices become so small they penetrate the reaction zone and completely disrupt it).
+(If this continues or if the rotation was cranked even higher, would the reactions get so disrupted that the burning is snuffed out, similar to what happens in combustion when the turbulent vortices become so small they penetrate the reaction zone and completely disrupt it?)
 \end{itemize}
@@
     \plotone{uv_plot_0_1}
-    \caption{\label{fig:uv}Phase plot of the horizontal $x$-velocity $u$, vertical $y$-velocity and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and 1000~Hz runs.}
+    \caption{\label{fig:uv}Phase plot of the horizontal $x$-velocity $u$, vertical $y$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and 1000~Hz runs.}
 \end{figure}

### 2020-08-04 xrb2/paper.tex
Added speed plots to xrb2

@@
 \newcommand{\omegadot}{\dot{\omega}}
+\newcommand{\enucdot}{\dot{e}_{nuc}}
 
@@
 
+\begin{figure}[t]
+	\centering
+	\plottwo{speeds_1_100th_max.pdf}{speeds_1_1000th_max.pdf}
+	\caption{Plots of flame front position vs.\ time for the 500 Hz and 1000 Hz runs.
+The left-hand plot shows the results when tracking the $\frac{1}{100} \max[\enucdot]$ contour, while the right-hand plot shows the results for the $\frac{1}{1000} \max[\enucdot]$ contour.}
+\end{figure}
+
 Looking at the phase plots in Figures~\ref{fig:urho}-\ref{fig:abar}, we see:

### 2020-08-07 simplified_sdc_intro/paper.tex
move this into an intro paper the detonation part will live on in a separate paper

--- 2020-04-30 sdc_applications/paper.tex
@@
-\documentclass[times,modern]{aastex62}
+\documentclass[times,modern]{aastex63}
 
@@
 
-\subsection{Flames}
-
-ECSNe flames
-
-\subsection{Detonations}
-
-Our focus here is modeling helium detonations.  Previously, we
-showed~\cite{katzzingale:2019} that accurate modeling the ignition of
-a detonations requires a resolution such that $\tau_s < \tau_e$, where
-$\tau_s = \Delta x/c_s$ is the sound crossing time of a zone and
-$\tau_e = e/\dot{e}$ is the nuclear energy injection timescale.  For
-the detonations explored here, we will work in that regime.
-
-Helium detonations can be important in different models of Type Ia
-supernova.  In the double detonation model for Type Ia supernovae, a
-helium detonation is used to trigger a carbon detonation in the
-underlying white dwarf.  White dwarf collisions or mergers can have
-helium detonations if there is a thin layer of helium on the surface
-of the white dwarfs.  We will model two densities, $2\times
-10^6~\gcc$, e.g., as used in \citet{glasner:2018} for double
-detonation models, and $10^7~\gcc$ to see the effects at higher
-densities.
-
-Effect of number of SDC iterations
-
-Effect of timestep
-
-Effect of resolution
-
-Difference densities?
-
-\begin{figure}[t]
-\centering
-\plotone{strang_timestep}
-\caption{\label{strang_timestep} Effect of timestep size on the detonation using Strang splitting.}
-\end{figure}
-
-\begin{figure}[t]
-\centering
-\plotone{sdc_iters}
-\caption{\label{sdc_iters} Effect of number of SDC iterations detonation.}
-\end{figure}
-
-\begin{figure}[t]
-\centering
-\plotone{sdc_timestep}
-\caption{\label{sdc_timestep} Effect of timestep size on the detonation using the SDC method.}
-\end{figure}
-
-\begin{figure}[t]
-\centering
-\plotone{strang_sdc_compare}
-\caption{\label{strang_vs_sdc} Comparison of Strang and SDC on the detonation problem.}
-\end{figure}
-
-Something about expense?  We can just compare CPU time.
-
-\subsection{Hydro and Gravity}
-
-Merging stars
-
 \section{Summary}

### 2020-08-07 sdc_detonations/notebook.tex
split the detonations off into a separate paper

% Default to the notebook output style

% Inherit from the specified cell style.

\documentclass[11pt]{article}

    \usepackage[T1]{fontenc}
    % Nicer default font (+ math font) than Computer Modern for most use cases
    \usepackage{mathpazo}

    % Basic figure setup, for now with no caption control since it's done
    % automatically by Pandoc (which extracts ![](path) syntax from Markdown).
    \usepackage{graphicx}
    % We will generate all images so they have a width \maxwidth. This means
    % that they will get their normal width if they fit onto the page, but
    % are scaled down if they would overflow the margins.
    \makeatletter
    \def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
    \else\Gin@nat@width\fi}
    \makeatother
    \let\Oldincludegraphics\includegraphics
    % Set max figure width to be 80% of text width, for now hardcoded.
    \renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=.8\maxwidth]{#1}}
    % Ensure that by default, figures have no caption (until we provide a
    % proper Figure object with a Caption API and a way to capture that
    % in the conversion process - todo).
    \usepackage{caption}
    \DeclareCaptionLabelFormat{nolabel}{}
    \captionsetup{labelformat=nolabel}

    \usepackage{adjustbox} % Used to constrain images to a maximum size 
    \usepackage{xcolor} % Allow colors to be defined
    \usepackage{enumerate} % Needed for markdown enumerations to work
    \usepackage{geometry} % Used to adjust the document margins
    \usepackage{amsmath} % Equations
    \usepackage{amssymb} % Equations
    \usepackage{textcomp} % defines textquotesingle
    % Hack from http://tex.stackexchange.com/a/47451/13684:
    \AtBeginDocument{%
        \def\PYZsq{\textquotesingle}% Upright quotes in Pygmentized code
    }
    \usepackage{upquote} % Upright quotes for verbatim code
    \usepackage{eurosym} % defines \euro
    \usepackage[mathletters]{ucs} % Extended unicode (utf-8) support
    \usepackage[utf8x]{inputenc} % Allow utf-8 characters in the tex document
    \usepackage{fancyvrb} % verbatim replacement that allows latex
    \usepackage{grffile} % extends the file name processing of package graphics 
                         % to support a larger range 
    % The hyperref package gives us a pdf with properly built
    % internal navigation ('pdf bookmarks' for the table of contents,
    % internal cross-reference links, web links for URLs, etc.)
    \usepackage{hyperref}
    \usepackage{longtable} % longtable support required by pandoc >1.10
    \usepackage{booktabs}  % table support for pandoc > 1.12.2
    \usepackage[inline]{enumitem} % IRkernel/repr support (it uses the enumerate* environment)
    \usepackage[normalem]{ulem} % ulem is needed to support strikethroughs (\sout)
                                % normalem makes italics be italics, not underlines

    % Colors for the hyperref package
    \definecolor{urlcolor}{rgb}{0,.145,.698}
    \definecolor{linkcolor}{rgb}{.71,0.21,0.01}
    \definecolor{citecolor}{rgb}{.12,.54,.11}

    % ANSI colors
    \definecolor{ansi-black}{HTML}{3E424D}
    \definecolor{ansi-black-intense}{HTML}{282C36}
    \definecolor{ansi-red}{HTML}{E75C58}
    \definecolor{ansi-red-intense}{HTML}{B22B31}
    \definecolor{ansi-green}{HTML}{00A250}
    \definecolor{ansi-green-intense}{HTML}{007427}
    \definecolor{ansi-yellow}{HTML}{DDB62B}
    \definecolor{ansi-yellow-intense}{HTML}{B27D12}
    \definecolor{ansi-blue}{HTML}{208FFB}
    \definecolor{ansi-blue-intense}{HTML}{0065CA}
    \definecolor{ansi-magenta}{HTML}{D160C4}
    \definecolor{ansi-magenta-intense}{HTML}{A03196}
    \definecolor{ansi-cyan}{HTML}{60C6C8}
    \definecolor{ansi-cyan-intense}{HTML}{258F8F}
    \definecolor{ansi-white}{HTML}{C5C1B4}
    \definecolor{ansi-white-intense}{HTML}{A1A6B2}

    % commands and environments needed by pandoc snippets
    % extracted from the output of `pandoc -s`
    \providecommand{\tightlist}{%
      \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
    \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
    % Add ',fontsize=\small' for more characters per line
    \newenvironment{Shaded}{}{}
    \newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
    \newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
    \newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
    \newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
    \newcommand{\RegionMarkerTok}[1]{{#1}}
    \newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\NormalTok}[1]{{#1}}

    % Additional commands for more recent versions of Pandoc
    \newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{{#1}}}
    \newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{{#1}}}
    \newcommand{\ImportTok}[1]{{#1}}
    \newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{{#1}}}}
    \newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{{#1}}}
    \newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{{#1}}}
    \newcommand{\BuiltInTok}[1]{{#1}}
    \newcommand{\ExtensionTok}[1]{{#1}}
    \newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{{#1}}}
    \newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{{#1}}}
    \newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}

    % Define a nice break command that doesn't care if a line doesn't already
    % exist.
    \def\br{\hspace*{\fill} \\* }
    % Math Jax compatability definitions
    \def\gt{>}
    \def\lt{<}
    % Document parameters
    \title{simplified\_sdc\_derivations}

    % Pygments definitions

\makeatletter
\def\PY@reset{\let\PY@it=\relax \let\PY@bf=\relax%
    \let\PY@ul=\relax \let\PY@tc=\relax%
    \let\PY@bc=\relax \let\PY@ff=\relax}
\def\PY@tok#1{\csname PY@tok@#1\endcsname}
\def\PY@toks#1+{\ifx\relax#1\empty\else%
    \PY@tok{#1}\expandafter\PY@toks\fi}
\def\PY@do#1{\PY@bc{\PY@tc{\PY@ul{%
    \PY@it{\PY@bf{\PY@ff{#1}}}}}}}
\def\PY#1#2{\PY@reset\PY@toks#1+\relax+\PY@do{#2}}

\expandafter\def\csname PY@tok@w\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
\expandafter\def\csname PY@tok@c\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@cp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.74,0.48,0.00}{##1}}}
\expandafter\def\csname PY@tok@k\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@kp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@kt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
\expandafter\def\csname PY@tok@o\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@ow\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@nb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@nc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@nn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@ne\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.82,0.25,0.23}{##1}}}
\expandafter\def\csname PY@tok@nv\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@no\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@nl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@ni\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.60,0.60,0.60}{##1}}}
\expandafter\def\csname PY@tok@na\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.49,0.56,0.16}{##1}}}
\expandafter\def\csname PY@tok@nt\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@s\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sd\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@si\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@se\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.13}{##1}}}
\expandafter\def\csname PY@tok@sr\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@ss\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@sx\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@m\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@gh\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@gu\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@gd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@gi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@gr\endcsname{\def\PY@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@ge\endcsname{\let\PY@it=\textit}
\expandafter\def\csname PY@tok@gs\endcsname{\let\PY@bf=\textbf}
\expandafter\def\csname PY@tok@gp\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@go\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.53,0.53}{##1}}}
\expandafter\def\csname PY@tok@gt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
\expandafter\def\csname PY@tok@err\endcsname{\def\PY@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
\expandafter\def\csname PY@tok@kc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@kd\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@kn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@kr\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@bp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@fm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@vc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@vg\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@vi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@vm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@sa\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@dl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@s2\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@sh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@s1\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@mb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@mf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@mh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@mi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@il\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@mo\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@ch\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@cm\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@cpf\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@c1\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@cs\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}

\def\PYZbs{\char`\\}
\def\PYZus{\char`\_}
\def\PYZob{\char`\{}
\def\PYZcb{\char`\}}
\def\PYZca{\char`\^}
\def\PYZam{\char`\&}
\def\PYZlt{\char`\<}
\def\PYZgt{\char`\>}
\def\PYZsh{\char`\#}
\def\PYZpc{\char`\%}
\def\PYZdl{\char`\$}
\def\PYZhy{\char`\-}
\def\PYZsq{\char`\'}
\def\PYZdq{\char`\"}
\def\PYZti{\char`\~}
% for compatibility with earlier versions
\def\PYZat{@}
\def\PYZlb{[}
\def\PYZrb{]}
\makeatother

    % Exact colors from NB
    \definecolor{incolor}{rgb}{0.0, 0.0, 0.5}
    \definecolor{outcolor}{rgb}{0.545, 0.0, 0.0}

    % Prevent overflowing lines due to hard-to-break entities
    \sloppy 
    % Setup hyperref package
    \hypersetup{
      breaklinks=true,  % so long urls are correctly broken across lines
      colorlinks=true,
      urlcolor=urlcolor,
      linkcolor=linkcolor,
      citecolor=citecolor,
      }
    % Slightly bigger margins than the latex defaults

    \geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

    \begin{document}

    \maketitle

    \hypertarget{derivations-for-the-castro-simplified-sdc-solver}{%
\section{Derivations for the Castro Simplified-SDC
Solver}\label{derivations-for-the-castro-simplified-sdc-solver}}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}1}]:} \PY{k+kn}{from} \PY{n+nn}{sympy} \PY{k}{import} \PY{n}{init\PYZus{}session}
        \PY{n}{init\PYZus{}session}\PY{p}{(}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
IPython console for SymPy 1.3 (Python 3.7.7-64-bit) (ground types: python)

These commands were executed:
>>> from \_\_future\_\_ import division
>>> from sympy import *
>>> x, y, z, t = symbols('x y z t')
>>> k, m, n = symbols('k m n', integer=True)
>>> f, g, h = symbols('f g h', cls=Function)
>>> init\_printing()

Documentation can be found at http://docs.sympy.org/1.3/

    \end{Verbatim}

    \hypertarget{jacobian}{%
\subsection{Jacobian}\label{jacobian}}

We want to compute the Jacobian \(\partial w/\partial \mathcal{U}\) with

\(\mathcal{U} = (\rho X_\alpha, \rho X_\beta, \rho E, \rho e)^\intercal\)

\(w = (\rho, X_\alpha, X_\beta, T)^\intercal\)

We'll start by writing down

\(A = \partial \mathcal{U}/\partial w\)

and then find its inverse

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}2}]:} \PY{n}{r}\PY{p}{,} \PY{n}{E}\PY{p}{,} \PY{n}{e}\PY{p}{,} \PY{n}{Xa}\PY{p}{,} \PY{n}{Xb}\PY{p}{,} \PY{n}{U}\PY{p}{,} \PY{n}{K} \PY{o}{=} \PY{n}{symbols}\PY{p}{(}\PY{l+s+sa}{r}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{rho E e X\PYZus{}}\PY{l+s+s2}{\PYZob{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{alpha\PYZcb{} X\PYZus{}}\PY{l+s+s2}{\PYZob{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{beta\PYZcb{} U K}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
        \PY{n}{dedr}\PY{p}{,} \PY{n}{dedT}\PY{p}{,} \PY{n}{dedXa}\PY{p}{,} \PY{n}{dedXb} \PY{o}{=} \PY{n}{symbols}\PY{p}{(}\PY{l+s+sa}{r}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{frac}\PY{l+s+s2}{\PYZob{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{partial}\PY{l+s+si}{\PYZob{}\PYZcb{}}\PY{l+s+s2}{e\PYZcb{}}\PY{l+s+s2}{\PYZob{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{partial}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{rho\PYZcb{} }\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{frac}\PY{l+s+s2}{\PYZob{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{partial}\PY{l+s+si}{\PYZob{}\PYZcb{}}\PY{l+s+s2}{e\PYZcb{}}\PY{l+s+s2}{\PYZob{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{partial}\PY{l+s+si}{\PYZob{}\PYZcb{}}\PY{l+s+s2}{T\PYZcb{} }\PY{l+s+s2}{\PYZdq{}} \PY{o}{+}
                                           \PY{l+s+sa}{r}\PY{l+s+s2}{\PYZdq{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{frac}\PY{l+s+s2}{\PYZob{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{partial}\PY{l+s+si}{\PYZob{}\PYZcb{}}\PY{l+s+s2}{e\PYZcb{}}\PY{l+s+s2}{\PYZob{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{partial}\PY{l+s+si}{\PYZob{}\PYZcb{}}\PY{l+s+s2}{X\PYZus{}}\PY{l+s+s2}{\PYZob{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{alpha\PYZcb{}\PYZcb{} }\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{frac}\PY{l+s+s2}{\PYZob{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{partial}\PY{l+s+si}{\PYZob{}\PYZcb{}}\PY{l+s+s2}{e\PYZcb{}}\PY{l+s+s2}{\PYZob{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{partial}\PY{l+s+si}{\PYZob{}\PYZcb{}}\PY{l+s+s2}{X\PYZus{}}\PY{l+s+s2}{\PYZob{}}\PY{l+s+s2}{\PYZbs{}}\PY{l+s+s2}{beta\PYZcb{}\PYZcb{}}\PY{l+s+s2}{\PYZdq{}}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}3}]:} \PY{n}{A} \PY{o}{=} \PY{n}{Matrix}\PY{p}{(}
            \PY{p}{[}\PY{p}{[}\PY{n}{Xa}\PY{p}{,} \PY{n}{r}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,}
             \PY{p}{[}\PY{n}{Xb}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{,} \PY{n}{r}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,}
             \PY{p}{[}\PY{n}{e} \PY{o}{+} \PY{n}{r}\PY{o}{*}\PY{n}{dedr} \PY{o}{+} \PY{n}{Rational}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{)}\PY{o}{*}\PY{n}{U}\PY{o}{*}\PY{o}{*}\PY{l+m+mi}{2}\PY{p}{,} \PY{n}{r}\PY{o}{*}\PY{n}{dedXa}\PY{p}{,} \PY{n}{r}\PY{o}{*}\PY{n}{dedXb}\PY{p}{,} \PY{n}{r}\PY{o}{*}\PY{n}{dedT}\PY{p}{]}\PY{p}{,}     
             \PY{p}{[}\PY{n}{e} \PY{o}{+} \PY{n}{r}\PY{o}{*}\PY{n}{dedr}\PY{p}{,} \PY{n}{r}\PY{o}{*}\PY{n}{dedXa}\PY{p}{,} \PY{n}{r}\PY{o}{*}\PY{n}{dedXb}\PY{p}{,} \PY{n}{r}\PY{o}{*}\PY{n}{dedT}\PY{p}{]}\PY{p}{]}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}4}]:} \PY{n}{A}
\end{Verbatim}

\texttt{\color{outcolor}Out[{\color{outcolor}4}]:}

    $$\left[\begin{matrix}X_{\alpha} & \rho & 0 & 0\\X_{\beta} & 0 & \rho & 0\\\frac{U^{2}}{2} + \frac{\partial{}e}{\partial\rho} \rho + e & \frac{\partial{}e}{\partial{}X_{\alpha}} \rho & \frac{\partial{}e}{\partial{}X_{\beta}} \rho & \frac{\partial{}e}{\partial{}T} \rho\\\frac{\partial{}e}{\partial\rho} \rho + e & \frac{\partial{}e}{\partial{}X_{\alpha}} \rho & \frac{\partial{}e}{\partial{}X_{\beta}} \rho & \frac{\partial{}e}{\partial{}T} \rho\end{matrix}\right]$$

    The inverse of \(A\) is \(\partial w/\partial \mathcal{U}\)

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}5}]:} \PY{n}{simplify}\PY{p}{(}\PY{n}{A}\PY{o}{.}\PY{n}{inv}\PY{p}{(}\PY{p}{)}\PY{p}{)}
\end{Verbatim}

\texttt{\color{outcolor}Out[{\color{outcolor}5}]:}

    $$\left[\begin{matrix}0 & 0 & \frac{2}{U^{2}} & - \frac{2}{U^{2}}\\\frac{1}{\rho} & 0 & - \frac{2 X_{\alpha}}{U^{2} \rho} & \frac{2 X_{\alpha}}{U^{2} \rho}\\0 & \frac{1}{\rho} & - \frac{2 X_{\beta}}{U^{2} \rho} & \frac{2 X_{\beta}}{U^{2} \rho}\\- \frac{\frac{\partial{}e}{\partial{}X_{\alpha}}}{\frac{\partial{}e}{\partial{}T} \rho} & - \frac{\frac{\partial{}e}{\partial{}X_{\beta}}}{\frac{\partial{}e}{\partial{}T} \rho} & \frac{2 \left(X_{\alpha} \frac{\partial{}e}{\partial{}X_{\alpha}} + X_{\beta} \frac{\partial{}e}{\partial{}X_{\beta}} - \frac{\partial{}e}{\partial\rho} \rho - e\right)}{U^{2} \frac{\partial{}e}{\partial{}T} \rho} & \frac{U^{2} - 2 X_{\alpha} \frac{\partial{}e}{\partial{}X_{\alpha}} - 2 X_{\beta} \frac{\partial{}e}{\partial{}X_{\beta}} + 2 \frac{\partial{}e}{\partial\rho} \rho + 2 e}{U^{2} \frac{\partial{}e}{\partial{}T} \rho}\end{matrix}\right]$$

    Compare to the form of the matrix with total energy in the full SDC
algorithm

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}6}]:} \PY{n}{A} \PY{o}{=} \PY{n}{Matrix}\PY{p}{(}
            \PY{p}{[}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,}
             \PY{p}{[}\PY{n}{Xa}\PY{p}{,} \PY{n}{r}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,}
             \PY{p}{[}\PY{n}{Xb}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{,} \PY{n}{r}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,}
             \PY{p}{[}\PY{n}{e} \PY{o}{+} \PY{n}{r}\PY{o}{*}\PY{n}{dedr} \PY{o}{+} \PY{n}{Rational}\PY{p}{(}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{2}\PY{p}{)}\PY{o}{*}\PY{n}{U}\PY{o}{*}\PY{o}{*}\PY{l+m+mi}{2}\PY{p}{,} \PY{n}{r}\PY{o}{*}\PY{n}{dedXa}\PY{p}{,} \PY{n}{r}\PY{o}{*}\PY{n}{dedXb}\PY{p}{,} \PY{n}{r}\PY{o}{*}\PY{n}{dedT}\PY{p}{]}\PY{p}{]}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}7}]:} \PY{n}{A}
\end{Verbatim}

\texttt{\color{outcolor}Out[{\color{outcolor}7}]:}

    $$\left[\begin{matrix}1 & 0 & 0 & 0\\X_{\alpha} & \rho & 0 & 0\\X_{\beta} & 0 & \rho & 0\\\frac{U^{2}}{2} + \frac{\partial{}e}{\partial\rho} \rho + e & \frac{\partial{}e}{\partial{}X_{\alpha}} \rho & \frac{\partial{}e}{\partial{}X_{\beta}} \rho & \frac{\partial{}e}{\partial{}T} \rho\end{matrix}\right]$$

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}8}]:} \PY{n}{simplify}\PY{p}{(}\PY{n}{A}\PY{o}{.}\PY{n}{inv}\PY{p}{(}\PY{p}{)}\PY{p}{)}
\end{Verbatim}

\texttt{\color{outcolor}Out[{\color{outcolor}8}]:}

    $$\left[\begin{matrix}1 & 0 & 0 & 0\\- \frac{X_{\alpha}}{\rho} & \frac{1}{\rho} & 0 & 0\\- \frac{X_{\beta}}{\rho} & 0 & \frac{1}{\rho} & 0\\\frac{- \frac{U^{2}}{2} + X_{\alpha} \frac{\partial{}e}{\partial{}X_{\alpha}} + X_{\beta} \frac{\partial{}e}{\partial{}X_{\beta}} - \frac{\partial{}e}{\partial\rho} \rho - e}{\frac{\partial{}e}{\partial{}T} \rho} & - \frac{\frac{\partial{}e}{\partial{}X_{\alpha}}}{\frac{\partial{}e}{\partial{}T} \rho} & - \frac{\frac{\partial{}e}{\partial{}X_{\beta}}}{\frac{\partial{}e}{\partial{}T} \rho} & \frac{1}{\frac{\partial{}e}{\partial{}T} \rho}\end{matrix}\right]$$

    \hypertarget{second-attempt}{%
\subsection{Second Attempt}\label{second-attempt}}

This adds rho to \(\mathcal{U}\) and the kinetic energy to \(w\):

\(\mathcal{U} = (\rho, \rho_\alpha, \rho X_\beta, \rho E, \rho e)^\intercal\)

\(w = (\rho, X_\alpha, X_\beta, K, T)^\intercal\)

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}9}]:} \PY{n}{A} \PY{o}{=} \PY{n}{Matrix}\PY{p}{(}
            \PY{p}{[}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,}
             \PY{p}{[}\PY{n}{Xa}\PY{p}{,} \PY{n}{r}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,}
             \PY{p}{[}\PY{n}{Xb}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{,} \PY{n}{r}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,}
             \PY{p}{[}\PY{n}{e} \PY{o}{+} \PY{n}{r}\PY{o}{*}\PY{n}{dedr} \PY{o}{+} \PY{n}{K}\PY{p}{,} \PY{n}{r}\PY{o}{*}\PY{n}{dedXa}\PY{p}{,} \PY{n}{r}\PY{o}{*}\PY{n}{dedXb}\PY{p}{,} \PY{n}{r}\PY{p}{,} \PY{n}{r}\PY{o}{*}\PY{n}{dedT}\PY{p}{]}\PY{p}{,}     
             \PY{p}{[}\PY{n}{e} \PY{o}{+} \PY{n}{r}\PY{o}{*}\PY{n}{dedr}\PY{p}{,} \PY{n}{r}\PY{o}{*}\PY{n}{dedXa}\PY{p}{,} \PY{n}{r}\PY{o}{*}\PY{n}{dedXb}\PY{p}{,} \PY{l+m+mi}{0}\PY{p}{,} \PY{n}{r}\PY{o}{*}\PY{n}{dedT}\PY{p}{]}\PY{p}{]}\PY{p}{)}
\end{Verbatim}

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}10}]:} \PY{n}{A}
\end{Verbatim}

\texttt{\color{outcolor}Out[{\color{outcolor}10}]:}

    $$\left[\begin{matrix}1 & 0 & 0 & 0 & 0\\X_{\alpha} & \rho & 0 & 0 & 0\\X_{\beta} & 0 & \rho & 0 & 0\\K + \frac{\partial{}e}{\partial\rho} \rho + e & \frac{\partial{}e}{\partial{}X_{\alpha}} \rho & \frac{\partial{}e}{\partial{}X_{\beta}} \rho & \rho & \frac{\partial{}e}{\partial{}T} \rho\\\frac{\partial{}e}{\partial\rho} \rho + e & \frac{\partial{}e}{\partial{}X_{\alpha}} \rho & \frac{\partial{}e}{\partial{}X_{\beta}} \rho & 0 & \frac{\partial{}e}{\partial{}T} \rho\end{matrix}\right]$$

    \begin{Verbatim}[commandchars=\\\{\}]
{\color{incolor}In [{\color{incolor}11}]:} \PY{n}{simplify}\PY{p}{(}\PY{n}{A}\PY{o}{.}\PY{n}{inv}\PY{p}{(}\PY{p}{)}\PY{p}{)}
\end{Verbatim}

\texttt{\color{outcolor}Out[{\color{outcolor}11}]:}

    $$\left[\begin{matrix}1 & 0 & 0 & 0 & 0\\- \frac{X_{\alpha}}{\rho} & \frac{1}{\rho} & 0 & 0 & 0\\- \frac{X_{\beta}}{\rho} & 0 & \frac{1}{\rho} & 0 & 0\\- \frac{K}{\rho} & 0 & 0 & \frac{1}{\rho} & - \frac{1}{\rho}\\\frac{X_{\alpha} \frac{\partial{}e}{\partial{}X_{\alpha}} + X_{\beta} \frac{\partial{}e}{\partial{}X_{\beta}} - \frac{\partial{}e}{\partial\rho} \rho - e}{\frac{\partial{}e}{\partial{}T} \rho} & - \frac{\frac{\partial{}e}{\partial{}X_{\alpha}}}{\frac{\partial{}e}{\partial{}T} \rho} & - \frac{\frac{\partial{}e}{\partial{}X_{\beta}}}{\frac{\partial{}e}{\partial{}T} \rho} & 0 & \frac{1}{\frac{\partial{}e}{\partial{}T} \rho}\end{matrix}\right]$$

    % Add a bibliography block to the postdoc

    \end{document}

### 2020-08-11 simplified_sdc_intro/paper.tex
new title
some more fixet s

@@
 %======================================================================
-%\title{A Simple Spectral Deferred Correction Method to Couple Nuclear Burning and Hydrodynamics for Detonations}
+\title{A Simplified Spectral Deferred Correction Method for Coupling Hydrodynamics with Reaction Networks and Nuclear Statistical Equilibrium}
 
-\title{So you've decided to detonate helium}
-
-\shorttitle{Detonations and NSE}
+\shorttitle{A Simplified SDC Method}
 \shortauthors{}
@@
 \begin{abstract}
-Helium detonations can be challenging to model because of the
-difficulty in accurately coupling hydrodynamics and reactions.  This
-is particularly acute in the ash of the detonation where nuclear
-statistical equilibrium is established.  We look a several different
-methods of coupling explicit hydrodynamics and stiff reaction sources,
-applied to detonations which reach NSE conditions.  This is all done
-in the framework of the \castro\ hydrodynamics code, and all algorithm
-implementations are freely available.
+Reacting astrophysical flows can be challenging to model because of
+the difficulty in accurately coupling hydrodynamics and reactions.
+This can be particularly acute at high temperatures where nuclear
+statistical equilibrium is established.  We develop a simplified
+approach to spectral deferred corrections coupling of explicit
+hydrodynamics and stiff reaction sources and apply it to some example
+problems.  This is all done in the framework of the
+\castro\ hydrodynamics code, and all algorithm implementations are
+freely available.
 \end{abstract}
@@
 with a table of NSE abundances and the zone's composition set through
-table look-ups\MarginPar{example? GCD?}.
+table look-ups (e.g.\ \citealt{ma:2013}) \MarginPar{need more here}
 
-In this paper, we look at some different approaches to coupling
-hydrodynamics and reactions, using a helium detonation that attains
-NSE as our testbed.  We use the \castro\ hydrodynamics code for all of
+In \cite{castro_sdc}, we developed second- and fourth-order accurate
+methods in space and time for coupling hydrodynamics and nuclear
+reaction networks and demonstrated the method of a variety of test
+problems.  The approach her is considerably simpler, reusing the main
+CTU hydrodynamics construction and a largely similar ODE integration
+scheme, making this method easier to add to existing simulation codes.
+However, it is restricted to second-order accuracy overall.  We term this
+algorithm the ``simplified SDC method''.
+We use the \castro\ hydrodynamics code for all of
 our numerical experiments, and all of the code to reproduce the results
@@
 We solve the Euler equations for compressible, reacting flow.  We'll focus on
-one-dimensional flow, and the system appears as:
+one-dimensional flow for this description, and the system appears as:
 \begin{equation}
@@
 \end{equation}
-where $\Uc = (\rho, (\rho X_k), (\rho u), (\rho E))^\intercal$ are
-the conserved fluid quantities: mass density, $\rho$, velocity, $u$,
+where
+\begin{equation}
+\Uc = \left ( \begin{array}{c}
+           \rho \\
+           \rho X_k \\
+           \rho u \\
+           \rho E \end{array}\right )
+\end{equation}
+are the conserved fluid quantities: mass density, $\rho$, velocity, $u$,
 specific total energy, $E$, and nuclear
@@
 \end{equation}
-In this paper, we will not consider any hydrodynamical sources, but we
-will carry the term $\Hb$ around for completeness.  The reactive
+ The reactive
 sources take the form:
@@
 methods for stabilty.  Operator splitting is traditionally employed
-here, and our goal in this paper is to look at several different
-approaches for splitting.
-
-We describe several time-integration techniques for coupling
-hydrodynamics and reactions in the next sections.
+here, and we compare to it in this paper.
 
 \subsection{Strang Splitting}
+
+In Strang splitting, we first evolve the system without advection
+through $\Delta t/2$, then do the advective update without reactions
+through $\Delta t$, and finally do the last reactive update through $\Delta t/2$.
+
+In the absence of advective terms, our system appears as:
+\begin{align}
+\odt{\rho} & = 0 \\
+\odt{(\rho X_k)} &= \rho \omegadot_k \\
+\odt{(\rho u)} &= 0 \\
+\odt{(\rho E)} &= \rho \Sdot 
+\end{align}
+Ww can write the energy equation as:
+\begin{equation}
+\odt{(\rho E)} = \odt{(\rho e)} + \odt{K} = \rho \Sdot
+\end{equation}
+where $K$ is the kinetic energy, $K = |u|^2/2$.  Since the velocity is
+unchanged by reactions, this means that
+\begin{equation}
+\odt{(\rho e)} = \rho \odt{e} = \rho \Sdot
+\end{equation}
+The reaction rates are typically expressed as $\omegadot_k(\rho, T,
+X_k)$, if we evolve this system.  This requires an EOS inversion each
+time we need to evaluate the reactive terms.
+
+Normally we want to integrate temperature temperature.  We can write the 
+energy equation as:
+\begin{equation}
+\rho \odt{e} = \rho \left [ \underbrace{\left . \frac{\partial e}{\partial T} \right |_\rho}_{c_v}
+                            \odt{T} +
+                            \left . \frac{\partial e}{\partial \rho} \right |_T
+                            \cancelto{0}{\odt{\rho}} \;\; \right ] = \rho \Sdot
+\end{equation}
+reducing to
+\begin{equation}
+\odt{T} = \frac{1}{c_v} \Sdot
+\end{equation}
+We can alternately derive this from an enthalpy equation, which is the correct
+formulation when we are at constant pressure, and we get the same temperature
+equation but with $c_v$ replaced by $c_p$.
+Technically this requires an EOS call to get $c_x$, but if the
+specific heat varies slowly, we can freeze it during the reaction
+solve.
 
@@
 the energy generation or abundance changes over a timestep.
+
+write these out
 
@@
 of methods has the most difficulty in integrating the detonation.
-
-\subsection{Balanced Splitting}
 
@@
         \rho(t) &= \rho^n + \Adv{\rho}^{n+1/2,(k)} \, t \\
-        (\rho \Ub)(t) &= (\rho \Ub)^n + \Adv{(\rho \Ub)}^{n+1/2,(k)} \, t
+        (\rho \Ub)(t) &= (\rho \Ub)^n + \Adv{\rho \Ub}^{n+1/2,(k)} \, t
       \end{align}
@@
 
-\subsection{Old Primitive Stuff}
-
-and our primitive system as:
-\begin{equation}
-\qb_t = \Advs{\qb} + \Rb(\qb)
-\end{equation}
-with
-\begin{equation}
-\Advs{\qb} = -\Ab^\xv(\qb) \qb_x - \Ab^\yv(\qb) \qb_y - \Ab^\zv(\qb) \qb_z + \Sqhydro
-\end{equation}
+\subsection{Nuclear Statistical Equilibrium}
 

### 2020-08-11 ecsn_flamespeed/paper.tex
Initial Commit

%\documentclass[12pt, preprint]{aastex}
\documentclass[preprint,times,tighten]{aastex63}

\usepackage{epsf,color,amsmath}

\usepackage{cancel}

\newcommand{\sfrac}[2]{\mathchoice%
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
    \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
    \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptscriptfont0 #1}\kern-.2em/
    \kern-.15em\lower.25ex\hbox{\the\scriptscriptfont0 #2}} {#1\!/#2}}

\newcommand{\myhalf}{\sfrac{1}{2}}
\newcommand{\nph}{{n+\myhalf}}
\newcommand{\nmh}{{n-\myhalf}}

\newcommand{\inp}{\mathrm{in}}
\newcommand{\outp}{\mathrm{out}}

% boldsymbol means bold italic
\newcommand{\eb}{{\bf{e}}}
\newcommand{\Ub}{{\bf{U}}}
\newcommand{\xb}{{\bf{x}}}
\newcommand{\kb}{{\bf{k}}}
\newcommand{\Vb}{{\bf{V}}_n}
\newcommand{\Vbhat}{{\bf{\widehat{V}}}_n}
\newcommand{\Omegab}{{\bf{\Omega}}}
\newcommand{\gb}{{\bf{g}}}
\newcommand{\rb}{{\bf{r}}}

\newcommand{\pb}{p_\mathrm{base}}
\newcommand{\epsdot}{\dot{\epsilon}}
\newcommand{\qburn}{q_\mathrm{burn}}
\newcommand{\rt}{\tilde{r}_0}

\newcommand{\nablab}{\mathbf{\nabla}}
\newcommand{\dt}{\Delta\ t}

\newcommand{\omegadot}{\dot{\omega}}

\newcommand{\Hext}{H_{\rm ext}}
\newcommand{\Hnuc}{H_{\rm nuc}}
\newcommand{\kth}{k_{\rm th}}

\newcommand{\Gammaonebar}{\overline{\Gamma}_1}
\newcommand{\Sbar}{\overline{S}}

\newcommand{\etarho}{\eta_\rho}
\newcommand{\etarhoh}{\eta_{\rho~h}}

\newcommand{\Ubt}{\widetilde{\Ub}}
\newcommand{\wt}{\widetilde{w}}

\newcommand{\He}{$^4$He}
\newcommand{\C}{$^{12}$C}
\newcommand{\Fe}{$^{56}$Fe}

\newcommand{\isot}[2]{$^{#2}\mathrm{#1}$}
\newcommand{\isotm}[2]{{}^{#2}\mathrm{#1}}

\newcommand{\maestro}{{\sf MAESTRO}}
\newcommand{\castro}{{\sf Castro}}
\newcommand{\amrex}{{\sf AMReX}}
\newcommand{\pynucastro}{{\sf pynucastro}}

\newcommand{\avg}[1]{\overline{#1}}
\newcommand{\avgtwod}[1]{\langle~#1 \rangle}
\newcommand{\rms}[2]{\left(\delta#1\right)_{r_{#2}}}
\newcommand{\mymax}[1]{\left(#1\right)_{\rm max}}

\newcommand{\Tb}{\ensuremath{T_\mathrm{base}}}
\newcommand{\gcc}{\mathrm{g~cm^{-3} }}
\newcommand{\cms}{\mathrm{cm~s^{-1} }}

\newcommand{\half}{\frac{1}{2}}

\setlength{\marginparwidth}{0.5in}

\newcommand{\MarginPar}[1]{
    \marginpar{\vskip-\baselineskip%
               \raggedright%
               \tiny\sffamily%
               {\color{red}\hrule%
               \smallskip%
               #1\par%
               \smallskip%
               \hrule}}%
}

\newcommand{\AssignTo}[1]{
    \marginpar{\vskip-\baselineskip%
               \raggedright%
               \tiny\sffamily%
               {\color{blue}\hrule%
               \smallskip%
               #1\par%
               \smallskip%
               \hrule}}%
}

\begin{document}
%======================================================================
% Title
%======================================================================
\title{Effect of numerical methods on Laminar Flame Speeds}

\shorttitle{ECSN Flame Speeds}
%% \shortauthors{Short Authors}

%% \author[ORCiD]{Name}
%% \affiliation{Affiliation}

%% \author[ORCiD]{Name}
%% \affiliation{Affiliation}

%% \author[ORCiD]{Name}
%% \affiliation{Affiliation}

%% \author[ORCiD]{Name}
%% \affiliation{Affiliation}

%% \correspondingauthor{}
%% \email{}

%======================================================================
% Abstract and Keywords
%======================================================================
\begin{abstract}
\end{abstract}

\keywords{}

%======================================================================
% Introduction
%======================================================================
\section{Introduction}\label{Sec:Introduction}
In astrophysical explosions, there are two types of burning that can occur, detonation and deflagration. The former is a supersonic blast, that travels faster than the speed of sound in a material. The latter however, is subsonic, which means the sound waves propogate ahead of the material flow. If the sound speed is much faster than the velocity of the fluid, this is known as a low Mach number flow. When this is the case, sound waves carry very little energy in this regime and to a good aproximation, can be filtered out from the simulation. This allows us to greatly increase the CFL condition, taking much longer time steps than were possible before when we had to model the sound waves as well. This type of modeling is extremely important in astrophysical regimes like convection and core collapse supernovae.

When a degenerate oxygen-neon core reaches a central density of $\approx 10^{10} \frac{g}{cm^3}$, the star is dense enough to facilitate electron capture onto $^{20}Ne$ leading to a deflagration wave that propogates away from the core of the star. As the flame propogates outwards, there is a competition between nuclear energy release from the deflagration wave and nuclear energy capture by the electron capture onto $^{20}Ne$ behind the wave. This will either result in an explosion, a supernovae, or an implosion, forming a neutron star.  \citep{1991ApJ...367L..19N} \citep{Zha_2019}

As shown in \cite{Schwab_2020} our goal is to produce tabulated laminar flame speed calculations for core collapse supernovae. Previously this was done in MESA \citep{MESA}, a 1D stellar evolution code. In this paper, we use MAESTROeX \citep{MAESTROeX}, a low Mach number stellar hydrodyamic code to produce the tabulated values. In particular, we will explore the low Mach number approximation and it's effects on flame speed, as well as spectral deferred corrections (SDC) versus traditional strang splitting. In addition, we will be running these simulations in 2D, which can bring in additional effects such as turbuluence not accounted for in a 1D model.

%======================================================================
% Numerical Approach
%======================================================================
\section{Numerical Approach}\label{Sec:numerics}
The simulations in this paper were all run using the open source code MAESTROeX, a compressible fluid dynamics code for low Mach number reacting flow. MAESTROeX evolves the following set of equations which are, conservation of mass, momentum, and energy and a constraint on the velocity.

\begin{equation}
  \frac{\partial\left(\rho X_{k}\right)}{\partial t}=-\nabla \cdot\left(\rho X_{k} \mathbf{U}\right)+\rho \dot{\omega}_{k}
\end{equation}

\begin{equation}
  \frac{\partial \mathbf{U}}{\partial t}=-\mathbf{U} \cdot \nabla \mathbf{U}-\frac{\beta_{0}}{\rho} \nabla\left(\frac{\pi}{\beta_{0}}\right)-\frac{\rho-\rho_{0}}{\rho} g \mathbf{e}_{r}
\end{equation}

\begin{equation}
  \frac{\partial(\rho h)}{\partial t}=-\nabla \cdot(\rho h \mathbf{U})+\frac{D p_{0}}{D t}+\rho H_{\mathrm{nuc}}
\end{equation}

\begin{equation}
  \frac{\partial \rho}{\partial t}=-\nabla \cdot(\rho \mathbf{U})
\end{equation}

where $\rho$ is mass density, $\mathbf{U}$ is fluid velocity, $h$ is specific enthalpy, $H_{\mathrm{nuc}}$ is the energy released per time per unit mass, $X_k$ is the mass fraction for species $k$, and $\dot{\omega}_{k}$ is the associated production rate of species $k$. For more about MAESTROeX and the low Mach number hydrodynamic assumptions, please see \cite{MAESTROeX}.

MAESTROeX solves these equations using adaptive mesh refinement (AMR), block-structured, Cartesian grid finite-volume, made possible by being built on AMReX \citep{amrex_joss}. MAESTROeX uses an explicit Godunov integration scheme for advection, VODE \citep{vode} ODE solver reactions.

Spectral Deferred Corrections (SDC) is an iterative scheme, like Strang splitting, used to integrate the thermodynamic variables in MAESTROeX. In \cite{SDC-old} SDC is shown to be more accurate and faster for low Mach number reacting flow. In Strang splitting, MAESTROeX first evolves reactions for one half a time step, then advects a full time step, then finishing evolving reactions for the final time step. SDC actually incoporates advective terms into the system of ODEs we are integrating with reactions, which over the course of a full timestep has proven to be faster and more accurate.

%======================================================================
% Results
%======================================================================
\section{Simulations and Results}\label{Sec:results}
\begin{enumerate}

  \item Get reverse rates working with pynucastro for a full comparison. This is needed for temperatures $> 10^{10} K$
  \item Past work:
  \item First step is to reproduce \cite{TW92} laminar flame speeds.    \cite{Schwab_2020} does this is both CO and ONe mixtures. This would probably be good to do as well.
  \item Along the same lines, \cite{Schwab_2020} showed flame speed had a strong dependence on $Y_e$ and isotope number and provided a fitting function.
  \item It'd be interesting to verify all of that but then be able to show how some of these new? methods come in to play:
  \item SDC vs strang
  \item Low mach hydro
  \item Multidimensional Effects?
  \item What can we say about these methods and how they work and how do they effect flame speed?
\end{enumerate}

\section{Conclusion}

\section{Convergence Studies}\label{Sec:results}

We'll want to run some convergence studies to show things are behaving well and that our tabulated values do indeed converge when we increase resolution.

%Not sure which of the below to exclude / include.
\acknowledgements MAESTROeX is open-source and freely available at
\url{http://github.com/AMReX-Astro/MAESTROeX}.  The problem setups used
here are available in the git repo under {\tt MAESTROeX/Exec/science/flame}.
The work at Stony Brook was supported by DOE/Office
of Nuclear Physics grant DE-FG02-87ER40317.  This research used
resources of the National Energy Research Scientific Computing Center,
a DOE Office of Science User Facility supported by the Office of
Science of the U.~S.\ Department of Energy under Contract
No.\ DE-AC02-05CH11231.  This research used resources of the Oak Ridge
Leadership Computing Facility at the Oak Ridge National Laboratory,
which is supported by the Office of Science of the U.S. Department of
Energy under Contract No. DE-AC05-00OR22725, awarded through the DOE
INCITE program.  We thank NVIDIA Corporation for the donation of a
Titan X and Titan V GPU through their academic grant program.  This
research has made use of NASA's Astrophysics Data System Bibliographic
Services.

\facilities{NERSC}

\software{AMReX \citep{amrex_joss},
          MAESTROeX \citep{MAESTROeX_joss},
          pynucastro \citep{pynucastro}
          GCC (\url{https://gcc.gnu.org/}),
          linux (\url{https://www.kernel.org/}),
          matplotlib (\citealt{Hunter:2007}, \url{http://matplotlib.org/}),
          NumPy \citep{numpy,numpy2},
          python (\url{https://www.python.org/}),
          valgrind \citep{valgrind},
          VODE \citep{vode},
          yt \citep{yt}}

%======================================================================
% References
%======================================================================

\bibliographystyle{aasjournal}
\bibliography{ws}

\end{document}

### 2020-08-14 xrb2/paper.tex
new SciDAC ack

@@
   flame\_wave}.  The work at Stony Brook was supported by DOE/Office
-of Nuclear Physics grant DE-FG02-87ER40317 and the SciDAC program DOE
-grant DE-SC0017955.  MZ acknowledges support from the Simons
+of Nuclear Physics grant DE-FG02-87ER40317.  This material is based upon work supported by the
+U.S. Department of Energy, Office of Science, Office of Advanced
+Scientific Computing Research and Office of Nuclear Physics, Scientific
+Discovery through Advanced Computing (SciDAC) program under Award
+Number DE-SC0017955.  MZ acknowledges support from the Simons
 Foundation.  This research used resources of the National Energy

### 2020-08-15 simplified_sdc_intro/paper.tex
new SciDAC ack
more work on this paper

@@
 statistical equilibrium is established.  We develop a simplified
-approach to spectral deferred corrections coupling of explicit
-hydrodynamics and stiff reaction sources and apply it to some example
-problems.  This is all done in the framework of the
-\castro\ hydrodynamics code, and all algorithm implementations are
-freely available.
+approach to spectral deferred corrections (SDC) coupling of explicit
+hydrodynamics and stiff reaction sources as an alternative to operator
+splitting or the more comprehensive SDC approach we demonstrated
+previously.  We apply the new method some example problems and show
+how to modify it to work with a hybrid network consisting of a
+reaction ODE system and a table for nuclear statistical equilibrium.
+This is all done in the framework of the \castro\ hydrodynamics code,
+and all algorithm implementations are freely available.
 \end{abstract}
@@
 methods in space and time for coupling hydrodynamics and nuclear
-reaction networks and demonstrated the method of a variety of test
-problems.  The approach her is considerably simpler, reusing the main
+reaction networks based on spectral deferred corrections (SDC)
+methods, and demonstrated the method of a variety of test problems.
+The approach presented here is considerably simpler, reusing the main
 CTU hydrodynamics construction and a largely similar ODE integration
 scheme, making this method easier to add to existing simulation codes.
-However, it is restricted to second-order accuracy overall.  We term this
-algorithm the ``simplified SDC method''.
-We use the \castro\ hydrodynamics code for all of
-our numerical experiments, and all of the code to reproduce the results
-in this paper are in the \castro\ github repository.
+However, it is restricted to second-order accuracy overall.  We term
+this algorithm the ``simplified SDC method''.  We use the
+\castro\ hydrodynamics code \cite{castro} for all of our numerical
+experiments, and all of the code to reproduce the results in this
+paper are in the \castro\ github
+repository\footnote{\url{https://github.com/amrex-astro/Castro/}}.
 
@@
 
-We solve the Euler equations for compressible, reacting flow.  We'll focus on
-one-dimensional flow for this description, and the system appears as:
+We solve the Euler equations for compressible, reacting flow.  We'll
+focus on one-dimensional flow for this description, and the system
+appears as:
 \begin{equation}
@@
 
-Finally, $\Sc(\Uc)$ are the source terms, which we decompose into
+We decompose the source terms, $\Sc(\Uc)$, into
 hydrodynamical sources, $\Hb$ (like gravity), and reactive sources,
@@
 
-Sometimes it is preferable to work with the primitive variables,  $\qb
-= (\rho, X_k, u, p, (\rho e))^\intercal$.  Here, the system appears
+Sometimes it is preferable to work with the primitive variables,
+\begin{equation}
+\qb = \left ( \begin{array}{c}
+  \rho \\
+  X_k \\
+  u \\
+  p \\
+  (\rho e) \\
+\end{array} \right )
+\end{equation}
+Here, the system appears
 as:
@@
 
-In the absence of advective terms, our system appears as:
+In the absence of advective terms, our reaction system appears as just
+$d\Uc/dt = \Rb(\Uc)$, or:
 \begin{align}
@@
 \end{align}
-Ww can write the energy equation as:
+We can write the energy equation as:
 \begin{equation}
@@
 The reaction rates are typically expressed as $\omegadot_k(\rho, T,
-X_k)$, if we evolve this system.  This requires an EOS inversion each
+X_k)$, when we evolve this system.  This requires an EOS inversion to get $T$ from $e$ each
 time we need to evaluate the reactive terms.
@@
 \begin{equation}
+\label{eq:strang:T}
 \odt{T} = \frac{1}{c_v} \Sdot
 \end{equation}
-We can alternately derive this from an enthalpy equation, which is the correct
-formulation when we are at constant pressure, and we get the same temperature
-equation but with $c_v$ replaced by $c_p$.
+We can alternately derive this from an enthalpy equation, which is the
+correct formulation when we are at constant pressure, and we get the
+same temperature equation but with $c_v$ replaced by $c_p$.
 Technically this requires an EOS call to get $c_x$, but if the
 specific heat varies slowly, we can freeze it during the reaction
-solve.
+solve or use a Taylor expansion in $T$ to capture slow evolution.
+
+We also typically integrate mass fraction itself, instead of partial
+densities:
+\begin{equation}
+\label{eq:strang:X}
+\odt{X_k} = \omegadot_k
+\end{equation}
+The system of equations \ref{eq:strang:T} and \ref{eq:strang:X} are
+solved using an implicit ODE solver designed for stiff systems of
+equations.  We use VODE~\citep{vode}.
+
+With Strang splitting, the state can drift significantly off of the
+smooth solution to the coupled reactive hydrodynamics equations, as
+shown graphically in \cite{astronum:2018} using an earlier version of
+the present algorithm.
+
+A variation on Strang splitting called (re-)balanced splitting was
+developed in \citet{speth:2013}.
 
@@
 
-In addition to the standard CFL timestep limiter for explicit
-hydrodynamics, \castro\ can also enforcing timestep limiters based on
-the energy generation or abundance changes over a timestep.
+Since this method is based off of the CTU hydrodynamics scheme, it
+benefits from the larger timestep that method can take (when done with
+full corner coupling) as compared to a method-of-lines approach (see
+\citealt{ppmunsplit}).  In addition to the standard CFL timestep
+limiter for explicit hydrodynamics, \castro\ can also enforcing
+timestep limiters based on the energy generation or abundance changes
+over a timestep.
 
@@
 of methods has the most difficulty in integrating the detonation.
-
-A variation on Strang splitting called (re-)balanced splitting was
-developed in \citet{speth:2013}.
 
@@
 \acknowledgements \castro\ is freely available at
-\url{http://github.com/AMReX-Astro/Castro}.  All of the code and problem
-setups used here are available in the git repo.  The work at Stony
-Brook was supported by DOE/Office of Nuclear Physics grant
-DE-FG02-87ER40317.
+\url{http://github.com/AMReX-Astro/Castro}.  All of the code and
+problem setups used here are available in the git repo.  The work at
+Stony Brook was supported by DOE/Office of Nuclear Physics grant
+DE-FG02-87ER40317.  This material is based upon work supported by the
+U.S. Department of Energy, Office of Science, Office of Advanced
+Scientific Computing Research and Office of Nuclear Physics, Scientific
+Discovery through Advanced Computing (SciDAC) program under Award
+Number DE-SC0017955.
 

### 2020-08-15 simplified_sdc_intro/paper.tex
add some NSE stuff

@@
 \citealt{ppmunsplit}).  In addition to the standard CFL timestep
-limiter for explicit hydrodynamics, \castro\ can also enforcing
+limiter for explicit hydrodynamics, \castro\ can also enforce
 timestep limiters based on the energy generation or abundance changes
-over a timestep.
-
-write these out
+over a timestep:
+\begin{align}
+\label{eq:dt:nuce}
+\Delta t &\le \mathtt{dtnuc\_e}\, \min_{i,j,k} \left\{\frac{e_{i,j,k}}{\dot{e}_{i,j,k}}\right\} \\
+%
+\label{eq:dt:nucX}
+\Delta t &\le \mathtt{dtnuc\_X}\, \min_{i,j,k} \left\{\min_n\frac{X^n_{i,j,k}}{\dot{X}^n_{i,j,k}}\right\}
+\end{align}
+where $\mathtt{dtnuc\_e}$ and $\mathtt{dtnuc\_X}$ are runtime parameters
+used to control the allowed change.
 
@@
 failing to converge in the implicit solve, or violation of one of the
-timestep limiters during the step.  At the moment, this retry
-mechanism only works with the CTU hydrodynamics + Strang splitting.
-We enable it here, since, as will be shown shortly, this combination
-of methods has the most difficulty in integrating the detonation.
+timestep limiters during the step.  This means that equations
+\ref{eq:dt:nuce} and \ref{eq:dt:nucX} are not reactive, but instead
+guaranteed to be met throughout the simulation, because the step is
+rejected if they are violated.  The retry mechanism in \castro\ works
+with both the Strang and simplified-SDC integration scheme.
 
@@
 
-The basic idea of spectral deferred corrections is break the timestep
-into a number of temporal nodes and use low order difference methods
-in a correction equation that, upon iteration, achieves a high-order
-solution coupling the different physics together.
+The basic idea of spectral deferred corrections is to express the
+update as an integral and divide the time-update into a number of
+discrete time nodes.  The integral is then approximated using a
+quadrature rule over these time nodes and low order approximations are
+used to update the state from one time node to the next.  The method
+uses iteration to successively improve the solution and the ultimate
+accuracy is determined by the quadrature method used to evaluate the
+integral, which is done using an iteratively-lagged solution.
 
-\subsection{Simplified-SDC Method}
-
-A second-order accurate time integration method using a simple
-deferred correction strategy was shown in \citep{SDC-old}.  A
+We start by writing our update as:
+\begin{equation}
+\Uc^{n+1} = \Uc^n + \int_t^{t+\Delta t} \left [ \Advs{\Uc} + \Rb(\Uc) \right ] dt
+\end{equation}
+For the simplified SDC method we explore here, we will make the
+advective term piecewise constant in time, using the value at the
+midpoint in time, to achieve second-order accuracy, giving us:
+\begin{equation}
+\label{eq:integral:simplesdc}
+\Uc^{n+1} = \Uc^n + \int_t^{t+\Delta t} \left \{ \Adv{\Uc}^{n+1/2} + \Rb(\Uc) \right \} dt
+\end{equation}
+This is the approach first shown in \citep{SDC-old}.  A
 version of this was also implemented in the \maestro\ low-Mach number
@@
 state.  We describe the application of this to the equations of
-compressible hydrodynamics below:
+compressible hydrodynamics below.
 
@@
   \begin{itemize}
-  \item Set the first iterate of the primitive-variable reactive source terms from the
-    last step:
+  \item We need an approximation to how much the reactions alone
+    changed the primitive variable state over the timestep, which we
+    will denote $\Ic_q$.  Since we do not have any information about
+    the current timestep in the first iteration,
+    we use the value from the last iteration of the previous timestep:
     \begin{equation}
-      \Ic^{n+1/2,(0)}_{\qb} = \Ic^{n-1/2,(\smax)}_{\qb}
+      \Ic^{n+1/2,(-1)}_{\qb} = \Ic^{n-1/2,(\smax-1)}_{\qb}
     \end{equation}
@@
 
-  Iterate from $k = 1, \ldots, \smax$.  For second-order accuracy,
-  $\smax = 2$ is sufficient.
+  Iterate from $k = 0, \ldots, \smax-1$.  For second-order accuracy,
+  $\smax = 2$ is sufficient.  In addition to denoting the time-level
+  with a superscript (like $n$ or $n+1$), we'll use a second subscript
+  in parentheses to keep track of the iteration.
 
   \begin{itemize}
-  \item {\em Create the Advective Update Term}
+  \item {\em Create the advective update term, $\Adv{\Uc}^{n+1/2,(k)}$}
 
@@
       terms accounting only for reactions, $\Ic_\qb^{n+1/2,(k-1)}$, as the
-      reactive source.  Any hydrodynamic source terms are time-centered
+      reactive source.  This is in contrast to Strang-splitting, where no
+      explicit reactive source terms are seen by the hydrodynamics update.
+      Any hydrodynamic source terms are time-centered
       using the previous iteration:
@@
     \item solve the Riemann problem at each interface to get a unique
-      conserved state on each interface, $\Uc^{n+1/2}_{i+1/2}$
+      conserved state on each interface, $\Uc^{n+1/2,(k)}_{i+1/2}$
 
     \item construct the advective update terms, first ignoring the hydrodynamics sources:
-      $\Advt{\Uc}^{n+1/2}_{i,j,k}$, e.g., as:
+      $\Advt{\Uc}^{n+1/2,(k)}_{i,j,k}$, e.g., as:
       \begin{align}
         \Advt{\Uc}^{n+1/2,(k)}_{i,j,k} =
-          &- \frac{\Fb^\xv(\Uc^{n+1/2}_{i+1/2}) - \Fb^\xv(\Uc^{n+1/2}_{i-1/2})}{\Delta x} 
+          &- \frac{\Fb^\xv(\Uc^{n+1/2,(k)}_{i+1/2}) - \Fb^\xv(\Uc^{n+1/2,(k)}_{i-1/2})}{\Delta x}
       \end{align}
@@
 
-    \begin{itemize}
-    \item Define the hydro source terms---these will be time-centered,
-      but piecewise-constant in time throughout the integration.
-
-    \item With the spatial discretization done, our system appears as a
-      system of ODEs: \MarginPar{what is the right form of acceleration
-        here? should we do piecewise linear in time?}
+    We update the state by doing the integral in equation
+    \ref{eq:integral:simplesdc}.  Since we are approximating the
+    advective term as piecewise constant in time, we can simply use an
+    ODE integrator to integrate this, just as we do with the reaction
+    system in Strang splitting.  The difference here being that we are
+    integrating the conserved variables and the state sees the effect
+    of advection as we integrate the reactions.  The ODE form of this
+    is
     \begin{equation}
-        \left . \odt{\Uc} \right |_{i,j,k} = \Adv{\Uc}^{n+1/2,(k)}_{i,j,k} + \Rb(\Uc)
+      \odt{\Uc} = \Adv{\Uc}^{n+1,(k)} + \Rb(\Uc)
     \end{equation}
@@
     of state.
+
+    Our integrator also needs the Jacobian of the system, in terms of
+    the conserved variables.  This is different than the form of the
+    Jacobian usually used in reaction networks.  We describe the form
+    of the Jacobian in appendix \ref{sec:app:jacobian}.
 
@@
     \end{align}
-      At the end of the integration, our new state is $\Uc^{n+1,(k)}$.
 
-    \end{itemize}
+    At the end of this integration, we have the state at the next iteration,
+    $\Uc^{n+1,(k+1)}$
 
@@
 
-\subsection{Nuclear Statistical Equilibrium}
+\section{Nuclear Statistical Equilibrium}
+
+For this study we will use the 19 isotope ``aprox19'' network together
+with a table that gives the NSE abundances in regions where the system
+is in NSE.  The table provides:
+\begin{align*}
+Y_e &= \sum_k \frac{Z_k X_k}{A_k} \\
+\bar{A} &= \left [ \sum_k \frac{X_k}{A_k} \right ]^{-1} \\
+\frac{B}{A} &= \sum_k \frac{B_k X_k}{A_k}
+\end{align*}
+where $B_k$ is the binding energy of nucleus $k$.  In our simulations,
+we store these 3 quantities as auxiliary data that is carried along
+with the rest of the fluid state in each zone.
+
+For Strang split coupling of hydro and reactions, $DX_k/Dt = 0$,
+and our evolution equations are:
+\begin{align*}
+\frac{DY_e}{Dt} &= \sum_k \frac{Z_k}{A_k} \frac{DX_k}{Dt} = 0 \\
+\frac{D}{Dt} \frac{1}{\bar{A}} &= - \frac{1}{\bar{A}^2} \frac{D\bar{A}}{Dt} = \sum_k \frac{1}{A_k} \frac{DX_k}{Dt} = 0 \rightarrow \frac{D\bar{A}}{Dt} = 0 \\
+\frac{D}{Dt} \left (\frac{B}{A} \right ) &= \sum_k \frac{B_k}{A_k} \frac{DX_k}{Dt} = 0
+\end{align*}
+Therefore each of these auxillar equations obeys an advection
+equation in the hydro part of the advancement.  In the SDC algorithm,
+there will be a reactive source (an $\Ic_q$) for each of these that is
+computed in the same manner as above.
+
+The NSE table was generated using a 125 nuclei reaction network and is
+the same as described in \cite{ma:2013}.  The compositional quantities
+it carries, $\bar{A}$ and $Y_e$ and not representable from the 19
+isotopes we carry in the main network. For this reason, when we are
+using the NSE network, we always take the composition quantities in
+the EOS directly from the auxiliary state in each zone instead of
+using the $X_k$ directly. Our equation of state needs $\bar{Z}$ in addition
+to the auxiliary quantities, which is is easily computed as
+\begin{equation}
+\bar{Z} = \bar{A} Y_e
+\end{equation}
+
+For Strang splitting, the basic flow of using the aprox19 + NSE network
+is as follows:
+\begin{itemize}
+\item initialize the problem, including $X_k$
+
+\item fill the initial aux data with $Y_e$, $\bar{A}$, and $(B/A)$.
+
+\item in hydro, update these quantities simply via advection
+
+\item for the reactive update:
+
+  \begin{itemize}
+  \item check if NSE applies (see below)
+
+  \item if we are in an NSE region:
+
+    \begin{itemize}
+    \item use $\rho$, $T$, and $Y_e$ to call the table. This
+      returns: $dY_e/dt$, $(B/A)_{\rm out}$, and $\bar{A}_{\rm out}$.
+
+    \item update $Y_e$:
+    \begin{equation}
+      (Y_e)_{\rm out} = (Y_e)_{\rm in} - \Delta t \frac{dY_e}{dt}
+    \end{equation}
+
+  \item update the auxiliary state value of $\bar{A}$ to be simply
+    $\bar{A}_{\rm out}$ from the table.
+
+  \item compute the energy generation rate, $\Sdot$ as:
+    \begin{equation}
+      \Sdot = \eta \left [ \left ( \frac{B}{A} \right )_{\rm out} -
+        \left ( \frac{B}{A} \right )_{\rm in} \right ] N_A \frac{1}{\Delta t}
+    \end{equation}
+    where $\eta$ is an inertia term < 1 to prevent the energy changing
+    too much in one set.
+
+  \item set the auxiliary state value of $(B/A)$ as:
+    \begin{equation}
+      \left (\frac{B}{A}\right )_{\rm out} = \left (\frac{B}{A}\right )_{\rm in} + \eta \left [ \left ( \frac{B}{A}\right )_{\rm out} - \left (\frac{B}{A} \right )_{\rm in} \right]
+    \end{equation}
+
+  \item update the mass fractions, $X_k$, using the values from the table
+    \end{itemize}
+
+  \item if we are not in NSE:
+
+    \begin{itemize}
+    \item integrate the aprox19 network as usual
+    \item update the aux quantities at the end of the burn
+    \end{itemize}
+  \end{itemize}
+\end{itemize}
+
+We use the following conditions to determine if we are in NSE:
+\begin{itemize}
+\item $\rho > \rho_\mathrm{nse}$
+\item $T > T_\mathrm{nse}$
+\item $X(\isotm{C}{12}) < A_\mathrm{nse}$
+\item $X(\isotm{He}{4}) + X(\isotm{Cr}{48}) + X(\isotm{Fe}{52}) + X(\isotm{Fe}{54}) + X(\isotm{Ni}{56}) > B_\mathrm{nse}$
+\end{itemize}
 
 \section{Simulations}
+
+\subsection{Reacting convergence test problem}
+
+\cite{castro_sdc} introduced a test problem for measuring convergence
+of a reacting hydrodynamic algorithm.  We run that same test here with the
+simplified SDC algorithm.
 
@@
 \section{Jacobian}
+
+\label{sec:app:jacobian}
 

### 2020-08-16 simplified_sdc_intro/paper.tex
update NSE flow

@@
 is in NSE.  The table provides:
-\begin{align*}
+\begin{align}
 Y_e &= \sum_k \frac{Z_k X_k}{A_k} \\
@@
 \frac{B}{A} &= \sum_k \frac{B_k X_k}{A_k}
-\end{align*}
+\end{align}
 where $B_k$ is the binding energy of nucleus $k$.  In our simulations,
@@
 and our evolution equations are:
-\begin{align*}
+\begin{align}
 \frac{DY_e}{Dt} &= \sum_k \frac{Z_k}{A_k} \frac{DX_k}{Dt} = 0 \\
@@
 \frac{D}{Dt} \left (\frac{B}{A} \right ) &= \sum_k \frac{B_k}{A_k} \frac{DX_k}{Dt} = 0
-\end{align*}
+\end{align}
 Therefore each of these auxillar equations obeys an advection
@@
 \begin{itemize}
-\item initialize the problem, including $X_k$
+\item {\em Initialize the problem}
 
-\item fill the initial aux data with $Y_e$, $\bar{A}$, and $(B/A)$.
+  In addition to initializing the mass fractions, $X_k$, we also fill
+  the auxiliary composition data, $Y_e$, $\bar{A}$, and $(B/A)$, using
+  these $X_k$.
 
-\item in hydro, update these quantities simply via advection
+\item {\em Hydrodynamics update}
 
-\item for the reactive update:
+  The hydrodynamics update proceeds as normal, but with an advection
+  equation for each of the auxiliary composition variables:
+  \begin{align}
+    \ddt{(\rho Y_e)} + \ddx{(\rho Y_e u)} &= 0 \\
+    \ddt{(\rho \bar{A})} + \ddx{(\rho \bar{A} u)} &= 0 \\
+    \ddt{[\rho (B/A)]} + \ddx{[\rho (B/A) u]} &= 0 
+  \end{align}
+
+\item {\em Reactive update}
 
   \begin{itemize}
-  \item check if NSE applies (see below)
-
-  \item if we are in an NSE region:
+  \item For a zone that is in NSE:
 
@@
     \item use $\rho$, $T$, and $Y_e$ to call the table. This
-      returns: $dY_e/dt$, $(B/A)_{\rm out}$, and $\bar{A}_{\rm out}$.
+      returns: $dY_e/dt$, $(B/A)_{\rm out}$, $\bar{A}_{\rm out}$,
+      and the $(X_k)_{\rm out}$ from the NSE network grouped into the 19 isotopes
+      we carry as part of aprox19.
 
@@
     \begin{equation}
-      (Y_e)_{\rm out} = (Y_e)_{\rm in} - \Delta t \frac{dY_e}{dt}
+      (Y_e)_{\rm out} = (Y_e)_{\rm in} + \Delta t \frac{dY_e}{dt}
     \end{equation}
@@
   \item update the auxiliary state value of $\bar{A}$ to be simply
-    $\bar{A}_{\rm out}$ from the table.
+    $\bar{A}_{\rm out}$ from the table and the mass fractions
+    $X_k$ to be the $(X_k)_{\rm out}$ returned from the table.
 
@@
 
-  \item update the mass fractions, $X_k$, using the values from the table
     \end{itemize}
 
-  \item if we are not in NSE:
+  \item For zones not in NSE:
 
@@
 We use the following conditions to determine if we are in NSE:
-\begin{itemize}
-\item $\rho > \rho_\mathrm{nse}$
-\item $T > T_\mathrm{nse}$
-\item $X(\isotm{C}{12}) < A_\mathrm{nse}$
-\item $X(\isotm{He}{4}) + X(\isotm{Cr}{48}) + X(\isotm{Fe}{52}) + X(\isotm{Fe}{54}) + X(\isotm{Ni}{56}) > B_\mathrm{nse}$
-\end{itemize}
+\begin{align}
+\rho &> \rho_\mathrm{nse} \\
+T &> T_\mathrm{nse} \\
+X(\isotm{C}{12}) &< A_\mathrm{nse} \\
+X(\isotm{He}{4}) + \sum_{k \in {\rm Fe-group}} X_k &> B_\mathrm{nse} 
+\end{align}
+where the Fe-group nuclei are \isot{Cr}{48}, \isot{Fe}{52}, \isot{Fe}{54}, and \isot{Ni}{56}.
 

### 2020-08-18 simplified_sdc_intro/paper.tex
some tweaks

@@
 \label{eq:dt:nuce}
-\Delta t &\le \mathtt{dtnuc\_e}\, \min_{i,j,k} \left\{\frac{e_{i,j,k}}{\dot{e}_{i,j,k}}\right\} \\
+\Delta t &\le f_e\, \min_{i} \left\{\frac{e_{i}}{\dot{e}_{i}}\right\} \\
 %
 \label{eq:dt:nucX}
-\Delta t &\le \mathtt{dtnuc\_X}\, \min_{i,j,k} \left\{\min_n\frac{X^n_{i,j,k}}{\dot{X}^n_{i,j,k}}\right\}
+\Delta t &\le f_X\, \min_{i} \left\{\min_{k,X_k > \epsilon_X}\frac{{X_k}_{i}}{{\omegadot_{k,i}}}\right\}
 \end{align}
-where $\mathtt{dtnuc\_e}$ and $\mathtt{dtnuc\_X}$ are runtime parameters
-used to control the allowed change.
+where $i$ is the zone index and $f_e$ and $f_X$ are runtime parameters
+used to control the allowed change, and only species for which $X_k > \epsilon_X$ are considered.
 
@@
     \item construct the advective update terms, first ignoring the hydrodynamics sources:
-      $\Advt{\Uc}^{n+1/2,(k)}_{i,j,k}$, e.g., as:
+      $\Advt{\Uc}^{n+1/2,(k)}_{i}$, e.g., as:
       \begin{align}
-        \Advt{\Uc}^{n+1/2,(k)}_{i,j,k} =
+        \Advt{\Uc}^{n+1/2,(k)}_{i} =
           &- \frac{\Fb^\xv(\Uc^{n+1/2,(k)}_{i+1/2}) - \Fb^\xv(\Uc^{n+1/2,(k)}_{i-1/2})}{\Delta x}
@@
     \begin{equation}
-      \Adv{\Uc}^{n+1/2,(k)}_{i,j,k} = \Advt{\Uc}^{n+1/2,(k)}_{i,j,k} + \Shydro^{n+1/2}
+      \Adv{\Uc}^{n+1/2,(k)}_{i} = \Advt{\Uc}^{n+1/2,(k)}_{i} + \Shydro^{n+1/2}
     \end{equation}
@@
 
-\section{Nuclear Statistical Equilibrium}
+\section{Reaction Networks and Nuclear Statistical Equilibrium}
 
-For this study we will use the 19 isotope ``aprox19'' network together
-with a table that gives the NSE abundances in regions where the system
-is in NSE.  The table provides:
+For this study we will use the 19 nuceli network containing
+\isot{H}{1}, \isot{He}{3}, \isot{He}{4}, \isot{C}{12}, \isot{N}{14},
+\isot{O}{16}, \isot{Ne}{20}, \isot{Mg}{24}, \isot{Si}{28},
+\isot{S}{32}, \isot{Ar}{36}, \isot{Ca}{40}, \isot{Ti}{44},
+\isot{Cr}{48}, \isot{Fe}{52}, \isot{Fe}{54}, \isot{Ni}{56}, protons
+(from photodisintegration), and neutrons.  This is based on the
+``aprox19'' network from \cite{aprox19} and originally described in
+\cite{Kepler}.  We combine this with a table that gives the nuclear
+statistical equilibrium (NSE) abundances in regions where the system
+is in NSE.  The NSE table was generated using a 127 nuclei reaction
+network and is the same as described in \cite{ma:2013}.  In our
+simulations, we carry all 19 isotopes in the main network in each zone
+and advect them in the hydrodynamics portion of the algorithm.  The
+composition of the larger 127 nuclei network is mapped into the 19
+isotopes we carry according to table~\ref{table:nuclei}.
+
+The NSE table provides:
 \begin{align}
@@
 
-The NSE table was generated using a 125 nuclei reaction network and is
-the same as described in \cite{ma:2013}.  The compositional quantities
+ The compositional quantities
 it carries, $\bar{A}$ and $Y_e$ and not representable from the 19
@@
 simplified SDC algorithm.
+
+\subsection{Detonation}
+
+The purpose here is to look at what timestep is taken when we use the nuclear burning limiters.
+Run with aprox19 only, Strang and SDC, then aprox19 + NSE, Strang and NSE.
+
+\subsection{Massive Star}
+
+Play around with the threshold where NSE kicks in
 

### 2020-08-20 xrb2/paper.tex
add some figure descriptions

@@
 \plotone{time_series_500}
-\caption{Time series of the 500 Hz run showing $\bar{A}$.}
+\caption{\label{fig:time_series_500} Time series of the 500 Hz run showing $\bar{A}$.}
 \end{figure}
@@
 \plotone{time_series_1000}
-\caption{Time series of the 1000 Hz run showing $\bar{A}$.}
+\caption{\label{fig:time_series_1000} Time series of the 1000 Hz run showing $\bar{A}$.}
 \end{figure}
+
+Figures~\ref{fig:time_series_500} and \ref{fig:time_series_1000} show
+time series of the mean molecular weight for the 500~Hz and 1000~Hz
+runs.  Compared to those in \cite{flame_wave1}, ashes behind the flame
+do not reach as high atomic weights.  This is not surprising, since
+those early runs boosted the reaction rates.  Comparing these two new
+runs, we see that the burning is much more evolved for the higher
+rotation rate.  We believe that this is because the rotation confines
+the initial perturbation and subsequent expansion from the burning
+better, allowing the reactions to progress further.
 
@@
 
-different initial models?
-
 different rotation rates
 
-look at flame speed, shape of flame, amount of mixing (e.g. it looks like slower rotation rates have less mixed flames as the hurricane is less strong.)
+shape of flame, amount of mixing (e.g. it looks like slower rotation rates have less mixed flames as the hurricane is less strong.)
+
+Figure~\ref{fig:flame_speeds} shows the position of the flame as a
+function of time.  Defining the position is difficult, so we \MarginPar{need explanation here}
 
@@
 	\plottwo{speeds_1_100th_max.pdf}{speeds_1_1000th_max.pdf}
-	\caption{Plots of flame front position vs.\ time for the 500 Hz and 1000 Hz runs.
+	\caption{\label{fig:flame_speeds} Plots of flame front position vs.\ time for the 500 Hz and 1000 Hz runs.
 The left-hand plot shows the results when tracking the $\frac{1}{100} \max[\enucdot]$ contour, while the right-hand plot shows the results for the $\frac{1}{1000} \max[\enucdot]$ contour.}

### 2020-08-24 xrb2/paper.tex
more notes

@@
 
-lateral profiles of X vs distance behind the flame
-
+MZ: lateral profiles of X vs distance behind the flame; 
 vertical profiles of the atmosphere behind the flame---how does it compare to isentropic?
@@
 shape of flame, amount of mixing (e.g. it looks like slower rotation rates have less mixed flames as the hurricane is less strong.)
+
+Don: look at ratio of enuc to e?
 

### 2020-08-25 xrb2/paper.tex
add some references

@@
 
-Insert some stuff here about previous work on rotation
+Previous studies have found that rotation can have a significant affect on the flame's propagation.
+As the rotation rate increases, the Coriolis force whips the spreading flame up into a hurricane-like structure \citep{spitkovsky2002,cavecchi:2013}.
+The stronger Coriolis force leads to greater confinement of the hot accreted matter, leading to easier ignition of the flame \citep{art-2015-cavecchi-etal}.
+
+Previous work has predicted that crustal heating is stronger a lower accretion rates and weaker at higher accretion rates, due to the effect of neutrino losses \citep{Cumming2006,johnston:2019}.
+In our models, we keep the crust at a constant temperature, so by increasing this temperature we can effectively increase the crustal heating and mimic the effects of lower accretion rates.
+A shallow heating mechanism of as yet unknown origin has been found necessary to reproduce observed properties of XRBs \citep{Deibel2015,Turlione2015,Keek2017}.
 
@@
           GCC (\url{https://gcc.gnu.org/}),
+          Jupyter \citep{Kluyver2016},
           linux (\url{https://www.kernel.org/}),

### 2020-09-26 simplified_sdc_intro/paper.tex
update
more notes
updates

@@
 
-For Strang split coupling of hydro and reactions, $DX_k/Dt = 0$,
 and our evolution equations are:
 \begin{align}
-\frac{DY_e}{Dt} &= \sum_k \frac{Z_k}{A_k} \frac{DX_k}{Dt} = 0 \\
-\frac{D}{Dt} \frac{1}{\bar{A}} &= - \frac{1}{\bar{A}^2} \frac{D\bar{A}}{Dt} = \sum_k \frac{1}{A_k} \frac{DX_k}{Dt} = 0 \rightarrow \frac{D\bar{A}}{Dt} = 0 \\
-\frac{D}{Dt} \left (\frac{B}{A} \right ) &= \sum_k \frac{B_k}{A_k} \frac{DX_k}{Dt} = 0
+\frac{DY_e}{Dt} &= \sum_k \frac{Z_k}{A_k} \frac{DX_k}{Dt} = \sum_k \frac{Z_k}{A_k} \omegadot_k \\
+\frac{D\bar{A}}{Dt} &= -\bar{A}^2 \sum_k \frac{1}{A_k} \frac{DX_k}{Dt} = -\bar{A}^2 \sum_k \frac{1}{A_k} \omegadot_k \\
+\frac{D}{Dt} \left (\frac{B}{A} \right ) &= \sum_k \frac{B_k}{A_k} \frac{DX_k}{Dt} = \sum_k \frac{B_k}{A_k} \omegadot_k
 \end{align}
-Therefore each of these auxillar equations obeys an advection
+For Strang split coupling of hydro and reactions, $DX_k/Dt = 0$,
+Therefore each of these auxillary equations obeys an advection
 equation in the hydro part of the advancement.  In the SDC algorithm,
@@
 
-  In addition to initializing the mass fractions, $X_k$, we also fill
-  the auxiliary composition data, $Y_e$, $\bar{A}$, and $(B/A)$, using
-  these $X_k$.
+  We initializing the mass fractions, $X_k$, and then compute the
+  electron fraction, $Y_e$, from them.  For the two other composition
+  quantities we carry, $\bar{A}$, and $(B/A)$, we need values that are
+  consistent with the value of $Y_e$ and the nuclei stored in the NSE
+  table.  Therefore, if the thermodynamic conditions put us in NSE
+  (using the conditions defined below), then we get $\bar{A}$ and
+  $(B/A)$ from the NSE table.  Otherwise, we compute these directly
+  from $X_k$.  $X_k$.
 
@@
 
+Also explore: PLM vs PPM, 1 vs. 2 vs. 3 iterations
+
+\subsection{NSE convergence test problem}
+
+To test the coupling of the NSE table to the hydrodynamics, we run a
+similar problem as above, except with the thermodynamic conditions
+appropriate for the matter to be in NSE.
+
 \subsection{Detonation}
@@
 Run with aprox19 only, Strang and SDC, then aprox19 + NSE, Strang and NSE.
+
+Look at number of SDC iterations too.
+
+How does the timestep limiter work in the middle of SDC iterating?
+
+aprox21
+
+look at number of RHS calls, wallclock time
+
+output every step and then for a single zone plot enuc every timestep to see if it oscillates
 

### 2020-10-07 simplified_sdc_intro/paper.tex
add ECP ack

--- 2020-08-18 simplified_sdc_intro/paper.tex
@@
 
+Also explore: PLM vs PPM, 1 vs. 2 vs. 3 iterations
+
 \subsection{Detonation}
@@
 U.S. Department of Energy, Office of Science, Office of Advanced
-Scientific Computing Research and Office of Nuclear Physics, Scientific
-Discovery through Advanced Computing (SciDAC) program under Award
-Number DE-SC0017955.
+Scientific Computing Research and Office of Nuclear Physics,
+Scientific Discovery through Advanced Computing (SciDAC) program under
+Award Number DE-SC0017955.  This research was supported by the
+Exascale Computing Project (17-SC-20-SC), a collaborative effort of
+the U.S. Department of Energy Office of Science and the National
+Nuclear Security Administration.
 

### 2020-10-16 simplified_sdc_intro/paper.tex
some updates

--- 2020-09-26 simplified_sdc_intro/paper.tex
@@
 \newcommand{\divg}[1]{{\nablab \cdot \left (#1\right)}}
-\newcommand{\inp}{{\mathrm{input}}}
 \newcommand{\dedr}{\left . {\partial{}e}/{\partial\rho}\right |_{T, X_k}}
@@
 \newcommand{\avg}[1]{{\left \langle #1 \right \rangle}}
+
+\newcommand{\nse}[1]{{\mathrm{NSE}( #1 )}}
+
+\newcommand{\out}{{\rm out}}
+\newcommand{\inp}{{\rm in}}
 
@@
 
-The NSE table provides:
+The NSE table requires:
+\begin{equation}
+  Y_e = \sum_k \frac{Z_k X_k}{A_k}
+\end{equation}
+(where $A_k$ and $Z_k$ are the atomic weight and atomic number of nucleus $k$) and provides
 \begin{align}
-Y_e &= \sum_k \frac{Z_k X_k}{A_k} \\
 \bar{A} &= \left [ \sum_k \frac{X_k}{A_k} \right ]^{-1} \\
-\frac{B}{A} &= \sum_k \frac{B_k X_k}{A_k}
+\left (\frac{B}{A} \right ) &= \sum_k \frac{B_k X_k}{A_k}
 \end{align}
@@
 we store these 3 quantities as auxiliary data that is carried along
-with the rest of the fluid state in each zone.
+with the rest of the fluid state in each zone.  The table also returns
+values of the mass fractions mapped onto the 19-isotopes we carry, $X_k$, and the time-derivative of $Y_e$, $dY_e/dt$.
+We'll use the notation:
+\begin{equation}
+\nse{\rho,T,Y_e} \rightarrow \bar{A}, X_k, (B/A), dY_e/dt
+\end{equation}
+to represent the NSE table call and its inputs and outputs.
 
-and our evolution equations are:
+We can derive evolution equations for each of these composition quantities as:
 \begin{align}
@@
 there will be a reactive source (an $\Ic_q$) for each of these that is
-computed in the same manner as above.
+computed in the same manner as above.  We note that our NSE table
+provides the evolution of $Y_e$ due to reactions ($DY_e/Dt$) directly.
 
@@
 
-For Strang splitting, the basic flow of using the aprox19 + NSE network
-is as follows:
-\begin{itemize}
-\item {\em Initialize the problem}
+\subsection{Initialization}
 
@@
 (using the conditions defined below), then we get $\bar{A}$ and
-  $(B/A)$ from the NSE table.  Otherwise, we compute these directly
-  from $X_k$.  $X_k$.
+$(B/A)$ from the NSE table.  Otherwise, we compute these directly from
+$X_k$.
+
+\subsection{NSE condition}
+
+We use the following conditions to determine if we are in NSE:
+\begin{align}
+\rho &> \rho_\mathrm{nse} \\
+T &> T_\mathrm{nse} \\
+X(\isotm{C}{12}) &< A_\mathrm{nse} \\
+X(\isotm{He}{4}) + \sum_{k \in {\rm Fe-group}} X_k &> B_\mathrm{nse} 
+\end{align}
+where the Fe-group nuclei are \isot{Cr}{48}, \isot{Fe}{52}, \isot{Fe}{54}, and \isot{Ni}{56}.
+
+\subsection{Strang-split algorithm for NSE}
+
+For Strang splitting, the basic flow of using the aprox19 + NSE network
+is as follows:
+\begin{itemize}
 
@@
   \begin{itemize}
+
     \item For a zone that is in NSE:
 
+      The goal is to update the composition and thermodynamics due to the 
+      change in the nuclei abundances over the (half-)timestep.  There are
+      a few approaches we can do here.
+
       \begin{itemize}
-    \item use $\rho$, $T$, and $Y_e$ to call the table. This
-      returns: $dY_e/dt$, $(B/A)_{\rm out}$, $\bar{A}_{\rm out}$,
-      and the $(X_k)_{\rm out}$ from the NSE network grouped into the 19 isotopes
-      we carry as part of aprox19.
+      \item Method I: {\em Use the time-level $n$ information for the update}.
 
-    \item update $Y_e$:
+        This is the approach used in \citet{ma:2013}.  We call the table with the 
+        input state:
+        \begin{equation}
+          \nse{\rho, T, (Y_e)_{\rm in}} \rightarrow \bar{A}_\out, (X_k)_\out, (B/A)_\out, dY_e/dt
+        \end{equation}
+
+        We can now update $Y_e$:
         \begin{equation}
@@
         \end{equation}
-
-  \item update the auxiliary state value of $\bar{A}$ to be simply
-    $\bar{A}_{\rm out}$ from the table and the mass fractions
+        and take the auxiliary state values of $\bar{A}$ and $(B/A)$ to be simply
+        the values from the table, $\bar{A}_\out$ and $(B/A)_\out$, and the mass fractions
         $X_k$ to be the $(X_k)_{\rm out}$ returned from the table.
 
-  \item compute the energy generation rate, $\Sdot$ as:
+        Finally, we compute the energy generation rate, $\Sdot$ as:
         \begin{equation}
-      \Sdot = \eta \left [ \left ( \frac{B}{A} \right )_{\rm out} -
+          \label{eq:nse_energy}
+          \Sdot = \left [ \left ( \frac{B}{A} \right )_{\rm out} -
             \left ( \frac{B}{A} \right )_{\rm in} \right ] N_A \frac{1}{\Delta t}
         \end{equation}
-    where $\eta$ is an inertia term < 1 to prevent the energy changing
-    too much in one set.
 
-  \item set the auxiliary state value of $(B/A)$ as:
+        We note that \citet{ma:2013} modified Eq.~\ref{eq:nse_energy}
+        to take use only a fraction (0.3) of the energy change from
+        the table, to avoid a potential instability that can arise if
+        too much energy is deposited in a zone in a single timestep
+        (they called this an inertia term).  We prefer to deal with
+        the issue of too much energy being deposited in a zone in one
+        step through the retry mechanism already built into \castro.
+
+      \item Method II: {\em predictor-corrector.}
+
+        This starts out the same as the above method:
         \begin{equation}
-      \left (\frac{B}{A}\right )_{\rm out} = \left (\frac{B}{A}\right )_{\rm in} + \eta \left [ \left ( \frac{B}{A}\right )_{\rm out} - \left (\frac{B}{A} \right )_{\rm in} \right]
+          \nse{\rho, T, (Y_e)_{\rm in}} \rightarrow \bar{A}^\star, (X_k)^\star, (B/A)^\star, dY_e/dt
         \end{equation}
+        but we indicate with a $\star$ that most of those values are
+        provisional, and we will seek a better value in the corrector step.
+        We only update $Y_e$ from this call:
+        \begin{equation}
+          (Y_e)_{\rm out} = (Y_e)_{\rm in} + \Delta t \frac{dY_e}{dt}
+        \end{equation}
+        and then we use the table again, but with the updated $Y_e$:
+        \begin{equation}
+          \nse{\rho, T, (Y_e)_\out} \rightarrow \bar{A}_\out, (X_k)_\out, (B/A)_\out, dY_e/dt
+        \end{equation}
+        This corrects the composition so that it is consistent with the updated
+        $Y_e$.  We use these values $\bar{A}_\out$, $(X_k)_\out$, and $(B/A)_\out$ to then complete the update, computing the energy as in Eq.~\ref{eq:nse_energy}.
 
@@
 
-We use the following conditions to determine if we are in NSE:
-\begin{align}
-\rho &> \rho_\mathrm{nse} \\
-T &> T_\mathrm{nse} \\
-X(\isotm{C}{12}) &< A_\mathrm{nse} \\
-X(\isotm{He}{4}) + \sum_{k \in {\rm Fe-group}} X_k &> B_\mathrm{nse} 
-\end{align}
-where the Fe-group nuclei are \isot{Cr}{48}, \isot{Fe}{52}, \isot{Fe}{54}, and \isot{Ni}{56}.
+\subsection{SDC-NSE Coupling}
 

### 2020-10-17 simplified_sdc_intro/paper.tex
add some aux stuff

@@
            \rho X_k \\
+           \rho \alpha_l \\
            \rho u \\
@@
 \end{equation}
-are the conserved fluid quantities: mass density, $\rho$, velocity, $u$,
-specific total energy, $E$, and nuclear
-species mass fractions, $X_k$.  The specific total energy relates to
-the specific internal energy as $E = e + u^2/2$.  The
-corresponding fluxes are
+are the conserved fluid quantities: mass density, $\rho$, velocity,
+$u$, specific total energy, $E$, and nuclear species mass fractions,
+$X_k$, and auxiliary composition variables, $a_l$.  The mass fractions
+are constrained to sum to 1, $\sum_k X_k = 1$, but no such constraint
+exists on the auxiliary variables.  The specific total energy relates
+to the specific internal energy as $E = e + u^2/2$.  The corresponding
+fluxes are
 \begin{equation}
 \Fb^\xv (\Uc) = \left ( \begin{array}{c}
-   \rho u \\ \rho X_k u \\ \rho u u + p \\  \rho u E + u p
+   \rho u \\
+   \rho X_k u \\
+   \rho \alpha_l u \\
+   \rho u u + p \\
+   \rho u E + u p
    \end{array} \right )
 \end{equation}
-Here the pressure, $p$, enters, and is found via the equation of state,
+Here the pressure, $p$, enters, and is found via the equation of state.  For a system
+where the composition is completely specified by the mass fractions, $X_k$, the equation
+of state would take the form:
 \begin{equation}
 p = p(\rho, X_k, e)
+\end{equation}
+However, as we'll see shortly, when we use the NSE table, we are using
+the results of a much larger network then can be represented by the
+mass fractions, and in this case, we rely on the auxiliary composition
+variables, $\alpha_l$, to describe the state of the composition, and
+our EOS has the form:
+\begin{equation}
+p = p(\rho, \alpha_l, e)
 \end{equation}

### 2020-10-17 simplified_sdc_intro/paper.tex
some updates
more updates on SDC method

--- 2020-10-16 simplified_sdc_intro/paper.tex
@@
 \newcommand{\nse}[1]{{\mathrm{NSE}( #1 )}}
+
+\newcommand{\rhonse}{{\rho_\mathrm{nse}}}
+\newcommand{\tnse}{{T_\mathrm{nse}}}
+\newcommand{\Anse}{{A_\mathrm{nse}}}
+\newcommand{\Bnse}{{B_\mathrm{nse}}}
 
@@
 
+\label{sec:initialization}
+
   \begin{itemize}
@@
 
-We use the following conditions to determine if we are in NSE:
+We treat a zone as being in NSE if the density and temperature exceed
+some threshold and the composition is mainly $\alpha$ and Fe-group
+nuclei, where the Fe-group nuclei are \isot{Cr}{48}, \isot{Fe}{52},
+\isot{Fe}{54}, and \isot{Ni}{56}.  The full condition is:
 \begin{align}
-\rho &> \rho_\mathrm{nse} \\
-T &> T_\mathrm{nse} \\
-X(\isotm{C}{12}) &< A_\mathrm{nse} \\
-X(\isotm{He}{4}) + \sum_{k \in {\rm Fe-group}} X_k &> B_\mathrm{nse} 
+\rho &> \rhonse \\
+T &> \tnse \\
+X(\isotm{C}{12}) &< \Anse \\
+X(\isotm{He}{4}) + \sum_{k \in \mathrm{Fe-group}} X_k &> \Bnse
 \end{align}
-where the Fe-group nuclei are \isot{Cr}{48}, \isot{Fe}{52}, \isot{Fe}{54}, and \isot{Ni}{56}.
+Typical values are $\rhonse = 3\times 10^8~\gcc$, $T_\mathrm{nse} =
+3\times 10^9~\mathrm{K}$, and $\Anse = 0.12$ and $\Bnse=
+0.88$---these are the values used in \citet{ma:2013}.
 
@@
 
+With SDC evolution, when we are in NSE, we need to do the advective
+and reactive updates together.  With the tabulated NSE reaction
+values, we get instantaneous values of the reaction sources, but the only
+time derivative we have is for $Y_e$.  We will
+use a basic predictor-corrector update:
+\begin{equation}
+\Uc^\star = \Uc^n + \Delta t \Adv{\Uc}^{n+1/2} + \Delta t \Rb (\Uc^n)
+\end{equation}
+where $\Rb(\Uc^n)$ is the reaction source vector evaluated by calling the NSE
+table with the time-level $n$ state.
+In particular, this will also operate on the auxiliary composition variables,
+including $Y_e$:
+\begin{equation}
+(\rho Y_e)^\star = (\rho Y_e)^n + \Delta t \Adv{\rho Y_e}^{n+1/2} + \Delta t \rho (dY_e/dt)^n
+\end{equation}
+From this, we get $\rho^\star$, $(Y_e)^\star$, $(X_k)^\star$, and
+$e^\star$.  In order call the get the temperature, we need to call the
+EOS which requires $\bar{A}^\star$.  Since we cannot use the NSE table
+without the temperature, we need to
+
 \section{Simulations}
@@
 similar problem as above, except with the thermodynamic conditions
-appropriate for the matter to be in NSE.
+appropriate for the matter to be in NSE.  The initial conditions are:
+\begin{align}
+\rho &= \rho_0 \\
+T &= T_0 \left [ 1 + (\delta T) e^{-(r/\lambda)^2} \cos^6(\pi r/L) \right ] \\
+Y_e &= (Y_e)_0 \left [ 1 + (\delta Y_e) e^{-(r/\lambda)^2} \cos^6(\pi r/L) \right ] 
+\end{align}
+where $r$ is the distance from the center of the domain and the
+remaining parameters are listed in Table~\ref{table:nse}.  This has
+$Y_e$ varying between $[0.47, 0.5]$ initially\MarginPar{check}.  The mass fractions
+and remaining composition variables are then initialized from the NSE table, as described in section~\ref{sec:initialization}.
+
+The domain has a size $[0, L]^2$, and the timestep is fixed as:
+\begin{equation}
+\Delta t = 5\times 10^{-4} \left ( \frac{128}{N} \right )~\mathrm{s}
+\end{equation}
+where $N$ is the number of zones in each direction.
+
+\begin{deluxetable}{lcc}
+\tablecaption{\label{table:nse} NSE convergence test problem.}
+\tablehead{\colhead{parameter} & \colhead{value}}
+\startdata
+$\rho_0$ & $5\times 10^8~\gcc$ \\
+$T_0$    & $4\times 10^9~\mathrm{K}$ \\
+$(\delta T)$ & $0.2$ \\
+$(Y_e)_0$   & $0.5$ \\
+$(\delta Y_e)$ & $-0.05$ \\
+$\lambda$   & $2\times 10^7~\mathrm{cm}$ \\
+$L$         & $10^8~\mathrm{cm}$ \\
+\enddata
+\end{deluxetable}
 

### 2020-10-17 simplified_sdc_intro/paper.tex

@@
   \Rb(\Uc) = \left ( \begin{array}{c}
-     0 \\ \rho \omegadot_k \\ 0 \\ \rho \dot{S}
+     0 \\
+     \rho \omegadot(X_k) \\
+     \rho \omegadot(\alpha_l) \\
+     0 \\
+     \rho \dot{S}
   \end{array} \right )
@@
   X_k \\
+  \alpha_l \\
   u \\
@@
 \begin{equation}
-\Ab^\xv(\qb) = \left ( \begin{array}{ccccc}
-    u & 0 & \rho & 0 & 0 \\
-    0 & u & 0    & 0 & 0 \\
-    0 & 0 & u    & 1/\rho & 0 \\
-    0 & 0 & \Gamma_1 p & u & 0 \\
-    0 & 0 & \rho h & 0 & u
+\Ab^\xv(\qb) = \left ( \begin{array}{cccccc}
+    u & 0 & 0 & \rho & 0 & 0 \\
+    0 & u & 0 & 0    & 0 & 0 \\
+    0 & 0 & u & 0    & 0 & 0 \\
+    0 & 0 & 0 & u    & 1/\rho & 0 \\
+    0 & 0 & 0 & \Gamma_1 p & u & 0 \\
+    0 & 0 & 0 & \rho h & 0 & u
   \end{array} \right )

### 2020-10-19 simplified_sdc_intro/paper.tex
more updates

@@
 \end{equation}
-where $\omegadot_k$ is the creation rate for species $k$ and $\dot{S}$
+where $\omegadot(X_k)$ is the creation rate for species $k$, $\omegadot(\alpha_l)$ is the creation rate for
+auxiliary composition variable $l$, and $\dot{S}$
 is the energy generation rate per unit mass.
@@
 \begin{equation}
+\label{eq:prim_sources}
 \Rb(q) = \left ( \begin{array}{c}
      0 \\
-     \omegadot_k \\
+     \omegadot(X_k) \\
+     \omegadot(\alpha_l) \\
      0 \\
@@
 \odt{\rho} & = 0 \\
-\odt{(\rho X_k)} &= \rho \omegadot_k \\
+\odt{(\rho X_k)} &= \rho \omegadot(X_k) \\
+\odt{(\rho \alpha_l)} &= \rho \omegadot(\alpha_l) \\
 \odt{(\rho u)} &= 0 \\
@@
 \label{eq:dt:nuce}
-\Delta t &\le f_e\, \min_{i} \left\{\frac{e_{i}}{\dot{e}_{i}}\right\} \\
+\Delta t &\le f_e\, \min_{i} \left\{\frac{e_{i}}{\dot{S}_{i}}\right\} \\
 %
 \label{eq:dt:nucX}
-\Delta t &\le f_X\, \min_{i} \left\{\min_{k,X_k > \epsilon_X}\frac{{X_k}_{i}}{{\omegadot_{k,i}}}\right\}
+\Delta t &\le f_X\, \min_{i} \left\{\min_{k,X_k > \epsilon_X}\frac{{X_k}_{i}}{{\omegadot(X_k)_i}}\right\}
 \end{align}
@@
 guaranteed to be met throughout the simulation, because the step is
-rejected if they are violated.  The retry mechanism in \castro\ works
-with both the Strang and simplified-SDC integration scheme.
+rejected if they are violated.  This contrasts to similar approaches
+used in other codes where conditions \ref{eq:dt:nuce} and
+\ref{eq:dt:nucX} are used to restrict the next timestep size, but the
+update over the current step is kept, which already violated the
+constraints. \MarginPar{would be nice to have a reference here} The retry mechanism in \castro\ works with both the
+Strang and simplified-SDC integration scheme.
 
@@
     but we need the advective update for $\qb$, which we have not
-    constructed.  Additionally, we cannot simply use the equation of
+    constructed.  We note that $\Ic^{(k)}_\qb$ is an approximation to the integral of 
+    Eq.~\ref{eq:prim_sources} over the timestep.  Additionally, we cannot simply use the equation of
     state on $\Ic^{(k)}_{\Uc}$ since this is a time-derivative and

### 2020-10-20 simplified_sdc_intro/paper.tex
some fixes
first pass at NSE-SDC coupling

@@
 \newcommand{\Ic}{{\boldsymbol{\mathcal{I}}}}
+\newcommand{\Ics}{{\mathcal{I}}}
 \newcommand{\smax}{{s_\mathrm{max}}}
@@
   with a superscript (like $n$ or $n+1$), we'll use a second subscript
-  in parentheses to keep track of the iteration.
+  in parentheses to keep track of the iteration.  A single iteration
+  starts with $\Uc^n$ and results in the new time-level state for that
+  iteration, $\Uc^{n+1,(k)}$.
 
@@
       \Shydro^{n+1/2} = \left ( \begin{array}{c}
-                     0 \\ 0 \\
+                     0 \\ 0 \\ 0 \\
                     {\bf S}_{\rho\Ub}^{n+1/2} \cdot \ex \\
@@
     \begin{equation}
-      \odt{\Uc} = \Adv{\Uc}^{n+1,(k)} + \Rb(\Uc)
-    \end{equation}
+      \odt{\Uc} = \Adv{\Uc}^{n+1/2,(k)} + \Rb(\Uc)
+    \end{equation}
+    This integration begins with $\Uc^n$ and results in $\Uc^{n+1,(k)}$.
 
@@
     \end{align}
-
-    At the end of this integration, we have the state at the next iteration,
-    $\Uc^{n+1,(k+1)}$
 
@@
 \end{equation}
-where $\Rb(\Uc^n)$ is the reaction source vector evaluated by calling the NSE
-table with the time-level $n$ state.
-In particular, this will also operate on the auxiliary composition variables,
-including $Y_e$:
+Calling the NSE table on the starting state gives use:
+\begin{equation}
+\nse{\rho^n, T^n, (Y_e)^n} \rightarrow \bar{A}^n, (X_k)^n, (B/A)^n, (dY_e/dt)^n
+\end{equation}
+To get the reactive source, we need to use the iteratively-lagged information
+from $\Ic_q^{n+1/2,(k-1)}$:
+\begin{equation}
+\Rb(\Uc^n) \approx \left (\begin{array}{c}
+  0 \\
+  \rho^n \Ics_q(\omegadot(X_k)) \\
+  \rho^n \Ics_q(\omegadot(\alpha_l)) \\
+  0 \\
+  \Ics_q(\rho e) 
+\end{array}
+\right )
+\end{equation}
+Since we have $(dY_e/dt)^n$ from the table, we replace the source for $Y_e$ in the above
+$\Rb(\Uc^n)$ as:
+\begin{equation}
+R(\rho Y_e) = \rho (dY_e/dt)^n
+\end{equation}
+which gives an update for the electron fraction as:
 \begin{equation}
@@
 \end{equation}
-From this, we get $\rho^\star$, $(Y_e)^\star$, $(X_k)^\star$, and
-$e^\star$.  In order call the get the temperature, we need to call the
-EOS which requires $\bar{A}^\star$.  Since we cannot use the NSE table
-without the temperature, we need to
+From this prediction, we get $\rho^\star$, $\alpha_l^\star$, and
+$e^\star$, which allows us to call the EOS and get an updated
+temperature, $T^\star$.
+We can then again call the NSE table:
+\begin{equation}
+\nse{\rho^\star, T^\star, (Y_e)^\star} \rightarrow \bar{A}^{n+1,(k)}, (X_k)^{n+1,(k)}, (B/A)^{n+1,(k)}, (dY_e/dt)
+\end{equation}
+where we use the outputs to define the new values of the composition variables.
+Since the density and momenta don't have reactive sources, we just use the $\Uc^\star$ state
+as their new values:
+\begin{align}
+\rho^{n+1,(k)} &= \rho^\star \\
+(\rho u)^{n+1,(k)} &= (\rho u)^\star
+\end{align}
+We then compute the energy release as
+\begin{equation}
+       (\rho \Sdot)^{n+1/2} = \left [ \rho^{n+1,(k)} \left ( \frac{B}{A} \right )^{n+1,(k)} -
+                            \rho^n \left ( \frac{B}{A} \right )^n \right ] N_A \frac{1}{\Delta t}
+\end{equation}\MarginPar{is this right?}
+and then do the final update of the total energy:
+\begin{equation}
+(\rho E)^{n+1,(k)} = (\rho E)^n + \Delta t \Adv{\rho E}^{n+1/2} + (\rho \Sdot)^{n+1/2}
+\end{equation}
 

### 2020-10-22 simplified_sdc_intro/paper.tex
fix the source terms (I think)

@@
     Now the conservative hydrodynamics source terms are computed by first updating to the
-    new state with advection only\footnote{For a source like gravity, this update can be done first for $\rho$ and then define the new momentum source using $\rho^{\star\star}$ and likewise for energy}, as:
-    \begin{equation}
-      \Uc^{\star\star} = \Uc^n + \Delta t \Advt{\Uc}^{n+1/2,(k)}
+    new state with advection and the old-time source term applied for the full $\Delta t$ as:
+    \begin{equation}
+      \Uc^{\star\star} = \Uc^n + \Delta t \Advt{\Uc}^{n+1/2,(k)} + \Delta t \Hb(\Uc^n)
     \end{equation}
@@
     %% \end{equation}
-
-    The final advective update term is then:
-    \begin{equation}
-      \Adv{\Uc}^{n+1/2,(k)}_{i} = \Advt{\Uc}^{n+1/2,(k)}_{i} + \Shydro^{n+1/2}
-    \end{equation}
-    with
-    \begin{equation}
-      \Shydro^{n+1/2} = \left ( \begin{array}{c}
-                     0 \\ 0 \\ 0 \\
-                    {\bf S}_{\rho\Ub}^{n+1/2} \cdot \ex \\
-                    S_{\rho E}^{n+1/2} \end{array} \right )
-    \end{equation}
+    We then evaluate the source terms with $\Uc^{\star\star}$ and
+    correct the advective term so that we have a time-centered
+    source: \MarginPar{are we actually doing this correctly for SDC?}
+    The final advective update term is then\footnote{For a source like gravity, this update can be done first for $\rho$ and then define the new momentum source using $\rho^{\star\star}$ and likewise for energy}:
+    \begin{equation}
+      \Adv{\Uc}^{n+1/2,(k)}_{i} = \Advt{\Uc}^{n+1/2,(k)}_{i} +
+      \frac{\Delta t}{2} \left (\Shydro(\Uc^n) + \Shydro(\Uc^{\star\star}) \right )
+    \end{equation}
+
+    %% with
+    %% \begin{equation}
+    %%   \Shydro^{n+1/2} = \left ( \begin{array}{c}
+    %%                  0 \\ 0 \\ 0 \\
+    %%                 {\bf S}_{\rho\Ub}^{n+1/2} \cdot \ex \\
+    %%                 S_{\rho E}^{n+1/2} \end{array} \right )
+    %% \end{equation}
 

### 2020-10-22 simplified_sdc_intro/paper.tex
add some tables
fixed typos

--- 2020-10-19 simplified_sdc_intro/paper.tex
@@
 \newcommand{\divg}[1]{{\nablab \cdot \left (#1\right)}}
+\newcommand{\inp}{{\mathrm{input}}}
 \newcommand{\dedr}{\left . {\partial{}e}/{\partial\rho}\right |_{T, X_k}}
@@
 \newcommand{\avg}[1]{{\left \langle #1 \right \rangle}}
-
-\newcommand{\nse}[1]{{\mathrm{NSE}( #1 )}}
-
-\newcommand{\rhonse}{{\rho_\mathrm{nse}}}
-\newcommand{\tnse}{{T_\mathrm{nse}}}
-\newcommand{\Anse}{{A_\mathrm{nse}}}
-\newcommand{\Bnse}{{B_\mathrm{nse}}}
-
-\newcommand{\out}{{\rm out}}
-\newcommand{\inp}{{\rm in}}
 
@@
 
-A particular difficult phase of evolution to model is the nuclear
-statistic equilibrium that sets in with helium burning for
+A particularly difficult phase of evolution to model is the nuclear
+statistical equilibrium that sets in with helium burning for
 temperatures in excess of $4\times 10^9~\mathrm{K}$.  Physically, the
-forward and reverse rates should balance leading to an equilibrium.
+forward and reverse rates of reaction should balance leading to an equilibrium.
 With operator splitting, an NSE region will have a large positive flow
@@
 reaction networks based on spectral deferred corrections (SDC)
-methods, and demonstrated the method of a variety of test problems.
+methods, and demonstrated the method in a variety of test problems.
 The approach presented here is considerably simpler, reusing the main
@@
 experiments, and all of the code to reproduce the results in this
-paper are in the \castro\ github
+paper is in the \castro\ github
 repository\footnote{\url{https://github.com/amrex-astro/Castro/}}.
@@
            \rho X_k \\
-           \rho \alpha_l \\
            \rho u \\
@@
 \end{equation}
-are the conserved fluid quantities: mass density, $\rho$, velocity,
-$u$, specific total energy, $E$, and nuclear species mass fractions,
-$X_k$, and auxiliary composition variables, $a_l$.  The mass fractions
-are constrained to sum to 1, $\sum_k X_k = 1$, but no such constraint
-exists on the auxiliary variables.  The specific total energy relates
-to the specific internal energy as $E = e + u^2/2$.  The corresponding
-fluxes are
+are the conserved fluid quantities: mass density, $\rho$, velocity, $u$,
+specific total energy, $E$, and nuclear
+species mass fractions, $X_k$.  The specific total energy relates to
+the specific internal energy as $E = e + u^2/2$.  The
+corresponding fluxes are
 \begin{equation}
 \Fb^\xv (\Uc) = \left ( \begin{array}{c}
-   \rho u \\
-   \rho X_k u \\
-   \rho \alpha_l u \\
-   \rho u u + p \\
-   \rho u E + u p
+   \rho u \\ \rho X_k u \\ \rho u u + p \\  \rho u E + u p
    \end{array} \right )
 \end{equation}
-Here the pressure, $p$, enters, and is found via the equation of state.  For a system
-where the composition is completely specified by the mass fractions, $X_k$, the equation
-of state would take the form:
+Here the pressure, $p$, enters, and is found via the equation of state,
 \begin{equation}
 p = p(\rho, X_k, e)
-\end{equation}
-However, as we'll see shortly, when we use the NSE table, we are using
-the results of a much larger network then can be represented by the
-mass fractions, and in this case, we rely on the auxiliary composition
-variables, $\alpha_l$, to describe the state of the composition, and
-our EOS has the form:
-\begin{equation}
-p = p(\rho, \alpha_l, e)
 \end{equation}
@@
   \Rb(\Uc) = \left ( \begin{array}{c}
-     0 \\
-     \rho \omegadot(X_k) \\
-     \rho \omegadot(\alpha_l) \\
-     0 \\
-     \rho \dot{S}
+     0 \\ \rho \omegadot_k \\ 0 \\ \rho \dot{S}
   \end{array} \right )
 \end{equation}
-where $\omegadot(X_k)$ is the creation rate for species $k$, $\omegadot(\alpha_l)$ is the creation rate for
-auxiliary composition variable $l$, and $\dot{S}$
+where $\omegadot_k$ is the creation rate for species $k$ and $\dot{S}$
 is the energy generation rate per unit mass.
@@
   X_k \\
-  \alpha_l \\
   u \\
@@
 \begin{equation}
-\Ab^\xv(\qb) = \left ( \begin{array}{cccccc}
-    u & 0 & 0 & \rho & 0 & 0 \\
-    0 & u & 0 & 0    & 0 & 0 \\
-    0 & 0 & u & 0    & 0 & 0 \\
-    0 & 0 & 0 & u    & 1/\rho & 0 \\
-    0 & 0 & 0 & \Gamma_1 p & u & 0 \\
-    0 & 0 & 0 & \rho h & 0 & u
+\Ab^\xv(\qb) = \left ( \begin{array}{ccccc}
+    u & 0 & \rho & 0 & 0 \\
+    0 & u & 0    & 0 & 0 \\
+    0 & 0 & u    & 1/\rho & 0 \\
+    0 & 0 & \Gamma_1 p & u & 0 \\
+    0 & 0 & \rho h & 0 & u
   \end{array} \right )
@@
 \begin{equation}
-\label{eq:prim_sources}
 \Rb(q) = \left ( \begin{array}{c}
      0 \\
-     \omegadot(X_k) \\
-     \omegadot(\alpha_l) \\
+     \omegadot_k \\
      0 \\
@@
 We use the \castro\ hydrodynamics code~\cite{castro}, together with
-the corner transport upwind piecewise parabolic method
+corner transport upwind piecewise parabolic method
 hydrodynamics~\citep{millercolella:2002}.  This is a finite-volume
@@
 \odt{\rho} & = 0 \\
-\odt{(\rho X_k)} &= \rho \omegadot(X_k) \\
-\odt{(\rho \alpha_l)} &= \rho \omegadot(\alpha_l) \\
+\odt{(\rho X_k)} &= \rho \omegadot_k \\
 \odt{(\rho u)} &= 0 \\
@@
 
-Normally we want to integrate temperature temperature.  We can write the
+Normally we want to integrate the temperature.  We can write the
 energy equation as:
@@
 \label{eq:dt:nuce}
-\Delta t &\le f_e\, \min_{i} \left\{\frac{e_{i}}{\dot{S}_{i}}\right\} \\
+\Delta t &\le f_e\, \min_{i} \left\{\frac{e_{i}}{\dot{e}_{i}}\right\} \\
 %
 \label{eq:dt:nucX}
-\Delta t &\le f_X\, \min_{i} \left\{\min_{k,X_k > \epsilon_X}\frac{{X_k}_{i}}{{\omegadot(X_k)_i}}\right\}
+\Delta t &\le f_X\, \min_{i} \left\{\min_{k,X_k > \epsilon_X}\frac{{X_k}_{i}}{{\omegadot_{k,i}}}\right\}
 \end{align}
@@
 required timestep).  Among the conditions that can trigger this are
-density failing below zero during advection, the ODE integration
+density falling below zero during advection, the ODE integration
 failing to converge in the implicit solve, or violation of one of the
@@
 guaranteed to be met throughout the simulation, because the step is
-rejected if they are violated.  This contrasts to similar approaches
-used in other codes where conditions \ref{eq:dt:nuce} and
-\ref{eq:dt:nucX} are used to restrict the next timestep size, but the
-update over the current step is kept, which already violated the
-constraints. \MarginPar{would be nice to have a reference here} The retry mechanism in \castro\ works with both the
-Strang and simplified-SDC integration scheme.
+rejected if they are violated.  The retry mechanism in \castro\ works
+with both the Strang and simplified-SDC integration scheme.
 
@@
 version of this was also implemented in the \maestro\ low-Mach number
-hydrodynamics code some time ago.  A compressible version, for use in
+hydrodynamics code some time ago.  The compressible version for use in
 \castro\ is slightly different due to the need to do some operations
@@
 
-\label{sec:initialization}
-
   \begin{itemize}
-  \item We need an approximation to how much the reactions alone
+  \item We need an approximation of how much the reactions alone
     changed the primitive variable state over the timestep, which we
@@
     but we need the advective update for $\qb$, which we have not
-    constructed.  We note that $\Ic^{(k)}_\qb$ is an approximation to the integral of 
-    Eq.~\ref{eq:prim_sources} over the timestep.  Additionally, we cannot simply use the equation of
+    constructed.  Additionally, we cannot simply use the equation of
     state on $\Ic^{(k)}_{\Uc}$ since this is a time-derivative and
@@
 
-For this study we will use the 19 nuceli network containing
+For this study we will use the 19 nuclei network containing
 \isot{H}{1}, \isot{He}{3}, \isot{He}{4}, \isot{C}{12}, \isot{N}{14},
@@
 
-The NSE table requires:
-\begin{equation}
-  Y_e = \sum_k \frac{Z_k X_k}{A_k}
-\end{equation}
-(where $A_k$ and $Z_k$ are the atomic weight and atomic number of nucleus $k$) and provides
+The NSE table provides:
 \begin{align}
+Y_e &= \sum_k \frac{Z_k X_k}{A_k} \\
 \bar{A} &= \left [ \sum_k \frac{X_k}{A_k} \right ]^{-1} \\
-\left (\frac{B}{A} \right ) &= \sum_k \frac{B_k X_k}{A_k}
+\frac{B}{A} &= \sum_k \frac{B_k X_k}{A_k}
 \end{align}
@@
 we store these 3 quantities as auxiliary data that is carried along
-with the rest of the fluid state in each zone.  The table also returns
-values of the mass fractions mapped onto the 19-isotopes we carry, $X_k$, and the time-derivative of $Y_e$, $dY_e/dt$.
-We'll use the notation:
-\begin{equation}
-\nse{\rho,T,Y_e} \rightarrow \bar{A}, X_k, (B/A), dY_e/dt
-\end{equation}
-to represent the NSE table call and its inputs and outputs.
+with the rest of the fluid state in each zone.
 
-We can derive evolution equations for each of these composition quantities as:
+For Strang split coupling of hydro and reactions, $DX_k/Dt = 0$,
+and our evolution equations are:
 \begin{align}
-\frac{DY_e}{Dt} &= \sum_k \frac{Z_k}{A_k} \frac{DX_k}{Dt} = \sum_k \frac{Z_k}{A_k} \omegadot_k \\
-\frac{D\bar{A}}{Dt} &= -\bar{A}^2 \sum_k \frac{1}{A_k} \frac{DX_k}{Dt} = -\bar{A}^2 \sum_k \frac{1}{A_k} \omegadot_k \\
-\frac{D}{Dt} \left (\frac{B}{A} \right ) &= \sum_k \frac{B_k}{A_k} \frac{DX_k}{Dt} = \sum_k \frac{B_k}{A_k} \omegadot_k
+\frac{DY_e}{Dt} &= \sum_k \frac{Z_k}{A_k} \frac{DX_k}{Dt} = 0 \\
+\frac{D}{Dt} \frac{1}{\bar{A}} &= - \frac{1}{\bar{A}^2} \frac{D\bar{A}}{Dt} = \sum_k \frac{1}{A_k} \frac{DX_k}{Dt} = 0 \rightarrow \frac{D\bar{A}}{Dt} = 0 \\
+\frac{D}{Dt} \left (\frac{B}{A} \right ) &= \sum_k \frac{B_k}{A_k} \frac{DX_k}{Dt} = 0
 \end{align}
-For Strang split coupling of hydro and reactions, $DX_k/Dt = 0$,
-Therefore each of these auxillary equations obeys an advection
+Therefore each of these auxiliary equations obeys an advection
 equation in the hydro part of the advancement.  In the SDC algorithm,
 there will be a reactive source (an $\Ic_q$) for each of these that is
-computed in the same manner as above.  We note that our NSE table
-provides the evolution of $Y_e$ due to reactions ($DY_e/Dt$) directly.
+computed in the same manner as above.
 
  The compositional quantities
-it carries, $\bar{A}$ and $Y_e$ and not representable from the 19
+it carries, $\bar{A}$ and $Y_e$, are not representable from the 19
 isotopes we carry in the main network. For this reason, when we are
@@
 
-\subsection{Initialization}
-
-We initializing the mass fractions, $X_k$, and then compute the
-electron fraction, $Y_e$, from them.  For the two other composition
-quantities we carry, $\bar{A}$, and $(B/A)$, we need values that are
-consistent with the value of $Y_e$ and the nuclei stored in the NSE
-table.  Therefore, if the thermodynamic conditions put us in NSE
-(using the conditions defined below), then we get $\bar{A}$ and
-$(B/A)$ from the NSE table.  Otherwise, we compute these directly from
-$X_k$.
-
-\subsection{NSE condition}
-
-We treat a zone as being in NSE if the density and temperature exceed
-some threshold and the composition is mainly $\alpha$ and Fe-group
-nuclei, where the Fe-group nuclei are \isot{Cr}{48}, \isot{Fe}{52},
-\isot{Fe}{54}, and \isot{Ni}{56}.  The full condition is:
-\begin{align}
-\rho &> \rhonse \\
-T &> \tnse \\
-X(\isotm{C}{12}) &< \Anse \\
-X(\isotm{He}{4}) + \sum_{k \in \mathrm{Fe-group}} X_k &> \Bnse
-\end{align}
-Typical values are $\rhonse = 3\times 10^8~\gcc$, $T_\mathrm{nse} =
-3\times 10^9~\mathrm{K}$, and $\Anse = 0.12$ and $\Bnse=
-0.88$---these are the values used in \citet{ma:2013}.
-
-\subsection{Strang-split algorithm for NSE}
-
 For Strang splitting, the basic flow of using the aprox19 + NSE network
@@
 \begin{itemize}
+\item {\em Initialize the problem}
+
+  In addition to initializing the mass fractions, $X_k$, we also fill
+  the auxiliary composition data, $Y_e$, $\bar{A}$, and $(B/A)$, using
+  these $X_k$.
 
@@
   \begin{itemize}
-
   \item For a zone that is in NSE:
 
-      The goal is to update the composition and thermodynamics due to the 
-      change in the nuclei abundances over the (half-)timestep.  There are
-      a few approaches we can do here.
+    \begin{itemize}
+    \item use $\rho$, $T$, and $Y_e$ to call the table. This
+      returns: $dY_e/dt$, $(B/A)_{\rm out}$, $\bar{A}_{\rm out}$,
+      and the $(X_k)_{\rm out}$ from the NSE network grouped into the 19 isotopes
+      we carry as part of aprox19.
 
-      \begin{itemize}
-      \item Method I: {\em Use the time-level $n$ information for the update}.
-
-        This is the approach used in \citet{ma:2013}.  We call the table with the 
-        input state:
-        \begin{equation}
-          \nse{\rho, T, (Y_e)_{\rm in}} \rightarrow \bar{A}_\out, (X_k)_\out, (B/A)_\out, dY_e/dt
-        \end{equation}
-
-        We can now update $Y_e$:
+    \item update $Y_e$:
     \begin{equation}
@@
     \end{equation}
-        and take the auxiliary state values of $\bar{A}$ and $(B/A)$ to be simply
-        the values from the table, $\bar{A}_\out$ and $(B/A)_\out$, and the mass fractions
+
+  \item update the auxiliary state value of $\bar{A}$ to be simply
+    $\bar{A}_{\rm out}$ from the table and the mass fractions
     $X_k$ to be the $(X_k)_{\rm out}$ returned from the table.
 
-        Finally, we compute the energy generation rate, $\Sdot$ as:
+  \item compute the energy generation rate, $\Sdot$ as:
     \begin{equation}
-          \label{eq:nse_energy}
-          \Sdot = \left [ \left ( \frac{B}{A} \right )_{\rm out} -
+      \Sdot = \eta \left [ \left ( \frac{B}{A} \right )_{\rm out} -
         \left ( \frac{B}{A} \right )_{\rm in} \right ] N_A \frac{1}{\Delta t}
     \end{equation}
+    where $\eta$ is an inertia term < 1 to prevent the energy changing
+    too much in one set.
 
-        We note that \citet{ma:2013} modified Eq.~\ref{eq:nse_energy}
-        to take use only a fraction (0.3) of the energy change from
-        the table, to avoid a potential instability that can arise if
-        too much energy is deposited in a zone in a single timestep
-        (they called this an inertia term).  We prefer to deal with
-        the issue of too much energy being deposited in a zone in one
-        step through the retry mechanism already built into \castro.
-
-      \item Method II: {\em predictor-corrector.}
-
-        This starts out the same as the above method:
+  \item set the auxiliary state value of $(B/A)$ as:
     \begin{equation}
-          \nse{\rho, T, (Y_e)_{\rm in}} \rightarrow \bar{A}^\star, (X_k)^\star, (B/A)^\star, dY_e/dt
+      \left (\frac{B}{A}\right )_{\rm out} = \left (\frac{B}{A}\right )_{\rm in} + \eta \left [ \left ( \frac{B}{A}\right )_{\rm out} - \left (\frac{B}{A} \right )_{\rm in} \right]
     \end{equation}
-        but we indicate with a $\star$ that most of those values are
-        provisional, and we will seek a better value in the corrector step.
-        We only update $Y_e$ from this call:
-        \begin{equation}
-          (Y_e)_{\rm out} = (Y_e)_{\rm in} + \Delta t \frac{dY_e}{dt}
-        \end{equation}
-        and then we use the table again, but with the updated $Y_e$:
-        \begin{equation}
-          \nse{\rho, T, (Y_e)_\out} \rightarrow \bar{A}_\out, (X_k)_\out, (B/A)_\out, dY_e/dt
-        \end{equation}
-        This corrects the composition so that it is consistent with the updated
-        $Y_e$.  We use these values $\bar{A}_\out$, $(X_k)_\out$, and $(B/A)_\out$ to then complete the update, computing the energy as in Eq.~\ref{eq:nse_energy}.
 
@@
 
-\subsection{SDC-NSE Coupling}
-
-With SDC evolution, when we are in NSE, we need to do the advective
-and reactive updates together.  With the tabulated NSE reaction
-values, we get instantaneous values of the reaction sources, but the only
-time derivative we have is for $Y_e$.  We will
-use a basic predictor-corrector update:
-\begin{equation}
-\Uc^\star = \Uc^n + \Delta t \Adv{\Uc}^{n+1/2} + \Delta t \Rb (\Uc^n)
-\end{equation}
-where $\Rb(\Uc^n)$ is the reaction source vector evaluated by calling the NSE
-table with the time-level $n$ state.
-In particular, this will also operate on the auxiliary composition variables,
-including $Y_e$:
-\begin{equation}
-(\rho Y_e)^\star = (\rho Y_e)^n + \Delta t \Adv{\rho Y_e}^{n+1/2} + \Delta t \rho (dY_e/dt)^n
-\end{equation}
-From this, we get $\rho^\star$, $(Y_e)^\star$, $(X_k)^\star$, and
-$e^\star$.  In order call the get the temperature, we need to call the
-EOS which requires $\bar{A}^\star$.  Since we cannot use the NSE table
-without the temperature, we need to
+We use the following conditions to determine if we are in NSE:
+\begin{align}
+\rho &> \rho_\mathrm{nse} \\
+T &> T_\mathrm{nse} \\
+X(\isotm{C}{12}) &< A_\mathrm{nse} \\
+X(\isotm{He}{4}) + \sum_{k \in {\rm Fe-group}} X_k &> B_\mathrm{nse} 
+\end{align}
+where the Fe-group nuclei are \isot{Cr}{48}, \isot{Fe}{52}, \isot{Fe}{54}, and \isot{Ni}{56}.
 
@@
 Also explore: PLM vs PPM, 1 vs. 2 vs. 3 iterations
-
-\subsection{NSE convergence test problem}
-
-To test the coupling of the NSE table to the hydrodynamics, we run a
-similar problem as above, except with the thermodynamic conditions
-appropriate for the matter to be in NSE.  The initial conditions are:
-\begin{align}
-\rho &= \rho_0 \\
-T &= T_0 \left [ 1 + (\delta T) e^{-(r/\lambda)^2} \cos^6(\pi r/L) \right ] \\
-Y_e &= (Y_e)_0 \left [ 1 + (\delta Y_e) e^{-(r/\lambda)^2} \cos^6(\pi r/L) \right ] 
-\end{align}
-where $r$ is the distance from the center of the domain and the
-remaining parameters are listed in Table~\ref{table:nse}.  This has
-$Y_e$ varying between $[0.47, 0.5]$ initially\MarginPar{check}.  The mass fractions
-and remaining composition variables are then initialized from the NSE table, as described in section~\ref{sec:initialization}.
-
-The domain has a size $[0, L]^2$, and the timestep is fixed as:
-\begin{equation}
-\Delta t = 5\times 10^{-4} \left ( \frac{128}{N} \right )~\mathrm{s}
-\end{equation}
-where $N$ is the number of zones in each direction.
-
-\begin{deluxetable}{lcc}
-\tablecaption{\label{table:nse} NSE convergence test problem.}
-\tablehead{\colhead{parameter} & \colhead{value}}
-\startdata
-$\rho_0$ & $5\times 10^8~\gcc$ \\
-$T_0$    & $4\times 10^9~\mathrm{K}$ \\
-$(\delta T)$ & $0.2$ \\
-$(Y_e)_0$   & $0.5$ \\
-$(\delta Y_e)$ & $-0.05$ \\
-$\lambda$   & $2\times 10^7~\mathrm{cm}$ \\
-$L$         & $10^8~\mathrm{cm}$ \\
-\enddata
-\end{deluxetable}
 
@@
           NumPy \citep{numpy,numpy2},
-          python (\url{https://www.python.org/})
+          python (\url{https://www.python.org/}), 
+          SymPy \citep{sympy}
          }
@@
 \begin{equation}
-{\bf J} = \frac{\partial \Rb}{\partial {\bf w }} \frac{\partial {\bf w}}{\partial \Uc}
-\end{equation}.
+{\bf J} = \frac{\partial \Rb}{\partial {\bf w }} \frac{\partial {\bf w}}{\partial \Uc}.
+\end{equation}
 
@@
 for the purposes of computing the Jacobian (we denote the conserved state used for the
-Jacobian $\Uc^\prime$.
+Jacobian $\Uc^\prime$).
 Writing this out for two species, $X_\alpha$ and $X_\beta$, we have

### 2020-10-23 simplified_sdc_intro/paper.tex
add more comparisons

--- 2020-10-22 simplified_sdc_intro/paper.tex
@@
 
+\begin{deluxetable}{lllllll}
+\tablecaption{\label{table:react_convert_strang} Convergence ($L_1$ norm) for the reacting convergence problem using Strang splitting.}
+\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
+           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
+           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
+\startdata
+ $\rho$                      & $2.781 \times 10^{18}$  & 2.051  & $6.711 \times 10^{17}$  & 2.580  & $1.122 \times 10^{17}$  \\
+ $\rho u$                    & $6.781 \times 10^{26}$  & 2.445  & $1.245 \times 10^{26}$  & 2.907  & $1.659 \times 10^{25}$  \\
+ $\rho v$                    & $6.781 \times 10^{26}$  & 2.445  & $1.245 \times 10^{26}$  & 2.907  & $1.659 \times 10^{25}$  \\
+ $\rho E$                    & $2.465 \times 10^{35}$  & 2.333  & $4.892 \times 10^{34}$  & 2.650  & $7.794 \times 10^{33}$  \\
+ $\rho e$                    & $2.268 \times 10^{35}$  & 2.298  & $4.610 \times 10^{34}$  & 2.722  & $6.988 \times 10^{33}$  \\
+ $T$                         & $2.248 \times 10^{21}$  & 1.682  & $7.003 \times 10^{20}$  & 2.439  & $1.291 \times 10^{20}$  \\
+ $\rho X(\isotm{He}{4})$     & $2.862 \times 10^{18}$  & 2.027  & $7.023 \times 10^{17}$  & 2.554  & $1.196 \times 10^{17}$  \\
+ $\rho X(\isotm{C}{12})$     & $1.717 \times 10^{17}$  & 1.945  & $4.460 \times 10^{16}$  & 2.194  & $9.745 \times 10^{15}$  \\
+ $\rho X(\isotm{O}{16})$     & $1.718 \times 10^{14}$  & 1.648  & $5.482 \times 10^{13}$  & 1.898  & $1.471 \times 10^{13}$  \\
+ $\rho X(\isotm{Fe}{56})$    & $2.781 \times 10^{8}$   & 2.051  & $6.711 \times 10^{7}$   & 2.580  & $1.122 \times 10^{7}$   \\
+\enddata
+\end{deluxetable}
+
+\begin{deluxetable}{lllllll}
+\tablecaption{\label{table:react_convert_strang} Convergence ($L_1$ norm) for the reacting convergence problem using Strang splitting with $T$ evolution disabled.}
+\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
+           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
+           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
+\startdata
+ $\rho$                      & $2.735 \times 10^{18}$  & 2.057  & $6.571 \times 10^{17}$  & 2.474  & $1.183 \times 10^{17}$  \\
+ $\rho u$                    & $6.729 \times 10^{26}$  & 2.416  & $1.260 \times 10^{26}$  & 2.699  & $1.941 \times 10^{25}$  \\
+ $\rho v$                    & $6.729 \times 10^{26}$  & 2.416  & $1.260 \times 10^{26}$  & 2.699  & $1.941 \times 10^{25}$  \\
+ $\rho E$                    & $2.513 \times 10^{35}$  & 2.226  & $5.372 \times 10^{34}$  & 2.169  & $1.195 \times 10^{34}$  \\
+ $\rho e$                    & $2.292 \times 10^{35}$  & 2.194  & $5.011 \times 10^{34}$  & 2.253  & $1.051 \times 10^{34}$  \\
+ $T$                         & $2.321 \times 10^{21}$  & 1.606  & $7.623 \times 10^{20}$  & 1.993  & $1.915 \times 10^{20}$  \\
+ $\rho X(\isotm{He}{4})$     & $2.806 \times 10^{18}$  & 2.028  & $6.880 \times 10^{17}$  & 2.432  & $1.275 \times 10^{17}$  \\
+ $\rho X(\isotm{C}{12})$     & $1.807 \times 10^{17}$  & 1.846  & $5.026 \times 10^{16}$  & 1.605  & $1.653 \times 10^{16}$  \\
+ $\rho X(\isotm{O}{16})$     & $1.905 \times 10^{14}$  & 1.500  & $6.737 \times 10^{13}$  & 1.439  & $2.484 \times 10^{13}$  \\
+ $\rho X(\isotm{Fe}{56})$    & $2.735 \times 10^{8}$   & 2.057  & $6.571 \times 10^{7}$   & 2.474  & $1.183 \times 10^{7}$   \\
+\enddata
+\end{deluxetable}
+
 \subsection{NSE convergence test problem}
@@
 $L$         & $10^8~\mathrm{cm}$ \\
+\enddata
+\end{deluxetable}
+
+\begin{deluxetable}{lllllll}
+\tablecaption{\label{table:nse_strang_methodI} Convergence ($L_1$ norm) for the NSE convergence
+problem using Strang splitting + Method I.}
+\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
+           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
+           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
+\startdata
+ $\rho$                      & $4.301 \times 10^{19}$  & 1.348  & $1.690 \times 10^{19}$  & 1.166  & $7.532 \times 10^{18}$  \\
+ $\rho u$                    & $1.866 \times 10^{28}$  & 1.429  & $6.929 \times 10^{27}$  & 1.101  & $3.230 \times 10^{27}$  \\
+ $\rho v$                    & $1.866 \times 10^{28}$  & 1.429  & $6.929 \times 10^{27}$  & 1.101  & $3.230 \times 10^{27}$  \\
+ $\rho E$                    & $1.781 \times 10^{38}$  & 1.053  & $8.585 \times 10^{37}$  & 1.021  & $4.231 \times 10^{37}$  \\
+ $\rho e$                    & $1.781 \times 10^{38}$  & 1.053  & $8.583 \times 10^{37}$  & 1.021  & $4.230 \times 10^{37}$  \\
+ $T$                         & $6.853 \times 10^{21}$  & 1.044  & $3.325 \times 10^{21}$  & 1.018  & $1.642 \times 10^{21}$  \\
+ $\rho X(\isotm{He}{4})$     & $1.070 \times 10^{18}$  & 1.186  & $4.701 \times 10^{17}$  & 1.051  & $2.269 \times 10^{17}$  \\
+ $\rho X(\isotm{C}{12})$     & $2.166 \times 10^{12}$  & 1.533  & $7.482 \times 10^{11}$  & 1.189  & $3.283 \times 10^{11}$  \\
+ $\rho X(\isotm{O}{16})$     & $1.244 \times 10^{13}$  & 1.396  & $4.727 \times 10^{12}$  & 1.117  & $2.179 \times 10^{12}$  \\
+ $\rho Y_e$                  & $2.442 \times 10^{19}$  & 1.442  & $8.990 \times 10^{18}$  & 1.212  & $3.880 \times 10^{18}$  \\
+ $\rho \bar{A}$              & $1.055 \times 10^{22}$  & 1.032  & $5.158 \times 10^{21}$  & 1.022  & $2.541 \times 10^{21}$  \\
+ $\rho (B/A)$                & $4.512 \times 10^{20}$  & 1.386  & $1.727 \times 10^{20}$  & 1.207  & $7.479 \times 10^{19}$  \\
+\enddata
+\end{deluxetable}
+
+\begin{deluxetable}{lllllll}
+\tablecaption{\label{table:nse_strang_methodII} Convergence ($L_1$ norm) for the NSE convergence
+problem using Strang splitting + Method II.}
+\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
+           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
+           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
+\startdata
+ $\rho$                      & $4.522 \times 10^{19}$  & 1.659  & $1.432 \times 10^{19}$  & 1.542  & $4.917 \times 10^{18}$  \\
+ $\rho u$                    & $1.883 \times 10^{28}$  & 1.681  & $5.873 \times 10^{27}$  & 1.524  & $2.042 \times 10^{27}$  \\
+ $\rho v$                    & $1.883 \times 10^{28}$  & 1.681  & $5.873 \times 10^{27}$  & 1.524  & $2.042 \times 10^{27}$  \\
+ $\rho E$                    & $5.933 \times 10^{37}$  & 1.613  & $1.940 \times 10^{37}$  & 1.411  & $7.293 \times 10^{36}$  \\
+ $\rho e$                    & $5.931 \times 10^{37}$  & 1.613  & $1.939 \times 10^{37}$  & 1.411  & $7.291 \times 10^{36}$  \\
+ $T$                         & $9.885 \times 10^{20}$  & 1.207  & $4.281 \times 10^{20}$  & 1.058  & $2.057 \times 10^{20}$  \\
+ $\rho X(\isotm{He}{4})$     & $5.086 \times 10^{17}$  & 1.430  & $1.887 \times 10^{17}$  & 1.187  & $8.288 \times 10^{16}$  \\
+ $\rho X(\isotm{C}{12})$     & $1.585 \times 10^{12}$  & 1.829  & $4.462 \times 10^{11}$  & 1.433  & $1.652 \times 10^{11}$  \\
+ $\rho X(\isotm{O}{16})$     & $8.060 \times 10^{12}$  & 1.706  & $2.470 \times 10^{12}$  & 1.347  & $9.710 \times 10^{11}$  \\
+ $\rho Y_e$                  & $2.379 \times 10^{19}$  & 1.678  & $7.431 \times 10^{18}$  & 1.578  & $2.489 \times 10^{18}$  \\
+ $\rho \bar{A}$              & $5.141 \times 10^{21}$  & 1.230  & $2.191 \times 10^{21}$  & 1.087  & $1.032 \times 10^{21}$  \\
+ $\rho (B/A)$                & $4.149 \times 10^{20}$  & 1.693  & $1.283 \times 10^{20}$  & 1.562  & $4.345 \times 10^{19}$  \\
 \enddata

### 2020-10-23 simplified_sdc_intro/paper.tex
remove composition for NSE test
some paper updates

@@
 \newcommand{\Advt}[1]{{\left [\boldsymbol{\mathcal{\tilde{A}}} \left(#1\right)\right]}}
+\newcommand{\Advss}[1]{{\left [{\mathcal{{A}}} \left(#1\right)\right]}}
 \newcommand{\Advs}[1]{\boldsymbol{\mathcal{A}} \left(#1\right)}
@@
 scheme, making this method easier to add to existing simulation codes.
-However, it is restricted to second-order accuracy overall.  We term
-this algorithm the ``simplified SDC method''.  We use the
-\castro\ hydrodynamics code \cite{castro} for all of our numerical
-experiments, and all of the code to reproduce the results in this
-paper are in the \castro\ github
+It also extends to adaptive mesh refinement with subcycling in a
+straightforward manner, avoiding the complications described in
+\citep{mccorquodalecolella} needed to fill ghost cells when using
+method-of-lines integration.  However, it is restricted to
+second-order accuracy overall.  We term this algorithm the
+``simplified SDC method''.  We use the \castro\ hydrodynamics code
+\citep{castro} for all of our numerical experiments, and all of the
+code to reproduce the results in this paper are in the \castro\ github
 repository\footnote{\url{https://github.com/amrex-astro/Castro/}}.
@@
 \begin{equation}
-  \Sq = \Sqhydro + \Rb(q)
+  \Sq = \Sqhydro + \Rb(\qb)
 \end{equation}
@@
 
-    Note that our advection terms are piecewise constant \MarginPar{need to think about this more}
-    approximations, so when we integrate our system, their
-    contributions will be linear in time.  Since $(\rho X_k)$ sees a
-    linear advective term, $\rho$ does as well.  $(\rho \Ub)$ and
-    $(\rho E)$ will also have linear contributions in time from
-    advection.  We can see that the internal energy will also have a
-    linear contribution by recognizing that the kinetic energy term,
-    $K = (\rho |\Ub|^2)/\rho$ is linear in time.
-    Since the reactions don't affect $\rho$ and $(\rho \Ub)$, we can
-    algebraically update these using this piecewise linear behavior:
+    Note that our advection terms are piecewise
+    constant \MarginPar{need to think about this more} approximations,
+    so when we integrate our system, their contributions will be
+    linear in time.  Since $(\rho X_k)$ sees a linear advective term,
+    $\rho$ does as well.  $(\rho u)$ will also have linear
+    contributions in time from advection.  Since the reactions don't
+    affect $\rho$ and $(\rho u)$, we can algebraically update these
+    using this piecewise linear behavior:
     \begin{align}
-      \rho(t) &= \rho^n + \Adv{\rho}^{n+1/2,(k)} \, t \\
-      (\rho \Ub)(t) &= (\rho \Ub)^n + \Adv{\rho \Ub}^{n+1/2,(k)} \, t
+      \rho(t) &= \rho^n + \Advss{\rho}^{n+1/2,(k)} \, t \\
+      (\rho u)(t) &= (\rho u)^n + \Advss{\rho u}^{n+1/2,(k)} \, t
     \end{align}
@@
         This corrects the composition so that it is consistent with the updated
-        $Y_e$.  We use these values $\bar{A}_\out$, $(X_k)_\out$, and $(B/A)_\out$ to then complete the update, computing the energy as in Eq.~\ref{eq:nse_energy}.
+        $Y_e$.
+We use these values $\bar{A}_\out$, $(X_k)_\out$, and $(B/A)_\out$ to then complete the update, computing the energy as in Eq.~\ref{eq:nse_energy}.
+\MarginPar{we could imagine looping over the EOS to get an updated T and then the table to get the updated $\bar{A}$, to see if that converges?}
 
@@
 \begin{equation}
-(\rho Y_e)^\star = (\rho Y_e)^n + \Delta t \Adv{\rho Y_e}^{n+1/2} + \Delta t \rho (dY_e/dt)^n
+(\rho Y_e)^\star = (\rho Y_e)^n + \Delta t \Advss{\rho Y_e}^{n+1/2} + \Delta t \rho (dY_e/dt)^n
 \end{equation}
@@
 \begin{equation}
-(\rho E)^{n+1,(k)} = (\rho E)^n + \Delta t \Adv{\rho E}^{n+1/2} + (\rho \Sdot)^{n+1/2}
+(\rho E)^{n+1,(k)} = (\rho E)^n + \Delta t \Advss{\rho E}^{n+1/2} + (\rho \Sdot)^{n+1/2}
 \end{equation}
@@
  $T$                         & $6.853 \times 10^{21}$  & 1.044  & $3.325 \times 10^{21}$  & 1.018  & $1.642 \times 10^{21}$  \\
- $\rho X(\isotm{He}{4})$     & $1.070 \times 10^{18}$  & 1.186  & $4.701 \times 10^{17}$  & 1.051  & $2.269 \times 10^{17}$  \\
- $\rho X(\isotm{C}{12})$     & $2.166 \times 10^{12}$  & 1.533  & $7.482 \times 10^{11}$  & 1.189  & $3.283 \times 10^{11}$  \\
- $\rho X(\isotm{O}{16})$     & $1.244 \times 10^{13}$  & 1.396  & $4.727 \times 10^{12}$  & 1.117  & $2.179 \times 10^{12}$  \\
  $\rho Y_e$                  & $2.442 \times 10^{19}$  & 1.442  & $8.990 \times 10^{18}$  & 1.212  & $3.880 \times 10^{18}$  \\
@@
  $T$                         & $9.885 \times 10^{20}$  & 1.207  & $4.281 \times 10^{20}$  & 1.058  & $2.057 \times 10^{20}$  \\
- $\rho X(\isotm{He}{4})$     & $5.086 \times 10^{17}$  & 1.430  & $1.887 \times 10^{17}$  & 1.187  & $8.288 \times 10^{16}$  \\
- $\rho X(\isotm{C}{12})$     & $1.585 \times 10^{12}$  & 1.829  & $4.462 \times 10^{11}$  & 1.433  & $1.652 \times 10^{11}$  \\
- $\rho X(\isotm{O}{16})$     & $8.060 \times 10^{12}$  & 1.706  & $2.470 \times 10^{12}$  & 1.347  & $9.710 \times 10^{11}$  \\
  $\rho Y_e$                  & $2.379 \times 10^{19}$  & 1.678  & $7.431 \times 10^{18}$  & 1.578  & $2.489 \times 10^{18}$  \\
@@
 
-We presented a simple spectral deferred corrections scheme for coupling
+We presented a simplified spectral deferred corrections scheme for coupling
 hydrodynamics and reactions.
+
+Future work is to extend this methodology to MHD.
 
@@
 \begin{equation}
-K = \frac{1}{2} |\Ub|^2
+K = \frac{1}{2} u^2
 \end{equation}
@@
 \end{equation}
-and write the kinetic energy as $K = |\Ub|^2/2$.  We get the inverse
+and write the kinetic energy as $K = u^2/2$.  We get the inverse
 (computed via SymPy, see the included Jupyter notebook) is:

### 2020-10-23 simplified_sdc_intro/paper.tex
add a figure
more tables

@@
 
+\begin{deluxetable}{lllllll}
+\tablecaption{\label{table:react_convert_strang} Convergence ($L_1$ norm) for the reacting convergence problem using SDC integration.}
+\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
+           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
+           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
+\startdata
+ $\rho$                      & $2.795 \times 10^{18}$  & 2.044  & $6.777 \times 10^{17}$  & 2.554  & $1.154 \times 10^{17}$  \\
+ $\rho u$                    & $6.798 \times 10^{26}$  & 2.449  & $1.245 \times 10^{26}$  & 2.889  & $1.680 \times 10^{25}$  \\
+ $\rho v$                    & $6.798 \times 10^{26}$  & 2.449  & $1.245 \times 10^{26}$  & 2.889  & $1.680 \times 10^{25}$  \\
+ $\rho E$                    & $2.452 \times 10^{35}$  & 2.352  & $4.804 \times 10^{34}$  & 2.742  & $7.179 \times 10^{33}$  \\
+ $\rho e$                    & $2.262 \times 10^{35}$  & 2.321  & $4.527 \times 10^{34}$  & 2.822  & $6.404 \times 10^{33}$  \\
+ $T$                         & $2.236 \times 10^{21}$  & 1.691  & $6.927 \times 10^{20}$  & 2.481  & $1.240 \times 10^{20}$  \\
+ $\rho X(\isotm{He}{4})$     & $2.879 \times 10^{18}$  & 2.020  & $7.096 \times 10^{17}$  & 2.529  & $1.229 \times 10^{17}$  \\
+ $\rho X(\isotm{C}{12})$     & $1.697 \times 10^{17}$  & 1.950  & $4.393 \times 10^{16}$  & 2.231  & $9.357 \times 10^{15}$  \\
+ $\rho X(\isotm{O}{16})$     & $1.686 \times 10^{14}$  & 1.659  & $5.338 \times 10^{13}$  & 1.957  & $1.375 \times 10^{13}$  \\
+ $\rho X(\isotm{Fe}{56})$    & $2.795 \times 10^{8}$   & 2.044  & $6.777 \times 10^{7}$   & 2.554  & $1.154 \times 10^{7}$   \\
+\enddata
+\end{deluxetable}
+
 \subsection{NSE convergence test problem}
@@
 \begin{equation}
-\Delta t = 5\times 10^{-4} \left ( \frac{128}{N} \right )~\mathrm{s}
+\Delta t = 10^{-3} \left ( \frac{64}{N} \right )~\mathrm{s}
 \end{equation}
@@
 \end{deluxetable}
+
+Figure~\ref{fig:nse_ye} shows the $Y_e$ profile for the Strang evolution case.
+
+\begin{figure}[t]
+\centering
+\plotone{ye_plot}
+\caption{\label{fig:nse_ye} Profile of $Y_e$ through the middle of the domain (slicing along $x$)
+for the $128^2$ Strang method II NSE test simulation.}
+\end{figure}
 

### 2020-10-23 simplified_sdc_intro/paper.tex
add note
update Strang no-NSE

@@
 \begin{equation}
+  \label{eq:aux:ye}
   Y_e = \sum_k \frac{Z_k X_k}{A_k}
@@
 \begin{align}
+\label{eq:aux:abar}
 \bar{A} &= \left [ \sum_k \frac{X_k}{A_k} \right ]^{-1} \\
+\label{eq:aux:bea}
 \left (\frac{B}{A} \right ) &= \sum_k \frac{B_k X_k}{A_k}
@@
     \begin{itemize}
-    \item integrate the aprox19 network as usual
-    \item update the aux quantities at the end of the burn
+    \item Integrate the full reaction network (Eqs.~\ref{eq:strang:T} and \ref{eq:strang:X}) as usual for Strang splitting
+    \item Update the aux quantities at the end of the burn using Eqs.~\ref{eq:aux:ye}, \ref{eq:aux:abar}, and \ref{eq:aux:bea} with the new mass fractions, $X_k$.
     \end{itemize}
@@
 $e^\star$, which allows us to call the EOS and get an updated
-temperature, $T^\star$.
+temperature, $T^\star$. \MarginPar{Andy suggests time-centering $dY_e/dt$}
 We can then again call the NSE table:

### 2020-10-26 simplified_sdc_intro/paper.tex
quick pass through abstract and intro

--- 2020-10-23 simplified_sdc_intro/paper.tex
@@
 \newcommand{\Ic}{{\boldsymbol{\mathcal{I}}}}
-\newcommand{\Ics}{{\mathcal{I}}}
 \newcommand{\smax}{{s_\mathrm{max}}}
@@
 \newcommand{\Advt}[1]{{\left [\boldsymbol{\mathcal{\tilde{A}}} \left(#1\right)\right]}}
-\newcommand{\Advss}[1]{{\left [{\mathcal{{A}}} \left(#1\right)\right]}}
 \newcommand{\Advs}[1]{\boldsymbol{\mathcal{A}} \left(#1\right)}
@@
 %======================================================================
-\title{A Simplified Spectral Deferred Correction Method for Coupling Hydrodynamics with Reaction Networks and Nuclear Statistical Equilibrium}
-
-\shorttitle{A Simplified SDC Method}
+%\title{A Simplified Spectral Deferred Correction Method for Coupling Hydrodynamics with Reaction Networks and Nuclear Statistical Equilibrium}
+\title{Improved Hydrodynamic/Reaction Coupling for Flows With Nuclear Statistical Equilibrium}
+
+%\shorttitle{A Simplified SDC Method}
+\shorttitle{Hydro/Reaction Coupling with NSE}
+
 \shortauthors{}
@@
 statistical equilibrium is established.  We develop a simplified
-approach to spectral deferred corrections (SDC) coupling of explicit
+approach based on spectral deferred corrections (SDC) coupling of explicit
 hydrodynamics and stiff reaction sources as an alternative to operator
 splitting or the more comprehensive SDC approach we demonstrated
-previously.  We apply the new method some example problems and show
+previously.  We apply the new method to some example problems and show
 how to modify it to work with a hybrid network consisting of a
@@
 
-Reactive flow can be challenging to model when the timescale for
-changes in the nuclear abundances due to reactions is comparable to
-the hydrodynamical timescales.  Traditional methods of coupling
+Reactive flow can be challenging to model when changes in the
+nuclear abundances due to reactions are integrated over hydrodynamic
+time scales.  Traditional methods of coupling
 hydrodynamics and reactions used in astrophysics use operator
 splitting---each physical process acts on the output of the previous
-process.  This makes it easy to add physics in a modular way to a
+process in alternating fashion.  This makes it easy to add physics in a modular way to a
 simulation code, but competition between physical processes can cause
-the coupling to breakdown.  In general, the error in the splitting
-controlled by the timestep.
-
-A particular difficult phase of evolution to model is the nuclear
-statistic equilibrium that sets in with helium burning for
+the coupling to breakdown.  These splitting errors can lead to loss of
+accuracy and further time step limitations.
+
+A particularly difficult phase of evolution to model is the nuclear
+statistical equilibrium that sets in with helium burning for
 temperatures in excess of $4\times 10^9~\mathrm{K}$.  Physically, the
-forward and reverse rates should balance leading to an equilibrium.
+forward and reverse rates of reaction should balance leading to an equilibrium.
 With operator splitting, an NSE region will have a large positive flow
@@
 flow over the next timestep, as the code struggles to produce an
-equilibrium.  This large change in abundances (and large alternately
+equilibrium.  These large changes in abundances (and large alternately
 positive and negative energy generation rates) can be a challenge for
@@
 reaction networks based on spectral deferred corrections (SDC)
-methods, and demonstrated the method of a variety of test problems.
-The approach presented here is considerably simpler, reusing the main
+methods, and demonstrated the method in a variety of test problems.
+The approach presented here is considerably simpler from a temporal
+integration standpoint, but allows is to incorporate modifications
+for regions in nuclear statistical equilibirum.
+Furthermore, we reuse the main
 CTU hydrodynamics construction and a largely similar ODE integration
-scheme, making this method easier to add to existing simulation codes.
-It also extends to adaptive mesh refinement with subcycling in a
-straightforward manner, avoiding the complications described in
-\citep{mccorquodalecolella} needed to fill ghost cells when using
-method-of-lines integration.  However, it is restricted to
-second-order accuracy overall.  We term this algorithm the
-``simplified SDC method''.  We use the \castro\ hydrodynamics code
-\citep{castro} for all of our numerical experiments, and all of the
-code to reproduce the results in this paper are in the \castro\ github
+scheme from the original Strang split approach,
+making this method easier to add to existing simulation codes.
+Overall, compared to the original Strang split approach, this new
+approach is more computationally efficient, allows for a larger
+time step, and reduces overall integration error.
+However, it is restricted to second-order accuracy overall.  We term
+this algorithm the ``simplified SDC method''.  We use the
+\castro\ hydrodynamics code \cite{castro} for all of our numerical
+experiments, and all of the code to reproduce the results in this
+paper is in the \castro\ github
 repository\footnote{\url{https://github.com/amrex-astro/Castro/}}.
@@
 \begin{equation}
-  \Sq = \Sqhydro + \Rb(\qb)
+  \Sq = \Sqhydro + \Rb(q)
 \end{equation}
@@
   with a superscript (like $n$ or $n+1$), we'll use a second subscript
-  in parentheses to keep track of the iteration.  A single iteration
-  starts with $\Uc^n$ and results in the new time-level state for that
-  iteration, $\Uc^{n+1,(k)}$.
+  in parentheses to keep track of the iteration.
 
@@
     Now the conservative hydrodynamics source terms are computed by first updating to the
-    new state with advection and the old-time source term applied for the full $\Delta t$ as:
-    \begin{equation}
-      \Uc^{\star\star} = \Uc^n + \Delta t \Advt{\Uc}^{n+1/2,(k)} + \Delta t \Hb(\Uc^n)
+    new state with advection only\footnote{For a source like gravity, this update can be done first for $\rho$ and then define the new momentum source using $\rho^{\star\star}$ and likewise for energy}, as:
+    \begin{equation}
+      \Uc^{\star\star} = \Uc^n + \Delta t \Advt{\Uc}^{n+1/2,(k)}
     \end{equation}
@@
     %% \end{equation}
-    We then evaluate the source terms with $\Uc^{\star\star}$ and
-    correct the advective term so that we have a time-centered
-    source: \MarginPar{are we actually doing this correctly for SDC?}
-    The final advective update term is then\footnote{For a source like gravity, this update can be done first for $\rho$ and then define the new momentum source using $\rho^{\star\star}$ and likewise for energy}:
-    \begin{equation}
-      \Adv{\Uc}^{n+1/2,(k)}_{i} = \Advt{\Uc}^{n+1/2,(k)}_{i} +
-      \frac{\Delta t}{2} \left (\Shydro(\Uc^n) + \Shydro(\Uc^{\star\star}) \right )
-    \end{equation}
-
-    %% with
-    %% \begin{equation}
-    %%   \Shydro^{n+1/2} = \left ( \begin{array}{c}
-    %%                  0 \\ 0 \\ 0 \\
-    %%                 {\bf S}_{\rho\Ub}^{n+1/2} \cdot \ex \\
-    %%                 S_{\rho E}^{n+1/2} \end{array} \right )
-    %% \end{equation}
+
+    The final advective update term is then:
+    \begin{equation}
+      \Adv{\Uc}^{n+1/2,(k)}_{i} = \Advt{\Uc}^{n+1/2,(k)}_{i} + \Shydro^{n+1/2}
+    \end{equation}
+    with
+    \begin{equation}
+      \Shydro^{n+1/2} = \left ( \begin{array}{c}
+                     0 \\ 0 \\
+                    {\bf S}_{\rho\Ub}^{n+1/2} \cdot \ex \\
+                    S_{\rho E}^{n+1/2} \end{array} \right )
+    \end{equation}
 
@@
     \begin{equation}
-      \odt{\Uc} = \Adv{\Uc}^{n+1/2,(k)} + \Rb(\Uc)
-    \end{equation}
-    This integration begins with $\Uc^n$ and results in $\Uc^{n+1,(k)}$.
+      \odt{\Uc} = \Adv{\Uc}^{n+1,(k)} + \Rb(\Uc)
+    \end{equation}
 
@@
 
-    Note that our advection terms are piecewise
-    constant \MarginPar{need to think about this more} approximations,
-    so when we integrate our system, their contributions will be
-    linear in time.  Since $(\rho X_k)$ sees a linear advective term,
-    $\rho$ does as well.  $(\rho u)$ will also have linear
-    contributions in time from advection.  Since the reactions don't
-    affect $\rho$ and $(\rho u)$, we can algebraically update these
-    using this piecewise linear behavior:
+    Note that our advection terms are piecewise constant \MarginPar{need to think about this more}
+    approximations, so when we integrate our system, their
+    contributions will be linear in time.  Since $(\rho X_k)$ sees a
+    linear advective term, $\rho$ does as well.  $(\rho \Ub)$ and
+    $(\rho E)$ will also have linear contributions in time from
+    advection.  We can see that the internal energy will also have a
+    linear contribution by recognizing that the kinetic energy term,
+    $K = (\rho |\Ub|^2)/\rho$ is linear in time.
+    Since the reactions don't affect $\rho$ and $(\rho \Ub)$, we can
+    algebraically update these using this piecewise linear behavior:
     \begin{align}
-      \rho(t) &= \rho^n + \Advss{\rho}^{n+1/2,(k)} \, t \\
-      (\rho u)(t) &= (\rho u)^n + \Advss{\rho u}^{n+1/2,(k)} \, t
+      \rho(t) &= \rho^n + \Adv{\rho}^{n+1/2,(k)} \, t \\
+      (\rho \Ub)(t) &= (\rho \Ub)^n + \Adv{\rho \Ub}^{n+1/2,(k)} \, t
     \end{align}
+
+    At the end of this integration, we have the state at the next iteration,
+    $\Uc^{n+1,(k+1)}$
 
@@
 we store these 3 quantities as auxiliary data that is carried along
-with the rest of the fluid state in each zone.  The table also returns
-values of the mass fractions mapped onto the 19-isotopes we carry, $X_k$, and the time-derivative of $Y_e$, $dY_e/dt$.
-We'll use the notation:
-\begin{equation}
-\nse{\rho,T,Y_e} \rightarrow \bar{A}, X_k, (B/A), dY_e/dt
-\end{equation}
-to represent the NSE table call and its inputs and outputs.
-
-We can derive evolution equations for each of these composition quantities as:
+with the rest of the fluid state in each zone.
+
+For Strang split coupling of hydro and reactions, $DX_k/Dt = 0$,
+and our evolution equations are:
 \begin{align}
-\frac{DY_e}{Dt} &= \sum_k \frac{Z_k}{A_k} \frac{DX_k}{Dt} = \sum_k \frac{Z_k}{A_k} \omegadot_k \\
-\frac{D\bar{A}}{Dt} &= -\bar{A}^2 \sum_k \frac{1}{A_k} \frac{DX_k}{Dt} = -\bar{A}^2 \sum_k \frac{1}{A_k} \omegadot_k \\
-\frac{D}{Dt} \left (\frac{B}{A} \right ) &= \sum_k \frac{B_k}{A_k} \frac{DX_k}{Dt} = \sum_k \frac{B_k}{A_k} \omegadot_k
+\frac{DY_e}{Dt} &= \sum_k \frac{Z_k}{A_k} \frac{DX_k}{Dt} = 0 \\
+\frac{D}{Dt} \frac{1}{\bar{A}} &= - \frac{1}{\bar{A}^2} \frac{D\bar{A}}{Dt} = \sum_k \frac{1}{A_k} \frac{DX_k}{Dt} = 0 \rightarrow \frac{D\bar{A}}{Dt} = 0 \\
+\frac{D}{Dt} \left (\frac{B}{A} \right ) &= \sum_k \frac{B_k}{A_k} \frac{DX_k}{Dt} = 0
 \end{align}
-For Strang split coupling of hydro and reactions, $DX_k/Dt = 0$,
-Therefore each of these auxillary equations obeys an advection
+Therefore each of these auxiliary equations obeys an advection
 equation in the hydro part of the advancement.  In the SDC algorithm,
 there will be a reactive source (an $\Ic_q$) for each of these that is
-computed in the same manner as above.  We note that our NSE table
-provides the evolution of $Y_e$ due to reactions ($DY_e/Dt$) directly.
+computed in the same manner as above.
 
  The compositional quantities
-it carries, $\bar{A}$ and $Y_e$ and not representable from the 19
+it carries, $\bar{A}$ and $Y_e$, are not representable from the 19
 isotopes we carry in the main network. For this reason, when we are
@@
   \begin{itemize}
-
   \item For a zone that is in NSE:
 
-      The goal is to update the composition and thermodynamics due to the 
-      change in the nuclei abundances over the (half-)timestep.  There are
-      a few approaches we can do here.
-
     \begin{itemize}
-      \item Method I: {\em Use the time-level $n$ information for the update}.
-
-        This is the approach used in \citet{ma:2013}.  We call the table with the 
-        input state:
-        \begin{equation}
-          \nse{\rho, T, (Y_e)_{\rm in}} \rightarrow \bar{A}_\out, (X_k)_\out, (B/A)_\out, dY_e/dt
-        \end{equation}
-
-        We can now update $Y_e$:
+    \item use $\rho$, $T$, and $Y_e$ to call the table. This
+      returns: $dY_e/dt$, $(B/A)_{\rm out}$, $\bar{A}_{\rm out}$,
+      and the $(X_k)_{\rm out}$ from the NSE network grouped into the 19 isotopes
+      we carry as part of aprox19.
+
+    \item update $Y_e$:
     \begin{equation}
@@
     \end{equation}
-        and take the auxiliary state values of $\bar{A}$ and $(B/A)$ to be simply
-        the values from the table, $\bar{A}_\out$ and $(B/A)_\out$, and the mass fractions
+
+  \item update the auxiliary state value of $\bar{A}$ to be simply
+    $\bar{A}_{\rm out}$ from the table and the mass fractions
     $X_k$ to be the $(X_k)_{\rm out}$ returned from the table.
 
-        Finally, we compute the energy generation rate, $\Sdot$ as:
-        \begin{equation}
-          \label{eq:nse_energy}
-          \Sdot = \left [ \left ( \frac{B}{A} \right )_{\rm out} -
+  \item compute the energy generation rate, $\Sdot$ as:
+    \begin{equation}
+      \Sdot = \eta \left [ \left ( \frac{B}{A} \right )_{\rm out} -
         \left ( \frac{B}{A} \right )_{\rm in} \right ] N_A \frac{1}{\Delta t}
     \end{equation}
-
-        We note that \citet{ma:2013} modified Eq.~\ref{eq:nse_energy}
-        to take use only a fraction (0.3) of the energy change from
-        the table, to avoid a potential instability that can arise if
-        too much energy is deposited in a zone in a single timestep
-        (they called this an inertia term).  We prefer to deal with
-        the issue of too much energy being deposited in a zone in one
-        step through the retry mechanism already built into \castro.
-
-      \item Method II: {\em predictor-corrector.}
-
-        This starts out the same as the above method:
-        \begin{equation}
-          \nse{\rho, T, (Y_e)_{\rm in}} \rightarrow \bar{A}^\star, (X_k)^\star, (B/A)^\star, dY_e/dt
-        \end{equation}
-        but we indicate with a $\star$ that most of those values are
-        provisional, and we will seek a better value in the corrector step.
-        We only update $Y_e$ from this call:
-        \begin{equation}
-          (Y_e)_{\rm out} = (Y_e)_{\rm in} + \Delta t \frac{dY_e}{dt}
-        \end{equation}
-        and then we use the table again, but with the updated $Y_e$:
-        \begin{equation}
-          \nse{\rho, T, (Y_e)_\out} \rightarrow \bar{A}_\out, (X_k)_\out, (B/A)_\out, dY_e/dt
-        \end{equation}
-        This corrects the composition so that it is consistent with the updated
-        $Y_e$.
-We use these values $\bar{A}_\out$, $(X_k)_\out$, and $(B/A)_\out$ to then complete the update, computing the energy as in Eq.~\ref{eq:nse_energy}.
-\MarginPar{we could imagine looping over the EOS to get an updated T and then the table to get the updated $\bar{A}$, to see if that converges?}
+    where $\eta$ is an inertia term < 1 to prevent the energy changing
+    too much in one set.
+
+  \item set the auxiliary state value of $(B/A)$ as:
+    \begin{equation}
+      \left (\frac{B}{A}\right )_{\rm out} = \left (\frac{B}{A}\right )_{\rm in} + \eta \left [ \left ( \frac{B}{A}\right )_{\rm out} - \left (\frac{B}{A} \right )_{\rm in} \right]
+    \end{equation}
 
@@
 
-\subsection{SDC-NSE Coupling}
-
-With SDC evolution, when we are in NSE, we need to do the advective
-and reactive updates together.  With the tabulated NSE reaction
-values, we get instantaneous values of the reaction sources, but the only
-time derivative we have is for $Y_e$.  We will
-use a basic predictor-corrector update:
-\begin{equation}
-\Uc^\star = \Uc^n + \Delta t \Adv{\Uc}^{n+1/2} + \Delta t \Rb (\Uc^n)
-\end{equation}
-Calling the NSE table on the starting state gives use:
-\begin{equation}
-\nse{\rho^n, T^n, (Y_e)^n} \rightarrow \bar{A}^n, (X_k)^n, (B/A)^n, (dY_e/dt)^n
-\end{equation}
-To get the reactive source, we need to use the iteratively-lagged information
-from $\Ic_q^{n+1/2,(k-1)}$:
-\begin{equation}
-\Rb(\Uc^n) \approx \left (\begin{array}{c}
-  0 \\
-  \rho^n \Ics_q(\omegadot(X_k)) \\
-  \rho^n \Ics_q(\omegadot(\alpha_l)) \\
-  0 \\
-  \Ics_q(\rho e) 
-\end{array}
-\right )
-\end{equation}
-Since we have $(dY_e/dt)^n$ from the table, we replace the source for $Y_e$ in the above
-$\Rb(\Uc^n)$ as:
-\begin{equation}
-R(\rho Y_e) = \rho (dY_e/dt)^n
-\end{equation}
-which gives an update for the electron fraction as:
-\begin{equation}
-(\rho Y_e)^\star = (\rho Y_e)^n + \Delta t \Advss{\rho Y_e}^{n+1/2} + \Delta t \rho (dY_e/dt)^n
-\end{equation}
-From this prediction, we get $\rho^\star$, $\alpha_l^\star$, and
-$e^\star$, which allows us to call the EOS and get an updated
-temperature, $T^\star$.
-We can then again call the NSE table:
-\begin{equation}
-\nse{\rho^\star, T^\star, (Y_e)^\star} \rightarrow \bar{A}^{n+1,(k)}, (X_k)^{n+1,(k)}, (B/A)^{n+1,(k)}, (dY_e/dt)
-\end{equation}
-where we use the outputs to define the new values of the composition variables.
-Since the density and momenta don't have reactive sources, we just use the $\Uc^\star$ state
-as their new values:
+We use the following conditions to determine if we are in NSE:
 \begin{align}
-\rho^{n+1,(k)} &= \rho^\star \\
-(\rho u)^{n+1,(k)} &= (\rho u)^\star
+\rho &> \rho_\mathrm{nse} \\
+T &> T_\mathrm{nse} \\
+X(\isotm{C}{12}) &< A_\mathrm{nse} \\
+X(\isotm{He}{4}) + \sum_{k \in {\rm Fe-group}} X_k &> B_\mathrm{nse} 
 \end{align}
-We then compute the energy release as
-\begin{equation}
-       (\rho \Sdot)^{n+1/2} = \left [ \rho^{n+1,(k)} \left ( \frac{B}{A} \right )^{n+1,(k)} -
-                            \rho^n \left ( \frac{B}{A} \right )^n \right ] N_A \frac{1}{\Delta t}
-\end{equation}\MarginPar{is this right?}
-and then do the final update of the total energy:
-\begin{equation}
-(\rho E)^{n+1,(k)} = (\rho E)^n + \Delta t \Advss{\rho E}^{n+1/2} + (\rho \Sdot)^{n+1/2}
-\end{equation}
+where the Fe-group nuclei are \isot{Cr}{48}, \isot{Fe}{52}, \isot{Fe}{54}, and \isot{Ni}{56}.
 
@@
 
-\begin{deluxetable}{lllllll}
-\tablecaption{\label{table:react_convert_strang} Convergence ($L_1$ norm) for the reacting convergence problem using Strang splitting.}
-\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
-\startdata
- $\rho$                      & $2.781 \times 10^{18}$  & 2.051  & $6.711 \times 10^{17}$  & 2.580  & $1.122 \times 10^{17}$  \\
- $\rho u$                    & $6.781 \times 10^{26}$  & 2.445  & $1.245 \times 10^{26}$  & 2.907  & $1.659 \times 10^{25}$  \\
- $\rho v$                    & $6.781 \times 10^{26}$  & 2.445  & $1.245 \times 10^{26}$  & 2.907  & $1.659 \times 10^{25}$  \\
- $\rho E$                    & $2.465 \times 10^{35}$  & 2.333  & $4.892 \times 10^{34}$  & 2.650  & $7.794 \times 10^{33}$  \\
- $\rho e$                    & $2.268 \times 10^{35}$  & 2.298  & $4.610 \times 10^{34}$  & 2.722  & $6.988 \times 10^{33}$  \\
- $T$                         & $2.248 \times 10^{21}$  & 1.682  & $7.003 \times 10^{20}$  & 2.439  & $1.291 \times 10^{20}$  \\
- $\rho X(\isotm{He}{4})$     & $2.862 \times 10^{18}$  & 2.027  & $7.023 \times 10^{17}$  & 2.554  & $1.196 \times 10^{17}$  \\
- $\rho X(\isotm{C}{12})$     & $1.717 \times 10^{17}$  & 1.945  & $4.460 \times 10^{16}$  & 2.194  & $9.745 \times 10^{15}$  \\
- $\rho X(\isotm{O}{16})$     & $1.718 \times 10^{14}$  & 1.648  & $5.482 \times 10^{13}$  & 1.898  & $1.471 \times 10^{13}$  \\
- $\rho X(\isotm{Fe}{56})$    & $2.781 \times 10^{8}$   & 2.051  & $6.711 \times 10^{7}$   & 2.580  & $1.122 \times 10^{7}$   \\
-\enddata
-\end{deluxetable}
-
-\begin{deluxetable}{lllllll}
-\tablecaption{\label{table:react_convert_strang} Convergence ($L_1$ norm) for the reacting convergence problem using Strang splitting with $T$ evolution disabled.}
-\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
-\startdata
- $\rho$                      & $2.735 \times 10^{18}$  & 2.057  & $6.571 \times 10^{17}$  & 2.474  & $1.183 \times 10^{17}$  \\
- $\rho u$                    & $6.729 \times 10^{26}$  & 2.416  & $1.260 \times 10^{26}$  & 2.699  & $1.941 \times 10^{25}$  \\
- $\rho v$                    & $6.729 \times 10^{26}$  & 2.416  & $1.260 \times 10^{26}$  & 2.699  & $1.941 \times 10^{25}$  \\
- $\rho E$                    & $2.513 \times 10^{35}$  & 2.226  & $5.372 \times 10^{34}$  & 2.169  & $1.195 \times 10^{34}$  \\
- $\rho e$                    & $2.292 \times 10^{35}$  & 2.194  & $5.011 \times 10^{34}$  & 2.253  & $1.051 \times 10^{34}$  \\
- $T$                         & $2.321 \times 10^{21}$  & 1.606  & $7.623 \times 10^{20}$  & 1.993  & $1.915 \times 10^{20}$  \\
- $\rho X(\isotm{He}{4})$     & $2.806 \times 10^{18}$  & 2.028  & $6.880 \times 10^{17}$  & 2.432  & $1.275 \times 10^{17}$  \\
- $\rho X(\isotm{C}{12})$     & $1.807 \times 10^{17}$  & 1.846  & $5.026 \times 10^{16}$  & 1.605  & $1.653 \times 10^{16}$  \\
- $\rho X(\isotm{O}{16})$     & $1.905 \times 10^{14}$  & 1.500  & $6.737 \times 10^{13}$  & 1.439  & $2.484 \times 10^{13}$  \\
- $\rho X(\isotm{Fe}{56})$    & $2.735 \times 10^{8}$   & 2.057  & $6.571 \times 10^{7}$   & 2.474  & $1.183 \times 10^{7}$   \\
-\enddata
-\end{deluxetable}
-
-\subsection{NSE convergence test problem}
-
-To test the coupling of the NSE table to the hydrodynamics, we run a
-similar problem as above, except with the thermodynamic conditions
-appropriate for the matter to be in NSE.  The initial conditions are:
-\begin{align}
-\rho &= \rho_0 \\
-T &= T_0 \left [ 1 + (\delta T) e^{-(r/\lambda)^2} \cos^6(\pi r/L) \right ] \\
-Y_e &= (Y_e)_0 \left [ 1 + (\delta Y_e) e^{-(r/\lambda)^2} \cos^6(\pi r/L) \right ] 
-\end{align}
-where $r$ is the distance from the center of the domain and the
-remaining parameters are listed in Table~\ref{table:nse}.  This has
-$Y_e$ varying between $[0.47, 0.5]$ initially\MarginPar{check}.  The mass fractions
-and remaining composition variables are then initialized from the NSE table, as described in section~\ref{sec:initialization}.
-
-The domain has a size $[0, L]^2$, and the timestep is fixed as:
-\begin{equation}
-\Delta t = 5\times 10^{-4} \left ( \frac{128}{N} \right )~\mathrm{s}
-\end{equation}
-where $N$ is the number of zones in each direction.
-
-\begin{deluxetable}{lcc}
-\tablecaption{\label{table:nse} NSE convergence test problem.}
-\tablehead{\colhead{parameter} & \colhead{value}}
-\startdata
-$\rho_0$ & $5\times 10^8~\gcc$ \\
-$T_0$    & $4\times 10^9~\mathrm{K}$ \\
-$(\delta T)$ & $0.2$ \\
-$(Y_e)_0$   & $0.5$ \\
-$(\delta Y_e)$ & $-0.05$ \\
-$\lambda$   & $2\times 10^7~\mathrm{cm}$ \\
-$L$         & $10^8~\mathrm{cm}$ \\
-\enddata
-\end{deluxetable}
-
-\begin{deluxetable}{lllllll}
-\tablecaption{\label{table:nse_strang_methodI} Convergence ($L_1$ norm) for the NSE convergence
-problem using Strang splitting + Method I.}
-\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
-\startdata
- $\rho$                      & $4.301 \times 10^{19}$  & 1.348  & $1.690 \times 10^{19}$  & 1.166  & $7.532 \times 10^{18}$  \\
- $\rho u$                    & $1.866 \times 10^{28}$  & 1.429  & $6.929 \times 10^{27}$  & 1.101  & $3.230 \times 10^{27}$  \\
- $\rho v$                    & $1.866 \times 10^{28}$  & 1.429  & $6.929 \times 10^{27}$  & 1.101  & $3.230 \times 10^{27}$  \\
- $\rho E$                    & $1.781 \times 10^{38}$  & 1.053  & $8.585 \times 10^{37}$  & 1.021  & $4.231 \times 10^{37}$  \\
- $\rho e$                    & $1.781 \times 10^{38}$  & 1.053  & $8.583 \times 10^{37}$  & 1.021  & $4.230 \times 10^{37}$  \\
- $T$                         & $6.853 \times 10^{21}$  & 1.044  & $3.325 \times 10^{21}$  & 1.018  & $1.642 \times 10^{21}$  \\
- $\rho Y_e$                  & $2.442 \times 10^{19}$  & 1.442  & $8.990 \times 10^{18}$  & 1.212  & $3.880 \times 10^{18}$  \\
- $\rho \bar{A}$              & $1.055 \times 10^{22}$  & 1.032  & $5.158 \times 10^{21}$  & 1.022  & $2.541 \times 10^{21}$  \\
- $\rho (B/A)$                & $4.512 \times 10^{20}$  & 1.386  & $1.727 \times 10^{20}$  & 1.207  & $7.479 \times 10^{19}$  \\
-\enddata
-\end{deluxetable}
-
-\begin{deluxetable}{lllllll}
-\tablecaption{\label{table:nse_strang_methodII} Convergence ($L_1$ norm) for the NSE convergence
-problem using Strang splitting + Method II.}
-\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
-\startdata
- $\rho$                      & $4.522 \times 10^{19}$  & 1.659  & $1.432 \times 10^{19}$  & 1.542  & $4.917 \times 10^{18}$  \\
- $\rho u$                    & $1.883 \times 10^{28}$  & 1.681  & $5.873 \times 10^{27}$  & 1.524  & $2.042 \times 10^{27}$  \\
- $\rho v$                    & $1.883 \times 10^{28}$  & 1.681  & $5.873 \times 10^{27}$  & 1.524  & $2.042 \times 10^{27}$  \\
- $\rho E$                    & $5.933 \times 10^{37}$  & 1.613  & $1.940 \times 10^{37}$  & 1.411  & $7.293 \times 10^{36}$  \\
- $\rho e$                    & $5.931 \times 10^{37}$  & 1.613  & $1.939 \times 10^{37}$  & 1.411  & $7.291 \times 10^{36}$  \\
- $T$                         & $9.885 \times 10^{20}$  & 1.207  & $4.281 \times 10^{20}$  & 1.058  & $2.057 \times 10^{20}$  \\
- $\rho Y_e$                  & $2.379 \times 10^{19}$  & 1.678  & $7.431 \times 10^{18}$  & 1.578  & $2.489 \times 10^{18}$  \\
- $\rho \bar{A}$              & $5.141 \times 10^{21}$  & 1.230  & $2.191 \times 10^{21}$  & 1.087  & $1.032 \times 10^{21}$  \\
- $\rho (B/A)$                & $4.149 \times 10^{20}$  & 1.693  & $1.283 \times 10^{20}$  & 1.562  & $4.345 \times 10^{19}$  \\
-\enddata
-\end{deluxetable}
-
 \subsection{Detonation}
@@
 
-We presented a simplified spectral deferred corrections scheme for coupling
+We presented a simple spectral deferred corrections scheme for coupling
 hydrodynamics and reactions.
-
-Future work is to extend this methodology to MHD.
 
@@
 \begin{equation}
-K = \frac{1}{2} u^2
+K = \frac{1}{2} |\Ub|^2
 \end{equation}
@@
 \end{equation}
-and write the kinetic energy as $K = u^2/2$.  We get the inverse
+and write the kinetic energy as $K = |\Ub|^2/2$.  We get the inverse
 (computed via SymPy, see the included Jupyter notebook) is:

### 2020-10-26 simplified_sdc_intro/paper.tex
starting to edit strang section

@@
 
-We solve the Euler equations for compressible, reacting flow.  We'll
-focus on one-dimensional flow for this description, and the system
-appears as:
+We solve the Euler equations for compressible, reacting flow.  For ease
+of exposition we describe the one-dimensional case;
+multidimensional extensions are a straightforward modification to
+include the CTU hydrodynamics scheme.
+The system appears as:
 \begin{equation}
@@
 \end{equation}
-where $\Gamma_1$ is an adiabatic index, $\Gamma_1 = d\log p/d\log\rho
-|_s$.  Note, the primitive state has two thermodynamic quantities, $p$
+where $\Gamma_1$ is an adiabatic index,
+$\Gamma_1 = d\log p/d\log\rho|_s$ at constant entropy.
+Note, the primitive state has two thermodynamic quantities, $p$
 and $(\rho e)$, to more efficiently handle the general equation of
@@
 \begin{equation}
-  \Sq = \Sqhydro + \Rb(q)
+  \Sq = \Sqhydro + \Rb(\qb)
 \end{equation}
+\MarginPar{Do you want to use $\Sc(\qb)$ and $\Hb(\qb)$ to keep notation consistent with the conserved variables, since you are already doing that with $\Rb(\qb)$?}
 with
 \begin{equation}
-\Rb(q) = \left ( \begin{array}{c}
+\Rb(\qb) = \left ( \begin{array}{c}
      0 \\
@@
 reaction sources are stiff, and need to be integrated using implicit
-methods for stabilty.  Operator splitting is traditionally employed
-here, and we compare to it in this paper.
+methods for stabilty.  Operator splitting (e.g., Strang) is traditionally employed
+here, and is used as a benchmark for comparison in this paper.
 
@@
 
-In Strang splitting, we first evolve the system without advection
-through $\Delta t/2$, then do the advective update without reactions
-through $\Delta t$, and finally do the last reactive update through $\Delta t/2$.
+In Strang splitting, we first integrate the system with reactions terms only (no advection)
+over $\Delta t/2$, then integrate the advection terms only (no reactions) over $\Delta t$,
+and finally integrate the reaction terms only over $\Delta t/2$.
 
@@
 \end{equation}
-where $K$ is the kinetic energy, $K = |u|^2/2$.  Since the velocity is
-unchanged by reactions, this means that
+where $K$ is the kinetic energy, $K = |u|^2/2$.  Since the density and velocity
+are unchanged by reactions, we consider the following:
 \begin{equation}
@@
 \end{equation}
-The reaction rates are typically expressed as $\omegadot_k(\rho, T,
-X_k)$, when we evolve this system.  This requires an EOS inversion to get $T$ from $e$ each
+The reaction rates are typically expressed as $\omegadot_k(\rho, T, X_k)$
+when we evolve this system.  This requires an EOS inversion to get $T$ from $e$ each
 time we need to evaluate the reactive terms.
-
-Normally we want to integrate the temperature.  We can write the
-energy equation as:
+We choose to instead integrate the temperature.
+We can write the energy equation as:
 \begin{equation}
@@
 \end{equation}
-The system of equations \ref{eq:strang:T} and \ref{eq:strang:X} are
-solved using an implicit ODE solver designed for stiff systems of
-equations.  We use VODE~\citep{vode}.
+We integrate Equations (\ref{eq:strang:T}) and (\ref{eq:strang:X}) using
+an implicit ODE solver designed for stiff systems of
+equations, VODE~\citep{vode}.
 
@@
 benefits from the larger timestep that method can take (when done with
-full corner coupling) as compared to a method-of-lines approach (see
+full corner coupling, the advective CFL condition is unity)
+as compared to a method-of-lines approach (see
 \citealt{ppmunsplit}).  In addition to the standard CFL timestep
@@
 timestep limiters during the step.  This means that equations
-\ref{eq:dt:nuce} and \ref{eq:dt:nucX} are not reactive, but instead
+(\ref{eq:dt:nuce}) and (\ref{eq:dt:nucX}) are not reactive, but instead
 guaranteed to be met throughout the simulation, because the step is
@@
     We update the state by doing the integral in equation
-    \ref{eq:integral:simplesdc}.  Since we are approximating the
+    (\ref{eq:integral:simplesdc}).  Since we are approximating the
     advective term as piecewise constant in time, we can simply use an
@@
     \rightarrow \qb^{n+1,(k)}$, we can construct $\Ic^{(k)}_{\qb}$ as given
-    in Eq.~\ref{eq:Iq}.
+    in Equation (\ref{eq:Iq}).
     Putting all of this together, we see:
@@
 composition of the larger 127 nuclei network is mapped into the 19
-isotopes we carry according to table~\ref{table:nuclei}.
+isotopes we carry according to Table \ref{table:nuclei}.
 

### 2020-10-26 simplified_sdc_intro/paper.tex
more comments

@@
 \begin{equation}
-  \Sc(\Uc) = \Hb(\Uc) + \Rb(\Uc)
+  \Sc(\Uc) = \Hb(\Uc) + \Rb(\Uc).
 \end{equation}
- The reactive
-sources take the form:
+\MarginPar{What is $\Hb(\Uc)$?}
+ The reactive sources take the form:
 \begin{equation}
@@
   \Sq = \Sqhydro + \Rb(\qb)
-\end{equation}
-\MarginPar{Do you want to use $\Sc(\qb)$ and $\Hb(\qb)$ to keep notation consistent with the conserved variables, since you are already doing that with $\Rb(\qb)$?}
+\end{equation}\MarginPar{Can we use $\Sc(\qb)$ instead of $\Sq$ and $\Hb(\qb)$ instead of $\Sqhydro$ to keep notation consistent with the conserved variables, since you are already doing that with $\Rb(\qb)$?}
 with
@@
 \begin{equation}
-\Uc^{n+1} = \Uc^n + \int_t^{t+\Delta t} \left [ \Advs{\Uc} + \Rb(\Uc) \right ] dt
+\Uc^{n+1} = \Uc^n + \int_t^{t+\Delta t} \left [ \Advs{\Uc} + \Rb(\Uc) \right ] dt,
 \end{equation}
+where $\Advs{\Uc} =  \frac{\partial \Fb^\xv (\Uc)}{\partial x} + \Hb(\Uc)$.
 For the simplified SDC method we explore here, we will make the
@@
     \begin{equation}
-      \Ic^{n+1/2,(-1)}_{\qb} = \Ic^{n-1/2,(\smax-1)}_{\qb}
+      \Ic^{n+1/2,(0)}_{\qb} = \Ic^{n-1/2,(\smax)}_{\qb}
     \end{equation}
+    \MarginPar{changed to 1-based $k$ iteration for aesthetics}
   %% \item Solve the Poisson problem for the initial gravitational potential:
@@
 
-  Iterate from $k = 0, \ldots, \smax-1$.  For second-order accuracy,
+  Iterate from $k = 1, \ldots, \smax$.  For second-order accuracy,
   $\smax = 2$ is sufficient.  In addition to denoting the time-level
@@
 
-    \item predict $\qb$ to the interfaces at $n+1/2$ using the CTU PPM
+    \item predict $\qb$ to the interfaces at $t^{n+1/2}$ using the CTU PPM
       method.  The source terms, $\Sq$, used in the prediction are:
@@
       using the previous iteration:
+\MarginPar{Formally both $t^n$ or $t^{n+1/2}$ sources give you 2nd-order since they are CTU predictor source terms.  I've never found this type of iteration for this term to amount to any benefit}
       \begin{equation}
@@
 
-    \item construct the advective update terms, first ignoring the hydrodynamics sources:
-      $\Advt{\Uc}^{n+1/2,(k)}_{i}$, e.g., as:
+    \item construct the advective update terms, $\Advt{\Uc}^{n+1/2,(k)}_{i}$,
+      by first ignoring the hydrodynamics sources,
       \begin{align}
@@
     new state with advection only\footnote{For a source like gravity, this update can be done first for $\rho$ and then define the new momentum source using $\rho^{\star\star}$ and likewise for energy}, as:
+\MarginPar{I'm confused what happens with the $\Uc^{\star\star}$ state; I don't see it below.  Is The $\Shydro^{n+1/2}$ used below suppose to be from the $\Uc^{\star\star}$ state?}
     \begin{equation}

### 2020-10-26 simplified_sdc_intro/paper.tex
more comments, typo fixes

@@
 \newcommand{\avg}[1]{{\left \langle #1 \right \rangle}}
+
+\newcommand{\nse}[1]{{\mathrm{NSE}( #1 )}}
+
+\newcommand{\rhonse}{{\rho_\mathrm{nse}}}
+\newcommand{\tnse}{{T_\mathrm{nse}}}
+\newcommand{\Anse}{{A_\mathrm{nse}}}
+\newcommand{\Bnse}{{B_\mathrm{nse}}}
+
+\newcommand{\out}{{\rm out}}
+\newcommand{\inp}{{\rm in}}
 
@@
 CTU hydrodynamics construction and a largely similar ODE integration
-scheme from the original Strang split approach,
-making this method easier to add to existing simulation codes.
-Overall, compared to the original Strang split approach, this new
-approach is more computationally efficient, allows for a larger
-time step, and reduces overall integration error.
-However, it is restricted to second-order accuracy overall.  We term
-this algorithm the ``simplified SDC method''.  We use the
-\castro\ hydrodynamics code \cite{castro} for all of our numerical
-experiments, and all of the code to reproduce the results in this
-paper is in the \castro\ github
+scheme, making this method easier to add to existing simulation codes.
+It also extends to adaptive mesh refinement with subcycling in a
+straightforward manner, avoiding the complications described in
+\citep{mccorquodalecolella} needed to fill ghost cells when using
+method-of-lines integration.  However, it is restricted to
+second-order accuracy overall.  We term this algorithm the
+``simplified SDC method''.  We use the \castro\ hydrodynamics code
+\citep{castro} for all of our numerical experiments, and all of the
+code to reproduce the results in this paper are in the \castro\ github
 repository\footnote{\url{https://github.com/amrex-astro/Castro/}}.
@@
 \end{equation}
-\MarginPar{What is $\Hb(\Uc)$?}
+\MarginPar{Detail out what is $\Hb(\Uc)$?}
  The reactive sources take the form:
@@
   \Rb(\Uc) = \left ( \begin{array}{c}
-     0 \\ \rho \omegadot_k \\ 0 \\ \rho \dot{S}
+     0 \\
+     \rho \omegadot(X_k) \\
+     \rho \omegadot(\alpha_l) \\
+     0 \\
+     \rho \dot{S}
   \end{array} \right )
 \end{equation}
-where $\omegadot_k$ is the creation rate for species $k$ and $\dot{S}$
+where $\omegadot(X_k)$ is the creation rate for species $k$, $\omegadot(\alpha_l)$ is the creation rate for
+auxiliary composition variable $l$, and $\dot{S}$
 is the energy generation rate per unit mass.
@@
 sources (now in terms of the primitive variables) and reaction terms,
+\MarginPar{Can we use $\Sc(\qb)$ instead of $\Sq$ and $\Hb(\qb)$ instead of $\Sqhydro$ to keep notation consistent with the conserved variables, since you are already doing that with $\Rb(\qb)$?}
 \begin{equation}
   \Sq = \Sqhydro + \Rb(\qb)
-\end{equation}\MarginPar{Can we use $\Sc(\qb)$ instead of $\Sq$ and $\Hb(\qb)$ instead of $\Sqhydro$ to keep notation consistent with the conserved variables, since you are already doing that with $\Rb(\qb)$?}
+\end{equation}
 with
 \begin{equation}
-\Rb(\qb) = \left ( \begin{array}{c}
+\label{eq:prim_sources}
+\Rb(q) = \left ( \begin{array}{c}
      0 \\
-     \omegadot_k \\
+     \omegadot(X_k) \\
+     \omegadot(\alpha_l) \\
      0 \\
@@
     \end{equation}
-    \MarginPar{changed to 1-based $k$ iteration for aesthetics}
+    \MarginPar{changed to 1-based $k$ iteration for aesthetics?  I can switch it back if you prefer ``-1'' to be the starting point instead of ``0''.}
   %% \item Solve the Poisson problem for the initial gravitational potential:
@@
     Now the conservative hydrodynamics source terms are computed by first updating to the
-    new state with advection only\footnote{For a source like gravity, this update can be done first for $\rho$ and then define the new momentum source using $\rho^{\star\star}$ and likewise for energy}, as:
+    new state with advection and the old-time source term applied for the full $\Delta t$ as:
 \MarginPar{I'm confused what happens with the $\Uc^{\star\star}$ state; I don't see it below.  Is The $\Shydro^{n+1/2}$ used below suppose to be from the $\Uc^{\star\star}$ state?}
     \begin{equation}
-      \Uc^{\star\star} = \Uc^n + \Delta t \Advt{\Uc}^{n+1/2,(k)}
+      \Uc^{\star\star} = \Uc^n + \Delta t \Advt{\Uc}^{n+1/2,(k)} + \Delta t \Hb(\Uc^n)
     \end{equation}
@@
     %% \end{equation}
-
-    The final advective update term is then:
-    \begin{equation}
-      \Adv{\Uc}^{n+1/2,(k)}_{i} = \Advt{\Uc}^{n+1/2,(k)}_{i} + \Shydro^{n+1/2}
-    \end{equation}
-    with
-    \begin{equation}
-      \Shydro^{n+1/2} = \left ( \begin{array}{c}
-                     0 \\ 0 \\
-                    {\bf S}_{\rho\Ub}^{n+1/2} \cdot \ex \\
-                    S_{\rho E}^{n+1/2} \end{array} \right )
-    \end{equation}
+    We then evaluate the source terms with $\Uc^{\star\star}$ and
+    correct the advective term so that we have a time-centered
+    source: \MarginPar{are we actually doing this correctly for SDC?}
+    The final advective update term is then\footnote{For a source like gravity, this update can be done first for $\rho$ and then define the new momentum source using $\rho^{\star\star}$ and likewise for energy}:
+    \begin{equation}
+      \Adv{\Uc}^{n+1/2,(k)}_{i} = \Advt{\Uc}^{n+1/2,(k)}_{i} +
+      \frac{\Delta t}{2} \left (\Shydro(\Uc^n) + \Shydro(\Uc^{\star\star}) \right )
+    \end{equation}
+
+    %% with
+    %% \begin{equation}
+    %%   \Shydro^{n+1/2} = \left ( \begin{array}{c}
+    %%                  0 \\ 0 \\ 0 \\
+    %%                 {\bf S}_{\rho\Ub}^{n+1/2} \cdot \ex \\
+    %%                 S_{\rho E}^{n+1/2} \end{array} \right )
+    %% \end{equation}
 
@@
     integrating the conserved variables and the state sees the effect
-    of advection as we integrate the reactions.  The ODE form of this
-    is
-    \begin{equation}
-      \odt{\Uc} = \Adv{\Uc}^{n+1,(k)} + \Rb(\Uc)
-    \end{equation}
+    of advection as we integrate the reactions.  So rather than use
+    $d\Uc/dt = \Rb(\Uc)$, the ODE form we use is
+    \begin{equation}
+      \odt{\Uc} = \Adv{\Uc}^{n+1/2,(k)} + \Rb(\Uc)
+    \end{equation}
+    This integration begins with $\Uc^n$ and results in $\Uc^{n+1,(k)}$.
 
@@
 
-    Note that our advection terms are piecewise constant \MarginPar{need to think about this more}
-    approximations, so when we integrate our system, their
-    contributions will be linear in time.  Since $(\rho X_k)$ sees a
-    linear advective term, $\rho$ does as well.  $(\rho \Ub)$ and
-    $(\rho E)$ will also have linear contributions in time from
-    advection.  We can see that the internal energy will also have a
-    linear contribution by recognizing that the kinetic energy term,
-    $K = (\rho |\Ub|^2)/\rho$ is linear in time.
-    Since the reactions don't affect $\rho$ and $(\rho \Ub)$, we can
-    algebraically update these using this piecewise linear behavior:
+    Note that our advection terms are piecewise
+    constant \MarginPar{need to think about this more} approximations,
+    so when we integrate our system, their contributions will be
+    linear in time.  Since $(\rho X_k)$ sees a linear advective term,
+    $\rho$ does as well.  $(\rho u)$ will also have linear
+    contributions in time from advection.  Since the reactions don't
+    affect $\rho$ and $(\rho u)$, we can algebraically update these
+    using this piecewise linear behavior:
     \begin{align}
-      \rho(t) &= \rho^n + \Adv{\rho}^{n+1/2,(k)} \, t \\
-      (\rho \Ub)(t) &= (\rho \Ub)^n + \Adv{\rho \Ub}^{n+1/2,(k)} \, t
+      \rho(t) &= \rho^n + \Advss{\rho}^{n+1/2,(k)} \, t \\
+      (\rho u)(t) &= (\rho u)^n + \Advss{\rho u}^{n+1/2,(k)} \, t
     \end{align}
-
-    At the end of this integration, we have the state at the next iteration,
-    $\Uc^{n+1,(k+1)}$
 
@@
 
-The NSE table provides:
+The NSE table requires:
+\begin{equation}
+  \label{eq:aux:ye}
+  Y_e = \sum_k \frac{Z_k X_k}{A_k}
+\end{equation}
+(where $A_k$ and $Z_k$ are the atomic weight and atomic number of nucleus $k$) and provides
 \begin{align}
-Y_e &= \sum_k \frac{Z_k X_k}{A_k} \\
+\label{eq:aux:abar}
 \bar{A} &= \left [ \sum_k \frac{X_k}{A_k} \right ]^{-1} \\
-\frac{B}{A} &= \sum_k \frac{B_k X_k}{A_k}
+\label{eq:aux:bea}
+\left (\frac{B}{A} \right ) &= \sum_k \frac{B_k X_k}{A_k}
 \end{align}
@@
 we store these 3 quantities as auxiliary data that is carried along
-with the rest of the fluid state in each zone.
-
+with the rest of the fluid state in each zone.  The table also returns
+values of the mass fractions mapped onto the 19-isotopes we carry, $X_k$, and the time-derivative of $Y_e$, $dY_e/dt$.
+We'll use the notation:
+\begin{equation}
+\nse{\rho,T,Y_e} \rightarrow \bar{A}, X_k, (B/A), dY_e/dt
+\end{equation}
+to represent the NSE table call and its inputs and outputs.
+
+We can derive evolution equations for each of these composition quantities as:
+\begin{align}
+\frac{DY_e}{Dt} &= \sum_k \frac{Z_k}{A_k} \frac{DX_k}{Dt} = \sum_k \frac{Z_k}{A_k} \omegadot_k \\
+\frac{D\bar{A}}{Dt} &= -\bar{A}^2 \sum_k \frac{1}{A_k} \frac{DX_k}{Dt} = -\bar{A}^2 \sum_k \frac{1}{A_k} \omegadot_k \\
+\frac{D}{Dt} \left (\frac{B}{A} \right ) &= \sum_k \frac{B_k}{A_k} \frac{DX_k}{Dt} = \sum_k \frac{B_k}{A_k} \omegadot_k
+\end{align}
 For Strang split coupling of hydro and reactions, $DX_k/Dt = 0$,
-and our evolution equations are:
-\begin{align}
-\frac{DY_e}{Dt} &= \sum_k \frac{Z_k}{A_k} \frac{DX_k}{Dt} = 0 \\
-\frac{D}{Dt} \frac{1}{\bar{A}} &= - \frac{1}{\bar{A}^2} \frac{D\bar{A}}{Dt} = \sum_k \frac{1}{A_k} \frac{DX_k}{Dt} = 0 \rightarrow \frac{D\bar{A}}{Dt} = 0 \\
-\frac{D}{Dt} \left (\frac{B}{A} \right ) &= \sum_k \frac{B_k}{A_k} \frac{DX_k}{Dt} = 0
-\end{align}
-Therefore each of these auxiliary equations obeys an advection
+therefore each of these auxillary equations obeys an advection
 equation in the hydro part of the advancement.  In the SDC algorithm,
 there will be a reactive source (an $\Ic_q$) for each of these that is
-computed in the same manner as above.
+computed in the same manner as above.  We note that our NSE table
+provides the evolution of $Y_e$ due to reactions ($DY_e/Dt$) directly.
 
 The compositional quantities
-it carries, $\bar{A}$ and $Y_e$, are not representable from the 19
+it carries, $\bar{A}$ and $Y_e$ are not representable from the 19
 isotopes we carry in the main network. For this reason, when we are
@@
 
+\subsection{Initialization}
+
+We initializing the mass fractions, $X_k$, and then compute the
+electron fraction, $Y_e$, from them.  For the two other composition
+quantities we carry, $\bar{A}$, and $(B/A)$, we need values that are
+consistent with the value of $Y_e$ and the nuclei stored in the NSE
+table.  Therefore, if the thermodynamic conditions put us in NSE
+(using the conditions defined below), then we get $\bar{A}$ and
+$(B/A)$ from the NSE table.  Otherwise, we compute these directly from
+$X_k$.
+
+\subsection{NSE condition}
+
+We treat a zone as being in NSE if the density and temperature exceed
+some threshold and the composition is mainly $\alpha$ and Fe-group
+nuclei, where the Fe-group nuclei are \isot{Cr}{48}, \isot{Fe}{52},
+\isot{Fe}{54}, and \isot{Ni}{56}.  The full condition is:
+\begin{align}
+\rho &> \rhonse \\
+T &> \tnse \\
+X(\isotm{C}{12}) &< \Anse \\
+X(\isotm{He}{4}) + \sum_{k \in \mathrm{Fe-group}} X_k &> \Bnse
+\end{align}
+Typical values are $\rhonse = 3\times 10^8~\gcc$, $T_\mathrm{nse} =
+3\times 10^9~\mathrm{K}$, and $\Anse = 0.12$ and $\Bnse=
+0.88$---these are the values used in \citet{ma:2013}.
+
+\subsection{Strang-split algorithm for NSE}
+
 For Strang splitting, the basic flow of using the aprox19 + NSE network
@@
 \begin{itemize}
-\item {\em Initialize the problem}
-
-  In addition to initializing the mass fractions, $X_k$, we also fill
-  the auxiliary composition data, $Y_e$, $\bar{A}$, and $(B/A)$, using
-  these $X_k$.
 
@@
   \begin{itemize}
+
     \item For a zone that is in NSE:
 
+      The goal is to update the composition and thermodynamics due to the 
+      change in the nuclei abundances over the (half-)timestep.  There are
+      a few approaches we can do here.
+
       \begin{itemize}
-    \item use $\rho$, $T$, and $Y_e$ to call the table. This
-      returns: $dY_e/dt$, $(B/A)_{\rm out}$, $\bar{A}_{\rm out}$,
-      and the $(X_k)_{\rm out}$ from the NSE network grouped into the 19 isotopes
-      we carry as part of aprox19.
-
-    \item update $Y_e$:
+      \item Method I: {\em Use the time-level $n$ information for the update}.
+
+        This is the approach used in \citet{ma:2013}.  We call the table with the 
+        input state:
+        \begin{equation}
+          \nse{\rho, T, (Y_e)_{\rm in}} \rightarrow \bar{A}_\out, (X_k)_\out, (B/A)_\out, dY_e/dt
+        \end{equation}
+
+        We can now update $Y_e$:
         \begin{equation}
@@
         \end{equation}
-
-  \item update the auxiliary state value of $\bar{A}$ to be simply
-    $\bar{A}_{\rm out}$ from the table and the mass fractions
+        and take the auxiliary state values of $\bar{A}$ and $(B/A)$ to be simply
+        the values from the table, $\bar{A}_\out$ and $(B/A)_\out$, and the mass fractions
         $X_k$ to be the $(X_k)_{\rm out}$ returned from the table.
 
-  \item compute the energy generation rate, $\Sdot$ as:
-    \begin{equation}
-      \Sdot = \eta \left [ \left ( \frac{B}{A} \right )_{\rm out} -
+        Finally, we compute the energy generation rate, $\Sdot$ as:
+        \begin{equation}
+          \label{eq:nse_energy}
+          \Sdot = \left [ \left ( \frac{B}{A} \right )_{\rm out} -
             \left ( \frac{B}{A} \right )_{\rm in} \right ] N_A \frac{1}{\Delta t}
         \end{equation}
-    where $\eta$ is an inertia term < 1 to prevent the energy changing
-    too much in one set.
-
-  \item set the auxiliary state value of $(B/A)$ as:
-    \begin{equation}
-      \left (\frac{B}{A}\right )_{\rm out} = \left (\frac{B}{A}\right )_{\rm in} + \eta \left [ \left ( \frac{B}{A}\right )_{\rm out} - \left (\frac{B}{A} \right )_{\rm in} \right]
-    \end{equation}
+
+        We note that \citet{ma:2013} modified Eq.~\ref{eq:nse_energy}
+        to take use only a fraction (0.3) of the energy change from
+        the table, to avoid a potential instability that can arise if
+        too much energy is deposited in a zone in a single timestep
+        (they called this an inertia term).  We prefer to deal with
+        the issue of too much energy being deposited in a zone in one
+        step through the retry mechanism already built into \castro.
+
+      \item Method II: {\em predictor-corrector.}
+
+        This starts out the same as the above method:
+        \begin{equation}
+          \nse{\rho, T, (Y_e)_{\rm in}} \rightarrow \bar{A}^\star, (X_k)^\star, (B/A)^\star, dY_e/dt
+        \end{equation}
+        but we indicate with a $\star$ that most of those values are
+        provisional, and we will seek a better value in the corrector step.
+        We only update $Y_e$ from this call:
+        \begin{equation}
+          (Y_e)_{\rm out} = (Y_e)_{\rm in} + \Delta t \frac{dY_e}{dt}
+        \end{equation}
+        and then we use the table again, but with the updated $Y_e$:
+        \begin{equation}
+          \nse{\rho, T, (Y_e)_\out} \rightarrow \bar{A}_\out, (X_k)_\out, (B/A)_\out, dY_e/dt
+        \end{equation}
+        This corrects the composition so that it is consistent with the updated
+        $Y_e$.
+We use these values $\bar{A}_\out$, $(X_k)_\out$, and $(B/A)_\out$ to then complete the update, computing the energy as in Eq.~\ref{eq:nse_energy}.
+\MarginPar{we could imagine looping over the EOS to get an updated T and then the table to get the updated $\bar{A}$, to see if that converges?}
 
@@
 
-We use the following conditions to determine if we are in NSE:
+\subsection{SDC-NSE Coupling}
+
+With SDC evolution, when we are in NSE, we need to do the advective
+and reactive updates together.  With the tabulated NSE reaction
+values, we get instantaneous values of the reaction sources, but the only
+time derivative we have is for $Y_e$.  We will
+use a basic predictor-corrector update:
+\begin{equation}
+\Uc^\star = \Uc^n + \Delta t \Adv{\Uc}^{n+1/2} + \Delta t \Rb (\Uc^n)
+\end{equation}
+Calling the NSE table on the starting state gives use:
+\begin{equation}
+\nse{\rho^n, T^n, (Y_e)^n} \rightarrow \bar{A}^n, (X_k)^n, (B/A)^n, (dY_e/dt)^n
+\end{equation}
+To get the reactive source, we need to use the iteratively-lagged information
+from $\Ic_q^{n+1/2,(k-1)}$:
+\begin{equation}
+\Rb(\Uc^n) \approx \left (\begin{array}{c}
+  0 \\
+  \rho^n \Ics_q(\omegadot(X_k)) \\
+  \rho^n \Ics_q(\omegadot(\alpha_l)) \\
+  0 \\
+  \Ics_q(\rho e) 
+\end{array}
+\right )
+\end{equation}
+Since we have $(dY_e/dt)^n$ from the table, we replace the source for $Y_e$ in the above
+$\Rb(\Uc^n)$ as:
+\begin{equation}
+R(\rho Y_e) = \rho (dY_e/dt)^n
+\end{equation}
+which gives an update for the electron fraction as:
+\begin{equation}
+(\rho Y_e)^\star = (\rho Y_e)^n + \Delta t \Advss{\rho Y_e}^{n+1/2} + \Delta t \rho (dY_e/dt)^n
+\end{equation}
+From this prediction, we get $\rho^\star$, $\alpha_l^\star$, and
+$e^\star$, which allows us to call the EOS and get an updated
+temperature, $T^\star$. \MarginPar{Andy suggests time-centering $dY_e/dt$}
+We can then again call the NSE table:
+\begin{equation}
+\nse{\rho^\star, T^\star, (Y_e)^\star} \rightarrow \bar{A}^{n+1,(k)}, (X_k)^{n+1,(k)}, (B/A)^{n+1,(k)}, (dY_e/dt)
+\end{equation}
+where we use the outputs to define the new values of the composition variables.
+Since the density and momenta don't have reactive sources, we just use the $\Uc^\star$ state
+as their new values:
 \begin{align}
-\rho &> \rho_\mathrm{nse} \\
-T &> T_\mathrm{nse} \\
-X(\isotm{C}{12}) &< A_\mathrm{nse} \\
-X(\isotm{He}{4}) + \sum_{k \in {\rm Fe-group}} X_k &> B_\mathrm{nse} 
+\rho^{n+1,(k)} &= \rho^\star \\
+(\rho u)^{n+1,(k)} &= (\rho u)^\star
 \end{align}
-where the Fe-group nuclei are \isot{Cr}{48}, \isot{Fe}{52}, \isot{Fe}{54}, and \isot{Ni}{56}.
+We then compute the energy release as
+\begin{equation}
+       (\rho \Sdot)^{n+1/2} = \left [ \rho^{n+1,(k)} \left ( \frac{B}{A} \right )^{n+1,(k)} -
+                            \rho^n \left ( \frac{B}{A} \right )^n \right ] N_A \frac{1}{\Delta t}
+\end{equation}\MarginPar{is this right?}
+and then do the final update of the total energy:
+\begin{equation}
+(\rho E)^{n+1,(k)} = (\rho E)^n + \Delta t \Advss{\rho E}^{n+1/2} + (\rho \Sdot)^{n+1/2}
+\end{equation}
 
@@
 
+\begin{deluxetable}{lllllll}
+\tablecaption{\label{table:react_convert_strang} Convergence ($L_1$ norm) for the reacting convergence problem using Strang splitting.}
+\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
+           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
+           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
+\startdata
+ $\rho$                      & $2.781 \times 10^{18}$  & 2.051  & $6.711 \times 10^{17}$  & 2.580  & $1.122 \times 10^{17}$  \\
+ $\rho u$                    & $6.781 \times 10^{26}$  & 2.445  & $1.245 \times 10^{26}$  & 2.907  & $1.659 \times 10^{25}$  \\
+ $\rho v$                    & $6.781 \times 10^{26}$  & 2.445  & $1.245 \times 10^{26}$  & 2.907  & $1.659 \times 10^{25}$  \\
+ $\rho E$                    & $2.465 \times 10^{35}$  & 2.333  & $4.892 \times 10^{34}$  & 2.650  & $7.794 \times 10^{33}$  \\
+ $\rho e$                    & $2.268 \times 10^{35}$  & 2.298  & $4.610 \times 10^{34}$  & 2.722  & $6.988 \times 10^{33}$  \\
+ $T$                         & $2.248 \times 10^{21}$  & 1.682  & $7.003 \times 10^{20}$  & 2.439  & $1.291 \times 10^{20}$  \\
+ $\rho X(\isotm{He}{4})$     & $2.862 \times 10^{18}$  & 2.027  & $7.023 \times 10^{17}$  & 2.554  & $1.196 \times 10^{17}$  \\
+ $\rho X(\isotm{C}{12})$     & $1.717 \times 10^{17}$  & 1.945  & $4.460 \times 10^{16}$  & 2.194  & $9.745 \times 10^{15}$  \\
+ $\rho X(\isotm{O}{16})$     & $1.718 \times 10^{14}$  & 1.648  & $5.482 \times 10^{13}$  & 1.898  & $1.471 \times 10^{13}$  \\
+ $\rho X(\isotm{Fe}{56})$    & $2.781 \times 10^{8}$   & 2.051  & $6.711 \times 10^{7}$   & 2.580  & $1.122 \times 10^{7}$   \\
+\enddata
+\end{deluxetable}
+
+\begin{deluxetable}{lllllll}
+\tablecaption{\label{table:react_convert_strang} Convergence ($L_1$ norm) for the reacting convergence problem using Strang splitting with $T$ evolution disabled.}
+\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
+           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
+           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
+\startdata
+ $\rho$                      & $2.735 \times 10^{18}$  & 2.057  & $6.571 \times 10^{17}$  & 2.474  & $1.183 \times 10^{17}$  \\
+ $\rho u$                    & $6.729 \times 10^{26}$  & 2.416  & $1.260 \times 10^{26}$  & 2.699  & $1.941 \times 10^{25}$  \\
+ $\rho v$                    & $6.729 \times 10^{26}$  & 2.416  & $1.260 \times 10^{26}$  & 2.699  & $1.941 \times 10^{25}$  \\
+ $\rho E$                    & $2.513 \times 10^{35}$  & 2.226  & $5.372 \times 10^{34}$  & 2.169  & $1.195 \times 10^{34}$  \\
+ $\rho e$                    & $2.292 \times 10^{35}$  & 2.194  & $5.011 \times 10^{34}$  & 2.253  & $1.051 \times 10^{34}$  \\
+ $T$                         & $2.321 \times 10^{21}$  & 1.606  & $7.623 \times 10^{20}$  & 1.993  & $1.915 \times 10^{20}$  \\
+ $\rho X(\isotm{He}{4})$     & $2.806 \times 10^{18}$  & 2.028  & $6.880 \times 10^{17}$  & 2.432  & $1.275 \times 10^{17}$  \\
+ $\rho X(\isotm{C}{12})$     & $1.807 \times 10^{17}$  & 1.846  & $5.026 \times 10^{16}$  & 1.605  & $1.653 \times 10^{16}$  \\
+ $\rho X(\isotm{O}{16})$     & $1.905 \times 10^{14}$  & 1.500  & $6.737 \times 10^{13}$  & 1.439  & $2.484 \times 10^{13}$  \\
+ $\rho X(\isotm{Fe}{56})$    & $2.735 \times 10^{8}$   & 2.057  & $6.571 \times 10^{7}$   & 2.474  & $1.183 \times 10^{7}$   \\
+\enddata
+\end{deluxetable}
+
+\begin{deluxetable}{lllllll}
+\tablecaption{\label{table:react_convert_strang} Convergence ($L_1$ norm) for the reacting convergence problem using SDC integration.}
+\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
+           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
+           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
+\startdata
+ $\rho$                      & $2.795 \times 10^{18}$  & 2.044  & $6.777 \times 10^{17}$  & 2.554  & $1.154 \times 10^{17}$  \\
+ $\rho u$                    & $6.798 \times 10^{26}$  & 2.449  & $1.245 \times 10^{26}$  & 2.889  & $1.680 \times 10^{25}$  \\
+ $\rho v$                    & $6.798 \times 10^{26}$  & 2.449  & $1.245 \times 10^{26}$  & 2.889  & $1.680 \times 10^{25}$  \\
+ $\rho E$                    & $2.452 \times 10^{35}$  & 2.352  & $4.804 \times 10^{34}$  & 2.742  & $7.179 \times 10^{33}$  \\
+ $\rho e$                    & $2.262 \times 10^{35}$  & 2.321  & $4.527 \times 10^{34}$  & 2.822  & $6.404 \times 10^{33}$  \\
+ $T$                         & $2.236 \times 10^{21}$  & 1.691  & $6.927 \times 10^{20}$  & 2.481  & $1.240 \times 10^{20}$  \\
+ $\rho X(\isotm{He}{4})$     & $2.879 \times 10^{18}$  & 2.020  & $7.096 \times 10^{17}$  & 2.529  & $1.229 \times 10^{17}$  \\
+ $\rho X(\isotm{C}{12})$     & $1.697 \times 10^{17}$  & 1.950  & $4.393 \times 10^{16}$  & 2.231  & $9.357 \times 10^{15}$  \\
+ $\rho X(\isotm{O}{16})$     & $1.686 \times 10^{14}$  & 1.659  & $5.338 \times 10^{13}$  & 1.957  & $1.375 \times 10^{13}$  \\
+ $\rho X(\isotm{Fe}{56})$    & $2.795 \times 10^{8}$   & 2.044  & $6.777 \times 10^{7}$   & 2.554  & $1.154 \times 10^{7}$   \\
+\enddata
+\end{deluxetable}
+
+\subsection{NSE convergence test problem}
+
+To test the coupling of the NSE table to the hydrodynamics, we run a
+similar problem as above, except with the thermodynamic conditions
+appropriate for the matter to be in NSE.  The initial conditions are:
+\begin{align}
+\rho &= \rho_0 \\
+T &= T_0 \left [ 1 + (\delta T) e^{-(r/\lambda)^2} \cos^6(\pi r/L) \right ] \\
+Y_e &= (Y_e)_0 \left [ 1 + (\delta Y_e) e^{-(r/\lambda)^2} \cos^6(\pi r/L) \right ] 
+\end{align}
+where $r$ is the distance from the center of the domain and the
+remaining parameters are listed in Table~\ref{table:nse}.  This has
+$Y_e$ varying between $[0.47, 0.5]$ initially\MarginPar{check}.  The mass fractions
+and remaining composition variables are then initialized from the NSE table, as described in section~\ref{sec:initialization}.
+
+The domain has a size $[0, L]^2$, and the timestep is fixed as:
+\begin{equation}
+\Delta t = 10^{-3} \left ( \frac{64}{N} \right )~\mathrm{s}
+\end{equation}
+where $N$ is the number of zones in each direction.
+
+\begin{deluxetable}{lcc}
+\tablecaption{\label{table:nse} NSE convergence test problem.}
+\tablehead{\colhead{parameter} & \colhead{value}}
+\startdata
+$\rho_0$ & $5\times 10^8~\gcc$ \\
+$T_0$    & $4\times 10^9~\mathrm{K}$ \\
+$(\delta T)$ & $0.2$ \\
+$(Y_e)_0$   & $0.5$ \\
+$(\delta Y_e)$ & $-0.05$ \\
+$\lambda$   & $2\times 10^7~\mathrm{cm}$ \\
+$L$         & $10^8~\mathrm{cm}$ \\
+\enddata
+\end{deluxetable}
+
+Figure~\ref{fig:nse_ye} shows the $Y_e$ profile for the Strang evolution case.
+
+\begin{figure}[t]
+\centering
+\plotone{ye_plot}
+\caption{\label{fig:nse_ye} Profile of $Y_e$ through the middle of the domain (slicing along $x$)
+for the $128^2$ Strang method II NSE test simulation.}
+\end{figure}
+
+\begin{deluxetable}{lllllll}
+\tablecaption{\label{table:nse_strang_methodI} Convergence ($L_1$ norm) for the NSE convergence
+problem using Strang splitting + Method I.}
+\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
+           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
+           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
+\startdata
+ $\rho$                      & $4.301 \times 10^{19}$  & 1.348  & $1.690 \times 10^{19}$  & 1.166  & $7.532 \times 10^{18}$  \\
+ $\rho u$                    & $1.866 \times 10^{28}$  & 1.429  & $6.929 \times 10^{27}$  & 1.101  & $3.230 \times 10^{27}$  \\
+ $\rho v$                    & $1.866 \times 10^{28}$  & 1.429  & $6.929 \times 10^{27}$  & 1.101  & $3.230 \times 10^{27}$  \\
+ $\rho E$                    & $1.781 \times 10^{38}$  & 1.053  & $8.585 \times 10^{37}$  & 1.021  & $4.231 \times 10^{37}$  \\
+ $\rho e$                    & $1.781 \times 10^{38}$  & 1.053  & $8.583 \times 10^{37}$  & 1.021  & $4.230 \times 10^{37}$  \\
+ $T$                         & $6.853 \times 10^{21}$  & 1.044  & $3.325 \times 10^{21}$  & 1.018  & $1.642 \times 10^{21}$  \\
+ $\rho Y_e$                  & $2.442 \times 10^{19}$  & 1.442  & $8.990 \times 10^{18}$  & 1.212  & $3.880 \times 10^{18}$  \\
+ $\rho \bar{A}$              & $1.055 \times 10^{22}$  & 1.032  & $5.158 \times 10^{21}$  & 1.022  & $2.541 \times 10^{21}$  \\
+ $\rho (B/A)$                & $4.512 \times 10^{20}$  & 1.386  & $1.727 \times 10^{20}$  & 1.207  & $7.479 \times 10^{19}$  \\
+\enddata
+\end{deluxetable}
+
+\begin{deluxetable}{lllllll}
+\tablecaption{\label{table:nse_strang_methodII} Convergence ($L_1$ norm) for the NSE convergence
+problem using Strang splitting + Method II.}
+\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
+           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
+           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
+\startdata
+ $\rho$                      & $4.522 \times 10^{19}$  & 1.659  & $1.432 \times 10^{19}$  & 1.542  & $4.917 \times 10^{18}$  \\
+ $\rho u$                    & $1.883 \times 10^{28}$  & 1.681  & $5.873 \times 10^{27}$  & 1.524  & $2.042 \times 10^{27}$  \\
+ $\rho v$                    & $1.883 \times 10^{28}$  & 1.681  & $5.873 \times 10^{27}$  & 1.524  & $2.042 \times 10^{27}$  \\
+ $\rho E$                    & $5.933 \times 10^{37}$  & 1.613  & $1.940 \times 10^{37}$  & 1.411  & $7.293 \times 10^{36}$  \\
+ $\rho e$                    & $5.931 \times 10^{37}$  & 1.613  & $1.939 \times 10^{37}$  & 1.411  & $7.291 \times 10^{36}$  \\
+ $T$                         & $9.885 \times 10^{20}$  & 1.207  & $4.281 \times 10^{20}$  & 1.058  & $2.057 \times 10^{20}$  \\
+ $\rho Y_e$                  & $2.379 \times 10^{19}$  & 1.678  & $7.431 \times 10^{18}$  & 1.578  & $2.489 \times 10^{18}$  \\
+ $\rho \bar{A}$              & $5.141 \times 10^{21}$  & 1.230  & $2.191 \times 10^{21}$  & 1.087  & $1.032 \times 10^{21}$  \\
+ $\rho (B/A)$                & $4.149 \times 10^{20}$  & 1.693  & $1.283 \times 10^{20}$  & 1.562  & $4.345 \times 10^{19}$  \\
+\enddata
+\end{deluxetable}
+
 \subsection{Detonation}

### 2020-10-29 simplified_sdc_intro/paper.tex
minor comments

@@
 values of the mass fractions mapped onto the 19-isotopes we carry, $X_k$, and the time-derivative of $Y_e$, $dY_e/dt$.
-We'll use the notation:
+We use the notation:
 \begin{equation}
@@
 to the auxiliary quantities, which is is easily computed as
+\MarginPar{might be worth spelling out exactly what inputs the NSE EOS needs to make things crystal clear}
 \begin{equation}
@@
 
-We initializing the mass fractions, $X_k$, and then compute the
-electron fraction, $Y_e$, from them.  For the two other composition
+We initialize the mass fractions, $X_k$, and then compute the
+electron fraction, $Y_e$, using (\ref{eq:aux:ye}).
+For the two other composition
 quantities we carry, $\bar{A}$, and $(B/A)$, we need values that are
@@
 table.  Therefore, if the thermodynamic conditions put us in NSE
-(using the conditions defined below), then we get $\bar{A}$ and
+(using the conditions defined below), then we obtain $\bar{A}$ and
 $(B/A)$ from the NSE table.  Otherwise, we compute these directly from
-$X_k$.
+$X_k$ using (\ref{eq:aux:abar}) and (\ref{eq:aux:bea}).
 

### 2020-10-29 simplified_sdc_intro/paper.tex
comments

@@
 $u$, specific total energy, $E$, and nuclear species mass fractions,
-$X_k$, and auxiliary composition variables, $a_l$.  The mass fractions
+$X_k$, and auxiliary composition variables, $\alpha_l$.  The mass fractions
 are constrained to sum to 1, $\sum_k X_k = 1$, but no such constraint
@@
 isotopes we carry in the main network. For this reason, when we are
-using the NSE network, we always take the composition quantities in
+using the NSE network, we always provide these two composition quantities as
+inputs to the EOS rather than using the $X_k$ directly.
 the EOS directly from the auxiliary state in each zone instead of
-using the $X_k$ directly. Our equation of state needs $\bar{Z}$ in addition
+using the $X_k$ directly.
+Our equation of state needs $\bar{Z}$ in addition
 to the auxiliary quantities, which is is easily computed as
-\MarginPar{might be worth spelling out exactly what inputs the NSE EOS needs to make things crystal clear}
 \begin{equation}
@@
 
-For Strang splitting, the basic flow of using the aprox19 + NSE network
-is as follows:
+For Strang splitting, the hydrodynamic and reactive substeps over
+the overall time-advancemenet scheme using the aprox19 + NSE network
+are as follows:
 \begin{itemize}
@@
 
+  At the beginning of each hydrodynamic update we have an input state,
+  $\Uc_{\rm in}$ that we wish to integrate over $\Delta t$ to obtain $\Uc_{\rm out}$
   The hydrodynamics update proceeds as normal, but with an advection
@@
 
+  At the beginning of each reactive update we have an input state,
+  $\Uc_{\rm in}$ that we wish to integrate over $\Delta t/2$ to obtain $\Uc_{\rm out}$.
+
   \begin{itemize}
@@
         \begin{equation}
-          \nse{\rho, T, (Y_e)_{\rm in}} \rightarrow \bar{A}_\out, (X_k)_\out, (B/A)_\out, dY_e/dt
+          \nse{\rho_{\rm in}, T_{\rm in}, (Y_e)_{\rm in}} \rightarrow \bar{A}_\out, (X_k)_\out, (B/A)_\out, dY_e/dt
         \end{equation}
@@
         \end{equation}
-        and then we use the table again, but with the updated $Y_e$:
-        \begin{equation}
-          \nse{\rho, T, (Y_e)_\out} \rightarrow \bar{A}_\out, (X_k)_\out, (B/A)_\out, dY_e/dt
+        and then we use the table again, but with the updated $Y_e$
+        (note that $\rho_{\rm out}=\rho_{\rm in}$ in the Strang reaction formulation):
+        \MarginPar{Indicate what $T$ is used.}
+        \begin{equation}
+          \nse{\rho_{\rm out}, T_{\rm ???}, (Y_e)_\out} \rightarrow \bar{A}_\out, (X_k)_\out, (B/A)_\out, dY_e/dt
         \end{equation}

### 2020-11-02 xrb2/paper.tex
add nicole
add new figure

@@
 
+\author{Nicole Ford}
+\affiliation{Lawrence Berkeley National Laboratory, Berkeley, CA}
+
 \author[0000-0003-2300-5165]{Donald Willcox}
@@
 
+\begin{deluxetable}{lrrr}
+	\tablecaption{\label{table:500Hz_runs} Temperature properties of the 500~Hz simulations.}
+	\tablehead{\colhead{run} & \colhead{$\delta T$ (K)} & \colhead{$T_\mathrm{hi}$ (K)} & \colhead{$T_\mathrm{star}$ (K)}} 
+	\startdata
+	{\tt flame\_wave\_500}      & $1.2\times 10^9$ & $2\times 10^8$ & $10^8$ \\
+	{\tt flame\_wave\_500\_hot} & $1.2\times 10^9$ & $2\times 10^8$ & $2\times 10^8$ \\
+	{\tt flame\_wave\_500\_3E8} & $1.1\times 10^9$ & $3\times 10^8$ & $3\times 10^8$ \\
+	{\tt flame\_wave\_500\_4E8} & $10^9$ & $4\times 10^8$ & $4\times 10^8$ \\
+	\enddata
+	%
+\end{deluxetable}
+
 Figures~\ref{fig:time_series_500} and \ref{fig:time_series_1000} show
@@
 
+\subsection{Effect of Temperature Structure}
+
+\begin{figure}[t]
+\centering
+\plotone{compare_500Hz_abar}
+\caption{\label{fig:compare_500Hz_abar} Comparison of $\bar{A}$ for 3 different 500~Hz models with different neutron star temperatures, $T_\star$, and resulting envelope
+structures.  We see that the hottest model has a faster propagating flame.}
+\end{figure}
+
+Figure~\ref{fig:compare_500Hz_abar} shows 3 different 500 Hz simulations with different initial temperature structures.
+
 \acknowledgements \castro\ is open-source and freely available at

### 2020-11-10 simplified_sdc_intro/paper.tex
tweak title again

@@
 %\title{A Simplified Spectral Deferred Correction Method for Coupling Hydrodynamics with Reaction Networks and Nuclear Statistical Equilibrium}
-\title{Improved Hydrodynamic/Reaction Coupling for Flows With Nuclear Statistical Equilibrium}
+\title{An Improved Method for Coupling Hydrodynamics with Reaction Networks and Nuclear Statistical Equilibrium}
 
@@
     reaction sources on the state variables for the next iteration.
-    For the conserved quantities, these would simply be: \MarginPar{fix signs on A}
+    For the conserved quantities, these would simply be:
     \begin{equation}
@@
 
-The NSE table requires:
+The NSE table requires: \MarginPar{need a statement about how the table was constructed and what NSE means}
 \begin{equation}

### 2020-11-11 xrb2/paper.tex
added my orcid
Updates to flame propagation section of XRB2 paper

@@
 \newcommand{\half}{\frac{1}{2}}
+
+% Approximately proportional to, from StackOverflow
+\def\app#1#2{%
+	\mathrel{%
+		\setbox0=\hbox{$#1\sim$}%
+		\setbox2=\hbox{%
+			\rlap{\hbox{$#1\propto$}}%
+			\lower1.1\ht0\box0%
+		}%
+		\raise0.25\ht2\box2%
+	}%
+}
+\def\appropto{\mathpalette\app\relax}
 
@@
 	{\tt flame\_wave\_500}      & $1.2\times 10^9$ & $2\times 10^8$ & $10^8$ \\
-	{\tt flame\_wave\_500\_hot} & $1.2\times 10^9$ & $2\times 10^8$ & $2\times 10^8$ \\
+	% Do we want the 100 Hz run in here as well, for completeness (and so we have a name to reference)?
+	{\tt flame\_wave\_500\_2E8} & $1.2\times 10^9$ & $2\times 10^8$ & $2\times 10^8$ \\
 	{\tt flame\_wave\_500\_3E8} & $1.1\times 10^9$ & $3\times 10^8$ & $3\times 10^8$ \\
@@
 
-different rotation rates
+different rotation rates. 250 Hz failed to ignite, could indicate that another source of confinement (e.g.\ magnetic fields) would need to take
+over at lower rotation rates to allow a burst to occur, at least for our current initial model. We've seen that increasing rotation rate can
+increase nuclear reaction rate already
 
@@
 
-Figure~\ref{fig:flame_speeds} shows the position of the flame as a
-function of time.  Defining the position is difficult, so we \MarginPar{need explanation here}
+\subsection{Flame Propagation}
+
+For the purpose of measuring the flame propagation speed and acceleration, we track the position of each of our flames
+as a function of time. We define the position in terms of a particular $\enucdot$ value, as we did in \citet{flame_wave1}.
+To recapitulate: we first reduce the 2D $\enucdot$ data for each simulation run to a set of 1D radial profiles by averaging
+over the vertical coordinate. After averaging, we take our reference $\enucdot$ value to be some fraction of the global $\enucdot$
+maximum across all of these profiles. Since the flames in our simulations propagate rightward, we then search the region of
+each profile to the right of the local $\enucdot$ maximum for the point where the $\enucdot$ first drops below this reference
+value. This point gives us the location of our flame front.
+
+In \citet{flame_wave1}, we used $0.1\%$ of the global $\enucdot$ maximum for our reference value. For the high temperature
+unboosted flames, however, we found that the $\enucdot$ profiles failed to reach that small a value across the domain at most times,
+which prevented us from getting reliable position measurements. We therefore use $1\%$ of the global $\enucdot$ maximum in this paper
+rather than $0.1\%$. This is still sufficiently small that our measurements are not overly sensitive to turbulence and other local fluid
+motions (the issue with simply tracking the local maximum), but allows us to avoid the pitfall encountered by the $0.1\%$ metric.
 
@@
 	\centering
-	\plottwo{speeds_1_100th_max.pdf}{speeds_1_1000th_max.pdf}
-	\caption{\label{fig:flame_speeds} Plots of flame front position vs.\ time for the 500 Hz and 1000 Hz runs.
-The left-hand plot shows the results when tracking the $\frac{1}{100} \max[\enucdot]$ contour, while the right-hand plot shows the results for the $\frac{1}{1000} \max[\enucdot]$ contour.}
+	\plotone{flame_speeds_rotation.pdf} % Not certain what style we want for the plot here
+	\caption{\label{fig:flame_speeds_1} Plots of flame front position vs.\ time for the 500 Hz and 1000 Hz runs. The dashed lines
+		show quadratic least-squares fits to the data for $t \gtrsim 20~\mathrm{ms}$.}
 \end{figure}
+
+\begin{figure}[t]
+	\centering
+	{\color{red} Placeholder for the flame speed plot that shows temperature dependence}
+	\caption{\label{fig:flame_speeds_2} Plots of flame front position vs.\ time for the three 500 Hz runs with
+		$T_{star} \leq 3 \times 10^8~\mathrm{K}$. The dashed lines show quadratic least-squares fits to the data
+		for $t \gtrsim 20~\mathrm{ms}$.}
+\end{figure}
+
+Figures~\ref{fig:flame_speeds_1} and \ref{fig:flame_speeds_2} give position vs.\ time for our runs, with the former showing dependence
+on rotation rate and the latter on the neutron star temperature. As some of the runs (in particular the $\tt{flame\_wave\_3E8}$ run) exhibit
+significant accelerations, we fit a quadratic curve to each set of data and calculate the acceleration and initial velocity from the fit
+parameters. These results are given in Table~\ref{table:flame_speeds}. We also include the flame speed at $t = 100~\mathrm{ms}$ extrapolated
+from the fit.
+
+We found in \citet{flame_wave1} that the flame speed $s$ obeys:
+\begin{equation}
+	s \appropto L_R (\epsdot)^{\half} \propto \frac{(\epsdot)^{\half}}{\Omega}
+\end{equation}
+where $\epsdot$ is the specific energy generation rate and $L_R$ is the Rossby length. This was consistent with the results of
+\citet{cavecchi:2013}, who noted that the flame speeds in their simulations scaled with the Rossby length. As seen in
+Figure~\ref{fig:flame_speeds_1}, there is no obvious inverse scaling with rotation rate from this set of runs. We observed earlier,
+however, that nuclear reactions progress more quickly at the higher rotation rate. This results in a higher $\enucdot$ --- up to 4
+to 6 times higher near the burning front after the flame ignites --- which may counteract the reduction in flame speed from the
+increased Coriolis confinement. We also observe that 500 Hz run accelerates faster than the 1000 Hz run, which appears to experience a
+small deceleration at early times. This disparity could be due to the 500 Hz flame taking more time to develop
+fully and reach higher burning rates, or a result of the difference in Coriolis force. \MarginPar{Do we have other ideas why?}
+
+% It might be worth plotting peak enucdot as a function of time for each run
+
+{\color{red} Maybe Nicole can talk a bit about the effect of the neutron star temperature here? I have a placeholder for the plot earlier, but
+feel free to move it. What's in this section now is just an initial draft/outline.}
+
+\begin{deluxetable}{lrrr}
+	\tablecaption{\label{table:flame_speeds} Flame speed and acceleration values measured for each simulation.
+$s_0$ is the speed at $t = 0~\mathrm{ms}$, and $s_{100}$ is the speed at $t = 100~\mathrm{ms}$.}
+	\tablehead{\colhead{run} & \colhead{$s_0$ (km s$^{-1}$)} & \colhead{$s_{100}$ (km s$^{-1}$)} & \colhead{$a$ (km s$^{-2}$)}} 
+	\startdata
+	{\tt flame\_wave\_500}      & $3.08 \pm 0.01$ & $3.62 \pm 0.02$ & $5.41 \pm 0.19$ \\
+	{\tt flame\_wave\_1000}     & $3.41 \pm 0.01$ & $3.31 \pm 0.02$ & $-1.03 \pm 0.13$ \\
+	{\tt flame\_wave\_500\_2E8} & $3.76 \pm 0.02$ & $\sim 3.93$ & $1.74 \pm 0.17$ \\
+	{\tt flame\_wave\_500\_3E8} & $5.29 \pm 0.10$ & $\sim 41.0$ & $357.4 \pm 1.5$ \\
+	\enddata
+	%
+\end{deluxetable}
 

### 2020-11-11 xrb2/paper.tex
Added CSGF acknowledgement

@@
 
-\author{Nicole Ford}
+\author[0000-0001-8921-3624]{Nicole Ford}
 \affiliation{Lawrence Berkeley National Laboratory, Berkeley, CA}
@@
 \citet{cavecchi:2013}, who noted that the flame speeds in their simulations scaled with the Rossby length. As seen in
-Figure~\ref{fig:flame_speeds_1}, there is no obvious inverse scaling with rotation rate from this set of runs. We observed earlier,
+Figure~\ref{fig:flame_speeds_1}, there is no obvious inverse scaling with rotation rate in this set of runs. We observed earlier,
 however, that nuclear reactions progress more quickly at the higher rotation rate. This results in a higher $\enucdot$ --- up to 4
@@
 increased Coriolis confinement. We also observe that 500 Hz run accelerates faster than the 1000 Hz run, which appears to experience a
-small deceleration at early times. This disparity could be due to the 500 Hz flame taking more time to develop
-fully and reach higher burning rates, or a result of the difference in Coriolis force. \MarginPar{Do we have other ideas why?}
+small deceleration at early times. This disparity could be due to the 500 Hz flame taking more time to develop fully and reach higher
+burning rates, resulting in an $\epsdot$ increase over time. It may also be a direct result of the difference in Coriolis force. \MarginPar{Do we have other ideas why?}
 
@@
 {\color{red} Maybe Nicole can talk a bit about the effect of the neutron star temperature here? I have a placeholder for the plot earlier, but
-feel free to move it. What's in this section now is just an initial draft/outline.}
+feel free to move/edit it. What's in this section now is just an initial draft/outline.}
 
@@
 Discovery through Advanced Computing (SciDAC) program under Award
-Number DE-SC0017955.  MZ acknowledges support from the Simons
-Foundation.  This research used resources of the National Energy
+Number DE-SC0017955.  This material is also based upon work supported by the U.S. Department
+of Energy, Office of Science, Office of Advanced Scientific Computing Research, Department of
+Energy Computational Science Graduate Fellowship under Award Number DE-SC0021110.
+MZ acknowledges support from the Simons Foundation. 
+This research used resources of the National Energy
 Research Scientific Computing Center, a DOE Office of Science User

### 2020-11-11 xrb2/paper.tex
Updates to XRB2 flame speed plots

@@
 	\plotone{flame_speeds_rotation.pdf} % Not certain what style we want for the plot here
-	\caption{\label{fig:flame_speeds_1} Plots of flame front position vs.\ time for the 500 Hz and 1000 Hz runs. The dashed lines
+	\caption{\label{fig:flame_speeds_1} Flame front position vs.\ time for the 500 Hz and 1000 Hz runs. The dashed lines
 		show quadratic least-squares fits to the data for $t \gtrsim 20~\mathrm{ms}$.}
@@
 	{\color{red} Placeholder for the flame speed plot that shows temperature dependence}
-	\caption{\label{fig:flame_speeds_2} Plots of flame front position vs.\ time for the three 500 Hz runs with
+	\caption{\label{fig:flame_speeds_2} Flame front position vs.\ time for the three 500 Hz runs with
 		$T_{star} \leq 3 \times 10^8~\mathrm{K}$. The dashed lines show quadratic least-squares fits to the data

### 2020-11-11 xrb2/paper.tex
Switch to CGS

@@
 	\centering
-	\plotone{flame_speeds_rotation.pdf} % Not certain what style we want for the plot here
-	\caption{\label{fig:flame_speeds_1} Flame front position vs.\ time for the 500 Hz and 1000 Hz runs. The dashed lines
-		show quadratic least-squares fits to the data for $t \gtrsim 20~\mathrm{ms}$.}
+	\plotone{flame_speeds_rotation.pdf}
+	\caption{\label{fig:flame_speeds_1} Flame front position vs.\ time for the standard ($10^8~\mathrm{K}$) 500 Hz and 1000 Hz runs. The dashed lines
+		show quadratic least-squares fits to the data for $t \gtrsim 0.02~\mathrm{s}$.}
 \end{figure}
@@
 	\caption{\label{fig:flame_speeds_2} Flame front position vs.\ time for the three 500 Hz runs with
-		$T_{star} \leq 3 \times 10^8~\mathrm{K}$. The dashed lines show quadratic least-squares fits to the data
-		for $t \gtrsim 20~\mathrm{ms}$.}
+		$T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$. The dashed lines show quadratic least-squares fits to the data
+		for $t \gtrsim 0.02~\mathrm{s}$.}
 \end{figure}
+
+\begin{deluxetable}{lrrr}
+	\tablecaption{\label{table:flame_speeds} Flame speed and acceleration values measured for each simulation. $s_0$ is the speed at $t = 0~\mathrm{ms}$.}
+	\tablehead{\colhead{run} & \colhead{$s_0$ (cm s$^{-1}$)} & \colhead{$a$ (cm s$^{-2}$)}} 
+	\startdata
+	{\tt flame\_wave\_500}      & $(3.08 \pm 0.01) \times 10^5$ & $(5.41 \pm 0.19) \times 10^5$ \\
+	{\tt flame\_wave\_1000}     & $(3.41 \pm 0.01) \times 10^5$ & $(-1.03 \pm 0.13) \times 10^5$ \\
+	{\tt flame\_wave\_500\_2E8} & $(3.76 \pm 0.02) \times 10^5$ & $(1.74 \pm 0.17) \times 10^5$ \\
+	{\tt flame\_wave\_500\_3E8} & $(5.29 \pm 0.10) \times 10^5$ & $(3.57 \pm 0.02) \times 10^7$ \\
+	\enddata
+	%
+\end{deluxetable}
 
@@
 significant accelerations, we fit a quadratic curve to each set of data and calculate the acceleration and initial velocity from the fit
-parameters. These results are given in Table~\ref{table:flame_speeds}. We also include the flame speed at $t = 100~\mathrm{ms}$ extrapolated
-from the fit.
+parameters. These results are given in Table~\ref{table:flame_speeds}.
 
@@
 increased Coriolis confinement. We also observe that 500 Hz run accelerates faster than the 1000 Hz run, which appears to experience a
-small deceleration at early times. This disparity could be due to the 500 Hz flame taking more time to develop fully and reach higher
-burning rates, resulting in an $\epsdot$ increase over time. It may also be a direct result of the difference in Coriolis force. \MarginPar{Do we have other ideas why?}
+small deceleration at early times. \MarginPar{This is speculative -- do we have a clear idea why?} This disparity may be a direct result
+of the difference in Coriolis force. It could also be due to the 500 Hz flame taking more time to develop fully and reach higher burning
+rates, resulting in an $\epsdot$ increase over time.
 
@@
 feel free to move/edit it. What's in this section now is just an initial draft/outline.}
-
-\begin{deluxetable}{lrrr}
-	\tablecaption{\label{table:flame_speeds} Flame speed and acceleration values measured for each simulation.
-$s_0$ is the speed at $t = 0~\mathrm{ms}$, and $s_{100}$ is the speed at $t = 100~\mathrm{ms}$.}
-	\tablehead{\colhead{run} & \colhead{$s_0$ (km s$^{-1}$)} & \colhead{$s_{100}$ (km s$^{-1}$)} & \colhead{$a$ (km s$^{-2}$)}} 
-	\startdata
-	{\tt flame\_wave\_500}      & $3.08 \pm 0.01$ & $3.62 \pm 0.02$ & $5.41 \pm 0.19$ \\
-	{\tt flame\_wave\_1000}     & $3.41 \pm 0.01$ & $3.31 \pm 0.02$ & $-1.03 \pm 0.13$ \\
-	{\tt flame\_wave\_500\_2E8} & $3.76 \pm 0.02$ & $\sim 3.93$ & $1.74 \pm 0.17$ \\
-	{\tt flame\_wave\_500\_3E8} & $5.29 \pm 0.10$ & $\sim 41.0$ & $357.4 \pm 1.5$ \\
-	\enddata
-	%
-\end{deluxetable}
 

### 2020-11-11 xrb2/paper.tex
Reword MarginPar
Added 500Hz flame speeds plot to paper

@@
 	\centering
-	{\color{red} Placeholder for the flame speed plot that shows temperature dependence}
+	\plotone{500Hz_stacked_enuc1_2order.pdf}
 	\caption{\label{fig:flame_speeds_2} Flame front position vs.\ time for the three 500 Hz runs with
@@
 \begin{deluxetable}{lrrr}
-	\tablecaption{\label{table:flame_speeds} Flame speed and acceleration values measured for each simulation. $s_0$ is the speed at $t = 0~\mathrm{ms}$.}
+	\tablecaption{\label{table:flame_speeds} Flame speed and acceleration values measured for each simulation. $s_0$ is the speed at $t = 0$.}
 	\tablehead{\colhead{run} & \colhead{$s_0$ (cm s$^{-1}$)} & \colhead{$a$ (cm s$^{-2}$)}} 
@@
 increased Coriolis confinement. We also observe that 500 Hz run accelerates faster than the 1000 Hz run, which appears to experience a
-small deceleration at early times. \MarginPar{This is speculative -- do we have a clear idea why?} This disparity may be a direct result
-of the difference in Coriolis force. It could also be due to the 500 Hz flame taking more time to develop fully and reach higher burning
-rates, resulting in an $\epsdot$ increase over time.
+small deceleration at early times. \MarginPar{This is speculation on my part -- does someone else have clear idea why?} This disparity
+may be a direct result of the difference in Coriolis force. It could also be due to the 500 Hz flame taking more time to develop fully
+and reach higher burning rates, resulting in an $\epsdot$ increase over time.
 

### 2020-11-12 xrb2/paper.tex
add note
Added 1000Hz and 250Hz info to Table 1 for completeness

@@
 
+In our latest set of runs, we explore a variety of initial rotation rate and temperature conditions for the flame. The initial conditions are given in Table \ref{table:500Hz_runs}.
+
 Things to show:
@@
 
-\begin{deluxetable}{lrrr}
-	\tablecaption{\label{table:500Hz_runs} Temperature properties of the 500~Hz simulations.}
-	\tablehead{\colhead{run} & \colhead{$\delta T$ (K)} & \colhead{$T_\mathrm{hi}$ (K)} & \colhead{$T_\mathrm{star}$ (K)}} 
+\begin{deluxetable}{lrrrrr}
+	\tablecaption{\label{table:500Hz_runs} Rotation rate and temperature properties of the simulations.}
+	\tablehead{\colhead{run} & \colhead{Rotation Rate (Hz)} & \colhead{$\delta T$ (K)} & \colhead{$T_\mathrm{hi}$ (K)} & \colhead{$T_\mathrm{star}$ (K)}} 
 	\startdata
-	{\tt flame\_wave\_500}      & $1.2\times 10^9$ & $2\times 10^8$ & $10^8$ \\
-	% Do we want the 100 Hz run in here as well, for completeness (and so we have a name to reference)?
-	{\tt flame\_wave\_500\_2E8} & $1.2\times 10^9$ & $2\times 10^8$ & $2\times 10^8$ \\
-	{\tt flame\_wave\_500\_3E8} & $1.1\times 10^9$ & $3\times 10^8$ & $3\times 10^8$ \\
-	{\tt flame\_wave\_500\_4E8} & $10^9$ & $4\times 10^8$ & $4\times 10^8$ \\
+	{\tt flame\_wave\_1000}     & $1000$ & $1.2\times 10^9$ & $2\times 10^8$ & $10^8$ \\
+	{\tt flame\_wave\_500}      & $500$ & $1.2\times 10^9$ & $2\times 10^8$ & $10^8$ \\
+	{\tt flame\_wave\_500\_2E8} & $500$ & $1.2\times 10^9$ & $2\times 10^8$ & $2\times 10^8$ \\
+	{\tt flame\_wave\_500\_3E8} & $500$ & $1.1\times 10^9$ & $3\times 10^8$ & $3\times 10^8$ \\
+	{\tt flame\_wave\_500\_4E8} & $500$ & $10^9$ & $4\times 10^8$ & $4\times 10^8$ \\
+	{\tt flame\_wave\_250}      & $250$ & $1.2\times 10^9$ & $2\times 10^8$ & $10^8$ \\
 	\enddata
-	%
 \end{deluxetable}

### 2020-11-13 xrb2/paper.tex
Reorganized subsections to be rotation & temperature-specific, added some temperature discussion

@@
 
-In our latest set of runs, we explore a variety of initial rotation rate and temperature conditions for the flame. The initial conditions are given in Table \ref{table:500Hz_runs}.
+In our latest set of runs, we explore a variety of initial rotation rate and temperature conditions for the flame. The initial conditions are provided in Table \ref{table:500Hz_runs}.
+
+\begin{deluxetable}{lrrrrr}
+	\tablecaption{\label{table:500Hz_runs} Rotation rate and temperature properties of the simulations.}
+	\tablehead{\colhead{run} & \colhead{Rotation Rate (Hz)} & \colhead{$\delta T$ (K)} & \colhead{$T_\mathrm{hi}$ (K)} & \colhead{$T_\mathrm{star}$ (K)}} 
+	\startdata
+	{\tt flame\_wave\_1000}     & $1000$ & $1.2\times 10^9$ & $2\times 10^8$ & $10^8$ \\
+	{\tt flame\_wave\_500}      & $500$ & $1.2\times 10^9$ & $2\times 10^8$ & $10^8$ \\
+	{\tt flame\_wave\_500\_2E8} & $500$ & $1.2\times 10^9$ & $2\times 10^8$ & $2\times 10^8$ \\
+	{\tt flame\_wave\_500\_3E8} & $500$ & $1.1\times 10^9$ & $3\times 10^8$ & $3\times 10^8$ \\
+	{\tt flame\_wave\_500\_4E8} & $500$ & $10^9$ & $4\times 10^8$ & $4\times 10^8$ \\
+	{\tt flame\_wave\_250}      & $250$ & $1.2\times 10^9$ & $2\times 10^8$ & $10^8$ \\
+	\enddata
+\end{deluxetable}
+
+\subsection{Effect of Rotation Rate on Flame Structure}\label{ssec:rot_structure}
 
@@
 \end{figure}
-
-\begin{deluxetable}{lrrrrr}
-	\tablecaption{\label{table:500Hz_runs} Rotation rate and temperature properties of the simulations.}
-	\tablehead{\colhead{run} & \colhead{Rotation Rate (Hz)} & \colhead{$\delta T$ (K)} & \colhead{$T_\mathrm{hi}$ (K)} & \colhead{$T_\mathrm{star}$ (K)}} 
-	\startdata
-	{\tt flame\_wave\_1000}     & $1000$ & $1.2\times 10^9$ & $2\times 10^8$ & $10^8$ \\
-	{\tt flame\_wave\_500}      & $500$ & $1.2\times 10^9$ & $2\times 10^8$ & $10^8$ \\
-	{\tt flame\_wave\_500\_2E8} & $500$ & $1.2\times 10^9$ & $2\times 10^8$ & $2\times 10^8$ \\
-	{\tt flame\_wave\_500\_3E8} & $500$ & $1.1\times 10^9$ & $3\times 10^8$ & $3\times 10^8$ \\
-	{\tt flame\_wave\_500\_4E8} & $500$ & $10^9$ & $4\times 10^8$ & $4\times 10^8$ \\
-	{\tt flame\_wave\_250}      & $250$ & $1.2\times 10^9$ & $2\times 10^8$ & $10^8$ \\
-	\enddata
-\end{deluxetable}
 
@@
 
-\subsection{Flame Propagation}
+\subsection{Effect of Rotation Rate on Flame Propagation}\label{ssec:rot_propagation}
 
@@
 
-\begin{figure}[t]
-	\centering
-	\plotone{500Hz_stacked_enuc1_2order.pdf}
-	\caption{\label{fig:flame_speeds_2} Flame front position vs.\ time for the three 500 Hz runs with
-		$T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$. The dashed lines show quadratic least-squares fits to the data
-		for $t \gtrsim 0.02~\mathrm{s}$.}
-\end{figure}
-
 \begin{deluxetable}{lrrr}
@@
 	\startdata
+	{\tt flame\_wave\_1000}     & $(3.41 \pm 0.01) \times 10^5$ & $(-1.03 \pm 0.13) \times 10^5$ \\
 	{\tt flame\_wave\_500}      & $(3.08 \pm 0.01) \times 10^5$ & $(5.41 \pm 0.19) \times 10^5$ \\
-	{\tt flame\_wave\_1000}     & $(3.41 \pm 0.01) \times 10^5$ & $(-1.03 \pm 0.13) \times 10^5$ \\
 	{\tt flame\_wave\_500\_2E8} & $(3.76 \pm 0.02) \times 10^5$ & $(1.74 \pm 0.17) \times 10^5$ \\
@@
 
-Figures~\ref{fig:flame_speeds_1} and \ref{fig:flame_speeds_2} give position vs.\ time for our runs, with the former showing dependence
-on rotation rate and the latter on the neutron star temperature. As some of the runs (in particular the $\tt{flame\_wave\_3E8}$ run) exhibit
-significant accelerations, we fit a quadratic curve to each set of data and calculate the acceleration and initial velocity from the fit
-parameters. These results are given in Table~\ref{table:flame_speeds}.
+Figure~\ref{fig:flame_speeds_1} gives position vs.\ time for the $\tt{flame\_wave\_500}$ and $\tt{flame\_wave\_1000}$ runs to show the dependence on rotation rate.
+As some of the hotter 500~Hz runs exhibit significant accelerations, we fit a quadratic curve to each set of data and calculate the acceleration and initial velocity from the fit parameters.
+These results are given in Table~\ref{table:flame_speeds}.
 
@@
 to 6 times higher near the burning front after the flame ignites --- which may counteract the reduction in flame speed from the
-increased Coriolis confinement. We also observe that 500 Hz run accelerates faster than the 1000 Hz run, which appears to experience a
+increased Coriolis confinement. We also observe that the 500 Hz run accelerates faster than the 1000 Hz run, which appears to experience a
 small deceleration at early times. \MarginPar{This is speculation on my part -- does someone else have clear idea why?} This disparity
@@
 % It might be worth plotting peak enucdot as a function of time for each run
-
-{\color{red} Maybe Nicole can talk a bit about the effect of the neutron star temperature here? I have a placeholder for the plot earlier, but
-feel free to move/edit it. What's in this section now is just an initial draft/outline.}
 
@@
 
-\subsection{Effect of Temperature Structure}
+\subsection{Effect of Temperature Structure}\label{ssec:temp_structure}
+
+To explore the effect of different initial temperature configurations, we run four simulations fixed at a rotation rate of 500 Hz with temperatures as shown in Table \ref{table:500Hz_runs}.
+For all the 500 Hz simulations (with the exception of the coolest run, $\tt{flame\_wave\_500}$), we set $T_\mathrm{star}$ = $T_\mathrm{hi}$, scaling $\delta T$ accordingly to maintain a consistent value of $T_\mathrm{hi} + \delta T$.
+If we let $T_\mathrm{star} < T_\mathrm{hi}$, the cooler neutron star surface might act as a heat sink and siphon away energy that would otherwise go into heating the burning layer.
 
@@
 \plotone{compare_500Hz_abar}
-\caption{\label{fig:compare_500Hz_abar} Comparison of $\bar{A}$ for 3 different 500~Hz models with different neutron star temperatures, $T_\star$, and resulting envelope
-structures.  We see that the hottest model has a faster propagating flame.}
+\caption{\label{fig:compare_500Hz_abar} Comparison of $\bar{A}$ for 3 different 500~Hz models with different neutron star temperatures, $T_\star$, and resulting envelope structures.
+} %This will be updated to show at 70ms
 \end{figure}
@@
 Figure~\ref{fig:compare_500Hz_abar} shows 3 different 500 Hz simulations with different initial temperature structures.
+The hottest model, $\tt{flame\_wave\_500\_3E8}$ has a faster propagating flame.
+It also appears to have a secondary burning region above the neutron star atmosphere layer (the atmosphere edge is located at $\sim 1.2\times 10^4$ cm).
+It is likely that a small amount of hot material broke through the top of the atmosphere.
+
+{\color{red} On the subject of things breaking out of the atmosphere, discuss interesting characteristics of the extreme 4E8 run here: maybe include a plot showing abar and temp at 40 ms, and another plot showing the momentum components?}
+
+\begin{figure}[t]
+	\centering
+	\plotone{500Hz_stacked_enuc1_2order.pdf}
+	\caption{\label{fig:flame_speeds_2} Flame front position vs.\ time for the three 500 Hz runs with
+		$T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$. The dashed lines show quadratic least-squares fits to the data
+		for $t \gtrsim 0.02~\mathrm{s}$.}
+\end{figure}
+
+{\color{red} Discuss flame acceleration here, it's sort of mysterious but could be due to energy released from burning dominating the flame propagation and increasing speed over time, or maybe enhanced mixing/diffusion is pulling in more fuel to burn, or maybe the average fuel layer temp is going up, making it easier for the flame to burn and propagate}
 

### 2020-11-13 xrb2/paper.tex
Added more discussion about the temperature acceleration plot

@@
 Figure~\ref{fig:compare_500Hz_abar} shows 3 different 500 Hz simulations with different initial temperature structures.
-The hottest model, $\tt{flame\_wave\_500\_3E8}$ has a faster propagating flame.
+The hottest model, $\tt{flame\_wave\_500\_3E8}$, has a faster propagating flame.
 It also appears to have a secondary burning region above the neutron star atmosphere layer (the atmosphere edge is located at $\sim 1.2\times 10^4$ cm).
@@
 
-{\color{red} On the subject of things breaking out of the atmosphere, discuss interesting characteristics of the extreme 4E8 run here: maybe include a plot showing abar and temp at 40 ms, and another plot showing the momentum components?}
+{\color{red} On the subject of things breaking out of the atmosphere, discuss interesting characteristics of the extreme 4E8 run here: maybe include a plot showing enuc at 40 ms comapared to one of the cooler runs, and another plot showing the momentum components?
+Seems like the flame moves very quickly across the domain, and then residual burning after the flame has passed eventually ignites the entire fuel layer.
+This total ignition can also occur in nature for a neutron star with a lot of dense and hot accreted material.}
 
@@
 
-{\color{red} Discuss flame acceleration here, it's sort of mysterious but could be due to energy released from burning dominating the flame propagation and increasing speed over time, or maybe enhanced mixing/diffusion is pulling in more fuel to burn, or maybe the average fuel layer temp is going up, making it easier for the flame to burn and propagate}
+The method for measuring flame propagation described in Section \ref{ssec:rot_propagation} is applied in Figure \ref{fig:flame_speeds_2} to the three 500 Hz runs with $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$.
+As the initial $T_{\mathrm{star}}$ is increased, the flame becomes greatly accelerated.
+The initial flame velocities and accelerations derived from a quadratic least-squares fit to each of the datasets are provided in Table \ref{table:flame_speeds}.
+The reason for the acceleration of the flame over time is not entirely clear.
+It could be that energy released from burning begins to dominate the flame's propagation as it evolves, increasing the flame speed over time.
+Another possibility is that the increased temperatures lead to enhanced turbulent mixing effects that pull in more fresh fuel for the flame to burn.
+Yet another possibility is that the higher initial $T_{\mathrm{star}}$ leads to a greater average temperature in the fuel layer over time, making it easier for the flame to burn the fuel and propagate.
 

### 2020-11-18 xrb2/paper.tex
add figures to paper

@@
 
+\begin{figure}[t]
+    \centering
+    \plotone{urho_plot_hotter}
+    \caption{\label{fig:urho_hot}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $2\times10^8$~K and $3\times10^8$~K.}
+\end{figure}
+
+\begin{figure}[t]
+    \centering
+    \plotone{uv_plot_hotter}
+    \caption{\label{fig:uv_hot}Phase plot of the horizontal $x$-velocity $u$, vertical $y$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $2\times10^8$~K and $3\times10^8$~K.}
+\end{figure}
+
+\begin{figure}[t]
+    \centering
+    \plotone{abar_plot_hotter}
+    \caption{\label{fig:abar_hot}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $2\times10^8$~K and $3\times10^8$~K.}
+\end{figure}
+
 \acknowledgements \castro\ is open-source and freely available at

### 2020-11-19 xrb2/paper.tex
added some text describing the temperature phase plots

@@
 
+Looking at the phase plots in Figures~\ref{fig:urho_hot}-\ref{fig:abar_hot}, we see:
+\begin{itemize}
+    \item The $u$-$\rho$ plot in Figure~\ref{fig:urho_hot} shows that the horizontal velocity when $T_{\rm star} = 3\times10^8$~K is sufficiently greater than for the two cooler runs.
+This is consistent with the plot of the flame front position, where we saw that the hottest flame was significantly faster at later times than the cooler flames.
+    \item The $u$-$v$ plot in Figure~\ref{fig:uv_hot} further demonstrates this, as again we see that the magnitude of the velocity of the flame for the hottest run is significantly larger.
+Also of interest in these plots is that the coolest run does not seem to have developed the characteristic vortex structure at the burning front (i.e.
+where $\dot{e}_{\rm nuc}$ is greatest) that can be clearly seen for the hotter runs.
+    \item The $\overline{A}$-$T$ plot in Figure~\ref{fig:abar_hot} suggests that the actual reactions happening in the hottest run may be different than for the cooler runs.
+Not only is the flame itself significantly hotter overall, but there is a large causally connected region across a wide range of $\overline{A}$.
+This indicates that the hotter temperature has facilitated significant burning in reaction pathways that were not favored at the lower temperatures.
+\end{itemize}
+
 \begin{figure}[t]
@@
 \end{figure}
+
+%======================================================================
+% Conclusions
+%======================================================================
+\section{Discussion and Conclusions}\label{Sec:conclusions}
+
+tl;dr: faster rotation and/or hotter surface temperature = more burning.
 

### 2020-11-21 xrb2/paper.tex
Added 4e8 flame plots to temperature discussion

@@
 Figure~\ref{fig:compare_500Hz_abar} shows 3 different 500 Hz simulations with different initial temperature structures.
-The hottest model, $\tt{flame\_wave\_500\_3E8}$, has a faster propagating flame.
+The hottest of the three models, $\tt{flame\_wave\_500\_3E8}$, has a faster propagating flame.
 It also appears to have a secondary burning region above the neutron star atmosphere layer (the atmosphere edge is located at $\sim 1.2\times 10^4$ cm).
@@
 This total ignition can also occur in nature for a neutron star with a lot of dense and hot accreted material.}
+
+\begin{figure}[t]
+	\centering
+	\plotone{multiplot_3x1_time_series_enuc_late.png}
+	\caption{\label{fig:4e8_stacked_enuc} Time series of the 500~Hz run $\tt{flame\_wave\_500\_4E8}$, with $T_{\mathrm{star}} = 4 \times 10^8~\mathrm{K}$, showing $\dot{e}_{nuc}$.
+This model burns so strongly that it is only run out to 40 ms; the snapshots shown here are at T = 30 ms, 35 ms, and 40 ms.}
+\end{figure}
+
+\begin{figure}[t]
+	\centering
+	\plotone{multiplot_3x1_time_series_abar_late.png}
+	\caption{\label{fig:4e8_stacked_abar} Time series of the 500~Hz run $\tt{flame\_wave\_500\_4E8}$, with $T_{\mathrm{star}} = 4 \times 10^8~\mathrm{K}$, showing $\bar{A}$.}
+\end{figure}
 

### 2020-11-23 xrb2/paper.tex
some methods

@@
 \section{Numerical Approach}\label{Sec:numerics}
+\MarginPar{mz}
 
-We use the Castro hydrodynamics code \citep{castro} and the simulation framework introduced in \citet{flame_wave1}.
-The main difference is that we no longer boost the flames by adjusting
-the reaction rates and conductivities.  By porting \castro\ to GPUs,
-we are able to run the unboosted flame simulations with fewer
-resources than the boosted simulations used in our earlier paper.
+We use the Castro hydrodynamics code \citep{castro} and the simulation
+framework introduced in \citet{flame_wave1}.  The main difference is
+that we no longer boost the flames by adjusting the reaction rates and
+conductivities.  By porting \castro\ to GPUs \citep{castro:sc20}, we
+are able to run the unboosted flame simulations with fewer resources
+than the boosted simulations used in our earlier paper.  For these
+axisymmetric simulations, we adding a missing geometric source terms
+from \citet{bernard-champmartin} that captures the effects of the
+divergence of the flux operating on the azimuthal unit vector.  This
+term is a small correction, but was missing from our earlier
+simulations.
 
-We add the geometric source terms from \citet{bernard-champmartin}
+The initial model is setup in the same fashion as described in
+\citet{flame_wave}.  In particular, we create a ``hot'' and ``cool''
+hydrostatic model representing the ash and fuel states and blend the
+two models laterally to create a hot region near the origin of
+coordinates and a smooth transition to the cooler region at larger
+radii.  In the calculations presented here, we explore the structure
+of the initial models by varying some of the parameters, in particular, the temperature
+of the neutron star and at the base of the fuel layer. \MarginPar{add plot showing models}
 

### 2020-11-23 xrb2/paper.tex
switch flame_wave to just F for brevity

@@
 	\startdata
-	{\tt flame\_wave\_1000}     & $1000$ & $1.2\times 10^9$ & $2\times 10^8$ & $10^8$ \\
-	{\tt flame\_wave\_500}      & $500$ & $1.2\times 10^9$ & $2\times 10^8$ & $10^8$ \\
-	{\tt flame\_wave\_500\_2E8} & $500$ & $1.2\times 10^9$ & $2\times 10^8$ & $2\times 10^8$ \\
-	{\tt flame\_wave\_500\_3E8} & $500$ & $1.1\times 10^9$ & $3\times 10^8$ & $3\times 10^8$ \\
-	{\tt flame\_wave\_500\_4E8} & $500$ & $10^9$ & $4\times 10^8$ & $4\times 10^8$ \\
-	{\tt flame\_wave\_250}      & $250$ & $1.2\times 10^9$ & $2\times 10^8$ & $10^8$ \\
+	{\tt F1000}     & $1000$ & $1.2\times 10^9$ & $2\times 10^8$ & $10^8$ \\
+	{\tt F500}      & $500$ & $1.2\times 10^9$ & $2\times 10^8$ & $10^8$ \\
+	{\tt F500\_2E8} & $500$ & $1.2\times 10^9$ & $2\times 10^8$ & $2\times 10^8$ \\
+	{\tt F500\_3E8} & $500$ & $1.1\times 10^9$ & $3\times 10^8$ & $3\times 10^8$ \\
+	{\tt F500\_4E8} & $500$ & $10^9$ & $4\times 10^8$ & $4\times 10^8$ \\
+	{\tt F250}      & $250$ & $1.2\times 10^9$ & $2\times 10^8$ & $10^8$ \\
 	\enddata
@@
 	\startdata
-	{\tt flame\_wave\_1000}     & $(3.41 \pm 0.01) \times 10^5$ & $(-1.03 \pm 0.13) \times 10^5$ \\
-	{\tt flame\_wave\_500}      & $(3.08 \pm 0.01) \times 10^5$ & $(5.41 \pm 0.19) \times 10^5$ \\
-	{\tt flame\_wave\_500\_2E8} & $(3.76 \pm 0.02) \times 10^5$ & $(1.74 \pm 0.17) \times 10^5$ \\
-	{\tt flame\_wave\_500\_3E8} & $(5.29 \pm 0.10) \times 10^5$ & $(3.57 \pm 0.02) \times 10^7$ \\
+	{\tt F1000}     & $(3.41 \pm 0.01) \times 10^5$ & $(-1.03 \pm 0.13) \times 10^5$ \\
+	{\tt F500}      & $(3.08 \pm 0.01) \times 10^5$ & $(5.41 \pm 0.19) \times 10^5$ \\
+	{\tt F500\_2E8} & $(3.76 \pm 0.02) \times 10^5$ & $(1.74 \pm 0.17) \times 10^5$ \\
+	{\tt F500\_3E8} & $(5.29 \pm 0.10) \times 10^5$ & $(3.57 \pm 0.02) \times 10^7$ \\
 	\enddata
@@
 
-Figure~\ref{fig:flame_speeds_1} gives position vs.\ time for the $\tt{flame\_wave\_500}$ and $\tt{flame\_wave\_1000}$ runs to show the dependence on rotation rate.
+Figure~\ref{fig:flame_speeds_1} gives position vs.\ time for the $\tt{F500}$ and $\tt{F1000}$ runs to show the dependence on rotation rate.
 As some of the hotter 500~Hz runs exhibit significant accelerations, we fit a quadratic curve to each set of data and calculate the acceleration and initial velocity from the fit parameters.
@@
 To explore the effect of different initial temperature configurations, we run four simulations fixed at a rotation rate of 500 Hz with temperatures as shown in Table \ref{table:500Hz_runs}.
-For all the 500 Hz simulations (with the exception of the coolest run, $\tt{flame\_wave\_500}$), we set $T_\mathrm{star}$ = $T_\mathrm{hi}$, scaling $\delta T$ accordingly to maintain a consistent value of $T_\mathrm{hi} + \delta T$.
+For all the 500 Hz simulations (with the exception of the coolest run, $\tt{F500}$), we set $T_\mathrm{star}$ = $T_\mathrm{hi}$, scaling $\delta T$ accordingly to maintain a consistent value of $T_\mathrm{hi} + \delta T$.
 If we let $T_\mathrm{star} < T_\mathrm{hi}$, the cooler neutron star surface might act as a heat sink and siphon away energy that would otherwise go into heating the burning layer.
@@
 Figure~\ref{fig:compare_500Hz_abar} shows 3 different 500 Hz simulations with different initial temperature structures.
-The hottest of the three models, $\tt{flame\_wave\_500\_3E8}$, has a faster propagating flame.
+The hottest of the three models, $\tt{F500\_3E8}$, has a faster propagating flame.
 It also appears to have a secondary burning region above the neutron star atmosphere layer (the atmosphere edge is located at $\sim 1.2\times 10^4$ cm).
@@
 	\plotone{multiplot_3x1_time_series_enuc_late.png}
-	\caption{\label{fig:4e8_stacked_enuc} Time series of the 500~Hz run $\tt{flame\_wave\_500\_4E8}$, with $T_{\mathrm{star}} = 4 \times 10^8~\mathrm{K}$, showing $\dot{e}_{nuc}$.
+	\caption{\label{fig:4e8_stacked_enuc} Time series of the 500~Hz run $\tt{F500\_4E8}$, with $T_{\mathrm{star}} = 4 \times 10^8~\mathrm{K}$, showing $\dot{e}_{nuc}$.
 This model burns so strongly that it is only run out to 40 ms; the snapshots shown here are at T = 30 ms, 35 ms, and 40 ms.}
@@
 	\plotone{multiplot_3x1_time_series_abar_late.png}
-	\caption{\label{fig:4e8_stacked_abar} Time series of the 500~Hz run $\tt{flame\_wave\_500\_4E8}$, with $T_{\mathrm{star}} = 4 \times 10^8~\mathrm{K}$, showing $\bar{A}$.}
+	\caption{\label{fig:4e8_stacked_abar} Time series of the 500~Hz run $\tt{F500\_4E8}$, with $T_{\mathrm{star}} = 4 \times 10^8~\mathrm{K}$, showing $\bar{A}$.}
 \end{figure}

### 2020-11-26 xrb2/paper.tex
add some more figures
update refs

@@
 
-We use the Castro hydrodynamics code \citep{castro} and the simulation
+We use the Castro hydrodynamics code \citep{castro,castro_joss} and the simulation
 framework introduced in \citet{flame_wave1}.  The main difference is
 that we no longer boost the flames by adjusting the reaction rates and
-conductivities.  By porting \castro\ to GPUs \citep{castro:sc20}, we
+conductivities.  By porting \castro\ to GPUs \citep{sc20_gpu}, we
 are able to run the unboosted flame simulations with fewer resources
@@
 of the initial models by varying some of the parameters, in particular, the temperature
-of the neutron star and at the base of the fuel layer. \MarginPar{add plot showing models}
+of the neutron star and at the base of the fuel layer.
+
+\begin{figure}[t]
+\centering
+\plotone{hot}
+\caption{\label{fig:hot_profiles} Initial temperature and density structure vs.\ height in the ``hot'' state.}
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\plotone{cold}
+\caption{\label{fig:cold_profiles} Initial temperature and density structure vs.\ height in the ``cool'' state.}
+\end{figure}
 
@@
 
+Next steps: 3-d, explore resolution more, H/He flames, MHD flames
+
 \acknowledgements \castro\ is open-source and freely available at

### 2020-12-01 xrb2/paper.tex
Edits to temperature section

@@
 The initial model is setup in the same fashion as described in
-\citet{flame_wave}.  In particular, we create a ``hot'' and ``cool''
+\citet{flame_wave1}.  In particular, we create a ``hot'' and ``cool''
 hydrostatic model representing the ash and fuel states and blend the
@@
 Figures~\ref{fig:time_series_500} and \ref{fig:time_series_1000} show
-time series of the mean molecular weight for the 500~Hz and 1000~Hz
+time series of the mean molecular weight ($\bar{A}$) for the 500~Hz and 1000~Hz
 runs.  Compared to those in \cite{flame_wave1}, ashes behind the flame
@@
 For the purpose of measuring the flame propagation speed and acceleration, we track the position of each of our flames
-as a function of time. We define the position in terms of a particular $\enucdot$ value, as we did in \citet{flame_wave1}.
+as a function of time. We define the position in terms of a particular energy generation rate ($\enucdot$) value, as we did in \citet{flame_wave1}.
 To recapitulate: we first reduce the 2D $\enucdot$ data for each simulation run to a set of 1D radial profiles by averaging
@@
 
-\subsection{Effect of Temperature Structure}\label{ssec:temp_structure}
+\subsection{Effect of Temperature on Flame Structure}\label{ssec:temp_structure}
 
@@
 
-Figure~\ref{fig:compare_500Hz_abar} shows 3 different 500 Hz simulations with different initial temperature structures.
+Figure~\ref{fig:compare_500Hz_abar} shows three 500 Hz simulations with different initial temperature structures.
 The hottest of the three models, $\tt{F500\_3E8}$, has a faster propagating flame.
 It also appears to have a secondary burning region above the neutron star atmosphere layer (the atmosphere edge is located at $\sim 1.2\times 10^4$ cm).
-It is likely that a small amount of hot material broke through the top of the atmosphere.
+It is likely that a small amount of hot material broke out of the flame front and rose through the atmosphere.
 
-{\color{red} On the subject of things breaking out of the atmosphere, discuss interesting characteristics of the extreme 4E8 run here: maybe include a plot showing enuc at 40 ms comapared to one of the cooler runs, and another plot showing the momentum components?
-Seems like the flame moves very quickly across the domain, and then residual burning after the flame has passed eventually ignites the entire fuel layer.
-This total ignition can also occur in nature for a neutron star with a lot of dense and hot accreted material.}
+In contrast to the models with $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$, $\tt{F500\_4E8}$ is so hot that the organized flame front structure is lost.
+The fuel layer essentially simmers in place, rather than forming a flame; this effect can occur in nature, for neutron stars with very high accretion rates.
+\MarginPar{NF: I remember Alice mentioning this a long time ago, but is there anything I can cite to support it?}.
+This model burns so strongly that it is only run out to $40$ ms.
+Due to the lack of a clear burning front, we neglect to analyze the flame propagation speed and acceleration for this run.
+After an initial period of burning moves across the domain, residual burning continues and eventually ignites the entire fuel layer at late times, as shown in Figures \ref{fig:4e8_stacked_enuc} and \ref{fig:4e8_stacked_abar} for three snapshots taken in the last ten seconds of the simulation.
+In Figure~\ref{fig:4e8_stacked_enuc}, $\enucdot$ reaches values of $10^{18} - 10^{20}~\frac{\mathrm{erg}}{\mathrm{g} \cdot \mathrm{s}}$ across the domain and at heights up to $\sim 0.5\times 10^4$ cm.
+There is still significant burning occuring even higher, with $\enucdot$ reaching $\sim 10^{15}~\frac{\mathrm{erg}}{\mathrm{g} \cdot \mathrm{s}}$ at heights up to $\sim 0.8\times 10^4$ cm.
+For comparison, the next hottest run ($\tt{F500\_3E8}$) only reaches maximum $\enucdot$ values of $10^{18}~\frac{\mathrm{erg}}{\mathrm{g} \cdot \mathrm{s}}$ and a maximum height of $\sim 0.5\times 10^4$ cm, even at the latest timesteps ($\sim 70~\mathrm{ms}$, when the flame is most developed).
+Significant $\enucdot$ values for $\tt{F500\_3E8}$ are confined to the flame front, rather than spanning the entire domain.
+\MarginPar{NF: should I include a plot for this or is it ok to just say it?}. 
+Figure~\ref{fig:4e8_stacked_abar} plots $\bar{A}$ for $\tt{F500\_4E8}$, and similarly shows burning that extends across the domain and high into the atmosphere and fuel layer, lacking the charactersitic flame structure shown in $\bar{A}$ plots for the lower temperature $500$ Hz runs (see Figure~\ref{fig:compare_500Hz_abar}).
+A distinct mass of material - resembling a more dramatic version of that shown in Figure~\ref{fig:compare_500Hz_abar} - appears to have broken out of the atmosphere near the axis of symmetry.
+%Can also add a note here about how high abar values are reached across the domain, maybe don or alice can help give a more detailed analysis of abar.
+Maybe I could also segue into Alice's phase plots
 
@@
 	\centering
-	\plotone{multiplot_3x1_time_series_enuc_late.png}
+	\plotone{compare_500Hz_enuc_4e8.pdf}
 	\caption{\label{fig:4e8_stacked_enuc} Time series of the 500~Hz run $\tt{F500\_4E8}$, with $T_{\mathrm{star}} = 4 \times 10^8~\mathrm{K}$, showing $\dot{e}_{nuc}$.
@@
 	\centering
-	\plotone{multiplot_3x1_time_series_abar_late.png}
+	\plotone{compare_500Hz_abar_4e8.pdf}
 	\caption{\label{fig:4e8_stacked_abar} Time series of the 500~Hz run $\tt{F500\_4E8}$, with $T_{\mathrm{star}} = 4 \times 10^8~\mathrm{K}$, showing $\bar{A}$.}
 \end{figure}
-
-\begin{figure}[t]
-	\centering
-	\plotone{500Hz_stacked_enuc1_2order.pdf}
-	\caption{\label{fig:flame_speeds_2} Flame front position vs.\ time for the three 500 Hz runs with
-		$T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$. The dashed lines show quadratic least-squares fits to the data
-		for $t \gtrsim 0.02~\mathrm{s}$.}
-\end{figure}
-
-The method for measuring flame propagation described in Section \ref{ssec:rot_propagation} is applied in Figure \ref{fig:flame_speeds_2} to the three 500 Hz runs with $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$.
-As the initial $T_{\mathrm{star}}$ is increased, the flame becomes greatly accelerated.
-The initial flame velocities and accelerations derived from a quadratic least-squares fit to each of the datasets are provided in Table \ref{table:flame_speeds}.
-The reason for the acceleration of the flame over time is not entirely clear.
-It could be that energy released from burning begins to dominate the flame's propagation as it evolves, increasing the flame speed over time.
-Another possibility is that the increased temperatures lead to enhanced turbulent mixing effects that pull in more fresh fuel for the flame to burn.
-Yet another possibility is that the higher initial $T_{\mathrm{star}}$ leads to a greater average temperature in the fuel layer over time, making it easier for the flame to burn the fuel and propagate.
 
@@
     \caption{\label{fig:abar_hot}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $2\times10^8$~K and $3\times10^8$~K.}
+\end{figure}
+
+\subsection{Effect of Temperature on Flame Propagation}\label{ssec:temp_prop}
+
+The method for measuring flame propagation described in Section \ref{ssec:rot_propagation} is applied in Figure \ref{fig:flame_speeds_2} to the three 500 Hz runs with $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$.
+As the initial $T_{\mathrm{star}}$ is increased, the flame becomes greatly accelerated.
+The initial flame velocities and accelerations derived from a quadratic least-squares fit to each of the datasets are provided in Table \ref{table:flame_speeds}.
+The reason for the acceleration of the flame over time is not entirely clear.
+It could be that energy released from burning begins to dominate the flame's propagation as it evolves, increasing the flame speed over time.
+Another possibility is that the increased temperatures lead to enhanced turbulent mixing effects that pull in more fresh fuel for the flame to burn.
+Yet another possibility is that the higher initial $T_{\mathrm{star}}$ leads to a greater average temperature in the fuel layer over time, making it easier for the flame to burn the fuel and propagate.
+
+\begin{figure}[t]
+	\centering
+	\plotone{500Hz_stacked_enuc1_2order.pdf}
+	\caption{\label{fig:flame_speeds_2} Flame front position vs.\ time for the three 500 Hz runs with
+		$T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$. The dashed lines show quadratic least-squares fits to the data
+		for $t \gtrsim 0.02~\mathrm{s}$. Note that, due to its rapid acceleration, $\tt{F500\_3E8}$ (orange) is only run out to $\sim 70~\mathrm{ms}$ to avoid surpassing the domain boundary.}
 \end{figure}

### 2020-12-02 xrb2/paper.tex
add more citations

@@
 \newcommand{\omegadot}{\dot{\omega}}
-\newcommand{\enucdot}{\dot{e}_{nuc}}
+\newcommand{\enucdot}{\dot{e}_{\rm nuc}}
 
@@
 We investigate the effect of rotation rate on the flame.
-With the exception of (citation to the weird 11 Hz flame), most observations of XRBs come from sources with rotation rates of 200-600Hz \citep{bilous:2019,galloway:2020} \MarginPar{check whether this is true for XRBs or just burst oscillations}.
+With the exception of (citation to the weird 11 Hz flame \citep{altamirano2010discovery}), most observations of XRBs come from sources with rotation rates of 200-600Hz \citep{bilous:2019,galloway:2020} \MarginPar{check whether this is true for XRBs or just burst oscillations}.
 There are a number of factors that could explain this lack of observations below 200 Hz.
@@
 As some of the hotter 500~Hz runs exhibit significant accelerations, we fit a quadratic curve to each set of data and calculate the acceleration and initial velocity from the fit parameters.
-These results are given in Table~\ref{table:flame_speeds}.
+These results are given in Table~\ref{table:flame_speeds}\MarginPar{How is the acceleration calculated for this table?
+I.e.
+Specifically, at what time / over what time window is it calculated?}.
 
@@
 may be a direct result of the difference in Coriolis force. It could also be due to the 500 Hz flame taking more time to develop fully
-and reach higher burning rates, resulting in an $\epsdot$ increase over time.
+and reach higher burning rates, resulting in an $\epsdot$ \MarginPar{Is this supposed to be enucdot???}increase over time.
 
@@
 In contrast to the models with $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$, $\tt{F500\_4E8}$ is so hot that the organized flame front structure is lost.
-The fuel layer essentially simmers in place, rather than forming a flame; this effect can occur in nature, for neutron stars with very high accretion rates.
-\MarginPar{NF: I remember Alice mentioning this a long time ago, but is there anything I can cite to support it?}.
-This model burns so strongly that it is only run out to $40$ ms.
+The fuel layer essentially simmers in place, rather than forming a flame; this effect can occur in nature, for neutron stars with very high accretion rates \citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}.
+\MarginPar{NF: I remember Alice mentioning this a long time ago, but is there anything I can cite to support it?} This model burns so strongly that it is only run out to $40$ ms.
 Due to the lack of a clear burning front, we neglect to analyze the flame propagation speed and acceleration for this run.
 After an initial period of burning moves across the domain, residual burning continues and eventually ignites the entire fuel layer at late times, as shown in Figures \ref{fig:4e8_stacked_enuc} and \ref{fig:4e8_stacked_abar} for three snapshots taken in the last ten seconds of the simulation.
-In Figure~\ref{fig:4e8_stacked_enuc}, $\enucdot$ reaches values of $10^{18} - 10^{20}~\frac{\mathrm{erg}}{\mathrm{g} \cdot \mathrm{s}}$ across the domain and at heights up to $\sim 0.5\times 10^4$ cm.
-There is still significant burning occuring even higher, with $\enucdot$ reaching $\sim 10^{15}~\frac{\mathrm{erg}}{\mathrm{g} \cdot \mathrm{s}}$ at heights up to $\sim 0.8\times 10^4$ cm.
-For comparison, the next hottest run ($\tt{F500\_3E8}$) only reaches maximum $\enucdot$ values of $10^{18}~\frac{\mathrm{erg}}{\mathrm{g} \cdot \mathrm{s}}$ and a maximum height of $\sim 0.5\times 10^4$ cm, even at the latest timesteps ($\sim 70~\mathrm{ms}$, when the flame is most developed).
+In Figure~\ref{fig:4e8_stacked_enuc}, $\enucdot$ reaches values of $10^{18} - 10^{20}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ across the domain and at heights up to $\sim 0.5\times 10^4$ cm.
+There is still significant burning occuring even higher, with $\enucdot$ reaching $\sim 10^{15}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ at heights up to $\sim 0.8\times 10^4$ cm.
+For comparison, the next hottest run ($\tt{F500\_3E8}$) only reaches maximum $\enucdot$ values of $10^{18}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ and a maximum height of $\sim 0.5\times 10^4$ cm, even at the latest timesteps ($\sim 70~\mathrm{ms}$, when the flame is most developed).
 Significant $\enucdot$ values for $\tt{F500\_3E8}$ are confined to the flame front, rather than spanning the entire domain.

### 2020-12-02 xrb2/paper.tex
Added info to acceleration table caption

@@
 \begin{deluxetable}{lrrr}
-	\tablecaption{\label{table:flame_speeds} Flame speed and acceleration values measured for each simulation. $s_0$ is the speed at $t = 0$.}
+	\tablecaption{\label{table:flame_speeds} Flame speed and acceleration values measured for each simulation.
+$s_0$ is the speed at $t = 0$.
+The initial flame velocities and accelerations are derived from a quadratic least-squares fit to each of the datasets for times $t \gtrsim 0.02~\mathrm{s}$.}
 	\tablehead{\colhead{run} & \colhead{$s_0$ (cm s$^{-1}$)} & \colhead{$a$ (cm s$^{-2}$)}} 
@@
 I.e.
-Specifically, at what time / over what time window is it calculated?}.
+Specifically, at what time / over what time window is it calculated?} \MarginPar{NF: I will update the error estimates}.
 
@@
 		$T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$. The dashed lines show quadratic least-squares fits to the data
-		for $t \gtrsim 0.02~\mathrm{s}$. Note that, due to its rapid acceleration, $\tt{F500\_3E8}$ (orange) is only run out to $\sim 70~\mathrm{ms}$ to avoid surpassing the domain boundary.}
+		for $t \gtrsim 0.02~\mathrm{s}$. Note that, due to its rapid acceleration, $\tt{F500\_3E8}$ (orange) is only run out to $\sim 0.07~\mathrm{s}$ to avoid surpassing the domain boundary.}
 \end{figure}

### 2020-12-08 xrb2/paper.tex
some intro work

@@
 
+\author[0000-0001-8921-3624]{Nicole Ford}
+\affiliation{Lawrence Berkeley National Laboratory, Berkeley, CA}
+
 \author[0000-0001-8401-030X]{Michael Zingale}
@@
 \affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010}
-
-\author[0000-0001-8921-3624]{Nicole Ford}
-\affiliation{Lawrence Berkeley National Laboratory, Berkeley, CA}
 
@@
 \begin{abstract}
-We continue to investigate laterally propagating flames in XRBs...
+We continue to investigate laterally propagating flames in XRBs using
+fully compressible hydrodynamics simulations.  In the current study we
+relax the approximations used previously and explore the effect of
+rotation rate and the thermal structure of the neutron star on the
+flame.  We find that higher temperature atmospheres lead to faster
+flames and observe acceleration of the flame front as it propagates
+through the atmosphere.  All of the software used for these
+simulations is freely available.
 
@@
 
-Unlike our previous paper, all runs are now conducted without artificially boosting the speed of the flame in order to reduce the computational cost.
-This is possible because of the work done to offload the code to GPUs, where it runs significantly faster.
-We can therefore run the tests without any boosting.
+Considerable evidence suggests that ignition in an X-ray burst starts
+in a localized region and then spreads across the surface of the
+neutron star~\citep{bhattacharyya:2007,chakraborty:2014}.  We continue
+our study of flame spreading through fully compressible hydrodynamics
+simulations of the flame.  Building on our previous
+paper~\citep{flame_wave1}, we relax the approximations we made
+previously (artificially boosting the speed of the flame in order to
+reduce the computational cost) and explore how the flame properties
+depend on rotation rate and the thermal structure of the neutron
+star. This new set of realistic simulations is possible because of the
+work done to offload our simulation code, \castro~\cite{castro_joss}, to GPUs, where it
+runs significantly faster.
 

### 2020-12-08 xrb2/paper.tex
more intro

@@
 star. This new set of realistic simulations is possible because of the
-work done to offload our simulation code, \castro~\cite{castro_joss}, to GPUs, where it
+work done to offload our simulation code, \castro~\citep{castro_joss}, to GPUs, where it
 runs significantly faster.
 
-We investigate the effect of rotation rate on the flame.
-With the exception of (citation to the weird 11 Hz flame \citep{altamirano2010discovery}), most observations of XRBs come from sources with rotation rates of 200-600Hz \citep{bilous:2019,galloway:2020} \MarginPar{check whether this is true for XRBs or just burst oscillations}.
-There are a number of factors that could explain this lack of observations below 200 Hz.
-It could be that there is some physical process which inhibits the flame ignition and/or spread at lower rotation rates.
-It could be that burst oscillations at lower rotation rates are smaller in amplitude and therefore more difficult to detect.
-It could be that it's not got anything to do with the flame at all, but that neutron stars in accreting low mass X-ray binaries rarely have rotation rates below 200 Hz \MarginPar{look into this?}.
+We investigate the effect of rotation rate on the flame. With the
+exception of \citet{altamirano2010discovery}, most observations of
+XRBs come from sources with rotation rates of 200-600Hz
+\citep{bilous:2019,galloway:2020} \MarginPar{check whether this is
+  true for XRBs or just burst oscillations}. There are a number of
+factors that could explain this lack of observations below 200 Hz. It
+could be that there is some physical process which inhibits the flame
+ignition and/or spread at lower rotation rates. It could be that burst
+oscillations at lower rotation rates are smaller in amplitude and
+therefore more difficult to detect. It could be that it's not got
+anything to do with the flame at all, but that neutron stars in
+accreting low mass X-ray binaries rarely have rotation rates below 200
+Hz \MarginPar{look into this?}.
 
-Previous studies have found that rotation can have a significant affect on the flame's propagation.
-As the rotation rate increases, the Coriolis force whips the spreading flame up into a hurricane-like structure \citep{spitkovsky2002,cavecchi:2013}.
-The stronger Coriolis force leads to greater confinement of the hot accreted matter, leading to easier ignition of the flame \citep{art-2015-cavecchi-etal}.
+Previous studies have found that rotation can have a significant
+affect on the flame's propagation. As the rotation rate increases, the
+Coriolis force whips the spreading flame up into a hurricane-like
+structure \citep{spitkovsky2002,cavecchi:2013}. The stronger Coriolis
+force leads to greater confinement of the hot accreted matter, leading
+to easier ignition of the flame \citep{art-2015-cavecchi-etal}.
 
-Previous work has predicted that crustal heating is stronger a lower accretion rates and weaker at higher accretion rates, due to the effect of neutrino losses \citep{Cumming2006,johnston:2019}.
-In our models, we keep the crust at a constant temperature, so by increasing this temperature we can effectively increase the crustal heating and mimic the effects of lower accretion rates.
-A shallow heating mechanism of as yet unknown origin has been found necessary to reproduce observed properties of XRBs \citep{Deibel2015,Turlione2015,Keek2017}.
+The temperature structure of the accreted fuel layer can also affect
+the flame propagation.  \citet{Timmes00} showed that laminar helium flames 
+have higher speeds when moving into hotter upstream fuel.
+It is known that crustal heating is stronger at lower
+accretion rates and weaker at higher accretion rates, due to the
+effect of neutrino losses \citep{Cumming2006,johnston:2019}. In our
+models, we keep the crust at a constant temperature, so by increasing
+this temperature we can effectively increase the crustal heating and
+mimic the effects of lower accretion rates. A shallow heating
+mechanism of as yet unknown origin has been found necessary to
+reproduce observed properties of XRBs
+\citep{Deibel2015,Turlione2015,Keek2017}.
 

### 2020-12-08 xrb2/paper.tex
acknowledgments
more work on the paper

@@
 
-We use the Castro hydrodynamics code \citep{castro,castro_joss} and the simulation
-framework introduced in \citet{flame_wave1}.  The main difference is
-that we no longer boost the flames by adjusting the reaction rates and
-conductivities.  By porting \castro\ to GPUs \citep{sc20_gpu}, we
-are able to run the unboosted flame simulations with fewer resources
-than the boosted simulations used in our earlier paper.  For these
-axisymmetric simulations, we adding a missing geometric source terms
-from \citet{bernard-champmartin} that captures the effects of the
-divergence of the flux operating on the azimuthal unit vector.  This
-term is a small correction, but was missing from our earlier
-simulations.
+We use the Castro hydrodynamics code \citep{castro,castro_joss} and
+the simulation framework introduced in \citet{flame_wave1}.  The
+current simulations are all performed in a two-dimensional
+axisymmetric geometry.  For these axisymmetric simulations, we add a
+missing geometric source terms from \citet{bernard-champmartin} that
+captures the effects of the divergence of the flux operating on the
+azimuthal unit vector.  This term is a small correction, but was
+missing from our earlier simulations.  The simulations framework
+initializes a fuel layer in hydrostatic equilibrium, laterally
+blending a hot model on the left side of the domain (the coordinate
+origin) and a cool model on the right.  The initial temperature
+gradient between the hot and cool models drives a laterally
+propagating flame through the cool fuel.  In our original set of
+calculations, we artificially boosted the flame speed by adjusting the
+conductivity and reaction rate to produce a flame moving 5--10$\times$
+faster than the nominal laminar flame speed.  We also used high
+rotation rates to reduce the lateral lengthscale at which the Coriolis
+force balances the lateral flame spreading.  For the current
+simulations, we no longer boost the flame speed---the true
+conductivities (taken from \citealt{Timmes00}) and reaction rates are
+used.  We continue to use a 13-isotope $\alpha$-chain to describe the
+helium burning.  The port of \castro\ to GPUs~\citep{sc20_gpu} enables
+us to run these simulations without the previous approximations, while
+continuing to resolve the burning front.
 
@@
 
-\acknowledgements \castro\ is open-source and freely available at
+\acknowledgments \castro\ is open-source and freely available at
 \url{http://github.com/AMReX-Astro/Castro}.  The problem setups used

### 2020-12-08 xrb2/paper.tex
more work

@@
 coordinates and a smooth transition to the cooler region at larger
-radii.  In the calculations presented here, we explore the structure
-of the initial models by varying some of the parameters, in particular, the temperature
-of the neutron star and at the base of the fuel layer.
+radii.  The cool initial characterized by 3 temperatures:
+$T_\mathrm{star}$ is the isothermal temperature of the underlying
+neutron star, $T_\mathrm{hi}$ is the temperature at the base of the
+fuel layer, and $T_\mathrm{lo}$ is the minimum temperature of the
+atmosphere---the atmosphere structure is isentropic as it falls from
+$T_\mathrm{hi}$ down to $T_\mathrm{lo}$.  For the hot model, we
+replace $T_\mathrm{hi}$ with $T_\mathrm{hi} + \delta T$.  In the
+calculations presented here, we explore the structure of the initial
+models by varying these parameters, in particular, the
+temperature of the neutron star and at the base of the fuel layer.
 
@@
 \plotone{hot}
-\caption{\label{fig:hot_profiles} Initial temperature and density structure vs.\ height in the ``hot'' state.}
+\caption{\label{fig:hot_profiles} Initial temperature and density
+  structure vs.\ height in the ``hot'' state.}
 \end{figure}
@@
 \plotone{cold}
-\caption{\label{fig:cold_profiles} Initial temperature and density structure vs.\ height in the ``cool'' state.}
+\caption{\label{fig:cool_profiles} Initial temperature and density
+  structure vs.\ height in the ``cool'' state.}
 \end{figure}
 
-%======================================================================
-% Results
-%======================================================================
-\section{Simulations and Results}\label{Sec:results}
-
-In our latest set of runs, we explore a variety of initial rotation rate and temperature conditions for the flame. The initial conditions are provided in Table \ref{table:500Hz_runs}.
+For the current simulations, we explore a variety of initial rotation
+rate and temperature conditions for the flame. The main parameters
+describing the models are provided in Table
+\ref{table:500Hz_runs}. Figure~\ref{fig:hot_profiles} shows the
+temperature and density structure for our hot models and
+figure~\ref{fig:cool_profiles} shows the temperature and density
+structure for the cool models.
 
@@
 \end{deluxetable}
+
+%======================================================================
+% Results
+%======================================================================
+\section{Simulations and Results}\label{Sec:results}
 

### 2020-12-08 xrb2/paper.tex
fix figures

@@
 describing the models are provided in Table
-\ref{table:500Hz_runs}. Figure~\ref{fig:hot_profiles} shows the
+\ref{table:runs}. Figure~\ref{fig:hot_profiles} shows the
 temperature and density structure for our hot models and
@@
 \begin{deluxetable}{lrrrrr}
-	\tablecaption{\label{table:500Hz_runs} Rotation rate and temperature properties of the simulations.}
+	\tablecaption{\label{table:runs} Rotation rate and temperature properties of the simulations.}
 	\tablehead{\colhead{run} & \colhead{Rotation Rate (Hz)} & \colhead{$\delta T$ (K)} & \colhead{$T_\mathrm{hi}$ (K)} & \colhead{$T_\mathrm{star}$ (K)}} 
@@
 \section{Simulations and Results}\label{Sec:results}
+
+We present 6 simulations in total, summarized in
+Table~\ref{table:runs}.  These simulations encompass 3 different
+rotation rates: 250~Hz, 500~Hz, and 1000~Hz, and for the 500~Hz run, 4
+different temperature profiles.  In the following subsections, we look
+at how the flame properties depends on the model parameters.
 

### 2020-12-08 xrb2/paper.tex
ack tweak
update table
update figure
more text

@@
 calculations presented here, we explore the structure of the initial
-models by varying these parameters, in particular, the
-temperature of the neutron star and at the base of the fuel layer.
+models by varying these parameters.  All models have the same
+peak temperature in the hot model, $T_\mathrm{star} + \delta T$.
 
@@
 	\tablecaption{\label{table:runs} Rotation rate and temperature properties of the simulations.}
-	\tablehead{\colhead{run} & \colhead{Rotation Rate (Hz)} & \colhead{$\delta T$ (K)} & \colhead{$T_\mathrm{hi}$ (K)} & \colhead{$T_\mathrm{star}$ (K)}} 
+	\tablehead{\colhead{run} & \colhead{Rotation Rate (Hz)} & \colhead{$\delta T$ (K)} & \colhead{$T_\mathrm{hi}$ (K)} & \colhead{$T_\mathrm{star}$ (K)} & \colhead{$T_\mathrm{lo}$ (K)}} 
 	\startdata
-	{\tt F1000}     & $1000$ & $1.2\times 10^9$ & $2\times 10^8$ & $10^8$ \\
-	{\tt F500}      & $500$ & $1.2\times 10^9$ & $2\times 10^8$ & $10^8$ \\
-	{\tt F500\_2E8} & $500$ & $1.2\times 10^9$ & $2\times 10^8$ & $2\times 10^8$ \\
-	{\tt F500\_3E8} & $500$ & $1.1\times 10^9$ & $3\times 10^8$ & $3\times 10^8$ \\
-	{\tt F500\_4E8} & $500$ & $10^9$ & $4\times 10^8$ & $4\times 10^8$ \\
-	{\tt F250}      & $250$ & $1.2\times 10^9$ & $2\times 10^8$ & $10^8$ \\
+	{\tt F1000}     & $1000$ & $1.2\times 10^9$ & $2\times 10^8$ & $10^8$ & $8\times 10^6$ \\
+	{\tt F500}      & $500$ & $1.2\times 10^9$ & $2\times 10^8$ & $10^8$ & $8\times 10^6$ \\
+	{\tt F500\_2E8} & $500$ & $1.2\times 10^9$ & $2\times 10^8$ & $2\times 10^8$ & $8\times 10^6$ \\
+	{\tt F500\_3E8} & $500$ & $1.1\times 10^9$ & $3\times 10^8$ & $3\times 10^8$ & $8\times 10^6$ \\
+	{\tt F500\_4E8} & $500$ & $10^9$ & $4\times 10^8$ & $4\times 10^8$ & $8\times 10^6$ \\
+	{\tt F250}      & $250$ & $1.2\times 10^9$ & $2\times 10^8$ & $10^8$ & $8\times 10^6$ \\
 	\enddata
@@
 
-Things to show:
+We run 3 models (F250, F500, and F1000) with the same initial model but differing rotation rates.
 
@@
 \caption{\label{fig:compare_500Hz_abar} Comparison of $\bar{A}$ for 3 different 500~Hz models with different neutron star temperatures, $T_\star$, and resulting envelope structures.
-} %This will be updated to show at 70ms
+Each flame is shown at 70~ms.} %This will be updated to show at 70ms
 \end{figure}
@@
 \acknowledgments \castro\ is open-source and freely available at
-\url{http://github.com/AMReX-Astro/Castro}.  The problem setups used
-here are available in the git repo as {\tt flame} and {\tt
+\url{http://github.com/AMReX-Astro/Castro}.  The problem setup used
+here is available in the git repo as {\tt
   flame\_wave}.  The work at Stony Brook was supported by DOE/Office

### 2020-12-09 xrb2/paper.tex
some domain info

@@
 different temperature profiles.  In the following subsections, we look
-at how the flame properties depends on the model parameters.
+at how the flame properties depends on the model parameters.  All
+simulations are run in a domain $1.8432\times 10^5~\mathrm{cm} \times
+3.072\times 10^4~\mathrm{cm}$ with a coarse grid of $1152 \times 192$
+zones and 2 levels of refinment (the first a jump by a factor of $4$
+and the second a jump by a factor of $2$ finer).  This gives a
+fine-grid resolution of $20~\mathrm{cm}$.  The entire atmosphere is
+refined in these simulations, with the refinement tagging on all zones
+with $\rho > 2.5\times 10^4~\gcc$.
 

### 2020-12-16 xrb2/paper.tex
Minor edits throughout, added to section 3.1

--- 2020-12-08 xrb2/paper.tex
@@
 models, we keep the crust at a constant temperature, so by increasing
-this temperature we can effectively increase the crustal heating and
-mimic the effects of lower accretion rates. A shallow heating
+this temperature we can effectively increase the crustal heating and/or
+mimic the effects of higher accretion rates. A shallow heating
 mechanism of as yet unknown origin has been found necessary to
@@
 
-In the following sections, we conduct a series of simulations at various rotation rates to investigate its effect on the flame.
+In the following sections, we conduct a series of simulations at various rotation rates and temperatures to investigate their effects on the flame.
 We find that at lower rotation rates, the flame itself becomes harder to ignite.
+At higher rotation rates, nuclear burning is enhanced and the flame propagates more quickly.
+At high temperatures, burning is greatly enhanced and the flame accelerates as it propagates.
 We discuss the implications that this may have for burst physics and observations.
@@
 
-The initial model is setup in the same fashion as described in
+The initial model is set up in the same fashion as described in
 \citet{flame_wave1}.  In particular, we create a ``hot'' and ``cool''
@@
 
-We present 6 simulations in total, summarized in
-Table~\ref{table:runs}.  These simulations encompass 3 different
-rotation rates: 250~Hz, 500~Hz, and 1000~Hz, and for the 500~Hz run, 4
+We present six simulations in total, summarized in
+Table~\ref{table:runs}.  These simulations encompass three different
+rotation rates: 250~Hz, 500~Hz, and 1000~Hz, and for the 500~Hz run, four
 different temperature profiles.  In the following subsections, we look
-at how the flame properties depends on the model parameters.
+at how the flame properties depend on the model parameters.  All
+simulations are run in a domain of $1.8432\times 10^5~\mathrm{cm} \times
+3.072\times 10^4~\mathrm{cm}$ with a coarse grid of $1152 \times 192$
+zones and two levels of refinment (the first a jump by a factor of four
+and the second a jump by a factor of two finer).  This gives a
+fine-grid resolution of $20~\mathrm{cm}$.  The entire atmosphere is
+refined in these simulations, with the refinement tagging on all zones
+with $\rho > 2.5\times 10^4~\gcc$.
 
@@
 
-We run 3 models (F250, F500, and F1000) with the same initial model but differing rotation rates.
+We run three models ({\tt F250}, {\tt F500}, and {\tt F1000}) with the same initial model but differing rotation rates.
 
@@
 \plotone{time_series_500}
-\caption{\label{fig:time_series_500} Time series of the 500 Hz run showing $\bar{A}$.}
+\caption{\label{fig:time_series_500} Time series of {\tt F500} showing $\bar{A}$.}
 \end{figure}
@@
 \plotone{time_series_1000}
-\caption{\label{fig:time_series_1000} Time series of the 1000 Hz run showing $\bar{A}$.}
+\caption{\label{fig:time_series_1000} Time series of {\tt F1000} showing $\bar{A}$.}
 \end{figure}
+\MarginPar{update both the abar figures to only show 50 ms and 100 ms snapshots}
 
-Figures~\ref{fig:time_series_500} and \ref{fig:time_series_1000} show
-time series of the mean molecular weight ($\bar{A}$) for the 500~Hz and 1000~Hz
-runs.  Compared to those in \cite{flame_wave1}, ashes behind the flame
-do not reach as high atomic weights.  This is not surprising, since
-those early runs boosted the reaction rates.  Comparing these two new
-runs, we see that the burning is much more evolved for the higher
-rotation rate.  We believe that this is because the rotation confines
-the initial perturbation and subsequent expansion from the burning
-better, allowing the reactions to progress further.
-
-MZ: lateral profiles of X vs distance behind the flame; 
-vertical profiles of the atmosphere behind the flame---how does it compare to isentropic?
-
-what is the composition on the photosphere across the flame
-
-different rotation rates. 250 Hz failed to ignite, could indicate that another source of confinement (e.g.\ magnetic fields) would need to take
-over at lower rotation rates to allow a burst to occur, at least for our current initial model. We've seen that increasing rotation rate can
-increase nuclear reaction rate already
-
-shape of flame, amount of mixing (e.g. it looks like slower rotation rates have less mixed flames as the hurricane is less strong.)
-
-Don: look at ratio of enuc to e?
-
-\subsection{Effect of Rotation Rate on Flame Propagation}\label{ssec:rot_propagation}
-
-For the purpose of measuring the flame propagation speed and acceleration, we track the position of each of our flames
-as a function of time. We define the position in terms of a particular energy generation rate ($\enucdot$) value, as we did in \citet{flame_wave1}.
-To recapitulate: we first reduce the 2D $\enucdot$ data for each simulation run to a set of 1D radial profiles by averaging
-over the vertical coordinate. After averaging, we take our reference $\enucdot$ value to be some fraction of the global $\enucdot$
-maximum across all of these profiles. Since the flames in our simulations propagate rightward, we then search the region of
-each profile to the right of the local $\enucdot$ maximum for the point where the $\enucdot$ first drops below this reference
-value. This point gives us the location of our flame front.
-
-In \citet{flame_wave1}, we used $0.1\%$ of the global $\enucdot$ maximum for our reference value. For the high temperature
-unboosted flames, however, we found that the $\enucdot$ profiles failed to reach that small a value across the domain at most times,
-which prevented us from getting reliable position measurements. We therefore use $1\%$ of the global $\enucdot$ maximum in this paper
-rather than $0.1\%$. This is still sufficiently small that our measurements are not overly sensitive to turbulence and other local fluid
-motions (the issue with simply tracking the local maximum), but allows us to avoid the pitfall encountered by the $0.1\%$ metric.
-
-\begin{figure}[t]
-	\centering
-	\plotone{flame_speeds_rotation.pdf}
-	\caption{\label{fig:flame_speeds_1} Flame front position vs.\ time for the standard ($10^8~\mathrm{K}$) 500 Hz and 1000 Hz runs. The dashed lines
-		show quadratic least-squares fits to the data for $t \gtrsim 0.02~\mathrm{s}$.}
-\end{figure}
-
-\begin{deluxetable}{lrrr}
-	\tablecaption{\label{table:flame_speeds} Flame speed and acceleration values measured for each simulation.
-$s_0$ is the speed at $t = 0$.
-The initial flame velocities and accelerations are derived from a quadratic least-squares fit to each of the datasets for times $t \gtrsim 0.02~\mathrm{s}$.}
-	\tablehead{\colhead{run} & \colhead{$s_0$ (cm s$^{-1}$)} & \colhead{$a$ (cm s$^{-2}$)}} 
-	\startdata
-	{\tt F1000}     & $(3.41 \pm 0.01) \times 10^5$ & $(-1.03 \pm 0.13) \times 10^5$ \\
-	{\tt F500}      & $(3.08 \pm 0.01) \times 10^5$ & $(5.41 \pm 0.19) \times 10^5$ \\
-	{\tt F500\_2E8} & $(3.76 \pm 0.02) \times 10^5$ & $(1.74 \pm 0.17) \times 10^5$ \\
-	{\tt F500\_3E8} & $(5.29 \pm 0.10) \times 10^5$ & $(3.57 \pm 0.02) \times 10^7$ \\
-	\enddata
-	%
-\end{deluxetable}
-
-Figure~\ref{fig:flame_speeds_1} gives position vs.\ time for the $\tt{F500}$ and $\tt{F1000}$ runs to show the dependence on rotation rate.
-As some of the hotter 500~Hz runs exhibit significant accelerations, we fit a quadratic curve to each set of data and calculate the acceleration and initial velocity from the fit parameters.
-These results are given in Table~\ref{table:flame_speeds}\MarginPar{How is the acceleration calculated for this table?
-I.e.
-Specifically, at what time / over what time window is it calculated?} \MarginPar{NF: I will update the error estimates}.
-
-We found in \citet{flame_wave1} that the flame speed $s$ obeys:
-\begin{equation}
-	s \appropto L_R (\epsdot)^{\half} \propto \frac{(\epsdot)^{\half}}{\Omega}
-\end{equation}
-where $\epsdot$ is the specific energy generation rate and $L_R$ is the Rossby length. This was consistent with the results of
-\citet{cavecchi:2013}, who noted that the flame speeds in their simulations scaled with the Rossby length. As seen in
-Figure~\ref{fig:flame_speeds_1}, there is no obvious inverse scaling with rotation rate in this set of runs. We observed earlier,
-however, that nuclear reactions progress more quickly at the higher rotation rate. This results in a higher $\enucdot$ --- up to 4
-to 6 times higher near the burning front after the flame ignites --- which may counteract the reduction in flame speed from the
-increased Coriolis confinement. We also observe that the 500 Hz run accelerates faster than the 1000 Hz run, which appears to experience a
-small deceleration at early times. \MarginPar{This is speculation on my part -- does someone else have clear idea why?} This disparity
-may be a direct result of the difference in Coriolis force. It could also be due to the 500 Hz flame taking more time to develop fully
-and reach higher burning rates, resulting in an $\epsdot$ \MarginPar{Is this supposed to be enucdot???}increase over time.
-
-% It might be worth plotting peak enucdot as a function of time for each run
+Figures~\ref{fig:time_series_500} and \ref{fig:time_series_1000} show time series of the mean molecular weight ($\bar{A}$) for the {\tt F500} and {\tt F1000} runs.
+Compared to those in \cite{flame_wave1}, ashes behind the flame do not reach as high atomic weights.
+This is not surprising, since those early runs boosted the reaction rates.
+Comparing these two new runs, the burning is much more evolved for the higher rotation rate.
+We believe that this is because the increased rotation confines the initial perturbation and subsequent expansion from the burning better, allowing the reactions to progress further.
+In contrast to {\tt F500} and {\tt F1000}, the lowest rotation run --- {\tt F250} --- failed to ignite.
+The lack of ignition for {\tt F250} also aligns with the reasoning given above, with the lower rotation in this case potentially leading to insufficient confinement of the burning.
+In this scenario, another source of confinement (e.g.\ magnetic fields) would need to take over at lower rotation rates to allow a burst to occur, at least for our current initial model.
 
@@
 
+MZ: lateral profiles of X vs distance behind the flame; 
+vertical profiles of the atmosphere behind the flame---how does it compare to isentropic?
+
+what is the composition on the photosphere across the flame
+
+shape of flame, amount of mixing (e.g. it looks like slower rotation rates have less mixed flames as the hurricane is less strong.)
+
+Don: look at ratio of enuc to e?
+
+\subsection{Effect of Rotation Rate on Flame Propagation}\label{ssec:rot_propagation}
+
+For the purpose of measuring the flame propagation speed and acceleration, we track the position of each of our flames
+as a function of time. We define the position in terms of a particular energy generation rate ($\enucdot$) value, as we did in \citet{flame_wave1}.
+To recapitulate: we first reduce the 2D $\enucdot$ data for each simulation run to a set of 1D radial profiles by averaging
+over the vertical coordinate. After averaging, we take our reference $\enucdot$ value to be some fraction of the global $\enucdot$
+maximum across all of these profiles. Since the flames in our simulations propagate rightward, we then search the region of
+each profile to the right of the local $\enucdot$ maximum for the point where the $\enucdot$ first drops below this reference
+value. This point gives us the location of our flame front.
+
+In \citet{flame_wave1}, we used $0.1\%$ of the global $\enucdot$ maximum for our reference value. For the high temperature
+unboosted flames, however, we found that the $\enucdot$ profiles failed to reach that small a value across the domain at most times,
+which prevented us from getting reliable position measurements. We therefore use $1\%$ of the global $\enucdot$ maximum in this paper
+rather than $0.1\%$. This is still sufficiently small that our measurements are not overly sensitive to turbulence and other local fluid
+motions (the issue with simply tracking the local maximum), but allows us to avoid the pitfall encountered by the $0.1\%$ metric.
+
+\begin{deluxetable}{lrrr}
+	\tablecaption{\label{table:flame_speeds} Flame speed and acceleration values measured for each simulation.
+$s_0$ is the speed at $t = 0$.
+The initial flame velocities and accelerations are derived from a quadratic least-squares fit to each of the datasets for times $t \gtrsim 0.02~\mathrm{s}$.}
+	\tablehead{\colhead{run} & \colhead{$s_0$ (cm s$^{-1}$)} & \colhead{$a$ (cm s$^{-2}$)}} 
+	\startdata
+	{\tt F1000}     & $(3.41 \pm 0.01) \times 10^5$ & $(-1.03 \pm 0.13) \times 10^5$ \\
+	{\tt F500}      & $(3.08 \pm 0.01) \times 10^5$ & $(5.41 \pm 0.19) \times 10^5$ \\
+	{\tt F500\_2E8} & $(3.76 \pm 0.02) \times 10^5$ & $(1.74 \pm 0.17) \times 10^5$ \\
+	{\tt F500\_3E8} & $(5.29 \pm 0.10) \times 10^5$ & $(3.57 \pm 0.02) \times 10^7$ \\
+	\enddata
+	%
+\end{deluxetable}
+
+\begin{figure}[t]
+	\centering
+	\plotone{flame_speeds_rotation.pdf}
+	\caption{\label{fig:flame_speeds_1} Flame front position vs.\ time for the standard ($10^8~\mathrm{K}$) 500 Hz and 1000 Hz runs. The dashed lines
+		show quadratic least-squares fits to the data for $t \gtrsim 0.02~\mathrm{s}$.}
+\end{figure}
+
+Figure~\ref{fig:flame_speeds_1} gives position vs.\ time for the $\tt{F500}$ and $\tt{F1000}$ runs to show the dependence on rotation rate.
+As some of the hotter 500~Hz runs exhibit significant accelerations, we fit a quadratic curve to each set of data and calculate the acceleration and initial velocity from the fit parameters.
+These results are given in Table~\ref{table:flame_speeds}.
+The initial flame velocities and accelerations are derived from a quadratic least-squares fit to each of the datasets for times $t \gtrsim 0.02~\mathrm{s}$.
+\MarginPar{How is the acceleration calculated for this table?
+I.e.
+Specifically, at what time / over what time window is it calculated?} \MarginPar{NF: I will update the error estimates}.
+
+We found in \citet{flame_wave1} that the flame speed $s$ obeys:
+\begin{equation}
+	s \appropto L_R (\epsdot)^{\half} \propto \frac{(\epsdot)^{\half}}{\Omega}
+\end{equation}
+where $\epsdot$ is the specific energy generation rate and $L_R$ is the Rossby length. This was consistent with the results of
+\citet{cavecchi:2013}, who noted that the flame speeds in their simulations scaled with the Rossby length. As seen in
+Figure~\ref{fig:flame_speeds_1}, there is no obvious inverse scaling with rotation rate in this set of runs. We observed earlier,
+however, that nuclear reactions progress more quickly at the higher rotation rate. This results in a higher $\enucdot$ --- up to four
+to six times higher near the burning front after the flame ignites --- which may counteract the reduction in flame speed from the
+increased Coriolis confinement. We also observe that the 500 Hz run accelerates faster than the 1000 Hz run, which appears to experience a
+small deceleration at early times. \MarginPar{This is speculation on my part -- does someone else have clear idea why?} This disparity
+may be a direct result of the difference in Coriolis force. It could also be due to the 500 Hz flame taking more time to develop fully
+and reach higher burning rates, resulting in an $\epsdot$ \MarginPar{Is this supposed to be enucdot???}increase over time.
+
+% It might be worth plotting peak enucdot as a function of time for each run
+
 \subsection{Effect of Temperature on Flame Structure}\label{ssec:temp_structure}
 
-To explore the effect of different initial temperature configurations, we run four simulations fixed at a rotation rate of 500 Hz with temperatures as shown in Table \ref{table:500Hz_runs}.
+To explore the effect of different initial temperature configurations, we run four simulations fixed at a rotation rate of 500 Hz with temperatures as shown in Table \ref{table:runs}.
 For all the 500 Hz simulations (with the exception of the coolest run, $\tt{F500}$), we set $T_\mathrm{star}$ = $T_\mathrm{hi}$, scaling $\delta T$ accordingly to maintain a consistent value of $T_\mathrm{hi} + \delta T$.
@@
 It is likely that a small amount of hot material broke out of the flame front and rose through the atmosphere.
+This effect, however, may be numerical rather than physical, resulting from the boundary conditions imposed at the axis of symmetry for these simulations.
 
 In contrast to the models with $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$, $\tt{F500\_4E8}$ is so hot that the organized flame front structure is lost.
-The fuel layer essentially simmers in place, rather than forming a flame; this effect can occur in nature, for neutron stars with very high accretion rates \citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}.
-\MarginPar{NF: I remember Alice mentioning this a long time ago, but is there anything I can cite to support it?} This model burns so strongly that it is only run out to $40$ ms.
+The fuel layer undergoes stable burning and essentially simmers in place, rather than forming a flame; this effect can occur in nature, for neutron stars with accretion that balances the Eddington luminosity \citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}.
+This model burns so strongly that it is only run out to $40$ ms.
 Due to the lack of a clear burning front, we neglect to analyze the flame propagation speed and acceleration for this run.
@@
 A distinct mass of material - resembling a more dramatic version of that shown in Figure~\ref{fig:compare_500Hz_abar} - appears to have broken out of the atmosphere near the axis of symmetry.
+Again, this may just be a result of the imposed boundary conditions rather than a physical effect.
 %Can also add a note here about how high abar values are reached across the domain, maybe don or alice can help give a more detailed analysis of abar.

### 2020-12-17 xrb2/paper.tex
made phase plot bullets into paragraphs

@@
 
+The {\tt F500} and {\tt F1000} runs both qualitatively resemble the flame structure in \citet{flame_wave1} --- a laterally propagating flame that is lifted off of the bottom of the fuel layer --- but they differ in their burning structures.
 Figures~\ref{fig:time_series_500} and \ref{fig:time_series_1000} show time series of the mean molecular weight ($\bar{A}$) for the {\tt F500} and {\tt F1000} runs.
@@
 This is not surprising, since those early runs boosted the reaction rates.
-Comparing these two new runs, the burning is much more evolved for the higher rotation rate.
+Comparing these two new runs, the burning is much more evolved for the higher rotation rate, and the ash is actually able to move ahead of the flame front (visible in the Figure  \ref{fig:time_series_1000} 100 ms snapshot).
 We believe that this is because the increased rotation confines the initial perturbation and subsequent expansion from the burning better, allowing the reactions to progress further.
@@
 
-Looking at the phase plots in Figures~\ref{fig:urho}-\ref{fig:abar}, we see:
-\begin{itemize}
-    \item From the $u$-$\rho$ and $u$-$v$ plots, we see that the burning occurs in a region of high density where the fluid is moving in a circular motion.
-From the previous plots (see: streamlines?), this most likely corresponds to the leading edge of the flame.
-This feature is not (yet?) developed in the 250~Hz flame.
-(will this ever develop, or will it just fizzle out completely?)
-    \item From the $\bar{A}$-$T$ plots in Figure \ref{fig:abar}, we can see that as the rotation rate increases, so does the peak temperature.
+Burning in the {\tt F500} and {\tt F1000} runs is concentrated in a dense region with circular motion.
+Figure \ref{fig:urho} compares the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for {\tt F250}, {\tt F500}, and {\tt F1000}.
+Most of the burning for each of the simulations occurs in a high density region $\rho > 3 \times 10^5 {\rm g/cm^3}$.
+The fluid in this dense, high energy generation region undergoes vortical motion, shown in the Figure \ref{fig:uv} phase plots comparing $u$, $v$ and $\dot{e}_{\rm nuc}$.
+From the previous plots (see: streamlines?) \MarginPar{NF: which plots?}, this most likely corresponds to the leading edge of the flame where fresh fuel is being entrained.
+This feature is not developed in the 250~Hz flame; it could potentially develop at later times, or the flame could just fizzle out.
+
+The mean molecular weight $\bar{A}$ within each of our simulations seems to grow along defined tracks confined to certain temperatures $T$, as shown in the Figure \ref{fig:abar} phase plots.
 We believe that the tracks in the plot correspond to reaction pathways in the reaction network.
-In the 1000~Hz plot, these tracks are somewhat disrupted, possibly due to the more vigorous mixing of the vortex at the flame front.
-(If this continues or if the rotation was cranked even higher, would the reactions get so disrupted that the burning is snuffed out, similar to what happens in combustion when the turbulent vortices become so small they penetrate the reaction zone and completely disrupt it?)
-\end{itemize}
+Comparing Figure \ref{fig:abar} to \citet{flame_wave1}, these tracks are much more neat and clearly defined.
+The ''messiness" of the tracks may be dependent on how mixed the flame interior is --- since these new simulations are un-boosted, they may be inherently less mixed than those in \citet{flame_wave1}.
+{\tt F1000} aligns with this interpretation: its $\bar{A}$ tracks are somewhat disrupted compared to the slower rotation runs, possibly due to the more vigorous mixing of the vortex at the flame front.
+Comparing the different runs, we also see that as the rotation rate increases, so does the peak temperature.
+This makes sense if higher rotation leads to a more confined, intense vortex near the flame front.
+It also agrees with our earlier interpretation of enhanced burning seen in Figure \ref{fig:time_series_1000} for {\tt F1000}.
+The disruption of reaction pathways with higher rotation rates and mixed flames may have interesting consequences; if the reactions are disrupted enough, it is possible that the burning could be snuffed out.
+This can occur in combustion scenarios when the turbulent vortices become so small they penetrate the reaction zone and completely disrupt the burning.
 
@@
     \centering
-    \plotone{urho_plot_0_1}
+    \plotone{urho_plot_0_1}{}
     \caption{\label{fig:urho}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and 1000~Hz runs.}

### 2020-12-18 xrb2/paper.tex
added to section 3.2

@@
 Comparing the different runs, we also see that as the rotation rate increases, so does the peak temperature.
-This makes sense if higher rotation leads to a more confined, intense vortex near the flame front.
+This makes sense if higher rotation leads to a more concentrated, intense vortex near the flame front.
 It also agrees with our earlier interpretation of enhanced burning seen in Figure \ref{fig:time_series_1000} for {\tt F1000}.
@@
 
-Figure~\ref{fig:flame_speeds_1} gives position vs.\ time for the $\tt{F500}$ and $\tt{F1000}$ runs to show the dependence on rotation rate.
-As some of the hotter 500~Hz runs exhibit significant accelerations, we fit a quadratic curve to each set of data and calculate the acceleration and initial velocity from the fit parameters.
+Figure~\ref{fig:flame_speeds_1} gives position vs.\ time for the $\tt{F500}$ and $\tt{F1000}$ runs (blue and orange, respectively) to show the dependence on rotation rate.
+As some of the hotter 500~Hz runs exhibit significant accelerations, we fit a quadratic curve to each set of data for times $t \gtrsim 0.02~\mathrm{s}$ and calculate the acceleration and initial velocity from the fit parameters.
 These results are given in Table~\ref{table:flame_speeds}.
-The initial flame velocities and accelerations are derived from a quadratic least-squares fit to each of the datasets for times $t \gtrsim 0.02~\mathrm{s}$.
 \MarginPar{How is the acceleration calculated for this table?
@@
 \begin{equation}
+	\label{eqn:rossby}
 	s \appropto L_R (\epsdot)^{\half} \propto \frac{(\epsdot)^{\half}}{\Omega}
 \end{equation}
-where $\epsdot$ is the specific energy generation rate and $L_R$ is the Rossby length. This was consistent with the results of
-\citet{cavecchi:2013}, who noted that the flame speeds in their simulations scaled with the Rossby length. As seen in
-Figure~\ref{fig:flame_speeds_1}, there is no obvious inverse scaling with rotation rate in this set of runs. We observed earlier,
-however, that nuclear reactions progress more quickly at the higher rotation rate. This results in a higher $\enucdot$ --- up to four
-to six times higher near the burning front after the flame ignites --- which may counteract the reduction in flame speed from the
-increased Coriolis confinement. We also observe that the 500 Hz run accelerates faster than the 1000 Hz run, which appears to experience a
-small deceleration at early times. \MarginPar{This is speculation on my part -- does someone else have clear idea why?} This disparity
-may be a direct result of the difference in Coriolis force. It could also be due to the 500 Hz flame taking more time to develop fully
-and reach higher burning rates, resulting in an $\epsdot$ \MarginPar{Is this supposed to be enucdot???}increase over time.
+where $\epsdot$ is the specific energy generation rate and $L_R$ is the Rossby length.
+This was consistent with the results of \citet{cavecchi:2013}, who noted that the flame speeds in their simulations scaled with the Rossby length.
+As seen in Figure~\ref{fig:flame_speeds_1}, there is no obvious inverse scaling of the flame speed with rotation rate in this set of runs.
+We observed earlier, however, that nuclear reactions progress more quickly at the higher rotation rate.
+This results in a higher $\enucdot$ --- up to four to six times higher near the burning front after the flame ignites --- which may counteract the reduction in flame speed from the increased Coriolis confinement.
+The initial position of $\tt{F1000}$ does appear to be reduced compared to the slower $\tt{F500}$.
+From Equation \ref{eqn:rossby}, increasing the rotation rate from 500 Hz to 1000 Hz should decrease $L_R$ by a factor of two, and the initial position of the flame should be reduced by a similar factor.
+Qualitatively, we think that the higher rotation $\tt{F1000}$ run starts with a tighter votex due to the increased Coriolis force, leading to a lower initial position for the flame front.
+$L_R$ is not, however, a perfect metric for the flame lengthscale.
+The scaling described in Equation \ref{eqn:rossby} seems to be more accurate at higher rotation rates (1000 Hz and up), based on \citet{flame_wave1}.
+We also observe that $\tt{F500}$ accelerates faster than $\tt{F1000}$, which appears to experience a small deceleration at early times.
+\MarginPar{This is speculation on my part -- does someone else have clear idea why?} This disparity may be a direct result of the difference in Coriolis force.
+It could also be due to the 500 Hz flame taking more time to develop fully and reach higher burning rates, resulting in an $\epsdot$ \MarginPar{Is this supposed to be enucdot???}increase over time.
 

### 2020-12-19 xrb2/paper.tex
Minor edits, added to section 3.3

@@
 This makes sense if higher rotation leads to a more concentrated, intense vortex near the flame front.
-It also agrees with our earlier interpretation of enhanced burning seen in Figure \ref{fig:time_series_1000} for {\tt F1000}.
+It also agrees with our earlier interpretation of the enhanced burning seen in Figure \ref{fig:time_series_1000} for {\tt F1000}.
 The disruption of reaction pathways with higher rotation rates and mixed flames may have interesting consequences; if the reactions are disrupted enough, it is possible that the burning could be snuffed out.
@@
 This results in a higher $\enucdot$ --- up to four to six times higher near the burning front after the flame ignites --- which may counteract the reduction in flame speed from the increased Coriolis confinement.
-The initial position of $\tt{F1000}$ does appear to be reduced compared to the slower $\tt{F500}$.
-From Equation \ref{eqn:rossby}, increasing the rotation rate from 500 Hz to 1000 Hz should decrease $L_R$ by a factor of two, and the initial position of the flame should be reduced by a similar factor.
-Qualitatively, we think that the higher rotation $\tt{F1000}$ run starts with a tighter votex due to the increased Coriolis force, leading to a lower initial position for the flame front.
+The initial position of $\tt{F1000}$ does appear to be reduced compared to the lower rotation $\tt{F500}$ run.
+From Equation \ref{eqn:rossby}, increasing the rotation rate from 500 Hz to 1000 Hz should decrease $L_R$ by a factor of two, and the initial position of the flame should be reduced by a similar factor (again, due to greater confinement from the Coriolis force).
 $L_R$ is not, however, a perfect metric for the flame lengthscale.
@@
 
+ \MarginPar{NF: I think we should plot peak enucdot as a function of time for each run, it would be useful for me to reference it in one of the later sections}
+
 To explore the effect of different initial temperature configurations, we run four simulations fixed at a rotation rate of 500 Hz with temperatures as shown in Table \ref{table:runs}.
@@
 If we let $T_\mathrm{star} < T_\mathrm{hi}$, the cooler neutron star surface might act as a heat sink and siphon away energy that would otherwise go into heating the burning layer.
+By choosing $T_\mathrm{star}$ = $T_\mathrm{hi}$, we can effectively explore simulations with greater surface heating.
+There are two physically distinct mechanisms that could each produce an increased temperature at the crust: crustal heating or a higher accretion rate.
+In these simulations, we cannot distinguish between whether we are generating heat due to a higher accretion rate or due to a shallow crustal heating mechanism, because we are not resolving the crust.
+Instead, we are modeling the resulting heating effect.
 
@@
 
-Figure~\ref{fig:compare_500Hz_abar} shows three 500 Hz simulations with different initial temperature structures.
-The hottest of the three models, $\tt{F500\_3E8}$, has a faster propagating flame.
-It also appears to have a secondary burning region above the neutron star atmosphere layer (the atmosphere edge is located at $\sim 1.2\times 10^4$ cm).
-It is likely that a small amount of hot material broke out of the flame front and rose through the atmosphere.
+Figure~\ref{fig:compare_500Hz_abar} shows $\bar{A}$ for three 500 Hz simulations with different initial temperature structures at $t = 70~\mathrm{ms}$.
+The hottest of the three models, $\tt{F500\_3E8}$, has a faster propagating flame (discussed further in Section \ref{ssec:temp_prop}).
+It also appears to have a secondary burning region (colored light orange) forming on the far left above the flame edge.
+It is likely that a small amount of hot material broke out of the flame and rose through the atmosphere.
 This effect, however, may be numerical rather than physical, resulting from the boundary conditions imposed at the axis of symmetry for these simulations.
 
-In contrast to the models with $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$, $\tt{F500\_4E8}$ is so hot that the organized flame front structure is lost.
+\begin{figure}[t]
+    \centering
+    \plotone{abar_plot_hotter}
+    \caption{\label{fig:abar_hot}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $2\times10^8$~K and $3\times10^8$~K.}
+\end{figure}
+
+Similar to the $\bar{A}$ trends found for different rotation rates in Figure \ref{fig:abar}, the $\bar{A}$-$T$ phase plot in Figure~\ref{fig:abar_hot} suggests that the actual reactions happening in the hottest run may be different than for the cooler runs.
+Not only is the $\tt{F500\_3E8}$ flame itself significantly hotter overall, but there is a large causally connected region across a wide range of $\bar{A}$.
+This indicates that the hotter temperature has facilitated significant burning in reaction pathways that were not favored at the lower temperatures.
+The reaction pathways are also very disrupted for $\tt{F500\_3E8}$ compared to the cooler runs, suggesting that the hotter temperature leads to more turbulent mixing.
+Compared to the highest rotation rate run, $\tt{F1000}$, $\tt{F500\_3E8}$ reaches a peak temperature roughly $1 \times 10^9~\mathrm{K}$ higher ($\sim 2.7 \times 10^9~\mathrm{K}$ versus $\sim 1.7 \times 10^9~\mathrm{K}$).
+Though increasing the rotation rate does seem to concentrate the burning and increase the peak temperature, it does not reach the same level as directly increasing the temperature.
+
+In contrast to the models with $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$, $\tt{F500\_4E8}$ is so hot that the organized flame structure is lost.
 The fuel layer undergoes stable burning and essentially simmers in place, rather than forming a flame; this effect can occur in nature, for neutron stars with accretion that balances the Eddington luminosity \citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}.
 This model burns so strongly that it is only run out to $40$ ms.
-Due to the lack of a clear burning front, we neglect to analyze the flame propagation speed and acceleration for this run.
 After an initial period of burning moves across the domain, residual burning continues and eventually ignites the entire fuel layer at late times, as shown in Figures \ref{fig:4e8_stacked_enuc} and \ref{fig:4e8_stacked_abar} for three snapshots taken in the last ten seconds of the simulation.
@@
 \MarginPar{NF: should I include a plot for this or is it ok to just say it?}. 
-Figure~\ref{fig:4e8_stacked_abar} plots $\bar{A}$ for $\tt{F500\_4E8}$, and similarly shows burning that extends across the domain and high into the atmosphere and fuel layer, lacking the charactersitic flame structure shown in $\bar{A}$ plots for the lower temperature $500$ Hz runs (see Figure~\ref{fig:compare_500Hz_abar}).
+Figure~\ref{fig:4e8_stacked_abar} shows $\bar{A}$ for $\tt{F500\_4E8}$.
+Again, burning extends across the domain and high into the atmosphere and fuel layer, lacking the charactersitic flame structure shown in $\bar{A}$ plots for the lower temperature $500$ Hz runs (see Figure~\ref{fig:compare_500Hz_abar}).
 A distinct mass of material - resembling a more dramatic version of that shown in Figure~\ref{fig:compare_500Hz_abar} - appears to have broken out of the atmosphere near the axis of symmetry.
-Again, this may just be a result of the imposed boundary conditions rather than a physical effect.
-%Can also add a note here about how high abar values are reached across the domain, maybe don or alice can help give a more detailed analysis of abar.
-Maybe I could also segue into Alice's phase plots
+The atmosphere edge is located at $\sim 1.2\times 10^4~\mathrm{cm}$.
+To reiterate, this may just be a result of the imposed boundary conditions rather than a physical effect.
+$\tt{F500\_4E8}$ clearly reaches much higher $\bar{A}$ values across the domain (especially at the latest snapshot, $t = 40~\mathrm{ms}$) compared to all the other runs described in this paper (see Figures \ref{fig:compare_500Hz_abar}, \ref{fig:time_series_500}, and \ref{fig:time_series_1000}).
+This is not surprising, since as we have seen through our varied temperature simulations, the burning is very temperature-sensitive.
 
@@
 
-Looking at the phase plots in Figures~\ref{fig:urho_hot}-\ref{fig:abar_hot}, we see:
-\begin{itemize}
-    \item The $u$-$\rho$ plot in Figure~\ref{fig:urho_hot} shows that the horizontal velocity when $T_{\rm star} = 3\times10^8$~K is sufficiently greater than for the two cooler runs.
-This is consistent with the plot of the flame front position, where we saw that the hottest flame was significantly faster at later times than the cooler flames.
-    \item The $u$-$v$ plot in Figure~\ref{fig:uv_hot} further demonstrates this, as again we see that the magnitude of the velocity of the flame for the hottest run is significantly larger.
-Also of interest in these plots is that the coolest run does not seem to have developed the characteristic vortex structure at the burning front (i.e.
-where $\dot{e}_{\rm nuc}$ is greatest) that can be clearly seen for the hotter runs.
-    \item The $\overline{A}$-$T$ plot in Figure~\ref{fig:abar_hot} suggests that the actual reactions happening in the hottest run may be different than for the cooler runs.
-Not only is the flame itself significantly hotter overall, but there is a large causally connected region across a wide range of $\overline{A}$.
-This indicates that the hotter temperature has facilitated significant burning in reaction pathways that were not favored at the lower temperatures.
-\end{itemize}
-
-\begin{figure}[t]
-    \centering
-    \plotone{urho_plot_hotter}
-    \caption{\label{fig:urho_hot}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $2\times10^8$~K and $3\times10^8$~K.}
-\end{figure}
-
-\begin{figure}[t]
-    \centering
-    \plotone{uv_plot_hotter}
-    \caption{\label{fig:uv_hot}Phase plot of the horizontal $x$-velocity $u$, vertical $y$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $2\times10^8$~K and $3\times10^8$~K.}
-\end{figure}
-
-\begin{figure}[t]
-    \centering
-    \plotone{abar_plot_hotter}
-    \caption{\label{fig:abar_hot}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $2\times10^8$~K and $3\times10^8$~K.}
-\end{figure}
-
 \subsection{Effect of Temperature on Flame Propagation}\label{ssec:temp_prop}
@@
 The method for measuring flame propagation described in Section \ref{ssec:rot_propagation} is applied in Figure \ref{fig:flame_speeds_2} to the three 500 Hz runs with $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$.
+Due to the lack of a clear burning front in $\tt{F500\_4E8}$, we neglect to analyze its propagation speed and acceleration.
 As the initial $T_{\mathrm{star}}$ is increased, the flame becomes greatly accelerated.
@@
 		for $t \gtrsim 0.02~\mathrm{s}$. Note that, due to its rapid acceleration, $\tt{F500\_3E8}$ (orange) is only run out to $\sim 0.07~\mathrm{s}$ to avoid surpassing the domain boundary.}
+\end{figure}
+
+Looking at the phase plots in Figures~\ref{fig:urho_hot}-\ref{fig:abar_hot}, we see:
+\begin{itemize}
+    \item The $u$-$\rho$ plot in Figure~\ref{fig:urho_hot} shows that the horizontal velocity when $T_{\rm star} = 3\times10^8$~K is sufficiently greater than for the two cooler runs.
+This is consistent with the plot of the flame front position, where we saw that the hottest flame was significantly faster at later times than the cooler flames.
+    \item The $u$-$v$ plot in Figure~\ref{fig:uv_hot} further demonstrates this, as again we see that the magnitude of the velocity of the flame for the hottest run is significantly larger.
+Also of interest in these plots is that the coolest run does not seem to have developed the characteristic vortex structure at the burning front (i.e.
+where $\dot{e}_{\rm nuc}$ is greatest) that can be clearly seen for the hotter runs.
+\end{itemize}
+
+\begin{figure}[t]
+    \centering
+    \plotone{urho_plot_hotter}
+    \caption{\label{fig:urho_hot}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $2\times10^8$~K and $3\times10^8$~K.}
+\end{figure}
+
+\begin{figure}[t]
+    \centering
+    \plotone{uv_plot_hotter}
+    \caption{\label{fig:uv_hot}Phase plot of the horizontal $x$-velocity $u$, vertical $y$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $2\times10^8$~K and $3\times10^8$~K.}
 \end{figure}

### 2020-12-23 xrb2/paper.tex
add new plots
authorshipo
added acknowledgments

@@
 
+\author[0000-0001-8921-3624]{Nicole Ford}
+\affiliation{Lawrence Berkeley National Laboratory, Berkeley, CA}
+
 \author[0000-0001-6191-4285]{Kiran Eiden}
@@
              Stony Brook, NY 11794-3800}
-
-\author[0000-0001-8921-3624]{Nicole Ford}
-\affiliation{Lawrence Berkeley National Laboratory, Berkeley, CA}
 
@@
 \caption{\label{fig:time_series_1000} Time series of {\tt F1000} showing $\bar{A}$.}
+\end{figure}
+\MarginPar{update both the abar figures to only show 50 ms and 100 ms snapshots}
+
+\begin{figure}[t]
+\centering
+\plotone{time_series_enuc_500}
+\caption{\label{fig:time_series_enuc_500} Time series of {\tt F500} showing $\dot{e}_\mathrm{nuc}$.}
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\plotone{time_series_enuc_1000}
+\caption{\label{fig:time_series_enuc_1000} Time series of {\tt F1000} showing $\dot{e}_\mathrm{nuc}$.}
 \end{figure}
@@
 
-Next steps: 3-d, explore resolution more, H/He flames, MHD flames
+Next steps: tracer particles, 3-d, explore resolution more, H/He flames, MHD flames
 
@@
 Energy Computational Science Graduate Fellowship under Award Number DE-SC0021110.
+This work was supported in part by the U.S.
+Department of Energy, Office of Science, Office of Workforce Development for Teachers and Scientists (WDTS) under the Science Undergraduate Laboratory Internship (SULI) program.
 MZ acknowledges support from the Simons Foundation.

### 2020-12-23 xrb2/paper.tex
updated flame speeds table
Updated notes and density phase plot caption

@@
 \section{Numerical Approach}\label{Sec:numerics}
-\MarginPar{mz}
 
@@
 \end{figure}
-\MarginPar{update both the abar figures to only show 50 ms and 100 ms snapshots}
 
@@
 The fluid in this dense, high energy generation region undergoes vortical motion, shown in the Figure \ref{fig:uv} phase plots comparing $u$, $v$ and $\dot{e}_{\rm nuc}$.
-From the previous plots (see: streamlines?) \MarginPar{NF: which plots?}, this most likely corresponds to the leading edge of the flame where fresh fuel is being entrained.
+From the previous plots (see: streamlines?) \MarginPar{NF: need to add streamline plots?}, this most likely corresponds to the leading edge of the flame where fresh fuel is being entrained.
 This feature is not developed in the 250~Hz flame; it could potentially develop at later times, or the flame could just fizzle out.
@@
     \plotone{urho_plot_0_1}{}
-    \caption{\label{fig:urho}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and 1000~Hz runs.}
+    \caption{\label{fig:urho}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and 1000~Hz runs.
+The slightly lower $\dot{e}_{\rm nuc}$ values along the $u = 0$ axis are most likely a numerical artifact related to the density gradient setup and finite resolution of these simulations.}
 \end{figure}
@@
 
-MZ: lateral profiles of X vs distance behind the flame; 
-vertical profiles of the atmosphere behind the flame---how does it compare to isentropic?
-
-what is the composition on the photosphere across the flame
-
-shape of flame, amount of mixing (e.g. it looks like slower rotation rates have less mixed flames as the hurricane is less strong.)
-
-Don: look at ratio of enuc to e?
+\MarginPar{Mike wants to check vertical profiles of atmosphere behind flame}
 
@@
 	\startdata
-	{\tt F1000}     & $(3.41 \pm 0.01) \times 10^5$ & $(-1.03 \pm 0.13) \times 10^5$ \\
-	{\tt F500}      & $(3.08 \pm 0.01) \times 10^5$ & $(5.41 \pm 0.19) \times 10^5$ \\
-	{\tt F500\_2E8} & $(3.76 \pm 0.02) \times 10^5$ & $(1.74 \pm 0.17) \times 10^5$ \\
-	{\tt F500\_3E8} & $(5.29 \pm 0.10) \times 10^5$ & $(3.57 \pm 0.02) \times 10^7$ \\
+	{\tt F1000}     & $(3.38 \pm 0.11) \times 10^5$ & $(-0.43 \pm 1.57) \times 10^5$ \\
+	{\tt F500}      & $(2.66 \pm 0.27) \times 10^5$ & $(10.97 \pm 3.22) \times 10^5$ \\
+	{\tt F500\_2E8} & $(3.98 \pm 0.19) \times 10^5$ & $(-1.56 \pm 2.63) \times 10^5$ \\
+	{\tt F500\_3E8} & $(-4.56 \pm 2.92) \times 10^5$ & $(3.39 \pm 0.60) \times 10^7$ \\
 	\enddata

### 2020-12-23 xrb2/paper.tex
higher resolution phase plots

@@
 Comparing Figure \ref{fig:abar} to \citet{flame_wave1}, these tracks are much more neat and clearly defined.
-The ''messiness" of the tracks may be dependent on how mixed the flame interior is --- since these new simulations are un-boosted, they may be inherently less mixed than those in \citet{flame_wave1}.
+The ''messiness" of the tracks may be dependent on how mixed the flame interior is.
+Since these new simulations are un-boosted, they may be inherently less mixed than those in \citet{flame_wave1}.
 {\tt F1000} aligns with this interpretation: its $\bar{A}$ tracks are somewhat disrupted compared to the slower rotation runs, possibly due to the more vigorous mixing of the vortex at the flame front.
@@
     \plotone{urho_plot_0_1}{}
-    \caption{\label{fig:urho}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and 1000~Hz runs.
+    \caption{\label{fig:urho}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and 1000~Hz runs at $t = 100$~ms.
 The slightly lower $\dot{e}_{\rm nuc}$ values along the $u = 0$ axis are most likely a numerical artifact related to the density gradient setup and finite resolution of these simulations.}
@@
     \plotone{uv_plot_0_1}
-    \caption{\label{fig:uv}Phase plot of the horizontal $x$-velocity $u$, vertical $y$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and 1000~Hz runs.}
+    \caption{\label{fig:uv}Phase plot of the horizontal $x$-velocity $u$, vertical $y$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and 1000~Hz runs at $t = 100$~ms.}
 \end{figure}
@@
     \plotone{abar_plot_0_1}
-    \caption{\label{fig:abar}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and 1000~Hz runs.}
+    \caption{\label{fig:abar}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and 1000~Hz runs at $t = 100$~ms.}
 \end{figure}
@@
     \plotone{abar_plot_hotter}
-    \caption{\label{fig:abar_hot}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $2\times10^8$~K and $3\times10^8$~K.}
+    \caption{\label{fig:abar_hot}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $2\times10^8$~K and $3\times10^8$~K at $t = 40$~ms.}
 \end{figure}
@@
     \plotone{urho_plot_hotter}
-    \caption{\label{fig:urho_hot}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $2\times10^8$~K and $3\times10^8$~K.}
+    \caption{\label{fig:urho_hot}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $2\times10^8$~K and $3\times10^8$~K at $t = 40$~ms.}
 \end{figure}
@@
     \plotone{uv_plot_hotter}
-    \caption{\label{fig:uv_hot}Phase plot of the horizontal $x$-velocity $u$, vertical $y$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $2\times10^8$~K and $3\times10^8$~K.}
+    \caption{\label{fig:uv_hot}Phase plot of the horizontal $x$-velocity $u$, vertical $y$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $2\times10^8$~K and $3\times10^8$~K at $t = 40$~ms.}
 \end{figure}

### 2020-12-28 xrb2/paper.tex
Update affiliation
Incorporate some suggestions from last meeting

@@
 \author[0000-0001-6191-4285]{Kiran Eiden}
+\affiliation{Dept.\ of Astronomy, University of California, Berkeley,
+	CA 94720-3411}
 \affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
@@
 
-For the purpose of measuring the flame propagation speed and acceleration, we track the position of each of our flames
-as a function of time. We define the position in terms of a particular energy generation rate ($\enucdot$) value, as we did in \citet{flame_wave1}.
-To recapitulate: we first reduce the 2D $\enucdot$ data for each simulation run to a set of 1D radial profiles by averaging
-over the vertical coordinate. After averaging, we take our reference $\enucdot$ value to be some fraction of the global $\enucdot$
-maximum across all of these profiles. Since the flames in our simulations propagate rightward, we then search the region of
-each profile to the right of the local $\enucdot$ maximum for the point where the $\enucdot$ first drops below this reference
+For the purpose of measuring the flame propagation speed and acceleration, we track the position of 
+each of our flames as a function of time. We define the position in terms of a particular energy 
+generation rate ($\enucdot$) value, as we did in \citet{flame_wave1}. To recapitulate: we first 
+reduce the 2D $\enucdot$ data for each simulation run to a set of 1D radial profiles by averaging
+over the vertical coordinate. After averaging, we take our reference $\enucdot$ value to be some 
+fraction of the global $\enucdot$ maximum across all of these profiles. Since the flames in our 
+simulations propagate rightward, we then search the region of each profile to the right of the 
+local $\enucdot$ maximum for the point where the $\enucdot$ first drops below this reference
 value. This point gives us the location of our flame front.
 
-In \citet{flame_wave1}, we used $0.1\%$ of the global $\enucdot$ maximum for our reference value. For the high temperature
-unboosted flames, however, we found that the $\enucdot$ profiles failed to reach that small a value across the domain at most times,
-which prevented us from getting reliable position measurements. We therefore use $1\%$ of the global $\enucdot$ maximum in this paper
-rather than $0.1\%$. This is still sufficiently small that our measurements are not overly sensitive to turbulence and other local fluid
-motions (the issue with simply tracking the local maximum), but allows us to avoid the pitfall encountered by the $0.1\%$ metric.
+In \citet{flame_wave1}, we used $0.1\%$ of the global $\enucdot$ maximum for our reference value. 
+For the high temperature unboosted flames, however, we found that the $\enucdot$ profiles failed to 
+reach that small a value across the domain at most times, which prevented us from getting reliable 
+position measurements. We therefore use $1\%$ of the global $\enucdot$ maximum in this paper
+rather than $0.1\%$. This is still sufficiently small that our measurements are not overly 
+sensitive to turbulence and other local fluid motions (the issue with simply tracking the local 
+maximum), but allows us to avoid the pitfall encountered by the $0.1\%$ metric.
 
 \begin{deluxetable}{lrrr}
-	\tablecaption{\label{table:flame_speeds} Flame speed and acceleration values measured for each simulation.
-$s_0$ is the speed at $t = 0$.
-The initial flame velocities and accelerations are derived from a quadratic least-squares fit to each of the datasets for times $t \gtrsim 0.02~\mathrm{s}$.}
+	\tablecaption{\label{table:flame_speeds} Flame speed and acceleration values measured for each 
+	simulation. $s_0$ is the speed at $t = 0$. The initial flame velocities and accelerations are 
+	derived from a quadratic least-squares fit to each of the datasets for times $t \gtrsim 
+	20~\mathrm{ms}$.}
 	\tablehead{\colhead{run} & \colhead{$s_0$ (cm s$^{-1}$)} & \colhead{$a$ (cm s$^{-2}$)}} 
@@
 	\caption{\label{fig:flame_speeds_1} Flame front position vs.\ time for the standard ($10^8~\mathrm{K}$) 500 Hz and 1000 Hz runs. The dashed lines
-		show quadratic least-squares fits to the data for $t \gtrsim 0.02~\mathrm{s}$.}
+		show quadratic least-squares fits to the data for $t \gtrsim 20~\mathrm{ms}$.}
 \end{figure}
 
-Figure~\ref{fig:flame_speeds_1} gives position vs.\ time for the $\tt{F500}$ and $\tt{F1000}$ runs (blue and orange, respectively) to show the dependence on rotation rate.
-As some of the hotter 500~Hz runs exhibit significant accelerations, we fit a quadratic curve to each set of data for times $t \gtrsim 0.02~\mathrm{s}$ and calculate the acceleration and initial velocity from the fit parameters.
-These results are given in Table~\ref{table:flame_speeds}.
-\MarginPar{How is the acceleration calculated for this table?
-I.e.
-Specifically, at what time / over what time window is it calculated?} \MarginPar{NF: I will update the error estimates}.
+Figure~\ref{fig:flame_speeds_1} gives position vs.\ time for the $\tt{F500}$ and $\tt{F1000}$ runs 
+(blue and orange, respectively) to show the dependence on rotation rate. As some of the hotter 
+500~Hz runs exhibit significant accelerations, we fit a quadratic curve to each set of data for 
+times $t \gtrsim 20~\mathrm{ms}$, with the fit parameters giving the flame acceleration and 
+initial velocity. \MarginPar{KE: Can add an explanation of the averaging method once that's 
+decided.} These results are displayed in Table~\ref{table:flame_speeds}. \MarginPar{NF: I will 
+update the error estimates.}
 
@@
 \begin{equation}
-	\label{eqn:rossby}
-	s \appropto L_R (\epsdot)^{\half} \propto \frac{(\epsdot)^{\half}}{\Omega}
+	\label{eqn:flame_speed_scaling}
+	s \appropto L_R (\enucdot)^{\half} \propto \frac{(\enucdot)^{\half}}{\Omega}
 \end{equation}
-where $\epsdot$ is the specific energy generation rate and $L_R$ is the Rossby length.
-This was consistent with the results of \citet{cavecchi:2013}, who noted that the flame speeds in their simulations scaled with the Rossby length.
-As seen in Figure~\ref{fig:flame_speeds_1}, there is no obvious inverse scaling of the flame speed with rotation rate in this set of runs.
-We observed earlier, however, that nuclear reactions progress more quickly at the higher rotation rate.
-This results in a higher $\enucdot$ --- up to four to six times higher near the burning front after the flame ignites --- which may counteract the reduction in flame speed from the increased Coriolis confinement.
-The initial position of $\tt{F1000}$ does appear to be reduced compared to the lower rotation $\tt{F500}$ run.
-From Equation \ref{eqn:rossby}, increasing the rotation rate from 500 Hz to 1000 Hz should decrease $L_R$ by a factor of two, and the initial position of the flame should be reduced by a similar factor (again, due to greater confinement from the Coriolis force).
-$L_R$ is not, however, a perfect metric for the flame lengthscale.
-The scaling described in Equation \ref{eqn:rossby} seems to be more accurate at higher rotation rates (1000 Hz and up), based on \citet{flame_wave1}.
-We also observe that $\tt{F500}$ accelerates faster than $\tt{F1000}$, which appears to experience a small deceleration at early times.
-\MarginPar{This is speculation on my part -- does someone else have clear idea why?} This disparity may be a direct result of the difference in Coriolis force.
-It could also be due to the 500 Hz flame taking more time to develop fully and reach higher burning rates, resulting in an $\epsdot$ \MarginPar{Is this supposed to be enucdot???}increase over time.
+where $\enucdot$ is the specific nuclear energy generation rate and $L_R$ is the Rossby length. 
+This was consistent with the results of \citet{cavecchi:2013}, who noted that the flame speeds in 
+their simulations scaled with the Rossby length. As seen in Figure~\ref{fig:flame_speeds_1}, there 
+is no obvious inverse scaling of the flame speed with rotation rate in this set of runs. We 
+observed earlier, however, that nuclear reactions progress more quickly at the higher rotation 
+rate. This results in a higher $\enucdot$ --- up to four to six times higher near the burning front 
+after the flame ignites --- which may counteract the reduction in flame speed from the increased 
+Coriolis confinement. The initial position of $\tt{F1000}$ does appear to be reduced compared to 
+the lower rotation $\tt{F500}$ run. From Equation \ref{eqn:flame_speed_scaling}, increasing the 
+rotation rate from 500 Hz to 1000 Hz should decrease $L_R$ by a factor of two, and the initial 
+position of the flame should be reduced by a similar factor (again, due to greater confinement from 
+the Coriolis force). $L_R$ is not, however, a perfect metric for the flame lengthscale. The scaling 
+described in Equation \ref{eqn:flame_speed_scaling} seems to be more accurate at higher rotation 
+rates (1000 Hz and up), based on \citet{flame_wave1}. We also observe that $\tt{F500}$ accelerates 
+faster than $\tt{F1000}$, which appears to experience a small deceleration at early times. 
+\MarginPar{KE: This is speculation on my part -- does someone else have clear idea why?} This 
+disparity may be a direct result of the difference in Coriolis force. It could also be due to the 
+500 Hz flame taking more time to develop fully and reach higher burning rates, resulting in an 
+$\enucdot$ increase over time.
 
@@
 		$T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$. The dashed lines show quadratic least-squares fits to the data
-		for $t \gtrsim 0.02~\mathrm{s}$. Note that, due to its rapid acceleration, $\tt{F500\_3E8}$ (orange) is only run out to $\sim 0.07~\mathrm{s}$ to avoid surpassing the domain boundary.}
+		for $t \gtrsim 20~\mathrm{ms}$. Note that due to its rapid acceleration, $\tt{F500\_3E8}$ 
+		(orange) is only run out to $\sim 0.07~\mathrm{s}$ to avoid surpassing the domain boundary.}
 \end{figure}

### 2020-12-29 xrb2/paper.tex
spelling fixes

@@
 
-\author[0000-0002-1530-781X]{Alice Harpole}
+\author[0000-0002-1530-781X]{A.\ Harpole}
 \affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
@@
 
-\author[0000-0001-8921-3624]{Nicole Ford}
+\author[0000-0001-8921-3624]{N.\ Ford}
 \affiliation{Lawrence Berkeley National Laboratory, Berkeley, CA}
 
-\author[0000-0001-6191-4285]{Kiran Eiden}
+\author[0000-0001-6191-4285]{K.\ Eiden}
 \affiliation{Dept.\ of Astronomy, University of California, Berkeley,
@@
 
-\author[0000-0001-8401-030X]{Michael Zingale}
+\author[0000-0001-8401-030X]{M.\ Zingale}
 \affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
@@
 
-\author[0000-0003-2300-5165]{Donald Willcox}
+\author[0000-0003-2300-5165]{D.\ E.\ Willcox}
 \affiliation{Lawrence Berkeley National Laboratory, Berkeley, CA}
 
-\author[0000-0002-6447-3603]{Yuri Cavecchi}
+\author[0000-0002-6447-3603]{Y.\ Cavecchi}
 \affiliation{Mathematical Sciences and STAG Research Centre, University of Southampton, SO17 1BJ}
 
-\author[0000-0003-0439-4556]{Max P.\ Katz}
+\author[0000-0003-0439-4556]{M.\ P.\ Katz}
 \affiliation{NVIDIA Corp}
 
-\correspondingauthor{Alice Harpole}
+\correspondingauthor{A.\ Harpole}
 \email{alice.harpole@stonybrook.edu}
@@
 oscillations at lower rotation rates are smaller in amplitude and
-therefore more difficult to detect. It could be that it's not got
+therefore more difficult to detect. It could be that it does not have
 anything to do with the flame at all, but that neutron stars in
@@
 Previous studies have found that rotation can have a significant
-affect on the flame's propagation. As the rotation rate increases, the
+effect on the flame's propagation. As the rotation rate increases, the
 Coriolis force whips the spreading flame up into a hurricane-like
@@
 accretion rates and weaker at higher accretion rates, due to the
-effect of neutrino losses \citep{Cumming2006,johnston:2019}. In our
+effect of neutrino losses \citep{Cumming2006,johnston:2019}. 
+A shallow heating
+mechanism of as yet unknown origin has been found necessary to
+reproduce observed properties of XRBs in 1D simulations
+\citep{Deibel2015,Turlione2015,Keek2017}.
+In our
 models, we keep the crust at a constant temperature, so by increasing
-this temperature we can effectively increase the crustal heating and/or
-mimic the effects of higher accretion rates. A shallow heating
-mechanism of as yet unknown origin has been found necessary to
-reproduce observed properties of XRBs
-\citep{Deibel2015,Turlione2015,Keek2017}.
+this temperature we can effectively increase the crustal heating, shallow heating
+and/or
+mimic the effects of lower accretion rates.
 
-In the following sections, we conduct a series of simulations at various rotation rates and temperatures to investigate their effects on the flame.
-We find that at lower rotation rates, the flame itself becomes harder to ignite.
-At higher rotation rates, nuclear burning is enhanced and the flame propagates more quickly.
-At high temperatures, burning is greatly enhanced and the flame accelerates as it propagates.
-We discuss the implications that this may have for burst physics and observations.
+In the following sections, we conduct a series of simulations at various rotation 
+rates and temperatures to investigate their effects on the flame. We find that 
+at lower rotation rates, the flame itself becomes harder to ignite. At higher 
+rotation rates, nuclear burning is enhanced and the flame propagates more quickly. 
+At higher crustal temperatures, burning is greatly enhanced and the flame accelerates as 
+it propagates. We discuss the implications that this may have for burst physics 
+and observations.
 
@@
 current simulations are all performed in a two-dimensional
-axisymmetric geometry.  For these axisymmetric simulations, we add a
-missing geometric source terms from \citet{bernard-champmartin} that
+axisymmetric geometry.  For these axisymmetric simulations, we add an
+additional geometric source term from \citet{bernard-champmartin} that
 captures the effects of the divergence of the flux operating on the
 azimuthal unit vector.  This term is a small correction, but was
-missing from our earlier simulations.  The simulations framework
+missing from our previous simulations.  The simulation framework
 initializes a fuel layer in hydrostatic equilibrium, laterally
@@
 rotation rates to reduce the lateral lengthscale at which the Coriolis
-force balances the lateral flame spreading.  For the current
+force balances the lateral flame spreading in order to reduce the size 
+of the simulation domain.  For the current
 simulations, we no longer boost the flame speed---the true
@@
 used.  We continue to use a 13-isotope $\alpha$-chain to describe the
-helium burning.  The port of \castro\ to GPUs~\citep{sc20_gpu} enables
-us to run these simulations without the previous approximations, while
+helium burning.  The port of \castro\ to GPUs~\citep{sc20_gpu} 
+significantly improved its overall performance, enabling
+us to run these simulations without the previous approximations while
 continuing to resolve the burning front.
@@
 coordinates and a smooth transition to the cooler region at larger
-radii.  The cool initial characterized by 3 temperatures:
+radii.  The cool initial model is characterized by 3 temperatures:
 $T_\mathrm{star}$ is the isothermal temperature of the underlying
@@
 fuel layer, and $T_\mathrm{lo}$ is the minimum temperature of the
-atmosphere---the atmosphere structure is isentropic as it falls from
+atmosphere. The atmosphere structure is isentropic as it falls from
 $T_\mathrm{hi}$ down to $T_\mathrm{lo}$.  For the hot model, we
@@
 3.072\times 10^4~\mathrm{cm}$ with a coarse grid of $1152 \times 192$
-zones and two levels of refinment (the first a jump by a factor of four
+zones and two levels of refinement (the first a jump by a factor of four
 and the second a jump by a factor of two finer).  This gives a
@@
 In Figure~\ref{fig:4e8_stacked_enuc}, $\enucdot$ reaches values of $10^{18} - 10^{20}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ across the domain and at heights up to $\sim 0.5\times 10^4$ cm.
-There is still significant burning occuring even higher, with $\enucdot$ reaching $\sim 10^{15}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ at heights up to $\sim 0.8\times 10^4$ cm.
+There is still significant burning occurring even higher, with $\enucdot$ reaching $\sim 10^{15}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ at heights up to $\sim 0.8\times 10^4$ cm.
 For comparison, the next hottest run ($\tt{F500\_3E8}$) only reaches maximum $\enucdot$ values of $10^{18}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ and a maximum height of $\sim 0.5\times 10^4$ cm, even at the latest timesteps ($\sim 70~\mathrm{ms}$, when the flame is most developed).
@@
 Figure~\ref{fig:4e8_stacked_abar} shows $\bar{A}$ for $\tt{F500\_4E8}$.
-Again, burning extends across the domain and high into the atmosphere and fuel layer, lacking the charactersitic flame structure shown in $\bar{A}$ plots for the lower temperature $500$ Hz runs (see Figure~\ref{fig:compare_500Hz_abar}).
+Again, burning extends across the domain and high into the atmosphere and fuel layer, lacking the characteristic flame structure shown in $\bar{A}$ plots for the lower temperature $500$ Hz runs (see Figure~\ref{fig:compare_500Hz_abar}).
 A distinct mass of material - resembling a more dramatic version of that shown in Figure~\ref{fig:compare_500Hz_abar} - appears to have broken out of the atmosphere near the axis of symmetry.

### 2020-12-30 xrb2/paper.tex
more rewording

@@
 \begin{deluxetable}{lrrrrr}
-	\tablecaption{\label{table:runs} Rotation rate and temperature properties of the simulations.}
+	\tablecaption{\label{table:runs} Rotation rate and temperature properties of the simulations.
+In the left-hand column we list the names we shall use to refer to each simulation throughout this paper.}
 	\tablehead{\colhead{run} & \colhead{Rotation Rate (Hz)} & \colhead{$\delta T$ (K)} & \colhead{$T_\mathrm{hi}$ (K)} & \colhead{$T_\mathrm{star}$ (K)} & \colhead{$T_\mathrm{lo}$ (K)}} 
@@
 3.072\times 10^4~\mathrm{cm}$ with a coarse grid of $1152 \times 192$
-zones and two levels of refinement (the first a jump by a factor of four
-and the second a jump by a factor of two finer).  This gives a
-fine-grid resolution of $20~\mathrm{cm}$.  The entire atmosphere is
-refined in these simulations, with the refinement tagging on all zones
+zones and two levels of refinement (the first level refining the resolution 
+by factor of four, and the second by a factor of two again).  This gives a
+fine-grid resolution of $20~\mathrm{cm}$.  In these simulations, refinement 
+is carried out in all zones within the atmosphere
 with $\rho > 2.5\times 10^4~\gcc$.
@@
 
-We run three models ({\tt F250}, {\tt F500}, and {\tt F1000}) with the same initial model but differing rotation rates.
-
 \begin{figure}[t]
@@
 \plotone{time_series_500}
-\caption{\label{fig:time_series_500} Time series of {\tt F500} showing $\bar{A}$.}
+\caption{\label{fig:time_series_500} Time series of the 500~Hz run {\tt F500} showing the mean molecular weight, $\bar{A}$.}
 \end{figure}
@@
 \plotone{time_series_1000}
-\caption{\label{fig:time_series_1000} Time series of {\tt F1000} showing $\bar{A}$.}
+\caption{\label{fig:time_series_1000} Time series of the 1000~Hz run {\tt F1000} showing the mean molecular weight, $\bar{A}$.}
 \end{figure}
@@
 \plotone{time_series_enuc_500}
-\caption{\label{fig:time_series_enuc_500} Time series of {\tt F500} showing $\dot{e}_\mathrm{nuc}$.}
+\caption{\label{fig:time_series_enuc_500} Time series of the 500~Hz run {\tt F500} showing the nuclear energy generation rate, $\dot{e}_\mathrm{nuc}$.}
 \end{figure}
@@
 \plotone{time_series_enuc_1000}
-\caption{\label{fig:time_series_enuc_1000} Time series of {\tt F1000} showing $\dot{e}_\mathrm{nuc}$.}
+\caption{\label{fig:time_series_enuc_1000} Time series of the 1000~Hz run {\tt F1000} showing the nuclear energy generation rate, $\dot{e}_\mathrm{nuc}$.}
 \end{figure}
 
+We run three models ({\tt F250}, {\tt F500}, and {\tt F1000}) with the same initial model in terms of temperature but differing rotation rates.
 The {\tt F500} and {\tt F1000} runs both qualitatively resemble the flame structure in \citet{flame_wave1} --- a laterally propagating flame that is lifted off of the bottom of the fuel layer --- but they differ in their burning structures.
-Figures~\ref{fig:time_series_500} and \ref{fig:time_series_1000} show time series of the mean molecular weight ($\bar{A}$) for the {\tt F500} and {\tt F1000} runs.
+Figures~\ref{fig:time_series_500} and \ref{fig:time_series_1000} show time series of the mean molecular weight, $\bar{A}$, for the {\tt F500} and {\tt F1000} runs.
 Compared to those in \cite{flame_wave1}, ashes behind the flame do not reach as high atomic weights.
-This is not surprising, since those early runs boosted the reaction rates.
+This is not surprising, since those previous runs artificially boosted the reaction rates.
 Comparing these two new runs, the burning is much more evolved for the higher rotation rate, and the ash is actually able to move ahead of the flame front (visible in the Figure  \ref{fig:time_series_1000} 100 ms snapshot).
-We believe that this is because the increased rotation confines the initial perturbation and subsequent expansion from the burning better, allowing the reactions to progress further.
+We believe that this is because the increased rotation better confines the initial perturbation and subsequent expansion from the burning, increasing the temperature and density in the flame front such that the reaction rate increases, which allows the reactions to progress further.
 In contrast to {\tt F500} and {\tt F1000}, the lowest rotation run --- {\tt F250} --- failed to ignite.
-The lack of ignition for {\tt F250} also aligns with the reasoning given above, with the lower rotation in this case potentially leading to insufficient confinement of the burning.
-In this scenario, another source of confinement (e.g.\ magnetic fields) would need to take over at lower rotation rates to allow a burst to occur, at least for our current initial model.
+The lack of ignition for {\tt F250} also aligns with the reasoning given above, with the lower rotation in this case potentially leading to insufficient confinement such that the temperature and density required for ignition is not achieved.
+In this scenario, another source of confinement (e.g.\ magnetic fields) would need to take over at lower rotation rates to allow a burst to occur, at least for the initial model used here.
 
 Burning in the {\tt F500} and {\tt F1000} runs is concentrated in a dense region with circular motion.
-Figure \ref{fig:urho} compares the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for {\tt F250}, {\tt F500}, and {\tt F1000}.
-Most of the burning for each of the simulations occurs in a high density region $\rho > 3 \times 10^5 {\rm g/cm^3}$.
+Figure \ref{fig:urho} compares the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the {\tt F250}, {\tt F500}, and {\tt F1000} runs.
+Most of the burning for each of the simulations occurs in a high density region $\rho > 3 \times 10^5 {\rm g\,cm^{-3}}$.
 The fluid in this dense, high energy generation region undergoes vortical motion, shown in the Figure \ref{fig:uv} phase plots comparing $u$, $v$ and $\dot{e}_{\rm nuc}$.
 From the previous plots (see: streamlines?) \MarginPar{NF: need to add streamline plots?}, this most likely corresponds to the leading edge of the flame where fresh fuel is being entrained.
-This feature is not developed in the 250~Hz flame; it could potentially develop at later times, or the flame could just fizzle out.
+This feature is not developed in the 250~Hz flame; it could potentially develop at later times (past the point at which we terminated our simulation), or the flame could just fizzle out and fail to ignite entirely.
 
@@
 Comparing Figure \ref{fig:abar} to \citet{flame_wave1}, these tracks are much more neat and clearly defined.
-The ''messiness" of the tracks may be dependent on how mixed the flame interior is.
+The ``messiness" of the tracks may be dependent on how mixed the flame interior is.
 Since these new simulations are un-boosted, they may be inherently less mixed than those in \citet{flame_wave1}.
@@
 It also agrees with our earlier interpretation of the enhanced burning seen in Figure \ref{fig:time_series_1000} for {\tt F1000}.
-The disruption of reaction pathways with higher rotation rates and mixed flames may have interesting consequences; if the reactions are disrupted enough, it is possible that the burning could be snuffed out.
-This can occur in combustion scenarios when the turbulent vortices become so small they penetrate the reaction zone and completely disrupt the burning.
 
@@
     \plotone{urho_plot_0_1}{}
-    \caption{\label{fig:urho}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and 1000~Hz runs at $t = 100$~ms.
+    \caption{\label{fig:urho}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and 1000~Hz runs ({\tt F250}, {\tt F500} and {\tt F1000}) at $t = 100$~ms.
 The slightly lower $\dot{e}_{\rm nuc}$ values along the $u = 0$ axis are most likely a numerical artifact related to the density gradient setup and finite resolution of these simulations.}
@@
     \plotone{uv_plot_0_1}
-    \caption{\label{fig:uv}Phase plot of the horizontal $x$-velocity $u$, vertical $y$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and 1000~Hz runs at $t = 100$~ms.}
+    \caption{\label{fig:uv}Phase plot of the horizontal $x$-velocity $u$, vertical $y$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and 1000~Hz runs ({\tt F250}, {\tt F500} and {\tt F1000}) at $t = 100$~ms.}
 \end{figure}
@@
     \plotone{abar_plot_0_1}
-    \caption{\label{fig:abar}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and 1000~Hz runs at $t = 100$~ms.}
+    \caption{\label{fig:abar}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and 1000~Hz runs ({\tt F250}, {\tt F500} and {\tt F1000}) at $t = 100$~ms.}
 \end{figure}
@@
 For the purpose of measuring the flame propagation speed and acceleration, we track the position of 
-each of our flames as a function of time. We define the position in terms of a particular energy 
-generation rate ($\enucdot$) value, as we did in \citet{flame_wave1}. To recapitulate: we first 
+each of our flames as a function of time. We define the position in terms of a specific value of the energy 
+generation rate, $\enucdot$, as we did in \citet{flame_wave1}. To recapitulate: we first 
 reduce the 2D $\enucdot$ data for each simulation run to a set of 1D radial profiles by averaging
@@
 fraction of the global $\enucdot$ maximum across all of these profiles. Since the flames in our 
-simulations propagate rightward, we then search the region of each profile to the right of the 
+simulations propagate in the positive radial direction, we then search the region of each profile at greater radius than the
 local $\enucdot$ maximum for the point where the $\enucdot$ first drops below this reference
@@
 	\plotone{flame_speeds_rotation.pdf}
-	\caption{\label{fig:flame_speeds_1} Flame front position vs.\ time for the standard ($10^8~\mathrm{K}$) 500 Hz and 1000 Hz runs. The dashed lines
+	\caption{\label{fig:flame_speeds_1} Flame front position vs.\ time for the standard ($10^8~\mathrm{K}$) 500 Hz and 1000 Hz runs ({\tt F500} and {\tt F1000}). The dashed lines
 		show quadratic least-squares fits to the data for $t \gtrsim 20~\mathrm{ms}$.}
@@
 (blue and orange, respectively) to show the dependence on rotation rate. As some of the hotter 
-500~Hz runs exhibit significant accelerations, we fit a quadratic curve to each set of data for 
+500~Hz runs exhibit significant acceleration, we fit a quadratic curve to each set of data for 
 times $t \gtrsim 20~\mathrm{ms}$, with the fit parameters giving the flame acceleration and 
@@
 
-We found in \citet{flame_wave1} that the flame speed $s$ obeys:
+We found in \citet{flame_wave1} that the flame speed, $s$, obeys the following relation:
 \begin{equation}
 	\label{eqn:flame_speed_scaling}
-	s \appropto L_R (\enucdot)^{\half} \propto \frac{(\enucdot)^{\half}}{\Omega}
+	s \appropto L_R (\enucdot)^{\half} \propto \frac{(\enucdot)^{\half}}{\Omega},
 \end{equation}
@@
 is no obvious inverse scaling of the flame speed with rotation rate in this set of runs. We 
-observed earlier, however, that nuclear reactions progress more quickly at the higher rotation 
+observed earlier, however, that nuclear reactions progress more quickly at higher rotation 
 rate. This results in a higher $\enucdot$ --- up to four to six times higher near the burning front 
@@
 By choosing $T_\mathrm{star}$ = $T_\mathrm{hi}$, we can effectively explore simulations with greater surface heating.
-There are two physically distinct mechanisms that could each produce an increased temperature at the crust: crustal heating or a higher accretion rate.
-In these simulations, we cannot distinguish between whether we are generating heat due to a higher accretion rate or due to a shallow crustal heating mechanism, because we are not resolving the crust.
-Instead, we are modeling the resulting heating effect.
+There are several physically distinct mechanisms that could produce an increased temperature at the crust: crustal heating, some other form of shallow heating or a higher accretion rate.
+In these simulations, we do not model the mechanism producing the heating effect, just the effect itself, so we do not distinguish between which of these mechanisms cause the heating.
 
@@
 \plotone{compare_500Hz_abar}
-\caption{\label{fig:compare_500Hz_abar} Comparison of $\bar{A}$ for 3 different 500~Hz models with different neutron star temperatures, $T_\star$, and resulting envelope structures.
+\caption{\label{fig:compare_500Hz_abar} Comparison of $\bar{A}$ for 3 different 500~Hz models with different neutron star temperatures, $T_\mathrm{star}$, and resulting envelope structures.
 Each flame is shown at 70~ms.} %This will be updated to show at 70ms
@@
 Figure~\ref{fig:compare_500Hz_abar} shows $\bar{A}$ for three 500 Hz simulations with different initial temperature structures at $t = 70~\mathrm{ms}$.
-The hottest of the three models, $\tt{F500\_3E8}$, has a faster propagating flame (discussed further in Section \ref{ssec:temp_prop}).
+The hottest of the three models, $\tt{F500\_3E8}$, has a faster propagating flame (this will be discussed further in Section \ref{ssec:temp_prop}).
 It also appears to have a secondary burning region (colored light orange) forming on the far left above the flame edge.
@@
 This effect, however, may be numerical rather than physical, resulting from the boundary conditions imposed at the axis of symmetry for these simulations.
-
-\begin{figure}[t]
-    \centering
-    \plotone{abar_plot_hotter}
-    \caption{\label{fig:abar_hot}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $2\times10^8$~K and $3\times10^8$~K at $t = 40$~ms.}
-\end{figure}
 
@@
 Compared to the highest rotation rate run, $\tt{F1000}$, $\tt{F500\_3E8}$ reaches a peak temperature roughly $1 \times 10^9~\mathrm{K}$ higher ($\sim 2.7 \times 10^9~\mathrm{K}$ versus $\sim 1.7 \times 10^9~\mathrm{K}$).
-Though increasing the rotation rate does seem to concentrate the burning and increase the peak temperature, it does not reach the same level as directly increasing the temperature.
+Though increasing the rotation rate does seem to concentrate the burning and increase the peak temperature, it does not reach the same level as directly increasing the temperature at the crust.
 
 In contrast to the models with $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$, $\tt{F500\_4E8}$ is so hot that the organized flame structure is lost.
-The fuel layer undergoes stable burning and essentially simmers in place, rather than forming a flame; this effect can occur in nature, for neutron stars with accretion that balances the Eddington luminosity \citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}.
+The fuel layer undergoes stable burning and essentially simmers in place, rather than forming a flame.
+This process has been observed neutron stars with accretion rates that exceed the Eddington limit \citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}; in this model with have essentially recreated the conditions found in such stars with high accretion rates.
 This model burns so strongly that it is only run out to $40$ ms.
@@
 Significant $\enucdot$ values for $\tt{F500\_3E8}$ are confined to the flame front, rather than spanning the entire domain.
-\MarginPar{NF: should I include a plot for this or is it ok to just say it?}. 
+
+\MarginPar{NF: should I include a plot for this or is it ok to just say it?}
 Figure~\ref{fig:4e8_stacked_abar} shows $\bar{A}$ for $\tt{F500\_4E8}$.
 Again, burning extends across the domain and high into the atmosphere and fuel layer, lacking the characteristic flame structure shown in $\bar{A}$ plots for the lower temperature $500$ Hz runs (see Figure~\ref{fig:compare_500Hz_abar}).
-A distinct mass of material - resembling a more dramatic version of that shown in Figure~\ref{fig:compare_500Hz_abar} - appears to have broken out of the atmosphere near the axis of symmetry.
+A distinct mass of material -- resembling a more dramatic version of that shown in Figure~\ref{fig:compare_500Hz_abar} -- appears to have broken out of the atmosphere near the axis of symmetry.
 The atmosphere edge is located at $\sim 1.2\times 10^4~\mathrm{cm}$.
-To reiterate, this may just be a result of the imposed boundary conditions rather than a physical effect.
+As mentioned before, this may just be a result of the imposed boundary conditions rather than a physical effect.
 $\tt{F500\_4E8}$ clearly reaches much higher $\bar{A}$ values across the domain (especially at the latest snapshot, $t = 40~\mathrm{ms}$) compared to all the other runs described in this paper (see Figures \ref{fig:compare_500Hz_abar}, \ref{fig:time_series_500}, and \ref{fig:time_series_1000}).
-This is not surprising, since as we have seen through our varied temperature simulations, the burning is very temperature-sensitive.
+This is not surprising, since as we have shown in our simulations with differing initial temperature structures, the burning is very temperature-sensitive.
 
@@
     \caption{\label{fig:uv_hot}Phase plot of the horizontal $x$-velocity $u$, vertical $y$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $2\times10^8$~K and $3\times10^8$~K at $t = 40$~ms.}
+\end{figure}
+
+\begin{figure}[t]
+    \centering
+    \plotone{abar_plot_hotter}
+    \caption{\label{fig:abar_hot}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $2\times10^8$~K and $3\times10^8$~K at $t = 40$~ms.}
 \end{figure}

### 2020-12-30 xrb2/paper.tex
conclusion

@@
 rate and temperature conditions for the flame. The main parameters
-describing the models are provided in Table
-\ref{table:runs}. Figure~\ref{fig:hot_profiles} shows the
+describing the models are provided in Table~\ref{table:runs}. Figure~\ref{fig:hot_profiles} shows the
 temperature and density structure for our hot models and
@@
 
-To explore the effect of different initial temperature configurations, we run four simulations fixed at a rotation rate of 500 Hz with temperatures as shown in Table \ref{table:runs}.
+To explore the effect of different initial temperature configurations, we run four simulations fixed at a rotation rate of 500 Hz with temperatures as shown in Table~\ref{table:runs}.
 For all the 500 Hz simulations (with the exception of the coolest run, $\tt{F500}$), we set $T_\mathrm{star}$ = $T_\mathrm{hi}$, scaling $\delta T$ accordingly to maintain a consistent value of $T_\mathrm{hi} + \delta T$.
@@
 The method for measuring flame propagation described in Section \ref{ssec:rot_propagation} is applied in Figure \ref{fig:flame_speeds_2} to the three 500 Hz runs with $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$.
-Due to the lack of a clear burning front in $\tt{F500\_4E8}$, we neglect to analyze its propagation speed and acceleration.
+Due to the lack of a clear burning front in $\tt{F500\_4E8}$, we do not analyze its propagation speed and acceleration.
 As the initial $T_{\mathrm{star}}$ is increased, the flame becomes greatly accelerated.
-The initial flame velocities and accelerations derived from a quadratic least-squares fit to each of the datasets are provided in Table \ref{table:flame_speeds}.
+The initial flame velocities and accelerations derived from a quadratic least-squares fit to each of the datasets are provided in Table~\ref{table:flame_speeds}.
 The reason for the acceleration of the flame over time is not entirely clear.
@@
 
-Looking at the phase plots in Figures~\ref{fig:urho_hot}-\ref{fig:abar_hot}, we see:
-\begin{itemize}
-    \item The $u$-$\rho$ plot in Figure~\ref{fig:urho_hot} shows that the horizontal velocity when $T_{\rm star} = 3\times10^8$~K is sufficiently greater than for the two cooler runs.
-This is consistent with the plot of the flame front position, where we saw that the hottest flame was significantly faster at later times than the cooler flames.
-    \item The $u$-$v$ plot in Figure~\ref{fig:uv_hot} further demonstrates this, as again we see that the magnitude of the velocity of the flame for the hottest run is significantly larger.
+The $u$-$\rho$ phase plot in Figure~\ref{fig:urho_hot} shows that the horizontal velocity for the $T_{\rm star} = 3\times10^8$~K is sufficiently greater in magnitude than for the two cooler runs.
+The $u$-$v$ plot in Figure~\ref{fig:uv_hot} further demonstrates this, as again we see that the magnitude of the velocity in the vertical direction of the flame for the hottest run is also significantly larger.
+This motion therefore most likely corresponds to the vortical motion rather than to the propagation of the flame itself, something which is further supported by the fact that the velocity magnitude is 2-3 orders of magnitude greater than the flame speeds listed in Table~\ref{table:flame_speeds}.
 Also of interest in these plots is that the coolest run does not seem to have developed the characteristic vortex structure at the burning front (i.e.
-where $\dot{e}_{\rm nuc}$ is greatest) that can be clearly seen for the hotter runs.
-\end{itemize}
+where $\dot{e}_{\rm nuc}$ is greatest) that can be clearly seen for the hotter runs at this time (40~ms).
+As can be seen in Figure~\ref{fig:uv} (which was plotted at 100~ms), this does develop more at later times.
+Similar to what saw when comparing the runs with different rotation rates, it would therefore appear that it takes longer for the flame to develop when $T_{\rm star}$ is cooler.
 
@@
 
-tl;dr: faster rotation and/or hotter surface temperature = more burning.
+We ran a number of simulations of laterally propagating flames in XRBs in order to explore the effects and rotation and thermal structure.
+We found that increasing the rotation rate increased the energy generation rate within the flame, enhancing its propagating.
+This is likely due to greater confinement of the flame from the increased Coriolis force.
+We explored several models with different crustal temperatures, to determine what effects mechanisms such as higher accretion rates, crustal heating and shallow heating may have on flame propagation.
+We found that increasing the temperature of the crust significantly enhanced the flame propagation.
+This we believe to be because a cooler crust allows heat to more efficiently be transferred from the flame itself, therefore reducing the flame's temperature, slowing burning and consequently the reducing its propagation speed.
+We saw that if the crust temperature was too high, then instead of a flame, the atmosphere would burn stably.
+This is reminiscent of what is seen for neutron stars with accretion rates exceeding the Eddington luminosity.
 
-Next steps: tracer particles, 3-d, explore resolution more, H/He flames, MHD flames
+In future simulations, we would like to improve and expand our simulations in order to better understand the processes at play and to include more physics.
+This includes adding tracer particles to the simulations so we can monitor the fluid motion; extending our simulations to 3D, which would hopefully alleviate some of the boundary effects we have observed in these simulations but will require significant computational resource; and exploring the resolution of our simulations more so that we can ensure that we have resolved all the necessary physical processes.
+We would also like to model H/He flames, \MarginPar{someone please describe here why these would be interesting as I can't remember what makes them different}.
+We recently added an MHD solver to \castro; this will allow us to explore the effects of magnetic fields on flame propagation in XRBs.
+
+% tl;dr: faster rotation and/or hotter surface temperature = more burning.
+
+% Next steps: tracer particles, 3-d, explore resolution more, H/He flames, MHD flames
 

### 2020-12-30 xrb2/paper.tex
added citation to Lupe's thesis
Migrate discussion of Rossby length to Section 3.1
The new additions to Section 3.1 are in blue.

@@
 
-We run three models ({\tt F250}, {\tt F500}, and {\tt F1000}) with the same initial model in terms of temperature but differing rotation rates.
-The {\tt F500} and {\tt F1000} runs both qualitatively resemble the flame structure in \citet{flame_wave1} --- a laterally propagating flame that is lifted off of the bottom of the fuel layer --- but they differ in their burning structures.
-Figures~\ref{fig:time_series_500} and \ref{fig:time_series_1000} show time series of the mean molecular weight, $\bar{A}$, for the {\tt F500} and {\tt F1000} runs.
-Compared to those in \cite{flame_wave1}, ashes behind the flame do not reach as high atomic weights.
-This is not surprising, since those previous runs artificially boosted the reaction rates.
-Comparing these two new runs, the burning is much more evolved for the higher rotation rate, and the ash is actually able to move ahead of the flame front (visible in the Figure  \ref{fig:time_series_1000} 100 ms snapshot).
-We believe that this is because the increased rotation better confines the initial perturbation and subsequent expansion from the burning, increasing the temperature and density in the flame front such that the reaction rate increases, which allows the reactions to progress further.
-In contrast to {\tt F500} and {\tt F1000}, the lowest rotation run --- {\tt F250} --- failed to ignite.
-The lack of ignition for {\tt F250} also aligns with the reasoning given above, with the lower rotation in this case potentially leading to insufficient confinement such that the temperature and density required for ignition is not achieved.
-In this scenario, another source of confinement (e.g.\ magnetic fields) would need to take over at lower rotation rates to allow a burst to occur, at least for the initial model used here.
+We run three models ({\tt F250}, {\tt F500}, and {\tt F1000}) with the same initial model in terms 
+of temperature but differing rotation rates. {\color{blue}We saw in \citet{flame_wave1} that 
+increasing the 
+neutron star rotation rate reduces the horizontal lengthscale of the flame. An estimate of 
+this lengthscale is given by the Rossby radius of deformation, $L_R$. The Rossby radius may be 
+thought of as the scale over which the balance between the Coriolis force and horizontal pressure 
+gradient becomes important. It is approximately:
+\begin{equation}
+	\label{eqn:rossby}
+	L_R \approx \frac{\sqrt{g H_0}}{\Omega},
+\end{equation}
+where $g$ is gravitational acceleration, $H_0$ is the atmospheric scale height, and $\Omega$ is the 
+neutron star rotation rate. Comparing Figure \ref{fig:time_series_enuc_500} and Figure 
+\ref{fig:time_series_enuc_1000}, the horizontal extent of the $\tt{F1000}$ flame does appear to be 
+reduced compared to the lower rotation $\tt{F500}$ run. From Equation \ref{eqn:rossby}, increasing 
+the rotation rate from 500 Hz to 1000 Hz should decrease $L_R$ by a factor of two, and the greater 
+confinement from the Coriolis force should reduce the horizontal extent of the flame by a similar 
+factor. The Rossby radius is not a perfect measure of this horizontal lengthscale, however, and 
+here the difference is not a factor of two. The simulations seem to follow the theoretical
+scaling described in Equation \ref{eqn:rossby} more closely at higher rotation rates (1000 Hz and 
+up), based on the results of \citet{flame_wave1}.}
+
+The {\tt F500} and {\tt F1000} runs both qualitatively resemble the flame structure in 
+\citet{flame_wave1} --- a laterally propagating flame that is lifted off of the bottom of the fuel 
+layer --- but they differ in their burning structures. Figures~\ref{fig:time_series_500} and 
+\ref{fig:time_series_1000} show time series of the mean molecular weight, $\bar{A}$, for the {\tt 
+F500} and {\tt F1000} runs.  Compared to those in \cite{flame_wave1}, ashes behind the flame do not 
+reach as high atomic weights. This is not surprising, since those previous runs artificially 
+boosted the reaction rates.  Comparing these two new runs, the burning is much more evolved for the 
+higher rotation rate, and the ash is actually able to move ahead of the flame front (visible in the 
+Figure  \ref{fig:time_series_1000} 100 ms snapshot). We believe that this is because the increased 
+rotation better confines the initial perturbation and subsequent expansion from the burning, 
+increasing the temperature and density in the flame front such that the reaction rate increases, 
+which allows the reactions to progress further. In contrast to {\tt F500} and {\tt F1000}, the 
+lowest rotation run --- {\tt F250} --- failed to ignite. The lack of ignition for {\tt F250} also 
+aligns with the reasoning given above, with the lower rotation in this case potentially leading to 
+insufficient confinement such that the temperature and density required for ignition is not 
+achieved. In this scenario, another source of confinement (e.g.\ magnetic fields) would need to 
+take over at lower rotation rates to allow a burst to occur, at least for the initial model used 
+here.
 
@@
 For the purpose of measuring the flame propagation speed and acceleration, we track the position of 
-each of our flames as a function of time. We define the position in terms of a specific value of the energy 
-generation rate, $\enucdot$, as we did in \citet{flame_wave1}. To recapitulate: we first 
+each of our flames as a function of time. We define the position in terms of a specific value of 
+the energy generation rate, $\enucdot$, as we did in \citet{flame_wave1}. To recapitulate: we first 
 reduce the 2D $\enucdot$ data for each simulation run to a set of 1D radial profiles by averaging
@@
 fraction of the global $\enucdot$ maximum across all of these profiles. Since the flames in our 
-simulations propagate in the positive radial direction, we then search the region of each profile at greater radius than the
-local $\enucdot$ maximum for the point where the $\enucdot$ first drops below this reference
-value. This point gives us the location of our flame front.
+simulations propagate in the positive radial direction, we then search the region of each profile 
+at greater radius than the local $\enucdot$ maximum for the point where the $\enucdot$ first drops 
+below this reference value. This point gives us the location of our flame front.
 
@@
 \end{equation}
-where $\enucdot$ is the specific nuclear energy generation rate and $L_R$ is the Rossby length. 
+where $\enucdot$ is the specific nuclear energy generation rate and $L_R$ is the Rossby length 
+described by Equation \ref{eqn:rossby}. 
 This was consistent with the results of \citet{cavecchi:2013}, who noted that the flame speeds in 
@@
 after the flame ignites --- which may counteract the reduction in flame speed from the increased 
-Coriolis confinement. The initial position of $\tt{F1000}$ does appear to be reduced compared to 
-the lower rotation $\tt{F500}$ run. From Equation \ref{eqn:flame_speed_scaling}, increasing the 
-rotation rate from 500 Hz to 1000 Hz should decrease $L_R$ by a factor of two, and the initial 
-position of the flame should be reduced by a similar factor (again, due to greater confinement from 
-the Coriolis force). $L_R$ is not, however, a perfect metric for the flame lengthscale. The scaling 
-described in Equation \ref{eqn:flame_speed_scaling} seems to be more accurate at higher rotation 
-rates (1000 Hz and up), based on \citet{flame_wave1}. We also observe that $\tt{F500}$ accelerates 
+Coriolis confinement. We also observe that $\tt{F500}$ accelerates 
 faster than $\tt{F1000}$, which appears to experience a small deceleration at early times. 
@@
 We would also like to model H/He flames, \MarginPar{someone please describe here why these would be interesting as I can't remember what makes them different}.
-We recently added an MHD solver to \castro; this will allow us to explore the effects of magnetic fields on flame propagation in XRBs.
+We recently added an MHD solver to \castro~\citep{sazo2020thesis}; this will allow us to explore the effects of magnetic fields on flame propagation in XRBs.
 

### 2020-12-30 xrb2/paper.tex
Edited flame scaling paragraph in Section 3.2

@@
 thought of as the scale over which the balance between the Coriolis force and horizontal pressure 
-gradient becomes important. It is approximately:
+gradient becomes important, and is approximately:
 \begin{equation}
@@
 where $\enucdot$ is the specific nuclear energy generation rate and $L_R$ is the Rossby length 
-described by Equation \ref{eqn:rossby}. 
-This was consistent with the results of \citet{cavecchi:2013}, who noted that the flame speeds in 
-their simulations scaled with the Rossby length. As seen in Figure~\ref{fig:flame_speeds_1}, there 
-is no obvious inverse scaling of the flame speed with rotation rate in this set of runs. We 
-observed earlier, however, that nuclear reactions progress more quickly at higher rotation 
-rate. This results in a higher $\enucdot$ --- up to four to six times higher near the burning front 
-after the flame ignites --- which may counteract the reduction in flame speed from the increased 
-Coriolis confinement. We also observe that $\tt{F500}$ accelerates 
-faster than $\tt{F1000}$, which appears to experience a small deceleration at early times. 
-\MarginPar{KE: This is speculation on my part -- does someone else have clear idea why?} This 
-disparity may be a direct result of the difference in Coriolis force. It could also be due to the 
-500 Hz flame taking more time to develop fully and reach higher burning rates, resulting in an 
-$\enucdot$ increase over time.
+described by Equation \ref{eqn:rossby}. This finding was consistent with the results of 
+\citet{cavecchi:2013}, who noted that the flame speeds in their simulations scaled with the Rossby 
+length. As seen in Figure~\ref{fig:flame_speeds_1}, there is no obvious inverse scaling of the 
+flame speed with rotation rate in this set of runs. We observed earlier, however, that nuclear 
+reactions progress more quickly at higher rotation rate. This results in a higher $\enucdot$ --- up 
+to four to six times higher near the burning front after the flame ignites --- which may counteract 
+the reduction in flame speed from the increased Coriolis confinement. We also observe that 
+$\tt{F500}$ accelerates faster than $\tt{F1000}$, which appears to experience a small deceleration 
+at early times. This disparity may be a direct result of the difference in Coriolis force.
 

### 2020-12-30 xrb2/paper.tex
some discussion

@@
 
-In future simulations, we would like to improve and expand our simulations in order to better understand the processes at play and to include more physics.
-This includes adding tracer particles to the simulations so we can monitor the fluid motion; extending our simulations to 3D, which would hopefully alleviate some of the boundary effects we have observed in these simulations but will require significant computational resource; and exploring the resolution of our simulations more so that we can ensure that we have resolved all the necessary physical processes.
-We would also like to model H/He flames, \MarginPar{someone please describe here why these would be interesting as I can't remember what makes them different}.
-We recently added an MHD solver to \castro~\citep{sazo2020thesis}; this will allow us to explore the effects of magnetic fields on flame propagation in XRBs.
+In future simulations, we would like to improve and expand our
+simulations in order to better understand the processes at play and to
+include more physics. This includes adding tracer particles to the
+simulations so we can monitor the fluid motion and perform more
+detailed nucleosynthesis; extending our simulations to 3D, which would
+hopefully alleviate some of the boundary effects we have observed in
+these simulations but will require significant computational resource;
+and exploring the resolution of our simulations more so that we can
+ensure that we have resolved all the necessary physical processes. We
+would also like to model H/He flames, as these are the sites of
+rp-process nucleosynthesis~\citep{rpprocess}.  Initially we will use the same reaction
+sequence explored in our previous convection
+calculations~\citep{xrb2}.   Finally, we
+recently added an MHD solver to \castro~\citep{sazo2020thesis}; this
+will allow us to explore the effects of magnetic fields on flame
+propagation in XRBs.
 
@@
 \software{AMReX \citep{amrex_joss},
-          Castro \citep{castro},
+          Castro \citep{castro,castro_joss},
           GCC (\url{https://gcc.gnu.org/}),

### 2020-12-31 simplified_sdc_intro/paper.tex
fix sign issue
some updates

@@
 timestep limiters based on the energy generation or abundance changes
-over a timestep:
+over a timestep: \MarginPar{do we enforce this in the NSE region?}
 \begin{align}
@@
 \end{equation}
-where $\Advs{\Uc} =  \frac{\partial \Fb^\xv (\Uc)}{\partial x} + \Hb(\Uc)$.
+where $\Advs{\Uc} =  -{\partial \Fb^\xv (\Uc)}/{\partial x} + \Hb(\Uc)$.
 For the simplified SDC method we explore here, we will make the
@@
 ``aprox19'' network from \cite{aprox19} and originally described in
-\cite{Kepler}.  We combine this with a table that gives the nuclear
+\cite{Kepler}.
+We use a modified version of the VODE~\citep{vode} integrator---ported to C++ with checks added to the timestep rejection logic that ensure that the species mass fractions stay between 0 and 1\footnote{This modified version of VODE is available in our Microphysics repo: \url{https://github.com/starkiller-astro/Microphysics}.}.\MarginPar{discuss tolerances}  We combine this with a table that gives the nuclear
 statistical equilibrium (NSE) abundances in regions where the system
@@
       The goal is to update the composition and thermodynamics due to the 
-      change in the nuclei abundances over the (half-)timestep.  There are
-      a few approaches we can do here.
-
-      \begin{itemize}
-      \item Method I: {\em Use the time-level $n$ information for the update}.
-
-        This is the approach used in \citet{ma:2013}.  We call the table with the 
-        input state:
-        \begin{equation}
-          \nse{\rho_{\rm in}, T_{\rm in}, (Y_e)_{\rm in}} \rightarrow \bar{A}_\out, (X_k)_\out, (B/A)_\out, dY_e/dt
-        \end{equation}
-
-        We can now update $Y_e$:
-        \begin{equation}
-          (Y_e)_{\rm out} = (Y_e)_{\rm in} + \Delta t \frac{dY_e}{dt}
-        \end{equation}
-        and take the auxiliary state values of $\bar{A}$ and $(B/A)$ to be simply
-        the values from the table, $\bar{A}_\out$ and $(B/A)_\out$, and the mass fractions
-        $X_k$ to be the $(X_k)_{\rm out}$ returned from the table.
-
-        Finally, we compute the energy generation rate, $\Sdot$ as:
-        \begin{equation}
-          \label{eq:nse_energy}
-          \Sdot = \left [ \left ( \frac{B}{A} \right )_{\rm out} -
-            \left ( \frac{B}{A} \right )_{\rm in} \right ] N_A \frac{1}{\Delta t}
-        \end{equation}
-
-        We note that \citet{ma:2013} modified Eq.~\ref{eq:nse_energy}
-        to take use only a fraction (0.3) of the energy change from
-        the table, to avoid a potential instability that can arise if
-        too much energy is deposited in a zone in a single timestep
-        (they called this an inertia term).  We prefer to deal with
-        the issue of too much energy being deposited in a zone in one
-        step through the retry mechanism already built into \castro.
-
-      \item Method II: {\em predictor-corrector.}
-
-        This starts out the same as the above method:
-        \begin{equation}
-          \nse{\rho, T, (Y_e)_{\rm in}} \rightarrow \bar{A}^\star, (X_k)^\star, (B/A)^\star, dY_e/dt
-        \end{equation}
-        but we indicate with a $\star$ that most of those values are
+      change in the nuclei abundances over the (half-)timestep.  \citet{ma:2013}
+      uses a first-order in time difference to get the new composition state
+      and evaluates the energy release from the change in binding energy in the NSE
+      state.  They only apply a fraction (0.3) of the energy release to the internal
+      energy in a zone, to avoid a potential instability that can arise if too much
+      energy is added to a zone in a single timestep.  We prefer to deal with this
+      issue through the retry mechanism already built into \castro.
+
+      The approach we use begins with calling the NSE table with the input state:
+      \begin{equation}
+        \nse{\rho_\inp, T_\inp, (Y_e)_\inp} \rightarrow \bar{A}^\star, (X_k)^\star, (B/A)^\star, dY_e/dt
+        \end{equation}
+      We indicate with a $\star$ that most of those values are
       provisional, and we will seek a better value in the corrector step.
@@
       \begin{equation}
-          (Y_e)_{\rm out} = (Y_e)_{\rm in} + \Delta t \frac{dY_e}{dt}
+        (Y_e)_\out = (Y_e)_\inp + \Delta t \frac{dY_e}{dt}
       \end{equation}
@@
       (note that $\rho_{\rm out}=\rho_{\rm in}$ in the Strang reaction formulation):
-        \MarginPar{Indicate what $T$ is used.}
-        \begin{equation}
-          \nse{\rho_{\rm out}, T_{\rm ???}, (Y_e)_\out} \rightarrow \bar{A}_\out, (X_k)_\out, (B/A)_\out, dY_e/dt
-        \end{equation}
+      \begin{equation}
+        \nse{\rho_\out, T_\inp, (Y_e)_\out} \rightarrow \bar{A}_\out, (X_k)_\out, (B/A)_\out, dY_e/dt
+      \end{equation}
+      In this formulation, we have not updated the temperature. \MarginPar{discuss this more, and perhaps switch to a mode where we try to predict T}
       This corrects the composition so that it is consistent with the updated
-        $Y_e$.
-We use these values $\bar{A}_\out$, $(X_k)_\out$, and $(B/A)_\out$ to then complete the update, computing the energy as in Eq.~\ref{eq:nse_energy}.
+      $Y_e$.  We use these values $\bar{A}_\out$, $(X_k)_\out$, and $(B/A)_\out$ to then complete the update, computing the energy release, $\Sdot$ as:
+      \begin{equation}
+        \label{eq:nse_energy}
+        \Sdot = \left [ \left ( \frac{B}{A} \right )_\out -
+          \left ( \frac{B}{A} \right )_\inp \right ] N_A \frac{1}{\Delta t}
+      \end{equation}
       \MarginPar{we could imagine looping over the EOS to get an updated T and then the table to get the updated $\bar{A}$, to see if that converges?}
-
-      \end{itemize}
 
@@
 \begin{equation}
-\Uc^\star = \Uc^n + \Delta t \Adv{\Uc}^{n+1/2} + \Delta t \Rb (\Uc^n)
-\end{equation}
-Calling the NSE table on the starting state gives use:
+\Uc^\star = \Uc^n + \Delta t \Adv{\Uc}^{n+1/2} + \Delta t \left [\Rb (\Uc) \right ]^{n+1/2}
+\end{equation}
+Calling the NSE table on the starting state gives us:
 \begin{equation}
@@
 \end{equation}
-To get the reactive source, we need to use the iteratively-lagged information
-from $\Ic_q^{n+1/2,(k-1)}$:
-\begin{equation}
-\Rb(\Uc^n) \approx \left (\begin{array}{c}
-  0 \\
-  \rho^n \Ics_q(\omegadot(X_k)) \\
-  \rho^n \Ics_q(\omegadot(\alpha_l)) \\
-  0 \\
-  \Ics_q(\rho e) 
-\end{array}
-\right )
-\end{equation}
+To get the reactive source, we need to use the iteratively-lagged time-centered information
+from $\Ic_q^{n+1/2,(k-1)}$.  We only need to update the internal energy and auxiliary data,
+since together with the density, that is all that is needed for the equation of state to get the
+updated temperature.    So our sources are:
+\begin{equation}
+\Rb(\rho e) \approx \Ics_q(\rho e)
+\end{equation}
+and
+\begin{equation}
+\Rb(\rho \alpha_l) \approx \rho^{n+1/2} \Ics_q(\alpha_l)
+\end{equation}
+were $\rho^{n+1/2} = (\rho^n + \rho^\star)/2$.
+
 Since we have $(dY_e/dt)^n$ from the table, we replace the source for $Y_e$ in the above
-$\Rb(\Uc^n)$ as:
-\begin{equation}
-R(\rho Y_e) = \rho (dY_e/dt)^n
+$\Rb(\Uc^n)$ as: \MarginPar{it might be better if we had $\Ics_\Uc$}
+\begin{equation}
+R(\rho Y_e) = \rho^{n+1/2} (dY_e/dt)^n
 \end{equation}
@@
 \begin{equation}
-(\rho Y_e)^\star = (\rho Y_e)^n + \Delta t \Advss{\rho Y_e}^{n+1/2} + \Delta t \rho (dY_e/dt)^n
+(\rho Y_e)^\star = (\rho Y_e)^n + \Delta t \Advss{\rho Y_e}^{n+1/2} + \Delta t \rho^{n+1/2} (dY_e/dt)^n
 \end{equation}

### 2020-12-31 simplified_sdc_intro/paper.tex
more details

@@
 \begin{equation}
-\Rb(\rho e) \approx \Ics_q(\rho e)
+R(\rho e) \approx \Ics_q(\rho e)
 \end{equation}
@@
 \begin{equation}
-\Rb(\rho \alpha_l) \approx \rho^{n+1/2} \Ics_q(\alpha_l)
-\end{equation}
-were $\rho^{n+1/2} = (\rho^n + \rho^\star)/2$.
+R(\rho \alpha_l) \approx \rho^{n+1/2} \Ics_q(\alpha_l)
+\end{equation}
+where $\rho^{n+1/2} = (\rho^n + \rho^\star)/2$.
 
@@
 \end{equation}
+We can then get $\alpha_l^\star = (\rho \alpha_l)^\star/\rho^\star$.
 From this prediction, we get $\rho^\star$, $\alpha_l^\star$, and
@@
 temperature, $T^\star$. \MarginPar{Andy suggests time-centering $dY_e/dt$}
+\begin{equation}
+T^\star = T(\rho^\star, \alpha_l^\star, e^\star)
+\end{equation}
 We can then again call the NSE table:
@@
 \begin{equation}
-(\rho E)^{n+1,(k)} = (\rho E)^n + \Delta t \Advss{\rho E}^{n+1/2} + (\rho \Sdot)^{n+1/2}
+(\rho E)^{n+1,(k)} = (\rho E)^n + \Delta t \Advss{\rho E}^{n+1/2} + \Delta t (\rho \Sdot)^{n+1/2}
 \end{equation}

### 2021-01-02 xrb2/paper.tex
added more sec 3.1 text (in blue)

@@
 
-\begin{figure}[t]
-\centering
-\plotone{time_series_500}
-\caption{\label{fig:time_series_500} Time series of the 500~Hz run {\tt F500} showing the mean molecular weight, $\bar{A}$.}
-\end{figure}
-
-\begin{figure}[t]
-\centering
-\plotone{time_series_1000}
-\caption{\label{fig:time_series_1000} Time series of the 1000~Hz run {\tt F1000} showing the mean molecular weight, $\bar{A}$.}
-\end{figure}
+We run three models ({\tt F250}, {\tt F500}, and {\tt F1000}) with the same initial model in terms 
+of temperature but differing rotation rates. {\color{blue}We saw in \citet{flame_wave1} that 
+increasing the 
+neutron star rotation rate reduces the horizontal lengthscale of the flame. An estimate of 
+this lengthscale is given by the Rossby radius of deformation, $L_R$. The Rossby radius may be 
+thought of as the scale over which the balance between the Coriolis force and horizontal pressure 
+gradient becomes important, and is approximately:
+\begin{equation}
+	\label{eqn:rossby}
+	L_R \approx \frac{\sqrt{g H_0}}{\Omega},
+\end{equation}
+where $g$ is gravitational acceleration, $H_0$ is the atmospheric scale height, and $\Omega$ is the 
+neutron star rotation rate. In Figure \ref{fig:time_series_enuc_500} and Figure 
+\ref{fig:time_series_enuc_1000}, we use the energy generation rate $\dot{e}_\mathrm{nuc}$ at $50~\mathrm{ms}$ and $100~\mathrm{ms}$ for {\tt F500} and {\tt F1000} to discern the horizontal extent of the flame at different rotation rates.
+Taking the rightmost edge of the bright teal/green region (where significant energy generation is occuring) as the leading edge of the flame in each plot, we see that the horizontal extent of the $\tt{F1000}$ flame appears to be
+reduced compared to the lower rotation $\tt{F500}$ run. From Equation \ref{eqn:rossby}, increasing 
+the rotation rate from 500 Hz to 1000 Hz should decrease $L_R$ by a factor of two, and the greater 
+confinement from the Coriolis force should reduce the horizontal extent of the flame by a similar 
+factor. The Rossby radius is not a perfect measure of this horizontal lengthscale, however, and 
+here the difference is not a factor of two. The simulations seem to follow the theoretical
+scaling described in Equation \ref{eqn:rossby} more closely at higher rotation rates (1000 Hz and 
+up), based on the results of \citet{flame_wave1}.}
 
@@
 
-We run three models ({\tt F250}, {\tt F500}, and {\tt F1000}) with the same initial model in terms 
-of temperature but differing rotation rates. {\color{blue}We saw in \citet{flame_wave1} that 
-increasing the 
-neutron star rotation rate reduces the horizontal lengthscale of the flame. An estimate of 
-this lengthscale is given by the Rossby radius of deformation, $L_R$. The Rossby radius may be 
-thought of as the scale over which the balance between the Coriolis force and horizontal pressure 
-gradient becomes important, and is approximately:
-\begin{equation}
-	\label{eqn:rossby}
-	L_R \approx \frac{\sqrt{g H_0}}{\Omega},
-\end{equation}
-where $g$ is gravitational acceleration, $H_0$ is the atmospheric scale height, and $\Omega$ is the 
-neutron star rotation rate. Comparing Figure \ref{fig:time_series_enuc_500} and Figure 
-\ref{fig:time_series_enuc_1000}, the horizontal extent of the $\tt{F1000}$ flame does appear to be 
-reduced compared to the lower rotation $\tt{F500}$ run. From Equation \ref{eqn:rossby}, increasing 
-the rotation rate from 500 Hz to 1000 Hz should decrease $L_R$ by a factor of two, and the greater 
-confinement from the Coriolis force should reduce the horizontal extent of the flame by a similar 
-factor. The Rossby radius is not a perfect measure of this horizontal lengthscale, however, and 
-here the difference is not a factor of two. The simulations seem to follow the theoretical
-scaling described in Equation \ref{eqn:rossby} more closely at higher rotation rates (1000 Hz and 
-up), based on the results of \citet{flame_wave1}.}
+\begin{figure}[t]
+\centering
+\plotone{time_series_500}
+\caption{\label{fig:time_series_500} Time series of the 500~Hz run {\tt F500} showing the mean molecular weight, $\bar{A}$.}
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\plotone{time_series_1000}
+\caption{\label{fig:time_series_1000} Time series of the 1000~Hz run {\tt F1000} showing the mean molecular weight, $\bar{A}$.}
+\end{figure}
 
@@
 increasing the temperature and density in the flame front such that the reaction rate increases, 
-which allows the reactions to progress further. In contrast to {\tt F500} and {\tt F1000}, the 
+which allows the reactions to progress further.
+{\color{blue} The $\dot{e}_\mathrm{nuc}$ plots in Figure \ref{fig:time_series_enuc_1000} also support this interpretation, with the interior of the {\tt F1000} flame reaching higher $\dot{e}_\mathrm{nuc}$ values than for the {\tt F500} run in Figure \ref{fig:time_series_enuc_500}.} In contrast to {\tt F500} and {\tt F1000}, the
 lowest rotation run --- {\tt F250} --- failed to ignite. The lack of ignition for {\tt F250} also 

### 2021-01-02 xrb2/paper.tex
small sec 3.3 edit (in blue)

@@
 \caption{\label{fig:compare_500Hz_abar} Comparison of $\bar{A}$ for 3 different 500~Hz models with different neutron star temperatures, $T_\mathrm{star}$, and resulting envelope structures.
-Each flame is shown at 70~ms.} %This will be updated to show at 70ms
+Each flame is shown at 70~ms.}
 \end{figure}
@@
 The hottest of the three models, $\tt{F500\_3E8}$, has a faster propagating flame (this will be discussed further in Section \ref{ssec:temp_prop}).
-It also appears to have a secondary burning region (colored light orange) forming on the far left above the flame edge.
-It is likely that a small amount of hot material broke out of the flame and rose through the atmosphere.
-This effect, however, may be numerical rather than physical, resulting from the boundary conditions imposed at the axis of symmetry for these simulations.
 
@@
 Again, burning extends across the domain and high into the atmosphere and fuel layer, lacking the characteristic flame structure shown in $\bar{A}$ plots for the lower temperature $500$ Hz runs (see Figure~\ref{fig:compare_500Hz_abar}).
-A distinct mass of material -- resembling a more dramatic version of that shown in Figure~\ref{fig:compare_500Hz_abar} -- appears to have broken out of the atmosphere near the axis of symmetry.
+A distinct mass of material appears to have broken out of the atmosphere near the axis of symmetry.
 The atmosphere edge is located at $\sim 1.2\times 10^4~\mathrm{cm}$.
-As mentioned before, this may just be a result of the imposed boundary conditions rather than a physical effect.
-$\tt{F500\_4E8}$ clearly reaches much higher $\bar{A}$ values across the domain (especially at the latest snapshot, $t = 40~\mathrm{ms}$) compared to all the other runs described in this paper (see Figures \ref{fig:compare_500Hz_abar}, \ref{fig:time_series_500}, and \ref{fig:time_series_1000}).
+{\color{blue}A similar effect is also visible (to a lesser extent) in the $\tt{F500\_3E8}$ plot in Figure \ref{fig:compare_500Hz_abar}, with a faint haze of material rising above the flame.
+This effect, however, may be numerical rather than physical, resulting from the boundary conditions imposed at the axis of symmetry for these simulations.} $\tt{F500\_4E8}$ clearly reaches much higher $\bar{A}$ values across the domain (especially at the latest snapshot, $t = 40~\mathrm{ms}$) compared to all the other runs described in this paper (see Figures \ref{fig:compare_500Hz_abar}, \ref{fig:time_series_500}, and \ref{fig:time_series_1000}).
 This is not surprising, since as we have shown in our simulations with differing initial temperature structures, the burning is very temperature-sensitive.

### 2021-01-05 xrb2/paper.tex
Updated Section 3.2
- Expanded explanation of position vs. time fitting procedure
- Switched s_0 to v_0 in flame speeds table (being somewhat pedantic,
  but s_0 was negative for the 3e8 K run)
- Added note on v_0 vs. actual initial velocity
- Tentatively suggested baroclinicity plot

@@
 
-Burning in the {\tt F500} and {\tt F1000} runs is concentrated in a dense region with circular motion.
-Figure \ref{fig:urho} compares the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the {\tt F250}, {\tt F500}, and {\tt F1000} runs.
-Most of the burning for each of the simulations occurs in a high density region $\rho > 3 \times 10^5 {\rm g\,cm^{-3}}$.
-The fluid in this dense, high energy generation region undergoes vortical motion, shown in the Figure \ref{fig:uv} phase plots comparing $u$, $v$ and $\dot{e}_{\rm nuc}$.
-From the previous plots (see: streamlines?) \MarginPar{NF: need to add streamline plots?}, this most likely corresponds to the leading edge of the flame where fresh fuel is being entrained.
-This feature is not developed in the 250~Hz flame; it could potentially develop at later times (past the point at which we terminated our simulation), or the flame could just fizzle out and fail to ignite entirely.
+Burning in the {\tt F500} and {\tt F1000} runs is concentrated in a dense region with circular 
+motion. Figure \ref{fig:urho} compares the horizontal $x$-velocity $u$, density $\rho$ and the 
+nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the {\tt F250}, {\tt F500}, and {\tt F1000} 
+runs. Most of the burning for each of the simulations occurs in a high density region $\rho > 3 
+\times 10^5 {\rm g\,cm^{-3}}$. The fluid in this dense, high energy generation region undergoes 
+vortical motion, shown in the Figure \ref{fig:uv} phase plots comparing $u$, $v$ and $\dot{e}_{\rm 
+nuc}$. From the previous plots (see: streamlines/baroclinicity?) \MarginPar{NF: need to add 
+streamline plots?}, this most likely corresponds to the leading edge of the flame where fresh fuel 
+is being entrained. This feature is not developed in the 250~Hz flame; it could potentially develop 
+at later times (past the point at which we terminated our simulation), or the flame could just 
+fizzle out and fail to ignite entirely.
 
@@
 	\tablecaption{\label{table:flame_speeds} Flame speed and acceleration values measured for each 
-	simulation. $s_0$ is the speed at $t = 0$. The initial flame velocities and accelerations are 
-	derived from a quadratic least-squares fit to each of the datasets for times $t \gtrsim 
-	20~\mathrm{ms}$.}
-	\tablehead{\colhead{run} & \colhead{$s_0$ (cm s$^{-1}$)} & \colhead{$a$ (cm s$^{-2}$)}} 
+	simulation. {\color{blue}$v_0$} is the flame velocity {\color{blue}in the $+r$ direction} at $t 
+	= 0$. The initial flame 
+	velocities and accelerations are derived from a quadratic least-squares fit to each of the 
+	datasets for times $t \gtrsim 20~\mathrm{ms}$.}
+	\tablehead{\colhead{run} & \colhead{$v_0$ (cm s$^{-1}$)} & \colhead{$a$ (cm s$^{-2}$)}} 
 	\startdata
@@
 Figure~\ref{fig:flame_speeds_1} gives position vs.\ time for the $\tt{F500}$ and $\tt{F1000}$ runs 
-(blue and orange, respectively) to show the dependence on rotation rate. As some of the hotter 
-500~Hz runs exhibit significant acceleration, we fit a quadratic curve to each set of data for 
-times $t \gtrsim 20~\mathrm{ms}$, with the fit parameters giving the flame acceleration and 
-initial velocity. \MarginPar{KE: Can add an explanation of the averaging method once that's 
-decided.} These results are displayed in Table~\ref{table:flame_speeds}. \MarginPar{NF: I will 
-update the error estimates.}
+(blue and orange, respectively) to show the dependence on rotation rate. {\color{blue}In 
+\citet{flame_wave1}, we 
+used linear least-squares fits to our position vs.\ time data to estimate the propagation velocity. 
+For this study we instead fit a quadratic curve to each set of data, as some of the hotter 500~Hz 
+runs exhibit significant acceleration. We discard the data points with $t \lesssim 20~\mathrm{ms}$, 
+since these correspond to the initial transient period before the flame has begun to propagate 
+steadily.} The fit parameters yield estimates of the flame acceleration and velocity at $t = 0$, 
+which are provided in Table~\ref{table:flame_speeds}. {\color{blue}Note that the listed velocity is 
+only 
+a parameter that may be used to calculate the flame speed at an arbitrary time. It is not an 
+estimate of the true initial velocity of the flame, since the flame has not achieved ignition yet 
+at $t = 0$.} \MarginPar{KE: Outside of updating the data in the table, is there anything else we 
+want to add to this para?}
 
@@
 		for $t \gtrsim 20~\mathrm{ms}$. Note that due to its rapid acceleration, $\tt{F500\_3E8}$ 
-		(orange) is only run out to $\sim 0.07~\mathrm{s}$ to avoid surpassing the domain boundary.}
+		(orange) is only run out to $\sim 70~\mathrm{ms}$ to avoid surpassing the domain boundary.}
 \end{figure}

### 2021-01-07 simplified_sdc_intro/paper.tex
update to match latest code

@@
 \begin{equation}
+       (\rho \Sdot)^{n+1/2} = \frac{\rho^{n+1,(k)} + \rho^n}{2}
+             \left [ \left ( \frac{B}{A} \right )^{n+1,(k)} -
+                     \left ( \frac{B}{A} \right )^n \right ] N_A \frac{1}{\Delta t}
+\end{equation}
+
+{\color{red}
+Note I tried this first, but it really doesn't work.  I'd like to understand better:
+\begin{equation}
        (\rho \Sdot)^{n+1/2} = \left [ \rho^{n+1,(k)} \left ( \frac{B}{A} \right )^{n+1,(k)} -
-                            \rho^n \left ( \frac{B}{A} \right )^n \right ] N_A \frac{1}{\Delta t}
-\end{equation}\MarginPar{is this right?}
+                            \rho^n \left ( \frac{B}{A} \right )^n \right ] N_A \frac{1}{\Delta }t
+\end{equation}
+}
 and then do the final update of the total energy:
@@
 
+% run on 12-31-20
+% Castro: 20.12-80-gfe3e9c8de8
 \begin{deluxetable}{lllllll}
-\tablecaption{\label{table:react_convert_strang} Convergence ($L_1$ norm) for the reacting convergence problem using SDC integration.}
+\tablecaption{\label{table:react_convert_strang} Convergence ($L_1$ norm) for the reacting convergence problem using SDC integration (2 iterations).}
 \tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
@@
 \startdata
- $\rho$                      & $2.795 \times 10^{18}$  & 2.044  & $6.777 \times 10^{17}$  & 2.554  & $1.154 \times 10^{17}$  \\
- $\rho u$                    & $6.798 \times 10^{26}$  & 2.449  & $1.245 \times 10^{26}$  & 2.889  & $1.680 \times 10^{25}$  \\
- $\rho v$                    & $6.798 \times 10^{26}$  & 2.449  & $1.245 \times 10^{26}$  & 2.889  & $1.680 \times 10^{25}$  \\
- $\rho E$                    & $2.452 \times 10^{35}$  & 2.352  & $4.804 \times 10^{34}$  & 2.742  & $7.179 \times 10^{33}$  \\
- $\rho e$                    & $2.262 \times 10^{35}$  & 2.321  & $4.527 \times 10^{34}$  & 2.822  & $6.404 \times 10^{33}$  \\
- $T$                         & $2.236 \times 10^{21}$  & 1.691  & $6.927 \times 10^{20}$  & 2.481  & $1.240 \times 10^{20}$  \\
- $\rho X(\isotm{He}{4})$     & $2.879 \times 10^{18}$  & 2.020  & $7.096 \times 10^{17}$  & 2.529  & $1.229 \times 10^{17}$  \\
- $\rho X(\isotm{C}{12})$     & $1.697 \times 10^{17}$  & 1.950  & $4.393 \times 10^{16}$  & 2.231  & $9.357 \times 10^{15}$  \\
- $\rho X(\isotm{O}{16})$     & $1.686 \times 10^{14}$  & 1.659  & $5.338 \times 10^{13}$  & 1.957  & $1.375 \times 10^{13}$  \\
- $\rho X(\isotm{Fe}{56})$    & $2.795 \times 10^{8}$   & 2.044  & $6.777 \times 10^{7}$   & 2.554  & $1.154 \times 10^{7}$   \\
+ $\rho$                      & $2.794 \times 10^{18}$  & 2.047  & $6.760 \times 10^{17}$  & 2.563  & $1.144 \times 10^{17}$  \\
+ $\rho u$                    & $6.796 \times 10^{26}$  & 2.450  & $1.243 \times 10^{26}$  & 2.898  & $1.669 \times 10^{25}$  \\
+ $\rho v$                    & $6.796 \times 10^{26}$  & 2.450  & $1.243 \times 10^{26}$  & 2.898  & $1.669 \times 10^{25}$  \\
+ $\rho E$                    & $2.450 \times 10^{35}$  & 2.357  & $4.783 \times 10^{34}$  & 2.743  & $7.144 \times 10^{33}$  \\
+ $\rho e$                    & $2.261 \times 10^{35}$  & 2.325  & $4.513 \times 10^{34}$  & 2.821  & $6.385 \times 10^{33}$  \\
+ $T$                         & $2.236 \times 10^{21}$  & 1.690  & $6.929 \times 10^{20}$  & 2.473  & $1.248 \times 10^{20}$  \\
+ $\rho X(\isotm{He}{4})$     & $2.877 \times 10^{18}$  & 2.023  & $7.078 \times 10^{17}$  & 2.536  & $1.220 \times 10^{17}$  \\
+ $\rho X(\isotm{C}{12})$     & $1.697 \times 10^{17}$  & 1.947  & $4.402 \times 10^{16}$  & 2.222  & $9.436 \times 10^{15}$  \\
+ $\rho X(\isotm{O}{16})$     & $1.688 \times 10^{14}$  & 1.658  & $5.349 \times 10^{13}$  & 1.950  & $1.385 \times 10^{13}$  \\
+ $\rho X(\isotm{Fe}{56})$    & $2.794 \times 10^{8}$   & 2.047  & $6.760 \times 10^{7}$   & 2.563  & $1.144 \times 10^{7}$   \\
 \enddata
@@
 \caption{\label{fig:nse_ye} Profile of $Y_e$ through the middle of the domain (slicing along $x$)
-for the $128^2$ Strang method II NSE test simulation.}
+for the $128^2$ Strang NSE test simulation.}
 \end{figure}
-
-\begin{deluxetable}{lllllll}
-\tablecaption{\label{table:nse_strang_methodI} Convergence ($L_1$ norm) for the NSE convergence
-problem using Strang splitting + Method I.}
-\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
-\startdata
- $\rho$                      & $4.301 \times 10^{19}$  & 1.348  & $1.690 \times 10^{19}$  & 1.166  & $7.532 \times 10^{18}$  \\
- $\rho u$                    & $1.866 \times 10^{28}$  & 1.429  & $6.929 \times 10^{27}$  & 1.101  & $3.230 \times 10^{27}$  \\
- $\rho v$                    & $1.866 \times 10^{28}$  & 1.429  & $6.929 \times 10^{27}$  & 1.101  & $3.230 \times 10^{27}$  \\
- $\rho E$                    & $1.781 \times 10^{38}$  & 1.053  & $8.585 \times 10^{37}$  & 1.021  & $4.231 \times 10^{37}$  \\
- $\rho e$                    & $1.781 \times 10^{38}$  & 1.053  & $8.583 \times 10^{37}$  & 1.021  & $4.230 \times 10^{37}$  \\
- $T$                         & $6.853 \times 10^{21}$  & 1.044  & $3.325 \times 10^{21}$  & 1.018  & $1.642 \times 10^{21}$  \\
- $\rho Y_e$                  & $2.442 \times 10^{19}$  & 1.442  & $8.990 \times 10^{18}$  & 1.212  & $3.880 \times 10^{18}$  \\
- $\rho \bar{A}$              & $1.055 \times 10^{22}$  & 1.032  & $5.158 \times 10^{21}$  & 1.022  & $2.541 \times 10^{21}$  \\
- $\rho (B/A)$                & $4.512 \times 10^{20}$  & 1.386  & $1.727 \times 10^{20}$  & 1.207  & $7.479 \times 10^{19}$  \\
-\enddata
-\end{deluxetable}
 
@@
 \tablecaption{\label{table:nse_strang_methodII} Convergence ($L_1$ norm) for the NSE convergence
-problem using Strang splitting + Method II.}
+problem using Strang splitting.}
 \tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 

### 2021-01-07 simplified_sdc_intro/paper.tex
some updates based on the current code
some updates

@@
 A particularly difficult phase of evolution to model is the nuclear
-statistical equilibrium that sets in with helium burning for
-temperatures in excess of $4\times 10^9~\mathrm{K}$.  Physically, the
+statistical equilibrium that sets in for
+temperatures in excess of $\mbox{few} \times 10^9~\mathrm{K}$.  Physically, the
 forward and reverse rates of reaction should balance leading to an equilibrium.
@@
 using the $X_k$ directly.
-Our equation of state needs $\bar{Z}$ in addition
-to the auxiliary quantities, which is is easily computed as
-\begin{equation}
-\bar{Z} = \bar{A} Y_e
-\end{equation}
+Our equation of state needs the mean charge per nucleus, $\bar{Z}$, in addition
+to the auxiliary quantities, which is computed as
+\begin{equation}
+\bar{Z} = \bar{A} \sum_k \frac{Z_k X_k}{A_k} = \bar{A} Y_e
+\end{equation}
+(see, e.g., \citealt{flash}).
 
@@
 
-For Strang splitting, the hydrodynamic and reactive substeps over
+For Strang splitting, we alternate the reaction and hydrodynamics,
+with each operation working on the output of the previous operation.
+If we define an advection operator over a timestep $\Delta t$ acting on $\Uc$ as
+$\mathbb{A}_{\Delta t}(\Uc)$ and a reaction operator over a timestep of
+$\Delta t/2$ acting on $\Uc$ as $\mathbb{R}_{\Delta t/2}(\Uc)$, then the advance appears
+as:
+\begin{align}
+  \Uc^\star &= \mathbb{R}_{\Delta t/2}(\Uc^n) \\
+  \Uc^{n+1,\star} &= \mathbb{A}_{\Delta t}(\Uc^\star) \\
+  \Uc^{n+1} &= \mathbb{R}_{\Delta t/2}(\Uc^{n+1,\star})
+\end{align}
+Tthe hydrodynamic and reactive substeps over
 the overall time-advancemenet scheme using the aprox19 + NSE network
@@
  $T$                         & $9.885 \times 10^{20}$  & 1.207  & $4.281 \times 10^{20}$  & 1.058  & $2.057 \times 10^{20}$  \\
+ $\rho X(\isotm{He}{4})$     & $5.086 \times 10^{17}$  & 1.430  & $1.887 \times 10^{17}$  & 1.187  & $8.288 \times 10^{16}$  \\
+ $\rho X(\isotm{C}{12})$     & $1.585 \times 10^{12}$  & 1.829  & $4.462 \times 10^{11}$  & 1.433  & $1.652 \times 10^{11}$  \\
+ $\rho X(\isotm{O}{16})$     & $8.060 \times 10^{12}$  & 1.706  & $2.470 \times 10^{12}$  & 1.347  & $9.710 \times 10^{11}$  \\
+ $\rho X(\isotm{Cr}{48})$    & $1.149 \times 10^{19}$  & 1.400  & $4.353 \times 10^{18}$  & 1.258  & $1.820 \times 10^{18}$  \\
+ $\rho X(\isotm{Fe}{52})$    & $9.018 \times 10^{19}$  & 1.173  & $4.001 \times 10^{19}$  & 1.127  & $1.832 \times 10^{19}$  \\
+ $\rho X(\isotm{Fe}{54})$    & $5.036 \times 10^{20}$  & 1.731  & $1.517 \times 10^{20}$  & 1.727  & $4.582 \times 10^{19}$  \\
+ $\rho X(\isotm{Ni}{56})$    & $5.686 \times 10^{20}$  & 1.622  & $1.847 \times 10^{20}$  & 1.563  & $6.252 \times 10^{19}$  \\
  $\rho Y_e$                  & $2.379 \times 10^{19}$  & 1.678  & $7.431 \times 10^{18}$  & 1.578  & $2.489 \times 10^{18}$  \\
@@
  $\rho (B/A)$                & $4.149 \times 10^{20}$  & 1.693  & $1.283 \times 10^{20}$  & 1.562  & $4.345 \times 10^{19}$  \\
+\enddata
+\end{deluxetable}
+
+\begin{deluxetable}{lllllll}
+\tablecaption{\label{table:nse_strang_methodII} Convergence ($L_1$ norm) for the NSE convergence
+problem using the simplified-SDC algorithm.}
+\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
+           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
+           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
+\startdata
+ $\rho$                      & $4.558 \times 10^{19}$  & 1.633  & $1.470 \times 10^{19}$  & 1.578  & $4.923 \times 10^{18}$  \\
+ $\rho u$                    & $1.945 \times 10^{28}$  & 1.648  & $6.208 \times 10^{27}$  & 1.513  & $2.176 \times 10^{27}$  \\
+ $\rho v$                    & $1.945 \times 10^{28}$  & 1.648  & $6.208 \times 10^{27}$  & 1.513  & $2.176 \times 10^{27}$  \\
+ $\rho E$                    & $5.710 \times 10^{37}$  & 1.620  & $1.857 \times 10^{37}$  & 1.552  & $6.333 \times 10^{36}$  \\
+ $\rho e$                    & $5.708 \times 10^{37}$  & 1.620  & $1.857 \times 10^{37}$  & 1.553  & $6.331 \times 10^{36}$  \\
+ $T$                         & $5.873 \times 10^{20}$  & 1.621  & $1.910 \times 10^{20}$  & 1.494  & $6.779 \times 10^{19}$  \\
+ $\rho X(\isotm{He}{4})$     & $2.453 \times 10^{17}$  & 1.835  & $6.876 \times 10^{16}$  & 1.657  & $2.181 \times 10^{16}$  \\
+ $\rho X(\isotm{C}{12})$     & $1.130 \times 10^{12}$  & 1.846  & $3.145 \times 10^{11}$  & 1.678  & $9.831 \times 10^{10}$  \\
+ $\rho X(\isotm{O}{16})$     & $5.339 \times 10^{12}$  & 1.830  & $1.502 \times 10^{12}$  & 1.667  & $4.730 \times 10^{11}$  \\
+ $\rho X(\isotm{Cr}{48})$    & $6.880 \times 10^{18}$  & 1.791  & $1.988 \times 10^{18}$  & 1.601  & $6.554 \times 10^{17}$  \\
+ $\rho X(\isotm{Fe}{52})$    & $3.470 \times 10^{19}$  & 1.606  & $1.140 \times 10^{19}$  & 1.608  & $3.740 \times 10^{18}$  \\
+ $\rho X(\isotm{Fe}{54})$    & $4.766 \times 10^{20}$  & 1.764  & $1.403 \times 10^{20}$  & 1.777  & $4.093 \times 10^{19}$  \\
+ $\rho X(\isotm{Ni}{56})$    & $4.835 \times 10^{20}$  & 1.781  & $1.407 \times 10^{20}$  & 1.788  & $4.075 \times 10^{19}$  \\
+ $\rho Y_e$                  & $2.370 \times 10^{19}$  & 1.644  & $7.582 \times 10^{18}$  & 1.580  & $2.536 \times 10^{18}$  \\
+ $\rho \bar{A}$              & $4.061 \times 10^{21}$  & 1.678  & $1.270 \times 10^{21}$  & 1.675  & $3.975 \times 10^{20}$  \\
+ $\rho (B/A)$                & $4.224 \times 10^{20}$  & 1.664  & $1.333 \times 10^{20}$  & 1.602  & $4.391 \times 10^{19}$  \\
 \enddata

### 2021-01-08 xrb2/paper.tex
Added some text and velocity at t=50 ms in flame propagation section

@@
 	velocities and accelerations are derived from a quadratic least-squares fit to each of the 
-	datasets for times $t \gtrsim 20~\mathrm{ms}$.}
-	\tablehead{\colhead{run} & \colhead{$v_0$ (cm s$^{-1}$)} & \colhead{$a$ (cm s$^{-2}$)}} 
+	datasets for times $t \gtrsim 20~\mathrm{ms}$.
+{\color{blue} Using these fit parameters we calculate the velocity at $t = 50~\mathrm{ms}$ ($v_{50}$), at which point the flames have reached steady propagation.}}
+	\tablehead{\colhead{run} & \colhead{$v_0$ (cm s$^{-1}$)} & \colhead{$a$ (cm s$^{-2}$)} & \colhead{$v_{50}$ (cm s$^{-1}$)}} 
 	\startdata
-  {\tt F1000}     & $(3.38 \pm 0.11) \times 10^5$ & $(-0.43 \pm 1.57) \times 10^5$ \\
-  {\tt F500}      & $(2.66 \pm 0.27) \times 10^5$ & $(10.97 \pm 3.22) \times 10^5$ \\
-  {\tt F500\_2E8} & $(3.98 \pm 0.19) \times 10^5$ & $(-1.56 \pm 2.63) \times 10^5$ \\
-  {\tt F500\_3E8} & $(-4.56 \pm 2.92) \times 10^5$ & $(3.39 \pm 0.60) \times 10^7$ \\
+  {\tt F1000}     & $(3.414 \pm 0.008) \times 10^5$ & $(-1.03 \pm 0.13) \times 10^5$ & $(3.362 \pm 0.010) \times 10^5$\\
+  {\tt F500}      & $(3.077 \pm 0.013) \times 10^5$ & $(5.41 \pm 0.19) \times 10^5$ & $(3.347 \pm 0.016) \times 10^5$\\
+  {\tt F500\_2E8} & $(3.760 \pm 0.015) \times 10^5$ & $(1.74 \pm 0.25) \times 10^5$ & $(3.847 \pm 0.020) \times 10^5$\\
+  {\tt F500\_3E8} & $(-5.293 \pm 0.100) \times 10^5$ & $(3.57 \pm 0.02) \times 10^7$ & $(1.258 \pm 0.015) \times 10^6$\\
 	\enddata
@@
 estimate of the true initial velocity of the flame, since the flame has not achieved ignition yet 
-at $t = 0$.} \MarginPar{KE: Outside of updating the data in the table, is there anything else we 
+at $t = 0$.
+We use the fit parameters to calculate the flame speeds at $t = 50~\mathrm{ms}$ (when the flame has reached steady propagation), given in the fourth column of Table \ref{table:flame_speeds}.} \MarginPar{KE: Outside of updating the data in the table, is there anything else we
 want to add to this para?}
@@
 As the initial $T_{\mathrm{star}}$ is increased, the flame becomes greatly accelerated.
-The initial flame velocities and accelerations derived from a quadratic least-squares fit to each of the datasets are provided in Table~\ref{table:flame_speeds}.
-The reason for the acceleration of the flame over time is not entirely clear.
+The initial flame velocities and accelerations derived from a quadratic least-squares fit to each of the datasets, {\color{blue} as well as the flame velocities at $t = 50~\mathrm{ms}$ calculated using the fit parameters,} are provided in Table~\ref{table:flame_speeds}.
+{\color{blue} Comparing the flame propagation at different initial temperatures, the most robust feature is the acceleration of the $\tt{F500\_3E8}$ run at $t \sim 40~\mathrm{ms}$.} The reason for the acceleration of the flame over time is not entirely clear.
 It could be that energy released from burning begins to dominate the flame's propagation as it evolves, increasing the flame speed over time.
@@
 		for $t \gtrsim 20~\mathrm{ms}$. Note that due to its rapid acceleration, $\tt{F500\_3E8}$ 
-		(orange) is only run out to $\sim 70~\mathrm{ms}$ to avoid surpassing the domain boundary.}
+		(green) is only run out to $\sim 70~\mathrm{ms}$ to avoid surpassing the domain boundary.}
 \end{figure}

### 2021-01-08 xrb2/paper.tex
small clarification comments in red

@@
 
-\author[0000-0001-8921-3624]{N.\ Ford}
+\author[0000-0001-8921-3624]{N.\ M.\ Ford}
 \affiliation{Lawrence Berkeley National Laboratory, Berkeley, CA}
@@
 The fuel layer undergoes stable burning and essentially simmers in place, rather than forming a flame.
-This process has been observed neutron stars with accretion rates that exceed the Eddington limit \citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}; in this model with have essentially recreated the conditions found in such stars with high accretion rates.
+This process has been observed in neutron stars with accretion rates that exceed the Eddington limit \citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}; in this model with have essentially recreated the conditions found in such stars with high accretion rates.
 This model burns so strongly that it is only run out to $40$ ms.
@@
 The initial flame velocities and accelerations derived from a quadratic least-squares fit to each of the datasets, {\color{blue} as well as the flame velocities at $t = 50~\mathrm{ms}$ calculated using the fit parameters,} are provided in Table~\ref{table:flame_speeds}.
-{\color{blue} Comparing the flame propagation at different initial temperatures, the most robust feature is the acceleration of the $\tt{F500\_3E8}$ run at $t \sim 40~\mathrm{ms}$.} The reason for the acceleration of the flame over time is not entirely clear.
+{\color{blue} Comparing the flame propagation at different initial temperatures, the most robust feature is the acceleration of the $\tt{F500\_3E8}$ run at $t \sim 40~\mathrm{ms}$.} The reason for the acceleration of the flame is not entirely clear.
 It could be that energy released from burning begins to dominate the flame's propagation as it evolves, increasing the flame speed over time.
@@
 
-The $u$-$\rho$ phase plot in Figure~\ref{fig:urho_hot} shows that the horizontal velocity for the $T_{\rm star} = 3\times10^8$~K is sufficiently greater in magnitude than for the two cooler runs.
+The $u$-$\rho$ phase plot in Figure~\ref{fig:urho_hot} shows that the horizontal velocity for the $T_{\rm star} = 3\times10^8$~K is sufficiently  \MarginPar{should this say "significantly"?} greater in magnitude than for the two cooler runs.
 The $u$-$v$ plot in Figure~\ref{fig:uv_hot} further demonstrates this, as again we see that the magnitude of the velocity in the vertical direction of the flame for the hottest run is also significantly larger.
@@
 We ran a number of simulations of laterally propagating flames in XRBs in order to explore the effects and rotation and thermal structure.
-We found that increasing the rotation rate increased the energy generation rate within the flame, enhancing its propagating.
+We found that increasing the rotation rate increased the energy generation rate within the flame, enhancing its propagating \MarginPar{I thought the increased rotation rate enhanced nuclear burning, but suppressed flame propagation?}.
 This is likely due to greater confinement of the flame from the increased Coriolis force.

### 2021-01-11 xrb2/paper.tex
fix phase plot figure captions

@@
     \plotone{urho_plot_hotter}
-    \caption{\label{fig:urho_hot}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $2\times10^8$~K and $3\times10^8$~K at $t = 40$~ms.}
+    \caption{\label{fig:urho_hot}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $3\times10^8$~K and $4\times10^8$~K at $t = 40$~ms.}
 \end{figure}
@@
     \plotone{uv_plot_hotter}
-    \caption{\label{fig:uv_hot}Phase plot of the horizontal $x$-velocity $u$, vertical $y$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $2\times10^8$~K and $3\times10^8$~K at $t = 40$~ms.}
+    \caption{\label{fig:uv_hot}Phase plot of the horizontal $x$-velocity $u$, vertical $y$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $3\times10^8$~K and $4\times10^8$~K at $t = 40$~ms.}
 \end{figure}
@@
     \plotone{abar_plot_hotter}
-    \caption{\label{fig:abar_hot}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $2\times10^8$~K and $3\times10^8$~K at $t = 40$~ms.}
+    \caption{\label{fig:abar_hot}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $3\times10^8$~K and $4\times10^8$~K at $t = 40$~ms.}
 \end{figure}

### 2021-01-11 xrb2/paper.tex
add tildes

@@
 exception of \citet{altamirano2010discovery}, most observations of
-XRBs come from sources with rotation rates of 200-600Hz
+XRBs come from sources with rotation rates of 200-$600~\mathrm{Hz}$
 \citep{bilous:2019,galloway:2020} \MarginPar{check whether this is
 true for XRBs or just burst oscillations}. There are a number of
-factors that could explain this lack of observations below 200 Hz. It
+factors that could explain this lack of observations below $200~\mathrm{Hz}$. It
 could be that there is some physical process which inhibits the flame
@@
 anything to do with the flame at all, but that neutron stars in
-accreting low mass X-ray binaries rarely have rotation rates below 200
-Hz \MarginPar{look into this?}.
+accreting low mass X-ray binaries rarely have rotation rates below $200~\mathrm{Hz}$ 
+\MarginPar{look into this?}.
 
@@
 Table~\ref{table:runs}.  These simulations encompass three different
-rotation rates: 250~Hz, 500~Hz, and 1000~Hz, and for the 500~Hz run, four
+rotation rates: $250~\mathrm{Hz}$, $500~\mathrm{Hz}$, and $1000~\mathrm{Hz}$, 
+and for the $500~\mathrm{Hz}$ run, four
 different temperature profiles.  In the following subsections, we look
@@
 reduced compared to the lower rotation $\tt{F500}$ run. From Equation \ref{eqn:rossby}, increasing 
-the rotation rate from 500 Hz to 1000 Hz should decrease $L_R$ by a factor of two, and the greater 
+the rotation rate from $500~\mathrm{Hz}$ to $1000~\mathrm{Hz}$ should decrease $L_R$ by a factor of two, and the greater 
 confinement from the Coriolis force should reduce the horizontal extent of the flame by a similar 
@@
 here the difference is not a factor of two. The simulations seem to follow the theoretical
-scaling described in Equation \ref{eqn:rossby} more closely at higher rotation rates (1000 Hz and 
+scaling described in Equation \ref{eqn:rossby} more closely at higher rotation rates ($1000~\mathrm{Hz}$ and 
 up), based on the results of \citet{flame_wave1}.}
@@
 \plotone{time_series_enuc_500}
-\caption{\label{fig:time_series_enuc_500} Time series of the 500~Hz run {\tt F500} showing the nuclear energy generation rate, $\dot{e}_\mathrm{nuc}$.}
+\caption{\label{fig:time_series_enuc_500} Time series of the $500~\mathrm{Hz}$ run {\tt F500} showing the nuclear energy generation rate, $\dot{e}_\mathrm{nuc}$.}
 \end{figure}
@@
 \plotone{time_series_enuc_1000}
-\caption{\label{fig:time_series_enuc_1000} Time series of the 1000~Hz run {\tt F1000} showing the nuclear energy generation rate, $\dot{e}_\mathrm{nuc}$.}
+\caption{\label{fig:time_series_enuc_1000} Time series of the $1000~\mathrm{Hz}$ run {\tt F1000} showing the nuclear energy generation rate, $\dot{e}_\mathrm{nuc}$.}
 \end{figure}
@@
 \plotone{time_series_500}
-\caption{\label{fig:time_series_500} Time series of the 500~Hz run {\tt F500} showing the mean molecular weight, $\bar{A}$.}
+\caption{\label{fig:time_series_500} Time series of the $500~\mathrm{Hz}$ run {\tt F500} showing the mean molecular weight, $\bar{A}$.}
 \end{figure}
@@
 \plotone{time_series_1000}
-\caption{\label{fig:time_series_1000} Time series of the 1000~Hz run {\tt F1000} showing the mean molecular weight, $\bar{A}$.}
+\caption{\label{fig:time_series_1000} Time series of the $1000~\mathrm{Hz}$ run {\tt F1000} showing the mean molecular weight, $\bar{A}$.}
 \end{figure}
@@
 higher rotation rate, and the ash is actually able to move ahead of the flame front (visible in the 
-Figure  \ref{fig:time_series_1000} 100 ms snapshot). We believe that this is because the increased 
+Figure  \ref{fig:time_series_1000} $100~\mathrm{ms}$ snapshot). We believe that this is because the increased 
 rotation better confines the initial perturbation and subsequent expansion from the burning, 
@@
 streamline plots?}, this most likely corresponds to the leading edge of the flame where fresh fuel 
-is being entrained. This feature is not developed in the 250~Hz flame; it could potentially develop 
+is being entrained. This feature is not developed in the $250~\mathrm{Hz}$ flame; it could potentially develop 
 at later times (past the point at which we terminated our simulation), or the flame could just 
@@
     \plotone{urho_plot_0_1}{}
-    \caption{\label{fig:urho}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and 1000~Hz runs ({\tt F250}, {\tt F500} and {\tt F1000}) at $t = 100$~ms.
+    \caption{\label{fig:urho}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and $1000~\mathrm{Hz}$ runs ({\tt F250}, {\tt F500} and {\tt F1000}) at $t = 100~\mathrm{ms}$.
 The slightly lower $\dot{e}_{\rm nuc}$ values along the $u = 0$ axis are most likely a numerical artifact related to the density gradient setup and finite resolution of these simulations.}
@@
     \plotone{uv_plot_0_1}
-    \caption{\label{fig:uv}Phase plot of the horizontal $x$-velocity $u$, vertical $y$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and 1000~Hz runs ({\tt F250}, {\tt F500} and {\tt F1000}) at $t = 100$~ms.}
+    \caption{\label{fig:uv}Phase plot of the horizontal $x$-velocity $u$, vertical $y$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and $1000~\mathrm{Hz}$ runs ({\tt F250}, {\tt F500} and {\tt F1000}) at $t = 100~\mathrm{ms}$.}
 \end{figure}
@@
     \plotone{abar_plot_0_1}
-    \caption{\label{fig:abar}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and 1000~Hz runs ({\tt F250}, {\tt F500} and {\tt F1000}) at $t = 100$~ms.}
+    \caption{\label{fig:abar}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and $1000~\mathrm{Hz}$ runs ({\tt F250}, {\tt F500} and {\tt F1000}) at $t = 100~\mathrm{ms}$.}
 \end{figure}
@@
 	\plotone{flame_speeds_rotation.pdf}
-	\caption{\label{fig:flame_speeds_1} Flame front position vs.\ time for the standard ($10^8~\mathrm{K}$) 500 Hz and 1000 Hz runs ({\tt F500} and {\tt F1000}). The dashed lines
+	\caption{\label{fig:flame_speeds_1} Flame front position vs.\ time for the standard ($10^8~\mathrm{K}$) $500~\mathrm{Hz}$ and $1000~\mathrm{Hz}$ runs ({\tt F500} and {\tt F1000}). The dashed lines
 		show quadratic least-squares fits to the data for $t \gtrsim 20~\mathrm{ms}$.}
@@
 used linear least-squares fits to our position vs.\ time data to estimate the propagation velocity. 
-For this study we instead fit a quadratic curve to each set of data, as some of the hotter 500~Hz 
+For this study we instead fit a quadratic curve to each set of data, as some of the hotter $500~\mathrm{Hz}$ 
 runs exhibit significant acceleration. We discard the data points with $t \lesssim 20~\mathrm{ms}$, 
@@
 
-To explore the effect of different initial temperature configurations, we run four simulations fixed at a rotation rate of 500 Hz with temperatures as shown in Table~\ref{table:runs}.
-For all the 500 Hz simulations (with the exception of the coolest run, $\tt{F500}$), we set $T_\mathrm{star}$ = $T_\mathrm{hi}$, scaling $\delta T$ accordingly to maintain a consistent value of $T_\mathrm{hi} + \delta T$.
+To explore the effect of different initial temperature configurations, we run four simulations fixed at a rotation rate of $500~\mathrm{Hz}$ with temperatures as shown in Table~\ref{table:runs}.
+For all the $500~\mathrm{Hz}$ simulations (with the exception of the coolest run, $\tt{F500}$), we set $T_\mathrm{star}$ = $T_\mathrm{hi}$, scaling $\delta T$ accordingly to maintain a consistent value of $T_\mathrm{hi} + \delta T$.
 If we let $T_\mathrm{star} < T_\mathrm{hi}$, the cooler neutron star surface might act as a heat sink and siphon away energy that would otherwise go into heating the burning layer.
@@
 \plotone{compare_500Hz_abar}
-\caption{\label{fig:compare_500Hz_abar} Comparison of $\bar{A}$ for 3 different 500~Hz models with different neutron star temperatures, $T_\mathrm{star}$, and resulting envelope structures.
-Each flame is shown at 70~ms.}
+\caption{\label{fig:compare_500Hz_abar} Comparison of $\bar{A}$ for 3 different $500~\mathrm{Hz}$ models with different neutron star temperatures, $T_\mathrm{star}$, and resulting envelope structures.
+Each flame is shown at $70~\mathrm{ms}$.}
 \end{figure}
 
-Figure~\ref{fig:compare_500Hz_abar} shows $\bar{A}$ for three 500 Hz simulations with different initial temperature structures at $t = 70~\mathrm{ms}$.
+Figure~\ref{fig:compare_500Hz_abar} shows $\bar{A}$ for three $500~\mathrm{Hz}$ simulations with different initial temperature structures at $t = 70~\mathrm{ms}$.
 The hottest of the three models, $\tt{F500\_3E8}$, has a faster propagating flame (this will be discussed further in Section \ref{ssec:temp_prop}).
@@
 This process has been observed in neutron stars with accretion rates that exceed the Eddington limit \citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}; in this model with have essentially recreated the conditions found in such stars with high accretion rates.
-This model burns so strongly that it is only run out to $40$ ms.
+This model burns so strongly that it is only run out to $40~\mathrm{ms}$.
 After an initial period of burning moves across the domain, residual burning continues and eventually ignites the entire fuel layer at late times, as shown in Figures \ref{fig:4e8_stacked_enuc} and \ref{fig:4e8_stacked_abar} for three snapshots taken in the last ten seconds of the simulation.
-In Figure~\ref{fig:4e8_stacked_enuc}, $\enucdot$ reaches values of $10^{18} - 10^{20}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ across the domain and at heights up to $\sim 0.5\times 10^4$ cm.
-There is still significant burning occurring even higher, with $\enucdot$ reaching $\sim 10^{15}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ at heights up to $\sim 0.8\times 10^4$ cm.
-For comparison, the next hottest run ($\tt{F500\_3E8}$) only reaches maximum $\enucdot$ values of $10^{18}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ and a maximum height of $\sim 0.5\times 10^4$ cm, even at the latest timesteps ($\sim 70~\mathrm{ms}$, when the flame is most developed).
+In Figure~\ref{fig:4e8_stacked_enuc}, $\enucdot$ reaches values of $10^{18} - 10^{20}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ across the domain and at heights up to $\sim 0.5\times 10^4~\mathrm{cm}$.
+There is still significant burning occurring even higher, with $\enucdot$ reaching $\sim 10^{15}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ at heights up to $\sim 0.8\times 10^4~\mathrm{cm}$.
+For comparison, the next hottest run ($\tt{F500\_3E8}$) only reaches maximum $\enucdot$ values of $10^{18}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ and a maximum height of $\sim 0.5\times 10^4~\mathrm{cm}$, even at the latest timesteps ($\sim 70~\mathrm{ms}$, when the flame is most developed).
 Significant $\enucdot$ values for $\tt{F500\_3E8}$ are confined to the flame front, rather than spanning the entire domain.
@@
 Figure~\ref{fig:4e8_stacked_abar} shows $\bar{A}$ for $\tt{F500\_4E8}$.
-Again, burning extends across the domain and high into the atmosphere and fuel layer, lacking the characteristic flame structure shown in $\bar{A}$ plots for the lower temperature $500$ Hz runs (see Figure~\ref{fig:compare_500Hz_abar}).
+Again, burning extends across the domain and high into the atmosphere and fuel layer, lacking the characteristic flame structure shown in $\bar{A}$ plots for the lower temperature $500~\mathrm{Hz}$ runs (see Figure~\ref{fig:compare_500Hz_abar}).
 A distinct mass of material appears to have broken out of the atmosphere near the axis of symmetry.
@@
 	\plotone{compare_500Hz_enuc_4e8.pdf}
-	\caption{\label{fig:4e8_stacked_enuc} Time series of the 500~Hz run $\tt{F500\_4E8}$, with $T_{\mathrm{star}} = 4 \times 10^8~\mathrm{K}$, showing $\dot{e}_{nuc}$.
-This model burns so strongly that it is only run out to 40 ms; the snapshots shown here are at T = 30 ms, 35 ms, and 40 ms.}
+	\caption{\label{fig:4e8_stacked_enuc} Time series of the $500~\mathrm{Hz}$ run $\tt{F500\_4E8}$, with $T_{\mathrm{star}} = 4 \times 10^8~\mathrm{K}$, showing $\dot{e}_{nuc}$.
+This model burns so strongly that it is only run out to $40~\mathrm{ms}$; the snapshots shown here are at $T = 30~\mathrm{ms}$, $35~\mathrm{ms}$, and $40~\mathrm{ms}$.}
 \end{figure}
@@
 	\plotone{compare_500Hz_abar_4e8.pdf}
-	\caption{\label{fig:4e8_stacked_abar} Time series of the 500~Hz run $\tt{F500\_4E8}$, with $T_{\mathrm{star}} = 4 \times 10^8~\mathrm{K}$, showing $\bar{A}$.}
+	\caption{\label{fig:4e8_stacked_abar} Time series of the $500~\mathrm{Hz}$ run $\tt{F500\_4E8}$, with $T_{\mathrm{star}} = 4 \times 10^8~\mathrm{K}$, showing $\bar{A}$.}
 \end{figure}
@@
 
-The method for measuring flame propagation described in Section \ref{ssec:rot_propagation} is applied in Figure \ref{fig:flame_speeds_2} to the three 500 Hz runs with $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$.
+The method for measuring flame propagation described in Section \ref{ssec:rot_propagation} is applied in Figure \ref{fig:flame_speeds_2} to the three $500~\mathrm{Hz}$ runs with $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$.
 Due to the lack of a clear burning front in $\tt{F500\_4E8}$, we do not analyze its propagation speed and acceleration.
@@
 	\plotone{500Hz_stacked_enuc1_2order.pdf}
-	\caption{\label{fig:flame_speeds_2} Flame front position vs.\ time for the three 500 Hz runs with
+	\caption{\label{fig:flame_speeds_2} Flame front position vs.\ time for the three $500~\mathrm{Hz}$ runs with
 		$T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$. The dashed lines show quadratic least-squares fits to the data
@@
 Also of interest in these plots is that the coolest run does not seem to have developed the characteristic vortex structure at the burning front (i.e.
-where $\dot{e}_{\rm nuc}$ is greatest) that can be clearly seen for the hotter runs at this time (40~ms).
-As can be seen in Figure~\ref{fig:uv} (which was plotted at 100~ms), this does develop more at later times.
+where $\dot{e}_{\rm nuc}$ is greatest) that can be clearly seen for the hotter runs at this time ($40~\mathrm{ms}$).
+As can be seen in Figure~\ref{fig:uv} (which was plotted at $100~\mathrm{ms}$), this does develop more at later times.
 Similar to what saw when comparing the runs with different rotation rates, it would therefore appear that it takes longer for the flame to develop when $T_{\rm star}$ is cooler.
@@
     \plotone{urho_plot_hotter}
-    \caption{\label{fig:urho_hot}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $3\times10^8$~K and $4\times10^8$~K at $t = 40$~ms.}
+    \caption{\label{fig:urho_hot}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ at $t = 40~\mathrm{ms}$.}
 \end{figure}
@@
     \plotone{uv_plot_hotter}
-    \caption{\label{fig:uv_hot}Phase plot of the horizontal $x$-velocity $u$, vertical $y$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $3\times10^8$~K and $4\times10^8$~K at $t = 40$~ms.}
+    \caption{\label{fig:uv_hot}Phase plot of the horizontal $x$-velocity $u$, vertical $y$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ at $t = 40~\mathrm{ms}$.}
 \end{figure}
@@
     \plotone{abar_plot_hotter}
-    \caption{\label{fig:abar_hot}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 500~Hz runs with $T_{\rm star}$ of $10^8$~K, $3\times10^8$~K and $4\times10^8$~K at $t = 40$~ms.}
+    \caption{\label{fig:abar_hot}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ at $t = 40~\mathrm{ms}$.}
 \end{figure}

### 2021-01-11 xrb2/paper.tex
some rephrasing

@@
 thought of as the scale over which the balance between the Coriolis force and horizontal pressure 
-gradient becomes important, and is approximately:
+gradient becomes important, and is approximately given by
 \begin{equation}
@@
 neutron star rotation rate. In Figure \ref{fig:time_series_enuc_500} and Figure 
-\ref{fig:time_series_enuc_1000}, we use the energy generation rate $\dot{e}_\mathrm{nuc}$ at $50~\mathrm{ms}$ and $100~\mathrm{ms}$ for {\tt F500} and {\tt F1000} to discern the horizontal extent of the flame at different rotation rates.
-Taking the rightmost edge of the bright teal/green region (where significant energy generation is occuring) as the leading edge of the flame in each plot, we see that the horizontal extent of the $\tt{F1000}$ flame appears to be
-reduced compared to the lower rotation $\tt{F500}$ run. From Equation \ref{eqn:rossby}, increasing 
-the rotation rate from $500~\mathrm{Hz}$ to $1000~\mathrm{Hz}$ should decrease $L_R$ by a factor of two, and the greater 
+\ref{fig:time_series_enuc_1000}, we use the energy generation rate $\dot{e}_\mathrm{nuc}$ measured at $50~\mathrm{ms}$ and $100~\mathrm{ms}$ to discern the horizontal extent of the flame at different rotation rates.
+Taking the edge at greatest radius of the bright teal/green region (where significant energy generation is occuring) as the leading edge of the flame in each plot, we see that the horizontal extent of the $1000~\mathrm{Hz}$ flame ($\tt{F1000}$) appears to be
+reduced compared to the lower rotation $500~\mathrm{Hz}$ run ($\tt{F500}$). From Equation \ref{eqn:rossby}, we can see that increasing 
+the rotation rate from $500~\mathrm{Hz}$ to $1000~\mathrm{Hz}$ should decrease $L_R$ by a factor of two, and that the greater 
 confinement from the Coriolis force should reduce the horizontal extent of the flame by a similar 
-factor. The Rossby radius is not a perfect measure of this horizontal lengthscale, however, and 
-here the difference is not a factor of two. The simulations seem to follow the theoretical
+factor.
+However, the Rossby radius is only an approximate measure of this horizontal lengthscale, and in our simulations we see that this scaling does not work so well for all rotation rates.
+The simulations seem to follow the theoretical
 scaling described in Equation \ref{eqn:rossby} more closely at higher rotation rates ($1000~\mathrm{Hz}$ and 
-up), based on the results of \citet{flame_wave1}.}
+higher), based on the results of \citet{flame_wave1}.}
 
@@
 which allows the reactions to progress further.
-{\color{blue} The $\dot{e}_\mathrm{nuc}$ plots in Figure \ref{fig:time_series_enuc_1000} also support this interpretation, with the interior of the {\tt F1000} flame reaching higher $\dot{e}_\mathrm{nuc}$ values than for the {\tt F500} run in Figure \ref{fig:time_series_enuc_500}.} In contrast to {\tt F500} and {\tt F1000}, the
+{\color{blue} The $\dot{e}_\mathrm{nuc}$ plots in Figure \ref{fig:time_series_enuc_1000} also support this interpretation, with the interior\MarginPar{What is meant by interior here??
+The edge?
+The part nearest the crust?} of the {\tt F1000} flame reaching higher $\dot{e}_\mathrm{nuc}$ values than for the {\tt F500} run in Figure \ref{fig:time_series_enuc_500}.} In contrast to {\tt F500} and {\tt F1000}, the
 lowest rotation run --- {\tt F250} --- failed to ignite. The lack of ignition for {\tt F250} also 
@@
 	simulation. {\color{blue}$v_0$} is the flame velocity {\color{blue}in the $+r$ direction} at $t 
-	= 0$. The initial flame 
+	= 0~\mathrm{ms}$. The initial flame 
 	velocities and accelerations are derived from a quadratic least-squares fit to each of the 
 	datasets for times $t \gtrsim 20~\mathrm{ms}$.
-{\color{blue} Using these fit parameters we calculate the velocity at $t = 50~\mathrm{ms}$ ($v_{50}$), at which point the flames have reached steady propagation.}}
+{\color{blue} Using these fit parameters we calculate the velocity at $t = 50~\mathrm{ms}$, $v_{50}$, at which point the flames have reached steady propagation.}}
 	\tablehead{\colhead{run} & \colhead{$v_0$ (cm s$^{-1}$)} & \colhead{$a$ (cm s$^{-2}$)} & \colhead{$v_{50}$ (cm s$^{-1}$)}} 
@@
 
-Figure~\ref{fig:flame_speeds_1} gives position vs.\ time for the $\tt{F500}$ and $\tt{F1000}$ runs 
+Figure~\ref{fig:flame_speeds_1} gives the position of the flame front as a function of time for the $\tt{F500}$ and $\tt{F1000}$ runs 
 (blue and orange, respectively) to show the dependence on rotation rate. {\color{blue}In 
 \citet{flame_wave1}, we 
-used linear least-squares fits to our position vs.\ time data to estimate the propagation velocity. 
+applied a linear least-squares fit to the flame front position as a function of time to estimate the propagation velocity. 
 For this study we instead fit a quadratic curve to each set of data, as some of the hotter $500~\mathrm{Hz}$ 
-runs exhibit significant acceleration. We discard the data points with $t \lesssim 20~\mathrm{ms}$, 
+runs exhibit significant acceleration. We discarded the data points with $t \lesssim 20~\mathrm{ms}$, 
 since these correspond to the initial transient period before the flame has begun to propagate 
-steadily.} The fit parameters yield estimates of the flame acceleration and velocity at $t = 0$, 
+steadily.} The fit parameters yield estimates of the flame acceleration and velocity at $t = 0~\mathrm{ms}$, 
 which are provided in Table~\ref{table:flame_speeds}. {\color{blue}Note that the listed velocity is 
@@
 estimate of the true initial velocity of the flame, since the flame has not achieved ignition yet 
-at $t = 0$.
+at $t = 0~\mathrm{ms}$.
 We use the fit parameters to calculate the flame speeds at $t = 50~\mathrm{ms}$ (when the flame has reached steady propagation), given in the fourth column of Table \ref{table:flame_speeds}.} \MarginPar{KE: Outside of updating the data in the table, is there anything else we
@@
 {\color{blue}A similar effect is also visible (to a lesser extent) in the $\tt{F500\_3E8}$ plot in Figure \ref{fig:compare_500Hz_abar}, with a faint haze of material rising above the flame.
-This effect, however, may be numerical rather than physical, resulting from the boundary conditions imposed at the axis of symmetry for these simulations.} $\tt{F500\_4E8}$ clearly reaches much higher $\bar{A}$ values across the domain (especially at the latest snapshot, $t = 40~\mathrm{ms}$) compared to all the other runs described in this paper (see Figures \ref{fig:compare_500Hz_abar}, \ref{fig:time_series_500}, and \ref{fig:time_series_1000}).
+However, this is likely to be a numerical artefact rather than a true physical effect, and a result of the boundary conditions imposed at the axis of symmetry for these simulations.} $\tt{F500\_4E8}$ clearly reaches much higher $\bar{A}$ values across the domain (especially at the latest snapshot, $t = 40~\mathrm{ms}$) compared to all the other runs described in this paper (see Figures \ref{fig:compare_500Hz_abar}, \ref{fig:time_series_500}, and \ref{fig:time_series_1000}).
 This is not surprising, since as we have shown in our simulations with differing initial temperature structures, the burning is very temperature-sensitive.

### 2021-01-12 xrb2/paper.tex
section 3.1 clarifcation in blue

@@
 which allows the reactions to progress further.
-{\color{blue} The $\dot{e}_\mathrm{nuc}$ plots in Figure \ref{fig:time_series_enuc_1000} also support this interpretation, with the interior\MarginPar{What is meant by interior here??
+{\color{blue} The $\dot{e}_\mathrm{nuc}$ plots in Figure \ref{fig:time_series_enuc_1000} also support this interpretation, with the interior of the flame nearest to the crust\MarginPar{What is meant by interior here??
 The edge?
-The part nearest the crust?} of the {\tt F1000} flame reaching higher $\dot{e}_\mathrm{nuc}$ values than for the {\tt F500} run in Figure \ref{fig:time_series_enuc_500}.} In contrast to {\tt F500} and {\tt F1000}, the
+The part nearest the crust?} in the {\tt F1000} run reaching higher $\dot{e}_\mathrm{nuc}$ values than for the {\tt F500} run in Figure \ref{fig:time_series_enuc_500}.} In contrast to {\tt F500} and {\tt F1000}, the
 lowest rotation run --- {\tt F250} --- failed to ignite. The lack of ignition for {\tt F250} also 

### 2021-01-13 xrb2/paper.tex
minor changes

--- 2021-01-11 xrb2/paper.tex
@@
 
-The $u$-$\rho$ phase plot in Figure~\ref{fig:urho_hot} shows that the horizontal velocity for the $T_{\rm star} = 3\times10^8$~K is sufficiently  \MarginPar{should this say "significantly"?} greater in magnitude than for the two cooler runs.
+The $u$-$\rho$ phase plot in Figure~\ref{fig:urho_hot} shows that the horizontal velocity for the $T_{\rm star} = 4\times10^8$~K is significantly greater in magnitude than for the cooler runs.
 The $u$-$v$ plot in Figure~\ref{fig:uv_hot} further demonstrates this, as again we see that the magnitude of the velocity in the vertical direction of the flame for the hottest run is also significantly larger.
-This motion therefore most likely corresponds to the vortical motion rather than to the propagation of the flame itself, something which is further supported by the fact that the velocity magnitude is 2-3 orders of magnitude greater than the flame speeds listed in Table~\ref{table:flame_speeds}.
+This motion therefore most likely corresponds to the vortical motion rather than to the propagation of the flame itself, something which is further supported by the fact that the velocity magnitude is two to three orders of magnitude greater than the flame speeds listed in Table~\ref{table:flame_speeds}.
 Also of interest in these plots is that the coolest run does not seem to have developed the characteristic vortex structure at the burning front (i.e.
@@
 We ran a number of simulations of laterally propagating flames in XRBs in order to explore the effects and rotation and thermal structure.
-We found that increasing the rotation rate increased the energy generation rate within the flame, enhancing its propagating \MarginPar{I thought the increased rotation rate enhanced nuclear burning, but suppressed flame propagation?}.
+We found that increasing the rotation rate increased the energy generation rate within the flame, enhancing its propagation \MarginPar{I thought the increased rotation rate enhanced nuclear burning, but suppressed flame propagation?}.
 This is likely due to greater confinement of the flame from the increased Coriolis force.

### 2021-01-13 xrb2/paper.tex
Added quadratic curve equation to section 3.2

--- 2021-01-12 xrb2/paper.tex
@@
 
-Figure~\ref{fig:flame_speeds_1} gives the position of the flame front as a function of time for the $\tt{F500}$ and $\tt{F1000}$ runs 
-(blue and orange, respectively) to show the dependence on rotation rate. {\color{blue}In 
-\citet{flame_wave1}, we 
-applied a linear least-squares fit to the flame front position as a function of time to estimate the propagation velocity. 
-For this study we instead fit a quadratic curve to each set of data, as some of the hotter $500~\mathrm{Hz}$ 
-runs exhibit significant acceleration. We discarded the data points with $t \lesssim 20~\mathrm{ms}$, 
-since these correspond to the initial transient period before the flame has begun to propagate 
-steadily.} The fit parameters yield estimates of the flame acceleration and velocity at $t = 0~\mathrm{ms}$, 
-which are provided in Table~\ref{table:flame_speeds}. {\color{blue}Note that the listed velocity is 
-only 
-a parameter that may be used to calculate the flame speed at an arbitrary time. It is not an 
-estimate of the true initial velocity of the flame, since the flame has not achieved ignition yet 
-at $t = 0~\mathrm{ms}$.
-We use the fit parameters to calculate the flame speeds at $t = 50~\mathrm{ms}$ (when the flame has reached steady propagation), given in the fourth column of Table \ref{table:flame_speeds}.} \MarginPar{KE: Outside of updating the data in the table, is there anything else we
-want to add to this para?}
+Figure~\ref{fig:flame_speeds_1} gives the radial position of the flame front as a function of time 
+for the $\tt{F500}$ and $\tt{F1000}$ runs (blue and orange, respectively) to show the dependence on 
+rotation rate. In \citet{flame_wave1}, we applied a linear least-squares fit to the flame front 
+position as a function of time to estimate the propagation velocity. As some of the flames in this 
+set of simulation runs exhibit significant acceleration, for this study we instead fit the data 
+with a quadratic curve of the form:
+\begin{equation}
+	\label{eq:quadratic_fit}
+	\color{blue} r(t) = \half a t^2 + v_0 t + r_0
+\end{equation}
+\noindent The parameter $a$ is the acceleration of the flame, $v_0$ is the velocity at $t = 
+0~\mathrm{ms}$, and $r_0$ is the position at $t = 0~\mathrm{ms}$. We do not include the data points 
+with $t \lesssim 20~\mathrm{ms}$ when performing the fit, since these correspond to the initial 
+transient period before the flame has begun to propagate steadily. The values of $a$ and $v_0$ for 
+the full suite of simulation runs are provided in Table~\ref{table:flame_speeds}. Note that $v_0$ 
+is only a parameter that may be used to calculate the flame speed at an arbitrary time. It is not 
+an estimate of the true initial velocity of the flame, since the flame has not achieved ignition 
+yet at $t = 0~\mathrm{ms}$. We use the fit parameters to calculate the flame speeds at $t = 
+50~\mathrm{ms}$ (when the flame has reached steady propagation), given in the fourth column of 
+Table \ref{table:flame_speeds}.
 

### 2021-01-14 xrb2/paper.tex
Switch back to km/s and km/s/s in flame speed table

@@
 	\tablecaption{\label{table:flame_speeds} Flame speed and acceleration values measured for each 
-	simulation. {\color{blue}$v_0$} is the flame velocity {\color{blue}in the $+r$ direction} at $t 
+	simulation. $v_0$ is the flame velocity in the $+r$ direction at $t 
 	= 0~\mathrm{ms}$. The initial flame 
 	velocities and accelerations are derived from a quadratic least-squares fit to each of the 
-	datasets for times $t \gtrsim 20~\mathrm{ms}$.
-{\color{blue} Using these fit parameters we calculate the velocity at $t = 50~\mathrm{ms}$, $v_{50}$, at which point the flames have reached steady propagation.}}
-	\tablehead{\colhead{run} & \colhead{$v_0$ (cm s$^{-1}$)} & \colhead{$a$ (cm s$^{-2}$)} & \colhead{$v_{50}$ (cm s$^{-1}$)}} 
+	datasets for times $t \gtrsim 20~\mathrm{ms}$. Using these fit parameters we calculate the 
+	velocity at $t = 50~\mathrm{ms}$, $v_{50}$, at which point the flames have reached steady 
+	propagation.}
+	\tablehead{\colhead{run} & \colhead{$v_0$ (km s$^{-1}$)} & \colhead{$a$ (km s$^{-2}$)} & 
+	\colhead{$v_{50}$ (km s$^{-1}$)}} 
 	\startdata
-  {\tt F1000}     & $(3.414 \pm 0.008) \times 10^5$ & $(-1.03 \pm 0.13) \times 10^5$ & $(3.362 \pm 0.010) \times 10^5$\\
-  {\tt F500}      & $(3.077 \pm 0.013) \times 10^5$ & $(5.41 \pm 0.19) \times 10^5$ & $(3.347 \pm 0.016) \times 10^5$\\
-  {\tt F500\_2E8} & $(3.760 \pm 0.015) \times 10^5$ & $(1.74 \pm 0.25) \times 10^5$ & $(3.847 \pm 0.020) \times 10^5$\\
-  {\tt F500\_3E8} & $(-5.293 \pm 0.100) \times 10^5$ & $(3.57 \pm 0.02) \times 10^7$ & $(1.258 \pm 0.015) \times 10^6$\\
+  {\tt F1000}     & $3.414 \pm 0.008$ & $-1.03 \pm 0.13$ & $3.362 \pm 0.010$\\
+  {\tt F500}      & $3.077 \pm 0.013$ & $5.41 \pm 0.19$ & $3.347 \pm 0.016$\\
+  {\tt F500\_2E8} & $3.760 \pm 0.015$ & $1.74 \pm 0.25$ & $3.847 \pm 0.020$\\
+  {\tt F500\_3E8} & $-5.293 \pm 0.100$ & $357 \pm 2$ & $12.58 \pm 0.15$\\
 	\enddata

### 2021-01-17 xrb2/paper.tex
Several edits
Update flame speeds table to be in km instead of cm
Add run names (as needed) to figure captions
Add Margin notes where phase plots analysis needs to be updated to reflect corrected fig captions

--- 2021-01-13 xrb2/paper.tex
@@
 
-\MarginPar{Mike wants to check vertical profiles of atmosphere behind flame}
-
 \subsection{Effect of Rotation Rate on Flame Propagation}\label{ssec:rot_propagation}
@@
 {\color{blue} Using these fit parameters we calculate the velocity at $t = 50~\mathrm{ms}$, $v_{50}$, at which point the flames have reached steady propagation.}}
-	\tablehead{\colhead{run} & \colhead{$v_0$ (cm s$^{-1}$)} & \colhead{$a$ (cm s$^{-2}$)} & \colhead{$v_{50}$ (cm s$^{-1}$)}} 
+	\tablehead{\colhead{run} & \colhead{$v_0$ (km s$^{-1}$)} & \colhead{$a$ (km s$^{-2}$)} & \colhead{$v_{50}$ (km s$^{-1}$)}} 
 	\startdata
-  {\tt F1000}     & $(3.414 \pm 0.008) \times 10^5$ & $(-1.03 \pm 0.13) \times 10^5$ & $(3.362 \pm 0.010) \times 10^5$\\
-  {\tt F500}      & $(3.077 \pm 0.013) \times 10^5$ & $(5.41 \pm 0.19) \times 10^5$ & $(3.347 \pm 0.016) \times 10^5$\\
-  {\tt F500\_2E8} & $(3.760 \pm 0.015) \times 10^5$ & $(1.74 \pm 0.25) \times 10^5$ & $(3.847 \pm 0.020) \times 10^5$\\
-  {\tt F500\_3E8} & $(-5.293 \pm 0.100) \times 10^5$ & $(3.57 \pm 0.02) \times 10^7$ & $(1.258 \pm 0.015) \times 10^6$\\
+  {\tt F1000}     & $(3.414 \pm 0.008)$ & $(-1.03 \pm 0.13)$ & $(3.36 \pm 0.01)$\\
+  {\tt F500}      & $(3.077 \pm 0.013)$ & $(5.41 \pm 0.19)$ & $(3.35 \pm 0.02)$\\
+  {\tt F500\_2E8} & $(3.760 \pm 0.015)$ & $(1.74 \pm 0.25)$ & $(3.85 \pm 0.02)$\\
+  {\tt F500\_3E8} & $(-5.293 \pm 0.100)$ & $(357.41 \pm 2.15)$ & $(12.58 \pm 0.15)$\\
 	\enddata
@@
 \plotone{compare_500Hz_abar}
-\caption{\label{fig:compare_500Hz_abar} Comparison of $\bar{A}$ for 3 different $500~\mathrm{Hz}$ models with different neutron star temperatures, $T_\mathrm{star}$, and resulting envelope structures.
+\caption{\label{fig:compare_500Hz_abar} Comparison of $\bar{A}$ for 3 different $500~\mathrm{Hz}$ models with neutron star temperatures $T_{\rm star}$ of $10^8~\mathrm{K}$, $2\times10^8~\mathrm{K}$ and $3\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_2e8}$, and $\tt{F500\_3e8}$, respectively), and resulting envelope structures.
 Each flame is shown at $70~\mathrm{ms}$.}
@@
 Figure~\ref{fig:compare_500Hz_abar} shows $\bar{A}$ for three $500~\mathrm{Hz}$ simulations with different initial temperature structures at $t = 70~\mathrm{ms}$.
-The hottest of the three models, $\tt{F500\_3E8}$, has a faster propagating flame (this will be discussed further in Section \ref{ssec:temp_prop}).
+The hottest of the three plotted models, $\tt{F500\_3E8}$, has a faster propagating flame (this will be discussed further in Section \ref{ssec:temp_prop}).
+{\color{blue} It also reaches slightly higher $\bar{A}$ values than the two cooler runs.
+We do not plot $\tt{F500\_4E8}$ here because it fails to form a clear burning front.}
 
+\MarginPar{NF: the discussion here was originally comparing $\tt{F1000}$ and $\tt{F500\_3e8}$ using the incorrect figure caption, but it's actually comparing $\tt{F500_\_4e8}$ and $\tt{F1000}$.
+need to update?}
 Similar to the $\bar{A}$ trends found for different rotation rates in Figure \ref{fig:abar}, the $\bar{A}$-$T$ phase plot in Figure~\ref{fig:abar_hot} suggests that the actual reactions happening in the hottest run may be different than for the cooler runs.
@@
 The fuel layer undergoes stable burning and essentially simmers in place, rather than forming a flame.
-This process has been observed in neutron stars with accretion rates that exceed the Eddington limit \citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}; in this model with have essentially recreated the conditions found in such stars with high accretion rates.
+This process has been observed in neutron stars with accretion rates that exceed the Eddington limit \citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}; in this model we have essentially recreated the conditions found in such stars with high accretion rates.
 This model burns so strongly that it is only run out to $40~\mathrm{ms}$.
@@
 	\caption{\label{fig:flame_speeds_2} Flame front position vs.\ time for the three $500~\mathrm{Hz}$ runs with
-		$T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$. The dashed lines show quadratic least-squares fits to the data
+		$T_{\rm star}$ of $10^8~\mathrm{K}$, $2\times10^8~\mathrm{K}$ and $3\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_2e8}$, and $\tt{F500\_3e8}$, respectively).
+The dashed lines show quadratic least-squares fits to the data
 		for $t \gtrsim 20~\mathrm{ms}$. Note that due to its rapid acceleration, $\tt{F500\_3E8}$ 
@@
 
-The $u$-$\rho$ phase plot in Figure~\ref{fig:urho_hot} shows that the horizontal velocity for the $T_{\rm star} = 4\times10^8$~K is significantly greater in magnitude than for the cooler runs.
+\MarginPar{NF: This paragraph also needs to be updated to reference $\tt{F500\_4E8}$ instead of $\tt{F500\_3E8}$}
+The $u$-$\rho$ phase plot in Figure~\ref{fig:urho_hot} shows that the horizontal velocity for the $T_{\rm star} = 3\times10^8$~K is significantly greater in magnitude than for the two cooler runs.
 The $u$-$v$ plot in Figure~\ref{fig:uv_hot} further demonstrates this, as again we see that the magnitude of the velocity in the vertical direction of the flame for the hottest run is also significantly larger.
-This motion therefore most likely corresponds to the vortical motion rather than to the propagation of the flame itself, something which is further supported by the fact that the velocity magnitude is two to three orders of magnitude greater than the flame speeds listed in Table~\ref{table:flame_speeds}.
+This motion therefore most likely corresponds to the vortical motion rather than to the propagation of the flame itself, something which is further supported by the fact that the velocity magnitude is 2-3 orders of magnitude greater than the flame speeds listed in Table~\ref{table:flame_speeds}.
 Also of interest in these plots is that the coolest run does not seem to have developed the characteristic vortex structure at the burning front (i.e.
@@
     \plotone{urho_plot_hotter}
-    \caption{\label{fig:urho_hot}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ at $t = 40~\mathrm{ms}$.}
+    \caption{\label{fig:urho_hot}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$ {\color{blue} (the latest time that $\tt{F500\_4E8}$ is run out to).
+The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.}
 \end{figure}
@@
     \plotone{uv_plot_hotter}
-    \caption{\label{fig:uv_hot}Phase plot of the horizontal $x$-velocity $u$, vertical $y$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ at $t = 40~\mathrm{ms}$.}
+    \caption{\label{fig:uv_hot}Phase plot of the horizontal $x$-velocity $u$, vertical $y$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$.
+{\color{blue} The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.}
 \end{figure}
@@
     \plotone{abar_plot_hotter}
-    \caption{\label{fig:abar_hot}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ at $t = 40~\mathrm{ms}$.}
+    \caption{\label{fig:abar_hot}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$.
+{\color{blue} The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.}
 \end{figure}
@@
 We ran a number of simulations of laterally propagating flames in XRBs in order to explore the effects and rotation and thermal structure.
-We found that increasing the rotation rate increased the energy generation rate within the flame, enhancing its propagation \MarginPar{I thought the increased rotation rate enhanced nuclear burning, but suppressed flame propagation?}.
+We found that increasing the rotation rate increased the energy generation rate within the flame, enhancing its propagating \MarginPar{I thought the increased rotation rate enhanced nuclear burning, but suppressed flame propagation?}.
 This is likely due to greater confinement of the flame from the increased Coriolis force.

### 2021-01-18 xrb2/paper.tex

@@
 
-\MarginPar{NF: This paragraph also needs to be updated to reference $\tt{F500\_4E8}$ instead of $\tt{F500\_3E8}$}
-The $u$-$\rho$ phase plot in Figure~\ref{fig:urho_hot} shows that the horizontal velocity for the $T_{\rm star} = 3\times10^8$~K is significantly greater in magnitude than for the two cooler runs.
-The $u$-$v$ plot in Figure~\ref{fig:uv_hot} further demonstrates this, as again we see that the magnitude of the velocity in the vertical direction of the flame for the hottest run is also significantly larger.
-This motion therefore most likely corresponds to the vortical motion rather than to the propagation of the flame itself, something which is further supported by the fact that the velocity magnitude is 2-3 orders of magnitude greater than the flame speeds listed in Table~\ref{table:flame_speeds}.
+The $u$-$\rho$ phase plot in Figure~\ref{fig:urho_hot} shows that the horizontal velocity for the $T_{\rm star} = 4\times10^8$~K is significantly greater in magnitude than for the two cooler runs.
+The $u$-$v$ plot in Figure~\ref{fig:uv_hot} further demonstrates this, as again we see that the magnitude of the velocity in the
+vertical direction of the flame for the hottest run is also significantly larger.
+This motion therefore most likely corresponds to the vortical motion rather than to the propagation of the flame itself, something which is further supported by the fact that the velocity magnitude is two to three orders of magnitude greater than the flame speeds listed in Table~\ref{table:flame_speeds}.
 Also of interest in these plots is that the coolest run does not seem to have developed the characteristic vortex structure at the burning front (i.e.

### 2021-01-20 xrb2/paper.tex
Updates to section 3.3

@@
 
- \MarginPar{NF: I think we should plot peak enucdot as a function of time for each run, it would be useful for me to reference it in one of the later sections}
-
 To explore the effect of different initial temperature configurations, we run four simulations fixed at a rotation rate of $500~\mathrm{Hz}$ with temperatures as shown in Table~\ref{table:runs}.
@@
 {\color{blue} It also reaches slightly higher $\bar{A}$ values than the two cooler runs.
-We do not plot $\tt{F500\_4E8}$ here because it fails to form a clear burning front.}
-
-\MarginPar{NF: the discussion here was originally comparing $\tt{F1000}$ and $\tt{F500\_3e8}$ using the incorrect figure caption, but it's actually comparing $\tt{F500_\_4e8}$ and $\tt{F1000}$.
-need to update?}
-Similar to the $\bar{A}$ trends found for different rotation rates in Figure \ref{fig:abar}, the $\bar{A}$-$T$ phase plot in Figure~\ref{fig:abar_hot} suggests that the actual reactions happening in the hottest run may be different than for the cooler runs.
-Not only is the $\tt{F500\_3E8}$ flame itself significantly hotter overall, but there is a large causally connected region across a wide range of $\bar{A}$.
-This indicates that the hotter temperature has facilitated significant burning in reaction pathways that were not favored at the lower temperatures.
-The reaction pathways are also very disrupted for $\tt{F500\_3E8}$ compared to the cooler runs, suggesting that the hotter temperature leads to more turbulent mixing.
-Compared to the highest rotation rate run, $\tt{F1000}$, $\tt{F500\_3E8}$ reaches a peak temperature roughly $1 \times 10^9~\mathrm{K}$ higher ($\sim 2.7 \times 10^9~\mathrm{K}$ versus $\sim 1.7 \times 10^9~\mathrm{K}$).
-Though increasing the rotation rate does seem to concentrate the burning and increase the peak temperature, it does not reach the same level as directly increasing the temperature at the crust.
+We do not plot $\tt{F500\_4E8}$ here because it fails to form a clear burning front.
+The Figure~\ref{fig:abar_hot} $\bar{A}$-$T$ phase plots of $\tt{F500}$ (left) and $\tt{F500\_3E8}$ (middle) also reflect these $\bar{A}$ features, with $\tt{F500\_3E8}$ reaching slightly higher $\bar{A}$ values.
+$\tt{F500\_3E8}$ also reaches higher $\enucdot$ values at the low end of the temperature range ($< 0.5 \times 10^9 \mathrm{K}$).
+There appear to be more causally connected regions across a range of $\bar{A}$ at low temperatures for $\tt{F500\_3E8}$ than for $\tt{F500}$, suggesting that the higher $\enucdot$ for $\tt{F500\_3E8}$ generates burning in certain reaction pathways that are not present in the cooler $\tt{F500}$ run.}
+\MarginPar{relate this section back to Kiran's max temperature and $\enucdot$ plots once they're added, note that $\tt{F500\_3E8}$ doesn't reach as high temperatures, less evolved abar than $\tt{F1000}$}
 
@@
 This model burns so strongly that it is only run out to $40~\mathrm{ms}$.
-After an initial period of burning moves across the domain, residual burning continues and eventually ignites the entire fuel layer at late times, as shown in Figures \ref{fig:4e8_stacked_enuc} and \ref{fig:4e8_stacked_abar} for three snapshots taken in the last ten seconds of the simulation.
+After an initial period when the burning moves across the domain, residual burning continues and eventually ignites the entire fuel layer at late times, as shown in Figures \ref{fig:4e8_stacked_enuc} and \ref{fig:4e8_stacked_abar} for three snapshots taken in the last ten seconds of the simulation.
 In Figure~\ref{fig:4e8_stacked_enuc}, $\enucdot$ reaches values of $10^{18} - 10^{20}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ across the domain and at heights up to $\sim 0.5\times 10^4~\mathrm{cm}$.
 There is still significant burning occurring even higher, with $\enucdot$ reaching $\sim 10^{15}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ at heights up to $\sim 0.8\times 10^4~\mathrm{cm}$.
-For comparison, the next hottest run ($\tt{F500\_3E8}$) only reaches maximum $\enucdot$ values of $10^{18}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ and a maximum height of $\sim 0.5\times 10^4~\mathrm{cm}$, even at the latest timesteps ($\sim 70~\mathrm{ms}$, when the flame is most developed).
-Significant $\enucdot$ values for $\tt{F500\_3E8}$ are confined to the flame front, rather than spanning the entire domain.
+For comparison, the next hottest run ($\tt{F500\_3E8}$) only reaches maximum $\enucdot$ values on the order of $10^{18}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ {\color{blue} (see Figure ??
+Kiran $\enucdot$ plot)}, even at the latest timesteps ($\sim 70~\mathrm{ms}$, when the flame is most developed).
+{\color{blue} Significant $\enucdot$ values for all other runs are confined to the flame front, rather than spanning the entire domain}.
 
-\MarginPar{NF: should I include a plot for this or is it ok to just say it?}
 Figure~\ref{fig:4e8_stacked_abar} shows $\bar{A}$ for $\tt{F500\_4E8}$.
@@
 This is not surprising, since as we have shown in our simulations with differing initial temperature structures, the burning is very temperature-sensitive.
+\MarginPar{is it generally temperature-sensitive, or only sensitive to the highest temperatures?}
+
+ {\color{blue} Though $\tt{F500\_4E8}$ does not form a distinct burning front, it does achieve greatly enhanced burning, as shown in the Figure~\ref{fig:abar_hot} $\bar{A}$-$T$ phase plot (right panel).
+Not only is $\tt{F500\_4E8}$ significantly hotter overall compared to all other runs}, but there is a large causally connected region across a wide range of $\bar{A}$.
+This indicates that the hotter temperature has facilitated significant burning in reaction pathways that were not favored at the lower temperatures.
+The reaction pathways are also very disrupted for $\tt{F500\_4E8}$ compared to the cooler runs, suggesting that the hotter temperature leads to more vigorous mixing.
+{\color{blue}Indeed, this appears to be the case looking at Figure~\ref{fig:4e8_stacked_abar} and the right panel of Figure~\ref{fig:uv_hot} (discussed further in section~\ref{ssec:temp_prop}).
+The disrupted reaction pathways resemble those found in Figure \ref{fig:abar} for the highest rotation $\tt{F1000}$ run, though they are even more dramatically disrupted for $\tt{F500\_4E8}$.
+$\tt{F500\_4E8}$ is also noticably hotter than $\tt{F1000}$, even though it is run for significantly less time ($t = 40~\mathrm{ms}$ vs $t = 100~\mathrm{ms}$).
+While the correlation between increased temperature, greater mixing, and enhanced burning seen in $\tt{F500\_4E8}$ agrees with our previous anaylses, without an increased rotation and Coriolis force the $\tt{F500\_4E8}$ run seems insufficiently confined to maintain a burning front.} \MarginPar{is this interpretation of the lack of Coriolis force correct?}
 

### 2021-01-20 xrb2/paper.tex
added a couple sentences to Section 3.3

@@
 However, this is likely to be a numerical artefact rather than a true physical effect, and a result of the boundary conditions imposed at the axis of symmetry for these simulations.} $\tt{F500\_4E8}$ clearly reaches much higher $\bar{A}$ values across the domain (especially at the latest snapshot, $t = 40~\mathrm{ms}$) compared to all the other runs described in this paper (see Figures \ref{fig:compare_500Hz_abar}, \ref{fig:time_series_500}, and \ref{fig:time_series_1000}).
-This is not surprising, since as we have shown in our simulations with differing initial temperature structures, the burning is very temperature-sensitive.
-\MarginPar{is it generally temperature-sensitive, or only sensitive to the highest temperatures?}
 
@@
 $\tt{F500\_4E8}$ is also noticably hotter than $\tt{F1000}$, even though it is run for significantly less time ($t = 40~\mathrm{ms}$ vs $t = 100~\mathrm{ms}$).
-While the correlation between increased temperature, greater mixing, and enhanced burning seen in $\tt{F500\_4E8}$ agrees with our previous anaylses, without an increased rotation and Coriolis force the $\tt{F500\_4E8}$ run seems insufficiently confined to maintain a burning front.} \MarginPar{is this interpretation of the lack of Coriolis force correct?}
+While the correlation between increased temperature, greater mixing, and enhanced burning seen in $\tt{F500\_4E8}$ is present throughout our simulations, $\tt{F500\_4E8}$ is unique in that it seems insufficiently confined to maintain a burning front.} \MarginPar{is it correct to say it's insufficiently confined?} {\color{blue} Enhanced burning in this hottest run aligns with results from the other simulations with differing initial temperature structures; as $T_{\mathrm{star}}$ is increased, the overall behavior and propagation of the flame is significantly altered, implying that burning is very temperature-sensitive.
+We explore flame propagation further in Section~\ref{ssec:temp_prop}.}
 

### 2021-01-21 xrb2/paper.tex
fix spelling

@@
 \ref{fig:time_series_enuc_1000}, we use the energy generation rate $\dot{e}_\mathrm{nuc}$ measured at $50~\mathrm{ms}$ and $100~\mathrm{ms}$ to discern the horizontal extent of the flame at different rotation rates.
-Taking the edge at greatest radius of the bright teal/green region (where significant energy generation is occuring) as the leading edge of the flame in each plot, we see that the horizontal extent of the $1000~\mathrm{Hz}$ flame ($\tt{F1000}$) appears to be
+Taking the edge at greatest radius of the bright teal/green region (where significant energy generation is occurring) as the leading edge of the flame in each plot, we see that the horizontal extent of the $1000~\mathrm{Hz}$ flame ($\tt{F1000}$) appears to be
 reduced compared to the lower rotation $500~\mathrm{Hz}$ run ($\tt{F500}$). From Equation \ref{eqn:rossby}, we can see that increasing 
@@
 The disrupted reaction pathways resemble those found in Figure \ref{fig:abar} for the highest rotation $\tt{F1000}$ run, though they are even more dramatically disrupted for $\tt{F500\_4E8}$.
-$\tt{F500\_4E8}$ is also noticably hotter than $\tt{F1000}$, even though it is run for significantly less time ($t = 40~\mathrm{ms}$ vs $t = 100~\mathrm{ms}$).
+$\tt{F500\_4E8}$ is also noticeably hotter than $\tt{F1000}$, even though it is run for significantly less time ($t = 40~\mathrm{ms}$ vs $t = 100~\mathrm{ms}$).
 While the correlation between increased temperature, greater mixing, and enhanced burning seen in $\tt{F500\_4E8}$ is present throughout our simulations, $\tt{F500\_4E8}$ is unique in that it seems insufficiently confined to maintain a burning front.} \MarginPar{is it correct to say it's insufficiently confined?} {\color{blue} Enhanced burning in this hottest run aligns with results from the other simulations with differing initial temperature structures; as $T_{\mathrm{star}}$ is increased, the overall behavior and propagation of the flame is significantly altered, implying that burning is very temperature-sensitive.

### 2021-01-21 xrb2/paper.tex
Added enuc and temperature maximum plot

--- 2021-01-20 xrb2/paper.tex
@@
 \newcommand{\avgtwod}[1]{\langle~#1 \rangle}
+\newcommand{\repavg}[1]{\langle \langle #1 \rangle \rangle}
 \newcommand{\rms}[2]{\left(\delta#1\right)_{r_{#2}}}
@@
 
+\begin{figure}
+	\plotone{maxima_vs_time.pdf}
+	\caption{\label{fig:maxima}Estimates of temperature (left panel) and nuclear energy generation 
+	rate (right panel) in the burning region as functions of time. The quantities on the vertical 
+	axes are the averages defined by Equation \ref{eqn:top_percentile_avg}.}
+\end{figure}
+
+{\color{blue}
+For some of our analysis, we would like to have a means of estimating the temperature and nuclear 
+energy generation rate in the burning region of each simulation. For this purpose, we define 
+$\repavg{Q}$ to be the mass weighted average
+\begin{equation}
+	\label{eqn:top_percentile_avg}
+	\repavg{Q} \equiv \frac{\sum_{c_i} m(c_i) Q(c_i)}{\sum_{c_i} m(c_i)}; ~c_i \in C_{99}(Q).
+\end{equation}
+\noindent Here, $C_{99}(Q)$ is the set of grid cells with $Q$ values in the top percentile, and 
+$Q(c_i)$ is the value of the quantity $Q$ in cell $c_i$. An alternative to doing this averaging is 
+to simply take the maximum of the quantity across the entire simulation domain, but using 
+$\repavg{Q}$ instead allows us to track changes over a region rather than at a single localized 
+point. Figure \ref{fig:maxima} shows $\repavg{T}$ and $\repavg{\enucdot}$ as functions of time for 
+the subset of our runs that achieve a propagating flame. This figure is referenced at several 
+points throughout the subsequent sections.
+}
+
 \subsection{Effect of Rotation Rate on Flame Structure}\label{ssec:rot_structure}
@@
 
+\MarginPar{Mike wants to check vertical profiles of atmosphere behind flame}
+
 \subsection{Effect of Rotation Rate on Flame Propagation}\label{ssec:rot_propagation}
@@
 	\tablecaption{\label{table:flame_speeds} Flame speed and acceleration values measured for each 
-	simulation. {\color{blue}$v_0$} is the flame velocity {\color{blue}in the $+r$ direction} at $t 
+	simulation. $v_0$ is the flame velocity in the $+r$ direction at $t 
 	= 0~\mathrm{ms}$. The initial flame 
 	velocities and accelerations are derived from a quadratic least-squares fit to each of the 
-	datasets for times $t \gtrsim 20~\mathrm{ms}$.
-{\color{blue} Using these fit parameters we calculate the velocity at $t = 50~\mathrm{ms}$, $v_{50}$, at which point the flames have reached steady propagation.}}
-	\tablehead{\colhead{run} & \colhead{$v_0$ (km s$^{-1}$)} & \colhead{$a$ (km s$^{-2}$)} & \colhead{$v_{50}$ (km s$^{-1}$)}} 
+	datasets for times $t \gtrsim 20~\mathrm{ms}$. Using these fit parameters we calculate the 
+	velocity at $t = 50~\mathrm{ms}$, $v_{50}$, at which point the flames have reached steady 
+	propagation.}
+	\tablehead{\colhead{run} & \colhead{$v_0$ (km s$^{-1}$)} & \colhead{$a$ (km s$^{-2}$)} & 
+	\colhead{$v_{50}$ (km s$^{-1}$)}} 
 	\startdata
-  {\tt F1000}     & $(3.414 \pm 0.008)$ & $(-1.03 \pm 0.13)$ & $(3.36 \pm 0.01)$\\
-  {\tt F500}      & $(3.077 \pm 0.013)$ & $(5.41 \pm 0.19)$ & $(3.35 \pm 0.02)$\\
-  {\tt F500\_2E8} & $(3.760 \pm 0.015)$ & $(1.74 \pm 0.25)$ & $(3.85 \pm 0.02)$\\
-  {\tt F500\_3E8} & $(-5.293 \pm 0.100)$ & $(357.41 \pm 2.15)$ & $(12.58 \pm 0.15)$\\
+  {\tt F1000}     & $3.414 \pm 0.008$ & $-1.03 \pm 0.13$ & $3.362 \pm 0.010$\\
+  {\tt F500}      & $3.077 \pm 0.013$ & $5.41 \pm 0.19$ & $3.347 \pm 0.016$\\
+  {\tt F500\_2E8} & $3.760 \pm 0.015$ & $1.74 \pm 0.25$ & $3.847 \pm 0.020$\\
+  {\tt F500\_3E8} & $-5.293 \pm 0.100$ & $357 \pm 2$ & $12.58 \pm 0.15$\\
 	\enddata
@@
 
-Figure~\ref{fig:flame_speeds_1} gives the position of the flame front as a function of time for the $\tt{F500}$ and $\tt{F1000}$ runs 
-(blue and orange, respectively) to show the dependence on rotation rate. {\color{blue}In 
-\citet{flame_wave1}, we 
-applied a linear least-squares fit to the flame front position as a function of time to estimate the propagation velocity. 
-For this study we instead fit a quadratic curve to each set of data, as some of the hotter $500~\mathrm{Hz}$ 
-runs exhibit significant acceleration. We discarded the data points with $t \lesssim 20~\mathrm{ms}$, 
-since these correspond to the initial transient period before the flame has begun to propagate 
-steadily.} The fit parameters yield estimates of the flame acceleration and velocity at $t = 0~\mathrm{ms}$, 
-which are provided in Table~\ref{table:flame_speeds}. {\color{blue}Note that the listed velocity is 
-only 
-a parameter that may be used to calculate the flame speed at an arbitrary time. It is not an 
-estimate of the true initial velocity of the flame, since the flame has not achieved ignition yet 
-at $t = 0~\mathrm{ms}$.
-We use the fit parameters to calculate the flame speeds at $t = 50~\mathrm{ms}$ (when the flame has reached steady propagation), given in the fourth column of Table \ref{table:flame_speeds}.} \MarginPar{KE: Outside of updating the data in the table, is there anything else we
-want to add to this para?}
+Figure~\ref{fig:flame_speeds_1} gives the radial position of the flame front as a function of time 
+for the $\tt{F500}$ and $\tt{F1000}$ runs (blue and orange, respectively) to show the dependence on 
+rotation rate. In \citet{flame_wave1}, we applied a linear least-squares fit to the flame front 
+position as a function of time to estimate the propagation velocity. As some of the flames in this 
+set of simulation runs exhibit significant acceleration, for this study we instead fit the data 
+with a quadratic curve of the form:
+\begin{equation}
+	\label{eqn:quadratic_fit}
+	\color{blue} r(t) = \half a t^2 + v_0 t + r_0
+\end{equation}
+\noindent The parameter $a$ is the acceleration of the flame, $v_0$ is the velocity at $t = 
+0~\mathrm{ms}$, and $r_0$ is the position at $t = 0~\mathrm{ms}$. We do not include the data points 
+with $t \lesssim 20~\mathrm{ms}$ when performing the fit, since these correspond to the initial 
+transient period before the flame has begun to propagate steadily. The values of $a$ and $v_0$ for 
+the full suite of simulation runs are provided in Table~\ref{table:flame_speeds}. Note that $v_0$ 
+is only a parameter that may be used to calculate the flame speed at an arbitrary time. It is not 
+an estimate of the true initial velocity of the flame, since the flame has not achieved ignition 
+yet at $t = 0~\mathrm{ms}$. We use the fit parameters to calculate the flame speeds at $t = 
+50~\mathrm{ms}$ (when the flame has reached steady propagation), given in the fourth column of 
+Table \ref{table:flame_speeds}.
 
@@
 flame speed with rotation rate in this set of runs. We observed earlier, however, that nuclear 
-reactions progress more quickly at higher rotation rate. This results in a higher $\enucdot$ --- up 
-to four to six times higher near the burning front after the flame ignites --- which may counteract 
-the reduction in flame speed from the increased Coriolis confinement. We also observe that 
-$\tt{F500}$ accelerates faster than $\tt{F1000}$, which appears to experience a small deceleration 
-at early times. This disparity may be a direct result of the difference in Coriolis force.
+reactions progress more quickly at higher rotation rate. {\color{blue}This results in a higher 
+$\enucdot$ --- up 
+to three to four times higher near the burning front after the flame ignites (see Figure 
+\ref{fig:maxima}) --- which may counteract the reduction in flame speed from the increased Coriolis 
+confinement.} We also observe that $\tt{F500}$ accelerates faster than $\tt{F1000}$, which appears 
+to experience a small deceleration at early times. This disparity may be a direct result of the 
+difference in Coriolis force.
 
@@
 \subsection{Effect of Temperature on Flame Structure}\label{ssec:temp_structure}
+
+ \MarginPar{NF: I think we should plot peak enucdot as a function of time for each run, it would be useful for me to reference it in one of the later sections}
 
@@
 \plotone{compare_500Hz_abar}
-\caption{\label{fig:compare_500Hz_abar} Comparison of $\bar{A}$ for 3 different $500~\mathrm{Hz}$ models with neutron star temperatures $T_{\rm star}$ of $10^8~\mathrm{K}$, $2\times10^8~\mathrm{K}$ and $3\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_2e8}$, and $\tt{F500\_3e8}$, respectively), and resulting envelope structures.
+\caption{\label{fig:compare_500Hz_abar} Comparison of $\bar{A}$ for 3 different $500~\mathrm{Hz}$ models with different neutron star temperatures, $T_\mathrm{star}$, and resulting envelope structures.
 Each flame is shown at $70~\mathrm{ms}$.}
@@
 Figure~\ref{fig:compare_500Hz_abar} shows $\bar{A}$ for three $500~\mathrm{Hz}$ simulations with different initial temperature structures at $t = 70~\mathrm{ms}$.
-The hottest of the three plotted models, $\tt{F500\_3E8}$, has a faster propagating flame (this will be discussed further in Section \ref{ssec:temp_prop}).
-{\color{blue} It also reaches slightly higher $\bar{A}$ values than the two cooler runs.
-We do not plot $\tt{F500\_4E8}$ here because it fails to form a clear burning front.
-The Figure~\ref{fig:abar_hot} $\bar{A}$-$T$ phase plots of $\tt{F500}$ (left) and $\tt{F500\_3E8}$ (middle) also reflect these $\bar{A}$ features, with $\tt{F500\_3E8}$ reaching slightly higher $\bar{A}$ values.
-$\tt{F500\_3E8}$ also reaches higher $\enucdot$ values at the low end of the temperature range ($< 0.5 \times 10^9 \mathrm{K}$).
-There appear to be more causally connected regions across a range of $\bar{A}$ at low temperatures for $\tt{F500\_3E8}$ than for $\tt{F500}$, suggesting that the higher $\enucdot$ for $\tt{F500\_3E8}$ generates burning in certain reaction pathways that are not present in the cooler $\tt{F500}$ run.}
-\MarginPar{relate this section back to Kiran's max temperature and $\enucdot$ plots once they're added, note that $\tt{F500\_3E8}$ doesn't reach as high temperatures, less evolved abar than $\tt{F1000}$}
+The hottest of the three models, $\tt{F500\_3E8}$, has a faster propagating flame (this will be discussed further in Section \ref{ssec:temp_prop}).
+
+Similar to the $\bar{A}$ trends found for different rotation rates in Figure \ref{fig:abar}, the $\bar{A}$-$T$ phase plot in Figure~\ref{fig:abar_hot} suggests that the actual reactions happening in the hottest run may be different than for the cooler runs.
+Not only is the $\tt{F500\_3E8}$ flame itself significantly hotter overall, but there is a large causally connected region across a wide range of $\bar{A}$.
+This indicates that the hotter temperature has facilitated significant burning in reaction pathways that were not favored at the lower temperatures.
+The reaction pathways are also very disrupted for $\tt{F500\_3E8}$ compared to the cooler runs, suggesting that the hotter temperature leads to more turbulent mixing.
+Compared to the highest rotation rate run, $\tt{F1000}$, $\tt{F500\_3E8}$ reaches a peak temperature roughly $1 \times 10^9~\mathrm{K}$ higher ($\sim 2.7 \times 10^9~\mathrm{K}$ versus $\sim 1.7 \times 10^9~\mathrm{K}$).
+Though increasing the rotation rate does seem to concentrate the burning and increase the peak temperature, it does not reach the same level as directly increasing the temperature at the crust.
 
@@
 The fuel layer undergoes stable burning and essentially simmers in place, rather than forming a flame.
-This process has been observed in neutron stars with accretion rates that exceed the Eddington limit \citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}; in this model we have essentially recreated the conditions found in such stars with high accretion rates.
+This process has been observed in neutron stars with accretion rates that exceed the Eddington limit \citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}; in this model with have essentially recreated the conditions found in such stars with high accretion rates.
 This model burns so strongly that it is only run out to $40~\mathrm{ms}$.
-After an initial period when the burning moves across the domain, residual burning continues and eventually ignites the entire fuel layer at late times, as shown in Figures \ref{fig:4e8_stacked_enuc} and \ref{fig:4e8_stacked_abar} for three snapshots taken in the last ten seconds of the simulation.
+After an initial period of burning moves across the domain, residual burning continues and eventually ignites the entire fuel layer at late times, as shown in Figures \ref{fig:4e8_stacked_enuc} and \ref{fig:4e8_stacked_abar} for three snapshots taken in the last ten seconds of the simulation.
 In Figure~\ref{fig:4e8_stacked_enuc}, $\enucdot$ reaches values of $10^{18} - 10^{20}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ across the domain and at heights up to $\sim 0.5\times 10^4~\mathrm{cm}$.
 There is still significant burning occurring even higher, with $\enucdot$ reaching $\sim 10^{15}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ at heights up to $\sim 0.8\times 10^4~\mathrm{cm}$.
-For comparison, the next hottest run ($\tt{F500\_3E8}$) only reaches maximum $\enucdot$ values on the order of $10^{18}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ {\color{blue} (see Figure ??
-Kiran $\enucdot$ plot)}, even at the latest timesteps ($\sim 70~\mathrm{ms}$, when the flame is most developed).
-{\color{blue} Significant $\enucdot$ values for all other runs are confined to the flame front, rather than spanning the entire domain}.
+For comparison, the next hottest run ($\tt{F500\_3E8}$) only reaches maximum $\enucdot$ values of $10^{18}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ and a maximum height of $\sim 0.5\times 10^4~\mathrm{cm}$, even at the latest timesteps ($\sim 70~\mathrm{ms}$, when the flame is most developed).
+Significant $\enucdot$ values for $\tt{F500\_3E8}$ are confined to the flame front, rather than spanning the entire domain.
 
+\MarginPar{NF: should I include a plot for this or is it ok to just say it?}
 Figure~\ref{fig:4e8_stacked_abar} shows $\bar{A}$ for $\tt{F500\_4E8}$.
@@
 This is not surprising, since as we have shown in our simulations with differing initial temperature structures, the burning is very temperature-sensitive.
-\MarginPar{is it generally temperature-sensitive, or only sensitive to the highest temperatures?}
-
- {\color{blue} Though $\tt{F500\_4E8}$ does not form a distinct burning front, it does achieve greatly enhanced burning, as shown in the Figure~\ref{fig:abar_hot} $\bar{A}$-$T$ phase plot (right panel).
-Not only is $\tt{F500\_4E8}$ significantly hotter overall compared to all other runs}, but there is a large causally connected region across a wide range of $\bar{A}$.
-This indicates that the hotter temperature has facilitated significant burning in reaction pathways that were not favored at the lower temperatures.
-The reaction pathways are also very disrupted for $\tt{F500\_4E8}$ compared to the cooler runs, suggesting that the hotter temperature leads to more vigorous mixing.
-{\color{blue}Indeed, this appears to be the case looking at Figure~\ref{fig:4e8_stacked_abar} and the right panel of Figure~\ref{fig:uv_hot} (discussed further in section~\ref{ssec:temp_prop}).
-The disrupted reaction pathways resemble those found in Figure \ref{fig:abar} for the highest rotation $\tt{F1000}$ run, though they are even more dramatically disrupted for $\tt{F500\_4E8}$.
-$\tt{F500\_4E8}$ is also noticably hotter than $\tt{F1000}$, even though it is run for significantly less time ($t = 40~\mathrm{ms}$ vs $t = 100~\mathrm{ms}$).
-While the correlation between increased temperature, greater mixing, and enhanced burning seen in $\tt{F500\_4E8}$ agrees with our previous anaylses, without an increased rotation and Coriolis force the $\tt{F500\_4E8}$ run seems insufficiently confined to maintain a burning front.} \MarginPar{is this interpretation of the lack of Coriolis force correct?}
 
@@
 	\caption{\label{fig:flame_speeds_2} Flame front position vs.\ time for the three $500~\mathrm{Hz}$ runs with
-		$T_{\rm star}$ of $10^8~\mathrm{K}$, $2\times10^8~\mathrm{K}$ and $3\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_2e8}$, and $\tt{F500\_3e8}$, respectively).
-The dashed lines show quadratic least-squares fits to the data
+		$T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$. The dashed lines show quadratic least-squares fits to the data
 		for $t \gtrsim 20~\mathrm{ms}$. Note that due to its rapid acceleration, $\tt{F500\_3E8}$ 
@@
 
-The $u$-$\rho$ phase plot in Figure~\ref{fig:urho_hot} shows that the horizontal velocity for the $T_{\rm star} = 4\times10^8$~K is significantly greater in magnitude than for the two cooler runs.
-The $u$-$v$ plot in Figure~\ref{fig:uv_hot} further demonstrates this, as again we see that the magnitude of the velocity in the
-vertical direction of the flame for the hottest run is also significantly larger.
-This motion therefore most likely corresponds to the vortical motion rather than to the propagation of the flame itself, something which is further supported by the fact that the velocity magnitude is two to three orders of magnitude greater than the flame speeds listed in Table~\ref{table:flame_speeds}.
+The $u$-$\rho$ phase plot in Figure~\ref{fig:urho_hot} shows that the horizontal velocity for the $T_{\rm star} = 3\times10^8$~K is sufficiently  \MarginPar{should this say "significantly"?} greater in magnitude than for the two cooler runs.
+The $u$-$v$ plot in Figure~\ref{fig:uv_hot} further demonstrates this, as again we see that the magnitude of the velocity in the vertical direction of the flame for the hottest run is also significantly larger.
+This motion therefore most likely corresponds to the vortical motion rather than to the propagation of the flame itself, something which is further supported by the fact that the velocity magnitude is 2-3 orders of magnitude greater than the flame speeds listed in Table~\ref{table:flame_speeds}.
 Also of interest in these plots is that the coolest run does not seem to have developed the characteristic vortex structure at the burning front (i.e.
@@
     \plotone{urho_plot_hotter}
-    \caption{\label{fig:urho_hot}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$ {\color{blue} (the latest time that $\tt{F500\_4E8}$ is run out to).
-The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.}
+    \caption{\label{fig:urho_hot}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ at $t = 40~\mathrm{ms}$.}
 \end{figure}
@@
     \plotone{uv_plot_hotter}
-    \caption{\label{fig:uv_hot}Phase plot of the horizontal $x$-velocity $u$, vertical $y$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$.
-{\color{blue} The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.}
+    \caption{\label{fig:uv_hot}Phase plot of the horizontal $x$-velocity $u$, vertical $y$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ at $t = 40~\mathrm{ms}$.}
 \end{figure}
@@
     \plotone{abar_plot_hotter}
-    \caption{\label{fig:abar_hot}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$.
-{\color{blue} The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.}
+    \caption{\label{fig:abar_hot}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ at $t = 40~\mathrm{ms}$.}
 \end{figure}

### 2021-01-21 xrb2/paper.tex

--- 2021-01-21 xrb2/paper.tex
@@
 	\tablecaption{\label{table:flame_speeds} Flame speed and acceleration values measured for each 
-	simulation. {\color{blue}$v_0$} is the flame velocity {\color{blue}in the $+r$ direction} at $t 
+	simulation. $v_0$ is the flame velocity in the $+r$ direction at $t 
 	= 0~\mathrm{ms}$. The initial flame 
@@
 	\startdata
-  {\tt F1000}     & $(3.414 \pm 0.008)$ & $(-1.03 \pm 0.13)$ & $(3.36 \pm 0.01)$\\
-  {\tt F500}      & $(3.077 \pm 0.013)$ & $(5.41 \pm 0.19)$ & $(3.35 \pm 0.02)$\\
-  {\tt F500\_2E8} & $(3.760 \pm 0.015)$ & $(1.74 \pm 0.25)$ & $(3.85 \pm 0.02)$\\
-  {\tt F500\_3E8} & $(-5.293 \pm 0.100)$ & $(357.41 \pm 2.15)$ & $(12.58 \pm 0.15)$\\
+  {\tt F1000}     & $3.414 \pm 0.008$ & $-1.03 \pm 0.13$ & $3.36 \pm 0.01$\\
+  {\tt F500}      & $3.077 \pm 0.013$ & $5.41 \pm 0.19$ & $3.35 \pm 0.02$\\
+  {\tt F500\_2E8} & $3.760 \pm 0.015$ & $1.74 \pm 0.25$ & $3.85 \pm 0.02$\\
+  {\tt F500\_3E8} & $-5.293 \pm 0.100$ & $357.41 \pm 2.15$ & $12.58 \pm 0.15$\\
 	\enddata

### 2021-01-21 xrb2/paper.tex
Unblue some things that have already been reviewed

@@
 We run three models ({\tt F250}, {\tt F500}, and {\tt F1000}) with the same initial model in terms 
-of temperature but differing rotation rates. {\color{blue}We saw in \citet{flame_wave1} that 
-increasing the 
-neutron star rotation rate reduces the horizontal lengthscale of the flame. An estimate of 
-this lengthscale is given by the Rossby radius of deformation, $L_R$. The Rossby radius may be 
-thought of as the scale over which the balance between the Coriolis force and horizontal pressure 
-gradient becomes important, and is approximately given by
+of temperature but differing rotation rates. We saw in \citet{flame_wave1} that 
+increasing the neutron star rotation rate reduces the horizontal lengthscale of the flame. An 
+estimate of this lengthscale is given by the Rossby radius of deformation, $L_R$. The Rossby radius 
+may be thought of as the scale over which the balance between the Coriolis force and horizontal 
+pressure gradient becomes important, and is approximately given by
 \begin{equation}
@@
 Taking the edge at greatest radius of the bright teal/green region (where significant energy generation is occurring) as the leading edge of the flame in each plot, we see that the horizontal extent of the $1000~\mathrm{Hz}$ flame ($\tt{F1000}$) appears to be
-reduced compared to the lower rotation $500~\mathrm{Hz}$ run ($\tt{F500}$). From Equation \ref{eqn:rossby}, we can see that increasing 
-the rotation rate from $500~\mathrm{Hz}$ to $1000~\mathrm{Hz}$ should decrease $L_R$ by a factor of two, and that the greater 
-confinement from the Coriolis force should reduce the horizontal extent of the flame by a similar 
-factor.
-However, the Rossby radius is only an approximate measure of this horizontal lengthscale, and in our simulations we see that this scaling does not work so well for all rotation rates.
-The simulations seem to follow the theoretical
-scaling described in Equation \ref{eqn:rossby} more closely at higher rotation rates ($1000~\mathrm{Hz}$ and 
-higher), based on the results of \citet{flame_wave1}.}
+reduced compared to the lower rotation $500~\mathrm{Hz}$ run ($\tt{F500}$). From Equation 
+\ref{eqn:rossby}, we can see that increasing the rotation rate from $500~\mathrm{Hz}$ to 
+$1000~\mathrm{Hz}$ should decrease $L_R$ by a factor of two, and that the greater confinement from 
+the Coriolis force should reduce the horizontal extent of the flame by a similar factor. However, 
+the Rossby radius is only an approximate measure of this horizontal lengthscale, and in our 
+simulations we see that this scaling does not work so well for all rotation rates. The simulations 
+seem to follow the theoretical scaling described in Equation \ref{eqn:rossby} more closely at 
+higher rotation rates ($1000~\mathrm{Hz}$ and higher), based on the results of \citet{flame_wave1}.
 
@@
 increasing the temperature and density in the flame front such that the reaction rate increases, 
-which allows the reactions to progress further.
-{\color{blue} The $\dot{e}_\mathrm{nuc}$ plots in Figure \ref{fig:time_series_enuc_1000} also support this interpretation, with the interior of the flame nearest to the crust\MarginPar{What is meant by interior here??
-The edge?
-The part nearest the crust?} in the {\tt F1000} run reaching higher $\dot{e}_\mathrm{nuc}$ values than for the {\tt F500} run in Figure \ref{fig:time_series_enuc_500}.} In contrast to {\tt F500} and {\tt F1000}, the
-lowest rotation run --- {\tt F250} --- failed to ignite. The lack of ignition for {\tt F250} also 
-aligns with the reasoning given above, with the lower rotation in this case potentially leading to 
-insufficient confinement such that the temperature and density required for ignition is not 
-achieved. In this scenario, another source of confinement (e.g.\ magnetic fields) would need to 
-take over at lower rotation rates to allow a burst to occur, at least for the initial model used 
-here.
+which allows the reactions to progress further. The $\dot{e}_\mathrm{nuc}$ plots in Figure 
+\ref{fig:time_series_enuc_1000} also support this interpretation, with the interior of the 
+flame nearest to the crust in the {\tt F1000} run reaching higher $\dot{e}_\mathrm{nuc}$ values 
+than for the {\tt F500} run in Figure \ref{fig:time_series_enuc_500}. In contrast to {\tt F500} 
+and {\tt F1000}, the lowest rotation run --- {\tt F250} --- failed to ignite. The lack of ignition 
+for {\tt F250} also aligns with the reasoning given above, with the lower rotation in this case 
+potentially leading to insufficient confinement such that the temperature and density required for 
+ignition is not achieved. In this scenario, another source of confinement (e.g.\ magnetic fields) 
+would need to take over at lower rotation rates to allow a burst to occur, at least for the initial 
+model used here.
 
@@
 	velocities and accelerations are derived from a quadratic least-squares fit to each of the 
-	datasets for times $t \gtrsim 20~\mathrm{ms}$.
-{\color{blue} Using these fit parameters we calculate the velocity at $t = 50~\mathrm{ms}$, $v_{50}$, at which point the flames have reached steady propagation.}}
+	datasets for times $t \gtrsim 20~\mathrm{ms}$. Using these fit parameters we calculate the 
+	velocity at $t = 50~\mathrm{ms}$, $v_{50}$, at which point the flames have reached steady 
+	propagation.}
 	\tablehead{\colhead{run} & \colhead{$v_0$ (km s$^{-1}$)} & \colhead{$a$ (km s$^{-2}$)} & \colhead{$v_{50}$ (km s$^{-1}$)}} 
@@
 
-Figure~\ref{fig:4e8_stacked_abar} shows $\bar{A}$ for $\tt{F500\_4E8}$.
-Again, burning extends across the domain and high into the atmosphere and fuel layer, lacking the characteristic flame structure shown in $\bar{A}$ plots for the lower temperature $500~\mathrm{Hz}$ runs (see Figure~\ref{fig:compare_500Hz_abar}).
-A distinct mass of material appears to have broken out of the atmosphere near the axis of symmetry.
-The atmosphere edge is located at $\sim 1.2\times 10^4~\mathrm{cm}$.
-{\color{blue}A similar effect is also visible (to a lesser extent) in the $\tt{F500\_3E8}$ plot in Figure \ref{fig:compare_500Hz_abar}, with a faint haze of material rising above the flame.
-However, this is likely to be a numerical artefact rather than a true physical effect, and a result of the boundary conditions imposed at the axis of symmetry for these simulations.} $\tt{F500\_4E8}$ clearly reaches much higher $\bar{A}$ values across the domain (especially at the latest snapshot, $t = 40~\mathrm{ms}$) compared to all the other runs described in this paper (see Figures \ref{fig:compare_500Hz_abar}, \ref{fig:time_series_500}, and \ref{fig:time_series_1000}).
+Figure~\ref{fig:4e8_stacked_abar} shows $\bar{A}$ for $\tt{F500\_4E8}$. Again, burning extends 
+across the domain and high into the atmosphere and fuel layer, lacking the characteristic flame 
+structure shown in $\bar{A}$ plots for the lower temperature $500~\mathrm{Hz}$ runs (see 
+Figure~\ref{fig:compare_500Hz_abar}). A distinct mass of material appears to have broken out of the 
+atmosphere near the axis of symmetry. The atmosphere edge is located at $\sim 1.2\times 
+10^4~\mathrm{cm}$. A similar effect is also visible (to a lesser extent) in the $\tt{F500\_3E8}$ 
+plot in Figure \ref{fig:compare_500Hz_abar}, with a faint haze of material rising above the flame. 
+However, this is likely to be a numerical artefact rather than a true physical effect, and a result 
+of the boundary conditions imposed at the axis of symmetry for these simulations. $\tt{F500\_4E8}$ 
+clearly reaches much higher $\bar{A}$ values across the domain (especially at the latest snapshot, 
+$t = 40~\mathrm{ms}$) compared to all the other runs described in this paper (see Figures 
+\ref{fig:compare_500Hz_abar}, \ref{fig:time_series_500}, and \ref{fig:time_series_1000}).
 
@@
 
-The method for measuring flame propagation described in Section \ref{ssec:rot_propagation} is applied in Figure \ref{fig:flame_speeds_2} to the three $500~\mathrm{Hz}$ runs with $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$.
-Due to the lack of a clear burning front in $\tt{F500\_4E8}$, we do not analyze its propagation speed and acceleration.
-As the initial $T_{\mathrm{star}}$ is increased, the flame becomes greatly accelerated.
-The initial flame velocities and accelerations derived from a quadratic least-squares fit to each of the datasets, {\color{blue} as well as the flame velocities at $t = 50~\mathrm{ms}$ calculated using the fit parameters,} are provided in Table~\ref{table:flame_speeds}.
-{\color{blue} Comparing the flame propagation at different initial temperatures, the most robust feature is the acceleration of the $\tt{F500\_3E8}$ run at $t \sim 40~\mathrm{ms}$.} The reason for the acceleration of the flame is not entirely clear.
-It could be that energy released from burning begins to dominate the flame's propagation as it evolves, increasing the flame speed over time.
-Another possibility is that the increased temperatures lead to enhanced turbulent mixing effects that pull in more fresh fuel for the flame to burn.
-Yet another possibility is that the higher initial $T_{\mathrm{star}}$ leads to a greater average temperature in the fuel layer over time, making it easier for the flame to burn the fuel and propagate.
+The method for measuring flame propagation described in Section \ref{ssec:rot_propagation} is 
+applied in Figure \ref{fig:flame_speeds_2} to the three $500~\mathrm{Hz}$ runs with 
+$T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$. Due to the lack of a clear burning front in 
+$\tt{F500\_4E8}$, we do not analyze its propagation speed and acceleration. As the initial 
+$T_{\mathrm{star}}$ is increased, the flame becomes greatly accelerated. The initial flame 
+velocities and accelerations derived from a quadratic least-squares fit to each of the datasets, as 
+well as the flame velocities at $t = 50~\mathrm{ms}$ calculated using the fit parameters, are 
+provided in Table~\ref{table:flame_speeds}. Comparing the flame propagation at different initial 
+temperatures, the most robust feature is the acceleration of the $\tt{F500\_3E8}$ run at $t \sim 
+40~\mathrm{ms}$. The reason for the acceleration of the flame is not entirely clear. It could be 
+that energy released from burning begins to dominate the flame's propagation as it evolves, 
+increasing the flame speed over time. Another possibility is that the increased temperatures lead 
+to enhanced turbulent mixing effects that pull in more fresh fuel for the flame to burn. Yet 
+another possibility is that the higher initial $T_{\mathrm{star}}$ leads to a greater average 
+temperature in the fuel layer over time, making it easier for the flame to burn the fuel and 
+propagate.
 

### 2021-01-21 xrb2/paper.tex
Insert figure reference

@@
 
-In contrast to the models with $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$, $\tt{F500\_4E8}$ is so hot that the organized flame structure is lost.
-The fuel layer undergoes stable burning and essentially simmers in place, rather than forming a flame.
-This process has been observed in neutron stars with accretion rates that exceed the Eddington limit \citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}; in this model we have essentially recreated the conditions found in such stars with high accretion rates.
-This model burns so strongly that it is only run out to $40~\mathrm{ms}$.
-After an initial period when the burning moves across the domain, residual burning continues and eventually ignites the entire fuel layer at late times, as shown in Figures \ref{fig:4e8_stacked_enuc} and \ref{fig:4e8_stacked_abar} for three snapshots taken in the last ten seconds of the simulation.
-In Figure~\ref{fig:4e8_stacked_enuc}, $\enucdot$ reaches values of $10^{18} - 10^{20}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ across the domain and at heights up to $\sim 0.5\times 10^4~\mathrm{cm}$.
-There is still significant burning occurring even higher, with $\enucdot$ reaching $\sim 10^{15}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ at heights up to $\sim 0.8\times 10^4~\mathrm{cm}$.
-For comparison, the next hottest run ($\tt{F500\_3E8}$) only reaches maximum $\enucdot$ values on the order of $10^{18}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ {\color{blue} (see Figure ??
-Kiran $\enucdot$ plot)}, even at the latest timesteps ($\sim 70~\mathrm{ms}$, when the flame is most developed).
-{\color{blue} Significant $\enucdot$ values for all other runs are confined to the flame front, rather than spanning the entire domain}.
+In contrast to the models with $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$, $\tt{F500\_4E8}$ 
+is so hot that the organized flame structure is lost. The fuel layer undergoes stable burning and 
+essentially simmers in place, rather than forming a flame. This process has been observed in 
+neutron stars with accretion rates that exceed the Eddington limit 
+\citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}; in this model we have essentially 
+recreated the conditions found in such stars with high accretion rates. This model burns so 
+strongly that it is only run out to $40~\mathrm{ms}$. After an initial period when the burning 
+moves across the domain, residual burning continues and eventually ignites the entire fuel layer at 
+late times, as shown in Figures \ref{fig:4e8_stacked_enuc} and \ref{fig:4e8_stacked_abar} for three 
+snapshots taken in the last ten seconds of the simulation. In Figure~\ref{fig:4e8_stacked_enuc}, 
+$\enucdot$ reaches values of $10^{18} - 10^{20}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ 
+across the domain and at heights up to $\sim 0.5\times 10^4~\mathrm{cm}$. There is still 
+significant burning occurring even higher, with $\enucdot$ reaching $\sim 
+10^{15}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ at heights up to $\sim 0.8\times 
+10^4~\mathrm{cm}$. For comparison, the next hottest run ($\tt{F500\_3E8}$) only reaches maximum 
+$\enucdot$ values on the order of $10^{18}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ 
+{\color{blue} (see Figure \ref{fig:maxima})}, even at the latest timesteps ($\sim 70~\mathrm{ms}$, 
+when the flame is most developed). {\color{blue} Significant $\enucdot$ values for all other runs 
+are confined to the flame front, rather than spanning the entire domain}.
 

### 2021-01-21 xrb2/paper.tex
modified section 3.3 paragraphs to reference maxima plots

--- 2021-01-21 xrb2/paper.tex
@@
 
-Figure~\ref{fig:compare_500Hz_abar} shows $\bar{A}$ for three $500~\mathrm{Hz}$ simulations with different initial temperature structures at $t = 70~\mathrm{ms}$.
-The hottest of the three plotted models, $\tt{F500\_3E8}$, has a faster propagating flame (this will be discussed further in Section \ref{ssec:temp_prop}).
+Figure~\ref{fig:compare_500Hz_abar} shows $\bar{A}$ for three $500~\mathrm{Hz}$ simulations with different initial temperature structures ($T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$) at $t = 70~\mathrm{ms}$.
+$\tt{F500\_3E8}$ {\color{blue} (Figure~\ref{fig:compare_500Hz_abar}, bottom panel) --- the hottest run to form a clear burning front ---} has a faster propagating flame (this will be discussed further in Section \ref{ssec:temp_prop}).
 {\color{blue} It also reaches slightly higher $\bar{A}$ values than the two cooler runs.
@@
 $\tt{F500\_3E8}$ also reaches higher $\enucdot$ values at the low end of the temperature range ($< 0.5 \times 10^9 \mathrm{K}$).
-There appear to be more causally connected regions across a range of $\bar{A}$ at low temperatures for $\tt{F500\_3E8}$ than for $\tt{F500}$, suggesting that the higher $\enucdot$ for $\tt{F500\_3E8}$ generates burning in certain reaction pathways that are not present in the cooler $\tt{F500}$ run.}
-\MarginPar{relate this section back to Kiran's max temperature and $\enucdot$ plots once they're added, note that $\tt{F500\_3E8}$ doesn't reach as high temperatures, less evolved abar than $\tt{F1000}$}
+There appear to be more causally connected regions across a range of $\bar{A}$ at low temperatures for $\tt{F500\_3E8}$ than for $\tt{F500}$, suggesting that the higher $\enucdot$ for $\tt{F500\_3E8}$ generates burning in certain reaction pathways that are not present in the cooler $\tt{F500}$ run.
+Note that, although $\tt{F500\_3E8}$ is the hottest run with a modified initial temperature configuration to form a distinct flame front, the highest rotation $\tt{F1000}$ flame actually reaches higher temperatures (see Figure~\ref{fig:maxima}, left panel) as well as higher $\bar{A}$ (see Figure~\ref{fig:time_series_1000}).}
 
@@
 There is still significant burning occurring even higher, with $\enucdot$ reaching $\sim 10^{15}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ at heights up to $\sim 0.8\times 10^4~\mathrm{cm}$.
-For comparison, the next hottest run ($\tt{F500\_3E8}$) only reaches maximum $\enucdot$ values on the order of $10^{18}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ {\color{blue} (see Figure ??
-Kiran $\enucdot$ plot)}, even at the latest timesteps ($\sim 70~\mathrm{ms}$, when the flame is most developed).
+For comparison, the next hottest run ($\tt{F500\_3E8}$) only reaches maximum $\enucdot$ values on the order of $10^{18}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ {\color{blue} (see Figure~\ref{fig:maxima}, right panel)}, even at the latest timesteps ($\sim 70~\mathrm{ms}$, when the flame is most developed).
 {\color{blue} Significant $\enucdot$ values for all other runs are confined to the flame front, rather than spanning the entire domain}.

### 2021-01-21 xrb2/paper.tex
Added streamline plot

--- 2021-01-21 xrb2/paper.tex
@@
 point. Figure \ref{fig:maxima} shows $\repavg{T}$ and $\repavg{\enucdot}$ as functions of time for 
-the subset of our runs that achieve a propagating flame. This figure is referenced at several 
-points throughout the subsequent sections.
+the subset of our runs that achieve a propagating flame. This figure is referenced throughout the 
+subsequent sections.
 }
+
+\begin{figure}
+	\plotone{streamlines_500Hz_temp.pdf}
+	\caption{\label{fig:streamlines}Streamlines showing the velocity field at time $t = 
+	100~\mathrm{ms}$ for the {\tt F500} run. Streamline segments are colored by transverse velocity 
+	(yellower implies faster flow), and segments with transverse velocity $v < 2 \times 
+	10^6~\mathrm{cm/s}$ are scrubbed out. Note that the plot has been stretched vertically to 
+	better show the streamlines -- the aspect ratio is not the aspect ratio of the simulation 
+	domain.}
+\end{figure}
+
+{\color{blue} Figure \ref{fig:streamlines} shows the velocity field. Not sure where we want to put 
+this.}
 

### 2021-01-21 xrb2/paper.tex

--- 2021-01-21 xrb2/paper.tex
@@
 
-In contrast to the models with $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$, $\tt{F500\_4E8}$ is so hot that the organized flame structure is lost.
-The fuel layer undergoes stable burning and essentially simmers in place, rather than forming a flame.
-This process has been observed in neutron stars with accretion rates that exceed the Eddington limit \citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}; in this model we have essentially recreated the conditions found in such stars with high accretion rates.
-This model burns so strongly that it is only run out to $40~\mathrm{ms}$.
-After an initial period when the burning moves across the domain, residual burning continues and eventually ignites the entire fuel layer at late times, as shown in Figures \ref{fig:4e8_stacked_enuc} and \ref{fig:4e8_stacked_abar} for three snapshots taken in the last ten seconds of the simulation.
-In Figure~\ref{fig:4e8_stacked_enuc}, $\enucdot$ reaches values of $10^{18} - 10^{20}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ across the domain and at heights up to $\sim 0.5\times 10^4~\mathrm{cm}$.
-There is still significant burning occurring even higher, with $\enucdot$ reaching $\sim 10^{15}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ at heights up to $\sim 0.8\times 10^4~\mathrm{cm}$.
-For comparison, the next hottest run ($\tt{F500\_3E8}$) only reaches maximum $\enucdot$ values on the order of $10^{18}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ {\color{blue} (see Figure~\ref{fig:maxima}, right panel)}, even at the latest timesteps ($\sim 70~\mathrm{ms}$, when the flame is most developed).
-{\color{blue} Significant $\enucdot$ values for all other runs are confined to the flame front, rather than spanning the entire domain}.
+In contrast to the models with $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$, $\tt{F500\_4E8}$ 
+is so hot that the organized flame structure is lost. The fuel layer undergoes stable burning and 
+essentially simmers in place, rather than forming a flame. This process has been observed in 
+neutron stars with accretion rates that exceed the Eddington limit 
+\citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}; in this model we have essentially 
+recreated the conditions found in such stars with high accretion rates. This model burns so 
+strongly that it is only run out to $40~\mathrm{ms}$. After an initial period when the burning 
+moves across the domain, residual burning continues and eventually ignites the entire fuel layer at 
+late times, as shown in Figures \ref{fig:4e8_stacked_enuc} and \ref{fig:4e8_stacked_abar} for three 
+snapshots taken in the last ten seconds of the simulation. In Figure~\ref{fig:4e8_stacked_enuc}, 
+$\enucdot$ reaches values of $10^{18} - 10^{20}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ 
+across the domain and at heights up to $\sim 0.5\times 10^4~\mathrm{cm}$. There is still 
+significant burning occurring even higher, with $\enucdot$ reaching $\sim 
+10^{15}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ at heights up to $\sim 0.8\times 
+10^4~\mathrm{cm}$. For comparison, the next hottest run ($\tt{F500\_3E8}$) only reaches maximum 
+$\enucdot$ values on the order of $10^{18}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ 
+{\color{blue} (see Figure~\ref{fig:maxima}, right panel)}, even at the latest timesteps ($\sim 
+70~\mathrm{ms}$, when the flame is most developed). {\color{blue} Significant $\enucdot$ values for 
+all other runs are confined to the flame front, rather than spanning the entire domain}.
 

### 2021-01-21 xrb2/paper.tex
Resolving some of Yuri's edits

@@
 \author[0000-0002-6447-3603]{Y.\ Cavecchi}
-\affiliation{Mathematical Sciences and STAG Research Centre, University of Southampton, SO17 1BJ}
+\affiliation{Universidad Nacional Aut\'{o}noma M\'{e}xico, Instituto de Astronom\'{i}a, Ciudad Universitaria, CDMX 04510, Mexico}
 
@@
 For the high temperature unboosted flames, however, we found that the $\enucdot$ profiles failed to 
-reach that small a value across the domain at most times, which prevented us from getting reliable 
+reach that small a value across the domain at most times, which prevented us from obtaining reliable 
 position measurements. We therefore use $1\%$ of the global $\enucdot$ maximum in this paper
@@
 Figure~\ref{fig:compare_500Hz_abar} shows $\bar{A}$ for three $500~\mathrm{Hz}$ simulations with different initial temperature structures ($T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$) at $t = 70~\mathrm{ms}$.
-$\tt{F500\_3E8}$ {\color{blue} (Figure~\ref{fig:compare_500Hz_abar}, bottom panel) --- the hottest run to form a clear burning front ---} has a faster propagating flame (this will be discussed further in Section \ref{ssec:temp_prop}).
+{\color{blue}We do not plot $\tt{F500\_4E8}$ here because it fails to form a clear burning front.
+$\tt{F500\_3E8}$ (Figure~\ref{fig:compare_500Hz_abar}, bottom panel) --- the hottest run to form a clear burning front ---} has a faster propagating flame (this will be discussed further in Section \ref{ssec:temp_prop}).
 {\color{blue} It also reaches slightly higher $\bar{A}$ values than the two cooler runs.
-We do not plot $\tt{F500\_4E8}$ here because it fails to form a clear burning front.
 The Figure~\ref{fig:abar_hot} $\bar{A}$-$T$ phase plots of $\tt{F500}$ (left) and $\tt{F500\_3E8}$ (middle) also reflect these $\bar{A}$ features, with $\tt{F500\_3E8}$ reaching slightly higher $\bar{A}$ values.
@@
 As can be seen in Figure~\ref{fig:uv} (which was plotted at $100~\mathrm{ms}$), this does develop more at later times.
-Similar to what saw when comparing the runs with different rotation rates, it would therefore appear that it takes longer for the flame to develop when $T_{\rm star}$ is cooler.
+Similar to what we saw when comparing the runs with different rotation rates, it would therefore appear that it takes longer for the flame to develop when $T_{\rm star}$ is cooler.
 
@@
 
-We ran a number of simulations of laterally propagating flames in XRBs in order to explore the effects and rotation and thermal structure.
+We ran a number of simulations of laterally propagating flames in XRBs in order to explore the effects of rotation and thermal structure.
 We found that increasing the rotation rate increased the energy generation rate within the flame, enhancing its propagation \MarginPar{I thought the increased rotation rate enhanced nuclear burning, but suppressed flame propagation?}.
@@
 We found that increasing the temperature of the crust significantly enhanced the flame propagation.
-This we believe to be because a cooler crust allows heat to more efficiently be transferred from the flame itself, therefore reducing the flame's temperature, slowing burning and consequently the reducing its propagation speed.
+This we believe to be because a cooler crust allows heat to more efficiently be transferred from the flame itself, therefore reducing the flame's temperature, slowing burning and consequently reducing its propagation speed.
 We saw that if the crust temperature was too high, then instead of a flame, the atmosphere would burn stably.

### 2021-01-21 xrb2/paper.tex
add citation

@@
 potentially leading to insufficient confinement such that the temperature and density required for 
-ignition is not achieved. In this scenario, another source of confinement (e.g.\ magnetic fields) 
+ignition is not achieved. In this scenario, another source of confinement (e.g.\ magnetic fields, see \citep{art-2016-cavecchi-etal}) 
 would need to take over at lower rotation rates to allow a burst to occur, at least for the initial 

### 2021-01-22 xrb2/paper.tex
Finished Yuri's edits

@@
 \begin{abstract}
-We continue to investigate laterally propagating flames in XRBs using
+We continue to investigate laterally propagating flames in type I X-ray bursts using
 fully compressible hydrodynamics simulations.  In the current study we
@@
 
-Considerable evidence suggests that ignition in an X-ray burst starts
+Considerable evidence suggests that ignition in an X-ray burst (XRB) starts
 in a localized region and then spreads across the surface of the
@@
 We investigate the effect of rotation rate on the flame. With the
-exception of \citet{altamirano2010discovery}, most observations of
+exception of IGR J17480-2446 (\citet{altamirano2010discovery}, spinning at $11~\mathrm{Hz}$), most observations of
 XRBs come from sources with rotation rates of 200-$600~\mathrm{Hz}$
@@
 origin) and a cool model on the right.  The initial temperature
-gradient between the hot and cool models drives a laterally
+gradient between the hot and cool fluids drives a laterally
 propagating flame through the cool fuel.  In our original set of
@@
 higher rotation rates ($1000~\mathrm{Hz}$ and higher), based on the results of \citet{flame_wave1}.
+{\color{blue} Perhaps these new un-boosted runs have not had sufficient time to reach steady propagation, in contrast to the boosted runs in \citet{flame_wave1} that reached steady propagation more quickly.}
 
@@
 which allows the reactions to progress further. The $\dot{e}_\mathrm{nuc}$ plots in Figure 
-\ref{fig:time_series_enuc_1000} also support this interpretation, with the interior of the 
-flame nearest to the crust in the {\tt F1000} run reaching higher $\dot{e}_\mathrm{nuc}$ values 
+\ref{fig:time_series_enuc_1000} also support this interpretation, with the region of the 
+flame front nearest to the crust in the {\tt F1000} run reaching higher $\dot{e}_\mathrm{nuc}$ values 
 than for the {\tt F500} run in Figure \ref{fig:time_series_enuc_500}. In contrast to {\tt F500} 
@@
 model used here.
+{\color{blue} Given that the size of our domain is $\sim 1~L_R$ for {\tt F250} (using Equation~\ref{eqn:rossby}), however, it is also possible that we simply cannot confine the flame sufficiently with our current domain width.}
 
 Burning in the {\tt F500} and {\tt F1000} runs is concentrated in a dense region with circular 
-motion. Figure \ref{fig:urho} compares the horizontal $x$-velocity $u$, density $\rho$ and the 
+motion. Figure \ref{fig:urho} compares the horizontal $r$-velocity $u$, density $\rho$ and the 
 nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the {\tt F250}, {\tt F500}, and {\tt F1000} 
@@
 \times 10^5 {\rm g\,cm^{-3}}$. The fluid in this dense, high energy generation region undergoes 
-vortical motion, shown in the Figure \ref{fig:uv} phase plots comparing $u$, $v$ and $\dot{e}_{\rm 
+vortical motion, shown in the Figure \ref{fig:uv} phase plots comparing $u$, vertical z-velocity $v$ and $\dot{e}_{\rm 
 nuc}$. From the previous plots (see: streamlines/baroclinicity?) \MarginPar{NF: need to add 
@@
     \plotone{urho_plot_0_1}{}
-    \caption{\label{fig:urho}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and $1000~\mathrm{Hz}$ runs ({\tt F250}, {\tt F500} and {\tt F1000}) at $t = 100~\mathrm{ms}$.
+    \caption{\label{fig:urho}Phase plot of the horizontal $r$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and $1000~\mathrm{Hz}$ runs ({\tt F250}, {\tt F500} and {\tt F1000}) at $t = 100~\mathrm{ms}$.
 The slightly lower $\dot{e}_{\rm nuc}$ values along the $u = 0$ axis are most likely a numerical artifact related to the density gradient setup and finite resolution of these simulations.}
@@
     \plotone{uv_plot_0_1}
-    \caption{\label{fig:uv}Phase plot of the horizontal $x$-velocity $u$, vertical $y$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and $1000~\mathrm{Hz}$ runs ({\tt F250}, {\tt F500} and {\tt F1000}) at $t = 100~\mathrm{ms}$.}
+    \caption{\label{fig:uv}Phase plot of the horizontal $r$-velocity $u$, vertical $z$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the 250, 500 and $1000~\mathrm{Hz}$ runs ({\tt F250}, {\tt F500} and {\tt F1000}) at $t = 100~\mathrm{ms}$.}
 \end{figure}
@@
 fraction of the global $\enucdot$ maximum across all of these profiles. Since the flames in our 
-simulations propagate in the positive radial direction, we then search the region of each profile 
+simulations propagate in the positive horizontal direction, we then search the region of each profile 
 at greater radius than the local $\enucdot$ maximum for the point where the $\enucdot$ first drops 
@@
 
-The $u$-$\rho$ phase plot in Figure~\ref{fig:urho_hot} shows that the horizontal velocity for the $T_{\rm star} = 4\times10^8$~K is significantly greater in magnitude than for the two cooler runs.
-The $u$-$v$ plot in Figure~\ref{fig:uv_hot} further demonstrates this, as again we see that the magnitude of the velocity in the
-vertical direction of the flame for the hottest run is also significantly larger.
+The $u$-$\rho$ phase plot in Figure~\ref{fig:urho_hot} shows that the horizontal velocity for the $T_{\rm star} = 4\times10^8$~K {\color{blue} ($\tt{F500\_4e8}$; right panel)} is significantly greater in magnitude than for the cooler {\color{blue}$\tt{F500\_3e8}$ (middle panel) and $\tt{F500}$ (left panel)} runs.
+The $u$-$v$ plot in Figure~\ref{fig:uv_hot} further demonstrates this, as again we see that the magnitude of the {\color{blue}$\tt{F500\_4e8}$ flame's vertical $z$-velocity $v$} is also significantly larger.
 This motion therefore most likely corresponds to the vortical motion rather than to the propagation of the flame itself, something which is further supported by the fact that the velocity magnitude is two to three orders of magnitude greater than the flame speeds listed in Table~\ref{table:flame_speeds}.
@@
     \plotone{urho_plot_hotter}
-    \caption{\label{fig:urho_hot}Phase plot of the horizontal $x$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$ {\color{blue} (the latest time that $\tt{F500\_4E8}$ is run out to).
+    \caption{\label{fig:urho_hot}Phase plot of the horizontal $r$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$ {\color{blue} (the latest time that $\tt{F500\_4E8}$ is run out to).
 The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.}
@@
     \plotone{uv_plot_hotter}
-    \caption{\label{fig:uv_hot}Phase plot of the horizontal $x$-velocity $u$, vertical $y$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$.
+    \caption{\label{fig:uv_hot}Phase plot of the horizontal $r$-velocity $u$, vertical $z$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$.
 {\color{blue} The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.}
@@
 We ran a number of simulations of laterally propagating flames in XRBs in order to explore the effects of rotation and thermal structure.
-We found that increasing the rotation rate increased the energy generation rate within the flame, enhancing its propagation \MarginPar{I thought the increased rotation rate enhanced nuclear burning, but suppressed flame propagation?}.
-This is likely due to greater confinement of the flame from the increased Coriolis force.
-We explored several models with different crustal temperatures, to determine what effects mechanisms such as higher accretion rates, crustal heating and shallow heating may have on flame propagation.
+We found that increasing the rotation rate increased the energy generation rate within the flame and enhanced nuclear burning.
+{\color{blue} Apart from the lowest rotation run (which failed to ignite), flame propagation was not noticeably impacted by rotation rate; by the time the different flame fronts reached steady propagation, they shared comparable velocities.
+These results are likely due to the rotation-dependent strength of the Coriolis force and confinement of the flame.} \MarginPar{NF: is the wording of the blue text correct?} We explored several models with different crustal temperatures, to determine what effects mechanisms such as higher accretion rates, crustal heating and shallow heating may have on flame propagation.
 We found that increasing the temperature of the crust significantly enhanced the flame propagation.

### 2021-01-22 xrb2/paper.tex
Update notation for average in paper body
started streamlines discussion

@@
 \newcommand{\avgtwod}[1]{\langle~#1 \rangle}
-\newcommand{\repavg}[1]{\langle \langle #1 \rangle \rangle}
+\newcommand{\repavg}[1]{\langle #1 \rangle_w}
 \newcommand{\rms}[2]{\left(\delta#1\right)_{r_{#2}}}
@@
 {\color{blue}
-For some of our analysis, we would like to have a means of estimating the temperature and nuclear 
-energy generation rate in the burning region of each simulation. For this purpose, we define 
+For some of our analysis, we would like to have a means of estimating the temperature ($T$) and nuclear 
+energy generation rate ($\enucdot$) in the burning region of each simulation. For this purpose, we define 
 $\repavg{Q}$ to be the mass weighted average
@@
 }
-
-\begin{figure}
-	\plotone{streamlines_500Hz_temp.pdf}
-	\caption{\label{fig:streamlines}Streamlines showing the velocity field at time $t = 
-	100~\mathrm{ms}$ for the {\tt F500} run. Streamline segments are colored by transverse velocity 
-	(yellower implies faster flow), and segments with transverse velocity $v < 2 \times 
-	10^6~\mathrm{cm/s}$ are scrubbed out. Note that the plot has been stretched vertically to 
-	better show the streamlines -- the aspect ratio is not the aspect ratio of the simulation 
-	domain.}
-\end{figure}
-
-{\color{blue} Figure \ref{fig:streamlines} shows the velocity field. Not sure where we want to put 
-this.}
 
@@
 neutron star rotation rate. In Figure \ref{fig:time_series_enuc_500} and Figure 
-\ref{fig:time_series_enuc_1000}, we use the energy generation rate $\dot{e}_\mathrm{nuc}$ measured at $50~\mathrm{ms}$ and $100~\mathrm{ms}$ to discern the horizontal extent of the flame at different rotation rates.
+\ref{fig:time_series_enuc_1000}, we use the $\dot{e}_\mathrm{nuc}$ measured at $50~\mathrm{ms}$ and $100~\mathrm{ms}$ to discern the horizontal extent of the flame at different rotation rates.
 Taking the edge at greatest radius of the bright teal/green region (where significant energy generation is occurring) as the leading edge of the flame in each plot, we see that the horizontal extent of the $1000~\mathrm{Hz}$ flame ($\tt{F1000}$) appears to be
@@
 higher rotation rates ($1000~\mathrm{Hz}$ and higher), based on the results of \citet{flame_wave1}.
-{\color{blue} Perhaps these new un-boosted runs have not had sufficient time to reach steady propagation, in contrast to the boosted runs in \citet{flame_wave1} that reached steady propagation more quickly.}
+{\color{blue} Perhaps these new un-boosted runs have not had sufficient time to reach steady propagation, in contrast to the boosted runs in \citet{flame_wave1} that reached steady propagation more quickly.} \MarginPar{does the equation for the Rossby radius only hold for flames with steady propagation?}
 
@@
 model used here.
-{\color{blue} Given that the size of our domain is $\sim 1~L_R$ for {\tt F250} (using Equation~\ref{eqn:rossby}), however, it is also possible that we simply cannot confine the flame sufficiently with our current domain width.}
+{\color{blue} Given that the size of our domain is $\sim 1~L_R$ for {\tt F250} (using Equation~\ref{eqn:rossby}), it is also possible that we simply cannot confine the flame sufficiently with our current domain width.}
 
-Burning in the {\tt F500} and {\tt F1000} runs is concentrated in a dense region with circular 
-motion. Figure \ref{fig:urho} compares the horizontal $r$-velocity $u$, density $\rho$ and the 
+\MarginPar{NF: Each of the figures referenced in this paragraph contributes to the idea that we are seeing burning/energy generation from fluid undergoing vortical motion/fuel being entrained around the flame edge...
+I just think the logical flow of the sentences could be a bit better.}
+Burning in the {\tt F500} and {\tt F1000} runs is concentrated in a hot, dense region with circular 
+motion.
+{\color{blue} We visualize the velocity field in the burning region for the {\tt F500} run in Figure~\ref{fig:streamlines}, colored by temperature.
+Velocities near the bottom edge of the high temperature/high density region (white) tend to have a significant $+z$ component.
+This boundary most likely corresponds to the leading edge of the flame, where colder fuel undergoes vortical motion and becomes entrained into the burning front.} In Figure~\ref{fig:urho}, which compares the horizontal $r$-velocity $u$, density $\rho$ and the
 nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the {\tt F250}, {\tt F500}, and {\tt F1000} 
-runs. Most of the burning for each of the simulations occurs in a high density region $\rho > 3 
+runs, most of the burning for each of the simulations occurs in a high density region $\rho > 3 
 \times 10^5 {\rm g\,cm^{-3}}$. The fluid in this dense, high energy generation region undergoes 
-vortical motion, shown in the Figure \ref{fig:uv} phase plots comparing $u$, vertical z-velocity $v$ and $\dot{e}_{\rm 
-nuc}$. From the previous plots (see: streamlines/baroclinicity?) \MarginPar{NF: need to add 
-streamline plots?}, this most likely corresponds to the leading edge of the flame where fresh fuel 
-is being entrained. This feature is not developed in the $250~\mathrm{Hz}$ flame; it could potentially develop 
+vortical motion, shown in the Figure~\ref{fig:uv} phase plots comparing $u$, vertical z-velocity $v$ and $\dot{e}_{\rm 
+nuc}$. {\color{blue} This agrees with the velocity streamlines shown in Figure~\ref{fig:streamlines}.}
+This feature is not developed in the $250~\mathrm{Hz}$ flame in Figure~\ref{fig:uv} {\color{blue} (left panel)}; it could potentially develop 
 at later times (past the point at which we terminated our simulation), or the flame could just 
 fizzle out and fail to ignite entirely.
+
+\begin{figure}
+  \plotone{streamlines_500Hz_temp.pdf}
+  \caption{\label{fig:streamlines}Streamlines showing the velocity field at time $t = 
+  100~\mathrm{ms}$ for the {\tt F500} run. Streamline segments are colored by transverse velocity 
+  (yellower implies faster flow), and segments with transverse velocity $v < 2 \times 
+  10^6~\mathrm{cm/s}$ are scrubbed out. Note that the plot has been stretched vertically to 
+  better show the streamlines -- the aspect ratio is not the aspect ratio of the simulation 
+  domain.}
+\end{figure}
 

### 2021-01-22 xrb2/paper.tex
Fixed Yuri's institution
Removed streamlines discussion

@@
 \author[0000-0002-6447-3603]{Y.\ Cavecchi}
-\affiliation{Universidad Nacional Aut\'{o}noma M\'{e}xico, Instituto de Astronom\'{i}a, Ciudad Universitaria, CDMX 04510, Mexico}
+\affiliation{Universidad Nacional Aut\'{o}noma de M\'{e}xico, Instituto de Astronom\'{i}a, Ciudad Universitaria, CDMX 04510, Mexico}
 
@@
 
-\MarginPar{NF: Each of the figures referenced in this paragraph contributes to the idea that we are seeing burning/energy generation from fluid undergoing vortical motion/fuel being entrained around the flame edge...
-I just think the logical flow of the sentences could be a bit better.}
-Burning in the {\tt F500} and {\tt F1000} runs is concentrated in a hot, dense region with circular 
-motion.
-{\color{blue} We visualize the velocity field in the burning region for the {\tt F500} run in Figure~\ref{fig:streamlines}, colored by temperature.
-Velocities near the bottom edge of the high temperature/high density region (white) tend to have a significant $+z$ component.
-This boundary most likely corresponds to the leading edge of the flame, where colder fuel undergoes vortical motion and becomes entrained into the burning front.} In Figure~\ref{fig:urho}, which compares the horizontal $r$-velocity $u$, density $\rho$ and the
+Burning in the {\tt F500} and {\tt F1000} runs is concentrated in a dense region with circular 
+motion. In Figure~\ref{fig:urho}, which compares the horizontal $r$-velocity $u$, density $\rho$ and the 
 nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the {\tt F250}, {\tt F500}, and {\tt F1000} 
@@
 vortical motion, shown in the Figure~\ref{fig:uv} phase plots comparing $u$, vertical z-velocity $v$ and $\dot{e}_{\rm 
-nuc}$. {\color{blue} This agrees with the velocity streamlines shown in Figure~\ref{fig:streamlines}.}
+nuc}$. This most likely corresponds to the leading edge of the flame where fresh fuel is being entrained. \MarginPar{not sure if we should keep this sentence about entrainment...}
 This feature is not developed in the $250~\mathrm{Hz}$ flame in Figure~\ref{fig:uv} {\color{blue} (left panel)}; it could potentially develop 
@@
 fizzle out and fail to ignite entirely.
-
-\begin{figure}
-  \plotone{streamlines_500Hz_temp.pdf}
-  \caption{\label{fig:streamlines}Streamlines showing the velocity field at time $t = 
-  100~\mathrm{ms}$ for the {\tt F500} run. Streamline segments are colored by transverse velocity 
-  (yellower implies faster flow), and segments with transverse velocity $v < 2 \times 
-  10^6~\mathrm{cm/s}$ are scrubbed out. Note that the plot has been stretched vertically to 
-  better show the streamlines -- the aspect ratio is not the aspect ratio of the simulation 
-  domain.}
-\end{figure}
 

### 2021-01-25 xrb2/paper.tex
more small edits

@@
 rotation rate and the thermal structure of the neutron star on the
-flame.  We find that higher temperature atmospheres lead to faster
+flame.  We find that at lower rotation rates the flame becomes harder to ignite,
+whereas at higher rotation rates the nuclear burning is enhanced and the flame propagates steadily.
+We also find that higher temperature atmospheres lead to faster
 flames and observe acceleration of the flame front as it propagates
@@
 exception of IGR J17480-2446 (\citet{altamirano2010discovery}, spinning at $11~\mathrm{Hz}$), most observations of
-XRBs come from sources with rotation rates of 200-$600~\mathrm{Hz}$
-\citep{bilous:2019,galloway:2020} \MarginPar{check whether this is
-true for XRBs or just burst oscillations}. There are a number of
+XRBs come from sources with rotation rates of $200-600~\mathrm{Hz}$
+\citep{bilous:2019,galloway:2020}. There are a number of
 factors that could explain this lack of observations below $200~\mathrm{Hz}$. It
@@
 at lower rotation rates, the flame itself becomes harder to ignite. At higher 
-rotation rates, nuclear burning is enhanced and the flame propagates more quickly. 
+rotation rates, nuclear burning is enhanced and the flame propagates steadily. 
 At higher crustal temperatures, burning is greatly enhanced and the flame accelerates as 
@@
 propagating flame through the cool fuel.  In our original set of
-calculations, we artificially boosted the flame speed by adjusting the
+calculations (in in \citet{flame_wave1}), in order to make the 
+simulations computationally feasible we artificially boosted the flame speed by adjusting the
 conductivity and reaction rate to produce a flame moving 5--10$\times$
@@
 force balances the lateral flame spreading in order to reduce the size 
-of the simulation domain.  For the current
+of the simulation domain. The port of \castro\ to GPUs~\citep{sc20_gpu} 
+significantly improved its overall performance, enabling
+us to run these new simulations without the previous approximations while
+continuing to resolve the burning front. For these
 simulations, we no longer boost the flame speed---the true
 conductivities (taken from \citealt{Timmes00}) and reaction rates are
-used.  We continue to use a 13-isotope $\alpha$-chain to describe the
-helium burning.  The port of \castro\ to GPUs~\citep{sc20_gpu} 
-significantly improved its overall performance, enabling
-us to run these simulations without the previous approximations while
-continuing to resolve the burning front.
+used. We are also able to use slower, more physically realistic rotation rates.
+  We continue to use a 13-isotope $\alpha$-chain to describe the
+helium burning.
 
@@
 hydrostatic model representing the ash and fuel states and blend the
-two models laterally to create a hot region near the origin of
+two models laterally to create a hot region near the origin of the
 coordinates and a smooth transition to the cooler region at larger
-radii.  The cool initial model is characterized by 3 temperatures:
+radii.  The cool initial model is characterized by three temperatures:
 $T_\mathrm{star}$ is the isothermal temperature of the underlying
@@
 rate and temperature conditions for the flame. The main parameters
-describing the models are provided in Table~\ref{table:runs}. Figure~\ref{fig:hot_profiles} shows the
+describing the models and the names by which we shall refer to them in this paper 
+are provided in Table~\ref{table:runs}. Figure~\ref{fig:hot_profiles} shows the
 temperature and density structure for our hot models and
-figure~\ref{fig:cool_profiles} shows the temperature and density
+Figure~\ref{fig:cool_profiles} shows the temperature and density
 structure for the cool models.
@@
 is carried out in all zones within the atmosphere
-with $\rho > 2.5\times 10^4~\gcc$.
+with $\rho > 2.5\times 10^4~\gcc$. We use an axisymmetric coordinate system, 
+with the horizontal $r$-direction pointing along the surface of the star and the 
+vertical $z$-direction pointing tangential to the surface.
 
@@
 	rate (right panel) in the burning region as functions of time. The quantities on the vertical 
-	axes are the averages defined by Equation \ref{eqn:top_percentile_avg}.}
+	axes are the mass-weighted averages defined in Equation~\ref{eqn:top_percentile_avg}.}
 \end{figure}
 
-{\color{blue}
 For some of our analysis, we would like to have a means of estimating the temperature ($T$) and nuclear 
 energy generation rate ($\enucdot$) in the burning region of each simulation. For this purpose, we define 
-$\repavg{Q}$ to be the mass weighted average
+the mass weighted average $\repavg{Q}$ of some quantity $Q$ to be
 \begin{equation}
@@
 \noindent Here, $C_{99}(Q)$ is the set of grid cells with $Q$ values in the top percentile, and 
-$Q(c_i)$ is the value of the quantity $Q$ in cell $c_i$. An alternative to doing this averaging is 
-to simply take the maximum of the quantity across the entire simulation domain, but using 
-$\repavg{Q}$ instead allows us to track changes over a region rather than at a single localized 
-point. Figure \ref{fig:maxima} shows $\repavg{T}$ and $\repavg{\enucdot}$ as functions of time for 
+$Q(c_i)$ is the value of $Q$ in cell $c_i$\MarginPar{Please define what m(c i) here!}. By using 
+$\repavg{Q}$ instead of simply taking the maximum of the quantity across the entire simulation domain 
+allows us to track changes over the domain as a whole rather than at a single localized 
+point. This will therefore be a better reflection of the overall behavior of the flame rather than of a 
+single localized fluctuation. Figure \ref{fig:maxima} shows $\repavg{T}$ and $\repavg{\enucdot}$ as functions of time for 
 the subset of our runs that achieve a propagating flame. This figure is referenced throughout the 
 subsequent sections.
-}
 
@@
 \end{equation}
-where $g$ is gravitational acceleration, $H_0$ is the atmospheric scale height, and $\Omega$ is the 
+where $g$ is the gravitational acceleration, $H_0$ is the atmospheric scale height, and $\Omega$ is the 
 neutron star rotation rate. In Figure \ref{fig:time_series_enuc_500} and Figure 
 \ref{fig:time_series_enuc_1000}, we use the $\dot{e}_\mathrm{nuc}$ measured at $50~\mathrm{ms}$ and $100~\mathrm{ms}$ to discern the horizontal extent of the flame at different rotation rates.
-Taking the edge at greatest radius of the bright teal/green region (where significant energy generation is occurring) as the leading edge of the flame in each plot, we see that the horizontal extent of the $1000~\mathrm{Hz}$ flame ($\tt{F1000}$) appears to be
+Taking the edge at greatest radius of the bright teal/green region where the most significant energy generation is occurring as the leading edge of the flame in each plot, we see that the horizontal extent of the $1000~\mathrm{Hz}$ flame ($\tt{F1000}$) appears to be
 reduced compared to the lower rotation $500~\mathrm{Hz}$ run ($\tt{F500}$). From Equation 
@@
 flame front nearest to the crust in the {\tt F1000} run reaching higher $\dot{e}_\mathrm{nuc}$ values 
-than for the {\tt F500} run in Figure \ref{fig:time_series_enuc_500}. In contrast to {\tt F500} 
+than for the {\tt F500} run in Figure~\ref{fig:time_series_enuc_500}. In contrast to {\tt F500} 
 and {\tt F1000}, the lowest rotation run --- {\tt F250} --- failed to ignite. The lack of ignition 
@@
 model used here.
-{\color{blue} Given that the size of our domain is $\sim 1~L_R$ for {\tt F250} (using Equation~\ref{eqn:rossby}), it is also possible that we simply cannot confine the flame sufficiently with our current domain width.}
+{\color{blue} Given that the size of our domain is $\sim 1~L_R$ for {\tt F250} (using Equation~\ref{eqn:rossby}), it is also possible that we simply cannot confine the flame sufficiently with our current domain width.} We see in Figure~\ref{fig:flame_speeds_1} that the {\tt F500} flame took longer to achieve steady propagation than the {\tt F1000} flame.
+It may therefore also be that we did not run our simulation for long enough to see the {\tt F250} flame achieve the conditions required for ignition and steady propagation.
 
@@
 \times 10^5 {\rm g\,cm^{-3}}$. The fluid in this dense, high energy generation region undergoes 
-vortical motion, shown in the Figure~\ref{fig:uv} phase plots comparing $u$, vertical z-velocity $v$ and $\dot{e}_{\rm 
+vortical motion, shown in the Figure~\ref{fig:uv} phase plots comparing $u$, the vertical $z$-velocity $v$ and $\dot{e}_{\rm 
 nuc}$. This most likely corresponds to the leading edge of the flame where fresh fuel is being entrained. \MarginPar{not sure if we should keep this sentence about entrainment...}
@@
 set of simulation runs exhibit significant acceleration, for this study we instead fit the data 
-with a quadratic curve of the form:
+with a quadratic curve of the form
 \begin{equation}
 	\label{eqn:quadratic_fit}
-	\color{blue} r(t) = \half a t^2 + v_0 t + r_0
+	r(t) = \half a t^2 + v_0 t + r_0,
 \end{equation}
-\noindent The parameter $a$ is the acceleration of the flame, $v_0$ is the velocity at $t = 
-0~\mathrm{ms}$, and $r_0$ is the position at $t = 0~\mathrm{ms}$. We do not include the data points 
+\noindent where the parameter $a$ is the acceleration of the flame, $v_0$ is the velocity at $t = 
+0~\mathrm{ms}$, and $r_0$ is the flame front position at $t = 0~\mathrm{ms}$. We do not include the data points 
 with $t \lesssim 20~\mathrm{ms}$ when performing the fit, since these correspond to the initial 
@@
 flame speed with rotation rate in this set of runs. We observed earlier, however, that nuclear 
-reactions progress more quickly at higher rotation rate. {\color{blue}This results in a higher 
+reactions progress more quickly at higher rotation rate. This results in a higher 
 $\enucdot$ --- up 
@@
 \ref{fig:maxima}) --- which may counteract the reduction in flame speed from the increased Coriolis 
-confinement.} We also observe that $\tt{F500}$ accelerates faster than $\tt{F1000}$, which appears 
+confinement. We also observe that $\tt{F500}$ accelerates faster than $\tt{F1000}$, which appears 
 to experience a small deceleration at early times. This disparity may be a direct result of the 
@@
 $t = 40~\mathrm{ms}$) compared to all the other runs described in this paper (see Figures 
-\ref{fig:compare_500Hz_abar}, \ref{fig:time_series_500}, and \ref{fig:time_series_1000}).
+\ref{fig:time_series_500}, \ref{fig:time_series_1000} and \ref{fig:compare_500Hz_abar}).
 
@@
 $\tt{F500\_4E8}$ is also noticeably hotter than $\tt{F1000}$, even though it is run for significantly less time ($t = 40~\mathrm{ms}$ vs $t = 100~\mathrm{ms}$).
-While the correlation between increased temperature, greater mixing, and enhanced burning seen in $\tt{F500\_4E8}$ is present throughout our simulations, $\tt{F500\_4E8}$ is unique in that it seems insufficiently confined to maintain a burning front.} \MarginPar{is it correct to say it's insufficiently confined?} {\color{blue} Enhanced burning in this hottest run aligns with results from the other simulations with differing initial temperature structures; as $T_{\mathrm{star}}$ is increased, the overall behavior and propagation of the flame is significantly altered, implying that burning is very temperature-sensitive.
+Although the $\tt{F500\_4E8}$ run is clearly a special case in that it develops steady burning across the domain rather than a propagating flame, enhanced burning in this hottest run aligns with results from the other simulations with differing initial temperature structures.
+As $T_{\mathrm{star}}$ is increased, the overall behavior and propagation of the flame is significantly altered, implying that burning is very temperature-sensitive.
 We explore flame propagation further in Section~\ref{ssec:temp_prop}.}
@@
 $\tt{F500\_4E8}$, we do not analyze its propagation speed and acceleration. As the initial 
-$T_{\mathrm{star}}$ is increased, the flame becomes greatly accelerated. The initial flame 
+$T_{\mathrm{star}}$ is increased beyond $\sim 2\times 10^8$, the flame becomes greatly accelerated. The initial flame 
 velocities and accelerations derived from a quadratic least-squares fit to each of the datasets, as 
@@
 temperatures, the most robust feature is the acceleration of the $\tt{F500\_3E8}$ run at $t \sim 
-40~\mathrm{ms}$. The reason for the acceleration of the flame is not entirely clear. It could be 
+40~\mathrm{ms}$. The reason for the acceleration of the flame is not entirely clear. 
+Whereas for the cooler runs, a state of steady flame propagation is achieved, for the $\tt{F500\_3E8}$ run
+the flame speed continues to increase, suggesting that some instability driving the 
+flame's propagation persists to later times. It could be 
 that energy released from burning begins to dominate the flame's propagation as it evolves, 
@@
 We found that increasing the rotation rate increased the energy generation rate within the flame and enhanced nuclear burning.
-{\color{blue} Apart from the lowest rotation run (which failed to ignite), flame propagation was not noticeably impacted by rotation rate; by the time the different flame fronts reached steady propagation, they shared comparable velocities.
-These results are likely due to the rotation-dependent strength of the Coriolis force and confinement of the flame.} \MarginPar{NF: is the wording of the blue text correct?} We explored several models with different crustal temperatures, to determine what effects mechanisms such as higher accretion rates, crustal heating and shallow heating may have on flame propagation.
+Apart from the lowest rotation run (which failed to ignite), flame propagation was not noticeably impacted by rotation rate; by the time the different flame fronts reached steady propagation, they shared comparable velocities.
+These results are likely due to the rotation-dependent strength of the Coriolis force its confinement of the flame balancing the enhanced nuclear burning.
+We explored several models with different crustal temperatures to determine what effects mechanisms such as higher accretion rates, crustal heating and shallow heating may have on flame propagation.
 We found that increasing the temperature of the crust significantly enhanced the flame propagation.
-This we believe to be because a cooler crust allows heat to more efficiently be transferred from the flame itself, therefore reducing the flame's temperature, slowing burning and consequently reducing its propagation speed.
-We saw that if the crust temperature was too high, then instead of a flame, the atmosphere would burn stably.
+This we believe to be because a cooler crust allows heat to more efficiently be transferred away from the flame itself, therefore reducing the flame's temperature, slowing burning and consequently reducing its propagation speed.
+At higher temperatures, we saw that the flame itself accelerated, the increased temperature and inability for heat to be transported away from the flame front driving unstable, accelerating flame propagation.
+We saw that if the crust temperature was too high, then instead of a flame the entire atmosphere would burn stably.
 This is reminiscent of what is seen for neutron stars with accretion rates exceeding the Eddington luminosity.
@@
 hopefully alleviate some of the boundary effects we have observed in
-these simulations but will require significant computational resource;
+these simulations but will require significant computational resources;
 and exploring the resolution of our simulations more so that we can
-ensure that we have resolved all the necessary physical processes. We
+ensure that we have resolved all of the necessary physical processes. We
 would also like to model H/He flames, as these are the sites of
@@
 recently added an MHD solver to \castro~\citep{sazo2020thesis}; this
-will allow us to explore the effects of magnetic fields on flame
+will allow us in the future to explore the effects of magnetic fields on flame
 propagation in XRBs.

### 2021-01-25 xrb2/paper.tex
Added clarification on m(c_i)

@@
 \noindent Here, $C_{99}(Q)$ is the set of grid cells with $Q$ values in the top percentile, and 
-$Q(c_i)$ is the value of $Q$ in cell $c_i$\MarginPar{Please define what m(c i) here!}. By using 
-$\repavg{Q}$ instead of simply taking the maximum of the quantity across the entire simulation domain 
-allows us to track changes over the domain as a whole rather than at a single localized 
-point. This will therefore be a better reflection of the overall behavior of the flame rather than of a 
-single localized fluctuation. Figure \ref{fig:maxima} shows $\repavg{T}$ and $\repavg{\enucdot}$ as functions of time for 
-the subset of our runs that achieve a propagating flame. This figure is referenced throughout the 
-subsequent sections.
+$Q(c_i)$ is the value of $Q$ in cell $c_i$ (with $m(c_i)$ being the total mass in cell $c_i$). By 
+using $\repavg{Q}$ instead of simply taking the maximum of the quantity across the entire 
+simulation domain allows us to track changes over the domain as a whole rather than at a single 
+localized point. This will therefore be a better reflection of the overall behavior of the flame 
+rather than of a single localized fluctuation. Figure \ref{fig:maxima} shows $\repavg{T}$ and 
+$\repavg{\enucdot}$ as functions of time for the subset of our runs that achieve a propagating 
+flame. This figure is referenced throughout the subsequent sections.
 

### 2021-01-25 xrb2/paper.tex
Updated plot of temp/enucdot maxima

@@
 \begin{abstract}
-We continue to investigate laterally propagating flames in type I X-ray bursts using
-fully compressible hydrodynamics simulations.  In the current study we
-relax the approximations used previously and explore the effect of
-rotation rate and the thermal structure of the neutron star on the
-flame.  We find that at lower rotation rates the flame becomes harder to ignite,
-whereas at higher rotation rates the nuclear burning is enhanced and the flame propagates steadily.
-We also find that higher temperature atmospheres lead to faster
-flames and observe acceleration of the flame front as it propagates
-through the atmosphere.  All of the software used for these
-simulations is freely available.
+We continue to investigate two-dimensional laterally propagating flames in type I X-ray bursts using fully compressible hydrodynamics simulations.
+In the current study we relax previous approximations where we artificially boosted the flames.
+We now use more physically realistic reaction rates, thermal conductivities, and rotation rates, exploring the effects of neutron star rotation rate and thermal structure on the flame.
+We find that at lower rotation rates the flame becomes harder to ignite, whereas at higher rotation rates the nuclear burning is enhanced by increased confinement from the Coriolis force and the flame propagates steadily.
+At higher crustal temperatures, the flame moves more quickly and accelerates as it propagates through the atmosphere.
+If the temperature is too high, , instead of a flame propagating across the surface the entire atmosphere burns steadily.
+All of the software used for these simulations is freely available.
 
@@
 We investigate the effect of rotation rate on the flame. With the
-exception of IGR J17480-2446 (\citet{altamirano2010discovery}, spinning at $11~\mathrm{Hz}$), most observations of
+exception of IGR J17480-2446 (\citealt{altamirano2010discovery}, spinning at $11~\mathrm{Hz}$), most observations of
 XRBs come from sources with rotation rates of $200-600~\mathrm{Hz}$
@@
 
-For some of our analysis, we would like to have a means of estimating the temperature ($T$) and nuclear 
-energy generation rate ($\enucdot$) in the burning region of each simulation. For this purpose, we define 
-the mass weighted average $\repavg{Q}$ of some quantity $Q$ to be
+For some of our analysis, we would like to have a means of estimating the temperature ($T$) and 
+nuclear energy generation rate ($\enucdot$) in the burning region of each simulation. For this 
+purpose, we define the mass-weighted average $\repavg{Q}$ of some quantity $Q$ to be
 \begin{equation}
@@
 At higher temperatures, we saw that the flame itself accelerated, the increased temperature and inability for heat to be transported away from the flame front driving unstable, accelerating flame propagation.
-We saw that if the crust temperature was too high, then instead of a flame the entire atmosphere would burn stably.
+We saw that if the crust temperature was too high, then instead of a flame the entire atmosphere would burn steadily.
 This is reminiscent of what is seen for neutron stars with accretion rates exceeding the Eddington luminosity.

### 2021-01-25 xrb2/paper.tex
Update abstract
update acks
Added to Fig 3 caption

@@
 At higher crustal temperatures, the flame moves more quickly and accelerates as it propagates through the atmosphere.
-If the temperature is too high, , instead of a flame propagating across the surface the entire atmosphere burns steadily.
+If the temperature is too high, instead of a flame propagating across the surface the entire atmosphere burns steadily.
 All of the software used for these simulations is freely available.
@@
 	rate (right panel) in the burning region as functions of time. The quantities on the vertical 
-	axes are the mass-weighted averages defined in Equation~\ref{eqn:top_percentile_avg}.}
+	axes are the mass-weighted averages defined in Equation~\ref{eqn:top_percentile_avg} for the temperature (left panel) and energy generation rate (right panel).}
 \end{figure}
@@
 Discovery through Advanced Computing (SciDAC) program under Award
-Number DE-SC0017955.  This material is also based upon work supported by the U.S. Department
+Number DE-SC0017955.
+This research was supported by the Exascale Computing Project (17-SC-20-SC), a collaborative effort of the U.S.
+Department of Energy Office of Science and the National Nuclear Security Administration.
+This material is also based upon work supported by the U.S. Department
 of Energy, Office of Science, Office of Advanced Scientific Computing Research, Department of

### 2021-01-25 xrb2/paper.tex
update conclusions

@@
 Apart from the lowest rotation run (which failed to ignite), flame propagation was not noticeably impacted by rotation rate; by the time the different flame fronts reached steady propagation, they shared comparable velocities.
-These results are likely due to the rotation-dependent strength of the Coriolis force its confinement of the flame balancing the enhanced nuclear burning.
+These results are likely due to the rotation-dependent strength of the Coriolis force and its confinement of the flame balancing the enhanced nuclear burning.
 We explored several models with different crustal temperatures to determine what effects mechanisms such as higher accretion rates, crustal heating and shallow heating may have on flame propagation.
@@
 This we believe to be because a cooler crust allows heat to more efficiently be transferred away from the flame itself, therefore reducing the flame's temperature, slowing burning and consequently reducing its propagation speed.
-At higher temperatures, we saw that the flame itself accelerated, the increased temperature and inability for heat to be transported away from the flame front driving unstable, accelerating flame propagation.
+At higher crustal temperatures, we saw that the inability for heat to be efficiently transported away from the flame front increased the flame temperature, driving unstable, accelerating flame propagation.
 We saw that if the crust temperature was too high, then instead of a flame the entire atmosphere would burn steadily.
@@
 
-In future simulations, we would like to improve and expand our
+In future work, we would like to improve and expand our
 simulations in order to better understand the processes at play and to

### 2021-01-25 xrb2/paper.tex
De-blue text

@@
 nuc}$. This most likely corresponds to the leading edge of the flame where fresh fuel is being entrained. \MarginPar{not sure if we should keep this sentence about entrainment...}
-This feature is not developed in the $250~\mathrm{Hz}$ flame in Figure~\ref{fig:uv} {\color{blue} (left panel)}; it could potentially develop 
+This feature is not developed in the $250~\mathrm{Hz}$ flame in Figure~\ref{fig:uv} (left panel); it could potentially develop 
 at later times (past the point at which we terminated our simulation), or the flame could just 
@@
 Figure~\ref{fig:compare_500Hz_abar} shows $\bar{A}$ for three $500~\mathrm{Hz}$ simulations with different initial temperature structures ($T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$) at $t = 70~\mathrm{ms}$.
-{\color{blue}We do not plot $\tt{F500\_4E8}$ here because it fails to form a clear burning front.
-$\tt{F500\_3E8}$ (Figure~\ref{fig:compare_500Hz_abar}, bottom panel) --- the hottest run to form a clear burning front ---} has a faster propagating flame (this will be discussed further in Section \ref{ssec:temp_prop}).
-{\color{blue} It also reaches slightly higher $\bar{A}$ values than the two cooler runs.
+We do not plot $\tt{F500\_4E8}$ here because it fails to form a clear burning front.
+$\tt{F500\_3E8}$ (Figure~\ref{fig:compare_500Hz_abar}, bottom panel) --- the hottest run to form a clear burning front --- has a faster propagating flame (this will be discussed further in Section \ref{ssec:temp_prop}).
+It also reaches slightly higher $\bar{A}$ values than the two cooler runs.
 The Figure~\ref{fig:abar_hot} $\bar{A}$-$T$ phase plots of $\tt{F500}$ (left) and $\tt{F500\_3E8}$ (middle) also reflect these $\bar{A}$ features, with $\tt{F500\_3E8}$ reaching slightly higher $\bar{A}$ values.
@@
 There appear to be more causally connected regions across a range of $\bar{A}$ at low temperatures for $\tt{F500\_3E8}$ than for $\tt{F500}$, suggesting that the higher $\enucdot$ for $\tt{F500\_3E8}$ generates burning in certain reaction pathways that are not present in the cooler $\tt{F500}$ run.
-Note that, although $\tt{F500\_3E8}$ is the hottest run with a modified initial temperature configuration to form a distinct flame front, the highest rotation $\tt{F1000}$ flame actually reaches higher temperatures (see Figure~\ref{fig:maxima}, left panel) as well as higher $\bar{A}$ (see Figure~\ref{fig:time_series_1000}).}
+Note that, although $\tt{F500\_3E8}$ is the hottest run with a modified initial temperature configuration to form a distinct flame front, the highest rotation $\tt{F1000}$ flame actually reaches higher temperatures (see Figure~\ref{fig:maxima}, left panel) as well as higher $\bar{A}$ (see Figure~\ref{fig:time_series_1000}).
 
@@
 $\enucdot$ values on the order of $10^{18}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ 
-{\color{blue} (see Figure~\ref{fig:maxima}, right panel)}, even at the latest timesteps ($\sim 
-70~\mathrm{ms}$, when the flame is most developed). {\color{blue} Significant $\enucdot$ values for 
-all other runs are confined to the flame front, rather than spanning the entire domain}.
+(see Figure~\ref{fig:maxima}, right panel), even at the latest timesteps ($\sim 
+70~\mathrm{ms}$, when the flame is most developed). Significant $\enucdot$ values for 
+all other runs are confined to the flame front, rather than spanning the entire domain.
 
@@
 
- {\color{blue} Though $\tt{F500\_4E8}$ does not form a distinct burning front, it does achieve greatly enhanced burning, as shown in the Figure~\ref{fig:abar_hot} $\bar{A}$-$T$ phase plot (right panel).
-Not only is $\tt{F500\_4E8}$ significantly hotter overall compared to all other runs}, but there is a large causally connected region across a wide range of $\bar{A}$.
+ Though $\tt{F500\_4E8}$ does not form a distinct burning front, it does achieve greatly enhanced burning, as shown in the Figure~\ref{fig:abar_hot} $\bar{A}$-$T$ phase plot (right panel).
+Not only is $\tt{F500\_4E8}$ significantly hotter overall compared to all other runs, but there is a large causally connected region across a wide range of $\bar{A}$.
 This indicates that the hotter temperature has facilitated significant burning in reaction pathways that were not favored at the lower temperatures.
 The reaction pathways are also very disrupted for $\tt{F500\_4E8}$ compared to the cooler runs, suggesting that the hotter temperature leads to more vigorous mixing.
-{\color{blue}Indeed, this appears to be the case looking at Figure~\ref{fig:4e8_stacked_abar} and the right panel of Figure~\ref{fig:uv_hot} (discussed further in section~\ref{ssec:temp_prop}).
+Indeed, this appears to be the case looking at Figure~\ref{fig:4e8_stacked_abar} and the right panel of Figure~\ref{fig:uv_hot} (discussed further in section~\ref{ssec:temp_prop}).
 The disrupted reaction pathways resemble those found in Figure \ref{fig:abar} for the highest rotation $\tt{F1000}$ run, though they are even more dramatically disrupted for $\tt{F500\_4E8}$.
@@
 As $T_{\mathrm{star}}$ is increased, the overall behavior and propagation of the flame is significantly altered, implying that burning is very temperature-sensitive.
-We explore flame propagation further in Section~\ref{ssec:temp_prop}.}
+We explore flame propagation further in Section~\ref{ssec:temp_prop}.
 
@@
 
-The $u$-$\rho$ phase plot in Figure~\ref{fig:urho_hot} shows that the horizontal velocity for the $T_{\rm star} = 4\times10^8$~K {\color{blue} ($\tt{F500\_4e8}$; right panel)} is significantly greater in magnitude than for the cooler {\color{blue}$\tt{F500\_3e8}$ (middle panel) and $\tt{F500}$ (left panel)} runs.
-The $u$-$v$ plot in Figure~\ref{fig:uv_hot} further demonstrates this, as again we see that the magnitude of the {\color{blue}$\tt{F500\_4e8}$ flame's vertical $z$-velocity $v$} is also significantly larger.
+The $u$-$\rho$ phase plot in Figure~\ref{fig:urho_hot} shows that the horizontal velocity for the $T_{\rm star} = 4\times10^8$~K ($\tt{F500\_4e8}$; right panel) is significantly greater in magnitude than for the cooler $\tt{F500\_3e8}$ (middle panel) and $\tt{F500}$ (left panel) runs.
+The $u$-$v$ plot in Figure~\ref{fig:uv_hot} further demonstrates this, as again we see that the magnitude of the $\tt{F500\_4e8}$ flame's vertical $z$-velocity $v$ is also significantly larger.
 This motion therefore most likely corresponds to the vortical motion rather than to the propagation of the flame itself, something which is further supported by the fact that the velocity magnitude is two to three orders of magnitude greater than the flame speeds listed in Table~\ref{table:flame_speeds}.
@@
     \plotone{urho_plot_hotter}
-    \caption{\label{fig:urho_hot}Phase plot of the horizontal $r$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$ {\color{blue} (the latest time that $\tt{F500\_4E8}$ is run out to).
-The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.}
+    \caption{\label{fig:urho_hot}Phase plot of the horizontal $r$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$ (the latest time that $\tt{F500\_4E8}$ is run out to).
+The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.
 \end{figure}
@@
     \caption{\label{fig:uv_hot}Phase plot of the horizontal $r$-velocity $u$, vertical $z$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$.
-{\color{blue} The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.}
+The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.
 \end{figure}
@@
     \caption{\label{fig:abar_hot}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$.
-{\color{blue} The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.}
+The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.
 \end{figure}

### 2021-01-26 xrb2/paper.tex
De-blued more text, removed margin comment

@@
 higher rotation rates ($1000~\mathrm{Hz}$ and higher), based on the results of \citet{flame_wave1}.
-{\color{blue} Perhaps these new un-boosted runs have not had sufficient time to reach steady propagation, in contrast to the boosted runs in \citet{flame_wave1} that reached steady propagation more quickly.} \MarginPar{does the equation for the Rossby radius only hold for flames with steady propagation?}
 
@@
 model used here.
-{\color{blue} Given that the size of our domain is $\sim 1~L_R$ for {\tt F250} (using Equation~\ref{eqn:rossby}), it is also possible that we simply cannot confine the flame sufficiently with our current domain width.} We see in Figure~\ref{fig:flame_speeds_1} that the {\tt F500} flame took longer to achieve steady propagation than the {\tt F1000} flame.
+Given that the size of our domain is $\sim 1~L_R$ for {\tt F250} (using Equation~\ref{eqn:rossby}), it is also possible that we simply cannot confine the flame sufficiently with our current domain width.
+We see in Figure~\ref{fig:flame_speeds_1} that the {\tt F500} flame took longer to achieve steady propagation than the {\tt F1000} flame.
 It may therefore also be that we did not run our simulation for long enough to see the {\tt F250} flame achieve the conditions required for ignition and steady propagation.
@@
 vortical motion, shown in the Figure~\ref{fig:uv} phase plots comparing $u$, the vertical $z$-velocity $v$ and $\dot{e}_{\rm 
-nuc}$. This most likely corresponds to the leading edge of the flame where fresh fuel is being entrained. \MarginPar{not sure if we should keep this sentence about entrainment...}
+nuc}$. This most likely corresponds to the leading edge of the flame where fresh fuel is being entrained.
 This feature is not developed in the $250~\mathrm{Hz}$ flame in Figure~\ref{fig:uv} (left panel); it could potentially develop 

### 2021-01-26 xrb2/paper.tex
Removed margin comment

@@
 anything to do with the flame at all, but that neutron stars in
-accreting low mass X-ray binaries rarely have rotation rates below $200~\mathrm{Hz}$ 
-\MarginPar{look into this?}.
+accreting low mass X-ray binaries rarely have rotation rates below $200~\mathrm{Hz}$.
 

### 2021-01-26 xrb2/paper.tex
Small edits

@@
 propagating flame through the cool fuel.  In our original set of
-calculations (in in \citet{flame_wave1}), in order to make the 
+calculations (in \citet{flame_wave1}), in order to make the 
 simulations computationally feasible we artificially boosted the flame speed by adjusting the
@@
 faster than the nominal laminar flame speed.  We also used high
-rotation rates to reduce the lateral lengthscale at which the Coriolis
+rotation rates ($\geq 1000~\mathrm{Hz}$) to reduce the lateral lengthscale at which the Coriolis
 force balances the lateral flame spreading in order to reduce the size 
@@
 	rate (right panel) in the burning region as functions of time. The quantities on the vertical 
-	axes are the mass-weighted averages defined in Equation~\ref{eqn:top_percentile_avg} for the temperature (left panel) and energy generation rate (right panel).}
+	axes are the mass-weighted averages defined in Equation~\ref{eqn:top_percentile_avg}.}
 \end{figure}
@@
 \noindent Here, $C_{99}(Q)$ is the set of grid cells with $Q$ values in the top percentile, and 
-$Q(c_i)$ is the value of $Q$ in cell $c_i$ (with $m(c_i)$ being the total mass in cell $c_i$). By 
-using $\repavg{Q}$ instead of simply taking the maximum of the quantity across the entire 
+$Q(c_i)$ is the value of $Q$ in cell $c_i$ (with $m(c_i)$ being the total mass in cell $c_i$). Using $\repavg{Q}$ instead of simply taking the maximum of the quantity across the entire 
 simulation domain allows us to track changes over the domain as a whole rather than at a single 
@@
 Given that the size of our domain is $\sim 1~L_R$ for {\tt F250} (using Equation~\ref{eqn:rossby}), it is also possible that we simply cannot confine the flame sufficiently with our current domain width.
-We see in Figure~\ref{fig:flame_speeds_1} that the {\tt F500} flame took longer to achieve steady propagation than the {\tt F1000} flame.
+We see in Figure~\ref{fig:flame_speeds_1} (discussed further in Section~\ref{ssec:rot_propagation}) that the {\tt F500} flame took longer to achieve steady propagation than the {\tt F1000} flame.
 It may therefore also be that we did not run our simulation for long enough to see the {\tt F250} flame achieve the conditions required for ignition and steady propagation.
@@
 	simulation. $v_0$ is the flame velocity in the $+r$ direction at $t 
-	= 0~\mathrm{ms}$. The initial flame 
+	= 0~\mathrm{ms}$ and $a$ is the acceleration of the flame. The initial flame 
 	velocities and accelerations are derived from a quadratic least-squares fit to each of the 
@@
 10^4~\mathrm{cm}$. A similar effect is also visible (to a lesser extent) in the $\tt{F500\_3E8}$ 
-plot in Figure \ref{fig:compare_500Hz_abar}, with a faint haze of material rising above the flame. 
+plot in Figure \ref{fig:compare_500Hz_abar}, with a faint haze of material rising above the flame near the axis of symmetry. 
 However, this is likely to be a numerical artefact rather than a true physical effect, and a result 
@@
 	\plotone{compare_500Hz_enuc_4e8.pdf}
-	\caption{\label{fig:4e8_stacked_enuc} Time series of the $500~\mathrm{Hz}$ run $\tt{F500\_4E8}$, with $T_{\mathrm{star}} = 4 \times 10^8~\mathrm{K}$, showing $\dot{e}_{nuc}$.
+	\caption{\label{fig:4e8_stacked_enuc} Time series of the $500~\mathrm{Hz}$ run $\tt{F500\_4E8}$, with $T_{\mathrm{star}} = 4 \times 10^8~\mathrm{K}$, showing $\enucdot$.
 This model burns so strongly that it is only run out to $40~\mathrm{ms}$; the snapshots shown here are at $T = 30~\mathrm{ms}$, $35~\mathrm{ms}$, and $40~\mathrm{ms}$.}
@@
 $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$. Due to the lack of a clear burning front in 
-$\tt{F500\_4E8}$, we do not analyze its propagation speed and acceleration. As the initial 
+$\tt{F500\_4E8}$, we do not analyze its propagation velocity and acceleration. As the initial 
 $T_{\mathrm{star}}$ is increased beyond $\sim 2\times 10^8$, the flame becomes greatly accelerated. The initial flame 
@@
     \caption{\label{fig:urho_hot}Phase plot of the horizontal $r$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$ (the latest time that $\tt{F500\_4E8}$ is run out to).
-The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.
+The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here.}
 \end{figure}
@@
     \caption{\label{fig:uv_hot}Phase plot of the horizontal $r$-velocity $u$, vertical $z$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$.
-The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.
+The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here.}
 \end{figure}
@@
     \caption{\label{fig:abar_hot}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$.
-The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.
+The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here.}
 \end{figure}

### 2021-01-26 xrb2/paper.tex
minor changes

--- 2021-01-26 xrb2/paper.tex
@@
 \begin{abstract}
-We continue to investigate two-dimensional laterally propagating flames in type I X-ray bursts using fully compressible hydrodynamics simulations.
-In the current study we relax previous approximations where we artificially boosted the flames.
-We now use more physically realistic reaction rates, thermal conductivities, and rotation rates, exploring the effects of neutron star rotation rate and thermal structure on the flame.
-We find that at lower rotation rates the flame becomes harder to ignite, whereas at higher rotation rates the nuclear burning is enhanced by increased confinement from the Coriolis force and the flame propagates steadily.
-At higher crustal temperatures, the flame moves more quickly and accelerates as it propagates through the atmosphere.
-If the temperature is too high, instead of a flame propagating across the surface the entire atmosphere burns steadily.
-All of the software used for these simulations is freely available.
+We continue to investigate laterally propagating flames in type I X-ray bursts using
+fully compressible hydrodynamics simulations.  In the current study we
+relax the approximations used previously and explore the effect of
+rotation rate and the thermal structure of the neutron star on the
+flame.  We find that at lower rotation rates the flame becomes harder to ignite,
+whereas at higher rotation rates the nuclear burning is enhanced and the flame propagates steadily.
+We also find that higher temperature atmospheres lead to faster
+flames and observe acceleration of the flame front as it propagates
+through the atmosphere.  All of the software used for these
+simulations is freely available.
 
@@
 We investigate the effect of rotation rate on the flame. With the
-exception of IGR J17480-2446 (\citealt{altamirano2010discovery}, spinning at $11~\mathrm{Hz}$), most observations of
+exception of IGR J17480-2446 (\citet{altamirano2010discovery}, spinning at $11~\mathrm{Hz}$), most observations of
 XRBs come from sources with rotation rates of $200-600~\mathrm{Hz}$
@@
 	rate (right panel) in the burning region as functions of time. The quantities on the vertical 
-	axes are the mass-weighted averages defined in Equation~\ref{eqn:top_percentile_avg} for the temperature (left panel) and energy generation rate (right panel).}
+	axes are the mass-weighted averages defined in Equation~\ref{eqn:top_percentile_avg}.}
 \end{figure}
 
-For some of our analysis, we would like to have a means of estimating the temperature ($T$) and 
-nuclear energy generation rate ($\enucdot$) in the burning region of each simulation. For this 
-purpose, we define the mass-weighted average $\repavg{Q}$ of some quantity $Q$ to be
+For some of our analysis, we would like to have a means of estimating the temperature ($T$) and nuclear 
+energy generation rate ($\enucdot$) in the burning region of each simulation. For this purpose, we define 
+the mass weighted average $\repavg{Q}$ of some quantity $Q$ to be
 \begin{equation}
@@
 \noindent Here, $C_{99}(Q)$ is the set of grid cells with $Q$ values in the top percentile, and 
-$Q(c_i)$ is the value of $Q$ in cell $c_i$ (with $m(c_i)$ being the total mass in cell $c_i$). By 
-using $\repavg{Q}$ instead of simply taking the maximum of the quantity across the entire 
-simulation domain allows us to track changes over the domain as a whole rather than at a single 
-localized point. This will therefore be a better reflection of the overall behavior of the flame 
-rather than of a single localized fluctuation. Figure \ref{fig:maxima} shows $\repavg{T}$ and 
-$\repavg{\enucdot}$ as functions of time for the subset of our runs that achieve a propagating 
-flame. This figure is referenced throughout the subsequent sections.
+$Q(c_i)$ is the value of $Q$ in cell $c_i$\MarginPar{Please define what m(c i) here!}. By using 
+$\repavg{Q}$ instead of simply taking the maximum of the quantity across the entire simulation domain 
+allows us to track changes over the domain as a whole rather than at a single localized 
+point. This will therefore be a better reflection of the overall behavior of the flame rather than of a 
+single localized fluctuation. Figure \ref{fig:maxima} shows $\repavg{T}$ and $\repavg{\enucdot}$ as functions of time for 
+the subset of our runs that achieve a propagating flame. This figure is referenced throughout the 
+subsequent sections.
 
@@
 higher rotation rates ($1000~\mathrm{Hz}$ and higher), based on the results of \citet{flame_wave1}.
+{\color{blue} Perhaps these new un-boosted runs have not had sufficient time to reach steady propagation, in contrast to the boosted runs in \citet{flame_wave1} that reached steady propagation more quickly.} \MarginPar{does the equation for the Rossby radius only hold for flames with steady propagation?}
 
@@
 model used here.
-Given that the size of our domain is $\sim 1~L_R$ for {\tt F250} (using Equation~\ref{eqn:rossby}), it is also possible that we simply cannot confine the flame sufficiently with our current domain width.
-We see in Figure~\ref{fig:flame_speeds_1} that the {\tt F500} flame took longer to achieve steady propagation than the {\tt F1000} flame.
+{\color{blue} Given that the size of our domain is $\sim 1~L_R$ for {\tt F250} (using Equation~\ref{eqn:rossby}), it is also possible that we simply cannot confine the flame sufficiently with our current domain width.} We see in Figure~\ref{fig:flame_speeds_1} that the {\tt F500} flame took longer to achieve steady propagation than the {\tt F1000} flame.
 It may therefore also be that we did not run our simulation for long enough to see the {\tt F250} flame achieve the conditions required for ignition and steady propagation.
@@
 vortical motion, shown in the Figure~\ref{fig:uv} phase plots comparing $u$, the vertical $z$-velocity $v$ and $\dot{e}_{\rm 
-nuc}$. This most likely corresponds to the leading edge of the flame where fresh fuel is being entrained.
-This feature is not developed in the $250~\mathrm{Hz}$ flame in Figure~\ref{fig:uv} (left panel); it could potentially develop 
+nuc}$. This most likely corresponds to the leading edge of the flame where fresh fuel is being entrained. \MarginPar{not sure if we should keep this sentence about entrainment...}
+This feature is not developed in the $250~\mathrm{Hz}$ flame in Figure~\ref{fig:uv} {\color{blue} (left panel)}; it could potentially develop 
 at later times (past the point at which we terminated our simulation), or the flame could just 
@@
 Figure~\ref{fig:compare_500Hz_abar} shows $\bar{A}$ for three $500~\mathrm{Hz}$ simulations with different initial temperature structures ($T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$) at $t = 70~\mathrm{ms}$.
-We do not plot $\tt{F500\_4E8}$ here because it fails to form a clear burning front.
-$\tt{F500\_3E8}$ (Figure~\ref{fig:compare_500Hz_abar}, bottom panel) --- the hottest run to form a clear burning front --- has a faster propagating flame (this will be discussed further in Section \ref{ssec:temp_prop}).
-It also reaches slightly higher $\bar{A}$ values than the two cooler runs.
+{\color{blue}We do not plot $\tt{F500\_4E8}$ here because it fails to form a clear burning front.
+$\tt{F500\_3E8}$ (Figure~\ref{fig:compare_500Hz_abar}, bottom panel) --- the hottest run to form a clear burning front ---} has a faster propagating flame (this will be discussed further in Section \ref{ssec:temp_prop}).
+{\color{blue} It also reaches slightly higher $\bar{A}$ values than the two cooler runs.
 The Figure~\ref{fig:abar_hot} $\bar{A}$-$T$ phase plots of $\tt{F500}$ (left) and $\tt{F500\_3E8}$ (middle) also reflect these $\bar{A}$ features, with $\tt{F500\_3E8}$ reaching slightly higher $\bar{A}$ values.
@@
 There appear to be more causally connected regions across a range of $\bar{A}$ at low temperatures for $\tt{F500\_3E8}$ than for $\tt{F500}$, suggesting that the higher $\enucdot$ for $\tt{F500\_3E8}$ generates burning in certain reaction pathways that are not present in the cooler $\tt{F500}$ run.
-Note that, although $\tt{F500\_3E8}$ is the hottest run with a modified initial temperature configuration to form a distinct flame front, the highest rotation $\tt{F1000}$ flame actually reaches higher temperatures (see Figure~\ref{fig:maxima}, left panel) as well as higher $\bar{A}$ (see Figure~\ref{fig:time_series_1000}).
+Note that, although $\tt{F500\_3E8}$ is the hottest run with a modified initial temperature configuration to form a distinct flame front, the highest rotation $\tt{F1000}$ flame actually reaches higher temperatures (see Figure~\ref{fig:maxima}, left panel) as well as higher $\bar{A}$ (see Figure~\ref{fig:time_series_1000}).}
 
@@
 $\enucdot$ values on the order of $10^{18}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ 
-(see Figure~\ref{fig:maxima}, right panel), even at the latest timesteps ($\sim 
-70~\mathrm{ms}$, when the flame is most developed). Significant $\enucdot$ values for 
-all other runs are confined to the flame front, rather than spanning the entire domain.
+{\color{blue} (see Figure~\ref{fig:maxima}, right panel)}, even at the latest timesteps ($\sim 
+70~\mathrm{ms}$, when the flame is most developed). {\color{blue} Significant $\enucdot$ values for 
+all other runs are confined to the flame front, rather than spanning the entire domain}.
 
@@
 
- Though $\tt{F500\_4E8}$ does not form a distinct burning front, it does achieve greatly enhanced burning, as shown in the Figure~\ref{fig:abar_hot} $\bar{A}$-$T$ phase plot (right panel).
-Not only is $\tt{F500\_4E8}$ significantly hotter overall compared to all other runs, but there is a large causally connected region across a wide range of $\bar{A}$.
+ {\color{blue} Though $\tt{F500\_4E8}$ does not form a distinct burning front, it does achieve greatly enhanced burning, as shown in the Figure~\ref{fig:abar_hot} $\bar{A}$-$T$ phase plot (right panel).
+Not only is $\tt{F500\_4E8}$ significantly hotter overall compared to all other runs}, but there is a large causally connected region across a wide range of $\bar{A}$.
 This indicates that the hotter temperature has facilitated significant burning in reaction pathways that were not favored at the lower temperatures.
 The reaction pathways are also very disrupted for $\tt{F500\_4E8}$ compared to the cooler runs, suggesting that the hotter temperature leads to more vigorous mixing.
-Indeed, this appears to be the case looking at Figure~\ref{fig:4e8_stacked_abar} and the right panel of Figure~\ref{fig:uv_hot} (discussed further in section~\ref{ssec:temp_prop}).
+{\color{blue}Indeed, this appears to be the case looking at Figure~\ref{fig:4e8_stacked_abar} and the right panel of Figure~\ref{fig:uv_hot} (discussed further in section~\ref{ssec:temp_prop}).
 The disrupted reaction pathways resemble those found in Figure \ref{fig:abar} for the highest rotation $\tt{F1000}$ run, though they are even more dramatically disrupted for $\tt{F500\_4E8}$.
@@
 As $T_{\mathrm{star}}$ is increased, the overall behavior and propagation of the flame is significantly altered, implying that burning is very temperature-sensitive.
-We explore flame propagation further in Section~\ref{ssec:temp_prop}.
+We explore flame propagation further in Section~\ref{ssec:temp_prop}.}
 
@@
 
-The $u$-$\rho$ phase plot in Figure~\ref{fig:urho_hot} shows that the horizontal velocity for the $T_{\rm star} = 4\times10^8$~K ($\tt{F500\_4e8}$; right panel) is significantly greater in magnitude than for the cooler $\tt{F500\_3e8}$ (middle panel) and $\tt{F500}$ (left panel) runs.
-The $u$-$v$ plot in Figure~\ref{fig:uv_hot} further demonstrates this, as again we see that the magnitude of the $\tt{F500\_4e8}$ flame's vertical $z$-velocity $v$ is also significantly larger.
+The $u$-$\rho$ phase plot in Figure~\ref{fig:urho_hot} shows that the horizontal velocity for the $T_{\rm star} = 4\times10^8$~K {\color{blue} ($\tt{F500\_4e8}$; right panel)} is significantly greater in magnitude than for the cooler {\color{blue}$\tt{F500\_3e8}$ (middle panel) and $\tt{F500}$ (left panel)} runs.
+The $u$-$v$ plot in Figure~\ref{fig:uv_hot} further demonstrates this, as again we see that the magnitude of the {\color{blue}$\tt{F500\_4e8}$ flame's vertical $z$-velocity $v$} is also significantly larger.
 This motion therefore most likely corresponds to the vortical motion rather than to the propagation of the flame itself, something which is further supported by the fact that the velocity magnitude is two to three orders of magnitude greater than the flame speeds listed in Table~\ref{table:flame_speeds}.
@@
     \plotone{urho_plot_hotter}
-    \caption{\label{fig:urho_hot}Phase plot of the horizontal $r$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$ (the latest time that $\tt{F500\_4E8}$ is run out to).
-The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.
+    \caption{\label{fig:urho_hot}Phase plot of the horizontal $r$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$ {\color{blue} (the latest time that $\tt{F500\_4E8}$ is run out to).
+The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.}
 \end{figure}
@@
     \caption{\label{fig:uv_hot}Phase plot of the horizontal $r$-velocity $u$, vertical $z$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$.
-The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.
+{\color{blue} The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.}
 \end{figure}
@@
     \caption{\label{fig:abar_hot}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$.
-The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.
+{\color{blue} The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.}
 \end{figure}
@@
 Apart from the lowest rotation run (which failed to ignite), flame propagation was not noticeably impacted by rotation rate; by the time the different flame fronts reached steady propagation, they shared comparable velocities.
-These results are likely due to the rotation-dependent strength of the Coriolis force and its confinement of the flame balancing the enhanced nuclear burning.
+These results are likely due to the rotation-dependent strength of the Coriolis force its confinement of the flame balancing the enhanced nuclear burning.
 We explored several models with different crustal temperatures to determine what effects mechanisms such as higher accretion rates, crustal heating and shallow heating may have on flame propagation.
@@
 This we believe to be because a cooler crust allows heat to more efficiently be transferred away from the flame itself, therefore reducing the flame's temperature, slowing burning and consequently reducing its propagation speed.
-At higher crustal temperatures, we saw that the inability for heat to be efficiently transported away from the flame front increased the flame temperature, driving unstable, accelerating flame propagation.
-We saw that if the crust temperature was too high, then instead of a flame the entire atmosphere would burn steadily.
+At higher temperatures, we saw that the flame itself accelerated, the increased temperature and inability for heat to be transported away from the flame front driving unstable, accelerating flame propagation.
+We saw that if the crust temperature was too high, then instead of a flame the entire atmosphere would burn stably.
 This is reminiscent of what is seen for neutron stars with accretion rates exceeding the Eddington luminosity.
 
-In future work, we would like to improve and expand our
+In future simulations, we would like to improve and expand our
 simulations in order to better understand the processes at play and to
@@
 Discovery through Advanced Computing (SciDAC) program under Award
-Number DE-SC0017955.
-This research was supported by the Exascale Computing Project (17-SC-20-SC), a collaborative effort of the U.S.
-Department of Energy Office of Science and the National Nuclear Security Administration.
-This material is also based upon work supported by the U.S. Department
+Number DE-SC0017955.  This material is also based upon work supported by the U.S. Department
 of Energy, Office of Science, Office of Advanced Scientific Computing Research, Department of

### 2021-01-26 xrb2/paper.tex
minor changes

--- 2021-01-26 xrb2/paper.tex
@@
 could be that there is some physical process which inhibits the flame
-ignition and/or spread at lower rotation rates. It could be that burst
-oscillations at lower rotation rates are smaller in amplitude and
+ignition and/or spread at lower rotation rates. It could be that bursts
+at lower rotation rates are smaller in amplitude and
 therefore more difficult to detect. It could be that it does not have
@@
 propagating flame through the cool fuel.  In our original set of
-calculations (in \citet{flame_wave1}), in order to make the 
+calculations (in \citep{flame_wave1}), in order to make the 
 simulations computationally feasible we artificially boosted the flame speed by adjusting the
@@
 is carried out in all zones within the atmosphere
-with $\rho > 2.5\times 10^4~\gcc$. We use an axisymmetric coordinate system, 
+with density $\rho > 2.5\times 10^4~\gcc$. We use an axisymmetric coordinate system, 
 with the horizontal $r$-direction pointing along the surface of the star and the 
@@
 \end{equation}
-\noindent Here, $C_{99}(Q)$ is the set of grid cells with $Q$ values in the top percentile, and 
-$Q(c_i)$ is the value of $Q$ in cell $c_i$ (with $m(c_i)$ being the total mass in cell $c_i$). Using $\repavg{Q}$ instead of simply taking the maximum of the quantity across the entire 
+\noindent Here, $C_{99}(Q)$ is the set of grid cells with $Q$ values in the top percentile, 
+$Q(c_i)$ is the value of $Q$ in cell $c_i$, and $m(c_i)$ is the total mass in cell $c_i$. 
+Using $\repavg{Q}$ instead of simply taking the maximum of the quantity across the entire 
 simulation domain allows us to track changes over the domain as a whole rather than at a single 
@@
 neutron star rotation rate. In Figure \ref{fig:time_series_enuc_500} and Figure 
-\ref{fig:time_series_enuc_1000}, we use the $\dot{e}_\mathrm{nuc}$ measured at $50~\mathrm{ms}$ and $100~\mathrm{ms}$ to discern the horizontal extent of the flame at different rotation rates.
+\ref{fig:time_series_enuc_1000}, we use $\dot{e}_\mathrm{nuc}$ measured at $50~\mathrm{ms}$ and $100~\mathrm{ms}$ to discern the horizontal extent of the flame at different rotation rates.
 Taking the edge at greatest radius of the bright teal/green region where the most significant energy generation is occurring as the leading edge of the flame in each plot, we see that the horizontal extent of the $1000~\mathrm{Hz}$ flame ($\tt{F1000}$) appears to be
@@
 potentially leading to insufficient confinement such that the temperature and density required for 
-ignition is not achieved. In this scenario, another source of confinement (e.g.\ magnetic fields, see \citep{art-2016-cavecchi-etal}) 
+ignition is not achieved. In this scenario, another source of confinement (e.g.\ magnetic fields, see \cite{art-2016-cavecchi-etal}) 
 would need to take over at lower rotation rates to allow a burst to occur, at least for the initial 
@@
 This feature is not developed in the $250~\mathrm{Hz}$ flame in Figure~\ref{fig:uv} (left panel); it could potentially develop 
-at later times (past the point at which we terminated our simulation), or the flame could just 
+at later times (past the point at which we terminated our simulation), or the burning could just 
 fizzle out and fail to ignite entirely.

### 2021-01-26 xrb2/paper.tex
minor changes

@@
 at later times (past the point at which we terminated our simulation), or the burning could just 
-fizzle out and fail to ignite entirely.
+fizzle out and the flame fail to ignite entirely.
 

### 2021-01-27 xrb2/paper.tex
fix accretion rate stuff so it's more agnostic on the high vs low accretion rate causing increased crustal heating front

@@
 have higher speeds when moving into hotter upstream fuel.
-It is known that crustal heating is stronger at lower
+It has been suggested that crustal heating may be stronger at lower
 accretion rates and weaker at higher accretion rates, due to the
-effect of neutrino losses \citep{Cumming2006,johnston:2019}. 
+effect of neutrino losses \citep{Cumming2006,johnston:2019}, however at 
+very high accretion rates it has been observed that the entire atmosphere 
+simmers in place rather than a flame propagating across the surface \citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}. 
 A shallow heating
@@
 and/or
-mimic the effects of lower accretion rates.
+mimic the effects of accretion-induced heating.
 
@@
 By choosing $T_\mathrm{star}$ = $T_\mathrm{hi}$, we can effectively explore simulations with greater surface heating.
-There are several physically distinct mechanisms that could produce an increased temperature at the crust: crustal heating, some other form of shallow heating or a higher accretion rate.
+There are several physically distinct mechanisms that could produce an increased temperature at the crust: crustal heating, some other form of shallow heating or accretion-induced heating.
 In these simulations, we do not model the mechanism producing the heating effect, just the effect itself, so we do not distinguish between which of these mechanisms cause the heating.

### 2021-01-28 xrb2/paper.tex
update affil

--- 2021-01-26 xrb2/paper.tex
@@
              Stony Brook, NY 11794-3800}
-\affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010}
 
@@
 could be that there is some physical process which inhibits the flame
-ignition and/or spread at lower rotation rates. It could be that bursts
-at lower rotation rates are smaller in amplitude and
+ignition and/or spread at lower rotation rates. It could be that burst
+oscillations at lower rotation rates are smaller in amplitude and
 therefore more difficult to detect. It could be that it does not have
 anything to do with the flame at all, but that neutron stars in
-accreting low mass X-ray binaries rarely have rotation rates below $200~\mathrm{Hz}$.
+accreting low mass X-ray binaries rarely have rotation rates below $200~\mathrm{Hz}$ 
+\MarginPar{look into this?}.
 
@@
 propagating flame through the cool fuel.  In our original set of
-calculations (in \citep{flame_wave1}), in order to make the 
+calculations (in in \citet{flame_wave1}), in order to make the 
 simulations computationally feasible we artificially boosted the flame speed by adjusting the
@@
 faster than the nominal laminar flame speed.  We also used high
-rotation rates ($\geq 1000~\mathrm{Hz}$) to reduce the lateral lengthscale at which the Coriolis
+rotation rates to reduce the lateral lengthscale at which the Coriolis
 force balances the lateral flame spreading in order to reduce the size 
@@
 is carried out in all zones within the atmosphere
-with density $\rho > 2.5\times 10^4~\gcc$. We use an axisymmetric coordinate system, 
+with $\rho > 2.5\times 10^4~\gcc$. We use an axisymmetric coordinate system, 
 with the horizontal $r$-direction pointing along the surface of the star and the 
@@
 \end{equation}
-\noindent Here, $C_{99}(Q)$ is the set of grid cells with $Q$ values in the top percentile, 
-$Q(c_i)$ is the value of $Q$ in cell $c_i$, and $m(c_i)$ is the total mass in cell $c_i$. 
-Using $\repavg{Q}$ instead of simply taking the maximum of the quantity across the entire 
+\noindent Here, $C_{99}(Q)$ is the set of grid cells with $Q$ values in the top percentile, and 
+$Q(c_i)$ is the value of $Q$ in cell $c_i$ (with $m(c_i)$ being the total mass in cell $c_i$). By 
+using $\repavg{Q}$ instead of simply taking the maximum of the quantity across the entire 
 simulation domain allows us to track changes over the domain as a whole rather than at a single 
@@
 neutron star rotation rate. In Figure \ref{fig:time_series_enuc_500} and Figure 
-\ref{fig:time_series_enuc_1000}, we use $\dot{e}_\mathrm{nuc}$ measured at $50~\mathrm{ms}$ and $100~\mathrm{ms}$ to discern the horizontal extent of the flame at different rotation rates.
+\ref{fig:time_series_enuc_1000}, we use the $\dot{e}_\mathrm{nuc}$ measured at $50~\mathrm{ms}$ and $100~\mathrm{ms}$ to discern the horizontal extent of the flame at different rotation rates.
 Taking the edge at greatest radius of the bright teal/green region where the most significant energy generation is occurring as the leading edge of the flame in each plot, we see that the horizontal extent of the $1000~\mathrm{Hz}$ flame ($\tt{F1000}$) appears to be
@@
 potentially leading to insufficient confinement such that the temperature and density required for 
-ignition is not achieved. In this scenario, another source of confinement (e.g.\ magnetic fields, see \cite{art-2016-cavecchi-etal}) 
+ignition is not achieved. In this scenario, another source of confinement (e.g.\ magnetic fields, see \citep{art-2016-cavecchi-etal}) 
 would need to take over at lower rotation rates to allow a burst to occur, at least for the initial 
 model used here.
-Given that the size of our domain is $\sim 1~L_R$ for {\tt F250} (using Equation~\ref{eqn:rossby}), it is also possible that we simply cannot confine the flame sufficiently with our current domain width.
-We see in Figure~\ref{fig:flame_speeds_1} (discussed further in Section~\ref{ssec:rot_propagation}) that the {\tt F500} flame took longer to achieve steady propagation than the {\tt F1000} flame.
+{\color{blue} Given that the size of our domain is $\sim 1~L_R$ for {\tt F250} (using Equation~\ref{eqn:rossby}), it is also possible that we simply cannot confine the flame sufficiently with our current domain width.} We see in Figure~\ref{fig:flame_speeds_1} that the {\tt F500} flame took longer to achieve steady propagation than the {\tt F1000} flame.
 It may therefore also be that we did not run our simulation for long enough to see the {\tt F250} flame achieve the conditions required for ignition and steady propagation.
@@
 vortical motion, shown in the Figure~\ref{fig:uv} phase plots comparing $u$, the vertical $z$-velocity $v$ and $\dot{e}_{\rm 
-nuc}$. This most likely corresponds to the leading edge of the flame where fresh fuel is being entrained.
+nuc}$. This most likely corresponds to the leading edge of the flame where fresh fuel is being entrained. \MarginPar{not sure if we should keep this sentence about entrainment...}
 This feature is not developed in the $250~\mathrm{Hz}$ flame in Figure~\ref{fig:uv} (left panel); it could potentially develop 
-at later times (past the point at which we terminated our simulation), or the burning could just 
+at later times (past the point at which we terminated our simulation), or the flame could just 
 fizzle out and fail to ignite entirely.
@@
 	simulation. $v_0$ is the flame velocity in the $+r$ direction at $t 
-	= 0~\mathrm{ms}$ and $a$ is the acceleration of the flame. The initial flame 
+	= 0~\mathrm{ms}$. The initial flame 
 	velocities and accelerations are derived from a quadratic least-squares fit to each of the 
@@
 10^4~\mathrm{cm}$. A similar effect is also visible (to a lesser extent) in the $\tt{F500\_3E8}$ 
-plot in Figure \ref{fig:compare_500Hz_abar}, with a faint haze of material rising above the flame near the axis of symmetry. 
+plot in Figure \ref{fig:compare_500Hz_abar}, with a faint haze of material rising above the flame. 
 However, this is likely to be a numerical artefact rather than a true physical effect, and a result 
@@
 	\plotone{compare_500Hz_enuc_4e8.pdf}
-	\caption{\label{fig:4e8_stacked_enuc} Time series of the $500~\mathrm{Hz}$ run $\tt{F500\_4E8}$, with $T_{\mathrm{star}} = 4 \times 10^8~\mathrm{K}$, showing $\enucdot$.
+	\caption{\label{fig:4e8_stacked_enuc} Time series of the $500~\mathrm{Hz}$ run $\tt{F500\_4E8}$, with $T_{\mathrm{star}} = 4 \times 10^8~\mathrm{K}$, showing $\dot{e}_{nuc}$.
 This model burns so strongly that it is only run out to $40~\mathrm{ms}$; the snapshots shown here are at $T = 30~\mathrm{ms}$, $35~\mathrm{ms}$, and $40~\mathrm{ms}$.}
@@
 $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$. Due to the lack of a clear burning front in 
-$\tt{F500\_4E8}$, we do not analyze its propagation velocity and acceleration. As the initial 
+$\tt{F500\_4E8}$, we do not analyze its propagation speed and acceleration. As the initial 
 $T_{\mathrm{star}}$ is increased beyond $\sim 2\times 10^8$, the flame becomes greatly accelerated. The initial flame 
@@
     \caption{\label{fig:urho_hot}Phase plot of the horizontal $r$-velocity $u$, density $\rho$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$ (the latest time that $\tt{F500\_4E8}$ is run out to).
-The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here.}
+The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.
 \end{figure}
@@
     \caption{\label{fig:uv_hot}Phase plot of the horizontal $r$-velocity $u$, vertical $z$-velocity $v$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$.
-The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here.}
+The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.
 \end{figure}
@@
     \caption{\label{fig:abar_hot}Phase plot of the mean molecular weight $\bar{A}$, temperature $T$ and the nuclear energy generation rate $\dot{e}_{\rm nuc}$ for the $500~\mathrm{Hz}$ runs with $T_{\rm star}$ of $10^8~\mathrm{K}$, $3\times10^8~\mathrm{K}$ and $4\times10^8~\mathrm{K}$ ($\tt{F500}$, $\tt{F500\_3e8}$, and $\tt{F500\_4e8}$) at $t = 40~\mathrm{ms}$.
-The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here.}
+The run with $T_{\rm star} = 2\times10^8~\mathrm{K}$ ($\tt{F500\_2E8}$) closely resembles the cooler $\tt{F500}$ run, and is not shown here}.
 \end{figure}

### 2021-01-28 xrb2/paper.tex
add unit

--- 2021-01-27 xrb2/paper.tex
@@
 $\tt{F500\_4E8}$, we do not analyze its propagation velocity and acceleration. As the initial 
-$T_{\mathrm{star}}$ is increased beyond $\sim 2\times 10^8$, the flame becomes greatly accelerated. The initial flame 
+$T_{\mathrm{star}}$ is increased beyond $\sim 2\times 10^8~\mathrm{K}$, the flame becomes greatly accelerated. The initial flame 
 velocities and accelerations derived from a quadratic least-squares fit to each of the datasets, as 

### 2021-01-29 xrb2/paper.tex
tangential -> perpendicular

@@
 with the horizontal $r$-direction pointing along the surface of the star and the 
-vertical $z$-direction pointing tangential to the surface.
+vertical $z$-direction pointing perpendicular to the surface.
 

### 2021-01-29 xrb2/paper.tex
small edits throughout

@@
 accretion rates and weaker at higher accretion rates, due to the
-effect of neutrino losses \citep{Cumming2006,johnston:2019}, however at 
-very high accretion rates it has been observed that the entire atmosphere 
-simmers in place rather than a flame propagating across the surface \citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}. 
+effect of neutrino losses \citep{Cumming2006,johnston:2019}.
+On the other hand, at very high accretion rates the atmosphere is so heated that it simmers in place rather than forming a propagating flame \citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}.
 A shallow heating
@@
 In the following sections, we conduct a series of simulations at various rotation 
-rates and temperatures to investigate their effects on the flame. We find that 
+rates and crustal temperatures to investigate their effects on the flame. We find that 
 at lower rotation rates, the flame itself becomes harder to ignite. At higher 
@@
 
-We use the Castro hydrodynamics code \citep{castro,castro_joss} and
+We use the \castro\ hydrodynamics code \citep{castro,castro_joss} and
 the simulation framework introduced in \citet{flame_wave1}.  The
@@
 propagating flame through the cool fuel.  In our original set of
-calculations (in \citep{flame_wave1}), in order to make the 
+calculations \citep{flame_wave1}, in order to make the 
 simulations computationally feasible we artificially boosted the flame speed by adjusting the
@@
     \plotone{hot}
-    \caption{\label{fig:hot_profiles} Initial temperature and density
-    structure vs.\ height in the ``hot'' state.}
+    \caption{\label{fig:hot_profiles} Initial temperature structure (left panel) and density
+    structure (right panel) as a function of height in the ``hot'' state.}
 \end{figure}
@@
     \plotone{cold}
-    \caption{\label{fig:cool_profiles} Initial temperature and density
-    structure vs.\ height in the ``cool'' state.}
+    \caption{\label{fig:cool_profiles} Initial temperature structure (left panel) and density
+    structure (right panel) as a function of height in the ``cool'' state.}
 \end{figure}
@@
 runs, most of the burning for each of the simulations occurs in a high density region $\rho > 3 
-\times 10^5 {\rm g\,cm^{-3}}$. The fluid in this dense, high energy generation region undergoes 
+\times 10^5~{\rm g\,cm^{-3}}$. The fluid in this dense, high energy generation region undergoes 
 vortical motion, shown in the Figure~\ref{fig:uv} phase plots comparing $u$, the vertical $z$-velocity $v$ and $\dot{e}_{\rm 
@@
 The Figure~\ref{fig:abar_hot} $\bar{A}$-$T$ phase plots of $\tt{F500}$ (left) and $\tt{F500\_3E8}$ (middle) also reflect these $\bar{A}$ features, with $\tt{F500\_3E8}$ reaching slightly higher $\bar{A}$ values.
-$\tt{F500\_3E8}$ also reaches higher $\enucdot$ values at the low end of the temperature range ($< 0.5 \times 10^9 \mathrm{K}$).
+$\tt{F500\_3E8}$ also reaches higher $\enucdot$ values at the low end of the temperature range ($< 0.5 \times 10^9~\mathrm{K}$).
 There appear to be more causally connected regions across a range of $\bar{A}$ at low temperatures for $\tt{F500\_3E8}$ than for $\tt{F500}$, suggesting that the higher $\enucdot$ for $\tt{F500\_3E8}$ generates burning in certain reaction pathways that are not present in the cooler $\tt{F500}$ run.
@@
 70~\mathrm{ms}$, when the flame is most developed). Significant $\enucdot$ values for 
-all other runs are confined to the flame front, rather than spanning the entire domain.
+all runs other than $\tt{F500\_4E8}$ are confined to the flame front, rather than spanning the entire domain.
 
@@
 
-The $u$-$\rho$ phase plot in Figure~\ref{fig:urho_hot} shows that the horizontal velocity for the $T_{\rm star} = 4\times10^8$~K ($\tt{F500\_4e8}$; right panel) is significantly greater in magnitude than for the cooler $\tt{F500\_3e8}$ (middle panel) and $\tt{F500}$ (left panel) runs.
-The $u$-$v$ plot in Figure~\ref{fig:uv_hot} further demonstrates this, as again we see that the magnitude of the $\tt{F500\_4e8}$ flame's vertical $z$-velocity $v$ is also significantly larger.
-This motion therefore most likely corresponds to the vortical motion rather than to the propagation of the flame itself, something which is further supported by the fact that the velocity magnitude is two to three orders of magnitude greater than the flame speeds listed in Table~\ref{table:flame_speeds}.
+In both the $u$-$\rho$ phase plot in Figure~\ref{fig:urho_hot} and the $u$-$v$ phase plot in Figure~\ref{fig:uv_hot}, the horizontal $u$-velocity for the $\tt{F500\_4e8}$ (right panel) burning region is slightly larger than that of the cooler runs.
+Additionally, the cooler runs' $u$-velocities are two to three orders of magnitude greater than the flame speeds listed in Table~\ref{table:flame_speeds}.
+These phase plot velocities therefore most likely correspond to vortical motion in the turbulent burning vortices rather than the propagation of a flame itself.
+The vertical $z$-velocity $v$ in Figure~\ref{fig:uv_hot} further suggests that $\tt{F500\_4e8}$ undergoes increased vortical motion, as we see that the velocity magnitude in the burning region is significantly larger than in the lower temperature runs.
 Also of interest in these plots is that the coolest run does not seem to have developed the characteristic vortex structure at the burning front (i.e.

### 2021-01-29 xrb2/paper.tex
Corp -> Corporation
T_star -> T_hi
luminosity -> limit

@@
 \author[0000-0003-0439-4556]{M.\ P.\ Katz}
-\affiliation{NVIDIA Corp}
+\affiliation{NVIDIA Corporation}
 
@@
 models by varying these parameters.  All models have the same
-peak temperature in the hot model, $T_\mathrm{star} + \delta T$.
+peak temperature in the hot model, $T_\mathrm{hi} + \delta T$.
 
@@
 We saw that if the crust temperature was too high, then instead of a flame the entire atmosphere would burn steadily.
-This is reminiscent of what is seen for neutron stars with accretion rates exceeding the Eddington luminosity.
+This is reminiscent of what is seen for neutron stars with accretion rates exceeding the Eddington limit.
 

### 2021-01-29 xrb2/paper.tex
reaction pathways -> burning tracjectories

@@
 The mean molecular weight $\bar{A}$ within each of our simulations seems to grow along defined tracks confined to certain temperatures $T$, as shown in the Figure \ref{fig:abar} phase plots.
-We believe that the tracks in the plot correspond to reaction pathways in the reaction network.
+We believe that the tracks in the plot correspond to burning trajectories in the reaction network.
 Comparing Figure \ref{fig:abar} to \citet{flame_wave1}, these tracks are much more neat and clearly defined.
@@
 $\tt{F500\_3E8}$ also reaches higher $\enucdot$ values at the low end of the temperature range ($< 0.5 \times 10^9~\mathrm{K}$).
-There appear to be more causally connected regions across a range of $\bar{A}$ at low temperatures for $\tt{F500\_3E8}$ than for $\tt{F500}$, suggesting that the higher $\enucdot$ for $\tt{F500\_3E8}$ generates burning in certain reaction pathways that are not present in the cooler $\tt{F500}$ run.
+There appear to be more causally connected regions across a range of $\bar{A}$ at low temperatures for $\tt{F500\_3E8}$ than for $\tt{F500}$, suggesting that the higher $\enucdot$ for $\tt{F500\_3E8}$ generates burning in certain burning trajectories that are not present in the cooler $\tt{F500}$ run.
 Note that, although $\tt{F500\_3E8}$ is the hottest run with a modified initial temperature configuration to form a distinct flame front, the highest rotation $\tt{F1000}$ flame actually reaches higher temperatures (see Figure~\ref{fig:maxima}, left panel) as well as higher $\bar{A}$ (see Figure~\ref{fig:time_series_1000}).
@@
 Not only is $\tt{F500\_4E8}$ significantly hotter overall compared to all other runs, but there is a large causally connected region across a wide range of $\bar{A}$.
-This indicates that the hotter temperature has facilitated significant burning in reaction pathways that were not favored at the lower temperatures.
-The reaction pathways are also very disrupted for $\tt{F500\_4E8}$ compared to the cooler runs, suggesting that the hotter temperature leads to more vigorous mixing.
+This indicates that the hotter temperature has facilitated significant burning in burning trajectories that were not favored at the lower temperatures.
+The burning trajectories are also very disrupted for $\tt{F500\_4E8}$ compared to the cooler runs, suggesting that the hotter temperature leads to more vigorous mixing.
 Indeed, this appears to be the case looking at Figure~\ref{fig:4e8_stacked_abar} and the right panel of Figure~\ref{fig:uv_hot} (discussed further in section~\ref{ssec:temp_prop}).
-The disrupted reaction pathways resemble those found in Figure \ref{fig:abar} for the highest rotation $\tt{F1000}$ run, though they are even more dramatically disrupted for $\tt{F500\_4E8}$.
+The disrupted burning trajectories resemble those found in Figure \ref{fig:abar} for the highest rotation $\tt{F1000}$ run, though they are even more dramatically disrupted for $\tt{F500\_4E8}$.
 $\tt{F500\_4E8}$ is also noticeably hotter than $\tt{F1000}$, even though it is run for significantly less time ($t = 40~\mathrm{ms}$ vs $t = 100~\mathrm{ms}$).

### 2021-01-29 xrb2/paper.tex
add clarifying sentence for F500 phase plot

@@
 The vertical $z$-velocity $v$ in Figure~\ref{fig:uv_hot} further suggests that $\tt{F500\_4e8}$ undergoes increased vortical motion, as we see that the velocity magnitude in the burning region is significantly larger than in the lower temperature runs.
-Also of interest in these plots is that the coolest run does not seem to have developed the characteristic vortex structure at the burning front (i.e.
+Also of interest in these plots is that the coolest run shows high velocity material comprising both high and low $\dot{e}_{\rm nuc}$ with horizontal velocity asymmetry, both of which suggest the burning vortex in the coolest case is less well defined than at higher temperatures.
+The coolest run thus does not seem to have developed the characteristic vortex structure at the burning front (i.e.
 where $\dot{e}_{\rm nuc}$ is greatest) that can be clearly seen for the hotter runs at this time ($40~\mathrm{ms}$).

### 2021-02-14 strang/paper.tex
new paper stub

--- 2021-01-07 simplified_sdc_intro/paper.tex
@@
 %\title{A Simplified Spectral Deferred Correction Method for Coupling Hydrodynamics with Reaction Networks and Nuclear Statistical Equilibrium}
-\title{An Improved Method for Coupling Hydrodynamics with Reaction Networks and Nuclear Statistical Equilibrium}
+\title{Practical Effects of Integrating Temperature / Energy with Strang Split Reactions}
 
 %\shorttitle{A Simplified SDC Method}
-\shorttitle{Hydro/Reaction Coupling with NSE}
+\shorttitle{Temperature Integration with Strang}
 
@@
 \begin{abstract}
-Reacting astrophysical flows can be challenging to model because of
-the difficulty in accurately coupling hydrodynamics and reactions.
-This can be particularly acute at high temperatures where nuclear
-statistical equilibrium is established.  We develop a simplified
-approach based on spectral deferred corrections (SDC) coupling of explicit
-hydrodynamics and stiff reaction sources as an alternative to operator
-splitting or the more comprehensive SDC approach we demonstrated
-previously.  We apply the new method to some example problems and show
-how to modify it to work with a hybrid network consisting of a
-reaction ODE system and a table for nuclear statistical equilibrium.
-This is all done in the framework of the \castro\ hydrodynamics code,
-and all algorithm implementations are freely available.
+Folks should integrate energy with their reactions!
 \end{abstract}
@@
 
-Reactive flow can be challenging to model when changes in the
-nuclear abundances due to reactions are integrated over hydrodynamic
-time scales.  Traditional methods of coupling
-hydrodynamics and reactions used in astrophysics use operator
-splitting---each physical process acts on the output of the previous
-process in alternating fashion.  This makes it easy to add physics in a modular way to a
-simulation code, but competition between physical processes can cause
-the coupling to breakdown.  These splitting errors can lead to loss of
-accuracy and further time step limitations.
-
-A particularly difficult phase of evolution to model is the nuclear
-statistical equilibrium that sets in with helium burning for
-temperatures in excess of $4\times 10^9~\mathrm{K}$.  Physically, the
-forward and reverse rates of reaction should balance leading to an equilibrium.
-With operator splitting, an NSE region will have a large positive flow
-through the network in a zone in one step followed by a large negative
-flow over the next timestep, as the code struggles to produce an
-equilibrium.  These large changes in abundances (and large alternately
-positive and negative energy generation rates) can be a challenge for
-a code.  The easiest way to improve the coupling is to cut the
-timestep, but this makes simulations prohibitively expensive.
-Sometimes the burning is simply halted on a zone-by-zone basis when
-NSE conditions are reached (e.g., as in \citealt{hedet}).
-Alternately, at high temperatures, a reaction network can be replaced
-with a table of NSE abundances and the zone's composition set through
-table look-ups (e.g.\ \citealt{ma:2013}) \MarginPar{need more here}
-
-In \cite{castro_sdc}, we developed second- and fourth-order accurate
-methods in space and time for coupling hydrodynamics and nuclear
-reaction networks based on spectral deferred corrections (SDC)
-methods, and demonstrated the method in a variety of test problems.
-The approach presented here is considerably simpler from a temporal
-integration standpoint, but allows is to incorporate modifications
-for regions in nuclear statistical equilibirum.
-Furthermore, we reuse the main
-CTU hydrodynamics construction and a largely similar ODE integration
-scheme, making this method easier to add to existing simulation codes.
-It also extends to adaptive mesh refinement with subcycling in a
-straightforward manner, avoiding the complications described in
-\citep{mccorquodalecolella} needed to fill ghost cells when using
-method-of-lines integration.  However, it is restricted to
-second-order accuracy overall.  We term this algorithm the
-``simplified SDC method''.  We use the \castro\ hydrodynamics code
-\citep{castro} for all of our numerical experiments, and all of the
-code to reproduce the results in this paper are in the \castro\ github
-repository\footnote{\url{https://github.com/amrex-astro/Castro/}}.
-
 \section{Numerical Methodology}
 
-We solve the Euler equations for compressible, reacting flow.  For ease
-of exposition we describe the one-dimensional case;
-multidimensional extensions are a straightforward modification to
-include the CTU hydrodynamics scheme.
-The system appears as:
-\begin{equation}
-\frac{\partial \Uc_t}{\partial t}  + \frac{\partial \Fb^\xv (\Uc)}{\partial x} = \Sc(\Uc)
-\end{equation}
-where
-\begin{equation}
-\Uc = \left ( \begin{array}{c}
-           \rho \\
-           \rho X_k \\
-           \rho \alpha_l \\
-           \rho u \\
-           \rho E \end{array}\right )
-\end{equation}
-are the conserved fluid quantities: mass density, $\rho$, velocity,
-$u$, specific total energy, $E$, and nuclear species mass fractions,
-$X_k$, and auxiliary composition variables, $\alpha_l$.  The mass fractions
-are constrained to sum to 1, $\sum_k X_k = 1$, but no such constraint
-exists on the auxiliary variables.  The specific total energy relates
-to the specific internal energy as $E = e + u^2/2$.  The corresponding
-fluxes are
-\begin{equation}
-\Fb^\xv (\Uc) = \left ( \begin{array}{c}
-   \rho u \\
-   \rho X_k u \\
-   \rho \alpha_l u \\
-   \rho u u + p \\
-   \rho u E + u p
-   \end{array} \right )
-\end{equation}
-Here the pressure, $p$, enters, and is found via the equation of state.  For a system
-where the composition is completely specified by the mass fractions, $X_k$, the equation
-of state would take the form:
-\begin{equation}
-p = p(\rho, X_k, e)
-\end{equation}
-However, as we'll see shortly, when we use the NSE table, we are using
-the results of a much larger network then can be represented by the
-mass fractions, and in this case, we rely on the auxiliary composition
-variables, $\alpha_l$, to describe the state of the composition, and
-our EOS has the form:
-\begin{equation}
-p = p(\rho, \alpha_l, e)
-\end{equation}
-
-We decompose the source terms, $\Sc(\Uc)$, into
-hydrodynamical sources, $\Hb$ (like gravity), and reactive sources,
-$\Rb$,
-\begin{equation}
-  \Sc(\Uc) = \Hb(\Uc) + \Rb(\Uc).
-\end{equation}
-\MarginPar{Detail out what is $\Hb(\Uc)$?}
- The reactive sources take the form:
-\begin{equation}
-  \Rb(\Uc) = \left ( \begin{array}{c}
-     0 \\
-     \rho \omegadot(X_k) \\
-     \rho \omegadot(\alpha_l) \\
-     0 \\
-     \rho \dot{S}
-  \end{array} \right )
-\end{equation}
-where $\omegadot(X_k)$ is the creation rate for species $k$, $\omegadot(\alpha_l)$ is the creation rate for
-auxiliary composition variable $l$, and $\dot{S}$
-is the energy generation rate per unit mass.
-
-Sometimes it is preferable to work with the primitive variables,
-\begin{equation}
-\qb = \left ( \begin{array}{c}
-  \rho \\
-  X_k \\
-  \alpha_l \\
-  u \\
-  p \\
-  (\rho e) \\
-\end{array} \right )
-\end{equation}
-Here, the system appears
-as:
-\begin{equation}
-\qb_t + \Ab^\xv(\qb) \qb_x  = \Sq
-\end{equation}
-with the matrix $\Ab^\xv$ giving the coefficients of the spatial derivatives
-of the primitive variables:
-\begin{equation}
-\Ab^\xv(\qb) = \left ( \begin{array}{cccccc}
-    u & 0 & 0 & \rho & 0 & 0 \\
-    0 & u & 0 & 0    & 0 & 0 \\
-    0 & 0 & u & 0    & 0 & 0 \\
-    0 & 0 & 0 & u    & 1/\rho & 0 \\
-    0 & 0 & 0 & \Gamma_1 p & u & 0 \\
-    0 & 0 & 0 & \rho h & 0 & u
-  \end{array} \right )
-\end{equation}
-where $\Gamma_1$ is an adiabatic index,
-$\Gamma_1 = d\log p/d\log\rho|_s$ at constant entropy.
-Note, the primitive state has two thermodynamic quantities, $p$
-and $(\rho e)$, to more efficiently handle the general equation of
-state in the Riemann solver, as described in \citet{castro}, but
-alternate formulations are possible \citep{colellaglaz:1985}.
-The source term vector, $\Sq$, can again be decomposed into hydrodynamic
-sources (now in terms of the primitive variables) and reaction terms,
-\MarginPar{Can we use $\Sc(\qb)$ instead of $\Sq$ and $\Hb(\qb)$ instead of $\Sqhydro$ to keep notation consistent with the conserved variables, since you are already doing that with $\Rb(\qb)$?}
-\begin{equation}
-  \Sq = \Sqhydro + \Rb(\qb)
-\end{equation}
-with
-\begin{equation}
-\label{eq:prim_sources}
-\Rb(q) = \left ( \begin{array}{c}
-     0 \\
-     \omegadot(X_k) \\
-     \omegadot(\alpha_l) \\
-     0 \\
-     \Gamma_1 p \sigma \Sdot \\
-     \rho \Sdot
-   \end{array} \right )
-\end{equation}
-where
-\begin{equation}
-\sigma \equiv \frac{\partial p/\partial T |_\rho}{\rho c_p \partial p/\partial \rho |_T}
-\end{equation}
-and $c_p$ is the specific heat at constant pressure, $c_p = \partial
-h/\partial T |_p$.  A derivation of this source for the pressure
-equation can be found in \cite{ABNZ:III}.  We note that this source is
-algebraically identical to that shown in Eq.~25 of \cite{castro}.
-
-We use the \castro\ hydrodynamics code~\cite{castro}, together with
-corner transport upwind piecewise parabolic method
-hydrodynamics~\citep{millercolella:2002}.  This is a finite-volume
-method that uses characteristic tracing to predict a time-centered
-flux through the interfaces of the grid zones.  For pure
-hydrodynamics, these time-centered fluxes result in second-order
-accurate time integration.  With reactions, we want to couple
-hydrodynamics and the reaction sources to second order.  Nuclear
-reaction sources are stiff, and need to be integrated using implicit
-methods for stabilty.  Operator splitting (e.g., Strang) is traditionally employed
-here, and is used as a benchmark for comparison in this paper.
-
-\subsection{Strang Splitting}
-
-In Strang splitting, we first integrate the system with reactions terms only (no advection)
-over $\Delta t/2$, then integrate the advection terms only (no reactions) over $\Delta t$,
-and finally integrate the reaction terms only over $\Delta t/2$.
-
-In the absence of advective terms, our reaction system appears as just
-$d\Uc/dt = \Rb(\Uc)$, or:
-\begin{align}
-\odt{\rho} & = 0 \\
-\odt{(\rho X_k)} &= \rho \omegadot(X_k) \\
-\odt{(\rho \alpha_l)} &= \rho \omegadot(\alpha_l) \\
-\odt{(\rho u)} &= 0 \\
-\odt{(\rho E)} &= \rho \Sdot
-\end{align}
-We can write the energy equation as:
-\begin{equation}
-\odt{(\rho E)} = \odt{(\rho e)} + \odt{K} = \rho \Sdot
-\end{equation}
-where $K$ is the kinetic energy, $K = |u|^2/2$.  Since the density and velocity
-are unchanged by reactions, we consider the following:
-\begin{equation}
-\odt{(\rho e)} = \rho \odt{e} = \rho \Sdot
-\end{equation}
-The reaction rates are typically expressed as $\omegadot_k(\rho, T, X_k)$
-when we evolve this system.  This requires an EOS inversion to get $T$ from $e$ each
-time we need to evaluate the reactive terms.
-We choose to instead integrate the temperature.
-We can write the energy equation as:
-\begin{equation}
-\rho \odt{e} = \rho \left [ \underbrace{\left . \frac{\partial e}{\partial T} \right |_\rho}_{c_v}
-                            \odt{T} +
-                            \left . \frac{\partial e}{\partial \rho} \right |_T
-                            \cancelto{0}{\odt{\rho}} \;\; \right ] = \rho \Sdot
-\end{equation}
-reducing to
-\begin{equation}
-\label{eq:strang:T}
-\odt{T} = \frac{1}{c_v} \Sdot
-\end{equation}
-We can alternately derive this from an enthalpy equation, which is the
-correct formulation when we are at constant pressure, and we get the
-same temperature equation but with $c_v$ replaced by $c_p$.
-Technically this requires an EOS call to get $c_x$, but if the
-specific heat varies slowly, we can freeze it during the reaction
-solve or use a Taylor expansion in $T$ to capture slow evolution.
-
-We also typically integrate mass fraction itself, instead of partial
-densities:
-\begin{equation}
-\label{eq:strang:X}
-\odt{X_k} = \omegadot_k
-\end{equation}
-We integrate Equations (\ref{eq:strang:T}) and (\ref{eq:strang:X}) using
-an implicit ODE solver designed for stiff systems of
-equations, VODE~\citep{vode}.
-
-With Strang splitting, the state can drift significantly off of the
-smooth solution to the coupled reactive hydrodynamics equations, as
-shown graphically in \cite{astronum:2018} using an earlier version of
-the present algorithm.
-
-A variation on Strang splitting called (re-)balanced splitting was
-developed in \citet{speth:2013}.
-
-\subsection{Timestep Limiters and Retry Mechanism}
-
-Since this method is based off of the CTU hydrodynamics scheme, it
-benefits from the larger timestep that method can take (when done with
-full corner coupling, the advective CFL condition is unity)
-as compared to a method-of-lines approach (see
-\citealt{ppmunsplit}).  In addition to the standard CFL timestep
-limiter for explicit hydrodynamics, \castro\ can also enforce
-timestep limiters based on the energy generation or abundance changes
-over a timestep: \MarginPar{do we enforce this in the NSE region?}
-\begin{align}
-\label{eq:dt:nuce}
-\Delta t &\le f_e\, \min_{i} \left\{\frac{e_{i}}{\dot{S}_{i}}\right\} \\
-%
-\label{eq:dt:nucX}
-\Delta t &\le f_X\, \min_{i} \left\{\min_{k,X_k > \epsilon_X}\frac{{X_k}_{i}}{{\omegadot(X_k)_i}}\right\}
-\end{align}
-where $i$ is the zone index and $f_e$ and $f_X$ are runtime parameters
-used to control the allowed change, and only species for which $X_k > \epsilon_X$ are considered.
-
-\castro\ has the ability to reject a timestep if it detects a failure
-and retry with smaller timesteps (subcycling to make up the original
-required timestep).  Among the conditions that can trigger this are
-density falling below zero during advection, the ODE integration
-failing to converge in the implicit solve, or violation of one of the
-timestep limiters during the step.  This means that equations
-(\ref{eq:dt:nuce}) and (\ref{eq:dt:nucX}) are not reactive, but instead
-guaranteed to be met throughout the simulation, because the step is
-rejected if they are violated.  This contrasts to similar approaches
-used in other codes where conditions \ref{eq:dt:nuce} and
-\ref{eq:dt:nucX} are used to restrict the next timestep size, but the
-update over the current step is kept, which already violated the
-constraints. \MarginPar{would be nice to have a reference here} The retry mechanism in \castro\ works with both the
-Strang and simplified-SDC integration scheme.
-
-\subsection{Spectral Deferred Corrections}
-
-The basic idea of spectral deferred corrections is to express the
-update as an integral and divide the time-update into a number of
-discrete time nodes.  The integral is then approximated using a
-quadrature rule over these time nodes and low order approximations are
-used to update the state from one time node to the next.  The method
-uses iteration to successively improve the solution and the ultimate
-accuracy is determined by the quadrature method used to evaluate the
-integral, which is done using an iteratively-lagged solution.
-
-We start by writing our update as:
-\begin{equation}
-\Uc^{n+1} = \Uc^n + \int_t^{t+\Delta t} \left [ \Advs{\Uc} + \Rb(\Uc) \right ] dt,
-\end{equation}
-where $\Advs{\Uc} =  -{\partial \Fb^\xv (\Uc)}/{\partial x} + \Hb(\Uc)$.
-For the simplified SDC method we explore here, we will make the
-advective term piecewise constant in time, using the value at the
-midpoint in time, to achieve second-order accuracy, giving us:
-\begin{equation}
-\label{eq:integral:simplesdc}
-\Uc^{n+1} = \Uc^n + \int_t^{t+\Delta t} \left \{ \Adv{\Uc}^{n+1/2} + \Rb(\Uc) \right \} dt
-\end{equation}
-This is the approach first shown in \citep{SDC-old}.  A
-version of this was also implemented in the \maestro\ low-Mach number
-hydrodynamics code some time ago.  The compressible version for use in
-\castro\ is slightly different due to the need to do some operations
-on the conserved variable state and some on the primitive variable
-state.  We describe the application of this to the equations of
-compressible hydrodynamics below.
-
-\begin{itemize}
-
-\item {\em Initialization}
-
-\label{sec:initialization}
-
-  \begin{itemize}
-  \item We need an approximation of how much the reactions alone
-    changed the primitive variable state over the timestep, which we
-    will denote $\Ic_q$.  Since we do not have any information about
-    the current timestep in the first iteration,
-    we use the value from the last iteration of the previous timestep:
-    \begin{equation}
-      \Ic^{n+1/2,(0)}_{\qb} = \Ic^{n-1/2,(\smax)}_{\qb}
-    \end{equation}
-    \MarginPar{changed to 1-based $k$ iteration for aesthetics?  I can switch it back if you prefer ``-1'' to be the starting point instead of ``0''.}
-  %% \item Solve the Poisson problem for the initial gravitational potential:
-  %%   \begin{equation}
-  %%     \nabla^2 \Phi^n = 4\pi G \rho^n
-  %%   \end{equation}
-
-  %% \item Set the first guess at the new time potential as
-  %%   $\Phi^{n+1,(0)} = \Phi^n$.
-
-  \end{itemize}
-
-\item {\em Iterate}
-
-  Iterate from $k = 1, \ldots, \smax$.  For second-order accuracy,
-  $\smax = 2$ is sufficient.  In addition to denoting the time-level
-  with a superscript (like $n$ or $n+1$), we'll use a second subscript
-  in parentheses to keep track of the iteration.  A single iteration
-  starts with $\Uc^n$ and results in the new time-level state for that
-  iteration, $\Uc^{n+1,(k)}$.
-
-  \begin{itemize}
-  \item {\em Create the advective update term, $\Adv{\Uc}^{n+1/2,(k)}$}
-
-    \begin{itemize}
-    \item convert $\Uc \rightarrow \qb$.  This is an algebraic transformation,
-      but will require the EOS.
-
-    \item predict $\qb$ to the interfaces at $t^{n+1/2}$ using the CTU PPM
-      method.  The source terms, $\Sq$, used in the prediction are:
-      \begin{equation}
-        \Sq = \Sqhydro^{n+1/2} + \Ic_\qb^{n+1/2,(k-1)}
-      \end{equation}
-      Here we use the iteratively lagged integrals of the primitive variable
-      terms accounting only for reactions, $\Ic_\qb^{n+1/2,(k-1)}$, as the
-      reactive source.  This is in contrast to Strang-splitting, where no
-      explicit reactive source terms are seen by the hydrodynamics update.
-      Any hydrodynamic source terms are time-centered
-      using the previous iteration:
-\MarginPar{Formally both $t^n$ or $t^{n+1/2}$ sources give you 2nd-order since they are CTU predictor source terms.  I've never found this type of iteration for this term to amount to any benefit}
-      \begin{equation}
-        \Sqhydro^{n+1/2} = \frac{1}{2} \left ( \Sqhydro^n + \Sqhydro^{n+1,(k-1)} \right )
-      \end{equation}
-
-    \item solve the Riemann problem at each interface to get a unique
-      conserved state on each interface, $\Uc^{n+1/2,(k)}_{i+1/2}$
-
-    \item construct the advective update terms, $\Advt{\Uc}^{n+1/2,(k)}_{i}$,
-      by first ignoring the hydrodynamics sources,
-      \begin{align}
-        \Advt{\Uc}^{n+1/2,(k)}_{i} =
-          &- \frac{\Fb^\xv(\Uc^{n+1/2,(k)}_{i+1/2}) - \Fb^\xv(\Uc^{n+1/2,(k)}_{i-1/2})}{\Delta x}
-      \end{align}
-    Now the conservative hydrodynamics source terms are computed by first updating to the
-    new state with advection and the old-time source term applied for the full $\Delta t$ as:
-\MarginPar{I'm confused what happens with the $\Uc^{\star\star}$ state; I don't see it below.  Is The $\Shydro^{n+1/2}$ used below suppose to be from the $\Uc^{\star\star}$ state?}
-    \begin{equation}
-      \Uc^{\star\star} = \Uc^n + \Delta t \Advt{\Uc}^{n+1/2,(k)} + \Delta t \Hb(\Uc^n)
-    \end{equation}
-    %% first updating the density to the new state with advection only as:
-    %% \begin{equation}
-    %%   \rho^{\star\star} = \rho^n + \Delta t \Advt{\rho}^{n+1/2,(k)}
-    %% \end{equation}
-    %% then constructing the momentum source term:
-    %% \begin{equation}
-    %%   {\bf S}_{\rho\Ub}^{n+1/2} = \frac{1}{2} (\rho^n \gb^n + \rho^{\star\star} \gb^{n+1,(k-1)})
-    %% \end{equation}
-    %% Then updating the momentum with only advection as:
-    %% \begin{equation}
-    %%   (\rho \Ub)^{\star\star} = (\rho \Ub)^n + \Delta t \Advt{\rho\Ub}^{n+1/2,(k)} + \Delta t {\bf S}_{\rho\Ub}^{n+1/2}
-    %% \end{equation}
-    %% and finally constructing the energy source:
-    %% \begin{equation}
-    %%   S_{\rho E}^{n+1/2} = \frac{1}{2} \left [ (\rho\Ub)^n \cdot \gb^n + (\rho\Ub)^{\star\star} \cdot \gb^{n+1,(k-1)}
-    %%     \right ]
-    %% \end{equation}
-    We then evaluate the source terms with $\Uc^{\star\star}$ and
-    correct the advective term so that we have a time-centered
-    source: \MarginPar{are we actually doing this correctly for SDC?}
-    The final advective update term is then\footnote{For a source like gravity, this update can be done first for $\rho$ and then define the new momentum source using $\rho^{\star\star}$ and likewise for energy}:
-    \begin{equation}
-      \Adv{\Uc}^{n+1/2,(k)}_{i} = \Advt{\Uc}^{n+1/2,(k)}_{i} +
-      \frac{\Delta t}{2} \left (\Shydro(\Uc^n) + \Shydro(\Uc^{\star\star}) \right )
-    \end{equation}
-
-    %% with
-    %% \begin{equation}
-    %%   \Shydro^{n+1/2} = \left ( \begin{array}{c}
-    %%                  0 \\ 0 \\ 0 \\
-    %%                 {\bf S}_{\rho\Ub}^{n+1/2} \cdot \ex \\
-    %%                 S_{\rho E}^{n+1/2} \end{array} \right )
-    %% \end{equation}
-
-    \MarginPar{if $\Shydro$ depended on the outcome of the burn, we might need to do things differently here}
-
-    \end{itemize}
-
-  \item {\em Update the System Using a Method of Lines Integration}
-
-    We update the state by doing the integral in equation
-    (\ref{eq:integral:simplesdc}).  Since we are approximating the
-    advective term as piecewise constant in time, we can simply use an
-    ODE integrator to integrate this, just as we do with the reaction
-    system in Strang splitting.  The difference here being that we are
-    integrating the conserved variables and the state sees the effect
-    of advection as we integrate the reactions.  So rather than use
-    $d\Uc/dt = \Rb(\Uc)$, the ODE form we use is
-    \begin{equation}
-      \odt{\Uc} = \Adv{\Uc}^{n+1/2,(k)} + \Rb(\Uc)
-    \end{equation}
-    This integration begins with $\Uc^n$ and results in $\Uc^{n+1,(k)}$.
-
-    As we are integrating this system we need to get the
-    temperature, $T$, for the rate evaluations.  We construct this
-    by subtracting the kinetic energy from the total energy to get
-    the specific internal energy, $e$, and then calling the equation
-    of state.
-
-    Our integrator also needs the Jacobian of the system, in terms of
-    the conserved variables.  This is different than the form of the
-    Jacobian usually used in reaction networks.  We describe the form
-    of the Jacobian in appendix \ref{sec:app:jacobian}.
-
-    Note that our advection terms are piecewise
-    constant \MarginPar{need to think about this more} approximations,
-    so when we integrate our system, their contributions will be
-    linear in time.  Since $(\rho X_k)$ sees a linear advective term,
-    $\rho$ does as well.  $(\rho u)$ will also have linear
-    contributions in time from advection.  Since the reactions don't
-    affect $\rho$ and $(\rho u)$, we can algebraically update these
-    using this piecewise linear behavior:
-    \begin{align}
-      \rho(t) &= \rho^n + \Advss{\rho}^{n+1/2,(k)} \, t \\
-      (\rho u)(t) &= (\rho u)^n + \Advss{\rho u}^{n+1/2,(k)} \, t
-    \end{align}
-
-  \item {\em Compute the Reactive Source Terms.}
-
-    We now seek the $\Ic$'s that capture the effect of just the
-    reaction sources on the state variables for the next iteration.
-    For the conserved quantities, these would simply be:
-    \begin{equation}
-      \Ic^{(k)}_{\Uc} = \frac{\Uc^{n+1,(k)} - \Uc^n}{\Delta t} - \Adv{\Uc}^{n+1/2,(k)}
-    \end{equation}
-    However, for our primitive variables, which are used in the
-    prediction, we need to construct the required source terms more
-    carefully.  We want:
-    \begin{equation}
-      \label{eq:Iq}
-      \Ic^{(k)}_{\qb} = \frac{\qb^{n+1,(k)} - \qb^n}{\Delta t} - \Adv{\qb}^{n+1/2,(k)}
-    \end{equation}
-    but we need the advective update for $\qb$, which we have not
-    constructed.  We note that $\Ic^{(k)}_\qb$ is an approximation to the integral of 
-    Eq.~\ref{eq:prim_sources} over the timestep.  Additionally, we cannot simply use the equation of
-    state on $\Ic^{(k)}_{\Uc}$ since this is a time-derivative and
-    does not represent a well-defined state in itself.  Instead, we
-    derive $\Ic^{(k)}_{\qb}$ via a multi-step process.  We first find
-    the conservative state as if it were updated only with advection:
-    \begin{equation}
-      \Uc^\star = \Uc^n + \Delta t \Adv{\Uc}^{n+1/2,(k)}
-    \end{equation}
-    and then construct the corresponding primitive variable state via an algebraic transform,
-    $\Uc^\star \rightarrow \qb^\star$.
-    This allows us to define the advective update for $\qb$ as:
-    \begin{equation}
-      \Adv{\qb}^{n+1/2,(k)} = \frac{\qb^\star - \qb^n}{\Delta t}
-    \end{equation}
-    Defining the primitive state corresponding to the fully-updated
-    conserved state via an algebraic transform, $\Uc^{n+1,(k)}
-    \rightarrow \qb^{n+1,(k)}$, we can construct $\Ic^{(k)}_{\qb}$ as given
-    in Equation (\ref{eq:Iq}).
-    Putting all of this together, we see:
-    \begin{equation}
-      \Ic^{(k)}_{\qb} = \frac{\qb^{n+1,(k)} - \qb^\star}{\Delta t}
-    \end{equation}
-
-  %% \item {\em Solve for the New Gravitational Potential.}
-
-  %%   We solve
-  %%   \begin{equation}
-  %%     \nabla^2 \Phi^{n+1,(k)} = 4\pi G \rho^{n+1,(k)}
-  %%   \end{equation}
-
-  \end{itemize}
-
-\end{itemize}
-
-\section{Reaction Networks and Nuclear Statistical Equilibrium}
-
-For this study we will use the 19 nuclei network containing
-\isot{H}{1}, \isot{He}{3}, \isot{He}{4}, \isot{C}{12}, \isot{N}{14},
-\isot{O}{16}, \isot{Ne}{20}, \isot{Mg}{24}, \isot{Si}{28},
-\isot{S}{32}, \isot{Ar}{36}, \isot{Ca}{40}, \isot{Ti}{44},
-\isot{Cr}{48}, \isot{Fe}{52}, \isot{Fe}{54}, \isot{Ni}{56}, protons
-(from photodisintegration), and neutrons.  This is based on the
-``aprox19'' network from \cite{aprox19} and originally described in
-\cite{Kepler}.
-We use a modified version of the VODE~\citep{vode} integrator---ported to C++ with checks added to the timestep rejection logic that ensure that the species mass fractions stay between 0 and 1\footnote{This modified version of VODE is available in our Microphysics repo: \url{https://github.com/starkiller-astro/Microphysics}.}.\MarginPar{discuss tolerances}  We combine this with a table that gives the nuclear
-statistical equilibrium (NSE) abundances in regions where the system
-is in NSE.  The NSE table was generated using a 127 nuclei reaction
-network and is the same as described in \cite{ma:2013}.  In our
-simulations, we carry all 19 isotopes in the main network in each zone
-and advect them in the hydrodynamics portion of the algorithm.  The
-composition of the larger 127 nuclei network is mapped into the 19
-isotopes we carry according to Table \ref{table:nuclei}.
-
-The NSE table requires: \MarginPar{need a statement about how the table was constructed and what NSE means}
-\begin{equation}
-  \label{eq:aux:ye}
-  Y_e = \sum_k \frac{Z_k X_k}{A_k}
-\end{equation}
-(where $A_k$ and $Z_k$ are the atomic weight and atomic number of nucleus $k$) and provides
-\begin{align}
-\label{eq:aux:abar}
-\bar{A} &= \left [ \sum_k \frac{X_k}{A_k} \right ]^{-1} \\
-\label{eq:aux:bea}
-\left (\frac{B}{A} \right ) &= \sum_k \frac{B_k X_k}{A_k}
-\end{align}
-where $B_k$ is the binding energy of nucleus $k$.  In our simulations,
-we store these 3 quantities as auxiliary data that is carried along
-with the rest of the fluid state in each zone.  The table also returns
-values of the mass fractions mapped onto the 19-isotopes we carry, $X_k$, and the time-derivative of $Y_e$, $dY_e/dt$.
-We use the notation:
-\begin{equation}
-\nse{\rho,T,Y_e} \rightarrow \bar{A}, X_k, (B/A), dY_e/dt
-\end{equation}
-to represent the NSE table call and its inputs and outputs.
-
-We can derive evolution equations for each of these composition quantities as:
-\begin{align}
-\frac{DY_e}{Dt} &= \sum_k \frac{Z_k}{A_k} \frac{DX_k}{Dt} = \sum_k \frac{Z_k}{A_k} \omegadot_k \\
-\frac{D\bar{A}}{Dt} &= -\bar{A}^2 \sum_k \frac{1}{A_k} \frac{DX_k}{Dt} = -\bar{A}^2 \sum_k \frac{1}{A_k} \omegadot_k \\
-\frac{D}{Dt} \left (\frac{B}{A} \right ) &= \sum_k \frac{B_k}{A_k} \frac{DX_k}{Dt} = \sum_k \frac{B_k}{A_k} \omegadot_k
-\end{align}
-For Strang split coupling of hydro and reactions, $DX_k/Dt = 0$,
-therefore each of these auxillary equations obeys an advection
-equation in the hydro part of the advancement.  In the SDC algorithm,
-there will be a reactive source (an $\Ic_q$) for each of these that is
-computed in the same manner as above.  We note that our NSE table
-provides the evolution of $Y_e$ due to reactions ($DY_e/Dt$) directly.
-
-The compositional quantities
-it carries, $\bar{A}$ and $Y_e$ are not representable from the 19
-isotopes we carry in the main network. For this reason, when we are
-using the NSE network, we always provide these two composition quantities as
-inputs to the EOS rather than using the $X_k$ directly.
-the EOS directly from the auxiliary state in each zone instead of
-using the $X_k$ directly.
-Our equation of state needs $\bar{Z}$ in addition
-to the auxiliary quantities, which is is easily computed as
-\begin{equation}
-\bar{Z} = \bar{A} Y_e
-\end{equation}
-
-\subsection{Initialization}
-
-We initialize the mass fractions, $X_k$, and then compute the
-electron fraction, $Y_e$, using (\ref{eq:aux:ye}).
-For the two other composition
-quantities we carry, $\bar{A}$, and $(B/A)$, we need values that are
-consistent with the value of $Y_e$ and the nuclei stored in the NSE
-table.  Therefore, if the thermodynamic conditions put us in NSE
-(using the conditions defined below), then we obtain $\bar{A}$ and
-$(B/A)$ from the NSE table.  Otherwise, we compute these directly from
-$X_k$ using (\ref{eq:aux:abar}) and (\ref{eq:aux:bea}).
-
-\subsection{NSE condition}
-
-We treat a zone as being in NSE if the density and temperature exceed
-some threshold and the composition is mainly $\alpha$ and Fe-group
-nuclei, where the Fe-group nuclei are \isot{Cr}{48}, \isot{Fe}{52},
-\isot{Fe}{54}, and \isot{Ni}{56}.  The full condition is:
-\begin{align}
-\rho &> \rhonse \\
-T &> \tnse \\
-X(\isotm{C}{12}) &< \Anse \\
-X(\isotm{He}{4}) + \sum_{k \in \mathrm{Fe-group}} X_k &> \Bnse
-\end{align}
-Typical values are $\rhonse = 3\times 10^8~\gcc$, $T_\mathrm{nse} =
-3\times 10^9~\mathrm{K}$, and $\Anse = 0.12$ and $\Bnse=
-0.88$---these are the values used in \citet{ma:2013}.
-
-\subsection{Strang-split algorithm for NSE}
-
-For Strang splitting, the hydrodynamic and reactive substeps over
-the overall time-advancemenet scheme using the aprox19 + NSE network
-are as follows:
-\begin{itemize}
-
-\item {\em Hydrodynamics update}
-
-  At the beginning of each hydrodynamic update we have an input state,
-  $\Uc_{\rm in}$ that we wish to integrate over $\Delta t$ to obtain $\Uc_{\rm out}$
-  The hydrodynamics update proceeds as normal, but with an advection
-  equation for each of the auxiliary composition variables:
-  \begin{align}
-    \ddt{(\rho Y_e)} + \ddx{(\rho Y_e u)} &= 0 \\
-    \ddt{(\rho \bar{A})} + \ddx{(\rho \bar{A} u)} &= 0 \\
-    \ddt{[\rho (B/A)]} + \ddx{[\rho (B/A) u]} &= 0 
-  \end{align}
-
-\item {\em Reactive update}
-
-  At the beginning of each reactive update we have an input state,
-  $\Uc_{\rm in}$ that we wish to integrate over $\Delta t/2$ to obtain $\Uc_{\rm out}$.
-
-  \begin{itemize}
-
-    \item For a zone that is in NSE:
-
-      The goal is to update the composition and thermodynamics due to the 
-      change in the nuclei abundances over the (half-)timestep.  \citet{ma:2013}
-      uses a first-order in time difference to get the new composition state
-      and evaluates the energy release from the change in binding energy in the NSE
-      state.  They only apply a fraction (0.3) of the energy release to the internal
-      energy in a zone, to avoid a potential instability that can arise if too much
-      energy is added to a zone in a single timestep.  We prefer to deal with this
-      issue through the retry mechanism already built into \castro.
-
-      The approach we use begins with calling the NSE table with the input state:
-      \begin{equation}
-        \nse{\rho_\inp, T_\inp, (Y_e)_\inp} \rightarrow \bar{A}^\star, (X_k)^\star, (B/A)^\star, dY_e/dt
-        \end{equation}
-      We indicate with a $\star$ that most of those values are
-      provisional, and we will seek a better value in the corrector step.
-      We only update $Y_e$ from this call:
-      \begin{equation}
-        (Y_e)_\out = (Y_e)_\inp + \Delta t \frac{dY_e}{dt}
-      \end{equation}
-      and then we use the table again, but with the updated $Y_e$
-      (note that $\rho_{\rm out}=\rho_{\rm in}$ in the Strang reaction formulation):
-      \begin{equation}
-        \nse{\rho_\out, T_\inp, (Y_e)_\out} \rightarrow \bar{A}_\out, (X_k)_\out, (B/A)_\out, dY_e/dt
-      \end{equation}
-      In this formulation, we have not updated the temperature. \MarginPar{discuss this more, and perhaps switch to a mode where we try to predict T}
-      This corrects the composition so that it is consistent with the updated
-      $Y_e$.  We use these values $\bar{A}_\out$, $(X_k)_\out$, and $(B/A)_\out$ to then complete the update, computing the energy release, $\Sdot$ as:
-      \begin{equation}
-        \label{eq:nse_energy}
-        \Sdot = \left [ \left ( \frac{B}{A} \right )_\out -
-          \left ( \frac{B}{A} \right )_\inp \right ] N_A \frac{1}{\Delta t}
-      \end{equation}
-      \MarginPar{we could imagine looping over the EOS to get an updated T and then the table to get the updated $\bar{A}$, to see if that converges?}
-
-    \item For zones not in NSE:
-
-    \begin{itemize}
-    \item Integrate the full reaction network (Eqs.~\ref{eq:strang:T} and \ref{eq:strang:X}) as usual for Strang splitting
-    \item Update the aux quantities at the end of the burn using Eqs.~\ref{eq:aux:ye}, \ref{eq:aux:abar}, and \ref{eq:aux:bea} with the new mass fractions, $X_k$.
-    \end{itemize}
-  \end{itemize}
-\end{itemize}
-
-\subsection{SDC-NSE Coupling}
-
-With SDC evolution, when we are in NSE, we need to do the advective
-and reactive updates together.  With the tabulated NSE reaction
-values, we get instantaneous values of the reaction sources, but the only
-time derivative we have is for $Y_e$.  We will
-use a basic predictor-corrector update:
-\begin{equation}
-\Uc^\star = \Uc^n + \Delta t \Adv{\Uc}^{n+1/2} + \Delta t \left [\Rb (\Uc) \right ]^{n+1/2}
-\end{equation}
-Calling the NSE table on the starting state gives us:
-\begin{equation}
-\nse{\rho^n, T^n, (Y_e)^n} \rightarrow \bar{A}^n, (X_k)^n, (B/A)^n, (dY_e/dt)^n
-\end{equation}
-To get the reactive source, we need to use the iteratively-lagged time-centered information
-from $\Ic_q^{n+1/2,(k-1)}$.  We only need to update the internal energy and auxiliary data,
-since together with the density, that is all that is needed for the equation of state to get the
-updated temperature.    So our sources are:
-\begin{equation}
-R(\rho e) \approx \Ics_q(\rho e)
-\end{equation}
-and
-\begin{equation}
-R(\rho \alpha_l) \approx \rho^{n+1/2} \Ics_q(\alpha_l)
-\end{equation}
-where $\rho^{n+1/2} = (\rho^n + \rho^\star)/2$.
-
-Since we have $(dY_e/dt)^n$ from the table, we replace the source for $Y_e$ in the above
-$\Rb(\Uc^n)$ as: \MarginPar{it might be better if we had $\Ics_\Uc$}
-\begin{equation}
-R(\rho Y_e) = \rho^{n+1/2} (dY_e/dt)^n
-\end{equation}
-which gives an update for the electron fraction as:
-\begin{equation}
-(\rho Y_e)^\star = (\rho Y_e)^n + \Delta t \Advss{\rho Y_e}^{n+1/2} + \Delta t \rho^{n+1/2} (dY_e/dt)^n
-\end{equation}
-We can then get $\alpha_l^\star = (\rho \alpha_l)^\star/\rho^\star$.
-From this prediction, we get $\rho^\star$, $\alpha_l^\star$, and
-$e^\star$, which allows us to call the EOS and get an updated
-temperature, $T^\star$. \MarginPar{Andy suggests time-centering $dY_e/dt$}
-\begin{equation}
-T^\star = T(\rho^\star, \alpha_l^\star, e^\star)
-\end{equation}
-We can then again call the NSE table:
-\begin{equation}
-\nse{\rho^\star, T^\star, (Y_e)^\star} \rightarrow \bar{A}^{n+1,(k)}, (X_k)^{n+1,(k)}, (B/A)^{n+1,(k)}, (dY_e/dt)
-\end{equation}
-where we use the outputs to define the new values of the composition variables.
-Since the density and momenta don't have reactive sources, we just use the $\Uc^\star$ state
-as their new values:
-\begin{align}
-\rho^{n+1,(k)} &= \rho^\star \\
-(\rho u)^{n+1,(k)} &= (\rho u)^\star
-\end{align}
-We then compute the energy release as
-\begin{equation}
-       (\rho \Sdot)^{n+1/2} = \frac{\rho^{n+1,(k)} + \rho^n}{2}
-             \left [ \left ( \frac{B}{A} \right )^{n+1,(k)} -
-                     \left ( \frac{B}{A} \right )^n \right ] N_A \frac{1}{\Delta t}
-\end{equation}
-
-{\color{red}
-Note I tried this first, but it really doesn't work.  I'd like to understand better:
-\begin{equation}
-       (\rho \Sdot)^{n+1/2} = \left [ \rho^{n+1,(k)} \left ( \frac{B}{A} \right )^{n+1,(k)} -
-                            \rho^n \left ( \frac{B}{A} \right )^n \right ] N_A \frac{1}{\Delta }t
-\end{equation}
-}
-and then do the final update of the total energy:
-\begin{equation}
-(\rho E)^{n+1,(k)} = (\rho E)^n + \Delta t \Advss{\rho E}^{n+1/2} + \Delta t (\rho \Sdot)^{n+1/2}
-\end{equation}
+We are using Castro
 
@@
 
-\cite{castro_sdc} introduced a test problem for measuring convergence
-of a reacting hydrodynamic algorithm.  We run that same test here with the
-simplified SDC algorithm.
+We use the reacting convergence test problem introduced in \cite{castro_sdc}.
 
-Also explore: PLM vs PPM, 1 vs. 2 vs. 3 iterations
+\clearpage
 
 \begin{deluxetable}{lllllll}
-\tablecaption{\label{table:react_convert_strang} Convergence ($L_1$ norm) for the reacting convergence problem using Strang splitting.}
+\tablecaption{\label{table:react_convert_no_T} Convergence ($L_1$ norm) for the reacting convergence with no $T$ integration.}
+\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
+           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
+           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
+\startdata
+ $\rho$                      & $2.735 \times 10^{18}$  & 2.057  & $6.571 \times 10^{17}$  & 2.474  & $1.183 \times 10^{17}$  \\
+ $\rho u$                    & $6.729 \times 10^{26}$  & 2.416  & $1.260 \times 10^{26}$  & 2.699  & $1.941 \times 10^{25}$  \\
+ $\rho v$                    & $6.729 \times 10^{26}$  & 2.416  & $1.260 \times 10^{26}$  & 2.699  & $1.941 \times 10^{25}$  \\
+ $\rho E$                    & $2.513 \times 10^{35}$  & 2.226  & $5.372 \times 10^{34}$  & 2.169  & $1.195 \times 10^{34}$  \\
+ $\rho e$                    & $2.292 \times 10^{35}$  & 2.194  & $5.011 \times 10^{34}$  & 2.253  & $1.051 \times 10^{34}$  \\
+ $T$                         & $2.321 \times 10^{21}$  & 1.606  & $7.623 \times 10^{20}$  & 1.993  & $1.915 \times 10^{20}$  \\
+ $\rho X(\isotm{He}{4})$     & $2.806 \times 10^{18}$  & 2.028  & $6.880 \times 10^{17}$  & 2.432  & $1.275 \times 10^{17}$  \\
+ $\rho X(\isotm{C}{12})$     & $1.807 \times 10^{17}$  & 1.846  & $5.026 \times 10^{16}$  & 1.605  & $1.653 \times 10^{16}$  \\
+ $\rho X(\isotm{O}{16})$     & $1.905 \times 10^{14}$  & 1.500  & $6.737 \times 10^{13}$  & 1.439  & $2.484 \times 10^{13}$  \\
+ $\rho X(\isotm{Fe}{56})$    & $2.735 \times 10^{8}$   & 2.057  & $6.571 \times 10^{7}$   & 2.474  & $1.183 \times 10^{7}$   \\
+\enddata
+\end{deluxetable}
+
+\begin{deluxetable}{lllllll}
+\tablecaption{\label{table:react_convert_no_T} Convergence ($L_1$ norm) for the reacting convergence with $T$ integration but frozen $c_x$.}
 \tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
@@
 \begin{deluxetable}{lllllll}
-\tablecaption{\label{table:react_convert_strang} Convergence ($L_1$ norm) for the reacting convergence problem using Strang splitting with $T$ evolution disabled.}
+\tablecaption{\label{table:react_convert_no_T} Convergence ($L_1$ norm) for the reacting convergence with no $T$ integration and calling the EOS.}
 \tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
@@
 \startdata
- $\rho$                      & $2.735 \times 10^{18}$  & 2.057  & $6.571 \times 10^{17}$  & 2.474  & $1.183 \times 10^{17}$  \\
- $\rho u$                    & $6.729 \times 10^{26}$  & 2.416  & $1.260 \times 10^{26}$  & 2.699  & $1.941 \times 10^{25}$  \\
- $\rho v$                    & $6.729 \times 10^{26}$  & 2.416  & $1.260 \times 10^{26}$  & 2.699  & $1.941 \times 10^{25}$  \\
- $\rho E$                    & $2.513 \times 10^{35}$  & 2.226  & $5.372 \times 10^{34}$  & 2.169  & $1.195 \times 10^{34}$  \\
- $\rho e$                    & $2.292 \times 10^{35}$  & 2.194  & $5.011 \times 10^{34}$  & 2.253  & $1.051 \times 10^{34}$  \\
- $T$                         & $2.321 \times 10^{21}$  & 1.606  & $7.623 \times 10^{20}$  & 1.993  & $1.915 \times 10^{20}$  \\
- $\rho X(\isotm{He}{4})$     & $2.806 \times 10^{18}$  & 2.028  & $6.880 \times 10^{17}$  & 2.432  & $1.275 \times 10^{17}$  \\
- $\rho X(\isotm{C}{12})$     & $1.807 \times 10^{17}$  & 1.846  & $5.026 \times 10^{16}$  & 1.605  & $1.653 \times 10^{16}$  \\
- $\rho X(\isotm{O}{16})$     & $1.905 \times 10^{14}$  & 1.500  & $6.737 \times 10^{13}$  & 1.439  & $2.484 \times 10^{13}$  \\
- $\rho X(\isotm{Fe}{56})$    & $2.735 \times 10^{8}$   & 2.057  & $6.571 \times 10^{7}$   & 2.474  & $1.183 \times 10^{7}$   \\
+ $\rho$                      & $2.781 \times 10^{18}$  & 2.051  & $6.711 \times 10^{17}$  & 2.580  & $1.122 \times 10^{17}$  \\
+ $\rho u$                    & $6.781 \times 10^{26}$  & 2.445  & $1.245 \times 10^{26}$  & 2.907  & $1.659 \times 10^{25}$  \\
+ $\rho v$                    & $6.781 \times 10^{26}$  & 2.445  & $1.245 \times 10^{26}$  & 2.907  & $1.659 \times 10^{25}$  \\
+ $\rho E$                    & $2.465 \times 10^{35}$  & 2.333  & $4.892 \times 10^{34}$  & 2.650  & $7.794 \times 10^{33}$  \\
+ $\rho e$                    & $2.268 \times 10^{35}$  & 2.298  & $4.610 \times 10^{34}$  & 2.722  & $6.988 \times 10^{33}$  \\
+ $T$                         & $2.248 \times 10^{21}$  & 1.682  & $7.003 \times 10^{20}$  & 2.439  & $1.291 \times 10^{20}$  \\
+ $\rho X(\isotm{He}{4})$     & $2.862 \times 10^{18}$  & 2.027  & $7.023 \times 10^{17}$  & 2.554  & $1.196 \times 10^{17}$  \\
+ $\rho X(\isotm{C}{12})$     & $1.717 \times 10^{17}$  & 1.945  & $4.460 \times 10^{16}$  & 2.194  & $9.745 \times 10^{15}$  \\
+ $\rho X(\isotm{O}{16})$     & $1.718 \times 10^{14}$  & 1.648  & $5.482 \times 10^{13}$  & 1.898  & $1.471 \times 10^{13}$  \\
+ $\rho X(\isotm{Fe}{56})$    & $2.781 \times 10^{8}$   & 2.051  & $6.711 \times 10^{7}$   & 2.580  & $1.122 \times 10^{7}$   \\
 \enddata
@@
 
-% run on 12-31-20
-% Castro: 20.12-80-gfe3e9c8de8
-\begin{deluxetable}{lllllll}
-\tablecaption{\label{table:react_convert_strang} Convergence ($L_1$ norm) for the reacting convergence problem using SDC integration (2 iterations).}
-\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
-\startdata
- $\rho$                      & $2.794 \times 10^{18}$  & 2.047  & $6.760 \times 10^{17}$  & 2.563  & $1.144 \times 10^{17}$  \\
- $\rho u$                    & $6.796 \times 10^{26}$  & 2.450  & $1.243 \times 10^{26}$  & 2.898  & $1.669 \times 10^{25}$  \\
- $\rho v$                    & $6.796 \times 10^{26}$  & 2.450  & $1.243 \times 10^{26}$  & 2.898  & $1.669 \times 10^{25}$  \\
- $\rho E$                    & $2.450 \times 10^{35}$  & 2.357  & $4.783 \times 10^{34}$  & 2.743  & $7.144 \times 10^{33}$  \\
- $\rho e$                    & $2.261 \times 10^{35}$  & 2.325  & $4.513 \times 10^{34}$  & 2.821  & $6.385 \times 10^{33}$  \\
- $T$                         & $2.236 \times 10^{21}$  & 1.690  & $6.929 \times 10^{20}$  & 2.473  & $1.248 \times 10^{20}$  \\
- $\rho X(\isotm{He}{4})$     & $2.877 \times 10^{18}$  & 2.023  & $7.078 \times 10^{17}$  & 2.536  & $1.220 \times 10^{17}$  \\
- $\rho X(\isotm{C}{12})$     & $1.697 \times 10^{17}$  & 1.947  & $4.402 \times 10^{16}$  & 2.222  & $9.436 \times 10^{15}$  \\
- $\rho X(\isotm{O}{16})$     & $1.688 \times 10^{14}$  & 1.658  & $5.349 \times 10^{13}$  & 1.950  & $1.385 \times 10^{13}$  \\
- $\rho X(\isotm{Fe}{56})$    & $2.794 \times 10^{8}$   & 2.047  & $6.760 \times 10^{7}$   & 2.563  & $1.144 \times 10^{7}$   \\
-\enddata
-\end{deluxetable}
-
-\subsection{NSE convergence test problem}
-
-To test the coupling of the NSE table to the hydrodynamics, we run a
-similar problem as above, except with the thermodynamic conditions
-appropriate for the matter to be in NSE.  The initial conditions are:
-\begin{align}
-\rho &= \rho_0 \\
-T &= T_0 \left [ 1 + (\delta T) e^{-(r/\lambda)^2} \cos^6(\pi r/L) \right ] \\
-Y_e &= (Y_e)_0 \left [ 1 + (\delta Y_e) e^{-(r/\lambda)^2} \cos^6(\pi r/L) \right ] 
-\end{align}
-where $r$ is the distance from the center of the domain and the
-remaining parameters are listed in Table~\ref{table:nse}.  This has
-$Y_e$ varying between $[0.47, 0.5]$ initially\MarginPar{check}.  The mass fractions
-and remaining composition variables are then initialized from the NSE table, as described in section~\ref{sec:initialization}.
-
-The domain has a size $[0, L]^2$, and the timestep is fixed as:
-\begin{equation}
-\Delta t = 10^{-3} \left ( \frac{64}{N} \right )~\mathrm{s}
-\end{equation}
-where $N$ is the number of zones in each direction.
-
-\begin{deluxetable}{lcc}
-\tablecaption{\label{table:nse} NSE convergence test problem.}
-\tablehead{\colhead{parameter} & \colhead{value}}
-\startdata
-$\rho_0$ & $5\times 10^8~\gcc$ \\
-$T_0$    & $4\times 10^9~\mathrm{K}$ \\
-$(\delta T)$ & $0.2$ \\
-$(Y_e)_0$   & $0.5$ \\
-$(\delta Y_e)$ & $-0.05$ \\
-$\lambda$   & $2\times 10^7~\mathrm{cm}$ \\
-$L$         & $10^8~\mathrm{cm}$ \\
-\enddata
-\end{deluxetable}
-
-Figure~\ref{fig:nse_ye} shows the $Y_e$ profile for the Strang evolution case.
-
-\begin{figure}[t]
-\centering
-\plotone{ye_plot}
-\caption{\label{fig:nse_ye} Profile of $Y_e$ through the middle of the domain (slicing along $x$)
-for the $128^2$ Strang NSE test simulation.}
-\end{figure}
-
-\begin{deluxetable}{lllllll}
-\tablecaption{\label{table:nse_strang_methodII} Convergence ($L_1$ norm) for the NSE convergence
-problem using Strang splitting.}
-\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
-\startdata
- $\rho$                      & $4.522 \times 10^{19}$  & 1.659  & $1.432 \times 10^{19}$  & 1.542  & $4.917 \times 10^{18}$  \\
- $\rho u$                    & $1.883 \times 10^{28}$  & 1.681  & $5.873 \times 10^{27}$  & 1.524  & $2.042 \times 10^{27}$  \\
- $\rho v$                    & $1.883 \times 10^{28}$  & 1.681  & $5.873 \times 10^{27}$  & 1.524  & $2.042 \times 10^{27}$  \\
- $\rho E$                    & $5.933 \times 10^{37}$  & 1.613  & $1.940 \times 10^{37}$  & 1.411  & $7.293 \times 10^{36}$  \\
- $\rho e$                    & $5.931 \times 10^{37}$  & 1.613  & $1.939 \times 10^{37}$  & 1.411  & $7.291 \times 10^{36}$  \\
- $T$                         & $9.885 \times 10^{20}$  & 1.207  & $4.281 \times 10^{20}$  & 1.058  & $2.057 \times 10^{20}$  \\
- $\rho Y_e$                  & $2.379 \times 10^{19}$  & 1.678  & $7.431 \times 10^{18}$  & 1.578  & $2.489 \times 10^{18}$  \\
- $\rho \bar{A}$              & $5.141 \times 10^{21}$  & 1.230  & $2.191 \times 10^{21}$  & 1.087  & $1.032 \times 10^{21}$  \\
- $\rho (B/A)$                & $4.149 \times 10^{20}$  & 1.693  & $1.283 \times 10^{20}$  & 1.562  & $4.345 \times 10^{19}$  \\
-\enddata
-\end{deluxetable}
-
-\subsection{Detonation}
-
-The purpose here is to look at what timestep is taken when we use the nuclear burning limiters.
-Run with aprox19 only, Strang and SDC, then aprox19 + NSE, Strang and NSE.
-
-Look at number of SDC iterations too.
-
-How does the timestep limiter work in the middle of SDC iterating?
-
-aprox21
-
-look at number of RHS calls, wallclock time
-
-output every step and then for a single zone plot enuc every timestep to see if it oscillates
-
-\subsection{Massive Star}
-
-Play around with the threshold where NSE kicks in
-
 \section{Summary}
-
-We presented a simplified spectral deferred corrections scheme for coupling
-hydrodynamics and reactions.
-
-Future work is to extend this methodology to MHD.
 
@@
 
-\appendix
-
-\section{Jacobian}
-
-\label{sec:app:jacobian}
-
-To solve the reaction system implicitly, the ODE solver needs the Jacobian,
-$\partial \Rb/\partial \Uc$.  We follow the method of \cite{castro_sdc}
-and factor this into two pieces,
-\begin{equation}
-{\bf J} = \frac{\partial \Rb}{\partial {\bf w }} \frac{\partial {\bf w}}{\partial \Uc}.
-\end{equation}
-
-Even though it has no reactive sources, we include $\rho$ in our conservative state
-for the purposes of computing the Jacobian (we denote the conserved state used for the
-Jacobian $\Uc^\prime$).
-Writing this out for two species, $X_\alpha$ and $X_\beta$, we have
-\begin{equation}
-\Uc^\prime = \left ( \begin{array}{c} \rho \\ \rho X_\alpha \\ \rho X_\beta \\ \rho E \\ \rho e \end{array} \right )
-\end{equation}
-We take the intermediate state to be ${\bf w} = (\rho, X_\alpha, X_\beta,
-K, T)^\intercal$, where $K$ is the kinetic energy:
-\begin{equation}
-K = \frac{1}{2} u^2
-\end{equation}
-
-The Jacobian transformation $\partial \Uc^\prime/\partial {\bf w}$ is:
-\begin{equation}
-\frac{\partial \Uc}{\partial {\bf w}} = \left (
-   \begin{array}{ccccc}
-       1 & 0 & 0 & 0 & 0 \\
-       X_\alpha & \rho & 0 & 0 & 0 \\
-       X_\beta & 0 & \rho & 0 & 0  \\
-       \rho e_\rho  + e + K &
-                 \rho  e_{X_\alpha} & \rho e_{X_\beta} & \rho &
-                 \rho e_T \\
-       \rho e_\rho  + e  &
-                 \rho  e_{X_\alpha} & \rho e_{X_\beta} & 0 &
-                 \rho e_T
-     \end{array}\right)
-\end{equation}
-where we use the following notation for compactness:
-\begin{equation}
-e_\rho = \dedrd \qquad
-e_T = \dedTd \qquad
-e_{X_k} = \dedXd
-\end{equation}
-and write the kinetic energy as $K = u^2/2$.  We get the inverse
-(computed via SymPy, see the included Jupyter notebook) is:
-
-\begin{equation}
-\renewcommand{\arraystretch}{1.5}
-\frac{\partial {\bf w}}{\partial \Uc^\prime} = \left (
-  \begin{array}{ccccc}
-   1  & 0 & 0 & 0 & 0 \\
-   - \frac{X_\alpha}{\rho} & \frac{1}{\rho} & 0 & 0 & 0 \\
-   - \frac{X_\beta}{\rho} & 0 & \frac{1}{\rho} & 0 & 0 \\
-   - \frac{K}{\rho} & 0 & 0 & \frac{1}{\rho} & -\frac{1}{\rho} \\
-   \frac{\sum_k X_{k} e_{X_k} - \rho e_\rho - e}{ \rho e_T} &
-    -\frac{e_{X_\alpha}}{\rho e_T} & -\frac{e_{X_\beta}}{\rho e_T} & 0 & \frac{1}{\rho e_T} \\
-   \end{array}\right)
-\renewcommand{\arraystretch}{1}
-\end{equation}
-
-The reaction vector is
-\begin{equation}
-\Rb(\Uc^\prime) = \left (  \begin{array}{c} 0 \\ \rho \omegadot_\alpha \\ \rho \omegadot_\beta \\ \rho \Sdot \\ \rho \Sdot \end{array} \right )
-\end{equation}
-and the Jacobian is computed as $\partial \Rb/\partial {\bf w}$:
-\begin{equation}
-\frac{\partial \Rb}{\partial {\bf w}} = \left (
-  \begin{array}{ccccc}
-     0 & 0 & 0 & 0 & 0 \\
-     \omegadot_\alpha + \rho \frac{\partial \omegadot_\alpha}{\partial \rho} &
-     \rho \frac{\partial \omegadot_\alpha}{\partial X_\alpha} &
-     \rho \frac{\partial \omegadot_\alpha}{\partial X_\beta} & 0 & \rho \frac{\partial \omegadot_\alpha}{\partial T} \\
-      %
-     \omegadot_\beta + \rho \frac{\partial \omegadot_\beta}{\partial \rho} &
-     \rho \frac{\partial \omegadot_\beta}{\partial X_\alpha} &
-     \rho \frac{\partial \omegadot_\beta}{\partial X_\beta} & 0 &  \rho \frac{\partial \omegadot_\beta}{\partial T} \\
-     %
-     \Sdot + \rho \frac{\partial \Sdot}{\partial \rho} & \rho \frac{\partial \Sdot}{\partial X_\alpha} & \rho \frac{\partial \Sdot}{\partial X_\beta} & 0 & \rho \frac{\partial \Sdot}{\partial T} \\
-     %
-     \Sdot + \rho \frac{\partial \Sdot}{\partial \rho} & \rho \frac{\partial \Sdot}{\partial X_\alpha} & \rho \frac{\partial \Sdot}{\partial X_\beta} & 0 & \rho \frac{\partial \Sdot}{\partial T} \\
-  \end{array}
-  \right )
-\end{equation}
-
 %======================================================================

### 2021-02-19 strang/paper.tex
add e integration results
some paper work

@@
 %\title{A Simplified Spectral Deferred Correction Method for Coupling Hydrodynamics with Reaction Networks and Nuclear Statistical Equilibrium}
-\title{Practical Effects of Integrating Temperature / Energy with Strang Split Reactions}
+\title{Practical Effects of Integrating Temperature with Strang Split Reactions}
 
@@
 \begin{abstract}
-Folks should integrate energy with their reactions!
+For astrophysical reacting flows, operator splitting is commonly used
+to couple hydrodynamics and reactions.  Each process operates
+independent of one another, but by staggering the updates in a
+symmetric fashion (via Strang splitting) second order accuracy in time
+can be achieved.  However, approximations are often made to the
+reacting system, including the choice of whether or not to integrate
+temperature with the species.  Here we demonstrate through a simple
+convergence test that integrating temperature together with reactions
+is needed for second-order accuracy.
 \end{abstract}
@@
 
-\clearpage
-
-\begin{deluxetable}{lllllll}
-\tablecaption{\label{table:react_convert_no_T} Convergence ($L_1$ norm) for the reacting convergence with no $T$ integration.}
-\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
-\startdata
- $\rho$                      & $2.735 \times 10^{18}$  & 2.057  & $6.571 \times 10^{17}$  & 2.474  & $1.183 \times 10^{17}$  \\
- $\rho u$                    & $6.729 \times 10^{26}$  & 2.416  & $1.260 \times 10^{26}$  & 2.699  & $1.941 \times 10^{25}$  \\
- $\rho v$                    & $6.729 \times 10^{26}$  & 2.416  & $1.260 \times 10^{26}$  & 2.699  & $1.941 \times 10^{25}$  \\
- $\rho E$                    & $2.513 \times 10^{35}$  & 2.226  & $5.372 \times 10^{34}$  & 2.169  & $1.195 \times 10^{34}$  \\
- $\rho e$                    & $2.292 \times 10^{35}$  & 2.194  & $5.011 \times 10^{34}$  & 2.253  & $1.051 \times 10^{34}$  \\
- $T$                         & $2.321 \times 10^{21}$  & 1.606  & $7.623 \times 10^{20}$  & 1.993  & $1.915 \times 10^{20}$  \\
- $\rho X(\isotm{He}{4})$     & $2.806 \times 10^{18}$  & 2.028  & $6.880 \times 10^{17}$  & 2.432  & $1.275 \times 10^{17}$  \\
- $\rho X(\isotm{C}{12})$     & $1.807 \times 10^{17}$  & 1.846  & $5.026 \times 10^{16}$  & 1.605  & $1.653 \times 10^{16}$  \\
- $\rho X(\isotm{O}{16})$     & $1.905 \times 10^{14}$  & 1.500  & $6.737 \times 10^{13}$  & 1.439  & $2.484 \times 10^{13}$  \\
- $\rho X(\isotm{Fe}{56})$    & $2.735 \times 10^{8}$   & 2.057  & $6.571 \times 10^{7}$   & 2.474  & $1.183 \times 10^{7}$   \\
-\enddata
-\end{deluxetable}
-
-\begin{deluxetable}{lllllll}
-\tablecaption{\label{table:react_convert_no_T} Convergence ($L_1$ norm) for the reacting convergence with $T$ integration but frozen $c_x$.}
-\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
-\startdata
- $\rho$                      & $2.781 \times 10^{18}$  & 2.051  & $6.711 \times 10^{17}$  & 2.580  & $1.122 \times 10^{17}$  \\
- $\rho u$                    & $6.781 \times 10^{26}$  & 2.445  & $1.245 \times 10^{26}$  & 2.907  & $1.659 \times 10^{25}$  \\
- $\rho v$                    & $6.781 \times 10^{26}$  & 2.445  & $1.245 \times 10^{26}$  & 2.907  & $1.659 \times 10^{25}$  \\
- $\rho E$                    & $2.465 \times 10^{35}$  & 2.333  & $4.892 \times 10^{34}$  & 2.650  & $7.794 \times 10^{33}$  \\
- $\rho e$                    & $2.268 \times 10^{35}$  & 2.298  & $4.610 \times 10^{34}$  & 2.722  & $6.988 \times 10^{33}$  \\
- $T$                         & $2.248 \times 10^{21}$  & 1.682  & $7.003 \times 10^{20}$  & 2.439  & $1.291 \times 10^{20}$  \\
- $\rho X(\isotm{He}{4})$     & $2.862 \times 10^{18}$  & 2.027  & $7.023 \times 10^{17}$  & 2.554  & $1.196 \times 10^{17}$  \\
- $\rho X(\isotm{C}{12})$     & $1.717 \times 10^{17}$  & 1.945  & $4.460 \times 10^{16}$  & 2.194  & $9.745 \times 10^{15}$  \\
- $\rho X(\isotm{O}{16})$     & $1.718 \times 10^{14}$  & 1.648  & $5.482 \times 10^{13}$  & 1.898  & $1.471 \times 10^{13}$  \\
- $\rho X(\isotm{Fe}{56})$    & $2.781 \times 10^{8}$   & 2.051  & $6.711 \times 10^{7}$   & 2.580  & $1.122 \times 10^{7}$   \\
-\enddata
-\end{deluxetable}
-
-\begin{deluxetable}{lllllll}
-\tablecaption{\label{table:react_convert_no_T} Convergence ($L_1$ norm) for the reacting convergence with no $T$ integration and calling the EOS.}
-\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
-\startdata
- $\rho$                      & $2.781 \times 10^{18}$  & 2.051  & $6.711 \times 10^{17}$  & 2.580  & $1.122 \times 10^{17}$  \\
- $\rho u$                    & $6.781 \times 10^{26}$  & 2.445  & $1.245 \times 10^{26}$  & 2.907  & $1.659 \times 10^{25}$  \\
- $\rho v$                    & $6.781 \times 10^{26}$  & 2.445  & $1.245 \times 10^{26}$  & 2.907  & $1.659 \times 10^{25}$  \\
- $\rho E$                    & $2.465 \times 10^{35}$  & 2.333  & $4.892 \times 10^{34}$  & 2.650  & $7.794 \times 10^{33}$  \\
- $\rho e$                    & $2.268 \times 10^{35}$  & 2.298  & $4.610 \times 10^{34}$  & 2.722  & $6.988 \times 10^{33}$  \\
- $T$                         & $2.248 \times 10^{21}$  & 1.682  & $7.003 \times 10^{20}$  & 2.439  & $1.291 \times 10^{20}$  \\
- $\rho X(\isotm{He}{4})$     & $2.862 \times 10^{18}$  & 2.027  & $7.023 \times 10^{17}$  & 2.554  & $1.196 \times 10^{17}$  \\
- $\rho X(\isotm{C}{12})$     & $1.717 \times 10^{17}$  & 1.945  & $4.460 \times 10^{16}$  & 2.194  & $9.745 \times 10^{15}$  \\
- $\rho X(\isotm{O}{16})$     & $1.718 \times 10^{14}$  & 1.648  & $5.482 \times 10^{13}$  & 1.898  & $1.471 \times 10^{13}$  \\
- $\rho X(\isotm{Fe}{56})$    & $2.781 \times 10^{8}$   & 2.051  & $6.711 \times 10^{7}$   & 2.580  & $1.122 \times 10^{7}$   \\
-\enddata
-\end{deluxetable}
+\begin{figure}[t]
+\centering
+\plotone{test}
+\caption{\label{thefigure} our results}
+\end{figure}
 

### 2021-02-19 strang/paper.tex
update text

@@
 
-We are using Castro
+We are using the freely-available \castro\ simulation
+code~\citep{castro,castro_joss}.  \castro\ solves the equations of
+hydrodynamics using an unsplit piecewise parabolic method coupled with
+gravity and reactions.  For reactive flow, either Strang splitting or
+spectral deferred corrections are used to couple the hydrodynamics and
+reactions~\citep{castro_sdc}.  Here we focus on the Strang splitting.
 
-\section{Simulations}
+One option to make temperature integration less expensive is to freeze
+the values of the thermodynamics quantities at the start of the
+reactive integration.  This approach was done in \citet{ABNZ:III}.
+This can work well in ``simmering'' phases of reactions where the
+thermodynamic state is not changing much over a timestep, but for
+explosive reactions, this may be a poor approximation.
 
-\subsection{Reacting convergence test problem}
-
-We use the reacting convergence test problem introduced in \cite{castro_sdc}.
+In \cite{castro_sdc}, we introduced a test problem where we could
+measure the convergence of a reacting hydrodynamics problem via
+Richardson extrapolation---this was an acoustic pulse with helium
+burning via $3-\alpha$ and
+$\isotm{C}{12}(\alpha,\gamma)\isotm{O}{16}$.  Those tests focused on
+SDC coupling and showed that we can get overall 4th order in space and
+time convergence.  Here we run the same test with Strang coupling,
+looking at the various approaches at incorporating a temperature
+equation in the reactive portion of the update.
 

### 2021-02-19 strang/paper.tex
more updates

@@
 Richardson extrapolation---this was an acoustic pulse with helium
-burning via $3-\alpha$ and
+burning via $3$-$\alpha$ and
 $\isotm{C}{12}(\alpha,\gamma)\isotm{O}{16}$.  Those tests focused on
@@
 equation in the reactive portion of the update.
+
+For each method, we run the \castro\ {\tt reacting\_convergence} test
+problem at 5 resolutions: $64^2$, $128^2$, $256^2$, $512^2$, and
+$1024^2$.  We then compute four errors between adjacent resolutions by
+coarsening the finer resolution run, and computing the $L_1$ norm over
+all zones.  Figure~\ref{thefigure} shows the norm of the error
+vs.\ the coarse run resolution.  Guidelines showing ideal first and
+second order convergence are shown.
 
@@
           python (\url{https://www.python.org/}),
-          SymPy \citep{sympy}
+          VODE \cite{vode}
          }

### 2021-02-22 strang/paper.tex
updates

@@
-\documentclass[times,modern]{aastex63}
+\documentclass[times,preprint]{aastex63}
 
@@
 
-One option to make temperature integration less expensive is to freeze
-the values of the thermodynamics quantities at the start of the
-reactive integration.  This approach was done in \citet{ABNZ:III}.
-This can work well in ``simmering'' phases of reactions where the
-thermodynamic state is not changing much over a timestep, but for
-explosive reactions, this may be a poor approximation.
+The equations of hydrodynamics with reacting sources are:
+\begin{align}
+\frac{\partial \rho}{\partial t} + \nabla \cdot (\rho \Ub) &= 0 \\
+\frac{\partial (\rho X_k)}{\partial t} + \nabla \cdot (\rho \Ub X_k) &= \rho \omegadot_k \\
+\frac{\partial (\rho \Ub)}{\partial t} + \nabla \cdot (\rho \Ub \Ub) + \nabla p &= 0 \\
+\frac{\partial (\rho E)}{\partial t} + \nabla \cdot \left [ (\rho E + p) \Ub \right ] &= \rho \Sdot
+\end{align}
+
+When we are reacting, we need the temperature, so we can look at a temperature evolution
+equation:
+\begin{equation}
+\rho c_v \frac{DT}{Dt} = \rho \left (\frac{p}{\rho^2} - e_\rho \right ) \frac{D\rho}{Dt} + \rho \Sdot
+\end{equation}
+(this form neglects the change in temperature due to composition
+changes from reactions, see \citealt{ABNZ:III}, which assumes chemical
+equilibrium).  During reactions, we neglect the hydrodynamics terms,
+so the reactive system updates according to:
+\begin{align}
+\frac{D\rho}{Dt} &= 0 \\
+\frac{DX_k}{Dt} &= \omegadot_k \label{eq:species} \\
+\frac{DT}{Dt} &= \frac{\Sdot}{c_v}
+\end{align}
+
+There are several different approximations that people make in the
+literature to this operator-split reacting system, in order to make
+integrating the reaction system computationally less expensive:
+\begin{itemize}
+\item Neglect temerature evolution completely in the burn, only
+  evolving Eq.~\ref{eq:species}.  This is the method used in
+  \citet{flash}.
+\item Evolve the temperature equation, but ``freeze'' the specific
+  heat, $c_v$, at the start of integration to avoid potentially
+  expensive equation of state calls.  This was discussed in
+  \citet{ABNZ:III} and is the default method in \castro.
+\item Evolve the full system, calling the equation of state each time
+  we evaluate the righthand side of the system to get an updated
+  $c_v$.
+\end{itemize}
+Depending on how vigorous the burning is and how much the temperature
+changes during a hydrodynamic timestep, one or more of these
+approximations may be reasonable.  For explosive reactions, we expect
+that evolving the full system will be required.  The goal of this
+note is to try to quantify the convergence of a reacting hydrodynamics
+code with these different approximations.
 
@@
 \section{Summary}
+
+Looking at global convergence of a reacting hydrodynamics problem we
+see that second order convergence is only realized when temperature is
+evolved alongside the composition when using an operator-split
+approach to reactions.
 

### 2021-02-22 strang/paper.tex
more work on the note

@@
 convergence test that integrating temperature together with reactions
-is needed for second-order accuracy.
+achieves the best convergence when modeling reactive flows with Strang
+splitting.
 \end{abstract}
@@
 \section{Introduction}\label{Sec:Introduction}
+
+Simulations of stellar flows require solving the equations of
+hydrodynamics coupled with a nuclear reaction network.  The equations
+of hydrodynamics with reacting sources are:
+\begin{align}
+\frac{\partial \rho}{\partial t} + \nabla \cdot (\rho \Ub) &= 0 \\
+\frac{\partial (\rho X_k)}{\partial t} + \nabla \cdot (\rho \Ub X_k) &= \rho \omegadot_k \\
+\frac{\partial (\rho \Ub)}{\partial t} + \nabla \cdot (\rho \Ub \Ub) + \nabla p &= 0 \\
+\frac{\partial (\rho E)}{\partial t} + \nabla \cdot \left [ (\rho E + p) \Ub \right ] &= \rho \Sdot
+\end{align}
+where $\rho$ is the density, $\Ub$ is the velocity vector, $X_k$ are the species mass fractions,
+$\omegadot_k$ are the species creation rates (from reactions), $p$ is the pressure, $E$ is the
+total energy / unit mass, and $\Sdot$ is the nuclear energy generation rate.
+
+When we are reacting, we need to evolve the temperature, which we can write as:
+equation:
+\begin{equation}
+\rho c_v \frac{DT}{Dt} = \rho \left (\frac{p}{\rho^2} - e_\rho \right ) \frac{D\rho}{Dt} + \rho \Sdot
+\end{equation}
+(this form neglects the change in temperature due to composition
+changes from reactions, see \citealt{ABNZ:III}, which assumes chemical
+equilibrium).
 
@@
 
-The equations of hydrodynamics with reacting sources are:
-\begin{align}
-\frac{\partial \rho}{\partial t} + \nabla \cdot (\rho \Ub) &= 0 \\
-\frac{\partial (\rho X_k)}{\partial t} + \nabla \cdot (\rho \Ub X_k) &= \rho \omegadot_k \\
-\frac{\partial (\rho \Ub)}{\partial t} + \nabla \cdot (\rho \Ub \Ub) + \nabla p &= 0 \\
-\frac{\partial (\rho E)}{\partial t} + \nabla \cdot \left [ (\rho E + p) \Ub \right ] &= \rho \Sdot
-\end{align}
+In a Strang split evolution~\cite{strang:1968}, we update the full
+hydrodynamics state, $\Uc$ as:
+\begin{equation}
+  \Uc^{n+1} = {\bf R}_{\Delta t/2} {\bf A}_{\Delta t} {\bf R}_{\Delta t/2} \Uc^n
+\end{equation}
+where ${\bf R}_{\Delta t/2}$ is the reaction update through a timestep
+$\Delta t/2$ and ${\bf A}_{\Delta t}$ is the advective update through
+$\Delta t$.  We see with this splitting, each process operates on the
+state left behind by the previous operation, and the staggering of the
+physics is done to give second order accuracy in time.
 
-When we are reacting, we need the temperature, so we can look at a temperature evolution
-equation:
-\begin{equation}
-\rho c_v \frac{DT}{Dt} = \rho \left (\frac{p}{\rho^2} - e_\rho \right ) \frac{D\rho}{Dt} + \rho \Sdot
-\end{equation}
-(this form neglects the change in temperature due to composition
-changes from reactions, see \citealt{ABNZ:III}, which assumes chemical
-equilibrium).  During reactions, we neglect the hydrodynamics terms,
+During reactions, we neglect the hydrodynamics terms,
 so the reactive system updates according to:
@@
   evolving Eq.~\ref{eq:species}.  This is the method used in
-  \citet{flash}.
+  \citet{flash}. \MarginPar{other codes?}
 \item Evolve the temperature equation, but ``freeze'' the specific
@@
   expensive equation of state calls.  This was discussed in
-  \citet{ABNZ:III} and is the default method in \castro.
+  \cite{Bell:2004} and is the default method in \castro.
 \item Evolve the full system, calling the equation of state each time
   we evaluate the righthand side of the system to get an updated
-  $c_v$.
+  $c_v$.  This is an option in \castro.
 \end{itemize}
@@
 
-In \cite{castro_sdc}, we introduced a test problem where we could
+In \citet{castro_sdc}, we introduced a test problem where we could
 measure the convergence of a reacting hydrodynamics problem via
-Richardson extrapolation---this was an acoustic pulse with helium
+Richardson extrapolation\MarginPar{ref?}---this was an acoustic pulse with helium
 burning via $3$-$\alpha$ and
@@
 problem at 5 resolutions: $64^2$, $128^2$, $256^2$, $512^2$, and
-$1024^2$.  We then compute four errors between adjacent resolutions by
+$1024^2$, with the timestep kept fixed in proportion to the grid resolution.
+We then compute four errors between adjacent resolutions by
 coarsening the finer resolution run, and computing the $L_1$ norm over
@@
 evolved alongside the composition when using an operator-split
-approach to reactions.
+approach to reactions.  This is just a single, rather simple problem,
+but this suggests that reactive hydrodynamics simulations should
+switch to integrating temperature or another energy equation together
+with reactions to yield better overall convergence and accuracy.  We
+expect that for problems with vigorous burning, such as detonation,s
+this will be especially important.
 

### 2021-02-23 strang/paper.tex
fix typo
new notebook

@@
 
-When we are reacting, we need to evolve the temperature, which we can write as:
-equation:
+When we are reacting, we can look at internal energy:
+\begin{equation}
+\rho \frac{De}{Dt} + p \nabla \cdot \Ub = \rho \Sdot
+\end{equation}
+or alternately, we can evolve the temperature, which we can write as:
 \begin{equation}
@@
 \frac{D\rho}{Dt} &= 0 \\
-\frac{DX_k}{Dt} &= \omegadot_k \label{eq:species} \\
-\frac{DT}{Dt} &= \frac{\Sdot}{c_v}
+\frac{DX_k}{Dt} &= \omegadot_k \label{eq:species}
 \end{align}
+with either
+\begin{equation}
+\frac{DT}{Dt} = \frac{\Sdot}{c_v}
+\end{equation}
+or
+\begin{equation}
+\frac{De}{Dt} = \Sdot
+\end{equation}
 
@@
 \begin{itemize}
-\item Neglect temerature evolution completely in the burn, only
-  evolving Eq.~\ref{eq:species}.  This is the method used in
-  \citet{flash}. \MarginPar{other codes?}
-\item Evolve the temperature equation, but ``freeze'' the specific
+\item Evolve $(X_k)$ only.  This neglects temerature evolution
+  completely in the burn, only evolving Eq.~\ref{eq:species}.  This is
+  the method used in \citet{flash}. \MarginPar{other codes?}
+
+\item Evolve $(X_k, T)$, but ``freeze'' the specific
   heat, $c_v$, at the start of integration to avoid potentially
@@
   \cite{Bell:2004} and is the default method in \castro.
-\item Evolve the full system, calling the equation of state each time
+
+\item Evolve $(X_k, T)$, calling the equation of state each time
   we evaluate the righthand side of the system to get an updated
   $c_v$.  This is an option in \castro.
+
+\item Evolve $(X_k, e)$, and get $T$ from $e$ using the equation of
+  state.
 \end{itemize}

### 2021-02-23 xrb2/paper.tex
clarify burning trajectories sentence
address referee comments

@@
 %\documentclass[12pt, preprint]{aastex}
-\documentclass[preprint,times,tighten]{aastex63}
+\documentclass[preprint,times,tighten,linenumbers,trackchanges]{aastex63}
 
@@
 We investigate the effect of rotation rate on the flame. With the
-exception of IGR J17480-2446 (\citealt{altamirano2010discovery}, spinning at $11~\mathrm{Hz}$), most observations of
-XRBs come from sources with rotation rates of $200-600~\mathrm{Hz}$
+exception of IGR J17480-2446 (\citealt{altamirano2010discovery}, spinning at $11~\mathrm{Hz}$), \replaced{most observations of
+XRBs come from sources with}{most observations of XRBs which come from sources with known rotation rates have} rotation rates of $200-600~\mathrm{Hz}$
 \citep{bilous:2019,galloway:2020}. There are a number of
@@
 coordinates and a smooth transition to the cooler region at larger
-radii.  The cool initial model is characterized by three temperatures:
+radii. \added{This hot region extends out to $r_\mathrm{pert} =4.096\times 10^4 \mathrm{cm}$.}
+ The cool initial model is characterized by three temperatures:
 $T_\mathrm{star}$ is the isothermal temperature of the underlying
@@
 atmosphere. The atmosphere structure is isentropic as it falls from
-$T_\mathrm{hi}$ down to $T_\mathrm{lo}$.  For the hot model, we
+$T_\mathrm{hi}$ down to $T_\mathrm{lo}$, \added{and so is marginally 
+unstable to convection}.  For the hot model, we
 replace $T_\mathrm{hi}$ with $T_\mathrm{hi} + \delta T$.  In the
@@
 The mean molecular weight $\bar{A}$ within each of our simulations seems to grow along defined tracks confined to certain temperatures $T$, as shown in the Figure \ref{fig:abar} phase plots.
-We believe that the tracks in the plot correspond to burning trajectories in the reaction network.
+We believe that the tracks in the plot correspond to different burning trajectories in phase space resulting from different thermodynamic conditions.
 Comparing Figure \ref{fig:abar} to \citet{flame_wave1}, these tracks are much more neat and clearly defined.
@@
 
-We ran a number of simulations of laterally propagating flames in XRBs in order to explore the effects of rotation and thermal structure.
-We found that increasing the rotation rate increased the energy generation rate within the flame and enhanced nuclear burning.
-Apart from the lowest rotation run (which failed to ignite), flame propagation was not noticeably impacted by rotation rate; by the time the different flame fronts reached steady propagation, they shared comparable velocities.
-These results are likely due to the rotation-dependent strength of the Coriolis force and its confinement of the flame balancing the enhanced nuclear burning.
-We explored several models with different crustal temperatures to determine what effects mechanisms such as higher accretion rates, crustal heating and shallow heating may have on flame propagation.
-We found that increasing the temperature of the crust significantly enhanced the flame propagation.
-This we believe to be because a cooler crust allows heat to more efficiently be transferred away from the flame itself, therefore reducing the flame's temperature, slowing burning and consequently reducing its propagation speed.
-At higher crustal temperatures, we saw that the inability for heat to be efficiently transported away from the flame front increased the flame temperature, driving unstable, accelerating flame propagation.
-We saw that if the crust temperature was too high, then instead of a flame the entire atmosphere would burn steadily.
-This is reminiscent of what is seen for neutron stars with accretion rates exceeding the Eddington limit.
+We ran a number of simulations of laterally propagating flames in XRBs in order 
+to explore the effects of rotation and thermal structure. We found that increasing 
+the rotation rate increased the energy generation rate within the flame and 
+enhanced nuclear burning. Apart from the lowest rotation run (which failed to 
+ignite), flame propagation was not noticeably impacted by rotation rate; by 
+the time the different flame fronts reached steady propagation, they shared 
+comparable velocities. These results are likely due to the rotation-dependent 
+strength of the Coriolis force and its confinement of the flame balancing 
+the enhanced nuclear burning. \added{This lack of dependence of the flame speed on 
+the rotation rate could be tested using observations, as it infers that the burst 
+rise time should have little dependence on the rotation rate. }
+
+We explored several models with different 
+crustal temperatures to determine what effects mechanisms such as higher 
+accretion rates, crustal heating and shallow heating may have on flame 
+propagation. We found that increasing the temperature of the crust 
+significantly enhanced the flame propagation. This we believe to be 
+because a cooler crust allows heat to more efficiently be transferred 
+away from the flame itself, therefore reducing the flame's temperature, 
+slowing burning and consequently reducing its propagation speed. At higher 
+crustal temperatures, we saw that the inability for heat to be efficiently 
+transported away from the flame front increased the flame temperature, 
+driving unstable, accelerating flame propagation. We saw that if the crust 
+temperature was too high, then instead of a flame the entire atmosphere 
+would burn steadily. This is reminiscent of what is seen for neutron stars 
+with accretion rates exceeding the Eddington limit. \added{This dependence of the 
+flame speed on the crustal temperature could also be tested using observations, 
+by looking to see how the burst rise time depends on the accretion rate (which 
+may affect the crustal heating).}
 
@@
 here is available in the git repo as {\tt
-  flame\_wave}.  The work at Stony Brook was supported by DOE/Office
+  flame\_wave}.  \added{We thank the anonymous referee for the helpful suggestions.}
+  The work at Stony Brook was supported by DOE/Office
 of Nuclear Physics grant DE-FG02-87ER40317.  This material is based upon work supported by the

### 2021-02-24 xrb2/paper.tex
addressed the last comment on the stable burning with high crustal temperatures

@@
 
-Figure~\ref{fig:compare_500Hz_abar} shows $\bar{A}$ for three $500~\mathrm{Hz}$ simulations with different initial temperature structures ($T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$) at $t = 70~\mathrm{ms}$.
-We do not plot $\tt{F500\_4E8}$ here because it fails to form a clear burning front.
-$\tt{F500\_3E8}$ (Figure~\ref{fig:compare_500Hz_abar}, bottom panel) --- the hottest run to form a clear burning front --- has a faster propagating flame (this will be discussed further in Section \ref{ssec:temp_prop}).
-It also reaches slightly higher $\bar{A}$ values than the two cooler runs.
-The Figure~\ref{fig:abar_hot} $\bar{A}$-$T$ phase plots of $\tt{F500}$ (left) and $\tt{F500\_3E8}$ (middle) also reflect these $\bar{A}$ features, with $\tt{F500\_3E8}$ reaching slightly higher $\bar{A}$ values.
-$\tt{F500\_3E8}$ also reaches higher $\enucdot$ values at the low end of the temperature range ($< 0.5 \times 10^9~\mathrm{K}$).
-There appear to be more causally connected regions across a range of $\bar{A}$ at low temperatures for $\tt{F500\_3E8}$ than for $\tt{F500}$, suggesting that the higher $\enucdot$ for $\tt{F500\_3E8}$ generates burning in certain burning trajectories that are not present in the cooler $\tt{F500}$ run.
-Note that, although $\tt{F500\_3E8}$ is the hottest run with a modified initial temperature configuration to form a distinct flame front, the highest rotation $\tt{F1000}$ flame actually reaches higher temperatures (see Figure~\ref{fig:maxima}, left panel) as well as higher $\bar{A}$ (see Figure~\ref{fig:time_series_1000}).
+Figure~\ref{fig:compare_500Hz_abar} shows $\bar{A}$ for three $500~\mathrm{Hz}$ simulations with 
+different initial temperature structures ($T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$) at 
+$t = 70~\mathrm{ms}$. We do not plot $\tt{F500\_4E8}$ here because it fails to form a clear burning 
+front. $\tt{F500\_3E8}$ (Figure~\ref{fig:compare_500Hz_abar}, bottom panel) --- the hottest run to 
+form a clear burning front --- has a faster propagating flame (this will be discussed further in 
+Section \ref{ssec:temp_prop}). It also reaches slightly higher $\bar{A}$ values than the two cooler 
+runs. The Figure~\ref{fig:abar_hot} $\bar{A}$-$T$ phase plots of $\tt{F500}$ (left) and 
+$\tt{F500\_3E8}$ (middle) also reflect these $\bar{A}$ features, with $\tt{F500\_3E8}$ reaching 
+slightly higher $\bar{A}$ values. $\tt{F500\_3E8}$ also reaches higher $\enucdot$ values at the 
+low end of the temperature range ($< 0.5 \times 10^9~\mathrm{K}$). There appear to be more causally 
+connected regions across a range of $\bar{A}$ at low temperatures for $\tt{F500\_3E8}$ than for 
+$\tt{F500}$, suggesting that the higher $\enucdot$ for $\tt{F500\_3E8}$ generates burning in certain 
+burning trajectories that are not present in the cooler $\tt{F500}$ run. Note that, although 
+$\tt{F500\_3E8}$ is the hottest run with a modified initial temperature configuration to form 
+a distinct flame front, the highest rotation $\tt{F1000}$ flame actually reaches higher temperatures 
+(see Figure~\ref{fig:maxima}, left panel) as well as higher $\bar{A}$ (see Figure~\ref{fig:time_series_1000}).
 
 In contrast to the models with $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$, $\tt{F500\_4E8}$ 
-is so hot that the organized flame structure is lost. The fuel layer undergoes stable burning and 
+is so hot that the organized flame structure is lost. \deleted{The fuel layer undergoes stable burning and 
 essentially simmers in place, rather than forming a flame. This process has been observed in 
@@
 \citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}; in this model we have essentially 
-recreated the conditions found in such stars with high accretion rates. This model burns so 
+recreated the conditions found in such stars with high accretion rates.} This model burns so 
 strongly that it is only run out to $40~\mathrm{ms}$. After an initial period when the burning 
@@
 all runs other than $\tt{F500\_4E8}$ are confined to the flame front, rather than spanning the entire domain.
+
+\added{Rather than the fuel igniting in a 
+localized hot spot and forming a flame which propagates across the domain, for this system the heat produced 
+by the high temperature crust alone is sufficient to ignite material across the entire domain. Uniform burning 
+across a large fraction of or even the entire surface of the neutron star has been observed in neutron stars 
+with accretion rates that exceed the Eddington limit 
+\citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}. However, it is important to note here that the
+conditions in our model vary from the conditions that would occur in a neutron star ocean undergoing stable burning 
+in a number of ways. Firstly, the model begins without any burning taking place, and initial model assumes that a 
+significant amount of unburnt fuel has been able to build up. For a stably burning system, freshly accreted material 
+would be continuously burnt, so there would be much less unburnt fuel available. Secondly, we are only modelling 
+a few tens of milliseconds, rather than an ongoing process. We therefore do not demonstrate that this system would 
+continue to burn stably if the high crustal temperature were maintained and sufficient fuel continued to be 
+available, e.g.~through the continued accretion of fresh material. However, this run does suggest that 
+a crust with a sufficiently high temperature can provide enough energy to heat the fuel above it to the point of 
+ignition, and it may be that if these conditions were maintained that it would be able to sustain stable burning.}
 

### 2021-02-25 xrb2/paper.tex
added new fig

@@
 Taking the edge at greatest radius of the bright teal/green region where the most significant energy generation is occurring as the leading edge of the flame in each plot, we see that the horizontal extent of the $1000~\mathrm{Hz}$ flame ($\tt{F1000}$) appears to be
-reduced compared to the lower rotation $500~\mathrm{Hz}$ run ($\tt{F500}$). From Equation 
+reduced compared to the lower rotation $500~\mathrm{Hz}$ run ($\tt{F500}$).
+\added{As the natural aspect ratio makes it hard to see the flame structure, we show the energy generation rate for the $\tt{F1000}$ simulation with the vertical extent stretched in Figure~\ref{fig:flame_stretch}.}  From Equation
 \ref{eqn:rossby}, we can see that increasing the rotation rate from $500~\mathrm{Hz}$ to 
@@
 \end{figure}
+
+\added{
+\begin{figure}[t]
+\centering
+\plotone{stretch}
+\caption{\label{fig:flame_stretch} Energy generation rate for run {\tt F1000} at 100 ms with the vertical extent stretched to show detail.}
+\end{figure}
+}
 

### 2021-02-25 xrb2/paper.tex
Updated discussion of flame speed scaling relation

@@
 
-We found in \citet{flame_wave1} that the flame speed, $s$, obeys the following relation:
+\added{For the simple case of a laminar flame front driven by conduction, the propagation rate, 
+$s_l$, of the flame is approximately
+\begin{equation}
+	\label{eqn:laminar_flame_speed}
+	s_l \approx \left( \frac{\mathcal{D} \enucdot}{E} \right)^{\half}.
+\end{equation}
+Here, $\mathcal{D} = k_{th} / (\rho c_v)$ is the thermal diffusivity, $\enucdot$ is the specific
+nuclear energy generation rate, and $E$ is the specific total energy \citep{timmeswoosley:1992}. 
+\citet{cavecchi:2013} explored the effects of rotation on the flame propagation, and noted that the 
+flame speed in their simulations scaled with the ratio $L_R/H_0$ of Rossby length to atmospheric 
+scale height. Combining this scaling with Equation \ref{eqn:laminar_flame_speed} yields a revised 
+estimate for the flame speed, $s$:
+\begin{equation}
+	s \approx \frac{L_R}{H_0} s_l \approx \frac{L_R}{H_0} \left( \frac{\mathcal{D} \enucdot}{E} 
+	\right)^{\half}.
+\end{equation}
+Observing that $L_R$ is inversely proportional to the angular rotation frequency $\Omega$ and 
+and neglecting variations in $H_0$ and $E$, we can then derive the following approximate scaling 
+relation:
 \begin{equation}
 	\label{eqn:flame_speed_scaling}
-	s \appropto L_R (\enucdot)^{\half} \propto \frac{(\enucdot)^{\half}}{\Omega},
+	s \appropto \frac{(\mathcal{D} \enucdot)^{\half}}{\Omega}
 \end{equation}
+We were able to test the validity of this scaling relation in \citet{flame_wave1}, where we studied 
+the effects of varying the thermal conductivity/diffusivity, energy generation rate, and rotation 
+rate. These results supplemented the findings of \citet{cavecchi:2013}. Despite this formulation of 
+the flame speed neglecting some relevant physics (e.g.\ turbulence), the relation proved
+consistent with our simulation results.}
+
+\deleted{We found in \citet{flame_wave1} that the flame speed, $s$, obeys the following relation:
+$s \appropto L_R (\enucdot)^{\half} \propto \frac{(\enucdot)^{\half}}{\Omega},$
 where $\enucdot$ is the specific nuclear energy generation rate and $L_R$ is the Rossby length 
@@
 \citet{cavecchi:2013}, who noted that the flame speeds in their simulations scaled with the Rossby 
-length. As seen in Figure~\ref{fig:flame_speeds_1}, there is no obvious inverse scaling of the 
-flame speed with rotation rate in this set of runs. We observed earlier, however, that nuclear 
+length.}
+
+\added{This study explores a more physical parameter space than \citet{flame_wave1}}, and as seen 
+in 
+Figure~\ref{fig:flame_speeds_1}, there is no clear inverse scaling of the 
+flame speed with rotation rate in the new set of runs. We observed earlier, however, that nuclear 
 reactions progress more quickly at higher rotation rate. This results in a higher 
-$\enucdot$ --- up 
-to three to four times higher near the burning front after the flame ignites (see Figure 
-\ref{fig:maxima}) --- which may counteract the reduction in flame speed from the increased Coriolis 
-confinement. We also observe that $\tt{F500}$ accelerates faster than $\tt{F1000}$, which appears 
-to experience a small deceleration at early times. This disparity may be a direct result of the 
-difference in Coriolis force.
+$\enucdot$ --- up to three to four times higher near the burning front after the flame ignites (see 
+Figure \ref{fig:maxima}) --- which may counteract the reduction in flame speed from the increased 
+Coriolis confinement. Comparing accelerations, we also observe that $\tt{F500}$ accelerates faster 
+than $\tt{F1000}$, which appears to experience a small deceleration at early times. This disparity 
+may be a direct result of the difference in Coriolis force.
 

### 2021-02-25 xrb2/paper.tex
address more referee comments

@@
 \begin{abstract}
-We continue to investigate two-dimensional laterally propagating flames in type I X-ray bursts using fully compressible hydrodynamics simulations.
-In the current study we relax previous approximations where we artificially boosted the flames.
-We now use more physically realistic reaction rates, thermal conductivities, and rotation rates, exploring the effects of neutron star rotation rate and thermal structure on the flame.
-We find that at lower rotation rates the flame becomes harder to ignite, whereas at higher rotation rates the nuclear burning is enhanced by increased confinement from the Coriolis force and the flame propagates steadily.
-At higher crustal temperatures, the flame moves more quickly and accelerates as it propagates through the atmosphere.
-If the temperature is too high, instead of a flame propagating across the surface the entire atmosphere burns steadily.
+We continue to investigate two-dimensional laterally propagating flames in type I X-ray 
+bursts using fully compressible hydrodynamics simulations.  In the current study we relax 
+previous approximations where we artificially boosted the flames. We now use more physically 
+realistic reaction rates, thermal conductivities, and rotation rates, exploring the effects 
+of neutron star rotation rate and thermal structure on the flame.  We find that at lower 
+rotation rates the flame becomes harder to ignite, whereas at higher rotation rates the 
+nuclear burning is enhanced by increased confinement from the Coriolis force and the 
+flame propagates steadily. At higher crustal temperatures, the flame moves more quickly 
+and accelerates as it propagates through the atmosphere. If the temperature is too high, 
+instead of a flame propagating across the surface the entire atmosphere burns 
+\replaced{steadily}{uniformly}. \added{Our findings could have implications for the 
+relationship between observed burst rise times and neutron star rotation and accretion rates.} 
 All of the software used for these simulations is freely available.
@@
 where $g$ is the gravitational acceleration, $H_0$ is the atmospheric scale height, and $\Omega$ is the 
-neutron star rotation rate. In Figure \ref{fig:time_series_enuc_500} and Figure 
+neutron star rotation rate.
+\added{If we take $g \sim 3.6 \times 10^{12}~\mathrm{cm}~\mathrm{s}^{-2}$ and $H_0\sim 10^3~\mathrm{cm}$, then for the $500~\mathrm{Hz}$ system we find the Rossby length to approximately be $L_R \sim 1.2 \times 10^5~\mathrm{cm}$.
+This estimate is likely to be an overestimate of the true value of the Rossby length in our simulations, as we saw that the flame is confined on a much smaller length scale.} In Figure \ref{fig:time_series_enuc_500} and Figure
 \ref{fig:time_series_enuc_1000}, we use $\dot{e}_\mathrm{nuc}$ measured at $50~\mathrm{ms}$ and $100~\mathrm{ms}$ to discern the horizontal extent of the flame at different rotation rates.
@@
 nuclear energy generation rate, and $E$ is the specific total energy \citep{timmeswoosley:1992}. 
-\citet{cavecchi:2013} explored the effects of rotation on the flame propagation, and noted that the 
+\citet{cavecchi:2013} explored the effects of rotation on the flame's propagation, and noted that the 
 flame speed in their simulations scaled with the ratio $L_R/H_0$ of Rossby length to atmospheric 
@@
 	\label{eqn:flame_speed_scaling}
-	s \appropto \frac{(\mathcal{D} \enucdot)^{\half}}{\Omega}
+	s \appropto \frac{(\mathcal{D} \enucdot)^{\half}}{\Omega}.
 \end{equation}
@@
 the effects of varying the thermal conductivity/diffusivity, energy generation rate, and rotation 
-rate. These results supplemented the findings of \citet{cavecchi:2013}. Despite this formulation of 
+rate. These results supported the findings of \citet{cavecchi:2013}. Despite this formulation of 
 the flame speed neglecting some relevant physics (e.g.\ turbulence), the relation proved
@@
 here is available in the git repo as {\tt
-  flame\_wave}.  \added{We thank the anonymous referee for the helpful suggestions.}
+  flame\_wave}.  \added{We thank the anonymous referee for their helpful suggestions.}
   The work at Stony Brook was supported by DOE/Office

### 2021-02-26 xrb2/paper.tex
added variable definitions, minor grammar edits

@@
 \end{equation}
-Here, $\mathcal{D} = k_{th} / (\rho c_v)$ is the thermal diffusivity, $\enucdot$ is the specific
+Here, $\mathcal{D} = k_{th} / (\rho c_v)$ is the thermal diffusivity with thermal conductivity $k_{th}$ and specific heat capacity $c_v$, $\enucdot$ is the specific
 nuclear energy generation rate, and $E$ is the specific total energy \citep{timmeswoosley:1992}. 
@@
 conditions in our model vary from the conditions that would occur in a neutron star ocean undergoing stable burning 
-in a number of ways. Firstly, the model begins without any burning taking place, and initial model assumes that a 
+in a number of ways. Firstly, the model begins without any burning taking place, and the initial model assumes that a 
 significant amount of unburnt fuel has been able to build up. For a stably burning system, freshly accreted material 
@@
 the enhanced nuclear burning. \added{This lack of dependence of the flame speed on 
-the rotation rate could be tested using observations, as it infers that the burst 
+the rotation rate could be tested using observations, as it implies that the burst 
 rise time should have little dependence on the rotation rate. }

### 2021-03-08 strang/paper.tex
update refs
update

@@
 
-%% \author[0000-0002-5749-334X]{J.~B.~Bell}
-%% \affiliation{Center for Computational Sciences and Engineering, Lawrence Berkeley National Laboratory, Berkeley, CA  94720}
+%% \author[0000-0002-1530-781X]{A.\ Harpole}
+%% \affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
+%%              Stony Brook, NY 11794-3800}
 
-%% \author[0000-0003-1791-0265]{A.~J.~Nonaka}
-%% \affiliation{Center for Computational Sciences and Engineering, Lawrence Berkeley National Laboratory, Berkeley, CA  94720}
-
-%% \author[0000-0001-8092-1974]{W.~Zhang}
-%% \affiliation{Center for Computational Sciences and Engineering, Lawrence Berkeley National Laboratory, Berkeley, CA  94720}
-
-%% \correspondingauthor{Michael Zingale}
-%% \email{michael.zingale@stonybrook.edu}
+%% \author[0000-0003-2300-5165]{D.\ E.\ Willcox}
+%% \affiliation{Lawrence Berkeley National Laboratory, Berkeley, CA}
 
@@
 
-We are using the freely-available \castro\ simulation
-code~\citep{castro,castro_joss}.  \castro\ solves the equations of
-hydrodynamics using an unsplit piecewise parabolic method coupled with
-gravity and reactions.  For reactive flow, either Strang splitting or
-spectral deferred corrections are used to couple the hydrodynamics and
+We use the freely-available \castro\ simulation
+code~\citep{castro,castro_joss} to solve the equations of
+hydrodynamics, using an unsplit piecewise parabolic method coupled with
+reactions.  \castro\ uses either Strang splitting or
+spectral deferred corrections (SDC) to couple the hydrodynamics and
 reactions~\citep{castro_sdc}.  Here we focus on the Strang splitting.
 
-In a Strang split evolution~\cite{strang:1968}, we update the full
-hydrodynamics state, $\Uc$ as:
+In a Strang split evolution~\citep{strang:1968}, we update the full
+hydrodynamics state, $\Uc$, as:
 \begin{equation}
@@
 
-\item Evolve $(X_k, T)$, but ``freeze'' the specific
-  heat, $c_v$, at the start of integration to avoid potentially
-  expensive equation of state calls.  This was discussed in
-  \cite{Bell:2004} and is the default method in \castro.
-
-\item Evolve $(X_k, T)$, calling the equation of state each time
-  we evaluate the righthand side of the system to get an updated
-  $c_v$.  This is an option in \castro.
+\item Evolve $(X_k, T)$.  To avoid expensive equation of state calls
+  in getting the specific heat, we can optionally ``freeze'' the value
+  of $c_v$ at the start of integration to avoid potentially expensive
+  equation of state calls.  This was discussed in \cite{Bell:2004} and
+  until recently was the default method in \castro.
 
@@
 \end{itemize}
+For all of these cases, density is constant during the reaction operation.
 Depending on how vigorous the burning is and how much the temperature
 changes during a hydrodynamic timestep, one or more of these
-approximations may be reasonable.  For explosive reactions, we expect
+methods may be reasonable.  For explosive reactions, we expect
 that evolving the full system will be required.  The goal of this
@@
 measure the convergence of a reacting hydrodynamics problem via
-Richardson extrapolation\MarginPar{ref?}---this was an acoustic pulse with helium
+Richardson extrapolation---this was an acoustic pulse with helium
 burning via $3$-$\alpha$ and
-$\isotm{C}{12}(\alpha,\gamma)\isotm{O}{16}$.  Those tests focused on
-SDC coupling and showed that we can get overall 4th order in space and
-time convergence.  Here we run the same test with Strang coupling,
-looking at the various approaches at incorporating a temperature
-equation in the reactive portion of the update.
+$\isotm{C}{12}(\alpha,\gamma)\isotm{O}{16}$.  Initially, the domain is
+pure $\isotm{He}{4}$, but both $\isotm{C}{12}$ and $\isotm{O}{16}$ are
+created as time evolves.  The published tests focused on SDC coupling
+and showed that we can get overall 4th order in space and time
+convergence.  Here we run the same test with Strang coupling, looking
+at the various approaches at incorporating a temperature / energy equation in
+the reactive portion of the update.
 
@@
 coarsening the finer resolution run, and computing the $L_1$ norm over
-all zones.  Figure~\ref{thefigure} shows the norm of the error
-vs.\ the coarse run resolution.  Guidelines showing ideal first and
-second order convergence are shown.
+all zones.
+
+Figure~\ref{thefigure} shows the norm of the error vs.\ the coarse run
+resolution.  The slope of these lines is a measure of the convergence
+rate~\citep{ORVV}.  Guidelines show ideal first and second order
+convergence.  We see that all methods converge at least second order
+for density, but for the thermodynamic quantities, $(\rho e)$ and $T$,
+the method where only $X_k$ is evolved during reactions has larger
+errors and much worse convergence than the other methods.  Looking at
+the trace nuclei generated in the burning, $\isotm{C}{12}$ and
+$\isotm{O}{16}$, we see a large difference between the two methods
+that evolve some sort of energy and the one method where only $X_k$ is
+evolved---the latter converging essentially first order at high
+resolution.
 
@@
 \plotone{test}
-\caption{\label{thefigure} our results}
+\caption{\label{thefigure} Convergence of fluid quantities as a function of resolution
+for 3 different Strang equation systems: just evolving $X_k$, evolving $(X_k, T)$, but fixing $c_v$ at the start of integration (``frozen''), and evolving $(X_k, e)$.}
 \end{figure}
@@
 Looking at global convergence of a reacting hydrodynamics problem we
-see that second order convergence is only realized when temperature is
-evolved alongside the composition when using an operator-split
+see that second order convergence is only realized when temperature or
+energy is evolved alongside the composition when using a Strang-split
 approach to reactions.  This is just a single, rather simple problem,
@@
 switch to integrating temperature or another energy equation together
-with reactions to yield better overall convergence and accuracy.  We
-expect that for problems with vigorous burning, such as detonation,s
-this will be especially important.
+with reactions to yield better overall convergence and accuracy.  This
+complements the work of \citep{muller:1986} which showed that when
+evolving near nuclear statistical equilibrium, that evolving entropy
+with the system is needed for stability.  We expect that for problems
+with vigorous burning, such as detonations, this will be especially
+important.
 

### 2021-03-09 strang/paper.tex
add fma

@@
 \frac{\partial \rho}{\partial t} + \nabla \cdot (\rho \Ub) &= 0 \\
-\frac{\partial (\rho X_k)}{\partial t} + \nabla \cdot (\rho \Ub X_k) &= \rho \omegadot_k \\
+\frac{\partial (\rho X_k)}{\partial t} + \nabla \cdot (\rho X_k \Ub) &= \rho \omegadot_k \\
 \frac{\partial (\rho \Ub)}{\partial t} + \nabla \cdot (\rho \Ub \Ub) + \nabla p &= 0 \\
@@
 \item Evolve $(X_k, e)$, and get $T$ from $e$ using the equation of
-  state.
+  state.  This was discussed in \citet{fma} and is the current default
+  method in \castro.
 \end{itemize}
@@
           python (\url{https://www.python.org/}),
-          VODE \cite{vode}
+          VODE \citep{vode}
          }

### 2021-03-10 strang/paper.tex
Add Gadget cite
Add Raskin cite

@@
 \begin{itemize}
-\item Evolve $(X_k)$ only.  This neglects temerature evolution
+\item Evolve $(X_k)$ only.  This neglects temperature evolution
   completely in the burn, only evolving Eq.~\ref{eq:species}.  This is
@@
 
-\item Evolve $(X_k, T)$.  To avoid expensive equation of state calls
+\item Evolve $(X_k, T)$.   This is used in \cite{Pakmor:2012}.  To avoid
+  expensive equation of state calls
   in getting the specific heat, we can optionally ``freeze'' the value
@@
 \end{itemize}
+\cite{Raskin:2010} also propose a hybrid system where the first approach is
+used in most cases, switching to the second approach only near NSE.
 For all of these cases, density is constant during the reaction operation.

### 2021-03-10 strang/paper.tex
Add Garcia-Senz
Reword

@@
 
-There are several different approximations that people make in the
-literature to this operator-split reacting system, in order to make
-integrating the reaction system computationally less expensive:
+There are several different approaches taken in the literature to
+this operator-split reacting system, including some approximations
+that make integrating the reaction system computationally less expensive:
 \begin{itemize}
@@
 
-\item Evolve $(X_k, T)$.   This is used in \cite{Pakmor:2012}.  To avoid
-  expensive equation of state calls
+\item Evolve $(X_k, T)$.   This is used in \cite{Pakmor:2012} and
+  \cite{Garcia-Senz:2013}.  To avoid expensive equation of state calls
   in getting the specific heat, we can optionally ``freeze'' the value

### 2021-03-12 strang/paper.tex
Clarify referent of this
remove redundant text

@@
   in getting the specific heat, we can optionally ``freeze'' the value
-  of $c_v$ at the start of integration to avoid potentially expensive
-  equation of state calls.  This was discussed in \cite{Bell:2004} and
+  of $c_v$ at the start of integration.  This was discussed in \cite{Bell:2004} and
   until recently was the default method in \castro.
@@
 with the system is needed for stability.  We expect that for problems
-with vigorous burning, such as detonations, this will be especially
-important.
+with vigorous burning, such as detonations, directly coupling the composition
+and thermodynamic evolution will be especially important.
 

### 2021-03-12 strang/paper.tex
fixed typo

@@
 complements the work of \citep{muller:1986} which showed that when
-evolving near nuclear statistical equilibrium, that evolving entropy
+evolving near nuclear statistical equilibrium, evolving entropy
 with the system is needed for stability.  We expect that for problems

### 2021-03-17 strang/paper.tex
final? version

@@
 
-%% \author[0000-0001-8401-030X]{M.~Zingale}
-%% \affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
-%%        Stony Brook, NY 11794-3800}
+\author[0000-0001-8401-030X]{M.~Zingale}
+\affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
+       Stony Brook, NY 11794-3800}
 
-%% \author[0000-0003-0439-4556]{M.~P.~Katz}
-%% \affiliation{Nvidia Corp}
+\author[0000-0003-0439-4556]{M.~P.~Katz}
+\affiliation{Nvidia Corp}
 
-%% \author[0000-0002-1530-781X]{A.\ Harpole}
-%% \affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
-%%              Stony Brook, NY 11794-3800}
+\author[0000-0003-2300-5165]{D.\ E.\ Willcox}
+\affiliation{Lawrence Berkeley National Laboratory, Berkeley, CA}
 
-%% \author[0000-0003-2300-5165]{D.\ E.\ Willcox}
-%% \affiliation{Lawrence Berkeley National Laboratory, Berkeley, CA}
+\author[0000-0002-1530-781X]{A.\ Harpole}
+\affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
+             Stony Brook, NY 11794-3800}
 
@@
   completely in the burn, only evolving Eq.~\ref{eq:species}.  This is
-  the method used in \citet{flash}. \MarginPar{other codes?}
+  the method used in \citet{flash}.
 
@@
 
-\acknowledgements \castro\ is freely available at
-\url{http://github.com/AMReX-Astro/Castro}.  All of the code and
-problem setups used here are available in the git repo.  The work at
+\acknowledgements \castro\ and all of the code used here is freely available at
+\url{https://github.com/AMReX-Astro/Castro}.   The work at
 Stony Brook was supported by DOE/Office of Nuclear Physics grant
@@
           GNU Compiler Collection (\url{https://gcc.gnu.org/}),
-          Linux (\url{https://www.kernel.org}),
           matplotlib (\citealt{Hunter:2007},  \url{http://matplotlib.org/})

### 2021-03-17 strang/paper.tex
update line styles

@@
 \caption{\label{thefigure} Convergence of fluid quantities as a function of resolution
-for 3 different Strang equation systems: just evolving $X_k$, evolving $(X_k, T)$, but fixing $c_v$ at the start of integration (``frozen''), and evolving $(X_k, e)$.}
+for 3 different Strang equation systems: just evolving $X_k$, evolving $(X_k, T)$ with $c_v$ held fixed, and evolving $(X_k, e)$.}
 \end{figure}

### 2021-03-18 strang/paper.tex
Make sure all variables are defined

@@
 $\omegadot_k$ are the species creation rates (from reactions), $p$ is the pressure, $E$ is the
-total energy / unit mass, and $\Sdot$ is the nuclear energy generation rate.
+specific total energy, and $\Sdot$ is the nuclear energy generation rate.
 
-When we are reacting, we can look at internal energy:
+When we are reacting, we can look at internal energy
 \begin{equation}
-\rho \frac{De}{Dt} + p \nabla \cdot \Ub = \rho \Sdot
+\rho \frac{De}{Dt} + p \nabla \cdot \Ub = \rho \Sdot,
 \end{equation}
-or alternately, we can evolve the temperature, which we can write as:
+where $e$ is the specific internal energy or alternately, we can evolve the temperature, 
+which we can write as
 \begin{equation}
-\rho c_v \frac{DT}{Dt} = \rho \left (\frac{p}{\rho^2} - e_\rho \right ) \frac{D\rho}{Dt} + \rho \Sdot
+\rho c_v \frac{DT}{Dt} = \rho \left (\frac{p}{\rho^2} - e_\rho \right ) \frac{D\rho}{Dt} + \rho \Sdot,
 \end{equation}
-(this form neglects the change in temperature due to composition
+where $c_v$ is the specific heat capacity at constant volume, $T$ is the temperature and $e_\rho$ is the derivative of 
+the specific internal energy with respect to density (this form neglects the change in 
+temperature due to composition
 changes from reactions, see \citealt{ABNZ:III}, which assumes chemical
@@
 \caption{\label{thefigure} Convergence of fluid quantities as a function of resolution
-for 3 different Strang equation systems: just evolving $X_k$, evolving $(X_k, T)$ with $c_v$ held fixed, and evolving $(X_k, e)$.}
+for 3 different Strang equation systems: just evolving $X_k$, evolving $(X_k, T)$ with $c_v$ held fixed, and evolving $(X_k, e)$. The dotted lines show ideal first and second order convergence.}
 \end{figure}

### 2021-03-18 xrb2/paper.tex
Make sure all variables are defined

@@
 neutron star rotation rate.
-\added{If we take $g \sim 3.6 \times 10^{12}~\mathrm{cm}~\mathrm{s}^{-2}$ and $H_0\sim 10^3~\mathrm{cm}$, then for the $500~\mathrm{Hz}$ system we find the Rossby length to approximately be $L_R \sim 1.2 \times 10^5~\mathrm{cm}$.
+\added{If we take $g \sim 1.5 \times 10^{14}~\mathrm{cm}~\mathrm{s}^{-2}$ and $H_0\sim 10^3~\mathrm{cm}$, then for the $500~\mathrm{Hz}$ system we find the Rossby length to approximately be $L_R \sim 1.2 \times 10^5~\mathrm{cm}$.
 This estimate is likely to be an overestimate of the true value of the Rossby length in our simulations, as we saw that the flame is confined on a much smaller length scale.} In Figure \ref{fig:time_series_enuc_500} and Figure

### 2021-03-18 strang/paper.tex
fix abstract

@@
 temperature with the species.  Here we demonstrate through a simple
-convergence test that integrating temperature together with reactions
+convergence test that integrating an energy equation together with reactions
 achieves the best convergence when modeling reactive flows with Strang

### 2021-03-18 strang/paper.tex
word count

@@
 \end{align}
-where $\rho$ is the density, $\Ub$ is the velocity vector, $X_k$ are the species mass fractions,
-$\omegadot_k$ are the species creation rates (from reactions), $p$ is the pressure, $E$ is the
+where $\rho$ is the density, $\Ub$ is the velocity vector, $X_k$ are the species mass fractions
+with creation rates $\omegadot_k$ from reactions, $p$ is the pressure, $E$ is the
 specific total energy, and $\Sdot$ is the nuclear energy generation rate.
@@
 \end{equation}
-where $e$ is the specific internal energy or alternately, we can evolve the temperature, 
-which we can write as
+where $e$ is the specific internal energy or alternately, we can evolve the temperature, $T$,
+via
 \begin{equation}
-\rho c_v \frac{DT}{Dt} = \rho \left (\frac{p}{\rho^2} - e_\rho \right ) \frac{D\rho}{Dt} + \rho \Sdot,
+\rho c_v \frac{DT}{Dt} = \rho \left (\frac{p}{\rho^2} - \left .\frac{\partial e}{\partial \rho} \right |_T \right ) \frac{D\rho}{Dt} + \rho \Sdot,
 \end{equation}
-where $c_v$ is the specific heat capacity at constant volume, $T$ is the temperature and $e_\rho$ is the derivative of 
-the specific internal energy with respect to density (this form neglects the change in 
-temperature due to composition
-changes from reactions, see \citealt{ABNZ:III}, which assumes chemical
+where $c_v = \partial e / \partial T |_\rho$ (this form neglects composition
+changes, see \citealt{ABNZ:III}, which assumes chemical
 equilibrium).
@@
 
-\acknowledgements \castro\ and all of the code used here is freely available at
+\acknowledgements \castro\ is freely available at
 \url{https://github.com/AMReX-Astro/Castro}.   The work at
@@
 \software{\amrex\ \citep{amrex_joss},
-          \castro\ \citep{castro},
-          GNU Compiler Collection (\url{https://gcc.gnu.org/}),
-          matplotlib (\citealt{Hunter:2007},  \url{http://matplotlib.org/})
-          NumPy \citep{numpy,numpy2},
-          python (\url{https://www.python.org/}),
+          \castro\ \citep{castro_joss},
+          matplotlib \citep{Hunter:2007},
+          NumPy \citep{numpy},
           VODE \citep{vode}

### 2021-03-22 strang/paper.tex
new aastex
Fix affiliation
Non-parenthetical cite

@@
-\documentclass[times,preprint]{aastex63}
+\documentclass[times,preprint]{aastex631}
 
@@
 \author[0000-0003-0439-4556]{M.~P.~Katz}
-\affiliation{Nvidia Corp}
+\affiliation{NVIDIA Corporation}
 
@@
 with reactions to yield better overall convergence and accuracy.  This
-complements the work of \citep{muller:1986} which showed that when
+complements the work of \cite{muller:1986} which showed that when
 evolving near nuclear statistical equilibrium, evolving entropy
@@
 
-\acknowledgements \castro\ is freely available at
+\begin{acknowledgments}
+
+\castro\ is freely available at
 \url{https://github.com/AMReX-Astro/Castro}.   The work at
@@
 Nuclear Security Administration.
+
+\end{acknowledgments}
 

### 2021-03-22 strang/paper.tex
word count

@@
 achieves the best convergence when modeling reactive flows with Strang
-splitting.
+splitting.  Additionally, second order convergence cannot be achieved without integrating an energy or temperature equation.
 \end{abstract}
@@
 where $\rho$ is the density, $\Ub$ is the velocity vector, $X_k$ are the species mass fractions
-with creation rates $\omegadot_k$ from reactions, $p$ is the pressure, $E$ is the
+with creation rates $\omegadot_k$, $p$ is the pressure, $E$ is the
 specific total energy, and $\Sdot$ is the nuclear energy generation rate.
@@
 where $e$ is the specific internal energy or alternately, we can evolve the temperature, $T$,
-via
 \begin{equation}
@@
 where $c_v = \partial e / \partial T |_\rho$ (this form neglects composition
-changes, see \citealt{ABNZ:III}, which assumes chemical
-equilibrium).
+changes, see \citealt{ABNZ:III}).
 
@@
 pure $\isotm{He}{4}$, but both $\isotm{C}{12}$ and $\isotm{O}{16}$ are
-created as time evolves.  The published tests focused on SDC coupling
-and showed that we can get overall 4th order in space and time
-convergence.  Here we run the same test with Strang coupling, looking
+created as time evolves.  The published tests 
+showed that we can get overall 4th order in space and time
+convergence with SDC coupling.  Here we run the same test with Strang coupling, looking
 at the various approaches at incorporating a temperature / energy equation in
@@
 resolution.  The slope of these lines is a measure of the convergence
-rate~\citep{ORVV}.  Guidelines show ideal first and second order
-convergence.  We see that all methods converge at least second order
+rate~\citep{ORVV}.  We see that all methods converge at least second order
 for density, but for the thermodynamic quantities, $(\rho e)$ and $T$,

### 2021-03-26 strang/paper.tex
address editor's comments

@@
 \begin{abstract}
-For astrophysical reacting flows, operator splitting is commonly used
-to couple hydrodynamics and reactions.  Each process operates
-independent of one another, but by staggering the updates in a
-symmetric fashion (via Strang splitting) second order accuracy in time
-can be achieved.  However, approximations are often made to the
-reacting system, including the choice of whether or not to integrate
-temperature with the species.  Here we demonstrate through a simple
-convergence test that integrating an energy equation together with reactions
-achieves the best convergence when modeling reactive flows with Strang
-splitting.  Additionally, second order convergence cannot be achieved without integrating an energy or temperature equation.
+Many astrophysical environments involve convective or explosive flows
+driven by thermonuclear reactions (Type Ia supernovae, classical
+novae, X-ray bursts, stellar evolution).  Simulation codes need to
+accurately capture the interactions between reactions and
+hydrodynamics to produce realistic models of these events.  For
+astrophysical reacting flows, operator splitting is commonly used to
+couple hydrodynamics and reactions.  Each process operates independent
+of one another, but by staggering the updates in a symmetric fashion
+(via Strang splitting) second order accuracy in time can be achieved.
+However, approximations are often made to the reacting system,
+including the choice of whether or not to integrate temperature with
+the species.  Here we demonstrate through a simple convergence test
+that integrating an energy equation together with reactions achieves
+the best convergence when modeling reactive flows with Strang
+splitting.  Additionally, second order convergence cannot be achieved
+without integrating an energy or temperature equation.
 \end{abstract}
@@
 
-Simulations of stellar flows require solving the equations of
+Type Ia supernovae, classical novae, X-ray bursts, and general stellar
+evolution reactive flows, where the nuclear energy release can drive
+strong convective or explosive flows.  These stellar flows require solving the equations of
 hydrodynamics coupled with a nuclear reaction network.  The equations
@@
 resolution.  The slope of these lines is a measure of the convergence
-rate~\citep{ORVV}.  We see that all methods converge at least second order
+rate~\citep{ORVV}.  We see that all methods converge to at least second order
 for density, but for the thermodynamic quantities, $(\rho e)$ and $T$,

### 2021-03-26 strang/paper.tex
add a verb

@@
 Type Ia supernovae, classical novae, X-ray bursts, and general stellar
-evolution reactive flows, where the nuclear energy release can drive
+evolution involve reactive flows, where the nuclear energy release can drive
 strong convective or explosive flows.  These stellar flows require solving the equations of

### 2021-04-06 simplified_sdc_intro/paper.tex
add RNAAS

@@
 where $K$ is the kinetic energy, $K = |u|^2/2$.  Since the density and velocity
-are unchanged by reactions, we consider the following:
-\begin{equation}
+are unchanged by reactions, our energy equation becomes:
+\begin{equation}
+\label{eq:strang:e}
 \odt{(\rho e)} = \rho \odt{e} = \rho \Sdot
@@
 The reaction rates are typically expressed as $\omegadot_k(\rho, T, X_k)$
-when we evolve this system.  This requires an EOS inversion to get $T$ from $e$ each
+when we evolve this system, which requires us to get temperature from
+the equation of state each
 time we need to evaluate the reactive terms.
-We choose to instead integrate the temperature.
-We can write the energy equation as:
-\begin{equation}
-\rho \odt{e} = \rho \left [ \underbrace{\left . \frac{\partial e}{\partial T} \right |_\rho}_{c_v}
-                            \odt{T} +
-                            \left . \frac{\partial e}{\partial \rho} \right |_T
-                            \cancelto{0}{\odt{\rho}} \;\; \right ] = \rho \Sdot
-\end{equation}
-reducing to
-\begin{equation}
-\label{eq:strang:T}
-\odt{T} = \frac{1}{c_v} \Sdot
-\end{equation}
-We can alternately derive this from an enthalpy equation, which is the
-correct formulation when we are at constant pressure, and we get the
-same temperature equation but with $c_v$ replaced by $c_p$.
-Technically this requires an EOS call to get $c_x$, but if the
-specific heat varies slowly, we can freeze it during the reaction
-solve or use a Taylor expansion in $T$ to capture slow evolution.
 
@@
 \end{equation}
-We integrate Equations (\ref{eq:strang:T}) and (\ref{eq:strang:X}) using
+We integrate Equations (\ref{eq:strang:e}) and (\ref{eq:strang:X}) using
 an implicit ODE solver designed for stiff systems of
@@
 
-With Strang splitting, the state can drift significantly off of the
-smooth solution to the coupled reactive hydrodynamics equations, as
-shown graphically in \cite{astronum:2018} using an earlier version of
-the present algorithm.
+We explored this and other approaches (including not evolving an
+energy / temperature equation during the reaction step) in
+\citep{strang_rnaas} and showed that the above formulation gets second
+order convergence.  However, for very strong reactions, when using
+Strang splitting the state can drift significantly off of the smooth
+solution to the coupled reactive hydrodynamics equations, as shown
+graphically in \cite{astronum:2018} using an earlier version of the
+present algorithm.
 

### 2021-04-06 simplified_sdc_intro/paper.tex
add NSE bailout description

@@
 
+\subsection{NSE Bailout}
+
+There is one additional part of the time integration algorithm.
+Because $T$ evolves during the reactions (for both Strang and SDC), it
+is possible for a zone to start out not in NSE but evolve into NSE
+during the reaction update.  When this happens, the ODE integrator may
+fail, because an excessive number of timesteps is required.  Rather
+than trigger our retry mechanism and throw away the entire timestep
+and start over with a smaller $\Delta t$, we first check if the zone
+evolved into NSE, and if so, we redo the zone update using the NSE
+prescription.
+
 \section{Simulations}
@@
 Also explore: PLM vs PPM, 1 vs. 2 vs. 3 iterations
+
+% run on 04-06-21
+% Castro       git describe: 21.04-dirty
+% AMReX        git describe: 21.03-58-g87c81d0ca
+% Microphysics git describe: 21.04-1-g33a4d0fe
 
@@
 
-\begin{deluxetable}{lllllll}
-\tablecaption{\label{table:react_convert_strang} Convergence ($L_1$ norm) for the reacting convergence problem using Strang splitting with $T$ evolution disabled.}
-\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
-\startdata
- $\rho$                      & $2.735 \times 10^{18}$  & 2.057  & $6.571 \times 10^{17}$  & 2.474  & $1.183 \times 10^{17}$  \\
- $\rho u$                    & $6.729 \times 10^{26}$  & 2.416  & $1.260 \times 10^{26}$  & 2.699  & $1.941 \times 10^{25}$  \\
- $\rho v$                    & $6.729 \times 10^{26}$  & 2.416  & $1.260 \times 10^{26}$  & 2.699  & $1.941 \times 10^{25}$  \\
- $\rho E$                    & $2.513 \times 10^{35}$  & 2.226  & $5.372 \times 10^{34}$  & 2.169  & $1.195 \times 10^{34}$  \\
- $\rho e$                    & $2.292 \times 10^{35}$  & 2.194  & $5.011 \times 10^{34}$  & 2.253  & $1.051 \times 10^{34}$  \\
- $T$                         & $2.321 \times 10^{21}$  & 1.606  & $7.623 \times 10^{20}$  & 1.993  & $1.915 \times 10^{20}$  \\
- $\rho X(\isotm{He}{4})$     & $2.806 \times 10^{18}$  & 2.028  & $6.880 \times 10^{17}$  & 2.432  & $1.275 \times 10^{17}$  \\
- $\rho X(\isotm{C}{12})$     & $1.807 \times 10^{17}$  & 1.846  & $5.026 \times 10^{16}$  & 1.605  & $1.653 \times 10^{16}$  \\
- $\rho X(\isotm{O}{16})$     & $1.905 \times 10^{14}$  & 1.500  & $6.737 \times 10^{13}$  & 1.439  & $2.484 \times 10^{13}$  \\
- $\rho X(\isotm{Fe}{56})$    & $2.735 \times 10^{8}$   & 2.057  & $6.571 \times 10^{7}$   & 2.474  & $1.183 \times 10^{7}$   \\
-\enddata
-\end{deluxetable}
-
-% run on 12-31-20
-% Castro: 20.12-80-gfe3e9c8de8
+% run on 04-06-21
+% Castro       git describe: 21.04-dirty
+% AMReX        git describe: 21.03-58-g87c81d0ca
+% Microphysics git describe: 21.04-1-g33a4d0fe
+
 \begin{deluxetable}{lllllll}

### 2021-04-24 simplified_sdc_intro/paper.tex
ome updates
update NSE algorithm

@@
 With SDC evolution, when we are in NSE, we need to do the advective
-and reactive updates together.  With the tabulated NSE reaction
-values, we get instantaneous values of the reaction sources, but the only
-time derivative we have is for $Y_e$.  We will
-use a basic predictor-corrector update:
-\begin{equation}
-\Uc^\star = \Uc^n + \Delta t \Adv{\Uc}^{n+1/2} + \Delta t \left [\Rb (\Uc) \right ]^{n+1/2}
-\end{equation}
+and reactive updates together.  We want to compute:
+\begin{equation}
+\Uc^{n+1,(k)} = \Uc^n + \Delta t \Adv{\Uc}^{n+1/2,(k)} + \Delta t \left [\Rb (\Uc) \right ]^{n+1/2,(k)}
+\end{equation}
+For density and momentum, we can do this update already, since there
+are no reactive sources.  That gives us $\rho^{n+1,(k)}$ and $(\rho
+\Ub)^{n+1,(k)}$.  For the other quantities, the NSE table gives only the
+instantaneous values. aside from $Y_e$.  We therefore do an iterative
+update.
+
 Calling the NSE table on the starting state gives us:
@@
 \end{equation}
-To get the reactive source, we need to use the iteratively-lagged time-centered information
-from $\Ic_q^{n+1/2,(k-1)}$.  We only need to update the internal energy and auxiliary data,
-since together with the density, that is all that is needed for the equation of state to get the
-updated temperature.    So our sources are:
-\begin{equation}
-R(\rho e) \approx \Ics_q(\rho e)
-\end{equation}
-and
-\begin{equation}
-R(\rho \alpha_l) \approx \rho^{n+1/2} \Ics_q(\alpha_l)
-\end{equation}
-where $\rho^{n+1/2} = (\rho^n + \rho^\star)/2$.
-
-Since we have $(dY_e/dt)^n$ from the table, we replace the source for $Y_e$ in the above
-$\Rb(\Uc^n)$ as: \MarginPar{it might be better if we had $\Ics_\Uc$}
-\begin{equation}
-R(\rho Y_e) = \rho^{n+1/2} (dY_e/dt)^n
-\end{equation}
-which gives an update for the electron fraction as:
-\begin{equation}
-(\rho Y_e)^\star = (\rho Y_e)^n + \Delta t \Advss{\rho Y_e}^{n+1/2} + \Delta t \rho^{n+1/2} (dY_e/dt)^n
-\end{equation}
-We can then get $\alpha_l^\star = (\rho \alpha_l)^\star/\rho^\star$.
-From this prediction, we get $\rho^\star$, $\alpha_l^\star$, and
-$e^\star$, which allows us to call the EOS and get an updated
-temperature, $T^\star$. \MarginPar{Andy suggests time-centering $dY_e/dt$}
-\begin{equation}
-T^\star = T(\rho^\star, \alpha_l^\star, e^\star)
-\end{equation}
-We can then again call the NSE table:
-\begin{equation}
-\nse{\rho^\star, T^\star, (Y_e)^\star} \rightarrow \bar{A}^{n+1,(k)}, (X_k)^{n+1,(k)}, (B/A)^{n+1,(k)}, (dY_e/dt)
-\end{equation}
-where we use the outputs to define the new values of the composition variables.
-Since the density and momenta don't have reactive sources, we just use the $\Uc^\star$ state
-as their new values:
+We now compute a first approximation to the reactive source:
 \begin{align}
-\rho^{n+1,(k)} &= \rho^\star \\
-(\rho u)^{n+1,(k)} &= (\rho u)^\star
+R^{(0)}(\rho e) &= 0 \\
+R^{(0)}(\rho Y_e) &= \rho^n (dY_e/dt)^n \\
+R^{(0)}(\rho \bar{A}) &= 0
 \end{align}
-We then compute the energy release as
-\begin{equation}
-       (\rho \Sdot)^{n+1/2} = \frac{\rho^{n+1,(k)} + \rho^n}{2}
-             \left [ \left ( \frac{B}{A} \right )^{n+1,(k)} -
+
+Now we iterate $\xi = 0, \ldots N-1$, doing the following:
+
+\begin{itemize}
+\item Update the energy as:
+   \begin{equation}
+     (\rho e)^{n+1,(\xi)} = (\rho e)^n + \Delta t \Advss{\rho e}^{n+1/2,(k)} + \Delta t R^{(\xi)}(\rho e)
+   \end{equation}
+
+\item Update the auxiliary data as:
+   \begin{equation}
+     (\rho \alpha_l)^{n+1,(\xi)} = (\rho \alpha_l)^n + \Delta t \Advss{\rho \alpha_l}^{n+1/2,(k)} + \Delta t R^{(\xi)}(\rho \alpha_l)
+   \end{equation}
+
+\item Compute the updated temperature, $T^{n+1,(k)}$ as
+  \begin{equation}
+    T^{n+1,(\xi)} = T(\rho^{n+1,(k)},  \alpha^{n+1,(\xi)},  e^{n+1,(\xi)})
+  \end{equation}
+  where $\alpha^{n+1,(\xi)} = (\rho  \alpha)^{n+1,(\xi)} / \rho^{n+1,(k)}$ and $e^{n+1,(\xi)} = (\rho e)^{n+1,(\xi)} / \rho^{n+1,(k)}$.
+
+\item Call the NSE table to get the new NSE state
+  \begin{equation}
+    \nse{\rho^{n+1,(k)}, T^{n+1,(\xi)}, (Y_e)^{n+1,(\xi)}} \rightarrow \bar{A}^{n+1,(\xi)}, (X_k)^{n+1,(\xi)}, (B/A)^{n+1,(\xi)}, (dY_e/dt)^{n+1,(\xi)}
+  \end{equation}
+
+\item Recompute the reactive source terms for the next iteration.
+
+  The energy generation rate is found as:
+  \begin{equation}
+    (\rho \Sdot)^{n+1/2,(\xi)} = \frac{\rho^{n+1,(k)} + \rho^n}{2}
+    \left [ \left ( \frac{B}{A} \right )^{n+1,(\xi)} -
       \left ( \frac{B}{A} \right )^n \right ] N_A \frac{1}{\Delta t}
   \end{equation}
-
-{\color{red}
-Note I tried this first, but it really doesn't work.  I'd like to understand better:
-\begin{equation}
-       (\rho \Sdot)^{n+1/2} = \left [ \rho^{n+1,(k)} \left ( \frac{B}{A} \right )^{n+1,(k)} -
-                            \rho^n \left ( \frac{B}{A} \right )^n \right ] N_A \frac{1}{\Delta }t
-\end{equation}
-}
-and then do the final update of the total energy:
-\begin{equation}
-(\rho E)^{n+1,(k)} = (\rho E)^n + \Delta t \Advss{\rho E}^{n+1/2} + \Delta t (\rho \Sdot)^{n+1/2}
-\end{equation}
+  and our new auxiliary sources are:
+\begin{align}
+R^{(\xi+1)}(\rho e) &= (\rho \Sdot)^{n+1/2,(\xi)} \\
+R^{(\xi+1)}(\rho Y_e) &= \frac{\rho^{n+1,(k)} + \rho^n}{2} \frac{1}{2} \left [ \left (\frac{dY_e}{dt} \right )^n + \left (\frac{dY_e}{dt} \right )^{n+1,(\xi)} \right ] \\
+R^{(\xi+1)}(\rho \bar{A}) &= \frac{\rho^{n+1,(k)} + \rho^n}{2} \frac{1}{\Delta t} \left [ \bar{A}^{n+1,(\xi)} - \frac{(\rho \bar{A})^n}{\rho^n} \right ]
+\end{align}
+
+\end{itemize}
+
+With the iteration complete, we can do the final update of the total energy using last iterations prediction of the source (for iteration $N$):
+\begin{equation}
+(\rho E)^{n+1,(k)} = (\rho E)^n + \Delta t \Advss{\rho E}^{n+1/2} + \Delta t (\rho \Sdot)^{n+1/2,(N)}
+\end{equation}
+and we take the values of the auxiliary state and mass fractions from the last call to the NSE table:
+\begin{align}
+(\rho Y_e)^{n+1,(k)} &= \rho^{n+1,(k)} (Y_e)^{n+1,(N-1)} \\
+(\rho \bar{A})^{n+1,(k)} &= \rho^{n+1,(k)} (\bar{A})^{n+1,(N-1)} \\
+(\rho (B/A))^{n+1,(k)} &= \rho^{n+1,(k)} (B/A)^{n+1,(N-1)} \\
+(\rho Y_e)^{n+1,(k)} &= \rho^{n+1,(k)} (X_k)^{n+1,(N-1)}
+\end{align}
+
+We note that this update is not fully second order, however, the
+treatment of $Y_e$ is, and we expect that to have the dominant effect
+in the thermodynamic evolution through NSE.
 
@@
 \startdata
- $\rho$                      & $2.781 \times 10^{18}$  & 2.051  & $6.711 \times 10^{17}$  & 2.580  & $1.122 \times 10^{17}$  \\
- $\rho u$                    & $6.781 \times 10^{26}$  & 2.445  & $1.245 \times 10^{26}$  & 2.907  & $1.659 \times 10^{25}$  \\
- $\rho v$                    & $6.781 \times 10^{26}$  & 2.445  & $1.245 \times 10^{26}$  & 2.907  & $1.659 \times 10^{25}$  \\
- $\rho E$                    & $2.465 \times 10^{35}$  & 2.333  & $4.892 \times 10^{34}$  & 2.650  & $7.794 \times 10^{33}$  \\
- $\rho e$                    & $2.268 \times 10^{35}$  & 2.298  & $4.610 \times 10^{34}$  & 2.722  & $6.988 \times 10^{33}$  \\
- $T$                         & $2.248 \times 10^{21}$  & 1.682  & $7.003 \times 10^{20}$  & 2.439  & $1.291 \times 10^{20}$  \\
- $\rho X(\isotm{He}{4})$     & $2.862 \times 10^{18}$  & 2.027  & $7.023 \times 10^{17}$  & 2.554  & $1.196 \times 10^{17}$  \\
- $\rho X(\isotm{C}{12})$     & $1.717 \times 10^{17}$  & 1.945  & $4.460 \times 10^{16}$  & 2.194  & $9.745 \times 10^{15}$  \\
- $\rho X(\isotm{O}{16})$     & $1.718 \times 10^{14}$  & 1.648  & $5.482 \times 10^{13}$  & 1.898  & $1.471 \times 10^{13}$  \\
- $\rho X(\isotm{Fe}{56})$    & $2.781 \times 10^{8}$   & 2.051  & $6.711 \times 10^{7}$   & 2.580  & $1.122 \times 10^{7}$   \\
+ $\rho$                      & $2.794 \times 10^{18}$  & 2.044  & $6.777 \times 10^{17}$  & 2.554  & $1.154 \times 10^{17}$  \\
+ $\rho u$                    & $6.796 \times 10^{26}$  & 2.448  & $1.245 \times 10^{26}$  & 2.889  & $1.681 \times 10^{25}$  \\
+ $\rho v$                    & $6.796 \times 10^{26}$  & 2.448  & $1.245 \times 10^{26}$  & 2.889  & $1.681 \times 10^{25}$  \\
+ $\rho E$                    & $2.451 \times 10^{35}$  & 2.351  & $4.803 \times 10^{34}$  & 2.742  & $7.179 \times 10^{33}$  \\
+ $\rho e$                    & $2.261 \times 10^{35}$  & 2.320  & $4.526 \times 10^{34}$  & 2.821  & $6.403 \times 10^{33}$  \\
+ $T$                         & $2.237 \times 10^{21}$  & 1.691  & $6.927 \times 10^{20}$  & 2.482  & $1.240 \times 10^{20}$  \\
+ $\rho X(\isotm{He}{4})$     & $2.878 \times 10^{18}$  & 2.020  & $7.096 \times 10^{17}$  & 2.529  & $1.229 \times 10^{17}$  \\
+ $\rho X(\isotm{C}{12})$     & $1.698 \times 10^{17}$  & 1.950  & $4.393 \times 10^{16}$  & 2.232  & $9.353 \times 10^{15}$  \\
+ $\rho X(\isotm{O}{16})$     & $1.687 \times 10^{14}$  & 1.660  & $5.338 \times 10^{13}$  & 1.957  & $1.375 \times 10^{13}$  \\
+ $\rho X(\isotm{Fe}{56})$    & $2.794 \times 10^{8}$   & 2.044  & $6.777 \times 10^{7}$   & 2.554  & $1.154 \times 10^{7}$   \\
 \enddata
@@
 \startdata
- $\rho$                      & $2.794 \times 10^{18}$  & 2.047  & $6.760 \times 10^{17}$  & 2.563  & $1.144 \times 10^{17}$  \\
- $\rho u$                    & $6.796 \times 10^{26}$  & 2.450  & $1.243 \times 10^{26}$  & 2.898  & $1.669 \times 10^{25}$  \\
- $\rho v$                    & $6.796 \times 10^{26}$  & 2.450  & $1.243 \times 10^{26}$  & 2.898  & $1.669 \times 10^{25}$  \\
- $\rho E$                    & $2.450 \times 10^{35}$  & 2.357  & $4.783 \times 10^{34}$  & 2.743  & $7.144 \times 10^{33}$  \\
- $\rho e$                    & $2.261 \times 10^{35}$  & 2.325  & $4.513 \times 10^{34}$  & 2.821  & $6.385 \times 10^{33}$  \\
- $T$                         & $2.236 \times 10^{21}$  & 1.690  & $6.929 \times 10^{20}$  & 2.473  & $1.248 \times 10^{20}$  \\
- $\rho X(\isotm{He}{4})$     & $2.877 \times 10^{18}$  & 2.023  & $7.078 \times 10^{17}$  & 2.536  & $1.220 \times 10^{17}$  \\
- $\rho X(\isotm{C}{12})$     & $1.697 \times 10^{17}$  & 1.947  & $4.402 \times 10^{16}$  & 2.222  & $9.436 \times 10^{15}$  \\
- $\rho X(\isotm{O}{16})$     & $1.688 \times 10^{14}$  & 1.658  & $5.349 \times 10^{13}$  & 1.950  & $1.385 \times 10^{13}$  \\
- $\rho X(\isotm{Fe}{56})$    & $2.794 \times 10^{8}$   & 2.047  & $6.760 \times 10^{7}$   & 2.563  & $1.144 \times 10^{7}$   \\
+ $\rho$                      & $2.795 \times 10^{18}$  & 2.044  & $6.777 \times 10^{17}$  & 2.554  & $1.154 \times 10^{17}$  \\
+ $\rho u$                    & $6.798 \times 10^{26}$  & 2.449  & $1.245 \times 10^{26}$  & 2.889  & $1.680 \times 10^{25}$  \\
+ $\rho v$                    & $6.798 \times 10^{26}$  & 2.449  & $1.245 \times 10^{26}$  & 2.889  & $1.680 \times 10^{25}$  \\
+ $\rho E$                    & $2.452 \times 10^{35}$  & 2.352  & $4.804 \times 10^{34}$  & 2.742  & $7.179 \times 10^{33}$  \\
+ $\rho e$                    & $2.262 \times 10^{35}$  & 2.321  & $4.527 \times 10^{34}$  & 2.822  & $6.404 \times 10^{33}$  \\
+ $T$                         & $2.236 \times 10^{21}$  & 1.691  & $6.927 \times 10^{20}$  & 2.481  & $1.240 \times 10^{20}$  \\
+ $\rho X(\isotm{He}{4})$     & $2.879 \times 10^{18}$  & 2.020  & $7.096 \times 10^{17}$  & 2.529  & $1.229 \times 10^{17}$  \\
+ $\rho X(\isotm{C}{12})$     & $1.697 \times 10^{17}$  & 1.950  & $4.393 \times 10^{16}$  & 2.231  & $9.357 \times 10^{15}$  \\
+ $\rho X(\isotm{O}{16})$     & $1.686 \times 10^{14}$  & 1.659  & $5.338 \times 10^{13}$  & 1.957  & $1.375 \times 10^{13}$  \\
+ $\rho X(\isotm{Fe}{56})$    & $2.795 \times 10^{8}$   & 2.044  & $6.777 \times 10^{7}$   & 2.554  & $1.154 \times 10^{7}$   \\
 \enddata
@@
 
+% run on 04-06-21
+% Castro       git describe: 21.04-dirty
+% AMReX        git describe: 21.03-58-g87c81d0ca
+% Microphysics git describe: 21.04-1-g33a4d0fe
+
 \begin{deluxetable}{lllllll}
@@
  $T$                         & $9.885 \times 10^{20}$  & 1.207  & $4.281 \times 10^{20}$  & 1.058  & $2.057 \times 10^{20}$  \\
+ $\rho X(\isotm{H}{1})$      & $4.522 \times 10^{-11}$ & 1.659  & $1.432 \times 10^{-11}$ & 1.542  & $4.917 \times 10^{-12}$ \\
  $\rho X(\isotm{He}{4})$     & $5.086 \times 10^{17}$  & 1.430  & $1.887 \times 10^{17}$  & 1.187  & $8.288 \times 10^{16}$  \\
@@
 \end{deluxetable}
+
+% run on 04-06-21
+% Castro       git describe: 21.04-dirty
+% AMReX        git describe: 21.03-58-g87c81d0ca
+% Microphysics git describe: 21.04-1-g33a4d0fe
 
@@
  $T$                         & $5.873 \times 10^{20}$  & 1.621  & $1.910 \times 10^{20}$  & 1.494  & $6.779 \times 10^{19}$  \\
+ $\rho X(\isotm{H}{1})$      & $4.558 \times 10^{-11}$ & 1.633  & $1.470 \times 10^{-11}$ & 1.578  & $4.923 \times 10^{-12}$ \\
  $\rho X(\isotm{He}{4})$     & $2.453 \times 10^{17}$  & 1.835  & $6.876 \times 10^{16}$  & 1.657  & $2.181 \times 10^{16}$  \\

### 2021-04-29 simplified_sdc_intro/paper.tex
update table
some discussion on interface states

@@
       \end{equation}
+
+      In the unsplit CTU method \citep{ppmunsplit}, the interface
+      states used for the final Riemann problem through the zone
+      interface consist of a normal predictor and a transverse flux
+      correction.  We can add the source terms either to the normal
+      predictor (for example, doing characteristic tracing as
+      described in \citealt{ppm}) or after all of the transverse flux
+      corrections are made.  Both are second-order accurate.  For the
+      hydrodynamics sources, we do those in the normal predictor.
+      However, for the reactive sources, we found that it is most
+      reliable to add them at the end of the interface state
+      construction, after the transverse flux corrections.  This is
+      because we want to ensure that sum over species of $\Ic_\qb$ is
+      zero, and characteristic tracing or the various flux corrections
+      may not preserve this.
 
@@
 \startdata
- $\rho$                      & $4.558 \times 10^{19}$  & 1.633  & $1.470 \times 10^{19}$  & 1.578  & $4.923 \times 10^{18}$  \\
- $\rho u$                    & $1.945 \times 10^{28}$  & 1.648  & $6.208 \times 10^{27}$  & 1.513  & $2.176 \times 10^{27}$  \\
- $\rho v$                    & $1.945 \times 10^{28}$  & 1.648  & $6.208 \times 10^{27}$  & 1.513  & $2.176 \times 10^{27}$  \\
- $\rho E$                    & $5.710 \times 10^{37}$  & 1.620  & $1.857 \times 10^{37}$  & 1.552  & $6.333 \times 10^{36}$  \\
- $\rho e$                    & $5.708 \times 10^{37}$  & 1.620  & $1.857 \times 10^{37}$  & 1.553  & $6.331 \times 10^{36}$  \\
- $T$                         & $5.873 \times 10^{20}$  & 1.621  & $1.910 \times 10^{20}$  & 1.494  & $6.779 \times 10^{19}$  \\
- $\rho X(\isotm{H}{1})$      & $4.558 \times 10^{-11}$ & 1.633  & $1.470 \times 10^{-11}$ & 1.578  & $4.923 \times 10^{-12}$ \\
- $\rho X(\isotm{He}{4})$     & $2.453 \times 10^{17}$  & 1.835  & $6.876 \times 10^{16}$  & 1.657  & $2.181 \times 10^{16}$  \\
- $\rho X(\isotm{C}{12})$     & $1.130 \times 10^{12}$  & 1.846  & $3.145 \times 10^{11}$  & 1.678  & $9.831 \times 10^{10}$  \\
- $\rho X(\isotm{O}{16})$     & $5.339 \times 10^{12}$  & 1.830  & $1.502 \times 10^{12}$  & 1.667  & $4.730 \times 10^{11}$  \\
- $\rho X(\isotm{Cr}{48})$    & $6.880 \times 10^{18}$  & 1.791  & $1.988 \times 10^{18}$  & 1.601  & $6.554 \times 10^{17}$  \\
- $\rho X(\isotm{Fe}{52})$    & $3.470 \times 10^{19}$  & 1.606  & $1.140 \times 10^{19}$  & 1.608  & $3.740 \times 10^{18}$  \\
- $\rho X(\isotm{Fe}{54})$    & $4.766 \times 10^{20}$  & 1.764  & $1.403 \times 10^{20}$  & 1.777  & $4.093 \times 10^{19}$  \\
- $\rho X(\isotm{Ni}{56})$    & $4.835 \times 10^{20}$  & 1.781  & $1.407 \times 10^{20}$  & 1.788  & $4.075 \times 10^{19}$  \\
- $\rho Y_e$                  & $2.370 \times 10^{19}$  & 1.644  & $7.582 \times 10^{18}$  & 1.580  & $2.536 \times 10^{18}$  \\
- $\rho \bar{A}$              & $4.061 \times 10^{21}$  & 1.678  & $1.270 \times 10^{21}$  & 1.675  & $3.975 \times 10^{20}$  \\
- $\rho (B/A)$                & $4.224 \times 10^{20}$  & 1.664  & $1.333 \times 10^{20}$  & 1.602  & $4.391 \times 10^{19}$  \\
+ $\rho$                      & $4.553 \times 10^{19}$  & 1.636  & $1.465 \times 10^{19}$  & 1.583  & $4.891 \times 10^{18}$  \\
+ $\rho u$                    & $1.949 \times 10^{28}$  & 1.650  & $6.210 \times 10^{27}$  & 1.513  & $2.176 \times 10^{27}$  \\
+ $\rho v$                    & $1.949 \times 10^{28}$  & 1.650  & $6.210 \times 10^{27}$  & 1.513  & $2.176 \times 10^{27}$  \\
+ $\rho E$                    & $5.691 \times 10^{37}$  & 1.622  & $1.848 \times 10^{37}$  & 1.561  & $6.266 \times 10^{36}$  \\
+ $\rho e$                    & $5.689 \times 10^{37}$  & 1.622  & $1.848 \times 10^{37}$  & 1.561  & $6.265 \times 10^{36}$  \\
+ $T$                         & $5.778 \times 10^{20}$  & 1.740  & $1.729 \times 10^{20}$  & 1.702  & $5.316 \times 10^{19}$  \\
+ $\rho X(\isotm{H}{1})$      & $4.553 \times 10^{-11}$ & 1.636  & $1.465 \times 10^{-11}$ & 1.583  & $4.891 \times 10^{-12}$ \\
+ $\rho X(\isotm{He}{4})$     & $2.541 \times 10^{17}$  & 1.807  & $7.262 \times 10^{16}$  & 1.634  & $2.339 \times 10^{16}$  \\
+ $\rho X(\isotm{C}{12})$     & $1.140 \times 10^{12}$  & 1.838  & $3.189 \times 10^{11}$  & 1.670  & $1.002 \times 10^{11}$  \\
+ $\rho X(\isotm{O}{16})$     & $5.413 \times 10^{12}$  & 1.819  & $1.535 \times 10^{12}$  & 1.658  & $4.864 \times 10^{11}$  \\
+ $\rho X(\isotm{Cr}{48})$    & $7.020 \times 10^{18}$  & 1.793  & $2.026 \times 10^{18}$  & 1.614  & $6.621 \times 10^{17}$  \\
+ $\rho X(\isotm{Fe}{52})$    & $3.631 \times 10^{19}$  & 1.579  & $1.215 \times 10^{19}$  & 1.597  & $4.018 \times 10^{18}$  \\
+ $\rho X(\isotm{Fe}{54})$    & $4.670 \times 10^{20}$  & 1.807  & $1.335 \times 10^{20}$  & 1.869  & $3.654 \times 10^{19}$  \\
+ $\rho X(\isotm{Ni}{56})$    & $4.858 \times 10^{20}$  & 1.808  & $1.387 \times 10^{20}$  & 1.860  & $3.820 \times 10^{19}$  \\
+ $\rho Y_e$                  & $2.372 \times 10^{19}$  & 1.649  & $7.564 \times 10^{18}$  & 1.587  & $2.518 \times 10^{18}$  \\
+ $\rho \bar{A}$              & $4.009 \times 10^{21}$  & 1.693  & $1.240 \times 10^{21}$  & 1.690  & $3.843 \times 10^{20}$  \\
+ $\rho (B/A)$                & $4.222 \times 10^{20}$  & 1.666  & $1.330 \times 10^{20}$  & 1.605  & $4.374 \times 10^{19}$  \\
 \enddata

### 2021-05-03 simplified_sdc_intro/paper.tex
update the Jacobian

@@
 \begin{equation}
+\label{eq:factored_jac}
 {\bf J} = \frac{\partial \Rb}{\partial {\bf w }} \frac{\partial {\bf w}}{\partial \Uc}.
@@
 We take the intermediate state to be ${\bf w} = (\rho, X_\alpha, X_\beta,
-K, T)^\intercal$, where $K$ is the kinetic energy:
+K, e)^\intercal$, where $K$ is the kinetic energy:
 \begin{equation}
@@
 \end{equation}
+Note: this intermediate state differs from what we used in \citet{castro_sdc} in that
+we are using $e$ instead of $T$ as an independent variable now.  This follows from
+the work shown in \citet{strang_rnaas} that explored integrating the Strang
+reaction system with an energy equation.  If the network gives derivatives in terms
+of temperature, we can convert as:
+\begin{equation}
+\left . \frac{\partial \phi}{\partial T} \right |_\rho =
+  \left . \frac{\partial \phi}{\partial e} \right |_\rho 
+  \left . \frac{\partial e}{\partial T} \right |_\rho =
+  c_v \left . \frac{\partial \phi}{\partial e} \right |_\rho
+\end{equation}
+for some quantity $\phi(\rho, T)$.  Here we identified the specific heat
+at constant volume, $c_v = \partial e/\partial T|_\rho$.
 
@@
        X_\beta & 0 & \rho & 0 & 0  \\
-       \rho e_\rho  + e + K &
+       e + K &
                  \rho  e_{X_\alpha} & \rho e_{X_\beta} & \rho &
-                 \rho e_T \\
-       \rho e_\rho  + e  &
+                 \rho \\
+       e  &
                  \rho  e_{X_\alpha} & \rho e_{X_\beta} & 0 &
-                 \rho e_T
+                 \rho
      \end{array}\right)
@@
 \begin{equation}
-e_\rho = \dedrd \qquad
-e_T = \dedTd \qquad
 e_{X_k} = \dedXd
 \end{equation}
-and write the kinetic energy as $K = u^2/2$.  We get the inverse
-(computed via SymPy, see the included Jupyter notebook) is:
-
+We get the inverse
+(computed via SymPy, see the included Jupyter notebook) as:
 \begin{equation}
@@
    - \frac{K}{\rho} & 0 & 0 & \frac{1}{\rho} & -\frac{1}{\rho} \\
-   \frac{\sum_k X_{k} e_{X_k} - \rho e_\rho - e}{ \rho e_T} &
-    -\frac{e_{X_\alpha}}{\rho e_T} & -\frac{e_{X_\beta}}{\rho e_T} & 0 & \frac{1}{\rho e_T} \\
+   \frac{\sum_k X_{k} e_{X_k}  - e}{ \rho} &
+    -\frac{e_{X_\alpha}}{\rho} & -\frac{e_{X_\beta}}{\rho} & 0 & \frac{1}{\rho} \\
    \end{array}\right)
@@
 
+While VODE can compute the entire Jacobian, ${\bf J}$ numerically via differencing, we found
+that does not give reliable results.  Instead, we compute $\partial {\bf w}/\partial {\bf U}^\prime$ analytically and $\partial {\bf R}/\partial {\bf w}$ numerically via one-sided differencing.
+The resulting Jacobian, as computed via equation~\ref{eq:factored_jac}, is larger than
+expected by the integrator because of the inclusion of the $\rho$
+terms in the intermediate steps, so we remove the row corresponding to
+$R(\rho)$ and the column corresponding to $\partial/\partial \rho$
+before passing the Jacobian to VODE.
+
 %======================================================================

### 2021-05-19 simplified_sdc_intro/paper.tex
some SdC updates

@@
 \newcommand{\Bnse}{{B_\mathrm{nse}}}
+\newcommand{\Cnse}{{C_\mathrm{nse}}}
 
@@
 ``aprox19'' network from \cite{aprox19} and originally described in
-\cite{Kepler}.
-We use a modified version of the VODE~\citep{vode} integrator---ported to C++ with checks added to the timestep rejection logic that ensure that the species mass fractions stay between 0 and 1\footnote{This modified version of VODE is available in our Microphysics repo: \url{https://github.com/starkiller-astro/Microphysics}.}.\MarginPar{discuss tolerances}  We combine this with a table that gives the nuclear
+\cite{Kepler}.  We use a modified version of the VODE~\citep{vode}
+integrator---ported to C++ with checks added to the timestep rejection
+logic that ensure that the species mass fractions stay between 0 and \MarginPar{we need to be clear that there is different logic for SDC}
+1\footnote{This modified version of VODE is available in our
+Microphysics repo:
+\url{https://github.com/starkiller-astro/Microphysics}.}.\MarginPar{discuss
+  tolerances} We combine this with a table that gives the nuclear
 statistical equilibrium (NSE) abundances in regions where the system
@@
 X(\isotm{C}{12}) &< \Anse \\
-X(\isotm{He}{4}) + \sum_{k \in \mathrm{Fe-group}} X_k &> \Bnse
+X(\isotm{He}{4}) + \sum_{k \in \mathrm{Fe-group}} X_k &> \Bnse \\
+X(\isotm{Si}{26}) &< \Cnse 
 \end{align}
-Typical values are $\rhonse = 3\times 10^8~\gcc$, $T_\mathrm{nse} =
-3\times 10^9~\mathrm{K}$, and $\Anse = 0.12$ and $\Bnse=
-0.88$---these are the values used in \citet{ma:2013}.
+Typical values are $\rhonse = 2\times 10^6~\gcc$, $T_\mathrm{nse} =
+3\times 10^9~\mathrm{K}$, $\Anse = 0.12$, $\Bnse=
+0.88$, and $\Cnse = 0.01$.  The values of $\Anse$ and $\Bnse$ are based on \citet{ma:2013}.
+The value of $\Cnse$ ensures that Si-burning has ended before invoking the table.
 

### 2021-07-13 simplified_sdc_intro/paper.tex

@@
 energy / temperature equation during the reaction step) in
-\citep{strang_rnaas} and showed that the above formulation gets second
+\citet{strang_rnaas} and showed that the above formulation gets second
 order convergence.  However, for very strong reactions, when using
@@
       may not preserve this.
+
+      {\color{red} We enforce that the species interface states remain in $[0, 1]$ after adding the reactive source.}
 
@@
 evolved into NSE, and if so, we redo the zone update using the NSE
-prescription.
+prescription.  {\color{red} We optionally apply a relaxation factor, typically, $0.9$, to the NSE conditions to allow a state that is close to NSE after an integrator failure enter NSE.}
 
@@
 While VODE can compute the entire Jacobian, ${\bf J}$ numerically via differencing, we found
-that does not give reliable results.  Instead, we compute $\partial {\bf w}/\partial {\bf U}^\prime$ analytically and $\partial {\bf R}/\partial {\bf w}$ numerically via one-sided differencing.
+that does not give reliable results.
+Instead, we compute $\partial {\bf w}/\partial {\bf U}^\prime$ analytically and $\partial {\bf R}/\partial {\bf w}$ numerically via one-sided differencing, following
+the algorithm in \cite{lsode} to minimize numerical noise.
 The resulting Jacobian, as computed via equation~\ref{eq:factored_jac}, is larger than

### 2021-07-13 simplified_sdc_intro/paper.tex
some more Jac work

@@
      \rho \frac{\partial \omegadot_\alpha}{\partial X_\alpha} &
-     \rho \frac{\partial \omegadot_\alpha}{\partial X_\beta} & 0 & \rho \frac{\partial \omegadot_\alpha}{\partial T} \\
+     \rho \frac{\partial \omegadot_\alpha}{\partial X_\beta} & 0 & \rho \frac{\partial \omegadot_\alpha}{\partial e} \\
       %
@@
      \rho \frac{\partial \omegadot_\beta}{\partial X_\alpha} &
-     \rho \frac{\partial \omegadot_\beta}{\partial X_\beta} & 0 &  \rho \frac{\partial \omegadot_\beta}{\partial T} \\
+     \rho \frac{\partial \omegadot_\beta}{\partial X_\beta} & 0 &  \rho \frac{\partial \omegadot_\beta}{\partial e} \\
      %
-     \Sdot + \rho \frac{\partial \Sdot}{\partial \rho} & \rho \frac{\partial \Sdot}{\partial X_\alpha} & \rho \frac{\partial \Sdot}{\partial X_\beta} & 0 & \rho \frac{\partial \Sdot}{\partial T} \\
+     \Sdot + \rho \frac{\partial \Sdot}{\partial \rho} & \rho \frac{\partial \Sdot}{\partial X_\alpha} & \rho \frac{\partial \Sdot}{\partial X_\beta} & 0 & \rho \frac{\partial \Sdot}{\partial e} \\
      %
-     \Sdot + \rho \frac{\partial \Sdot}{\partial \rho} & \rho \frac{\partial \Sdot}{\partial X_\alpha} & \rho \frac{\partial \Sdot}{\partial X_\beta} & 0 & \rho \frac{\partial \Sdot}{\partial T} \\
+     \Sdot + \rho \frac{\partial \Sdot}{\partial \rho} & \rho \frac{\partial \Sdot}{\partial X_\alpha} & \rho \frac{\partial \Sdot}{\partial X_\beta} & 0 & \rho \frac{\partial \Sdot}{\partial e} \\
   \end{array}
@@
 that does not give reliable results.
-Instead, we compute $\partial {\bf w}/\partial {\bf U}^\prime$ analytically and $\partial {\bf R}/\partial {\bf w}$ numerically via one-sided differencing, following
+Instead, we compute $\partial {\bf w}/\partial {\bf \Uc}^\prime$ analytically and $\partial {\bf R}/\partial {\bf w}$ numerically via one-sided differencing, following
 the algorithm in \cite{lsode} to minimize numerical noise.
-The resulting Jacobian, as computed via equation~\ref{eq:factored_jac}, is larger than
-expected by the integrator because of the inclusion of the $\rho$
-terms in the intermediate steps, so we remove the row corresponding to
-$R(\rho)$ and the column corresponding to $\partial/\partial \rho$
-before passing the Jacobian to VODE.
+
+The final Jacobian is found by multiplying these two and removing the row corresponding to
+$R(\rho)$ and the column corresponding to $\partial/\partial \rho$:
+\begin{equation}
+\frac{\partial \Rb}{\partial {\bf \Uc}} = \left (
+  \begin{array}{ccccc}
+    \frac{\partial \omegadot_\alpha}{\partial X_\alpha} -\frac{\partial \omegadot_\alpha}{\partial e} e_{X_\alpha} &
+    \frac{\partial \omegadot_\alpha}{\partial X_\beta} -\frac{\partial \omegadot_\alpha}{\partial e}  e_{X_\beta} &
+    0 &
+    \frac{\partial \omegadot_\alpha}{\partial e} \\
+     %
+    \frac{\partial \omegadot_\beta}{\partial X_\alpha} -\frac{\partial \omegadot_\beta}{\partial e} e_{X_\alpha} &
+    \frac{\partial \omegadot_\beta}{\partial X_\beta} -\frac{\partial \omegadot_\beta}{\partial e}  e_{X_\beta} &
+    0 &
+    \frac{\partial \omegadot_\beta}{\partial e} \\
+     %
+     \frac{\partial \Sdot}{\partial X_\alpha} - \frac{\partial \Sdot}{\partial e} e_{X_\alpha} &
+     \frac{\partial \Sdot}{\partial X_\beta} - \frac{\partial \Sdot}{\partial e} e_{X_\beta} &
+     0 &
+     \frac{\partial \Sdot}{\partial e} \\
+     %
+     \frac{\partial \Sdot}{\partial X_\alpha} - \frac{\partial \Sdot}{\partial e} e_{X_\alpha} &
+     \frac{\partial \Sdot}{\partial X_\beta} - \frac{\partial \Sdot}{\partial e} e_{X_\beta} &
+     0 &
+     \frac{\partial \Sdot}{\partial e} \\
+  \end{array}
+  \right )
+\end{equation}
 

### 2021-07-20 simplified_sdc_intro/paper.tex
some updates

@@
            \rho u \\
-           \rho E \end{array}\right )
+           \rho E \\
+           \rho e \end{array}\right )
 \end{equation}
@@
 $u$, specific total energy, $E$, and nuclear species mass fractions,
-$X_k$, and auxiliary composition variables, $\alpha_l$.  The mass fractions
+$X_k$, and auxiliary composition variables, $\alpha_l$.  We also carry
+the specific internal energy, $e$, as part of a dual energy
+formulation (see \citealt{bryan:1995,wdmergerI}). The mass fractions
 are constrained to sum to 1, $\sum_k X_k = 1$, but no such constraint
@@
    \rho u u + p \\
-   \rho u E + u p
+   \rho u E + u p \\
+   \rho u e
    \end{array} \right )
@@
 \end{equation}
-\MarginPar{Detail out what is $\Hb(\Uc)$?}
+The hydrodynamic sources include the ``pdV'' work for internal energy
+as well as gravity
+\begin{equation}
+  \Hb(\Ub) = \left ( \begin{array}{c}
+    0 \\
+    0 \\
+    0 \\
+    \rho g \\
+    \rho u g \\
+    -p \partial u/\partial x \end{array} \right )
+\end{equation}
  The reactive sources take the form:
@@
      0 \\
+     \rho \dot{S} \\
      \rho \dot{S} 
@@
 \label{eq:prim_sources}
-\Rb(q) = \left ( \begin{array}{c}
+\Rb(\qb) = \left ( \begin{array}{c}
      0 \\
@@
     Now the conservative hydrodynamics source terms are computed by first updating to the
-    new state with advection and the old-time source term applied for the full $\Delta t$ as:
-\MarginPar{I'm confused what happens with the $\Uc^{\star\star}$ state; I don't see it below.  Is The $\Shydro^{n+1/2}$ used below suppose to be from the $\Uc^{\star\star}$ state?}
+    new state with advection and the old-time source term applied for the full $\Delta t$ as: \MarginPar{little more complicated than this---we don't include the pdV source here}
     \begin{equation}

### 2021-07-20 simplified_sdc_intro/paper.tex
update Jacobian

@@
 
-To solve the reaction system implicitly, the ODE solver needs the Jacobian,
-$\partial \Rb/\partial \Uc$.  We follow the method of \cite{castro_sdc}
-and factor this into two pieces,
+To solve the reaction system implicitly, the ODE solver needs the
+Jacobian, $\partial \Rb/\partial \Uc^\prime$, where $\Uc^\prime =
+(\rho X_k, \rho e)^\intercal$ is the subset of the conserved variables we are
+integrating.  We follow the method of \cite{castro_sdc} and factor
+this into two pieces,
 \begin{equation}
 \label{eq:factored_jac}
-{\bf J} = \frac{\partial \Rb}{\partial {\bf w }} \frac{\partial {\bf w}}{\partial \Uc}.
-\end{equation}
-
-Even though it has no reactive sources, we include $\rho$ in our conservative state
-for the purposes of computing the Jacobian (we denote the conserved state used for the
-Jacobian $\Uc^\prime$).
+{\bf J} = \frac{\partial \Rb}{\partial {\bf w }} \frac{\partial {\bf w}}{\partial \Uc^\prime}.
+\end{equation}
 Writing this out for two species, $X_\alpha$ and $X_\beta$, we have
 \begin{equation}
-\Uc^\prime = \left ( \begin{array}{c} \rho \\ \rho X_\alpha \\ \rho X_\beta \\ \rho E \\ \rho e \end{array} \right )
-\end{equation}
-We take the intermediate state to be ${\bf w} = (\rho, X_\alpha, X_\beta,
-K, e)^\intercal$, where $K$ is the kinetic energy:
-\begin{equation}
-K = \frac{1}{2} u^2
-\end{equation}
-Note: this intermediate state differs from what we used in \citet{castro_sdc} in that
-we are using $e$ instead of $T$ as an independent variable now.  This follows from
-the work shown in \citet{strang_rnaas} that explored integrating the Strang
-reaction system with an energy equation.  If the network gives derivatives in terms
+\Uc^\prime = \left ( \begin{array}{c} \rho X_\alpha \\ \rho X_\beta \\ \rho e \end{array} \right )
+\end{equation}
+For interfacing with the reaction network, we use a simpler set of variables,
+\begin{equation}
+{\bf w} = \left ( \begin{array}{c} X_\alpha \\  X_\beta \\ e \end{array} \right )
+\end{equation}
+Note: this intermediate state differs from what we used in \citet{castro_sdc} but agrees with that from \cite{strang_rnaas} in that
+we use $e$ instead of $T$ as the independent variable.  If the network gives derivatives in terms
 of temperature, we can convert as:
@@
 \begin{equation}
-\frac{\partial \Uc}{\partial {\bf w}} = \left (
-   \begin{array}{ccccc}
-       1 & 0 & 0 & 0 & 0 \\
-       X_\alpha & \rho & 0 & 0 & 0 \\
-       X_\beta & 0 & \rho & 0 & 0  \\
-       e + K &
-                 \rho  e_{X_\alpha} & \rho e_{X_\beta} & \rho &
-                 \rho \\
-       e  &
-                 \rho  e_{X_\alpha} & \rho e_{X_\beta} & 0 &
-                 \rho
+\frac{\partial \Uc^\prime}{\partial {\bf w}} = \left (
+   \begin{array}{ccc}
+       \rho & 0 & 0 \\
+       0 & \rho & 0  \\
+       \rho  e_{X_\alpha} & \rho e_{X_\beta} & \rho \\
      \end{array}\right)
@@
 \frac{\partial {\bf w}}{\partial \Uc^\prime} = \left (
-  \begin{array}{ccccc}
-   1  & 0 & 0 & 0 & 0 \\
-   - \frac{X_\alpha}{\rho} & \frac{1}{\rho} & 0 & 0 & 0 \\
-   - \frac{X_\beta}{\rho} & 0 & \frac{1}{\rho} & 0 & 0 \\
-   - \frac{K}{\rho} & 0 & 0 & \frac{1}{\rho} & -\frac{1}{\rho} \\
-   \frac{\sum_k X_{k} e_{X_k}  - e}{ \rho} &
-    -\frac{e_{X_\alpha}}{\rho} & -\frac{e_{X_\beta}}{\rho} & 0 & \frac{1}{\rho} \\
+  \begin{array}{ccc}
+   \frac{1}{\rho} & 0 & 0 \\
+   0 & \frac{1}{\rho} & 0 \\
+    -\frac{e_{X_\alpha}}{\rho} & -\frac{e_{X_\beta}}{\rho} & \frac{1}{\rho} \\
    \end{array}\right)
@@
 \begin{equation}
-\Rb(\Uc^\prime) = \left (  \begin{array}{c} 0 \\ \rho \omegadot_\alpha \\ \rho \omegadot_\beta \\ \rho \Sdot \\ \rho \Sdot \end{array} \right )
-\end{equation}
-and the Jacobian is computed as $\partial \Rb/\partial {\bf w}$:
-\begin{equation}
+\Rb(\Uc^\prime) = \left (  \begin{array}{c} \rho \omegadot_\alpha \\ \rho \omegadot_\beta \\ \rho \Sdot \end{array} \right )
+\end{equation}
+We take all the quantities to be functions of $\rho$, $e$, and $X_k$,
+but since $\rho$ doesn't change from the reactions (the reactive
+source of the continuity equation is zero), it is held constant in
+these derivatives.  The Jacobian is computed as $\partial \Rb/\partial
+{\bf w}$:
+\begin{equation}
+\renewcommand{\arraystretch}{1.5}
 \frac{\partial \Rb}{\partial {\bf w}} = \left (
-  \begin{array}{ccccc}
-     0 & 0 & 0 & 0 & 0 \\
-     \omegadot_\alpha + \rho \frac{\partial \omegadot_\alpha}{\partial \rho} &
+  \begin{array}{ccc}
      \rho \frac{\partial \omegadot_\alpha}{\partial X_\alpha} &
-     \rho \frac{\partial \omegadot_\alpha}{\partial X_\beta} & 0 & \rho \frac{\partial \omegadot_\alpha}{\partial e} \\
+     \rho \frac{\partial \omegadot_\alpha}{\partial X_\beta} & 
+     \rho \frac{\partial \omegadot_\alpha}{\partial e} \\
       %
-     \omegadot_\beta + \rho \frac{\partial \omegadot_\beta}{\partial \rho} &
      \rho \frac{\partial \omegadot_\beta}{\partial X_\alpha} &
-     \rho \frac{\partial \omegadot_\beta}{\partial X_\beta} & 0 &  \rho \frac{\partial \omegadot_\beta}{\partial e} \\
+     \rho \frac{\partial \omegadot_\beta}{\partial X_\beta} &
+     \rho \frac{\partial \omegadot_\beta}{\partial e} \\
      %
-     \Sdot + \rho \frac{\partial \Sdot}{\partial \rho} & \rho \frac{\partial \Sdot}{\partial X_\alpha} & \rho \frac{\partial \Sdot}{\partial X_\beta} & 0 & \rho \frac{\partial \Sdot}{\partial e} \\
-     %
-     \Sdot + \rho \frac{\partial \Sdot}{\partial \rho} & \rho \frac{\partial \Sdot}{\partial X_\alpha} & \rho \frac{\partial \Sdot}{\partial X_\beta} & 0 & \rho \frac{\partial \Sdot}{\partial e} \\
+     \rho \frac{\partial \Sdot}{\partial X_\alpha} &
+     \rho \frac{\partial \Sdot}{\partial X_\beta} &
+     \rho \frac{\partial \Sdot}{\partial e} \\
   \end{array}
   \right )
+\renewcommand{\arraystretch}{1}
 \end{equation}
@@
 the algorithm in \cite{lsode} to minimize numerical noise.
-
-The final Jacobian is found by multiplying these two and removing the row corresponding to
-$R(\rho)$ and the column corresponding to $\partial/\partial \rho$:
-\begin{equation}
-\frac{\partial \Rb}{\partial {\bf \Uc}} = \left (
-  \begin{array}{ccccc}
+The final Jacobian is found by multiplying these two:
+\begin{equation}
+\renewcommand{\arraystretch}{1.5}
+\frac{\partial \Rb}{\partial \Uc^\prime} = \left (
+  \begin{array}{ccc}
     \frac{\partial \omegadot_\alpha}{\partial X_\alpha} -\frac{\partial \omegadot_\alpha}{\partial e} e_{X_\alpha} &
     \frac{\partial \omegadot_\alpha}{\partial X_\beta} -\frac{\partial \omegadot_\alpha}{\partial e}  e_{X_\beta} &
-    0 &
     \frac{\partial \omegadot_\alpha}{\partial e} \\
@@
     \frac{\partial \omegadot_\beta}{\partial X_\beta} -\frac{\partial \omegadot_\beta}{\partial e}  e_{X_\beta} &
-    0 &
     \frac{\partial \omegadot_\beta}{\partial e} \\
@@
      \frac{\partial \Sdot}{\partial X_\beta} - \frac{\partial \Sdot}{\partial e} e_{X_\beta} &
-     0 &
-     \frac{\partial \Sdot}{\partial e} \\
-     %
-     \frac{\partial \Sdot}{\partial X_\alpha} - \frac{\partial \Sdot}{\partial e} e_{X_\alpha} &
-     \frac{\partial \Sdot}{\partial X_\beta} - \frac{\partial \Sdot}{\partial e} e_{X_\beta} &
-     0 &
      \frac{\partial \Sdot}{\partial e} \\
@@
   \right )
+\renewcommand{\arraystretch}{1}
 \end{equation}

### 2021-07-20 simplified_sdc_intro/paper.tex
update to reflect the new formulation

@@
 
-\newcommand{\nse}[1]{{\mathrm{NSE}( #1 )}}
+\newcommand{\nse}[1]{{\mathtt{NSE}( #1 )}}
 
@@
 \begin{equation}
+  \label{eq:react_source}
   \Rb(\Uc) = \left ( \begin{array}{c}
@@
     \end{equation}
-    This integration begins with $\Uc^n$ and results in $\Uc^{n+1,(k)}$.
-
-    As we are integrating this system we need to get the
-    temperature, $T$, for the rate evaluations.  We construct this
-    by subtracting the kinetic energy from the total energy to get
-    the specific internal energy, $e$, and then calling the equation
-    of state.
+
+    Looking at the form of Eq.~\ref{eq:react_source}, we see that only
+    the species, auxiliary quantities, and energies have source terms.
+    Since we are reacting, we will use the $X_k$ as the primary
+    composition variable and compute $\alpha_l$ as needed from them,
+    so we don't directly integrate the auxiliary quantities.
+    Likewise, the total and internal energies both provide the same
+    information, and integrating both overconstrains the system, so we
+    just integrate the internal energy.  We define the subset of variables
+    that are directly integrated as:
+    \begin{equation}
+      \Uc^\prime = \left ( \begin{array}{c} \rho X_k \\ \rho e \end{array} \right )
+    \end{equation}
+    and we integrate
+    \begin{equation}
+      \odt{\Uc^\prime} = \Adv{\Uc^\prime}^{n+1/2,(k)} + \Rb(\Uc^\prime)
+    \end{equation}
+    This integration begins with $\Uc^{\prime,n}$ and results in $\Uc^{\prime,n+1,(k)}$.
+
+    We will need the density at times during the integration, which we construct as:
+    \begin{equation}
+      \rho(t) = \rho^n + \Advss{\rho}^{n+1/2,(k)} \, t
+    \end{equation}
+    As we are integrating this system we need to get the temperature,
+    $T$, for the rate evaluations.  We obtain this by directly from
+    internal energy, composition, and density using the equation of
+    state.
 
     Our integrator also needs the Jacobian of the system, in terms of
-    the conserved variables.  This is different than the form of the
+    the $\Uc^\prime$.  This is different than the form of the
     Jacobian usually used in reaction networks.  We describe the form
@@
     $\rho$ does as well.  $(\rho u)$ will also have linear
-    contributions in time from advection.  Since the reactions don't
-    affect $\rho$ and $(\rho u)$, we can algebraically update these
-    using this piecewise linear behavior:
-    \begin{align}
-      \rho(t) &= \rho^n + \Advss{\rho}^{n+1/2,(k)} \, t \\
-      (\rho u)(t) &= (\rho u)^n + \Advss{\rho u}^{n+1/2,(k)} \, t
-    \end{align}
+    contributions in time from advection.
+
+    At the end of the integration, we can do the conservative update of momentum
+    and energy.  Momentum is straightforward, since there are no reactive sources:
+    \begin{equation}
+      (\rho u)^{n+1} = (\rho u)^n + \Delta t \Advss{\rho u}^{n+1/2,(k)}
+    \end{equation}
+    For total energy, we first need to isolate the reactive source for energy:
+    \begin{equation}
+      (\rho \Sdot)^{n+1/2} = \frac{(\rho e)^{n+1} - (\rho e)^n}{\Delta t} - \Advss{\rho e}^{n+1/2,(k)}
+    \end{equation}
+    Then the update is
+    \begin{equation}
+      (\rho E)^{n+1} = (\rho E)^n + \Delta t \Advss{\rho E}^{n+1/2,(k)} + \Delta t (\rho \Sdot)^{n+1/2}
+    \end{equation}
 
@@
     \begin{itemize}
-    \item Integrate the full reaction network (Eqs.~\ref{eq:strang:T} and \ref{eq:strang:X}) as usual for Strang splitting
+    \item Integrate the full reaction network (Eqs.~\ref{eq:strang:e} and \ref{eq:strang:X}) as usual for Strang splitting
     \item Update the aux quantities at the end of the burn using Eqs.~\ref{eq:aux:ye}, \ref{eq:aux:abar}, and \ref{eq:aux:bea} with the new mass fractions, $X_k$.
@@
 \begin{deluxetable}{lllllll}
-\tablecaption{\label{table:react_convert_strang} Convergence ($L_1$ norm) for the reacting convergence problem using SDC integration (2 iterations).}
+\tablecaption{\label{table:react_convert_sdc} Convergence ($L_1$ norm) for the reacting convergence problem using SDC integration (2 iterations).}
 \tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
@@
 \begin{deluxetable}{lllllll}
-\tablecaption{\label{table:nse_strang_methodII} Convergence ($L_1$ norm) for the NSE convergence
+\tablecaption{\label{table:nse_sdc_methodII} Convergence ($L_1$ norm) for the NSE convergence
 problem using the simplified-SDC algorithm.}

### 2021-08-28 simplified_sdc_intro/paper.tex
add section
some more updates

@@
 evolved into NSE, and if so, we redo the zone update using the NSE
-prescription.  {\color{red} We optionally apply a relaxation factor, typically, $0.9$, to the NSE conditions to allow a state that is close to NSE after an integrator failure enter NSE.}
+prescription.  To ensure that this is not triggered at the very start
+of integration, we require a minimum number of steps (10) to have
+elapsed before checking for NSE.  After we leave the integrator, the
+state is returned to the burner driver where we finish the integration
+via NSE if the conditions satisfy the NSE criteria.  We optionally
+apply a relaxation factor, typically, $0.9$, to the NSE conditions to
+allow a state that is close to NSE after an integrator failure enter
+NSE.
 
@@
 
+\subsection{Reacting Buoyant Bubble}
+
+We next consider a buoyant, reacting bubble in a stratified,
+plane-parallel atmosphere.  This test exercises the treatment of
+hydrodynamical sources in the SDC integration.  We use the set of
+initial conditions that was first shown in \cite{ABNZ:III}.
+
 \subsection{Massive Star}
@@
 \end{equation}
+where the state ${\bf w}$ is chosen to match the set of variables used to evaluate the reaction rates.
 Writing this out for two species, $X_\alpha$ and $X_\beta$, we have
@@
 \end{equation}
-For interfacing with the reaction network, we use a simpler set of variables,
-\begin{equation}
-{\bf w} = \left ( \begin{array}{c} X_\alpha \\  X_\beta \\ e \end{array} \right )
-\end{equation}
-Note: this intermediate state differs from what we used in \citet{castro_sdc} but agrees with that from \cite{strang_rnaas} in that
-we use $e$ instead of $T$ as the independent variable.  If the network gives derivatives in terms
-of temperature, we can convert as:
-\begin{equation}
-\left . \frac{\partial \phi}{\partial T} \right |_\rho =
-  \left . \frac{\partial \phi}{\partial e} \right |_\rho 
-  \left . \frac{\partial e}{\partial T} \right |_\rho =
-  c_v \left . \frac{\partial \phi}{\partial e} \right |_\rho
-\end{equation}
-for some quantity $\phi(\rho, T)$.  Here we identified the specific heat
-at constant volume, $c_v = \partial e/\partial T|_\rho$.
+For interfacing with the reaction network, we use
+\begin{equation}
+{\bf w} = \left ( \begin{array}{c} X_\alpha \\  X_\beta \\ T \end{array} \right )
+\end{equation}
+Note: even though we are using $T$ here instead of $e$, we still do the overall ODE integration
+in terms of $(\rho e)$, consistent with the Strang method described in \cite{strang_rnaas}.
+%%  in that
+%% we use $e$ instead of $T$ as the independent variable.  If the network gives derivatives in terms
+%% of temperature, we can convert as:
+%% \begin{equation}
+%% \left . \frac{\partial \phi}{\partial T} \right |_\rho =
+%%   \left . \frac{\partial \phi}{\partial e} \right |_\rho 
+%%   \left . \frac{\partial e}{\partial T} \right |_\rho =
+%%   c_v \left . \frac{\partial \phi}{\partial e} \right |_\rho
+%% \end{equation}
+%% for some quantity $\phi(\rho, T)$.  Here we identified the specific heat
+%% at constant volume, $c_v = \partial e/\partial T|_\rho$.
 
@@
        0 & \rho & 0  \\
-       \rho  e_{X_\alpha} & \rho e_{X_\beta} & \rho \\
+       \rho  e_{X_\alpha} & \rho e_{X_\beta} & \rho c_v \\
      \end{array}\right)
@@
 \end{equation}
+and the specific heat at constant volume is
+\begin{equation}
+c_v = \left . \frac{\partial e}{\partial T} \right |_{\rho, X_k}
+\end{equation}
 We get the inverse
-(computed via SymPy, see the included Jupyter notebook) as:
+(computed via SymPy) as:
 \begin{equation}
@@
    0 & \frac{1}{\rho} & 0 \\
-    -\frac{e_{X_\alpha}}{\rho} & -\frac{e_{X_\beta}}{\rho} & \frac{1}{\rho} \\
+    -\frac{e_{X_\alpha}}{\rho c_v} & -\frac{e_{X_\beta}}{\rho c_v} & \frac{1}{\rho c_v} \\
    \end{array}\right)
@@
      \rho \frac{\partial \omegadot_\alpha}{\partial X_\beta} & 
-     \rho \frac{\partial \omegadot_\alpha}{\partial e} \\
+     \rho \frac{\partial \omegadot_\alpha}{\partial T} \\
       %
@@
      \rho \frac{\partial \omegadot_\beta}{\partial X_\beta} &
-     \rho \frac{\partial \omegadot_\beta}{\partial e} \\
+     \rho \frac{\partial \omegadot_\beta}{\partial T} \\
      %
@@
      \rho \frac{\partial \Sdot}{\partial X_\beta} &
-     \rho \frac{\partial \Sdot}{\partial e} \\
+     \rho \frac{\partial \Sdot}{\partial T} \\
   \end{array}
@@
 
-While VODE can compute the entire Jacobian, ${\bf J}$ numerically via differencing, we found
-that does not give reliable results.
-Instead, we compute $\partial {\bf w}/\partial {\bf \Uc}^\prime$ analytically and $\partial {\bf R}/\partial {\bf w}$ numerically via one-sided differencing, following
-the algorithm in \cite{lsode} to minimize numerical noise.
 The final Jacobian is found by multiplying these two:
@@
   \begin{array}{ccc}
-    \frac{\partial \omegadot_\alpha}{\partial X_\alpha} -\frac{\partial \omegadot_\alpha}{\partial e} e_{X_\alpha} &
-    \frac{\partial \omegadot_\alpha}{\partial X_\beta} -\frac{\partial \omegadot_\alpha}{\partial e}  e_{X_\beta} &
-    \frac{\partial \omegadot_\alpha}{\partial e} \\
+    \frac{\partial \omegadot_\alpha}{\partial X_\alpha} - \frac{e_{X_\alpha}}{c_v} \frac{\partial \omegadot_\alpha}{\partial T} &
+    \frac{\partial \omegadot_\alpha}{\partial X_\beta} - \frac{e_{X_\beta}}{c_v} \frac{\partial \omegadot_\alpha}{\partial T} &
+    \frac{1}{c_v} \frac{\partial \omegadot_\alpha}{\partial T} \\
      %
-    \frac{\partial \omegadot_\beta}{\partial X_\alpha} -\frac{\partial \omegadot_\beta}{\partial e} e_{X_\alpha} &
-    \frac{\partial \omegadot_\beta}{\partial X_\beta} -\frac{\partial \omegadot_\beta}{\partial e}  e_{X_\beta} &
-    \frac{\partial \omegadot_\beta}{\partial e} \\
+    \frac{\partial \omegadot_\beta}{\partial X_\alpha} - \frac{e_{X_\alpha}}{c_v} \frac{\partial \omegadot_\beta}{\partial T} &
+    \frac{\partial \omegadot_\beta}{\partial X_\beta} - \frac{e_{X_\beta}}{c_v} \frac{\partial \omegadot_\beta}{\partial T} &
+    \frac{1}{c_v} \frac{\partial \omegadot_\beta}{\partial T} \\
      %
-     \frac{\partial \Sdot}{\partial X_\alpha} - \frac{\partial \Sdot}{\partial e} e_{X_\alpha} &
-     \frac{\partial \Sdot}{\partial X_\beta} - \frac{\partial \Sdot}{\partial e} e_{X_\beta} &
-     \frac{\partial \Sdot}{\partial e} \\
+     \frac{\partial \Sdot}{\partial X_\alpha} -  \frac{e_{X_\alpha}}{c_v} \frac{\partial \Sdot}{\partial T} &
+     \frac{\partial \Sdot}{\partial X_\beta} -  \frac{e_{X_\beta}}{c_v} \frac{\partial \Sdot}{\partial T} &
+     \frac{1}{c_v} \frac{\partial \Sdot}{\partial T} \\
   \end{array}
@@
 \end{equation}
+We note that the form of these entries is the same as one would arrive at if you start with the rates expressed as
+$\omega_k(\rho, T(\rho, X_j, e), X_j)$ and note that constant $e$ implies that
+\begin{equation}
+\left . \frac{\partial T}{\partial X_k} \right |_{\rho, e} = - \frac{e_{X_k}}{c_v}
+\end{equation}
+
+While VODE can compute the entire Jacobian, ${\bf J}$ numerically via
+differencing, we found that does not give reliable results.  Instead,
+we compute compute the derivatives with respect to $X_k$ and $T$
+one-sided differencing, following the algorithm in \cite{lsode} to
+minimize numerical noise.  We then use the equation of state to
+compute $e_{X_k}$ and $c_v$ and construct the entries of the final
+Jacobian.
 

### 2021-09-26 simplified_sdc_intro/paper.tex
update
some restructuring of the first 2 sections

@@
 \newcommand{\ddx}[1]{{\frac{{\partial#1}}{\partial x}}}
+\newcommand{\ddxs}[1]{{\frac{{\partial}}{\partial x}}#1}
 \newcommand{\ddt}[1]{{\frac{{\partial#1}}{\partial t}}}
@@
 
-%% \author[0000-0001-8401-030X]{M.~Zingale}
-%% \affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
-%%        Stony Brook, NY 11794-3800}
-
-%% \author[0000-0003-0439-4556]{M.~P.~Katz}
-%% \affiliation{Nvidia Corp}
-
-%% \author[0000-0002-5749-334X]{J.~B.~Bell}
-%% \affiliation{Center for Computational Sciences and Engineering, Lawrence Berkeley National Laboratory, Berkeley, CA  94720}
-
-%% \author[0000-0003-1791-0265]{A.~J.~Nonaka}
-%% \affiliation{Center for Computational Sciences and Engineering, Lawrence Berkeley National Laboratory, Berkeley, CA  94720}
-
-%% \author[0000-0001-8092-1974]{W.~Zhang}
-%% \affiliation{Center for Computational Sciences and Engineering, Lawrence Berkeley National Laboratory, Berkeley, CA  94720}
-
-%% \correspondingauthor{Michael Zingale}
-%% \email{michael.zingale@stonybrook.edu}
+\author[0000-0001-8401-030X]{M.~Zingale}
+\affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
+  Stony Brook, NY 11794-3800}
+
+\author[0000-0003-0439-4556]{M.~P.~Katz}
+\affiliation{Nvidia Corp}
+
+\author[0000-0003-1791-0265]{A.~J.~Nonaka}
+\affiliation{Center for Computational Sciences and Engineering, Lawrence Berkeley National Laboratory, Berkeley, CA  94720}
+
+\author[0000-0002-1530-781X]{Alice Harpole}
+\affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
+             Stony Brook, NY 11794-3800}
+
+\correspondingauthor{Michael Zingale}
+\email{michael.zingale@stonybrook.edu}
 
@@
 the difficulty in accurately coupling hydrodynamics and reactions.
-This can be particularly acute at high temperatures where nuclear
-statistical equilibrium is established.  We develop a simplified
-approach based on spectral deferred corrections (SDC) coupling of explicit
-hydrodynamics and stiff reaction sources as an alternative to operator
-splitting or the more comprehensive SDC approach we demonstrated
-previously.  We apply the new method to some example problems and show
-how to modify it to work with a hybrid network consisting of a
-reaction ODE system and a table for nuclear statistical equilibrium.
-This is all done in the framework of the \castro\ hydrodynamics code,
-and all algorithm implementations are freely available.
+This can be particularly acute during explosive burning or at high
+temperatures where nuclear statistical equilibrium is established.  We
+develop a simplified approach based on the ideas of spectral deferred
+corrections (SDC) coupling of explicit hydrodynamics and stiff
+reaction sources as an alternative to operator splitting or the more
+comprehensive SDC approach we demonstrated previously.  We apply the
+new method to some example problems and show how to modify it to work
+with a hybrid network consisting of a reaction ODE system and a table
+for nuclear statistical equilibrium.  This is all done in the
+framework of the \castro\ hydrodynamics code, and all algorithm
+implementations are freely available.
 \end{abstract}
@@
 
-Reactive flow can be challenging to model when changes in the
-nuclear abundances due to reactions are integrated over hydrodynamic
-time scales.  Traditional methods of coupling
+Modeling astrophysical reacting flows can be challenging because of
+the disparity better the nuclear and hydrodynamics timescales.
+Reaction networks tend to be stiff, requiring implicit integration
+techniques to stably integrate the system.  In contrast, compressible
+hydrodynamics flows are limited by the (often much longer)
+sound-crossing time over a computational cell and can be solved using
+explicit time integration. Traditional methods of coupling
 hydrodynamics and reactions used in astrophysics use operator
 splitting---each physical process acts on the output of the previous
-process in alternating fashion.  This makes it easy to add physics in a modular way to a
-simulation code, but competition between physical processes can cause
-the coupling to breakdown.  These splitting errors can lead to loss of
-accuracy and further time step limitations.
+process in alternating fashion.  This makes it easy to use different
+time-integration methods for the different physics, and to build a
+simulation code in a modular way.  However, competition between the
+different physical processes can cause the coupling to breakdown.
+These splitting errors can lead to loss of accuracy and further time
+step limitations.
 
@@
 
-In \cite{castro_sdc}, we developed second- and fourth-order accurate
-methods in space and time for coupling hydrodynamics and nuclear
-reaction networks based on spectral deferred corrections (SDC)
-methods, and demonstrated the method in a variety of test problems.
-The approach presented here is considerably simpler from a temporal
-integration standpoint, but allows is to incorporate modifications
-for regions in nuclear statistical equilibirum.
-Furthermore, we reuse the main
-CTU hydrodynamics construction and a largely similar ODE integration
+The \castro\ hydrodynamics code \citep{castro,castro_joss} is used for
+all of our numerical experiments.  \castro\ is a compressible
+(magneto-, radiation-) hydrodynamics code built on the AMReX adaptive
+mesh refinement (AMR) framework \citep{amrex_joss}.  \castro\ has been
+designed to be performance portable and runs on massively parallel CPU
+and GPU architectures \citep{castro_gpu}.  For hydrodynamics, the
+corner transport upwind (CTU) \citep{ppmunsplit} method with the
+piecewise parabolic method (PPM) \citep{ppm,millercolella:2002} is
+used.  \castro\ includes self-gravity, rotation, arbitrary equations
+of state and reaction networks, and has been used for modeling X-ray
+bursts and different models of thermonuclear, core-collapse, and
+pair-instability supernovae.  Recently in \citet{castro_sdc}, we
+developed second- and fourth-order accurate methods in space and time
+for coupling hydrodynamics and nuclear reaction networks based on
+spectral deferred corrections (SDC) methods, and demonstrated the
+method in a variety of test problems.
+
+The time-integration approach presented here is considerably simpler than
+the SDC method of \citet{castro_sdc}, but allows us to reuse the main CTU
+hydrodynamics construction and a largely similar ODE integration
 scheme, making this method easier to add to existing simulation codes.
-It also extends to adaptive mesh refinement with subcycling in a
-straightforward manner, avoiding the complications described in
-\citep{mccorquodalecolella} needed to fill ghost cells when using
-method-of-lines integration.  However, it is restricted to
+Furthermore, it also extends to adaptive mesh refinement with
+subcycling in a straightforward manner, avoiding the complications
+described in \citep{mccorquodalecolella} needed to fill ghost cells
+when using method-of-lines integration.  However, it is restricted to
 second-order accuracy overall.  We term this algorithm the
-``simplified SDC method''.  We use the \castro\ hydrodynamics code
-\citep{castro} for all of our numerical experiments, and all of the
+``simplified SDC method''.  In this paper we describe the overall method, show its extension to NSE,
+and demonstrate it on several test problems using \castro.  All of the
 code to reproduce the results in this paper are in the \castro\ github
@@
 The system appears as:
-\begin{equation}
-\frac{\partial \Uc_t}{\partial t}  + \frac{\partial \Fb^\xv (\Uc)}{\partial x} = \Sc(\Uc)
-\end{equation}
-where
+\begin{align}
+\label{eq:conserved_system}
+\ddt{\rho} + \ddxs{(\rho u)} &= 0 \\
+\ddt{(\rho X_k)} + \ddxs{(\rho X_k u)} &= \rho \omegadot(X_k) \\
+\ddt{(\rho \alpha_k)} + \ddxs{(\rho \alpha_k u)} &= \rho \omegadot(\alpha_k) \\
+\ddt{(\rho u)} + \ddxs{(\rho u^2 + p)} &= \rho g \\
+\ddt{(\rho E)} + \ddxs{(\rho E u + p u)} &= \rho ug + \rho \Sdot 
+\end{align}
+where $\rho$ is the mass density, $u$ is velocity, $E$ is specific
+total energy, $p$ is the pressure, and we carry nuclear species mass
+fractions, $X_k$, and auxiliary composition variables, $\alpha_l$.  The specific total
+energy relates to the specific internal energy, $e$, as $E = e + u^2/2$,
+and we also separately evolve $e$, as part of a dual
+energy formulation (see \citealt{bryan:1995,wdmergerI}):
+\begin{equation}
+\ddt{(\rho e)} + \ddxs{(\rho e u)} + p \ddx{u} = \rho \Sdot
+\end{equation}
+ The mass fractions are constrained to sum to 1, $\sum_k X_k = 1$, but
+ no such constraint exists on the auxiliary variables.  Gravity,
+ represented by the gravitational acceleration $g$ appears as a source
+ term.  Finally, $\omegadot(X_k)$ is the creation rate for species
+ $k$, $\omegadot(\alpha_l)$ is the creation rate for auxiliary
+ composition variable $l$, and $\dot{S}$ is the energy generation rate
+ per unit mass.  Aside from the $p\partial u/\partial x$ (``pdV'')
+ term in the internal energy equation, this system is in conservative
+ form with source terms.
+
+To close the system, we need the equation of state.  For a system
+where the composition is completely specified by the mass fractions,
+$X_k$, the equation of state would take the form:
+\begin{equation}
+p = p(\rho, X_k, e)
+\end{equation}
+However, as we'll see shortly, when we use the NSE table, we are using
+the results of a much larger network then can be represented by the
+mass fractions, and in this case, we rely on the auxiliary composition
+variables, $\alpha_l$, to describe the state of the composition, and
+our EOS has the form:
+\begin{equation}
+p = p(\rho, \alpha_l, e)
+\end{equation}
+
+For convenience, we'll use $\Uc$ to represent the conserved state
+and split the sources up into pure hydrodynamic and reactive parts:
 \begin{equation}
@@
            \rho e \end{array}\right )
-\end{equation}
-are the conserved fluid quantities: mass density, $\rho$, velocity,
-$u$, specific total energy, $E$, and nuclear species mass fractions,
-$X_k$, and auxiliary composition variables, $\alpha_l$.  We also carry
-the specific internal energy, $e$, as part of a dual energy
-formulation (see \citealt{bryan:1995,wdmergerI}). The mass fractions
-are constrained to sum to 1, $\sum_k X_k = 1$, but no such constraint
-exists on the auxiliary variables.  The specific total energy relates
-to the specific internal energy as $E = e + u^2/2$.  The corresponding
-fluxes are
-\begin{equation}
-\Fb^\xv (\Uc) = \left ( \begin{array}{c}
-   \rho u \\
-   \rho X_k u \\
-   \rho \alpha_l u \\
-   \rho u u + p \\
-   \rho u E + u p \\
-   \rho u e
-   \end{array} \right )
-\end{equation}
-Here the pressure, $p$, enters, and is found via the equation of state.  For a system
-where the composition is completely specified by the mass fractions, $X_k$, the equation
-of state would take the form:
-\begin{equation}
-p = p(\rho, X_k, e)
-\end{equation}
-However, as we'll see shortly, when we use the NSE table, we are using
-the results of a much larger network then can be represented by the
-mass fractions, and in this case, we rely on the auxiliary composition
-variables, $\alpha_l$, to describe the state of the composition, and
-our EOS has the form:
-\begin{equation}
-p = p(\rho, \alpha_l, e)
-\end{equation}
-
-We decompose the source terms, $\Sc(\Uc)$, into
-hydrodynamical sources, $\Hb$ (like gravity), and reactive sources,
-$\Rb$,
-\begin{equation}
-  \Sc(\Uc) = \Hb(\Uc) + \Rb(\Uc).
-\end{equation}
-The hydrodynamic sources include the ``pdV'' work for internal energy
-as well as gravity
-\begin{equation}
-  \Hb(\Ub) = \left ( \begin{array}{c}
+  \qquad
+  \Hb(\Uc) = \left ( \begin{array}{c}
     0 \\
@@
     \rho u g \\
-    -p \partial u/\partial x \end{array} \right )
-\end{equation}
- The reactive sources take the form:
-\begin{equation}
-  \label{eq:react_source}
+    0 \end{array} \right )
+  \qquad
   \Rb(\Uc) = \left ( \begin{array}{c}
@@
 \end{equation}
-where $\omegadot(X_k)$ is the creation rate for species $k$, $\omegadot(\alpha_l)$ is the creation rate for
-auxiliary composition variable $l$, and $\dot{S}$
-is the energy generation rate per unit mass.
+We will represent the total source as $\Sc(\Uc)$:
+\begin{equation}
+  \Sc(\Uc) = \Hb(\Uc) + \Rb(\Uc).
+\end{equation}
+We note that the internal energy ``pdV'' work is not treated as a source
+term but is instead constructed with the hydrodynamical fluxes are computed in the CTU method.
 
@@
 \begin{equation}
-\qb_t + \Ab^\xv(\qb) \qb_x  = \Sq
+\qb_t + \Ab^\xv(\qb) \qb_x  = \Sc(\qb)
 \end{equation}
@@
 \end{equation}
-where $\Gamma_1$ is an adiabatic index,
+where $h$ is the specific enthalpy and $\Gamma_1$ is an adiabatic index,
 $\Gamma_1 = d\log p/d\log\rho|_s$ at constant entropy.
+The CTU+PPM algorithm uses the characteristic wave structure of $\Ab$
+to collect the information that makes it to an interface over a timestep
+in order to compute the fluxes through the interface.
 Note, the primitive state has two thermodynamic quantities, $p$
@@
 alternate formulations are possible \citep{colellaglaz:1985}.
-The source term vector, $\Sq$, can again be decomposed into hydrodynamic
+The source term vector, $\Sc(\qb)$, can again be decomposed into hydrodynamic
 sources (now in terms of the primitive variables) and reaction terms,
-\MarginPar{Can we use $\Sc(\qb)$ instead of $\Sq$ and $\Hb(\qb)$ instead of $\Sqhydro$ to keep notation consistent with the conserved variables, since you are already doing that with $\Rb(\qb)$?}
-\begin{equation}
-  \Sq = \Sqhydro + \Rb(\qb)
+\begin{equation}
+  \Sc(\qb) = \Hb(\qb) + \Rb(\qb)
 \end{equation}
@@
 \label{eq:prim_sources}
+\Hb(\qb) = \left ( \begin{array}{c}
+     0 \\
+     0 \\
+     0 \\
+     g \\
+     0 \\
+     0 \\
+   \end{array} \right )
+\qquad
 \Rb(\qb) = \left ( \begin{array}{c}
@@
 
-We use the \castro\ hydrodynamics code~\cite{castro}, together with
-corner transport upwind piecewise parabolic method
-hydrodynamics~\citep{millercolella:2002}.  This is a finite-volume
-method that uses characteristic tracing to predict a time-centered
-flux through the interfaces of the grid zones.  For pure
-hydrodynamics, these time-centered fluxes result in second-order
-accurate time integration.  With reactions, we want to couple
-hydrodynamics and the reaction sources to second order.  Nuclear
+The CTU+PPM method for hydrodynamics is second-order accurate in space
+and time.  We want to couple the reactive sources to the hydrodynamics
+to be second-order in time as well.  As discussed above, nuclear
 reaction sources are stiff, and need to be integrated using implicit
-methods for stabilty.  Operator splitting (e.g., Strang) is traditionally employed
-here, and is used as a benchmark for comparison in this paper.
+methods for stabilty.  Operator splitting (e.g., Strang) is
+traditionally employed here, and is used as a benchmark for comparison
+in this paper.  We'll discuss this traditional approach next before moving
+on to our new time-coupling method.
 
@@
 
-The basic idea of spectral deferred corrections is to express the
+The basic idea of spectral deferred corrections \MarginPar{ref} is to express the
 update as an integral and divide the time-update into a number of
@@
 accuracy is determined by the quadrature method used to evaluate the
-integral, which is done using an iteratively-lagged solution.
-
-We start by writing our update as:
+integral, which is done using an iteratively-lagged solution.  This is the
+classic approach that we implemented in \cite{castro_sdc}.
+
+Our simplified-SDC approach is based on the method described in \citet{SDC-old} and
+a similar implementation in the \maestro\ simulation code.
+We again start by considering the update
+in integral form:
 \begin{equation}
@@
 \end{equation}
-where $\Advs{\Uc} =  -{\partial \Fb^\xv (\Uc)}/{\partial x} + \Hb(\Uc)$.
-For the simplified SDC method we explore here, we will make the
-advective term piecewise constant in time, using the value at the
-midpoint in time, to achieve second-order accuracy, giving us:
+where $\Advs{\Uc}$ includes all terms in system
+\ref{eq:conserved_system} except for the reactive source terms,
+$\Rb(\Ub)$ (i.e., it includes the flux divergences, ``pdV'' term for
+internal energy, and the pure hydrodynamical sources).  In this approach, we 
+approximate the advective term
+piecewise constant in time, using the value at the midpoint in time,
+to achieve second-order accuracy, giving us:
 \begin{equation}
@@
 \end{equation}
-This is the approach first shown in \citep{SDC-old}.  A
-version of this was also implemented in the \maestro\ low-Mach number
-hydrodynamics code some time ago.  The compressible version for use in
-\castro\ is slightly different due to the need to do some operations
+The \castro\ version of the simplified-SDC algorithm
+differs from the previous versions due to the need to do some operations
 on the conserved variable state and some on the primitive variable
-state.  We describe the application of this to the equations of
-compressible hydrodynamics below.
+state.
+
+The basic time update algorithm proceeds as:
 
@@
     changed the primitive variable state over the timestep, which we
-    will denote $\Ic_q$.  Since we do not have any information about
-    the current timestep in the first iteration,
-    we use the value from the last iteration of the previous timestep:
+    will denote $\Ic_q$ (this is basically an approximation of
+    $\Rb(q)$).
+
+    Since we do not have any information about the current
+    timestep in the first iteration, we use the value from the last
+    iteration of the previous timestep:
     \begin{equation}
@@
     \end{equation}
-    \MarginPar{changed to 1-based $k$ iteration for aesthetics?  I can switch it back if you prefer ``-1'' to be the starting point instead of ``0''.}
   %% \item Solve the Poisson problem for the initial gravitational potential:
@@
       \begin{equation}
-        \Sq = \Sqhydro^{n+1/2} + \Ic_\qb^{n+1/2,(k-1)}
+        \Sc(\qb) = \Hb(\qb) + \Ic_\qb^{n+1/2,(k-1)}
       \end{equation}
@@
 
+Do this in 1-d
+Show the amount of work needed for Strang vs. SDC
+
 Play around with the threshold where NSE kicks in

### 2021-09-27 simplified_sdc_intro/paper.tex
more updates

@@
       explicit reactive source terms are seen by the hydrodynamics update.
-      Any hydrodynamic source terms are time-centered
-      using the previous iteration:
-\MarginPar{Formally both $t^n$ or $t^{n+1/2}$ sources give you 2nd-order since they are CTU predictor source terms.  I've never found this type of iteration for this term to amount to any benefit}
-      \begin{equation}
-        \Sqhydro^{n+1/2} = \frac{1}{2} \left ( \Sqhydro^n + \Sqhydro^{n+1,(k-1)} \right )
-      \end{equation}
+%%       Any hydrodynamic source terms are time-centered
+%%       using the previous iteration:
+%% \MarginPar{Formally both $t^n$ or $t^{n+1/2}$ sources give you 2nd-order since they are CTU predictor source terms.  I've never found this type of iteration for this term to amount to any benefit}
+%%       \begin{equation}
+%%         \Sqhydro^{n+1/2} = \frac{1}{2} \left ( \Sqhydro^n + \Sqhydro^{n+1,(k-1)} \right )
+%%       \end{equation}
 
@@
       zero, and characteristic tracing or the various flux corrections
-      may not preserve this.
-
-      {\color{red} We enforce that the species interface states remain in $[0, 1]$ after adding the reactive source.}
+      may not preserve this.  We enforce that the species interface
+      states remain in $[0, 1]$ after adding the reactive source.
 

### 2021-09-27 simplified_sdc_intro/paper.tex
more styling

@@
 \newcommand{\Advss}[1]{{\left [{\mathcal{{A}}} \left(#1\right)\right]}}
+\newcommand{\Advsst}[1]{{\left [{\mathcal{\tilde{A}}} \left(#1\right)\right]}}
 \newcommand{\Advs}[1]{\boldsymbol{\mathcal{A}} \left(#1\right)}
@@
 multidimensional extensions are a straightforward modification to
-include the CTU hydrodynamics scheme.
-The system appears as:
-\begin{align}
-\label{eq:conserved_system}
-\ddt{\rho} + \ddxs{(\rho u)} &= 0 \\
-\ddt{(\rho X_k)} + \ddxs{(\rho X_k u)} &= \rho \omegadot(X_k) \\
-\ddt{(\rho \alpha_k)} + \ddxs{(\rho \alpha_k u)} &= \rho \omegadot(\alpha_k) \\
-\ddt{(\rho u)} + \ddxs{(\rho u^2 + p)} &= \rho g \\
-\ddt{(\rho E)} + \ddxs{(\rho E u + p u)} &= \rho ug + \rho \Sdot 
-\end{align}
-where $\rho$ is the mass density, $u$ is velocity, $E$ is specific
-total energy, $p$ is the pressure, and we carry nuclear species mass
-fractions, $X_k$, and auxiliary composition variables, $\alpha_l$.  The specific total
-energy relates to the specific internal energy, $e$, as $E = e + u^2/2$,
-and we also separately evolve $e$, as part of a dual
-energy formulation (see \citealt{bryan:1995,wdmergerI}):
-\begin{equation}
-\ddt{(\rho e)} + \ddxs{(\rho e u)} + p \ddx{u} = \rho \Sdot
-\end{equation}
- The mass fractions are constrained to sum to 1, $\sum_k X_k = 1$, but
- no such constraint exists on the auxiliary variables.  Gravity,
- represented by the gravitational acceleration $g$ appears as a source
- term.  Finally, $\omegadot(X_k)$ is the creation rate for species
- $k$, $\omegadot(\alpha_l)$ is the creation rate for auxiliary
- composition variable $l$, and $\dot{S}$ is the energy generation rate
- per unit mass.  Aside from the $p\partial u/\partial x$ (``pdV'')
- term in the internal energy equation, this system is in conservative
- form with source terms.
-
-To close the system, we need the equation of state.  For a system
-where the composition is completely specified by the mass fractions,
-$X_k$, the equation of state would take the form:
-\begin{equation}
-p = p(\rho, X_k, e)
-\end{equation}
-However, as we'll see shortly, when we use the NSE table, we are using
-the results of a much larger network then can be represented by the
-mass fractions, and in this case, we rely on the auxiliary composition
-variables, $\alpha_l$, to describe the state of the composition, and
-our EOS has the form:
-\begin{equation}
-p = p(\rho, \alpha_l, e)
-\end{equation}
-
-For convenience, we'll use $\Uc$ to represent the conserved state
-and split the sources up into pure hydrodynamic and reactive parts:
+include the CTU hydrodynamics scheme.  Our conserved variables are
 \begin{equation}
@@
            \rho e \end{array}\right )
-  \qquad
+\end{equation}
+where $\rho$ is the mass density, $u$ is velocity, $E$ is specific
+total energy, $p$ is the pressure, and we carry nuclear species mass
+fractions, $X_k$, and auxiliary composition variables, $\alpha_l$.  The specific total
+energy relates to the specific internal energy, $e$, as $E = e + u^2/2$,
+and we also separately evolve $e$, as part of a dual
+energy formulation (see \citealt{bryan:1995,wdmergerI}).
+ The mass fractions are constrained to sum to 1, $\sum_k X_k = 1$, but
+ no such constraint exists on the auxiliary variables. 
+Defining the hydrodynamical fluxes:
+\begin{equation}
+  \Fb(\Uc) = \left ( \begin{array}{c}
+         \rho u \\
+         \rho X_k u \\
+         \rho \alpha_l u \\
+         \rho u^2 p \\
+         (\rho E + p) u \\
+         \rho e u \end{array}\right )
+\end{equation}
+We can write the system in conservative form for all state variables aside from $(\rho e)$ as:
+\begin{equation}
+  \ddt{\Uc} + \ddx{\Fb(\Uc)} = \Sc(\Uc)
+\end{equation}
+where for the special case of $(\rho e)$, we have an additional ``pdV'' term:
+\begin{equation}
+\ddt{(\rho e)} + \ddx{F(\rho e)} + p \ddx{u} = S(\rho e)
+\end{equation}
+
+We'll split the source term, $\Sc(\Uc)$ into pure hydrodynamic and reactive parts:
+\begin{equation}
   \Hb(\Uc) = \left ( \begin{array}{c}
@@
 \end{equation}
-We will represent the total source as $\Sc(\Uc)$:
+with 
 \begin{equation}
@@
 \end{equation}
-We note that the internal energy ``pdV'' work is not treated as a source
-term but is instead constructed with the hydrodynamical fluxes are computed in the CTU method.
+Here, gravity, represented by the gravitational acceleration $g$
+appears as a pure hydrodynamical source term.  From reactions,
+$\omegadot(X_k)$ is the creation rate for species $k$,
+$\omegadot(\alpha_l)$ is the creation rate for auxiliary composition
+variable $l$, and $\dot{S}$ is the energy generation rate per unit
+mass.  We note that the internal energy ``pdV'' work is not treated as
+a source term but is instead constructed with the hydrodynamical
+fluxes are computed in the CTU method.
+
+Aside from the $p\partial u/\partial x$ (``pdV'') term in the internal
+energy equation, this system is in conservative form with source
+terms.
+
+We'll define the advective terms with hydrodynamic sources, $\Advs{\Uc}$ as
+\begin{equation}
+\Advs{\Uc} = -\ddx{\Fb(\Uc)} + \Hb(\Uc)
+\end{equation}
+for the general case, with the $(\rho e)$ component again having the extra
+``pdV'' term:
+\begin{equation}
+\mathcal{A}(\rho e) = -\ddx{F(\rho e)} -p \ddx{u} + H(\rho e)
+\end{equation}
+
+To close the system, we need the equation of state.  For a system
+where the composition is completely specified by the mass fractions,
+$X_k$, the equation of state would take the form:
+\begin{equation}
+p = p(\rho, X_k, e)
+\end{equation}
+However, as we'll see shortly, when we use the NSE table, we are using
+the results of a much larger network then can be represented by the
+mass fractions, and in this case, we rely on the auxiliary composition
+variables, $\alpha_l$, to describe the state of the composition, and
+our EOS has the form:
+\begin{equation}
+p = p(\rho, \alpha_l, e)
+\end{equation}
 
@@
 \end{equation}
-where $\Advs{\Uc}$ includes all terms in system
-\ref{eq:conserved_system} except for the reactive source terms,
-$\Rb(\Ub)$ (i.e., it includes the flux divergences, ``pdV'' term for
-internal energy, and the pure hydrodynamical sources).  In this approach, we 
+In this approach, we 
 approximate the advective term
@@
       by first ignoring the hydrodynamics sources,
+      \begin{equation}
+        \Advt{\Uc}^{n+1/2,(k)}_{i} =
+          - \frac{\Fb^\xv(\Uc^{n+1/2,(k)}_{i+1/2}) - \Fb^\xv(\Uc^{n+1/2,(k)}_{i-1/2})}{\Delta x}
+      \end{equation}
+      for the general case, and
       \begin{align}
-        \Advt{\Uc}^{n+1/2,(k)}_{i} =
-          &- \frac{\Fb^\xv(\Uc^{n+1/2,(k)}_{i+1/2}) - \Fb^\xv(\Uc^{n+1/2,(k)}_{i-1/2})}{\Delta x}
+        \Advsst{\rho e}^{n+1/2,(k)}_{i} =
+          &- \frac{F\left((\rho e)^{n+1/2,(k)}_{i+1/2}\right) - F\left((\rho e)^{n+1/2,(k)}_{i-1/2}\right)}{\Delta x} \\
+          &- \left (\frac{p_{i+1/2}^{n+1/2,(k)} + p_{i-1/2}^{n+1/2,(k)}}{2} \right )\frac{u_{i+1/2}^{n+1/2,(k)} - u_{i-1/2}^{n+1/2,(k)}}{\Delta x}
       \end{align}
+      for $(\rho e)$.
+
     Now the conservative hydrodynamics source terms are computed by first updating to the
-    new state with advection and the old-time source term applied for the full $\Delta t$ as: \MarginPar{little more complicated than this---we don't include the pdV source here}
+    new state with advection and the old-time source term applied for the full $\Delta t$ as: 
     \begin{equation}
@@
     correct the advective term so that we have a time-centered
-    source: \MarginPar{are we actually doing this correctly for SDC?}
+    source.
     The final advective update term is then\footnote{For a source like gravity, this update can be done first for $\rho$ and then define the new momentum source using $\rho^{\star\star}$ and likewise for energy}:
@@
 
-    \MarginPar{if $\Shydro$ depended on the outcome of the burn, we might need to do things differently here}
+    We note that this procedure works because the source terms here do
+    not depend on the outcome of the reactions.
 
@@
     of the Jacobian in appendix \ref{sec:app:jacobian}.
-
-    Note that our advection terms are piecewise
-    constant \MarginPar{need to think about this more} approximations,
-    so when we integrate our system, their contributions will be
-    linear in time.  Since $(\rho X_k)$ sees a linear advective term,
-    $\rho$ does as well.  $(\rho u)$ will also have linear
-    contributions in time from advection.
 

### 2021-10-17 simplified_sdc_intro/paper.tex
new figure

@@
 \end{equation}
-where $K$ is the kinetic energy, $K = |u|^2/2$.  Since the density and velocity
-are unchanged by reactions, our energy equation becomes:
+where $K$ is the kinetic energy, $K = \rho |u|^2/2$.  Since the density and velocity
+are unchanged by reactions (when Strang splitting), our energy equation becomes:
 \begin{equation}

### 2021-10-17 simplified_sdc_intro/sdc_strang_flowchart.py

@@ -1,0 +1,67 @@
+import matplotlib
+import matplotlib.pyplot as plt
+import numpy as np
+
+matplotlib.rc('text', usetex=True) 
+matplotlib.rcParams.update({"font.size": 18})
+matplotlib.rcParams['text.latex.preamble']=r"""
+\usepackage{amsmath}
+\newcommand{\pluseq}{\mathrel{+}=}"""
+
+# first strang
+
+fig = plt.figure()
+ax = fig.add_subplot(111)
+ax.axis("off")
+
+
+dy = 1.0
+dx = 1.0
+
+eps = 0.075
+
+# top line
+
+y = 2*dy
+
+ax.arrow(0, y, dx, 0, head_width=0.075, length_includes_head=True,
+         linewidth=2.0, color="C0")
+
+ax.text(0-eps, y, r"$\boldsymbol{\mathcal{U}}^n$", verticalalignment="center", horizontalalignment="right", color="C0")
+ax.text(dx+eps, y, r"$\boldsymbol{\mathcal{U}}^\star$", verticalalignment="center", horizontalalignment="left", color="C0")
+
+ax.text(dx/2, y+eps, r"$\frac{d\boldsymbol{\mathcal{U}}}{dt} = {\bf R}$", verticalalignment="bottom", horizontalalignment="center", color="C0")
+
+# middle line
+
+y = dy
+
+ax.arrow(0, y, 2*dx, 0, head_width=0.075, length_includes_head=True, linewidth=2.0, color="C1")
+
+ax.text(0-eps, y, r"$\boldsymbol{\mathcal{U}}^\star$", verticalalignment="center", horizontalalignment="right", color="C1")
+ax.text(2*dx+eps, y, r"$\boldsymbol{\mathcal{U}}^{n+1,\star}$", verticalalignment="center", horizontalalignment="left", color="C1")
+
+ax.text(dx, y+eps, r"$\boldsymbol{\mathcal{U}^\star} \pluseq \boldsymbol{\mathcal{A}}(\boldsymbol{\mathcal{U}^{\star,n+1/2}})$", verticalalignment="bottom", horizontalalignment="center", color="C1")
+
+# bottom line
+
+y = 0
+
+ax.arrow(dx, y, dx, 0, head_width=0.075, length_includes_head=True, linewidth=2.0, color="C0")
+
[…]

### 2021-11-19 simplified_sdc_intro/paper.tex
some stripping of NSE

@@
 %\title{A Simplified Spectral Deferred Correction Method for Coupling Hydrodynamics with Reaction Networks and Nuclear Statistical Equilibrium}
-\title{An Improved Method for Coupling Hydrodynamics with Reaction Networks and Nuclear Statistical Equilibrium}
+\title{An Improved Method for Coupling Hydrodynamics with Astrophysical Reaction Networks}
 
@@
 comprehensive SDC approach we demonstrated previously.  We apply the
-new method to some example problems and show how to modify it to work
-with a hybrid network consisting of a reaction ODE system and a table
-for nuclear statistical equilibrium.  This is all done in the
+new method to some example problems with small and moderately-sized
+astrophysical nuclear reaction networks.  This is all done in the
 framework of the \castro\ hydrodynamics code, and all algorithm
@@
 A particularly difficult phase of evolution to model is the nuclear
-statistical equilibrium that sets in for
-temperatures in excess of $\mbox{few} \times 10^9~\mathrm{K}$.  Physically, the
-forward and reverse rates of reaction should balance leading to an equilibrium.
+statistical equilibrium that sets in for temperatures in excess of
+$\mbox{few} \times 10^9~\mathrm{K}$.  Physically, the forward and
+reverse rates of reaction should balance leading to an equilibrium.
 With operator splitting, an NSE region will have a large positive flow
@@
 with a table of NSE abundances and the zone's composition set through
-table look-ups (e.g.\ \citealt{ma:2013}) \MarginPar{need more here}
+table look-ups (e.g.\ \citealt{ma:2013}).  The present work focuses on
+networks alone, but in a follow-on paper we will explore hybrid
+burning models consisting of networks and NSE tables.
 
@@
          \rho X_k u \\
-         \rho \alpha_l u \\
          \rho u^2 p \\
@@
   \Hb(\Uc) = \left ( \begin{array}{c}
-    0 \\
     0 \\
@@
      \rho \omegadot(X_k) \\
-     \rho \omegadot(\alpha_l) \\
      0 \\
@@
 appears as a pure hydrodynamical source term.  From reactions,
-$\omegadot(X_k)$ is the creation rate for species $k$,
-$\omegadot(\alpha_l)$ is the creation rate for auxiliary composition
-variable $l$, and $\dot{S}$ is the energy generation rate per unit
+$\omegadot(X_k)$ is the creation rate for species $k$
+ and $\dot{S}$ is the energy generation rate per unit
 mass.  We note that the internal energy ``pdV'' work is not treated as
@@
 \end{equation}
-However, as we'll see shortly, when we use the NSE table, we are using
-the results of a much larger network then can be represented by the
-mass fractions, and in this case, we rely on the auxiliary composition
-variables, $\alpha_l$, to describe the state of the composition, and
-our EOS has the form:
-\begin{equation}
-p = p(\rho, \alpha_l, e)
-\end{equation}
 
@@
 \begin{equation}
-\Ab^\xv(\qb) = \left ( \begin{array}{cccccc}
-    u & 0 & 0 & \rho & 0 & 0 \\
-    0 & u & 0 & 0    & 0 & 0 \\
-    0 & 0 & u & 0    & 0 & 0 \\
-    0 & 0 & 0 & u    & 1/\rho & 0 \\
-    0 & 0 & 0 & \Gamma_1 p & u & 0 \\
-    0 & 0 & 0 & \rho h & 0 & u
+\Ab^\xv(\qb) = \left ( \begin{array}{ccccc}
+    u & 0 &  \rho & 0 & 0 \\
+    0 & u &  0    & 0 & 0 \\
+    0 & 0 &  u    & 1/\rho & 0 \\
+    0 & 0 &  \Gamma_1 p & u & 0 \\
+    0 & 0 &  \rho h & 0 & u
   \end{array} \right )
@@
      0 \\
-     0 \\
      g \\
@@
      \omegadot(X_k) \\
-     \omegadot(\alpha_l) \\
      0 \\
@@
 in this paper.  We'll discuss this traditional approach next before moving
-on to our new time-coupling method.
+on to our new time-coupling method.  Figure~\ref{fig:flowchart} shows a graphical
+flowchart of the reaction and hydrodynamics advances in the two methods.
+
+\begin{figure}[t]
+\plottwo{sdc_strang_flowchart_I}{sdc_strang_flowchart_II}
+\caption{\label{fig:flowchart} flowchart}
+\end{figure}
 
@@
 \odt{(\rho X_k)} &= \rho \omegadot(X_k) \\
-\odt{(\rho \alpha_l)} &= \rho \omegadot(\alpha_l) \\
 \odt{(\rho u)} &= 0 \\
@@
 \end{equation}
-We integrate Equations (\ref{eq:strang:e}) and (\ref{eq:strang:X}) using
-an implicit ODE solver designed for stiff systems of
-equations, VODE~\citep{vode}.
+We integrate Equations (\ref{eq:strang:e}) and (\ref{eq:strang:X})
+using an implicit ODE solver designed for stiff systems of equations,
+based on VODE~\citep{vode}.  We use a modified version of the
+VODE~\citep{vode} integrator---ported to C++ with checks added to the
+timestep rejection logic that ensure that the species mass fractions
+stay between 0 and \MarginPar{we need to be clear that there is
+  different logic for SDC} 1\footnote{This modified version of VODE is
+available in our Microphysics repo:
+\url{https://github.com/starkiller-astro/Microphysics}.}.\MarginPar{discuss
+  tolerances}
 
@@
 timestep limiters based on the energy generation or abundance changes
-over a timestep: \MarginPar{do we enforce this in the NSE region?}
+over a timestep: 
 \begin{align}
@@
 
-\section{Reaction Networks and Nuclear Statistical Equilibrium}
-
-For this study we will use the 19 nuclei network containing
-\isot{H}{1}, \isot{He}{3}, \isot{He}{4}, \isot{C}{12}, \isot{N}{14},
-\isot{O}{16}, \isot{Ne}{20}, \isot{Mg}{24}, \isot{Si}{28},
-\isot{S}{32}, \isot{Ar}{36}, \isot{Ca}{40}, \isot{Ti}{44},
-\isot{Cr}{48}, \isot{Fe}{52}, \isot{Fe}{54}, \isot{Ni}{56}, protons
-(from photodisintegration), and neutrons.  This is based on the
-``aprox19'' network from \cite{aprox19} and originally described in
-\cite{Kepler}.  We use a modified version of the VODE~\citep{vode}
-integrator---ported to C++ with checks added to the timestep rejection
-logic that ensure that the species mass fractions stay between 0 and \MarginPar{we need to be clear that there is different logic for SDC}
-1\footnote{This modified version of VODE is available in our
-Microphysics repo:
-\url{https://github.com/starkiller-astro/Microphysics}.}.\MarginPar{discuss
-  tolerances} We combine this with a table that gives the nuclear
-statistical equilibrium (NSE) abundances in regions where the system
-is in NSE.  The NSE table was generated using a 127 nuclei reaction
-network and is the same as described in \cite{ma:2013}.  In our
-simulations, we carry all 19 isotopes in the main network in each zone
-and advect them in the hydrodynamics portion of the algorithm.  The
-composition of the larger 127 nuclei network is mapped into the 19
-isotopes we carry according to Table \ref{table:nuclei}.
-
-The NSE table requires: \MarginPar{need a statement about how the table was constructed and what NSE means}
-\begin{equation}
-  \label{eq:aux:ye}
-  Y_e = \sum_k \frac{Z_k X_k}{A_k}
-\end{equation}
-(where $A_k$ and $Z_k$ are the atomic weight and atomic number of nucleus $k$) and provides
-\begin{align}
-\label{eq:aux:abar}
-\bar{A} &= \left [ \sum_k \frac{X_k}{A_k} \right ]^{-1} \\
-\label{eq:aux:bea}
-\left (\frac{B}{A} \right ) &= \sum_k \frac{B_k X_k}{A_k}
-\end{align}
-where $B_k$ is the binding energy of nucleus $k$.  In our simulations,
-we store these 3 quantities as auxiliary data that is carried along
-with the rest of the fluid state in each zone.  The table also returns
-values of the mass fractions mapped onto the 19-isotopes we carry, $X_k$, and the time-derivative of $Y_e$, $dY_e/dt$.
-We use the notation:
-\begin{equation}
-\nse{\rho,T,Y_e} \rightarrow \bar{A}, X_k, (B/A), dY_e/dt
-\end{equation}
-to represent the NSE table call and its inputs and outputs.
-
-We can derive evolution equations for each of these composition quantities as:
-\begin{align}
-\frac{DY_e}{Dt} &= \sum_k \frac{Z_k}{A_k} \frac{DX_k}{Dt} = \sum_k \frac{Z_k}{A_k} \omegadot_k \\
-\frac{D\bar{A}}{Dt} &= -\bar{A}^2 \sum_k \frac{1}{A_k} \frac{DX_k}{Dt} = -\bar{A}^2 \sum_k \frac{1}{A_k} \omegadot_k \\
-\frac{D}{Dt} \left (\frac{B}{A} \right ) &= \sum_k \frac{B_k}{A_k} \frac{DX_k}{Dt} = \sum_k \frac{B_k}{A_k} \omegadot_k
-\end{align}
-For Strang split coupling of hydro and reactions, $DX_k/Dt = 0$,
-therefore each of these auxillary equations obeys an advection
-equation in the hydro part of the advancement.  In the SDC algorithm,
-there will be a reactive source (an $\Ic_q$) for each of these that is
-computed in the same manner as above.  We note that our NSE table
-provides the evolution of $Y_e$ due to reactions ($DY_e/Dt$) directly.
-
-The compositional quantities
-it carries, $\bar{A}$ and $Y_e$ are not representable from the 19
-isotopes we carry in the main network. For this reason, when we are
-using the NSE network, we always provide these two composition quantities as
-inputs to the EOS rather than using the $X_k$ directly.
-the EOS directly from the auxiliary state in each zone instead of
-using the $X_k$ directly.
-Our equation of state needs the mean charge per nucleus, $\bar{Z}$, in addition
-to the auxiliary quantities, which is computed as
-\begin{equation}
-\bar{Z} = \bar{A} \sum_k \frac{Z_k X_k}{A_k} = \bar{A} Y_e
-\end{equation}
-(see, e.g., \citealt{flash}).
-
-\subsection{Initialization}
-
-We initialize the mass fractions, $X_k$, and then compute the
-electron fraction, $Y_e$, using (\ref{eq:aux:ye}).
-For the two other composition
-quantities we carry, $\bar{A}$, and $(B/A)$, we need values that are
-consistent with the value of $Y_e$ and the nuclei stored in the NSE
-table.  Therefore, if the thermodynamic conditions put us in NSE
-(using the conditions defined below), then we obtain $\bar{A}$ and
-$(B/A)$ from the NSE table.  Otherwise, we compute these directly from
-$X_k$ using (\ref{eq:aux:abar}) and (\ref{eq:aux:bea}).
-
-\subsection{NSE condition}
-
-We treat a zone as being in NSE if the density and temperature exceed
-some threshold and the composition is mainly $\alpha$ and Fe-group
-nuclei, where the Fe-group nuclei are \isot{Cr}{48}, \isot{Fe}{52},
-\isot{Fe}{54}, and \isot{Ni}{56}.  The full condition is:
-\begin{align}
-\rho &> \rhonse \\
-T &> \tnse \\
-X(\isotm{C}{12}) &< \Anse \\
-X(\isotm{He}{4}) + \sum_{k \in \mathrm{Fe-group}} X_k &> \Bnse \\
-X(\isotm{Si}{26}) &< \Cnse 
-\end{align}
-Typical values are $\rhonse = 2\times 10^6~\gcc$, $T_\mathrm{nse} =
-3\times 10^9~\mathrm{K}$, $\Anse = 0.12$, $\Bnse=
-0.88$, and $\Cnse = 0.01$.  The values of $\Anse$ and $\Bnse$ are based on \citet{ma:2013}.
-The value of $\Cnse$ ensures that Si-burning has ended before invoking the table.
-
-\subsection{Strang-split algorithm for NSE}
-
-For Strang splitting, we alternate the reaction and hydrodynamics,
-with each operation working on the output of the previous operation.
-If we define an advection operator over a timestep $\Delta t$ acting on $\Uc$ as
-$\mathbb{A}_{\Delta t}(\Uc)$ and a reaction operator over a timestep of
-$\Delta t/2$ acting on $\Uc$ as $\mathbb{R}_{\Delta t/2}(\Uc)$, then the advance appears
-as:
-\begin{align}
-  \Uc^\star &= \mathbb{R}_{\Delta t/2}(\Uc^n) \\
-  \Uc^{n+1,\star} &= \mathbb{A}_{\Delta t}(\Uc^\star) \\
-  \Uc^{n+1} &= \mathbb{R}_{\Delta t/2}(\Uc^{n+1,\star})
-\end{align}
-Tthe hydrodynamic and reactive substeps over
-the overall time-advancemenet scheme using the aprox19 + NSE network
-are as follows:
-\begin{itemize}
-
-\item {\em Hydrodynamics update}
-
-  At the beginning of each hydrodynamic update we have an input state,
-  $\Uc_{\rm in}$ that we wish to integrate over $\Delta t$ to obtain $\Uc_{\rm out}$
-  The hydrodynamics update proceeds as normal, but with an advection
-  equation for each of the auxiliary composition variables:
-  \begin{align}
-    \ddt{(\rho Y_e)} + \ddx{(\rho Y_e u)} &= 0 \\
-    \ddt{(\rho \bar{A})} + \ddx{(\rho \bar{A} u)} &= 0 \\
-    \ddt{[\rho (B/A)]} + \ddx{[\rho (B/A) u]} &= 0 
-  \end{align}
-
-\item {\em Reactive update}
-
-  At the beginning of each reactive update we have an input state,
-  $\Uc_{\rm in}$ that we wish to integrate over $\Delta t/2$ to obtain $\Uc_{\rm out}$.
-
-  \begin{itemize}
-
-    \item For a zone that is in NSE:
-
-      The goal is to update the composition and thermodynamics due to the 
-      change in the nuclei abundances over the (half-)timestep.  \citet{ma:2013}
-      uses a first-order in time difference to get the new composition state
-      and evaluates the energy release from the change in binding energy in the NSE
-      state.  They only apply a fraction (0.3) of the energy release to the internal
-      energy in a zone, to avoid a potential instability that can arise if too much
-      energy is added to a zone in a single timestep.  We prefer to deal with this
-      issue through the retry mechanism already built into \castro.
-
-      The approach we use begins with calling the NSE table with the input state:
-      \begin{equation}
-        \nse{\rho_\inp, T_\inp, (Y_e)_\inp} \rightarrow \bar{A}^\star, (X_k)^\star, (B/A)^\star, dY_e/dt
-        \end{equation}
-      We indicate with a $\star$ that most of those values are
-      provisional, and we will seek a better value in the corrector step.
-      We only update $Y_e$ from this call:
-      \begin{equation}
-        (Y_e)_\out = (Y_e)_\inp + \Delta t \frac{dY_e}{dt}
-      \end{equation}
-      and then we use the table again, but with the updated $Y_e$
-      (note that $\rho_{\rm out}=\rho_{\rm in}$ in the Strang reaction formulation):
-      \begin{equation}
-        \nse{\rho_\out, T_\inp, (Y_e)_\out} \rightarrow \bar{A}_\out, (X_k)_\out, (B/A)_\out, dY_e/dt
-      \end{equation}
-      In this formulation, we have not updated the temperature. \MarginPar{discuss this more, and perhaps switch to a mode where we try to predict T}
-      This corrects the composition so that it is consistent with the updated
-      $Y_e$.  We use these values $\bar{A}_\out$, $(X_k)_\out$, and $(B/A)_\out$ to then complete the update, computing the energy release, $\Sdot$ as:
-      \begin{equation}
-        \label{eq:nse_energy}
-        \Sdot = \left [ \left ( \frac{B}{A} \right )_\out -
-          \left ( \frac{B}{A} \right )_\inp \right ] N_A \frac{1}{\Delta t}
-      \end{equation}
-      \MarginPar{we could imagine looping over the EOS to get an updated T and then the table to get the updated $\bar{A}$, to see if that converges?}
-
-    \item For zones not in NSE:
-
-    \begin{itemize}
-    \item Integrate the full reaction network (Eqs.~\ref{eq:strang:e} and \ref{eq:strang:X}) as usual for Strang splitting
-    \item Update the aux quantities at the end of the burn using Eqs.~\ref{eq:aux:ye}, \ref{eq:aux:abar}, and \ref{eq:aux:bea} with the new mass fractions, $X_k$.
-    \end{itemize}
-  \end{itemize}
-\end{itemize}
-
-\subsection{SDC-NSE Coupling}
-
-With SDC evolution, when we are in NSE, we need to do the advective
-and reactive updates together.  We want to compute:
-\begin{equation}
-\Uc^{n+1,(k)} = \Uc^n + \Delta t \Adv{\Uc}^{n+1/2,(k)} + \Delta t \left [\Rb (\Uc) \right ]^{n+1/2,(k)}
-\end{equation}
-For density and momentum, we can do this update already, since there
-are no reactive sources.  That gives us $\rho^{n+1,(k)}$ and $(\rho
-\Ub)^{n+1,(k)}$.  For the other quantities, the NSE table gives only the
-instantaneous values. aside from $Y_e$.  We therefore do an iterative
-update.
-
-Calling the NSE table on the starting state gives us:
-\begin{equation}
-\nse{\rho^n, T^n, (Y_e)^n} \rightarrow \bar{A}^n, (X_k)^n, (B/A)^n, (dY_e/dt)^n
-\end{equation}
-We now compute a first approximation to the reactive source:
-\begin{align}
-R^{(0)}(\rho e) &= 0 \\
-R^{(0)}(\rho Y_e) &= \rho^n (dY_e/dt)^n \\
-R^{(0)}(\rho \bar{A}) &= 0
-\end{align}
-
-Now we iterate $\xi = 0, \ldots N-1$, doing the following:
-
-\begin{itemize}
-\item Update the energy as:
-   \begin{equation}
-     (\rho e)^{n+1,(\xi)} = (\rho e)^n + \Delta t \Advss{\rho e}^{n+1/2,(k)} + \Delta t R^{(\xi)}(\rho e)
-   \end{equation}
-
-\item Update the auxiliary data as:
-   \begin{equation}
-     (\rho \alpha_l)^{n+1,(\xi)} = (\rho \alpha_l)^n + \Delta t \Advss{\rho \alpha_l}^{n+1/2,(k)} + \Delta t R^{(\xi)}(\rho \alpha_l)
-   \end{equation}
-
-\item Compute the updated temperature, $T^{n+1,(k)}$ as
-  \begin{equation}
-    T^{n+1,(\xi)} = T(\rho^{n+1,(k)},  \alpha^{n+1,(\xi)},  e^{n+1,(\xi)})
-  \end{equation}
-  where $\alpha^{n+1,(\xi)} = (\rho  \alpha)^{n+1,(\xi)} / \rho^{n+1,(k)}$ and $e^{n+1,(\xi)} = (\rho e)^{n+1,(\xi)} / \rho^{n+1,(k)}$.
-
-\item Call the NSE table to get the new NSE state
-  \begin{equation}
-    \nse{\rho^{n+1,(k)}, T^{n+1,(\xi)}, (Y_e)^{n+1,(\xi)}} \rightarrow \bar{A}^{n+1,(\xi)}, (X_k)^{n+1,(\xi)}, (B/A)^{n+1,(\xi)}, (dY_e/dt)^{n+1,(\xi)}
-  \end{equation}
-
-\item Recompute the reactive source terms for the next iteration.
-
-  The energy generation rate is found as:
-  \begin{equation}
-    (\rho \Sdot)^{n+1/2,(\xi)} = \frac{\rho^{n+1,(k)} + \rho^n}{2}
-    \left [ \left ( \frac{B}{A} \right )^{n+1,(\xi)} -
-      \left ( \frac{B}{A} \right )^n \right ] N_A \frac{1}{\Delta t}
-  \end{equation}
-  and our new auxiliary sources are:
-\begin{align}
-R^{(\xi+1)}(\rho e) &= (\rho \Sdot)^{n+1/2,(\xi)} \\
-R^{(\xi+1)}(\rho Y_e) &= \frac{\rho^{n+1,(k)} + \rho^n}{2} \frac{1}{2} \left [ \left (\frac{dY_e}{dt} \right )^n + \left (\frac{dY_e}{dt} \right )^{n+1,(\xi)} \right ] \\
-R^{(\xi+1)}(\rho \bar{A}) &= \frac{\rho^{n+1,(k)} + \rho^n}{2} \frac{1}{\Delta t} \left [ \bar{A}^{n+1,(\xi)} - \frac{(\rho \bar{A})^n}{\rho^n} \right ]
-\end{align}
-
-\end{itemize}
-
-With the iteration complete, we can do the final update of the total energy using last iterations prediction of the source (for iteration $N$):
-\begin{equation}
-(\rho E)^{n+1,(k)} = (\rho E)^n + \Delta t \Advss{\rho E}^{n+1/2} + \Delta t (\rho \Sdot)^{n+1/2,(N)}
-\end{equation}
-and we take the values of the auxiliary state and mass fractions from the last call to the NSE table:
-\begin{align}
-(\rho Y_e)^{n+1,(k)} &= \rho^{n+1,(k)} (Y_e)^{n+1,(N-1)} \\
-(\rho \bar{A})^{n+1,(k)} &= \rho^{n+1,(k)} (\bar{A})^{n+1,(N-1)} \\
-(\rho (B/A))^{n+1,(k)} &= \rho^{n+1,(k)} (B/A)^{n+1,(N-1)} \\
-(\rho Y_e)^{n+1,(k)} &= \rho^{n+1,(k)} (X_k)^{n+1,(N-1)}
-\end{align}
-
-We note that this update is not fully second order, however, the
-treatment of $Y_e$ is, and we expect that to have the dominant effect
-in the thermodynamic evolution through NSE.
-
-\subsection{NSE Bailout}
-
-There is one additional part of the time integration algorithm.
-Because $T$ evolves during the reactions (for both Strang and SDC), it
-is possible for a zone to start out not in NSE but evolve into NSE
-during the reaction update.  When this happens, the ODE integrator may
-fail, because an excessive number of timesteps is required.  Rather
-than trigger our retry mechanism and throw away the entire timestep
-and start over with a smaller $\Delta t$, we first check if the zone
-evolved into NSE, and if so, we redo the zone update using the NSE
-prescription.  To ensure that this is not triggered at the very start
-of integration, we require a minimum number of steps (10) to have
-elapsed before checking for NSE.  After we leave the integrator, the
-state is returned to the burner driver where we finish the integration
-via NSE if the conditions satisfy the NSE criteria.  We optionally
-apply a relaxation factor, typically, $0.9$, to the NSE conditions to
-allow a state that is close to NSE after an integrator failure enter
-NSE.
-
 \section{Simulations}
@@
 
-\subsection{NSE convergence test problem}
-
-To test the coupling of the NSE table to the hydrodynamics, we run a
-similar problem as above, except with the thermodynamic conditions
-appropriate for the matter to be in NSE.  The initial conditions are:
-\begin{align}
-\rho &= \rho_0 \\
-T &= T_0 \left [ 1 + (\delta T) e^{-(r/\lambda)^2} \cos^6(\pi r/L) \right ] \\
-Y_e &= (Y_e)_0 \left [ 1 + (\delta Y_e) e^{-(r/\lambda)^2} \cos^6(\pi r/L) \right ] 
-\end{align}
-where $r$ is the distance from the center of the domain and the
-remaining parameters are listed in Table~\ref{table:nse}.  This has
-$Y_e$ varying between $[0.47, 0.5]$ initially\MarginPar{check}.  The mass fractions
-and remaining composition variables are then initialized from the NSE table, as described in section~\ref{sec:initialization}.
-
-The domain has a size $[0, L]^2$, and the timestep is fixed as:
-\begin{equation}
-\Delta t = 10^{-3} \left ( \frac{64}{N} \right )~\mathrm{s}
-\end{equation}
-where $N$ is the number of zones in each direction.
-
-\begin{deluxetable}{lcc}
-\tablecaption{\label{table:nse} NSE convergence test problem.}
-\tablehead{\colhead{parameter} & \colhead{value}}
-\startdata
-$\rho_0$ & $5\times 10^8~\gcc$ \\
-$T_0$    & $4\times 10^9~\mathrm{K}$ \\
-$(\delta T)$ & $0.2$ \\
-$(Y_e)_0$   & $0.5$ \\
-$(\delta Y_e)$ & $-0.05$ \\
-$\lambda$   & $2\times 10^7~\mathrm{cm}$ \\
-$L$         & $10^8~\mathrm{cm}$ \\
-\enddata
-\end{deluxetable}
-
-Figure~\ref{fig:nse_ye} shows the $Y_e$ profile for the Strang evolution case.
-
-\begin{figure}[t]
-\centering
-\plotone{ye_plot}
-\caption{\label{fig:nse_ye} Profile of $Y_e$ through the middle of the domain (slicing along $x$)
-for the $128^2$ Strang NSE test simulation.}
-\end{figure}
-
-% run on 04-06-21
-% Castro       git describe: 21.04-dirty
-% AMReX        git describe: 21.03-58-g87c81d0ca
-% Microphysics git describe: 21.04-1-g33a4d0fe
-
-\begin{deluxetable}{lllllll}
-\tablecaption{\label{table:nse_strang_methodII} Convergence ($L_1$ norm) for the NSE convergence
-problem using Strang splitting.}
-\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
-\startdata
- $\rho$                      & $4.522 \times 10^{19}$  & 1.659  & $1.432 \times 10^{19}$  & 1.542  & $4.917 \times 10^{18}$  \\
- $\rho u$                    & $1.883 \times 10^{28}$  & 1.681  & $5.873 \times 10^{27}$  & 1.524  & $2.042 \times 10^{27}$  \\
- $\rho v$                    & $1.883 \times 10^{28}$  & 1.681  & $5.873 \times 10^{27}$  & 1.524  & $2.042 \times 10^{27}$  \\
- $\rho E$                    & $5.933 \times 10^{37}$  & 1.613  & $1.940 \times 10^{37}$  & 1.411  & $7.293 \times 10^{36}$  \\
- $\rho e$                    & $5.931 \times 10^{37}$  & 1.613  & $1.939 \times 10^{37}$  & 1.411  & $7.291 \times 10^{36}$  \\
- $T$                         & $9.885 \times 10^{20}$  & 1.207  & $4.281 \times 10^{20}$  & 1.058  & $2.057 \times 10^{20}$  \\
- $\rho X(\isotm{H}{1})$      & $4.522 \times 10^{-11}$ & 1.659  & $1.432 \times 10^{-11}$ & 1.542  & $4.917 \times 10^{-12}$ \\
- $\rho X(\isotm{He}{4})$     & $5.086 \times 10^{17}$  & 1.430  & $1.887 \times 10^{17}$  & 1.187  & $8.288 \times 10^{16}$  \\
- $\rho X(\isotm{C}{12})$     & $1.585 \times 10^{12}$  & 1.829  & $4.462 \times 10^{11}$  & 1.433  & $1.652 \times 10^{11}$  \\
- $\rho X(\isotm{O}{16})$     & $8.060 \times 10^{12}$  & 1.706  & $2.470 \times 10^{12}$  & 1.347  & $9.710 \times 10^{11}$  \\
- $\rho X(\isotm{Cr}{48})$    & $1.149 \times 10^{19}$  & 1.400  & $4.353 \times 10^{18}$  & 1.258  & $1.820 \times 10^{18}$  \\
- $\rho X(\isotm{Fe}{52})$    & $9.018 \times 10^{19}$  & 1.173  & $4.001 \times 10^{19}$  & 1.127  & $1.832 \times 10^{19}$  \\
- $\rho X(\isotm{Fe}{54})$    & $5.036 \times 10^{20}$  & 1.731  & $1.517 \times 10^{20}$  & 1.727  & $4.582 \times 10^{19}$  \\
- $\rho X(\isotm{Ni}{56})$    & $5.686 \times 10^{20}$  & 1.622  & $1.847 \times 10^{20}$  & 1.563  & $6.252 \times 10^{19}$  \\
- $\rho Y_e$                  & $2.379 \times 10^{19}$  & 1.678  & $7.431 \times 10^{18}$  & 1.578  & $2.489 \times 10^{18}$  \\
- $\rho \bar{A}$              & $5.141 \times 10^{21}$  & 1.230  & $2.191 \times 10^{21}$  & 1.087  & $1.032 \times 10^{21}$  \\
- $\rho (B/A)$                & $4.149 \times 10^{20}$  & 1.693  & $1.283 \times 10^{20}$  & 1.562  & $4.345 \times 10^{19}$  \\
-\enddata
-\end{deluxetable}
-
-% run on 04-06-21
-% Castro       git describe: 21.04-dirty
-% AMReX        git describe: 21.03-58-g87c81d0ca
-% Microphysics git describe: 21.04-1-g33a4d0fe
-
-\begin{deluxetable}{lllllll}
-\tablecaption{\label{table:nse_sdc_methodII} Convergence ($L_1$ norm) for the NSE convergence
-problem using the simplified-SDC algorithm.}
-\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
-\startdata
- $\rho$                      & $4.553 \times 10^{19}$  & 1.636  & $1.465 \times 10^{19}$  & 1.583  & $4.891 \times 10^{18}$  \\
- $\rho u$                    & $1.949 \times 10^{28}$  & 1.650  & $6.210 \times 10^{27}$  & 1.513  & $2.176 \times 10^{27}$  \\
- $\rho v$                    & $1.949 \times 10^{28}$  & 1.650  & $6.210 \times 10^{27}$  & 1.513  & $2.176 \times 10^{27}$  \\
- $\rho E$                    & $5.691 \times 10^{37}$  & 1.622  & $1.848 \times 10^{37}$  & 1.561  & $6.266 \times 10^{36}$  \\
- $\rho e$                    & $5.689 \times 10^{37}$  & 1.622  & $1.848 \times 10^{37}$  & 1.561  & $6.265 \times 10^{36}$  \\
- $T$                         & $5.778 \times 10^{20}$  & 1.740  & $1.729 \times 10^{20}$  & 1.702  & $5.316 \times 10^{19}$  \\
- $\rho X(\isotm{H}{1})$      & $4.553 \times 10^{-11}$ & 1.636  & $1.465 \times 10^{-11}$ & 1.583  & $4.891 \times 10^{-12}$ \\
- $\rho X(\isotm{He}{4})$     & $2.541 \times 10^{17}$  & 1.807  & $7.262 \times 10^{16}$  & 1.634  & $2.339 \times 10^{16}$  \\
- $\rho X(\isotm{C}{12})$     & $1.140 \times 10^{12}$  & 1.838  & $3.189 \times 10^{11}$  & 1.670  & $1.002 \times 10^{11}$  \\
- $\rho X(\isotm{O}{16})$     & $5.413 \times 10^{12}$  & 1.819  & $1.535 \times 10^{12}$  & 1.658  & $4.864 \times 10^{11}$  \\
- $\rho X(\isotm{Cr}{48})$    & $7.020 \times 10^{18}$  & 1.793  & $2.026 \times 10^{18}$  & 1.614  & $6.621 \times 10^{17}$  \\
- $\rho X(\isotm{Fe}{52})$    & $3.631 \times 10^{19}$  & 1.579  & $1.215 \times 10^{19}$  & 1.597  & $4.018 \times 10^{18}$  \\
- $\rho X(\isotm{Fe}{54})$    & $4.670 \times 10^{20}$  & 1.807  & $1.335 \times 10^{20}$  & 1.869  & $3.654 \times 10^{19}$  \\
- $\rho X(\isotm{Ni}{56})$    & $4.858 \times 10^{20}$  & 1.808  & $1.387 \times 10^{20}$  & 1.860  & $3.820 \times 10^{19}$  \\
- $\rho Y_e$                  & $2.372 \times 10^{19}$  & 1.649  & $7.564 \times 10^{18}$  & 1.587  & $2.518 \times 10^{18}$  \\
- $\rho \bar{A}$              & $4.009 \times 10^{21}$  & 1.693  & $1.240 \times 10^{21}$  & 1.690  & $3.843 \times 10^{20}$  \\
- $\rho (B/A)$                & $4.222 \times 10^{20}$  & 1.666  & $1.330 \times 10^{20}$  & 1.605  & $4.374 \times 10^{19}$  \\
-\enddata
-\end{deluxetable}
-
-\subsection{Detonation}
-
-The purpose here is to look at what timestep is taken when we use the nuclear burning limiters.
-Run with aprox19 only, Strang and SDC, then aprox19 + NSE, Strang and NSE.
-
-Look at number of SDC iterations too.
-
-How does the timestep limiter work in the middle of SDC iterating?
-
-aprox21
-
-look at number of RHS calls, wallclock time
-
-output every step and then for a single zone plot enuc every timestep to see if it oscillates
-
 \subsection{Reacting Buoyant Bubble}
@@
 
-\subsection{Massive Star}
+\subsection{Sub-Chandra Double Detonation}
 

### 2021-11-19 simplified_sdc_intro/sdc_strang_flowchart.py

@@ -6,7 +6,9 @@ matplotlib.rc('text', usetex=True)
 matplotlib.rcParams.update({"font.size": 18})
 matplotlib.rcParams['text.latex.preamble']=r"""
 \usepackage{amsmath}
-\newcommand{\pluseq}{\mathrel{+}=}"""
+\newcommand{\pluseq}{\mathrel{+}=}
+\newcommand{\Uc}{\boldsymbol{\mathcal{U}}}
+"""
 
 # first strang
 
@@ -27,10 +29,10 @@ y = 2*dy
 ax.arrow(0, y, dx, 0, head_width=0.075, length_includes_head=True,
          linewidth=2.0, color="C0")
 
-ax.text(0-eps, y, r"$\boldsymbol{\mathcal{U}}^n$", verticalalignment="center", horizontalalignment="right", color="C0")
-ax.text(dx+eps, y, r"$\boldsymbol{\mathcal{U}}^\star$", verticalalignment="center", horizontalalignment="left", color="C0")
+ax.text(0-eps, y, r"$\Uc^n$", verticalalignment="center", horizontalalignment="right", color="C0")
+ax.text(dx+eps, y, r"$\Uc^\star$", verticalalignment="center", horizontalalignment="left", color="C0")
 
-ax.text(dx/2, y+eps, r"$\frac{d\boldsymbol{\mathcal{U}}}{dt} = {\bf R}$", verticalalignment="bottom", horizontalalignment="center", color="C0")
+ax.text(dx/2, y+eps, r"$\frac{d\Uc}{dt} = {\bf R}(\Uc)$", verticalalignment="bottom", horizontalalignment="center", color="C0")
 
 # middle line
 
@@ -38,10 +40,10 @@ y = dy
 
 ax.arrow(0, y, 2*dx, 0, head_width=0.075, length_includes_head=True, linewidth=2.0, color="C1")
 
-ax.text(0-eps, y, r"$\boldsymbol{\mathcal{U}}^\star$", verticalalignment="center", horizontalalignment="right", color="C1")
-ax.text(2*dx+eps, y, r"$\boldsymbol{\mathcal{U}}^{n+1,\star}$", verticalalignment="center", horizontalalignment="left", color="C1")
+ax.text(0-eps, y, r"$\Uc^\star$", verticalalignment="center", horizontalalignment="right", color="C1")
[…]

### 2021-11-21 simplified_sdc_intro/paper.tex
add figure
more updates

@@
 \newcommand{\maestro}{{\sf Maestro}}
+\newcommand{\pynucastro}{{\sf pynucastro}}
 \newcommand{\flash}{{\sf Flash}}
@@
 \begin{figure}[t]
-\plottwo{sdc_strang_flowchart_I}{sdc_strang_flowchart_II}
-\caption{\label{fig:flowchart} flowchart}
+\plotone{sdc_strang_flowchart_I}
+\caption{\label{fig:strang_flowchart} flowchart}
+\end{figure}
+
+\begin{figure}[t]
+\plotone{sdc_strang_flowchart_II}
+\caption{\label{fig:sdc_flowchart} flowchart}
 \end{figure}
@@
 
-\subsection{Reacting Buoyant Bubble}
-
-We next consider a buoyant, reacting bubble in a stratified,
-plane-parallel atmosphere.  This test exercises the treatment of
-hydrodynamical sources in the SDC integration.  We use the set of
-initial conditions that was first shown in \cite{ABNZ:III}.
+\subsection{X-ray Burst}
+
+We next consider a flame propagating through an accreted helium layer on a neutron star,
+as described in \cite{harpole:2021}.
 
 \subsection{Sub-Chandra Double Detonation}
+
+We use a network with 28 nuclei and 96 rates that captures He burning
+with links for $(\alpha,p)(p,\gamma)$ reactions and the rates
+involving \isot{N}{14} to bypass
+$\isotm{C}{12}(\alpha,\gamma)\isotm{O}{16}$ as described in
+\cite{shenbildsten}.  The rates are taken from the ReacLib library
+\citep{reaclib} and the network is written by
+\pynucastro~\citep{pynucastro}.  The nuclei and links in the network are
+shown in Figure~\ref{fig:subch_network}.  The many reaction pathways in this network
+make it a good test for the SDC algorithm.
+
+The initial model is generated following the procedure described in \citep{subchandra}.
+For the present work, we are only concerned with the behavior of the time-integration strategy
+and not on the astrophysical implications of the physical setup we explore.
+
+\begin{figure}[t]
+\centering
+\plotone{subch2}
+\caption{\label{fig:subch_network} Reaction network for the double detonation problem.}
+\end{figure}
 
@@
           python (\url{https://www.python.org/}), 
+          \pynucastro\ \citep{pynucastro},
           SymPy \citep{sympy}

### 2021-12-22 simplified_sdc_intro/paper.tex
some more text

@@
 \newcommand{\inp}{{\rm in}}
+
+\newcommand{\atolspec}{{\epsilon_\mathrm{abs,spec}}}
+\newcommand{\rtolspec}{{\epsilon_\mathrm{rel,spec}}}
+\newcommand{\atolener}{{\epsilon_\mathrm{abs,ener}}}
+\newcommand{\rtolener}{{\epsilon_\mathrm{rel,ener}}}
 
@@
 
-Do this in 1-d
-Show the amount of work needed for Strang vs. SDC
-
-Play around with the threshold where NSE kicks in
+We do a suite of runs with both Strang and the simplified-SDC
+integration, using CFL numbers of 0.2, 0.4, and 0.8 and reaction
+network tolerances of $\atolspec = \rtolspec = 10^{-5}$ and $\atolener
+= \rtolener = 10^{-6}$.  We also run two sets with tighter tolerances.
+The runs labeled ``mixed\_tols'' have $\atolspec = 10^{-8}$ and the runs
+labeled ``tol1.e-6'' have $\atolspec = \rtolspec = 10^{-6}$, with the other
+tolerances unchanged.
+
+Overview plots of CFL = 0.8
+
+Nonaka plots
+
+Zoom panels showing different runs
+
+Mass of different species (as a plot)
+
+1-d slices along x=0?
 

### 2021-12-27 simplified_sdc_intro/paper.tex
more updates

@@
 new method to some example problems with small and moderately-sized
-astrophysical nuclear reaction networks.  This is all done in the
-framework of the \castro\ hydrodynamics code, and all algorithm
-implementations are freely available.
+astrophysical nuclear reaction networks and explore the timestep size
+and reaction network tolerances to show that the simplified-SDC
+approach compares favorably with traditional Strang operator
+splitting.  This is all done in the framework of the
+\castro\ hydrodynamics code, and all algorithm implementations are
+freely available.
 \end{abstract}
@@
 Reaction networks tend to be stiff, requiring implicit integration
-techniques to stably integrate the system.  In contrast, compressible
+techniques to stably integrate the system~\citep{BYRNE19871}.  In contrast, compressible
 hydrodynamics flows are limited by the (often much longer)
@@
 a code.  The easiest way to improve the coupling is to cut the
-timestep, but this makes simulations prohibitively expensive.
-Sometimes the burning is simply halted on a zone-by-zone basis when
-NSE conditions are reached (e.g., as in \citealt{hedet}).
-Alternately, at high temperatures, a reaction network can be replaced
-with a table of NSE abundances and the zone's composition set through
-table look-ups (e.g.\ \citealt{ma:2013}).  The present work focuses on
-networks alone, but in a follow-on paper we will explore hybrid
-burning models consisting of networks and NSE tables.
+timestep (see, e.g., \citealt{couch:2015}), but this makes simulations
+prohibitively expensive.  Sometimes the burning is simply halted on a
+zone-by-zone basis when NSE conditions are reached (e.g., as in
+\citealt{hedet}).  Alternately, at high temperatures, a reaction
+network can be replaced with a table of NSE abundances and the zone's
+composition set through table look-ups (e.g.\ \citealt{ma:2013}).  The
+present work focuses on networks alone, but in a follow-on paper we
+will explore hybrid burning models consisting of networks and NSE
+tables.
 
@@
 and GPU architectures \citep{castro_gpu}.  For hydrodynamics, the
-corner transport upwind (CTU) \citep{ppmunsplit} method with the
-piecewise parabolic method (PPM) \citep{ppm,millercolella:2002} is
+corner transport upwind (CTU; \citealt{ppmunsplit}) method with the
+piecewise parabolic method (PPM; \citealt{ppm,millercolella:2002}) is
 used.  \castro\ includes self-gravity, rotation, arbitrary equations
@@
 
-The time-integration approach presented here is considerably simpler than
-the SDC method of \citet{castro_sdc}, but allows us to reuse the main CTU
-hydrodynamics construction and a largely similar ODE integration
-scheme, making this method easier to add to existing simulation codes.
-Furthermore, it also extends to adaptive mesh refinement with
-subcycling in a straightforward manner, avoiding the complications
-described in \citep{mccorquodalecolella} needed to fill ghost cells
-when using method-of-lines integration.  However, it is restricted to
+The time-integration approach presented here is considerably simpler
+than the SDC method of \citet{castro_sdc}, but allows us to reuse the
+main CTU \MarginPar{any cites to CCSE papers?}  hydrodynamics
+construction and a largely similar ODE integration scheme, making this
+method easier to add to existing simulation codes.  Furthermore, it
+also extends to adaptive mesh refinement with subcycling in a
+straightforward manner, avoiding the complications described in
+\citep{mccorquodalecolella} needed to fill ghost cells when using
+method-of-lines integration.  However, it is restricted to
 second-order accuracy overall.  We term this algorithm the
-``simplified SDC method''.  In this paper we describe the overall method, show its extension to NSE,
-and demonstrate it on several test problems using \castro.  All of the
-code to reproduce the results in this paper are in the \castro\ github
+``simplified-SDC method''.  In this paper we describe the overall
+method and demonstrate it on a test problem and astrophysical
+simulation using \castro.  All of the code to reproduce the results in
+this paper are in the \castro\ github
 repository\footnote{\url{https://github.com/amrex-astro/Castro/}}.
@@
 total energy, $p$ is the pressure, and we carry nuclear species mass
-fractions, $X_k$, and auxiliary composition variables, $\alpha_l$.  The specific total
+fractions, $X_k$.  The specific total
 energy relates to the specific internal energy, $e$, as $E = e + u^2/2$,
@@
 fluxes are computed in the CTU method.
-
 Aside from the $p\partial u/\partial x$ (``pdV'') term in the internal
@@
 in this paper.  We'll discuss this traditional approach next before moving
-on to our new time-coupling method.  Figure~\ref{fig:flowchart} shows a graphical
+on to our new time-coupling method.  Figures~\ref{fig:flowchart_I} and \ref{fig:flowchart_II} show a graphical
 flowchart of the reaction and hydrodynamics advances in the two methods.
@@
 \plotone{sdc_strang_flowchart_I}
-\caption{\label{fig:strang_flowchart} flowchart}
+\caption{\label{fig:strang_flowchart} Flowchart of the operator split,
+  Strang, time integration.  We first react for $\Delta t/2$, then
+  advect for $\Delta t$, and finally react for $\Delta t/2$.  Each
+  process uses the state left behind from the previous process.}
 \end{figure}
@@
 \plotone{sdc_strang_flowchart_II}
-\caption{\label{fig:sdc_flowchart} flowchart}
+\caption{\label{fig:sdc_flowchart} Flowchart of the simplified-SDC
+  approach.  An iteratively-lagged advection term is used to update
+  the system to the new time, with each iteration improving the
+  coupling.}
 \end{figure}
@@
 We integrate Equations (\ref{eq:strang:e}) and (\ref{eq:strang:X})
-using an implicit ODE solver designed for stiff systems of equations,
-based on VODE~\citep{vode}.  We use a modified version of the
-VODE~\citep{vode} integrator---ported to C++ with checks added to the
-timestep rejection logic that ensure that the species mass fractions
-stay between 0 and \MarginPar{we need to be clear that there is
-  different logic for SDC} 1\footnote{This modified version of VODE is
-available in our Microphysics repo:
-\url{https://github.com/starkiller-astro/Microphysics}.}.\MarginPar{discuss
-  tolerances}
+using an implicit ODE solver described below.
 
@@
 
+\subsection{Stiff ODE Integrator}
+
+We use a modified version of the VODE~\citep{vode} integrator,
+designed for stiff ODEs.  Our version has been ported C++, as a
+header-only library enabling it to run on GPUs.  We have also added
+checks to the timestep rejection logic that help prevent VODE from
+wandering too far off the solution\footnote{This modified version of
+VODE is available in our Microphysics repo:
+\url{https://github.com/starkiller-astro/Microphysics}.}  These have
+been shown empirically to improve the performance of VODE.
+
+VODE allows us to specific both relative and absolute tolerances
+for the species and energy during the integration.  They are combined
+into a weight that VODE uses to assess convergence of the step:
+\begin{equation}
+w_i = \epsilon_\mathrm{rel} |y_i| + \epsilon_\mathrm{abs}
+\end{equation}
+for integration variable $y_i$.  We will denote the species tolerances
+as $\rtolspec$ and $\atolspec$ and the energy tolerances as $\rtolener$ and
+$\atolener$.
+
+For Strang, we require that
+an individual mass fraction not change too much over a step:
+\begin{equation}
+\label{eq:vode_species_reject}
+\begin{array}{c}|X_k^{m+1}| < f |X_k^m| \\ |X_k^{m+1}| > \frac{1}{f} |X_k^m| \end{array} ~  \mbox{if}~ |X_k^m| > \eta \atolspec ~\mbox{and}~ |X_k^{m+1}| > \eta \atolspec 
+\end{equation}
+Here, $m$ is the current VODE solution and $m+1$ is the potential
+new-time solution.  We use $f = 4$ for the results here.  If these
+conditions are violated, then VODE will reject the timestep and retry
+with a smaller step.  We also require that the mass fractions are
+contained in $[0, 1]$ to a tolerance of $10^{-2}$.
+
+For SDC, we enforce these same constraints, but for
+Eq.~\ref{eq:vode_species_reject}, we only use the change from $m$ to
+$m+1$ due to reactions, by subtracting off the advective contribution
+over that substep.  We also require that:
+\begin{align}
+\rho^{m+1} &> 0 \\
+(\rho e)^{m+1} &> 0
+\end{align}
+
 \section{Simulations}

### 2021-12-27 simplified_sdc_intro/paper.tex

@@
 \citet{strang_rnaas} and showed that the above formulation gets second
-order convergence.  However, for very strong reactions, when using
-Strang splitting the state can drift significantly off of the smooth
-solution to the coupled reactive hydrodynamics equations, as shown
-graphically in \cite{astronum:2018} using an earlier version of the
-present algorithm.
-
-A variation on Strang splitting called (re-)balanced splitting was
-developed in \citet{speth:2013}.
+order convergence and showed that integrating an energy / temperature
+equation is needed to get second order accuracy (see also
+\citealt{muller:1986} for a discussion on stability).  We note however
+that some astrophysical simulation codes only evolve the species
+equations.  For very strong reactions, when using Strang splitting the
+state can drift significantly off of the smooth solution to the
+coupled reactive hydrodynamics equations, as shown graphically in
+\citet{astronum:2018} (using an earlier version of the present
+algorithm).  A variation on Strang splitting called (re-)balanced
+splitting was developed in \citet{speth:2013}, but we do not explore
+that here. \MarginPar{keep this sentence?}
 
@@
 timestep limiters based on the energy generation or abundance changes
-over a timestep: 
-\begin{align}
-\label{eq:dt:nuce}
-\Delta t &\le f_e\, \min_{i} \left\{\frac{e_{i}}{\dot{S}_{i}}\right\} \\
-%
-\label{eq:dt:nucX}
-\Delta t &\le f_X\, \min_{i} \left\{\min_{k,X_k > \epsilon_X}\frac{{X_k}_{i}}{{\omegadot(X_k)_i}}\right\}
-\end{align}
-where $i$ is the zone index and $f_e$ and $f_X$ are runtime parameters
-used to control the allowed change, and only species for which $X_k > \epsilon_X$ are considered.
+over a timestep.  An energy limiter takes the form of either:
+\begin{equation}
+\label{eq:dt:nuce_I}
+\Delta t \le f_e\, \min_{i} \left\{\frac{e^n_{i}}{|\dot{S}_{i}^n|}\right\} \\
+\end{equation}
+or
+\begin{equation}
+\label{eq:dt:nuce_II}
+\Delta t \le f_e\, \min_{i} \left\{ \frac{e^n_i}{\int \dot{S}_i(t) dt} \right\}
+\end{equation}
+where $i$ is the zone index and $f_e$ is a parameter used to control
+the allowed change (e.g., $f_e = 0.5$ will limit the step such that
+reactions can only increase the internal energy by 50\%).  The two
+formulations differ in that Eq.~\ref{eq:dt:nuce_I} uses the
+instantaneous evaluation of the energy generation rate at the start of
+the timestep while Eq.~\ref{eq:dt:nuce_II} uses the integral of the
+energy generation rate over the entire step (or last Strang half).
 

### 2021-12-28 simplified_sdc_intro/paper.tex
more work

@@
-\documentclass[times,modern]{aastex63}
+%\documentclass[times,modern]{aastex63}
+\documentclass[times,preprint]{aastex63}
 
@@
 density falling below zero during advection, the ODE integration
-failing to converge in the implicit solve, or violation of one of the
-timestep limiters during the step.  This means that equations
-(\ref{eq:dt:nuce}) and (\ref{eq:dt:nucX}) are not reactive, but instead
-guaranteed to be met throughout the simulation, because the step is
-rejected if they are violated.  This contrasts to similar approaches
-used in other codes where conditions \ref{eq:dt:nuce} and
-\ref{eq:dt:nucX} are used to restrict the next timestep size, but the
-update over the current step is kept, which already violated the
-constraints. \MarginPar{would be nice to have a reference here} The retry mechanism in \castro\ works with both the
-Strang and simplified-SDC integration scheme.
+failing to converge in the implicit solve (due to too many steps or
+the internal timestep falling below a minimum), or violation of one of
+the timestep limiters during the step.  This means that equation
+(\ref{eq:dt:nuce_I}) or (\ref{eq:dt:nuce_II}) are not reactive, but
+instead guaranteed to be met throughout the simulation, because the
+step is rejected if they are violated.  This contrasts to similar
+approaches used in other codes where conditions \ref{eq:dt:nuce_I} or
+\ref{eq:dt:nuce_II} are used to restrict the next timestep size, but
+the update over the current step is kept, which already violated the
+constraints. \MarginPar{would be nice to have a reference here} The
+retry mechanism in \castro\ works with both the Strang and
+simplified-SDC integration scheme.  Retrying the step is much more
+stringent that simply adjusting the next step, so we put a cap on the
+number of retries allowed, aborting if we exceed the cap.
 
@@
 We note that the form of these entries is the same as one would arrive at if you start with the rates expressed as
-$\omega_k(\rho, T(\rho, X_j, e), X_j)$ and note that constant $e$ implies that
+$\omega_k(\rho, T(\rho, X_j, e), X_j)$ and recognize that constant $e$ implies that
 \begin{equation}

### 2021-12-28 simplified_sdc_intro/paper.tex
more work

@@
 \newcommand{\castro}{{\sf Castro}}
-\newcommand{\maestro}{{\sf Maestro}}
+\newcommand{\maestroex}{{\sf MAESTROeX}}
 \newcommand{\pynucastro}{{\sf pynucastro}}
@@
 \begin{equation}
+  \label{eq:cons_sources}
   \Hb(\Uc) = \left ( \begin{array}{c}
@@
 benefits from the larger timestep that method can take (when done with
-full corner coupling, the advective CFL condition is unity)
-as compared to a method-of-lines approach (see
-\citealt{ppmunsplit}).  In addition to the standard CFL timestep
-limiter for explicit hydrodynamics, \castro\ can also enforce
-timestep limiters based on the energy generation or abundance changes
-over a timestep.  An energy limiter takes the form of either:
+full corner coupling, the advective CFL condition is unity) as
+compared to a method-of-lines approach (see \citealt{ppmunsplit}).  In
+addition to the standard CFL timestep limiter for explicit
+hydrodynamics, timestep limiters based on
+the energy generation or abundance changes over a timestep, such as
+those introduced in \cite{prometheus}, are often used.  An energy limiter takes the
+form of either:
 \begin{equation}
@@
 \label{eq:dt:nuce_II}
-\Delta t \le f_e\, \min_{i} \left\{ \frac{e^n_i}{\int \dot{S}_i(t) dt} \right\}
+\Delta t \le f_e\, \min_{i} \left\{ \frac{e^n_i}{|\int \dot{S}_i(t) dt|} \right\}
 \end{equation}
@@
 Our simplified-SDC approach is based on the method described in \citet{SDC-old} and
-a similar implementation in the \maestro\ simulation code.
+a similar (but unpublished) implementation in the \maestroex\ simulation code \cite{maestroex}.
 We again start by considering the update
@@
     will denote $\Ic_q$ (this is basically an approximation of
-    $\Rb(q)$).
+    $\Rb(\qb)$).
 
@@
   with a superscript (like $n$ or $n+1$), we'll use a second subscript
-  in parentheses to keep track of the iteration.  A single iteration
+  in parentheses to keep track of the iteration.  A single iteration, $k$,
   starts with $\Uc^n$ and results in the new time-level state for that
@@
 
-    Looking at the form of Eq.~\ref{eq:react_source}, we see that only
-    the species, auxiliary quantities, and energies have source terms.
-    Since we are reacting, we will use the $X_k$ as the primary
-    composition variable and compute $\alpha_l$ as needed from them,
-    so we don't directly integrate the auxiliary quantities.
-    Likewise, the total and internal energies both provide the same
-    information, and integrating both overconstrains the system, so we
-    just integrate the internal energy.  We define the subset of variables
-    that are directly integrated as:
+    Looking at the form of Eq.~\ref{eq:cons_sources}, we see that only
+    the species and energies have source terms.  The total and
+    internal energies both provide the same information, and
+    integrating both overconstrains the system, so we just integrate
+    the internal energy.  We define the subset of variables that are
+    directly integrated as:
     \begin{equation}
@@
     \begin{equation}
-      \rho(t) = \rho^n + \Advss{\rho}^{n+1/2,(k)} \, t
+      \rho(t) = \rho^n + \Advss{\rho}^{n+1/2,(k)} \, (t - t^n)
     \end{equation}
@@
     However, for our primitive variables, which are used in the
-    prediction, we need to construct the required source terms more
+    interface state prediction, we need to construct the required source terms more
     carefully.  We want:
@@
 \end{itemize}
+
+This completes the update of a single iteration.
 
@@
 Here, $m$ is the current VODE solution and $m+1$ is the potential
-new-time solution.  We use $f = 4$ for the results here.  If these
-conditions are violated, then VODE will reject the timestep and retry
-with a smaller step.  We also require that the mass fractions are
-contained in $[0, 1]$ to a tolerance of $10^{-2}$.
+new-time solution.  The parameter $f$ is the factor a mass fraction is
+allowed to change per VODE step.  We use $f = 4$ for the results here.
+The parameter $\eta$ allows us to only use this condition for species
+with mass fractions above a threshold.  We use $\eta = 1$ by default,
+but $\eta = f$ sometimes helps prevent VODE from worrying about
+species that dip below the threshold during integration, since VODE
+uses a norm over all of the integrated variables to compute the total
+error.  If these conditions are violated, then VODE will reject the
+timestep and retry with a smaller step.  We also require that the mass
+fractions are contained in $[0, 1]$ to a tolerance of $10^{-2}$.
 

### 2021-12-28 simplified_sdc_intro/data/initial_model/initial_model.py
fix plotting

@@ -1,0 +1,91 @@
+# take a raw inputs file for aprox21 and convert to the aprox19
+# nuclei.  This means getting rid of Cr56 and Fe56 (by lumping them
+# into Ni56)
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+def find_r_for_rho(r, rho, rho_want):
+    idx = np.where(rho < rho_want)[0][0]
+    return r[idx]
+
+
+#file = "../15m_500_sec.aprox19.hse.6400"
+file = "sub_chandra.M_WD-1.10.M_He-0.050.hse.C.N14.5.000km"
+Lx = 1.024e9
+
+data = np.loadtxt(file)
+
+print(data.shape)
+
+# now manually read to get the variable names
+
+# the first column is position
+
+names = ["r"]
+
+with open(file) as f:
+    for n, line in enumerate(f):
+        if line.startswith("# num"):
+            continue
+
+        if line.startswith("# npts"):
+            continue
+
+        if not line.startswith("#"):
+            break
+
+        names.append(line.split()[-1].strip())
+
+# now make plots
+
+idens = names.index("density")
+itemp = names.index("temperature")
+
+fig = plt.figure()
+ax = fig.add_subplot(211)
+
+l1 = ax.plot(data[:,0], data[:,idens], label="density", lw=2)
+l2 = ax.plot(data[:,0], data[:,itemp], label="temperature", lw=2)
+
+ax.set_xscale("log")
+ax.set_yscale("log")
+
+ax.set_xlabel("r [cm]")
+ax.set_ylabel(r"$\rho~[\rm{g/cm^3}]$, $T~[K]$")
+
+ax.legend(frameon=False)
+
+
+ax = fig.add_subplot(212)
+
+threshold = 0.001
+
+
+for n, var in enumerate(names):
+    print(n, var)
+
+    if var in ["r", "density", "temperature", "pressure"]:
+        continue
+
+    Xmax = data[:,n].max()
+
+    lw = 2
+    if Xmax > threshold:
+        ax.plot(data[:,0], data[:,n], label=var, lw=lw)
+
+
+ax.set_xscale("log")
+ax.set_yscale("log")
+
+ax.set_xlabel("r [cm]")
+ax.set_ylabel("mass fraction")
+
+ax.legend(frameon=False)
+
+fig.set_size_inches((8, 10))
+
+fig.tight_layout()
+
+fig.savefig("initial_model.pdf")
+

### 2021-12-28 simplified_sdc_intro/paper.tex
more updates

@@
 
-\section{Simulations}
+\section{Tests}
 
@@
 \cite{castro_sdc} introduced a test problem for measuring convergence
-of a reacting hydrodynamic algorithm.  We run that same test here with the
-simplified SDC algorithm.
-
-Also explore: PLM vs PPM, 1 vs. 2 vs. 3 iterations
-
-% run on 04-06-21
-% Castro       git describe: 21.04-dirty
-% AMReX        git describe: 21.03-58-g87c81d0ca
-% Microphysics git describe: 21.04-1-g33a4d0fe
+of a reacting hydrodynamic algorithm.  This is based on the acoustic
+pulse problem from \citep{mccorquodalecolella}, and was also used in
+\citep{strang_rnaas}.  A periodic domain with a uniform entropy is
+initialized with a pressure profile with a small perturbation.  This
+drives a low amplitude acoustic wave radially outward from the center.
+A simple reaction network with the triple-alpha and
+$\isotm{C}{12}(\alpha,\gamma)\isotm{O}{16}$ rates releases energy.  By
+running at different resolutions, with a fixed timestep that scales
+with resolution, we can compute the convergence rate.  In order to
+demonstrate second-order convergence, the problem must be smooth (so
+the slope limiters don't kick in severely), which means that the
+energy release cannot be so vigorous as to drive shocks.  test here
+with the simplified SDC algorithm.
+
+% EOS: /home/zingale/development/Microphysics/EOS/helmholtz
+% NETWORK: /home/zingale/development/Microphysics/networks/triple_alpha_plus_cago
+
+% Castro       git describe: 21.12-9-gd7a27ccd5
+% AMReX        git describe: 21.12-64-gfde72340e
+% Microphysics git describe: 21.12-12-g9595a34f
 
 \begin{deluxetable}{lllllll}
-\tablecaption{\label{table:react_convert_strang} Convergence ($L_1$ norm) for the reacting convergence problem using Strang splitting.}
+\tablecaption{\label{table:react_converge_strang} Convergence ($L_1$ norm) for the reacting convergence problem using Strang splitting.}
 \tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
@@
 
-% run on 04-06-21
-% Castro       git describe: 21.04-dirty
-% AMReX        git describe: 21.03-58-g87c81d0ca
-% Microphysics git describe: 21.04-1-g33a4d0fe
-
 \begin{deluxetable}{lllllll}
-\tablecaption{\label{table:react_convert_sdc} Convergence ($L_1$ norm) for the reacting convergence problem using SDC integration (2 iterations).}
+\tablecaption{\label{table:react_converge_sdc} Convergence ($L_1$ norm) for the reacting convergence problem using SDC integration (2 iterations).}
 \tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
@@
 \startdata
- $\rho$                      & $2.795 \times 10^{18}$  & 2.044  & $6.777 \times 10^{17}$  & 2.554  & $1.154 \times 10^{17}$  \\
- $\rho u$                    & $6.798 \times 10^{26}$  & 2.449  & $1.245 \times 10^{26}$  & 2.889  & $1.680 \times 10^{25}$  \\
- $\rho v$                    & $6.798 \times 10^{26}$  & 2.449  & $1.245 \times 10^{26}$  & 2.889  & $1.680 \times 10^{25}$  \\
- $\rho E$                    & $2.452 \times 10^{35}$  & 2.352  & $4.804 \times 10^{34}$  & 2.742  & $7.179 \times 10^{33}$  \\
- $\rho e$                    & $2.262 \times 10^{35}$  & 2.321  & $4.527 \times 10^{34}$  & 2.822  & $6.404 \times 10^{33}$  \\
- $T$                         & $2.236 \times 10^{21}$  & 1.691  & $6.927 \times 10^{20}$  & 2.481  & $1.240 \times 10^{20}$  \\
- $\rho X(\isotm{He}{4})$     & $2.879 \times 10^{18}$  & 2.020  & $7.096 \times 10^{17}$  & 2.529  & $1.229 \times 10^{17}$  \\
- $\rho X(\isotm{C}{12})$     & $1.697 \times 10^{17}$  & 1.950  & $4.393 \times 10^{16}$  & 2.231  & $9.357 \times 10^{15}$  \\
- $\rho X(\isotm{O}{16})$     & $1.686 \times 10^{14}$  & 1.659  & $5.338 \times 10^{13}$  & 1.957  & $1.375 \times 10^{13}$  \\
- $\rho X(\isotm{Fe}{56})$    & $2.795 \times 10^{8}$   & 2.044  & $6.777 \times 10^{7}$   & 2.554  & $1.154 \times 10^{7}$   \\
+ $\rho$                      & $2.784 \times 10^{18}$  & 2.048  & $6.734 \times 10^{17}$  & 2.558  & $1.143 \times 10^{17}$  \\
+ $\rho u$                    & $6.779 \times 10^{26}$  & 2.447  & $1.243 \times 10^{26}$  & 2.895  & $1.671 \times 10^{25}$  \\
+ $\rho v$                    & $6.779 \times 10^{26}$  & 2.447  & $1.243 \times 10^{26}$  & 2.895  & $1.671 \times 10^{25}$  \\
+ $\rho E$                    & $2.450 \times 10^{35}$  & 2.353  & $4.795 \times 10^{34}$  & 2.737  & $7.193 \times 10^{33}$  \\
+ $\rho e$                    & $2.256 \times 10^{35}$  & 2.320  & $4.518 \times 10^{34}$  & 2.817  & $6.414 \times 10^{33}$  \\
+ $T$                         & $2.232 \times 10^{21}$  & 1.695  & $6.893 \times 10^{20}$  & 2.484  & $1.233 \times 10^{20}$  \\
+ $\rho X(\isotm{He}{4})$     & $2.866 \times 10^{18}$  & 2.024  & $7.049 \times 10^{17}$  & 2.534  & $1.217 \times 10^{17}$  \\
+ $\rho X(\isotm{C}{12})$     & $1.695 \times 10^{17}$  & 1.959  & $4.357 \times 10^{16}$  & 2.236  & $9.250 \times 10^{15}$  \\
+ $\rho X(\isotm{O}{16})$     & $1.681 \times 10^{14}$  & 1.662  & $5.313 \times 10^{13}$  & 1.963  & $1.363 \times 10^{13}$  \\
+ $\rho X(\isotm{Fe}{56})$    & $2.784 \times 10^{8}$   & 2.048  & $6.734 \times 10^{7}$   & 2.558  & $1.143 \times 10^{7}$   \\
 \enddata
@@
 
-\subsection{X-ray Burst}
-
-We next consider a flame propagating through an accreted helium layer on a neutron star,
-as described in \cite{harpole:2021}.
+Tables \ref{table:react_converge_strang} and
+\ref{table:react_converge_sdc} show the Strang and simplified-SDC
+convergence respectively.  We see that both methods converge nearly
+second order, and agree well with each other.
 
 \subsection{Sub-Chandra Double Detonation}
+
+We next model a double detonation Type Ia
+supernova.  In this model, a detonation begins in the accreted helium
+layer on a sub-Chandrasekhar mass white dwarf and that detonation can
+either drive a compression wave into the core igniting a second
+detonation in the carbon / oxygen core, or burn through the interface
+at the base of the accreted layer producing an inward propagating
+carbon detonation.  Our goal here is not to explore the feasibility of
+the model or understand whether a detonation propagating through the
+He-C interface is physical, but rather to look at the coupling of the
+hydrodynamics and reactions.  For that reason, we explore only a
+single model and start off with a rather large temperature
+perturbation in the He layer.
 
@@
 \citep{reaclib} and the network is written by
-\pynucastro~\citep{pynucastro}.  The nuclei and links in the network are
-shown in Figure~\ref{fig:subch_network}.  The many reaction pathways in this network
-make it a good test for the SDC algorithm.
-
-The initial model is generated following the procedure described in \citep{subchandra}.
-For the present work, we are only concerned with the behavior of the time-integration strategy
-and not on the astrophysical implications of the physical setup we explore.
+\pynucastro~\citep{pynucastro} directly in the C++ format that our
+code requires.  The nuclei and links in the network are shown in
+Figure~\ref{fig:subch_network}.  This moderate-sized network with many
+reaction pathways (forward and reverse) make it a good test for the
+coupling between hydro and reactions.  An overview of the network is
+shown in Figure~\ref{fig:subch_network}.
 
@@
 
+The initial model is generated following the procedure described in
+\citep{subchandra}.  We use a $1.1~M_\odot$ carbon white dwarf with a
+$0.05~M_\odot$ He layer.  We seed the He layer with 1\% (by mass) of
+\isot{N}{14}.  The white dwarf is isothermal and the temperature ramps
+up at the base of the He layer and is then isentropic until the
+surface.  Figure~\ref{fig:subch_initial_model} shows the structure of
+the initial model.  We use 2d axisymmetric coordinates with a domain
+size of $1.024\times 10^9~\mathrm{cm} \times 2.048\times
+10^9~\mathrm{cm}$, a base grid of $256 \times 512$ zones and two
+levels of refinement (both jumping by $2\times$), giving a fine-grid
+resolution of $10~\mathrm{km}$.  Gravity is modeled as a monopole.
+
+\begin{figure}[t]
+\centering
+\plotone{initial_model}
+\caption{\label{fig:subch_initial_model} Initial model for the double detonation test problem.}
+\end{figure}
+
+An initial perturbation is placed on the symmetry axis of the form:
+
 We do a suite of runs with both Strang and the simplified-SDC
@@
 = \rtolener = 10^{-6}$.  We also run two sets with tighter tolerances.
-The runs labeled ``mixed\_tols'' have $\atolspec = 10^{-8}$ and the runs
-labeled ``tol1.e-6'' have $\atolspec = \rtolspec = 10^{-6}$, with the other
-tolerances unchanged.
+The runs labeled ``mixed\_tols'' have $\atolspec = 10^{-8}$ and the
+runs labeled ``tol1.e-6'' have $\atolspec = \rtolspec = 10^{-6}$, with
+the other tolerances unchanged.  For the step rejection part of VODE,
+we use $f = 4$ and $\eta = 4$.
 
@@
 1-d slices along x=0?
+
+We have not focused on performance or load-balancing the reactions in
+the tests that we ran, but these limited runs show that the
+simplified-SDC method runs as fast (or faster) than the Strang method
+for the same tolerances and CFL, despite having to do the advection
+twice.  We would expect similar behavior for any simulations that are
+dominated by the cost of reactions.
 

### 2021-12-29 simplified_sdc_intro/paper.tex
more work

@@
 resolution of $10~\mathrm{km}$.  Gravity is modeled as a monopole.
+Our choice of initial perturbation has the detonation propagate inward
+into the carbon white dwarf.  We run the simulation for 0.1 s---this
+is enough time for the carbon detonation to be well developed.
 
@@
 
-Overview plots of CFL = 0.8
+Figure \ref{fig:subch_strang_cfl08_I} shows the temperature, mean
+molecular weight, and nuclear energy generation rate for the strang
+CFL = 0.8 simulation ({\tt strang\_subch2\_cfl0.8}).  We see that the
+mean molecular weight appears mottled behind the inward propagating
+detonation in the white dwarf.  Comparing to the energy generation
+rate, we see that region is characterized by large energy releases of
+alternating signs checkerboarding thoughout the region.  This is
+characteristic of the nucleosynthesis struggling to come into
+equilibrium at high temperatures when the reverse rates can be
+important.  In Figure~\ref{fig:subch_strang_cfl08_II} we look at some
+important mass fractions for the same simulation.  Again we see that
+the abundances are not smooth behind the inward detonation.
+
+\begin{figure}
+\centering
+\plotone{strang_subch2_cfl0.8_state}
+\caption{\label{fig:subch_strang_cfl08_I}}
+\end{figure}
+
+\begin{figure}
+\centering
+\plotone{strang_subch2_cfl0.8_mass_fractions}
+\caption{\label{fig:subch_strang_cfl08_II}}
+\end{figure}
+
+We next look at the simplified-SDC version of this same case, {\tt
+  sdc\_subch2\_cfl0.8}.  Figure~\ref{fig:subch_sdc_cfl08_I} shows the
+temperature, mean molecular weight, and energy generation rate.
+Compared to Figure~\ref{fig:subch_strang_cfl08_I}, we see that the
+composition and energy generation rates are smooth, without any of the
+checkerboarding that plagued the Strang simulation.  The compositions
+plot (figure~\ref{fig:subch_sdc_cfl08_II}) shows the same behavior.
+This suggests that the simplified-SDC algorithm has an easier time
+attaining the correct equilibrium in these conditions.
+
+\begin{figure}
+\centering
+\plotone{sdc_subch2_cfl0.8_state}
+\caption{\label{fig:subch_sdc_cfl08_I}}
+\end{figure}
+
+\begin{figure}
+\centering
+\plotone{sdc_subch2_cfl0.8_mass_fractions}
+\caption{\label{fig:subch_sdc_cfl08_II}}
+\end{figure}
 
@@
 
-Mass of different species (as a plot)
+For supernova calculations, we often want to know how much mass of
+\isot{Ni}{56} was produced.  Table \ref{table:masses} shows the total
+mass of \isot{He}{4}, \isot{Si}{28}, and \isot{Ni}{56} for each of the
+simulations.
 

### 2021-12-29 simplified_sdc_intro/paper.tex
more work

@@
-%\documentclass[times,modern]{aastex63}
+\documentclass[times,modern]{aastex63}
 \documentclass[times,preprint]{aastex63}
@@
 
+For both Strang and simplified-SDC, we will use a numerical Jacobian.
+VODE computes it internally for Strang and we use the method from
+Appendix \ref{app:jacobian} for the simplified-SDC method.  VODE
+employs Jacobian-caching, so the Jacobian does not need to be
+re-evaluated each time it is used.
+
 \section{Tests}
@@
 
-We use a network with 28 nuclei and 96 rates that captures He burning
+We use a network with 28 nuclei and 96 rates that captures He burning \MarginPar{should I do a few aprox13 runs for comparison?}
 with links for $(\alpha,p)(p,\gamma)$ reactions and the rates
@@
 coupling between hydro and reactions.  An overview of the network is
-shown in Figure~\ref{fig:subch_network}.
+shown in Figure~\ref{fig:subch_network}.  This network is available 
+as {\tt subch2} in \citep{microphysics}.
 
@@
 
-Nonaka plots
+We now look at the evolution of the state while burning by storing the
+thermodynamic state internal to VODE each substep as it evaluates the
+righthand side of the ODE system it is integrating.  We pick a point
+on the symmetry axis a few zones behind the inward moving shock.
+Figure~\ref{fig:nonaka_plot} shows the evolution of \isot{Ni}{56} over
+a single coarse-grid timestep (4 fine-grid timesteps).
 
@@
 1-d slices along x=0?
+
+\section{Summary}
+
+We presented a simplified spectral deferred corrections scheme for coupling
+hydrodynamics and reactions.  We demonstrated that for a moderate-sized reaction
+network, the simplified-SDC method
 
@@
 twice.  We would expect similar behavior for any simulations that are
-dominated by the cost of reactions.
-
-\section{Summary}
-
-We presented a simplified spectral deferred corrections scheme for coupling
-hydrodynamics and reactions.
-
-Future work is to extend this methodology to MHD.
+dominated by the cost of reactions.  If the results above a robust across
+different problems and networks, then this suggests that we can run 
+the SDC simulation with larger CFL and/or less strict reaction tolerances,
+which would translate into a large performance boost.
+
+This new time-integration method is freely available in \castro, and
+we will continue to explore it on other problems.  We are particularly
+interested in massive star evolution with this method, where the
+vigorous shell burning leading up to core-collapse would be an ideal
+application for the simplified-SDC method.  In a follow-on paper, we
+will show how to couple this new integration method to a table for
+evaluating an NSE distribution in the iron core of a massive star.
+
+Future work is to extend this methodology to MHD.  This is very
+straightforward since the simplified-SDC method operates solely on the
+thermodynamic state.  We will also explore how to couple radiation
+hydrodynamics, where we need to do an implicit update for the
+radiation energy.
 

### 2021-12-29 simplified_sdc_intro/paper.tex
more updates

@@
-\documentclass[times,modern]{aastex63}
+%\documentclass[times,modern]{aastex63}
 \documentclass[times,preprint]{aastex63}
@@
 righthand side of the ODE system it is integrating.  We pick a point
-on the symmetry axis a few zones behind the inward moving shock.
+on the symmetry axis a few zones behind the inward moving shock and
+ran both a Strang and simplified-SDC simulation for a single
+coarse-grid timestep.  To ensure that we start off the same, both
+simulations were started from a checkpoint file from the
+simplified-SDC simulation a few steps before 0.1~s.
 Figure~\ref{fig:nonaka_plot} shows the evolution of \isot{Ni}{56} over
-a single coarse-grid timestep (4 fine-grid timesteps).
+a single coarse-grid timestep (4 fine-grid timesteps).  We see that
+the Strang solution is much more discontinuous, as the reactions and
+advection push the solution in different ways, because the processes
+are completely independent of one another.  If an equilibrium state is
+important to the nucleosynthesis, the large deviations in the Strang
+solution mean that we are evaluating the reaction network with a state
+out of equilibrium for much of the integration time. \MarginPar{need to double check the data---why are there points in the vertical part of Strang?}
+
+\begin{figure}
+\plotone{sdc_compare}
+\caption{\label{fig:nonaka_plot} Evolution of $X(\isotm{Ni}{56})$ over
+  4 fine-grid timesteps in a single zone behind the inward propagating
+  shock.  The Strang points (blue) show the first $\Delta t/2$ burn
+  bring the abundance away from equilibrium, then the advection
+  overcorrecting it, and finally the second $\Delta t/2$ bringing it
+  back toward the equilibrium.  The two simplified-SDC iterations
+  appear much smoother.  For the SDC runs, there are a few points
+  evaluated beyond the stop time---VODE can use data from outside the
+  integration range to interpolate to the final solution.}
+\end{figure}
 

### 2021-12-29 simplified_sdc_intro/paper.tex
more figures

@@
 
-Zoom panels showing different runs
-
-For supernova calculations, we often want to know how much mass of
-\isot{Ni}{56} was produced.  Table \ref{table:masses} shows the total
+We next look at how the structure seen in
+Figures~\ref{fig:subch_strang_cfl08_I} and \ref{fig:subch_sdc_cfl08_I}
+vary with CFL number and network tolerance.  We focus just on the
+region behind the inward propagating shock.
+Figures~\ref{fig:strang_enuc_summary} and
+\ref{fig:strang_abar_summary} show the nuclear energy generation rate
+and mean molecular weight for all of the Strang runs.  We see that
+changing the CFL number does little to affect the state behind the
+shock---it continues to take on the mottled appearance characteristic
+of not reaching a proper equilibrium.  For the {\tt mixedtol} case, we
+see that the state is greatly improved.  We were unable to run the
+{\tt tol1.e-6} case for Strang, as the code aborted with too many
+subcycles.  While reducing the tolerances does help, it greatly
+increases the cost of the simulation.
+
+\begin{figure}
+\plotone{strang_enuc_summary_plot.pdf}
+\caption{\label{fig:strang_enuc_summary} Comparison of Strang runs showing
+the nuclear energy generation rate behind the inward propagating
+shock.}
+\end{figure}
+
+\begin{figure}
+\plotone{strang_abar_summary_plot.pdf}
+\caption{\label{fig:strang_abar_summary} Comparison of Strang runs showing
+the mean molecular weight behind the inward propagating
+shock.}
+\end{figure}
+
+The SDC comparisons are shown in Figures~\ref{fig:sdc_enuc_summary}
+and \ref{fig:sdc_abar_summary}.  We see that the state behind the
+shock looks great regardless of the CFL or integrator tolerances.
+
+\begin{figure}
+\plotone{sdc_enuc_summary_plot.pdf}
+\caption{\label{fig:sdc_enuc_summary} Comparison of simplified-SDC runs showing
+the nuclear energy generation rate behind the inward propagating
+shock.}
+\end{figure}
+
+\begin{figure}
+\plotone{sdc_abar_summary_plot.pdf}
+\caption{\label{fig:sdc_abar_summary} Comparison of simplified-SDC runs showing
+the mean molecular weight behind the inward propagating
+shock.}
+\end{figure}
+
+To quantify this effect, we next looks at the total mass of some key species.
+This is often a desired diagnostic for supernova calculations, since 
+\isot{Ni}{56} powers the lightcurve.  Table \ref{table:masses} shows the total
 mass of \isot{He}{4}, \isot{Si}{28}, and \isot{Ni}{56} for each of the

### 2021-12-29 simplified_sdc_intro/paper.tex
finish first cut at the sub ch

@@
 
-To quantify this effect, we next looks at the total mass of some key species.
-This is often a desired diagnostic for supernova calculations, since 
-\isot{Ni}{56} powers the lightcurve.  Table \ref{table:masses} shows the total
-mass of \isot{He}{4}, \isot{Si}{28}, and \isot{Ni}{56} for each of the
-simulations.
+To quantify this effect, we next looks at the total mass of some key
+species.  This is often a desired diagnostic for supernova
+calculations, since \isot{Ni}{56} powers the lightcurve.  Table
+\ref{table:masses} shows the total mass of \isot{He}{4},
+\isot{Si}{28}, and \isot{Ni}{56} for each of the simulations.  For the
+Strang runs, there are large variations between the different
+simulations while the SDC simulations all agree to within a few
+percent.  Further, the Strang run with the tighter tolerances is in
+good agreement with the masses found by the SDC runs.  This supports
+the ideas described above that all of the SDC runs agree while only
+the Strang run with the tighter tolerances is reliable.
+
+\begin{deluxetable}{llll}
+\tablecaption{\label{table:masses} Species masses.}
+\tablehead{\colhead{simulation} & \colhead{$M_{\isotm{He}{4}}$} & \colhead{$M_{\isotm{Si}{28}}$} & \colhead{$M_{\isotm{Ni}{56}}$}}
+\startdata
+\multicolumn{4}{c}{Strang runs} \\
+\hline
+{\tt strang\_subch2\_cfl0.2}           & $6.10\times 10^{31}$ & $6.72\times 10^{31}$ & $5.29\times 10^{31}$ \\
+{\tt strang\_subch2\_cfl0.4}           & $5.20\times 10^{31}$ & $6.70\times 10^{31}$ & $6.02\times 10^{31}$ \\
+{\tt strang\_subch2\_cfl0.4\_mixedtol} & $3.05\times 10^{31}$ & $7.29\times 10^{31}$ & $7.80\times 10^{31}$ \\
+{\tt strang\_subch2\_cfl0.8}           & $4.59\times 10^{31}$ & $6.88\times 10^{31}$ & $6.27\times 10^{31}$ \\
+\hline
+\multicolumn{4}{c}{SDC runs} \\
+\hline
+{\tt sdc\_subch2\_cfl0.2}           & $3.35\times 10^{31}$  & $7.26\times 10^{31}$  & $8.27\times 10^{31}$ \\
+{\tt sdc\_subch2\_cfl0.4}           & $3.30\times 10^{31}$  & $7.27\times 10^{31}$  & $8.13\times 10^{31}$ \\
+{\tt sdc\_subch2\_cfl0.4\_mixedtol} & $3.09\times 10^{31}$  & $7.36\times 10^{31}$  & $8.08\times 10^{31}$ \\
+{\tt sdc\_subch2\_cfl0.4\_tol1.e-6} & $3.25\times 10^{31}$  & $7.27\times 10^{31}$  & $8.25\times 10^{31}$ \\
+{\tt sdc\_subch2\_cfl0.8}           & $3.18\times 10^{31}$  & $7.34\times 10^{31}$  & $7.83\times 10^{31}$ \\
+\enddata
+\end{deluxetable}
 

### 2022-01-25 xrb3/paper.tex
more reorg

@@
-%\documentclass[12pt, preprint]{aastex}
-\documentclass[preprint,times,tighten]{aastex63}
+\documentclass[preprint,times,tighten]{aastex631}
 
@@
 %======================================================================
-\title{Dynamics of Laterally Propagating Flames in X-ray Bursts. III. Mixed H/He Flames}
+\title{Dynamics of Laterally Propagating Flames in X-ray Bursts. III. Three-Dimensional Flames}
 
@@
 
+We use the \castro\ \citep{castro,castro_joss} simulation code, following
+the same numerical approach described in \citet{eiden:2020,harpole:2021}.
+
 %======================================================================
@@
 
-\acknowledgements \castro\ is open-source and freely available at
+\begin{acknowledgements}
+\castro\ is open-source and freely available at
 \url{http://github.com/AMReX-Astro/Castro}.  The problem setups used
@@
 Services.
+\end{acknowledgements}
 

### 2022-01-26 xrb3/paper.tex
update acks
more drafting

@@
 \begin{abstract}
-We continue to investigate laterally propagating flames in XRBs...
+We continue to investigate laterally propagating flames in XRBs,
+comparing the dynamics between two- and three-dimensions.
 \end{abstract}
@@
 
-%======================================================================
-% Numerical Approach
-%======================================================================
-\section{Numerical Approach}\label{Sec:numerics}
+X-ray bursts result from thermonuclear burning of an accreted H/He or
+He layer on a neutron star \citep{galloway:2017}.  Observations of
+brightness oscillations in the rise of the lightcurve suggest that the
+burning begins localized and then spreads across the neutron star.
+One-dimensional simulations of X-ray bursts reproduce the lightcurve
+and recurrence time \citep{woosley-xrb} and can use large reaction
+networks to explore rp-process nucleosynthesis.
+
+To understand how the burning front propagates across the surface, we
+need to do multidimensional simulations.  The work of
+\citet{cavecchi:2013,art-2015-cavecchi-etal,art-2016-cavecchi-etal}
+
+The three-dimensional study by \citet{Cavecchi2019} showed...
+
+We have been performing simulations of flame spreading on a neutron
+star resolving the flame structure using realistic reaction networks.
+Our two-dimensional studies \citep{eiden:2020,harpole:2021} demonstrated our open-source framework for modeling X-ray bursts using
 
@@
 
+In \citet{harpole:2021}, we explored several different rotation rates
+and initial model thermal structures.  We found that a model with a
+neutron star crust temperature of $3\times 10^8~\mathrm{K}$ resulted
+in a flame that accelerated quickly as it spread across the neutron
+star surface.  We adopt this model for the 3D simulation presented
+here, since it requires the least amount of simulation time to yield a
+developed flame.  Moving to 3D is very expensive, so to further reduce
+the computational expense, we use a slightly lower resolution of 25 cm
+instead of 20 cm from \citet{harpole:2021}.  We pick a rotation rate
+of 1000 Hz.
+
+Our 3D simulation used a base grid of $1024^2 \times 128$ zones with 2
+levels of refinement, the first jumping by a factor of 4 and the next
+by a factor of 2.  The domain has a size $(2.048\times
+10^5~\mathrm{cm})^2 \times 2.56\times 10^4~\mathrm{cm}$, giving a fine
+grid resolution of $25~\mathrm{cm}$.  The initial model and other
+simulation parameters are identical to those used in
+\citet{harpole:2021}.  We used static mesh refinement, to achieve
+better load balancing, fully refining the atmosphere below a height of
+$3600~\mathrm{cm}$.  These simulations used the 7-isotope He-burning
+network used in \citet{eiden:2020}.  The simulations were run on the
+OLCF Summit supercomputer, on 342 to 1366 nodes, with 6 NVIDIA V100
+GPUs per node.  The entire computation was offloaded to GPUs following
+the strategy described in \citet{castro_gpu}.
+
+Things to show:
+
+Slice plot compared to 2-d simulation
+
+Phase plot comparisons
+
+Volume renderings
+
+\section{Summary}
+
+We have extended our simulation framework to three-dimensions, and
+explored the differences between a two-dimensional axisymmetric
+simulation and a fully three-dimensional simulation.
+
+Future work will explore mixed H/He bursts, using this same framework.
+
 \begin{acknowledgements}
 \castro\ is open-source and freely available at
-\url{http://github.com/AMReX-Astro/Castro}.  The problem setups used
-here are available in the git repo as {\tt flame} and {\tt
-  flame\_wave}.  The work at Stony Brook was supported by DOE/Office
-of Nuclear Physics grant DE-FG02-87ER40317.  This research used
-resources of the National Energy Research Scientific Computing Center,
-a DOE Office of Science User Facility supported by the Office of
-Science of the U.~S.\ Department of Energy under Contract
-No.\ DE-AC02-05CH11231.  This research used resources of the Oak Ridge
-Leadership Computing Facility at the Oak Ridge National Laboratory,
-which is supported by the Office of Science of the U.S. Department of
-Energy under Contract No. DE-AC05-00OR22725, awarded through the DOE
-INCITE program.  We thank NVIDIA Corporation for the donation of a
-Titan X and Titan V GPU through their academic grant program.  This
-research has made use of NASA's Astrophysics Data System Bibliographic
-Services.
+\url{http://github.com/AMReX-Astro/Castro}.  The problem setup used
+here is available in the git repo as {\tt flame\_wave}.  The work at
+Stony Brook was supported by DOE/Office of Nuclear Physics grant
+DE-FG02-87ER40317.  This research used resources of the National
+Energy Research Scientific Computing Center, a DOE Office of Science
+User Facility supported by the Office of Science of the
+U.~S.\ Department of Energy under Contract No.\ DE-AC02-05CH11231.
+This research used resources of the Oak Ridge Leadership Computing
+Facility at the Oak Ridge National Laboratory, which is supported by
+the Office of Science of the U.S. Department of Energy under Contract
+No. DE-AC05-00OR22725, awarded through the DOE INCITE program.  We
+thank NVIDIA Corporation for the donation of a Titan X and Titan V GPU
+through their academic grant program.  This research has made use of
+NASA's Astrophysics Data System Bibliographic Services.
 \end{acknowledgements}

### 2022-02-05 xrb3/plots/volume_sum/mass_plot.py
add a plot of mass of species

@@ -1,0 +1,54 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+
+# Use LaTeX for rendering
+mpl.rcParams["text.usetex"] = True
+# load the xfrac package
+mpl.rcParams["text.latex.preamble"] += r'\newcommand{\isotm}[2]{{}^{#2}\mathrm{#1}}'
+mpl.rcParams['mathtext.fontset'] = 'cm'
+mpl.rcParams['mathtext.rm'] = 'serif'
+
+
+fig, ax = plt.subplots()
+
+data = np.loadtxt("species_masses_3d.txt")
+total_mass = np.sum(data, axis=1)
+
+print(f"total mass 3d = {total_mass[0]}")
+print(f"total He mass 3d = {data[0, 1]}, {data[-1, 1]}")
+
+mass_He0 = data[0, 1]
+
+ax.plot(data[:,0], data[:,1] / mass_He0, label=r"$\isotm{He}{4}$", color="C0")
+ax.plot(data[:,0], data[:,2] / mass_He0, label=r"$\isotm{C}{12}$", color="C1")
+ax.plot(data[:,0], data[:,3] / mass_He0, label=r"$\isotm{O}{16}$", color="C2")
+ax.plot(data[:,0], data[:,4] / mass_He0, label=r"$\isotm{Ne}{20}$", color="C3")
+ax.plot(data[:,0], data[:,5] / mass_He0, label=r"$\isotm{Mg}{24}$", color="C4")
+ax.plot(data[:,0], data[:,6] / mass_He0, label=r"$\isotm{Si}{28}$", color="C5")
+ax.plot(data[:,0], data[:,7] / mass_He0, label=r"$\isotm{Ni}{56}$", color="C6")
+
+
+data2d = np.loadtxt("species_masses_2d.txt")
+mass_He0 = data2d[0, 1]
+
+
+ax.plot(data2d[:,0], data2d[:,1] / mass_He0, linestyle="--", color="C0")
+ax.plot(data2d[:,0], data2d[:,2] / mass_He0, linestyle="--", color="C1")
+ax.plot(data2d[:,0], data2d[:,3] / mass_He0, linestyle="--", color="C2")
+ax.plot(data2d[:,0], data2d[:,4] / mass_He0, linestyle="--", color="C3")
+ax.plot(data2d[:,0], data2d[:,5] / mass_He0, linestyle="--", color="C4")
[…]

### 2022-02-05 xrb3/paper.tex
more work on the paper

@@
 %======================================================================
-\title{Dynamics of Laterally Propagating Flames in X-ray Bursts. III. Three-Dimensional Flames}
+\title{Dynamics of Laterally Propagating Flames in X-ray Bursts. III. Comparing Hotspot Evolution in Two and Three Dimensions}
 
@@
 GPUs per node.  The entire computation was offloaded to GPUs following
-the strategy described in \citet{castro_gpu}.
+the strategy described in \citet{castro_gpu}.  Overall, about 370,000
+node-hours were used for the calculation.  The full state of the
+calculation was output only every 0.005 s, as each output file is 2.7 TB
+in size (for single precision data).  We output a few fields more
+frequently for visualization.  We ran a 2D simulation with the same
+resolution and refinement strategy to compare with here.
 
@@
 Volume renderings
+
+Because the 3D flame does not stay perfectly circular, it is difficult
+to define the flame speed using the same method employed in
+\citet{eiden:2020}.  Instead, we will look at the mass of the ash
+material to assess how quickly the burning is taking place.
+Figure~\ref{fig:mass_plot} shows the mass of the different species as
+a function of time, for the 2D and 3D calculation.  The axisymmetric
+2D calculation has slightly less volume than the 3D domain, since
+rotating about the symmetry axis produces a cylindrical domain
+inscribed in the 3D domain.  To compensate for this, we scale the
+masses by the initial \isot{He}{4} mass.  This plot shows that the
+masses of the heavy species , especially \isot{C}{12} and
+\isot{Si}{28} grow quickly with time, but the 2D simulation appears to
+burn faster than the 3D simulation.  This is because the assumption of
+axisymmetry there does not allow for the complex structure we see in
+the evolution of composition in the 3D flame.  This suggests that the
+3D flame speed is slower than the 2D flame speed.
+
+\begin{figure}[t]
+\centering
+\plotone{mass_plot}
+\caption{\label{fig:mass_plot} Mass of species scaled to initial He mass as a function of time.  The 3D simulation is shown as the solid lines and the 2D simulation is shown as the dashed lines.}
+\end{figure}
 

### 2022-02-05 xrb3/plots/volume_sum/mass_plot.py

@@ -50,5 +50,5 @@ ax.set_ylabel("mass / initial He mass")
 fig.set_size_inches(7.0, 7.0)
 
 fig.tight_layout()
-fig.savefig("mass_plot.png")
+fig.savefig("mass_plot.pdf")
 

### 2022-02-06 xrb3/paper.tex
more paper writing

@@
 We continue to investigate laterally propagating flames in XRBs,
-comparing the dynamics between two- and three-dimensions.
+comparing the early evolution and structure of a burning hotspot
+between two- and three-dimensions.
 \end{abstract}
@@
 here, since it requires the least amount of simulation time to yield a
-developed flame.  Moving to 3D is very expensive, so to further reduce
-the computational expense, we use a slightly lower resolution of 25 cm
-instead of 20 cm from \citet{harpole:2021}.  We pick a rotation rate
-of 1000 Hz.
+developed flame.
+
+Moving to 3D is very expensive, so to further reduce the computational
+expense, we use a slightly lower resolution of 25 cm instead of 20 cm
+from \citet{harpole:2021}.  We pick a rotation rate of 1000 Hz.
+Because the hotspot is placed in the center of the domain, the size of
+the domain also becomes a constraint, and as a result, the evolution
+time we can reach will become limited by the time it takes the burning
+to approach the edge of the domain.  As a result, we focus here on the
+early evolution.
 
@@
 explored the differences between a two-dimensional axisymmetric
-simulation and a fully three-dimensional simulation.
+simulation and a fully three-dimensional hydrodynamical simulation.
+
+The main limitation of this study is that computational resources
+limited the size of the domain and resolution we could use.  To
+improve simulations for the future, there are several avenues to
+pursue.  First, the main expense of the current simulations is the
+reaction network.  We need to work more on optimization of the
+network.  Second, while additional resolution will be expensive, we
+can go higher order, as demonstrated with the 4th order accurate
+coupling of hydro, diffusion, and reactions in \cite{castro-sdc}.  The
+main outstanding issue with applying that work to the present problem
+is the multilevel time integration.
 

### 2022-03-26 simplified_sdc_intro/paper.tex
rename plots
fix conserved state
add melissa
add some net details

@@
              Stony Brook, NY 11794-3800}
+
+\author{Melissa Rasmussen}
 
@@
 \centering
-\plotone{strang_subch2_cfl0.8_state}
+\plotone{strang_subch2_cfl0_8_state}
 \caption{\label{fig:subch_strang_cfl08_I}}
@@
 \centering
-\plotone{strang_subch2_cfl0.8_mass_fractions}
+\plotone{strang_subch2_cfl0_8_mass_fractions}
 \caption{\label{fig:subch_strang_cfl08_II}}
@@
 \centering
-\plotone{sdc_subch2_cfl0.8_state}
+\plotone{sdc_subch2_cfl0_8_state}
 \caption{\label{fig:subch_sdc_cfl08_I}}
@@
 \centering
-\plotone{sdc_subch2_cfl0.8_mass_fractions}
+\plotone{sdc_subch2_cfl0_8_mass_fractions}
 \caption{\label{fig:subch_sdc_cfl08_II}}
@@
 
+\section{Reaction Network}
+
+Our reaction network includes the nuclei: $\mathrm{n}$,
+${}^{1}\mathrm{H}$, ${}^{4}\mathrm{He}$, ${}^{12}\mathrm{C}$,
+${}^{14}\mathrm{C}$, ${}^{13}\mathrm{N}$, ${}^{14}\mathrm{N}$,
+${}^{16}\mathrm{O}$, ${}^{18}\mathrm{O}$, ${}^{18}\mathrm{F}$,
+${}^{20}\mathrm{Ne}$, ${}^{21}\mathrm{Ne}$, ${}^{22}\mathrm{Ne}$,
+${}^{22}\mathrm{Na}$, ${}^{23}\mathrm{Na}$, ${}^{23}\mathrm{Mg}$,
+${}^{24}\mathrm{Mg}$, ${}^{27}\mathrm{Al}$, ${}^{27}\mathrm{Si}$,
+${}^{28}\mathrm{Si}$, ${}^{31}\mathrm{P}$, ${}^{31}\mathrm{S}$,
+${}^{32}\mathrm{S}$, ${}^{35}\mathrm{Cl}$, ${}^{36}\mathrm{Ar}$,
+${}^{39}\mathrm{K}$, ${}^{40}\mathrm{Ca}$, ${}^{43}\mathrm{Sc}$,
+${}^{44}\mathrm{Ti}$, ${}^{47}\mathrm{V}$, ${}^{48}\mathrm{Cr}$,
+${}^{51}\mathrm{Mn}$, ${}^{52}\mathrm{Fe}$, ${}^{55}\mathrm{Co}$,
+${}^{56}\mathrm{Ni}$.
+
+The rates linking them are:
+
+\startlongtable
+\begin{deluxetable}{ll}
+\tablecaption{\label{table:react_converge_strang} Convergence ($L_1$ norm) for the reacting convergence problem using Strang splitting.}
+\tablehead{\colhead{forward reaction} & \colhead{reverse reaction}}
+\startdata
+${}^{4}\mathrm{He} + {}^{4}\mathrm{He} + {}^{4}\mathrm{He} \rightarrow {}^{12}\mathrm{C}$ & 
+  ${}^{12}\mathrm{C} \rightarrow {}^{4}\mathrm{He} + {}^{4}\mathrm{He} + {}^{4}\mathrm{He}$ \\
+${}^{12}\mathrm{C} + {}^{1}\mathrm{H} \rightarrow {}^{13}\mathrm{N}$ & 
+  ${}^{13}\mathrm{N} \rightarrow {}^{1}\mathrm{H} + {}^{12}\mathrm{C}$ \\
+${}^{12}\mathrm{C} + {}^{4}\mathrm{He} \rightarrow {}^{16}\mathrm{O}$ & 
+  ${}^{16}\mathrm{O} \rightarrow {}^{4}\mathrm{He} + {}^{12}\mathrm{C}$ \\
+${}^{12}\mathrm{C} + {}^{12}\mathrm{C} \rightarrow {}^{4}\mathrm{He} + {}^{20}\mathrm{Ne}$ & 
+  ${}^{20}\mathrm{Ne} + {}^{4}\mathrm{He} \rightarrow {}^{12}\mathrm{C} + {}^{12}\mathrm{C}$ \\
+${}^{12}\mathrm{C} + {}^{12}\mathrm{C} \rightarrow {}^{1}\mathrm{H} + {}^{23}\mathrm{Na}$ & 
+  ${}^{23}\mathrm{Na} + {}^{1}\mathrm{H} \rightarrow {}^{12}\mathrm{C} + {}^{12}\mathrm{C}$ \\
+${}^{14}\mathrm{C} + {}^{4}\mathrm{He} \rightarrow {}^{18}\mathrm{O}$ & 
+  ${}^{18}\mathrm{O} \rightarrow {}^{4}\mathrm{He} + {}^{14}\mathrm{C}$ \\
+${}^{14}\mathrm{C} \rightarrow {}^{14}\mathrm{N}$ & 
+                                         \\
+${}^{13}\mathrm{N} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{16}\mathrm{O}$ & 
+  ${}^{16}\mathrm{O} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{13}\mathrm{N}$ \\
+${}^{14}\mathrm{N} + {}^{4}\mathrm{He} \rightarrow {}^{18}\mathrm{F}$ & 
+  ${}^{18}\mathrm{F} \rightarrow {}^{4}\mathrm{He} + {}^{14}\mathrm{N}$ \\
+${}^{16}\mathrm{O} + {}^{4}\mathrm{He} \rightarrow {}^{20}\mathrm{Ne}$ & 
+  ${}^{20}\mathrm{Ne} \rightarrow {}^{4}\mathrm{He} + {}^{16}\mathrm{O}$ \\
+${}^{16}\mathrm{O} + {}^{12}\mathrm{C} \rightarrow {}^{4}\mathrm{He} + {}^{24}\mathrm{Mg}$ & 
+  ${}^{24}\mathrm{Mg} + {}^{4}\mathrm{He} \rightarrow {}^{12}\mathrm{C} + {}^{16}\mathrm{O}$ \\
+${}^{16}\mathrm{O} + {}^{12}\mathrm{C} \rightarrow {}^{1}\mathrm{H} + {}^{27}\mathrm{Al}$ & 
+  ${}^{27}\mathrm{Al} + {}^{1}\mathrm{H} \rightarrow {}^{12}\mathrm{C} + {}^{16}\mathrm{O}$ \\
+${}^{16}\mathrm{O} + {}^{16}\mathrm{O} \rightarrow {}^{4}\mathrm{He} + {}^{28}\mathrm{Si}$ & 
+  ${}^{28}\mathrm{Si} + {}^{4}\mathrm{He} \rightarrow {}^{16}\mathrm{O} + {}^{16}\mathrm{O}$ \\
+${}^{16}\mathrm{O} + {}^{16}\mathrm{O} \rightarrow {}^{1}\mathrm{H} + {}^{31}\mathrm{P}$ & 
+  ${}^{31}\mathrm{P} + {}^{1}\mathrm{H} \rightarrow {}^{16}\mathrm{O} + {}^{16}\mathrm{O}$ \\
+${}^{16}\mathrm{O} + {}^{16}\mathrm{O} \rightarrow \mathrm{n} + {}^{31}\mathrm{S}$ & 
+                                         \\
+${}^{18}\mathrm{O} + {}^{4}\mathrm{He} \rightarrow {}^{22}\mathrm{Ne}$ & 
+  ${}^{22}\mathrm{Ne} \rightarrow {}^{4}\mathrm{He} + {}^{18}\mathrm{O}$ \\
+${}^{18}\mathrm{F} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{21}\mathrm{Ne}$ & 
+  ${}^{21}\mathrm{Ne} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{18}\mathrm{F}$ \\
+${}^{18}\mathrm{F} + {}^{4}\mathrm{He} \rightarrow {}^{22}\mathrm{Na}$ & 
+  ${}^{22}\mathrm{Na} \rightarrow {}^{4}\mathrm{He} + {}^{18}\mathrm{F}$ \\
+${}^{18}\mathrm{F} \rightarrow {}^{18}\mathrm{O}$ & 
+                                         \\
+${}^{20}\mathrm{Ne} + {}^{4}\mathrm{He} \rightarrow {}^{24}\mathrm{Mg}$ & 
+  ${}^{24}\mathrm{Mg} \rightarrow {}^{4}\mathrm{He} + {}^{20}\mathrm{Ne}$ \\
+${}^{20}\mathrm{Ne} + {}^{12}\mathrm{C} \rightarrow {}^{4}\mathrm{He} + {}^{28}\mathrm{Si}$ & 
+  ${}^{28}\mathrm{Si} + {}^{4}\mathrm{He} \rightarrow {}^{12}\mathrm{C} + {}^{20}\mathrm{Ne}$ \\
+${}^{20}\mathrm{Ne} + {}^{12}\mathrm{C} \rightarrow {}^{1}\mathrm{H} + {}^{31}\mathrm{P}$ & 
+  ${}^{31}\mathrm{P} + {}^{1}\mathrm{H} \rightarrow {}^{12}\mathrm{C} + {}^{20}\mathrm{Ne}$ \\
+${}^{21}\mathrm{Ne} + {}^{1}\mathrm{H} \rightarrow {}^{22}\mathrm{Na}$ & 
+  ${}^{22}\mathrm{Na} \rightarrow {}^{1}\mathrm{H} + {}^{21}\mathrm{Ne}$ \\
+${}^{22}\mathrm{Ne} + {}^{1}\mathrm{H} \rightarrow {}^{23}\mathrm{Na}$ & 
+  ${}^{23}\mathrm{Na} \rightarrow {}^{1}\mathrm{H} + {}^{22}\mathrm{Ne}$ \\
+${}^{22}\mathrm{Na} \rightarrow {}^{22}\mathrm{Ne}$ & 
+                                         \\
+${}^{23}\mathrm{Na} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{20}\mathrm{Ne}$ & 
+  ${}^{20}\mathrm{Ne} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{23}\mathrm{Na}$ \\
+${}^{23}\mathrm{Na} + {}^{1}\mathrm{H} \rightarrow {}^{24}\mathrm{Mg}$ & 
+  ${}^{24}\mathrm{Mg} \rightarrow {}^{1}\mathrm{H} + {}^{23}\mathrm{Na}$ \\
+${}^{23}\mathrm{Na} + {}^{4}\mathrm{He} \rightarrow {}^{27}\mathrm{Al}$ & 
+  ${}^{27}\mathrm{Al} \rightarrow {}^{4}\mathrm{He} + {}^{23}\mathrm{Na}$ \\
+${}^{23}\mathrm{Mg} + \mathrm{n} \rightarrow {}^{24}\mathrm{Mg}$ & 
+                                         \\
+${}^{24}\mathrm{Mg} + {}^{4}\mathrm{He} \rightarrow {}^{28}\mathrm{Si}$ & 
+  ${}^{28}\mathrm{Si} \rightarrow {}^{4}\mathrm{He} + {}^{24}\mathrm{Mg}$ \\
+${}^{27}\mathrm{Al} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{24}\mathrm{Mg}$ & 
+  ${}^{24}\mathrm{Mg} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{27}\mathrm{Al}$ \\
+${}^{27}\mathrm{Al} + {}^{1}\mathrm{H} \rightarrow {}^{28}\mathrm{Si}$ & 
+  ${}^{28}\mathrm{Si} \rightarrow {}^{1}\mathrm{H} + {}^{27}\mathrm{Al}$ \\
+${}^{27}\mathrm{Al} + {}^{4}\mathrm{He} \rightarrow {}^{31}\mathrm{P}$ & 
+  ${}^{31}\mathrm{P} \rightarrow {}^{4}\mathrm{He} + {}^{27}\mathrm{Al}$ \\
+${}^{27}\mathrm{Si} + \mathrm{n} \rightarrow {}^{28}\mathrm{Si}$ & 
+                                         \\
+${}^{28}\mathrm{Si} + {}^{4}\mathrm{He} \rightarrow {}^{32}\mathrm{S}$ & 
+  ${}^{32}\mathrm{S} \rightarrow {}^{4}\mathrm{He} + {}^{28}\mathrm{Si}$ \\
+${}^{31}\mathrm{P} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{28}\mathrm{Si}$ & 
+  ${}^{28}\mathrm{Si} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{31}\mathrm{P}$ \\
+${}^{31}\mathrm{P} + {}^{1}\mathrm{H} \rightarrow {}^{32}\mathrm{S}$ & 
+  ${}^{32}\mathrm{S} \rightarrow {}^{1}\mathrm{H} + {}^{31}\mathrm{P}$ \\
+${}^{31}\mathrm{P} + {}^{4}\mathrm{He} \rightarrow {}^{35}\mathrm{Cl}$ & 
+  ${}^{35}\mathrm{Cl} \rightarrow {}^{4}\mathrm{He} + {}^{31}\mathrm{P}$ \\
+${}^{31}\mathrm{S} + \mathrm{n} \rightarrow {}^{32}\mathrm{S}$ & 
+                                         \\
+${}^{32}\mathrm{S} + {}^{4}\mathrm{He} \rightarrow {}^{36}\mathrm{Ar}$ & 
+  ${}^{36}\mathrm{Ar} \rightarrow {}^{4}\mathrm{He} + {}^{32}\mathrm{S}$ \\
+${}^{35}\mathrm{Cl} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{32}\mathrm{S}$ & 
+  ${}^{32}\mathrm{S} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{35}\mathrm{Cl}$ \\
+${}^{35}\mathrm{Cl} + {}^{1}\mathrm{H} \rightarrow {}^{36}\mathrm{Ar}$ & 
+  ${}^{36}\mathrm{Ar} \rightarrow {}^{1}\mathrm{H} + {}^{35}\mathrm{Cl}$ \\
+${}^{35}\mathrm{Cl} + {}^{4}\mathrm{He} \rightarrow {}^{39}\mathrm{K}$ & 
+  ${}^{39}\mathrm{K} \rightarrow {}^{4}\mathrm{He} + {}^{35}\mathrm{Cl}$ \\
+${}^{36}\mathrm{Ar} + {}^{4}\mathrm{He} \rightarrow {}^{40}\mathrm{Ca}$ & 
+  ${}^{40}\mathrm{Ca} \rightarrow {}^{4}\mathrm{He} + {}^{36}\mathrm{Ar}$ \\
+${}^{39}\mathrm{K} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{36}\mathrm{Ar}$ & 
+  ${}^{36}\mathrm{Ar} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{39}\mathrm{K}$ \\
+${}^{39}\mathrm{K} + {}^{1}\mathrm{H} \rightarrow {}^{40}\mathrm{Ca}$ & 
+  ${}^{40}\mathrm{Ca} \rightarrow {}^{1}\mathrm{H} + {}^{39}\mathrm{K}$ \\
+${}^{39}\mathrm{K} + {}^{4}\mathrm{He} \rightarrow {}^{43}\mathrm{Sc}$ & 
+  ${}^{43}\mathrm{Sc} \rightarrow {}^{4}\mathrm{He} + {}^{39}\mathrm{K}$ \\
+${}^{40}\mathrm{Ca} + {}^{4}\mathrm{He} \rightarrow {}^{44}\mathrm{Ti}$ & 
+  ${}^{44}\mathrm{Ti} \rightarrow {}^{4}\mathrm{He} + {}^{40}\mathrm{Ca}$ \\
+${}^{43}\mathrm{Sc} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{40}\mathrm{Ca}$ & 
+  ${}^{40}\mathrm{Ca} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{43}\mathrm{Sc}$ \\
+${}^{43}\mathrm{Sc} + {}^{1}\mathrm{H} \rightarrow {}^{44}\mathrm{Ti}$ & 
+  ${}^{44}\mathrm{Ti} \rightarrow {}^{1}\mathrm{H} + {}^{43}\mathrm{Sc}$ \\
+${}^{43}\mathrm{Sc} + {}^{4}\mathrm{He} \rightarrow {}^{47}\mathrm{V}$ & 
+  ${}^{47}\mathrm{V} \rightarrow {}^{4}\mathrm{He} + {}^{43}\mathrm{Sc}$ \\
+${}^{44}\mathrm{Ti} + {}^{4}\mathrm{He} \rightarrow {}^{48}\mathrm{Cr}$ & 
+  ${}^{48}\mathrm{Cr} \rightarrow {}^{4}\mathrm{He} + {}^{44}\mathrm{Ti}$ \\
+${}^{47}\mathrm{V} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{44}\mathrm{Ti}$ & 
+  ${}^{44}\mathrm{Ti} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{47}\mathrm{V}$ \\
+${}^{47}\mathrm{V} + {}^{1}\mathrm{H} \rightarrow {}^{48}\mathrm{Cr}$ & 
+  ${}^{48}\mathrm{Cr} \rightarrow {}^{1}\mathrm{H} + {}^{47}\mathrm{V}$ \\
+${}^{47}\mathrm{V} + {}^{4}\mathrm{He} \rightarrow {}^{51}\mathrm{Mn}$ & 
+  ${}^{51}\mathrm{Mn} \rightarrow {}^{4}\mathrm{He} + {}^{47}\mathrm{V}$ \\
+${}^{48}\mathrm{Cr} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{51}\mathrm{Mn}$ & 
+  ${}^{51}\mathrm{Mn} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{48}\mathrm{Cr}$ \\
+${}^{48}\mathrm{Cr} + {}^{4}\mathrm{He} \rightarrow {}^{52}\mathrm{Fe}$ & 
+  ${}^{52}\mathrm{Fe} \rightarrow {}^{4}\mathrm{He} + {}^{48}\mathrm{Cr}$ \\
+${}^{51}\mathrm{Mn} + {}^{1}\mathrm{H} \rightarrow {}^{52}\mathrm{Fe}$ & 
+  ${}^{52}\mathrm{Fe} \rightarrow {}^{1}\mathrm{H} + {}^{51}\mathrm{Mn}$ \\
+${}^{51}\mathrm{Mn} + {}^{4}\mathrm{He} \rightarrow {}^{55}\mathrm{Co}$ & 
+  ${}^{55}\mathrm{Co} \rightarrow {}^{4}\mathrm{He} + {}^{51}\mathrm{Mn}$ \\
+${}^{52}\mathrm{Fe} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{55}\mathrm{Co}$ & 
+  ${}^{55}\mathrm{Co} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{52}\mathrm{Fe}$ \\
+${}^{52}\mathrm{Fe} + {}^{4}\mathrm{He} \rightarrow {}^{56}\mathrm{Ni}$ & 
+  ${}^{56}\mathrm{Ni} \rightarrow {}^{4}\mathrm{He} + {}^{52}\mathrm{Fe}$ \\
+${}^{55}\mathrm{Co} + {}^{1}\mathrm{H} \rightarrow {}^{56}\mathrm{Ni}$ & 
+  ${}^{56}\mathrm{Ni} \rightarrow {}^{1}\mathrm{H} + {}^{55}\mathrm{Co}$ \\
+ &  ${}^{12}\mathrm{C} + {}^{12}\mathrm{C} \rightarrow \mathrm{n} + {}^{23}\mathrm{Mg}$ \\
+ &  ${}^{16}\mathrm{O} + {}^{12}\mathrm{C} \rightarrow \mathrm{n} + {}^{27}\mathrm{Si}$ \\
+\enddata
+\end{deluxetable}
+
 %======================================================================

### 2022-03-27 simplified_sdc_intro/paper.tex
some net information

@@
 %\documentclass[times,modern]{aastex63}
-\documentclass[times,preprint]{aastex63}
+%\documentclass[times,preprint]{aastex63}
+\documentclass{aastex63}
 
@@
 
-\begin{figure}[t]
-\centering
-\plotone{subch2}
-\caption{\label{fig:subch_network} Reaction network for the double detonation problem.}
-\end{figure}
-
 The initial model is generated following the procedure described in
@@
 
-Our reaction network includes the nuclei: $\mathrm{n}$,
-${}^{1}\mathrm{H}$, ${}^{4}\mathrm{He}$, ${}^{12}\mathrm{C}$,
-${}^{14}\mathrm{C}$, ${}^{13}\mathrm{N}$, ${}^{14}\mathrm{N}$,
-${}^{16}\mathrm{O}$, ${}^{18}\mathrm{O}$, ${}^{18}\mathrm{F}$,
-${}^{20}\mathrm{Ne}$, ${}^{21}\mathrm{Ne}$, ${}^{22}\mathrm{Ne}$,
-${}^{22}\mathrm{Na}$, ${}^{23}\mathrm{Na}$, ${}^{23}\mathrm{Mg}$,
-${}^{24}\mathrm{Mg}$, ${}^{27}\mathrm{Al}$, ${}^{27}\mathrm{Si}$,
-${}^{28}\mathrm{Si}$, ${}^{31}\mathrm{P}$, ${}^{31}\mathrm{S}$,
-${}^{32}\mathrm{S}$, ${}^{35}\mathrm{Cl}$, ${}^{36}\mathrm{Ar}$,
-${}^{39}\mathrm{K}$, ${}^{40}\mathrm{Ca}$, ${}^{43}\mathrm{Sc}$,
-${}^{44}\mathrm{Ti}$, ${}^{47}\mathrm{V}$, ${}^{48}\mathrm{Cr}$,
-${}^{51}\mathrm{Mn}$, ${}^{52}\mathrm{Fe}$, ${}^{55}\mathrm{Co}$,
-${}^{56}\mathrm{Ni}$.
-
-The rates linking them are:
+We build a reaction network that approximates an $\alpha$-chain,
+including all $\isotm{He}{4}$, and all of the $\alpha$ nuclei from
+$\isotm{C}{12}$ to $\isotm{Ni}{56}$.  We want to capture both the
+$(\alpha,\gamma)$ and $(\alpha,p)(p,\gamma)$ links, so we include the
+intermediate nuclei from the $(\alpha,p)$ rate instead of
+approximating these links assuming proton equilibriation.  
+Following \citet{shenbildsten}, we include the sequences
+$\isotm{C}{13}(\alpha,\gamma)\isotm{O}{18}(\alpha,\gamma)\isotm{Ne}{22}$
+and $\isotm{N}{14}(\alpha,\gamma)\isotm{F}{18}(\alpha,
+p)\isotm{Ne}{21}$, which produce the protons that can allow for
+$\isotm{C}{12}(p, \gamma)\isotm{N}{13}(\alpha, p)\isotm{O}{16}$ to compete
+with $\isotm{C}{12}(\alpha,\gamma)\isotm{O}{16}$.  We add $\isotm{Na}{22}$
+to link the $\isotm{Ne}{21}$ produced back to the $\alpha$-chain.  We grab
+all of the ReacLib \citep{reaclib} reaction rates linking these nuclei using \pynucastro~\citep{pynucastro}.
+We are missing one rate each for $\isotm{C}{12} + \isotm{C}{12}$,
+$\isotm{C}{12} + \isotm{O}{16}$, and $\isotm{O}{16} + \isotm{O}{16}$
+    involving neutron production followed by a capture back to one of
+    our $\alpha$-chain nuclei, so we explicitly add these rates as
+    well.
+
+Our final set of nuclei are $\mathrm{n}$, ${}^{1}\mathrm{H}$,
+${}^{4}\mathrm{He}$, ${}^{12}\mathrm{C}$, ${}^{14}\mathrm{C}$,
+${}^{13}\mathrm{N}$, ${}^{14}\mathrm{N}$, ${}^{16}\mathrm{O}$,
+${}^{18}\mathrm{O}$, ${}^{18}\mathrm{F}$, ${}^{20}\mathrm{Ne}$,
+${}^{21}\mathrm{Ne}$, ${}^{22}\mathrm{Ne}$, ${}^{22}\mathrm{Na}$,
+${}^{23}\mathrm{Na}$, ${}^{23}\mathrm{Mg}$, ${}^{24}\mathrm{Mg}$,
+${}^{27}\mathrm{Al}$, ${}^{27}\mathrm{Si}$, ${}^{28}\mathrm{Si}$,
+${}^{31}\mathrm{P}$, ${}^{31}\mathrm{S}$, ${}^{32}\mathrm{S}$,
+${}^{35}\mathrm{Cl}$, ${}^{36}\mathrm{Ar}$, ${}^{39}\mathrm{K}$,
+${}^{40}\mathrm{Ca}$, ${}^{43}\mathrm{Sc}$, ${}^{44}\mathrm{Ti}$,
+${}^{47}\mathrm{V}$, ${}^{48}\mathrm{Cr}$, ${}^{51}\mathrm{Mn}$,
+${}^{52}\mathrm{Fe}$, ${}^{55}\mathrm{Co}$, ${}^{56}\mathrm{Ni}$.  The 119
+rates linking them are given in table ~\ref{table:rates}, with the
+forward rates having a positive $Q$ value and the reverse having a
+negative $Q$ value.  We add screening to the rates using the prescription
+from The C++ code for the network is output by \pynucastro\
+directly for \castro.  A graphical overview of the network showing the links
+between the nuclei is shown in figure~\ref{fig:subch_network}.
 
@@
 \begin{deluxetable}{ll}
-\tablecaption{\label{table:react_converge_strang} Convergence ($L_1$ norm) for the reacting convergence problem using Strang splitting.}
+\tablecaption{\label{table:rates} Reaction rates included in our network.}
 \tablehead{\colhead{forward reaction} & \colhead{reverse reaction}}
@@
 
+\begin{figure}[t]
+\centering
+\plotone{subch2}
+\caption{\label{fig:subch_network} Reaction network for the double detonation problem.}
+\end{figure}
+
 %======================================================================

### 2022-03-27 simplified_sdc_intro/paper.tex
more net details

@@
 
-Our final set of nuclei are $\mathrm{n}$, ${}^{1}\mathrm{H}$,
+Our final set of 35 nuclei are $\mathrm{n}$, ${}^{1}\mathrm{H}$,
 ${}^{4}\mathrm{He}$, ${}^{12}\mathrm{C}$, ${}^{14}\mathrm{C}$,
@@
 ${}^{52}\mathrm{Fe}$, ${}^{55}\mathrm{Co}$, ${}^{56}\mathrm{Ni}$.  The 119
-rates linking them are given in table ~\ref{table:rates}, with the
+rates linking them are given in table~\ref{table:rates}, with the
 forward rates having a positive $Q$ value and the reverse having a
 negative $Q$ value.  We add screening to the rates using the prescription
-from The C++ code for the network is output by \pynucastro\
+from \citet{graboske:1973,alastuey:1978,itoh:1979}.  The C++ code for the network is output by \pynucastro\
 directly for \castro.  A graphical overview of the network showing the links

### 2022-04-06 simplified_sdc_intro/paper.tex
update network

@@
 
-Our final set of 35 nuclei are $\mathrm{n}$, ${}^{1}\mathrm{H}$,
-${}^{4}\mathrm{He}$, ${}^{12}\mathrm{C}$, ${}^{14}\mathrm{C}$,
-${}^{13}\mathrm{N}$, ${}^{14}\mathrm{N}$, ${}^{16}\mathrm{O}$,
-${}^{18}\mathrm{O}$, ${}^{18}\mathrm{F}$, ${}^{20}\mathrm{Ne}$,
-${}^{21}\mathrm{Ne}$, ${}^{22}\mathrm{Ne}$, ${}^{22}\mathrm{Na}$,
-${}^{23}\mathrm{Na}$, ${}^{23}\mathrm{Mg}$, ${}^{24}\mathrm{Mg}$,
-${}^{27}\mathrm{Al}$, ${}^{27}\mathrm{Si}$, ${}^{28}\mathrm{Si}$,
-${}^{31}\mathrm{P}$, ${}^{31}\mathrm{S}$, ${}^{32}\mathrm{S}$,
-${}^{35}\mathrm{Cl}$, ${}^{36}\mathrm{Ar}$, ${}^{39}\mathrm{K}$,
-${}^{40}\mathrm{Ca}$, ${}^{43}\mathrm{Sc}$, ${}^{44}\mathrm{Ti}$,
-${}^{47}\mathrm{V}$, ${}^{48}\mathrm{Cr}$, ${}^{51}\mathrm{Mn}$,
-${}^{52}\mathrm{Fe}$, ${}^{55}\mathrm{Co}$, ${}^{56}\mathrm{Ni}$.  The 119
+Our final set of 28 nuclei are ${}^{1}\mathrm{H}$, ${}^{4}\mathrm{He}$, ${}^{12}\mathrm{C}$, ${}^{13}\mathrm{N}$, ${}^{14}\mathrm{N}$, ${}^{16}\mathrm{O}$, ${}^{18}\mathrm{F}$, ${}^{20}\mathrm{Ne}$, ${}^{21}\mathrm{Ne}$, ${}^{22}\mathrm{Na}$, ${}^{23}\mathrm{Na}$, ${}^{24}\mathrm{Mg}$, ${}^{27}\mathrm{Al}$, ${}^{28}\mathrm{Si}$, ${}^{31}\mathrm{P}$, ${}^{32}\mathrm{S}$, ${}^{35}\mathrm{Cl}$, ${}^{36}\mathrm{Ar}$, ${}^{39}\mathrm{K}$, ${}^{40}\mathrm{Ca}$, ${}^{43}\mathrm{Sc}$, ${}^{44}\mathrm{Ti}$, ${}^{47}\mathrm{V}$, ${}^{48}\mathrm{Cr}$, ${}^{51}\mathrm{Mn}$, ${}^{52}\mathrm{Fe}$, ${}^{55}\mathrm{Co}$, ${}^{56}\mathrm{Ni}$.
+The 107
 rates linking them are given in table~\ref{table:rates}, with the
@@
   ${}^{23}\mathrm{Na} + {}^{1}\mathrm{H} \rightarrow {}^{12}\mathrm{C} + {}^{12}\mathrm{C}$ \\
-${}^{14}\mathrm{C} + {}^{4}\mathrm{He} \rightarrow {}^{18}\mathrm{O}$ & 
-  ${}^{18}\mathrm{O} \rightarrow {}^{4}\mathrm{He} + {}^{14}\mathrm{C}$ \\
-${}^{14}\mathrm{C} \rightarrow {}^{14}\mathrm{N}$ & 
+${}^{12}\mathrm{C} + {}^{12}\mathrm{C} \rightarrow {}^{24}\mathrm{Mg}$ & 
                                          \\
@@
   ${}^{27}\mathrm{Al} + {}^{1}\mathrm{H} \rightarrow {}^{12}\mathrm{C} + {}^{16}\mathrm{O}$ \\
+${}^{16}\mathrm{O} + {}^{12}\mathrm{C} \rightarrow {}^{28}\mathrm{Si}$ & 
+                                         \\
 ${}^{16}\mathrm{O} + {}^{16}\mathrm{O} \rightarrow {}^{4}\mathrm{He} + {}^{28}\mathrm{Si}$ & 
@@
   ${}^{31}\mathrm{P} + {}^{1}\mathrm{H} \rightarrow {}^{16}\mathrm{O} + {}^{16}\mathrm{O}$ \\
-${}^{16}\mathrm{O} + {}^{16}\mathrm{O} \rightarrow \mathrm{n} + {}^{31}\mathrm{S}$ & 
+${}^{16}\mathrm{O} + {}^{16}\mathrm{O} \rightarrow {}^{32}\mathrm{S}$ & 
                                          \\
-${}^{18}\mathrm{O} + {}^{4}\mathrm{He} \rightarrow {}^{22}\mathrm{Ne}$ & 
-  ${}^{22}\mathrm{Ne} \rightarrow {}^{4}\mathrm{He} + {}^{18}\mathrm{O}$ \\
 ${}^{18}\mathrm{F} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{21}\mathrm{Ne}$ & 
@@
   ${}^{22}\mathrm{Na} \rightarrow {}^{4}\mathrm{He} + {}^{18}\mathrm{F}$ \\
-${}^{18}\mathrm{F} \rightarrow {}^{18}\mathrm{O}$ & 
-                                         \\
 ${}^{20}\mathrm{Ne} + {}^{4}\mathrm{He} \rightarrow {}^{24}\mathrm{Mg}$ & 
@@
   ${}^{22}\mathrm{Na} \rightarrow {}^{1}\mathrm{H} + {}^{21}\mathrm{Ne}$ \\
-${}^{22}\mathrm{Ne} + {}^{1}\mathrm{H} \rightarrow {}^{23}\mathrm{Na}$ & 
-  ${}^{23}\mathrm{Na} \rightarrow {}^{1}\mathrm{H} + {}^{22}\mathrm{Ne}$ \\
-${}^{22}\mathrm{Na} \rightarrow {}^{22}\mathrm{Ne}$ & 
-                                         \\
 ${}^{23}\mathrm{Na} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{20}\mathrm{Ne}$ & 
@@
   ${}^{27}\mathrm{Al} \rightarrow {}^{4}\mathrm{He} + {}^{23}\mathrm{Na}$ \\
-${}^{23}\mathrm{Mg} + \mathrm{n} \rightarrow {}^{24}\mathrm{Mg}$ & 
-                                         \\
 ${}^{24}\mathrm{Mg} + {}^{4}\mathrm{He} \rightarrow {}^{28}\mathrm{Si}$ & 
@@
   ${}^{31}\mathrm{P} \rightarrow {}^{4}\mathrm{He} + {}^{27}\mathrm{Al}$ \\
-${}^{27}\mathrm{Si} + \mathrm{n} \rightarrow {}^{28}\mathrm{Si}$ & 
-                                         \\
 ${}^{28}\mathrm{Si} + {}^{4}\mathrm{He} \rightarrow {}^{32}\mathrm{S}$ & 
@@
   ${}^{35}\mathrm{Cl} \rightarrow {}^{4}\mathrm{He} + {}^{31}\mathrm{P}$ \\
-${}^{31}\mathrm{S} + \mathrm{n} \rightarrow {}^{32}\mathrm{S}$ & 
-                                         \\
 ${}^{32}\mathrm{S} + {}^{4}\mathrm{He} \rightarrow {}^{36}\mathrm{Ar}$ & 
@@
   ${}^{56}\mathrm{Ni} \rightarrow {}^{1}\mathrm{H} + {}^{55}\mathrm{Co}$ \\
- &  ${}^{12}\mathrm{C} + {}^{12}\mathrm{C} \rightarrow \mathrm{n} + {}^{23}\mathrm{Mg}$ \\
- &  ${}^{16}\mathrm{O} + {}^{12}\mathrm{C} \rightarrow \mathrm{n} + {}^{27}\mathrm{Si}$ \\
 \enddata

### 2022-04-06 simplified_sdc_intro/paper.tex
updates

@@
 intermediate nuclei from the $(\alpha,p)$ rate instead of
-approximating these links assuming proton equilibriation.  
-Following \citet{shenbildsten}, we include the sequences
-$\isotm{C}{13}(\alpha,\gamma)\isotm{O}{18}(\alpha,\gamma)\isotm{Ne}{22}$
-and $\isotm{N}{14}(\alpha,\gamma)\isotm{F}{18}(\alpha,
-p)\isotm{Ne}{21}$, which produce the protons that can allow for
-$\isotm{C}{12}(p, \gamma)\isotm{N}{13}(\alpha, p)\isotm{O}{16}$ to compete
-with $\isotm{C}{12}(\alpha,\gamma)\isotm{O}{16}$.  We add $\isotm{Na}{22}$
-to link the $\isotm{Ne}{21}$ produced back to the $\alpha$-chain.  We grab
-all of the ReacLib \citep{reaclib} reaction rates linking these nuclei using \pynucastro~\citep{pynucastro}.
-We are missing one rate each for $\isotm{C}{12} + \isotm{C}{12}$,
-$\isotm{C}{12} + \isotm{O}{16}$, and $\isotm{O}{16} + \isotm{O}{16}$
-    involving neutron production followed by a capture back to one of
-    our $\alpha$-chain nuclei, so we explicitly add these rates as
-    well.
+approximating these links assuming proton equilibriation.  Following
+\citet{shenbildsten}, we include the sequence
+$\isotm{N}{14}(\alpha,\gamma)\isotm{F}{18}(\alpha, p)\isotm{Ne}{21}$,
+which produce the protons that can allow for $\isotm{C}{12}(p,
+\gamma)\isotm{N}{13}(\alpha, p)\isotm{O}{16}$ to compete with
+$\isotm{C}{12}(\alpha,\gamma)\isotm{O}{16}$.  We add $\isotm{Na}{22}$
+to link the $\isotm{Ne}{21}$ produced back to the $\alpha$-chain.  We
+grab all of the ReacLib \citep{reaclib} reaction rates linking these
+nuclei using \pynucastro~\citep{pynucastro}.  We are missing one rate
+each for $\isotm{C}{12} + \isotm{C}{12}$, $\isotm{C}{12} +
+\isotm{O}{16}$, and $\isotm{O}{16} + \isotm{O}{16}$.  These are
+the sequences $\isotm{C}{12}(\isotm{C}{12}, \mathrm{n})\isotm{Mg}{23}(n, \gamma)\isotm{Mg}{24}$,
+$\isotm{O}{16}(\isotm{C}{12}, \mathrm{n})\isotm{Si}{27}(n, \gamma)\isotm{Si}{28}$, and
+$\isotm{O}{16}(\isotm{O}{16}, \mathrm{n})\isotm{S}{31}(n, \gamma)\isotm{S}{32}$,involving neutron
+production followed by a capture back to one of our $\alpha$-chain
+nuclei.  These are the only rates we approximate here, assuming
+that the subsequent neutron capture is instantaneous.
 

### 2022-05-08 simplified_sdc_intro/paper.tex
update table with latest runs

@@
 \hline
-{\tt strang\_subch2\_cfl0.2}           & $6.10\times 10^{31}$ & $6.72\times 10^{31}$ & $5.29\times 10^{31}$ \\
-{\tt strang\_subch2\_cfl0.4}           & $5.20\times 10^{31}$ & $6.70\times 10^{31}$ & $6.02\times 10^{31}$ \\
-{\tt strang\_subch2\_cfl0.4\_mixedtol} & $3.05\times 10^{31}$ & $7.29\times 10^{31}$ & $7.80\times 10^{31}$ \\
-{\tt strang\_subch2\_cfl0.8}           & $4.59\times 10^{31}$ & $6.88\times 10^{31}$ & $6.27\times 10^{31}$ \\
+{\tt strang\_subch2\_cfl0.1}           & 1.059\times 10^{32}  &  3.804\times 10^{31}  &  1.092\times 10^{32} \\
+{\tt strang\_subch2\_cfl0.2}           & 9.114\times 10^{31}  &  4.174\times 10^{31}  &  1.078\times 10^{32} \\
+{\tt strang\_subch2\_cfl0.2\_tol1.e-8} & 7.222\times 10^{31}  &  8.623\times 10^{31}  &  4.041\times 10^{31} \\
+{\tt strang\_subch2\_cfl0.4}           & 7.819\times 10^{31}  &  4.208\times 10^{31}  &  1.045\times 10^{32} \\
 \hline
@@
 \hline
-{\tt sdc\_subch2\_cfl0.2}           & $3.35\times 10^{31}$  & $7.26\times 10^{31}$  & $8.27\times 10^{31}$ \\
-{\tt sdc\_subch2\_cfl0.4}           & $3.30\times 10^{31}$  & $7.27\times 10^{31}$  & $8.13\times 10^{31}$ \\
-{\tt sdc\_subch2\_cfl0.4\_mixedtol} & $3.09\times 10^{31}$  & $7.36\times 10^{31}$  & $8.08\times 10^{31}$ \\
-{\tt sdc\_subch2\_cfl0.4\_tol1.e-6} & $3.25\times 10^{31}$  & $7.27\times 10^{31}$  & $8.25\times 10^{31}$ \\
-{\tt sdc\_subch2\_cfl0.8}           & $3.18\times 10^{31}$  & $7.34\times 10^{31}$  & $7.83\times 10^{31}$ \\
+{\tt sdc\_subch2\_cfl0.1}           & 7.319\times 10^{31}  &  9.147\times 10^{31}  &  3.765\times 10^{31} \\
+{\tt sdc\_subch2\_cfl0.2}           & 7.208\times 10^{31}  &  9.151\times 10^{31}  &  3.783\times 10^{31} \\
+{\tt sdc\_subch2\_cfl0.2\_tol1.e-8} & 7.376\times 10^{31}  &  9.220\times 10^{31}  &  3.513\times 10^{31} \\
+{\tt sdc\_subch2\_cfl0.4}           & 7.203\times 10^{31}  &  8.843\times 10^{31}  &  3.754\times 10^{31} \\
+\hline
 \enddata

### 2022-05-09 simplified_sdc_intro/paper.tex
lots of updates

@@
 a code.  The easiest way to improve the coupling is to cut the
-timestep (see, e.g., \citealt{couch:2015}), but this makes simulations
+timestep (see, e.g., \citealt{couch:2015}), but this can make simulations
 prohibitively expensive.  Sometimes the burning is simply halted on a
 zone-by-zone basis when NSE conditions are reached (e.g., as in
-\citealt{hedet}).  Alternately, at high temperatures, a reaction
-network can be replaced with a table of NSE abundances and the zone's
-composition set through table look-ups (e.g.\ \citealt{ma:2013}).  The
+\citealt{hedet}).  The
 present work focuses on networks alone, but in a follow-on paper we
-will explore hybrid burning models consisting of networks and NSE
-tables.
+will explore hybrid burning models consisting of networks and tables for
+nuclear statistical equilibrium.
 
@@
 method-of-lines integration.  However, it is restricted to
-second-order accuracy overall.  We term this algorithm the
+second-order accuracy in time overall.  We term this algorithm the
 ``simplified-SDC method''.  In this paper we describe the overall
@@
 multidimensional extensions are a straightforward modification to
-include the CTU hydrodynamics scheme.  Our conserved variables are
+include in the CTU hydrodynamics scheme.  Our conserved variables are
 \begin{equation}
@@
 energy formulation (see \citealt{bryan:1995,wdmergerI}).
- The mass fractions are constrained to sum to 1, $\sum_k X_k = 1$, but
- no such constraint exists on the auxiliary variables. 
+ The mass fractions are constrained to sum to 1, $\sum_k X_k = 1$.
 Defining the hydrodynamical fluxes:
@@
 
-To close the system, we need the equation of state.  For a system
-where the composition is completely specified by the mass fractions,
-$X_k$, the equation of state would take the form:
+To close the system, we need the equation of state of the form:
 \begin{equation}
@@
 in this paper.  We'll discuss this traditional approach next before moving
-on to our new time-coupling method.  Figures~\ref{fig:flowchart_I} and \ref{fig:flowchart_II} show a graphical
-flowchart of the reaction and hydrodynamics advances in the two methods.
+on to our new time-coupling method.
+
+\subsection{Strang Splitting}
+
+In Strang splitting, we first integrate the system with reactions terms only (no advection)
+over $\Delta t/2$, then integrate the advection terms only (no reactions) over $\Delta t$,
+and finally integrate the reaction terms only over $\Delta t/2$.  This flow is illustrated
+in Figure~\ref{fig:strang_flowchart}.
 
@@
   \end{figure}
-
-\begin{figure}[t]
-\plotone{sdc_strang_flowchart_II}
-\caption{\label{fig:sdc_flowchart} Flowchart of the simplified-SDC
-  approach.  An iteratively-lagged advection term is used to update
-  the system to the new time, with each iteration improving the
-  coupling.}
-\end{figure}
-
-\subsection{Strang Splitting}
-
-In Strang splitting, we first integrate the system with reactions terms only (no advection)
-over $\Delta t/2$, then integrate the advection terms only (no reactions) over $\Delta t$,
-and finally integrate the reaction terms only over $\Delta t/2$.
 
@@
 energy generation rate over the entire step (or last Strang half).
-
-\castro\ has the ability to reject a timestep if it detects a failure
-and retry with smaller timesteps (subcycling to make up the original
-required timestep).  Among the conditions that can trigger this are
-density falling below zero during advection, the ODE integration
-failing to converge in the implicit solve (due to too many steps or
-the internal timestep falling below a minimum), or violation of one of
-the timestep limiters during the step.  This means that equation
-(\ref{eq:dt:nuce_I}) or (\ref{eq:dt:nuce_II}) are not reactive, but
-instead guaranteed to be met throughout the simulation, because the
-step is rejected if they are violated.  This contrasts to similar
-approaches used in other codes where conditions \ref{eq:dt:nuce_I} or
-\ref{eq:dt:nuce_II} are used to restrict the next timestep size, but
-the update over the current step is kept, which already violated the
-constraints. \MarginPar{would be nice to have a reference here} The
-retry mechanism in \castro\ works with both the Strang and
-simplified-SDC integration scheme.  Retrying the step is much more
-stringent that simply adjusting the next step, so we put a cap on the
-number of retries allowed, aborting if we exceed the cap.
+Usually these limiters are used in a reactive manner---the next step will see
+a smaller timestep based on what happened during the current step, but the
+current step is still accepted, even if it violated these conditions.
+
+\castro\ has the ability to reject a timestep if it detects a failure and retry
+with smaller timesteps (subcycling to make up the original required timestep).
+Among the conditions that can trigger this are density falling below zero during
+advection, the ODE integration failing to converge in the implicit solve (due to
+too many steps or the internal timestep falling below a minimum), or violation
+of one of the timestep limiters during the step.  This means that the timestep
+constraints are proactive instead of reactive.  The retry mechanism in \castro\
+works with both the Strang and simplified-SDC integration scheme.  Retrying the
+step is much more stringent that simply adjusting the next step, so we put a cap
+on the number of retries allowed, aborting if we exceed the cap.
 
@@
 on the conserved variable state and some on the primitive variable
-state.
+state.   The overall flow of the simplified-SDC algorithm is shown
+in Figure~\ref{fig:flowchart_II}.
+
+\begin{figure}[t]
+\plotone{sdc_strang_flowchart_II}
+\caption{\label{fig:sdc_flowchart} Flowchart of the simplified-SDC
+  approach.  An iteratively-lagged advection term is used to update
+  the system to the new time, with each iteration improving the
+  coupling.}
+\end{figure}
 
@@
 VODE is available in our Microphysics repo:
-\url{https://github.com/starkiller-astro/Microphysics}.}  These have
+\url{https://github.com/amrex-astro/Microphysics}.}  These have
 been shown empirically to improve the performance of VODE.
@@
 fractions are contained in $[0, 1]$ to a tolerance of $10^{-2}$.
-
 For SDC, we enforce these same constraints, but for
-Eq.~\ref{eq:vode_species_reject}, we only use the change from $m$ to
-$m+1$ due to reactions, by subtracting off the advective contribution
-over that substep.  We also require that:
+Eq.~\ref{eq:vode_species_reject}, we only use the change from $m$ to $m+1$ due
+to reactions, by subtracting off the advective contribution over that substep.
+We also require that:
 \begin{align}
@@
 
-For both Strang and simplified-SDC, we will use a numerical Jacobian.
-VODE computes it internally for Strang and we use the method from
-Appendix \ref{app:jacobian} for the simplified-SDC method.  VODE
-employs Jacobian-caching, so the Jacobian does not need to be
-re-evaluated each time it is used.
+Finally, some of the rates in \citet{reaclib} increase by 100s of orders of
+magnitude at low temperature, presumably because the interpolant fit was done at
+higher temperatures.  Since VODE can take exploratory righthand side evaluations
+outside of the nominal time range, it can encounter these poorly-behaved rates
+and it can cause the integration to fail (by generating {\tt inf}s).  To prevent
+this behavior, we set a lower temperature limit of the reaction rates, below
+which we zero them out.  For the simulations here, we choose a cutoff of
+$5\times 10^7~\mathrm{K}$.
+
+For both Strang and simplified-SDC, we will use a numerical Jacobian that VODE
+computes internally (following the algorithm in \citealt{lsode}).  For
+completeness, we derive the analytic form of the Jacobian in Appendix
+\ref{sec:app:jacobian} for the simplified-SDC method. VODE employs
+Jacobian-caching, so the Jacobian does not need to be re-evaluated each time it
+is used.
 
@@
 the slope limiters don't kick in severely), which means that the
-energy release cannot be so vigorous as to drive shocks.  test here
-with the simplified SDC algorithm.
+energy release cannot be so vigorous as to drive shocks.
 
@@
 convergence respectively.  We see that both methods converge nearly
-second order, and agree well with each other.
+second order, and agree well with each other.  This demonstrates that both
+methods are working as designed.
 
@@
 
-We next model a double detonation Type Ia
-supernova.  In this model, a detonation begins in the accreted helium
-layer on a sub-Chandrasekhar mass white dwarf and that detonation can
-either drive a compression wave into the core igniting a second
-detonation in the carbon / oxygen core, or burn through the interface
-at the base of the accreted layer producing an inward propagating
-carbon detonation.  Our goal here is not to explore the feasibility of
-the model or understand whether a detonation propagating through the
-He-C interface is physical, but rather to look at the coupling of the
-hydrodynamics and reactions.  For that reason, we explore only a
-single model and start off with a rather large temperature
-perturbation in the He layer.
-
-We use a network with 28 nuclei and 96 rates that captures He burning \MarginPar{should I do a few aprox13 runs for comparison?}
-with links for $(\alpha,p)(p,\gamma)$ reactions and the rates
-involving \isot{N}{14} to bypass
-$\isotm{C}{12}(\alpha,\gamma)\isotm{O}{16}$ as described in
+We next model a double detonation Type Ia supernova.  In this model, a
+detonation begins in the accreted helium layer on a sub-Chandrasekhar mass white
+dwarf and that detonation can either drive a compression wave into the core
+igniting a second detonation in the carbon / oxygen core, or burn through the
+interface at the base of the accreted layer producing an inward propagating
+carbon detonation (see, e.g., \citealt{fink:2007}).  Our goal here is not to
+explore the feasibility of the model or understand whether a detonation
+propagating through the He-C interface is physical, but rather to look at the
+coupling of the hydrodynamics and reactions.  We want to drive vigorous burning
+and directly compare Strang and simplified-SDC time integration on this problem.
+For that reason, we explore only a single model and start off with a rather
+large temperature perturbation in the He layer.
+
+We use a network with 28 nuclei and 107 rates that captures He burning with
+links for $(\alpha,p)(p,\gamma)$ reactions and the rates involving \isot{N}{14}
+to bypass $\isotm{C}{12}(\alpha,\gamma)\isotm{O}{16}$ as described in
 \cite{shenbildsten}.  The rates are taken from the ReacLib library
-\citep{reaclib} and the network is written by
-\pynucastro~\citep{pynucastro} directly in the C++ format that our
-code requires.  The nuclei and links in the network are shown in
-Figure~\ref{fig:subch_network}.  This moderate-sized network with many
-reaction pathways (forward and reverse) make it a good test for the
-coupling between hydro and reactions.  An overview of the network is
-shown in Figure~\ref{fig:subch_network}.  This network is available 
-as {\tt subch2} in \citep{microphysics}.
+\citep{reaclib} and the network is written by \pynucastro~\citep{pynucastro}
+directly in the C++ format that our code requires.  The details of the network
+are given in Appendix \ref{sec:app:reactionnet}. This moderate-sized network
+with many reaction pathways (forward and reverse) make it a good test for the
+coupling between hydro and reactions.
 
@@
 An initial perturbation is placed on the symmetry axis of the form:
-
-We do a suite of runs with both Strang and the simplified-SDC
-integration, using CFL numbers of 0.2, 0.4, and 0.8 and reaction
-network tolerances of $\atolspec = \rtolspec = 10^{-5}$ and $\atolener
-= \rtolener = 10^{-6}$.  We also run two sets with tighter tolerances.
-The runs labeled ``mixed\_tols'' have $\atolspec = 10^{-8}$ and the
-runs labeled ``tol1.e-6'' have $\atolspec = \rtolspec = 10^{-6}$, with
-the other tolerances unchanged.  For the step rejection part of VODE,
-we use $f = 4$ and $\eta = 4$.
+\begin{equation}
+  T = T_0 \left [ 1 + X(\isotm{He}{4}) f (1 + \tanh(2 - r1)) \right ]
+\end{equation}
+where
+\begin{equation}
+  r_1 = \left [ x^2 + (y - r_0)^2 \right ]^{1/2} / \lambda
+\end{equation}
+and
+\begin{equation}
+  r_0 = r_\mathrm{pert} + r_\mathrm{base}
+\end{equation}
+Here, $r_\mathrm{base}$ is the radius at which the helium layer begins
+and $r_\mathrm{pert}$ is the distance above the base to put the perturbation.  We
+choose $r_\mathrm{pert} = 100~\mathrm{km}$.  The temperature is pertubed above
+the initial model value, denoted as $T_0$ here.  The amplitude of the perturbation
+is $f = 12$ and the scale of the perturbation is $\lambda = 200~\mathrm{km}$.
+
+We do a suite of runs with both Strang and the simplified-SDC integration, using
+CFL numbers of 0.1, 0.2, and 0.4 and reaction network tolerances of $\atolspec =
+\rtolspec = 10^{-5}$ and $\atolener = \rtolener = 10^{-65$.  We also run the CFL
+= 0.2 case with a tighter tolerance, using  $\atolspec = 10^{-8}$.  For the step
+rejection part of VODE, we use $f = 4$ and $\eta = 4$.
 
@@
 \hline
-{\tt strang\_subch2\_cfl0.1}           & 1.059\times 10^{32}  &  3.804\times 10^{31}  &  1.092\times 10^{32} \\
-{\tt strang\_subch2\_cfl0.2}           & 9.114\times 10^{31}  &  4.174\times 10^{31}  &  1.078\times 10^{32} \\
-{\tt strang\_subch2\_cfl0.2\_tol1.e-8} & 7.222\times 10^{31}  &  8.623\times 10^{31}  &  4.041\times 10^{31} \\
-{\tt strang\_subch2\_cfl0.4}           & 7.819\times 10^{31}  &  4.208\times 10^{31}  &  1.045\times 10^{32} \\
+{\tt strang\_subch2\_cfl0.1}           & $1.059\times 10^{32}$  &  $3.804\times 10^{31}$  &  $1.092\times 10^{32}$ \\
+{\tt strang\_subch2\_cfl0.2}           & $9.114\times 10^{31}$  &  $4.174\times 10^{31}$  &  $1.078\times 10^{32}$ \\
+{\tt strang\_subch2\_cfl0.2\_tol1.e-8} & $7.222\times 10^{31}$  &  $8.623\times 10^{31}$  &  $4.041\times 10^{31}$ \\
+{\tt strang\_subch2\_cfl0.4}           & $7.819\times 10^{31}$  &  $4.208\times 10^{31}$  &  $1.045\times 10^{32}$ \\
 \hline
@@
 \hline
-{\tt sdc\_subch2\_cfl0.1}           & 7.319\times 10^{31}  &  9.147\times 10^{31}  &  3.765\times 10^{31} \\
-{\tt sdc\_subch2\_cfl0.2}           & 7.208\times 10^{31}  &  9.151\times 10^{31}  &  3.783\times 10^{31} \\
-{\tt sdc\_subch2\_cfl0.2\_tol1.e-8} & 7.376\times 10^{31}  &  9.220\times 10^{31}  &  3.513\times 10^{31} \\
-{\tt sdc\_subch2\_cfl0.4}           & 7.203\times 10^{31}  &  8.843\times 10^{31}  &  3.754\times 10^{31} \\
+{\tt sdc\_subch2\_cfl0.1}           & $7.319\times 10^{31}$  &  $9.147\times 10^{31}$  &  $3.765\times 10^{31}$ \\
+{\tt sdc\_subch2\_cfl0.2}           & $7.208\times 10^{31}$  &  $9.151\times 10^{31}$  &  $3.783\times 10^{31}$ \\
+{\tt sdc\_subch2\_cfl0.2\_tol1.e-8} & $7.376\times 10^{31}$  &  $9.220\times 10^{31}$  &  $3.513\times 10^{31}$ \\
+{\tt sdc\_subch2\_cfl0.4}           & $7.203\times 10^{31}$  &  $8.843\times 10^{31}$  &  $3.754\times 10^{31}$ \\
 \hline
@@
 \end{deluxetable}
-
-1-d slices along x=0?
 
@@
 \section{Reaction Network}
-
+\label{sec:app:reactionnet}
 We build a reaction network that approximates an $\alpha$-chain,
@@
 
-Our final set of 28 nuclei are ${}^{1}\mathrm{H}$, ${}^{4}\mathrm{He}$, ${}^{12}\mathrm{C}$, ${}^{13}\mathrm{N}$, ${}^{14}\mathrm{N}$, ${}^{16}\mathrm{O}$, ${}^{18}\mathrm{F}$, ${}^{20}\mathrm{Ne}$, ${}^{21}\mathrm{Ne}$, ${}^{22}\mathrm{Na}$, ${}^{23}\mathrm{Na}$, ${}^{24}\mathrm{Mg}$, ${}^{27}\mathrm{Al}$, ${}^{28}\mathrm{Si}$, ${}^{31}\mathrm{P}$, ${}^{32}\mathrm{S}$, ${}^{35}\mathrm{Cl}$, ${}^{36}\mathrm{Ar}$, ${}^{39}\mathrm{K}$, ${}^{40}\mathrm{Ca}$, ${}^{43}\mathrm{Sc}$, ${}^{44}\mathrm{Ti}$, ${}^{47}\mathrm{V}$, ${}^{48}\mathrm{Cr}$, ${}^{51}\mathrm{Mn}$, ${}^{52}\mathrm{Fe}$, ${}^{55}\mathrm{Co}$, ${}^{56}\mathrm{Ni}$.
-The 107
-rates linking them are given in table~\ref{table:rates}, with the
-forward rates having a positive $Q$ value and the reverse having a
-negative $Q$ value.  We add screening to the rates using the prescription
-from \citet{graboske:1973,alastuey:1978,itoh:1979}.  The C++ code for the network is output by \pynucastro\
-directly for \castro.  A graphical overview of the network showing the links
-between the nuclei is shown in figure~\ref{fig:subch_network}.
+Our final set of 28 nuclei are ${}^{1}\mathrm{H}$, ${}^{4}\mathrm{He}$,
+${}^{12}\mathrm{C}$, ${}^{13}\mathrm{N}$, ${}^{14}\mathrm{N}$,
+${}^{16}\mathrm{O}$, ${}^{18}\mathrm{F}$, ${}^{20}\mathrm{Ne}$,
+${}^{21}\mathrm{Ne}$, ${}^{22}\mathrm{Na}$, ${}^{23}\mathrm{Na}$,
+${}^{24}\mathrm{Mg}$, ${}^{27}\mathrm{Al}$, ${}^{28}\mathrm{Si}$,
+${}^{31}\mathrm{P}$, ${}^{32}\mathrm{S}$, ${}^{35}\mathrm{Cl}$,
+${}^{36}\mathrm{Ar}$, ${}^{39}\mathrm{K}$, ${}^{40}\mathrm{Ca}$,
+${}^{43}\mathrm{Sc}$, ${}^{44}\mathrm{Ti}$, ${}^{47}\mathrm{V}$,
+${}^{48}\mathrm{Cr}$, ${}^{51}\mathrm{Mn}$, ${}^{52}\mathrm{Fe}$,
+${}^{55}\mathrm{Co}$, ${}^{56}\mathrm{Ni}$.  The 107 rates linking them are
+given in table~\ref{table:rates}, with the forward rates having a positive $Q$
+value and the reverse having a negative $Q$ value.  We add screening to the
+rates using the prescription from \citet{graboske:1973,alastuey:1978,itoh:1979}.
+The C++ code for the network is output by \pynucastro\ directly for \castro.  A
+graphical overview of the network showing the links between the nuclei is shown
+in figure~\ref{fig:subch_network}.
 

### 2022-05-09 simplified_sdc_intro/paper.tex
more work

@@
         \Advsst{\rho e}^{n+1/2,(k)}_{i} =
-          &- \frac{F\left((\rho e)^{n+1/2,(k)}_{i+1/2}\right) - F\left((\rho e)^{n+1/2,(k)}_{i-1/2}\right)}{\Delta x} \\
+          &- \frac{F\left((\rho e)^{n+1/2,(k)}_{i+1/2}\right) - F\left((\rho e)^{n+1/2,(k)}_{i-1/2}\right)}{\Delta x} \nonumber \\
           &- \left (\frac{p_{i+1/2}^{n+1/2,(k)} + p_{i-1/2}^{n+1/2,(k)}}{2} \right )\frac{u_{i+1/2}^{n+1/2,(k)} - u_{i-1/2}^{n+1/2,(k)}}{\Delta x}
@@
 size of $1.024\times 10^9~\mathrm{cm} \times 2.048\times
-10^9~\mathrm{cm}$, a base grid of $256 \times 512$ zones and two
-levels of refinement (both jumping by $2\times$), giving a fine-grid
-resolution of $10~\mathrm{km}$.  Gravity is modeled as a monopole.
+10^9~\mathrm{cm}$, a base grid of $256 \times 512$ zones and a single
+level of refinement (jumping by $2\times$), giving a fine-grid
+resolution of $20~\mathrm{km}$.  Gravity is modeled as a monopole.
 Our choice of initial perturbation has the detonation propagate inward
@@
 \begin{equation}
-  T = T_0 \left [ 1 + X(\isotm{He}{4}) f (1 + \tanh(2 - r1)) \right ]
+  T = T_0 \left [ 1 + X(\isotm{He}{4}) f (1 + \tanh(2 - r_1)) \right ]
 \end{equation}
@@
 CFL numbers of 0.1, 0.2, and 0.4 and reaction network tolerances of $\atolspec =
-\rtolspec = 10^{-5}$ and $\atolener = \rtolener = 10^{-65$.  We also run the CFL
+\rtolspec = 10^{-5}$ and $\atolener = \rtolener = 10^{-5}$.  We also run the CFL
 = 0.2 case with a tighter tolerance, using  $\atolspec = 10^{-8}$.  For the step
@@
 
-Figure \ref{fig:subch_strang_cfl08_I} shows the temperature, mean
+Figure \ref{fig:subch_strang_cfl02} shows the temperature, mean
 molecular weight, and nuclear energy generation rate for the strang
-CFL = 0.8 simulation ({\tt strang\_subch2\_cfl0.8}).  We see that the
+CFL = 0.82 simulation ({\tt strang\_subch2\_cfl0.2}).  We see that the
 mean molecular weight appears mottled behind the inward propagating
@@
 equilibrium at high temperatures when the reverse rates can be
-important.  In Figure~\ref{fig:subch_strang_cfl08_II} we look at some
-important mass fractions for the same simulation.  Again we see that
-the abundances are not smooth behind the inward detonation.
+important.
 
@@
 \centering
-\plotone{strang_subch2_cfl0_8_state}
-\caption{\label{fig:subch_strang_cfl08_I}}
+\plotone{strang_subch2_cfl0.2_tols_slice}
+\caption{\label{fig:subch_strang_cfl02}}
 \end{figure}
+
+We next look at the simplified-SDC version of this same case, {\tt
+  sdc\_subch2\_cfl0.2}.  Figure~\ref{fig:subch_sdc_cfl02} shows the
+temperature, mean molecular weight, and energy generation rate.
+Compared to Figure~\ref{fig:subch_strang_cfl02}, we see that the
+composition and energy generation rates are smooth, without any of the
+checkerboarding that plagued the Strang simulation.
+This suggests that the simplified-SDC algorithm has an easier time
+attaining the correct equilibrium in these conditions.
 
@@
 \centering
-\plotone{strang_subch2_cfl0_8_mass_fractions}
-\caption{\label{fig:subch_strang_cfl08_II}}
-\end{figure}
-
-We next look at the simplified-SDC version of this same case, {\tt
-  sdc\_subch2\_cfl0.8}.  Figure~\ref{fig:subch_sdc_cfl08_I} shows the
-temperature, mean molecular weight, and energy generation rate.
-Compared to Figure~\ref{fig:subch_strang_cfl08_I}, we see that the
-composition and energy generation rates are smooth, without any of the
-checkerboarding that plagued the Strang simulation.  The compositions
-plot (figure~\ref{fig:subch_sdc_cfl08_II}) shows the same behavior.
-This suggests that the simplified-SDC algorithm has an easier time
-attaining the correct equilibrium in these conditions.
-
-\begin{figure}
-\centering
-\plotone{sdc_subch2_cfl0_8_state}
-\caption{\label{fig:subch_sdc_cfl08_I}}
-\end{figure}
-
-\begin{figure}
-\centering
-\plotone{sdc_subch2_cfl0_8_mass_fractions}
-\caption{\label{fig:subch_sdc_cfl08_II}}
+\plotone{sdc_subch2_cfl0.2_tols_slice}
+\caption{\label{fig:subch_sdc_cfl02}}
 \end{figure}
@@
 We next look at how the structure seen in
-Figures~\ref{fig:subch_strang_cfl08_I} and \ref{fig:subch_sdc_cfl08_I}
+Figures~\ref{fig:subch_strang_cfl02} and \ref{fig:subch_sdc_cfl02}
 vary with CFL number and network tolerance.  We focus just on the
@@
 shock---it continues to take on the mottled appearance characteristic
-of not reaching a proper equilibrium.  For the {\tt mixedtol} case, we
-see that the state is greatly improved.  We were unable to run the
-{\tt tol1.e-6} case for Strang, as the code aborted with too many
-subcycles.  While reducing the tolerances does help, it greatly
-increases the cost of the simulation.
+of not reaching a proper equilibrium.  For the case with tight tolerances, we
+see that the state is greatly improved.
 
 \begin{figure}
-\plotone{strang_enuc_summary_plot.pdf}
+\plotone{strang_enuc_summary_plot}
 \caption{\label{fig:strang_enuc_summary} Comparison of Strang runs showing
@@
 \begin{figure}
-\plotone{strang_abar_summary_plot.pdf}
+\plotone{strang_abar_summary_plot}
 \caption{\label{fig:strang_abar_summary} Comparison of Strang runs showing
@@
 \begin{figure}
-\plotone{sdc_enuc_summary_plot.pdf}
+\plotone{sdc_enuc_summary_plot}
 \caption{\label{fig:sdc_enuc_summary} Comparison of simplified-SDC runs showing
@@
 \begin{figure}
-\plotone{sdc_abar_summary_plot.pdf}
+\plotone{sdc_abar_summary_plot}
 \caption{\label{fig:sdc_abar_summary} Comparison of simplified-SDC runs showing
@@
 hydrodynamics and reactions.  We demonstrated that for a moderate-sized reaction
-network, the simplified-SDC method
+network, the simplified-SDC method compares well against Strang for the double
+detonation problem---getting a better solution at looser reaction tolerances,
+and consistent results across a range of CFL numbers.  The benefits of this new
+integration scheme are likely problem and network dependent, and we will explore
+it on other science problemsin future papers.
 
 We have not focused on performance or load-balancing the reactions in
-the tests that we ran, but these limited runs show that the
-simplified-SDC method runs as fast (or faster) than the Strang method
-for the same tolerances and CFL, despite having to do the advection
+the tests that we ran, but for the simulations shown here,
+the simplified-SDC method runs as fast (or faster) than the Strang method
+for the same tolerances and CFL, despite having to do the advection and reactions
 twice.  We would expect similar behavior for any simulations that are
@@
 \acknowledgements \castro\ is freely available at
-\url{http://github.com/AMReX-Astro/Castro}.  All of the code and
-problem setups used here are available in the git repo.  The work at
-Stony Brook was supported by DOE/Office of Nuclear Physics grant
-DE-FG02-87ER40317.  This material is based upon work supported by the
-U.S. Department of Energy, Office of Science, Office of Advanced
-Scientific Computing Research and Office of Nuclear Physics,
-Scientific Discovery through Advanced Computing (SciDAC) program under
-Award Number DE-SC0017955.  This research was supported by the
-Exascale Computing Project (17-SC-20-SC), a collaborative effort of
-the U.S. Department of Energy Office of Science and the National
-Nuclear Security Administration.
+\url{http://github.com/AMReX-Astro/Castro}.  All of the code and problem setups
+used here are available in the git repo.  The double detonation problem is in
+{\tt Castro/Exec/science/subchandra}.  The work at Stony Brook was supported by
+DOE/Office of Nuclear Physics grant DE-FG02-87ER40317.  This material is based
+upon work supported by the U.S. Department of Energy, Office of Science, Office
+of Advanced Scientific Computing Research and Office of Nuclear Physics,
+Scientific Discovery through Advanced Computing (SciDAC) program under Award
+Number DE-SC0017955.  This research was supported by the Exascale Computing
+Project (17-SC-20-SC), a collaborative effort of the U.S. Department of Energy
+Office of Science and the National Nuclear Security Administration.  This
+research used resources of the National Energy Research Scientific Computing
+Center (NERSC), a U.S. Department of Energy Office of Science User Facility
+located at Lawrence Berkeley National Laboratory, operated under Contract No.
+DE-AC02-05CH11231 using NERSC award NP-ERCAP0020354.
 
@@
 \begin{equation}
-\left . \frac{\partial T}{\partial X_k} \right |_{\rho, e} = - \frac{e_{X_k}}{c_v}
-\end{equation}
-
-While VODE can compute the entire Jacobian, ${\bf J}$ numerically via
-differencing, we found that does not give reliable results.  Instead,
-we compute compute the derivatives with respect to $X_k$ and $T$
-one-sided differencing, following the algorithm in \cite{lsode} to
-minimize numerical noise.  We then use the equation of state to
-compute $e_{X_k}$ and $c_v$ and construct the entries of the final
-Jacobian.
+\left . \frac{\partial T}{\partial X_k} \right |_{\rho, e,X_{j,j\ne k}} = - \frac{e_{X_k}}{c_v}
+\end{equation}
+
+Finally we note that one implication of this formulation is that the Jacobian is no longer sparse.  Future
+work will explore alternate formulations and approximate Jacobians.
+
+% While VODE can compute the entire Jacobian, ${\bf J}$ numerically via
+% differencing, we found that does not give reliable results.  Instead,
+% we compute compute the derivatives with respect to $X_k$ and $T$
+% one-sided differencing, following the algorithm in \cite{lsode} to
+% minimize numerical noise.  We then use the equation of state to
+% compute $e_{X_k}$ and $c_v$ and construct the entries of the final
+% Jacobian.
 

### 2022-05-09 simplified_sdc_intro/paper.tex
some updates

@@
 %\documentclass[times,modern]{aastex63}
-%\documentclass[times,preprint]{aastex63}
-\documentclass{aastex63}
+\documentclass[times,preprint]{aastex63}
+%\documentclass{aastex63}
 
@@
 
-\author[0000-0002-1530-781X]{Alice Harpole}
+\author[0000-0002-1530-781X]{A.~Harpole}
 \affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
@@
 
-\author{Melissa Rasmussen}
+\author{M.~Rasmussen}
+\affiliation{Utah State University, Logan UT 84322}
 
@@
 Project (17-SC-20-SC), a collaborative effort of the U.S. Department of Energy
-Office of Science and the National Nuclear Security Administration.  This
+Office of Science and the National Nuclear Security Administration.  MR was supported
+via an NSF REU grant to Stony Brook, NSF 1852143.  This
 research used resources of the National Energy Research Scientific Computing
 Center (NERSC), a U.S. Department of Energy Office of Science User Facility
-located at Lawrence Berkeley National Laboratory, operated under Contract No.
+located at Lawrence Berkeley National Laboratory, operated under Contract No.\
 DE-AC02-05CH11231 using NERSC award NP-ERCAP0020354.
@@
           Linux (\url{https://www.kernel.org}),
-          matplotlib (\citealt{Hunter:2007},  \url{http://matplotlib.org/})
+          matplotlib (\citealt{Hunter:2007},  \url{http://matplotlib.org/}),
+          NetworkX \citep{networkx},
           NumPy \citep{numpy,numpy2},
@@
           \pynucastro\ \citep{pynucastro},
-          SymPy \citep{sympy}
+          pytest \citep{pytest},
+          SymPy \citep{sympy},
+          yt \citep{yt}
          }

### 2022-05-10 simplified_sdc_intro/paper.tex
more edits

@@
 comprehensive SDC approach we demonstrated previously.  We apply the
-new method to some example problems with small and moderately-sized
-astrophysical nuclear reaction networks and explore the timestep size
+new method to a double detonation problem with a moderately-sized
+astrophysical nuclear reaction network and explore the timestep size
 and reaction network tolerances to show that the simplified-SDC
@@
 
-In Strang splitting, we first integrate the system with reactions terms only (no advection)
+In the Strang splitting~\cite{strang:1968} flavor of operator splitting, we first integrate the system with reactions terms only (no advection)
 over $\Delta t/2$, then integrate the advection terms only (no reactions) over $\Delta t$,
@@
 \citet{strang_rnaas} and showed that the above formulation gets second
-order convergence and showed that integrating an energy / temperature
+order convergence and that in particular, integrating an energy / temperature
 equation is needed to get second order accuracy (see also
@@
 a smaller timestep based on what happened during the current step, but the
-current step is still accepted, even if it violated these conditions.
+current step is still accepted, even if it violated these conditions.  Our
+goal is to avoid the need for these limiters by doing the coupling of hydrodynamics
+and reactions better.
 
@@
 step is much more stringent that simply adjusting the next step, so we put a cap
-on the number of retries allowed, aborting if we exceed the cap.
+on the number of retries allowed, aborting if we exceed the cap. \MarginPar{describe in more detail?}
 
@@
 integral, which is done using an iteratively-lagged solution.  This is the
-classic approach that we implemented in \cite{castro_sdc}.
-
-Our simplified-SDC approach is based on the method described in \citet{SDC-old} and
-a similar (but unpublished) implementation in the \maestroex\ simulation code \cite{maestroex}.
-We again start by considering the update
-in integral form:
+classic approach that we implemented in Castro in \citet{castro_sdc}.
+
+Our simplified-SDC approach is based on the method described in \citet{SDC-old}
+and a similar (but unpublished) implementation in the \maestroex\ simulation
+code \citep{maestroex}. We again start by considering the update in integral
+form:
 \begin{equation}
@@
 state.   The overall flow of the simplified-SDC algorithm is shown
-in Figure~\ref{fig:flowchart_II}.
+in Figure~\ref{sdc_flowchart}.
 
@@
 to reactions, by subtracting off the advective contribution over that substep.
-We also require that:
+Finally, we enforce that:
 \begin{align}
@@
 
-Finally, some of the rates in \citet{reaclib} increase by 100s of orders of
+Some of the rates in \citet{reaclib} increase by 100s of orders of
 magnitude at low temperature, presumably because the interpolant fit was done at
@@
 this behavior, we set a lower temperature limit of the reaction rates, below
-which we zero them out.  For the simulations here, we choose a cutoff of
+which we zero them out.  For the simulations in section~\ref{sec:dd}, we choose a cutoff of
 $5\times 10^7~\mathrm{K}$.
@@
 
-\cite{castro_sdc} introduced a test problem for measuring convergence
+\citet{castro_sdc} introduced a test problem for measuring convergence
 of a reacting hydrodynamic algorithm.  This is based on the acoustic
-pulse problem from \citep{mccorquodalecolella}, and was also used in
-\citep{strang_rnaas}.  A periodic domain with a uniform entropy is
+pulse problem from \citet{mccorquodalecolella}, and was also used in
+\citet{strang_rnaas}.  A periodic domain with a uniform entropy is
 initialized with a pressure profile with a small perturbation.  This
@@
 
-Tables \ref{table:react_converge_strang} and
-\ref{table:react_converge_sdc} show the Strang and simplified-SDC
-convergence respectively.  We see that both methods converge nearly
+Tables \ref{table:react_converge_strang} and \ref{table:react_converge_sdc} show
+the Strang and simplified-SDC convergence respectively.  The convergence rate is
+is computed by taking log-base-2 of the ratio of adjacent errors.  A rate of 2.0
+indicates second order convergence.  We see that both methods converge nearly
 second order, and agree well with each other.  This demonstrates that both
@@
 \subsection{Sub-Chandra Double Detonation}
+\label{sec:dd}
 
@@
 = 0.2 case with a tighter tolerance, using  $\atolspec = 10^{-8}$.  For the step
-rejection part of VODE, we use $f = 4$ and $\eta = 4$.
+rejection part of VODE, we use $f = 4$ and $\eta = 4$.  The runs are labeled with
+the prefix {\tt sdc\_} or {\tt strang\_} to detonate the integrator, and also include
+the CFL number in the name.  Finally, the runs with the tighter tolerances have
+the suffix {\tt \_tol1.e-8}.  We note that both the Strang and the SDC runs encounter
+VODE integration failures occassionally and utilize the \castro\ step retry functionality.
 
@@
 molecular weight, and nuclear energy generation rate for the strang
-CFL = 0.82 simulation ({\tt strang\_subch2\_cfl0.2}).  We see that the
+CFL = 0.2 simulation ({\tt strang\_subch2\_cfl0.2}).  At this point
+in the evolution, the helium detonation has wrapped more than halfway
+across the surface and the ingoing carbon detonation is approaching the
+center of the star.  We see that the
 mean molecular weight appears mottled behind the inward propagating
@@
 \plotone{strang_subch2_cfl0.2_tols_slice}
-\caption{\label{fig:subch_strang_cfl02}}
+\caption{\label{fig:subch_strang_cfl02} Temperature, mean molecular weight, and
+energy generation rate for the Strang CFL = 0.2 run with normal tolerances ({\tt
+strang_subch2\_cfl0.2}) at 0.1 s.}
 \end{figure}
@@
 We next look at the simplified-SDC version of this same case, {\tt
-  sdc\_subch2\_cfl0.2}.  Figure~\ref{fig:subch_sdc_cfl02} shows the
-temperature, mean molecular weight, and energy generation rate.
-Compared to Figure~\ref{fig:subch_strang_cfl02}, we see that the
-composition and energy generation rates are smooth, without any of the
-checkerboarding that plagued the Strang simulation.
-This suggests that the simplified-SDC algorithm has an easier time
-attaining the correct equilibrium in these conditions.
+sdc\_subch2\_cfl0.2}.  Figure~\ref{fig:subch_sdc_cfl02} shows the temperature,
+mean molecular weight, and energy generation rate. Compared to
+Figure~\ref{fig:subch_strang_cfl02}, we see that the composition and energy
+generation rates are smooth, without any of the checkerboarding that plagued the
+Strang simulation.  This suggests that the simplified-SDC algorithm has an
+easier time attaining the correct equilibrium in these conditions. More
+strikingly, we see that the inward propagating carbon detonation has not
+progressed as far into the white dwarf at this time.  The two integration
+methods are giving very different results here.
 
@@
 \plotone{sdc_subch2_cfl0.2_tols_slice}
-\caption{\label{fig:subch_sdc_cfl02}}
+\caption{\label{fig:subch_sdc_cfl02} Temperature, mean molecular weight, and
+energy generation rate for the simplified-SDC CFL = 0.2 run with normal tolerances ({\tt
+sdc_subch2\_cfl0.2}) at 0.1 s.}
 \end{figure}
 
-We now look at the evolution of the state while burning by storing the
-thermodynamic state internal to VODE each substep as it evaluates the
-righthand side of the ODE system it is integrating.  We pick a point
-on the symmetry axis a few zones behind the inward moving shock and
-ran both a Strang and simplified-SDC simulation for a single
-coarse-grid timestep.  To ensure that we start off the same, both
-simulations were started from a checkpoint file from the
-simplified-SDC simulation a few steps before 0.1~s.
-Figure~\ref{fig:nonaka_plot} shows the evolution of \isot{Ni}{56} over
-a single coarse-grid timestep (4 fine-grid timesteps).  We see that
-the Strang solution is much more discontinuous, as the reactions and
-advection push the solution in different ways, because the processes
-are completely independent of one another.  If an equilibrium state is
-important to the nucleosynthesis, the large deviations in the Strang
-solution mean that we are evaluating the reaction network with a state
-out of equilibrium for much of the integration time. \MarginPar{need to double check the data---why are there points in the vertical part of Strang?}
-
-\begin{figure}
-\plotone{sdc_compare}
-\caption{\label{fig:nonaka_plot} Evolution of $X(\isotm{Ni}{56})$ over
-  4 fine-grid timesteps in a single zone behind the inward propagating
-  shock.  The Strang points (blue) show the first $\Delta t/2$ burn
-  bring the abundance away from equilibrium, then the advection
-  overcorrecting it, and finally the second $\Delta t/2$ bringing it
-  back toward the equilibrium.  The two simplified-SDC iterations
-  appear much smoother.  For the SDC runs, there are a few points
-  evaluated beyond the stop time---VODE can use data from outside the
-  integration range to interpolate to the final solution.}
-\end{figure}
+% We now look at the evolution of the state while burning by storing the
+% thermodynamic state internal to VODE each substep as it evaluates the
+% righthand side of the ODE system it is integrating.  We pick a point
+% on the symmetry axis a few zones behind the inward moving shock and
+% ran both a Strang and simplified-SDC simulation for a single
+% coarse-grid timestep.  To ensure that we start off the same, both
+% simulations were started from a checkpoint file from the
+% simplified-SDC simulation a few steps before 0.1~s.
+% Figure~\ref{fig:nonaka_plot} shows the evolution of \isot{Ni}{56} over
+% a single coarse-grid timestep (4 fine-grid timesteps).  We see that
+% the Strang solution is much more discontinuous, as the reactions and
+% advection push the solution in different ways, because the processes
+% are completely independent of one another.  If an equilibrium state is
+% important to the nucleosynthesis, the large deviations in the Strang
+% solution mean that we are evaluating the reaction network with a state
+% out of equilibrium for much of the integration time. \MarginPar{need to double check the data---why are there points in the vertical part of Strang?}
+
+% \begin{figure}
+% \plotone{sdc_compare}
+% \caption{\label{fig:nonaka_plot} Evolution of $X(\isotm{Ni}{56})$ over
+%   4 fine-grid timesteps in a single zone behind the inward propagating
+%   shock.  The Strang points (blue) show the first $\Delta t/2$ burn
+%   bring the abundance away from equilibrium, then the advection
+%   overcorrecting it, and finally the second $\Delta t/2$ bringing it
+%   back toward the equilibrium.  The two simplified-SDC iterations
+%   appear much smoother.  For the SDC runs, there are a few points
+%   evaluated beyond the stop time---VODE can use data from outside the
+%   integration range to interpolate to the final solution.}
+% \end{figure}
 
@@
 of not reaching a proper equilibrium.  For the case with tight tolerances, we
-see that the state is greatly improved.
+see that the state is greatly improved.  Furthermore, it gets the inward
+propagating carbon detonation in the same position as the SDC runs.
 
@@
 shock looks great regardless of the CFL or integrator tolerances.
+This demonstrates that the simplified-SDC integration algorithm is more
+robust than Strang integration for this problem.
 
@@
 integration scheme are likely problem and network dependent, and we will explore
-it on other science problemsin future papers.
+it on other science problems in future papers.
 
@@
 the simplified-SDC method runs as fast (or faster) than the Strang method
-for the same tolerances and CFL, despite having to do the advection and reactions
+for the same tolerances and CFL number, despite having to do the advection and reactions
 twice.  We would expect similar behavior for any simulations that are
-dominated by the cost of reactions.  If the results above a robust across
+dominated by the cost of reactions.  If the results above are robust across
 different problems and networks, then this suggests that we can run
@@
 This new time-integration method is freely available in \castro, and
-we will continue to explore it on other problems.  We are particularly
+we will continue to explore it on other problems.  Although these simulations
+were run on CPUs, the entire simplified-SDC algorithm also runs on GPUs.
 interested in massive star evolution with this method, where the

### 2022-05-10 simplified_sdc_intro/paper.tex
builds

@@
 energy generation rate for the Strang CFL = 0.2 run with normal tolerances ({\tt
-strang_subch2\_cfl0.2}) at 0.1 s.}
+strang\_subch2\_cfl0.2}) at 0.1 s.}
 \end{figure}
@@
 energy generation rate for the simplified-SDC CFL = 0.2 run with normal tolerances ({\tt
-sdc_subch2\_cfl0.2}) at 0.1 s.}
+sdc\_subch2\_cfl0.2}) at 0.1 s.}
 \end{figure}

### 2022-06-01 simplified_sdc_intro/paper.tex
update to latest aastex
put in the overleaf version

@@
 %\documentclass[times,modern]{aastex63}
-\documentclass[times,preprint]{aastex63}
+\documentclass[times,preprint]{aastex631}
 %\documentclass{aastex63}
@@
 \newcommand{\Ics}{{\mathcal{I}}}
-\newcommand{\smax}{{s_\mathrm{max}}}
+\newcommand{\kmax}{{k_\mathrm{max}}}
 \newcommand{\kth}{k_\mathrm{th}}
@@
 
-\newcommand{\Shydro}{{{\bf H}}}
-\newcommand{\Hb}{{\bf H}}
+\newcommand{\Gb}{{\bf G}}
 \newcommand{\Rb}{{\bf R}}
@@
 \author[0000-0003-0439-4556]{M.~P.~Katz}
-\affiliation{Nvidia Corp}
-
-\author[0000-0003-1791-0265]{A.~J.~Nonaka}
-\affiliation{Center for Computational Sciences and Engineering, Lawrence Berkeley National Laboratory, Berkeley, CA  94720}
-
-\author[0000-0002-1530-781X]{A.~Harpole}
 \affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
@@
 
-\author{M.~Rasmussen}
+\author[0000-0003-1791-0265]{A.~Nonaka}
+\affiliation{Center for Computational Sciences and Engineering, Lawrence Berkeley National Laboratory, Berkeley, CA  94720}
+
+\author[0000-0002-0297-0313]{M.~Rasmussen}
 \affiliation{Utah State University, Logan UT 84322}
@@
 temperatures where nuclear statistical equilibrium is established.  We
-develop a simplified approach based on the ideas of spectral deferred
+develop a new approach based on the ideas of spectral deferred
 corrections (SDC) coupling of explicit hydrodynamics and stiff
-reaction sources as an alternative to operator splitting or the more
+reaction sources as an alternative to operator splitting that is simpler
+than the more
 comprehensive SDC approach we demonstrated previously.  We apply the
@@
 and reaction network tolerances to show that the simplified-SDC
-approach compares favorably with traditional Strang operator
+approach provides improved coupling with decreased computational
+expense compared to traditional Strang operator
 splitting.  This is all done in the framework of the
@@
 Modeling astrophysical reacting flows can be challenging because of
-the disparity better the nuclear and hydrodynamics timescales.
+the disparity between the nuclear and hydrodynamics timescales.
 Reaction networks tend to be stiff, requiring implicit integration
@@
 flow over the next timestep, as the code struggles to produce an
-equilibrium.  These large changes in abundances (and large alternately
-positive and negative energy generation rates) can be a challenge for
+equilibrium. 
+These large changes in abundances (and large alternately positive and negative energy generation rates) can be a challenge for
 a code.  The easiest way to improve the coupling is to cut the
-timestep (see, e.g., \citealt{couch:2015}), but this can make simulations
+timestep (see, e.g., \citealt{couch:2015,rivas:2022}), but this can make simulations
 prohibitively expensive.  Sometimes the burning is simply halted on a
@@
 mesh refinement (AMR) framework \citep{amrex_joss}.  \castro\ has been
-designed to be performance portable and runs on massively parallel CPU
-and GPU architectures \citep{castro_gpu}.  For hydrodynamics, the
+designed to be performance portable and runs on massively parallel CPU,
+multicore, and GPU architectures \citep{castro_gpu}.  For hydrodynamics, the
 corner transport upwind (CTU; \citealt{ppmunsplit}) method with the
@@
 pair-instability supernovae.  Recently in \citet{castro_sdc}, we
-developed second- and fourth-order accurate methods in space and time
-for coupling hydrodynamics and nuclear reaction networks based on
-spectral deferred corrections (SDC) methods, and demonstrated the
-method in a variety of test problems.
+developed second- and fourth-order accurate in space and time
+method-of-lines approaches for coupling hydrodynamics and nuclear reaction networks based on
+spectral deferred corrections (SDC), and demonstrated these
+methods using a variety of test problems.
 
@@
 than the SDC method of \citet{castro_sdc}, but allows us to reuse the
-main CTU \MarginPar{any cites to CCSE papers?}  hydrodynamics
-construction and a largely similar ODE integration scheme, making this
+piecewise-linear or piecewise-parabolic CTU hydrodynamics
+construction \citep{saltzman1994,millercolella:2002}
+used in the the original \castro\ paper
+%\MarginPar{any cites to CCSE papers? - you would want Colella90 (2D CTU), Saltzman94 (3D CTU), Miller/Colella02 (2D/3D PPM)}
+and a largely similar ODE integration scheme, making this
 method easier to add to existing simulation codes.  Furthermore, it
@@
 straightforward manner, avoiding the complications described in
-\citep{mccorquodalecolella} needed to fill ghost cells when using
+\cite{mccorquodalecolella} needed to fill ghost cells when using
 method-of-lines integration.  However, it is restricted to
@@
 simulation using \castro.  All of the code to reproduce the results in
-this paper are in the \castro\ github
+this paper are in the \castro\ GitHub
 repository\footnote{\url{https://github.com/amrex-astro/Castro/}}.
@@
 \end{equation}
-where for the special case of $(\rho e)$, we have an additional ``pdV'' term:
-\begin{equation}
-\ddt{(\rho e)} + \ddx{F(\rho e)} + p \ddx{u} = S(\rho e)
-\end{equation}
-
-We'll split the source term, $\Sc(\Uc)$ into pure hydrodynamic and reactive parts:
+where for the special case of $(\rho e)$, we have an additional ``$p\, dV$'' term:
+\begin{equation}
+\ddt{(\rho e)} + \ddx{F_{\rho e}} + p \ddx{u} = S_{\rho e}
+\end{equation}
+We split the source term into gravitational and reactive parts, $\Sc(\Uc) = \Gb(\Uc) + \Rb(\Uc)$, with
 \begin{equation}
   \label{eq:cons_sources}
-  \Hb(\Uc) = \left ( \begin{array}{c}
+  \Gb(\Uc) = \left ( \begin{array}{c}
     0 \\
@@
     \rho u g \\
-    0 \end{array} \right )
+    0 \end{array} \right ),
   \qquad
@@
      0 \\
-     \rho \omegadot(X_k) \\
+     \rho \dot\omega_{X_k} \\
      0 \\
@@
      \rho \dot{S}
-  \end{array} \right )
-\end{equation}
-with
-\begin{equation}
-  \Sc(\Uc) = \Hb(\Uc) + \Rb(\Uc).
-\end{equation}
-Here, gravity, represented by the gravitational acceleration $g$
-appears as a pure hydrodynamical source term.  From reactions,
-$\omegadot(X_k)$ is the creation rate for species $k$
+  \end{array} \right ).
+\end{equation}
+From reactions,
+$\dot\omega_{X_k}$ is the creation rate for species $k$
  and $\dot{S}$ is the energy generation rate per unit
-mass.  We note that the internal energy ``pdV'' work is not treated as
+mass.  We note that the internal energy $p\partial u/\partial x$ (``$p\, dV$'')
+work is not treated as
 a source term but is instead constructed with the hydrodynamical
-fluxes are computed in the CTU method.
-Aside from the $p\partial u/\partial x$ (``pdV'') term in the internal
+fluxes that are computed in the CTU method.
+Aside from the ``$p\, dV$'' term in the internal
 energy equation, this system is in conservative form with source
@@
 
-We'll define the advective terms with hydrodynamic sources, $\Advs{\Uc}$ as
-\begin{equation}
-\Advs{\Uc} = -\ddx{\Fb(\Uc)} + \Hb(\Uc)
+We define the advective terms with gravitational sources, $\Advs{\Uc}$, as
+\begin{equation}
+\Advs{\Uc} = -\ddx{\Fb(\Uc)} + \Gb(\Uc)
 \end{equation}
 for the general case, with the $(\rho e)$ component again having the extra
-``pdV'' term:
-\begin{equation}
-\mathcal{A}(\rho e) = -\ddx{F(\rho e)} -p \ddx{u} + H(\rho e)
-\end{equation}
-
-To close the system, we need the equation of state of the form:
-\begin{equation}
-p = p(\rho, X_k, e)
-\end{equation}
-
-Sometimes it is preferable to work with the primitive variables,
+``$p\, dV$'' term:
+\begin{equation}
+\mathcal{A}_{\rho e} = -\ddx{F_{\rho e}} -p \ddx{u} + G_{\rho e}
+\end{equation}
+To close the system, we need an equation of state of the form:
+\begin{equation}
+p = p(\rho, {\bf X}, e).
+\end{equation}
+Sometimes it is preferable to work with the primitive variables:
 \begin{equation}
@@
 alternate formulations are possible \citep{colellaglaz:1985}.
-The source term vector, $\Sc(\qb)$, can again be decomposed into hydrodynamic
+The source term vector, $\Sc(\qb)$, can again be decomposed into gravitational
 sources (now in terms of the primitive variables) and reaction terms,
 \begin{equation}
-  \Sc(\qb) = \Hb(\qb) + \Rb(\qb)
+  \Sc(\qb) = \Gb(\qb) + \Rb(\qb),
 \end{equation}
@@
 \label{eq:prim_sources}
-\Hb(\qb) = \left ( \begin{array}{c}
+\Gb(\qb) = \left ( \begin{array}{c}
      0 \\
@@
      0 \\
-   \end{array} \right )
+   \end{array} \right ),
 \qquad
@@
      0 \\
-     \omegadot(X_k) \\
+     \dot\omega_{X_k} \\
      0 \\
@@
      \rho \Sdot
-   \end{array} \right )
+   \end{array} \right ),
 \end{equation}
@@
 equation can be found in \cite{ABNZ:III}.  We note that this source is
-algebraically identical to that shown in Eq.~25 of \cite{castro}.
+algebraically identical to that shown in Equation (25) of \cite{castro}.
 
@@
 traditionally employed here, and is used as a benchmark for comparison
-in this paper.  We'll discuss this traditional approach next before moving
+in this paper.  We discuss this traditional approach next before moving
 on to our new time-coupling method.
@@
 
-In the Strang splitting~\cite{strang:1968} flavor of operator splitting, we first integrate the system with reactions terms only (no advection)
+In the Strang splitting~\citep{strang:1968} flavor of operator splitting, we first integrate the system with reactions terms only (no advection)
 over $\Delta t/2$, then integrate the advection terms only (no reactions) over $\Delta t$,
@@
   \plotone{sdc_strang_flowchart_I}
-  \caption{\label{fig:strang_flowchart} Flowchart of the operator split,
-    Strang, time integration.  We first react for $\Delta t/2$, then
+  \caption{\label{fig:strang_flowchart} Flowchart of the operator split (Strang) time integration.  We first react for $\Delta t/2$, then
     advect for $\Delta t$, and finally react for $\Delta t/2$.  Each
@@
 \odt{\rho} & = 0 \\
-\odt{(\rho X_k)} &= \rho \omegadot(X_k) \\
+\odt{(\rho X_k)} &= \rho \dot\omega_{X_k} \\
 \odt{(\rho u)} &= 0 \\
@@
 \end{align}
+Notice that in the Strang formulation density is held constant when reacting.
 We can write the energy equation as:
@@
 \end{equation}
-The reaction rates are typically expressed as $\omegadot_k(\rho, T, X_k)$
+The reaction rates are typically expressed as $\omegadot_k(\rho, T, {\bf X})$
 when we evolve this system, which requires us to get temperature from
@@
 
-We also typically integrate mass fraction itself, instead of partial
+We also typically integrate mass fractions, instead of partial
 densities:
@@
 \citet{astronum:2018} (using an earlier version of the present
-algorithm).  A variation on Strang splitting called (re-)balanced
-splitting was developed in \citet{speth:2013}, but we do not explore
-that here. \MarginPar{keep this sentence?}
+algorithm).  
+%A variation on Strang splitting called (re-)balanced
+%splitting was developed in \citet{speth:2013}, but we do not explore
+%that here. \MarginPar{keep this sentence? MK: Don't think it's needed.}
 
@@
 full corner coupling, the advective CFL condition is unity) as
-compared to a method-of-lines approach (see \citealt{ppmunsplit}).  In
+compared to a method-of-lines approach; see \cite{ppmunsplit}.  In
 addition to the standard CFL timestep limiter for explicit
@@
 reactions can only increase the internal energy by 50\%).  The two
-formulations differ in that Eq.~\ref{eq:dt:nuce_I} uses the
+formulations differ in that Equation (\ref{eq:dt:nuce_I}) uses the
 instantaneous evaluation of the energy generation rate at the start of
-the timestep while Eq.~\ref{eq:dt:nuce_II} uses the integral of the
+the timestep while Equation (\ref{eq:dt:nuce_II}) uses the integral of the
 energy generation rate over the entire step (or last Strang half).
@@
 current step is still accepted, even if it violated these conditions.  Our
-goal is to avoid the need for these limiters by doing the coupling of hydrodynamics
-and reactions better.
+goal is to avoid the need for these limiters by improving the coupling of hydrodynamics
+and reactions.
 
@@
 step is much more stringent that simply adjusting the next step, so we put a cap
-on the number of retries allowed, aborting if we exceed the cap. \MarginPar{describe in more detail?}
+on the number of retries allowed, aborting if we exceed the cap. 
+%\MarginPar{describe in more detail?
+MK: While I do want us to evangelize the retry scheme, it %seems slightly out of place in this paper.
+MZ: I do note later that the retry kicked in for %both Strang and SDC, so that's why I added it here.}
 
@@
 
-The basic idea of spectral deferred corrections \MarginPar{ref} is to express the
-update as an integral and divide the time-update into a number of
-discrete time nodes.  The integral is then approximated using a
-quadrature rule over these time nodes and low order approximations are
-used to update the state from one time node to the next.  The method
-uses iteration to successively improve the solution and the ultimate
-accuracy is determined by the quadrature method used to evaluate the
-integral, which is done using an iteratively-lagged solution.  This is the
-classic approach that we implemented in Castro in \citet{castro_sdc}.
-
-Our simplified-SDC approach is based on the method described in \citet{SDC-old}
+The basic idea of spectral deferred corrections \citep{dutt:2000} is to divide each time step into sub-steps defined by high-order quadrature points (e.g.~Gauss-Lobatto) and iteratively correct the solution at each temporal node in order to reduce the integration error.
+Each correction equation can be formulated using low-order integration techniques, such as forward or backward-Euler, with additional terms on each right-hand-side arising from integrating the residual over the sub-step from the previous iteration using high-order temporal integration.
+Each iteration over all sub-steps increases the overall order of accuracy of the method by one, up to the underlying order of accuracy of the quadrature over the entire time step.
+A number of SDC approaches have been developed
+over the years targeting tighter multiphysics coupling, such as 
+\cite{bourlioux2003high,SDC-old}.
+In these approaches, different physical processes can be treated with different approaches (such as forward-Euler for advection and backward-Euler for reactions), and furthermore different physical processes can be sub-stepped between temporal nodes.
+
+The approach that we implemented in \citet{castro_sdc} is based off a
+multi-implicit, method of lines approach in \citet{bourlioux2003high}.
+In contrast to those works, the focus of this paper is not higher-order integration, but tighter coupling
+between advection and reactions in a second-order framework.
+Thus, our simplified-SDC approach is based on the method described in \citet{SDC-old}
 and a similar (but unpublished) implementation in the \maestroex\ simulation
-code \citep{maestroex}. We again start by considering the update in integral
-form:
-\begin{equation}
-\Uc^{n+1} = \Uc^n + \int_t^{t+\Delta t} \left [ \Advs{\Uc} + \Rb(\Uc) \right ] dt,
-\end{equation}
-In this approach, we
-approximate the advective term
-piecewise constant in time, using the value at the midpoint in time,
-to achieve second-order accuracy, giving us:
-\begin{equation}
-\label{eq:integral:simplesdc}
-\Uc^{n+1} = \Uc^n + \int_t^{t+\Delta t} \left \{ \Adv{\Uc}^{n+1/2} + \Rb(\Uc) \right \} dt
-\end{equation}
-The \castro\ version of the simplified-SDC algorithm
-differs from the previous versions due to the need to do some operations
-on the conserved variable state and some on the primitive variable
-state.   The overall flow of the simplified-SDC algorithm is shown
-in Figure~\ref{sdc_flowchart}.
-
+code \citep{maestroex}.
+The key point in this approach is that we are re-using the same numerical kernels 
+for advection and reactions that are used in a Strang splitting approach,
+but we are able to include, rather than exclude, the effects of the other 
+physical processes in each correction equation solve.
+
+We begin with the general form of a multi-implicit SDC correction equation with two physical processes, as originally presented in \cite{bourlioux2003high}.
+Using $(k)$ to denote the iterate, we have
+\begin{eqnarray}
+\Uc^{(k)}(t) = \Uc^n &+& \int_{t^n}^t \left [ \Advs{\Uc^{(k)}} + \Rb\left(\Uc^{(k)}\right) - \Advs{\Uc^{(k-1)}} - \Rb\left(\Uc^{(k-1)}\right) \right ] dt \nonumber\\
+&+& \int_{t^n}^t \left [ \Advs{\Uc^{(k-1)}} + \Rb\left(\Uc^{(k-1)}\right) \right ] dt.
+\end{eqnarray}
+In the spirit of \cite{SDC-old}, we treat each advection term as piecewise constant over the time step,
+and equal to a time-centered advection term evaluated with the Godunov integrator.
+We note that this is a departure from the approach in \cite{bourlioux2003high} that uses a method of lines approach, where 
+the advection terms are computed at each node, rather than the midpoint between nodes, using spatial extrapolation only.
+Using $\boldsymbol{\mathcal{A}}(\Uc^{n+1/2})$ to denote the advection term, we have
+\begin{eqnarray}
+\Uc^{(k)}(t) = \Uc^n &+& \int_{t^n}^t \left [ \Advs{\Uc^{n+1/2,(k)}} + \Rb\left(\Uc^{(k)}\right) - \Advs{\Uc^{n+1/2,(k-1)}} - \Rb\left(\Uc^{(k-1)}\right) \right ] dt \nonumber\\
+&+& \int_{t^n}^t \left [ \Advs{\Uc^{n+1/2,(k-1)}} + \Rb\left(\Uc^{(k-1)}\right) \right ] dt.
+\end{eqnarray}
+Also following \cite{SDC-old}, we differentiate this equation in time
+\begin{eqnarray}
+\frac{\partial\Uc^{(k)}}{\partial t} = && \left [ \Advs{\Uc^{n+1/2,(k)}} + \Rb\left(\Uc^{(k)}\right) - \Advs{\Uc^{n+1/2,(k-1)}} - \Rb\left(\Uc^{(k-1)}\right) \right]\nonumber\\
+&+& \left [ \Advs{\Uc^{n+1/2,(k-1)}} + \Rb\left(\Uc^{(k-1)}\right) \right].
+\end{eqnarray}
+We combine terms, and arrive at a system of ODEs at each cell,
+\begin{equation}
+\frac{\partial\Uc^{(k)}}{\partial t} =  \left [ \Advs{\Uc^{n+1/2,(k)}} + \Rb\left(\Uc^{(k)}\right) \right].\label{eq:sdcupdate}
+\end{equation}
+We integrate this system from $t^n$ to $t^{n+1}$ using a modified version
+of the VODE~\citep{vode} integrator, as described in Section \ref{sec:VODE}.
+Thus, we are effectively using two temporal nodes ($t^n$ and $t^{n+1}$) 
+and the overall integration is second-order accurate.
+In order to properly couple reactions into the computation of the advection terms, 
+we use an iteratively lagged reaction source term,
+denoted $\Ic_q$ (this is basically an approximation of $\Rb(\qb)$).
+The overall flow of the simplified-SDC algorithm is shown in Figure~\ref{fig:sdc_flowchart}.
+
+%We start by considering the update in integral form:
+%\begin{equation}
+%\Uc^{n+1} = \Uc^n + \int_t^{t+\Delta t} \left [ \Advs{\Uc} + \Rb(\Uc) \right ] dt,
+%\end{equation}
+%In this approach, we approximate the advective term piecewise constant in time, using the value at the midpoint in time, to achieve second-order accuracy, giving us:
+%\begin{equation}
+%\label{eq:integral:simplesdc}
+%\Uc^{n+1} = \Uc^n + \int_t^{t+\Delta t} \left \{ \Adv{\Uc^{n+1/2}} + \Rb(\Uc) \right \} dt
+%\end{equation}
+%The \castro\ version of the simplified-SDC algorithm differs from the previous versions due to the need to do some operations on the conserved variable state and some on the primitive variable state.
 \begin{figure}[t]
@@
 
-The basic time update algorithm proceeds as:
+\subsection{Time Advancement Scheme}
+The basic time update algorithm proceeds as follows.
+We begin with the state at $t^n$, denoted $\Uc^n$, and proceed as follows:
 
@@
 
-\item {\em Initialization}
-
-\label{sec:initialization}
-
-  \begin{itemize}
-  \item We need an approximation of how much the reactions alone
-    changed the primitive variable state over the timestep, which we
-    will denote $\Ic_q$ (this is basically an approximation of
-    $\Rb(\qb)$).
-
-    Since we do not have any information about the current
-    timestep in the first iteration, we use the value from the last
-    iteration of the previous timestep:
-    \begin{equation}
-      \Ic^{n+1/2,(0)}_{\qb} = \Ic^{n-1/2,(\smax)}_{\qb}
-    \end{equation}
+%\item {\em Initialization}
+
+%\label{sec:initialization}
+
+%  \begin{itemize}
+%  \item We need an approximation of how much the reactions alone
+%    changed the primitive variable state over the timestep, which we
+%    will denote $\Ic_q$ (this is basically an approximation of
+%    $\Rb(\qb)$).
+
+%    Since we do not have any information about the current
+%    timestep in the first iteration, we use the value from the last
+%    iteration of the previous timestep:
+%    \begin{equation}
+%      \Ic^{n+1/2,(0)}_{\qb} = \Ic^{n-1/2,(\kmax)}_{\qb}
+%    \end{equation}
+
   %% \item Solve the Poisson problem for the initial gravitational potential:
@@
 
-  \end{itemize}
+  %\end{itemize}
 
@@
 
-  Iterate from $k = 1, \ldots, \smax$.  For second-order accuracy,
-  $\smax = 2$ is sufficient.  In addition to denoting the time-level
-  with a superscript (like $n$ or $n+1$), we'll use a second subscript
+  Iterate from $k = 1, \ldots, \kmax$.  For second-order accuracy,
+  $\kmax = 2$ is sufficient; further SDC iterations will continue to decrease the splitting error,
+  but not increase the formal order of accuracy of the method.  In addition to denoting the time-level
+  with a superscript (like $n$ or $n+1$), we use a second superscript
   in parentheses to keep track of the iteration.  A single iteration, $k$,
@@
   \begin{itemize}
-  \item {\em Create the advective update term, $\Adv{\Uc}^{n+1/2,(k)}$}
+  \item {\em Step 1: Create the advective update term, $\Adv{\Uc}^{n+1/2,(k)}$}
 
     \begin{itemize}
-    \item convert $\Uc \rightarrow \qb$.  This is an algebraic transformation,
-      but will require the EOS.
-
-    \item predict $\qb$ to the interfaces at $t^{n+1/2}$ using the CTU PPM
+    \item Convert $\Uc \rightarrow \qb$.  This is an algebraic transformation,
+      that utilizes the equation of state.
+
+    \item Predict $\qb$ to the interfaces at $t^{n+1/2}$ using the CTU PPM
       method.  The source terms, $\Sq$, used in the prediction are:
       \begin{equation}
-        \Sc(\qb) = \Hb(\qb) + \Ic_\qb^{n+1/2,(k-1)}
-      \end{equation}
-      Here we use the iteratively lagged integrals of the primitive variable
-      terms accounting only for reactions, $\Ic_\qb^{n+1/2,(k-1)}$, as the
-      reactive source.  This is in contrast to Strang-splitting, where no
-      explicit reactive source terms are seen by the hydrodynamics update.
+        \Sc(\qb) = \Gb(\qb) + \Ic_\qb^{n+1/2,(k-1)}
+      \end{equation}
+      $\Ic_\qb^{n+1/2,(k-1)}$ is a numerical approximation for the effect of
+      reactions from the previous iterate, which is computed at the end of each iterate 
+      as described below.
+      The use of this term contrasts with Strang-splitting, where no
+      reactive source terms are included in the hydrodynamics update.
+      If $k=1$, we instead use the value from the last iteration of the previous
+      time step, $\Ic^{n-1/2,(\kmax)}_{\qb}$.
 %%       Any hydrodynamic source terms are time-centered
@@
       corrections are made.  Both are second-order accurate.  For the
-      hydrodynamics sources, we do those in the normal predictor.
+      gravitational sources, we do those in the normal predictor.
       However, for the reactive sources, we found that it is most
@@
       zero, and characteristic tracing or the various flux corrections
-      may not preserve this.  We enforce that the species interface
+      in general do not preserve this.  We enforce that the species interface
       states remain in $[0, 1]$ after adding the reactive source.
 
-    \item solve the Riemann problem at each interface to get a unique
-      conserved state on each interface, $\Uc^{n+1/2,(k)}_{i+1/2}$
-
-    \item construct the advective update terms, $\Advt{\Uc}^{n+1/2,(k)}_{i}$,
-      by first ignoring the hydrodynamics sources,
+    \item Solve the Riemann problem at each interface to get a unique
+      conserved state on each interface, $\Uc^{n+1/2,(k)}_{i+1/2}$.
+
+    \item Construct the advective update terms, $\Advt{\Uc}^{n+1/2,(k)}_{i}$,
+      first without the gravitational sources,
       \begin{equation}
@@
       for the general case, and
-      \begin{align}
-        \Advsst{\rho e}^{n+1/2,(k)}_{i} =
-          &- \frac{F\left((\rho e)^{n+1/2,(k)}_{i+1/2}\right) - F\left((\rho e)^{n+1/2,(k)}_{i-1/2}\right)}{\Delta x} \nonumber \\
-          &- \left (\frac{p_{i+1/2}^{n+1/2,(k)} + p_{i-1/2}^{n+1/2,(k)}}{2} \right )\frac{u_{i+1/2}^{n+1/2,(k)} - u_{i-1/2}^{n+1/2,(k)}}{\Delta x}
-      \end{align}
+      \begin{eqnarray}
+        \left(\tilde{\mathcal{A}}_{\rho e}^{n+1/2,(k)}\right)_{i} &=&
+          - \frac{\left(F_{\rho e}^{n+1/2,(k)}\right)_{i+1/2} - \left(F_{\rho e}^{n+1/2,(k)}\right)_{i-1/2}}{\Delta x} \nonumber \\
+          &&- \left (\frac{p_{i+1/2}^{n+1/2,(k)} + p_{i-1/2}^{n+1/2,(k)}}{2} \right )
+          \left(\frac{u_{i+1/2}^{n+1/2,(k)} - u_{i-1/2}^{n+1/2,(k)}}{\Delta x}\right)
+      \end{eqnarray}
       for $(\rho e)$.
 
-    Now the conservative hydrodynamics source terms are computed by first updating to the
+    Now the gravitational source terms, $\Gb(\Uc)$, are computed by first updating to the
     new state with advection and the old-time source term applied for the full $\Delta t$ as:
     \begin{equation}
-      \Uc^{\star\star} = \Uc^n + \Delta t \Advt{\Uc}^{n+1/2,(k)} + \Delta t \Hb(\Uc^n)
+      \Uc^{\star\star} = \Uc^n + \Delta t \Advt{\Uc}^{n+1/2,(k)} + \Delta t \Gb(\Uc^n)\label{eq:grav_pred}
     \end{equation}
@@
     source.
-    The final advective update term is then\footnote{For a source like gravity, this update can be done first for $\rho$ and then define the new momentum source using $\rho^{\star\star}$ and likewise for energy}:
+    The final advective update term is then\footnote{For a source like gravity, this update can be done first for $\rho$ and then define the new momentum source using $\rho^{\star\star}$ and likewise for energy.}:
     \begin{equation}
       \Adv{\Uc}^{n+1/2,(k)}_{i} = \Advt{\Uc}^{n+1/2,(k)}_{i} +
-      \frac{\Delta t}{2} \left (\Shydro(\Uc^n) + \Shydro(\Uc^{\star\star}) \right )
+      \frac{\Delta t}{2} \left [\Gb(\Uc^n) + \Gb(\Uc^{\star\star}) \right ]\label{eq:grav_corr}
     \end{equation}
@@
     %% \begin{equation}
-    %%   \Shydro^{n+1/2} = \left ( \begin{array}{c}
+    %%   \Gb^{n+1/2} = \left ( \begin{array}{c}
     %%                  0 \\ 0 \\ 0 \\
@@
 
-    We note that this procedure works because the source terms here do
-    not depend on the outcome of the reactions.
+    Formally, per \cite{SDC-old}, an alternative strategy would be to skip Equation (\ref{eq:grav_pred}) and instead
+    perform only Equation (\ref{eq:grav_corr}) but with $\Gb(\Uc^{n+1,(k-1)})$ replacing $\Gb(\Uc^{\star\star})$.
+    In the future we will explore the efficacy of both approaches.
+    %\MarginPar{AJN: rewrote this; ok?  If the alternative works it's actually much better since %fewer gravity solves are required.}
+    %We note that this procedure works because the source terms here do not depend on the outcome of the reactions.
+In \cite{SDC-old}, instead of using $\Gb(\Uc^{\star\star})$, $\Gb(\Uc^{n+1,(k-1)})$ is used.
+For the conservative gravity formulation in \cite{wdmergerI}, we prefer the former formulation since the source terms are computed using the result of the Riemann solver.\MarginPar{MK: I think this should be written to say that we are *assuming* the source terms do not depend on the outcome of the reactions, or clarify that it is in fact true specifically in Castro that we only use source terms where this is true.
+But actually, wouldn't thermal diffusion depend on the outcome of reactions?
+MZ: I am not sure anymore what the intent was here.
+I think we should just remove this sentence.
+I don't think that it affects diffusion.
+I think it meant that $\omegadot$ does not explicitly appear..}
 
@@
 
-  \item {\em Update the System Using a Method of Lines Integration}
-
-    We update the state by doing the integral in equation
-    (\ref{eq:integral:simplesdc}).  Since we are approximating the
-    advective term as piecewise constant in time, we can simply use an
+  \item {\em Step 2: Integrate Using VODE}
+  %\MarginPar{Need to change this.  The point of this paper is we are NOT using method of lines (that was the older castro sdc paper).  This simplified approach is NOT method of lines.}
+
+    We update the state by integrating equation
+    (\ref{eq:sdcupdate}) over the full time step.  Since we are approximating
+    $\boldsymbol{\mathcal{A}}(\Uc)$ as piecewise constant in time, we can use an
     ODE integrator to integrate this, just as we do with the reaction
-    system in Strang splitting.  The difference here being that we are
+    system in Strang splitting.  The difference here is that we are
     integrating the conserved variables and the state sees the effect
-    of advection as we integrate the reactions.  So rather than use
-    $d\Uc/dt = \Rb(\Uc)$, the ODE form we use is
+    of advection and gravity as we integrate the reactions.  So rather than use
+    $d\Uc/dt = \Rb(\Uc)$ as in the Strang case, the ODE form we use is
     \begin{equation}
@@
     \end{equation}
-
-    Looking at the form of Eq.~\ref{eq:cons_sources}, we see that only
-    the species and energies have source terms.  The total and
+    The details of the VODE~\citep{vode} integrator we use are
+    described in Section \ref{sec:VODE}.
+
+    Looking at the form of Equation (\ref{eq:cons_sources}), we see that only
+    the species and energies have non-zero terms in $\Rb$.  The total and
     internal energies both provide the same information, and
@@
     \begin{equation}
-      \rho(t) = \rho^n + \Advss{\rho}^{n+1/2,(k)} \, (t - t^n)
+      \rho(t) = \rho^n + \mathcal{A}_\rho^{n+1/2,(k)} \, (t - t^n)
     \end{equation}
     As we are integrating this system we need to get the temperature,
-    $T$, for the rate evaluations.  We obtain this by directly from
+    $T$, for the rate evaluations.  We obtain this directly from
     internal energy, composition, and density using the equation of
@@
     Our integrator also needs the Jacobian of the system, in terms of
-    the $\Uc^\prime$.  This is different than the form of the
+    $\Uc^\prime$.  This is different than the form of the
     Jacobian usually used in reaction networks.  We describe the form
-    of the Jacobian in appendix \ref{sec:app:jacobian}.
+    of the Jacobian in Appendix \ref{sec:app:jacobian}.
 
@@
     \begin{equation}
-      (\rho u)^{n+1} = (\rho u)^n + \Delta t \Advss{\rho u}^{n+1/2,(k)}
+      (\rho u)^{n+1,(k)} = (\rho u)^n + \Delta t \mathcal{A}_{\rho u}^{n+1/2,(k)}
     \end{equation}
@@
     \begin{equation}
-      (\rho \Sdot)^{n+1/2} = \frac{(\rho e)^{n+1} - (\rho e)^n}{\Delta t} - \Advss{\rho e}^{n+1/2,(k)}
+      (\rho \Sdot)^{n+1/2,(k)} = \frac{(\rho e)^{n+1,(k)} - (\rho e)^n}{\Delta t} - \mathcal{A}_{\rho e}^{n+1/2,(k)}
     \end{equation}
@@
     \begin{equation}
-      (\rho E)^{n+1} = (\rho E)^n + \Delta t \Advss{\rho E}^{n+1/2,(k)} + \Delta t (\rho \Sdot)^{n+1/2}
-    \end{equation}
-
-  \item {\em Compute the Reactive Source Terms.}
+      (\rho E)^{n+1,(k)} = (\rho E)^n + \Delta t \mathcal{A}_{\rho E}^{n+1/2,(k)} + \Delta t (\rho \Sdot)^{n+1/2,(k)}
+    \end{equation}
+
+  \item {\em Step 3: Compute the Reactive Source Terms, $\Ic^{n+1/2,(k)}$}
 
@@
     \begin{equation}
-      \Ic^{(k)}_{\Uc} = \frac{\Uc^{n+1,(k)} - \Uc^n}{\Delta t} - \Adv{\Uc}^{n+1/2,(k)}
+      \Ic^{n+1/2,(k)}_{\Uc} = \frac{\Uc^{n+1,(k)} - \Uc^n}{\Delta t} - \Adv{\Uc}^{n+1/2,(k)}
     \end{equation}
@@
     \begin{equation}
-      \label{eq:Iq}
-      \Ic^{(k)}_{\qb} = \frac{\qb^{n+1,(k)} - \qb^n}{\Delta t} - \Adv{\qb}^{n+1/2,(k)}
+      \Ic^{n+1/2,(k)}_{\qb} = \frac{\qb^{n+1,(k)} - \qb^n}{\Delta t} - \Adv{\qb}^{n+1/2,(k)}\label{eq:Iq_prim}
     \end{equation}
     but we need the advective update for $\qb$, which we have not
-    constructed.  We note that $\Ic^{(k)}_\qb$ is an approximation to the integral of
-    Eq.~\ref{eq:prim_sources} over the timestep.  Additionally, we cannot simply use the equation of
-    state on $\Ic^{(k)}_{\Uc}$ since this is a time-derivative and
+    constructed.  We note that $\Ic^{n+1/2,(k)}_\qb$ is an approximation to the integral of
+    Equation (\ref{eq:prim_sources}) over the timestep.  Additionally, we cannot simply use the equation of
+    state on $\Ic^{n+1/2,(k)}_{\Uc}$ since this is a time-derivative and
     does not represent a well-defined state in itself.  Instead, we
-    derive $\Ic^{(k)}_{\qb}$ via a multi-step process.  We first find
+    derive $\Ic^{n+1/2,(k)}_{\qb}$ via a multi-step process.  We first find
     the conservative state as if it were updated only with advection:
@@
     \begin{equation}
-      \Adv{\qb}^{n+1/2,(k)} = \frac{\qb^\star - \qb^n}{\Delta t}
+      \Adv{\qb}^{n+1/2,(k)} = \frac{\qb^\star - \qb^n}{\Delta t}\label{eq:Aq_prim}
     \end{equation}
@@
     conserved state via an algebraic transform, $\Uc^{n+1,(k)}
-    \rightarrow \qb^{n+1,(k)}$, we can construct $\Ic^{(k)}_{\qb}$ as given
-    in Equation (\ref{eq:Iq}).
-    Putting all of this together, we see:
-    \begin{equation}
-      \Ic^{(k)}_{\qb} = \frac{\qb^{n+1,(k)} - \qb^\star}{\Delta t}
+    \rightarrow \qb^{n+1,(k)}$, we can construct $\Ic^{n+1/2,(k)}_{\qb}$ by
+    combining Equations (\ref{eq:Iq_prim}) and (\ref{eq:Aq_prim}),
+    \begin{equation}
+      \Ic^{n+1/2,(k)}_{\qb} = \frac{\qb^{n+1,(k)} - \qb^\star}{\Delta t}.
     \end{equation}
@@
 
-\subsection{Stiff ODE Integrator}
+\subsection{Stiff ODE Integrator}\label{sec:VODE}
 
 We use a modified version of the VODE~\citep{vode} integrator,
-designed for stiff ODEs.  Our version has been ported C++, as a
+designed for stiff ODEs.  Our version has been ported to C++, as a
 header-only library enabling it to run on GPUs.  We have also added
@@
 VODE is available in our Microphysics repo:
-\url{https://github.com/amrex-astro/Microphysics}.}  These have
-been shown empirically to improve the performance of VODE.
-
-VODE allows us to specific both relative and absolute tolerances
+\url{https://github.com/AMReX-Astro/Microphysics}.}.  We have found empirically that these checks improve the performance of VODE.
+
+VODE allows us to specify both relative and absolute tolerances
 for the species and energy during the integration.  They are combined
@@
 
-For Strang, we require that
+For Strang splitting, we require that
 an individual mass fraction not change too much over a step:
@@
 For SDC, we enforce these same constraints, but for
-Eq.~\ref{eq:vode_species_reject}, we only use the change from $m$ to $m+1$ due
+Equation (\ref{eq:vode_species_reject}), we only use the change from $m$ to $m+1$ due
 to reactions, by subtracting off the advective contribution over that substep.
@@
 
-Some of the rates in \citet{reaclib} increase by 100s of orders of
+Some of the rates in \citet{reaclib} increase by hundreds of orders of
 magnitude at low temperature, presumably because the interpolant fit was done at
-higher temperatures.  Since VODE can take exploratory righthand side evaluations
+higher temperatures.  Since VODE can take exploratory right-hand side evaluations
 outside of the nominal time range, it can encounter these poorly-behaved rates
@@
 
-For both Strang and simplified-SDC, we will use a numerical Jacobian that VODE
+For both Strang splitting and simplified-SDC, we use a numerical Jacobian that VODE
 computes internally (following the algorithm in \citealt{lsode}).  For
@@
 $\isotm{C}{12}(\alpha,\gamma)\isotm{O}{16}$ rates releases energy.  By
-running at different resolutions, with a fixed timestep that scales
-with resolution, we can compute the convergence rate.  In order to
-demonstrate second-order convergence, the problem must be smooth (so
+running at different spatial resolutions with a fixed CFL number so that the timestep scales
+with resolution, we can compute the convergence rate in space and time.
+In order to demonstrate second-order convergence, the problem must be smooth (so
 the slope limiters don't kick in severely), which means that the
@@
 Tables \ref{table:react_converge_strang} and \ref{table:react_converge_sdc} show
-the Strang and simplified-SDC convergence respectively.  The convergence rate is
+the Strang splitting and simplified-SDC convergence respectively.  The convergence rate 
 is computed by taking log-base-2 of the ratio of adjacent errors.  A rate of 2.0
@@
 
+%\MarginPar{AJN: do you have any wall-clock, or RHS-count, or Nonaka-plot performance metrics?
+SDC should not only give better scientific results, but should be faster as well...
+I address this with the new plots in the last section now.}
 We next model a double detonation Type Ia supernova.  In this model, a
@@
 coupling of the hydrodynamics and reactions.  We want to drive vigorous burning
-and directly compare Strang and simplified-SDC time integration on this problem.
+and directly compare Strang splitting and simplified-SDC time integration on this problem.
 For that reason, we explore only a single model and start off with a rather
@@
 \citep{subchandra}.  We use a $1.1~M_\odot$ carbon white dwarf with a
-$0.05~M_\odot$ He layer.  We seed the He layer with 1\% (by mass) of
+$0.05~M_\odot$ He layer.
+We choose a pure \isot{C}{12} white dwarf to make the reactions more vigorous, consistent with our focus on the coupling of hydrodynamics and reactions.
+We seed the He layer with 1\% (by mass) of
 \isot{N}{14}.  The white dwarf is isothermal and the temperature ramps
@@
 surface.  Figure~\ref{fig:subch_initial_model} shows the structure of
-the initial model.  We use 2d axisymmetric coordinates with a domain
-size of $1.024\times 10^9~\mathrm{cm} \times 2.048\times
+the initial model.  We use 2D axisymmetric coordinates with a domain
+size of $1.024\times 10^9~\mathrm{cm}$ by $2.048\times
 10^9~\mathrm{cm}$, a base grid of $256 \times 512$ zones and a single
@@
 and $r_\mathrm{pert}$ is the distance above the base to put the perturbation.  We
-choose $r_\mathrm{pert} = 100~\mathrm{km}$.  The temperature is pertubed above
+choose $r_\mathrm{pert} = 100~\mathrm{km}$.  The temperature is perturbed above
 the initial model value, denoted as $T_0$ here.  The amplitude of the perturbation
@@
 
-We do a suite of runs with both Strang and the simplified-SDC integration, using
+We do a suite of runs with both Strang splitting and the simplified-SDC integration, using
 CFL numbers of 0.1, 0.2, and 0.4 and reaction network tolerances of $\atolspec =
@@
 rejection part of VODE, we use $f = 4$ and $\eta = 4$.  The runs are labeled with
-the prefix {\tt sdc\_} or {\tt strang\_} to detonate the integrator, and also include
+the prefix {\tt sdc\_} or {\tt strang\_} to denote the integrator, and also include
 the CFL number in the name.  Finally, the runs with the tighter tolerances have
 the suffix {\tt \_tol1.e-8}.  We note that both the Strang and the SDC runs encounter
-VODE integration failures occassionally and utilize the \castro\ step retry functionality.
+VODE integration failures occasionally and utilize the \castro\ step retry functionality.
 
 Figure \ref{fig:subch_strang_cfl02} shows the temperature, mean
-molecular weight, and nuclear energy generation rate for the strang
+molecular weight, and nuclear energy generation rate for the Strang splitting
 CFL = 0.2 simulation ({\tt strang\_subch2\_cfl0.2}).  At this point
@@
 generation rates are smooth, without any of the checkerboarding that plagued the
-Strang simulation.  This suggests that the simplified-SDC algorithm has an
-easier time attaining the correct equilibrium in these conditions. More
+Strang simulation.  This suggests that the simplified-SDC algorithm more accurately
+attains the correct equilibrium in these conditions. More
 strikingly, we see that the inward propagating carbon detonation has not
@@
 Figures~\ref{fig:subch_strang_cfl02} and \ref{fig:subch_sdc_cfl02}
-vary with CFL number and network tolerance.  We focus just on the
+varies with CFL number and network tolerance.  We focus just on the
 region behind the inward propagating shock.
@@
 of not reaching a proper equilibrium.  For the case with tight tolerances, we
-see that the state is greatly improved.  Furthermore, it gets the inward
-propagating carbon detonation in the same position as the SDC runs.
+see that the state is greatly improved with a smoother gradient.  Furthermore, the inwardly
+propagating carbon detonation reaches the same position as in the SDC runs.
 
@@
 and \ref{fig:sdc_abar_summary}.  We see that the state behind the
-shock looks great regardless of the CFL or integrator tolerances.
+shock looks smoother and more consistent regardless of the CFL number or integrator tolerances.
 This demonstrates that the simplified-SDC integration algorithm is more
@@
 
-To quantify this effect, we next looks at the total mass of some key
+To quantify this effect, we next look at the total mass of some key
 species.  This is often a desired diagnostic for supernova
@@
 
+Finally, we can ask how many right-hand side evaluations each integration method required, as a measure of computational cost; when an integration uses many RHS evaluations, it often implies difficulty around equilibrium.
+We stored this in the plotfiles every $0.001~\mathrm{s}$, and on the finest level, for zones that were burning, we compute the L1, L2, and L-inf norms of the number of right-hand side calls (the L1 norm is the same as the average and the L-inf is the same as the maximum here).
+This includes the calls that were made to evaluate the numerical Jacobian.
+Figure~\ref{fig:strang_nrhs} shows the results for the {\tt strang\_subch2\_cfl0.2} run.
+We see a steady
+increase in the cost of the reactions as time progresses.
+Note that the L-inf norm curve is much higher than the other curves, but this is dominated by only a few zones right at the detonation.
+The other norms give a better sense of the total work in reactions.
+Figure~\ref{fig:sdc_nrhs} shows the corresponding plot for the simplified-SDC case ({\tt sdc\_subch2\_cfl0.2}).  While the L-inf
+norm is of the same magnitude, it fluctuates much more.  However,
+both the L1 and L2 norms are smaller, indicating that the integration does not need to work as hard with the simplified-SDC integration as compared to Strang.
+%\MarginPar{AJN: add a note about the 1e-8 tol strang run (that gives same looking results as %SDC) probably is even MORE expensive?}
+
+\begin{figure}[t]
+\plotone{strang_nrhs}
+\caption{\label{fig:strang_nrhs} Number of righthand side evaluations for {\tt strang\_subch2\_cfl0.2} showing the L1 norm/average (solid), L2 norm (dashed), and L-inf/maximum (dotted)
+for both halves of the Strang integration.}
+\end{figure}
+
+\begin{figure}[t]
+\plotone{sdc_nrhs}
+\caption{\label{fig:sdc_nrhs} Number of righthand side evaluations for {\tt sdc\_subch2\_cfl0.2} showing the L1 norm/average (solid), L2 norm (dashed), and L-inf/maximum (dotted)
+for 2 iterations of the simplified-SDC integration.}
+\end{figure}
 \section{Summary}
@@
 detonation problem---getting a better solution at looser reaction tolerances,
-and consistent results across a range of CFL numbers.  The benefits of this new
+consistent results across a range of CFL numbers, and is more computationally efficient.
+The benefits of this new
 integration scheme are likely problem and network dependent, and we will explore
@@
 the tests that we ran, but for the simulations shown here,
-the simplified-SDC method runs as fast (or faster) than the Strang method
-for the same tolerances and CFL number, despite having to do the advection and reactions
-twice.  We would expect similar behavior for any simulations that are
+the simplified-SDC method runs as fast (or faster) than the Strang splitting method
+for the same tolerances and CFL number.
+%, despite having to do the advection and reactions
+%twice.
+%\MarginPar{AJN: I would delete "despite having to to the advection and reactions twice".  %Strang already does reactions twice, but advection once.  But advection is cheap, right?}
+We would expect similar behavior for any simulations that are
 dominated by the cost of reactions.  If the results above are robust across
 different problems and networks, then this suggests that we can run
-the SDC simulation with larger CFL and/or less strict reaction tolerances,
+the SDC simulation with larger CFL number and/or less strict reaction tolerances,
 which would translate into a large performance boost.
@@
 were run on CPUs, the entire simplified-SDC algorithm also runs on GPUs.
-interested in massive star evolution with this method, where the
+We are interested in massive star evolution with this method, where the
 vigorous shell burning leading up to core-collapse would be an ideal
@@
 
-\acknowledgements \castro\ is freely available at
-\url{http://github.com/AMReX-Astro/Castro}.  All of the code and problem setups
+\begin{acknowledgments} 
+\castro\ is freely available at
+\url{https://github.com/AMReX-Astro/Castro}.  All of the code and problem setups
 used here are available in the git repo.  The double detonation problem is in
-{\tt Castro/Exec/science/subchandra}.  The work at Stony Brook was supported by
-DOE/Office of Nuclear Physics grant DE-FG02-87ER40317.  This material is based
+{\tt Castro/Exec/science/subchandra}.  The reaction network used
+here is available at \url{https://github.com/AMReX-Astro/Microphysics} as the {\tt subch2} network.  The work at Stony Brook was supported by
+DOE/Office of Nuclear Physics grant DE-FG02-87ER40317.  We thank Alice Harpole
+for her contributions to the AMReX Astrophysics suite.  This material is based
 upon work supported by the U.S. Department of Energy, Office of Science, Office
@@
 DE-AC02-05CH11231 using NERSC award NP-ERCAP0020354.
+\end{acknowledgments}
 
@@
 $\isotm{O}{16}(\isotm{C}{12}, \mathrm{n})\isotm{Si}{27}(n, \gamma)\isotm{Si}{28}$, and
-$\isotm{O}{16}(\isotm{O}{16}, \mathrm{n})\isotm{S}{31}(n, \gamma)\isotm{S}{32}$,involving neutron
+$\isotm{O}{16}(\isotm{O}{16}, \mathrm{n})\isotm{S}{31}(n, \gamma)\isotm{S}{32}$, involving neutron
 production followed by a capture back to one of our $\alpha$-chain
@@
 given in table~\ref{table:rates}, with the forward rates having a positive $Q$
-value and the reverse having a negative $Q$ value.  We add screening to the
+value and the reverse having a negative $Q$ value.  We do not recompute the reverse
+rates via detailed balance or correct them with high-temperature partition functions---this will be considered in a future science-focused paper.  We add screening to the
 rates using the prescription from \citet{graboske:1973,alastuey:1978,itoh:1979}.

### 2022-06-01 simplified_sdc_intro/submit/sdc_strang_flowchart.py
submitted version

@@ -1,0 +1,114 @@
+import matplotlib
+import matplotlib.pyplot as plt
+import numpy as np
+
+matplotlib.rc('text', usetex=True) 
+matplotlib.rcParams.update({"font.size": 18})
+matplotlib.rcParams['text.latex.preamble']=r"""
+\usepackage{amsmath}
+\newcommand{\pluseq}{\mathrel{+}=}
+\newcommand{\Uc}{\boldsymbol{\mathcal{U}}}
+"""
+
+# first strang
+
+fig = plt.figure()
+ax = fig.add_subplot(111)
+ax.axis("off")
+
+
+dy = 1.0
+dx = 1.0
+
+eps = 0.075
+
+# top line
+
+y = 2*dy
+
+ax.arrow(0, y, dx, 0, head_width=0.075, length_includes_head=True,
+         linewidth=2.0, color="C0")
+
+ax.text(0-eps, y, r"$\Uc^n$", verticalalignment="center", horizontalalignment="right", color="C0")
+ax.text(dx+eps, y, r"$\Uc^\star$", verticalalignment="center", horizontalalignment="left", color="C0")
+
+ax.text(dx/2, y+eps, r"$\frac{d\Uc}{dt} = {\bf R}(\Uc)$", verticalalignment="bottom", horizontalalignment="center", color="C0")
+
+# middle line
+
+y = dy
+
+ax.arrow(0, y, 2*dx, 0, head_width=0.075, length_includes_head=True, linewidth=2.0, color="C1")
+
+ax.text(0-eps, y, r"$\Uc^\star$", verticalalignment="center", horizontalalignment="right", color="C1")
+ax.text(2*dx+eps, y, r"$\Uc^{n+1,\star}$", verticalalignment="center", horizontalalignment="left", color="C1")
+
+ax.text(dx, y+eps, r"$\Uc^\star \pluseq \boldsymbol{\mathcal{A}}(\Uc^{\star,n+1/2})$", verticalalignment="bottom", horizontalalignment="center", color="C1")
+
+# bottom line
+
+y = 0
+
+ax.arrow(dx, y, dx, 0, head_width=0.075, length_includes_head=True, linewidth=2.0, color="C0")
+
+ax.text(dx-eps, y, r"$\Uc^{n+1,\star}$", verticalalignment="center", horizontalalignment="right", color="C0")
[…]

### 2022-07-12 simplified_sdc_intro/paper.tex
first pass at referee

@@
 %\documentclass[times,modern]{aastex63}
-\documentclass[times,preprint]{aastex631}
+\documentclass[linenumbers,trackchanges]{aastex631}
 %\documentclass{aastex63}
@@
 simulation code in a modular way.  However, competition between the
-different physical processes can cause the coupling to breakdown.
+different physical processes can cause the coupling to \replaced{breakdown}{break down, since the reactions do not see the effects of the hydrodynamics and vice versa}.
 These splitting errors can lead to loss of accuracy and further time
@@
 These large changes in abundances (and large alternately positive and negative energy generation rates) can be a challenge for
-a code.  The easiest way to improve the coupling is to cut the
+\replaced{a code}{the integration method---it may take more steps than allowed, require a timestep below floating point accuracy, or fail to meet the tolerenaces}.
+The easiest way to improve the coupling is to cut the
 timestep (see, e.g., \citealt{couch:2015,rivas:2022}), but this can make simulations
@@
 
-In the Strang splitting~\citep{strang:1968} flavor of operator splitting, we first integrate the system with reactions terms only (no advection)
-over $\Delta t/2$, then integrate the advection terms only (no reactions) over $\Delta t$,
-and finally integrate the reaction terms only over $\Delta t/2$.  This flow is illustrated
-in Figure~\ref{fig:strang_flowchart}.
+In the Strang splitting~\citep{strang:1968} flavor of operator
+splitting, we first integrate the system with reactions terms only (no
+advection) over $\Delta t/2$, then integrate the advection terms only
+(no reactions) over $\Delta t$, and finally integrate the reaction
+terms only over $\Delta t/2$.  \added{The staggering of the reactive update
+means the hydro implicitly sees a time-centgered reactive source, making
+the update second-order accurate in time.}  This flow is illustrated in
+Figure~\ref{fig:strang_flowchart}.
 
@@
 energy / temperature equation during the reaction step) in
-\citet{strang_rnaas} and showed that the above formulation gets second
+\citet{strang_rnaas} and showed that the above formulation \replaced{gets}{achieved} second
 order convergence and that in particular, integrating an energy / temperature
@@
 
-We begin with the general form of a multi-implicit SDC correction equation with two physical processes, as originally presented in \cite{bourlioux2003high}.
+We begin with the general form of a multi-implicit SDC correction equation with two physical processes\deleted{, as originally presented in} \citep{bourlioux2003high}.
 Using $(k)$ to denote the iterate, we have
@@
 
-    \item Predict $\qb$ to the interfaces at $t^{n+1/2}$ using the CTU PPM
-      method.  The source terms, $\Sq$, used in the prediction are:
+    \item Predict $\qb$ to the interfaces at $t^{n+1/2}$ using the CTU
+      PPM method\added{---this involves taking the cell-centered
+        primitive state and tracking all of the information that can
+        reach the interface over a timestep, following the procedure
+        from \citet{ppmunsplit,millercolella:2002}}.  The source
+      terms, $\Sq$, used in the prediction are:
       \begin{equation}
@@
     Our integrator also needs the Jacobian of the system, in terms of
-    $\Uc^\prime$.  This is different than the form of the
-    Jacobian usually used in reaction networks.  We describe the form
-    of the Jacobian in Appendix \ref{sec:app:jacobian}.
+    $\Uc^\prime$.  This is different than the form of the Jacobian
+    usually used in reaction networks \added{(we depend on $e$ instead of
+    $T$)}.  We describe the form of the Jacobian in Appendix
+    \ref{sec:app:jacobian}.
 
@@
 We use a modified version of the VODE~\citep{vode} integrator,
-designed for stiff ODEs.  Our version has been ported to C++, as a
-header-only library enabling it to run on GPUs.  We have also added
+designed for stiff ODEs.  Our version has been ported to \replaced{C++, as a
+header-only library enabling it to run on GPUs}{modern C++ and designed to run on GPUs}.  We have also added
 checks to the timestep rejection logic that help prevent VODE from
@@
 
-Tables \ref{table:react_converge_strang} and \ref{table:react_converge_sdc} show
-the Strang splitting and simplified-SDC convergence respectively.  The convergence rate 
-is computed by taking log-base-2 of the ratio of adjacent errors.  A rate of 2.0
-indicates second order convergence.  We see that both methods converge nearly
-second order, and agree well with each other.  This demonstrates that both
-methods are working as designed.
+Tables \ref{table:react_converge_strang} and
+\ref{table:react_converge_sdc} show the Strang splitting and
+simplified-SDC convergence respectively.
+\added{We ran each integration method on a domain with $64^2$, $128^2$, $256^2$, $512^2$ zones.
+An error is defined between successive resolutions by coarsening the higher resolution run and taking the L1 norm of the zone-by-zone difference of the two simulations---for example, $\epsilon_{64\rightarrow 128}$ is the error between the $64^2$ and $128^2$ simulations.}  \deleted{The convergence rate is
+computed by taking log-base-2 of the ratio of adjacent errors.}  \added{The convergence
+rate from $\epsilon_{64\rightarrow 128}$ to $\epsilon_{128\rightarrow 256}$ is computed as $\log_2 \epsilon_{64\rightarrow 128}/\epsilon_{128\rightarrow 256}$ and is given in the column between those errors, and likewise for the higher-resolution case.}  A rate
+of 2.0 indicates second order convergence.  We see that both methods
+\replaced{converge nearly second order, and agree well with each
+  other}{exhibit convergence at roughly second-order and agree well.}.
+This demonstrates that both methods are working as designed.
 

### 2022-07-12 simplified_sdc_intro/submit/sdc_strang_flowchart.py

@@ -1,114 +1,0 @@
-import matplotlib
-import matplotlib.pyplot as plt
-import numpy as np
-
-matplotlib.rc('text', usetex=True) 
-matplotlib.rcParams.update({"font.size": 18})
-matplotlib.rcParams['text.latex.preamble']=r"""
-\usepackage{amsmath}
-\newcommand{\pluseq}{\mathrel{+}=}
-\newcommand{\Uc}{\boldsymbol{\mathcal{U}}}
-"""
-
-# first strang
-
-fig = plt.figure()
-ax = fig.add_subplot(111)
-ax.axis("off")
-
-
-dy = 1.0
-dx = 1.0
-
-eps = 0.075
-
-# top line
-
-y = 2*dy
-
-ax.arrow(0, y, dx, 0, head_width=0.075, length_includes_head=True,
-         linewidth=2.0, color="C0")
-
-ax.text(0-eps, y, r"$\Uc^n$", verticalalignment="center", horizontalalignment="right", color="C0")
-ax.text(dx+eps, y, r"$\Uc^\star$", verticalalignment="center", horizontalalignment="left", color="C0")
-
-ax.text(dx/2, y+eps, r"$\frac{d\Uc}{dt} = {\bf R}(\Uc)$", verticalalignment="bottom", horizontalalignment="center", color="C0")
-
-# middle line
-
-y = dy
-
-ax.arrow(0, y, 2*dx, 0, head_width=0.075, length_includes_head=True, linewidth=2.0, color="C1")
-
-ax.text(0-eps, y, r"$\Uc^\star$", verticalalignment="center", horizontalalignment="right", color="C1")
-ax.text(2*dx+eps, y, r"$\Uc^{n+1,\star}$", verticalalignment="center", horizontalalignment="left", color="C1")
-
-ax.text(dx, y+eps, r"$\Uc^\star \pluseq \boldsymbol{\mathcal{A}}(\Uc^{\star,n+1/2})$", verticalalignment="bottom", horizontalalignment="center", color="C1")
-
-# bottom line
-
-y = 0
-
-ax.arrow(dx, y, dx, 0, head_width=0.075, length_includes_head=True, linewidth=2.0, color="C0")
-
-ax.text(dx-eps, y, r"$\Uc^{n+1,\star}$", verticalalignment="center", horizontalalignment="right", color="C0")
[…]

### 2022-07-15 simplified_sdc_intro/paper.tex
update report

@@
 simulation code in a modular way.  However, competition between the
-different physical processes can cause the coupling to \replaced{breakdown}{break down, since the reactions do not see the effects of the hydrodynamics and vice versa}.
+different physical processes can cause the coupling to \replaced{breakdown}{break down, since the reactions do not directly incorporate the effects of the hydrodynamics and vice versa}.
 These splitting errors can lead to loss of accuracy and further time
@@
 These large changes in abundances (and large alternately positive and negative energy generation rates) can be a challenge for
-\replaced{a code}{the integration method---it may take more steps than allowed, require a timestep below floating point accuracy, or fail to meet the tolerenaces}.
+\replaced{a code}{the integration method---it may take more steps than allowed by the ODE integrator, require a timestep below floating point accuracy, or fail to meet the tolerenaces}.
 The easiest way to improve the coupling is to cut the
@@
 terms only over $\Delta t/2$.  \added{The staggering of the reactive update
-means the hydro implicitly sees a time-centgered reactive source, making
+means the hydro implicitly sees a time-centered reactive source, making
 the update second-order accurate in time.}  This flow is illustrated in
@@
       corrections are made.  Both are second-order accurate.  For the
-      gravitational sources, we do those in the normal predictor.
+      gravitational sources, we do those in the normal predictor\added{, consistent
+      with the formulation in \citet{ppm}.}
       However, for the reactive sources, we found that it is most

### 2022-07-18 simplified_sdc_intro/paper.tex

@@
 These large changes in abundances (and large alternately positive and negative energy generation rates) can be a challenge for
-\replaced{a code}{the integration method---it may take more steps than allowed by the ODE integrator, require a timestep below floating point accuracy, or fail to meet the tolerenaces}.
+\replaced{a code}{the integration method---it may take more steps than allowed by the ODE integrator, require a timestep below floating point accuracy, or fail to meet the tolerances}.
 The easiest way to improve the coupling is to cut the

### 2022-07-21 simplified_sdc_intro/paper.tex
add new plot

@@
 equilibrium at high temperatures when the reverse rates can be
-important.
+important.  \added{To highlight this, Figure~\ref{fig:strang_steps} shows the nuclear
+energy generation rate in a highly-zoomed region around the detonation over four
+consecutive coarse timesteps.  We see that on the detonation front itself, the
+sign of the energy generation rate changes from one step to the next in many zones.}
 
@@
 strang\_subch2\_cfl0.2}) at 0.1 s.}
+\end{figure}
+
+\begin{figure}
+\centering
+\plotone{strang_step_diff_summary_plot}
+\caption{\label{fig:strang_steps} Energy generation rate for the
+  Strang CFL = 0.2 run with normal tolerances ({\tt
+    strang\_subch2\_cfl0.2}) shown at 4 consecutive timesteps around
+  $t = 0.1~\mathrm{s}$ zoomed in at the inward propagating detonation.}
 \end{figure}

### 2022-11-09 retries/paper.tex
Check in template for retry paper

--- 2022-05-10 simplified_sdc_intro/paper.tex
@@
-%\documentclass[times,modern]{aastex63}
-\documentclass[times,preprint]{aastex63}
-%\documentclass{aastex63}
+\documentclass[times,modern]{elsarticle}
 
@@
 
+\newcommand{\Sb}{{\bf S}}
+
 \newcommand{\sfrac}[2]{\mathchoice%
@@
 \newcommand{\castro}{{\sf Castro}}
-\newcommand{\maestroex}{{\sf MAESTROeX}}
-\newcommand{\pynucastro}{{\sf pynucastro}}
-\newcommand{\flash}{{\sf Flash}}
 \newcommand{\amrex}{{\sf AMReX}}
 
-\newcommand{\isot}[2]{$^{#2}\mathrm{#1}$}
-\newcommand{\isotm}[2]{{}^{#2}\mathrm{#1}}
-
-\newcommand{\gcc}{\mathrm{g~cm^{-3} }}
-\newcommand{\cms}{\mathrm{cm~s^{-1} }}
-
-\newcommand{\nablab}{{\mathbf{\nabla}}}
-\newcommand{\Ub}{\mathbf{U}}
-\newcommand{\gb}{\mathbf{g}}
-\newcommand{\omegadot}{\dot{\omega}}
-\newcommand{\Sdot}{\dot{S}}
-\newcommand{\ddx}[1]{{\frac{{\partial#1}}{\partial x}}}
-\newcommand{\ddxs}[1]{{\frac{{\partial}}{\partial x}}#1}
-\newcommand{\ddt}[1]{{\frac{{\partial#1}}{\partial t}}}
-\newcommand{\odt}[1]{{\frac{{d#1}}{dt}}}
-\newcommand{\divg}[1]{{\nablab \cdot \left (#1\right)}}
-\newcommand{\dedr}{\left . {\partial{}e}/{\partial\rho}\right |_{T, X_k}}
-\newcommand{\dedrd}{\left . \frac{\partial{}e}{\partial\rho}\right |_{T, X_k}}
-\newcommand{\dedX}{\left . {\partial{}e}/{\partial{}X_k} \right |_{\rho, T}}
-\newcommand{\dedXd}{\left . \frac{\partial{}e}{\partial{}X_k} \right |_{\rho, T, X_{j,j\ne k}}}
-\newcommand{\dedT}{\left . {\partial{}e}/{\partial{}T} \right |_{\rho,X_k}}
-\newcommand{\dedTd}{\left . \frac{\partial{}e}{\partial{}T} \right |_{\rho,X_k}}
-
-\newcommand{\Ic}{{\boldsymbol{\mathcal{I}}}}
-\newcommand{\Ics}{{\mathcal{I}}}
-\newcommand{\smax}{{s_\mathrm{max}}}
-\newcommand{\kth}{k_\mathrm{th}}
 \usepackage{bm}
-
-\newcommand{\Uc}{{\,\bm{\mathcal{U}}}}
-\newcommand{\Fb}{\mathbf{F}}
-\newcommand{\Sc}{\mathbf{S}}
-
-\newcommand{\xv}{{(x)}}
-\newcommand{\yv}{{(y)}}
-\newcommand{\zv}{{(z)}}
-
-\newcommand{\ex}{{\bf e}_x}
-\newcommand{\ey}{{\bf e}_y}
-\newcommand{\ez}{{\bf e}_z}
-
-\newcommand{\Ab}{{\bf A}}
-\newcommand{\Sq}{{\bf S}_\qb}
-\newcommand{\Sqhydro}{{\Sq^{\mathrm{hydro}}}}
-\newcommand{\qb}{{\bf q}}
-
-\newcommand{\Shydro}{{{\bf H}}}
-\newcommand{\Hb}{{\bf H}}
-\newcommand{\Rb}{{\bf R}}
-\newcommand{\Rq}{{\bf R}}
-\newcommand{\Adv}[1]{{\left [\boldsymbol{\mathcal{A}} \left(#1\right)\right]}}
-\newcommand{\Advt}[1]{{\left [\boldsymbol{\mathcal{\tilde{A}}} \left(#1\right)\right]}}
-\newcommand{\Advss}[1]{{\left [{\mathcal{{A}}} \left(#1\right)\right]}}
-\newcommand{\Advsst}[1]{{\left [{\mathcal{\tilde{A}}} \left(#1\right)\right]}}
-\newcommand{\Advs}[1]{\boldsymbol{\mathcal{A}} \left(#1\right)}
-
-\newcommand{\avg}[1]{{\left \langle #1 \right \rangle}}
-
-\newcommand{\nse}[1]{{\mathtt{NSE}( #1 )}}
-
-\newcommand{\rhonse}{{\rho_\mathrm{nse}}}
-\newcommand{\tnse}{{T_\mathrm{nse}}}
-\newcommand{\Anse}{{A_\mathrm{nse}}}
-\newcommand{\Bnse}{{B_\mathrm{nse}}}
-\newcommand{\Cnse}{{C_\mathrm{nse}}}
-
-\newcommand{\out}{{\rm out}}
-\newcommand{\inp}{{\rm in}}
-
-\newcommand{\atolspec}{{\epsilon_\mathrm{abs,spec}}}
-\newcommand{\rtolspec}{{\epsilon_\mathrm{rel,spec}}}
-\newcommand{\atolener}{{\epsilon_\mathrm{abs,ener}}}
-\newcommand{\rtolener}{{\epsilon_\mathrm{rel,ener}}}
 
@@
 %======================================================================
-%\title{A Simplified Spectral Deferred Correction Method for Coupling Hydrodynamics with Reaction Networks and Nuclear Statistical Equilibrium}
-\title{An Improved Method for Coupling Hydrodynamics with Astrophysical Reaction Networks}
+\title{Timestep Retries for Robust Simulations}
 
-%\shorttitle{A Simplified SDC Method}
-\shorttitle{Hydro/Reaction Coupling with NSE}
+\author{Max P. Katz}
+\author{Michael Zingale}
 
-\shortauthors{}
-
-\author[0000-0001-8401-030X]{M.~Zingale}
-\affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
-  Stony Brook, NY 11794-3800}
-
-\author[0000-0003-0439-4556]{M.~P.~Katz}
-\affiliation{Nvidia Corp}
-
-\author[0000-0003-1791-0265]{A.~J.~Nonaka}
-\affiliation{Center for Computational Sciences and Engineering, Lawrence Berkeley National Laboratory, Berkeley, CA  94720}
-
-\author[0000-0002-1530-781X]{A.~Harpole}
-\affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
-  Stony Brook, NY 11794-3800}
-
-\author{M.~Rasmussen}
-\affiliation{Utah State University, Logan UT 84322}
-
-\correspondingauthor{Michael Zingale}
-\email{michael.zingale@stonybrook.edu}
+\maketitle
 
@@
 \begin{abstract}
-Reacting astrophysical flows can be challenging to model because of
-the difficulty in accurately coupling hydrodynamics and reactions.
-This can be particularly acute during explosive burning or at high
-temperatures where nuclear statistical equilibrium is established.  We
-develop a simplified approach based on the ideas of spectral deferred
-corrections (SDC) coupling of explicit hydrodynamics and stiff
-reaction sources as an alternative to operator splitting or the more
-comprehensive SDC approach we demonstrated previously.  We apply the
-new method to a double detonation problem with a moderately-sized
-astrophysical nuclear reaction network and explore the timestep size
-and reaction network tolerances to show that the simplified-SDC
-approach compares favorably with traditional Strang operator
-splitting.  This is all done in the framework of the
-\castro\ hydrodynamics code, and all algorithm implementations are
-freely available.
 \end{abstract}
-
-\keywords{hydrodynamics---methods: numerical}
 
@@
 
-Modeling astrophysical reacting flows can be challenging because of
-the disparity better the nuclear and hydrodynamics timescales.
-Reaction networks tend to be stiff, requiring implicit integration
-techniques to stably integrate the system~\citep{BYRNE19871}.  In contrast, compressible
-hydrodynamics flows are limited by the (often much longer)
-sound-crossing time over a computational cell and can be solved using
-explicit time integration. Traditional methods of coupling
-hydrodynamics and reactions used in astrophysics use operator
-splitting---each physical process acts on the output of the previous
-process in alternating fashion.  This makes it easy to use different
-time-integration methods for the different physics, and to build a
-simulation code in a modular way.  However, competition between the
-different physical processes can cause the coupling to breakdown.
-These splitting errors can lead to loss of accuracy and further time
-step limitations.
-
-A particularly difficult phase of evolution to model is the nuclear
-statistical equilibrium that sets in for temperatures in excess of
-$\mbox{few} \times 10^9~\mathrm{K}$.  Physically, the forward and
-reverse rates of reaction should balance leading to an equilibrium.
-With operator splitting, an NSE region will have a large positive flow
-through the network in a zone in one step followed by a large negative
-flow over the next timestep, as the code struggles to produce an
-equilibrium.  These large changes in abundances (and large alternately
-positive and negative energy generation rates) can be a challenge for
-a code.  The easiest way to improve the coupling is to cut the
-timestep (see, e.g., \citealt{couch:2015}), but this can make simulations
-prohibitively expensive.  Sometimes the burning is simply halted on a
-zone-by-zone basis when NSE conditions are reached (e.g., as in
-\citealt{hedet}).  The
-present work focuses on networks alone, but in a follow-on paper we
-will explore hybrid burning models consisting of networks and tables for
-nuclear statistical equilibrium.
-
-The \castro\ hydrodynamics code \citep{castro,castro_joss} is used for
-all of our numerical experiments.  \castro\ is a compressible
-(magneto-, radiation-) hydrodynamics code built on the AMReX adaptive
-mesh refinement (AMR) framework \citep{amrex_joss}.  \castro\ has been
-designed to be performance portable and runs on massively parallel CPU
-and GPU architectures \citep{castro_gpu}.  For hydrodynamics, the
-corner transport upwind (CTU; \citealt{ppmunsplit}) method with the
-piecewise parabolic method (PPM; \citealt{ppm,millercolella:2002}) is
-used.  \castro\ includes self-gravity, rotation, arbitrary equations
-of state and reaction networks, and has been used for modeling X-ray
-bursts and different models of thermonuclear, core-collapse, and
-pair-instability supernovae.  Recently in \citet{castro_sdc}, we
-developed second- and fourth-order accurate methods in space and time
-for coupling hydrodynamics and nuclear reaction networks based on
-spectral deferred corrections (SDC) methods, and demonstrated the
-method in a variety of test problems.
-
-The time-integration approach presented here is considerably simpler
-than the SDC method of \citet{castro_sdc}, but allows us to reuse the
-main CTU \MarginPar{any cites to CCSE papers?}  hydrodynamics
-construction and a largely similar ODE integration scheme, making this
-method easier to add to existing simulation codes.  Furthermore, it
-also extends to adaptive mesh refinement with subcycling in a
-straightforward manner, avoiding the complications described in
-\citep{mccorquodalecolella} needed to fill ghost cells when using
-method-of-lines integration.  However, it is restricted to
-second-order accuracy in time overall.  We term this algorithm the
-``simplified-SDC method''.  In this paper we describe the overall
-method and demonstrate it on a test problem and astrophysical
-simulation using \castro.  All of the code to reproduce the results in
-this paper are in the \castro\ github
-repository\footnote{\url{https://github.com/amrex-astro/Castro/}}.
-
-\section{Numerical Methodology}
-
-We solve the Euler equations for compressible, reacting flow.  For ease
-of exposition we describe the one-dimensional case;
-multidimensional extensions are a straightforward modification to
-include in the CTU hydrodynamics scheme.  Our conserved variables are
-\begin{equation}
-  \Uc = \left ( \begin{array}{c}
-           \rho \\
-           \rho X_k \\
-           \rho u \\
-           \rho E \\
-           \rho e \end{array}\right )
-\end{equation}
-where $\rho$ is the mass density, $u$ is velocity, $E$ is specific
-total energy, $p$ is the pressure, and we carry nuclear species mass
-fractions, $X_k$.  The specific total
-energy relates to the specific internal energy, $e$, as $E = e + u^2/2$,
-and we also separately evolve $e$, as part of a dual
-energy formulation (see \citealt{bryan:1995,wdmergerI}).
- The mass fractions are constrained to sum to 1, $\sum_k X_k = 1$.
-Defining the hydrodynamical fluxes:
-\begin{equation}
-  \Fb(\Uc) = \left ( \begin{array}{c}
-         \rho u \\
-         \rho X_k u \\
-         \rho u^2 p \\
-         (\rho E + p) u \\
-         \rho e u \end{array}\right )
-\end{equation}
-We can write the system in conservative form for all state variables aside from $(\rho e)$ as:
-\begin{equation}
-  \ddt{\Uc} + \ddx{\Fb(\Uc)} = \Sc(\Uc)
-\end{equation}
-where for the special case of $(\rho e)$, we have an additional ``pdV'' term:
-\begin{equation}
-\ddt{(\rho e)} + \ddx{F(\rho e)} + p \ddx{u} = S(\rho e)
-\end{equation}
-
-We'll split the source term, $\Sc(\Uc)$ into pure hydrodynamic and reactive parts:
-\begin{equation}
-  \label{eq:cons_sources}
-  \Hb(\Uc) = \left ( \begin{array}{c}
-    0 \\
-    0 \\
-    \rho g \\
-    \rho u g \\
-    0 \end{array} \right )
-  \qquad
-  \Rb(\Uc) = \left ( \begin{array}{c}
-     0 \\
-     \rho \omegadot(X_k) \\
-     0 \\
-     \rho \dot{S} \\
-     \rho \dot{S}
-  \end{array} \right )
-\end{equation}
-with
-\begin{equation}
-  \Sc(\Uc) = \Hb(\Uc) + \Rb(\Uc).
-\end{equation}
-Here, gravity, represented by the gravitational acceleration $g$
-appears as a pure hydrodynamical source term.  From reactions,
-$\omegadot(X_k)$ is the creation rate for species $k$
- and $\dot{S}$ is the energy generation rate per unit
-mass.  We note that the internal energy ``pdV'' work is not treated as
-a source term but is instead constructed with the hydrodynamical
-fluxes are computed in the CTU method.
-Aside from the $p\partial u/\partial x$ (``pdV'') term in the internal
-energy equation, this system is in conservative form with source
-terms.
-
-We'll define the advective terms with hydrodynamic sources, $\Advs{\Uc}$ as
-\begin{equation}
-\Advs{\Uc} = -\ddx{\Fb(\Uc)} + \Hb(\Uc)
-\end{equation}
-for the general case, with the $(\rho e)$ component again having the extra
-``pdV'' term:
-\begin{equation}
-\mathcal{A}(\rho e) = -\ddx{F(\rho e)} -p \ddx{u} + H(\rho e)
-\end{equation}
-
-To close the system, we need the equation of state of the form:
-\begin{equation}
-p = p(\rho, X_k, e)
-\end{equation}
-
-Sometimes it is preferable to work with the primitive variables,
-\begin{equation}
-\qb = \left ( \begin{array}{c}
-  \rho \\
-  X_k \\
-  u \\
-  p \\
-  (\rho e) \\
-\end{array} \right )
-\end{equation}
-Here, the system appears
-as:
-\begin{equation}
-\qb_t + \Ab^\xv(\qb) \qb_x  = \Sc(\qb)
-\end{equation}
-with the matrix $\Ab^\xv$ giving the coefficients of the spatial derivatives
-of the primitive variables:
-\begin{equation}
-\Ab^\xv(\qb) = \left ( \begin{array}{ccccc}
-    u & 0 &  \rho & 0 & 0 \\
-    0 & u &  0    & 0 & 0 \\
-    0 & 0 &  u    & 1/\rho & 0 \\
-    0 & 0 &  \Gamma_1 p & u & 0 \\
-    0 & 0 &  \rho h & 0 & u
-  \end{array} \right )
-\end{equation}
-where $h$ is the specific enthalpy and $\Gamma_1$ is an adiabatic index,
-$\Gamma_1 = d\log p/d\log\rho|_s$ at constant entropy.
-The CTU+PPM algorithm uses the characteristic wave structure of $\Ab$
-to collect the information that makes it to an interface over a timestep
-in order to compute the fluxes through the interface.
-Note, the primitive state has two thermodynamic quantities, $p$
-and $(\rho e)$, to more efficiently handle the general equation of
-state in the Riemann solver, as described in \citet{castro}, but
-alternate formulations are possible \citep{colellaglaz:1985}.
-The source term vector, $\Sc(\qb)$, can again be decomposed into hydrodynamic
-sources (now in terms of the primitive variables) and reaction terms,
-\begin{equation}
-  \Sc(\qb) = \Hb(\qb) + \Rb(\qb)
-\end{equation}
-with
-\begin{equation}
-\label{eq:prim_sources}
-\Hb(\qb) = \left ( \begin{array}{c}
-     0 \\
-     0 \\
-     g \\
-     0 \\
-     0 \\
-   \end{array} \right )
-\qquad
-\Rb(\qb) = \left ( \begin{array}{c}
-     0 \\
-     \omegadot(X_k) \\
-     0 \\
-     \Gamma_1 p \sigma \Sdot \\
-     \rho \Sdot
-   \end{array} \right )
-\end{equation}
-where
-\begin{equation}
-\sigma \equiv \frac{\partial p/\partial T |_\rho}{\rho c_p \partial p/\partial \rho |_T}
-\end{equation}
-and $c_p$ is the specific heat at constant pressure, $c_p = \partial
-h/\partial T |_p$.  A derivation of this source for the pressure
-equation can be found in \cite{ABNZ:III}.  We note that this source is
-algebraically identical to that shown in Eq.~25 of \cite{castro}.
-
-The CTU+PPM method for hydrodynamics is second-order accurate in space
-and time.  We want to couple the reactive sources to the hydrodynamics
-to be second-order in time as well.  As discussed above, nuclear
-reaction sources are stiff, and need to be integrated using implicit
-methods for stabilty.  Operator splitting (e.g., Strang) is
-traditionally employed here, and is used as a benchmark for comparison
-in this paper.  We'll discuss this traditional approach next before moving
-on to our new time-coupling method.
-
-\subsection{Strang Splitting}
-
-In the Strang splitting~\cite{strang:1968} flavor of operator splitting, we first integrate the system with reactions terms only (no advection)
-over $\Delta t/2$, then integrate the advection terms only (no reactions) over $\Delta t$,
-and finally integrate the reaction terms only over $\Delta t/2$.  This flow is illustrated
-in Figure~\ref{fig:strang_flowchart}.
-
-\begin{figure}[t]
-  \plotone{sdc_strang_flowchart_I}
-  \caption{\label{fig:strang_flowchart} Flowchart of the operator split,
-    Strang, time integration.  We first react for $\Delta t/2$, then
-    advect for $\Delta t$, and finally react for $\Delta t/2$.  Each
-    process uses the state left behind from the previous process.}
-  \end{figure}
-
-In the absence of advective terms, our reaction system appears as just
-$d\Uc/dt = \Rb(\Uc)$, or:
-\begin{align}
-\odt{\rho} & = 0 \\
-\odt{(\rho X_k)} &= \rho \omegadot(X_k) \\
-\odt{(\rho u)} &= 0 \\
-\odt{(\rho E)} &= \rho \Sdot
-\end{align}
-We can write the energy equation as:
-\begin{equation}
-\odt{(\rho E)} = \odt{(\rho e)} + \odt{K} = \rho \Sdot
-\end{equation}
-where $K$ is the kinetic energy, $K = \rho |u|^2/2$.  Since the density and velocity
-are unchanged by reactions (when Strang splitting), our energy equation becomes:
-\begin{equation}
-\label{eq:strang:e}
-\odt{(\rho e)} = \rho \odt{e} = \rho \Sdot
-\end{equation}
-The reaction rates are typically expressed as $\omegadot_k(\rho, T, X_k)$
-when we evolve this system, which requires us to get temperature from
-the equation of state each
-time we need to evaluate the reactive terms.
-
-We also typically integrate mass fraction itself, instead of partial
-densities:
-\begin{equation}
-\label{eq:strang:X}
-\odt{X_k} = \omegadot_k
-\end{equation}
-We integrate Equations (\ref{eq:strang:e}) and (\ref{eq:strang:X})
-using an implicit ODE solver described below.
-
-We explored this and other approaches (including not evolving an
-energy / temperature equation during the reaction step) in
-\citet{strang_rnaas} and showed that the above formulation gets second
-order convergence and that in particular, integrating an energy / temperature
-equation is needed to get second order accuracy (see also
-\citealt{muller:1986} for a discussion on stability).  We note however
-that some astrophysical simulation codes only evolve the species
-equations.  For very strong reactions, when using Strang splitting the
-state can drift significantly off of the smooth solution to the
-coupled reactive hydrodynamics equations, as shown graphically in
-\citet{astronum:2018} (using an earlier version of the present
-algorithm).  A variation on Strang splitting called (re-)balanced
-splitting was developed in \citet{speth:2013}, but we do not explore
-that here. \MarginPar{keep this sentence?}
-
-\subsection{Timestep Limiters and Retry Mechanism}
-
-Since this method is based off of the CTU hydrodynamics scheme, it
-benefits from the larger timestep that method can take (when done with
-full corner coupling, the advective CFL condition is unity) as
-compared to a method-of-lines approach (see \citealt{ppmunsplit}).  In
-addition to the standard CFL timestep limiter for explicit
-hydrodynamics, timestep limiters based on
-the energy generation or abundance changes over a timestep, such as
-those introduced in \cite{prometheus}, are often used.  An energy limiter takes the
-form of either:
-\begin{equation}
-\label{eq:dt:nuce_I}
-\Delta t \le f_e\, \min_{i} \left\{\frac{e^n_{i}}{|\dot{S}_{i}^n|}\right\} \\
-\end{equation}
-or
-\begin{equation}
-\label{eq:dt:nuce_II}
-\Delta t \le f_e\, \min_{i} \left\{ \frac{e^n_i}{|\int \dot{S}_i(t) dt|} \right\}
-\end{equation}
-where $i$ is the zone index and $f_e$ is a parameter used to control
-the allowed change (e.g., $f_e = 0.5$ will limit the step such that
-reactions can only increase the internal energy by 50\%).  The two
-formulations differ in that Eq.~\ref{eq:dt:nuce_I} uses the
-instantaneous evaluation of the energy generation rate at the start of
-the timestep while Eq.~\ref{eq:dt:nuce_II} uses the integral of the
-energy generation rate over the entire step (or last Strang half).
-Usually these limiters are used in a reactive manner---the next step will see
-a smaller timestep based on what happened during the current step, but the
-current step is still accepted, even if it violated these conditions.  Our
-goal is to avoid the need for these limiters by doing the coupling of hydrodynamics
-and reactions better.
-
-\castro\ has the ability to reject a timestep if it detects a failure and retry
-with smaller timesteps (subcycling to make up the original required timestep).
-Among the conditions that can trigger this are density falling below zero during
-advection, the ODE integration failing to converge in the implicit solve (due to
-too many steps or the internal timestep falling below a minimum), or violation
-of one of the timestep limiters during the step.  This means that the timestep
-constraints are proactive instead of reactive.  The retry mechanism in \castro\
-works with both the Strang and simplified-SDC integration scheme.  Retrying the
-step is much more stringent that simply adjusting the next step, so we put a cap
-on the number of retries allowed, aborting if we exceed the cap. \MarginPar{describe in more detail?}
-
-\subsection{Spectral Deferred Corrections}
-
-The basic idea of spectral deferred corrections \MarginPar{ref} is to express the
-update as an integral and divide the time-update into a number of
-discrete time nodes.  The integral is then approximated using a
-quadrature rule over these time nodes and low order approximations are
-used to update the state from one time node to the next.  The method
-uses iteration to successively improve the solution and the ultimate
-accuracy is determined by the quadrature method used to evaluate the
-integral, which is done using an iteratively-lagged solution.  This is the
-classic approach that we implemented in Castro in \citet{castro_sdc}.
-
-Our simplified-SDC approach is based on the method described in \citet{SDC-old}
-and a similar (but unpublished) implementation in the \maestroex\ simulation
-code \citep{maestroex}. We again start by considering the update in integral
-form:
-\begin{equation}
-\Uc^{n+1} = \Uc^n + \int_t^{t+\Delta t} \left [ \Advs{\Uc} + \Rb(\Uc) \right ] dt,
-\end{equation}
-In this approach, we
-approximate the advective term
-piecewise constant in time, using the value at the midpoint in time,
-to achieve second-order accuracy, giving us:
-\begin{equation}
-\label{eq:integral:simplesdc}
-\Uc^{n+1} = \Uc^n + \int_t^{t+\Delta t} \left \{ \Adv{\Uc}^{n+1/2} + \Rb(\Uc) \right \} dt
-\end{equation}
-The \castro\ version of the simplified-SDC algorithm
-differs from the previous versions due to the need to do some operations
-on the conserved variable state and some on the primitive variable
-state.   The overall flow of the simplified-SDC algorithm is shown
-in Figure~\ref{sdc_flowchart}.
-
-\begin{figure}[t]
-\plotone{sdc_strang_flowchart_II}
-\caption{\label{fig:sdc_flowchart} Flowchart of the simplified-SDC
-  approach.  An iteratively-lagged advection term is used to update
-  the system to the new time, with each iteration improving the
-  coupling.}
-\end{figure}
-
-The basic time update algorithm proceeds as:
-
-\begin{itemize}
-
-\item {\em Initialization}
-
-\label{sec:initialization}
-
-  \begin{itemize}
-  \item We need an approximation of how much the reactions alone
-    changed the primitive variable state over the timestep, which we
-    will denote $\Ic_q$ (this is basically an approximation of
-    $\Rb(\qb)$).
-
-    Since we do not have any information about the current
-    timestep in the first iteration, we use the value from the last
-    iteration of the previous timestep:
-    \begin{equation}
-      \Ic^{n+1/2,(0)}_{\qb} = \Ic^{n-1/2,(\smax)}_{\qb}
-    \end{equation}
-  %% \item Solve the Poisson problem for the initial gravitational potential:
-  %%   \begin{equation}
-  %%     \nabla^2 \Phi^n = 4\pi G \rho^n
-  %%   \end{equation}
-
-  %% \item Set the first guess at the new time potential as
-  %%   $\Phi^{n+1,(0)} = \Phi^n$.
-
-  \end{itemize}
-
-\item {\em Iterate}
-
-  Iterate from $k = 1, \ldots, \smax$.  For second-order accuracy,
-  $\smax = 2$ is sufficient.  In addition to denoting the time-level
-  with a superscript (like $n$ or $n+1$), we'll use a second subscript
-  in parentheses to keep track of the iteration.  A single iteration, $k$,
-  starts with $\Uc^n$ and results in the new time-level state for that
-  iteration, $\Uc^{n+1,(k)}$.
-
-  \begin{itemize}
-  \item {\em Create the advective update term, $\Adv{\Uc}^{n+1/2,(k)}$}
-
-    \begin{itemize}
-    \item convert $\Uc \rightarrow \qb$.  This is an algebraic transformation,
-      but will require the EOS.
-
-    \item predict $\qb$ to the interfaces at $t^{n+1/2}$ using the CTU PPM
-      method.  The source terms, $\Sq$, used in the prediction are:
-      \begin{equation}
-        \Sc(\qb) = \Hb(\qb) + \Ic_\qb^{n+1/2,(k-1)}
-      \end{equation}
-      Here we use the iteratively lagged integrals of the primitive variable
-      terms accounting only for reactions, $\Ic_\qb^{n+1/2,(k-1)}$, as the
-      reactive source.  This is in contrast to Strang-splitting, where no
-      explicit reactive source terms are seen by the hydrodynamics update.
-%%       Any hydrodynamic source terms are time-centered
-%%       using the previous iteration:
-%% \MarginPar{Formally both $t^n$ or $t^{n+1/2}$ sources give you 2nd-order since they are CTU predictor source terms.  I've never found this type of iteration for this term to amount to any benefit}
-%%       \begin{equation}
-%%         \Sqhydro^{n+1/2} = \frac{1}{2} \left ( \Sqhydro^n + \Sqhydro^{n+1,(k-1)} \right )
-%%       \end{equation}
-
-      In the unsplit CTU method \citep{ppmunsplit}, the interface
-      states used for the final Riemann problem through the zone
-      interface consist of a normal predictor and a transverse flux
-      correction.  We can add the source terms either to the normal
-      predictor (for example, doing characteristic tracing as
-      described in \citealt{ppm}) or after all of the transverse flux
-      corrections are made.  Both are second-order accurate.  For the
-      hydrodynamics sources, we do those in the normal predictor.
-      However, for the reactive sources, we found that it is most
-      reliable to add them at the end of the interface state
-      construction, after the transverse flux corrections.  This is
-      because we want to ensure that sum over species of $\Ic_\qb$ is
-      zero, and characteristic tracing or the various flux corrections
-      may not preserve this.  We enforce that the species interface
-      states remain in $[0, 1]$ after adding the reactive source.
-
-    \item solve the Riemann problem at each interface to get a unique
-      conserved state on each interface, $\Uc^{n+1/2,(k)}_{i+1/2}$
-
-    \item construct the advective update terms, $\Advt{\Uc}^{n+1/2,(k)}_{i}$,
-      by first ignoring the hydrodynamics sources,
-      \begin{equation}
-        \Advt{\Uc}^{n+1/2,(k)}_{i} =
-          - \frac{\Fb^\xv(\Uc^{n+1/2,(k)}_{i+1/2}) - \Fb^\xv(\Uc^{n+1/2,(k)}_{i-1/2})}{\Delta x}
-      \end{equation}
-      for the general case, and
-      \begin{align}
-        \Advsst{\rho e}^{n+1/2,(k)}_{i} =
-          &- \frac{F\left((\rho e)^{n+1/2,(k)}_{i+1/2}\right) - F\left((\rho e)^{n+1/2,(k)}_{i-1/2}\right)}{\Delta x} \nonumber \\
-          &- \left (\frac{p_{i+1/2}^{n+1/2,(k)} + p_{i-1/2}^{n+1/2,(k)}}{2} \right )\frac{u_{i+1/2}^{n+1/2,(k)} - u_{i-1/2}^{n+1/2,(k)}}{\Delta x}
-      \end{align}
-      for $(\rho e)$.
-
-    Now the conservative hydrodynamics source terms are computed by first updating to the
-    new state with advection and the old-time source term applied for the full $\Delta t$ as:
-    \begin{equation}
-      \Uc^{\star\star} = \Uc^n + \Delta t \Advt{\Uc}^{n+1/2,(k)} + \Delta t \Hb(\Uc^n)
-    \end{equation}
-    %% first updating the density to the new state with advection only as:
-    %% \begin{equation}
-    %%   \rho^{\star\star} = \rho^n + \Delta t \Advt{\rho}^{n+1/2,(k)}
-    %% \end{equation}
-    %% then constructing the momentum source term:
-    %% \begin{equation}
-    %%   {\bf S}_{\rho\Ub}^{n+1/2} = \frac{1}{2} (\rho^n \gb^n + \rho^{\star\star} \gb^{n+1,(k-1)})
-    %% \end{equation}
-    %% Then updating the momentum with only advection as:
-    %% \begin{equation}
-    %%   (\rho \Ub)^{\star\star} = (\rho \Ub)^n + \Delta t \Advt{\rho\Ub}^{n+1/2,(k)} + \Delta t {\bf S}_{\rho\Ub}^{n+1/2}
-    %% \end{equation}
-    %% and finally constructing the energy source:
-    %% \begin{equation}
-    %%   S_{\rho E}^{n+1/2} = \frac{1}{2} \left [ (\rho\Ub)^n \cdot \gb^n + (\rho\Ub)^{\star\star} \cdot \gb^{n+1,(k-1)}
-    %%     \right ]
-    %% \end{equation}
-    We then evaluate the source terms with $\Uc^{\star\star}$ and
-    correct the advective term so that we have a time-centered
-    source.
-    The final advective update term is then\footnote{For a source like gravity, this update can be done first for $\rho$ and then define the new momentum source using $\rho^{\star\star}$ and likewise for energy}:
-    \begin{equation}
-      \Adv{\Uc}^{n+1/2,(k)}_{i} = \Advt{\Uc}^{n+1/2,(k)}_{i} +
-      \frac{\Delta t}{2} \left (\Shydro(\Uc^n) + \Shydro(\Uc^{\star\star}) \right )
-    \end{equation}
-
-    %% with
-    %% \begin{equation}
-    %%   \Shydro^{n+1/2} = \left ( \begin{array}{c}
-    %%                  0 \\ 0 \\ 0 \\
-    %%                 {\bf S}_{\rho\Ub}^{n+1/2} \cdot \ex \\
-    %%                 S_{\rho E}^{n+1/2} \end{array} \right )
-    %% \end{equation}
-
-    We note that this procedure works because the source terms here do
-    not depend on the outcome of the reactions.
-
-    \end{itemize}
-
-  \item {\em Update the System Using a Method of Lines Integration}
-
-    We update the state by doing the integral in equation
-    (\ref{eq:integral:simplesdc}).  Since we are approximating the
-    advective term as piecewise constant in time, we can simply use an
-    ODE integrator to integrate this, just as we do with the reaction
-    system in Strang splitting.  The difference here being that we are
-    integrating the conserved variables and the state sees the effect
-    of advection as we integrate the reactions.  So rather than use
-    $d\Uc/dt = \Rb(\Uc)$, the ODE form we use is
-    \begin{equation}
-      \odt{\Uc} = \Adv{\Uc}^{n+1/2,(k)} + \Rb(\Uc)
-    \end{equation}
-
-    Looking at the form of Eq.~\ref{eq:cons_sources}, we see that only
-    the species and energies have source terms.  The total and
-    internal energies both provide the same information, and
-    integrating both overconstrains the system, so we just integrate
-    the internal energy.  We define the subset of variables that are
-    directly integrated as:
-    \begin{equation}
-      \Uc^\prime = \left ( \begin{array}{c} \rho X_k \\ \rho e \end{array} \right )
-    \end{equation}
-    and we integrate
-    \begin{equation}
-      \odt{\Uc^\prime} = \Adv{\Uc^\prime}^{n+1/2,(k)} + \Rb(\Uc^\prime)
-    \end{equation}
-    This integration begins with $\Uc^{\prime,n}$ and results in $\Uc^{\prime,n+1,(k)}$.
-
-    We will need the density at times during the integration, which we construct as:
-    \begin{equation}
-      \rho(t) = \rho^n + \Advss{\rho}^{n+1/2,(k)} \, (t - t^n)
-    \end{equation}
-    As we are integrating this system we need to get the temperature,
-    $T$, for the rate evaluations.  We obtain this by directly from
-    internal energy, composition, and density using the equation of
-    state.
-
-    Our integrator also needs the Jacobian of the system, in terms of
-    the $\Uc^\prime$.  This is different than the form of the
-    Jacobian usually used in reaction networks.  We describe the form
-    of the Jacobian in appendix \ref{sec:app:jacobian}.
-
-    At the end of the integration, we can do the conservative update of momentum
-    and energy.  Momentum is straightforward, since there are no reactive sources:
-    \begin{equation}
-      (\rho u)^{n+1} = (\rho u)^n + \Delta t \Advss{\rho u}^{n+1/2,(k)}
-    \end{equation}
-    For total energy, we first need to isolate the reactive source for energy:
-    \begin{equation}
-      (\rho \Sdot)^{n+1/2} = \frac{(\rho e)^{n+1} - (\rho e)^n}{\Delta t} - \Advss{\rho e}^{n+1/2,(k)}
-    \end{equation}
-    Then the update is
-    \begin{equation}
-      (\rho E)^{n+1} = (\rho E)^n + \Delta t \Advss{\rho E}^{n+1/2,(k)} + \Delta t (\rho \Sdot)^{n+1/2}
-    \end{equation}
-
-  \item {\em Compute the Reactive Source Terms.}
-
-    We now seek the $\Ic$'s that capture the effect of just the
-    reaction sources on the state variables for the next iteration.
-    For the conserved quantities, these would simply be:
-    \begin{equation}
-      \Ic^{(k)}_{\Uc} = \frac{\Uc^{n+1,(k)} - \Uc^n}{\Delta t} - \Adv{\Uc}^{n+1/2,(k)}
-    \end{equation}
-    However, for our primitive variables, which are used in the
-    interface state prediction, we need to construct the required source terms more
-    carefully.  We want:
-    \begin{equation}
-      \label{eq:Iq}
-      \Ic^{(k)}_{\qb} = \frac{\qb^{n+1,(k)} - \qb^n}{\Delta t} - \Adv{\qb}^{n+1/2,(k)}
-    \end{equation}
-    but we need the advective update for $\qb$, which we have not
-    constructed.  We note that $\Ic^{(k)}_\qb$ is an approximation to the integral of
-    Eq.~\ref{eq:prim_sources} over the timestep.  Additionally, we cannot simply use the equation of
-    state on $\Ic^{(k)}_{\Uc}$ since this is a time-derivative and
-    does not represent a well-defined state in itself.  Instead, we
-    derive $\Ic^{(k)}_{\qb}$ via a multi-step process.  We first find
-    the conservative state as if it were updated only with advection:
-    \begin{equation}
-      \Uc^\star = \Uc^n + \Delta t \Adv{\Uc}^{n+1/2,(k)}
-    \end{equation}
-    and then construct the corresponding primitive variable state via an algebraic transform,
-    $\Uc^\star \rightarrow \qb^\star$.
-    This allows us to define the advective update for $\qb$ as:
-    \begin{equation}
-      \Adv{\qb}^{n+1/2,(k)} = \frac{\qb^\star - \qb^n}{\Delta t}
-    \end{equation}
-    Defining the primitive state corresponding to the fully-updated
-    conserved state via an algebraic transform, $\Uc^{n+1,(k)}
-    \rightarrow \qb^{n+1,(k)}$, we can construct $\Ic^{(k)}_{\qb}$ as given
-    in Equation (\ref{eq:Iq}).
-    Putting all of this together, we see:
-    \begin{equation}
-      \Ic^{(k)}_{\qb} = \frac{\qb^{n+1,(k)} - \qb^\star}{\Delta t}
-    \end{equation}
-
-  %% \item {\em Solve for the New Gravitational Potential.}
-
-  %%   We solve
-  %%   \begin{equation}
-  %%     \nabla^2 \Phi^{n+1,(k)} = 4\pi G \rho^{n+1,(k)}
-  %%   \end{equation}
-
-  \end{itemize}
-
-\end{itemize}
-
-This completes the update of a single iteration.
-
-\subsection{Stiff ODE Integrator}
-
-We use a modified version of the VODE~\citep{vode} integrator,
-designed for stiff ODEs.  Our version has been ported C++, as a
-header-only library enabling it to run on GPUs.  We have also added
-checks to the timestep rejection logic that help prevent VODE from
-wandering too far off the solution\footnote{This modified version of
-VODE is available in our Microphysics repo:
-\url{https://github.com/amrex-astro/Microphysics}.}  These have
-been shown empirically to improve the performance of VODE.
-
-VODE allows us to specific both relative and absolute tolerances
-for the species and energy during the integration.  They are combined
-into a weight that VODE uses to assess convergence of the step:
-\begin{equation}
-w_i = \epsilon_\mathrm{rel} |y_i| + \epsilon_\mathrm{abs}
-\end{equation}
-for integration variable $y_i$.  We will denote the species tolerances
-as $\rtolspec$ and $\atolspec$ and the energy tolerances as $\rtolener$ and
-$\atolener$.
-
-For Strang, we require that
-an individual mass fraction not change too much over a step:
-\begin{equation}
-\label{eq:vode_species_reject}
-\begin{array}{c}|X_k^{m+1}| < f |X_k^m| \\ |X_k^{m+1}| > \frac{1}{f} |X_k^m| \end{array} ~  \mbox{if}~ |X_k^m| > \eta \atolspec ~\mbox{and}~ |X_k^{m+1}| > \eta \atolspec
-\end{equation}
-Here, $m$ is the current VODE solution and $m+1$ is the potential
-new-time solution.  The parameter $f$ is the factor a mass fraction is
-allowed to change per VODE step.  We use $f = 4$ for the results here.
-The parameter $\eta$ allows us to only use this condition for species
-with mass fractions above a threshold.  We use $\eta = 1$ by default,
-but $\eta = f$ sometimes helps prevent VODE from worrying about
-species that dip below the threshold during integration, since VODE
-uses a norm over all of the integrated variables to compute the total
-error.  If these conditions are violated, then VODE will reject the
-timestep and retry with a smaller step.  We also require that the mass
-fractions are contained in $[0, 1]$ to a tolerance of $10^{-2}$.
-For SDC, we enforce these same constraints, but for
-Eq.~\ref{eq:vode_species_reject}, we only use the change from $m$ to $m+1$ due
-to reactions, by subtracting off the advective contribution over that substep.
-Finally, we enforce that:
-\begin{align}
-\rho^{m+1} &> 0 \\
-(\rho e)^{m+1} &> 0
-\end{align}
-
-Some of the rates in \citet{reaclib} increase by 100s of orders of
-magnitude at low temperature, presumably because the interpolant fit was done at
-higher temperatures.  Since VODE can take exploratory righthand side evaluations
-outside of the nominal time range, it can encounter these poorly-behaved rates
-and it can cause the integration to fail (by generating {\tt inf}s).  To prevent
-this behavior, we set a lower temperature limit of the reaction rates, below
-which we zero them out.  For the simulations in section~\ref{sec:dd}, we choose a cutoff of
-$5\times 10^7~\mathrm{K}$.
-
-For both Strang and simplified-SDC, we will use a numerical Jacobian that VODE
-computes internally (following the algorithm in \citealt{lsode}).  For
-completeness, we derive the analytic form of the Jacobian in Appendix
-\ref{sec:app:jacobian} for the simplified-SDC method. VODE employs
-Jacobian-caching, so the Jacobian does not need to be re-evaluated each time it
-is used.
-
-\section{Tests}
-
-\subsection{Reacting convergence test problem}
-
-\citet{castro_sdc} introduced a test problem for measuring convergence
-of a reacting hydrodynamic algorithm.  This is based on the acoustic
-pulse problem from \citet{mccorquodalecolella}, and was also used in
-\citet{strang_rnaas}.  A periodic domain with a uniform entropy is
-initialized with a pressure profile with a small perturbation.  This
-drives a low amplitude acoustic wave radially outward from the center.
-A simple reaction network with the triple-alpha and
-$\isotm{C}{12}(\alpha,\gamma)\isotm{O}{16}$ rates releases energy.  By
-running at different resolutions, with a fixed timestep that scales
-with resolution, we can compute the convergence rate.  In order to
-demonstrate second-order convergence, the problem must be smooth (so
-the slope limiters don't kick in severely), which means that the
-energy release cannot be so vigorous as to drive shocks.
-
-% EOS: /home/zingale/development/Microphysics/EOS/helmholtz
-% NETWORK: /home/zingale/development/Microphysics/networks/triple_alpha_plus_cago
-
-% Castro       git describe: 21.12-9-gd7a27ccd5
-% AMReX        git describe: 21.12-64-gfde72340e
-% Microphysics git describe: 21.12-12-g9595a34f
-
-\begin{deluxetable}{lllllll}
-\tablecaption{\label{table:react_converge_strang} Convergence ($L_1$ norm) for the reacting convergence problem using Strang splitting.}
-\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} &
-           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} &
-           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
-\startdata
- $\rho$                      & $2.794 \times 10^{18}$  & 2.044  & $6.777 \times 10^{17}$  & 2.554  & $1.154 \times 10^{17}$  \\
- $\rho u$                    & $6.796 \times 10^{26}$  & 2.448  & $1.245 \times 10^{26}$  & 2.889  & $1.681 \times 10^{25}$  \\
- $\rho v$                    & $6.796 \times 10^{26}$  & 2.448  & $1.245 \times 10^{26}$  & 2.889  & $1.681 \times 10^{25}$  \\
- $\rho E$                    & $2.451 \times 10^{35}$  & 2.351  & $4.803 \times 10^{34}$  & 2.742  & $7.179 \times 10^{33}$  \\
- $\rho e$                    & $2.261 \times 10^{35}$  & 2.320  & $4.526 \times 10^{34}$  & 2.821  & $6.403 \times 10^{33}$  \\
- $T$                         & $2.237 \times 10^{21}$  & 1.691  & $6.927 \times 10^{20}$  & 2.482  & $1.240 \times 10^{20}$  \\
- $\rho X(\isotm{He}{4})$     & $2.878 \times 10^{18}$  & 2.020  & $7.096 \times 10^{17}$  & 2.529  & $1.229 \times 10^{17}$  \\
- $\rho X(\isotm{C}{12})$     & $1.698 \times 10^{17}$  & 1.950  & $4.393 \times 10^{16}$  & 2.232  & $9.353 \times 10^{15}$  \\
- $\rho X(\isotm{O}{16})$     & $1.687 \times 10^{14}$  & 1.660  & $5.338 \times 10^{13}$  & 1.957  & $1.375 \times 10^{13}$  \\
- $\rho X(\isotm{Fe}{56})$    & $2.794 \times 10^{8}$   & 2.044  & $6.777 \times 10^{7}$   & 2.554  & $1.154 \times 10^{7}$   \\
-\enddata
-\end{deluxetable}
-
-\begin{deluxetable}{lllllll}
-\tablecaption{\label{table:react_converge_sdc} Convergence ($L_1$ norm) for the reacting convergence problem using SDC integration (2 iterations).}
-\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} &
-           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} &
-           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
-\startdata
- $\rho$                      & $2.784 \times 10^{18}$  & 2.048  & $6.734 \times 10^{17}$  & 2.558  & $1.143 \times 10^{17}$  \\
- $\rho u$                    & $6.779 \times 10^{26}$  & 2.447  & $1.243 \times 10^{26}$  & 2.895  & $1.671 \times 10^{25}$  \\
- $\rho v$                    & $6.779 \times 10^{26}$  & 2.447  & $1.243 \times 10^{26}$  & 2.895  & $1.671 \times 10^{25}$  \\
- $\rho E$                    & $2.450 \times 10^{35}$  & 2.353  & $4.795 \times 10^{34}$  & 2.737  & $7.193 \times 10^{33}$  \\
- $\rho e$                    & $2.256 \times 10^{35}$  & 2.320  & $4.518 \times 10^{34}$  & 2.817  & $6.414 \times 10^{33}$  \\
- $T$                         & $2.232 \times 10^{21}$  & 1.695  & $6.893 \times 10^{20}$  & 2.484  & $1.233 \times 10^{20}$  \\
- $\rho X(\isotm{He}{4})$     & $2.866 \times 10^{18}$  & 2.024  & $7.049 \times 10^{17}$  & 2.534  & $1.217 \times 10^{17}$  \\
- $\rho X(\isotm{C}{12})$     & $1.695 \times 10^{17}$  & 1.959  & $4.357 \times 10^{16}$  & 2.236  & $9.250 \times 10^{15}$  \\
- $\rho X(\isotm{O}{16})$     & $1.681 \times 10^{14}$  & 1.662  & $5.313 \times 10^{13}$  & 1.963  & $1.363 \times 10^{13}$  \\
- $\rho X(\isotm{Fe}{56})$    & $2.784 \times 10^{8}$   & 2.048  & $6.734 \times 10^{7}$   & 2.558  & $1.143 \times 10^{7}$   \\
-\enddata
-\end{deluxetable}
-
-Tables \ref{table:react_converge_strang} and \ref{table:react_converge_sdc} show
-the Strang and simplified-SDC convergence respectively.  The convergence rate is
-is computed by taking log-base-2 of the ratio of adjacent errors.  A rate of 2.0
-indicates second order convergence.  We see that both methods converge nearly
-second order, and agree well with each other.  This demonstrates that both
-methods are working as designed.
-
-\subsection{Sub-Chandra Double Detonation}
-\label{sec:dd}
-
-We next model a double detonation Type Ia supernova.  In this model, a
-detonation begins in the accreted helium layer on a sub-Chandrasekhar mass white
-dwarf and that detonation can either drive a compression wave into the core
-igniting a second detonation in the carbon / oxygen core, or burn through the
-interface at the base of the accreted layer producing an inward propagating
-carbon detonation (see, e.g., \citealt{fink:2007}).  Our goal here is not to
-explore the feasibility of the model or understand whether a detonation
-propagating through the He-C interface is physical, but rather to look at the
-coupling of the hydrodynamics and reactions.  We want to drive vigorous burning
-and directly compare Strang and simplified-SDC time integration on this problem.
-For that reason, we explore only a single model and start off with a rather
-large temperature perturbation in the He layer.
-
-We use a network with 28 nuclei and 107 rates that captures He burning with
-links for $(\alpha,p)(p,\gamma)$ reactions and the rates involving \isot{N}{14}
-to bypass $\isotm{C}{12}(\alpha,\gamma)\isotm{O}{16}$ as described in
-\cite{shenbildsten}.  The rates are taken from the ReacLib library
-\citep{reaclib} and the network is written by \pynucastro~\citep{pynucastro}
-directly in the C++ format that our code requires.  The details of the network
-are given in Appendix \ref{sec:app:reactionnet}. This moderate-sized network
-with many reaction pathways (forward and reverse) make it a good test for the
-coupling between hydro and reactions.
-
-The initial model is generated following the procedure described in
-\citep{subchandra}.  We use a $1.1~M_\odot$ carbon white dwarf with a
-$0.05~M_\odot$ He layer.  We seed the He layer with 1\% (by mass) of
-\isot{N}{14}.  The white dwarf is isothermal and the temperature ramps
-up at the base of the He layer and is then isentropic until the
-surface.  Figure~\ref{fig:subch_initial_model} shows the structure of
-the initial model.  We use 2d axisymmetric coordinates with a domain
-size of $1.024\times 10^9~\mathrm{cm} \times 2.048\times
-10^9~\mathrm{cm}$, a base grid of $256 \times 512$ zones and a single
-level of refinement (jumping by $2\times$), giving a fine-grid
-resolution of $20~\mathrm{km}$.  Gravity is modeled as a monopole.
-Our choice of initial perturbation has the detonation propagate inward
-into the carbon white dwarf.  We run the simulation for 0.1 s---this
-is enough time for the carbon detonation to be well developed.
-
-\begin{figure}[t]
-\centering
-\plotone{initial_model}
-\caption{\label{fig:subch_initial_model} Initial model for the double detonation test problem.}
-\end{figure}
-
-An initial perturbation is placed on the symmetry axis of the form:
-\begin{equation}
-  T = T_0 \left [ 1 + X(\isotm{He}{4}) f (1 + \tanh(2 - r_1)) \right ]
-\end{equation}
-where
-\begin{equation}
-  r_1 = \left [ x^2 + (y - r_0)^2 \right ]^{1/2} / \lambda
-\end{equation}
-and
-\begin{equation}
-  r_0 = r_\mathrm{pert} + r_\mathrm{base}
-\end{equation}
-Here, $r_\mathrm{base}$ is the radius at which the helium layer begins
-and $r_\mathrm{pert}$ is the distance above the base to put the perturbation.  We
-choose $r_\mathrm{pert} = 100~\mathrm{km}$.  The temperature is pertubed above
-the initial model value, denoted as $T_0$ here.  The amplitude of the perturbation
-is $f = 12$ and the scale of the perturbation is $\lambda = 200~\mathrm{km}$.
-
-We do a suite of runs with both Strang and the simplified-SDC integration, using
-CFL numbers of 0.1, 0.2, and 0.4 and reaction network tolerances of $\atolspec =
-\rtolspec = 10^{-5}$ and $\atolener = \rtolener = 10^{-5}$.  We also run the CFL
-= 0.2 case with a tighter tolerance, using  $\atolspec = 10^{-8}$.  For the step
-rejection part of VODE, we use $f = 4$ and $\eta = 4$.  The runs are labeled with
-the prefix {\tt sdc\_} or {\tt strang\_} to detonate the integrator, and also include
-the CFL number in the name.  Finally, the runs with the tighter tolerances have
-the suffix {\tt \_tol1.e-8}.  We note that both the Strang and the SDC runs encounter
-VODE integration failures occassionally and utilize the \castro\ step retry functionality.
-
-Figure \ref{fig:subch_strang_cfl02} shows the temperature, mean
-molecular weight, and nuclear energy generation rate for the strang
-CFL = 0.2 simulation ({\tt strang\_subch2\_cfl0.2}).  At this point
-in the evolution, the helium detonation has wrapped more than halfway
-across the surface and the ingoing carbon detonation is approaching the
-center of the star.  We see that the
-mean molecular weight appears mottled behind the inward propagating
-detonation in the white dwarf.  Comparing to the energy generation
-rate, we see that region is characterized by large energy releases of
-alternating signs checkerboarding thoughout the region.  This is
-characteristic of the nucleosynthesis struggling to come into
-equilibrium at high temperatures when the reverse rates can be
-important.
-
-\begin{figure}
-\centering
-\plotone{strang_subch2_cfl0.2_tols_slice}
-\caption{\label{fig:subch_strang_cfl02} Temperature, mean molecular weight, and
-energy generation rate for the Strang CFL = 0.2 run with normal tolerances ({\tt
-strang_subch2\_cfl0.2}) at 0.1 s.}
-\end{figure}
-
-We next look at the simplified-SDC version of this same case, {\tt
-sdc\_subch2\_cfl0.2}.  Figure~\ref{fig:subch_sdc_cfl02} shows the temperature,
-mean molecular weight, and energy generation rate. Compared to
-Figure~\ref{fig:subch_strang_cfl02}, we see that the composition and energy
-generation rates are smooth, without any of the checkerboarding that plagued the
-Strang simulation.  This suggests that the simplified-SDC algorithm has an
-easier time attaining the correct equilibrium in these conditions. More
-strikingly, we see that the inward propagating carbon detonation has not
-progressed as far into the white dwarf at this time.  The two integration
-methods are giving very different results here.
-
-\begin{figure}
-\centering
-\plotone{sdc_subch2_cfl0.2_tols_slice}
-\caption{\label{fig:subch_sdc_cfl02} Temperature, mean molecular weight, and
-energy generation rate for the simplified-SDC CFL = 0.2 run with normal tolerances ({\tt
-sdc_subch2\_cfl0.2}) at 0.1 s.}
-\end{figure}
-
-% We now look at the evolution of the state while burning by storing the
-% thermodynamic state internal to VODE each substep as it evaluates the
-% righthand side of the ODE system it is integrating.  We pick a point
-% on the symmetry axis a few zones behind the inward moving shock and
-% ran both a Strang and simplified-SDC simulation for a single
-% coarse-grid timestep.  To ensure that we start off the same, both
-% simulations were started from a checkpoint file from the
-% simplified-SDC simulation a few steps before 0.1~s.
-% Figure~\ref{fig:nonaka_plot} shows the evolution of \isot{Ni}{56} over
-% a single coarse-grid timestep (4 fine-grid timesteps).  We see that
-% the Strang solution is much more discontinuous, as the reactions and
-% advection push the solution in different ways, because the processes
-% are completely independent of one another.  If an equilibrium state is
-% important to the nucleosynthesis, the large deviations in the Strang
-% solution mean that we are evaluating the reaction network with a state
-% out of equilibrium for much of the integration time. \MarginPar{need to double check the data---why are there points in the vertical part of Strang?}
-
-% \begin{figure}
-% \plotone{sdc_compare}
-% \caption{\label{fig:nonaka_plot} Evolution of $X(\isotm{Ni}{56})$ over
-%   4 fine-grid timesteps in a single zone behind the inward propagating
-%   shock.  The Strang points (blue) show the first $\Delta t/2$ burn
-%   bring the abundance away from equilibrium, then the advection
-%   overcorrecting it, and finally the second $\Delta t/2$ bringing it
-%   back toward the equilibrium.  The two simplified-SDC iterations
-%   appear much smoother.  For the SDC runs, there are a few points
-%   evaluated beyond the stop time---VODE can use data from outside the
-%   integration range to interpolate to the final solution.}
-% \end{figure}
-
-We next look at how the structure seen in
-Figures~\ref{fig:subch_strang_cfl02} and \ref{fig:subch_sdc_cfl02}
-vary with CFL number and network tolerance.  We focus just on the
-region behind the inward propagating shock.
-Figures~\ref{fig:strang_enuc_summary} and
-\ref{fig:strang_abar_summary} show the nuclear energy generation rate
-and mean molecular weight for all of the Strang runs.  We see that
-changing the CFL number does little to affect the state behind the
-shock---it continues to take on the mottled appearance characteristic
-of not reaching a proper equilibrium.  For the case with tight tolerances, we
-see that the state is greatly improved.  Furthermore, it gets the inward
-propagating carbon detonation in the same position as the SDC runs.
-
-\begin{figure}
-\plotone{strang_enuc_summary_plot}
-\caption{\label{fig:strang_enuc_summary} Comparison of Strang runs showing
-the nuclear energy generation rate behind the inward propagating
-shock.}
-\end{figure}
-
-\begin{figure}
-\plotone{strang_abar_summary_plot}
-\caption{\label{fig:strang_abar_summary} Comparison of Strang runs showing
-the mean molecular weight behind the inward propagating
-shock.}
-\end{figure}
-
-The SDC comparisons are shown in Figures~\ref{fig:sdc_enuc_summary}
-and \ref{fig:sdc_abar_summary}.  We see that the state behind the
-shock looks great regardless of the CFL or integrator tolerances.
-This demonstrates that the simplified-SDC integration algorithm is more
-robust than Strang integration for this problem.
-
-\begin{figure}
-\plotone{sdc_enuc_summary_plot}
-\caption{\label{fig:sdc_enuc_summary} Comparison of simplified-SDC runs showing
-the nuclear energy generation rate behind the inward propagating
-shock.}
-\end{figure}
-
-\begin{figure}
-\plotone{sdc_abar_summary_plot}
-\caption{\label{fig:sdc_abar_summary} Comparison of simplified-SDC runs showing
-the mean molecular weight behind the inward propagating
-shock.}
-\end{figure}
-
-To quantify this effect, we next looks at the total mass of some key
-species.  This is often a desired diagnostic for supernova
-calculations, since \isot{Ni}{56} powers the lightcurve.  Table
-\ref{table:masses} shows the total mass of \isot{He}{4},
-\isot{Si}{28}, and \isot{Ni}{56} for each of the simulations.  For the
-Strang runs, there are large variations between the different
-simulations while the SDC simulations all agree to within a few
-percent.  Further, the Strang run with the tighter tolerances is in
-good agreement with the masses found by the SDC runs.  This supports
-the ideas described above that all of the SDC runs agree while only
-the Strang run with the tighter tolerances is reliable.
-
-\begin{deluxetable}{llll}
-\tablecaption{\label{table:masses} Species masses.}
-\tablehead{\colhead{simulation} & \colhead{$M_{\isotm{He}{4}}$} & \colhead{$M_{\isotm{Si}{28}}$} & \colhead{$M_{\isotm{Ni}{56}}$}}
-\startdata
-\multicolumn{4}{c}{Strang runs} \\
-\hline
-{\tt strang\_subch2\_cfl0.1}           & $1.059\times 10^{32}$  &  $3.804\times 10^{31}$  &  $1.092\times 10^{32}$ \\
-{\tt strang\_subch2\_cfl0.2}           & $9.114\times 10^{31}$  &  $4.174\times 10^{31}$  &  $1.078\times 10^{32}$ \\
-{\tt strang\_subch2\_cfl0.2\_tol1.e-8} & $7.222\times 10^{31}$  &  $8.623\times 10^{31}$  &  $4.041\times 10^{31}$ \\
-{\tt strang\_subch2\_cfl0.4}           & $7.819\times 10^{31}$  &  $4.208\times 10^{31}$  &  $1.045\times 10^{32}$ \\
-\hline
-\multicolumn{4}{c}{SDC runs} \\
-\hline
-{\tt sdc\_subch2\_cfl0.1}           & $7.319\times 10^{31}$  &  $9.147\times 10^{31}$  &  $3.765\times 10^{31}$ \\
-{\tt sdc\_subch2\_cfl0.2}           & $7.208\times 10^{31}$  &  $9.151\times 10^{31}$  &  $3.783\times 10^{31}$ \\
-{\tt sdc\_subch2\_cfl0.2\_tol1.e-8} & $7.376\times 10^{31}$  &  $9.220\times 10^{31}$  &  $3.513\times 10^{31}$ \\
-{\tt sdc\_subch2\_cfl0.4}           & $7.203\times 10^{31}$  &  $8.843\times 10^{31}$  &  $3.754\times 10^{31}$ \\
-\hline
-\enddata
-\end{deluxetable}
-
-\section{Summary}
-
-We presented a simplified spectral deferred corrections scheme for coupling
-hydrodynamics and reactions.  We demonstrated that for a moderate-sized reaction
-network, the simplified-SDC method compares well against Strang for the double
-detonation problem---getting a better solution at looser reaction tolerances,
-and consistent results across a range of CFL numbers.  The benefits of this new
-integration scheme are likely problem and network dependent, and we will explore
-it on other science problems in future papers.
-
-We have not focused on performance or load-balancing the reactions in
-the tests that we ran, but for the simulations shown here,
-the simplified-SDC method runs as fast (or faster) than the Strang method
-for the same tolerances and CFL number, despite having to do the advection and reactions
-twice.  We would expect similar behavior for any simulations that are
-dominated by the cost of reactions.  If the results above are robust across
-different problems and networks, then this suggests that we can run
-the SDC simulation with larger CFL and/or less strict reaction tolerances,
-which would translate into a large performance boost.
-
-This new time-integration method is freely available in \castro, and
-we will continue to explore it on other problems.  Although these simulations
-were run on CPUs, the entire simplified-SDC algorithm also runs on GPUs.
-interested in massive star evolution with this method, where the
-vigorous shell burning leading up to core-collapse would be an ideal
-application for the simplified-SDC method.  In a follow-on paper, we
-will show how to couple this new integration method to a table for
-evaluating an NSE distribution in the iron core of a massive star.
-
-Future work is to extend this methodology to MHD.  This is very
-straightforward since the simplified-SDC method operates solely on the
-thermodynamic state.  We will also explore how to couple radiation
-hydrodynamics, where we need to do an implicit update for the
-radiation energy.
-
-\acknowledgements \castro\ is freely available at
-\url{http://github.com/AMReX-Astro/Castro}.  All of the code and problem setups
-used here are available in the git repo.  The double detonation problem is in
-{\tt Castro/Exec/science/subchandra}.  The work at Stony Brook was supported by
-DOE/Office of Nuclear Physics grant DE-FG02-87ER40317.  This material is based
-upon work supported by the U.S. Department of Energy, Office of Science, Office
-of Advanced Scientific Computing Research and Office of Nuclear Physics,
-Scientific Discovery through Advanced Computing (SciDAC) program under Award
-Number DE-SC0017955.  This research was supported by the Exascale Computing
-Project (17-SC-20-SC), a collaborative effort of the U.S. Department of Energy
-Office of Science and the National Nuclear Security Administration.  MR was supported
-via an NSF REU grant to Stony Brook, NSF 1852143.  This
-research used resources of the National Energy Research Scientific Computing
-Center (NERSC), a U.S. Department of Energy Office of Science User Facility
-located at Lawrence Berkeley National Laboratory, operated under Contract No.\
-DE-AC02-05CH11231 using NERSC award NP-ERCAP0020354.
-
-\software{\amrex\ \citep{amrex_joss},
-          \castro\ \citep{castro},
-          GNU Compiler Collection (\url{https://gcc.gnu.org/}),
-          Linux (\url{https://www.kernel.org}),
-          matplotlib (\citealt{Hunter:2007},  \url{http://matplotlib.org/}),
-          NetworkX \citep{networkx},
-          NumPy \citep{numpy,numpy2},
-          python (\url{https://www.python.org/}),
-          \pynucastro\ \citep{pynucastro},
-          pytest \citep{pytest},
-          SymPy \citep{sympy},
-          yt \citep{yt}
-         }
-
-\appendix
-
-\section{Jacobian}
-
-\label{sec:app:jacobian}
-
-To solve the reaction system implicitly, the ODE solver needs the
-Jacobian, $\partial \Rb/\partial \Uc^\prime$, where $\Uc^\prime =
-(\rho X_k, \rho e)^\intercal$ is the subset of the conserved variables we are
-integrating.  We follow the method of \cite{castro_sdc} and factor
-this into two pieces,
-\begin{equation}
-\label{eq:factored_jac}
-{\bf J} = \frac{\partial \Rb}{\partial {\bf w }} \frac{\partial {\bf w}}{\partial \Uc^\prime}.
-\end{equation}
-where the state ${\bf w}$ is chosen to match the set of variables used to evaluate the reaction rates.
-Writing this out for two species, $X_\alpha$ and $X_\beta$, we have
-\begin{equation}
-\Uc^\prime = \left ( \begin{array}{c} \rho X_\alpha \\ \rho X_\beta \\ \rho e \end{array} \right )
-\end{equation}
-For interfacing with the reaction network, we use
-\begin{equation}
-{\bf w} = \left ( \begin{array}{c} X_\alpha \\  X_\beta \\ T \end{array} \right )
-\end{equation}
-Note: even though we are using $T$ here instead of $e$, we still do the overall ODE integration
-in terms of $(\rho e)$, consistent with the Strang method described in \cite{strang_rnaas}.
-%%  in that
-%% we use $e$ instead of $T$ as the independent variable.  If the network gives derivatives in terms
-%% of temperature, we can convert as:
-%% \begin{equation}
-%% \left . \frac{\partial \phi}{\partial T} \right |_\rho =
-%%   \left . \frac{\partial \phi}{\partial e} \right |_\rho
-%%   \left . \frac{\partial e}{\partial T} \right |_\rho =
-%%   c_v \left . \frac{\partial \phi}{\partial e} \right |_\rho
-%% \end{equation}
-%% for some quantity $\phi(\rho, T)$.  Here we identified the specific heat
-%% at constant volume, $c_v = \partial e/\partial T|_\rho$.
-
-The Jacobian transformation $\partial \Uc^\prime/\partial {\bf w}$ is:
-\begin{equation}
-\frac{\partial \Uc^\prime}{\partial {\bf w}} = \left (
-   \begin{array}{ccc}
-       \rho & 0 & 0 \\
-       0 & \rho & 0  \\
-       \rho  e_{X_\alpha} & \rho e_{X_\beta} & \rho c_v \\
-     \end{array}\right)
-\end{equation}
-where we use the following notation for compactness:
-\begin{equation}
-e_{X_k} = \dedXd
-\end{equation}
-and the specific heat at constant volume is
-\begin{equation}
-c_v = \left . \frac{\partial e}{\partial T} \right |_{\rho, X_k}
-\end{equation}
-We get the inverse
-(computed via SymPy) as:
-\begin{equation}
-\renewcommand{\arraystretch}{1.5}
-\frac{\partial {\bf w}}{\partial \Uc^\prime} = \left (
-  \begin{array}{ccc}
-   \frac{1}{\rho} & 0 & 0 \\
-   0 & \frac{1}{\rho} & 0 \\
-    -\frac{e_{X_\alpha}}{\rho c_v} & -\frac{e_{X_\beta}}{\rho c_v} & \frac{1}{\rho c_v} \\
-   \end{array}\right)
-\renewcommand{\arraystretch}{1}
-\end{equation}
-
-The reaction vector is
-\begin{equation}
-\Rb(\Uc^\prime) = \left (  \begin{array}{c} \rho \omegadot_\alpha \\ \rho \omegadot_\beta \\ \rho \Sdot \end{array} \right )
-\end{equation}
-We take all the quantities to be functions of $\rho$, $e$, and $X_k$,
-but since $\rho$ doesn't change from the reactions (the reactive
-source of the continuity equation is zero), it is held constant in
-these derivatives.  The Jacobian is computed as $\partial \Rb/\partial
-{\bf w}$:
-\begin{equation}
-\renewcommand{\arraystretch}{1.5}
-\frac{\partial \Rb}{\partial {\bf w}} = \left (
-  \begin{array}{ccc}
-     \rho \frac{\partial \omegadot_\alpha}{\partial X_\alpha} &
-     \rho \frac{\partial \omegadot_\alpha}{\partial X_\beta} &
-     \rho \frac{\partial \omegadot_\alpha}{\partial T} \\
-      %
-     \rho \frac{\partial \omegadot_\beta}{\partial X_\alpha} &
-     \rho \frac{\partial \omegadot_\beta}{\partial X_\beta} &
-     \rho \frac{\partial \omegadot_\beta}{\partial T} \\
-     %
-     \rho \frac{\partial \Sdot}{\partial X_\alpha} &
-     \rho \frac{\partial \Sdot}{\partial X_\beta} &
-     \rho \frac{\partial \Sdot}{\partial T} \\
-  \end{array}
-  \right )
-\renewcommand{\arraystretch}{1}
-\end{equation}
-
-The final Jacobian is found by multiplying these two:
-\begin{equation}
-\renewcommand{\arraystretch}{1.5}
-\frac{\partial \Rb}{\partial \Uc^\prime} = \left (
-  \begin{array}{ccc}
-    \frac{\partial \omegadot_\alpha}{\partial X_\alpha} - \frac{e_{X_\alpha}}{c_v} \frac{\partial \omegadot_\alpha}{\partial T} &
-    \frac{\partial \omegadot_\alpha}{\partial X_\beta} - \frac{e_{X_\beta}}{c_v} \frac{\partial \omegadot_\alpha}{\partial T} &
-    \frac{1}{c_v} \frac{\partial \omegadot_\alpha}{\partial T} \\
-     %
-    \frac{\partial \omegadot_\beta}{\partial X_\alpha} - \frac{e_{X_\alpha}}{c_v} \frac{\partial \omegadot_\beta}{\partial T} &
-    \frac{\partial \omegadot_\beta}{\partial X_\beta} - \frac{e_{X_\beta}}{c_v} \frac{\partial \omegadot_\beta}{\partial T} &
-    \frac{1}{c_v} \frac{\partial \omegadot_\beta}{\partial T} \\
-     %
-     \frac{\partial \Sdot}{\partial X_\alpha} -  \frac{e_{X_\alpha}}{c_v} \frac{\partial \Sdot}{\partial T} &
-     \frac{\partial \Sdot}{\partial X_\beta} -  \frac{e_{X_\beta}}{c_v} \frac{\partial \Sdot}{\partial T} &
-     \frac{1}{c_v} \frac{\partial \Sdot}{\partial T} \\
-  \end{array}
-  \right )
-\renewcommand{\arraystretch}{1}
-\end{equation}
-We note that the form of these entries is the same as one would arrive at if you start with the rates expressed as
-$\omega_k(\rho, T(\rho, X_j, e), X_j)$ and recognize that constant $e$ implies that
-\begin{equation}
-\left . \frac{\partial T}{\partial X_k} \right |_{\rho, e,X_{j,j\ne k}} = - \frac{e_{X_k}}{c_v}
-\end{equation}
-
-Finally we note that one implication of this formulation is that the Jacobian is no longer sparse.  Future
-work will explore alternate formulations and approximate Jacobians.
-
-% While VODE can compute the entire Jacobian, ${\bf J}$ numerically via
-% differencing, we found that does not give reliable results.  Instead,
-% we compute compute the derivatives with respect to $X_k$ and $T$
-% one-sided differencing, following the algorithm in \cite{lsode} to
-% minimize numerical noise.  We then use the equation of state to
-% compute $e_{X_k}$ and $c_v$ and construct the entries of the final
-% Jacobian.
-
-\section{Reaction Network}
-\label{sec:app:reactionnet}
-We build a reaction network that approximates an $\alpha$-chain,
-including all $\isotm{He}{4}$, and all of the $\alpha$ nuclei from
-$\isotm{C}{12}$ to $\isotm{Ni}{56}$.  We want to capture both the
-$(\alpha,\gamma)$ and $(\alpha,p)(p,\gamma)$ links, so we include the
-intermediate nuclei from the $(\alpha,p)$ rate instead of
-approximating these links assuming proton equilibriation.  Following
-\citet{shenbildsten}, we include the sequence
-$\isotm{N}{14}(\alpha,\gamma)\isotm{F}{18}(\alpha, p)\isotm{Ne}{21}$,
-which produce the protons that can allow for $\isotm{C}{12}(p,
-\gamma)\isotm{N}{13}(\alpha, p)\isotm{O}{16}$ to compete with
-$\isotm{C}{12}(\alpha,\gamma)\isotm{O}{16}$.  We add $\isotm{Na}{22}$
-to link the $\isotm{Ne}{21}$ produced back to the $\alpha$-chain.  We
-grab all of the ReacLib \citep{reaclib} reaction rates linking these
-nuclei using \pynucastro~\citep{pynucastro}.  We are missing one rate
-each for $\isotm{C}{12} + \isotm{C}{12}$, $\isotm{C}{12} +
-\isotm{O}{16}$, and $\isotm{O}{16} + \isotm{O}{16}$.  These are
-the sequences $\isotm{C}{12}(\isotm{C}{12}, \mathrm{n})\isotm{Mg}{23}(n, \gamma)\isotm{Mg}{24}$,
-$\isotm{O}{16}(\isotm{C}{12}, \mathrm{n})\isotm{Si}{27}(n, \gamma)\isotm{Si}{28}$, and
-$\isotm{O}{16}(\isotm{O}{16}, \mathrm{n})\isotm{S}{31}(n, \gamma)\isotm{S}{32}$,involving neutron
-production followed by a capture back to one of our $\alpha$-chain
-nuclei.  These are the only rates we approximate here, assuming
-that the subsequent neutron capture is instantaneous.
-
-Our final set of 28 nuclei are ${}^{1}\mathrm{H}$, ${}^{4}\mathrm{He}$,
-${}^{12}\mathrm{C}$, ${}^{13}\mathrm{N}$, ${}^{14}\mathrm{N}$,
-${}^{16}\mathrm{O}$, ${}^{18}\mathrm{F}$, ${}^{20}\mathrm{Ne}$,
-${}^{21}\mathrm{Ne}$, ${}^{22}\mathrm{Na}$, ${}^{23}\mathrm{Na}$,
-${}^{24}\mathrm{Mg}$, ${}^{27}\mathrm{Al}$, ${}^{28}\mathrm{Si}$,
-${}^{31}\mathrm{P}$, ${}^{32}\mathrm{S}$, ${}^{35}\mathrm{Cl}$,
-${}^{36}\mathrm{Ar}$, ${}^{39}\mathrm{K}$, ${}^{40}\mathrm{Ca}$,
-${}^{43}\mathrm{Sc}$, ${}^{44}\mathrm{Ti}$, ${}^{47}\mathrm{V}$,
-${}^{48}\mathrm{Cr}$, ${}^{51}\mathrm{Mn}$, ${}^{52}\mathrm{Fe}$,
-${}^{55}\mathrm{Co}$, ${}^{56}\mathrm{Ni}$.  The 107 rates linking them are
-given in table~\ref{table:rates}, with the forward rates having a positive $Q$
-value and the reverse having a negative $Q$ value.  We add screening to the
-rates using the prescription from \citet{graboske:1973,alastuey:1978,itoh:1979}.
-The C++ code for the network is output by \pynucastro\ directly for \castro.  A
-graphical overview of the network showing the links between the nuclei is shown
-in figure~\ref{fig:subch_network}.
-
-\startlongtable
-\begin{deluxetable}{ll}
-\tablecaption{\label{table:rates} Reaction rates included in our network.}
-\tablehead{\colhead{forward reaction} & \colhead{reverse reaction}}
-\startdata
-${}^{4}\mathrm{He} + {}^{4}\mathrm{He} + {}^{4}\mathrm{He} \rightarrow {}^{12}\mathrm{C}$ &
-  ${}^{12}\mathrm{C} \rightarrow {}^{4}\mathrm{He} + {}^{4}\mathrm{He} + {}^{4}\mathrm{He}$ \\
-${}^{12}\mathrm{C} + {}^{1}\mathrm{H} \rightarrow {}^{13}\mathrm{N}$ &
-  ${}^{13}\mathrm{N} \rightarrow {}^{1}\mathrm{H} + {}^{12}\mathrm{C}$ \\
-${}^{12}\mathrm{C} + {}^{4}\mathrm{He} \rightarrow {}^{16}\mathrm{O}$ &
-  ${}^{16}\mathrm{O} \rightarrow {}^{4}\mathrm{He} + {}^{12}\mathrm{C}$ \\
-${}^{12}\mathrm{C} + {}^{12}\mathrm{C} \rightarrow {}^{4}\mathrm{He} + {}^{20}\mathrm{Ne}$ &
-  ${}^{20}\mathrm{Ne} + {}^{4}\mathrm{He} \rightarrow {}^{12}\mathrm{C} + {}^{12}\mathrm{C}$ \\
-${}^{12}\mathrm{C} + {}^{12}\mathrm{C} \rightarrow {}^{1}\mathrm{H} + {}^{23}\mathrm{Na}$ &
-  ${}^{23}\mathrm{Na} + {}^{1}\mathrm{H} \rightarrow {}^{12}\mathrm{C} + {}^{12}\mathrm{C}$ \\
-${}^{12}\mathrm{C} + {}^{12}\mathrm{C} \rightarrow {}^{24}\mathrm{Mg}$ &
-                                         \\
-${}^{13}\mathrm{N} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{16}\mathrm{O}$ &
-  ${}^{16}\mathrm{O} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{13}\mathrm{N}$ \\
-${}^{14}\mathrm{N} + {}^{4}\mathrm{He} \rightarrow {}^{18}\mathrm{F}$ &
-  ${}^{18}\mathrm{F} \rightarrow {}^{4}\mathrm{He} + {}^{14}\mathrm{N}$ \\
-${}^{16}\mathrm{O} + {}^{4}\mathrm{He} \rightarrow {}^{20}\mathrm{Ne}$ &
-  ${}^{20}\mathrm{Ne} \rightarrow {}^{4}\mathrm{He} + {}^{16}\mathrm{O}$ \\
-${}^{16}\mathrm{O} + {}^{12}\mathrm{C} \rightarrow {}^{4}\mathrm{He} + {}^{24}\mathrm{Mg}$ &
-  ${}^{24}\mathrm{Mg} + {}^{4}\mathrm{He} \rightarrow {}^{12}\mathrm{C} + {}^{16}\mathrm{O}$ \\
-${}^{16}\mathrm{O} + {}^{12}\mathrm{C} \rightarrow {}^{1}\mathrm{H} + {}^{27}\mathrm{Al}$ &
-  ${}^{27}\mathrm{Al} + {}^{1}\mathrm{H} \rightarrow {}^{12}\mathrm{C} + {}^{16}\mathrm{O}$ \\
-${}^{16}\mathrm{O} + {}^{12}\mathrm{C} \rightarrow {}^{28}\mathrm{Si}$ &
-                                         \\
-${}^{16}\mathrm{O} + {}^{16}\mathrm{O} \rightarrow {}^{4}\mathrm{He} + {}^{28}\mathrm{Si}$ &
-  ${}^{28}\mathrm{Si} + {}^{4}\mathrm{He} \rightarrow {}^{16}\mathrm{O} + {}^{16}\mathrm{O}$ \\
-${}^{16}\mathrm{O} + {}^{16}\mathrm{O} \rightarrow {}^{1}\mathrm{H} + {}^{31}\mathrm{P}$ &
-  ${}^{31}\mathrm{P} + {}^{1}\mathrm{H} \rightarrow {}^{16}\mathrm{O} + {}^{16}\mathrm{O}$ \\
-${}^{16}\mathrm{O} + {}^{16}\mathrm{O} \rightarrow {}^{32}\mathrm{S}$ &
-                                         \\
-${}^{18}\mathrm{F} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{21}\mathrm{Ne}$ &
-  ${}^{21}\mathrm{Ne} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{18}\mathrm{F}$ \\
-${}^{18}\mathrm{F} + {}^{4}\mathrm{He} \rightarrow {}^{22}\mathrm{Na}$ &
-  ${}^{22}\mathrm{Na} \rightarrow {}^{4}\mathrm{He} + {}^{18}\mathrm{F}$ \\
-${}^{20}\mathrm{Ne} + {}^{4}\mathrm{He} \rightarrow {}^{24}\mathrm{Mg}$ &
-  ${}^{24}\mathrm{Mg} \rightarrow {}^{4}\mathrm{He} + {}^{20}\mathrm{Ne}$ \\
-${}^{20}\mathrm{Ne} + {}^{12}\mathrm{C} \rightarrow {}^{4}\mathrm{He} + {}^{28}\mathrm{Si}$ &
-  ${}^{28}\mathrm{Si} + {}^{4}\mathrm{He} \rightarrow {}^{12}\mathrm{C} + {}^{20}\mathrm{Ne}$ \\
-${}^{20}\mathrm{Ne} + {}^{12}\mathrm{C} \rightarrow {}^{1}\mathrm{H} + {}^{31}\mathrm{P}$ &
-  ${}^{31}\mathrm{P} + {}^{1}\mathrm{H} \rightarrow {}^{12}\mathrm{C} + {}^{20}\mathrm{Ne}$ \\
-${}^{21}\mathrm{Ne} + {}^{1}\mathrm{H} \rightarrow {}^{22}\mathrm{Na}$ &
-  ${}^{22}\mathrm{Na} \rightarrow {}^{1}\mathrm{H} + {}^{21}\mathrm{Ne}$ \\
-${}^{23}\mathrm{Na} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{20}\mathrm{Ne}$ &
-  ${}^{20}\mathrm{Ne} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{23}\mathrm{Na}$ \\
-${}^{23}\mathrm{Na} + {}^{1}\mathrm{H} \rightarrow {}^{24}\mathrm{Mg}$ &
-  ${}^{24}\mathrm{Mg} \rightarrow {}^{1}\mathrm{H} + {}^{23}\mathrm{Na}$ \\
-${}^{23}\mathrm{Na} + {}^{4}\mathrm{He} \rightarrow {}^{27}\mathrm{Al}$ &
-  ${}^{27}\mathrm{Al} \rightarrow {}^{4}\mathrm{He} + {}^{23}\mathrm{Na}$ \\
-${}^{24}\mathrm{Mg} + {}^{4}\mathrm{He} \rightarrow {}^{28}\mathrm{Si}$ &
-  ${}^{28}\mathrm{Si} \rightarrow {}^{4}\mathrm{He} + {}^{24}\mathrm{Mg}$ \\
-${}^{27}\mathrm{Al} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{24}\mathrm{Mg}$ &
-  ${}^{24}\mathrm{Mg} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{27}\mathrm{Al}$ \\
-${}^{27}\mathrm{Al} + {}^{1}\mathrm{H} \rightarrow {}^{28}\mathrm{Si}$ &
-  ${}^{28}\mathrm{Si} \rightarrow {}^{1}\mathrm{H} + {}^{27}\mathrm{Al}$ \\
-${}^{27}\mathrm{Al} + {}^{4}\mathrm{He} \rightarrow {}^{31}\mathrm{P}$ &
-  ${}^{31}\mathrm{P} \rightarrow {}^{4}\mathrm{He} + {}^{27}\mathrm{Al}$ \\
-${}^{28}\mathrm{Si} + {}^{4}\mathrm{He} \rightarrow {}^{32}\mathrm{S}$ &
-  ${}^{32}\mathrm{S} \rightarrow {}^{4}\mathrm{He} + {}^{28}\mathrm{Si}$ \\
-${}^{31}\mathrm{P} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{28}\mathrm{Si}$ &
-  ${}^{28}\mathrm{Si} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{31}\mathrm{P}$ \\
-${}^{31}\mathrm{P} + {}^{1}\mathrm{H} \rightarrow {}^{32}\mathrm{S}$ &
-  ${}^{32}\mathrm{S} \rightarrow {}^{1}\mathrm{H} + {}^{31}\mathrm{P}$ \\
-${}^{31}\mathrm{P} + {}^{4}\mathrm{He} \rightarrow {}^{35}\mathrm{Cl}$ &
-  ${}^{35}\mathrm{Cl} \rightarrow {}^{4}\mathrm{He} + {}^{31}\mathrm{P}$ \\
-${}^{32}\mathrm{S} + {}^{4}\mathrm{He} \rightarrow {}^{36}\mathrm{Ar}$ &
-  ${}^{36}\mathrm{Ar} \rightarrow {}^{4}\mathrm{He} + {}^{32}\mathrm{S}$ \\
-${}^{35}\mathrm{Cl} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{32}\mathrm{S}$ &
-  ${}^{32}\mathrm{S} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{35}\mathrm{Cl}$ \\
-${}^{35}\mathrm{Cl} + {}^{1}\mathrm{H} \rightarrow {}^{36}\mathrm{Ar}$ &
-  ${}^{36}\mathrm{Ar} \rightarrow {}^{1}\mathrm{H} + {}^{35}\mathrm{Cl}$ \\
-${}^{35}\mathrm{Cl} + {}^{4}\mathrm{He} \rightarrow {}^{39}\mathrm{K}$ &
-  ${}^{39}\mathrm{K} \rightarrow {}^{4}\mathrm{He} + {}^{35}\mathrm{Cl}$ \\
-${}^{36}\mathrm{Ar} + {}^{4}\mathrm{He} \rightarrow {}^{40}\mathrm{Ca}$ &
-  ${}^{40}\mathrm{Ca} \rightarrow {}^{4}\mathrm{He} + {}^{36}\mathrm{Ar}$ \\
-${}^{39}\mathrm{K} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{36}\mathrm{Ar}$ &
-  ${}^{36}\mathrm{Ar} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{39}\mathrm{K}$ \\
-${}^{39}\mathrm{K} + {}^{1}\mathrm{H} \rightarrow {}^{40}\mathrm{Ca}$ &
-  ${}^{40}\mathrm{Ca} \rightarrow {}^{1}\mathrm{H} + {}^{39}\mathrm{K}$ \\
-${}^{39}\mathrm{K} + {}^{4}\mathrm{He} \rightarrow {}^{43}\mathrm{Sc}$ &
-  ${}^{43}\mathrm{Sc} \rightarrow {}^{4}\mathrm{He} + {}^{39}\mathrm{K}$ \\
-${}^{40}\mathrm{Ca} + {}^{4}\mathrm{He} \rightarrow {}^{44}\mathrm{Ti}$ &
-  ${}^{44}\mathrm{Ti} \rightarrow {}^{4}\mathrm{He} + {}^{40}\mathrm{Ca}$ \\
-${}^{43}\mathrm{Sc} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{40}\mathrm{Ca}$ &
-  ${}^{40}\mathrm{Ca} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{43}\mathrm{Sc}$ \\
-${}^{43}\mathrm{Sc} + {}^{1}\mathrm{H} \rightarrow {}^{44}\mathrm{Ti}$ &
-  ${}^{44}\mathrm{Ti} \rightarrow {}^{1}\mathrm{H} + {}^{43}\mathrm{Sc}$ \\
-${}^{43}\mathrm{Sc} + {}^{4}\mathrm{He} \rightarrow {}^{47}\mathrm{V}$ &
-  ${}^{47}\mathrm{V} \rightarrow {}^{4}\mathrm{He} + {}^{43}\mathrm{Sc}$ \\
-${}^{44}\mathrm{Ti} + {}^{4}\mathrm{He} \rightarrow {}^{48}\mathrm{Cr}$ &
-  ${}^{48}\mathrm{Cr} \rightarrow {}^{4}\mathrm{He} + {}^{44}\mathrm{Ti}$ \\
-${}^{47}\mathrm{V} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{44}\mathrm{Ti}$ &
-  ${}^{44}\mathrm{Ti} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{47}\mathrm{V}$ \\
-${}^{47}\mathrm{V} + {}^{1}\mathrm{H} \rightarrow {}^{48}\mathrm{Cr}$ &
-  ${}^{48}\mathrm{Cr} \rightarrow {}^{1}\mathrm{H} + {}^{47}\mathrm{V}$ \\
-${}^{47}\mathrm{V} + {}^{4}\mathrm{He} \rightarrow {}^{51}\mathrm{Mn}$ &
-  ${}^{51}\mathrm{Mn} \rightarrow {}^{4}\mathrm{He} + {}^{47}\mathrm{V}$ \\
-${}^{48}\mathrm{Cr} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{51}\mathrm{Mn}$ &
-  ${}^{51}\mathrm{Mn} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{48}\mathrm{Cr}$ \\
-${}^{48}\mathrm{Cr} + {}^{4}\mathrm{He} \rightarrow {}^{52}\mathrm{Fe}$ &
-  ${}^{52}\mathrm{Fe} \rightarrow {}^{4}\mathrm{He} + {}^{48}\mathrm{Cr}$ \\
-${}^{51}\mathrm{Mn} + {}^{1}\mathrm{H} \rightarrow {}^{52}\mathrm{Fe}$ &
-  ${}^{52}\mathrm{Fe} \rightarrow {}^{1}\mathrm{H} + {}^{51}\mathrm{Mn}$ \\
-${}^{51}\mathrm{Mn} + {}^{4}\mathrm{He} \rightarrow {}^{55}\mathrm{Co}$ &
-  ${}^{55}\mathrm{Co} \rightarrow {}^{4}\mathrm{He} + {}^{51}\mathrm{Mn}$ \\
-${}^{52}\mathrm{Fe} + {}^{4}\mathrm{He} \rightarrow {}^{1}\mathrm{H} + {}^{55}\mathrm{Co}$ &
-  ${}^{55}\mathrm{Co} + {}^{1}\mathrm{H} \rightarrow {}^{4}\mathrm{He} + {}^{52}\mathrm{Fe}$ \\
-${}^{52}\mathrm{Fe} + {}^{4}\mathrm{He} \rightarrow {}^{56}\mathrm{Ni}$ &
-  ${}^{56}\mathrm{Ni} \rightarrow {}^{4}\mathrm{He} + {}^{52}\mathrm{Fe}$ \\
-${}^{55}\mathrm{Co} + {}^{1}\mathrm{H} \rightarrow {}^{56}\mathrm{Ni}$ &
-  ${}^{56}\mathrm{Ni} \rightarrow {}^{1}\mathrm{H} + {}^{55}\mathrm{Co}$ \\
-\enddata
-\end{deluxetable}
-
-\begin{figure}[t]
-\centering
-\plotone{subch2}
-\caption{\label{fig:subch_network} Reaction network for the double detonation problem.}
-\end{figure}
+Retries are great \cite{katzzingale:2019}.
 

### 2022-11-10 retries/paper.tex
Start writing some text for the retry paper

@@
 
-Retries are great \cite{katzzingale:2019}.
+Multiphysics simulation software is hard to implement robustly. One reason is that
+physics modules usually have one or more failure modes, and the code may have
+additional failure modes related to the coupling between modules. Some examples of
+simulation failure modes in our field of astrophysics include:
+\begin{itemize}
+  \item Numerical integration failures in the system of ordinary differential equations (ODEs) representing nuclear fusion
+  \item Numerical integration successes in coupled ODE systems that nevertheless result in nonphysical outcomes, like mass fraction > 1
+  \item Convergence failures in iterative linear solvers, such as those used to solve the Poisson equation
+  \item Convergence failures in iterative non-linear solvers, such as those used for radiation diffusion
+  \item Failures in non-linear root finding, such as in non-monotonic equations of state
+  \item Zones whose local velocity exceeds the CFL hydrodynamic stability criterion given the selected timestep
+  \item Negative density in a zone after a hydrodynamic update
+\end{itemize}
+All of these failure modes (and more) are represented in the simulation code Castro \citep{castro_joss}.
+In our experience, any useful science simulation will almost inevitably hit one of these failure modes.
+
+There are (at least) three possible responses to such failures. First, the application can abort and tell the user
+to restart from the latest checkpoint with some modified parameters in the hope of making the failure less likely.
+This is undesirable because it can be very disruptive to a long-running simulation, and also because it may not be
+clear what step will actually resolve the problem. Second, the application can attempt to patch over the issue.
+For example, a negative density may be replaced by a small positive density, or a flux limiter could be applied
+to prevent it from occurring in the first place \citep{positivity_preserving}. While this can often keep the
+simulation going, it creates unphysical distortions in the simulation. Our anecdotal experience is that most
+simulation codes use these first two approaches to deal with such failures.
+
+A third approach, which we believe is generally more robust and less distortionary, is to retry a timestep
+that encounters one of these failure modes. Typically one will use a smaller $\Delta t$ on the retry, although
+other options exist, such as modifying the spatial resolution when using an adaptive mesh. One astrophysics
+application that successfully does this, which we drew inspiration from for this work, is the stellar evolution
+code MESA \citep{MESA}.
 

### 2022-12-14 simplified_sdc_nse_notes/paper.tex
notes

--- 2021-09-27 simplified_sdc_intro/paper.tex
@@
 \usepackage{cancel}
+\usepackage{tcolorbox}
 
@@
 %\title{A Simplified Spectral Deferred Correction Method for Coupling Hydrodynamics with Reaction Networks and Nuclear Statistical Equilibrium}
-\title{An Improved Method for Coupling Hydrodynamics with Reaction Networks and Nuclear Statistical Equilibrium}
-
-%\shorttitle{A Simplified SDC Method}
-\shorttitle{Hydro/Reaction Coupling with NSE}
-
-\shortauthors{}
-
-\author[0000-0001-8401-030X]{M.~Zingale}
-\affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
-  Stony Brook, NY 11794-3800}
-
-\author[0000-0003-0439-4556]{M.~P.~Katz}
-\affiliation{Nvidia Corp}
-
-\author[0000-0003-1791-0265]{A.~J.~Nonaka}
-\affiliation{Center for Computational Sciences and Engineering, Lawrence Berkeley National Laboratory, Berkeley, CA  94720}
-
-\author[0000-0002-1530-781X]{Alice Harpole}
-\affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
-             Stony Brook, NY 11794-3800}
-
-\correspondingauthor{Michael Zingale}
-\email{michael.zingale@stonybrook.edu}
+\title{Notes on Extending Simplified SDC to Nuclear Statistical Equilibrium}
 
@@
 \begin{abstract}
-Reacting astrophysical flows can be challenging to model because of
-the difficulty in accurately coupling hydrodynamics and reactions.
-This can be particularly acute during explosive burning or at high
-temperatures where nuclear statistical equilibrium is established.  We
-develop a simplified approach based on the ideas of spectral deferred
-corrections (SDC) coupling of explicit hydrodynamics and stiff
-reaction sources as an alternative to operator splitting or the more
-comprehensive SDC approach we demonstrated previously.  We apply the
-new method to some example problems and show how to modify it to work
-with a hybrid network consisting of a reaction ODE system and a table
-for nuclear statistical equilibrium.  This is all done in the
-framework of the \castro\ hydrodynamics code, and all algorithm
-implementations are freely available.
 \end{abstract}
@@
 
-%======================================================================
-% Introduction
-%======================================================================
-\section{Introduction}\label{Sec:Introduction}
-
-Modeling astrophysical reacting flows can be challenging because of
-the disparity better the nuclear and hydrodynamics timescales.
-Reaction networks tend to be stiff, requiring implicit integration
-techniques to stably integrate the system.  In contrast, compressible
-hydrodynamics flows are limited by the (often much longer)
-sound-crossing time over a computational cell and can be solved using
-explicit time integration. Traditional methods of coupling
-hydrodynamics and reactions used in astrophysics use operator
-splitting---each physical process acts on the output of the previous
-process in alternating fashion.  This makes it easy to use different
-time-integration methods for the different physics, and to build a
-simulation code in a modular way.  However, competition between the
-different physical processes can cause the coupling to breakdown.
-These splitting errors can lead to loss of accuracy and further time
-step limitations.
-
-A particularly difficult phase of evolution to model is the nuclear
-statistical equilibrium that sets in for
-temperatures in excess of $\mbox{few} \times 10^9~\mathrm{K}$.  Physically, the
-forward and reverse rates of reaction should balance leading to an equilibrium.
-With operator splitting, an NSE region will have a large positive flow
-through the network in a zone in one step followed by a large negative
-flow over the next timestep, as the code struggles to produce an
-equilibrium.  These large changes in abundances (and large alternately
-positive and negative energy generation rates) can be a challenge for
-a code.  The easiest way to improve the coupling is to cut the
-timestep, but this makes simulations prohibitively expensive.
-Sometimes the burning is simply halted on a zone-by-zone basis when
-NSE conditions are reached (e.g., as in \citealt{hedet}).
-Alternately, at high temperatures, a reaction network can be replaced
-with a table of NSE abundances and the zone's composition set through
-table look-ups (e.g.\ \citealt{ma:2013}) \MarginPar{need more here}
-
-The \castro\ hydrodynamics code \citep{castro,castro_joss} is used for
-all of our numerical experiments.  \castro\ is a compressible
-(magneto-, radiation-) hydrodynamics code built on the AMReX adaptive
-mesh refinement (AMR) framework \citep{amrex_joss}.  \castro\ has been
-designed to be performance portable and runs on massively parallel CPU
-and GPU architectures \citep{castro_gpu}.  For hydrodynamics, the
-corner transport upwind (CTU) \citep{ppmunsplit} method with the
-piecewise parabolic method (PPM) \citep{ppm,millercolella:2002} is
-used.  \castro\ includes self-gravity, rotation, arbitrary equations
-of state and reaction networks, and has been used for modeling X-ray
-bursts and different models of thermonuclear, core-collapse, and
-pair-instability supernovae.  Recently in \citet{castro_sdc}, we
-developed second- and fourth-order accurate methods in space and time
-for coupling hydrodynamics and nuclear reaction networks based on
-spectral deferred corrections (SDC) methods, and demonstrated the
-method in a variety of test problems.
-
-The time-integration approach presented here is considerably simpler than
-the SDC method of \citet{castro_sdc}, but allows us to reuse the main CTU
-hydrodynamics construction and a largely similar ODE integration
-scheme, making this method easier to add to existing simulation codes.
-Furthermore, it also extends to adaptive mesh refinement with
-subcycling in a straightforward manner, avoiding the complications
-described in \citep{mccorquodalecolella} needed to fill ghost cells
-when using method-of-lines integration.  However, it is restricted to
-second-order accuracy overall.  We term this algorithm the
-``simplified SDC method''.  In this paper we describe the overall method, show its extension to NSE,
-and demonstrate it on several test problems using \castro.  All of the
-code to reproduce the results in this paper are in the \castro\ github
-repository\footnote{\url{https://github.com/amrex-astro/Castro/}}.
-
-\section{Numerical Methodology}
-
-We solve the Euler equations for compressible, reacting flow.  For ease
-of exposition we describe the one-dimensional case;
-multidimensional extensions are a straightforward modification to
-include the CTU hydrodynamics scheme.  Our conserved variables are
-\begin{equation}
-  \Uc = \left ( \begin{array}{c}
-           \rho \\
-           \rho X_k \\
-           \rho \alpha_l \\
-           \rho u \\
-           \rho E \\
-           \rho e \end{array}\right )
-\end{equation}
-where $\rho$ is the mass density, $u$ is velocity, $E$ is specific
-total energy, $p$ is the pressure, and we carry nuclear species mass
-fractions, $X_k$, and auxiliary composition variables, $\alpha_l$.  The specific total
-energy relates to the specific internal energy, $e$, as $E = e + u^2/2$,
-and we also separately evolve $e$, as part of a dual
-energy formulation (see \citealt{bryan:1995,wdmergerI}).
- The mass fractions are constrained to sum to 1, $\sum_k X_k = 1$, but
- no such constraint exists on the auxiliary variables. 
-Defining the hydrodynamical fluxes:
-\begin{equation}
-  \Fb(\Uc) = \left ( \begin{array}{c}
-         \rho u \\
-         \rho X_k u \\
-         \rho \alpha_l u \\
-         \rho u^2 p \\
-         (\rho E + p) u \\
-         \rho e u \end{array}\right )
-\end{equation}
-We can write the system in conservative form for all state variables aside from $(\rho e)$ as:
-\begin{equation}
-  \ddt{\Uc} + \ddx{\Fb(\Uc)} = \Sc(\Uc)
-\end{equation}
-where for the special case of $(\rho e)$, we have an additional ``pdV'' term:
-\begin{equation}
-\ddt{(\rho e)} + \ddx{F(\rho e)} + p \ddx{u} = S(\rho e)
-\end{equation}
-
-We'll split the source term, $\Sc(\Uc)$ into pure hydrodynamic and reactive parts:
-\begin{equation}
-  \Hb(\Uc) = \left ( \begin{array}{c}
-    0 \\
-    0 \\
-    0 \\
-    \rho g \\
-    \rho u g \\
-    0 \end{array} \right )
-  \qquad
-  \Rb(\Uc) = \left ( \begin{array}{c}
-     0 \\
-     \rho \omegadot(X_k) \\
-     \rho \omegadot(\alpha_l) \\
-     0 \\
-     \rho \dot{S} \\
-     \rho \dot{S} 
-  \end{array} \right )
-\end{equation}
-with 
-\begin{equation}
-  \Sc(\Uc) = \Hb(\Uc) + \Rb(\Uc).
-\end{equation}
-Here, gravity, represented by the gravitational acceleration $g$
-appears as a pure hydrodynamical source term.  From reactions,
-$\omegadot(X_k)$ is the creation rate for species $k$,
-$\omegadot(\alpha_l)$ is the creation rate for auxiliary composition
-variable $l$, and $\dot{S}$ is the energy generation rate per unit
-mass.  We note that the internal energy ``pdV'' work is not treated as
-a source term but is instead constructed with the hydrodynamical
-fluxes are computed in the CTU method.
-
-Aside from the $p\partial u/\partial x$ (``pdV'') term in the internal
-energy equation, this system is in conservative form with source
-terms.
-
-We'll define the advective terms with hydrodynamic sources, $\Advs{\Uc}$ as
-\begin{equation}
-\Advs{\Uc} = -\ddx{\Fb(\Uc)} + \Hb(\Uc)
-\end{equation}
-for the general case, with the $(\rho e)$ component again having the extra
-``pdV'' term:
-\begin{equation}
-\mathcal{A}(\rho e) = -\ddx{F(\rho e)} -p \ddx{u} + H(\rho e)
-\end{equation}
-
-To close the system, we need the equation of state.  For a system
-where the composition is completely specified by the mass fractions,
-$X_k$, the equation of state would take the form:
-\begin{equation}
-p = p(\rho, X_k, e)
-\end{equation}
-However, as we'll see shortly, when we use the NSE table, we are using
-the results of a much larger network then can be represented by the
-mass fractions, and in this case, we rely on the auxiliary composition
-variables, $\alpha_l$, to describe the state of the composition, and
-our EOS has the form:
-\begin{equation}
-p = p(\rho, \alpha_l, e)
-\end{equation}
-
-Sometimes it is preferable to work with the primitive variables,
-\begin{equation}
-\qb = \left ( \begin{array}{c}
-  \rho \\
-  X_k \\
-  \alpha_l \\
-  u \\
-  p \\
-  (\rho e) \\
-\end{array} \right )
-\end{equation}
-Here, the system appears
-as:
-\begin{equation}
-\qb_t + \Ab^\xv(\qb) \qb_x  = \Sc(\qb)
-\end{equation}
-with the matrix $\Ab^\xv$ giving the coefficients of the spatial derivatives
-of the primitive variables:
-\begin{equation}
-\Ab^\xv(\qb) = \left ( \begin{array}{cccccc}
-    u & 0 & 0 & \rho & 0 & 0 \\
-    0 & u & 0 & 0    & 0 & 0 \\
-    0 & 0 & u & 0    & 0 & 0 \\
-    0 & 0 & 0 & u    & 1/\rho & 0 \\
-    0 & 0 & 0 & \Gamma_1 p & u & 0 \\
-    0 & 0 & 0 & \rho h & 0 & u
-  \end{array} \right )
-\end{equation}
-where $h$ is the specific enthalpy and $\Gamma_1$ is an adiabatic index,
-$\Gamma_1 = d\log p/d\log\rho|_s$ at constant entropy.
-The CTU+PPM algorithm uses the characteristic wave structure of $\Ab$
-to collect the information that makes it to an interface over a timestep
-in order to compute the fluxes through the interface.
-Note, the primitive state has two thermodynamic quantities, $p$
-and $(\rho e)$, to more efficiently handle the general equation of
-state in the Riemann solver, as described in \citet{castro}, but
-alternate formulations are possible \citep{colellaglaz:1985}.
-The source term vector, $\Sc(\qb)$, can again be decomposed into hydrodynamic
-sources (now in terms of the primitive variables) and reaction terms,
-\begin{equation}
-  \Sc(\qb) = \Hb(\qb) + \Rb(\qb)
-\end{equation}
-with
-\begin{equation}
-\label{eq:prim_sources}
-\Hb(\qb) = \left ( \begin{array}{c}
-     0 \\
-     0 \\
-     0 \\
-     g \\
-     0 \\
-     0 \\
-   \end{array} \right )
-\qquad
-\Rb(\qb) = \left ( \begin{array}{c}
-     0 \\
-     \omegadot(X_k) \\
-     \omegadot(\alpha_l) \\
-     0 \\
-     \Gamma_1 p \sigma \Sdot \\
-     \rho \Sdot
-   \end{array} \right )
-\end{equation}
-where
-\begin{equation}
-\sigma \equiv \frac{\partial p/\partial T |_\rho}{\rho c_p \partial p/\partial \rho |_T}
-\end{equation}
-and $c_p$ is the specific heat at constant pressure, $c_p = \partial
-h/\partial T |_p$.  A derivation of this source for the pressure
-equation can be found in \cite{ABNZ:III}.  We note that this source is
-algebraically identical to that shown in Eq.~25 of \cite{castro}.
-
-The CTU+PPM method for hydrodynamics is second-order accurate in space
-and time.  We want to couple the reactive sources to the hydrodynamics
-to be second-order in time as well.  As discussed above, nuclear
-reaction sources are stiff, and need to be integrated using implicit
-methods for stabilty.  Operator splitting (e.g., Strang) is
-traditionally employed here, and is used as a benchmark for comparison
-in this paper.  We'll discuss this traditional approach next before moving
-on to our new time-coupling method.
-
-\subsection{Strang Splitting}
-
-In Strang splitting, we first integrate the system with reactions terms only (no advection)
-over $\Delta t/2$, then integrate the advection terms only (no reactions) over $\Delta t$,
-and finally integrate the reaction terms only over $\Delta t/2$.
-
-In the absence of advective terms, our reaction system appears as just
-$d\Uc/dt = \Rb(\Uc)$, or:
-\begin{align}
-\odt{\rho} & = 0 \\
-\odt{(\rho X_k)} &= \rho \omegadot(X_k) \\
-\odt{(\rho \alpha_l)} &= \rho \omegadot(\alpha_l) \\
-\odt{(\rho u)} &= 0 \\
-\odt{(\rho E)} &= \rho \Sdot
-\end{align}
-We can write the energy equation as:
-\begin{equation}
-\odt{(\rho E)} = \odt{(\rho e)} + \odt{K} = \rho \Sdot
-\end{equation}
-where $K$ is the kinetic energy, $K = |u|^2/2$.  Since the density and velocity
-are unchanged by reactions, our energy equation becomes:
-\begin{equation}
-\label{eq:strang:e}
-\odt{(\rho e)} = \rho \odt{e} = \rho \Sdot
-\end{equation}
-The reaction rates are typically expressed as $\omegadot_k(\rho, T, X_k)$
-when we evolve this system, which requires us to get temperature from
-the equation of state each
-time we need to evaluate the reactive terms.
-
-We also typically integrate mass fraction itself, instead of partial
-densities:
-\begin{equation}
-\label{eq:strang:X}
-\odt{X_k} = \omegadot_k
-\end{equation}
-We integrate Equations (\ref{eq:strang:e}) and (\ref{eq:strang:X}) using
-an implicit ODE solver designed for stiff systems of
-equations, VODE~\citep{vode}.
-
-We explored this and other approaches (including not evolving an
-energy / temperature equation during the reaction step) in
-\citet{strang_rnaas} and showed that the above formulation gets second
-order convergence.  However, for very strong reactions, when using
-Strang splitting the state can drift significantly off of the smooth
-solution to the coupled reactive hydrodynamics equations, as shown
-graphically in \cite{astronum:2018} using an earlier version of the
-present algorithm.
-
-A variation on Strang splitting called (re-)balanced splitting was
-developed in \citet{speth:2013}.
-
-\subsection{Timestep Limiters and Retry Mechanism}
-
-Since this method is based off of the CTU hydrodynamics scheme, it
-benefits from the larger timestep that method can take (when done with
-full corner coupling, the advective CFL condition is unity)
-as compared to a method-of-lines approach (see
-\citealt{ppmunsplit}).  In addition to the standard CFL timestep
-limiter for explicit hydrodynamics, \castro\ can also enforce
-timestep limiters based on the energy generation or abundance changes
-over a timestep: \MarginPar{do we enforce this in the NSE region?}
-\begin{align}
-\label{eq:dt:nuce}
-\Delta t &\le f_e\, \min_{i} \left\{\frac{e_{i}}{\dot{S}_{i}}\right\} \\
-%
-\label{eq:dt:nucX}
-\Delta t &\le f_X\, \min_{i} \left\{\min_{k,X_k > \epsilon_X}\frac{{X_k}_{i}}{{\omegadot(X_k)_i}}\right\}
-\end{align}
-where $i$ is the zone index and $f_e$ and $f_X$ are runtime parameters
-used to control the allowed change, and only species for which $X_k > \epsilon_X$ are considered.
-
-\castro\ has the ability to reject a timestep if it detects a failure
-and retry with smaller timesteps (subcycling to make up the original
-required timestep).  Among the conditions that can trigger this are
-density falling below zero during advection, the ODE integration
-failing to converge in the implicit solve, or violation of one of the
-timestep limiters during the step.  This means that equations
-(\ref{eq:dt:nuce}) and (\ref{eq:dt:nucX}) are not reactive, but instead
-guaranteed to be met throughout the simulation, because the step is
-rejected if they are violated.  This contrasts to similar approaches
-used in other codes where conditions \ref{eq:dt:nuce} and
-\ref{eq:dt:nucX} are used to restrict the next timestep size, but the
-update over the current step is kept, which already violated the
-constraints. \MarginPar{would be nice to have a reference here} The retry mechanism in \castro\ works with both the
-Strang and simplified-SDC integration scheme.
-
-\subsection{Spectral Deferred Corrections}
-
-The basic idea of spectral deferred corrections \MarginPar{ref} is to express the
-update as an integral and divide the time-update into a number of
-discrete time nodes.  The integral is then approximated using a
-quadrature rule over these time nodes and low order approximations are
-used to update the state from one time node to the next.  The method
-uses iteration to successively improve the solution and the ultimate
-accuracy is determined by the quadrature method used to evaluate the
-integral, which is done using an iteratively-lagged solution.  This is the
-classic approach that we implemented in \cite{castro_sdc}.
-
-Our simplified-SDC approach is based on the method described in \citet{SDC-old} and
-a similar implementation in the \maestro\ simulation code.
-We again start by considering the update
-in integral form:
-\begin{equation}
-\Uc^{n+1} = \Uc^n + \int_t^{t+\Delta t} \left [ \Advs{\Uc} + \Rb(\Uc) \right ] dt,
-\end{equation}
-In this approach, we 
-approximate the advective term
-piecewise constant in time, using the value at the midpoint in time,
-to achieve second-order accuracy, giving us:
-\begin{equation}
-\label{eq:integral:simplesdc}
-\Uc^{n+1} = \Uc^n + \int_t^{t+\Delta t} \left \{ \Adv{\Uc}^{n+1/2} + \Rb(\Uc) \right \} dt
-\end{equation}
-The \castro\ version of the simplified-SDC algorithm
-differs from the previous versions due to the need to do some operations
-on the conserved variable state and some on the primitive variable
-state.
-
-The basic time update algorithm proceeds as:
-
-\begin{itemize}
-
-\item {\em Initialization}
-
-\label{sec:initialization}
-
-  \begin{itemize}
-  \item We need an approximation of how much the reactions alone
-    changed the primitive variable state over the timestep, which we
-    will denote $\Ic_q$ (this is basically an approximation of
-    $\Rb(q)$).
-
-    Since we do not have any information about the current
-    timestep in the first iteration, we use the value from the last
-    iteration of the previous timestep:
-    \begin{equation}
-      \Ic^{n+1/2,(0)}_{\qb} = \Ic^{n-1/2,(\smax)}_{\qb}
-    \end{equation}
-  %% \item Solve the Poisson problem for the initial gravitational potential:
-  %%   \begin{equation}
-  %%     \nabla^2 \Phi^n = 4\pi G \rho^n
-  %%   \end{equation}
-
-  %% \item Set the first guess at the new time potential as
-  %%   $\Phi^{n+1,(0)} = \Phi^n$.
-
-  \end{itemize}
-
-\item {\em Iterate}
-
-  Iterate from $k = 1, \ldots, \smax$.  For second-order accuracy,
-  $\smax = 2$ is sufficient.  In addition to denoting the time-level
-  with a superscript (like $n$ or $n+1$), we'll use a second subscript
-  in parentheses to keep track of the iteration.  A single iteration
-  starts with $\Uc^n$ and results in the new time-level state for that
-  iteration, $\Uc^{n+1,(k)}$.
-
-  \begin{itemize}
-  \item {\em Create the advective update term, $\Adv{\Uc}^{n+1/2,(k)}$}
-
-    \begin{itemize}
-    \item convert $\Uc \rightarrow \qb$.  This is an algebraic transformation,
-      but will require the EOS.
-
-    \item predict $\qb$ to the interfaces at $t^{n+1/2}$ using the CTU PPM
-      method.  The source terms, $\Sq$, used in the prediction are:
-      \begin{equation}
-        \Sc(\qb) = \Hb(\qb) + \Ic_\qb^{n+1/2,(k-1)}
-      \end{equation}
-      Here we use the iteratively lagged integrals of the primitive variable
-      terms accounting only for reactions, $\Ic_\qb^{n+1/2,(k-1)}$, as the
-      reactive source.  This is in contrast to Strang-splitting, where no
-      explicit reactive source terms are seen by the hydrodynamics update.
-%%       Any hydrodynamic source terms are time-centered
-%%       using the previous iteration:
-%% \MarginPar{Formally both $t^n$ or $t^{n+1/2}$ sources give you 2nd-order since they are CTU predictor source terms.  I've never found this type of iteration for this term to amount to any benefit}
-%%       \begin{equation}
-%%         \Sqhydro^{n+1/2} = \frac{1}{2} \left ( \Sqhydro^n + \Sqhydro^{n+1,(k-1)} \right )
-%%       \end{equation}
-
-      In the unsplit CTU method \citep{ppmunsplit}, the interface
-      states used for the final Riemann problem through the zone
-      interface consist of a normal predictor and a transverse flux
-      correction.  We can add the source terms either to the normal
-      predictor (for example, doing characteristic tracing as
-      described in \citealt{ppm}) or after all of the transverse flux
-      corrections are made.  Both are second-order accurate.  For the
-      hydrodynamics sources, we do those in the normal predictor.
-      However, for the reactive sources, we found that it is most
-      reliable to add them at the end of the interface state
-      construction, after the transverse flux corrections.  This is
-      because we want to ensure that sum over species of $\Ic_\qb$ is
-      zero, and characteristic tracing or the various flux corrections
-      may not preserve this.  We enforce that the species interface
-      states remain in $[0, 1]$ after adding the reactive source.
-
-    \item solve the Riemann problem at each interface to get a unique
-      conserved state on each interface, $\Uc^{n+1/2,(k)}_{i+1/2}$
-
-    \item construct the advective update terms, $\Advt{\Uc}^{n+1/2,(k)}_{i}$,
-      by first ignoring the hydrodynamics sources,
-      \begin{equation}
-        \Advt{\Uc}^{n+1/2,(k)}_{i} =
-          - \frac{\Fb^\xv(\Uc^{n+1/2,(k)}_{i+1/2}) - \Fb^\xv(\Uc^{n+1/2,(k)}_{i-1/2})}{\Delta x}
-      \end{equation}
-      for the general case, and
-      \begin{align}
-        \Advsst{\rho e}^{n+1/2,(k)}_{i} =
-          &- \frac{F\left((\rho e)^{n+1/2,(k)}_{i+1/2}\right) - F\left((\rho e)^{n+1/2,(k)}_{i-1/2}\right)}{\Delta x} \\
-          &- \left (\frac{p_{i+1/2}^{n+1/2,(k)} + p_{i-1/2}^{n+1/2,(k)}}{2} \right )\frac{u_{i+1/2}^{n+1/2,(k)} - u_{i-1/2}^{n+1/2,(k)}}{\Delta x}
-      \end{align}
-      for $(\rho e)$.
-
-    Now the conservative hydrodynamics source terms are computed by first updating to the
-    new state with advection and the old-time source term applied for the full $\Delta t$ as: 
-    \begin{equation}
-      \Uc^{\star\star} = \Uc^n + \Delta t \Advt{\Uc}^{n+1/2,(k)} + \Delta t \Hb(\Uc^n)
-    \end{equation}
-    %% first updating the density to the new state with advection only as:
-    %% \begin{equation}
-    %%   \rho^{\star\star} = \rho^n + \Delta t \Advt{\rho}^{n+1/2,(k)}
-    %% \end{equation}
-    %% then constructing the momentum source term:
-    %% \begin{equation}
-    %%   {\bf S}_{\rho\Ub}^{n+1/2} = \frac{1}{2} (\rho^n \gb^n + \rho^{\star\star} \gb^{n+1,(k-1)})
-    %% \end{equation}
-    %% Then updating the momentum with only advection as:
-    %% \begin{equation}
-    %%   (\rho \Ub)^{\star\star} = (\rho \Ub)^n + \Delta t \Advt{\rho\Ub}^{n+1/2,(k)} + \Delta t {\bf S}_{\rho\Ub}^{n+1/2}
-    %% \end{equation}
-    %% and finally constructing the energy source:
-    %% \begin{equation}
-    %%   S_{\rho E}^{n+1/2} = \frac{1}{2} \left [ (\rho\Ub)^n \cdot \gb^n + (\rho\Ub)^{\star\star} \cdot \gb^{n+1,(k-1)}
-    %%     \right ]
-    %% \end{equation}
-    We then evaluate the source terms with $\Uc^{\star\star}$ and
-    correct the advective term so that we have a time-centered
-    source.
-    The final advective update term is then\footnote{For a source like gravity, this update can be done first for $\rho$ and then define the new momentum source using $\rho^{\star\star}$ and likewise for energy}:
-    \begin{equation}
-      \Adv{\Uc}^{n+1/2,(k)}_{i} = \Advt{\Uc}^{n+1/2,(k)}_{i} +
-      \frac{\Delta t}{2} \left (\Shydro(\Uc^n) + \Shydro(\Uc^{\star\star}) \right )
-    \end{equation}
-
-    %% with
-    %% \begin{equation}
-    %%   \Shydro^{n+1/2} = \left ( \begin{array}{c}
-    %%                  0 \\ 0 \\ 0 \\
-    %%                 {\bf S}_{\rho\Ub}^{n+1/2} \cdot \ex \\
-    %%                 S_{\rho E}^{n+1/2} \end{array} \right )
-    %% \end{equation}
-
-    We note that this procedure works because the source terms here do
-    not depend on the outcome of the reactions.
-
-    \end{itemize}
-
-  \item {\em Update the System Using a Method of Lines Integration}
-
-    We update the state by doing the integral in equation
-    (\ref{eq:integral:simplesdc}).  Since we are approximating the
-    advective term as piecewise constant in time, we can simply use an
-    ODE integrator to integrate this, just as we do with the reaction
-    system in Strang splitting.  The difference here being that we are
-    integrating the conserved variables and the state sees the effect
-    of advection as we integrate the reactions.  So rather than use
-    $d\Uc/dt = \Rb(\Uc)$, the ODE form we use is
-    \begin{equation}
-      \odt{\Uc} = \Adv{\Uc}^{n+1/2,(k)} + \Rb(\Uc)
-    \end{equation}
-
-    Looking at the form of Eq.~\ref{eq:react_source}, we see that only
-    the species, auxiliary quantities, and energies have source terms.
-    Since we are reacting, we will use the $X_k$ as the primary
-    composition variable and compute $\alpha_l$ as needed from them,
-    so we don't directly integrate the auxiliary quantities.
-    Likewise, the total and internal energies both provide the same
-    information, and integrating both overconstrains the system, so we
-    just integrate the internal energy.  We define the subset of variables
-    that are directly integrated as:
-    \begin{equation}
-      \Uc^\prime = \left ( \begin{array}{c} \rho X_k \\ \rho e \end{array} \right )
-    \end{equation}
-    and we integrate
-    \begin{equation}
-      \odt{\Uc^\prime} = \Adv{\Uc^\prime}^{n+1/2,(k)} + \Rb(\Uc^\prime)
-    \end{equation}
-    This integration begins with $\Uc^{\prime,n}$ and results in $\Uc^{\prime,n+1,(k)}$.
-
-    We will need the density at times during the integration, which we construct as:
-    \begin{equation}
-      \rho(t) = \rho^n + \Advss{\rho}^{n+1/2,(k)} \, t
-    \end{equation}
-    As we are integrating this system we need to get the temperature,
-    $T$, for the rate evaluations.  We obtain this by directly from
-    internal energy, composition, and density using the equation of
-    state.
-
-    Our integrator also needs the Jacobian of the system, in terms of
-    the $\Uc^\prime$.  This is different than the form of the
-    Jacobian usually used in reaction networks.  We describe the form
-    of the Jacobian in appendix \ref{sec:app:jacobian}.
-
-    At the end of the integration, we can do the conservative update of momentum
-    and energy.  Momentum is straightforward, since there are no reactive sources:
-    \begin{equation}
-      (\rho u)^{n+1} = (\rho u)^n + \Delta t \Advss{\rho u}^{n+1/2,(k)}
-    \end{equation}
-    For total energy, we first need to isolate the reactive source for energy:
-    \begin{equation}
-      (\rho \Sdot)^{n+1/2} = \frac{(\rho e)^{n+1} - (\rho e)^n}{\Delta t} - \Advss{\rho e}^{n+1/2,(k)}
-    \end{equation}
-    Then the update is
-    \begin{equation}
-      (\rho E)^{n+1} = (\rho E)^n + \Delta t \Advss{\rho E}^{n+1/2,(k)} + \Delta t (\rho \Sdot)^{n+1/2}
-    \end{equation}
-
-  \item {\em Compute the Reactive Source Terms.}
-
-    We now seek the $\Ic$'s that capture the effect of just the
-    reaction sources on the state variables for the next iteration.
-    For the conserved quantities, these would simply be:
-    \begin{equation}
-      \Ic^{(k)}_{\Uc} = \frac{\Uc^{n+1,(k)} - \Uc^n}{\Delta t} - \Adv{\Uc}^{n+1/2,(k)}
-    \end{equation}
-    However, for our primitive variables, which are used in the
-    prediction, we need to construct the required source terms more
-    carefully.  We want:
-    \begin{equation}
-      \label{eq:Iq}
-      \Ic^{(k)}_{\qb} = \frac{\qb^{n+1,(k)} - \qb^n}{\Delta t} - \Adv{\qb}^{n+1/2,(k)}
-    \end{equation}
-    but we need the advective update for $\qb$, which we have not
-    constructed.  We note that $\Ic^{(k)}_\qb$ is an approximation to the integral of 
-    Eq.~\ref{eq:prim_sources} over the timestep.  Additionally, we cannot simply use the equation of
-    state on $\Ic^{(k)}_{\Uc}$ since this is a time-derivative and
-    does not represent a well-defined state in itself.  Instead, we
-    derive $\Ic^{(k)}_{\qb}$ via a multi-step process.  We first find
-    the conservative state as if it were updated only with advection:
-    \begin{equation}
-      \Uc^\star = \Uc^n + \Delta t \Adv{\Uc}^{n+1/2,(k)}
-    \end{equation}
-    and then construct the corresponding primitive variable state via an algebraic transform,
-    $\Uc^\star \rightarrow \qb^\star$.
-    This allows us to define the advective update for $\qb$ as:
-    \begin{equation}
-      \Adv{\qb}^{n+1/2,(k)} = \frac{\qb^\star - \qb^n}{\Delta t}
-    \end{equation}
-    Defining the primitive state corresponding to the fully-updated
-    conserved state via an algebraic transform, $\Uc^{n+1,(k)}
-    \rightarrow \qb^{n+1,(k)}$, we can construct $\Ic^{(k)}_{\qb}$ as given
-    in Equation (\ref{eq:Iq}).
-    Putting all of this together, we see:
-    \begin{equation}
-      \Ic^{(k)}_{\qb} = \frac{\qb^{n+1,(k)} - \qb^\star}{\Delta t}
-    \end{equation}
-
-  %% \item {\em Solve for the New Gravitational Potential.}
-
-  %%   We solve
-  %%   \begin{equation}
-  %%     \nabla^2 \Phi^{n+1,(k)} = 4\pi G \rho^{n+1,(k)}
-  %%   \end{equation}
-
-  \end{itemize}
-
-\end{itemize}
-
 \section{Reaction Networks and Nuclear Statistical Equilibrium}
 
-For this study we will use the 19 nuclei network containing
-\isot{H}{1}, \isot{He}{3}, \isot{He}{4}, \isot{C}{12}, \isot{N}{14},
-\isot{O}{16}, \isot{Ne}{20}, \isot{Mg}{24}, \isot{Si}{28},
-\isot{S}{32}, \isot{Ar}{36}, \isot{Ca}{40}, \isot{Ti}{44},
-\isot{Cr}{48}, \isot{Fe}{52}, \isot{Fe}{54}, \isot{Ni}{56}, protons
-(from photodisintegration), and neutrons.  This is based on the
-``aprox19'' network from \cite{aprox19} and originally described in
-\cite{Kepler}.  We use a modified version of the VODE~\citep{vode}
-integrator---ported to C++ with checks added to the timestep rejection
-logic that ensure that the species mass fractions stay between 0 and \MarginPar{we need to be clear that there is different logic for SDC}
-1\footnote{This modified version of VODE is available in our
-Microphysics repo:
-\url{https://github.com/starkiller-astro/Microphysics}.}.\MarginPar{discuss
-  tolerances} We combine this with a table that gives the nuclear
-statistical equilibrium (NSE) abundances in regions where the system
-is in NSE.  The NSE table was generated using a 127 nuclei reaction
-network and is the same as described in \cite{ma:2013}.  In our
-simulations, we carry all 19 isotopes in the main network in each zone
-and advect them in the hydrodynamics portion of the algorithm.  The
-composition of the larger 127 nuclei network is mapped into the 19
-isotopes we carry according to Table \ref{table:nuclei}.
-
-The NSE table requires: \MarginPar{need a statement about how the table was constructed and what NSE means}
+The NSE table requires:
 \begin{equation}
@@
 (see, e.g., \citealt{flash}).
-
-\subsection{Initialization}
-
-We initialize the mass fractions, $X_k$, and then compute the
-electron fraction, $Y_e$, using (\ref{eq:aux:ye}).
-For the two other composition
-quantities we carry, $\bar{A}$, and $(B/A)$, we need values that are
-consistent with the value of $Y_e$ and the nuclei stored in the NSE
-table.  Therefore, if the thermodynamic conditions put us in NSE
-(using the conditions defined below), then we obtain $\bar{A}$ and
-$(B/A)$ from the NSE table.  Otherwise, we compute these directly from
-$X_k$ using (\ref{eq:aux:abar}) and (\ref{eq:aux:bea}).
-
-\subsection{NSE condition}
-
-We treat a zone as being in NSE if the density and temperature exceed
-some threshold and the composition is mainly $\alpha$ and Fe-group
-nuclei, where the Fe-group nuclei are \isot{Cr}{48}, \isot{Fe}{52},
-\isot{Fe}{54}, and \isot{Ni}{56}.  The full condition is:
-\begin{align}
-\rho &> \rhonse \\
-T &> \tnse \\
-X(\isotm{C}{12}) &< \Anse \\
-X(\isotm{He}{4}) + \sum_{k \in \mathrm{Fe-group}} X_k &> \Bnse \\
-X(\isotm{Si}{26}) &< \Cnse 
-\end{align}
-Typical values are $\rhonse = 2\times 10^6~\gcc$, $T_\mathrm{nse} =
-3\times 10^9~\mathrm{K}$, $\Anse = 0.12$, $\Bnse=
-0.88$, and $\Cnse = 0.01$.  The values of $\Anse$ and $\Bnse$ are based on \citet{ma:2013}.
-The value of $\Cnse$ ensures that Si-burning has ended before invoking the table.
 
@@
       \end{equation}
-      In this formulation, we have not updated the temperature. \MarginPar{discuss this more, and perhaps switch to a mode where we try to predict T}
+      In this formulation, we have not updated the temperature. 
       This corrects the composition so that it is consistent with the updated
@@
       \end{equation}
-      \MarginPar{we could imagine looping over the EOS to get an updated T and then the table to get the updated $\bar{A}$, to see if that converges?}
 
@@
 
-\subsection{SDC-NSE Coupling}
+\subsection{SDC-NSE Coupling Approach I}
+
+This is the approach on the {\tt development} branch.
 
@@
 
-\subsection{NSE Bailout}
-
-There is one additional part of the time integration algorithm.
-Because $T$ evolves during the reactions (for both Strang and SDC), it
-is possible for a zone to start out not in NSE but evolve into NSE
-during the reaction update.  When this happens, the ODE integrator may
-fail, because an excessive number of timesteps is required.  Rather
-than trigger our retry mechanism and throw away the entire timestep
-and start over with a smaller $\Delta t$, we first check if the zone
-evolved into NSE, and if so, we redo the zone update using the NSE
-prescription.  To ensure that this is not triggered at the very start
-of integration, we require a minimum number of steps (10) to have
-elapsed before checking for NSE.  After we leave the integrator, the
-state is returned to the burner driver where we finish the integration
-via NSE if the conditions satisfy the NSE criteria.  We optionally
-apply a relaxation factor, typically, $0.9$, to the NSE conditions to
-allow a state that is close to NSE after an integrator failure enter
-NSE.
-
-\section{Simulations}
-
-\subsection{Reacting convergence test problem}
-
-\cite{castro_sdc} introduced a test problem for measuring convergence
-of a reacting hydrodynamic algorithm.  We run that same test here with the
-simplified SDC algorithm.
-
-Also explore: PLM vs PPM, 1 vs. 2 vs. 3 iterations
-
-% run on 04-06-21
-% Castro       git describe: 21.04-dirty
-% AMReX        git describe: 21.03-58-g87c81d0ca
-% Microphysics git describe: 21.04-1-g33a4d0fe
-
-\begin{deluxetable}{lllllll}
-\tablecaption{\label{table:react_convert_strang} Convergence ($L_1$ norm) for the reacting convergence problem using Strang splitting.}
-\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
-\startdata
- $\rho$                      & $2.794 \times 10^{18}$  & 2.044  & $6.777 \times 10^{17}$  & 2.554  & $1.154 \times 10^{17}$  \\
- $\rho u$                    & $6.796 \times 10^{26}$  & 2.448  & $1.245 \times 10^{26}$  & 2.889  & $1.681 \times 10^{25}$  \\
- $\rho v$                    & $6.796 \times 10^{26}$  & 2.448  & $1.245 \times 10^{26}$  & 2.889  & $1.681 \times 10^{25}$  \\
- $\rho E$                    & $2.451 \times 10^{35}$  & 2.351  & $4.803 \times 10^{34}$  & 2.742  & $7.179 \times 10^{33}$  \\
- $\rho e$                    & $2.261 \times 10^{35}$  & 2.320  & $4.526 \times 10^{34}$  & 2.821  & $6.403 \times 10^{33}$  \\
- $T$                         & $2.237 \times 10^{21}$  & 1.691  & $6.927 \times 10^{20}$  & 2.482  & $1.240 \times 10^{20}$  \\
- $\rho X(\isotm{He}{4})$     & $2.878 \times 10^{18}$  & 2.020  & $7.096 \times 10^{17}$  & 2.529  & $1.229 \times 10^{17}$  \\
- $\rho X(\isotm{C}{12})$     & $1.698 \times 10^{17}$  & 1.950  & $4.393 \times 10^{16}$  & 2.232  & $9.353 \times 10^{15}$  \\
- $\rho X(\isotm{O}{16})$     & $1.687 \times 10^{14}$  & 1.660  & $5.338 \times 10^{13}$  & 1.957  & $1.375 \times 10^{13}$  \\
- $\rho X(\isotm{Fe}{56})$    & $2.794 \times 10^{8}$   & 2.044  & $6.777 \times 10^{7}$   & 2.554  & $1.154 \times 10^{7}$   \\
-\enddata
-\end{deluxetable}
-
-% run on 04-06-21
-% Castro       git describe: 21.04-dirty
-% AMReX        git describe: 21.03-58-g87c81d0ca
-% Microphysics git describe: 21.04-1-g33a4d0fe
-
-\begin{deluxetable}{lllllll}
-\tablecaption{\label{table:react_convert_sdc} Convergence ($L_1$ norm) for the reacting convergence problem using SDC integration (2 iterations).}
-\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
-\startdata
- $\rho$                      & $2.795 \times 10^{18}$  & 2.044  & $6.777 \times 10^{17}$  & 2.554  & $1.154 \times 10^{17}$  \\
- $\rho u$                    & $6.798 \times 10^{26}$  & 2.449  & $1.245 \times 10^{26}$  & 2.889  & $1.680 \times 10^{25}$  \\
- $\rho v$                    & $6.798 \times 10^{26}$  & 2.449  & $1.245 \times 10^{26}$  & 2.889  & $1.680 \times 10^{25}$  \\
- $\rho E$                    & $2.452 \times 10^{35}$  & 2.352  & $4.804 \times 10^{34}$  & 2.742  & $7.179 \times 10^{33}$  \\
- $\rho e$                    & $2.262 \times 10^{35}$  & 2.321  & $4.527 \times 10^{34}$  & 2.822  & $6.404 \times 10^{33}$  \\
- $T$                         & $2.236 \times 10^{21}$  & 1.691  & $6.927 \times 10^{20}$  & 2.481  & $1.240 \times 10^{20}$  \\
- $\rho X(\isotm{He}{4})$     & $2.879 \times 10^{18}$  & 2.020  & $7.096 \times 10^{17}$  & 2.529  & $1.229 \times 10^{17}$  \\
- $\rho X(\isotm{C}{12})$     & $1.697 \times 10^{17}$  & 1.950  & $4.393 \times 10^{16}$  & 2.231  & $9.357 \times 10^{15}$  \\
- $\rho X(\isotm{O}{16})$     & $1.686 \times 10^{14}$  & 1.659  & $5.338 \times 10^{13}$  & 1.957  & $1.375 \times 10^{13}$  \\
- $\rho X(\isotm{Fe}{56})$    & $2.795 \times 10^{8}$   & 2.044  & $6.777 \times 10^{7}$   & 2.554  & $1.154 \times 10^{7}$   \\
-\enddata
-\end{deluxetable}
-
-\subsection{NSE convergence test problem}
-
-To test the coupling of the NSE table to the hydrodynamics, we run a
-similar problem as above, except with the thermodynamic conditions
-appropriate for the matter to be in NSE.  The initial conditions are:
-\begin{align}
-\rho &= \rho_0 \\
-T &= T_0 \left [ 1 + (\delta T) e^{-(r/\lambda)^2} \cos^6(\pi r/L) \right ] \\
-Y_e &= (Y_e)_0 \left [ 1 + (\delta Y_e) e^{-(r/\lambda)^2} \cos^6(\pi r/L) \right ] 
-\end{align}
-where $r$ is the distance from the center of the domain and the
-remaining parameters are listed in Table~\ref{table:nse}.  This has
-$Y_e$ varying between $[0.47, 0.5]$ initially\MarginPar{check}.  The mass fractions
-and remaining composition variables are then initialized from the NSE table, as described in section~\ref{sec:initialization}.
-
-The domain has a size $[0, L]^2$, and the timestep is fixed as:
+\begin{tcolorbox}
+{\bf Issues:} I think that when we call 
 \begin{equation}
-\Delta t = 10^{-3} \left ( \frac{64}{N} \right )~\mathrm{s}
-\end{equation}
-where $N$ is the number of zones in each direction.
-
-\begin{deluxetable}{lcc}
-\tablecaption{\label{table:nse} NSE convergence test problem.}
-\tablehead{\colhead{parameter} & \colhead{value}}
-\startdata
-$\rho_0$ & $5\times 10^8~\gcc$ \\
-$T_0$    & $4\times 10^9~\mathrm{K}$ \\
-$(\delta T)$ & $0.2$ \\
-$(Y_e)_0$   & $0.5$ \\
-$(\delta Y_e)$ & $-0.05$ \\
-$\lambda$   & $2\times 10^7~\mathrm{cm}$ \\
-$L$         & $10^8~\mathrm{cm}$ \\
-\enddata
-\end{deluxetable}
-
-Figure~\ref{fig:nse_ye} shows the $Y_e$ profile for the Strang evolution case.
-
-\begin{figure}[t]
-\centering
-\plotone{ye_plot}
-\caption{\label{fig:nse_ye} Profile of $Y_e$ through the middle of the domain (slicing along $x$)
-for the $128^2$ Strang NSE test simulation.}
-\end{figure}
-
-% run on 04-06-21
-% Castro       git describe: 21.04-dirty
-% AMReX        git describe: 21.03-58-g87c81d0ca
-% Microphysics git describe: 21.04-1-g33a4d0fe
-
-\begin{deluxetable}{lllllll}
-\tablecaption{\label{table:nse_strang_methodII} Convergence ($L_1$ norm) for the NSE convergence
-problem using Strang splitting.}
-\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
-\startdata
- $\rho$                      & $4.522 \times 10^{19}$  & 1.659  & $1.432 \times 10^{19}$  & 1.542  & $4.917 \times 10^{18}$  \\
- $\rho u$                    & $1.883 \times 10^{28}$  & 1.681  & $5.873 \times 10^{27}$  & 1.524  & $2.042 \times 10^{27}$  \\
- $\rho v$                    & $1.883 \times 10^{28}$  & 1.681  & $5.873 \times 10^{27}$  & 1.524  & $2.042 \times 10^{27}$  \\
- $\rho E$                    & $5.933 \times 10^{37}$  & 1.613  & $1.940 \times 10^{37}$  & 1.411  & $7.293 \times 10^{36}$  \\
- $\rho e$                    & $5.931 \times 10^{37}$  & 1.613  & $1.939 \times 10^{37}$  & 1.411  & $7.291 \times 10^{36}$  \\
- $T$                         & $9.885 \times 10^{20}$  & 1.207  & $4.281 \times 10^{20}$  & 1.058  & $2.057 \times 10^{20}$  \\
- $\rho X(\isotm{H}{1})$      & $4.522 \times 10^{-11}$ & 1.659  & $1.432 \times 10^{-11}$ & 1.542  & $4.917 \times 10^{-12}$ \\
- $\rho X(\isotm{He}{4})$     & $5.086 \times 10^{17}$  & 1.430  & $1.887 \times 10^{17}$  & 1.187  & $8.288 \times 10^{16}$  \\
- $\rho X(\isotm{C}{12})$     & $1.585 \times 10^{12}$  & 1.829  & $4.462 \times 10^{11}$  & 1.433  & $1.652 \times 10^{11}$  \\
- $\rho X(\isotm{O}{16})$     & $8.060 \times 10^{12}$  & 1.706  & $2.470 \times 10^{12}$  & 1.347  & $9.710 \times 10^{11}$  \\
- $\rho X(\isotm{Cr}{48})$    & $1.149 \times 10^{19}$  & 1.400  & $4.353 \times 10^{18}$  & 1.258  & $1.820 \times 10^{18}$  \\
- $\rho X(\isotm{Fe}{52})$    & $9.018 \times 10^{19}$  & 1.173  & $4.001 \times 10^{19}$  & 1.127  & $1.832 \times 10^{19}$  \\
- $\rho X(\isotm{Fe}{54})$    & $5.036 \times 10^{20}$  & 1.731  & $1.517 \times 10^{20}$  & 1.727  & $4.582 \times 10^{19}$  \\
- $\rho X(\isotm{Ni}{56})$    & $5.686 \times 10^{20}$  & 1.622  & $1.847 \times 10^{20}$  & 1.563  & $6.252 \times 10^{19}$  \\
- $\rho Y_e$                  & $2.379 \times 10^{19}$  & 1.678  & $7.431 \times 10^{18}$  & 1.578  & $2.489 \times 10^{18}$  \\
- $\rho \bar{A}$              & $5.141 \times 10^{21}$  & 1.230  & $2.191 \times 10^{21}$  & 1.087  & $1.032 \times 10^{21}$  \\
- $\rho (B/A)$                & $4.149 \times 10^{20}$  & 1.693  & $1.283 \times 10^{20}$  & 1.562  & $4.345 \times 10^{19}$  \\
-\enddata
-\end{deluxetable}
-
-% run on 04-06-21
-% Castro       git describe: 21.04-dirty
-% AMReX        git describe: 21.03-58-g87c81d0ca
-% Microphysics git describe: 21.04-1-g33a4d0fe
-
-\begin{deluxetable}{lllllll}
-\tablecaption{\label{table:nse_sdc_methodII} Convergence ($L_1$ norm) for the NSE convergence
-problem using the simplified-SDC algorithm.}
-\tablehead{\colhead{field} & \colhead{$\epsilon_{64 \rightarrow 128}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{128\rightarrow 256}$} & 
-           \colhead{rate} & \colhead{$\epsilon_{256\rightarrow 512}$}}
-\startdata
- $\rho$                      & $4.553 \times 10^{19}$  & 1.636  & $1.465 \times 10^{19}$  & 1.583  & $4.891 \times 10^{18}$  \\
- $\rho u$                    & $1.949 \times 10^{28}$  & 1.650  & $6.210 \times 10^{27}$  & 1.513  & $2.176 \times 10^{27}$  \\
- $\rho v$                    & $1.949 \times 10^{28}$  & 1.650  & $6.210 \times 10^{27}$  & 1.513  & $2.176 \times 10^{27}$  \\
- $\rho E$                    & $5.691 \times 10^{37}$  & 1.622  & $1.848 \times 10^{37}$  & 1.561  & $6.266 \times 10^{36}$  \\
- $\rho e$                    & $5.689 \times 10^{37}$  & 1.622  & $1.848 \times 10^{37}$  & 1.561  & $6.265 \times 10^{36}$  \\
- $T$                         & $5.778 \times 10^{20}$  & 1.740  & $1.729 \times 10^{20}$  & 1.702  & $5.316 \times 10^{19}$  \\
- $\rho X(\isotm{H}{1})$      & $4.553 \times 10^{-11}$ & 1.636  & $1.465 \times 10^{-11}$ & 1.583  & $4.891 \times 10^{-12}$ \\
- $\rho X(\isotm{He}{4})$     & $2.541 \times 10^{17}$  & 1.807  & $7.262 \times 10^{16}$  & 1.634  & $2.339 \times 10^{16}$  \\
- $\rho X(\isotm{C}{12})$     & $1.140 \times 10^{12}$  & 1.838  & $3.189 \times 10^{11}$  & 1.670  & $1.002 \times 10^{11}$  \\
- $\rho X(\isotm{O}{16})$     & $5.413 \times 10^{12}$  & 1.819  & $1.535 \times 10^{12}$  & 1.658  & $4.864 \times 10^{11}$  \\
- $\rho X(\isotm{Cr}{48})$    & $7.020 \times 10^{18}$  & 1.793  & $2.026 \times 10^{18}$  & 1.614  & $6.621 \times 10^{17}$  \\
- $\rho X(\isotm{Fe}{52})$    & $3.631 \times 10^{19}$  & 1.579  & $1.215 \times 10^{19}$  & 1.597  & $4.018 \times 10^{18}$  \\
- $\rho X(\isotm{Fe}{54})$    & $4.670 \times 10^{20}$  & 1.807  & $1.335 \times 10^{20}$  & 1.869  & $3.654 \times 10^{19}$  \\
- $\rho X(\isotm{Ni}{56})$    & $4.858 \times 10^{20}$  & 1.808  & $1.387 \times 10^{20}$  & 1.860  & $3.820 \times 10^{19}$  \\
- $\rho Y_e$                  & $2.372 \times 10^{19}$  & 1.649  & $7.564 \times 10^{18}$  & 1.587  & $2.518 \times 10^{18}$  \\
- $\rho \bar{A}$              & $4.009 \times 10^{21}$  & 1.693  & $1.240 \times 10^{21}$  & 1.690  & $3.843 \times 10^{20}$  \\
- $\rho (B/A)$                & $4.222 \times 10^{20}$  & 1.666  & $1.330 \times 10^{20}$  & 1.605  & $4.374 \times 10^{19}$  \\
-\enddata
-\end{deluxetable}
-
-\subsection{Detonation}
-
-The purpose here is to look at what timestep is taken when we use the nuclear burning limiters.
-Run with aprox19 only, Strang and SDC, then aprox19 + NSE, Strang and NSE.
-
-Look at number of SDC iterations too.
-
-How does the timestep limiter work in the middle of SDC iterating?
-
-aprox21
-
-look at number of RHS calls, wallclock time
-
-output every step and then for a single zone plot enuc every timestep to see if it oscillates
-
-\subsection{Reacting Buoyant Bubble}
-
-We next consider a buoyant, reacting bubble in a stratified,
-plane-parallel atmosphere.  This test exercises the treatment of
-hydrodynamical sources in the SDC integration.  We use the set of
-initial conditions that was first shown in \cite{ABNZ:III}.
-
-\subsection{Massive Star}
-
-Do this in 1-d
-Show the amount of work needed for Strang vs. SDC
-
-Play around with the threshold where NSE kicks in
-
-\section{Summary}
-
-We presented a simplified spectral deferred corrections scheme for coupling
-hydrodynamics and reactions.
-
-Future work is to extend this methodology to MHD.
-
-\acknowledgements \castro\ is freely available at
-\url{http://github.com/AMReX-Astro/Castro}.  All of the code and
-problem setups used here are available in the git repo.  The work at
-Stony Brook was supported by DOE/Office of Nuclear Physics grant
-DE-FG02-87ER40317.  This material is based upon work supported by the
-U.S. Department of Energy, Office of Science, Office of Advanced
-Scientific Computing Research and Office of Nuclear Physics,
-Scientific Discovery through Advanced Computing (SciDAC) program under
-Award Number DE-SC0017955.  This research was supported by the
-Exascale Computing Project (17-SC-20-SC), a collaborative effort of
-the U.S. Department of Energy Office of Science and the National
-Nuclear Security Administration.
-
-\software{\amrex\ \citep{amrex_joss},
-          \castro\ \citep{castro},
-          GNU Compiler Collection (\url{https://gcc.gnu.org/}),
-          Linux (\url{https://www.kernel.org}),
-          matplotlib (\citealt{Hunter:2007},  \url{http://matplotlib.org/})
-          NumPy \citep{numpy,numpy2},
-          python (\url{https://www.python.org/}), 
-          SymPy \citep{sympy}
-         }
-
-\appendix
-
-\section{Jacobian}
-
-\label{sec:app:jacobian}
-
-To solve the reaction system implicitly, the ODE solver needs the
-Jacobian, $\partial \Rb/\partial \Uc^\prime$, where $\Uc^\prime =
-(\rho X_k, \rho e)^\intercal$ is the subset of the conserved variables we are
-integrating.  We follow the method of \cite{castro_sdc} and factor
-this into two pieces,
-\begin{equation}
-\label{eq:factored_jac}
-{\bf J} = \frac{\partial \Rb}{\partial {\bf w }} \frac{\partial {\bf w}}{\partial \Uc^\prime}.
-\end{equation}
-where the state ${\bf w}$ is chosen to match the set of variables used to evaluate the reaction rates.
-Writing this out for two species, $X_\alpha$ and $X_\beta$, we have
-\begin{equation}
-\Uc^\prime = \left ( \begin{array}{c} \rho X_\alpha \\ \rho X_\beta \\ \rho e \end{array} \right )
-\end{equation}
-For interfacing with the reaction network, we use
-\begin{equation}
-{\bf w} = \left ( \begin{array}{c} X_\alpha \\  X_\beta \\ T \end{array} \right )
-\end{equation}
-Note: even though we are using $T$ here instead of $e$, we still do the overall ODE integration
-in terms of $(\rho e)$, consistent with the Strang method described in \cite{strang_rnaas}.
-%%  in that
-%% we use $e$ instead of $T$ as the independent variable.  If the network gives derivatives in terms
-%% of temperature, we can convert as:
-%% \begin{equation}
-%% \left . \frac{\partial \phi}{\partial T} \right |_\rho =
-%%   \left . \frac{\partial \phi}{\partial e} \right |_\rho 
-%%   \left . \frac{\partial e}{\partial T} \right |_\rho =
-%%   c_v \left . \frac{\partial \phi}{\partial e} \right |_\rho
-%% \end{equation}
-%% for some quantity $\phi(\rho, T)$.  Here we identified the specific heat
-%% at constant volume, $c_v = \partial e/\partial T|_\rho$.
-
-The Jacobian transformation $\partial \Uc^\prime/\partial {\bf w}$ is:
-\begin{equation}
-\frac{\partial \Uc^\prime}{\partial {\bf w}} = \left (
-   \begin{array}{ccc}
-       \rho & 0 & 0 \\
-       0 & \rho & 0  \\
-       \rho  e_{X_\alpha} & \rho e_{X_\beta} & \rho c_v \\
-     \end{array}\right)
-\end{equation}
-where we use the following notation for compactness:
-\begin{equation}
-e_{X_k} = \dedXd
-\end{equation}
-and the specific heat at constant volume is
-\begin{equation}
-c_v = \left . \frac{\partial e}{\partial T} \right |_{\rho, X_k}
-\end{equation}
-We get the inverse
-(computed via SymPy) as:
-\begin{equation}
-\renewcommand{\arraystretch}{1.5}
-\frac{\partial {\bf w}}{\partial \Uc^\prime} = \left (
-  \begin{array}{ccc}
-   \frac{1}{\rho} & 0 & 0 \\
-   0 & \frac{1}{\rho} & 0 \\
-    -\frac{e_{X_\alpha}}{\rho c_v} & -\frac{e_{X_\beta}}{\rho c_v} & \frac{1}{\rho c_v} \\
-   \end{array}\right)
-\renewcommand{\arraystretch}{1}
+    \nse{\rho^{n+1,(k)}, T^{n+1,(\xi)}, (Y_e)^{n+1,(\xi)}} \rightarrow \bar{A}^{n+1,(\xi)}, (X_k)^{n+1,(\xi)}, (B/A)^{n+1,(\xi)}, (dY_e/dt)^{n+1,(\xi)}
 \end{equation}
 
-The reaction vector is
+The inputs already have seen advection, so that means that the outputs know about advection.
+So when we use $(B/A)^{n+1,(\xi)}$ to compute the energy release, are we also including the
+change due to advection?
+\end{tcolorbox}
+
+\subsection{SDC-NSE Coupling Approach II}
+
+This is on {\tt older\_sdc\_fix} branch, PR 1092.
+
+It differs from above as in that it attempts to remove the advective
+part from the energy generation rate, addressing the issue above.
+
+Now we do compute a $(B/A)$ without advection:
 \begin{equation}
-\Rb(\Uc^\prime) = \left (  \begin{array}{c} \rho \omegadot_\alpha \\ \rho \omegadot_\beta \\ \rho \Sdot \end{array} \right )
+   \widetilde{\rho \left ( \frac{B}{A} \right ) } = \left [ \rho \left ( \frac{B}{A} \right ) \right ]^{n+1,(\xi)} - \Advss{\rho (B/A)}^{n+1/2,(k)}
 \end{equation}
-We take all the quantities to be functions of $\rho$, $e$, and $X_k$,
-but since $\rho$ doesn't change from the reactions (the reactive
-source of the continuity equation is zero), it is held constant in
-these derivatives.  The Jacobian is computed as $\partial \Rb/\partial
-{\bf w}$:
+and do the energy generation as:
 \begin{equation}
-\renewcommand{\arraystretch}{1.5}
-\frac{\partial \Rb}{\partial {\bf w}} = \left (
-  \begin{array}{ccc}
-     \rho \frac{\partial \omegadot_\alpha}{\partial X_\alpha} &
-     \rho \frac{\partial \omegadot_\alpha}{\partial X_\beta} & 
-     \rho \frac{\partial \omegadot_\alpha}{\partial T} \\
-      %
-     \rho \frac{\partial \omegadot_\beta}{\partial X_\alpha} &
-     \rho \frac{\partial \omegadot_\beta}{\partial X_\beta} &
-     \rho \frac{\partial \omegadot_\beta}{\partial T} \\
-     %
-     \rho \frac{\partial \Sdot}{\partial X_\alpha} &
-     \rho \frac{\partial \Sdot}{\partial X_\beta} &
-     \rho \frac{\partial \Sdot}{\partial T} \\
-  \end{array}
-  \right )
-\renewcommand{\arraystretch}{1}
+   (\rho \Sdot)^{n+1/2,(\xi)} = 
+     \left \{ \widetilde{\rho \left ( \frac{B}{A} \right )} -
+       \left [ \rho \left ( \frac{B}{A} \right ) \right ]^n \right \} N_A \frac{1}{\Delta t}
+\end{equation}
+This will not include the advection.
+
+\subsection{SDC-NSE Coupling Approach III}
+
+This is the branch {\tt fix\_sdc\_nse} in PR 1090.
+
+This follows the approach from the original SDC paper, for $\mathcal{I}_q$,  by first
+predicting the update of $(B/A)$ with just advection, and using this to compute
+the change in $(B/A)$ from just reactions.
+\begin{equation}
+\left [ \rho \left ( \frac{B}{A} \right ) \right ]^\star = 
+   \left [ \rho \left ( \frac{B}{A} \right ) \right ]^n + \Delta t \Advss{\rho (B/A)}^{n+1/2}
+\end{equation}
+then we compute
+\begin{equation}
+\left (\frac{B}{A} \right )^\star = \left [ \rho \left ( \frac{B}{A} \right ) \right ]^\star  / \rho^{n+1}
+\end{equation}
+Then the energy generation is:
+\begin{equation}
+   (\rho \Sdot)^{n+1/2,(\xi)} =  \rho^{n+1/2}
+     \left [ \left ( \frac{B}{A} \right )^{n+1} -
+              \left ( \frac{B}{A} \right )^\star \right ] N_A \frac{1}{\Delta t}
 \end{equation}
 
-The final Jacobian is found by multiplying these two:
+\subsection{Thoughts}
+
+\begin{itemize}
+
+\item Imagine that we evolve in a zone such that the NSE state does
+  not change, but density does (this would mean that the temperature
+  must have changed).  In that case the energy generation should be
+  zero, so writing is like:
   \begin{equation}
-\renewcommand{\arraystretch}{1.5}
-\frac{\partial \Rb}{\partial \Uc^\prime} = \left (
-  \begin{array}{ccc}
-    \frac{\partial \omegadot_\alpha}{\partial X_\alpha} - \frac{e_{X_\alpha}}{c_v} \frac{\partial \omegadot_\alpha}{\partial T} &
-    \frac{\partial \omegadot_\alpha}{\partial X_\beta} - \frac{e_{X_\beta}}{c_v} \frac{\partial \omegadot_\alpha}{\partial T} &
-    \frac{1}{c_v} \frac{\partial \omegadot_\alpha}{\partial T} \\
-     %
-    \frac{\partial \omegadot_\beta}{\partial X_\alpha} - \frac{e_{X_\alpha}}{c_v} \frac{\partial \omegadot_\beta}{\partial T} &
-    \frac{\partial \omegadot_\beta}{\partial X_\beta} - \frac{e_{X_\beta}}{c_v} \frac{\partial \omegadot_\beta}{\partial T} &
-    \frac{1}{c_v} \frac{\partial \omegadot_\beta}{\partial T} \\
-     %
-     \frac{\partial \Sdot}{\partial X_\alpha} -  \frac{e_{X_\alpha}}{c_v} \frac{\partial \Sdot}{\partial T} &
-     \frac{\partial \Sdot}{\partial X_\beta} -  \frac{e_{X_\beta}}{c_v} \frac{\partial \Sdot}{\partial T} &
-     \frac{1}{c_v} \frac{\partial \Sdot}{\partial T} \\
-  \end{array}
-  \right )
-\renewcommand{\arraystretch}{1}
+    \rho \Sdot = \rho^{n+1/2} \left [ \left ( \frac{B}{A} \right )^{n+1} - \left ( \frac{B}{A} \right )^n \right ] \frac{N_A}{\Delta t}
   \end{equation}
-We note that the form of these entries is the same as one would arrive at if you start with the rates expressed as
-$\omega_k(\rho, T(\rho, X_j, e), X_j)$ and note that constant $e$ implies that
-\begin{equation}
-\left . \frac{\partial T}{\partial X_k} \right |_{\rho, e} = - \frac{e_{X_k}}{c_v}
-\end{equation}
-
-While VODE can compute the entire Jacobian, ${\bf J}$ numerically via
-differencing, we found that does not give reliable results.  Instead,
-we compute compute the derivatives with respect to $X_k$ and $T$
-one-sided differencing, following the algorithm in \cite{lsode} to
-minimize numerical noise.  We then use the equation of state to
-compute $e_{X_k}$ and $c_v$ and construct the entries of the final
-Jacobian.
+  captures that.
+\end{itemize}
 

### 2022-12-14 retries/paper.tex

@@
 \end{itemize}
-All of these failure modes (and more) are represented in the simulation code Castro \citep{castro_joss}.
+All of these failure modes (and more) are represented in the simulation code \castro\ \citep{castro_joss}.
 In our experience, any useful science simulation will almost inevitably hit one of these failure modes.
@@
 application that successfully does this, which we drew inspiration from for this work, is the stellar evolution
-code MESA \citep{MESA}.
+code MESA \citep{MESA}. MESA also has the notion of a ``backup'': a retry can be defined as a second attempt at
+the current timestep, while a backup involves returning to an earlier timestep.
+
+\section{Implementation Details}\label{Sec:Implementation}
+
+A functional retry mechanism requires a decision scheme and associated code infrastructure for when to perform
+a second attempt at the timestep. A retry scheme effectively turns a timestep into a \texttt{while} loop: attempt
+an advance by $\Delta t$, and if it fails then attempt an advance by $\Delta t / 2$, and keep going until you have
+reached $t_{i} + \Delta t$ where $t_{i}$ is the starting simulation time. So in both principle and practice,
+implementing a retry mechanism is actually quite straightforward, with two complications as noted next.
+
+First, depending on context, this scheme may require saving application state and then restoring that application
+state at a later point. For example, \castro\ is a multi-level mesh refinement code where the fine levels are
+subcycled with respect to the coarse levels \citep{castro,berger_colella}, and so fine levels typically need to
+interpolate coarse data between time levels $t_{i}$ and $t_{i} + \Delta t$, where $\Delta t$ is the coarse grid
+timestep. \castro\ does this by maintaining both an ``old'' state and a ``new'' state corresponding to those two
+time levels. But subcycling within a given level due to retries will mean that at the end of the coarse step, the
+``old'' state data will be at some later time, such as $t_{i} + 3\Delta t / 4$ in the case of two failures. This
+needs to be reset back to the initial ``old'' state at $t_{i}$ for valid interpolation onto the fine grid. In \castro\
+we handle this by making a backup copy of the data\footnote{For the GPU build of \castro, we save this backup copy
+in host pinned memory since we are typically GPU memory constrained} at $t_{i}$ if (and only if) we encounter an
+advance failure on the first try.
+
+Second, this scheme may involve refactoring physics modules that perform irrevocable changes to application
+state. One way to guarantee that the scheme works is for the sequence of physics modules corresponding to an
+advance to be written as an idempotent operator: given an input ``old'' state at $t_{i}$, the advance updates
+only the ``new'' state at $t_{i} + \Delta t$, and only in a way that if you re-run the timestep you get the same
+result in the ``new'' state. This makes a retry scheme trivial: if you detect a failure, exit from the advance
+immediately, and then retry the advance with a smaller timestep.
 

### 2022-12-14 simplified_sdc_intro/paper.tex

@@
 simulation code in a modular way.  However, competition between the
-different physical processes can cause the coupling to \replaced{breakdown}{break down, since the reactions do not directly incorporate the effects of the hydrodynamics and vice versa}.
+different physical processes can cause the coupling to \edit1{break down, since the reactions do not directly incorporate the effects of the hydrodynamics and vice versa}.
 These splitting errors can lead to loss of accuracy and further time
@@
 These large changes in abundances (and large alternately positive and negative energy generation rates) can be a challenge for
-\replaced{a code}{the integration method---it may take more steps than allowed by the ODE integrator, require a timestep below floating point accuracy, or fail to meet the tolerances}.
+\edit1{the integration method---it may take more steps than allowed by the ODE integrator, require a timestep below floating point accuracy, or fail to meet the tolerances}.
 The easiest way to improve the coupling is to cut the
@@
 (no reactions) over $\Delta t$, and finally integrate the reaction
-terms only over $\Delta t/2$.  \added{The staggering of the reactive update
+terms only over $\Delta t/2$.  \edit1{The staggering of the reactive update
 means the hydro implicitly sees a time-centered reactive source, making
@@
 energy / temperature equation during the reaction step) in
-\citet{strang_rnaas} and showed that the above formulation \replaced{gets}{achieved} second
+\citet{strang_rnaas} and showed that the above formulation \edit1{achieved} second
 order convergence and that in particular, integrating an energy / temperature
@@
 
-We begin with the general form of a multi-implicit SDC correction equation with two physical processes\deleted{, as originally presented in} \citep{bourlioux2003high}.
+We begin with the general form of a multi-implicit SDC correction equation with two physical processes.
 Using $(k)$ to denote the iterate, we have
@@
     \item Predict $\qb$ to the interfaces at $t^{n+1/2}$ using the CTU
-      PPM method\added{---this involves taking the cell-centered
+      PPM method\edit1{---this involves taking the cell-centered
         primitive state and tracking all of the information that can
@@
       corrections are made.  Both are second-order accurate.  For the
-      gravitational sources, we do those in the normal predictor\added{, consistent
+      gravitational sources, we do those in the normal predictor\edit1{, consistent
       with the formulation in \citet{ppm}.}
@@
     $\Uc^\prime$.  This is different than the form of the Jacobian
-    usually used in reaction networks \added{(we depend on $e$ instead of
+    usually used in reaction networks \edit1{(we depend on $e$ instead of
     $T$)}.  We describe the form of the Jacobian in Appendix
@@
 We use a modified version of the VODE~\citep{vode} integrator,
-designed for stiff ODEs.  Our version has been ported to \replaced{C++, as a
-header-only library enabling it to run on GPUs}{modern C++ and designed to run on GPUs}.  We have also added
+designed for stiff ODEs.  Our version has been ported to \edit1{modern C++ and designed to run on GPUs}.  We have also added
 checks to the timestep rejection logic that help prevent VODE from
@@
 simplified-SDC convergence respectively.
-\added{We ran each integration method on a domain with $64^2$, $128^2$, $256^2$, $512^2$ zones.
-An error is defined between successive resolutions by coarsening the higher resolution run and taking the L1 norm of the zone-by-zone difference of the two simulations---for example, $\epsilon_{64\rightarrow 128}$ is the error between the $64^2$ and $128^2$ simulations.}  \deleted{The convergence rate is
-computed by taking log-base-2 of the ratio of adjacent errors.}  \added{The convergence
+\edit1{We ran each integration method on a domain with $64^2$, $128^2$, $256^2$, $512^2$ zones.
+An error is defined between successive resolutions by coarsening the higher resolution run and taking the L1 norm of the zone-by-zone difference of the two simulations---for example, $\epsilon_{64\rightarrow 128}$ is the error between the $64^2$ and $128^2$ simulations.}  \edit1{The convergence
 rate from $\epsilon_{64\rightarrow 128}$ to $\epsilon_{128\rightarrow 256}$ is computed as $\log_2 \epsilon_{64\rightarrow 128}/\epsilon_{128\rightarrow 256}$ and is given in the column between those errors, and likewise for the higher-resolution case.}  A rate
 of 2.0 indicates second order convergence.  We see that both methods
-\replaced{converge nearly second order, and agree well with each
-  other}{exhibit convergence at roughly second-order and agree well.}.
+\edit1{exhibit convergence at roughly second-order and agree well.}.
 This demonstrates that both methods are working as designed.
@@
 equilibrium at high temperatures when the reverse rates can be
-important.  \added{To highlight this, Figure~\ref{fig:strang_steps} shows the nuclear
+important.  \edit1{To highlight this, Figure~\ref{fig:strang_steps} shows the nuclear
 energy generation rate in a highly-zoomed region around the detonation over four

### 2022-12-14 xrb3/paper.tex

@@
 
+\author[0000-0001-8401-030X]{Michael Zingale}
+\affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
+             Stony Brook, NY 11794-3800}
+
 %% \author[0000-0002-1530-781X]{Alice Harpole}
@@
 
-%% \author[0000-0001-6191-4285]{Kiran Eiden}
-%% \affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
-%%              Stony Brook, NY 11794-3800}
-
-%% \author[0000-0001-8401-030X]{Michael Zingale}
-%% \affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
-%%              Stony Brook, NY 11794-3800}
+\author[0000-0001-6191-4285]{Kiran Eiden}
+\affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
+             Stony Brook, NY 11794-3800}
 
@@
 
-%% \author{Yuri Cavecchi}
-%% \affiliation{Mathematical Sciences and STAG Research Centre, University of Southampton, SO17 1BJ}
+\author[0000-0003-0439-4556]{Max P.\ Katz}
+\affiliation{NVIDIA Corp}
 
-%% \author[0000-0003-0439-4556]{Max P.\ Katz}
-%% \affiliation{NVIDIA Corp}
-
-%% \author[0000-0001-8092-1974]{Weiqun Zhang}
-%% \affiliation{Lawrence Berkeley National Laboratory, Berkeley, CA}
-
-%% \correspondingauthor{Alice Harpole}
-%% \email{alice.harpolee@stonybrook.edu}
+\correspondingauthor{Michael Zingale}
+\email{michael.zingale@stonybrook.edu}
 
@@
 \begin{abstract}
-We continue to investigate laterally propagating flames in XRBs,
-comparing the early evolution and structure of a burning hotspot
-between two- and three-dimensions.
-\end{abstract}
+We explore the early evolution of a hot spot and a flame ignition and
+spreading on the surface of a neutron star in three-dimensions.  We
+focus on the difference in the structure of the flame compared to
+previous two-dimensional simulations.\end{abstract}
 
@@
 
-X-ray bursts result from thermonuclear burning of an accreted H/He or
-He layer on a neutron star \citep{galloway:2017}.  Observations of
-brightness oscillations in the rise of the lightcurve suggest that the
-burning begins localized and then spreads across the neutron star.
-One-dimensional simulations of X-ray bursts reproduce the lightcurve
-and recurrence time \citep{woosley-xrb} and can use large reaction
-networks to explore rp-process nucleosynthesis.
+X-ray bursts (XRBs) result from thermonuclear burning of an accreted
+H/He or He layer on a neutron star \citep{galloway:2017}.
+Observations of brightness oscillations in the rise of the lightcurve
+suggest that the burning begins localized and then spreads across the
+neutron star.  This spreading is inherently a multi-dimensional
+phenomena, and hydrodynamic simulations that resolve the reactive zone
+and capture the scales over which rotation is important are needed to
+understand the flame propagation.  We show a first attempt at modeling
+the full three-dimensional ignition and spreading of a hot spot in an
+XRB.  This builds on our earlier two dimensional work
+\citep{eiden:2020,harpole:2021} that developed our simulation
+framework and explored how the acceleration of the burning front
+depends on the initial model structure.  As with those simulations, we
+use the freely-available \castro\ simulation code
+\citep{castro,castro_joss} simulation code.
 
@@
 
-We have been performing simulations of flame spreading on a neutron
-star resolving the flame structure using realistic reaction networks.
-Our two-dimensional studies \citep{eiden:2020,harpole:2021} demonstrated our open-source framework for modeling X-ray bursts using
-
-We use the \castro\ \citep{castro,castro_joss} simulation code, following
-the same numerical approach described in \citet{eiden:2020,harpole:2021}.
+The goal of this paper is to understand the challenges of a fully-resolved
+three-dimensional simulation and explore what is possible.  This will inform
+our follow-on simulations.
 

### 2023-02-19 xrb3/data/mass_plot.py
update the mass plot for the new simulation

@@ -1,0 +1,59 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+
+# Use LaTeX for rendering
+mpl.rcParams["text.usetex"] = True
+# load the xfrac package
+mpl.rcParams["text.latex.preamble"] += r'\newcommand{\isotm}[2]{{}^{#2}\mathrm{#1}}'
+mpl.rcParams['mathtext.fontset'] = 'cm'
+mpl.rcParams['mathtext.rm'] = 'serif'
+
+###
+### NOTE: the data in these files is in solar masses!
+###
+
+fig, ax = plt.subplots()
+
+data3d = np.loadtxt("flame_wave_3d_nonsquare_cool/species_diag.out")
+total_mass = np.sum(data3d[:,2:8])
+
+print(f"total mass 3d = {total_mass}")
+print(f"total He mass 3d = {data3d[0, 2]}, {data3d[-1, 2]}")
+
+mass_He0 = data3d[0, 2]
+
+ax.plot(data3d[:,1], data3d[:,2] / mass_He0, label=r"$\isotm{He}{4}$", color="C0")
+ax.plot(data3d[:,1], data3d[:,3] / mass_He0, label=r"$\isotm{C}{12}$", color="C1")
+ax.plot(data3d[:,1], data3d[:,4] / mass_He0, label=r"$\isotm{O}{16}$", color="C2")
+ax.plot(data3d[:,1], data3d[:,5] / mass_He0, label=r"$\isotm{Ne}{20}$", color="C3")
+ax.plot(data3d[:,1], data3d[:,6] / mass_He0, label=r"$\isotm{Mg}{24}$", color="C4")
+ax.plot(data3d[:,1], data3d[:,7] / mass_He0, label=r"$\isotm{Si}{28}$", color="C5")
+ax.plot(data3d[:,1], data3d[:,8] / mass_He0, label=r"$\isotm{Ni}{56}$", color="C6")
+
+
+data2d = np.loadtxt("flame_wave_2d_nonsquare_cool_new/species_diag.out")
+mass_He0 = data2d[0, 2]
+
+
+ax.plot(data2d[:,1], data2d[:,2] / mass_He0, linestyle="--", color="C0")
+ax.plot(data2d[:,1], data2d[:,3] / mass_He0, linestyle="--", color="C1")
[…]

### 2023-02-20 xrb3/paper.tex
new figures in place
add some more figs
some fig adjustments
some more details on resolution

@@
-\documentclass[preprint,times,tighten]{aastex631}
+\documentclass[twocolumn,times,tighten]{aastex631}
 
@@
 In \citet{harpole:2021}, we explored several different rotation rates
-and initial model thermal structures.  We found that a model with a
-neutron star crust temperature of $3\times 10^8~\mathrm{K}$ resulted
-in a flame that accelerated quickly as it spread across the neutron
-star surface.  We adopt this model for the 3D simulation presented
-here, since it requires the least amount of simulation time to yield a
-developed flame.
+and initial model thermal structures.  We explored several different
+temperature structures in our initial model.  We choose a model with a
+neutron star crust temperature of $2\times 10^8~\mathrm{K}$.  We adopt
+this model for the 3D simulation presented here, since it gives a clean
+well-defined flame.
 
 Moving to 3D is very expensive, so to further reduce the computational
-expense, we use a slightly lower resolution of 25 cm instead of 20 cm
-from \citet{harpole:2021}.  We pick a rotation rate of 1000 Hz.
+expense, we use an aniostropic resolution, with 32 cm laterally and 16 cm resolution
+vertically.  This is finer vertically than done in 
+\citet{harpole:2021}.  We pick a rotation rate of 1000 Hz.
 Because the hotspot is placed in the center of the domain, the size of
@@
 
-Our 3D simulation used a base grid of $1024^2 \times 128$ zones with 2
+Our 3D simulation used a base grid of $768^2 \times 192$ zones with 2
 levels of refinement, the first jumping by a factor of 4 and the next
-by a factor of 2.  The domain has a size $(2.048\times
-10^5~\mathrm{cm})^2 \times 2.56\times 10^4~\mathrm{cm}$, giving a fine
-grid resolution of $25~\mathrm{cm}$.  The initial model and other
+by a factor of 2.  The domain has a size $(1.96608\times
+10^5~\mathrm{cm})^2 \times 2.4576\times 10^4~\mathrm{cm}$, giving a
+fine grid resolution of $32~\mathrm{cm}$ laterally and
+$16~\mathrm{cm}$ vertically.  The vertical resolution is finer to help
+maintain hydrostatic equilibrium.  The initial model and other
 simulation parameters are identical to those used in
@@
 GPUs per node.  The entire computation was offloaded to GPUs following
-the strategy described in \citet{castro_gpu}.  Overall, about 370,000
+the strategy described in \citet{castro_gpu}.  Overall, about 250,000
 node-hours were used for the calculation.  The full state of the
-calculation was output only every 0.005 s, as each output file is 2.7 TB
-in size (for single precision data).  We output a few fields more
+calculation was output only every 0.005 s, as each output file is 2.4 
+TB in size (for single precision data).  We output a few fields more
 frequently for visualization.  We ran a 2D simulation with the same
@@
 
-Slice plot compared to 2-d simulation
+\begin{figure*}[t]
+\centering
+\plotone{time_series}
+\caption{\label{fig:2d_abar} stuff}
+\end{figure*}
 
-Phase plot comparisons
+\begin{figure}[t]
+\centering
+%\epsscale{0.75}
+\plotone{abar_top_stack}
+%\epsscale{1.0}
+\caption{\label{fig:vr_abar} stuff}
+\end{figure}
+
+%% \begin{figure}[t]
+%% \centering
+%% \plotone{flame_wave_1000Hz_25cm_smallplt88081_Hnuc_annotated_top} \\
+%% \plotone{flame_wave_1000Hz_25cm_smallplt174562_Hnuc_annotated_top.png}
+%% \caption{\label{fig:vr_hnuc} stuff}
+%% \end{figure}
 

### 2023-02-26 xrb3/paper.tex
more work on the paper

@@
 %======================================================================
-\title{Dynamics of Laterally Propagating Flames in X-ray Bursts. III. Comparing Hotspot Evolution in Two and Three Dimensions}
+\title{Dynamics of Laterally Propagating Flames in X-ray Bursts. III. Comparing Early Hotspot Evolution in Two and Three Dimensions}
 
@@
 spreading on the surface of a neutron star in three-dimensions.  We
-focus on the difference in the structure of the flame compared to
-previous two-dimensional simulations.\end{abstract}
+look at the nucleosynthesis and morphology of the burning front and
+compare to two-dimensional axisymmetric simulations.  Finally, we discuss
+the progress toward full-star resolved flame simulations.\end{abstract}
 
@@
 neutron star.  This spreading is inherently a multi-dimensional
-phenomena, and hydrodynamic simulations that resolve the reactive zone
-and capture the scales over which rotation is important are needed to
-understand the flame propagation.  We show a first attempt at modeling
-the full three-dimensional ignition and spreading of a hot spot in an
-XRB.  This builds on our earlier two dimensional work
+phenomena, and hydrodynamic simulations that resolve the reactive
+zone, use realistic reaction networks, and capture the scales over
+which rotation is important are needed to understand the flame
+propagation and nucleosynthesis.  We show a first attempt at modeling
+the early evolution of a spreading hotspot in three dimensions.  This
+builds on our earlier two dimensional work
 \citep{eiden:2020,harpole:2021} that developed our simulation
@@
 use the freely-available \castro\ simulation code
-\citep{castro,castro_joss} simulation code.
+\citep{castro,castro_joss} simulation code, with all of the code
+needed to run these simulations in our public github repos.
 
-To understand how the burning front propagates across the surface, we
-need to do multidimensional simulations.  The work of
-\citet{cavecchi:2013,art-2015-cavecchi-etal,art-2016-cavecchi-etal}
-
-The three-dimensional study by \citet{Cavecchi2019} showed...
+Our simulations complement the multidimensional models of flames of
+\citet{cavecchi:2013,art-2015-cavecchi-etal,art-2016-cavecchi-etal,Cavecchi2019}
+by focusing on resolving the reaction zone of the flame and modeling
+the vertical structure hydrodynamically instead of hydrostatically.
+We are also modeling flames starting from a small hot spot.  This
+however means that we are confined to smaller domains, with similar
+computational resources, so one of the goals of this study is to
+understand how far we can push resolved XRB flame simulations.  The
+work by \citet{goodwin:2021} also explores multidimensional evolution,
+looking at the thermal transport and how affects the location where
+the hotspot ignites.  Together, all three approaches help build a
+comprehensive multi-dimensioanl picture of X-ray bursts.
 
@@
 this model for the 3D simulation presented here, since it gives a clean
-well-defined flame.
+well-defined flame.  The model is constructed following the procedure
+described in \citet{eiden:2020,harpole:2021}.
 
 Moving to 3D is very expensive, so to further reduce the computational
-expense, we use an aniostropic resolution, with 32 cm laterally and 16 cm resolution
-vertically.  This is finer vertically than done in 
-\citet{harpole:2021}.  We pick a rotation rate of 1000 Hz.
-Because the hotspot is placed in the center of the domain, the size of
-the domain also becomes a constraint, and as a result, the evolution
-time we can reach will become limited by the time it takes the burning
-to approach the edge of the domain.  As a result, we focus here on the
+expense, we use an aniostropic resolution, with 32 cm laterally and 16
+cm resolution vertically.  This is finer vertically than done in
+\citet{harpole:2021}.  We pick a rotation rate of 1000 Hz.  Because
+the hotspot is placed in the center of the domain, the size of the
+domain also becomes a constraint, and as a result, the evolution time
+we can reach will become limited by the time it takes the burning to
+approach the edge of the domain.  As a result, we focus here on the
 early evolution.
@@
 $16~\mathrm{cm}$ vertically.  The vertical resolution is finer to help
-maintain hydrostatic equilibrium.  The initial model and other
-simulation parameters are identical to those used in
-\citet{harpole:2021}.  We used static mesh refinement, to achieve
-better load balancing, fully refining the atmosphere below a height of
+maintain hydrostatic equilibrium.  The simulation methodology and
+parameters are mostly identical to those used in \citet{harpole:2021}.
+We use an unsplit piecewise parabolic method \cite{ppmunsplit} for the
+advection and operator (Strang) splitting for the reactions, evolving
+an internal energy evolution equation during the burn, as described in
+\citet{strang_rnaas}.  The main difference with the previous
+simulations is that we switched from a hydrostatic to reflecting
+boundary at the bottom of the domain and used a simple well-balanced
+scheme (similar to \citealt{kappeli:2016} but adapted to deal with
+characteristic tracing in PPM as described in \citealt{ppm-hse}).
+Tests demonstrated that this boundary does a better job than our
+previous approach at supporting the atmosphere as the flame propagates
+for long times.
+
+We used static mesh refinement, to achieve better
+load balancing, fully refining the atmosphere below a height of
 $3600~\mathrm{cm}$.  These simulations used the 7-isotope He-burning
@@
 
-Things to show:
-
-\begin{figure*}[t]
-\centering
-\plotone{time_series}
-\caption{\label{fig:2d_abar} stuff}
-\end{figure*}
-
 \begin{figure}[t]
@@
 %\epsscale{1.0}
-\caption{\label{fig:vr_abar} stuff}
+\caption{\label{fig:vr_abar} Volume rendering of the three-dimensional XRB
+simulation showing the mean molecular weight, $\bar{A}$.  The flame is viewed
+from above.}
 \end{figure}
+
+Figure~\ref{fig:vr_abar} shows the a three-dimensional simulation via
+top-down volume rendering of the mean-molecular weight,
+\begin{equation}
+\bar{A} = \sum_k \frac{X_k}{A_k}
+\end{equation}
+where $X_k$ is the mass fraction of species $k$ and $A_k$ is the
+atomic weight (in mass units).  The structure is shown at 10, 20, and
+40~ms of simulation time.  While initially quite symmetric, the axial
+symmetry is broken and the flame structure takes on a more tenuous
+form.  The panels show the flame spreading considerably through the
+domain as time evolves, and the simulation is halted once the flame
+nears the boundaries.
+
+Figure~\ref{fig:2d_abar} shows the corresponding two-dimensional
+axisymmetric simulation.  The initial model and simulation parameters
+are identical to the three-dimensional case.  This simulation uses
+zones that are 32 cm wide and 16 cm tall, analogous to the lateral and
+vertical resolution of the three-dimensional simulation.  We see that
+the overall structure and evolution looks identical to that reported
+in \cite{harpole:2021}.
+
+\begin{figure*}[t]
+\centering
+\plotone{time_series}
+\caption{\label{fig:2d_abar} Time-sequence of the two-dimensional axisymmetric
+simualtion using the same setup as the three-dimensional simulation shown
+in Figure~\ref{fig:vr_abar}.  The domain is zoomed in, showing only the left
+two-thirds of the surface.}
+\end{figure*}
 
@@
 %% \end{figure}
-
-Volume renderings
 
@@
 masses by the initial \isot{He}{4} mass.  This plot shows that the
-masses of the heavy species , especially \isot{C}{12} and
-\isot{Si}{28} grow quickly with time, but the 2D simulation appears to
-burn faster than the 3D simulation.  This is because the assumption of
+masses of the heavy species, especially \isot{C}{12}, grow quickly
+with time.  The differences between the 2D and 3D simulations appear
+small---the burning is slightly faster in 2D, but the trends track
+very well.  This slight difference is likely because the assumption of
 axisymmetry there does not allow for the complex structure we see in
-the evolution of composition in the 3D flame.  This suggests that the
-3D flame speed is slower than the 2D flame speed.
+the evolution of composition in the 3D flame.  This strong agreement
+suggests that using 2D axisymmetric calculations is a good model for
+the early flame evolution.  And since they are so much less
+computationally expensive, we this will allow us to explore much more
+complex reaction networks and understand the nucleosynthesis in more
+detail.
 
@@
 \plotone{mass_plot}
-\caption{\label{fig:mass_plot} Mass of species scaled to initial He mass as a function of time.  The 3D simulation is shown as the solid lines and the 2D simulation is shown as the dashed lines.}
+\caption{\label{fig:mass_plot} Mass of species scaled to initial He
+  mass as a function of time.  The 3D simulation is shown as the solid
+  lines and the 2D simulation is shown as the dashed lines.}
 \end{figure}
@@
 simulation and a fully three-dimensional hydrodynamical simulation.
+Overall, the early evolution of the flame spreading from a hotspot
+behaves similarly in two- and three-dimensions.  As this
+three-dimensional simulation pushed the limits of available
+computational resources, this result suggests that we can continue to
+use two-dimensional axisymmetric simulations to explore the initial
+flame structure and spend our computational resources on increasing
+the size of our reaction networks to explore the sensitivity to
+network size and understand the overall nucleosynthesis.
 
-The main limitation of this study is that computational resources
-limited the size of the domain and resolution we could use.  To
-improve simulations for the future, there are several avenues to
-pursue.  First, the main expense of the current simulations is the
-reaction network.  We need to work more on optimization of the
-network.  Second, while additional resolution will be expensive, we
-can go higher order, as demonstrated with the 4th order accurate
-coupling of hydro, diffusion, and reactions in \cite{castro-sdc}.  The
+There is still a role for three-dimensional simulations---the
+convective burning in the accreted layer should create turbulence that
+the flame will encounter as it propagates.  Although we have shown
+that for the flame in a Type Ia supernovae, the turbulence from the
+era of convection is not strong enough to disrupt the
+flame~\cite{wdturb}, the flame in an XRB is considerably slower and
+thicker, so it remains to be seen what effect it might have.  This can
+be assessed in a different geometry, like a long channel, which would
+allow us to trade domain size for resolution to increase the numerical
+Reynolds number.  We also will consider higher-order simulation
+methodologies which could allow us to capture these effects at lower
+resolution.  We've already demonstrated a fully fourth-order accruate
+(in space and time) algorithm for coupling hydro, diffusion, and
+reactions in Castro \cite{castro-sdc} that could be used here.  The
 main outstanding issue with applying that work to the present problem
 is the multilevel time integration.
-
-Future work will explore mixed H/He bursts, using this same framework.
 

### 2023-02-27 xrb3/paper.tex
more work

@@
 suggest that the burning begins localized and then spreads across the
-neutron star.  This spreading is inherently a multi-dimensional
+neutron star \citep{bhattacharyya:2007}.  This spreading is inherently a multi-dimensional
 phenomena, and hydrodynamic simulations that resolve the reactive
@@
 which rotation is important are needed to understand the flame
-propagation and nucleosynthesis.  We show a first attempt at modeling
+propagation and nucleosynthesis (especially rp-process in mixed H/He bursts \citealt{rpprocess}).  We show a first attempt at modeling
 the early evolution of a spreading hotspot in three dimensions.  This
@@
 \begin{equation}
-\bar{A} = \sum_k \frac{X_k}{A_k}
+\frac{1}{\bar{A}} = \sum_k \frac{X_k}{A_k}
 \end{equation}
@@
 era of convection is not strong enough to disrupt the
-flame~\cite{wdturb}, the flame in an XRB is considerably slower and
+flame~\citep{wdturb}, the flame in an XRB is considerably slower and
 thicker, so it remains to be seen what effect it might have.  This can
@@
 (in space and time) algorithm for coupling hydro, diffusion, and
-reactions in Castro \cite{castro-sdc} that could be used here.  The
+reactions in Castro \citep{castro-sdc} that could be used here.  The
 main outstanding issue with applying that work to the present problem

### 2023-03-02 xrb3/paper.tex
update paper title
add metadata link

@@
 %======================================================================
-\title{Dynamics of Laterally Propagating Flames in X-ray Bursts. III. Comparing Early Hotspot Evolution in Two and Three Dimensions}
+\title{Comparing Early Evolution of Flames in X-ray Bursts in Two and Three Dimensions}
 
-\shorttitle{Lateral Flame Dynamics III}
+\shorttitle{Early XRB Flame Evolution}
 %% \shortauthors{Harpole et al.}
@@
 \url{http://github.com/AMReX-Astro/Castro}.  The problem setup used
-here is available in the git repo as {\tt flame\_wave}.  The work at
-Stony Brook was supported by DOE/Office of Nuclear Physics grant
-DE-FG02-87ER40317.  This research used resources of the National
-Energy Research Scientific Computing Center, a DOE Office of Science
-User Facility supported by the Office of Science of the
+here is available in the git repo as {\tt flame\_wave}. The metadata
+describing the build environment and the global diagnostic output
+files are available on Zenodo at \citet{xrb_data}.  We thank Alice
+Harpole and Max Katz for their contributions to the AMReX Astrophysics
+suite.
+
+The work at Stony Brook was supported by DOE/Office of Nuclear
+Physics grant DE-FG02-87ER40317.  This research used resources of the
+National Energy Research Scientific Computing Center, a DOE Office of
+Science User Facility supported by the Office of Science of the
 U.~S.\ Department of Energy under Contract No.\ DE-AC02-05CH11231.
+This research was supported by the Exascale Computing Project
+(17-SC-20-SC), a collaborative effort of the U.S. Department of Energy
+Office of Science and the National Nuclear Security Administration.
 This research used resources of the Oak Ridge Leadership Computing

### 2023-03-03 xrb3/paper.tex
some updates

@@
 \begin{abstract}
-We explore the early evolution of a hot spot and a flame ignition and
-spreading on the surface of a neutron star in three-dimensions.  We
-look at the nucleosynthesis and morphology of the burning front and
-compare to two-dimensional axisymmetric simulations.  Finally, we discuss
-the progress toward full-star resolved flame simulations.\end{abstract}
+We explore the early evolution of flame ignition and spreading on the
+surface of a neutron star in three-dimensions, in the context of X-ray
+bursts.  We look at the nucleosynthesis and morphology of the burning
+front and compare to two-dimensional axisymmetric simulations to gauge
+how important a full three-dimensional treatment of the flame is for
+the early dynamics.  Finally, we discuss the progress toward full-star
+resolved flame simulations.\end{abstract}
 
@@
 use the freely-available \castro\ simulation code
-\citep{castro,castro_joss} simulation code, with all of the code
+\citep{castro,castro_joss}, with all of the code
 needed to run these simulations in our public github repos.
@@
 
-The goal of this paper is to understand the challenges of a fully-resolved
-three-dimensional simulation and explore what is possible.  This will inform
-our follow-on simulations.
+The goal of this paper is to understand the challenges of resolved
+three-dimensional hydrodynamical simulation and explore what is
+possible with today's resources.  This will inform our follow-on
+simulations.
 
@@
 
-In \citet{harpole:2021}, we explored several different rotation rates
-and initial model thermal structures.  We explored several different
-temperature structures in our initial model.  We choose a model with a
-neutron star crust temperature of $2\times 10^8~\mathrm{K}$.  We adopt
-this model for the 3D simulation presented here, since it gives a clean
-well-defined flame.  The model is constructed following the procedure
-described in \citet{eiden:2020,harpole:2021}.
+In \citet{eiden:2020}, we developed the simulation methodology for
+modeling laterally spreading flames through the accreted layer on the
+surface of a neutron star.  Building on this, in \citet{harpole:2021},
+we explored several different rotation rates and initial model thermal
+structures to understand how the flame speed depends on the thermal
+structure of the envelope.  Both of these papers used a
+two-dimensional axisymmetric geometry.  The three-dimensional
+simulations we present here build on these works.  We use a neutron
+star crust temperature of $2\times 10^8~\mathrm{K}$ and rotation rate
+of 1000~Hz, as followed in \citet{harpole:2021}.  This model was
+chosen here since it gives a clean well-defined flame.  The model is
+constructed following the same procedure described in
+\citet{eiden:2020,harpole:2021}.
 
 Moving to 3D is very expensive, so to further reduce the computational
-expense, we use an aniostropic resolution, with 32 cm laterally and 16
+expense, we use an aniostropic resolution, with 32 cm resolution laterally and 16
 cm resolution vertically.  This is finer vertically than done in
-\citet{harpole:2021}.  We pick a rotation rate of 1000 Hz.  Because
+\citet{harpole:2021}.  Because
 the hotspot is placed in the center of the domain, the size of the
@@
 by a factor of 2.  The domain has a size $(1.96608\times
-10^5~\mathrm{cm})^2 \times 2.4576\times 10^4~\mathrm{cm}$, giving a
-fine grid resolution of $32~\mathrm{cm}$ laterally and
-$16~\mathrm{cm}$ vertically.  The vertical resolution is finer to help
-maintain hydrostatic equilibrium.  The simulation methodology and
+10^5~\mathrm{cm})^2 \times 2.4576\times 10^4~\mathrm{cm}$.  The simulation methodology and
 parameters are mostly identical to those used in \citet{harpole:2021}.
-We use an unsplit piecewise parabolic method \cite{ppmunsplit} for the
+We use an unsplit piecewise parabolic method \citep{ppmunsplit,millercolella:2002} for the
 advection and operator (Strang) splitting for the reactions, evolving
 an internal energy evolution equation during the burn, as described in
-\citet{strang_rnaas}.  The main difference with the previous
+\citet{strang_rnaas}.  These simulations used the 7-isotope He-burning
+network used in \citet{eiden:2020}.  The main difference with the previous
 simulations is that we switched from a hydrostatic to reflecting
@@
 
-We used static mesh refinement, to achieve better
-load balancing, fully refining the atmosphere below a height of
-$3600~\mathrm{cm}$.  These simulations used the 7-isotope He-burning
-network used in \citet{eiden:2020}.  The simulations were run on the
-OLCF Summit supercomputer, on 342 to 1366 nodes, with 6 NVIDIA V100
-GPUs per node.  The entire computation was offloaded to GPUs following
-the strategy described in \citet{castro_gpu}.  Overall, about 250,000
-node-hours were used for the calculation.  The full state of the
-calculation was output only every 0.005 s, as each output file is 2.4
-TB in size (for single precision data).  We output a few fields more
-frequently for visualization.  We ran a 2D simulation with the same
-resolution and refinement strategy to compare with here.
+We used static mesh refinement, to achieve better load balancing,
+fully refining the atmosphere below a height of $3600~\mathrm{cm}$.
+The simulations were run on the OLCF Summit supercomputer, on 342 to
+1366 nodes, with 6 NVIDIA V100 GPUs per node.  The entire computation
+was offloaded to GPUs following the strategy described in
+\citet{castro_gpu}.  GPU-offloading was critical to being able to
+perform these simulations, since we run more than an order of
+magnitude faster using all the GPUs on a node as compared to all of
+the CPUs on the node.  Overall, about 250,000 node-hours were used for
+the calculation.  The full state of the calculation was output only
+every 0.005 s, as each output file is 2.4 TB in size (for single
+precision data).  We output a few fields more frequently for
+visualization, and global diagnostics were output every timestep.  We
+ran a 2D simulation with the same resolution and refinement strategy
+to compare with here.
 
@@
 Figure~\ref{fig:2d_abar} shows the corresponding two-dimensional
-axisymmetric simulation.  The initial model and simulation parameters
-are identical to the three-dimensional case.  This simulation uses
-zones that are 32 cm wide and 16 cm tall, analogous to the lateral and
-vertical resolution of the three-dimensional simulation.  We see that
+axisymmetric simulation.  In the axisymemtric geometry, each zone is essentially
+a torus, rotated about the vertical axis.  The initial model and simulation parameters
+are identical to the three-dimensional case.  We see that
 the overall structure and evolution looks identical to that reported
-in \cite{harpole:2021}.
+in \cite{harpole:2021}.  We see that there is a well-defined flame in the simulation plane
+that spreads outward with time.
 
@@
 inscribed in the 3D domain.  To compensate for this, we scale the
-masses by the initial \isot{He}{4} mass.  This plot shows that the
+masses by the total mass on the grid.  This plot shows that the
 masses of the heavy species, especially \isot{C}{12}, grow quickly
@@
 the size of our reaction networks to explore the sensitivity to
-network size and understand the overall nucleosynthesis.
+network size and understand the overall nucleosynthesis.  We can also
+do initial explorations of full star bursts in two-dimensions, perhaps
+using axisymmetric spherical coordinates (modeling an annular region
+in $r$-$\theta$).  In this geometry, point ignition requires starting
+at the poles, but this would allow us to see the effect of varying
+gravity with latitude (due to centrifugal forces).
 

### 2023-03-03 xrb3/data/mass_plot.py

@@ -16,39 +16,41 @@ mpl.rcParams['mathtext.rm'] = 'serif'
 fig, ax = plt.subplots()
 
 data3d = np.loadtxt("flame_wave_3d_nonsquare_cool/species_diag.out")
-total_mass = np.sum(data3d[:,2:8])
+total_mass = np.sum(data3d[:,2:9], axis=1)
 
 print(f"total mass 3d = {total_mass}")
 print(f"total He mass 3d = {data3d[0, 2]}, {data3d[-1, 2]}")
 
 mass_He0 = data3d[0, 2]
 
-ax.plot(data3d[:,1], data3d[:,2] / mass_He0, label=r"$\isotm{He}{4}$", color="C0")
-ax.plot(data3d[:,1], data3d[:,3] / mass_He0, label=r"$\isotm{C}{12}$", color="C1")
-ax.plot(data3d[:,1], data3d[:,4] / mass_He0, label=r"$\isotm{O}{16}$", color="C2")
-ax.plot(data3d[:,1], data3d[:,5] / mass_He0, label=r"$\isotm{Ne}{20}$", color="C3")
-ax.plot(data3d[:,1], data3d[:,6] / mass_He0, label=r"$\isotm{Mg}{24}$", color="C4")
-ax.plot(data3d[:,1], data3d[:,7] / mass_He0, label=r"$\isotm{Si}{28}$", color="C5")
-ax.plot(data3d[:,1], data3d[:,8] / mass_He0, label=r"$\isotm{Ni}{56}$", color="C6")
+ax.plot(data3d[:,1], data3d[:,2] / total_mass[0], label=r"$\isotm{He}{4}$", color="C0")
+ax.plot(data3d[:,1], data3d[:,3] / total_mass[0], label=r"$\isotm{C}{12}$", color="C1")
+ax.plot(data3d[:,1], data3d[:,4] / total_mass[0], label=r"$\isotm{O}{16}$", color="C2")
+ax.plot(data3d[:,1], data3d[:,5] / total_mass[0], label=r"$\isotm{Ne}{20}$", color="C3")
+ax.plot(data3d[:,1], data3d[:,6] / total_mass[0], label=r"$\isotm{Mg}{24}$", color="C4")
+ax.plot(data3d[:,1], data3d[:,7] / total_mass[0], label=r"$\isotm{Si}{28}$", color="C5")
+ax.plot(data3d[:,1], data3d[:,8] / total_mass[0], label=r"$\isotm{Ni}{56}$", color="C6")
 
 
[…]

### 2023-03-04 xrb3/paper.tex
add new plot
update conclusions

@@
 
-%% \author[0000-0002-1530-781X]{Alice Harpole}
-%% \affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
-%%              Stony Brook, NY 11794-3800}
-
 \author[0000-0001-6191-4285]{Kiran Eiden}
-\affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
-             Stony Brook, NY 11794-3800}
-
-%% \author[0000-0003-2300-5165]{Donald Willcox}
-%% \affiliation{Lawrence Berkeley National Laboratory, Berkeley, CA}
-
-\author[0000-0003-0439-4556]{Max P.\ Katz}
-\affiliation{NVIDIA Corp}
+\affiliation{Department of Astronomy,
+University of California, Berkeley,
+Berkeley, CA 94720-3411, USA}
 
@@
 suggest that the burning begins localized and then spreads across the
-neutron star \citep{bhattacharyya:2007}.  This spreading is inherently a multi-dimensional
-phenomena, and hydrodynamic simulations that resolve the reactive
-zone, use realistic reaction networks, and capture the scales over
-which rotation is important are needed to understand the flame
-propagation and nucleosynthesis (especially rp-process in mixed H/He bursts \citealt{rpprocess}).  We show a first attempt at modeling
-the early evolution of a spreading hotspot in three dimensions.  This
-builds on our earlier two dimensional work
-\citep{eiden:2020,harpole:2021} that developed our simulation
-framework and explored how the acceleration of the burning front
-depends on the initial model structure.  As with those simulations, we
-use the freely-available \castro\ simulation code
-\citep{castro,castro_joss}, with all of the code
-needed to run these simulations in our public github repos.
+neutron star \citep{bhattacharyya:2007}.  This spreading is inherently
+a multi-dimensional phenomena, and hydrodynamic simulations that
+resolve the reactive zone, use realistic reaction networks, and
+capture the scales over which rotation is important are needed to
+understand the flame propagation and nucleosynthesis (especially
+rp-process in mixed H/He bursts, as discussed in \citealt{rpprocess}).
+We show a first attempt at modeling the early evolution of a spreading
+hotspot in three dimensions.  This builds on our earlier two
+dimensional work \citep{eiden:2020,harpole:2021} that developed our
+simulation framework and explored how the acceleration of the burning
+front depends on the initial model structure.  As with those
+simulations, we use the freely-available \castro\ simulation code
+\citep{castro,castro_joss}, with all of the code needed to run these
+simulations in our public github repos.
 
-Our simulations complement the multidimensional models of flames of
+Our simulations complement the two- and three-dimensional models of flames of
 \citet{cavecchi:2013,art-2015-cavecchi-etal,art-2016-cavecchi-etal,Cavecchi2019}
@@
 the vertical structure hydrodynamically instead of hydrostatically.
-We are also modeling flames starting from a small hot spot.  This
+We also focus on modeling flames starting from a small hot spot.  This
 however means that we are confined to smaller domains, with similar
@@
 
-In \citet{eiden:2020}, we developed the simulation methodology for
-modeling laterally spreading flames through the accreted layer on the
-surface of a neutron star.  Building on this, in \citet{harpole:2021},
-we explored several different rotation rates and initial model thermal
-structures to understand how the flame speed depends on the thermal
-structure of the envelope.  Both of these papers used a
-two-dimensional axisymmetric geometry.  The three-dimensional
-simulations we present here build on these works.  We use a neutron
-star crust temperature of $2\times 10^8~\mathrm{K}$ and rotation rate
-of 1000~Hz, as followed in \citet{harpole:2021}.  This model was
-chosen here since it gives a clean well-defined flame.  The model is
-constructed following the same procedure described in
+In \citet{eiden:2020}, we developed the simulation methodology in
+\castro\ for modeling laterally spreading flames through the accreted
+layer on the surface of a neutron star.  Building on this, in
+\citet{harpole:2021}, we explored several different rotation rates and
+initial model thermal structures to understand how the flame speed
+depends on the thermal structure of the envelope.  Both of these
+papers used a two-dimensional axisymmetric geometry.  The
+three-dimensional simulations we present here build on these works.
+We use a neutron star crust temperature of $2\times 10^8~\mathrm{K}$
+and rotation rate of 1000~Hz, as followed in \citet{harpole:2021}.
+This model was chosen here since it gives a clean well-defined flame.
+The model is constructed following the same procedure described in
 \citet{eiden:2020,harpole:2021}.
@@
 
-Our 3D simulation used a base grid of $768^2 \times 192$ zones with 2
-levels of refinement, the first jumping by a factor of 4 and the next
-by a factor of 2.  The domain has a size $(1.96608\times
-10^5~\mathrm{cm})^2 \times 2.4576\times 10^4~\mathrm{cm}$.  The simulation methodology and
+\castro\ uses the the \amrex\ adaptive mesh refinement library
+\citep{amrex_joss} to manage the discretization and domain
+decomposition.  Our 3D simulation used a base grid of $768^2 \times
+192$ zones with 2 levels of refinement, the first jumping by a factor
+of 4 and the next by a factor of 2.  We subcycle in time, so the
+coarser grids take a large timestep than the fine grids.  We used
+static mesh refinement, to achieve better load balancing, fully
+refining the atmosphere below a height of $3600~\mathrm{cm}$.  The
+domain has a size $(1.96608\times 10^5~\mathrm{cm})^2 \times
+2.4576\times 10^4~\mathrm{cm}$.  The simulation methodology and
 parameters are mostly identical to those used in \citet{harpole:2021}.
-We use an unsplit piecewise parabolic method \citep{ppmunsplit,millercolella:2002} for the
-advection and operator (Strang) splitting for the reactions, evolving
-an internal energy evolution equation during the burn, as described in
+We use an unsplit piecewise parabolic method
+\citep{ppmunsplit,millercolella:2002} for the advection and operator
+(Strang) splitting for the reactions, evolving an internal energy
+evolution equation during the burn, as described in
 \citet{strang_rnaas}.  These simulations used the 7-isotope He-burning
-network used in \citet{eiden:2020}.  The main difference with the previous
+network described in \citet{iso7}.  This was one of the networks used
+in \citet{eiden:2020}.  The main difference with the previous
 simulations is that we switched from a hydrostatic to reflecting
@@
 
-We used static mesh refinement, to achieve better load balancing,
-fully refining the atmosphere below a height of $3600~\mathrm{cm}$.
 The simulations were run on the OLCF Summit supercomputer, on 342 to
 1366 nodes, with 6 NVIDIA V100 GPUs per node.  The entire computation
-was offloaded to GPUs following the strategy described in
+was offloaded to GPUs using the \amrex\ C++ lambda-capturing mechanism, as described in
 \citet{castro_gpu}.  GPU-offloading was critical to being able to
@@
 Figure~\ref{fig:2d_abar} shows the corresponding two-dimensional
-axisymmetric simulation.  In the axisymemtric geometry, each zone is essentially
-a torus, rotated about the vertical axis.  The initial model and simulation parameters
-are identical to the three-dimensional case.  We see that
-the overall structure and evolution looks identical to that reported
-in \cite{harpole:2021}.  We see that there is a well-defined flame in the simulation plane
-that spreads outward with time.
+axisymmetric simulation.  In the axisymemtric geometry, each zone is
+essentially a torus, rotated about the vertical axis.  The initial
+model and simulation parameters are identical to the three-dimensional
+case.  We see that the overall structure and evolution looks identical
+to that reported in \cite{harpole:2021}.  We see that there is a
+well-defined flame in the simulation plane that spreads outward with
+time.  A vertical slice through the three-dimensional simulation (in
+the $x$-$z$ plane) is shown in Figure~\ref{fig:3dslice} at the final
+time, 40~ms.  Aside from the lack of reflection symmetry from the
+absence of axisymmetric geometry, it looks very similar to the two-dimensional case.
 
@@
 two-thirds of the surface.}
+\end{figure*}
+
+\begin{figure*}[t]
+\centering
+%\epsscale{0.75}
+\plotone{flame_wave_1000Hz_25cm_smallplt174562_slice}
+%\epsscale{1.0}
+\caption{\label{fig:3dslice} A vertical slice through the three-dimensional
+simulation showing the entire domain from the side.  The flame structure
+is clearly seen.}
 \end{figure*}
@@
 use two-dimensional axisymmetric simulations to explore the initial
-flame structure and spend our computational resources on increasing
-the size of our reaction networks to explore the sensitivity to
-network size and understand the overall nucleosynthesis.  We can also
-do initial explorations of full star bursts in two-dimensions, perhaps
-using axisymmetric spherical coordinates (modeling an annular region
-in $r$-$\theta$).  In this geometry, point ignition requires starting
-at the poles, but this would allow us to see the effect of varying
-gravity with latitude (due to centrifugal forces).
+flame structure and spend our computational resources on larger networks
+and bigger domains.
+
+These results also suggest that we can do initial explorations of full
+star bursts in two-dimensions, perhaps using axisymmetric spherical
+coordinates (modeling an annular region in $r$-$\theta$).  In this
+geometry, point ignition requires starting at the poles, but this
+would allow us to see the effect of varying gravity with latitude (due
+to centrifugal forces), as well as begin to consider lightcurve modeling.
 
@@
 (in space and time) algorithm for coupling hydro, diffusion, and
-reactions in Castro \citep{castro-sdc} that could be used here.  The
+reactions in \castro~\citep{castro-sdc} that could be used here.  The
 main outstanding issue with applying that work to the present problem
 is the multilevel time integration.
+
+We are currently exploring mixed H/He bursts as well as the
+sensitivity of the flame properties to the size of the reaction
+network used.  Both of these initial studies use our two-dimensional
+axisymmetric geometry.
 
@@
 
-\software{AMReX \citep{amrex_joss},
-          Castro \citep{castro,castro_joss},
+\software{\amrex~\citep{amrex_joss},
+          \castro~\citep{castro,castro_joss},
           GCC (\url{https://gcc.gnu.org/}),

### 2023-03-05 xrb3/paper.tex
turn on line numbers
a bit more code description

@@
-\documentclass[twocolumn,times,tighten]{aastex631}
+\documentclass[twocolumn,linenumbers,times,tighten]{aastex631}
 
@@
 \castro\ for modeling laterally spreading flames through the accreted
-layer on the surface of a neutron star.  Building on this, in
-\citet{harpole:2021}, we explored several different rotation rates and
-initial model thermal structures to understand how the flame speed
-depends on the thermal structure of the envelope.  Both of these
-papers used a two-dimensional axisymmetric geometry.  The
-three-dimensional simulations we present here build on these works.
-We use a neutron star crust temperature of $2\times 10^8~\mathrm{K}$
-and rotation rate of 1000~Hz, as followed in \citet{harpole:2021}.
-This model was chosen here since it gives a clean well-defined flame.
-The model is constructed following the same procedure described in
-\citet{eiden:2020,harpole:2021}.
+layer on the surface of a neutron star.  For these models,
+\castro\ evolves the compressible Euler equations with a Coriolis
+force arising from rotation, constant gravity (the plane-parallel
+approximation), reactive sources, and thermal diffusion (integrated
+explicitly).  We use a general stellar equation of state with
+radiation, ions as a perfect gas, and degenerate/relativistic
+electrons described in \citet{timmes_swesty:2000}.  Building on these
+first simulations, in \citet{harpole:2021}, we explored several
+different rotation rates and initial model thermal structures to
+understand how the flame speed depends on the thermal structure of the
+envelope.  Both of these papers used a two-dimensional axisymmetric
+geometry.  The three-dimensional simulations we present here build on
+these works.  We use a neutron star crust temperature of $2\times
+10^8~\mathrm{K}$ and rotation rate of 1000~Hz, as followed in
+\citet{harpole:2021}.  This model was chosen here since it gives a
+clean well-defined flame.  The model is constructed following the same
+procedure described in \citet{eiden:2020,harpole:2021}.
 
@@
 
-Figure~\ref{fig:2d_abar} shows the corresponding two-dimensional
-axisymmetric simulation.  In the axisymemtric geometry, each zone is
-essentially a torus, rotated about the vertical axis.  The initial
-model and simulation parameters are identical to the three-dimensional
-case.  We see that the overall structure and evolution looks identical
-to that reported in \cite{harpole:2021}.  We see that there is a
-well-defined flame in the simulation plane that spreads outward with
-time.  A vertical slice through the three-dimensional simulation (in
-the $x$-$z$ plane) is shown in Figure~\ref{fig:3dslice} at the final
-time, 40~ms.  Aside from the lack of reflection symmetry from the
-absence of axisymmetric geometry, it looks very similar to the two-dimensional case.
-
 \begin{figure*}[t]
@@
 \end{figure*}
+
+Figure~\ref{fig:2d_abar} shows the corresponding two-dimensional
+axisymmetric simulation.  In the axisymemtric geometry, each zone is
+essentially a torus, rotated about the vertical axis.  The initial
+model and simulation parameters are identical to the three-dimensional
+case.  We see that the overall structure and evolution looks identical
+to that reported in \cite{harpole:2021}.  We see that there is a
+well-defined flame in the simulation plane that spreads outward with
+time.  A vertical slice through the three-dimensional simulation (in
+the $x$-$z$ plane) is shown in Figure~\ref{fig:3dslice} at the final
+time, 40~ms.  Aside from the lack of reflection symmetry from the
+absence of axisymmetric geometry, it looks very similar to the two-dimensional case.
 

### 2023-03-06 xrb3/paper.tex
Comments on first draft of 3D XRB paper

@@
 suggest that the burning begins localized and then spreads across the
-neutron star \citep{bhattacharyya:2007}.  This spreading is inherently
-a multi-dimensional phenomena, and hydrodynamic simulations that
+neutron star \citep{bhattacharyya:2007}. {\color{red} The next sentence
+is long and it was hard to read on my initial pass over it. Could consider
+splitting the parenthetical note off as a separate sentence, and switching
+to active voice (write ``and we need hydrodynamic simulations" and replace
+``are needed" with ``in order").} 
+This spreading is inherently
+a multi-dimensional phenomenon, and hydrodynamic simulations that
 resolve the reactive zone, use realistic reaction networks, and
@@
 rp-process in mixed H/He bursts, as discussed in \citealt{rpprocess}).
+{\color{red} Could split paragraph and start next sentence with
+"Here we show...".}
 We show a first attempt at modeling the early evolution of a spreading
-hotspot in three dimensions.  This builds on our earlier two
-dimensional work \citep{eiden:2020,harpole:2021} that developed our
+hotspot in three dimensions.  This builds on our earlier two-dimensional
+work \citep{eiden:2020,harpole:2021} that developed our
 simulation framework and explored how the acceleration of the burning
@@
 \citep{castro,castro_joss}, with all of the code needed to run these
-simulations in our public github repos.
+simulations in our public GitHub repositories.
 
@@
 the hotspot ignites.  Together, all three approaches help build a
-comprehensive multi-dimensioanl picture of X-ray bursts.
+comprehensive multidimensional picture of X-ray bursts.
 
@@
 approach the edge of the domain.  As a result, we focus here on the
-early evolution.
+early evolution. {\color{red} Very minor: could start previous sentence with
+``consequently" or ``therefore" since ``as a result" appears in the previous sentence.}
 
@@
 simulation showing the mean molecular weight, $\bar{A}$.  The flame is viewed
-from above.}
+from above. {\color{red} It might be helpful to have the time for each panel either
+on the plots or in this caption.}
+}
 \end{figure}
@@
 nears the boundaries.
-
 \begin{figure*}[t]
@@
 \caption{\label{fig:2d_abar} Time-sequence of the two-dimensional axisymmetric
-simualtion using the same setup as the three-dimensional simulation shown
+simulation using the same setup as the three-dimensional simulation shown
 in Figure~\ref{fig:vr_abar}.  The domain is zoomed in, showing only the left
@@
 \caption{\label{fig:3dslice} A vertical slice through the three-dimensional
-simulation showing the entire domain from the side.  The flame structure
-is clearly seen.}
+simulation {\color{red} at 40~ms} showing the entire domain from the side.  The
+flame structure is clearly seen. {\color{red} The flame structure could probably
+be made a little clearer if this only showed the lower $10^4~\mathrm{cm}$ of the
+domain like the 2D plots. Could also try further reducing the peak $\bar{A}$ on the
+colorbar (e.g. to 5.5 or 5.25) both here and in the 2D plots to bring out the flame
+even more. I think the plots are still pretty clear with the current colorbar though.}}
 \end{figure*}
@@
 Figure~\ref{fig:2d_abar} shows the corresponding two-dimensional
-axisymmetric simulation.  In the axisymemtric geometry, each zone is
+axisymmetric simulation.  In the axisymmetric geometry, each zone is
 essentially a torus, rotated about the vertical axis.  The initial
@@
 case.  We see that the overall structure and evolution looks identical
-to that reported in \cite{harpole:2021}.  We see that there is a
+to that reported in \cite{harpole:2021}.  We also see that there is a
 well-defined flame in the simulation plane that spreads outward with
@@
 methodologies which could allow us to capture these effects at lower
-resolution.  We've already demonstrated a fully fourth-order accruate
+resolution.  We've already demonstrated a fully fourth-order accurate
 (in space and time) algorithm for coupling hydro, diffusion, and

### 2023-03-14 xrb3/paper.tex
address Kiran's comments

@@
 suggest that the burning begins localized and then spreads across the
-neutron star \citep{bhattacharyya:2007}. {\color{red} The next sentence
-is long and it was hard to read on my initial pass over it. Could consider
-splitting the parenthetical note off as a separate sentence, and switching
-to active voice (write ``and we need hydrodynamic simulations" and replace
-``are needed" with ``in order").} 
+neutron star \citep{bhattacharyya:2007}. 
 This spreading is inherently
-a multi-dimensional phenomenon, and hydrodynamic simulations that
-resolve the reactive zone, use realistic reaction networks, and
-capture the scales over which rotation is important are needed to
-understand the flame propagation and nucleosynthesis (especially
-rp-process in mixed H/He bursts, as discussed in \citealt{rpprocess}).
-{\color{red} Could split paragraph and start next sentence with
-"Here we show...".}
-We show a first attempt at modeling the early evolution of a spreading
+a multi-dimensional phenomenon, requiring hydrodynamic simulations that
+resolve the reactive zone, realistic reaction networks, and
+domain sizes that capture the scales over which rotation in order to
+understand the flame propagation and nucleosynthesis.  Detailed
+nuclear physics is especially important for understanding the 
+rp-process in mixed H/He bursts, as discussed in \citealt{rpprocess}.
+
+Here we show a first attempt at modeling the early evolution of a spreading
 hotspot in three dimensions.  This builds on our earlier two-dimensional
@@
 Moving to 3D is very expensive, so to further reduce the computational
-expense, we use an aniostropic resolution, with 32 cm resolution laterally and 16
-cm resolution vertically.  This is finer vertically than done in
-\citet{harpole:2021}.  Because
-the hotspot is placed in the center of the domain, the size of the
-domain also becomes a constraint, and as a result, the evolution time
-we can reach will become limited by the time it takes the burning to
-approach the edge of the domain.  As a result, we focus here on the
-early evolution. {\color{red} Very minor: could start previous sentence with
-``consequently" or ``therefore" since ``as a result" appears in the previous sentence.}
+expense, we use an aniostropic resolution, with 32 cm resolution
+laterally and 16 cm resolution vertically.  This is finer vertically
+than done in \citet{harpole:2021}.  Because the hotspot is placed in
+the center of the domain, the size of the domain also becomes a
+constraint, and as a result, the evolution time we can reach will
+become limited by the time it takes the burning to approach the edge
+of the domain.  Consequently, we focus here on the early evolution.
 
@@
 simulation showing the mean molecular weight, $\bar{A}$.  The flame is viewed
-from above. {\color{red} It might be helpful to have the time for each panel either
-on the plots or in this caption.}
-}
+from above. The panels show the state at 10 ms, 20 ms, and 40 ms (from top to bottom).}
 \end{figure}
@@
 simulation {\color{red} at 40~ms} showing the entire domain from the side.  The
-flame structure is clearly seen. {\color{red} The flame structure could probably
-be made a little clearer if this only showed the lower $10^4~\mathrm{cm}$ of the
-domain like the 2D plots. Could also try further reducing the peak $\bar{A}$ on the
-colorbar (e.g. to 5.5 or 5.25) both here and in the 2D plots to bring out the flame
-even more. I think the plots are still pretty clear with the current colorbar though.}}
+flame structure is clearly seen. }
 \end{figure*}

### 2023-03-14 xrb3/paper.tex
fix caption

@@
 \caption{\label{fig:3dslice} A vertical slice through the three-dimensional
-simulation {\color{red} at 40~ms} showing the entire domain from the side.  The
+simulation at 40~ms showing a region centered on the flame viewed from the side.  The
 flame structure is clearly seen. }

### 2023-03-15 xrb3/paper.tex
A pair of edits to first paragraph

@@
 resolve the reactive zone, realistic reaction networks, and
-domain sizes that capture the scales over which rotation in order to
+domain sizes that capture the scales over which rotation is important in order to
 understand the flame propagation and nucleosynthesis.  Detailed
 nuclear physics is especially important for understanding the 
-rp-process in mixed H/He bursts, as discussed in \citealt{rpprocess}.
+rp-process in mixed H/He bursts, as discussed in \citet{rpprocess}.
 

### 2023-03-17 xrb3/paper.tex
Add me as co-author
add helmeos to softwar

@@
 Berkeley, CA 94720-3411, USA}
+
+\author[0000-0003-0439-4556]{Max Katz}
+\affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
+             Stony Brook, NY 11794-3800}
 
@@
 files are available on Zenodo at \citet{xrb_data}.  We thank Alice
-Harpole and Max Katz for their contributions to the AMReX Astrophysics
-suite.
+Harpole for contributions to the AMReX Astrophysics suite.
 
@@
           GCC (\url{https://gcc.gnu.org/}),
+          helmeos \citep{timmes_swesty:2000},
           linux (\url{https://www.kernel.org/}),

### 2023-03-21 xrb3/paper.tex
add fontenc

--- 2023-03-14 xrb3/paper.tex
@@
 \usepackage{cancel}
+\usepackage[T1]{fontenc}
 
@@
 suggest that the burning begins localized and then spreads across the
-neutron star \citep{bhattacharyya:2007}. 
+neutron star \citep{bhattacharyya:2007}. {\color{red} The next sentence
+is long and it was hard to read on my initial pass over it. Could consider
+splitting the parenthetical note off as a separate sentence, and switching
+to active voice (write ``and we need hydrodynamic simulations" and replace
+``are needed" with ``in order").} 
 This spreading is inherently
-a multi-dimensional phenomenon, requiring hydrodynamic simulations that
-resolve the reactive zone, realistic reaction networks, and
-domain sizes that capture the scales over which rotation in order to
-understand the flame propagation and nucleosynthesis.  Detailed
-nuclear physics is especially important for understanding the 
-rp-process in mixed H/He bursts, as discussed in \citealt{rpprocess}.
-
-Here we show a first attempt at modeling the early evolution of a spreading
+a multi-dimensional phenomenon, and hydrodynamic simulations that
+resolve the reactive zone, use realistic reaction networks, and
+capture the scales over which rotation is important are needed to
+understand the flame propagation and nucleosynthesis (especially
+rp-process in mixed H/He bursts, as discussed in \citealt{rpprocess}).
+{\color{red} Could split paragraph and start next sentence with
+"Here we show...".}
+We show a first attempt at modeling the early evolution of a spreading
 hotspot in three dimensions.  This builds on our earlier two-dimensional
@@
 Moving to 3D is very expensive, so to further reduce the computational
-expense, we use an aniostropic resolution, with 32 cm resolution
-laterally and 16 cm resolution vertically.  This is finer vertically
-than done in \citet{harpole:2021}.  Because the hotspot is placed in
-the center of the domain, the size of the domain also becomes a
-constraint, and as a result, the evolution time we can reach will
-become limited by the time it takes the burning to approach the edge
-of the domain.  Consequently, we focus here on the early evolution.
+expense, we use an aniostropic resolution, with 32 cm resolution laterally and 16
+cm resolution vertically.  This is finer vertically than done in
+\citet{harpole:2021}.  Because
+the hotspot is placed in the center of the domain, the size of the
+domain also becomes a constraint, and as a result, the evolution time
+we can reach will become limited by the time it takes the burning to
+approach the edge of the domain.  As a result, we focus here on the
+early evolution. {\color{red} Very minor: could start previous sentence with
+``consequently" or ``therefore" since ``as a result" appears in the previous sentence.}
 
@@
 simulation showing the mean molecular weight, $\bar{A}$.  The flame is viewed
-from above. The panels show the state at 10 ms, 20 ms, and 40 ms (from top to bottom).}
+from above. {\color{red} It might be helpful to have the time for each panel either
+on the plots or in this caption.}
+}
 \end{figure}
@@
 \caption{\label{fig:3dslice} A vertical slice through the three-dimensional
-simulation at 40~ms showing a region centered on the flame viewed from the side.  The
-flame structure is clearly seen. }
+simulation {\color{red} at 40~ms} showing the entire domain from the side.  The
+flame structure is clearly seen. {\color{red} The flame structure could probably
+be made a little clearer if this only showed the lower $10^4~\mathrm{cm}$ of the
+domain like the 2D plots. Could also try further reducing the peak $\bar{A}$ on the
+colorbar (e.g. to 5.5 or 5.25) both here and in the 2D plots to bring out the flame
+even more. I think the plots are still pretty clear with the current colorbar though.}}
 \end{figure*}

### 2023-03-24 xrb3/paper.tex
spelling and grammar

--- 2023-03-17 xrb3/paper.tex
@@
 Moving to 3D is very expensive, so to further reduce the computational
-expense, we use an aniostropic resolution, with 32 cm resolution
+expense, we use an anisotropic resolution, with 32 cm resolution
 laterally and 16 cm resolution vertically.  This is finer vertically
@@
 1366 nodes, with 6 NVIDIA V100 GPUs per node.  The entire computation
-was offloaded to GPUs using the \amrex\ C++ lambda-capturing mechanism, as described in
-\citet{castro_gpu}.  GPU-offloading was critical to being able to
+was offloaded to GPUs using the \amrex\ C++ abstraction layer for parallel \texttt{for} loops, as described in
+\citet{castro_gpu}.  GPU offloading was critical to being able to
 perform these simulations, since we run more than an order of
 magnitude faster using all the GPUs on a node as compared to all of
-the CPUs on the node.  Overall, about 250,000 node-hours were used for
+the CPU cores on the node.  Overall, about 250,000 node-hours were used for
 the calculation.  The full state of the calculation was output only
@@
 the early flame evolution.  And since they are so much less
-computationally expensive, we this will allow us to explore much more
+computationally expensive, this will allow us to explore much more
 complex reaction networks and understand the nucleosynthesis in more
@@
 the flame will encounter as it propagates.  Although we have shown
-that for the flame in a Type Ia supernovae, the turbulence from the
+that for the flame in a Type Ia supernova, the turbulence from the
 era of convection is not strong enough to disrupt the

### 2023-03-26 xrb3/paper.tex
manually make the figure captions bigger

@@
 simulation showing the mean molecular weight, $\bar{A}$.  The flame is viewed
-from above. The panels show the state at 10 ms, 20 ms, and 40 ms (from top to bottom).}
+from above. The panels show the state at 10 ms, 20 ms, and 40 ms (from top to bottom).
+The colorbar shows the value of $\bar{A}$ together with the opacity assigned for
+that value---this allows us to bring out some structure in the rendering.}
 \end{figure}

### 2023-03-28 xrb3/paper.tex
submit version

--- 2023-03-24 xrb3/paper.tex
@@
-\documentclass[twocolumn,linenumbers,times,tighten]{aastex631}
+\documentclass[linenumbers]{aastex631}
 
@@
 Moving to 3D is very expensive, so to further reduce the computational
-expense, we use an anisotropic resolution, with 32 cm resolution
+expense, we use an aniostropic resolution, with 32 cm resolution
 laterally and 16 cm resolution vertically.  This is finer vertically
@@
 1366 nodes, with 6 NVIDIA V100 GPUs per node.  The entire computation
-was offloaded to GPUs using the \amrex\ C++ abstraction layer for parallel \texttt{for} loops, as described in
-\citet{castro_gpu}.  GPU offloading was critical to being able to
+was offloaded to GPUs using the \amrex\ C++ lambda-capturing mechanism, as described in
+\citet{castro_gpu}.  GPU-offloading was critical to being able to
 perform these simulations, since we run more than an order of
 magnitude faster using all the GPUs on a node as compared to all of
-the CPU cores on the node.  Overall, about 250,000 node-hours were used for
+the CPUs on the node.  Overall, about 250,000 node-hours were used for
 the calculation.  The full state of the calculation was output only
@@
 the early flame evolution.  And since they are so much less
-computationally expensive, this will allow us to explore much more
+computationally expensive, we this will allow us to explore much more
 complex reaction networks and understand the nucleosynthesis in more
@@
 the flame will encounter as it propagates.  Although we have shown
-that for the flame in a Type Ia supernova, the turbulence from the
+that for the flame in a Type Ia supernovae, the turbulence from the
 era of convection is not strong enough to disrupt the
@@
           GCC (\url{https://gcc.gnu.org/}),
-          helmeos \citep{timmes_swesty:2000},
           linux (\url{https://www.kernel.org/}),

### 2023-05-17 xrb3/paper.tex
archive
start addressing referee

--- 2023-03-24 xrb3/paper.tex
@@
-\documentclass[twocolumn,linenumbers,times,tighten]{aastex631}
+\documentclass[linenumbers]{aastex631}
 
-\usepackage{epsf,color,amsmath}
+%\usepackage{xcolor,amsmath}
 
-\usepackage{cancel}
+\usepackage[T1]{fontenc}
 
@@
 \centering
-%\epsscale{0.75}
-\plotone{abar_top_stack}
-%\epsscale{1.0}
+\plotone{full}
 \caption{\label{fig:vr_abar} Volume rendering of the three-dimensional XRB
-simulation showing the mean molecular weight, $\bar{A}$.  The flame is viewed
-from above. The panels show the state at 10 ms, 20 ms, and 40 ms (from top to bottom).}
+simulation showing the mean molecular weight, $\bar{A}$.  Two views
+are shown: the left column is viewed from above while the right column
+shows it from a shallow angle above the surface.  The panels show the state at 10 ms, 20 ms, and 40 ms (from top to bottom).
+The colorbar shows the value of $\bar{A}$ together with the opacity assigned for
+that value---this allows us to bring out some structure in the rendering.  A
+small triad in the upper right of the top panels shows the orientation,
+with red = $x$, green = $y$, and blue = $z$, our vertical direction.}
 \end{figure}

### 2023-05-17 xrb3/paper.tex
some updates for the refree

@@
-\documentclass[linenumbers]{aastex631}
+\documentclass[linenumbers,trackchanges]{aastex631}
 
@@
 approximation), reactive sources, and thermal diffusion (integrated
-explicitly).  We use a general stellar equation of state with
+explicitly)\edit1{:
+\begin{eqnarray}
+\frac{\partial \rho}{\partial t} + \nabla \cdot (\rho \Ub) &=& 0 \\
+\frac{\partial (\rho X_k)}{\partial t} + \nabla \cdot (\rho \Ub X_k) &=& \rho \omegadot_k \\
+\frac{\partial (\rho \Ub)}{\partial t} + \nabla \cdot (\rho \Ub \Ub) + \nabla p &=& \rho {\bf g} - 2 \rho {\bf \Omega} \times \Ub \\
+%% no centrifugal- \rho {\bf \Omega} \times ({\bf \Omega} \times {\bf r}) \\
+\frac{\partial (\rho E)}{\partial t} + \nabla \cdot (\rho \Ub E + p \Ub) &=& \rho \Ub \cdot {\bf g} + \nabla \cdot k_\mathrm{th} \nabla T + \rho \dot{\epsilon}
+%% no centrifugal - \rho (\Omegab \cdot \rb)(\Omegab \cdot \Ub) + \rho | \Omegab |^2 (\Ub \cdot r)
+\end{eqnarray}
+Here, $\rho$ is the mass density, $X_k$ are the mass fractions of the
+nuclear species with associated creation rates $\omegadot_k$, $\Ub$
+the velocity, $E$ is the specific total energy, $p$ is the pressure,
+${\bf g}$ is the gravitational
+acceleration, $\Omegab$ is the rotation vector, $k_\mathrm{th}$ is the thermal conductivity, and $T$ is the temperature.
+}
+ We use a general stellar equation of state with
 radiation, ions as a perfect gas, and degenerate/relativistic
-electrons described in \citet{timmes_swesty:2000}.  Building on these
+electrons described in \citet{timmes_swesty:2000} \edit1{which closes
+  the system:}
+\begin{eqnarray}
+E &=& e + \frac{1}{2} |\Ub|^2 \\
+p &=& p(\rho, e, X_k) \\
+T &=& T(\rho, e, X_k)
+\end{eqnarray}
+Building on these
 first simulations, in \citet{harpole:2021}, we explored several
@@
 network described in \citet{iso7}.  This was one of the networks used
-in \citet{eiden:2020}.  The main difference with the previous
+in \citet{eiden:2020}.  \edit1{Thermal diffusion is treated explicitly, using the conductivities from \citet{Timmes00}.
+The overall integration strategy is second-order accurate in space and time.}  The main difference with the previous
 simulations is that we switched from a hydrostatic to reflecting
@@
 \caption{\label{fig:vr_abar} Volume rendering of the three-dimensional XRB
-simulation showing the mean molecular weight, $\bar{A}$.  Two views
+simulation showing the mean molecular weight, $\bar{A}$.  \edit1{Two views
 are shown: the left column is viewed from above while the right column
@@
 small triad in the upper right of the top panels shows the orientation,
-with red = $x$, green = $y$, and blue = $z$, our vertical direction.}
+with red = $x$, green = $y$, and blue = $z$, our vertical direction.}}
 \end{figure}
@@
 40~ms of simulation time.  While initially quite symmetric, the axial
-symmetry is broken and the flame structure takes on a more tenuous
+symmetry is broken\edit1{, likely due to roundoff error,}  and the flame structure takes on a more tenuous
 form.  The panels show the flame spreading considerably through the
@@
 %\epsscale{0.75}
-\plotone{flame_wave_1000Hz_25cm_smallplt174562_slice}
+\plotone{time_series_3D}
 %\epsscale{1.0}
-\caption{\label{fig:3dslice} A vertical slice through the three-dimensional
-simulation at 40~ms showing a region centered on the flame viewed from the side.  The
+\caption{\label{fig:3dslice} Vertical slices through the three-dimensional
+simulation \edit1{at the same times as in Figure~\ref{fig:2d_abar}}, showing a region centered on the flame viewed from the side.  The
 flame structure is clearly seen. }

### 2023-05-17 xrb3/paper.tex
update

@@
 work by \citet{goodwin:2021} also explores multidimensional evolution,
-looking at the thermal transport and how affects the location where
+looking at the thermal transport and how \edit1{it} affects the location where
 the hotspot ignites.  Together, all three approaches help build a
@@
 \plotone{mass_plot}
-\caption{\label{fig:mass_plot} Mass of species scaled to initial He
+\caption{\label{fig:mass_plot} Mass of species scaled to \edit1{total}
   mass as a function of time.  The 3D simulation is shown as the solid
@@
 star bursts in two-dimensions, perhaps using axisymmetric spherical
-coordinates (modeling an annular region in $r$-$\theta$).  In this
+coordinates (modeling an annular region in \edit1{neutron star radius and colatitude}).  In this
 geometry, point ignition requires starting at the poles, but this

### 2023-05-17 xrb5/paper.tex
add xrb5 paper

%\documentclass[12pt, preprint]{aastex}
\documentclass[preprint,times,tighten]{aastex631}

\usepackage{epsf,color,amsmath}

\usepackage{cancel}

\newcommand{\sfrac}[2]{\mathchoice%
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
    \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptfont0 #1}\kern-.15em/
    \kern-.15em\lower.25ex\hbox{\the\scriptfont0 #2}}
  {\kern0em\raise.5ex\hbox{\the\scriptscriptfont0 #1}\kern-.2em/
    \kern-.15em\lower.25ex\hbox{\the\scriptscriptfont0 #2}} {#1\!/#2}}

\newcommand{\myhalf}{\sfrac{1}{2}}
\newcommand{\nph}{{n+\myhalf}}
\newcommand{\nmh}{{n-\myhalf}}

\newcommand{\inp}{\mathrm{in}}
\newcommand{\outp}{\mathrm{out}}

% boldsymbol means bold italic
\newcommand{\eb}{{\bf{e}}}
\newcommand{\Ub}{{\bf{U}}}
\newcommand{\xb}{{\bf{x}}}
\newcommand{\kb}{{\bf{k}}}
\newcommand{\Vb}{{\bf{V}}_n}
\newcommand{\Vbhat}{{\bf{\widehat{V}}}_n}
\newcommand{\Omegab}{{\bf{\Omega}}}
\newcommand{\gb}{{\bf{g}}}
\newcommand{\rb}{{\bf{r}}}

\newcommand{\pb}{p_\mathrm{base}}
\newcommand{\epsdot}{\dot{\epsilon}}
\newcommand{\qburn}{q_\mathrm{burn}}
\newcommand{\rt}{\tilde{r}_0}

\newcommand{\nablab}{\mathbf{\nabla}}
\newcommand{\dt}{\Delta\ t}

\newcommand{\omegadot}{\dot{\omega}}

\newcommand{\Hext}{H_{\rm ext}}
\newcommand{\Hnuc}{H_{\rm nuc}}
\newcommand{\kth}{k_{\rm th}}

\newcommand{\Gammaonebar}{\overline{\Gamma}_1}
\newcommand{\Sbar}{\overline{S}}

\newcommand{\etarho}{\eta_\rho}
\newcommand{\etarhoh}{\eta_{\rho~h}}

\newcommand{\Ubt}{\widetilde{\Ub}}
\newcommand{\wt}{\widetilde{w}}

\newcommand{\He}{$^4$He}
\newcommand{\C}{$^{12}$C}
\newcommand{\Fe}{$^{56}$Fe}
\newcommand{\Ni}{$^{56}$Ni}
\newcommand{\Oxygen}{$^{16}$O}

\newcommand{\isot}[2]{$^{#2}\mathrm{#1}$}
\newcommand{\isotm}[2]{{}^{#2}\mathrm{#1}}

\newcommand{\maestro}{{\sf MAESTRO}}
\newcommand{\castro}{{\sf Castro}}
\newcommand{\amrex}{{\sf AMReX}}
\newcommand{\pynucastro}{{\sf pynucastro}}
\newcommand{\microphysics}{{\sf Microphysics}}

\newcommand{\avg}[1]{\overline{#1}}
\newcommand{\avgtwod}[1]{\langle~#1 \rangle}
\newcommand{\rms}[2]{\left(\delta#1\right)_{r_{#2}}}
\newcommand{\mymax}[1]{\left(#1\right)_{\rm max}}

\newcommand{\Tb}{\ensuremath{T_\mathrm{base}}}
\newcommand{\gcc}{\mathrm{g~cm^{-3} }}
\newcommand{\cms}{\mathrm{cm~s^{-1} }}

\newcommand{\half}{\frac{1}{2}}

\setlength{\marginparwidth}{0.5in}

\usepackage{comment}
\usepackage{ulem}

\newcommand{\changeto}[2]{{{\color{red}\sout{#1} }}{{\color{blue} #2}}}
\newcommand{\remove}[1]{{\color{red}\sout{#1}}}
\newcommand{\addto}[1]{{\color{blue} #1}}

\newcommand{\MarginPar}[1]{
    \marginpar{\vskip-\baselineskip%
               \raggedright%
               \tiny\sffamily%
               {\color{red}\hrule%
               \smallskip%
               #1\par%
               \smallskip%
               \hrule}}%
}

\newcommand{\AssignTo}[1]{
    \marginpar{\vskip-\baselineskip%
               \raggedright%
               \tiny\sffamily%
               {\color{blue}\hrule%
               \smallskip%
               #1\par%
               \smallskip%
               \hrule}}%
}

\begin{document}
%======================================================================
% Title
%======================================================================
\title{Sensitivity of He Flames in X-ray Bursts to Nuclear Physics}

\shorttitle{XRB Flame
Sensitivity}
\shortauthors{Chen et al.}

\author[0000-0002-2839-107X]{Zhi Chen}
\affiliation{Department of Physics and Astronomy,
Stony Brook University,
Stony Brook, NY 11794-3800, USA}

\author[0000-0001-8401-030X]{Michael Zingale}
\affiliation{Department of Physics and Astronomy, Stony Brook University,
             Stony Brook, NY 11794-3800, USA}

%% \author[0000-0003-2300-5165]{Donald Willcox}
%% \affiliation{Lawrence Berkeley National Laboratory, Berkeley, CA}

%% \author[0000-0003-0439-4556]{Max P.\ Katz}
%% \affiliation{NVIDIA Corp}

\correspondingauthor{Zhi Chen}
\email{zhi.chen.3@stonybrook.edu}

%======================================================================
% Abstract and Keywords
%======================================================================
\begin{abstract}
Through the use of axisymmetric 2D hydrodynamic simulations, we further investigate laterally propagating flames in X-ray bursts (XRBs). Our aim is to investigate the sensitivity of a propagating helium flame to different nuclear physics.  Using the \castro\ simulation code, we confirm the phenomenon of enhanced energy generation shortly after a flame is established after by adding ${}^{12}\mbox{C}(\mbox{p}, \gamma) {}^{13}\mbox{N}(\alpha, \mbox{p}){}^{16}\mbox{O}$ to the network, in agreement with the past literature. This sudden outburst of energy leads to a short accelerating phase, causing a drastic alteration in the overall dynamics of the flame in XRBs. % that is crucial to accurately modeling the flame dynamics in XRBs. 
%We discover that the approximations made in the traditional networks like {\tt aprox13} are still accurate in the context of XRB, but insufficient to correctly model the flame propagation due to the missing rates. 
Furthermore, we investigate the influence of different plasma screening routines on the propagation of the XRB flame. We finally examine the performance of simplified-SDC, a novel approach to hydrodynamics and reaction coupling incorporated in \castro, as an alternative to operator-splitting.

\end{abstract}

\keywords{convection---hydrodynamics---methods: numerical---stars: neutron---X-rays: bursts}

%======================================================================
% Introduction
%======================================================================
\section{Introduction}\label{Sec:Introduction}

An X-ray burst (XRB) is a thermonuclear runaway caused by the ignition of the accreted fuel on the surface of a neutron star. Through Roche-lobe overflow, accreted matter is transferred from the companion star's envelope to the surface of the neutron star. The companion star is likely to have a similar composition to the Sun, i.e.\ a mixture of hydrogen, helium, and a small fraction of carbon, nitrogen, and oxygen \citep{Galloway_2020_basics_of_xrb}. Nuclear explosions are directly affected by the initial composition of the accreted layer and the accretion rate. XRBs found with a mixture of hydrogen and helium fuel layers show bursts with $\sim 5$ sec rise time, while a pure helium layer is typically more explosive and shows bursts with shorter rise times ($\sim 1$ sec). If the shell is enriched with carbon rather than helium, then the explosion is likely to have an extended rise duration, from minutes to hours known as a superburst \citep{Kuulkers_2002,Cumming_2001,Gupta_2007}. 
% This is in contrast to a few seconds in the case of regular type I XRBs, which are XRBs that have a rapid rise time followed by a gradual decline in the light curve. \MarginPar{this last sentence seems redundant}

Since the timescale of accretion between each burst is in the order of $10^4$ secs while the rise time is in the order of $\sim 1 - 10$ sec \citep{Parikh_2013}, it is unlikely for the same thermodynamic condition to exist on the entire surface of the neutron star to start burning simultaneously \citep{Shara_1982}. Therefore, nuclear ignition is likely to begin in a localized region, and spread to the rest of the neutron star \citep{Spitkovsky_2002}. This asymmetrical burning leads to modulations in the observed flux during the rotation of neutron stars, which causes asymmetrical surface brightness during XRBs \citep{strohmayer_2009}. This explains burst oscillation behavior, a millisecond period variation of XRB intensity during rise time, and the double-peak light curve for non-photospheric-radius-expansion XRBs discovered and supported by observational data \citep{Altamirano_2010,Chakraborty_2014,Bhattacharyya_2006, Kaaret_2007,Smith_1997}.

% On-sky observations and numerical simulations of XRBs are two approaches for studying the properties of the neutron star and the mechanism of XRBs. Even though there are many successes in constraining the equation of state and the structure of neutron stars through observational data of XRBs \citep{Ozel2016,Nattila2017,lattimer2014constraints,Steiner_2010}, numerical simulations are also critical to assist and complement the study of XRBs and neutron stars due to the limited events of XRBs and the limitations of observational techniques.\MarginPar{not sure we need this sentence justifying why we should do simulations}

In the past, numerous successes have been achieved in studying the properties of neutron stars and the mechanisms of XRBs using numerical simulations.
One-dimensional simulations can determine the nucleosynthesis, study the rp-process, estimate burst duration, and predict the light curve of XRBs by assuming spherical symmetry \citep{Woosley_2004,Johnston_2018,Meisel_2018,Johnston_2020}. On the other hand, multi-dimensional simulations are conducted to study the behavior of lateral flame propagation and flame structure for their nontrivial contributions to the structure of the XRB light curve \citep{eiden:2020,harpole:2021,cavecchi:2013,Cavecchi_2015,Cavecchi_2016}.

% Propagating flames can be classified into two categories, detonation and deflagration. Detonation XRB models \cite{Zingale_2001} were conducted in the early times since they're relatively computationally inexpensive. However, deflagration models are more accurate and realistic because detonation is only likely to occur with extreme thermodynamical conditions (A cite here probably).

% The effect of Coriolis force on the spread of the deflagrating nuclear flame was investigated in \cite{cavecchi:2013} to demonstrate the dependence of rotational frequency, $\nu$ and heat conductivity of the fluid, $\kappa_c$, on the flame speed as $1/(\nu \sqrt{\kappa_c})$. \cite{Cavecchi_2015} showed that flame can pass the equator if the flame originated from the pole. It also showed the differences in flame propagation depending on the origin and the destination (from pole to equator vs. from the equator to pole) on the neutron star. The effect of the magnetic field and its interaction with the Coriolis force on flame propagation is explored in \cite{Cavecchi_2016}. It is discovered that a strong magnetic field can effectively slow down the flame speed, while a weak magnetic field can cause the flame to travel at a faster speed compared to no magnetic field at all.

In our first work, \cite{eiden:2020}, we studied the overall behavior of the burning front propagation using a 2D simulation with various approximations like artificially boosted flame speed and simple networks to minimize the computational cost, as well as high rotation rate to reduce the lateral flame length scale for greater flame confinement. In \cite{harpole:2021}, we continued our work to explore the effect of rotation rate and crust temperature on flame propagation. We discovered that flame propagation was not greatly affected by the rotation rate, which is likely due to the balance between the confinement of the flame and the enhanced nuclear reaction rates caused by the increased rotation rate. It is also demonstrated that a cooler crust temperature drives a reduced flame speed and vice versa. 
Finally, in \citet{xrb_flame_3d} we showed
that 2D axisymmetric simulations compare well to full 3D simulations in capturing the nucleosynthesis of the early flame propagation.
Continuing our work, we investigate the effects of different reaction networks, plasma screening methods, and time integration methods on the simulation of lateral flame propagation in XRBs.

%======================================================================
% Numerical Approach
%======================================================================
\section{Numerical Approach}\label{Sec:numerics}
% Here briefly describe castro and the physics

We use {\castro} \citep{castro,castro_joss}, an open-source, adaptive mesh, astrophysical radiation hydrodynamics simulation code, to perform all the simulations discussed in this manuscript. The reaction networks,
integrators, equation of state \citep{timmes_swesty:2000}, thermal neutrino losses \citep{neutrino}, and conductivities
\cite{Timmes00} are contained in the related {\microphysics} package \citep{starkiller}.

Our simulation setup is the same as described in \citep{eiden:2020, harpole:2021}, so here we only reiterate the essential points. The simulations are performed using a 2D r-z cylindrical geometry, assuming azimuthal symmetry. 
The hydrodynamics is evolved using an unsplit piecewise parabolic method \citep{ppm,ppmunsplit,millercolella:2002}.
We work with a corotating frame, and assume a rotational frequency $\Omega = 1000$ Hz for the neutron star.  Gravity is assumed to be constant, since the accreted layer is thin.  Our initial model is also unchanged from our previous papers.  We employ a hydrostatic initial model to represent the neutron star's initial thermodynamic conditions, with a hot model in the left part of the domain and a cool model in the right---this drives a rightward propagating flame (a spreading hotspot for our geometry).  By default, reactions are incorporated using
operator-splitting.

%the atmosphere divided into distinct regions based on height, $z$. The isothermal base region ($0 < z < H_*$) represents the underlying neutron star and consists of the pure ${}^{56}$Ni with a temperature of $T_*$. The hyperbolic tangent transition phase ($H_* < z < H_* + 3 \delta_{\textrm{atm}}$) is characterized by a prescribed temperature and composition profile that transitions from the base region to the pure ${}^{4}$He isentropic atmosphere. The isentropic atmosphere ($z > H_* + 3 \delta_{\textrm{atm}}$) is solved for the corresponding temperature and density, subject to the hydrostatic and isentropic constraints, with the minimum temperature and density set as $T_{\textrm{lo}}$ and $\rho_{\textrm{cutoff}}$, respectively. The prescribed hyperbolic tangent transition profile on the left side of the domain ($r < r_{\textrm{pert}}$) is higher than the other side ($r > r_{\textrm{pert}}$), representing the initial temperature perturbation where the thermonuclear flame initiates.

This paper investigates the effect of reaction networks on the flame.  We explore four different networks including a standard 13-isotope alpha chain as our reference network (see Sec.\ \ref{Sec:network}), two
different plasma screening implementations (Sec.\ \ref{Sec:screening}), and how the reactions and hydro are coupled
together (Sec.\ \ref{Sec:integration}).

\begin{comment}
We use {\castro} \citep{castro,castro_joss}, an adaptive mesh, astrophysical radiation hydrodynamics simulation code, to perform all the simulations discussed in this manuscript. {\castro} evolves a general system of fully-compressible Euler's equations:

\begin{align}
    \frac{\partial \rho}{\partial t} + \nabla \cdot (\rho \mathbf{U}) &= S_{\textrm{ext},\rho} \label{Eq:castro_mass}\\
    \frac{\partial (\rho \mathbf{U})}{\partial t} + \nabla \cdot (\rho \mathbf{U}\mathbf{U}) + \nabla p &= S_{\textrm{ext},\rho \mathbf{U}} \label{Eq:castro_momen}\\
    \frac{\partial (\rho E)}{\partial t} + \nabla \cdot (\rho \mathbf{U} E + \rho E)  &= \nabla \cdot k_{\textrm{th}} \nabla T +\rho \dot{\epsilon} + S_{\textrm{ext}, \rho E} \label{Eq:castro_energy} \\
    \frac{\partial(\rho X_k)}{\partial t} + \nabla \cdot (\rho \mathbf{U}X_k) &=  \rho \dot{\omega}_k + S_{\textrm{ext}, \rho X_k} \label{Eq:castro_massfrac}
\end{align}

where $\rho$ is the density, $\mathbf{U}$ is the velocity vector, $p$ is the pressure, $X_k$ is the nuclei mass fraction with the constraint $\sum_k X_k = 1$, and $E$ is the total energy per unit mass, which can be related to the internal energy per unit mass, $e$, and the kinetic energy per unit mass as $e = E -  |\mathbf{U}|^2 / 2$. This system of equations are closed by supplying the Equation of State (EOS) such that $p = p(\rho, e, X_k)$. We follow the same convention introduced in \cite{eiden:2020}, where $(r, z, \theta)$ coordinate system is used instead of the traditional cylindrical coordinate system. Therefore, the velocity vector, $\mathbf{U}$, is now decomposed into $(u, v, w)$, where positive $w$ represents the velocity in the $\theta$ direction pointing out of the page. Thermal diffusion is modelled by incorporating an explicit source term, $\nabla \cdot k_{\textrm{th}} \nabla T$, characterized by the thermal conductivity, $k_{\textrm{th}}$, and temperature, $T$, into the energy equation (Eq. \ref{Eq:castro_energy}). The change in reacting isotopes is characterized by $\dot{\omega}_k$, the creation rates for the $k$-th isotope. The energy input term from the nuclear burning process, $\rho \dot{\epsilon}$, is added to the energy equation where $\dot{\epsilon}$ is the specific energy generation rate. Lastly, $S_{\textrm{ext}}$ represents all possible external sources terms.

In this manuscript, the same setup is used and discussed in our previous XRB papers \cite{eiden:2020, harpole:2021}. A detailed description is provided in \cite{eiden:2020}, and here we only reiterate the essential points. The simulation is performed using a 2D r-z cylindrical coordinate system by assuming azimuthal symmetry. It is convenient to work with a co-rotating frame since we're working with a rotating neutron star, which introduces Coriolis force term, $ -2\rho \mathbf{\Omega} \times \mathbf{U}$, to Euler's equation along with the gravity term, $\rho \mathbf{g}$.
%In general, the gravitational acceleration, $\mathbf{g}$, is found by solving the Poisson equation, $\nabla^2 \Phi = 4\pi G \rho$, where $\mathbf{g} = -\nabla\Phi$. 
We considered a constant gravity, $\mathbf{g}$. Considering the timescale of the XRB, the angular velocity is set to be fixed in the z-axis, $\mathbf{\Omega} = \Omega_0 \hat{z}$ to avoid extra source ter, $- \rho d\mathbf{\Omega}/dt \times \mathbf{r}$. The centrifugal force, $- \rho \mathbf{\Omega} \times (\mathbf{\Omega} \times \mathbf{r})$, is also neglected for the simplicity of the problem.
We also added geometric source terms, $\frac{1}{r}\rho w^2 \hat{r} - \frac{1}{r}\rho uw \hat{\theta}$, suggested by \cite{BERNARDCHAMPMARTIN2013170}, which arise from applying the cylindrical divergence term, $\nabla \cdot (\rho \mathbf{U}\mathbf{U})$ in Eq. \ref{Eq:castro_momen} for axisymmetric geometry.

Finally, we incorporated a density sponge mechanism \citep{ABNZ:III, Malone2014a,xrb2} with the objective of reducing the fluid velocity in the initial phase of the simulation when the flame is being established. This helps prevent the violent ejection of materials that can create unnecessary timestep constraints when studying flame propagation. The sponge mechanism adds an extra damping external source term, $\rho \mathbf{U} f/\Delta t$ and $\rho \mathbf{U}^2 f/\Delta t$, to the momentum and energy equations, where $f$ is the density dependent damping coefficient that equals to 0 once $\rho > 10^2 \ \mathrm{g} \ \mathrm{cm}^{-3}$ and $\Delta t$ is the timestep. The sponge mechanism is designed so that the damping mechanism has a greater damping effect when $\tau_{\mathrm{sponge}} < \Delta t$ with $\tau_{\mathrm{sponge}} = 10^{-7} $ s. The overall external source terms for our current setup is the following:

\begin{subequations}\label{Eq:castro_source}
\begin{equation}
    S_{\textrm{ext},\rho} = 0
\end{equation}

\begin{equation}
    S_{\textrm{ext},\rho \mathbf{U}} = -2\rho \mathbf{\Omega} \times \mathbf{U} + \frac{1}{r}\rho w^2 \hat{r} - \frac{1}{r}\rho uw \hat{\theta} + \rho \mathbf{g} + \rho \mathbf{U} \frac{f}{\Delta t}
\end{equation}

\begin{equation}
    S_{\textrm{ext}, \rho E} = \rho \mathbf{U} \cdot \mathbf{g} + \rho |\mathbf{U}|^2 \frac{f}{\Delta t}
\end{equation}

\begin{equation}
    S_{\textrm{ext}, \rho X_k} = 0. 
\end{equation}
\end{subequations}

Here we re-emphasize that Coriolis force plays an important role at limiting the horizontal length scale for flame confinement suggested by \cite{Spitkovsky_2002}. It can be shown that \citep{eiden:2020} the momentum equation (Eq. \ref{Eq:castro_momen}) has an explicit $w$ component, Eq. \ref{Eq:w_term}, a simple advection equation with the source term, $2 \Omega_0 u$ that is also solved during the simulation.

\begin{equation}\label{Eq:w_term}
    \frac{\partial w}{\partial t} + \mathbf{U} \cdot \nabla w = 2\Omega_0 u.
\end{equation}

The overall nuclear burning process is incorporated using the {\microphysics} \citep{starkiller} package. {\sf Helmholtz} EOS, a general EOS based on a table interpolation of the Helmholtz free energy described in \cite{timmes_swesty:2000}, is used. A general routine described in appendix A of \cite{Timmes00} is used to model thermal conductivity. We follow the procedure described in \cite{neutrino} to model thermal neutrino losses.

We selected 4 different reaction networks to study the dependence of reaction networks on the laterally propagating flame. We used the {\tt aprox13} network as the reference network for comparison with the other 3 networks. The details of these reactions networks are discussed in Sec. \ref{Sec:network}. We used two different plasma screening routines, {\tt SCREEN5} and {\tt CHUGUNOV2007}, to model plasma screening strengthening effects on the reaction rates. The details of the two screening methods are discussed in Sec. \ref{Sec:screening}. The RHS of the network is automatically generated using {\pynucastro} \citep{pynucastro, pynucastro2, the_pynucastro_development_2022_7239007} and ported to {\microphysics}. The RHS of the network is then solved along with the EOS, plasma screening method, thermal conductivity, and neutrino loss using VODE \citep{vode} in {\microphysics}. Lastly, the coupling between hydrodynamics and nuclear reactions are performed via either Strang-splitting \citep{strang:1968} or the simplified SDC approach \citep{Zingale_2022}. The details of these time integration methods are discussed in Sec. \ref{Sec:integration}.

{\castro} uses the adaptive mesh refinement system from {\amrex} \citep{amrex_joss}. The same simulation domain of $r = 1.8432 \times 10^5$ cm and $z = 3.072 \times 10^4$ cm is used for all of our models. At the coarse grid, we again used $1152 \times 192$ zones with two additional refinement levels with refinement ratios of 4 and 2. This means that the first refinement level is 4 times finer than the coarse level and the finest level is 2 times finer than the first level. This enables us to achieve a finest level with $9216 \times 1536$ zones, corresponding to a resolution of 20 cm. Grid refinement is limited to regions where $\rho > 2.5 \times 10^4 \ \textrm{g} \ \textrm{cm}^{-3}$ in order to avoid unncessary computational costs.

A complete description of the simulation domain necessitates the specification of a set of boundary conditions. For the left boundary, $r=0$, we have employed a symmetry or reflecting boundary condition. This choice ensures that the normal velocity of the fluid on the other side of the boundary has an opposite sign, allowing for laterally propagating flame from $z=0$. An outflow boundary condition is implemented for the right boundary. For the bottom boundary, we have employed a hydrostatic equilibrium (HSE) boundary condition described in \cite{ppm-hse}, which ensures that the HSE condition of the last interior zone of the boundary is integrated to the ghost cell. The ghost cells at the bottom of the simulation domain have a constant temperature of  $T= 1 \times 10^8$ K, consistent with the temperature of the neutron star. Moreover, the normal velocity at the bottom boundary is reflected with a zero-gradient transverse velocity. Finally, an ambient-fill boundary condition has been employed along with a standard normal velocity outflow for the top boundary. This boundary condition utilizes the initial model conditions and allows for a zero-gradient outflow for the velocity.

\subsection{Initial Models}\label{Sec:models}
% Here describe the initial setup of the problem.
% Show initial profiles?

This study begins by characterizing the atmospheric structure and initial thermodynamic conditions of the neutron star before the burning process begins. The initial model used in this simulation is identical to the ones used in our previous papers, and is discussed in detail in \cite{eiden:2020}. To construct the initial model, two 1D hydrostatic atmospheric vertical profiles are created as a function of $z$ to represent temperature-perturbed and unperturbed regions. These profiles incorporate both isentropic and isothermal conditions. The next step is to blend these two profiles as a function of $r$, resulting in the final initial model as a function of $(r,z)$. For this manuscript, we assume a pure ${}^{4} \mathrm{He}$ fuel layer on the surface of the star with a pure ${}^{56} \mathrm{Ni}$ core. Table \ref{Tab:initial_model} lists all input parameters.

\begin{table*}
\caption{\label{Tab:initial_model}
Input parameters for constructing the initial model.
}
\begin{ruledtabular}
\begin{tabular}{cc}
Input Parameters &
Values
\\ 
\colrule

$\mathbf{\Omega}$ & 1000 Hz $\hat{z}$\\
$T_*$    & $10^8$ K \\
$T_{\mathrm{hi}}$  & $2 \times 10^8$ K \\
$T_{\mathrm{lo}}$  & $8 \times 10^6$ K \\
$\delta T$  & $1.2 \times 10^9 $ K \\
$r_{\mathrm{pert}}$  & $4.096 \times 10^4$ cm \\
$\delta_{\mathrm{blend}}$ & 2048 cm \\
$\rho_{\mathrm{base}}$ & $3.43 \times 10^6 $ $\mathrm{g \ cm^{-3}}$ \\
$H_*$  & 2000 cm \\
$\delta_{\mathrm{atm}}$ & 50 cm \\
$X_{k_{\mathrm{base}}}({}^{56} \mathrm{Ni})$ & 1.0 \\
$X_{k_{\mathrm{fuel}}}({}^{4} \mathrm{He})$ & 1.0 \\
$\rho_{\mathrm{cutoff}}$ & $10^{-4}$  $\mathrm{g \ cm^{-3}}$ \\
$\mathbf{g}$ & $-1.5\times 10^{14}$ $\mathrm{g \ s^{-2}}$ $\hat{z}$\\

\end{tabular}
\end{ruledtabular}
\end{table*}

To generate the 1D atmospheric profile, we first define the base of neutron star for $z < H_*$, where $H_*$ denotes length scale of the base layer. 
Subsequently, we set the initial temperature and composition profiles, contingent upon whether $z > H_*$ or $z < H_*$. For the atmosphere region, $z > H_*$, both the temperature and composition follow a hyperbolic tangent transition that follows Eq. \ref{Eq:tanh-eqs}. In this equation, $T_*$ represents the temperature of the star, $T_{\mathrm{hi}}$ represents the highest temperature of the profile, $X_{k_{\mathrm{fuel}}}$ represents the mass fractions of the fuel layer, $X_{k_{\mathrm{base}}}$ represents the mass fractions of the base layer, and $\tilde{z} = z - H_* - 3/2 \delta_{\mathrm{atm}}$, where $\delta_{\mathrm{atm}}$ denotes a characteristic length scale. For the base region, $z < H_*$, a uniform temperature and composition profile is used, where $T = T_*$ and $X_k = X_{k_{\mathrm{base}}}$. Initially, the density and pressure for all regions are set to be the values at the base, $\rho = \rho_{\mathrm{base}}$ and $p = p_{\mathrm{base}}$, where pressure at the base, $p_{\mathrm{base}}$, is determined via EOS. These steps complete our preliminary description of the profile.

\begin{subequations}\label{Eq:tanh-eqs}

\begin{equation}
    T(z) = T_* + \frac{1}{2}\left(T_{\mathrm{hi}} - T_*\right)\left[1 + \tanh{\left( \frac{2\tilde{z}}{\delta_{\mathrm{atm}}} \right)}\right]
\end{equation}

\begin{equation}
    X_k(z) = X_{k_{\mathrm{fuel}}} + \frac{1}{2}\left(X_{k_{\mathrm{base}}} - X_{k_\mathrm{fuel}}\right)\left[1 + \tanh{\left( \frac{2\tilde{z}}{\delta_{\mathrm{atm}}} \right)}\right]
\end{equation}

\end{subequations}

We used the discretized HSE equations, Eq. \ref{Eq:hse_up} and \ref{Eq:hse_down}, to integrate the HSE states of the star starting from the base at $z=H_*$. Eq. \ref{Eq:hse_up} calculates the pressure at the current zone, $p_i$, by assuming that the pressure from the previous zone, $p_{i-1}$, is known and integrating upwards. Conversely, Eq. \ref{Eq:hse_down} follows a similar process by integrating downwards to obtain the pressure at the current zone, $p_i$, with the assumption that the pressure at the subsequent zone, $p_{i+1}$ is known. Using the HSE condition described in Eq. \ref{Eq:hse_up}, we perform Newton-Raphson iterations to obtain HSE states by integrating upwards from $z = H_*$. An isothermal profile is employed when $H_* < z < H_* + 3 \delta_{\mathrm{atm}}$. In this region, the initial temperature profile is held constant and only density is solved for, subject to the HSE constraint. The entropy and pressure are both updated using the EOS. At $z = H_* + 3 \delta_{\mathrm{atm}}$, the entropy, $s$, provides an additional constraint for the isentropic atmosphere profile. This is the region in which the hyperbolic tangent transition phase for temperature occurs.

\begin{subequations}
\begin{equation}\label{Eq:hse_up}
    p_i = p_{i-1} + \frac{1}{2}\Delta z( \rho_i + \rho_{i-1})\mathbf{g} \cdot \hat{\mathbf{z}}
\end{equation}

\begin{equation}\label{Eq:hse_down}
    p_i = p_{i+1} + \frac{1}{2}\Delta z( \rho_i + \rho_{i+1})\mathbf{g} \cdot \hat{\mathbf{z}}
\end{equation}
\end{subequations}

When $z > H_* + 3\delta_{\mathrm{atm}}$, the atmosphere transitions to an isentropic state which provides an additional entropy constraint, allowing for solving both density and temperature. However, when  $T < T_{\mathrm{lo}}$, another isothermal region is established. In this region, a uniform profile of $T = T_{\mathrm{lo}}$ is employed, while density is still solved for. As the atmosphere becomes thinner such that $\rho < \rho_{\mathrm{cutoff}}$, a constant profile of $\rho = \rho_{\mathrm{cutoff}}$ and $T = T_{\mathrm{lo}}$ is used. Once the atmosphere profile for $z > H_*$ is fully determined, we integrate downwards from $z = H_*$ using Eq. \ref{Eq:hse_down} as the constraint equation to update states in the base region. The base region is defined entirely using an isothermal profile, with density being determined by setting $T = T_*$. Different regions of the 1D hydrostatic model can then be summarized into the following:

\begin{enumerate}
    \item an isothermal base with $T = T_*$ when $0 < z < H_*$.
    \item a transition phase with prescribed temperature and composition profiles that blend the base and the atmosphere when $ H_* < z < H_* + 3\delta_{\mathrm{atm}}$.
    \item an isentropic atmosphere for flame propagation when $z > H_* + 3\delta_{\mathrm{atm}}$, $T > T_{\mathrm{lo}}$ and $\rho > \rho_{\mathrm{cutoff}}$.
    \item an isothermal low temperature buffer region when $T < T_{\mathrm{lo}}$.
    \item an isothermal and constant low density buffer region when $\rho < \rho_{\mathrm{cutoff}}$.
\end{enumerate}

The 2D initial model is obtained by combining two 1D hydrostatic vertical profiles, one representing the ``hot" model with temperature perturbations, and the other representing the ``cold" model with unperturbed-thermodynamic conditions. The construction of both models follows the description provided above, with the only difference being that $T_{\mathrm{hi}}$ is replaced with $T_{\mathrm{hi}} + \delta T$ in the ``hot" model. The ``hot" model is placed at $r < r_{\mathrm{pert}}$, where the flame is expected to start, while the ``cold" model is placed at  $r > r_{\mathrm{pert}} + \delta_{\mathrm{pert}}$. A smooth linear transition region, $r_{\mathrm{pert}} \leq r \leq r_{\mathrm{pert}} + \delta_{\mathrm{pert}}$, exists between the two models, characterized by $\delta_{\mathrm{pert}}$, which represents the length scale of the transition region. Finally, the EOS is called to update all thermodynamic quantities within the transition region. 
\end{comment}

% The blending process is described by Eq. \ref{Eq:blend}, where $r_{\mathrm{perturbed}} = 4.096 \times 10^4$ cm and the blending length scale, $\delta_{\mathrm{pert}} = 2048$ cm.

% \begin{subequations}\label{Eq:blend}
% \begin{equation}
%     f(r) = 1 - \frac{r - r_{\mathrm{pert}}}{\delta_{\mathrm{blend}}}
% \end{equation}

% \begin{equation}
%     \mathcal{Q}(r, z) = \left\{\begin{array}{lr}
%          \mathcal{Q}_{\mathrm{hot}}(z)  & \quad r < r_{\mathrm{pert}} \\

%         f(r)\mathcal{Q}_{\mathrm{hot}}(z) + \left[1 - f(r)\right]\mathcal{Q}_{\mathrm{cool}}(z) & \quad r_{\mathrm{pert}} \leq r \leq r_{\mathrm{pert}} + \delta_{\mathrm{pert}}\\

%         \mathcal{Q}_{\mathrm{cool}}(z)  & \quad r > r_{\mathrm{pert}} + \delta_{\mathrm{pert}}
%         \end{array}\right\}, \quad\mathcal{Q} \in \{p, \rho, X_k\}
% \end{equation}
% \end{subequations}

\subsection{Nuclear Reaction Networks}\label{Sec:network}
% Discuss APROX13, subch_full, subch_full_with_2rates_turned_off, subch_simple

During an XRB, nucleosynthesis involves more than a thousand isotopes \citep{Woosley_2004,Koike_2004}. However, due to computational constraints, most multi-dimensional simulations incorporate fewer than 20 isotopes. Thus, this manuscript aims to understand how the different approximations used in the nuclear physics impact on the lateral thermonuclear flame in XRBs.

\subsubsection{\tt aprox13}

In \citet{eiden:2020,harpole:2021}, we used the 13-isotope $\alpha$-chain network, {\tt aprox13} \citep{timmes_aprox13} network.  We'll
use that as the reference network for comparison here.  
%Visuals can be found in Figure \ref{fig:networks}. 
The main feature of {\tt aprox13} is the $(\alpha, \mbox{p})(\mbox{p}, \gamma)$ approximation that eliminates the intermediate nuclei involved in $(\alpha, \mbox{p})(\mbox{p}, \gamma)$ by
assuming proton equilibrium in the reactive flow.
As a result, {\tt aprox13} has a total of 13 isotopes and 31 rates. Nevertheless, the $(\alpha, \mbox{p})(\mbox{p}, \gamma)$ approximation becomes less accurate for temperatures $\gtrsim 2.5 \times 10^9$ K, potentially affecting the energy generation rates.  An illustration of this can be found in \cite{pynucastro2} (see Figure 6 there).

% Figure \ref{fig:aprox_comparison} depicts the evolution of the 13 {\tt aprox13} species under constant thermodynamic conditions, $\rho = 10^6$ g $\mathrm{cm}^{-3}$ and $T = 2 \times 10^9$ K. The solid lines represent the evolution with the $(\alpha, \mbox{p})(\mbox{p}, \gamma)$ approximation, while the dashed lines do not include this approximation. At $T = 2 \times 10^9$ K, we observe minor differences in the mass fractions of the heavier species (${}^{36}$Ar, ${}^{40}$Ca, ${}^{44}$Ti, ${}^{48}$Cr) that involved in $(\alpha, \mbox{p})(\mbox{p}, \gamma)$ approximation.

% \begin{figure}
% \centering
% \plotone{aprox_comparison.pdf}
% \caption{\label{fig:aprox_comparison} A comparison of the evolution of 13 species in {\tt aprox13} with and without $(\alpha, \mbox{p})(\mbox{p}, \gamma)$ approximation. The solid lines and the dashed lines show the evolution with and without  $(\alpha, \mbox{p})(\mbox{p}, \gamma)$ approximation, respectively. The RHS of the mass fractions are obtained using the {\pynucastro} package and integrated with $\rho = 10^6$ g $\mathrm{cm}^{-3}$ and $T = 2.0 \times 10^9$ K.}
% \end{figure}

% \MarginPar{I think for this figure, we can just refer to the similar one in the pynucastro paper}

As XRBs can reach temperatures close to the aforementioned limit ($T \gtrsim 2.5\times 10^9$ K), it important to
understand if this approximation affects the flame.  Therefore, we employ three more intricate networks, built with \pynucastro\ using the latest rates from the {\sf REACLIB} \citep{reaclib} library.  These networks are described below.

\subsubsection{{\tt subch\_full} and {\tt subch\_full\_mod}}

The {\tt subch\_full} was introduced in Appendix B of \cite{castro_simple_sdc} for simulating He burning in sub-Chandrasekhar mass white dwarfs (hence, the {\tt subch} prefix). {\tt subch\_full} network has three main modifications compared to {\tt aprox13} network. First {\tt subch\_full} explicitly includes the intermediate nuclei for the $(\alpha, \mbox{p})(\mbox{p}, \gamma)$ sequence.  Second, it has
a better representation of carbon and oxygen burning by including the endpoint nuclei of the difference branches: ${}^{12}\mbox{C}({}^{12}\mbox{C}, \mbox{p}) {}^{23}\mbox{Na} (\mbox{p}, \gamma) {}^{24}\mbox{Mg}$, ${}^{12}\mbox{C} ({}^{12}\mbox{C}, \mbox{n}){}^{23}\mbox{Mg} (\mbox{n}, \gamma) {}^{24}\mbox{Mg}$,  ${}^{16}\mbox{O} ({}^{16}\mbox{O}, \mbox{n}){}^{31}\mbox{S} (\mbox{n}, \gamma) {}^{32}\mbox{S}$, and ${}^{16}\mbox{O} ({}^{12}\mbox{C}, \mbox{n}){}^{27}\mbox{Si} (\mbox{n}, \gamma) {}^{28}\mbox{Si}$.  Since the neutron capture processes on the intermediate nuclei happen at a fast timescale, the rates
including neutrons are approximated by assuming the subsequent neutron capture is instantaneous as ${}^{12}\mbox{C} ({}^{12}\mbox{C}, \gamma) {}^{24}\mbox{Mg}$,  ${}^{16}\mbox{O} ({}^{16}\mbox{O}, \gamma) {}^{32}\mbox{S}$, and ${}^{16}\mbox{O} ({}^{12}\mbox{C}, \gamma) {}^{28}\mbox{Si}$, and the network does not include neutrons. We also note that their reverse rates are not included due to their negligible effects at $T \sim 10^9$ K.  Finally, we included the additional rates ${}^{14}\mbox{N}(\alpha,\gamma){}^{18}\mbox{F}(\alpha, \mbox{p}){}^{21}\mbox{Ne}$ and ${}^{12}\mbox{C}(\mbox{p}, \gamma) {}^{13}\mbox{N}(\alpha, \mbox{p}){}^{16}\mbox{O}$, discussed in \cite{Shen_2009, Weinberg_2006,Karakas_2008} to bypass the comparatively slow ${}^{12}$C $\alpha$-capture process, ${}^{12}\mbox{C} (\alpha, \gamma) {}^{16}\mbox{O}$, when $T \gtrsim 10^9$ K. The first $\alpha$-capture process on ${}^{14}\mbox{N}$ and ${}^{18}\mbox{F}$ led to the production of protons that can be used for the $\alpha$ chain process on ${}^{12}\mbox{C}$. When $T \gtrsim 10^9$ K, ${}^{12}\mbox{C}(\mbox{p}, \gamma) {}^{13}\mbox{N}(\alpha, \mbox{p}){}^{16}\mbox{O}$ is expected to dominate the reactive flow towards the heavy $\alpha$ chain nuclei, which can lead to a completely different end-stage composition \citep{Weinberg_2006}. Since these additional rates leave ${}^{21}\mbox{Ne}$ as the endpoint, ${}^{22}\mbox{Na}$ is added to connect all the nuclei. With all the modifications, {\tt subch\_full} has 28 isotopes and 107 rates.

{\tt subch\_full\_mod} is identical to the {\tt subch\_full} network, except that the ${}^{12}\mbox{C}(\mbox{p}, \gamma) {}^{13}\mbox{N}(\alpha, \mbox{p}){}^{16}\mbox{O}$ and its reverse reactions are turned off. This is done to explore the significance of these two rates, as described in \cite{Shen_2009, Weinberg_2006}.

\begin{comment}
This gives {\tt subch\_full\_mod} 27 isotopes and 103 rates in total. The second and third panel of Figure \ref{fig:networks} gives an overview of {\tt subch\_full} and {\tt subch\_full\_mod} networks. Here we note that \cite{Shen_2009} also discussed that the $\alpha$-capture chain, ${}^{14}\mbox{C}(\alpha,\gamma){}^{18}\mbox{O}(\alpha,\gamma){}^{22}\mbox{Ne}$, will take place at temperatures $\gtrsim 10^9$ K. However, this process is currently not considered since the energy contribution from this rate is likely to be negligible due to the small abundance of ${}^{14}\mbox{C}$ during nucleosynthesis. 
\end{comment}

\subsubsection{\tt subch\_simple}

Lastly, we present {\tt subch\_simple} network, a simplification of {\tt subch\_full} network that would resemble a more similar network to {\tt aprox13}. The network omits a total of six nuclei, namely ${}^{35}\mbox{Cl}$, ${}^{39}\mbox{K}$, ${}^{43}\mbox{Sc}$, ${}^{47}\mbox{V}$, ${}^{51}\mbox{Mn}$, and ${}^{55}\mbox{Co}$, using the $(\alpha, \mbox{p})(\mbox{p}, \gamma)$ approximation. The reverse rates of all ${}^{12}\mbox{C} + {}^{12}\mbox{C}$,  ${}^{16}\mbox{O} + {}^{16}\mbox{O}$, and ${}^{16}\mbox{O} + {}^{12}\mbox{C}$ are removed since they are not present in {\tt aprox13}. All the forward and reverse rates of ${}^{12}\mbox{C} + {}^{20}\mbox{Ne}$ , ${}^{23}\mbox{Na}(\alpha, \gamma){}^{27}\mbox{Al}$, and ${}^{27}\mbox{Al}(\alpha, \gamma){}^{31}\mbox{P}$ are also removed to simplify the network.After the simplifications, we now have 22 isotopes and 57 rates. See the bottom panel of Figure \ref{fig:networks} for visualization.

\begin{figure}
\centering
%\plotone{aprox13.pdf}
\plotone{subch_full.pdf}
%\includegraphics[width=0.88\linewidth]{subch_full_mod.pdf}
\plotone{subch_simple.pdf}
\caption{\label{fig:networks} A simple visualization {\tt subch\_full} (top) and {\tt subch\_simple} (bottom) using the \pynucastro\ package. The color bar shows the reaction rates with solar composition, $\rho = 10^6$ g $\mathrm{cm}^{-3}$, and $T = 2 \times 10^9$ K. The horizontal axis shows the atomic number, $Z$, the vertical axis shows the extra number of neutrons compared to protons for the isotopes. The gray nodes and dotted gray lines represent the $(\alpha, \mbox{p})(\mbox{p}, \gamma)$ approximation, which are not directly in the network. All reactions that have the form, $A(X, \alpha)B$ or $A(\alpha, X)B$, are hidden for better clarity. }
\end{figure}

% \begin{center}
% \plotone{aprox13.pdf}
% \plotone{subch_full.pdf}
% \plotone{subch_full_mod.pdf}
% \plotone{subch_simple.pdf}
% \end{center}

\subsection{Plasma Screening Methods}\label{Sec:screening}
% Discuss screen5 vs chugunov2007

Plasma screening is the enhancement of nuclear reaction rates due to the Coulomb coupling of the surrounding plasma electrons and ions. Plasma screening plays an important role in accurately calculating nuclear reaction rates in dense regions,
and there are many approximations for it in the literature, but these are not
often explored in real simulations.

A reaction rate cross section, $\sigma$, takes the form:
\begin{equation}
    \sigma(E) = \frac{S(E)}{E} e^{-2\pi \eta}
\end{equation}
where $S(E)$ contains the details of the nuclear physics, $E$ is the energy of particle collisions in the center-of-mass frame, and $\eta$, accounts for Coulomb barrier penetration due to quantum effects \citep{Newton2007}. The reaction rate, $R_{\textrm{th}}$, is found by integrating the cross-section over the Maxwellian velocity distribution.  If the nuclei involved in the reaction are present in a high-density region that is permeated with plasma particles, the strength of the Coulomb barrier would be reduced due to the plasma screening effect since the effective charge of the fusing nuclei are reduced. The screening enhancement factor is usually expressed as
\begin{equation}
    F_{\textrm{scr}} = \exp{(h)}
\end{equation}
where $h$ is a function that characterizes the screening magnitude. Therefore, the overall screened nuclear reaction rate is defined as
\begin{equation}\label{Eq:screened_rate}
    R_{\textrm{scr}} = F_{\textrm{scr}}R_{\textrm{th}}
\end{equation}

The plasma screening effect exhibits varying behavior depending on the specific thermodynamic conditions. 
\begin{comment}
The screening effects can be broadly classified into three burning regimes: thermonuclear burning, pycnonuclear burning, and an intermediate regime situated between the other two burning regimes. Pycnonuclear burning pertains to a temperature-independent burning process that occurs in cold and dense regions such as the core of a white dwarf. On the other hand, thermonuclear burning is a temperature-dependent process that takes place during thermonuclear explosion events such as supernovae and XRBs.
%(Add a footnote: it can be further classified into two sub regimes called zero-temperature and thermally enhanced burning). 
\end{comment}
In the thermonuclear burning regime present in XRBs, we can divide the plasma
into weak and strong plasma screening regimes depending on the Coulomb coupling parameter of ions, $\Gamma$. A rough estimate is to consider weak plasma screening when $\Gamma \ll 1$ and strong plasma screening when $\Gamma \gtrsim 1$. In this manuscript, we explore and compare the effect of two different screening routines, {\tt SCREEN5} and {\tt CHUGUNOV2007}, available in {\microphysics} on the spread of the lateral flame propagation in XRBs.

% Notes:

% 5 different regimes depend on the temperature of the fluid:

% 1) thermonuclear weak screening ($T >> T_l$ or $\Gamma << 1$): Coulomb coupling is weak and the quantum tunneling effect is affected slightly by the plasma screening. Generally, there is only one part to the enhancement factor, as $F_{scr} = \exp{(h)}$.
% 2) thermonuclear strong screening ($T_p \leq T \leq T_l$ or $\Gamma >> 1$): The plasma screening strongly enhances the reaction rate. The function $h$ can be split into two parts, $h_0$ and $h_1$.
% 3) thermo-pycnonuclear screening ($0.5T_p \leq T \leq T_p$): Intermediate regime between thermonuclear and pycnonuclear regimes.
% 4) thermally enhanced pycnonuclear ($T_q \leq T \leq 0.5T_p$): These temperatures are still so low that the majority of ions occupy their ground states. Some of them populate higher bound energy levels in the potential wells.
% 5) T=0 pycnonuclear ($T \leq T_q$). Assumes all plasma ions occupy ground states in their potential wells. Quantum tunneling and nuclear fusion occur mainly between close neighbors owning to zero-point ion vibrations. Reaction rates become temperature-independent and instead grows with density.

% In thermonuclear regimes, plasma screening is often described by the enhancement factor $F_{scr}$. In strong screening, the form of $h_0$ is well-defined but $h_1$ is not. The exact lower temperature limit of this description is uncertain. The screening factor for pycnonuclear regimes are still unknown.

\subsubsection{{\tt SCREEN5}}

{\tt SCREEN5} is the name of a widely-circulated screening routine (originally written in Fortran) that was the original screening routine used by \castro.
The overall procedure of {\tt SCREEN5} is summarized in the appendix of \cite{Wallace:1982}. In {\tt SCREEN5}, we define $\Gamma = (2/(Z_1+Z_2))^{1/3}Z_1 Z_2 \Gamma_e$, where $\Gamma_e = e^2\sqrt[3]{(4\pi n_e/3)}/(k_BT)$ describes the thermodynamic condition at which the reaction takes place, $Z$ is the charge of the fusing nuclei, $\mu_{12}$ is the reduced mass for the two fusing nuclei, $e$ is the electron charge, $n_e$ is the electron number density, and $k_B$ is Boltzmann's constant. When the plasma screening effect is weak or $\Gamma < 0.3$, {\tt SCREEN5} utilizes the equation proposed by \cite{Graboske_1973, Dewitt_1973}, which assumes that the interacting nuclei are separated by zero distance. To account for the spatial dependence of the screening enhancement factor, a more precise description of the strong plasma screening limit is utilized when $\Gamma > 0.8$. In \cite{jancovici:1977}, a quadratic dependence of the separation distance between the interacting nuclei was shown. This idea was applied to a one-component plasma by \cite{alastuey:1978}. By following a similar procedure outlined in \cite{itoh:1979}, the screening routine for one-component plasma can be extended to a multi-component plasma, which is suitable for a general mixture of ions. Finally, a weighted average between the weak and strong screening enhancement functions is used when $0.3 < \Gamma < 0.8$, the criteria for the intermediate screening regime.

%Finally, we utilize both the weak and strong screening equations to formulate a screening equation that accounts for the intermediate screening regime.

%{\tt SCREEN5} routine combines screening routines proposed by \cite{Graboske_1973, Dewitt_1973} for the weak plasma screening limit, \cite{alastuey:1978, jancovici:1977, itoh:1979} for the strong plasma screening limit, and the combination of the two for the intermediate screening, which is defined when $0.3 \leq \Gamma \leq 0.8$. We follow the procedure summarized in the appendix of \cite{Wallace:1982}

% The screening functions are obtained in the terms of chemical potential energy using the first-order approximation of the pair distribution function, Eq. \ref{Eq:pair_distribution_func}, where $H_{12}(r)$ is the screening function of interest with relation $\lim_{r \rightarrow 0} H_{12}(r) = h$

% \begin{equation}\label{Eq:pair_distribution_func}
%     g_{12}(r) = \exp{\left[-\frac{Z_1Z_2\beta e^2}{r} + H_{12}(r)\right]}
% \end{equation}

% The weak screening enhancement function, $h^{\textbf{weak}}_{12}$, proposed by \cite{Graboske_1973, Dewitt_1973} is given by Eq. \ref{Eq:weak_limit_scr}, where $Z_1$ and $Z_2$ are the charges of the two fusing reactants, $Y_i = X_i/A_i$ is the molar fraction of the nuclei, $\rho$ is density, and $T$ is temperature. \MarginPar{I don't think we need to put in all of the equations here.  We can just refer to the paper}

% \begin{equation}\label{Eq:weak_limit_scr}
%     h^{\textbf{weak}}_{12} = 1.88 \times 10^8 Z_1 Z_2 \sqrt{\left(\sum_i Z_i^2Y_i + \sum_i Z_iY_i\right)\rho/T^3}
% \end{equation}

% The strong screening enhancement function, $h^{\textbf{str}}_{12}$, in the dense matter proposed by \cite{alastuey:1978, jancovici:1977, Wallace:1982} is given by Eq. \ref{Eq:strong_limit_scr},

% \begin{align}\label{Eq:strong_limit_scr}
%     h^{\textbf{str}}_{12} =& C - \tau/3 \left[(5/32)(3\Gamma_{\textbf{eff}}/\tau)^3 - 0.014(3\Gamma_{\textbf{eff}}/\tau)^4 - 0.0128(3\Gamma_{\textbf{eff}}/\tau)^5\right] \nonumber \\
%     & -\Gamma_{\textbf{eff}} \left[0.0055(3\Gamma_{\textbf{eff}}/\tau)^4 - 0.0098(3\Gamma_{\textbf{eff}}/\tau)^5 + 0.0048(3\Gamma_{\textbf{eff}}/\tau)^6 \right]
% \end{align}

% where

% \begin{align}\label{Eq:C}
%     C = & 0.896434\Gamma_{\textbf{eff}}\left[(Z_1 + Z_2)^{5/3} - Z_1^{5/3} - Z_2^{5/3}\right] 
%      - 3.4474\Gamma_{\textbf{eff}}^{1/4} \left[(Z_1+Z_2)^{5/12} - Z_1^{5/12} - Z_2^{5/12}\right] \nonumber \\
%     & - 0.5551\left(\ln{\Gamma_{\textbf{eff}}} + (5/3)\ln{\left[Z_1Z_2/(Z_1+Z_2)\right]} \right) - 2.996
% \end{align}

% $\tau = ((27\pi^2/4)2\mu_{12}Z_1^2Z_2^2e^4/(k_BT\hbar^2))^{1/3} $ is a temperature parameter, $\Gamma_{\textbf{eff}} = (2/(Z_1+Z_2))^{1/3}Z_1 Z_2 \Gamma_e$ is the effective Coulomb coupling parameter where $\Gamma_e = e^2\sqrt[3]{(4\pi n_e/3)}/(k_BT)$, $\mu_{12}$ is the reduced mass for the two fusing nuclei, $e$ is the electron charge, $n_e$ is the electron number density, and $k_B$ is Boltzmann's constant.

% In order to account for various quantum effects of the reaction, an additional factor given by Eq. \ref{Eq:additional_fac} is added to the strong screening function, $h^{\textbf{str}}_{12}$. This factor is capped at $0.77$ to avoid the nonphysical phenomenon \cite{alastuey:1978}.

% \begin{equation}\label{Eq:additional_fac}
%     F = \ln{\left(1 - 0.0562(3\Gamma_{\textbf{eff}}/\tau)^3\right)}
% \end{equation}

% The weak screening routine is applied when $\Gamma_{\mbox{eff}} < 0.3$, while the strong screening routine is used when $\Gamma_{\mbox{eff}} > 0.8$. However, when $\Gamma_{\mbox{eff}}$ is in the intermediate plasma screening regime, that is, $0.3 < \Gamma_{\mbox{eff}} < 0.8$, we need to use both weak and strong screening equations to derive the intermediate screening function, $h^{\textbf{int}}{12}$ (Eq. \ref{Eq:intermediate_screening}), as the previously mentioned equations are only applicable in the limits of weak and strong screening. Here, $\Gamma_{\textbf{weak}} = 0.3$ and $\Gamma_{\textbf{str}} = 0.8$ are the threshold values that define the weak and strong screening regimes.

% \begin{equation}\label{Eq:intermediate_screening}
%     h^{\textbf{int}}_{12} = h^{\textbf{weak}}_{12} \left(\frac{\Gamma_{\textbf{str}} - \Gamma}{\Gamma_{\textbf{str}} - \Gamma_{\textbf{weak}}}\right)
%     + h^{\textbf{str}}_{12} \left(\frac{\Gamma - \Gamma_{\textbf{weak}}}{\Gamma_{\textbf{str}} - \Gamma_{\textbf{weak}}}\right)
% \end{equation}

\subsubsection{{\tt CHUGUNOV2007}}
%$\Gamma = (Z_1Z_2e^2)(a_{12} k_BT)$, where $Z_1$ and $Z_2$ are the charge number for each fusing nuclei, $e$ is the electron charge, $k_B$ is Boltzmann constant, $T$ is temperature, $a_{12} = (a_1 + a_2)/2$ is the average of the ion-sphere radius of the two nuclei, $a_1 = \left[3Z_1/(4\pi n_e) \right]^{1/3}$ is the ion-sphere radius for one of the reactants, $n_e = \rho/(\mu_e m_u) $ is the electron number density, $\rho$ is density, $\mu_e$ is the mean molecular weight per free electron, and $m_u$ is the atomic mass unit.

The overall implementation of {\tt CHUGUNOV2007} screening routine follows \cite{Chugunov_2007} with some modifications based on \cite{Yakovlev_2006} to extend calculations in a one-component plasma to multi-component plasma. The screening function, $h$, proposed in \cite{Chugunov_2007} used semi-classical calculations by assuming WKB Coulomb barrier penetration through the radial mean-field potential. Unlike {\tt SCREEN5} that uses separate expressions for different screening regimes, {\tt CHUGUNOV2007} employs a single expression that takes into account all screening limits up to $\Gamma \sim 600$. Additionally, we note that {\microphysics} includes other screening routines proposed by \cite{Chugunov_2009} and \cite{Chabrier:1998, Calder:2007}, although they are not covered in this discussion.

% They found that $h$ can be approximated using Eq. \ref{Eq:chugunov2007}, where $A_1 = 2.7822$, $A_2 = 98.34$, $A_3 = \sqrt{3} - A_1/\sqrt{A_2}$, $B_1 = -1.7476$, $B_2 = 66.07$, $B_3 = 1.12$, and $B_4 = 65$. $\tilde{\Gamma}$ is the modified Coulomb coupling parameter given by Eq. \ref{Eq:tildegamma}, where $\Gamma = (Z_1Z_2e^2)/(a_{12} k_BT)$ is the Coulomb coupling parameter,  $a_{12} = (a_1 + a_2)/2$ is the average of the ion-sphere radius of the two nuclei, $a_j = \sqrt[3]{(3 Z_j)/(4\pi n_e)}$ is the ion-sphere radius for one of the reactants, $\alpha = 0.022$, $\beta = 0.41 - 0.6/\Gamma$, $\gamma = 0.06 + 2.2/\Gamma$, $\zeta = \left(4\Gamma^2a_B/(\pi^2a_{12})\right)^{1/3}$, $a_B = \hbar^2/(Z_1Z_2e^22\mu_{12})$ is the Bohr radius, and $\mu_{12}$ is the reduced mass of the two reactants.

% \begin{equation}\label{Eq:chugunov2007}
%     h^{\textbf{fit}}_{MF}(\tilde{\Gamma}) = \tilde{\Gamma}^{3/2} \left(\frac{A_1}{\sqrt{A_2 + \tilde{\Gamma}}} + \frac{A_3}{1+\tilde{\Gamma}}\right) + \frac{B_1\tilde{\Gamma}^2}{B_2+\Gamma} + \frac{B_3\tilde{\Gamma}^2}{B_4 + \tilde{\Gamma}^2} 
% \end{equation}

% \begin{equation}\label{Eq:tildegamma}
%     \tilde{\Gamma} = \frac{\Gamma}{(1 + \alpha\zeta +\beta\zeta^2 + \gamma\zeta^3)^{1/3}}
% \end{equation}

% Unlike {\tt SCREEN5} that uses separate expressions for different screening regimes, {\tt CHUGUNOV2007} employs a single expression that takes into account all screening limits up to $\Gamma \sim 600$. Additionally, we note that {\microphysics} includes other screening routines proposed by \cite{Chugunov_2009} and \cite{Chabrier:1998, Calder:2007}, although they are not covered in this discussion.

% - Semi-classical calculations assuming WKB Coulomb barrier penetration through the radial mean-field potential.
% - Compare with Path Integral Monte Carlo calculations done by Miltizer and Pollock of Coulomb tunneling in nuclear reactions in dense matter.
% - Introduces dimensionless mean-field potential $u(r)$ (Eq. 12-14), using this mean-field potential, one can construct the mean-field reaction rate given by Eq. 16, $R^{MF}$. If we take the mean-field potential, $H(r) = 0$, then $R^{MF} = R_{th}$. In WKB approximation, the enhancement factor is then $F^{MF}_{scr} = R^{MF}(H)/R^{MF}(H=0)$
% - The final screening factor is given by $F^{MF}_{scr} = \exp{(h_0(\tilde{\Gamma}))}$.

\subsection{Time Evolution Methods}\label{Sec:integration}
%Discuss default(STRANG+CTU) vs simplified sdc

The default method for coupling hydrodynamics and nuclear reactions in {\castro} is the classic Strang-splitting method \citep{strang:1968}. This operator-splitting approach considers the hydrodynamics and nuclear reactions to be independent processes. 
The overall procedure of Strang-splitting is as follows: first, the reactive part of the system is integrated over half of the timestep, $\Delta t/2$, using a standard ODE solver to determine the solution of an intermediate state, $\mathcal{U}^\star$, which is centered in time. Advection then
evolves $\mathcal{U}^\star$ through a full timestep $\Delta t$
yielding the state $\mathcal{U}^{\star\star}$.  Finally,
the second half-timestep of burning is done to obtain the final state $\mathcal{U}^{n+1}$. By applying the advection and hydrodynamic operator to an intermediate state that has already incorporated the nuclear reaction effect over $\Delta t/2$, an indirect coupling is achieved between the two operations and achieves a second-order accuracy in time.  The Strang-splitting implementation
in \castro\ was described in \cite{castro_strang}.

\begin{comment}
If $\mathcal{A}$ represents the operator that updates both advection and hydrodynamic sources and $\mathcal{R}$ represents the operator that updates nuclear reactions, then the overall procedure of updating the state, $\mathcal{U}^n$, with time index, $n$, to a new state in the next time step, $\mathcal{U}^{n+1}$, can be summarized as:
\begin{equation}\label{Eq:strang-update}
    \mathcal{U}^{n+1} = \mathcal{R}_{\Delta t/2} \mathcal{A}_{\Delta t} \mathcal{R}_{\Delta t/2} \mathcal{U}^n
\end{equation}
\end{comment}

Spectral deferred correction (SDC) algorithms  \citep{dutt2000sdc, bourlioux2003reaction_sdc, castro-sdc} are iterative schemes for constructing higher-order accuracy solutions for ODEs by solving correction terms using low-order accuracy solvers like forward and backward Euler solvers. Correction terms are computed during each iteration to give improved solutions which are used as a source term in the next iteration. An arbitrarily high-order accuracy solution is then achieved after a series of correction sweeps.

%The simplified SDC method is a technique that leverages the general SDC idea and uses two iterations to achieve second-order accuracy in time. The algorithm involves three main steps. Firstly, a modified hydrodynamic update is performed on $\mathcal{U}^{n,(k)}$ using CTU and PPM, along with an additional source correction term. Here, the iteration counter is denoted by $k$, which takes values 0 and 1. The correction term captures the nuclear reaction effect for direct coupling, and the outcome is an intermediate state $\mathcal{U}^{n+1/2, (k)}$. Secondly, assuming that $\mathcal{U}^{n+1/2, (k)}$ is piecewise constant in time, it is used as an additional source term and integrated together with the reaction source term $\mathcal{R}$, using a standard ODE solver. Lastly, the correction term used in the first step is computed to complete the first iteration. A more detailed description of the simplified SDC method can be found in \cite{Zingale_2022}.

A simplfiied-SDC algorithm was introduced in \castro\ in \cite{castro_simple_sdc}.  The simplified-SDC algorithm explicitly couples reactions and hydrodynamics by including a reactive source in the hydrodynamics
interface state prediction and including an advective source in the reaction ODE integration.  
The simplified-SDC scheme offers advantages over the traditional Strang-splitting method even though they both achieve second-order accuracy in time. Specifically, the simplified-SDC scheme provides a direct coupling between hydrodynamics and nuclear reactions, which eliminates the splitting error associated with operator splitting. It also reduces the stiffness of solving reaction equations, leading to reduced computational expenses under extreme thermodynamic conditions.  For
explosive reactive flows, it was observed that the simplified-SDC method
provides a more accurate evolution than the Strang-split method.  In this paper, we seek to understand what advantages, if any, it has for the more moderate burning in XRBs.

%======================================================================
% Results
%======================================================================
\section{Simulations and Results}\label{Sec:results}

We present a total of seven XRB simulations models using different combinations of reaction networks, time evolution methods, and screening routines. In order to investigate the sensitivity of the propagating flame to nuclear reactions, the default screening routine, {\tt SCREEN5}, and time evolution methods, Strang-splitting, are used with a variation in the networks: {\tt aprox13}, {\tt subch\_full}, {\tt subch\_full\_mod}, {\tt subch\_simple}. We replaced the {\tt SCREEN5} screening routine with {\tt CHUGUNOV2007} for the {\tt aprox13} model with Strang-splitting to investigate the difference in their performance. Lastly, in order to test the performance of the simplified SDC method, we ran two additional simulations using the two selected reaction networks, {\tt aprox13} and {\tt subch\_full}, with simplified SDC instead of Strang-splitting. The overall summary of the simulation settings is shown in Table \ref{Tab:settings}, and we will now refer to each specific model using its name.

\begin{table*}
\caption{\label{Tab:settings}
Various settings used for each simulation.
}
\begin{ruledtabular}
\begin{tabular}{cccc}
Name &
Network &
Integration &
Screening
\\

\colrule

{\tt aprox13} & {\tt aprox13} & Strang-splitting & {\tt SCREEN5} \\
{\tt subch\_full} & {\tt subch\_full} & Strang-splitting & {\tt SCREEN5} \\
{\tt subch\_full\_mod} & {\tt subch\_full\_mod} & Strang-splitting & {\tt SCREEN5} \\
{\tt subch\_simple} & {\tt subch\_simple} & Strang-splitting & {\tt SCREEN5} \\
{\tt aprox13\_sdc} & {\tt aprox13} & simplified SDC & {\tt SCREEN5} \\
{\tt subch\_full\_sdc} & {\tt subch\_full} & simplified SDC & {\tt SCREEN5} \\
{\tt aprox13\_chu} & {\tt aprox13} & Strang-splitting & {\tt CHUGUNOV2007}

\end{tabular}
\end{ruledtabular}
\end{table*}

\subsection{Reaction Network Comparison}\label{Sec:result_network}
% aprox13, subch_full, subch_simple, subch_full_mod

% Column density?
% double peak structure in enuc?
% Include a figure showing the computational efficiency of these networks?

\begin{figure*}
\centering
\plotone{network_abar_50ms.pdf}
\caption{\label{fig:network_abar_50ms} Slice plots comparing $\Bar{A}$ for {\tt aprox13} (top panel), {\tt subch\_full} (second panel from top), {\tt subch\_full\_mod} (third panel), and {\tt subch\_simple} (last panel) at 50 ms. }
\end{figure*}

\begin{figure*}
\centering
\plotone{network_enuc_50ms.pdf}
\caption{\label{fig:network_enuc_50ms} Slice plots comparing $\dot{e}_{\textrm{nuc}}$ for {\tt aprox13} (top panel), {\tt subch\_full} (second panel from top), {\tt subch\_full\_mod} (third panel), and {\tt subch\_simple} (last panel) at 50 ms.}
\end{figure*}

% \begin{figure*}
% \centering
% \plotone{network_z_50ms.pdf}
% \caption{\label{fig:network_z_50ms} Slice plots comparing $w$ (velocity in $\theta$ direction) for {\tt aprox13} (top panel), {\tt subch\_full} (second panel from top), {\tt subch\_full\_mod} (third panel), and {\tt subch\_simple} (last panel) at 50 ms.}
% \end{figure*}

\subsubsection{Global behavior}

Figure \ref{fig:network_abar_50ms} shows a comparison of the mean molecular weight, $\Bar{A}$, for the four networks at $t = 50$ ms. Regions with a larger $\Bar{A}$ represents the ash of burning process which gives a rough estimate of where the burning process has already taken place. One immediate feature is the similarity of the ash structure between {\tt aprox13} and {\tt subch\_full\_mod}, as well as between {\tt subch\_full} and {\tt subch\_simple}. Unlike {\tt aprox13} and {\tt subch\_full\_mod}, there is a thicker overall ash structure on the surface indicating a much more rigorous burning process in general. The sudden increase in the ash height from $r = 5 \times 10^4$ cm to $10^5$ cm further implies a non-uniform burning has taken place, unlike the other two models. The darker color means that the ash in {\tt subch\_full} and {\tt subch\_simple} is composed of heavier nuclei suggesting a faster reactive flow burning to the heavier nuclei compared to the other two models. The ash also extends further out suggesting a faster flame speed compared to {\tt aprox13} and {\tt subch\_full\_mod}. 
The energy generation rate (shown in Figure \ref{fig:network_enuc_50ms}) shows the same trends: {\tt subch\_full} and {\tt subch\_simple} clearly possess a higher $\dot{e}_{\textrm{nuc}}$ in both magnitude and region coverage compared to the other two. 
% We continue to see the similar behavior of the 4 different simulation models in Figure \ref{fig:network_z_50ms}, which shows the velocity induced by the Coriolis force going out of the plane.

\begin{figure*}
\centering
\plotone{network_time_profile.pdf}
\caption{\label{fig:network_time_profile} Time profiles showing the weighted temperature (left panel) and energy generation rate (right panel) of the burning front for the 4 simulations models: {\tt aprox13}, {\tt subch\_full}, {\tt subch\_full\_mod}, and {\tt subch\_simple}.}
\end{figure*}

In order to quantify how the flame changes over time, we look at the
density-weighted profile of temperature and energy generation rate in the burning, similar to the procedure in \citet{harpole:2021}. 
In order to filter out the burning regions, we selected cells that are in the 99th percentile for temperature or energy generation rate. A density-weighted average is then applied to temperature and energy generation rate using Equation \ref{Eq:weighted_equation}, where $\left<Q\right>_w$ is the weighted quantity, $C_{99}(Q)$ are cells where the quantity, $Q$, is in the 99th percentile, and $\rho(c_i)$ and $Q(c_i)$ are the density and the quantity in $c_i$ cell.

\begin{equation}\label{Eq:weighted_equation}
    \left<Q\right>_w = \frac{\sum\limits_{c_i}^{} \rho(c_i) Q(c_i)}{\sum\limits_{c_i} \rho(c_i)}; c_i \in C_{99}(Q)
\end{equation}

The weighted results are shown in Figure \ref{fig:network_time_profile}. The profiles of {\tt subch\_full} and {\tt subch\_simple} are nearly identical profiles in both weighted temperature and energy generation rate. As the primary distinction between these two is the $(\alpha, \mbox{p})(\mbox{p}, \gamma)$ approximation for heavier nuclei, we can conclude that this approximation remains accurate in the context of XRB. In general, there is no uniform shape in both T and $\dot{e}_{\textrm{nuc}}$ for {\tt subch\_full} and {\tt subch\_simple}. After the flame-establishing phase at $\sim 3$ ms, there is a slight acceleration in temperature followed by a small decline, corresponding to the first spike in $\dot{e}_{\textrm{nuc}}$. However, there is a sudden burst of energy production output from $\sim 10$ ms after the first spike for {\tt subch\_full} and {\tt subch\_simple}. The sudden spike in energy production in {\tt subch\_full}, as compared to {\tt subch\_full\_mod}, is attributed to the inclusion of ${}^{12}\mbox{C}(\mbox{p}, \gamma) {}^{13}\mbox{N}(\alpha, \mbox{p}){}^{16}\mbox{O}$ rates, as it is the only distinction between these two networks. By looking at the weighted $\dot{e}_{\textrm{nuc}}$ profile, we determined that the burning acceleration phase had ended before 50 ms, as anticipated from previous slice plots. In general, the changes in $\dot{e}_{\textrm{nuc}}$ are well reflected on the weighted-temperature profile.

Unlike {\tt subch\_full} or {\tt subch\_simple}, {\tt aprox13} and {\tt subch\_full\_mod} have a much more steady burning process. Initially, {\tt subch\_full\_mod} appears to keep pace with {\tt subch\_full} and {\tt subch\_simple}, but $\dot{e}_{\textrm{nuc}}$ declines rapidly after reaching its peak following the flame-establishing phase. This is the consequence of the missing ${}^{12}\mbox{C}(\mbox{p}, \gamma) {}^{13}\mbox{N}(\alpha, \mbox{p}){}^{16}\mbox{O}$, which puts a heavy limitation on the subsequent $\alpha$-chain burning processes toward the heavy elements. However, $\dot{e}_{\textrm{nuc}}$ for {\tt subch\_full\_mod} gradually accelerates, reaching its peak at 120 ms similar to {\tt aprox13}, but with a faster rate. In contrast, $\dot{e}_{\textrm{nuc}}$ for {\tt subch\_full} and {\tt subch\_simple} gradually decrease during the later stages of burning after the burst.

\begin{figure*}
    \centering
    \plotone{network_aprox13_nuc_frac.pdf}
    \caption{\label{fig:network_aprox13_nuc} Slice plots showing the mass fractions of ${}^{12}$C, ${}^{16}$O, and ${}^{28}$Si for {\tt aprox13} at 20 ms.}
\end{figure*}

\begin{figure*}
    \centering
    \plotone{network_subch_nuc_frac.pdf}
    \caption{\label{fig:network_subch_nuc} Slice plots showing the mass fractions of ${}^{12}$C, ${}^{16}$O, and ${}^{28}$Si for {\tt subch\_full} at 20 ms.}
\end{figure*}

\begin{figure*}
    \centering
    \plotone{network_subch_mod_nuc_frac.pdf}
    \caption{\label{fig:network_subch_mod_nuc} Slice plots showing the mass fractions of ${}^{12}$C, ${}^{16}$O, and ${}^{28}$Si for {\tt subch\_full\_mod} at 20 ms.}
\end{figure*}

\subsubsection{Nucleosynthesis}

In order to investigate the burning process in detail, Figure \ref{fig:network_aprox13_nuc}, \ref{fig:network_subch_nuc}, and \ref{fig:network_subch_mod_nuc} show the mass fractions of ${}^{12}$C, ${}^{16}$O, and ${}^{28}$Si for {\tt aprox13}, {\tt subch\_full}, and {\tt subch\_full\_mod} at 20 ms, respectively. We exclude {\tt subch\_simple} due to its similarity to {\tt subch\_full}.  For {\tt aprox13}, all the burning products are concentrated in a thin region in the left of the domain, as we would expect
given the high temperature behind the flame.  Notably, there is an abundance of ${}^{12}$C, as {\tt aprox13} relies on the relatively slow $\alpha$-capture rate to convert ${}^{12}$C to ${}^{16}$O. Additionally, ${}^{16}$O is predominately transformed into ${}^{28}$Si, which is the primary end product of the network.

With {\tt subch\_full}, we see that the ${}^{12}$C is much more depleted and there is much more production of heavy $\alpha$ chain isotopes through the $\alpha$-chain, particularly ${}^{28}$Si, at $t = 20$ ms. This phenomenon is consistent with the results of \cite{Weinberg_2006}. The burning path, ${}^{12}\mbox{C}(\mbox{p}, \gamma) {}^{13}\mbox{N}(\alpha, \mbox{p}){}^{16}\mbox{O}$, provides an efficient route for converting ${}^{12}$C to ${}^{16}$O compared to the $\alpha$-capture rate. 
%Consequently, when $T \sim 1.3 \times 10^9$ K, the flow from ${}^{12}$C to ${}^{16}$O can surpass the triple-$\alpha$ process, resulting in a depletion of ${}^{12}$C and an amplification in the nuclear energy generation rates. 
%Consequently, it leads to the abundant production of heavy elements such as ${}^{28}$Si, as shown in Figure \ref{fig:network_subch_nuc}. 
%The early depletion of ${}^{12}$C also causes a shortage of burning fuels during the later burning stage, corresponding to an overall decline in the energy generation rate following the outburst.

Similar to {\tt aprox13}, {\tt subch\_full\_mod} exhibits a considerable amount of ${}^{12}$C due to the absence of ${}^{12}\mbox{C}(\mbox{p}, \gamma) {}^{13}\mbox{N}(\alpha, \mbox{p}){}^{16}\mbox{O}$ rate. However, unlike {\tt aprox13}, higher concentration of ${}^{16}$O is observed in the upper vertical regions in {\tt subch\_full\_mod}. This is likely the result of incorporating various additional rates, including more up-to-date rates from {\sf REACLIB} library, in {\tt subch\_full\_mod} compared to {\tt aprox13}. Nevertheless, the quantity of ${}^{16}$O in the bottom regions remains similar to those seen in {\tt aprox13}, as the slightly higher temperature in the bottom region facilitates the burning of these fuels into ${}^{28}$Si.

\begin{figure*}
    \centering
    \plotone{network_species_summary.pdf}
    \caption{\label{fig:network_species_summary} The overall evolution of the total mass for ${}^{12}$C, ${}^{16}$O, ${}^{20}$Ne, ${}^{24}$Mg, ${}^{28}$Si, and ${}^{32}$S for the 4 simulations models: {\tt aprox13}, {\tt subch\_full}, {\tt subch\_full\_mod}, and {\tt subch\_simple}.}
\end{figure*}

Figure \ref{fig:network_species_summary} shows the total mass of the key $\alpha$-chain isotopes to track their evolution. The initial spike in the production of ${}^{16}$O, ${}^{20}$Ne, ${}^{24}$Mg, corresponds to the initial peak in energy generation rates for all {\tt subch} networks shown in Figure \ref{fig:network_time_profile}. Although ${}^{12}\mbox{C}(\mbox{p}, \gamma) {}^{13}\mbox{N}(\alpha, \mbox{p}){}^{16}\mbox{O}$ efficiently burns ${}^{12}$C into ${}^{16}$O, resulting in a lower abundance of ${}^{12}$C in {\tt subch\_full} and {\tt subch\_simple}, a considerable amount of ${}^{12}$C still accumulates before $t \sim 18$ ms. Nevertheless, when $t \gtrsim 18$ms, corresponding to $T \sim 1.3 \times 10^9$ K, the flow from ${}^{12}$C to ${}^{16}$O can surpass the triple-$\alpha$ process, resulting in a depletion of ${}^{12}$C. This phenomenon is consistent with \cite{Weinberg2007}, which claims a similar behavior at $t \sim 1.2 \times 10^9$ K. Consequently, there is an amplification in the nuclear energy generation rates and the mass production of heavy isotopes such as ${}^{28}$Si. The early exhaustion of ${}^{12}$C also causes a shortage of burning fuels during the subsequent burning stage for {\tt subch\_full} and {\tt subch\_simple}. This is because the nuclear burning process is now bottle-necked by the triple-$\alpha$ process, corresponding to an overall decline in the energy generation rate following the outburst. In contrast, both {\tt aprox13} and {\tt subch\_full\_mod} exhibit a continuous accumulation of ${}^{12}$C since the networks are bottle-necked by the inefficient $\alpha$-capture rate on ${}^{12}$C to ${}^{16}$O. However, unlike {\tt aprox13}, {\tt subch\_full\_mod} demonstrates slightly more effective burning paths for burning ${}^{12}$C at $t \gtrsim 70$ ms. This phenomenon likely accelerates the overall nuclear burning process, resulting in significantly higher nuclear energy generation compared to {\tt aprox13} during the late-stage burning, as shown in Figure \ref{fig:network_time_profile}.

\begin{figure*}
    \centering
    \plotone{network_u_v_enuc_25ms.pdf}
    \plotone{network_u_v_enuc_100ms.pdf}
    \caption{\label{fig:network_u_v_enuc} $u-v$ phase plots for {\tt aprox13} (panels in the first column on the left), {\tt subch\_full} (second column), {\tt subch\_full\_mod} (third column), {\tt subch\_simple} (fourth column) at $t = 25$ ms (top 4 panels) and $t = 100$ ms (bottom 4 panels). The x-axis, $u$, shows the velocity in the $r$ direction, whereas the y-axis, $v$, shows the velocity in the $z$ direction. The color bar shows $\dot{e}_{\textrm{nuc}}$.}
\end{figure*}

\subsubsection{Dynamics}

Figure \ref{fig:network_u_v_enuc} shows the $u$-$v$ phase plot (radial velocity vs. vertical velocity) for the four simulations models at $t = 25$ ms and $t = 100$ ms, colored by $\dot{e}_{\textrm{nuc}}$. Notably, the peak of the energy spike occurs at $t \sim 25$ ms for {\tt subch\_full} and {\tt subch\_simple},  resulting in the comparatively larger distribution in their $u$-$v$ phase space plots.  By 100~ms, the distribution appears to shrink considerably, likely due to decrease in the energy output as the flame is
established. In contrast, the $u$-$v$ phase space distribution for {\tt aprox13} and {\tt subch\_full\_mod} increased from $t = 25$ ms to $t = 100$ ms. However, there is one common feature slowly forming in the late stage where higher $\dot{e}_{\textrm{nuc}}$ are preferably located in regions with a velocity opposite to the flame propagation. This a common feature observed in \cite{eiden:2020, harpole:2021}.

\begin{figure*}
\centering
\plotone{network_front.pdf}
\caption{\label{fig:network_front} Flame front position as a function time for {\tt aprox13}, {\tt subch\_full}, {\tt subch\_full\_mod}, and {\tt subch\_simple}. The results of the fitting function, Eq. \ref{Eq:quadratic_fit} and \ref{Eq:tanh_fit}, is also shown in the dashed lines.}
\end{figure*}

To examine an overall impact of different reaction networks on the dynamics of the laterally propagating flame, Figure \ref{fig:network_front} shows the radial flame front position as a function of time. We follow the approach outlined in \cite{eiden:2020}, where the position of the flame front is defined as the location at which $\dot{e}_{\textrm{nuc}}$ of the 1D radial profile first drops to $0.1\%$ of the global maximum $\dot{e}_{\textrm{nuc}}$ for $r > r_{\mbox{max}(\dot{e}_{nuc})}$. As expected, {\tt subch\_full} and {\tt subch\_simple} have nearly identical burning front positions, with a sudden burst of acceleration followed by a gradual decline in velocity. These behaviors are consistent with the temperature and $\dot{e}_{\textrm{nuc}}$ profiles we examined previously. On the other hand, {\tt subch\_full\_mod} and {\tt aprox13} exhibit a similar gradually accelerating burning front position, with a faster acceleration for {\tt subch\_full\_mod}.

\begin{table*}
\caption{\label{Tab:network_fitted_param}
Fitted parameters of the fitting functions Eq. \ref{Eq:quadratic_fit} and \ref{Eq:tanh_fit} for {\tt aprox13}, {\tt subch\_full}, {\tt subch\_full\_mod}, and {\tt aprox13\_chu}. The fitting function is applied for $t > 8$ ms.
}

\begin{ruledtabular}
\footnotesize
\centering
%\resizebox{1\textwidth}{!}{
\begin{tabular}{ccccccc}
\small 
Name &
$a_0$ [$\mbox{km} \ \mbox{s}^{-2}$] &
$v_0$ [$\mbox{km} \ \mbox{s}^{-1}$] &
$r_0$ [km] &
A [km]&
B [s]&
C

\\ 
\colrule
{\tt aprox13} & $24.22 \pm 0.23$ & $2.812 \pm 0.015$ & $0.7680\pm 0.0004 $ & N/A & N/A & N/A\\

{\tt subch\_full} & $9.22 \pm 0.82$ & $4.489 \pm 0.067$ & $0.8646 \pm 0.0014$ & $0.150 \pm 0.001$ & $0.0093 \pm 0.0001$ & $-2.551 \pm 0.035$ \\

{\tt subch\_full\_mod} & $58.53 \pm 0.24$ & $2.122 \pm 0.016$ & $0.7773 \pm 0.0004$ & N/A & N/A & N/A \\

{\tt subch\_simple} & $15.60 \pm 0.93$ & $3.961 \pm 0.076$ & $0.8745 \pm 0.0016$ & $0.153 \pm 0.002$ & $0.0091 \pm 0.0001$ & $-2.575 \pm 0.040$\\

{\tt aprox13\_chu} & $22.96 \pm 0.20$ & $2.578 \pm 0.0133$ & $0.7728 \pm 0.0004$ & N/A & N/A & N/A\\

% {\tt aprox13\_sdc} & $28.09 \pm 0.20$ & $2.392 \pm 0.013$ & $0.7836\pm 0.0004 $ & N/A & N/A & N/A\\

% {\tt subch\_full\_sdc} & $2.34 \pm 0.74$ & $4.947 \pm 0.061$ & $0.8601 \pm 0.0013$ & $0.146 \pm 0.001$ & $0.0098 \pm 0.0001$ & $-2.424 \pm 0.030$ \\

\end{tabular}
%}
\end{ruledtabular}
\end{table*}

We fit the flame front position with a simple function to estimate
the flame speed.  For the {\tt aprox13} and {\tt subch\_full\_mod} networks, we use the same expression
as in \citet{eiden:2020}:
\begin{equation}\label{Eq:quadratic_fit}
    r(t) = \frac{1}{2}a_0 t^2 + v_0 t + r_0
\end{equation}
For the {\tt subch\_full} and {\tt subch\_simple}, a hyperbolic tangent function is added to the fitting function to account for the burst of acceleration at $10 \mbox{ms} \lesssim t \lesssim 25 \mbox{ms}$.
\begin{equation}\label{Eq:tanh_fit}
    r(t) = A\tanh{\left(\frac{t}{B} + C\right)} + \frac{1}{2}a_0 t^2 + v_0 t + r_0
\end{equation}
Both fitting functions are applied for $t > 8$ ms, and the fitted parameters along with their respective errors are shown in Table \ref{Tab:network_fitted_param}. The errors are calculated by taking the square root of the diagonal of the covariance matrix.

\begin{table*}
\caption{\label{Tab:network_instan_vel}
Instantaneous flame propagation speed at $t = 23$ ms and $t = 100$ ms for {\tt aprox13}, {\tt subch\_full}, {\tt subch\_full\_mod}, {\tt subch\_simple}, and {\tt aprox13\_chu}. $t = 23$ ms and $t=100$ ms represent the acceleration phase for {\tt subch\_full} and {\tt subch\_simple} and the steady phase at the late-stage, respectively. $t_{10}$ represents the theoretical time for the flame to reach 10 km.
}

\begin{ruledtabular}
\footnotesize
\centering
%\resizebox{1\textwidth}{!}{
\begin{tabular}{cccc}
\small 
Name &
$v_{23}$ [$\mbox{km} \ \mbox{s}^{-1}$] &
$v_{100}$ [$\mbox{km} \ \mbox{s}^{-1}$] &
$t_{10}$ [s]
\\
\colrule

{\tt aprox13} & $3.369 \pm 0.016$ & $5.234 \pm 0.027$ & 0.7647 \\
{\tt subch\_full} & $20.732 \pm 0.284$ & $5.411 \pm 0.105$ & 0.9917 \\
{\tt subch\_full\_mod} & $3.468 \pm 0.017$ & $7.975 \pm 0.029$ & 0.4873 \\
{\tt subch\_simple} & $21.095 \pm 0.332$ & $5.521 \pm 0.120$ & 0.8483 \\
{\tt aprox13\_chu} & $3.106 \pm 0.014$ & $4.874 \pm 0.024$ & 0.7912 \\
% {\tt aprox13\_sdc} & $3.039 \pm 0.014$ & $5.201 \pm 0.024$ & $5.764 \pm 0.027$\\
% {\tt subch\_full\_sdc} & $19.811 \pm 0.248$ & $5.181 \pm 0.096$ & $5.228 \pm 0.108$\\
\end{tabular}
%}
\end{ruledtabular}
\end{table*}

Using the fitted parameters, the instantaneous speed of the flame front at different times can be calculated, as shown in Table \ref{Tab:network_instan_vel}. We observe that during the acceleration burst phase at $t = 23$ ms, the difference in the instantaneous speed between {\tt subch\_full} and {\tt subch\_simple} can be as much as 7 times higher compared to the other two simulations. Furthermore, {\tt subch\_full\_mod} accelerates towards the end since it still has a sufficient amount of ${}^{12}$C fuel due to the inefficient reaction flows to ${}^{16}$O. Even though the flame front for {\tt subch\_full\_mod} is expected to surpass {\tt subch\_full} and {\tt subch\_simple} at $t \sim 160$ ms, based on Figure \ref{fig:network_front}, $\dot{e}_{\textrm{nuc}}$ from Figure \ref{fig:network_time_profile} appears to reach its peak at $120$ ms. Therefore, there is no guarantee that the lateral propagating flame for {\tt subch\_full\_mod} would exceed its counterparts.

Assuming that the fitting functions accurately describe the flame propagation, the expected time for the flame to reach a typical neutron star radius of 10 km, $t_{10}$, can be calculated, as shown in Table \ref{Tab:network_instan_vel}. $t_{10}$ serves as a rough prediction for the rise time of XRBs with a pure ${}^{4}$He accretion layer. $t_{10}$ for all models are $\lesssim 1$ sec, whereas $t_{10}$ for {\tt subch\_full} is $\sim 1$ sec, consistent with previous observational studies \citep{galloway:2008}.

\begin{comment}
\begin{figure*}
\centering
\plotone{integration_wallclock.pdf}
\caption{\label{fig:wallclock_time} Wallclock time used per simulation time as function of simulation time. Wallclock time analysis is only conducted for $t > 40$ ms as some simulation models used CPU while others used GPU for $t < 40$ ms.}
\end{figure*}

In order to give an overview of the computational expenses needed for each simulation models, Figure \ref{fig:wallclock_time} shows the wallclock time used per simulation time as a function of simulation time for all models. As some models utilized CPU while others used GPU for $t < 40$ ms, a wallclock time analysis fails to provide an accurate estimate of the computational expenses incurred by each model for during this time frame. Therefore, we have limited the presentation of our data to $t > 40$ ms. The results indicate that the four simulation models, which differ only in terms of their reaction networks, exhibit the expected trends. Specifically, the simulation model with the smallest network, {\tt aprox13}, that has the fewest isotopes and reaction rates requires the least amount of wallclock hours for $1$ ms of simulation time. Conversely, the largest network, {\tt subch\_full}, requires $\sim 3-4$ times the wallclock time needed by {\tt aprox13} to complete the same simulation. Notably, the computational cost for {\tt subch\_full\_mod} gradually increases over time, ultimately exceeding {\tt subch\_full}.

\end{comment}

\subsection{Plasma Screening Routine Comparison}\label{Sec:result_screening}
%aprox13+screen5 aprox13+chugunov2007
% Show gamma, a plots at various times
% u-v enuc phase plot
% T, enuc profiles
% flame speed.

% \begin{figure*}
% \centering
% \plotone{screen_abar_50ms.pdf}
% \caption{\label{fig:screen_abar} Slice plots comparing $\Bar{A}$ for {\tt aprox13} (top panel) and {\tt aprox13\_chu} (bottom panel) at $t = 50$ ms.}
% \end{figure*}

% \begin{figure*}
% \centering
% \plotone{screen_density_50ms.pdf}
% \caption{\label{fig:screen_density} Slice plots comparing density for {\tt aprox13} (top panel) and {\tt aprox13\_chu} (bottom panel) at $t = 50$ ms.}
% \end{figure*}

\begin{figure*}
\centering
\plotone{screen_temp_50ms.pdf}
\caption{\label{fig:screen_temp} Slice plots comparing temperature for {\tt aprox13} (top panel) and {\tt aprox13\_chu} (bottom panel) at $t = 50$ ms.}
\end{figure*}

\begin{figure*}
\centering
\plotone{screen_gamma_50ms.pdf}
\caption{\label{fig:screen_gamma_50ms} Slice plots comparing $\Gamma_e$ for {\tt aprox13} (top panel) and {\tt aprox13\_chu} (bottom panel) at $t = 50$ ms.}
\end{figure*}

\begin{figure*}
\centering
\plotone{screen_gamma_100ms.pdf}
\caption{\label{fig:screen_gamma_100ms} Slice plots comparing $\Gamma_e$ for {\tt aprox13} (top panel) and {\tt aprox13\_chu} (bottom panel) at $t = 100$ ms.}
\end{figure*}

In this section, we present a comparison between the effects of {\tt SCREEN5} and {\tt CHUGUNOV2007} screening routines on the dynamics of the propagating flame in XRBs. Figure \ref{fig:screen_temp} shows the temperature for {\tt aprox13} and {\tt aprox13\_chu} at $t = 50$ ms, which suggests that there are no significant differences in the flame structure between the two models. Figure \ref{fig:screen_gamma_50ms} and \ref{fig:screen_gamma_100ms} show $\Gamma_e$, a measure to determine the approximate screening regime for the flame, at $t = 50$ ms and $t = 100$ ms.

Although a general trend of decreasing $\Gamma_e$ is observed as the flame progress, it is noteworthy that there are more fusion processes involving heavier nuclei at later stages of burning. Assuming triple-$\alpha$ process, it can inferred that $\Gamma \sim 0.05$ and $0.025$ for $t = 50$ ms and $t = 100$ ms. Meanwhile, for oxygen burning at $t = 50$ ms and 100 ms, $\Gamma \sim 0.48$ and $0.32$, respectively. These results indicate that helium burning occurs in the weak screening regime, while heavier nuclei burning processes occur in the intermediate screening regime and gradually transition into the weak regime.

\begin{figure*}
\centering
\plotone{screen_time_profile.pdf}
\caption{\label{fig:screen_profile} Time profiles showing the weighted temperature and energy generation rate of the burning front for {\tt aprox13} and {\tt aprox13\_chu}.}
\end{figure*}

The comparison between {\tt aprox13} and {\tt aprox13\_chu} regarding the weighted temperature and energy generate rate is illustrated in Figure \ref{fig:screen_profile}. This plot illustrates that overall the nuclear energy generation rate of {\tt aprox13} is higher and exhibits a faster increasing rate compared to {\tt aprox13\_chu}. At $t \sim 10$ ms, when the temperature of the two models are approximately equal, the difference in the nuclear energy generation rate indicates that {\tt SCREEN5} provides a stronger screening effect during the initial flame propagation phase compared to {\tt CHUGUNOV2007}.

% In order to assess the impact of plasma screening from the two models, we examine the time period $t \sim 10$ ms, when the temperature of the two models are approximately equal to mitigate the influence of temperature on the nuclear energy generation rate. Even during this time interval, the nuclear energy generation rate of {\tt aprox13} remains slightly higher than that of {\tt aprox13\_chu}, indicating that {\tt SCREEN5} provides a stronger screening effect during the initial flame propagation compared to {\tt CHUGUNOV2007}.

This finding is in agreement with the results from \cite{Chugunov_2007}, where it was shown that the screening effect calculated by \cite{Alastuey_Jancovici:1978} is always greater than \cite{Chugunov_2007} when $\Gamma \lesssim 40$ (See Figure 4 in \cite{Chugunov_2007}). 
%As $\Gamma$ increases, the difference between the two methods gradually decreases (see Figure 4 in \cite{Chugunov_2007}).
It should be emphasized that {\tt SCREEN5} uses calculation routines from \cite{Alastuey_Jancovici:1978} to formulate the intermediate screening function along with the calculations from \cite{Graboske_1973}. Therefore, reactions that fall within the intermediate screening regime experience a slightly higher screening effect from {\tt SCREEN5} compared to {\tt CHUGUNOV2007}. Consequently, the overall energy generation rates in the early time period are slightly higher for {\tt aprox13} compared to {\tt aprox13\_chu}. This phenomenon accelerates the increasing rate of temperature and the energy generation rate in the later stages of burning, amplifying the discrepancy between these two models as time progresses.

%As the system transitions from the weak screening regime in the early stage, the discrepancy in the screening effect becomes more pronounced as we gradually approach the intermediate screening regime, where calculations from \cite{Alastuey_Jancovici:1978} are used.

\begin{figure*}
\centering
\plotone{screen_front.pdf}
\caption{\label{fig:screen_front} Flame front position as a function of time for {\tt aprox13} and {\tt aprox13\_chu}. The dashed lines are the fitted curves using Eq. \ref{Eq:quadratic_fit}.}
\end{figure*}

Figure \ref{fig:screen_front} depicts the evolution of the flame front position over time. It is observed that the flame speed for {\tt aprox13\_chu} is slower compared to {\tt aprox13} due to the smaller value of $\dot{e}_{\textrm{nuc}}$ at later times. However, the influence of the two screening methods on the overall flame dynamics is minimal. After fitting the data using Eq. \ref{Eq:quadratic_fit}, the fitted parameters and the instantaneous velocity at various times are shown in Table \ref{Tab:network_fitted_param}.

% One interesting phenomenon to {\tt aprox13\_chu} is the $u-v$ phase plot for {\tt aprox13\_chu} at $t = 50$ ms and $t = 100$ ms shown in Figure \ref{fig:screen_u_v_enuc}. A similar distribution is observed for {\tt aprox13} at $t = 50$ ms (top left first panel in Figure \ref{fig:network_u_v_enuc}), where the region with high $\dot{e}_{\textrm{nuc}}$ is uniformly distributed around the center. However, high $\dot{e}_{\textrm{nuc}}$ appears to favor regions with positive $u$ as opposed to negative $u$ at $t = 100$ ms, contrary to all the previous models. This phenomenon is suspected to occur due to the comparatively slow instantaneous propagation speed of {\tt aprox13\_chu}. At $t = 120$ ms, $v_{120}$ for {\tt aprox13\_chu} begins to match $v_{100}$ for the other models (see Table \ref{Tab:network_instan_vel}), resulting in a shift of high $\dot{e}_{\textrm{nuc}}$ from the positive $u$ to the negative $u$ at $t=120$ ms.

% The use of different plasma screening methods has minimal impact on computational costs, as evidenced by the negligible differences observed between {\tt aprox13} and {\tt aprox13\_chu} in Figure \ref{fig:wallclock_time}. It is worth noting that screening factor calculations are not expected to be computationally intensive, and thus any variations in computational expenses are likely due to the subsequent effects of minor differences in $\dot{e}_{\textrm{nuc}}$ caused by the different screening methods.

% \begin{figure*}
% \centering
% \plotone{screen_u_v_enuc_chu.pdf}
% \caption{\label{fig:screen_u_v_enuc} Time series $u-v$ phase plots for {\tt aprox13\_chu} at $t = 50$ms (left panel), $t = 100$ ms (middle panel) and $t = 120$ ms (right panel).}
% \end{figure*}

 \subsection{Time Evolution Method Comparison}\label{Sec:result_integration}
%aprox13+Strang, aprox13+sdc, subch_full+Strang, subch_full+SDC

% Show details plots of burning at different times, close-up looks

% \begin{figure*}
% \centering
% \plotone{integration_abar_50ms.pdf}
% \caption{\label{fig:integration_abar} Slice plots comparing $\Bar{A}$ for {\tt aprox13\_sdc} (top panel) and {\tt subch\_full\_sdc} (bottom panel) at $t = 50$ ms.}
% \end{figure*}

\begin{figure*}
\centering
\plotone{integration_enuc_50ms.pdf}
\caption{\label{fig:integration_enuc} Slice plots comparing $\dot{e}_{\textrm{nuc}}$ for {\tt aprox13\_sdc} (top panel) and {\tt subch\_full\_sdc} (bottom panel) at $t = 50$ ms.}
\end{figure*}

Finally, we compare the simplified-SDC scheme with the traditional Strang-splitting. The overall results using the simplified-SDC scheme are nearly identical to the models that employed Strang-splitting.
% Figure \ref{fig:integration_abar} shows the mean molecular weight of the flame for {\tt aprox13\_sdc} and {\tt subch\_full\_sdc}, which used the simplified-SDC scheme. This result is nearly identical to the models that employed Strang-splitting as shown in Figure \ref{fig:network_abar_50ms}. 
One consequence of the strong coupling between advection and reactions is that to make the energy generation plot, we need to derive the nuclear energy release by subtracting off the advection contribution over a timestep.  In regions where there is not much burning, roundoff error can introduce some noise into the energy generation plot, which is seen as the multiple dotted regions in the $\dot{e}_{\textrm{nuc}}$ plot for {\tt aprox13\_sdc} and {\tt subch\_full\_sdc}.  This is simply an artifact of how we do
the derivation.

\begin{figure*}
\centering
\plotone{integration_u_v_enuc_100ms.pdf}
\caption{\label{fig:integration_u_v_enuc} $u-v$ phase plots for {\tt aprox13\_sdc} (left panel) and {\tt subch\_full\_sdc} (right panel) at $t = 100$ms.}
\end{figure*}

The main advantage of the simplified-SDC integration lies in its stronger coupling between the reaction and hydrodynamics. This feature is demonstrated in the $u-v$ phase plot for {\tt aprox13\_sdc} and {\tt subch\_full\_sdc} (Figure \ref{fig:integration_u_v_enuc}). Comparing to the $u-v$ phase plots for {\tt aprox13} and {\tt subch\_full} at $100$ ms (Figure \ref{fig:network_u_v_enuc}), Figure \ref{fig:integration_u_v_enuc} shows a smoother blend in the border of the regions with higher and lower $\dot{e}_{\textrm{nuc}}$, corresponding to negative and positive $u$, respectively. The smoother transition suggests this scheme is capable to resolve the flame in a volatile condition.

\begin{comment}
\begin{figure*}
\centering
\plotone{integration_time_profile.pdf}
\caption{\label{fig:integration_profile} Time profiles showing weighted temperature and energy generation rate of the burning front for {\tt aprox13}, {\tt aprox13\_sdc}, {\tt subch\_full}, and {\tt subch\_full\_sdc}.}
\end{figure*}

\begin{figure*}
\centering
\plotone{integration_front.pdf}
\caption{\label{fig:integration_front} Flame front position as a function of time for {\tt aprox13}, {\tt aprox13\_sdc}, {\tt subch\_full}, and {\tt subch\_full\_sdc}. The dashed lines are the fitted curves using Eq. \ref{Eq:quadratic_fit} and Eq.\ref{Eq:tanh_fit}.}
\end{figure*}
\end{comment}

As shown in \citet{castro_simple_sdc}, in regions where the burning
is vigorous, the simplified-SDC method provides a better solution
than Strang splitting.  However, the XRB flame we simulate here
are not very demanding, so the benefit is minimal.  
\begin{comment}
is not extreme enough for nuclear statistical equilibrium to occur, any difference between simplified-SDC and Strang-splitting schemes is expected to be negligible. This is supported by the temperature and $\dot{e}_{\textrm{nuc}}$ profiles for {\tt aprox13\_sdc} and {\tt subch\_full\_sdc} shown in Figure \ref{fig:integration_profile}, which are almost identical, except for minor variations in $\dot{e}_{\textrm{nuc}}$ for {\tt subch\_full} and {\tt subch\_full\_sdc}. Similarly, the flame front position plots shown in Figure \ref{fig:integration_front} for the simplified-SDC simulations are nearly identical compared to their counterparts, with minor variations (Table \ref{Tab:network_instan_vel}) in the instantaneous flame speed calculated using the the fitting function. 
\end{comment}
As a result, the two simplified-SDC simulations were also more
computationally-expensive than their Strang counterparts.
This differs than the case in \citet{castro_simple_sdc} where 
the simplified-SDC algorithm reduced computational expenses in extreme thermodynamic conditions by mitigating the stiffness of solving reaction equations.

% \subsection{Computational Expenses}

% \begin{figure*}
% \centering
% \plotone{integration_wallclock.pdf}
% \caption{\label{fig:wallclock_time} Wallclock time used per simulation time as function of simulation time. Wallclock time analysis is only conducted for $t > 40$ ms as our simulation models utilized a combination of CPU and GPU for $t < 40$ ms.}
% \end{figure*}

% In order to give an overview of the computational expenses needed for each simulation models, Figure \ref{fig:wallclock_time} shows the wallclock time used per simulation time as a function of simulation time for all models. As some models utilized CPU while others used GPU for $t < 40$ ms, a wallclock time analysis fails to provide an accurate estimate of the computational expenses incurred by each model for during this time frame. Therefore, we have limited the presentation of our data to $t > 40$ ms. The results indicate that the four simulation models, which differ only in terms of their reaction networks, exhibit the expected trends. Specifically, the simulation model with the smallest network, {\tt aprox13}, that has the fewest isotopes and reaction rates requires the least amount of wallclock hours for $1$ ms of simulation time. Conversely, the largest network, {\tt subch\_full}, requires $\sim 3-4$ times the wallclock time needed by {\tt aprox13} to complete the same simulation. Notably, the computational cost for {\tt subch\_full]\_mod} gradually increases over time, ultimately exceeding {\tt subch\_full}. {\tt aprox13} and {\tt aprox13\_chu} only show negligible differences in the computational cost shown in Figure \ref{fig:wallclock_time}. Regarding the simulation models that utilized different integration methods, both {\tt aprox13\_sdc} and {\tt subch\_full\_sdc} models, which employed simplified-SDC, were found to be more computationally expensive compared to their counterparts, {\tt aprox13} and {\tt subch\_full}, which utilized Strang-splitting. Despite the fact that simplified-SDC is expected to reduce computational expenses in extreme thermodynamic conditions by mitigating the stiffness of solving reaction equations, Figure \ref{fig:wallclock_time} suggests that it may increase computational costs under non-extreme conditions.

\section{Summary}

We explored the sensitivity of an XRB flame to the details of the
nuclear physics: size of, and approximations in a reaction network, screening methods, and time-integration strategies. The main
differences observed with reaction network are:

\begin{itemize}
    \item The $(\alpha, \mbox{p})(\mbox{p}, \gamma)$ approximation continues to be an accurate approach in simulating thermonuclear flames in XRBs. Up to $t = 120$ ms, the attained temperature during propagation of the thermonuclear flame is $\lesssim 2.5 \times 10^9$ K, and any minor errors associated with the approximation do not significantly affect the overall flame propagation. This conclusion is supported by the similar profiles observed between {\tt subch\_full} and {\tt subch\_simple} networks.

    \item The ${}^{12}\mbox{C}(\mbox{p}, \gamma) {}^{13}\mbox{N}(\alpha, \mbox{p}){}^{16}\mbox{O}$ rates are critical in accurately modelling nuclear burning, nucleosynthesis, and flame propagation in XRBs. At $T \gtrsim 10^9$ K, these reactions dominate over the triple-$\alpha$ and the slow $\alpha$ capture processes from ${}^{12}$C to ${}^{16}$O. This allows an instant depletion of ${}^{12}$C, leading to a burst of energy once the temperature reaches $\sim 1.3 \times 10^9$ K. This finding is consistent with the work of \cite{Weinberg_2006}, which claims a similar effect at $1.2 \times 10^9$ K. Upon incorporating these rates into the network, we have successfully simulated an accelerating phase for the laterally propagating flame.

    \item Even though there is not a significant difference between {\tt aprox13} and {\tt subch\_full\_mod}, {\tt subch\_full\_mod} demonstrates an increasingly higher $\dot{e}_{\textrm{nuc}}$. Given that the $(\alpha, \mbox{p})(\mbox{p}, \gamma)$ approximation is accurate, the additional rates must have gradually increased the overall $\dot{e}_{\textrm{nuc}}$ in the long run.
    Another possibility that caused the disparity between {\tt subch\_full\_mod} and {\tt aprox13} is the utilization of updated rates from {\sf REACLIB} library in {\tt subch\_full\_mod}, whereas {\tt aprox13} didn't employ the most up-to-date rates. It is plausible that the contemporary adjustments made to the various reaction rates have contributed to the discrepancy between the two networks. However, further investigations are necessary to confirm this hypothesis.

    \item  Among the four reaction networks used to simulate He flame propagation in XRBs, the {\tt subch\_simple} network proved to be the most effective. It is the smallest network that captures the initial acceleration of the propagating flame, which drastically alters the overall flame dynamics.

    %the additional rates such as ${}^{14}\mbox{N}(\alpha,\gamma){}^{18}\mbox{F}(\alpha, \mbox{p}){}^{21}\mbox{Ne}$ and could be the leading cause, although further investigation is needed.
\end{itemize}

Comparing the two screening routines, {\tt SCREEN5} and {\tt CHUNGUNOV2007}, we find that {\tt CHUGUNOV2007} has a slightly weaker screening effect in the weak and intermediate screening regimes. As a consequence, the weaker screening effect from {\tt CHUGUNOV2007} leads to a slightly slower flame compared to the {\tt SCREEN5} model. This result matches our expectations and is in agreement with \cite{Chugunov_2007}.

Finally, we investigated the performance of the simplified-SDC scheme in comparison to the traditional Strang-splitting.  For this problem, since the burning is not very vigorous, there is no strong benefit of using simplified-SDC over Strang-splitting.

Overall, this study gives us confidence that, by using the {\tt subch\_simple} network for our future simulations, we can accurately capture the dynamics of the flame. Our next step is to adapt the current simulation methodology to model a full star flame propagation model. A full star flame propagation simulation allows us to explore how flame dynamics changes subject to the geometric influence, such as the variations in Coriolis force. As the flame encounters the strongest Coriolis force at the pole and the weakest at the equator, its behavior can alter significantly depending on its position. Additionally, a full star simulation provides a more precise estimate of the time required for the flame to engulf the neutron star, which serves as a better approximation of the XRB's rise time.

%In the future, we would extend this model into 3D... run models with mixed hydrogen and helium to see the difference on flame dynamics between a pure helium fuel layer and also study the rapid proton capture processes during XRB...

\begin{acknowledgements}
\castro\ is open-source and freely available at
\url{http://github.com/AMReX-Astro/Castro}.  The work at Stony Brook was supported by DOE/Office
of Nuclear Physics grant DE-FG02-87ER40317. This research was supported by the Exascale Computing 
Project (17-SC-20-SC), a collaborative effort of the U.S. Department of Energy
Office of Science and the National Nuclear Security Administration. This research used
resources of the National Energy Research Scientific Computing Center,
a DOE Office of Science User Facility supported by the Office of
Science of the U.~S.\ Department of Energy under Contract
No.\ DE-AC02-05CH11231.  This research used resources of the Oak Ridge
Leadership Computing Facility at the Oak Ridge National Laboratory,
which is supported by the Office of Science of the U.S. Department of
Energy under Contract No. DE-AC05-00OR22725, awarded through the DOE
INCITE program.  We thank NVIDIA Corporation for the donation of a
Titan X and Titan V GPU through their academic grant program.  This
research has made use of NASA's Astrophysics Data System Bibliographic
Services.
\end{acknowledgements}

\facilities{NERSC, OLCF}

\software{AMReX \citep{amrex_joss},
          Castro \citep{castro},
          GCC (\url{https://gcc.gnu.org/}),
          linux (\url{https://www.kernel.org/}),
          matplotlib (\citealt{Hunter:2007}, \url{http://matplotlib.org/}),
          NumPy \citep{numpy,numpy2},
          python (\url{https://www.python.org/}),
          valgrind \citep{valgrind},
          VODE \citep{vode},
          yt \citep{yt}
          pynucastro \citep{pynucastro, pynucastro2,the_pynucastro_development_2022_7239007}}
% Appendix
%\appendix

%======================================================================
% References
%======================================================================

\bibliographystyle{aasjournal}
\bibliography{ws}

\end{document}

### 2023-05-17 xrb5/slides/defense_ppt.tex
add xrb5 paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Beamer Presentation
% LaTeX Template
% Version 2.0 (March 8, 2022)
%
% This template originates from:
% https://www.LaTeXTemplates.com
%
% Author:
% Vel (vel@latextemplates.com)
%
% License:
% CC BY-NC-SA 4.0 (https://creativecommons.org/licenses/by-nc-sa/4.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[
	11pt, % Set the default font size, options include: 8pt, 9pt, 10pt, 11pt, 12pt, 14pt, 17pt, 20pt
	%t, % Uncomment to vertically align all slide content to the top of the slide, rather than the default centered
	%aspectratio=169, % Uncomment to set the aspect ratio to a 16:9 ratio which matches the aspect ratio of 1080p and 4K screens and projectors
]{beamer}

\graphicspath{{./}{./}} % Specifies where to look for included images (trailing slash required)

\usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule for better rules in tables

\setbeamertemplate{bibliography item}{\insertbiblabel}
\usepackage[style=ieee]{biblatex}
\addbibresource{ws.bib}

%----------------------------------------------------------------------------------------
%	SELECT LAYOUT THEME
%----------------------------------------------------------------------------------------

% Beamer comes with a number of default layout themes which change the colors and layouts of slides. Below is a list of all themes available, uncomment each in turn to see what they look like.

%\usetheme{default}
%\usetheme{AnnArbor}
%\usetheme{Antibes}
%\usetheme{Bergen}
%\usetheme{Berkeley}
%\usetheme{Berlin}
%\usetheme{Boadilla}
%\usetheme{CambridgeUS}
%\usetheme{Copenhagen}
%\usetheme{Darmstadt}
%\usetheme{Dresden}
%\usetheme{Frankfurt}
%\usetheme{Goettingen}
%\usetheme{Hannover}
%\usetheme{Ilmenau}
%\usetheme{JuanLesPins}
%\usetheme{Luebeck}
\usetheme{Madrid}
%\usetheme{Malmoe}
%\usetheme{Marburg}
%\usetheme{Montpellier}
%\usetheme{PaloAlto}
%\usetheme{Pittsburgh}
%\usetheme{Rochester}
%\usetheme{Singapore}
%\usetheme{Szeged}
%\usetheme{Warsaw}

%----------------------------------------------------------------------------------------
%	SELECT COLOR THEME
%----------------------------------------------------------------------------------------

% Beamer comes with a number of color themes that can be applied to any layout theme to change its colors. Uncomment each of these in turn to see how they change the colors of your selected layout theme.

%\usecolortheme{albatross}
%\usecolortheme{beaver}
%\usecolortheme{beetle}
%\usecolortheme{crane}
%\usecolortheme{dolphin}
%\usecolortheme{dove}
%\usecolortheme{fly}
%\usecolortheme{lily}
%\usecolortheme{monarca}
%\usecolortheme{seagull}
%\usecolortheme{seahorse}
%\usecolortheme{spruce}
%\usecolortheme{whale}
%\usecolortheme{wolverine}

%----------------------------------------------------------------------------------------
%	SELECT FONT THEME & FONTS
%----------------------------------------------------------------------------------------

% Beamer comes with several font themes to easily change the fonts used in various parts of the presentation. Review the comments beside each one to decide if you would like to use it. Note that additional options can be specified for several of these font themes, consult the beamer documentation for more information.

\usefonttheme{default} % Typeset using the default sans serif font
%\usefonttheme{serif} % Typeset using the default serif font (make sure a sans font isn't being set as the default font if you use this option!)
%\usefonttheme{structurebold} % Typeset important structure text (titles, headlines, footlines, sidebar, etc) in bold
%\usefonttheme{structureitalicserif} % Typeset important structure text (titles, headlines, footlines, sidebar, etc) in italic serif
%\usefonttheme{structuresmallcapsserif} % Typeset important structure text (titles, headlines, footlines, sidebar, etc) in small caps serif

%------------------------------------------------

%\usepackage{mathptmx} % Use the Times font for serif text
\usepackage{palatino} % Use the Palatino font for serif text

%\usepackage{helvet} % Use the Helvetica font for sans serif text
\usepackage[default]{opensans} % Use the Open Sans font for sans serif text
%\usepackage[default]{FiraSans} % Use the Fira Sans font for sans serif text
%\usepackage[default]{lato} % Use the Lato font for sans serif text

%----------------------------------------------------------------------------------------
%	SELECT INNER THEME
%----------------------------------------------------------------------------------------

% Inner themes change the styling of internal slide elements, for example: bullet points, blocks, bibliography entries, title pages, theorems, etc. Uncomment each theme in turn to see what changes it makes to your presentation.

%\useinnertheme{default}
\useinnertheme{circles}
%\useinnertheme{rectangles}
%\useinnertheme{rounded}
%\useinnertheme{inmargin}

%----------------------------------------------------------------------------------------
%	SELECT OUTER THEME
%----------------------------------------------------------------------------------------

% Outer themes change the overall layout of slides, such as: header and footer lines, sidebars and slide titles. Uncomment each theme in turn to see what changes it makes to your presentation.

%\useoutertheme{default}
%\useoutertheme{infolines}
%\useoutertheme{miniframes}
%\useoutertheme{smoothbars}
%\useoutertheme{sidebar}
%\useoutertheme{split}
%\useoutertheme{shadow}
%\useoutertheme{tree}
%\useoutertheme{smoothtree}

%\setbeamertemplate{footline} % Uncomment this line to remove the footer line in all slides
%\setbeamertemplate{footline}[page number] % Uncomment this line to replace the footer line in all slides with a simple slide count

%\setbeamertemplate{navigation symbols}{} % Uncomment this line to remove the navigation symbols from the bottom of all slides
%----------------------------------------------------------------------------------------
%	PRESENTATION INFORMATION
%----------------------------------------------------------------------------------------

\title[He Flame Simulation in X-ray Bursts]{Sensitivity of He Flames in X-ray Bursts} % The short title in the optional parameter appears at the bottom of every slide, the full title in the main parameter is only on the title page

\subtitle{to Nuclear Physics}%, Plasma Screening Methods, and Time Integration Methods} % Presentation subtitle, remove this command if a subtitle isn't required

\author[Zhi Chen]{Zhi Chen} % Presenter name(s), the optional parameter can contain a shortened version to appear on the bottom of every slide, while the main parameter will appear on the title slide

\institute[SBU]{Stony Brook University \\ \smallskip \textit{}} % Your institution, the optional parameter can be used for the institution shorthand and will appear on the bottom of every slide after author names, while the required parameter is used on the title slide and can include your email address or additional information on separate lines

\date[\today]{Advisor: Mike Zingale \\ \today} % Presentation date or conference/meeting name, the optional parameter can contain a shortened version to appear on the bottom of every slide, while the required parameter value is output to the title slide

%----------------------------------------------------------------------------------------

\begin{document}
%----------------------------------------------------------------------------------------
%	TITLE SLIDE
%----------------------------------------------------------------------------------------

\begin{frame}
	\titlepage % Output the title slide, automatically created using the text entered in the PRESENTATION INFORMATION block above
\end{frame}

%----------------------------------------------------------------------------------------
%	TABLE OF CONTENTS SLIDE
%----------------------------------------------------------------------------------------

% The table of contents outputs the sections and subsections that appear in your presentation, specified with the standard \section and \subsection commands. You may either display all sections and subsections on one slide with \tableofcontents, or display each section at a time on subsequent slides with \tableofcontents[pausesections]. The latter is useful if you want to step through each section and mention what you will discuss.

\begin{frame}
	\frametitle{Presentation Overview} % Slide title, remove this command for no title

	\tableofcontents % Output the table of contents (all sections on one slide)
	%\tableofcontents[pausesections] % Output the table of contents (break sections up across separate slides)
\end{frame}

%----------------------------------------------------------------------------------------
%	PRESENTATION BODY SLIDES
%----------------------------------------------------------------------------------------

\section{Introduction/Motivation}

\begin{frame}
    \frametitle{Research Question}

    \begin{center}
        \large How does reaction networks, plasma screening routines, and time evolution methods affect the simulation of the laterally propagating He flame in X-ray bursts?
    \end{center}

\end{frame}

\begin{frame}
    \frametitle{X-ray Burst Overview}

    \begin{figure}
        \centering
        \includegraphics[width=0.6\linewidth]{x-ray-burst.jpg}
        \caption{Low mass neutron star binaries: Neutron Star accreting mass from its main sequence companion star due to Roche Lobe Overflow. [JINA WEB]}
    \end{figure}

\end{frame}

\begin{frame}
    \frametitle{Motivation}

    \begin{block}{Why?}
        \large What is the motivation behind flame propagation simulations for X-ray bursts? Why is it important?
    \end{block}

    \begin{itemize}
        \item Accretion timescale: $\sim 10^4$ s vs. burst timescale: $\sim 5$ s.
        \item Burst Oscillation Behavior
    \end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Burst Oscillation Behavior}

      \begin{figure}
        \centering
        \includegraphics[width=0.6\linewidth]{x-ray-burst-lightcurve.png}
        \caption{An X-ray burst light curve from 4U 1728–34. \cite{strohmayer_2009}}
    \end{figure}

\end{frame}

\section{Numerical Approach}

% \begin{frame}{Numerical Approach}
%     \begin{center}
%         \Huge {\it Softwares}
%     \end{center}
% \end{frame}

\begin{frame}
    \frametitle{CASTRO}

    \begin{block}{CASTRO}
    An adaptive mesh, astrophysical compressible hydrodynamics simulation code.
    \end{block}

    \begin{block}{General Simulation Domain}
        \begin{itemize}
            \item 2-D $r-z$ cylindrical coordinate system assuming azimuthal symmetry.
            \item Corotating Frame
        \end{itemize}
    \end{block}

\end{frame}

\begin{frame}{CASTRO: continued}

    \begin{block}{Euler's Equations}
    \small
        \begin{subequations}
        \begin{equation}
            \frac{\partial \rho}{\partial t} + \nabla \cdot (\rho \mathbf{U}) = 0
        \end{equation}

        \begin{equation}
            \frac{\partial (\rho \mathbf{U})}{\partial t} + \nabla \cdot (\rho \mathbf{U}\mathbf{U}) + \nabla p = -2\rho \Omega \times \mathbf{U} + \frac{1}{r}\rho w^2 \hat{r} - \frac{1}{r}\rho uw \hat{\theta} + \rho \mathbf{g}% + \rho \mathbf{U} \frac{f}{\Delta t}
        \end{equation}

        \begin{equation}
            \frac{\partial (\rho E)}{\partial t} + \nabla \cdot (\rho \mathbf{U} E + \rho E)  = \nabla \cdot k_{\textrm{th}}\nabla T + \rho \dot{\epsilon} + \rho \mathbf{U} \cdot \mathbf{g}% + \rho \left|\mathbf{U}\right|^2 \frac{f}{\Delta t}
        \end{equation}

        \begin{equation}
            \frac{\partial(\rho X_k)}{\partial t} + \nabla \cdot (\rho \mathbf{U}X_k) =  \rho \dot{\omega}_k
        \end{equation}

        \end{subequations}
        \end{block}

     \begin{block}{Note:}
        \begin{itemize}
            \item $\mathbf{r} = (r, z, \theta)$
            \item $\mathbf{U} = (u, v, w)$
        \end{itemize}
    \end{block}
\end{frame}

\begin{frame}
    \frametitle{Microphysics}

    \begin{block}{Microphysics}
    A software that contains a collection of microphysics routines for stellar explosions.
    \end{block}

    \begin{block}{It contains:}
        \begin{itemize}
            \item Equation of State \cite{timmes_swesty:2000}, of the form, $p=p(\rho,X_k,e)$.
            \item RHS for nuclear reaction networks
            \item Thermal conductivity \cite{Timmes00}
            \item Thermal Neutrino losses\cite{neutrino}
            \item Plasma screening routines
        \end{itemize}
    \end{block}
\end{frame}

% \begin{frame}
% \frametitle{Simulation Domain}

% \begin{itemize}
%     \item 2D cylindrical $r-z$ coordinate.
%     \item $r = [0, 1.8432 \times 10^5]$, $z = [0, 3.072 \times 10^4]$
%     \item Refinement levels: 4, 2, corresponding to $9216\times 1536$ zones at finest level.
% \end{itemize}
% \end{frame}

\subsection{Initial Model}

\begin{frame}{Initial Model}
    \begin{block}{Overview}
    The initial model characterizes the atmospheric structure and initial thermodynamic conditions of the neutron star.
    \end{block}

    \begin{block}{Procedure}
        \begin{enumerate}
            \item Generate two 1D atmospheric vertical HSE profiles, one with temperature perturbation ("hot" model), one without ("cold" model).
            \item Blend the two models as a function of $r$ to give a smooth transition in $r$ between two models.
        \end{enumerate}
    \end{block}

\end{frame}

\begin{frame}[allowframebreaks]
\frametitle{1D Hydrostatic Profile}
    \begin{columns}
        \column{0.4\linewidth}
            \begin{figure}
        		\includegraphics[width=1.0\linewidth]{1dprofile.pdf}
                %\caption{}
    	    \end{figure}

    	\column{0.6\linewidth}
    	    \begin{block}{Tanh transition equations:}
    	       \begin{equation}
                    \scriptsize T(z) = T_* + \frac{1}{2}\left(T_{\mathrm{hi}} - T_*\right)\left[1 + \tanh{\left( \frac{2\tilde{z}}{\delta_{\mathrm{atm}}} \right)}\right]
                \end{equation}

                \begin{equation}
                    \scriptsize X_k(z) = X_{k_{\mathrm{fuel}}} + \frac{1}{2}\left(X_{k_{\mathrm{base}}} - X_{k_\mathrm{fuel}}\right)\left[1 + \tanh{\left( \frac{2\tilde{z}}{\delta_{\mathrm{atm}}} \right)}\right]
                \end{equation}
    	    \end{block}

            \begin{block}{HSE Equations:}
                \begin{equation}
                    \scriptsize p_i = p_{i-1} + \frac{1}{2}\Delta z( \rho_i + \rho_{i-1})\mathbf{g} \cdot \hat{\mathbf{z}}
                \end{equation}

                \begin{equation}
                    \scriptsize p_i = p_{i+1} + \frac{1}{2}\Delta z( \rho_i + \rho_{i+1})\mathbf{g} \cdot \hat{\mathbf{z}}
                \end{equation}
            \end{block}

    \end{columns}

    % \begin{block}{Summary}
    % \begin{itemize}
    %     \item an isothermal base with $T = T_*$ when $0 < z < H_*$
    %     \item a transition phase with an isothermal profile that blends the base and the atmosphere when $H_* < z < H_* + 3\delta_{\textrm{atm}}$
    %     \item an isentropic atmosphere for flame propagation when $z > H_* + 3\delta_{\textrm{atm}}$, $T > T_{lo}$, and $\rho > \rho_{\textrm{cutoff}}$.
    %     \item an isothermal low temperature buffer region when $T < T_{lo}$
    %     \item an isothermal and constant low density buffer region when $\rho < \rho_{\textrm{cutoff}}$
    % \end{itemize}
    % \end{block}

    \begin{columns}
        \column{0.65\linewidth}
            \begin{figure}
                \includegraphics[height=0.85\textheight]{initial_profile.pdf}
                % \caption{This figure shows the initial 1D profile of the model. Dashed line represents the cold model while the solid line represents the hot model.}
            \end{figure}

        \column{0.35\linewidth}
            \small Figure: Initial 1D profile of the model. Dashed line represents the hot model while the solid line represents the cold model.
    \end{columns}
\end{frame}

\begin{frame}[allowframebreaks]
\frametitle{2D Profile Blending}

\begin{block}{(r,z) Blending}
    % \begin{equation}
    %     \mathcal{Q}(r,z) \in \{\rho, T, X_k\}
    % \end{equation}

    \begin{equation}
        \scriptsize \mathcal{Q}(r, z) = \left\{\begin{array}{lr}
                 \mathcal{Q}_{\mathrm{hot}}(z)  & \quad r < r_{\mathrm{pert}} \\

            f(r)\mathcal{Q}_{\mathrm{hot}}(z) + \left[1 - f(r)\right]\mathcal{Q}_{\mathrm{cool}}(z) & \quad r_{\mathrm{pert}} \leq r \leq r_{\mathrm{pert}} + \delta_{\mathrm{blend}}\\

            \mathcal{Q}_{\mathrm{cool}}(z)  & \quad r > r_{\mathrm{pert}} + \delta_{\mathrm{blend}}
                \end{array}\right\}, \quad\mathcal{Q} \in \{p, \rho, X_k\}
    \end{equation}

    where

    \begin{equation}
        f(r) = 1 - \frac{r - r_{\mathrm{pert}}}{\delta_{\mathrm{blend}}}
    \end{equation}

    for $r_{\mathrm{pert}} = 4.096 \times 10^4$ cm and $\delta_{\mathrm{blend}} = 2048$ cm.
\end{block}
    \begin{itemize}
        \item   $T(r,z)$ and $\rho e(r,z)$ are calculated via EOS after blending.
    \end{itemize}

    \begin{figure}
        \centering
        \includegraphics[width=1\linewidth]{init_rho.pdf}
        \includegraphics[width=1\linewidth]{init_temp.pdf}
        \caption{Initial temperature and density profile.}
    \end{figure}

\end{frame}

\begin{frame}{Boundary Conditions}

    \begin{block}{Boundary Conditions:}
        \begin{itemize}
            \item {\it Left}$(r=0 \textrm{cm})$: reflecting or symmetry boundary condition.
            \item {\it Right}$(r=1.8432\times 10^5 \textrm{cm})$: zero-gradient outflow boundary condition
            \item {\it Bottom}$(z=0 \textrm{cm})$: hydrostatic boundary condition with reflecting velocity.
            \item {\it Top}$(z=3.072\times10^4 \textrm{cm})$: ambient-fill boundary condition with zero-gradient velocity outflow.
        \end{itemize}
    \end{block}
\end{frame}

% \begin{frame}
% \frametitle{Initial Model Parameters}
%     \begin{table}
%     \begin{tabular}{cc}
%     \toprule
%     \textbf{Input Parameters} &
%     \textbf{Values}
%     \\ 
%     \midrule
%     $f$ & 1000 Hz \\
%     $T_*$    & $10^8$ K \\
%     $T_{\mathrm{hi}}$  & $2 \times 10^8$ K \\
%     $T_{\mathrm{lo}}$  & $8 \times 10^6$ K \\
%     $\delta T$  & $1.2 \times 10^9 $ K \\
%     $r_{\mathrm{pert}}$  & $4.096 \times 10^4$ cm \\
%     $\delta_{\mathrm{blend}}$ & 2048 cm \\
%     $\rho_{\mathrm{base}}$ & $3.43 \times 10^6 $ $\mathrm{g \ cm^{-3}}$ \\
%     $H_*$  & 2000 cm \\
%     $\delta_{\mathrm{atm}}$ & 50 cm \\
%     $X_{k_{\mathrm{base}}}({}^{56} \mathrm{Ni})$ & 1.0 \\
%     $X_{k_{\mathrm{fuel}}}({}^{4} \mathrm{He})$ & 1.0 \\
%     $\rho_{\mathrm{cutoff}}$ & $10^{-4}$  $\mathrm{g \ cm^{-3}}$ \\
%     $\mathbf{g}$ & $-1.5\times 10^{14}$ $\mathrm{g \ s^{-2}}$ $\hat{z}$\\
%     \bottomrule

%     \end{tabular}
%     %\caption{\small This table shows the input parameters for constructing the initial model.}
%     \end{table}
% \end{frame}

% \begin{frame}{Results}
%     \begin{center}
%         \Huge {\it Discussions / Results}
%     \end{center}
% \end{frame}

\section{Nuclear Reaction Networks}

\begin{frame}{Networks}
    \begin{center}
        \Huge {\it Effect of different networks}
    \end{center}
\end{frame}

\subsection{Overview of reaction networks}

\begin{frame}
\frametitle{Network: {\tt aprox13}}
    % \begin{columns}
    %     \column{0.7\linewidth}
    %         \begin{figure}
    %             \centering
    %             \includegraphics[width=1\linewidth]{aprox13.pdf}
    %         \end{figure}

    %     \column{0.3\linewidth}
    %         Figure: Overview of {\tt aprox13}, rates are integrated with $\rho = 10^6$ g $\mathrm{cm}^{-3}$ and $T = 2.0 \times 10^9$ K.
    % \end{columns}
    \begin{figure}
        \centering
        \includegraphics[width=1\linewidth]{aprox13.pdf}
        \caption{\scriptsize Overview of {\tt aprox13}, rates are integrated with $\rho = 10^6$ g $\mathrm{cm}^{-3}$ and $T = 2.0 \times 10^9$ K}
    \end{figure}
    \begin{block}{Feature}
        \begin{itemize}
            \item $(\alpha, \mbox{p})(\mbox{p}, \gamma)$ approximation
            \item 13 isotopes, 31 rates
        \end{itemize}
    \end{block}

\end{frame}

\begin{frame}
\frametitle{Network: {\tt subch full}}
    \begin{figure}
        \centering
        \includegraphics[width=1\linewidth]{subch_full.pdf}
        \caption{\scriptsize Overview of {\tt subch full}, integrated with $\rho = 10^6$ g $\mathrm{cm}^{-3}$ and $T = 2.0 \times 10^9$ K}
    \end{figure}

    \begin{block}{Feature}
        \begin{itemize}
            \item \scriptsize No $(\alpha, \mbox{p})(\mbox{p}, \gamma)$ approximation
            \item \scriptsize Additional rates, such as ${}^{12}\mbox{C} ({}^{12}\mbox{C}, p){}^{23}\mbox{Na}$ to give complete representation on carbon and oxygen burning.
            \item \scriptsize Additional rates, ${}^{14}\mbox{N}(\alpha,\gamma){}^{18}\mbox{F}(\alpha, \mbox{p}){}^{21}\mbox{Ne}$ and ${}^{12}\mbox{C}(\mbox{p}, \gamma) {}^{13}\mbox{N}(\alpha, \mbox{p}){}^{16}\mbox{O}$, discussed in Shen \& Bildsten 2009 and Weinberg 2006 \cite{Shen_2009,Weinberg_2006}.
            \item 28 isotopes, 107 rates
        \end{itemize}
    \end{block}
\end{frame}

\begin{frame}
\frametitle{Network: {\tt subch full mod}}
    \begin{figure}
        \centering
        \includegraphics[width=1\linewidth]{subch_full_mod.pdf}
        \caption{Overview of {\tt subch full mod}, integrated with $\rho = 10^6$ g $\mathrm{cm}^{-3}$ and $T = 2.0 \times 10^9$ K}
    \end{figure}
    \begin{block}{Feature}
        \begin{itemize}
            \item Identical to {\tt subch full} but ${}^{12}\mbox{C}(\mbox{p}, \gamma) {}^{13}\mbox{N}(\alpha, \mbox{p}){}^{16}\mbox{O}$ and its reverse rate are disabled.
            \item 27 isotopes, 103 rates
        \end{itemize}
    \end{block}
\end{frame}

\begin{frame}
\frametitle{Network: {\tt subch simple}}
\begin{figure}
    \centering
    \includegraphics[width=1\linewidth]{subch_simple.pdf}
    \caption{\scriptsize Overview of {\tt subch simple}, rates are integrated with $\rho = 10^6$ g $\mathrm{cm}^{-3}$ and $T = 2.0 \times 10^9$ K}
\end{figure}
\begin{block}{Feature}
    \begin{itemize}
        \item \footnotesize $(\alpha, \mbox{p})(\mbox{p}, \gamma)$ approximation for heavy isotopes
        \item \footnotesize The reverse rates of all ${}^{12}\mbox{C} + {}^{12}\mbox{C}$,  ${}^{16}\mbox{O} + {}^{16}\mbox{O}$, and ${}^{16}\mbox{O} + {}^{12}\mbox{C}$ are removed.
        \item \footnotesize Forward and reverse rates of ${}^{12}\mbox{C} + {}^{20}\mbox{Ne}$ , ${}^{23}\mbox{Na}(\alpha, \gamma){}^{27}\mbox{Al}$, and ${}^{27}\mbox{Al}(\alpha, \gamma){}^{31}\mbox{P}$ are removed
        \item 22 isotopes, 57 rates
    \end{itemize}
\end{block}
\end{frame}

\begin{frame}{Network: Models}
    % \begin{center}
    %     \LARGE Plasma Screening
    % \end{center}

    \begin{table}
        \resizebox{1\linewidth}{!}{
        \begin{tabular}{cccc}
        \toprule
        Name &
        Network &
        Integration &
        Screening\\ 
        \midrule

        {\tt aprox13} & {\tt aprox13} & strang-splitting & {\tt SCREEN5} \\
        {\tt subch full} & {\tt subch full} & strang-splitting & {\tt SCREEN5} \\
        {\tt subch full mod} & {\tt subch full mod} & strang-splitting & {\tt SCREEN5} \\
        {\tt subch simple} & {\tt subch simple} & strang-splitting & {\tt SCREEN5} \\
        % {\tt aprox13 sdc} & {\tt aprox13} & simplified SDC & {\tt SCREEN5} \\
        % {\tt subch full sdc} & {\tt subch full} & simplified SDC & {\tt SCREEN5} \\
        % {\tt aprox13 chu} & {\tt aprox13} & strang-splitting & {\tt CHUGUNOV2007} \\
        \bottomrule
        \end{tabular}}
        \caption{This table shows the various settings used for each simulation.}
    \end{table}

\end{frame}

\subsection{Simulation Results}

\begin{frame}
\frametitle{Network: $\Bar{A}$ Comparison}
    \begin{figure}
        \centering
        \includegraphics[width=1\linewidth]{network_abar_50ms.pdf}
        \caption{\scriptsize Slice plots of the flame propagation comparing average atomic weight, $\Bar{A}$, for {\tt aprox13} (top panel), {\tt subch full} (second panel from top), {\tt subch full mod} (third panel), and {\tt subch simple} (last panel) at 50ms.}
    \end{figure}
\end{frame}

% \begin{frame}
% \frametitle{Network: $\dot{e}_{nuc}$ Comparison}
%     \begin{figure}
%         \centering
%         \includegraphics[width=1\linewidth]{network_enuc_50ms.pdf}
%         \caption{This figure shows 4 slice plots of the flame propagation comparing specific energy generation rate, $\dot{e}_{nuc}$, for {\tt aprox13} (top panel), {\tt subch full} (second panel from top), {\tt subch full mod} (third panel), and {\tt subch simple} (last panel) at 50ms.}
%     \end{figure}
% \end{frame}

\begin{frame}
\frametitle{Network: Weighted $T$ and $\dot{e}_{nuc}$ Time Profiles}
    \begin{figure}
        \centering
        \includegraphics[width=1\linewidth]{network_time_profile.pdf}
        \caption{{\it Left}: The weighted temperature time profile. {\it Right}: The weighted specific energy generation rate time profile.}
    \end{figure}
\end{frame}

\begin{frame}[allowframebreaks]
\frametitle{Network: Mass fraction Comparison}
    \begin{figure}
        \centering
        \includegraphics[width=0.8\linewidth]{network_aprox13_nuc_frac.pdf}
        \caption{This figure shows the mass fractions of ${}^{12}$C, ${}^{16}$O, ${}^{20}$Ne,  ${}^{24}$Mg, ${}^{28}$Si, and ${}^{32}$S for {\tt aprox13} at 20 ms.}
    \end{figure}

    \begin{figure}
        \centering
        \includegraphics[width=0.8\linewidth]{network_subch_nuc_frac.pdf}
        \caption{This figure shows the mass fractions of ${}^{12}$C, ${}^{16}$O, ${}^{20}$Ne,  ${}^{24}$Mg, ${}^{28}$Si, and ${}^{32}$S for {\tt subch full} at 20 ms.}
    \end{figure}

    \begin{figure}
        \centering
        \includegraphics[width=0.8\linewidth]{network_subch_mod_nuc_frac.pdf}
        \caption{This figure shows the mass fractions of ${}^{12}$C, ${}^{16}$O, ${}^{20}$Ne, ${}^{24}$Mg, ${}^{28}$Si, and ${}^{32}$S for {\tt subch full mod} at 20 ms.}
    \end{figure}

\end{frame}

% \begin{frame}
% \frametitle{\small Network: Weighted $T$ and $\dot{e}_{nuc}$ Time Profiles (Revisit)}
%     \begin{figure}
%         \centering
%         \includegraphics[width=1\linewidth]{network_time_profile.pdf}
%         \caption{{\it Left}: The weighted temperature time profile. {\it Right}: The weighted specific energy generation rate time profile.}
%     \end{figure}
% \end{frame}

\begin{frame}{Network: Front Position vs. Time}
%\begin{columns}
    %\column{1\linewidth}
    \begin{figure}
        \centering
        \includegraphics[width=1\linewidth]{network_front.pdf}
        \caption{The flame front position as a function time for {\tt aprox13}, {\tt subch full}, {\tt subch full mod}, and {\tt subch simple}. {\it Solid} Lines: Data. {\it Dashed} lines: fit.}
    \end{figure}

    % \column{0.2\linewidth}
    % \footnotesize The flame front position as a function time for {\tt aprox13}, {\tt subch full}, {\tt subch full mod}, and {\tt subch simple}. {\it Solid} Lines: Data, {\it Dashed} lines: fit.
%\end{columns}

\end{frame}

\begin{frame}[allowframebreaks]
\frametitle{Network: Flame Speed Comparison}

    \begin{block}{ Fitting Functions}
        \begin{subequations}

        \begin{equation}
             r(t) = \frac{1}{2}a_0 t^2 + v_0 t + r_0
        \end{equation}

        \begin{equation}
             r(t) = A\tanh{\left(\frac{t}{B} + C\right)} + \frac{1}{2}a_0 t^2 + v_0 t + r_0
        \end{equation}
        \end{subequations}
    \end{block}

    \begin{table}
        \resizebox{1\linewidth}{!}{
        \begin{tabular}{ccccccc}
        \toprule
        Name &
        $a_0$ [$\mbox{km} \ \mbox{s}^{-2}$] &
        $v_0$ [$\mbox{km} \ \mbox{s}^{-1}$] &
        $r_0$ [km] &
        A [$\mbox{km}$]&
        B [s]&
        C \\ 
        \midrule
        {\tt aprox13} & $24.22 \pm 0.23$ & $2.812 \pm 0.015$ & $0.7680\pm 0.0004 $ & N/A & N/A & N/A\\

        {\tt subch full} & $9.22 \pm 0.82$ & $4.489 \pm 0.067$ & $0.8646 \pm 0.0014$ & $0.150 \pm 0.001$ & $0.0093 \pm 0.0001$ & $-2.551 \pm 0.035$ \\

        {\tt subch full mod} & $58.53 \pm 0.24$ & $2.122 \pm 0.016$ & $0.7773 \pm 0.0004$ & N/A & N/A & N/A \\

        {\tt subch simple} & $15.60 \pm 0.93$ & $3.961 \pm 0.076$ & $0.8745 \pm 0.0016$ & $0.153 \pm 0.002$ & $0.0091 \pm 0.0001$ & $-2.575 \pm 0.040$\\

        %{\tt aprox13 chu} & $22.96 \pm 0.20$ & $2.578 \pm 0.0133$ & $0.7728 \pm 0.0004$ & N/A & N/A & N/A\\

        %{\tt aprox13 sdc} & $28.09 \pm 0.20$ & $2.392 \pm 0.013$ & $0.7836\pm 0.0004 $ & N/A & N/A & N/A\\

        %{\tt subch full sdc} & $2.34 \pm 0.74$ & $4.947 \pm 0.061$ & $0.8601 \pm 0.0013$ & $0.146 \pm 0.001$ & $0.0098 \pm 0.0001$ & $-2.424 \pm 0.030$ \\
        \bottomrule
        \end{tabular}}
        \caption{\scriptsize This table shows the fitted parameters of the fitting functions. The fitting function is applied for $t > 8$ ms.}
    \end{table}

    \begin{table}
        \resizebox{1\linewidth}{!}{
        \begin{tabular}{cccc}
        \toprule
        Name &
        $v_{23}$ [$\mbox{km} \ \mbox{s}^{-1}$] &
        $v_{100}$ [$\mbox{km} \ \mbox{s}^{-1}$] &
        $v_{120}$ [$\mbox{km} \ \mbox{s}^{-1}$] \\
        \midrule
        {\tt aprox13} & $3.369 \pm 0.016$ & $5.234 \pm 0.027$ & $5.718 \pm 0.031$\\
        {\tt subch full} & $20.732 \pm 0.284$ & $5.411 \pm 0.105$ & $5.595 \pm 0.119$\\
        {\tt subch full mod} & $3.468 \pm 0.017$ & $7.975 \pm 0.029$ & $9.146 \pm 0.033$\\
        {\tt subch simple} & $21.095 \pm 0.332$ & $5.521 \pm 0.120$ & $5.833 \pm 0.135$\\
        %{\tt aprox13 chu} & $3.106 \pm 0.014$ & $4.874 \pm 0.024$ & $5.333 \pm 0.027$ \\
        %{\tt aprox13 sdc} & $3.039 \pm 0.014$ & $5.201 \pm 0.024$ & $5.764 \pm 0.027$\\
        %{\tt subch full sdc} & $19.811 \pm 0.248$ & $5.181 \pm 0.096$ & $5.228 \pm 0.108$\\
        \bottomrule
        \end{tabular}}
        \caption{This table shows the instantaneous flame propagation speed at $t = 23$ ms, $t = 100$ ms and $t = 120$ ms. $t = 23$ ms represents the acceleration for {\tt subch full} and {\tt subch simple}. $t = 100$ ms and $t = 120$ ms represent the steady phase at the late-stage.}
    \end{table}

\end{frame}

\section{Plasma Screening Routines}
\subsection{Overview of Screening Routines}

\begin{frame}{Plasma Screening}
    \begin{center}
        \Huge {\it Effect of different plasma screening routines}
    \end{center}
\end{frame}

\begin{frame}{Plasma Screening Introduction}
    \begin{block}{What is plasma screening?}
        \begin{itemize}
            \item Plasma screening enhances the nuclear reaction rate by reducing the Coulomb barrier that reacting nuclei must overcome. 
            \item The overall screening regime depends on the Coulomb coupling parameter, $\Gamma \propto \frac{((Z_1 +Z_2))^{1/3}}{Z_1Z_2} \Gamma_{e}$ with $\Gamma_e \propto \frac{n_e^{1/3}}{T}$ 
        \end{itemize}
    \end{block}

    \begin{block}{Screening Regimes}
    \begin{itemize}
        \item Weak Screening: $\Gamma \ll 1$
        \item Strong Screening: $\Gamma \gtrsim 1$
    \end{itemize}
    \end{block}

    \begin{equation}\label{Eq:screening}
        R_{scr} = F_{scr} R_{th}
    \end{equation}
\end{frame}

\begin{frame}
\frametitle{{\tt SCREEN5} vs. {\tt CHUGUNOV2007}}

    \begin{block}{\tt SCREEN5}
        \begin{itemize}
            \item {\it Weak} Regime $(\Gamma < 0.3)$: Follows Graboske 1973\cite{Graboske_1973}. Assumes interacting nuclei are separated by zero distance.
            \item {\it Strong} Regime $(\Gamma > 0.8)$: Follows Alastuey \& Jancovici (AJ) 1978 \cite{Alastuey_Jancovici:1978}. Assumes a quadratic dependence of the separation distance between interacting nuclei.
            \item {\it Intermediate} Regime $(0.3 < \Gamma < 0.8)$: A weighted value between both weak and strong regime calculations.
        \end{itemize}
    \end{block}

    \begin{block}{\tt CHUGUNOV2007}
        \begin{itemize}
            \item Follows Chugunov 2007 \cite{Chugunov_2007}.
            %\item Assumes a higher order polynomial dependence of the separation distance between interacting nuclei.
            \item Assumes WKB Coulomb barrier penetration through the radial mean-field potential.
            \item A single expression for all screening regimes.
        \end{itemize}
    \end{block}
\end{frame}

\begin{frame}{Plasma Screening: Models}
    % \begin{center}
    %     \LARGE Plasma Screening
    % \end{center}

    \begin{table}
        \resizebox{1\linewidth}{!}{
        \begin{tabular}{cccc}
        \toprule
        Name &
        Network &
        Integration &
        Screening\\ 
        \midrule

        {\tt aprox13} & {\tt aprox13} & strang-splitting & {\tt SCREEN5} \\
        % {\tt subch full} & {\tt subch full} & strang-splitting & {\tt SCREEN5} \\
        % {\tt subch full mod} & {\tt subch full mod} & strang-splitting & {\tt SCREEN5} \\
        % {\tt subch simple} & {\tt subch simple} & strang-splitting & {\tt SCREEN5} \\
        % {\tt aprox13 sdc} & {\tt aprox13} & simplified SDC & {\tt SCREEN5} \\
        % {\tt subch full sdc} & {\tt subch full} & simplified SDC & {\tt SCREEN5} \\
        {\tt aprox13 chu} & {\tt aprox13} & strang-splitting & {\tt CHUGUNOV2007} \\
        \bottomrule
        \end{tabular}}
        \caption{This table shows the various settings used for each simulation.}
    \end{table}

\end{frame}

\subsection{Simulation Results}

\begin{frame}{Screening: $\Bar{A}$ and $T$ Comparison}
    \begin{figure}
        \centering
        \includegraphics[width=1\linewidth]{screen_abar_50ms.pdf}
        \caption{{\it Top}: $\Bar{A}$ for {\tt aprox13}. {\it Bot}: $\Bar{A}$ for {\tt aprox13 chu}}
    \end{figure}

    \begin{figure}
        \centering
        \includegraphics[width=1\linewidth]{screen_temp_50ms.pdf}
        \caption{{\it Top}: $T$ for {\tt aprox13}. {\it Bot}: $T$ for {\tt aprox13 chu}}
    \end{figure}
\end{frame}

\begin{frame}{Screening: $\Gamma$ Comparison}
\begin{columns}
    \column{0.75\linewidth}

    \begin{figure}
        \centering
        \includegraphics[width=1\linewidth]{screen_gamma_50ms.pdf}
        \caption{\scriptsize {\it Top}: $\Gamma_e$ for {\tt aprox13}. {\it Bot}: $\Gamma_e$ for {\tt aprox13 chu}. Current time is $t = 50$ ms.}
    \end{figure}

    \begin{figure}
        \centering
        \includegraphics[width=1\linewidth]{screen_gamma_100ms.pdf}
        \caption{\scriptsize {\it Top}: $\Gamma_e$ for {\tt aprox13}. {\it Bot}: $\Gamma_e$ for {\tt aprox13 chu}. Current time is $t = 100$ ms.}
    \end{figure}

    \column{0.25\linewidth}
    \begin{block}{\scriptsize $\Gamma$ estimation}
        \begin{itemize}
            \item \scriptsize Triple-$\alpha$: $\Gamma \sim 0.05$ for $t = 50$ ms, $\Gamma \sim 0.025$ for $t = 100$ ms.
            \item \scriptsize ${}^{16}$O + ${}^{16}$O:             $\Gamma \sim 0.48$ for $t = 50$ ms, $\Gamma \sim 0.32$ for $t = 100$ ms.
        \end{itemize}
    \end{block}
    \end{columns}
\end{frame}

\begin{frame}{Screening: Weighted $T$ and $\dot{e}_{nuc}$ Time Profile}

    \begin{figure}
        \centering
        \includegraphics[width=1\linewidth]{screen_time_profile.pdf}
        \caption{{\it Left}: Weighted temperature time profile. {\it Right}: Weighted specific nuclear energy generation rate time profile.}
    \end{figure}

\end{frame}

\begin{frame}{Screening: Chugunov 2007 Comparison}
    \begin{figure}
        \centering
        \includegraphics[width=0.33\linewidth]{chugunov2007.png}
    \end{figure}

\scriptsize Figure 4 in Chugunov 2007 \cite{Chugunov_2007}. The vertical axis represents the screening strength. The horizontal axis is a temperature parameter.

\end{frame}

\begin{frame}{Screening: Flame Front Position vs. Time}

    \begin{figure}
        \centering
        \includegraphics[width=1\linewidth]{screen_front.pdf}
        \caption{This figure shows the flame front position as a function of time for {\tt aprox13} and {\tt aprox13 chu}.}
    \end{figure}
\end{frame}

\begin{frame}[allowframebreaks]
\frametitle{Screening: Flame Speed Comparison}

    \begin{block}{Fitting Functions}
        \begin{equation}
            r(t) = \frac{1}{2}a_0 t^2 + v_0 t + r_0
        \end{equation}
    \end{block}

    \begin{table}
        \resizebox{1\linewidth}{!}{
        \begin{tabular}{cccc}
        \toprule
        Name &
        $a_0$ [$\mbox{km} \ \mbox{s}^{-2}$] &
        $v_0$ [$\mbox{km} \ \mbox{s}^{-1}$] &
        $r_0$ [km] \\
        \midrule
        {\tt aprox13} & $24.22 \pm 0.23$ & $2.812 \pm 0.015$ & $0.7680\pm 0.0004 $\\

        {\tt aprox13 chu} & $22.96 \pm 0.20$ & $2.578 \pm 0.0133$ & $0.7728 \pm 0.0004$\\
        \bottomrule
        \end{tabular}}
        \caption{This table shows the fitted parameters of the fitting functions. The fitting function is applied for $t > 8$ ms.}
    \end{table}

    \begin{table}
        \resizebox{1\linewidth}{!}{
        \begin{tabular}{cccc}
        \toprule
        Name &
        $v_{23}$ [$\mbox{km} \ \mbox{s}^{-1}$] &
        $v_{100}$ [$\mbox{km} \ \mbox{s}^{-1}$] &
        $v_{120}$ [$\mbox{km} \ \mbox{s}^{-1}$] \\
        \midrule
        {\tt aprox13} & $3.369 \pm 0.016$ & $5.234 \pm 0.027$ & $5.718 \pm 0.031$\\
        {\tt aprox13 chu} & $3.106 \pm 0.014$ & $4.874 \pm 0.024$ & $5.333 \pm 0.027$ \\
        \bottomrule
        \end{tabular}}
        \caption{This table shows the instantaneous flame propagation speed at $t = 23$ ms, $t = 100$ ms and $t = 120$ ms. $t = 23$ ms represents the acceleration for {\tt subch full} and {\tt subch simple}. $t = 100$ ms and $t = 120$ ms represent the steady phase at the late-stage.}
    \end{table}

\end{frame}

\section{Time Integration Methods}
\subsection{Overview Time Integration Methods}

\begin{frame}{Integration}
    \begin{center}
        \Huge {\it Effect of different time integration methods}
    \end{center}
\end{frame}

% \begin{frame}{Introduction to Time Evolution Methods}
%     \begin{block}
%         Source terms for Euler's equation can be decomposed into hydrodynamic sources and reactive sources.
%     \end{block}
% \end{frame}

\begin{frame}{strang-splitting vs. simplified-SDC}

    \begin{block}{Strang-splitting}
        \begin{itemize}
            \item Consider hydrodynamics and nuclear reactions to be independent processes.
            \item $\mathcal{U}^{n+1} = \mathcal{R}_{\Delta t/2} \mathcal{A}_{\Delta t} \mathcal{R}_{\Delta t/2} \mathcal{U}^n$
            \item Indirect coupling between hydro-update and reaction.
            \item Achieves second-order accuracy in time.

        \end{itemize}
    \end{block}

    \begin{block}{simplified-SDC}
        \begin{itemize}
            \item Uses iterations (2) to construct higher-order accuracy solutions.
            \item Direct Coupling between hydro-update and reaction.
            \item Achieves second-order accuracy in time.
        \end{itemize}
    \end{block}

\end{frame}

\begin{frame}{Time Integration Methods: Models}

    % \begin{center}
    %     \LARGE Time Evolution Methods
    % \end{center}

    \begin{table}
        \resizebox{1\linewidth}{!}{
        \begin{tabular}{cccc}
        \toprule
        Name &
        Network &
        Integration &
        Screening\\ 
        \midrule

        {\tt aprox13} & {\tt aprox13} & strang-splitting & {\tt SCREEN5} \\
        {\tt subch full} & {\tt subch full} & strang-splitting & {\tt SCREEN5} \\
        % {\tt subch full mod} & {\tt subch full mod} & strang-splitting & {\tt SCREEN5} \\
        % {\tt subch simple} & {\tt subch simple} & strang-splitting & {\tt SCREEN5} \\
        {\tt aprox13 sdc} & {\tt aprox13} & simplified SDC & {\tt SCREEN5} \\
        {\tt subch full sdc} & {\tt subch full} & simplified SDC & {\tt SCREEN5} \\
        % {\tt aprox13 chu} & {\tt aprox13} & strang-splitting & {\tt CHUGUNOV2007} \\
        \bottomrule
        \end{tabular}}
        \caption{This table shows the various settings used for each simulation.}
    \end{table}

\end{frame}

\subsection{Simulation Results}

\begin{frame}{Integration: $\dot{e}_{nuc}$ Comparison}
    \begin{figure}
        \centering
        \includegraphics[width=1\linewidth]{pres_integration_enuc_50ms.pdf}
        \caption{{\it First (Top)}: $\dot{e}_{nuc}$ for {\tt aprox13}. {\it Second}: $\dot{e}_{nuc}$ for {\tt aprox13 sdc}. {\it Third}: $\dot{e}_{nuc}$ for {\tt subch full}. {\it Fourth (Bot)}: $\dot{e}_{nuc}$ for {\tt subch full sdc}.}
    \end{figure}
\end{frame}

\begin{frame}{Integration: Weighted $T$ and $\dot{e}_{nuc}$ Time Profiles}

    \begin{figure}
        \centering
        \includegraphics[width=0.95\linewidth]{integration_time_profile.pdf}
        \caption{\scriptsize {\it Left}: weighted temperature for {\tt aprox13}, {\tt aprox13 sdc}, {\tt subch full}, and {\tt subch full sdc}. {\it Right}: specific energy generation rate for {\tt aprox13}, {\tt aprox13 sdc}, {\tt subch full}, and {\tt subch full sdc}.}
    \end{figure}
\end{frame}

\begin{frame}{Integration: Front Position vs. Time}

    \begin{figure}
        \centering
        \includegraphics[width=1\linewidth]{integration_front.pdf}
        \caption{Flame front position as a function of time for {\tt aprox13}, {\tt aprox13 sdc}, {\tt subch full}, and {\tt subch full sdc}. {\it Solid} Lines: Data. {\it Dashed} lines: fit.}
    \end{figure}
\end{frame}

% \begin{frame}[allowframebreaks]
% \frametitle{Integration: Flame Speed Comparison}
%     \begin{block}{ Fitting Functions}
%         \begin{subequations}

%         \begin{equation}
%              r(t) = \frac{1}{2}a_0 t^2 + v_0 t + r_0
%         \end{equation}

%         \begin{equation}
%              r(t) = A\tanh{\left(\frac{t}{B} + C\right)} + \frac{1}{2}a_0 t^2 + v_0 t + r_0
%         \end{equation}
%         \end{subequations}
%     \end{block}

%     \begin{table}
%         \resizebox{1\linewidth}{!}{
%         \begin{tabular}{ccccccc}
%         \toprule
%         Name &
%         $a_0$ [$\mbox{km} \ \mbox{s}^{-2}$] &
%         $v_0$ [$\mbox{km} \ \mbox{s}^{-1}$] &
%         $r_0$ [km] &
%         A [$\mbox{km}$]&
%         B [s]&
%         C \\ 
%         \midrule
%         {\tt aprox13} & $24.22 \pm 0.23$ & $2.812 \pm 0.015$ & $0.7680\pm 0.0004 $ & N/A & N/A & N/A\\

%         {\tt aprox13 sdc} & $28.09 \pm 0.20$ & $2.392 \pm 0.013$ & $0.7836\pm 0.0004 $ & N/A & N/A & N/A\\

%         {\tt subch full} & $9.22 \pm 0.82$ & $4.489 \pm 0.067$ & $0.8646 \pm 0.0014$ & $0.150 \pm 0.001$ & $0.0093 \pm 0.0001$ & $-2.551 \pm 0.035$ \\

%         {\tt subch full sdc} & $2.34 \pm 0.74$ & $4.947 \pm 0.061$ & $0.8601 \pm 0.0013$ & $0.146 \pm 0.001$ & $0.0098 \pm 0.0001$ & $-2.424 \pm 0.030$ \\
%         \bottomrule
%         \end{tabular}}
%         \caption{\scriptsize This table shows the fitted parameters of the fitting functions. The fitting function is applied for $t > 8$ ms.}
%     \end{table}

%     \begin{table}
%         \resizebox{1\linewidth}{!}{
%         \begin{tabular}{cccc}
%         \toprule
%         Name &
%         $v_{23}$ [$\mbox{km} \ \mbox{s}^{-1}$] &
%         $v_{100}$ [$\mbox{km} \ \mbox{s}^{-1}$] &
%         $v_{120}$ [$\mbox{km} \ \mbox{s}^{-1}$] \\
%         \midrule
%         {\tt aprox13} & $3.369 \pm 0.016$ & $5.234 \pm 0.027$ & $5.718 \pm 0.031$\\
%         {\tt subch full} & $20.732 \pm 0.284$ & $5.411 \pm 0.105$ & $5.595 \pm 0.119$\\
%         {\tt aprox13 sdc} & $3.039 \pm 0.014$ & $5.201 \pm 0.024$ & $5.764 \pm 0.027$\\
%         {\tt subch full sdc} & $19.811 \pm 0.248$ & $5.181 \pm 0.096$ & $5.228 \pm 0.108$\\
%         \bottomrule
%         \end{tabular}}
%         \caption{This table shows the instantaneous flame propagation speed at $t = 23$ ms, $t = 100$ ms and $t = 120$ ms. $t = 23$ ms represents the acceleration for {\tt subch full}, {\tt subch simple}, and {\tt subch full sdc}. $t = 100$ ms and $t = 120$ ms represent the steady phase at the late-stage.}
%     \end{table}

% \end{frame}

\begin{frame}{Integration: $u-v$ Phase Plot}
    \begin{figure}
        \centering
        \includegraphics[width=0.95\linewidth]{pres_integration_u_v_enuc_100ms.pdf}
        \caption{\scriptsize {\it First (left)}: $u-v$ phase plot for {\tt aprox13}. {\it Second}: $u-v$ phase plot for {\tt aprox13 sdc}. {\it Third}: $u-v$ phase plot for {\tt subch full}. {\it Fourth (Right)}: $u-v$ phase plot for {\tt subch full sdc}. $t = 100$ ms}
    \end{figure}
\end{frame}

\section{Conclusion}

\begin{frame}[allowframebreaks]
\frametitle{Conclusion}
    \begin{block}{Takeaways: Network Study}
    \begin{itemize}
        \item  Due to the temperature constraints, $(\alpha, \mbox{p})(\mbox{p}, \gamma)$ approximation continues to be an accurate approach in simulating thermonuclear flames propagations in XRBs.

        \item The ${}^{12}\mbox{C}(\mbox{p}, \gamma) {}^{13}\mbox{N}(\alpha, \mbox{p}){}^{16}\mbox{O}$ rates are critical in accurately modelling flame propagation in XRBs. At $T \gtrsim 10^9$ K, these reactions dominate over the triple-$\alpha$ and the slow $\alpha$ capture processes from ${}^{12}$C to ${}^{16}$O. This allows an instant depletion of ${}^{12}$C, leading to a burst of energy and flame acceleration once temperature reaches $\sim 1.3 \times 10^9$ K.

        \item Additional rates and more modern rates added to {\tt subch full mod} have gradually increased the overall $\dot{e}_{nuc}$ compared to {\tt aprox13}.
    \end{itemize}
    \end{block}

    \begin{block}{Takeaways: Screening Study}
    \begin{itemize}
        \item Thermonuclear flames in XRBs are likely in either weak or the intermediate screening regime.
        \item {\tt SCREEN5} had a slightly stronger overall screening effect compared to {\tt CHUGUNOV2007}.
    \end{itemize}
    \end{block}

    \begin{block}{Takeaways: Integration Study}
    \begin{itemize}
        \item strang-splitting and simplified-SDC had a similar overall performance.
        \item Simplified-SDC had a stronger coupling between hydrodynamics and nuclear reactions. This is expected to be more significant at more extreme thermodynamic conditions.
        % \item There is a minor computational artifact when plotting $\dot{e}_{nuc}$ for simplified-SDC.
    \end{itemize}
    \end{block}
\end{frame}

%------------------------------------------------

\begin{frame}[allowframebreaks]% Use [allowframebreaks] to allow automatic splitting across slides if the content is too long
	\frametitle{References}
    \printbibliography
\end{frame}

%----------------------------------------------------------------------------------------
%	ACKNOWLEDGMENTS SLIDE
%----------------------------------------------------------------------------------------

%----------------------------------------------------------------------------------------
%	CLOSING SLIDE
%----------------------------------------------------------------------------------------

\begin{frame}[plain] % The optional argument 'plain' hides the headline and footline
	\begin{center}
		{\Huge The End}

		\bigskip\bigskip % Vertical whitespace

		{\LARGE Questions? Comments?}
	\end{center}
\end{frame}

%----------------------------------------------------------------------------------------

\end{document}

### 2023-06-04 xrb3/plots/initial_model/initial_model.py
add plots

@@ -1,0 +1,101 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+
+def initial_model():
+
+    cool_model = "flame_wave_new_cool.out"
+    hot_model = "flame_wave_new_hot.out"
+
+    cool_data = np.loadtxt(cool_model)
+    hot_data = np.loadtxt(hot_model)
+
+
+    # make the plots
+    r_min = 0
+    r_max = 10000.0
+
+    r_min = 0
+    r_max = 4500
+
+    #-------------------------------------------------------------------------
+    # density and temperature plots
+    #-------------------------------------------------------------------------
+    dens_min = 1.e2
+    dens_max = 5.e8
+
+    sp = plt.subplot(121)
+
+    sp.set_yscale('log')
+
+
+    # density
+    plt.plot(cool_data[:,0], cool_data[:,1], color="C0")
+    plt.plot(hot_data[:,0], hot_data[:,1], color="C0", ls="--")
+
+    plt.xlabel(r"z (cm)")
+    plt.ylabel(r"density (g cm$^{-3}$)", color="C0")
+
+    plt.xlim(r_min, r_max)
+    plt.ylim(dens_min, dens_max)
+
+    ax = plt.gca()
+
+    # the offset text is the 1.e8 that appears under the axis labels when
+    # doing scientific notation
+    #ax.tick_params(labeltop='off')
+    #ax.tick_params(labelbottom='off')
+    ax.xaxis.offsetText.set_visible(False)
+
+    # temperature
+    sp2 = plt.twinx()
+    sp2.set_yscale('log')
+
+    # original
+    plt.plot(cool_data[:,0], cool_data[:,2], color="C1")
+    plt.plot(hot_data[:,0], hot_data[:,2], color="C1", ls="--")
+
+    sp2.yaxis.tick_right()
+    #plt.axis(color="C1")
+    plt.ylabel(r"temperature (K)", color="C1")
+
+    plt.xlabel(r"z (cm)")
+
+    plt.xlim(r_min, r_max)
+    plt.ylim(8.e6, 2.e9)
+
+    #-------------------------------------------------------------------------
+    # species
+    #-------------------------------------------------------------------------
+
+    sp = plt.subplot(122)
+
+    sp.set_yscale('log')
+
[…]

### 2023-06-04 xrb3/paper.tex
finish report

@@
 clean well-defined flame.  The model is constructed following the same
-procedure described in \citet{eiden:2020,harpole:2021}.
+procedure described in \citet{eiden:2020,harpole:2021}.  \edit1{We
+generate a hot and a cool model, and blend them laterally to setup
+an initial perturbation that is in hydrostatic equilibrium.  The
+structure of these models is shown
+shown in Figure~\ref{fig:initial model}.}
+
+\begin{figure}[t]
+\centering
+\plotone{initial_model_paper}
+\caption{\label{fig:initial model} Initial model structure for both
+the hot (dashed) and cool (solid) models, showing the density
+and temperature structure as well as the initial composition.}
+\end{figure}
 

### 2023-06-05 xrb5/paper.tex
Add suggestions to xrb5 paper

@@
 
-Since the timescale of accretion between each burst is in the order of $10^4$ secs while the rise time is in the order of $\sim 1 - 10$ sec \citep{Parikh_2013}, it is unlikely for the same thermodynamic condition to exist on the entire surface of the neutron star to start burning simultaneously \citep{Shara_1982}.
-Therefore, nuclear ignition is likely to begin in a localized region, and spread to the rest of the neutron star \citep{Spitkovsky_2002}.
-This asymmetrical burning leads to modulations in the observed flux during the rotation of neutron stars, which causes asymmetrical surface brightness during XRBs \citep{strohmayer_2009}.
-This explains burst oscillation behavior, a millisecond period variation of XRB intensity during rise time, and the double-peak light curve for non-photospheric-radius-expansion XRBs discovered and supported by observational data \citep{Altamirano_2010,Chakraborty_2014,Bhattacharyya_2006, Kaaret_2007,Smith_1997}.
+Since the timescale of accretion between each burst is in the order of $10^4$ secs while the rise 
+time is in the order of $\sim 1 - 10$ sec \citep{Parikh_2013}, it is unlikely for the same 
+thermodynamic condition to exist on the entire surface of the neutron star to start burning 
+simultaneously \citep{Shara_1982}. {\color{red}Previous sentence seems to be missing a few words. 
+Could maybe add ``and for the entire surface" before ``to start burning"?} Therefore, nuclear 
+ignition is likely to begin in a 
+localized region, and spread to the rest of the neutron star \citep{Spitkovsky_2002}. This 
+asymmetrical burning leads to modulations in the observed flux during the rotation of neutron 
+stars, which causes asymmetrical surface brightness during XRBs \citep{strohmayer_2009}. This 
+explains burst oscillation behavior, a millisecond period variation of XRB intensity during rise 
+time, and the double-peak light curve for non-photospheric-radius-expansion XRBs discovered and 
+supported by observational data \citep{Altamirano_2010,Chakraborty_2014,Bhattacharyya_2006, 
+Kaaret_2007,Smith_1997}.
 
@@
 We explore four different networks including a standard 13-isotope alpha chain as our reference network (see Sec.\ \ref{Sec:network}), two
-different plasma screening implementations (Sec.\ \ref{Sec:screening}), and how the reactions and hydro are coupled
+different plasma screening implementations (Sec.\ \ref{Sec:screening}), and how the reactions and 
+hydrodynamics are coupled
 together (Sec.\ \ref{Sec:integration}).
@@
 
-During an XRB, nucleosynthesis involves more than a thousand isotopes \citep{Woosley_2004,Koike_2004}.
-However, due to computational constraints, most multi-dimensional simulations incorporate fewer than 20 isotopes.
-Thus, this manuscript aims to understand how the different approximations used in the nuclear physics impact on the lateral thermonuclear flame in XRBs.
+During an XRB, nucleosynthesis involves more than a thousand isotopes 
+\citep{Woosley_2004,Koike_2004}. However, due to computational constraints, most multi-dimensional 
+simulations incorporate fewer than 20 isotopes. Thus, this manuscript aims to understand how the 
+different approximations used in the nuclear physics impact the lateral thermonuclear flame in XRBs.
 
@@
 density-weighted profile of temperature and energy generation rate in the burning, similar to the procedure in \citet{harpole:2021}. 
-In order to filter out the burning regions, we selected cells that are in the 99th percentile for temperature or energy generation rate.
-A density-weighted average is then applied to temperature and energy generation rate using Equation \ref{Eq:weighted_equation}, where $\left<Q\right>_w$ is the weighted quantity, $C_{99}(Q)$ are cells where the quantity, $Q$, is in the 99th percentile, and $\rho(c_i)$ and $Q(c_i)$ are the density and the quantity in $c_i$ cell.
+In order to filter out the burning regions, we selected cells that are in the 99th percentile 
+{\color{red} or higher? (assuming same method)} for temperature or energy generation rate. A 
+density-weighted average is then applied to temperature and energy generation rate using Equation 
+\ref{Eq:weighted_equation}, where $\left<Q\right>_w$ is the weighted quantity, $C_{99}(Q)$ are 
+cells where the quantity, $Q$, is in the 99th percentile {\color{red} or higher}, and $\rho(c_i)$ 
+and $Q(c_i)$ are the 
+density and the quantity in $c_i$ cell.
 
@@
 
-With {\tt subch\_full}, we see that the ${}^{12}$C is much more depleted and there is much more production of heavy $\alpha$ chain isotopes through the $\alpha$-chain, particularly ${}^{28}$Si, at $t = 20$ ms.
-This phenomenon is consistent with the results of \cite{Weinberg_2006}.
-The burning path, ${}^{12}\mbox{C}(\mbox{p}, \gamma) {}^{13}\mbox{N}(\alpha, \mbox{p}){}^{16}\mbox{O}$, provides an efficient route for converting ${}^{12}$C to ${}^{16}$O compared to the $\alpha$-capture rate.
+With {\tt subch\_full}, we see that the ${}^{12}$C is much more depleted {\color{red} in the 
+burning region} and there is much more production of heavy $\alpha$ chain isotopes through the 
+$\alpha$-chain, particularly ${}^{28}$Si, at $t = 20$ ms. This phenomenon is consistent with the 
+results of \cite{Weinberg_2006}. The burning path, ${}^{12}\mbox{C}(\mbox{p}, \gamma) 
+{}^{13}\mbox{N}(\alpha, \mbox{p}){}^{16}\mbox{O}$, provides an efficient route for converting 
+${}^{12}$C to ${}^{16}$O compared to the $\alpha$-capture rate. 
 %Consequently, when $T \sim 1.3 \times 10^9$ K, the flow from ${}^{12}$C to ${}^{16}$O can surpass the triple-$\alpha$ process, resulting in a depletion of ${}^{12}$C and an amplification in the nuclear energy generation rates.
@@
 the flame speed.  For the {\tt aprox13} and {\tt subch\_full\_mod} networks, we use the same expression
-as in \citet{eiden:2020}:
+as in \citet{harpole:2021}:
 \begin{equation}\label{Eq:quadratic_fit}
@@
 \end{equation}
-Both fitting functions are applied for $t > 8$ ms, and the fitted parameters along with their respective errors are shown in Table \ref{Tab:network_fitted_param}.
-The errors are calculated by taking the square root of the diagonal of the covariance matrix.
+Both fitting functions are applied for $t > 8$ ms, and the fitted parameters along with their 
+respective errors are shown in Table \ref{Tab:network_fitted_param}. The errors are calculated by 
+taking the square root of the diagonal of the covariance matrix. {\color{red} It might be worth 
+mentioning which fitting function/method was used if that isn't stated anywhere.}
 
@@
 
-Although a general trend of decreasing $\Gamma_e$ is observed as the flame progress, it is noteworthy that there are more fusion processes involving heavier nuclei at later stages of burning.
-Assuming triple-$\alpha$ process, it can inferred that $\Gamma \sim 0.05$ and $0.025$ for $t = 50$ ms and $t = 100$ ms.
-Meanwhile, for oxygen burning at $t = 50$ ms and 100 ms, $\Gamma \sim 0.48$ and $0.32$, respectively.
-These results indicate that helium burning occurs in the weak screening regime, while heavier nuclei burning processes occur in the intermediate screening regime and gradually transition into the weak regime.
+Although a general trend of decreasing $\Gamma_e$ is observed as the flame progress, it is 
+noteworthy that there are more fusion processes involving heavier nuclei at later stages of 
+burning. Assuming triple-$\alpha$ process, it can inferred that {\color{red}the Coulomb coupling 
+parameter (just for clarity)} $\Gamma \sim 0.05$ and $0.025$ for $t = 50$ ms and $t = 100$ ms. 
+Meanwhile, for oxygen 
+burning at $t = 50$ ms and 100 ms, $\Gamma \sim 0.48$ and $0.32$, respectively. These results 
+indicate that helium burning occurs in the weak screening regime, while heavier nuclei burning 
+processes occur in the intermediate screening regime and gradually transition into the weak regime.
 
@@
 
-This finding is in agreement with the results from \cite{Chugunov_2007}, where it was shown that the screening effect calculated by \cite{Alastuey_Jancovici:1978} is always greater than \cite{Chugunov_2007} when $\Gamma \lesssim 40$ (See Figure 4 in \cite{Chugunov_2007}).
+This finding is in agreement with the results from \cite{Chugunov_2007}, where it was shown that 
+the screening effect calculated by \cite{Alastuey_Jancovici:1978} is always greater than 
+\cite{Chugunov_2007} when $\Gamma \lesssim 40$ (See Figure 4 in \citealt{Chugunov_2007}). 
 %As $\Gamma$ increases, the difference between the two methods gradually decreases (see Figure 4 in \cite{Chugunov_2007}).

### 2023-06-06 xrb5/paper.tex
update and address comments

--- 2023-05-17 xrb5/paper.tex
@@
 
+\author[0000-0001-6191-4285]{Kiran Eiden}
+\affiliation{Department of Astronomy,
+University of California, Berkeley,
+Berkeley, CA 94720-3411, USA}
+
 %% \author[0000-0003-2300-5165]{Donald Willcox}
@@
 Through the use of axisymmetric 2D hydrodynamic simulations, we further investigate laterally propagating flames in X-ray bursts (XRBs).
-Our aim is to investigate the sensitivity of a propagating helium flame to different nuclear physics.
+Our aim is to understand the sensitivity of a propagating helium flame to different nuclear physics.
 Using the \castro\ simulation code, we confirm the phenomenon of enhanced energy generation shortly after a flame is established after by adding ${}^{12}\mbox{C}(\mbox{p}, \gamma) {}^{13}\mbox{N}(\alpha, \mbox{p}){}^{16}\mbox{O}$ to the network, in agreement with the past literature.
@@
 
-Since the timescale of accretion between each burst is in the order of $10^4$ secs while the rise time is in the order of $\sim 1 - 10$ sec \citep{Parikh_2013}, it is unlikely for the same thermodynamic condition to exist on the entire surface of the neutron star to start burning simultaneously \citep{Shara_1982}.
+Since the timescale of accretion between each burst is in the order of $10^4$ secs while the rise time is in the order of $\sim 1 - 10$ sec \citep{Parikh_2013}, it is unlikely for the same thermodynamic condition to exist on the entire surface of the neutron star and for the entire surface to start burning simultaneously \citep{Shara_1982}.
 Therefore, nuclear ignition is likely to begin in a localized region, and spread to the rest of the neutron star \citep{Spitkovsky_2002}.
@@
 
-In the past, numerous successes have been achieved in studying the properties of neutron stars and the mechanisms of XRBs using numerical simulations.
+Numerous successful numerical simulations have explored the properties of neutron stars and XRBs.
 One-dimensional simulations can determine the nucleosynthesis, study the rp-process, estimate burst duration, and predict the light curve of XRBs by assuming spherical symmetry \citep{Woosley_2004,Johnston_2018,Meisel_2018,Johnston_2020}.
@@
 
-We use {\castro} \citep{castro,castro_joss}, an open-source, adaptive mesh, astrophysical radiation hydrodynamics simulation code, to perform all the simulations discussed in this manuscript.
-The reaction networks,
+We use {\castro} \citep{castro,castro_joss}, an open-source, adaptive mesh, astrophysical simulation code, to perform all the simulations discussed in this manuscript. The reaction networks,
 integrators, equation of state \citep{timmes_swesty:2000}, thermal neutrino losses \citep{neutrino}, and conductivities
@@
 However, due to computational constraints, most multi-dimensional simulations incorporate fewer than 20 isotopes.
-Thus, this manuscript aims to understand how the different approximations used in the nuclear physics impact on the lateral thermonuclear flame in XRBs.
+Thus, this paper aims to understand how the different approximations used in the nuclear physics impact on the lateral thermonuclear flame in XRBs.
 
@@
 As a result, {\tt aprox13} has a total of 13 isotopes and 31 rates.
-Nevertheless, the $(\alpha, \mbox{p})(\mbox{p}, \gamma)$ approximation becomes less accurate for temperatures $\gtrsim 2.5 \times 10^9$ K, potentially affecting the energy generation rates.
+Nevertheless, the $(\alpha, \mbox{p})(\mbox{p}, \gamma)$ approximation can become less accurate for temperatures $\gtrsim 2.5 \times 10^9$ K, potentially affecting the energy generation rates.
 An illustration of this can be found in \cite{pynucastro2} (see Figure 6 there).
@@
 
-The {\tt subch\_full} was introduced in Appendix B of \cite{castro_simple_sdc} for simulating He burning in sub-Chandrasekhar mass white dwarfs (hence, the {\tt subch} prefix).
+The {\tt subch\_full} network was introduced in Appendix B of \cite{castro_simple_sdc} for simulating He burning in sub-Chandrasekhar mass white dwarfs (hence, the {\tt subch} prefix).
 {\tt subch\_full} network has three main modifications compared to {\tt aprox13} network.
@@
 We also note that their reverse rates are not included due to their negligible effects at $T \sim 10^9$ K.
-Finally, we included the additional rates ${}^{14}\mbox{N}(\alpha,\gamma){}^{18}\mbox{F}(\alpha, \mbox{p}){}^{21}\mbox{Ne}$ and ${}^{12}\mbox{C}(\mbox{p}, \gamma) {}^{13}\mbox{N}(\alpha, \mbox{p}){}^{16}\mbox{O}$, discussed in \cite{Shen_2009, Weinberg_2006,Karakas_2008} to bypass the comparatively slow ${}^{12}$C $\alpha$-capture process, ${}^{12}\mbox{C} (\alpha, \gamma) {}^{16}\mbox{O}$, when $T \gtrsim 10^9$ K.
+Finally, we included the additional rates ${}^{14}\mbox{N}(\alpha,\gamma){}^{18}\mbox{F}(\alpha, \mbox{p}){}^{21}\mbox{Ne}$ and ${}^{12}\mbox{C}(\mbox{p}, \gamma) {}^{13}\mbox{N}(\alpha, \mbox{p}){}^{16}\mbox{O}$, discussed in \cite{Shen_2009, Weinberg_2006,Karakas_2008, fisker:2008} to bypass the comparatively slow ${}^{12}$C $\alpha$-capture process, ${}^{12}\mbox{C} (\alpha, \gamma) {}^{16}\mbox{O}$, when $T \gtrsim 10^9$ K.
 The first $\alpha$-capture process on ${}^{14}\mbox{N}$ and ${}^{18}\mbox{F}$ led to the production of protons that can be used for the $\alpha$ chain process on ${}^{12}\mbox{C}$.
-When $T \gtrsim 10^9$ K, ${}^{12}\mbox{C}(\mbox{p}, \gamma) {}^{13}\mbox{N}(\alpha, \mbox{p}){}^{16}\mbox{O}$ is expected to dominate the reactive flow towards the heavy $\alpha$ chain nuclei, which can lead to a completely different end-stage composition \citep{Weinberg_2006}.
+When $T \gtrsim 10^9$ K, ${}^{12}\mbox{C}(\mbox{p}, \gamma) {}^{13}\mbox{N}(\alpha, \mbox{p}){}^{16}\mbox{O}$ is expected to dominate the reactive flow towards the heavy $\alpha$ chain nuclei, which can lead to a completely different end-stage composition \citep{Weinberg_2006, fisker:2008}.
 Since these additional rates leave ${}^{21}\mbox{Ne}$ as the endpoint, ${}^{22}\mbox{Na}$ is added to connect all the nuclei.
 With all the modifications, {\tt subch\_full} has 28 isotopes and 107 rates.
+Figure~\ref{fig:networks} shows a visualization
+of this network.
 
 {\tt subch\_full\_mod} is identical to the {\tt subch\_full} network, except that the ${}^{12}\mbox{C}(\mbox{p}, \gamma) {}^{13}\mbox{N}(\alpha, \mbox{p}){}^{16}\mbox{O}$ and its reverse reactions are turned off.
-This is done to explore the significance of these two rates, as described in \cite{Shen_2009, Weinberg_2006}.
+This is done to explore the significance of these two rates, as described in \cite{Shen_2009, Weinberg_2006, fisker:2008}.
 
@@
 The reverse rates of all ${}^{12}\mbox{C} + {}^{12}\mbox{C}$,  ${}^{16}\mbox{O} + {}^{16}\mbox{O}$, and ${}^{16}\mbox{O} + {}^{12}\mbox{C}$ are removed since they are not present in {\tt aprox13}.
-All the forward and reverse rates of ${}^{12}\mbox{C} + {}^{20}\mbox{Ne}$ , ${}^{23}\mbox{Na}(\alpha, \gamma){}^{27}\mbox{Al}$, and ${}^{27}\mbox{Al}(\alpha, \gamma){}^{31}\mbox{P}$ are also removed to simplify the network.After the simplifications, we now have 22 isotopes and 57 rates.
-See the bottom panel of Figure \ref{fig:networks} for visualization.
+All the forward and reverse rates of ${}^{12}\mbox{C} + {}^{20}\mbox{Ne}$ , ${}^{23}\mbox{Na}(\alpha, \gamma){}^{27}\mbox{Al}$, and ${}^{27}\mbox{Al}(\alpha, \gamma){}^{31}\mbox{P}$ are also removed to simplify the network.
+After the simplifications, we now have 22 isotopes and 57 rates. See the bottom panel of Figure \ref{fig:networks} for visualization.
 
@@
 Plasma screening is the enhancement of nuclear reaction rates due to the Coulomb coupling of the surrounding plasma electrons and ions.
-Plasma screening plays an important role in accurately calculating nuclear reaction rates in dense regions,
-and there are many approximations for it in the literature, but these are not
-often explored in real simulations.
+Plasma screening plays an important role in accurately calculating nuclear reaction rates in dense regions.
+Depending on the thermodynamic conditions, 
+screening can enhance the rate by several orders
+of magnitude (see, e.g., \citealt{woosley-ignition}).
+There are many approximations for screening in the literature, but these are not
+often explored and compared in real simulations.
+We consider two different screening approximations
+for our XRB simulations.
 
@@
 The reaction rate, $R_{\textrm{th}}$, is found by integrating the cross-section over the Maxwellian velocity distribution.
-If the nuclei involved in the reaction are present in a high-density region that is permeated with plasma particles, the strength of the Coulomb barrier would be reduced due to the plasma screening effect since the effective charge of the fusing nuclei are reduced.
+If the nuclei involved in the reaction are present in a high-density region that is permeated with plasma particles, then the strength of the Coulomb barrier would be reduced due to the plasma screening effect since the effective charge of the fusing nuclei are reduced.
 The screening enhancement factor is usually expressed as
@@
 By following a similar procedure outlined in \cite{itoh:1979}, the screening routine for one-component plasma can be extended to a multi-component plasma, which is suitable for a general mixture of ions.
-Finally, a weighted average between the weak and strong screening enhancement functions is used when $0.3 < \Gamma < 0.8$, the criteria for the intermediate screening regime.
+Finally, in
+the intermediate screening regime, $0.3 < \Gamma < 0.8$, a weighted average between the weak and strong screening enhancement functions is used.
 
@@
 yielding the state $\mathcal{U}^{\star\star}$.  Finally,
-the second half-timestep of burning is done to obtain the final state $\mathcal{U}^{n+1}$.
+the second half-timestep of burning is done for $\Delta t/2$ starting
+with $\mathcal{U}^{\star\star}$ to obtain the final state $\mathcal{U}^{n+1}$.
 By applying the advection and hydrodynamic operator to an intermediate state that has already incorporated the nuclear reaction effect over $\Delta t/2$, an indirect coupling is achieved between the two operations and achieves a second-order accuracy in time.
@@
 It also reduces the stiffness of solving reaction equations, leading to reduced computational expenses under extreme thermodynamic conditions.
-For
-explosive reactive flows, it was observed that the simplified-SDC method
-provides a more accurate evolution than the Strang-split method.  In this paper, we seek to understand what advantages, if any, it has for the more moderate burning in XRBs.
+In \cite{castro_simple_sdc}, it was observed that the simplified-SDC method
+provides a much more accurate evolution than the Strang-split method when evolving He and C detonations
+in white dwarfs.
 
@@
 
-We present a total of seven XRB simulations models using different combinations of reaction networks, time evolution methods, and screening routines.
-In order to investigate the sensitivity of the propagating flame to nuclear reactions, the default screening routine, {\tt SCREEN5}, and time evolution methods, Strang-splitting, are used with a variation in the networks: {\tt aprox13}, {\tt subch\_full}, {\tt subch\_full\_mod}, {\tt subch\_simple}.
+We present a total of seven XRB simulations using different combinations of reaction networks, time evolution methods, and screening routines.
+In order to investigate the sensitivity of the propagating flame to nuclear reactions, the default screening routine, {\tt SCREEN5}, and Strang-splitting are used with the four different networks: {\tt aprox13}, {\tt subch\_full}, {\tt subch\_full\_mod}, {\tt subch\_simple}.
 We replaced the {\tt SCREEN5} screening routine with {\tt CHUGUNOV2007} for the {\tt aprox13} model with Strang-splitting to investigate the difference in their performance.
-Lastly, in order to test the performance of the simplified SDC method, we ran two additional simulations using the two selected reaction networks, {\tt aprox13} and {\tt subch\_full}, with simplified SDC instead of Strang-splitting.
-The overall summary of the simulation settings is shown in Table \ref{Tab:settings}, and we will now refer to each specific model using its name.
+Lastly, in order to test the performance of the simplified SDC method, we ran two additional simulations using {\tt aprox13} and {\tt subch\_full} with simplified SDC instead of Strang-splitting.
+The overall summary of the simulations is shown in Table \ref{Tab:settings}.
+We will
+refer to the simulations by the names in the table
+in the following discussion.
 
@@
 
-Figure \ref{fig:network_abar_50ms} shows a comparison of the mean molecular weight, $\Bar{A}$, for the four networks at $t = 50$ ms.
-Regions with a larger $\Bar{A}$ represents the ash of burning process which gives a rough estimate of where the burning process has already taken place.
+Figure \ref{fig:network_abar_50ms} shows a comparison of the mean molecular weight, $\Bar{A}$, for the four networks at $t = 50$ ms. Regions with a larger $\Bar{A}$ represent the ash, tracing
+the burning in the accreted layer.
 One immediate feature is the similarity of the ash structure between {\tt aprox13} and {\tt subch\_full\_mod}, as well as between {\tt subch\_full} and {\tt subch\_simple}.
-Unlike {\tt aprox13} and {\tt subch\_full\_mod}, there is a thicker overall ash structure on the surface indicating a much more rigorous burning process in general.
-The sudden increase in the ash height from $r = 5 \times 10^4$ cm to $10^5$ cm further implies a non-uniform burning has taken place, unlike the other two models.
+For {\tt subch\_full} and {\tt subch\_simple}, there is a thicker overall ash structure on the surface indicating much more vigorous burning.
+The sudden increase in the ash height from $r = 5 \times 10^4$ cm to $10^5$ cm for these two simulations further implies a non-uniform burning has taken place, unlike in {\tt aprox13} and {\tt subch\_full\_mod}.
 The darker color means that the ash in {\tt subch\_full} and {\tt subch\_simple} is composed of heavier nuclei suggesting a faster reactive flow burning to the heavier nuclei compared to the other two models.
@@
 In order to quantify how the flame changes over time, we look at the
-density-weighted profile of temperature and energy generation rate in the burning, similar to the procedure in \citet{harpole:2021}. 
-In order to filter out the burning regions, we selected cells that are in the 99th percentile for temperature or energy generation rate.
-A density-weighted average is then applied to temperature and energy generation rate using Equation \ref{Eq:weighted_equation}, where $\left<Q\right>_w$ is the weighted quantity, $C_{99}(Q)$ are cells where the quantity, $Q$, is in the 99th percentile, and $\rho(c_i)$ and $Q(c_i)$ are the density and the quantity in $c_i$ cell.
-
+density-weighted profile of temperature and energy generation rate in the burning, following the procedure in \citet{harpole:2021}. 
+We selected cells that are in the 99th percentile or higher for temperature or energy generation rate and compute a density-weighted average of temperature and energy generation rate:
 \begin{equation}\label{Eq:weighted_equation}
@@
 \end{equation}
-
-The weighted results are shown in Figure \ref{fig:network_time_profile}.
-The profiles of {\tt subch\_full} and {\tt subch\_simple} are nearly identical profiles in both weighted temperature and energy generation rate.
+where $\left<Q\right>_w$ is the weighted quantity, $C_{99}(Q)$ are cells where the quantity, $Q$, is in the 99th percentile or higher, and $\rho(c_i)$ and $Q(c_i)$ are the density and the quantity in $c_i$ cell.
+Figure \ref{fig:network_time_profile} shows the resulting profiles.
+
+The temperature and energy generation rate profiles for {\tt subch\_full} and {\tt subch\_simple} are nearly identical.
 As the primary distinction between these two is the $(\alpha, \mbox{p})(\mbox{p}, \gamma)$ approximation for heavier nuclei, we can conclude that this approximation remains accurate in the context of XRB.
+
 In general, there is no uniform shape in both T and $\dot{e}_{\textrm{nuc}}$ for {\tt subch\_full} and {\tt subch\_simple}.
@@
 
-Unlike {\tt subch\_full} or {\tt subch\_simple}, {\tt aprox13} and {\tt subch\_full\_mod} have a much more steady burning process.
+Unlike {\tt subch\_full} or {\tt subch\_simple}, {\tt aprox13} and {\tt subch\_full\_mod} have a much more steady burning.
 Initially, {\tt subch\_full\_mod} appears to keep pace with {\tt subch\_full} and {\tt subch\_simple}, but $\dot{e}_{\textrm{nuc}}$ declines rapidly after reaching its peak following the flame-establishing phase.
@@
 
-With {\tt subch\_full}, we see that the ${}^{12}$C is much more depleted and there is much more production of heavy $\alpha$ chain isotopes through the $\alpha$-chain, particularly ${}^{28}$Si, at $t = 20$ ms.
-This phenomenon is consistent with the results of \cite{Weinberg_2006}.
+With {\tt subch\_full}, we see that the ${}^{12}$C is much more depleted in the burning region and there is much more production of heavy $\alpha$ chain isotopes through the $\alpha$-chain, particularly ${}^{28}$Si, at $t = 20$ ms.
+This phenomenon is consistent with the results of \cite{Weinberg_2006, fisker:2008}.
 The burning path, ${}^{12}\mbox{C}(\mbox{p}, \gamma) {}^{13}\mbox{N}(\alpha, \mbox{p}){}^{16}\mbox{O}$, provides an efficient route for converting ${}^{12}$C to ${}^{16}$O compared to the $\alpha$-capture rate.
@@
 Similar to {\tt aprox13}, {\tt subch\_full\_mod} exhibits a considerable amount of ${}^{12}$C due to the absence of ${}^{12}\mbox{C}(\mbox{p}, \gamma) {}^{13}\mbox{N}(\alpha, \mbox{p}){}^{16}\mbox{O}$ rate.
-However, unlike {\tt aprox13}, higher concentration of ${}^{16}$O is observed in the upper vertical regions in {\tt subch\_full\_mod}.
+However, unlike {\tt aprox13}, a higher concentration of ${}^{16}$O is observed in the upper vertical regions in {\tt subch\_full\_mod}.
 This is likely the result of incorporating various additional rates, including more up-to-date rates from {\sf REACLIB} library, in {\tt subch\_full\_mod} compared to {\tt aprox13}.
@@
     \centering
-    \plotone{network_species_summary.pdf}
+    \plotone{network_species_summary_log.pdf}
     \caption{\label{fig:network_species_summary} The overall evolution of the total mass for ${}^{12}$C, ${}^{16}$O, ${}^{20}$Ne, ${}^{24}$Mg, ${}^{28}$Si, and ${}^{32}$S for the 4 simulations models: {\tt aprox13}, {\tt subch\_full}, {\tt subch\_full\_mod}, and {\tt subch\_simple}.}
@@
 Nevertheless, when $t \gtrsim 18$ms, corresponding to $T \sim 1.3 \times 10^9$ K, the flow from ${}^{12}$C to ${}^{16}$O can surpass the triple-$\alpha$ process, resulting in a depletion of ${}^{12}$C.
-This phenomenon is consistent with \cite{Weinberg2007}, which claims a similar behavior at $t \sim 1.2 \times 10^9$ K.
+This phenomenon is consistent with results in \cite{Weinberg2007, fisker:2008}.
 Consequently, there is an amplification in the nuclear energy generation rates and the mass production of heavy isotopes such as ${}^{28}$Si.
@@
 
-Figure \ref{fig:network_u_v_enuc} shows the $u$-$v$ phase plot (radial velocity vs.
-vertical velocity) for the four simulations models at $t = 25$ ms and $t = 100$ ms, colored by $\dot{e}_{\textrm{nuc}}$.
+Figure \ref{fig:network_u_v_enuc} shows the $u$-$v$ phase plot (radial velocity vs.\ vertical velocity) for the four simulations models at $t = 25$ ms and $t = 100$ ms, colored by $\dot{e}_{\textrm{nuc}}$.
 Notably, the peak of the energy spike occurs at $t \sim 25$ ms for {\tt subch\_full} and {\tt subch\_simple},  resulting in the comparatively larger distribution in their $u$-$v$ phase space plots.
@@
 
-To examine an overall impact of different reaction networks on the dynamics of the laterally propagating flame, Figure \ref{fig:network_front} shows the radial flame front position as a function of time.
+To examine the overall impact of different reaction networks on the dynamics of the laterally propagating flame, Figure \ref{fig:network_front} shows the radial flame front position as a function of time.
 We follow the approach outlined in \cite{eiden:2020}, where the position of the flame front is defined as the location at which $\dot{e}_{\textrm{nuc}}$ of the 1D radial profile first drops to $0.1\%$ of the global maximum $\dot{e}_{\textrm{nuc}}$ for $r > r_{\mbox{max}(\dot{e}_{nuc})}$.
@@
 We fit the flame front position with a simple function to estimate
-the flame speed.  For the {\tt aprox13} and {\tt subch\_full\_mod} networks, we use the same expression
-as in \citet{eiden:2020}:
+the flame speed.  For the {\tt aprox13} and {\tt subch\_full\_mod} networks, we use the same expression, Eq. \ref{Eq:quadratic_fit}, 
+as in \citet{harpole:2021}.
 \begin{equation}\label{Eq:quadratic_fit}
@@
 \end{equation}
-For the {\tt subch\_full} and {\tt subch\_simple}, a hyperbolic tangent function is added to the fitting function to account for the burst of acceleration at $10 \mbox{ms} \lesssim t \lesssim 25 \mbox{ms}$.
+On the other hand, Eq.
+\ref{Eq:tanh_fit} is used for {\tt subch\_full} and {\tt subch\_simple} networks.
+A hyperbolic tangent function is added to the fitting function to account for the burst of acceleration at $10 \ \mbox{ms} \lesssim t \lesssim 25 \ \mbox{ms}$.
 \begin{equation}\label{Eq:tanh_fit}
@@
 
-In this section, we present a comparison between the effects of {\tt SCREEN5} and {\tt CHUGUNOV2007} screening routines on the dynamics of the propagating flame in XRBs.
+We now present a comparison between the effects of {\tt SCREEN5} and {\tt CHUGUNOV2007} screening routines on the dynamics of the propagating flame in XRBs.
 Figure \ref{fig:screen_temp} shows the temperature for {\tt aprox13} and {\tt aprox13\_chu} at $t = 50$ ms, which suggests that there are no significant differences in the flame structure between the two models.
@@
 Although a general trend of decreasing $\Gamma_e$ is observed as the flame progress, it is noteworthy that there are more fusion processes involving heavier nuclei at later stages of burning.
-Assuming triple-$\alpha$ process, it can inferred that $\Gamma \sim 0.05$ and $0.025$ for $t = 50$ ms and $t = 100$ ms.
+Assuming triple-$\alpha$ process, it can inferred that the Coulomb coupling parameter, $\Gamma \sim 0.05$ and $0.025$ for $t = 50$ ms and $t = 100$ ms.
 Meanwhile, for oxygen burning at $t = 50$ ms and 100 ms, $\Gamma \sim 0.48$ and $0.32$, respectively.
@@
 This result is nearly identical to the models that employed Strang-splitting as shown in Figure \ref{fig:network_abar_50ms}.
-One consequence of the strong coupling between advection and reactions is that to make the energy generation plot, we need to derive the nuclear energy release by subtracting off the advection contribution over a timestep.
+One consequence of the strong coupling between advection and reactions is, that to make the energy generation plot, we need to derive the nuclear energy release by subtracting off the advection contribution over a timestep.
 In regions where there is not much burning, roundoff error can introduce some noise into the energy generation plot, which is seen as the multiple dotted regions in the $\dot{e}_{\textrm{nuc}}$ plot for {\tt aprox13\_sdc} and {\tt subch\_full\_sdc}.
@@
 The main advantage of the simplified-SDC integration lies in its stronger coupling between the reaction and hydrodynamics.
-This feature is demonstrated in the $u-v$ phase plot for {\tt aprox13\_sdc} and {\tt subch\_full\_sdc} (Figure \ref{fig:integration_u_v_enuc}).
-Comparing to the $u-v$ phase plots for {\tt aprox13} and {\tt subch\_full} at $100$ ms (Figure \ref{fig:network_u_v_enuc}), Figure \ref{fig:integration_u_v_enuc} shows a smoother blend in the border of the regions with higher and lower $\dot{e}_{\textrm{nuc}}$, corresponding to negative and positive $u$, respectively.
+This feature is demonstrated in the $u$-$v$ phase plot for {\tt aprox13\_sdc} and {\tt subch\_full\_sdc} (Figure \ref{fig:integration_u_v_enuc}).
+Comparing to the $u$-$v$ phase plots for {\tt aprox13} and {\tt subch\_full} at $100$ ms (Figure \ref{fig:network_u_v_enuc}), Figure \ref{fig:integration_u_v_enuc} shows a smoother blend in the border of the regions with higher and lower $\dot{e}_{\textrm{nuc}}$, corresponding to negative and positive $u$, respectively.
 The smoother transition suggests this scheme is capable to resolve the flame in a volatile condition.

### 2023-06-13 subch_tols/paper.tex
start of subch tol paper

--- 2023-05-17 xrb3/paper.tex
@@
 
+\newcommand{\ddx}[1]{{\frac{{\partial#1}}{\partial x}}}
+\newcommand{\ddxs}[1]{{\frac{{\partial}}{\partial x}}#1}
+\newcommand{\ddt}[1]{{\frac{{\partial#1}}{\partial t}}}
+\newcommand{\odt}[1]{{\frac{{d#1}}{dt}}}
+\newcommand{\divg}[1]{{\nablab \cdot \left (#1\right)}}
+\newcommand{\dedr}{\left . {\partial{}e}/{\partial\rho}\right |_{T, X_k}}
+\newcommand{\dedrd}{\left . \frac{\partial{}e}{\partial\rho}\right |_{T, X_k}}
+\newcommand{\dedX}{\left . {\partial{}e}/{\partial{}X_k} \right |_{\rho, T}}
+\newcommand{\dedXd}{\left . \frac{\partial{}e}{\partial{}X_k} \right |_{\rho, T, X_{j,j\ne k}}}
+\newcommand{\dedT}{\left . {\partial{}e}/{\partial{}T} \right |_{\rho,X_k}}
+\newcommand{\dedTd}{\left . \frac{\partial{}e}{\partial{}T} \right |_{\rho,X_k}}
+
+\newcommand{\Ic}{{\boldsymbol{\mathcal{I}}}}
+\newcommand{\Ics}{{\mathcal{I}}}
+\newcommand{\kmax}{{k_\mathrm{max}}}
+\usepackage{bm}
+
+\newcommand{\Uc}{{\,\bm{\mathcal{U}}}}
+\newcommand{\Fb}{\mathbf{F}}
+\newcommand{\Sc}{\mathbf{S}}
+
+\newcommand{\xv}{{(x)}}
+\newcommand{\yv}{{(y)}}
+\newcommand{\zv}{{(z)}}
+
+\newcommand{\ex}{{\bf e}_x}
+\newcommand{\ey}{{\bf e}_y}
+\newcommand{\ez}{{\bf e}_z}
+
+\newcommand{\Ab}{{\bf A}}
+\newcommand{\Sq}{{\bf S}_\qb}
+\newcommand{\Sqhydro}{{\Sq^{\mathrm{hydro}}}}
+\newcommand{\qb}{{\bf q}}
+
+\newcommand{\Gb}{{\bf G}}
+\newcommand{\Rb}{{\bf R}}
+\newcommand{\Rq}{{\bf R}}
+\newcommand{\Adv}[1]{{\left [\boldsymbol{\mathcal{A}} \left(#1\right)\right]}}
+\newcommand{\Advt}[1]{{\left [\boldsymbol{\mathcal{\tilde{A}}} \left(#1\right)\right]}}
+\newcommand{\Advss}[1]{{\left [{\mathcal{{A}}} \left(#1\right)\right]}}
+\newcommand{\Advsst}[1]{{\left [{\mathcal{\tilde{A}}} \left(#1\right)\right]}}
+\newcommand{\Advs}[1]{\boldsymbol{\mathcal{A}} \left(#1\right)}
+
 \setlength{\marginparwidth}{0.5in}
@@
 %======================================================================
-\title{Comparing Early Evolution of Flames in X-ray Bursts in Two and Three Dimensions}
+\title{Influence of Time Integration Methodology on Double Detonation Type
+Ia Supernova Nucleosynthesis}
 
-\shorttitle{Early XRB Flame Evolution}
-%% \shortauthors{Harpole et al.}
+\shorttitle{Integration Methods and SNe Ia Nucleosynthesis}
 
@@
 \affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
-             Stony Brook, NY 11794-3800}
+             Stony Brook, NY 11794-3800, USA}
 
-\author[0000-0001-6191-4285]{Kiran Eiden}
-\affiliation{Department of Astronomy,
-University of California, Berkeley,
-Berkeley, CA 94720-3411, USA}
+\author[0000-0002-2839-107X]{Zhi Chen}
+\affiliation{Department of Physics and Astronomy,
+Stony Brook University,
+Stony Brook, NY 11794-3800, USA}
 
@@
 \affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
-             Stony Brook, NY 11794-3800}
+             Stony Brook, NY 11794-3800, USA}
+
+\author[0000-0001-5961-1680]{Alexander Smith Clark}
+\affiliation{Department of Physics and Astronomy,
+Stony Brook University,
+Stony Brook, NY 11794-3800, USA}
+
+\author[0000-0003-3603-6868]{Eric T. Johnson}
+\affiliation{Department of Physics and Astronomy,
+Stony Brook University,
+Stony Brook, NY 11794-3800, USA}
 
@@
 
-X-ray bursts (XRBs) result from thermonuclear burning of an accreted
-H/He or He layer on a neutron star \citep{galloway:2017}.
-Observations of brightness oscillations in the rise of the lightcurve
-suggest that the burning begins localized and then spreads across the
-neutron star \citep{bhattacharyya:2007}. 
-This spreading is inherently
-a multi-dimensional phenomenon, requiring hydrodynamic simulations that
-resolve the reactive zone, realistic reaction networks, and
-domain sizes that capture the scales over which rotation is important in order to
-understand the flame propagation and nucleosynthesis.  Detailed
-nuclear physics is especially important for understanding the 
-rp-process in mixed H/He bursts, as discussed in \citet{rpprocess}.
+The double detonation model
 
-Here we show a first attempt at modeling the early evolution of a spreading
-hotspot in three dimensions.  This builds on our earlier two-dimensional
-work \citep{eiden:2020,harpole:2021} that developed our
-simulation framework and explored how the acceleration of the burning
-front depends on the initial model structure.  As with those
-simulations, we use the freely-available \castro\ simulation code
-\citep{castro,castro_joss}, with all of the code needed to run these
-simulations in our public GitHub repositories.
+Detonations can be difficult to model numerically.  Time step cuts \MarginPar{ref}.
+For the double detonation scenario, the ignition of the second detonation
+can be very sensitive to how the simulation is run \MarginPar{refs}.
+Artifically limiting the rates \MarginPar{ref}
 
-Our simulations complement the two- and three-dimensional models of flames of
-\citet{cavecchi:2013,art-2015-cavecchi-etal,art-2016-cavecchi-etal,Cavecchi2019}
-by focusing on resolving the reaction zone of the flame and modeling
-the vertical structure hydrodynamically instead of hydrostatically.
-We also focus on modeling flames starting from a small hot spot.  This
-however means that we are confined to smaller domains, with similar
-computational resources, so one of the goals of this study is to
-understand how far we can push resolved XRB flame simulations.  The
-work by \citet{goodwin:2021} also explores multidimensional evolution,
-looking at the thermal transport and how affects the location where
-the hotspot ignites.  Together, all three approaches help build a
-comprehensive multidimensional picture of X-ray bursts.
+Integration of reaction networks takes various forms in astrophysical
+simulation codes.
 
-The goal of this paper is to understand the challenges of resolved
-three-dimensional hydrodynamical simulation and explore what is
-possible with today's resources.  This will inform our follow-on
-simulations.
+When using operator-splitting, the density remains constant, since there
+is no reaction source in the mass continuity equation.  The mass
+fractions, $X_k$, evolve according to:
+\begin{equation}
+\frac{dX_k}{dt} = \omegadot_k
+\end{equation}
+where $\omegadot_k$ is the creation rate of species $k$ due to nuclear
+reaction.  In the simplest approximation of operator splitting, the
+evolution of the mass fractions is solved alone, without integrating
+the temperature or energy.  This is the form that was used in FLASH
+\cite{flash}.  A more accurate operator splitting also includes the
+energy evolution (perhaps in terms of temperature).  This will require
+calling the equation of state each time we evaluate the righthand side
+of the ODE system, increasing the cost.  However, as we showed in
+\citet{strang_rnaas}, this is needed to get second-order convergence.
+
+In \citet{castro_simple_sdc}, we introduced a method based on the
+ideas of spectral deferred corrections for coupling reactions and
+hydrodyamics (that we called ``simplified-SDC'').  Here, the overall
+time-integration is done iteratively, with the hydrodynamics seeing an
+explicit reactive source, and the reaction update evolving an ODE
+system that includes a piecewise-constant-in-time advective source,
+$\Adv{\Uc}^{n+1/2}$,
+\begin{equation}
+\ddt{\Uc} = \Rb(\Uc) + \Adv{\Uc}^{n+1/2}
+\end{equation}
+
+In addition to the time-integration strategy, integrating the reaction
+ODE system requires specifying tolerances.  Unfortunately, the tolerances
+used are not normally reported in papers, and in many cases, may not
+be easily controlled by a code user at runtime.
+
+\citet{castro_simple_sdc} looked at an extreme version of the 
+double detonation scenario---a very large perturbation was applied
+to drive a detonation directly into the underlying 
+carbon-oxygen white dwarf.  We showed in that paper that the Strang
+split integration method had difficulty with the integration unless
+we used tighter tolerances
+
+Summary of the SDC subch results form the original paper
 
@@
 
-In \citet{eiden:2020}, we developed the simulation methodology in
-\castro\ for modeling laterally spreading flames through the accreted
-layer on the surface of a neutron star.  For these models,
-\castro\ evolves the compressible Euler equations with a Coriolis
-force arising from rotation, constant gravity (the plane-parallel
-approximation), reactive sources, and thermal diffusion (integrated
-explicitly)\edit1{:
-\begin{eqnarray}
-\frac{\partial \rho}{\partial t} + \nabla \cdot (\rho \Ub) &=& 0 \\
-\frac{\partial (\rho X_k)}{\partial t} + \nabla \cdot (\rho \Ub X_k) &=& \rho \omegadot_k \\
-\frac{\partial (\rho \Ub)}{\partial t} + \nabla \cdot (\rho \Ub \Ub) + \nabla p &=& \rho {\bf g} - 2 \rho {\bf \Omega} \times \Ub \\
-%% no centrifugal- \rho {\bf \Omega} \times ({\bf \Omega} \times {\bf r}) \\
-\frac{\partial (\rho E)}{\partial t} + \nabla \cdot (\rho \Ub E + p \Ub) &=& \rho \Ub \cdot {\bf g} + \nabla \cdot k_\mathrm{th} \nabla T + \rho \dot{\epsilon}
-%% no centrifugal - \rho (\Omegab \cdot \rb)(\Omegab \cdot \Ub) + \rho | \Omegab |^2 (\Ub \cdot r)
-\end{eqnarray}
-Here, $\rho$ is the mass density, $X_k$ are the mass fractions of the
-nuclear species with associated creation rates $\omegadot_k$, $\Ub$
-the velocity, $E$ is the specific total energy, $p$ is the pressure,
-${\bf g}$ is the gravitational
-acceleration, $\Omegab$ is the rotation vector, $k_\mathrm{th}$ is the thermal conductivity, and $T$ is the temperature.
-}
- We use a general stellar equation of state with
-radiation, ions as a perfect gas, and degenerate/relativistic
-electrons described in \citet{timmes_swesty:2000} \edit1{which closes
-  the system:}
-\begin{eqnarray}
-E &=& e + \frac{1}{2} |\Ub|^2 \\
-p &=& p(\rho, e, X_k) \\
-T &=& T(\rho, e, X_k)
-\end{eqnarray}
-Building on these
-first simulations, in \citet{harpole:2021}, we explored several
-different rotation rates and initial model thermal structures to
-understand how the flame speed depends on the thermal structure of the
-envelope.  Both of these papers used a two-dimensional axisymmetric
-geometry.  The three-dimensional simulations we present here build on
-these works.  We use a neutron star crust temperature of $2\times
-10^8~\mathrm{K}$ and rotation rate of 1000~Hz, as followed in
-\citet{harpole:2021}.  This model was chosen here since it gives a
-clean well-defined flame.  The model is constructed following the same
-procedure described in \citet{eiden:2020,harpole:2021}.
+We use a 2d axisymmetric domain with a size of $5.12\times 10^9~\mathrm{cm}$ by $1.024\times 10^{10}~\mathrm{cm}$.  This is
+much larger than the initial size of the star, giving it plenty of
+room to expand.  The coarse grid is $640\times 1280$ zones and we use
+2 levels of refinement, giving a maximum resolution of
+$20~\mathrm{km}$.  The refinement strategy is picked to refine on the
+star and any regions where the temperature is greater than
+$10^8~\mathrm{K}$.  \castro\ uses subcycling, so finer grids are
+advanced at a smaller timestep than the coarse grids.  Self-gravity is
+done using a full Poisson solve using geometric multigrid, with
+Dirichlet boundary conditions on the domain boundary computed via a
+multipole expansion with a maximum order of 6.  For the low density
+regions outside of the star, we use a sponge term on the momentum
+equation to prevent the very low density material that is not in
+hydrostatic equilibrium from raining down on the star.
 
-Moving to 3D is very expensive, so to further reduce the computational
-expense, we use an anisotropic resolution, with 32 cm resolution
-laterally and 16 cm resolution vertically.  This is finer vertically
-than done in \citet{harpole:2021}.  Because the hotspot is placed in
-the center of the domain, the size of the domain also becomes a
-constraint, and as a result, the evolution time we can reach will
-become limited by the time it takes the burning to approach the edge
-of the domain.  Consequently, we focus here on the early evolution.
+If during the advance of a timestep an error is generated (negative
+density, ODE integration fails, mass fractions don't sum to 1), the
+step is thrown-out and retied with a smaller timestep.  This is done
+on a level-by-level basis in the overall AMR subcycling hierarchy.
 
-\castro\ uses the the \amrex\ adaptive mesh refinement library
-\citep{amrex_joss} to manage the discretization and domain
-decomposition.  Our 3D simulation used a base grid of $768^2 \times
-192$ zones with 2 levels of refinement, the first jumping by a factor
-of 4 and the next by a factor of 2.  We subcycle in time, so the
-coarser grids take a large timestep than the fine grids.  We used
-static mesh refinement, to achieve better load balancing, fully
-refining the atmosphere below a height of $3600~\mathrm{cm}$.  The
-domain has a size $(1.96608\times 10^5~\mathrm{cm})^2 \times
-2.4576\times 10^4~\mathrm{cm}$.  The simulation methodology and
-parameters are mostly identical to those used in \citet{harpole:2021}.
-We use an unsplit piecewise parabolic method
-\citep{ppmunsplit,millercolella:2002} for the advection and operator
-(Strang) splitting for the reactions, evolving an internal energy
-evolution equation during the burn, as described in
-\citet{strang_rnaas}.  These simulations used the 7-isotope He-burning
-network described in \citet{iso7}.  This was one of the networks used
-in \citet{eiden:2020}.  \edit1{Thermal diffusion is treated explicitly, using the conductivities from \citet{Timmes00}.
-The overall integration strategy is second-order accurate in space and time.}  The main difference with the previous
-simulations is that we switched from a hydrostatic to reflecting
-boundary at the bottom of the domain and used a simple well-balanced
-scheme (similar to \citealt{kappeli:2016} but adapted to deal with
-characteristic tracing in PPM as described in \citealt{ppm-hse}).
-Tests demonstrated that this boundary does a better job than our
-previous approach at supporting the atmosphere as the flame propagates
-for long times.
+The reaction system is integrating using a version of the VODE ODE integrator
+\citep{vode} ported to C++.  Our modifications to this integrator are described
+in \citet{castro_simple_sdc}.
 
-The simulations were run on the OLCF Summit supercomputer, on 342 to
-1366 nodes, with 6 NVIDIA V100 GPUs per node.  The entire computation
-was offloaded to GPUs using the \amrex\ C++ abstraction layer for parallel \texttt{for} loops, as described in
-\citet{castro_gpu}.  GPU offloading was critical to being able to
-perform these simulations, since we run more than an order of
-magnitude faster using all the GPUs on a node as compared to all of
-the CPU cores on the node.  Overall, about 250,000 node-hours were used for
-the calculation.  The full state of the calculation was output only
-every 0.005 s, as each output file is 2.4 TB in size (for single
-precision data).  We output a few fields more frequently for
-visualization, and global diagnostics were output every timestep.  We
-ran a 2D simulation with the same resolution and refinement strategy
-to compare with here.
+We will run a suite of simulations all using the same initial model.
+The initial model is exactly the same as in \citep{castro_simple_sdc}.
 
-\begin{figure}[t]
-\centering
-\plotone{full}
-\caption{\label{fig:vr_abar} Volume rendering of the three-dimensional XRB
-simulation showing the mean molecular weight, $\bar{A}$.  \edit1{Two views
-are shown: the left column is viewed from above while the right column
-shows it from a shallow angle above the surface.  The panels show the state at 10 ms, 20 ms, and 40 ms (from top to bottom).
-The colorbar shows the value of $\bar{A}$ together with the opacity assigned for
-that value---this allows us to bring out some structure in the rendering.  A
-small triad in the upper right of the top panels shows the orientation,
-with red = $x$, green = $y$, and blue = $z$, our vertical direction.}}
-\end{figure}
+The simulations are all run on the OLCF Frontier machine, using 4
+nodes / 32 AMD GPUs.  The data is moved to the GPUs at the start of
+the simulation and all computation is done there.  Our GPU offloading
+strategy takes advantage of the \amrex\ C++ lambda-capturing
+functionality to be performance portable.
 
-Figure~\ref{fig:vr_abar} shows the a three-dimensional simulation via
-top-down volume rendering of the mean-molecular weight,
-\begin{equation}
-\frac{1}{\bar{A}} = \sum_k \frac{X_k}{A_k}
-\end{equation}
-where $X_k$ is the mass fraction of species $k$ and $A_k$ is the
-atomic weight (in mass units).  The structure is shown at 10, 20, and
-40~ms of simulation time.  While initially quite symmetric, the axial
-symmetry is broken\edit1{, likely due to roundoff error,}  and the flame structure takes on a more tenuous
-form.  The panels show the flame spreading considerably through the
-domain as time evolves, and the simulation is halted once the flame
-nears the boundaries.
-\begin{figure*}[t]
-\centering
-\plotone{time_series}
-\caption{\label{fig:2d_abar} Time-sequence of the two-dimensional axisymmetric
-simulation using the same setup as the three-dimensional simulation shown
-in Figure~\ref{fig:vr_abar}.  The domain is zoomed in, showing only the left
-two-thirds of the surface.}
-\end{figure*}
+We run 3 different types of integration methods: traditional operator
+splitting, using Strang splitting, both with integration of the energy
+equation and without any temperature evolution during the burn.  \MarginPar{comment about the EOS}.  These methods are compared to the simplified-SDC integtation.
 
-\begin{figure*}[t]
-\centering
-%\epsscale{0.75}
-\plotone{time_series_3D}
-%\epsscale{1.0}
-\caption{\label{fig:3dslice} Vertical slices through the three-dimensional
-simulation \edit1{at the same times as in Figure~\ref{fig:2d_abar}}, showing a region centered on the flame viewed from the side.  The
-flame structure is clearly seen. }
-\end{figure*}
+We also vary the CFL number and the integration tolerances.
 
-Figure~\ref{fig:2d_abar} shows the corresponding two-dimensional
-axisymmetric simulation.  In the axisymmetric geometry, each zone is
-essentially a torus, rotated about the vertical axis.  The initial
-model and simulation parameters are identical to the three-dimensional
-case.  We see that the overall structure and evolution looks identical
-to that reported in \cite{harpole:2021}.  We also see that there is a
-well-defined flame in the simulation plane that spreads outward with
-time.  A vertical slice through the three-dimensional simulation (in
-the $x$-$z$ plane) is shown in Figure~\ref{fig:3dslice} at the final
-time, 40~ms.  Aside from the lack of reflection symmetry from the
-absence of axisymmetric geometry, it looks very similar to the two-dimensional case.
+Show Ni56 mass, intermediate mass
 
-%% \begin{figure}[t]
-%% \centering
-%% \plotone{flame_wave_1000Hz_25cm_smallplt88081_Hnuc_annotated_top} \\
-%% \plotone{flame_wave_1000Hz_25cm_smallplt174562_Hnuc_annotated_top.png}
-%% \caption{\label{fig:vr_hnuc} stuff}
-%% \end{figure}
+Show time-sequence + compression plot
 
-Because the 3D flame does not stay perfectly circular, it is difficult
-to define the flame speed using the same method employed in
-\citet{eiden:2020}.  Instead, we will look at the mass of the ash
-material to assess how quickly the burning is taking place.
-Figure~\ref{fig:mass_plot} shows the mass of the different species as
-a function of time, for the 2D and 3D calculation.  The axisymmetric
-2D calculation has slightly less volume than the 3D domain, since
-rotating about the symmetry axis produces a cylindrical domain
-inscribed in the 3D domain.  To compensate for this, we scale the
-masses by the total mass on the grid.  This plot shows that the
-masses of the heavy species, especially \isot{C}{12}, grow quickly
-with time.  The differences between the 2D and 3D simulations appear
-small---the burning is slightly faster in 2D, but the trends track
-very well.  This slight difference is likely because the assumption of
-axisymmetry there does not allow for the complex structure we see in
-the evolution of composition in the 3D flame.  This strong agreement
-suggests that using 2D axisymmetric calculations is a good model for
-the early flame evolution.  And since they are so much less
-computationally expensive, this will allow us to explore much more
-complex reaction networks and understand the nucleosynthesis in more
-detail.
-
-\begin{figure}[t]
-\centering
-\plotone{mass_plot}
-\caption{\label{fig:mass_plot} Mass of species scaled to initial He
-  mass as a function of time.  The 3D simulation is shown as the solid
-  lines and the 2D simulation is shown as the dashed lines.}
-\end{figure}
+Show cost plot
 
@@
 
-We have extended our simulation framework to three-dimensions, and
-explored the differences between a two-dimensional axisymmetric
-simulation and a fully three-dimensional hydrodynamical simulation.
-Overall, the early evolution of the flame spreading from a hotspot
-behaves similarly in two- and three-dimensions.  As this
-three-dimensional simulation pushed the limits of available
-computational resources, this result suggests that we can continue to
-use two-dimensional axisymmetric simulations to explore the initial
-flame structure and spend our computational resources on larger networks
-and bigger domains.
+Our results suggest that exploring the time-integration method and
+integrator tolerances is an important part of demonstrating
+convergence of simulations involving explosive reacting flows in
+astrophysics.
 
-These results also suggest that we can do initial explorations of full
-star bursts in two-dimensions, perhaps using axisymmetric spherical
-coordinates (modeling an annular region in $r$-$\theta$).  In this
-geometry, point ignition requires starting at the poles, but this
-would allow us to see the effect of varying gravity with latitude (due
-to centrifugal forces), as well as begin to consider lightcurve modeling.
+In the near future we will show how to include various nuclear
+statistical equilibrium approximations into the simplified-SDC
+formalism while retaining second-order accuracy.
 
-There is still a role for three-dimensional simulations---the
-convective burning in the accreted layer should create turbulence that
-the flame will encounter as it propagates.  Although we have shown
-that for the flame in a Type Ia supernova, the turbulence from the
-era of convection is not strong enough to disrupt the
-flame~\citep{wdturb}, the flame in an XRB is considerably slower and
-thicker, so it remains to be seen what effect it might have.  This can
-be assessed in a different geometry, like a long channel, which would
-allow us to trade domain size for resolution to increase the numerical
-Reynolds number.  We also will consider higher-order simulation
-methodologies which could allow us to capture these effects at lower
-resolution.  We've already demonstrated a fully fourth-order accurate
-(in space and time) algorithm for coupling hydro, diffusion, and
-reactions in \castro~\citep{castro-sdc} that could be used here.  The
-main outstanding issue with applying that work to the present problem
-is the multilevel time integration.
-
-We are currently exploring mixed H/He bursts as well as the
-sensitivity of the flame properties to the size of the reaction
-network used.  Both of these initial studies use our two-dimensional
-axisymmetric geometry.
+We encourage authors to include a discussion about the reaction integration algorithm
+and tolerances in papers to permit comparisons in the future.
 

### 2023-06-14 subch_tols/paper.tex
add a bit more

@@
 
+All simulations use the same network, a 22 nucleus network designed for
+He and C burning, using the ReacLib rates \cite{reaclib} and produced with
+\pynucastro \citep{pynucastro2}.  This network, called {\tt subch\_simple} was
+described in \cite{zhi2023}.
+
+The reaction system is integrated using a version of the VODE ODE integrator
+\citep{vode} ported to C++.  Our modifications to this integrator are described
+in \citet{castro_simple_sdc}.
+
 If during the advance of a timestep an error is generated (negative
@@
 
-The reaction system is integrating using a version of the VODE ODE integrator
-\citep{vode} ported to C++.  Our modifications to this integrator are described
-in \citet{castro_simple_sdc}.
+We will run a suite of simulations all using the same initial model.
+The initial model is constructed following the methodology in
+\citet{subchandra}.  We use an isothermal core of C/O of $1.1~M_\odot$
+with $T = 10^7~\mathrm{K}$ and a thin transition region where the
+temperature ramps up to $1.75\times 10^8~\mathrm{K}$ and the
+composition changes to 99\% \isot{He}{4} and 1\% \isot{N}{14}.  This
+added \isot{N}{14} helps produce protons via
+$\isotm{N}{14}(\alpha,\gamma)\isotm{F}{18}(\alpha,p)\isotm{Ne}{21}$
+which then allow for
+$\isotm{C}{12}(p,\gamma)\isotm{N}{13}(\alpha,p)\isotm{O}{16}$.  This
+sequence can be faster than
+$\isotm{C}{12}(\alpha,\gamma)\isotm{O}{16}$, as pointed out by
+\citet{shenbildsten} and is important for getting the detonation speed
+correctly.  This He envelope is then integrated isentropically.  The
+density of the transition from the underlying CO white dwarf to the He
+envelope was selected to yield an envelope mass of $0.05~M_\odot$.
 
-We will run a suite of simulations all using the same initial model.
-The initial model is exactly the same as in \citep{castro_simple_sdc}.
+We place a small temperature perturbation in the He layer using the
+same form as in \citet{castro_simple_sdc}:
+\begin{equation}
+  T = T_0 \left [ 1 + X(\isotm{He}{4}) f (1 + \tanh(2 - r_1)) \right ]
+\end{equation}
+where
+\begin{equation}
+  r_1 = \left [ x^2 + (y - r_0)^2 \right ]^{1/2} / \lambda
+\end{equation}
+and
+\begin{equation}
+  r_0 = r_\mathrm{pert} + r_\mathrm{base}
+\end{equation}
+Here, $r_\mathrm{base}$ is the radius at which the helium layer begins
+and $r_\mathrm{pert}$ is the distance above the base to put the
+perturbation.  We choose $r_\mathrm{pert} = 100~\mathrm{km}$.  The
+temperature is perturbed above the initial model value, denoted as
+$T_0$ here.  The amplitude of the perturbation is $f = 3$ and the
+scale of the perturbation is $\lambda = 12.5~\mathrm{km}$, and
+$r_\mathrm{pert} = 100~\mathrm{km}$.  This is a very small
+perturbation, but enough to seed the initial He detonation in the
+envelope.
 
@@
 equation and without any temperature evolution during the burn.  \MarginPar{comment about the EOS}.  These methods are compared to the simplified-SDC integtation.
+
+\begin{figure*}[t]
+\centering
+\plotone{subch_Temp_sequence}
+\caption{\label{fig:temp_sequence} Time-sequence of the SDC run showing the temperature.}
+\end{figure*}
+
+\begin{figure*}[t]
+\centering
+\plotone{subch_lap_rho_sequence}
+\caption{\label{fig:lap_rho_sequence} Time-sequence of the SDC run showing the compression.}
+\end{figure*}
 
@@
 
+The ignition of the carbon detonation remains numerically challenging for the ODE integrator, especially with large timesteps.  \MarginPar{add some info on why it fails}
 In the near future we will show how to include various nuclear

### 2023-06-20 subch_tols/paper.tex
add more figs
update

@@
 \begin{abstract}
-We explore the early evolution of flame ignition and spreading on the
-surface of a neutron star in three-dimensions, in the context of X-ray
-bursts.  We look at the nucleosynthesis and morphology of the burning
-front and compare to two-dimensional axisymmetric simulations to gauge
-how important a full three-dimensional treatment of the flame is for
-the early dynamics.  Finally, we discuss the progress toward full-star
-resolved flame simulations.\end{abstract}
+We explore the coupling of hydrodynamics and reactions in simulations of
+the double detonation model for Type Ia supernovae.
+\end{abstract}
 
-\keywords{convection---hydrodynamics---methods: numerical---stars: neutron---X-rays: bursts}
+\keywords{convection---hydrodynamics---methods: numerical}
 
@@
 
-Detonations can be difficult to model numerically.  Time step cuts \MarginPar{ref}.
-For the double detonation scenario, the ignition of the second detonation
-can be very sensitive to how the simulation is run \MarginPar{refs}.
-Artifically limiting the rates \MarginPar{ref}
+These events are modeled using the Euler equations with reactive and
+gravitational sources:
+\begin{eqnarray*}
+\ddt{\rho} + \nabla \cdot (\rho \Ub) &=& 0 \\
+\ddt{(\rho \Ub)} + \nabla \cdot (\rho \Ub \Ub) + \nabla p &=& \rho \gb \\
+\ddt{(\rho E)} + \nabla \cdot (\rho \Ub E + \Ub p) &=& \rho \Ub \cdot \gb + \rho \dot{S} \\
+\ddt{(\rho X_k)} + \nabla \cdot (\rho \Ub X_k) &=& \rho \omegadot_k
+\end{eqnarray*}
 
 Integration of reaction networks takes various forms in astrophysical
-simulation codes.
+simulation codes.  The most common method is via operator splitting:
+the advection and reactive terms are treated independently, with each
+process working on the result of the other.  Often, Strang-splitting
+\citep{strang:1968} is used, which alternates advection and reaction
+to yield second-order accuracy in time.
 
@@
 \citet{strang_rnaas}, this is needed to get second-order convergence.
+
+Detonations can be difficult to model numerically.  Time step cuts \MarginPar{ref}.
+For the double detonation scenario, the ignition of the second detonation
+can be very sensitive to how the simulation is run \MarginPar{refs}.
+Artifically limiting the rates \MarginPar{ref}
 
@@
 
-All simulations use the same network, a 22 nucleus network designed for
-He and C burning, using the ReacLib rates \cite{reaclib} and produced with
-\pynucastro \citep{pynucastro2}.  This network, called {\tt subch\_simple} was
-described in \cite{zhi2023}.
+Our default CFL number is 0.2 for these simulations, and the CFL constraint
+is the only timestep constraint used.  For larger CFL numbers, the ODE integration
+sometimes encounters problems \MarginPar{more}.
+
+All simulations use the same network, the {\tt subch\_simple} as
+described in \cite{zhi2023}.  This includes 22 nuclei and 94 rates
+(from ReacLib \citealt{reaclib}) and is produced with \pynucastro
+\citep{pynucastro2}.  Importantly, we include
+$\isotm{N}{14}(\alpha,\gamma)\isotm{F}{18}(\alpha,p)\isotm{Ne}{21}$
+which creates protons to allow for
+$\isotm{C}{12}(p,\gamma)\isotm{N}{13}(\alpha,p)\isotm{O}{16}$.  This
+sequence can be faster than
+$\isotm{C}{12}(\alpha,\gamma)\isotm{O}{16}$, as pointed out by
+\citet{shenbildsten} and is important for getting the detonation speed
+correctly.  To reduce the size of the network, some
+$(\alpha,p)(p,\gamma)$ and $(\alpha,\gamma)$ rates are combined into
+an effective $(\alpha,\gamma)$ rate.  Screening is provided following
+the procedure in \citet{wallace:1982}, combining the screening
+functions of \citet{graboske:1973,alastuey:1978,itoh:1979}.
 
@@
 \citep{vode} ported to C++.  Our modifications to this integrator are described
-in \citet{castro_simple_sdc}.
+in \citet{castro_simple_sdc}.  We use an analytic approximation to the Jacobian
+and found that we get best results when we disable the caching of the Jacobian
+in the integrator.
 
 If during the advance of a timestep an error is generated (negative
-density, ODE integration fails, mass fractions don't sum to 1), the
+density, ODE integration fails, mass fractions don't sum to 1, CFL constraint violated at the new time), the
 step is thrown-out and retied with a smaller timestep.  This is done
@@
 temperature ramps up to $1.75\times 10^8~\mathrm{K}$ and the
-composition changes to 99\% \isot{He}{4} and 1\% \isot{N}{14}.  This
-added \isot{N}{14} helps produce protons via
-$\isotm{N}{14}(\alpha,\gamma)\isotm{F}{18}(\alpha,p)\isotm{Ne}{21}$
-which then allow for
-$\isotm{C}{12}(p,\gamma)\isotm{N}{13}(\alpha,p)\isotm{O}{16}$.  This
-sequence can be faster than
-$\isotm{C}{12}(\alpha,\gamma)\isotm{O}{16}$, as pointed out by
-\citet{shenbildsten} and is important for getting the detonation speed
-correctly.  This He envelope is then integrated isentropically.  The
+composition changes to 99\% \isot{He}{4} and 1\% \isot{N}{14}.    This He envelope is then integrated isentropically.  The
 density of the transition from the underlying CO white dwarf to the He
@@
 
+Figure~\ref{fig:lap_rho_sequence} shows a Schlieren-style plot, which
+highlights density gradients.  It is constructed by plotting
+$\log_{10}(|\rho^{-1}\nabla^2\rho|)$.  This clearly shows the
+compression wave launched by the He detonation propagating in toward
+the center of the white dwarf.  The middle row shows this compression
+wave converging slightly off-center which then leads to the ignition
+of the C detonation.  This also highlights how the speed of the He
+detonation is important in determining how off-center the C detonation
+ignition will be.  Our choice of network was motivated by ensuring
+that we capture the energy release from He burning accurately, in
+particular the inclusion of the
+$\isotm{C}{12}(p,\gamma)\isotm{N}{13}(\alpha,p)\isotm{O}{16}$
+sequence.
+
 We also vary the CFL number and the integration tolerances.
@@
 
+\begin{figure}[t]
+\centering
+\plotone{subch_ni56}
+\caption{\label{fig:ni56} \isot{Ni}{56} mass vs.\ time for the different simulations.}
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\plotone{subch_cpu}
+\caption{\label{fig:cpu} MPI-hours of GPU time for the different simulations.}
+\end{figure}
+
 Show time-sequence + compression plot
 
-Show cost plot
+Figure~\ref{fig:cpu} shows the computational expense of the
+simulations, in terms of node hours.  All simulations were run on 4
+nodes (32 AMD GPUs total).  We note that we have not spent any time on
+load-balancing these simulations, so we should consider this to be
+only a guide.  At the start of the simulation, there are 56 boxes on
+the finest grid (representing 0.77\% of the domain) and at 1 s of
+evolution there is are 60 boxes covering 5.64\% of the domain.  The
+figure shows that the Strang simulation with a CFL of 0.2 is the
+cheapest, but as we saw above, this does not get the nucleosynthesis
+correct.  All of the other Strang simulations are more costly than the
+baseline SDC simulation.  This shows that not only does the
+simplified-SDC simulation accurately evolve the nucleosynthesis, it
+does so in a very efficient fashion.
 
@@
 
-The ignition of the carbon detonation remains numerically challenging for the ODE integrator, especially with large timesteps.  \MarginPar{add some info on why it fails}
-In the near future we will show how to include various nuclear
-statistical equilibrium approximations into the simplified-SDC
-formalism while retaining second-order accuracy.
+The ignition of the carbon detonation remains numerically challenging
+for the ODE integrator, especially with large
+timesteps.  \MarginPar{add some info on why it fails} In the near
+future we will show how to include various nuclear statistical
+equilibrium approximations into the simplified-SDC formalism while
+retaining second-order accuracy.  We will also explore different
+integrators and auto-differentiation for the Jacobian (in particular,
+the compositional contribution from the screening)
 
-We encourage authors to include a discussion about the reaction integration algorithm
-and tolerances in papers to permit comparisons in the future.
+We encourage authors to include a discussion about the reaction
+integration algorithm and tolerances in papers to permit comparisons
+in the future.
 

### 2023-06-20 xrb3/paper.tex

@@
 ${\bf g}$ is the gravitational
-acceleration, $\Omegab$ is the rotation vector, $k_\mathrm{th}$ is the thermal conductivity, and $T$ is the temperature.
-}
+acceleration, $\Omegab$ is the rotation vector, $k_\mathrm{th}$ is the thermal conductivity, $T$ is the temperature,} \edit2{and $\epsdot$ is the energy release from the reaction network}.
+
  We use a general stellar equation of state with
@@
 structure of these models is shown
-shown in Figure~\ref{fig:initial model}.}
+in Figure~\ref{fig:initial model}.}
 

### 2023-06-22 subch_tols/paper.tex
more updates

@@
-\documentclass[linenumbers,trackchanges]{aastex631}
+\documentclass[modern]{aastex631}
 
@@
 We explore the coupling of hydrodynamics and reactions in simulations of
-the double detonation model for Type Ia supernovae.
+the double detonation model for Type Ia supernovae.  When assessing the
+convergence of simulations, the focus is usually on spatial resolution.
+However, the method of coupling the physics together as well as the
+tolerances used in integrating a reaction network also play an important role.
+In this paper, we explore how the choices made in integrating the reaction
+portion of a simulation of the double detonation model of Type Ia supernovae
+influences the nucleosynthesis.  We find that the stronger coupling
+afforded by the simplified spectral deferred corrections method introduced
+previously provides a more accurate and efficient result than standard
+operator splitting.
 \end{abstract}
@@
 
-The double detonation model
+There are many different progenitor models for Type Ia supernova that
+are currently being explored.  The double detonation model
+\citep{fink:2007} involves a sub-Chandrasekhar mass white dwarf with a
+thin accreted He layer.  A detonation in the He layer propagates
+around the star while sending a compression wave converging toward the
+center.  This compression wave ignites a carbon detonation that then
+propagates outward from the center of the white dwarf.  Many groups
+have modeled this scenario, varying the initial conditions,
+resolution \citep{rivas:2022}, and reaction networks to understand the robustness of this
+model as well as its connection to observations.
+
+\citep{townsley:2019} \citep{gronow:2020}, \citep{gronow:2021}, \citep{glasner:2018}, \citep{roy:2022}, \citep{shenbildsten:2014},
 
@@
 \end{eqnarray*}
+where $\rho$ is the mass density, $\Ub$ is the velocity, $E$ is the specific
+total energy, $p$ is the pressure, and $X_k$ are the mass fractions
+of the nuclei that react.  The sources are the gravitational acceleration,
+$\gb$, which is obtained by solving the the Poisson equation
+\begin{equation}
+\nabla^2 \Phi = 4 \pi G \rho
+\end{equation}
+(where $G$ is Newton's constant) for the gravitational potential, $\Phi$, and defining $\gb = -\nabla \Phi$.
+The reaction network provides the species creation rate, $\omegadot$ and the energy generation rate, $\dot{S}$.
+Finally, the system is closed via the equation of state:
+\begin{equation}
+p = p(\rho, e, X_k)
+\end{equation}
+where $e$ is the specific internal energy, obtained as $e = E - |\Ub|^2/2$.
 
-Integration of reaction networks takes various forms in astrophysical
+Simulation codes use different methods to advance this system of equations.  For hydrodynamics,
+explicit-in-time integration is usually used, with the timestep, $\Delta t$, restricted by the Courant limit:
+\begin{equation}
+\Delta t = C \min_{i,j,k} \left \{ \frac{\Delta x}{|\Ub| + c} \right \}
+\end{equation}
+where $\Delta x$ is the width of a grid zone and $c$ is the sound
+speed, and the minimum is taken over all zones on the grid.  Here, the
+Courant number, $C$ is required to be $C < 1$ for stability.
+Reactions, on the other hand, tend to be stiff, and are usually evolved
+using implicit-in-time integration, with many small steps taken to
+make up the hydrodynamics timestep $\Delta t$.
+
+A key concern in modeling astrophysical reacting flows is the coupling
+of hydrodynamics and reactions.  If a large amount of energy is dumped
+into a zone, then a large flow will result to carry away this energy,
+requiring tight communication between the reaction and hydrodynamics
+solvers.  This coupling takes various forms in astrophysical
 simulation codes.  The most common method is via operator splitting:
@@
 \citep{strang:1968} is used, which alternates advection and reaction
-to yield second-order accuracy in time.
+to yield second-order accuracy in time.  The primary appeal of operator
+splitting is that it is easy to implement---the hydrodynamics and reaction
+solvers are largely independent of one another.
 
@@
 is no reaction source in the mass continuity equation.  The mass
-fractions, $X_k$, evolve according to:
-\begin{equation}
-\frac{dX_k}{dt} = \omegadot_k
-\end{equation}
+fractions, $X_k$, and energy evolve according to:
+\begin{eqnarray}
+\frac{dX_k}{dt} &= \omegadot_k \\
+\frac{de}{dt} &= \dot{S} \\
+\end{eqnarray}
 where $\omegadot_k$ is the creation rate of species $k$ due to nuclear
@@
 
+\subsection{Simulation setup}
+
 We use a 2d axisymmetric domain with a size of $5.12\times 10^9~\mathrm{cm}$ by $1.024\times 10^{10}~\mathrm{cm}$.  This is
@@
 described in \cite{zhi2023}.  This includes 22 nuclei and 94 rates
-(from ReacLib \citealt{reaclib}) and is produced with \pynucastro
+from ReacLib \citep{reaclib} and is produced with \pynucastro\
 \citep{pynucastro2}.  Importantly, we include
@@
 
-We run 3 different types of integration methods: traditional operator
-splitting, using Strang splitting, both with integration of the energy
-equation and without any temperature evolution during the burn.  \MarginPar{comment about the EOS}.  These methods are compared to the simplified-SDC integtation.
-
 \begin{figure*}[t]
@@
 \end{figure*}
+
+\subsection{Time-integration and Nucleosynthesis}
+
+We run 3 different types of integration methods: traditional operator
+splitting, using Strang splitting, both with integration of the energy
+equation and without any temperature evolution during the burn.  \MarginPar{comment about the EOS}.  These methods are compared to the simplified-SDC integtation.
 
@@
 
-We also vary the CFL number and the integration tolerances.
-
-Show Ni56 mass, intermediate mass
-
 \begin{figure}[t]
@@
 
+Figure~\ref{fig:ni56} shows the total amount of \isot{Ni}{56} as a function of time for
+a suite of simulations.
+
+We also vary the CFL number and the integration tolerances.
+
 \begin{figure}[t]
@@
 \plotone{subch_cpu}
-\caption{\label{fig:cpu} MPI-hours of GPU time for the different simulations.}
+\caption{\label{fig:cpu} node hours of GPU time for the different simulations.}
 \end{figure}
-
-Show time-sequence + compression plot
 
@@
 does so in a very efficient fashion.
+
+\subsection{Spatial resolution}
+
+\begin{figure}[t]
+\centering
+\plotone{subch_ni56_res}
+\caption{\label{fig:res} \isot{Ni}{56} mass vs.\ time for 3 different spatial resolutions, using the simplified-SDC time integration method.}
+\end{figure}
 
@@
 \url{http://github.com/AMReX-Astro/Castro}.  The problem setup used
-here is available in the git repo as {\tt flame\_wave}. The metadata
-describing the build environment and the global diagnostic output
-files are available on Zenodo at \citet{xrb_data}.  We thank Alice
-Harpole for contributions to the AMReX Astrophysics suite.
+here is available in the git repo as {\tt subchandra}.
 
@@
           NumPy \citep{numpy,numpy2},
+          pynucastro \citep{pynucastro,pynucastro2},
           python (\url{https://www.python.org/}),
+          SymPy \citep{sympy},
           valgrind \citep{valgrind},

### 2023-06-22 subch_tols/plots/plot_cpu.py

@@ -1,0 +1,32 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+sdc_files = [
+#    ("subch_sdc_cfl0.4/cpu_times.out", "SDC CFL=0.4", "C0", "+"),
+    ("subch_sdc/cpu_times.out", "SDC CFL=0.2", "k", None),
+    ]
+
+strang_files = [
+    ("subch_strang/cpu_times.out", "Strang CFL=0.2", "C0", "o"),
+#    ("subch_strang_cfl0.1/cpu_times.out", "Strang CFL=0.1", "C1", "x"),
+    ("subch_strang_noTevolve_fix/cpu_times.out", "Strang, CFL=0.2, no T evolution", "C2", "s"),
+    ("subch_strang_cfl0.05/cpu_times.out", "Strang CFL=0.05", "C0", "+"),
+    ("subch_strang_tol1e-6/cpu_times.out", "Strang CFL=0.2, tols=1.e-6", "C1", "^"),
+    ("subch_strang_tol1.e-8/cpu_times.out", "Strang CFL=0.2, atol_spec=1.e-8", "C1", "x")]
+
+fig, ax = plt.subplots()
+
+for diag_file, label, color, ls in sdc_files + strang_files:
+    data = np.genfromtxt("subch_diags/" + diag_file)
+
+    ax.plot(data[:,1], data[:,2]/8, label=label, color=color, marker=ls, markevery=0.025)
+
+ax.legend(fontsize="small")
+ax.set_xlim(0.0, 1.0)
+
+ax.grid(linestyle=":", which="both")
+
+ax.set_xlabel("time (s)")
+ax.set_ylabel("node hours")
+fig.set_size_inches((8, 8))
+fig.savefig("subch_cpu.pdf")

### 2023-06-22 subch_tols/plots/plot_subch_intermediate.py

@@ -1,0 +1,44 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+sdc_files = [
+#    ("subch_sdc_cfl0.4/species_diag.out", "SDC CFL=0.4", "C0", "+"),
+    ("subch_sdc/species_diag.out", "SDC CFL=0.2", "k", None),
+    ]
+
+strang_files = [
+    ("subch_strang/species_diag.out", "Strang CFL=0.2", "C0", "o"),
+#    ("subch_strang_cfl0.1/species_diag.out", "Strang CFL=0.1", "C1", "x"),
+    ("subch_strang_cfl0.05/species_diag.out", "Strang CFL=0.05", "C0", "+"),
+    ("subch_strang_tol1e-6/species_diag.out", "Strang CFL=0.2, tols=1.e-6", "C1", "^"),
+    ("subch_strang_tol1.e-8/species_diag.out", "Strang CFL=0.2, atol_spec=1.e-8", "C1", "x")]
+
+fig, ax = plt.subplots()
+
+def monotonize(arr):
+    new_rows = []
+    max_t = -1
+    for row in arr:
+        t = row[0]
+        if t > max_t:
+            new_rows.append(row)
+        max_t = max(t, max_t)
+
+    return np.asarray(new_rows)
+
+for diag_file, label, color, ls in sdc_files + strang_files:
+    data = monotonize(np.loadtxt("subch_diags/" + diag_file))
+
+    mass = data[:,16] + data[:,17] + data[:,18] + data[:,19] + data[:,20] + data[:,21]
+    ax.plot(data[:,1], mass, label=label, color=color, marker=ls, markevery=0.025)
+
+ax.legend(fontsize="small")
+ax.set_xlim(0.0, 1.0)
+ax.set_ylim(1.e-3, 1.0)
+ax.set_yscale("log")
+
+ax.set_xlabel("time (s)")
+ax.set_ylabel("intermediate nuclei mass (solar masses)")
+
+fig.set_size_inches((8, 8))
+fig.savefig("subch_intermediate.pdf")

### 2023-06-22 subch_tols/plots/plot_subch_ni.py

@@ -1,0 +1,46 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+sdc_files = [
+#    ("subch_sdc_cfl0.4/species_diag.out", "SDC CFL=0.4", "C0", "+"),
+    ("subch_sdc/species_diag.out", "SDC CFL=0.2", "k", None),
+    ]
+
+strang_files = [
+    ("subch_strang/species_diag.out", "Strang CFL=0.2", "C0", "o"),
+#    ("subch_strang_cfl0.1/species_diag.out", "Strang CFL=0.1", "C1", "x"),
+    ("subch_strang_noTevolve_fix/species_diag.out", "Strang, CFL=0.2, no T evolution", "C2", "s"),
+    ("subch_strang_cfl0.05/species_diag.out", "Strang CFL=0.05", "C0", "+"),
+    ("subch_strang_tol1e-6/species_diag.out", "Strang CFL=0.2, tols=1.e-6", "C1", "^"),
+    ("subch_strang_tol1.e-8/species_diag.out", "Strang CFL=0.2, atol_spec=1.e-8", "C1", "x")]
+
+fig, ax = plt.subplots()
+
+def monotonize(arr):
+    new_rows = []
+    max_t = -1
+    for row in arr:
+        t = row[0]
+        if t > max_t:
+            new_rows.append(row)
+        max_t = max(t, max_t)
+
+    return np.asarray(new_rows)
+
+for diag_file, label, color, ls in sdc_files + strang_files:
+    data = monotonize(np.loadtxt("subch_diags/" + diag_file))
+
+    ax.plot(data[:,1], data[:,23], label=label, color=color, marker=ls, markevery=0.025)
+
+ax.legend(fontsize="small")
+ax.set_xlim(0.0, 1.0)
+ax.set_ylim(1.e-2, 1.0)
+ax.set_yscale("log")
+
+ax.grid(linestyle=":", which="both")
+
+ax.set_xlabel("time (s)")
+ax.set_ylabel("Ni56 mass (solar masses)")
+
+fig.set_size_inches((8, 8))
+fig.savefig("subch_ni56.pdf")

### 2023-06-22 subch_tols/plots/plot_subch_ni_resolution.py

@@ -1,0 +1,40 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+sdc_files = [
+    ("subch_sdc_40km/species_diag.out", "SDC 40 km"), 
+    ("subch_sdc/species_diag.out", "SDC 20 km"), 
+    ("subch_sdc_10km_3lev/species_diag.out", "SDC 10 km"), 
+    ]
+
+
+fig, ax = plt.subplots()
+
+def monotonize(arr):
+    new_rows = []
+    max_t = -1
+    for row in arr:
+        t = row[0]
+        if t > max_t:
+            new_rows.append(row)
+        max_t = max(t, max_t)
+
+    return np.asarray(new_rows)
+
+for diag_file, label in sdc_files:
+    data = monotonize(np.loadtxt("subch_diags/" + diag_file))
+
+    ax.plot(data[:,1], data[:,23], label=label)
+
+ax.legend(fontsize="small")
+ax.set_xlim(0.0, 1.0)
+ax.set_ylim(1.e-2, 1.0)
+ax.set_yscale("log")
+
+ax.grid(linestyle=":", which="both")
+
+ax.set_xlabel("time (s)")
+ax.set_ylabel("Ni56 mass (solar masses)")
+
+fig.set_size_inches((8, 8))
+fig.savefig("subch_ni56_res.pdf")

### 2023-06-23 subch_tols/paper.tex
more work

@@
 
-\subsection{Time-integration and Nucleosynthesis}
-
-We run 3 different types of integration methods: traditional operator
-splitting, using Strang splitting, both with integration of the energy
-equation and without any temperature evolution during the burn.  \MarginPar{comment about the EOS}.  These methods are compared to the simplified-SDC integtation.
-
 Figure~\ref{fig:lap_rho_sequence} shows a Schlieren-style plot, which
@@
 
+\subsection{Time-integration and Nucleosynthesis}
+
+We run 3 different types of integration methods: traditional operator
+splitting, using Strang splitting, both with integration of the energy
+equation and without any temperature evolution during the
+burn.  \MarginPar{comment about the EOS}.  These methods are compared
+to the simplified-SDC integtation.  We also vary the CFL number and
+the integration tolerances.
+
 \begin{figure}[t]
@@
 
-We also vary the CFL number and the integration tolerances.
+To understand the difference in the \isot{Ni}{56} mass between the SDC and Strang simulations at 
+CFL = 0.2,
+
+\begin{figure}[t]
+\centering
+\plotone{sdc_strang_compare_abar_summary_plot}
+\caption{\label{fig:abar_sdc_strang} Comparison of the mean molecular weight between SDC and Strang (both run with CFL = 0.2) at $t = 0.2~\mathrm{s}$.}
+\end{figure}
 

### 2023-06-23 subch_tols/plots/plot_subch_ni_resolution.py
add inset

@@ -1,5 +1,7 @@
 import matplotlib.pyplot as plt
 import numpy as np
+from mpl_toolkits.axes_grid1.inset_locator import (InsetPosition, inset_axes,
+                                                   mark_inset)
 
 sdc_files = [
     ("subch_sdc_40km/species_diag.out", "SDC 40 km"), 
@@ -35,6 +37,25 @@ ax.grid(linestyle=":", which="both")
 
 ax.set_xlabel("time (s)")
 ax.set_ylabel("Ni56 mass (solar masses)")
+
+# Create a set of inset Axes: these should fill the bounding box allocated to
+# them.
+ax2 = plt.axes([0,0,1,1])
+# Manually set the position and relative size of the inset axes within ax1
+ip = InsetPosition(ax, [0.15,0.25,0.35,0.35])
+ax2.set_axes_locator(ip)
+# Mark the region corresponding to the inset axes on ax1 and draw lines
+# in grey linking the two axes.
+mark_inset(ax, ax2, loc1=1, loc2=3, fc="none", ec='0.5')
+
+for diag_file, label in sdc_files:
+    data = monotonize(np.loadtxt("subch_diags/" + diag_file))
+
+    ax2.plot(data[:,1], data[:,23], label=label)
+
+ax2.set_xlim(0.77, 0.7975)
+ax2.set_yscale("log")
+ax2.set_ylim(0.0175, 0.04)
 
 fig.set_size_inches((8, 8))
 fig.savefig("subch_ni56_res.pdf")

### 2023-06-23 subch_tols/plots/plot_subch_ni.py
add inset

@@ -1,5 +1,7 @@
 import matplotlib.pyplot as plt
 import numpy as np
+from mpl_toolkits.axes_grid1.inset_locator import (InsetPosition, inset_axes,
+                                                   mark_inset)
 
 sdc_files = [
 #    ("subch_sdc_cfl0.4/species_diag.out", "SDC CFL=0.4", "C0", "+"),
@@ -41,6 +43,26 @@ ax.grid(linestyle=":", which="both")
 
 ax.set_xlabel("time (s)")
 ax.set_ylabel("Ni56 mass (solar masses)")
+
+# Create a set of inset Axes: these should fill the bounding box allocated to
+# them.
+ax2 = plt.axes([0,0,1,1])
+# Manually set the position and relative size of the inset axes within ax1
+ip = InsetPosition(ax, [0.15,0.30,0.4,0.35])
+ax2.set_axes_locator(ip)
+# Mark the region corresponding to the inset axes on ax1 and draw lines
+# in grey linking the two axes.
+mark_inset(ax, ax2, loc1=1, loc2=3, fc="none", ec='0.5')
+
+for diag_file, label, color, ls in sdc_files + strang_files:
+    data = monotonize(np.loadtxt("subch_diags/" + diag_file))
+
+    ax2.plot(data[:,1], data[:,23], label=label, color=color, marker=ls, markevery=0.15)
+
+ax2.set_xlim(0.73, 0.7875)
+ax2.set_yscale("log")
+ax2.set_ylim(0.0175, 0.04)
+
 
 fig.set_size_inches((8, 8))
 fig.savefig("subch_ni56.pdf")

### 2023-06-25 subch_tols/paper.tex
add plot
update paper

@@
 \begin{abstract}
-We explore the coupling of hydrodynamics and reactions in simulations of
+We study the coupling of hydrodynamics and reactions in simulations of
 the double detonation model for Type Ia supernovae.  When assessing the
-convergence of simulations, the focus is usually on spatial resolution.
-However, the method of coupling the physics together as well as the
+convergence of simulations, the focus is usually on spatial resolution, 
+however, the method of coupling the physics together as well as the
 tolerances used in integrating a reaction network also play an important role.
-In this paper, we explore how the choices made in integrating the reaction
-portion of a simulation of the double detonation model of Type Ia supernovae
-influences the nucleosynthesis.  We find that the stronger coupling
+In this paper, we explore how the choices made in both coupling and integrating the reaction
+portion of a simulation influences the accuracy, efficiency, and the nucleosynthesis
+of simulations of the double detonation model of Type Ia supernovae.
+We find that the stronger coupling between hydrodynamics and reactions,
 afforded by the simplified spectral deferred corrections method introduced
-previously provides a more accurate and efficient result than standard
-operator splitting.
+previousl,y provides a more accurate and efficient result than standard
+operator splitting.  We also find no need to limit reactions rates or reduce the
+simulation timestep to the reaction timescale.  The entire simulation
+methodology used here is freely available as part of the \castro\ simulation
+code.
 \end{abstract}
@@
 
-There are many different progenitor models for Type Ia supernova that
-are currently being explored.  The double detonation model
+There are many different progenitor models for Type Ia supernova (SN Ia) that
+are currently being explored to explain the observed diversity of SN Ia explosions \cite{taubenberger:2017}.  The double detonation model
 \citep{fink:2007} involves a sub-Chandrasekhar mass white dwarf with a
@@
 
-\citep{townsley:2019} \citep{gronow:2020}, \citep{gronow:2021}, \citep{glasner:2018}, \citep{roy:2022}, \citep{shenbildsten:2014},
+\citep{townsley:2019} \citep{gronow:2020}, \citep{gronow:2021}, \citep{glasner:2018}, \citep{roy:2022}, \citep{shenbildsten:2014}, \cite{polin:2019}, \citep{kushnirkatz:2020}
 
@@
 gravitational sources:
-\begin{eqnarray*}
+\begin{eqnarray}
 \ddt{\rho} + \nabla \cdot (\rho \Ub) &=& 0 \\
@@
 \ddt{(\rho X_k)} + \nabla \cdot (\rho \Ub X_k) &=& \rho \omegadot_k
-\end{eqnarray*}
+\end{eqnarray}
 where $\rho$ is the mass density, $\Ub$ is the velocity, $E$ is the specific
@@
 Simulation codes use different methods to advance this system of equations.  For hydrodynamics,
-explicit-in-time integration is usually used, with the timestep, $\Delta t$, restricted by the Courant limit:
+explicit-in-time integration is usually used, with the timestep, $\Delta t$, restricted by the Courant limit.  In two-dimensions, this is:
 \begin{equation}
-\Delta t = C \min_{i,j,k} \left \{ \frac{\Delta x}{|\Ub| + c} \right \}
+\Delta t = C \min_{i,j} \left \{ \frac{\Delta x}{|u_{i,j}| + c_{i,j}}, \frac{\Delta y}{|v_{i,j}| + c_{i,j}}  \right \}
 \end{equation}
-where $\Delta x$ is the width of a grid zone and $c$ is the sound
+where $\Delta x$ and $\Delta y$ are the grid width in the $x$- and $y$-directions, $u$ and $v$
+are the $x$- and $y$-components of the velocity and $c$ is the sound
 speed, and the minimum is taken over all zones on the grid.  Here, the
-Courant number, $C$ is required to be $C < 1$ for stability.
+Courant number, $C$ is required to be $C < 1$ for stability.  We note that
+this is the form of the stability criterion for the corner transport upwind
+formulation \citep{ppmunsplit} of hydrodynamics we use.  Dimensionally-split
+or method-of-lines formulations will have a slighly more restrictive criteria.
 Reactions, on the other hand, tend to be stiff, and are usually evolved
@@
 \begin{eqnarray}
-\frac{dX_k}{dt} &= \omegadot_k \\
-\frac{de}{dt} &= \dot{S} \\
+\frac{dX_k}{dt} &= \omegadot_k(\rho, T, X_k) \label{eq:stranga} \\
+\frac{de}{dt} &= \dot{S}(\rho, T, X_k) \\
+T &= T(\rho, e, X_k) \label{eq:strangc}
 \end{eqnarray}
-where $\omegadot_k$ is the creation rate of species $k$ due to nuclear
-reaction.  In the simplest approximation of operator splitting, the
+In the simplest approximation of operator splitting, the
 evolution of the mass fractions is solved alone, without integrating
-the temperature or energy.  This is the form that was used in FLASH
-\cite{flash}.  A more accurate operator splitting also includes the
-energy evolution (perhaps in terms of temperature).  This will require
+the temperature or energy.  This is the form that was used, e.g., in FLASH
+\citep{flash}.  A more accurate operator splitting also includes the
+energy evolution (perhaps in terms of temperature), solving the system Eqs.~\ref{eq:stranga}--\ref{eq:strangc}.  This will require
 calling the equation of state each time we evaluate the righthand side
-of the ODE system, increasing the cost.  However, as we showed in
+of the ODE system, potentially increasing the computational expense.  However, as we showed in
 \citet{strang_rnaas}, this is needed to get second-order convergence.
 
-Detonations can be difficult to model numerically.  Time step cuts \MarginPar{ref}.
-For the double detonation scenario, the ignition of the second detonation
-can be very sensitive to how the simulation is run \MarginPar{refs}.
-Artifically limiting the rates \MarginPar{ref}
+Detonations can be difficult to model numerically.  One common
+approach is to use a simultion timestep based on the reaction
+timescale, as discussed in \citet{prometheus}, instead of the
+hydrodynamics timescale.  For the double detonation scenario, the
+ignition of the second detonation can be very sensitive to how the
+simulation is run \MarginPar{refs}.  Artifically limiting the
+rates \MarginPar{ref}
 
@@
 \begin{equation}
+\label{eq:simplesdc}
 \ddt{\Uc} = \Rb(\Uc) + \Adv{\Uc}^{n+1/2}
 \end{equation}
+Here, $\Uc$ is the conserved state, $\Uc = (\rho, \rho \Ub, \rho E,
+\rho X_k)^\intercal$, $\Rb(\Uc)$ are the reactive sources, and
+$\Adv{\Uc}^{n+1/2}$ is an approximation to the advective update over
+the timestep, as computed by the corner transport upwind formulation
+of the piecewise parabolic method
+\citep{ppmunsplit,millercolella:2002,ppm}. Crucially, this advective
+term includes an explicit reaction source term and the update,
+Eq.~\ref{eq:simplesdc} means that the reactions will ``see'' what
+advection is doing as the reaction network is integrated.
 
 In addition to the time-integration strategy, integrating the reaction
-ODE system requires specifying tolerances.  Unfortunately, the tolerances
-used are not normally reported in papers, and in many cases, may not
-be easily controlled by a code user at runtime.
+ODE system requires specifying tolerances, usually both an absolute, $\epsilon_\mathrm{abs}$, and relative tolerance, $\epsilon_\mathrm{rel}$, that are combined together into a weight of the form:
+\begin{equation}
+w_i = \epsilon_{\mathrm{abs},i} + \epsilon_{\mathrm{rel},i} |y_i|
+\end{equation}
+where the $y_i$ is one of the variables being integrated by the ODE
+integrator and $i$ is the index, indicating that different tolerances
+can be used for each variable.  Traditionally one set of tolerances is
+used for the mass fractions and another for the energy (if it is
+integrated).  Unfortunately, the tolerances used by simulations are
+not normally reported in papers, and in many cases, may not be easily
+controlled by a code user at runtime (and therefore, users may not be
+aware of what they are using).  Some examples can be found in publicly
+available networks.  For example, the freely available version of the
+{\tt aprox13} network uses $\epsilon_\mathrm{rel} = 10^{-5}$ and a
+floor on the mass fractions corresponding to $\epsilon_\mathrm{abs} =
+10^{-6}$.  Since the simplified-SDC update works in terms of the
+conserved variables, $\Uc$, we apply a density weighting to the absolute tolerances when comparing to the values used in Strang integration.
 
@@
 \centering
+\plotone{subch_Temp_res_compare}
+\caption{\label{fig:res_panel} Temperature evolution for SDC simulations with CFL=0.2 at 40, 20, and 10 km.  For the earliest time
+  in each resolution, the AMR grid structure is shown.}
+\end{figure}
+
+\begin{figure}[t]
+\centering
 \plotone{subch_ni56_res}
@@
           matplotlib (\citealt{Hunter:2007}, \url{http://matplotlib.org/}),
+          NetworkX \citep{networkx},
           NumPy \citep{numpy,numpy2},

### 2023-06-25 subch_tols/paper.tex
more work

@@
 
-\citet{castro_simple_sdc} looked at an extreme version of the
-double detonation scenario---a very large perturbation was applied
-to drive a detonation directly into the underlying
-carbon-oxygen white dwarf.  We showed in that paper that the Strang
-split integration method had difficulty with the integration unless
-we used tighter tolerances
-
-Summary of the SDC subch results form the original paper
+We considered an extreme version of the double detonation scenario in
+\citet{castro_simple_sdc}---a very large perturbation was applied to
+drive a detonation directly into the underlying carbon-oxygen white
+dwarf.  We showed that the Strang split integration method had
+difficulty with the integration unless we used tighter tolerances,
+while the simplified-SDC method was well-behaved.  Here we look at a
+more realistic setup for a double detonation, with a small
+perturbation that drives a He detonation around the star and a
+compression wave that results in the ignition of a C detonation near
+the center of the star.  We focus on the nucleosynthesis and the
+efficiency of the method.
 
@@
 
-We use a 2d axisymmetric domain with a size of $5.12\times 10^9~\mathrm{cm}$ by $1.024\times 10^{10}~\mathrm{cm}$.  This is
-much larger than the initial size of the star, giving it plenty of
-room to expand.  The coarse grid is $640\times 1280$ zones and we use
-2 levels of refinement, giving a maximum resolution of
-$20~\mathrm{km}$.  The refinement strategy is picked to refine on the
-star and any regions where the temperature is greater than
-$10^8~\mathrm{K}$.  \castro\ uses subcycling, so finer grids are
-advanced at a smaller timestep than the coarse grids.  Self-gravity is
-done using a full Poisson solve using geometric multigrid, with
-Dirichlet boundary conditions on the domain boundary computed via a
-multipole expansion with a maximum order of 6.  For the low density
-regions outside of the star, we use a sponge term on the momentum
-equation to prevent the very low density material that is not in
-hydrostatic equilibrium from raining down on the star.
+We use the freely-available \castro\ simulation code for all the
+simulations shown here.  Our domain uses a 2D axisymmetric geomety
+with a size of $5.12\times 10^9~\mathrm{cm}$ by $1.024\times
+10^{10}~\mathrm{cm}$.  This is much larger than the initial size of
+the star, giving it plenty of room to expand (we note that the figures
+shown in the results section are all zoomed in on the star).  The
+coarse grid is $640\times 1280$ zones and we use 2 levels of
+refinement, each a jump of $2\times$, giving a maximum resolution of
+$20~\mathrm{km}$.  \castro\ uses the \amrex\ library
+\citep{amrex_joss} to manage the grid hierarchy.  The refinement
+strategy is picked to refine on the star and any regions where the
+temperature is greater than $10^8~\mathrm{K}$.  Finally, \castro\ uses
+subcycling, so finer grids are advanced at a smaller timestep than the
+coarse grids.
 
-Our default CFL number is 0.2 for these simulations, and the CFL constraint
-is the only timestep constraint used.  For larger CFL numbers, the ODE integration
-sometimes encounters problems \MarginPar{more}.
+As noted above, we use the unsplit PPM algorithm for advection, with
+the Riemann solver from \citet{colellaglaz:1985} and the general
+stellar equation of state of \citet{timmes_swesty:2000}.  In
+all simulations, the CFL constraint is the
+only timestep constraint used.  Self-gravity is done using a full
+Poisson solve using geometric multigrid, with Dirichlet boundary
+conditions on the domain boundary computed via a multipole expansion
+with a maximum order of 6.  For the low density regions outside of the
+star, we use a sponge term on the momentum equation to prevent the
+very low density material that is not in hydrostatic equilibrium from
+raining down on the star.
 
-All simulations use the same network, the {\tt subch\_simple} as
-described in \cite{zhi2023}.  This includes 22 nuclei and 94 rates
-from ReacLib \citep{reaclib} and is produced with \pynucastro\
-\citep{pynucastro2}.  Importantly, we include
+All simulations use the same reaction network: the {\tt subch\_simple}
+as described in \cite{zhi2023}.  This includes 22 nuclei and 94 rates
+from ReacLib \citep{reaclib} and is produced with
+\pynucastro\ \citep{pynucastro2}.  \MarginPar{show I show the
+  network?} Importantly, we include
 $\isotm{N}{14}(\alpha,\gamma)\isotm{F}{18}(\alpha,p)\isotm{Ne}{21}$
@@
 \citet{shenbildsten} and is important for getting the detonation speed
-correctly.  To reduce the size of the network, some
-$(\alpha,p)(p,\gamma)$ and $(\alpha,\gamma)$ rates are combined into
-an effective $(\alpha,\gamma)$ rate.  Screening is provided following
-the procedure in \citet{wallace:1982}, combining the screening
-functions of \citet{graboske:1973,alastuey:1978,itoh:1979}.
+correctly.  To reduce the size of the network, we have
+\pynucastro\ combine some $(\alpha,p)(p,\gamma)$ and $(\alpha,\gamma)$
+rates into an effective $(\alpha,\gamma)$ rate.  Screening is provided
+following the procedure in \citet{wallace:1982}, combining the
+screening functions of \citet{graboske:1973,alastuey:1978,itoh:1979}.
 
-The reaction system is integrated using a version of the VODE ODE integrator
-\citep{vode} ported to C++.  Our modifications to this integrator are described
-in \citet{castro_simple_sdc}.  We use an analytic approximation to the Jacobian
-and found that we get best results when we disable the caching of the Jacobian
-in the integrator.
+The reaction system is integrated using a version of the VODE ODE
+integrator \citep{vode} ported to C++.  Our modifications to this
+integrator are described in \citet{castro_simple_sdc}.  We use an
+analytic approximation to the Jacobian (the main approximation is that
+the species derivatives of the screening function are not included)
+and found that we get best results when we disable the caching of the
+Jacobian in the integrator.
 
-If during the advance of a timestep an error is generated (negative
-density, ODE integration fails, mass fractions don't sum to 1, CFL constraint violated at the new time), the
-step is thrown-out and retied with a smaller timestep.  This is done
-on a level-by-level basis in the overall AMR subcycling hierarchy.
+If during the advance of a timestep an
+error is generated (negative density, ODE integration fails, mass
+fractions don't sum to 1, CFL constraint violated at the new time),
+the step is thrown-out and retied with a smaller timestep.  This is
+done on a level-by-level basis in the overall AMR subcycling
+hierarchy.
 
-We will run a suite of simulations all using the same initial model.
-The initial model is constructed following the methodology in
-\citet{subchandra}.  We use an isothermal core of C/O of $1.1~M_\odot$
-with $T = 10^7~\mathrm{K}$ and a thin transition region where the
-temperature ramps up to $1.75\times 10^8~\mathrm{K}$ and the
-composition changes to 99\% \isot{He}{4} and 1\% \isot{N}{14}.    This He envelope is then integrated isentropically.  The
-density of the transition from the underlying CO white dwarf to the He
-envelope was selected to yield an envelope mass of $0.05~M_\odot$.
+We will run a suite of simulations all using the same initial model,
+constructed following the methodology in \citet{subchandra}.  We use
+an isothermal core of C/O of $1.1~M_\odot$ with $T = 10^7~\mathrm{K}$
+and a thin transition region where the temperature ramps up to
+$1.75\times 10^8~\mathrm{K}$ and the composition changes to 99\%
+\isot{He}{4} and 1\% \isot{N}{14}.  This He envelope is then
+integrated isentropically.  The density of the transition from the
+underlying CO white dwarf to the He envelope was selected to yield an
+envelope mass of $0.05~M_\odot$.  The code for generating the model is
+freely available at
+\url{https://github.com/amrex-astro/initial_models}.  \MarginPar{should
+  we do a zenodo release of this?}
 
@@
 \begin{equation}
-  T = T_0 \left [ 1 + X(\isotm{He}{4}) f (1 + \tanh(2 - r_1)) \right ]
+  T = T_0 \left \{ 1 + X(\isotm{He}{4}) f \left [1 + \tanh(2 - \sigma) \right ] \right \}
 \end{equation}
@@
 \begin{equation}
-  r_1 = \left [ x^2 + (y - r_0)^2 \right ]^{1/2} / \lambda
+  \sigma = \left [ x^2 + (y - R_0)^2 \right ]^{1/2} / \lambda
 \end{equation}
@@
 \begin{equation}
-  r_0 = r_\mathrm{pert} + r_\mathrm{base}
+  R_0 = R_\mathrm{pert} + R_\mathrm{base}
 \end{equation}
-Here, $r_\mathrm{base}$ is the radius at which the helium layer begins
-and $r_\mathrm{pert}$ is the distance above the base to put the
-perturbation.  We choose $r_\mathrm{pert} = 100~\mathrm{km}$.  The
+where, $R_\mathrm{base}$ is the radius at which the helium layer
+begins and $R_\mathrm{pert}$ is the distance above the base to put the
+perturbation.  We choose $R_\mathrm{pert} = 100~\mathrm{km}$.  The
 temperature is perturbed above the initial model value, denoted as
 $T_0$ here.  The amplitude of the perturbation is $f = 3$ and the
-scale of the perturbation is $\lambda = 12.5~\mathrm{km}$, and
-$r_\mathrm{pert} = 100~\mathrm{km}$.  This is a very small
-perturbation, but enough to seed the initial He detonation in the
-envelope.
+scale of the perturbation is $\lambda = 12.5~\mathrm{km}$.  This is a
+very small perturbation, but enough to seed the initial He detonation
+in the envelope.  \MarginPar{give the peak temperature}
+
+Our base simulation uses the simplified-SDC integration, a CFL number
+of $0.2$, and sets $\epsilon_\mathrm{rel} = \epsilon_\mathrm{abs} =
+10^{-5}$ for both the species and energy.  Overall, the simulation
+requires 21,784 coarse grid timesteps to evolve to $1.0~\mathrm{s}$ of
+simulation time.
 
@@
 the simulation and all computation is done there.  Our GPU offloading
-strategy takes advantage of the \amrex\ C++ lambda-capturing
-functionality to be performance portable.
+strategy \citep{castro_gpu} takes advantage of the \amrex\ C++
+lambda-capturing functionality to be performance portable.
+All of the simulation code is on github at
+\url{https://github.com/amrex-astro/} and the inputs files, global
+diagnostics used to make the line plots shown below, and metadata
+describing the git hashes, compiler environment, and runtime
+parameters for all of the simulaitons is available at \MarginPar{add
+  Zenodo ref}.
 
@@
 
+Figure~\ref{fig:temp_sequence} shows the evolution of the temperature field for the base simulation.
+The initial perturbation is placed at the north pole at $t=0$ and leads to a detonation in the He layer that begins to wrap around the star.
+By $0.1~\mathrm{s}$ the detonation is almost halfway around the star and by $0.2~\mathrm{s}$ is almost at the south pole.
+Figure~\ref{fig:lapl_rho_sequence} shows shows a Schlieren-style plot, which
+highlights density gradients.  It is constructed by plotting
+$\log_{10}(|\rho^{-1}\nabla^2\rho|)$.  This clearly shows a
+compression wave launched by the early evolution of the He detonation propagating inward toward
+the center of the white dwarf.
+
 \begin{figure*}[t]
@@
 
-Figure~\ref{fig:lap_rho_sequence} shows a Schlieren-style plot, which
-highlights density gradients.  It is constructed by plotting
-$\log_{10}(|\rho^{-1}\nabla^2\rho|)$.  This clearly shows the
-compression wave launched by the He detonation propagating in toward
-the center of the white dwarf.  The middle row shows this compression
-wave converging slightly off-center which then leads to the ignition
-of the C detonation.  This also highlights how the speed of the He
-detonation is important in determining how off-center the C detonation
-ignition will be.  Our choice of network was motivated by ensuring
-that we capture the energy release from He burning accurately, in
-particular the inclusion of the
+Between $t = 0.3$ and $t = 0.7~\mathrm{s}$ the burning is confined to
+the outer layer, but the compression wave continues to advance inward,
+converging slightly off-center at around $t = 0.7~\mathrm{s}$.  This
+highlights how the speed of the He detonation is important in
+determining how off-center the C detonation ignition will be.  Our
+choice of network was motivated by ensuring that we capture the energy
+release from He burning accurately, in particular the inclusion of the
 $\isotm{C}{12}(p,\gamma)\isotm{N}{13}(\alpha,p)\isotm{O}{16}$
@@
 
+In the bottom row of Figure~\ref{fig:temp_sequence}, we see
+the second detonation ignite near the center of the star and
+begin to propagate outward.  By $1.0~\mathrm{s}$ it has evolved
+past the initial radius of the star and is into the extended
+ash layer left behind by the previous He detonation.
+
 \subsection{Time-integration and Nucleosynthesis}
+
+We now look at how Strang integration to this base simulation.
 

### 2023-06-25 subch_tols/paper.tex
more updates

@@
 Stony Brook, NY 11794-3800, USA}
+
+\author{Melissa Rasmussen}
+\affiliation{Stony Brook? Utah?}
+
+\author{Abigail Polin}
+\affiliation{Caltech}
 
@@
 scale of the perturbation is $\lambda = 12.5~\mathrm{km}$.  This is a
-very small perturbation, but enough to seed the initial He detonation
-in the envelope.  \MarginPar{give the peak temperature}
+very small perturbation, with a peak temperature of about $1.1\times
+10^9~\mathrm{K}$, but enough to seed the initial He detonation in the
+envelope.
 
@@
 
+Figure~\ref{fig:res_panel} shows a time sequence of the simplified-SDC
+simulation with CFL = 0.2 and our standard tolerances
+($\epsilon_\mathrm{abs} = \epsilon_\mathrm{rel} = 10^{-5}$ at 3
+diferent spatial resolutions: 40 km, 20 km, and 10 km.  For the 40 km
+run, we simply use one fewer levels of refinement compared to our base
+simulation.  For the 10 km run, we add an additional level of
+refinement that captures only the high temperature regions (including
+the initial He layer) but with the very center always refined by
+adding an additional refinement tagging criteria on densities $\rho >
+7.5\times 10^7~\gcc$.  The first column of the figure shows the
+initial grid structure---darker colors are higher levels of
+refinement.  Each of the boxes shown is a multiple of 32 zones in each
+dimension, with a maximum of 256 zones in any dimension.  This
+refinement for the highest resolution run ensures that we capture the
+convergence of the pressure wave at the center.
+
 \begin{figure}[t]
@@
 
+As we see from the time-evolution, the three different resolutions all
+behave largely the same.  The timescale for the He detonaiton to
+propagate around the star and the timescale at which the second C
+detonation ignite match well.  Figure~\ref{fig:res} shows the
+\isot{Ni}{56} mass produced from the three different simulations.
+Again, the agreement is quite strong.
+
 \begin{figure}[t]
@@
 \end{figure}
+
+We note that these resolution results differ from the conclusions of
+\citet{rivas:2022}.  There are a number of reasons why this might be
+the case.  First the initial models differ \MarginPar{details}.
+Second, we use a much larger reaction network here that better
+captures the energetics.  This means that the He detonation moves
+faster and the subsequent convergence of the pressure wave will be
+more on-center, meaning that the ignition of the C detonation occurs
+at a higher temperature.  Finally, we are using the simplified-SDC time-integration, which that work used Strang splitting \MarginPar{do they give details about e evolution?}
 

### 2023-06-27 subch_tols/plots/plot_cpu.py
update plots

@@ -2,17 +2,18 @@ import matplotlib.pyplot as plt
 import numpy as np
 
 sdc_files = [
-#    ("subch_sdc_cfl0.4/cpu_times.out", "SDC CFL=0.4", "C0", "+"),
-    ("subch_sdc/cpu_times.out", "SDC CFL=0.2", "k", None),
+#    ("subch_sdc_cfl0.4/species_diag.out", "SDC CFL=0.4", "C0", "+"),
+    ("subch_sdc/species_diag.out", "SDC-0.2", "k", None),
+    ("subch_sdc_cfl0.4/species_diag.out", "SDC-0.4", "0.5", None),
     ]
 
 strang_files = [
-    ("subch_strang/cpu_times.out", "Strang CFL=0.2", "C0", "o"),
-#    ("subch_strang_cfl0.1/cpu_times.out", "Strang CFL=0.1", "C1", "x"),
-    ("subch_strang_noTevolve_fix/cpu_times.out", "Strang, CFL=0.2, no T evolution", "C2", "s"),
-    ("subch_strang_cfl0.05/cpu_times.out", "Strang CFL=0.05", "C0", "+"),
-    ("subch_strang_tol1e-6/cpu_times.out", "Strang CFL=0.2, tols=1.e-6", "C1", "^"),
-    ("subch_strang_tol1.e-8/cpu_times.out", "Strang CFL=0.2, atol_spec=1.e-8", "C1", "x")]
+    ("subch_strang/species_diag.out", "Strang-0.2", "C0", "o"),
+#    ("subch_strang_cfl0.1/species_diag.out", "Strang CFL=0.1", "C1", "x"),
+#    ("subch_strang_noTevolve_fix/species_diag.out", "Strang, CFL=0.2, no T evolution", "C2", "s"),
+    ("subch_strang_cfl0.05/species_diag.out", "Strang-0.05", "C0", "+"),
+    ("subch_strang_tol1e-6/species_diag.out", "Strang-tol1", "C1", "^"),
+    ("subch_strang_tol1.e-8/species_diag.out", "Strang-tol2", "C1", "x")]
 
 fig, ax = plt.subplots()
 

### 2023-06-27 subch_tols/plots/plot_subch_ni.py

@@ -5,16 +5,17 @@ from mpl_toolkits.axes_grid1.inset_locator import (InsetPosition, inset_axes,
 
 sdc_files = [
 #    ("subch_sdc_cfl0.4/species_diag.out", "SDC CFL=0.4", "C0", "+"),
-    ("subch_sdc/species_diag.out", "SDC CFL=0.2", "k", None),
+    ("subch_sdc/species_diag.out", "SDC-0.2", "k", None),
+    ("subch_sdc_cfl0.4/species_diag.out", "SDC-0.4", "0.5", None),
     ]
 
 strang_files = [
-    ("subch_strang/species_diag.out", "Strang CFL=0.2", "C0", "o"),
+    ("subch_strang/species_diag.out", "Strang-0.2", "C0", "o"),
 #    ("subch_strang_cfl0.1/species_diag.out", "Strang CFL=0.1", "C1", "x"),
-    ("subch_strang_noTevolve_fix/species_diag.out", "Strang, CFL=0.2, no T evolution", "C2", "s"),
-    ("subch_strang_cfl0.05/species_diag.out", "Strang CFL=0.05", "C0", "+"),
-    ("subch_strang_tol1e-6/species_diag.out", "Strang CFL=0.2, tols=1.e-6", "C1", "^"),
-    ("subch_strang_tol1.e-8/species_diag.out", "Strang CFL=0.2, atol_spec=1.e-8", "C1", "x")]
+#    ("subch_strang_noTevolve_fix/species_diag.out", "Strang, CFL=0.2, no T evolution", "C2", "s"),
+    ("subch_strang_cfl0.05/species_diag.out", "Strang-0.05", "C0", "+"),
+    ("subch_strang_tol1e-6/species_diag.out", "Strang-tol1", "C1", "^"),
+    ("subch_strang_tol1.e-8/species_diag.out", "Strang-tol2", "C1", "x")]
 
 fig, ax = plt.subplots()
 

### 2023-06-27 subch_tols/plots/plot_subch_ni_resolution.py

@@ -4,9 +4,9 @@ from mpl_toolkits.axes_grid1.inset_locator import (InsetPosition, inset_axes,
                                                    mark_inset)
 
 sdc_files = [
-    ("subch_sdc_40km/species_diag.out", "SDC 40 km"), 
-    ("subch_sdc/species_diag.out", "SDC 20 km"), 
-    ("subch_sdc_10km_3lev/species_diag.out", "SDC 10 km"), 
+    ("subch_sdc_40km/species_diag.out", "SDC-40km"), 
+    ("subch_sdc/species_diag.out", "SDC-0.2"), 
+    ("subch_sdc_10km_3lev/species_diag.out", "SDC-10km"), 
     ]
 
 

### 2023-06-27 subch_tols/plots/plot_cpu.py
fix plot

@@ -2,18 +2,18 @@ import matplotlib.pyplot as plt
 import numpy as np
 
 sdc_files = [
-#    ("subch_sdc_cfl0.4/species_diag.out", "SDC CFL=0.4", "C0", "+"),
-    ("subch_sdc/species_diag.out", "SDC-0.2", "k", None),
-    ("subch_sdc_cfl0.4/species_diag.out", "SDC-0.4", "0.5", None),
+#    ("subch_sdc_cfl0.4/cpu_times.out", "SDC CFL=0.4", "C0", "+"),
+    ("subch_sdc/cpu_times.out", "SDC-0.2", "k", None),
+    ("subch_sdc_cfl0.4/cpu_times.out", "SDC-0.4", "0.5", None),
     ]
 
 strang_files = [
-    ("subch_strang/species_diag.out", "Strang-0.2", "C0", "o"),
-#    ("subch_strang_cfl0.1/species_diag.out", "Strang CFL=0.1", "C1", "x"),
-#    ("subch_strang_noTevolve_fix/species_diag.out", "Strang, CFL=0.2, no T evolution", "C2", "s"),
-    ("subch_strang_cfl0.05/species_diag.out", "Strang-0.05", "C0", "+"),
-    ("subch_strang_tol1e-6/species_diag.out", "Strang-tol1", "C1", "^"),
-    ("subch_strang_tol1.e-8/species_diag.out", "Strang-tol2", "C1", "x")]
+    ("subch_strang/cpu_times.out", "Strang-0.2", "C0", "o"),
+#    ("subch_strang_cfl0.1/cpu_times.out", "Strang CFL=0.1", "C1", "x"),
+#    ("subch_strang_noTevolve_fix/cpu_times.out", "Strang, CFL=0.2, no T evolution", "C2", "s"),
+    ("subch_strang_cfl0.05/cpu_times.out", "Strang-0.05", "C0", "+"),
+    ("subch_strang_tol1e-6/cpu_times.out", "Strang-tol1", "C1", "^"),
+    ("subch_strang_tol1.e-8/cpu_times.out", "Strang-tol2", "C1", "x")]
 
 fig, ax = plt.subplots()
 

### 2023-07-19 xrb5/paper.tex
update

@@
 %\documentclass[12pt, preprint]{aastex}
-\documentclass[preprint,times,tighten]{aastex631}
+\documentclass[preprint,times,tighten,linenumbers,trackchanges]{aastex631}
 
@@
 \section{Introduction}\label{Sec:Introduction}
-
 An X-ray burst (XRB) is a thermonuclear runaway caused by the ignition of the accreted fuel on the surface of a neutron star.
@@
 Nuclear explosions are directly affected by the initial composition of the accreted layer and the accretion rate.
-XRBs found with a mixture of hydrogen and helium fuel layers show bursts with $\sim 5$ sec rise time, while a pure helium layer is typically more explosive and shows bursts with shorter rise times ($\sim 1$ sec).
+XRBs found with a mixture of hydrogen and helium fuel layers show bursts with $\sim 5$ s rise time, while a pure helium layer is typically more explosive and shows bursts with shorter rise times ($\sim 1$ s).
 If the shell is enriched with carbon rather than helium, then the explosion is likely to have an extended rise duration, from minutes to hours known as a superburst \citep{Kuulkers_2002,Cumming_2001,Gupta_2007}.
@@
 
-Since the timescale of accretion between each burst is in the order of $10^4$ secs while the rise time is in the order of $\sim 1 - 10$ sec \citep{Parikh_2013}, it is unlikely for the same thermodynamic condition to exist on the entire surface of the neutron star and for the entire surface to start burning simultaneously \citep{Shara_1982}.
+Since the timescale of accretion between each burst is in the order of $10^4$ s while the rise time is in the order of $\sim 1 - 10$ s \citep{Parikh_2013}, it is unlikely for the same thermodynamic condition to exist on the entire surface of the neutron star and for the entire surface to start burning simultaneously \citep{Shara_1982}.
 Therefore, nuclear ignition is likely to begin in a localized region, and spread to the rest of the neutron star \citep{Spitkovsky_2002}.
@@
 We employ a hydrostatic initial model to represent the neutron star's initial thermodynamic conditions, with a hot model in the left part of the domain and a cool model in the right---this drives a rightward propagating flame (a spreading hotspot for our geometry).
-By default, reactions are incorporated using
+\edit1{The initial composition can be broadly divided into two large regions.
+The base region ($0 < z < 2000$ cm) is composed of pure ${}^{56}$Ni, representing the underlying neutron star.
+On the other hand, the region beyond $z > 2000$ cm consists of pure ${}^{4}$He to represent the accretion layer.} By default, reactions are incorporated using
 operator-splitting.
+
+\edit1{{\castro} uses the adaptive mesh refinement system from {\amrex} \citep{amrex_joss}.
+The same simulation domain of $r = 1.8432 \times 10^5$ cm and $z = 3.072 \times 10^4$ cm is used for all of our models.
+At the coarse grid, we again used $1152 \times 192$ zones or a coarse resolution of $160$ cm with two additional refinement levels with refinement ratios of 4 and 2.
+This means that the first refinement level is 4 times finer than the coarse level and the finest level is 2 times finer than the first level.
+This enables us to achieve a finest level with $9216 \times 1536$ zones, corresponding to a resolution of 20 cm.
+Grid refinement is limited to regions where $\rho > 2.5 \times 10^4 \ \textrm{g} \ \textrm{cm}^{-3}$ in order to avoid unnecessary computational costs.
+The total simulation time is $\sim 120$ ms to avoid flame propagating outside of the domain.}
 
@@
 The reaction rate, $R_{\textrm{th}}$, is found by integrating the cross-section over the Maxwellian velocity distribution.
-If the nuclei involved in the reaction are present in a high-density region that is permeated with plasma particles, then the strength of the Coulomb barrier would be reduced due to the plasma screening effect since the effective charge of the fusing nuclei are reduced.
+If the nuclei involved in the reaction are present in a high-density region that is permeated with plasma particles, then the strength of the Coulomb barrier would be reduced due to the plasma screening effect since the effective charge of the fusing nuclei \deleted{are} \edit1{is} reduced.
 The screening enhancement factor is usually expressed as
@@
 {\tt SCREEN5} is the name of a widely-circulated screening routine (originally written in Fortran) that was the original screening routine used by \castro.
-The overall procedure of {\tt SCREEN5} is summarized in the appendix of \cite{Wallace:1982}.
-In {\tt SCREEN5}, we define $\Gamma = (2/(Z_1+Z_2))^{1/3}Z_1 Z_2 \Gamma_e$, where $\Gamma_e = e^2\sqrt[3]{(4\pi n_e/3)}/(k_BT)$ describes the thermodynamic condition at which the reaction takes place, $Z$ is the charge of the fusing nuclei, $\mu_{12}$ is the reduced mass for the two fusing nuclei, $e$ is the electron charge, $n_e$ is the electron number density, and $k_B$ is Boltzmann's constant.
+The overall procedure of {\tt SCREEN5} is summarized in the appendix of \cite{Wallace:1982}. In {\tt SCREEN5}, \edit1{the Coulomb coupling parameter, $\Gamma$, is defined in Eq. \ref{Eq:gamma_eq}},
+
+\begin{equation}\label{Eq:gamma_eq}
+    \Gamma = \sqrt[3]{\frac{2}{Z_1+Z_2}}Z_1 Z_2 \Gamma_e
+\end{equation}
+
+\edit1{where $\Gamma_e$ describes the thermodynamic condition at which the reaction takes place and is defined in Eq. \ref{Eq:gamma_e_eq},}
+
+\begin{equation}\label{Eq:gamma_e_eq}
+    \Gamma_e = \sqrt[3]{\frac{4\pi n_e}{3}}\frac{e^2}{k_BT}
+\end{equation}
+
+\deleted{we define $\Gamma = (2/(Z_1+Z_2))^{1/3}Z_1 Z_2 \Gamma_e$, where $\Gamma_e = e^2\sqrt[3]{(4\pi n_e/3)}/(k_BT)$ describes the thermodynamic condition at which the reaction takes place,} $Z$ is the charge of the fusing nuclei, \deleted{$\mu_{12} = $ is the reduced mass for the two fusing nuclei,} $e$ is the electron charge, $n_e$ is the electron number density, and $k_B$ is Boltzmann's constant.
 When the plasma screening effect is weak or $\Gamma < 0.3$, {\tt SCREEN5} utilizes the equation proposed by \cite{Graboske_1973, Dewitt_1973}, which assumes that the interacting nuclei are separated by zero distance.
@@
 {\tt aprox13} & {\tt aprox13} & Strang-splitting & {\tt SCREEN5} \\
+{\tt subch\_full\_mod} & {\tt subch\_full\_mod} & Strang-splitting & {\tt SCREEN5} \\
 {\tt subch\_full} & {\tt subch\_full} & Strang-splitting & {\tt SCREEN5} \\
-{\tt subch\_full\_mod} & {\tt subch\_full\_mod} & Strang-splitting & {\tt SCREEN5} \\
 {\tt subch\_simple} & {\tt subch\_simple} & Strang-splitting & {\tt SCREEN5} \\
+{\tt aprox13\_chu} & {\tt aprox13} & Strang-splitting & {\tt CHUGUNOV2007}\\
 {\tt aprox13\_sdc} & {\tt aprox13} & simplified SDC & {\tt SCREEN5} \\
 {\tt subch\_full\_sdc} & {\tt subch\_full} & simplified SDC & {\tt SCREEN5} \\
-{\tt aprox13\_chu} & {\tt aprox13} & Strang-splitting & {\tt CHUGUNOV2007}
 
@@
 \plotone{network_abar_50ms.pdf}
-\caption{\label{fig:network_abar_50ms} Slice plots comparing $\Bar{A}$ for {\tt aprox13} (top panel), {\tt subch\_full} (second panel from top), {\tt subch\_full\_mod} (third panel), and {\tt subch\_simple} (last panel) at 50 ms.
-}
+\caption{\label{fig:network_abar_50ms} Slice plots comparing $\Bar{A}$ for {\tt aprox13} (top panel), {\tt subch\_full\_mod} (second panel from top), {\tt subch\_full} (third panel), and {\tt subch\_simple} (last panel) at 50 ms.
+\edit1{The vertical scaling is stretched to provide better clarity in the vertical flame structure.}}
 \end{figure*}
@@
 \plotone{network_enuc_50ms.pdf}
-\caption{\label{fig:network_enuc_50ms} Slice plots comparing $\dot{e}_{\textrm{nuc}}$ for {\tt aprox13} (top panel), {\tt subch\_full} (second panel from top), {\tt subch\_full\_mod} (third panel), and {\tt subch\_simple} (last panel) at 50 ms.}
+\caption{\label{fig:network_enuc_50ms} Slice plots comparing $\dot{e}_{\textrm{nuc}}$ for various models at 50 ms.
+\edit1{The vertical scaling is stretched to provide better clarity in the vertical flame structure.}}
 \end{figure*}
@@
     \plotone{network_aprox13_nuc_frac.pdf}
-    \caption{\label{fig:network_aprox13_nuc} Slice plots showing the mass fractions of ${}^{12}$C, ${}^{16}$O, and ${}^{28}$Si for {\tt aprox13} at 20 ms.}
+    \caption{\label{fig:network_aprox13_nuc} Slice plots showing the mass fractions of ${}^{12}$C, ${}^{16}$O, and ${}^{28}$Si for {\tt aprox13} at 20 ms.
+\edit1{The vertical scaling is stretched to provide better clarity in the vertical flame structure.}}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+    \plotone{network_subch_mod_nuc_frac.pdf}
+    \caption{\label{fig:network_subch_mod_nuc} Slice plots showing the mass fractions of ${}^{12}$C, ${}^{16}$O, and ${}^{28}$Si for {\tt subch\_full\_mod} at 20 ms.
+\edit1{The vertical scaling is stretched to provide better clarity in the vertical flame structure.}}
 \end{figure*}
@@
     \plotone{network_subch_nuc_frac.pdf}
-    \caption{\label{fig:network_subch_nuc} Slice plots showing the mass fractions of ${}^{12}$C, ${}^{16}$O, and ${}^{28}$Si for {\tt subch\_full} at 20 ms.}
-\end{figure*}
-
-\begin{figure*}
-    \centering
-    \plotone{network_subch_mod_nuc_frac.pdf}
-    \caption{\label{fig:network_subch_mod_nuc} Slice plots showing the mass fractions of ${}^{12}$C, ${}^{16}$O, and ${}^{28}$Si for {\tt subch\_full\_mod} at 20 ms.}
+    \caption{\label{fig:network_subch_nuc} Slice plots showing the mass fractions of ${}^{12}$C, ${}^{16}$O, and ${}^{28}$Si for {\tt subch\_full} at 20 ms.
+\edit1{The vertical scaling is stretched to provide better clarity in the vertical flame structure.}}
 \end{figure*}
@@
     \plotone{network_species_summary_log.pdf}
-    \caption{\label{fig:network_species_summary} The overall evolution of the total mass for ${}^{12}$C, ${}^{16}$O, ${}^{20}$Ne, ${}^{24}$Mg, ${}^{28}$Si, and ${}^{32}$S for the 4 simulations models: {\tt aprox13}, {\tt subch\_full}, {\tt subch\_full\_mod}, and {\tt subch\_simple}.}
+    \caption{\label{fig:network_species_summary} The overall evolution of the total mass for ${}^{12}$C, ${}^{16}$O, ${}^{20}$Ne, ${}^{24}$Mg, ${}^{28}$Si, and ${}^{32}$S for the 4 simulations models: {\tt aprox13},  {\tt subch\_full\_mod}, {\tt subch\_full}, and {\tt subch\_simple}.}
 \end{figure*}
@@
     \centering
-    \plotone{network_u_v_enuc_25ms.pdf}
-    \plotone{network_u_v_enuc_100ms.pdf}
-    \caption{\label{fig:network_u_v_enuc} $u-v$ phase plots for {\tt aprox13} (panels in the first column on the left), {\tt subch\_full} (second column), {\tt subch\_full\_mod} (third column), {\tt subch\_simple} (fourth column) at $t = 25$ ms (top 4 panels) and $t = 100$ ms (bottom 4 panels).
+    \plotone{network_u_v_enuc_aprox13.pdf}
+    \plotone{network_u_v_enuc_subch_full_mod.pdf}
+    \plotone{network_u_v_enuc_subch_full.pdf}
+    \plotone{network_u_v_enuc_subch_simple.pdf}
+    %\plotone{network_u_v_enuc_25ms.pdf}
+    %\plotone{network_u_v_enuc_100ms.pdf}
+    \caption{\label{fig:network_u_v_enuc} $u-v$ phase plots for {\tt aprox13} \edit1{(panels in the first row), {\tt subch\_full\_mod} (second row), {\tt subch\_full} (third row), {\tt subch\_simple} (fourth row) at $t = 25$ ms (4 panels on the first column) and $t = 100$ ms (4 panels on the second column)}.
 The x-axis, $u$, shows the velocity in the $r$ direction, whereas the y-axis, $v$, shows the velocity in the $z$ direction.
@@
 Fitted parameters of the fitting functions Eq.
-\ref{Eq:quadratic_fit} and \ref{Eq:tanh_fit} for {\tt aprox13}, {\tt subch\_full}, {\tt subch\_full\_mod}, and {\tt aprox13\_chu}.
+\ref{Eq:quadratic_fit} and \ref{Eq:tanh_fit} for {\tt aprox13}, \edit1{{\tt subch\_full\_mod}, {\tt subch\_full}}, and {\tt aprox13\_chu}.
 The fitting function is applied for $t > 8$ ms.
@@
 
+{\tt subch\_full\_mod} & $58.53 \pm 0.24$ & $2.122 \pm 0.016$ & $0.7773 \pm 0.0004$ & N/A & N/A & N/A\\
+
 {\tt subch\_full} & $9.22 \pm 0.82$ & $4.489 \pm 0.067$ & $0.8646 \pm 0.0014$ & $0.150 \pm 0.001$ & $0.0093 \pm 0.0001$ & $-2.551 \pm 0.035$ \\
-
-{\tt subch\_full\_mod} & $58.53 \pm 0.24$ & $2.122 \pm 0.016$ & $0.7773 \pm 0.0004$ & N/A & N/A & N/A \\
 
@@
 \caption{\label{Tab:network_instan_vel}
-Instantaneous flame propagation speed at $t = 23$ ms and $t = 100$ ms for {\tt aprox13}, {\tt subch\_full}, {\tt subch\_full\_mod}, {\tt subch\_simple}, and {\tt aprox13\_chu}.
-$t = 23$ ms and $t=100$ ms represent the acceleration phase for {\tt subch\_full} and {\tt subch\_simple} and the steady phase at the late-stage, respectively.
-$t_{10}$ represents the theoretical time for the flame to reach 10 km.
+Instantaneous flame propagation speed at $t = 23$ ms and $t = 100$ ms for {\tt aprox13}, \edit1{{\tt subch\_full\_mod}, {\tt subch\_full}}, {\tt subch\_simple}, \edit1{and {\tt aprox13\_chu} are represented by $v_{23}$ and $v_{100}$, respectively.} \edit1{The acceleration phase for {\tt subch\_full} and {\tt subch\_simple} and the steady phase at the late-stage are represented by $t = 23$ ms and $t = 100$ ms, respectively.} \deleted{$t = 23$ ms and $t=100$ ms represent the acceleration phase for {\tt subch\_full} and {\tt subch\_simple} and the steady phase at the late-stage, respectively.} \edit1{The theoretical time for the flame to reach 30 km is represented by $t_{30}$.} \deleted{$t_{10}$ represents the theoretical time for the flame to reach 10 km.}
 }
@@
 $v_{100}$ [$\mbox{km} \ \mbox{s}^{-1}$] &
-$t_{10}$ [s]
+$t_{30}$ [s]
 \\
@@
 
-{\tt aprox13} & $3.369 \pm 0.016$ & $5.234 \pm 0.027$ & 0.7647 \\
-{\tt subch\_full} & $20.732 \pm 0.284$ & $5.411 \pm 0.105$ & 0.9917 \\
-{\tt subch\_full\_mod} & $3.468 \pm 0.017$ & $7.975 \pm 0.029$ & 0.4873 \\
-{\tt subch\_simple} & $21.095 \pm 0.332$ & $5.521 \pm 0.120$ & 0.8483 \\
-{\tt aprox13\_chu} & $3.106 \pm 0.014$ & $4.874 \pm 0.024$ & 0.7912 \\
+{\tt aprox13} & $3.369 \pm 0.016$ & $5.234 \pm 0.027$ & $1.4419 \pm 0.0064$\\
+{\tt subch\_full\_mod} & $3.468 \pm 0.017$ & $7.975 \pm 0.029$ & $0.9697 \pm 0.0019$ \\
+{\tt subch\_full} & $20.732 \pm 0.284$ & $5.411 \pm 0.105$ & $2.0717 \pm 0.0749$ \\
+{\tt subch\_simple} & $21.095 \pm 0.332$ & $5.521 \pm 0.120$ & $1.6918 \pm 0.0441$ \\
+{\tt aprox13\_chu} & $3.106 \pm 0.014$ & $4.874 \pm 0.024$ & $1.4872 \pm 0.0060$ \\
 % {\tt aprox13\_sdc} & $3.039 \pm 0.014$ & $5.201 \pm 0.024$ & $5.764 \pm 0.027$\\
@@
 Furthermore, {\tt subch\_full\_mod} accelerates towards the end since it still has a sufficient amount of ${}^{12}$C fuel due to the inefficient reaction flows to ${}^{16}$O.
-Even though the flame front for {\tt subch\_full\_mod} is expected to surpass {\tt subch\_full} and {\tt subch\_simple} at $t \sim 160$ ms, based on Figure \ref{fig:network_front}, $\dot{e}_{\textrm{nuc}}$ from Figure \ref{fig:network_time_profile} appears to reach its peak at $120$ ms.
-Therefore, there is no guarantee that the lateral propagating flame for {\tt subch\_full\_mod} would exceed its counterparts.
-
-Assuming that the fitting functions accurately describe the flame propagation, the expected time for the flame to reach a typical neutron star radius of 10 km, $t_{10}$, can be calculated, as shown in Table \ref{Tab:network_instan_vel}.
-$t_{10}$ serves as a rough prediction for the rise time of XRBs with a pure ${}^{4}$He accretion layer.
-$t_{10}$ for all models are $\lesssim 1$ sec, whereas $t_{10}$ for {\tt subch\_full} is $\sim 1$ sec, consistent with previous observational studies \citep{galloway:2008}.
+Even though the flame front for {\tt subch\_full\_mod} is expected to surpass {\tt subch\_full} and {\tt subch\_simple} at $t \sim 160$ ms, based on Figure \ref{fig:network_front}, $\dot{e}_{\textrm{nuc}}$ from Figure \ref{fig:network_time_profile} appears to reach its peak at $\sim 120$ ms.
+\deleted{Therefore, there is no guarantee that the lateral propagating flame for {\tt subch\_full\_mod} would exceed its counterparts} \edit1{In the event of extreme downturns in $\dot{e}_{\textrm{nuc}}$ in the future, there is no assurance that the lateral propagating flame for {\tt subch\_full\_mod} would surpass its counterparts.}
+
+Assuming that the fitting functions accurately describe the flame propagation, the expected time for the flame to \deleted{reach a typical neutron star radius of 10 km} \edit1{propagate laterally across a neutron star with a radius of 10 km or $\pi R \sim 30$ km}, $t_{30}$, can be calculated, as shown in Table \ref{Tab:network_instan_vel}.
+\edit1{The quantity $t_{30}$} \deleted{$t_{10}$} serves as a rough prediction for the rise time of XRBs with a pure ${}^{4}$He accretion layer.
+\edit1{A basic calculation shows that $t_{30}$} \deleted{$t_{10}$} for all models are \deleted{$\lesssim 1$ s} \edit1{$\lesssim 2$ s}, whereas \deleted{$t_{10}$} \edit1{$t_{30}$} for {\tt subch\_full} is \deleted{$\sim 1$ s} \edit1{$\sim 2$ s}, consistent with previous observational studies \citep{galloway:2008}.
+\edit1{Here we emphasize that $t_{30}$ is extrapolated beyond the data with the assumption that the flame has a sustained constant acceleration.
+However, this assumption may not hold true.
+For instance, in the case of {\tt subch\_full\_mod}, $\dot{e}_{\textrm{nuc}}$ already reaches its peak at 120 ms, indicating that the flame's acceleration is not constant.
+Therefore, $t_{30}$ only serves as an order-of-magnitude calculation.}
 
@@
 \plotone{screen_temp_50ms.pdf}
-\caption{\label{fig:screen_temp} Slice plots comparing temperature for {\tt aprox13} (top panel) and {\tt aprox13\_chu} (bottom panel) at $t = 50$ ms.}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\plotone{screen_gamma_50ms.pdf}
-\caption{\label{fig:screen_gamma_50ms} Slice plots comparing $\Gamma_e$ for {\tt aprox13} (top panel) and {\tt aprox13\_chu} (bottom panel) at $t = 50$ ms.}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\plotone{screen_gamma_100ms.pdf}
-\caption{\label{fig:screen_gamma_100ms} Slice plots comparing $\Gamma_e$ for {\tt aprox13} (top panel) and {\tt aprox13\_chu} (bottom panel) at $t = 100$ ms.}
+\caption{\label{fig:screen_temp} Slice plots comparing temperature for {\tt aprox13} (top panel) and {\tt aprox13\_chu} (bottom panel) at $t = 50$ ms.
+\edit1{The vertical scaling is stretched to provide better clarity in the vertical flame structure.}}
+\end{figure*}
+
+\begin{figure*}
+\centering
+\plotone{screen_gamma_e_50ms.pdf}
+\caption{\label{fig:screen_gamma_50ms} Slice plots comparing $\Gamma_e$ for {\tt aprox13} (top panel) and {\tt aprox13\_chu} (bottom panel) at $t = 50$ ms.
+\edit1{The vertical scaling is stretched to provide better clarity in the vertical flame structure.}}
+\end{figure*}
+
+\begin{figure*}
+\centering
+\plotone{screen_gamma_e_100ms.pdf}
+\caption{\label{fig:screen_gamma_100ms} Slice plots comparing $\Gamma_e$ for {\tt aprox13} (top panel) and {\tt aprox13\_chu} (bottom panel) at $t = 100$ ms.
+\edit1{The vertical scaling is stretched to provide better clarity in the vertical flame structure.}}
 \end{figure*}
@@
 Figure \ref{fig:screen_temp} shows the temperature for {\tt aprox13} and {\tt aprox13\_chu} at $t = 50$ ms, which suggests that there are no significant differences in the flame structure between the two models.
-Figure \ref{fig:screen_gamma_50ms} and \ref{fig:screen_gamma_100ms} show $\Gamma_e$, a measure to determine the approximate screening regime for the flame, at $t = 50$ ms and $t = 100$ ms.
+Figure \ref{fig:screen_gamma_50ms} and \ref{fig:screen_gamma_100ms} show $\Gamma_e$ \edit1{(defined in Eq.
+\ref{Eq:gamma_e_eq})}, \edit1{which} provides a measure to determine the approximate screening regime for the flame, at $t = 50$ ms and $t = 100$ ms.
 
 Although a general trend of decreasing $\Gamma_e$ is observed as the flame progress, it is noteworthy that there are more fusion processes involving heavier nuclei at later stages of burning.
-Assuming triple-$\alpha$ process, it can inferred that the Coulomb coupling parameter, $\Gamma \sim 0.05$ and $0.025$ for $t = 50$ ms and $t = 100$ ms.
+Assuming triple-$\alpha$ process, it can inferred that the Coulomb coupling parameter \edit1{$\Gamma$ (defined in Eq.
+\ref{Eq:gamma_eq})} $\sim 0.05$ for $t = 50$ ms and $0.025$ for $t = 100$ ms.
 Meanwhile, for oxygen burning at $t = 50$ ms and 100 ms, $\Gamma \sim 0.48$ and $0.32$, respectively.
@@
 
-\begin{figure*}
-\centering
-\plotone{screen_time_profile.pdf}
-\caption{\label{fig:screen_profile} Time profiles showing the weighted temperature and energy generation rate of the burning front for {\tt aprox13} and {\tt aprox13\_chu}.}
-\end{figure*}
-
-The comparison between {\tt aprox13} and {\tt aprox13\_chu} regarding the weighted temperature and energy generate rate is illustrated in Figure \ref{fig:screen_profile}.
+% \begin{figure*}
+% \centering
+% \plotone{screen_time_profile.pdf}
+% \caption{\label{fig:screen_profile} Time profiles showing the weighted temperature and energy generation rate of the burning front for {\tt aprox13} and {\tt aprox13\_chu}.}
+% \end{figure*}
+
+The comparison between {\tt aprox13} and {\tt aprox13\_chu} regarding the weighted temperature and energy generate rate is illustrated in Figure \edit1{\ref{fig:network_time_profile}}.
 This plot illustrates that overall the nuclear energy generation rate of {\tt aprox13} is higher and exhibits a faster increasing rate compared to {\tt aprox13\_chu}.
@@
 
-\begin{figure*}
-\centering
-\plotone{screen_front.pdf}
-\caption{\label{fig:screen_front} Flame front position as a function of time for {\tt aprox13} and {\tt aprox13\_chu}. The dashed lines are the fitted curves using Eq. \ref{Eq:quadratic_fit}.}
-\end{figure*}
-
-Figure \ref{fig:screen_front} depicts the evolution of the flame front position over time.
+% \begin{figure*}
+% \centering
+% \plotone{screen_front.pdf}
+% \caption{\label{fig:screen_front} Flame front position as a function of time for {\tt aprox13} and {\tt aprox13\_chu}. The dashed lines are the fitted curves using Eq. \ref{Eq:quadratic_fit}.}
+% \end{figure*}
+
+Figure \ref{fig:network_front} depicts the evolution of the flame front position over time \edit1{for {\tt aprox13\_chu} comparing to other models}.
 It is observed that the flame speed for {\tt aprox13\_chu} is slower compared to {\tt aprox13} due to the smaller value of $\dot{e}_{\textrm{nuc}}$ at later times.
@@
 
-\begin{figure*}
-\centering
-\plotone{integration_enuc_50ms.pdf}
-\caption{\label{fig:integration_enuc} Slice plots comparing $\dot{e}_{\textrm{nuc}}$ for {\tt aprox13\_sdc} (top panel) and {\tt subch\_full\_sdc} (bottom panel) at $t = 50$ ms.}
-\end{figure*}
+% \begin{figure*}
+% \centering
+% \plotone{integration_enuc_50ms.pdf}
+% \caption{\label{fig:integration_enuc} Slice plots comparing $\dot{e}_{\textrm{nuc}}$ for {\tt aprox13\_sdc} (top panel) and {\tt subch\_full\_sdc} (bottom panel) at $t = 50$ ms.}
+% \end{figure*}
 
@@
 This result is nearly identical to the models that employed Strang-splitting as shown in Figure \ref{fig:network_abar_50ms}.
-One consequence of the strong coupling between advection and reactions is, that to make the energy generation plot, we need to derive the nuclear energy release by subtracting off the advection contribution over a timestep.
+One consequence of the strong coupling between advection and reactions is, that to make the energy generation plot \edit1{shown in Fig \ref{fig:network_enuc_50ms}}, we need to derive the nuclear energy release by subtracting off the advection contribution over a timestep.
 In regions where there is not much burning, roundoff error can introduce some noise into the energy generation plot, which is seen as the multiple dotted regions in the $\dot{e}_{\textrm{nuc}}$ plot for {\tt aprox13\_sdc} and {\tt subch\_full\_sdc}.
-This is simply an artifact of how we do
-the derivation.
+This is simply an artifact of how we do the derivation.
 
@@
 
+\deleted{As shown in \citet{castro_simple_sdc}, in regions where the burning
+is vigorous, the simplified-SDC method provides a better solution
+than Strang splitting.  However, the XRB flame we simulate here
+are not very demanding, so the benefit is minimal.  
+%is not extreme enough for nuclear statistical equilibrium to occur, any difference between simplified-SDC and Strang-splitting schemes is expected to be negligible.
+This is supported by the temperature and $\dot{e}_{\textrm{nuc}}$ profiles for {\tt aprox13\_sdc} and {\tt subch\_full\_sdc} shown in Figure \ref{fig:integration_profile}, which are almost identical, except for minor variations in $\dot{e}_{\textrm{nuc}}$ for {\tt subch\_full} and {\tt subch\_full\_sdc}.
+Similarly, the flame front position plots shown in Figure \ref{fig:integration_front} for the simplified-SDC simulations are nearly identical compared to their counterparts, with minor variations (Table \ref{Tab:network_instan_vel}) in the instantaneous flame speed calculated using the the fitting function.
+As a result, the two simplified-SDC simulations were also more
+computationally-expensive than their Strang counterparts.
+This differs from the case in \citet{castro_simple_sdc} where 
+the simplified-SDC algorithm reduced computational expenses in extreme thermodynamic conditions by mitigating the stiffness of solving reaction equations.}
+
+\subsection{Computational Expenses}
+
+\begin{figure*}
+\centering
+\plotone{integration_wallclock.pdf}
+\caption{\label{fig:wallclock_time} Wallclock time used per ms of simulation time as a function of simulation time.
+Wallclock time analysis is only conducted for $t > 40$ ms as our simulation models utilized a combination of CPU and GPU for $t < 40$ ms.}
+\end{figure*}
+
+\edit1{In order to give an overview of the computational expenses needed for each simulation models, Figure \ref{fig:wallclock_time} shows the wallclock time used per ms of simulation time as a function of simulation time for all models.
+As some models utilized CPU while others used GPU for $t < 40$ ms, a wallclock time analysis fails to provide an accurate estimate of the computational expenses incurred by each model during this time frame.
+Therefore, we have limited the presentation of our data to $t > 40$ ms.
+The results indicate that the four simulation models, which differ only in terms of their reaction networks, exhibit the expected trends.
+Specifically, the simulation model with the smallest network ({\tt aprox13}) that has the fewest isotopes and reaction rates requires the smallest number of wallclock hours for $1$ ms of simulation time.
+Conversely, the largest network, {\tt subch\_full}, requires $\sim 3-4$ times the wallclock time needed by {\tt aprox13} to complete the same simulation.
+Notably, the computational cost for {\tt subch\_full\_mod} gradually increases over time, ultimately exceeding {\tt subch\_full}.
+{\tt aprox13} and {\tt aprox13\_chu} only show negligible differences in the computational cost shown in Figure \ref{fig:wallclock_time}.
+}
+
+\edit1{Regarding the simulation models that utilized different integration methods, both {\tt aprox13\_sdc} and {\tt subch\_full\_sdc} models, which employed simplified-SDC, were found to be more computationally expensive compared to their counterparts, {\tt aprox13} and {\tt subch\_full}, which utilized Strang-splitting.
 As shown in \citet{castro_simple_sdc}, in regions where the burning
@@
 than Strang splitting.  However, the XRB flame we simulate here
-are not very demanding, so the benefit is minimal.  
-\begin{comment}
-is not extreme enough for nuclear statistical equilibrium to occur, any difference between simplified-SDC and Strang-splitting schemes is expected to be negligible.
-This is supported by the temperature and $\dot{e}_{\textrm{nuc}}$ profiles for {\tt aprox13\_sdc} and {\tt subch\_full\_sdc} shown in Figure \ref{fig:integration_profile}, which are almost identical, except for minor variations in $\dot{e}_{\textrm{nuc}}$ for {\tt subch\_full} and {\tt subch\_full\_sdc}.
-Similarly, the flame front position plots shown in Figure \ref{fig:integration_front} for the simplified-SDC simulations are nearly identical compared to their counterparts, with minor variations (Table \ref{Tab:network_instan_vel}) in the instantaneous flame speed calculated using the the fitting function.
-\end{comment}
-As a result, the two simplified-SDC simulations were also more
-computationally-expensive than their Strang counterparts.
-This differs than the case in \citet{castro_simple_sdc} where 
-the simplified-SDC algorithm reduced computational expenses in extreme thermodynamic conditions by mitigating the stiffness of solving reaction equations.
-
-% \subsection{Computational Expenses}
-
-% \begin{figure*}
-% \centering
-% \plotone{integration_wallclock.pdf}
-% \caption{\label{fig:wallclock_time} Wallclock time used per simulation time as function of simulation time.
-Wallclock time analysis is only conducted for $t > 40$ ms as our simulation models utilized a combination of CPU and GPU for $t < 40$ ms.}
-% \end{figure*}
-
-% In order to give an overview of the computational expenses needed for each simulation models, Figure \ref{fig:wallclock_time} shows the wallclock time used per simulation time as a function of simulation time for all models.
-As some models utilized CPU while others used GPU for $t < 40$ ms, a wallclock time analysis fails to provide an accurate estimate of the computational expenses incurred by each model for during this time frame.
-Therefore, we have limited the presentation of our data to $t > 40$ ms.
-The results indicate that the four simulation models, which differ only in terms of their reaction networks, exhibit the expected trends.
-Specifically, the simulation model with the smallest network, {\tt aprox13}, that has the fewest isotopes and reaction rates requires the least amount of wallclock hours for $1$ ms of simulation time.
-Conversely, the largest network, {\tt subch\_full}, requires $\sim 3-4$ times the wallclock time needed by {\tt aprox13} to complete the same simulation.
-Notably, the computational cost for {\tt subch\_full]\_mod} gradually increases over time, ultimately exceeding {\tt subch\_full}.
-{\tt aprox13} and {\tt aprox13\_chu} only show negligible differences in the computational cost shown in Figure \ref{fig:wallclock_time}.
-Regarding the simulation models that utilized different integration methods, both {\tt aprox13\_sdc} and {\tt subch\_full\_sdc} models, which employed simplified-SDC, were found to be more computationally expensive compared to their counterparts, {\tt aprox13} and {\tt subch\_full}, which utilized Strang-splitting.
-Despite the fact that simplified-SDC is expected to reduce computational expenses in extreme thermodynamic conditions by mitigating the stiffness of solving reaction equations, Figure \ref{fig:wallclock_time} suggests that it may increase computational costs under non-extreme conditions.
+is not very demanding, so the benefit is minimal. 
+%As a result, the two simplified-SDC simulations were also more computationally-expensive than their Strang counterparts.
+This differs from the case in \citet{castro_simple_sdc} where 
+the simplified-SDC algorithm reduced computational expenses in extreme thermodynamic conditions by mitigating the stiffness of solving reaction equations.}
+
+%Despite the fact that simplified-SDC is expected to reduce computational expenses in extreme thermodynamic conditions by mitigating the stiffness of solving reaction equations, Figure \ref{fig:wallclock_time} suggests that it may increase computational costs under non-extreme conditions.
 
@@
 We explored the sensitivity of an XRB flame to the details of the
-nuclear physics: size of, and approximations in a reaction network, screening methods, and time-integration strategies. The main
-differences observed with reaction network are:
+nuclear physics: size of, and approximations in a reaction network, screening methods, and time-integration strategies. \edit1{The main takeaways from this study are:}
+
+%The main differences observed with reaction network are:
 
@@
     %the additional rates such as ${}^{14}\mbox{N}(\alpha,\gamma){}^{18}\mbox{F}(\alpha, \mbox{p}){}^{21}\mbox{Ne}$ and could be the leading cause, although further investigation is needed.
+
+    \item \edit1{Comparing the two screening routines, {\tt SCREEN5} and {\tt CHUNGUNOV2007}, we find that {\tt CHUGUNOV2007} has a slightly weaker screening effect in the weak and intermediate screening regimes.
+As a consequence, the weaker screening effect from {\tt CHUGUNOV2007} leads to a slightly slower flame compared to the {\tt SCREEN5} model.
+This result matches our expectations and is in agreement with \cite{Chugunov_2007}.}
+
+    \item \edit1{We investigated the performance of the simplified-SDC scheme in comparison to the traditional Strang-splitting.
+For this problem, since the burning is not very vigorous, there is no strong benefit of using simplified-SDC over Strang-splitting.}
 \end{itemize}
 
-Comparing the two screening routines, {\tt SCREEN5} and {\tt CHUNGUNOV2007}, we find that {\tt CHUGUNOV2007} has a slightly weaker screening effect in the weak and intermediate screening regimes.
+\deleted{Comparing the two screening routines, {\tt SCREEN5} and {\tt CHUNGUNOV2007}, we find that {\tt CHUGUNOV2007} has a slightly weaker screening effect in the weak and intermediate screening regimes.
 As a consequence, the weaker screening effect from {\tt CHUGUNOV2007} leads to a slightly slower flame compared to the {\tt SCREEN5} model.
-This result matches our expectations and is in agreement with \cite{Chugunov_2007}.
-
-Finally, we investigated the performance of the simplified-SDC scheme in comparison to the traditional Strang-splitting.
-For this problem, since the burning is not very vigorous, there is no strong benefit of using simplified-SDC over Strang-splitting.
+This result matches our expectations and is in agreement with \cite{Chugunov_2007}.}
+
+\deleted{Finally, we investigated the performance of the simplified-SDC scheme in comparison to the traditional Strang-splitting.
+For this problem, since the burning is not very vigorous, there is no strong benefit of using simplified-SDC over Strang-splitting.}
 
@@
 \castro\ is open-source and freely available at
-\url{http://github.com/AMReX-Astro/Castro}.  The work at Stony Brook was supported by DOE/Office
+\url{http://github.com/AMReX-Astro/Castro}.
+\edit1{Metadata, input files, scripts to generate plots, etc, are available at \citep{metadata_for_xrb5}.} The work at Stony Brook was supported by DOE/Office
 of Nuclear Physics grant DE-FG02-87ER40317. This research was supported by the Exascale Computing 
@@
           yt \citep{yt}
-          pynucastro \citep{pynucastro, pynucastro2,the_pynucastro_development_2022_7239007}}
+          pynucastro \citep{pynucastro, pynucastro2,the_pynucastro_development_2022_7239007},
+          uncertainties \citep{uncertainties}
+          }
          % Appendix

### 2023-07-19 xrb5/paper.tex
remove \deleted{} since aas is not happy about it

@@
 The reaction rate, $R_{\textrm{th}}$, is found by integrating the cross-section over the Maxwellian velocity distribution.
-If the nuclei involved in the reaction are present in a high-density region that is permeated with plasma particles, then the strength of the Coulomb barrier would be reduced due to the plasma screening effect since the effective charge of the fusing nuclei \deleted{are} \edit1{is} reduced.
+If the nuclei involved in the reaction are present in a high-density region that is permeated with plasma particles, then the strength of the Coulomb barrier would be reduced due to the plasma screening effect since the effective charge of the fusing nuclei \edit1{is} reduced.
 The screening enhancement factor is usually expressed as
@@
 
-\deleted{we define $\Gamma = (2/(Z_1+Z_2))^{1/3}Z_1 Z_2 \Gamma_e$, where $\Gamma_e = e^2\sqrt[3]{(4\pi n_e/3)}/(k_BT)$ describes the thermodynamic condition at which the reaction takes place,} $Z$ is the charge of the fusing nuclei, \deleted{$\mu_{12} = $ is the reduced mass for the two fusing nuclei,} $e$ is the electron charge, $n_e$ is the electron number density, and $k_B$ is Boltzmann's constant.
+$Z$ is the charge of the fusing nuclei, $e$ is the electron charge, $n_e$ is the electron number density, and $k_B$ is Boltzmann's constant.
 When the plasma screening effect is weak or $\Gamma < 0.3$, {\tt SCREEN5} utilizes the equation proposed by \cite{Graboske_1973, Dewitt_1973}, which assumes that the interacting nuclei are separated by zero distance.
@@
 \caption{\label{Tab:network_instan_vel}
-Instantaneous flame propagation speed at $t = 23$ ms and $t = 100$ ms for {\tt aprox13}, \edit1{{\tt subch\_full\_mod}, {\tt subch\_full}}, {\tt subch\_simple}, \edit1{and {\tt aprox13\_chu} are represented by $v_{23}$ and $v_{100}$, respectively.} \edit1{The acceleration phase for {\tt subch\_full} and {\tt subch\_simple} and the steady phase at the late-stage are represented by $t = 23$ ms and $t = 100$ ms, respectively.} \deleted{$t = 23$ ms and $t=100$ ms represent the acceleration phase for {\tt subch\_full} and {\tt subch\_simple} and the steady phase at the late-stage, respectively.} \edit1{The theoretical time for the flame to reach 30 km is represented by $t_{30}$.} \deleted{$t_{10}$ represents the theoretical time for the flame to reach 10 km.}
+Instantaneous flame propagation speed at $t = 23$ ms and $t = 100$ ms for {\tt aprox13}, \edit1{{\tt subch\_full\_mod}, {\tt subch\_full}}, {\tt subch\_simple}, \edit1{and {\tt aprox13\_chu} are represented by $v_{23}$ and $v_{100}$, respectively.} \edit1{The acceleration phase for {\tt subch\_full} and {\tt subch\_simple} and the steady phase at the late-stage are represented by $t = 23$ ms and $t = 100$ ms, respectively.} \edit1{The theoretical time for the flame to reach 30 km is represented by $t_{30}$.}
 }
@@
 Even though the flame front for {\tt subch\_full\_mod} is expected to surpass {\tt subch\_full} and {\tt subch\_simple} at $t \sim 160$ ms, based on Figure \ref{fig:network_front}, $\dot{e}_{\textrm{nuc}}$ from Figure \ref{fig:network_time_profile} appears to reach its peak at $\sim 120$ ms.
-\deleted{Therefore, there is no guarantee that the lateral propagating flame for {\tt subch\_full\_mod} would exceed its counterparts} \edit1{In the event of extreme downturns in $\dot{e}_{\textrm{nuc}}$ in the future, there is no assurance that the lateral propagating flame for {\tt subch\_full\_mod} would surpass its counterparts.}
-
-Assuming that the fitting functions accurately describe the flame propagation, the expected time for the flame to \deleted{reach a typical neutron star radius of 10 km} \edit1{propagate laterally across a neutron star with a radius of 10 km or $\pi R \sim 30$ km}, $t_{30}$, can be calculated, as shown in Table \ref{Tab:network_instan_vel}.
-\edit1{The quantity $t_{30}$} \deleted{$t_{10}$} serves as a rough prediction for the rise time of XRBs with a pure ${}^{4}$He accretion layer.
-\edit1{A basic calculation shows that $t_{30}$} \deleted{$t_{10}$} for all models are \deleted{$\lesssim 1$ s} \edit1{$\lesssim 2$ s}, whereas \deleted{$t_{10}$} \edit1{$t_{30}$} for {\tt subch\_full} is \deleted{$\sim 1$ s} \edit1{$\sim 2$ s}, consistent with previous observational studies \citep{galloway:2008}.
+\edit1{In the event of extreme downturns in $\dot{e}_{\textrm{nuc}}$ in the future, there is no assurance that the lateral propagating flame for {\tt subch\_full\_mod} would surpass its counterparts.}
+
+Assuming that the fitting functions accurately describe the flame propagation, the expected time for the flame to \edit1{propagate laterally across a neutron star with a radius of 10 km or $\pi R \sim 30$ km}, $t_{30}$, can be calculated, as shown in Table \ref{Tab:network_instan_vel}.
+\edit1{The quantity $t_{30}$}  serves as a rough prediction for the rise time of XRBs with a pure ${}^{4}$He accretion layer.
+\edit1{A basic calculation shows that $t_{30}$} for all models are \edit1{$\lesssim 2$ s}, whereas \edit1{$t_{30}$} for {\tt subch\_full} is \edit1{$\sim 2$ s}, consistent with previous observational studies \citep{galloway:2008}.
 \edit1{Here we emphasize that $t_{30}$ is extrapolated beyond the data with the assumption that the flame has a sustained constant acceleration.
@@
 
-\deleted{As shown in \citet{castro_simple_sdc}, in regions where the burning
+\begin{comment}
+As shown in \citet{castro_simple_sdc}, in regions where the burning
 is vigorous, the simplified-SDC method provides a better solution
@@
 This differs from the case in \citet{castro_simple_sdc} where 
-the simplified-SDC algorithm reduced computational expenses in extreme thermodynamic conditions by mitigating the stiffness of solving reaction equations.}
+the simplified-SDC algorithm reduced computational expenses in extreme thermodynamic conditions by mitigating the stiffness of solving reaction equations.
+\end{comment}
 
@@
 \end{itemize}
-
-\deleted{Comparing the two screening routines, {\tt SCREEN5} and {\tt CHUNGUNOV2007}, we find that {\tt CHUGUNOV2007} has a slightly weaker screening effect in the weak and intermediate screening regimes.
-As a consequence, the weaker screening effect from {\tt CHUGUNOV2007} leads to a slightly slower flame compared to the {\tt SCREEN5} model.
-This result matches our expectations and is in agreement with \cite{Chugunov_2007}.}
-
-\deleted{Finally, we investigated the performance of the simplified-SDC scheme in comparison to the traditional Strang-splitting.
-For this problem, since the burning is not very vigorous, there is no strong benefit of using simplified-SDC over Strang-splitting.}
 

### 2023-09-02 subch_tols/paper.tex
a bit more res
overleaf version

--- 2023-06-25 subch_tols/paper.tex
@@
-\documentclass[modern]{aastex631}
+\documentclass[preprint,linenumbers]{aastex631}
 
@@
 \newcommand{\pynucastro}{{\sf pynucastro}}
+\newcommand{\microphysics}{{\sf Microphysics}}
+\newcommand{\flash}{{\sf Flash}}
 
@@
 
+\newcommand{\add}[1]{{\color{red} #1}}
+
 \setlength{\marginparwidth}{0.5in}
@@
 %======================================================================
-\title{Influence of Time Integration Methodology on Double Detonation Type
-Ia Supernova Nucleosynthesis}
+\title{Sensitivity of Simulations of Double Detonation Type
+Ia Supernova to Integration Methodology}
 
-\shorttitle{Integration Methods and SNe Ia Nucleosynthesis}
+\shorttitle{Double Detonation SNe Ia and Integration Methods}
 
 \author[0000-0001-8401-030X]{Michael Zingale}
-\affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
+\affiliation{Department of Physics and Astronomy, Stony Brook University,
              Stony Brook, NY 11794-3800, USA}
@@
 
+\author[0000-0002-0297-0313]{Melissa Rasmussen}
+\affiliation{Department of Physics and Astronomy,
+Stony Brook University,
+Stony Brook, NY 11794-3800, USA}
+
+\author[0000-0002-1633-6495]{Abigail Polin}
+\affiliation{The Observatories of the Carnegie Institution for Science, 813 Santa Barbara St., Pasadena, CA 91101, USA}
+\affiliation{TAPIR, Walter Burke Institute for Theoretical Physics, 350-17, Caltech, Pasadena, CA 91125, USA}
+
 \author[0000-0003-0439-4556]{Max Katz}
-\affiliation{Dept.\ of Physics and Astronomy, Stony Brook University,
+\affiliation{Department of Physics and Astronomy, Stony Brook University,
              Stony Brook, NY 11794-3800, USA}
@@
 \begin{abstract}
-We study the coupling of hydrodynamics and reactions in simulations of
-the double detonation model for Type Ia supernovae.  When assessing the
-convergence of simulations, the focus is usually on spatial resolution, 
+We study the coupling of hydrodynamics and reactions in simulations of the double detonation model for Type Ia supernovae.  When assessing the
+convergence of simulations, the focus is usually on spatial resolution; 
 however, the method of coupling the physics together as well as the
@@
 In this paper, we explore how the choices made in both coupling and integrating the reaction
-portion of a simulation influences the accuracy, efficiency, and the nucleosynthesis
-of simulations of the double detonation model of Type Ia supernovae.
-We find that the stronger coupling between hydrodynamics and reactions,
-afforded by the simplified spectral deferred corrections method introduced
-previousl,y provides a more accurate and efficient result than standard
-operator splitting.  We also find no need to limit reactions rates or reduce the
+portion of a simulation (operator / Strang splitting vs.\ the simplified spectral deferred corrections method we introduced previously) influences the accuracy, efficiency, and the nucleosynthesis
+of simulations of double detonations.
+We find no need to limit reaction rates or reduce the
 simulation timestep to the reaction timescale.  The entire simulation
-methodology used here is freely available as part of the \castro\ simulation
+methodology used here is GPU-accelerated and made freely available as part of the \castro\ simulation
 code.
@@
 
-There are many different progenitor models for Type Ia supernova (SN Ia) that
-are currently being explored to explain the observed diversity of SN Ia explosions \cite{taubenberger:2017}.  The double detonation model
-\citep{fink:2007} involves a sub-Chandrasekhar mass white dwarf with a
-thin accreted He layer.  A detonation in the He layer propagates
+Many different progenitor models for Type Ia supernovae (SNe Ia) 
+are currently being explored to explain the observed spectroscopy diversity of SNe Ia explosions \citep{taubenberger:2017}. One model, the double detonation model
+\citep{NOMOTO82,WOO_WEAV84}, involves a sub-Chandrasekhar mass white dwarf with an accreted He layer. A detonation in the $\mathrm{He}$ layer propagates
 around the star while sending a compression wave converging toward the
 center.  This compression wave ignites a carbon detonation that then
-propagates outward from the center of the white dwarf.  Many groups
-have modeled this scenario, varying the initial conditions,
-resolution \citep{rivas:2022}, and reaction networks to understand the robustness of this
-model as well as its connection to observations.
+propagates outward from the center of the white dwarf, releasing enough
+nuclear binding energy to gravitationally unbind the star.
 
-\citep{townsley:2019} \citep{gronow:2020}, \citep{gronow:2021}, \citep{glasner:2018}, \citep{roy:2022}, \citep{shenbildsten:2014}, \cite{polin:2019}, \citep{kushnirkatz:2020}
+Many simulations have been performed on this model, varying the initial conditions, reaction network sizes, and resolution. Recent works show
+that, particularly in the case of a thin helium shell, a double detonation is viable progenitor for some SNe Ia \citep{fink:2007,mollwoosley:2013,shenbildsten:2014,glasner:2018,shen:2018,polin:2019,townsley:2019,kushnirkatz:2020,gronow:2020,gronow:2021,boos:2021,roy:2022,rivas:2022}.
+These works use slightly different methods of integrating the reaction network and controlling the timestep, which we discuss in the sections below.
+Our goal is to understand how sensitive a simulation of double detonations is to the details of how the reaction
+network is integrated and how it is coupled to hydrodynamics.
 
-These events are modeled using the Euler equations with reactive and
+All of these works solve the compressible Euler equations with reactive and
 gravitational sources:
 \begin{eqnarray}
+\label{eq:euler1}
 \ddt{\rho} + \nabla \cdot (\rho \Ub) &=& 0 \\
@@
 \ddt{(\rho E)} + \nabla \cdot (\rho \Ub E + \Ub p) &=& \rho \Ub \cdot \gb + \rho \dot{S} \\
+\label{eq:euler4}
 \ddt{(\rho X_k)} + \nabla \cdot (\rho \Ub X_k) &=& \rho \omegadot_k
@@
 total energy, $p$ is the pressure, and $X_k$ are the mass fractions
-of the nuclei that react.  The sources are the gravitational acceleration,
-$\gb$, which is obtained by solving the the Poisson equation
+of the nuclei that react.  The gravitational acceleration,
+$\gb$, is obtained by solving the Poisson equation,
 \begin{equation}
-\nabla^2 \Phi = 4 \pi G \rho
+\nabla^2 \Phi = 4 \pi G \rho \enskip ,
 \end{equation}
 (where $G$ is Newton's constant) for the gravitational potential, $\Phi$, and defining $\gb = -\nabla \Phi$.
-The reaction network provides the species creation rate, $\omegadot$ and the energy generation rate, $\dot{S}$.
+We note that $\Phi$ is often approximated as a monopole or via a multipole expansion, which works well so long as the star
+stays reasonably spherically symmetric.
+The reaction network provides the species creation rate, $\omegadot_k$, and the energy generation rate, $\dot{S}$.
 Finally, the system is closed via the equation of state:
-\begin{equation}
-p = p(\rho, e, X_k)
-\end{equation}
+\begin{eqnarray}
+p &= p(\rho, e, X_k) \\
+T &= T(\rho, e, X_k)
+\end{eqnarray}
 where $e$ is the specific internal energy, obtained as $e = E - |\Ub|^2/2$.
 
-Simulation codes use different methods to advance this system of equations.  For hydrodynamics,
-explicit-in-time integration is usually used, with the timestep, $\Delta t$, restricted by the Courant limit.  In two-dimensions, this is:
-\begin{equation}
-\Delta t = C \min_{i,j} \left \{ \frac{\Delta x}{|u_{i,j}| + c_{i,j}}, \frac{\Delta y}{|v_{i,j}| + c_{i,j}}  \right \}
-\end{equation}
-where $\Delta x$ and $\Delta y$ are the grid width in the $x$- and $y$-directions, $u$ and $v$
-are the $x$- and $y$-components of the velocity and $c$ is the sound
-speed, and the minimum is taken over all zones on the grid.  Here, the
-Courant number, $C$ is required to be $C < 1$ for stability.  We note that
-this is the form of the stability criterion for the corner transport upwind
-formulation \citep{ppmunsplit} of hydrodynamics we use.  Dimensionally-split
-or method-of-lines formulations will have a slighly more restrictive criteria.
-Reactions, on the other hand, tend to be stiff, and are usually evolved
+This is a multiphysics system of equations and the different physical processes have different associated timescales.
+Correspondingly, simulation codes use different methods to advance this system of equations.  For hydrodynamics,
+explicit-in-time integration is usually used, with the timestep, $\Delta t$, restricted by the Courant limit (the precise form depends on whether we are doing
+unsplit hydrodynamics,  dimensionally-split, or method of lines time-integration, see \citealt{ppmunsplit}).
+%In two-dimensions, this is: \MarginPar{I thinks we can skip this! -alex}
+%\begin{equation}
+%\Delta t = C \min_{i,j} \left \{ \frac{\Delta x}{|u_{i,j}| + c_{i,j}}, \frac{\Delta y}{|v_{i,j}| + c_{i,j}}  \right \}
+%\end{equation}
+%where $\Delta x$ and $\Delta y$ are the grid width in the $x$- and $y$-directions, $u$ and $v$
+%are the $x$- and $y$-components of the velocity and $c$ is the sound
+%speed, and the minimum is taken over all zones on the grid.  Here, the
+%Courant number, $C$ is required to be $C < 1$ for stability.  We note that
+%this is the form of the stability criterion for the corner transport upwind
+%formulation \citep{ppmunsplit} of hydrodynamics we use here.  Dimensionally-split
+%or method-of-lines formulations will have a different, slightly more restrictive %criterion.
+In contrast, reactions are usually evolved
 using implicit-in-time integration, with many small steps taken to
@@
 
-When using operator-splitting, the density remains constant, since there
+When using operator splitting, the density remains constant during the reactive update, since there
 is no reaction source in the mass continuity equation.  The mass
@@
 \frac{dX_k}{dt} &= \omegadot_k(\rho, T, X_k) \label{eq:stranga} \\
-\frac{de}{dt} &= \dot{S}(\rho, T, X_k) \\
+\frac{de}{dt} &= \dot{S}(\rho, T, X_k)  \label{eq:strang_energy}\\
 T &= T(\rho, e, X_k) \label{eq:strangc}
@@
 evolution of the mass fractions is solved alone, without integrating
-the temperature or energy.  This is the form that was used, e.g., in FLASH
-\citep{flash}.  A more accurate operator splitting also includes the
-energy evolution (perhaps in terms of temperature), solving the system Eqs.~\ref{eq:stranga}--\ref{eq:strangc}.  This will require
-calling the equation of state each time we evaluate the righthand side
+the temperature or energy, Eq.~\ref{eq:strang_energy}.  This is the form that is used, e.g., in \flash\
+(\citealt{flash}, see also \citealt{townsley:2016}).  A more accurate operator splitting also includes the
+energy evolution (perhaps in terms of temperature), solving the system Eqs.~\ref{eq:stranga}--\ref{eq:strangc}.  This is the default in \castro~\citep{castro,castro_joss},
+and temperature evolution is also used in the works of \citet{gronow:2020,gronow:2021} following the implementation of \citet{pakmor:2012}. This evolution requires
+calling the equation of state each time we evaluate the right-hand side
 of the ODE system, potentially increasing the computational expense.  However, as we showed in
-\citet{strang_rnaas}, this is needed to get second-order convergence.
+\citet{strang_rnaas}, this is needed to get second-order convergence.  In our \amrex-Astro \microphysics\ library \citep{microphysics},
+we use C++ templating heavily to reduce the computational
+expense of the EOS calls by only computing the thermodynamic quantities needed by the integrator.
 
-Detonations can be difficult to model numerically.  One common
-approach is to use a simultion timestep based on the reaction
-timescale, as discussed in \citet{prometheus}, instead of the
-hydrodynamics timescale.  For the double detonation scenario, the
-ignition of the second detonation can be very sensitive to how the
-simulation is run \MarginPar{refs}.  Artifically limiting the
-rates \MarginPar{ref}
+Detonations can be difficult to model numerically:
+the explosive energy released from reactions occurs
+on a much shorter timescale than
+the sound-crossing time of a zone.  Many
+simulations use a simulation timestep based on the reaction
+timescale.
+For example, \citet{prometheus} advocated for two constraints, one that limits the change in $T$ over a timestep and another that limits the change of each mass fraction, $X_k$ over a timestep.
+These can both be much more
+restrictive than the hydrodynamics timescale.  Many current works instead use a reaction timestep based on the change in internal energy, $e$, over the timestep. 
+ This is an option in \flash.  This energy-based reaction timestep limiter
+ is used, for example, in the double detonation simulations of \citet{kushnirkatz:2020,rivas:2022}.  This helps keep the reactions
+and hydrodynamics coupled, but can greatly increase the expense of the
+simulation.  It is also the case that since the timestep is so small, $T$ does not change much over the simulation timestep, so evolving only the mass fractions may be reasonable.  
+Artificially limiting the
+rates \citep{kushnir:2013, shen:2018, kushnirkatz:2020, boos:2021} is also sometimes employed to prevent the reactions from from releasing energy faster than the sound crossing timescale of a cell.
+Once the detonation ignition region is spatially resolved, this should not be needed.
+As we will see in the simulations presented here, if the reactions explicitly know about the advection that is taking place during the burn, neither of these limiting techniques appear to be needed.
+
+In this paper, we focus on nucleosynthesis, exploring how reactions and hydrodynamics are coupled and the role of tolerances, the ignition of the second detonation,  and the
+efficiency of the overall method.  We also look at how the results behave with resolution.  Finally, we discuss some of the difficulties encountered during the simulations.
+
+%======================================================================
+% Results
+%======================================================================
+\section{Simulation Setup}\label{Sec:setup}
+
+\subsection{Numerical method}
+
+We use the freely-available \castro\ simulation code for all the
+simulations shown here.  
+\castro\ solves the compressible Euler equations, Eqs. \ref{eq:euler1}--\ref{eq:euler4}, using an unsplit PPM algorithm for advection \citep{ppmunsplit,millercolella:2002,ppm}, with
+the Riemann solver from \citet{colellaglaz:1985} and the general
+stellar equation of state of \citet{timmes_swesty:2000}.  \castro\ also advects
+internal energy together with total energy, in the dual-energy formalism introduced in \citet{bryan:1995} (see \citealt{wdmergerI} for details on the implementation).  In
+all simulations, the CFL constraint is the
+only timestep constraint used.  Self-gravity is done using a full
+Poisson solve using geometric multigrid, with Dirichlet boundary
+conditions on the domain boundary computed via a multipole expansion
+with a maximum order of 6.  \castro\ uses adaptive mesh refinement (AMR) via the \amrex\ library
+\citep{amrex_joss}, with 
+subcycling throughout the AMR hierarchy, so finer grids are advanced at a smaller timestep than the
+coarse grids.
 
 In \citet{castro_simple_sdc}, we introduced a method based on the
-ideas of spectral deferred corrections for coupling reactions and
-hydrodyamics (that we called ``simplified-SDC'').  Here, the overall
+ideas of spectral deferred corrections \citep{dutt:2000, bourlioux:2003} for coupling reactions and
+hydrodynamics.  We termed the method ``simplified-SDC.''  In
+the simplified-SDC method, the overall
 time-integration is done iteratively, with the hydrodynamics seeing an
-explicit reactive source, and the reaction update evolving an ODE
+explicit reactive source and the reaction update evolving an ODE
 system that includes a piecewise-constant-in-time advective source,
@@
 the timestep, as computed by the corner transport upwind formulation
-of the piecewise parabolic method
-\citep{ppmunsplit,millercolella:2002,ppm}. Crucially, this advective
-term includes an explicit reaction source term and the update,
-Eq.~\ref{eq:simplesdc} means that the reactions will ``see'' what
+of the piecewise parabolic method. Crucially, this advective
+term includes an explicit reaction source term and, likewise, the reaction update,
+Eq.~\ref{eq:simplesdc}, means that the reactions will ``see'' what
 advection is doing as the reaction network is integrated.
+
+The reaction system is integrated using a version of the VODE ODE
+integrator \citep{vode} ported to C++.  Our modifications to this
+integrator are described in \citet{castro_simple_sdc}.  We use an
+analytic approximation to the Jacobian (the main approximation is that
+the species derivatives of the screening function are not included)
+and found that we get the best results when we disable the caching of the
+Jacobian in the integrator.
 
@@
 where the $y_i$ is one of the variables being integrated by the ODE
-integrator and $i$ is the index, indicating that different tolerances
+integrator and $i$ its the index, indicating that different tolerances
 can be used for each variable.  Traditionally one set of tolerances is
 used for the mass fractions and another for the energy (if it is
-integrated).  Unfortunately, the tolerances used by simulations are
+integrated).
+We will refer to the species tolerances as $\epsilon_\mathrm{abs}(X_k)$ and $\epsilon_\mathrm{ref}(X_k)$ and the energy tolerances as $\epsilon_\mathrm{abs}(e)$ and $\epsilon_\mathrm{rel}(e)$.
+Unfortunately, the tolerances used by simulations are
 not normally reported in papers, and in many cases, may not be easily
-controlled by a code user at runtime (and therefore, users may not be
-aware of what they are using).  Some examples can be found in publicly
+controlled by a code user at runtime. Some examples can be found in publicly
 available networks.  For example, the freely available version of the
@@
 floor on the mass fractions corresponding to $\epsilon_\mathrm{abs} =
-10^{-6}$.  Since the simplified-SDC update works in terms of the
-conserved variables, $\Uc$, we apply a density weighting to the absolute tolerances when comparing to the values used in Strang integration.
-
-We considered an extreme version of the double detonation scenario in
-\citet{castro_simple_sdc}---a very large perturbation was applied to
-drive a detonation directly into the underlying carbon-oxygen white
-dwarf.  We showed that the Strang split integration method had
-difficulty with the integration unless we used tighter tolerances,
-while the simplified-SDC method was well-behaved.  Here we look at a
-more realistic setup for a double detonation, with a small
-perturbation that drives a He detonation around the star and a
-compression wave that results in the ignition of a C detonation near
-the center of the star.  We focus on the nucleosynthesis and the
-efficiency of the method.
-
-%======================================================================
-% Results
-%======================================================================
-\section{Simulations and Results}\label{Sec:results}
-
-\subsection{Simulation setup}
-
-We use the freely-available \castro\ simulation code for all the
-simulations shown here.  Our domain uses a 2D axisymmetric geomety
-with a size of $5.12\times 10^9~\mathrm{cm}$ by $1.024\times
-10^{10}~\mathrm{cm}$.  This is much larger than the initial size of
-the star, giving it plenty of room to expand (we note that the figures
-shown in the results section are all zoomed in on the star).  The
-coarse grid is $640\times 1280$ zones and we use 2 levels of
-refinement, each a jump of $2\times$, giving a maximum resolution of
-$20~\mathrm{km}$.  \castro\ uses the \amrex\ library
-\citep{amrex_joss} to manage the grid hierarchy.  The refinement
-strategy is picked to refine on the star and any regions where the
-temperature is greater than $10^8~\mathrm{K}$.  Finally, \castro\ uses
-subcycling, so finer grids are advanced at a smaller timestep than the
-coarse grids.
-
-As noted above, we use the unsplit PPM algorithm for advection, with
-the Riemann solver from \citet{colellaglaz:1985} and the general
-stellar equation of state of \citet{timmes_swesty:2000}.  In
-all simulations, the CFL constraint is the
-only timestep constraint used.  Self-gravity is done using a full
-Poisson solve using geometric multigrid, with Dirichlet boundary
-conditions on the domain boundary computed via a multipole expansion
-with a maximum order of 6.  For the low density regions outside of the
-star, we use a sponge term on the momentum equation to prevent the
-very low density material that is not in hydrostatic equilibrium from
-raining down on the star.
+10^{-6}$.  We caution that different integrators
+may use different error measures or norms so comparing the tolerance values across integrators can be difficult. 
+ Since the simplified-SDC update works in terms of the
+conserved variables, $\Uc$, we apply a density weighting to the absolute tolerances when comparing to the values used in Strang integration (see \citealt{castro_simple_sdc}).
 
@@
 from ReacLib \citep{reaclib} and is produced with
-\pynucastro\ \citep{pynucastro2}.  \MarginPar{show I show the
-  network?} Importantly, we include
+\pynucastro\ \citep{pynucastro2}.  Importantly, we include
 $\isotm{N}{14}(\alpha,\gamma)\isotm{F}{18}(\alpha,p)\isotm{Ne}{21}$
@@
 \citet{shenbildsten} and is important for getting the detonation speed
-correctly.  To reduce the size of the network, we have
+correct.  To reduce the size of the network, we have
 \pynucastro\ combine some $(\alpha,p)(p,\gamma)$ and $(\alpha,\gamma)$
-rates into an effective $(\alpha,\gamma)$ rate.  Screening is provided
+rates into an effective $(\alpha,\gamma)$ rate.  We also approximate
+the neutron captures for $\isotm{C}{12}(\isotm{C}{12},\isotm{n}{})\isotm{Mg}{23}(\isotm{n}{},\gamma)\isotm{Mg}{24}$, $\isotm{O}{16}(\isotm{C}{12},\isotm{n}{})\isotm{S}{31}(\isotm{n}{},\gamma)\isotm{S}{32}$, $\isotm{O}{16}(\isotm{O}{16}, \isotm{n}{})\isotm{S}{31}(\isotm{n}{},\gamma)\isotm{S}{32}$, eliminating
+the intermediate nucleus in each sequence.  Together these approximations make the 22 nuclei network approximate a 31 nuclei network.  Screening is provided
 following the procedure in \citet{wallace:1982}, combining the
@@
 
-The reaction system is integrated using a version of the VODE ODE
-integrator \citep{vode} ported to C++.  Our modifications to this
-integrator are described in \citet{castro_simple_sdc}.  We use an
-analytic approximation to the Jacobian (the main approximation is that
-the species derivatives of the screening function are not included)
-and found that we get best results when we disable the caching of the
-Jacobian in the integrator.
+Our choice of reaction network size falls in the middle of those reported in the literature.
+\citet{townsley:2019,boos:2021} use a 55-isotope network, but do not describe whether energy / temperature is integrated together with the rates (although, based on \citealt{townsley:2016} it likely is not).
+\citet{gronow:2020,gronow:2021} use more than 30 isotopes, and use the integration method from \citet{pakmor:2012}. 
+\citet{rivas:2022} use the 13-isotope {\tt aprox13} network, and likewise do not state whether energy / temperature is integrated together with the rates, although \flash\ does not usually include energy/temperature evolution with the reactions.
+\citet{mollwoosley:2013} used the {\tt aprox19} network together with a table from a large network to better describe the nucleosynthesis.   While they used \castro,
+the temperature evolution during the reactions was simpler then than in the current versions of \castro.  \cite{roy:2022} also use {\tt aprox19}.  
+The {\tt aprox13} and {\tt aprox19} networks do not have the reaction sequences discussed in \citet{shenbildsten} and will lead to an underestimate of the He detonation speed, which in turn affects
+where the compressiion ignites the C detonation.
+None of these works discuss the integration tolerances.
+Several works also disable burning in shocks (this includes \citealt{kushnirkatz:2020,gronow:2020,gronow:2021,rivas:2022} and likely others as this can be the default behavior in some codes).
+We do not take that approach here in part because the burning will explicitly see what is happening from advection as we integrate the network (see also \citealt{papatheodore:2014}).
 
 If during the advance of a timestep an
-error is generated (negative density, ODE integration fails, mass
+error is generated (negative density, ODE integration fails because it cannot meet the tolerances required or takes too many steps, mass
 fractions don't sum to 1, CFL constraint violated at the new time),
-the step is thrown-out and retied with a smaller timestep.  This is
+the step is thrown-out and retried with a smaller timestep.  This is
 done on a level-by-level basis in the overall AMR subcycling
-hierarchy.
+hierarchy.  This is discussed further in \citet{castro_simple_sdc}.
+
+\subsection{Initial model}
+
+We considered an extreme version of the double detonation scenario in
+\citet{castro_simple_sdc}---a very large perturbation was applied to
+drive a detonation directly in a pure carbon
+white dwarf.  We showed that the Strang-splitting integration method (including energy evolution) had
+difficulty with the integration unless we used tighter tolerances,
+while the simplified-SDC method was well-behaved.  Here we look at a
+much more realistic setup for a double detonation
+and look at convergence with integration method and resolution.  Our new model uses a much smaller
+perturbation that drives a He detonation around the C/O white dwarf and a
+compression wave that results in the ignition of a C detonation near
+the center of the star.
 
@@
 envelope mass of $0.05~M_\odot$.  The code for generating the model is
-freely available at
-\url{https://github.com/amrex-astro/initial_models}.  \MarginPar{should
-  we do a zenodo release of this?}
+freely available \citep{initial_models}.
+Our choice of white dwarf mass and
+envelope mass was inspired by our previous work on modeling the convective runaway in double detonations \citep{jacobs:2016}, where we found larger envelopes and/or core masses are needed to get the localized runaway that could lead to a double detonation.
 
@@
 where, $R_\mathrm{base}$ is the radius at which the helium layer
-begins and $R_\mathrm{pert}$ is the distance above the base to put the
+begins (determined as the radius where $X(\isotm{He}{4})$ first goes beyond $0.5$) and $R_\mathrm{pert}$ is the distance above the base to put the
 perturbation.  We choose $R_\mathrm{pert} = 100~\mathrm{km}$.  The
@@
 scale of the perturbation is $\lambda = 12.5~\mathrm{km}$.  This is a
-very small perturbation, but enough to seed the initial He detonation
-in the envelope.  \MarginPar{give the peak temperature}
+very small perturbation, with a peak temperature of about $1.1\times
+10^9~\mathrm{K}$, spread over roughly 2 zones (in our references 20 km resolution simulations).  This
+is large enough to seed the initial He detonation in the
+envelope.
 
-Our base simulation uses the simplified-SDC integration, a CFL number
+\begin{figure}
+\centering
+\plotone{subch_Temp_zoom}
+\caption{\label{fig:initial_hotspot} Zoom-in on the initial hotspot 100 km above the He-C interface.  The dotted contours show densities
+of $10^4$, $10^5$, and $10^6~\gcc$}
+\end{figure}
+
+\subsection{Suite of Runs}
+
+All simulations use a 2D axisymmetric geometry
+with a size of $5.12\times 10^9~\mathrm{cm}$ by $1.024\times
+10^{10}~\mathrm{cm}$.  This is much larger than the initial size of
+the star, giving it plenty of room to expand (we note that the figures
+shown in the results section are all zoomed in on the star).  AMR is used, with a coarse grid of $640\times 1280$ zones.
+For most simulations, 2 levels of
+refinement are used (each a jump of $2\times$) giving a maximum resolution of
+$20~\mathrm{km}$.  The refinement
+strategy is picked to refine the star and any regions where the
+temperature is greater than $10^8~\mathrm{K}$.  For the low-density regions outside of the
+star, we use a sponge term on the momentum equation to prevent the
+very low-density material that is not in hydrostatic equilibrium from
+raining down on the star (see \citealt{eiden:2020} for a discussion of
+the sponge term in \castro).
+
+Our base simulation, which we refer to as SDC-0.2, uses the simplified-SDC integration, a CFL number
 of $0.2$, and sets $\epsilon_\mathrm{rel} = \epsilon_\mathrm{abs} =
@@
 requires 21,784 coarse grid timesteps to evolve to $1.0~\mathrm{s}$ of
-simulation time.
+simulation time.  Table~\ref{table:simulations} summarizes the
+suite of different simulations we consider.  We also show the
+total amount of $\isotm{Ni}{56}$ produced at 1.0~s of simulation time for
+each simulation.
+
+\begin{deluxetable*}{lllllllll}
+\tablecaption{\label{table:simulations} Parameters for our simulation suite.}
+\tablehead{\colhead{name} &
+\colhead{coupling} &
+\colhead{CFL} &
+\colhead{fine grid res} &
+\colhead{$\epsilon_\mathrm{atol}(X_k)$} &
+\colhead{$\epsilon_\mathrm{rtol}(X_k)$} &
+\colhead{$\epsilon_\mathrm{atol}(e)$} &
+\colhead{$\epsilon_\mathrm{rtol}(e)$} &
+\colhead{$M_{\isotm{Ni}{56}}(t = 1.0~\mathrm{s}) / M_\odot$}}
+\startdata
+SDC-0.2 & simplified-SDC & 0.2 & 20 km & $10^{-5}$ &  $10^{-5}$ &  $10^{-5}$ &  $10^{-5}$ & 0.652\\
+SDC-0.4 & simplified-SDC & 0.4 & 20 km & $10^{-5}$ &  $10^{-5}$ &  $10^{-5}$ &  $10^{-5}$ & 0.654 \\
+\tableline
+Strang-0.2 & Strang & 0.2 & 20 km & $10^{-5}$ &  $10^{-5}$ &  $10^{-5}$ &  $10^{-5}$ & 0.734 \\
+Strang-0.05 & Strang & 0.05 & 20 km &$10^{-5}$ &  $10^{-5}$ &  $10^{-5}$ &  $10^{-5}$ & 0.747 \\
+Strang-tol & Strang & 0.2 & 20 km & $10^{-8}$ &  $10^{-5}$ &  $10^{-5}$ &  $10^{-5}$ & 0.665 \\
+\tableline
+SDC-40km & simplified-SDC & 0.2 & 40 km & $10^{-5}$ &  $10^{-5}$ &  $10^{-5}$ &  $10^{-5}$ & 0.647 \\
+SDC-10km & simplified-SDC & 0.2 & 10 km & $10^{-5}$ &  $10^{-5}$ &  $10^{-5}$ &  $10^{-5}$ & 0.652 \\
+SDC-5km & simplified-SDC & 0.2 & 5 km & $10^{-5}$ &  $10^{-5}$ &  $10^{-5}$ &  $10^{-5}$ & 0.652 \\
+\enddata
+\end{deluxetable*}
 
@@
 strategy \citep{castro_gpu} takes advantage of the \amrex\ C++
-lambda-capturing functionality to be performance portable.
-All of the simulation code is on github at
-\url{https://github.com/amrex-astro/} and the inputs files, global
+parallel--for abstraction layer to be performance portable.
+All of the simulation code is on GitHub\footnote{
+\url{https://github.com/amrex-astro/}} and the inputs files, global
 diagnostics used to make the line plots shown below, and metadata
 describing the git hashes, compiler environment, and runtime
-parameters for all of the simulaitons is available at \MarginPar{add
-  Zenodo ref}.
+parameters for all of the simulations is available on Zenodo
+at \citet{metadata}.
+
+\section{Results}
 
@@
 By $0.1~\mathrm{s}$ the detonation is almost halfway around the star and by $0.2~\mathrm{s}$ is almost at the south pole.
-Figure~\ref{fig:lapl_rho_sequence} shows shows a Schlieren-style plot, which
-highlights density gradients.  It is constructed by plotting
-$\log_{10}(|\rho^{-1}\nabla^2\rho|)$.  This clearly shows a
+Figure~\ref{fig:lap_rho_sequence} shows a Schlieren-style plot, which
+highlights density gradients (see, e.g., \citealt{svakhine:2005} a discussion of
+this type of visualization).  We plot
+$\log_{10}(|\rho^{-1}\nabla^2\rho|)$---this clearly shows a
 compression wave launched by the early evolution of the He detonation propagating inward toward
@@
 highlights how the speed of the He detonation is important in
-determining how off-center the C detonation ignition will be.  Our
+determining how off-center the C detonation ignition will be.  
+A faster He detonation means the convergence happens
+at higher density, making the C ignition easier.
+Our
 choice of network was motivated by ensuring that we capture the energy
@@
 In the bottom row of Figure~\ref{fig:temp_sequence}, we see
-the second detonation ignite near the center of the star and
-begin to propagate outward.  By $1.0~\mathrm{s}$ it has evolved
-past the initial radius of the star and is into the extended
+the second detonation ignites near the center of the star and
+begin to propagate outward.  By $1.0~\mathrm{s}$, it has evolved
+past the initial radius of the star and into the extended
 ash layer left behind by the previous He detonation.
@@
 
-We now look at how Strang integration to this base simulation.
-
-We run 3 different types of integration methods: traditional operator
-splitting, using Strang splitting, both with integration of the energy
-equation and without any temperature evolution during the
-burn.  \MarginPar{comment about the EOS}.  These methods are compared
-to the simplified-SDC integtation.  We also vary the CFL number and
-the integration tolerances.
+We now look at the details of the nucleosynthesis
+and how Strang integration compares to our base simplified-SDC simulation.
+We will focus on the case where the energy equation is integrated together with the species while doing Strang-splitting.
+For the Strang comparisons, we will vary the CFL number and the integration tolerances.
+We note that we were not able to run the case where the temperature is held fixed during the Strang burn past the C detonation ignition (for the CFL numbers we consider) without the integrator encountering problems (as we note below, the temperature became unphysically high).
+This is likely why many works taking this approach use a reaction-based timestep.
 
@@
 Figure~\ref{fig:ni56} shows the total amount of \isot{Ni}{56} as a function of time for
-a suite of simulations.
+the suite of simulations.
+The black solid line is for our SDC-0.2 simulation discussed above.
+The general trend we see is that the \isot{Ni}{56} mass increases quickly until 0.2~s, corresponding to the timescale for the He detonation to wrap around the star.
+About $0.02 M_\odot$ of \isot{Ni}{56} is produced during this phase.
+Around 0.7~s, we see the \isot{Ni}{56} mass steeply increase, corresponding to the ignition of the C detonation and then level off at 1.0~s, corresponding to the detonation making it through the entire star.
+We tabulate the \isot{Ni}{56} mass in Table~\ref{table:simulations} at 1.0~s as well.
 
-To understand the difference in the \isot{Ni}{56} mass between the SDC and Strang simulations at 
-CFL = 0.2,
+The blue curves on Figure~\ref{fig:ni56} show the Strang runs with the same default integrator tolerances, but one with CFL = 0.2 (Strang-0.2) and the other with CFL = 0.05 (Strang-0.05).
+We see both of them over-produce \isot{Ni}{56} at the end, but the smaller CFL run agrees with the amount of \isot{Ni}{56} produced during the He det phase.
+Both lead to an earlier ignition of the C detonation.
 
-\begin{figure}[t]
+The orange curve on Figure~\ref{fig:ni56} shows the Strang runs with the CFL = 0.2, but now different integrator tolerances.  The tolerances for
+energy, unchanged: $\epsilon_\mathrm{abs}(e) = \epsilon_\mathrm{rel}(e) = 10^{-5}$, but now the simulation
+labeled Strang-tol uses $\epsilon_\mathrm{abs}(X_k) = 10^{-8}$ and $\epsilon_\mathrm{rel}(X_k) = 10^{-5}$, allowing it to track lower abundances of trace species.
+We see that Strang-tol matches the SDC-0.2 calculation well---the curves are nearly on top of one another.
+This agrees with what we observed with the more extreme model considered in \citet{castro_simple_sdc}, and gives us confidence that the simplified-SDC method works well.
+
+\begin{figure}
 \centering
-\plotone{sdc_strang_compare_abar_summary_plot}
-\caption{\label{fig:abar_sdc_strang} Comparison of the mean molecular weight between SDC and Strang (both run with CFL = 0.2) at $t = 0.2~\mathrm{s}$.}
+\plotone{subch_intermediate}
+\caption{\label{fig:subch_intermediate} Total mass of the
+the nuclei from \isot{Si}{28} to \isot{Ti}{44} vs.\ time for the different simulations.}
 \end{figure}
+
+Figure~\ref{fig:subch_intermediate} shows the total mass 
+of the nuclei from \isot{Si}{28} to \isot{Ti}{44} vs.
+time.
+We see the same general trend as with the \isot{Ni}{56} plot, with the Strang-tol run agreeing with the SDC runs.
+However, in this case, Strang-0.2 and Strang-0.05 underproduce these elements after the C detonation is established.
+
+\begin{figure}
+\centering
+\plotone{subch_temp}
+\caption{\label{fig:subch_temp} Peak temperature vs.\ time
+for the different simulations.}
+\end{figure}
+
+Figure~\ref{fig:subch_temp} shows the peak temperature vs.\ time.
+From the initial hotspot, we see a quick increase to $\sim 3.3\times 10^9~\mathrm{K}$ as the He detonation propagates around the star.
+The peak temperature then
+begins to drop as the main burning in the He layer is over.
+After $0.75~\mathrm{s}$ there is a sharp increase in the temperature to over $7\times 10^9~\mathrm{K}$---this is the ignition of the central C detonation.
+We note that at this extremely high temperature, the nuclei are almost certainly in nuclear statistical equilibrium (NSE), so replacing the burning with an NSE solver would make the simulations more efficient (as done, e.g., in \citealt{ma:2013,kushnirkatz:2020}).
+This will be explored in the future.
+It is also interesting to
+note here that the Strang-0.05 run agrees with the temperature curve for the SDC runs, so reducing the CFL number with Strang coupling does bring the temperature under control.  We also note that if 
+we try to use Strang coupling without any $T$ or energy evolution, the peak temperature reaches $1.2\times 10^{10}~\mathrm{K}$ during the C detonation ignition,
+and we are unable to run further (the ODE integrator encounters problems).
+
+%To understand the difference in the \isot{Ni}{56} mass \MarginPar{should I keep this?}
+%between the SDC and Strang simulations at 
+%CFL = 0.2, Figure~\ref{fig:abar_sdc_strang} shows the mean molecular weight of the composition for the base SDC-0.2 and  Strang-0.2 simulations.
+We see in the SDC-0.2 simulation, the highest mass ash (Ni) is confined to a thin layer on the surface of the white dwarf, while for the Strang-0.2 case, it is much more extended.
+\MarginPar{zhi: will reducing the tolerance for strang (Strang-tol model) help to converge to the sdc model for this abar plot?
+yes, it does, which is what the curves show} The detonation in the Strang case also appears to have moved slightly faster, likely explaining why the ignition of the C detonation for this run was earlier than the base simplified-SDC simulation.
+
+%\begin{figure}[t]
+%\centering
+%\plotone{sdc_strang_compare_abar_summary_plot}
+%\caption{\label{fig:abar_sdc_strang} Comparison of the mean molecular weight %between SDC and Strang (both run with CFL = 0.2) at $t = 0.2~\mathrm{s}$.}
+%\end{figure}
+
+Finally, we consider what happens if we use a longer timestep with the simplied-SDC algorithm.  SDC-0.4 uses a CFL number of 0.4,
+and the \isot{Ni}{56} mass is shown as the gray line in Figure~\ref{fig:ni56}.
+As we see, it is virtually on top of the line corresponding to SDC-0.2, demonstrating that we remain converged when taking a larger timestep.
+We note however, that the ODE integrator does have a harder time evolving the coupled system when the C detonation ignites (usually isolated to just a few zones each timestep).
+As a result, at the point of the C detonation ignition, we switched the integrator to using a difference-approximation to the Jacobian \citep{lsode} instead of the analytic Jacobian,
+and the simulation ran without difficulties afterwards.
+This highlights however that the integrator does work hard at times, and suggests a focus on exploring different integrators or techniques to improve the robustness of the ODE integration.
+
+\subsection{Computational Efficiency}
 
@@
 
+We next want to ask whether the simplified-SDC method is more
+expensive.
 Figure~\ref{fig:cpu} shows the computational expense of the
 simulations, in terms of node hours.  All simulations were run on 4
-nodes (32 AMD GPUs total).  We note that we have not spent any time on
-load-balancing these simulations, so we should consider this to be
-only a guide.  At the start of the simulation, there are 56 boxes on
-the finest grid (representing 0.77\% of the domain) and at 1 s of
-evolution there is are 60 boxes covering 5.64\% of the domain.  The
-figure shows that the Strang simulation with a CFL of 0.2 is the
+nodes (32 AMD GPUs total).  We note that we have not spent much time on
+optimizing the load-balancing of these simulations, so we should consider this to be
+only a guide.  \amrex\ divides the domain up into
+boxes at each level.  On a level, these boxes can vary in size, but all have the same resolution.  We require 
+each box to be a multiple of 32 zones on a side, with a maximum length in a dimension of 256 zones.  At the start of the simulation, there are 56 boxes on
+the finest level (representing 0.77\% of the domain) and at 1 s of
+evolution, there are 60 boxes covering 5.64\% of the domain.  The
+figure shows that the Strang-0.2 simulation is the
 cheapest, but as we saw above, this does not get the nucleosynthesis
 correct.  All of the other Strang simulations are more costly than the
-baseline SDC simulation.  This shows that not only does the
-simplified-SDC simulation accurately evolve the nucleosynthesis, it
+baseline SDC simulation.
+In particular, the Strang run with the tightest tolerances, Strang-tol, that agreed with the \isot{Ni}{56} produced by SDC-0.2 is about $3\times$ more costly.
+As we also see, the SDC-0.4 simulation runs very quickly---about as fast as the Strang-0.2 simulation.
+This shows that not only does the
+simplified-SDC simulation accurately evolves the nucleosynthesis, it
 does so in a very efficient fashion.
@@
 
-\begin{figure}[t]
+Figure~\ref{fig:res_panel} shows a time sequence of the simplified-SDC
+simulation with CFL = 0.2 and our standard tolerances
+($\epsilon_\mathrm{abs} = \epsilon_\mathrm{rel} = 10^{-5}$) at 4
+different maximum spatial resolutions: 40 km, 20 km, 10 km, and 5 km (these are SDC-40km, our original SDC-0.2 simulation, SDC-10km, and SDC-5km in Table~\ref{table:simulations}).  For the SDC-40km
+run, we simply use one fewer level of refinement compared to our SDC-0.2
+simulation.  For the SDC-10km run, we add an additional level of
+refinement (another jump of $2\times$, giving 3 refinement levels total) that captures only the high temperature regions (including
+the initial He layer) but with the very center always refined by
+adding an additional refinement tagging criteria on densities $\rho >
+7.5\times 10^7~\gcc$.  For the 5 km run, we add a fourth
+level of refinement (again a jump of $2\times$), but only
+allow the refinement at the center to go up to this finest level of refinement.  The idea is to understand the effect of increased resolution on the C detonation ignition.  We confirmed
+that for the SDC-5km simulation. the C detonation ignited in the region of the domain refined to 5~km.
+The first column of the figure shows the
+initial grid structure---the colors corresponding to
+each level are the same for the different runs.  Recall
+that the boxes at each level can vary in size, but are a multiple of 32 zones in each
+dimension, with a maximum of 256 zones in any dimension.  For the SDC-10km resolution
+run, we needed to switch from an analytic approximation to the Jacobian
+to a numerical difference approximation once the C detonation ignited (just like with SDC-0.4), because
+the evolution became too challenging for the ODE integrator otherwise.
+
+\begin{figure*}[t]
 \centering
@@
   in each resolution, the AMR grid structure is shown.}
-\end{figure}
+\end{figure*}
+
+%\begin{figure}
+%\centering
+%\plotone{subch_plt17035_slice}
+%\caption{\label{fig:zoom-5km} Close-up of the SDC-5km simulation just after the C detonation ignited.
+% The grid structure is overlayed and shows that the initial detonation is completely contained in the %highest resolution region of the domain.}
+%\end{figure}
+
+%Figure~\ref{fig:zoom-5km} shows the SDC-5km run just
+%around to the point when the C detonation ignites.  The
+%grid structure is also overlaid, and we see that the
+%initial detonation and compression wave are completely
+%contained in the highest-resolution portion of the domain.
 
@@
 \plotone{subch_ni56_res}
-\caption{\label{fig:res} \isot{Ni}{56} mass vs.\ time for 3 different spatial resolutions, using the simplified-SDC time integration method.}
+\caption{\label{fig:res} \isot{Ni}{56} mass vs.\ time for 4 different spatial resolutions, using the simplified-SDC time integration method.}
 \end{figure}
+
+As we see from the time-evolution, the three different resolutions all
+behave largely the same.  The timescale for the He detonation to
+propagate around the star and the timescale at which the second C
+detonation ignite match well.  Figure~\ref{fig:res} shows the
+\isot{Ni}{56} mass produced from the three different simulations.
+Again, the agreement is quite strong.
+
+\begin{figure*}
+\centering
+\epsscale{0.9}
+\plotone{det_structure}
+\caption{\label{fig:det_structure} Structure of the
+C detonation at $t = 0.8~\mathrm{s}$ for the 3 different resolution runs.}
+\epsscale{1.0}
+\end{figure*}
+
+Figure~\ref{fig:det_structure} shows a one-dimensional
+slice (along the horizontal, $r$ coordinate) through the detonation at 0.8~s.  Both the energy generation
+rate and temperature are shown.  For the energy generation
+rate, we see both positive and negative regions, indicating that we are attempting to find the NSE state as we evolve.  The temperature plot shows that the detonation
+is very steep, particularly for the highest-resolution
+simulations.  The slightly different detonation positions
+arise because of the slightly different timings of the He detonation and convergence of the compression wave, which vary with resolution.
+
+We note that these resolution results differ from the conclusions of
+\citet{rivas:2022}.  There are a number of reasons why this might be
+the case.  First, the initial models differ---they use a slightly lower mass white dwarf ($0.98~M_\odot$).  Second, we use a much larger reaction network here that better
+captures the energetics.  This means that the He detonation moves
+faster in our simulations and the subsequent convergence of the pressure wave will be
+more on-center, meaning that the ignition of the C detonation occurs
+at a higher density.  Finally, we are using our new simplified-SDC time-integration, while that work used Strang-splitting.
+We also have a slightly lower resolution than the finest scale used in \citet{townsley:2019}.
 
@@
 Our results suggest that exploring the time-integration method and
-integrator tolerances is an important part of demonstrating
+integrator tolerances are an important part of demonstrating
 convergence of simulations involving explosive reacting flows in
-astrophysics.
+astrophysics.  We showed that the simplified-SDC method that we introduced
+in \citet{castro_simple_sdc} provides an efficient and accurate means to
+model double detonation Type Ia supernovae.  In particular, for converged Ni masses, the simplified-SDC simulations required less computer time than Strang simulations. 
+ And for comparable computer time, the simplified-SDC simulations were much more accurate.
+
+Exploring the integration algorithm and tolerances, in addition to the traditional focus on resolution, was very helpful in
+understanding convergence.   Our
+suite of simulations show that our nucleosynthesis and dynamics appear to be converged.
 
@@
 for the ODE integrator, especially with large
-timesteps.  \MarginPar{add some info on why it fails} In the near
-future we will show how to include various nuclear statistical
-equilibrium approximations into the simplified-SDC formalism while
-retaining second-order accuracy.  We will also explore different
-integrators and auto-differentiation for the Jacobian (in particular,
-the compositional contribution from the screening)
+timesteps.
+At the ignition of the C detonation, it is not uncommon for the temperature in a zone during the burn to increase from $3\times 10^9~\mathrm{K}$ to $6\times 10^9~\mathrm{K}$ during a single hydrodynamics timestep.
+At these temperatures, nuclear statistical equilibrium sets in, and the integrator can spend a lot of time trying to find the equilibrium, reducing its internal timestep in the process.
+The idea behind simplified-SDC coupling is that the flow should respond to a large amount of energy generated on the grid so the burning is tempered in later iterations, but this may not always be possible over a single timestep.
+We will explore an alternate formulations of energy-based timestep limiters that are compatible with simplified-SDC that might help in this regime.
+Sometimes, as we saw, the integrator works better in this regime if we use a numerical approximation to the Jacobian.
+We are exploring heuristics to switch between analytic and numerical Jacobians based on the thermodynamics state as well as auto-differentiation methods for the Jacobian that might work more efficiently.
+Additionally, we are looking at
+different classes of ODE integrators that could be more efficient here.
 
-We encourage authors to include a discussion about the reaction
-integration algorithm and tolerances in papers to permit comparisons
-in the future.
+In the near
+future, we will demonstrate how to include various nuclear statistical
+equilibrium approximations into the simplified-SDC formalism. 
+\citet{kushnirkatz:2020} showed how enforcing NSE can help with the efficiency of the integration.
+For our integration methods, where temperature varies in the reaction network as we integrate, the procedure is more complicated.
+We can
+begin a burn with the full network and then need to switch to NSE midway through, once the temperature increases to the point where NSE is established.
+Care must be taken to do this in a second-order fashion in the simplified-SDC framework.
+
+We will use the simplified-SDC simulation framework demonstrated here to explore the science case of the double detonation SN Ia model.
+The initial state of the atmosphere is not still, but there will be a convective velocity field built up to the ignition of the detonation.
+This has been modeled \citep{jacobs:2016,glasner:2018} and we can explore how the detonation propagation is affected by this state.
+It is straightforward to consider larger
+networks in our framework, as needed, using \pynucastro.
+The follow-on goal is to add more iron-group nuclei and the weak rates that link them to our current network and explore the distribution of iron and nickel produced in these explosions.
+
+The main three-dimensional effect to explore is multi-point ignition.
+This was looked at in \citet{mollwoosley:2013}.
+The question is whether the ignition of the central detonation still hapopens when multiple He detonations send compression waves inward.
+They found that it still can occur, but this should be revisited with larger networks and our new integration methodology.
+On the OLCF Frontier machine, a 20~km resolution simulation in 3D requires about $100\times$ more compute time, which translates into $\mathcal{O}(10^4)$ node hours, spread across 1024 GPUs initially (more will be needed as it refines).
+This is feasible with our simulation framework.
 
@@
 \castro\ is open-source and freely available at
-\url{http://github.com/AMReX-Astro/Castro}.  The problem setup used
-here is available in the git repo as {\tt subchandra}.
+\url{https://github.com/AMReX-Astro/Castro}.  The problem setup used
+here is available in the git repo as {\tt subchandra}.  The reaction network infrastructure
+is contained in the AMReX-Astro Microphysics repository at \url{https://github.com/AMReX-Astro/Microphysics}.  The initial model routines
+are available at \url{https://github.com/AMReX-Astro/initial_models}.
 
